---
title: "Preparation 01: Import and align light, sleep, and wear data"
subtitle: "From fixed site releases to checked one-minute measurement records"
date: last-modified
engine: knitr
bibliography: ../../bibliography.bib
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-01
#| include: false

library(digest)
library(dplyr)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)

stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )

  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

manifest_registry <- tibble::tribble(
  ~Stage, ~Manifest,
  "Imported and state-aligned measurements",
  "artifacts/12_manifests/import_alignment_artifacts.csv",
  "Prepared sleep and wear intervals",
  "artifacts/12_manifests/state_interval_artifacts.csv"
)

manifest_checks <- vector("list", nrow(manifest_registry))
manifest_objects <- vector("list", nrow(manifest_registry))

for (index in seq_len(nrow(manifest_registry))) {
  manifest <- read_stored_csv(manifest_registry$Manifest[index])
  assert_columns(
    manifest,
    c("path", "sha256", "bytes"),
    manifest_registry$Manifest[index]
  )
  listed_paths <- absolute_path(manifest$path)
  exists <- file.exists(listed_paths)
  observed_bytes <- rep(NA_real_, length(listed_paths))
  observed_sha256 <- rep(NA_character_, length(listed_paths))
  observed_bytes[exists] <- file.info(listed_paths[exists])$size
  observed_sha256[exists] <- vapply(
    listed_paths[exists],
    digest::digest,
    character(1),
    algo = "sha256",
    file = TRUE
  )

  manifest_checks[[index]] <- tibble::tibble(
    Stage = manifest_registry$Stage[index],
    `Files listed` = nrow(manifest),
    `Files present` = sum(exists),
    `Size checks` = sum(exists & observed_bytes == manifest$bytes),
    `SHA-256 checks` = sum(
      exists & observed_sha256 == manifest$sha256
    ),
    Check = ifelse(
      all(
        exists &
          observed_bytes == manifest$bytes &
          observed_sha256 == manifest$sha256
      ),
      "PASS",
      "FAIL"
    )
  )
  manifest_objects[[index]] <- manifest
}

manifest_summary <- dplyr::bind_rows(manifest_checks)
stopifnot(all(manifest_summary$Check == "PASS"))

pinned_downloads <- read_stored_csv(
  "artifacts/12_manifests/pinned_downloads.csv"
)
assert_columns(
  pinned_downloads,
  c(
    "site", "modality", "local_path", "sha256", "bytes", "downloaded"
  ),
  "pinned source-file inventory"
)
source_paths <- absolute_path(pinned_downloads$local_path)
source_exists <- file.exists(source_paths)
source_bytes <- rep(NA_real_, length(source_paths))
source_sha256 <- rep(NA_character_, length(source_paths))
source_bytes[source_exists] <- file.info(source_paths[source_exists])$size
source_sha256[source_exists] <- vapply(
  source_paths[source_exists],
  digest::digest,
  character(1),
  algo = "sha256",
  file = TRUE
)
source_identity_check <- tibble::tibble(
  `Source files listed` = nrow(pinned_downloads),
  `Files present` = sum(source_exists),
  `Size checks` = sum(source_exists & source_bytes == pinned_downloads$bytes),
  `SHA-256 checks` = sum(
    source_exists & source_sha256 == pinned_downloads$sha256
  ),
  `Retrieved during the accepted reuse run` = sum(
    pinned_downloads$downloaded %in% TRUE
  ),
  Check = ifelse(
    all(
      source_exists &
        source_bytes == pinned_downloads$bytes &
        source_sha256 == pinned_downloads$sha256
    ) && !any(pinned_downloads$downloaded %in% TRUE),
    "PASS",
    "FAIL"
  )
)
stopifnot(source_identity_check$Check == "PASS")
source_identity_display <- tibble::tibble(
  `Identity check` = c(
    "Fixed source files listed",
    "Files present",
    "Byte-size checks",
    "SHA-256 checks",
    "Files retrieved during the accepted reuse run",
    "Overall result"
  ),
  `Stored outcome` = c(
    source_identity_check$`Source files listed`,
    source_identity_check$`Files present`,
    source_identity_check$`Size checks`,
    source_identity_check$`SHA-256 checks`,
    source_identity_check$`Retrieved during the accepted reuse run`,
    source_identity_check$Check
  )
)

site_registry <- read_stored_csv("config/site_display_registry.csv")
assert_columns(
  site_registry,
  c("site", "display_order", "display_name", "color_hex"),
  "site display registry"
)

source_audit <- read_stored_csv("artifacts/02_aligned/source_audit.csv")
key_audit <- read_stored_csv("artifacts/02_aligned/key_audit.csv")
time_audit <- read_stored_csv("artifacts/02_aligned/time_audit.csv")
stream_epoch_audit <- read_stored_csv(
  "artifacts/02_aligned/stream_epoch_audit.csv"
)
join_audit <- read_stored_csv(
  "artifacts/02_aligned/join_cardinality_audit.csv"
)
state_conflict_audit <- read_stored_csv(
  "artifacts/02_aligned/state_conflict_audit.csv"
)
state_interval_summary <- read_stored_csv(
  "artifacts/02_aligned/state_interval_summary.csv"
)
state_interval_audit <- read_stored_csv(
  "artifacts/02_aligned/state_interval_audit.csv"
)
saturation_audit <- read_stored_csv(
  "artifacts/02_aligned/saturation_audit.csv"
)
sample_audit <- read_stored_csv("artifacts/02_aligned/sample_audit.csv")
dst_fold_audit <- read_stored_csv("artifacts/02_aligned/dst_fold_audit.csv")

assert_columns(
  source_audit,
  c(
    "site", "location", "modality", "commit", "doi", "source_rows",
    "source_participants", "source_start_utc", "source_end_utc"
  ),
  "source audit"
)
assert_columns(
  sample_audit,
  c(
    "site", "placement", "stage", "rows", "participants",
    "participant_days", "finite_medi", "finite_light",
    "finite_medi_light_pairs", "complete_native_minute_schedules"
  ),
  "sample audit"
)

source_identity <- source_audit |>
  dplyr::group_by(
    .data$site,
    .data$location,
    .data$commit,
    .data$doi
  ) |>
  dplyr::summarise(
    `Files imported` = dplyr::n(),
    `Light placements` = paste(
      dplyr::recode(
        sort(.data$modality[grepl("^light_", .data$modality)]),
        light_glasses = "Near-eye",
        light_chest = "Chest"
      ),
      collapse = ", "
    ),
    .groups = "drop"
  ) |>
  dplyr::left_join(site_registry, by = "site", relationship = "one-to-one") |>
  dplyr::arrange(.data$display_order) |>
  dplyr::transmute(
    Site = .data$display_name,
    `Recorded location` = .data$location,
    `Fixed release` = paste0(
      "**DOI:** ", .data$doi,
      "<br>**Revision:** ", substr(.data$commit, 1L, 10L)
    ),
    .data$`Files imported`,
    .data$`Light placements`
  )

aggregation_regimes <- stream_epoch_audit |>
  dplyr::group_by(
    .data$stream_epoch_seconds,
    .data$expected_subepochs_per_minute
  ) |>
  dplyr::summarise(
    `Participant-device streams` = dplyr::n(),
    Sites = dplyr::n_distinct(.data$site),
    `One-minute rows` = sum(.data$minute_rows),
    .groups = "drop"
  ) |>
  dplyr::arrange(.data$stream_epoch_seconds) |>
  dplyr::transmute(
    `Original recording interval` = paste0(
      .data$stream_epoch_seconds,
      " s"
    ),
    `Expected readings per minute` = .data$expected_subepochs_per_minute,
    .data$`Participant-device streams`,
    .data$Sites,
    .data$`One-minute rows`
  )

aligned_sample <- sample_audit |>
  dplyr::filter(.data$stage == "state_aligned_one_minute")

placement_summary <- aligned_sample |>
  dplyr::group_by(.data$placement) |>
  dplyr::summarise(
    Participants = sum(.data$participants),
    `Participant-days` = sum(.data$participant_days),
    `One-minute grid rows` = sum(.data$rows),
    `Grid hours` = sum(.data$rows) / 60,
    `Valid melEDI minutes` = sum(.data$finite_medi),
    `Valid illuminance minutes` = sum(.data$finite_light),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    )
  ) |>
  dplyr::select(.data$Placement, dplyr::everything(), -.data$placement)

placement_summary_display <- placement_summary |>
  tidyr::pivot_longer(
    cols = -Placement,
    names_to = "Measure",
    values_to = "Value"
  ) |>
  tidyr::pivot_wider(names_from = Placement, values_from = Value)

site_sample <- aligned_sample |>
  dplyr::select(
    .data$site,
    .data$placement,
    .data$participants,
    .data$participant_days,
    minute_rows = .data$rows
  ) |>
  tidyr::pivot_wider(
    names_from = .data$placement,
    values_from = c(
      .data$participants,
      .data$participant_days,
      .data$minute_rows
    ),
    names_glue = "{placement}_{.value}"
  ) |>
  dplyr::left_join(site_registry, by = "site", relationship = "one-to-one") |>
  dplyr::arrange(.data$display_order) |>
  dplyr::transmute(
    Site = .data$display_name,
    `Near-eye participants` = .data$glasses_participants,
    `Near-eye participant-days` = .data$glasses_participant_days,
    `Near-eye minutes` = .data$glasses_minute_rows,
    `Chest participants` = .data$chest_participants,
    `Chest participant-days` = .data$chest_participant_days,
    `Chest minutes` = .data$chest_minute_rows
  )

state_interval_display <- state_interval_summary |>
  dplyr::left_join(site_registry, by = "site", relationship = "one-to-one") |>
  dplyr::arrange(.data$display_order) |>
  dplyr::transmute(
    Site = .data$display_name,
    `Sleep intervals` = .data$sleep_intervals,
    `Sleep participants` = .data$sleep_participants,
    `Sleep entries reviewed` = .data$sleep_audit_rows,
    `Wear intervals` = .data$wear_intervals,
    `Wear participants` = .data$wear_participants,
    `Wear entries reviewed` = .data$wear_audit_rows
  )

state_outcomes <- state_conflict_audit |>
  dplyr::group_by(.data$placement) |>
  dplyr::summarise(
    `One-minute rows` = sum(.data$rows),
    `Diary-defined sleep` = sum(.data$diary_sleep_minutes),
    `Sleep without a wear-log label` = sum(
      .data$diary_sleep_without_wear_label
    ),
    `Wear log says off during sleep` = sum(
      .data$wear_off_during_diary_sleep
    ),
    `Wear log says off while awake` = sum(
      .data$wear_off_outside_diary_sleep
    ),
    `Minutes without a wear-log label` = sum(.data$wear_state_missing),
    `Bedside sleep minutes retained` = sum(
      .data$retained_bedside_sleep_minutes
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    )
  ) |>
  dplyr::select(.data$Placement, dplyr::everything(), -.data$placement)

state_outcomes_display <- state_outcomes |>
  tidyr::pivot_longer(
    cols = -Placement,
    names_to = "Stored outcome",
    values_to = "Minutes"
  ) |>
  tidyr::pivot_wider(names_from = Placement, values_from = Minutes)

saturation_summary <- saturation_audit |>
  dplyr::group_by(.data$placement) |>
  dplyr::summarise(
    `Minutes at or above 100,000 lx melEDI` = sum(.data$saturated_minutes),
    `Affected participants` = sum(.data$affected_participants),
    `Affected participant-days` = sum(.data$affected_participant_days),
    `Lowest affected melEDI` = min(
      .data$minimum_saturated_medi,
      na.rm = TRUE
    ),
    `Highest affected melEDI` = max(
      .data$maximum_saturated_medi,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    )
  ) |>
  dplyr::select(.data$Placement, dplyr::everything(), -.data$placement)

validation_summary <- tibble::tribble(
  ~Check, ~`Stored outcome`, ~Evidence,
  "Source-file identity",
  paste0(
    nrow(pinned_downloads),
    " of ",
    nrow(pinned_downloads),
    " local source files matched their recorded size and SHA-256; none was retrieved during the accepted reuse run"
  ),
  "artifacts/12_manifests/pinned_downloads.csv",
  "Participant-time identifiers",
  paste0(
    sum(key_audit$key_is_unique),
    " of ",
    nrow(key_audit),
    " stage-by-stream keys were unique; ",
    sum(key_audit$missing_key_rows),
    " rows lacked a key"
  ),
  "artifacts/02_aligned/key_audit.csv",
  "Native recording schedules",
  paste0(
    sum(stream_epoch_audit$incomplete_native_minute_schedules),
    " one-minute schedules lacked an expected original reading"
  ),
  "artifacts/02_aligned/stream_epoch_audit.csv",
  "Sleep and wear matching",
  paste0(
    sum(join_audit$cardinality_preserved),
    " of ",
    nrow(join_audit),
    " site-placement joins retained every light row"
  ),
  "artifacts/02_aligned/join_cardinality_audit.csv",
  "Repeated local clock times",
  paste0(
    nrow(dst_fold_audit),
    " repeated wall-clock minutes were retained as distinct UTC instants"
  ),
  "artifacts/02_aligned/dst_fold_audit.csv",
  "State intervals requiring review",
  paste0(
    nrow(state_interval_audit),
    " sleep-derived interval records were excluded from interval use and listed with a reason; no wear interval required review"
  ),
  "artifacts/02_aligned/state_interval_audit.csv",
  "Accepted artifact identities",
  paste0(
    sum(manifest_summary$`Files listed`),
    " manifest entries passed existence, size, and SHA-256 checks"
  ),
  "artifacts/12_manifests/import_alignment_artifacts.csv and state_interval_artifacts.csv"
)

state_rules <- tibble::tribble(
  ~`Recorded information`, ~`Rule applied`,
  ~`Effect on the analytical signal`,
  "Sleep diary",
  "The interval from sleep preparation to wake is treated as sleep, with the start included and the end excluded.",
  "The measurement describes the bedside sleep environment and is retained.",
  "Wear log: off during diary-defined sleep",
  "The diary-defined sleep interval takes precedence for the measurement context.",
  "The bedside sleep measurement is retained.",
  "Wear log: off while awake",
  "This is treated as true non-wear.",
  "melEDI and illuminance are unavailable for analysis during that minute.",
  "Wear log: sleep or site leave while awake",
  "The label is retained as provenance but does not by itself define non-wear.",
  "The recorded light values remain available if otherwise valid.",
  "No wear-log label",
  "Missing wear information is not converted into an off-body assumption.",
  "The recorded light values remain available if otherwise valid.",
  "melEDI at or above 100,000 lx",
  "The raw value and operating-limit flag are retained.",
  "The analytical melEDI value is unavailable; illuminance is not removed solely by this rule."
)

script_map <- tibble::tribble(
  ~Step, ~Producing_code, ~Reads_operation_and_output,
  1L,
  "scripts/pipeline/build_import_alignment.R<br>scripts/pipeline/import_sources.R",
  "Reads `config/site_sources.csv`, resolves each fixed repository revision and DOI, checks the local source-file fingerprint, and loads the named object. This is separate so source identity is established before any time or state transformation. **Outputs:** cached fixed source files plus `artifacts/02_aligned/source_audit.csv`.",
  2L,
  "scripts/pipeline/time_axes.R<br>scripts/pipeline/aggregation_coverage.R",
  "Reads the original logger records, preserves UTC as actual time and local clock time as wall time, infers each participant-device recording interval, and creates a one-minute value only when every expected original reading is present and finite. This is separate so temporal completeness is checked before diary or wear information can affect validity. **Outputs:** `artifacts/01_imported/minute/` and the time and recording-interval audits.",
  3L,
  "scripts/pipeline/state_alignment.R",
  "Reads the fixed sleep diaries and wear logs, checks interval bounds and overlap, and stores valid half-open intervals. Unusable entries are recorded rather than silently repaired. **Outputs:** `artifacts/02_aligned/state_intervals/`, `state_interval_audit.csv`, and `state_interval_summary.csv`.",
  4L,
  "scripts/pipeline/state_alignment.R<br>scripts/pipeline/state_interval_projection.R",
  "Matches the checked intervals to one-minute light rows without changing row cardinality, derives bedside-sleep, worn, and true non-wear contexts, and applies the melEDI operating limit. Kept separate from interval preparation so matching can be audited independently. **Outputs:** per-site and combined files under `artifacts/02_aligned/` plus state, join, saturation, and sample audits.",
  5L,
  "scripts/pipeline/verify_import_alignment_artifacts.R",
  "Reads the completed outputs and independently checks manifest identity, source-fingerprint propagation, unique keys, signal-validity flags, and the accepted placement totals. It does not rebuild data. **Output:** a pass/fail verification result; no scientific artifact is written."
) |>
  dplyr::mutate(
    `Step and producing code` = paste0(
      "**", .data$Step, "**<br>",
      wrap_provenance_path(.data$Producing_code)
    )
  ) |>
  dplyr::select(
    .data$`Step and producing code`,
    `Reads, operation, separation, and output` =
      .data$Reads_operation_and_output
  )

handoff_artifacts <- tibble::tribble(
  ~Stored_artifact, ~Exact_path, ~Use_in_Preparation_02,
  "Combined near-eye one-minute data",
  "artifacts/02_aligned/light_glasses_aligned.rds",
  "Construct complete participant-day grids and assess coverage",
  "Combined chest one-minute data",
  "artifacts/02_aligned/light_chest_aligned.rds",
  "Construct the complementary-placement grids and assess coverage separately",
  "Site-specific sleep and wear intervals",
  "artifacts/02_aligned/state_intervals/",
  "Assign known states to explicitly inserted missing-minute rows before calculating coverage",
  "Preparation 01 audits",
  "artifacts/02_aligned/*.csv",
  "Trace source identity, time handling, joins, state decisions, operating-limit exclusions, and sample accounting",
  "Artifact inventories",
  "artifacts/12_manifests/import_alignment_artifacts.csv and state_interval_artifacts.csv",
  "Confirm the exact files and digital fingerprints passed forward"
)

environment_summary <- tibble::tibble(
  Component = c(
    "R",
    "Quarto profile",
    "Table package",
    "Identity check",
    "Scientific recomputation during render"
  ),
  Value = c(
    paste(R.version$major, R.version$minor, sep = "."),
    "nathealth (HTML review render)",
    paste0("gt ", as.character(utils::packageVersion("gt"))),
    "File size and SHA-256 with digest",
    "None"
  )
)
```

## Purpose and position in the analysis chain

Preparation 01 establishes the measurement record that all later preparation
steps use. The production workflow reads the fixed light-logger, sleep-diary,
and wear-log releases from nine study sites. It checks their identities,
converts the logger observations to complete one-minute values, preserves both
actual and local clock time, and assigns a measurement context to each minute.
Its outputs enter Preparation 02, where hourly and daily coverage is assessed.

The exact inputs are the 35 fixed local source files recorded in
[`pinned_downloads.csv`](../../artifacts/12_manifests/pinned_downloads.csv).
Their repository revisions, DOIs, URLs, byte sizes, and SHA-256 digital
fingerprints are propagated to
[`source_audit.csv`](../../artifacts/02_aligned/source_audit.csv). A SHA-256
fingerprint is used here to establish file identity; it does not describe the
scientific contents of the file.

::: {.callout-note appearance="simple" title="What is calculated during this render?"}
This report reads already stored CSV manifests and audits. During rendering it
checks the existence, byte size, and SHA-256 fingerprint of every listed
Preparation 01 artifact and fixed source file; checks the columns needed for
the displays; and calculates only compact descriptive totals from stored audit
rows. It does not retrieve a source file, import logger data, aggregate a
minute, align a diary or wear period, replace a signal value, rebuild an
artifact, or run the production verifier.
:::

```{mermaid}
%%| label: fig-preparation01-chain
%%| fig-cap: "Position of Preparation 01 in the data-to-analysis chain. The report reads the stored outputs of these operations but does not execute them."
%%| fig-alt: "Flow diagram from fixed site releases through source-identity checks, one-minute aggregation, sleep and wear interval preparation, and state alignment to the one-minute files used for coverage assessment in Preparation 02."
flowchart LR
  A["Fixed site releases<br/>light, sleep diary, wear log"] --> B["Verify source identity<br/>revision, DOI, SHA-256"]
  B --> C["Retain actual and<br/>local clock time"]
  C --> D["Require complete original<br/>readings for each minute"]
  A --> E["Prepare checked sleep<br/>and wear intervals"]
  D --> F["Align measurement context<br/>and validity flags"]
  E --> F
  F --> G["Stored one-minute files<br/>Preparation 02 coverage checks"]
```

## 1. Establish the exact source files

The production import uses one recorded release per site. Table
@tbl-preparation01-source-identity shows the human-readable site names in the
submitted-manuscript order. The source audit contains four inputs per site
(sleep diary, wear log, near-eye light, and chest light), except Tübingen,
where no chest logger file is part of the study input.

```{r}
#| label: tbl-preparation01-source-identity
#| tbl-cap: "Fixed source releases used for the Preparation 01 production build. The displayed revision is shortened for readability; the stored audit retains the complete revision, URL, byte size, and SHA-256 fingerprint."

source_identity |>
  prep_gt(
    source_note = paste0(
      "Complete source identities: ",
      "[artifacts/02_aligned/source_audit.csv]",
      "(../../artifacts/02_aligned/source_audit.csv)."
    )
  ) |>
  gt::fmt_markdown(columns = `Fixed release`) |>
  gt::cols_label(
    `Files imported` = "Files",
    `Light placements` = "Light measurements"
  ) |>
  gt::cols_width(
    Site ~ gt::pct(16),
    `Recorded location` ~ gt::pct(22),
    `Fixed release` ~ gt::pct(30),
    `Files imported` ~ gt::pct(8),
    `Light placements` ~ gt::pct(24)
  )
```

```{r}
#| label: tbl-preparation01-source-file-check
#| tbl-cap: "Identity check for the fixed local source files."

source_identity_display |>
  prep_gt(
    source_note = paste0(
      "The accepted production run reused the already stored fixed files; ",
      "rendering also reads them only for identity checks."
    )
  ) |>
  gt::cols_width(
    `Identity check` ~ gt::pct(68),
    `Stored outcome` ~ gt::pct(32)
  )
```

## 2. Preserve actual time and local clock time

Two time coordinates are retained because they answer different questions.
`datetime_utc` identifies the actual instant and is used to order readings and
measure elapsed time. `datetime_wall` represents the local clock reading, so
08:00 has the same time-of-day meaning at every site. At an autumn clock
change, one local minute can correspond to two actual instants. The stored
audit contains 360 such participant-level wall-clock minutes—120 for Madrid
near-eye data, 120 for Madrid chest data, and 120 for Tübingen near-eye data.
They remain distinct in UTC and are not discarded or averaged at this stage.

The original logger interval is inferred separately for each participant and
placement. A one-minute value is retained only when all expected original
readings in that minute are present and finite. Table
@tbl-preparation01-aggregation-regimes shows the resulting regimes; the
production audits recorded no incomplete original-reading schedule.

```{r}
#| label: tbl-preparation01-aggregation-regimes
#| tbl-cap: "Original logger recording intervals and the completeness requirement used to create one-minute values."

aggregation_regimes |>
  prep_gt(
    source_note = paste0(
      "Stored stream-level evidence: ",
      "[stream_epoch_audit.csv]",
      "(../../artifacts/02_aligned/stream_epoch_audit.csv)."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Expected readings per minute`,
      `Participant-device streams`,
      Sites,
      `One-minute rows`
    ),
    use_seps = TRUE
  ) |>
  gt::cols_width(
    `Original recording interval` ~ gt::pct(23),
    `Expected readings per minute` ~ gt::pct(22),
    `Participant-device streams` ~ gt::pct(23),
    Sites ~ gt::pct(12),
    `One-minute rows` ~ gt::pct(20)
  )
```

## 3. Prepare sleep and wear periods

The sleep diary defines the interval from sleep preparation to wake. Its start
is included and its end is excluded. These measurements characterize the
bedside sleep environment and remain part of the intended hybrid 24-hour
record. The wear log supplies additional context. An `off` label while awake
is treated as true non-wear; `sleep` and `site_leave` labels are retained for
provenance but do not independently remove a light value.

```{r}
#| label: tbl-preparation01-state-rules
#| tbl-cap: "Rules used to translate sleep-diary and wear-log information into measurement context and signal availability."

state_rules |>
  prep_gt() |>
  gt::cols_width(
    `Recorded information` ~ gt::pct(22),
    `Rule applied` ~ gt::pct(36),
    `Effect on the analytical signal` ~ gt::pct(42)
  )
```

Valid intervals are stored for later projection onto complete participant-day
grids. Twenty-two sleep-derived records could not define a supported wake
interval and are listed with participant, source rows, bounds, and reason in
[`state_interval_audit.csv`](../../artifacts/02_aligned/state_interval_audit.csv).
They were not silently repaired. No wear-log interval required review.

```{r}
#| label: tbl-preparation01-state-intervals
#| tbl-cap: "Usable sleep and wear periods stored for each study site, with source entries retained for review."

state_interval_display |>
  prep_gt(
    source_note = paste0(
      "Exact reviewed entries: ",
      "[state_interval_audit.csv]",
      "(../../artifacts/02_aligned/state_interval_audit.csv)."
    )
  ) |>
  gt::tab_spanner(
    label = "Sleep diary",
    columns = c(
      `Sleep intervals`,
      `Sleep participants`,
      `Sleep entries reviewed`
    )
  ) |>
  gt::tab_spanner(
    label = "Wear log",
    columns = c(
      `Wear intervals`,
      `Wear participants`,
      `Wear entries reviewed`
    )
  ) |>
  gt::fmt_integer(columns = -Site, use_seps = TRUE) |>
  gt::cols_label(
    `Sleep intervals` = "Periods",
    `Sleep participants` = "Participants",
    `Sleep entries reviewed` = "Reviewed",
    `Wear intervals` = "Periods",
    `Wear participants` = "Participants",
    `Wear entries reviewed` = "Reviewed"
  ) |>
  gt::cols_width(
    Site ~ gt::pct(18),
    gt::everything() ~ gt::pct(13.7)
  )
```

## 4. Align measurement context without losing rows

Sleep and wear periods are matched to each one-minute light record using
participant and actual time. All 17 site-placement joins retained their input
row count. A missing wear-log label is not interpreted as non-wear. Table
@tbl-preparation01-state-outcomes therefore reports missing wear information
as provenance, not as an automatic exclusion.

```{r}
#| label: tbl-preparation01-state-outcomes
#| tbl-cap: "Stored sleep and wear outcomes after alignment to one-minute light records. Counts are minutes."

state_outcomes_display |>
  prep_gt(
    source_note = paste0(
      "These categories document how the two state sources overlap; they are ",
      "not competing estimates of sleep."
    )
  ) |>
  gt::fmt_integer(columns = c(`Near-eye`, Chest), use_seps = TRUE) |>
  gt::cols_width(
    `Stored outcome` ~ gt::pct(60),
    `Near-eye` ~ gt::pct(20),
    Chest ~ gt::pct(20)
  )
```

## 5. Apply the melEDI operating limit

The valid melanopic equivalent daylight illuminance operating range is
strictly below 100,000 lx melEDI [@condor2023actlumus]. Values at or above
that boundary remain in the stored raw channel and receive an explicit flag,
but the analytical melEDI channel is unavailable for that minute. The rule
does not remove illuminance solely because melEDI reached its limit.

```{r}
#| label: tbl-preparation01-saturation
#| tbl-cap: "One-minute values at or above the 100,000 lx melEDI operating limit."

saturation_summary |>
  prep_gt(
    source_note = paste0(
      "No hour or participant-day is removed in Preparation 01; coverage is ",
      "assessed in Preparation 02."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Minutes at or above 100,000 lx melEDI`,
      `Affected participants`,
      `Affected participant-days`
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_number(
    columns = c(`Lowest affected melEDI`, `Highest affected melEDI`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_label(
    `Minutes at or above 100,000 lx melEDI` = "Affected minutes",
    `Lowest affected melEDI` = "Lowest melEDI (lx)",
    `Highest affected melEDI` = "Highest melEDI (lx)"
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(15),
    gt::everything() ~ gt::pct(17)
  )
```

## 6. Stored sample after import and alignment

The result is a complete set of one-minute grid rows for every imported
participant stream. A grid row can contain an unavailable signal because of
missing original data, non-wear, or the melEDI operating limit; therefore grid
rows and valid signal minutes are reported separately.

```{r}
#| label: tbl-preparation01-placement-sample
#| tbl-cap: "Overall measurement sample stored after one-minute aggregation and state alignment."

placement_summary_display |>
  prep_gt() |>
  gt::fmt_integer(
    columns = c(`Near-eye`, Chest),
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Measure ~ gt::pct(60),
    `Near-eye` ~ gt::pct(20),
    Chest ~ gt::pct(20)
  )
```

```{r}
#| label: tbl-preparation01-site-sample
#| tbl-cap: "Participants, participant-days, and one-minute grid rows by site and measurement placement."

site_sample |>
  prep_gt() |>
  gt::tab_spanner(
    label = "Near-eye",
    columns = starts_with("Near-eye")
  ) |>
  gt::tab_spanner(
    label = "Chest",
    columns = starts_with("Chest")
  ) |>
  gt::fmt_integer(columns = -Site, use_seps = TRUE) |>
  gt::cols_label(
    `Near-eye participants` = "Participants",
    `Near-eye participant-days` = "Participant-days",
    `Near-eye minutes` = "Minutes",
    `Chest participants` = "Participants",
    `Chest participant-days` = "Participant-days",
    `Chest minutes` = "Minutes"
  ) |>
  gt::cols_width(
    Site ~ gt::pct(17),
    gt::everything() ~ gt::pct(13.8)
  )
```

Near-eye and chest measurements remain separate. Chest measurements are a
complementary placement and are not pooled with the near-eye data.

## 7. Checks that had to pass

The stored audits expose the checks performed during production. Rendering
does not rerun their producing verifier; it confirms the identities of the
stored evidence and summarizes the recorded outcomes.

```{r}
#| label: tbl-preparation01-validation
#| tbl-cap: "Validation outcomes recorded for the accepted Preparation 01 outputs."

validation_summary |>
  dplyr::mutate(Evidence = wrap_provenance_path(.data$Evidence)) |>
  prep_gt() |>
  gt::fmt_markdown(columns = Evidence) |>
  gt::cols_width(
    Check ~ gt::pct(22),
    `Stored outcome` ~ gt::pct(52),
    Evidence ~ gt::pct(26)
  )
```

```{r}
#| label: tbl-preparation01-manifest-checks
#| tbl-cap: "File-identity checks performed during this report render for accepted Preparation 01 artifacts."

manifest_summary |>
  prep_gt(
    source_note = paste0(
      "Every listed artifact was hashed during render. No manifest or ",
      "scientific artifact was written."
    )
  ) |>
  gt::cols_width(
    Stage ~ gt::pct(35),
    gt::everything() ~ gt::pct(13)
  )
```

## Production scripts and why the operations are separate

The producing code is listed in execution order in Table
@tbl-preparation01-script-map. Separating source identity, temporal
aggregation, interval preparation, state matching, and verification leaves a
stored checkpoint at each scientifically consequential boundary.

```{r}
#| label: tbl-preparation01-script-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

script_map |>
  prep_gt() |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

## Exact artifacts passed to Preparation 02

Preparation 02 receives the two combined one-minute files and the checked
state intervals. The intervals are needed because Preparation 02 inserts
explicit missing-minute rows before calculating coverage; known sleep, wake,
and non-wear periods must also be assigned to those inserted rows.

```{r}
#| label: tbl-preparation01-handoff
#| tbl-cap: "Stored Preparation 01 artifacts passed to coverage assessment."

handoff_artifacts |>
  dplyr::mutate(Exact_path = wrap_provenance_path(.data$Exact_path)) |>
  prep_gt() |>
  gt::fmt_markdown(columns = Exact_path) |>
  gt::cols_label(
    Stored_artifact = "Stored artifact",
    Exact_path = "Exact path",
    Use_in_Preparation_02 = "Use in Preparation 02"
  ) |>
  gt::cols_width(
    Stored_artifact ~ gt::pct(25),
    Exact_path ~ gt::pct(34),
    Use_in_Preparation_02 ~ gt::pct(41)
  )
```

## Render environment

```{r}
#| label: tbl-preparation01-environment
#| tbl-cap: "Environment used for this documentation-only HTML render."

environment_summary |>
  prep_gt() |>
  gt::cols_width(
    Component ~ gt::pct(38),
    Value ~ gt::pct(62)
  )
```
