---
title: "Preparation 01: Import light data and align diary and wear periods"
execute:
  echo: true
  warning: true
  message: true
  error: false
  freeze: false
---

## What this step does

This notebook imports the exact recorded version of each site's data, combines
the original logger readings into one-minute values, and labels each minute
using the sleep diary and wear log. Preparation 02 then checks hourly and daily
data coverage.

Two time variables are retained. `datetime_utc` records the actual instant and
is used to order measurements and calculate elapsed time and gaps.
`datetime_wall` records local clock time in a shared format, so 08:00 means
08:00 local time at every site. During the autumn daylight-saving change,
repeated local minutes remain separate observations in actual time.

## Rules applied

- Each site's source-file version is identified by repository revision, DOI,
  download address and SHA-256 checksum. A SHA-256 checksum is a digital
  fingerprint used to confirm that a file has not changed.
- The logger's original recording interval is checked separately for every
  site, participant and placement. A one-minute value is calculated only when
  all readings expected within that minute are present and valid.
- Sleep is defined by the diary interval from `sleepprep` to `wake`, with
  the start included and the end excluded. Measurements during that interval
  represent the bedside sleep environment and remain part of the intentional
  hybrid 24-hour measurement.
- Wear-log entries labelled `sleep` or `site_leave` are retained for
  traceability. Only `off` outside diary-defined sleep is treated as true
  non-wear. For those minutes, both melEDI and illuminance are excluded from
  analysis.
- The valid melanopic EDI operating range is strictly below
  \(10^5\) lx melEDI. Values at or above 100,000 lx melEDI are retained in
  `MEDI_raw`, flagged, and excluded from the analytical `MEDI` channel after
  one-minute aggregation [@condor2023actlumus].

## R code used on this page

```{r}
#| label: setup-import-alignment

root_hint <- Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd())
source(file.path(root_hint, "scripts/pipeline/paths_io.R"))
root <- project_root(root_hint)

dependencies <- c(
  "assertions.R",
  "import_sources.R",
  "time_axes.R",
  "state_alignment.R",
  "aggregation_coverage.R",
  "build_import_alignment.R",
  "verify_import_alignment_artifacts.R"
)
for (dependency in dependencies) {
  source(file.path(root, "scripts", "pipeline", dependency))
}

site_argument <- Sys.getenv("NATHEALTH_IMPORT_SITES", unset = "")
requested_sites <- if (nzchar(site_argument)) {
  strsplit(site_argument, ",", fixed = TRUE)[[1L]]
} else {
  NULL
}
cache_policy <- Sys.getenv("NATHEALTH_CACHE_POLICY", unset = "reuse")
```

## Import and align the data

```{r}
#| label: build-import-alignment

import_alignment <- build_import_alignment(
  root = root,
  requested_sites = requested_sites,
  cache_policy = cache_policy,
  saturation_threshold = 100000
)

import_alignment_verification <- if (isTRUE(import_alignment$complete_run)) {
  verify_import_alignment_artifacts(
    root = root,
    saturation_threshold = 100000,
    expected_summary = tibble::tribble(
      ~placement,
      ~rows,
      ~participants,
      ~participant_days,
      ~saturated_minutes,
      ~saturated_participants,
      ~saturated_participant_days,
      "glasses", 1635960L, 143L, 1136L, 39L, 10L, 13L,
      "chest", 1798560L, 157L, 1249L, 87L, 16L, 21L
    )
  )
} else {
  list(
    status = "NOT_RUN_FOR_SUBSET",
    run_label = import_alignment$run_label
  )
}
```

The run label is `r import_alignment$run_label`. A normal run processes all
sites and writes the combined near-eye and chest files used in the main
analysis. If specific sites are requested, the notebook labels the result as a
test run and writes it separately, so it cannot replace the full dataset. The
post-build check is
`r import_alignment_verification$status`.

## Exact input versions

```{r}
#| label: tbl-import-source-audit
#| tbl-cap: "Exact source versions used in Preparation 01."

source_summary <- import_alignment$audits$source_audit |>
  dplyr::select(
    .data$site,
    .data$location,
    .data$modality,
    .data$commit,
    .data$source_rows,
    .data$source_participants,
    .data$source_start_utc,
    .data$source_end_utc
  )
knitr::kable(
  source_summary,
  col.names = c(
    "Site",
    "Location",
    "Data type",
    "Source revision",
    "Rows",
    "Participants",
    "First measurement (UTC)",
    "Last measurement (UTC)"
  )
)
```

The complete source addresses, file sizes and digital fingerprints are stored
in [`source_audit.csv`](../../artifacts/02_aligned/source_audit.csv).

## Check identifiers and time

```{r}
#| label: tbl-import-key-audit
#| tbl-cap: "Checks for duplicate or missing participant-time identifiers before and after conversion to one-minute data."

knitr::kable(
  import_alignment$audits$key_audit,
  col.names = c(
    "Site",
    "Placement",
    "Preparation stage",
    "Identifier",
    "Rows",
    "Participants",
    "Rows missing an identifier",
    "Rows duplicated beyond expectation",
    "Identifier is unique"
  )
)
```

```{r}
#| label: tbl-import-time-audit
#| tbl-cap: "Checks of actual time, original recording intervals, and repeated local clock times."

time_summary <- import_alignment$audits$time_audit |>
  dplyr::select(
    .data$site,
    .data$placement,
    .data$source_rows,
    .data$minute_rows,
    .data$participants,
    .data$native_epoch_seconds,
    .data$expected_subepochs_per_minute,
    .data$minutes_with_incomplete_native_schedule,
    .data$wall_minutes_with_multiple_real_instants
  )
knitr::kable(
  time_summary,
  col.names = c(
    "Site",
    "Placement",
    "Original readings",
    "One-minute rows",
    "Participants",
    "Original recording interval (seconds)",
    "Expected readings per minute",
    "Minutes missing an expected reading",
    "Local minutes occurring twice"
  )
)
```

The exact inferred epoch and completeness requirement for every participant
stream is exported in `stream_epoch_audit.csv`. The compact regime count is:

```{r}
#| label: tbl-import-epoch-regimes
#| tbl-cap: "Participant-device data streams by original recording interval and expected readings per minute."

epoch_regimes <- import_alignment$audits$stream_epoch_audit |>
  dplyr::count(
    .data$site,
    .data$position,
    .data$stream_epoch_seconds,
    .data$expected_subepochs_per_minute,
    name = "participant_streams"
  )
knitr::kable(
  epoch_regimes,
  col.names = c(
    "Site",
    "Placement",
    "Original recording interval (seconds)",
    "Expected readings per minute",
    "Participant-device streams"
  )
)
```

Repeated wall-clock minutes, if present, are recorded with both true instants
and offsets in `dst_fold_audit.csv`. No repeated hour is discarded or averaged
in the real-time representation.

## Add sleep and wear information

```{r}
#| label: tbl-import-join-audit
#| tbl-cap: "Checks that each light record matched the expected number of sleep-diary and wear-log intervals."

knitr::kable(
  import_alignment$audits$join_cardinality_audit,
  col.names = c(
    "Site",
    "Placement",
    "Light rows before matching",
    "Rows after adding sleep diary",
    "Rows after adding wear log",
    "Rows matched to a sleep period",
    "Rows matched to a wear period",
    "No rows duplicated or lost"
  )
)
```

```{r}
#| label: tbl-import-state-conflicts
#| tbl-cap: "Agreement between sleep diary and wear log, and retained bedside sleep measurements."

knitr::kable(
  import_alignment$audits$state_conflict_audit,
  col.names = c(
    "Site",
    "Placement",
    "One-minute rows",
    "Diary-defined sleep minutes",
    "Sleep minutes without a wear-log label",
    "Wear log says sleep outside the diary sleep period",
    "Wear log says off during diary-defined sleep",
    "Wear log says off while awake",
    "Minutes without a wear-log label",
    "Retained bedside sleep minutes"
  )
)
```

Diary or wear entries that cannot define a valid interval are listed with the
participant, source rows, start and end, and reason in
`state_interval_audit.csv`; they are not silently repaired.

The valid sleep and wear intervals are saved for each site under
`artifacts/02_aligned/state_intervals/`, together with their source version and
digital fingerprint. Preparation 02 applies these intervals to every minute,
including explicitly inserted missing-minute rows, before coverage is
calculated. This ensures that missing readings inside a known sleep, wake or
non-wear period are counted in the correct time window.

```{r}
#| label: tbl-import-interval-audit
#| tbl-cap: "Prepared sleep and wear intervals, including records that could not be used and why."

knitr::kable(
  import_alignment$audits$state_interval_summary,
  col.names = c(
    "Site",
    "Usable sleep intervals",
    "Participants with a usable sleep interval",
    "Sleep entries listed for review",
    "Usable wear intervals",
    "Participants with a usable wear interval",
    "Wear entries listed for review"
  )
)
```

## Operating limit and sample accounting

```{r}
#| label: tbl-import-saturation-audit
#| tbl-cap: "One-minute melEDI observations at or above the 100,000 lx melEDI operating limit."

knitr::kable(
  import_alignment$audits$saturation_audit,
  col.names = c(
    "Site",
    "Placement",
    "Operating limit (lx melEDI)",
    "Minutes at or above the limit",
    "Affected participants",
    "Affected participant-days",
    "Lowest affected value (lx melEDI)",
    "Highest affected value (lx melEDI)"
  )
)
```

```{r}
#| label: tbl-import-sample-audit
#| tbl-cap: "Measurement rows and usable signals retained through Preparation 01, by participants and participant-days."

knitr::kable(
  import_alignment$audits$sample_audit,
  col.names = c(
    "Site",
    "Placement",
    "Preparation stage",
    "Rows",
    "Participants",
    "Participant-days",
    "Rows with valid melEDI",
    "Rows with valid illuminance",
    "Rows with both signals valid",
    "Inserted original-interval rows",
    "One-minute rows with every expected original reading"
  )
)
```

The affected high-light observations remain listed for traceability, but no
hour or participant-day is removed here. Preparation 02 applies the approved
50% hourly and 80% daily coverage rule. The daily calculation uses the full
1,440-minute hybrid day, including bedside sleep-environment measurements.
This differs from the preregistered sleep-excluded denominator (technical
record `DEV-055`).

## Files produced

Each site and placement is saved as an R data file (`.rds`) containing one
data object. The file records participant, site, placement, source version,
actual time, local clock time, original recording interval, raw signals,
sleep/wear state and reasons for exclusion. The combined near-eye and chest
files are written only after all requested sites pass the identifier, data
matching, state and operating-limit checks. File paths and digital
fingerprints are recorded in:

`r import_alignment$artifact_manifest_path`

The sleep and wear interval files have their own file inventory at:

`r import_alignment$state_interval_manifest_path`
