---
title: "Preparation 02: Assess coverage and document sample flow"
subtitle: "From aligned one-minute measurements to eligible participant-days"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-02
#| include: false

library(digest)
library(dplyr)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

coverage_manifest <- read_stored_csv(
  "artifacts/12_manifests/coverage_artifacts.csv"
)
assert_columns(
  coverage_manifest,
  c("path", "sha256", "bytes", "artifact_type", "producer"),
  "coverage artifact manifest"
)
manifest_paths <- absolute_path(coverage_manifest$path)
manifest_exists <- file.exists(manifest_paths)
manifest_bytes <- rep(NA_real_, length(manifest_paths))
manifest_sha256 <- rep(NA_character_, length(manifest_paths))
manifest_bytes[manifest_exists] <- file.info(manifest_paths[manifest_exists])$size
manifest_sha256[manifest_exists] <- vapply(
  manifest_paths[manifest_exists],
  digest::digest,
  character(1),
  algo = "sha256",
  file = TRUE
)
manifest_summary <- tibble::tibble(
  `Files listed` = nrow(coverage_manifest),
  `Files present` = sum(manifest_exists),
  `Size checks` = sum(
    manifest_exists & manifest_bytes == coverage_manifest$bytes
  ),
  `SHA-256 checks` = sum(
    manifest_exists & manifest_sha256 == coverage_manifest$sha256
  ),
  Check = ifelse(
    all(
      manifest_exists &
        manifest_bytes == coverage_manifest$bytes &
        manifest_sha256 == coverage_manifest$sha256
    ),
    "PASS",
    "FAIL"
  )
)
stopifnot(manifest_summary$Check == "PASS")

site_registry <- read_stored_csv("config/site_display_registry.csv")
settings <- read_stored_csv("artifacts/03_coverage/coverage_settings.csv")
daily <- dplyr::bind_rows(
  read_stored_csv(
    "artifacts/03_coverage/light_glasses_daily_coverage.csv"
  ),
  read_stored_csv(
    "artifacts/03_coverage/light_chest_daily_coverage.csv"
  )
)
hourly <- dplyr::bind_rows(
  read_stored_csv(
    "artifacts/03_coverage/light_glasses_hourly_coverage.csv"
  ),
  read_stored_csv(
    "artifacts/03_coverage/light_chest_hourly_coverage.csv"
  )
)
gap_runs <- dplyr::bind_rows(
  read_stored_csv("artifacts/03_coverage/light_glasses_gap_runs.csv"),
  read_stored_csv("artifacts/03_coverage/light_chest_gap_runs.csv")
)
sample_flow <- read_stored_csv("artifacts/03_coverage/sample_flow.csv")

assert_columns(
  settings,
  c(
    "placement", "input_sites", "input_participants",
    "input_participant_days", "input_true_utc_minutes",
    "output_true_utc_minutes", "coverage_rule_id", "coverage_signal",
    "expected_wall_minutes_per_hour", "expected_wall_minutes_per_day",
    "minimum_hour_coverage", "minimum_day_coverage", "eligible_hours",
    "ineligible_hours", "eligible_days", "ineligible_days",
    "all_zero_medi_days_excluded", "medi_eligible_true_utc_minutes",
    "light_eligible_true_utc_minutes", "dst_fold_wall_minutes"
  ),
  "coverage settings"
)
assert_columns(
  daily,
  c(
    "site", "position", "local_date", "day_valid_fraction_raw",
    "day_all_zero_medi_excluded", "day_eligible",
    "day_eligible_after_hour", "day_dst_fold_minutes"
  ),
  "daily coverage data"
)
assert_columns(
  hourly,
  c("site", "position", "hour_valid_fraction", "hour_eligible"),
  "hourly coverage data"
)
assert_columns(
  sample_flow,
  c(
    "placement", "scope", "stage_order", "stage", "participants",
    "participant_days", "true_utc_minutes", "wall_minutes",
    "finite_medi_minutes", "finite_light_minutes"
  ),
  "coverage sample flow"
)

stopifnot(
  all(settings$coverage_rule_id == "A"),
  all(settings$coverage_signal == "MEDI"),
  all(settings$minimum_hour_coverage == 0.50),
  all(settings$minimum_day_coverage == 0.80),
  all(settings$expected_wall_minutes_per_hour == 60L),
  all(settings$expected_wall_minutes_per_day == 1440L),
  all(settings$input_true_utc_minutes == settings$output_true_utc_minutes)
)

coverage_rules <- tibble::tribble(
  ~Decision, ~Requirement, ~`Scientific consequence`,
  "Daily coverage",
  "At least 80% of 1,440 local-clock minutes: 1,152 valid melEDI minutes",
  "The participant-day remains available to later metric-specific checks.",
  "Hourly coverage",
  "At least 50% of 60 local-clock minutes: 30 valid melEDI minutes",
  "The hour can support an hourly summary; a failed hour does not erase its observed minutes from daily or minute-based metrics.",
  "Coverage signal",
  "Finite melEDI values",
  "Illuminance retains its own observed and missing values; a missing melEDI value does not remove an observed illuminance value.",
  "Sleep-period denominator",
  "Diary-defined sleep remains in the complete 24-hour measurement cycle",
  "Bedside sleep-environment measurements contribute to coverage.",
  "Day-long flat-zero screen",
  "Exclude an otherwise passing day only when every finite melEDI value is exactly 0 lx",
  "Individual zero readings remain valid; the original day and an inclusive sensitivity flag remain stored.",
  "Autumn clock change",
  "Repeated local-clock minutes are combined only for the coverage decision",
  "Both real observations remain separate in actual time."
)

placement_labels <- c(glasses = "Near-eye", chest = "Chest")
settings_display <- settings |>
  dplyr::mutate(Placement = unname(placement_labels[.data$placement])) |>
  dplyr::select(
    .data$Placement,
    `Study sites` = .data$input_sites,
    `Input participants` = .data$input_participants,
    `Input participant-days` = .data$input_participant_days,
    `Actual-time rows before coverage` = .data$input_true_utc_minutes,
    `Actual-time rows after coverage` = .data$output_true_utc_minutes,
    `Hours passing` = .data$eligible_hours,
    `Hours failing` = .data$ineligible_hours,
    `Days passing` = .data$eligible_days,
    `Days failing` = .data$ineligible_days,
    `All-zero days excluded` = .data$all_zero_medi_days_excluded,
    `Usable melEDI minutes` = .data$medi_eligible_true_utc_minutes,
    `Usable illuminance minutes` = .data$light_eligible_true_utc_minutes,
    `Repeated autumn clock minutes` = .data$dst_fold_wall_minutes
  ) |>
  tidyr::pivot_longer(
    cols = -Placement,
    names_to = "Stored outcome",
    values_to = "Value"
  ) |>
  tidyr::pivot_wider(names_from = Placement, values_from = Value)

site_day_summary <- daily |>
  dplyr::group_by(.data$site, .data$position) |>
  dplyr::summarise(
    `Participant-days` = dplyr::n(),
    `Days passing` = sum(.data$day_eligible),
    `Days failing` = sum(!.data$day_eligible),
    `Passing days (%)` = 100 * mean(.data$day_eligible),
    `All-zero days excluded` = sum(.data$day_all_zero_medi_excluded),
    `Days retained by direct daily rule` = sum(
      .data$day_eligible & !.data$day_eligible_after_hour
    ),
    .groups = "drop"
  ) |>
  dplyr::left_join(site_registry, by = "site", relationship = "many-to-one") |>
  dplyr::arrange(.data$display_order, .data$position) |>
  dplyr::transmute(
    Site = .data$display_name,
    Placement = unname(placement_labels[.data$position]),
    .data$`Participant-days`,
    .data$`Days passing`,
    .data$`Days failing`,
    .data$`Passing days (%)`,
    .data$`All-zero days excluded`,
    .data$`Days retained by direct daily rule`
  )

day_coverage_distribution <- daily |>
  dplyr::group_by(.data$position) |>
  dplyr::summarise(
    Minimum = min(.data$day_valid_fraction_raw),
    `25th percentile` = stats::quantile(
      .data$day_valid_fraction_raw,
      0.25,
      names = FALSE
    ),
    Median = stats::median(.data$day_valid_fraction_raw),
    `75th percentile` = stats::quantile(
      .data$day_valid_fraction_raw,
      0.75,
      names = FALSE
    ),
    Maximum = max(.data$day_valid_fraction_raw),
    .groups = "drop"
  ) |>
  dplyr::mutate(Placement = unname(placement_labels[.data$position])) |>
  dplyr::select(-.data$position) |>
  tidyr::pivot_longer(
    cols = -Placement,
    names_to = "Daily valid-fraction summary",
    values_to = "Fraction"
  ) |>
  tidyr::pivot_wider(names_from = Placement, values_from = Fraction)

gap_labels <- c(
  raw_absence = "No one-minute record",
  signal_invalidity = "Recorded signal unavailable",
  all_zero_medi_day = "All finite melEDI values were 0 lx",
  day_failure = "Day below 80% valid melEDI"
)
signal_labels <- c(MEDI = "melEDI", LIGHT = "Illuminance", row = "Whole record")
gap_summary <- gap_runs |>
  dplyr::group_by(.data$position, .data$gap_reason, .data$signal) |>
  dplyr::summarise(
    `Continuous periods` = dplyr::n(),
    `Participant-days` = dplyr::n_distinct(
      paste(.data$site, .data$Id, .data$local_date)
    ),
    `Local-clock minutes` = sum(.data$wall_minutes),
    `Longest period (minutes)` = max(.data$wall_minutes),
    .groups = "drop"
  ) |>
  dplyr::transmute(
    Placement = unname(placement_labels[.data$position]),
    `Recorded reason` = unname(gap_labels[.data$gap_reason]),
    Signal = unname(signal_labels[.data$signal]),
    .data$`Continuous periods`,
    .data$`Participant-days`,
    .data$`Local-clock minutes`,
    .data$`Longest period (minutes)`
  ) |>
  dplyr::arrange(.data$Placement, .data$`Recorded reason`, .data$Signal)

stage_labels <- c(
  aligned_real_minutes = "All aligned one-minute rows",
  precoverage_medi_finite = "Rows with valid melEDI before coverage",
  precoverage_light_finite = "Rows with valid illuminance before coverage",
  all_zero_medi_inclusive_sensitivity_real_minutes =
    "Rows retained when all-zero days are included",
  primary_day_eligible_real_minutes = "Rows in days that passed",
  medi_eligible_real_minutes = "Usable melEDI rows after coverage",
  light_eligible_real_minutes = "Usable illuminance rows after coverage",
  hourly_metric_eligible_real_minutes =
    "Rows in hours that support hourly summaries",
  hour_screened_sensitivity_real_minutes =
    "Rows retained by the hour-screened comparison"
)
sample_flow_display <- sample_flow |>
  dplyr::filter(.data$scope == "overall") |>
  dplyr::arrange(.data$placement, .data$stage_order) |>
  dplyr::transmute(
    Placement = unname(placement_labels[.data$placement]),
    Stage = unname(stage_labels[.data$stage]),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Actual-time minutes` = .data$true_utc_minutes,
    `Valid melEDI minutes` = .data$finite_medi_minutes,
    `Valid illuminance minutes` = .data$finite_light_minutes
  )

validation_summary <- tibble::tribble(
  ~Check, ~`Stored outcome`, ~Evidence,
  "Actual-time row preservation",
  paste0(
    format(sum(settings$input_true_utc_minutes), big.mark = ","),
    " rows entered and the same number left the coverage step"
  ),
  "artifacts/03_coverage/coverage_settings.csv",
  "Complete daily grid",
  paste0(
    nrow(daily),
    " participant-days each used a fixed 1,440-minute local-clock denominator"
  ),
  "artifacts/03_coverage/light_*_daily_coverage.csv",
  "Complete hourly grid",
  paste0(
    nrow(hourly),
    " hourly decisions were stored (24 per participant-day)"
  ),
  "artifacts/03_coverage/light_*_hourly_coverage.csv",
  "Direct daily decision",
  paste0(
    sum(daily$day_eligible & !daily$day_eligible_after_hour),
    " days passed directly but would have failed after first removing sub-50% hours"
  ),
  "artifacts/03_coverage/light_*_daily_coverage.csv",
  "Day-long flat-zero screen",
  paste0(
    sum(daily$day_all_zero_medi_excluded),
    " otherwise passing days were excluded and remain explicitly flagged"
  ),
  "artifacts/03_coverage/light_*_daily_coverage.csv",
  "Accepted artifact identities",
  paste0(
    manifest_summary$`Files listed`,
    " of ",
    manifest_summary$`Files listed`,
    " files passed existence, size, and SHA-256 checks"
  ),
  "artifacts/12_manifests/coverage_artifacts.csv"
)

script_map <- tibble::tribble(
  ~Step, ~Producing_code, ~Reads_operation_and_output,
  1L,
  "scripts/pipeline/build_coverage_sample_flow.R",
  "Reads the two aligned one-minute files from Preparation 01 and confirms their recorded SHA-256 identities. It coordinates placements separately so near-eye and chest measurements cannot be pooled. **Outputs:** placement-specific coverage files and shared audit tables under `artifacts/03_coverage/`.",
  2L,
  "scripts/pipeline/aggregation_coverage.R",
  "Constructs a fixed 1,440-minute local-clock grid for each observed participant-date, combines repeated autumn wall-clock minutes only for the coverage decision, and evaluates the 50%-per-hour and 80%-per-day rules. This is separate from actual-time storage so both real observations in a repeated hour remain intact. **Outputs:** hourly and daily coverage tables plus coverage flags mapped back to every actual-time row.",
  3L,
  "scripts/pipeline/build_coverage_sample_flow.R",
  "Applies the day-long flat-zero screen, keeps separate melEDI and illuminance availability, records continuous missing-data periods and their consequences, and counts participants, participant-days, and minutes at each branch. **Outputs:** `sample_flow.csv`, placement gap-run tables, and `coverage_settings.csv`.",
  4L,
  "scripts/pipeline/verify_coverage_artifacts.R",
  "Reads completed outputs and checks input fingerprint propagation, row preservation, fixed denominators, thresholds, all-zero flags, actual-time ordering, coverage masks, sample flow, and manifest identities. It is separate from production and writes no scientific artifact."
) |>
  dplyr::mutate(
    `Step and producing code` = paste0(
      "**", .data$Step, "**<br>",
      wrap_provenance_path(.data$Producing_code)
    )
  ) |>
  dplyr::select(
    .data$`Step and producing code`,
    `Reads, operation, separation, and output` =
      .data$Reads_operation_and_output
  )

handoff_artifacts <- tibble::tribble(
  ~Stored_artifact, ~Exact_path, ~Next_use,
  "Near-eye coverage-annotated actual-time rows",
  "artifacts/03_coverage/light_glasses_coverage.rds",
  "Reference-profile learning and metric derivation for the primary placement",
  "Chest coverage-annotated actual-time rows",
  "artifacts/03_coverage/light_chest_coverage.rds",
  "Complementary-placement reference profiles and metrics",
  "Hourly coverage decisions",
  "artifacts/03_coverage/light_{glasses,chest}_hourly_coverage.csv",
  "Determine which hours can support hourly summaries",
  "Daily coverage decisions",
  "artifacts/03_coverage/light_{glasses,chest}_daily_coverage.csv",
  "Identify participant-days entering later metric-specific checks",
  "Reason-coded missing-data periods",
  "artifacts/03_coverage/light_{glasses,chest}_gap_runs.csv",
  "Describe the timing and duration of missing or unusable measurements",
  "Sample flow and settings",
  "artifacts/03_coverage/sample_flow.csv and coverage_settings.csv",
  "Audit denominators, thresholds, retained samples, and sensitivity flags",
  "Artifact inventory",
  "artifacts/12_manifests/coverage_artifacts.csv",
  "Confirm exact file identities passed forward"
)

environment_summary <- tibble::tibble(
  Component = c(
    "R",
    "Quarto profile",
    "Table package",
    "Identity check",
    "Scientific recomputation during render"
  ),
  Value = c(
    paste(R.version$major, R.version$minor, sep = "."),
    "nathealth (HTML review render)",
    paste0("gt ", as.character(utils::packageVersion("gt"))),
    "File size and SHA-256 with digest",
    "None"
  )
)
```

## Purpose and position in the analysis chain

Preparation 02 evaluates whether the aligned one-minute records from
Preparation 01 contain enough valid melEDI to represent an hour or a complete
participant-day. It reads the stored near-eye and chest files separately,
constructs a complete local-clock grid for every participant-date present in
those files, and records hourly and daily coverage decisions. The accepted
outputs enter Preparation 03 for reference-profile learning and Preparation
04 for metric derivation.

The primary dataset uses the full 1,440-minute hybrid measurement day,
including bedside sleep-environment measurements. A day passes when at least
80% of those minutes have valid melEDI. The 50% hourly rule is applied only
when an hourly summary is needed; a failed hour does not erase its observed
minutes from the daily decision or from minute-based metrics.

The separately predefined **gap-timing-unaware dataset** still passed the
general 50%-per-hour and 80%-per-day coverage rules. Its name means that the
timing of remaining missing observations was not used for an additional
metric-specific adjustment. For contrast, the primary dataset could be
interpreted as a “time-sensitive primary metric dataset.” This report presents
the general coverage step for the primary dataset; the separate dataset is
constructed and identified later in the preparation chain.

::: {.callout-note appearance="simple" title="What is calculated during this render?"}
This report reads the stored coverage manifest, settings, hourly decisions,
daily decisions, reason-coded missing-data periods, and sample-flow table.
During rendering it checks file identities and required columns and calculates
only compact descriptive summaries of those stored rows. It does not construct
a participant-day grid, reassess an hour or day, change a coverage flag,
rebuild a coverage file, derive a metric, or run the production verifier.
:::

```{mermaid}
%%| label: fig-preparation02-chain
%%| fig-cap: "Position of Preparation 02 in the data-to-analysis chain. Rendering reads the stored coverage outputs but does not recreate the grid or decisions."
%%| fig-alt: "Flow diagram from aligned near-eye and chest one-minute records through a complete local-clock grid, hourly and daily melEDI coverage decisions, a day-long flat-zero check, and stored coverage and sample-flow files passed to reference-profile and metric preparation."
flowchart LR
  A["Aligned actual-time minutes<br/>Preparation 01"] --> B["Complete local-clock grid<br/>for each observed date"]
  B --> C["Hourly support<br/>30 of 60 valid minutes"]
  B --> D["Daily support<br/>1,152 of 1,440 valid minutes"]
  D --> E["Exclude otherwise passing<br/>day-long flat-zero signal"]
  C --> F["Stored hourly decisions"]
  E --> G["Stored day flags, gap periods,<br/>and sample flow"]
  F --> H["Preparation 03 profiles<br/>Preparation 04 metrics"]
  G --> H
```

## 1. The decisions applied

Coverage uses melEDI as the common availability signal. Illuminance retains
its own missingness, so an unavailable melEDI value does not erase an observed
illuminance value. Table @tbl-preparation02-rules separates the hourly and
daily decisions and shows when each affects later use.

```{r}
#| label: tbl-preparation02-rules
#| tbl-cap: "Coverage and plausibility decisions recorded in Preparation 02."

coverage_rules |>
  prep_gt() |>
  gt::cols_width(
    Decision ~ gt::pct(20),
    Requirement ~ gt::pct(36),
    `Scientific consequence` ~ gt::pct(44)
  )
```

The daily denominator differs from the preregistered sleep-excluded
denominator. Sleep-excluded rules remain prespecified sensitivities; they are
not substituted for the primary decision on this page. Preparation 04 later
checks whether the timing and amount of remaining support are adequate for
each specific metric.

## 2. Overall input and retained coverage

No observation ordered in actual time is deleted. Coverage decisions are
stored as flags and eligible signal columns. The aligned values, raw values,
time variables, and sleep/wear provenance remain available in the placement
files.

```{r}
#| label: tbl-preparation02-overall
#| tbl-cap: "Overall inputs and stored coverage outcomes by measurement placement."

settings_display |>
  prep_gt() |>
  gt::fmt_integer(columns = c(`Near-eye`, Chest), use_seps = TRUE) |>
  gt::cols_width(
    `Stored outcome` ~ gt::pct(60),
    `Near-eye` ~ gt::pct(20),
    Chest ~ gt::pct(20)
  )
```

The primary day rule retained 816 of 1,136 near-eye participant-days and 902
of 1,249 chest participant-days. The corresponding eligible files contain
1,131,060 valid near-eye melEDI minutes and 1,253,505 valid chest melEDI
minutes. Near-eye and chest measurements remain separate throughout.

## 3. Coverage by site

Table @tbl-preparation02-site-days shows both the numerator and denominator at
each site. “Days retained by direct daily rule” are days that reached 80%
coverage across the day but would have failed if sub-50% hours had first been
discarded. The direct daily rule retained seven such near-eye days and eight
such chest days.

```{r}
#| label: tbl-preparation02-site-days
#| tbl-cap: "Daily coverage outcomes by site and measurement placement."

site_day_summary |>
  prep_gt() |>
  gt::fmt_integer(
    columns = c(
      `Participant-days`,
      `Days passing`,
      `Days failing`,
      `All-zero days excluded`,
      `Days retained by direct daily rule`
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_number(columns = `Passing days (%)`, decimals = 1) |>
  gt::cols_label(
    `Participant-days` = "Days",
    `Days passing` = "Pass",
    `Days failing` = "Fail",
    `Passing days (%)` = "Pass (%)",
    `All-zero days excluded` = "Flat-zero excluded",
    `Days retained by direct daily rule` = "Direct-rule retained"
  ) |>
  gt::cols_width(
    Site ~ gt::pct(19),
    Placement ~ gt::pct(12),
    gt::everything() ~ gt::pct(13.8)
  )
```

Daily valid melEDI fractions were not concentrated only near the 80% boundary.
Their stored distribution ranged from near-zero coverage to complete coverage,
with medians of 94.7% for near-eye days and 95.8% for chest days.

```{r}
#| label: tbl-preparation02-day-distribution
#| tbl-cap: "Distribution of the valid melEDI fraction across all participant-days before applying the 80% decision."

day_coverage_distribution |>
  prep_gt() |>
  gt::fmt_percent(columns = c(`Near-eye`, Chest), decimals = 1) |>
  gt::cols_width(
    `Daily valid-fraction summary` ~ gt::pct(60),
    `Near-eye` ~ gt::pct(20),
    Chest ~ gt::pct(20)
  )
```

## 4. Why measurements were unavailable

The stored missing-data records distinguish a missing one-minute record, an
observed record whose signal was unusable, an otherwise passing day with only
zero melEDI values, and a day below 80% valid melEDI. These categories can
overlap: missing readings are the underlying event, while hourly or daily
failure is a consequence for a later summary.

```{r}
#| label: tbl-preparation02-gaps
#| tbl-cap: "Continuous missing or unusable periods and daily coverage consequences."

gap_summary |>
  prep_gt(
    source_note = paste0(
      "An hour below 50% is not labelled as a gap because its observed ",
      "one-minute values remain available for daily and minute-based metrics."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Continuous periods`,
      `Participant-days`,
      `Local-clock minutes`,
      `Longest period (minutes)`
    ),
    use_seps = TRUE
  ) |>
  gt::cols_label(
    `Continuous periods` = "Periods",
    `Participant-days` = "Days",
    `Local-clock minutes` = "Minutes",
    `Longest period (minutes)` = "Longest (min)"
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(12),
    `Recorded reason` ~ gt::pct(31),
    Signal ~ gt::pct(13),
    gt::everything() ~ gt::pct(11)
  )
```

Five otherwise passing days were excluded by the flat-zero screen: two
near-eye and two chest days at Izmir, plus one chest day at Delft. Individual
zero readings remain valid measurements, and the inclusive sensitivity flag
remains stored for those days.

## 5. Sample flow through each branch

The full sample-flow file reports each stage both overall and by site. Table
@tbl-preparation02-sample-flow shows the overall branches without treating
actual-time rows, local-clock rows, and valid signal minutes as interchangeable
denominators.

```{r}
#| label: tbl-preparation02-sample-flow
#| tbl-cap: "Participants, participant-days, actual-time minutes, and valid signal minutes at each coverage branch."

sample_flow_display |>
  prep_gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(
      Participants,
      `Participant-days`,
      `Actual-time minutes`,
      `Valid melEDI minutes`,
      `Valid illuminance minutes`
    ),
    use_seps = TRUE
  ) |>
  gt::cols_label(
    `Participant-days` = "Days",
    `Actual-time minutes` = "Actual-time min",
    `Valid melEDI minutes` = "Valid melEDI",
    `Valid illuminance minutes` = "Valid illuminance"
  ) |>
  gt::cols_width(
    Stage ~ gt::pct(42),
    gt::everything() ~ gt::pct(14.5)
  )
```

The all-zero-inclusive and hour-screened rows are stored comparisons, not the
primary dataset. They do not replace the accepted direct daily rule.

## 6. Checks that passed

```{r}
#| label: tbl-preparation02-validation
#| tbl-cap: "Validation outcomes recorded for the accepted Preparation 02 outputs."

validation_summary |>
  dplyr::mutate(Evidence = wrap_provenance_path(.data$Evidence)) |>
  prep_gt() |>
  gt::fmt_markdown(columns = Evidence) |>
  gt::cols_width(
    Check ~ gt::pct(23),
    `Stored outcome` ~ gt::pct(52),
    Evidence ~ gt::pct(25)
  )
```

```{r}
#| label: tbl-preparation02-manifest
#| tbl-cap: "File-identity checks performed during this render for accepted Preparation 02 artifacts."

manifest_summary |>
  prep_gt(
    source_note = paste0(
      "Every listed file was hashed during render. No manifest or scientific ",
      "artifact was written."
    )
  ) |>
  gt::cols_width(gt::everything() ~ gt::pct(20))
```

## Production scripts and execution order

```{r}
#| label: tbl-preparation02-script-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

script_map |>
  prep_gt() |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

## Exact artifacts passed forward

```{r}
#| label: tbl-preparation02-handoff
#| tbl-cap: "Stored coverage and sample-flow artifacts passed to later preparation steps."

handoff_artifacts |>
  dplyr::mutate(Exact_path = wrap_provenance_path(.data$Exact_path)) |>
  prep_gt() |>
  gt::fmt_markdown(columns = Exact_path) |>
  gt::cols_label(
    Stored_artifact = "Stored artifact",
    Exact_path = "Exact path",
    Next_use = "Next use"
  ) |>
  gt::cols_width(
    Stored_artifact ~ gt::pct(28),
    Exact_path ~ gt::pct(34),
    Next_use ~ gt::pct(38)
  )
```

## Render environment

```{r}
#| label: tbl-preparation02-environment
#| tbl-cap: "Environment used for this documentation-only HTML render."

environment_summary |>
  prep_gt() |>
  gt::cols_width(
    Component ~ gt::pct(38),
    Value ~ gt::pct(62)
  )
```
