---
title: "Preparation 02: Apply coverage rules and count retained data"
execute:
  echo: true
  warning: true
  message: true
  error: false
  freeze: false
---

## Purpose

This notebook applies the approved primary coverage rule A to the one-minute data
from Preparation 01. An hour passes when at least 50% of its 60 local-clock
minutes have valid melanopic equivalent daylight illuminance (melEDI) data.
After this hourly rule has been applied, a day passes when valid melEDI covers
at least 80% of the full 1,440-minute measurement day. The shared local-clock
representation means that, for example, 08:00 denotes 08:00 local time at
every site.

The full-day decision intentionally includes bedside measurements during the
sleep-diary window, together with near-eye or chest measurements while awake.
This differs from the preregistered denominator, which excluded sleep. Rules B
and C exclude diary-defined sleep at the daily stage or at both the hourly and
daily stages, respectively, and are reserved as prespecified sensitivity
analyses. Preparation 04 later checks whether each metric has enough data at
the times relevant to that metric.

No observation ordered in actual time is deleted. The output keeps the
pre-coverage analysis variables (`MEDI`, `LIGHT`, `MEDI_precoverage`, and
`LIGHT_precoverage`), the raw signals, the actual-time and local-clock
variables, the diary and wear information, and flags showing whether each
hour and day passed. `MEDI_eligible` and `LIGHT_eligible` are set to missing
outside a failed hour or day. Within a passing period, melEDI and illuminance
retain their own missing values, so a missing melEDI value does not erase an
otherwise observed illuminance value.

## What this step uses, produces, and checks

- Coverage is evaluated using melEDI as the approved common coverage
  signal.
- Primary coverage rule A uses the full local-clock day. Explicit diary sleep
  is not removed from its hourly or daily denominator.
- Each local-clock hour has a fixed denominator of 60 minutes and
  passes at coverage \(\geq 0.50\).
- The hourly decision is applied before the local-clock day is compared with
  its fixed denominator of 1,440 minutes; a day passes at coverage
  \(\geq 0.80\).
- During the autumn daylight-saving change, a repeated local-clock minute is
  averaged only for the coverage decision. Both observations remain separate
  in the data ordered by actual time.
- Participant-days are defined by dates present in the aligned input. Missing
  minutes within those dates are represented explicitly in the 1,440-minute
  local-clock grid. A completely absent calendar date is not created without
  an external record that the participant-day existed.
- A day that passes remains available to all later metrics. Too little data at
  relevant times can make one metric unavailable without discarding the day
  from every other metric.

## Setup

```{r}
#| label: setup-coverage

root_hint <- Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd())
source(file.path(root_hint, "scripts/pipeline/paths_io.R"))
root <- project_root(root_hint)

dependencies <- c(
  "assertions.R",
  "aggregation_coverage.R",
  "build_coverage_sample_flow.R"
)
for (dependency in dependencies) {
  source(file.path(root, "scripts", "pipeline", dependency))
}

run_label <- Sys.getenv(
  "NATHEALTH_COVERAGE_RUN_LABEL",
  unset = "full"
)
aligned_root_argument <- Sys.getenv(
  "NATHEALTH_ALIGNED_RUN_ROOT",
  unset = ""
)
coverage_root_argument <- Sys.getenv(
  "NATHEALTH_COVERAGE_RUN_ROOT",
  unset = ""
)
placement_argument <- Sys.getenv(
  "NATHEALTH_COVERAGE_PLACEMENTS",
  unset = ""
)
placements <- if (nzchar(placement_argument)) {
  strsplit(placement_argument, ",", fixed = TRUE)[[1L]]
} else {
  NULL
}
```

For a quick test run, set `NATHEALTH_COVERAGE_RUN_LABEL` to the Preparation 01
subset label. By default, the test reads
`artifacts/02_aligned/runs/<label>/` and writes only to
`artifacts/03_coverage/runs/<label>/`. The two explicit run-root environment
variables can instead point to isolated input and output directories. A
test run therefore cannot overwrite the full files used in the main analysis.

## Apply the coverage rules and prepare sample-flow counts

```{r}
#| label: build-coverage-sample-flow

coverage_build <- build_coverage_sample_flow(
  root = root,
  run_label = run_label,
  aligned_run_root = if (nzchar(aligned_root_argument)) {
    aligned_root_argument
  } else {
    NULL
  },
  coverage_run_root = if (nzchar(coverage_root_argument)) {
    coverage_root_argument
  } else {
    NULL
  },
  placements = placements,
  minimum_hour_coverage = 0.50,
  minimum_day_coverage = 0.80
)
```

The run label is `r coverage_build$run_label`. Its input root is
`r coverage_build$aligned_run_root`, and its output root is
`r coverage_build$coverage_run_root`.

## Coverage settings and row-count check

```{r}
#| label: tbl-coverage-settings
#| tbl-cap: "Coverage rules and check that no actual-time rows were lost."

coverage_rule_summary <- coverage_build$settings |>
  dplyr::transmute(
    Placement = .data$placement,
    Rule = .data$coverage_rule_id,
    `Coverage signal` = dplyr::if_else(
      .data$coverage_signal == "MEDI",
      "melEDI",
      .data$coverage_signal
    ),
    Sites = .data$input_sites,
    Participants = .data$input_participants,
    `Participant-days` = .data$input_participant_days,
    `Hourly requirement` = paste0(
      round(100 * .data$minimum_hour_coverage),
      "% of ",
      .data$expected_wall_minutes_per_hour,
      " minutes"
    ),
    `Daily requirement` = paste0(
      round(100 * .data$minimum_day_coverage),
      "% of ",
      .data$expected_wall_minutes_per_day,
      " minutes"
    ),
    `Diary-defined sleep included` = dplyr::if_else(
      .data$diary_sleep_excluded_from_denominator,
      "No",
      "Yes"
    )
  )
knitr::kable(coverage_rule_summary)

coverage_row_check <- coverage_build$settings |>
  dplyr::transmute(
    Placement = .data$placement,
    `Rows before` = .data$input_true_utc_minutes,
    `Rows after` = .data$output_true_utc_minutes,
    `Rows lost` = (
      .data$input_true_utc_minutes - .data$output_true_utc_minutes
    ),
    `Hours passing` = .data$eligible_hours,
    `Hours failing` = .data$ineligible_hours,
    `Days passing` = .data$eligible_days,
    `Days failing` = .data$ineligible_days,
    `Minutes with usable melEDI` = .data$medi_eligible_true_utc_minutes,
    `Minutes with usable illuminance` =
      .data$light_eligible_true_utc_minutes
  )
knitr::kable(
  coverage_row_check,
  caption = "Row preservation and retained hours, days, and signals."
)
```

The number of rows ordered in actual time must agree before and after this
step for every placement. The exact Preparation 01 input files and their
SHA-256 digital fingerprints are retained in the output-file inventory.

## Days retained by site and sensor placement

```{r}
#| label: tbl-coverage-day-summary
#| tbl-cap: "Participant-days passing the coverage rule, by site and sensor placement."

day_summary <- dplyr::bind_rows(coverage_build$daily) |>
  dplyr::group_by(.data$placement, .data$site) |>
  dplyr::summarise(
    participant_days = dplyr::n(),
    eligible_days = sum(.data$day_eligible),
    ineligible_days = sum(!.data$day_eligible),
    median_valid_fraction_after_hour = stats::median(
      .data$day_valid_fraction_after_hour
    ),
    minimum_valid_fraction_after_hour = min(
      .data$day_valid_fraction_after_hour
    ),
    maximum_valid_fraction_after_hour = max(
      .data$day_valid_fraction_after_hour
    ),
    dst_fold_wall_minutes = sum(.data$day_dst_fold_minutes),
    .groups = "drop"
  )
knitr::kable(
  day_summary,
  col.names = c(
    "Placement",
    "Site",
    "Participant-days",
    "Days passing",
    "Days failing",
    "Median valid fraction",
    "Minimum valid fraction",
    "Maximum valid fraction",
    "Repeated autumn clock minutes"
  )
)
```

The downloadable hourly and daily coverage tables retain the number of
expected and observed minutes, the observed fraction, and whether the
corresponding hour or day passed.

## Why data are missing

```{r}
#| label: tbl-coverage-gap-summary
#| tbl-cap: "Continuous missing-data periods and the hourly or daily failures they caused."

gap_summary <- dplyr::bind_rows(coverage_build$gap_runs) |>
  dplyr::group_by(
    .data$position,
    .data$gap_reason,
    .data$reason_class,
    .data$signal
  ) |>
  dplyr::summarise(
    runs = dplyr::n(),
    participant_days = dplyr::n_distinct(
      paste(.data$site, .data$Id, .data$local_date)
    ),
    wall_minutes = sum(.data$wall_minutes),
    maximum_run_minutes = max(.data$wall_minutes),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    position = dplyr::recode(
      .data$position,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    gap_reason = dplyr::recode(
      .data$gap_reason,
      raw_absence = "No one-minute record",
      signal_invalidity = "Recorded signal could not be used",
      hour_failure = "Hour had less than 50% valid melEDI",
      day_failure = "Day had less than 80% valid melEDI"
    ),
    reason_class = dplyr::recode(
      .data$reason_class,
      availability = "Missing record",
      measurement = "Invalid measurement",
      coverage = "Coverage rule"
    ),
    signal = dplyr::recode(
      .data$signal,
      MEDI = "melEDI",
      LIGHT = "Illuminance",
      row = "Whole record"
    )
  )
knitr::kable(
  gap_summary,
  col.names = c(
    "Placement",
    "Recorded reason",
    "Reason type",
    "Signal",
    "Continuous periods",
    "Participant-days",
    "Local-clock minutes",
    "Longest period (minutes)"
  )
)
```

The recorded reasons distinguish a missing one-minute record, an observed
record whose light signal could not be used, an hour that failed the 50%
requirement, and a day that failed the 80% requirement after the hourly rule
was applied. The matching internal reason codes remain in the downloadable
data.

These reasons can overlap. For example, missing readings can cause both an
hourly and a daily failure. Keeping each reason visible shows both the original
missing data and its later consequence for coverage.

## Sample counts after each coverage step

```{r}
#| label: tbl-coverage-sample-flow
#| tbl-cap: "Sample counts after each coverage step, by sensor placement."

sample_flow_summary <- coverage_build$sample_flow |>
  dplyr::filter(.data$scope == "overall") |>
  dplyr::mutate(
    stage_label = dplyr::recode(
      .data$stage,
      aligned_real_minutes = "All aligned one-minute rows",
      precoverage_medi_finite = "Rows with valid melEDI before coverage",
      precoverage_light_finite =
        "Rows with valid illuminance before coverage",
      hour_eligible_real_minutes = "Rows in hours that passed",
      day_eligible_real_minutes = "Rows in days that passed",
      hour_and_day_eligible_real_minutes =
        "Rows in hours and days that both passed",
      medi_eligible_real_minutes = "Usable melEDI rows after coverage",
      light_eligible_real_minutes =
        "Usable illuminance rows after coverage"
    )
  ) |>
  dplyr::select(
    .data$placement,
    .data$stage_label,
    .data$participants,
    .data$participant_days,
    .data$true_utc_minutes,
    .data$wall_minutes,
    .data$finite_medi_minutes,
    .data$finite_light_minutes
  )
knitr::kable(
  sample_flow_summary,
  col.names = c(
    "Placement",
    "Preparation stage",
    "Participants",
    "Participant-days",
    "Actual-time minutes",
    "Local-clock minutes",
    "Minutes with valid melEDI",
    "Minutes with valid illuminance"
  )
)
```

The downloadable sample-flow table contains the same stages by site and
overall. Participants are identified by both site and participant label, so
the same label used at different sites cannot be mistaken for one person.

## Files produced

For each sensor placement, Preparation 02 writes one R data file containing
all rows in actual-time order, the unmodified and analysis light signals,
the hourly and daily coverage decisions, and the reasons for missing values.
It also writes downloadable tables for hourly coverage, daily coverage,
continuous missing-data periods, sample flow, and settings.

The output-file inventory includes a SHA-256 digital fingerprint for every
file and is stored at:

`r coverage_build$artifact_manifest_path`
