---
title: "Preparation 03: Learn fixed time-of-day reference profiles"
subtitle: "From eligible one-minute records to metric-specific temporal-support maps"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-03
#| include: false

library(digest)
library(dplyr)
library(ggplot2)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

wrap_hash <- function(value) {
  gsub("(.{8})", "\\1\u200b", value, perl = TRUE)
}

profile_manifest_path <- absolute_path(
  "artifacts/12_manifests/reference_profile_artifacts.csv"
)
profile_manifest <- read_stored_csv(
  "artifacts/12_manifests/reference_profile_artifacts.csv"
)
assert_columns(
  profile_manifest,
  c("path", "sha256", "bytes", "artifact_type", "producer", "rows"),
  "reference-profile artifact manifest"
)

current_manifest_sha256 <- digest::digest(
  file = profile_manifest_path,
  algo = "sha256"
)
expected_current_manifest_sha256 <-
  "5e08099602682a4b951b0304764f341932ef2f03ccd4c62f32793ea82084e062"
preceding_verified_manifest_sha256 <-
  "c8e02302521360d3a5cb18f49e0a97aed4a1f0ea64343ead68bde274cddce10d"
stopifnot(identical(
  current_manifest_sha256,
  expected_current_manifest_sha256
))

manifest_paths <- absolute_path(profile_manifest$path)
manifest_exists <- file.exists(manifest_paths)
manifest_bytes <- rep(NA_real_, length(manifest_paths))
manifest_sha256 <- rep(NA_character_, length(manifest_paths))
manifest_bytes[manifest_exists] <-
  file.info(manifest_paths[manifest_exists])$size
manifest_sha256[manifest_exists] <- vapply(
  manifest_paths[manifest_exists],
  function(path) digest::digest(file = path, algo = "sha256"),
  character(1)
)

manifest_summary <- tibble::tibble(
  `Files listed` = nrow(profile_manifest),
  `Files present` = sum(manifest_exists),
  `Size checks` = sum(
    manifest_exists & manifest_bytes == profile_manifest$bytes
  ),
  `SHA-256 checks` = sum(
    manifest_exists & manifest_sha256 == profile_manifest$sha256
  ),
  Check = ifelse(
    all(
      manifest_exists &
        manifest_bytes == profile_manifest$bytes &
        manifest_sha256 == profile_manifest$sha256
    ),
    "PASS",
    "FAIL"
  )
)
stopifnot(manifest_summary$Check == "PASS")

settings <- read_stored_csv(
  "artifacts/04_reference_profiles/reference_profile_settings.csv"
)
input_provenance <- read_stored_csv(
  "artifacts/04_reference_profiles/reference_profile_input_provenance.csv"
)
profile_support <- read_stored_csv(
  "artifacts/04_reference_profiles/reference_profile_support.csv"
)
timing_support <- read_stored_csv(
  "artifacts/04_reference_profiles/timing_exceedance_distribution_support.csv"
)
map_support <- read_stored_csv(
  "artifacts/04_reference_profiles/relevance_map_support.csv"
)
reference_profiles <- read_stored_csv(
  "artifacts/04_reference_profiles/reference_profiles.csv"
)
timing_distributions <- read_stored_csv(
  "artifacts/04_reference_profiles/timing_exceedance_distributions.csv"
)
relevance_maps <- read_stored_csv(
  "artifacts/04_reference_profiles/metric_relevance_maps.csv"
)

assert_columns(
  settings,
  c(
    "run_label", "input_coverage_rule_id",
    "input_coverage_settings_sha256", "bin_minutes",
    "participant_balance", "timing_distribution",
    "timing_distribution_threshold_lx",
    "timing_distribution_comparison", "profile_variants",
    "state_domains", "signals", "pooled_full_day_min",
    "leave_one_site_out_full_day_min", "state_specific_min",
    "site_specific_min", "unsupported_bin_rule", "l5_permitted",
    "profile_rows", "timing_distribution_rows", "relevance_map_rows"
  ),
  "reference-profile settings"
)
assert_columns(
  input_provenance,
  c(
    "placement", "signal", "input_path", "input_sha256",
    "input_sites", "input_participants", "input_participant_days",
    "input_true_utc_minutes", "eligible_true_utc_minutes",
    "learning_wall_minutes", "eligible_learning_wall_minutes",
    "repeated_wall_minutes_averaged",
    "repeated_wall_minutes_with_mixed_state",
    "maximum_true_rows_per_wall_minute"
  ),
  "reference-profile input provenance"
)
assert_columns(
  profile_support,
  c(
    "profile_scope", "state_domain", "profile_variant",
    "minimum_required_participants", "clock_bins", "supported_bins",
    "unsupported_bins", "profile_estimable"
  ),
  "reference-profile support"
)
assert_columns(
  timing_support,
  c(
    "profile_scope", "placement", "clock_bins", "supported_bins",
    "unsupported_bins", "nonzero_probability_bins",
    "profile_estimable"
  ),
  "threshold-timing support"
)
assert_columns(
  map_support,
  c(
    "metric_map", "map_application", "signal", "map_estimable",
    "value_correction_allowed", "paired_channel_required"
  ),
  "metric relevance-map support"
)
assert_columns(
  reference_profiles,
  c(
    "placement", "state_domain", "signal", "clock_bin",
    "supported_reference_value", "profile_scope", "profile_supported"
  ),
  "fixed reference profiles"
)

stopifnot(
  nrow(settings) == 1L,
  settings$run_label == "full",
  settings$input_coverage_rule_id == "A",
  settings$bin_minutes == 30L,
  settings$timing_distribution_threshold_lx == 250,
  settings$timing_distribution_comparison == "strict_greater_than",
  !settings$l5_permitted,
  nrow(reference_profiles) == settings$profile_rows,
  nrow(timing_distributions) == settings$timing_distribution_rows,
  nrow(relevance_maps) == settings$relevance_map_rows
)

manifest_csv_rows <- profile_manifest |>
  dplyr::filter(!is.na(.data$rows)) |>
  dplyr::transmute(
    file = basename(.data$path),
    rows = as.integer(.data$rows)
  )
stored_row_counts <- tibble::tibble(
  file = c(
    "reference_profiles.csv",
    "timing_exceedance_distributions.csv",
    "metric_relevance_maps.csv",
    "reference_profile_input_provenance.csv",
    "reference_profile_settings.csv",
    "reference_profile_support.csv",
    "relevance_map_support.csv",
    "timing_exceedance_distribution_support.csv"
  ),
  current_rows = c(
    nrow(reference_profiles),
    nrow(timing_distributions),
    nrow(relevance_maps),
    nrow(input_provenance),
    nrow(settings),
    nrow(profile_support),
    nrow(map_support),
    nrow(timing_support)
  )
) |>
  dplyr::left_join(manifest_csv_rows, by = "file")
stopifnot(all(stored_row_counts$current_rows == stored_row_counts$rows))

verification_decision <- paste(
  readLines(
    absolute_path(
      "audit/decisions/preparation03_version_specific_verification.md"
    ),
    warn = FALSE,
    encoding = "UTF-8"
  ),
  collapse = "\n"
)
preceding_verification_record <- paste(
  readLines(
    absolute_path(
      "audit/findings/threshold_timing_zero_relevance_map.md"
    ),
    warn = FALSE,
    encoding = "UTF-8"
  ),
  collapse = "\n"
)
finding_register <- read_stored_csv("audit/ledgers/finding_register.csv")
assert_columns(
  finding_register,
  c("finding_id", "status", "scope", "decision_id", "resolution"),
  "finding register"
)
open_verification_item <- finding_register |>
  dplyr::filter(.data$finding_id == "FIND-043")
stopifnot(
  nrow(open_verification_item) == 1L,
  open_verification_item$status == "documented_open",
  open_verification_item$decision_id == "PREP-002",
  grepl(
    expected_current_manifest_sha256,
    verification_decision,
    fixed = TRUE
  ),
  grepl(
    preceding_verified_manifest_sha256,
    verification_decision,
    fixed = TRUE
  ),
  grepl("394 of 394", preceding_verification_record, fixed = TRUE)
)

learning_decisions <- tibble::tribble(
  ~Decision, ~`Recorded operation`, ~`Why it is needed`,
  "Eligible input",
  "Use the one-minute near-eye and chest files that passed Preparation 02 Rule A. The complete hybrid 24-hour day includes bedside measurements during diary-defined sleep.",
  "The reference must represent the same eligible measurement days that enter metric derivation.",
  "Time coordinate",
  "Retain every observation in actual time and make a separate local-clock copy for learning 48 fixed 30-minute periods.",
  "Actual time preserves chronology; local clock aligns observations that occurred at the same time of day.",
  "Participant balance",
  "Within each 30-minute period, take each participant's median and then the median across participants.",
  "Participants with more recorded days or minutes cannot dominate the population reference.",
  "Threshold timing",
  "Within every participant-day and period, calculate the fraction of valid melEDI minutes strictly above 250 lx; average days within participant and then participants within profile.",
  "A population median may remain below 250 lx even when individual days contain threshold events.",
  "Profile variants",
  "Store pooled, site-specific, and leave-one-site-out versions for full-day, wake, three hours before sleep, and sleep periods.",
  "Preparation 04 can use the pooled profile and examine prespecified site-related sensitivity inputs without relearning profiles.",
  "Insufficient support",
  "Require 20 participants for pooled and leave-one-site-out full-day profiles and five for site- or diary-period-specific profiles. Retain unsupported periods without interpolation.",
  "Missing support remains explicit and cannot be replaced by an invented time-of-day pattern."
)

stored_settings_display <- tibble::tribble(
  ~Setting, ~`Stored value`,
  "Run", settings$run_label,
  "Coverage rule", settings$input_coverage_rule_id,
  "Time-of-day resolution", paste(settings$bin_minutes, "minutes"),
  "Signals", "melEDI and illuminance",
  "Profile periods", "Full day, wake, three hours before sleep, and sleep",
  "Profile versions", "Pooled, site-specific, and leave-one-site-out",
  "Strict threshold distribution",
  paste0("> ", settings$timing_distribution_threshold_lx, " lx melEDI"),
  "Minimum participants: pooled full day",
  as.character(settings$pooled_full_day_min),
  "Minimum participants: leave one site out, full day",
  as.character(settings$leave_one_site_out_full_day_min),
  "Minimum participants: site or diary period",
  as.character(settings$state_specific_min),
  "Unsupported 30-minute period",
  "Retained without a reference value; no interpolation",
  "Five-hour darkest-window profile",
  "Not produced; the analysis uses L10"
)

placement_labels <- c(glasses = "Near-eye", chest = "Chest")
signal_labels <- c(MEDI = "melEDI", LIGHT = "Illuminance")
input_display <- input_provenance |>
  dplyr::mutate(
    Placement = unname(placement_labels[.data$placement]),
    Signal = unname(signal_labels[.data$signal]),
    `Study coverage` = paste0(
      scales::comma(.data$input_sites), " sites; ",
      scales::comma(.data$input_participants), " participants; ",
      scales::comma(.data$input_participant_days), " participant-days"
    ),
    `Actual-time minute rows` = paste0(
      scales::comma(.data$input_true_utc_minutes), " total; ",
      scales::comma(.data$eligible_true_utc_minutes), " usable for ",
      .data$Signal
    ),
    `Input SHA-256` = wrap_hash(.data$input_sha256)
  ) |>
  dplyr::select(
    .data$Placement,
    .data$Signal,
    .data$`Study coverage`,
    .data$`Actual-time minute rows`,
    .data$`Input SHA-256`
  ) |>
  dplyr::arrange(
    match(.data$Placement, c("Near-eye", "Chest")),
    match(.data$Signal, c("melEDI", "Illuminance"))
  )

dst_display <- input_provenance |>
  dplyr::distinct(
    .data$placement,
    .data$learning_wall_minutes,
    .data$repeated_wall_minutes_averaged,
    .data$repeated_wall_minutes_with_mixed_state,
    .data$maximum_true_rows_per_wall_minute
  ) |>
  dplyr::mutate(
    Placement = unname(placement_labels[.data$placement]),
    `Local-clock learning check` = paste0(
      scales::comma(.data$learning_wall_minutes),
      " local-clock rows; ",
      scales::comma(.data$repeated_wall_minutes_averaged),
      " repeated rows averaged; at most ",
      .data$maximum_true_rows_per_wall_minute,
      " actual observations per local minute."
    ),
    `Diary-period boundary check` = ifelse(
      .data$repeated_wall_minutes_with_mixed_state > 0,
      paste0(
        scales::comma(.data$repeated_wall_minutes_with_mixed_state),
        " repeated rows crossed diary-period labels. The diary period was ",
        "selected in actual time before local-clock averaging."
      ),
      "No repeated row crossed a diary-period label."
    )
  ) |>
  dplyr::select(
    .data$Placement,
    .data$`Local-clock learning check`,
    .data$`Diary-period boundary check`
  ) |>
  dplyr::arrange(match(.data$Placement, c("Near-eye", "Chest")))

scope_labels <- c(
  pooled = "Pooled across sites",
  site_specific = "Each site separately",
  leave_one_site_out = "Leave one site out"
)
period_labels <- c(
  full_day = "Full day",
  wake = "Wake",
  `pre-sleep` = "Three hours before sleep",
  sleep = "Sleep"
)
profile_support_display <- profile_support |>
  dplyr::group_by(.data$profile_scope, .data$state_domain) |>
  dplyr::summarise(
    profiles = dplyr::n(),
    supported = sum(.data$supported_bins),
    unsupported = sum(.data$unsupported_bins),
    estimable = sum(.data$profile_estimable),
    minimum = paste(
      sort(unique(.data$minimum_required_participants)),
      collapse = "/"
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    `Profile and period` = paste0(
      unname(scope_labels[.data$profile_scope]),
      " — ",
      unname(period_labels[.data$state_domain])
    ),
    `Minimum participants` = .data$minimum,
    `Supported half-hour periods` = paste0(
      scales::comma(.data$supported),
      " of ",
      scales::comma(.data$supported + .data$unsupported)
    ),
    `Estimable profiles` = paste0(
      scales::comma(.data$estimable),
      " of ",
      scales::comma(.data$profiles)
    )
  ) |>
  dplyr::arrange(
    match(
      .data$profile_scope,
      c("pooled", "site_specific", "leave_one_site_out")
    ),
    match(
      .data$state_domain,
      c("full_day", "wake", "pre-sleep", "sleep")
    )
  ) |>
  dplyr::select(
    .data$`Profile and period`,
    .data$`Minimum participants`,
    .data$`Supported half-hour periods`,
    .data$`Estimable profiles`
  )

timing_support_display <- timing_support |>
  dplyr::group_by(.data$profile_scope, .data$placement) |>
  dplyr::summarise(
    distributions = dplyr::n(),
    supported = sum(.data$supported_bins),
    unsupported = sum(.data$unsupported_bins),
    nonzero_min = min(.data$nonzero_probability_bins),
    nonzero_max = max(.data$nonzero_probability_bins),
    estimable = sum(.data$profile_estimable),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    `Profile and placement` = paste0(
      unname(scope_labels[.data$profile_scope]),
      " — ",
      unname(placement_labels[.data$placement])
    ),
    `Supported half-hour periods` = paste0(
      scales::comma(.data$supported),
      " of ",
      scales::comma(.data$supported + .data$unsupported)
    ),
    `Periods with non-zero probability per distribution` = ifelse(
      .data$nonzero_min == .data$nonzero_max,
      as.character(.data$nonzero_min),
      paste0(.data$nonzero_min, "–", .data$nonzero_max)
    ),
    `Estimable distributions` = paste0(
      .data$estimable,
      " of ",
      .data$distributions
    )
  ) |>
  dplyr::arrange(
    match(
      .data$profile_scope,
      c("pooled", "site_specific", "leave_one_site_out")
    ),
    match(.data$placement, c("glasses", "chest"))
  ) |>
  dplyr::select(
    .data$`Profile and placement`,
    .data$`Supported half-hour periods`,
    .data$`Periods with non-zero probability per distribution`,
    .data$`Estimable distributions`
  )

map_names <- c(
  dose = "melEDI dose",
  M10 = "Brightest 10 hours (M10)",
  L10 = "Darkest 10 hours (L10)",
  paired_channel_coverage = "Historical paired-channel map",
  timing_above_250 = "Timing above 250 lx melEDI"
)
map_use <- c(
  dose_correction =
    "May estimate additive dose in unobserved supported periods.",
  support_only =
    "Checks whether enough relevant times were observed; does not change the value.",
  paired_coverage_only =
    paste(
      "Retained historical support record; it is not used by the current",
      "MDER method."
    )
)
map_support_display <- map_support |>
  dplyr::group_by(
    .data$metric_map,
    .data$map_application,
    .data$signal
  ) |>
  dplyr::summarise(
    maps = dplyr::n(),
    estimable = sum(.data$map_estimable),
    value_change = all(.data$value_correction_allowed),
    paired = all(.data$paired_channel_required),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    `Metric map` = paste0(
      unname(map_names[.data$metric_map]),
      ifelse(
        .data$metric_map == "paired_channel_coverage",
        paste0(" — ", unname(signal_labels[.data$signal])),
        ""
      )
    ),
    `Use in Preparation 04` = unname(map_use[.data$map_application]),
    `Estimable maps` = paste0(.data$estimable, " of ", .data$maps),
    `May change a metric value` = ifelse(.data$value_change, "Yes", "No")
  ) |>
  dplyr::arrange(
    match(
      .data$metric_map,
      c(
        "dose", "M10", "L10", "timing_above_250",
        "paired_channel_coverage"
      )
    ),
    match(.data$signal, c("MEDI", "LIGHT"))
  ) |>
  dplyr::select(
    .data$`Metric map`,
    .data$`Use in Preparation 04`,
    .data$`Estimable maps`,
    .data$`May change a metric value`
  )

profile_plot_data <- reference_profiles |>
  dplyr::filter(
    .data$profile_scope == "pooled",
    .data$state_domain == "full_day"
  ) |>
  dplyr::mutate(
    Placement = unname(placement_labels[.data$placement]),
    Signal = factor(
      unname(signal_labels[.data$signal]),
      levels = c("melEDI", "Illuminance")
    ),
    `Local clock time` = .data$clock_bin / 60,
    `Reference value` = .data$supported_reference_value
  )
stopifnot(
  nrow(profile_plot_data) == 192L,
  all(profile_plot_data$profile_supported),
  !anyNA(profile_plot_data$`Reference value`)
)

current_validation <- tibble::tribble(
  ~Check, ~`Current evidence`, ~Result,
  "Manifest identity",
  wrap_hash(current_manifest_sha256),
  "IDENTIFIED",
  "Manifest members",
  paste0(
    manifest_summary$`Files present`, " of ",
    manifest_summary$`Files listed`, " present; ",
    manifest_summary$`Size checks`, " size and ",
    manifest_summary$`SHA-256 checks`, " SHA-256 checks match"
  ),
  manifest_summary$Check,
  "Stored table dimensions",
  paste0(
    scales::comma(nrow(reference_profiles)), " profile rows; ",
    scales::comma(nrow(timing_distributions)),
    " threshold-distribution rows; ",
    scales::comma(nrow(relevance_maps)), " relevance-map rows"
  ),
  "PASS",
  "Median-profile support",
  paste0(
    scales::comma(sum(profile_support$profile_estimable)),
    " of ",
    scales::comma(nrow(profile_support)),
    " stored profiles are estimable; unsupported periods remain explicit"
  ),
  "PASS",
  "Strict-threshold timing support",
  paste0(
    sum(timing_support$profile_estimable),
    " of ",
    nrow(timing_support),
    " stored distributions are estimable"
  ),
  "PASS",
  "Metric relevance maps",
  paste0(
    scales::comma(sum(map_support$map_estimable)),
    " of ",
    scales::comma(nrow(map_support)),
    " stored maps are estimable"
  ),
  "PASS",
  "Independent reconstruction of current scientific values",
  "No stored record is tied to the current manifest.",
  "OPEN — FIND-043"
)

script_map <- tibble::tribble(
  ~`Step and producing code`, ~`Reads, operation, separation, and output`,
  paste0(
    "**1 — Validate fixed inputs**<br>",
    wrap_provenance_path(
      "scripts/pipeline/build_reference_profiles.R"
    )
  ),
  paste0(
    "Reads the Rule A coverage settings and the near-eye and chest ",
    "one-minute files. It checks the run, placements, signals, time axes, ",
    "coverage decisions, and input fingerprints before profile learning. ",
    "This input gate is separate so a changed or incompatible Preparation ",
    "02 artifact cannot silently enter the learning functions."
  ),
  paste0(
    "**2 — Build the local-clock learning plane**<br>",
    wrap_provenance_path(
      "scripts/pipeline/build_reference_profiles.R"
    )
  ),
  paste0(
    "Keeps the actual-time records unchanged, selects diary periods in ",
    "actual time, and then averages repeated autumn local-clock values only ",
    "for profile learning. It produces checked training rows and the stored ",
    "input-provenance table."
  ),
  paste0(
    "**3 — Learn profiles and maps**<br>",
    wrap_provenance_path("scripts/pipeline/reference_profiles.R")
  ),
  paste0(
    "Learns participant-balanced melEDI and illuminance profiles, the ",
    "participant-day-then-participant-balanced strict >250 lx timing ",
    "distribution, and the metric-specific relevance maps. Scientific ",
    "formulas are isolated here from file discovery and writing."
  ),
  paste0(
    "**4 — Check support and write artifacts**<br>",
    wrap_provenance_path(
      "scripts/pipeline/build_reference_profiles.R"
    )
  ),
  paste0(
    "Checks minimum-participant rules, completeness, map normalization, and ",
    "whether a map may alter a value. It writes the CSV and RDS artifacts, ",
    "support summaries, settings, provenance, and their manifest."
  ),
  paste0(
    "**5 — Independent reconstruction (separate audit)**<br>",
    wrap_provenance_path(
      "scripts/pipeline/verify_reference_profile_artifacts.R"
    )
  ),
  paste0(
    "Can reconstruct profiles and maps from fixed inputs and compare every ",
    "field and fingerprint. It is deliberately separate from production. ",
    "This report does not execute it; current-manifest reconstruction ",
    "remains open as FIND-043."
  )
)

artifact_handoff <- tibble::tribble(
  ~`Stored artifact`, ~`Scientific contents and next use`,
  wrap_provenance_path(
    "artifacts/04_reference_profiles/reference_profiles.csv"
  ),
  paste0(
    "All pooled, site-specific, and leave-one-site-out melEDI and ",
    "illuminance profiles, including support counts. Preparation 04 reads ",
    "the same content from reference_profiles.rds."
  ),
  wrap_provenance_path(
    paste0(
      "artifacts/04_reference_profiles/",
      "timing_exceedance_distributions.csv"
    )
  ),
  paste0(
    "Strict >250 lx melEDI probability by 30-minute period, balanced first ",
    "across days and then participants. Preparation 04 reads the same ",
    "content from the paired RDS file for timing-support checks."
  ),
  wrap_provenance_path(
    "artifacts/04_reference_profiles/metric_relevance_maps.csv"
  ),
  paste0(
    "Metric-specific temporal weights and permissions for dose, M10, L10, ",
    "and threshold timing. Historical paired-channel MDER maps remain in ",
    "the file for provenance but are not used by the current MDER method."
  ),
  wrap_provenance_path(
    "artifacts/04_reference_profiles/reference_profile_support.csv"
  ),
  "Support and estimability for every intensity-profile group.",
  wrap_provenance_path(
    paste0(
      "artifacts/04_reference_profiles/",
      "timing_exceedance_distribution_support.csv"
    )
  ),
  "Support and estimability for all strict-threshold timing distributions.",
  wrap_provenance_path(
    "artifacts/04_reference_profiles/relevance_map_support.csv"
  ),
  "Support, estimability, and value-change permissions for every relevance map.",
  wrap_provenance_path(
    paste0(
      "artifacts/04_reference_profiles/",
      "reference_profile_input_provenance.csv"
    )
  ),
  "Exact Preparation 02 inputs, fingerprints, dimensions, and clock checks.",
  wrap_provenance_path(
    "artifacts/12_manifests/reference_profile_artifacts.csv"
  ),
  "Eleven output paths with byte sizes and SHA-256 fingerprints."
)

runtime_display <- tibble::tibble(
  Component = c("R", "gt", "dplyr", "ggplot2", "readr"),
  Version = c(
    R.version.string,
    as.character(utils::packageVersion("gt")),
    as.character(utils::packageVersion("dplyr")),
    as.character(utils::packageVersion("ggplot2")),
    as.character(utils::packageVersion("readr"))
  )
)
```

## Purpose and position in the preparation chain

Preparation 03 learned fixed time-of-day reference profiles from the
participant-days retained in [Preparation 02](02_coverage_sample_flow.qmd).
These profiles describe when melEDI and illuminance were typically observed
across the day while giving each participant equal influence. Preparation 04
uses them to judge whether a participant-day was observed at the times that
matter for a particular metric. The profiles do not replace an observed
measurement and, except for the explicitly permitted additive dose
correction, do not change a metric value.

The exact scientific inputs were:

- [near-eye Rule A one-minute records](../../artifacts/03_coverage/light_glasses_coverage.rds);
- [chest Rule A one-minute records](../../artifacts/03_coverage/light_chest_coverage.rds);
- [the stored Rule A settings](../../artifacts/03_coverage/coverage_settings.csv).

The outputs are the fixed intensity profiles, the strict-threshold timing
distribution, metric-specific relevance maps, their support summaries, and
the provenance records under
`artifacts/04_reference_profiles/`. They are passed to
[Preparation 04](04_metric_derivation.qmd).

```{mermaid}
%%| label: fig-reference-profile-chain
%%| fig-cap: "Preparation 03 data flow. The production scripts ran before this report; rendering reads their stored outputs only."
%%| fig-alt: "Flow diagram from near-eye and chest one-minute records that passed Preparation 02 Rule A, through an unchanged actual-time record and a separate local-clock learning copy, into participant-balanced 30-minute intensity profiles and strict greater-than-250-lux timing distributions. These produce metric-specific support maps and stored artifacts used in Preparation 04."

flowchart TB
  A["Preparation 02<br/>Rule A one-minute records"] --> B["Preserve actual time<br/>make local-clock learning copy"]
  B --> C["30-minute participant-balanced<br/>intensity profiles"]
  B --> D["30-minute strict >250 lx<br/>timing distributions"]
  C --> E["Metric-specific<br/>relevance maps"]
  D --> E
  E --> F["Stored profiles, maps,<br/>support, and provenance"]
  F --> G["Preparation 04<br/>metric derivation"]
```

## How the reference profiles are defined

The profiles separate time alignment, participant weighting, and support
requirements so that each scientific decision is explicit.
@tbl-learning-decisions states the rules in the order in which they affected
the data.

```{r}
#| label: tbl-learning-decisions
#| tbl-cap: "Scientific decisions used to learn the fixed reference profiles."

learning_decisions |>
  prep_gt(
    source_note = paste0(
      "These are stored production decisions. No profile-learning ",
      "operation is executed during this render."
    )
  ) |>
  gt::cols_width(
    Decision ~ gt::pct(20),
    `Recorded operation` ~ gt::pct(40),
    `Why it is needed` ~ gt::pct(40)
  )
```

The complete full day is the intended hybrid 24-hour measurement cycle,
including bedside measurements during diary-defined sleep. Wake, sleep, and
the three hours before sleep are separate profile domains; they do not redefine
which days passed the general coverage rule.

```{r}
#| label: tbl-stored-profile-settings
#| tbl-cap: "Exact settings recorded with the current Preparation 03 artifacts."

stored_settings_display |>
  prep_gt(
    source_note = paste0(
      "Source: artifacts/04_reference_profiles/",
      "reference_profile_settings.csv."
    )
  ) |>
  gt::cols_width(
    Setting ~ gt::pct(38),
    `Stored value` ~ gt::pct(62)
  )
```

## What the pooled profiles look like

@fig-pooled-reference-profiles shows the pooled full-day profiles that provide
the clearest overview of the learned time-of-day pattern. Each point is one
stored 30-minute reference value. The figure is descriptive: its lines connect
stored values and do not interpolate missing support.

```{r}
#| label: fig-pooled-reference-profiles
#| fig-cap: "Pooled participant-balanced full-day reference profiles for near-eye and chest melEDI and illuminance. Each line connects the 48 stored 30-minute values; all displayed pooled full-day periods met the 20-participant support requirement."
#| fig-alt: "Two vertically stacked line charts show pooled reference values across local clock time. The upper panel shows melEDI and the lower panel illuminance. Each panel has separate near-eye and chest lines, with low values overnight, increasing values during daytime, and declining values toward evening. All 48 half-hour periods are displayed for both placements."
#| fig-width: 7.6
#| fig-height: 5.4
#| fig-align: center
#| fig-format: svg

ggplot(
  profile_plot_data,
  aes(
    x = .data$`Local clock time`,
    y = .data$`Reference value`,
    color = .data$Placement
  )
) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 1.2) +
  facet_wrap(
    vars(.data$Signal),
    ncol = 1,
    scales = "free_y"
  ) +
  scale_color_manual(
    values = c("Near-eye" = "#0072B2", "Chest" = "#D55E00"),
    breaks = c("Near-eye", "Chest")
  ) +
  scale_x_continuous(
    breaks = seq(0, 21, by = 3),
    labels = sprintf("%02d:00", seq(0, 21, by = 3)),
    limits = c(0, 23.5),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    labels = scales::label_number(
      accuracy = 1,
      big.mark = ","
    ),
    expand = expansion(mult = c(0.03, 0.08))
  ) +
  labs(
    x = "Local clock time",
    y = "Reference value (lx)",
    color = "Placement"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text = element_text(size = 9, color = "#333333"),
    axis.title = element_text(size = 10),
    legend.position = "top",
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.minor = element_blank(),
    plot.margin = margin(8, 12, 8, 8)
  ) +
  guides(
    color = guide_legend(
      override.aes = list(linewidth = 1.2, size = 2)
    )
  ) +
  coord_cartesian(clip = "off")
```

[Download the source data for @fig-pooled-reference-profiles](../../artifacts/04_reference_profiles/reference_profiles.csv).

## Exact inputs and clock handling

The near-eye and chest input files retained their actual-time rows and were
identified before profile learning. @tbl-profile-inputs shows the participant
and minute support for each signal. The slight
signal-specific differences in usable minutes reflect the stored missing
values; one signal was not used to erase an observed value in the other.

```{r}
#| label: tbl-profile-inputs
#| tbl-cap: "Exact Rule A inputs recorded for profile learning."

input_display |>
  prep_gt(
    source_note = paste0(
      "Exact paths and unbroken fingerprints are stored in ",
      "artifacts/04_reference_profiles/",
      "reference_profile_input_provenance.csv."
    )
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(10),
    Signal ~ gt::pct(10),
    `Study coverage` ~ gt::pct(23),
    `Actual-time minute rows` ~ gt::pct(27),
    `Input SHA-256` ~ gt::pct(30)
  )
```

Autumn daylight-saving changes can assign the same local clock label to two
different actual observations. Preparation 03 averaged those repeated labels
only in the separate local-clock copy used for learning. It kept both
observations in the fixed actual-time input. For diary-period profiles, the
period was selected in actual time before repeated local labels were averaged,
so a repeated clock label could not move an observation across a sleep or wake
boundary.

```{r}
#| label: tbl-dst-profile-handling
#| tbl-cap: "Checks for repeated autumn local-clock values in the learning copy."

dst_display |>
  prep_gt(
    source_note = paste0(
      "A repeated local-clock value may have at most two actual-time ",
      "observations. Missing values were not replaced by zero."
    )
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(12),
    `Local-clock learning check` ~ gt::pct(43),
    `Diary-period boundary check` ~ gt::pct(45)
  )
```

## Support for the intensity profiles

Each profile retains all 48 half-hour periods. A period without the required
number of participants remains in the stored table with no usable reference
value; it is never interpolated. Consequently, a profile can be estimable
overall while containing explicitly unsupported periods. @tbl-profile-support
summarizes all 288 stored profile groups.

```{r}
#| label: tbl-profile-support
#| tbl-cap: "Stored participant support for the melEDI and illuminance profiles."

profile_support_display |>
  prep_gt(
    source_note = paste0(
      "Counts combine near-eye and chest placement-signal profiles. ",
      "Unsupported half-hour periods remain present without interpolation."
    )
  ) |>
  gt::cols_width(
    `Profile and period` ~ gt::pct(40),
    `Minimum participants` ~ gt::pct(16),
    `Supported half-hour periods` ~ gt::pct(24),
    `Estimable profiles` ~ gt::pct(20)
  )
```

Full-day and pooled wake profiles had support in every stored half-hour
period. Support was sparser within sleep and the three hours before sleep,
especially in individual-site profiles. Five site-specific sleep profiles
were not estimable; the stored diagnostics preserve that limitation rather
than supplying a replacement.

## Probability of exceeding 250 lx by time of day

The timing of strict melEDI values above 250 lx cannot be inferred from whether
the population median profile itself exceeds 250 lx. Individual
participant-days may cross the threshold even when the population median
remains lower. Preparation 03 therefore learned a separate probability for
each half-hour period:

1. within every participant-day, it calculated the fraction of valid minutes
   with melEDI strictly greater than 250 lx;
2. it gave days equal weight within each participant; and
3. it then gave participants equal weight within the pooled, site-specific,
   or leave-one-site-out distribution.

@tbl-threshold-timing-support shows that all 36 stored distributions were
estimable and had positive probability in at least some periods. Preparation
04 expands each half-hour probability across its 30 constituent minutes only
to measure temporal support. It does not weight or change an observed timing.

```{r}
#| label: tbl-threshold-timing-support
#| tbl-cap: "Stored support for strict melEDI >250 lx timing distributions."

timing_support_display |>
  prep_gt(
    source_note = paste0(
      "The primary support cutoff is 0.80; the 0.70 and 0.90 ",
      "sensitivities reuse these same fixed distributions."
    )
  ) |>
  gt::cols_width(
    `Profile and placement` ~ gt::pct(31),
    `Supported half-hour periods` ~ gt::pct(22),
    `Periods with non-zero probability per distribution` ~ gt::pct(29),
    `Estimable distributions` ~ gt::pct(18)
  )
```

## How Preparation 04 may use each map

The intensity and threshold distributions were converted into
metric-specific relevance maps. These maps state which times of day matter
for checking the observation support of a metric. They do not all have
permission to change a value. @tbl-map-use makes that distinction explicit.

```{r}
#| label: tbl-map-use
#| tbl-cap: "Stored relevance maps, their use, and whether they may change a metric value."

map_support_display |>
  prep_gt(
    source_note = paste0(
      "Only the additive missing-dose correction may alter a metric ",
      "value. M10, L10, and threshold-timing maps are support-only; the ",
      "historical paired-channel maps are not used by current MDER."
    )
  ) |>
  gt::cols_width(
    `Metric map` ~ gt::pct(28),
    `Use in Preparation 04` ~ gt::pct(43),
    `Estimable maps` ~ gt::pct(15),
    `May change a metric value` ~ gt::pct(14)
  )
```

The M10 and L10 maps check whether enough relevant times were observed for the
brightest and darkest 10-hour windows. The timing map checks support for the
first, last, and mean timing above 250 lx. None of these active support-only
maps scales a level, duration, or timing, and no L5 profile or map was
produced.

For current MDER, the daily value is the arithmetic mean of viable one-minute
melEDI-to-illuminance ratios. No reference profile weights, scales, or gates
that calculation. The stored paired-channel maps describe an earlier
ratio-of-integrals method and are retained only as historical provenance; they
are not active Preparation 04 inputs for MDER.

## Limitation of current independent verification

Independent reconstruction of the current reference-profile input remains incomplete.

The current files pass the identity and internal-consistency checks described
below, but those checks do not independently reconstruct the scientific
values. This incomplete provenance check is not evidence that a current
profile, metric, or downstream result is incorrect.

## Technical provenance and file checks

The current Preparation 03 artifact inventory is
[`reference_profile_artifacts.csv`](../../artifacts/12_manifests/reference_profile_artifacts.csv).

::: {.callout-note appearance="simple" title="What is calculated during this render?"}
This page reads already stored Preparation 03 artifacts. During rendering, it
checks their paths, byte sizes, SHA-256 fingerprints, required columns, stored
row counts, and recorded support summaries. It also filters the stored pooled
profiles for the displayed figure and calculates compact descriptive totals
from the stored support tables. Rendering does **not** relearn a profile,
rebuild a relevance map, reconstruct scientific values from one-minute data,
write an analytical artifact, or run any Preparation 03 verifier.
:::

The current manifest and every file it lists passed the bounded identity
checks in @tbl-current-validation. Stored dimensions and support fields are
also internally consistent with the current settings. These checks establish
which artifacts this report read; they do not independently reconstruct the
scientific values.

```{r}
#| label: tbl-current-validation
#| tbl-cap: "Checks demonstrably tied to the current Preparation 03 manifest."

current_validation |>
  prep_gt(
    source_note = paste0(
      "Current manifest: artifacts/12_manifests/",
      "reference_profile_artifacts.csv. PASS refers only to the stated ",
      "identity, dimension, or stored-support check."
    )
  ) |>
  gt::fmt_markdown(columns = `Current evidence`) |>
  gt::cols_width(
    Check ~ gt::pct(29),
    `Current evidence` ~ gt::pct(56),
    Result ~ gt::pct(15)
  )
```

### Producing scripts in execution order

@tbl-script-execution-map identifies the code that originally produced and
checked these artifacts. The two-column layout keeps the scientific operation
and its reason together instead of forcing explanatory prose into a narrow
table column.

```{r}
#| label: tbl-script-execution-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

script_map |>
  prep_gt(
    source_note = paste0(
      "Production and independent verification are separate by design. ",
      "The current independent-reconstruction item remains open."
    )
  ) |>
  gt::fmt_markdown(columns = `Step and producing code`) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

### Artifacts passed to metric derivation

Preparation 04 reads the stored RDS forms of the fixed profiles and relevance
maps. The CSV forms provide the paired reader- and machine-readable source
data. @tbl-artifact-handoff lists the exact durable paths.

```{r}
#| label: tbl-artifact-handoff
#| tbl-cap: "Stored Preparation 03 artifacts and their role in later metric derivation."

artifact_handoff |>
  prep_gt(
    source_note = paste0(
      "Every listed file is covered by the current 11-file artifact ",
      "manifest. Rendering writes none of them."
    )
  ) |>
  gt::fmt_markdown(columns = `Stored artifact`) |>
  gt::cols_width(
    `Stored artifact` ~ gt::pct(38),
    `Scientific contents and next use` ~ gt::pct(62)
  )
```

```{r}
#| label: tbl-preparation-03-runtime
#| tbl-cap: "R runtime and packages used for this documentation-only render."

runtime_display |>
  prep_gt(
    source_note = paste0(
      "The production artifacts record R 4.6.1. This page was rendered ",
      "with the project's R 4.6.1 library and nathealth profile."
    )
  ) |>
  gt::cols_width(
    Component ~ gt::pct(38),
    Version ~ gt::pct(62)
  )
```

### Handoff to Preparation 04

Preparation 03 passes forward fixed, versioned temporal reference information:
13,824 intensity-profile rows, 1,728 strict-threshold distribution rows, and
36,288 metric-relevance rows, together with their support diagnostics,
settings, input fingerprints, and artifact manifest. Preparation 04 may read
and apply those stored objects under their recorded permissions. It may not
relearn, interpolate, or silently substitute a profile during rendering.

::: {.callout-note appearance="simple" title="Version-specific verification record"}
The complete independent reconstruction that passed **394 of 394 checks**
applied to the **preceding** manifest with SHA-256
`c8e02302521360d3a5cb18f49e0a97aed4a1f0ea64343ead68bde274cddce10d`.
It reconstructed 288 median-profile groups and 36 strict-threshold
distributions and matched that manifest's 11 file fingerprints. It must not
be treated as verification of the current manifest
`5e08099602682a4b951b0304764f341932ef2f03ccd4c62f32793ea82084e062`.

[PREP-002](../../audit/decisions/preparation03_version_specific_verification.md)
records independent reconstruction of the current manifest as open audit item
**FIND-043**. This report does not run that verifier.
:::
