---
title: "Preparation 03: Learn fixed time-of-day reference profiles"
execute:
  echo: true
  warning: true
  message: true
  error: false
  freeze: false
---

## Purpose

Before participant metrics are calculated, this notebook learns when light
exposure typically occurs across the day. A reference profile summarizes the
median melEDI or illuminance at each time of day while giving every
participant equal influence. A separate profile records the probability that
melEDI is strictly above 250 lx at each time. The profiles are learned once
and are used only to judge whether a participant-day has enough data at
relevant times; they never replace, weight, or change an observed value.

## What this step uses, produces, and checks

- Inputs are the one-minute files that passed Rule A in Preparation 02,
  identified by SHA-256 checksums. A SHA-256 checksum is a digital fingerprint
  used to confirm that a file has not changed. The full-day period is the
  intended hybrid 24-hour cycle, including bedside measurements during
  diary-defined sleep.
- Outputs include the main profiles that pool all sites, profiles calculated
  separately by site, and leave-one-site-out profiles that omit one site at a
  time. The files also include the distributions for timing above 250 lx,
  time-of-day relevance maps, and checks of how much data support them under
  `artifacts/04_reference_profiles/`.
- Profiles use fixed 30-minute periods on the shared local-clock scale and are
  calculated separately by sensor placement, signal, and full-day, wake,
  pre-sleep, or sleep period.
- For each 30-minute period, the intensity profile first takes each
  participant's median and then takes the median across participants. The
  number of contributing participants remains visible.
- For timing above 250 lx, the calculation first finds the fraction of valid
  minutes strictly above 250 lx melEDI within every participant-day and
  30-minute period. Days receive equal weight within a participant, and
  participants then receive equal weight in the profile.
- During the autumn daylight-saving change, repeated local-clock values are
  averaged only while learning the full-day profile. All observations remain
  separate in the fixed input ordered by actual time. For wake, pre-sleep, or
  sleep profiles, the diary period is selected in actual time before repeated
  local-clock values are averaged. This prevents a repeated clock label that
  crosses a diary boundary from being assigned to the wrong period.
- A full-day period in the pooled or leave-one-site-out profile requires at
  least 20 participants. A period in a site-specific or diary-period-specific
  profile requires at least five participants.
- A 30-minute period with too few participants remains visible but has no
  reference value. It is not interpolated and no unweighted replacement is
  used.
- Later notebooks may verify these fixed profiles and expand each 30-minute
  weight to one-minute rows, but may not relearn or interpolate the profiles.
  No L5 profile or relevance map is produced.

## Setup

```{r}
#| label: setup-reference-profiles

root_hint <- Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd())
source(file.path(root_hint, "scripts/pipeline/paths_io.R"))
root <- project_root(root_hint)

dependencies <- c(
  "assertions.R",
  "reference_profiles.R",
  "build_reference_profiles.R"
)
for (dependency in dependencies) {
  source(file.path(root, "scripts", "pipeline", dependency))
}

run_label <- Sys.getenv(
  "NATHEALTH_PROFILE_RUN_LABEL",
  unset = "full"
)
coverage_root_argument <- Sys.getenv(
  "NATHEALTH_COVERAGE_RUN_ROOT",
  unset = ""
)
profile_root_argument <- Sys.getenv(
  "NATHEALTH_PROFILE_RUN_ROOT",
  unset = ""
)
placement_argument <- Sys.getenv(
  "NATHEALTH_PROFILE_PLACEMENTS",
  unset = ""
)
placements <- if (nzchar(placement_argument)) {
  strsplit(placement_argument, ",", fixed = TRUE)[[1L]]
} else {
  NULL
}
```

For a quick test run, set `NATHEALTH_PROFILE_RUN_LABEL` to the Preparation 02
subset label. By default, the test reads
`artifacts/03_coverage/runs/<label>/` and writes only to
`artifacts/04_reference_profiles/runs/<label>/`. Explicit input and output
roots can instead be supplied through `NATHEALTH_COVERAGE_RUN_ROOT` and
`NATHEALTH_PROFILE_RUN_ROOT`. A test run cannot overwrite the full profiles
used in the main analysis. Render this notebook through the project profile, for
example:

```sh
NATHEALTH_PROFILE_RUN_LABEL=subset_baua \
quarto render notebooks/preparation/03_reference_profiles.qmd \
  --profile nathealth --to html
```

## Learn the profiles once

```{r}
#| label: build-reference-profiles

profile_build <- build_reference_profiles(
  root = root,
  run_label = run_label,
  coverage_run_root = if (nzchar(coverage_root_argument)) {
    coverage_root_argument
  } else {
    NULL
  },
  profile_run_root = if (nzchar(profile_root_argument)) {
    profile_root_argument
  } else {
    NULL
  },
  placements = placements,
  bin_minutes = 30L,
  pooled_full_day_min = 20L,
  leave_one_site_out_full_day_min = 20L,
  state_specific_min = 5L,
  site_specific_min = 5L,
  zero_offset = 0.1
)
```

The run label is `r profile_build$run_label`. The exact Preparation 02 inputs,
identified by their digital fingerprints, are under
`r profile_build$coverage_run_root`, and the fixed outputs are under
`r profile_build$profile_run_root`.

## Settings and exact input versions

```{r}
#| label: tbl-reference-profile-settings
#| tbl-cap: "Settings used to learn the fixed reference profiles."

profile_settings <- profile_build$settings
profile_settings_summary <- data.frame(
  Setting = c(
    "Run",
    "Coverage rule used",
    "Period used for daily coverage",
    "Diary-defined sleep excluded",
    "Time-of-day period",
    "Unit given equal influence",
    "Threshold for the timing distribution",
    "Main and sensitivity profile versions",
    "Minimum participants: pooled full day",
    "Minimum participants: leave one site out",
    "Minimum participants: within a diary period",
    "Minimum participants: within a site",
    "Rule for periods with too few participants",
    "Five-hour darkest-window (L5) profile allowed"
  ),
  Value = c(
    profile_settings$run_label,
    profile_settings$input_coverage_rule_id,
    "Full 24-hour hybrid measurement day",
    ifelse(
      profile_settings$input_diary_sleep_excluded_from_denominator,
      "Yes",
      "No"
    ),
    paste(profile_settings$bin_minutes, "minutes"),
    "Participant",
    paste0(
      "> ",
      profile_settings$timing_distribution_threshold_lx,
      " lx melEDI"
    ),
    "All sites; each site; leave one site out",
    profile_settings$pooled_full_day_min,
    profile_settings$leave_one_site_out_full_day_min,
    profile_settings$state_specific_min,
    profile_settings$site_specific_min,
    "Keep the period, but do not assign a reference value",
    ifelse(profile_settings$l5_permitted, "Yes", "No; the analysis uses L10")
  )
)
knitr::kable(profile_settings_summary)
```

```{r}
#| label: tbl-reference-profile-provenance
#| tbl-cap: "Exact input files and checks that actual and local clock times agree."

profile_input_summary <- profile_build$provenance |>
  dplyr::transmute(
    Placement = .data$placement,
    Signal = dplyr::recode(
      .data$signal,
      MEDI = "melEDI",
      LIGHT = "Illuminance"
    ),
    `Study sites` = .data$input_sites,
    Participants = .data$input_participants,
    `Participant-days` = .data$input_participant_days,
    `Actual-time rows` = .data$input_true_utc_minutes,
    `Usable actual-time rows` = .data$eligible_true_utc_minutes,
    `Local-clock rows used for learning` = .data$learning_wall_minutes,
    `Usable local-clock rows` = .data$eligible_learning_wall_minutes,
    `Repeated local-clock rows averaged` =
      .data$repeated_wall_minutes_averaged,
    `Repeated rows spanning diary periods` =
      .data$repeated_wall_minutes_with_mixed_state,
    `Maximum actual rows for one local minute` =
      .data$maximum_true_rows_per_wall_minute,
    `Input digital fingerprint` = .data$input_sha256
  )
knitr::kable(profile_input_summary)
```

The input digital fingerprints identify the complete files ordered in actual
time. Repeated local-clock minutes during the autumn daylight-saving change
are averaged only in the local-clock copy used to learn the profiles. If one
of the repeated observations is missing and the other has a valid value, the
valid value is retained rather than treating the missing value as zero. The
table records repeated clock labels that cross diary periods and checks the
total and valid actual-time rows contributing to every full-day and
diary-period-specific profile.

## How much data supports each profile

```{r}
#| label: tbl-reference-profile-support
#| tbl-cap: "Time periods with enough participants to estimate each main or sensitivity profile."

profile_support_summary <- profile_build$profile_support |>
  dplyr::select(
    profile_scope,
    placement,
    state_domain,
    signal,
    clock_bins,
    supported_bins,
    unsupported_bins,
    minimum_required_participants,
    profile_estimable
  ) |>
  dplyr::group_by(
    .data$profile_scope,
    .data$placement,
    .data$state_domain,
    .data$signal,
    .data$minimum_required_participants
  ) |>
  dplyr::summarise(
    profile_variants = dplyr::n(),
    supported_bins = sum(.data$supported_bins),
    unsupported_bins = sum(.data$unsupported_bins),
    estimable_profiles = sum(.data$profile_estimable),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    profile_scope = dplyr::recode(
      .data$profile_scope,
      pooled = "All sites",
      site_specific = "Each site",
      leave_one_site_out = "Leave one site out"
    ),
    placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    state_domain = dplyr::recode(
      .data$state_domain,
      full_day = "Full 24-hour day",
      wake = "Wake",
      `pre-sleep` = "Three hours before sleep",
      sleep = "Sleep"
    ),
    signal = dplyr::recode(
      .data$signal,
      MEDI = "melEDI",
      LIGHT = "Illuminance"
    )
  )
knitr::kable(
  profile_support_summary,
  col.names = c(
    "Profile type",
    "Placement",
    "Diary period",
    "Signal",
    "Minimum participants",
    "Profile versions",
    "Supported periods",
    "Unsupported periods",
    "Profiles that could be calculated"
  )
)
```

Every downloadable profile table retains all 48 half-hour periods for each
combination of placement, diary period, and light signal. A period with too
few participants remains in the table but has no reference value. The internal
support flag records why it was not used.

## Probability of exceeding 250 lx by time of day

The probability of timing above 250 lx cannot be learned from the amount by
which the median melEDI profile exceeds 250 lx. The population median can
remain below 250 lx even when individual participant-days cross the threshold.
Preparation 03 therefore estimates the probability that melEDI is strictly
above 250 lx at each time of day using the full-day minute records that passed
Rule A.

At each 30-minute period, every valid participant-day has equal weight within
its participant, and every contributing participant then has equal weight in
the profile. The pooled, site-specific, and leave-one-site-out versions are
learned once and use the same minimum participant counts as the corresponding
full-day profiles. The pooled, placement-specific distribution is used in the
main analysis. Site-specific and leave-one-site-out versions are fixed
sensitivity inputs.

Later code gives each of the 30 minutes within a period its learned probability
weight. This map is used only to determine whether enough relevant times were
observed for the first, last, or mean timing above 250 lx. It never weights or
changes an observed timing value. The main support cutoff is 0.80; the 0.70 and
0.90 sensitivities reuse the same learned distribution. If the total learned
probability is zero, timing support cannot be calculated; no uniform or
median-based replacement is used.

## Time-of-day relevance used for each metric

```{r}
#| label: tbl-reference-map-support
#| tbl-cap: "Whether each metric-specific time-of-day profile could be calculated."

map_support_summary <- profile_build$map_support |>
  dplyr::group_by(
    .data$metric_map,
    .data$map_application,
    .data$signal
  ) |>
  dplyr::summarise(
    maps = dplyr::n(),
    estimable_maps = sum(.data$map_estimable),
    minimum_weight_sum = min(.data$relevance_weight_sum),
    maximum_weight_sum = max(.data$relevance_weight_sum),
    value_correction_allowed = all(.data$value_correction_allowed),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    metric_map = dplyr::recode(
      .data$metric_map,
      dose = "melEDI dose",
      M10 = "Brightest 10 hours",
      L10 = "Darkest 10 hours",
      paired_channel_coverage = "MDER",
      timing_above_250 = "Timing above 250 lx melEDI"
    ),
    map_application = dplyr::recode(
      .data$map_application,
      dose_correction = "Estimate missing dose",
      paired_coverage_only =
        "Check paired melEDI and illuminance coverage",
      support_only = "Check whether enough relevant times were observed"
    ),
    signal = dplyr::recode(
      .data$signal,
      MEDI = "melEDI",
      LIGHT = "Illuminance"
    )
  )
knitr::kable(
  map_support_summary,
  col.names = c(
    "Metric profile",
    "How the profile is used",
    "Signal",
    "Profiles",
    "Profiles that could be calculated",
    "Smallest total relevance across the day",
    "Largest total relevance across the day",
    "May change a metric value"
  )
)
```

The totals in the two relevance columns sum the learned importance assigned
across the 48 half-hour periods. They are checks that each fixed profile is
well defined; they are not participant exposure values.

Only the additive correction for missing dose can alter a metric value. The
M10, L10, timing-above-threshold, and paired-signal profiles only check whether
enough relevant times were observed. They do not scale levels, durations,
timings, or ratios.

## Files produced

Preparation 03 writes:

- fixed profiles as `reference_profiles.rds` and
  `reference_profiles.csv`;
- the fixed threshold-timing distribution as
  `timing_exceedance_distributions.rds` and
  `timing_exceedance_distributions.csv`;
- metric-specific maps as `metric_relevance_maps.rds` and
  `metric_relevance_maps.csv`;
- checks showing how much data support each profile, probability
  distribution, and relevance map;
- fixed settings and a record of the exact input versions; and
- an output-file inventory with SHA-256 digital fingerprints.

The file inventory is:

`r profile_build$artifact_manifest_path`

Focused R 4.6.1 tests passed for the probability-above-threshold calculation
and its pooled, site-specific, and leave-one-site-out versions. The complete
independent check also reconstructed all 288 median-profile groups and all 36
probability-above-threshold groups: 394 of 394 checks passed, and all 11 file
fingerprints matched. Preparation 03 is therefore complete; the Preparation 04
metric values and their differences from the submitted analysis remain
pending.
