---
title: "Preparation 04: Derive light-exposure metrics"
subtitle: "From eligible one-minute records to metric-specific analysis inputs"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-04
#| include: false

library(digest)
library(dplyr)
library(ggplot2)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

relative_path <- function(path) {
  sub(paste0("^", root, "/"), "", path)
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE, progress = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

wrap_hash <- function(value) {
  gsub("(.{8})", "\\1\u200b", value, perl = TRUE)
}

placement_label <- function(value) {
  dplyr::recode(value, glasses = "Near-eye", chest = "Chest")
}

format_integer <- function(value) {
  scales::comma(value, accuracy = 1)
}

format_count_fraction <- function(calculated, possible) {
  paste0(
    format_integer(calculated),
    " / ",
    format_integer(possible),
    " (",
    scales::percent(calculated / possible, accuracy = 0.1),
    ")"
  )
}

check_manifest <- function(path, family, expected_sha256) {
  manifest_path <- absolute_path(path)
  manifest <- read_stored_csv(path)
  assert_columns(
    manifest,
    c("path", "sha256", "bytes", "producer", "artifact_type"),
    paste(family, "manifest")
  )

  manifest_sha256 <- digest::digest(
    file = manifest_path,
    algo = "sha256"
  )
  stopifnot(identical(manifest_sha256, expected_sha256))

  listed_paths <- absolute_path(manifest$path)
  present <- file.exists(listed_paths)
  current_bytes <- rep(NA_real_, length(listed_paths))
  current_sha256 <- rep(NA_character_, length(listed_paths))
  current_bytes[present] <- file.info(listed_paths[present])$size
  current_sha256[present] <- vapply(
    listed_paths[present],
    function(file) digest::digest(file = file, algo = "sha256"),
    character(1)
  )

  summary <- tibble::tibble(
    Family = family,
    `Manifest path` = path,
    `Manifest SHA-256` = manifest_sha256,
    `Files listed` = nrow(manifest),
    `Files present` = sum(present),
    `Size checks` = sum(present & current_bytes == manifest$bytes),
    `SHA-256 checks` = sum(present & current_sha256 == manifest$sha256),
    Check = ifelse(
      all(
        present &
          current_bytes == manifest$bytes &
          current_sha256 == manifest$sha256
      ),
      "PASS",
      "FAIL"
    )
  )
  stopifnot(summary$Check == "PASS")

  list(data = manifest, summary = summary)
}

check_file_hash <- function(path, expected_sha256) {
  observed <- digest::digest(
    file = absolute_path(path),
    algo = "sha256"
  )
  stopifnot(identical(observed, expected_sha256))
  observed
}

current_metric_manifest_sha256 <-
  "028bce108339c1277df4a070597430ea20b49c34c2888f4eef42741c1fe76a5e"
mder_audit_metric_manifest_sha256 <-
  "7bb2830c695e5600f0877acca500a187a040fe727a029bdb069f69099abe4f43"
current_mder_audit_sha256 <-
  "5bd34fd4ac26dc2761540393d40e096cde835d8dd4b4edda9253243c9083d4bb"
current_gap_manifest_sha256 <-
  "4ed62fbe58de65a6d05d8cfc6b5d870d7c74a3c83fe89bcd72bd1dd838698935"
current_gap_participant_day_rds_sha256 <-
  "7561b5dd47cb5e23ca94ba59bd57648f11fe492f86af53a566c5c9cf42d932f1"
current_gap_support_rds_sha256 <-
  "a0bc5d7ea2142709412da2253158086416ddb730b0f60ff616bba9d6ccb1edc5"
current_gap_repair_evidence_sha256 <-
  "81e3aa9439b88cf239d6669dc4343cfb146cd2f45790bae412ab837156315018"
current_state_manifest_sha256 <-
  "755508076e42ed74ce0d40f48de192b69ea81b04532db2d5a950c451236e2619"
historical_mder_gate_sha256 <-
  "9ca9b97dac5efd9ec4e7b43a1fb2227b7a13cf98ac4b92318606c4874f9e460b"
mder_decision_sha256 <-
  "1664347de976057807fcb5ac6e24bd4b66af1fe32378fabd1f2acafcfc9266de"
l10_zero_decision_sha256 <-
  "23b9f70d1d16f7fd3ebbdbc57aaf78c0a701fe1f9d22bd667d926cd320d40797"
l10_zero_evidence_sha256 <-
  "a37efd3449a8d1a6065d0eb8964bd6cf8b241f1683a26c065ea1a946e23214fb"

metric_bundle <- check_manifest(
  "artifacts/12_manifests/metric_artifacts.csv",
  "Metric outputs",
  current_metric_manifest_sha256
)
state_bundle <- check_manifest(
  "artifacts/12_manifests/state_support_gate_artifacts.csv",
  "Diary-period support gate",
  current_state_manifest_sha256
)
gap_bundle <- check_manifest(
  "artifacts/12_manifests/manuscript_prepared_data_artifacts.csv",
  "Gap-timing-unaware prepared data",
  current_gap_manifest_sha256
)

metric_manifest <- metric_bundle$data
state_manifest <- state_bundle$data
gap_manifest <- gap_bundle$data

mder_audit_path <- "artifacts/08_diagnostics/mder_METRIC-010/audit_manifest.csv"
mder_audit_manifest <- read_stored_csv(mder_audit_path)
assert_columns(
  mder_audit_manifest,
  c("role", "path", "sha256", "bytes", "decision_id", "r_version"),
  "METRIC-010 MDER audit manifest"
)
stopifnot(
  identical(
    digest::digest(file = absolute_path(mder_audit_path), algo = "sha256"),
    current_mder_audit_sha256
  ),
  all(mder_audit_manifest$decision_id == "METRIC-010"),
  all(mder_audit_manifest$r_version == "4.6.1")
)

# The independent audit records its exact historical input snapshot. During
# render, only its stored output records are checked against that snapshot;
# none of the scientific calculations is repeated.
mder_audit_outputs <- mder_audit_manifest |>
  dplyr::filter(.data$role == "output")
mder_output_paths <- absolute_path(mder_audit_outputs$path)
mder_output_present <- file.exists(mder_output_paths)
mder_output_bytes <- rep(NA_real_, length(mder_output_paths))
mder_output_sha256 <- rep(NA_character_, length(mder_output_paths))
mder_output_bytes[mder_output_present] <-
  file.info(mder_output_paths[mder_output_present])$size
mder_output_sha256[mder_output_present] <- vapply(
  mder_output_paths[mder_output_present],
  function(file) digest::digest(file = file, algo = "sha256"),
  character(1)
)
stopifnot(
  all(mder_output_present),
  all(mder_output_bytes == mder_audit_outputs$bytes),
  all(mder_output_sha256 == mder_audit_outputs$sha256)
)

mder_audit_summary <- tibble::tibble(
  Family = "Independent MDER audit outputs",
  `Manifest path` = mder_audit_path,
  `Manifest SHA-256` = current_mder_audit_sha256,
  `Files listed` = nrow(mder_audit_outputs),
  `Files present` = sum(mder_output_present),
  `Size checks` = sum(mder_output_bytes == mder_audit_outputs$bytes),
  `SHA-256 checks` = sum(mder_output_sha256 == mder_audit_outputs$sha256),
  Check = "PASS"
)

gap_repair_manifest_path <-
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_repair_evidence_manifest.csv"
gap_repair_manifest <- read_stored_csv(gap_repair_manifest_path)
assert_columns(
  gap_repair_manifest,
  c(
    "artifact", "path", "sha256", "bytes", "producer", "r_version",
    "status"
  ),
  "gap-timing-unaware MDER repair evidence manifest"
)
stopifnot(
  identical(
    check_file_hash(
      gap_repair_manifest_path,
      current_gap_repair_evidence_sha256
    ),
    current_gap_repair_evidence_sha256
  ),
  all(gap_repair_manifest$r_version == "4.6.1"),
  all(gap_repair_manifest$status == "PASS")
)
gap_repair_paths <- absolute_path(gap_repair_manifest$path)
gap_repair_present <- file.exists(gap_repair_paths)
gap_repair_bytes <- rep(NA_real_, length(gap_repair_paths))
gap_repair_sha256 <- rep(NA_character_, length(gap_repair_paths))
gap_repair_bytes[gap_repair_present] <-
  file.info(gap_repair_paths[gap_repair_present])$size
gap_repair_sha256[gap_repair_present] <- vapply(
  gap_repair_paths[gap_repair_present],
  function(file) digest::digest(file = file, algo = "sha256"),
  character(1)
)
stopifnot(
  all(gap_repair_present),
  all(gap_repair_bytes == gap_repair_manifest$bytes),
  all(gap_repair_sha256 == gap_repair_manifest$sha256)
)
gap_repair_audit_summary <- tibble::tibble(
  Family = "Gap-timing-unaware MDER repair evidence",
  `Manifest path` = gap_repair_manifest_path,
  `Manifest SHA-256` = current_gap_repair_evidence_sha256,
  `Files listed` = nrow(gap_repair_manifest),
  `Files present` = sum(gap_repair_present),
  `Size checks` = sum(gap_repair_bytes == gap_repair_manifest$bytes),
  `SHA-256 checks` = sum(gap_repair_sha256 == gap_repair_manifest$sha256),
  Check = "PASS"
)

l10_zero_manifest_path <-
  "audit/reconciliation/l10_METRIC-011/METRIC-011_evidence_manifest.csv"
l10_zero_manifest <- read_stored_csv(l10_zero_manifest_path)
assert_columns(
  l10_zero_manifest,
  c(
    "artifact", "path", "sha256", "bytes", "producer", "r_version",
    "status"
  ),
  "METRIC-011 numerical-zero evidence manifest"
)
stopifnot(
  identical(
    check_file_hash(l10_zero_manifest_path, l10_zero_evidence_sha256),
    l10_zero_evidence_sha256
  ),
  all(l10_zero_manifest$r_version == "4.6.1"),
  all(l10_zero_manifest$status == "PASS")
)
l10_zero_paths <- absolute_path(l10_zero_manifest$path)
l10_zero_present <- file.exists(l10_zero_paths)
l10_zero_bytes <- rep(NA_real_, length(l10_zero_paths))
l10_zero_sha256 <- rep(NA_character_, length(l10_zero_paths))
l10_zero_bytes[l10_zero_present] <-
  file.info(l10_zero_paths[l10_zero_present])$size
l10_zero_sha256[l10_zero_present] <- vapply(
  l10_zero_paths[l10_zero_present],
  function(file) digest::digest(file = file, algo = "sha256"),
  character(1)
)
stopifnot(
  all(l10_zero_present),
  all(l10_zero_bytes == l10_zero_manifest$bytes),
  all(l10_zero_sha256 == l10_zero_manifest$sha256)
)
l10_zero_audit_summary <- tibble::tibble(
  Family = "L10 numerical-zero evidence",
  `Manifest path` = l10_zero_manifest_path,
  `Manifest SHA-256` = l10_zero_evidence_sha256,
  `Files listed` = nrow(l10_zero_manifest),
  `Files present` = sum(l10_zero_present),
  `Size checks` = sum(l10_zero_bytes == l10_zero_manifest$bytes),
  `SHA-256 checks` = sum(l10_zero_sha256 == l10_zero_manifest$sha256),
  Check = "PASS"
)

check_file_hash(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
  current_gap_participant_day_rds_sha256
)
check_file_hash(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/mder_support.rds",
  current_gap_support_rds_sha256
)

stopifnot(
  identical(
    digest::digest(
      file = absolute_path(
        "artifacts/12_manifests/mder_support_gate_artifacts.csv"
      ),
      algo = "sha256"
    ),
    historical_mder_gate_sha256
  ),
  identical(
    digest::digest(
      file = absolute_path(
        "audit/decisions/mder_mean_of_viable_ratios.md"
      ),
      algo = "sha256"
    ),
    mder_decision_sha256
  ),
  identical(
    check_file_hash(
      "audit/decisions/l10_numerical_zero_normalization.md",
      l10_zero_decision_sha256
    ),
    l10_zero_decision_sha256
  )
)

manifest_validation <- dplyr::bind_rows(
  metric_bundle$summary,
  mder_audit_summary,
  gap_bundle$summary,
  gap_repair_audit_summary,
  l10_zero_audit_summary,
  state_bundle$summary
)

metric_registry <- read_stored_csv("config/metric_display_registry.csv")
settings <- read_stored_csv(
  "artifacts/05_metrics/metric_derivation_settings.csv"
)
state_candidates <- read_stored_csv(
  "artifacts/05_metrics/state_support_candidate_diagnostics.csv"
)
state_inputs <- read_stored_csv(
  "artifacts/05_metrics/metric_state_interval_inputs.csv"
)
mder_verification <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/verification_summary.csv"
)
mder_distribution <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/distribution_summary.csv"
)
non_mder_invariance <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/non_mder_invariance.csv"
)
gap_mder_summary <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_mder_summary.csv"
)
gap_mder_support <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_mder_support_summary.csv"
)
gap_non_mder_invariance <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_non_mder_invariance.csv"
)
l10_zero_reclassifications <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/all_numerical_zero_reclassifications.csv"
)
l10_positive_roundoff <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/positive_roundoff_reclassifications.csv"
)
l10_primary_changes <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/primary_scientific_cell_changes.csv"
)
l10_invariance <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/invariance_summary.csv"
)
l10_manifest_transition <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/manifest_transition.csv"
)

assert_columns(
  metric_registry,
  c(
    "metric_id", "manuscript_name", "manuscript_category",
    "analysis_unit", "display_unit", "analytical_role", "variant_label"
  ),
  "metric display registry"
)
assert_columns(
  settings,
  c(
    "placement", "coverage_path", "coverage_sha256",
    "coverage_settings_path", "coverage_settings_sha256",
    "state_interval_manifest_path", "state_interval_manifest_sha256",
    "reference_profiles_path", "reference_profiles_sha256",
    "timing_exceedance_distributions_path",
    "timing_exceedance_distributions_sha256", "relevance_maps_path",
    "relevance_maps_sha256", "verified_minimum_hour_coverage",
    "verified_minimum_day_coverage", "state_support_cutoff",
    "mder_support_cutoff", "mder_support_decision_id",
    "mder_support_status", "mder_support_rule", "mder_failure_scope",
    "mder_ratio_definition", "mder_pair_resolution_minutes",
    "mder_zero_pair_rule", "mder_nonfinite_pair_rule",
    "mder_ratio_scaled_or_weighted", "minimum_window_support",
    "zero_aware_offset_lx", "numerical_zero_decision_id",
    "numerical_zero_status", "numerical_zero_scope",
    "numerical_zero_rule", "numerical_zero_tolerance_multiplier",
    "numerical_zero_positive_requires_all_source_zero",
    "numerical_zero_raw_value_preserved",
    "numerical_zero_reclassified_cells",
    "numerical_zero_zero_capable_model_rule",
    "numerical_zero_two_part_model_rule",
    "dose_minimum_relevance_support", "maximum_dose_correction_factor",
    "participants", "eligible_participant_days", "thirty_minute_rows",
    "hourly_rows"
  ),
  "metric settings"
)
assert_columns(
  state_candidates,
  c(
    "position", "metric", "state_domain",
    "candidate_state_support_cutoff", "eligible_participant_days",
    "retained_metric_instances", "retained_fraction",
    "incomplete_state_domain_instances", "no_state_window_instances",
    "insufficient_state_support_instances"
  ),
  "diary-period support summary"
)
assert_columns(
  mder_verification,
  c(
    "placement", "eligible_participant_days", "estimable_mder_days",
    "excluded_mder_days", "below_viable_ratio_fraction",
    "no_viable_momentary_ratio", "status", "metric",
    "minimum_viable_ratio_fraction", "metric_manifest_sha256",
    "immutable_inputs", "independently_recalculated", "r_version"
  ),
  "independent MDER verification summary"
)
assert_columns(
  mder_distribution,
  c(
    "dataset", "estimand", "position", "participant_days_available",
    "participant_days_estimable", "participant_days_missing",
    "participants_available", "participants_estimable", "mean", "median"
  ),
  "stored MDER distribution summary"
)
assert_columns(
  non_mder_invariance,
  c(
    "placement", "shared_non_mder_columns", "columns_with_any_difference",
    "differing_cells", "exact_non_mder_invariance"
  ),
  "stored non-MDER invariance summary"
)
assert_columns(
  gap_mder_summary,
  c(
    "position", "participant_days", "participants", "old_estimable_days",
    "new_estimable_days", "new_mean", "new_median"
  ),
  "repaired gap-timing-unaware MDER summary"
)
assert_columns(
  gap_mder_support,
  c(
    "position", "failure_reason", "participant_days", "participants",
    "viable_minutes_minimum", "viable_minutes_median",
    "viable_minutes_maximum"
  ),
  "repaired gap-timing-unaware MDER support summary"
)
assert_columns(
  gap_non_mder_invariance,
  c("check", "value"),
  "gap-timing-unaware non-MDER invariance summary"
)
assert_columns(
  l10_zero_reclassifications,
  c(
    "position", "metric", "raw_backtransformed_value_lx",
    "normalized_value_lx", "numerical_zero_tolerance_lx",
    "source_all_zero", "source_valid_minutes", "source_zero_minutes",
    "source_positive_minutes", "source_missing_minutes",
    "numerical_zero_decision_id", "numerical_zero_reason"
  ),
  "METRIC-011 numerical-zero reclassifications"
)
assert_columns(
  l10_positive_roundoff,
  c(
    "position", "metric", "raw_backtransformed_value_lx",
    "normalized_value_lx", "numerical_zero_tolerance_lx",
    "source_all_zero", "source_valid_minutes", "source_zero_minutes",
    "source_positive_minutes", "source_missing_minutes"
  ),
  "METRIC-011 positive-roundoff reclassifications"
)
assert_columns(
  l10_primary_changes,
  c("position", "metric", "old_value_lx", "new_value_lx"),
  "METRIC-011 primary scientific cell changes"
)
assert_columns(
  l10_invariance,
  c("check", "status", "detail"),
  "METRIC-011 invariance summary"
)
assert_columns(
  l10_manifest_transition,
  c("artifact", "path", "sha256_before", "sha256_after", "changed"),
  "METRIC-011 manifest transition"
)

stopifnot(
  nrow(settings) == 2L,
  setequal(settings$placement, c("glasses", "chest")),
  all(settings$verified_minimum_hour_coverage == 0.50),
  all(settings$verified_minimum_day_coverage == 0.80),
  all(settings$state_support_cutoff == 0.80),
  all(settings$mder_support_cutoff == 0.50),
  all(settings$mder_support_decision_id == "METRIC-010"),
  all(settings$mder_support_status == "author_approved"),
  all(
    settings$mder_support_rule ==
      "positive_finite_one_minute_ratio_fraction_gte_0.50"
  ),
  all(settings$mder_failure_scope == "metric_specific_only_day_retained"),
  all(
    settings$mder_ratio_definition ==
      "arithmetic_mean_of_positive_finite_one_minute_ratios"
  ),
  all(settings$mder_pair_resolution_minutes == 1),
  all(settings$mder_zero_pair_rule == "exclude_if_either_channel_zero"),
  all(
    settings$mder_nonfinite_pair_rule ==
      "exclude_if_source_or_ratio_nonfinite"
  ),
  all(!settings$mder_ratio_scaled_or_weighted),
  all(settings$zero_aware_offset_lx == 0.1),
  all(settings$numerical_zero_decision_id == "METRIC-011"),
  all(settings$numerical_zero_status == "author_approved"),
  all(
    settings$numerical_zero_scope ==
      "offset_geometric_mean_backtransforms"
  ),
  all(settings$numerical_zero_tolerance_multiplier == 100),
  all(settings$numerical_zero_positive_requires_all_source_zero),
  all(settings$numerical_zero_raw_value_preserved),
  all(settings$numerical_zero_reclassified_cells == 6),
  all(
    settings$numerical_zero_zero_capable_model_rule ==
      "retain_participant_day_as_zero"
  ),
  all(
    settings$numerical_zero_two_part_model_rule ==
      paste0(
        "retain_in_zero_occurrence_component;",
        "exclude_only_from_strictly_positive_magnitude_component"
      )
  ),
  all(settings$minimum_window_support == 0.80),
  all(settings$dose_minimum_relevance_support == 0.80),
  all(settings$maximum_dose_correction_factor == 1.25),
  all(settings$thirty_minute_rows == 48 * settings$eligible_participant_days),
  all(settings$hourly_rows == 24 * settings$eligible_participant_days),
  setequal(
    unique(state_candidates$candidate_state_support_cutoff),
    c(0.70, 0.80, 0.90)
  ),
  all(mder_verification$status == "PASS"),
  all(mder_verification$metric == "mder_mean_of_viable_ratios"),
  all(mder_verification$minimum_viable_ratio_fraction == 0.50),
  all(
    mder_verification$metric_manifest_sha256 ==
      mder_audit_metric_manifest_sha256
  ),
  all(mder_verification$immutable_inputs),
  all(mder_verification$independently_recalculated),
  all(mder_verification$r_version == "4.6.1"),
  all(non_mder_invariance$shared_non_mder_columns == 43L),
  all(non_mder_invariance$columns_with_any_difference == 0L),
  all(non_mder_invariance$differing_cells == 0L),
  all(non_mder_invariance$exact_non_mder_invariance),
  identical(
    gap_non_mder_invariance$value[
      gap_non_mder_invariance$check == "non_mder_rows_after"
    ],
    "25620"
  ),
  identical(
    gap_non_mder_invariance$value[
      gap_non_mder_invariance$check == "non_mder_cells_exactly_unchanged"
    ],
    "TRUE"
  ),
  nrow(l10_zero_reclassifications) == 12L,
  identical(
    as.integer(table(factor(
      l10_zero_reclassifications$position,
      levels = c("glasses", "chest")
    ))),
    c(6L, 6L)
  ),
  all(l10_zero_reclassifications$metric == "l10_mean_medi"),
  all(l10_zero_reclassifications$normalized_value_lx == 0),
  all(l10_zero_reclassifications$numerical_zero_decision_id == "METRIC-011"),
  nrow(l10_positive_roundoff) == 8L,
  identical(
    as.integer(table(factor(
      l10_positive_roundoff$position,
      levels = c("glasses", "chest")
    ))),
    c(3L, 5L)
  ),
  isTRUE(all.equal(
    l10_positive_roundoff$raw_backtransformed_value_lx,
    rep(4.163336342344337e-17, 8L),
    tolerance = 0
  )),
  all(l10_positive_roundoff$normalized_value_lx == 0),
  all(l10_positive_roundoff$source_all_zero),
  all(l10_positive_roundoff$source_positive_minutes == 0L),
  sum(l10_positive_roundoff$source_valid_minutes == 600L) == 7L,
  sum(
    l10_positive_roundoff$source_valid_minutes == 547L &
      l10_positive_roundoff$source_missing_minutes == 53L
  ) == 1L,
  all(
    l10_positive_roundoff$source_zero_minutes ==
      l10_positive_roundoff$source_valid_minutes
  ),
  all(
    l10_positive_roundoff$numerical_zero_tolerance_lx ==
      2.220446049250313e-14
  ),
  nrow(l10_primary_changes) == 8L,
  all(l10_primary_changes$metric == "l10_mean_medi"),
  isTRUE(all.equal(
    l10_primary_changes$old_value_lx,
    rep(4.163336342344337e-17, 8L),
    tolerance = 0
  )),
  all(l10_primary_changes$new_value_lx == 0),
  all(
    l10_invariance$status[l10_invariance$check != "hypothesis_models"] ==
      "PASS"
  ),
  identical(
    l10_invariance$status[l10_invariance$check == "hypothesis_models"],
    "NOT_RUN"
  ),
  identical(
    l10_manifest_transition$sha256_after[
      l10_manifest_transition$artifact == "metric"
    ],
    current_metric_manifest_sha256
  ),
  identical(
    l10_manifest_transition$sha256_after[
      l10_manifest_transition$artifact == "manuscript_prepared"
    ],
    current_gap_manifest_sha256
  )
)

grid_data <- dplyr::bind_rows(
  read_stored_csv("artifacts/05_metrics/metrics_glasses_30_minute.csv") |>
    dplyr::mutate(placement = "glasses", grid = "30-minute"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_30_minute.csv") |>
    dplyr::mutate(placement = "chest", grid = "30-minute"),
  read_stored_csv("artifacts/05_metrics/metrics_glasses_one_hour.csv") |>
    dplyr::mutate(placement = "glasses", grid = "One hour"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_one_hour.csv") |>
    dplyr::mutate(placement = "chest", grid = "One hour")
)
assert_columns(
  grid_data,
  c(
    "site", "Id", "local_date", "placement", "grid",
    "bin_admissible", "valid_medi_wall_minutes"
  ),
  "stored 30-minute and hourly grids"
)

admissibility <- dplyr::bind_rows(
  read_stored_csv("artifacts/05_metrics/metrics_glasses_admissibility.csv") |>
    dplyr::mutate(placement = "glasses"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_admissibility.csv") |>
    dplyr::mutate(placement = "chest")
)
assert_columns(
  admissibility,
  c(
    "placement", "analysis_unit", "metric", "state_domain",
    "estimable", "failure_reason"
  ),
  "metric-specific availability data"
)

gap_data <- dplyr::bind_rows(
  read_stored_csv("artifacts/05_metrics/metrics_glasses_gap_diagnostics.csv") |>
    dplyr::mutate(placement = "glasses"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_gap_diagnostics.csv") |>
    dplyr::mutate(placement = "chest")
)
assert_columns(
  gap_data,
  c(
    "site", "Id", "local_date", "placement", "expected_real_minutes",
    "represented_wall_minutes", "valid_medi_real_minutes",
    "unknown_diary_state_real_minutes", "missing_medi_real_minutes",
    "missing_medi_runs", "maximum_missing_medi_run_minutes",
    "dst_fold_real_minutes", "structural_nonexistent_wall_minutes"
  ),
  "gap and daylight-saving diagnostics"
)

sample_summary <- settings |>
  dplyr::transmute(
    Placement = placement_label(.data$placement),
    Sites = vapply(
      .data$placement,
      function(value) {
        dplyr::n_distinct(gap_data$site[gap_data$placement == value])
      },
      integer(1)
    ),
    Participants = .data$participants,
    `Eligible participant-days` = .data$eligible_participant_days,
    `Thirty-minute rows` = .data$thirty_minute_rows,
    `Hourly rows` = .data$hourly_rows
  ) |>
  dplyr::arrange(factor(.data$Placement, levels = c("Near-eye", "Chest")))

bin_support <- grid_data |>
  dplyr::group_by(.data$placement, .data$grid) |>
  dplyr::summarise(
    Values = dplyr::n(),
    `Values calculated` = sum(.data$bin_admissible),
    `Too little data` = sum(!.data$bin_admissible),
    `Valid melEDI minutes` = sum(.data$valid_medi_wall_minutes),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    Placement = placement_label(.data$placement),
    `Time grid and summary` = dplyr::recode(
      .data$grid,
      `30-minute` = "30-minute arithmetic mean",
      `One hour` = "One-hour zero-aware geometric mean"
    )
  ) |>
  dplyr::select(
    .data$Placement,
    .data$`Time grid and summary`,
    .data$Values,
    .data$`Values calculated`,
    .data$`Too little data`,
    .data$`Valid melEDI minutes`
  ) |>
  dplyr::arrange(
    factor(.data$Placement, levels = c("Near-eye", "Chest")),
    .data$`Time grid and summary`
  )

state_support_display <- state_candidates |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(
        metric = .data$metric_id,
        `Diary-period metric` = .data$manuscript_name
      ),
    by = "metric"
  ) |>
  dplyr::mutate(
    Placement = placement_label(.data$position),
    cutoff = paste0(round(100 * .data$candidate_state_support_cutoff), "%")
  ) |>
  dplyr::select(
    .data$Placement,
    .data$`Diary-period metric`,
    `Eligible days` = .data$eligible_participant_days,
    .data$cutoff,
    retained = .data$retained_metric_instances,
    retention = .data$retained_fraction
  ) |>
  tidyr::pivot_wider(names_from = .data$cutoff, values_from = .data$retained) |>
  dplyr::left_join(
    state_candidates |>
      dplyr::filter(.data$candidate_state_support_cutoff == 0.80) |>
      dplyr::left_join(
        metric_registry |>
          dplyr::select(
            metric = .data$metric_id,
            `Diary-period metric` = .data$manuscript_name
          ),
        by = "metric"
      ) |>
      dplyr::transmute(
        Placement = placement_label(.data$position),
        .data$`Diary-period metric`,
        `Primary retention` = scales::percent(
          .data$retained_fraction,
          accuracy = 0.1
        )
      ),
    by = c("Placement", "Diary-period metric")
  ) |>
  dplyr::arrange(
    factor(.data$Placement, levels = c("Near-eye", "Chest")),
    .data$`Diary-period metric`
  )

primary_mder_display <- mder_distribution |>
  dplyr::filter(.data$dataset == "primary") |>
  dplyr::transmute(
    Dataset = "Primary dataset",
    Placement = placement_label(.data$position),
    `Eligible days` = .data$participant_days_available,
    `Days with MDER` = .data$participant_days_estimable,
    `Days without MDER` = .data$participant_days_missing,
    `Participants with MDER` = .data$participants_estimable,
    `Mean MDER` = .data$mean,
    `Median MDER` = .data$median
  )

gap_participants_with_mder <- gap_mder_support |>
  dplyr::filter(is.na(.data$failure_reason)) |>
  dplyr::select(
    .data$position,
    `Participants with MDER` = .data$participants
  )

gap_mder_display <- gap_mder_summary |>
  dplyr::left_join(gap_participants_with_mder, by = "position") |>
  dplyr::transmute(
    Dataset = "Gap-timing-unaware dataset",
    Placement = placement_label(.data$position),
    `Eligible days` = .data$participant_days,
    `Days with MDER` = .data$new_estimable_days,
    `Days without MDER` = .data$participant_days - .data$new_estimable_days,
    .data$`Participants with MDER`,
    `Mean MDER` = .data$new_mean,
    `Median MDER` = .data$new_median
  )

mder_availability_display <- dplyr::bind_rows(
  primary_mder_display,
  gap_mder_display
) |>
  dplyr::arrange(
    factor(
      .data$Dataset,
      levels = c("Primary dataset", "Gap-timing-unaware dataset")
    ),
    factor(.data$Placement, levels = c("Near-eye", "Chest"))
  )

primary_mder <- mder_availability_display |>
  dplyr::filter(.data$Dataset == "Primary dataset")
stopifnot(
  identical(primary_mder$`Eligible days`, c(816, 902)),
  identical(primary_mder$`Days with MDER`, c(702, 732)),
  identical(primary_mder$`Days without MDER`, c(114, 170)),
  identical(primary_mder$`Participants with MDER`, c(137, 152)),
  isTRUE(all.equal(primary_mder$`Mean MDER`, c(0.7238529, 0.7565438),
    tolerance = 1e-7
  )),
  isTRUE(all.equal(primary_mder$`Median MDER`, c(0.7239000, 0.7494261),
    tolerance = 1e-7
  ))
)

gap_mder <- mder_availability_display |>
  dplyr::filter(.data$Dataset == "Gap-timing-unaware dataset")
stopifnot(
  identical(gap_mder$`Eligible days`, c(811, 897)),
  identical(gap_mder$`Days with MDER`, c(687, 723)),
  identical(gap_mder$`Days without MDER`, c(124, 174)),
  identical(gap_mder$`Participants with MDER`, c(137, 152)),
  isTRUE(all.equal(gap_mder$`Mean MDER`, c(0.7242573, 0.7567377),
    tolerance = 1e-7
  )),
  isTRUE(all.equal(gap_mder$`Median MDER`, c(0.7238676, 0.7495175),
    tolerance = 1e-7
  )),
  identical(
    gap_mder_summary$old_estimable_days[
      match(c("glasses", "chest"), gap_mder_summary$position)
    ],
    c(725, 729)
  )
)

mder_plot_data <- mder_verification |>
  dplyr::transmute(
    Placement = placement_label(.data$placement),
    `MDER available` = .data$estimable_mder_days,
    `MDER unavailable` = .data$excluded_mder_days
  ) |>
  tidyr::pivot_longer(
    cols = c(.data$`MDER available`, .data$`MDER unavailable`),
    names_to = "Status",
    values_to = "Participant-days"
  ) |>
  dplyr::mutate(
    Placement = factor(.data$Placement, levels = c("Near-eye", "Chest")),
    Status = factor(
      .data$Status,
      levels = c("MDER unavailable", "MDER available")
    )
  )

availability_summary <- admissibility |>
  dplyr::filter(.data$metric %in% metric_registry$metric_id) |>
  dplyr::group_by(.data$placement, .data$analysis_unit, .data$metric) |>
  dplyr::summarise(
    possible = dplyr::n(),
    calculated = sum(.data$estimable),
    missing = sum(!.data$estimable),
    .groups = "drop"
  ) |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(-.data$analysis_unit),
    by = c("metric" = "metric_id")
  ) |>
  dplyr::mutate(
    Placement = placement_label(.data$placement),
    `Metric family` = dplyr::recode(
      .data$manuscript_category,
      `level-based` = "Level",
      `duration-based` = "Duration",
      `timing-based` = "Timing",
      `exposure-history-based` = "Exposure history",
      `spectrum-based` = "Spectral composition",
      `dynamics-based` = "Temporal dynamics"
    ),
    Metric = dplyr::case_when(
      .data$metric %in% c(
        "dose_observed_medi",
        "dose_time_sensitive_corrected_medi",
        "longest_bout_above_250_exact_only_sensitivity"
      ) ~ paste0(.data$manuscript_name, " — ", .data$variant_label),
      TRUE ~ .data$manuscript_name
    ),
    `One value per` = dplyr::recode(
      .data$analysis_unit,
      participant_day = "Participant-day",
      participant = "Participant"
    ),
    availability = format_count_fraction(.data$calculated, .data$possible)
  ) |>
  dplyr::select(
    .data$`Metric family`,
    .data$Metric,
    .data$`One value per`,
    .data$Placement,
    .data$availability
  ) |>
  tidyr::pivot_wider(
    names_from = .data$Placement,
    values_from = .data$availability
  ) |>
  dplyr::mutate(
    `Metric family` = factor(
      .data$`Metric family`,
      levels = c(
        "Level", "Duration", "Timing", "Exposure history",
        "Spectral composition", "Temporal dynamics"
      )
    )
  ) |>
  dplyr::arrange(.data$`Metric family`, .data$Metric)

l10_availability <- admissibility |>
  dplyr::filter(.data$metric == "l10_mean_medi") |>
  dplyr::group_by(.data$placement) |>
  dplyr::summarise(
    possible = dplyr::n(),
    calculated = sum(.data$estimable),
    .groups = "drop"
  ) |>
  dplyr::arrange(factor(.data$placement, levels = c("glasses", "chest")))
stopifnot(
  identical(l10_availability$placement, c("glasses", "chest")),
  identical(l10_availability$possible, c(816L, 902L)),
  identical(l10_availability$calculated, c(816L, 902L))
)

l10_zero_display <- tibble::tibble(
  Placement = c("Near-eye", "Chest", "Total"),
  `New exact-zero L10 cells` = c(3L, 5L, 8L),
  `Source-minute evidence in selected 600-minute windows` = c(
    "Three windows: 600 observed zero minutes and no missing minutes",
    paste(
      "Four windows: 600 observed zero minutes; one window:",
      "547 observed zero minutes and 53 missing minutes"
    ),
    paste(
      "Seven windows: 600 observed zero minutes; one window:",
      "547 observed zero minutes and 53 missing minutes"
    )
  ),
  `Current L10 values available` = c("816 / 816", "902 / 902", "1,718 / 1,718")
)

gap_summary <- gap_data |>
  dplyr::group_by(.data$placement) |>
  dplyr::summarise(
    `Participant-days` = dplyr::n(),
    `Expected actual-time hours` = sum(.data$expected_real_minutes) / 60,
    `Valid melEDI hours` = sum(.data$valid_medi_real_minutes) / 60,
    `Missing melEDI hours` = sum(.data$missing_medi_real_minutes) / 60,
    `Continuous missing periods` = sum(.data$missing_medi_runs),
    `Longest missing period (min)` = max(
      .data$maximum_missing_medi_run_minutes
    ),
    `Unknown diary-state hours` = sum(
      .data$unknown_diary_state_real_minutes
    ) / 60,
    `Repeated autumn minutes` = sum(.data$dst_fold_real_minutes),
    `Absent spring minutes` = sum(
      .data$structural_nonexistent_wall_minutes
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(Placement = placement_label(.data$placement)) |>
  dplyr::select(-.data$placement) |>
  dplyr::relocate(.data$Placement) |>
  dplyr::arrange(factor(.data$Placement, levels = c("Near-eye", "Chest")))

gap_support_summary <- gap_summary |>
  dplyr::select(
    .data$Placement,
    .data$`Participant-days`,
    .data$`Expected actual-time hours`,
    .data$`Valid melEDI hours`,
    .data$`Missing melEDI hours`,
    .data$`Continuous missing periods`,
    .data$`Longest missing period (min)`,
    .data$`Unknown diary-state hours`
  )

daylight_saving_summary <- gap_summary |>
  dplyr::select(
    .data$Placement,
    .data$`Participant-days`,
    .data$`Repeated autumn minutes`,
    .data$`Absent spring minutes`
  )

path_hash_markdown <- function(path, sha256) {
  paste0(
    wrap_provenance_path(relative_path(path)),
    "<br>SHA-256: ",
    wrap_hash(sha256)
  )
}

coverage_lines <- settings |>
  dplyr::arrange(factor(.data$placement, levels = c("glasses", "chest"))) |>
  dplyr::transmute(
    line = paste0(
      "**", placement_label(.data$placement), ":** ",
      path_hash_markdown(.data$coverage_path, .data$coverage_sha256)
    )
  ) |>
  dplyr::pull(.data$line) |>
  paste(collapse = "<br><br>")

first_settings <- settings[1, , drop = FALSE]
input_display <- tibble::tibble(
  `Input and exact stored artifact` = c(
    paste0("**Rule A one-minute coverage**<br>", coverage_lines),
    paste0(
      "**Sleep and wear intervals**<br>",
      path_hash_markdown(
        first_settings$state_interval_manifest_path,
        first_settings$state_interval_manifest_sha256
      ),
      "<br>", nrow(state_inputs), " site-by-interval files are listed in ",
      "artifacts/05_metrics/metric_state_interval_inputs.csv."
    ),
    paste0(
      "**Fixed intensity profiles**<br>",
      path_hash_markdown(
        first_settings$reference_profiles_path,
        first_settings$reference_profiles_sha256
      )
    ),
    paste0(
      "**Probability of melEDI >250 lx by time of day**<br>",
      path_hash_markdown(
        first_settings$timing_exceedance_distributions_path,
        first_settings$timing_exceedance_distributions_sha256
      )
    ),
    paste0(
      "**Metric-specific relevance maps**<br>",
      path_hash_markdown(
        first_settings$relevance_maps_path,
        first_settings$relevance_maps_sha256
      )
    )
  ),
  `How Preparation 04 uses it` = c(
    paste(
      "Provides the eligible one-minute near-eye and chest records from",
      "Preparation 02. Eligibility has already applied the 50%-per-hour",
      "and 80%-per-day rules; hourly support affects hourly summaries only."
    ),
    paste(
      "Projects diary-defined wake, the three hours before sleep, sleep,",
      "wear, and bedside context onto the complete actual-time minute grid",
      "before light values are joined."
    ),
    paste(
      "Provides the fixed participant-balanced melEDI and illuminance",
      "profiles learned in Preparation 03 for non-MDER metrics. The current",
      "MDER calculation does not use profile weights, scales, or gates."
    ),
    paste(
      "Checks whether observed minutes cover the times when exposure above",
      "250 lx is expected; it never weights or changes an observed timing."
    ),
    paste(
      "Checks time-of-day support for dose, M10, L10, and timing metrics.",
      "Only the approved dose correction may change a metric value; these",
      "maps do not enter the current MDER calculation."
    )
  )
)

script_map <- tibble::tibble(
  `Step and producing code` = c(
    paste0(
      "**1. Reconstruct diary and wear state on the minute grid**<br>",
      "scripts/pipeline/build_metric_derivation.R<br>",
      "scripts/pipeline/state_interval_projection.R"
    ),
    paste0(
      "**2. Derive metrics and metric-specific availability**<br>",
      "scripts/pipeline/metric_derivation.R<br>",
      "scripts/pipeline/time_support.R"
    ),
    paste0(
      "**3. Store values, diagnostics, and inventories**<br>",
      "scripts/pipeline/build_metric_derivation.R"
    ),
    paste0(
      "**4. Reconstruct current MDER independently**<br>",
      "scripts/pipeline/verify_metric_derivation_mder.R"
    ),
    paste0(
      "**5. Preserve the amendment audit record**<br>",
      "audit/scripts/finalize_mder_METRIC_010.R"
    ),
    paste0(
      "**6. Reconstruct and verify the gap-timing-unaware MDER**<br>",
      "audit/scripts/repair_gap_mder_METRIC_010.R<br>",
      "audit/scripts/finalize_gap_mder_METRIC_010_evidence.R"
    ),
    paste0(
      "**7. Verify and seal numerical-zero normalization**<br>",
      "audit/scripts/finalize_l10_METRIC_011.R<br>",
      "scripts/pipeline/verify_metric_derivation_core.R"
    )
  ),
  `Reads, operation, separation, and output` = c(
    paste(
      "Reads the exact sleep and wear interval files from Preparation 01.",
      "It applies half-open [start, end) intervals in actual UTC time, then",
      "checks state, wear, and measurement context where an original record",
      "exists. Keeping this projection separate prevents a missing light row",
      "from erasing a known diary period."
    ),
    paste(
      "Reads the reconstructed minute grid, eligible light values, and fixed",
      "profiles for the metrics that need them. For MDER, time_support.R",
      "constructs the complete 1,440-position local-clock grid, averages",
      "fall-back duplicates channel-wise, forms only finite positive",
      "one-minute melEDI/illuminance ratios, and averages them when at least",
      "720 are viable. For offset geometric means, it preserves the raw",
      "back-transform and applies the source-verified METRIC-011",
      "numerical-zero rule before storing a non-negative value. It records",
      "metric-specific failure reasons."
    ),
    paste(
      "Writes participant-day and participant R objects, complete 30-minute",
      "and hourly grids, rectangular CSV mirrors, availability and gap",
      "diagnostics, settings, exact input inventories, placement-specific",
      "numerical-zero audits, and the 31-file metric",
      "manifest. This storage step is the hand-off to Preparation 06."
    ),
    paste(
      "Reads the stored primary metric files and independently reconstructs",
      "MDER from the stored one-minute inputs. Its separate implementation",
      "checks the local-clock grid, duplicate-minute averaging, viable-pair",
      "count, value, and failure reason. It writes the stored PASS summary."
    ),
    paste(
      "Reads the verified metric manifest and the preserved historical MDER",
      "snapshot. It records current distributions, placement and site support,",
      "paired comparisons, key-level changes, and the exact invariance of 43",
      "non-MDER fields. It writes the six-output METRIC-010 audit bundle and",
      "its manifest; this report reads those files without recreating them."
    ),
    paste(
      "Reads the pinned one-minute melEDI and illuminance inputs for the",
      "gap-timing-unaware dataset and applies the same viable-pair, local-grid,",
      "daylight-saving, and inclusive 720-minute MDER rule. The separate",
      "finalizer records the corrected availability, support reasons, exact",
      "25,620-cell non-MDER invariance, input/output hashes, and a seven-file",
      "PASS evidence manifest under audit/reconciliation/. These scripts are",
      "named for provenance only and are not executed during rendering."
    ),
    paste(
      "The finalizer reads the rebuilt primary metric layer and calls the",
      "independent core verifier before checking the",
      "offset geometric-mean back-transforms, selected L10 source minutes,",
      "exact-zero decisions, and before/after scientific cells. It records",
      "the eight newly changed L10 cells, manifest transitions, invariance",
      "checks, and a seven-file METRIC-011 evidence manifest. It also reseals",
      "only preparation outputs that inherit L10; no hypothesis model is run."
    )
  )
)

core_rules <- tibble::tibble(
  `Scientific operation` = c(
    "Measurement represented",
    "Actual time and local clock",
    "Thirty-minute and hourly values",
    "Numerical zeros after geometric means",
    "Thresholds and missing minutes",
    "Diary-period metrics",
    "Dose and MDER",
    "Brightest and darkest 10 hours",
    "Timing and longest continuous period",
    "Interdaily stability and intradaily variability"
  ),
  `Rule and reason` = c(
    paste(
      "The full day is a placement-specific hybrid: near-eye or chest",
      "measurement while awake, plus bedside sleep-environment measurement",
      "during diary-defined sleep. It is not continuous 24-hour ocular",
      "exposure."
    ),
    paste(
      "Actual UTC time preserves order, elapsed duration, gaps, adjacency,",
      "and both repeated autumn hours. Local clock aligns time-of-day bins,",
      "M10/L10, timing, and interdaily stability."
    ),
    paste(
      "A 30-minute arithmetic mean needs 15 of 30 valid minutes. An hourly",
      "zero-aware geometric mean needs 30 of 60 and uses a 0.1 lx offset so",
      "true zeros remain valid. Unsupported bins stay as missing rows."
    ),
    paste(
      "After adding 0.1 lx, averaging on the log10 scale, and transforming",
      "back, production subtracts the offset. A result within the unit-aware",
      "floating-point tolerance is stored as exact zero only when every",
      "finite source value is exact zero; negative light beyond tolerance",
      "stops production. Missing minutes stay missing and never become zero."
    ),
    paste(
      "Comparisons are strictly >1,000, >250, <10, and <1 lx melEDI.",
      "Missing or invalid minutes are never converted to zero or darkness and",
      "never bridge a continuous period."
    ),
    paste(
      "Wake, pre-sleep, and sleep values require 80% support in that specific",
      "diary period; 70% and 90% are fixed sensitivities. An incomplete diary",
      "domain, absent period, or inadequate support affects only that metric."
    ),
    paste(
      "Observed dose integrates actual durations. Only dose may be corrected",
      "for when missing data occurred, with at least 80% profile relevance and",
      "a maximum factor of 1.25. MDER is the arithmetic mean of viable",
      "one-minute melEDI/illuminance ratios on the 1,440-position local-clock",
      "grid. Both channels must be finite and strictly positive, and at least",
      "720 ratios are required. It is not weighted, scaled, profile-gated, or",
      "calculated as a ratio of daily integrals."
    ),
    paste(
      "M10 tests all 841 non-wrapping one-minute starts; L10 tests all 1,440",
      "starts and may cross midnight. Each window is 600 minutes, uses at",
      "least 80% ordinary and profile support, and returns a geometric mean",
      "and midpoint. The same source-verified numerical-zero rule applies to",
      "the geometric mean. No L5 metric is produced."
    ),
    paste(
      "First, last, and mean timing above 250 lx require sufficient support",
      "where such events are expected. The longest observed uninterrupted",
      "period is a lower bound; a possible upper bound and exact-only",
      "sensitivity preserve uncertainty caused by missing data."
    ),
    paste(
      "Both need at least three eligible days. Interdaily stability needs 20",
      "clock hours represented on at least two days; intradaily variability",
      "needs 24 truly adjacent hourly pairs across at least three days."
    )
  )
)

output_file_lines <- function(types) {
  selected <- metric_manifest |>
    dplyr::filter(.data$artifact_type %in% types) |>
    dplyr::arrange(.data$placement, .data$path)
  paste(
    wrap_provenance_path(relative_path(selected$path)),
    collapse = "<br>"
  )
}

output_display <- tibble::tibble(
  `Stored layer` = c(
    "Participant-day metrics",
    "Participant-level IS and IV",
    "Complete 30-minute grid",
    "Complete hourly grid",
    "Availability, gap, support, and uncertainty records",
    "Primary MDER verification and amendment record",
    "Gap-timing-unaware participant-day metrics and MDER support",
    "Gap-timing-unaware MDER repair evidence",
    "L10 numerical-zero normalization evidence"
  ),
  `Role in the next step` = c(
    "Daily metric values supplied to Preparation 06",
    "Repeated-day metrics supplied to Preparation 06",
    "Thirty-minute analysis rows supplied to Preparation 06",
    "Hourly analysis rows supplied to Preparation 06",
    "Audit and sample-flow evidence; not substituted for metric values",
    "Independent verification and provenance for primary MDER",
    "Repaired sensitivity inputs passed to Preparation 06",
    "Independent verification, corrected sample flow, and invariance evidence",
    "Source-minute audit, eight-cell change record, and invariance evidence"
  ),
  `Exact stored artifacts` = c(
    output_file_lines("participant_day_metrics_rds"),
    output_file_lines("participant_is_iv_rds"),
    output_file_lines("complete_30_minute_arithmetic_medi_grid_rds"),
    output_file_lines("complete_one_hour_geometric_medi_grid_rds"),
    paste(
      wrap_provenance_path(
        c(
          "artifacts/05_metrics/metrics_glasses_admissibility.csv",
          "artifacts/05_metrics/metrics_chest_admissibility.csv",
          "artifacts/05_metrics/metrics_glasses_gap_diagnostics.csv",
          "artifacts/05_metrics/metrics_chest_gap_diagnostics.csv",
          "artifacts/05_metrics/metrics_glasses_support_diagnostics.csv",
          "artifacts/05_metrics/metrics_chest_support_diagnostics.csv"
        )
      ),
      collapse = "<br>"
    ),
    paste(
      wrap_provenance_path(
        c(
          "artifacts/08_diagnostics/mder_METRIC-010/verification_summary.csv",
          "artifacts/08_diagnostics/mder_METRIC-010/distribution_summary.csv",
          "artifacts/08_diagnostics/mder_METRIC-010/support_by_site.csv",
          "artifacts/08_diagnostics/mder_METRIC-010/paired_comparisons.csv",
          "artifacts/08_diagnostics/mder_METRIC-010/key_level_comparison.csv",
          "artifacts/08_diagnostics/mder_METRIC-010/non_mder_invariance.csv",
          "artifacts/08_diagnostics/mder_METRIC-010/audit_manifest.csv"
        )
      ),
      collapse = "<br>"
    ),
    paste(
      wrap_provenance_path(
        c(
          "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
          "artifacts/06_model_data/scenarios/manuscript_prepared_data/mder_support.rds",
          "artifacts/12_manifests/manuscript_prepared_data_artifacts.csv"
        )
      ),
      collapse = "<br>"
    ),
    paste(
      wrap_provenance_path(
        c(
          "audit/reconciliation/mder_METRIC-010_gap_repair/gap_mder_summary.csv",
          "audit/reconciliation/mder_METRIC-010_gap_repair/gap_mder_support_summary.csv",
          "audit/reconciliation/mder_METRIC-010_gap_repair/gap_non_mder_invariance.csv",
          "audit/reconciliation/mder_METRIC-010_gap_repair/gap_repair_input_output_hashes.csv",
          "audit/reconciliation/mder_METRIC-010_gap_repair/gap_repair_evidence_manifest.csv"
        )
      ),
      collapse = "<br>"
    ),
    paste(
      wrap_provenance_path(
        c(
          "audit/reconciliation/l10_METRIC-011/all_numerical_zero_reclassifications.csv",
          "audit/reconciliation/l10_METRIC-011/positive_roundoff_reclassifications.csv",
          "audit/reconciliation/l10_METRIC-011/primary_scientific_cell_changes.csv",
          "audit/reconciliation/l10_METRIC-011/manifest_transition.csv",
          "audit/reconciliation/l10_METRIC-011/invariance_summary.csv",
          "audit/reconciliation/l10_METRIC-011/METRIC-011_evidence_manifest.csv"
        )
      ),
      collapse = "<br>"
    )
  )
)

verification_history <- tibble::tibble(
  `Artifact version` = c(
    paste0(
      "**Current METRIC-011 metric outputs**<br>",
      wrap_hash(current_metric_manifest_sha256),
      "<br>Evidence: ", wrap_hash(l10_zero_evidence_sha256)
    ),
    paste0(
      "**Preceding METRIC-010 MDER audit**<br>",
      "Metric manifest: ", wrap_hash(mder_audit_metric_manifest_sha256),
      "<br>Audit manifest: ", wrap_hash(current_mder_audit_sha256)
    ),
    paste0(
      "**Current repaired gap-timing-unaware MDER**<br>",
      "Preparation manifest: ", wrap_hash(current_gap_manifest_sha256),
      "<br>Repair evidence: ",
      wrap_hash(current_gap_repair_evidence_sha256),
      "<br>Participant-day RDS: ",
      wrap_hash(current_gap_participant_day_rds_sha256),
      "<br>MDER-support RDS: ", wrap_hash(current_gap_support_rds_sha256)
    ),
    paste0(
      "**Current diary-period support gate**<br>",
      wrap_hash(current_state_manifest_sha256)
    ),
    paste0(
      "**Historical ratio-of-integrals support gate**<br>",
      wrap_hash(historical_mder_gate_sha256)
    )
  ),
  `What is documented for that exact version` = c(
    paste(
      "The current metric manifest lists the stored metric layer for 816",
      "near-eye and 902 chest participant-days. METRIC-011 independently",
      "verified the rebuilt layer and records exactly eight primary L10 mean",
      "cells changed from positive numerical roundoff to exact zero. Every",
      "non-L10 value and all samples are unchanged; all seven evidence files",
      "match the sealed evidence manifest under R 4.6.1."
    ),
    paste(
      "The six-output METRIC-010 audit is tied to the explicitly identified",
      "preceding metric manifest. It independently reconstructed MDER values,",
      "viable-minute counts, and failure reasons. METRIC-011 did not change",
      "MDER, and its finalizer independently reran the MDER verifier against",
      "the current rebuild without fitting a model."
    ),
    paste(
      "The repaired prepared-data manifest pins the participant-day and",
      "MDER-support RDS files. Independent evidence records 687 of 811",
      "near-eye and 723 of 897 chest days with MDER, exact preservation of",
      "all 25,620 non-MDER cells, and PASS identities for all seven evidence",
      "files. The earlier 725/729 availability record is superseded."
    ),
    paste(
      "Every listed file is present and matches its stored size and SHA-256.",
      "Exact independent reconstruction of this current diary-period support",
      "manifest remains the sole Preparation 04 item recorded as FIND-044."
    ),
    paste(
      "This gate used profile support and a ratio of paired-channel daily",
      "integrals. It is retained only as historical evidence of the",
      "superseded METRIC-003 method and is not an input to current MDER."
    )
  )
)
```

## Purpose and position in the preparation chain

Preparation 04 converts the eligible one-minute light records from
Preparation 02 into the daily, diary-period, 30-minute, hourly, and
participant-level metrics used by later analyses. It also records whether
each individual metric had enough relevant data to be interpreted. A failed
support check makes only that metric unavailable; it does not discard the
participant-day from unrelated analyses. For offset geometric means, it also
records when floating-point roundoff is normalized to a source-verified exact
zero.

The exact inputs are the Preparation 02 Rule A coverage files, the
Preparation 01 sleep and wear intervals, and the fixed time-of-day profiles
and relevance maps learned in Preparation 03. The main outputs are eight R
data files containing participant-day, participant, 30-minute, and hourly
values for the near-eye and chest placements. [Preparation 06](06_model_ready_datasets.qmd)
combines those files with site and participant context; this report does not
inspect any hypothesis result.

```{mermaid}
flowchart TB
  P1["Preparation 01<br/>sleep and wear intervals"] --> D["Preparation 04<br/>metric derivation"]
  P2["Preparation 02<br/>eligible one-minute records"] --> D
  P2 --> M["Current MDER<br/>viable one-minute ratios"]
  P3["Preparation 03<br/>fixed temporal profiles"] --> N["Non-MDER metrics that<br/>need time-profile support"]
  N --> D
  M --> D
  D --> O["Daily, participant,<br/>30-minute, and hourly metrics"]
  O --> V["Independent MDER<br/>verification record"]
  O --> P6["Preparation 06<br/>model-ready data"]
```

::: {.callout-important title="Interpret the full-day metrics as a hybrid measurement"}
During wake and the three hours before sleep, the record represents the
logger's worn placement: near-eye or chest. During diary-defined sleep, it
represents the bedside sleep environment. A metric spanning the full day is
therefore an environmental-light summary of this planned hybrid record, not a
continuous 24-hour measure of ocular exposure.
:::

## Exact inputs

```{r}
#| label: tbl-preparation04-inputs
#| tbl-cap: "Exact stored inputs to metric derivation and the scientific role of each."

prep_gt(
  input_display |>
    dplyr::mutate(
      `Input and exact stored artifact` = gsub(
        "<br>SHA-256: [^<]*",
        "",
        .data$`Input and exact stored artifact`
      )
    ),
  source_note = paste(
    "Input paths are read from the current stored settings and inventories;",
    "no input is rebuilt during rendering."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Input and exact stored artifact` ~ gt::pct(46),
    `How Preparation 04 uses it` ~ gt::pct(54)
  )
```

Preparation 02 had already retained only participant-days with valid melEDI
for at least 80% of the fixed 1,440-minute hybrid day. The 50% requirement
within an hour governs whether an hourly value can be calculated; it does not
remove otherwise valid one-minute measurements from daily or minute-based
metrics. All-finite, all-zero melEDI days had also been removed from the
primary coverage files while remaining available as a predefined sensitivity.

::: {.callout-note appearance="simple" title="Short glossary"}
- **Metric support:** the observed data available at the times relevant to a
  particular metric; insufficient support makes that metric unavailable but
  does not remove the participant-day.
- **M10 and L10:** the mean melEDI during, respectively, the brightest and
  darkest 10-hour windows of a participant-day.
- **MDER:** the melanopic daylight efficacy ratio, calculated here as the
  arithmetic mean of viable one-minute melEDI-to-illuminance ratios.
- **Transformed and back-transformed quantities:** calculations may be made on
  a scale chosen for statistical or numerical reasons; back-transformation
  returns the result to its original practical unit, such as lx.
:::

## Scientific rules applied during production

```{r}
#| label: tbl-metric-rules
#| tbl-cap: "Metric construction, temporal support, and missing-data rules applied by the production scripts."

prep_gt(core_rules) |>
  gt::cols_width(
    `Scientific operation` ~ gt::pct(31),
    `Rule and reason` ~ gt::pct(69)
  )
```

### Source-verified zeros after geometric means

Geometric means can include true zero measurements because production first
adds 0.1 lx, averages on the log10 scale, transforms back, and subtracts the
same 0.1 lx. In exact arithmetic, an all-zero source window returns zero.
Floating-point arithmetic can instead leave a minute positive remainder: the
eight affected L10 means were
`4.163336342344337e-17` lx, far below the source-verified numerical tolerance
of `2.220446049250313e-14` lx.

The active numerical-zero rule stores a positive remainder within that
tolerance as exact zero only when every finite source minute is itself exactly
zero. Missing minutes remain missing; they are counted separately and never
imputed as zeros. A materially negative result stops production, and a small
positive result with any positive source minute is retained.

```{r}
#| label: tbl-l10-numerical-zero
#| tbl-cap: "Primary L10 values changed from positive floating-point roundoff to source-verified exact zero."

prep_gt(
  l10_zero_display,
  source_note = paste(
    "Counts and source-minute evidence are read from the stored independent",
    "numerical-zero record; the report does not recalculate an L10 window."
  )
) |>
  gt::fmt_number(
    columns = .data$`New exact-zero L10 cells`,
    decimals = 0
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(13),
    `New exact-zero L10 cells` ~ gt::pct(16),
    `Source-minute evidence in selected 600-minute windows` ~ gt::pct(48),
    `Current L10 values available` ~ gt::pct(23)
  )
```

Exactly eight primary L10 mean cells changed: three near-eye and five chest.
Seven selected 600-minute windows contained 600 observed zero minutes. The
remaining chest window contained 547 observed zero minutes and 53 missing
minutes; all 547 observed values were zero, and the 53 missing minutes stayed
missing. L10 remains available for all 816 near-eye and 902 chest
participant-days, so neither the sample nor its support rule changed.
Every non-L10 value and all samples are unchanged. No hypothesis model is run
or rebuilt by this report.

The sealed audit contains 12 numerical-zero normalizations in total: four
previously documented negative roundoff residuals and the eight positive
residuals above. A zero-capable analysis retains these participant-days as
exact zeros; a later two-part model retains them in its zero/occurrence part
and excludes them only from a strictly positive magnitude part by definition.

The central distinction is between participant-day eligibility and
metric-specific availability. A day that passed the general coverage rule
remains in the prepared data even when, for example, its pre-sleep period is
unknown, its M10 window has insufficient support, or its paired illuminance
data cannot support MDER.

## Cohort and time-grid outputs

```{r}
#| label: tbl-current-metric-sample
#| tbl-cap: "Participants, participant-days, and complete time-grid rows in the current stored metric outputs."

prep_gt(
  sample_summary,
  source_note = paste(
    "Every participant-day contributes 48 planned 30-minute rows and 24",
    "planned hourly rows, including rows whose metric value is unavailable."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$Sites, .data$Participants, .data$`Eligible participant-days`,
      .data$`Thirty-minute rows`, .data$`Hourly rows`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(18),
    Sites ~ gt::pct(12),
    Participants ~ gt::pct(17),
    `Eligible participant-days` ~ gt::pct(19),
    `Thirty-minute rows` ~ gt::pct(17),
    `Hourly rows` ~ gt::pct(17)
  )
```

Near-eye metrics contain 141 participants from nine sites and 816
participant-days. Chest metrics contain 154 participants from eight sites and
902 participant-days. These are the current stored Preparation 04 samples,
not downstream model-specific complete-case samples.

### Diary-period support

Wake, the three hours before sleep, and sleep each use a metric-specific 80%
support rule. The 70% and 90% columns show the fixed sensitivity cutoffs; they
do not change general participant-day eligibility.

```{r}
#| label: tbl-state-support
#| tbl-cap: "Diary-period metric values retained at the fixed 70%, 80%, and 90% support cutoffs."

prep_gt(
  state_support_display,
  source_note = paste(
    "The 80% column is primary. Incomplete diary domains and absent target",
    "periods remain distinct from insufficient observed melEDI support."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$`Eligible days`, .data$`70%`, .data$`80%`, .data$`90%`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    `Diary-period metric` ~ gt::pct(35),
    `Eligible days` ~ gt::pct(11),
    `70%` ~ gt::pct(9),
    `80%` ~ gt::pct(9),
    `90%` ~ gt::pct(9),
    `Primary retention` ~ gt::pct(13)
  )
```

At the primary cutoff, the near-eye output retains 737 wake, 655 pre-sleep,
and 778 sleep-environment values. The chest output retains 818, 743, and 861,
respectively. An unavailable value carries one of three distinct explanations:
the diary could not bound the day's state domains, the target period did not
occur, or the target period contained too little valid melEDI.

### MDER from viable one-minute channel pairs

The melanopic daylight efficacy ratio (MDER) describes melEDI relative to
photopic illuminance. For each participant-day, production first represents
the complete 1,440-position local wall-clock grid. If the autumn clock change
creates two observations for one local minute, melEDI and illuminance are each
averaged within that clock minute before their ratio is formed. A local minute
that does not exist during the spring clock change remains missing.

A minute contributes only when both channels are finite and strictly greater
than zero. This restriction defines a viable proportional comparison: a zero
illuminance denominator is undefined, and a zero in either channel does not
represent a meaningful positive channel ratio for this estimand. MDER is the
arithmetic mean of the viable one-minute melEDI/illuminance ratios. At least
720 of the 1,440 positions must be viable; the boundary is inclusive, so
exactly 720 passes. If fewer are viable, only MDER is unavailable. The
participant-day and all other metrics remain eligible.

No time profile weights, gates, or scales enter current MDER, and it is not a
ratio of daily channel integrals. Records describing a profile-supported
ratio-of-integrals method and 70%, 80%, or 90% MDER gates document an earlier
method; they are not active.

The **gap-timing-unaware dataset** still passed the general 50%-per-hour and
80%-per-day coverage rules. Its name means that the timing of remaining
missing observations was not used for an additional metric-specific
adjustment. For this first explanation only, the primary dataset could be
interpreted as a **time-sensitive primary metric dataset**; hereafter it is
called simply the **primary dataset**. Both datasets now use the same
momentary-ratio MDER definition.

```{r}
#| label: tbl-mder-support
#| tbl-cap: "Current MDER availability and distribution summaries under the same viable one-minute-ratio estimand."

prep_gt(
  mder_availability_display,
  groupname_col = "Dataset",
  source_note = paste(
    "Primary and gap-timing-unaware rows are read from their stored",
    "independent-verification summaries; rendering does not reconstruct MDER."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$`Eligible days`, .data$`Days with MDER`,
      .data$`Days without MDER`, .data$`Participants with MDER`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::fmt_number(
    columns = c(.data$`Mean MDER`, .data$`Median MDER`),
    decimals = 4
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(15),
    `Eligible days` ~ gt::pct(12),
    `Days with MDER` ~ gt::pct(14),
    `Days without MDER` ~ gt::pct(15),
    `Participants with MDER` ~ gt::pct(18),
    `Mean MDER` ~ gt::pct(13),
    `Median MDER` ~ gt::pct(13)
  )
```

In the primary dataset, MDER is available for 702 of 816 near-eye
participant-days from 137 participants (mean 0.7239; median 0.7239) and 732 of
902 chest participant-days from 152 participants (mean 0.7565; median
0.7494). In the gap-timing-unaware dataset, it is available for 687 of 811
near-eye participant-days from 137 participants (mean 0.7243; median 0.7239)
and 723 of 897 chest participant-days from 152 participants (mean 0.7567;
median 0.7495).

The stored comparison found all 25,620 non-MDER participant-day cells exactly
unchanged. A failed MDER support rule therefore removes only MDER, not the day
or any other metric.

```{r}
#| label: fig-mder-availability
#| fig-cap: "Availability of the current MDER in the primary dataset. At least 720 viable one-minute ratios are required. Counts are read from the stored independent-verification summary; rendering does not reconstruct MDER."
#| fig-alt: "Two horizontal stacked bars show MDER availability in the primary dataset. Near-eye has 702 participant-days with MDER and 114 without, out of 816. Chest has 732 with MDER and 170 without, out of 902."
#| fig-width: 7.2
#| fig-height: 4.2

ggplot(
  mder_plot_data,
  aes(
    x = .data$`Participant-days`,
    y = .data$Placement,
    fill = .data$Status
  )
) +
  geom_col(width = 0.58, colour = "white", linewidth = 0.4) +
  geom_text(
    aes(label = scales::comma(.data$`Participant-days`)),
    position = position_stack(vjust = 0.5),
    colour = "white",
    size = 3.2,
    fontface = "bold"
  ) +
  scale_x_continuous(
    labels = scales::label_comma(),
    expand = expansion(mult = c(0, 0.03))
  ) +
  scale_fill_manual(
    values = c(
      "MDER available" = "#0072B2",
      "MDER unavailable" = "#D55E00"
    ),
    breaks = c("MDER available", "MDER unavailable")
  ) +
  labs(
    x = "Eligible participant-days",
    y = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text = element_text(size = 9, colour = "#333333"),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(7, 10, 7, 7)
  )
```

[Download the source data for @fig-mder-availability](../../artifacts/08_diagnostics/mder_METRIC-010/verification_summary.csv).

### Thirty-minute and hourly values

```{r}
#| label: tbl-bin-support
#| tbl-cap: "Stored 30-minute and hourly values with and without enough valid melEDI data."

prep_gt(
  bin_support,
  source_note = paste(
    "A row is retained even when it has too little support to calculate the",
    "requested summary."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$Values, .data$`Values calculated`, .data$`Too little data`,
      .data$`Valid melEDI minutes`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    `Time grid and summary` ~ gt::pct(31),
    Values ~ gt::pct(12),
    `Values calculated` ~ gt::pct(15),
    `Too little data` ~ gt::pct(13),
    `Valid melEDI minutes` ~ gt::pct(15)
  )
```

The regular grids preserve unsupported periods as explicit missing values.
This matters because deleting rows would compress time and could create false
adjacency, change a window's apparent duration, or hide where coverage was
lost.

## Which metric values are available?

```{r}
#| label: tbl-metric-availability
#| tbl-cap: "Metric values calculated in the current stored outputs, shown as calculated / possible (percentage)."

prep_gt(
  availability_summary,
  groupname_col = "Metric family",
  source_note = paste(
    "Counts describe Preparation 04 availability before any",
    "hypothesis-specific complete-case filtering. Internal M10/L10 onset and",
    "offset diagnostics are not reader-facing analytical metrics."
  )
) |>
  gt::cols_width(
    Metric ~ gt::pct(45),
    `One value per` ~ gt::pct(17),
    `Near-eye` ~ gt::pct(19),
    Chest ~ gt::pct(19)
  )
```

A known absence and an unknown value remain different. For example, an
adequately observed day with no exposure above a threshold has a duration of
zero, but first, last, and mean timing do not exist. For the longest period
above 250 lx, the observed duration remains a lower bound when a missing
interval could conceal a longer period; the separate exact-only sensitivity
is available only when the lower and possible upper bounds agree.

## Gaps and daylight-saving changes

```{r}
#| label: tbl-gap-diagnostics
#| tbl-cap: "Actual-time coverage, missing periods, diary-state uncertainty, and daylight-saving diagnostics in the current metric outputs."

prep_gt(
  gap_support_summary,
  source_note = paste(
    "Expected and missing hours use the actual UTC minute grid. Repeated",
    "autumn and absent spring minutes describe the separate local-clock grid."
  )
) |>
  gt::fmt_number(
    columns = where(is.numeric),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(12),
    `Participant-days` ~ gt::pct(10),
    `Expected actual-time hours` ~ gt::pct(13),
    `Valid melEDI hours` ~ gt::pct(11),
    `Missing melEDI hours` ~ gt::pct(11),
    `Continuous missing periods` ~ gt::pct(12),
    `Longest missing period (min)` ~ gt::pct(11),
    `Unknown diary-state hours` ~ gt::pct(19)
  )
```

```{r}
#| label: tbl-daylight-saving-diagnostics
#| tbl-cap: "Daylight-saving minutes represented on the separate actual-time and local-clock grids."

prep_gt(
  daylight_saving_summary,
  source_note = paste(
    "Repeated autumn minutes remain as two actual-time observations but one",
    "averaged local-clock position. Absent spring minutes remain structural",
    "gaps on the local-clock grid."
  )
) |>
  gt::fmt_number(
    columns = where(is.numeric),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(25),
    `Participant-days` ~ gt::pct(25),
    `Repeated autumn minutes` ~ gt::pct(25),
    `Absent spring minutes` ~ gt::pct(25)
  )
```

Actual-time metrics retain both occurrences of a repeated autumn hour and
respect the true elapsed gap during the spring transition. Clock-aligned
metrics use a fixed 24-hour local-clock grid; repeated autumn values are
averaged only on that grid, and structurally absent spring minutes remain
absent. Missing light, unknown diary state, and a daylight-saving transition
are therefore not treated as the same event.

## Limitation of current independent verification

Independent reconstruction of current diary-period support remains incomplete.

This qualification applies to the current diary-period support files. It
does not apply to current MDER and is not evidence that the diary-period data
or any downstream result is wrong. The file checks below do not attempt that
scientific reconstruction.

## Technical provenance and version-specific verification

::: {.callout-note title="What is calculated during this render?"}
This report reads already stored CSV outputs. During rendering it checks file
presence, byte size, SHA-256 identity, required columns, and fixed row-count
relationships. It also formats lightweight summaries already stored by the
independent MDER audits and the numerical-zero evidence. It does **not** run a
preparation builder, change a metric, independently reconstruct MDER, L10, or
another support classification, fit a model, or overwrite an accepted
analytical artifact.
:::

### Production sequence and stored hand-offs

The producing scripts are shown in execution order. They remain separate so
that support decisions, state reconstruction, metric calculation, and file
storage each have an inspectable input and output boundary.

```{r}
#| label: tbl-script-execution-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

prep_gt(
  script_map,
  source_note = paste(
    "The visible report translates each module's role; the paths preserve",
    "the exact production provenance."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

### Exact artifacts passed forward

```{r}
#| label: tbl-output-artifacts
#| tbl-cap: "Preparation 04 artifacts stored for Preparation 06 and for metric-specific audit."

prep_gt(
  output_display,
  source_note = paste(
    "All listed analytical files were produced by",
    "scripts/pipeline/build_metric_derivation.R; rendering reads them only."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Stored layer` ~ gt::pct(22),
    `Role in the next step` ~ gt::pct(29),
    `Exact stored artifacts` ~ gt::pct(49)
  )
```

The eight R data files in the first four rows are the metric layer used to
construct the primary dataset in Preparation 06. The repaired
gap-timing-unaware participant-day and MDER-support files use the same
momentary-ratio MDER estimand and are passed forward separately. Historical
internal filenames remain for provenance but are not visible dataset labels.
The two audit bundles document the exact primary verification and the repaired
gap-timing-unaware sample flow, support reasons, and non-MDER invariance. The
METRIC-011 bundle separately records the source minutes, raw back-transforms,
eight changed L10 cells, manifest transition, and exact non-L10 invariance.

Preparation 04 creates the primary dataset's metric layer. Preparation 05
acquires questionnaire and diary inputs, and Preparation 06 brings the metric
and non-light inputs together. The gap-timing-unaware dataset remains a
predefined prepared-data sensitivity and is not substituted into the primary
files listed above.

### File checks and version-specific evidence

```{r}
#| label: tbl-current-manifest-validation
#| tbl-cap: "Bounded file-identity checks for the current metric, repaired gap-timing-unaware, diary-period, MDER-audit, and L10 numerical-zero artifacts."

manifest_validation_display <- manifest_validation |>
  dplyr::mutate(
    `Manifest path` = wrap_provenance_path(.data$`Manifest path`),
    `Manifest SHA-256` = wrap_hash(.data$`Manifest SHA-256`)
  )

prep_gt(
  manifest_validation_display,
  source_note = paste(
    "For the metric, repaired gap-timing-unaware, and diary-period manifests,",
    "PASS means that every listed file matches its stored size and SHA-256.",
    "The primary MDER audit, seven-file gap repair evidence bundle, and",
    "seven-file METRIC-011 evidence bundle also match their independently",
    "produced manifests.",
    "Rendering performs identity checks only."
  )
) |>
  gt::fmt_markdown(columns = c(.data$`Manifest path`)) |>
  gt::fmt_number(
    columns = c(
      .data$`Files listed`, .data$`Files present`, .data$`Size checks`,
      .data$`SHA-256 checks`
    ),
    decimals = 0
  ) |>
  gt::cols_width(
    Family ~ gt::pct(16),
    `Manifest path` ~ gt::pct(24),
    `Manifest SHA-256` ~ gt::pct(25),
    `Files listed` ~ gt::pct(7),
    `Files present` ~ gt::pct(7),
    `Size checks` ~ gt::pct(7),
    `SHA-256 checks` ~ gt::pct(8),
    Check ~ gt::pct(6)
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(columns = .data$Check)
  )
```

Independent verification is version-specific. The current metric manifest
and numerical-zero results belong to the METRIC-011 evidence bundle. The
six-output METRIC-010 MDER audit belongs to the explicitly identified
preceding metric manifest; METRIC-011 changed no MDER value and independently
reran its verifier against the current rebuild. The repaired
gap-timing-unaware MDER verification belongs to its prepared-data manifest and
repair evidence bundle. The remaining qualification applies only to the
current diary-period support manifest.

```{r}
#| label: tbl-version-specific-verification
#| tbl-cap: "Which verification statements apply to the current and earlier Preparation 04 artifact versions."

prep_gt(
  verification_history,
  source_note = paste(
    "The controlling METRIC-011 decision has SHA-256",
    wrap_hash(l10_zero_decision_sha256),
    "; METRIC-010 remains the active MDER estimand decision with SHA-256",
    wrap_hash(mder_decision_sha256), "."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Artifact version` ~ gt::pct(38),
    `What is documented for that exact version` ~ gt::pct(62)
  )
```

The active metric manifest is
`028bce108339c1277df4a070597430ea20b49c34c2888f4eef42741c1fe76a5e`,
and the METRIC-011 evidence manifest is
`a37efd3449a8d1a6065d0eb8964bd6cf8b241f1683a26c065ea1a946e23214fb`.
They record the current geometric-mean zero rule, all eight primary L10 cell
changes, and exact invariance outside L10. The preceding metric manifest
`7bb2830c695e5600f0877acca500a187a040fe727a029bdb069f69099abe4f43`
and independent MDER audit manifest
`5bd34fd4ac26dc2761540393d40e096cde835d8dd4b4edda9253243c9083d4bb`
remain identified because that stored six-output audit is tied to their exact
version. METRIC-011 changed no MDER value and independently reran the MDER
verifier against the current rebuild.

The repaired gap-timing-unaware preparation manifest is
`4ed62fbe58de65a6d05d8cfc6b5d870d7c74a3c83fe89bcd72bd1dd838698935`,
and its independent repair evidence manifest is
`81e3aa9439b88cf239d6669dc4343cfb146cd2f45790bae412ab837156315018`.
They pin the participant-day RDS
`7561b5dd47cb5e23ca94ba59bd57648f11fe492f86af53a566c5c9cf42d932f1`
and MDER-support RDS
`a0bc5d7ea2142709412da2253158086416ddb730b0f60ff616bba9d6ccb1edc5`.

Under [PREP-003](../../audit/decisions/preparation04_version_specific_verification.md),
exact independent reconstruction of the current diary-period support manifest
`755508076e42ed74ce0d40f48de192b69ea81b04532db2d5a950c451236e2619`
remains open audit item **FIND-044**. This remaining qualification does not
apply to current MDER and is not evidence that the diary-period data or any
downstream result is wrong. Rendering does not attempt that scientific check.
