---
title: "Preparation 04: Derive light-exposure metrics"
subtitle: "From eligible one-minute records to metric-specific analysis inputs"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-04
#| include: false

library(digest)
library(dplyr)
library(ggplot2)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

relative_path <- function(path) {
  sub(paste0("^", root, "/"), "", path)
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE, progress = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

wrap_hash <- function(value) {
  gsub("(.{8})", "\\1\u200b", value, perl = TRUE)
}

placement_label <- function(value) {
  dplyr::recode(value, glasses = "Near-eye", chest = "Chest")
}

format_integer <- function(value) {
  scales::comma(value, accuracy = 1)
}

format_count_fraction <- function(calculated, possible) {
  paste0(
    format_integer(calculated),
    " / ",
    format_integer(possible),
    " (",
    scales::percent(calculated / possible, accuracy = 0.1),
    ")"
  )
}

check_manifest <- function(path, family, expected_sha256) {
  manifest_path <- absolute_path(path)
  manifest <- read_stored_csv(path)
  assert_columns(
    manifest,
    c("path", "sha256", "bytes", "producer", "artifact_type"),
    paste(family, "manifest")
  )

  manifest_sha256 <- digest::digest(
    file = manifest_path,
    algo = "sha256"
  )
  stopifnot(identical(manifest_sha256, expected_sha256))

  listed_paths <- absolute_path(manifest$path)
  present <- file.exists(listed_paths)
  current_bytes <- rep(NA_real_, length(listed_paths))
  current_sha256 <- rep(NA_character_, length(listed_paths))
  current_bytes[present] <- file.info(listed_paths[present])$size
  current_sha256[present] <- vapply(
    listed_paths[present],
    function(file) digest::digest(file = file, algo = "sha256"),
    character(1)
  )

  summary <- tibble::tibble(
    Family = family,
    `Manifest path` = path,
    `Manifest SHA-256` = manifest_sha256,
    `Files listed` = nrow(manifest),
    `Files present` = sum(present),
    `Size checks` = sum(present & current_bytes == manifest$bytes),
    `SHA-256 checks` = sum(present & current_sha256 == manifest$sha256),
    Check = ifelse(
      all(
        present &
          current_bytes == manifest$bytes &
          current_sha256 == manifest$sha256
      ),
      "PASS",
      "FAIL"
    )
  )
  stopifnot(summary$Check == "PASS")

  list(data = manifest, summary = summary)
}

current_metric_manifest_sha256 <-
  "6ec3185620d921e1f81b8464e850249613deb823988187d58794b3471303f4e8"
current_mder_manifest_sha256 <-
  "9ca9b97dac5efd9ec4e7b43a1fb2227b7a13cf98ac4b92318606c4874f9e460b"
current_state_manifest_sha256 <-
  "755508076e42ed74ce0d40f48de192b69ea81b04532db2d5a950c451236e2619"
earlier_metric_manifest_sha256 <-
  "ac0325716421f2d90bb83b445ffb6ad965b131e382373e45f4ad6157eac5937d"
intervening_metric_manifest_sha256 <-
  "944f395e00735e6f8200798d4cde387454441d39b9defda68f583bd8778cd34b"
earlier_mder_manifest_sha256 <-
  "c4ecfab41892e44b38dd78097277ccef5f94a379b44c7b309de281e9215a7b3c"

metric_bundle <- check_manifest(
  "artifacts/12_manifests/metric_artifacts.csv",
  "Metric outputs",
  current_metric_manifest_sha256
)
mder_bundle <- check_manifest(
  "artifacts/12_manifests/mder_support_gate_artifacts.csv",
  "MDER support gate",
  current_mder_manifest_sha256
)
state_bundle <- check_manifest(
  "artifacts/12_manifests/state_support_gate_artifacts.csv",
  "Diary-period support gate",
  current_state_manifest_sha256
)

metric_manifest <- metric_bundle$data
mder_manifest <- mder_bundle$data
state_manifest <- state_bundle$data
manifest_validation <- dplyr::bind_rows(
  metric_bundle$summary,
  mder_bundle$summary,
  state_bundle$summary
)

metric_registry <- read_stored_csv("config/metric_display_registry.csv")
settings <- read_stored_csv(
  "artifacts/05_metrics/metric_derivation_settings.csv"
)
state_candidates <- read_stored_csv(
  "artifacts/05_metrics/state_support_candidate_diagnostics.csv"
)
mder_candidates <- read_stored_csv(
  paste0(
    "artifacts/08_diagnostics/mder_support_gate/",
    "mder_support_candidate_summary.csv"
  )
)
state_inputs <- read_stored_csv(
  "artifacts/05_metrics/metric_state_interval_inputs.csv"
)
mder_inputs <- read_stored_csv(
  paste0(
    "artifacts/08_diagnostics/mder_support_gate/",
    "mder_support_gate_inputs.csv"
  )
)

assert_columns(
  metric_registry,
  c(
    "metric_id", "manuscript_name", "manuscript_category",
    "analysis_unit", "display_unit", "analytical_role", "variant_label"
  ),
  "metric display registry"
)
assert_columns(
  settings,
  c(
    "placement", "coverage_path", "coverage_sha256",
    "coverage_settings_path", "coverage_settings_sha256",
    "state_interval_manifest_path", "state_interval_manifest_sha256",
    "reference_profiles_path", "reference_profiles_sha256",
    "timing_exceedance_distributions_path",
    "timing_exceedance_distributions_sha256", "relevance_maps_path",
    "relevance_maps_sha256", "verified_minimum_hour_coverage",
    "verified_minimum_day_coverage", "state_support_cutoff",
    "mder_support_cutoff", "minimum_window_support",
    "dose_minimum_relevance_support", "maximum_dose_correction_factor",
    "participants", "eligible_participant_days", "thirty_minute_rows",
    "hourly_rows"
  ),
  "metric settings"
)
assert_columns(
  state_candidates,
  c(
    "position", "metric", "state_domain",
    "candidate_state_support_cutoff", "eligible_participant_days",
    "retained_metric_instances", "retained_fraction",
    "incomplete_state_domain_instances", "no_state_window_instances",
    "insufficient_state_support_instances"
  ),
  "diary-period support summary"
)
assert_columns(
  mder_candidates,
  c(
    "position", "candidate_support_cutoff",
    "eligible_participant_days", "retained_participant_days",
    "retained_participants", "sites_with_retained_days",
    "no_paired_observation_days",
    "nonpositive_paired_light_integral_days", "below_any_support_days",
    "retained_fraction"
  ),
  "MDER support summary"
)

stopifnot(
  nrow(settings) == 2L,
  setequal(settings$placement, c("glasses", "chest")),
  all(settings$verified_minimum_hour_coverage == 0.50),
  all(settings$verified_minimum_day_coverage == 0.80),
  all(settings$state_support_cutoff == 0.80),
  all(settings$mder_support_cutoff == 0.80),
  all(settings$minimum_window_support == 0.80),
  all(settings$dose_minimum_relevance_support == 0.80),
  all(settings$maximum_dose_correction_factor == 1.25),
  all(settings$thirty_minute_rows == 48 * settings$eligible_participant_days),
  all(settings$hourly_rows == 24 * settings$eligible_participant_days),
  setequal(
    unique(state_candidates$candidate_state_support_cutoff),
    c(0.70, 0.80, 0.90)
  ),
  setequal(
    unique(mder_candidates$candidate_support_cutoff),
    c(0.70, 0.80, 0.90)
  )
)

grid_data <- dplyr::bind_rows(
  read_stored_csv("artifacts/05_metrics/metrics_glasses_30_minute.csv") |>
    dplyr::mutate(placement = "glasses", grid = "30-minute"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_30_minute.csv") |>
    dplyr::mutate(placement = "chest", grid = "30-minute"),
  read_stored_csv("artifacts/05_metrics/metrics_glasses_one_hour.csv") |>
    dplyr::mutate(placement = "glasses", grid = "One hour"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_one_hour.csv") |>
    dplyr::mutate(placement = "chest", grid = "One hour")
)
assert_columns(
  grid_data,
  c(
    "site", "Id", "local_date", "placement", "grid",
    "bin_admissible", "valid_medi_wall_minutes"
  ),
  "stored 30-minute and hourly grids"
)

admissibility <- dplyr::bind_rows(
  read_stored_csv("artifacts/05_metrics/metrics_glasses_admissibility.csv") |>
    dplyr::mutate(placement = "glasses"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_admissibility.csv") |>
    dplyr::mutate(placement = "chest")
)
assert_columns(
  admissibility,
  c(
    "placement", "analysis_unit", "metric", "state_domain",
    "estimable", "failure_reason"
  ),
  "metric-specific availability data"
)

gap_data <- dplyr::bind_rows(
  read_stored_csv("artifacts/05_metrics/metrics_glasses_gap_diagnostics.csv") |>
    dplyr::mutate(placement = "glasses"),
  read_stored_csv("artifacts/05_metrics/metrics_chest_gap_diagnostics.csv") |>
    dplyr::mutate(placement = "chest")
)
assert_columns(
  gap_data,
  c(
    "site", "Id", "local_date", "placement", "expected_real_minutes",
    "represented_wall_minutes", "valid_medi_real_minutes",
    "unknown_diary_state_real_minutes", "missing_medi_real_minutes",
    "missing_medi_runs", "maximum_missing_medi_run_minutes",
    "dst_fold_real_minutes", "structural_nonexistent_wall_minutes"
  ),
  "gap and daylight-saving diagnostics"
)

sample_summary <- settings |>
  dplyr::transmute(
    Placement = placement_label(.data$placement),
    Sites = vapply(
      .data$placement,
      function(value) {
        dplyr::n_distinct(gap_data$site[gap_data$placement == value])
      },
      integer(1)
    ),
    Participants = .data$participants,
    `Eligible participant-days` = .data$eligible_participant_days,
    `Thirty-minute rows` = .data$thirty_minute_rows,
    `Hourly rows` = .data$hourly_rows
  ) |>
  dplyr::arrange(factor(.data$Placement, levels = c("Near-eye", "Chest")))

bin_support <- grid_data |>
  dplyr::group_by(.data$placement, .data$grid) |>
  dplyr::summarise(
    Values = dplyr::n(),
    `Values calculated` = sum(.data$bin_admissible),
    `Too little data` = sum(!.data$bin_admissible),
    `Valid melEDI minutes` = sum(.data$valid_medi_wall_minutes),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    Placement = placement_label(.data$placement),
    `Time grid and summary` = dplyr::recode(
      .data$grid,
      `30-minute` = "30-minute arithmetic mean",
      `One hour` = "One-hour zero-aware geometric mean"
    )
  ) |>
  dplyr::select(
    .data$Placement,
    .data$`Time grid and summary`,
    .data$Values,
    .data$`Values calculated`,
    .data$`Too little data`,
    .data$`Valid melEDI minutes`
  ) |>
  dplyr::arrange(
    factor(.data$Placement, levels = c("Near-eye", "Chest")),
    .data$`Time grid and summary`
  )

state_support_display <- state_candidates |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(
        metric = .data$metric_id,
        `Diary-period metric` = .data$manuscript_name
      ),
    by = "metric"
  ) |>
  dplyr::mutate(
    Placement = placement_label(.data$position),
    cutoff = paste0(round(100 * .data$candidate_state_support_cutoff), "%")
  ) |>
  dplyr::select(
    .data$Placement,
    .data$`Diary-period metric`,
    `Eligible days` = .data$eligible_participant_days,
    .data$cutoff,
    retained = .data$retained_metric_instances,
    retention = .data$retained_fraction
  ) |>
  tidyr::pivot_wider(names_from = .data$cutoff, values_from = .data$retained) |>
  dplyr::left_join(
    state_candidates |>
      dplyr::filter(.data$candidate_state_support_cutoff == 0.80) |>
      dplyr::left_join(
        metric_registry |>
          dplyr::select(
            metric = .data$metric_id,
            `Diary-period metric` = .data$manuscript_name
          ),
        by = "metric"
      ) |>
      dplyr::transmute(
        Placement = placement_label(.data$position),
        .data$`Diary-period metric`,
        `Primary retention` = scales::percent(
          .data$retained_fraction,
          accuracy = 0.1
        )
      ),
    by = c("Placement", "Diary-period metric")
  ) |>
  dplyr::arrange(
    factor(.data$Placement, levels = c("Near-eye", "Chest")),
    .data$`Diary-period metric`
  )

mder_support_display <- mder_candidates |>
  dplyr::mutate(
    Placement = placement_label(.data$position),
    cutoff = paste0(round(100 * .data$candidate_support_cutoff), "%")
  ) |>
  dplyr::select(
    .data$Placement,
    `Eligible days` = .data$eligible_participant_days,
    .data$cutoff,
    retained = .data$retained_participant_days
  ) |>
  tidyr::pivot_wider(names_from = .data$cutoff, values_from = .data$retained) |>
  dplyr::left_join(
    mder_candidates |>
      dplyr::filter(.data$candidate_support_cutoff == 0.80) |>
      dplyr::transmute(
        Placement = placement_label(.data$position),
        `Participants at 80%` = .data$retained_participants,
        `Sites at 80%` = .data$sites_with_retained_days,
        `Primary retention` = scales::percent(
          .data$retained_fraction,
          accuracy = 0.1
        )
      ),
    by = "Placement"
  ) |>
  dplyr::arrange(factor(.data$Placement, levels = c("Near-eye", "Chest")))

availability_summary <- admissibility |>
  dplyr::filter(.data$metric %in% metric_registry$metric_id) |>
  dplyr::group_by(.data$placement, .data$analysis_unit, .data$metric) |>
  dplyr::summarise(
    possible = dplyr::n(),
    calculated = sum(.data$estimable),
    missing = sum(!.data$estimable),
    .groups = "drop"
  ) |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(-.data$analysis_unit),
    by = c("metric" = "metric_id")
  ) |>
  dplyr::mutate(
    Placement = placement_label(.data$placement),
    `Metric family` = dplyr::recode(
      .data$manuscript_category,
      `level-based` = "Level",
      `duration-based` = "Duration",
      `timing-based` = "Timing",
      `exposure-history-based` = "Exposure history",
      `spectrum-based` = "Spectral composition",
      `dynamics-based` = "Temporal dynamics"
    ),
    Metric = dplyr::case_when(
      .data$metric %in% c(
        "dose_observed_medi",
        "dose_time_sensitive_corrected_medi",
        "longest_bout_above_250_exact_only_sensitivity"
      ) ~ paste0(.data$manuscript_name, " — ", .data$variant_label),
      TRUE ~ .data$manuscript_name
    ),
    `One value per` = dplyr::recode(
      .data$analysis_unit,
      participant_day = "Participant-day",
      participant = "Participant"
    ),
    availability = format_count_fraction(.data$calculated, .data$possible)
  ) |>
  dplyr::select(
    .data$`Metric family`,
    .data$Metric,
    .data$`One value per`,
    .data$Placement,
    .data$availability
  ) |>
  tidyr::pivot_wider(
    names_from = .data$Placement,
    values_from = .data$availability
  ) |>
  dplyr::mutate(
    `Metric family` = factor(
      .data$`Metric family`,
      levels = c(
        "Level", "Duration", "Timing", "Exposure history",
        "Spectral composition", "Temporal dynamics"
      )
    )
  ) |>
  dplyr::arrange(.data$`Metric family`, .data$Metric)

gap_summary <- gap_data |>
  dplyr::group_by(.data$placement) |>
  dplyr::summarise(
    `Participant-days` = dplyr::n(),
    `Expected actual-time hours` = sum(.data$expected_real_minutes) / 60,
    `Valid melEDI hours` = sum(.data$valid_medi_real_minutes) / 60,
    `Missing melEDI hours` = sum(.data$missing_medi_real_minutes) / 60,
    `Continuous missing periods` = sum(.data$missing_medi_runs),
    `Longest missing period (min)` = max(
      .data$maximum_missing_medi_run_minutes
    ),
    `Unknown diary-state hours` = sum(
      .data$unknown_diary_state_real_minutes
    ) / 60,
    `Repeated autumn minutes` = sum(.data$dst_fold_real_minutes),
    `Absent spring minutes` = sum(
      .data$structural_nonexistent_wall_minutes
    ),
    .groups = "drop"
  ) |>
  dplyr::mutate(Placement = placement_label(.data$placement)) |>
  dplyr::select(-.data$placement) |>
  dplyr::relocate(.data$Placement) |>
  dplyr::arrange(factor(.data$Placement, levels = c("Near-eye", "Chest")))

gap_support_summary <- gap_summary |>
  dplyr::select(
    .data$Placement,
    .data$`Participant-days`,
    .data$`Expected actual-time hours`,
    .data$`Valid melEDI hours`,
    .data$`Missing melEDI hours`,
    .data$`Continuous missing periods`,
    .data$`Longest missing period (min)`,
    .data$`Unknown diary-state hours`
  )

daylight_saving_summary <- gap_summary |>
  dplyr::select(
    .data$Placement,
    .data$`Participant-days`,
    .data$`Repeated autumn minutes`,
    .data$`Absent spring minutes`
  )

path_hash_markdown <- function(path, sha256) {
  paste0(
    wrap_provenance_path(relative_path(path)),
    "<br>SHA-256: ",
    wrap_hash(sha256)
  )
}

coverage_lines <- settings |>
  dplyr::arrange(factor(.data$placement, levels = c("glasses", "chest"))) |>
  dplyr::transmute(
    line = paste0(
      "**", placement_label(.data$placement), ":** ",
      path_hash_markdown(.data$coverage_path, .data$coverage_sha256)
    )
  ) |>
  dplyr::pull(.data$line) |>
  paste(collapse = "<br><br>")

first_settings <- settings[1, , drop = FALSE]
input_display <- tibble::tibble(
  `Input and exact stored artifact` = c(
    paste0("**Rule A one-minute coverage**<br>", coverage_lines),
    paste0(
      "**Sleep and wear intervals**<br>",
      path_hash_markdown(
        first_settings$state_interval_manifest_path,
        first_settings$state_interval_manifest_sha256
      ),
      "<br>", nrow(state_inputs), " site-by-interval files are listed in ",
      "artifacts/05_metrics/metric_state_interval_inputs.csv."
    ),
    paste0(
      "**Fixed intensity profiles**<br>",
      path_hash_markdown(
        first_settings$reference_profiles_path,
        first_settings$reference_profiles_sha256
      )
    ),
    paste0(
      "**Probability of melEDI >250 lx by time of day**<br>",
      path_hash_markdown(
        first_settings$timing_exceedance_distributions_path,
        first_settings$timing_exceedance_distributions_sha256
      )
    ),
    paste0(
      "**Metric-specific relevance maps**<br>",
      path_hash_markdown(
        first_settings$relevance_maps_path,
        first_settings$relevance_maps_sha256
      )
    )
  ),
  `How Preparation 04 uses it` = c(
    paste(
      "Provides the eligible one-minute near-eye and chest records from",
      "Preparation 02. Eligibility has already applied the 50%-per-hour",
      "and 80%-per-day rules; hourly support affects hourly summaries only."
    ),
    paste(
      "Projects diary-defined wake, the three hours before sleep, sleep,",
      "wear, and bedside context onto the complete actual-time minute grid",
      "before light values are joined."
    ),
    paste(
      "Provides the fixed participant-balanced melEDI and illuminance",
      "profiles learned in Preparation 03. They are not relearned for a",
      "participant-day."
    ),
    paste(
      "Checks whether observed minutes cover the times when exposure above",
      "250 lx is expected; it never weights or changes an observed timing."
    ),
    paste(
      "Checks time-of-day support for dose, M10, L10, timing, and MDER.",
      "Only the approved dose correction may change a metric value."
    )
  )
)

script_map <- tibble::tibble(
  `Step and producing code` = c(
    paste0(
      "**1. Screen paired-channel support**<br>",
      "scripts/pipeline/build_mder_support_gate.R<br>",
      "scripts/pipeline/mder_support_gate.R"
    ),
    paste0(
      "**2. Reconstruct diary and wear state on the minute grid**<br>",
      "scripts/pipeline/build_metric_derivation.R<br>",
      "scripts/pipeline/state_interval_projection.R"
    ),
    paste0(
      "**3. Derive metrics and metric-specific availability**<br>",
      "scripts/pipeline/metric_derivation.R<br>",
      "scripts/pipeline/time_support.R"
    ),
    paste0(
      "**4. Store values, diagnostics, and inventories**<br>",
      "scripts/pipeline/build_metric_derivation.R"
    )
  ),
  `Reads, operation, separation, and output` = c(
    paste(
      "Reads Preparation 02 coverage plus Preparation 03 melEDI and",
      "illuminance profiles. It classifies support at the fixed 70%, 80%,",
      "and 90% cutoffs without calculating MDER. This separate step made the",
      "80% choice inspectable before any ratio or hypothesis result existed.",
      "It produced the five-file MDER support-gate bundle."
    ),
    paste(
      "Reads the exact sleep and wear interval files from Preparation 01.",
      "It applies half-open [start, end) intervals in actual UTC time, then",
      "checks state, wear, and measurement context where an original record",
      "exists. Keeping this projection separate prevents a missing light row",
      "from erasing a known diary period."
    ),
    paste(
      "Reads the reconstructed minute grid, eligible light values, and fixed",
      "profiles. It calculates daily, diary-period, window, timing, spectral,",
      "and repeated-day metrics; records why any one value is unavailable;",
      "and retains the participant-day for unrelated metrics. Time-support",
      "functions isolate gap, window, circular-time, and boundary rules."
    ),
    paste(
      "Writes participant-day and participant R objects, complete 30-minute",
      "and hourly grids, rectangular CSV mirrors, availability and gap",
      "diagnostics, settings, exact input inventories, and the 29-file metric",
      "manifest. This storage step is the hand-off to Preparation 06."
    )
  )
)

core_rules <- tibble::tibble(
  `Scientific operation` = c(
    "Measurement represented",
    "Actual time and local clock",
    "Thirty-minute and hourly values",
    "Thresholds and missing minutes",
    "Diary-period metrics",
    "Dose and MDER",
    "Brightest and darkest 10 hours",
    "Timing and longest continuous period",
    "Interdaily stability and intradaily variability"
  ),
  `Rule and reason` = c(
    paste(
      "The full day is a placement-specific hybrid: near-eye or chest",
      "measurement while awake, plus bedside sleep-environment measurement",
      "during diary-defined sleep. It is not continuous 24-hour ocular",
      "exposure."
    ),
    paste(
      "Actual UTC time preserves order, elapsed duration, gaps, adjacency,",
      "and both repeated autumn hours. Local clock aligns time-of-day bins,",
      "M10/L10, timing, and interdaily stability."
    ),
    paste(
      "A 30-minute arithmetic mean needs 15 of 30 valid minutes. An hourly",
      "zero-aware geometric mean needs 30 of 60 and uses a 0.1 lx offset so",
      "true zeros remain valid. Unsupported bins stay as missing rows."
    ),
    paste(
      "Comparisons are strictly >1,000, >250, <10, and <1 lx melEDI.",
      "Missing or invalid minutes are never converted to zero or darkness and",
      "never bridge a continuous period."
    ),
    paste(
      "Wake, pre-sleep, and sleep values require 80% support in that specific",
      "diary period; 70% and 90% are fixed sensitivities. An incomplete diary",
      "domain, absent period, or inadequate support affects only that metric."
    ),
    paste(
      "Observed dose integrates actual durations. Only dose may be corrected",
      "for when missing data occurred, with at least 80% profile relevance and",
      "a maximum factor of 1.25. MDER is an unscaled ratio of paired melEDI",
      "and illuminance integrals and requires all three 80% support checks."
    ),
    paste(
      "M10 tests all 841 non-wrapping one-minute starts; L10 tests all 1,440",
      "starts and may cross midnight. Each window is 600 minutes, uses at",
      "least 80% ordinary and profile support, and returns a geometric mean",
      "and midpoint. No L5 metric is produced."
    ),
    paste(
      "First, last, and mean timing above 250 lx require sufficient support",
      "where such events are expected. The longest observed uninterrupted",
      "period is a lower bound; a possible upper bound and exact-only",
      "sensitivity preserve uncertainty caused by missing data."
    ),
    paste(
      "Both need at least three eligible days. Interdaily stability needs 20",
      "clock hours represented on at least two days; intradaily variability",
      "needs 24 truly adjacent hourly pairs across at least three days."
    )
  )
)

output_file_lines <- function(types) {
  selected <- metric_manifest |>
    dplyr::filter(.data$artifact_type %in% types) |>
    dplyr::arrange(.data$placement, .data$path)
  paste(
    wrap_provenance_path(relative_path(selected$path)),
    collapse = "<br>"
  )
}

output_display <- tibble::tibble(
  `Stored layer` = c(
    "Participant-day metrics",
    "Participant-level IS and IV",
    "Complete 30-minute grid",
    "Complete hourly grid",
    "Availability, gap, support, and uncertainty records"
  ),
  `Role in the next step` = c(
    "Daily metric values supplied to Preparation 06",
    "Repeated-day metrics supplied to Preparation 06",
    "Thirty-minute analysis rows supplied to Preparation 06",
    "Hourly analysis rows supplied to Preparation 06",
    "Audit and sample-flow evidence; not substituted for metric values"
  ),
  `Exact stored artifacts` = c(
    output_file_lines("participant_day_metrics_rds"),
    output_file_lines("participant_is_iv_rds"),
    output_file_lines("complete_30_minute_arithmetic_medi_grid_rds"),
    output_file_lines("complete_one_hour_geometric_medi_grid_rds"),
    paste(
      wrap_provenance_path(
        c(
          "artifacts/05_metrics/metrics_glasses_admissibility.csv",
          "artifacts/05_metrics/metrics_chest_admissibility.csv",
          "artifacts/05_metrics/metrics_glasses_gap_diagnostics.csv",
          "artifacts/05_metrics/metrics_chest_gap_diagnostics.csv",
          "artifacts/05_metrics/metrics_glasses_support_diagnostics.csv",
          "artifacts/05_metrics/metrics_chest_support_diagnostics.csv"
        )
      ),
      collapse = "<br>"
    )
  )
)

verification_history <- tibble::tibble(
  `Artifact version` = c(
    paste0(
      "**Current metric outputs**<br>",
      wrap_hash(current_metric_manifest_sha256)
    ),
    paste0(
      "**Earlier complete metric reconstruction**<br>",
      wrap_hash(earlier_metric_manifest_sha256)
    ),
    paste0(
      "**Current MDER and diary-period support gates**<br>",
      wrap_hash(current_mder_manifest_sha256),
      "<br>", wrap_hash(current_state_manifest_sha256)
    ),
    paste0(
      "**Earlier complete MDER support reconstruction**<br>",
      wrap_hash(earlier_mder_manifest_sha256)
    )
  ),
  `What is documented for that exact version` = c(
    paste(
      "The stored settings contain 816 near-eye and 902 chest",
      "participant-days. Every file listed by the current manifest is",
      "present and matches its stored size and SHA-256. Complete independent",
      "reconstruction of the current values remains FIND-044."
    ),
    paste(
      "The complete independent metric check applied only to this earlier",
      "manifest and its 811 near-eye and 897 chest participant-days. It must",
      "not be presented as a reconstruction of the current files."
    ),
    paste(
      "The current MDER summary retains 760 of 816 near-eye and 850 of 902",
      "chest participant-days at 80%. Current diary-period counts are shown",
      "above. File identities match their manifests; complete independent",
      "reconstruction of both current classification families remains",
      "FIND-044."
    ),
    paste(
      "The complete independent MDER check applied only to this earlier",
      "manifest: 733 of 811 near-eye and 825 of 897 chest participant-days",
      "were retained at 80%. It must not be transferred to the current gate."
    )
  )
)
```

## Purpose and position in the preparation chain

Preparation 04 converts the eligible one-minute light records from
Preparation 02 into the daily, diary-period, 30-minute, hourly, and
participant-level metrics used by later analyses. It also records whether
each individual metric had enough relevant data to be interpreted. A failed
support check makes only that metric unavailable; it does not discard the
participant-day from unrelated analyses.

The exact inputs are the Preparation 02 Rule A coverage files, the
Preparation 01 sleep and wear intervals, and the fixed time-of-day profiles
and relevance maps learned in Preparation 03. The main outputs are eight R
data files containing participant-day, participant, 30-minute, and hourly
values for the near-eye and chest placements. Preparation 06 combines those
files with site and participant context; this report does not inspect any
hypothesis result.

```{mermaid}
flowchart LR
  P1["Preparation 01<br/>sleep and wear intervals"] --> D["Preparation 04<br/>metric derivation"]
  P2["Preparation 02<br/>eligible one-minute records"] --> G["MDER and diary-period<br/>support checks"]
  P2 --> D
  P3["Preparation 03<br/>fixed temporal profiles"] --> G
  P3 --> D
  G --> D
  D --> O["Daily, participant,<br/>30-minute, and hourly metrics"]
  O --> P6["Preparation 06<br/>model-ready data"]
```

::: {.callout-note title="What is calculated during this render?"}
This report reads already stored CSV outputs. During rendering it checks file
presence, byte size, SHA-256 identity, required columns, and fixed row-count
relationships, then calculates only lightweight descriptive totals for the
tables and figure below. It does **not** run a preparation builder, derive or
change a metric, independently reconstruct a support classification, fit a
model, or overwrite an accepted analytical artifact.
:::

::: {.callout-important title="Interpret the full-day metrics as a hybrid measurement"}
During wake and the three hours before sleep, the record represents the
logger's worn placement: near-eye or chest. During diary-defined sleep, it
represents the bedside sleep environment. A metric spanning the full day is
therefore an environmental-light summary of this planned hybrid record, not a
continuous 24-hour measure of ocular exposure.
:::

## Exact inputs

```{r}
#| label: tbl-preparation04-inputs
#| tbl-cap: "Exact stored inputs to metric derivation and the scientific role of each."

prep_gt(
  input_display,
  source_note = paste(
    "Paths and SHA-256 values are read from the current metric settings and",
    "input inventories; no input is rebuilt during rendering."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Input and exact stored artifact` ~ gt::pct(46),
    `How Preparation 04 uses it` ~ gt::pct(54)
  )
```

Preparation 02 had already retained only participant-days with valid melEDI
for at least 80% of the fixed 1,440-minute hybrid day. The 50% requirement
within an hour governs whether an hourly value can be calculated; it does not
remove otherwise valid one-minute measurements from daily or minute-based
metrics. All-finite, all-zero melEDI days had also been removed from the
primary coverage files while remaining available as a predefined sensitivity.

## Production sequence and stored hand-offs

The producing scripts are shown in execution order. They remain separate so
that support decisions, state reconstruction, metric calculation, and file
storage each have an inspectable input and output boundary.

```{r}
#| label: tbl-script-execution-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

prep_gt(
  script_map,
  source_note = paste(
    "The visible report translates each module's role; the paths preserve",
    "the exact production provenance."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

## Scientific rules applied during production

```{r}
#| label: tbl-metric-rules
#| tbl-cap: "Metric construction, temporal support, and missing-data rules applied by the production scripts."

prep_gt(core_rules) |>
  gt::cols_width(
    `Scientific operation` ~ gt::pct(31),
    `Rule and reason` ~ gt::pct(69)
  )
```

The central distinction is between participant-day eligibility and
metric-specific availability. A day that passed the general coverage rule
remains in the prepared data even when, for example, its pre-sleep period is
unknown, its M10 window has insufficient support, or its paired illuminance
data cannot support MDER.

## Cohort and time-grid outputs

```{r}
#| label: tbl-current-metric-sample
#| tbl-cap: "Participants, participant-days, and complete time-grid rows in the current stored metric outputs."

prep_gt(
  sample_summary,
  source_note = paste(
    "Every participant-day contributes 48 planned 30-minute rows and 24",
    "planned hourly rows, including rows whose metric value is unavailable."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$Sites, .data$Participants, .data$`Eligible participant-days`,
      .data$`Thirty-minute rows`, .data$`Hourly rows`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(18),
    Sites ~ gt::pct(12),
    Participants ~ gt::pct(17),
    `Eligible participant-days` ~ gt::pct(19),
    `Thirty-minute rows` ~ gt::pct(17),
    `Hourly rows` ~ gt::pct(17)
  )
```

Near-eye metrics contain 141 participants from nine sites and 816
participant-days. Chest metrics contain 154 participants from eight sites and
902 participant-days. These are the current stored Preparation 04 samples,
not downstream model-specific complete-case samples.

### Diary-period support

Wake, the three hours before sleep, and sleep each use a metric-specific 80%
support rule. The 70% and 90% columns show the fixed sensitivity cutoffs; they
do not change general participant-day eligibility.

```{r}
#| label: tbl-state-support
#| tbl-cap: "Diary-period metric values retained at the fixed 70%, 80%, and 90% support cutoffs."

prep_gt(
  state_support_display,
  source_note = paste(
    "The 80% column is primary. Incomplete diary domains and absent target",
    "periods remain distinct from insufficient observed melEDI support."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$`Eligible days`, .data$`70%`, .data$`80%`, .data$`90%`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    `Diary-period metric` ~ gt::pct(35),
    `Eligible days` ~ gt::pct(11),
    `70%` ~ gt::pct(9),
    `80%` ~ gt::pct(9),
    `90%` ~ gt::pct(9),
    `Primary retention` ~ gt::pct(13)
  )
```

At the primary cutoff, the near-eye output retains 737 wake, 655 pre-sleep,
and 778 sleep-environment values. The chest output retains 818, 743, and 861,
respectively. An unavailable value carries one of three distinct explanations:
the diary could not bound the day's state domains, the target period did not
occur, or the target period contained too little valid melEDI.

### Paired support for MDER

MDER compares melEDI with photopic illuminance over exactly the same valid
minutes. Ordinary paired-minute coverage, melEDI-profile support, and
illuminance-profile support must each meet the cutoff; the paired illuminance
integral must also be positive. The profile checks assess when data are
missing but never scale or weight the ratio.

```{r}
#| label: tbl-mder-support
#| tbl-cap: "Participant-days retained for MDER at the fixed support cutoffs in the current stored gate."

prep_gt(
  mder_support_display,
  source_note = paste(
    "The current stored gate calculated support classifications only; it did",
    "not calculate MDER ratios or hypothesis results."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$`Eligible days`, .data$`70%`, .data$`80%`, .data$`90%`,
      .data$`Participants at 80%`, .data$`Sites at 80%`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    `Eligible days` ~ gt::pct(12),
    `70%` ~ gt::pct(9),
    `80%` ~ gt::pct(9),
    `90%` ~ gt::pct(9),
    `Participants at 80%` ~ gt::pct(15),
    `Sites at 80%` ~ gt::pct(13),
    `Primary retention` ~ gt::pct(19)
  )
```

```{r}
#| label: fig-mder-support-retention
#| fig-cap: "Participant-day retention for MDER at the three fixed paired-data support cutoffs. The 80% cutoff is primary; 70% and 90% are predefined sensitivities. Values are read from the current stored MDER support-gate summary, not recalculated from minute-level data."
#| fig-alt: "Line chart showing MDER participant-day retention at 70%, 80%, and 90% support. Near-eye retention decreases from about 99% to 93% to 78%; chest retention decreases from about 98% to 94% to 80%."
#| fig-width: 7.2
#| fig-height: 4.2

mder_plot_data <- mder_candidates |>
  dplyr::mutate(Placement = placement_label(.data$position))

ggplot(
  mder_plot_data,
  aes(
    x = .data$candidate_support_cutoff,
    y = .data$retained_fraction,
    colour = .data$Placement,
    linetype = .data$Placement,
    shape = .data$Placement,
    group = .data$Placement
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.6, stroke = 0.8) +
  scale_x_continuous(
    breaks = c(0.70, 0.80, 0.90),
    labels = scales::label_percent(accuracy = 1),
    expand = expansion(mult = c(0.04, 0.04))
  ) +
  scale_y_continuous(
    limits = c(0.70, 1.00),
    breaks = seq(0.70, 1.00, by = 0.05),
    labels = scales::label_percent(accuracy = 1),
    expand = expansion(mult = c(0, 0.02))
  ) +
  scale_colour_manual(
    values = c("Near-eye" = "#0072B2", "Chest" = "#D55E00")
  ) +
  scale_linetype_manual(
    values = c("Near-eye" = "solid", "Chest" = "longdash")
  ) +
  scale_shape_manual(values = c("Near-eye" = 16, "Chest" = 17)) +
  labs(
    x = "Required paired-data support",
    y = "Eligible participant-days retained",
    colour = "Placement",
    linetype = "Placement",
    shape = "Placement"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text = element_text(size = 9, colour = "#333333"),
    axis.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    plot.margin = margin(7, 10, 7, 7)
  )
```

[Download the source data for @fig-mder-support-retention](../../artifacts/08_diagnostics/mder_support_gate/mder_support_candidate_summary.csv).

The current 80% gate retains 760 of 816 near-eye participant-days and 850 of
902 chest participant-days. Every near-eye and chest participant remains
represented at the primary cutoff, across nine and eight sites,
respectively.

### Thirty-minute and hourly values

```{r}
#| label: tbl-bin-support
#| tbl-cap: "Stored 30-minute and hourly values with and without enough valid melEDI data."

prep_gt(
  bin_support,
  source_note = paste(
    "A row is retained even when it has too little support to calculate the",
    "requested summary."
  )
) |>
  gt::fmt_number(
    columns = c(
      .data$Values, .data$`Values calculated`, .data$`Too little data`,
      .data$`Valid melEDI minutes`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    `Time grid and summary` ~ gt::pct(31),
    Values ~ gt::pct(12),
    `Values calculated` ~ gt::pct(15),
    `Too little data` ~ gt::pct(13),
    `Valid melEDI minutes` ~ gt::pct(15)
  )
```

The regular grids preserve unsupported periods as explicit missing values.
This matters because deleting rows would compress time and could create false
adjacency, change a window's apparent duration, or hide where coverage was
lost.

## Which metric values are available?

```{r}
#| label: tbl-metric-availability
#| tbl-cap: "Metric values calculated in the current stored outputs, shown as calculated / possible (percentage)."

prep_gt(
  availability_summary,
  groupname_col = "Metric family",
  source_note = paste(
    "Counts describe Preparation 04 availability before any",
    "hypothesis-specific complete-case filtering. Internal M10/L10 onset and",
    "offset diagnostics are not reader-facing analytical metrics."
  )
) |>
  gt::cols_width(
    Metric ~ gt::pct(45),
    `One value per` ~ gt::pct(17),
    `Near-eye` ~ gt::pct(19),
    Chest ~ gt::pct(19)
  )
```

A known absence and an unknown value remain different. For example, an
adequately observed day with no exposure above a threshold has a duration of
zero, but first, last, and mean timing do not exist. For the longest period
above 250 lx, the observed duration remains a lower bound when a missing
interval could conceal a longer period; the separate exact-only sensitivity
is available only when the lower and possible upper bounds agree.

## Gaps and daylight-saving changes

```{r}
#| label: tbl-gap-diagnostics
#| tbl-cap: "Actual-time coverage, missing periods, diary-state uncertainty, and daylight-saving diagnostics in the current metric outputs."

prep_gt(
  gap_support_summary,
  source_note = paste(
    "Expected and missing hours use the actual UTC minute grid. Repeated",
    "autumn and absent spring minutes describe the separate local-clock grid."
  )
) |>
  gt::fmt_number(
    columns = where(is.numeric),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(12),
    `Participant-days` ~ gt::pct(10),
    `Expected actual-time hours` ~ gt::pct(13),
    `Valid melEDI hours` ~ gt::pct(11),
    `Missing melEDI hours` ~ gt::pct(11),
    `Continuous missing periods` ~ gt::pct(12),
    `Longest missing period (min)` ~ gt::pct(11),
    `Unknown diary-state hours` ~ gt::pct(19)
  )
```

```{r}
#| label: tbl-daylight-saving-diagnostics
#| tbl-cap: "Daylight-saving minutes represented on the separate actual-time and local-clock grids."

prep_gt(
  daylight_saving_summary,
  source_note = paste(
    "Repeated autumn minutes remain as two actual-time observations but one",
    "averaged local-clock position. Absent spring minutes remain structural",
    "gaps on the local-clock grid."
  )
) |>
  gt::fmt_number(
    columns = where(is.numeric),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(25),
    `Participant-days` ~ gt::pct(25),
    `Repeated autumn minutes` ~ gt::pct(25),
    `Absent spring minutes` ~ gt::pct(25)
  )
```

Actual-time metrics retain both occurrences of a repeated autumn hour and
respect the true elapsed gap during the spring transition. Clock-aligned
metrics use a fixed 24-hour local-clock grid; repeated autumn values are
averaged only on that grid, and structurally absent spring minutes remain
absent. Missing light, unknown diary state, and a daylight-saving transition
are therefore not treated as the same event.

## Exact artifacts passed forward

```{r}
#| label: tbl-output-artifacts
#| tbl-cap: "Preparation 04 artifacts stored for Preparation 06 and for metric-specific audit."

prep_gt(
  output_display,
  source_note = paste(
    "All listed analytical files were produced by",
    "scripts/pipeline/build_metric_derivation.R; rendering reads them only."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Stored layer` ~ gt::pct(22),
    `Role in the next step` ~ gt::pct(29),
    `Exact stored artifacts` ~ gt::pct(49)
  )
```

The eight R data files in the first four rows are the metric layer used to
construct the primary dataset in Preparation 06. The predefined
**gap-timing-unaware dataset** still passed the general 50%-per-hour and
80%-per-day coverage rules; its name means that the timing of remaining
missing observations was not used for an additional metric-specific
adjustment. It does not mean that gaps or coverage were ignored. For this
first contrast only, the primary dataset could be interpreted as a
**time-sensitive primary metric dataset**, because its approved metric rules
use the timing of remaining missing observations where scientifically
relevant. Hereafter, it is simply the **primary dataset**.

Preparation 04 creates the primary dataset's metric layer. Preparation 05
acquires questionnaire and diary inputs, and Preparation 06 brings the metric
and non-light inputs together. The gap-timing-unaware dataset remains a
predefined prepared-data sensitivity and is not substituted into the primary
files listed above.

## Current file checks and version-specific verification

```{r}
#| label: tbl-current-manifest-validation
#| tbl-cap: "Bounded file-identity checks for the three current Preparation 04 manifests."

manifest_validation_display <- manifest_validation |>
  dplyr::mutate(
    `Manifest path` = wrap_provenance_path(.data$`Manifest path`),
    `Manifest SHA-256` = wrap_hash(.data$`Manifest SHA-256`)
  )

prep_gt(
  manifest_validation_display,
  source_note = paste(
    "PASS means that every listed file exists and matches the stored byte",
    "size and SHA-256. It does not mean that metric values or support",
    "classifications were independently reconstructed during rendering."
  )
) |>
  gt::fmt_markdown(columns = c(.data$`Manifest path`)) |>
  gt::fmt_number(
    columns = c(
      .data$`Files listed`, .data$`Files present`, .data$`Size checks`,
      .data$`SHA-256 checks`
    ),
    decimals = 0
  ) |>
  gt::cols_width(
    Family ~ gt::pct(16),
    `Manifest path` ~ gt::pct(24),
    `Manifest SHA-256` ~ gt::pct(25),
    `Files listed` ~ gt::pct(7),
    `Files present` ~ gt::pct(7),
    `Size checks` ~ gt::pct(7),
    `SHA-256 checks` ~ gt::pct(8),
    Check ~ gt::pct(6)
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(columns = .data$Check)
  )
```

The complete independent-verification records available for Preparation 04
apply to earlier artifact versions. PREP-003 requires those claims to remain
attached to their exact manifests and samples.

```{r}
#| label: tbl-version-specific-verification
#| tbl-cap: "Which verification statements apply to the current and earlier Preparation 04 artifact versions."

prep_gt(
  verification_history,
  source_note = paste(
    "CHG-046 also records the intervening metric manifest",
    wrap_hash(intervening_metric_manifest_sha256),
    "; it is not the current metric manifest."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Artifact version` ~ gt::pct(38),
    `What is documented for that exact version` ~ gt::pct(62)
  )
```

Under [PREP-003](../../audit/decisions/preparation04_version_specific_verification.md),
the earlier complete metric reconstruction applies only to metric manifest
`ac0325716421f2d90bb83b445ffb6ad965b131e382373e45f4ad6157eac5937d`
and its 811 near-eye and 897 chest participant-days. The earlier complete MDER
support reconstruction applies only to manifest
`c4ecfab41892e44b38dd78097277ccef5f94a379b44c7b309de281e9215a7b3c`
and its 733-of-811 near-eye and 825-of-897 chest retention at 80%.

Those results must not be transferred to the three current manifests. Exact
independent reconstruction of the current metric values, MDER classifications,
and diary-period classifications remains open audit item **FIND-044**. This
provenance gap is not evidence that the current artifacts or any downstream
result is wrong. It states precisely which scientific check has not yet been
documented for these exact file versions; this bounded reader-facing render
does not attempt to perform that check.
