---
title: "Preparation 04: Calculate light-exposure metrics and check data sufficiency"
execute:
  echo: true
  warning: true
  message: true
  error: false
  freeze: false
---

## Purpose

This notebook calculates the approved personal light-exposure metrics. A
participant-day means one participant observed on one calendar day. For every
participant-day and metric, the notebook records whether enough relevant data
were available to calculate the value reliably. If one time period, diary
window, timing value, ratio, or corrected dose cannot be calculated, that day
remains available for all other metrics.

The full-day metrics describe an intentional, placement-specific hybrid
record. For glasses, this combines near-eye measurements during wake and
pre-sleep with bedside sleep-environment measurements during the sleep-diary
window. For chest loggers, the waking component is chest-level rather than
near-eye, while the sleep component remains the bedside sleep environment.
Neither construct is described as continuous 24-hour ocular exposure.

## What this step uses, produces, and checks

- Inputs are the exact coverage files from Preparation 02, the sleep and wear
  interval files from Preparation 01, and the fixed profiles and time-of-day
  relevance maps learned in Preparation 03. SHA-256 checksums, which are
  digital fingerprints, identify the exact input files and confirm that they
  have not changed.
- Preparation 04 applies the Preparation 01 sleep and wear intervals to every
  minute ordered in Coordinated Universal Time (UTC) before joining the light
  values that passed coverage. UTC preserves the true order and elapsed time
  of observations across daylight-saving changes. Intervals use `[start, end)`
  bounds: the start is included and the end is excluded. The earlier state
  labels are not simply copied. Wherever Preparation 01 contained a source
  measurement, the recalculated diary period, wear record, and measurement
  context must match it exactly.
- Before calculating any metric, Preparation 04 verifies that Preparation 02
  used approved primary Rule A: valid melEDI for at least 80% of the full
  1,440-minute hybrid measurement day, including bedside measurements during
  diary-defined sleep. A local-clock hour needs at least 50% valid melEDI only
  when an hourly value is calculated. A sparse hour does not remove its valid
  one-minute measurements from daily or other minute-based metrics.
  Preparation 02 also excludes an otherwise passing participant-day when every
  finite one-minute melEDI value is exactly 0 lx. The original values remain
  auditable, and an all-zero-inclusive sensitivity is retained.
- Actual UTC time determines the expected minute sequence, elapsed durations,
  dose, continuous periods above threshold, gaps, and which observations are
  adjacent for intradaily variability (IV). It also determines the matched
  melEDI and illuminance values whose exposure is accumulated over time for
  MDER.
- The shared local-clock scale determines the 30-minute and hourly
  time-of-day values, the brightest and darkest 10-hour means (M10 and L10),
  timing metrics, and interdaily stability (IS). M10 and L10 test a 600-minute
  window beginning at every eligible local-clock minute. Their fixed
  30-minute relevance profiles are expanded to one-minute values without
  being relearned. Repeated minutes during the autumn daylight-saving change
  are averaged only on this local-clock scale; both actual observations remain
  available to metrics based on elapsed time.
- Melanopic equivalent daylight illuminance (melEDI) is the reader-facing name
  for the light quantity stored as `MEDI` in the analysis files. A 30-minute
  arithmetic-mean melEDI value requires at least 15 valid
  local-clock minutes. A one-hour zero-aware geometric-mean melEDI value
  requires at least 30 valid local-clock minutes. The geometric mean adds
  0.1 lx before taking logarithms and subtracts it afterwards, which allows
  true zero values to be included. A period with too little data is missing
  without removing the participant-day.
- Threshold endpoints are strict: greater than 1,000 lx, greater than 250 lx,
  less than 10 lx, and less than 1 lx.
- Only dose can be corrected for missing values using their time-of-day
  relevance. For each minute, Preparation 03 learned how much that time
  normally contributes to daily melEDI. The represented fraction is the sum
  of those learned contributions for observed minutes divided by their sum
  across the full day. Correction is allowed only when this fraction is at
  least 80% and is capped at a factor of 1.25.
  Means, durations, continuous periods, timings, 10-hour windows, and MDER are
  not scaled.
- The melanopic daylight efficacy ratio (MDER) is the ratio of melEDI and
  illuminance exposure accumulated over exactly the same observed minutes.
  Three requirements must each reach 80%: the fraction of expected minutes
  with both values present; the fraction of the learned daily melEDI profile
  represented by those minutes; and the equivalent fraction of the learned
  illuminance profile. The accumulated illuminance must also be positive.
  MDER is not scaled or weighted by a profile.
- For each time of day, Preparation 03 learned the probability that melEDI
  exceeds 250 lx. First, last, and mean timing above 250 lx use this fixed
  probability distribution only to check whether observed minutes represent
  enough of the times when an event would be expected. The distribution does
  not change or weight the observed timing value.
- The 10 darkest hours wrap across midnight. No five-hour darkest-window
  metric or output is permitted.
- The full-day duration above 250 lx is reported descriptively but is not used
  to test a preregistered hypothesis.
- The main longest-period value is the longest observed uninterrupted period
  above 250 lx and is interpreted as a lower bound: missing or invalid minutes
  break an observed period, so the true period may have been longer. The
  possible upper bound, a flag showing whether missing data could hide a
  longer period, a sensitivity restricted to exactly known periods, onset and
  offset of the selected period, and contact with the day boundary remain in
  the outputs. If several periods share the maximum duration, the earliest
  onset and then the earliest offset select the reported period. Separate
  checks distinguish boundary contact by that selected period from boundary
  contact by any tied maximum period. Uncertainty about the exact duration
  never removes the participant-day.

## Minimum data required for wake, pre-sleep, and sleep metrics

A wake, pre-sleep, or sleep metric requires valid melEDI for at least 80% of
that diary-defined period. The fixed 70% and 90% alternatives are separate,
metric-specific sensitivity analyses. The technical decision record is
`STATE-005`.

Failure affects only the metric for that diary period. It never excludes the
participant-day or its full-day metrics. The output also reports retained and
missing counts at all three fixed cutoffs. The cutoff was approved from checks
of data support before final state-duration values or hypothesis results were
inspected.

The output distinguishes three reasons why a diary-period metric can be
missing: the diary cannot identify the period across that day; the day has no
such period; or the period exists but contains too little valid melEDI. The
technical output retains the corresponding reason codes
`incomplete_state_domain`, `no_state_window`, and
`insufficient_state_support`. Each failure remains specific to the affected
metric; the participant-day and all full-day metrics are retained.

## Minimum paired data required for MDER

MDER is calculated only when each of the three requirements described above
reaches 80% and accumulated illuminance over the matched minutes is positive.
Otherwise, MDER is missing with a recorded reason; the participant-day and its
other metrics remain available. Fixed 70% and 90% scenarios are sensitivity
analyses. The learned profiles only check whether the relevant times were
observed and never scale, weight, or correct the ratio. The technical decision
record is `METRIC-003`.

The exact retained counts at the three fixed cutoffs are recalculated and
independently verified below from the revised full-day data. This comparison
checks only whether enough paired melEDI and illuminance data are available;
it does not calculate MDER itself. The complete methods, input fingerprints,
and checks are recorded in
[the MDER data-sufficiency audit](../../audit/findings/mder_support_cutoff_gate.md).

## Minimum data required for timing above 250 lx

Timing support is not based on the amount by which the median melEDI profile
exceeds 250 lx: a population median can remain below 250 lx even when
individual participant-days cross that threshold. Preparation 04 instead uses
the distribution learned in Preparation 03 for the probability of being
strictly above 250 lx at each time. That distribution gives equal influence
first to days within a participant and then to participants. It is fixed at
30-minute resolution and each learned value is assigned to its 30
local-clock minutes.

The primary cutoff is 80%. For a given timing metric, the numerator is the sum
of the learned exceedance probabilities for observed minutes in its search
period; the denominator is their sum for all expected minutes in that same
period. The fixed 70% and 90% scenarios change only this cutoff and reuse the
same learned distribution; they do not relearn a profile or change an observed
value. Data across the whole day determine support for mean timing. Data in the
early and late search periods determine support for first and last timing,
respectively. A timing value with too little support is missing with a recorded
reason, while its participant-day and unrelated metrics remain available.

## How the brightest and darkest 10-hour periods are found

M10 and L10 each use exactly 600 consecutive one-minute local-clock intervals.
M10 tests all 841 possible starts from minute 0 through minute 840, so its
10-hour period remains within the same calendar day. L10 tests all 1,440
possible starts and may cross midnight. The fixed 30-minute M10/L10 relevance
profiles are assigned to their one-minute periods; learning the profiles at
30-minute resolution does not restrict the candidate start times. L10 is
retained and L5 is not calculated.

## How missing data affect the longest period above 250 lx

For the longest period above 250 lx, missing or invalid minutes, gaps in
actual time, changes in measurement context, and the day boundary break an
observed period. The main output is the longest observed period and is labelled
as a lower bound. A separate possible upper bound treats every missing or
invalid minute as above 250 lx while retaining the same rules for consecutive
minutes and measurement context.

When the possible upper bound exceeds the observed lower bound, the lower
bound remains available with a flag showing that missing data could hide a
longer qualifying period. The sensitivity analysis restricted to exactly
known durations is then missing; it is reported only when both bounds agree.
Therefore, an observed lower bound of zero on a partly unobserved day is not
interpreted as proof that no qualifying period occurred. No day is removed
solely because its exact longest-period duration is uncertain.

If several observed periods share the maximum duration, the reported onset
and offset come from the period with the earliest onset and then the earliest
offset. One check records whether that selected period touches the day
boundary; a separate check records whether any tied maximum period touches
the boundary.

## Setup

```{r}
#| label: setup-metric-derivation

root_hint <- Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd())
source(file.path(root_hint, "scripts/pipeline/paths_io.R"))
root <- project_root(root_hint)

dependencies <- c(
  "assertions.R",
  "state_alignment.R",
  "time_axes.R",
  "time_support.R",
  "reference_profiles.R",
  "state_interval_projection.R",
  "mder_support_gate.R",
  "build_mder_support_gate.R",
  "verify_mder_support_gate_artifacts.R",
  "metric_derivation.R",
  "build_metric_derivation.R",
  "metric_display_registry.R"
)
for (dependency in dependencies) {
  source(file.path(root, "scripts", "pipeline", dependency))
}

metric_display_registry <- read_metric_display_registry(root)
metric_display_names <- stats::setNames(
  metric_display_registry$manuscript_name,
  metric_display_registry$metric_id
)
metric_display_names <- c(
  metric_display_names,
  m10_onset = "M10 onset (calculation check only)",
  m10_offset = "M10 offset (calculation check only)",
  l10_onset = "L10 onset (calculation check only)",
  l10_offset = "L10 offset (calculation check only)"
)
display_metric_name <- function(metric) {
  display_name <- unname(metric_display_names[metric])
  if (anyNA(display_name)) {
    stop(
      "A reader-facing metric name is missing for: ",
      paste(unique(metric[is.na(display_name)]), collapse = ", ")
    )
  }
  display_name
}

run_label <- Sys.getenv(
  "NATHEALTH_METRIC_RUN_LABEL",
  unset = "full"
)
coverage_root_argument <- Sys.getenv(
  "NATHEALTH_COVERAGE_RUN_ROOT",
  unset = ""
)
state_interval_root_argument <- Sys.getenv(
  "NATHEALTH_STATE_INTERVAL_RUN_ROOT",
  unset = ""
)
state_interval_manifest_argument <- Sys.getenv(
  "NATHEALTH_STATE_INTERVAL_MANIFEST",
  unset = ""
)
profile_root_argument <- Sys.getenv(
  "NATHEALTH_PROFILE_RUN_ROOT",
  unset = ""
)
metric_root_argument <- Sys.getenv(
  "NATHEALTH_METRIC_RUN_ROOT",
  unset = ""
)
placement_argument <- Sys.getenv(
  "NATHEALTH_METRIC_PLACEMENTS",
  unset = ""
)
placements <- if (nzchar(placement_argument)) {
  strsplit(placement_argument, ",", fixed = TRUE)[[1L]]
} else {
  NULL
}
profile_variant <- Sys.getenv(
  "NATHEALTH_METRIC_PROFILE_VARIANT",
  unset = "pooled"
)
state_support_argument <- Sys.getenv(
  "NATHEALTH_STATE_SUPPORT_CUTOFF",
  unset = ""
)
state_support_cutoff <- if (nzchar(state_support_argument)) {
  suppressWarnings(as.numeric(state_support_argument))
} else if (identical(run_label, "full")) {
  0.80
} else {
  NA_real_
}
if (nzchar(state_support_argument) && !is.finite(state_support_cutoff)) {
  abort_pipeline("NATHEALTH_STATE_SUPPORT_CUTOFF must be numeric")
}
mder_support_argument <- Sys.getenv(
  "NATHEALTH_MDER_SUPPORT_CUTOFF",
  unset = ""
)
mder_support_cutoff <- if (nzchar(mder_support_argument)) {
  suppressWarnings(as.numeric(mder_support_argument))
} else if (identical(run_label, "full")) {
  0.80
} else {
  NA_real_
}
if (nzchar(mder_support_argument) && !is.finite(mder_support_cutoff)) {
  abort_pipeline("NATHEALTH_MDER_SUPPORT_CUTOFF must be numeric")
}
```

The main and sensitivity outputs are stored separately and all reuse the fixed
profiles learned in Preparation 03. The main calculation uses 80% for
diary-period and MDER data sufficiency. Separate 70% and 90% sensitivity
calculations change only the relevant cutoff. Site-specific and
leave-one-site-out sensitivities likewise reuse the fixed profiles; they do
not learn new profiles after results are known.

## Participant-days retained at each candidate cutoff

```{r}
#| label: build-state-support-gate

state_support_gate <- build_state_support_gate(
  root = root,
  run_label = run_label,
  coverage_run_root = if (nzchar(coverage_root_argument)) {
    coverage_root_argument
  } else {
    NULL
  },
  state_interval_run_root = if (nzchar(state_interval_root_argument)) {
    state_interval_root_argument
  } else {
    NULL
  },
  state_interval_manifest_path = if (
    nzchar(state_interval_manifest_argument)
  ) {
    state_interval_manifest_argument
  } else {
    NULL
  },
  metric_run_root = if (nzchar(metric_root_argument)) {
    metric_root_argument
  } else {
    NULL
  },
  placements = placements,
  profile_variant = profile_variant,
  candidate_cutoffs = c(0.70, 0.80, 0.90)
)
```

```{r}
#| label: tbl-state-support-gate
#| tbl-cap: "State-specific metric values retained at each candidate data-coverage cutoff, calculated before 80% was selected."

state_candidate_display <- function(data) {
  data |>
    dplyr::transmute(
      Placement = dplyr::recode(
        .data$position,
        glasses = "Near-eye",
        chest = "Chest"
      ),
      Metric = display_metric_name(.data$metric),
      `Coverage cutoff` = paste0(
        round(100 * .data$candidate_state_support_cutoff),
        "%"
      ),
      `Eligible days` = .data$eligible_participant_days,
      `Values retained` = .data$retained_metric_instances,
      `Retained (%)` = round(100 * .data$retained_fraction, 1),
      `Incomplete diary period` =
        .data$incomplete_state_domain_instances,
      `No diary period` = .data$no_state_window_instances,
      `Below coverage cutoff` =
        .data$insufficient_state_support_instances
    )
}
knitr::kable(
  state_candidate_display(state_support_gate$candidates)
)
```

This table is created without selecting a cutoff or calculating a final
state-specific metric. It informed the author's decision and must not be used
to choose the cutoff that preserves a preferred estimate or sample size.
The complete candidate table, coverage quantiles, participant counts,
additional losses at each stricter cutoff, and results by site are written to
`state_support_candidate_diagnostics.csv` and
`state_support_candidate_by_site.csv`.

## Participant-days retained for MDER at each candidate cutoff

```{r}
#| label: build-mder-support-gate

mder_support_gate <- build_mder_support_gate(
  root = root,
  run_label = run_label,
  coverage_run_root = if (nzchar(coverage_root_argument)) {
    coverage_root_argument
  } else {
    NULL
  },
  profile_run_root = if (nzchar(profile_root_argument)) {
    profile_root_argument
  } else {
    NULL
  },
  placements = placements,
  candidate_cutoffs = c(0.70, 0.80, 0.90)
)
mder_support_verification <- verify_mder_support_gate_artifacts(
  root = root,
  run_label = run_label,
  coverage_run_root = if (nzchar(coverage_root_argument)) {
    coverage_root_argument
  } else {
    NULL
  },
  profile_run_root = if (nzchar(profile_root_argument)) {
    profile_root_argument
  } else {
    NULL
  }
)
stopifnot(identical(mder_support_verification$status, "PASS"))
```

```{r}
#| label: tbl-mder-support-gate
#| tbl-cap: "Participant-days with enough paired melEDI and illuminance data for MDER at each fixed cutoff."

mder_candidate_display <- mder_support_gate$candidates |>
  dplyr::mutate(
    Placement = dplyr::recode(
      .data$position,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    `Coverage cutoff` = paste0(
      round(100 * .data$candidate_support_cutoff),
      "%"
    )
  ) |>
  dplyr::select(
    .data$Placement,
    .data$`Coverage cutoff`,
    `Eligible participant-days` = .data$eligible_participant_days,
    `Participant-days retained` = .data$retained_participant_days,
    `Participants retained` = .data$retained_participants,
    `Sites represented` = .data$sites_with_retained_days,
    `No paired measurements` = .data$no_paired_observation_days,
    `Non-positive illuminance exposure` =
      .data$nonpositive_paired_light_integral_days,
    `Below at least one coverage requirement` =
      .data$below_any_support_days
  )
knitr::kable(mder_candidate_display)
```

The independent check reproduced the daily classifications and all summaries
from the verified one-minute coverage files and the fixed time-of-day
profiles. The 80% row is used for the main MDER calculation; the 70% and 90%
rows define the fixed sensitivity analyses.

## Derive metrics

```{r}
#| label: build-metric-derivation

if (!is.finite(state_support_cutoff)) {
  abort_pipeline(
    paste0(
      "The approved primary cutoff is 0.80. Set ",
      "NATHEALTH_STATE_SUPPORT_CUTOFF explicitly for a namespaced ",
      "non-full sensitivity run."
    )
  )
}
if (!is.finite(mder_support_cutoff)) {
  abort_pipeline(
    paste0(
      "The approved primary MDER cutoff is 0.80. Set ",
      "NATHEALTH_MDER_SUPPORT_CUTOFF explicitly for a namespaced ",
      "non-full sensitivity run."
    )
  )
}
metric_build <- build_metric_derivation(
  root = root,
  run_label = run_label,
  coverage_run_root = if (nzchar(coverage_root_argument)) {
    coverage_root_argument
  } else {
    NULL
  },
  state_interval_run_root = if (nzchar(state_interval_root_argument)) {
    state_interval_root_argument
  } else {
    NULL
  },
  state_interval_manifest_path = if (
    nzchar(state_interval_manifest_argument)
  ) {
    state_interval_manifest_argument
  } else {
    NULL
  },
  profile_run_root = if (nzchar(profile_root_argument)) {
    profile_root_argument
  } else {
    NULL
  },
  metric_run_root = if (nzchar(metric_root_argument)) {
    metric_root_argument
  } else {
    NULL
  },
  placements = placements,
  profile_variant = profile_variant,
  zero_offset = 0.1,
  minimum_window_support = 0.80,
  minimum_relevance_support = 0.80,
  minimum_state_support = state_support_cutoff,
  minimum_mder_support = mder_support_cutoff,
  provisional_state_support = FALSE,
  synthetic_test = FALSE,
  minimum_circular_resultant = 0.10
)
```

This is the `r metric_build$run_label` metric run. Technical input inventories
identify the exact Preparation 01 sleep and wear intervals, the Preparation 02
coverage file, and both Preparation 03 profile files. Too little relevant data
can make one metric missing but cannot change whether the participant-day
passed the overall coverage rule.

## Settings and retained time-of-day values

```{r}
#| label: tbl-metric-settings
#| tbl-cap: "Settings and size of each metric output."

metric_settings_summary <- purrr::map_dfr(
  seq_len(nrow(metric_build$settings)),
  function(row) {
    settings <- metric_build$settings[row, , drop = FALSE]
    tibble::tibble(
      Placement = dplyr::recode(
        settings$placement,
        glasses = "Near-eye",
        chest = "Chest"
      ),
      Setting = c(
        "Profile version",
        "Minimum hourly coverage",
        "Minimum daily coverage",
        "Minimum data in a diary period",
        "Minimum paired data for MDER",
        "Time step used for metric calculation",
        "Offset used for zero-aware geometric means",
        "Sensor operating boundary",
        "Threshold comparisons",
        "Minimum data in a 30-minute value",
        "Minimum data in an hourly value",
        "M10 and L10 duration",
        "Minimum data in an M10 or L10 window",
        "Minimum timing relevance represented",
        "Minimum dose relevance represented",
        "Largest permitted dose correction",
        "M10 may cross midnight",
        "L10 may cross midnight",
        "L5 calculated",
        "Primary longest-period result",
        "Measurement represented",
        "Participants",
        "Eligible participant-days",
        "Thirty-minute rows",
        "Hourly rows",
        "Input files verified"
      ),
      Value = as.character(c(
        ifelse(
          settings$profile_variant == "pooled",
          "All sites combined",
          settings$profile_variant
        ),
        paste0(100 * settings$verified_minimum_hour_coverage, "%"),
        paste0(100 * settings$verified_minimum_day_coverage, "%"),
        paste0(100 * settings$state_support_cutoff, "%"),
        paste0(100 * settings$mder_support_cutoff, "%"),
        paste(settings$minute_epoch_seconds, "seconds (one minute)"),
        paste(settings$zero_aware_offset_lx, "lx"),
        "Values below 100,000 lx melEDI",
        "> 1,000; > 250; < 10; < 1 lx melEDI",
        paste(settings$thirty_minute_minimum, "of 30 minutes"),
        paste(settings$hourly_minimum, "of 60 minutes"),
        paste(settings$window_minutes, "minutes"),
        paste0(100 * settings$minimum_window_support, "%"),
        paste0(100 * settings$dose_minimum_relevance_support, "%"),
        paste0(100 * settings$dose_minimum_relevance_support, "%"),
        paste0(
          "Factor of ",
          format(settings$maximum_dose_correction_factor, trim = TRUE)
        ),
        ifelse(settings$m10_wraps_midnight, "Yes", "No"),
        ifelse(settings$l10_wraps_midnight, "Yes", "No"),
        "No",
        "Longest observed uninterrupted period (lower bound)",
        ifelse(
          settings$placement == "glasses",
          paste(
            "Near-eye while awake plus bedside sleep-environment",
            "measurement"
          ),
          paste(
            "Chest while awake plus bedside sleep-environment",
            "measurement"
          )
        ),
        settings$participants,
        settings$eligible_participant_days,
        settings$thirty_minute_rows,
        settings$hourly_rows,
        "Yes"
      ))
    )
  }
) |>
  tidyr::pivot_wider(
    names_from = .data$Placement,
    values_from = .data$Value
  )
knitr::kable(metric_settings_summary)
```

Every eligible participant-day produces exactly 48 rows of 30-minute
arithmetic-mean values and 24 rows of hourly zero-aware geometric-mean values.
A period with too few valid minutes cannot produce a value, but its row and
participant-day remain present. The technical reason code is
`insufficient_bin_support`. Exact input-file fingerprints remain available in
the [metric artifact inventory](../../artifacts/12_manifests/metric_artifacts.csv).

```{r}
#| label: tbl-metric-bin-support
#| tbl-cap: "Thirty-minute and hourly values that had enough valid data to calculate."

bin_support <- purrr::imap_dfr(
  metric_build$results,
  function(result, placement) {
    dplyr::bind_rows(
      result$thirty_minute |>
        dplyr::summarise(
          placement = placement,
          grid = "30_minute_arithmetic",
          rows = dplyr::n(),
          admissible_rows = sum(.data$bin_admissible),
          unsupported_rows = sum(!.data$bin_admissible),
          retained_days = dplyr::n_distinct(
            paste(.data$site, .data$Id, .data$local_date)
          )
        ),
      result$hourly |>
        dplyr::summarise(
          placement = placement,
          grid = "one_hour_zero_aware_geometric",
          rows = dplyr::n(),
          admissible_rows = sum(.data$bin_admissible),
          unsupported_rows = sum(!.data$bin_admissible),
          retained_days = dplyr::n_distinct(
            paste(.data$site, .data$Id, .data$local_date)
          )
      )
    )
  }
) |>
  dplyr::mutate(
    placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    grid = dplyr::recode(
      .data$grid,
      `30_minute_arithmetic` = "30-minute arithmetic mean",
      one_hour_zero_aware_geometric = "Hourly zero-aware geometric mean"
    )
  )
knitr::kable(
  bin_support,
  col.names = c(
    "Placement",
    "Time grid",
    "Values",
    "Values calculated",
    "Values with too little data",
    "Participant-days retained"
  )
)
```

## Metrics that could and could not be calculated

```{r}
#| label: tbl-metric-admissibility
#| tbl-cap: "Number of values that could and could not be calculated for each metric, with reasons."

admissibility_summary <- purrr::imap_dfr(
  metric_build$results,
  function(result, placement) {
    result$metric_values |>
      dplyr::group_by(
        .data$analysis_unit,
        .data$metric,
        .data$state_domain
      ) |>
      dplyr::summarise(
        placement = placement,
        metric_instances = dplyr::n(),
        estimable_instances = sum(.data$estimable),
        metric_na_instances = sum(!.data$estimable),
        .groups = "drop"
      )
  }
) |>
  dplyr::mutate(
    analysis_unit = dplyr::recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    metric = display_metric_name(.data$metric),
    state_domain = dplyr::recode(
      .data$state_domain,
      full_day = "Full 24-hour day",
      wake = "Wake",
      `pre-sleep` = "Three hours before sleep",
      sleep = "Sleep"
    ),
    placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    )
  )
knitr::kable(
  admissibility_summary,
  col.names = c(
    "One value per",
    "Metric",
    "Diary period",
    "Placement",
    "Possible values",
    "Values calculated",
    "Values missing"
  )
)
```

The placement-specific value and data-sufficiency files retain one row for
each possible metric value. When enough data show that melEDI never crossed
the threshold, duration is zero, while first, last, and mean timing do not
exist and are therefore missing. The technical reason code is
`no_threshold_event`. If missing data could hide a longer qualifying period,
the observed lower bound, possible upper bound, and uncertainty flag remain
available; only the sensitivity restricted to exactly known periods is
missing. A timing result is also left missing when the event may have occurred
outside its permitted search period.

## Check of candidate state-window cutoffs

```{r}
#| label: tbl-state-support-candidates
#| tbl-cap: "Check that retained counts at 70%, 80%, and 90% match the earlier cutoff comparison."

knitr::kable(
  state_candidate_display(metric_build$state_support_candidates)
)
```

This table, calculated after metric derivation, must match the earlier
candidate-cutoff table for the same coverage inputs. It remains descriptive
and does not retrospectively justify the approved cutoff.

## Checks for gaps and daylight-saving changes

```{r}
#| label: tbl-metric-gap-diagnostics
#| tbl-cap: "Checks of actual-time and local-clock data, gaps, and daylight-saving changes, by placement."

gap_summary <- purrr::imap_dfr(
  metric_build$results,
  function(result, placement) {
    result$gap |>
      dplyr::summarise(
        placement = placement,
        participant_days = dplyr::n(),
        true_utc_minutes = sum(.data$expected_real_minutes),
        represented_wall_minutes = sum(.data$represented_wall_minutes),
        unknown_diary_state_real_minutes = sum(
          .data$unknown_diary_state_real_minutes
        ),
        missing_medi_real_minutes = sum(.data$missing_medi_real_minutes),
        missing_medi_runs = sum(.data$missing_medi_runs),
        maximum_missing_run_minutes = max(
          .data$maximum_missing_medi_run_minutes
        ),
        dst_fold_real_minutes = sum(.data$dst_fold_real_minutes),
        nonexistent_spring_wall_minutes = sum(
          .data$structural_nonexistent_wall_minutes
        )
      )
  }
) |>
  dplyr::mutate(
    placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    )
  )
knitr::kable(
  gap_summary,
  col.names = c(
    "Placement",
    "Participant-days",
    "Expected actual-time minutes",
    "Local-clock minutes represented",
    "Minutes with unknown diary period",
    "Minutes with missing melEDI",
    "Continuous missing periods",
    "Longest missing period (minutes)",
    "Repeated autumn clock minutes",
    "Absent spring clock minutes"
  )
)
```

The gap and uncertainty files distinguish several situations. A known zero
means sufficient observed data showed that no qualifying event occurred. A
zero lower bound means no event was observed, but missing data could hide one.
The files also show whether a period touched the day boundary, was interrupted
by missing data, or whether a timing event may have fallen outside the
permitted search period. Missing diary periods are recorded separately.
Participant-level interdaily stability (IS) and intradaily variability (IV)
rows record the actual number of valid one-minute melEDI values and the
expected number of minutes ordered in actual time. They do not assume that
every retained hourly summary represents 60 observed minutes.

## Files produced

For each placement, Preparation 04 writes:

- participant-day metrics and participant-level IS/IV as R data files
  containing one object and as source-data-ready CSVs;
- complete 30-minute arithmetic-mean and one-hour zero-aware geometric-mean
  melEDI values as R data files and CSVs;
- one-row-per-value CSVs for metric values and whether they could be
  calculated; and
- checks of metric coverage, uncertainty caused by missing data, and gaps,
  including the observed longest-period lower bound, possible upper bound,
  uncertainty flag, and sensitivity restricted to exactly known periods.

The run also writes fixed settings, an exact inventory of the Preparation 01
sleep and wear inputs, the diagnostic 70%/80%/90% data-sufficiency table, and
an output-file inventory with digital fingerprints. These technical files are
linked from the metric artifact inventory above.

Focused R 4.6.1 tests passed for timing above 250 lx, the every-minute M10/L10
search, and the longest-period lower-bound rules. An independent
minute-by-minute calculation also reproduced M10 and L10. The full-data
outputs have been built; their independent verification and comparison with
the manuscript-prepared-data sensitivity remain in progress.
