---
title: "Preparation 05: Download fixed participant and site data"
---

## Purpose

This notebook downloads the exact questionnaire, diary, and site-data versions
needed for the analysis. It records where each file came from and which
variables it contains. It does not clean, recalculate, filter, or combine
scientific variables.

## What this step uses, produces, and checks

- Inputs: `config/site_sources.csv`,
  `config/model_input_availability.csv`, and the verified Preparation 01
  inventory of exact downloaded files.
- Outputs: separate local copies of the model inputs, checks of the object and
  columns stored in each file, and
  `artifacts/12_manifests/model_input_acquisition.csv`.
- Required checks: all 63 expected combinations of site and data type remain
  visible; download addresses contain the complete commit identifier and never
  `/main/`; paths to local copies are recorded relative to the project; each
  source file contains exactly its declared single object; and the nine
  sleep-diary files are reused only after their Preparation 01 download
  address, commit, SHA-256 digital fingerprint, and byte size are independently
  rechecked. A SHA-256 digital fingerprint confirms that a file has not
  changed.

The file inventory can also record an expected missing source and its reason,
although every source in the current release is expected to be present.

## Setup

```{r}
#| label: setup-model-input-acquisition
#| eval: false

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
source(file.path(root, "scripts/pipeline/paths_io.R"))
source(file.path(root, "scripts/pipeline/assertions.R"))
source(file.path(root, "scripts/pipeline/import_sources.R"))
source(file.path(root, "scripts/pipeline/model_input_acquisition.R"))
source(file.path(root, "scripts/pipeline/build_model_input_acquisition.R"))
source(file.path(root, "scripts/pipeline/verify_model_input_acquisition.R"))
```

## Download and save the exact inputs

This is the only step that connects to the network. Later hypothesis notebooks
use the checked local R data files (`.rds`) and do not download source data.
Use `cache_policy = "refresh"` for the first complete download and whenever an
interrupted run has left local files without a complete file inventory. Use
`cache_policy = "reuse"` only after the file inventory has passed the
independent check; the notebook then stops if a download address, commit,
digital fingerprint, byte size, or path differs.

```{r}
#| label: acquire-model-inputs
#| eval: false

model_input_acquisition <- build_model_input_acquisition(
  root = root,
  cache_policy = "refresh"
)
```

## Check downloaded files and their recorded sources

```{r}
#| label: verify-model-input-acquisition
#| eval: false

model_input_verification <- verify_model_input_acquisition(root = root)
stopifnot(identical(model_input_verification$status, "PASS"))
knitr::kable(
  data.frame(
    Check = c(
      "Overall verification status",
      "Expected source files",
      "Present source files",
      "Expected missing source files",
      "Sleep diaries reused from Preparation 01",
      "Fixed source versions",
      "Files downloaded in this run",
      "Source-version inventory fingerprint",
      "Downloaded-file inventory fingerprint",
      "Object check fingerprint",
      "Column check fingerprint"
    ),
    Result = unlist(
      model_input_verification[c(
        "status",
        "expected_sources",
        "present_sources",
        "expected_absences",
        "preparation01_reused",
        "source_pins",
        "downloaded_in_manifest",
        "source_pins_sha256",
        "manifest_sha256",
        "object_audit_sha256",
        "column_audit_sha256"
      )],
      use.names = FALSE
    )
  ),
  caption = "Overall result of the independent file check."
)
```

Preparation 06 cleans and interprets variables, checks participant
identifiers, and combines the files needed for each hypothesis.
