---
title: "Preparation 05: Acquire fixed questionnaire and diary inputs"
subtitle: "From site-specific data releases to locally verified analysis inputs"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-05
#| include: false

library(digest)
library(dplyr)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

relative_path <- function(path) {
  sub(paste0("^", root, "/"), "", path)
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE, progress = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

wrap_hash <- function(value) {
  gsub("(.{8})", "\\1\u200b", value, perl = TRUE)
}

use_reader_site_names <- function(data) {
  data |>
    dplyr::mutate(
      dplyr::across(
        where(is.character),
        ~ gsub("\\bTUM\\b", "Munich (DE)", .x, perl = TRUE)
      )
    )
}

modality_labels <- c(
  demographics = "Demographics",
  chronotype = "Chronotype",
  leba = "Light Exposure Behaviour Assessment (LEBA)",
  vlsq8 = "Visual Light Sensitivity Questionnaire-8 (VLSQ-8)",
  exercisediary = "Exercise diary",
  lightexposurediary = "Light-exposure diary",
  sleepdiaries = "Sleep diary"
)

modality_roles <- c(
  demographics = paste(
    "Participant characteristics such as age, sex, gender, and employment",
    "status."
  ),
  chronotype = paste(
    "Questionnaire-derived circadian preference and sleep-timing measures,",
    "including sleep-corrected midsleep on free days."
  ),
  leba = "Habitual light-exposure behaviour scores and questionnaire items.",
  vlsq8 = "Self-reported visual light sensitivity and component responses.",
  exercisediary = "Repeated exercise and activity-context records.",
  lightexposurediary = "Repeated self-reported lighting-context intervals.",
  sleepdiaries = paste(
    "Daily sleep timing and duration records; the same exact files were",
    "already acquired in Preparation 01."
  )
)

acquisition_manifest_path <- absolute_path(
  "artifacts/12_manifests/model_input_acquisition.csv"
)
expected_acquisition_manifest_sha256 <-
  "ac79a83c2d60dde1ac563f6e2cb09c58d4d7d3b06549d9b9262e4acd5a50a58b"
expected_source_pin_registry_sha256 <-
  "3fbf9f40125c60e1d0779d2b2d524e79534597586c60ec236a1ce16299e18831"
expected_tum_exercise_commit <-
  "618fda8521f3cf581661ceb2c026e3de7cd9ba82"
expected_tum_exercise_sha256 <-
  "0db56324e9826d427c42984fdfc43eb1e7487211365f222a1c0b2ecff618f1ef"
acquisition_manifest_sha256 <- digest::digest(
  file = acquisition_manifest_path,
  algo = "sha256"
)
stopifnot(identical(
  acquisition_manifest_sha256,
  expected_acquisition_manifest_sha256
))
source_pin_registry_sha256 <- digest::digest(
  file = absolute_path("config/model_input_source_pins.csv"),
  algo = "sha256"
)
stopifnot(identical(
  source_pin_registry_sha256,
  expected_source_pin_registry_sha256
))

acquisition <- read_stored_csv(
  "artifacts/12_manifests/model_input_acquisition.csv"
)
availability <- read_stored_csv("config/model_input_availability.csv")
source_pins <- read_stored_csv("config/model_input_source_pins.csv")
site_sources <- read_stored_csv("config/site_sources.csv")
site_display <- read_stored_csv("config/site_display_registry.csv")
preparation01_downloads <- read_stored_csv(
  "artifacts/12_manifests/pinned_downloads.csv"
)
object_audit <- read_stored_csv(
  "artifacts/01_imported/model_inputs/object_audit.csv"
)
column_audit <- read_stored_csv(
  "artifacts/01_imported/model_inputs/column_audit.csv"
)

assert_columns(
  acquisition,
  c(
    "site", "modality", "expected_status", "observed_status",
    "repository", "commit", "doi", "relative_source_path",
    "source_url", "object_name", "source_pin_applied",
    "source_pin_reason_code", "expected_sha256", "expected_bytes",
    "cache_path", "sha256", "bytes", "cache_origin", "downloaded"
  ),
  "model-input acquisition manifest"
)
assert_columns(
  availability,
  c("site", "modality", "expected_status", "absence_reason_code"),
  "model-input availability registry"
)
assert_columns(
  source_pins,
  c(
    "site", "modality", "repository", "commit", "doi",
    "relative_source_path", "object_name", "expected_sha256",
    "expected_bytes", "reason_code"
  ),
  "model-input source-pin registry"
)
assert_columns(
  site_sources,
  c("site", "repository", "commit", "doi", "timezone"),
  "site-source registry"
)
assert_columns(
  site_display,
  c("site", "display_order", "display_name", "color_hex"),
  "site display registry"
)
assert_columns(
  preparation01_downloads,
  c(
    "site", "modality", "repository", "commit", "doi", "source_url",
    "local_path", "sha256", "bytes"
  ),
  "Preparation 01 fixed-download inventory"
)
assert_columns(
  object_audit,
  c(
    "site", "modality", "object_name", "object_class", "object_type",
    "rows", "columns"
  ),
  "stored object audit"
)
assert_columns(
  column_audit,
  c(
    "site", "modality", "column_order", "column_name", "column_class",
    "column_type"
  ),
  "stored column audit"
)

cache_paths <- absolute_path(acquisition$cache_path)
cache_present <- file.exists(cache_paths)
cache_bytes <- rep(NA_real_, length(cache_paths))
cache_sha256 <- rep(NA_character_, length(cache_paths))
cache_bytes[cache_present] <- file.info(cache_paths[cache_present])$size
cache_sha256[cache_present] <- vapply(
  cache_paths[cache_present],
  function(path) digest::digest(file = path, algo = "sha256"),
  character(1)
)

expected_grid <- availability |>
  dplyr::select(
    .data$site,
    .data$modality,
    registered_status = .data$expected_status
  )
acquisition_grid <- acquisition |>
  dplyr::select(
    .data$site,
    .data$modality,
    acquired_status = .data$expected_status
  )
grid_check <- dplyr::full_join(
  expected_grid,
  acquisition_grid,
  by = c("site", "modality")
)

site_release_check <- acquisition |>
  dplyr::left_join(
    site_sources |>
      dplyr::select(
        .data$site,
        site_repository = .data$repository,
        site_commit = .data$commit,
        site_doi = .data$doi
      ),
    by = "site"
  ) |>
  dplyr::left_join(
    source_pins |>
      dplyr::select(
        .data$site,
        .data$modality,
        pin_repository = .data$repository,
        pin_commit = .data$commit,
        pin_doi = .data$doi,
        pin_sha256 = .data$expected_sha256,
        pin_bytes = .data$expected_bytes
      ),
    by = c("site", "modality")
  ) |>
  dplyr::mutate(
    expected_repository = dplyr::if_else(
      .data$source_pin_applied,
      .data$pin_repository,
      .data$site_repository
    ),
    expected_commit = dplyr::if_else(
      .data$source_pin_applied,
      .data$pin_commit,
      .data$site_commit
    ),
    expected_doi = dplyr::if_else(
      .data$source_pin_applied,
      .data$pin_doi,
      .data$site_doi
    )
  )

sleep_acquisition <- acquisition |>
  dplyr::filter(.data$modality == "sleepdiaries") |>
  dplyr::select(
    .data$site,
    acquisition_repository = .data$repository,
    acquisition_commit = .data$commit,
    acquisition_doi = .data$doi,
    acquisition_url = .data$source_url,
    acquisition_path = .data$cache_path,
    acquisition_sha256 = .data$sha256,
    acquisition_bytes = .data$bytes
  )
sleep_preparation01 <- preparation01_downloads |>
  dplyr::filter(.data$modality == "sleepdiaries") |>
  dplyr::transmute(
    .data$site,
    preparation01_repository = .data$repository,
    preparation01_commit = .data$commit,
    preparation01_doi = .data$doi,
    preparation01_url = .data$source_url,
    preparation01_path = relative_path(.data$local_path),
    preparation01_sha256 = .data$sha256,
    preparation01_bytes = .data$bytes
  )
sleep_identity <- dplyr::inner_join(
  sleep_acquisition,
  sleep_preparation01,
  by = "site"
) |>
  dplyr::mutate(
    exact_match =
      .data$acquisition_repository == .data$preparation01_repository &
      .data$acquisition_commit == .data$preparation01_commit &
      .data$acquisition_doi == .data$preparation01_doi &
      .data$acquisition_url == .data$preparation01_url &
      .data$acquisition_path == .data$preparation01_path &
      .data$acquisition_sha256 == .data$preparation01_sha256 &
      .data$acquisition_bytes == .data$preparation01_bytes
  )

stopifnot(
  nrow(acquisition) == 63L,
  nrow(availability) == 63L,
  nrow(grid_check) == 63L,
  all(grid_check$registered_status == grid_check$acquired_status),
  all(acquisition$expected_status == "present"),
  all(acquisition$observed_status %in% c("present_reused", "present_downloaded")),
  all(cache_present),
  all(cache_bytes == acquisition$bytes),
  all(cache_sha256 == acquisition$sha256),
  all(!grepl("/main/", acquisition$source_url, fixed = TRUE)),
  all(nchar(acquisition$commit) == 40L),
  all(mapply(
    function(commit, url) grepl(commit, url, fixed = TRUE),
    acquisition$commit,
    acquisition$source_url,
    USE.NAMES = FALSE
  )),
  all(!grepl("^/", acquisition$cache_path)),
  all(acquisition$repository == site_release_check$expected_repository),
  all(acquisition$commit == site_release_check$expected_commit),
  all(acquisition$doi == site_release_check$expected_doi),
  nrow(source_pins) == 1L,
  source_pins$site == "TUM",
  source_pins$modality == "exercisediary",
  source_pins$commit == expected_tum_exercise_commit,
  source_pins$expected_sha256 == expected_tum_exercise_sha256,
  source_pins$expected_bytes == 2662,
  sum(acquisition$source_pin_applied) == 1L,
  nrow(sleep_identity) == 9L,
  all(sleep_identity$exact_match),
  nrow(object_audit) == 63L,
  nrow(column_audit) == 963L,
  nrow(dplyr::distinct(object_audit, .data$site, .data$modality)) == 63L,
  nrow(dplyr::distinct(column_audit, .data$site, .data$modality,
                       .data$column_order)) == 963L
)

object_manifest_check <- object_audit |>
  dplyr::inner_join(
    acquisition |>
      dplyr::select(
        .data$site,
        .data$modality,
        manifest_object_name = .data$object_name
      ),
    by = c("site", "modality")
  )
stopifnot(
  nrow(object_manifest_check) == 63L,
  all(object_manifest_check$object_name ==
        object_manifest_check$manifest_object_name)
)

input_display <- tibble::tibble(
  `Stored input registry` = c(
    "config/site_sources.csv",
    "config/model_input_availability.csv",
    "config/model_input_source_pins.csv",
    "artifacts/12_manifests/pinned_downloads.csv"
  ),
  `What it fixes before acquisition` = c(
    paste(
      "Nine site repositories, complete 40-character commit identifiers,",
      "data-release DOIs, and time zones."
    ),
    paste(
      "The expected 9-site × 7-modality grid. A declared absence could be",
      "recorded with a reason, but all 63 entries are expected to be present."
    ),
    paste(
      "A narrowly scoped release override for the corrected TUM exercise",
      "diary, including its expected SHA-256 and byte size."
    ),
    paste(
      "The exact sleep-diary files already acquired in Preparation 01, so",
      "Preparation 05 can reuse rather than download them again."
    )
  )
)

site_release_display <- site_sources |>
  dplyr::inner_join(site_display, by = "site") |>
  dplyr::left_join(
    acquisition |>
      dplyr::count(.data$site, name = "available_files"),
    by = "site"
  ) |>
  dplyr::arrange(.data$display_order) |>
  dplyr::transmute(
    Site = paste0(
      "<span style=\"color:", .data$color_hex,
      "\" aria-hidden=\"true\">●</span> ",
      .data$display_name
    ),
    `Fixed release and acquired files` = paste0(
      "**", .data$repository, "**<br>",
      "Commit: ", wrap_hash(.data$commit), "<br>",
      "[DOI ", .data$doi, "](https://doi.org/", .data$doi, ")<br>",
      .data$available_files, " of 7 expected files present"
    )
  )

modality_display <- object_audit |>
  dplyr::group_by(.data$modality, .data$object_name) |>
  dplyr::summarise(
    Sites = dplyr::n_distinct(.data$site),
    `Stored objects` = dplyr::n(),
    `Rows across sites` = sum(.data$rows),
    `Columns per object` = paste(sort(unique(.data$columns)), collapse = ", "),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    Modality = unname(modality_labels[.data$modality]),
    `Scientific role before normalization` = unname(
      modality_roles[.data$modality]
    )
  ) |>
  dplyr::select(
    .data$Modality,
    .data$`Scientific role before normalization`,
    .data$Sites,
    .data$`Stored objects`,
    .data$`Rows across sites`,
    .data$`Columns per object`
  ) |>
  dplyr::arrange(match(.data$Modality, unname(modality_labels)))

acquisition_outcomes <- tibble::tibble(
  `How the exact local file was obtained` = c(
    "Reused from Preparation 01",
    "Reused from the verified model-input cache",
    "Downloaded from its fixed release"
  ),
  Files = c(
    sum(acquisition$cache_origin == "preparation01"),
    sum(acquisition$cache_origin == "model_input_cache"),
    sum(acquisition$cache_origin == "model_input_download")
  ),
  `Why this was acceptable` = c(
    paste(
      "All nine sleep diaries matched the Preparation 01 repository, commit,",
      "DOI, address, local path, SHA-256, and byte size."
    ),
    paste(
      "The stored file already matched the exact fixed release and its",
      "manifest fingerprint, so no network request was needed."
    ),
    paste(
      "Only the corrected TUM exercise diary required acquisition in the",
      "recorded run; its release, fingerprint, and byte size were fixed in",
      "advance."
    )
  )
)

pin <- source_pins[1, , drop = FALSE]
tum_pin <- tibble::tibble(
  Item = c(
    "Site and modality", "Repository", "Fixed commit", "DOI",
    "Stored object", "Expected file SHA-256", "Expected bytes",
    "Reason"
  ),
  Value = c(
    paste(pin$site, "exercise diary"),
    pin$repository,
    wrap_hash(pin$commit),
    paste0("[", pin$doi, "](https://doi.org/", pin$doi, ")"),
    pin$object_name,
    wrap_hash(pin$expected_sha256),
    scales::comma(pin$expected_bytes),
    paste(
      "Use the corrected upstream participant identity; no local",
      "identifier rewrite is applied."
    )
  )
)

script_map <- tibble::tibble(
  `Step and producing code` = c(
    paste0(
      "**1. Resolve the expected files**<br>",
      "scripts/pipeline/model_input_acquisition.R<br>",
      "scripts/pipeline/import_sources.R"
    ),
    paste0(
      "**2. Reuse or acquire the exact releases**<br>",
      "scripts/pipeline/build_model_input_acquisition.R"
    ),
    paste0(
      "**3. Record object and column structure**<br>",
      "scripts/pipeline/model_input_acquisition.R"
    ),
    paste0(
      "**4. Independently check the acquisition bundle**<br>",
      "scripts/pipeline/verify_model_input_acquisition.R"
    )
  ),
  `Reads, operation, separation, and output` = c(
    paste(
      "Reads the site-release, availability, and modality-specific pin",
      "registries. It constructs all 63 expected site–modality records and",
      "their commit-specific addresses before any network or cache decision.",
      "This keeps intended availability separate from what happens to be",
      "found locally."
    ),
    paste(
      "Reads each fixed specification. Sleep diaries are reused only after",
      "matching Preparation 01; another cached file is reused only when its",
      "release identity and fingerprint match; otherwise the script retrieves",
      "the commit-specific file. It writes the acquisition manifest and local",
      "copies without cleaning or combining scientific variables."
    ),
    paste(
      "Loads each acquired RData file in isolation during production, requires",
      "the declared single object, and records object class, dimensions,",
      "column order, names, classes, and storage types. Keeping this as an",
      "audit prevents silent schema changes from reaching Preparation 06."
    ),
    paste(
      "Reconstructs the 63-file specification, verifies fixed commits and",
      "pins, rechecks every local fingerprint and byte size, confirms the nine",
      "Preparation 01 reuses, and validates the object/column audits. The",
      "stored verification record applies to the exact current manifest; this",
      "report reads that record and does not rerun the verifier."
    )
  )
)

artifact_display <- tibble::tibble(
  `Stored artifact` = c(
    "63 fixed questionnaire and diary files",
    "Object audit",
    "Column audit",
    "Acquisition manifest"
  ),
  `Exact path and content` = c(
    paste(
      "Exact site–modality paths are listed in",
      "artifacts/12_manifests/model_input_acquisition.csv; 54 files are under",
      "artifacts/01_imported/model_inputs/cache/ and nine sleep diaries remain",
      "under artifacts/01_imported/cache/."
    ),
    paste(
      "artifacts/01_imported/model_inputs/object_audit.csv — one row for each",
      "site–modality object, including the declared object name, class, rows,",
      "and columns."
    ),
    paste(
      "artifacts/01_imported/model_inputs/column_audit.csv — 963 ordered",
      "column records with name, class, and storage type."
    ),
    paste0(
      "artifacts/12_manifests/model_input_acquisition.csv — the complete ",
      "site–modality inventory (SHA-256 ",
      wrap_hash(acquisition_manifest_sha256), ")."
    )
  ),
  `What Preparation 06 receives` = c(
    paste(
      "The untouched local RData objects and their exact release provenance",
      "for normalization."
    ),
    "A stored check of object identity and dimensions.",
    "A stored check of the available schema before any variable interpretation.",
    "The authoritative site–modality-to-file map and fingerprints."
  )
)

validation_display <- tibble::tibble(
  Check = c(
    "Current acquisition-manifest identity",
    "Expected and present site–modality files",
    "Local file size and SHA-256 agreement",
    "Fixed source addresses",
    "Preparation 01 sleep-diary identity",
    "Single declared objects",
    "Recorded source columns",
    "Modality-specific release pins"
  ),
  Result = c(
    "PASS",
    "63 / 63",
    paste0(
      sum(cache_present & cache_bytes == acquisition$bytes &
            cache_sha256 == acquisition$sha256),
      " / 63"
    ),
    paste0(
      sum(nchar(acquisition$commit) == 40L),
      " commit-specific; 0 use /main/"
    ),
    paste0(sum(sleep_identity$exact_match), " / 9"),
    paste0(nrow(object_manifest_check), " / 63"),
    scales::comma(nrow(column_audit)),
    paste0(sum(acquisition$source_pin_applied), " / 1")
  ),
  `What the check establishes` = c(
    paste0(
      "The manifest SHA-256 is ",
      wrap_hash(acquisition_manifest_sha256), "."
    ),
    "Every file declared in the 9 × 7 availability grid is present.",
    "Each local file is byte-identical to the file recorded in the manifest.",
    paste(
      "Every address embeds a complete commit identifier rather than a",
      "moving branch name."
    ),
    paste(
      "Repository, commit, DOI, address, path, SHA-256, and bytes all match",
      "the earlier acquisition."
    ),
    "Every site–modality file has the declared single object name.",
    "The stored audit covers every ordered source column across 63 objects.",
    "The corrected TUM exercise release is the only scoped override."
  )
)

verification_evidence <- tibble::tibble(
  `Stored verification evidence` = c(
    "Focused acquisition and corruption tests",
    "TUM source-transition test",
    "Independent acquisition-to-normalization check",
    "Scientific boundary"
  ),
  `Result tied to the current acquisition manifest` = c(
    paste(
      "Passed release resolution, cache reuse, absent-input, pin-expiry,",
      "duplicate-pin, and manifest-tamper cases."
    ),
    paste(
      "Confirmed the fixed TUM exercise object contains seven TUM_S001 dates,",
      "no TUM_S101 row, and no duplicate participant-date key."
    ),
    paste(
      "Under R 4.6.1, reconstructed all 963 source columns and confirmed exact",
      "values, missingness, scores, factor levels, units, labels, release",
      "provenance, and row traces. The recorded acquisition manifest is",
      "ac79a83c2d60dde1ac563f6e2cb09c58d4d7d3b06549d9b9262e4acd5a50a58b."
    ),
    paste(
      "These checks establish immutable acquisition and lossless hand-off.",
      "They do not authorize a hypothesis-specific join, exclusion, recoding,",
      "model, estimate, or claim."
    )
  )
)
```

## Purpose and position in the preparation chain

Preparation 05 obtains the fixed questionnaire and diary files needed for
participant characteristics, chronotype, habitual light-exposure behaviour,
visual light sensitivity, exercise, self-reported lighting context, and
sleep. Each site's files come from a fixed data release. The result is a local,
verified copy of every expected file, plus stored descriptions of each
object's rows and columns. Exact repository, release, and file-identity details
are collected in the technical provenance section.

This step does not clean variables, calculate scores, harmonize factor levels,
interpret timestamps, join participants to light metrics, or filter a model
sample. [Preparation 06](06_model_ready_datasets.qmd) performs those operations
from the fixed files recorded here. The light-exposure metrics from Preparation
04 form a separate input branch into Preparation 06.

```{mermaid}
flowchart LR
  S["Fixed questionnaire<br/>and diary releases"] --> A["Preparation 05<br/>acquire and verify files"]
  P1["Preparation 01<br/>nine verified sleep diaries"] --> A
  A --> L["63 local questionnaire<br/>and diary objects"]
  A --> U["Object and column audits"]
  L --> P6["Preparation 06<br/>normalize and join inputs"]
  U --> P6
  P4["Preparation 04<br/>light-exposure metrics"] --> P6
```

## Files acquired and why

The availability record contains all 63 combinations of nine sites and
seven modalities. Its explicit grid matters because a missing download cannot
be mistaken for an intentionally unavailable questionnaire. The record can
carry a reason for an expected absence, although no current entry is absent.

## Access, privacy, and acquired contents

The acquisition audit records structure rather than displaying participant
responses. Row counts below describe the unchanged objects across all nine
sites. Preparation 06 later decides how columns are interpreted and which
ones may enter an analysis.

```{r}
#| label: tbl-acquired-modalities
#| tbl-cap: "Questionnaire and diary objects acquired at all nine sites."

prep_gt(
  modality_display,
  source_note = paste(
    "Rows are summed across nine site files. Column counts are identical",
    "within each modality."
  )
) |>
  gt::fmt_number(
    columns = c(.data$Sites, .data$`Stored objects`, .data$`Rows across sites`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Modality ~ gt::pct(24),
    `Scientific role before normalization` ~ gt::pct(39),
    Sites ~ gt::pct(7),
    `Stored objects` ~ gt::pct(9),
    `Rows across sites` ~ gt::pct(11),
    `Columns per object` ~ gt::pct(10)
  )
```

The 63 objects contain 963 recorded columns in total. That count is a schema
audit, not a claim that all columns are suitable for analysis. In particular,
raw free-text fields remain in the untouched source objects but are not shown
here; Preparation 06 explicitly excludes them from analytical outputs.

## Technical provenance

::: {.callout-note title="What is calculated during this render?"}
Rendering reads the stored registries, acquisition manifest, object audit, and
column audit. It checks their schemas, hashes the 63 already local files,
compares the nine sleep diaries with the Preparation 01 inventory, and
calculates compact row and file totals. It does **not** connect to the network,
open raw questionnaire values, run the acquisition builder or verifier,
normalize a variable, construct a model input, or overwrite any stored
artifact.
:::

### Input specification records

```{r}
#| label: tbl-preparation05-inputs
#| tbl-cap: "Stored records that fix which questionnaire and diary files Preparation 05 acquires."

prep_gt(
  use_reader_site_names(input_display),
  source_note = paste(
    "All four registries are read during rendering; none is modified."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_label(
    `Stored input registry` = "Input record",
    `What it fixes before acquisition` = "Role before acquisition"
  ) |>
  gt::cols_width(
    `Stored input registry` ~ gt::pct(38),
    `What it fixes before acquisition` ~ gt::pct(62)
  )
```

### Fixed source releases

```{r}
#| label: tbl-site-releases
#| tbl-cap: "Site data releases in submitted-manuscript order and colours."

prep_gt(
  site_release_display,
  source_note = paste(
    "Every default release uses a complete 40-character commit identifier.",
    "The Munich (DE) exercise-diary exception is shown separately below."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    Site ~ gt::pct(32),
    `Fixed release and acquired files` ~ gt::pct(68)
  )
```

All download addresses embed the recorded commit. None contains `/main/` or
another moving branch label. Consequently, revisiting the same address refers
to the same repository state rather than whatever happens to be current at a
later date.

#### Corrected Munich (DE) exercise source

One file deliberately uses a modality-specific commit instead of the default
Munich (DE) release. The upstream exercise diary corrects the participant
identity at the source; Preparation 05 does not rewrite that identifier
locally.

```{r}
#| label: tbl-tum-exercise-pin
#| tbl-cap: "Exact fixed release for the corrected Munich (DE) exercise diary."

prep_gt(
  use_reader_site_names(tum_pin),
  source_note = paste(
    "The release exception is restricted to one site and one modality; all",
    "other Munich (DE) files retain the site-wide release shown above."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    Item ~ gt::pct(38),
    Value ~ gt::pct(62)
  )
```

### How files were obtained in the recorded run

```{r}
#| label: tbl-acquisition-outcomes
#| tbl-cap: "How the 63 exact local files were obtained without changing their scientific contents."

prep_gt(use_reader_site_names(acquisition_outcomes)) |>
  gt::fmt_number(columns = .data$Files, decimals = 0) |>
  gt::cols_width(
    `How the exact local file was obtained` ~ gt::pct(32),
    Files ~ gt::pct(10),
    `Why this was acceptable` ~ gt::pct(58)
  )
```

The recorded run reused 62 files and downloaded one. The reused files comprise
nine sleep diaries from Preparation 01 and 53 files from the verified
model-input cache. Reuse is not based only on a filename: the release identity,
local path, byte size, and SHA-256 must all agree. A SHA-256 is a digital
fingerprint; changing even one byte changes the fingerprint with overwhelming
probability.

### Production sequence

```{r}
#| label: tbl-script-execution-map
#| tbl-cap: "Producing scripts and modules in execution order; none is executed while this report renders."

prep_gt(
  script_map,
  source_note = paste(
    "The production acquisition may contact commit-specific network",
    "addresses. This reader-facing render never does."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

Separating the expected-file specification from cache or download decisions
prevents local availability from defining the scientific input. Separating
object/column inspection from later normalization likewise ensures that a
changed upstream schema is visible before values are interpreted or joined.

### Exact stored artifacts passed to Preparation 06

```{r}
#| label: tbl-acquisition-artifacts
#| tbl-cap: "Stored Preparation 05 artifacts and their role in the next preparation step."

prep_gt(
  use_reader_site_names(artifact_display),
  source_note = paste(
    "The acquisition manifest lists every exact local path, source release,",
    "object name, SHA-256, and byte size."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Stored artifact` ~ gt::pct(24),
    `Exact path and content` ~ gt::pct(48),
    `What Preparation 06 receives` ~ gt::pct(28)
  )
```

The local RData objects remain separated by site and modality. Preparation 05
does not combine sites, rename values, recalculate questionnaire scores,
select categories, or discard incomplete intervals. The manifest and audits
allow Preparation 06 to perform those operations against known, immutable
inputs.

### Checks tied to the current stored files

```{r}
#| label: tbl-current-acquisition-validation
#| tbl-cap: "Bounded checks of the current acquisition manifest, files, and stored structural audits."

prep_gt(
  use_reader_site_names(validation_display),
  source_note = paste(
    "File-identity and schema checks are performed during rendering from",
    "already stored outputs; raw response values are not opened."
  )
) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    Check ~ gt::pct(31),
    Result ~ gt::pct(20),
    `What the check establishes` ~ gt::pct(49)
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(columns = .data$Result)
  )
```

All 63 local files match the acquisition manifest with SHA-256
`ac79a83c2d60dde1ac563f6e2cb09c58d4d7d3b06549d9b9262e4acd5a50a58b`.
The nine sleep diaries also match the Preparation 01 inventory independently
on repository, commit, DOI, address, local path, fingerprint, and bytes.

#### Stored independent verification

```{r}
#| label: tbl-stored-acquisition-verification
#| tbl-cap: "Stored verification evidence associated with the current acquisition manifest."

prep_gt(
  use_reader_site_names(verification_evidence),
  source_note = paste(
    "The verification results are read from the Preparation 05 provenance",
    "finding and the Preparation 06 normalization gate. They are not rerun",
    "during this render."
  )
) |>
  gt::cols_width(
    `Stored verification evidence` ~ gt::pct(36),
    `Result tied to the current acquisition manifest` ~ gt::pct(64)
  )
```

The exact-current verification record is documented in the
[detailed input-provenance record](../../audit/findings/preparation05_input_provenance_gaps.md)
and the
[normalization verification record](../../audit/reconciliation/preparation06/normalization_gate.md).
The verifier is not rerun during this reader-facing render.
It establishes that the fixed source objects reached normalization without an
unrecorded release or schema change. It does not establish a hypothesis
result, model sample, effect estimate, or scientific claim.
