---
title: "Preparation 06: Build model-ready datasets"
subtitle: "From checked metrics and participant information to hypothesis-specific inputs"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-06
#| include: false

library(digest)
library(dplyr)
library(ggplot2)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)

stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE)
}

read_stored_rds <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readRDS(full_path)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )

  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

manifest_registry <- tibble::tribble(
  ~stage_order, ~stage, ~manifest,
  1L, "Standardized participant inputs",
  "artifacts/12_manifests/model_input_normalization.csv",
  2L, "Gap-timing-unaware prepared data",
  "artifacts/12_manifests/manuscript_prepared_data_artifacts.csv",
  3L, "Site and daylight context",
  "artifacts/12_manifests/site_solar_context_artifacts.csv",
  4L, "Local-clock and elapsed-time links",
  "artifacts/06_model_data/temporal_provenance/artifact_manifest.csv",
  5L, "Shared model inputs",
  "artifacts/12_manifests/base_model_data_artifacts.csv",
  6L, "H01 primary prepared frames",
  "artifacts/12_manifests/H01_model_data_artifacts.csv",
  7L, "H01 gap-timing-unaware prepared frames",
  "artifacts/12_manifests/H01_manuscript_prepared_data_artifacts.csv",
  8L, "Stored descriptive comparison",
  "artifacts/12_manifests/preanalysis_comparison_artifacts.csv"
)

manifest_objects <- list()
manifest_file_checks <- list()

for (index in seq_len(nrow(manifest_registry))) {
  registry_row <- manifest_registry[index, ]
  manifest_data <- read_stored_csv(registry_row$manifest)
  assert_columns(
    manifest_data,
    c("path", "sha256", "bytes"),
    registry_row$manifest
  )

  listed_paths <- absolute_path(manifest_data$path)
  exists <- file.exists(listed_paths)
  actual_bytes <- rep(NA_real_, length(listed_paths))
  actual_sha256 <- rep(NA_character_, length(listed_paths))
  actual_bytes[exists] <- file.info(listed_paths[exists])$size
  actual_sha256[exists] <- vapply(
    listed_paths[exists],
    digest::digest,
    character(1),
    algo = "sha256",
    file = TRUE
  )

  manifest_file_checks[[index]] <- tibble::tibble(
    stage_order = registry_row$stage_order,
    stage = registry_row$stage,
    path = manifest_data$path,
    exists = exists,
    bytes_match = exists & actual_bytes == as.numeric(manifest_data$bytes),
    sha256_match = exists & actual_sha256 == manifest_data$sha256
  )
  manifest_objects[[registry_row$stage]] <- manifest_data
}

manifest_file_checks <- dplyr::bind_rows(manifest_file_checks)
stopifnot(
  all(manifest_file_checks$exists),
  all(manifest_file_checks$bytes_match),
  all(manifest_file_checks$sha256_match)
)

manifest_summary <- manifest_file_checks |>
  dplyr::group_by(.data$stage_order, .data$stage) |>
  dplyr::summarise(
    `Files listed` = dplyr::n(),
    `Files present` = sum(.data$exists),
    `Size checks` = sum(.data$bytes_match),
    `SHA-256 checks` = sum(.data$sha256_match),
    Check = ifelse(
      all(.data$exists & .data$bytes_match & .data$sha256_match),
      "PASS",
      "FAIL"
    ),
    .groups = "drop"
  ) |>
  dplyr::arrange(.data$stage_order) |>
  dplyr::select(-.data$stage_order)

normalization_manifest <-
  manifest_objects[["Standardized participant inputs"]]
gap_manifest <-
  manifest_objects[["Gap-timing-unaware prepared data"]]
site_manifest <-
  manifest_objects[["Site and daylight context"]]

normalized_labels <- c(
  normalized_demographics = "Participant characteristics",
  normalized_chronotype = "Chronotype questionnaire",
  normalized_leba = "Light-exposure behaviour questionnaire (LEBA)",
  normalized_vlsq8 = "Visual light sensitivity questionnaire (VLSQ-8)",
  normalized_exercisediary = "Exercise diary",
  normalized_lightexposurediary = "Light-exposure diary",
  normalized_sleepdiaries = "Sleep diary"
)

normalized_summary <- normalization_manifest |>
  dplyr::filter(.data$artifact_id %in% names(normalized_labels)) |>
  dplyr::mutate(
    `Input type` = unname(normalized_labels[.data$artifact_id]),
    `Stored file` = .data$path
  ) |>
  dplyr::select(
    .data$`Input type`,
    Rows = .data$rows,
    Columns = .data$columns,
    .data$`Stored file`
  )

interval_issues <- read_stored_csv(
  "artifacts/06_model_data/normalized_inputs/audits/interval_issue_audit.csv"
)
assert_columns(
  interval_issues,
  c("modality", "site", "issue_code"),
  "interval issue audit"
)
interval_issue_summary <- interval_issues |>
  dplyr::count(.data$modality, .data$site, .data$issue_code, name = "Records") |>
  dplyr::mutate(
    `Prepared input` = dplyr::recode(
      .data$modality,
      lightexposurediary = "Light-exposure diary",
      sleepdiaries = "Sleep diary"
    ),
    `Why interval calculations cannot use the record` = dplyr::recode(
      .data$issue_code,
      missing_start_and_end = "Start and end time are missing",
      missing_wake = "Wake time is missing"
    )
  ) |>
  dplyr::select(
    .data$`Prepared input`,
    Site = .data$site,
    .data$Records,
    .data$`Why interval calculations cannot use the record`
  )

free_text_audit <- read_stored_csv(
  "artifacts/06_model_data/normalized_inputs/audits/free_text_audit.csv"
)
value_preservation_audit <- read_stored_csv(
  "artifacts/06_model_data/normalized_inputs/audits/value_preservation_audit.csv"
)
stopifnot(
  all(free_text_audit$excluded_from_analytic_rds),
  !any(free_text_audit$content_exported),
  all(value_preservation_audit$status == "PASS")
)

gap_output_labels <- c(
  participant_metrics_csv = "Participant-level metrics",
  participant_day_metrics_csv = "Participant-day metrics",
  thirty_minute_data_csv = "30-minute values",
  one_hour_data_csv = "One-hour values"
)

gap_artifact_summary <- gap_manifest |>
  dplyr::filter(.data$artifact_id %in% names(gap_output_labels)) |>
  dplyr::mutate(
    `Stored data level` = unname(gap_output_labels[.data$artifact_id]),
    `Stored file` = .data$path
  ) |>
  dplyr::select(
    .data$`Stored data level`,
    Rows = .data$rows,
    Columns = .data$columns,
    .data$`Stored file`
  )

correction_record <- read_stored_csv(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/correction_manifest.csv"
) |>
  dplyr::transmute(
    `Correction record` = .data$correction_id,
    Site = .data$affected_site,
    `Diary rows` = .data$affected_rows,
    `Recorded identifier` = .data$erroneous_value,
    `Verified identifier` = .data$corrected_value,
    `Scientific light values changed` = ifelse(
      .data$scientific_light_values_changed,
      "Yes",
      "No"
    )
  )

site_registry <- read_stored_csv("config/site_display_registry.csv")
assert_columns(
  site_registry,
  c("site", "display_order", "display_name", "color_hex"),
  "site display registry"
)

site_join_audit <- read_stored_csv(
  "artifacts/06_model_data/context/site_solar_context_join_audit.csv"
)
stopifnot(
  all(site_join_audit$status == "PASS"),
  all(site_join_audit$input_rows == site_join_audit$joined_rows),
  all(site_join_audit$unmatched_rows == 0L)
)

site_join_summary <- site_join_audit |>
  dplyr::transmute(
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    Resolution = dplyr::recode(
      .data$resolution,
      daily = "Daily",
      `30_minute` = "30 minutes",
      one_hour = "One hour"
    ),
    `Rows before join` = .data$input_rows,
    `Rows after join` = .data$joined_rows,
    `Participant-days` = .data$participant_days,
    `Unmatched rows` = .data$unmatched_rows,
    Check = .data$status
  )

true_utc_bins <- read_stored_rds(
  "artifacts/06_model_data/temporal_provenance/true_utc_source_bins.rds"
)
wall_outcome_links <- read_stored_rds(
  "artifacts/06_model_data/temporal_provenance/wall_outcome_links.rds"
)
assert_columns(
  true_utc_bins,
  c(
    "site", "Id", "local_date", "local_day_type",
    "relationship_type", "sequence_eligible",
    "true_elapsed_sequence_start"
  ),
  "true elapsed-time source bins"
)
assert_columns(
  wall_outcome_links,
  c(
    "site", "Id", "local_date", "relationship_type",
    "spring_gap_provenance"
  ),
  "local-clock outcome links"
)

count_participant_days <- function(data, day_type) {
  data |>
    dplyr::filter(.data$local_day_type == day_type) |>
    dplyr::distinct(.data$site, .data$Id, .data$local_date) |>
    nrow()
}

temporal_summary <- tibble::tibble(
  `Stored check or outcome` = c(
    "Original UTC intervals",
    "Rows on the shared local-clock grid",
    "Participant-days spanning the autumn clock change",
    "Local-clock rows averaging two real intervals",
    "Participant-days spanning the spring clock change",
    "Structural local-clock rows in the spring gap",
    "Original intervals without a one-to-one local-clock row",
    "Intervals eligible for continuous-sequence models",
    "Starts of continuous elapsed-time sequences"
  ),
  Value = c(
    nrow(true_utc_bins),
    nrow(wall_outcome_links),
    count_participant_days(true_utc_bins, "fall_back_25h"),
    sum(
      wall_outcome_links$relationship_type ==
        "two_to_one_averaged_fall_back",
      na.rm = TRUE
    ),
    count_participant_days(true_utc_bins, "spring_forward_23h"),
    sum(
      wall_outcome_links$spring_gap_provenance ==
        "structural_nonexistent_wall_bin",
      na.rm = TRUE
    ),
    sum(
      true_utc_bins$relationship_type != "one_to_one_true_elapsed",
      na.rm = TRUE
    ),
    sum(true_utc_bins$sequence_eligible %in% TRUE),
    sum(true_utc_bins$true_elapsed_sequence_start %in% TRUE)
  )
)

near_eye_grid <- read_stored_rds(
  "artifacts/06_model_data/base/metrics_glasses_30_minute_context.rds"
)
chest_grid <- read_stored_rds(
  "artifacts/06_model_data/base/metrics_chest_30_minute_context.rds"
)
assert_columns(
  near_eye_grid,
  c("site", "Id", "local_date", "bin_admissible"),
  "near-eye 30-minute model input"
)
assert_columns(
  chest_grid,
  c("site", "Id", "local_date", "bin_admissible"),
  "chest 30-minute model input"
)

summarise_grid <- function(data, prefix) {
  output <- data |>
    dplyr::group_by(.data$site) |>
    dplyr::summarise(
      participants = dplyr::n_distinct(.data$Id),
      participant_days = dplyr::n_distinct(
        paste(.data$Id, .data$local_date)
      ),
      grid_rows = dplyr::n(),
      admissible_rows = sum(.data$bin_admissible %in% TRUE),
      .groups = "drop"
    )
  names(output)[-1] <- paste(prefix, names(output)[-1], sep = "_")
  output
}

site_samples <- site_registry |>
  dplyr::left_join(
    summarise_grid(near_eye_grid, "near_eye"),
    by = "site",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    summarise_grid(chest_grid, "chest"),
    by = "site",
    relationship = "one-to-one"
  ) |>
  dplyr::arrange(.data$display_order) |>
  dplyr::transmute(
    Site = .data$display_name,
    site_code = .data$site,
    color_hex = .data$color_hex,
    `Near-eye participants` = .data$near_eye_participants,
    `Near-eye participant-days` = .data$near_eye_participant_days,
    `Near-eye 30-min rows` = .data$near_eye_grid_rows,
    `Near-eye admissible rows` = .data$near_eye_admissible_rows,
    `Chest participants` = .data$chest_participants,
    `Chest participant-days` = .data$chest_participant_days,
    `Chest 30-min rows` = .data$chest_grid_rows,
    `Chest admissible rows` = .data$chest_admissible_rows
  )

participant_availability <- read_stored_csv(
  "artifacts/06_model_data/base/participant_availability_audit.csv"
)
modality_columns <- c(
  "Participant characteristics" = "has_demographics",
  "Chronotype questionnaire" = "has_chronotype",
  "Light-exposure behaviour questionnaire (LEBA)" = "has_leba",
  "Visual light sensitivity questionnaire (VLSQ-8)" = "has_vlsq8",
  "Exercise diary" = "has_exercisediary",
  "Light-exposure diary" = "has_lightexposurediary",
  "Sleep diary" = "has_sleepdiaries"
)
assert_columns(
  participant_availability,
  c(
    "has_glasses_metrics", "has_chest_metrics",
    unname(modality_columns)
  ),
  "participant availability audit"
)

summarise_availability <- function(metric_column, placement_label) {
  metric_participants <- participant_availability[[metric_column]] %in% TRUE
  matching_counts <- vapply(
    unname(modality_columns),
    function(column) {
      sum(participant_availability[[column]][metric_participants] %in% TRUE)
    },
    integer(1)
  )
  tibble::tibble(
    Placement = placement_label,
    `Questionnaire or diary input` = names(modality_columns),
    `Participants with light metrics` = sum(metric_participants),
    `Participants with matching input` = matching_counts,
    `Participants without matching input` =
      sum(metric_participants) - matching_counts
  )
}

availability_summary <- dplyr::bind_rows(
  summarise_availability("has_glasses_metrics", "Near-eye"),
  summarise_availability("has_chest_metrics", "Chest")
)

h01_primary_flow <- read_stored_csv(
  "artifacts/06_model_data/H01/sample_flow.csv"
)
h01_gap_flow <- read_stored_csv(
  paste0(
    "artifacts/06_model_data/H01/scenarios/",
    "manuscript_prepared_data/sample_flow.csv"
  )
)
h01_primary_contract <- read_stored_csv(
  "artifacts/06_model_data/H01/metric_contract.csv"
)
h01_gap_contract <- read_stored_csv(
  paste0(
    "artifacts/06_model_data/H01/scenarios/",
    "manuscript_prepared_data/metric_contract.csv"
  )
)

make_h01_sample <- function(sample_flow, metric_contract) {
  assert_columns(
    sample_flow,
    c(
      "placement", "scenario", "metric_order", "metric_id", "stage",
      "scope", "model_observations", "participants",
      "contributing_participant_days", "metric_support_valid_hours",
      "sites"
    ),
    "H01 sample-flow file"
  )
  assert_columns(
    metric_contract,
    c("metric_id", "manuscript_name"),
    "H01 metric contract"
  )

  output <- sample_flow |>
    dplyr::filter(
      .data$placement == "glasses",
      .data$scenario == "all_available",
      .data$stage == "metric_estimable",
      .data$scope == "overall"
    ) |>
    dplyr::left_join(
      metric_contract |>
        dplyr::select(.data$metric_id, .data$manuscript_name),
      by = "metric_id",
      relationship = "many-to-one"
    ) |>
    dplyr::arrange(.data$metric_order) |>
    dplyr::transmute(
      Metric = .data$manuscript_name,
      Participants = .data$participants,
      `Rows in prepared frame` = .data$model_observations,
      `Contributing participant-days` =
        .data$contributing_participant_days,
      `Valid measurement hours` = .data$metric_support_valid_hours,
      Sites = .data$sites
    )

  stopifnot(nrow(output) == 17L)
  output
}

h01_primary_sample <- make_h01_sample(
  h01_primary_flow,
  h01_primary_contract
)
h01_gap_sample <- make_h01_sample(
  h01_gap_flow,
  h01_gap_contract
)

h01_primary_status <- read_stored_csv(
  "artifacts/06_model_data/H01/scenario_status.csv"
)
h01_gap_status <- read_stored_csv(
  paste0(
    "artifacts/06_model_data/H01/scenarios/",
    "manuscript_prepared_data/scenario_status.csv"
  )
)
paired_unavailable <- dplyr::bind_rows(
  h01_primary_status |>
    dplyr::mutate(Dataset = "Primary dataset"),
  h01_gap_status |>
    dplyr::mutate(Dataset = "Gap-timing-unaware dataset")
) |>
  dplyr::filter(!.data$available)
stopifnot(nrow(paired_unavailable) == 8L)

figure_source_path <- paste0(
  "artifacts/08_diagnostics/preanalysis_comparison/",
  "categorical_levels.csv"
)
figure_source <- read_stored_csv(figure_source_path)
assert_columns(
  figure_source,
  c("variable_id", "placement", "series", "level", "n"),
  "stored categorical figure source data"
)

figure_site_data <- figure_source |>
  dplyr::filter(
    .data$variable_id == "participant_site",
    .data$placement == "glasses"
  ) |>
  dplyr::mutate(
    Dataset = dplyr::recode(
      .data$series,
      canonical = "Primary dataset",
      baseline = "Gap-timing-unaware dataset"
    )
  ) |>
  dplyr::inner_join(
    site_registry,
    by = c("level" = "site"),
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    Dataset = factor(
      .data$Dataset,
      levels = c("Primary dataset", "Gap-timing-unaware dataset")
    ),
    display_name = factor(
      .data$display_name,
      levels = rev(site_registry$display_name)
    )
  )
stopifnot(nrow(figure_site_data) == 18L)

render_boundary <- tibble::tribble(
  ~During_render, ~What_it_means,
  "Read stored RDS and CSV files",
  "No source acquisition, normalization, metric derivation, or model-frame construction",
  "Check file size and SHA-256 identity",
  "Every file listed by the eight preparation manifests must match its stored record",
  "Check required columns",
  "The page stops if a stored file no longer has the fields needed for its displays",
  "Calculate lightweight counts and draw one plot",
  "Only descriptive summaries of stored rows are calculated; no analytical artifact is written"
)

script_map <- tibble::tribble(
  ~Step, ~Producing_script, ~Function_module, ~Reads_and_operation, ~Stored_output,
  1L,
  "scripts/pipeline/build_model_input_normalization.R",
  "scripts/pipeline/model_input_normalization.R",
  "Reads the fixed local files acquired in Preparation 05; aligns field types, labels, identifiers, and time representations while preserving recorded values. Kept separate so source-value and free-text checks are explicit.",
  "artifacts/06_model_data/normalized_inputs/",
  2L,
  "scripts/pipeline/build_manuscript_prepared_data.R",
  "scripts/pipeline/manuscript_prepared_data.R",
  "Reads the retained prepared measurements and checked participant inputs; assembles the predefined comparison dataset and records its single verified identifier correction. Kept separate so its preparation remains distinguishable from the primary dataset.",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/",
  3L,
  "scripts/pipeline/build_site_solar_context.R",
  "scripts/pipeline/site_solar_context.R",
  "Reads site coordinates, time zones, and all dates needed by both datasets; creates civil dawn, civil dusk, photoperiod, and descriptive solar-noon fields. Kept separate so every site/date join can be checked once.",
  "artifacts/06_model_data/context/",
  4L,
  "scripts/pipeline/build_temporal_sequence_provenance.R",
  "scripts/pipeline/temporal_sequence_provenance.R",
  "Reads accepted coverage and metric time series; links local-clock outcomes to their true UTC intervals and marks continuous elapsed-time sequences. Kept separate because time-of-day and elapsed-time questions require different coordinates.",
  "artifacts/06_model_data/temporal_provenance/",
  5L,
  "scripts/pipeline/build_base_model_data.R",
  "scripts/pipeline/base_model_data.R",
  "Reads checked metrics, standardized participant inputs, and site context; attaches one matching context record while retaining metric rows and recording participant-input availability. Kept separate from hypothesis-specific exclusions.",
  "artifacts/06_model_data/base/",
  6L,
  "scripts/pipeline/build_h01_model_data.R",
  "scripts/pipeline/h01_model_data.R",
  "Reads the shared model inputs and approved H01 metric contract; stores outcomes, predictors, metric-specific inclusion records, and sample flow for 17 metrics. No model is fitted.",
  "artifacts/06_model_data/H01.rds and artifacts/06_model_data/H01/",
  7L,
  "scripts/pipeline/build_h01_manuscript_prepared_data.R",
  "scripts/pipeline/h01_manuscript_prepared_adapter.R",
  "Reads the gap-timing-unaware data and applies the same H01 implementation contract. Kept separate so shared model code can be checked without erasing the preparation difference.",
  "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/",
  8L,
  "scripts/pipeline/build_preanalysis_comparison.R",
  "scripts/pipeline/preanalysis_comparison.R",
  "Reads both stored preparations; writes descriptive count, distribution, reconciliation, and figure-source records before hypothesis fitting. It does not test a scientific hypothesis.",
  "artifacts/08_diagnostics/preanalysis_comparison/"
)

handoff_artifacts <- tibble::tribble(
  ~Stored_artifact, ~Exact_path, ~Next_use,
  "Standardized participant inputs",
  "artifacts/06_model_data/normalized_inputs/",
  "Joined to checked metrics in the shared model inputs",
  "Gap-timing-unaware prepared measurements",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/",
  "Predefined prepared-data comparison in hypothesis analyses",
  "Site and daylight context",
  "artifacts/06_model_data/context/site_solar_context.rds",
  "Site/date predictors and verified many-to-one joins",
  "Elapsed-time provenance",
  "artifacts/06_model_data/temporal_provenance/true_utc_source_bins.rds",
  "Continuity and autocorrelation coordinates for H02 and H11",
  "Shared near-eye and chest inputs",
  "artifacts/06_model_data/base/",
  "Starting point for each approved hypothesis-specific build",
  "H01 primary prepared frames",
  "artifacts/06_model_data/H01.rds",
  "H01 analysis preparation and fitting",
  "H01 gap-timing-unaware prepared frames",
  "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/H01.rds",
  "H01 prepared-data comparison",
  "Stored descriptive comparison data",
  "artifacts/08_diagnostics/preanalysis_comparison/",
  "Reader-facing preparation comparisons and provenance checks"
)

environment_summary <- tibble::tibble(
  Component = c(
    "R",
    "Quarto profile",
    "Table package",
    "Manifest hashing",
    "Scientific recomputation during render"
  ),
  Value = c(
    paste(R.version$major, R.version$minor, sep = "."),
    "nathealth (HTML review render)",
    paste0("gt ", as.character(utils::packageVersion("gt"))),
    "SHA-256 with digest",
    "None"
  )
)
```

## Purpose and position in the analysis chain

Preparation 06 turns the checked light-exposure metrics from Preparation 04
and the fixed participant, questionnaire, and diary files acquired in
Preparation 05 into inputs that hypothesis analyses can use. It first makes
participant files structurally consistent, then adds participant and site
context to the metric files, preserves the distinction between local clock
time and actual elapsed time, and finally stores a separate input when a
hypothesis method has been approved. H01 is the first such build. This step
does not fit a statistical model.

The predefined **gap-timing-unaware dataset** still passed the general
50%-per-hour and 80%-per-day coverage rules. Its name means that the timing of
remaining missing observations was not used for an additional metric-specific
adjustment. For contrast, the primary dataset could be interpreted as a
“time-sensitive primary metric dataset.” From this point onward, this report
calls it simply the **primary dataset**.

The exact inputs used by the producing scripts are recorded in the eight
manifests checked on this page. They cover standardized participant inputs,
the separate prepared-data comparison, site and daylight context,
elapsed-time provenance, shared model inputs, both H01 preparations, and the
stored descriptive comparison. The corresponding paths are reported in
@tbl-manifest-identity and @tbl-script-execution-map.

::: {.callout-note appearance="simple" title="What is calculated during this render?"}
This report reads already stored RDS and CSV outputs. During rendering it
checks that every manifest-listed file has the recorded size and SHA-256
digital fingerprint, checks the columns needed for the displays, calculates
only lightweight descriptive counts from stored rows, and redraws one
reader-facing plot from its stored source-data CSV. It does not source or run
any producing script, retrieve data, standardize inputs, derive metrics,
construct or overwrite a model frame, fit a model, estimate autocorrelation,
or write an analytical artifact.
:::

```{mermaid}
%%| label: fig-preparation06-chain
%%| fig-cap: "Position of Preparation 06 in the data-to-analysis chain. Solid arrows show stored data passed forward; rendering reads these outputs but does not recreate them."
%%| fig-alt: "Flow diagram from fixed participant files and checked light-exposure metrics, through standardized participant inputs, site and elapsed-time context, shared model inputs, and hypothesis-specific prepared datasets, to hypothesis analyses. A separate branch carries the gap-timing-unaware dataset through the same hypothesis implementation."
flowchart LR
  A["Fixed participant, questionnaire, and diary files<br/>Preparation 05"] --> C["Standardize structures<br/>and audit source-value preservation"]
  B["Checked light-exposure metrics<br/>Preparation 04"] --> D["Add site, daylight, and<br/>elapsed-time context"]
  C --> E["Shared model inputs"]
  D --> E
  E --> F["Primary hypothesis-specific<br/>prepared dataset"]
  G["Gap-timing-unaware<br/>prepared dataset"] --> H["Same hypothesis<br/>implementation"]
  F --> I["Hypothesis analyses"]
  H --> I
```

```{r}
#| label: tbl-render-boundary
#| tbl-cap: "Bounded operations permitted while this report renders."

render_boundary |>
  dplyr::rename(
    `During render` = .data$During_render,
    `Interpretation` = .data$What_it_means
  ) |>
  prep_gt(
    source_note = paste0(
      "This table describes the page's execution boundary; producing ",
      "scripts are not sourced or called."
    )
  ) |>
  gt::cols_width(
    `During render` ~ gt::pct(27),
    Interpretation ~ gt::pct(73)
  )
```

## 1. Standardize participant, questionnaire, and diary structures

The first producing script,
`scripts/pipeline/build_model_input_normalization.R`, reads only the fixed
local files acquired in Preparation 05. Its function module,
`scripts/pipeline/model_input_normalization.R`, places site identifiers,
participant identifiers, labels, column types, and time representations into
a consistent structure. Elapsed time is represented in Coordinated Universal
Time (UTC); time-of-day fields retain the site's local clock time. This
structural work is kept separate from later joins so that recorded values,
missing values, and excluded free text can be checked before metrics and
participant information meet.

The procedure did not recalculate questionnaire totals or factor scores. The
963 field-level value-preservation records all passed, and all 63 audited
free-text fields were excluded from analytical RDS files without exporting
their content.

```{r}
#| label: tbl-normalized-participant-inputs
#| tbl-cap: "Standardized participant, questionnaire, and diary inputs stored before any model-specific selection."

normalized_summary |>
  prep_gt(
    source_note = paste0(
      "Stored manifest: `artifacts/12_manifests/",
      "model_input_normalization.csv`."
    )
  ) |>
  gt::fmt_integer(columns = c(Rows, Columns)) |>
  gt::cols_width(
    `Input type` ~ gt::pct(28),
    Rows ~ gt::pct(9),
    Columns ~ gt::pct(9),
    `Stored file` ~ gt::pct(54)
  )
```

Records without a complete interval remain in their prepared input file, but
they cannot contribute to a calculation that requires both interval
endpoints. The stored audit identifies 27 such light-exposure-diary records
(one at RISE and 26 at THUAS) and one MPI sleep-diary record without a wake
time.

```{r}
#| label: tbl-incomplete-diary-intervals
#| tbl-cap: "Stored diary records that remain available descriptively but cannot support interval-based calculations."

interval_issue_summary |>
  prep_gt(
    source_note = paste0(
      "Read from `artifacts/06_model_data/normalized_inputs/audits/",
      "interval_issue_audit.csv`; no interval is reconstructed during render."
    )
  ) |>
  gt::fmt_integer(columns = Records) |>
  gt::cols_width(
    `Prepared input` ~ gt::pct(22),
    Site ~ gt::pct(12),
    Records ~ gt::pct(10),
    `Why interval calculations cannot use the record` ~ gt::pct(56)
  )
```

## 2. Keep the prepared-data comparison separate

The next script, `scripts/pipeline/build_manuscript_prepared_data.R`, and its
function module, `scripts/pipeline/manuscript_prepared_data.R`, assemble the
gap-timing-unaware dataset independently of the primary dataset. That
separation prevents a preparation difference from being erased when both are
later passed through the same hypothesis implementation.

The stored data cover participant-level, participant-day, 30-minute, and
one-hour values. Exact metric-support hours cannot be recovered from these
retained prepared artifacts; later tables therefore show those hours as
unavailable rather than zero or a reconstructed estimate.

```{r}
#| label: tbl-gap-timing-unaware-artifacts
#| tbl-cap: "Accepted data levels stored for the gap-timing-unaware dataset."

gap_artifact_summary |>
  prep_gt(
    source_note = paste0(
      "Stored manifest: `artifacts/12_manifests/",
      "manuscript_prepared_data_artifacts.csv`. The internal filename is ",
      "retained for provenance."
    )
  ) |>
  gt::fmt_integer(columns = c(Rows, Columns)) |>
  gt::cols_width(
    `Stored data level` ~ gt::pct(22),
    Rows ~ gt::pct(10),
    Columns ~ gt::pct(10),
    `Stored file` ~ gt::pct(58)
  )
```

One objectively verified participant-identifier error is documented rather
than silently changed: record DEV-056 assigns seven TUM exercise-diary rows
to `TUM_S001` instead of the recorded `TUM_S101`. The correction did not
change a scientific light value.

```{r}
#| label: tbl-gap-timing-unaware-correction
#| tbl-cap: "Verified identifier correction recorded for the gap-timing-unaware dataset."

correction_record |>
  prep_gt(
    source_note = paste0(
      "Exact record: `artifacts/06_model_data/scenarios/",
      "manuscript_prepared_data/correction_manifest.csv`."
    )
  ) |>
  gt::fmt_integer(columns = `Diary rows`)
```

## 3. Add site and daylight context

`scripts/pipeline/build_site_solar_context.R` reads the approved site
coordinates and time zones and the local dates required by both prepared
datasets. Its separate function module,
`scripts/pipeline/site_solar_context.R`, stores civil dawn, civil dusk,
photoperiod, and descriptive solar-noon information for 618 site-date
combinations across nine sites, including four site-dates with a
daylight-saving transition.

Every merge uses site and local date as a many-to-one key: many measurement
rows may use one site-date context record. All six joins retained their input
row counts, and none lacked a matching site-date record.

```{r}
#| label: tbl-site-daylight-joins
#| tbl-cap: "Row-preservation checks when site and daylight context was added to the primary metric files."

site_join_summary |>
  prep_gt(
    source_note = paste0(
      "Read from `artifacts/06_model_data/context/",
      "site_solar_context_join_audit.csv`. Relationship: many measurement ",
      "rows to one site-date record."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Rows before join`, `Rows after join`,
      `Participant-days`, `Unmatched rows`
    )
  ) |>
  gt::tab_spanner(
    label = "Join size",
    columns = c(`Rows before join`, `Rows after join`)
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(12),
    Resolution ~ gt::pct(13),
    `Rows before join` ~ gt::pct(15),
    `Rows after join` ~ gt::pct(15),
    `Participant-days` ~ gt::pct(17),
    `Unmatched rows` ~ gt::pct(15),
    Check ~ gt::pct(13)
  )
```

## 4. Preserve time of day and actual elapsed time

Local clock time and elapsed time answer different questions. A 30-minute row
labelled 08:00 represents 08:00 at every study site, which is useful for
time-of-day comparisons. UTC records how much real time separated two rows,
which is necessary for continuity and autocorrelation. The script
`scripts/pipeline/build_temporal_sequence_provenance.R` and function module
`scripts/pipeline/temporal_sequence_provenance.R` preserve both coordinates
without changing melEDI values.

When clocks move back, a local-clock interval may represent two real
intervals; its stored local-clock outcome is their average. When clocks move
forward, a local-clock interval can have no corresponding real interval.
Only an unambiguous, usable real interval can continue an elapsed-time
sequence. A real gap, unusable outcome, repeated autumn interval, or
structural spring gap starts or interrupts a sequence.

```{r}
#| label: tbl-temporal-sequence-provenance
#| tbl-cap: "Stored links between local-clock outcomes and actual elapsed-time intervals."

temporal_summary |>
  prep_gt(
    source_note = paste0(
      "Read from the RDS files in `artifacts/06_model_data/",
      "temporal_provenance/`; the render does not rebuild the links."
    )
  ) |>
  gt::fmt_integer(columns = Value) |>
  gt::cols_width(
    `Stored check or outcome` ~ gt::pct(75),
    Value ~ gt::pct(25)
  )
```

## 5. Combine checked data without choosing a hypothesis sample

`scripts/pipeline/build_base_model_data.R` reads the checked metric files,
standardized participant inputs, and site context. Its function module,
`scripts/pipeline/base_model_data.R`, adds one matching participant record and
one matching site/date record to each measurement row. It does not apply a
complete-case filter or a hypothesis-specific exclusion. Instead, it records
which questionnaire and diary inputs are available, allowing each hypothesis
to state its own requirements later.

The shared primary near-eye input contains
`r dplyr::n_distinct(near_eye_grid$Id)` participants,
`r dplyr::n_distinct(paste(near_eye_grid$Id, near_eye_grid$local_date))`
participant-days, and `r format(nrow(near_eye_grid), big.mark = ",")`
30-minute grid rows. The complementary chest input contains
`r dplyr::n_distinct(chest_grid$Id)` participants,
`r dplyr::n_distinct(paste(chest_grid$Id, chest_grid$local_date))`
participant-days, and `r format(nrow(chest_grid), big.mark = ",")` rows. Chest
measurements remain complementary; they are not pooled with the primary
near-eye measurements.

```{r}
#| label: tbl-shared-site-samples
#| tbl-cap: "Participants, participant-days, and 30-minute grid rows in the shared primary inputs, by submitted-manuscript site."

site_sample_gt <- site_samples |>
  dplyr::select(-.data$site_code, -.data$color_hex) |>
  prep_gt(
    source_note = paste0(
      "Calculated from the stored near-eye and chest RDS files in ",
      "`artifacts/06_model_data/base/`. Admissible rows passed the stored ",
      "30-minute support decision; no admissibility rule is rerun here."
    )
  ) |>
  gt::fmt_integer(columns = -Site) |>
  gt::tab_spanner(
    label = "Primary near-eye input",
    columns = gt::starts_with("Near-eye")
  ) |>
  gt::tab_spanner(
    label = "Complementary chest input",
    columns = gt::starts_with("Chest")
  ) |>
  gt::cols_width(
    Site ~ gt::pct(18),
    gt::starts_with("Near-eye") ~ gt::pct(10.25),
    gt::starts_with("Chest") ~ gt::pct(10.25)
  )

for (row_index in seq_len(nrow(site_samples))) {
  site_sample_gt <- site_sample_gt |>
    gt::tab_style(
      style = gt::cell_borders(
        sides = "left",
        color = site_samples$color_hex[row_index],
        weight = gt::px(5)
      ),
      locations = gt::cells_body(rows = row_index)
    )
}

site_sample_gt
```

All participants with near-eye or chest metrics have matching participant
characteristics, chronotype, LEBA, VLSQ-8, and sleep-diary inputs. Exercise
diary data are absent for three participants at each placement, and
light-exposure diary data are absent for one participant at each placement.
These are availability flags, not exclusions at this stage.

```{r}
#| label: tbl-participant-input-availability
#| tbl-cap: "Availability of questionnaire and diary inputs among participants with stored light-exposure metrics."

availability_summary |>
  prep_gt(
    groupname_col = "Placement",
    source_note = paste0(
      "Read from `artifacts/06_model_data/base/",
      "participant_availability_audit.csv`; filtering_applied is FALSE for ",
      "every participant record."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Participants with light metrics`,
      `Participants with matching input`,
      `Participants without matching input`
    )
  ) |>
  gt::cols_width(
    `Questionnaire or diary input` ~ gt::pct(43),
    `Participants with light metrics` ~ gt::pct(19),
    `Participants with matching input` ~ gt::pct(19),
    `Participants without matching input` ~ gt::pct(19)
  )
```

## 6. Store the first hypothesis-specific inputs

Once the H01 method was approved,
`scripts/pipeline/build_h01_model_data.R` used the shared inputs and the H01
metric contract to store outcomes, site and daylight predictors, participant
identifiers, metric-specific inclusion records, and sample flow. Its function
module is `scripts/pipeline/h01_model_data.R`. The corresponding adapter,
`scripts/pipeline/build_h01_manuscript_prepared_data.R`, applies the same H01
implementation contract to the gap-timing-unaware dataset through
`scripts/pipeline/h01_manuscript_prepared_adapter.R`. Neither operation fits a
model.

H01 covers 17 metric contracts. The following tables report the all-available
near-eye prepared frames. “Rows in prepared frame” is the potential input at
this boundary, not a claim about the final fitted-model sample. Any later
transformation or model-specific missing-value removal is rechecked and
stored by the H01 analysis. Valid measurement hours describe the support from
which a metric was derived; they are not additional model observations.

```{r}
#| label: tbl-h01-primary-prepared-samples
#| tbl-cap: "Available records for the 17 primary near-eye H01 prepared frames."

h01_primary_sample |>
  prep_gt(
    source_note = paste0(
      "Read from `artifacts/06_model_data/H01/sample_flow.csv` and ",
      "`artifacts/06_model_data/H01/metric_contract.csv`."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      Participants, `Rows in prepared frame`,
      `Contributing participant-days`, Sites
    )
  ) |>
  gt::fmt_number(columns = `Valid measurement hours`, decimals = 1) |>
  gt::cols_width(
    Metric ~ gt::pct(31),
    Participants ~ gt::pct(11),
    `Rows in prepared frame` ~ gt::pct(14),
    `Contributing participant-days` ~ gt::pct(17),
    `Valid measurement hours` ~ gt::pct(17),
    Sites ~ gt::pct(10)
  )
```

For the brightest and darkest 10-hour means and their midpoints, the primary
metric files do not export the exact contributing minutes; the corresponding
hour cells remain unavailable. The gap-timing-unaware artifacts do not retain
exact metric-support hours for any H01 metric, so the full column is shown as
unavailable.

```{r}
#| label: tbl-h01-gap-prepared-samples
#| tbl-cap: "Available records for the 17 gap-timing-unaware near-eye H01 prepared frames."

h01_gap_sample |>
  prep_gt(
    source_note = paste0(
      "Read from the sample-flow and metric-contract files under ",
      "`artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/`. ",
      "Support hours are unavailable in the retained prepared artifacts."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      Participants, `Rows in prepared frame`,
      `Contributing participant-days`, Sites
    )
  ) |>
  gt::fmt_number(columns = `Valid measurement hours`, decimals = 1) |>
  gt::cols_width(
    Metric ~ gt::pct(31),
    Participants ~ gt::pct(11),
    `Rows in prepared frame` ~ gt::pct(14),
    `Contributing participant-days` ~ gt::pct(17),
    `Valid measurement hours` ~ gt::pct(17),
    Sites ~ gt::pct(10)
  )
```

The H01 files also prepare paired participant-day comparisons between
near-eye and chest measurements where possible. Interdaily stability and
intradaily variability each produce one participant-level value from a series
of days. Their paired versions would require recomputation from the same
minute-level day set at both placements. They are therefore explicitly marked
unavailable for both placements in both datasets—eight scenario records in
total—instead of being approximated from unequal day sets.

## 7. Describe composition before hypothesis fitting

`scripts/pipeline/build_preanalysis_comparison.R` reads both stored
preparations through `scripts/pipeline/preanalysis_comparison.R`. It records
counts, distributions, key reconciliation, and paired descriptive
comparisons. These outputs help readers see preparation differences before a
hypothesis is tested; they are not hypothesis tests themselves.

@fig-site-composition shows the near-eye participant composition used by this
stored comparison. It is drawn during render from the paired source-data CSV,
using the submitted-manuscript site names, order, and colours. The two panels
use the same horizontal scale.

```{r}
#| label: fig-site-composition
#| fig-cap: "Near-eye participants by study site in the primary and gap-timing-unaware datasets. Counts are descriptive inputs to the stored pre-analysis comparison, not fitted-model sample sizes."
#| fig-alt: "Two stacked horizontal bar panels compare near-eye participant counts by study site. Each site has its submitted-manuscript colour and a count printed at the end of the bar. The primary and gap-timing-unaware datasets use the same horizontal scale."
#| fig-width: 7.5
#| fig-height: 6.8
#| fig-dpi: 144
#| out-width: "100%"

ggplot(
  figure_site_data,
  aes(x = .data$n, y = .data$display_name, fill = .data$level)
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = .data$n),
    hjust = -0.22,
    size = 3.2,
    colour = "#222222"
  ) +
  facet_wrap(vars(.data$Dataset), ncol = 1, scales = "fixed") +
  scale_fill_manual(
    values = stats::setNames(site_registry$color_hex, site_registry$site),
    guide = "none"
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 5),
    expand = expansion(mult = c(0, 0.16))
  ) +
  labs(
    x = "Participants with near-eye metrics",
    y = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text = element_text(size = 9, colour = "#222222"),
    axis.title = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 10, face = "bold", hjust = 0),
    strip.background = element_rect(fill = "#F2F2F2", colour = NA),
    plot.margin = margin(8, 16, 8, 8)
  )
```

[Download the source data for @fig-site-composition](../../artifacts/08_diagnostics/preanalysis_comparison/categorical_levels.csv).

The figure uses a 7.5-inch by 6.8-inch canvas and 9–10 pt axis, tick, facet,
and count labels before HTML scaling. Final-display typography, clipping,
label wrapping, panel balance, bar distinguishability, and colour use are
checked after the HTML render under REPORT-011.

## 8. Verify provenance and hand off stored artifacts

The page checked every file listed in the eight preparation manifests before
displaying any result. A PASS means that the file existed and both its byte
size and SHA-256 digital fingerprint matched the manifest; it does not mean
that the render reran the producing calculation.

```{r}
#| label: tbl-manifest-identity
#| tbl-cap: "File-identity checks read from the eight stored preparation manifests."

manifest_summary |>
  prep_gt(
    source_note = paste0(
      "Every listed file was hashed during render. No manifest or ",
      "analytical artifact was written."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Files listed`, `Files present`, `Size checks`, `SHA-256 checks`
    )
  ) |>
  gt::tab_spanner(
    label = "Files matching the manifest",
    columns = c(`Files present`, `Size checks`, `SHA-256 checks`)
  )
```

The producing scripts are listed below in execution order. Their separation
is scientific provenance: each step has a narrower responsibility and a
stored output that can be checked before the next step begins. Rendering this
report reads those outputs only.

```{r}
#| label: tbl-script-execution-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

script_stage_labels <- c(
  "Standardize participant inputs",
  "Assemble gap-timing-unaware inputs",
  "Add site and daylight context",
  "Preserve elapsed-time provenance",
  "Create shared model inputs",
  "Store H01 primary frames",
  "Store gap-timing-unaware H01 frames",
  "Store pre-analysis displays"
)

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

script_map |>
  dplyr::mutate(
    `Step and producing code` = paste0(
      "**", .data$Step, ". ", script_stage_labels[.data$Step], "**<br>",
      "**Script:** <code>",
      wrap_provenance_path(.data$Producing_script),
      "</code><br>",
      "**Module:** <code>",
      wrap_provenance_path(.data$Function_module),
      "</code>"
    ),
    `Reads, operation, separation, and output` = paste0(
      .data$Reads_and_operation,
      "<br><br>**Stored output:** <code>",
      wrap_provenance_path(.data$Stored_output),
      "</code>"
    )
  ) |>
  dplyr::select(
    .data$`Step and producing code`,
    .data$`Reads, operation, separation, and output`
  ) |>
  prep_gt(
    source_note = paste0(
      "Paths identify the external production code and accepted outputs; ",
      "the render does not source these files."
    )
  ) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

The following files and directories are the durable handoff from Preparation
06. They—not values recalculated by this report—enter the hypothesis-specific
analysis-preparation companions and model-fitting scripts.

```{r}
#| label: tbl-preparation06-handoff
#| tbl-cap: "Stored Preparation 06 artifacts passed to later analysis stages."

handoff_artifacts |>
  dplyr::rename(
    `Stored artifact` = .data$Stored_artifact,
    `Exact path` = .data$Exact_path,
    `Used next for` = .data$Next_use
  ) |>
  prep_gt(
    source_note = paste0(
      "H02–H11 inputs are added only after each hypothesis method is ",
      "approved and are documented in their own analysis-preparation ",
      "companions."
    )
  ) |>
  gt::cols_width(
    `Stored artifact` ~ gt::pct(25),
    `Exact path` ~ gt::pct(43),
    `Used next for` ~ gt::pct(32)
  )
```

```{r}
#| label: tbl-preparation06-environment
#| tbl-cap: "Environment and execution boundary recorded for this review render."

environment_summary |>
  prep_gt(
    source_note = paste0(
      "Rendering uses the project's R 4.6.1 library and the nathealth ",
      "Quarto profile."
    )
  ) |>
  gt::cols_width(
    Component ~ gt::pct(38),
    Value ~ gt::pct(62)
  )
```
