---
title: "Preparation 06: Prepare data for each hypothesis"
---

## What this step does

This step brings questionnaire and diary files into a consistent structure,
combines them with the checked light-exposure metrics and participant and site
information, and prepares the checked inputs needed to build a separate
dataset for each hypothesis. Making the files consistent means using the same
column types, labels and time representations across sites; it does not mean
recalculating questionnaire scores or changing source values. A
hypothesis-specific file is built only after that hypothesis's method has been
approved.

## What goes in and what comes out

- **Inputs:** the fixed source files retrieved in Preparation 05, the checked
  light-exposure metrics, participant and site information, and the analysis
  decisions already approved.
- **Outputs:** checked shared input files and, as each method is approved, one
  R data file (`.rds`) for that hypothesis. H01 is the first
  hypothesis-specific build. Files for H02 through H11, descriptive analyses,
  and the paired sensor-placement comparison are added in their own audit
  stages. Supporting files record which rows were included, how files were
  combined, and what each variable means.
- **Safeguards:** every data merge states whether one or several rows are
  expected to match; every hypothesis has an explicit inclusion flag; free
  text is excluded from analysis files; and no hypothesis chooses its outcomes
  because another hypothesis produced a statistically significant result.

## R code used on this page

```{r}
#| label: setup-model-data

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)
source(file.path(root, "scripts/pipeline/paths_io.R"))
source(file.path(root, "scripts/pipeline/assertions.R"))
source(file.path(root, "scripts/pipeline/import_sources.R"))
source(file.path(root, "scripts/pipeline/model_input_acquisition.R"))
source(file.path(root, "scripts/pipeline/verify_model_input_acquisition.R"))
source(file.path(root, "scripts/pipeline/model_input_normalization.R"))
source(file.path(root, "scripts/pipeline/build_model_input_normalization.R"))
source(file.path(root, "scripts/pipeline/verify_model_input_normalization.R"))
source(file.path(root, "scripts/pipeline/site_solar_context.R"))
source(file.path(root, "scripts/pipeline/build_site_solar_context.R"))
source(file.path(
  root,
  "scripts/pipeline/verify_site_solar_context_artifacts.R"
))
source(file.path(
  root,
  "scripts/pipeline/temporal_sequence_provenance.R"
))
source(file.path(
  root,
  "scripts/pipeline/build_temporal_sequence_provenance.R"
))
source(file.path(
  root,
  "scripts/pipeline/verify_temporal_sequence_provenance_artifacts.R"
))
source(file.path(root, "scripts/pipeline/base_model_data.R"))
source(file.path(root, "scripts/pipeline/build_base_model_data.R"))
source(file.path(
  root,
  "scripts/pipeline/verify_base_model_data_artifacts.R"
))
source(file.path(root, "scripts/pipeline/metric_display_registry.R"))
source(file.path(root, "scripts/pipeline/h01_model_data.R"))
source(file.path(root, "scripts/pipeline/build_h01_model_data.R"))
source(file.path(
  root,
  "scripts/pipeline/verify_h01_model_data_artifacts.R"
))
source(file.path(root, "scripts/pipeline/manuscript_prepared_data.R"))
source(file.path(
  root,
  "scripts/pipeline/build_manuscript_prepared_data.R"
))
source(file.path(
  root,
  "scripts/pipeline/verify_manuscript_prepared_data_artifacts.R"
))
source(file.path(
  root,
  "scripts/pipeline/h01_manuscript_prepared_adapter.R"
))
source(file.path(
  root,
  "scripts/pipeline/build_h01_manuscript_prepared_data.R"
))
source(file.path(
  root,
  "scripts/pipeline/verify_h01_manuscript_prepared_data_artifacts.R"
))
source(file.path(root, "scripts/pipeline/preanalysis_comparison.R"))
source(file.path(
  root,
  "scripts/pipeline/build_preanalysis_comparison.R"
))
source(file.path(
  root,
  "scripts/pipeline/verify_preanalysis_comparison_artifacts.R"
))

paths <- pipeline_paths(root)
```

## Check and standardize questionnaire and diary files

Only the exact source-file versions retrieved and checked in Preparation 05
are used. Each source-file version is fixed to a specific repository revision,
so a later change to an online file cannot silently change this analysis. A
file inventory records the source revision, file size and SHA-256 checksum.
A SHA-256 checksum is a digital fingerprint: identical files have the same
fingerprint.

Before files from different sites are combined, each row receives its site and
source-version labels. Time intervals are stored both as Coordinated Universal
Time (UTC), which is used to calculate elapsed time, and as local clock time,
which is used to compare times of day across sites. The time at which a form
was submitted is kept as descriptive information but is not treated as an
exposure interval. Free-text responses are checked for presence and missing
values but are not copied into the analysis files.

```{r}
#| label: build-normalized-model-inputs

normalization <- build_model_input_normalization(root = root)
normalization_verification <- verify_model_input_normalization(root = root)
stopifnot(identical(normalization_verification$status, "PASS"))
```

```{r}
#| label: normalized-model-input-summary

prepared_file_labels <- c(
  normalized_demographics = "Participant characteristics",
  normalized_chronotype = "Chronotype questionnaire",
  normalized_leba = "Light-exposure behaviour questionnaire",
  normalized_vlsq8 = "Visual light sensitivity questionnaire",
  normalized_exercisediary = "Exercise diary",
  normalized_lightexposurediary = "Light-exposure diary",
  normalized_sleepdiaries = "Sleep diary",
  audit_source_schema = "Check of source columns and types",
  audit_labels = "Check of labels",
  audit_keys = "Check of participant and record identifiers",
  audit_missingness = "Check of missing values",
  audit_intervals = "Check of diary time intervals",
  audit_interval_issues = "Diary intervals requiring review",
  audit_free_text = "Check that free text is excluded",
  audit_value_preservation = "Check that source values are unchanged"
)

normalized_file_summary <- normalization_verification$artifact_summary |>
  dplyr::transmute(
    `Prepared file` = dplyr::recode(
      .data$artifact_id,
      !!!prepared_file_labels
    ),
    `File type` = toupper(.data$artifact_type),
    Rows = .data$rows,
    Columns = .data$columns,
    `Digital fingerprint (SHA-256)` = .data$sha256
  )
knitr::kable(
  normalized_file_summary
)

interval_summary <- normalization_verification$interval_summary |>
  dplyr::transmute(
    Source = dplyr::recode(
      .data$modality,
      exercisediary = "Exercise diary",
      lightexposurediary = "Light-exposure diary",
      sleepdiaries = "Sleep diary"
    ),
    `Use of timestamp` = dplyr::recode(
      .data$timestamp_role,
      analysis_interval = "Start and end of an analysis interval",
      form_metadata = "Time when the form was submitted"
    ),
    Rows = .data$rows,
    `Usable interval rows` = .data$analysis_eligible,
    `Rows missing an interval endpoint` = .data$quarantined,
    `Form-submission timestamps excluded` =
      .data$excluded_as_form_metadata
  )
knitr::kable(
  interval_summary
)
```

Twenty-seven light-diary records do not contain both a start and an end time:
one record from RISE and 26 from THUAS. These records remain in the prepared
file, but they are excluded from calculations that require a complete time
interval. One MPI sleep-diary record without a wake time is handled in the
same way. No questionnaire total or factor score is recalculated.

The checked manuscript-prepared sensitivity files are also assembled here,
before daylight information is calculated. This lets the daylight table cover
the dates required by both datasets while keeping the light-exposure values in
the two datasets separate.

```{r}
#| label: build-manuscript-prepared-inputs

manuscript_prepared_build <- build_manuscript_prepared_data(
  root = root,
  output_root = root
)
manuscript_prepared_verification <-
  verify_manuscript_prepared_data_artifacts(
    root = root,
    output_root = root
  )
stopifnot(
  manuscript_prepared_verification$status == "PASS",
  manuscript_prepared_verification$model_implementation_id ==
    "new_h01_h11"
)

supplemental_site_date_paths <- stats::setNames(
  manuscript_prepared_build$paths$rds[["participant_day_metrics"]],
  "manuscript_prepared_participant_day"
)
```

## Add site and daylight information

The analysis uses the checked location and time-zone information for all nine
study sites. Daylight information is calculated for every site and local date
required by either the main data or the manuscript-prepared sensitivity. Two
IZTECH dates occur only in the latter because the main analysis excludes their
complete day-long flat-zero melEDI records. Keeping those dates in the
daylight table allows the registered all-zero-inclusive comparison to be run;
it does not return them to the main data. Before these values are saved, the
main daily, 30-minute and one-hour files must contain the same sets of
participant-days. Civil dawn, civil dusk and photoperiod are defined using a
solar altitude of −6 degrees.

Solar noon is calculated from midnight in each site's local time zone. This
avoids accidentally assigning solar noon to the preceding UTC date for sites
west of Greenwich. Solar noon is retained only as descriptive site
information, not as an analysis variable. The actual instant is stored in UTC;
the corresponding local time, UTC offset and daylight-saving status are
stored separately.

```{r}
#| label: build-site-solar-context

site_solar_build <- build_site_solar_context_artifacts(
  root = root,
  supplemental_date_paths = supplemental_site_date_paths
)
site_solar_verification <- verify_site_solar_context_artifacts(
  root = root,
  supplemental_date_paths = supplemental_site_date_paths
)
stopifnot(identical(site_solar_verification$status, "PASS"))
```

```{r}
#| label: site-solar-context-summary

site_solar_summary <- data.frame(
  check = c(
    "Overall check",
    "Site-date combinations",
    "Study sites",
    "Site-dates containing a daylight-saving change",
    "Data files checked before adding daylight information",
    "Rows without matching site/date information",
    "Files in the output inventory"
  ),
  value = unlist(site_solar_verification[c(
    "status",
    "site_dates",
    "sites",
    "dst_transition_site_dates",
    "join_audit_rows",
    "unmatched_rows",
    "manifest_rows"
  )])
)
knitr::kable(site_solar_summary)

site_join_summary <- site_solar_build$join_audit |>
  dplyr::transmute(
    Input = dplyr::recode(
      .data$input_role,
      glasses_daily = "Near-eye daily metrics",
      glasses_30_minute = "Near-eye 30-minute values",
      glasses_one_hour = "Near-eye one-hour values",
      chest_daily = "Chest daily metrics",
      chest_30_minute = "Chest 30-minute values",
      chest_one_hour = "Chest one-hour values"
    ),
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    `Time resolution` = dplyr::recode(
      .data$resolution,
      daily = "One value per participant-day",
      `30_minute` = "30 minutes",
      one_hour = "One hour"
    ),
    `Matching fields` = dplyr::recode(
      .data$join_key,
      `site|local_date` = "Site and local date"
    ),
    `Expected match` = dplyr::recode(
      .data$relationship,
      `many-to-one` = "One site-date record for each input row"
    ),
    `Rows before` = .data$input_rows,
    `Rows after` = .data$joined_rows,
    `Participant-days` = .data$participant_days,
    `Site-date combinations` = .data$distinct_site_dates,
    `Rows without a match` = .data$unmatched_rows,
    Check = .data$status
  )
knitr::kable(
  site_join_summary
)
```

## Preserve local clock time and actual elapsed time

The H02 and H11 data use a shared 30-minute local-clock grid: for example,
08:00 means 08:00 local time at every site. This is the appropriate scale for
comparing time of day, but it does not always show how much real time passed
between two rows.

Daylight-saving changes create the important exceptions. When clocks move
back in autumn, one local half-hour occurs twice; the local-clock value is the
average of those two real intervals. When clocks move forward in spring, some
local-clock rows have no corresponding real interval. The link created below
preserves both views without changing any exposure value: local clock time for
time-of-day comparisons and UTC for gaps, continuity and autocorrelation.

```{r}
#| label: build-temporal-sequence-provenance

temporal_provenance <- build_temporal_sequence_provenance_artifacts(
  root = root,
  run_label = "full"
)
temporal_verification <-
  verify_temporal_sequence_provenance_artifacts(
    root = root,
    stop_on_failure = TRUE
  )
stopifnot(identical(temporal_verification$status, "PASS"))
```

```{r}
#| label: tbl-temporal-sequence-provenance
#| tbl-cap: "Checks linking local-clock rows to actual elapsed time."

temporal_summary <- tibble::tibble(
  check = c(
    "Original intervals stored in UTC",
    "Rows on the shared local-clock grid",
    "Participant-days spanning the autumn clock change",
    "Local-clock rows that average two real intervals",
    "Participant-days spanning the spring clock change",
    "Local-clock rows with no real interval during the spring gap",
    "Original intervals without a one-to-one local-clock row",
    "Intervals usable in continuous-sequence models",
    "Starts of continuous sequences"
  ),
  value = c(
    temporal_verification$source_bin_rows,
    temporal_verification$wall_link_rows,
    temporal_verification$fold_participant_days,
    temporal_verification$fold_wall_rows,
    temporal_verification$spring_gap_participant_days,
    temporal_verification$spring_gap_wall_rows,
    temporal_verification$non_one_to_one_source_bins,
    temporal_verification$sequence_eligible_bins,
    temporal_verification$sequence_starts
  )
)

knitr::kable(temporal_summary)
```

Only rows with one unambiguous real interval receive a position in a continuous
sequence. A real gap, a missing spring-clock interval, an unusable value or an
averaged autumn-clock row starts a new sequence. H02 and H11 may still use
local clock time as a predictor, but any autocorrelation term---which describes
dependence between neighbouring observations---must follow actual elapsed
time. This step prepares that information; it does not fit a model.

## Combine data without choosing model samples

The light-exposure metric files keep their original rows and order. Each
measurement row receives one matching set of site/date information and one
matching set of participant questionnaire information. The checks stop if a
merge would unexpectedly duplicate or lose rows.

No complete-case filtering or hypothesis-specific exclusion happens here.
The diary files contribute only indicators showing whether data are available;
the diary variables needed for a particular hypothesis are added later when
that hypothesis dataset is prepared.

```{r}
#| label: build-base-model-data

base_model_build <- build_base_model_data_artifacts(root = root)
base_model_verification <- verify_base_model_data_artifacts(root = root)
stopifnot(identical(base_model_verification$status, "PASS"))
```

```{r}
#| label: base-model-data-summary

base_file_labels <- c(
  participant_metadata = "Participant information",
  glasses_participant_day_context = "Near-eye daily metrics with site context",
  chest_participant_day_context = "Chest daily metrics with site context",
  glasses_participant = "Near-eye participant metrics",
  chest_participant = "Chest participant metrics",
  glasses_30_minute_context = "Near-eye 30-minute values with site context",
  chest_30_minute_context = "Chest 30-minute values with site context",
  glasses_one_hour_context = "Near-eye one-hour values with site context",
  chest_one_hour_context = "Chest one-hour values with site context",
  glasses_participant_day_enriched =
    "Near-eye daily metrics with participant information",
  chest_participant_day_enriched =
    "Chest daily metrics with participant information",
  glasses_participant_enriched =
    "Near-eye participant metrics with participant information",
  chest_participant_enriched =
    "Chest participant metrics with participant information",
  input_provenance = "Exact input-file record",
  metric_registry_firewall = "Check of metrics allowed into models",
  metric_key_join = "Check of metric identifiers",
  modality_key_join = "Check of questionnaire and diary identifiers",
  participant_availability =
    "Availability of questionnaire and diary data by participant",
  metadata_fields = "Data dictionary for participant information",
  metric_metadata_availability =
    "Availability of participant information for each metric file",
  contract = "Checks required before hypothesis datasets are prepared"
)

base_file_summary <- base_model_verification$artifact_summary |>
  dplyr::transmute(
    `Prepared dataset or check` = dplyr::recode(
      .data$artifact_id,
      !!!base_file_labels
    ),
    Rows = .data$rows,
    Columns = .data$columns,
    Check = .data$status
  )
knitr::kable(
  base_file_summary
)

participant_data_availability <- base_model_verification$anomaly_summary |>
  dplyr::transmute(
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    `Questionnaire or diary source` = dplyr::recode(
      .data$modality,
      demographics = "Participant characteristics",
      chronotype = "Chronotype questionnaire",
      leba = "Light-exposure behaviour questionnaire",
      vlsq8 = "Visual light sensitivity questionnaire",
      exercisediary = "Exercise diary",
      lightexposurediary = "Light-exposure diary",
      sleepdiaries = "Sleep diary"
    ),
    `Participants with light metrics` = .data$metric_participants,
    `Participants with matching source data` =
      .data$metric_participants_present,
    `Participants without matching source data` =
      .data$metric_participants_absent
  )
knitr::kable(
  participant_data_availability
)
```

## Prepare the H01 data

H01 compares 17 light-exposure metrics across sites. The file prepared here
contains the outcome, site, day length, latitude and participant identifier
needed for each H01 model. It also records why a row is excluded from a
particular metric. No model is fitted in this step.

Two analysis samples are prepared from the main dataset:

- **all available near-eye data**, used for the primary analysis; and
- **paired participant-days**, used to compare near-eye and chest measurements
  on days available at both positions.

IS and IV have one value per participant rather than one value per day. Their
paired versions must therefore be recalculated from the same set of days at
both positions; they are marked as unavailable until that calculation is
implemented, rather than being approximated from unequal day sets.

```{r}
#| label: build-h01-model-data

h01_build <- build_h01_model_data_artifacts(
  root = root,
  output_root = root
)
h01_verification <- verify_h01_model_data_artifacts(
  root = root,
  output_root = root
)
stopifnot(
  h01_verification$status ==
    "PASS_WITH_DECLARED_UNAVAILABLE_SCENARIO"
)
```

```{r}
#| label: tbl-h01-primary-sample
#| tbl-cap: "Data available for each primary near-eye H01 model."

h01_primary_sample <- h01_verification$primary_model_sample |>
  dplyr::left_join(
    h01_build$object$metric_contract |>
      dplyr::select("metric_id", "manuscript_name"),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::arrange(
    match(
      .data$metric_id,
      h01_build$object$metric_contract$metric_id
    )
  ) |>
  dplyr::transmute(
    Metric = .data$manuscript_name,
    Participants = .data$participants,
    `Rows used by the model` = .data$model_observations,
    `Participant-days contributing to the metric` =
      .data$contributing_participant_days,
    `Valid measurement hours contributing to the metric` =
      round(.data$metric_support_valid_hours, 1),
    `Expected measurement hours for that metric` =
      round(.data$metric_support_expected_hours, 1),
    Sites = .data$sites
  )

knitr::kable(h01_primary_sample)
```

The measurement hours in this table show how much recorded light data
contributed to a derived outcome. They are not additional observations in the
model. The H01 models use either participants or participant-days, as stated
in the “Rows used by the model” column. The metric files do not currently
export the exact contributing minutes for the brightest and darkest 10-hour
means and midpoints, so those hour counts are shown as unavailable instead of
being replaced with all recorded hours.

## Prepare the manuscript-prepared-data sensitivity

The sensitivity dataset keeps the light-exposure values that were prepared
for the manuscript and corrects only an objectively verified participant
identifier in the exercise diary. It preserves the manuscript-prepared
definitions of dose and MDER, because changing those definitions here would
remove the preparation difference that this sensitivity is designed to test.
The metric crosswalk states those differences explicitly.

Both this dataset and the main dataset will be analysed with the same new
H01--H11 implementation. The manuscript-generating model code is not part of
this sensitivity.

| Role | Prepared data | H01--H11 models | What the comparison answers |
|---|---|---|---|
| Main analysis | Data prepared with the audited rules | New implementation | Primary estimates and conclusions |
| Manuscript-prepared-data sensitivity | Data prepared for the manuscript, with only indisputable data errors corrected | The same new implementation | How much the changed data preparation and metric calculation affect the new results |
| Audit discussion only | Reported manuscript values and their generating code | Manuscript-generating implementation | Why a reported result differs; this is not a formal sensitivity scenario |

```{r}
#| label: build-manuscript-prepared-data

stopifnot(
  manuscript_prepared_verification$status == "PASS",
  manuscript_prepared_verification$model_implementation_id ==
    "new_h01_h11"
)

h01_manuscript_prepared_build <-
  build_h01_manuscript_prepared_data_artifacts(
    root = root,
    output_root = root
  )
h01_manuscript_prepared_verification <-
  verify_h01_manuscript_prepared_data_artifacts(
    root = root,
    output_root = root
  )
stopifnot(
  h01_manuscript_prepared_verification$status ==
    paste0(
      "PASS_WITH_DECLARED_UNAVAILABLE_SUPPORT_",
      "AND_PAIRED_PARTICIPANT_METRICS"
    ),
  h01_manuscript_prepared_verification$model_implementation_id ==
    h01_build$object$metadata$model_implementation_id,
  h01_manuscript_prepared_verification$implementation_contract_sha256 ==
    h01_build$object$metadata$implementation_contract_sha256,
  h01_manuscript_prepared_verification$shared_implementation_sha256 ==
    h01_build$object$metadata$shared_implementation_sha256
)
```

```{r}
#| label: tbl-manuscript-prepared-data
#| tbl-cap: "Rows retained in the checked manuscript-prepared-data sensitivity inputs."

manuscript_prepared_counts <- tibble::tibble(
  `Type of value` = c(
    "Participant-level metrics",
    "Participant-day metrics",
    "30-minute light values",
    "One-hour light values"
  ),
  Rows = unname(
    manuscript_prepared_verification$output_rows[c(
      "participant_metrics",
      "participant_day_metrics",
      "thirty_minute_data",
      "one_hour_data"
    )]
  )
)
knitr::kable(manuscript_prepared_counts)
```

```{r}
#| label: tbl-h01-manuscript-prepared-sample
#| tbl-cap: "Data available for each primary near-eye H01 model in the manuscript-prepared-data sensitivity."

h01_manuscript_prepared_sample <-
  h01_manuscript_prepared_verification$model_samples |>
  dplyr::filter(
    .data$placement == "glasses",
    .data$scenario == "all_available"
  ) |>
  dplyr::left_join(
    h01_manuscript_prepared_build$object$metric_contract |>
      dplyr::select("metric_id", "manuscript_name"),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::arrange(
    match(
      .data$metric_id,
      h01_manuscript_prepared_build$object$metric_contract$metric_id
    )
  ) |>
  dplyr::transmute(
    Metric = .data$manuscript_name,
    Participants = .data$participants,
    `Rows available to the model` = .data$model_observations,
    `Participant-days contributing to the metric` =
      .data$contributing_participant_days,
    `Valid measurement hours contributing to the metric` =
      "Not available from retained prepared artifacts",
    Sites = .data$sites
  )

knitr::kable(h01_manuscript_prepared_sample)
```

The table reports the prepared model frames, not the final fitted-model
sample. Every fitted model will store its exact analysis frame and recheck its
participants, participant-days or contributing days, observations and sites
after any transformation or missing-value removal. Exact metric-support hours
cannot be recovered from the retained manuscript-prepared metric artifacts,
so they are shown as unavailable rather than as zero.

When daylight saving time ends, the same local clock hour occurs twice.
Following the manuscript preparation, measurements from those repeated hours
are averaged into one local-hour value. This keeps every site on the same
shared local-clock scale and prevents the repeated hour from contributing
twice to a clock-time analysis. The underlying occurrence information is kept
only to document how that average was formed.

## Metric names used in the manuscript

The code uses short internal identifiers to avoid ambiguity. Tables and figures
use the familiar metric names and categories from the manuscript. Details such
as coverage threshold, correction method, censoring or sensitivity analysis
are shown in separate columns rather than being added to the metric name.

```{r}
#| label: tbl-metric-display-registry
#| tbl-cap: "Light-exposure metric names and categories used in the manuscript."

metric_display <- read_metric_display_registry(root) |>
  dplyr::mutate(
    analysis_unit = dplyr::recode(
      .data$analysis_unit,
      `participant-30-minute` = "30-minute observation",
      `participant-hour` = "One-hour observation",
      `participant-day` = "Participant-day",
      participant = "Participant"
    ),
    analytical_role = dplyr::recode(
      .data$analytical_role,
      primary_outcome = "Primary outcome",
      descriptive_only = "Descriptive result",
      sensitivity_outcome = "Sensitivity result",
      diagnostic_outcome = "Diagnostic result"
    )
  )
knitr::kable(
  metric_display[c(
    "manuscript_name",
    "abbreviation",
    "manuscript_category",
    "analysis_unit",
    "display_unit",
    "analytical_role",
    "variant_label"
  )],
  col.names = c(
    "Metric",
    "Abbreviation",
    "Category",
    "One value per",
    "Unit",
    "How used",
    "Analysis version"
  )
)
```

## Compare the manuscript-prepared and main datasets

Before fitting any hypothesis model, we compare the data produced by the
manuscript preparation with the newly prepared data for the main analysis.
This is a descriptive comparison of available observations and distributions;
it does not test a scientific hypothesis.

A direct mapping means only that two variables have the same intended subject
matter. It does not prove that they were calculated in exactly the same way.
If a metric was newly defined or changed enough that a direct comparison would
be misleading, the manuscript-prepared value is shown as “not applicable”
rather than invented.

The analysis stores a digital fingerprint for each input-file inventory. The
comparison stops if either fingerprint changes, so results from a changed
input cannot be mistaken for the version reviewed here.

```{r}
#| label: build-preanalysis-comparison

expected_normalization_manifest_sha256 <-
  "e3d484711abb54f69d63ff302e5a0329efda3fd0a8c85feaf9cc4f850005c9ab"
expected_base_manifest_sha256 <-
  "142d1044dd7e6a566e45e3a5c56033b1e8251e100c0d7c7523564c824950a0e4"

stopifnot(
  identical(
    normalization_verification$normalization_manifest_sha256,
    expected_normalization_manifest_sha256
  ),
  identical(
    base_model_verification$manifest_sha256,
    expected_base_manifest_sha256
  )
)

preanalysis_build <- build_preanalysis_comparison_artifacts(
  root = root,
  output_root = root,
  expected_normalization_manifest_sha256 =
    expected_normalization_manifest_sha256,
  expected_base_manifest_sha256 = expected_base_manifest_sha256
)
preanalysis_verification <- verify_preanalysis_comparison_artifacts(
  root = root,
  output_root = root,
  expected_normalization_manifest_sha256 =
    expected_normalization_manifest_sha256,
  expected_base_manifest_sha256 = expected_base_manifest_sha256
)
stopifnot(identical(preanalysis_verification$status, "PASS"))
```

```{r}
#| label: tbl-preanalysis-gate-summary
#| tbl-cap: "Checks completed before comparing the manuscript-prepared and main datasets."

preanalysis_gate_summary <- data.frame(
  check = c(
    "Overall check",
    "Variables compared",
    "Light-exposure metrics",
    "Variable-by-placement combinations",
    "Variables directly matched to the manuscript-prepared data",
    "Variables without a meaningful manuscript-prepared counterpart",
    "Files in the comparison inventory",
    "Comparison-inventory digital fingerprint (SHA-256)"
  ),
  value = c(
    preanalysis_verification$status,
    preanalysis_verification$variables,
    nrow(metric_display),
    preanalysis_verification$overview_rows,
    preanalysis_verification$applicable_variables,
    preanalysis_verification$not_applicable_variables,
    preanalysis_verification$artifacts,
    preanalysis_verification$manifest_sha256
  )
)
knitr::kable(preanalysis_gate_summary)
```

The complete comparison can be downloaded as
[`comparison_overview.csv`](../../artifacts/08_diagnostics/preanalysis_comparison/comparison_overview.csv).
It includes every metric and key analysis variable, with numbers of
observations, participants and participant-days. For numerical variables it
also reports the mean, median, spread and quantiles; for categorical variables
it reports the proportion in each category. Complete numerical quantiles are in
[`numeric_quantiles.csv`](../../artifacts/08_diagnostics/preanalysis_comparison/numeric_quantiles.csv),
and complete category proportions are in
[`categorical_levels.csv`](../../artifacts/08_diagnostics/preanalysis_comparison/categorical_levels.csv).

```{r}
#| label: tbl-preanalysis-count-comparison
#| tbl-cap: "Available observations, participants and participant-days in the manuscript-prepared and main datasets."

comparison_overview <- readr::read_csv(
  preanalysis_build$paths$csv[["comparison_overview"]],
  show_col_types = FALSE
)

display_placement <- function(value) {
  dplyr::recode(
    value,
    all = "All placements",
    glasses = "Near-eye",
    chest = "Chest"
  )
}

display_analysis_unit <- function(value) {
  dplyr::recode(
    value,
    `30_minute` = "30-minute observation",
    diary_interval = "Diary interval",
    one_hour = "One-hour observation",
    participant = "Participant",
    participant_day = "Participant-day"
  )
}

display_value_type <- function(value) {
  dplyr::recode(
    value,
    numeric = "Numerical",
    categorical = "Categorical"
  )
}

display_comparability <- function(value) {
  dplyr::recode(
    value,
    exact_field_key_and_scale_mapping =
      "Yes: same field, records, and scale",
    same_nominal_construct_changed_computation_admissibility_descriptive_only =
      "Descriptive comparison only: calculation or data support changed",
    changed_estimand_or_no_direct_mapping_no_numerical_comparison =
      "No: definition changed or no matching manuscript-prepared value"
  )
}

display_variant <- function(value) {
  value
}

format_integer <- function(value) {
  ifelse(is.na(value), "\u2014", format(value, trim = TRUE, scientific = FALSE))
}

format_number <- function(value) {
  vapply(
    value,
    function(item) {
      if (is.na(item)) {
        return("\u2014")
      }
      if (item != 0 && (abs(item) >= 10000 || abs(item) < 0.001)) {
        return(formatC(item, format = "e", digits = 2))
      }
      format(round(item, 3), trim = TRUE, scientific = FALSE)
    },
    character(1)
  )
}

format_transition <- function(submitted, audited, delta, formatter) {
  paste0(
    formatter(submitted),
    " \u2192 ",
    formatter(audited),
    " (\u0394 ",
    formatter(delta),
    ")"
  )
}

count_comparison <- comparison_overview |>
  dplyr::transmute(
    Variable = .data$manuscript_name,
    Variant = display_variant(.data$variant_label),
    Placement = display_placement(.data$placement),
    Unit = display_analysis_unit(.data$analysis_unit),
    Type = display_value_type(.data$value_type),
    estimable_n = format_transition(
      .data$baseline_n_estimable,
      .data$canonical_n_estimable,
      .data$delta_n_estimable,
      format_integer
    ),
    participants = format_transition(
      .data$baseline_n_participants_observed,
      .data$canonical_n_participants_observed,
      .data$delta_n_participants_observed,
      format_integer
    ),
    participant_days = format_transition(
      .data$baseline_n_participant_days_observed,
      .data$canonical_n_participant_days_observed,
      .data$delta_n_participant_days_observed,
      format_integer
    ),
    Comparability = display_comparability(.data$comparability_class)
  )

knitr::kable(
  count_comparison,
  col.names = c(
    "Variable",
    "Variant",
    "Placement",
    "Unit",
    "Type",
    "Available values: manuscript-prepared \u2192 main",
    "Participants: manuscript-prepared \u2192 main",
    "Participant-days: manuscript-prepared \u2192 main",
    "Can the values be compared directly?"
  )
)
```

```{r}
#| label: tbl-preanalysis-numeric-comparison
#| tbl-cap: "Means, medians and distributions in the manuscript-prepared and main datasets."

numeric_comparison <- comparison_overview |>
  dplyr::filter(.data$value_type == "numeric") |>
  dplyr::transmute(
    Variable = .data$manuscript_name,
    Variant = display_variant(.data$variant_label),
    Placement = display_placement(.data$placement),
    display_unit = .data$display_unit,
    mean = paste0(
      format_number(.data$baseline_mean),
      " \u2192 ",
      format_number(.data$canonical_mean)
    ),
    median = paste0(
      format_number(.data$baseline_median),
      " \u2192 ",
      format_number(.data$canonical_median)
    ),
    iqr = paste0(
      "[",
      format_number(.data$baseline_q25),
      ", ",
      format_number(.data$baseline_q75),
      "] \u2192 [",
      format_number(.data$canonical_q25),
      ", ",
      format_number(.data$canonical_q75),
      "]"
    ),
    largest_cumulative_difference = format_number(
      .data$ecdf_max_distance
    ),
    average_shift_relative_to_iqr = format_number(
      .data$wasserstein_1_iqr_scaled
    )
  )

knitr::kable(
  numeric_comparison,
  col.names = c(
    "Variable",
    "Variant",
    "Placement",
    "Display unit",
    "Mean: manuscript-prepared \u2192 main",
    "Median: manuscript-prepared \u2192 main",
    "Middle 50%: manuscript-prepared \u2192 main",
    "Largest cumulative-distribution difference (0 = none)",
    "Average distribution shift relative to the manuscript-prepared middle 50%"
  )
)
```

```{r}
#| label: tbl-preanalysis-categorical-comparison
#| tbl-cap: "Category frequencies in the manuscript-prepared and main datasets."

categorical_comparison <- comparison_overview |>
  dplyr::filter(.data$value_type == "categorical") |>
  dplyr::transmute(
    Variable = .data$manuscript_name,
    Variant = display_variant(.data$variant_label),
    Placement = display_placement(.data$placement),
    estimable_n = format_transition(
      .data$baseline_n_estimable,
      .data$canonical_n_estimable,
      .data$delta_n_estimable,
      format_integer
    ),
    total_variation_distance = format_number(
      .data$total_variation_distance
    ),
    maximum_level_proportion_change = format_number(
      .data$maximum_absolute_proportion_difference
    )
  )

knitr::kable(
  categorical_comparison,
  col.names = c(
    "Variable",
    "Variant",
    "Placement",
    "Available values: manuscript-prepared \u2192 main",
    "Overall difference in category proportions",
    "Largest change for any one category"
  )
)
```

For numerical variables, the largest cumulative-distribution difference
describes the greatest separation between the two distributions at any value;
it ranges from 0 (no observed difference) to 1 (complete separation). The
average distribution shift is expressed relative to the manuscript-prepared
interquartile range, which contains the middle 50% of those values.
For categorical variables, the overall difference in category proportions
ranges from 0 (identical proportions) to 1 (no overlap). These measures are
descriptive flags for closer inspection, not hypothesis tests.

The following panels show the same comparisons visually. They are analysis
checks rather than submission figures, and the exact plotted values are saved
beside each figure as a CSV file.

```{r}
#| label: fig-preanalysis-numeric-distributions
#| fig-cap: "Numerical distributions in the manuscript-prepared and main datasets, shown on a common standardized scale."
#| out-width: "100%"

numeric_distribution_path <- normalizePath(
  preanalysis_build$paths$figures[["numeric_distribution_overview"]],
  winslash = "/",
  mustWork = TRUE
)
stopifnot(file.exists(numeric_distribution_path))
knitr::include_graphics(
  numeric_distribution_path,
  error = FALSE
)
```

```{r}
#| label: fig-preanalysis-categorical-distributions
#| fig-cap: "Category proportions in the manuscript-prepared and main datasets."
#| out-width: "100%"

categorical_distribution_path <- normalizePath(
  preanalysis_build$paths$figures[["categorical_distribution_overview"]],
  winslash = "/",
  mustWork = TRUE
)
stopifnot(file.exists(categorical_distribution_path))
knitr::include_graphics(
  categorical_distribution_path,
  error = FALSE
)
```

## What happens next

1. Check column structure, labels, identifiers, missing values, usable time
   intervals and exact preservation of source values.
2. Confirm that participant, site, participant-day and repeated-observation
   identifiers are unique where expected.
3. Add demographics, chronotype, VLSQ-8, LEBA, experience-sampling, exercise
   and sleep variables, while checking that no merge duplicates or loses rows.
4. Record variable definitions, units, categories and reference groups in a
   data dictionary.
5. Save one analysis dataset and one sample-flow table for each hypothesis.
6. Create a paired sensor-placement dataset for participants with both
   placements. This comparison does not pool chest measurements into the
   primary near-eye analysis.
