---
title: "Preparation 06: Build model-ready datasets"
subtitle: "From checked metrics and participant information to hypothesis-specific inputs"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-06
#| include: false

library(digest)
library(dplyr)
library(ggplot2)
library(gt)
library(readr)
library(scales)
library(tibble)
library(tidyr)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR"),
  winslash = "/",
  mustWork = TRUE
)

stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE)
}

read_stored_rds <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readRDS(full_path)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

prep_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )

  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_hash <- function(value) {
  gsub("(.{8})", "\\1\u200b", value, perl = TRUE)
}

wrap_provenance_path <- function(path) {
  gsub("/", "/<wbr>", path, fixed = TRUE)
}

check_file_hash <- function(path, expected_sha256) {
  observed <- digest::digest(
    file = absolute_path(path),
    algo = "sha256"
  )
  stopifnot(identical(observed, expected_sha256))
  observed
}

check_evidence_manifest <- function(path, expected_sha256, family) {
  manifest <- read_stored_csv(path)
  assert_columns(
    manifest,
    c("path", "sha256", "bytes", "producer", "r_version", "status"),
    family
  )
  check_file_hash(path, expected_sha256)
  listed_paths <- absolute_path(manifest$path)
  present <- file.exists(listed_paths)
  actual_bytes <- rep(NA_real_, length(listed_paths))
  actual_sha256 <- rep(NA_character_, length(listed_paths))
  actual_bytes[present] <- file.info(listed_paths[present])$size
  actual_sha256[present] <- vapply(
    listed_paths[present],
    function(file) digest::digest(file = file, algo = "sha256"),
    character(1)
  )
  stopifnot(
    all(manifest$r_version == "4.6.1"),
    all(manifest$status == "PASS"),
    all(present),
    all(actual_bytes == manifest$bytes),
    all(actual_sha256 == manifest$sha256)
  )
  list(
    data = manifest,
    summary = tibble::tibble(
      stage = family,
      `Manifest SHA-256` = expected_sha256,
      `Files listed` = nrow(manifest),
      `Files present` = sum(present),
      `Size checks` = sum(actual_bytes == manifest$bytes),
      `SHA-256 checks` = sum(actual_sha256 == manifest$sha256),
      Check = "PASS"
    )
  )
}

manifest_registry <- tibble::tribble(
  ~stage_order, ~stage, ~manifest, ~expected_sha256,
  1L, "Standardized participant inputs",
  "artifacts/12_manifests/model_input_normalization.csv",
  "e3d484711abb54f69d63ff302e5a0329efda3fd0a8c85feaf9cc4f850005c9ab",
  2L, "Gap-timing-unaware prepared data",
  "artifacts/12_manifests/manuscript_prepared_data_artifacts.csv",
  "4ed62fbe58de65a6d05d8cfc6b5d870d7c74a3c83fe89bcd72bd1dd838698935",
  3L, "Site and daylight context",
  "artifacts/12_manifests/site_solar_context_artifacts.csv",
  "c0c6d7f97782c0d880a213050f576213f5d56b3204a02732f6399e7e4aeb8518",
  4L, "Local-clock and elapsed-time links",
  "artifacts/06_model_data/temporal_provenance/artifact_manifest.csv",
  "9355f7ca4f249059cf49808a3fb1caf9e764a6160f5d61beba8234a2bfbdef7d",
  5L, "Shared model inputs",
  "artifacts/12_manifests/base_model_data_artifacts.csv",
  "8344bdc0339a53079bf9eeb7d86de1ad7c15373641c3a0a5a01d040b418895ce",
  6L, "H01 primary prepared frames",
  "artifacts/12_manifests/H01_model_data_artifacts.csv",
  "25978c5d6903e6e835c1aff6a9bc2e7552b85295e8d840d3dd65bbf9b1eb6b72",
  7L, "H01 gap-timing-unaware prepared frames",
  "artifacts/12_manifests/H01_manuscript_prepared_data_artifacts.csv",
  "e0d98178ede61b74353e3a654e7f1b53d7e1c915833a2cda74eea45ed02f383b",
  8L, "Stored descriptive comparison",
  "artifacts/12_manifests/preanalysis_comparison_artifacts.csv",
  "f3c4bfbf120d2023c045b44e3c7c04c58bce1f8d11621956bca1e5ede90c5623"
)

manifest_objects <- list()
manifest_file_checks <- list()

for (index in seq_len(nrow(manifest_registry))) {
  registry_row <- manifest_registry[index, ]
  manifest_data <- read_stored_csv(registry_row$manifest)
  assert_columns(
    manifest_data,
    c("path", "sha256", "bytes"),
    registry_row$manifest
  )

  manifest_sha256 <- digest::digest(
    file = absolute_path(registry_row$manifest),
    algo = "sha256"
  )
  stopifnot(identical(manifest_sha256, registry_row$expected_sha256))

  listed_paths <- absolute_path(manifest_data$path)
  exists <- file.exists(listed_paths)
  actual_bytes <- rep(NA_real_, length(listed_paths))
  actual_sha256 <- rep(NA_character_, length(listed_paths))
  actual_bytes[exists] <- file.info(listed_paths[exists])$size
  actual_sha256[exists] <- vapply(
    listed_paths[exists],
    digest::digest,
    character(1),
    algo = "sha256",
    file = TRUE
  )

  manifest_file_checks[[index]] <- tibble::tibble(
    stage_order = registry_row$stage_order,
    stage = registry_row$stage,
    manifest_sha256 = manifest_sha256,
    path = manifest_data$path,
    exists = exists,
    bytes_match = exists & actual_bytes == as.numeric(manifest_data$bytes),
    sha256_match = exists & actual_sha256 == manifest_data$sha256
  )
  manifest_objects[[registry_row$stage]] <- manifest_data
}

manifest_file_checks <- dplyr::bind_rows(manifest_file_checks)
stopifnot(
  all(manifest_file_checks$exists),
  all(manifest_file_checks$bytes_match),
  all(manifest_file_checks$sha256_match)
)

manifest_summary <- manifest_file_checks |>
  dplyr::group_by(.data$stage_order, .data$stage) |>
  dplyr::summarise(
    `Manifest SHA-256` = dplyr::first(.data$manifest_sha256),
    `Files listed` = dplyr::n(),
    `Files present` = sum(.data$exists),
    `Size checks` = sum(.data$bytes_match),
    `SHA-256 checks` = sum(.data$sha256_match),
    Check = ifelse(
      all(.data$exists & .data$bytes_match & .data$sha256_match),
      "PASS",
      "FAIL"
    ),
    .groups = "drop"
  ) |>
  dplyr::arrange(.data$stage_order) |>
  dplyr::select(-.data$stage_order)

normalization_manifest <-
  manifest_objects[["Standardized participant inputs"]]
gap_manifest <-
  manifest_objects[["Gap-timing-unaware prepared data"]]
site_manifest <-
  manifest_objects[["Site and daylight context"]]
base_manifest <-
  manifest_objects[["Shared model inputs"]]

current_metric_manifest_sha256 <-
  "028bce108339c1277df4a070597430ea20b49c34c2888f4eef42741c1fe76a5e"
mder_audit_metric_manifest_sha256 <-
  "7bb2830c695e5600f0877acca500a187a040fe727a029bdb069f69099abe4f43"
current_mder_audit_sha256 <-
  "5bd34fd4ac26dc2761540393d40e096cde835d8dd4b4edda9253243c9083d4bb"
current_gap_manifest_sha256 <-
  "4ed62fbe58de65a6d05d8cfc6b5d870d7c74a3c83fe89bcd72bd1dd838698935"
current_gap_participant_day_rds_sha256 <-
  "7561b5dd47cb5e23ca94ba59bd57648f11fe492f86af53a566c5c9cf42d932f1"
current_gap_support_rds_sha256 <-
  "a0bc5d7ea2142709412da2253158086416ddb730b0f60ff616bba9d6ccb1edc5"
current_gap_repair_evidence_sha256 <-
  "81e3aa9439b88cf239d6669dc4343cfb146cd2f45790bae412ab837156315018"
current_site_manifest_sha256 <-
  "c0c6d7f97782c0d880a213050f576213f5d56b3204a02732f6399e7e4aeb8518"
current_downstream_rebuild_manifest_sha256 <-
  "408087d420999322628066caae31efc288a5a5213b8b57d4b9a4ad77c9ec9a77"
mder_repin_site_manifest_sha256 <-
  "959e7a1ff1e659ad3eb673c4d4b827da0b02b8a0a779e2647459638c4dac1045"
mder_repin_base_manifest_sha256 <-
  "b6fa22836faee5243bb6ce1cc9dceab5d5403f94d688f8f74472a19dbf6e3e09"
mder_repin_preanalysis_manifest_sha256 <-
  "f90ea36334b59821101ef36de50d5d84dcfc6e8bee23f1f918be10fb2247724e"
current_base_manifest_sha256 <-
  "8344bdc0339a53079bf9eeb7d86de1ad7c15373641c3a0a5a01d040b418895ce"
current_base_input_bundle_sha256 <-
  "e840ce9d2a7f653bc5ebbfe020ce087017dfdaf0276df09da578bda30a539916"
current_base_glasses_rds_sha256 <-
  "013afe75e9b75b141b4cb48d04111a9e6c688630085a24c9fe07ca3d142a4e9a"
current_base_chest_rds_sha256 <-
  "497466ccd1a35635cca40bb8f9ce51efb97321ab04b42ee1bb76785e466a2057"
l10_zero_decision_sha256 <-
  "23b9f70d1d16f7fd3ebbdbc57aaf78c0a701fe1f9d22bd667d926cd320d40797"
l10_zero_evidence_sha256 <-
  "a37efd3449a8d1a6065d0eb8964bd6cf8b241f1683a26c065ea1a946e23214fb"
mder_decision_sha256 <-
  "1664347de976057807fcb5ac6e24bd4b66af1fe32378fabd1f2acafcfc9266de"
base_gate_decision_sha256 <-
  "63f17f1b9b3a91d437a5964cde063770d0fb4fac3f9c81e588fc374d9cc71f04"
historical_mder_gate_sha256 <-
  "9ca9b97dac5efd9ec4e7b43a1fb2227b7a13cf98ac4b92318606c4874f9e460b"

stopifnot(
  identical(
    manifest_registry$expected_sha256[
      manifest_registry$stage == "Gap-timing-unaware prepared data"
    ],
    current_gap_manifest_sha256
  ),
  identical(
    manifest_registry$expected_sha256[
      manifest_registry$stage == "Site and daylight context"
    ],
    current_site_manifest_sha256
  ),
  identical(
    manifest_registry$expected_sha256[
      manifest_registry$stage == "Shared model inputs"
    ],
    current_base_manifest_sha256
  )
)
check_file_hash(
  "artifacts/12_manifests/metric_artifacts.csv",
  current_metric_manifest_sha256
)
check_file_hash(
  "artifacts/08_diagnostics/mder_METRIC-010/audit_manifest.csv",
  current_mder_audit_sha256
)
check_file_hash(
  "artifacts/12_manifests/mder_support_gate_artifacts.csv",
  historical_mder_gate_sha256
)
check_file_hash(
  "audit/decisions/mder_mean_of_viable_ratios.md",
  mder_decision_sha256
)
check_file_hash(
  "audit/decisions/preparation06_current_base_model_gate.md",
  base_gate_decision_sha256
)
check_file_hash(
  "audit/decisions/l10_numerical_zero_normalization.md",
  l10_zero_decision_sha256
)
check_file_hash(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
  current_gap_participant_day_rds_sha256
)
check_file_hash(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/mder_support.rds",
  current_gap_support_rds_sha256
)
check_file_hash(
  "artifacts/06_model_data/base/metrics_glasses_participant_day_context.rds",
  current_base_glasses_rds_sha256
)
check_file_hash(
  "artifacts/06_model_data/base/metrics_chest_participant_day_context.rds",
  current_base_chest_rds_sha256
)

gap_repair_bundle <- check_evidence_manifest(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_repair_evidence_manifest.csv",
  current_gap_repair_evidence_sha256,
  "Gap-timing-unaware MDER repair evidence"
)
downstream_rebuild_bundle <- check_evidence_manifest(
  "artifacts/08_diagnostics/mder_METRIC-010/downstream_rebuild/rebuild_manifest.csv",
  current_downstream_rebuild_manifest_sha256,
  "Preceding MDER provenance-repin evidence"
)
l10_zero_bundle <- check_evidence_manifest(
  "audit/reconciliation/l10_METRIC-011/METRIC-011_evidence_manifest.csv",
  l10_zero_evidence_sha256,
  "Current L10 numerical-zero evidence"
)
manifest_summary <- dplyr::bind_rows(
  manifest_summary,
  gap_repair_bundle$summary,
  downstream_rebuild_bundle$summary,
  l10_zero_bundle$summary
)

assert_columns(
  base_manifest,
  c(
    "input_bundle_sha256", "metric_manifest_sha256",
    "normalization_manifest_sha256", "site_context_manifest_sha256",
    "r_version", "status"
  ),
  "current shared-model-input manifest"
)
stopifnot(
  all(base_manifest$input_bundle_sha256 == current_base_input_bundle_sha256),
  all(base_manifest$metric_manifest_sha256 == current_metric_manifest_sha256),
  all(base_manifest$site_context_manifest_sha256 == current_site_manifest_sha256),
  identical(
    base_manifest$sha256[
      base_manifest$artifact_id == "glasses_participant_day_context"
    ],
    current_base_glasses_rds_sha256
  ),
  identical(
    base_manifest$sha256[
      base_manifest$artifact_id == "chest_participant_day_context"
    ],
    current_base_chest_rds_sha256
  ),
  all(base_manifest$r_version == "4.6.1"),
  all(base_manifest$status == "PASS")
)

mder_audit_manifest <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/audit_manifest.csv"
)
assert_columns(
  mder_audit_manifest,
  c("role", "path", "sha256", "bytes", "decision_id", "r_version"),
  "METRIC-010 audit manifest"
)
mder_audit_outputs <- mder_audit_manifest |>
  dplyr::filter(.data$role == "output")
mder_audit_paths <- absolute_path(mder_audit_outputs$path)
mder_audit_present <- file.exists(mder_audit_paths)
mder_audit_bytes <- rep(NA_real_, length(mder_audit_paths))
mder_audit_sha256 <- rep(NA_character_, length(mder_audit_paths))
mder_audit_bytes[mder_audit_present] <-
  file.info(mder_audit_paths[mder_audit_present])$size
mder_audit_sha256[mder_audit_present] <- vapply(
  mder_audit_paths[mder_audit_present],
  function(path) digest::digest(file = path, algo = "sha256"),
  character(1)
)
stopifnot(
  all(mder_audit_present),
  all(mder_audit_bytes == mder_audit_outputs$bytes),
  all(mder_audit_sha256 == mder_audit_outputs$sha256)
)

mder_verification <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/verification_summary.csv"
)
mder_distribution <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/distribution_summary.csv"
)
non_mder_invariance <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/non_mder_invariance.csv"
)
gap_mder_summary <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_mder_summary.csv"
)
gap_mder_support <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_mder_support_summary.csv"
)
gap_non_mder_invariance <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_non_mder_invariance.csv"
)
gap_repair_input_output_hashes <- read_stored_csv(
  "audit/reconciliation/mder_METRIC-010_gap_repair/gap_repair_input_output_hashes.csv"
)
site_context_value_check <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/downstream_rebuild/site_context_value_check.csv"
)
downstream_manifest_comparison <- read_stored_csv(
  "artifacts/08_diagnostics/mder_METRIC-010/downstream_rebuild/manifest_comparison.csv"
)
l10_zero_reclassifications <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/all_numerical_zero_reclassifications.csv"
)
l10_positive_roundoff <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/positive_roundoff_reclassifications.csv"
)
l10_primary_changes <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/primary_scientific_cell_changes.csv"
)
l10_invariance <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/invariance_summary.csv"
)
l10_manifest_transition <- read_stored_csv(
  "audit/reconciliation/l10_METRIC-011/manifest_transition.csv"
)
assert_columns(
  mder_verification,
  c(
    "placement", "eligible_participant_days", "estimable_mder_days",
    "excluded_mder_days", "status", "minimum_viable_ratio_fraction",
    "metric_manifest_sha256", "immutable_inputs",
    "independently_recalculated", "r_version"
  ),
  "current MDER verification summary"
)
assert_columns(
  mder_distribution,
  c(
    "dataset", "position", "participant_days_available",
    "participant_days_estimable", "participant_days_missing",
    "participants_estimable", "mean", "median"
  ),
  "current MDER distribution summary"
)
assert_columns(
  non_mder_invariance,
  c(
    "placement", "shared_non_mder_columns", "columns_with_any_difference",
    "differing_cells", "exact_non_mder_invariance"
  ),
  "non-MDER invariance summary"
)
assert_columns(
  gap_mder_summary,
  c(
    "position", "participant_days", "participants", "old_estimable_days",
    "new_estimable_days", "new_mean", "new_median"
  ),
  "repaired gap-timing-unaware MDER summary"
)
assert_columns(
  gap_mder_support,
  c(
    "position", "failure_reason", "participant_days", "participants",
    "viable_minutes_minimum", "viable_minutes_median",
    "viable_minutes_maximum"
  ),
  "repaired gap-timing-unaware MDER support summary"
)
assert_columns(
  gap_non_mder_invariance,
  c("check", "value"),
  "gap-timing-unaware non-MDER invariance summary"
)
assert_columns(
  gap_repair_input_output_hashes,
  c("artifact", "path", "phase", "sha256"),
  "gap-timing-unaware repair input/output hashes"
)
assert_columns(
  site_context_value_check,
  c("check", "value"),
  "site-context value invariance summary"
)
assert_columns(
  downstream_manifest_comparison,
  c("artifact", "path", "sha256_after", "changed"),
  "preceding MDER downstream manifest comparison"
)
assert_columns(
  l10_zero_reclassifications,
  c(
    "position", "metric", "raw_backtransformed_value_lx",
    "normalized_value_lx", "numerical_zero_tolerance_lx",
    "source_all_zero", "source_valid_minutes", "source_zero_minutes",
    "source_positive_minutes", "source_missing_minutes",
    "numerical_zero_decision_id"
  ),
  "METRIC-011 numerical-zero reclassifications"
)
assert_columns(
  l10_positive_roundoff,
  c(
    "position", "metric", "raw_backtransformed_value_lx",
    "normalized_value_lx", "numerical_zero_tolerance_lx",
    "source_all_zero", "source_valid_minutes", "source_zero_minutes",
    "source_positive_minutes", "source_missing_minutes"
  ),
  "METRIC-011 positive-roundoff reclassifications"
)
assert_columns(
  l10_primary_changes,
  c("position", "metric", "old_value_lx", "new_value_lx"),
  "METRIC-011 primary scientific cell changes"
)
assert_columns(
  l10_invariance,
  c("check", "status", "detail"),
  "METRIC-011 invariance summary"
)
assert_columns(
  l10_manifest_transition,
  c("artifact", "path", "sha256_before", "sha256_after", "changed"),
  "METRIC-011 manifest transition"
)
stopifnot(
  all(mder_verification$status == "PASS"),
  all(mder_verification$minimum_viable_ratio_fraction == 0.50),
  all(
    mder_verification$metric_manifest_sha256 ==
      mder_audit_metric_manifest_sha256
  ),
  all(mder_verification$immutable_inputs),
  all(mder_verification$independently_recalculated),
  all(mder_verification$r_version == "4.6.1"),
  all(non_mder_invariance$shared_non_mder_columns == 43L),
  all(non_mder_invariance$columns_with_any_difference == 0L),
  all(non_mder_invariance$differing_cells == 0L),
  all(non_mder_invariance$exact_non_mder_invariance),
  identical(
    gap_non_mder_invariance$value[
      gap_non_mder_invariance$check == "non_mder_rows_after"
    ],
    "25620"
  ),
  identical(
    gap_non_mder_invariance$value[
      gap_non_mder_invariance$check == "non_mder_cells_exactly_unchanged"
    ],
    "TRUE"
  ),
  identical(
    site_context_value_check$value[
      site_context_value_check$check == "site_context_rows"
    ],
    618
  ),
  identical(
    site_context_value_check$value[
      site_context_value_check$check == "site_context_columns"
    ],
    47
  ),
  identical(
    site_context_value_check$value[
      site_context_value_check$check == "site_context_values_unchanged"
    ],
    1
  ),
  identical(
    gap_repair_input_output_hashes$sha256[
      gap_repair_input_output_hashes$artifact == "post_repair_manifest"
    ],
    current_gap_manifest_sha256
  ),
  identical(
    gap_repair_input_output_hashes$sha256[
      gap_repair_input_output_hashes$artifact ==
        "post_repair_participant_day_rds"
    ],
    current_gap_participant_day_rds_sha256
  ),
  identical(
    gap_repair_input_output_hashes$sha256[
      gap_repair_input_output_hashes$artifact ==
        "post_repair_mder_support_rds"
    ],
    current_gap_support_rds_sha256
  ),
  identical(
    downstream_manifest_comparison$sha256_after[
      downstream_manifest_comparison$artifact == "site_context"
    ],
    mder_repin_site_manifest_sha256
  ),
  identical(
    downstream_manifest_comparison$sha256_after[
      downstream_manifest_comparison$artifact == "base_model"
    ],
    mder_repin_base_manifest_sha256
  ),
  identical(
    downstream_manifest_comparison$sha256_after[
      downstream_manifest_comparison$artifact == "preanalysis"
    ],
    mder_repin_preanalysis_manifest_sha256
  ),
  nrow(l10_zero_reclassifications) == 12L,
  identical(
    as.integer(table(factor(
      l10_zero_reclassifications$position,
      levels = c("glasses", "chest")
    ))),
    c(6L, 6L)
  ),
  all(l10_zero_reclassifications$metric == "l10_mean_medi"),
  all(l10_zero_reclassifications$normalized_value_lx == 0),
  all(l10_zero_reclassifications$numerical_zero_decision_id == "METRIC-011"),
  nrow(l10_positive_roundoff) == 8L,
  identical(
    as.integer(table(factor(
      l10_positive_roundoff$position,
      levels = c("glasses", "chest")
    ))),
    c(3L, 5L)
  ),
  isTRUE(all.equal(
    l10_positive_roundoff$raw_backtransformed_value_lx,
    rep(4.163336342344337e-17, 8L),
    tolerance = 0
  )),
  all(l10_positive_roundoff$normalized_value_lx == 0),
  all(l10_positive_roundoff$source_all_zero),
  all(l10_positive_roundoff$source_positive_minutes == 0L),
  sum(l10_positive_roundoff$source_valid_minutes == 600L) == 7L,
  sum(
    l10_positive_roundoff$source_valid_minutes == 547L &
      l10_positive_roundoff$source_missing_minutes == 53L
  ) == 1L,
  all(
    l10_positive_roundoff$source_zero_minutes ==
      l10_positive_roundoff$source_valid_minutes
  ),
  all(
    l10_positive_roundoff$numerical_zero_tolerance_lx ==
      2.220446049250313e-14
  ),
  nrow(l10_primary_changes) == 8L,
  all(l10_primary_changes$metric == "l10_mean_medi"),
  all(l10_primary_changes$new_value_lx == 0),
  all(
    l10_invariance$status[l10_invariance$check != "hypothesis_models"] ==
      "PASS"
  ),
  identical(
    l10_invariance$status[l10_invariance$check == "hypothesis_models"],
    "NOT_RUN"
  ),
  identical(
    l10_manifest_transition$sha256_after[
      match(
        c("metric", "manuscript_prepared", "site", "base", "h01", "h01_gap", "preanalysis"),
        l10_manifest_transition$artifact
      )
    ],
    c(
      current_metric_manifest_sha256,
      current_gap_manifest_sha256,
      current_site_manifest_sha256,
      current_base_manifest_sha256,
      manifest_registry$expected_sha256[
        manifest_registry$stage == "H01 primary prepared frames"
      ],
      manifest_registry$expected_sha256[
        manifest_registry$stage == "H01 gap-timing-unaware prepared frames"
      ],
      manifest_registry$expected_sha256[
        manifest_registry$stage == "Stored descriptive comparison"
      ]
    )
  )
)

normalized_labels <- c(
  normalized_demographics = "Participant characteristics",
  normalized_chronotype = "Chronotype questionnaire",
  normalized_leba = "Light-exposure behaviour questionnaire (LEBA)",
  normalized_vlsq8 = "Visual light sensitivity questionnaire (VLSQ-8)",
  normalized_exercisediary = "Exercise diary",
  normalized_lightexposurediary = "Light-exposure diary",
  normalized_sleepdiaries = "Sleep diary"
)

normalized_summary <- normalization_manifest |>
  dplyr::filter(.data$artifact_id %in% names(normalized_labels)) |>
  dplyr::mutate(
    `Input type` = unname(normalized_labels[.data$artifact_id]),
    `Stored file` = .data$path
  ) |>
  dplyr::select(
    .data$`Input type`,
    Rows = .data$rows,
    Columns = .data$columns,
    .data$`Stored file`
  )

interval_issues <- read_stored_csv(
  "artifacts/06_model_data/normalized_inputs/audits/interval_issue_audit.csv"
)
assert_columns(
  interval_issues,
  c("modality", "site", "issue_code"),
  "interval issue audit"
)
interval_issue_summary <- interval_issues |>
  dplyr::count(.data$modality, .data$site, .data$issue_code, name = "Records") |>
  dplyr::mutate(
    `Prepared input` = dplyr::recode(
      .data$modality,
      lightexposurediary = "Light-exposure diary",
      sleepdiaries = "Sleep diary"
    ),
    `Why interval calculations cannot use the record` = dplyr::recode(
      .data$issue_code,
      missing_start_and_end = "Start and end time are missing",
      missing_wake = "Wake time is missing"
    )
  ) |>
  dplyr::select(
    .data$`Prepared input`,
    Site = .data$site,
    .data$Records,
    .data$`Why interval calculations cannot use the record`
  )

free_text_audit <- read_stored_csv(
  "artifacts/06_model_data/normalized_inputs/audits/free_text_audit.csv"
)
value_preservation_audit <- read_stored_csv(
  "artifacts/06_model_data/normalized_inputs/audits/value_preservation_audit.csv"
)
stopifnot(
  all(free_text_audit$excluded_from_analytic_rds),
  !any(free_text_audit$content_exported),
  all(value_preservation_audit$status == "PASS")
)

gap_output_labels <- c(
  participant_metrics_csv = "Participant-level metrics",
  participant_day_metrics_csv = "Participant-day metrics",
  thirty_minute_data_csv = "30-minute values",
  one_hour_data_csv = "One-hour values"
)

gap_artifact_summary <- gap_manifest |>
  dplyr::filter(.data$artifact_id %in% names(gap_output_labels)) |>
  dplyr::mutate(
    `Stored data level` = unname(gap_output_labels[.data$artifact_id]),
    `Stored file` = .data$path
  ) |>
  dplyr::select(
    .data$`Stored data level`,
    Rows = .data$rows,
    Columns = .data$columns,
    .data$`Stored file`
  )

correction_record <- read_stored_csv(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/correction_manifest.csv"
) |>
  dplyr::transmute(
    `Correction record` = .data$correction_id,
    Site = .data$affected_site,
    `Diary rows` = .data$affected_rows,
    `Recorded identifier` = .data$erroneous_value,
    `Verified identifier` = .data$corrected_value,
    `Scientific light values changed` = ifelse(
      .data$scientific_light_values_changed,
      "Yes",
      "No"
    )
  )

site_registry <- read_stored_csv("config/site_display_registry.csv")
assert_columns(
  site_registry,
  c("site", "display_order", "display_name", "color_hex"),
  "site display registry"
)

site_join_audit <- read_stored_csv(
  "artifacts/06_model_data/context/site_solar_context_join_audit.csv"
)
stopifnot(
  all(site_join_audit$status == "PASS"),
  all(site_join_audit$input_rows == site_join_audit$joined_rows),
  all(site_join_audit$unmatched_rows == 0L)
)

site_join_summary <- site_join_audit |>
  dplyr::transmute(
    Placement = dplyr::recode(
      .data$placement,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    Resolution = dplyr::recode(
      .data$resolution,
      daily = "Daily",
      `30_minute` = "30 minutes",
      one_hour = "One hour"
    ),
    `Rows before join` = .data$input_rows,
    `Rows after join` = .data$joined_rows,
    `Participant-days` = .data$participant_days,
    `Unmatched rows` = .data$unmatched_rows,
    Check = .data$status
  )

true_utc_bins <- read_stored_rds(
  "artifacts/06_model_data/temporal_provenance/true_utc_source_bins.rds"
)
wall_outcome_links <- read_stored_rds(
  "artifacts/06_model_data/temporal_provenance/wall_outcome_links.rds"
)
assert_columns(
  true_utc_bins,
  c(
    "site", "Id", "local_date", "local_day_type",
    "relationship_type", "sequence_eligible",
    "true_elapsed_sequence_start"
  ),
  "true elapsed-time source bins"
)
assert_columns(
  wall_outcome_links,
  c(
    "site", "Id", "local_date", "relationship_type",
    "spring_gap_provenance"
  ),
  "local-clock outcome links"
)

count_participant_days <- function(data, day_type) {
  data |>
    dplyr::filter(.data$local_day_type == day_type) |>
    dplyr::distinct(.data$site, .data$Id, .data$local_date) |>
    nrow()
}

temporal_summary <- tibble::tibble(
  `Stored check or outcome` = c(
    "Original UTC intervals",
    "Rows on the shared local-clock grid",
    "Participant-days spanning the autumn clock change",
    "Local-clock rows averaging two real intervals",
    "Participant-days spanning the spring clock change",
    "Structural local-clock rows in the spring gap",
    "Original intervals without a one-to-one local-clock row",
    "Intervals eligible for continuous-sequence models",
    "Starts of continuous elapsed-time sequences"
  ),
  Value = c(
    nrow(true_utc_bins),
    nrow(wall_outcome_links),
    count_participant_days(true_utc_bins, "fall_back_25h"),
    sum(
      wall_outcome_links$relationship_type ==
        "two_to_one_averaged_fall_back",
      na.rm = TRUE
    ),
    count_participant_days(true_utc_bins, "spring_forward_23h"),
    sum(
      wall_outcome_links$spring_gap_provenance ==
        "structural_nonexistent_wall_bin",
      na.rm = TRUE
    ),
    sum(
      true_utc_bins$relationship_type != "one_to_one_true_elapsed",
      na.rm = TRUE
    ),
    sum(true_utc_bins$sequence_eligible %in% TRUE),
    sum(true_utc_bins$true_elapsed_sequence_start %in% TRUE)
  )
)

near_eye_grid <- read_stored_rds(
  "artifacts/06_model_data/base/metrics_glasses_30_minute_context.rds"
)
chest_grid <- read_stored_rds(
  "artifacts/06_model_data/base/metrics_chest_30_minute_context.rds"
)
assert_columns(
  near_eye_grid,
  c("site", "Id", "local_date", "bin_admissible"),
  "near-eye 30-minute model input"
)
assert_columns(
  chest_grid,
  c("site", "Id", "local_date", "bin_admissible"),
  "chest 30-minute model input"
)

summarise_grid <- function(data, prefix) {
  output <- data |>
    dplyr::group_by(.data$site) |>
    dplyr::summarise(
      participants = dplyr::n_distinct(.data$Id),
      participant_days = dplyr::n_distinct(
        paste(.data$Id, .data$local_date)
      ),
      grid_rows = dplyr::n(),
      admissible_rows = sum(.data$bin_admissible %in% TRUE),
      .groups = "drop"
    )
  names(output)[-1] <- paste(prefix, names(output)[-1], sep = "_")
  output
}

site_samples <- site_registry |>
  dplyr::left_join(
    summarise_grid(near_eye_grid, "near_eye"),
    by = "site",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    summarise_grid(chest_grid, "chest"),
    by = "site",
    relationship = "one-to-one"
  ) |>
  dplyr::arrange(.data$display_order) |>
  dplyr::transmute(
    Site = .data$display_name,
    site_code = .data$site,
    color_hex = .data$color_hex,
    `Near-eye participants` = .data$near_eye_participants,
    `Near-eye participant-days` = .data$near_eye_participant_days,
    `Near-eye 30-min rows` = .data$near_eye_grid_rows,
    `Near-eye admissible rows` = .data$near_eye_admissible_rows,
    `Chest participants` = .data$chest_participants,
    `Chest participant-days` = .data$chest_participant_days,
    `Chest 30-min rows` = .data$chest_grid_rows,
    `Chest admissible rows` = .data$chest_admissible_rows
  )

participant_availability <- read_stored_csv(
  "artifacts/06_model_data/base/participant_availability_audit.csv"
)
modality_columns <- c(
  "Participant characteristics" = "has_demographics",
  "Chronotype questionnaire" = "has_chronotype",
  "Light-exposure behaviour questionnaire (LEBA)" = "has_leba",
  "Visual light sensitivity questionnaire (VLSQ-8)" = "has_vlsq8",
  "Exercise diary" = "has_exercisediary",
  "Light-exposure diary" = "has_lightexposurediary",
  "Sleep diary" = "has_sleepdiaries"
)
assert_columns(
  participant_availability,
  c(
    "has_glasses_metrics", "has_chest_metrics",
    unname(modality_columns)
  ),
  "participant availability audit"
)

summarise_availability <- function(metric_column, placement_label) {
  metric_participants <- participant_availability[[metric_column]] %in% TRUE
  matching_counts <- vapply(
    unname(modality_columns),
    function(column) {
      sum(participant_availability[[column]][metric_participants] %in% TRUE)
    },
    integer(1)
  )
  tibble::tibble(
    Placement = placement_label,
    `Questionnaire or diary input` = names(modality_columns),
    `Participants with light metrics` = sum(metric_participants),
    `Participants with matching input` = matching_counts,
    `Participants without matching input` =
      sum(metric_participants) - matching_counts
  )
}

availability_summary <- dplyr::bind_rows(
  summarise_availability("has_glasses_metrics", "Near-eye"),
  summarise_availability("has_chest_metrics", "Chest")
)

h01_primary_flow <- read_stored_csv(
  "artifacts/06_model_data/H01/sample_flow.csv"
)
h01_gap_flow <- read_stored_csv(
  paste0(
    "artifacts/06_model_data/H01/scenarios/",
    "manuscript_prepared_data/sample_flow.csv"
  )
)
h01_primary_contract <- read_stored_csv(
  "artifacts/06_model_data/H01/metric_contract.csv"
)
h01_gap_contract <- read_stored_csv(
  paste0(
    "artifacts/06_model_data/H01/scenarios/",
    "manuscript_prepared_data/metric_contract.csv"
  )
)

make_h01_sample <- function(sample_flow, metric_contract) {
  assert_columns(
    sample_flow,
    c(
      "placement", "scenario", "metric_order", "metric_id", "stage",
      "scope", "model_observations", "participants",
      "contributing_participant_days", "metric_support_valid_hours",
      "sites"
    ),
    "H01 sample-flow file"
  )
  assert_columns(
    metric_contract,
    c("metric_id", "manuscript_name"),
    "H01 metric contract"
  )

  output <- sample_flow |>
    dplyr::filter(
      .data$placement == "glasses",
      .data$scenario == "all_available",
      .data$stage == "metric_estimable",
      .data$scope == "overall"
    ) |>
    dplyr::left_join(
      metric_contract |>
        dplyr::select(.data$metric_id, .data$manuscript_name),
      by = "metric_id",
      relationship = "many-to-one"
    ) |>
    dplyr::arrange(.data$metric_order) |>
    dplyr::transmute(
      Metric = .data$manuscript_name,
      Participants = .data$participants,
      `Rows in prepared frame` = .data$model_observations,
      `Contributing participant-days` =
        .data$contributing_participant_days,
      `Valid measurement hours` = .data$metric_support_valid_hours,
      Sites = .data$sites
    )

  stopifnot(nrow(output) == 17L)
  output
}

h01_primary_sample <- make_h01_sample(
  h01_primary_flow,
  h01_primary_contract
)
h01_gap_sample <- make_h01_sample(
  h01_gap_flow,
  h01_gap_contract
)

h01_primary_status <- read_stored_csv(
  "artifacts/06_model_data/H01/scenario_status.csv"
)
h01_gap_status <- read_stored_csv(
  paste0(
    "artifacts/06_model_data/H01/scenarios/",
    "manuscript_prepared_data/scenario_status.csv"
  )
)
paired_unavailable <- dplyr::bind_rows(
  h01_primary_status |>
    dplyr::mutate(Dataset = "Primary dataset"),
  h01_gap_status |>
    dplyr::mutate(Dataset = "Gap-timing-unaware dataset")
) |>
  dplyr::filter(!.data$available)
stopifnot(nrow(paired_unavailable) == 8L)

figure_source_path <- paste0(
  "artifacts/08_diagnostics/preanalysis_comparison/",
  "categorical_levels.csv"
)
figure_source <- read_stored_csv(figure_source_path)
assert_columns(
  figure_source,
  c("variable_id", "placement", "series", "level", "n"),
  "stored categorical figure source data"
)

figure_site_data <- figure_source |>
  dplyr::filter(
    .data$variable_id == "participant_site",
    .data$placement == "glasses"
  ) |>
  dplyr::mutate(
    Dataset = dplyr::recode(
      .data$series,
      canonical = "Primary dataset",
      baseline = "Gap-timing-unaware dataset"
    )
  ) |>
  dplyr::inner_join(
    site_registry,
    by = c("level" = "site"),
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    Dataset = factor(
      .data$Dataset,
      levels = c("Primary dataset", "Gap-timing-unaware dataset")
    ),
    display_name = factor(
      .data$display_name,
      levels = rev(site_registry$display_name)
    )
  )
stopifnot(nrow(figure_site_data) == 18L)

primary_mder_display <- mder_distribution |>
  dplyr::filter(.data$dataset == "primary") |>
  dplyr::transmute(
    Dataset = "Primary dataset",
    Placement = dplyr::recode(
      .data$position,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    `Eligible days` = .data$participant_days_available,
    `Days with MDER` = .data$participant_days_estimable,
    `Days without MDER` = .data$participant_days_missing,
    `Participants with MDER` = .data$participants_estimable,
    `Mean MDER` = .data$mean,
    `Median MDER` = .data$median
  )

gap_participants_with_mder <- gap_mder_support |>
  dplyr::filter(is.na(.data$failure_reason)) |>
  dplyr::select(
    .data$position,
    `Participants with MDER` = .data$participants
  )

gap_mder_display <- gap_mder_summary |>
  dplyr::left_join(gap_participants_with_mder, by = "position") |>
  dplyr::transmute(
    Dataset = "Gap-timing-unaware dataset",
    Placement = dplyr::recode(
      .data$position,
      glasses = "Near-eye",
      chest = "Chest"
    ),
    `Eligible days` = .data$participant_days,
    `Days with MDER` = .data$new_estimable_days,
    `Days without MDER` = .data$participant_days - .data$new_estimable_days,
    .data$`Participants with MDER`,
    `Mean MDER` = .data$new_mean,
    `Median MDER` = .data$new_median
  )

mder_availability_display <- dplyr::bind_rows(
  primary_mder_display,
  gap_mder_display
) |>
  dplyr::arrange(
    factor(
      .data$Dataset,
      levels = c("Primary dataset", "Gap-timing-unaware dataset")
    ),
    factor(.data$Placement, levels = c("Near-eye", "Chest"))
  )

primary_mder <- mder_availability_display |>
  dplyr::filter(.data$Dataset == "Primary dataset")
stopifnot(
  identical(primary_mder$`Eligible days`, c(816, 902)),
  identical(primary_mder$`Days with MDER`, c(702, 732)),
  identical(primary_mder$`Days without MDER`, c(114, 170)),
  identical(primary_mder$`Participants with MDER`, c(137, 152)),
  isTRUE(all.equal(primary_mder$`Mean MDER`, c(0.7238529, 0.7565438),
    tolerance = 1e-7
  )),
  isTRUE(all.equal(primary_mder$`Median MDER`, c(0.7239000, 0.7494261),
    tolerance = 1e-7
  ))
)

gap_mder <- mder_availability_display |>
  dplyr::filter(.data$Dataset == "Gap-timing-unaware dataset")
stopifnot(
  identical(gap_mder$`Eligible days`, c(811, 897)),
  identical(gap_mder$`Days with MDER`, c(687, 723)),
  identical(gap_mder$`Days without MDER`, c(124, 174)),
  identical(gap_mder$`Participants with MDER`, c(137, 152)),
  isTRUE(all.equal(gap_mder$`Mean MDER`, c(0.7242573, 0.7567377),
    tolerance = 1e-7
  )),
  isTRUE(all.equal(gap_mder$`Median MDER`, c(0.7238676, 0.7495175),
    tolerance = 1e-7
  )),
  identical(
    gap_mder_summary$old_estimable_days[
      match(c("glasses", "chest"), gap_mder_summary$position)
    ],
    c(725, 729)
  )
)

l10_zero_display <- tibble::tibble(
  Placement = c("Near-eye", "Chest", "Total"),
  `New exact-zero L10 cells` = c(3L, 5L, 8L),
  `Stored source-minute evidence` = c(
    "Three 600-minute windows: 600 observed zeros in each",
    paste(
      "Four windows: 600 observed zeros; one window:",
      "547 observed zeros + 53 missing minutes"
    ),
    paste(
      "Seven windows: 600 observed zeros; one window:",
      "547 observed zeros + 53 missing minutes"
    )
  ),
  `Effect on prepared samples` = c(
    "None; all 816 primary participant-days remain",
    "None; all 902 primary participant-days remain",
    "No sample or metric-support contract changed"
  )
)

l10_provenance_display <- tibble::tibble(
  `Stored record` = c(
    "Numerical-zero decision",
    "Independent numerical-zero evidence manifest",
    "Current metric manifest",
    "Near-eye shared participant-day RDS",
    "Chest shared participant-day RDS",
    "Current site/daylight-context manifest",
    "Current shared-model-input manifest",
    "Current shared-model input bundle",
    "Current H01 primary preparation manifest",
    "Current H01 gap-timing-unaware preparation manifest",
    "Current preanalysis comparison manifest"
  ),
  `Exact path or identity` = c(
    "audit/decisions/l10_numerical_zero_normalization.md",
    "audit/reconciliation/l10_METRIC-011/METRIC-011_evidence_manifest.csv",
    "artifacts/12_manifests/metric_artifacts.csv",
    "artifacts/06_model_data/base/metrics_glasses_participant_day_context.rds",
    "artifacts/06_model_data/base/metrics_chest_participant_day_context.rds",
    "artifacts/12_manifests/site_solar_context_artifacts.csv",
    "artifacts/12_manifests/base_model_data_artifacts.csv",
    "Recorded in every row of the shared-model-input manifest",
    "artifacts/12_manifests/H01_model_data_artifacts.csv",
    "artifacts/12_manifests/H01_manuscript_prepared_data_artifacts.csv",
    "artifacts/12_manifests/preanalysis_comparison_artifacts.csv"
  ),
  `SHA-256` = c(
    l10_zero_decision_sha256,
    l10_zero_evidence_sha256,
    current_metric_manifest_sha256,
    current_base_glasses_rds_sha256,
    current_base_chest_rds_sha256,
    current_site_manifest_sha256,
    current_base_manifest_sha256,
    current_base_input_bundle_sha256,
    manifest_registry$expected_sha256[
      manifest_registry$stage == "H01 primary prepared frames"
    ],
    manifest_registry$expected_sha256[
      manifest_registry$stage == "H01 gap-timing-unaware prepared frames"
    ],
    manifest_registry$expected_sha256[
      manifest_registry$stage == "Stored descriptive comparison"
    ]
  ),
  Status = c(
    "Active", "PASS", "Current", "Current", "Current", "Current",
    "Current", "PASS", "Current", "Current", "Current"
  )
)

mder_provenance_display <- tibble::tibble(
  `Stored record` = c(
    "Author-approved MDER decision",
    "Current metric manifest",
    "Metric manifest tied to the stored MDER audit",
    "Independent primary MDER audit manifest",
    "Gap-timing-unaware preparation manifest",
    "Gap-timing-unaware participant-day RDS",
    "Gap-timing-unaware MDER-support RDS",
    "Gap-timing-unaware repair evidence manifest",
    "Site/daylight context manifest",
    "Current shared-model-input manifest",
    "Shared-model input bundle",
    "Preceding METRIC-010 provenance-repin evidence manifest",
    "Current Preparation 06 gate decision",
    "Historical ratio-of-integrals support gate"
  ),
  `Exact path or identity` = c(
    "audit/decisions/mder_mean_of_viable_ratios.md",
    "artifacts/12_manifests/metric_artifacts.csv",
    "Recorded in artifacts/08_diagnostics/mder_METRIC-010/verification_summary.csv",
    "artifacts/08_diagnostics/mder_METRIC-010/audit_manifest.csv",
    "artifacts/12_manifests/manuscript_prepared_data_artifacts.csv",
    "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
    "artifacts/06_model_data/scenarios/manuscript_prepared_data/mder_support.rds",
    "audit/reconciliation/mder_METRIC-010_gap_repair/gap_repair_evidence_manifest.csv",
    "artifacts/12_manifests/site_solar_context_artifacts.csv",
    "artifacts/12_manifests/base_model_data_artifacts.csv",
    "Recorded in every row of the shared-model-input manifest",
    "artifacts/08_diagnostics/mder_METRIC-010/downstream_rebuild/rebuild_manifest.csv",
    "audit/decisions/preparation06_current_base_model_gate.md",
    "artifacts/12_manifests/mder_support_gate_artifacts.csv"
  ),
  `SHA-256` = c(
    mder_decision_sha256,
    current_metric_manifest_sha256,
    mder_audit_metric_manifest_sha256,
    current_mder_audit_sha256,
    current_gap_manifest_sha256,
    current_gap_participant_day_rds_sha256,
    current_gap_support_rds_sha256,
    current_gap_repair_evidence_sha256,
    current_site_manifest_sha256,
    current_base_manifest_sha256,
    current_base_input_bundle_sha256,
    current_downstream_rebuild_manifest_sha256,
    base_gate_decision_sha256,
    historical_mder_gate_sha256
  ),
  Status = c(
    "Active",
    "Current",
    "Version-bound",
    "PASS",
    "Current",
    "Current",
    "Current",
    "PASS",
    "Current",
    "Current",
    "PASS",
    "Historical evidence",
    "Current",
    "Historical only"
  )
)

render_boundary <- tibble::tribble(
  ~During_render, ~What_it_means,
  "Read stored RDS and CSV files",
  "No source acquisition, normalization, metric derivation, or model-frame construction",
  "Check file size and SHA-256 identity",
  "The eight preparation manifests, current metric manifest, MDER evidence bundles, and METRIC-011 evidence must match",
  "Check required columns",
  "The page stops if a stored file no longer has the fields needed for its displays",
  "Calculate lightweight counts and draw one plot",
  "Only descriptive summaries of stored rows are calculated; no analytical artifact is written"
)

script_map <- tibble::tribble(
  ~Step, ~Producing_script, ~Function_module, ~Reads_and_operation, ~Stored_output,
  1L,
  "scripts/pipeline/build_metric_derivation.R",
  "scripts/pipeline/metric_derivation.R; scripts/pipeline/time_support.R",
  "Reads the eligible one-minute records and reconstructs the current metric layer. For MDER, it uses the complete 1,440-position local-clock grid, averages fall-back duplicates channel-wise, averages viable positive finite one-minute melEDI/illuminance ratios, and requires at least 720 ratios. For offset geometric means, it preserves the raw back-transform and applies the source-verified METRIC-011 numerical-zero rule before storing a non-negative value. This upstream step is shown because its current manifest is an exact input to the shared model data.",
  "artifacts/05_metrics/ and artifacts/12_manifests/metric_artifacts.csv",
  2L,
  "scripts/pipeline/verify_metric_derivation_mder.R",
  "Independent METRIC-010 implementation",
  "Reads stored one-minute inputs and the metric version identified in its verification summary; independently checks the local-clock grid, viable-pair count, MDER value, and failure reason. The stored METRIC-010 audit is tied to preceding metric manifest 7bb2830c…f43; METRIC-011 separately verified that MDER did not change in the current metric layer.",
  "artifacts/08_diagnostics/mder_METRIC-010/verification_summary.csv",
  3L,
  "audit/scripts/finalize_mder_METRIC_010.R",
  "Stored amendment-audit finalizer",
  "Reads the METRIC-010-verified metric layer and preserved historical snapshot; records distributions, availability, comparisons, and exact non-MDER invariance. This version-bound bundle produces documentation evidence, not a new model result.",
  "artifacts/08_diagnostics/mder_METRIC-010/ and its audit_manifest.csv",
  4L,
  "scripts/pipeline/build_model_input_normalization.R",
  "scripts/pipeline/model_input_normalization.R",
  "Reads the fixed local files acquired in Preparation 05; aligns field types, labels, identifiers, and time representations while preserving recorded values. Kept separate so source-value and free-text checks are explicit.",
  "artifacts/06_model_data/normalized_inputs/",
  5L,
  "scripts/pipeline/build_manuscript_prepared_data.R",
  "scripts/pipeline/manuscript_prepared_data.R",
  "Reads the retained prepared measurements and checked participant inputs; assembles the predefined comparison dataset, applies the active MDER estimand to its pinned one-minute inputs, and records its verified identifier correction. Kept separate so its preparation remains distinguishable from the primary dataset.",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/",
  6L,
  "audit/scripts/repair_gap_mder_METRIC_010.R; audit/scripts/finalize_gap_mder_METRIC_010_evidence.R",
  "Independent gap-timing-unaware MDER reconstruction and evidence finalizer",
  "Reads the pinned one-minute gap-timing-unaware inputs and reconstructed daily outputs; independently records viable-minute support, corrected MDER availability and values, exact preservation of 25,620 non-MDER cells, and all input/output hashes. These scripts produced stored repair evidence and are not run by this report.",
  "audit/reconciliation/mder_METRIC-010_gap_repair/ and its gap_repair_evidence_manifest.csv",
  7L,
  "scripts/pipeline/build_site_solar_context.R",
  "scripts/pipeline/site_solar_context.R",
  "Reads site coordinates, time zones, and all dates needed by both datasets; creates civil dawn, civil dusk, photoperiod, and descriptive solar-noon fields. Kept separate so every site/date join can be checked once.",
  "artifacts/06_model_data/context/",
  8L,
  "scripts/pipeline/build_temporal_sequence_provenance.R",
  "scripts/pipeline/temporal_sequence_provenance.R",
  "Reads accepted coverage and metric time series; links local-clock outcomes to their true UTC intervals and marks continuous elapsed-time sequences. Kept separate because time-of-day and elapsed-time questions require different coordinates.",
  "artifacts/06_model_data/temporal_provenance/",
  9L,
  "scripts/pipeline/build_base_model_data.R",
  "scripts/pipeline/base_model_data.R",
  "Reads the current metric manifest, standardized participant inputs, and site context; attaches one matching context record while retaining metric rows and recording participant-input availability. The base manifest pins the current METRIC-011 metric manifest and input bundle. Kept separate from hypothesis-specific exclusions.",
  "artifacts/06_model_data/base/",
  10L,
  "scripts/pipeline/build_h01_model_data.R",
  "scripts/pipeline/h01_model_data.R",
  "Reads the shared model inputs and approved H01 metric contract; stores outcomes, predictors, metric-specific inclusion records, and sample flow for 17 metrics. No model is fitted.",
  "artifacts/06_model_data/H01.rds and artifacts/06_model_data/H01/",
  11L,
  "scripts/pipeline/build_h01_manuscript_prepared_data.R",
  "scripts/pipeline/h01_manuscript_prepared_adapter.R",
  "Reads the gap-timing-unaware data and applies the same H01 implementation contract. Kept separate so shared model code can be checked without erasing the preparation difference.",
  "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/",
  12L,
  "scripts/pipeline/build_preanalysis_comparison.R",
  "scripts/pipeline/preanalysis_comparison.R",
  "Reads both stored preparations; writes descriptive count, distribution, reconciliation, and figure-source records before hypothesis fitting. It does not test a scientific hypothesis.",
  "artifacts/08_diagnostics/preanalysis_comparison/",
  13L,
  "audit/scripts/rebuild_mder_METRIC_010_downstream.R",
  "Preceding downstream repin and invariance record",
  "Records the METRIC-010 manifest identities after the repaired gap-timing-unaware input moved through site context, shared inputs, H01 preparations, and the descriptive comparison. It verifies that all 618 by 47 site-context values were exact. METRIC-011 supersedes those identities but not this version-specific invariance record.",
  "artifacts/08_diagnostics/mder_METRIC-010/downstream_rebuild/ and its rebuild_manifest.csv",
  14L,
  "audit/scripts/finalize_l10_METRIC_011.R",
  "scripts/pipeline/verify_metric_derivation_core.R; scripts/pipeline/verify_metric_derivation_mder.R",
  "Reads the rebuilt primary metric layer, its preserved pre-repair snapshot, and the exact L10 source windows. It verifies the numerical-zero decisions and current metric layer, proves that only eight primary L10 means changed, confirms that the gap-timing-unaware scientific values did not change, and reseals dependent preparation manifests. No hypothesis model is fitted or refitted.",
  "audit/reconciliation/l10_METRIC-011/ and its METRIC-011_evidence_manifest.csv"
)

handoff_artifacts <- tibble::tribble(
  ~Stored_artifact, ~Exact_path, ~Next_use,
  "Current metric manifest",
  "artifacts/12_manifests/metric_artifacts.csv",
  "Pins the refreshed metric layer, including active MDER and the numerical-zero rule",
  "Independent MDER audit",
  "artifacts/08_diagnostics/mder_METRIC-010/",
  "Version-bound MDER verification; METRIC-011 separately confirms MDER is unchanged",
  "L10 numerical-zero verification evidence",
  "audit/reconciliation/l10_METRIC-011/",
  "Eight changed L10 cells, exact source-minute evidence, invariance, and current manifest identities",
  "Standardized participant inputs",
  "artifacts/06_model_data/normalized_inputs/",
  "Joined to checked metrics in the shared model inputs",
  "Gap-timing-unaware prepared measurements",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/",
  "Predefined prepared-data comparison in hypothesis analyses",
  "Gap-timing-unaware MDER repair evidence",
  "audit/reconciliation/mder_METRIC-010_gap_repair/",
  "Corrected MDER sample flow, support reasons, non-MDER invariance, and hashes",
  "Site and daylight context",
  "artifacts/06_model_data/context/site_solar_context.rds",
  "Site/date predictors and verified many-to-one joins",
  "Elapsed-time provenance",
  "artifacts/06_model_data/temporal_provenance/true_utc_source_bins.rds",
  "Continuity and autocorrelation coordinates for H02 and H11",
  "Shared near-eye and chest inputs",
  "artifacts/06_model_data/base/",
  "Starting point for each approved hypothesis-specific build",
  "H01 primary prepared frames",
  "artifacts/06_model_data/H01.rds",
  "H01 analysis preparation and fitting",
  "H01 gap-timing-unaware prepared frames",
  "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/H01.rds",
  "H01 prepared-data comparison",
  "Stored descriptive comparison data",
  "artifacts/08_diagnostics/preanalysis_comparison/",
  "Reader-facing preparation comparisons and provenance checks",
  "Preceding MDER provenance-repin evidence",
  "artifacts/08_diagnostics/mder_METRIC-010/downstream_rebuild/",
  "METRIC-010 identities and exact site-context invariance for that version"
)

environment_summary <- tibble::tibble(
  Component = c(
    "R",
    "Quarto profile",
    "Table package",
    "Manifest hashing",
    "Scientific recomputation during render"
  ),
  Value = c(
    paste(R.version$major, R.version$minor, sep = "."),
    "nathealth (HTML review render)",
    paste0("gt ", as.character(utils::packageVersion("gt"))),
    "SHA-256 with digest",
    "None"
  )
)
```

## Purpose and position in the analysis chain

Preparation 06 turns the checked light-exposure metrics from Preparation 04
and the fixed participant, questionnaire, and diary files acquired in
Preparation 05 into inputs that hypothesis analyses can use. It first makes
participant files structurally consistent, then adds participant and site
context to the metric files, preserves the distinction between local clock
time and actual elapsed time, and finally stores a separate input when a
hypothesis method has been approved. H01 is the first such build. This step
does not fit a statistical model.

The predefined **gap-timing-unaware dataset** still passed the general
50%-per-hour and 80%-per-day coverage rules. Its name means that the timing of
remaining missing observations was not used for an additional metric-specific
adjustment. For contrast, the primary dataset could be interpreted as a
“time-sensitive primary metric dataset.” From this point onward, this report
calls it simply the **primary dataset**.

The exact stored inputs are the current light-metric files, the fixed
questionnaire and diary files, site and daylight context, elapsed-time
records, and the separately prepared gap-timing-unaware dataset. Their file
identities, producing scripts, and verification records are collected in the
technical provenance section after the scientific sample consequences.

```{mermaid}
%%| label: fig-preparation06-chain
%%| fig-cap: "Position of Preparation 06 in the data-to-analysis chain. Solid arrows show stored data passed forward; rendering reads these outputs but does not recreate them."
%%| fig-alt: "Flow diagram from fixed participant files and checked light-exposure metrics, through standardized participant inputs, site and elapsed-time context, shared model inputs, and hypothesis-specific prepared datasets, to hypothesis analyses. A separate branch carries the gap-timing-unaware dataset through the same hypothesis implementation."
flowchart TB
  A["Fixed participant, questionnaire, and diary files<br/>Preparation 05"] --> C["Standardize structures<br/>and audit source-value preservation"]
  V["Independent MDER<br/>verification"] -. verifies .-> B["Checked light-exposure metrics<br/>Preparation 04"]
  N["Independent numerical-zero<br/>verification"] -. verifies .-> B
  B --> D["Add site, daylight, and<br/>elapsed-time context"]
  C --> E["Shared model inputs"]
  D --> E
  E --> F["Primary hypothesis-specific<br/>prepared dataset"]
  G["Gap-timing-unaware<br/>prepared dataset"] --> H["Same hypothesis<br/>implementation"]
  F --> I["Hypothesis analyses"]
  H --> I
```

## 1. Standardize participant, questionnaire, and diary structures

The first operation reads only the fixed local files acquired in Preparation
05 and places site identifiers, participant identifiers, labels, column types,
and time representations into a consistent structure. Elapsed time is
represented in Coordinated Universal Time (UTC); time-of-day fields retain the
site's local clock time. This structural work is kept separate from later joins
so that recorded values, missing values, and excluded free text can be checked
before metrics and participant information meet.

The procedure did not recalculate questionnaire totals or factor scores. The
963 field-level value-preservation records all passed, and all 63 audited
free-text fields were excluded from analytical RDS files without exporting
their content.

```{r}
#| label: tbl-normalized-participant-inputs
#| tbl-cap: "Standardized participant, questionnaire, and diary inputs stored before any model-specific selection."

normalized_summary |>
  prep_gt(
    source_note = "Read from the stored standardization inventory."
  ) |>
  gt::fmt_integer(columns = c(Rows, Columns)) |>
  gt::cols_width(
    `Input type` ~ gt::pct(28),
    Rows ~ gt::pct(9),
    Columns ~ gt::pct(9),
    `Stored file` ~ gt::pct(54)
  )
```

Records without a complete interval remain in their prepared input file, but
they cannot contribute to a calculation that requires both interval
endpoints. The stored audit identifies 27 such light-exposure-diary records
(one at Borås (SE) and 26 at Delft (NL)) and one Tübingen (DE) sleep-diary
record without a wake time.

```{r}
#| label: tbl-incomplete-diary-intervals
#| tbl-cap: "Stored diary records that remain available descriptively but cannot support interval-based calculations."

interval_issue_summary |>
  dplyr::mutate(
    Site = site_registry$display_name[
      match(.data$Site, site_registry$site)
    ]
  ) |>
  prep_gt(
    source_note = paste0(
      "Read from `artifacts/06_model_data/normalized_inputs/audits/",
      "interval_issue_audit.csv`; no interval is reconstructed during render."
    )
  ) |>
  gt::fmt_integer(columns = Records) |>
  gt::cols_width(
    `Prepared input` ~ gt::pct(22),
    Site ~ gt::pct(12),
    Records ~ gt::pct(10),
    `Why interval calculations cannot use the record` ~ gt::pct(56)
  )
```

## 2. Keep the prepared-data comparison separate

The gap-timing-unaware dataset is assembled independently of the primary
dataset. That separation prevents a preparation difference from being erased
when both are later passed through the same hypothesis implementation.

The stored data cover participant-level, participant-day, 30-minute, and
one-hour values. Exact metric-support hours cannot be recovered from these
retained prepared artifacts; later tables therefore show those hours as
unavailable rather than zero or a reconstructed estimate.

```{r}
#| label: tbl-gap-timing-unaware-artifacts
#| tbl-cap: "Accepted data levels stored for the gap-timing-unaware dataset."

gap_artifact_summary |>
  prep_gt(
    source_note = paste(
      "Read from the stored gap-timing-unaware artifact inventory;",
      "historical internal filenames are retained only in technical provenance."
    )
  ) |>
  gt::fmt_integer(columns = c(Rows, Columns)) |>
  gt::cols_width(
    `Stored data level` ~ gt::pct(22),
    Rows ~ gt::pct(10),
    Columns ~ gt::pct(10),
    `Stored file` ~ gt::pct(58)
  )
```

One objectively verified participant-identifier error is documented rather
than silently changed: record [DEV-056](../preregistration_deviations.qmd#dev-056) assigns seven TUM exercise-diary rows
to `TUM_S001` instead of the recorded `TUM_S101`. The correction did not
change a scientific light value.

```{r}
#| label: tbl-gap-timing-unaware-correction
#| tbl-cap: "Verified identifier correction recorded for the gap-timing-unaware dataset."

correction_record |>
  dplyr::mutate(
    Site = site_registry$display_name[
      match(.data$Site, site_registry$site)
    ]
  ) |>
  prep_gt(
    source_note = paste0(
      "Exact record: `artifacts/06_model_data/scenarios/",
      "manuscript_prepared_data/correction_manifest.csv`."
    )
  ) |>
  gt::fmt_integer(columns = `Diary rows`)
```

## 3. Add site and daylight context

The site-context operation reads the approved site coordinates and time zones
and the local dates required by both prepared datasets. It stores civil dawn,
civil dusk, photoperiod, and descriptive solar-noon information for 618
site-date combinations across nine sites, including four site-dates with a
daylight-saving transition.

The gap-timing-unaware input supplies supplemental site dates. The stored
comparison found all 618 rows and all 47 context fields exactly unchanged
across the documented input transition.

Every merge uses site and local date as a many-to-one key: many measurement
rows may use one site-date context record. All six joins retained their input
row counts, and none lacked a matching site-date record.

```{r}
#| label: tbl-site-daylight-joins
#| tbl-cap: "Row-preservation checks when site and daylight context was added to the primary metric files."

site_join_summary |>
  prep_gt(
    source_note = paste0(
      "Read from `artifacts/06_model_data/context/",
      "site_solar_context_join_audit.csv`. Relationship: many measurement ",
      "rows to one site-date record."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Rows before join`, `Rows after join`,
      `Participant-days`, `Unmatched rows`
    )
  ) |>
  gt::tab_spanner(
    label = "Join size",
    columns = c(`Rows before join`, `Rows after join`)
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(12),
    Resolution ~ gt::pct(13),
    `Rows before join` ~ gt::pct(15),
    `Rows after join` ~ gt::pct(15),
    `Participant-days` ~ gt::pct(17),
    `Unmatched rows` ~ gt::pct(15),
    Check ~ gt::pct(13)
  )
```

## 4. Preserve time of day and actual elapsed time

Local clock time and elapsed time answer different questions. A 30-minute row
labelled 08:00 represents 08:00 at every study site, which is useful for
time-of-day comparisons. UTC records how much real time separated two rows,
which is necessary for continuity and autocorrelation. A separate
temporal-sequence operation preserves both coordinates without changing melEDI
values.

When clocks move back, a local-clock interval may represent two real
intervals; its stored local-clock outcome is their average. When clocks move
forward, a local-clock interval can have no corresponding real interval.
Only an unambiguous, usable real interval can continue an elapsed-time
sequence. A real gap, unusable outcome, repeated autumn interval, or
structural spring gap starts or interrupts a sequence.

```{r}
#| label: tbl-temporal-sequence-provenance
#| tbl-cap: "Stored links between local-clock outcomes and actual elapsed-time intervals."

temporal_summary |>
  prep_gt(
    source_note = paste0(
      "Read from the RDS files in `artifacts/06_model_data/",
      "temporal_provenance/`; the render does not rebuild the links."
    )
  ) |>
  gt::fmt_integer(columns = Value) |>
  gt::cols_width(
    `Stored check or outcome` ~ gt::pct(75),
    Value ~ gt::pct(25)
  )
```

## 5. Combine checked data without choosing a hypothesis sample

The combination step reads the checked metric files, standardized participant
inputs, and site context. It adds one matching participant record and one
matching site/date record to each measurement row. It does not apply a
complete-case filter or a hypothesis-specific exclusion. Instead, it records
which questionnaire and diary inputs are available, allowing each hypothesis
to state its own requirements later.

The shared primary near-eye input contains
`r dplyr::n_distinct(near_eye_grid$Id)` participants,
`r dplyr::n_distinct(paste(near_eye_grid$Id, near_eye_grid$local_date))`
participant-days, and `r format(nrow(near_eye_grid), big.mark = ",")`
30-minute grid rows. The complementary chest input contains
`r dplyr::n_distinct(chest_grid$Id)` participants,
`r dplyr::n_distinct(paste(chest_grid$Id, chest_grid$local_date))`
participant-days, and `r format(nrow(chest_grid), big.mark = ",")` rows. Chest
measurements remain complementary; they are not pooled with the primary
near-eye measurements.

```{r}
#| label: tbl-shared-site-samples
#| tbl-cap: "Participants, participant-days, and 30-minute grid rows in the shared primary inputs, by submitted-manuscript site."

site_sample_gt <- site_samples |>
  dplyr::select(-.data$site_code, -.data$color_hex) |>
  prep_gt(
    source_note = paste0(
      "Calculated from the stored near-eye and chest RDS files in ",
      "`artifacts/06_model_data/base/`. Admissible rows passed the stored ",
      "30-minute support decision; no admissibility rule is rerun here."
    )
  ) |>
  gt::fmt_integer(columns = -Site) |>
  gt::tab_spanner(
    label = "Primary near-eye input",
    columns = gt::starts_with("Near-eye")
  ) |>
  gt::tab_spanner(
    label = "Complementary chest input",
    columns = gt::starts_with("Chest")
  ) |>
  gt::cols_width(
    Site ~ gt::pct(18),
    gt::starts_with("Near-eye") ~ gt::pct(10.25),
    gt::starts_with("Chest") ~ gt::pct(10.25)
  )

for (row_index in seq_len(nrow(site_samples))) {
  site_sample_gt <- site_sample_gt |>
    gt::tab_style(
      style = gt::cell_borders(
        sides = "left",
        color = site_samples$color_hex[row_index],
        weight = gt::px(5)
      ),
      locations = gt::cells_body(rows = row_index)
    )
}

site_sample_gt
```

All participants with near-eye or chest metrics have matching participant
characteristics, chronotype, LEBA, VLSQ-8, and sleep-diary inputs. Exercise
diary data are absent for three participants at each placement, and
light-exposure diary data are absent for one participant at each placement.
These are availability flags, not exclusions at this point.

```{r}
#| label: tbl-participant-input-availability
#| tbl-cap: "Availability of questionnaire and diary inputs among participants with stored light-exposure metrics."

availability_summary |>
  prep_gt(
    groupname_col = "Placement",
    source_note = paste0(
      "Read from `artifacts/06_model_data/base/",
      "participant_availability_audit.csv`; filtering_applied is FALSE for ",
      "every participant record."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Participants with light metrics`,
      `Participants with matching input`,
      `Participants without matching input`
    )
  ) |>
  gt::cols_width(
    `Questionnaire or diary input` ~ gt::pct(43),
    `Participants with light metrics` ~ gt::pct(19),
    `Participants with matching input` ~ gt::pct(19),
    `Participants without matching input` ~ gt::pct(19)
  )
```

### Current MDER in the shared inputs

The shared inputs use the current metric version. In these files, daily MDER
is the arithmetic mean of viable one-minute melEDI/illuminance ratios on a complete
1,440-position local wall-clock grid. Both channels must be finite and
strictly greater than zero; a zero-denominator ratio is undefined, and a zero
in either channel does not define the positive proportional comparison used by
this estimand. When the autumn clock change produces duplicate local minutes,
the two channels are averaged separately within that clock minute before the
ratio is formed. An absent spring-forward minute remains missing.

At least 720 viable ratios are required, and exactly 720 passes the inclusive
50% rule. Failing that rule makes MDER alone unavailable: the participant-day,
its other metrics, and its attached participant and site context remain in the
shared inputs. Current MDER uses no time-profile weighting, exclusion, or
scaling and is not a ratio of daily integrals. The older ratio-of-integrals and
profile-based exclusion records document an earlier method and are not active.

```{r}
#| label: tbl-current-mder-availability
#| tbl-cap: "Current MDER availability and distribution summaries in the two stored prepared datasets."

mder_availability_display |>
  prep_gt(
    groupname_col = "Dataset",
    source_note = paste(
      "Primary and gap-timing-unaware rows are read from their stored",
      "independent-verification summaries; rendering does not reconstruct MDER."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Eligible days`, `Days with MDER`, `Days without MDER`,
      `Participants with MDER`
    )
  ) |>
  gt::fmt_number(
    columns = c(`Mean MDER`, `Median MDER`),
    decimals = 4
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(15),
    `Eligible days` ~ gt::pct(12),
    `Days with MDER` ~ gt::pct(14),
    `Days without MDER` ~ gt::pct(15),
    `Participants with MDER` ~ gt::pct(18),
    `Mean MDER` ~ gt::pct(13),
    `Median MDER` ~ gt::pct(13)
  )
```

In the primary dataset, MDER is available for 702 of 816 near-eye
participant-days from 137 participants (mean 0.7239; median 0.7239) and 732 of
902 chest participant-days from 152 participants (mean 0.7565; median
0.7494). In the gap-timing-unaware dataset, the same momentary-ratio estimand
is available for 687 of 811 near-eye participant-days from 137 participants
(mean 0.7243; median 0.7239) and 723 of 897 chest participant-days from 152
participants (mean 0.7567; median 0.7495).

The stored comparison found all 25,620 non-MDER participant-day cells exactly
unchanged and retained all 618 × 47 site/daylight context values exactly.
Thus no non-MDER or site-context scientific value changed.

### Source-verified L10 zeros in the shared inputs

L10 is the mean melEDI during the darkest 10-hour window of a participant-day.
Its geometric mean can represent true zeros because production adds 0.1 lx,
averages `log10(melEDI + 0.1)`, transforms back, and subtracts 0.1 lx. In exact
arithmetic, an all-zero source window therefore returns zero. Floating-point
arithmetic left eight primary L10 means at
`4.163336342344337e-17` lx instead. This remainder is far below the
source-verified tolerance of `2.220446049250313e-14` lx.

The active numerical-zero rule stores a positive remainder within that
tolerance as exact zero only when every finite source minute is exactly zero.
Missing minutes remain missing; they are counted separately and are never
imputed as zeros. A small positive remainder is retained if any finite source
minute is positive, and a materially negative result stops production.

```{r}
#| label: tbl-l10-zero-handoff
#| tbl-cap: "Primary L10 values passed into the shared inputs as source-verified exact zeros."

l10_zero_display |>
  dplyr::transmute(
    .data$Placement,
    .data$`New exact-zero L10 cells`,
    `Stored evidence and effect on the prepared sample` = paste0(
      .data$`Stored source-minute evidence`,
      "<br>",
      .data$`Effect on prepared samples`
    )
  ) |>
  prep_gt(
    source_note = paste(
      "Counts and source-minute evidence are read from the stored independent",
      "numerical-zero record; rendering does not recalculate an L10 window."
    )
  ) |>
  gt::fmt_integer(columns = `New exact-zero L10 cells`) |>
  gt::fmt_markdown(
    columns = `Stored evidence and effect on the prepared sample`
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(15),
    `New exact-zero L10 cells` ~ gt::pct(18),
    `Stored evidence and effect on the prepared sample` ~ gt::pct(67)
  )
```

Exactly eight primary L10 mean cells changed: three near-eye and five chest.
Seven 600-minute windows contained 600 observed zero minutes. The remaining
chest window contained 547 observed zero minutes and 53 missing minutes; all
547 observed values were zero, while the 53 missing minutes stayed missing.
All 816 near-eye and 902 chest participant-days remain in the shared inputs.

Every non-L10 scientific value, every sample definition, all 30-minute and
hourly values, and all gap-timing-unaware scientific values are unchanged.
The shared primary inputs inherit the eight exact-zero L10 values without a
sample change. No hypothesis model was fitted or refitted.

For an analysis that permits zero outcomes, these participant-days remain as
zeros. In a two-part analysis, they remain in the zero/occurrence component
and are excluded only from the strictly positive magnitude component by that
component's definition.

## 6. Store the first hypothesis-specific inputs

Once the H01 method was approved, the shared inputs and H01 metric contract
were used to store outcomes, site and daylight predictors, participant
identifiers, metric-specific inclusion records, and sample flow. The same H01
implementation contract was applied separately to the gap-timing-unaware
dataset. Neither operation fits a model.

H01 covers 17 metric contracts. The following tables report the all-available
near-eye prepared frames. “Rows in prepared frame” is the potential input at
this boundary, not a claim about the final fitted-model sample. Any later
transformation or model-specific missing-value removal is rechecked and
stored by the H01 analysis. Valid measurement hours describe the support from
which a metric was derived; they are not additional model observations.

```{r}
#| label: tbl-h01-primary-prepared-samples
#| tbl-cap: "Available records for the 17 primary near-eye H01 prepared frames."

h01_primary_sample |>
  prep_gt(
    source_note = paste0(
      "Read from `artifacts/06_model_data/H01/sample_flow.csv` and ",
      "`artifacts/06_model_data/H01/metric_contract.csv`."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      Participants, `Rows in prepared frame`,
      `Contributing participant-days`, Sites
    )
  ) |>
  gt::fmt_number(columns = `Valid measurement hours`, decimals = 1) |>
  gt::cols_width(
    Metric ~ gt::pct(31),
    Participants ~ gt::pct(11),
    `Rows in prepared frame` ~ gt::pct(14),
    `Contributing participant-days` ~ gt::pct(17),
    `Valid measurement hours` ~ gt::pct(17),
    Sites ~ gt::pct(10)
  )
```

For the brightest and darkest 10-hour means and their midpoints, the primary
metric files do not export the exact contributing minutes; the corresponding
hour cells remain unavailable. The gap-timing-unaware artifacts do not retain
exact metric-support hours for any H01 metric, so the full column is shown as
unavailable.

```{r}
#| label: tbl-h01-gap-prepared-samples
#| tbl-cap: "Available records for the 17 gap-timing-unaware near-eye H01 prepared frames."

h01_gap_sample |>
  prep_gt(
    source_note = paste0(
      "Read from the sample-flow and metric-contract files under ",
      "`artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/`. ",
      "Support hours are unavailable in the retained prepared artifacts."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      Participants, `Rows in prepared frame`,
      `Contributing participant-days`, Sites
    )
  ) |>
  gt::fmt_number(columns = `Valid measurement hours`, decimals = 1) |>
  gt::cols_width(
    Metric ~ gt::pct(31),
    Participants ~ gt::pct(11),
    `Rows in prepared frame` ~ gt::pct(14),
    `Contributing participant-days` ~ gt::pct(17),
    `Valid measurement hours` ~ gt::pct(17),
    Sites ~ gt::pct(10)
  )
```

The H01 files also prepare comparisons using the same participants and
participant-days at both sensor positions where possible. Interdaily stability
and intradaily variability each produce one participant-level value from a
series of days. Their paired versions would require recomputation from the
same minute-level day set at both placements. They are therefore explicitly
marked unavailable for both placements in both datasets—eight scenario records
in total—instead of being approximated from unequal day sets.

## 7. Describe composition before hypothesis fitting

A stored pre-analysis comparison reads both prepared datasets and records
counts, distributions, key reconciliation, and paired descriptive comparisons.
These outputs help readers see preparation differences before a hypothesis is
tested; they are not hypothesis tests themselves.

@fig-site-composition shows the near-eye participant composition used by this
stored comparison. It is drawn during render from the paired source-data CSV,
using the submitted-manuscript site names, order, and colours. The two panels
use the same horizontal scale.

```{r}
#| label: fig-site-composition
#| fig-cap: "Near-eye participants by study site in the primary and gap-timing-unaware datasets. Counts are descriptive inputs to the stored pre-analysis comparison, not fitted-model sample sizes."
#| fig-alt: "Two stacked horizontal bar panels compare near-eye participant counts by study site. Each site has its submitted-manuscript colour and a count printed at the end of the bar. The primary and gap-timing-unaware datasets use the same horizontal scale."
#| fig-width: 7.5
#| fig-height: 6.8
#| fig-dpi: 144
#| out-width: "100%"

ggplot(
  figure_site_data,
  aes(x = .data$n, y = .data$display_name, fill = .data$level)
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = .data$n),
    hjust = -0.22,
    size = 3.2,
    colour = "#222222"
  ) +
  facet_wrap(vars(.data$Dataset), ncol = 1, scales = "fixed") +
  scale_fill_manual(
    values = stats::setNames(site_registry$color_hex, site_registry$site),
    guide = "none"
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 5),
    expand = expansion(mult = c(0, 0.16))
  ) +
  labs(
    x = "Participants with near-eye metrics",
    y = NULL
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text = element_text(size = 9, colour = "#222222"),
    axis.title = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 10, face = "bold", hjust = 0),
    strip.background = element_rect(fill = "#F2F2F2", colour = NA),
    plot.margin = margin(8, 16, 8, 8)
  )
```

[Download the source data for @fig-site-composition](../../artifacts/08_diagnostics/preanalysis_comparison/categorical_levels.csv).

The figure uses a 7.5-inch by 6.8-inch canvas and 9–10 pt axis, tick, facet,
and count labels before HTML scaling. Its final-display check covers
typography, clipping, label wrapping, panel balance, bar distinguishability,
and colour use.

## Limitations carried into the model-ready inputs

Independent reconstruction of the current reference-profile input remains incomplete.

Independent reconstruction of current diary-period support remains incomplete.

These qualifications carry forward because the model-ready files use the
current reference profiles and diary-period support classifications. They do
not indicate that the current data or a downstream result is wrong. Readers
can inspect the state, daylight, and placement information in representative
one-minute profiles in [Preparation 07](07_example_days.qmd).

## 8. Technical provenance and stored-artifact handoff

::: {.callout-note appearance="simple" title="What is calculated during this render?"}
This report reads already stored RDS and CSV outputs. During rendering it
checks that every manifest-listed file has the recorded size and SHA-256
digital fingerprint, checks the stored MDER and numerical-zero evidence,
checks the columns needed for the displays, calculates only lightweight
descriptive counts from stored rows, and redraws one reader-facing plot from
its stored source-data CSV. It does not source or run any producing script,
retrieve data, standardize inputs, derive metrics, construct or overwrite a
model frame, fit a model, estimate autocorrelation, or write an analytical
artifact.
:::

### Render boundary

```{r}
#| label: tbl-render-boundary
#| tbl-cap: "Bounded operations permitted while this report renders."

render_boundary |>
  dplyr::rename(
    `During render` = .data$During_render,
    `Interpretation` = .data$What_it_means
  ) |>
  prep_gt(
    source_note = paste0(
      "This table describes the page's execution boundary; producing ",
      "scripts are not sourced or called."
    )
  ) |>
  gt::cols_width(
    `During render` ~ gt::pct(27),
    Interpretation ~ gt::pct(73)
  )
```

### File-identity checks

The page checked every file listed in the eight preparation manifests, the two
MDER repair/rebuild evidence manifests, and the METRIC-011 evidence manifest
before displaying any result. A PASS means that the file existed and both its
byte size and SHA-256 digital fingerprint matched the manifest; it does not
mean that the render reran the producing calculation.

```{r}
#| label: tbl-manifest-identity
#| tbl-cap: "File-identity checks read from eight preparation manifests, two stored MDER evidence manifests, and the METRIC-011 evidence manifest."

manifest_summary |>
  dplyr::mutate(`Manifest SHA-256` = wrap_hash(.data$`Manifest SHA-256`)) |>
  prep_gt(
    source_note = paste0(
      "Each manifest file first matched its expected SHA-256; every listed ",
      "file was then checked. No manifest or analytical artifact was written."
    )
  ) |>
  gt::cols_label(stage = "Preparation stage") |>
  gt::fmt_integer(
    columns = c(
      `Files listed`, `Files present`, `Size checks`, `SHA-256 checks`
    )
  ) |>
  gt::tab_spanner(
    label = "Files matching the manifest",
    columns = c(`Files present`, `Size checks`, `SHA-256 checks`)
  ) |>
  gt::cols_width(
    stage ~ gt::pct(22),
    `Manifest SHA-256` ~ gt::pct(29),
    `Files listed` ~ gt::pct(8),
    `Files present` ~ gt::pct(9),
    `Size checks` ~ gt::pct(9),
    `SHA-256 checks` ~ gt::pct(11),
    Check ~ gt::pct(8)
  )
```

### MDER artifact identities

```{r}
#| label: tbl-current-mder-provenance
#| tbl-cap: "Exact decisions, manifests, and bundle identities governing MDER in the current shared inputs."

mder_provenance_display |>
  dplyr::mutate(
    `Exact path or identity` = wrap_provenance_path(
      .data$`Exact path or identity`
    ),
    `SHA-256` = wrap_hash(.data$`SHA-256`)
  ) |>
  prep_gt(
    source_note = paste0(
      "Current records were checked by SHA-256 during render. The historical ",
      "support gate is displayed only to prevent its superseded rule from ",
      "being mistaken for the active method."
    )
  ) |>
  gt::fmt_markdown(columns = c(`Exact path or identity`)) |>
  gt::cols_width(
    `Stored record` ~ gt::pct(24),
    `Exact path or identity` ~ gt::pct(35),
    `SHA-256` ~ gt::pct(31),
    Status ~ gt::pct(10)
  )
```

### Numerical-zero artifact identities

```{r}
#| label: tbl-l10-preparation-provenance
#| tbl-cap: "Exact METRIC-011 decisions, evidence, shared inputs, and preparation-manifest identities."

l10_provenance_display |>
  dplyr::transmute(
    `Stored record` = paste0(
      "**", .data$`Stored record`, "**<br>", .data$Status
    ),
    `Exact path and SHA-256` = paste0(
      "<code>",
      wrap_provenance_path(.data$`Exact path or identity`),
      "</code><br>SHA-256: ",
      wrap_hash(.data$`SHA-256`)
    )
  ) |>
  prep_gt(
    source_note = paste0(
      "The current files and all seven files listed by the METRIC-011 ",
      "evidence manifest were checked by byte size and SHA-256 during render."
    )
  ) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Stored record` ~ gt::pct(32),
    `Exact path and SHA-256` ~ gt::pct(68)
  )
```

### Producing scripts in execution order

The producing scripts are listed below in execution order. Their separation
is scientific provenance: each step has a narrower responsibility and a
stored output that can be checked before the next step begins. Rendering this
report reads those outputs only.

```{r}
#| label: tbl-script-execution-map
#| tbl-cap: "Producing scripts and function modules in execution order; none is executed while this report renders."

script_stage_labels <- c(
  "Derive the current metric layer",
  "Verify current MDER independently",
  "Record the MDER amendment audit",
  "Standardize participant inputs",
  "Assemble gap-timing-unaware inputs",
  "Verify repaired gap-timing-unaware MDER",
  "Add site and daylight context",
  "Preserve elapsed-time provenance",
  "Create shared model inputs",
  "Store H01 primary frames",
  "Store gap-timing-unaware H01 frames",
  "Store pre-analysis displays",
  "Record the preceding MDER repin and invariance",
  "Verify and reseal numerical-zero normalization"
)

script_map |>
  dplyr::mutate(
    `Step and producing code` = paste0(
      "**", .data$Step, ". ", script_stage_labels[.data$Step], "**<br>",
      "**Script:** <code>",
      wrap_provenance_path(.data$Producing_script),
      "</code><br>",
      "**Module:** <code>",
      wrap_provenance_path(.data$Function_module),
      "</code>"
    ),
    `Reads, operation, separation, and output` = paste0(
      .data$Reads_and_operation,
      "<br><br>**Stored output:** <code>",
      wrap_provenance_path(.data$Stored_output),
      "</code>"
    )
  ) |>
  dplyr::select(
    .data$`Step and producing code`,
    .data$`Reads, operation, separation, and output`
  ) |>
  prep_gt(
    source_note = paste0(
      "Paths identify the external production code and accepted outputs; ",
      "the render does not source these files."
    )
  ) |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_width(
    `Step and producing code` ~ gt::pct(38),
    `Reads, operation, separation, and output` ~ gt::pct(62)
  )
```

The following files and directories are the durable handoff from Preparation
06. They—not values recalculated by this report—enter the hypothesis-specific
analysis-preparation companions and model-fitting scripts.

```{r}
#| label: tbl-preparation06-handoff
#| tbl-cap: "Stored Preparation 06 artifacts passed to later analysis stages."

handoff_artifacts |>
  dplyr::rename(
    `Stored artifact` = .data$Stored_artifact,
    `Exact path` = .data$Exact_path,
    `Used next for` = .data$Next_use
  ) |>
  prep_gt(
    source_note = paste0(
      "H02–H11 inputs are added only after each hypothesis method is ",
      "approved and are documented in their own analysis-preparation ",
      "companions."
    )
  ) |>
  gt::cols_width(
    `Stored artifact` ~ gt::pct(25),
    `Exact path` ~ gt::pct(43),
    `Used next for` ~ gt::pct(32)
  )
```

```{r}
#| label: tbl-preparation06-environment
#| tbl-cap: "Environment and execution boundary recorded for this review render."

environment_summary |>
  prep_gt(
    source_note = paste0(
      "Rendering uses the project's R 4.6.1 library and the nathealth ",
      "Quarto profile."
    )
  ) |>
  gt::cols_width(
    Component ~ gt::pct(38),
    Value ~ gt::pct(62)
  )
```
