---
title: "Preparation 07: Inspect example days from each study site"
subtitle: "A display-only check of prepared one-minute light exposure, daylight, diary periods, and declared non-wear"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show supporting R code"
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-preparation-07
#| include: false

library(digest)
library(dplyr)
library(ggplot2)
library(gt)
library(readr)
library(scales)
library(tibble)

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()),
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0")
)

absolute_path <- function(path) {
  ifelse(grepl("^/", path), path, file.path(root, path))
}

read_stored_csv <- function(path) {
  full_path <- absolute_path(path)
  stopifnot(file.exists(full_path))
  readr::read_csv(full_path, show_col_types = FALSE, progress = FALSE)
}

assert_columns <- function(data, required, label) {
  missing_columns <- setdiff(required, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      label,
      " is missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  invisible(data)
}

file_sha256 <- function(path) {
  digest::digest(file = absolute_path(path), algo = "sha256")
}

prep_gt <- function(data, source_note = NULL) {
  output <- gt::gt(data) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(13),
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

wrap_path <- function(value) {
  gsub("/", "/\u200b", value, fixed = TRUE)
}

wrap_hash <- function(value) {
  gsub("(.{8})", "\\1\u200b", value, perl = TRUE)
}

format_minutes <- function(value) {
  paste0(scales::comma(value, accuracy = 1), " min")
}

equivalence_audit_path <-
  "audit/reconciliation/preparation07/site_solar_context_equivalence_audit.md"
equivalence_evidence_path <-
  "audit/reconciliation/preparation07/site_solar_context_equivalence_evidence.csv"
equivalence_seal_path <-
  "audit/reconciliation/preparation07/site_solar_context_equivalence_manifest.csv"
expected_equivalence_audit_sha256 <-
  "870cbb5d5f9414f61ee7db05b61a12eb51d56c9273e9ae7dfd05094984c47164"
expected_equivalence_evidence_sha256 <-
  "0757fc175c44fd33e9d8b06e4f9352699c18bfe50bbcf1c65d141f3f36a1cb97"
expected_equivalence_seal_sha256 <-
  "01a07b87799faeb8291bb6187ced96ab7abdcf2399ee77c5dcfd19d608b1194c"
historical_solar_rds_sha256 <-
  "b0e8de539572ee595cc91027e7a2d919ba01237f780e50a76e5e4468d497c4bf"
current_solar_rds_sha256 <-
  "39ffe488de86f5d7cdc56d65c582c9de31f9054f8565936b4ec74e01491f26d0"
current_solar_csv_sha256 <-
  "7dc64cc5026947ef96d6e4ab112bb767414420be97817f022082248db68d7028"

stopifnot(
  identical(
    file_sha256(equivalence_audit_path),
    expected_equivalence_audit_sha256
  ),
  identical(
    file_sha256(equivalence_evidence_path),
    expected_equivalence_evidence_sha256
  ),
  identical(
    file_sha256(equivalence_seal_path),
    expected_equivalence_seal_sha256
  )
)

equivalence_evidence <- read_stored_csv(equivalence_evidence_path)
equivalence_seal <- read_stored_csv(equivalence_seal_path)
assert_columns(
  equivalence_evidence,
  c(
    "check_id", "status", "evidence_level", "artifact_or_source",
    "expected", "observed", "interpretation"
  ),
  "stored site/daylight-context equivalence evidence"
)
assert_columns(
  equivalence_seal,
  c("path", "role", "bytes", "sha256", "producer", "r_version", "status"),
  "stored site/daylight-context equivalence seal"
)

evidence_pair_matches <- function(check_id, value) {
  evidence_row <- equivalence_evidence |>
    dplyr::filter(.data$check_id == .env$check_id)
  nrow(evidence_row) == 1L &&
    identical(evidence_row$expected[[1]], value) &&
    identical(evidence_row$observed[[1]], value)
}

expected_verifier_result <- "17 PASS and manifest::input_hashes FAIL"
expected_verifier_row <- equivalence_evidence |>
  dplyr::filter(.data$check_id == "independent_verifier_total")
equivalence_evidence_checks <- c(
  nrow(equivalence_evidence) == 17L,
  !anyDuplicated(equivalence_evidence$check_id),
  sum(equivalence_evidence$status == "PASS") == 16L,
  nrow(expected_verifier_row) == 1L,
  identical(
    expected_verifier_row$status[[1]],
    "EXPECTED_PROVENANCE_FAIL_ONLY"
  ),
  identical(expected_verifier_row$expected[[1]], expected_verifier_result),
  identical(expected_verifier_row$observed[[1]], expected_verifier_result),
  evidence_pair_matches("historical_solar_rds_pin", historical_solar_rds_sha256),
  evidence_pair_matches("current_solar_rds_pin", current_solar_rds_sha256),
  evidence_pair_matches(
    "context_csv_historical_current",
    current_solar_csv_sha256
  ),
  evidence_pair_matches(
    "full_context_frame",
    "618 rows; 47 columns; zero changed cells"
  ),
  evidence_pair_matches(
    "selected_day_context_strings",
    "9 keys and 4 fields exactly equal as stored strings"
  ),
  identical(
    file_sha256("artifacts/06_model_data/context/site_solar_context.csv"),
    current_solar_csv_sha256
  )
)
stopifnot(all(equivalence_evidence_checks))

expected_seal_paths <- c(equivalence_audit_path, equivalence_evidence_path)
seal_order <- match(expected_seal_paths, equivalence_seal$path)
equivalence_seal_checks <- c(
  nrow(equivalence_seal) == 2L,
  identical(sort(equivalence_seal$path), sort(expected_seal_paths)),
  !anyNA(seal_order),
  identical(
    as.numeric(equivalence_seal$bytes[seal_order]),
    c(9342, 4824)
  ),
  identical(
    equivalence_seal$sha256[seal_order],
    c(
      expected_equivalence_audit_sha256,
      expected_equivalence_evidence_sha256
    )
  ),
  all(equivalence_seal$r_version == "4.6.1"),
  all(equivalence_seal$status == "PASS")
)
sealed_files <- absolute_path(equivalence_seal$path)
equivalence_seal_checks <- c(
  equivalence_seal_checks,
  all(file.exists(sealed_files)),
  all(file.info(sealed_files)$size == equivalence_seal$bytes),
  all(
    vapply(
      equivalence_seal$path,
      file_sha256,
      character(1)
    ) == equivalence_seal$sha256
  )
)
stopifnot(all(equivalence_seal_checks))
equivalence_evidence_verified <-
  all(equivalence_evidence_checks) && all(equivalence_seal_checks)

manifest_path <- "artifacts/12_manifests/prepared_day_showcase_artifacts.csv"
expected_manifest_sha256 <-
  "c5ca66b2a6ec4fabbe57d08135db7f365bcd123365caab6c987bcb2b62f7a322"
current_manifest_sha256 <- file_sha256(manifest_path)
stopifnot(identical(current_manifest_sha256, expected_manifest_sha256))

manifest <- read_stored_csv(manifest_path)
assert_columns(
  manifest,
  c(
    "artifact_type", "path", "sha256", "bytes", "rows", "columns",
    "producer", "seed", "coverage_input_sha256",
    "solar_context_input_sha256", "site_metadata_input_sha256",
    "r_version", "lightlogr_version", "ggplot2_version"
  ),
  "prepared-day showcase manifest"
)
stopifnot(
  nrow(manifest) == 6L,
  !anyDuplicated(manifest$artifact_type),
  length(unique(manifest$seed)) == 1L,
  unique(manifest$seed) == 20260730L
)

listed_paths <- absolute_path(manifest$path)
listed_present <- file.exists(listed_paths)
listed_bytes <- rep(NA_real_, nrow(manifest))
listed_hashes <- rep(NA_character_, nrow(manifest))
listed_bytes[listed_present] <- file.info(listed_paths[listed_present])$size
listed_hashes[listed_present] <- vapply(
  listed_paths[listed_present],
  function(path) digest::digest(file = path, algo = "sha256"),
  character(1)
)

artifact_identity <- manifest |>
  dplyr::mutate(
    present = listed_present,
    bytes_match = .data$present & listed_bytes == .data$bytes,
    sha256_match = .data$present & listed_hashes == .data$sha256
  )
stopifnot(
  all(artifact_identity$present),
  all(artifact_identity$bytes_match),
  all(artifact_identity$sha256_match)
)

selected_days <- read_stored_csv(
  "artifacts/08_diagnostics/prepared_day_showcase/selected_days.csv"
)
eligible_counts <- read_stored_csv(
  "artifacts/08_diagnostics/prepared_day_showcase/eligible_day_counts.csv"
)
selection_settings <- read_stored_csv(
  "artifacts/08_diagnostics/prepared_day_showcase/selection_settings.csv"
)
source_data <- read_stored_csv(
  "artifacts/11_source_data/prepared_day_showcase.csv"
)
site_display <- read_stored_csv("config/site_display_registry.csv")

assert_columns(
  selected_days,
  c(
    "site", "location", "participant", "local_date", "wake_time",
    "sleep_preparation_time", "civil_dawn", "civil_dusk",
    "valid_melEDI_minutes", "valid_melEDI_percent",
    "declared_nonwear_minutes", "other_unavailable_minutes",
    "eligible_days_at_site", "selection_seed"
  ),
  "stored selected participant-days"
)
assert_columns(
  eligible_counts,
  c("site", "eligible_participant_days"),
  "stored eligible participant-day counts"
)
assert_columns(
  selection_settings,
  c("setting", "value"),
  "stored selection settings"
)
assert_columns(
  source_data,
  c(
    "site", "participant", "local_date", "clock_minute", "melEDI_lx",
    "availability", "declared_nonwear", "brown_period",
    "brown_reference_lx", "civil_dawn_wall_minute",
    "civil_dusk_wall_minute", "solar_depression_deg", "observed_run",
    "selection_seed"
  ),
  "stored one-minute figure source data"
)
assert_columns(
  site_display,
  c("site", "display_order", "display_name", "color_hex"),
  "submitted site display registry"
)

csv_objects <- list(
  selected_days_csv = selected_days,
  eligible_day_counts_csv = eligible_counts,
  selection_settings_csv = selection_settings,
  plot_source_data_csv = source_data
)
csv_dimensions <- tibble::tibble(
  artifact_type = names(csv_objects),
  current_rows = vapply(csv_objects, nrow, integer(1)),
  current_columns = vapply(csv_objects, ncol, integer(1))
) |>
  dplyr::left_join(
    manifest |>
      dplyr::select(
        .data$artifact_type,
        recorded_rows = .data$rows,
        recorded_columns = .data$columns
      ),
    by = "artifact_type"
  ) |>
  dplyr::mutate(
    dimensions_match =
      .data$current_rows == .data$recorded_rows &
      .data$current_columns == .data$recorded_columns
  )
stopifnot(all(csv_dimensions$dimensions_match))

input_identity <- tibble::tribble(
  ~input_key, ~Input, ~Path, ~`Manifest-recorded SHA-256`,
  "prepared_minutes",
  "Prepared near-eye minute data",
  "artifacts/03_coverage/light_glasses_coverage.rds",
  unique(manifest$coverage_input_sha256),
  "site_daylight",
  "Site- and date-specific daylight context",
  "artifacts/06_model_data/context/site_solar_context.rds",
  unique(manifest$solar_context_input_sha256),
  "site_metadata",
  "Site names, coordinates, and time zones",
  "config/site_metadata.csv",
  unique(manifest$site_metadata_input_sha256)
) |>
  dplyr::mutate(
    `Current SHA-256` = vapply(.data$Path, file_sha256, character(1)),
    identity_verified = dplyr::case_when(
      .data$input_key %in% c("prepared_minutes", "site_metadata") ~
        .data$`Manifest-recorded SHA-256` == .data$`Current SHA-256`,
      .data$input_key == "site_daylight" ~
        .data$`Manifest-recorded SHA-256` == historical_solar_rds_sha256 &
        .data$`Current SHA-256` == current_solar_rds_sha256 &
        equivalence_evidence_verified,
      .default = FALSE
    ),
    Verification = dplyr::case_when(
      !.data$identity_verified ~ "Check failed",
      .data$input_key == "site_daylight" ~
        "Equivalent values, different file version",
      .default = "Exact file identity"
    )
  )
stopifnot(all(input_identity$identity_verified))

eligible_consistency <- selected_days |>
  dplyr::select(
    .data$site,
    selected_record = .data$eligible_days_at_site
  ) |>
  dplyr::inner_join(eligible_counts, by = "site")
stopifnot(
  nrow(selected_days) == 9L,
  nrow(eligible_counts) == 9L,
  nrow(eligible_consistency) == 9L,
  all(
    eligible_consistency$selected_record ==
      eligible_consistency$eligible_participant_days
  ),
  nrow(source_data) == 12960L,
  nrow(source_data) == 1440L * nrow(selected_days),
  all(
    as.integer(
      table(
        interaction(
          source_data$site,
          source_data$participant,
          source_data$local_date,
          drop = TRUE
        )
      )
    ) == 1440L
  ),
  identical(sort(unique(selected_days$site)), sort(unique(site_display$site))),
  length(unique(selected_days$selection_seed)) == 1L,
  unique(selected_days$selection_seed) == 20260730L,
  length(unique(source_data$selection_seed)) == 1L,
  unique(source_data$selection_seed) == 20260730L
)

site_order <- site_display |>
  dplyr::arrange(.data$display_order)
selection_with_display <- selected_days |>
  dplyr::inner_join(site_order, by = "site") |>
  dplyr::arrange(.data$display_order) |>
  dplyr::mutate(
    panel_label = paste0(
      .data$display_name,
      "\n",
      .data$participant,
      " · ",
      as.character(.data$local_date)
    )
  )
panel_levels <- selection_with_display$panel_label

plot_data <- source_data |>
  dplyr::inner_join(
    selection_with_display |>
      dplyr::select(
        .data$site,
        .data$display_order,
        .data$display_name,
        .data$color_hex,
        .data$panel_label
      ),
    by = "site"
  ) |>
  dplyr::mutate(
    panel_label = factor(.data$panel_label, levels = panel_levels),
    state_display = dplyr::if_else(
      .data$declared_nonwear,
      "Declared non-wear",
      .data$brown_period
    )
  )

solar_panels <- plot_data |>
  dplyr::distinct(
    .data$site,
    .data$panel_label,
    .data$civil_dawn_wall_minute,
    .data$civil_dusk_wall_minute
  )
civil_night <- dplyr::bind_rows(
  solar_panels |>
    dplyr::transmute(
      .data$site,
      .data$panel_label,
      xmin = 0,
      xmax = .data$civil_dawn_wall_minute
    ),
  solar_panels |>
    dplyr::transmute(
      .data$site,
      .data$panel_label,
      xmin = .data$civil_dusk_wall_minute,
      xmax = 1440
    )
)

state_colours <- c(
  "Waking daytime" = "#80AFC4",
  "Three hours before sleep" = "#F2B35F",
  "Sleep" = "#2F4054",
  "Declared non-wear" = "#E13B2D"
)

make_showcase_plot <- function(sites, show_legend = FALSE) {
  display_data <- plot_data |>
    dplyr::filter(.data$site %in% sites)
  display_night <- civil_night |>
    dplyr::filter(.data$site %in% sites)

  ggplot2::ggplot(display_data) +
    ggplot2::geom_rect(
      data = display_night,
      ggplot2::aes(
        xmin = .data$xmin,
        xmax = .data$xmax,
        ymin = -Inf,
        ymax = Inf
      ),
      inherit.aes = FALSE,
      fill = "#244B78",
      alpha = 0.10
    ) +
    ggplot2::geom_tile(
      ggplot2::aes(
        x = .data$clock_minute,
        y = -0.25,
        fill = .data$state_display
      ),
      width = 1.05,
      height = 0.50
    ) +
    ggplot2::geom_step(
      ggplot2::aes(
        x = .data$clock_minute,
        y = .data$brown_reference_lx,
        group = .data$site
      ),
      colour = "#555555",
      linewidth = 0.45,
      linetype = "22",
      alpha = 0.9
    ) +
    ggplot2::geom_line(
      ggplot2::aes(
        x = .data$clock_minute,
        y = .data$melEDI_lx,
        group = interaction(.data$site, .data$observed_run)
      ),
      colour = "#202020",
      linewidth = 0.30,
      na.rm = TRUE
    ) +
    ggplot2::geom_point(
      ggplot2::aes(
        x = .data$clock_minute,
        y = .data$melEDI_lx
      ),
      colour = "#202020",
      size = 0.22,
      alpha = 0.65,
      na.rm = TRUE
    ) +
    ggplot2::facet_wrap(
      ggplot2::vars(.data$panel_label),
      nrow = 1,
      drop = TRUE
    ) +
    ggplot2::scale_x_continuous(
      breaks = seq(0, 1200, by = 240),
      labels = sprintf("%02d:00", seq(0, 20, by = 4)),
      limits = c(0, 1440),
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = LightLogR::symlog_trans(base = 10, thr = 1, scale = 1),
      breaks = c(0, 1, 10, 250, 1000, 10000, 100000),
      labels = scales::label_number(big.mark = "\u2009", accuracy = 1),
      expand = ggplot2::expansion(mult = c(0.01, 0.03))
    ) +
    ggplot2::scale_fill_manual(
      values = state_colours,
      breaks = names(state_colours),
      name = "Diary period and declared non-wear"
    ) +
    ggplot2::coord_cartesian(ylim = c(-0.5, 100000)) +
    ggplot2::labs(
      x = "Local clock time",
      y = "melEDI (lx; one-minute values)"
    ) +
    ggplot2::guides(
      fill = ggplot2::guide_legend(nrow = 1, byrow = TRUE)
    ) +
    ggplot2::theme_minimal(base_size = 10.5) +
    ggplot2::theme(
      legend.position = if (show_legend) "bottom" else "none",
      legend.title = ggplot2::element_text(size = 10),
      legend.text = ggplot2::element_text(size = 10),
      strip.background = ggplot2::element_rect(
        fill = "#F3F4F6",
        colour = "#D1D5DB",
        linewidth = 0.4
      ),
      strip.text = ggplot2::element_text(
        size = 10,
        face = "bold",
        lineheight = 1.05,
        margin = ggplot2::margin(4, 3, 4, 3)
      ),
      axis.text = ggplot2::element_text(size = 10),
      axis.title = ggplot2::element_text(size = 10),
      panel.grid.minor = ggplot2::element_blank(),
      panel.spacing = grid::unit(8, "pt"),
      plot.margin = ggplot2::margin(5, 7, 5, 5, "pt")
    )
}

current_checks <- tibble::tribble(
  ~Check, ~Result, ~Status,
  "Showcase manifest identity",
  "The current manifest SHA-256 matches the accepted value.",
  "PASS",
  "Scientific input dispositions",
  paste(
    "Two inputs retain exact file identity; site/daylight values are",
    "verified equivalent across file versions."
  ),
  ifelse(all(input_identity$identity_verified), "PASS", "FAIL"),
  "Sealed site/daylight reconciliation",
  paste(
    "The stored evidence contains 16 successful checks and one expected",
    "historical/current file-identity difference."
  ),
  ifelse(equivalence_evidence_verified, "PASS", "FAIL"),
  "Six manifest-listed display files",
  paste0(sum(artifact_identity$present), " / 6 files are present."),
  ifelse(all(artifact_identity$present), "PASS", "FAIL"),
  "Stored byte sizes",
  paste0(sum(artifact_identity$bytes_match), " / 6 sizes match the manifest."),
  ifelse(all(artifact_identity$bytes_match), "PASS", "FAIL"),
  "Stored file fingerprints",
  paste0(sum(artifact_identity$sha256_match), " / 6 SHA-256 checks match."),
  ifelse(all(artifact_identity$sha256_match), "PASS", "FAIL"),
  "Four stored CSV dimensions",
  paste0(sum(csv_dimensions$dimensions_match), " / 4 row-and-column checks match."),
  ifelse(all(csv_dimensions$dimensions_match), "PASS", "FAIL"),
  "Selected-day structure",
  "Nine sites × 1,440 local-clock minutes = 12,960 rows.",
  "PASS",
  "Fixed seed recorded in stored outputs",
  "Seed 20260730 is consistent in the selection and one-minute source data.",
  "PASS"
)
stopifnot(all(current_checks$Status == "PASS"))
```

## Purpose and position in the preparation chain

This final preparation page is a **display-only inspection** of the prepared
one-minute data. It shows one participant-day from each study site before any
hypothesis model is fitted. The profiles combine melanopic equivalent daylight
illuminance (melEDI; lx), local daylight, diary-defined periods, and declared
non-wear so that readers can see what the accepted minute-level records look
like.

The exact scientific inputs are the prepared near-eye minute data from
Preparation 02, the site- and date-specific daylight context assembled in
Preparation 06, and the site metadata used to interpret local time. The fixed
selection and its display files are illustrative only: none enters an H01–H11
model. Hypothesis analyses receive their stored inputs from Preparation 06,
so there is no hypothesis-analysis handoff from this page.

```{mermaid}
%%| label: fig-example-day-chain
%%| fig-cap: "Position of Preparation 07 in the data-to-analysis chain. It makes a display from accepted preparation artifacts but supplies no model input."
%%| fig-alt: "Prepared near-eye minute data, stored daylight context, and site metadata feed a reproducible display builder. Its stored selection, source data, and figures feed only the Preparation 07 inspection page; an explicit endpoint states that there is no hypothesis-analysis handoff."
flowchart TB
  A["Prepared near-eye minutes<br/>Preparation 02"] --> D["Reproducible display build"]
  B["Site and date daylight context<br/>Preparation 06"] --> D
  C["Site metadata<br/>local time interpretation"] --> D
  D --> E["Stored selection, source data,<br/>PNG, SVG, and manifest"]
  E --> F["Preparation 07<br/>reader-facing inspection"]
  F -. "display only" .-> G["No hypothesis-analysis handoff"]
```

## Inputs and selection requirements

A participant-day could enter the selection pool only if it passed the general
coverage rules from Preparation 02: at least 50% valid observations within
each summarized hour and at least 80% valid observations across the full
24-hour day. It also needed all 1,440 local-clock minutes and one complete
diary sleep interval with a visible sleep start and wake transition. One day
per site was then selected reproducibly; the selection was not repeated to
find especially tidy or favourable profiles.

The 80% daily rule covers the complete 24-hour measurement cycle. It combines
near-eye measurements while participants were awake with bedside measurements
during the diary-defined sleep interval. Retaining both contexts preserves the
full day; it does not imply that the bedside sleep measurement was made at the
eye.

## Selected participant-days

The two compact tables separate timing from data availability so that neither
has to be compressed into narrow cells. Submitted site names, order, and colour
markers come from `config/site_display_registry.csv`.

```{r}
#| label: tbl-example-day-selection
#| tbl-cap: "Participant-days selected reproducibly, in submitted site order."

selection_table <- selection_with_display |>
  dplyr::transmute(
    Site = paste0(
      "<span style=\"color:",
      .data$color_hex,
      "\" aria-hidden=\"true\">●</span> ",
      .data$display_name
    ),
    Participant = .data$participant,
    Date = as.character(.data$local_date),
    `Diary sleep start` = .data$sleep_preparation_time,
    Wake = .data$wake_time,
    `Civil dawn–dusk` = paste0(.data$civil_dawn, "–", .data$civil_dusk)
  )

selection_table |>
  prep_gt(
    source_note = paste(
      "A coloured dot identifies the registered site colour; sleep and",
      "daylight times are local wall-clock times."
    )
  ) |>
  gt::fmt_markdown(columns = Site) |>
  gt::cols_align(
    align = "center",
    columns = c(Date, `Diary sleep start`, Wake, `Civil dawn–dusk`)
  ) |>
  gt::cols_width(
    Site ~ gt::pct(22),
    Participant ~ gt::pct(19),
    Date ~ gt::pct(15),
    `Diary sleep start` ~ gt::pct(15),
    Wake ~ gt::pct(11),
    `Civil dawn–dusk` ~ gt::pct(18)
  )
```

```{r}
#| label: tbl-example-day-availability
#| tbl-cap: "Observed melEDI and unavailable minutes in the nine selected days."

availability_table <- selection_with_display |>
  dplyr::transmute(
    Site = paste0(
      "<span style=\"color:",
      .data$color_hex,
      "\" aria-hidden=\"true\">●</span> ",
      .data$display_name
    ),
    `Valid melEDI` = paste0(
      scales::comma(.data$valid_melEDI_minutes, accuracy = 1),
      " min (",
      scales::percent(.data$valid_melEDI_percent / 100, accuracy = 0.1),
      ")"
    ),
    `Declared non-wear` = format_minutes(
      .data$declared_nonwear_minutes
    ),
    `Other unavailable` = format_minutes(
      .data$other_unavailable_minutes
    ),
    `Eligible pool at site` = scales::comma(
      .data$eligible_days_at_site,
      accuracy = 1
    )
  )

availability_table |>
  prep_gt(
    source_note = paste0(
      "Across the displayed days, ",
      scales::comma(sum(selected_days$valid_melEDI_minutes), accuracy = 1),
      " of 12,960 minutes have valid melEDI; ",
      scales::comma(
        sum(selected_days$declared_nonwear_minutes),
        accuracy = 1
      ),
      " are declared non-wear and ",
      scales::comma(
        sum(selected_days$other_unavailable_minutes),
        accuracy = 1
      ),
      " are unavailable for another reason."
    )
  ) |>
  gt::fmt_markdown(columns = Site) |>
  gt::cols_align(
    align = "center",
    columns = c(
      `Valid melEDI`,
      `Declared non-wear`,
      `Other unavailable`,
      `Eligible pool at site`
    )
  ) |>
  gt::cols_width(
    Site ~ gt::pct(24),
    `Valid melEDI` ~ gt::pct(23),
    `Declared non-wear` ~ gt::pct(18),
    `Other unavailable` ~ gt::pct(18),
    `Eligible pool at site` ~ gt::pct(17)
  )
```

The valid share ranges from
`r scales::percent(min(selected_days$valid_melEDI_percent) / 100, accuracy = 0.1)`
to
`r scales::percent(max(selected_days$valid_melEDI_percent) / 100, accuracy = 0.1)`.
All nine selected days therefore remain above the 80% daily threshold. A zero
melEDI value is a measurement and remains visible. Declared non-wear and other
unavailable minutes remain missing, so the exposure line is not connected
across them. The 50% hourly rule governs hourly summaries; it does not erase
an observed one-minute value from this inspection.

## How to read the one-minute profiles

- Civil night is the shaded part of the day when the Sun is more than 6°
  below the horizon; the unshaded interval is civil day.
- The strip below zero shows the diary-defined waking daytime, the three hours
  before sleep, and sleep. Red overwrites the strip where the wear log records
  “off” outside diary-defined sleep; wear-log entries do not define sleep.
- The dashed step is a contextual reference rather than a fitted value: 250 lx
  during waking daytime, 10 lx during the three hours before sleep, and 1 lx
  during sleep.
- All panels use the same transformed y-axis so that zero, dim exposure, and
  high daytime exposure can appear together. The transformation changes only
  display spacing; the source CSV retains the stored melEDI values in lx.

## One-minute profiles

The exact 12,960 rows shown below are available as the paired
[figure source-data CSV](../../artifacts/11_source_data/prepared_day_showcase.csv).
For this HTML page, those stored source data are split into three three-panel
figures so that axes, times, participant labels, and annotations remain
readable at the final display width.

```{r}
#| label: fig-example-days-1
#| fig-cap: "Example days 1–3 in submitted site order. The black trace and points are stored one-minute melEDI; dark-blue shading marks civil night; the strip below zero marks waking daytime, the three hours before sleep, diary-defined sleep, and declared non-wear. The dashed step gives contextual values of 250 lx, 10 lx, and 1 lx for those three diary periods."
#| fig-alt: "Three panels show one-minute melEDI for selected participant-days in Borås (SE), Delft (NL), and Dortmund (DE). Each panel spans a local 24-hour day. Black exposure traces vary over the day and contain gaps where values are unavailable. Dark-blue background marks civil night, and a coloured strip below zero marks waking daytime, the three hours before sleep, sleep, and declared non-wear."
#| fig-width: 9.5
#| fig-height: 4.2
#| fig-dpi: 150
#| out-width: "100%"

make_showcase_plot(c("RISE", "THUAS", "BAUA"), show_legend = TRUE)
```

```{r}
#| label: fig-example-days-2
#| fig-cap: "Example days 4–6 in submitted site order. Values, shading, bands, and the contextual dashed step have the same meaning as in @fig-example-days-1."
#| fig-alt: "Three panels show one-minute melEDI for selected participant-days in Tübingen (DE), Munich (DE), and Madrid (ES). Each panel spans a local 24-hour day with black exposure traces, dark-blue civil-night shading, coloured diary-period and non-wear bands below zero, and blank segments where melEDI is unavailable."
#| fig-width: 9.5
#| fig-height: 3.7
#| fig-dpi: 150
#| out-width: "100%"

make_showcase_plot(c("MPI", "TUM", "FUSPCEU"))
```

```{r}
#| label: fig-example-days-3
#| fig-cap: "Example days 7–9 in submitted site order. Values, shading, bands, and the contextual dashed step have the same meaning as in @fig-example-days-1."
#| fig-alt: "Three panels show one-minute melEDI for selected participant-days in Izmir (TR), San José (CR), and Kumasi (GH). Each panel spans a local 24-hour day with black exposure traces, dark-blue civil-night shading, coloured diary-period and non-wear bands below zero, and blank segments where melEDI is unavailable."
#| fig-width: 9.5
#| fig-height: 3.7
#| fig-dpi: 150
#| out-width: "100%"

make_showcase_plot(c("IZTECH", "UCR", "KNUST"))
```

## Technical provenance and checks on the current stored bundle

::: {.callout-note title="What is calculated during this render?"}
This report reads the stored selection, settings, one-minute figure source
data, display manifest, and site display registry. It calculates lightweight
counts for the tables and redraws the profiles from the stored source-data CSV
for a more readable page layout. It does not repeat the selection, rebuild an
accepted file, recalculate a metric, or fit a model. File-identity checks are
described in the technical provenance section.
:::

### Exact input identities

The display manifest has SHA-256
`r current_manifest_sha256`. It truthfully records the three file versions
used when the example days were selected and plotted.

The stored example-day display was created from an earlier version of the
site/daylight-context file. The accepted current file contains the same
site/daylight values used here; only file-level provenance metadata changed.

The table distinguishes exact file identity from independently sealed value
equivalence without loading the two RDS data objects.

The stored reconciliation is pinned by audit report
`r wrap_path(equivalence_audit_path)` (SHA-256
`r wrap_hash(expected_equivalence_audit_sha256)`), evidence file
`r wrap_path(equivalence_evidence_path)` (SHA-256
`r wrap_hash(expected_equivalence_evidence_sha256)`), and non-circular seal
`r wrap_path(equivalence_seal_path)` (SHA-256
`r wrap_hash(expected_equivalence_seal_sha256)`).

```{r}
#| label: tbl-example-day-inputs
#| tbl-cap: "Historical producer and current file identities for the example-day scientific inputs."

input_identity |>
  dplyr::transmute(
    .data$Input,
    Path = wrap_path(.data$Path),
    `Manifest-recorded SHA-256` = wrap_hash(
      .data$`Manifest-recorded SHA-256`
    ),
    `Current SHA-256` = wrap_hash(.data$`Current SHA-256`),
    .data$Verification
  ) |>
  prep_gt(
    source_note = paste(
      "The historical manifest remains unchanged. Exact identity is required",
      "for the near-eye and site-metadata rows; the site/daylight row uses",
      "the sealed reconciliation. Neither RDS file is opened during render."
    )
  ) |>
  gt::cols_width(
    Input ~ gt::pct(19),
    Path ~ gt::pct(22),
    `Manifest-recorded SHA-256` ~ gt::pct(20),
    `Current SHA-256` ~ gt::pct(20),
    Verification ~ gt::pct(19)
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold", color = "#1B6E3C"),
    locations = gt::cells_body(columns = Verification)
  )
```

### Complete stored selection and display rules

```{r}
#| label: tbl-example-day-rules
#| tbl-cap: "Stored rules used to define and display eligible example days."

rule_labels <- c(
  selection_seed = "Reproducible selection",
  placement = "Measurement placement",
  coverage_rule = "General coverage requirement",
  sleep_timing_requirement = "Diary timing requirement",
  sleep_authority = "Sleep definition",
  nonwear_display_rule = "Declared non-wear",
  photoperiod_definition = "Daylight definition",
  plot_time_basis = "Time axis"
)

rule_explanations <- c(
  selection_seed = "One draw per site with fixed random seed 20260730.",
  placement = "Near-eye measurements are shown while awake; the prepared sleep interval uses the bedside context.",
  coverage_rule = "At least 50% valid within each summarized hour and at least 80% valid across the 24-hour day.",
  sleep_timing_requirement = "Exactly one visible diary sleep start and one wake transition from a complete diary interval.",
  sleep_authority = "The sleep diary defines the interval from sleep preparation to wake.",
  nonwear_display_rule = "Wear-log ‘off’ outside diary-defined sleep is displayed as declared non-wear.",
  photoperiod_definition = "Civil dawn to civil dusk, when the Sun is no more than 6° below the horizon.",
  plot_time_basis = "Local wall-clock minute for display; the corresponding UTC time remains in the source data."
)

selection_settings |>
  dplyr::filter(.data$setting %in% names(rule_labels)) |>
  dplyr::mutate(
    Rule = unname(rule_labels[.data$setting]),
    `Scientific meaning` = unname(rule_explanations[.data$setting])
  ) |>
  dplyr::select(.data$Rule, .data$`Scientific meaning`) |>
  prep_gt(
    source_note = paste(
      "These explanations translate the exact stored settings in",
      "artifacts/08_diagnostics/prepared_day_showcase/selection_settings.csv."
    )
  ) |>
  gt::cols_width(
    Rule ~ gt::pct(28),
    `Scientific meaning` ~ gt::pct(72)
  )
```

### Production and verification scripts

The display has its own narrow production workflow so that an illustrative
choice cannot silently alter the prepared data. The producer reads accepted
inputs and writes only display artifacts. The independent verifier is kept in
a separate module and does not source the producer, allowing it to reconstruct
the selection and compare the stored outputs by a different route.

```{r}
#| label: tbl-example-day-scripts
#| tbl-cap: "Producing and verification scripts in execution order; none is executed while this report renders."

script_map <- tibble::tribble(
  ~Step, ~Script, ~`What it does and why it is separate`,
  "1", "scripts/pipeline/build_prepared_day_showcase.R",
  paste(
    "Production driver. It starts the producer, then the independent",
    "verifier, and stops if verification does not pass."
  ),
  "2", "scripts/pipeline/prepared_day_showcase.R",
  paste(
    "Reads the accepted near-eye coverage object, daylight context, and",
    "site metadata; identifies eligible days; selects one per site with",
    "seed 20260730; and writes the selection, settings, source data, PNG,",
    "SVG, and manifest. Its output is illustrative, not analytical input."
  ),
  "3", "scripts/pipeline/verify_prepared_day_showcase_artifacts.R",
  paste(
    "Independently reopens the scientific inputs and display files; checks",
    "the artifact set and fingerprints; reproduces the fixed-seed choice;",
    "and compares minute values, non-wear, daylight, and figure files. It",
    "writes nothing and is not called by this report."
  )
)

script_map |>
  dplyr::mutate(Script = wrap_path(.data$Script)) |>
  prep_gt(
    source_note = paste(
      "The full production verifier is deliberately not rerun here because",
      "it would reload prepared data and reconstruct the selection."
    )
  ) |>
  gt::cols_width(
    Step ~ gt::px(48),
    Script ~ gt::pct(32),
    `What it does and why it is separate` ~ gt::pct(60)
  ) |>
  gt::cols_align(align = "center", columns = Step)
```

### Current bundle checks

The production verifier is designed to reopen the prepared inputs, reproduce
the fixed-seed selection, and compare the selected minutes, state masks,
daylight context, and figure files. That result-producing verifier was not run
for this documentation render. The table below is intentionally narrower: it
reports only file identity, the sealed site/daylight equivalence evidence,
schema, and lightweight structural checks performed on the current bundle
while this page rendered.

```{r}
#| label: tbl-example-day-current-checks
#| tbl-cap: "Bounded identity, equivalence, and structure checks performed on the current example-day bundle during this render."

current_checks |>
  prep_gt(
    source_note = paste(
      "PASS means exact file identity where required or sealed scientific",
      "equivalence for site/daylight context. It is not a rerun of the",
      "independent fixed-seed reconstruction verifier."
    )
  ) |>
  gt::cols_width(
    Check ~ gt::pct(34),
    Result ~ gt::pct(56),
    Status ~ gt::pct(10)
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold", color = "#1B6E3C"),
    locations = gt::cells_body(columns = Status)
  )
```

### Stored display artifacts and analytical handoff

All six display artifacts were produced by
`scripts/pipeline/prepared_day_showcase.R` with R
`r unique(manifest$r_version)`, LightLogR
`r unique(manifest$lightlogr_version)`, and ggplot2
`r unique(manifest$ggplot2_version)`. Their current byte sizes and SHA-256
fingerprints match the manifest.
The complete production display is stored as
[PNG](../../artifacts/10_figures/prepared_day_showcase.png) and
[vector SVG](../../artifacts/10_figures/prepared_day_showcase.svg).

```{r}
#| label: tbl-example-day-artifacts
#| tbl-cap: "Exact stored artifacts produced for the example-day inspection."

artifact_labels <- c(
  selected_days_csv = "Selected participant-days",
  eligible_day_counts_csv = "Eligible pool sizes by site",
  selection_settings_csv = "Selection and display settings",
  plot_source_data_csv = "Paired one-minute figure source data",
  figure_png = "Accepted high-resolution figure",
  figure_svg = "Accepted vector figure"
)
artifact_roles <- c(
  selected_days_csv = "Identifies the fixed participant-day selected at each site and summarizes its timing and availability.",
  eligible_day_counts_csv = "Records how many participant-days were eligible before the one-per-site draw.",
  selection_settings_csv = "Records the seed, coverage, diary, non-wear, daylight, and time-axis rules.",
  plot_source_data_csv = "Contains all 12,960 one-minute rows displayed in the profiles.",
  figure_png = "Stores the production raster display for review and reuse.",
  figure_svg = "Stores the same production display in resolution-independent form."
)

manifest |>
  dplyr::mutate(
    Artifact = unname(artifact_labels[.data$artifact_type]),
    Purpose = unname(artifact_roles[.data$artifact_type]),
    Shape = dplyr::if_else(
      !is.na(.data$rows),
      paste0(
        scales::comma(.data$rows, accuracy = 1),
        " rows × ",
        .data$columns,
        " columns"
      ),
      toupper(tools::file_ext(.data$path))
    ),
    Path = wrap_path(.data$path)
  ) |>
  dplyr::select(.data$Artifact, .data$Purpose, .data$Shape, .data$Path) |>
  prep_gt(
    source_note = paste0(
      "Current manifest SHA-256: ",
      wrap_hash(current_manifest_sha256),
      "."
    )
  ) |>
  gt::cols_width(
    Artifact ~ gt::pct(22),
    Purpose ~ gt::pct(38),
    Shape ~ gt::pct(16),
    Path ~ gt::pct(24)
  )
```

These artifacts document and visualize already prepared data. They do not
change an exposure value, state assignment, inclusion decision, sample, model
frame, or hypothesis result, and they are not inputs to the hypothesis
analyses.
