---
title: "Preparation 07: Example days from each study site"
subtitle: "One-minute melEDI, daylight, diary periods, and declared non-wear"
execute:
  echo: true
  warning: true
  message: true
  error: false
  freeze: false
---

## Purpose

This page shows what the prepared one-minute data look like before any
hypothesis model is fitted. It displays one participant-day from each study
site, together with local daylight, diary-defined periods, and declared
non-wear.

The examples were chosen with a fixed random seed. A day could be selected
only if it passed the main coverage rule, contained all 1,440 local-clock
minutes, and had a complete diary sleep window with one visible sleep-window
start and one wake time. The random selection was not repeated to find
especially tidy or favourable examples.

## Prepare the review page

```{r}
#| label: setup-example-days

root_hint <- Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd())
source(file.path(root_hint, "scripts/pipeline/paths_io.R"))
root <- project_root(root_hint)
source(file.path(
  root,
  "scripts/pipeline/verify_prepared_day_showcase_artifacts.R"
))

showcase_paths <- prepared_day_showcase_verifier_paths(root)
verification <- verify_prepared_day_showcase_artifacts(root = root)
if (!identical(verification$status, "PASS")) {
  print(verification$failures)
  stop("Example-day files did not pass verification.", call. = FALSE)
}

selected_days <- readr::read_csv(
  showcase_paths$selected_days,
  show_col_types = FALSE,
  progress = FALSE
)
eligible_counts <- readr::read_csv(
  showcase_paths$eligible_day_counts,
  show_col_types = FALSE,
  progress = FALSE
)
```

## Selected participant-days

The 80% daily rule applies to the complete 24-hour measurement cycle. It
combines near-eye measurements while participants were awake with bedside
measurements during the diary-defined sleep window. This is an intentional
way to retain information across the full day; it does not imply that the
bedside sleep measurement was made at the eye.

```{r}
#| label: tbl-example-day-selection
#| tbl-cap: "Participant-days selected with the fixed random seed."

selection_table <- selected_days |>
  dplyr::transmute(
    Site = paste0(.data$location, " (", .data$site, ")"),
    Participant = .data$participant,
    Date = as.character(.data$local_date),
    `Diary sleep-window start` = .data$sleep_preparation_time,
    Wake = .data$wake_time,
    `Civil dawn–dusk` = paste0(.data$civil_dawn, "–", .data$civil_dusk),
    `Valid melEDI` = sprintf(
      "%s min (%.1f%%)",
      .data$valid_melEDI_minutes,
      .data$valid_melEDI_percent
    ),
    `Declared non-wear` = paste0(
      .data$declared_nonwear_minutes,
      " min"
    ),
    `Other unavailable` = paste0(
      .data$other_unavailable_minutes,
      " min"
    ),
    `Eligible days at site` = .data$eligible_days_at_site
  )

knitr::kable(selection_table, align = "llcccccccc")
```

The table reports the actual number of one-minute melEDI values shown for
each day. Declared non-wear and other unavailable minutes remain missing; the
plot does not connect the line across those gaps. The 50% hourly requirement
does not remove observed one-minute values from this figure. It is used only
when an hourly summary is calculated.

## One-minute profiles

![One randomly selected qualifying participant-day from each site. The black
trace is the prepared one-minute melEDI series. Dark blue background shading
marks civil night. The opaque strip below zero shows waking daytime, the three
hours before sleep, diary-defined sleep, and declared non-wear. The dashed step
shows the 250 lx, 10 lx, and 1 lx contextual reference values for those three
diary periods.](../../artifacts/10_figures/prepared_day_showcase.png){#fig-example-days
fig-alt="Nine panels, one per study site, show one-minute melEDI across a local
24-hour day. A coloured strip below each panel marks waking daytime in blue,
the three hours before sleep in orange, sleep in dark navy, and declared
non-wear in red. Translucent dark blue background marks civil night. Black
exposure traces and gaps differ across sites and participant-days."
width="100%"}

### How to read the figure

- The black line and points are one-minute melEDI values. Zero is a measured
  value and is retained; a blank segment means that no analysis-ready melEDI
  value was available because the record or signal was missing or invalid, or
  because the complete day failed the 80% rule. A partly observed hour is not
  blanked merely because it contains fewer than 30 valid minutes.
- The background marks photoperiod: the unshaded interval is civil day and the
  translucent dark-blue interval is civil night, when the Sun was more than
  6° below the horizon.
- The opaque strip below zero shows the daily periods used in the analysis:
  waking daytime, the three hours before sleep, and the diary-defined sleep
  window. “Brown period” is the technical variable name for these three
  periods.
- Red sections overwrite the same strip where the wear log recorded “off”
  outside the diary-defined sleep window. Wear-log entries do not determine
  sleep.
- The dashed step is a contextual reference, not a fitted value: 250 lx
  during waking daytime, 10 lx in the three hours before sleep, and 1 lx
  during sleep.

## Checks completed

The display was checked against the prepared one-minute data rather than
against a manually copied figure. A SHA-256 checksum is a digital fingerprint
that changes if a file changes.

```{r}
#| label: tbl-example-day-verification
#| tbl-cap: "Independent checks of the example-day files."

reported_checks <- c(
  "files::all_required",
  "manifest::file_hashes",
  "selection::one_day_per_site",
  "selection::fixed_seed_reproduction",
  "source_data::complete_minute_grid",
  "source_data::prepared_minute_parity",
  "source_data::nonwear_mask",
  "source_data::solar_context_parity",
  "figures::nonempty_png_and_svg"
)
check_names <- c(
  "All required inputs and outputs are present",
  "Digital fingerprints match the files",
  "Exactly one day is shown for every site",
  "The fixed random selection can be reproduced",
  "Every displayed day contains 1,440 clock minutes",
  "Displayed values and states match the prepared data",
  "Declared non-wear has no analysis-ready melEDI",
  "Dawn, dusk, and photoperiod match the site and date",
  "Both high-resolution PNG and vector SVG exist"
)
verification_table <- verification$checks |>
  dplyr::filter(.data$check_id %in% reported_checks) |>
  dplyr::mutate(
    Check = check_names[match(.data$check_id, reported_checks)]
  ) |>
  dplyr::select(Check, Status = status)

knitr::kable(verification_table, align = "lc")
```

All `r nrow(verification$checks)` programmed checks passed. The source-data
file contains `r format(verification$summary$source_rows, big.mark = ",")`
one-minute rows: 1,440 rows for each of the nine selected participant-days.

The exact values behind the figure are stored in
`artifacts/11_source_data/prepared_day_showcase.csv`; the selection table and
selection settings are stored in
`artifacts/08_diagnostics/prepared_day_showcase/`.
