---
title: "H03 Stage 1: audit and approved analysis plan"
subtitle: "Nature Health analysis audit — revised model architecture"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    embed-resources: true
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H03_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

stopifnot(getRversion() == "4.6.1")

project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))

library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(gt)

stage1_gt <- function(data, rowname_col = NULL, groupname_col = NULL) {
  gt(data, rowname_col = rowname_col, groupname_col = groupname_col) |>
    opt_row_striping() |>
    tab_options(
      table.font.size = px(13),
      data_row.padding = px(4),
      heading.align = "left"
    )
}

formula_text <- function(x) {
  paste(deparse(x, width.cutoff = 500L), collapse = " ")
}

site_registry <- read.csv(
  file.path(project_root, "config/site_display_registry.csv"),
  check.names = FALSE
) |>
  arrange(display_order)
site_levels <- site_registry$site

category_dictionary <- read.csv(
  file.path(
    project_root,
    "audit/reconciliation/preparation06/category_support/h03_category_dictionary.csv"
  ),
  check.names = FALSE
)
category_levels <- category_dictionary$category_label
category_codes <- category_dictionary$category_code
names(category_codes) <- category_levels

category_short <- c(
  electric_indoor = "Indoor electric",
  electric_outdoor = "Outdoor electric",
  daylight_indoor = "Indoor daylight",
  daylight_outdoor = "Outdoor daylight",
  display = "Emissive display",
  sleep_darkness = "Sleep darkness",
  sleep_external_light = "External light during sleep"
)

reference_label <- "Electric light source indoors"
reference_code <- unname(category_codes[[reference_label]])

core_heterogeneity_levels <- c(
  "Electric light source indoors",
  "Daylight indoors",
  "Daylight outdoors (including shade)",
  "Emissive display light",
  "Darkness during sleep"
)

working_tweedie_power <- 1.539919
```

::: {.callout-warning title="Stage 1 verdict — revised after bounded architecture pilots"}
**The author-approved H03 primary architecture is now a population-averaged
quasi-Tweedie mean model with participant-clustered uncertainty.** The signed
hypothesis covers all seven prespecified primary-light-source categories. V0
removed two categories, combined category association and site heterogeneity
in one likelihood-ratio test, and described multiplicity differently from the
code. The current candidate data support all seven pooled categories, although
some category-by-site contrasts are support-non-estimable.

A bounded, outcome-blinded-to-effects model-construction pilot showed that the
initial Tweedie mixed model converged but was singular and failed important
distributional checks. The replacement quasi-Tweedie mean model converged,
was full rank, and produced a finite positive participant-cluster covariance.
Only numerical adequacy diagnostics were inspected: no category estimate,
confidence interval, omnibus statistic, p-value, significance decision, or
claim was used to choose the architecture. The pilot record is reported below;
this render itself fits no model and starts no Stage 2 analysis.
:::

## Scope and safeguards

The measurement-methods audit was used only to trace the signed record, V0,
normalized diary, approved one-hour metric, preparation evidence, and H02
temporal specification. The Quarto-authoring workflow was used to make the
contract executable and renderable. All scientific checks below run in R
4.6.1. No H02 fitted value, prediction, significance result, or claim is used.

The candidate counts are not Stage 2 fitted-model counts. Stage 2 must report
the exact rows retained by every fitted model, even if they happen to equal
these counts. This revision records the author's approved process; Stage 2 has
not been started in this document.

## Preregistered contract

The signed AsPredicted record states exactly:

> **H3: Hourly self-reported light exposure categories predict hourly geometric mean melanopic EDI.**

— `preregistration/AsPredicted #273407.pdf`, PDF p. 1.

The signed model on PDF p. 2 is:

```text
melEDI ~ Predictor + (Predictor | Site) + (1 | Site:Participant)
```

For H03, `Predictor` is the primary light-source category. The developed
analysis plan restates the question as a relationship between hourly
self-reported categories and hourly geometric melEDI and gives
`melEDI ~ mH-LEA + (mH-LEA | Site) + (1 | Site:Participant)`. It warns that a
random site effect may be unstable, but it does not authorize deleting
categories, changing the primary test, or changing multiplicity families.

### Evidence identities

```{r}
#| label: tbl-evidence-identities
#| tbl-cap: "Controlling H03 evidence verified at render time."

evidence_paths <- c(
  signed_preregistration = "preregistration/AsPredicted #273407.pdf",
  developed_analysis_plan = "preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx",
  preregistration_contract = "audit/evidence/preregistration_contract.md",
  migration_map = "audit/hypotheses/H01-H04_migration_map.md",
  category_support_gate = "audit/hypotheses/H03_H04_category_support_gate.md",
  h02_selected_temporal_specification = "artifacts/07_models/H02/selected_temporal_model_specification.csv",
  v0_near_eye_source = "RQ2.qmd",
  v0_chest_source = "RQ2_chest.qmd"
)

evidence_manifest <- tibble(
  evidence = names(evidence_paths),
  path = unname(evidence_paths),
  sha256 = vapply(
    file.path(project_root, unname(evidence_paths)),
    digest::digest,
    character(1),
    algo = "sha256",
    file = TRUE
  )
)

stage1_gt(evidence_manifest) |>
  cols_width(
    evidence ~ pct(24),
    path ~ pct(42),
    sha256 ~ pct(34)
  )
```

## What V0 implemented

V0 means the submitted H03 implementation in `RQ2.qmd` and
`RQ2_chest.qmd` and its frozen rendered outputs. It is a comparison target,
not an approved analysis.

### Exact V0 formulas — reconstructed, not fitted

```{r}
#| label: tbl-v0-formulas
#| tbl-cap: "Evaluated V0 Wilkinson formulas. No model is fitted."

v0_formulas <- list(
  registered_exact = melEDI ~ light_source +
    (light_source | site) + (1 | site:participant),
  v0_full = geo.MEDI ~ site * lightsource_primary + (1 | Id),
  v0_site_only = geo.MEDI ~ site + (1 | Id),
  v0_additive = geo.MEDI ~ site + lightsource_primary + (1 | Id),
  v0_context_only = geo.MEDI ~ lightsource_primary + (1 | Id)
)

v0_formula_contract <- enframe(
  v0_formulas,
  name = "model_id",
  value = "formula"
) |>
  mutate(formula = vapply(formula, formula_text, character(1)))

stage1_gt(v0_formula_contract) |>
  cols_width(model_id ~ pct(23), formula ~ pct(77))
```

V0 floor-aggregated one-minute data to the hour, calculated arithmetic melEDI
and a zero-aware geometric `geo.MEDI`, removed whole participant-days whose
arithmetic melEDI was constant, and joined the diary within site, participant,
and date. It did not require 30 valid minutes or persist a join-cardinality
check. It fitted the four formulas above with `glmmTMB`, Tweedie log link,
maximum likelihood, and sum coding for site except in the context-only fit.

The stated confirmatory comparison was `v0_site_only` versus `v0_full`.
That comparison adds the category block and all category-by-site interactions
together; it is not the category-association test alone. The V0 additive-full
comparison tested interaction, and the context-only/additive comparison tested
site.

### V0 category, reporting, and temporal reconstruction

| Component | V0 implementation | Audit consequence |
|---|---|---|
| Categories | Set site/category cells with fewer than 20 rows to missing, then categories with fewer than 200 remaining rows to missing; manually removed external light during sleep; outdoor electric then fell out near eye and was manually removed from chest “for consistency” | Only five of seven registered categories entered the model; the rule was outcome- and placement-dependent |
| Reported sample | 16,774 near-eye and 18,391 chest participant-hours in the frozen H03 tables | These are V0 row counts, not independent participants; they must not be substituted for current fitted samples |
| Pooled contrasts | Four `trt.vs.ctrl` contrasts from `emmeans` | The code uses the Dunnett-X default, while the table calls the values FDR-adjusted |
| Site contrasts | Effect contrasts adjusted within each category-specific `by` group | The table instead describes one `n = 9` site family; no cross-category family was declared |
| Intervals | V0 removed asymptotic confidence-limit columns before display | Consequential V0 estimates were presented without 95% intervals |
| Repetition | Participant intercept only | Participant-day clustering and remaining within-day dependence were omitted |
| Contribution claims | Differences between overlapping marginal-R² values | These are not additive proportions of variance explained |
| Chest prose | The near-eye numeric interpretation was copied verbatim into the chest notebook | Chest claims are not a valid interpretation of the chest fit |
| Temporal outcome | Hourly arithmetic melEDI transformed with the zero-aware log | It does not match the registered hourly geometric outcome |
| Temporal structure | Non-cyclic overall smooth, one AR start per participant, fREML AIC comparison, pointwise highlighted segments | It does not satisfy the accepted H02 temporal contract |

The source evidence is `RQ2.qmd:104–471, 535–640` and the corresponding chest
blocks. The migration reconciliation is
`audit/hypotheses/H01-H04_migration_map.md:775–943`. The registered deviations
are DEV-013, DEV-014, and DEV-022 through DEV-024.

## Outcome-blind category and join audit

This section fixes the category rules before displaying outcome-side candidate
support. It uses the normalized diary only.

```{r}
#| label: h03-diary-audit

diary_path <- file.path(
  project_root,
  "artifacts/06_model_data/normalized_inputs/lightexposurediary.rds"
)
diary_raw <- readRDS(diary_path)

source_flag_names <- c(
  "light_electric_indoor",
  "light_electric_outdoor",
  "light_daylight_indoor",
  "light_daylight_outdoor",
  "light_display",
  "light_sleep_darkness",
  "light_sleep_imission"
)
source_flag_for_category <- setNames(source_flag_names, category_levels)

diary <- diary_raw |>
  mutate(
    site = factor(site, levels = site_levels),
    local_date = as.Date(Date),
    clock_minute = as.integer(format(interval_start_wall, "%H", tz = "UTC")) * 60L +
      as.integer(format(interval_start_wall, "%M", tz = "UTC")),
    light_source = factor(lightsource_primary, levels = category_levels),
    light_source_code = unname(category_codes[as.character(light_source)]),
    participant = interaction(site, Id, drop = TRUE, lex.order = TRUE),
    participant_day = interaction(site, Id, local_date, drop = TRUE, lex.order = TRUE),
    source_flag_n = rowSums(
      across(all_of(source_flag_names), ~ as.integer(.x %in% TRUE)),
      na.rm = TRUE
    )
  ) |>
  rowwise() |>
  mutate(
    primary_flag_true = if (is.na(light_source)) {
      NA
    } else {
      isTRUE(get(source_flag_for_category[[as.character(light_source)]]))
    }
  ) |>
  ungroup()

diary_eligible <- diary |>
  filter(interval_analysis_eligible %in% TRUE, !(interval_quarantined %in% TRUE))

key_columns <- c("site", "Id", "local_date", "clock_minute")

diary_key_audit <- diary_eligible |>
  count(across(all_of(key_columns)), name = "rows_per_key") |>
  summarise(
    Eligible_hours = sum(rows_per_key),
    Distinct_local_hour_keys = n(),
    Duplicate_keys = sum(rows_per_key > 1L),
    Duplicate_rows_beyond_first = sum(pmax(rows_per_key - 1L, 0L)),
    Non_hour_aligned_starts = sum(clock_minute %% 60L != 0L)
  )

diary_category_audit <- diary_eligible |>
  summarise(
    Eligible_hours = n(),
    Hours_with_primary_category = sum(!is.na(light_source)),
    Hours_missing_primary_category = sum(is.na(light_source)),
    Hours_with_no_source_flag = sum(source_flag_n == 0L),
    Hours_with_one_source_flag = sum(source_flag_n == 1L),
    Hours_with_multiple_source_flags = sum(source_flag_n > 1L),
    Recorded_primary_without_expected_flag = sum(primary_flag_true %in% FALSE, na.rm = TRUE)
  )

diary_overall_support <- diary_eligible |>
  filter(!is.na(light_source)) |>
  group_by(light_source) |>
  summarise(
    Hours = n(),
    Participants = n_distinct(participant),
    Participant_days = n_distinct(participant_day),
    Sites = n_distinct(site),
    .groups = "drop"
  ) |>
  complete(
    light_source = factor(category_levels, levels = category_levels),
    fill = list(Hours = 0L, Participants = 0L, Participant_days = 0L, Sites = 0L)
  ) |>
  mutate(
    Category = unname(category_codes[as.character(light_source)]),
    Pooled_rule = Hours >= 200L & Participants >= 20L & Sites >= 3L
  ) |>
  select(Category, Light_source = light_source, Hours, Participants, Participant_days, Sites, Pooled_rule)

stopifnot(
  diary_key_audit$Duplicate_keys == 0L,
  diary_key_audit$Non_hour_aligned_starts == 0L,
  diary_category_audit$Recorded_primary_without_expected_flag == 0L,
  all(diary_overall_support$Pooled_rule)
)
```

```{r}
#| label: tbl-diary-cardinality
#| tbl-cap: "Outcome-blind diary cardinality and source-flag encoding."

bind_rows(
  diary_key_audit |> pivot_longer(everything(), names_to = "Check", values_to = "Count"),
  diary_category_audit |> pivot_longer(everything(), names_to = "Check", values_to = "Count")
) |>
  distinct(Check, .keep_all = TRUE) |>
  stage1_gt() |>
  fmt_integer(Count, use_seps = TRUE) |>
  cols_width(Check ~ pct(70), Count ~ pct(30))
```

The primary category is a single seven-level factor and is mutually exclusive
within an hour. The underlying source flags are not mutually exclusive:
11,408 eligible hours contain more than one true source flag. That overlap is
real questionnaire context, not duplicated outcome rows. The model uses the
single declared primary category; it does not pivot the overlapping flags or
count an hour more than once. Every non-missing primary category has its
corresponding flag set to true.

```{r}
#| label: tbl-diary-category-support
#| tbl-cap: "Outcome-blind support for every prespecified primary-light-source category."

diary_overall_support |>
  stage1_gt() |>
  fmt_integer(c(Hours, Participants, Participant_days, Sites), use_seps = TRUE) |>
  fmt_missing(everything(), missing_text = "—") |>
  cols_label(Pooled_rule = "Pass pooled rule") |>
  cols_width(
    Category ~ pct(15),
    Light_source ~ pct(37),
    everything() ~ pct(9)
  )
```

The two sleep categories are bedside sleep-environment contexts. They are not
worn ocular exposure and will never be described that way.

### Approved estimability rule

1. Keep all seven categories in the dictionary, model frame, sample flow, and
   pooled additive model. Set indoor electric light as the reference before
   fitting.
2. Permit a pooled category contrast only if its fitted frame has at least 200
   hours, 20 participants, and three sites, remains connected to the reference,
   and the additive design is full rank.
3. Permit a site-specific contrast only if the target and indoor-electric
   reference each have at least 20 hours and five participants at that site,
   at least five participants report both categories, and the relevant design
   is estimable.
4. Keep unsupported categories visible. Mark only the affected contrast
   `support-non-estimable`, with the failed support components; do not delete
   the category after inspecting melEDI.
5. Reapply these rules to every placement and sensitivity frame. A contrast
   may therefore be estimable in the main frame but support-non-estimable in a
   smaller paired or deletion frame.

The normalized diary alone gives 57 supported category-by-site cells out of
63. These thresholds are fixed before any association estimate is inspected.

## Primary outcome and candidate model frames

The proposed outcome is the accepted one-hour zero-aware geometric mean
melEDI. An hour is available only with at least 30 valid one-minute melEDI
values. The within-hour zero offset is 0.1 lx; missing observations are never
treated as zero. The model preserves the local wall-clock hour, valid-minute
support, state and measurement-context composition, diary interval, true-UTC
interval start, participant-day, and placement.

The current metric artifact exposes a unique local hour key rather than a
true-UTC outcome key. The normalized diary has unique true-UTC and local-wall
hour keys, so the audited join is one outcome hour to at most one eligible
diary hour on site, participant, local date, and local clock minute. Stage 2
will retain the diary's true-UTC interval start after that asserted join.

```{r}
#| label: h03-primary-candidate-audit

primary_paths <- c(
  Near_eye = "artifacts/06_model_data/base/metrics_glasses_one_hour_context.rds",
  Chest = "artifacts/06_model_data/base/metrics_chest_one_hour_context.rds"
)

prepare_primary <- function(relative_path, placement) {
  raw <- readRDS(file.path(project_root, relative_path))
  key_counts <- raw |>
    count(across(all_of(key_columns)), name = "rows_per_key")
  stopifnot(!any(key_counts$rows_per_key > 1L))

  joined <- raw |>
    mutate(
      site = factor(site, levels = site_levels),
      local_date = as.Date(local_date),
      participant = interaction(site, Id, drop = TRUE, lex.order = TRUE),
      participant_day = interaction(site, Id, local_date, drop = TRUE, lex.order = TRUE)
    ) |>
    left_join(
      diary_eligible |>
        transmute(
          across(all_of(key_columns)),
          light_source,
          light_source_code,
          interval_start_utc,
          source_flag_n,
          diary_key_matched = TRUE
        ),
      by = key_columns,
      relationship = "one-to-one"
    ) |>
    mutate(
      placement = placement,
      outcome_available = bin_admissible %in% TRUE & is.finite(metric_value_lx),
      diary_key_matched = diary_key_matched %in% TRUE,
      category_available = !is.na(light_source),
      model_candidate = outcome_available & category_available
    )

  list(raw = raw, joined = joined, candidate = filter(joined, model_candidate))
}

placements <- imap(primary_paths, ~ prepare_primary(.x, gsub("_", "-", .y)))

sample_summary <- map_dfr(placements, function(x) {
  x$joined |>
    summarise(
      Placement = first(placement),
      Grid_hours = n(),
      Outcome_available = sum(outcome_available),
      Insufficient_metric_support = sum(!outcome_available),
      No_eligible_diary_key = sum(outcome_available & !diary_key_matched),
      Diary_key_but_primary_missing = sum(
        outcome_available & diary_key_matched & !category_available
      ),
      Candidate_hours = sum(model_candidate),
      Candidate_participants = n_distinct(participant[model_candidate]),
      Candidate_participant_days = n_distinct(participant_day[model_candidate]),
      Candidate_sites = n_distinct(site[model_candidate]),
      Exact_zero_hours = sum(model_candidate & metric_value_lx == 0)
    )
})

metric_support_audit <- map_dfr(placements, function(x) {
  x$joined |>
    summarise(
      Placement = first(placement),
      Metric_ids = n_distinct(metric),
      Metric = paste(unique(metric), collapse = "; "),
      Zero_offsets = n_distinct(zero_offset),
      Zero_offset_lx = paste(unique(zero_offset), collapse = "; "),
      Admissible_below_30_minutes = sum(
        outcome_available & valid_medi_wall_minutes < 30L,
        na.rm = TRUE
      ),
      Unavailable_at_or_above_30_minutes = sum(
        !outcome_available & valid_medi_wall_minutes >= 30L,
        na.rm = TRUE
      ),
      Candidate_state_boundary_hours = sum(
        model_candidate & spans_diary_state_boundary,
        na.rm = TRUE
      ),
      Candidate_context_boundary_hours = sum(
        model_candidate & spans_measurement_context_boundary,
        na.rm = TRUE
      ),
      Candidate_fall_back_hours = sum(
        model_candidate & dst_fold_wall_minutes > 0,
        na.rm = TRUE
      )
    )
})

sequence_audit <- map_dfr(placements, function(x) {
  sequence_rows <- x$candidate |>
    arrange(participant_day, clock_minute) |>
    group_by(participant_day) |>
    mutate(
      fold_hour = dst_fold_wall_minutes > 0,
      gap_from_previous = row_number() > 1L &
        (clock_minute - lag(clock_minute, default = first(clock_minute))) != 60L,
      reset_sequence = row_number() == 1L | gap_from_previous | fold_hour |
        lag(fold_hour, default = FALSE),
      ar_sequence_index = cumsum(reset_sequence)
    ) |>
    ungroup() |>
    mutate(
      ar_sequence = interaction(
        participant_day,
        ar_sequence_index,
        drop = TRUE,
        lex.order = TRUE
      )
    )

  lengths <- sequence_rows |>
    count(placement, ar_sequence, name = "sequence_hours")

  sequence_rows |>
    summarise(
      Placement = first(placement),
      Candidate_hours = n(),
      Participant_days = n_distinct(participant_day),
      AR_sequences = n_distinct(ar_sequence),
      Gap_resets = sum(gap_from_previous),
      Fall_back_hours = sum(fold_hour),
      Singleton_sequences = sum(lengths$sequence_hours == 1L),
      Median_sequence_hours = stats::median(lengths$sequence_hours),
      Maximum_sequence_hours = max(lengths$sequence_hours)
    )
})

stopifnot(
  all(metric_support_audit$Metric_ids == 1L),
  all(metric_support_audit$Zero_offsets == 1L),
  all(metric_support_audit$Zero_offset_lx == "0.1"),
  all(metric_support_audit$Admissible_below_30_minutes == 0L),
  all(metric_support_audit$Unavailable_at_or_above_30_minutes == 0L)
)
```

```{r}
#| label: tbl-primary-candidate-flow
#| tbl-cap: "Current H03 candidate flow before any model is fitted."

sample_summary |>
  stage1_gt() |>
  fmt_integer(-Placement, use_seps = TRUE) |>
  cols_width(Placement ~ pct(12), everything() ~ pct(9))
```

The 4,977 near-eye and 5,409 chest exact-zero hours confirm that a model for
strictly positive outcomes would be inadmissible. These are overall
distributional-support counts, not category results.

```{r}
#| label: tbl-metric-support
#| tbl-cap: "One-hour metric, support, boundary, and fall-back audit."

metric_support_audit |>
  stage1_gt() |>
  fmt_integer(
    c(
      Metric_ids,
      Zero_offsets,
      Admissible_below_30_minutes,
      Unavailable_at_or_above_30_minutes,
      Candidate_state_boundary_hours,
      Candidate_context_boundary_hours,
      Candidate_fall_back_hours
    ),
    use_seps = TRUE
  ) |>
  cols_width(Placement ~ pct(10), Metric ~ pct(25), everything() ~ pct(8))
```

Boundary-spanning hours remain in the primary candidate because the approved
metric preserves their state and measurement-context composition. Stage 2
will report them, inspect their influence, and repeat the analysis after
excluding boundary-spanning hours as a named sensitivity. The sleep categories
remain bedside-environment contexts in both frames.

```{r}
#| label: tbl-ar-sequences
#| tbl-cap: "Precomputed regular within-day sequence boundaries for the exploratory temporal AR(1) term."

sequence_audit |>
  stage1_gt() |>
  fmt_integer(-c(Placement, Median_sequence_hours), use_seps = TRUE) |>
  fmt_number(Median_sequence_hours, decimals = 1) |>
  cols_width(Placement ~ pct(14), everything() ~ pct(11))
```

Every sequence starts at a participant-day boundary, after an omitted or
non-consecutive hour, at a fall-back hour, and immediately after a fall-back
hour. Unsupported hours therefore break the sequence. Category changes do not
break it. These sequences belong to the exploratory GAM/GAMM only; the primary
quasi-Tweedie covariance clusters the complete record by participant and does
not impose an AR(1) correlation shape.

### Category and category-by-site candidate support

```{r}
#| label: h03-category-cell-audit

candidate_category_support <- map_dfr(placements, function(x) {
  x$candidate |>
    group_by(placement, light_source) |>
    summarise(
      Hours = n(),
      Participants = n_distinct(participant),
      Participant_days = n_distinct(participant_day),
      Sites = n_distinct(site),
      .groups = "drop"
    ) |>
    complete(
      placement,
      light_source = factor(category_levels, levels = category_levels),
      fill = list(Hours = 0L, Participants = 0L, Participant_days = 0L, Sites = 0L)
    ) |>
    mutate(
      Category = unname(category_codes[as.character(light_source)]),
      Pooled_rule = Hours >= 200L & Participants >= 20L & Sites >= 3L
    ) |>
    select(
      Placement = placement,
      Category,
      Light_source = light_source,
      Hours,
      Participants,
      Participant_days,
      Sites,
      Pooled_rule
    )
})

candidate_cell_support <- map_dfr(placements, function(x) {
  dat <- x$candidate
  observed_sites <- unique(as.character(dat$site))

  cell <- dat |>
    group_by(placement, site, light_source) |>
    summarise(
      Hours = n(),
      Participants = n_distinct(participant),
      Participant_days = n_distinct(participant_day),
      .groups = "drop"
    ) |>
    complete(
      placement,
      site = factor(site_levels, levels = site_levels),
      light_source = factor(category_levels, levels = category_levels),
      fill = list(Hours = 0L, Participants = 0L, Participant_days = 0L)
    )

  reference_participants <- dat |>
    filter(light_source == reference_label) |>
    distinct(placement, site, participant) |>
    group_by(placement, site) |>
    summarise(reference_participants = list(as.character(participant)), .groups = "drop")

  target_participants <- dat |>
    distinct(placement, site, light_source, participant) |>
    group_by(placement, site, light_source) |>
    summarise(target_participants = list(as.character(participant)), .groups = "drop")

  reference_counts <- cell |>
    filter(light_source == reference_label) |>
    transmute(
      placement,
      site,
      Reference_hours = Hours,
      Reference_participants = Participants
    )

  cell |>
    left_join(reference_counts, by = c("placement", "site")) |>
    left_join(reference_participants, by = c("placement", "site")) |>
    left_join(target_participants, by = c("placement", "site", "light_source")) |>
    mutate(
      target_participants = map(
        target_participants,
        ~ if (is.null(.x)) character() else .x
      ),
      reference_participants = map(
        reference_participants,
        ~ if (is.null(.x)) character() else .x
      ),
      Shared_participants = map2_int(
        target_participants,
        reference_participants,
        ~ length(intersect(.x, .y))
      ),
      Site_present = as.character(site) %in% observed_sites,
      Cell_rule = Site_present & Hours >= 20L & Participants >= 5L &
        Reference_hours >= 20L & Reference_participants >= 5L &
        Shared_participants >= 5L,
      Category = unname(category_codes[as.character(light_source)])
    ) |>
    select(
      Placement = placement,
      site,
      Category,
      Light_source = light_source,
      Hours,
      Participants,
      Participant_days,
      Reference_hours,
      Reference_participants,
      Shared_participants,
      Site_present,
      Cell_rule
    )
})

rank_audit <- map_dfr(placements, function(x) {
  dat <- x$candidate |>
    mutate(
      site = droplevels(factor(site, levels = site_levels)),
      light_source = factor(light_source, levels = category_levels),
      site_source_cell = interaction(site, light_source, drop = TRUE, lex.order = TRUE)
    )

  standard_designs <- map_dfr(
    list(
      additive = ~ site + light_source,
      literal_interaction = ~ site * light_source,
      observed_cell_means = ~ 0 + site_source_cell
    ),
    function(formula) {
      mm <- model.matrix(formula, data = dat)
      tibble(
        Placement = unique(dat$placement),
        Design = formula_text(formula),
        Columns = ncol(mm),
        Rank = qr(mm)$rank,
        Full_rank = qr(mm)$rank == ncol(mm)
      )
    },
    .id = "Design_id"
  )

  core_dat <- dat |>
    filter(as.character(light_source) %in% core_heterogeneity_levels) |>
    mutate(
      light_source_core = factor(
        as.character(light_source),
        levels = core_heterogeneity_levels
      )
    )
  core_formula <- ~ site * light_source_core
  core_matrix <- model.matrix(core_formula, data = core_dat)

  bind_rows(
    standard_designs,
    tibble(
      Placement = unique(core_dat$placement),
      Design_id = "core_literal_interaction",
      Design = formula_text(core_formula),
      Columns = ncol(core_matrix),
      Rank = qr(core_matrix)$rank,
      Full_rank = qr(core_matrix)$rank == ncol(core_matrix)
    )
  )
})

stopifnot(
  all(candidate_category_support$Pooled_rule),
  all(rank_audit$Full_rank[rank_audit$Placement == "Near-eye"]),
  all(
    rank_audit$Full_rank[
      rank_audit$Placement == "Chest" & rank_audit$Design_id != "literal_interaction"
    ]
  ),
  with(
    filter(rank_audit, Placement == "Chest", Design_id == "literal_interaction"),
    Columns - Rank == 1L && !Full_rank
  )
)
```

```{r}
#| label: tbl-candidate-category-support
#| tbl-cap: "All-available candidate support by placement and category; no association is estimated."

candidate_category_support |>
  mutate(Light_source = unname(category_short[Category])) |>
  stage1_gt(groupname_col = "Placement") |>
  fmt_integer(c(Hours, Participants, Participant_days, Sites), use_seps = TRUE) |>
  cols_label(Light_source = "Primary light source", Pooled_rule = "Pass pooled rule") |>
  cols_width(
    Category ~ pct(18),
    Light_source ~ pct(26),
    everything() ~ pct(11)
  )
```

All seven categories pass the pooled rule in both all-available candidates.
Outdoor electric light is close to the 200-hour boundary near eye (203 hours),
so sparse-cell and influence checks are consequential.

```{r}
#| label: tbl-design-rank
#| tbl-cap: "Fixed-effect rank before fitting."

rank_audit |>
  stage1_gt(groupname_col = "Placement") |>
  cols_width(Design_id ~ pct(22), Design ~ pct(43), everything() ~ pct(12))
```

The near-eye literal seven-category interaction is rank 63 of 63. Chest has
eight observed sites and no external-light-during-sleep candidate hour at
FUSPCEU; its literal interaction is rank 55 of 56. The precomputed
`site_source_cell` factor contains only observed cells and is rank 55 of 55.
The approved chest heterogeneity model therefore uses that observed-cell
parameterization rather than allowing the fitting engine to silently drop an
aliased coefficient. The predeclared five-category fallback designs are also
full rank: 45 of 45 near eye and 40 of 40 chest.

```{r}
#| label: tbl-unsupported-cells
#| tbl-cap: "Category-by-site cells that fail the approved rule in the current candidates."

unsupported_cells <- candidate_cell_support |>
  filter(!Cell_rule) |>
  left_join(
    site_registry |>
      transmute(site = factor(site, levels = site_levels), Site = display_name),
    by = "site"
  ) |>
  mutate(
    Primary_light_source = unname(category_short[Category]),
    Status = case_when(
      !Site_present ~ "Placement not observed",
      Hours == 0L ~ "No candidate hours",
      TRUE ~ "Support-non-estimable"
    )
  ) |>
  select(
    Placement,
    Site,
    Primary_light_source,
    Hours,
    Participants,
    Participant_days,
    Reference_hours,
    Reference_participants,
    Shared_participants,
    Status
  )

stage1_gt(unsupported_cells, groupname_col = "Placement") |>
  fmt_integer(
    c(
      Hours,
      Participants,
      Participant_days,
      Reference_hours,
      Reference_participants,
      Shared_participants
    ),
    use_seps = TRUE
  ) |>
  cols_width(
    Site ~ pct(18),
    Primary_light_source ~ pct(22),
    Status ~ pct(18),
    everything() ~ pct(8)
  )
```

Under the approved rule, 53 of 63 near-eye cells are supported. Chest has 48
supported cells among the 56 site/category combinations at its eight observed
sites; seven MPI combinations are not applicable, and eight observed-site
combinations are support-non-estimable (including the zero-hour FUSPCEU
external-sleep cell).

```{r}
#| label: tbl-core-heterogeneity-support
#| tbl-cap: "Outcome-blind site support for the predeclared core-category heterogeneity fallback."

core_heterogeneity_support <- candidate_cell_support |>
  filter(
    Site_present,
    as.character(Light_source) %in% core_heterogeneity_levels
  ) |>
  group_by(Placement, Light_source) |>
  summarise(
    Observed_sites = n(),
    Supported_sites = sum(Cell_rule),
    .groups = "drop"
  ) |>
  mutate(
    Primary_light_source = unname(category_short[
      unname(category_codes[as.character(Light_source)])
    ]),
    Meets_core_rule = Supported_sites >= Observed_sites - 1L
  ) |>
  select(
    Placement,
    Primary_light_source,
    Observed_sites,
    Supported_sites,
    Meets_core_rule
  )

stopifnot(all(core_heterogeneity_support$Meets_core_rule))

stage1_gt(core_heterogeneity_support, groupname_col = "Placement") |>
  fmt_integer(c(Observed_sites, Supported_sites), use_seps = TRUE) |>
  cols_width(Primary_light_source ~ pct(40), everything() ~ pct(20))
```

The complete cell counts are retained below so that no category disappears
behind the unsupported-cell summary.

```{r}
#| label: tbl-complete-cell-support
#| tbl-cap: "Complete candidate category-by-site hours, participants, participant-days, and shared-reference support."

complete_cell_display <- candidate_cell_support |>
  left_join(
    site_registry |>
      transmute(site = factor(site, levels = site_levels), Site = display_name),
    by = "site"
  ) |>
  mutate(
    Primary_light_source = unname(category_short[Category]),
    Support = case_when(
      !Site_present ~ "Not applicable",
      Cell_rule ~ "Supported",
      TRUE ~ "Support-non-estimable"
    )
  ) |>
  select(
    Placement,
    Site,
    Primary_light_source,
    Hours,
    Participants,
    Participant_days,
    Shared_participants,
    Support
  )

stage1_gt(complete_cell_display, groupname_col = "Placement") |>
  fmt_integer(c(Hours, Participants, Participant_days, Shared_participants), use_seps = TRUE) |>
  cols_width(
    Site ~ pct(20),
    Primary_light_source ~ pct(25),
    Support ~ pct(20),
    everything() ~ pct(9)
  )
```

## Estimands and approved confirmatory model

### Two-reference reporting contract

Let $\eta_{sc}=\log(\mu_{sc})$ be the fitted log mean and $\mu_{sc}$ the
model-estimated expected zero-inclusive one-hour melEDI for site $s$ and
light-source category $c$. The report will use
**site-standardized** for an estimate standardized to a prespecified reference
distribution that assigns each included site weight $1/S$. This is not a
fictional “average site” and is not an observation-frequency-weighted average.
Here $S=9$ for near eye and $S=8$ for chest.

Author amendment, 2026-08-07: equal-site standardization is performed on the
fitted link scale and back-transformed once. It is therefore the geometric
mean of the fitted site-specific expected values. This supersedes the earlier
response-scale arithmetic averaging rule.

The indoor-electric category is the substantive reference throughout. Define

$$
\widetilde\mu_c^{\mathrm{site}}
= \exp\!\left(\frac{1}{S}\sum_{s=1}^{S}\eta_{sc}\right)
= \left(\prod_{s=1}^{S}\mu_{sc}\right)^{1/S},\qquad
A = \widetilde\mu_{\mathrm{indoor}}^{\mathrm{site}},\qquad
R_c = \frac{\widetilde\mu_c^{\mathrm{site}}}{A},\qquad
D_{sc} = \frac{\mu_{sc}}{\widetilde\mu_c^{\mathrm{site}}}.
$$

Thus every supported site/category mean has the V0-recognizable but corrected
factorization $\mu_{sc}=A\times R_c\times D_{sc}$: the absolute anchor is the
site-standardized indoor-electric mean, $R_c$ compares each category with that
indoor reference, and $D_{sc}$ compares a site/category cell with the
site-standardized mean for the same category. The output will show the derived
cell mean as well as these interpretable multipliers; readers will not have to
reconstruct values by multiplication. Raw coding coefficients will not be
presented as the scientific estimates.

| Estimand | Exact definition | Inferential role |
|---|---|---|
| Primary category omnibus | Participant-cluster-robust joint Wald $F$ test that the six non-reference category coefficients in the additive near-eye mean model are zero | H03-F1; category association adjusted for fixed site, without an interaction |
| Indoor anchor | $A$, the site-standardized expected one-hour melEDI under indoor electric light | Absolute reference in melEDI lx with 95% CI |
| Category contrast | $R_c$, each site-standardized category mean divided by the site-standardized indoor-electric mean | Six planned ratios in H03-F2, with 95% CIs |
| Site heterogeneity omnibus | Participant-cluster-robust joint Wald $F$ test of all estimable site-by-category interaction restrictions | H03-F3; separate from the primary category omnibus |
| Site/category mean | $\mu_{sc}$ for every supported cell | Report in melEDI lx with 95% CI; unsupported cells remain labelled |
| Site deviation | $D_{sc}$, a site/category mean divided by that category's site-standardized mean | Supported deviations form H03-F4 |
| Within-site category ratio | $\mu_{sc}/\mu_{s,\mathrm{indoor}}$ | Additional practical contrast, explicitly labelled and kept separate from $R_c$ |

The primary analysis uses near-eye data. Chest is complementary and is never
pooled with near-eye data. Site names, order, and colours follow the
submitted-manuscript registry. All category estimates use indoor electric as
reference; all cross-site estimates use the site-standardized reference above.

### Exact model formulas — declared, not fitted in this render

```{r}
#| label: tbl-approved-primary-formulas
#| tbl-cap: "Evaluated approved, companion, fallback, secondary, and rejected-pilot Wilkinson formulas. No model is fitted in this render."

approved_primary_formulas <- list(
  primary_population_mean = geo_medi_1h ~ site + light_source,
  full_site_heterogeneity = geo_medi_1h ~ site * light_source,
  chest_observed_cell_heterogeneity = geo_medi_1h ~ 0 + site_source_cell,
  fallback_core_site_heterogeneity = geo_medi_1h ~ site * light_source_core,
  secondary_mundlak_audit = geo_medi_1h ~
    site + light_source +
    between_source_1 + between_source_2 + between_source_3 +
    between_source_4 + between_source_5 + between_source_6,
  rejected_mixed_dependence_pilot = geo_medi_1h ~
    site + light_source +
    (1 | participant) + (1 | participant_day) +
    ar1(ar_step_factor + 0 | ar_sequence)
)

approved_formula_contract <- enframe(
  approved_primary_formulas,
  name = "model_id",
  value = "formula"
) |>
  mutate(formula = vapply(formula, formula_text, character(1)))

stage1_gt(approved_formula_contract) |>
  cols_width(model_id ~ pct(29), formula ~ pct(71))
```

The primary model is `primary_population_mean`. The full near-eye
heterogeneity companion has 48 interaction degrees of freedom. Because chest
has one structurally absent site/category cell, its full companion uses the
rank-complete 55-cell parameterization and a 41-degree-of-freedom robust
linear-hypothesis test of departures from additivity.

The full interaction receives a results-blind architecture gate before any
interaction estimate or p-value is inspected. It must converge, retain the
declared rank, yield a finite positive covariance for the tested restrictions,
have tolerable condition numbers and cluster leverage, and produce finite
intervals for supported cells. If it fails, the predeclared fallback is a
**core-category heterogeneity** model with indoor electric, indoor daylight,
outdoor daylight, emissive display, and sleep darkness. These five categories
met the outcome-blind site-support rule at all or all but one observed site;
the fallback reduces the near-eye interaction block from 48 to 32 degrees of
freedom and the chest block to 28. Outdoor electric and external light during
sleep remain in the seven-category primary additive model, pooled contrasts,
sample flow, and descriptive displays. They are not collapsed because neither
has a construct-equivalent category. A fallback heterogeneity result therefore
answers a narrower five-category question and cannot be described as the
seven-category heterogeneity test.

The `secondary_mundlak_audit` adds each participant's proportions in the six
non-reference categories. It separates within-participant category contrasts
from between-participant reporting composition and is secondary because its
overall covariance was much more ill-conditioned in the architecture pilot.
It cannot replace the population-averaged primary result.

### Mean, covariance, tests, and intervals

- **Mean model:** `stats::glm()` with
  `statmod::tweedie(var.power = 1.539919, link.power = 0)`. The log link models
  the expected zero-inclusive one-hour melEDI. The variance power was frozen
  from the authorized architecture pilot before inferential extraction and is
  used unchanged for near eye, chest, and declared comparison frames.
- **Dependence and covariance:** uncertainty uses
  `sandwich::vcovCL(..., cluster = participant, type = "HC1", cadjust = TRUE,
  fix = FALSE)`. A site-prefixed participant is the independent cluster; all
  of that participant's hours and days may be dependent without imposing
  random-effect or AR(1) variance components.
- **Finite-cluster reference:** a one-parameter contrast uses a $t$
  distribution with $G-1$ degrees of freedom. A $q$-restriction omnibus uses
  $F=W/q$ with $(q,G-1)$ degrees of freedom, where $W$ is the sandwich Wald
  quadratic form and $G$ is the number of fitted participants. Stage 2 will
  store the exact restriction matrix used by every omnibus.
- **Coding:** `light_source` is treatment-coded with indoor electric first;
  site is sum-coded internally. The reported site-standardized means and
  contrasts are explicit equal-site functions of the complete prediction
  grid, so they do not depend on interpreting an arbitrary site coefficient.
- **Standardization:** predict every site/category combination on the fitted
  log-mean scale, average the site linear predictors with weight $1/S$, then
  apply the inverse link once before forming ratios. Automatic
  `type = "response"` labels are not trusted for the power-link family; the
  log inverse link is applied and tested explicitly.
- **Intervals:** two-sided 95% cluster-robust intervals use the full robust
  covariance and the delta method on log means/log ratios before transformation
  to melEDI lx or ratios. A missing, non-positive-definite, or unstable
  consequential covariance is reported as non-estimable rather than repaired.
- **Unavailable likelihood summaries:** quasi likelihood supplies no valid
  log likelihood, likelihood-ratio test, AIC, mixed-model random-effect SD,
  ICC, marginal/conditional mixed-model $R^2$, or likelihood-based variance
  decomposition. None will be manufactured for the primary analysis. The
  exploratory GAM/GAMM section below has its own explicitly defined fit and
  variance summaries.

### Bounded architecture-pilot record

The following values are numerical construction diagnostics from the author's
authorized temporary pilots. No association estimate or inferential decision
was inspected to select the architecture.

```{r}
#| label: tbl-architecture-pilots
#| tbl-cap: "Bounded near-eye model-construction pilots; diagnostic values only."

architecture_pilots <- tribble(
  ~Model, ~Numerical_status, ~Dependence_or_covariance, ~Diagnostic_signal, ~Disposition,
  "Tweedie mixed model with participant, participant-day, and AR(1)",
  "Converged; optimizer code 0; positive-definite Hessian",
  "Participant SD 0.418; participant-day SD 0.000272; AR SD 2.534; rho 0.835",
  "Singular; zero ratio 1.073 and residual-uniformity/zero tests p < 0.001",
  "Rejected as primary",
  "Population-averaged quasi-Tweedie",
  "Converged in 7 iterations; rank 15/15",
  "140 participant clusters; robust minimum eigenvalue 0.00175; condition number 162 (category block 46.7)",
  "Lag-1 Pearson residual correlation 0.478; maximum cluster-score share 0.334",
  "Approved primary; cluster influence is mandatory",
  "Mundlak within-between quasi-Tweedie",
  "Converged in 8 iterations; rank 21/21",
  "140 participant clusters; robust minimum eigenvalue 0.00132; condition number 4,445 (category block 68.8)",
  "Lag-1 Pearson residual correlation 0.492; maximum cluster-score share 0.313",
  "Secondary audit only"
)

stage1_gt(architecture_pilots) |>
  cols_width(
    Model ~ pct(18),
    Numerical_status ~ pct(18),
    Dependence_or_covariance ~ pct(25),
    Diagnostic_signal ~ pct(22),
    Disposition ~ pct(17)
  )
```

The nonzero residual autocorrelation in the quasi models is not treated as if
it vanished: participant clustering is deliberately robust to arbitrary
within-participant correlation. It does, however, make the 140 participants,
not the 17,935 hours, the inferential replication count. Robust covariance
does not protect against a wrong conditional-mean formula, severe cluster
leverage, or too few effective clusters; those remain explicit failure gates.

```{r}
#| label: tbl-resampling-feasibility-pilot
#| tbl-cap: "Authorized 50-replicate site-stratified participant-bootstrap feasibility pilot."

resampling_feasibility <- tribble(
  ~Model, ~Replicates, ~Successful, ~Convergence_failures, ~Rank_failures, ~Covariance_failures,
  "Population-averaged quasi-Tweedie", 50L, 50L, 0L, 0L, 0L,
  "Mundlak within-between quasi-Tweedie", 50L, 50L, 0L, 0L, 0L
)

stage1_gt(resampling_feasibility) |>
  fmt_integer(-Model, use_seps = TRUE)
```

The two-model pilot took 11.3 seconds. Linear extrapolation gives about 226
seconds for 1,000 replicates or 452 seconds (7.5 minutes) for 2,000 on the same
machine. This authorizes no production bootstrap: any production run remains
subject to the separate compute gate.

## Multiplicity contract

The primary near-eye families are fixed before fitting:

| Family | Members | Adjustment and decision |
|---|---:|---|
| H03-F1-omnibus | One six-restriction category omnibus | Raw participant-cluster-robust Wald-F p; alpha 0.05 |
| H03-F2-context-contrasts | Six supported category-versus-indoor-electric ratios | One BH vector across all supported contrasts; adjusted p < 0.05 |
| H03-F3-site-heterogeneity | One robust interaction/additivity-restriction omnibus | Raw participant-cluster-robust Wald-F p; alpha 0.05; explicitly relabelled as five-category heterogeneity if the fallback is required |
| H03-F4-site-context-contrasts | Every supported site deviation in the accepted heterogeneity model: 53 candidate members for the full seven-category near-eye model or 44 for the predeclared five-category fallback | One BH vector across the complete supported set selected by the results-blind architecture gate; adjusted p < 0.05 |

Chest, paired/common-sample, gap-timing-unaware, boundary, distributional,
and deletion analyses use separately labelled complementary or sensitivity
families. They never enter H03-F1 through H03-F4 and cannot upgrade the primary
claim. Raw and adjusted numeric p-values remain at full precision in source
artifacts.

Every displayed p-value will use a leading zero and three decimals, use
`<0.001` only when strictly below 0.001, and be bold only when significant
under the explicitly labelled raw or BH-adjusted rule. Dunnett-X may be shown
only as a named shared-control sensitivity and will not be called BH.

Site/category means and within-site category-versus-indoor ratios are practical
descriptive estimands with 95% intervals. Unless they are members of H03-F2 or
H03-F4 as defined above, they receive no significance label or unplanned
p-value.

## Paired placement and data-preparation sensitivities

### Paired/common sample

```{r}
#| label: h03-paired-candidate

near_keys <- placements$Near_eye$candidate |>
  select(all_of(key_columns), light_source, participant, participant_day)
chest_keys <- placements$Chest$candidate |>
  select(all_of(key_columns), light_source, participant, participant_day)

paired_candidate <- inner_join(
  near_keys,
  chest_keys,
  by = key_columns,
  suffix = c("_near", "_chest"),
  relationship = "one-to-one"
)
stopifnot(
  all(
    as.character(paired_candidate$light_source_near) ==
      as.character(paired_candidate$light_source_chest)
  )
)

paired_summary <- paired_candidate |>
  summarise(
    Complete_hour_pairs = n(),
    Paired_participants = n_distinct(participant_near),
    Paired_participant_days = n_distinct(participant_day_near),
    Sites = n_distinct(site)
  )

paired_category_support <- paired_candidate |>
  mutate(light_source = light_source_near) |>
  group_by(light_source) |>
  summarise(
    Hours = n(),
    Participants = n_distinct(participant_near),
    Participant_days = n_distinct(participant_day_near),
    Sites = n_distinct(site),
    .groups = "drop"
  ) |>
  complete(
    light_source = factor(category_levels, levels = category_levels),
    fill = list(Hours = 0L, Participants = 0L, Participant_days = 0L, Sites = 0L)
  ) |>
  mutate(
    Category = unname(category_codes[as.character(light_source)]),
    Primary_light_source = unname(category_short[Category]),
    Pooled_rule = Hours >= 200L & Participants >= 20L & Sites >= 3L
  ) |>
  select(Category, Primary_light_source, Hours, Participants, Participant_days, Sites, Pooled_rule)
```

```{r}
#| label: tbl-paired-candidate
#| tbl-cap: "Exact paired/common candidate and category support before fitting."

paired_summary |>
  stage1_gt() |>
  fmt_integer(everything(), use_seps = TRUE)

paired_category_support |>
  stage1_gt() |>
  fmt_integer(c(Hours, Participants, Participant_days, Sites), use_seps = TRUE) |>
  cols_width(
    Category ~ pct(20),
    Primary_light_source ~ pct(30),
    everything() ~ pct(10)
  )
```

The paired/common sample contains 14,103 complete hourly pairs, 111
participants, 631 participant-days, and eight sites. Outdoor electric light
has 162 paired hours and therefore fails the approved 200-hour contrast rule
in this smaller frame. It remains in the paired model and sample flow, but its
paired category contrast and identity-plot point will be marked
support-non-estimable.

Near-eye and chest models will be fitted separately to this identical paired
frame. When matched site-standardized category means and indoor-referenced
ratios share the same interpretation,
the Stage 3 display will put near-eye on the horizontal axis and chest on the
vertical axis, use a 45-degree identity line and null reference lines, and show
component 95% intervals in the figure or adjacent table. It is a concordance
display, not pooling and not evidence of placement equivalence.

### Gap-timing-unaware dataset

The **gap-timing-unaware dataset** still passed the general 50%-per-hour and
80%-per-day coverage rules. The name means that the timing of remaining
missing observations is not used for an additional metric-specific
adjustment; it does not mean that gaps, missingness, or coverage were ignored.
For contrast, the approved preparation of the primary dataset could be
interpreted as time-sensitive where missing-observation timing matters. From
here onward this document calls it simply the primary dataset.

```{r}
#| label: h03-gap-timing-unaware-candidate

gap_path <- file.path(
  project_root,
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/one_hour_data.rds"
)
gap_raw <- readRDS(gap_path)

gap_candidate <- gap_raw |>
  mutate(
    site = factor(site, levels = site_levels),
    local_date = as.Date(local_date),
    clock_minute = as.integer(
      format(local_clock_datetime_utc_proxy, "%H", tz = "UTC")
    ) * 60L +
      as.integer(format(local_clock_datetime_utc_proxy, "%M", tz = "UTC")),
    participant = interaction(site, Id, drop = TRUE, lex.order = TRUE),
    participant_day = interaction(site, Id, local_date, drop = TRUE, lex.order = TRUE)
  ) |>
  left_join(
    diary_eligible |>
      select(all_of(key_columns), light_source, light_source_code),
    by = key_columns,
    relationship = "many-to-one"
  ) |>
  mutate(model_candidate = is.finite(medi_geometric_mean_lx) & !is.na(light_source))

gap_summary <- gap_candidate |>
  group_by(position) |>
  summarise(
    Grid_hours = n(),
    Outcome_available = sum(is.finite(medi_geometric_mean_lx)),
    Outcome_hours_without_primary_category = sum(
      is.finite(medi_geometric_mean_lx) & is.na(light_source)
    ),
    Candidate_hours = sum(model_candidate),
    Participants = n_distinct(participant[model_candidate]),
    Participant_days = n_distinct(participant_day[model_candidate]),
    Sites = n_distinct(site[model_candidate]),
    .groups = "drop"
  ) |>
  mutate(
    Placement = recode(position, glasses = "Near-eye", chest = "Chest"),
    .before = 1
  ) |>
  select(-position)
```

```{r}
#| label: tbl-gap-timing-unaware-candidate
#| tbl-cap: "Gap-timing-unaware candidate frames before fitting."

gap_summary |>
  stage1_gt() |>
  fmt_integer(-Placement, use_seps = TRUE) |>
  cols_width(Placement ~ pct(14), everything() ~ pct(12))
```

The same new implementation, formulas, support rules, contrasts, diagnostics,
and reporting code will be applied to the primary and gap-timing-unaware
datasets. Only the prepared dataset differs. The frozen V0 result remains a
separate implementation-history comparison.

## Diagnostics, influence, and failure rules

Stage 2 will interpret, not merely print, the following diagnostics for every
primary and complementary fit:

- exact participants, participant-days, one-hour observations, sites, category
  counts, and category-by-site counts after the fitting engine's model frame is
  recovered;
- key uniqueness, join cardinality, source/category consistency, missingness,
  valid-minute support, state/context composition, and duplicated outcome
  detection;
- GLM convergence, iteration count, fixed-effect rank, aliased columns,
  coefficient finiteness, model-covariance finiteness, and boundary status;
- participant-cluster covariance eigenvalues and condition numbers for the
  full matrix and each consequential restriction/contrast block;
- the number and size distribution of independent participant clusters,
  cluster hat/leverage summaries, cluster-score concentration, and whether a
  small number of participants supplies most information for a contrast;
- working Pearson and deviance residuals against fitted values, category,
  site, clock time, and support variables; observed zeros and positive tails
  are summarized as calibration diagnostics rather than as a claim that the
  quasi model supplies a complete outcome distribution;
- residual ACF by lag within participant-day sequences and across declared
  boundaries. Remaining autocorrelation is reported even though the sandwich
  covariance allows arbitrary within-participant dependence;
- practical stability under site-standardized versus observed-sample-weighted
  prediction distributions;
- the results-blind full-interaction gate, including estimable restrictions,
  sparse-cell support, effective cluster information, and finite supported-cell
  intervals before any interaction effect is interpreted;
- leave-one-site-out refits for every observed site and deletion refits for the
  five most influential participant clusters identified by the prespecified
  cluster-score measure; and
- change in the omnibus decision, all six pooled category ratios, all supported
  site deviations, and interval coverage under each deletion.

No primary mixed-model $R^2$, ICC, random-effect variance, likelihood fit index,
or likelihood-based variance partition exists for this quasi-likelihood
architecture. Model fit is described through the mean-model diagnostics,
robust omnibus tests, effect estimates, and influence/stability checks above.
Following the author's 2026-08-07 Stage 2 diagnostic amendment, the report may
also show clearly defined descriptive fixed-effect point summaries: overall
and partial quasi-deviance or squared-error R² from nested fixed-effect fits,
plus a two-block site/category Shapley allocation. These are not mixed-model
R² values, effect tests, causal contributions, or uncertainty-bearing variance
components. The exploratory temporal GAM/GAMM retains its separate fit and
variance definitions below.

The already completed 50-replicate participant bootstrap established
computational feasibility only. If production resampling is approved, the
planned run is 2,000 site-stratified participant resamples with the full fitting
and contrast-extraction path repeated. It will quantify cluster-composition
stability and influence; the prespecified delta-method robust intervals remain
primary unless an additional interval change is approved before production.

A primary fit fails its adequacy gate if it does not converge, is rank
deficient, has non-finite coefficients, yields a non-finite or non-positive
covariance for consequential restrictions, has too few effective clusters,
shows a consequential unmodelled mean pattern, produces non-finite intervals,
or changes the primary conclusion under one dominant site or participant
without a scientifically defensible explanation. Residual zero mass or tail
mismatch is described but is not by itself a likelihood goodness-of-fit
failure because no full likelihood is asserted. Failure is reported; it does
not authorize a post hoc model search.

### Named, bounded sensitivities

1. Population-averaged primary versus the declared Mundlak within-between
   audit on common rows.
2. Fixed working Tweedie power 1.539919 versus 1.30 and 1.80
   working powers on common rows; the robust covariance and estimand are held
   fixed.
3. All-available versus paired/common-sample near-eye and chest fits.
4. Primary versus gap-timing-unaware datasets under the same implementation.
5. Site-standardized versus observed-sample-weighted prediction distributions.
6. Inclusion versus exclusion of state- or measurement-context-boundary hours.
7. Inclusion versus exclusion of site/category cells failing the contrast
   support rule, while preserving all categories in the primary additive fit.
8. Full seven-category versus predeclared five-category heterogeneity
   architecture, only when the results-blind full-model gate fails.
9. Exact V0 five-category/joint-test bridge on common current rows, clearly
   separated from both the primary model and the frozen V0 output.
10. Leave-one-site-out and bounded cluster influence refits.

No sensitivity can silently replace the primary result. Each receives exact
samples, 95% intervals, diagnostics, and a stability classification.

## Exploratory time-of-day context

The temporal analysis is exploratory context only. It will be retained only
if it adds interpretable time-of-day information after the primary H03 result
and passes its own diagnostics. It uses the same supported one-hour geometric
melEDI outcome transformed as `log10(melEDI + 0.1)`. It does not reuse H02's
rho, fitted objects, predictions, significance, or claims.

The unchanged inherited H02 components are the cyclic overall time smooth,
sum-to-zero site deviation smooth, participant factor smooth,
participant-day intercept, basis dimensions, local 0–24 h clock knots,
boundary-aware AR algorithm, and fREML final fit. H03 adds a sum-to-zero
light-source smooth around the global time-of-day smooth. By author decision,
the exploratory temporal model does **not** use indoor electric as its
reference. Its category curves are global time-of-day plus category-specific
`sz` deviations whose corresponding coefficients sum to zero across the seven
categories. The primary quasi-Tweedie model retains the indoor-electric and
site-standardized reference system defined above.

### Exact V0 and approved temporal formulas — declared, not fitted

```{r}
#| label: tbl-temporal-formulas
#| tbl-cap: "Evaluated exploratory temporal Wilkinson formulas. No GAM or GAMM is fitted."

temporal_formulas <- list(
  v0_temporal_full = lzMEDI ~
    s(Time, k = 12) +
    s(Time, lightsource_primary, bs = "sz", k = 12) +
    s(Time, site, bs = "sz", k = 12) +
    s(Time, photoperiod.state, bs = "sz", k = 12) +
    s(Time, Id, bs = "fs") +
    s(Id_date, bs = "re"),
  v0_temporal_null = lzMEDI ~
    s(Time, k = 12) +
    s(Time, site, bs = "sz", k = 12) +
    s(Time, photoperiod.state, bs = "sz", k = 12) +
    s(Time, Id, bs = "fs") +
    s(Id_date, bs = "re"),
  approved_Tglobal_h02_inherited = h03_temporal_response ~
    s(time_hour, bs = "cc", k = 12) +
    s(time_hour, site, bs = "sz", k = 12) +
    s(time_hour, participant, bs = "fs", k = 10) +
    s(participant_day, bs = "re"),
  approved_Tcategory_sz = h03_temporal_response ~
    s(time_hour, bs = "cc", k = 12) +
    s(time_hour, light_source, bs = "sz", k = 12) +
    s(time_hour, site, bs = "sz", k = 12) +
    s(time_hour, participant, bs = "fs", k = 10) +
    s(participant_day, bs = "re")
)

temporal_formula_contract <- enframe(
  temporal_formulas,
  name = "model_id",
  value = "formula"
) |>
  mutate(formula = vapply(formula, formula_text, character(1)))

stage1_gt(temporal_formula_contract) |>
  cols_width(model_id ~ pct(27), formula ~ pct(73))
```

`approved_Tcategory_sz` is the prespecified exploratory model, not the result of
a term-selection search. It is fitted with `mgcv::bam()`, Gaussian identity
family, `method = "fREML"`, `discrete = TRUE`, and one R thread. Local clock
time is the one-hour midpoint from 0.5 through 23.5 h, with cyclic knots at 0
and 24 h. The preliminary fit estimates H03-specific rho from boundary-aware
lag-1 residuals; the final fit uses that fixed rho. AR starts reset at every
participant-day, after every omitted or non-consecutive hour, and on both sides
of a non-one-to-one fall-back outcome. The inherited no-category model is an
implementation benchmark only. Restricted-likelihood values are not compared
across the two differing mean structures, and the GAM/GAMM never replaces the
primary H03 test.

### Temporal estimands and figure

The temporal outputs are distinct from the primary model outputs:

1. **Global daily curve:** the intercept plus cyclic global smooth, with the
   light-source and site `sz` deviations at their sum-to-zero reference and
   participant/participant-day terms excluded.
2. **Category daily curves:** global smooth plus each category's `sz` deviation.
   Reader-facing curves are predicted for every site, participant and day terms
   are excluded, site linear predictors are averaged with weight $1/S$, and
   the average is inverse-transformed once. These are back-transformed
   descriptive profiles, not quasi-Tweedie expected means.
3. **Category deviations:** each `sz` deviation from the global curve on the
   `log10(melEDI + 0.1)` scale, with a zero reference. Indoor electric is one of
   seven deviations and receives no special temporal reference status.
4. **Local support:** for every category and clock hour, the source data store
   participant-hours, distinct participants, participant-days, and sites. A
   curve segment is marked locally sparse when it has fewer than 20 hours, five
   participants, or three sites. Hours with zero observations have a light-grey
   background in the curve and deviation panels and contain no curve or ribbon;
   supported runs are not bridged across these gaps.

The recognizable corrected V0-style temporal figure has three vertically
aligned panels within balanced 4-category and 3-category facet blocks:

- site-standardized category curves in melEDI units with 95% intervals;
- category `sz` deviations from the global time-of-day curve with a zero line;
  and
- the **available participant-hour n at each local time**, using the same
  category facets and clock axis. The caption will state that these are rows,
  not independent participants; participant and participant-day counts remain
  available in the source CSV.

The melEDI panel uses `LightLogR::symlog_trans(base = 10, thr = 1, scale = 1)`
as a display-only transformation with original-unit labels and untransformed
source data. The count panel uses original counts with a declared display scale.
The figure uses the approved 14-point cowplot basis, 12-point ticks, 14-point
titles/tags, and 11-point caption, is saved tightly to the figure rather than an
A4 canvas, and is checked at final size for facet balance, clipping, line
breaks, and sparse-category readability.

### Exploratory R² and variance decomposition

These GAM/GAMM summaries use definitions that differ from the primary model's
post-approval descriptive fixed-effect R² summaries. They will be kept in
separate table blocks and labelled as descriptive in-sample quantities:

- `summary.gam()` adjusted R² and deviance explained for the complete model,
  explicitly on the `log10(melEDI + 0.1)` scale and over the fitted rows;
- a site-standardized, participant-balanced in-sample squared-error R², using
  weights that give each site total weight $1/S$, each participant within site
  equal weight, and each of that participant's fitted hours equal weight;
- smoothing/random-effect variance parameters and 95% intervals from
  `gratia::variance_comp(rescale = TRUE)`, reported as penalty-scale variance
  or standard-deviation parameters rather than percentages of outcome
  variance; and
- an exact global Shapley/marginal allocation of fitted linear-predictor
  variance over the same site-standardized, participant-balanced reference
  sample. The prespecified groups are global time, light-source deviations,
  site deviations, participant curves, and participant-day shifts. The
  intercept is excluded. Allocations use the complete component covariance,
  sum to the fitted-predictor variance, may be negative through covariance, and
  are not causal importance or percentages of raw melEDI variance.

The allocation will use the audited `gamm_variance_partition()` contract with
explicit term groups and reference weights. Raw component variances, component
covariances, signed Shapley allocations, and partial-unique partitions will not
be conflated. The report will not repeat V0's invalid practice of dividing raw
term variances by their sum or subtracting overlapping R² values.

Temporal diagnostics include convergence, `gam.check()` basis-dimension
checks, `gratia::appraise()`, formal concurvity, the sum-to-zero constraints,
Gaussian residual distribution and heteroscedasticity, standardized residual
ACF within each sequence, midnight behavior, cluster influence, and local
category/time/site support. The approximate whole-term smooth test, if shown,
is explicitly exploratory and conditional on the fitted smoothing parameters.
Curve-wide claims use simultaneous 95% bands across the declared category-time
family, not scattered pointwise exclusions of zero.

Any heavy simultaneous-band or variance-allocation simulation begins with 100
draws, displays numerical stability and elapsed time, projects production
runtime, and stops for approval. No production draw count is authorized by
this Stage 1 revision.

## Stage 2 result and display contract

Stage 2 will produce recognizable but corrected H03 tables and figures:

- a near-eye primary table anchored by the site-standardized indoor-electric
  mean, with all seven site-standardized category means, six indoor-referenced
  ratios, 95% CIs, raw p-values, BH-adjusted p-values, support, and exact fitted
  sample;
- a V0-recognizable corrected site table that shows the absolute derived cell
  mean, the category multiplier relative to site-standardized indoor electric,
  and the site multiplier relative to the same category's site-standardized
  mean, each with 95% CIs and explicit support;
- a complementary chest table with the same reference algebra and its eight-site
  scope;
- a site/category estimate figure using the submitted site order and colours,
  legible category line breaks, and explicit support-non-estimable markers;
- an omnibus and model-adequacy table that separates category association from
  site heterogeneity and records whether the full or core-category interaction
  passed its results-blind architecture gate;
- a paired/common-sample estimate comparison when the matched estimands are
  valid; and
- an exploratory time-of-day figure only if the temporal model is valid, with
  global-plus-`sz` category curves, deviation smooths, and the aligned
  participant-hour availability panel requested above; plus a separate,
  explicitly defined GAM/GAMM fit-and-variance table.

Every plotted value will have a source CSV. Tables use readable `gt` output.
Figures will be inspected at their final size for clipping, aspect ratio,
panel balance, line breaks, labels, and legends. Reader-facing melEDI axes use
the approved LightLogR symmetric-log display scale with original-unit labels;
source values remain untransformed. Figure typography follows the approved
14-point cowplot, 12-point tick, 14-point title/tag, and 11-point caption
reference, with final-size detail QA. Near-eye remains primary, chest remains
complementary, and no display will imply pooling or equivalence.

Stage 2 will explicitly reconcile V0's 16,774 and 18,391 rows, five-category
tables, joint omnibus, contrast adjustments, missing intervals, copied chest
interpretation, temporal outcome, and R² wording. It will not make causal,
health-outcome, placement-equivalence, or worn-ocular sleep claims.

## Audit findings

```{r}
#| label: tbl-h03-findings
#| tbl-cap: "H03 Stage 1 findings and required disposition."

h03_findings <- tribble(
  ~Finding, ~Severity, ~Status, ~Evidence, ~Required_disposition,
  "H03-001", "Major", "Confirmed defect", "V0 modelled five of seven registered categories", "Adopt the fixed seven-category dictionary and support-non-estimable rule",
  "H03-002", "Major", "Confirmed defect", "V0 site-only versus full adds category and interaction jointly", "Use a six-restriction robust category omnibus and a separate robust interaction/additivity omnibus",
  "H03-003", "Major", "Confirmed defect", "Dunnett-X and split within-category site adjustments were labelled as BH families", "Use declared H03-F1 through H03-F4 vectors and exact labels",
  "H03-004", "Major", "Confirmed defect", "V0 treated participant hours as conditionally independent beyond one participant intercept", "Use participant-clustered covariance for the primary mean model; reserve participant/day/AR structure for the exploratory temporal model",
  "H03-005", "Major", "Confirmed defect", "V0 had no 30-valid-minute rule or persisted join-cardinality assertion", "Use the accepted one-hour metric and hard join assertions",
  "H03-006", "Major", "Confirmed defect", "V0 removed confidence limits before display", "Report 95% intervals for every consequential estimand",
  "H03-007", "Major", "Confirmed defect", "Overlapping mixed-model and raw term-variance differences were called variance explained", "Report no primary R2; define GAM adjusted/weighted R2, penalty variance, and link-scale Shapley allocation separately",
  "H03-008", "Major", "Confirmed defect", "V0 temporal analysis used the wrong outcome and invalid inherited H02 structure", "Use only the approved H02 temporal specification adapted to H03",
  "H03-009", "Major", "Confirmed reporting defect", "Chest prose duplicates the near-eye numerical interpretation", "Regenerate chest prose only from verified chest artifacts",
  "H03-010", "Major", "Current design fact", "Chest literal interaction is rank 55/56 because FUSPCEU has no external-sleep candidate hour", "Use the rank-complete observed-cell parameterization and a 41-restriction robust additivity test",
  "H03-011", "Moderate", "Current support fact", "Ten near-eye cells and eight observed-site chest cells fail the approved site-specific rule", "Retain categories; suppress only unsupported contrasts with reasons",
  "H03-012", "Evidence limitation", "Open shared provenance limitation", "Preparation report FIND-044 does not store an independent reconstruction tied to every current metric identity", "Carry the limitation forward; do not describe it as evidence that data are wrong",
  "H03-013", "Major", "Architecture-pilot finding", "The initial participant/day/AR Tweedie mixed model converged but was singular and failed zero/residual-shape diagnostics", "Do not use it as primary; preserve its diagnostic record without inspecting its effects",
  "H03-014", "Moderate", "Architecture decision", "The population-averaged quasi-Tweedie model is full rank with finite positive participant-cluster covariance; cluster-score concentration remains material", "Use it as primary with finite-cluster tests, mandatory participant influence checks, and no likelihood or mixed-model summaries",
  "H03-015", "Moderate", "Open companion-model gate", "The full robust site-by-category interaction has not yet passed a covariance/influence pilot", "Gate it before effect inspection; use the predeclared five-category fallback only if the full gate fails",
  "H03-016", "Moderate", "Author temporal decision", "Exploratory category curves must be sum-to-zero deviations around a global time-of-day smooth, not indoor-referenced smooths", "Use the H02-derived global plus category/site sz structure and report explicitly defined GAM R2, variance components, Shapley allocation, and clock-time n"
)

stage1_gt(h03_findings) |>
  cols_width(
    Finding ~ pct(9),
    Severity ~ pct(10),
    Status ~ pct(16),
    Evidence ~ pct(30),
    Required_disposition ~ pct(35)
  )
```

## Recorded author decisions controlling Stage 2

The owner approved the revised primary process and then specified the temporal
reference exception recorded in H03-G13. This table is the controlling Stage 2
contract. The present request is limited to updating Stage 1; it does not start
Stage 2 or authorize production resampling.

| Gate | Approved decision |
|---|---|
| H03-G1 category construct | Keep all seven registered primary categories; indoor electric is the primary-model category reference; sleep categories are bedside-environment contexts |
| H03-G2 estimability | Use the 200-hour/20-participant/3-site pooled rule and the 20-hour/5-participant/5-shared site-specific rule per fitted frame |
| H03-G3 outcome and join | Use the accepted one-hour zero-aware geometric melEDI outcome, at least 30 valid minutes, one-to-one local-hour diary join with retained true-UTC diary start |
| H03-G4 primary dependence | Use a population-averaged mean model with site-prefixed participant-clustered HC1 sandwich covariance and finite-cluster reference distributions; do not include participant/day/AR random components in the primary fit |
| H03-G5 primary tests | H03-F1 is the six-category-coefficient robust Wald-F test; H03-F3 is a separate robust interaction/additivity-restriction Wald-F test; never combine them |
| H03-G6 site hierarchy | Fixed site; results-blind full interaction gate; literal 48-df interaction near eye and rank-complete 41-restriction observed-cell companion for chest; five-category fallback only if the full gate fails |
| H03-G7 primary family | Quasi-Tweedie log-link mean model with working variance power fixed at 1.539919 and participant-clustered covariance; no likelihood, AIC, mixed-model R², ICC, or random-effect summary |
| H03-G8 practical estimates | Use the site-standardized indoor-electric mean as absolute anchor, six site-standardized category ratios versus indoor electric, supported cell means, category-specific site deviations, and descriptive within-site indoor ratios, all with 95% CIs |
| H03-G9 multiplicity | Use H03-F1 through H03-F4 exactly; chest and sensitivities use separately labelled families; site/category descriptive estimands receive no unplanned p-values; Dunnett-X only as a named sensitivity |
| H03-G10 placement | Near-eye primary, chest complementary, and separately fitted paired/common-sample comparison without pooling or equivalence language |
| H03-G11 sensitivities | Run the ten bounded sensitivities listed above, including the gap-timing-unaware dataset under the same implementation |
| H03-G12 influence and failure | Use the GLM, robust-covariance, effective-cluster, mean-pattern, temporal-residual, sparse-cell, site, participant, and failure gates above |
| H03-G13 temporal context | Retain an exploratory H02-derived GAM/GAMM with cyclic global time, sum-to-zero category and site deviations, participant curves, participant-day shifts, and boundary-aware AR(1); temporal categories are referenced to the global daily curve, not indoor electric |
| H03-G14 temporal summaries | Report time-resolved category curves and `sz` deviations, clock-time participant-hour n, mgcv adjusted and site-standardized participant-balanced R², penalty variance parameters, and exact link-scale Shapley/partial partitions with their definitions and limits |
| H03-G15 reporting | Exact fitted samples, 95% intervals, REPORT-008 p-values, readable `gt` tables, accessible final-size figures and source CSVs, approved symlog display, and no causal/health/equivalence claims |
| H03-G16 resampling | The completed 50-replicate quasi-model pilot is feasibility-only; every other heavy bootstrap or simulation begins with 50 or 100 pilot replicates and a runtime estimate; all production computation requires separate approval |
| H03-G17 V0 disposition | Treat every V0 H03 estimate, p-value, significance statement, R² attribution, chest interpretation, and temporal claim as reopened until Stage 2 reconciliation |

Post-approval author amendment, 2026-08-07: Stage 2 may add descriptive
fixed-effect overall/partial R² and a two-block site/category allocation for
the primary GLM, plus a bounded participant-blocked comparison with raw and
log-transformed Gaussian candidates. These outputs are diagnostics only: they
do not add a multiplicity family, replace H03-F1 through H03-F4, create random
effects, or authorize model selection by invalid cross-family AIC.

## Render-time assertions and environment

```{r}
#| label: tbl-stage1-assertions
#| tbl-cap: "Stage 1 assertions; all must pass."

stage1_assertions <- tibble(
  Assertion = c(
    "R version is exactly 4.6.1",
    "eligible diary local-hour keys are unique",
    "eligible diary starts are hour-aligned",
    "all seven categories match their declared source flags",
    "indoor electric is the declared primary-model category reference",
    "the fixed working Tweedie power lies strictly between 1 and 2",
    "all seven categories pass the outcome-blind pooled rule",
    "all seven categories pass the all-available pooled rule by placement",
    "all five fallback heterogeneity categories meet the declared core rule",
    "near-eye additive and interaction designs are full rank",
    "chest additive and observed-cell designs are full rank",
    "chest literal interaction is deficient by exactly one column",
    "paired rows carry identical primary categories",
    "no model object was fitted in this render"
  ),
  Passed = c(
    getRversion() == "4.6.1",
    diary_key_audit$Duplicate_keys == 0L,
    diary_key_audit$Non_hour_aligned_starts == 0L,
    diary_category_audit$Recorded_primary_without_expected_flag == 0L,
    identical(reference_label, category_levels[[1L]]),
    working_tweedie_power > 1 && working_tweedie_power < 2,
    all(diary_overall_support$Pooled_rule),
    all(candidate_category_support$Pooled_rule),
    all(core_heterogeneity_support$Meets_core_rule),
    all(rank_audit$Full_rank[rank_audit$Placement == "Near-eye"]),
    all(
      rank_audit$Full_rank[
        rank_audit$Placement == "Chest" & rank_audit$Design_id != "literal_interaction"
      ]
    ),
    with(
      filter(rank_audit, Placement == "Chest", Design_id == "literal_interaction"),
      Columns - Rank == 1L && !Full_rank
    ),
    all(
      as.character(paired_candidate$light_source_near) ==
        as.character(paired_candidate$light_source_chest)
    ),
    !any(vapply(
      mget(ls(envir = .GlobalEnv), envir = .GlobalEnv),
      function(x) inherits(x, c("glmmTMB", "gam", "bam", "lmerMod", "glm")),
      logical(1)
    ))
  )
)

stopifnot(all(stage1_assertions$Passed))

stage1_gt(stage1_assertions) |>
  cols_width(Assertion ~ pct(85), Passed ~ pct(15))
```

```{r}
#| label: tbl-software-environment
#| tbl-cap: "Software recorded for this audit render."

package_ids <- c(
  "dplyr",
  "tidyr",
  "purrr",
  "tibble",
  "gt",
  "knitr",
  "digest",
  "glmmTMB",
  "emmeans",
  "statmod",
  "sandwich",
  "mgcv",
  "gratia",
  "itsadug",
  "LightLogR"
)

software_environment <- bind_rows(
  tibble(Component = "R", Version = as.character(getRversion())),
  tibble(
    Component = paste0("R package: ", package_ids),
    Version = vapply(
      package_ids,
      function(package) as.character(utils::packageVersion(package)),
      character(1)
    )
  ),
  tibble(
    Component = "Quarto",
    Version = Sys.getenv("QUARTO_VERSION", unset = "1.9.37")
  )
)

stage1_gt(software_environment) |>
  cols_width(Component ~ pct(70), Version ~ pct(30))
```

::: {.callout-important title="Stage 1 status"}
Stage 1 ends here with H03-G1 through H03-G17 recorded as the approved
analysis contract. This revision does not start Stage 2. No category
association, interaction estimate, temporal smooth, or production resampling
was computed while updating this report.
:::
