---
title: "H03 Stage 2: implementation, diagnostics, and V0 comparison"
subtitle: "Nature Health analysis audit — author result gate"
date: "2026-08-07"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
editor: source
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H03_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- locate_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

stopifnot(identical(as.character(getRversion()), "4.6.1"))
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))

library(dplyr)
library(tidyr)
library(readr)
library(gt)

source(file.path(project_root, "scripts/pipeline/p_value_display.R"))

artifact <- function(...) file.path(project_root, "artifacts", ...)
read_h03 <- function(...) {
  readr::read_csv(artifact(...), show_col_types = FALSE, na = "")
}

approvals <- read_h03("06_model_data", "H03", "H03_author_approvals.csv")
input_audit <- read_h03("06_model_data", "H03", "H03_input_audit.csv")
formulas <- read_h03("06_model_data", "H03", "H03_formula_registry.csv")
samples <- read_h03("06_model_data", "H03", "H03_model_frame_index.csv")
category_support <- read_h03("06_model_data", "H03", "H03_category_support.csv")
cell_support <- read_h03("06_model_data", "H03", "H03_site_category_support.csv")
gate <- read_h03("08_diagnostics", "H03", "H03_interaction_architecture_gate.csv")
selection <- read_h03("08_diagnostics", "H03", "H03_interaction_architecture_selection.csv")
primary <- read_h03("09_tables", "H03", "H03_primary_category_estimands.csv")
omnibus <- read_h03("09_tables", "H03", "H03_primary_omnibus_tests.csv")
site_context <- read_h03("09_tables", "H03", "H03_site_context_estimands.csv")
diagnostics <- read_h03("08_diagnostics", "H03", "H03_model_diagnostics.csv")
zero_mass <- read_h03("08_diagnostics", "H03", "H03_zero_mass_diagnostics.csv")
paired <- read_h03("09_tables", "H03", "H03_paired_placement_comparison.csv")
sensitivity <- read_h03("09_tables", "H03", "H03_sensitivity_comparison.csv")
sensitivity_omnibus <- read_h03("09_tables", "H03", "H03_sensitivity_omnibus_tests.csv")
influence <- read_h03("09_tables", "H03", "H03_influence_category_refits.csv")
influence_site <- read_h03("09_tables", "H03", "H03_influence_site_context_refits.csv")
influence_jobs <- read_h03("08_diagnostics", "H03", "H03_influence_refit_registry.csv")
v0_models <- read_h03("09_tables", "H03", "H03_v0_bridge_models.csv")
v0_tests <- read_h03("09_tables", "H03", "H03_v0_bridge_tests.csv")
v0_ratios <- read_h03("09_tables", "H03", "H03_v0_bridge_ratios.csv")
v0_comparison <- read_h03("09_tables", "H03", "H03_v0_denominator_and_test_comparison.csv")
deferred <- read_h03("09_tables", "H03", "H03_deferred_computation.csv")
standardization_reconciliation <- read_h03(
  "09_tables",
  "H03",
  "H03_standardization_reconciliation.csv"
)
glm_candidate_registry <- read_h03(
  "09_tables",
  "H03",
  "H03_glm_candidate_model_registry.csv"
)
glm_r_squared <- read_h03(
  "09_tables",
  "H03",
  "H03_glm_r_squared_and_effect_partition.csv"
)
glm_cross_validation <- read_h03(
  "09_tables",
  "H03",
  "H03_glm_candidate_cross_validation.csv"
)

temporal_summary <- read_h03("09_tables", "H03", "H03_temporal_model_summary.csv")
temporal_r2 <- read_h03("09_tables", "H03", "H03_temporal_weighted_r_squared.csv")
temporal_partition <- read_h03("09_tables", "H03", "H03_temporal_shapley_allocation.csv")
temporal_k <- read_h03("08_diagnostics", "H03", "H03_temporal_k_check.csv")
temporal_concurvity <- read_h03("08_diagnostics", "H03", "H03_temporal_concurvity.csv")
temporal_constraints <- read_h03("08_diagnostics", "H03", "H03_temporal_sz_constraints.csv")
temporal_midnight <- read_h03("08_diagnostics", "H03", "H03_temporal_midnight_diagnostics.csv")
temporal_acf <- read_h03("08_diagnostics", "H03", "H03_temporal_residual_acf.csv")
temporal_smooth_tests <- read_h03("09_tables", "H03", "H03_temporal_smooth_tests.csv")
temporal_basis_registry <- read_h03(
  "09_tables", "H03", "H03_temporal_basis_model_registry.csv"
)
temporal_basis_fit <- read_h03(
  "09_tables", "H03", "H03_temporal_basis_fit_comparison.csv"
)
temporal_basis_delta <- read_h03(
  "09_tables", "H03", "H03_temporal_basis_comparison_deltas.csv"
)

format_p <- function(value) nh_format_p_value(value)
format_p_cell <- function(value, significant = FALSE) {
  ifelse(
    is.na(value),
    "—",
    {
      display <- nh_p_value_display(value, significant = significant)
      ifelse(
        display$p_bold,
        paste0("**", display$p_display, "**"),
        display$p_display
      )
    }
  )
}
format_mean_ci <- function(estimate, low, high) {
  sprintf("%.1f (%.1f to %.1f)", estimate, low, high)
}
format_ratio_ci <- function(estimate, low, high) {
  sprintf("%.3f (%.3f to %.3f)", estimate, low, high)
}
format_number <- function(value, digits = 3L) {
  formatC(value, digits = digits, format = "f", big.mark = ",")
}
audit_gt <- function(data, title, subtitle = NULL, groupname_col = NULL) {
  gt::gt(data, groupname_col = groupname_col) |>
    gt::tab_header(title = title, subtitle = subtitle) |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(4),
      heading.align = "left",
      table.width = gt::pct(100)
    )
}

main_samples <- samples |>
  filter(.data$run_id %in% c("main__near_eye", "main__chest")) |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
near_primary <- primary |>
  filter(.data$placement == "Near-eye")
near_category_test <- omnibus |>
  filter(.data$placement == "Near-eye", .data$test_id == "category_omnibus")
near_heterogeneity_test <- omnibus |>
  filter(
    .data$placement == "Near-eye",
    .data$test_id == "site_heterogeneity_omnibus"
  )
near_indoor_standardization <- standardization_reconciliation |>
  filter(
    .data$placement == "Near-eye",
    .data$category_code == "electric_indoor"
  )
near_cv <- glm_cross_validation |>
  filter(.data$placement == "Near-eye")
chest_cv <- glm_cross_validation |>
  filter(.data$placement == "Chest")
near_quasi_cv <- near_cv |>
  filter(.data$model_id == "quasi_tweedie_log")
near_gaussian_cv <- near_cv |>
  filter(.data$model_id == "gaussian_identity_raw")
near_log_gaussian_cv <- near_cv |>
  filter(.data$model_id == "gaussian_log10_response")
chest_quasi_cv <- chest_cv |>
  filter(.data$model_id == "quasi_tweedie_log")
chest_gaussian_cv <- chest_cv |>
  filter(.data$model_id == "gaussian_identity_raw")
chest_log_gaussian_cv <- chest_cv |>
  filter(.data$model_id == "gaussian_log10_response")

stopifnot(
  nrow(approvals) == 17L,
  all(approvals$approved),
  all(input_audit$hash_verified),
  nrow(primary) == 14L,
  nrow(omnibus) == 4L,
  nrow(near_category_test) == 1L,
  nrow(near_heterogeneity_test) == 1L,
  nrow(near_indoor_standardization) == 1L,
  nrow(glm_candidate_registry) == 3L,
  nrow(glm_r_squared) == 18L,
  nrow(glm_cross_validation) == 6L,
  all(glm_cross_validation$folds == 5L),
  all(glm_cross_validation$all_converged),
  all(glm_cross_validation$all_full_rank),
  abs(
    near_indoor_standardization$stage2_minus_approved_link_scale_lx
  ) < 1e-8,
  all(omnibus$status == "ESTIMABLE"),
  all(gate$gate_pass),
  all(temporal_summary$converged),
  all(temporal_constraints$passes),
  nrow(temporal_basis_registry) == 2L,
  nrow(temporal_basis_fit) == 4L,
  nrow(temporal_basis_delta) == 2L,
  all(temporal_basis_fit$converged),
  all(temporal_basis_fit$rank == temporal_basis_fit$coefficients)
)
```

::: {.callout-important title="Result at the Stage 2 gate"}
The signed hypothesis is:

> **H3: Hourly self-reported light exposure categories predict hourly geometric mean melanopic EDI.**

The seven self-reported primary-light-source categories are associated with
measured one-hour near-eye melEDI in the prespecified primary omnibus,
*F*(`r near_category_test$restrictions`,
`r near_category_test$denominator_df`) =
`r sprintf("%.2f", near_category_test$f_statistic)`,
*p* `r format_p(near_category_test$p_raw)`. Relative to indoor electric
light and after the six-contrast BH correction, near-eye melEDI was higher
for indoor daylight and outdoor daylight and lower for emissive display,
sleep darkness, and external light during sleep. Outdoor electric light was
not distinguishable from indoor electric light.

The full seven-category site-heterogeneity architecture passed its
predeclared, effect-blind estimability gate. Site heterogeneity was strong,
*F*(`r near_heterogeneity_test$restrictions`,
`r near_heterogeneity_test$denominator_df`) =
`r sprintf("%.2f", near_heterogeneity_test$f_statistic)`,
*p* `r format_p(near_heterogeneity_test$p_raw)`, so the site-standardized
ratios are cross-site summaries, not uniform effects that apply identically
at every site.

The primary model is numerically stable and full rank, but it is not a
well-calibrated Tweedie probability model: residual variance remains strongly
fitted-value dependent, repeated hours remain correlated, and the working
Tweedie zero mass substantially exceeds the observed zero fraction. Its
defensible role is therefore a participant-cluster-robust population-mean
analysis. It does not support causal, health-outcome, or placement-equivalence
claims.
:::

## What was fitted

Near-eye is primary. Chest is complementary. The exact one-hour response is
the accepted zero-aware geometric melEDI outcome; no H02 fitted value,
prediction, significance result, or claim was reused.

The primary model has fixed site and light-source terms, a log link, and the
working quasi-Tweedie variance power 1.539919. Site uses submitted-order sum
coding. Light source uses indoor electric light as the treatment reference.
Participant-clustered HC1 covariance, with finite cluster *t* and *F*
reference distributions, carries the repeated-hour and repeated-day
uncertainty. This is a population-mean model, so an lme-style random-effect
singularity parameter does not exist; full design rank and covariance
positive-definiteness are the relevant numerical checks.

The category means and ratios use the submitted site set with each site
receiving the same weight on the fitted log-mean scale, followed by one inverse
link transformation. “Site-standardized” is used below for this geometric mean
of the fitted site-specific expected values.

### Exact evaluated formulas

```{r}
#| label: tbl-stage2-formulas
#| echo: true

formulas |>
  filter(.data$formula_id %in% c(
    "primary_population_mean",
    "full_site_heterogeneity",
    "chest_observed_cell_heterogeneity",
    "secondary_mundlak_audit",
    "v0_bridge_full",
    "v0_bridge_site_only",
    "temporal_category"
  )) |>
  transmute(
    Model = .data$formula_id,
    `Exact Wilkinson formula supplied to R` = .data$formula
  ) |>
  audit_gt(
    "Evaluated H03 formulas",
    "Displayed from the run registry; this report reads saved artifacts and fits no model"
  ) |>
  cols_width(Model ~ pct(25), `Exact Wilkinson formula supplied to R` ~ pct(75))
```

### Exact analytical samples

```{r}
#| label: tbl-primary-samples
#| tbl-cap: "Exact primary and complementary fitted samples."

main_samples |>
  transmute(
    Placement = as.character(.data$placement),
    `Participant-hours` = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Exact-zero hours` = .data$exact_zero_hours,
    `Boundary hours` = .data$boundary_hours
  ) |>
  audit_gt(
    "Fitted sample",
    "One observation is one accepted participant-hour"
  ) |>
  fmt_integer(columns = where(is.numeric), use_seps = TRUE)
```

The near-eye model contains 17,935 participant-hours from 140 participants,
801 participant-days, and nine sites. The chest model contains 19,512
participant-hours from 151 participants, 880 participant-days, and eight
sites. These are row, participant, day, and site denominators respectively;
they are not interchangeable.

## Category association and practical contrasts

```{r}
#| label: tbl-primary-category-results
#| tbl-cap: "Site-standardized category estimates and contrasts to indoor electric light."

primary |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Category = .data$short_label,
    `Expected melEDI, lx (95% CI)` = format_mean_ci(
      .data$expected_mel_edi_lx,
      .data$expected_conf_low_lx,
      .data$expected_conf_high_lx
    ),
    `Ratio to indoor electric (95% CI)` = if_else(
      .data$category_order == 1L,
      "Reference",
      format_ratio_ci(
        .data$ratio_to_indoor,
        .data$ratio_conf_low,
        .data$ratio_conf_high
      )
    ),
    `Raw p` = format_p_cell(.data$p_raw, FALSE),
    `BH-adjusted p` = format_p_cell(
      .data$p_adjusted,
      !is.na(.data$p_adjusted) & .data$p_adjusted <= 0.05
    )
  ) |>
  select(
    Placement,
    Category,
    `Expected melEDI, lx (95% CI)`,
    `Ratio to indoor electric (95% CI)`,
    `Raw p`,
    `BH-adjusted p`
  ) |>
  audit_gt(
    "Primary near-eye and complementary chest estimates",
    paste(
      "Participant-cluster-robust 95% confidence intervals.",
      "Six non-reference contrasts per placement; BH-adjusted p is bold only",
      "when significant at 0.050. Raw p has no bolding rule."
    ),
    groupname_col = "Placement"
  ) |>
  fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
```

For near-eye melEDI, the site-standardized indoor-electric anchor is
`r sprintf("%.1f", near_primary$expected_mel_edi_lx[near_primary$category_code == "electric_indoor"])`
lx. The estimated ratios are
`r sprintf("%.2f", near_primary$ratio_to_indoor[near_primary$category_code == "daylight_indoor"])`
for indoor daylight,
`r sprintf("%.2f", near_primary$ratio_to_indoor[near_primary$category_code == "daylight_outdoor"])`
for outdoor daylight,
`r sprintf("%.2f", near_primary$ratio_to_indoor[near_primary$category_code == "display"])`
for emissive display,
`r sprintf("%.3f", near_primary$ratio_to_indoor[near_primary$category_code == "sleep_darkness"])`
for sleep darkness, and
`r sprintf("%.2f", near_primary$ratio_to_indoor[near_primary$category_code == "sleep_external_light"])`
for external light during sleep. These are observational category
associations relative to indoor electric light.

::: {.callout-note title="Why the preliminary 88.9 lx value is retained"}
The volatile explainer used the same 17,935-hour frame and the same additive
quasi-Tweedie mean model. Its 88.9 lx indoor-electric value was obtained by
averaging the nine site linear predictors and then back-transforming:

$$
\exp\left(S^{-1}\sum_s \eta_{sc}\right).
$$

That is the geometric mean of the site-specific expected values. Following the
author's 2026-08-07 amendment, this is also the approved Stage 2 estimand. The
intermediate Stage 2 render instead averaged the expected values in melEDI
units:

$$
S^{-1}\sum_s \exp(\eta_{sc}).
$$

For indoor electric light, the nine fitted site means range from
`r sprintf("%.1f", near_indoor_standardization$minimum_site_mean_lx)` to
`r sprintf("%.1f", near_indoor_standardization$maximum_site_mean_lx)` lx. Their
geometric mean is
`r sprintf("%.1f", near_indoor_standardization$link_scale_equal_site_backtransform_lx)`
lx, reproducing the preliminary 88.9 lx after rounding; their equal-site
arithmetic mean is
`r sprintf("%.1f", near_indoor_standardization$response_scale_equal_site_mean_lx)`
lx. The arithmetic mean is
`r sprintf("%.1f", near_indoor_standardization$response_minus_link_scale_lx)`
lx, or
`r sprintf("%.1f", near_indoor_standardization$response_relative_to_link_scale_percent)`%,
higher. The now-superseded 97.1 lx value therefore reflected only a different
standardization/back-transformation order, not a sample or fitted-model change.
Because site and category are additive on the log scale, this common site
factor cancels from category ratios; the fitted category coefficients, ratios,
and omnibus test are unchanged. Absolute-mean confidence intervals use the
delta-method gradient for the approved link-scale standardization.
:::

```{r}
#| label: fig-primary-category-results
#| fig-cap: "Site-standardized one-hour melEDI estimates and ratios to indoor electric light. The melEDI transformation is display-only."
#| fig-alt: "Two-panel plot. The upper panel shows site-standardized expected one-hour melEDI with 95% intervals for seven light-source categories at near-eye and chest placements. Outdoor daylight is highest and sleep darkness is lowest. The lower panel shows six category ratios to indoor electric light; outdoor electric crosses one, while daylight ratios are above one and display and sleep-related ratios are below one."
#| out-width: "100%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_primary_category_estimates.png"
))
```

### Omnibus and heterogeneity tests

```{r}
#| label: tbl-primary-omnibus

omnibus |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Test = recode(
      .data$test_id,
      category_omnibus = "Category association",
      site_heterogeneity = "Category-by-site heterogeneity"
    ),
    `F (numerator df, denominator df)` = sprintf(
      "%.2f (%d, %d)",
      .data$f_statistic,
      .data$restrictions,
      .data$denominator_df
    ),
    p = format_p_cell(.data$p_raw, .data$p_raw <= 0.05),
    Architecture = .data$architecture
  ) |>
  select(Placement, Test, `F (numerator df, denominator df)`, p, Architecture) |>
  audit_gt(
    "Prespecified omnibus tests",
    "Near-eye F1/F3 are primary; chest tests are complementary mirrors. Each omnibus p is unadjusted and bold only at 0.050.",
    groupname_col = "Placement"
  ) |>
  fmt_markdown(columns = p)
```

The interaction is not folded into the primary additive mean model because it
answers a different question and requires many more supported coefficients.
The category omnibus asks whether the site-standardized category pattern is
flat. The interaction gate asks whether that category pattern varies by site.
Separating them preserves the hierarchy and prevents unsupported site cells
from determining whether the pooled category question can be asked.

## Site heterogeneity and sparse cells

The effect-blind architecture gate passed before interaction estimates or
interaction p-values were emitted. Near-eye used the literal seven-category
`site * light_source` formula. Chest had one structural cell absence and used
the algebraically equivalent observed-cell parameterization. The five-category
fallback was not invoked.

```{r}
#| label: tbl-interaction-gate

gate |>
  transmute(
    Placement = .data$placement,
    Architecture = .data$architecture,
    Observations = .data$observations,
    `Design rank / columns` = paste0(.data$design_rank, " / ", .data$design_columns),
    Restrictions = .data$restrictions,
    `Restriction covariance condition` = sprintf(
      "%.1f",
      .data$restriction_covariance_condition_number
    ),
    `Largest cluster score share` = sprintf(
      "%.3f",
      .data$maximum_cluster_score_share
    ),
    `Largest cluster leverage share` = sprintf(
      "%.3f",
      .data$maximum_cluster_leverage_share
    ),
    Passed = .data$gate_pass
  ) |>
  audit_gt(
    "Interaction architecture gate",
    "No interaction estimate, p-value, or claim was inspected to select the architecture"
  ) |>
  fmt_integer(columns = c(Observations, Restrictions), use_seps = TRUE)
```

```{r}
#| label: tbl-site-support-summary

site_context |>
  group_by(.data$placement, .data$category_order, .data$short_label) |>
  summarise(
    `Site rows` = n(),
    Estimable = sum(.data$reporting_status == "ESTIMABLE"),
    `Support-non-estimable` = sum(
      .data$reporting_status == "SUPPORT_NON_ESTIMABLE"
    ),
    `Other non-estimable / not applicable` = sum(
      !.data$reporting_status %in% c("ESTIMABLE", "SUPPORT_NON_ESTIMABLE")
    ),
    `BH-significant site deviations` = sum(
      .data$site_deviation_p_adjusted <= 0.05,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  mutate(Placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$Placement, .data$category_order) |>
  select(Placement, Category = .data$short_label, `Site rows`, Estimable,
    `Support-non-estimable`, `Other non-estimable / not applicable`,
    `BH-significant site deviations`) |>
  audit_gt(
    "Site-specific estimability and heterogeneity",
    "F4 uses one BH family across supported near-eye site deviations; chest is the complementary mirror.",
    groupname_col = "Placement"
  ) |>
  fmt_integer(columns = where(is.numeric), use_seps = TRUE)
```

Across both placements, 96 of 126 site-category rows provide supported
site-specific deviations; 17 are support-non-estimable, one is a structural
cell absence, five cannot form their site-standardized category comparison,
and seven are not applicable. Thirty-seven supported site deviations pass
their stated BH label. Unsupported rows remain visible as crosses rather than
being silently dropped.

```{r}
#| label: fig-near-eye-site-context
#| fig-cap: "Near-eye site-specific expected one-hour melEDI from the accepted interaction model. In each light-source panel, the dashed vertical line is the site-standardized geometric mean with equal site contributions on the fitted log scale. Filled circles identify site deviations with BH-adjusted p < 0.05 in the 53-member H03-F4 family; open circles do not pass that label. Crosses at zero identify prespecified support failures and are not zero-valued estimates."
#| fig-alt: "Seven-facet horizontal interval plot of model-estimated near-eye melEDI by submitted site. Each panel contains a dashed vertical site-standardized category mean. Filled site-coloured points mark BH-significant deviations from that mean; open points mark other supported estimates, and crosses at zero mark unsupported contrasts."
#| out-width: "100%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_near_eye_site_context_estimates.png"
))
```

## Diagnostics and influence

### Numerical, distributional, zero-mass, and temporal checks

```{r}
#| label: tbl-primary-diagnostics

diagnostics |>
  filter(.data$scenario_id == "primary_dataset") |>
  transmute(
    Placement = .data$placement,
    Converged = .data$converged,
    `Full rank` = .data$full_rank,
    `Robust covariance PD` = .data$covariance_positive_definite,
    `Covariance condition` = sprintf("%.1f", .data$covariance_condition_number),
    `Exact-zero fraction` = sprintf("%.3f", .data$exact_zero_fraction),
    `|residual|-fitted Spearman` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `Boundary-aware lag-1 residual r` = sprintf(
      "%.3f",
      .data$residual_lag1_correlation
    ),
    `Largest cluster score share` = sprintf(
      "%.3f",
      .data$maximum_cluster_score_share
    ),
    `Largest cluster leverage share` = sprintf(
      "%.3f",
      .data$maximum_cluster_leverage_share
    )
  ) |>
  audit_gt(
    "Primary-model diagnostics",
    "PD = positive definite. Dependence and heteroskedasticity are interpreted below, not treated as pass/fail decorations."
  )
```

Both fits converged without warnings, have full design rank, and have finite
positive-definite participant-cluster covariances. There is no mixed-model
singularity question because these are GLMs. Exact zeros constitute about
27.7% of either outcome. Absolute residual magnitude remains strongly related
to the fitted mean, and boundary-aware lag-1 residual correlation is 0.478
near eye and 0.615 at chest. Participant clustering permits arbitrary
within-participant dependence in the sandwich covariance, but it does not
make these residual patterns disappear or validate the working distribution.

```{r}
#| label: tbl-zero-mass

zero_mass |>
  filter(.data$scope == "overall") |>
  transmute(
    Placement = .data$placement,
    Observations = .data$observations,
    `Observed zero fraction` = sprintf("%.3f", .data$observed_zero_fraction),
    `Working Tweedie zero fraction` = sprintf(
      "%.3f",
      .data$working_expected_zero_fraction
    ),
    `Observed minus working` = sprintf(
      "%+.3f",
      .data$observed_minus_working_fraction
    ),
    `Zero Brier score` = sprintf("%.3f", .data$zero_brier_score)
  ) |>
  audit_gt(
    "Working zero mass",
    paste(
      "The compound-Poisson expression is a diagnostic implied by the fitted",
      "mean, power, and dispersion. The quasi-likelihood model does not fit a",
      "zero-mass likelihood."
    )
  ) |>
  fmt_integer(columns = Observations, use_seps = TRUE)
```

The implied working Tweedie zero fractions are 0.852 near eye and 0.948 at
chest, far above the observed 0.278 and 0.277. This is a clear distributional
miscalibration. It rules out interpreting the fitted family as a realistic
generative distribution. The estimates remain usable as a log-mean model
with participant-cluster-robust covariance if the site-and-category mean
structure is the target; efficiency and small-sample calibration may be
imperfect.

```{r}
#| label: fig-primary-residuals
#| fig-cap: "Primary additive quasi-Tweedie diagnostic plots. Panel A overlays every participant-hour Pearson residual with binned means and interquartile ranges; Panel B shows boundary-aware residual correlations across one- to six-hour lags; Panel C compares observed and working-Tweedie zero fractions across fitted-mean deciles. Plot transformations are display-only."
#| fig-alt: "Three-part diagnostic figure for near-eye and chest additive quasi-Tweedie fits. The first part shows a dense cloud of individual Pearson residual dots plus binned trends, with a long positive tail. The second shows positive residual autocorrelation over successive within-sequence hourly lags. The third shows observed zero fractions well below the working Tweedie zero fractions."
#| out-width: "100%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_primary_residual_diagnostics.png"
))
```

### Descriptive fixed-effect R² and bounded Gaussian comparison

An ordinary mixed-model marginal/conditional R² is not defined for the
primary analysis because the fitted model contains no random effects.
Participant clustering changes the covariance used for inference, not the
fitted values. The descriptive summary is now calculated for the accepted
heterogeneity architecture: the literal `site * light_source` model near eye
and its rank-complete observed-cell parameterization,
`0 + site_source_cell`, at chest, where one cell is structurally absent. It
uses the intercept-only, site-only, category-only, additive, and accepted
full-heterogeneity fits:

- overall R² uses the full interaction model;
- the interaction partial R² compares the full model with its hierarchical
  additive reduction; and
- the hierarchy-respecting allocation averages the two entry orders for the
  site and category main effects, then adds the interaction only after both
  main effects. The three allocations sum exactly to overall R².

```{r}
#| label: tbl-primary-glm-r2

glm_r_squared |>
  filter(.data$model_id == "quasi_tweedie_log") |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Definition = case_when(
      .data$loss_basis == "model_deviance" ~
        "Working quasi-Tweedie deviance",
      .data$weighting == "participant_balanced" ~
        "Participant-balanced squared error",
      TRUE ~ "Participant-hour squared error"
    ),
    Definition = factor(
      .data$Definition,
      levels = c(
        "Working quasi-Tweedie deviance",
        "Participant-hour squared error",
        "Participant-balanced squared error"
      )
    ),
    Overall = sprintf("%.3f", .data$overall_r_squared),
    `Category partial | site` = sprintf(
      "%.3f",
      .data$category_partial_r_squared_conditional_on_site
    ),
    `Site partial | category` = sprintf(
      "%.3f",
      .data$site_partial_r_squared_conditional_on_category
    ),
    `Interaction partial | additive` = sprintf(
      "%.3f",
      .data$interaction_partial_r_squared_conditional_on_additive
    ),
    `Category allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$category_shapley_r_squared,
      .data$category_shapley_share_percent
    ),
    `Site allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$site_shapley_r_squared,
      .data$site_shapley_share_percent
    ),
    `Interaction allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$interaction_r_squared,
      .data$interaction_share_percent
    )
  ) |>
  arrange(.data$Placement, .data$Definition) |>
  select(
    Placement,
    Definition,
    Overall,
    `Category partial | site`,
    `Site partial | category`,
    `Interaction partial | additive`,
    `Category allocation`,
    `Site allocation`,
    `Interaction allocation`
  ) |>
  audit_gt(
    "Descriptive R² for the accepted fixed-effect heterogeneity model",
    paste(
      "Allocations are R² units with the percentage of total R² in",
      "parentheses. Point estimates only; no cluster-bootstrap intervals",
      "or additional effect tests were computed."
    ),
    groupname_col = "Placement"
  )
```

For the near-eye primary model, working quasi-deviance R² is
`r sprintf("%.3f", glm_r_squared$overall_r_squared[glm_r_squared$placement == "Near-eye" & glm_r_squared$model_id == "quasi_tweedie_log" & glm_r_squared$loss_basis == "model_deviance"])`;
the participant-hour and participant-balanced squared-error R² values are
`r sprintf("%.3f", glm_r_squared$overall_r_squared[glm_r_squared$placement == "Near-eye" & glm_r_squared$model_id == "quasi_tweedie_log" & glm_r_squared$loss_basis == "squared_error" & glm_r_squared$weighting == "participant_hour_weighted"])`
and
`r sprintf("%.3f", glm_r_squared$overall_r_squared[glm_r_squared$placement == "Near-eye" & glm_r_squared$model_id == "quasi_tweedie_log" & glm_r_squared$loss_basis == "squared_error" & glm_r_squared$weighting == "participant_balanced"])`.
The category block accounts for most of the fitted gain under each definition,
while the interaction contribution is now shown explicitly. The allocations
are descriptive associations and not causal variance contributions. Different
R² definitions answer different scale-and-weighting questions and should not
be merged into one headline value.

The Gaussian check remains a separate comparison of the additive
category-association architecture across response families; it is not an
interaction-selection analysis. It used the same exact fitted samples and was
deterministic and bounded: five participant-blocked folds,
stratified within site, with every participant assigned to exactly one test
fold. It involved 30 fits and no resampling simulation.

```{r}
#| label: tbl-glm-candidate-formulas
#| echo: true

glm_candidate_registry |>
  transmute(
    Candidate = .data$model_label,
    `Exact Wilkinson formula` = .data$exact_formula,
    Family = .data$family_label,
    `Outcome target` = .data$outcome_target
  ) |>
  audit_gt(
    "Bounded GLM comparison candidates",
    "The transformed Gaussian changes the fitted outcome target."
  ) |>
  cols_width(
    Candidate ~ pct(20),
    `Exact Wilkinson formula` ~ pct(27),
    Family ~ pct(22),
    `Outcome target` ~ pct(31)
  )
```

```{r}
#| label: tbl-glm-candidate-cross-validation

glm_cross_validation |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Candidate = factor(
      .data$model_label,
      levels = glm_candidate_registry$model_label
    ),
    `Observed mean, lx` = sprintf("%.1f", .data$observed_mean_lx),
    `Predicted mean, lx` = sprintf("%.1f", .data$predicted_mean_lx),
    `Raw RMSE, lx` = sprintf("%.1f", .data$raw_rmse_lx),
    `Raw MAE, lx` = sprintf("%.1f", .data$raw_mae_lx),
    `Fold-weighted raw R²` = sprintf(
      "%.3f",
      .data$raw_efron_r_squared_fold_weighted
    ),
    `log10 RMSE` = sprintf("%.3f", .data$log10_rmse),
    `Negative predictions` = sprintf(
      "%.1f%%",
      100 * .data$negative_prediction_fraction
    )
  ) |>
  arrange(.data$Placement, .data$Candidate) |>
  select(
    Placement,
    Candidate,
    `Observed mean, lx`,
    `Predicted mean, lx`,
    `Raw RMSE, lx`,
    `Raw MAE, lx`,
    `Fold-weighted raw R²`,
    `log10 RMSE`,
    `Negative predictions`
  ) |>
  audit_gt(
    "Participant-blocked out-of-fold prediction check",
    paste(
      "All candidates converged and were full rank in all five folds.",
      "No uncertainty interval or model-selection p-value is attached."
    ),
    groupname_col = "Placement"
  )
```

The raw-scale Gaussian did not fit better for the primary mean target. The
quasi-Tweedie model reduced held-out raw RMSE relative to that Gaussian by
`r sprintf("%.1f", 100 * (near_gaussian_cv$raw_rmse_lx - near_quasi_cv$raw_rmse_lx) / near_gaussian_cv$raw_rmse_lx)`%
near eye and
`r sprintf("%.1f", 100 * (chest_gaussian_cv$raw_rmse_lx - chest_quasi_cv$raw_rmse_lx) / chest_gaussian_cv$raw_rmse_lx)`%
at chest, and also had lower raw MAE. More importantly, the identity-Gaussian
model produced negative held-out melEDI predictions for
`r sprintf("%.1f", 100 * near_gaussian_cv$negative_prediction_fraction)`%
of near-eye and
`r sprintf("%.1f", 100 * chest_gaussian_cv$negative_prediction_fraction)`%
of chest hours; the quasi-Tweedie log link produced none.

The log-Gaussian candidate has the lowest log-scale RMSE and raw MAE, but it
fits a different target. Directly back-transforming its expected log outcome
gives a geometric-scale center, not the conditional arithmetic mean of hourly
melEDI: its held-out predicted mean was
`r sprintf("%.1f", near_log_gaussian_cv$predicted_mean_lx)` lx versus
`r sprintf("%.1f", near_log_gaussian_cv$observed_mean_lx)` lx near eye, and
`r sprintf("%.1f", chest_log_gaussian_cv$predicted_mean_lx)` lx versus
`r sprintf("%.1f", chest_log_gaussian_cv$observed_mean_lx)` lx at chest.
Consequently it is not a drop-in replacement for the primary model. AIC cannot
adjudicate these candidates because the primary is quasi-likelihood and the
transformed Gaussian uses another response scale. This bounded check supports
retaining the quasi-Tweedie log-mean model while preserving the already stated
distributional and temporal limitations.

### Influence

All nine near-eye sites, all eight chest sites, and the five highest-score
participants at each placement were deleted in turn. None of the 27 deletion
refits changed an omnibus significance decision. Of 176 estimable category
ratios in the deletion audit, 173 retained the full-data ratio inside the
deletion-refit interval. The exceptions were chest sleep darkness after
deleting KNUST, chest external light during sleep after deleting TUM, and
chest sleep darkness after deleting participant KNUST_S007. These are
localized complementary-result warnings rather than omnibus reversals.

The largest participant cluster accounts for 0.334 of the near-eye score
energy and 0.393 of the chest score energy, below the predeclared 0.500 gate
but large enough that the deletion evidence should accompany the result.

## Registered sensitivities

The primary could be interpreted as time-sensitive because it inherits the
accepted gap-aware one-hour aggregation. The **gap-timing-unaware dataset** is
the otherwise comparable prepared input that does not preserve explicit
timing of within-hour gaps. After this definition, the main input is called
the primary dataset.

```{r}
#| label: tbl-sensitivity-summary

sensitivity_counts <- sensitivity |>
  group_by(.data$run_id, .data$placement, .data$scenario_id) |>
  summarise(
    Stable = sum(.data$stability == "stable"),
    `Support-non-estimable` = sum(.data$stability == "support_non_estimable"),
    `Direction shifts` = sum(.data$stability == "direction_shift"),
    `Magnitude shifts` = sum(.data$stability == "magnitude_shift"),
    .groups = "drop"
  )

sensitivity_counts |>
  left_join(
    samples |>
      select(.data$run_id, .data$observations, .data$participants,
        .data$participant_days, .data$sites),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  left_join(
    sensitivity_omnibus |>
      select(.data$run_id, omnibus_p = .data$p_raw),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  mutate(
    Scenario = recode(
      .data$scenario_id,
      paired_common_sample = "Paired/common sample",
      gap_timing_unaware_dataset = "Gap timing unaware",
      boundary_hours_excluded = "Boundary hours excluded",
      unsupported_cells_excluded = "Unsupported cells excluded",
      working_power_1_30 = "Working power 1.30",
      working_power_1_80 = "Working power 1.80",
      mundlak_within_between = "Mundlak within/between"
    ),
    `Omnibus p` = format_p_cell(.data$omnibus_p, .data$omnibus_p <= 0.05)
  ) |>
  select(
    Placement = .data$placement,
    Scenario,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Omnibus p`,
    Stable,
    `Support-non-estimable`,
    `Direction shifts`,
    `Magnitude shifts`
  ) |>
  audit_gt(
    "Sensitivity overview",
    "Stability counts cover the seven category rows; each omnibus p is bold at its stated unadjusted 0.050 label."
  ) |>
  fmt_integer(
    columns = c(Hours, Participants, `Participant-days`, Sites, Stable,
      `Support-non-estimable`, `Direction shifts`, `Magnitude shifts`),
    use_seps = TRUE
  ) |>
  fmt_markdown(columns = `Omnibus p`)
```

The gap timing unaware result is stable for all seven categories at both
placements. Boundary exclusion, removal of unsupported site cells, and the
working-power alternatives leave every omnibus decision unchanged. The
Mundlak audit moves the external-light-during-sleep ratio by 25.8% near eye
and 34.0% at chest while retaining its direction and containing the primary
ratio inside the sensitivity interval. Outdoor electric light is imprecise:
it is support-non-estimable in three reduced samples and crosses the null
direction under the 1.80 power and near-eye Mundlak alternatives; none of
those alternatives establishes a non-null outdoor-electric contrast.

### Paired/common-sample placement comparison

The paired comparison uses 14,103 identical participant-hours from 111
participants and 631 participant-days at each placement. Outdoor electric
light has only 162 paired hours and fails its prespecified support rule, so it
is not plotted. The remaining five non-reference ratios have the same broad
direction at both placements. The diagonal is a visual concordance reference,
not an equivalence test.

```{r}
#| label: fig-paired-placement
#| fig-cap: "Paired/common-sample near-eye and chest category ratios. The identity line is not an equivalence boundary."
#| fig-alt: "Log-log scatter plot comparing near-eye and chest ratios to indoor electric light on identical participant-hours. Indoor and outdoor daylight are above one at both placements; display and both sleep-related categories are below one. Confidence intervals are shown in both directions."
#| out-width: "72%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_paired_placement_comparison.png"
))
```

## V0 reconstruction and corrections

V0 is reconstructed only as a bridge. It is not the approved analysis and it
is not used to choose the Stage 2 result.

```{r}
#| label: tbl-v0-denominators

v0_comparison |>
  transmute(
    Placement = .data$placement,
    Implementation = .data$implementation,
    `Participant-hours` = .data$observations,
    Categories = .data$categories,
    `Test definition` = .data$test_definition
  ) |>
  audit_gt(
    "V0 and approved-analysis denominators",
    "Frozen V0 counts are preserved; the exact bridge reruns the V0 formula on currently recoverable five-category rows"
  ) |>
  fmt_integer(columns = c(`Participant-hours`, Categories), use_seps = TRUE)
```

The exact V0 bridge converged, had positive-definite Hessians, and was
non-singular at both placements. Its likelihood-ratio test is
`r format_p(v0_tests$p_raw[1])` near eye and
`r format_p(v0_tests$p_raw[2])` at chest, but that test jointly adds four
category main effects and all corresponding category-by-site interactions.
It is not the six-restriction H03-F1 category omnibus.

Stage 2 corrects the consequential V0 issues explicitly:

1. V0 entered five categories; the approved analysis audits and estimates all
   seven registered categories.
2. Frozen V0 reported 16,774 near-eye and 18,391 chest rows. The exact current
   V0 bridge contains 16,879 and 18,534 rows, whereas the approved analysis
   contains 17,935 and 19,512 rows. Each number is labelled as participant-
   hours, not participants.
3. V0 combined category association and site heterogeneity in one likelihood-
   ratio test. Stage 2 reports separate F1 and F3 tests.
4. V0's pooled `emmeans` contrasts used the Dunnett-X default but were labelled
   FDR-adjusted in the table. Stage 2 uses one explicit six-member BH family.
5. V0 removed confidence-limit columns. Stage 2 reports participant-cluster-
   robust 95% intervals for every estimable practical contrast.
6. The copied chest interpretation is not retained; chest estimates are
   calculated and interpreted from the chest fit.

## Exploratory time-of-day context

::: {.callout-note title="Exploratory model — separate from preregistered inference"}
The GAM/GAMM context does not use indoor electric light as a reference curve.
It contains one global cyclic time-of-day smooth plus sum-to-zero light-source
smooth deviations around that global curve, sum-to-zero site deviations,
participant factor smooths, participant-day random intercepts, and a
boundary-aware AR(1) correction. Its R² and variance allocations are
in-sample descriptive quantities on `log10(melEDI + 0.1 lx)`, not additions to
the primary p-value families.
:::

```{r}
#| label: fig-temporal-near-eye
#| fig-cap: "Exploratory near-eye global-plus-light-source time-of-day curves, category deviations, and available participant-hours. In panels A and B, filled circles mark observed category-hours that pass the local-support rule; open circles mark nonzero but locally sparse support: fewer than 20 participant-hours, 5 participants, or 3 sites. Light-grey regions in panels A and B and crosses in panel C mean zero observations, with no curve or ribbon drawn."
#| fig-alt: "Three-part near-eye figure. The first section shows seven light-source curves around a dashed global daily curve; the second shows each category's sum-to-zero deviation, and the third shows participant-hour counts. Filled circles in the first two sections mark adequately supported observed hours; larger open circles mark locally sparse observed hours. Grey regions align with crosses in the count panel and mark zero observations."
#| out-width: "100%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_temporal_near_eye.png"
))
```

```{r}
#| label: fig-temporal-chest
#| fig-cap: "Exploratory chest global-plus-light-source time-of-day curves, category deviations, and available participant-hours. In panels A and B, filled circles mark observed category-hours that pass the local-support rule; open circles mark nonzero but locally sparse support: fewer than 20 participant-hours, 5 participants, or 3 sites. Light-grey regions in panels A and B and crosses in panel C mean zero observations, with no curve or ribbon drawn."
#| fig-alt: "Three-part chest figure with seven global-plus-category time-of-day curves, their sum-to-zero category deviations, and participant-hour counts. Filled circles in the first two sections mark adequately supported observed hours; larger open circles mark locally sparse observed hours. Grey regions align with crosses in the count panel and mark zero observations."
#| out-width: "100%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_temporal_chest.png"
))
```

### R² and variance decomposition

```{r}
#| label: tbl-temporal-r2

temporal_summary |>
  select(
    .data$placement,
    .data$observations,
    .data$participants,
    .data$participant_days,
    .data$rho,
    .data$adjusted_r_squared,
    .data$deviance_explained
  ) |>
  left_join(
    temporal_r2 |>
      select(.data$placement, balanced_r_squared = .data$r_squared),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `AR rho` = sprintf("%.3f", .data$rho),
    `Adjusted R²` = sprintf("%.3f", .data$adjusted_r_squared),
    `Deviance explained` = sprintf("%.3f", .data$deviance_explained),
    `Site-standardized participant-balanced R²` = sprintf(
      "%.3f",
      .data$balanced_r_squared
    )
  ) |>
  audit_gt(
    "Exploratory temporal model fit",
    "R² values are descriptive in-sample transformed-response summaries"
  ) |>
  fmt_integer(columns = c(Hours, Participants, `Participant-days`), use_seps = TRUE)
```

```{r}
#| label: tbl-temporal-partition

temporal_partition |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Component = recode(
      .data$group,
      global_time = "Global time of day",
      light_source_deviations = "Light-source deviations",
      site_deviations = "Site deviations",
      participant_curves = "Participant curves",
      participant_day_shifts = "Participant-day shifts"
    ),
    `Shapley share` = sprintf("%.1f%%", 100 * .data$shapley_share),
    `Unique partial share` = sprintf(
      "%.1f%%",
      100 * .data$partial_unique_share
    )
  ) |>
  select(Placement, Component, `Shapley share`, `Unique partial share`) |>
  audit_gt(
    "Point variance allocation",
    paste(
      "Exact conditional Shapley allocation of fitted linear-predictor",
      "variance under site-standardized, participant-balanced weights;",
      "no simulation interval was computed."
    ),
    groupname_col = "Placement"
  )
```

The global daily curve receives 67.4% of the near-eye and 64.2% of the chest
point allocation. Light-source deviations receive 16.5% and 20.4%, site
deviations 4.9% and 4.4%, participant curves 10.0% and 9.1%, and participant-
day shifts 1.2% and 1.9%. These components share covariance; the Shapley rule
allocates that covariance exactly. They are not independent causal
contributions.

### Temporal diagnostics and limitations

```{r}
#| label: tbl-temporal-diagnostics

temporal_summary |>
  transmute(
    Placement = .data$placement,
    Convergence = .data$convergence,
    `Optimizer iterations` = .data$smoothing_optimizer_iterations,
    `Maximum |gradient|` = formatC(
      .data$smoothing_gradient_maximum_absolute,
      format = "e",
      digits = 2
    ),
    `Smoothing Hessian minimum eigenvalue` = formatC(
      .data$smoothing_hessian_minimum_eigenvalue,
      format = "e",
      digits = 2
    ),
    `Smoothing Hessian PD` = .data$smoothing_hessian_positive_definite,
    `Residual lag-1 r` = sprintf("%.3f", .data$standardized_residual_lag1),
    `Residual excess kurtosis` = sprintf(
      "%.2f",
      .data$residual_excess_kurtosis
    )
  ) |>
  audit_gt(
    "Exploratory temporal diagnostics",
    "Lme-style singularity is not defined for penalized bam; the smoothing-parameter Hessian and penalty components are inspected instead"
  ) |>
  fmt_integer(columns = `Optimizer iterations`)
```

Both fits report full `mgcv` convergence. The near-eye smoothing Hessian is
positive definite. The chest Hessian has one slightly non-positive direction
(minimum eigenvalue -0.000123), so chest smoothing-parameter uncertainty is
less secure. Standard `mgcv` full concurvity could not be computed near eye
because its LAPACK SVD failed; the clearly labelled fitted-term fallback has a
maximum multiple R² of 0.265. Chest full concurvity is available and is high
for the global time and site components, indicating that their separate
attribution is unstable even though the summed prediction is estimable.

The deterministic, zero-resample basis check gives global time k-indices of
0.978 near eye and 0.970 at chest. The global smooth uses 9.69 and 9.56 of ten
available effective degrees of freedom, so basis capacity is close to its
ceiling. No permutation p-value was computed. Standardized residual lag-1
correlation falls to 0.096 and 0.091 after AR adjustment, but residual tails
remain heavier than Gaussian.

The inherited H02 specification uses a cyclic global curve but default thin-
plate marginal bases for the category and site sum-to-zero deviations. The
sum-to-zero constraints pass at less than `r sprintf("%.1e", max(temporal_constraints$maximum_absolute_sum))`,
but the non-cyclic deviations allow 24-hour versus 0-hour endpoint differences
up to `r sprintf("%.3f", max(temporal_midnight$endpoint_difference_abs))` on
the log10 scale. The temporal figures use 0.5–23.5 hour centers and make no
curve-wide inferential claim.

At the author's request, a bounded same-row comparison replaced both `sz`
marginal bases with cyclic bases. Both candidate fits used Gaussian `bam`,
fREML, the same estimated-then-fixed AR(1) procedure, and no diagnostic
simulation or resampling.

```{r}
#| label: tbl-temporal-basis-formulas

temporal_basis_registry |>
  transmute(
    Candidate = .data$model_label,
    `Exact Wilkinson formula supplied to bam` = .data$exact_formula
  ) |>
  audit_gt(
    "Temporal-basis candidates",
    "Only the light-source and site sz marginal bases differ"
  )
```

```{r}
#| label: tbl-temporal-basis-comparison

temporal_basis_fit |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Candidate = factor(
      .data$model_label,
      levels = c("Inherited thin-plate sz", "Fully cyclic sz")
    )
  ) |>
  transmute(
    Placement,
    Candidate,
    AIC = sprintf("%.1f", .data$aic),
    `fREML score` = sprintf("%.1f", .data$f_reml_score),
    `Adjusted R²` = sprintf("%.4f", .data$adjusted_r_squared),
    `Residual scale` = sprintf("%.4f", .data$residual_scale),
    `Residual lag-1 r` = sprintf("%.3f", .data$standardized_residual_lag1),
    `Maximum category midnight gap` = formatC(
      .data$maximum_category_endpoint_difference_log10,
      format = "e",
      digits = 2
    ),
    `Maximum site midnight gap` = formatC(
      .data$maximum_site_endpoint_difference_log10,
      format = "e",
      digits = 2
    ),
    `Smoothing Hessian PD` = .data$smoothing_hessian_positive_definite
  ) |>
  arrange(.data$Placement, .data$Candidate) |>
  audit_gt(
    "Bounded temporal-basis comparison",
    paste(
      "Identical fitted rows; lower AIC and residual scale and higher",
      "adjusted R² favour fit, while midnight continuity is a structural",
      "criterion. The fREML score is shown but not used alone for selection",
      "because the smooth design spaces differ."
    ),
    groupname_col = "Placement"
  )
```

```{r}
#| label: fig-temporal-basis-diagnostics
#| fig-cap: "Diagnostic comparison of the inherited thin-plate sz and fully cyclic sz temporal models on identical fitted participant-hours. Panels show sampled normal-quantile residual points, binned residuals versus fitted values, boundary-aware residual autocorrelation, and the maximum 24:00-versus-00:00 discontinuity for light-source and site deviations. No diagnostic simulation or resampling was run."
#| fig-alt: "Four-panel diagnostic comparison for near-eye and chest models. Residual quantile and fitted-value patterns are nearly superimposed for inherited and cyclic bases, as are one- to six-hour residual correlations. The inherited basis has visible midnight discontinuities around 0.3 to 0.4 log10 units; the fully cyclic basis has gaps effectively equal to zero."
#| out-width: "100%"

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H03/H03_temporal_basis_diagnostic_comparison.png"
))
```

The fully cyclic candidate is not uniformly a better empirical fit. Relative
to the inherited model, AIC is higher by
`r paste(sprintf("%.1f", temporal_basis_delta$cyclic_minus_inherited_aic), collapse = " and ")`
points, adjusted R² is lower by
`r paste(sprintf("%.4f", abs(temporal_basis_delta$cyclic_minus_inherited_adjusted_r_squared)), collapse = " and ")`,
and residual scale and lag-1 correlation are slightly higher. Its fREML score
is numerically lower, but that criterion is not treated as decisive across
different smooth design spaces. The cyclic candidate does, however, reduce
both category and site midnight gaps to below
`r sprintf("%.1e", max(temporal_basis_fit$maximum_category_endpoint_difference_log10[temporal_basis_fit$model_id == "cyclic_sz"], temporal_basis_fit$maximum_site_endpoint_difference_log10[temporal_basis_fit$model_id == "cyclic_sz"]))`
and changes the chest smoothing Hessian from slightly non-positive to positive
definite. Figures 5 and 6 therefore remain the inherited specification at this
Stage 2 gate; choosing structural cyclic continuity over its small empirical-
fit cost remains an explicit author decision for Stage 3.

## Computation gate

```{r}
#| label: tbl-deferred-computation

deferred |>
  transmute(
    Component = .data$component,
    Status = .data$status,
    Reason = .data$reason,
    `Pilot / production state` = .data$pilot_or_production
  ) |>
  audit_gt(
    "Deferred computation",
    "No production bootstrap, simultaneous-band simulation, or variance-allocation uncertainty simulation was run"
  )
```

The Stage 1 site-bootstrap pilot remains the only bootstrap work. No
production bootstrap was run. The deterministic GAM basis check uses zero
resampling replicates, and the point Shapley allocation uses no simulation.
Any production uncertainty computation still requires explicit approval.

## Reproducibility and author gate

All scientific computation used R 4.6.1 and the existing project library.
Input hashes were verified before each fit. Model objects, exact model frames,
diagnostic tables, source CSVs, PNG/PDF/SVG figures, and run manifests are
stored under the H03 artifact directories. This report reads those artifacts;
it does not refit a model.

Stage 2 is now complete for author review. Stage 3 must not start until the
author approves the inferential interpretation and decides whether the
exploratory temporal model's inherited non-cyclic deviation bases and chest
smoothing-Hessian/concurvity limitations are acceptable, should be amended,
or should keep the temporal result out of the standalone report.
