---
title: "H03 analysis preparation and provenance"
subtitle: "From verified one-hour records and reported light sources to site-average exposure contrasts"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
    lightbox: true
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-h03-preparation
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(knitr)
  library(readr)
  library(scales)
  library(tibble)
  library(tidyr)
})

root_candidates <- unique(c(
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(
    file.path(getwd(), "../../.."),
    winslash = "/",
    mustWork = FALSE
  )
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(
        path,
        "scripts/hypotheses/H03/h03_contract.R"
      ))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(root.dir = root)

source(file.path(root, "scripts/pipeline/paths_io.R"))

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H03 preparation requires R 4.6.1", call. = FALSE)
}

read_h03_csv <- function(relative_path) {
  readr::read_csv(file.path(root, relative_path), show_col_types = FALSE)
}

h03_gt <- function(table, font_size = 12) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

format_p <- function(value, significant = FALSE) {
  displayed <- ifelse(
    is.na(value),
    "—",
    ifelse(value < 0.001, "<0.001", sprintf("%.3f", value))
  )
  ifelse(significant & !is.na(value), paste0("**", displayed, "**"), displayed)
}

placement_colors <- c(
  "Near-eye" = "#0072B2",
  "Chest" = "#D55E00"
)

input_audit <- read_h03_csv(
  "artifacts/06_model_data/H03/H03_input_audit.csv"
)
frame_index <- read_h03_csv(
  "artifacts/06_model_data/H03/H03_model_frame_index.csv"
)
multiplicity_registry <- read_h03_csv(
  "artifacts/06_model_data/H03/H03_multiplicity_registry.csv"
)
frame_integrity <- read_h03_csv(
  "artifacts/11_source_data/H03/H03_preparation_frame_integrity.csv"
)
zero_summary <- read_h03_csv(
  "artifacts/11_source_data/H03/H03_preparation_zero_summary.csv"
)
response_distribution <- read_h03_csv(
  paste0(
    "artifacts/11_source_data/H03/",
    "H03_preparation_positive_response_distribution.csv"
  )
)
category_support <- read_h03_csv(
  "artifacts/11_source_data/H03/H03_preparation_category_support.csv"
)
site_category_support <- read_h03_csv(
  paste0(
    "artifacts/11_source_data/H03/",
    "H03_preparation_site_category_support.csv"
  )
)
clock_support <- read_h03_csv(
  paste0(
    "artifacts/11_source_data/H03/",
    "H03_preparation_clock_category_support.csv"
  )
)
participant_day_support <- read_h03_csv(
  paste0(
    "artifacts/11_source_data/H03/",
    "H03_preparation_participant_day_support.csv"
  )
)
primary_tests <- read_h03_csv(
  "artifacts/09_tables/H03/H03_primary_omnibus_tests.csv"
)
interaction_gate <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_interaction_architecture_gate.csv"
)
diagnostics <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_model_diagnostics.csv"
)
zero_diagnostics <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_zero_mass_diagnostics.csv"
)
residual_acf <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_primary_residual_acf.csv"
)
influence_registry <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_influence_refit_registry.csv"
)
sensitivity_tests <- read_h03_csv(
  "artifacts/09_tables/H03/H03_sensitivity_omnibus_tests.csv"
)
glm_r_squared <- read_h03_csv(
  paste0(
    "artifacts/09_tables/H03/",
    "H03_glm_r_squared_and_effect_partition.csv"
  )
)
temporal_summary <- read_h03_csv(
  "artifacts/09_tables/H03/H03_reader_temporal_model_summary.csv"
)
temporal_r_squared <- read_h03_csv(
  "artifacts/09_tables/H03/H03_reader_temporal_weighted_r_squared.csv"
)
temporal_allocation <- read_h03_csv(
  paste0(
    "artifacts/09_tables/H03/",
    "H03_reader_temporal_variance_allocation.csv"
  )
)
latitude_diagnostics <- read_h03_csv(
  paste0(
    "artifacts/08_diagnostics/H03/",
    "H03_reader_latitude_model_diagnostics.csv"
  )
)
execution_environment <- read_h03_csv(
  "artifacts/12_manifests/H03/H03_execution_environment.csv"
)
temporal_environment <- read_h03_csv(
  paste0(
    "artifacts/12_manifests/H03/",
    "H03_temporal_execution_environment.csv"
  )
)
participant_random_intercept_summary <- read_h03_csv(
  paste0(
    "artifacts/09_tables/H03/",
    "H03_near_eye_participant_random_intercept_summary.csv"
  )
)
participant_random_intercept_shapley <- read_h03_csv(
  paste0(
    "artifacts/09_tables/H03/",
    "H03_near_eye_participant_random_intercept_marginal_r2_shapley.csv"
  )
)
participant_random_intercept_model_checks <- read_h03_csv(
  paste0(
    "artifacts/08_diagnostics/H03/",
    "H03_near_eye_participant_random_intercept_shapley_models.csv"
  )
)
participant_random_intercept_diagnostics <- read_h03_csv(
  paste0(
    "artifacts/08_diagnostics/H03/",
    "H03_near_eye_participant_random_intercept_diagnostics.csv"
  )
)
participant_random_intercept_environment <- read_h03_csv(
  paste0(
    "artifacts/12_manifests/H03/",
    "H03_near_eye_participant_random_intercept_environment.csv"
  )
)
participant_random_intercept_manifest <- read_h03_csv(
  paste0(
    "artifacts/12_manifests/H03/",
    "H03_near_eye_participant_random_intercept_manifest.csv"
  )
)

participant_random_intercept_expected_files <- c(
  "H03_near_eye_participant_random_intercept_assessment.rds",
  "H03_near_eye_participant_random_intercept_diagnostics.csv",
  "H03_near_eye_participant_random_intercept_shapley_models.csv",
  "H03_near_eye_participant_random_intercept_marginal_r2_shapley.csv",
  "H03_near_eye_participant_random_intercept_summary.csv",
  "H03_near_eye_participant_random_intercept_environment.csv"
)
participant_random_intercept_paths <- normalizePath(
  participant_random_intercept_manifest$path,
  winslash = "/",
  mustWork = TRUE
)
stopifnot(
  nrow(participant_random_intercept_manifest) == 6L,
  setequal(
    basename(participant_random_intercept_paths),
    participant_random_intercept_expected_files
  ),
  all(vapply(
    participant_random_intercept_paths,
    artifact_sha256,
    character(1)
  ) == participant_random_intercept_manifest$sha256),
  all(
    file.info(participant_random_intercept_paths)$size ==
      participant_random_intercept_manifest$bytes
  ),
  all(participant_random_intercept_manifest$r_version == "4.6.1")
)

model_frame_archive <- readRDS(file.path(
  root,
  "artifacts/06_model_data/H03/H03_model_frames.rds"
))
stopifnot(
  nrow(model_frame_archive$main$near_eye) == 17935L,
  nrow(model_frame_archive$main$chest) == 19512L
)
```

## About this analysis record

H03 asks whether all seven prespecified self-reported primary-light-source
categories are associated with measured one-hour personal light exposure.
The primary near-eye sensor position measures light close to the eyes. The
complementary chest sensor position measures light at the chest and is not a
measure of ocular exposure. The outcome is melanopic equivalent daylight
illuminance (melEDI). The central estimand is the site-average geometric
expected melEDI for each category and its ratio to indoor electric light. A
site-average estimate gives each included site equal weight on the fitted
log-mean scale; one back-transformation then returns the estimate to melEDI in
lux.

This page identifies the data and code that produced the result, explains why
each analytical operation was needed, and shows the recorded evidence used to
check it. The analysis is observational and does not identify causal effects,
health outcomes, or equivalence between measurement placements.

The [H03 results report](../../../notebooks/hypotheses/H03.qmd) presents the
scientific findings and its [linked registration-record summary](../../../notebooks/hypotheses/H03.qmd#h03-preregistration-deviations).
The metric and model-ready input definitions are also
documented in [Preparation 04](../../../notebooks/preparation/04_metric_derivation.qmd)
and [Preparation 06](../../../notebooks/preparation/06_model_ready_datasets.qmd).

::: {.callout-note title="Scope of this analysis record"}
This page reads accepted estimates, 95% confidence intervals (95% CIs),
omnibus tests, model checks, influence refits, sensitivity results,
exploratory smooths, and latitude results from their recorded files. Its
bounded display work does not
fit or refit a model, calculate predictions, re-estimate residual
autocorrelation, rerun a leave-one-site-out analysis, bootstrap, simulate
data, or recalculate a variance allocation.
:::

## Analysis map

```{mermaid}
flowchart TD
  A["Verified near-eye and chest<br/>one-hour geometric melEDI"] --> D["One-to-one local-hour join"]
  B["Normalized primary-light-source diary<br/>seven ordered categories"] --> D
  C["Shared country-coded site registry,<br/>order, and colours"] --> D
  D --> E["Eligibility, key, value,<br/>missingness and overlap checks"]
  E --> F["Primary model frames<br/>participant-hours within days"]
  F --> G["Additive quasi-Tweedie GLMs<br/>site plus light source"]
  F --> H["Category-by-site interaction<br/>supported observed cells"]
  G --> I["Indoor-electric ratios,<br/>95% CIs and omnibus test"]
  H --> J["Site-average category means,<br/>site deviations and within-site ratios"]
  G --> K["Residual, zero-mass,<br/>temporal and influence checks"]
  F --> L["Same-hour, gap, boundary,<br/>support, power and within/between checks"]
  F --> M["Exploratory nonlinear time-of-day GAMs<br/>and linear absolute-latitude GLMs"]
  I --> N["H03 results report"]
  J --> N
  K --> N
  L --> N
  M --> N
  O["Automated structural and identity checks"] -. verify .-> E
  O -. verify .-> F
  O -. verify .-> N
```

The arrows describe dependencies. They do not mean that this page repeats
model estimation.

## 1. Verified inputs

### File identity

Six inputs determine the current analysis frames and displays: two accepted
one-hour outcome files, the normalized diary, the predefined alternative
preparation, the category dictionary, and the shared country-coded site
registry. Their complete SHA-256 identities were recorded before estimation.

```{r}
#| label: tbl-h03-prep-input-identities
#| tbl-cap: "Verified inputs used by the H03 result-producing analysis."

result_input_ids <- c(
  "main_near_eye",
  "main_chest",
  "normalized_diary",
  "gap_timing_unaware",
  "category_dictionary",
  "site_registry"
)
input_labels <- c(
  main_near_eye = "Primary near-eye one-hour outcome",
  main_chest = "Complementary chest one-hour outcome",
  normalized_diary = "Normalized light-source diary",
  gap_timing_unaware = "Gap-timing-unaware one-hour outcomes",
  category_dictionary = "Light-source category dictionary",
  site_registry = "Shared country-coded site registry"
)

verified_inputs <- input_audit |>
  filter(.data$input_id %in% result_input_ids) |>
  mutate(
    current_sha256 = vapply(.data$path, artifact_sha256, character(1)),
    Input = unname(input_labels[.data$input_id]),
    Path = if_else(
      startsWith(.data$path, paste0(root, "/")),
      substring(.data$path, nchar(root) + 2L),
      basename(.data$path)
    ),
    `SHA-256 prefix` = substr(.data$current_sha256, 1L, 16L),
    verification_state = if_else(
      .data$current_sha256 == .data$sha256,
      "PASS",
      "FAIL"
    ),
    Verification = recode(
      .data$verification_state,
      PASS = "Verified",
      FAIL = "Review needed"
    )
  )
stopifnot(
  nrow(verified_inputs) == 6L,
  all(verified_inputs$verification_state == "PASS")
)

verified_inputs |>
  select("Input", "Path", "SHA-256 prefix", "Verification") |>
  gt::gt() |>
  gt::cols_width(
    Input ~ gt::pct(27),
    Path ~ gt::pct(51),
    `SHA-256 prefix` ~ gt::pct(15),
    Verification ~ gt::pct(7)
  ) |>
  gt::tab_source_note(
    "Complete 64-character identities are retained in artifacts/06_model_data/H03/H03_input_audit.csv."
  ) |>
  h03_gt(11)
```

### Schema, keys, missingness, and category overlap

Outcome and diary records were joined one-to-one by site, participant,
local date, and local clock hour. The primary model frames contain only
admissible outcome hours with a non-missing primary-light-source category.
Each retained record has one analysis category. The normalized diary also
retains seven source-indicator flags: 7,109 near-eye hours and 8,012 chest
hours had more than one true source flag, but the declared primary category
was one of those true flags and remained the single mutually exclusive
analysis label. Thus the overlap information was audited without duplicating
an hour across categories.

```{r}
#| label: tbl-h03-prep-integrity
#| tbl-cap: "Bounded integrity checks of the accepted primary H03 model frames."

integrity_display <- frame_integrity |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    `Duplicate hour keys` = .data$duplicate_hour_keys,
    `Missing outcomes` = .data$missing_outcome,
    `Missing categories` = .data$missing_category,
    `Invalid outcomes` = .data$invalid_response,
    `Hours below 30 valid minutes` = .data$valid_wall_minutes_below_30,
    `One source flag` = .data$source_flag_one,
    `Multiple source flags` = .data$source_flag_multiple,
    `Category-code conflicts` = .data$category_mapping_conflicts
  )
stopifnot(
  all(integrity_display$`Duplicate hour keys` == 0L),
  all(integrity_display$`Missing outcomes` == 0L),
  all(integrity_display$`Missing categories` == 0L),
  all(integrity_display$`Invalid outcomes` == 0L),
  all(integrity_display$`Hours below 30 valid minutes` == 0L),
  all(integrity_display$`Category-code conflicts` == 0L)
)

integrity_display |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(columns = where(is.numeric), sep_mark = ",") |>
  h03_gt(11)
```

### Outcome definition and exact zeros

The outcome is the exact one-hour zero-aware geometric mean of melEDI in lux,
calculated from at least 30 valid minutes. No logarithmic offset is added to
this primary response. Exact zero is meaningful and remained in the
quasi-Tweedie models: 4,977 of 17,935 near-eye hours (27.8%) and 5,409 of
19,512 chest hours (27.7%).

```{r}
#| label: tbl-h03-prep-zero-support
#| tbl-cap: "Exact-zero and positive melEDI support in the fitted primary frames."

zero_summary |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    `Exact-zero hours` = .data$exact_zero_hours,
    `Exact-zero fraction` = .data$exact_zero_fraction,
    `Positive median (lx)` = .data$positive_median_lx,
    `Positive 95th percentile (lx)` =
      .data$positive_95th_percentile_lx,
    `Maximum (lx)` = .data$maximum_lx,
    `Zeros retained` = if_else(
      .data$zeros_retained_in_models,
      "Yes",
      "No"
    )
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Hours, `Exact-zero hours`),
    sep_mark = ","
  ) |>
  gt::fmt_percent(columns = `Exact-zero fraction`, decimals = 1) |>
  gt::fmt_number(
    columns = c(
      `Positive median (lx)`,
      `Positive 95th percentile (lx)`,
      `Maximum (lx)`
    ),
    decimals = 1,
    sep_mark = ","
  ) |>
  h03_gt(11)
```

```{r}
#| label: fig-h03-prep-positive-distribution
#| fig-cap: "Distribution of positive one-hour geometric melEDI in the primary model frames. Bins have width 0.25 on the log10 scale, while labels remain in lux. Exact-zero observations are reported separately above and remained in the fitted models."
#| fig-alt: "Two stacked histograms show strongly right-skewed positive one-hour melEDI for near-eye and chest measurements. Both span values below 0.001 lux to above 10,000 lux and have most positive observations between about 1 and 1,000 lux. Exact zeros are not drawn in this positive-only histogram."
#| fig-width: 10
#| fig-height: 5.8

ggplot(
  response_distribution |>
    mutate(
      placement = factor(
        .data$placement,
        levels = c("Near-eye", "Chest")
      )
    ),
  aes(
    x = .data$log10_midpoint,
    y = .data$positive_hours,
    fill = .data$placement
  )
) +
  geom_col(width = 0.235, colour = "white", linewidth = 0.15) +
  facet_wrap(vars(.data$placement), ncol = 1, scales = "free_y") +
  scale_fill_manual(values = placement_colors, guide = "none") +
  scale_x_continuous(
    breaks = c(-4, -2, 0, 2, 4),
    labels = c("0.0001", "0.01", "1", "100", "10,000"),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    labels = label_number(big.mark = ","),
    expand = expansion(mult = c(0, 0.06))
  ) +
  labs(
    x = "Positive one-hour geometric melEDI (lx; log10-spaced bins)",
    y = "Participant-hours"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold")
  )
```

The exact binned counts are available as [figure source
data](../../../artifacts/11_source_data/H03/H03_preparation_positive_response_distribution.csv).

## 2. Model frames and support

### Primary samples and repeated observations

A participant-hour is one accepted participant-by-local-hour record; a
participant-day comprises one participant's accepted hours on one local date.
The primary near-eye frame contains 17,935 participant-hours from 140
participants, 801 participant-days, and nine sites. The complementary chest
frame contains 19,512 participant-hours from 151 participants, 880
participant-days, and eight sites. Participant-hours repeat within
participant-days and participants; therefore the nominal hour count is not
treated as the number of independent units.

```{r}
#| label: tbl-h03-prep-primary-samples
#| tbl-cap: "Exact primary model-frame samples and temporal sequence boundaries."

frame_integrity |>
  transmute(
    Placement = .data$placement,
    `Participant-hours` = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Exact-zero hours` = .data$exact_zero_hours,
    `Boundary-spanning hours` = .data$either_boundary_hours,
    `Contiguous sequences` = .data$ar_sequences
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(columns = where(is.numeric), sep_mark = ",") |>
  gt::tab_source_note(
    "A new contiguous sequence begins at the first retained hour of a participant-day, after a true-UTC gap, or around a daylight-saving fall-back fold."
  ) |>
  h03_gt()
```

```{r}
#| label: tbl-h03-prep-participant-day-support
#| tbl-cap: "Observed hours, categories, zeros, and contiguous sequences within participant-days."

participant_day_summary <- participant_day_support |>
  group_by(.data$placement) |>
  summarise(
    `Participant-days` = n(),
    `Median hours` = stats::median(.data$hours),
    `Hours Q1–Q3` = paste0(
      stats::quantile(.data$hours, 0.25, names = FALSE),
      "–",
      stats::quantile(.data$hours, 0.75, names = FALSE)
    ),
    `Median categories` = stats::median(.data$categories),
    `Days with exact zeros` = sum(.data$exact_zero_hours > 0L),
    `Days with >1 sequence` = sum(.data$ar_sequences > 1L),
    `Days with a boundary hour` = sum(.data$boundary_hours > 0L),
    .groups = "drop"
  )

participant_day_summary |>
  rename(Placement = .data$placement) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(
      `Participant-days`,
      `Days with exact zeros`,
      `Days with >1 sequence`,
      `Days with a boundary hour`
    ),
    sep_mark = ","
  ) |>
  gt::fmt_number(
    columns = c(`Median hours`, `Median categories`),
    decimals = 1
  ) |>
  h03_gt()
```

The participant-day rows used for this descriptive table are available as
[source data](../../../artifacts/11_source_data/H03/H03_preparation_participant_day_support.csv).

### Category coding and pooled estimability

The seven registered labels are retained in their submitted order. Indoor
electric light is the reference for every reported category ratio. A pooled
category was predeclared estimable only if it had at least 200 hours, 20
participants, and three sites. All seven categories pass this pooled rule in
both placements; this does not imply that every site-specific cell is
supported.

```{r}
#| label: tbl-h03-prep-category-support
#| tbl-cap: "Registered light-source categories and pooled support."

category_support |>
  transmute(
    Placement = .data$placement,
    Order = .data$category_order,
    Category = .data$light_source,
    Code = .data$category_code,
    Hours = .data$hours,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Pooled estimable` = if_else(
      .data$pooled_estimable,
      "Yes",
      "No"
    )
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Order, Hours, Participants, `Participant-days`, Sites),
    sep_mark = ","
  ) |>
  gt::cols_width(Category ~ gt::pct(29), Code ~ gt::pct(17)) |>
  h03_gt(10.5)
```

```{r}
#| label: fig-h03-prep-category-support
#| fig-cap: "Observed category support in the primary frames. Hours and participants use separate horizontal scales; every category passed the predefined pooled estimability thresholds of at least 200 hours, 20 participants, and three sites."
#| fig-alt: "Two-panel horizontal bar chart shows participant-hours and participants for seven light-source categories, with near-eye bars in blue and chest bars in orange. Sleep darkness, indoor daylight, and indoor electric have the largest hour counts; outdoor electric has the smallest support."
#| fig-width: 10
#| fig-height: 6.6

category_plot <- category_support |>
  select(
    .data$placement,
    .data$category_order,
    .data$short_label,
    .data$hours,
    .data$participants
  ) |>
  pivot_longer(
    cols = c("hours", "participants"),
    names_to = "measure",
    values_to = "value"
  ) |>
  mutate(
    placement = factor(
      .data$placement,
      levels = c("Near-eye", "Chest")
    ),
    measure = recode(
      .data$measure,
      hours = "Participant-hours",
      participants = "Participants"
    ),
    short_label = factor(
      .data$short_label,
      levels = rev(
        category_support |>
          filter(.data$placement == "Near-eye") |>
          arrange(.data$category_order) |>
          pull(.data$short_label)
      )
    )
  )

ggplot(
  category_plot,
  aes(
    x = .data$value,
    y = .data$short_label,
    fill = .data$placement
  )
) +
  geom_col(
    position = position_dodge2(width = 0.78, preserve = "single"),
    width = 0.68
  ) +
  facet_wrap(vars(.data$measure), scales = "free_x", ncol = 2) +
  scale_fill_manual(values = placement_colors) +
  scale_x_continuous(
    labels = label_number(big.mark = ","),
    expand = expansion(mult = c(0, 0.07))
  ) +
  labs(x = NULL, y = NULL, fill = "Placement") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.text = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12),
    legend.position = "top",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
```

The exact bar values are available as [figure source
data](../../../artifacts/11_source_data/H03/H03_preparation_category_support.csv).

### Site-specific support

A site-category contrast was reportable only when the category had at least
20 hours, five participants, and five participants also observed in the
indoor-electric reference cell at that site. Near eye has 53 supported and 10
observed-but-sparse cells. Chest has 48 supported cells, seven
observed-but-sparse cells, and eight cells with no observations; seven of the
latter are from the near-eye-only site. Unsupported cells are not assigned
site-specific estimates or p-values.

```{r}
#| label: tbl-h03-prep-site-cell-support
#| tbl-cap: "Site-category support status by placement."

site_category_support |>
  count(.data$placement, .data$support_status, name = "cells") |>
  complete(
    placement = c("Near-eye", "Chest"),
    support_status = c(
      "Supported",
      "Observed but sparse",
      "No observations"
    ),
    fill = list(cells = 0L)
  ) |>
  mutate(
    placement = factor(
      .data$placement,
      levels = c("Near-eye", "Chest")
    ),
    support_status = factor(
      .data$support_status,
      levels = c(
        "Supported",
        "Observed but sparse",
        "No observations"
      )
    )
  ) |>
  arrange(.data$placement, .data$support_status) |>
  rename(
    Placement = .data$placement,
    `Cell status` = .data$support_status,
    Cells = .data$cells
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(columns = Cells) |>
  h03_gt()
```

```{r}
#| label: fig-h03-prep-site-category-support
#| fig-cap: "Participant-hour support for every site-by-light-source cell. Fill shows the observed hour count; an open circle marks an observed cell that failed the predefined 20-hour, five-participant, or five-shared-participant rule, and × marks no observations."
#| fig-alt: "Two stacked heatmaps show seven light-source categories by country-coded study site. Darker cells have more participant-hours. Outdoor electric and external light during sleep contain most open-circle sparse cells. Chest has no observations for the near-eye-only Munich (DE) site, shown by crosses."
#| fig-width: 11
#| fig-height: 8.2

site_plot <- site_category_support |>
  mutate(
    placement = factor(
      .data$placement,
      levels = c("Near-eye", "Chest")
    ),
    site_name = factor(
      .data$display_name,
      levels = unique(
        site_category_support |>
          arrange(.data$display_order) |>
          pull(.data$display_name)
      )
    ),
    category = factor(
      .data$short_label,
      levels = rev(
        category_support |>
          filter(.data$placement == "Near-eye") |>
          arrange(.data$category_order) |>
          pull(.data$short_label)
      )
    )
  )

ggplot(
  site_plot,
  aes(x = .data$site_name, y = .data$category, fill = .data$hours)
) +
  geom_tile(colour = "white", linewidth = 0.45) +
  geom_point(
    data = site_plot |>
      filter(.data$support_status == "Observed but sparse"),
    shape = 21,
    size = 3.0,
    stroke = 0.8,
    fill = "white",
    colour = "black"
  ) +
  geom_text(
    data = site_plot |>
      filter(.data$support_status == "No observations"),
    aes(label = "×"),
    colour = "black",
    size = 5
  ) +
  facet_wrap(vars(.data$placement), ncol = 1) +
  scale_fill_viridis_c(
    option = "C",
    trans = "sqrt",
    breaks = c(0, 100, 400, 800),
    labels = label_number(big.mark = ","),
    guide = guide_colorbar(
      barwidth = grid::unit(7, "cm"),
      barheight = grid::unit(0.45, "cm")
    )
  ) +
  labs(x = NULL, y = NULL, fill = "Participant-hours") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold", size = 14),
    axis.text.x = element_text(
      angle = 30,
      hjust = 1,
      vjust = 1,
      size = 12
    ),
    axis.text.y = element_text(size = 12),
    legend.position = "top",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
```

The exact tile and marker values are available as [figure source
data](../../../artifacts/11_source_data/H03/H03_preparation_site_category_support.csv).

### Local-time availability and sequence boundaries

The exploratory temporal displays use the same participant-hours. A
category-hour is called locally sparse when it is nonzero but has fewer than
20 participant-hours, fewer than five participants, or fewer than three
sites. No curve is displayed where there are no observations, and sparse
times are marked without being treated as unsupported for the primary
all-hours GLM.

```{r}
#| label: fig-h03-prep-clock-category-support
#| fig-cap: "Available participant-hours by light source and local clock hour. Fill shows hours; an open circle denotes nonzero but locally sparse support (<20 participant-hours, <5 participants, or <3 sites), and × denotes no observations."
#| fig-alt: "Two stacked heatmaps show local hours zero through 23 for seven light-source categories. Indoor electric and daylight categories cover broad daytime ranges, while outdoor electric is sparse through much of the day. Sleep darkness is concentrated overnight, and external light during sleep is also concentrated at night."
#| fig-width: 11
#| fig-height: 8

clock_plot <- clock_support |>
  mutate(
    placement = factor(
      .data$placement,
      levels = c("Near-eye", "Chest")
    ),
    category = factor(
      .data$short_label,
      levels = rev(
        category_support |>
          filter(.data$placement == "Near-eye") |>
          arrange(.data$category_order) |>
          pull(.data$short_label)
      )
    )
  )

ggplot(
  clock_plot,
  aes(x = .data$local_hour, y = .data$category, fill = .data$hours)
) +
  geom_tile(colour = "white", linewidth = 0.25) +
  geom_point(
    data = clock_plot |>
      filter(.data$support_status == "Locally sparse"),
    shape = 21,
    size = 2.8,
    stroke = 0.75,
    fill = "white",
    colour = "black"
  ) +
  geom_text(
    data = clock_plot |>
      filter(.data$support_status == "No observations"),
    aes(label = "×"),
    colour = "black",
    size = 4
  ) +
  facet_wrap(vars(.data$placement), ncol = 1) +
  scale_x_continuous(
    breaks = c(0, 6, 12, 18, 23),
    expand = c(0, 0)
  ) +
  scale_fill_viridis_c(
    option = "D",
    trans = "sqrt",
    breaks = c(0, 100, 300, 600),
    labels = label_number(big.mark = ","),
    guide = guide_colorbar(
      barwidth = grid::unit(7, "cm"),
      barheight = grid::unit(0.45, "cm")
    )
  ) +
  labs(x = "Local time (hour)", y = NULL, fill = "Participant-hours") +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.position = "top",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
```

The exact tile and marker values are available as [figure source
data](../../../artifacts/11_source_data/H03/H03_preparation_clock_category_support.csv).

## 3. Population-mean and category-by-site interaction models

### Additive population-mean model

The prespecified primary question is tested with a population-mean GLM:

```r
geo_medi_1h ~ site + light_source
```

`geo_medi_1h` is the one-hour zero-aware geometric melEDI in lux. The model
uses a log link and a quasi-Tweedie variance function
`Var(Y | X) = phi * mu^1.539919`. Site is a fixed factor with sum contrasts;
light source is a fixed seven-level factor with indoor electric light as the
treatment reference. The formula contains no random effect. Repeated
participant-hours are addressed for inference with participant-clustered HC1
sandwich covariance, finite denominator degrees of freedom equal to the
number of participant clusters minus one, and two-sided 95% confidence
intervals (95% CIs). This interval calculation accounts for the fact that
observations from the same participant may be related.

Cluster-robust covariance protects coefficient uncertainty against unspecified
within-participant dependence under its large-cluster assumptions; it does
not remove serial structure from residuals or change the fitted conditional
mean. Participant-day repetition and within-day temporal dependence are
therefore also examined directly in model checks and sensitivity analyses.

```{r}
#| label: tbl-h03-prep-primary-specification
#| tbl-cap: "Primary and complementary population-mean model settings."

tibble::tribble(
  ~Component, ~Near_eye, ~Chest,
  "Formula", "geo_medi_1h ~ site + light_source", "geo_medi_1h ~ site + light_source",
  "Response", "One-hour zero-aware geometric melEDI (lx)", "One-hour zero-aware geometric melEDI (lx)",
  "Mean and variance", "Log link; quasi-Tweedie p = 1.539919", "Log link; quasi-Tweedie p = 1.539919",
  "Reference", "Indoor electric light", "Indoor electric light",
  "Site", "Nine-level fixed factor; sum contrasts", "Eight-level fixed factor; sum contrasts",
  "Uncertainty", "Participant-clustered HC1; df = 139", "Participant-clustered HC1; df = 150",
  "Reported category mean", "Site-average estimate on log-mean scale, back-transformed once", "Site-average estimate on log-mean scale, back-transformed once",
  "Role", "Primary inference", "Complementary evidence"
) |>
  gt::gt() |>
  gt::cols_label(Near_eye = "Near eye", Chest = "Chest") |>
  gt::cols_width(
    Component ~ gt::pct(21),
    Near_eye ~ gt::pct(39.5),
    Chest ~ gt::pct(39.5)
  ) |>
  h03_gt(11)
```

### Omnibus test, category contrasts, and site-average estimates

The primary omnibus is a six-restriction Wald F test that all six
non-reference light-source coefficients equal zero. The near-eye test is the
preregistered primary inference; the chest test is complementary. The six
non-reference ratios to indoor electric light form one multiplicity family
within placement and receive a false-discovery-rate (FDR) adjustment using the
Benjamini–Hochberg method. Category means and ratios are site-average
estimates: each site receives equal weight on the fitted log-mean scale before
one back-transformation to melEDI in lux. This is a geometric expectation, not
the arithmetic mean of response-scale site predictions.

```{r}
#| label: tbl-h03-prep-primary-tests
#| tbl-cap: "Recorded additive-model category omnibus tests."

primary_tests |>
  filter(.data$test_id == "category_omnibus") |>
  transmute(
    Placement = .data$placement,
    Restrictions = .data$restrictions,
    `Participant clusters` = .data$clusters,
    `Denominator df` = .data$denominator_df,
    `Wald F` = .data$f_statistic,
    `Raw p` = format_p(.data$p_raw, .data$p_raw < 0.05),
    Status = .data$status
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_number(columns = `Wald F`, decimals = 1) |>
  gt::fmt_markdown(columns = `Raw p`) |>
  h03_gt()
```

```{r}
#| label: tbl-h03-prep-estimand-contract
#| tbl-cap: "Reported H03 estimands and their reference distributions."

tibble::tribble(
  ~Estimand, ~Source_model, ~Reference_distribution, ~Inferential_role,
  "Primary category omnibus", "Additive site + light-source model", "All retained near-eye participant-hours", "Preregistered six-restriction test",
  "Category mean melEDI", "Additive model for the primary omnibus; category-by-site interaction model for descriptive summaries", "Site-average estimate on the fitted log-mean scale", "Absolute geometric expectation with 95% CI",
  "Category ratio", "Same model as corresponding category mean", "Indoor electric light; same site-average weights", "Six non-reference ratios; FDR within placement",
  "Category-by-site interaction omnibus", "Category-by-site interaction model", "All estimable observed site-category deviations", "One additivity-restriction test",
  "Site deviation", "Category-by-site interaction model", "Site-specific category mean divided by the site-average category mean", "Supported cells only; FDR within placement",
  "Within-site category ratio", "Category-by-site interaction model", "Indoor electric light in the same site", "Supported cells only; descriptive"
) |>
  gt::gt() |>
  gt::cols_label(
    Source_model = "Source model",
    Reference_distribution = "Reference or comparison",
    Inferential_role = "Inferential role"
  ) |>
  gt::cols_width(
    Estimand ~ gt::pct(19),
    Source_model ~ gt::pct(25),
    Reference_distribution ~ gt::pct(29),
    Inferential_role ~ gt::pct(27)
  ) |>
  h03_gt(10.5)
```

### Category-by-site interaction

The association between light-source category and melEDI is allowed to differ
by study site in a model examined separately from the additive primary
omnibus. The near-eye frame supports the literal interaction:

```r
geo_medi_1h ~ site * light_source
```

The chest frame lacks all categories at the near-eye-only Munich (DE) site and is
therefore represented by one coefficient for each observed site-category
cell:

```r
geo_medi_1h ~ 0 + site_source_cell
```

The second expression is an equivalent saturated observed-cell
parameterization, not a different scientific estimand. Its restriction matrix
tests whether the observed cells can be represented additively by site plus
light source. The models use the same log link, quasi-Tweedie variance,
participant-clustered HC1 covariance, and geometric site-average estimation as
the additive models.

```{r}
#| label: tbl-h03-prep-interaction-gate
#| tbl-cap: "Estimability checks for the complete category-by-site interaction models."

interaction_gate |>
  transmute(
    Placement = .data$placement,
    Architecture = recode(
      .data$architecture,
      full_literal = "Literal site × light-source interaction",
      full_observed_cell = "Saturated observed site-category cells"
    ),
    Formula = .data$formula,
    `Design rank` = paste0(
      .data$design_rank,
      "/",
      .data$design_columns
    ),
    Restrictions = .data$restrictions,
    `Restriction rank` = .data$restriction_covariance_rank,
    `Covariance condition number` =
      .data$restriction_covariance_condition_number,
    `Maximum score share` = .data$maximum_cluster_score_share,
    `Maximum leverage share` = .data$maximum_cluster_leverage_share,
    `Estimability result` = if_else(
      .data$gate_pass,
      "Estimable",
      "Not estimable"
    )
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_number(
    columns = c(
      `Covariance condition number`,
      `Maximum score share`,
      `Maximum leverage share`
    ),
    decimals = 3
  ) |>
  gt::cols_width(Architecture ~ gt::pct(19), Formula ~ gt::pct(21)) |>
  h03_gt(10)
```

Both full architectures converged, were full rank, had finite robust
covariance, retained all planned restriction ranks, and passed the
predeclared condition-number and cluster-influence thresholds. The recorded
category-by-site interaction tests were F(48, 139) = 80.0 near eye and F(41, 150) = 13.1 at
chest, both p < 0.001. Supported site-specific estimates were emitted only
after these checks passed.

### Multiplicity families

```{r}
#| label: tbl-h03-prep-multiplicity
#| tbl-cap: "Predefined H03 multiplicity families."

multiplicity_registry |>
  transmute(
    Family = recode(
      .data$family_id,
      `H03-F1-omnibus` = "Primary category omnibus",
      `H03-F2-context-contrasts` = "Site-average category ratios",
      `H03-F3-site-heterogeneity` =
        "Category-by-site interaction omnibus",
      `H03-F4-site-context-contrasts` = "Supported site deviations"
    ),
    Scope = recode(
      .data$scope,
      primary_near_eye = "Primary near-eye analysis"
    ),
    Members = .data$member_definition,
    Adjustment = recode(.data$method, none = "None", BH = "FDR"),
    Alpha = .data$decision_alpha
  ) |>
  gt::gt() |>
  gt::fmt_number(columns = Alpha, decimals = 2) |>
  gt::cols_width(
    Family ~ gt::pct(23),
    Scope ~ gt::pct(19),
    Members ~ gt::pct(43),
    Adjustment ~ gt::pct(8),
    Alpha ~ gt::pct(7)
  ) |>
  h03_gt(10.5)
```

### Descriptive model performance

Squared-error and working-deviance summaries were calculated from the full
category-by-site interaction models, so the interaction contribution is retained. These
quantities describe in-sample fit and are not inferential substitutes for the
omnibus tests. The participant-hour version weights every retained hour
equally; the participant-balanced version first gives each participant equal
total weight. The working-deviance version uses the quasi-Tweedie
mean-variance relationship. Because these loss functions differ, their
R-squared values are not interchangeable. The Shapley allocation averages a
component's contribution across all orders in which components could be
added, allocating shared fitted-model information rather than counting it
repeatedly; these are point summaries without uncertainty intervals.

```{r}
#| label: tbl-h03-prep-glm-r-squared
#| tbl-cap: "Stored descriptive fit and effect partitions for the full quasi-Tweedie category-by-site interaction models."

glm_r_squared |>
  filter(.data$model_id == "quasi_tweedie_log") |>
  transmute(
    Placement = .data$placement,
    Scale = recode(
      .data$weighting,
      participant_hour_weighted = "Participant-hour squared error",
      participant_balanced = "Participant-balanced squared error",
      `model-defined observation weighting` = "Working quasi-Tweedie deviance"
    ),
    `Overall R²` = .data$overall_r_squared,
    `Site partial R²` =
      .data$site_partial_r_squared_conditional_on_category,
    `Category partial R²` =
      .data$category_partial_r_squared_conditional_on_site,
    `Interaction partial R²` =
      .data$interaction_partial_r_squared_conditional_on_additive,
    `Site allocation (%)` = .data$site_shapley_share_percent,
    `Category allocation (%)` = .data$category_shapley_share_percent,
    `Interaction allocation (%)` = .data$interaction_share_percent
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_number(
    columns = c(
      `Overall R²`,
      `Site partial R²`,
      `Category partial R²`,
      `Interaction partial R²`
    ),
    decimals = 3
  ) |>
  gt::fmt_number(columns = ends_with("(%)"), decimals = 1) |>
  h03_gt(9.8)
```

### Exploratory participant random-intercept assessment {#sec-h03-prep-participant-random-intercept}

This separate exploratory model lets each participant have a stable overall
exposure level while study site, light source, and their category-by-site
interaction remain fixed effects. It was fitted to the accepted near-eye
sample of 17,935 participant-hours from 140 participants, 801 participant-days,
nine sites, and seven light-source categories. The stored formula is
`geo_medi_1h ~ site * light_source + (1 | participant)`. The model used a
Tweedie response with a log link, maximum-likelihood estimation, and fixed
power 1.539919.

```{r}
#| label: tbl-h03-prep-participant-random-intercept
#| tbl-cap: "Stored exploratory participant random-intercept assessment for near-eye melEDI."

participant_random_intercept_fit_checks <-
  participant_random_intercept_model_checks |>
  summarise(
    models = dplyr::n(),
    converged = sum(.data$convergence_code == 0L & .data$converged),
    no_warnings = sum(.data$warning_count == 0L),
    positive_hessian = sum(.data$positive_definite_hessian),
    nonsingular = sum(!.data$singular)
  )

participant_random_intercept_display <- bind_rows(
  tibble::tribble(
    ~Section, ~Quantity, ~Estimate,
    "Model", "Formula", participant_random_intercept_summary$formula,
    "Model", "Family and link", paste0(
      participant_random_intercept_summary$family,
      "; ", participant_random_intercept_summary$link, " link"
    ),
    "Model", "Estimation", paste0(
      participant_random_intercept_summary$fitting_method,
      "; fixed power ",
      participant_random_intercept_summary$working_power_fixed
    ),
    "Model", "Accepted sample", paste0(
      format(participant_random_intercept_summary$observations, big.mark = ","),
      " hours; ", participant_random_intercept_summary$participants,
      " participants; ", participant_random_intercept_summary$participant_days,
      " participant-days; ", participant_random_intercept_summary$sites,
      " sites; ", participant_random_intercept_summary$light_source_categories,
      " categories"
    ),
    "Model-based R²", "Marginal R²", sprintf(
      "%.6f", participant_random_intercept_summary$marginal_r_squared
    ),
    "Model-based R²", "Conditional R²", sprintf(
      "%.6f", participant_random_intercept_summary$conditional_r_squared
    ),
    "Participant-level variation", "Participant R² increment", sprintf(
      "%.6f",
      participant_random_intercept_summary$participant_r_squared_increment
    ),
    "Participant-level variation", "Adjusted participant ICC", sprintf(
      "%.6f", participant_random_intercept_summary$adjusted_participant_icc
    ),
    "Participant-level variation", "Unadjusted participant ICC", sprintf(
      "%.6f", participant_random_intercept_summary$unadjusted_participant_icc
    ),
    "Participant-level variation", "Random-intercept SD (log scale)", sprintf(
      "%.6f", participant_random_intercept_summary$participant_intercept_sd_log
    ),
    "Participant-level variation", "Factor per random-intercept SD", sprintf(
      "%.6f", participant_random_intercept_summary$participant_factor_per_sd
    )
  ),
  participant_random_intercept_shapley |>
    transmute(
      Section = "Marginal R² allocation",
      Quantity = recode(
        .data$component,
        `Study site × light source` = "Category-by-site interaction"
      ),
      Estimate = sprintf(
        "%.6f (%.1f%%)",
        .data$marginal_r_squared_component,
        .data$share_of_full_marginal_r_squared_percent
      )
    ),
  tibble::tribble(
    ~Section, ~Quantity, ~Estimate,
    "Fit checks", "Nested fits converged", paste0(
      participant_random_intercept_fit_checks$converged,
      "/", participant_random_intercept_fit_checks$models,
      "; convergence code 0"
    ),
    "Fit checks", "Nested fits without warnings", paste0(
      participant_random_intercept_fit_checks$no_warnings,
      "/", participant_random_intercept_fit_checks$models
    ),
    "Fit checks", "Positive-definite Hessians", paste0(
      participant_random_intercept_fit_checks$positive_hessian,
      "/", participant_random_intercept_fit_checks$models
    ),
    "Fit checks", "Non-singular nested fits", paste0(
      participant_random_intercept_fit_checks$nonsingular,
      "/", participant_random_intercept_fit_checks$models
    ),
    "Fit checks", "Lag-one Pearson residual correlation", sprintf(
      "%.3f",
      participant_random_intercept_diagnostics$lag1_pearson_residual_correlation
    ),
    "Fit checks", "Observed / implied zero fraction", sprintf(
      "%.3f / %.3f",
      participant_random_intercept_diagnostics$observed_zero_fraction,
      participant_random_intercept_diagnostics$tweedie_implied_zero_fraction
    )
  )
)

stopifnot(
  nrow(participant_random_intercept_summary) == 1L,
  nrow(participant_random_intercept_diagnostics) == 1L,
  participant_random_intercept_fit_checks$models == 5L,
  participant_random_intercept_fit_checks$converged == 5L,
  participant_random_intercept_fit_checks$no_warnings == 5L,
  participant_random_intercept_fit_checks$positive_hessian == 5L,
  participant_random_intercept_fit_checks$nonsingular == 5L
)

participant_random_intercept_display |>
  gt::gt(groupname_col = "Section") |>
  gt::cols_width(Quantity ~ gt::pct(48), Estimate ~ gt::pct(52)) |>
  gt::tab_source_note(
    "R², ICC, and the hierarchy-respecting Shapley allocation are point summaries without uncertainty intervals."
  ) |>
  h03_gt(10.5)
```

All five hierarchy-valid nested fits converged without warnings, had
positive-definite Hessians, and were non-singular. The model contains no
random light-source slopes, participant-day effect, or AR(1) term. Residual
dependence and the zero process remain imperfectly represented: the lag-one
Pearson residual correlation was 0.288, while the observed and Tweedie-implied
zero fractions were 0.278 and 0.398.

These point summaries are descriptive, model-dependent, non-causal, do not
have uncertainty intervals, and are not a unique allocation of exposure
variation. The assessment does not replace the accepted population-mean
primary model or change its omnibus test or multiplicity families. It is
distinct from the separate within-participant and between-participant
sensitivity, which is a population-mean quasi-Tweedie augmentation with
participant-clustered covariance. That sensitivity is not a mixed model,
random-intercept model, random-slope model, or variance-component analysis.
The corresponding [reader-facing
auxiliary result](../../../notebooks/hypotheses/H03.qmd) retains the same
boundary.

## 4. Model checks and influence

### Convergence, rank, and robust covariance

```{r}
#| label: tbl-h03-prep-primary-diagnostics
#| tbl-cap: "Recorded model checks for the additive primary and complementary GLMs."

diagnostics |>
  filter(.data$run_id %in% c("main__near_eye", "main__chest")) |>
  transmute(
    Placement = .data$placement,
    Converged = if_else(.data$converged, "Yes", "No"),
    Iterations = .data$iterations,
    `Design rank` = paste0(
      .data$design_rank,
      "/",
      .data$design_columns
    ),
    `Robust covariance positive definite` = if_else(
      .data$covariance_positive_definite,
      "Yes",
      "No"
    ),
    `Covariance condition number` =
      .data$covariance_condition_number,
    Dispersion = .data$dispersion,
    `Pearson residual SD` = .data$pearson_sd,
    `Residual lag-1 correlation` =
      .data$residual_lag1_correlation,
    `Maximum cluster score share` =
      .data$maximum_cluster_score_share,
    `Maximum cluster leverage share` =
      .data$maximum_cluster_leverage_share,
    Warnings = if_else(
      is.na(.data$fit_warnings) & is.na(.data$covariance_warnings),
      "None",
      paste(.data$fit_warnings, .data$covariance_warnings)
    )
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_number(
    columns = c(
      `Covariance condition number`,
      Dispersion,
      `Pearson residual SD`
    ),
    decimals = 1
  ) |>
  gt::fmt_number(
    columns = c(
      `Residual lag-1 correlation`,
      `Maximum cluster score share`,
      `Maximum cluster leverage share`
    ),
    decimals = 3
  ) |>
  h03_gt(9.5)
```

Both additive models converged without warnings, were full rank, and had
finite positive-definite participant-clustered covariance. Large dispersion
and residual spread show that the working variance is only an approximation.
Hours close together in time still show similar unexplained deviations: the
technical lag-one Pearson residual correlations were 0.478 near eye and 0.615
at chest within contiguous participant-day sequences. Clustered covariance
allows within-participant residual dependence when estimating uncertainty,
but does not make these correlations vanish.

### Zero mass and residual shape

The fixed-power Tweedie working model substantially overpredicted exact zeros:
the observed overall fractions were 0.278 near eye and 0.277 at chest,
compared with working expectations of 0.852 and 0.948. This is a
distributional limitation, not evidence that observed zeros should be
removed. The analysis therefore treats the Tweedie family as a working
mean-variance model, uses robust covariance, and avoids probability claims
about the zero process.

```{r}
#| label: tbl-h03-prep-zero-diagnostics
#| tbl-cap: "Overall observed and working-model exact-zero fractions."

zero_diagnostics |>
  filter(.data$scope == "overall") |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    `Observed zero fraction` = .data$observed_zero_fraction,
    `Working zero fraction` = .data$working_expected_zero_fraction,
    `Observed minus working` = .data$observed_minus_working_fraction,
    `Observed / working` = .data$observed_to_working_ratio,
    `Working power` = .data$working_power,
    Dispersion = .data$dispersion
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(columns = Hours, sep_mark = ",") |>
  gt::fmt_number(
    columns = c(
      `Observed zero fraction`,
      `Working zero fraction`,
      `Observed minus working`,
      `Observed / working`,
      `Working power`
    ),
    decimals = 3
  ) |>
  gt::fmt_number(columns = Dispersion, decimals = 1) |>
  h03_gt()
```

```{r}
#| label: tbl-h03-prep-residual-acf
#| tbl-cap: "Recorded Pearson residual autocorrelation within contiguous participant-day sequences."

residual_acf |>
  filter(.data$lag <= 3L) |>
  transmute(
    Placement = .data$placement,
    Lag = .data$lag,
    Correlation = .data$correlation,
    `Eligible pairs` = .data$eligible_pairs
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_number(columns = Correlation, decimals = 3) |>
  gt::fmt_integer(columns = `Eligible pairs`, sep_mark = ",") |>
  gt::tab_source_note(
    "Pairs never cross a participant-day, a true-UTC gap, or a daylight-saving fold boundary."
  ) |>
  h03_gt()
```

The [results report model-check
figure](../../../notebooks/hypotheses/H03.qmd#model-checks) shows individual
residual points, binned residual summaries, autocorrelation, and observed
versus working zero fractions. Its exact source data are stored in
`artifacts/11_source_data/H03/`.

### Influence checks

Every site was removed once within each placement. The five participants with
the largest cluster-score contribution were also removed once within each
placement. The site deletions examine dependence on a whole site; the
participant deletions examine the most influential independent covariance
clusters. These deterministic refits preserve the model family, formula,
contrast definitions, and site-average rule.

```{r}
#| label: tbl-h03-prep-influence
#| tbl-cap: "Recorded deterministic deletion-refit plan."

influence_registry |>
  count(.data$placement, .data$deletion_type, name = "refits") |>
  transmute(
    Placement = .data$placement,
    `Deletion unit` = recode(
      .data$deletion_type,
      site = "Site",
      participant = "High-score participant"
    ),
    Refits = .data$refits,
    Purpose = if_else(
      .data$deletion_type == "site",
      "Assess dependence on each whole site",
      "Assess the five largest cluster-score contributions"
    )
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(columns = Refits) |>
  h03_gt()
```

All 27 deletion fits are retained in
[the refit registry](../../../artifacts/08_diagnostics/H03/H03_influence_refit_registry.csv);
category-ratio changes are in
[the category refits](../../../artifacts/09_tables/H03/H03_influence_category_refits.csv),
and site-context changes are in
[the site-context refits](../../../artifacts/09_tables/H03/H03_influence_site_context_refits.csv).
The category omnibus decision did not change in any deletion.

## 5. Predefined sensitivities

One sensitivity uses the **gap-timing-unaware dataset**. It still passes the
general requirement of at least 50% valid minutes per hour and at least 80% valid hours per day,
but the timing of the remaining missing observations is not used for an
additional metric-specific adjustment. At this first explanation only, the
primary dataset could be interpreted as a time-sensitive primary metric dataset.
Hereafter it is called simply the primary dataset.

Other predefined branches use the same participants and participant-hours
available at both sensor positions; exclude hours that span diary-state or
measurement-context boundaries; exclude site-category cells that fail the
predefined support rule; vary the quasi-Tweedie working power to 1.30 and
1.80; and add participant-level category-proportion terms to separate
within-participant from between-participant information. The same-participant,
same-hour comparison uses separately fitted placement estimates and is not an
equivalence test or a direct test of a sensor-position effect. These branches
do not replace the primary estimand.

```{r}
#| label: tbl-h03-prep-sensitivity-samples
#| tbl-cap: "Exact fitted samples for the predefined H03 sensitivity branches."

scenario_labels <- c(
  paired_common_sample = "Same-participant, same-hour near-eye/chest sample",
  gap_timing_unaware_dataset = "Gap-timing-unaware dataset",
  boundary_hours_excluded = "Boundary-spanning hours excluded",
  unsupported_cells_excluded = "Unsupported site-category cells excluded",
  working_power_1_30 = "Quasi-Tweedie working power 1.30",
  working_power_1_80 = "Quasi-Tweedie working power 1.80",
  mundlak_within_between = "Within-/between-participant category terms"
)

frame_index |>
  filter(.data$scenario_id %in% names(scenario_labels)) |>
  transmute(
    Sensitivity = unname(scenario_labels[.data$scenario_id]),
    Placement = .data$placement,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Working power` = .data$working_power
  ) |>
  arrange(
    .data$Sensitivity,
    factor(.data$Placement, levels = c("Near-eye", "Chest"))
  ) |>
  gt::gt(groupname_col = "Sensitivity") |>
  gt::fmt_integer(
    columns = c(Hours, Participants, `Participant-days`, Sites),
    sep_mark = ","
  ) |>
  gt::fmt_number(columns = `Working power`, decimals = 2) |>
  h03_gt(10)
```

```{r}
#| label: tbl-h03-prep-sensitivity-tests
#| tbl-cap: "Recorded six-restriction category omnibus tests across sensitivity branches."

sensitivity_tests |>
  mutate(
    Sensitivity = unname(scenario_labels[.data$scenario_id])
  ) |>
  transmute(
    Sensitivity = .data$Sensitivity,
    Placement = .data$placement,
    `Wald F` = .data$f_statistic,
    `Denominator df` = .data$denominator_df,
    `Raw p` = format_p(.data$p_raw, .data$p_raw < 0.05),
    Status = .data$status
  ) |>
  gt::gt(groupname_col = "Sensitivity") |>
  gt::fmt_number(columns = `Wald F`, decimals = 1) |>
  gt::fmt_markdown(columns = `Raw p`) |>
  h03_gt(10)
```

Every sensitivity omnibus was estimable and p < 0.001. The recorded
[category-level sensitivity
comparison](../../../artifacts/09_tables/H03/H03_sensitivity_comparison.csv)
contains the ratios, 95% CIs, relative changes, direction checks, and
whether the primary point estimate falls inside each sensitivity interval.

## 6. Exploratory nonlinear time-of-day and linear-latitude context

### Nonlinear time-of-day GAM analysis

The nonlinear generalized additive model (GAM) analysis is exploratory and
separate from the preregistered GLM inference. The association with clock time
is allowed to bend across the day rather than follow a straight line. It
models the raw one-hour geometric melEDI outcome with a fixed-power Tweedie
mean, log link, global cyclic time smooth, sum-to-zero light-source and site
deviation smooths, participant-specific smooths, and participant-day random
intercepts:

```r
geo_medi_1h ~
  s(time_hour, bs = "cc", k = 12) +
  s(time_hour, light_source, bs = "sz", k = 12) +
  s(time_hour, site, bs = "sz", k = 12) +
  s(time_hour, participant, bs = "fs", k = 10) +
  s(participant_day, bs = "re")
```

The sum-to-zero smooths add category- and site-specific deviations to the
global cyclic time-of-day smooth; indoor electric light is not a smooth
reference. `bam()` used fREML, discrete computation, cyclic knots at 0 and 24
hours, and a second fit with an AR(1) working error parameter estimated from
the preliminary standardized residual lag-1 correlation. Participant,
participant-day, and site smooths remain in the fitted model but are excluded
from the displayed population curves. The stored figure curves form a
site-average by giving sites equal weight and then omit unsupported time
points from the display.

```{r}
#| label: tbl-h03-prep-temporal-models
#| tbl-cap: "Recorded final exploratory time-of-day model settings and checks."

temporal_summary |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    Categories = .data$categories,
    `Working power` = .data$working_power,
    Method = .data$method,
    `AR(1) rho` = .data$rho,
    `Adjusted R²` = .data$adjusted_r_squared,
    `Deviance explained` = .data$deviance_explained,
    `Residual lag-1` = .data$standardized_residual_lag1,
    Converged = if_else(.data$converged, "Yes", "No")
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Hours, Participants, `Participant-days`, Sites, Categories),
    sep_mark = ","
  ) |>
  gt::fmt_number(
    columns = c(
      `Working power`,
      `AR(1) rho`,
      `Adjusted R²`,
      `Deviance explained`,
      `Residual lag-1`
    ),
    decimals = 3
  ) |>
  h03_gt(9.8)
```

```{r}
#| label: tbl-h03-prep-temporal-performance
#| tbl-cap: "Stored site-average temporal-model performance and variance allocation."

temporal_performance <- temporal_r_squared |>
  transmute(
    Placement = .data$placement,
    Quantity = "Participant-balanced raw-scale R²",
    Value = .data$r_squared
  ) |>
  bind_rows(
    temporal_allocation |>
      transmute(
        Placement = .data$placement,
        Quantity = recode(
          .data$group,
          global_time = "Global time Shapley share",
          light_source_deviations = "Light-source-deviation Shapley share",
          site_deviations = "Site-deviation Shapley share",
          participant_curves = "Participant-curve Shapley share",
          participant_day_shifts = "Participant-day-shift Shapley share"
        ),
        Value = .data$shapley_share
      )
  )

temporal_performance |>
  mutate(
    Placement = factor(
      .data$Placement,
      levels = c("Near-eye", "Chest")
    )
  ) |>
  arrange(.data$Placement, .data$Quantity) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_number(columns = Value, decimals = 3) |>
  gt::tab_source_note(
    "The R² is on the raw melEDI scale with site-average, participant-within-site, and hour-within-participant weights. Shapley allocation averages a component's contribution across all addition orders so shared fitted-model information is not counted repeatedly; these are point allocations without uncertainty intervals."
  ) |>
  h03_gt()
```

The [near-eye](../../../artifacts/10_figures/H03/H03_reader_temporal_near_eye.png)
and [chest](../../../artifacts/10_figures/H03/H03_reader_temporal_chest.png)
figures show the fitted category curves, category-to-global ratios, and
available observations by local hour. Model checks, basis dimension,
concurvity, residual autocorrelation, zero calibration, coefficient
covariance, penalty-scale variances, weighted R-squared, and variance
allocation remain in their recorded model-check and table files.

### Linear absolute-latitude context

The exploratory ecological replacement of site uses centered absolute
latitude and its category interaction:

```r
geo_medi_1h ~ light_source * absolute_latitude_centered_10deg
```

It uses the same quasi-Tweedie working power and log link, but uncertainty is
clustered by site because latitude varies only between sites. Seven
category-specific slopes form one FDR family within placement. Leave-one-site-
out fits check direction and point-estimate range; they do not create a
larger effective site sample.

```{r}
#| label: tbl-h03-prep-latitude
#| tbl-cap: "Recorded exploratory absolute-latitude model construction and support."

latitude_diagnostics |>
  transmute(
    Placement = .data$placement,
    Formula = .data$formula,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Latitude centre (°)` = .data$center_absolute_latitude_deg,
    `Design rank` = paste0(
      .data$design_rank,
      "/",
      .data$coefficients
    ),
    `Site-cluster covariance rank` = paste0(
      .data$covariance_rank,
      "/",
      .data$covariance_dimension
    ),
    Converged = if_else(.data$converged, "Yes", "No"),
    Warnings = if_else(.data$warning_count == 0L, "None", .data$warnings)
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Hours, Participants, `Participant-days`, Sites),
    sep_mark = ","
  ) |>
  gt::fmt_number(columns = `Latitude centre (°)`, decimals = 1) |>
  gt::cols_width(Formula ~ gt::pct(27)) |>
  h03_gt(9.8)
```

The small number of independent sites limits ecological precision. Latitude
may stand in for season, climate, built environment, behaviour, or other site
differences; no causal latitude interpretation is made.

## 7. Technical provenance and reproduction

### Code and output map

```{r}
#| label: tbl-h03-prep-script-map
#| tbl-cap: "H03 code modules and scripts in analytical order."

tibble::tribble(
  ~Order, ~File, ~Purpose, ~Principal_outputs, ~Run_on_this_page,
  1L, "h03_contract.R", "Defines accepted inputs, category/site registries, formulas, support rules, contrasts, working power, and model-check thresholds.", "In-memory contract used by every H03 runner", "No",
  2L, "h03_data.R", "Verifies one-to-one local-hour joins, builds primary and sensitivity frames, factorizes categories and sites, and creates contiguous time sequences.", "H03_model_frames.rds and frame index", "No",
  3L, "h03_modeling.R", "Implements quasi-Tweedie GLMs, participant-clustered HC1 inference, geometric site-average estimation, category-by-site restrictions, model checks, and influence summaries.", "Models, estimates, tests and model checks", "No",
  4L, "run_h03_interaction_gate.R", "Checks the complete category-by-site interaction models before emitting site-context effects.", "Interaction check and fitted category-by-site models", "No",
  5L, "run_h03_stage2.R", "Fits the additive, category-by-site interaction, sensitivity, and deterministic influence analyses from the accepted frames.", "Model archives, result tables and run manifests", "No",
  6L, "run_h03_postfit_diagnostics.R", "Calculates residual autocorrelation, zero-mass, residual-point, and grouped model-check outputs from stored fits.", "Model-check CSV and plot-source data", "No",
  7L, "run_h03_glm_fit_assessment.R", "Calculates bounded descriptive R-squared and fixed-effect partitions for the full category-by-site interaction models.", "R-squared and effect-partition table", "No",
  8L, "run_h03_participant_random_intercept_assessment.R", "Fits the sealed descriptive participant random-intercept assessment and hierarchy-valid nested models.", "Auxiliary model, summary, R² allocation, model checks, environment and manifest", "No",
  9L, "h03_temporal.R", "Defines exploratory raw-outcome GAM fitting, AR sequences, model checks, smooth extraction, R-squared, and variance allocations.", "Temporal functions used by the dedicated runner", "No",
  10L, "run_h03_temporal_raw_mean_pilot.R", "Produces the retained raw-outcome time-of-day GAM fits and their stored curves and model checks.", "Temporal model archive, curves and model checks", "No",
  11L, "build_h03_stage3_assets.R", "Builds the final temporal figures and linked source data from stored fits.", "Near-eye/chest temporal figures and source CSVs", "No",
  12L, "build_h03_stage3_revision.R", "Produces category-by-site interaction summaries and the bounded exploratory latitude analysis.", "Category-by-site summary and latitude outputs", "No",
  13L, "h03_reporting.R", "Defines accepted tables, figure themes, labels, scales, and export routines.", "Reader-facing display functions", "No",
  14L, "build_h03_preparation_artifacts.R", "Builds bounded descriptive source data shown on this page from the accepted model frames.", "Seven preparation source-data CSVs", "Before render"
) |>
  gt::gt() |>
  gt::fmt_integer(columns = Order) |>
  gt::cols_label(
    Principal_outputs = "Principal outputs",
    Run_on_this_page = "Run on this page"
  ) |>
  gt::cols_width(
    Order ~ gt::pct(5),
    File ~ gt::pct(18),
    Purpose ~ gt::pct(42),
    Principal_outputs ~ gt::pct(27),
    Run_on_this_page ~ gt::pct(8)
  ) |>
  h03_gt(9.3)
```

Only the last row creates the lightweight descriptive inputs for this page.
All inferential and exploratory model outputs were produced before rendering
and are read by identity.

```{r}
#| label: tbl-h03-prep-output-map
#| tbl-cap: "Stored H03 outputs supporting the result report."

tibble::tribble(
  ~Output_group, ~Location, ~Contents, ~Reader_use,
  "Stored model frames", "artifacts/06_model_data/H03/H03_model_frames.rds", "Primary, same-participant/same-hour, gap-timing-unaware, boundary-excluded, and supported-cell frames", "Exact samples and descriptive support",
  "Model objects", "artifacts/07_models/H03/", "Additive, category-by-site interaction, temporal, and bounded comparison fits", "Reproducible fitted state; not rendered directly",
  "Model checks", "artifacts/08_diagnostics/H03/", "Convergence, rank, covariance, residual, zero-mass, temporal, sparse-cell, influence, and latitude checks", "Model adequacy and limitations",
  "Result tables", "artifacts/09_tables/H03/", "Omnibus tests, category means and ratios, site context, sensitivities, R², temporal and latitude results", "Exact reported numerical outputs",
  "Figures", "artifacts/10_figures/H03/", "Tightly bounded PNG, PDF, and SVG result figures", "Reader-facing visual evidence",
  "Figure/table source data", "artifacts/11_source_data/H03/", "Exact rows used by result and preparation figure layers", "Independent display reconstruction",
  "Manifests", "artifacts/12_manifests/H03/", "R/package versions, hashes, byte sizes, run settings, and figure QA", "Identity and reproducibility checks",
  "Auxiliary model", "artifacts/07_models/H03/H03_near_eye_participant_random_intercept_assessment.rds", "Sealed near-eye participant random-intercept model", "Descriptive participant-level variation assessment; not rendered directly",
  "Auxiliary summary", "artifacts/09_tables/H03/H03_near_eye_participant_random_intercept_summary.csv", "Formula, sample, marginal and conditional R², ICC, and random-intercept scale", "Exact auxiliary point summaries",
  "Auxiliary R² allocation", "artifacts/09_tables/H03/H03_near_eye_participant_random_intercept_marginal_r2_shapley.csv", "Hierarchy-respecting Shapley allocation of marginal R²", "Descriptive fixed-effect allocation",
  "Auxiliary nested-model checks", "artifacts/08_diagnostics/H03/H03_near_eye_participant_random_intercept_shapley_models.csv", "Five nested-model formulas, R² values, convergence, Hessian, and singularity checks", "Allocation provenance and fit checks",
  "Auxiliary diagnostics", "artifacts/08_diagnostics/H03/H03_near_eye_participant_random_intercept_diagnostics.csv", "Residual dependence, zero fractions, and numerical checks", "Auxiliary model limitations",
  "Auxiliary environment", "artifacts/12_manifests/H03/H03_near_eye_participant_random_intercept_environment.csv", "R and consequential package versions", "Execution context",
  "Auxiliary manifest", "artifacts/12_manifests/H03/H03_near_eye_participant_random_intercept_manifest.csv", "Six-file SHA-256, byte-size, and R-version inventory", "Sealed auxiliary identity",
  "Result report", "notebooks/hypotheses/H03.qmd", "Scientific narrative, tables, figures, model checks, sensitivities, and limitations", "Primary reader-facing synthesis",
  "Preparation companion", "audit/hypotheses/H03/H03_analysis_preparation.qmd", "Data-to-result construction, checks, and provenance", "This page"
) |>
  gt::gt() |>
  gt::cols_label(
    Output_group = "Output group",
    Reader_use = "Reader use"
  ) |>
  gt::cols_width(
    Output_group ~ gt::pct(18),
    Location ~ gt::pct(30),
    Contents ~ gt::pct(31),
    Reader_use ~ gt::pct(21)
  ) |>
  h03_gt(9.8)
```

### Commands, identities, and environment

```{r}
#| label: tbl-h03-prep-boundary
#| tbl-cap: "Calculations performed when this page is rendered."

tibble::tribble(
  ~Category, ~Calculated_on_render, ~Explanation,
  "File identity and schema", "Yes", "Checks recorded SHA-256 identities, required columns, unique local-hour keys, and accepted-frame dimensions.",
  "Descriptive support", "Yes", "Summarises stored category, site, local-hour, participant-day, exact-zero, and sequence support.",
  "Model construction and estimation", "No", "Reads the formula registry, fitted-model archives, estimates, intervals, and omnibus tests.",
  "Model checks and sensitivities", "No", "Reads recorded model checks, influence refits, model-performance summaries, and sensitivity outputs.",
  "Reader figures", "Display only", "Draws four descriptive figures from linked exact source-data CSV files."
) |>
  gt::gt() |>
  gt::cols_label(Calculated_on_render = "Calculated on render") |>
  gt::cols_width(
    Category ~ gt::pct(24),
    Calculated_on_render ~ gt::pct(18),
    Explanation ~ gt::pct(58)
  ) |>
  h03_gt()
```

The preparation-report manifest records the authoring source, byte-identical
website source copy, rendered HTML, H03 scripts and tests, stored scientific
outputs, figure source data, and environment identities. It deliberately
excludes itself from its checksum inventory.

The analysis and this companion use R 4.6.1 and the existing project library.
The GLM environment records `statmod`, `sandwich`, `ggplot2`, `LightLogR`,
and `gt`; the temporal environment additionally records `mgcv` and `gratia`.

```{r}
#| label: tbl-h03-prep-environment
#| tbl-cap: "Recorded R and consequential package versions."

bind_rows(
  execution_environment |>
    mutate(Context = "Primary GLM and reporting"),
  temporal_environment |>
    filter(.data$component %in% c(
      "R",
      "R package: mgcv",
      "R package: gratia"
    )) |>
    mutate(Context = "Exploratory temporal model"),
  participant_random_intercept_environment |>
    mutate(Context = "Exploratory participant random intercept")
) |>
  distinct(.data$Context, .data$component, .data$version) |>
  transmute(
    Context = .data$Context,
    Component = .data$component,
    Version = .data$version
  ) |>
  gt::gt(groupname_col = "Context") |>
  h03_gt()
```

### Reproduction guidance

For a bounded verification that does not recompute a model:

1. verify the six input identities in
   `artifacts/06_model_data/H03/H03_input_audit.csv`;
2. inspect `H03_model_frames.rds` and `H03_model_frame_index.csv` for the
   exact fitted samples;
3. compare formulas and support rules with `h03_contract.R`, `h03_data.R`,
   and `h03_modeling.R`;
4. trace each reported number to `artifacts/09_tables/H03/` and each
   model-check statement to `artifacts/08_diagnostics/H03/`;
5. rebuild displays only from their paired CSVs in
   `artifacts/11_source_data/H03/`; and
6. verify the final identities in `H03_preparation_report_manifest.csv`.

The lightweight preparation inputs are regenerated with:

```bash
Rscript --vanilla scripts/hypotheses/H03/build_h03_preparation_artifacts.R
```

The companion itself is rendered directly with the Nature Health profile
after it is registered beside the H03 results page. A focused H03 test then
checks reciprocal links, navigation order, source-copy identity, figure
captions and alt text, paired source data, compact `gt` tables, prohibited
scientific calls, sample assertions, figure QA, and manifest hashes.

The [H03 results report](../../../notebooks/hypotheses/H03.qmd) is the
appropriate page for scientific interpretation. This companion records how
that report was constructed and how its evidence can be checked.
