---
title: "H03: Self-reported light source and measured personal light exposure"
subtitle: "Associations with one-hour melanopic equivalent daylight illuminance"
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h03-reader
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

root <- Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
if (!nzchar(root) || !file.exists(file.path(root, "renv.lock"))) {
  root <- locate_project_root()
}
root <- normalizePath(root, winslash = "/", mustWork = TRUE)
stopifnot(identical(as.character(getRversion()), "4.6.1"))
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))
source(file.path(root, "scripts/pipeline/p_value_display.R"))

read_h03 <- function(...) {
  readr::read_csv(
    file.path(root, "artifacts", ...),
    show_col_types = FALSE,
    na = ""
  )
}

samples <- read_h03(
  "06_model_data", "H03", "H03_model_frame_index.csv"
)
category_support <- read_h03(
  "06_model_data", "H03", "H03_category_support.csv"
)
primary <- read_h03(
  "09_tables", "H03", "H03_primary_category_estimands.csv"
)
omnibus <- read_h03(
  "09_tables", "H03", "H03_primary_omnibus_tests.csv"
)
site_context <- read_h03(
  "09_tables", "H03", "H03_site_context_estimands.csv"
)
interaction_category <- read_h03(
  "09_tables", "H03", "H03_reader_heterogeneity_category_estimands.csv"
)
latitude_slopes <- read_h03(
  "09_tables", "H03", "H03_reader_latitude_category_slopes.csv"
)
latitude_diagnostics <- read_h03(
  "08_diagnostics", "H03", "H03_reader_latitude_model_diagnostics.csv"
)
interaction_gate <- read_h03(
  "08_diagnostics", "H03", "H03_interaction_architecture_gate.csv"
)
diagnostics <- read_h03(
  "08_diagnostics", "H03", "H03_model_diagnostics.csv"
)
zero_mass <- read_h03(
  "08_diagnostics", "H03", "H03_zero_mass_diagnostics.csv"
)
glm_r_squared <- read_h03(
  "09_tables", "H03", "H03_glm_r_squared_and_effect_partition.csv"
)
sensitivity <- read_h03(
  "09_tables", "H03", "H03_sensitivity_comparison.csv"
)
sensitivity_omnibus <- read_h03(
  "09_tables", "H03", "H03_sensitivity_omnibus_tests.csv"
)
paired <- read_h03(
  "09_tables", "H03", "H03_paired_placement_comparison.csv"
)
influence <- read_h03(
  "09_tables", "H03", "H03_influence_category_refits.csv"
)
influence_jobs <- read_h03(
  "08_diagnostics", "H03", "H03_influence_refit_registry.csv"
)
temporal_summary <- read_h03(
  "09_tables", "H03", "H03_reader_temporal_model_summary.csv"
)
temporal_r_squared <- read_h03(
  "09_tables", "H03", "H03_reader_temporal_weighted_r_squared.csv"
)
temporal_allocation <- read_h03(
  "09_tables", "H03", "H03_reader_temporal_variance_allocation.csv"
)
temporal_k <- read_h03(
  "08_diagnostics", "H03", "H03_reader_temporal_k_check.csv"
)
temporal_acf <- read_h03(
  "08_diagnostics", "H03", "H03_reader_temporal_residual_acf.csv"
)
site_registry <- readr::read_csv(
  file.path(root, "config", "site_display_registry.csv"),
  show_col_types = FALSE
) |>
  arrange(.data$display_order)

format_p <- function(value) nh_format_p_value(value)
format_p_cell <- function(value, significant = FALSE) {
  ifelse(
    is.na(value),
    "—",
    {
      display <- nh_p_value_display(value, significant = significant)
      ifelse(
        display$p_bold,
        paste0("**", display$p_display, "**"),
        display$p_display
      )
    }
  )
}
format_mean_ci <- function(estimate, low, high, digits = 1L) {
  paste0(
    formatC(estimate, digits = digits, format = "f", big.mark = ","),
    " (",
    formatC(low, digits = digits, format = "f", big.mark = ","),
    "–",
    formatC(high, digits = digits, format = "f", big.mark = ","),
    ")"
  )
}
format_ratio_ci <- function(estimate, low, high, digits = 3L) {
  paste0(
    formatC(estimate, digits = digits, format = "f"),
    " (",
    formatC(low, digits = digits, format = "f"),
    "–",
    formatC(high, digits = digits, format = "f"),
    ")"
  )
}
h03_gt <- function(table, font_size = 13) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

main_samples <- samples |>
  filter(.data$run_id %in% c("main__near_eye", "main__chest")) |>
  mutate(Placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$Placement)
near_primary <- primary |>
  filter(.data$placement == "Near-eye")
chest_primary <- primary |>
  filter(.data$placement == "Chest")
near_interaction <- interaction_category |>
  filter(.data$placement == "Near-eye") |>
  arrange(.data$category_order)
chest_interaction <- interaction_category |>
  filter(.data$placement == "Chest") |>
  arrange(.data$category_order)
near_category_test <- omnibus |>
  filter(.data$placement == "Near-eye", .data$test_id == "category_omnibus")
chest_category_test <- omnibus |>
  filter(.data$placement == "Chest", .data$test_id == "category_omnibus")
near_heterogeneity_test <- omnibus |>
  filter(
    .data$placement == "Near-eye",
    .data$test_id == "site_heterogeneity_omnibus"
  )
chest_heterogeneity_test <- omnibus |>
  filter(
    .data$placement == "Chest",
    .data$test_id == "site_heterogeneity_omnibus"
  )

paired_near_sample <- samples |>
  filter(.data$run_id == "paired__near_eye")
gap_near_sample <- samples |>
  filter(.data$run_id == "gap__near_eye")
gap_chest_sample <- samples |>
  filter(.data$run_id == "gap__chest")

influence_estimable <- influence |>
  filter(
    .data$category_order != 1L,
    .data$estimability_status == "ESTIMABLE"
  )
influence_exceptions <- influence_estimable |>
  filter(!.data$full_ratio_inside_deletion_interval)

noteworthy_higher <- site_context |>
  filter(
    .data$placement == "Near-eye",
    .data$reporting_status == "ESTIMABLE",
    is.finite(.data$site_deviation_p_adjusted),
    .data$site_deviation_p_adjusted <= 0.05,
    .data$site_deviation_ratio > 1
  ) |>
  slice_max(.data$site_deviation_ratio, n = 3L, with_ties = FALSE)
noteworthy_lower <- site_context |>
  filter(
    .data$placement == "Near-eye",
    .data$reporting_status == "ESTIMABLE",
    is.finite(.data$site_deviation_p_adjusted),
    .data$site_deviation_p_adjusted <= 0.05,
    .data$site_deviation_ratio < 1
  ) |>
  slice_min(.data$site_deviation_ratio, n = 3L, with_ties = FALSE)
noteworthy_sites <- bind_rows(noteworthy_higher, noteworthy_lower)

answer_category_line <- function(category_order) {
  row <- near_interaction |>
    filter(.data$category_order == .env$category_order)
  stopifnot(nrow(row) == 1L)
  ratio <- if (category_order == 1L) {
    "1.000 (reference)"
  } else {
    format_ratio_ci(
      row$ratio_to_indoor,
      row$ratio_conf_low,
      row$ratio_conf_high
    )
  }
  paste0(
    "**", row$short_label, ":** mean ",
    format_mean_ci(
      row$expected_mel_edi_lx,
      row$expected_conf_low_lx,
      row$expected_conf_high_lx
    ),
    " lx; ratio to indoor electric ", ratio, "."
  )
}

answer_site_line <- function(index) {
  row <- noteworthy_sites[index, , drop = FALSE]
  stopifnot(nrow(row) == 1L)
  paste0(
    "**", row$site_display_name, ", ", row$short_label, ":** mean ",
    format_mean_ci(
      row$cell_mean_lx,
      row$cell_conf_low_lx,
      row$cell_conf_high_lx
    ),
    " lx; deviation factor ",
    format_ratio_ci(
      row$site_deviation_ratio,
      row$site_deviation_conf_low,
      row$site_deviation_conf_high
    ),
    "; BH-adjusted *p* ", format_p(row$site_deviation_p_adjusted), "."
  )
}

stopifnot(
  nrow(main_samples) == 2L,
  nrow(primary) == 14L,
  nrow(interaction_category) == 14L,
  nrow(near_interaction) == 7L,
  all(near_interaction$standardization_status == "ESTIMABLE"),
  nrow(latitude_slopes) == 14L,
  all(latitude_diagnostics$converged),
  all(latitude_diagnostics$design_rank == 14L),
  all(latitude_diagnostics$warning_count == 0L),
  nrow(noteworthy_sites) == 6L,
  nrow(near_category_test) == 1L,
  nrow(chest_category_test) == 1L,
  nrow(near_heterogeneity_test) == 1L,
  nrow(chest_heterogeneity_test) == 1L,
  all(omnibus$status == "ESTIMABLE"),
  all(interaction_gate$gate_pass),
  nrow(temporal_summary) == 2L,
  all(temporal_summary$converged),
  all(temporal_summary$smoothing_hessian_positive_definite),
  nrow(temporal_allocation) == 10L,
  all(abs(temporal_allocation$shapley_efficiency_error) < 1e-10),
  nrow(influence_jobs) == 27L,
  !any(influence$omnibus_decision_changed),
  nrow(influence_estimable) == 149L,
  sum(influence_estimable$full_ratio_inside_deletion_interval) == 146L
)
```

The companion [H03 analysis preparation and
provenance](../../audit/hypotheses/H03/H03_analysis_preparation.html) traces
the verified inputs, model frames, formulas, diagnostics, sensitivities, and
stored outputs underlying this report.

## Question

The preregistered hypothesis was:

> **H3: Hourly self-reported light exposure categories predict hourly geometric mean melanopic EDI.**

For each accepted participant-hour, participants reported one primary light
source. This analysis asks whether the seven registered categories are
associated with measured one-hour melEDI after accounting for study site.

::: {.callout-note title="Answer in brief"}
The seven light-source categories were associated with near-eye melEDI,
*F*(6, 139) = `r sprintf("%.2f", near_category_test$f_statistic)`,
*p* **`r format_p(near_category_test$p_raw)`**. Because category-by-site
heterogeneity was supported, all descriptive means and ratios below come from
the interaction model; only the preregistered primary omnibus comes from the
additive model. Estimates are site-standardized and have 95% confidence
intervals:

- `r answer_category_line(1L)`
- `r answer_category_line(2L)`
- `r answer_category_line(3L)`
- `r answer_category_line(4L)`
- `r answer_category_line(5L)`
- `r answer_category_line(6L)`
- `r answer_category_line(7L)`

The three largest higher and three largest lower supported near-eye site
deviations among those passing the predeclared BH label were:

- `r answer_site_line(1L)`
- `r answer_site_line(2L)`
- `r answer_site_line(3L)`
- `r answer_site_line(4L)`
- `r answer_site_line(5L)`
- `r answer_site_line(6L)`

Complementary chest summaries had the same broad ordering for the six
categories that could be standardized across all eight chest sites. External
light during sleep could not be standardized across all chest sites because
one site-category cell was absent. The gap-timing-unaware dataset did not
change the primary omnibus conclusion or any primary category classification.
:::

## What was analysed

Near-eye measurements are primary because they more closely represent light
near the eyes. Chest measurements are analysed separately as complementary
evidence; the placements are not pooled and similarity is not interpreted as
equivalence. Sleep-category measurements describe the bedside sleep
environment rather than light measured at a worn near-eye or chest position.

The response is the accepted zero-aware geometric mean melEDI within each
hour, including genuine zero values. The primary model contains 17,935
near-eye participant-hours from 140 participants, 801 participant-days, and
nine sites. The complementary model contains 19,512 chest participant-hours
from 151 participants, 880 participant-days, and eight sites.

A predefined sensitivity uses the **gap-timing-unaware dataset**. This dataset
still passed the general 50%-per-hour and 80%-per-day coverage rules. The term
means that the timing of the remaining missing observations is not used for an
additional metric-specific adjustment; it does not mean that gaps,
missingness, or coverage were ignored. For this contrast only, the approved
preparation could be interpreted as a time-sensitive primary metric dataset.
Below, it is called simply the primary dataset.

```{r}
#| label: tbl-h03-overall-samples
#| tbl-cap: "Exact fitted samples for the primary and complementary models."

main_samples |>
  transmute(
    Placement = as.character(.data$Placement),
    `Participant-hours` = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Exact-zero hours` = .data$exact_zero_hours
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  h03_gt()
```

```{r}
#| label: tbl-h03-category-support
#| tbl-cap: "Exact fitted support for every light-source category."

primary |>
  mutate(Placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$Placement, .data$category_order) |>
  transmute(
    Placement,
    Category = .data$short_label,
    `Participant-hours` = .data$hours,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(`Participant-hours`, Participants, `Participant-days`, Sites),
    use_seps = TRUE
  ) |>
  h03_gt(12)
```

All seven categories meet the pooled support rule. Outdoor electric light is
the sparsest category, with 203 near-eye and 219 chest hours. Sparse
category-by-site cells remain in the pooled models but are labelled
non-estimable for site-specific contrasts when they contain fewer than 20
participant-hours, five participants, or five participants shared with the
site's indoor-electric reference.

### Model specification and operational clarifications

The primary analysis is a population-mean quasi-Tweedie GLM with a log link,
fixed site effects, fixed light-source effects, and variance power 1.539919.
Indoor electric light is the category reference. Participant-clustered HC1
covariance and finite-cluster *t* and *F* reference distributions account for
repeated participant-hours and participant-days. This changes uncertainty;
it does not remove temporal patterns from residuals.

The registered random-effects architecture was not retained because its
participant-day component collapsed and its response-distribution diagnostics
were inadequate. The population-mean formulation preserves all seven
categories and gives an estimable category omnibus. Site heterogeneity is
assessed separately rather than folded into the primary question: near eye
uses the full `site * light_source` interaction, while chest uses an equivalent
observed-cell parameterization because one site-category cell is absent.

```{r}
#| label: tbl-h03-formulas
#| tbl-cap: "Exact formulas supplied to the primary, heterogeneity, and exploratory temporal models."

tibble::tibble(
  Model = c(
    "Primary population mean",
    "Near-eye site heterogeneity",
    "Chest site heterogeneity",
    "Exploratory latitude replacement",
    "Exploratory time of day"
  ),
  `Exact Wilkinson formula supplied to R` = c(
    "geo_medi_1h ~ site + light_source",
    "geo_medi_1h ~ site * light_source",
    "geo_medi_1h ~ 0 + site_source_cell",
    "geo_medi_1h ~ 0 + light_source + light_source:absolute_latitude_10deg_centered",
    temporal_summary$formula[[1L]]
  )
) |>
  gt::gt() |>
  gt::cols_width(
    Model ~ gt::pct(23),
    `Exact Wilkinson formula supplied to R` ~ gt::pct(77)
  ) |>
  h03_gt(12)
```

Cross-site estimates are **site-standardized**: every included site
contributes equally on the fitted log-mean scale, followed by one
back-transformation. The resulting absolute estimate is the geometric mean of
the fitted site-specific expected values. It is not weighted by the number of
observations at each site.

## Category summaries from the heterogeneity model

The additive primary model and the interaction model do not impose the same
mean structure. The additive model forces every site to share one category
ratio and estimates that common ratio from all participant-hours; its
site-standardized indoor-electric estimate is 88.9 lx. The interaction model
first estimates each site-category cell and then gives every site equal weight
on the fitted log-mean scale; its corresponding estimate is 85.1 lx. The
difference is therefore a consequence of the supported category-by-site
heterogeneity, not a different sample or outcome.

The preregistered additive-model omnibus remains the primary inferential test.
Because that common-effect model is descriptively restrictive, all category
means, ratios, confidence intervals, and site summaries from this point use
the accepted heterogeneity model.

```{r}
#| label: tbl-h03-primary-results
#| tbl-cap: "Heterogeneity-model site-standardized category means and ratios to indoor electric light."

interaction_category |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Category = .data$short_label,
    `Expected melEDI, lx (95% CI)` = if_else(
      .data$standardization_status == "ESTIMABLE",
      format_mean_ci(
        .data$expected_mel_edi_lx,
        .data$expected_conf_low_lx,
        .data$expected_conf_high_lx
      ),
      "Not estimable across all sites"
    ),
    `Ratio to indoor electric (95% CI)` = case_when(
      .data$standardization_status != "ESTIMABLE" ~
        "Not estimable across all sites",
      .data$category_order == 1L ~ "1.000 (reference)",
      TRUE ~ format_ratio_ci(
          .data$ratio_to_indoor,
          .data$ratio_conf_low,
          .data$ratio_conf_high
        )
    )
  ) |>
  arrange(.data$Placement, .data$category_order) |>
  select(
    Placement,
    Category,
    `Expected melEDI, lx (95% CI)`,
    `Ratio to indoor electric (95% CI)`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "These descriptive estimates use the accepted heterogeneity model and",
      "participant-cluster-robust confidence intervals. No p-value is attached",
      "to a different-model estimate; the additive primary omnibus is reported",
      "separately. Chest external light during sleep lacks one site-category",
      "cell and cannot be standardized across all eight chest sites."
    ))
  ) |>
  h03_gt(12)
```

For near-eye melEDI, the heterogeneity-model site-standardized outdoor-
daylight mean was 959.6 lx (95% CI 786.8–1,170.2), or 11.274 times the
indoor-electric mean (95% CI 8.709–14.595). Outdoor electric was instead 9.4
lx (95% CI 6.9–12.8), or 0.110 times the reference (95% CI 0.078–0.156).
The latter differs markedly from the additive-model summary because the
outdoor-electric contrast varies strongly by site and is sparse in several
site-category cells.

The complementary chest interaction model gave the same broad ordering for
the six fully standardizable categories. These separate placement fits do not
establish that the two sensor positions are interchangeable.

```{r}
#| label: fig-h03-primary-estimates
#| fig-cap: "Heterogeneity-model site-standardized one-hour melEDI and ratios to indoor electric light; bars are participant-cluster-robust 95% confidence intervals."
#| fig-alt: "Two-panel plot of site-standardized heterogeneity-model estimates for seven light-source categories. Outdoor daylight is highest and sleep darkness lowest. Near-eye outdoor electric is about one tenth of indoor electric. The complementary chest external-light estimate is absent because it cannot be standardized across all eight chest sites."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_heterogeneity_category_estimates.png"
))
```

[Result source data](../../artifacts/09_tables/H03/H03_reader_heterogeneity_category_estimands.csv)
and [figure source data](../../artifacts/11_source_data/H03/H03_reader_heterogeneity_category_figure_data.csv)
retain the unrounded estimates and untransformed melEDI values.

### Omnibus tests

```{r}
#| label: tbl-h03-omnibus
#| tbl-cap: "Category-association and site-heterogeneity omnibus tests."

omnibus |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Test = recode(
      .data$test_id,
      category_omnibus = "Category association",
      site_heterogeneity_omnibus = "Category-by-site heterogeneity"
    ),
    `F (numerator df, denominator df)` = sprintf(
      "%.2f (%d, %d)",
      .data$f_statistic,
      .data$restrictions,
      .data$denominator_df
    ),
    `Raw p` = format_p_cell(.data$p_raw, .data$p_raw <= 0.05)
  ) |>
  arrange(.data$Placement, .data$test_id) |>
  select(Placement, Test, `F (numerator df, denominator df)`, `Raw p`) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_markdown(columns = `Raw p`) |>
  gt::tab_source_note(
    gt::md("Each omnibus p-value is unadjusted and bold at its stated 0.050 label.")
  ) |>
  h03_gt()
```

The site-heterogeneity tests were also supported near eye,
*F*(48, 139) = `r sprintf("%.2f", near_heterogeneity_test$f_statistic)`,
*p* **`r format_p(near_heterogeneity_test$p_raw)`**, and at chest,
*F*(41, 150) = `r sprintf("%.2f", chest_heterogeneity_test$f_statistic)`,
*p* **`r format_p(chest_heterogeneity_test$p_raw)`**. The pooled category
ratios are therefore cross-site summaries rather than effects expected to be
identical at every site.

### Site-specific context

The interaction model can be read as

$$
\mu_{sc} = A \times R_c \times D_{sc},
$$

where (A) is the site-standardized indoor-electric mean, (R_c) is the
category ratio to indoor electric light, and (D_{sc}) is a site's deviation
from the site-standardized mean for that category. The following near-eye
table keeps those two reference systems distinct. Its overall row and site
rows all come from the heterogeneity model; the primary additive-model
omnibus remains the preregistered inferential result reported above.

```{r}
#| label: tbl-h03-near-site-factorization
#| tbl-cap: "Near-eye category effects and site-specific deviations in one factorized display."

category_levels <- site_context |>
  filter(.data$placement == "Near-eye") |>
  distinct(.data$category_order, .data$short_label) |>
  arrange(.data$category_order) |>
  pull(.data$short_label)

overall_context <- site_context |>
  filter(.data$placement == "Near-eye") |>
  distinct(
    .data$category_order,
    .data$short_label,
    .data$site_standardized_category_mean_lx,
    .data$category_mean_conf_low_lx,
    .data$category_mean_conf_high_lx,
    .data$category_ratio_to_indoor,
    .data$category_ratio_conf_low,
    .data$category_ratio_conf_high
  ) |>
  mutate(
    Location = "Site-standardized result",
    Value = paste0(
      sprintf("%.1f lx", .data$site_standardized_category_mean_lx),
      "<br><small>",
      sprintf(
        "mean CI %.1f–%.1f",
        .data$category_mean_conf_low_lx,
        .data$category_mean_conf_high_lx
      ),
      "</small><br>",
      if_else(
        .data$category_order == 1L,
        "1.000× reference",
        sprintf("%.3f× indoor", .data$category_ratio_to_indoor)
      ),
      if_else(
        .data$category_order == 1L,
        "",
        paste0(
          "<br><small>",
          sprintf(
            "ratio CI %.3f–%.3f",
            .data$category_ratio_conf_low,
            .data$category_ratio_conf_high
          ),
          "</small>"
        )
      )
    ),
    row_order = 0,
    row_type = "overall",
    site_color_hex = NA_character_
  ) |>
  select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$site_color_hex,
    .data$short_label,
    .data$Value
  )

separator_context <- tidyr::expand_grid(short_label = category_levels) |>
  mutate(
    Location = "&nbsp;",
    row_order = 0.5,
    row_type = "separator",
    site_color_hex = NA_character_,
    Value = "&nbsp;",
    .before = 1
  )

site_context_rows <- site_context |>
  filter(.data$placement == "Near-eye") |>
  mutate(
    Location = paste0(
      "<span style='color:", .data$site_color_hex,
      ";font-size:1.15em'>●</span>&nbsp;", .data$site_display_name
    ),
    row_order = .data$display_order,
    row_type = "site",
    Value = case_when(
      .data$reporting_status != "ESTIMABLE" ~
        "<small>Not estimable</small>",
      TRUE ~ paste0(
        if_else(
          !is.na(.data$site_deviation_p_adjusted) &
            .data$site_deviation_p_adjusted <= 0.05,
          "**",
          ""
        ),
        sprintf("%.2f×", .data$site_deviation_ratio),
        if_else(
          !is.na(.data$site_deviation_p_adjusted) &
            .data$site_deviation_p_adjusted <= 0.05,
          "**",
          ""
        ),
        "<br><small>",
        sprintf(
          "95%% CI %.2f–%.2f",
          .data$site_deviation_conf_low,
          .data$site_deviation_conf_high
        ),
        "</small>"
      )
    )
  ) |>
  select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$site_color_hex,
    .data$short_label,
    .data$Value
  )

site_factor_table_data <- bind_rows(
  overall_context,
  separator_context,
  site_context_rows
) |>
  mutate(
    short_label = factor(.data$short_label, levels = category_levels)
  ) |>
  arrange(.data$row_order) |>
  select(-.data$row_order, -.data$row_type, -.data$site_color_hex) |>
  pivot_wider(names_from = .data$short_label, values_from = .data$Value) |>
  select(.data$Location, all_of(category_levels))

site_factor_table <- site_factor_table_data |>
  gt::gt() |>
  gt::fmt_markdown(columns = everything()) |>
  gt::cols_label(Location = "Site") |>
  gt::cols_label(
    `Indoor electric` = gt::md("Indoor<br>electric"),
    `Outdoor electric` = gt::md("Outdoor<br>electric"),
    `Indoor daylight` = gt::md("Indoor<br>daylight"),
    `Outdoor daylight` = gt::md("Outdoor<br>daylight"),
    `Emissive display` = gt::md("Emissive<br>display"),
    `Sleep darkness` = gt::md("Sleep<br>darkness"),
    `External light during sleep` = gt::md("External light<br>during sleep")
  ) |>
  gt::cols_width(Location ~ gt::px(165)) |>
  h03_gt(10.5) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#E8EEF3"),
      gt::cell_text(weight = "700"),
      gt::cell_borders(
        sides = "bottom",
        color = "#8192A3",
        weight = gt::px(2)
      )
    ),
    locations = gt::cells_body(rows = 1)
  ) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#FFFFFF"),
      gt::cell_text(color = "#FFFFFF", size = gt::px(1)),
      gt::cell_borders(
        sides = c("top", "bottom"),
        color = "#FFFFFF",
        weight = gt::px(0)
      )
    ),
    locations = gt::cells_body(rows = 2)
  ) |>
  gt::tab_options(data_row.padding = gt::px(5)) |>
  gt::tab_source_note(
    gt::md(paste(
      "The blue-grey site-standardized row gives both the expected mean and",
      "ratio to indoor electric for every category; the empty row separates",
      "it from the site-specific deviations. Coloured circles use the submitted",
      "site palette. Site cells are deviation ratios around the corresponding",
      "category mean. Bold site ratios pass BH adjustment across 53 supported",
      "deviations."
    ))
  )

site_factor_table
```

```{r}
#| label: fig-h03-near-site-context
#| fig-cap: "Near-eye site-specific expected melEDI. Filled points pass BH adjustment; crosses denote unsupported contrasts."
#| fig-alt: "Seven-facet interval plot of near-eye melEDI by site and light source on one common axis spanning zero through at least one thousand lux. Dashed lines show site-standardized category means. Site-coloured filled points identify BH-significant site deviations, site-coloured open points other supported estimates, and crosses unsupported contrasts."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_near_eye_site_context_estimates.png"
))
```

[Site-context source data](../../artifacts/09_tables/H03/H03_site_context_estimands.csv)
contain every site-category estimate, interval, support label, and adjusted
site-deviation *p*-value.

### Descriptive fixed-effect R²

The primary analysis has no random effects, so mixed-model marginal and
conditional R² are not defined. Descriptive R² summaries are instead based on
the accepted heterogeneity model, including its category-by-site interaction.

```{r}
#| label: tbl-h03-fixed-r2
#| tbl-cap: "Descriptive R² and hierarchy-respecting allocation for the heterogeneity model."

glm_r_squared |>
  filter(.data$model_id == "quasi_tweedie_log") |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Definition = case_when(
      .data$loss_basis == "model_deviance" ~ "Working quasi-Tweedie deviance",
      .data$weighting == "participant_balanced" ~
        "Participant-balanced squared error",
      TRUE ~ "Participant-hour squared error"
    ),
    Definition = factor(
      .data$Definition,
      levels = c(
        "Working quasi-Tweedie deviance",
        "Participant-hour squared error",
        "Participant-balanced squared error"
      )
    ),
    `Overall R²` = sprintf("%.3f", .data$overall_r_squared),
    `Category allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$category_shapley_r_squared,
      .data$category_shapley_share_percent
    ),
    `Site allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$site_shapley_r_squared,
      .data$site_shapley_share_percent
    ),
    `Interaction allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$interaction_r_squared,
      .data$interaction_share_percent
    ),
    `Interaction partial R²` = sprintf(
      "%.3f",
      .data$interaction_partial_r_squared_conditional_on_additive
    )
  ) |>
  arrange(.data$Placement, .data$Definition) |>
  select(
    Placement,
    Definition,
    `Overall R²`,
    `Category allocation`,
    `Site allocation`,
    `Interaction allocation`,
    `Interaction partial R²`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "Allocations are R² units with shares of overall R² in parentheses.",
      "They are point descriptions without resampling intervals or additional",
      "effect tests."
    ))
  ) |>
  h03_gt(11)
```

Under the near-eye working quasi-deviance definition, overall R² was 0.522;
category, site, and interaction allocations were 86.3%, 8.7%, and 5.0% of
that fitted improvement. These are scale-dependent descriptions of the fitted
mean structure, not causal shares of exposure variance.

## Model assessment

Both population-mean fits converged without warnings, retained full design
rank, and had finite positive-definite participant-cluster covariances. Their
working distribution nevertheless does not reproduce important features of
individual-hour data.

```{r}
#| label: tbl-h03-primary-diagnostics
#| tbl-cap: "Numerical, residual, temporal, and zero-mass diagnostics."

diagnostics |>
  filter(.data$scenario_id == "primary_dataset") |>
  left_join(
    zero_mass |>
      filter(.data$scope == "overall") |>
      select(
        .data$placement,
        working_zero = .data$working_expected_zero_fraction
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Converged = .data$converged,
    `Full rank` = .data$full_rank,
    `Robust covariance positive definite` =
      .data$covariance_positive_definite,
    `|Residual|-fitted Spearman` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `Lag-1 residual correlation` = sprintf(
      "%.3f",
      .data$residual_lag1_correlation
    ),
    `Observed zero fraction` = sprintf("%.3f", .data$exact_zero_fraction),
    `Working zero fraction` = sprintf("%.3f", .data$working_zero)
  ) |>
  gt::gt() |>
  h03_gt(11)
```

Residual spread increases with the fitted mean, and residual correlation at a
one-hour lag remains 0.478 near eye and 0.615 at chest. Participant clustering
protects coefficient uncertainty against arbitrary dependence within a
participant, but it does not whiten residuals or turn the working Tweedie
family into a calibrated probability model.

The discrepancy is clearest for zeros. Exact zeros comprise 27.8% of hours,
whereas the working Tweedie relationship implies 85.2% near eye and 94.8% at
chest. The model is therefore interpreted only as a log-mean model with
participant-robust inference, not as an individual-hour prediction model or a
model of zero probability.

```{r}
#| label: fig-h03-primary-diagnostics
#| fig-cap: "Primary-model residual, temporal-dependence, and zero-mass diagnostics. The melEDI transformation is display-only."
#| fig-alt: "Three-panel diagnostic plot. Individual residuals show mean-dependent spread, residual correlations remain positive across hourly lags, and observed zero fractions are much lower than the working Tweedie fractions."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_primary_residual_diagnostics.png"
))
```

Deleting each available site and the five participants with the largest score
contributions at each placement produced 27 refits. No category-omnibus
decision changed, and 146 of 149 estimable non-reference category ratios retained the
full-data ratio within the deletion-refit interval. The three exceptions were
localized chest results: sleep darkness after excluding Kumasi, external
light during sleep after excluding Munich, and sleep darkness after excluding
one Kumasi participant.

## Sensitivity analyses

```{r}
#| label: tbl-h03-sensitivities
#| tbl-cap: "Samples, omnibus tests, and category stability across predefined sensitivities."

sensitivity_counts <- sensitivity |>
  group_by(.data$run_id, .data$placement, .data$scenario_id) |>
  summarise(
    Stable = sum(.data$stability == "stable"),
    `Support-non-estimable` = sum(
      .data$stability == "support_non_estimable"
    ),
    `Direction shifts` = sum(.data$stability == "direction_shift"),
    `Magnitude shifts` = sum(.data$stability == "magnitude_shift"),
    .groups = "drop"
  )

sensitivity_counts |>
  left_join(
    samples |>
      select(
        .data$run_id,
        .data$observations,
        .data$participants,
        .data$participant_days,
        .data$sites
      ),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  left_join(
    sensitivity_omnibus |>
      select(.data$run_id, omnibus_p = .data$p_raw),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Scenario = recode(
      .data$scenario_id,
      paired_common_sample = "Paired/common sample",
      gap_timing_unaware_dataset = "Gap-timing-unaware dataset",
      boundary_hours_excluded = "Boundary hours excluded",
      unsupported_cells_excluded = "Unsupported cells excluded",
      working_power_1_30 = "Working power 1.30",
      working_power_1_80 = "Working power 1.80",
      mundlak_within_between = "Within/between participant"
    ),
    `Raw omnibus p` = format_p_cell(
      .data$omnibus_p,
      .data$omnibus_p <= 0.05
    )
  ) |>
  arrange(.data$Placement, .data$Scenario) |>
  select(
    Placement,
    Scenario,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Raw omnibus p`,
    Stable,
    `Support-non-estimable`,
    `Direction shifts`,
    `Magnitude shifts`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(
      Hours, Participants, `Participant-days`, Sites, Stable,
      `Support-non-estimable`, `Direction shifts`, `Magnitude shifts`
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_markdown(columns = `Raw omnibus p`) |>
  gt::tab_source_note(
    gt::md("Each raw omnibus p-value is bold at its stated 0.050 label.")
  ) |>
  h03_gt(10.5)
```

The gap-timing-unaware dataset retained 17,616 near-eye and 19,220 chest
hours; all seven category rows were stable at both placements. Excluding
boundary hours or unsupported cells and changing the working variance power
also left every omnibus decision unchanged. The sparse outdoor-electric
contrast became non-estimable in three reduced samples and crossed the null
direction under two alternatives, but no alternative established a non-null
outdoor-electric contrast. The within/between-participant sensitivity changed
the magnitude of the external-light-during-sleep contrast without reversing
its direction.

### Common-sample placement comparison

The common-sample models use 14,103 matched participant-hours from 111
participants, 631 participant-days, and eight sites at both placements.
Outdoor electric light has only 162 matched hours and fails its prespecified
support rule. The remaining contrasts have the same broad direction at both
placements. The identity line is a visual concordance reference, not an
equivalence boundary.

```{r}
#| label: fig-h03-paired-placement
#| fig-cap: "Common-sample near-eye and chest category ratios. The identity line does not define equivalence."
#| fig-alt: "Log-log scatter plot of near-eye against chest category ratios to indoor electric light on identical participant-hours. Daylight categories lie above one at both placements, while display and sleep categories lie below one."
#| out-width: "72%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_paired_placement_comparison.png"
))
```

[Common-sample source data](../../artifacts/09_tables/H03/H03_paired_placement_comparison.csv)
give both component intervals and exact category-specific support.

## Exploratory time-of-day context

::: {.callout-note title="Exploratory analysis"}
This analysis is separate from the preregistered category inference. It uses
one global cyclic time-of-day smooth plus sum-to-zero light-source and site
deviations, participant curves, participant-day shifts, and a boundary-aware
AR(1) correction. Light-source curves are ratios to the global daily smooth,
not to indoor electric light, and no curve-wide test is made.
:::

The exploratory model uses the untransformed one-hour geometric melEDI outcome
with a fixed-power Tweedie mean and log link. This targets the conditional
arithmetic mean rather than a back-transformed mean of a log-transformed
outcome, which is why its daytime curves are substantially higher than curves
for a geometric-scale centre. Site, participant, and participant-day smooths
remain in the fitted model but are set to zero in the displayed global and
category curves.

The temporal model is closer to the primary model's raw-outcome mean target,
but the two displays are not numerically interchangeable. The primary scalar
averages over all contributed times and standardizes sites on its fitted
log-mean scale. The temporal curves condition on clock time and show only the
global and light-source components. Their values therefore need not average to
the primary category estimates.

```{r}
#| label: fig-h03-temporal-near-eye
#| fig-cap: "Exploratory near-eye time-of-day means, ratios to the global smooth, and available participant-hours."
#| fig-alt: "Three-row, seven-category near-eye figure. The first row shows time-specific melEDI means and a global dashed curve, the second category-to-global ratios, and the third participant-hour counts. Filled circles mark supported observed hours, open circles sparse observed hours, and grey gaps zero observations."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_temporal_near_eye.png"
))
```

[Near-eye curve data](../../artifacts/11_source_data/H03/H03_reader_temporal_near_eye_curves.csv),
[ratio data](../../artifacts/11_source_data/H03/H03_reader_temporal_near_eye_ratios.csv),
and [clock-time support](../../artifacts/11_source_data/H03/H03_reader_temporal_near_eye_support.csv)
contain the plotted values and untransformed melEDI.

```{r}
#| label: fig-h03-temporal-chest
#| fig-cap: "Exploratory chest time-of-day means, ratios to the global smooth, and available participant-hours."
#| fig-alt: "Three-row, seven-category chest figure showing time-specific melEDI means, category-to-global ratios, and participant-hour counts. Filled and open circles distinguish adequate from sparse observed support; grey gaps have no observations or fitted curve."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_temporal_chest.png"
))
```

[Chest curve data](../../artifacts/11_source_data/H03/H03_reader_temporal_chest_curves.csv),
[ratio data](../../artifacts/11_source_data/H03/H03_reader_temporal_chest_ratios.csv),
and [clock-time support](../../artifacts/11_source_data/H03/H03_reader_temporal_chest_support.csv)
contain the plotted values and untransformed melEDI.

### R² and variance allocation

```{r}
#| label: tbl-h03-temporal-fit
#| tbl-cap: "Exploratory temporal-model fit summaries."

temporal_summary |>
  select(
    .data$placement,
    .data$observations,
    .data$participants,
    .data$participant_days,
    .data$rho,
    .data$adjusted_r_squared,
    .data$deviance_explained
  ) |>
  left_join(
    temporal_r_squared |>
      select(.data$placement, balanced_r_squared = .data$r_squared),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `AR rho` = sprintf("%.3f", .data$rho),
    `Adjusted R²` = sprintf("%.3f", .data$adjusted_r_squared),
    `Deviance explained` = sprintf("%.3f", .data$deviance_explained),
    `Site-standardized participant-balanced R²` = sprintf(
      "%.3f",
      .data$balanced_r_squared
    )
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = c(Hours, Participants, `Participant-days`), use_seps = TRUE) |>
  gt::tab_source_note(
    gt::md(paste(
      "The participant-balanced R² is calculated on raw melEDI.",
      "All R² values are descriptive in-sample point estimates without",
      "resampling intervals."
    ))
  ) |>
  h03_gt(11)
```

```{r}
#| label: tbl-h03-temporal-allocation
#| tbl-cap: "Point allocation of fitted temporal linear-predictor variance."

temporal_allocation |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Component = recode(
      .data$group,
      global_time = "Global time of day",
      light_source_deviations = "Light-source deviations",
      site_deviations = "Site deviations",
      participant_curves = "Participant curves",
      participant_day_shifts = "Participant-day shifts"
    ),
    `Allocated share` = sprintf("%.1f%%", 100 * .data$shapley_share),
    `Unique partial share` = sprintf(
      "%.1f%%",
      100 * .data$partial_unique_share
    )
  ) |>
  select(Placement, Component, `Allocated share`, `Unique partial share`) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "Exact point allocation on the natural-log conditional-mean predictor",
      "under site-standardized, participant-balanced weights. Components",
      "share covariance; no simulation interval was computed."
    ))
  ) |>
  h03_gt()
```

Global time of day received 60.4% of the near-eye and 58.3% of the chest point
allocation. Light-source deviations received 24.2% and 26.4%, site deviations
5.2% and 6.5%, participant curves 7.6% and 5.7%, and participant-day shifts
2.5% and 3.1%. These are allocations of fitted linear-predictor variation,
not independent or causal percentages of raw melEDI variance.

### Temporal model assessment

```{r}
#| label: tbl-h03-temporal-diagnostics
#| tbl-cap: "Convergence, basis-capacity, residual, and zero-mass diagnostics for the temporal models."

temporal_global_k <- temporal_k |>
  filter(.data$term == "s(time_hour)") |>
  select(
    .data$placement,
    global_edf = .data$effective_df,
    global_k_prime = .data$k_prime,
    global_k_index = .data$k_index
  )

temporal_summary |>
  left_join(
    temporal_global_k,
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Converged = .data$converged,
    `Smoothing Hessian positive definite` =
      .data$smoothing_hessian_positive_definite,
    `Maximum |gradient|` = formatC(
      .data$smoothing_gradient_maximum_absolute,
      digits = 2,
      format = "e"
    ),
    `Global edf / k′` = sprintf(
      "%.2f / %d",
      .data$global_edf,
      .data$global_k_prime
    ),
    `Global k-index` = sprintf("%.3f", .data$global_k_index),
    `Lag-1 residual correlation` = sprintf(
      "%.3f",
      .data$standardized_residual_lag1
    ),
    `|Residual|-fitted Spearman` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `Observed / working zero fraction` = sprintf(
      "%.3f / %.3f",
      .data$observed_zero_fraction,
      .data$working_expected_zero_fraction
    )
  ) |>
  gt::gt() |>
  gt::tab_source_note(
    gt::md("The deterministic k-index check used zero permutation replicates.")
  ) |>
  h03_gt(10.5)
```

Both final fits converged at full coefficient rank with positive-definite
smoothing-parameter Hessians. The chest fit with zero AR correlation, used
only to estimate the fixed AR value, did not converge; the final fixed-AR fit
did. The global smooths used 9.36 and 9.47 effective degrees of freedom out of
ten available, with deterministic k-indices of 0.890 and 0.884. The smooths
are therefore close to their basis capacity, and no permutation *p*-value was
computed.

The AR adjustment reduces but does not eliminate short-lag dependence:
lag-one residual correlations are 0.146 near eye and 0.144 at chest. Residual
spread remains fitted-value dependent, and the working zero fractions of
0.465 and 0.508 exceed the observed fractions of 0.278 and 0.277. These
limitations confine the temporal analysis to descriptive conditional-mean
context. The ribbons are pointwise 95% intervals; they are not simultaneous
bands and do not support curve-wide significance claims.

```{r}
#| label: fig-h03-temporal-diagnostics
#| fig-cap: "Exploratory temporal-model residual, autocorrelation, and zero-mass diagnostics."
#| fig-alt: "Three-part diagnostic figure. Residual spread changes with fitted melEDI, lag-one residual correlation remains about 0.14 at both placements, and working Tweedie zero fractions exceed observed fractions across fitted-value deciles."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_temporal_diagnostics.png"
))
```

[Residual points](../../artifacts/11_source_data/H03/H03_reader_temporal_residual_points.csv),
[residual bins](../../artifacts/11_source_data/H03/H03_reader_temporal_residual_bins.csv),
[autocorrelations](../../artifacts/08_diagnostics/H03/H03_reader_temporal_residual_acf.csv),
and [zero calibration](../../artifacts/11_source_data/H03/H03_reader_temporal_zero_calibration.csv)
provide the exact diagnostic source values.

## Exploratory linear latitude replacement

Latitude cannot be estimated alongside a saturated fixed-site term because
each site has only one latitude. A separate exploratory model therefore
replaces site with category-specific linear **absolute-latitude** slopes:

```r
geo_medi_1h ~ 0 + light_source +
  light_source:absolute_latitude_10deg_centered
```

The coefficient for each category is the multiplicative change in expected
one-hour melEDI per 10° farther from the equator. Centering at the equal-site
mean absolute latitude—39.2° near eye and 38.1° at chest—changes the
intercepts but not the slopes. Uncertainty is clustered at site, not
participant, because latitude varies only across nine near-eye and eight chest
sites. Seven slope p-values are BH-adjusted within each placement.

```{r}
#| label: tbl-h03-latitude-slopes
#| tbl-cap: "Exploratory category-specific linear absolute-latitude associations."

latitude_slopes |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Category = factor(
      .data$short_label,
      levels = near_interaction$short_label
    ),
    `Ratio per +10° (95% CI)` = format_ratio_ci(
      .data$ratio_per_10deg,
      .data$ratio_conf_low,
      .data$ratio_conf_high
    ),
    `Raw p` = format_p_cell(.data$p_raw, FALSE),
    `BH-adjusted p` = format_p_cell(
      .data$p_adjusted,
      .data$p_adjusted < 0.05
    ),
    `Leave-one-site-out sign` = paste0(
      .data$sign_agreement,
      "/",
      .data$omissions
    ),
    `Leave-one-site-out ratio range` = sprintf(
      "%.2f–%.2f",
      .data$ratio_min,
      .data$ratio_max
    )
  ) |>
  arrange(.data$Placement, .data$Category) |>
  select(
    Placement,
    Category,
    `Ratio per +10° (95% CI)`,
    `Raw p`,
    `BH-adjusted p`,
    Sites = .data$category_sites,
    `Leave-one-site-out sign`,
    `Leave-one-site-out ratio range`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`)) |>
  gt::fmt_integer(columns = Sites) |>
  gt::tab_source_note(
    gt::md(paste(
      "Ratios and site-cluster HC1 95% intervals are coefficient-wise",
      "ecological summaries. BH adjustment forms one seven-slope family",
      "within each placement. The omission columns show how often the slope",
      "kept its full-model direction and its point-estimate range."
    ))
  ) |>
  h03_gt(10.5)
```

Near eye showed BH-labelled positive gradients for indoor electric light
(ratio 1.197 per +10°, 95% CI 1.081–1.326), indoor daylight (1.253,
1.131–1.388), outdoor daylight (1.354, 1.105–1.661), and sleep darkness
(1.428, 1.076–1.895). Each retained a positive slope in all nine
leave-one-site-out fits. At chest, only indoor electric light retained the BH
label (1.142, 1.061–1.229) and remained positive in all eight omissions.

Both latitude models converged without warnings and retained design rank
14/14. Their site-cluster covariance ranks were necessarily limited to eight
and seven because only nine and eight independent site clusters were
available. Accordingly, the report gives coefficient-wise slope tests and no
joint seven-slope Wald test.

```{r}
#| label: fig-h03-latitude-slopes
#| fig-cap: "Exploratory factor change in expected melEDI per 10° increase in absolute latitude; points filled in their light-source colour pass the seven-slope within-placement BH label."
#| fig-alt: "Two-panel forest plot of category-specific ratios per ten degrees farther from the equator. Point outlines and intervals use the corresponding light-source colour. Four near-eye slopes—indoor electric, indoor daylight, outdoor daylight, and sleep darkness—are filled in that colour because their BH-adjusted p-values are below 0.050. Only indoor electric is filled at chest. Outdoor electric, display, and external-light intervals are wider."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_latitude_category_slopes.png"
))
```

[Latitude result data](../../artifacts/09_tables/H03/H03_reader_latitude_category_slopes.csv),
[site latitudes](../../artifacts/11_source_data/H03/H03_reader_latitude_site_support.csv),
[leave-one-site-out fits](../../artifacts/08_diagnostics/H03/H03_reader_latitude_leave_one_site_out.csv),
and [model diagnostics](../../artifacts/08_diagnostics/H03/H03_reader_latitude_model_diagnostics.csv)
provide the exact exploratory outputs.

These are ecological associations across a small number of sites. Absolute
latitude may stand in for season, climate, built environment, behaviour, or
other site differences; the analysis does not identify a causal latitude
effect or show that latitude adequately replaces the supported categorical
site heterogeneity.

## Interpretation

Self-reported primary light source is strongly associated with measured
one-hour personal light exposure. Daylight, especially outdoors, corresponds
to substantially higher melEDI than indoor electric light, whereas display
and sleep-environment categories correspond to lower values. The magnitude of
these contrasts varies materially by site. Outdoor electric is especially
sparse and heterogeneous, so its cross-site descriptive value depends
strongly on whether a common category contrast is imposed.

The result is observational. It does not show that changing a reported light
source causes a particular exposure, predict a health outcome, or establish
equivalence between near-eye and chest measurements. Participant-clustered
inference, sensitivity analyses, and deletion checks support the overall
category conclusion, while residual and zero-mass diagnostics limit claims
about the working Tweedie distribution. The time-of-day models add useful
descriptive context but remain separate from the preregistered inference. The
linear absolute-latitude replacement identifies several exploratory ecological
gradients, but the small number of sites and site-level confounding preclude a
causal latitude interpretation.
