---
title: "H03: Self-reported light source and measured personal light exposure"
subtitle: "Associations with one-hour melanopic equivalent daylight illuminance"
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h03-reader
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

root <- Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
if (!nzchar(root) || !file.exists(file.path(root, "renv.lock"))) {
  root <- locate_project_root()
}
root <- normalizePath(root, winslash = "/", mustWork = TRUE)
stopifnot(identical(as.character(getRversion()), "4.6.1"))
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))
source(file.path(root, "scripts/pipeline/p_value_display.R"))

read_h03 <- function(...) {
  readr::read_csv(
    file.path(root, "artifacts", ...),
    show_col_types = FALSE,
    na = ""
  )
}

samples <- read_h03(
  "06_model_data", "H03", "H03_model_frame_index.csv"
)
category_support <- read_h03(
  "06_model_data", "H03", "H03_category_support.csv"
)
primary <- read_h03(
  "09_tables", "H03", "H03_primary_category_estimands.csv"
)
omnibus <- read_h03(
  "09_tables", "H03", "H03_primary_omnibus_tests.csv"
)
site_context <- read_h03(
  "09_tables", "H03", "H03_site_context_estimands.csv"
)
interaction_category <- read_h03(
  "09_tables", "H03", "H03_reader_heterogeneity_category_estimands.csv"
)
latitude_slopes <- read_h03(
  "09_tables", "H03", "H03_reader_latitude_category_slopes.csv"
)
latitude_diagnostics <- read_h03(
  "08_diagnostics", "H03", "H03_reader_latitude_model_diagnostics.csv"
)
interaction_gate <- read_h03(
  "08_diagnostics", "H03", "H03_interaction_architecture_gate.csv"
)
diagnostics <- read_h03(
  "08_diagnostics", "H03", "H03_model_diagnostics.csv"
)
zero_mass <- read_h03(
  "08_diagnostics", "H03", "H03_zero_mass_diagnostics.csv"
)
glm_r_squared <- read_h03(
  "09_tables", "H03", "H03_glm_r_squared_and_effect_partition.csv"
)
participant_random_intercept <- read_h03(
  "09_tables", "H03",
  "H03_near_eye_participant_random_intercept_summary.csv"
)
participant_random_intercept_diagnostics <- read_h03(
  "08_diagnostics", "H03",
  "H03_near_eye_participant_random_intercept_diagnostics.csv"
)
participant_random_intercept_shapley <- read_h03(
  "09_tables", "H03",
  paste0(
    "H03_near_eye_participant_random_intercept_",
    "marginal_r2_shapley.csv"
  )
)
participant_random_intercept_shapley_models <- read_h03(
  "08_diagnostics", "H03",
  paste0(
    "H03_near_eye_participant_random_intercept_",
    "shapley_models.csv"
  )
)
sensitivity <- read_h03(
  "09_tables", "H03", "H03_sensitivity_comparison.csv"
)
sensitivity_omnibus <- read_h03(
  "09_tables", "H03", "H03_sensitivity_omnibus_tests.csv"
)
paired <- read_h03(
  "09_tables", "H03", "H03_paired_placement_comparison.csv"
)
influence <- read_h03(
  "09_tables", "H03", "H03_influence_category_refits.csv"
)
influence_jobs <- read_h03(
  "08_diagnostics", "H03", "H03_influence_refit_registry.csv"
)
temporal_summary <- read_h03(
  "09_tables", "H03", "H03_reader_temporal_model_summary.csv"
)
temporal_r_squared <- read_h03(
  "09_tables", "H03", "H03_reader_temporal_weighted_r_squared.csv"
)
temporal_allocation <- read_h03(
  "09_tables", "H03", "H03_reader_temporal_variance_allocation.csv"
)
temporal_k <- read_h03(
  "08_diagnostics", "H03", "H03_reader_temporal_k_check.csv"
)
temporal_acf <- read_h03(
  "08_diagnostics", "H03", "H03_reader_temporal_residual_acf.csv"
)
site_registry <- readr::read_csv(
  file.path(root, "config", "site_display_registry.csv"),
  show_col_types = FALSE
) |>
  arrange(.data$display_order)

format_p <- function(value) nh_format_p_value(value)
format_p_cell <- function(value, significant = FALSE) {
  ifelse(
    is.na(value),
    "—",
    {
      display <- nh_p_value_display(value, significant = significant)
      ifelse(
        display$p_bold,
        paste0("**", display$p_display, "**"),
        display$p_display
      )
    }
  )
}
format_mean_ci <- function(estimate, low, high, digits = 1L) {
  paste0(
    formatC(estimate, digits = digits, format = "f", big.mark = ","),
    " (",
    formatC(low, digits = digits, format = "f", big.mark = ","),
    "–",
    formatC(high, digits = digits, format = "f", big.mark = ","),
    ")"
  )
}
format_ratio_ci <- function(estimate, low, high, digits = 3L) {
  paste0(
    formatC(estimate, digits = digits, format = "f"),
    " (",
    formatC(low, digits = digits, format = "f"),
    "–",
    formatC(high, digits = digits, format = "f"),
    ")"
  )
}
h03_gt <- function(table, font_size = 13) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

main_samples <- samples |>
  filter(.data$run_id %in% c("main__near_eye", "main__chest")) |>
  mutate(Placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$Placement)
near_primary <- primary |>
  filter(.data$placement == "Near-eye")
chest_primary <- primary |>
  filter(.data$placement == "Chest")
near_interaction <- interaction_category |>
  filter(.data$placement == "Near-eye") |>
  arrange(.data$category_order)
chest_interaction <- interaction_category |>
  filter(.data$placement == "Chest") |>
  arrange(.data$category_order)
near_category_test <- omnibus |>
  filter(.data$placement == "Near-eye", .data$test_id == "category_omnibus")
chest_category_test <- omnibus |>
  filter(.data$placement == "Chest", .data$test_id == "category_omnibus")
near_heterogeneity_test <- omnibus |>
  filter(
    .data$placement == "Near-eye",
    .data$test_id == "site_heterogeneity_omnibus"
  )
chest_heterogeneity_test <- omnibus |>
  filter(
    .data$placement == "Chest",
    .data$test_id == "site_heterogeneity_omnibus"
  )

paired_near_sample <- samples |>
  filter(.data$run_id == "paired__near_eye")
gap_near_sample <- samples |>
  filter(.data$run_id == "gap__near_eye")
gap_chest_sample <- samples |>
  filter(.data$run_id == "gap__chest")

influence_estimable <- influence |>
  filter(
    .data$category_order != 1L,
    .data$estimability_status == "ESTIMABLE"
  )
influence_exceptions <- influence_estimable |>
  filter(!.data$full_ratio_inside_deletion_interval)

noteworthy_higher <- site_context |>
  filter(
    .data$placement == "Near-eye",
    .data$reporting_status == "ESTIMABLE",
    is.finite(.data$site_deviation_p_adjusted),
    .data$site_deviation_p_adjusted <= 0.05,
    .data$site_deviation_ratio > 1
  ) |>
  slice_max(.data$site_deviation_ratio, n = 3L, with_ties = FALSE)
noteworthy_lower <- site_context |>
  filter(
    .data$placement == "Near-eye",
    .data$reporting_status == "ESTIMABLE",
    is.finite(.data$site_deviation_p_adjusted),
    .data$site_deviation_p_adjusted <= 0.05,
    .data$site_deviation_ratio < 1
  ) |>
  slice_min(.data$site_deviation_ratio, n = 3L, with_ties = FALSE)
noteworthy_sites <- bind_rows(noteworthy_higher, noteworthy_lower)

answer_category_line <- function(category_order) {
  row <- near_interaction |>
    filter(.data$category_order == .env$category_order)
  stopifnot(nrow(row) == 1L)
  ratio <- if (category_order == 1L) {
    "1.000 (reference)"
  } else {
    format_ratio_ci(
      row$ratio_to_indoor,
      row$ratio_conf_low,
      row$ratio_conf_high
    )
  }
  paste0(
    "**", row$short_label, ":** mean ",
    format_mean_ci(
      row$expected_mel_edi_lx,
      row$expected_conf_low_lx,
      row$expected_conf_high_lx
    ),
    " lx; ratio to indoor electric ", ratio, "."
  )
}

answer_site_line <- function(index) {
  row <- noteworthy_sites[index, , drop = FALSE]
  stopifnot(nrow(row) == 1L)
  paste0(
    "**", row$site_display_name, ", ", row$short_label, ":** mean ",
    format_mean_ci(
      row$cell_mean_lx,
      row$cell_conf_low_lx,
      row$cell_conf_high_lx
    ),
    " lx; deviation factor ",
    format_ratio_ci(
      row$site_deviation_ratio,
      row$site_deviation_conf_low,
      row$site_deviation_conf_high
    ),
    "; FDR-adjusted *p* ", format_p(row$site_deviation_p_adjusted), "."
  )
}

stopifnot(
  nrow(main_samples) == 2L,
  nrow(primary) == 14L,
  nrow(interaction_category) == 14L,
  nrow(near_interaction) == 7L,
  all(near_interaction$standardization_status == "ESTIMABLE"),
  nrow(latitude_slopes) == 14L,
  all(latitude_diagnostics$converged),
  all(latitude_diagnostics$design_rank == 14L),
  all(latitude_diagnostics$warning_count == 0L),
  nrow(noteworthy_sites) == 6L,
  nrow(near_category_test) == 1L,
  nrow(chest_category_test) == 1L,
  nrow(near_heterogeneity_test) == 1L,
  nrow(chest_heterogeneity_test) == 1L,
  all(omnibus$status == "ESTIMABLE"),
  all(interaction_gate$gate_pass),
  nrow(temporal_summary) == 2L,
  all(temporal_summary$converged),
  all(temporal_summary$smoothing_hessian_positive_definite),
  nrow(temporal_allocation) == 10L,
  all(abs(temporal_allocation$shapley_efficiency_error) < 1e-10),
  nrow(influence_jobs) == 27L,
  !any(influence$omnibus_decision_changed),
  nrow(influence_estimable) == 149L,
  sum(influence_estimable$full_ratio_inside_deletion_interval) == 146L
)
```

## Question

The preregistered hypothesis was:

> **H3: Hourly self-reported light exposure categories predict hourly geometric mean melanopic EDI.**

For each accepted participant-hour, participants reported one primary light
source. This analysis asks whether the seven registered categories are
associated with measured one-hour melanopic equivalent daylight illuminance
(melEDI) after accounting for study site.

::: {.callout-note title="Answer in brief"}
The seven light-source categories were associated with near-eye melEDI,
*F*(6, 139) = `r sprintf("%.2f", near_category_test$f_statistic)`,
*p* **`r format_p(near_category_test$p_raw)`**. The association between
light-source category and melEDI was allowed to differ by study site in a
light-source-category-by-site interaction model. Because that interaction was
supported, all descriptive means and ratios below come from this model; only
the preregistered primary omnibus comes from the additive population-mean
model. These are site-average estimates: each site receives equal weight on
the fitted log-mean scale, followed by one back-transformation to melEDI in
lux. They are shown with 95% confidence intervals (95% CIs):

- `r answer_category_line(1L)`
- `r answer_category_line(2L)`
- `r answer_category_line(3L)`
- `r answer_category_line(4L)`
- `r answer_category_line(5L)`
- `r answer_category_line(6L)`
- `r answer_category_line(7L)`

The three largest higher and three largest lower supported near-eye site
deviations among those labelled after the predeclared false-discovery-rate
(FDR) adjustment were:

- `r answer_site_line(1L)`
- `r answer_site_line(2L)`
- `r answer_site_line(3L)`
- `r answer_site_line(4L)`
- `r answer_site_line(5L)`
- `r answer_site_line(6L)`

Complementary chest summaries had the same broad ordering for the six
categories for which a site-average estimate could be calculated across all
eight chest sites. A site-average estimate for external light during sleep
could not be calculated because one chest site-category cell was absent. The
gap-timing-unaware dataset did not change the primary omnibus conclusion or
any primary category classification.
:::

## What was analysed

The primary near-eye sensor position measures light close to the eyes. The
complementary chest sensor position measures light at the chest and is not a
measure of ocular exposure. Chest measurements are analysed separately; the
positions are not pooled and similarity is not interpreted as equivalence.
Sleep-category measurements describe the bedside sleep environment rather
than light measured at a worn near-eye or chest position.

The response is the accepted zero-aware geometric mean melEDI within each
hour, including genuine zero values. A participant-hour is one accepted
participant-by-local-hour record; a participant-day comprises one
participant's accepted hours on one local date. The primary model contains
17,935 near-eye participant-hours from 140 participants, 801 participant-days,
and nine sites. The complementary model contains 19,512 chest
participant-hours from 151 participants, 880 participant-days, and eight
sites. The outcome derivation and model-ready inputs are documented in
[Preparation 04](../preparation/04_metric_derivation.qmd) and
[Preparation 06](../preparation/06_model_ready_datasets.qmd).

A predefined sensitivity uses the **gap-timing-unaware dataset**. This dataset
still passed the general 50%-per-hour and 80%-per-day coverage rules. The term
means that the timing of the remaining missing observations is not used for an
additional metric-specific adjustment; it does not mean that gaps,
missingness, or coverage were ignored. For this contrast only, the approved
preparation could be interpreted as a time-sensitive primary metric dataset.
Below, it is called simply the primary dataset.

```{r}
#| label: tbl-h03-overall-samples
#| tbl-cap: "Exact fitted samples for the primary and complementary models."

main_samples |>
  transmute(
    Placement = as.character(.data$Placement),
    `Participant-hours` = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Exact-zero hours` = .data$exact_zero_hours
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  h03_gt()
```

```{r}
#| label: tbl-h03-category-support
#| tbl-cap: "Exact fitted support for every light-source category."

primary |>
  mutate(Placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$Placement, .data$category_order) |>
  transmute(
    Placement,
    Category = .data$short_label,
    `Participant-hours` = .data$hours,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(`Participant-hours`, Participants, `Participant-days`, Sites),
    use_seps = TRUE
  ) |>
  h03_gt(12)
```

All seven categories meet the pooled support rule. Outdoor electric light is
the sparsest category, with 203 near-eye and 219 chest hours. Sparse
category-by-site cells remain in the pooled models but are labelled
non-estimable for site-specific contrasts when they contain fewer than 20
participant-hours, five participants, or five participants shared with the
site's indoor-electric reference.

### Primary population-mean and category-by-site interaction models

The primary analysis is a population-mean quasi-Tweedie GLM with a log link,
fixed site effects, fixed light-source effects, and variance power 1.539919.
Indoor electric light is the category reference. The primary omnibus asks
whether the six non-reference light-source coefficients are jointly zero.

Participant-clustered HC1 covariance and finite-cluster *t* and *F* reference
distributions account for the fact that observations from the same participant
may be related. This changes uncertainty; it does not remove temporal patterns
from residuals.

A separate light-source-category-by-site interaction model allows the
association between light-source category and melEDI to differ by study site.
Near eye uses the full `site * light_source` interaction, while chest uses an
equivalent observed-cell parameterization because one site-category cell is
absent. This model supplies the descriptive category and site summaries; the
additive population-mean model supplies the preregistered primary omnibus.

Cross-site results are reported as **site-average estimates**. Each included
site contributes equally on the fitted log-mean scale, followed by one
back-transformation that returns the estimate to melEDI in lux. The resulting
absolute estimate is the geometric mean of the fitted site-specific expected
values and is not weighted by the number of observations at each site.

## Principal results

The additive primary model and the category-by-site interaction model do not
impose the same mean structure. The additive model forces every site to share
one category ratio and estimates that common ratio from all participant-hours;
its site-average indoor-electric estimate is 88.9 lx. The interaction model
first estimates each site-category cell and then gives every site equal weight
on the fitted log-mean scale; its corresponding estimate is 85.1 lx. The
difference reflects the supported category-by-site interaction, not a
different sample or outcome.

The preregistered additive-model omnibus remains the primary inferential test.
Because that common-effect model is descriptively restrictive, all category
means, ratios, 95% CIs, and site summaries from this point use the accepted
category-by-site interaction model. The principal display is
@fig-h03-primary-estimates, followed by the detailed values in
@tbl-h03-primary-results.

```{r}
#| label: fig-h03-primary-estimates
#| fig-cap: "Site-average one-hour melEDI and ratios to indoor electric light from the light-source-category-by-site interaction model; bars are participant-cluster-robust 95% CIs."
#| fig-alt: "Two-panel plot of site-average estimates from the light-source-category-by-site interaction model for seven light-source categories. Outdoor daylight is highest and sleep darkness lowest. Near-eye outdoor electric is about one tenth of indoor electric. The complementary chest external-light estimate is absent because a site-average estimate cannot be calculated across all eight chest sites."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_heterogeneity_category_estimates.png"
))
```

```{r}
#| label: tbl-h03-primary-results
#| tbl-cap: "Site-average category means and ratios to indoor electric light from the light-source-category-by-site interaction model."

interaction_category |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Category = .data$short_label,
    `Expected melEDI, lx (95% CI)` = if_else(
      .data$standardization_status == "ESTIMABLE",
      format_mean_ci(
        .data$expected_mel_edi_lx,
        .data$expected_conf_low_lx,
        .data$expected_conf_high_lx
      ),
      "Site-average estimate not estimable"
    ),
    `Ratio to indoor electric (95% CI)` = case_when(
      .data$standardization_status != "ESTIMABLE" ~
        "Site-average estimate not estimable",
      .data$category_order == 1L ~ "1.000 (reference)",
      TRUE ~ format_ratio_ci(
          .data$ratio_to_indoor,
          .data$ratio_conf_low,
          .data$ratio_conf_high
        )
    )
  ) |>
  arrange(.data$Placement, .data$category_order) |>
  select(
    Placement,
    Category,
    `Expected melEDI, lx (95% CI)`,
    `Ratio to indoor electric (95% CI)`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "These descriptive site-average estimates use the accepted",
      "light-source-category-by-site interaction model. Each site receives",
      "equal weight on the fitted log-mean scale before back-transformation",
      "to melEDI in lux. The 95% CIs account for related observations within",
      "participants using participant-clustered HC1 covariance. No p-value is",
      "attached to a different-model estimate; the additive primary omnibus",
      "is reported separately. The complementary chest external light during",
      "sleep category lacks one site-category cell, so its site-average estimate",
      "across all eight chest sites remains non-estimable."
    ))
  ) |>
  h03_gt(12)
```

For near-eye melEDI, the category-by-site interaction model gave a site-average
outdoor-daylight mean of 959.6 lx (95% CI 786.8–1,170.2), or 11.274 times the
indoor-electric mean (95% CI 8.709–14.595). Outdoor electric was instead 9.4
lx (95% CI 6.9–12.8), or 0.110 times the reference (95% CI 0.078–0.156).
The latter differs markedly from the additive-model summary because the
outdoor-electric contrast varies strongly by site and is sparse in several
site-category cells.

The complementary chest interaction model gave the same broad ordering for
the six categories with estimable site-average results. The external-light-
during-sleep site-average estimate remains absent because one chest
site-category cell was unavailable. These separate placement fits do not
establish that the two sensor positions are interchangeable.

[Result source data](../../artifacts/09_tables/H03/H03_reader_heterogeneity_category_estimands.csv)
and [figure source data](../../artifacts/11_source_data/H03/H03_reader_heterogeneity_category_figure_data.csv)
retain the unrounded estimates and untransformed melEDI values.

### Omnibus tests

```{r}
#| label: tbl-h03-omnibus
#| tbl-cap: "Category-association and category-by-site interaction omnibus tests."

omnibus |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Test = recode(
      .data$test_id,
      category_omnibus = "Category association",
      site_heterogeneity_omnibus = "Category-by-site interaction"
    ),
    `F (numerator df, denominator df)` = sprintf(
      "%.2f (%d, %d)",
      .data$f_statistic,
      .data$restrictions,
      .data$denominator_df
    ),
    `Raw p` = format_p_cell(.data$p_raw, .data$p_raw <= 0.05)
  ) |>
  arrange(.data$Placement, .data$test_id) |>
  select(Placement, Test, `F (numerator df, denominator df)`, `Raw p`) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_markdown(columns = `Raw p`) |>
  gt::tab_source_note(
    gt::md("Each omnibus p-value is unadjusted and bold at its stated 0.050 label.")
  ) |>
  h03_gt()
```

The formal category-by-site interaction tests were also supported near eye,
*F*(48, 139) = `r sprintf("%.2f", near_heterogeneity_test$f_statistic)`,
*p* **`r format_p(near_heterogeneity_test$p_raw)`**, and at chest,
*F*(41, 150) = `r sprintf("%.2f", chest_heterogeneity_test$f_statistic)`,
*p* **`r format_p(chest_heterogeneity_test$p_raw)`**. The pooled category
ratios are therefore cross-site summaries rather than effects expected to be
identical at every site.

### Site-specific context

The interaction model can be read as

$$
\mu_{sc} = A \times R_c \times D_{sc},
$$

where (A) is the site-average indoor-electric mean, (R_c) is the
category ratio to indoor electric light, and (D_{sc}) is a site's deviation
from the site-average mean for that category. The following near-eye
table keeps those two reference systems distinct. Its overall row and site
rows all come from the category-by-site interaction model; the primary additive-model
omnibus remains the preregistered inferential result reported above.

```{r}
#| label: tbl-h03-near-site-factorization
#| tbl-cap: "Near-eye category effects and site-specific deviations in one factorized display."

category_levels <- site_context |>
  filter(.data$placement == "Near-eye") |>
  distinct(.data$category_order, .data$short_label) |>
  arrange(.data$category_order) |>
  pull(.data$short_label)

overall_context <- site_context |>
  filter(.data$placement == "Near-eye") |>
  distinct(
    .data$category_order,
    .data$short_label,
    .data$site_standardized_category_mean_lx,
    .data$category_mean_conf_low_lx,
    .data$category_mean_conf_high_lx,
    .data$category_ratio_to_indoor,
    .data$category_ratio_conf_low,
    .data$category_ratio_conf_high
  ) |>
  mutate(
    Location = "Site-average result",
    Value = paste0(
      sprintf("%.1f lx", .data$site_standardized_category_mean_lx),
      "<br><small>",
      sprintf(
        "mean CI %.1f–%.1f",
        .data$category_mean_conf_low_lx,
        .data$category_mean_conf_high_lx
      ),
      "</small><br>",
      if_else(
        .data$category_order == 1L,
        "1.000× reference",
        sprintf("%.3f× indoor", .data$category_ratio_to_indoor)
      ),
      if_else(
        .data$category_order == 1L,
        "",
        paste0(
          "<br><small>",
          sprintf(
            "ratio CI %.3f–%.3f",
            .data$category_ratio_conf_low,
            .data$category_ratio_conf_high
          ),
          "</small>"
        )
      )
    ),
    row_order = 0,
    row_type = "overall",
    site_color_hex = NA_character_
  ) |>
  select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$site_color_hex,
    .data$short_label,
    .data$Value
  )

separator_context <- tidyr::expand_grid(short_label = category_levels) |>
  mutate(
    Location = "&nbsp;",
    row_order = 0.5,
    row_type = "separator",
    site_color_hex = NA_character_,
    Value = "&nbsp;",
    .before = 1
  )

site_context_rows <- site_context |>
  filter(.data$placement == "Near-eye") |>
  mutate(
    Location = paste0(
      "<span style='color:", .data$site_color_hex,
      ";font-size:1.15em'>●</span>&nbsp;", .data$site_display_name
    ),
    row_order = .data$display_order,
    row_type = "site",
    Value = case_when(
      .data$reporting_status != "ESTIMABLE" ~
        "<small>Not estimable</small>",
      TRUE ~ paste0(
        if_else(
          !is.na(.data$site_deviation_p_adjusted) &
            .data$site_deviation_p_adjusted <= 0.05,
          "**",
          ""
        ),
        sprintf("%.2f×", .data$site_deviation_ratio),
        if_else(
          !is.na(.data$site_deviation_p_adjusted) &
            .data$site_deviation_p_adjusted <= 0.05,
          "**",
          ""
        ),
        "<br><small>",
        sprintf(
          "95%% CI %.2f–%.2f",
          .data$site_deviation_conf_low,
          .data$site_deviation_conf_high
        ),
        "</small>"
      )
    )
  ) |>
  select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$site_color_hex,
    .data$short_label,
    .data$Value
  )

site_factor_table_data <- bind_rows(
  overall_context,
  separator_context,
  site_context_rows
) |>
  mutate(
    short_label = factor(.data$short_label, levels = category_levels)
  ) |>
  arrange(.data$row_order) |>
  select(-.data$row_order, -.data$row_type, -.data$site_color_hex) |>
  pivot_wider(names_from = .data$short_label, values_from = .data$Value) |>
  select(.data$Location, all_of(category_levels))

site_factor_table <- site_factor_table_data |>
  gt::gt() |>
  gt::fmt_markdown(columns = everything()) |>
  gt::cols_label(Location = "Site") |>
  gt::cols_label(
    `Indoor electric` = gt::md("Indoor<br>electric"),
    `Outdoor electric` = gt::md("Outdoor<br>electric"),
    `Indoor daylight` = gt::md("Indoor<br>daylight"),
    `Outdoor daylight` = gt::md("Outdoor<br>daylight"),
    `Emissive display` = gt::md("Emissive<br>display"),
    `Sleep darkness` = gt::md("Sleep<br>darkness"),
    `External light during sleep` = gt::md("External light<br>during sleep")
  ) |>
  gt::cols_width(Location ~ gt::px(165)) |>
  h03_gt(10.5) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#E8EEF3"),
      gt::cell_text(weight = "700"),
      gt::cell_borders(
        sides = "bottom",
        color = "#8192A3",
        weight = gt::px(2)
      )
    ),
    locations = gt::cells_body(rows = 1)
  ) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#FFFFFF"),
      gt::cell_text(color = "#FFFFFF", size = gt::px(1)),
      gt::cell_borders(
        sides = c("top", "bottom"),
        color = "#FFFFFF",
        weight = gt::px(0)
      )
    ),
    locations = gt::cells_body(rows = 2)
  ) |>
  gt::tab_options(data_row.padding = gt::px(5)) |>
  gt::tab_source_note(
    gt::md(paste(
      "The blue-grey site-average row gives both the expected mean and",
      "ratio to indoor electric for every category; the empty row separates",
      "it from the site-specific deviations. Coloured circles use the submitted",
      "site palette. Site cells are deviation ratios around the corresponding",
      "category mean. Bold site ratios pass the FDR adjustment across 53",
      "supported deviations."
    ))
  )

site_factor_table
```

```{r}
#| label: fig-h03-near-site-context
#| fig-cap: "Near-eye site-specific expected melEDI. Filled points pass the FDR adjustment; crosses denote unsupported contrasts."
#| fig-alt: "Seven-facet interval plot of near-eye melEDI by country-coded study site and light source on one common axis spanning zero through at least one thousand lux. Dashed lines show site-average category means. Site-coloured filled points identify FDR-labelled site deviations, site-coloured open points show other supported estimates, and crosses show unsupported contrasts."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_near_eye_site_context_estimates.png"
))
```

[Site-context source data](../../artifacts/09_tables/H03/H03_site_context_estimands.csv)
contain every site-category estimate, interval, support label, and adjusted
site-deviation *p*-value.

### Descriptive fixed-effect R²

The primary analysis has no random effects, so mixed-model marginal and
conditional R² are not defined. Descriptive R² summaries are instead based on
the accepted category-by-site interaction model.

```{r}
#| label: tbl-h03-fixed-r2
#| tbl-cap: "Descriptive R² and hierarchy-respecting allocation for the category-by-site interaction model."

glm_r_squared |>
  filter(.data$model_id == "quasi_tweedie_log") |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Definition = case_when(
      .data$loss_basis == "model_deviance" ~ "Working quasi-Tweedie deviance",
      .data$weighting == "participant_balanced" ~
        "Participant-balanced squared error",
      TRUE ~ "Participant-hour squared error"
    ),
    Definition = factor(
      .data$Definition,
      levels = c(
        "Working quasi-Tweedie deviance",
        "Participant-hour squared error",
        "Participant-balanced squared error"
      )
    ),
    `Overall R²` = sprintf("%.3f", .data$overall_r_squared),
    `Category allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$category_shapley_r_squared,
      .data$category_shapley_share_percent
    ),
    `Site allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$site_shapley_r_squared,
      .data$site_shapley_share_percent
    ),
    `Interaction allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$interaction_r_squared,
      .data$interaction_share_percent
    ),
    `Interaction partial R²` = sprintf(
      "%.3f",
      .data$interaction_partial_r_squared_conditional_on_additive
    )
  ) |>
  arrange(.data$Placement, .data$Definition) |>
  select(
    Placement,
    Definition,
    `Overall R²`,
    `Category allocation`,
    `Site allocation`,
    `Interaction allocation`,
    `Interaction partial R²`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "Allocations are R² units with shares of overall R² in parentheses.",
      "They are point descriptions without resampling intervals or additional",
      "effect tests."
    ))
  ) |>
  h03_gt(11)
```

Under the near-eye working quasi-deviance definition, overall R² was 0.522;
category, site, and interaction allocations were 86.3%, 8.7%, and 5.0% of
that fitted improvement. These are scale-dependent descriptions of the fitted
mean structure, not causal shares of exposure variance.

### Exploratory participant random-intercept assessment

An auxiliary near-eye model quantified stable participant-level differences:
`geo_medi_1h ~ site * light_source + (1 | participant)`. It used the accepted
17,935 participant-hours from 140 participants, a Tweedie distribution with a
log link, and the primary working power fixed at 1.539919. The model contains
neither a participant-day intercept nor a temporal correlation term. It was
fitted only for variance description and does not replace the accepted
population-mean analysis or estimate participant-specific light-source
slopes.

```{r}
#| label: tbl-h03-participant-random-intercept
#| tbl-cap: "Exploratory participant random-intercept variance assessment for the near-eye outcome."

participant_random_intercept_table <- tibble::tribble(
  ~Component, ~Quantity, ~Estimate, ~Interpretation,
  "Model-based R²", "Marginal R²",
  participant_random_intercept$marginal_r_squared[[1L]],
  "Site, light source, and their interaction",
  "Model-based R²", "Conditional R²",
  participant_random_intercept$conditional_r_squared[[1L]],
  "Fixed effects plus the participant intercept",
  "Model-based R²", "Participant increment",
  participant_random_intercept$participant_r_squared_increment[[1L]],
  "Conditional minus marginal R²"
) |>
  dplyr::bind_rows(
    participant_random_intercept_shapley |>
      dplyr::transmute(
        Component = "Marginal R² Shapley allocation",
        Quantity = .data$component,
        Estimate = .data$marginal_r_squared_component,
        Interpretation = sprintf(
          "%.1f%% of full marginal R²",
          .data$share_of_full_marginal_r_squared_percent
        )
      )
  ) |>
  dplyr::bind_rows(tibble::tribble(
  ~Component, ~Quantity, ~Estimate, ~Interpretation,
  "Participant heterogeneity", "Adjusted participant ICC",
  participant_random_intercept$adjusted_participant_icc[[1L]],
  "Participant share after excluding fixed-effect variance",
  "Participant heterogeneity", "Unadjusted participant ICC",
  participant_random_intercept$unadjusted_participant_icc[[1L]],
  "Participant share of total model-based variance",
  "Participant heterogeneity", "Participant SD on log scale",
  participant_random_intercept$participant_intercept_sd_log[[1L]],
  "Spread of stable participant intercepts",
  "Participant heterogeneity", "One-SD multiplicative factor",
  participant_random_intercept$participant_factor_per_sd[[1L]],
  "Expected-mean factor for a participant one SD above average",
  "Participant heterogeneity", "Participant/fixed variance ratio",
  participant_random_intercept$participant_to_fixed_variance_ratio[[1L]],
  "Participant-intercept variance divided by fixed-predictor variance"
  )) |>
  mutate(
    Estimate = if_else(
      .data$Quantity == "One-SD multiplicative factor",
      paste0(sprintf("%.2f", .data$Estimate), "×"),
      sprintf("%.3f", .data$Estimate)
    )
  )

participant_random_intercept_table |>
  gt::gt(groupname_col = "Component") |>
  gt::cols_align(align = "right", columns = "Estimate") |>
  gt::tab_source_note(
    gt::md(paste(
      "Point estimates use Nakagawa's model-based variance decomposition",
      "with a lognormal distribution-specific variance approximation; no",
      "bootstrap intervals were calculated. The decomposition uses the",
      "observed participant-hour frame rather than a site-average grid.",
      "The Shapley allocation refits hierarchy-valid nested models and",
      "averages both entry orders for site and light source; the interaction",
      "enters only after both main effects."
    ))
  ) |>
  h03_gt(11)
```

Marginal R² was `r sprintf("%.3f", participant_random_intercept$marginal_r_squared[[1L]])`
for the combined site, light-source, and interaction terms. Conditional R² was
`r sprintf("%.3f", participant_random_intercept$conditional_r_squared[[1L]])`
after adding the participant intercept, a difference of
`r sprintf("%.3f", participant_random_intercept$participant_r_squared_increment[[1L]])`.
The adjusted ICC was
`r sprintf("%.3f", participant_random_intercept$adjusted_participant_icc[[1L]])`
because it excludes fixed-effect variance from its denominator; the
unadjusted ICC was
`r sprintf("%.3f", participant_random_intercept$unadjusted_participant_icc[[1L]])`
on the total model-based variance scale. A one-SD participant intercept was a
`r sprintf("%.2f", participant_random_intercept$participant_factor_per_sd[[1L]])`-fold
conditional expected mean relative to a participant at the model-average
intercept while holding site and light source fixed. This is a common
participant multiplier across categories, not a participant-specific category
response. The variance decomposition is evaluated over observed
participant-hours and is not a site-average estimand.

Marginal R² has no unique term decomposition. In the hierarchy-respecting
Shapley analysis, the study-site component was
`r sprintf("%.3f", participant_random_intercept_shapley$marginal_r_squared_component[participant_random_intercept_shapley$component_id == "site"])`
(`r sprintf("%.1f", participant_random_intercept_shapley$share_of_full_marginal_r_squared_percent[participant_random_intercept_shapley$component_id == "site"])`%
of the full marginal R²), the light-source component was
`r sprintf("%.3f", participant_random_intercept_shapley$marginal_r_squared_component[participant_random_intercept_shapley$component_id == "light_source"])`
(`r sprintf("%.1f", participant_random_intercept_shapley$share_of_full_marginal_r_squared_percent[participant_random_intercept_shapley$component_id == "light_source"])`%),
and the site-by-light-source interaction component was
`r sprintf("%.3f", participant_random_intercept_shapley$marginal_r_squared_component[participant_random_intercept_shapley$component_id == "site_by_light_source"])`
(`r sprintf("%.1f", participant_random_intercept_shapley$share_of_full_marginal_r_squared_percent[participant_random_intercept_shapley$component_id == "site_by_light_source"])`%).
The three point allocations sum to the full marginal R². The analysis averages
both hierarchy-valid entry orders for site and light source and introduces the
interaction only after both main effects. It is consequently invariant to the
chosen reference site and light-source category. It remains a descriptive
allocation tied to this model-based R² definition and nested-model set, not a
unique or causal partition of exposure variance.

The fit returned convergence code 0 without warnings, had a positive-definite
Hessian, and was not singular. Its lag-one Pearson residual correlation was
`r sprintf("%.3f", participant_random_intercept_diagnostics$lag1_pearson_residual_correlation[[1L]])`,
so a participant intercept did not remove all within-participant temporal
dependence. Exact zeros comprised
`r sprintf("%.1f", 100 * participant_random_intercept_diagnostics$observed_zero_fraction[[1L]])`%,
compared with
`r sprintf("%.1f", 100 * participant_random_intercept_diagnostics$tweedie_implied_zero_fraction[[1L]])`%
under the fitted Tweedie distribution. The variance assessment is therefore
descriptive, not evidence that the model predicts individual participant-hours
well. Its Nakagawa R² values are not numerically interchangeable with the
quasi-deviance R² and fixed-effect allocations in the preceding table.
All five models used for the Shapley value function likewise converged without
warnings, had positive-definite Hessians, and were non-singular.

[Random-intercept assessment estimates](../../artifacts/09_tables/H03/H03_near_eye_participant_random_intercept_summary.csv)
and [model diagnostics](../../artifacts/08_diagnostics/H03/H03_near_eye_participant_random_intercept_diagnostics.csv),
together with the [hierarchy-respecting Shapley allocation](../../artifacts/09_tables/H03/H03_near_eye_participant_random_intercept_marginal_r2_shapley.csv)
and [nested-model checks](../../artifacts/08_diagnostics/H03/H03_near_eye_participant_random_intercept_shapley_models.csv),
provide the complete stored point estimates and checks.

## Model checks {#model-checks}

Both population-mean fits converged without warnings, retained full design
rank, and had finite positive-definite participant-cluster covariances. Their
working distribution nevertheless does not reproduce important features of
individual-hour data.

```{r}
#| label: tbl-h03-primary-diagnostics
#| tbl-cap: "Numerical, residual, temporal, and zero-mass model checks."

diagnostics |>
  filter(.data$scenario_id == "primary_dataset") |>
  left_join(
    zero_mass |>
      filter(.data$scope == "overall") |>
      select(
        .data$placement,
        working_zero = .data$working_expected_zero_fraction
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Converged = .data$converged,
    `Full rank` = .data$full_rank,
    `Robust covariance positive definite` =
      .data$covariance_positive_definite,
    `|Residual|-fitted Spearman` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `Lag-1 residual correlation` = sprintf(
      "%.3f",
      .data$residual_lag1_correlation
    ),
    `Observed zero fraction` = sprintf("%.3f", .data$exact_zero_fraction),
    `Working zero fraction` = sprintf("%.3f", .data$working_zero)
  ) |>
  gt::gt() |>
  h03_gt(11)
```

Hours close together in time still show similar unexplained deviations: the
technical lag-one residual correlations are 0.478 near eye and 0.615 at chest.
Participant clustering protects coefficient uncertainty against arbitrary
dependence within a participant, but it does not whiten residuals or turn the
working Tweedie family into a calibrated probability model.

The discrepancy is clearest for zeros. Exact zeros comprise 27.8% of hours,
whereas the working Tweedie relationship implies 85.2% near eye and 94.8% at
chest. The model is therefore interpreted only as a log-mean model with
participant-robust inference, not as an individual-hour prediction model or a
model of zero probability.

```{r}
#| label: fig-h03-primary-diagnostics
#| fig-cap: "Primary-model residual, temporal-dependence, and zero-mass checks. The melEDI transformation is display-only."
#| fig-alt: "Three-panel model-check plot. Individual residuals show mean-dependent spread, residual correlations remain positive across hourly lags, and observed zero fractions are much lower than the working Tweedie fractions."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_primary_residual_diagnostics.png"
))
```

Deleting each available site and the five participants with the largest score
contributions at each placement produced 27 refits. No category-omnibus
decision changed, and 146 of 149 estimable non-reference category ratios retained the
full-data ratio within the deletion-refit interval. The three exceptions were
localized chest results: sleep darkness after excluding Kumasi (GH), external
light during sleep after excluding Munich (DE), and sleep darkness after
excluding one Kumasi (GH) participant.

## Sensitivity analyses

The sensitivity analyses make one defined change at a time: they use the same
participants and participant-hours available at both sensor positions; use the
gap-timing-unaware dataset defined above; exclude hours spanning a diary-state
or measurement-context boundary; exclude site-category cells that fail the
predefined support rule; set the quasi-Tweedie working variance power to 1.30
or 1.80; or add participant-level category proportions to separate within-
participant from between-participant information.

```{r}
#| label: tbl-h03-sensitivities
#| tbl-cap: "Samples, omnibus tests, and category stability across predefined sensitivities."

sensitivity_counts <- sensitivity |>
  group_by(.data$run_id, .data$placement, .data$scenario_id) |>
  summarise(
    Stable = sum(.data$stability == "stable"),
    `Support-non-estimable` = sum(
      .data$stability == "support_non_estimable"
    ),
    `Direction shifts` = sum(.data$stability == "direction_shift"),
    `Magnitude shifts` = sum(.data$stability == "magnitude_shift"),
    .groups = "drop"
  )

sensitivity_counts |>
  left_join(
    samples |>
      select(
        .data$run_id,
        .data$observations,
        .data$participants,
        .data$participant_days,
        .data$sites
      ),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  left_join(
    sensitivity_omnibus |>
      select(.data$run_id, omnibus_p = .data$p_raw),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Scenario = recode(
      .data$scenario_id,
      paired_common_sample = "Same-participant, same-hour sample",
      gap_timing_unaware_dataset = "Gap-timing-unaware dataset",
      boundary_hours_excluded = "Boundary hours excluded",
      unsupported_cells_excluded = "Unsupported cells excluded",
      working_power_1_30 = "Working power 1.30",
      working_power_1_80 = "Working power 1.80",
      mundlak_within_between = "Within/between participant"
    ),
    `Raw omnibus p` = format_p_cell(
      .data$omnibus_p,
      .data$omnibus_p <= 0.05
    )
  ) |>
  arrange(.data$Placement, .data$Scenario) |>
  select(
    Placement,
    Scenario,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Raw omnibus p`,
    Stable,
    `Support-non-estimable`,
    `Direction shifts`,
    `Magnitude shifts`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(
      Hours, Participants, `Participant-days`, Sites, Stable,
      `Support-non-estimable`, `Direction shifts`, `Magnitude shifts`
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_markdown(columns = `Raw omnibus p`) |>
  gt::tab_source_note(
    gt::md("Each raw omnibus p-value is bold at its stated 0.050 label.")
  ) |>
  h03_gt(10.5)
```

The gap-timing-unaware dataset retained 17,616 near-eye and 19,220 chest
hours; all seven category rows were stable at both placements. Excluding
boundary hours or unsupported cells and changing the working variance power
also left every omnibus decision unchanged. The sparse outdoor-electric
contrast became non-estimable in three reduced samples and crossed the null
direction under two alternatives, but no alternative established a non-null
outdoor-electric contrast. The within/between-participant sensitivity changed
the magnitude of the external-light-during-sleep contrast without reversing
its direction.

### Same-participant, same-hour placement comparison

This comparison uses the same participants and participant-hours at both
sensor positions: 14,103 matched participant-hours from 111 participants, 631
participant-days, and eight sites. Separately fitted placement estimates do
not constitute an equivalence test or a direct test of a sensor-position
effect. Outdoor electric light has only 162 matched hours and fails its
prespecified support rule. The remaining contrasts have the same broad
direction at both positions. The identity line is a visual concordance
reference, not an equivalence boundary.

```{r}
#| label: fig-h03-paired-placement
#| fig-cap: "Same-participant, same-hour near-eye and chest category ratios. The identity line does not define equivalence."
#| fig-alt: "Log-log scatter plot of near-eye against chest category ratios to indoor electric light using the same participants and participant-hours at both sensor positions. Daylight categories lie above one at both positions, while display and sleep categories lie below one."
#| out-width: "72%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_paired_placement_comparison.png"
))
```

[Same-participant, same-hour source data](../../artifacts/09_tables/H03/H03_paired_placement_comparison.csv)
give both component intervals and exact category-specific support.

## Exploratory nonlinear time-of-day context

::: {.callout-note title="Exploratory analysis"}
This nonlinear generalized additive model (GAM) analysis is separate from the
preregistered category inference. The association with clock time is allowed
to bend across the day rather than follow a straight line. The model uses one
global cyclic time-of-day smooth plus sum-to-zero light-source and site
deviations, participant curves, participant-day shifts, and a boundary-aware
AR(1) correction. Light-source curves are ratios to the global daily smooth,
not to indoor electric light, and no curve-wide test is made.
:::

The exploratory model uses the untransformed one-hour geometric melEDI outcome
with a fixed-power Tweedie mean and log link. This targets the conditional
arithmetic mean rather than a back-transformed mean of a log-transformed
outcome, which is why its daytime curves are substantially higher than curves
for a geometric-scale centre. Site, participant, and participant-day smooths
remain in the fitted model but are set to zero in the displayed global and
category curves.

The temporal model is closer to the primary model's raw-outcome mean target,
but the two displays are not numerically interchangeable. The primary scalar
averages over all contributed times and forms a site-average on its fitted
log-mean scale. The temporal curves condition on clock time and show only the
global and light-source components. Their values therefore need not average
to the primary category estimates.

```{r}
#| label: fig-h03-temporal-near-eye
#| fig-cap: "Exploratory near-eye time-of-day means, ratios to the global smooth, and available participant-hours."
#| fig-alt: "Three-row, seven-category near-eye figure. The first row shows time-specific melEDI means and a global dashed curve, the second category-to-global ratios, and the third participant-hour counts. Filled circles mark supported observed hours, open circles sparse observed hours, and grey gaps zero observations."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_temporal_near_eye.png"
))
```

[Near-eye curve data](../../artifacts/11_source_data/H03/H03_reader_temporal_near_eye_curves.csv),
[ratio data](../../artifacts/11_source_data/H03/H03_reader_temporal_near_eye_ratios.csv),
and [clock-time support](../../artifacts/11_source_data/H03/H03_reader_temporal_near_eye_support.csv)
contain the plotted values and untransformed melEDI.

```{r}
#| label: fig-h03-temporal-chest
#| fig-cap: "Exploratory chest time-of-day means, ratios to the global smooth, and available participant-hours."
#| fig-alt: "Three-row, seven-category chest figure showing time-specific melEDI means, category-to-global ratios, and participant-hour counts. Filled and open circles distinguish adequate from sparse observed support; grey gaps have no observations or fitted curve."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_temporal_chest.png"
))
```

[Chest curve data](../../artifacts/11_source_data/H03/H03_reader_temporal_chest_curves.csv),
[ratio data](../../artifacts/11_source_data/H03/H03_reader_temporal_chest_ratios.csv),
and [clock-time support](../../artifacts/11_source_data/H03/H03_reader_temporal_chest_support.csv)
contain the plotted values and untransformed melEDI.

### R² and variance allocation

```{r}
#| label: tbl-h03-temporal-fit
#| tbl-cap: "Exploratory temporal-model fit summaries."

temporal_summary |>
  select(
    .data$placement,
    .data$observations,
    .data$participants,
    .data$participant_days,
    .data$rho,
    .data$adjusted_r_squared,
    .data$deviance_explained
  ) |>
  left_join(
    temporal_r_squared |>
      select(.data$placement, balanced_r_squared = .data$r_squared),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Hours = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `AR rho` = sprintf("%.3f", .data$rho),
    `Adjusted R²` = sprintf("%.3f", .data$adjusted_r_squared),
    `Deviance explained` = sprintf("%.3f", .data$deviance_explained),
    `Site-average participant-balanced R²` = sprintf(
      "%.3f",
      .data$balanced_r_squared
    )
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = c(Hours, Participants, `Participant-days`), use_seps = TRUE) |>
  gt::tab_source_note(
    gt::md(paste(
      "The participant-balanced R² is calculated on raw melEDI.",
      "All R² values are descriptive in-sample point estimates without",
      "resampling intervals."
    ))
  ) |>
  h03_gt(11)
```

```{r}
#| label: tbl-h03-temporal-allocation
#| tbl-cap: "Point allocation of fitted temporal linear-predictor variance."

temporal_allocation |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Component = recode(
      .data$group,
      global_time = "Global time of day",
      light_source_deviations = "Light-source deviations",
      site_deviations = "Site deviations",
      participant_curves = "Participant curves",
      participant_day_shifts = "Participant-day shifts"
    ),
    `Allocated share` = sprintf("%.1f%%", 100 * .data$shapley_share),
    `Unique partial share` = sprintf(
      "%.1f%%",
      100 * .data$partial_unique_share
    )
  ) |>
  select(Placement, Component, `Allocated share`, `Unique partial share`) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "Exact point allocation on the natural-log conditional-mean predictor",
      "under site-average, participant-balanced weights. Components",
      "share covariance; no simulation interval was computed."
    ))
  ) |>
  h03_gt()
```

Global time of day received 60.4% of the near-eye and 58.3% of the chest point
allocation. Light-source deviations received 24.2% and 26.4%, site deviations
5.2% and 6.5%, participant curves 7.6% and 5.7%, and participant-day shifts
2.5% and 3.1%. The Shapley allocation averages a component's contribution
across all orders in which components could be added, allocating shared
fitted-model information rather than counting it repeatedly. These are point
summaries without uncertainty intervals and are not independent or causal
percentages of raw melEDI variance.

### Temporal model checks

```{r}
#| label: tbl-h03-temporal-diagnostics
#| tbl-cap: "Convergence, basis-capacity, residual, and zero-mass checks for the temporal models."

temporal_global_k <- temporal_k |>
  filter(.data$term == "s(time_hour)") |>
  select(
    .data$placement,
    global_edf = .data$effective_df,
    global_k_prime = .data$k_prime,
    global_k_index = .data$k_index
  )

temporal_summary |>
  left_join(
    temporal_global_k,
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = .data$placement,
    Converged = .data$converged,
    `Smoothing Hessian positive definite` =
      .data$smoothing_hessian_positive_definite,
    `Maximum |gradient|` = formatC(
      .data$smoothing_gradient_maximum_absolute,
      digits = 2,
      format = "e"
    ),
    `Global edf / k′` = sprintf(
      "%.2f / %d",
      .data$global_edf,
      .data$global_k_prime
    ),
    `Global k-index` = sprintf("%.3f", .data$global_k_index),
    `Lag-1 residual correlation` = sprintf(
      "%.3f",
      .data$standardized_residual_lag1
    ),
    `|Residual|-fitted Spearman` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `Observed / working zero fraction` = sprintf(
      "%.3f / %.3f",
      .data$observed_zero_fraction,
      .data$working_expected_zero_fraction
    )
  ) |>
  gt::gt() |>
  gt::tab_source_note(
    gt::md("The deterministic k-index check used zero permutation replicates.")
  ) |>
  h03_gt(10.5)
```

Both final fits converged at full coefficient rank with positive-definite
smoothing-parameter Hessians. The chest fit with zero AR correlation, used
only to estimate the fixed AR value, did not converge; the final fixed-AR fit
did. The global smooths used 9.36 and 9.47 effective degrees of freedom out of
ten available, with deterministic k-indices of 0.890 and 0.884. The smooths
are therefore close to their basis capacity, and no permutation *p*-value was
computed.

The AR adjustment reduces but does not eliminate the tendency for nearby hours
to retain similar unexplained deviations: the technical lag-one residual
correlations are 0.146 near eye and 0.144 at chest. Residual spread remains
fitted-value dependent, and the working zero fractions of 0.465 and 0.508
exceed the observed fractions of 0.278 and 0.277. These limitations confine
the temporal analysis to descriptive conditional-mean context. The ribbons
are pointwise 95% intervals; they are not simultaneous bands and do not
support curve-wide significance claims.

```{r}
#| label: fig-h03-temporal-diagnostics
#| fig-cap: "Exploratory temporal-model residual, autocorrelation, and zero-mass checks."
#| fig-alt: "Three-part model-check figure. Residual spread changes with fitted melEDI, lag-one residual correlation remains about 0.14 at both placements, and working Tweedie zero fractions exceed observed fractions across fitted-value deciles."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_temporal_diagnostics.png"
))
```

[Residual points](../../artifacts/11_source_data/H03/H03_reader_temporal_residual_points.csv),
[residual bins](../../artifacts/11_source_data/H03/H03_reader_temporal_residual_bins.csv),
[autocorrelations](../../artifacts/08_diagnostics/H03/H03_reader_temporal_residual_acf.csv),
and [zero calibration](../../artifacts/11_source_data/H03/H03_reader_temporal_zero_calibration.csv)
provide the exact model-check source values.

## Exploratory linear latitude context

Latitude cannot be estimated alongside a saturated fixed-site term because
each site has only one latitude. A separate exploratory model therefore
replaces site with category-specific linear **absolute-latitude** slopes:

```r
geo_medi_1h ~ 0 + light_source +
  light_source:absolute_latitude_10deg_centered
```

The coefficient for each category is the multiplicative change in expected
one-hour melEDI per 10° farther from the equator. Centering at the site-average
mean absolute latitude—39.2° near eye and 38.1° at chest—changes the
intercepts but not the slopes. Uncertainty is clustered at site, not
participant, because latitude varies only across nine near-eye and eight chest
sites. Seven slope p-values receive an FDR adjustment within each placement.

```{r}
#| label: tbl-h03-latitude-slopes
#| tbl-cap: "Exploratory category-specific linear absolute-latitude associations."

latitude_slopes |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Category = factor(
      .data$short_label,
      levels = near_interaction$short_label
    ),
    `Ratio per +10° (95% CI)` = format_ratio_ci(
      .data$ratio_per_10deg,
      .data$ratio_conf_low,
      .data$ratio_conf_high
    ),
    `Raw p` = format_p_cell(.data$p_raw, FALSE),
    `FDR-adjusted p` = format_p_cell(
      .data$p_adjusted,
      .data$p_adjusted < 0.05
    ),
    `Leave-one-site-out sign` = paste0(
      .data$sign_agreement,
      "/",
      .data$omissions
    ),
    `Leave-one-site-out ratio range` = sprintf(
      "%.2f–%.2f",
      .data$ratio_min,
      .data$ratio_max
    )
  ) |>
  arrange(.data$Placement, .data$Category) |>
  select(
    Placement,
    Category,
    `Ratio per +10° (95% CI)`,
    `Raw p`,
    `FDR-adjusted p`,
    Sites = .data$category_sites,
    `Leave-one-site-out sign`,
    `Leave-one-site-out ratio range`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_markdown(columns = c(`Raw p`, `FDR-adjusted p`)) |>
  gt::fmt_integer(columns = Sites) |>
  gt::tab_source_note(
    gt::md(paste(
      "Ratios and site-cluster HC1 95% intervals are coefficient-wise",
      "ecological summaries. FDR adjustment forms one seven-slope family",
      "within each placement. The omission columns show how often the slope",
      "kept its full-model direction and its point-estimate range."
    ))
  ) |>
  h03_gt(10.5)
```

Near eye showed FDR-labelled positive gradients for indoor electric light
(ratio 1.197 per +10°, 95% CI 1.081–1.326), indoor daylight (1.253,
1.131–1.388), outdoor daylight (1.354, 1.105–1.661), and sleep darkness
(1.428, 1.076–1.895). Each retained a positive slope in all nine
leave-one-site-out fits. At chest, only indoor electric light retained the FDR
label (1.142, 1.061–1.229) and remained positive in all eight omissions.

Both latitude models converged without warnings and retained design rank
14/14. Their site-cluster covariance ranks were necessarily limited to eight
and seven because only nine and eight independent site clusters were
available. Accordingly, the report gives coefficient-wise slope tests and no
joint seven-slope Wald test.

```{r}
#| label: fig-h03-latitude-slopes
#| fig-cap: "Exploratory factor change in expected melEDI per 10° increase in absolute latitude; points filled in their light-source colour pass the seven-slope within-placement FDR label."
#| fig-alt: "Two-panel forest plot of category-specific ratios per ten degrees farther from the equator. Point outlines and intervals use the corresponding light-source colour. Four near-eye slopes—indoor electric, indoor daylight, outdoor daylight, and sleep darkness—are filled in that colour because their FDR-adjusted p-values are below 0.050. Only indoor electric is filled at chest. Outdoor electric, display, and external-light intervals are wider."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_reader_latitude_category_slopes.png"
))
```

[Latitude result data](../../artifacts/09_tables/H03/H03_reader_latitude_category_slopes.csv),
[site latitudes](../../artifacts/11_source_data/H03/H03_reader_latitude_site_support.csv),
[leave-one-site-out fits](../../artifacts/08_diagnostics/H03/H03_reader_latitude_leave_one_site_out.csv),
and [model checks](../../artifacts/08_diagnostics/H03/H03_reader_latitude_model_diagnostics.csv)
provide the exact exploratory outputs.

These are ecological associations across a small number of sites. Absolute
latitude may stand in for season, climate, built environment, behaviour, or
other site differences; the analysis does not identify a causal latitude
effect or show that latitude adequately replaces the supported
category-by-site interaction.

## Interpretation

Self-reported primary light source is strongly associated with measured
one-hour personal light exposure. Daylight, especially outdoors, corresponds
to substantially higher melEDI than indoor electric light, whereas display
and sleep-environment categories correspond to lower values. The magnitude of
these contrasts varies materially by site. Outdoor electric is especially
sparse and site-dependent, so its cross-site descriptive value depends
strongly on whether a common category contrast is imposed.

The result is observational. It does not show that changing a reported light
source causes a particular exposure, predict a health outcome, or establish
equivalence between near-eye and chest measurements. Participant-clustered
inference, sensitivity analyses, and deletion checks support the overall
category conclusion, while residual and zero-mass model checks limit claims
about the working Tweedie distribution. The nonlinear time-of-day GAMs add
useful descriptive context but remain separate from the preregistered
inference. The linear absolute-latitude context identifies several exploratory
ecological gradients, but the small number of sites and site-level confounding
preclude a causal latitude interpretation.

## Preregistration deviations {#h03-preregistration-deviations}

- **Study-site structure ([DEV-013](../preregistration_deviations.qmd#dev-013)).**
  The registered random-effects architecture was not retained because site
  support did not justify the registered random-slope formulation and the
  participant-day component collapsed. The accepted analysis uses fixed
  study-site adjustment for the population-average category association and
  a separate category-by-site interaction model.
- **Response model ([DEV-014](../preregistration_deviations.qmd#dev-014)).**
  The registered Gaussian response formulation had inadequate distributional
  diagnostics. The accepted fixed-power quasi-Tweedie log-mean models with
  participant-clustered covariance accommodate the non-negative, right-skewed
  response and exact zeros. They are mean models, not calibrated
  zero-generating distributions.
- **Light-source categories ([DEV-022](../preregistration_deviations.qmd#dev-022)).**
  The analysis retains all seven prespecified categories under one
  placement-independent support rule and labels unsupported site-category
  cells rather than silently deleting categories.
- **Primary and site-interaction questions ([DEV-023](../preregistration_deviations.qmd#dev-023)).**
  The additive model supplies the primary six-restriction category omnibus;
  the separate category-by-site interaction model supplies descriptive means,
  ratios, and site-specific variation.
- **Multiple-testing adjustment ([DEV-024](../preregistration_deviations.qmd#dev-024)).**
  False-discovery-rate adjustment is applied to each explicitly assembled
  complete contrast family; category and site-interaction questions remain
  separate families.

## Sources, technical formulas, and provenance

```{r}
#| label: tbl-h03-formulas
#| tbl-cap: "Exact formulas supplied to the primary, category-by-site interaction, and exploratory models."

tibble::tibble(
  Model = c(
    "Primary population mean",
    "Near-eye category-by-site interaction",
    "Chest category-by-site interaction",
    "Exploratory latitude context",
    "Exploratory nonlinear time of day"
  ),
  `Exact Wilkinson formula supplied to R` = c(
    "geo_medi_1h ~ site + light_source",
    "geo_medi_1h ~ site * light_source",
    "geo_medi_1h ~ 0 + site_source_cell",
    "geo_medi_1h ~ 0 + light_source + light_source:absolute_latitude_10deg_centered",
    temporal_summary$formula[[1L]]
  )
) |>
  gt::gt() |>
  gt::cols_width(
    Model ~ gt::pct(23),
    `Exact Wilkinson formula supplied to R` ~ gt::pct(77)
  ) |>
  h03_gt(12)
```

Metric derivation and model-ready dataset construction are documented in
[Preparation 04](../preparation/04_metric_derivation.qmd) and
[Preparation 06](../preparation/06_model_ready_datasets.qmd). [How the H03
analysis was prepared](../../audit/hypotheses/H03/H03_analysis_preparation.qmd)
traces the verified inputs, model frames, formulas, model checks,
sensitivities, stored outputs, and reproducibility details underlying this
report.
