---
title: "H04: Self-reported activity and measured personal light exposure"
subtitle: "Associations with one-hour melanopic equivalent daylight illuminance"
page-layout: full
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
  error: false
---

```{r}
#| label: setup-h04-reader
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

root <- Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
if (!nzchar(root) || !file.exists(file.path(root, "renv.lock"))) {
  root <- locate_project_root()
}
root <- normalizePath(root, winslash = "/", mustWork = TRUE)
stopifnot(identical(as.character(getRversion()), "4.6.1"))

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

source(file.path(root, "scripts/pipeline/p_value_display.R"))

artifact <- function(...) file.path(root, "artifacts", ...)
read_h04 <- function(...) {
  readr::read_csv(artifact(...), show_col_types = FALSE, na = "")
}

formulas <- read_h04("06_model_data", "H04", "H04_formula_registry.csv")
samples <- read_h04("06_model_data", "H04", "H04_model_frame_index.csv")
sample_flow <- read_h04("06_model_data", "H04", "H04_sample_flow.csv")
category_support <- read_h04(
  "06_model_data", "H04", "H04_category_support.csv"
)
primary <- read_h04(
  "09_tables", "H04", "H04_primary_category_estimands.csv"
)
tests <- read_h04(
  "09_tables", "H04", "H04_primary_and_heterogeneity_tests.csv"
)
site_activity <- read_h04(
  "09_tables", "H04", "H04_site_activity_estimands.csv"
)
diagnostics <- read_h04(
  "08_diagnostics", "H04", "H04_model_diagnostics.csv"
)
assessments <- read_h04(
  "08_diagnostics", "H04", "H04_diagnostic_assessments.csv"
)
sensitivity <- read_h04(
  "09_tables", "H04", "H04_sensitivity_comparison.csv"
)
sensitivity_tests <- read_h04(
  "09_tables", "H04", "H04_sensitivity_omnibus_tests.csv"
)
mundlak_between <- read_h04(
  "09_tables", "H04", "H04_mundlak_between_participant_estimands.csv"
)
mundlak_between_omnibus <- read_h04(
  "09_tables", "H04", "H04_mundlak_between_participant_omnibus.csv"
)
mundlak_support <- read_h04(
  "08_diagnostics", "H04", "H04_mundlak_activity_support.csv"
)
paired <- read_h04(
  "09_tables", "H04", "H04_paired_placement_estimands.csv"
)
temporal_summary <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_model_summary.csv"
)
temporal_retention <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_retention_decision.csv"
)
temporal_uncertainty <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_uncertainty_contract.csv"
)
heterogeneity_category <- read_h04(
  "09_tables", "H04", "H04_reader_heterogeneity_category_estimands.csv"
)
reader_category <- read_h04(
  "09_tables", "H04", "H04_reader_category_estimands.csv"
)
heterogeneity_r_squared <- read_h04(
  "09_tables", "H04", "H04_reader_heterogeneity_r_squared.csv"
)
participant_random_intercept <- read_h04(
  "09_tables", "H04", "H04_participant_random_intercept_summary.csv"
)
participant_random_intercept_shapley <- read_h04(
  "09_tables", "H04",
  "H04_participant_random_intercept_marginal_r2_shapley.csv"
)
participant_random_intercept_diagnostics <- read_h04(
  "08_diagnostics", "H04",
  "H04_participant_random_intercept_diagnostics.csv"
)
participant_random_intercept_models <- read_h04(
  "08_diagnostics", "H04",
  "H04_participant_random_intercept_shapley_models.csv"
)
primary_zero_mass <- read_h04(
  "08_diagnostics", "H04", "H04_reader_primary_zero_mass.csv"
)
temporal_r_squared <- read_h04(
  "09_tables", "H04", "H04_reader_temporal_weighted_r_squared.csv"
)
temporal_allocation <- read_h04(
  "09_tables", "H04", "H04_reader_temporal_variance_allocation.csv"
)
temporal_zero_mass <- read_h04(
  "08_diagnostics", "H04", "H04_reader_temporal_zero_mass.csv"
)

format_p <- function(value) nh_format_p_value(value)
format_p_cell <- function(value, significant = FALSE) {
  ifelse(
    is.na(value),
    "—",
    {
      display <- nh_p_value_display(value, significant = significant)
      ifelse(
        display$p_bold,
        paste0("**", display$p_display, "**"),
        display$p_display
      )
    }
  )
}
format_mean_ci <- function(estimate, low, high, digits = 1L) {
  paste0(
    formatC(estimate, digits = digits, format = "f", big.mark = ","),
    " (",
    formatC(low, digits = digits, format = "f", big.mark = ","),
    "–",
    formatC(high, digits = digits, format = "f", big.mark = ","),
    ")"
  )
}
format_ratio_ci <- function(estimate, low, high, digits = 3L) {
  paste0(
    formatC(estimate, digits = digits, format = "f"),
    " (",
    formatC(low, digits = digits, format = "f"),
    "–",
    formatC(high, digits = digits, format = "f"),
    ")"
  )
}
format_difference_ci <- function(estimate, low, high, digits = 1L) {
  paste0(
    formatC(estimate, digits = digits, format = "f", big.mark = ","),
    " (",
    formatC(low, digits = digits, format = "f", big.mark = ","),
    "–",
    formatC(high, digits = digits, format = "f", big.mark = ","),
    ")"
  )
}
h04_gt <- function(table, font_size = 13) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

main_samples <- samples |>
  filter(.data$scenario_id == "primary_dataset") |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
main_flow <- sample_flow |>
  filter(!is.na(.data$outcome_grid_hours)) |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
primary_tests <- tests |>
  filter(.data$test_id %in% c("H04-F1", "H04-F1b"))
heterogeneity_tests <- tests |>
  filter(.data$test_id == "H04-F3")
near_primary_test <- primary_tests |>
  filter(.data$placement == "Near-eye", .data$test_id == "H04-F1")
chest_primary_test <- primary_tests |>
  filter(.data$placement == "Chest", .data$test_id == "H04-F1")
main_diagnostics <- diagnostics |>
  filter(.data$scenario_id == "primary_dataset") |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
overall_assessment <- assessments |>
  filter(.data$diagnostic == "Overall primary mean-model assessment") |>
  select("placement", "assessment")
paired_samples <- samples |>
  filter(.data$scenario_id == "paired_common_sample") |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
temporal_activity <- temporal_summary |>
  filter(.data$formula_id == "temporal_activity_long") |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)

reader_activity_codes <- c(
  "home", "working_indoor", "outdoors", "road_vehicle", "sleeping", "other"
)
reader_activity_labels <- c(
  home = "At home",
  working_indoor = "Office/home working",
  outdoors = "Outdoors",
  road_vehicle = "Vehicle/public transport",
  sleeping = "Sleeping",
  other = "Other/unspecified activity"
)
reader_activity_order <- stats::setNames(
  seq_along(reader_activity_codes),
  reader_activity_codes
)
named_codes <- c("working_indoor", "outdoors", "road_vehicle", "sleeping")
named_sensitivity <- sensitivity |>
  filter(.data$activity_code %in% .env$named_codes)
sensitivity_summary <- named_sensitivity |>
  group_by(.data$scenario_id, .data$placement) |>
  summarise(
    stable = sum(.data$stability == "stable"),
    maximum_absolute_ratio_change_percent = max(
      abs(.data$ratio_relative_change_percent),
      na.rm = TRUE
    ),
    all_directions_concordant = all(.data$direction_concordant),
    primary_ratio_inside_sensitivity_interval = sum(
      .data$primary_ratio_inside_sensitivity_interval
    ),
    .groups = "drop"
  )
sensitivity_f1 <- sensitivity_tests |>
  filter(.data$test_id == "H04-F1") |>
  select(
    "scenario_id",
    "placement",
    "f_statistic",
    "denominator_df",
    "p_raw",
    "status"
  )
mundlak_within <- named_sensitivity |>
  filter(.data$scenario_id == "mundlak_within_between")
mundlak_display <- mundlak_within |>
  select(
    "placement",
    "activity_code",
    "primary_ratio",
    "primary_conf_low",
    "primary_conf_high",
    within_ratio = "ratio_to_home",
    within_conf_low = "ratio_conf_low",
    within_conf_high = "ratio_conf_high",
    within_p_adjusted = "p_adjusted",
    "ratio_relative_change_percent"
  ) |>
  left_join(
    mundlak_between |>
      filter(.data$inferential_role == "NAMED_COMPOSITION_VERSUS_HOME") |>
      select(
        "placement",
        "activity_code",
        between_ratio = "ratio_per_change",
        between_conf_low = "ratio_conf_low",
        between_conf_high = "ratio_conf_high",
        between_p_adjusted = "p_adjusted"
      ),
    by = c("placement", "activity_code"),
    relationship = "one-to-one"
  )

scenario_labels <- c(
  exactly_one_category = "Exactly one category",
  retain_coselected_other = "Retain co-selected Other",
  exclude_other_only = "Exclude Other-only hours",
  unweighted_long_rows = "Unweighted long rows",
  gap_timing_unaware = "Gap-timing-unaware data",
  paired_common_sample = "Placement-matched common sample",
  working_power_1_30 = "Working power 1.30",
  working_power_1_80 = "Working power 1.80",
  mundlak_within_between = "Within/between participant"
)

answer_category_line <- function(activity_code) {
  row <- reader_category |>
    filter(
      .data$placement == "Near-eye",
      .data$activity_code == .env$activity_code
    )
  stopifnot(nrow(row) == 1L)
  comparison <- if (identical(activity_code, "home")) {
    "reference category"
  } else {
    paste0(
      "ratio to At home ",
      format_ratio_ci(
        row$ratio_to_home,
        row$ratio_conf_low,
        row$ratio_conf_high
      )
    )
  }
  paste0(
    "**", unname(reader_activity_labels[[activity_code]]), ":** ",
    format_mean_ci(
      row$standardized_mean_lx,
      row$mean_conf_low_lx,
      row$mean_conf_high_lx
    ),
    " lx; ", comparison, "."
  )
}

figure_files <- file.path(
  artifact("10_figures", "H04"),
  c(
    "H04_reader_heterogeneity_category_estimates.png",
    "H04_site_activity_estimates.png",
    "H04_primary_diagnostics.png",
    "H04_paired_placement_comparison.png",
    "H04_temporal_near_eye.png",
    "H04_temporal_chest.png",
    "H04_temporal_diagnostics.png"
  )
)

stopifnot(
  nrow(main_samples) == 2L,
  identical(as.integer(main_samples$participants), c(126L, 150L)),
  identical(as.integer(main_samples$participant_days), c(724L, 875L)),
  identical(
    as.integer(main_samples$unique_participant_hours),
    c(16526L, 20128L)
  ),
  identical(as.integer(main_samples$long_rows), c(17266L, 21071L)),
  all(abs(
    main_samples$effective_weighted_hours -
      main_samples$unique_participant_hours
  ) < 1e-8),
  nrow(primary) == 12L,
  all(table(primary$placement) == 6L),
  nrow(near_primary_test) == 1L,
  nrow(chest_primary_test) == 1L,
  all(primary_tests$status == "ESTIMABLE"),
  all(heterogeneity_tests$status == "ESTIMABLE"),
  all(named_sensitivity$stability == "stable"),
  nrow(mundlak_within) == 8L,
  nrow(mundlak_display) == 8L,
  nrow(mundlak_between_omnibus) == 2L,
  all(mundlak_between_omnibus$status == "ESTIMABLE"),
  nrow(mundlak_support) == 10L,
  nrow(paired_samples) == 2L,
  all(paired_samples$unique_participant_hours == 14308L),
  nrow(temporal_activity) == 2L,
  nrow(heterogeneity_category) == 10L,
  nrow(reader_category) == 12L,
  all(table(reader_category$placement) == 6L),
  sum(reader_category$model_source == "site_heterogeneity") == 10L,
  sum(reader_category$model_source == "additive_display_only") == 2L,
  nrow(heterogeneity_r_squared) == 6L,
  nrow(participant_random_intercept) == 2L,
  nrow(participant_random_intercept_shapley) == 6L,
  nrow(participant_random_intercept_diagnostics) == 2L,
  nrow(participant_random_intercept_models) == 10L,
  all(participant_random_intercept_diagnostics$converged),
  all(participant_random_intercept_diagnostics$warning_count == 0L),
  all(participant_random_intercept_models$converged),
  all(participant_random_intercept_models$warning_count == 0L),
  nrow(primary_zero_mass) == 2L,
  nrow(temporal_r_squared) == 2L,
  nrow(temporal_allocation) == 10L,
  nrow(temporal_zero_mass) == 2L,
  all(temporal_activity$converged),
  all(temporal_retention$retained_for_context),
  all(temporal_uncertainty$resampling_replicates == 0L),
  all(!temporal_uncertainty$simultaneous_band),
  all(!temporal_uncertainty$curve_wide_inference),
  all(file.exists(figure_files))
)
```

## Question

The preregistered hypothesis was:

> **H4: Hourly self-reported activity categories predict hourly geometric mean melanopic EDI.**

This analysis asks whether five interpretable reported activity categories are
associated with measured one-hour melanopic equivalent daylight illuminance
(melEDI), a melanopsin-weighted measure of illuminance expressed in lux, after
accounting for study site. The categories are not mutually exclusive
behavioural states: when an hour has several retained labels, its contribution
is shared equally among them.

::: {.callout-note title="Answer in brief"}
Reported activity category was associated with near-eye melEDI,
*F*(`r near_primary_test$restrictions`, `r near_primary_test$denominator_df`) =
`r sprintf("%.2f", near_primary_test$f_statistic)`, raw *p*
**`r format_p(near_primary_test$p_raw)`**. Site-average estimates give each
study site equal weight. These estimates and 95% confidence intervals (95%
CIs), constructed to allow observations from the same participant to be
related, were:

- `r answer_category_line("home")`
- `r answer_category_line("working_indoor")`
- `r answer_category_line("outdoors")`
- `r answer_category_line("road_vehicle")`
- `r answer_category_line("sleeping")`

Because the activity-by-site interaction was supported, these descriptive
means and ratios come from the activity-by-site interaction model; only the
primary five-category omnibus comes from the additive population-mean model.

Chest estimates showed the same broad ordering. All four named-versus-home
contrasts retained their direction and classification in every prespecified
sensitivity. Other/unspecified activity remains visible as a heterogeneous
display category but supports no scientific claim. These are observational
associations, not causal effects of changing activity.
:::

## Data and activity representation

The near-eye sensor position is primary because it measures light near the
eyes, although it is not a retinal measurement. The chest sensor position is
complementary and is not a measure of ocular exposure. The positions are
fitted separately, and similarity is not interpreted as equivalence.

The response is the zero-aware geometric mean melEDI within each
hour, including genuine zero values. Before the activity data were reshaped,
active travel, outdoor work, and outdoor free time were combined into
**Outdoors**. Duplicate collapsed labels were removed within each
participant-hour. If Other was co-selected with a named category, only the
named category or categories were retained; Other was kept when it was the
sole selection. No diary free text was inspected or recoded.

An hour with $k$ retained distinct labels contributes $k$ rows, each with
weight $1/k$. The total analytical weight of every fitted participant-hour is
therefore one. Hours with no retained category are not interpreted as “no
activity” and are excluded from the category model.

A participant-hour is one original hourly record for one participant. A
participant-day groups that participant's records within one study day.

```{r}
#| label: tbl-h04-overall-samples
#| tbl-cap: "Exact fitted samples for the primary and complementary models."

main_samples |>
  transmute(
    Placement = as.character(.data$placement),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Unique participant-hours` = .data$unique_participant_hours,
    `Generated long rows` = .data$long_rows,
    `Effective weighted hours` = .data$effective_weighted_hours,
    Sites = .data$sites,
    `Exact-zero hours` = .data$exact_zero_unique_hours
  ) |>
  gt::gt() |>
  gt::fmt_integer(
    columns = c(
      Participants, `Participant-days`, `Unique participant-hours`,
      `Generated long rows`, Sites, `Exact-zero hours`
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_number(
    columns = `Effective weighted hours`,
    decimals = 0,
    use_seps = TRUE
  ) |>
  h04_gt(11.5)
```

The primary model contains 126 participants, 724 participant-days, 16,526
unique participant-hours, 17,266 long rows, and 16,526 effective weighted
hours. The complementary model contains 150 participants, 875
participant-days, 20,128 unique participant-hours, 21,071 long rows, and
20,128 effective weighted hours.

```{r}
#| label: tbl-h04-sample-flow
#| tbl-cap: "Outcome availability and explicit accounting of hours outside the category models."

main_flow |>
  transmute(
    Placement = as.character(.data$placement),
    `Outcome-grid hours` = .data$outcome_grid_hours,
    `Outcome unavailable` = .data$outcome_unavailable_hours,
    `No eligible diary key` = .data$outcome_hours_without_eligible_diary_key,
    `No retained activity` = .data$outcome_hours_with_no_retained_activity,
    `Fitted participant-hours` = .data$model_candidate_unique_hours
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  gt::tab_source_note(
    gt::md(paste(
      "No retained activity is not a behavioural category and is not fitted",
      "as a zero or reference activity."
    ))
  ) |>
  h04_gt()
```

```{r}
#| label: tbl-h04-category-support
#| tbl-cap: "Observed support and weighted totals for every displayed activity category."

category_support |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    reader_order = unname(
      .env$reader_activity_order[.data$activity_code]
    ),
    reader_label = unname(
      .env$reader_activity_labels[.data$activity_code]
    )
  ) |>
  arrange(.data$placement, .data$reader_order) |>
  transmute(
    Placement = as.character(.data$placement),
    Category = .data$reader_label,
    `Unique hours` = .data$unique_participant_hours,
    `Long rows` = .data$long_rows,
    `Weighted hours` = .data$effective_weighted_hours,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(
      `Unique hours`, `Long rows`, Participants, `Participant-days`, Sites
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_number(columns = `Weighted hours`, decimals = 1, use_seps = TRUE) |>
  gt::cols_width(
    Category ~ gt::pct(31),
    everything() ~ gt::pct(11.5)
  ) |>
  h04_gt(11.5)
```

Weighted hours can be smaller than category rows because a multi-select
hour's weight is divided among its retained labels. Long-row weighting is part
of the estimand: replacing $1/k$ by unit weight would instead analyse reported
selections and let multi-select hours contribute more.

## Model and estimand

The primary population-mean model estimates the average association between
reported activity and melEDI after accounting for site. It uses fixed site and
activity terms, a quasi-Tweedie variance function with power 1.539919, a log
link, and the exact $1/k$ prior weights. A separate activity-by-site
interaction model allows the association between each of the five
interpretable activity categories and melEDI to differ by study site. It is
used for descriptive category and site summaries, while the additive model
remains the basis for the primary inferential test. Other/unspecified activity
remains unrestricted in the additive model and is display-only.

Participant-cluster-robust uncertainty allows concurrent category memberships
and longitudinal observations from the same participant to be related. The
technical implementation uses HC1 covariance and finite-cluster *t* and *F*
reference distributions, with the number of participant clusters minus one as
denominator degrees of freedom.

The primary equality test compares Sleeping, At home, vehicle or
public-transport travel, indoor or home working, and Outdoors. The fitted
Other coefficient remains unrestricted and does not enter that four-row joint
restriction. A full six-category omnibus is reported secondarily.

Category means are reported as site-average estimates across the sites
observed for that sensor position. Every site contributes equally on the
fitted log-mean scale. The resulting average is then back-transformed, meaning
that it is returned to melEDI in lux. It is therefore the geometric mean of
the fitted site-specific means, not an observation-weighted arithmetic mean.
At home is the reader-facing comparison reference.

The estimand is average melEDI associated with each reported collapsed
activity category, with an hour's contribution shared across concurrently
reported categories. It is not a transition effect, an individual-specific
effect, or a causal contrast between mutually exclusive behaviours.

### Preregistration changes {#h04-preregistration-deviations}

- **Study-site structure ([DEV-013](../preregistration_deviations.qmd#dev-013)).**
  The accepted population-average category model uses fixed study-site
  adjustment, while a separate activity-by-site interaction model evaluates
  variation among sites. The site interaction does not redefine the primary
  omnibus test.
- **Response model ([DEV-014](../preregistration_deviations.qmd#dev-014)).**
  Fixed-power quasi-Tweedie log-mean models with participant-clustered
  covariance replace the registered Gaussian mixed model. They describe
  multiplicative mean comparisons and are not treated as calibrated
  zero-generating distributions.
- **Multi-label activity representation ([DEV-025](../preregistration_deviations.qmd#dev-025)).**
  The analysis collapses the three outdoor flags before pivoting, removes
  duplicate labels within an hour, gives each of $k$ concurrent retained
  categories weight $1/k$, and retains Other only when it is the sole
  selected category for descriptive display.
- **Primary and site-interaction questions ([DEV-026](../preregistration_deviations.qmd#dev-026)).**
  The additive site-plus-activity model supplies the primary activity
  omnibus; the separate activity-by-site interaction model supplies
  descriptive category means, ratios, and site-specific variation.
- **Multiple-testing adjustment ([DEV-027](../preregistration_deviations.qmd#dev-027)).**
  False-discovery-rate adjustment is applied to each explicitly assembled
  complete activity or site-interaction contrast family.

## Principal activity-category estimates

The additive primary model forces every site to share one activity ratio,
whereas the activity-by-site interaction model estimates site-specific
category means before giving every observed site equal weight on the fitted
log-mean scale. The preregistered additive-model omnibus remains the primary
inferential test, but the descriptive means, ratios, differences, 95% CIs,
and site summaries below use the less restrictive interaction model.
Other/unspecified activity is the sole exception: it is excluded from the
interaction model and remains visible only as a quiet additive-model
display-only estimate.

```{r}
#| label: fig-h04-primary-estimates
#| fig-cap: "Site-average one-hour melEDI estimates from the activity-by-site interaction model and named-category ratios versus At home; bars are participant-cluster-robust 95% CIs. Other/unspecified activity is an additive-model display-only estimate shown with an open point."
#| fig-alt: "Two-panel point-and-interval plot ordered from the At home reference through Office/home working, Outdoors, Vehicle/public transport, and Sleeping, with Other/unspecified activity last. The upper panel shows site-average melEDI estimates from the activity-by-site interaction model at the primary near-eye and complementary chest sensor positions: Sleeping is lowest and Outdoors highest. Other/unspecified activity is marked by an open point and comes from the additive model for display only. The lower panel shows interaction-model ratios for the four named categories versus At home on a logarithmic scale; Sleeping is below one and the other three categories are above one."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_reader_heterogeneity_category_estimates.png"
))
```

```{r}
#| label: tbl-h04-primary-results
#| tbl-cap: "Site-average activity-category means, ratios, and differences with participant-cluster-robust 95% CIs."

primary_results <- reader_category |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    mean_ci = format_mean_ci(
      .data$standardized_mean_lx,
      .data$mean_conf_low_lx,
      .data$mean_conf_high_lx
    ),
    ratio_ci = case_when(
      .data$activity_code == "home" ~ "1.000 (reference)",
      .data$activity_code == "other" ~ "Display only",
      TRUE ~ format_ratio_ci(
        .data$ratio_to_home,
        .data$ratio_conf_low,
        .data$ratio_conf_high
      )
    ),
    difference_ci = case_when(
      .data$activity_code == "home" ~ "0.0 (reference)",
      .data$activity_code == "other" ~ "Display only",
      TRUE ~ format_difference_ci(
        .data$difference_from_home_lx,
        .data$difference_conf_low_lx,
        .data$difference_conf_high_lx
      )
    ),
    reader_order = unname(
      .env$reader_activity_order[.data$activity_code]
    ),
    reader_label = unname(
      .env$reader_activity_labels[.data$activity_code]
    )
  ) |>
  arrange(.data$placement, .data$reader_order) |>
  transmute(
    Placement = as.character(.data$placement),
    Category = .data$reader_label,
    `Mean melEDI, lx (95% CI)` = .data$mean_ci,
    `Ratio to At home (95% CI)` = .data$ratio_ci,
    `Difference from At home, lx (95% CI)` = .data$difference_ci
  )

primary_results |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_style(
    style = gt::cell_text(color = "#666666", style = "italic"),
    locations = gt::cells_body(
      rows = Category == "Other/unspecified activity"
    )
  ) |>
  gt::tab_source_note(
    gt::md(paste(
      "Each named-category estimate gives every observed site equal weight",
      "on the fitted log-mean scale. Ratios compare geometric means with At",
      "home; differences are absolute melEDI contrasts in lux.",
      "Other/unspecified activity is displayed quietly, its ratio and",
      "difference cells say Display only, and its mean comes from the",
      "additive model rather than the activity-by-site interaction model.",
      "No scientific contrast or claim is based on that row."
    ))
  ) |>
  gt::cols_width(
    Category ~ gt::pct(26),
    everything() ~ gt::pct(24.7)
  ) |>
  h04_gt(11.5)
```

At the near-eye sensor position, the interaction-model site-average estimate
was 76.4 lx at home, 198.9 lx during office or home working, 714.2 lx
outdoors, 331.2 lx during vehicle or public-transport travel, and 4.3 lx while
sleeping.
Relative to At home, the four named ratios in that order were 2.604, 9.354,
4.338, and 0.056. Chest means in the same category order were 78.0, 202.9,
998.9, 512.9, and 4.7 lx; the corresponding ratios were 2.600, 12.800, 6.573,
and 0.060.

[Estimate data](../../artifacts/09_tables/H04/H04_reader_category_estimands.csv)
and [figure source data](../../artifacts/11_source_data/H04/H04_reader_heterogeneity_category_figure.csv)
retain the unrounded values.

## Supporting contrasts and omnibus tests

The four named comparisons with At home use a false-discovery-rate (FDR)
adjustment as one complete four-member family within each sensor position.

```{r}
#| label: tbl-h04-named-contrasts
#| tbl-cap: "Additive-model named-category comparisons with At home; FDR adjustment forms one four-member family within each sensor position."

primary |>
  filter(.data$activity_code %in% .env$named_codes) |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    ratio_ci = format_ratio_ci(
      .data$ratio_to_home,
      .data$ratio_conf_low,
      .data$ratio_conf_high
    ),
    raw_p = format_p_cell(.data$p_raw, significant = FALSE),
    adjusted_p = format_p_cell(
      .data$p_adjusted,
      significant = .data$p_adjusted < 0.05
    ),
    reader_order = unname(
      .env$reader_activity_order[.data$activity_code]
    ),
    reader_label = unname(
      .env$reader_activity_labels[.data$activity_code]
    )
  ) |>
  arrange(.data$placement, .data$reader_order) |>
  transmute(
    Placement = as.character(.data$placement),
    Category = .data$reader_label,
    `Ratio to At home (95% CI)` = .data$ratio_ci,
    `t statistic` = .data$statistic,
    `Raw p` = .data$raw_p,
    `FDR-adjusted p` = .data$adjusted_p
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_number(columns = `t statistic`, decimals = 2) |>
  gt::fmt_markdown(columns = c(`Raw p`, `FDR-adjusted p`)) |>
  gt::tab_source_note(
    gt::md(paste(
      "Bold values pass the explicitly labelled within-placement",
      "FDR-adjusted 0.050 rule. These inferential contrasts remain from the",
      "preregistered additive model and are separate from the",
      "activity-by-site interaction-model descriptive summaries."
    ))
  ) |>
  gt::cols_width(
    Category ~ gt::pct(31),
    `Ratio to At home (95% CI)` ~ gt::pct(28),
    everything() ~ gt::pct(13.7)
  ) |>
  h04_gt(11.5)
```

### Omnibus tests

```{r}
#| label: tbl-h04-omnibus
#| tbl-cap: "Primary five-named-category and secondary six-category robust omnibus tests."

primary_tests |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Test = ifelse(
      .data$test_id == "H04-F1",
      "Five named categories; Other unrestricted",
      "All six categories (secondary)"
    ),
    p = format_p_cell(
      .data$p_raw,
      significant = .data$test_id == "H04-F1" & .data$p_raw < 0.05
    )
  ) |>
  arrange(.data$placement, .data$test_id) |>
  transmute(
    Placement = as.character(.data$placement),
    Test,
    `F statistic` = .data$f_statistic,
    `Numerator df` = .data$restrictions,
    `Denominator df` = .data$denominator_df,
    `Raw p` = .data$p
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_number(columns = `F statistic`, decimals = 2) |>
  gt::fmt_integer(columns = c(`Numerator df`, `Denominator df`)) |>
  gt::fmt_markdown(columns = `Raw p`) |>
  gt::tab_source_note(
    gt::md(paste(
      "Bold p-values pass the primary raw 0.050 omnibus rule.",
      "The six-category row is secondary and is not used to reinterpret",
      "the primary five-category test."
    ))
  ) |>
  h04_gt(11.5)
```

The complementary chest primary test also rejected equality of the five named
categories, *F*(4, 149) = `r sprintf("%.2f", chest_primary_test$f_statistic)`,
raw *p* `r format_p(chest_primary_test$p_raw)`. This separate result supports
the broad placement consistency; it is not an independent replication or an
equivalence test.

## Site-specific context

The additive population-mean omnibus remains the primary inferential analysis.
The supported activity-by-site interaction model supplies the descriptive
category summaries above and tests whether the association between activity
category and melEDI differs among study sites. It excludes Other and uses only
observed, supported cells; an unsupported site-category contrast is labelled
non-estimable rather than forced into the model. Reader-facing site labels use
the submitted country-coded names: Borås (SE), Delft (NL), Dortmund (DE),
Tübingen (DE), Munich (DE), Madrid (ES), Izmir (TR), San José (CR), and
Kumasi (GH).

```{r}
#| label: tbl-h04-heterogeneity
#| tbl-cap: "Support-qualified activity-by-site interaction test and complete site-deviation family accounting."

heterogeneity_family <- site_activity |>
  group_by(.data$placement) |>
  summarise(
    family_n = max(.data$family_n, na.rm = TRUE),
    bh_significant = sum(
      .data$site_deviation_p_adjusted < 0.05,
      na.rm = TRUE
    ),
    support_non_estimable = sum(
      .data$reporting_status == "SUPPORT_NON_ESTIMABLE"
    ),
    .groups = "drop"
  )

heterogeneity_tests |>
  left_join(heterogeneity_family, by = "placement") |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    p = format_p_cell(.data$p_raw, significant = .data$p_raw < 0.05)
  ) |>
  arrange(.data$placement) |>
  transmute(
    Placement = as.character(.data$placement),
    Architecture = "Five named categories",
    `F statistic` = .data$f_statistic,
    `Numerator df` = .data$restrictions,
    `Denominator df` = .data$denominator_df,
    `Raw p` = .data$p,
    `FDR family members` = .data$family_n,
    `FDR-labelled deviations` = .data$bh_significant,
    `Support non-estimable` = .data$support_non_estimable
  ) |>
  gt::gt() |>
  gt::fmt_number(columns = `F statistic`, decimals = 2) |>
  gt::fmt_integer(
    columns = c(
      `Numerator df`, `Denominator df`, `FDR family members`,
      `FDR-labelled deviations`, `Support non-estimable`
    )
  ) |>
  gt::fmt_markdown(columns = `Raw p`) |>
  gt::tab_source_note(
    gt::md(paste(
      "Bold omnibus p-values pass the explicitly labelled exploratory raw",
      "0.050 rule. Site-deviation labels use a complete FDR family within",
      "each sensor position; Other is excluded."
    ))
  ) |>
  h04_gt(11)
```

The activity-by-site interaction was supported at the near-eye sensor
position, *F*(32, 125) = 8.03, raw *p* <0.001, and at the chest sensor
position, *F*(28, 149) = 11.62, raw *p* <0.001. This indicates that
activity-associated means vary among sites; it is not a basis for post-hoc
site ranking. The complete site-deviation families contain 44 near-eye and 40
chest comparisons, of which 17 and 9 carry the FDR-adjusted label.

```{r}
#| label: tbl-h04-near-site-factorization
#| tbl-cap: "Near-eye site-average category estimates and site-specific deviation ratios from the activity-by-site interaction model."

site_category_levels <- c(
  "At home",
  "Working in the office/from home",
  "Outdoors",
  "On the road with public transport/car",
  "Sleeping"
)

overall_context <- heterogeneity_category |>
  filter(.data$placement == "Near-eye") |>
  mutate(
    ratio_p_display = gsub(
      "<", "&lt;", format_p(.data$ratio_p_adjusted), fixed = TRUE
    )
  ) |>
  transmute(
    Location = "**Site-average estimate**",
    row_order = 0,
    row_type = "overall",
    activity = .data$activity,
    Value = paste0(
      "**",
      format_mean_ci(
        .data$standardized_mean_lx,
        .data$mean_conf_low_lx,
        .data$mean_conf_high_lx
      ),
      " lx**<br><small>",
      if_else(
        .data$activity == "At home",
        "Reference category<br>FDR p (reference)",
        paste0(
          format_ratio_ci(
            .data$ratio_to_home,
            .data$ratio_conf_low,
            .data$ratio_conf_high
          ),
          "× At home<br>FDR p ",
          .data$ratio_p_display
        )
      ),
      "</small>"
    )
  )

separator_context <- tibble::tibble(
  Location = "&nbsp;",
  row_order = 1,
  row_type = "separator",
  activity = site_category_levels,
  Value = "&nbsp;"
)

site_context_rows <- site_activity |>
  filter(
    .data$placement == "Near-eye",
    .data$activity %in% .env$site_category_levels
  ) |>
  mutate(
    Location = paste0(
      "<span style='color:", .data$site_color_hex,
      ";font-size:1.15em'>●</span>&nbsp;", .data$site_display_name
    ),
    row_order = .data$site_display_order + 1,
    row_type = "site",
    adjusted_p_display = gsub(
      "<", "&lt;", format_p(.data$site_deviation_p_adjusted), fixed = TRUE
    ),
    adjusted_p_display = if_else(
      .data$site_deviation_p_adjusted < 0.05,
      paste0("**", .data$adjusted_p_display, "**"),
      .data$adjusted_p_display
    ),
    Value = case_when(
      .data$reporting_status != "ESTIMABLE" ~
        "<small>Not estimable</small>",
      TRUE ~ paste0(
        if_else(.data$site_deviation_p_adjusted < 0.05, "**", ""),
        sprintf("%.2f×", .data$site_deviation_ratio),
        if_else(.data$site_deviation_p_adjusted < 0.05, "**", ""),
        "<br><small>",
        sprintf(
          "95%% CI %.2f–%.2f",
          .data$site_deviation_conf_low,
          .data$site_deviation_conf_high
        ),
        "</small><br><small>FDR p ",
        .data$adjusted_p_display,
        "</small>"
      )
    )
  ) |>
  select(Location, row_order, row_type, activity, Value)

site_factor_table_data <- bind_rows(
  overall_context,
  separator_context,
  site_context_rows
) |>
  mutate(
    activity = factor(.data$activity, levels = site_category_levels)
  ) |>
  arrange(.data$row_order) |>
  select(-.data$row_order, -.data$row_type) |>
  pivot_wider(names_from = .data$activity, values_from = .data$Value) |>
  select(.data$Location, all_of(site_category_levels))

site_factor_table_data |>
  gt::gt() |>
  gt::fmt_markdown(columns = everything()) |>
  gt::cols_label(
    Location = "Site",
    `At home` = "At home",
    `Working in the office/from home` =
      gt::md("Office/home<br>working"),
    Outdoors = "Outdoors",
    `On the road with public transport/car` =
      gt::md("Vehicle/public<br>transport"),
    Sleeping = "Sleeping"
  ) |>
  gt::cols_width(Location ~ gt::px(165)) |>
  h04_gt(10.5) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#E8EEF3"),
      gt::cell_text(weight = "700"),
      gt::cell_borders(
        sides = "bottom",
        color = "#8192A3",
        weight = gt::px(2)
      )
    ),
    locations = gt::cells_body(rows = 1)
  ) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#FFFFFF"),
      gt::cell_text(color = "#FFFFFF", size = gt::px(1)),
      gt::cell_borders(
        sides = c("top", "bottom"),
        color = "#FFFFFF",
        weight = gt::px(0)
      )
    ),
    locations = gt::cells_body(rows = 2)
  ) |>
  gt::tab_source_note(
    gt::md(paste(
      "The blue-grey row reports site-average category estimates and ratios",
      "to At home from the activity-by-site interaction model, with",
      "FDR-adjusted p-values",
      "for the four category comparisons. Site cells are deviation ratios",
      "around the corresponding category mean and report only the",
      "FDR-adjusted deviation p-value. Bold site ratios and p-values pass",
      "the complete near-eye multiplicity adjustment."
    ))
  )
```

```{r}
#| label: fig-h04-site-context
#| fig-cap: "Support-qualified site-specific activity-associated melEDI. Dashed lines are the site-average geometric means for each sensor position and activity; filled points pass the complete sensor-position-specific FDR adjustment, open points are other supported estimates, and crosses denote unsupported contrasts."
#| fig-alt: "Faceted point-and-interval plot of expected melEDI by country-coded study site, sensor position, and five named activity categories. A dashed vertical line in every facet gives the corresponding site-average geometric mean from the activity-by-site interaction model. Site colours are fixed across panels. Filled points identify FDR-adjusted site deviations, open points show other supported estimates, and a cross marks the support-non-estimable near-eye San José (CR) Outdoors cell. Other/unspecified activity is absent."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_site_activity_estimates.png"
))
```

[Site-estimate data](../../artifacts/09_tables/H04/H04_site_activity_estimands.csv)
and [figure source data](../../artifacts/11_source_data/H04/H04_site_activity_figure.csv)
contain every supported and non-estimable cell.

### Descriptive fixed-effect R²

The primary analysis has no random effects, so mixed-model marginal and
conditional R² are not defined. The following point summaries instead use the
activity-by-site interaction model and preserve the $1/k$ hour weights. The
hierarchy-respecting Shapley allocation distributes shared fitted-model
information across activity, site, and their interaction rather than counting
overlapping contributions more than once. It is a point description without
an uncertainty interval.

```{r}
#| label: tbl-h04-fixed-r2
#| tbl-cap: "Descriptive R² and hierarchy-respecting allocation for the activity-by-site interaction model."

heterogeneity_r_squared |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Definition = case_when(
      .data$loss_basis == "model_deviance" ~
        "Working quasi-Tweedie deviance",
      .data$weighting == "participant_balanced" ~
        "Participant-balanced squared error",
      TRUE ~ "Participant-hour squared error"
    ),
    Definition = factor(
      .data$Definition,
      levels = c(
        "Working quasi-Tweedie deviance",
        "Participant-hour squared error",
        "Participant-balanced squared error"
      )
    ),
    `Overall R²` = sprintf("%.3f", .data$overall_r_squared),
    `Activity allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$category_shapley_r_squared,
      .data$category_shapley_share_percent
    ),
    `Site allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$site_shapley_r_squared,
      .data$site_shapley_share_percent
    ),
    `Interaction allocation` = sprintf(
      "%.3f (%.1f%%)",
      .data$interaction_r_squared,
      .data$interaction_share_percent
    ),
    `Interaction partial R²` = sprintf(
      "%.3f",
      .data$interaction_partial_r_squared_conditional_on_additive
    )
  ) |>
  arrange(.data$Placement, .data$Definition) |>
  select(
    Placement,
    Definition,
    `Overall R²`,
    `Activity allocation`,
    `Site allocation`,
    `Interaction allocation`,
    `Interaction partial R²`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "Allocations are R² units with shares of overall R² in parentheses.",
      "Every multi-select hour is split by 1/k. These are point descriptions",
      "without resampling intervals or additional effect tests."
    ))
  ) |>
  h04_gt(11)
```

Under the near-eye working quasi-deviance definition, overall R² was
`r sprintf("%.3f", heterogeneity_r_squared$overall_r_squared[heterogeneity_r_squared$placement == "Near-eye" & heterogeneity_r_squared$loss_basis == "model_deviance"])`;
activity, site, and interaction allocations were
`r sprintf("%.1f%%", heterogeneity_r_squared$category_shapley_share_percent[heterogeneity_r_squared$placement == "Near-eye" & heterogeneity_r_squared$loss_basis == "model_deviance"])`,
`r sprintf("%.1f%%", heterogeneity_r_squared$site_shapley_share_percent[heterogeneity_r_squared$placement == "Near-eye" & heterogeneity_r_squared$loss_basis == "model_deviance"])`,
and
`r sprintf("%.1f%%", heterogeneity_r_squared$interaction_share_percent[heterogeneity_r_squared$placement == "Near-eye" & heterogeneity_r_squared$loss_basis == "model_deviance"])`
of that fitted improvement. They are scale-dependent descriptions of the
fitted mean structure, not causal shares of exposure variance.

### Exploratory participant random-intercept decomposition

The accepted population-mean models use participant-clustered covariance and
do not contain random effects. A separate descriptive mixed model was fitted
to quantify how much model-based variation is associated with the fixed
activity-by-site structure and with stable participant intercepts. The random
intercept lets participants have different overall exposure levels while
retaining the fixed activity-by-site structure:

```r
geo_medi_1h ~ site * activity_named + (1 | participant)
```

The model used a Tweedie distribution with a log link, maximum likelihood, and
the same fixed working power of 1.539919. It retained the supported five named
categories, so Other-only hours were not included. Every retained
participant-hour still contributed exactly one unit through its $1/k$
membership weights. Five hierarchy-valid models allocated marginal R² to
activity, study site, and their interaction. The participant random intercept
then supplied the difference between conditional and marginal R².

```{r}
#| label: tbl-h04-participant-random-intercept
#| tbl-cap: "Exploratory fractionally weighted participant random-intercept R² decomposition."

mixed_r2_table <- participant_random_intercept |>
  select(
    .data$placement,
    .data$marginal_r_squared,
    .data$conditional_r_squared,
    .data$participant_r_squared_increment,
    .data$residual_variance_share,
    .data$adjusted_participant_icc,
    .data$participant_factor_per_sd
  ) |>
  pivot_longer(
    cols = -placement,
    names_to = "quantity_id",
    values_to = "estimate"
  ) |>
  mutate(
    component = case_when(
      .data$quantity_id %in% c(
        "marginal_r_squared",
        "conditional_r_squared",
        "participant_r_squared_increment",
        "residual_variance_share"
      ) ~ "Model-based R²",
      TRUE ~ "Participant-level variation"
    ),
    quantity = recode(
      .data$quantity_id,
      marginal_r_squared = "Marginal R²",
      conditional_r_squared = "Conditional R²",
      participant_r_squared_increment = "Participant-intercept increment",
      residual_variance_share = "Distribution-specific remainder",
      adjusted_participant_icc = "Adjusted participant ICC",
      participant_factor_per_sd = "One-SD participant factor"
    ),
    interpretation = recode(
      .data$quantity_id,
      marginal_r_squared = "Activity, site, and their interaction",
      conditional_r_squared = "Fixed effects plus participant intercept",
      participant_r_squared_increment = "Conditional minus marginal R²",
      residual_variance_share = "Remainder under the lognormal approximation",
      adjusted_participant_icc =
        "Participant share after excluding fixed-effect variance",
      participant_factor_per_sd =
        "Conditional expected-mean factor one participant SD above average"
    ),
    order = match(
      .data$quantity_id,
      c(
        "marginal_r_squared",
        "conditional_r_squared",
        "participant_r_squared_increment",
        "residual_variance_share",
        "adjusted_participant_icc",
        "participant_factor_per_sd"
      )
    )
  ) |>
  bind_rows(
    participant_random_intercept_shapley |>
      transmute(
        placement = .data$placement,
        quantity_id = .data$component_id,
        estimate = .data$marginal_r_squared_component,
        component = "Marginal R² allocation",
        quantity = .data$component,
        interpretation = sprintf(
          "%.1f%% of full marginal R²",
          .data$share_of_full_marginal_r_squared_percent
        ),
        order = match(
          .data$component_id,
          c("activity", "site", "site_by_activity")
        ) + 6L
      )
  ) |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Component = factor(
      .data$component,
      levels = c(
        "Model-based R²",
        "Marginal R² allocation",
        "Participant-level variation"
      )
    ),
    Quantity = .data$quantity,
    Estimate = if_else(
      .data$quantity_id == "participant_factor_per_sd",
      paste0(sprintf("%.2f", .data$estimate), "×"),
      sprintf("%.3f", .data$estimate)
    ),
    Interpretation = .data$interpretation
  ) |>
  arrange(.data$Placement, .data$order) |>
  select(Placement, Component, Quantity, Estimate, Interpretation)

mixed_r2_table |>
  gt::gt(groupname_col = "Placement") |>
  gt::cols_align(align = "right", columns = Estimate) |>
  gt::tab_source_note(
    gt::md(paste(
      "Point estimates use Nakagawa's lognormal distribution-specific",
      "variance convention. The fixed linear predictor uses a",
      "fractional-frequency sample variance with exact 1/k weights, so",
      "each participant-hour sums to one. No bootstrap intervals were",
      "calculated. The allocation averages the two valid entry orders for",
      "activity and site, and adds their interaction only after both main",
      "effects."
    ))
  ) |>
  h04_gt(10.2)
```

Marginal R² was
`r sprintf("%.3f", participant_random_intercept$marginal_r_squared[participant_random_intercept$placement == "Near-eye"])`
near eye and
`r sprintf("%.3f", participant_random_intercept$marginal_r_squared[participant_random_intercept$placement == "Chest"])`
at chest. Conditional R² was
`r sprintf("%.3f", participant_random_intercept$conditional_r_squared[participant_random_intercept$placement == "Near-eye"])`
and
`r sprintf("%.3f", participant_random_intercept$conditional_r_squared[participant_random_intercept$placement == "Chest"])`,
respectively. The participant intercept therefore added about 0.092 at each
sensor position. Activity accounted for 80.7% of near-eye and 86.1% of chest
marginal R², compared with 12.5% and 5.9% for study site and 6.9% and 8.0% for
the interaction. A one-SD participant intercept corresponded to a 2.02-fold
near-eye and 2.05-fold chest conditional expected mean while holding activity
and site fixed. This common multiplier is not a participant-specific activity
slope.

All ten nested fits converged without warnings, had positive-definite Hessians,
and were non-singular. Aggregated participant-hour Pearson residuals retained
lag-one correlations of 0.317 near eye and 0.301 at chest. Observed exact-zero
fractions were 0.294 and 0.301, compared with mixed-model Tweedie-implied
fractions of 0.410 and 0.429. The decomposition is consequently an exploratory
point description, not evidence of calibrated individual-hour prediction or
a causal partition. The conventional unweighted expanded-row marginal R² was
0.768 and 0.770, only slightly higher than the required weighted values, but
it gives multi-select hours excess influence and is not the reported
estimand. These mixed-model R² values are not interchangeable with the
quasi-deviance and squared-error summaries above.

[Random-intercept assessment estimates](../../artifacts/09_tables/H04/H04_participant_random_intercept_summary.csv),
[model diagnostics](../../artifacts/08_diagnostics/H04/H04_participant_random_intercept_diagnostics.csv),
[hierarchy-respecting allocation](../../artifacts/09_tables/H04/H04_participant_random_intercept_marginal_r2_shapley.csv),
and [nested-model checks](../../artifacts/08_diagnostics/H04/H04_participant_random_intercept_shapley_models.csv)
provide the complete stored point estimates and checks.

[Technical mixed-model preparation and variance definitions](../../audit/hypotheses/H04/H04_analysis_preparation.qmd#sec-h04-prep-participant-random-intercept)
document the exact weighting and decomposition.

## Model checks

```{r}
#| label: tbl-h04-primary-diagnostics
#| tbl-cap: "Numerical, influence, zero-mass, variance-pattern, and serial-dependence model checks."

main_diagnostics |>
  left_join(overall_assessment, by = "placement") |>
  left_join(
    primary_zero_mass |>
      select(
        .data$placement,
        reader_observed_zero_fraction = .data$observed_zero_fraction,
        reader_working_zero_fraction = .data$working_zero_fraction
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = as.character(.data$placement),
    `Converged; rank` = paste0(
      .data$converged, "; ", .data$design_rank, "/", .data$design_columns
    ),
    `Exact-zero hours` = paste0(
      formatC(
        as.integer(.data$exact_zero_unique_hours),
        format = "d",
        big.mark = ","
      ),
      " (", sprintf("%.1f%%", 100 * .data$exact_zero_fraction), ")"
    ),
    `Observed / working zero fraction` = sprintf(
      "%.3f / %.3f",
      .data$reader_observed_zero_fraction,
      .data$reader_working_zero_fraction
    ),
    `|Residual|-fitted Spearman` =
      .data$hour_absolute_residual_fitted_spearman,
    `Lag-1 residual correlation` = .data$hour_residual_lag1_correlation,
    `Maximum cluster score share` = .data$maximum_cluster_score_share,
    `Maximum cluster leverage share` = .data$maximum_cluster_leverage_share,
    Assessment = .data$assessment
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(
      `|Residual|-fitted Spearman`, `Lag-1 residual correlation`,
      `Maximum cluster score share`, `Maximum cluster leverage share`
    ),
    decimals = 3
  ) |>
  gt::tab_source_note(
    gt::md(paste(
      "Residual correlations use one aggregated residual per unique",
      "participant-hour; fractional-weight checks passed exactly."
    ))
  ) |>
  h04_gt(11)
```

Both fits converged at full design rank, all registered robust restrictions
were estimable, and no participant crossed the predeclared influence limits.
The mean models are nevertheless **acceptable with limitation**. Exact zeros
comprise 28.9% of near-eye and 29.4% of chest hours, whereas the corresponding
working Tweedie zero fractions are 90.3% and 93.5%. The quasi mean model
retains observed zeros but is not a separate model of zero probability.
Residual spread changes with fitted magnitude, and lag-one residual
correlations remain 0.445 and 0.498. Participant clustering is the basis for
inference across the full longitudinal record; it does not make those model
discrepancies disappear.

```{r}
#| label: fig-h04-primary-diagnostics
#| fig-cap: "Primary-model residual, temporal-dependence, and zero-mass checks."
#| fig-alt: "Three-panel model-check figure for the primary additive models. Individual and binned Pearson residuals show fitted-value-dependent spread at both sensor positions; residual correlations are positive at the first several hourly lags; and working Tweedie zero fractions substantially exceed observed zero fractions across fitted-mean deciles. Concurrent category memberships are aggregated back to one weighted original participant-hour."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_primary_diagnostics.png"
))
```

[Residual points](../../artifacts/11_source_data/H04/H04_reader_primary_residual_points.csv),
[residual bins](../../artifacts/11_source_data/H04/H04_reader_primary_residual_bins.csv),
[autocorrelations](../../artifacts/08_diagnostics/H04/H04_reader_primary_residual_acf.csv),
and [zero calibration](../../artifacts/11_source_data/H04/H04_reader_primary_zero_calibration.csv)
provide the exact plotted values.

## Sensitivity analyses

The original sensitivity analyses each change a stated part of the accepted
analysis: restrict to exactly-one-category hours; retain co-selected Other as
a fractionally weighted category; exclude Other-only hours; replace $1/k$
weights with unit-weight long rows; use gap-timing-unaware data; use the same
participants and participant-hours at both sensor positions; change the
quasi-Tweedie working power to 1.30 or 1.80; or use the complementary chest
sensor position. An additional Mundlak-style sensitivity separates the
hourly category contrast from differences in participants' usual reported
activity composition.

The **gap-timing-unaware dataset** still passed the general per-hour and
per-day coverage requirements. The term means that the timing of remaining
missing observations is not used for an additional metric-specific
adjustment; it does not mean that gaps or coverage were ignored.

```{r}
#| label: tbl-h04-sensitivities
#| tbl-cap: "Exact fitted samples, five-category omnibus tests, and named-contrast stability across recorded sensitivities."

samples |>
  filter(.data$scenario_id != "primary_dataset") |>
  left_join(
    sensitivity_summary,
    by = c("scenario_id", "placement")
  ) |>
  left_join(
    sensitivity_f1,
    by = c("scenario_id", "placement")
  ) |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    scenario_order = match(.data$scenario_id, names(.env$scenario_labels)),
    Scenario = unname(.env$scenario_labels[.data$scenario_id]),
    sample_footprint = sprintf(
      "%s / %s / %s / %s / %s",
      formatC(as.integer(.data$participants), format = "d", big.mark = ","),
      formatC(
        as.integer(.data$participant_days),
        format = "d",
        big.mark = ","
      ),
      formatC(
        as.integer(.data$unique_participant_hours),
        format = "d",
        big.mark = ","
      ),
      formatC(as.integer(.data$long_rows), format = "d", big.mark = ","),
      formatC(
        as.integer(round(.data$effective_weighted_hours)),
        format = "d",
        big.mark = ","
      )
    ),
    stable_display = paste0(.data$stable, "/4"),
    f_display = sprintf(
      "%.2f (%d df)",
      .data$f_statistic,
      .data$denominator_df
    ),
    p_display = format_p_cell(.data$p_raw, significant = FALSE)
  ) |>
  arrange(.data$scenario_order, .data$placement) |>
  transmute(
    Scenario,
    Placement = as.character(.data$placement),
    `Participants / days / hours / rows / weighted hours` =
      .data$sample_footprint,
    `Stable named contrasts` = .data$stable_display,
    `Maximum ratio change, %` =
      .data$maximum_absolute_ratio_change_percent,
    `F statistic (denominator df)` = .data$f_display,
    `Raw p` = .data$p_display
  ) |>
  gt::gt(groupname_col = "Scenario") |>
  gt::fmt_number(columns = `Maximum ratio change, %`, decimals = 1) |>
  gt::fmt_markdown(columns = `Raw p`) |>
  gt::tab_source_note(
    gt::md(paste(
      "Sensitivity p-values are unadjusted model-check results and are not",
      "bolded. Stable means the direction and prespecified classification",
      "were retained. Unit-weight long rows change the estimand by allowing",
      "multi-select hours to contribute more than one unit. The within/between",
      "participant analysis is an additional exploratory sensitivity."
    ))
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(10),
    `Participants / days / hours / rows / weighted hours` ~ gt::pct(31),
    everything() ~ gt::pct(14.7)
  ) |>
  h04_gt(10.5)
```

All four named contrasts were stable in every sensitivity at both placements.
The deletion audit also removed each site and each of the five highest-score
participants in turn; no primary decision changed. These checks support the
direction and broad magnitude ordering, but they do not remove the model
limitations or upgrade the observational claim.

### Within- and between-participant activity patterns

For this additional sensitivity, each participant's usual category profile is
the fraction of their retained participant-hours assigned to each category
after the same $1/k$ sharing of multi-select hours. The model retains the
current reported category and adds the five non-reference participant-level
category proportions. The current-category ratios therefore compare reported
activity categories after adjustment for participants' usual activity
composition. The between-participant ratios compare a 10 percentage-point
higher usual share of the named category with the corresponding lower share
of At home, while holding current category, study site, and the other category
shares fixed. Other remains in the model but is display-only and is not part of
either four-comparison FDR family.

```{r}
#| label: tbl-h04-mundlak
#| tbl-cap: "Population-average and Mundlak within-participant activity ratios to At home, plus between-participant composition ratios per 10 percentage points; all intervals are 95% CIs."

mundlak_display |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Activity = factor(
      unname(.env$reader_activity_labels[.data$activity_code]),
      levels = unname(.env$reader_activity_labels[.env$named_codes])
    ),
    `Population-average ratio (95% CI)` = format_ratio_ci(
      .data$primary_ratio,
      .data$primary_conf_low,
      .data$primary_conf_high
    ),
    `Within-participant ratio (95% CI)` = format_ratio_ci(
      .data$within_ratio,
      .data$within_conf_low,
      .data$within_conf_high
    ),
    `Within FDR p` = format_p_cell(
      .data$within_p_adjusted,
      .data$within_p_adjusted < 0.05
    ),
    `Between ratio per 10 percentage points (95% CI)` = format_ratio_ci(
      .data$between_ratio,
      .data$between_conf_low,
      .data$between_conf_high
    ),
    `Between FDR p` = format_p_cell(
      .data$between_p_adjusted,
      .data$between_p_adjusted < 0.05
    )
  ) |>
  arrange(.data$Placement, .data$Activity) |>
  select(
    Placement,
    Activity,
    `Population-average ratio (95% CI)`,
    `Within-participant ratio (95% CI)`,
    `Within FDR p`,
    `Ratio change, %` = .data$ratio_relative_change_percent,
    `Between ratio per 10 percentage points (95% CI)`,
    `Between FDR p`
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_spanner(
    label = "Hourly category contrast",
    columns = c(
      `Population-average ratio (95% CI)`,
      `Within-participant ratio (95% CI)`,
      `Within FDR p`,
      `Ratio change, %`
    )
  ) |>
  gt::tab_spanner(
    label = "Usual category share",
    columns = c(
      `Between ratio per 10 percentage points (95% CI)`,
      `Between FDR p`
    )
  ) |>
  gt::fmt_number(columns = `Ratio change, %`, decimals = 1) |>
  gt::fmt_markdown(columns = c(`Within FDR p`, `Between FDR p`)) |>
  gt::tab_source_note(gt::md(paste(
    "FDR is applied separately to the four named within-participant and",
    "between-participant comparisons at each sensor position."
  ))) |>
  gt::cols_width(
    Activity ~ gt::pct(14),
    `Population-average ratio (95% CI)` ~ gt::pct(19),
    `Within-participant ratio (95% CI)` ~ gt::pct(19),
    `Within FDR p` ~ gt::pct(9),
    `Ratio change, %` ~ gt::pct(10),
    `Between ratio per 10 percentage points (95% CI)` ~ gt::pct(20),
    `Between FDR p` ~ gt::pct(9)
  ) |>
  h04_gt(9.2)
```

The within-participant category omnibus remained clear near eye,
$F(4,125)=67.17$, raw $p<0.001$, and at chest, $F(4,149)=77.86$, raw
$p<0.001$. Every primary ratio lay inside its Mundlak sensitivity interval.
The largest change was the near-eye Outdoors ratio, which decreased by 16.8%
from 11.58 to 9.64, with a 95% CI of 6.79 to 13.68. This indicates that the
strong category ordering is not explained solely by stable differences in
participants' usual activity profiles.

The four named between-participant composition terms were jointly associated
with near-eye melEDI, $F(4,125)=3.81$, raw $p=0.006$, but not clearly with chest
melEDI, $F(4,149)=2.34$, raw $p=0.058$. Near eye, a 10 percentage-point higher
usual Outdoors share was associated with a ratio of 1.42 (95% CI 1.09 to 1.86;
FDR $p=0.039$). The corresponding chest estimate was 1.30 (95% CI 1.06 to
1.59; FDR $p=0.054$). No other named between-participant comparison passed its
four-member FDR rule. These are observational composition associations, not
effects of changing activity, and they do not control time-varying differences
such as time of day, season, weather, or purpose of travel.

### Same-participant, same-hour sensor-position comparison

This comparison uses the same 110 participants and the same 14,308
participant-hours at both sensor positions. The exact frame contains 625
participant-days, 15,001 long rows, and 14,308 effective weighted hours at
each position. The positions are still fitted separately, so this is neither
an equivalence test nor a direct sensor-effect test.

```{r}
#| label: fig-h04-paired-placement
#| fig-cap: "Common-sample near-eye-versus-chest concordance of category ratios to At home; the identity line does not define equivalence."
#| fig-alt: "Log-log concordance plot of four common-sample activity ratios. The horizontal coordinate is the near-eye ratio to At home and the vertical coordinate is the chest ratio, with horizontal and vertical 95 percent confidence intervals. Sleeping lies near 0.1 at both placements; working lies near 2.5; travel and Outdoors are higher at chest than near eye. A dashed identity line is a visual concordance reference, not an equivalence test."
#| out-width: "72%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_paired_placement_comparison.png"
))
```

[Common-sample estimates](../../artifacts/09_tables/H04/H04_paired_placement_estimands.csv)
and [figure source data](../../artifacts/11_source_data/H04/H04_paired_placement_figure.csv)
retain the exact placement-specific values. The display uses identical
participant-hour keys but supplies no unplanned position-difference *p*-value.

## Exploratory time-of-day context

::: {.callout-note title="Exploratory analysis"}
This nonlinear generalized additive model (GAM) analysis is separate from the
preregistered category inference. It allows the association with clock time
to bend across the day using one cyclic global time-of-day smooth plus
thin-plate sum-to-zero activity and site deviations, participant curves,
participant-day shifts, and a boundary-aware AR(1) correction. Activity
curves are also shown as ratios to the global daily smooth, not to At home,
and no curve-wide test is made.
:::

The exploratory model retains the same activity-long rows and $1/k$ weights.
It contains a cyclic global time smooth, thin-plate sum-to-zero activity and
site smooths, participant-specific time smooths, and participant-day random
intercepts. Only the global smooth is cyclic. Activity and site deviations
may therefore separate at midnight; the open endpoints make this visible and
no line wraps from 24:00 to 00:00.

Activity-specific autoregressive runs restart at participant-day, activity,
repeated-hour, non-consecutive-hour, and gap boundaries. Consequently,
concurrent long rows from the same hour never become lag neighbours.

```{r}
#| label: tbl-h04-temporal-fit
#| tbl-cap: "Exploratory temporal-model samples, fit, and added-context summaries."

temporal_activity |>
  left_join(
    temporal_retention,
    by = c("placement" = "placement")
  ) |>
  left_join(
    temporal_r_squared |>
      select(
        .data$placement,
        balanced_r_squared = .data$r_squared
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  transmute(
    Placement = as.character(.data$placement),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Unique hours` = .data$unique_participant_hours,
    `Long rows` = .data$observations_long_rows,
    `Weighted hours` = .data$effective_weighted_hours,
    `AR rho` = .data$rho,
    `Adjusted R²` = .data$adjusted_r_squared,
    `Deviance explained` = .data$deviance_explained,
    `Site-average participant-balanced R²` = .data$balanced_r_squared
  ) |>
  gt::gt() |>
  gt::fmt_integer(
    columns = c(
      Participants, `Participant-days`, `Unique hours`, `Long rows`
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_number(columns = `Weighted hours`, decimals = 0, use_seps = TRUE) |>
  gt::fmt_number(
    columns = c(
      `AR rho`, `Adjusted R²`, `Deviance explained`,
      `Site-average participant-balanced R²`
    ),
    decimals = 3
  ) |>
  gt::tab_source_note(
    gt::md(paste(
      "The participant-balanced R² is calculated on raw melEDI with sites",
      "weighted equally, participants equally within site, hours equally",
      "within participant, and concurrent memberships split by 1/k. All R²",
      "values are descriptive in-sample point estimates without resampling",
      "intervals."
    ))
  ) |>
  h04_gt(10.5)
```

The activity smooth improved AIC by 3,215 near eye and 4,984 at chest, while
deviance explained increased from 81.8% to 84.8% and from 79.5% to 83.2%.
This supports retaining the profiles as descriptive context, not as a second
primary activity test.

The activity curves use a qualitative palette distinct from the light-source
colours used elsewhere in this report series, so colour does not imply a
mapping between different category systems.

```{r}
#| label: fig-h04-temporal-near-eye
#| fig-cap: "Exploratory near-eye time-of-day means, ratios to the global smooth, and effective weighted support."
#| fig-alt: "Taller three-row, six-category near-eye figure using an activity palette distinct from the light-source colours used elsewhere in this report series. Six single-line facet strips are ordered At home, Office/home working, Outdoors, Vehicle/public transport, Sleeping, and Other/unspecified. The first row shows site-average activity-associated melEDI curves with pointwise confidence ribbons and a dashed global cyclic curve; the second row shows each activity curve divided by that global curve on a common 0.01-to-50 ratio scale with 1-2-5 logarithmic breaks; and the third row shows effective weighted hours by clock hour. Sleeping is lowest, Outdoors highest during daytime, open circles mark sparse support, and grey gaps mark no observations. Other is display-only."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_temporal_near_eye.png"
))
```

```{r}
#| label: fig-h04-temporal-chest
#| fig-cap: "Exploratory chest time-of-day means, ratios to the global smooth, and effective weighted support."
#| fig-alt: "Taller three-row, six-category chest figure using an activity palette distinct from the light-source colours used elsewhere in this report series. Six single-line facet strips are ordered At home, Office/home working, Outdoors, Vehicle/public transport, Sleeping, and Other/unspecified. It shows site-average activity-associated melEDI curves and the dashed global cyclic curve, activity-to-global ratios on a common 0.01-to-50 scale with 1-2-5 logarithmic breaks, and effective weighted hours by clock time. Outdoors and travel have the highest daytime profiles; Sleeping is lowest overnight. Filled and open circles distinguish adequate from sparse support, grey gaps have no observations, and Other is display-only."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_temporal_chest.png"
))
```

[Near-eye curves](../../artifacts/11_source_data/H04/H04_reader_temporal_near_eye_curves.csv),
[near-eye activity/global ratios](../../artifacts/11_source_data/H04/H04_reader_temporal_near_eye_ratios.csv),
[near-eye global curve](../../artifacts/11_source_data/H04/H04_reader_temporal_near_eye_global.csv),
and [near-eye support](../../artifacts/11_source_data/H04/H04_reader_temporal_near_eye_support.csv)
contain the primary plotted values. The corresponding
[chest curves](../../artifacts/11_source_data/H04/H04_reader_temporal_chest_curves.csv),
[ratios](../../artifacts/11_source_data/H04/H04_reader_temporal_chest_ratios.csv),
[global curve](../../artifacts/11_source_data/H04/H04_reader_temporal_chest_global.csv),
and [support](../../artifacts/11_source_data/H04/H04_reader_temporal_chest_support.csv)
provide the complementary placement values.

### Temporal linear-predictor variance allocation

This hierarchy-respecting Shapley allocation distributes shared fitted-model
information across the temporal components rather than counting overlapping
contributions more than once. The percentages are point estimates and have no
uncertainty intervals.

```{r}
#| label: tbl-h04-temporal-allocation
#| tbl-cap: "Point allocation of fitted temporal linear-predictor variance."

temporal_allocation |>
  mutate(
    Placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    Component = recode(
      .data$group,
      global_time = "Global time of day",
      activity_deviations = "Activity deviations",
      site_deviations = "Site deviations",
      participant_curves = "Participant curves",
      participant_day_shifts = "Participant-day shifts"
    ),
    `Allocated share` = sprintf("%.1f%%", 100 * .data$shapley_share),
    `Unique partial share` = sprintf(
      "%.1f%%",
      100 * .data$partial_unique_share
    )
  ) |>
  select(Placement, Component, `Allocated share`, `Unique partial share`) |>
  gt::gt(groupname_col = "Placement") |>
  gt::tab_source_note(
    gt::md(paste(
      "Exact point allocation on the natural-log conditional-mean predictor",
      "under site-average, participant-balanced, and 1/k within-hour weights.",
      "Components share covariance; no simulation interval was computed."
    ))
  ) |>
  h04_gt()
```

Global time of day received 52.5% of the near-eye and 50.2% of the chest
point allocation. Activity deviations received 30.7% and 32.6%, site
deviations 5.6% and 6.5%, participant curves 8.4% and 7.1%, and
participant-day shifts 2.8% and 3.5%. These are allocations of fitted
linear-predictor variation, not independent or causal percentages of raw
melEDI variance.

### Temporal model checks

```{r}
#| label: tbl-h04-temporal-diagnostics
#| tbl-cap: "Convergence, basis capacity, residual dependence, support, and midnight checks for the temporal models."

temporal_retention |>
  left_join(
    temporal_zero_mass,
    by = "placement",
    relationship = "one-to-one"
  ) |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest"))
  ) |>
  arrange(.data$placement) |>
  transmute(
    Placement = as.character(.data$placement),
    `Minimum k-index` = .data$minimum_finite_k_index,
    `Residual lag 1` = .data$residual_lag1,
    `Observed / working zero fraction` = sprintf(
      "%.3f / %.3f",
      .data$observed_zero_fraction,
      .data$working_zero_fraction
    ),
    `Sparse clock/category cells` =
      .data$locally_sparse_clock_activity_cells,
    `Maximum activity midnight ratio` =
      .data$maximum_activity_endpoint_ratio,
    `Maximum site midnight ratio` = .data$maximum_site_endpoint_ratio,
    Assessment = .data$assessment
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(`Minimum k-index`, `Residual lag 1`),
    decimals = 3
  ) |>
  gt::fmt_number(
    columns = c(
      `Maximum activity midnight ratio`, `Maximum site midnight ratio`
    ),
    decimals = 2
  ) |>
  gt::fmt_integer(columns = `Sparse clock/category cells`) |>
  h04_gt(11)
```

Both temporal fits converged at full coefficient rank with positive-definite
smoothing-parameter Hessians and no final warnings. Their minimum finite
k-indices were 0.872 and 0.863. Residual lag-one correlations remained 0.176
and 0.167, and 41 near-eye plus 40 chest clock/category cells were locally
sparse. The non-cyclic deviations can show large endpoint ratios.
The working zero fractions of 0.452 and 0.502 exceed the observed fractions
of 0.289 and 0.294. The profiles are therefore assessed **acceptable with
limitation**.

For each time-by-activity profile, sites are averaged equally on the fitted
log-mean scale after participant and participant-day contributions are set to
zero. Uncertainty uses the fitted coefficient covariance from
`stats::vcov(fit, unconditional = TRUE)`, with the fitted `Vp` matrix only as
an error fallback, followed by one back-transformation. Ribbons are
model-based pointwise 95% CIs: each interval applies to one displayed clock
time rather than to the whole curve simultaneously. No time-specific test,
significant period, or curve-wide inference is reported. The positive
lag-one residual correlations mean that neighbouring hourly residuals remain
related despite the working AR(1) correction. No resampling enters the
displayed estimates or intervals.

```{r}
#| label: fig-h04-temporal-diagnostics
#| fig-cap: "Exploratory temporal-model residual, autocorrelation, and zero-mass checks."
#| fig-alt: "Three-panel model-check figure for the exploratory temporal models. Weighted one-hour residuals show changing spread with fitted melEDI, activity-run residual autocorrelation remains positive at lag one for both sensor positions, and the working Tweedie zero fraction exceeds the observed fraction across fitted-value deciles. The residual and zero panels combine concurrent activity memberships with their 1 over k weights."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_temporal_diagnostics.png"
))
```

[Residual points](../../artifacts/11_source_data/H04/H04_reader_temporal_residual_points.csv),
[residual bins](../../artifacts/11_source_data/H04/H04_reader_temporal_residual_bins.csv),
[activity-run autocorrelations](../../artifacts/08_diagnostics/H04/H04_reader_temporal_residual_acf.csv),
and [zero calibration](../../artifacts/11_source_data/H04/H04_reader_temporal_zero_calibration.csv)
provide the exact model-check source values.

## Interpretation

Reported activity is strongly associated with one-hour personal light
exposure. Sleeping corresponds to markedly lower melEDI than At home, whereas
vehicle or public-transport travel, indoor or home working, and especially
Outdoors correspond to higher values. The ordering is similar at chest, and
it persists in the exact placement-matched sample and across the predefined
sensitivity set. It also persists after adjustment for participants' usual
reported activity composition. Activity-associated means vary materially by
site.

The result remains observational. Fractional membership is a transparent way
to preserve one total contribution per participant-hour; it does not make
simultaneously reported labels mutually exclusive or remove ambiguity in
self-report timing. Other/unspecified activity is heterogeneous and supports
no substantive interpretation. The site-average estimates describe the
observed site set and do not automatically generalize to other populations.
Participant-clustered inference accommodates within-hour membership and the
longitudinal record under the stated mean model, but it does not cure residual
serial dependence, fitted-value-dependent spread, sparse site cells, or the
model's limited representation of exact zeros. The time-of-day profiles add
descriptive context only, and the separate sensor-position fits do not
establish equivalence.

## Sources and technical details

The exact sample-flow and model-ready dataset definitions are documented in
[Preparation 02](../preparation/02_coverage_sample_flow.qmd) and
[Preparation 06](../preparation/06_model_ready_datasets.qmd). The companion
[How the H04 analysis was prepared](../../audit/hypotheses/H04/H04_analysis_preparation.qmd)
traces the verified inputs, multi-select transformation, model frames,
estimands, model checks, sensitivities, temporal construction, and recorded
output chain behind this report.

### Exact model formulas

The following technical formulas are the exact Wilkinson formulas recorded
for the fitted analyses. They are shown for provenance; the population-mean
and activity-by-site roles are explained above.

```{r}
#| label: tbl-h04-formulas
#| tbl-cap: "Exact Wilkinson formulas recorded for the population-mean, activity-by-site interaction, and exploratory temporal models."

formula_labels <- c(
  primary_full = "Primary population mean",
  primary_five_named_null = "Primary five-category null",
  secondary_six_category_null = "Secondary six-category null",
  heterogeneity_five_named_full = "Named-category activity-by-site interaction",
  temporal_activity_long = "Exploratory time of day"
)

formulas |>
  filter(.data$formula_id %in% names(.env$formula_labels)) |>
  mutate(
    Model = unname(.env$formula_labels[.data$formula_id]),
    order = match(.data$formula_id, names(.env$formula_labels))
  ) |>
  arrange(.data$order) |>
  transmute(
    Model,
    `Exact Wilkinson formula supplied to R` = .data$formula
  ) |>
  gt::gt() |>
  gt::cols_width(
    Model ~ gt::pct(25),
    `Exact Wilkinson formula supplied to R` ~ gt::pct(75)
  ) |>
  h04_gt(11.5)
```
