---
title: "H05: Habitual light-related behaviour and personal light exposure"
subtitle: "Associations between LEBA questionnaire factors and 17 light-exposure metrics"
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h05-reader

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()),
  winslash = "/",
  mustWork = TRUE
)
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)
source(file.path(root, "scripts", "pipeline", "p_value_display.R"))

read_h05 <- function(...) {
  readr::read_csv(file.path(root, ...), show_col_types = FALSE)
}

near_results <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_near_eye_results.csv"
)
chest_results <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_chest_results.csv"
)
near_samples <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_near_eye_samples.csv"
)
chest_samples <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_chest_samples.csv"
)
master <- read_h05(
  "artifacts", "09_tables", "H05", "H05_model_results_master.csv"
)
diagnostics <- read_h05(
  "artifacts", "08_diagnostics", "H05", "H05_model_diagnostics.csv"
)
formula_registry <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_formula_registry.csv"
)
factor_registry <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_factor_registry.csv"
)
metric_registry <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_metric_registry.csv"
)
family_audit <- read_h05(
  "artifacts", "09_tables", "H05", "H05_family_audit.csv"
)
site_evidence <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_site_structure_evidence.csv"
)
random_site <- read_h05(
  "artifacts", "09_tables", "H05", "H05_random_site_sensitivity.csv"
)
leave_one_site_out <- read_h05(
  "artifacts", "09_tables", "H05", "H05_leave_one_site_out_summary.csv"
)
paired <- read_h05(
  "artifacts", "09_tables", "H05", "H05_paired_placement_comparison.csv"
)
paired_display <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_paired_effect_comparison_data.csv"
)
preparation_sensitivity <- read_h05(
  "artifacts", "09_tables", "H05", "H05_manuscript_prepared_comparison.csv"
)
exact_period <- read_h05(
  "artifacts", "09_tables", "H05",
  "H05_exactly_identified_longest_bout_sensitivity.csv"
)
descriptive_spearman <- read_h05(
  "artifacts", "09_tables", "H05", "H05_descriptive_spearman.csv"
)

near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
paired_near_id <- "main__glasses__paired_common_sample"
paired_chest_id <- "main__chest__paired_common_sample"
preparation_id <- "manuscript_prepared_data__glasses__all_available"

factor_code <- function(factor_id) {
  toupper(sub("leba_", "", factor_id))
}

format_number <- function(value, digits = 3L) {
  formatC(value, digits = digits, format = "f", big.mark = ",")
}

format_p <- function(value) {
  nh_format_p_value(value)
}

format_p_cell <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  if (display$p_bold[[1L]]) {
    paste0("**", display$p_display[[1L]], "**")
  } else {
    display$p_display[[1L]]
  }
}

format_effect <- function(estimate, low, high, type, unit, digits = 3L) {
  if (type == "odds_ratio") {
    return(paste0(
      "OR ", format_number(estimate, digits), " (",
      format_number(low, digits), "–", format_number(high, digits), ")"
    ))
  }
  if (type == "ratio") {
    return(paste0(
      "×", format_number(estimate, digits), " (",
      format_number(low, digits), "–", format_number(high, digits), ")"
    ))
  }
  suffix <- if (unit %in% c("h", "clock time")) " h" else ""
  paste0(
    sprintf(paste0("%+.", digits, "f"), estimate), suffix, " (",
    sprintf(paste0("%+.", digits, "f"), low), "–",
    sprintf(paste0("%+.", digits, "f"), high), suffix, ")"
  )
}

h05_gt <- function(table, font_size = 13) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

metric_group <- function(metric_order) {
  dplyr::case_when(
    metric_order <= 5 ~ "Stability and light level",
    metric_order <= 10 ~ "Duration and continuous periods",
    metric_order <= 15 ~ "Timing",
    TRUE ~ "Exposure history and spectrum"
  )
}

sample_table <- function(data) {
  data |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      `Analysis level` = if_else(
        .data$analysis_unit == "participant",
        "Participant",
        "Participant-day"
      ),
      `Model rows` = .data$observations,
      Participants = .data$participants,
      `Participant-days` = .data$participant_days,
      `Represented days` = .data$represented_days,
      Sites = .data$sites
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$`Analysis level`,
      .data$`Model rows`,
      .data$Participants,
      .data$`Participant-days`,
      .data$`Represented days`,
      .data$Sites
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::fmt_integer(
      columns = c(
        `Model rows`, Participants, `Participant-days`,
        `Represented days`, Sites
      )
    ) |>
    gt::cols_width(
      `Analysis level` ~ gt::pct(17),
      `Model rows` ~ gt::pct(12),
      Participants ~ gt::pct(12),
      `Participant-days` ~ gt::pct(14),
      `Represented days` ~ gt::pct(14),
      Sites ~ gt::pct(8)
    ) |>
    h05_gt(12)
}

result_matrix <- function(data, metric_orders) {
  data |>
    filter(.data$metric_order %in% .env$metric_orders) |>
    left_join(
      metric_registry |>
        select(.data$metric_id, .data$display_unit),
      by = "metric_id",
      relationship = "many-to-one"
    ) |>
    rowwise() |>
    mutate(
      Metric = .data$manuscript_name,
      Factor = factor_code(.data$factor_id),
      Result = if_else(
        .data$reader_inference_status == "unfit_for_inference",
        paste0(
          "**Unfit for inference**<br><small>Estimate and p-values ",
          "suppressed from the reader display</small>"
        ),
        paste0(
          format_effect(
            .data$estimate_practical_per_sd,
            .data$conf_low_practical_per_sd,
            .data$conf_high_practical_per_sd,
            .data$effect_type,
            .data$display_unit
          ),
          "<br><small>Raw p = ",
          format_p_cell(.data$p_raw, FALSE),
          "; BH-adjusted p = ",
          format_p_cell(.data$p_adjusted, .data$p_adjusted <= 0.05),
          "</small>"
        )
      )
    ) |>
    ungroup() |>
    select(.data$metric_order, .data$Metric, .data$Factor, .data$Result) |>
    tidyr::pivot_wider(names_from = .data$Factor, values_from = .data$Result) |>
    arrange(.data$metric_order) |>
    select(-.data$metric_order) |>
    gt::gt(rowname_col = "Metric") |>
    gt::fmt_markdown(columns = c(F2, F3, F4, F5)) |>
    gt::cols_width(
      F2 ~ gt::pct(18), F3 ~ gt::pct(18),
      F4 ~ gt::pct(18), F5 ~ gt::pct(18)
    ) |>
    gt::tab_source_note(
      source_note = paste(
        "Effects are per participant SD of the LEBA score with 95% Wald",
        "confidence intervals. Raw p has no separate bolding rule;",
        "BH-adjusted p is bold only at alpha = 0.050 across all 68 tests.",
        "Sleep-environment cells are unfit for inference and suppressed."
      )
    ) |>
    h05_gt(11)
}

adequacy_count_table <- function(data) {
  data |>
    mutate(
      reader_assessment = if_else(
        .data$reader_inference_status == "unfit_for_inference",
        "unfit_for_inference",
        .data$model_adequacy
      )
    ) |>
    count(.data$reader_assessment, name = "Models") |>
    mutate(
      Assessment = recode(
        .data$reader_assessment,
        acceptable = "Acceptable",
        acceptable_with_specified_limitations =
          "Acceptable with specified limitations",
        unfit_for_inference = "Unfit for inference",
        not_acceptable = "Not acceptable"
      )
    ) |>
    select(.data$Assessment, .data$Models) |>
    gt::gt(rowname_col = "Assessment") |>
    gt::fmt_integer(columns = Models) |>
    h05_gt()
}

limitation_count_table <- function(data) {
  data |>
    mutate(
      Limitation = case_when(
        is.na(.data$specified_limitations) ~ "None specified",
        grepl("WARN_STRONG_TWEEDIE_MISFIT", .data$specified_limitations) ~
          "Unfit sleep-environment model: simulated-residual and response-support failure",
        grepl("WARN_TWEEDIE_DIAGNOSTIC", .data$specified_limitations) ~
          "Ordinary Tweedie simulation-diagnostic warning",
        grepl("WARN_GAUSSIAN_DIAGNOSTIC", .data$specified_limitations) ~
          "Ordinary Gaussian residual warning",
        TRUE ~ .data$specified_limitations
      )
    ) |>
    count(.data$Limitation, name = "Models") |>
    arrange(desc(.data$Models)) |>
    gt::gt(rowname_col = "Limitation") |>
    gt::fmt_integer(columns = Models) |>
    h05_gt()
}

sleep_diagnostic_table <- function(run_id) {
  diagnostics |>
    filter(
      .data$run_id == .env$run_id,
      .data$metric_id == "duration_below_1_sleep_environment"
    ) |>
    transmute(
      Factor = factor_code(.data$factor_id),
      `Uniformity raw p` = nh_format_p_value(.data$dharma_uniformity_p),
      `Zero-mass raw p` = nh_format_p_value(.data$dharma_zero_inflation_p),
      `Outlier raw p` = nh_format_p_value(.data$dharma_outlier_p),
      `Observed zero fraction` = .data$observed_zero_fraction,
      `Simulated zero fraction` = .data$simulated_zero_fraction,
      `Predictions above support` = .data$predicted_above_bound_n
    ) |>
    gt::gt(rowname_col = "Factor") |>
    gt::fmt_number(
      columns = c(`Observed zero fraction`, `Simulated zero fraction`),
      decimals = 4
    ) |>
    gt::fmt_integer(columns = `Predictions above support`) |>
    gt::tab_source_note(
      source_note = paste(
        "These are raw simulation-diagnostic p-values. No standalone alpha",
        "or bolding rule is applied; they are interpreted jointly with the",
        "response-support failures that make these models unfit for inference."
      )
    ) |>
    h05_gt(12)
}

paired_component_table <- function(factor_id) {
  paired_display |>
    filter(.data$factor_id == .env$factor_id) |>
    arrange(.data$metric_order) |>
    rowwise() |>
    transmute(
      Metric = .data$manuscript_name,
      `Near-eye model-scale estimate (95% CI)` = paste0(
        format_number(.data$estimate_model_per_sd__glasses), " (",
        format_number(.data$conf_low_model_per_sd__glasses), "–",
        format_number(.data$conf_high_model_per_sd__glasses), ")"
      ),
      `Chest model-scale estimate (95% CI)` = paste0(
        format_number(.data$estimate_model_per_sd__chest), " (",
        format_number(.data$conf_low_model_per_sd__chest), "–",
        format_number(.data$conf_high_model_per_sd__chest), ")"
      ),
      `Exact matched sample` = if_else(
        .data$analysis_unit__near_eye == "participant",
        paste0(
          .data$participants__near_eye,
          " participants (",
          .data$observations__near_eye,
          " rows); ",
          .data$sites__near_eye,
          " sites"
        ),
        paste0(
          .data$participants__near_eye,
          " participants; ",
          .data$participant_days__near_eye,
          " participant-days; ",
          .data$sites__near_eye,
          " sites"
        )
      )
    ) |>
    ungroup() |>
    gt::gt(rowname_col = "Metric") |>
    gt::cols_width(
      `Near-eye model-scale estimate (95% CI)` ~ gt::pct(24),
      `Chest model-scale estimate (95% CI)` ~ gt::pct(24),
      `Exact matched sample` ~ gt::pct(25)
    ) |>
    gt::tab_source_note(
      source_note = paste(
        "These are the component 95% confidence intervals for the exact",
        "matched estimands plotted above. The identity display describes",
        "concordance and is not an equivalence analysis."
      )
    ) |>
    h05_gt(11)
}

near_top <- near_results |>
  filter(
    .data$factor_id == "leba_f2",
    .data$metric_id %in% c(
      "duration_above_1000",
      "dose_time_sensitive_corrected_medi"
    )
  ) |>
  arrange(.data$metric_order)
near_duration <- near_top |>
  filter(.data$metric_id == "duration_above_1000")
near_dose <- near_top |>
  filter(.data$metric_id == "dose_time_sensitive_corrected_medi")

chest_duration <- chest_results |>
  filter(
    .data$factor_id == "leba_f2",
    .data$metric_id == "duration_above_1000"
  )
chest_dose <- chest_results |>
  filter(
    .data$factor_id == "leba_f2",
    .data$metric_id == "dose_time_sensitive_corrected_medi"
  )

near_adequacy <- near_results |> count(.data$model_adequacy)
chest_adequacy <- chest_results |> count(.data$model_adequacy)

paired_near <- master |>
  filter(.data$run_id == paired_near_id)
paired_chest <- master |>
  filter(.data$run_id == paired_chest_id)
preparation_results <- master |>
  filter(.data$run_id == preparation_id)

leading_sensitivity <- bind_rows(
  near_top |> mutate(Scenario = "Primary near eye"),
  paired_near |>
    filter(
      .data$factor_id == "leba_f2",
      .data$metric_id %in% near_top$metric_id
    ) |>
    mutate(Scenario = "Common-sample near eye"),
  paired_chest |>
    filter(
      .data$factor_id == "leba_f2",
      .data$metric_id %in% near_top$metric_id
    ) |>
    mutate(Scenario = "Common-sample chest"),
  preparation_results |>
    filter(
      .data$factor_id == "leba_f2",
      .data$metric_id %in% near_top$metric_id
    ) |>
    mutate(Scenario = "Gap-timing-unaware dataset")
) |>
  left_join(
    metric_registry |> select(.data$metric_id, .data$display_unit),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  rowwise() |>
  mutate(
    `Effect per LEBA SD (95% CI)` = format_effect(
      .data$estimate_practical_per_sd,
      .data$conf_low_practical_per_sd,
      .data$conf_high_practical_per_sd,
      .data$effect_type,
      .data$display_unit
    ),
    `Fitted sample` = paste0(
      .data$participants, " participants; ",
      if_else(
        is.na(.data$participant_days),
        paste0(.data$observations, " participant rows"),
        paste0(.data$participant_days, " participant-days")
      ),
      "; ", .data$sites, " sites"
    )
  ) |>
  ungroup() |>
  select(
    .data$Scenario,
    Metric = .data$manuscript_name,
    .data$`Fitted sample`,
    .data$`Effect per LEBA SD (95% CI)`
  )

random_summary <- random_site |>
  count(.data$placement, .data$random_site_status, name = "Models") |>
  mutate(
    Placement = recode(.data$placement, glasses = "Near eye", chest = "Chest"),
    Status = recode(
      .data$random_site_status,
      DESCRIPTIVE_PASS = "Fit passed",
      DESCRIPTIVE_UNSTABLE = "Fit unstable"
    )
  ) |>
  select(.data$Placement, .data$Status, .data$Models)

loo_summary <- leave_one_site_out |>
  count(.data$stability_class, name = "Associations") |>
  mutate(
    Classification = recode(
      .data$stability_class,
      stable = "Stable",
      direction_stable_magnitude_sensitive = "Direction stable; magnitude sensitive",
      direction_unstable = "Direction changed after omitting at least one site"
    )
  ) |>
  select(.data$Classification, .data$Associations)

descriptive_top <- descriptive_spearman |>
  filter(
    .data$run_id == near_id,
    .data$factor_id == "leba_f2",
    .data$metric_id %in% near_top$metric_id
  ) |>
  transmute(
    Metric = .data$manuscript_name,
    `Spearman rho (95% CI)` = paste0(
      format_number(.data$spearman_rho, 3), " (",
      format_number(.data$conf_low, 3), "–",
      format_number(.data$conf_high, 3), ")"
    ),
    Participants = .data$pairs,
    Sites = .data$sites,
    `Represented days` = .data$represented_days
  )

stopifnot(
  nrow(near_results) == 68L,
  nrow(chest_results) == 68L,
  nrow(near_samples) == 17L,
  nrow(chest_samples) == 17L,
  all(family_audit$registry_rows == 68L),
  all(family_audit$vector_bh_verified),
  sum(near_results$p_adjusted <= 0.05) == 0L,
  sum(chest_results$p_adjusted <= 0.05) == 0L,
  sum(
    near_results$reader_inference_status == "unfit_for_inference"
  ) == 4L,
  sum(
    chest_results$reader_inference_status == "unfit_for_inference"
  ) == 4L,
  nrow(paired_display) == 68L,
  all(paired_display$exact_sample_match),
  sum(near_results$model_adequacy == "not_acceptable") == 0L,
  sum(chest_results$model_adequacy == "not_acceptable") == 0L,
  nrow(exact_period) == 4L,
  all(exact_period$sign_concordant),
  sum(random_site$random_site_status == "DESCRIPTIVE_UNSTABLE") == 15L,
  sum(leave_one_site_out$successful_refits) == 612L
)
```

## Question

The preregistered hypothesis was:

> “H5: LEBA questionnaire factors correlate with selected personal light exposure metrics.”

The Light Exposure Behaviour Assessment (LEBA) asks about habitual
light-related behaviours. This analysis tests whether four LEBA factor scores
are associated with 17 personal light-exposure metrics after accounting for
study site. Higher scores mean that the behaviour described by a factor was
reported more often.

::: {.callout-tip title="Answer in brief"}
None of the 68 primary near-eye associations remained statistically supported
after the complete Benjamini--Hochberg correction. The two largest positive
F2 estimates—time above 1,000 lx melEDI and melEDI dose—were consistent across
several sensitivity analyses, but their adjusted values were both
BH-adjusted *p* = `r format_p(near_duration$p_adjusted)`. They are therefore
descriptive, not multiplicity-retained findings. Separately, the four planned
H05 sleep-environment models are unfit for inference and support no estimate
or claim under this response definition and model structure.
:::

The exact data preparation, model construction, diagnostic sequence, and
output lineage are documented in the companion
[H05 analysis preparation and provenance](../../audit/hypotheses/H05/H05_analysis_preparation.html).

## What was analysed

Near-eye measurements are primary because they more closely represent light
near the eyes. All available chest measurements are analysed separately as
complementary evidence. A common-sample analysis then fits near-eye and chest
models to the same participants and participant-days, so placement differences
are not confused with differences in who contributed data. The placements are
never pooled as independent observations.

A predefined sensitivity uses the **gap-timing-unaware dataset**. It still
passed the general 50%-per-hour and 80%-per-day coverage rules. The term means
that the timing of the remaining missing observations is not used for an
additional metric-specific adjustment; it does not mean that gaps,
missingness, or coverage were ignored. For this contrast only, the approved
primary preparation could be interpreted as a **time-sensitive primary metric
dataset**, because it uses the timing of remaining missing observations where
that timing is relevant to the metric. Below, it is called simply the primary
dataset.

```{r}
#| label: tbl-h05-factors
#| tbl-cap: "The four analysed LEBA factors and their score interpretation."

factor_registry |>
  transmute(
    Factor = factor_code(.data$factor_id),
    Behaviour = .data$factor_label,
    `Possible score` = paste0(.data$possible_min, "–", .data$possible_max),
    Interpretation = .data$direction_text
  ) |>
  gt::gt(rowname_col = "Factor") |>
  gt::cols_width(
    Behaviour ~ gt::pct(27),
    `Possible score` ~ gt::pct(14),
    Interpretation ~ gt::pct(48)
  ) |>
  h05_gt()
```

F2 includes one reverse-coded item; F3--F5 are sums of their ordered items.
Every factor score was complete. Within each exact model sample, the score was
centred and expressed both per raw point and per participant-level standard
deviation. The report uses the latter so estimates can be compared across the
four differently ranged factor scores.

### Deviations and operational clarifications

The preregistration gives the broad model `Metric ~ LEBA + (1|Site)` but does
not fully define the repeated-measure structure, response families, metric
set, multiplicity family, clock handling, or placement roles. The following
choices make those elements explicit.

```{r}
#| label: tbl-h05-model-deviations
#| tbl-cap: "H05 model and inferential deviations or operational clarifications."

tibble::tribble(
  ~Aspect, ~Preregistered_or_unspecified, ~Analysis, ~Why_it_matters,
  "Site adjustment",
  "Random site intercept",
  "Submitted-order fixed site effects; registered random-site model retained as a sensitivity",
  "The fixed-site structure was more reliably estimable with nine sites.",
  "Repeated observations",
  "Not operationally specified",
  "Participant-day responses for 15 metrics with a participant random intercept; participant responses for IS and IV",
  "Daily information is retained without treating days from one person as independent.",
  "Response model",
  "Generic linear model",
  "Metric-specific Gaussian or Tweedie model and approved response transform",
  "The outcome scale and support differ across the 17 metrics.",
  "LEBA scale",
  "Not specified",
  "Centred score; effects reported per one participant-level SD",
  "This retains the continuous score while giving a comparable effect unit.",
  "Inference and multiplicity",
  "Correlation matrices plus linear models; adjustment family not operationalized",
  "Full-versus-reduced LEBA comparison and one complete 68-test BH family per inferential placement/scenario",
  "All four factors and 17 metrics remain in the correction."
) |>
  gt::gt(rowname_col = "Aspect") |>
  gt::cols_label(
    Preregistered_or_unspecified = "Preregistered or unspecified",
    Analysis = "Analysed",
    Why_it_matters = "Why it matters"
  ) |>
  gt::cols_width(
    Preregistered_or_unspecified ~ gt::pct(24),
    Analysis ~ gt::pct(34),
    Why_it_matters ~ gt::pct(30)
  ) |>
  h05_gt(12)
```

```{r}
#| label: tbl-h05-data-deviations
#| tbl-cap: "H05 metric, clock, placement, and descriptive-output clarifications."

tibble::tribble(
  ~Aspect, ~Preregistered_or_unspecified, ~Analysis, ~Why_it_matters,
  "Metric set",
  "Selected metrics not enumerated",
  "Seventeen prespecified metrics, including corrected dose and ratio-of-integrals MDER",
  "The 17 × 4 cross-product fixes the 68 tests before inference.",
  "Primary placement",
  "Not specified in H5",
  "All-available near eye primary; all-available chest complementary; common-sample placements as sensitivity",
  "Separates the scientific measurement role from sample composition.",
  "Clock outcomes",
  "Not operationally specified",
  "Continuous supported ranges; darkest-10-hour midpoint values strictly after 16:00 are shifted to the preceding clock scale",
  "Avoids averaging across an artificial midnight discontinuity.",
  "Longest period",
  "No identifiability rule stated",
  "All-available value treated as an observed lower bound; an exactly identified period is a sensitivity",
  "Gaps cannot be mistaken for a continuously observed period.",
  "Participant correlations",
  "Correlation matrices requested",
  "Spearman coefficients and intervals are descriptive only, without a second significance screen",
  "The site-adjusted model family remains the inferential analysis."
) |>
  gt::gt(rowname_col = "Aspect") |>
  gt::cols_label(
    Preregistered_or_unspecified = "Preregistered or unspecified",
    Analysis = "Analysed",
    Why_it_matters = "Why it matters"
  ) |>
  gt::cols_width(
    Preregistered_or_unspecified ~ gt::pct(24),
    Analysis ~ gt::pct(34),
    Why_it_matters ~ gt::pct(30)
  ) |>
  h05_gt(12)
```

## Models

For participant-day outcomes, the full model asks whether LEBA improves the
site-adjusted description while accounting for repeated days from the same
participant. The reduced model removes only LEBA. Interdaily stability (IS)
and intradaily variability (IV) have one response per participant, so they do
not use a participant random intercept. Fixed site effects use the registered
site order and sum-to-zero contrasts; the LEBA coefficient is therefore
adjusted for site without being tied to one reference site.

The following evaluated cell reads and prints the exact formula objects used
for the full, reduced, and registered random-site sensitivity models.

```{r}
#| label: tbl-h05-formulas
#| tbl-cap: "Exact evaluated formulas supplied to the H05 full, reduced, and random-site sensitivity models."
#| echo: true
#| code-fold: false

registered_formula <- function(analysis_unit, formula_id) {
  row <- formula_registry |>
    dplyr::filter(
      .data$analysis_unit == .env$analysis_unit,
      .data$formula_id == .env$formula_id
    )
  stopifnot(nrow(row) == 1L)
  stats::as.formula(row$formula[[1L]])
}

h05_formulas <- list(
  participant_day_full = registered_formula("participant_day", "fixed_full"),
  participant_day_reduced = registered_formula("participant_day", "fixed_reduced"),
  participant_full = registered_formula("participant", "fixed_full"),
  participant_reduced = registered_formula("participant", "fixed_reduced"),
  participant_day_random_site = registered_formula("participant_day", "random_site"),
  participant_random_site = registered_formula("participant", "random_site")
)

tibble::tibble(
  Model = c(
    "Participant-day primary full",
    "Participant-day primary reduced",
    "Participant primary full",
    "Participant primary reduced",
    "Participant-day random-site sensitivity",
    "Participant random-site sensitivity"
  ),
  `Exact Wilkinson formula` = paste0(
    "`",
    vapply(h05_formulas, deparse1, character(1)),
    "`"
  )
) |>
  gt::gt(rowname_col = "Model") |>
  gt::fmt_markdown(columns = `Exact Wilkinson formula`) |>
  gt::cols_width(`Exact Wilkinson formula` ~ gt::pct(67)) |>
  gt::tab_source_note(
    source_note = paste(
      "The table is evaluated directly from the registered formula objects;",
      "the reduced model removes only LEBA from the corresponding full model."
    )
  ) |>
  h05_gt(12)
```

Gaussian full and reduced models were compared using maximum likelihood.
Tweedie models used a log link and maximum likelihood. Effects and 95% Wald
confidence intervals are then presented on a practical response scale: ratios
for log-scale and Tweedie outcomes, an odds ratio for IS, and differences for
identity-scale outcomes. The response specifications are listed below.

```{r}
#| label: tbl-h05-response-specifications
#| tbl-cap: "Metric-specific analysis level, response model, transform, and reported effect scale."

metric_registry |>
  mutate(
    Group = metric_group(.data$metric_order),
    Metric = .data$manuscript_name,
    Level = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    Model = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    Transform = recode(
      .data$response_transform,
      logit = "Logit",
      identity = "None",
      log10_offset_0.1 = "log10(value + 0.1)",
      clock_hours = "Continuous clock hours",
      clock_hours_midnight_after_16 = "Clock hours; strict after-16:00 shift"
    ),
    `Reported effect` = recode(
      .data$effect_scale,
      odds_ratio = "Odds ratio",
      difference = "Difference",
      ratio = "Ratio"
    )
  ) |>
  select(.data$Group, .data$Metric, .data$Level, .data$Model, .data$Transform, .data$`Reported effect`) |>
  gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
  gt::cols_width(
    Level ~ gt::pct(16), Model ~ gt::pct(17),
    Transform ~ gt::pct(27), `Reported effect` ~ gt::pct(16)
  ) |>
  h05_gt(12)
```

## Primary near-eye result

Depending on the metric, the primary models used 139--141 participants,
655--816 participant-days for participant-day outcomes, and all nine sites.
IS and IV used 141 participant rows and represented 816 observed days. Exact
metric-specific samples are shown below.

```{r}
#| label: tbl-h05-near-samples
#| tbl-cap: "Exact fitted near-eye sample for each light-exposure metric."

sample_table(near_samples)
```

The exact sample table is available as
[source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_samples.csv).

### Complete association pattern

```{r}
#| label: fig-h05-near-effects
#| fig-cap: "Near-eye effects per participant SD of each LEBA factor. Grey sleep-environment cells are unfit for inference and their estimates are suppressed."
#| fig-alt: "Heatmap with 17 light-exposure metrics as rows and LEBA factors F2 to F5 as columns. Cells show practical-scale effects per participant standard deviation, except four grey sleep-environment cells labelled unfit. No cell is marked as retained after the 68-test adjustment."
#| out-width: "96%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_near_eye_effects.png"
))
```

The heatmap colours show the signed effect on each model's fitted scale, while
the printed value is on the practical response scale. Colours therefore show
direction and relative within-outcome magnitude, not a common physical unit
across all 17 rows. The tables provide every estimate, 95% confidence
interval, Raw p, and BH-adjusted p.

```{r}
#| label: tbl-h05-near-results-a
#| tbl-cap: "Near-eye associations for stability, light level, duration, and continuous-period metrics."

result_matrix(near_results, 1:10)
```

```{r}
#| label: tbl-h05-near-results-b
#| tbl-cap: "Near-eye associations for timing, exposure-history, and spectrum metrics."

result_matrix(near_results, 11:17)
```

Across the complete planned family, zero of 68 BH-adjusted p-values were at or
below 0.050. The four sleep-environment models remain in that archived family
so the correction is not changed after seeing the diagnostics, but they are
unfit for inference and their estimates and p-values are suppressed from the
reader-facing matrices.
The two smallest raw values were both for F2, spending time outdoors:

- time above 1,000 lx melEDI increased by a factor of
  `r format_number(near_duration$estimate_practical_per_sd)` per participant
  SD (95% CI `r format_number(near_duration$conf_low_practical_per_sd)`–`r format_number(near_duration$conf_high_practical_per_sd)`),
  Raw *p* = `r format_p(near_duration$p_raw)`, BH-adjusted
  *p* = `r format_p(near_duration$p_adjusted)`; and
- melEDI dose increased by a factor of
  `r format_number(near_dose$estimate_practical_per_sd)` (95% CI
  `r format_number(near_dose$conf_low_practical_per_sd)`–`r format_number(near_dose$conf_high_practical_per_sd)`),
  Raw *p* = `r format_p(near_dose$p_raw)`, BH-adjusted
  *p* = `r format_p(near_dose$p_adjusted)`.

These estimates remain useful for describing magnitude and uncertainty, but
neither is a multiplicity-retained result.

Complete near-eye numerical results and figure data are available as
[result source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_results.csv)
and [figure source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_effect_figure_data.csv).

### Near-eye model assessment

```{r}
#| label: fig-h05-near-adequacy
#| fig-cap: "Near-eye model-adequacy classifications for all 68 associations."
#| fig-alt: "Heatmap of model assessments for 17 near-eye metrics and four LEBA factors. Nineteen cells are green for acceptable, 45 are orange for acceptable with specified limitations, and the four sleep-environment cells are red for unfit for inference."
#| out-width: "94%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_near_eye_adequacy.png"
))
```

```{r}
#| label: tbl-h05-near-adequacy-counts
#| tbl-cap: "Near-eye model-adequacy counts."

adequacy_count_table(near_results)
```

```{r}
#| label: tbl-h05-near-limitations
#| tbl-cap: "Specified near-eye diagnostic limitations."

limitation_count_table(near_results)
```

All fixed-site fits completed, and all applicable Hessian checks were positive
definite. That numerical completion does not make every response model
scientifically defensible. Nineteen models were acceptable, 45 were acceptable
with a specified limitation, and the four sleep-environment models were
classified **unfit for inference**. Most retained limitations were ordinary
Gaussian residual warnings. They qualify how closely the working residual
distribution represents these data but do not indicate numerical fit failure.

The following representative screens show F2 for the two leading estimates
and the sleep-environment outcome. The smooth line in the
residual-versus-fitted plot helps reveal fitted-dependent residual structure; the quantile
plot compares standardized residuals with a normal reference. For the
Tweedie sleep model, the simulation-based tests in the subsequent table—not
the normal-reference plot—govern the assessment.

```{r}
#| label: fig-h05-near-residual-fitted
#| fig-cap: "Selected near-eye standardized residuals against fitted values."
#| fig-alt: "Three vertically stacked residual-versus-fitted plots for F2 models of time above 1,000 lx melEDI, melEDI dose, and time below 1 lx melEDI during sleep. Points show standardized Pearson residuals and blue smooths show their average pattern."
#| out-width: "88%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05",
  "H05_reader_near_eye_residual_fitted.png"
))
```

```{r}
#| label: fig-h05-near-residual-qq
#| fig-cap: "Selected near-eye residual quantile checks."
#| fig-alt: "Three vertically stacked normal-reference quantile plots for the same F2 models. The Gaussian melEDI-dose model has tail departures, while simulation diagnostics and response-support failures make the Tweedie sleep model unfit for inference."
#| out-width: "88%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05",
  "H05_reader_near_eye_residual_qq.png"
))
```

#### H05 sleep-environment models are unfit for inference

The four models for time below 1 lx melEDI during sleep require a stronger
qualification than the ordinary residual warnings. Their simulated residuals
showed pronounced non-uniformity, zero-mass disagreement, and outlier
evidence. They also produced 183--187 fitted values above the
participant-specific response support. These models are retained because the
prespecified response-family assessment did not identify an adequate common
replacement. They are therefore shown for audit completeness but classified
**unfit for inference**: their estimates, confidence intervals, and p-values
are not used to support any scientific claim.

This classification is specific to this H05 sleep-environment response and
the model structure used here. It does not determine whether the underlying
metric can be analysed in another hypothesis that uses a different response
variable or model structure.

```{r}
#| label: tbl-h05-near-sleep-diagnostics
#| tbl-cap: "Diagnostic evidence for classifying the near-eye sleep-environment models as unfit for inference."

sleep_diagnostic_table(near_id)
```

The near-eye family supports the overall null conclusion through the 64
reader-interpretable cells: 19 are acceptable and 45 have specified residual
limitations. The four additional planned sleep-environment cells remain in
the archived multiplicity family but are unfit for inference.

Exact source data for the selected diagnostic figures are available
[here](../../artifacts/11_source_data/H05/H05_reader_near_eye_selected_diagnostics.csv).
The model-assessment heatmap has its own
[paired source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_adequacy_figure_data.csv).

## Complementary chest result

The chest analysis repeats the same factor scores, metric definitions,
response models, fixed-site adjustment, effect scaling, and 68-test
correction. Depending on the metric, it used 153--154 participants, 743--902
participant-days for daily outcomes, and eight sites. IS and IV used 153
participant rows and represented 900 observed days.

```{r}
#| label: tbl-h05-chest-samples
#| tbl-cap: "Exact fitted chest sample for each light-exposure metric."

sample_table(chest_samples)
```

The exact chest sample table is available as
[source data](../../artifacts/11_source_data/H05/H05_reader_chest_samples.csv).

```{r}
#| label: fig-h05-chest-effects
#| fig-cap: "Chest effects per participant SD of each LEBA factor. Grey sleep-environment cells are unfit for inference and their estimates are suppressed."
#| fig-alt: "Heatmap with 17 chest light-exposure metrics as rows and LEBA factors F2 to F5 as columns. Cells show practical-scale effects per participant standard deviation, except four grey sleep-environment cells labelled unfit. No cell is retained after the chest-specific 68-test adjustment."
#| out-width: "96%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_chest_effects.png"
))
```

```{r}
#| label: tbl-h05-chest-results-a
#| tbl-cap: "Chest associations for stability, light level, duration, and continuous-period metrics."

result_matrix(chest_results, 1:10)
```

```{r}
#| label: tbl-h05-chest-results-b
#| tbl-cap: "Chest associations for timing, exposure-history, and spectrum metrics."

result_matrix(chest_results, 11:17)
```

Zero of 68 planned chest associations remained after its separate correction;
the smallest value was BH-adjusted *p* =
`r format_p(min(chest_results$p_adjusted))`. The F2 estimates for time above
1,000 lx melEDI and melEDI dose were again positive—ratios
`r format_number(chest_duration$estimate_practical_per_sd)` and
`r format_number(chest_dose$estimate_practical_per_sd)`, respectively—but
neither was multiplicity-retained.

Complete chest numerical results and figure data are available as
[result source data](../../artifacts/11_source_data/H05/H05_reader_chest_results.csv)
and [figure source data](../../artifacts/11_source_data/H05/H05_reader_chest_effect_figure_data.csv).

### Chest model assessment

```{r}
#| label: fig-h05-chest-adequacy
#| fig-cap: "Chest model-adequacy classifications for all 68 associations."
#| fig-alt: "Heatmap of model assessments for 17 chest metrics and four LEBA factors. Fourteen cells are green for acceptable, 50 are orange for acceptable with specified limitations, and the four sleep-environment cells are red for unfit for inference."
#| out-width: "94%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_chest_adequacy.png"
))
```

```{r}
#| label: tbl-h05-chest-adequacy-counts
#| tbl-cap: "Chest model-adequacy counts."

adequacy_count_table(chest_results)
```

```{r}
#| label: tbl-h05-chest-limitations
#| tbl-cap: "Specified chest diagnostic limitations."

limitation_count_table(chest_results)
```

Fourteen chest models were acceptable and 50 were acceptable with specified
limitations. The four chest sleep-environment models were **unfit for
inference**. They show the same material problem as near eye, with 241--244
predictions above participant-specific support and strong simulated-residual
disagreement; their estimates and p-values are suppressed from the
reader-facing matrices.

```{r}
#| label: tbl-h05-chest-sleep-diagnostics
#| tbl-cap: "Diagnostic evidence for classifying the chest sleep-environment models as unfit for inference."

sleep_diagnostic_table(chest_id)
```

The 64 reader-interpretable chest cells provide complementary evidence for the
overall null result, subject to their residual warnings. The four additional
planned sleep-environment cells remain in the archived family but are unfit
for inference.

The chest model-assessment heatmap has
[paired source data](../../artifacts/11_source_data/H05/H05_reader_chest_adequacy_figure_data.csv).

## Sensitivity analyses

### Common near-eye and chest sample

Every plotted point compares the same metric, factor, response definition,
model, transformation, and exact sample at both placements. Participant-day
metrics used 110--112 participants and 505--643 matched participant-days at
eight sites; IS and IV used the same 112 participant rows at eight sites.
Fifty-nine of 68 associations had the same direction at both placements and
overlapping component intervals; nine differed in direction but still had
overlapping component intervals.
Because no participant-cluster interval was calculated for the direct
chest-minus-near-eye difference, these are component comparisons rather than formal
tests of placement differences.

```{r}
#| label: fig-h05-paired-placement
#| fig-cap: "Exact common-sample model-scale effects, with near eye on x and chest on y. The dashed identity line describes concordance, grey lines mark the null, and equal x/y geometry preserves the 45-degree comparison. Samples contain 110–112 participants, 505–643 matched participant-days, and eight sites; IS and IV use 112 participant rows. Component 95% confidence intervals are in the adjacent tables. This is not an equivalence analysis."
#| fig-alt: "Four equal-axis scatterplot panels, one per LEBA factor, compare 17 exactly matched estimands with near-eye effects on the horizontal axis and chest effects on the vertical axis. A dashed 45-degree identity line and grey horizontal and vertical null lines are visible. Samples contain 110 to 112 participants, 505 to 643 matched participant-days, and eight sites; IS and IV use 112 participant rows. Most points have the same sign, while nine cross an axis; proximity to identity does not establish equivalence."
#| out-width: "90%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_paired_placement_effects.png"
))
```

The component estimates and 95% confidence intervals corresponding to every
plotted point are shown below.

::: {.panel-tabset}

### F2 — spending time outdoors

```{r}
#| label: tbl-h05-paired-f2
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F2."

paired_component_table("leba_f2")
```

### F3 — devices in bed

```{r}
#| label: tbl-h05-paired-f3
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F3."

paired_component_table("leba_f3")
```

### F4 — ambient light before bedtime

```{r}
#| label: tbl-h05-paired-f4
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F4."

paired_component_table("leba_f4")
```

### F5 — morning/daytime light use

```{r}
#| label: tbl-h05-paired-f5
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F5."

paired_component_table("leba_f5")
```

:::

Paired figure source data are available
[here](../../artifacts/11_source_data/H05/H05_paired_effect_comparison_data.csv).

### Stability of the two leading F2 estimates

```{r}
#| label: tbl-h05-leading-sensitivities
#| tbl-cap: "The two leading F2 estimates across placement, sample, and preparation sensitivities."

leading_sensitivity |>
  gt::gt(rowname_col = "Scenario", groupname_col = "Metric") |>
  gt::cols_width(
    `Fitted sample` ~ gt::pct(32),
    `Effect per LEBA SD (95% CI)` ~ gt::pct(30)
  ) |>
  h05_gt(12)
```

Both leading F2 directions remain positive on the common near-eye and chest
sample and in the gap-timing-unaware dataset. These checks support the
stability of the descriptive pattern; they do not create a new significance
screen.

As a direct participant-level description, the Spearman coefficients were:

```{r}
#| label: tbl-h05-descriptive-spearman
#| tbl-cap: "Descriptive participant-level F2 rank correlations for the two leading metrics."

descriptive_top |>
  gt::gt(rowname_col = "Metric") |>
  gt::fmt_integer(columns = c(Participants, Sites, `Represented days`)) |>
  gt::tab_source_note(
    source_note = paste(
      "No p-values or significance screen are attached to these descriptive",
      "correlations. Intervals use participant resampling."
    )
  ) |>
  h05_gt(12)
```

### Site adjustment and site influence

Fixed site was selected as primary because the inherited 17-metric
site-structure evaluation fitted all fixed-site models successfully, whereas
the corresponding random-site descriptions included two unstable, two
non-estimable, and two singular fits. In the H05 sensitivity itself, 15 of 136
main all-available random-site fits were unstable—seven near eye and eight
chest—while 121 passed. This instability is why random site is a sensitivity
rather than the primary adjustment.

```{r}
#| label: tbl-h05-random-site-summary
#| tbl-cap: "Random-site sensitivity fit status across near-eye and chest associations."

random_summary |>
  gt::gt(rowname_col = "Status", groupname_col = "Placement") |>
  gt::fmt_integer(columns = Models) |>
  h05_gt()
```

All 612 required leave-one-site-out refits completed. Twenty-one associations
were stable, 27 retained direction but were magnitude-sensitive, and 20
changed direction after at least one site was omitted. The two leading F2
associations had no site-omission sign reversal and were classified stable.
The broader pattern confirms that some small, near-zero estimates depend on
which of the nine sites is represented, reinforcing the restrained
family-wide interpretation.

```{r}
#| label: tbl-h05-leave-site-summary
#| tbl-cap: "Primary leave-one-site-out stability classifications."

loo_summary |>
  gt::gt(rowname_col = "Classification") |>
  gt::fmt_integer(columns = Associations) |>
  h05_gt()
```

### Gap-timing-unaware dataset

The same model implementation was applied to the gap-timing-unaware dataset,
with only confirmed data errors corrected. This sensitivity gave no
multiplicity-retained association; its smallest value was BH-adjusted *p* =
`r format_p(min(preparation_results$p_adjusted, na.rm = TRUE))`. Sixty-four of
68 estimates retained their sign and had overlapping component intervals;
four changed sign, also with overlapping component intervals. The main null
conclusion is unchanged.

### Exactly identified longest period

The all-available longest continuous period above 250 lx melEDI is an
observed lower bound because a data gap can interrupt identifiability. The
strict sensitivity retained only participant-days whose longest period was
exactly identified: 500 participant-days from 132 participants at all nine
sites. All four factor directions agreed with the all-available analysis.

```{r}
#| label: tbl-h05-exact-period
#| tbl-cap: "Exactly identified longest-period sensitivity."

exact_period |>
  transmute(
    Factor = factor_code(.data$factor_id),
    `All-available ratio` = format_number(
      .data$all_available_estimate_practical_per_sd, 3
    ),
    `Exactly identified ratio (95% CI)` = paste0(
      format_number(.data$estimate_practical_per_sd, 3), " (",
      format_number(.data$conf_low_practical_per_sd, 3), "–",
      format_number(.data$conf_high_practical_per_sd, 3), ")"
    ),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Direction retained` = if_else(.data$sign_concordant, "Yes", "No")
  ) |>
  gt::gt(rowname_col = "Factor") |>
  gt::fmt_integer(columns = c(Participants, `Participant-days`, Sites)) |>
  h05_gt(12)
```

For F2, the ratio attenuated from 1.117 to 1.046 (95% CI 0.930--1.175),
supporting the decision not to promote this lower-bound metric as a positive
association.

Complete sensitivity sources are available for the
[common-sample placement comparison](../../artifacts/09_tables/H05/H05_paired_placement_comparison.csv),
[random-site models](../../artifacts/09_tables/H05/H05_random_site_sensitivity.csv),
[leave-one-site-out analysis](../../artifacts/09_tables/H05/H05_leave_one_site_out_summary.csv),
[gap-timing-unaware dataset](../../artifacts/11_source_data/H05/H05_gap_timing_unaware_dataset.csv),
and [exactly identified longest-period analysis](../../artifacts/09_tables/H05/H05_exactly_identified_longest_bout_sensitivity.csv).

## Interpretation

The strongest defensible conclusion is that, after fixed-site adjustment and
one complete 68-test Benjamini--Hochberg correction, these data provide no
multiplicity-retained evidence that any of the four LEBA factors is associated
with the 17 selected personal light-exposure metrics.

The positive F2 estimates for time above 1,000 lx melEDI and melEDI dose are
consistent across the principal stability checks. Their confidence intervals
describe non-zero associations before multiplicity adjustment, but their
adjusted values do not support confirmatory claims. They should be treated as
descriptive patterns that may motivate future work, not as established
effects of spending time outdoors.

This is an observational association analysis. LEBA scores are self-reported,
site adjustment cannot remove all person-level confounding, only nine sites
contribute to the primary analysis, and several small estimates are sensitive
to site omission. Near-eye and chest measurements also represent related but
non-identical light fields. Finally, the sleep-environment Tweedie models are
unfit for inference because of material response-support and
simulated-residual failures; their estimates and p-values are suppressed from
the reader-facing results and cannot support a claim in H05. This judgment is
specific to the response variable and model structure used here and does not
preclude analysing the metric in another hypothesis with a different response
or model.

```{r}
#| label: session-info-h05-reader
#| include: false

sessionInfo()
```
