---
title: "H05: Habitual light-related behaviour and personal light exposure"
subtitle: "Associations between habitual light-related behaviours and 17 light-exposure metrics"
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
    lightbox: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h05-reader

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()),
  winslash = "/",
  mustWork = TRUE
)
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)
source(file.path(root, "scripts", "pipeline", "p_value_display.R"))

read_h05 <- function(...) {
  readr::read_csv(file.path(root, ...), show_col_types = FALSE)
}

near_results <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_near_eye_results.csv"
)
chest_results <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_chest_results.csv"
)
near_samples <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_near_eye_samples.csv"
)
chest_samples <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_reader_chest_samples.csv"
)
master <- read_h05(
  "artifacts", "09_tables", "H05", "H05_model_results_master.csv"
)
diagnostics <- read_h05(
  "artifacts", "08_diagnostics", "H05", "H05_model_diagnostics.csv"
)
formula_registry <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_formula_registry.csv"
)
factor_registry <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_factor_registry.csv"
)
metric_registry <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_metric_registry.csv"
)
family_audit <- read_h05(
  "artifacts", "09_tables", "H05", "H05_family_audit.csv"
)
site_evidence <- read_h05(
  "artifacts", "06_model_data", "H05", "H05_site_structure_evidence.csv"
)
random_site <- read_h05(
  "artifacts", "09_tables", "H05", "H05_random_site_sensitivity.csv"
)
leave_one_site_out <- read_h05(
  "artifacts", "09_tables", "H05", "H05_leave_one_site_out_summary.csv"
)
paired <- read_h05(
  "artifacts", "09_tables", "H05", "H05_paired_placement_comparison.csv"
)
paired_display <- read_h05(
  "artifacts", "11_source_data", "H05",
  "H05_paired_effect_comparison_data.csv"
)
preparation_sensitivity <- read_h05(
  "artifacts", "09_tables", "H05", "H05_manuscript_prepared_comparison.csv"
)
exact_period <- read_h05(
  "artifacts", "09_tables", "H05",
  "H05_exactly_identified_longest_bout_sensitivity.csv"
)
descriptive_spearman <- read_h05(
  "artifacts", "09_tables", "H05", "H05_descriptive_spearman.csv"
)
mder_upper_tail <- read_h05(
  "artifacts", "08_diagnostics", "H05",
  "H05_mder_metric010_upper_tail_summary.csv"
)
mder_influence <- read_h05(
  "artifacts", "08_diagnostics", "H05",
  "H05_mder_metric010_influence_refits.csv"
)
mder_gap_influence <- read_h05(
  "artifacts", "08_diagnostics", "H05",
  "H05_mder_metric010_gap_influence_refits.csv"
)
mder_gap_paired <- read_h05(
  "artifacts", "09_tables", "H05",
  "H05_mder_metric010_gap_paired_placement_comparison.csv"
)
mder_gap_reconciliation <- read_h05(
  "artifacts", "12_manifests", "H05",
  "H05_metric010_gap_reseal_reconciliation.csv"
)
l10_frame_audit <- read_h05(
  "artifacts", "12_manifests", "H05",
  "H05_metric011_primary_l10_frame_audit.csv"
)
l10_bh_audit <- read_h05(
  "artifacts", "12_manifests", "H05",
  "H05_metric011_bh_change_audit.csv"
)
l10_result_change <- read_h05(
  "artifacts", "12_manifests", "H05",
  "H05_metric011_l10_result_change_audit.csv"
)
l10_input_cells <- read_h05(
  "artifacts", "12_manifests", "H05",
  "H05_metric011_input_cell_audit.csv"
)
l10_reconciliation <- read_h05(
  "artifacts", "12_manifests", "H05",
  "H05_metric011_reconciliation.csv"
)

near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
paired_near_id <- "main__glasses__paired_common_sample"
paired_chest_id <- "main__chest__paired_common_sample"
preparation_id <- "manuscript_prepared_data__glasses__all_available"
mder_id <- "mder_mean_of_viable_ratios"

factor_code <- function(factor_id) {
  toupper(sub("leba_", "", factor_id))
}

format_number <- function(value, digits = 3L) {
  formatC(value, digits = digits, format = "f", big.mark = ",")
}

format_p <- function(value) {
  nh_format_p_value(value)
}

format_p_cell <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  if (display$p_bold[[1L]]) {
    paste0("**", display$p_display[[1L]], "**")
  } else {
    display$p_display[[1L]]
  }
}

format_effect <- function(estimate, low, high, type, unit, digits = 3L) {
  if (type == "odds_ratio") {
    return(paste0(
      "OR ", format_number(estimate, digits), " (",
      format_number(low, digits), "–", format_number(high, digits), ")"
    ))
  }
  if (type == "ratio") {
    return(paste0(
      "×", format_number(estimate, digits), " (",
      format_number(low, digits), "–", format_number(high, digits), ")"
    ))
  }
  suffix <- if (unit %in% c("h", "clock time")) " h" else ""
  paste0(
    sprintf(paste0("%+.", digits, "f"), estimate), suffix, " (",
    sprintf(paste0("%+.", digits, "f"), low), "–",
    sprintf(paste0("%+.", digits, "f"), high), suffix, ")"
  )
}

h05_gt <- function(table, font_size = 12) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

metric_group <- function(metric_order) {
  dplyr::case_when(
    metric_order <= 5 ~ "Stability and light level",
    metric_order <= 10 ~ "Duration and continuous periods",
    metric_order <= 15 ~ "Timing",
    TRUE ~ "Exposure history and spectrum"
  )
}

sample_table <- function(data) {
  data |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      `Analysis level` = if_else(
        .data$analysis_unit == "participant",
        "Participant",
        "Participant-day"
      ),
      `Model rows` = .data$observations,
      Participants = .data$participants,
      `Participant-days` = .data$participant_days,
      `Represented days` = .data$represented_days,
      Sites = .data$sites
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$`Analysis level`,
      .data$`Model rows`,
      .data$Participants,
      .data$`Participant-days`,
      .data$`Represented days`,
      .data$Sites
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::fmt_integer(
      columns = c(
        `Model rows`, Participants, `Participant-days`,
        `Represented days`, Sites
      )
    ) |>
    gt::cols_width(
      `Analysis level` ~ gt::pct(17),
      `Model rows` ~ gt::pct(12),
      Participants ~ gt::pct(12),
      `Participant-days` ~ gt::pct(14),
      `Represented days` ~ gt::pct(14),
      Sites ~ gt::pct(8)
    ) |>
    h05_gt(12)
}

result_matrix <- function(
  data,
  metric_orders,
  source_note_lead = "Effects are per one participant-level SD of the LEBA score with"
) {
  data |>
    filter(.data$metric_order %in% .env$metric_orders) |>
    left_join(
      metric_registry |>
        select(.data$metric_id, .data$display_unit),
      by = "metric_id",
      relationship = "many-to-one"
    ) |>
    rowwise() |>
    mutate(
      Metric = .data$manuscript_name,
      Factor = factor_code(.data$factor_id),
      Result = if_else(
        .data$reader_inference_status == "unfit_for_inference",
        paste0(
          "**Unfit for inference**<br><small>Estimate and p-values ",
          "suppressed from the reader display</small>"
        ),
        paste0(
          format_effect(
            .data$estimate_practical_per_sd,
            .data$conf_low_practical_per_sd,
            .data$conf_high_practical_per_sd,
            .data$effect_type,
            .data$display_unit
          ),
          "<br><small>Raw p = ",
          format_p_cell(.data$p_raw, FALSE),
          "; FDR-adjusted p = ",
          format_p_cell(.data$p_adjusted, .data$p_adjusted <= 0.05),
          "</small>"
        )
      )
    ) |>
    ungroup() |>
    select(.data$metric_order, .data$Metric, .data$Factor, .data$Result) |>
    tidyr::pivot_wider(names_from = .data$Factor, values_from = .data$Result) |>
    arrange(.data$metric_order) |>
    select(-.data$metric_order) |>
    gt::gt(rowname_col = "Metric") |>
    gt::fmt_markdown(columns = c(F2, F3, F4, F5)) |>
    gt::cols_width(
      F2 ~ gt::pct(18), F3 ~ gt::pct(18),
      F4 ~ gt::pct(18), F5 ~ gt::pct(18)
    ) |>
    gt::tab_source_note(
      source_note = paste(
        source_note_lead,
        "95% CIs. Raw p has no separate bolding rule; FDR-adjusted p is",
        "bold only at alpha = 0.050 across all 68 tests.",
        "Sleep-environment cells are unfit for inference and suppressed."
      )
    ) |>
    h05_gt(12)
}

adequacy_count_table <- function(data) {
  data |>
    mutate(
      reader_assessment = if_else(
        .data$reader_inference_status == "unfit_for_inference",
        "unfit_for_inference",
        .data$model_adequacy
      )
    ) |>
    count(.data$reader_assessment, name = "Models") |>
    mutate(
      Assessment = recode(
        .data$reader_assessment,
        acceptable = "Acceptable",
        acceptable_with_specified_limitations =
          "Acceptable with specified limitations",
        unfit_for_inference = "Unfit for inference",
        not_acceptable = "Not acceptable"
      )
    ) |>
    select(.data$Assessment, .data$Models) |>
    gt::gt(rowname_col = "Assessment") |>
    gt::fmt_integer(columns = Models) |>
    h05_gt()
}

limitation_count_table <- function(data) {
  data |>
    mutate(
      Limitation = case_when(
        is.na(.data$specified_limitations) ~ "None specified",
        grepl("WARN_STRONG_TWEEDIE_MISFIT", .data$specified_limitations) ~
          "Unfit sleep-environment model: simulated-residual and response-support failure",
        grepl("WARN_TWEEDIE_DIAGNOSTIC", .data$specified_limitations) ~
          "Ordinary Tweedie simulation-based model-check warning",
        grepl("WARN_GAUSSIAN_DIAGNOSTIC", .data$specified_limitations) ~
          "Ordinary Gaussian residual warning",
        TRUE ~ .data$specified_limitations
      )
    ) |>
    count(.data$Limitation, name = "Models") |>
    arrange(desc(.data$Models)) |>
    gt::gt(rowname_col = "Limitation") |>
    gt::fmt_integer(columns = Models) |>
    h05_gt()
}

sleep_diagnostic_table <- function(run_id) {
  diagnostics |>
    filter(
      .data$run_id == .env$run_id,
      .data$metric_id == "duration_below_1_sleep_environment"
    ) |>
    transmute(
      Factor = factor_code(.data$factor_id),
      `Uniformity raw p` = nh_format_p_value(.data$dharma_uniformity_p),
      `Zero-mass raw p` = nh_format_p_value(.data$dharma_zero_inflation_p),
      `Outlier raw p` = nh_format_p_value(.data$dharma_outlier_p),
      `Observed zero fraction` = .data$observed_zero_fraction,
      `Simulated zero fraction` = .data$simulated_zero_fraction,
      `Predictions above support` = .data$predicted_above_bound_n
    ) |>
    gt::gt(rowname_col = "Factor") |>
    gt::fmt_number(
      columns = c(`Observed zero fraction`, `Simulated zero fraction`),
      decimals = 4
    ) |>
    gt::fmt_integer(columns = `Predictions above support`) |>
    gt::tab_source_note(
      source_note = paste(
        "These are raw simulation-based model-check p-values. No standalone alpha",
        "or bolding rule is applied; they are interpreted jointly with the",
        "response-support failures that make these models unfit for inference."
      )
    ) |>
    h05_gt(12)
}

paired_component_table <- function(factor_id) {
  paired_display |>
    filter(.data$factor_id == .env$factor_id) |>
    arrange(.data$metric_order) |>
    rowwise() |>
    transmute(
      Metric = .data$manuscript_name,
      `Near-eye model-scale estimate (95% CI)` = paste0(
        format_number(.data$estimate_model_per_sd__glasses), " (",
        format_number(.data$conf_low_model_per_sd__glasses), "–",
        format_number(.data$conf_high_model_per_sd__glasses), ")"
      ),
      `Chest model-scale estimate (95% CI)` = paste0(
        format_number(.data$estimate_model_per_sd__chest), " (",
        format_number(.data$conf_low_model_per_sd__chest), "–",
        format_number(.data$conf_high_model_per_sd__chest), ")"
      ),
      `Exact matched sample` = if_else(
        .data$analysis_unit__near_eye == "participant",
        paste0(
          .data$participants__near_eye,
          " participants (",
          .data$observations__near_eye,
          " rows); ",
          .data$sites__near_eye,
          " sites"
        ),
        paste0(
          .data$participants__near_eye,
          " participants; ",
          .data$participant_days__near_eye,
          " participant-days; ",
          .data$sites__near_eye,
          " sites"
        )
      )
    ) |>
    ungroup() |>
    gt::gt(rowname_col = "Metric") |>
    gt::cols_width(
      `Near-eye model-scale estimate (95% CI)` ~ gt::pct(24),
      `Chest model-scale estimate (95% CI)` ~ gt::pct(24),
      `Exact matched sample` ~ gt::pct(25)
    ) |>
    gt::tab_source_note(
      source_note = paste(
        "These are the component 95% confidence intervals for the exact",
        "matched estimands plotted above. The identity display describes",
        "concordance and is not an equivalence analysis."
      )
    ) |>
    h05_gt(12)
}

mder_scenario_labels <- c(
  `main__glasses__all_available` = "All-available near eye",
  `main__chest__all_available` = "All-available chest",
  `main__glasses__paired_common_sample` = "Common-sample near eye",
  `main__chest__paired_common_sample` = "Common-sample chest",
  `manuscript_prepared_data__glasses__all_available` =
    "All-available near eye",
  `manuscript_prepared_data__chest__all_available` =
    "All-available chest",
  `manuscript_prepared_data__glasses__paired_common_sample` =
    "Common-sample near eye",
  `manuscript_prepared_data__chest__paired_common_sample` =
    "Common-sample chest"
)

mder_scenarios <- master |>
  filter(.data$metric_id == mder_id) |>
  mutate(
    Scenario = unname(mder_scenario_labels[.data$run_id]),
    Scenario = factor(
      .data$Scenario,
      levels = c(
        "All-available near eye", "All-available chest",
        "Common-sample near eye", "Common-sample chest"
      )
    )
  ) |>
  arrange(.data$data_scenario_id, .data$Scenario, .data$factor_order)
mder_primary_f5 <- mder_scenarios |>
  filter(
    .data$run_id == near_id,
    .data$factor_id == "leba_f5"
  )
mder_gap_min_adjusted <- mder_scenarios |>
  filter(.data$run_id == preparation_id) |>
  summarise(value = min(.data$p_adjusted, na.rm = TRUE)) |>
  pull(.data$value)

mder_scenario_table <- function(data_scenario_id) {
  mder_scenarios |>
    filter(.data$data_scenario_id == .env$data_scenario_id) |>
    rowwise() |>
    transmute(
      Scenario = .data$Scenario,
      Factor = factor_code(.data$factor_id),
      `MDER difference per LEBA SD (95% CI)` = format_effect(
        .data$estimate_practical_per_sd,
        .data$conf_low_practical_per_sd,
        .data$conf_high_practical_per_sd,
        .data$effect_type,
        ""
      ),
      `Raw p` = if_else(
        .data$inferential_family,
        format_p_cell(.data$p_raw, FALSE),
        "—"
      ),
      `FDR-adjusted p` = if_else(
        .data$inferential_family,
        format_p_cell(.data$p_adjusted, .data$p_adjusted <= 0.05),
        "—"
      ),
      `Exact fitted sample` = paste0(
        .data$participants, " participants; ",
        .data$participant_days, " participant-days; ",
        .data$sites, " sites"
      )
    ) |>
    ungroup() |>
    gt::gt(groupname_col = "Scenario", rowname_col = "Factor") |>
    gt::fmt_markdown(columns = c(`Raw p`, `FDR-adjusted p`)) |>
    gt::cols_width(
      `MDER difference per LEBA SD (95% CI)` ~ gt::pct(29),
      `Exact fitted sample` ~ gt::pct(31)
    ) |>
    gt::tab_source_note(
      source_note = paste(
        "Raw and FDR-adjusted p-values are shown only for declared",
        "68-test inferential families. Common-sample placement analyses",
        "are descriptive sensitivities and have no separate significance screen."
      )
    ) |>
    h05_gt(12)
}

mder_influence_summary <- bind_rows(
  mder_influence |>
    mutate(Dataset = "Primary dataset"),
  mder_gap_influence |>
    mutate(Dataset = "Gap-timing-unaware dataset")
) |>
  group_by(
    .data$Dataset,
    .data$run_id,
    .data$factor_order,
    .data$factor_id
  ) |>
  summarise(
    `Candidate refits` = n(),
    `Maximum absolute change (full-model SE)` = max(
      abs(.data$change_in_full_standard_errors)
    ),
    `Any sign reversal` = any(.data$sign_reversal),
    `All sensitivity intervals include zero` = all(
      .data$sensitivity_interval_contains_zero
    ),
    .groups = "drop"
  ) |>
  mutate(
    Placement = if_else(
      grepl("glasses", .data$run_id, fixed = TRUE),
      "Near eye",
      "Chest"
    ),
    Analysis = paste(.data$Dataset, .data$Placement, sep = ": "),
    Factor = factor_code(.data$factor_id)
  )

near_top <- near_results |>
  filter(
    .data$factor_id == "leba_f2",
    .data$metric_id %in% c(
      "duration_above_1000",
      "dose_time_sensitive_corrected_medi"
    )
  ) |>
  arrange(.data$metric_order)
near_duration <- near_top |>
  filter(.data$metric_id == "duration_above_1000")
near_dose <- near_top |>
  filter(.data$metric_id == "dose_time_sensitive_corrected_medi")

chest_duration <- chest_results |>
  filter(
    .data$factor_id == "leba_f2",
    .data$metric_id == "duration_above_1000"
  )
chest_dose <- chest_results |>
  filter(
    .data$factor_id == "leba_f2",
    .data$metric_id == "dose_time_sensitive_corrected_medi"
  )

near_adequacy <- near_results |> count(.data$model_adequacy)
chest_adequacy <- chest_results |> count(.data$model_adequacy)

paired_near <- master |>
  filter(.data$run_id == paired_near_id)
paired_chest <- master |>
  filter(.data$run_id == paired_chest_id)
preparation_results <- master |>
  filter(.data$run_id == preparation_id)

leading_sensitivity <- bind_rows(
  near_top |> mutate(Scenario = "Primary near eye"),
  paired_near |>
    filter(
      .data$factor_id == "leba_f2",
      .data$metric_id %in% near_top$metric_id
    ) |>
    mutate(Scenario = "Common-sample near eye"),
  paired_chest |>
    filter(
      .data$factor_id == "leba_f2",
      .data$metric_id %in% near_top$metric_id
    ) |>
    mutate(Scenario = "Common-sample chest"),
  preparation_results |>
    filter(
      .data$factor_id == "leba_f2",
      .data$metric_id %in% near_top$metric_id
    ) |>
    mutate(Scenario = "Gap-timing-unaware dataset")
) |>
  left_join(
    metric_registry |> select(.data$metric_id, .data$display_unit),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  rowwise() |>
  mutate(
    `Effect per LEBA SD (95% CI)` = format_effect(
      .data$estimate_practical_per_sd,
      .data$conf_low_practical_per_sd,
      .data$conf_high_practical_per_sd,
      .data$effect_type,
      .data$display_unit
    ),
    `Fitted sample` = paste0(
      .data$participants, " participants; ",
      if_else(
        is.na(.data$participant_days),
        paste0(.data$observations, " participant rows"),
        paste0(.data$participant_days, " participant-days")
      ),
      "; ", .data$sites, " sites"
    )
  ) |>
  ungroup() |>
  select(
    .data$Scenario,
    Metric = .data$manuscript_name,
    .data$`Fitted sample`,
    .data$`Effect per LEBA SD (95% CI)`
  )

random_summary <- random_site |>
  count(.data$placement, .data$random_site_status, name = "Models") |>
  mutate(
    Placement = recode(.data$placement, glasses = "Near eye", chest = "Chest"),
    Status = recode(
      .data$random_site_status,
      DESCRIPTIVE_PASS = "Fit passed",
      DESCRIPTIVE_UNSTABLE = "Fit unstable"
    )
  ) |>
  select(.data$Placement, .data$Status, .data$Models)

loo_summary <- leave_one_site_out |>
  count(.data$stability_class, name = "Associations") |>
  mutate(
    Classification = recode(
      .data$stability_class,
      stable = "Stable",
      direction_stable_magnitude_sensitive = "Direction stable; magnitude sensitive",
      direction_unstable = "Direction changed after omitting at least one site"
    )
  ) |>
  select(.data$Classification, .data$Associations)

descriptive_top <- descriptive_spearman |>
  filter(
    .data$run_id == near_id,
    .data$factor_id == "leba_f2",
    .data$metric_id %in% near_top$metric_id
  ) |>
  transmute(
    Metric = .data$manuscript_name,
    `Spearman rho (95% CI)` = paste0(
      format_number(.data$spearman_rho, 3), " (",
      format_number(.data$conf_low, 3), "–",
      format_number(.data$conf_high, 3), ")"
    ),
    Participants = .data$pairs,
    Sites = .data$sites,
    `Represented days` = .data$represented_days
  )

stopifnot(
  nrow(near_results) == 68L,
  nrow(chest_results) == 68L,
  nrow(near_samples) == 17L,
  nrow(chest_samples) == 17L,
  all(family_audit$registry_rows == 68L),
  all(family_audit$vector_bh_verified),
  sum(near_results$p_adjusted <= 0.05) == 0L,
  sum(chest_results$p_adjusted <= 0.05) == 0L,
  sum(
    near_results$reader_inference_status == "unfit_for_inference"
  ) == 4L,
  sum(
    chest_results$reader_inference_status == "unfit_for_inference"
  ) == 4L,
  nrow(paired_display) == 68L,
  all(paired_display$exact_sample_match),
  sum(near_results$model_adequacy == "not_acceptable") == 0L,
  sum(chest_results$model_adequacy == "not_acceptable") == 0L,
  nrow(exact_period) == 4L,
  all(exact_period$sign_concordant),
  sum(random_site$random_site_status == "DESCRIPTIVE_UNSTABLE") == 15L,
  sum(leave_one_site_out$successful_refits) == 612L,
  nrow(mder_scenarios) == 32L,
  nrow(mder_upper_tail) == 8L,
  nrow(mder_influence) == 44L,
  all(mder_influence$refit_status == "PASS"),
  sum(mder_influence$sensitivity_interval_contains_zero) == 40L,
  all(
    mder_influence$sensitivity_interval_contains_zero[
      mder_influence$run_id == chest_id
    ]
  ),
  nrow(mder_gap_influence) == 44L,
  all(mder_gap_influence$refit_status == "PASS"),
  sum(mder_gap_influence$sensitivity_interval_contains_zero) == 40L,
  nrow(mder_gap_paired) == 4L,
  all(mder_gap_paired$exact_sample_match),
  all(mder_gap_paired$sign_concordant),
  all(mder_gap_paired$component_intervals_overlap),
  nrow(mder_gap_reconciliation) == 16L,
  all(mder_gap_reconciliation$invariant_verified),
  all(
    mder_scenarios$participant_days[
      mder_scenarios$run_id == near_id
    ] == 702L
  ),
  all(
    mder_scenarios$participants[
      mder_scenarios$run_id == near_id
    ] == 137L
  ),
  all(
    mder_scenarios$participant_days[
      mder_scenarios$run_id == chest_id
    ] == 732L
  ),
  all(
    mder_scenarios$participants[
      mder_scenarios$run_id == chest_id
    ] == 152L
  ),
  all(
    paired$sign_concordant[paired$metric_id == mder_id]
  ),
  all(
    preparation_sensitivity$sign_concordant[
      preparation_sensitivity$metric_id == mder_id
    ]
  ),
  all(
    mder_scenarios$participant_days[
      mder_scenarios$run_id == preparation_id
    ] == 687L
  ),
  all(
    mder_scenarios$participants[
      mder_scenarios$run_id == preparation_id
    ] == 137L
  ),
  nrow(l10_frame_audit) == 4L,
  all(l10_frame_audit$non_value_fields_identical),
  all(l10_frame_audit$changed_keys_verified),
  nrow(l10_input_cells) == 16L,
  nrow(distinct(
    l10_input_cells,
    .data$position,
    .data$participant_key,
    .data$local_date
  )) == 8L,
  sum(l10_input_cells$scenario == "all_available") == 8L,
  sum(l10_input_cells$position == "glasses") == 6L,
  sum(l10_input_cells$position == "chest") == 10L,
  all(l10_input_cells$new_value_lx == 0),
  all(l10_input_cells$current_value_lx == 0),
  nrow(l10_result_change) == 16L,
  max(abs(l10_result_change$estimate_change_per_point)) < 4e-12,
  !any(l10_result_change$bh_retained_after, na.rm = TRUE),
  nrow(l10_bh_audit) == 204L,
  sum(l10_bh_audit$raw_p_changed) == 7L,
  sum(l10_bh_audit$adjusted_p_changed) == 2L,
  sum(l10_bh_audit$family_rank_changed) == 0L,
  nrow(l10_reconciliation) == 22L,
  all(l10_reconciliation$invariant_verified)
)
```

## Question

The Light Exposure Behaviour Assessment (LEBA) asks about habitual
light-related behaviours. The preregistered hypothesis was:

> “H5: LEBA questionnaire factors correlate with selected personal light exposure metrics.”

This analysis tests whether four LEBA factor scores
are associated with 17 personal light-exposure metrics after accounting for
study site. Higher scores mean that the behaviour described by a factor was
reported more often.

::: {.callout-note title="Answer in brief"}
None of the 68 primary near-eye associations remained statistically supported
after false-discovery-rate (FDR) adjustment across the complete family, and
the four planned sleep-environment cells are unfit for inference.
The leading positive near-eye estimates remained descriptive, the
complementary chest analysis reached the same family-wide conclusion, and the
planned sensitivity analyses did not change that interpretation. Important
model-check and site-influence limitations are summarized below.
:::

## Orientation

The four scores describe spending time outdoors (F2), devices in bed (F3),
ambient light before bedtime (F4), and morning/daytime light use (F5).
F2 includes one reverse-coded item; F3--F5 are sums of their ordered items.
Every factor score was complete. Within each exact model sample, the score was
centred and expressed both per raw point and per participant-level standard
deviation. The report uses one participant-level SD so estimates can be
compared across the four differently ranged factor scores.

Near-eye measurements are primary because they more closely represent light
near the eyes. All available chest measurements are analysed separately as
complementary evidence and are not interpreted as ocular exposure. A
**participant-day** is one eligible local calendar day from one participant;
daily metrics can therefore contribute repeated days from the same person. A
common-sample analysis fits the near-eye and chest models to the same
participants and participant-days at both sensor positions, so separately
fitted placement estimates are not confused with differences in who
contributed data. This is neither an equivalence test nor a direct test of a
placement effect, and the placements are never pooled as independent
observations.

The 95% confidence interval (CI) describes uncertainty around each reported
effect. False-discovery-rate (FDR) adjustment limits the expected proportion
of false discoveries across the complete planned family. The main models treat
study site as a fixed effect, meaning that each study site receives its own
adjustment. A random effect instead represents variation attributed to a
grouping variable, such as participant or study site, rather than estimating a
separate coefficient for each group. A sensitivity analysis repeats an
accepted comparison under a predefined alternative to assess whether the
interpretation is stable.

## Primary near-eye result

### Complete association pattern

```{r}
#| label: fig-h05-near-effects
#| fig-cap: "Near-eye effects per one participant-level SD of each LEBA factor. None of the 68 tests retained FDR-adjusted support. The four grey sleep-environment cells are labelled Unfit, and their estimates are suppressed."
#| fig-alt: "Heatmap with 17 light-exposure metrics as rows and LEBA factors F2 to F5 as columns. Cells show practical-scale effects per one participant-level standard deviation, except four grey sleep-environment cells labelled Unfit. None of the 68 tests retained false-discovery-rate-adjusted support."
#| out-width: "96%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_near_eye_effects.png"
))
```

The heatmap colours show the signed effect on each model's fitted scale, while
the printed value is on the practical response scale. Colours therefore show
direction and relative within-outcome magnitude, not a common physical unit
across all 17 rows. Tables @tbl-h05-near-results-a and
@tbl-h05-near-results-b form one logical continued table and provide every
estimate, 95% CI, raw *p*-value, and FDR-adjusted *p*-value.

```{r}
#| label: tbl-h05-near-results-a
#| tbl-cap: "Primary near-eye associations across all 68 factor–metric cells."

result_matrix(
  near_results,
  1:10,
  source_note_lead = paste(
    "Associations are reported per one participant-level SD of the LEBA",
    "score with"
  )
)
```

```{r}
#| label: tbl-h05-near-results-b
#| tbl-cap: "Primary near-eye associations (continued)."

result_matrix(
  near_results,
  11:17,
  source_note_lead = paste(
    "Associations are reported per one participant-level SD of the LEBA",
    "score with"
  )
)
```

The primary
near-eye sensor position measures light close to the eyes. For a score one
standard deviation (SD) higher on the factor describing spending time outdoors,
the two largest positive estimates were ratios of
`r format_number(near_duration$estimate_practical_per_sd)` for time above
1,000 lx melanopic equivalent daylight illuminance (melEDI); its 95%
confidence interval (95% CI) was
`r format_number(near_duration$conf_low_practical_per_sd)`–`r format_number(near_duration$conf_high_practical_per_sd)`.
The corresponding ratio for melEDI dose was
`r format_number(near_dose$estimate_practical_per_sd)` (95% CI
`r format_number(near_dose$conf_low_practical_per_sd)`–`r format_number(near_dose$conf_high_practical_per_sd)`)
and both had FDR-adjusted *p* =
`r format_p(near_duration$p_adjusted)` and therefore remain descriptive rather
than multiplicity-retained.

Across the complete planned family, zero of 68 FDR-adjusted *p*-values were at or
below 0.050. The four sleep-environment models remain in that complete family
so the correction is not changed after seeing the model checks, but they are
unfit for inference and their estimates and p-values are suppressed from the
reader-facing matrices.
The two smallest raw values were both for F2, spending time outdoors:

- time above 1,000 lx melEDI increased by a factor of
  `r format_number(near_duration$estimate_practical_per_sd)` per participant
  SD (95% CI `r format_number(near_duration$conf_low_practical_per_sd)`–`r format_number(near_duration$conf_high_practical_per_sd)`),
  raw *p* = `r format_p(near_duration$p_raw)`, FDR-adjusted
  *p* = `r format_p(near_duration$p_adjusted)`; and
- melEDI dose increased by a factor of
  `r format_number(near_dose$estimate_practical_per_sd)` (95% CI
  `r format_number(near_dose$conf_low_practical_per_sd)`–`r format_number(near_dose$conf_high_practical_per_sd)`),
  raw *p* = `r format_p(near_dose$p_raw)`, FDR-adjusted
  *p* = `r format_p(near_dose$p_adjusted)`.

These estimates remain useful for describing magnitude and uncertainty, but
neither is a multiplicity-retained result.

Complete near-eye numerical results and figure data are available as
[result source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_results.csv)
and [figure source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_effect_figure_data.csv).

## Complementary chest evidence

The chest analysis repeats the same factor scores, metric definitions,
response models, fixed-site adjustment, effect scaling, and 68-test
correction. Depending on the metric, it used 152--154 participants, 732--902
participant-days for daily outcomes, and eight sites. IS and IV used 153
participant rows and represented 900 observed days.

The complementary chest sensor position, which is
not a measure of ocular exposure, likewise retained zero of 68 associations
but reproduced the positive direction of both leading estimates.

The chest measurements remain complementary, non-ocular evidence and are not
interpreted as light reaching the retina.
<details>
<summary>Show complementary chest results and model checks</summary>

```{r}
#| label: fig-h05-chest-effects
#| fig-cap: "Chest effects per one participant-level SD of each LEBA factor. None of the 68 tests retained FDR-adjusted support. The four grey sleep-environment cells are labelled Unfit, and their estimates are suppressed."
#| fig-alt: "Heatmap with 17 chest light-exposure metrics as rows and LEBA factors F2 to F5 as columns. Cells show practical-scale effects per one participant-level standard deviation, except four grey sleep-environment cells labelled Unfit. None of the 68 tests retained false-discovery-rate-adjusted support."
#| out-width: "96%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_chest_effects.png"
))
```

```{r}
#| label: tbl-h05-chest-results-a
#| tbl-cap: "Chest associations for stability, light level, duration, and continuous-period metrics."

result_matrix(chest_results, 1:10)
```

```{r}
#| label: tbl-h05-chest-results-b
#| tbl-cap: "Chest associations for timing, exposure-history, and spectrum metrics."

result_matrix(chest_results, 11:17)
```

Zero of 68 planned chest associations remained after its separate correction;
the smallest value was FDR-adjusted *p* =
`r format_p(min(chest_results$p_adjusted))`. The F2 estimates for time above
1,000 lx melEDI and melEDI dose were again positive, with ratios
`r format_number(chest_duration$estimate_practical_per_sd)` and
`r format_number(chest_dose$estimate_practical_per_sd)`, respectively, but
neither was multiplicity-retained.

Complete chest numerical results and figure data are available as
[result source data](../../artifacts/11_source_data/H05/H05_reader_chest_results.csv)
and [figure source data](../../artifacts/11_source_data/H05/H05_reader_chest_effect_figure_data.csv).

### Chest model checks

```{r}
#| label: fig-h05-chest-adequacy
#| fig-cap: "Chest model-check classifications for all 68 associations."
#| fig-alt: "Heatmap of model-check classifications for 17 chest metrics and four LEBA factors. Fourteen cells are green for acceptable, 50 are orange for acceptable with specified limitations, and the four sleep-environment cells are red for unfit for inference."
#| out-width: "94%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_chest_adequacy.png"
))
```

```{r}
#| label: tbl-h05-chest-adequacy-counts
#| tbl-cap: "Chest model-adequacy counts."

adequacy_count_table(chest_results)
```

```{r}
#| label: tbl-h05-chest-limitations
#| tbl-cap: "Specified limitations identified by the chest model checks."

limitation_count_table(chest_results)
```

Fourteen chest models were acceptable and 50 were acceptable with specified
limitations. The four chest sleep-environment models were **unfit for
inference**. They show the same material problem as near eye, with 241--244
predictions above participant-specific support and strong simulated-residual
disagreement; their estimates and p-values are suppressed from the
reader-facing matrices.

```{r}
#| label: tbl-h05-chest-sleep-diagnostics
#| tbl-cap: "Model-check evidence for classifying the chest sleep-environment models as unfit for inference."

sleep_diagnostic_table(chest_id)
```

The 64 reader-interpretable chest cells provide complementary context for the
zero-retained result, subject to their residual warnings. The four additional
planned sleep-environment cells remain in the complete family but are unfit
for inference.

The chest model-check heatmap has
[paired source data](../../artifacts/11_source_data/H05/H05_reader_chest_adequacy_figure_data.csv).

</details>

## Model checks and the unfit sleep-environment boundary

Most near-eye and complementary chest models were acceptable or acceptable
with specified limitations. The four sleep-environment cells at each placement
are **unfit for H05 inference** because their simulated residuals and fitted
response support show material failures. Their estimates, confidence
intervals, and p-values are suppressed. This judgment applies only to the H05
response and model structure. It does not rule out analysing the underlying
metric in another hypothesis with a different response variable, estimand, or
model structure.

A Tweedie model is a positive-response regression model that can accommodate
right-skewed outcomes and exact zeros. For the sleep-environment outcome, that
working model did not provide a defensible inferential representation.

<details>
<summary>Show detailed near-eye model checks</summary>

```{r}
#| label: fig-h05-near-adequacy
#| fig-cap: "Near-eye model-check classifications for all 68 associations."
#| fig-alt: "Heatmap of model-check classifications for 17 near-eye metrics and four LEBA factors. Nineteen cells are green for acceptable, 45 are orange for acceptable with specified limitations, and the four sleep-environment cells are red for unfit for inference."
#| out-width: "94%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05", "H05_reader_near_eye_adequacy.png"
))
```

```{r}
#| label: tbl-h05-near-adequacy-counts
#| tbl-cap: "Near-eye model-adequacy counts."

adequacy_count_table(near_results)
```

```{r}
#| label: tbl-h05-near-limitations
#| tbl-cap: "Specified limitations identified by the near-eye model checks."

limitation_count_table(near_results)
```

The model checks (model diagnostics) covered numerical fitting, residual
behaviour, response support, and influence. All fixed-site fits completed, and
all applicable Hessian checks were positive definite. That numerical
completion does not make every response model
scientifically defensible. Nineteen models were acceptable, 45 were acceptable
with a specified limitation, and the four sleep-environment models were
classified **unfit for inference**. Most retained limitations were ordinary
Gaussian residual warnings. They qualify how closely the working residual
distribution represents these data but do not indicate numerical fit failure.

The following representative screens show F2 for the two leading estimates,
the accepted MDER outcome, and the sleep-environment outcome. The smooth line in the
residual-versus-fitted plot helps reveal fitted-dependent residual structure; the quantile
plot compares standardized residuals with a normal reference. For the
Tweedie sleep model, the simulation-based tests in the subsequent table,
rather than the normal-reference plot, govern the assessment.

```{r}
#| label: fig-h05-near-residual-fitted
#| fig-cap: "Selected near-eye standardized residuals against fitted values."
#| fig-alt: "Four vertically stacked residual-versus-fitted plots for F2 models of time above 1,000 lx melEDI, melEDI dose, MDER, and time below 1 lx melEDI during sleep. Points show standardized Pearson residuals and blue smooths show their average pattern."
#| out-width: "88%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05",
  "H05_reader_near_eye_residual_fitted.png"
))
```

```{r}
#| label: fig-h05-near-residual-qq
#| fig-cap: "Selected near-eye residual quantile checks."
#| fig-alt: "Four vertically stacked normal-reference quantile plots for the same F2 models, including MDER. The Gaussian melEDI-dose model has tail departures, while simulation-based model checks and response-support failures make the Tweedie sleep model unfit for inference."
#| out-width: "88%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05",
  "H05_reader_near_eye_residual_qq.png"
))
```

#### H05 sleep-environment models are unfit for inference

The four models for time below 1 lx melEDI during sleep require a stronger
qualification than the ordinary residual warnings. Their simulated residuals
showed pronounced non-uniformity, zero-mass disagreement, and outlier
evidence. They also produced 183--187 fitted values above the
participant-specific response support. These models are retained because the
prespecified response-family assessment did not identify an adequate common
replacement. They remain members of the complete planned family but are
classified **unfit for inference**: their estimates, confidence intervals,
and p-values are not used to support any scientific claim.

This classification is specific to this H05 sleep-environment response and
the model structure used here. It does not determine whether the underlying
metric can be analysed in another hypothesis that uses a different response
variable or model structure.

```{r}
#| label: tbl-h05-near-sleep-diagnostics
#| tbl-cap: "Model-check evidence for classifying the near-eye sleep-environment models as unfit for inference."

sleep_diagnostic_table(near_id)
```

The conclusion that zero associations retained FDR-adjusted support is based
on 64 reader-interpretable near-eye cells: 19 are acceptable and 45 have
specified residual limitations. The four additional planned
sleep-environment cells remain in the complete multiplicity family but are
unfit for inference.

Exact source data for the selected model-check figures are available
[here](../../artifacts/11_source_data/H05/H05_reader_near_eye_selected_diagnostics.csv).
The model-check heatmap has its own
[paired source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_adequacy_figure_data.csv).

</details>

## MDER results and upper-tail checks

A predefined sensitivity uses the **gap-timing-unaware dataset**. It applies
the same general 50%-per-hour and 80%-per-day coverage rules, but does not use
the timing of remaining missing observations for metric-specific adjustment.
For this contrast only, the approved primary preparation could be interpreted
as a **time-sensitive primary metric dataset**. Below, it is called simply the
primary dataset.

MDER is the arithmetic mean of viable one-minute melEDI/illuminance ratios.
Both channels must be finite and strictly positive, and at least 720 of the
complete 1,440 local wall-clock minutes must provide a viable ratio. Repeated
local minutes on a daylight-saving fall-back day are averaged channel by
channel before the ratio is formed; absent spring-forward minutes remain
missing. Each viable minute ratio has equal weight, so MDER is not a ratio of
daily integrals. Failure of this support rule makes only MDER missing.

The gap-timing-unaware dataset left the family-wide conclusion unchanged; its
exact 478-day MDER placement comparison retained all four directions with
overlapping component intervals.

No MDER association retained FDR-supported evidence. The small complementary
chest coefficients were more sensitive to upper-tail observations, but their
sensitivity intervals still crossed zero.
<details>
<summary>Show MDER details and upper-tail checks</summary>

MDER was available for 702 participant-days from 137 participants at all nine
near-eye sites and 732 participant-days from 152 participants at eight chest
sites in the primary dataset. In the gap-timing-unaware dataset, it was
available for 687 near-eye days from 137 participants and 723 chest days from
152 participants. The exact paired/common gap comparison used 478 days from
107 participants at eight sites. The tables below give all four MDER
associations in every accepted all-available and common-sample analysis. The
all-available primary near-eye, primary chest, and gap-timing-unaware near-eye
analyses retain their own complete 68-test families; common-sample rows are
descriptive placement sensitivities.

```{r}
#| label: tbl-h05-mder-primary-scenarios
#| tbl-cap: "MDER estimates, 95% confidence intervals, and exact fitted samples in the primary dataset."

mder_scenario_table("main")
```

```{r}
#| label: tbl-h05-mder-gap-scenarios
#| tbl-cap: "MDER estimates, 95% confidence intervals, and exact fitted samples in the gap-timing-unaware dataset."

mder_scenario_table("manuscript_prepared_data")
```

None of the MDER associations was retained after its applicable 68-test FDR
correction. In the primary near-eye analysis, the four differences per LEBA SD
ranged from -0.020 to +0.004 MDER units. The F5 interval was below zero before
multiplicity adjustment (Raw *p* = `r format_p(mder_primary_f5$p_raw)`,
FDR-adjusted *p* = `r format_p(mder_primary_f5$p_adjusted)`) and is not a
multiplicity-retained association. All four paired/common-sample MDER
directions agreed between near eye and chest and their component 95% confidence
intervals overlapped. All four gap-timing-unaware near-eye directions also
agreed with the primary near-eye analysis, again with overlapping component
intervals. In the gap-timing-unaware paired/common sample, all four near-eye and
chest directions likewise agreed and all component intervals overlapped; this
is concordance, not an equivalence test.

Because a mean of momentary ratios can be influenced by small positive
photopic denominators, the accepted MDER upper tail was screened without
automatically excluding observations. One near-eye day and three chest days
were above the Tukey outer fence in each all-available dataset. The maximum was
1.857 near eye in both datasets, 3.574 at the primary chest, and 3.645 at the
gap-timing-unaware chest.

```{r}
#| label: tbl-h05-mder-upper-tail
#| tbl-cap: "Accepted MDER upper-tail screen in the all-available analyses."

mder_upper_tail |>
  filter(.data$sample_scenario == "all_available") |>
  transmute(
    Dataset = if_else(
      .data$data_scenario_id == "main",
      "Primary dataset",
      "Gap-timing-unaware dataset"
    ),
    Placement = recode(
      .data$placement,
      glasses = "Near eye",
      chest = "Chest"
    ),
    Observations = .data$observations,
    Participants = .data$participants,
    Sites = .data$sites,
    Median = .data$median,
    `Outer fence` = .data$tukey_outer_fence,
    Maximum = .data$maximum,
    `Days above outer fence` = .data$outer_tail_days
  ) |>
  gt::gt(groupname_col = "Dataset", rowname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Observations, Participants, Sites, `Days above outer fence`)
  ) |>
  gt::fmt_number(columns = c(Median, `Outer fence`, Maximum), decimals = 3) |>
  gt::tab_source_note(
    source_note = paste(
      "The Tukey Q3 + 3 × IQR rule is an influence-screening rule, not an",
      "automatic exclusion rule; all screened observations remain in their fitted analyses."
    )
  ) |>
  h05_gt(12)
```

```{r}
#| label: tbl-h05-mder-influence
#| tbl-cap: "Accepted MDER upper-tail day and participant influence refits."

mder_influence_summary |>
  transmute(
    .data$Analysis,
    .data$Factor,
    .data$`Candidate refits`,
    .data$`Maximum absolute change (full-model SE)`,
    `Any sign reversal` = if_else(.data$`Any sign reversal`, "Yes", "No"),
    `All sensitivity intervals include zero` = if_else(
      .data$`All sensitivity intervals include zero`, "Yes", "No"
    )
  ) |>
  gt::gt(groupname_col = "Analysis", rowname_col = "Factor") |>
  gt::fmt_integer(columns = `Candidate refits`) |>
  gt::fmt_number(
    columns = `Maximum absolute change (full-model SE)`,
    decimals = 3
  ) |>
  h05_gt(12)
```

All 44 primary-dataset bounded day- and participant-deletion refits passed; 40
of their 44 sensitivity intervals included zero. The four near-eye coefficients changed
by at most 0.234 of their full-model
standard error and never reversed direction. The four near-eye F5 deletion
intervals remained below zero; the other near-eye sensitivity intervals
included zero. At the chest, the tiny F2
coefficient reversed direction in one deletion refit, and the maximum changes
for F2 and F4 were 0.823 and 0.828 standard errors. These small complementary
chest coefficients are therefore upper-tail-sensitive, but all chest
sensitivity intervals included zero. No observation was removed from the
primary result, and none of these checks changes the MDER conclusion after
multiplicity correction.

The gap-timing-unaware branch also had 44 successful bounded refits;
40 intervals included zero. Its four near-eye coefficients changed by at most
0.233 full-model standard errors without reversing direction. Its small chest
F2 and F4 coefficients changed by at most 0.823 and 0.841 standard errors,
respectively; F2 reversed direction in two deletions, and every chest interval
included zero. The gap F5 near-eye deletion intervals remained below zero,
but the corresponding near-eye sensitivity result was not retained after FDR
adjustment.

Exact screening and refit records are available in the
[upper-tail summary](../../artifacts/08_diagnostics/H05/H05_mder_metric010_upper_tail_summary.csv)
and [influence source data](../../artifacts/08_diagnostics/H05/H05_mder_metric010_influence_refits.csv).
The gap-timing-unaware sensitivity has separate [influence data](../../artifacts/08_diagnostics/H05/H05_mder_metric010_gap_influence_refits.csv) and an [exact paired/common MDER comparison](../../artifacts/09_tables/H05/H05_mder_metric010_gap_paired_placement_comparison.csv).

</details>

## Sensitivity analyses

The common-sample placement comparison showed broad directional concordance
without testing equivalence or a direct placement effect. Fixed study-site
adjustment remained the primary specification because the registered
random-site alternative included unstable fits. Leave-one-site-out,
gap-timing-unaware, and exact-period analyses supported the same restrained
family-wide interpretation while showing that some small estimates were
site-sensitive.

<details>
<summary>Show detailed sensitivity results</summary>

### Common near-eye and chest sample

Every plotted point compares the same metric, factor, response definition,
model, transformation, and exact sample at both placements. Participant-day
metrics used 107--112 participants and 489--643 matched participant-days at
eight sites; IS and IV used the same 112 participant rows at eight sites.
Fifty-nine of 68 associations had the same direction at both placements and
overlapping component intervals; nine differed in direction but still had
overlapping component intervals.
Because no participant-cluster interval was calculated for the direct
chest-minus-near-eye difference, these are component comparisons rather than formal
tests of placement differences.

```{r}
#| label: fig-h05-paired-placement
#| fig-cap: "Exact common-sample model-scale effects, with near eye on x and chest on y. The dashed identity line describes concordance, grey lines mark the null, and equal x/y geometry preserves the 45-degree comparison. Samples contain 107–112 participants, 489–643 matched participant-days, and eight sites; IS and IV use 112 participant rows. Component 95% confidence intervals are in the adjacent tables. This is not an equivalence analysis."
#| fig-alt: "Four equal-axis scatterplot panels, one per LEBA factor, compare 17 exactly matched estimands with near-eye effects on the horizontal axis and chest effects on the vertical axis. A dashed 45-degree identity line and grey horizontal and vertical null lines are visible. Samples contain 107 to 112 participants, 489 to 643 matched participant-days, and eight sites; IS and IV use 112 participant rows. Most points have the same sign, while nine cross an axis; proximity to identity does not establish equivalence."
#| out-width: "90%"
#| fig-align: center

knitr::include_graphics(file.path(
  root, "artifacts", "10_figures", "H05",
  "H05_reader_paired_placement_effects.png"
))
```

The component estimates and 95% confidence intervals corresponding to every
plotted point are shown below.

::: {.panel-tabset}

### F2: spending time outdoors

```{r}
#| label: tbl-h05-paired-f2
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F2."

paired_component_table("leba_f2")
```

### F3: devices in bed

```{r}
#| label: tbl-h05-paired-f3
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F3."

paired_component_table("leba_f3")
```

### F4: ambient light before bedtime

```{r}
#| label: tbl-h05-paired-f4
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F4."

paired_component_table("leba_f4")
```

### F5: morning/daytime light use

```{r}
#| label: tbl-h05-paired-f5
#| tbl-cap: "Exact paired/common near-eye and chest component estimates for F5."

paired_component_table("leba_f5")
```

:::

Paired figure source data are available
[here](../../artifacts/11_source_data/H05/H05_paired_effect_comparison_data.csv).

### Stability of the two leading F2 estimates

```{r}
#| label: tbl-h05-leading-sensitivities
#| tbl-cap: "The two leading F2 estimates across placement, sample, and preparation sensitivities."

leading_sensitivity |>
  gt::gt(rowname_col = "Scenario", groupname_col = "Metric") |>
  gt::cols_width(
    `Fitted sample` ~ gt::pct(32),
    `Effect per LEBA SD (95% CI)` ~ gt::pct(30)
  ) |>
  h05_gt(12)
```

Both leading F2 directions remain positive on the common near-eye and chest
sample and in the gap-timing-unaware dataset. These checks support the
stability of the descriptive pattern; they do not create a new significance
screen.

As a direct participant-level description, the Spearman coefficients were:

```{r}
#| label: tbl-h05-descriptive-spearman
#| tbl-cap: "Descriptive participant-level F2 rank correlations for the two leading metrics."

descriptive_top |>
  gt::gt(rowname_col = "Metric") |>
  gt::fmt_integer(columns = c(Participants, Sites, `Represented days`)) |>
  gt::tab_source_note(
    source_note = paste(
      "No p-values or significance screen are attached to these descriptive",
      "correlations. Intervals use participant resampling."
    )
  ) |>
  h05_gt(12)
```

### Site adjustment and site influence

Fixed site was selected as primary because the inherited 17-metric
site-structure evaluation fitted all fixed-site models successfully, whereas
the corresponding random-site descriptions included two unstable, two
non-estimable, and two singular fits. In the H05 sensitivity itself, 15 of 136
main all-available random-site fits were unstable: seven near eye and eight
chest, while 121 passed. This instability is why random site is a sensitivity
rather than the primary adjustment.

```{r}
#| label: tbl-h05-random-site-summary
#| tbl-cap: "Random-site sensitivity fit status across near-eye and chest associations."

random_summary |>
  gt::gt(rowname_col = "Status", groupname_col = "Placement") |>
  gt::fmt_integer(columns = Models) |>
  h05_gt()
```

All 612 required leave-one-site-out refits completed. Twenty associations
were stable, 27 retained direction but were magnitude-sensitive, and 21
changed direction after at least one site was omitted. The two leading F2
associations had no site-omission sign reversal and were classified stable.
The broader pattern confirms that some small, near-zero estimates depend on
which of the nine sites is represented, reinforcing the restrained
family-wide interpretation.

```{r}
#| label: tbl-h05-leave-site-summary
#| tbl-cap: "Primary leave-one-site-out stability classifications."

loo_summary |>
  gt::gt(rowname_col = "Classification") |>
  gt::fmt_integer(columns = Associations) |>
  h05_gt()
```

### Gap-timing-unaware dataset

The same model implementation was applied to the gap-timing-unaware dataset,
including the same mean-of-viable-momentary-ratios MDER definition. This
sensitivity gave no
multiplicity-retained association; its smallest value was FDR-adjusted *p* =
`r format_p(min(preparation_results$p_adjusted, na.rm = TRUE))`. Sixty-four of
68 estimates retained their sign and had overlapping component intervals;
four changed sign, also with overlapping component intervals. The conclusion
that zero associations retained FDR-adjusted support is unchanged. For MDER
specifically, 687 near-eye days from 137
participants entered the near-eye sensitivity family, and its smallest MDER
value was FDR-adjusted *p* =
`r format_p(mder_gap_min_adjusted)`. The matched placement sensitivity used
478 days from 107 participants at
eight sites; all four directions agreed and component intervals overlapped.

### Exactly identified longest period

The all-available longest continuous period above 250 lx melEDI is an
observed lower bound because a data gap can interrupt identifiability. The
strict sensitivity retained only participant-days whose longest period was
exactly identified: 500 participant-days from 132 participants at all nine
sites. All four factor directions agreed with the all-available analysis.

```{r}
#| label: tbl-h05-exact-period
#| tbl-cap: "Exactly identified longest-period sensitivity."

exact_period |>
  transmute(
    Factor = factor_code(.data$factor_id),
    `All-available ratio` = format_number(
      .data$all_available_estimate_practical_per_sd, 3
    ),
    `Exactly identified ratio (95% CI)` = paste0(
      format_number(.data$estimate_practical_per_sd, 3), " (",
      format_number(.data$conf_low_practical_per_sd, 3), "–",
      format_number(.data$conf_high_practical_per_sd, 3), ")"
    ),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Direction retained` = if_else(.data$sign_concordant, "Yes", "No")
  ) |>
  gt::gt(rowname_col = "Factor") |>
  gt::fmt_integer(columns = c(Participants, `Participant-days`, Sites)) |>
  h05_gt(12)
```

For F2, the ratio attenuated from 1.117 to 1.046 (95% CI 0.930--1.175),
supporting the decision not to promote this lower-bound metric as a positive
association.

Complete sensitivity sources are available for the
[common-sample placement comparison](../../artifacts/09_tables/H05/H05_paired_placement_comparison.csv),
[random-site models](../../artifacts/09_tables/H05/H05_random_site_sensitivity.csv),
[leave-one-site-out analysis](../../artifacts/09_tables/H05/H05_leave_one_site_out_summary.csv),
[gap-timing-unaware dataset](../../artifacts/11_source_data/H05/H05_gap_timing_unaware_dataset.csv),
[gap-timing-unaware matched MDER placement comparison](../../artifacts/09_tables/H05/H05_mder_metric010_gap_paired_placement_comparison.csv),
and [exactly identified longest-period analysis](../../artifacts/09_tables/H05/H05_exactly_identified_longest_bout_sensitivity.csv).

</details>

## Interpretation and limitations

The strongest defensible conclusion is that, after fixed-site adjustment and
one complete 68-test FDR adjustment, these data provide no
multiplicity-retained evidence that any of the four LEBA factors is associated
with the 17 selected personal light-exposure metrics.

The positive F2 estimates for time above 1,000 lx melEDI and melEDI dose are
consistent across the principal stability checks. Their confidence intervals
describe non-zero associations before multiplicity adjustment, but their
adjusted values do not support confirmatory claims. They should be treated as
descriptive patterns that may motivate future work, not as established
effects of spending time outdoors.

This is an observational association analysis. LEBA scores are self-reported,
site adjustment cannot remove all person-level confounding, only nine sites
contribute to the primary analysis, and several small estimates are sensitive
to site omission. Near-eye and chest measurements also represent related but
non-identical light fields. Finally, the sleep-environment Tweedie models are
unfit for inference because of material response-support and
simulated-residual failures; their estimates and p-values are suppressed from
the reader-facing results and cannot support a claim in H05. This judgment is
specific to the response variable and model structure used here and does not
preclude analysing the metric in another hypothesis with a different response
or model.

Uncertainty remains: several small estimates
changed direction when a site was omitted, the small complementary chest MDER
coefficients were sensitive to upper-tail observations but retained
zero-crossing sensitivity intervals, and the four planned sleep-environment
models are unfit for H05 inference under this response definition and model
structure.

## Detailed analysis record

This record preserves the factor construction, exact samples, model formulas,
response specifications, registration links, source links, and technical
provenance used to produce the reader-facing results above.

### Questionnaire factors and score scale

The factor table records the four behavioural scores and their construction.

```{r}
#| label: tbl-h05-factors
#| tbl-cap: "The four analysed LEBA factors and their score interpretation."

factor_registry |>
  transmute(
    Factor = factor_code(.data$factor_id),
    Behaviour = .data$factor_label,
    `Possible score` = paste0(.data$possible_min, "–", .data$possible_max),
    Interpretation = .data$direction_text
  ) |>
  gt::gt(rowname_col = "Factor") |>
  gt::cols_width(
    Behaviour ~ gt::pct(27),
    `Possible score` ~ gt::pct(14),
    Interpretation ~ gt::pct(48)
  ) |>
  h05_gt()
```

### Exact analytical samples

Depending on the metric, the primary models used 139--141 participants,
655--816 participant-days for participant-day outcomes, and all nine sites.
IS and IV used 141 participant rows and represented 816 observed days. Exact
metric-specific samples are shown below.

```{r}
#| label: tbl-h05-near-samples
#| tbl-cap: "Exact fitted near-eye sample for each light-exposure metric."

sample_table(near_samples)
```

The exact sample table is available as
[source data](../../artifacts/11_source_data/H05/H05_reader_near_eye_samples.csv).

```{r}
#| label: tbl-h05-chest-samples
#| tbl-cap: "Exact fitted chest sample for each light-exposure metric."

sample_table(chest_samples)
```

The exact chest sample table is available as
[source data](../../artifacts/11_source_data/H05/H05_reader_chest_samples.csv).

### Models, response specifications, and inference

For participant-day outcomes, the full model asks whether LEBA improves the
site-adjusted description while accounting for repeated days from the same
participant. Its participant random effect represents remaining
between-participant variation after LEBA and site are considered. The reduced
model removes only LEBA. Interdaily stability (IS) and intradaily variability
(IV) have one response per participant, so they do not use a participant
random effect. Fixed site effects use the registered site order and
sum-to-zero contrasts; the LEBA coefficient is therefore adjusted for site
without being tied to one reference site. This adjustment describes site
differences and does not imply that site causes an exposure outcome. The
registered random-site structure is retained as a sensitivity analysis.

The following evaluated cell reads and prints the exact formula objects used
for the full, reduced, and registered random-site sensitivity models.

```{r}
#| label: tbl-h05-formulas
#| tbl-cap: "Exact evaluated formulas supplied to the H05 full, reduced, and random-site sensitivity models."
#| echo: true
#| code-fold: false

registered_formula <- function(analysis_unit, formula_id) {
  row <- formula_registry |>
    dplyr::filter(
      .data$analysis_unit == .env$analysis_unit,
      .data$formula_id == .env$formula_id
    )
  stopifnot(nrow(row) == 1L)
  stats::as.formula(row$formula[[1L]])
}

h05_formulas <- list(
  participant_day_full = registered_formula("participant_day", "fixed_full"),
  participant_day_reduced = registered_formula("participant_day", "fixed_reduced"),
  participant_full = registered_formula("participant", "fixed_full"),
  participant_reduced = registered_formula("participant", "fixed_reduced"),
  participant_day_random_site = registered_formula("participant_day", "random_site"),
  participant_random_site = registered_formula("participant", "random_site")
)

tibble::tibble(
  Model = c(
    "Participant-day primary full",
    "Participant-day primary reduced",
    "Participant primary full",
    "Participant primary reduced",
    "Participant-day random-site sensitivity",
    "Participant random-site sensitivity"
  ),
  `Exact Wilkinson formula` = paste0(
    "`",
    vapply(h05_formulas, deparse1, character(1)),
    "`"
  )
) |>
  gt::gt(rowname_col = "Model") |>
  gt::fmt_markdown(columns = `Exact Wilkinson formula`) |>
  gt::cols_width(`Exact Wilkinson formula` ~ gt::pct(67)) |>
  gt::tab_source_note(
    source_note = paste(
      "The table is evaluated directly from the registered formula objects;",
      "the reduced model removes only LEBA from the corresponding full model."
    )
  ) |>
  h05_gt(12)
```

Gaussian full and reduced models were compared using maximum likelihood.
Tweedie models used a log link and maximum likelihood. Effects are
back-transformed from the fitted scale to a practical response scale: a ratio
for log-scale and Tweedie outcomes, an odds ratio for IS, a difference in the
stated metric unit for identity-scale outcomes, and an hours difference for
clock-time outcomes.  The response specifications are listed below.

```{r}
#| label: tbl-h05-response-specifications
#| tbl-cap: "Metric-specific analysis level, response model, transform, and reported effect scale."

metric_registry |>
  mutate(
    Group = metric_group(.data$metric_order),
    Metric = .data$manuscript_name,
    Level = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    Model = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    Transform = recode(
      .data$response_transform,
      logit = "Logit",
      identity = "None",
      log10_offset_0.1 = "log10(value + 0.1)",
      clock_hours = "Continuous clock hours",
      clock_hours_midnight_after_16 = "Clock hours; strict after-16:00 shift"
    ),
    `Reported effect` = recode(
      .data$effect_scale,
      odds_ratio = "Odds ratio",
      difference = "Difference",
      ratio = "Ratio"
    )
  ) |>
  select(.data$Group, .data$Metric, .data$Level, .data$Model, .data$Transform, .data$`Reported effect`) |>
  gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
  gt::cols_width(
    Level ~ gt::pct(16), Model ~ gt::pct(17),
    Transform ~ gt::pct(27), `Reported effect` ~ gt::pct(16)
  ) |>
  h05_gt(12)
```

For each metric and factor, the raw *p*-value compares the full model with the
corresponding reduced model. The primary multiplicity family contains every
combination of four factors and 17 metrics: 68 tests fixed before inference.
FDR adjustment is applied to that complete vector, and support is retained
only when the adjusted *p*-value is at or below 0.050.

### Metric construction and model-ready preparation

For darkest-10-hour mean melEDI, a geometric-mean back-transformation residual
within the unit-aware floating-point tolerance is treated as exact zero only
when every finite source minute in the selected window is exactly zero;
missing minutes remain missing. Metric definitions are documented in
[Preparation 04](../preparation/04_metric_derivation.qmd), and their
model-ready implementation is documented in
[Preparation 06](../preparation/06_model_ready_datasets.qmd).

## Preregistration and operational context

### Deviations and operational clarifications {#h05-preregistration-deviations}

The preregistration gives the broad model `Metric ~ LEBA + (1|Site)` but does
not fully define the repeated-measure structure, response families, metric
set, multiplicity family, clock handling, or placement roles. The following
choices make those elements explicit.

```{r}
#| label: tbl-h05-model-deviations
#| tbl-cap: "H05 model and inferential changes or operational clarifications."

tibble::tribble(
  ~Aspect, ~Preregistered_or_unspecified, ~Analysis, ~Why_it_matters,
  "Site adjustment",
  "Random site intercept",
  "Country-coded study-site-order fixed site effects; registered random-site model retained as a sensitivity",
  "The fixed-site structure was more reliably estimable with nine sites.",
  "Repeated observations",
  "Not operationally specified",
  "Participant-day responses for 15 metrics with a participant random intercept; participant responses for IS and IV",
  "Daily information is retained without treating days from one person as independent.",
  "Response model",
  "Generic linear model",
  "Metric-specific Gaussian or Tweedie model and approved response transform",
  "The outcome scale and support differ across the 17 metrics.",
  "LEBA scale",
  "Not specified",
  "Centred score; effects reported per one participant-level SD",
  "This retains the continuous score while giving a comparable effect unit.",
  "Inference and multiplicity",
  "Correlation matrices plus linear models; adjustment family not operationalized",
  "Full-versus-reduced LEBA comparison and one complete 68-test FDR family per inferential placement/scenario",
  "All four factors and 17 metrics remain in the correction."
) |>
  gt::gt(rowname_col = "Aspect") |>
  gt::cols_label(
    Preregistered_or_unspecified = "Preregistered or unspecified",
    Analysis = "Analysed",
    Why_it_matters = "Why it matters"
  ) |>
  gt::cols_width(
    Preregistered_or_unspecified ~ gt::pct(24),
    Analysis ~ gt::pct(34),
    Why_it_matters ~ gt::pct(30)
  ) |>
  h05_gt(12)
```

```{r}
#| label: tbl-h05-data-deviations
#| tbl-cap: "H05 metric, clock, placement, and descriptive-output clarifications."

tibble::tribble(
  ~Aspect, ~Preregistered_or_unspecified, ~Analysis, ~Why_it_matters,
  "Metric set",
  "Selected metrics not enumerated",
  paste0(
    "Seventeen prespecified metrics, including corrected dose and MDER as ",
    "the arithmetic mean of viable one-minute melEDI/illuminance ratios; ",
    "both channels must be finite and positive, and at least 720 of the ",
    "complete 1,440 local wall-clock minutes must be viable"
  ),
  "The 17 × 4 cross-product fixes the 68 tests before inference.",
  "Primary placement",
  "Not specified in H5",
  "All-available near eye primary; all-available chest complementary; common-sample placements as sensitivity",
  "Separates the scientific measurement role from sample composition.",
  "Clock outcomes",
  "Not operationally specified",
  "Continuous supported ranges; darkest-10-hour midpoint values strictly after 16:00 are shifted to the preceding clock scale",
  "Avoids averaging across an artificial midnight discontinuity.",
  "Numerical zeros in darkest-10-hour mean melEDI",
  "Not operationally specified",
  "A geometric-mean back-transformation residual within the unit-aware floating-point tolerance is set to exact zero only when every finite source minute in the selected window is exactly zero; missing minutes remain missing",
  "Preserves the non-negative measurement domain without treating missing observations as zeros.",
  "Longest period",
  "No identifiability rule stated",
  "All-available value treated as an observed lower bound; an exactly identified period is a sensitivity",
  "Gaps cannot be mistaken for a continuously observed period.",
  "Participant correlations",
  "Correlation matrices requested",
  "Spearman coefficients and intervals are descriptive only, without a second significance screen",
  "The site-adjusted model family remains the inferential analysis."
) |>
  gt::gt(rowname_col = "Aspect") |>
  gt::cols_label(
    Preregistered_or_unspecified = "Preregistered or unspecified",
    Analysis = "Analysed",
    Why_it_matters = "Why it matters"
  ) |>
  gt::cols_width(
    Preregistered_or_unspecified ~ gt::pct(24),
    Analysis ~ gt::pct(34),
    Why_it_matters ~ gt::pct(30)
) |>
  h05_gt(12)
```

**Current scientific deviations**

- Placement roles and matched comparison:
  [DEV-001](../preregistration_deviations.qmd#dev-001),
  [REP-001](../preregistration_deviations.qmd#rep-001), and
  [REP-002](../preregistration_deviations.qmd#rep-002).
- Site-adjusted H05 association models rather than pooled correlations:
  [DEV-028](../preregistration_deviations.qmd#dev-028).
- Four factors × 17 metrics and complete 68-test FDR families:
  [DEV-029](../preregistration_deviations.qmd#dev-029).
- Current window, timing, and continuous-period rules:
  [DEV-054](../preregistration_deviations.qmd#dev-054).
- Current mean of viable one-minute MDER ratios:
  [DEV-058](../preregistration_deviations.qmd#dev-058).

**Resolved implementation history**

- Complete-vector FDR adjustment replaced the historical scalar adjustment:
  [IMP-001](../preregistration_deviations.qmd#imp-001).
- Site-adjusted model inference replaced the historical inconsistent
  correlation display:
  [IMP-006](../preregistration_deviations.qmd#imp-006).

The two IMP entries are resolved history, not additional current methods or
open qualifications.

## Source data and technical provenance

Exact source-data links accompany each result, model-check display, and
sensitivity analysis above. The full data-preparation, model-construction,
model-check, and output lineage is documented in the
[H05 analysis preparation and provenance](../../audit/hypotheses/H05/H05_analysis_preparation.qmd).

<details>
<summary>Implementation history for the current MDER and numerical-zero rules</summary>

The current MDER outputs were produced by bounded MDER-only updates that
preserved all non-MDER raw tests and fitted objects. The primary and
gap-timing-unaware preservation records are the
[MDER reconciliation](../../artifacts/12_manifests/H05/H05_metric010_reconciliation.csv)
and [gap MDER reconciliation](../../artifacts/12_manifests/H05/H05_metric010_gap_reseal_reconciliation.csv).

For darkest-10-hour mean melEDI, applying the current numerical-zero rule
changed three primary near-eye and five primary chest participant-days from
approximately 4.16 × 10^-17^ lx to exactly 0 lx without changing a fitted
sample. Only the corresponding primary models, model checks, influence
checks, matched comparison, and mathematically dependent complete FDR
families were updated. Coefficients changed by at most 3.61 × 10^-12^ per LEBA
point; two adjusted *p*-values changed below three-decimal display precision,
no family rank changed, and all three families retained zero of 68
associations. The gap-timing-unaware L10 values were already exact zeros and
were not refitted.

</details>

```{r}
#| label: session-info-h05-reader
#| include: false

sessionInfo()
```
