---
title: "H06: Day type, exercise, sleep, and personal light exposure"
subtitle: "Associations with mean hourly melanopic equivalent daylight illuminance among observed hours that met support criteria"
date: "2026-08-11"
format:
  html:
    toc: true
    toc-depth: 2
    lightbox: true
    embed-resources: true
    css: H06.css
    code-fold: true
    code-summary: "Show analysis code"
execute:
  echo: false
  warning: false
  message: false
resources:
  - ../../artifacts/06_model_data/H06/H06_category_cells.csv
  - ../../artifacts/09_tables/H06/H06_stage3_site_specific_significance_screen.csv
  - ../../artifacts/11_source_data/H06/H06_reader_primary_effects_figure.csv
  - ../../artifacts/11_source_data/H06/H06_stage3_site_specific_significance_screen_figure.csv
  - ../../artifacts/11_source_data/H06/H06_paired_placement_effects_figure.csv
  - ../../artifacts/11_source_data/H06/H06_primary_residual_clock_figure.csv
  - ../../artifacts/11_source_data/H06/H06_reader_temporal_day_type_curves.csv
  - ../../artifacts/11_source_data/H06/H06_reader_temporal_day_type_ratios.csv
  - ../../artifacts/11_source_data/H06/H06_reader_temporal_day_type_support.csv
  - ../../artifacts/11_source_data/H06/H06_reader_temporal_activity_curves.csv
  - ../../artifacts/11_source_data/H06/H06_reader_temporal_activity_ratios.csv
  - ../../artifacts/11_source_data/H06/H06_reader_temporal_activity_support.csv
  - ../../artifacts/11_source_data/H06/H06_primary_work_free_site_deviations.csv
  - ../../artifacts/06_model_data/H06/employment_eligibility_sensitivity/H06_employment_eligibility_sample_flow.csv
  - ../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_effect_comparison.csv
  - ../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_heterogeneity_comparison.csv
editor: source
---

```{r}
#| label: setup

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

h06_root <- locate_project_root()
source(file.path(h06_root, "scripts", "pipeline", "paths_io.R"))
source(file.path(h06_root, "scripts", "pipeline", "p_value_display.R"))
source(file.path(
  h06_root, "scripts", "hypotheses", "H06", "h06_contract.R"
))
knitr::opts_knit$set(
  root.dir = file.path(h06_root, "notebooks", "hypotheses")
)

options(dplyr.summarise.inform = FALSE)

artifact <- function(...) file.path(h06_root, "artifacts", ...)
read_h06 <- function(...) {
  readr::read_csv(artifact(...), show_col_types = FALSE, na = "")
}

samples <- read_h06("06_model_data", "H06", "H06_exact_samples.csv")
multiplicity <- read_h06(
  "06_model_data", "H06", "H06_multiplicity_registry.csv"
)
package_versions <- read_h06(
  "06_model_data", "H06", "H06_package_versions.csv"
)
primary_contrasts <- read_h06(
  "09_tables", "H06", "H06_robust_F3_practical_contrasts.csv"
)
wald_tests <- read_h06(
  "09_tables", "H06", "H06_robust_wald_tests.csv"
)
sensitivity <- read_h06(
  "09_tables", "H06", "H06_robust_sensitivity_comparisons.csv"
)
employment_eligibility_flow <- read_h06(
  "06_model_data", "H06", "employment_eligibility_sensitivity",
  "H06_employment_eligibility_sample_flow.csv"
)
employment_eligibility_effects <- read_h06(
  "09_tables", "H06", "employment_eligibility_sensitivity",
  "H06_employment_eligibility_effect_comparison.csv"
)
employment_eligibility_heterogeneity <- read_h06(
  "09_tables", "H06", "employment_eligibility_sensitivity",
  "H06_employment_eligibility_heterogeneity_comparison.csv"
)
exploratory_effects <- read_h06(
  "09_tables", "H06", "H06_robust_exploratory_effects.csv"
)
fit_diagnostics <- read_h06(
  "08_diagnostics", "H06", "H06_robust_fit_diagnostics.csv"
)
calibration <- read_h06(
  "08_diagnostics", "H06", "H06_robust_fitted_decile_calibration.csv"
)
residual_acf <- read_h06(
  "08_diagnostics", "H06", "H06_robust_residual_acf.csv"
)
influence <- read_h06(
  "08_diagnostics", "H06", "H06_robust_influence_deletions.csv"
)
two_part_fit <- read_h06(
  "08_diagnostics", "H06", "H06_exploratory_two_part_fit_summary.csv"
)
two_part_k <- read_h06(
  "08_diagnostics", "H06", "H06_exploratory_two_part_k_sensitivity.csv"
)
two_part_k_check <- read_h06(
  "08_diagnostics", "H06", "H06_exploratory_two_part_k_check.csv"
)
two_part_formula_registry <- read_h06(
  "06_model_data", "H06", "H06_exploratory_two_part_formula_registry.csv"
)
two_part_day_type_predictions <- read_h06(
  "11_source_data", "H06",
  "H06_exploratory_two_part_day_type_predictions.csv"
)
two_part_activity_predictions <- read_h06(
  "11_source_data", "H06",
  "H06_exploratory_two_part_activity_predictions.csv"
)
temporal_contrasts <- read_h06(
  "11_source_data", "H06",
  "H06_exploratory_two_part_temporal_contrasts.csv"
)
work_free_site_deviations <- read_h06(
  "11_source_data", "H06",
  "H06_primary_work_free_site_deviations.csv"
)
site_specific_screen <- read_h06(
  "09_tables", "H06",
  "H06_stage3_site_specific_significance_screen.csv"
)

stage2_figure_qa <- read_h06(
  "12_manifests", "H06", "H06_stage2_figure_readability_qa.csv"
)
reader_display_qa <- read_h06(
  "12_manifests", "H06",
  "H06_stage3_reader_display_figure_readability_qa.csv"
)
site_screen_figure_qa <- read_h06(
  "12_manifests", "H06",
  "H06_stage3_site_specific_significance_figure_readability_qa.csv"
)

predictor_labels <- c(
  work_free_day = "Free day versus work day",
  activity_status = "Active versus sedentary",
  previous_sleep_duration_centered_h = "Previous sleep duration (per hour)"
)

display_predictor <- function(x) {
  out <- unname(predictor_labels[x])
  out[is.na(out)] <- x[is.na(out)]
  out
}

format_effect <- function(estimate, low, high, digits = 2L) {
  sprintf(
    paste0("%.", digits, "f × (%.", digits, "f–%.", digits, "f)"),
    estimate, low, high
  )
}

format_p_cell <- function(value, significant = FALSE) {
  display <- nh_p_value_display(value, significant = significant)
  ifelse(
    display$p_bold,
    paste0("**", display$p_display, "**"),
    display$p_display
  )
}

compact_gt <- function(
  data,
  title = NULL,
  subtitle = NULL,
  source_note = NULL,
  groupname_col = NULL,
  font_size = 12
) {
  output <- if (inherits(data, "gt_tbl")) {
    data
  } else {
    gt::gt(data, groupname_col = groupname_col)
  }
  output <- output |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      heading.align = "left",
      source_notes.font.size = gt::px(11)
    )
  if (!is.null(title) || !is.null(subtitle)) {
    output <- gt::tab_header(output, title = title, subtitle = subtitle)
  }
  if (!is.null(source_note)) {
    output <- gt::tab_source_note(output, source_note = source_note)
  }
  output
}

primary_run <- "main__glasses__all_available"
primary_results <- primary_contrasts |>
  dplyr::filter(.data$run_id == primary_run) |>
  dplyr::mutate(
    Predictor = display_predictor(.data$predictor_id),
    `Ratio (95% CI)` = format_effect(
      .data$estimate_ratio,
      .data$conf_low_ratio,
      .data$conf_high_ratio
    ),
    `FDR-adjusted p` = format_p_cell(
      .data$p_adjusted,
      .data$p_adjusted <= 0.05
    )
  )

primary_heterogeneity <- wald_tests |>
  dplyr::filter(
    .data$run_id == primary_run,
    .data$family_id == multiplicity$family_id[
      multiplicity$content ==
        "Primary near-eye site-heterogeneity tests"
    ]
  ) |>
  dplyr::mutate(
    Predictor = display_predictor(.data$predictor_id),
    Test = sprintf(
      "F(%d, %d) = %.2f",
      .data$restrictions,
      .data$denominator_df,
      .data$f_statistic
    ),
    `FDR-adjusted p` = format_p_cell(
      .data$p_adjusted,
      .data$p_adjusted <= 0.05
    )
  )

calibration_ranges <- calibration |>
  dplyr::group_by(.data$run_id, .data$model_role) |>
  dplyr::summarise(
    calibration_low = min(.data$observed_to_fitted_ratio),
    calibration_high = max(.data$observed_to_fitted_ratio),
    .groups = "drop"
  )

lag_summary <- residual_acf |>
  dplyr::filter(
    .data$model_role == "additive",
    .data$lag_hours %in% c(1, 6)
  ) |>
  dplyr::select(
    .data$run_id,
    .data$lag_hours,
    .data$residual_correlation
  ) |>
  tidyr::pivot_wider(
    names_from = .data$lag_hours,
    values_from = .data$residual_correlation,
    names_prefix = "lag_"
  )

diagnostic_summary <- fit_diagnostics |>
  dplyr::filter(.data$model_role == "additive") |>
  dplyr::left_join(calibration_ranges, by = c("run_id", "model_role")) |>
  dplyr::left_join(lag_summary, by = "run_id")

influence_summary <- influence |>
  dplyr::mutate(
    ci_excludes_one =
      .data$deletion_conf_low_ratio > 1 |
        .data$deletion_conf_high_ratio < 1
  ) |>
  dplyr::group_by(.data$deletion_type, .data$predictor_id) |>
  dplyr::summarise(
    Deletions = dplyr::n(),
    `Largest shift (HC3 SE)` = max(.data$absolute_shift_full_hc3_se),
    `CIs excluding 1` = sum(.data$ci_excludes_one),
    .groups = "drop"
  )

reader_figure_qa <- dplyr::bind_rows(
  stage2_figure_qa |>
    dplyr::filter(.data$figure_id %in% c(
      "H06_primary_residual_clock",
      "H06_paired_placement_effects"
    )),
  reader_display_qa,
  site_screen_figure_qa
)

input_contract <- h06_input_contract(h06_root) |>
  dplyr::mutate(
    observed_sha256 = vapply(.data$path, artifact_sha256, character(1)),
    hash_verified = .data$observed_sha256 == .data$expected_sha256
  )

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  all(input_contract$hash_verified),
  nrow(samples) == 6L,
  all(fit_diagnostics$numerical_gate_pass),
  all(wald_tests$status == "ESTIMABLE"),
  all(primary_contrasts$planned_size == 3L),
  all(reader_figure_qa$report_011_status == "PASS"),
  all(reader_figure_qa$effective_final_text_pt >= 7),
  nrow(reader_figure_qa) == 6L,
  all(two_part_fit$converged),
  all(two_part_fit$discrete),
  all(two_part_formula_registry$selected_basis_k == 16L),
  !any(two_part_formula_registry$sum_to_zero_smooths),
  !any(two_part_formula_registry$site_specific_clock_smooth),
  nrow(two_part_day_type_predictions) == 194L,
  nrow(two_part_activity_predictions) == 194L,
  nrow(temporal_contrasts) == 194L,
  identical(
    sort(unique(temporal_contrasts$contrast_id)),
    c("active_vs_sedentary", "free_vs_work")
  ),
  all(temporal_contrasts$basis_k == 16L),
  all(temporal_contrasts$sites_standardized == 9L),
  all(temporal_contrasts$null_reference_ratio == 1),
  all(temporal_contrasts$pointwise_highlighting == "none"),
  all(is.finite(temporal_contrasts$expected_melEDI_ratio)),
  all(temporal_contrasts$pointwise_low_ratio > 0),
  all(
    temporal_contrasts$pointwise_low_ratio <=
      temporal_contrasts$expected_melEDI_ratio
  ),
  all(
    temporal_contrasts$expected_melEDI_ratio <=
      temporal_contrasts$pointwise_high_ratio
  ),
  nrow(work_free_site_deviations) == 9L,
  isTRUE(all.equal(
    unique(work_free_site_deviations$equal_site_heterogeneity_model_ratio),
    1.15495727277359,
    tolerance = 1e-12
  )),
  isTRUE(all.equal(
    unique(
      work_free_site_deviations[[
        "equal_site_heterogeneity_model_conf_low_ratio"
      ]]
    ),
    0.926410139391625,
    tolerance = 1e-12
  )),
  isTRUE(all.equal(
    unique(
      work_free_site_deviations[[
        "equal_site_heterogeneity_model_conf_high_ratio"
      ]]
    ),
    1.43988741618114,
    tolerance = 1e-12
  )),
  identical(
    work_free_site_deviations$display_name[
      work_free_site_deviations$pointwise_deviation == "above_average"
    ],
    c("Borås (SE)", "Dortmund (DE)")
  ),
  !any(
    work_free_site_deviations$pointwise_deviation == "below_average"
  ),
  nrow(site_specific_screen) == 27L,
  dplyr::n_distinct(site_specific_screen$family_id) == 3L,
  all(!is.na(site_specific_screen$family_id)),
  all(site_specific_screen$planned_size == 9L),
  all(site_specific_screen$observed_size == 9L),
  all(site_specific_screen$family_method == "Benjamini-Hochberg"),
  dplyr::n_distinct(site_specific_screen$reference_family_id) == 1L,
  all(site_specific_screen$reference_planned_size == 9L),
  all(site_specific_screen$reference_observed_size == 9L),
  all(site_specific_screen$reference_family_method == "Benjamini-Hochberg"),
  all(is.finite(
    site_specific_screen$reference_vs_equal_site_p_adjusted
  )),
  !any(
    site_specific_screen$
      reference_vs_equal_site_adjusted_significant_0_05
  ),
  identical(
    paste(
      site_specific_screen$predictor_id[
        site_specific_screen$adjusted_significant_0_05
      ],
      site_specific_screen$site[
        site_specific_screen$adjusted_significant_0_05
      ],
      sep = "__"
    ),
    c(
      "work_free_day__RISE",
      "work_free_day__BAUA",
      "activity_status__MPI"
    )
  )
)
```

## Scientific question

The preregistered hypothesis was:

> **H6: Day type, daily exercise, and sleep variables predict daily light exposure metrics.**

The main H06 analysis deliberately narrows that broad question to the
zero-aware geometric mean of melanopic equivalent daylight illuminance
(melEDI) in each observed one-hour interval that met the support criteria. It
asks whether estimated mean hourly near-eye melEDI differs for free versus
work days, active versus sedentary day status, and per one additional hour of
previous sleep. It averages over retained hours rather than estimating daily
light-exposure metrics. Time of day is examined separately and exploratorily.
Near-eye measurements are primary; chest measurements provide complementary,
non-ocular evidence. Neither position is a direct retinal measure.

The [complementary participant-day analysis](H06_daily.qmd), which treats one
calendar day from one participant as the analytical unit, addresses daily
light-exposure metrics. It is complementary evidence, while this hourly report
remains the main H06 analysis. The two reports answer related but different
questions, so their estimates are not interchangeable.

[How the hourly H06 analysis was prepared](../../audit/hypotheses/H06/H06_analysis_preparation.qmd)
traces the inputs, fitted samples, formulas, model checks, sensitivity
analyses, and source data underlying this report.

::: {.callout-note title="Answer in brief"}
- **Primary result.** After false-discovery-rate adjustment, activity status
  (**FDR-adjusted p < 0.001**) and day type (**FDR-adjusted p = 0.004**) met
  the adjusted 0.050 threshold. Previous sleep did not (FDR-adjusted p =
  0.745).

- **Qualification.** The day-type association is site- and model-dependent.
  Chest, paired-day, dataset-sensitivity, and model-check results provide
  complementary qualifications. The primary conclusions were also stable in
  a near-eye-only sensitivity that excluded six participants outside the
  employment requirement.

- **Timing context.** The nonlinear clock-time analysis is exploratory. Its
  intervals describe individual displayed times, not a simultaneous or
  multiplicity-controlled whole-curve test.

All reported associations are observational.
:::

## Reader orientation

A participant-hour is one observed one-hour interval from one participant
that met the support criteria; a participant-day is one calendar day
contributed by one participant. The analysis reports the estimated mean
hourly melEDI among observed hours that met the support criteria. Because
participant-days contribute different numbers of retained participant-hours,
the main estimand gives more weight to days with more observed hours.

Near-eye measurements are primary because the sensor position is closer to
ocular light exposure. Chest measurements are complementary, non-ocular
evidence. Neither sensor is a direct retinal measure.

A participant-cluster-robust 95% confidence interval (95% CI) allows
observations from the same participant to be related and uses the HC3
small-sample correction. False-discovery-rate adjustment is used for each
declared set of related tests and is abbreviated FDR below. A
predictor-by-site interaction model allows an association to differ by study
site. Its site-average estimate is an average across sites that gives each
site equal weight. A sensitivity analysis asks whether a conclusion changes
when a defined feature of the dataset, sample, covariance correction, working
variance, site weighting, calendar label, participant or site inclusion, or
response model is changed.


## Principal hourly result

@fig-h06-core-effects summarizes the common-effect estimates and
their supporting sensor-position, paired-day, and dataset comparisons.

```{r}
#| label: fig-h06-core-effects
#| fig-alt: "Three aligned forest-plot panels show ratios and participant-cluster-robust 95% confidence intervals for free versus work day, active versus sedentary day status, and each additional hour of previous sleep. Rows compare the primary near-eye estimate with complementary chest, paired-day near-eye and chest, and gap-timing-unaware near-eye and chest estimates. Activity ratios remain above the null ratio of one across the displayed analyses, whereas previous-sleep intervals include one."
#| fig-cap: "Estimated mean hourly melEDI among observed hours that met the support criteria. Points and bars are ratios and participant-cluster-robust 95% CIs; the dashed line is the null ratio of 1. Near-eye is the primary sensor position, and chest estimates are complementary non-ocular comparisons."
#| out-width: 170mm

knitr::include_graphics(artifact(
  "10_figures", "H06", "H06_reader_primary_effects.png"
))
```

[Source CSV](../../artifacts/11_source_data/H06/H06_reader_primary_effects_figure.csv)

@tbl-h06-primary-effects gives the three primary common-effect tests.

```{r}
#| label: tbl-h06-primary-effects
#| tbl-cap: "Primary near-eye associations."

primary_results |>
  dplyr::select(
    .data$Predictor,
    .data$`Ratio (95% CI)`,
    .data$`FDR-adjusted p`
  ) |>
  compact_gt(
    source_note = paste0(
      "Directions are free versus work day, active versus sedentary day ",
      "status, and the ratio per one additional hour of previous sleep. ",
      "Participant-cluster-robust 95% CIs use HC3 with 137 participants. ",
      "The three p-values form one FDR-adjusted family; the null ratio is 1."
    )
  ) |>
  gt::fmt_markdown(columns = `FDR-adjusted p`)
```

Estimated mean hourly near-eye melEDI among observed hours that met the support
criteria was 1.45 times as high on free as on work days (95% CI 1.14–1.85;
FDR-adjusted p = 0.004) and 2.06 times as high on active as on sedentary days
(1.54–2.76; FDR-adjusted p < 0.001). The ratio per one additional hour of
previous sleep was 0.98 (0.85–1.12; FDR-adjusted p = 0.745). That interval
allows a modest decrease or increase; it does not establish the absence of an
association.

## Site dependence

The work/free association differed across study sites (FDR-adjusted p =
0.017). Activity and previous sleep had no FDR-retained predictor-by-site
interaction. In the interaction model, the site-average work/free ratio,
calculated with equal weight for each site, was 1.15 (95% CI 0.93–1.44),
compared with 1.45 in the common-effect model. The interaction model's
site-average day-type association did not meet the adjusted threshold
(three-predictor FDR-adjusted p = 0.298). Day type also did not meet the
adjusted threshold when the working variance changed (FDR-adjusted p = 0.260).
The day-type conclusion is therefore site- and model-dependent.

<details>
<summary>Show site-specific estimates and interaction details</summary>

### Site-specific estimates and interaction details

```{r}
#| label: tbl-h06-primary-site-interactions
#| tbl-cap: "Predictor-by-site interaction tests for the primary near-eye analysis."

primary_heterogeneity |>
  dplyr::select(
    .data$Predictor,
    .data$Test,
    .data$`FDR-adjusted p`
  ) |>
  compact_gt(
    source_note = paste0(
      "The three predictor-by-site interaction blocks form one FDR-adjusted ",
      "family. Only adjusted p-values at alpha = 0.050 ",
      "are bolded."
    )
  ) |>
  gt::fmt_markdown(columns = `FDR-adjusted p`)
```

The work/free association differed across sites (FDR-adjusted p = 0.017).
Activity and previous sleep had no FDR-retained site interaction. The primary
1.45 work/free ratio comes from a model that constrains this association to be
common across sites; it is not the average of the ratios from the interaction
model. In the interaction model, the site-average ratio is the geometric mean
of the nine site ratios: 1.15 (95% CI 0.93–1.44). Three sites are above that
average and six are below it because three larger positive log-scale
departures balance six smaller negative departures. This is not an algebraic
inconsistency, but the attenuation from 1.45 to 1.15 shows that the work/free
conclusion depends on model structure. The interaction-model site averages
remain close to the common-effect results for activity (1.93 versus 2.06) and
previous sleep (0.97 versus 0.98).

```{r}
#| label: fig-h06-primary-site-associations
#| fig-alt: "Three aligned site-specific forest-plot panels show free versus work day, active versus sedentary day status, and previous-sleep ratios with pointwise 95% confidence intervals. Sites are Borås (SE), Delft (NL), Dortmund (DE), Tübingen (DE), Munich (DE), Madrid (ES), Izmir (TR), San José (CR), and Kumasi (GH). Filled points mark associations retained after a separate nine-site FDR adjustment within that predictor: free versus work day at Borås (SE) and Dortmund (DE), and active versus sedentary at Tübingen (DE). Hollow points were not retained. A solid grey line marks ratio 1 and a dashed line marks the site-average estimate."
#| fig-cap: "Site-specific estimates from the predictor-by-site interaction model. Filled and hollow points distinguish associations that did and did not meet alpha 0.050 after a separate nine-site FDR adjustment within each predictor. Pointwise 95% CIs are unadjusted; the dashed line is the site-average estimate, calculated with equal weight for each of the nine sites."
#| out-width: 170mm

knitr::include_graphics(artifact(
  "10_figures", "H06",
  "H06_stage3_site_specific_significance_screen.png"
))
```

[Source CSV](../../artifacts/11_source_data/H06/H06_stage3_site_specific_significance_screen_figure.csv)

```{r}
#| label: tbl-h06-site-specific-associations
#| tbl-cap: "Site-specific near-eye associations with separate nine-site FDR adjustments."

site_columns <- c(
  "Reference profile",
  "Free / Work",
  "Active / Sedentary",
  "Previous sleep / hour"
)

reference_sleep_h <- unique(site_specific_screen$reference_sleep_duration_h)
stopifnot(length(reference_sleep_h) == 1L, is.finite(reference_sleep_h))

site_display_cells <- site_specific_screen |>
  dplyr::distinct(
    .data$site,
    .data$display_name,
    .data$color_hex,
    .data$display_order
  ) |>
  dplyr::mutate(
    Location = paste0(
      "<strong><span title='Study-site colour marker' ",
      "aria-label='Study-site colour marker' style='color:",
      .data$color_hex,
      ";font-size:1.15em'>●</span>&nbsp;",
      .data$display_name,
      "</strong>"
    )
  )

stopifnot(
  nrow(site_display_cells) == 9L,
  all(grepl("●", site_display_cells$Location, fixed = TRUE)),
  all(grepl("<strong>", site_display_cells$Location, fixed = TRUE))
)

overall_reference_context <- site_specific_screen |>
  dplyr::slice(1L) |>
  dplyr::transmute(
    Location = "Site-average estimate<br><small>predictor-by-site model</small>",
    row_order = 0,
    row_type = "overall",
    column_label = "Reference profile",
    cell = paste0(
      "<span style='white-space:nowrap'>",
      sprintf(
        "%.0f lx <small>(95%% CI %.0f–%.0f lx)</small>",
        .data$interaction_equal_site_reference_expected_melEDI_lx,
        .data$interaction_equal_site_reference_conf_low_melEDI_lx,
        .data$interaction_equal_site_reference_conf_high_melEDI_lx
      ),
      "</span>"
    )
  )

overall_screen_context <- site_specific_screen |>
  dplyr::distinct(
    .data$predictor_id,
    .data$interaction_equal_site_ratio,
    .data$interaction_equal_site_conf_low_ratio,
    .data$interaction_equal_site_conf_high_ratio,
    .data$interaction_equal_site_p_adjusted,
    .data$interaction_equal_site_comparison_expected_melEDI_lx,
    .data$interaction_equal_site_comparison_conf_low_melEDI_lx,
    .data$interaction_equal_site_comparison_conf_high_melEDI_lx
  ) |>
  dplyr::mutate(
    column_label = dplyr::recode(
      .data$predictor_id,
      work_free_day = "Free / Work",
      activity_status = "Active / Sedentary",
      previous_sleep_duration_centered_h = "Previous sleep / hour"
    ),
    estimate_ci = sprintf(
      "%.2f× <small>(95%% CI %.2f–%.2f)</small>",
      .data$interaction_equal_site_ratio,
      .data$interaction_equal_site_conf_low_ratio,
      .data$interaction_equal_site_conf_high_ratio
    ),
    cell = paste0(
      dplyr::if_else(.data$interaction_equal_site_p_adjusted <= 0.05, "**", ""),
      "<span style='white-space:nowrap'>",
      .data$estimate_ci,
      "</span>",
      dplyr::if_else(.data$interaction_equal_site_p_adjusted <= 0.05, "**", ""),
      "<br><small>3-predictor FDR p ",
      format_p_cell(
        .data$interaction_equal_site_p_adjusted,
        .data$interaction_equal_site_p_adjusted <= 0.05
      ),
      "</small><br><span style='white-space:nowrap'>",
      sprintf(
        "%.0f lx <small>(95%% CI %.0f–%.0f lx)</small>",
        .data$interaction_equal_site_comparison_expected_melEDI_lx,
        .data$interaction_equal_site_comparison_conf_low_melEDI_lx,
        .data$interaction_equal_site_comparison_conf_high_melEDI_lx
      ),
      "</span>"
    ),
    Location = "Site-average estimate<br><small>predictor-by-site model</small>",
    row_order = 0,
    row_type = "overall"
  ) |>
  dplyr::select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$column_label,
    .data$cell
  )

separator_screen_context <- tidyr::expand_grid(
  column_label = site_columns
) |>
  dplyr::mutate(
    Location = "&nbsp;",
    row_order = 0.5,
    row_type = "separator",
    cell = "&nbsp;",
    .before = 1
  )

site_reference_context <- site_specific_screen |>
  dplyr::filter(.data$predictor_id == "work_free_day") |>
  dplyr::left_join(
    dplyr::select(site_display_cells, .data$site, .data$Location),
    by = "site"
  ) |>
  dplyr::mutate(
    row_order = .data$display_order,
    row_type = "site",
    column_label = "Reference profile",
    estimate_ci = sprintf(
      "%.0f lx <small>(95%% CI %.0f–%.0f lx)</small>",
      .data$reference_expected_melEDI_lx,
      .data$reference_conf_low_melEDI_lx,
      .data$reference_conf_high_melEDI_lx
    ),
    cell = paste0(
      dplyr::if_else(
        .data$reference_vs_equal_site_adjusted_significant_0_05,
        "**",
        ""
      ),
      "<span style='white-space:nowrap'>",
      .data$estimate_ci,
      "</span>",
      dplyr::if_else(
        .data$reference_vs_equal_site_adjusted_significant_0_05,
        "**",
        ""
      ),
      "<br><small>vs site average FDR p ",
      format_p_cell(
        .data$reference_vs_equal_site_p_adjusted,
        .data$reference_vs_equal_site_adjusted_significant_0_05
      ),
      "</small>"
    )
  ) |>
  dplyr::select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$column_label,
    .data$cell
  )

site_screen_context <- site_specific_screen |>
  dplyr::left_join(
    dplyr::select(site_display_cells, .data$site, .data$Location),
    by = "site"
  ) |>
  dplyr::mutate(
    row_order = .data$display_order,
    row_type = "site",
    column_label = dplyr::recode(
      .data$predictor_id,
      work_free_day = "Free / Work",
      activity_status = "Active / Sedentary",
      previous_sleep_duration_centered_h = "Previous sleep / hour"
    ),
    estimate_ci = sprintf(
      "%.2f× <small>(95%% CI %.2f–%.2f)</small>",
      .data$estimate_ratio,
      .data$conf_low_ratio,
      .data$conf_high_ratio
    ),
    cell = paste0(
      dplyr::if_else(.data$adjusted_significant_0_05, "**", ""),
      "<span style='white-space:nowrap'>",
      .data$estimate_ci,
      "</span>",
      dplyr::if_else(.data$adjusted_significant_0_05, "**", ""),
      "<br><small>9-site FDR p ",
      format_p_cell(
        .data$p_adjusted,
        .data$adjusted_significant_0_05
      ),
      "</small><br><span style='white-space:nowrap'>",
      sprintf(
        "%.0f lx <small>(95%% CI %.0f–%.0f lx)</small>",
        .data$comparison_expected_melEDI_lx,
        .data$comparison_conf_low_melEDI_lx,
        .data$comparison_conf_high_melEDI_lx
      ),
      "</span>"
    )
  ) |>
  dplyr::select(
    .data$Location,
    .data$row_order,
    .data$row_type,
    .data$column_label,
    .data$cell
  )

site_screen_display <- dplyr::bind_rows(
  overall_reference_context,
  overall_screen_context,
  separator_screen_context,
  site_reference_context,
  site_screen_context
) |>
  dplyr::mutate(
    column_label = factor(.data$column_label, levels = site_columns)
  ) |>
  dplyr::arrange(.data$row_order) |>
  dplyr::select(-.data$row_order, -.data$row_type) |>
  tidyr::pivot_wider(names_from = .data$column_label, values_from = .data$cell) |>
  dplyr::select(
    .data$Location,
    `Reference profile`,
    `Free / Work`,
    `Active / Sedentary`,
    `Previous sleep / hour`
  )

site_screen_display |>
  gt::gt() |>
  gt::fmt_markdown(columns = gt::everything()) |>
  gt::cols_label(
    Location = "Site",
    `Reference profile` = gt::md(paste0(
      "Reference<br><small>Work day · sedentary · ",
      sprintf("%.1f", reference_sleep_h),
      " h sleep</small>"
    )),
    `Free / Work` = gt::md("Free /<br>Work"),
    `Active / Sedentary` = gt::md("Active /<br>Sedentary"),
    `Previous sleep / hour` = gt::md("Previous sleep /<br>hour")
  ) |>
  gt::cols_width(
    Location ~ gt::px(130),
    `Reference profile` ~ gt::px(180),
    `Free / Work` ~ gt::px(180),
    `Active / Sedentary` ~ gt::px(180),
    `Previous sleep / hour` ~ gt::px(180)
  ) |>
  compact_gt(
    font_size = 12,
    source_note = paste0(
      "The reference is a work day with sedentary status and ",
      sprintf("%.1f", reference_sleep_h),
      " h previous sleep, the participant-day mean in the fitted sample. ",
      "Reference and comparison melEDI values are response-scale point ",
      "predictions in lx. The final lx line in each effect cell is the ",
      "comparison condition: free day, active status, or one additional ",
      "hour of previous sleep. All displayed 95% CIs use participant-cluster ",
      "HC3. ",
      "The blue-grey row gives site-average estimates from the same ",
      "predictor-by-site interaction model and adjusts its three predictor ",
      "contrasts together. Site-ratio tests use three separate nine-site FDR ",
      "adjustments, one per predictor. Reference-profile site effects use a ",
      "fourth nine-site FDR adjustment against the site-average reference. Only ",
      "adjusted p-values are shown. ",
      "Bold ratios meet alpha = 0.050 against ratio 1; these are not ",
      "site-to-average or between-site tests. Bold site labels and filled ",
      "site-colour circles identify the study sites; circle fill does not ",
      "encode significance. The reference-column adjusted p-values show the ",
      "separate reference-profile site-to-average tests."
    )
  ) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#E8EEF3"),
      gt::cell_text(weight = "700"),
      gt::cell_borders(
        sides = "bottom",
        color = "#8192A3",
        weight = gt::px(2)
      )
    ),
    locations = gt::cells_body(rows = 1)
  ) |>
  gt::tab_style(
    style = list(
      gt::cell_fill(color = "#FFFFFF"),
      gt::cell_text(color = "#FFFFFF", size = gt::px(1)),
      gt::cell_borders(
        sides = c("top", "bottom"),
        color = "#FFFFFF",
        weight = gt::px(0)
      )
    ),
    locations = gt::cells_body(rows = 2)
  )
```

[Site-specific results CSV](../../artifacts/09_tables/H06/H06_stage3_site_specific_significance_screen.csv)

[Site-to-average deviation source CSV](../../artifacts/11_source_data/H06/H06_primary_work_free_site_deviations.csv)

</details>

## Sensitivity and complementary sensor evidence

The required **gap-timing-unaware dataset** uses the same 50%-per-hour and
80%-per-day coverage rules as the primary dataset, but does not use the timing
of the remaining within-hour and within-day gaps for a metric-specific
adjustment. The primary analysis could therefore be interpreted as
time-sensitive. The near-eye estimates were stable under this dataset change,
including on exact common hours. Complementary chest estimates had similar
directions.

The paired-day comparison used the same 109 participants and 553
participant-days at both sensor positions, but three hourly keys differed at
each position. It is not an observation-level identity comparison, an
equivalence analysis, or a direct sensor-position effect. Activity and
previous-sleep conclusions were stable across the declared checks. Day type
remained sensitive to the working variance and to omission of some sites.

A separate near-eye-only sensitivity excluded participants recorded as not
employed or marginally employed. It retained all nine sites and classified all
three primary associations as stable within model uncertainty.

<details>
<summary>Show complementary placement and sensitivity results</summary>

### Required dataset sensitivity

The gap-timing-unaware sensitivity applies the comparison defined above to the
same three predictors and sensor positions.

The all-available near-eye comparison contains 16,329 hours from 702
participant-days, 137 participants, and nine sites. On the exact common-hour
frame, 16,329 hour keys remain and only 12 response values differ. The three
near-eye ratios are stable: 1.44 for free/work, 2.08 for active/sedentary, and
0.96 per additional hour of sleep. The work/free site interaction also remains
FDR-retained (FDR-adjusted p = 0.015).

```{r}
#| label: tbl-h06-gap-sensitivity-comparisons
#| tbl-cap: "Primary and gap-sensitivity estimates on all available and exact common hours."

sensitivity |>
  dplyr::filter(.data$comparison_id %in% c(
    "primary_vs_gap_timing_unaware_all_available",
    "primary_vs_gap_timing_unaware_exact_common_hours"
  )) |>
  dplyr::mutate(
    Frame = dplyr::recode(
      .data$comparison_id,
      primary_vs_gap_timing_unaware_all_available = "All available",
      primary_vs_gap_timing_unaware_exact_common_hours = "Exact common hours"
    ),
    Predictor = display_predictor(.data$predictor_id),
    Primary = format_effect(
      .data$reference_ratio,
      .data$reference_conf_low,
      .data$reference_conf_high
    ),
    `Gap sensitivity` = format_effect(
      .data$alternative_ratio,
      .data$alternative_conf_low,
      .data$alternative_conf_high
    )
  ) |>
  dplyr::select(
    .data$Frame,
    .data$Predictor,
    .data$Primary,
    .data$`Gap sensitivity`
  ) |>
  compact_gt(
    groupname_col = "Frame",
    source_note = "Exact-common models use identical hour keys."
  )
```

### Employment-eligibility sensitivity (near-eye only)

This sensitivity excluded six participants recorded as not employed or
marginally employed. Students and trainees were retained, and age above 65
years was not an exclusion criterion. The restricted sample contained 15,871
supported hours from 684 participant-days and 131 participants at all nine
sites, compared with 16,596 hours, 715 participant-days, and 137 participants
in the primary near-eye sample. The same additive and predictor-by-site models,
participant-cluster HC3 inference, and three-test false-discovery-rate families
were used.

```{r}
#| label: tbl-h06-employment-eligibility-sensitivity
#| tbl-cap: "Near-eye associations before and after the employment-eligibility restriction."

employment_eligibility_effects |>
  dplyr::mutate(
    Predictor = display_predictor(.data$predictor_id),
    Primary = format_effect(
      .data$reference_ratio,
      .data$reference_conf_low,
      .data$reference_conf_high
    ),
    `Employment eligible` = format_effect(
      .data$alternative_ratio,
      .data$alternative_conf_low,
      .data$alternative_conf_high
    ),
    `Employment-eligible FDR-adjusted p` = format_p_cell(
      .data$alternative_p_adjusted,
      .data$alternative_p_adjusted <= 0.05
    ),
    Stability = dplyr::recode(
      .data$detailed_stability_classification,
      `stable within model uncertainty` =
        "Stable within model uncertainty"
    )
  ) |>
  dplyr::select(
    .data$Predictor,
    .data$Primary,
    .data$`Employment eligible`,
    .data$`Employment-eligible FDR-adjusted p`,
    .data$Stability
  ) |>
  compact_gt(
    font_size = 11,
    source_note = paste(
      "Ratios compare free with work days, active with sedentary days, and",
      "one additional hour of previous sleep. The three employment-eligible",
      "p-values form one FDR-adjusted family; the null ratio is 1."
    )
  ) |>
  gt::fmt_markdown(columns = `Employment-eligible FDR-adjusted p`)
```

The restricted estimates were 1.46 for free versus work days (95% CI
1.11–1.91; **FDR-adjusted p = 0.010**), 1.88 for active versus sedentary days
(1.42–2.48; **FDR-adjusted p < 0.001**), and 0.95 per additional hour of
previous sleep (0.83–1.09; FDR-adjusted p = 0.437). Day-type heterogeneity
across sites also remained supported, F(8, 130) = 3.59
(**FDR-adjusted p = 0.003**); activity and previous-sleep heterogeneity remained
unsupported. Both restricted fits passed the stored numerical checks. The
existing residual mean-variance, calibration, and temporal-correlation
qualifications remain.

[Sample-flow CSV](../../artifacts/06_model_data/H06/employment_eligibility_sensitivity/H06_employment_eligibility_sample_flow.csv) ·
[association-comparison CSV](../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_effect_comparison.csv) ·
[site-heterogeneity CSV](../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_heterogeneity_comparison.csv)

### Complementary chest and paired-day results

The all-available chest ratios are 1.51 (95% CI 1.21–1.89), 1.81
(1.28–2.54), and 0.95 (0.84–1.07) for free/work, active/sedentary, and sleep,
respectively. Chest is a complementary non-ocular sensor position. The
paired-day comparison uses the same 109 participants and 553 participant-days
at both sensor positions. Near-eye ratios are 1.59, 1.96, and 0.96; chest
ratios are 1.53, 1.82, and 0.95. Each frame contains 12,842 participant-hours,
but three hourly keys differ at each sensor position. An observation-level
identity scatter is therefore not valid, and similar directions do not show
equivalence or estimate a direct sensor-position effect.

```{r}
#| label: fig-h06-paired-placement
#| fig-alt: "Side-by-side forest plot comparing near-eye and chest ratios with 95% confidence intervals on the paired-day sample. Both placements show ratios above one for free versus work and active versus sedentary days, while the sleep intervals include one."
#| fig-cap: "Near-eye and complementary chest estimates for the same participants and participant-days. Three hourly keys differ at each sensor position, so an observation-level identity scatter is not valid."
#| out-width: 170mm

knitr::include_graphics(artifact(
  "10_figures", "H06", "H06_paired_placement_effects.png"
))
```

[Source CSV](../../artifacts/11_source_data/H06/H06_paired_placement_effects_figure.csv)

### Working-variance and influence sensitivity analyses

Changing the working variance from power 1 to fixed power 1.8 changes the
work/free estimate to 1.15 (95% CI 0.91–1.45; FDR-adjusted p = 0.260).
Activity remains 1.91 (1.49–2.45; FDR-adjusted p < 0.001), and sleep is 0.95
(0.88–1.04; FDR-adjusted p = 0.260). Changing the participant-cluster
small-sample correction from HC3 to HC1 or HC2 changes uncertainty only
slightly. Giving each site equal weight or weighting sites by the observed
sample gives the same common-effect estimates. Replacing diary-defined day
type with weekday/weekend gives 1.23 (0.92–1.66) for weekend versus weekday
and has no new p-value family.

```{r}
#| label: tbl-h06-key-sensitivities
#| tbl-cap: "Selected declared sensitivities for the three primary predictors."

sensitivity |>
  dplyr::filter(.data$comparison_id %in% c(
    "quasi_poisson_vs_fixed_p1_8",
    "HC3_vs_HC1",
    "HC3_vs_HC2",
    "equal_site_vs_observed_sample"
  )) |>
  dplyr::mutate(
    Comparison = dplyr::recode(
      .data$comparison_id,
      quasi_poisson_vs_fixed_p1_8 = "Quasi-Poisson vs p = 1.8",
      HC3_vs_HC1 = "HC3 vs HC1",
      HC3_vs_HC2 = "HC3 vs HC2",
      equal_site_vs_observed_sample =
        "Site-average weighting compared with observed-site weighting"
    ),
    Predictor = display_predictor(.data$predictor_id),
    Reference = format_effect(
      .data$reference_ratio,
      .data$reference_conf_low,
      .data$reference_conf_high
    ),
    Alternative = format_effect(
      .data$alternative_ratio,
      .data$alternative_conf_low,
      .data$alternative_conf_high
    ),
    Classification = tools::toTitleCase(gsub(
      "_",
      " ",
      .data$stability_classification
    ))
  ) |>
  dplyr::select(
    .data$Comparison,
    .data$Predictor,
    .data$Reference,
    .data$Alternative,
    .data$Classification
  ) |>
  compact_gt(groupname_col = "Comparison")
```

Seven score-ranked participants and all nine sites were deleted one at a time;
all 16 refit sets passed the numerical model checks. Activity intervals exclude one
in all 16 checks, and sleep intervals do so in none. Work/free intervals
exclude one after all seven participant deletions but only seven of nine site
deletions: omitting Dortmund (DE) gives 1.26 (0.98–1.62), and omitting
Borås (SE) gives 1.34 (0.99–1.80).

```{r}
#| label: tbl-h06-influence-checks
#| tbl-cap: "Targeted participant and site deletion checks."

influence_summary |>
  dplyr::mutate(
    Type = dplyr::recode(
      .data$deletion_type,
      participant = "Participant deletion",
      site = "Site deletion"
    ),
    Predictor = display_predictor(.data$predictor_id),
    `CIs excluding 1` = paste0(
      .data$`CIs excluding 1`,
      "/",
      .data$Deletions
    )
  ) |>
  dplyr::select(
    .data$Type,
    .data$Predictor,
    .data$`Largest shift (HC3 SE)`,
    .data$`CIs excluding 1`
  ) |>
  compact_gt() |>
  gt::fmt_number(columns = `Largest shift (HC3 SE)`, decimals = 3)
```

</details>

## Model-check qualification

All common-effect and predictor-by-site interaction fits met the stored
numerical checks. The primary response nevertheless has substantial
dispersion, imperfect fitted-decile calibration, increasing residual spread,
and residual correlation across nearby hours. Participant-cluster HC3
supports the reported uncertainty but does not remove those patterns or
repair the mean structure.

<details>
<summary>Show detailed model checks and influence analyses</summary>

### Detailed model checks and influence analyses

All stored additive and full fits converge, have full-rank fixed-effect
designs, and have finite, positive-definite unmodified HC3 covariance
matrices. These checks establish numerical estimability, not adequate fit.

```{r}
#| label: tbl-h06-model-checks
#| tbl-cap: "Model checks for the six common-effect mean models."

diagnostic_summary |>
  dplyr::transmute(
    Scenario = dplyr::recode(
      .data$run_id,
      main__glasses__all_available = "Primary near-eye",
      main__chest__all_available = "Chest, all available",
      main__glasses__paired_common = "Near-eye, paired days",
      main__chest__paired_common = "Chest, paired days",
      gap_timing_unaware__glasses__all_available =
        "Gap sensitivity, near-eye",
      gap_timing_unaware__chest__all_available =
        "Gap sensitivity, chest"
    ),
    `Zero hours` = .data$exact_zero_fraction,
    Dispersion = .data$dispersion,
    `Observed/fitted decile range` = sprintf(
      "%.2f–%.2f",
      .data$calibration_low,
      .data$calibration_high
    ),
    `Residual–fitted Spearman` = .data$residual_fitted_spearman,
    `|Residual|–fitted Spearman` =
      .data$absolute_residual_fitted_spearman,
    `Residual lag 1` = .data$lag_1,
    `Residual lag 6` = .data$lag_6,
    Checks = .data$numerical_gate_pass
  ) |>
  compact_gt(
    source_note = paste0(
      "Calibration is in-sample and descriptive. Dispersion is the ",
      "Pearson working-dispersion estimate."
    )
  ) |>
  gt::fmt_percent(columns = `Zero hours`, decimals = 1) |>
  gt::fmt_number(
    columns = c(
      Dispersion,
      `Residual–fitted Spearman`,
      `|Residual|–fitted Spearman`,
      `Residual lag 1`,
      `Residual lag 6`
    ),
    decimals = 3
  ) |>
  gt::fmt(
    columns = Checks,
    fns = function(x) ifelse(
      x,
      "Meets numerical checks",
      "Review needed"
    )
  )
```

The primary response contains 28.3% exact-zero hours and has working
dispersion 3,618. Fitted-decile observed/fitted ratios range from 0.78 to
1.61. Residual–fitted Spearman correlation is -0.564, and absolute
residual–fitted correlation is 0.553. Residual correlation is 0.464 at lag 1,
0.233 at lag 2, and 0.013 at lag 6. Participant-cluster HC3 accommodates
within-participant dependence for uncertainty but does not whiten residuals or
repair the mean structure. The residual-clock display shows remaining daytime
structure.

```{r}
#| label: fig-h06-residual-clock
#| fig-alt: "Residual model-check plot across local clock hour for the primary near-eye model. Mean residuals vary over the day, and a residual-versus-fitted panel shows imperfect mean calibration and increasing spread."
#| fig-cap: "Primary residual structure by local clock and fitted value. Participant-cluster HC3 does not remove these patterns."
#| out-width: 170mm

knitr::include_graphics(artifact(
  "10_figures", "H06", "H06_primary_residual_clock.png"
))
```

[Source CSV](../../artifacts/11_source_data/H06/H06_primary_residual_clock_figure.csv)

</details>

## Exploratory diary and clock-time analyses

Exercise location, active travel, sedentary time, sleep onset, and final wake
time are descriptive extensions fitted one at a time. They do not select or
rescue the primary model. The nonlinear generalized additive model (GAM)
analysis allows the association with local clock time to bend across the day
and is separate from the primary average associations. Its pointwise 95% CI
applies to one displayed time. It is not a simultaneous band or a
multiplicity-controlled whole-curve test.

Active status was higher from about 07:15 to 19:00 and lower from about 23:30
to 05:15 across midnight, when estimated melEDI was low. The free-day curve
was lower from about 06:45 to 09:15 and from 18:45 to 19:15. These windows
exclude the null ratio 1 on the local 15-minute grid, but they do not define
confirmatory significant time segments.

<details>
<summary>Show exploratory clock-time and diary analyses</summary>

### Detailed exploratory analyses

### Additional diary measures

Exercise location, active travel, sedentary time, sleep onset, and final wake
time were fitted one at a time on their own exact samples. They are descriptive
extensions, have 95% CIs but no confirmatory p-values, and do not select or
rescue the primary model.

```{r}
#| label: tbl-h06-exploratory-diary-associations
#| tbl-cap: "Exploratory one-at-a-time diary associations."

exploratory_effects |>
  dplyr::mutate(
    Effect = dplyr::recode(
      .data$effect_id,
      Indoors_vs_Outdoors = "Exercise indoors vs outdoors",
      `Indoors and outdoors_vs_Outdoors` = "Both locations vs outdoors",
      per_hour__active_commute_h = "Active travel, per hour",
      per_hour__sedentary_h = "Sedentary time, per hour",
      per_hour__previous_sleep_onset_centered_h =
        "Sleep onset, per hour later",
      per_hour__wake_centered_h = "Wake time, per hour later"
    ),
    `Ratio (95% CI)` = format_effect(
      .data$estimate_ratio,
      .data$conf_low_ratio,
      .data$conf_high_ratio
    )
  ) |>
  dplyr::select(.data$Effect, .data$`Ratio (95% CI)`) |>
  compact_gt(
    source_note = paste0(
      "Exploratory quasi-Poisson log means with participant-cluster HC3. ",
      "No p-values are calculated or displayed."
    )
  )
```

### Nonlinear time-of-day GAM analysis

The exploratory nonlinear generalized additive model (GAM) analysis allows
the associations with local clock time to bend across the day. It separates
occurrence (melEDI greater than zero, binomial-logit) from positive magnitude
(Gamma-log); multiplying their response-scale means gives estimated hourly
melEDI. Both components use four cyclic factor-by clock smooths for work/free
day × activity status, with one smoothing parameter shared across the four
curves. Study site enters as a fixed adjustment only; there is no
site-specific clock smooth and no `sz` component. Participant and
participant-day random-effect smooths, true-sequence `AR.start`, and residual
working correlations remain. The chosen basis is `k = 16`, fitted with
`bam(method = "fREML", discrete = TRUE)`.

```{r}
#| label: tbl-h06-exploratory-two-part-formulas
#| tbl-cap: "Exploratory two-part formulas."
#| echo: true
#| code-fold: false
#| code-summary: "Exploratory two-part formulas (always shown)"

two_part_formula_registry |>
  dplyr::select(
    Component = .data$component,
    Family = .data$response_family,
    `Evaluated Wilkinson formula` = .data$selected_formula,
    Method = .data$method,
    Discrete = .data$discrete,
    `Basis k` = .data$selected_basis_k
  ) |>
  compact_gt(
    source_note = paste0(
      "Exploratory only. Positive-magnitude sequences are rebuilt after ",
      "zero hours are removed."
    )
  )
```

The occurrence and positive-magnitude models converge with full-rank
fixed-smooth designs. Their residual working correlations are 0.418 and
0.440. Occurrence calibration overpredicts the lowest fitted-probability
decile (observed/fitted ratio 0.829); positive-magnitude calibration ranges
from about 0.89 to 1.21. The four occurrence smooths have `k`-check p-values
0.013–0.023, although their edf remain below the basis ceiling; this is a
residual model-structure warning. Positive-magnitude `k`-check p-values are
0.18–0.21. Moving from `k = 12` to `k = 16` changes the combined curves by
15.1% at the 95th percentile, so the curves provide timing context rather
than confirmatory whole-curve tests.

#### Reader displays and temporal contrasts

Each figure has three aligned parts: (A) expected curves, (B) the
corresponding complete response-scale
ratio with a null reference of one, and (C) participant-hour support. The
contrast panels are derived from the fitted two-part models without refitting;
they compare complete estimated melEDI rather than one smooth term.
Free day is divided by work day after equal weighting over activity status and
the nine sites. Active is divided by sedentary after equal weighting over
work/free day and the nine sites.

The expected-value panels use a symmetric logarithmic, or symlog, display
that is linear from 0 to 1 lx and logarithmic above 1 lx. Tick labels remain
in lux.

```{r}
#| label: fig-h06-exploratory-day-type-time
#| fig-alt: "Three vertically aligned panels summarize exploratory near-eye melEDI by day type and local clock time. The top panel overlays work-day and free-day expected curves with pointwise 95% bands and dots at integer hours supported in all nine sites. The middle panel shows free day divided by work day with a dashed null line at ratio one; integer-hour points are filled when the pointwise interval excludes one and hollow when it includes one. The bottom panel shows available participant-hours for each day type at every integer hour."
#| fig-cap: "Exploratory nonlinear day-type timing display: overlaid estimated curves, the complete response-scale free-day/work-day contrast, and participant-hour support. The other binary factor and all nine sites receive equal weight. A pointwise 95% CI applies at one displayed hour, not to the whole curve. Filled and hollow points distinguish intervals that exclude and include the null ratio of 1; this is not a multiplicity-controlled or simultaneous whole-curve test."
#| out-width: 170mm

knitr::include_graphics(artifact(
  "10_figures", "H06", "H06_reader_temporal_day_type.png"
))
```

[Curve source CSV](../../artifacts/11_source_data/H06/H06_reader_temporal_day_type_curves.csv) ·
[Ratio source CSV](../../artifacts/11_source_data/H06/H06_reader_temporal_day_type_ratios.csv) ·
[Support source CSV](../../artifacts/11_source_data/H06/H06_reader_temporal_day_type_support.csv)

```{r}
#| label: fig-h06-exploratory-activity-time
#| fig-alt: "Three vertically aligned panels summarize exploratory near-eye melEDI by activity status and local clock time. The top panel overlays sedentary and active expected curves with pointwise 95% bands and dots at integer hours supported in all nine sites. The middle panel shows active divided by sedentary with a dashed null line at ratio one; integer-hour points are filled when the pointwise interval excludes one and hollow when it includes one. The bottom panel shows available participant-hours for each activity status at every integer hour."
#| fig-cap: "Exploratory nonlinear activity timing display: overlaid estimated curves, the complete response-scale active/sedentary contrast, and participant-hour support. Work and free days and all nine sites receive equal weight. A pointwise 95% CI applies at one displayed hour, not to the whole curve. Filled and hollow points distinguish intervals that exclude and include the null ratio of 1; this is not a multiplicity-controlled or simultaneous whole-curve test."
#| out-width: 170mm

knitr::include_graphics(artifact(
  "10_figures", "H06", "H06_reader_temporal_activity.png"
))
```

[Curve source CSV](../../artifacts/11_source_data/H06/H06_reader_temporal_activity_curves.csv) ·
[Ratio source CSV](../../artifacts/11_source_data/H06/H06_reader_temporal_activity_ratios.csv) ·
[Support source CSV](../../artifacts/11_source_data/H06/H06_reader_temporal_activity_support.csv)

The paired prediction gradients retain covariance between numerator and
denominator within each component. The interval calculation treats smoothing
parameters as fixed and sets cross-component occurrence–magnitude covariance
to zero. It also inherits the residual working-correlation, occurrence
calibration, and basis-sensitivity limitations above. Overnight ratios and
intervals widen where numerator and denominator expectations approach zero.
No time segment is labelled significant, and no global or simultaneous
whole-curve inference is made.

</details>

## Interpretation and limitations

The common-effect results describe observational associations with estimated
mean hourly near-eye melEDI among observed hours that met the support criteria.
They do not establish causal effects and do not answer the broader
preregistered question about daily light-exposure metrics. Participant-days
with more retained participant-hours contribute more weight.

The free/work result is qualified by its predictor-by-site interaction, the
smaller site-average estimate from that interaction model, the alternative
working-variance result, and leave-one-site-out checks. Chest and paired-day
analyses provide complementary non-ocular context, not sensor-position
equivalence or a direct placement effect. Participant-cluster-robust inference
allows within-participant dependence, but it does not remove residual
time-of-day structure or repair an imperfect mean model. The nonlinear GAM
analysis is exploratory; its pointwise intervals apply to individual displayed
hours and do not support simultaneous whole-curve claims.

## Preregistration deviations {#h06-preregistration-deviations}

- [`DEV-015`](../preregistration_deviations.qmd#dev-015) records the documented
  hourly geometric-mean melEDI estimand rather than the preregistered daily
  metric set.
- [`DEV-030`](../preregistration_deviations.qmd#dev-030) limits the primary
  family to day type, active/sedentary status, and previous-night sleep
  duration.
- [`DEV-031`](../preregistration_deviations.qmd#dev-031) records the
  zero-inclusive quasi-Poisson common-effect model, participant-cluster HC3
  inference, predictor-by-site interactions, and separate exploratory
  nonlinear clock-time analysis.
- [`DEV-032`](../preregistration_deviations.qmd#dev-032) records the separate
  FDR families and the distinction between average associations,
  predictor-by-site interactions, and exploratory site screens.

## Detailed analysis record

This section retains the exact fitted samples, model formulas, declared FDR
families, linked source data, and technical provenance for auditing.

### Outcome, predictors, sensor positions, and fitted samples

The primary outcome is the zero-aware geometric mean melEDI for each retained
participant-hour at the near-eye sensor position. A participant-hour is one
observed one-hour interval from one participant that met the support criteria
defined in [Preparation 06](../preparation/06_model_ready_datasets.qmd).
Activity is binary: sedentary means no reported exercise, while active
includes light, moderate, or vigorous reported exercise. Day type and
activity are participant-day diary variables, where a participant-day is one
calendar day contributed by one participant; previous sleep duration is
linked from the preceding sleep period. Because participant-days contribute
19–24 supported participant-hours, the estimand is participant-hour weighted
rather than participant-day weighted.

```{r}
#| label: tbl-h06-exact-samples
#| tbl-cap: "Exact fitted samples for the six core H06 analyses."

samples |>
  dplyr::transmute(
    Scenario = dplyr::recode(
      .data$run_id,
      main__glasses__all_available = "Primary near-eye",
      main__chest__all_available = "Chest, all available",
      main__glasses__paired_common = "Near-eye, paired days",
      main__chest__paired_common = "Chest, paired days",
      gap_timing_unaware__glasses__all_available =
        "Gap sensitivity, near-eye",
      gap_timing_unaware__chest__all_available =
        "Gap sensitivity, chest"
    ),
    Hours = .data$one_hour_observations,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Supported hours/day` = paste0(
      .data$minimum_supported_hours_per_day,
      "–",
      .data$maximum_supported_hours_per_day
    )
  ) |>
  compact_gt(
    source_note = paste0(
      "All rows are exact fitted samples. The complete site × category ",
      "registry is available below."
    )
  ) |>
  gt::fmt_integer(columns = c(Hours, `Participant-days`, Participants, Sites))
```

The primary fit contains 16,596 supported hours from 715 participant-days,
137 participants, and nine study sites. The smallest primary site ×
category cells contain 185 hours, 8 participant-days, and 4 participants;
these minima do not all occur in the same cell. [Download the exact category
cell registry](../../artifacts/06_model_data/H06/H06_category_cells.csv).

The TUM exercise-diary identity check uses the pinned immutable
`melidosData` 1.0.6 object and finds 7 S001 rows and 0 S101 rows, with no local
identifier rewrite. The `KNUST_S005` numeral 3,600 is interpreted as 3,600
seconds, or one sedentary hour, only for eligible exploratory analyses; it is
not added to the primary analysis.

For the required sensitivity analysis, the **gap-timing-unaware dataset**
applies the same 50%-per-hour and 80%-per-day coverage rules but does not use
the timing of remaining missing observations for a metric-specific
adjustment. The primary analysis could therefore be interpreted as
time-sensitive. Both datasets are defined in
[Preparation 06](../preparation/06_model_ready_datasets.qmd). Hereafter, the
corrected input is called the **primary dataset**.

### Models and inference

The common-effect model is a zero-inclusive quasi-Poisson generalized linear
model with a log link and fixed study-site terms. A participant-cluster-robust
95% confidence interval (95% CI) allows observations from the same participant
to be related and uses the HC3 small-sample correction. Participants, rather
than participant-hours, are therefore the inferential replicates.
One-coefficient tests use a *t* reference with 136 degrees of freedom.

The predictor-by-site interaction model allows the association of day type,
activity, or previous sleep with hourly melEDI to differ by study site. Its
site-interaction blocks use Wald *F* tests with restriction-rank and 136
denominator degrees of freedom. A site-average estimate averages across the
nine sites with equal weight for each site on the linear-predictor scale
before back-transformation.

```{r}
#| label: tbl-h06-exact-confirmatory-formulas
#| tbl-cap: "Exact evaluated Wilkinson formulas."
#| echo: true
#| code-fold: false
#| code-summary: "Exact evaluated Wilkinson formulas (always shown)"

formula_text <- function(x) paste(deparse(x), collapse = " ")
selected_formulas <- h06_formula_set()[c("additive", "full")]

tibble::enframe(
  selected_formulas,
  name = "model_id",
  value = "formula"
) |>
  dplyr::mutate(
    Role = dplyr::recode(
      .data$model_id,
      additive = "Primary population-mean associations",
      full = "Predictor-by-site interaction blocks"
    ),
    `Evaluated Wilkinson formula` = vapply(
      .data$formula,
      formula_text,
      character(1)
    )
  ) |>
  dplyr::select(Model = .data$model_id, .data$Role,
                .data$`Evaluated Wilkinson formula`) |>
  compact_gt(
    source_note = paste0(
      "All comparisons use the same prepared rows. Quasi-likelihood AIC and ",
      "likelihood-ratio tests are not reported."
    )
  )
```

False-discovery-rate (FDR) adjustment is applied separately to the three
primary association tests, the three site-interaction tests, and the three
practical contrasts. Only FDR-adjusted p-values are displayed. The exploratory
site-specific screen uses three additional, separate nine-site FDR
adjustments, one within each predictor. Chest, paired-position,
weekday/weekend, influence, and other exploratory outputs do not create post
hoc p-value families.

```{r}
#| label: tbl-h06-fdr-adjustment
#| tbl-cap: "Declared sets of tests and FDR adjustment."

multiplicity |>
  dplyr::transmute(
    `Set of tests` = .data$content,
    Tests = .data$planned_size,
    Adjustment = dplyr::recode(.data$method, BH = "FDR")
  ) |>
  compact_gt()
```

<details>
<summary>Show technical source and figure checks</summary>

### Technical source and figure checks


### Figure reproducibility

All six reader-facing figures were assessed at 170 mm on an A4 portrait page
with 20-mm side margins. Native and intended widths are equal, and the
smallest essential text remains 7.5 pt. Physical-size inspection found no
clipping, overlap, harmful wrapping, distortion, or materially imbalanced
data region. Every figure has accessible alt text and linked source data.

```{r}
#| label: tbl-h06-figure-readability-checks
#| tbl-cap: "Physical-size checks for all reader-facing H06 figures."

reader_figure_qa |>
  dplyr::transmute(
    Figure = .data$figure_id,
    `Native width (mm)` = .data$native_width_mm,
    `Intended width (mm)` = .data$intended_print_display_width_mm,
    `Scale factor` = .data$scale_factor,
    `Smallest nominal text (pt)` =
      .data$smallest_essential_nominal_text_pt,
    `Effective final text (pt)` = .data$effective_final_text_pt,
    Status = dplyr::if_else(
      .data$report_011_status == "PASS",
      "Verified",
      "Review needed"
    )
  ) |>
  compact_gt()
```

The analysis used R 4.6.1 and the project library. Scientific transformations,
model fits, model checks, contrasts, tables, and figures are implemented in
H06-scoped R scripts. The temporal contrasts use the stored model objects and
did not refit a GAM. No bootstrap was required, no package was installed, and
no shared input was edited.

[How the hourly H06 analysis was prepared](../../audit/hypotheses/H06/H06_analysis_preparation.qmd)
provides the data, model, model-check, sensitivity, source-data, and
file-identity record for these results.

</details>
