---
title: "H07: Photoperiod and qualifying transitions in personal light exposure"
subtitle: "Pooled photoperiod associations across nine personal light-exposure metrics"
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h07-reader

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()),
  winslash = "/",
  mustWork = TRUE
)
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

read_h07 <- function(...) {
  readr::read_csv(file.path(root, ...), show_col_types = FALSE)
}

table_dir <- c("artifacts", "09_tables", "H07")
figure_dir <- file.path("artifacts", "08_figures", "H07")

main_samples <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_samples.csv"
)
plateau_all <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_plateau_summary.csv"
)
derivative_points <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_derivative_points.csv"
)
curve_points <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_curve_points.csv"
)
formula_registry <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_formula_registry.csv"
)
diagnostics <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_diagnostics.csv"
)
pairwise_concurvity <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_pairwise_concurvity.csv"
)
residual_summary <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_residual_distribution_summary.csv"
)
tweedie_pilot <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_tweedie_distribution_pilot.csv"
)
sensitivity_results <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_sensitivity_plateau_comparison.csv"
)
sensitivity_samples <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_sensitivity_samples.csv"
)
model_form_results <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_model_form_plateau_comparison.csv"
)
loso_results <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_loso_plateau_influence_summary.csv"
)
site_support <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_site_support.csv"
)
figure_settings <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_paired_figure_settings.csv"
)
session_record <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_session.csv"
)
site_registry <- read_h07("config", "site_display_registry.csv") |>
  arrange(.data$display_order)

derivative_method <- "REVISED_CENTRAL_UNCONDITIONAL_POINTWISE"
plateau <- plateau_all |>
  filter(.data$method_id == .env$derivative_method) |>
  arrange(.data$placement, .data$metric_order)
derivative_reader <- derivative_points |>
  filter(.data$method_id == .env$derivative_method)

metric_registry <- plateau |>
  distinct(
    .data$metric_id,
    .data$metric_order,
    .data$manuscript_name,
    .data$display_unit,
    .data$response_family,
    .data$response_transform
  ) |>
  arrange(.data$metric_order)

sample_display <- main_samples |>
  mutate(placement = sub("^primary__", "", .data$run_id)) |>
  left_join(
    metric_registry |>
      select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  arrange(.data$placement, .data$metric_order)

adapted_diagnostics <- diagnostics |>
  filter(.data$model_id == "adapted_photoperiod_smooth") |>
  mutate(placement = sub("^primary__", "", .data$run_id)) |>
  left_join(
    metric_registry |>
      select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name
      ),
    by = "metric_id",
    relationship = "many-to-one"
  )

target_concurvity <- pairwise_concurvity |>
  filter(
    .data$model_id == "adapted_photoperiod_smooth",
    .data$measure == "estimate",
    .data$supplier_term == "s(site_participant)",
    .data$target_term == "s(photoperiod_hours)"
  )

placement_label <- c(
  near_eye = "Near eye — primary",
  chest = "Chest — complementary"
)

fmt_number <- function(value, digits = 2L) {
  formatC(value, digits = digits, format = "f", big.mark = ",")
}

fmt_range <- function(lower, upper, digits = 2L, suffix = "") {
  paste0(
    fmt_number(lower, digits), "–", fmt_number(upper, digits), suffix
  )
}

fmt_ci <- function(estimate, lower, upper, digits = 3L) {
  paste0(
    fmt_number(estimate, digits), " (95% CI ",
    fmt_number(lower, digits), " to ", fmt_number(upper, digits), ")"
  )
}

metric_group <- function(metric_order) {
  dplyr::case_when(
    metric_order <= 3 ~ "Light level",
    metric_order <= 8 ~ "Duration and continuous period",
    TRUE ~ "Exposure history"
  )
}

h07_gt <- function(table, font_size = 12) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

sample_table <- function(placement) {
  sample_display |>
    filter(.data$placement == .env$placement) |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      Participants = .data$participants,
      `Participant-days` = .data$participant_days,
      Observations = .data$observations,
      Sites = .data$sites,
      `Civil photoperiod (h)` = fmt_range(
        .data$photoperiod_min,
        .data$photoperiod_max
      )
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$Participants,
      .data$`Participant-days`,
      .data$Observations,
      .data$Sites,
      .data$`Civil photoperiod (h)`
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::fmt_integer(
      columns = c(Participants, `Participant-days`, Observations, Sites)
    ) |>
    gt::cols_width(
      Participants ~ gt::pct(12),
      `Participant-days` ~ gt::pct(14),
      Observations ~ gt::pct(12),
      Sites ~ gt::pct(8),
      `Civil photoperiod (h)` ~ gt::pct(18)
    ) |>
    h07_gt(12)
}

result_reason <- function(disposition) {
  dplyr::case_when(
    disposition == "DERIVATIVE_DEFINED_PLATEAU_PATTERN" ~
      "Detected increase followed by a sustained zero-compatible tail",
    disposition == "NO_POINTWISE_DETECTED_PRIOR_INCREASE" ~
      "No preceding detected increase",
    disposition == "POINTWISE_DETECTED_INCREASE_AT_RECORDED_END" ~
      "Increase remained detected at the recorded maximum",
    TRUE ~ disposition
  )
}

result_table <- function(placement) {
  plateau |>
    filter(.data$placement == .env$placement) |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      Pattern = if_else(.data$revised_plateau_pattern, "Yes", "No"),
      `Transition bracket (h)` = if_else(
        .data$revised_plateau_pattern,
        fmt_range(
          .data$plateau_transition_lower,
          .data$plateau_start
        ),
        "—"
      ),
      `Derivative at maximum` = fmt_ci(
        .data$derivative_at_recorded_end,
        .data$derivative_at_recorded_end_lower,
        .data$derivative_at_recorded_end_upper
      ),
      Interpretation = result_reason(.data$disposition)
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$Pattern,
      .data$`Transition bracket (h)`,
      .data$`Derivative at maximum`,
      .data$Interpretation
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::tab_style(
      style = gt::cell_fill(color = "#E8F3F8"),
      locations = gt::cells_body(
        columns = Pattern,
        rows = Pattern == "Yes"
      )
    ) |>
    gt::cols_label(
      Pattern = "Qualifying transition",
      `Transition bracket (h)` = "Qualifying transition bracket (h)",
      `Derivative at maximum` = "Endpoint derivative"
    ) |>
    gt::cols_width(
      Pattern ~ gt::pct(8),
      `Transition bracket (h)` ~ gt::pct(16),
      `Derivative at maximum` ~ gt::pct(25),
      Interpretation ~ gt::pct(31)
    ) |>
    gt::tab_source_note(
      gt::md(
        paste0(
          "A ‘Yes’ is the formal derivative-defined classification: the ",
          "immediately preceding point has an interval wholly above zero, ",
          "the next contains zero, and every later interval through the ",
          "recorded maximum remains zero-compatible. Endpoint derivatives ",
          "and 95% confidence intervals are on each model's ",
          "linear-predictor scale, in model-scale units per hour of civil ",
          "photoperiod."
        )
      )
    ) |>
    h07_gt(12)
}

sensitivity_label <- function(sensitivity, data_scenario) {
  dplyr::case_when(
    sensitivity == "paired_common_placement" ~
      "Paired near-eye/chest common sample",
    sensitivity == "gap_timing_unaware" ~
      "Gap-timing-unaware dataset",
    sensitivity == "preparation_exact_common" &
      data_scenario == "primary" ~
      "Preparation exact-common rows: primary dataset",
    sensitivity == "preparation_exact_common" &
      data_scenario == "gap_timing_unaware" ~
      "Preparation exact-common rows: gap-timing-unaware dataset",
    sensitivity == "longest_period_exact_only" ~
      "Exactly identified longest-period rows",
    sensitivity == "observed_dose_common_sample" ~
      "Observed/corrected dose on common rows",
    TRUE ~ sensitivity
  )
}

sensitivity_summary <- sensitivity_results |>
  group_by(.data$data_scenario, .data$placement, .data$sensitivity) |>
  summarise(
    Evaluable = sum(!is.na(.data$classification_agrees)),
    Agreements = sum(.data$classification_agrees, na.rm = TRUE),
    changed = paste(
      unique(
        .data$manuscript_name[
          !is.na(.data$classification_agrees) &
            !.data$classification_agrees
        ]
      ),
      collapse = "; "
    ),
    .groups = "drop"
  ) |>
  mutate(
    Scenario = sensitivity_label(.data$sensitivity, .data$data_scenario),
    Placement = unname(placement_label[.data$placement]),
    Agreement = paste0(.data$Agreements, "/", .data$Evaluable),
    `Changed classifications` = if_else(
      .data$changed == "",
      "None",
      .data$changed
    ),
    sensitivity_order = match(
      .data$sensitivity,
      c(
        "gap_timing_unaware",
        "paired_common_placement",
        "preparation_exact_common",
        "longest_period_exact_only",
        "observed_dose_common_sample"
      )
    ),
    scenario_order = if_else(
      .data$data_scenario == "primary",
      1L,
      2L
    ),
    placement_order = if_else(.data$placement == "near_eye", 1L, 2L)
  ) |>
  arrange(
    .data$sensitivity_order,
    .data$scenario_order,
    .data$placement_order
  )

sensitivity_sample_labels <- tibble::tribble(
  ~run_id, ~Scenario, ~placement_order, ~scenario_order,
  "gap_timing_unaware__near_eye", "Gap-timing-unaware dataset", 1L, 1L,
  "gap_timing_unaware__chest", "Gap-timing-unaware dataset", 2L, 1L,
  "paired__near_eye", "Paired near-eye/chest common sample", 1L, 2L,
  "paired__chest", "Paired near-eye/chest common sample", 2L, 2L,
  "prep_common_primary__near_eye", "Preparation exact-common rows", 1L, 3L,
  "prep_common_primary__chest", "Preparation exact-common rows", 2L, 3L,
  "exact_period__near_eye", "Exactly identified longest-period rows", 1L, 4L,
  "exact_period__chest", "Exactly identified longest-period rows", 2L, 4L,
  "dose_common_corrected__near_eye", "Observed/corrected dose common rows", 1L, 5L,
  "dose_common_corrected__chest", "Observed/corrected dose common rows", 2L, 5L
)

range_text <- function(value) {
  if (min(value) == max(value)) {
    return(formatC(min(value), format = "d", big.mark = ","))
  }
  paste0(
    formatC(min(value), format = "d", big.mark = ","),
    "–",
    formatC(max(value), format = "d", big.mark = ",")
  )
}

sensitivity_sample_summary <- sensitivity_samples |>
  inner_join(sensitivity_sample_labels, by = "run_id") |>
  group_by(
    .data$run_id,
    .data$Scenario,
    .data$placement_order,
    .data$scenario_order
  ) |>
  summarise(
    Participants = range_text(.data$participants),
    `Participant-days` = range_text(.data$participant_days),
    Observations = range_text(.data$observations),
    Sites = range_text(.data$sites),
    .groups = "drop"
  ) |>
  mutate(
    Placement = if_else(.data$placement_order == 1L, "Near eye", "Chest")
  ) |>
  arrange(.data$scenario_order, .data$placement_order)

model_form_summary <- model_form_results |>
  group_by(.data$placement, .data$model_id) |>
  summarise(
    Evaluable = sum(!is.na(.data$classification_agrees)),
    Agreements = sum(.data$classification_agrees, na.rm = TRUE),
    failures = sum(grepl("^FAIL", .data$fit_status)),
    changed = paste(
      unique(
        .data$manuscript_name[
          !is.na(.data$classification_agrees) &
            !.data$classification_agrees
        ]
      ),
      collapse = "; "
    ),
    .groups = "drop"
  ) |>
  mutate(
    Model = recode(
      .data$model_id,
      adapted_photoperiod_expanded_basis = "Expanded smooth basis (k = 10)",
      adapted_photoperiod_fixed_site = "Site as a fixed effect"
    ),
    Placement = unname(placement_label[.data$placement]),
    Agreement = paste0(.data$Agreements, "/", .data$Evaluable),
    `Failed fits` = .data$failures,
    `Changed classifications` = if_else(
      .data$changed == "",
      "None",
      .data$changed
    ),
    placement_order = if_else(.data$placement == "near_eye", 1L, 2L)
  ) |>
  arrange(.data$Model, .data$placement_order)

diagnostic_summary <- adapted_diagnostics |>
  group_by(.data$placement) |>
  summarise(
    Converged = paste0(sum(.data$converged), "/", dplyr::n()),
    `Maximum absolute gradient` = max(.data$gradient_max_abs),
    `Minimum Hessian eigenvalue` = min(.data$hessian_min_eigenvalue),
    `Basis-check flags` = sum(.data$k_p_value < 0.05, na.rm = TRUE),
    `Maximum absolute lag-1` = max(
      abs(.data$pooled_consecutive_day_lag1),
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  left_join(
    target_concurvity |>
      group_by(.data$placement) |>
      summarise(
        concurvity_min = min(.data$concurvity),
        concurvity_max = max(.data$concurvity),
        .groups = "drop"
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  left_join(
    residual_summary |>
      group_by(.data$placement) |>
      summarise(
        deviance_min = min(.data$deviance_explained),
        deviance_max = max(.data$deviance_explained),
        qq_min = min(.data$normal_qq_correlation),
        qq_max = max(.data$normal_qq_correlation),
        .groups = "drop"
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  mutate(
    Placement = unname(placement_label[.data$placement]),
    placement_order = if_else(.data$placement == "near_eye", 1L, 2L),
    `Target/participant concurvity` = fmt_range(
      .data$concurvity_min,
      .data$concurvity_max,
      digits = 4L
    ),
    `Deviance explained` = paste0(
      fmt_number(100 * .data$deviance_min, 1L),
      "%–",
      fmt_number(100 * .data$deviance_max, 1L),
      "%"
    ),
    `Normal-reference QQ correlation` = fmt_range(
      .data$qq_min,
      .data$qq_max,
      digits = 3L
    )
  ) |>
  arrange(.data$placement_order)

site_display_lookup <- stats::setNames(
  site_registry$display_name,
  site_registry$site
)

display_omissions <- function(value) {
  vapply(
    value,
    function(one) {
      if (is.na(one) || one == "None") {
        return("None")
      }
      codes <- strsplit(one, "; ", fixed = TRUE)[[1L]]
      labels <- unname(site_display_lookup[codes])
      labels[is.na(labels)] <- codes[is.na(labels)]
      paste(labels, collapse = "; ")
    },
    character(1L)
  )
}

loso_table <- function(placement) {
  loso_results |>
    filter(.data$placement == .env$placement) |>
    arrange(.data$metric_order) |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      `Matching omissions` = paste0(
        .data$omissions_matching_main,
        "/",
        .data$omissions
      ),
      `Discordant site omissions` = display_omissions(
        .data$discordant_omissions
      ),
      `Transition range when present (h)` = if_else(
        is.na(.data$plateau_start_min),
        "—",
        fmt_range(.data$plateau_start_min, .data$plateau_start_max)
      )
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$`Matching omissions`,
      .data$`Discordant site omissions`,
      .data$`Transition range when present (h)`
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::cols_width(
      `Matching omissions` ~ gt::pct(14),
      `Discordant site omissions` ~ gt::pct(45),
      `Transition range when present (h)` ~ gt::pct(22)
    ) |>
    h07_gt(12)
}

near_patterns <- plateau |>
  filter(.data$placement == "near_eye", .data$revised_plateau_pattern)
chest_patterns <- plateau |>
  filter(.data$placement == "chest", .data$revised_plateau_pattern)

stopifnot(
  nrow(plateau) == 18L,
  nrow(near_patterns) == 6L,
  nrow(chest_patterns) == 7L,
  all(table(derivative_reader$placement, derivative_reader$metric_id) == 100L),
  nrow(sample_display) == 18L,
  all(sample_display$participant_days == sample_display$observations),
  nrow(curve_points) == 1872L,
  nrow(figure_settings) == 2L,
  all(adapted_diagnostics$converged),
  all(
    file.exists(
      file.path(root, figure_dir, figure_settings$filename)
    )
  ),
  identical(
    sensitivity_samples |>
      filter(grepl("^prep_common_primary__", .data$run_id)) |>
      arrange(.data$run_id, .data$metric_id) |>
      select(-.data$run_id, -.data$frame_sha256),
    sensitivity_samples |>
      filter(grepl("^prep_common_gap__", .data$run_id)) |>
      mutate(
        run_id = sub(
          "^prep_common_gap__",
          "prep_common_primary__",
          .data$run_id
        )
      ) |>
      arrange(.data$run_id, .data$metric_id) |>
      select(-.data$run_id, -.data$frame_sha256)
  )
)
```

## Scientific question

The preregistered hypothesis was:

> “H7: There is a ceiling (nonlinear) effect of absolute latitude and photoperiod on level-, duration-, and exposure-history-based light metrics.”

**Melanopic equivalent daylight illuminance (melEDI)** is an illuminance-like
measure weighted for melanopic photoreception and expressed in lux (lx). The
**primary near-eye sensor position** places the wearable sensor near the eyes;
the **complementary chest sensor position** is torso-worn and is not a measure
of ocular exposure.

This report asks the part of that question that these data can address: is a
pooled association between civil photoperiod and each planned personal
light-exposure metric visible, and does the fitted association change from a
locally supported increase to a slope that remains statistically compatible
with zero through the longest recorded photoperiod?

::: {.callout-note title="Answer in brief"}
A qualifying transition from increase to a sustained near-flat tail was
visible for **six of nine primary near-eye metrics**: mean melEDI, brightest
10 h mean, darkest 10 h mean, time above 1,000 lx melEDI, time above 250 lx
melEDI during wake, and melEDI dose. Their transition brackets ranged from
`r fmt_range(min(near_patterns$plateau_transition_lower), max(near_patterns$plateau_start))`
h of civil photoperiod. The complementary chest analysis showed the pattern
for **seven of nine metrics**: mean melEDI, darkest 10 h mean, both
above-threshold durations, time below 1 lx melEDI during sleep, the longest
continuous period above 250 lx melEDI, and melEDI dose. Their transition
brackets ranged from
`r fmt_range(min(chest_patterns$plateau_transition_lower), max(chest_patterns$plateau_start))`
h. These are visible pooled-data patterns under the qualifying-transition
rule, not evidence that a physiological or environmental ceiling exists.
No physiological or environmental ceiling was identified. Absolute latitude
cannot be separated from site in this design, and site, collection period,
incomplete photoperiod overlap, model form, and influential sites materially
limit interpretation.
:::

## Data, sensor positions, metrics, and support

All nine planned metrics were analysed independently of fitted results from
other hypotheses. Near-eye measurements are primary because they more closely
represent light near the eyes; chest measurements provide complementary
evidence. The placements were modelled separately. A common-sample sensitivity
uses the same participants and participant-days for both placements, but it
does not test or establish placement equivalence.

The primary dataset retains the approved metric-specific treatment of gaps.
A predefined sensitivity uses the **gap-timing-unaware dataset**. It still
passed the general 50%-per-hour and 80%-per-day coverage rules. The term means
that the timing of the remaining missing observations is not used for an
additional metric-specific adjustment; it does not mean that gaps,
missingness, or coverage were ignored. For this contrast only, the approved
primary preparation could be interpreted as a **time-sensitive primary metric
dataset**, because it uses the timing of remaining missing observations where
that timing is relevant to the metric. Below, it is called simply the primary
dataset.

A **participant-day** is one participant's available metric value on one local
calendar date at one sensor position. Repeated days were handled with a
participant random-effect smooth nested within site. Depending on the metric,
the primary near-eye analysis contains 139–141 participants and 655–816
participant-days, whereas the complementary chest analysis contains 153–154
participants and 743–902 participant-days. The recorded civil-photoperiod
support used by the metric-specific fits spans 10.33–20.52 h overall. The main
near-eye samples span 15 August 2023 to 19 October 2025; chest samples span 14
May 2024 to 19 October 2025. Exact metric-specific counts are reported with
each placement below.

```{r}
#| label: tbl-h07-response-specifications
#| tbl-cap: "Response specification for each planned personal light-exposure metric."

metric_registry |>
  mutate(
    Group = metric_group(.data$metric_order),
    Metric = .data$manuscript_name,
    Unit = .data$display_unit,
    `Model distribution` = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie with log link"
    ),
    `Response transformation` = recode(
      .data$response_transform,
      log10_offset_0.1 = "log10(metric + 0.1)",
      identity = "None"
    )
  ) |>
  select(
    .data$Group,
    .data$Metric,
    .data$Unit,
    .data$`Model distribution`,
    .data$`Response transformation`
  ) |>
  gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
  gt::cols_width(
    Unit ~ gt::pct(10),
    `Model distribution` ~ gt::pct(24),
    `Response transformation` ~ gt::pct(25)
  ) |>
  h07_gt(12)
```

Metric definitions and derivations are documented in
[Preparation 04](../preparation/04_metric_derivation.qmd).

## Reading guide

- A **nonlinear generalized additive model (GAM) analysis** allows the
  association between civil photoperiod and a light metric to bend rather than
  follow a straight line. The resulting line is called the fitted curve.
- The **derivative** is the slope of that fitted curve: how quickly the fitted
  outcome changes for each additional hour of photoperiod at each point. Here
  it is evaluated on the model's linear-predictor scale, in model-scale units
  per hour; the left-hand fitted curves are displayed in the named metric
  units.
- A **pointwise 95% confidence interval (95% CI)** describes uncertainty at
  one photoperiod value. It is not a simultaneous uncertainty band for the
  whole curve.
- **Overlap among nonlinear model terms** can make their separate
  contributions harder to distinguish. This is reported below using the
  technical term *concurvity*.

## Nonlinear photoperiod model and qualifying-transition rule

Each metric and placement used a separate nonlinear GAM fitted by restricted
maximum likelihood. The reported model was:

```{r}
#| label: code-h07-reported-formula
#| echo: true
#| output: asis

formula_registry |>
  filter(.data$model_id == "adapted_photoperiod_smooth") |>
  pull(.data$formula) |>
  cat()
```

Here, `response_value` is the response shown in the specification table,
`photoperiod_hours` is civil photoperiod, `site` is the study site, and
`site_participant` identifies a participant within site. The smooth used a
thin-plate regression spline with basis dimension *k* = 6. A broader *k* = 10
smooth and a model treating site as a fixed effect were fitted as model-form
sensitivities.

Absolute latitude is fixed within site. It is therefore structurally
confounded with site, and a latitude-by-photoperiod surface cannot separately
identify a latitude association while accounting for site. The planned joint
surface was inspected as an identifiability check but is not used for
inference or for the qualifying-transition classification. No temperature or
other unplanned environmental predictor was added.

For each reported smooth, the first derivative was evaluated at 100 equally
spaced values over that metric's exact recorded photoperiod range. Central
finite differences used a step of 0.01 h, with one-sided differences at the
two boundaries. Unconditional model-coefficient covariance produced
pointwise 95% confidence intervals on the model's linear-predictor scale.

A **qualifying transition from increase to a sustained near-flat tail**[^h07-formal-classification]
required all three conditions:

1. the immediately preceding grid point had a derivative interval wholly
   above zero;
2. the next grid point had a derivative interval containing zero; and
3. every later grid point through the longest recorded photoperiod also had a
   zero-compatible derivative interval.

The reported transition is the bracket from the last pointwise detected
increase to the first point of the sustained zero-compatible tail. This is the
requested change from a statistically supported positive derivative to a
non-significant derivative that remains non-significant to the end of the
recorded range. The rule does not define a scientifically negligible slope,
prove equivalence to zero, locate a true asymptote, or demonstrate a mechanism.
The 100 pointwise checks and the nine cross-metric classifications were not
multiplicity-adjusted; the classification is therefore descriptive.

[^h07-formal-classification]: The stored formal method label is
    **derivative-defined plateau pattern**. It denotes only the three-part
    pointwise classification above; it is not a biological-ceiling,
    permanent-plateau, or equivalence claim.

## Primary near-eye result

Depending on the metric, the primary near-eye models used 139–141
participants, 655–816 participant-days and observations, and all nine sites.
The modelled civil-photoperiod range was 10.33–20.52 h for every metric.

```{r}
#| label: tbl-h07-near-samples
#| tbl-cap: "Exact fitted near-eye sample for each personal light-exposure metric."

sample_table("near_eye")
```

The exact sample counts are available as
[source data](../../artifacts/09_tables/H07/H07_main_samples.csv).

How to read the main figure: the left panel in each row shows the fitted
quantity converted back to the displayed metric unit, and the right panel
shows its derivative on the model scale. Grey ribbons are pointwise 95% CIs,
not simultaneous bands. Blue derivative segments identify the qualifying
transition and sustained zero-compatible tail; the same region is marked by
tail shading and a dashed vertical transition line, providing non-colour
cues. The rugs show the observed photoperiod support. These marks classify the
fitted association; they do not identify a biological ceiling.

```{r}
#| label: fig-h07-near-smooth-derivative-pairs
#| fig-cap: "Primary near-eye associations with civil photoperiod. Each row pairs the fitted response-scale smooth (left) with its first derivative on the model scale (right). Grey ribbons are pointwise 95% confidence intervals; blue tail shading and a dashed transition line mark a qualifying transition and the sustained zero-compatible tail. Rugs show observed photoperiods."
#| fig-alt: "Nine paired rows show near-eye metric values against civil photoperiod on the left and first derivatives on the right. Mean melEDI, brightest and darkest 10-hour means, time above 1,000 lx, time above 250 lx during wake, and melEDI dose change from a positive derivative to intervals containing zero through the recorded maximum. Pre-sleep and sleep low-light durations have no preceding detected increase, while the longest continuous period above 250 lx remains increasing at the recorded maximum."
#| out-width: "100%"
#| fig-align: center
#| column: screen-inset

knitr::include_graphics(file.path(
  root,
  figure_dir,
  "H07_revised_smooth_derivative_pairs_near_eye.png"
))
```

The fitted smooths answer whether an association is visible in the pooled
data; the derivative panels apply the stated pattern rule. The figures do not
require every site to cover every photoperiod, but the rugs make the uneven
empirical support visible.

```{r}
#| label: tbl-h07-near-results
#| tbl-cap: "Primary near-eye qualifying-transition classifications and derivative estimates at the longest recorded photoperiod."

result_table("near_eye")
```

The endpoint derivative for the longest continuous period above 250 lx melEDI
remained positive, so this metric did not satisfy the pattern rule. Neither
low-light duration had a preceding interval with a pointwise detected positive
derivative. For time below 10 lx melEDI before sleep, the endpoint derivative
was negative rather than zero-compatible; this is not a positive-to-flat
pattern.

The plotted response-scale smooths and derivative calculations are available
as [smooth source data](../../artifacts/09_tables/H07/H07_main_curve_points.csv),
[derivative source data](../../artifacts/09_tables/H07/H07_revised_derivative_points.csv),
and [figure settings](../../artifacts/09_tables/H07/H07_revised_paired_figure_settings.csv).

## Complementary chest evidence

Depending on the metric, the chest models used 153–154 participants, 743–902
participant-days and observations, and eight sites. The modelled
civil-photoperiod range was again 10.33–20.52 h. Tübingen (DE) contributed
near-eye but not chest data, which is one reason the chest result is
complementary rather than a direct replication on an identical sample.

```{r}
#| label: tbl-h07-chest-samples
#| tbl-cap: "Exact fitted chest sample for each personal light-exposure metric."

sample_table("chest")
```

```{r}
#| label: fig-h07-chest-smooth-derivative-pairs
#| fig-cap: "Complementary chest associations with civil photoperiod. Each row pairs the fitted response-scale smooth (left) with its first derivative on the model scale (right). Grey ribbons are pointwise 95% confidence intervals; blue tail shading and a dashed transition line mark a qualifying transition and the sustained zero-compatible tail. Rugs show observed photoperiods."
#| fig-alt: "Nine paired rows show chest metric values against civil photoperiod on the left and first derivatives on the right. Mean melEDI, darkest 10-hour mean, time above 1,000 lx, time above 250 lx during wake, time below 1 lx during sleep, the longest continuous period above 250 lx, and melEDI dose change from a positive derivative to intervals containing zero through the recorded maximum. Brightest 10-hour mean remains increasing at the maximum, and pre-sleep low-light duration has no preceding detected increase."
#| out-width: "100%"
#| fig-align: center
#| column: screen-inset

knitr::include_graphics(file.path(
  root,
  figure_dir,
  "H07_revised_smooth_derivative_pairs_chest.png"
))
```

```{r}
#| label: tbl-h07-chest-results
#| tbl-cap: "Complementary chest qualifying-transition classifications and derivative estimates at the longest recorded photoperiod."

result_table("chest")
```

Brightest 10 h mean remained pointwise increasing at the recorded maximum.
Time below 10 lx melEDI before sleep again lacked a preceding pointwise
detected increase. The late transition for time below 1 lx melEDI during sleep
is especially uncertain because its response-distribution check failed, as
described below.

## Model checks

All 18 reported models converged. Gradients were small and all reported
Hessian minima were positive. **Residual autocorrelation** means that
observations close in sequence retain more similar residuals than observations
farther apart. Consecutive-day residual autocorrelation was modest in the
pooled descriptive check, with maximum absolute lag-1 correlations of
`r fmt_number(max(abs(adapted_diagnostics$pooled_consecutive_day_lag1[adapted_diagnostics$placement == "near_eye"])), 3)`
near eye and
`r fmt_number(max(abs(adapted_diagnostics$pooled_consecutive_day_lag1[adapted_diagnostics$placement == "chest"])), 3)`
at the chest. This does not prove temporal independence, but it did not
indicate a strong residual day-to-day process after participant clustering.

```{r}
#| label: tbl-h07-diagnostic-summary
#| tbl-cap: "Summary of convergence, smooth-basis, temporal, overlap among nonlinear terms (concurvity), and residual model checks."

diagnostic_summary |>
  select(
    .data$Placement,
    .data$Converged,
    .data$`Maximum absolute gradient`,
    .data$`Minimum Hessian eigenvalue`,
    .data$`Basis-check flags`,
    .data$`Maximum absolute lag-1`,
    .data$`Target/participant concurvity`,
    .data$`Deviance explained`,
    .data$`Normal-reference QQ correlation`
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_number(
    columns = c(
      `Maximum absolute gradient`,
      `Maximum absolute lag-1`
    ),
    decimals = 4
  ) |>
  gt::fmt_number(
    columns = `Minimum Hessian eigenvalue`,
    decimals = 5
  ) |>
  gt::fmt_integer(columns = `Basis-check flags`) |>
  gt::cols_label(
    `Basis-check flags` = gt::html("Basis flags<br><small>of 9</small>"),
    `Maximum absolute lag-1` = gt::html("Maximum |lag 1|"),
    `Target/participant concurvity` = gt::html(
      "Photoperiod/participant<br>concurvity"
    ),
    `Normal-reference QQ correlation` = gt::html(
      "QQ correlation<br>range"
    )
  ) |>
  gt::cols_width(
    Converged ~ gt::pct(10),
    `Basis-check flags` ~ gt::pct(10),
    `Target/participant concurvity` ~ gt::pct(18),
    `Deviance explained` ~ gt::pct(14),
    `Normal-reference QQ correlation` ~ gt::pct(16)
  ) |>
  h07_gt(12)
```

The basis-dimension check flagged three near-eye smooths and one chest smooth.
The expanded-basis sensitivity therefore carries more weight than a single
diagnostic threshold: it reproduced 16 of 17 evaluable classifications, with
one near-eye fit failing its Hessian check and the chest sleep low-light
classification changing.

Residual normal-reference correlations were generally high, but residuals
still showed fitted-value dependence and tail departures for some outcomes.
The Tweedie response checks were adequate for the two above-threshold duration
metrics. They were not adequate for time below 1 lx melEDI during sleep: each
placement had four observed zeros, the fitted model expected fewer than 0.0001
zeros, and none of 100 pilot simulations contained a zero. This distributional
mismatch makes the chest sleep low-light pattern particularly weak evidence;
no larger simulation can repair the response model's failure to reproduce the
observed zero mass.

The photoperiod smooth also had estimated overlap with the participant
random-effect smooth (concurvity) of approximately 0.9985–0.9988. This reflects
the clustered, repeated-day design and makes their separate contributions
harder to distinguish. It limits separation of between-person photoperiod
coverage from the pooled smooth. Smooth-term p-values are therefore not used
or displayed. The derivative intervals remain conditional on the fitted model
and should be read with this identifiability limitation.

## Sensitivity analyses

Each sensitivity changed a named feature. The gap-timing-unaware dataset
omitted the additional metric-specific adjustment based on when remaining
gaps occurred while retaining the coverage rules. The paired/common analysis
used only participant-days available at both sensor positions. The
cross-dataset common-row analyses restricted the primary and
gap-timing-unaware definitions to identical rows. The longest-period analysis
retained only days with an exactly identified period, and the dose analysis
compared observed and corrected melEDI dose on identical rows. Model-form
sensitivities increased photoperiod-smooth flexibility from *k* = 6 to *k* =
10 or represented site with fixed rather than random effects.
Leave-one-site-out analyses omitted one named site at a time. Agreement below
concerns the yes/no qualifying-transition classification, not an identical
transition point.

The gap-timing-unaware dataset reproduced all nine near-eye classifications
and eight of nine chest classifications; the chest melEDI-dose pattern was the
exception. The paired common sample used 110–112 participants, 505–643
participant-days and observations, and eight sites. It reproduced four of nine
near-eye and six of nine chest classifications, showing that several pooled
patterns depend on which participant-days and sites contribute rather than on
placement alone.

```{r}
#| label: tbl-h07-sensitivity-samples
#| tbl-cap: "Fitted sample sizes for the main data and metric-definition sensitivities. Ranges indicate metric-specific counts."

sensitivity_sample_summary |>
  select(
    .data$Scenario,
    .data$Placement,
    .data$Participants,
    .data$`Participant-days`,
    .data$Observations,
    .data$Sites
  ) |>
  gt::gt(groupname_col = "Scenario") |>
  gt::cols_width(
    Placement ~ gt::pct(16),
    Participants ~ gt::pct(18),
    `Participant-days` ~ gt::pct(21),
    Observations ~ gt::pct(18),
    Sites ~ gt::pct(10)
  ) |>
  gt::tab_source_note(
    gt::md(
      paste0(
        "Preparation exact-common samples were identical for the primary ",
        "and gap-timing-unaware datasets. Observed and corrected dose ",
        "definitions used the same common rows."
      )
    )
  ) |>
  h07_gt(12)
```

```{r}
#| label: tbl-h07-sensitivity-classifications
#| tbl-cap: "Agreement of sensitivity qualifying-transition classifications with the reported pooled photoperiod analysis."

sensitivity_summary |>
  select(
    .data$Scenario,
    .data$Placement,
    .data$Agreement,
    .data$`Changed classifications`
  ) |>
  gt::gt(groupname_col = "Scenario") |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    Agreement ~ gt::pct(12),
    `Changed classifications` ~ gt::pct(56)
  ) |>
  gt::tab_source_note(
    gt::md(
      paste0(
        "Agreement concerns the yes/no qualifying-transition classification only; transition ",
        "brackets can move even when the classification agrees."
      )
    )
  ) |>
  h07_gt(12)
```

Restricting the longest continuous period above 250 lx melEDI to days whose
period duration was exactly identified changed the near-eye classification
from no pattern to a pattern at 15.38–15.48 h; the chest classification was
unchanged. Comparing the observed and corrected melEDI-dose definitions on
identical rows left both placement classifications unchanged. Restricting the
two preparation variants to exact common rows preserved every near-eye
classification but removed the chest sleep low-light pattern in both variants.

```{r}
#| label: tbl-h07-model-form-sensitivities
#| tbl-cap: "Qualifying-transition stability under broader photoperiod smooths and fixed site effects."

model_form_summary |>
  select(
    .data$Model,
    .data$Placement,
    .data$Agreement,
    .data$`Failed fits`,
    .data$`Changed classifications`
  ) |>
  gt::gt(groupname_col = "Model") |>
  gt::fmt_integer(columns = `Failed fits`) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    Agreement ~ gt::pct(12),
    `Failed fits` ~ gt::pct(12),
    `Changed classifications` ~ gt::pct(48)
  ) |>
  h07_gt(12)
```

The broader smooth reproduced eight of eight evaluable near-eye and eight of
nine chest classifications; the near-eye longest-period fit failed its Hessian
check, and the chest sleep low-light pattern disappeared. Treating site as a
fixed effect reproduced eight of nine classifications at each placement: the
near-eye longest-period pattern appeared, while the chest sleep low-light
pattern disappeared. These changes reinforce that both metrics sit near a
classification boundary.

## Site influence and support

Omitting one site at a time did not act as a new eligibility rule; it tested
whether a pooled pattern depended strongly on one site's observations. Near-eye
mean melEDI and both low-light metrics reproduced their main classification in
all nine omissions. Five positive near-eye patterns changed only when
Tübingen (DE) was omitted. The near-eye longest-period classification changed
in four of nine omissions. Chest results were less stable for several
outcomes, most notably time below 1 lx melEDI during sleep, which matched the
main classification in only three of eight omissions.

```{r}
#| label: tbl-h07-near-loso
#| tbl-cap: "Leave-one-site-out influence on primary near-eye qualifying-transition classifications."

loso_table("near_eye")
```

```{r}
#| label: tbl-h07-chest-loso
#| tbl-cap: "Leave-one-site-out influence on complementary chest qualifying-transition classifications."

loso_table("chest")
```

The observed civil-photoperiod range was broad overall but uneven across
sites. For example, near-eye observations ranged from 10.33 to 12.35 h in
Madrid (ES), 16.41 to 17.45 h in Munich (DE), and 12.49 to 12.71 h in
Kumasi (GH); San José (CR) contributed only a narrow range around 13.47 h,
while Borås (SE) covered values up to 20.52 h. Sites were also measured in
different collection periods and, in some cases, different years. Thus the
pooled smooth can answer whether an association is visible in this dataset,
but it cannot separate photoperiod from every site-specific or
collection-period difference. Leave-one-site-out changes show where this
limitation affects the descriptive classification.

## Interpretation and limitations

The pooled analysis shows several fitted photoperiod associations that rise
and then meet the prespecified pointwise rule for a sustained
zero-compatible derivative. The recurring patterns for mean melEDI, darkest
10 h mean, both above-threshold duration metrics, and melEDI dose across the
two placements are the most consistent descriptive findings. Brightest 10 h
mean differs by placement, the longest-period result is sensitive to sample
definition and model form, and the chest sleep low-light result is undermined
by distributional failure, common-row analyses, model-form changes, and site
influence.

The strongest defensible conclusion is therefore that **qualifying
transitions from increase to a sustained near-flat tail are visible for
several personal light-exposure metrics in the pooled data**, subject to
substantial design and model limitations. The results do not establish a
ceiling, an equivalence region, a causal effect of photoperiod, a distinct
latitude effect, a health or clinical consequence, or equivalence between
near-eye and chest placement.

## Preregistration deviations {#h07-preregistration-deviations}

- [`DEV-016`](../preregistration_deviations.qmd#dev-016): the registered
  latitude-by-photoperiod surface is non-identifiable here, so the accepted
  report retains all nine outcomes and presents a separately labelled
  site-adjusted nonlinear photoperiod analysis as descriptive evidence.
- [`DEV-033`](../preregistration_deviations.qmd#dev-033): all nine eligible
  outcomes remain visible rather than being selected using H01 significance
  or H07 AIC.
- [`DEV-034`](../preregistration_deviations.qmd#dev-034): response
  specifications and model checks are metric-specific rather than one
  universal zero-aware log-Gaussian model.

## Source data and technical provenance

Exact input identities, metric-specific model frames, evaluated formulas,
model-check provenance, code order, and output identities are documented in
the [H07 analysis preparation and provenance](../../audit/hypotheses/H07/H07_analysis_preparation.qmd).
Metric definitions are available in
[Preparation 04](../preparation/04_metric_derivation.qmd); the plotted smooth,
derivative, and figure-setting source files are linked beside the relevant
outputs above.

Models were fitted with R `r session_record$r_version[[1]]`, mgcv
`r session_record$mgcv_version[[1]]`, and gratia
`r session_record$gratia_version[[1]]`.

```{r}
#| label: session-h07-reader
#| include: false

sessionInfo()
```
