---
title: "H07: Photoperiod and possible plateaus in personal light exposure"
subtitle: "Pooled photoperiod associations across nine personal light-exposure metrics"
format:
  html:
    toc: true
    toc-depth: 2
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h07-reader

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(tibble)
  library(tidyr)
})

root <- normalizePath(
  Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()),
  winslash = "/",
  mustWork = TRUE
)
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

read_h07 <- function(...) {
  readr::read_csv(file.path(root, ...), show_col_types = FALSE)
}

table_dir <- c("artifacts", "09_tables", "H07")
figure_dir <- file.path("artifacts", "08_figures", "H07")

main_samples <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_samples.csv"
)
plateau_all <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_plateau_summary.csv"
)
derivative_points <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_derivative_points.csv"
)
curve_points <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_curve_points.csv"
)
formula_registry <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_formula_registry.csv"
)
diagnostics <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_diagnostics.csv"
)
pairwise_concurvity <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_pairwise_concurvity.csv"
)
residual_summary <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_residual_distribution_summary.csv"
)
tweedie_pilot <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_tweedie_distribution_pilot.csv"
)
sensitivity_results <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_sensitivity_plateau_comparison.csv"
)
sensitivity_samples <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_sensitivity_samples.csv"
)
model_form_results <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_model_form_plateau_comparison.csv"
)
loso_results <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_loso_plateau_influence_summary.csv"
)
site_support <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_main_site_support.csv"
)
figure_settings <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_revised_paired_figure_settings.csv"
)
session_record <- read_h07(
  table_dir[[1]], table_dir[[2]], table_dir[[3]],
  "H07_session.csv"
)
site_registry <- read_h07("config", "site_display_registry.csv") |>
  arrange(.data$display_order)

derivative_method <- "REVISED_CENTRAL_UNCONDITIONAL_POINTWISE"
plateau <- plateau_all |>
  filter(.data$method_id == .env$derivative_method) |>
  arrange(.data$placement, .data$metric_order)
derivative_reader <- derivative_points |>
  filter(.data$method_id == .env$derivative_method)

metric_registry <- plateau |>
  distinct(
    .data$metric_id,
    .data$metric_order,
    .data$manuscript_name,
    .data$display_unit,
    .data$response_family,
    .data$response_transform
  ) |>
  arrange(.data$metric_order)

sample_display <- main_samples |>
  mutate(placement = sub("^primary__", "", .data$run_id)) |>
  left_join(
    metric_registry |>
      select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  arrange(.data$placement, .data$metric_order)

adapted_diagnostics <- diagnostics |>
  filter(.data$model_id == "adapted_photoperiod_smooth") |>
  mutate(placement = sub("^primary__", "", .data$run_id)) |>
  left_join(
    metric_registry |>
      select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name
      ),
    by = "metric_id",
    relationship = "many-to-one"
  )

target_concurvity <- pairwise_concurvity |>
  filter(
    .data$model_id == "adapted_photoperiod_smooth",
    .data$measure == "estimate",
    .data$supplier_term == "s(site_participant)",
    .data$target_term == "s(photoperiod_hours)"
  )

placement_label <- c(near_eye = "Near eye", chest = "Chest")

fmt_number <- function(value, digits = 2L) {
  formatC(value, digits = digits, format = "f", big.mark = ",")
}

fmt_range <- function(lower, upper, digits = 2L, suffix = "") {
  paste0(
    fmt_number(lower, digits), "–", fmt_number(upper, digits), suffix
  )
}

fmt_ci <- function(estimate, lower, upper, digits = 3L) {
  paste0(
    fmt_number(estimate, digits), " (95% CI ",
    fmt_number(lower, digits), " to ", fmt_number(upper, digits), ")"
  )
}

metric_group <- function(metric_order) {
  dplyr::case_when(
    metric_order <= 3 ~ "Light level",
    metric_order <= 8 ~ "Duration and continuous period",
    TRUE ~ "Exposure history"
  )
}

h07_gt <- function(table, font_size = 13) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(5),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

sample_table <- function(placement) {
  sample_display |>
    filter(.data$placement == .env$placement) |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      Participants = .data$participants,
      `Participant-days` = .data$participant_days,
      Observations = .data$observations,
      Sites = .data$sites,
      `Civil photoperiod (h)` = fmt_range(
        .data$photoperiod_min,
        .data$photoperiod_max
      )
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$Participants,
      .data$`Participant-days`,
      .data$Observations,
      .data$Sites,
      .data$`Civil photoperiod (h)`
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::fmt_integer(
      columns = c(Participants, `Participant-days`, Observations, Sites)
    ) |>
    gt::cols_width(
      Participants ~ gt::pct(12),
      `Participant-days` ~ gt::pct(14),
      Observations ~ gt::pct(12),
      Sites ~ gt::pct(8),
      `Civil photoperiod (h)` ~ gt::pct(18)
    ) |>
    h07_gt(12)
}

result_reason <- function(disposition) {
  dplyr::case_when(
    disposition == "DERIVATIVE_DEFINED_PLATEAU_PATTERN" ~
      "Positive derivative followed by a zero-compatible tail",
    disposition == "NO_POINTWISE_DETECTED_PRIOR_INCREASE" ~
      "No preceding pointwise detected increase",
    disposition == "POINTWISE_DETECTED_INCREASE_AT_RECORDED_END" ~
      "Increase remained pointwise detected at the recorded maximum",
    TRUE ~ disposition
  )
}

result_table <- function(placement) {
  plateau |>
    filter(.data$placement == .env$placement) |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      Pattern = if_else(.data$revised_plateau_pattern, "Yes", "No"),
      `Transition bracket (h)` = if_else(
        .data$revised_plateau_pattern,
        fmt_range(
          .data$plateau_transition_lower,
          .data$plateau_start
        ),
        "—"
      ),
      `Derivative at maximum` = fmt_ci(
        .data$derivative_at_recorded_end,
        .data$derivative_at_recorded_end_lower,
        .data$derivative_at_recorded_end_upper
      ),
      Interpretation = result_reason(.data$disposition)
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$Pattern,
      .data$`Transition bracket (h)`,
      .data$`Derivative at maximum`,
      .data$Interpretation
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::tab_style(
      style = gt::cell_fill(color = "#E8F3F8"),
      locations = gt::cells_body(
        columns = Pattern,
        rows = Pattern == "Yes"
      )
    ) |>
    gt::cols_width(
      Pattern ~ gt::pct(8),
      `Transition bracket (h)` ~ gt::pct(16),
      `Derivative at maximum` ~ gt::pct(25),
      Interpretation ~ gt::pct(31)
    ) |>
    gt::tab_source_note(
      gt::md(
        paste0(
          "Derivatives and their 95% confidence intervals are on each ",
          "model's linear-predictor scale, per hour of civil photoperiod."
        )
      )
    ) |>
    h07_gt(12)
}

sensitivity_label <- function(sensitivity, data_scenario) {
  dplyr::case_when(
    sensitivity == "paired_common_placement" ~
      "Paired near-eye/chest common sample",
    sensitivity == "gap_timing_unaware" ~
      "Gap-timing-unaware dataset",
    sensitivity == "preparation_exact_common" &
      data_scenario == "primary" ~
      "Preparation exact-common rows: primary dataset",
    sensitivity == "preparation_exact_common" &
      data_scenario == "gap_timing_unaware" ~
      "Preparation exact-common rows: gap-timing-unaware dataset",
    sensitivity == "longest_period_exact_only" ~
      "Exactly identified longest-period rows",
    sensitivity == "observed_dose_common_sample" ~
      "Observed/corrected dose on common rows",
    TRUE ~ sensitivity
  )
}

sensitivity_summary <- sensitivity_results |>
  group_by(.data$data_scenario, .data$placement, .data$sensitivity) |>
  summarise(
    Evaluable = sum(!is.na(.data$classification_agrees)),
    Agreements = sum(.data$classification_agrees, na.rm = TRUE),
    changed = paste(
      unique(
        .data$manuscript_name[
          !is.na(.data$classification_agrees) &
            !.data$classification_agrees
        ]
      ),
      collapse = "; "
    ),
    .groups = "drop"
  ) |>
  mutate(
    Scenario = sensitivity_label(.data$sensitivity, .data$data_scenario),
    Placement = unname(placement_label[.data$placement]),
    Agreement = paste0(.data$Agreements, "/", .data$Evaluable),
    `Changed classifications` = if_else(
      .data$changed == "",
      "None",
      .data$changed
    ),
    sensitivity_order = match(
      .data$sensitivity,
      c(
        "gap_timing_unaware",
        "paired_common_placement",
        "preparation_exact_common",
        "longest_period_exact_only",
        "observed_dose_common_sample"
      )
    ),
    scenario_order = if_else(
      .data$data_scenario == "primary",
      1L,
      2L
    ),
    placement_order = if_else(.data$placement == "near_eye", 1L, 2L)
  ) |>
  arrange(
    .data$sensitivity_order,
    .data$scenario_order,
    .data$placement_order
  )

sensitivity_sample_labels <- tibble::tribble(
  ~run_id, ~Scenario, ~placement_order, ~scenario_order,
  "gap_timing_unaware__near_eye", "Gap-timing-unaware dataset", 1L, 1L,
  "gap_timing_unaware__chest", "Gap-timing-unaware dataset", 2L, 1L,
  "paired__near_eye", "Paired near-eye/chest common sample", 1L, 2L,
  "paired__chest", "Paired near-eye/chest common sample", 2L, 2L,
  "prep_common_primary__near_eye", "Preparation exact-common rows", 1L, 3L,
  "prep_common_primary__chest", "Preparation exact-common rows", 2L, 3L,
  "exact_period__near_eye", "Exactly identified longest-period rows", 1L, 4L,
  "exact_period__chest", "Exactly identified longest-period rows", 2L, 4L,
  "dose_common_corrected__near_eye", "Observed/corrected dose common rows", 1L, 5L,
  "dose_common_corrected__chest", "Observed/corrected dose common rows", 2L, 5L
)

range_text <- function(value) {
  if (min(value) == max(value)) {
    return(formatC(min(value), format = "d", big.mark = ","))
  }
  paste0(
    formatC(min(value), format = "d", big.mark = ","),
    "–",
    formatC(max(value), format = "d", big.mark = ",")
  )
}

sensitivity_sample_summary <- sensitivity_samples |>
  inner_join(sensitivity_sample_labels, by = "run_id") |>
  group_by(
    .data$run_id,
    .data$Scenario,
    .data$placement_order,
    .data$scenario_order
  ) |>
  summarise(
    Participants = range_text(.data$participants),
    `Participant-days` = range_text(.data$participant_days),
    Observations = range_text(.data$observations),
    Sites = range_text(.data$sites),
    .groups = "drop"
  ) |>
  mutate(
    Placement = if_else(.data$placement_order == 1L, "Near eye", "Chest")
  ) |>
  arrange(.data$scenario_order, .data$placement_order)

model_form_summary <- model_form_results |>
  group_by(.data$placement, .data$model_id) |>
  summarise(
    Evaluable = sum(!is.na(.data$classification_agrees)),
    Agreements = sum(.data$classification_agrees, na.rm = TRUE),
    failures = sum(grepl("^FAIL", .data$fit_status)),
    changed = paste(
      unique(
        .data$manuscript_name[
          !is.na(.data$classification_agrees) &
            !.data$classification_agrees
        ]
      ),
      collapse = "; "
    ),
    .groups = "drop"
  ) |>
  mutate(
    Model = recode(
      .data$model_id,
      adapted_photoperiod_expanded_basis = "Expanded smooth basis (k = 10)",
      adapted_photoperiod_fixed_site = "Site as a fixed effect"
    ),
    Placement = unname(placement_label[.data$placement]),
    Agreement = paste0(.data$Agreements, "/", .data$Evaluable),
    `Failed fits` = .data$failures,
    `Changed classifications` = if_else(
      .data$changed == "",
      "None",
      .data$changed
    ),
    placement_order = if_else(.data$placement == "near_eye", 1L, 2L)
  ) |>
  arrange(.data$Model, .data$placement_order)

diagnostic_summary <- adapted_diagnostics |>
  group_by(.data$placement) |>
  summarise(
    Converged = paste0(sum(.data$converged), "/", dplyr::n()),
    `Maximum absolute gradient` = max(.data$gradient_max_abs),
    `Minimum Hessian eigenvalue` = min(.data$hessian_min_eigenvalue),
    `Basis-check flags` = sum(.data$k_p_value < 0.05, na.rm = TRUE),
    `Maximum absolute lag-1` = max(
      abs(.data$pooled_consecutive_day_lag1),
      na.rm = TRUE
    ),
    .groups = "drop"
  ) |>
  left_join(
    target_concurvity |>
      group_by(.data$placement) |>
      summarise(
        concurvity_min = min(.data$concurvity),
        concurvity_max = max(.data$concurvity),
        .groups = "drop"
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  left_join(
    residual_summary |>
      group_by(.data$placement) |>
      summarise(
        deviance_min = min(.data$deviance_explained),
        deviance_max = max(.data$deviance_explained),
        qq_min = min(.data$normal_qq_correlation),
        qq_max = max(.data$normal_qq_correlation),
        .groups = "drop"
      ),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  mutate(
    Placement = unname(placement_label[.data$placement]),
    placement_order = if_else(.data$placement == "near_eye", 1L, 2L),
    `Target/participant concurvity` = fmt_range(
      .data$concurvity_min,
      .data$concurvity_max,
      digits = 4L
    ),
    `Deviance explained` = paste0(
      fmt_number(100 * .data$deviance_min, 1L),
      "%–",
      fmt_number(100 * .data$deviance_max, 1L),
      "%"
    ),
    `Normal-reference QQ correlation` = fmt_range(
      .data$qq_min,
      .data$qq_max,
      digits = 3L
    )
  ) |>
  arrange(.data$placement_order)

site_display_lookup <- stats::setNames(
  site_registry$display_name,
  site_registry$site
)

display_omissions <- function(value) {
  vapply(
    value,
    function(one) {
      if (is.na(one) || one == "None") {
        return("None")
      }
      codes <- strsplit(one, "; ", fixed = TRUE)[[1L]]
      labels <- unname(site_display_lookup[codes])
      labels[is.na(labels)] <- codes[is.na(labels)]
      paste(labels, collapse = "; ")
    },
    character(1L)
  )
}

loso_table <- function(placement) {
  loso_results |>
    filter(.data$placement == .env$placement) |>
    arrange(.data$metric_order) |>
    mutate(
      Group = metric_group(.data$metric_order),
      Metric = .data$manuscript_name,
      `Matching omissions` = paste0(
        .data$omissions_matching_main,
        "/",
        .data$omissions
      ),
      `Discordant site omissions` = display_omissions(
        .data$discordant_omissions
      ),
      `Transition range when present (h)` = if_else(
        is.na(.data$plateau_start_min),
        "—",
        fmt_range(.data$plateau_start_min, .data$plateau_start_max)
      )
    ) |>
    select(
      .data$Group,
      .data$Metric,
      .data$`Matching omissions`,
      .data$`Discordant site omissions`,
      .data$`Transition range when present (h)`
    ) |>
    gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
    gt::cols_width(
      `Matching omissions` ~ gt::pct(14),
      `Discordant site omissions` ~ gt::pct(45),
      `Transition range when present (h)` ~ gt::pct(22)
    ) |>
    h07_gt(12)
}

near_patterns <- plateau |>
  filter(.data$placement == "near_eye", .data$revised_plateau_pattern)
chest_patterns <- plateau |>
  filter(.data$placement == "chest", .data$revised_plateau_pattern)

stopifnot(
  nrow(plateau) == 18L,
  nrow(near_patterns) == 6L,
  nrow(chest_patterns) == 7L,
  all(table(derivative_reader$placement, derivative_reader$metric_id) == 100L),
  nrow(sample_display) == 18L,
  all(sample_display$participant_days == sample_display$observations),
  nrow(curve_points) == 1872L,
  nrow(figure_settings) == 2L,
  all(adapted_diagnostics$converged),
  all(
    file.exists(
      file.path(root, figure_dir, figure_settings$filename)
    )
  ),
  identical(
    sensitivity_samples |>
      filter(grepl("^prep_common_primary__", .data$run_id)) |>
      arrange(.data$run_id, .data$metric_id) |>
      select(-.data$run_id, -.data$frame_sha256),
    sensitivity_samples |>
      filter(grepl("^prep_common_gap__", .data$run_id)) |>
      mutate(
        run_id = sub(
          "^prep_common_gap__",
          "prep_common_primary__",
          .data$run_id
        )
      ) |>
      arrange(.data$run_id, .data$metric_id) |>
      select(-.data$run_id, -.data$frame_sha256)
  )
)
```

## Question

The preregistered hypothesis was:

> “H7: There is a ceiling (nonlinear) effect of absolute latitude and photoperiod on level-, duration-, and exposure-history-based light metrics.”

This report asks the part of that question that these data can address: is a
pooled association between civil photoperiod and each planned personal
light-exposure metric visible, and does the fitted association change from a
pointwise detected increase to a derivative that remains statistically
compatible with zero through the longest recorded photoperiod?

::: {.callout-note title="Answer in brief"}
A derivative-defined plateau pattern was visible for **six of nine primary
near-eye metrics**: mean melEDI, brightest 10 h mean, darkest 10 h mean, time
above 1,000 lx melEDI, time above 250 lx melEDI during wake, and melEDI dose.
Their transition brackets ranged from
`r fmt_range(min(near_patterns$plateau_transition_lower), max(near_patterns$plateau_start))`
h of civil photoperiod. The complementary chest analysis showed the pattern
for **seven of nine metrics**: mean melEDI, darkest 10 h mean, both
above-threshold durations, time below 1 lx melEDI during sleep, the longest
continuous period above 250 lx melEDI, and melEDI dose. Their transition
brackets ranged from
`r fmt_range(min(chest_patterns$plateau_transition_lower), max(chest_patterns$plateau_start))`
h. These are visible pooled-data patterns under a pointwise derivative rule,
not evidence that a physiological or environmental ceiling exists. Absolute
latitude cannot be separated from site in this design, and site, collection
period, incomplete photoperiod overlap, model form, and influential sites
materially limit interpretation.
:::

## What was analysed

All nine planned metrics were analysed independently of fitted results from
other hypotheses. Near-eye measurements are primary because they more closely
represent light near the eyes; chest measurements provide complementary
evidence. The placements were modelled separately. A common-sample sensitivity
uses the same participants and participant-days for both placements, but it
does not test or establish placement equivalence.

The primary dataset retains the approved metric-specific treatment of gaps.
A predefined sensitivity uses the **gap-timing-unaware dataset**. It still
passed the general 50%-per-hour and 80%-per-day coverage rules. The term means
that the timing of the remaining missing observations is not used for an
additional metric-specific adjustment; it does not mean that gaps,
missingness, or coverage were ignored. For this contrast only, the approved
primary preparation could be interpreted as a **time-sensitive primary metric
dataset**, because it uses the timing of remaining missing observations where
that timing is relevant to the metric. Below, it is called simply the primary
dataset.

The analysis unit was a participant-day, with one observation for each
available metric, participant, day, and placement. Repeated days were handled
with a participant random-effect smooth nested within site. The main near-eye
samples span 15 August 2023 to 19 October 2025; chest samples span 14 May 2024
to 19 October 2025. Exact metric-specific counts are reported with each
placement below.

```{r}
#| label: tbl-h07-response-specifications
#| tbl-cap: "Response specification for each planned personal light-exposure metric."

metric_registry |>
  mutate(
    Group = metric_group(.data$metric_order),
    Metric = .data$manuscript_name,
    Unit = .data$display_unit,
    `Model distribution` = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie with log link"
    ),
    `Response transformation` = recode(
      .data$response_transform,
      log10_offset_0.1 = "log10(metric + 0.1)",
      identity = "None"
    )
  ) |>
  select(
    .data$Group,
    .data$Metric,
    .data$Unit,
    .data$`Model distribution`,
    .data$`Response transformation`
  ) |>
  gt::gt(rowname_col = "Metric", groupname_col = "Group") |>
  gt::cols_width(
    Unit ~ gt::pct(10),
    `Model distribution` ~ gt::pct(24),
    `Response transformation` ~ gt::pct(25)
  ) |>
  h07_gt(12)
```

## Model and plateau rule

Each metric and placement used a separate generalized additive model fitted by
restricted maximum likelihood. The reported model was:

```{r}
#| label: code-h07-reported-formula
#| echo: true
#| output: asis

formula_registry |>
  filter(.data$model_id == "adapted_photoperiod_smooth") |>
  pull(.data$formula) |>
  cat()
```

Here, `response_value` is the response shown in the specification table,
`photoperiod_hours` is civil photoperiod, `site` is the study site, and
`site_participant` identifies a participant within site. The smooth used a
thin-plate regression spline with basis dimension *k* = 6. A broader *k* = 10
smooth and a model treating site as a fixed effect were fitted as model-form
sensitivities.

Absolute latitude is fixed within site. It is therefore structurally
confounded with site, and a latitude-by-photoperiod surface cannot separately
identify a latitude association while accounting for site. The planned joint
surface was inspected as an identifiability check but is not used for
inference or for the plateau classification. No temperature or other
unplanned environmental predictor was added.

For each reported smooth, the first derivative was evaluated at 100 equally
spaced values over that metric's exact recorded photoperiod range. Central
finite differences used a step of 0.01 h, with one-sided differences at the
two boundaries. Unconditional model-coefficient covariance produced
pointwise 95% confidence intervals on the model's linear-predictor scale.

A **derivative-defined plateau pattern** required all three conditions:

1. the lower pointwise 95% confidence limit for the derivative was above zero
   at one or more photoperiod values;
2. a later grid point had a 95% confidence interval containing zero; and
3. every subsequent interval through the longest recorded photoperiod also
   contained zero.

The reported transition is the bracket from the last pointwise detected
increase to the first point of the sustained zero-compatible tail. This is the
requested change from a statistically supported positive derivative to a
non-significant derivative that remains non-significant to the end of the
recorded range. The rule does not define a scientifically negligible slope,
prove equivalence to zero, locate a true asymptote, or demonstrate a mechanism.
The 100 pointwise checks and the nine cross-metric classifications were not
multiplicity-adjusted; the classification is therefore descriptive.

## Primary near-eye result

Depending on the metric, the primary near-eye models used 139–141
participants, 655–816 participant-days and observations, and all nine sites.
The modelled civil-photoperiod range was 10.33–20.52 h for every metric.

```{r}
#| label: tbl-h07-near-samples
#| tbl-cap: "Exact fitted near-eye sample for each personal light-exposure metric."

sample_table("near_eye")
```

The exact sample counts are available as
[source data](../../artifacts/09_tables/H07/H07_main_samples.csv).

```{r}
#| label: fig-h07-near-smooth-derivative-pairs
#| fig-cap: "Primary near-eye associations with civil photoperiod. Each row pairs the fitted response-scale smooth (left) with its first derivative on the model scale (right). Grey ribbons are pointwise 95% confidence intervals; blue derivative segments mark a qualifying transition and the sustained zero-compatible tail. Rugs show observed photoperiods."
#| fig-alt: "Nine paired rows show near-eye metric values against civil photoperiod on the left and first derivatives on the right. Mean melEDI, brightest and darkest 10-hour means, time above 1,000 lx, time above 250 lx during wake, and melEDI dose change from a positive derivative to intervals containing zero through the recorded maximum. Pre-sleep and sleep low-light durations have no preceding detected increase, while the longest continuous period above 250 lx remains increasing at the recorded maximum."
#| out-width: "100%"
#| fig-align: center
#| column: screen-inset

knitr::include_graphics(file.path(
  root,
  figure_dir,
  "H07_revised_smooth_derivative_pairs_near_eye.png"
))
```

The fitted smooths answer whether an association is visible in the pooled
data; the derivative panels apply the stated pattern rule. The figures do not
require every site to cover every photoperiod, but the rugs make the uneven
empirical support visible.

```{r}
#| label: tbl-h07-near-results
#| tbl-cap: "Near-eye derivative-defined plateau classifications and derivative estimates at the longest recorded photoperiod."

result_table("near_eye")
```

The endpoint derivative for the longest continuous period above 250 lx melEDI
remained positive, so this metric did not satisfy the pattern rule. Neither
low-light duration had a preceding interval with a pointwise detected positive
derivative. For time below 10 lx melEDI before sleep, the endpoint derivative
was negative rather than zero-compatible; this is not a positive-to-flat
pattern.

The plotted response-scale smooths and derivative calculations are available
as [smooth source data](../../artifacts/09_tables/H07/H07_main_curve_points.csv),
[derivative source data](../../artifacts/09_tables/H07/H07_revised_derivative_points.csv),
and [figure settings](../../artifacts/09_tables/H07/H07_revised_paired_figure_settings.csv).

## Complementary chest result

Depending on the metric, the chest models used 153–154 participants, 743–902
participant-days and observations, and eight sites. The modelled
civil-photoperiod range was again 10.33–20.52 h. Tübingen contributed near-eye
but not chest data, which is one reason the chest result is complementary
rather than a direct replication on an identical sample.

```{r}
#| label: tbl-h07-chest-samples
#| tbl-cap: "Exact fitted chest sample for each personal light-exposure metric."

sample_table("chest")
```

```{r}
#| label: fig-h07-chest-smooth-derivative-pairs
#| fig-cap: "Complementary chest associations with civil photoperiod. Each row pairs the fitted response-scale smooth (left) with its first derivative on the model scale (right). Grey ribbons are pointwise 95% confidence intervals; blue derivative segments mark a qualifying transition and the sustained zero-compatible tail. Rugs show observed photoperiods."
#| fig-alt: "Nine paired rows show chest metric values against civil photoperiod on the left and first derivatives on the right. Mean melEDI, darkest 10-hour mean, time above 1,000 lx, time above 250 lx during wake, time below 1 lx during sleep, the longest continuous period above 250 lx, and melEDI dose change from a positive derivative to intervals containing zero through the recorded maximum. Brightest 10-hour mean remains increasing at the maximum, and pre-sleep low-light duration has no preceding detected increase."
#| out-width: "100%"
#| fig-align: center
#| column: screen-inset

knitr::include_graphics(file.path(
  root,
  figure_dir,
  "H07_revised_smooth_derivative_pairs_chest.png"
))
```

```{r}
#| label: tbl-h07-chest-results
#| tbl-cap: "Chest derivative-defined plateau classifications and derivative estimates at the longest recorded photoperiod."

result_table("chest")
```

Brightest 10 h mean remained pointwise increasing at the recorded maximum.
Time below 10 lx melEDI before sleep again lacked a preceding pointwise
detected increase. The late transition for time below 1 lx melEDI during sleep
is especially uncertain because its response-distribution check failed, as
described below.

## Model checks

All 18 reported models converged. Gradients were small and all reported
Hessian minima were positive. Consecutive-day residual dependence was modest
in the pooled descriptive check, with maximum absolute lag-1 correlations of
`r fmt_number(max(abs(adapted_diagnostics$pooled_consecutive_day_lag1[adapted_diagnostics$placement == "near_eye"])), 3)`
near eye and
`r fmt_number(max(abs(adapted_diagnostics$pooled_consecutive_day_lag1[adapted_diagnostics$placement == "chest"])), 3)`
at the chest. This does not prove temporal independence, but it did not
indicate a strong residual day-to-day process after participant clustering.

```{r}
#| label: tbl-h07-diagnostic-summary
#| tbl-cap: "Summary of convergence, smooth-basis, temporal, concurvity, and residual diagnostics for the reported models."

diagnostic_summary |>
  select(
    .data$Placement,
    .data$Converged,
    .data$`Maximum absolute gradient`,
    .data$`Minimum Hessian eigenvalue`,
    .data$`Basis-check flags`,
    .data$`Maximum absolute lag-1`,
    .data$`Target/participant concurvity`,
    .data$`Deviance explained`,
    .data$`Normal-reference QQ correlation`
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_number(
    columns = c(
      `Maximum absolute gradient`,
      `Maximum absolute lag-1`
    ),
    decimals = 4
  ) |>
  gt::fmt_number(
    columns = `Minimum Hessian eigenvalue`,
    decimals = 5
  ) |>
  gt::fmt_integer(columns = `Basis-check flags`) |>
  gt::cols_label(
    `Basis-check flags` = gt::html("Basis flags<br><small>of 9</small>"),
    `Maximum absolute lag-1` = gt::html("Maximum |lag 1|"),
    `Target/participant concurvity` = gt::html(
      "Photoperiod/participant<br>concurvity"
    ),
    `Normal-reference QQ correlation` = gt::html(
      "QQ correlation<br>range"
    )
  ) |>
  gt::cols_width(
    Converged ~ gt::pct(10),
    `Basis-check flags` ~ gt::pct(10),
    `Target/participant concurvity` ~ gt::pct(18),
    `Deviance explained` ~ gt::pct(14),
    `Normal-reference QQ correlation` ~ gt::pct(16)
  ) |>
  h07_gt(11)
```

The basis-dimension check flagged three near-eye smooths and one chest smooth.
The expanded-basis sensitivity therefore carries more weight than a single
diagnostic threshold: it reproduced 16 of 17 evaluable classifications, with
one near-eye fit failing its Hessian check and the chest sleep low-light
classification changing.

Residual normal-reference correlations were generally high, but residuals
still showed fitted-value dependence and tail departures for some outcomes.
The Tweedie response checks were adequate for the two above-threshold duration
metrics. They were not adequate for time below 1 lx melEDI during sleep: each
placement had four observed zeros, the fitted model expected fewer than 0.0001
zeros, and none of 100 pilot simulations contained a zero. This distributional
mismatch makes the chest sleep low-light pattern particularly weak evidence;
no larger simulation can repair the response model's failure to reproduce the
observed zero mass.

The photoperiod smooth also had estimated concurvity of approximately 0.9985–
0.9988 with the participant random-effect smooth. This reflects the clustered,
repeated-day design and limits separation of between-person photoperiod
coverage from the pooled smooth. Smooth-term p-values are therefore not used
or displayed. The derivative intervals remain conditional on the fitted model
and should be read with this identifiability limitation.

## Sensitivity analyses

The gap-timing-unaware dataset reproduced all nine near-eye classifications
and eight of nine chest classifications; the chest melEDI-dose pattern was the
exception. The paired common sample used 110–112 participants, 505–643
participant-days and observations, and eight sites. It reproduced four of nine
near-eye and six of nine chest classifications, showing that several pooled
patterns depend on which participant-days and sites contribute rather than on
placement alone.

```{r}
#| label: tbl-h07-sensitivity-samples
#| tbl-cap: "Fitted sample sizes for the main data and metric-definition sensitivities. Ranges indicate metric-specific counts."

sensitivity_sample_summary |>
  select(
    .data$Scenario,
    .data$Placement,
    .data$Participants,
    .data$`Participant-days`,
    .data$Observations,
    .data$Sites
  ) |>
  gt::gt(groupname_col = "Scenario") |>
  gt::cols_width(
    Placement ~ gt::pct(16),
    Participants ~ gt::pct(18),
    `Participant-days` ~ gt::pct(21),
    Observations ~ gt::pct(18),
    Sites ~ gt::pct(10)
  ) |>
  gt::tab_source_note(
    gt::md(
      paste0(
        "Preparation exact-common samples were identical for the primary ",
        "and gap-timing-unaware datasets. Observed and corrected dose ",
        "definitions used the same common rows."
      )
    )
  ) |>
  h07_gt(12)
```

```{r}
#| label: tbl-h07-sensitivity-classifications
#| tbl-cap: "Agreement of sensitivity classifications with the reported pooled photoperiod analysis."

sensitivity_summary |>
  select(
    .data$Scenario,
    .data$Placement,
    .data$Agreement,
    .data$`Changed classifications`
  ) |>
  gt::gt(groupname_col = "Scenario") |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    Agreement ~ gt::pct(12),
    `Changed classifications` ~ gt::pct(56)
  ) |>
  gt::tab_source_note(
    gt::md(
      paste0(
        "Agreement concerns the yes/no derivative pattern only; transition ",
        "brackets can move even when the classification agrees."
      )
    )
  ) |>
  h07_gt(12)
```

Restricting the longest continuous period above 250 lx melEDI to days whose
period duration was exactly identified changed the near-eye classification
from no pattern to a pattern at 15.38–15.48 h; the chest classification was
unchanged. Comparing the observed and corrected melEDI-dose definitions on
identical rows left both placement classifications unchanged. Restricting the
two preparation variants to exact common rows preserved every near-eye
classification but removed the chest sleep low-light pattern in both variants.

```{r}
#| label: tbl-h07-model-form-sensitivities
#| tbl-cap: "Classification stability under broader photoperiod smooths and fixed site effects."

model_form_summary |>
  select(
    .data$Model,
    .data$Placement,
    .data$Agreement,
    .data$`Failed fits`,
    .data$`Changed classifications`
  ) |>
  gt::gt(groupname_col = "Model") |>
  gt::fmt_integer(columns = `Failed fits`) |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    Agreement ~ gt::pct(12),
    `Failed fits` ~ gt::pct(12),
    `Changed classifications` ~ gt::pct(48)
  ) |>
  h07_gt(12)
```

The broader smooth reproduced eight of eight evaluable near-eye and eight of
nine chest classifications; the near-eye longest-period fit failed its Hessian
check, and the chest sleep low-light pattern disappeared. Treating site as a
fixed effect reproduced eight of nine classifications at each placement: the
near-eye longest-period pattern appeared, while the chest sleep low-light
pattern disappeared. These changes reinforce that both metrics sit near a
classification boundary.

## Site influence and support

Omitting one site at a time did not act as a new eligibility rule; it tested
whether a pooled pattern depended strongly on one site's observations. Near-eye
mean melEDI and both low-light metrics reproduced their main classification in
all nine omissions. Five positive near-eye patterns changed only when Tübingen
was omitted. The near-eye longest-period classification changed in four of
nine omissions. Chest results were less stable for several outcomes, most
notably time below 1 lx melEDI during sleep, which matched the main
classification in only three of eight omissions.

```{r}
#| label: tbl-h07-near-loso
#| tbl-cap: "Leave-one-site-out influence on near-eye derivative classifications."

loso_table("near_eye")
```

```{r}
#| label: tbl-h07-chest-loso
#| tbl-cap: "Leave-one-site-out influence on chest derivative classifications."

loso_table("chest")
```

The observed civil-photoperiod range was broad overall but uneven across
sites. For example, near-eye observations ranged from 10.33 to 12.35 h in
Madrid, 16.41 to 17.45 h in Munich, and 12.49 to 12.71 h in Kumasi; San José
contributed only a narrow range around 13.47 h, while Borås covered values up
to 20.52 h. Sites were also measured in different collection periods and, in some
cases, different years. Thus the pooled smooth can answer whether an
association is visible in this dataset, but it cannot separate photoperiod
from every site-specific or collection-period difference. Leave-one-site-out
changes show where this limitation affects the descriptive classification.

## Interpretation

The pooled analysis shows several fitted photoperiod associations that rise
and then meet the prespecified pointwise rule for a sustained
zero-compatible derivative. The recurring patterns for mean melEDI, darkest
10 h mean, both above-threshold duration metrics, and melEDI dose across the
two placements are the most consistent descriptive findings. Brightest 10 h
mean differs by placement, the longest-period result is sensitive to sample
definition and model form, and the chest sleep low-light result is undermined
by distributional failure, common-row analyses, model-form changes, and site
influence.

The strongest defensible conclusion is therefore that **derivative-defined
plateau patterns are visible for several personal light-exposure metrics in
the pooled data**, subject to substantial design and model limitations. The
results do not establish a ceiling, an equivalence region, a causal effect of
photoperiod, a distinct latitude effect, a health or clinical consequence, or
equivalence between near-eye and chest placement.

Exact input identities, metric-specific model frames, evaluated formulas,
diagnostic provenance, code order, and output identities are documented in
the [H07 analysis preparation and provenance companion](../../audit/hypotheses/H07/H07_analysis_preparation.html).

Models were fitted with R `r session_record$r_version[[1]]`, mgcv
`r session_record$mgcv_version[[1]]`, and gratia
`r session_record$gratia_version[[1]]`.

```{r}
#| label: session-h07-reader
#| include: false

sessionInfo()
```
