---
title: "H08: Visual light sensitivity and personal light exposure"
subtitle: "Associations between VLSQ-8 scores and nine light-exposure metrics"
format:
  html:
    toc: true
    toc-depth: 3
    page-layout: full
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h08-reader
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(stringr)
  library(tibble)
  library(tidyr)
})

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

configured_root <- Sys.getenv(
  "NATHEALTH_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (
  nzchar(configured_root) &&
    file.exists(file.path(configured_root, "renv.lock"))
) {
  root <- normalizePath(configured_root, winslash = "/", mustWork = TRUE)
} else {
  root <- locate_project_root()
}
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("The H08 reader report requires R 4.6.1", call. = FALSE)
}

source(file.path(root, "scripts", "pipeline", "p_value_display.R"))

read_h08 <- function(area, name) {
  readr::read_csv(
    file.path(root, "artifacts", area, "H08", name),
    show_col_types = FALSE,
    progress = FALSE,
    na = ""
  )
}

metric_registry <- read_h08("06_model_data", "H08_metric_registry.csv")
formula_registry <- read_h08("06_model_data", "H08_formula_registry.csv")
score_audit <- read_h08("08_diagnostics", "H08_vlsq_score_audit.csv")
sample_index <- read_h08("06_model_data", "H08_model_frame_index.csv")
paired_audit <- read_h08("06_model_data", "H08_paired_sample_audit.csv")
main_gap_audit <- read_h08(
  "06_model_data",
  "H08_main_gap_common_sample_audit.csv"
)
master <- read_h08("09_tables", "H08_model_results_master.csv")
family_audit <- read_h08("09_tables", "H08_family_audit.csv")
diagnostics <- read_h08("08_diagnostics", "H08_model_diagnostics.csv")
response_gate <- read_h08("08_diagnostics", "H08_response_family_gate.csv")
loo_summary <- read_h08(
  "08_diagnostics",
  "H08_leave_one_site_out_summary.csv"
)
predictions <- read_h08("09_tables", "H08_centered_predictions.csv")
site_slopes <- read_h08("09_tables", "H08_site_specific_slopes.csv")
gap_sensitivity <- read_h08(
  "09_tables",
  "H08_gap_timing_unaware_sensitivity.csv"
)
photoperiod <- read_h08("09_tables", "H08_photoperiod_sensitivity.csv")
participant_summary <- read_h08(
  "09_tables",
  "H08_participant_summary_sensitivity.csv"
)
exact_longest <- read_h08(
  "09_tables",
  "H08_exactly_identified_longest_period_sensitivity.csv"
)
observed_dose <- read_h08(
  "09_tables",
  "H08_observed_dose_sensitivity.csv"
)

near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
near_paired_id <- "main__glasses__paired_common_sample"
chest_paired_id <- "main__chest__paired_common_sample"

near <- master |>
  filter(run_id == near_id) |>
  arrange(metric_order)
chest <- master |>
  filter(run_id == chest_id) |>
  arrange(metric_order)

format_effect <- function(estimate, low, high, effect_type) {
  ifelse(
    effect_type == "ratio",
    sprintf("×%.3f (%.3f–%.3f)", estimate, low, high),
    sprintf("%+.3f h (%+.3f–%+.3f h)", estimate, low, high)
  )
}

format_p_cell <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  ifelse(
    display$p_bold,
    paste0("**", display$p_display, "**"),
    display$p_display
  )
}

format_sample <- function(participants, participant_days, sites) {
  sprintf("%d / %d / %d", participants, participant_days, sites)
}

model_to_practical <- function(value, response_transform, effect_scale) {
  case_when(
    effect_scale == "difference" ~ value,
    response_transform == "log10_offset_0.1" ~ 10^value,
    effect_scale == "ratio" ~ exp(value),
    TRUE ~ value
  )
}

h08_gt <- function(data, title = NULL, note = NULL, size = 12) {
  output <- data |>
    gt() |>
    opt_row_striping() |>
    sub_missing(missing_text = "—") |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(size),
      data_row.padding = px(4),
      heading.align = "left",
      column_labels.font.weight = "600",
      source_notes.font.size = px(max(10, size - 2))
    )
  if (!is.null(title)) output <- output |> tab_header(title = md(title))
  if (!is.null(note)) output <- output |> tab_source_note(md(note))
  output
}

result_table <- function(data) {
  data |>
    transmute(
      metric_order,
      Metric = manuscript_name,
      `Effect per VLSQ-8 SD (95% CI)` = format_effect(
        estimate_practical_per_sd,
        conf_low_practical_per_sd,
        conf_high_practical_per_sd,
        effect_type
      ),
      `Raw p` = format_p_cell(
        average_p_raw,
        average_adjusted_significant
      ),
      `BH-adjusted p` = format_p_cell(
        average_p_adjusted,
        average_adjusted_significant
      ),
      `Participants / participant-days / sites` = format_sample(
        participants,
        participant_days,
        sites
      )
    ) |>
    arrange(metric_order) |>
    select(-metric_order) |>
    h08_gt(
      note = paste(
        "Effects compare scores separated by one participant SD (5.540",
        "VLSQ-8 points) and include 95% Wald confidence intervals.",
        "BH adjustment is across all nine metrics; adjusted p-values would",
        "be bold at 0.050."
      )
    ) |>
    fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
}

stopifnot(
  nrow(near) == 9L,
  nrow(chest) == 9L,
  nrow(family_audit) == 8L,
  all(family_audit$complete_nine_member_family),
  all(family_audit$independent_recalculation_matches),
  sum(family_audit$adjusted_significant_n) == 0L,
  sum(near$average_adjusted_significant) == 0L,
  sum(near$interaction_adjusted_significant) == 0L,
  sum(chest$average_adjusted_significant) == 0L,
  sum(chest$interaction_adjusted_significant) == 0L,
  score_audit$scoring_rule_verified[[1L]],
  all(paired_audit$exact_row_keys_match),
  all(main_gap_audit$exact_row_keys_match),
  sum(loo_summary$successful_refits) == 153L
)

near_dose <- near |>
  filter(metric_id == "dose_time_sensitive_corrected_medi")
stopifnot(
  nrow(near_dose) == 1L,
  isTRUE(all.equal(
    near_dose$estimate_practical_per_sd,
    0.8457605,
    tolerance = 1e-6
  )),
  isTRUE(all.equal(
    near_dose$average_p_adjusted,
    0.1600391,
    tolerance = 1e-6
  ))
)
```

## Hypothesis and analytical question

The preregistered hypothesis was:

> **H8: Duration-, exposure-history-, and level-based metrics are associated
> with VLSQ-8 light sensitivity scores.**

The analytical question is whether participant-level self-reported visual
light sensitivity is associated with repeated participant-day personal
light-exposure metrics after adjustment for study site. Near-eye measurements
are primary because they more closely represent light near the eyes during
wear. Chest measurements provide complementary evidence and are not pooled
with the near-eye measurements.

::: {.callout-note title="Results in brief"}
Across nine light-exposure metrics, neither the primary near-eye average
associations nor their site-specific heterogeneity met the multiplicity-adjusted
criterion. The strongest near-eye directional pattern was lower corrected
melEDI dose per VLSQ-8 SD (ratio 0.846, 95% CI 0.716–0.999; raw *p* = 0.051,
BH-adjusted *p* = 0.160), so it is not a multiplicity-retained finding.
Complementary chest results and the central sensitivity analyses did not
change that conclusion, although site influence and specified model
limitations require caution.
:::

## What was analysed

VLSQ-8 was available for all 184 participants across nine sites. The stored
score equalled the sum of eight ordered 1–5 item codes plus 5 for every
participant. Scores ranged from 13 to 39, with a mean of 21.598 and a
participant SD of 5.540. The raw score was centred at the observed mean; no
within-site standardisation was used.

Nine participant-day outcomes covered light level, duration, continuous
periods, and exposure history. All use manuscript metric names and melEDI
units. The longest continuous period is an observed lower bound unless its
start and end are both exactly identified; that stricter definition is
examined as a sensitivity.

The predefined preparation sensitivity is called the
**gap-timing-unaware dataset**. It still passed the general 50%-per-hour and
80%-per-day coverage rules. The name means that the timing of the remaining
missing observations is not used for an additional metric-specific adjustment;
it does not mean that gaps, missingness, or coverage were ignored. For this
one-time first explanation, the primary could be interpreted as a
time-sensitive primary metric dataset because its approved preparation uses
the timing of remaining missing observations where that timing is relevant.
From here onward, it is called simply the **primary dataset**.

```{r}
#| label: tbl-h08-metrics
#| tbl-cap: "H08 light-exposure metrics, response specifications, and reported effect scales."

metric_registry |>
  arrange(metric_order) |>
  transmute(
    Metric = manuscript_name,
    Category = manuscript_category,
    Unit = display_unit,
    Model = recode(
      response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    Transform = recode(
      response_transform,
      log10_offset_0.1 = "log10(value + 0.1)",
      identity = "None"
    ),
    `Reported effect` = recode(
      effect_scale,
      ratio = "Ratio",
      difference = "Difference in hours"
    )
  ) |>
  h08_gt()
```

## Statistical models

Participant-day models include a participant random intercept nested within
site. Fixed site effects use the submitted site order and explicit sum
contrasts, so the additive VLSQ-8 coefficient is the site-adjusted average
slope rather than the slope at one reference site. The primary likelihood-ratio
test compares the site-only and additive models. A separate test compares the
additive and interaction models to ask whether the slope differs by site.

The following evaluated cell constructs and prints every exact Wilkinson
formula used for the principal comparisons and the two structured
sensitivities.

```{r}
#| label: tbl-h08-formulas
#| tbl-cap: "Exact evaluated Wilkinson formulas used in H08."
#| echo: true
#| code-fold: false

site_only <- stats::as.formula(
  "response_value ~ site + (1 | site:Id)"
)
additive <- stats::as.formula(
  "response_value ~ site + VLSQ8_c + (1 | site:Id)"
)
interaction <- stats::as.formula(
  "response_value ~ site * VLSQ8_c + (1 | site:Id)"
)
photoperiod_site_only <- stats::as.formula(
  "response_value ~ site + photoperiod_c + (1 | site:Id)"
)
photoperiod_additive <- stats::as.formula(
  "response_value ~ site + photoperiod_c + VLSQ8_c + (1 | site:Id)"
)
photoperiod_interaction <- stats::as.formula(
  "response_value ~ site * VLSQ8_c + photoperiod_c + (1 | site:Id)"
)
participant_site_only <- stats::as.formula(
  "participant_response ~ site"
)
participant_additive <- stats::as.formula(
  "participant_response ~ site + VLSQ8_c"
)
participant_interaction <- stats::as.formula(
  "participant_response ~ site * VLSQ8_c"
)

site_only
additive
interaction
photoperiod_site_only
photoperiod_additive
photoperiod_interaction
participant_site_only
participant_additive
participant_interaction
```

Average-association and site-heterogeneity tests were adjusted separately.
Near-eye and chest placements and the primary and gap-timing-unaware datasets
also had separate families, giving eight complete nine-test Benjamini–Hochberg
families. Paired-placement and common-sample sensitivities report estimates
and 95% confidence intervals without creating additional *p*-value screens.

```{r}
#| label: tbl-h08-families
#| tbl-cap: "Complete H08 multiplicity families."

family_audit |>
  transmute(
    Family = family_id,
    Role = role,
    `Observed / planned tests` = sprintf(
      "%d / %d",
      observed_raw_p,
      planned_n
    ),
    Method = multiplicity_method,
    `Raw p < 0.05` = raw_significant_n,
    `BH-adjusted p < 0.05` = adjusted_significant_n
  ) |>
  h08_gt(
    note = paste(
      "All eight families were complete and independently verified.",
      "No adjusted p-value met the 0.050 criterion."
    )
  )
```

## Exact fitted samples

The primary near-eye models used 139–141 participants, 655–816
participant-days, and all nine sites, depending on the metric. Complementary
chest models used 153–154 participants, 743–902 participant-days, and eight
sites. Every model used its metric-specific complete frame.

```{r}
#| label: tbl-h08-primary-samples
#| tbl-cap: "Exact primary-dataset fitted samples by metric and placement."

sample_index |>
  filter(run_id %in% c(near_id, chest_id)) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Placement = placement_label,
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Placement) |>
  pivot_wider(names_from = Placement, values_from = Sample) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Cells are participants / participant-days / sites.",
      "Exact row-key and model-frame hashes are retained in the source data."
    )
  )
```

The complete sample index is available as
[source data](../../artifacts/06_model_data/H08/H08_model_frame_index.csv),
with [site-specific cells](../../artifacts/06_model_data/H08/H08_model_frame_by_site.csv).

## Primary near-eye results

None of the nine average near-eye associations remained statistically
supported after the complete BH correction. Several point estimates suggested
lower bright-light exposure at higher VLSQ-8 scores, but their uncertainty and
family-adjusted evidence do not support a confirmatory claim.

The corrected melEDI-dose ratio was 0.846 per VLSQ-8 SD (95% CI
0.716–0.999). Its likelihood-ratio raw *p* was 0.051 and BH-adjusted *p* was
0.160. The Wald interval and likelihood-ratio test use different inferential
constructions; the complete-family conclusion is therefore inconclusive, not
statistically significant.

```{r}
#| label: fig-h08-near-eye-effects
#| fig-cap: "Primary near-eye associations between VLSQ-8 and nine personal light-exposure metrics. Effects compare scores separated by one participant SD; horizontal bars are 95% Wald confidence intervals. The additive pre-sleep outcome and ratio outcomes use separate horizontal scales."
#| fig-alt: "Two stacked forest-plot panels show nine primary near-eye associations per VLSQ-8 standard deviation. The top panel shows an hour difference for time below 10 lx melEDI before sleep, whose interval crosses zero. The lower panel shows percentage changes for eight ratio outcomes; most estimates are negative, but no BH-adjusted test is statistically supported."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H08",
  "H08_near_eye_effects.png"
))
```

```{r}
#| label: tbl-h08-near-eye-results
#| tbl-cap: "Primary near-eye VLSQ-8 associations, 95% confidence intervals, p-values, and exact fitted samples."

result_table(near)
```

Figure and result sources are available as
[figure source data](../../artifacts/11_source_data/H08/H08_near_eye_effects_data.csv)
and the [complete numerical result table](../../artifacts/09_tables/H08/H08_model_results_master.csv).

### Centred prediction contrast

Predictions at half an SD below and above the observed VLSQ-8 mean provide a
natural-scale description of the fitted contrast. They are not additional
hypothesis tests.

```{r}
#| label: tbl-h08-near-eye-predictions
#| tbl-cap: "Primary near-eye predictions at half an SD below and above the VLSQ-8 mean."

predictions |>
  filter(run_id == near_id) |>
  mutate(
    Point = recode(
      prediction_id,
      centre_minus_half_sd = "Lower score",
      centre_plus_half_sd = "Higher score"
    ),
    Prediction = sprintf(
      "%.3f (%.3f–%.3f)",
      estimate,
      conf_low,
      conf_high
    )
  ) |>
  select(metric_order, Metric = manuscript_name, Point, VLSQ8, Prediction) |>
  arrange(metric_order, Point) |>
  pivot_wider(
    names_from = Point,
    values_from = c(VLSQ8, Prediction),
    names_glue = "{Point}: {.value}"
  ) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Predictions are on each metric's natural scale with 95% delta-method",
      "confidence intervals and physical lower bounds applied."
    )
  )
```

## Complementary chest results

The complementary chest analysis repeated the same score, metric definitions,
response models, fixed-site adjustment, and nine-test family. No average chest
association remained after its separate BH correction. The directional
pattern broadly resembled the near-eye result but does not create a second
primary finding.

```{r}
#| label: fig-h08-chest-effects
#| fig-cap: "Complementary chest associations between VLSQ-8 and nine personal light-exposure metrics. Effects compare scores separated by one participant SD; horizontal bars are 95% Wald confidence intervals."
#| fig-alt: "Two stacked forest-plot panels show nine complementary chest associations per VLSQ-8 standard deviation. The additive pre-sleep outcome has an interval crossing zero. Most ratio estimates are below one, but no BH-adjusted test is statistically supported."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H08",
  "H08_chest_effects.png"
))
```

```{r}
#| label: tbl-h08-chest-results
#| tbl-cap: "Complementary chest VLSQ-8 associations, 95% confidence intervals, p-values, and exact fitted samples."

result_table(chest)
```

[Chest figure source data](../../artifacts/11_source_data/H08/H08_chest_effects_data.csv)
contain the plotted values and intervals.

## Site-specific heterogeneity

The separate interaction tests ask whether the VLSQ-8 slope differs across
submitted sites. None of the nine near-eye or nine chest interaction tests
remained after their placement-specific BH corrections. Site-specific slopes
are therefore descriptive components of globally inconclusive heterogeneity
tests, not independently supported site findings.

```{r}
#| label: tbl-h08-interactions
#| tbl-cap: "Site-by-VLSQ-8 heterogeneity tests."

bind_rows(near, chest) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `LRT statistic (df)` = sprintf(
      "%.3f (%d)",
      interaction_lrt_statistic,
      interaction_lrt_df
    ),
    `Raw p` = format_p_cell(
      interaction_p_raw,
      interaction_adjusted_significant
    ),
    `BH-adjusted p` = format_p_cell(
      interaction_p_adjusted,
      interaction_adjusted_significant
    )
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Near-eye and chest heterogeneity tests form separate complete",
      "nine-metric families. No adjusted p-value met 0.050."
    )
  ) |>
  fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
```

All site-specific slopes and 95% intervals, using submitted site names and
order, are available in the
[site-slope source table](../../artifacts/09_tables/H08/H08_site_specific_slopes.csv).

## Exact paired-placement comparison

Near-eye and chest models were fitted separately on identical participant-day
keys. The comparison does not pool placements, estimate a within-day placement
difference, or establish equivalence. Ratio outcomes are shown on a common
natural-log scale; the additive pre-sleep hour difference remains in the
adjacent table because the two units must not share an axis.

```{r}
#| label: fig-h08-paired-placement
#| fig-cap: "Near-eye and chest VLSQ-8 associations on identical participant-day samples. The dashed diagonal marks equal point estimates and grey lines mark the null; it is not an equivalence analysis."
#| fig-alt: "Equal-axis scatter plot compares eight ratio-scale near-eye estimates on the horizontal axis with chest estimates on the vertical axis. Component 95% confidence intervals, horizontal and vertical null lines, and a dashed identity line are shown. Most estimates lie below zero on the log-ratio scale, but proximity to the diagonal does not establish equivalence."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H08",
  "H08_paired_placement_effects.png"
))
```

```{r}
#| label: tbl-h08-paired-placement
#| tbl-cap: "Exactly paired near-eye and chest component estimates."

master |>
  filter(run_id %in% c(near_paired_id, chest_paired_id)) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Placement = placement_label,
    Effect = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Placement) |>
  pivot_wider(
    names_from = Placement,
    values_from = c(Effect, Sample),
    names_glue = "{Placement}: {.value}"
  ) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Effects compare scores separated by one VLSQ-8 SD and include 95%",
      "Wald confidence intervals. Sample cells are participants /",
      "participant-days / sites."
    )
  )
```

[Paired-placement figure and table source data](../../artifacts/11_source_data/H08/H08_paired_placement_effects_data.csv)
include the exact sample hashes.

## Model adequacy and influence

All 36 primary-dataset additive and interaction fits converged, had
positive-definite Hessians, were nonsingular, and retained full-rank fixed
effects. No primary or complementary effect was non-estimable. The common
response-family gate retained all nine prespecified specifications; review
flags require qualification, not a data-dependent change of model family.

```{r}
#| label: tbl-h08-response-gate
#| tbl-cap: "Response-family decisions across the four primary-dataset inferential targets."

response_gate |>
  arrange(metric_order) |>
  transmute(
    Metric = manuscript_name,
    `Response / transform` = paste(
      recode(
        response_family,
        gaussian = "Gaussian",
        tweedie_log = "Tweedie, log link"
      ),
      recode(
        response_transform,
        log10_offset_0.1 = "log10(value + 0.1)",
        identity = "untransformed"
      ),
      sep = " / "
    ),
    `Targets reviewed` = inferential_targets,
    `Review flags` = review_targets,
    Decision = str_to_sentence(str_replace_all(
      family_gate_status,
      "_",
      " "
    ))
  ) |>
  h08_gt(
    note = paste(
      "The targets are near-eye and chest average and heterogeneity fits.",
      "No major failure or non-estimable target occurred."
    )
  )
```

Three limitations are material. Darkest 10 h mean showed increasing residual
spread with fitted values (absolute-residual versus fitted Spearman
correlations 0.341 near eye and 0.349 chest). Time below 1 lx melEDI during
sleep contained four exact zeros at each placement, whereas the fitted
Tweedie models assigned near-zero expected counts. Corrected melEDI dose
passed the numerical checks, but its value depends on the timing of remaining
missing observations and is interpreted with the dataset and observed-dose
sensitivities below.

```{r}
#| label: tbl-h08-flagged-diagnostics
#| tbl-cap: "Primary-dataset diagnostic checks requiring explicit qualification."

diagnostics |>
  filter(
    run_id %in% c(near_id, chest_id),
    diagnostic_status != "PASS"
  ) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `Residual |fit| Spearman` = sprintf(
      "%+.3f",
      residual_absolute_fitted_spearman
    ),
    `Observed / expected zeros` = if_else(
      response_family == "tweedie_log",
      sprintf("%d / %.3g", observed_zero_n, expected_zero_n),
      "—"
    ),
    Decision = str_to_sentence(str_replace_all(
      diagnostic_issues,
      "_",
      " "
    ))
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "All convergence, Hessian, singularity, fixed-rank, serial-correlation,",
      "and prediction-bound fields are retained in the complete source table."
    )
  )
```

```{r}
#| label: fig-h08-model-adequacy
#| fig-cap: "Selected primary near-eye model-adequacy checks for all nine metrics. Panels retain their model-specific transformed or Pearson-residual scales and are diagnostic rather than effect displays."
#| fig-alt: "Multi-panel diagnostic figure for nine primary near-eye models shows residual-versus-fitted patterns and distribution checks. Darkest 10 hour mean has the clearest increase in residual spread, and the sleep-below-1-lux model has a small observed zero mass not reproduced by its Tweedie fit."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H08",
  "H08_near_eye_model_adequacy.png"
))
```

The [complete diagnostic table](../../artifacts/08_diagnostics/H08/H08_model_diagnostics.csv)
and [adequacy-figure source data](../../artifacts/11_source_data/H08/H08_near_eye_model_adequacy_data.csv)
contain the full checks.

All 153 planned leave-one-site-out refits succeeded. Some estimates changed
direction or magnitude when one site was omitted; this reinforces the
inconclusive interpretation but does not identify a site for exclusion. No
site or participant was removed after influence review.

```{r}
#| label: tbl-h08-site-influence
#| tbl-cap: "Leave-one-site-out stability of the primary-dataset average associations."

loo_summary |>
  arrange(metric_order, placement_label) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    Refits = sprintf("%d / %d", successful_refits, refits),
    `Sign reversal` = if_else(sign_reversal_any, "Yes", "No"),
    `Largest relative |change|` = sprintf(
      "%.2f",
      maximum_relative_absolute_change
    ),
    `Most influential omitted site` = most_influential_omitted_site,
    Interpretation = influence_status
  ) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Relative changes can be large when the full-sample estimate is close",
      "to zero; they do not by themselves identify an erroneous site."
    ),
    size = 10
  )
```

The ranked
[participant-influence screen](../../artifacts/08_diagnostics/H08/H08_participant_influence_screen.csv)
is an audit list rather than a deletion rule.

## Dataset sensitivity

The strongest comparison fits the primary and gap-timing-unaware values on
identical participant-day keys for every metric and placement. All 18 of
these exact-common-sample comparisons were stable within model uncertainty.
The scenario-specific all-available families also retained the same
BH-adjusted conclusion. Near-eye wake time above 250 lx melEDI and corrected
melEDI dose were precision-sensitive in the all-available comparison because
zero-exclusion by a component 95% interval changed; all chest comparisons
were stable. Neither classification creates a new hypothesis test.

```{r}
#| label: tbl-h08-gap-common-sample
#| tbl-cap: "Primary and gap-timing-unaware associations on exact common participant-day samples."

gap_common_display <- gap_sensitivity |>
  filter(sample_scenario == "main_gap_common_sample") |>
  left_join(
    metric_registry |>
      select(metric_id, response_transform, effect_scale),
    by = "metric_id"
  ) |>
  mutate(
    estimate_main_display = model_to_practical(
      estimate_main,
      response_transform,
      effect_scale
    ),
    conf_low_main_display = model_to_practical(
      conf_low_main,
      response_transform,
      effect_scale
    ),
    conf_high_main_display = model_to_practical(
      conf_high_main,
      response_transform,
      effect_scale
    ),
    estimate_gap_display = model_to_practical(
      estimate_gap_timing_unaware,
      response_transform,
      effect_scale
    ),
    conf_low_gap_display = model_to_practical(
      conf_low_gap_timing_unaware,
      response_transform,
      effect_scale
    ),
    conf_high_gap_display = model_to_practical(
      conf_high_gap_timing_unaware,
      response_transform,
      effect_scale
    )
  )

stopifnot(
  nrow(gap_common_display) == 18L,
  all(gap_common_display$exact_common_keys),
  all(
    gap_common_display$stability_classification ==
      "stable within model uncertainty"
  )
)

gap_common_display |>
  transmute(
    metric_order,
    Placement = if_else(placement == "glasses", "Near eye", "Chest"),
    Metric = manuscript_name,
    `Primary effect (95% CI)` = format_effect(
      estimate_main_display,
      conf_low_main_display,
      conf_high_main_display,
      effect_scale
    ),
    `Gap-timing-unaware effect (95% CI)` = format_effect(
      estimate_gap_display,
      conf_low_gap_display,
      conf_high_gap_display,
      effect_scale
    ),
    Sample = format_sample(
      participants_main,
      participant_days_main,
      sites_main
    ),
    Stability = stability_classification
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Effects compare scores separated by one VLSQ-8 SD and include 95%",
      "Wald intervals. Samples are participants / participant-days / sites."
    ),
    size = 10
  )
```

```{r}
#| label: fig-h08-gap-common-sample
#| fig-cap: "Primary and gap-timing-unaware VLSQ-8 associations on identical participant-day samples, shown separately for near-eye and chest placements."
#| fig-alt: "Paired comparison of primary and gap-timing-unaware estimates on identical samples for nine metrics at each placement. Estimates are generally close; all eighteen comparisons are classified as stable within model uncertainty."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H08",
  "H08_gap_common_sample_effects.png"
))
```

[Dataset-sensitivity figure source data](../../artifacts/11_source_data/H08/H08_gap_common_sample_effects_data.csv)
and the [complete comparison table](../../artifacts/09_tables/H08/H08_gap_timing_unaware_sensitivity.csv)
retain the exact row hashes and all-available classifications.

## Other structured sensitivities

Linear centred photoperiod was added only as a sensitivity covariate, with
site retained and no photoperiod interaction. Participant-summary models give
each participant one outcome value and use the corresponding fixed-effects
model. All 36 additive sensitivity fits converged and retained full-rank
fixed effects. Their estimates were broadly compatible with the
participant-day results and do not form new multiplicity families.

```{r}
#| label: tbl-h08-photoperiod-participant-summary
#| tbl-cap: "Primary, photoperiod-adjusted, and participant-summary estimates."

primary_sensitivity_display <- bind_rows(near, chest) |>
  select(
    placement,
    placement_label,
    metric_order,
    metric_id,
    manuscript_name,
    effect_type,
    primary_estimate = estimate_practical_per_sd,
    primary_low = conf_low_practical_per_sd,
    primary_high = conf_high_practical_per_sd
  ) |>
  left_join(
    photoperiod |>
      select(
        placement,
        metric_id,
        photo_estimate = estimate_practical_per_sd,
        photo_low = conf_low_practical_per_sd,
        photo_high = conf_high_practical_per_sd,
        photo_converged = converged,
        photo_full_rank = fixed_full_rank
      ),
    by = c("placement", "metric_id")
  ) |>
  left_join(
    participant_summary |>
      select(
        placement,
        metric_id,
        participant_estimate = estimate_practical_per_sd,
        participant_low = conf_low_practical_per_sd,
        participant_high = conf_high_practical_per_sd,
        participant_converged = converged,
        participant_full_rank = fixed_full_rank
      ),
    by = c("placement", "metric_id")
  )

stopifnot(
  nrow(primary_sensitivity_display) == 18L,
  all(primary_sensitivity_display$photo_converged),
  all(primary_sensitivity_display$photo_full_rank),
  all(primary_sensitivity_display$participant_converged),
  all(primary_sensitivity_display$participant_full_rank)
)

primary_sensitivity_display |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `Primary effect (95% CI)` = format_effect(
      primary_estimate,
      primary_low,
      primary_high,
      effect_type
    ),
    `Photoperiod-adjusted effect (95% CI)` = format_effect(
      photo_estimate,
      photo_low,
      photo_high,
      effect_type
    ),
    `Participant-summary effect (95% CI)` = format_effect(
      participant_estimate,
      participant_low,
      participant_high,
      effect_type
    )
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "All entries are effects per one VLSQ-8 SD with 95% Wald intervals.",
      "Sensitivity-specific samples and hashes are retained in source data."
    ),
    size = 10
  )
```

The
[photoperiod-adjusted](../../artifacts/09_tables/H08/H08_photoperiod_sensitivity.csv)
and
[participant-summary](../../artifacts/09_tables/H08/H08_participant_summary_sensitivity.csv)
tables contain exact samples and fitted-frame hashes.

### Metric-definition sensitivities

The primary longest-period metric is an observed lower bound. Requiring exact
identification reduced the sample and produced imprecise estimates: near-eye
ratio 0.958 (95% CI 0.856–1.072) and chest ratio 0.939 (95% CI
0.855–1.032). On the exactly paired placement sample, the ratios were 0.943
near eye and 0.952 chest, with both intervals crossing one.

The observed-dose comparison holds the corrected-dose fitted sample fixed and
replaces the value with uncorrected observed dose. It is descriptive and has
no additional *p*-value family. The all-available near-eye ratio was 0.842
(95% CI 0.713–0.995); its paired-sample interval crossed one, and the chest
all-available interval included one. These component intervals do not
override the primary BH family.

```{r}
#| label: tbl-h08-metric-definition-sensitivities
#| tbl-cap: "Exactly identified longest-period and observed-dose sensitivities."

bind_rows(
  exact_longest |>
    mutate(Sensitivity = "Exactly identified longest period"),
  observed_dose |>
    mutate(Sensitivity = "Observed dose on corrected-dose sample")
) |>
  transmute(
    metric_order,
    Sensitivity,
    Placement = placement_label,
    `Sample basis` = recode(
      sample_scenario,
      all_available = "All available",
      paired_common_sample = "Paired placement"
    ),
    `Effect per VLSQ-8 SD (95% CI)` = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Sensitivity, `Sample basis`, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Intervals are 95% Wald intervals. These descriptive sensitivities",
      "do not create additional p-value screens."
    )
  )
```

Exact samples and hashes are available for the
[longest-period](../../artifacts/09_tables/H08/H08_exactly_identified_longest_period_sensitivity.csv)
and [observed-dose](../../artifacts/09_tables/H08/H08_observed_dose_sensitivity.csv)
sensitivities.

## Interpretation

On the accepted score, metric, sample, and model definitions, the observed
data provide no multiplicity-adjusted evidence that VLSQ-8 score is associated
with any of the nine primary near-eye personal light-exposure metrics. The
complementary chest analyses and the central sensitivity analyses do not
materially strengthen that evidence. Several estimates point toward lower
bright-light exposure at higher scores, but their uncertainty, complete-family
adjustment, and site influence make that directional pattern inconclusive.

This conclusion is not evidence that every true association is exactly zero,
and the paired-placement comparison is not an equivalence test. The analysis
is observational: it does not establish that visual light sensitivity changes
personal light exposure, nor does it support physiological or health-effect
claims.

## Reproducibility

The [H08 analysis preparation and provenance companion](../../audit/hypotheses/H08/H08_analysis_preparation.html)
documents the verified path from analytical inputs to these reported outputs.

All displayed estimates are read from the accepted, hash-audited H08 outputs;
this report does not refit an inferential model. Consequential
calculations were produced in R 4.6.1. The
[master result table](../../artifacts/09_tables/H08/H08_model_results_master.csv),
and [figure manifest and alt text](../../artifacts/12_manifests/H08/H08_figure_manifest.csv)
provide the numerical audit trail. The H08 reader-report manifest retains the
file-level identities.

```{r}
#| label: h08-session-information
#| code-fold: true
#| code-summary: "R session information"

sessionInfo()
```
