---
title: "H09: Chronotype and timing of personal light exposure"
subtitle: "MCTQ and MEQ associations with five registered timing metrics"
format:
  html:
    toc: true
    toc-depth: 3
    page-layout: full
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup-h09-reader
#| include: false

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

configured_root <- Sys.getenv(
  "NATHEALTH_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (
  nzchar(configured_root) &&
    file.exists(file.path(configured_root, "renv.lock"))
) {
  root <- normalizePath(configured_root, winslash = "/", mustWork = TRUE)
} else {
  root <- locate_project_root()
}
knitr::opts_knit$set(
  root.dir = file.path(root, "notebooks", "hypotheses")
)

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
if (dir.exists(project_library)) {
  .libPaths(c(project_library, .libPaths()))
}

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(stringr)
  library(tibble)
  library(tidyr)
})

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("The H09 reader report requires R 4.6.1", call. = FALSE)
}

source(file.path(root, "scripts", "pipeline", "p_value_display.R"))

read_h09 <- function(area, name) {
  readr::read_csv(
    file.path(root, "artifacts", area, "H09", name),
    show_col_types = FALSE,
    progress = FALSE,
    na = ""
  )
}

metric_registry <- read_h09("06_model_data", "H09_metric_registry.csv")
predictor_registry <- read_h09(
  "06_model_data",
  "H09_predictor_registry.csv"
)
score_audit <- read_h09("08_diagnostics", "H09_chronotype_score_audit.csv")
master <- read_h09("09_tables", "H09_model_results_master.csv")
family_audit <- read_h09("09_tables", "H09_family_audit.csv")
diagnostic_registry <- read_h09(
  "08_diagnostics",
  "H09_diagnostic_assessment_registry.csv"
)
diagnostic_summary <- read_h09(
  "08_diagnostics",
  "H09_diagnostic_target_summary.csv"
)
diagnostic_adjudication <- read_h09(
  "08_diagnostics",
  "H09_diagnostic_author_adjudication.csv"
)
gap_sensitivity <- read_h09(
  "09_tables",
  "H09_gap_timing_unaware_sensitivity.csv"
)
photoperiod <- read_h09(
  "09_tables",
  "H09_photoperiod_sensitivity.csv"
)
participant_summary <- read_h09(
  "09_tables",
  "H09_participant_summary_sensitivity.csv"
)
ar1 <- read_h09("09_tables", "H09_ar1_sensitivity.csv")
l10_cut <- read_h09("09_tables", "H09_l10_cut_sensitivity.csv")
paired_effects <- read_h09(
  "09_tables",
  "H09_paired_placement_effects.csv"
)
mean_timing <- read_h09(
  "09_tables",
  "H09_fifth_outcome_sensitivity.csv"
) |>
  filter(.data$metric_id == "mean_timing_above_250")

primary_registered <- master |>
  filter(
    .data$data_scenario_id == "primary",
    .data$sample_scenario == "all_available",
    .data$primary_family_member
  ) |>
  arrange(.data$placement, .data$metric_order, .data$instrument_id)
primary_near <- primary_registered |>
  filter(.data$placement == "glasses")
primary_chest <- primary_registered |>
  filter(.data$placement == "chest")

format_effect <- function(estimate, low, high) {
  ifelse(
    is.finite(estimate) & is.finite(low) & is.finite(high),
    sprintf("%+.3f (%+.3f to %+.3f)", estimate, low, high),
    "—"
  )
}

format_p_cell <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  ifelse(
    display$p_bold,
    paste0("**", display$p_display, "**"),
    display$p_display
  )
}

format_sample <- function(participants, days, observations, hours, sites) {
  ifelse(
    is.finite(participants) & is.finite(days) & is.finite(observations) &
      is.finite(hours) & is.finite(sites),
    sprintf(
      "%d / %d / %d / %.1f h / %d",
      participants,
      days,
      observations,
      hours,
      sites
    ),
    "—"
  )
}

h09_gt <- function(data, title = NULL, note = NULL, size = 11) {
  output <- data |>
    gt() |>
    fmt_markdown(columns = where(is.character)) |>
    opt_row_striping() |>
    sub_missing(missing_text = "—") |>
    opt_align_table_header(align = "left") |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(size),
      data_row.padding = px(4),
      heading.align = "left",
      column_labels.font.weight = "600",
      source_notes.font.size = px(max(9, size - 2))
    )
  if (!is.null(title)) output <- output |> tab_header(title = md(title))
  if (!is.null(note)) output <- output |> tab_source_note(md(note))
  output
}

result_table <- function(data) {
  data |>
    transmute(
      .data$metric_order,
      Metric = .data$manuscript_name,
      Instrument = .data$instrument_name,
      `Effect in local clock hours (95% CI)` = format_effect(
        .data$estimate,
        .data$conf_low,
        .data$conf_high
      ),
      `Raw p` = format_p_cell(
        .data$main_p_raw,
        .data$main_raw_significant
      ),
      `BH-adjusted p` = format_p_cell(
        .data$main_p_adjusted,
        .data$main_adjusted_significant
      ),
      `Participants / days / observations / hours / sites` = format_sample(
        .data$participants,
        .data$participant_days,
        .data$observations,
        .data$derivation_hours,
        .data$sites
      )
    ) |>
    arrange(.data$metric_order, .data$Instrument) |>
    select(-.data$metric_order) |>
    h09_gt(
      note = paste(
        "MCTQ effects are per one-hour later MSFsc; MEQ effects are per",
        "10 points greater morning preference. Intervals are 95% Wald",
        "confidence intervals. Raw p-values are bold when raw p < 0.050;",
        "adjusted p-values are bold when BH-adjusted p < 0.050 within the",
        "instrument-specific five-metric family."
      ),
      size = 10
    )
}

stopifnot(
  nrow(primary_near) == 10L,
  nrow(primary_chest) == 10L,
  nrow(family_audit |>
    filter(.data$run_id %in% c(
      "primary__glasses__all_available",
      "primary__chest__all_available"
    ))) == 8L,
  all(family_audit$complete_registered_family),
  all(family_audit$independent_recalculation_matches),
  sum(primary_registered$main_adjusted_significant) == 10L,
  !any(primary_registered$interaction_adjusted_significant),
  sum(diagnostic_summary$overall_assessment == "acceptable") == 18L,
  sum(diagnostic_summary$overall_assessment == "not acceptable") == 6L,
  all(
    diagnostic_registry$assessment[
      diagnostic_registry$domain %in% c(
        "Response and residual distribution",
        "Residual heteroscedasticity"
      )
    ] == "acceptable"
  ),
  all(diagnostic_adjudication$author_final_assessment == "acceptable"),
  all(paired_effects$exact_sample_match),
  nrow(mean_timing) == 4L
)
```

## Hypothesis and analytical question

The preregistered hypothesis was:

> **H9: Timing-based metrics are associated with chronotype (MCTQ, MEQ).**

The analytical question is whether participant-level chronotype is associated
with the local clock timing of personal light exposure across repeated
participant-days, after accounting for study site. Near-eye measurements are
primary because they more closely represent light near the eyes during wear.
Chest measurements provide complementary evidence; placements are not pooled.

::: {.callout-note title="Answer in brief"}
Later MCTQ corrected midsleep and lower MEQ morning preference were associated
with later primary near-eye M10 midpoint, L10 midpoint, and first timing above
250 lx melEDI after five-metric BH adjustment. The clearest near-eye estimate
was for first timing above 250 lx melEDI: +0.381 h per one-hour later MCTQ MSFsc
(95% CI +0.146 to +0.616; BH-adjusted *p* = 0.003) and −0.444 h per 10 points
greater MEQ morning preference (95% CI −0.699 to −0.189; BH-adjusted *p* =
0.001). Last timing and the midpoint of the longest continuous period were not
supported. Complementary chest results broadly reinforced the first-timing
pattern, and no site interaction survived multiplicity adjustment. Structured
sensitivities supported the main directional interpretation, with explicit
limitations for the unavailable longest-period dataset sensitivity and a few
near-zero or precision-sensitive estimates.
:::

## Chronotype constructs and timing outcomes

MCTQ MSFsc and MEQ represent related but distinct constructs. MCTQ MSFsc
describes corrected sleep timing: a larger value denotes later corrected
midsleep. MEQ describes morningness-eveningness preference: a larger score
denotes greater morning preference. The available analytical inputs contained
the pinned aggregate calculated fields for 186 participants. MCTQ MSFsc was
complete for 185 participants and missing for one; MEQ was complete for all
186. The aggregate values were preserved exactly from their pinned source,
but item-level questionnaire responses were unavailable for independent score
reconstruction.

The five registered participant-day outcomes were the midpoint of the
brightest 10 hours (M10), midpoint of the darkest 10 hours (L10), first and
last timing above 250 lx melEDI, and midpoint of the longest continuous period
above 250 lx melEDI. M10 and L10 onset and offset were not analysed. Clock
values were linearized because their observed support was adequately linear;
night-time values used the specified negative-hour conversion where needed.
No circular distribution was imposed solely because an outcome was
clock-valued.

```{r}
#| label: tbl-h09-metrics
#| tbl-cap: "Registered H09 timing outcomes."

metric_registry |>
  filter(.data$primary_family_member) |>
  arrange(.data$metric_order) |>
  transmute(
    Metric = .data$manuscript_name,
    Unit = .data$display_unit,
    `Analysis unit` = str_replace_all(.data$analysis_unit, "_", " "),
    Definition = .data$value_definition
  ) |>
  h09_gt(size = 10)
```

The **gap-timing-unaware dataset** is the dataset applying the 50%-per-hour
and 80%-per-day coverage rules but not using the remaining gaps' time of day
in metric-specific support decisions. It can be contrasted once with the
time-sensitive primary dataset; below, the latter is called simply the
primary dataset.

## Statistical approach

Each chronotype instrument was analysed separately. Participant-day Gaussian
mixed models included fixed study-site effects and a participant random
intercept nested within site. The additive model estimated a site-adjusted
between-participant association; it was compared with a site-only model. A
separate interaction model tested whether the chronotype slope differed by
site. Site-specific slopes were treated as descriptive unless the global
interaction survived adjustment.

```{r}
#| label: tbl-h09-formulas
#| tbl-cap: "Exact Wilkinson formulas used for the principal H09 comparisons."

mctq_site_only <- stats::as.formula(
  "timing_hour ~ site + (1 | site:Id)"
)
mctq_additive <- stats::as.formula(
  "timing_hour ~ site + mctq_hour_centered + (1 | site:Id)"
)
mctq_interaction <- stats::as.formula(
  "timing_hour ~ site * mctq_hour_centered + (1 | site:Id)"
)
meq_site_only <- stats::as.formula(
  "timing_hour ~ site + (1 | site:Id)"
)
meq_additive <- stats::as.formula(
  "timing_hour ~ site + meq_10_centered + (1 | site:Id)"
)
meq_interaction <- stats::as.formula(
  "timing_hour ~ site * meq_10_centered + (1 | site:Id)"
)

tibble(
  Instrument = rep(c("MCTQ MSFsc", "MEQ"), each = 3L),
  Role = rep(c("Site only", "Additive", "Site interaction"), 2L),
  Formula = vapply(
    list(
      mctq_site_only,
      mctq_additive,
      mctq_interaction,
      meq_site_only,
      meq_additive,
      meq_interaction
    ),
    deparse1,
    character(1)
  )
) |>
  h09_gt(size = 10)
```

The centered predictor names encode simple changes of origin and scale.
`mctq_hour_centered` is each participant's MCTQ MSFsc expressed in clock hours
minus the participant mean (`r sprintf("%.3f", score_audit$mctq_center_hour)`
h); an increase of one unit therefore means one hour later corrected midsleep.
`meq_10_centered` is the MEQ score minus the participant mean
(`r sprintf("%.3f", score_audit$meq_center_score)`) and divided by 10; an
increase of one unit means 10 MEQ points toward greater morning preference.
Centering places the model intercept at the observed mean chronotype. It does
not change the chronotype slopes, their confidence intervals, or their tests.

Four inferential families were kept distinct for each placement: MCTQ main
effects, MEQ main effects, MCTQ-by-site interactions, and MEQ-by-site
interactions. Each family used Benjamini–Hochberg adjustment across the five
registered timing outcomes. Statistical significance was decided at
BH-adjusted *p* < 0.050 before formatting. All displayed intervals are 95%
Wald confidence intervals. The mean-timing sensitivity reported later was not
added to these families.

## Exact fitted samples

Primary near-eye fits used 131–141 participants, 478–816 participant-days and
observations, 11,325.5–18,851.0 derivation hours, and nine sites, depending on
the outcome and instrument. Complementary chest fits used 149–154
participants, 547–902 participant-days and observations, 12,980.0–20,891.8
derivation hours, and eight sites. Every estimate below reports its exact
fitted sample.

```{r}
#| label: tbl-h09-primary-samples
#| tbl-cap: "Exact primary-dataset fitted samples by registered outcome, instrument, and placement."

primary_registered |>
  transmute(
    .data$metric_order,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    Placement = .data$placement_label,
    `Participants / days / observations / hours / sites` = format_sample(
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$derivation_hours,
      .data$sites
    )
  ) |>
  arrange(.data$metric_order, .data$Instrument, .data$Placement) |>
  select(-.data$metric_order) |>
  h09_gt(
    note = "Participant-days and observations coincide because each fitted row is one participant-day outcome.",
    size = 10
  )
```

## Primary near-eye results

Both chronotype instruments showed the same substantive pattern in opposite
score directions. Later MCTQ MSFsc was associated with later M10, L10, and
first-above-250 timing. Greater MEQ morning preference was associated with
earlier timing for those same three outcomes. The adjusted evidence did not
support associations with last-above-250 timing or the midpoint of the longest
continuous period.

```{r}
#| label: tbl-h09-near-eye-results
#| tbl-cap: "Primary near-eye chronotype associations, 95% confidence intervals, p-values, and exact fitted samples."

result_table(primary_near)
```

```{r}
#| label: fig-h09-primary-effects
#| fig-cap: "Site-adjusted chronotype associations with the five registered personal light-exposure timing metrics. Near-eye estimates are primary and chest estimates complementary; horizontal bars are 95% confidence intervals."
#| fig-alt: "Two-panel forest plot separates MCTQ MSFsc and MEQ. Each panel shows near-eye and chest estimates with 95% confidence intervals for M10 midpoint, L10 midpoint, first and last timing above 250 lx melEDI, and midpoint of the longest continuous period. MCTQ estimates for M10, L10, and first timing tend positive, while MEQ estimates for those outcomes tend negative."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H09",
  "H09_primary_effects.png"
))
```

[Primary effect figure source data](../../artifacts/11_source_data/H09/H09_primary_effects_data.csv)
and the [complete numerical results](../../artifacts/09_tables/H09/H09_model_results_master.csv)
retain the plotted estimates, intervals, tests, and exact samples.

## Complementary chest results

Chest measurements were analysed with the same outcome definitions and
instrument-specific five-metric adjustment, without pooling with near-eye
measurements. The first-above-250 association was supported for both
instruments. MEQ associations were also supported for M10 and L10 midpoint;
the chest MCTQ L10 result was close to but did not meet the adjusted rule
(BH-adjusted *p* = 0.055). Last timing and the longest-period midpoint were not
supported.

```{r}
#| label: tbl-h09-chest-results
#| tbl-cap: "Complementary chest chronotype associations, 95% confidence intervals, p-values, and exact fitted samples."

result_table(primary_chest)
```

## Site-specific heterogeneity

No MCTQ-by-site or MEQ-by-site interaction survived its separate five-metric
BH family at either placement. The smallest interaction-adjusted *p*-value was
0.070 for the near-eye longest-period midpoint with MEQ. Consequently,
site-specific trends remain descriptive rather than separate inferential
findings.

```{r}
#| label: tbl-h09-interactions
#| tbl-cap: "Site-by-chronotype interaction tests for the primary dataset."

primary_registered |>
  transmute(
    .data$metric_order,
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    `LRT statistic (df)` = sprintf(
      "%.3f (%d)",
      .data$interaction_chi_square,
      .data$interaction_df
    ),
    `Raw p` = format_p_cell(
      .data$interaction_p_raw,
      .data$interaction_raw_significant
    ),
    `BH-adjusted p` = format_p_cell(
      .data$interaction_p_adjusted,
      .data$interaction_adjusted_significant
    )
  ) |>
  arrange(.data$metric_order, .data$Instrument, .data$Placement) |>
  select(-.data$metric_order) |>
  h09_gt(
    note = paste(
      "Raw p-values are bold when raw p < 0.050; adjusted p-values are bold",
      "when BH-adjusted p < 0.050 within the relevant five-metric",
      "instrument-by-placement interaction family. No adjusted value met",
      "the rule."
    ),
    size = 9
  )
```

The [descriptive site-specific slopes and 95% intervals](../../artifacts/09_tables/H09/H09_site_specific_slopes.csv)
use the submitted site names and ordering.

## Placement-matched evidence

Near-eye and chest models were fitted separately on identical participant-day
keys for every registered metric and instrument. The display compares their
component estimates and component 95% confidence intervals. It does not
estimate a between-placement difference interval, pool placements, or test
equivalence; no equivalence margin was prespecified.

```{r}
#| label: fig-h09-paired-placement
#| fig-cap: "Near-eye and chest chronotype associations fitted separately on identical participant-day samples. Horizontal and vertical bars are component 95% confidence intervals; the dashed diagonal is an identity reference and is not an equivalence boundary."
#| fig-alt: "Equal-axis scatter plot compares near-eye estimates horizontally with chest estimates vertically for five timing metrics in separate MCTQ and MEQ panels. Each point has horizontal and vertical 95% confidence intervals, with zero lines and a dashed identity line."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H09",
  "H09_paired_placement_effects.png"
))
```

```{r}
#| label: tbl-h09-paired-placement
#| tbl-cap: "Exactly matched near-eye and chest component estimates."

paired_effects |>
  arrange(.data$metric_order, .data$instrument_id) |>
  transmute(
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    `Near-eye effect (95% CI), h` = format_effect(
      .data$estimate_glasses,
      .data$conf_low_glasses,
      .data$conf_high_glasses
    ),
    `Chest effect (95% CI), h` = format_effect(
      .data$estimate_chest,
      .data$conf_low_chest,
      .data$conf_high_chest
    ),
    `Exact paired participants / days / observations` = sprintf(
      "%d / %d / %d",
      .data$participants_glasses,
      .data$participant_days_glasses,
      .data$observations_glasses
    )
  ) |>
  h09_gt(
    note = paste(
      "All sample identities matched exactly. The intervals belong to the",
      "two separately fitted component effects; no interval for their",
      "difference was estimated."
    ),
    size = 10
  )
```

[Paired-placement source data](../../artifacts/11_source_data/H09/H09_paired_placement_effects_data.csv)
retain exact sample hashes and component intervals.

## Model adequacy, dependence, and influence

The diagnostic registry assessed 16 domains for each of 24 primary targets.
Eighteen targets were acceptable across every domain; six retained at least
one explicit limitation. Visual review of the full near-eye and chest panels
supported acceptable final assessments for response/residual distribution and
residual heteroscedasticity. Conservative numerical screen flags remain in the
diagnostic source data and are not erased by that visual assessment.

```{r}
#| label: tbl-h09-qualified-diagnostics
#| tbl-cap: "Primary targets retaining at least one explicit diagnostic limitation."

diagnostic_summary |>
  filter(.data$overall_assessment == "not acceptable") |>
  left_join(
    metric_registry |>
      select(.data$metric_id, .data$metric_order, .data$manuscript_name),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  arrange(.data$metric_order, .data$instrument_id, .data$placement_label) |>
  transmute(
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_id,
    `Not-acceptable domain(s)` = str_replace_all(
      .data$not_acceptable_domain_names,
      "[|]",
      ";"
    ),
    `Participants / days / observations / hours / sites` = format_sample(
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$derivation_hours,
      .data$sites
    )
  ) |>
  h09_gt(
    note = paste(
      "Distribution and heteroscedasticity are not among these final",
      "limitations. A target can have more than one remaining domain."
    ),
    size = 9
  )
```

The two near-eye last-timing slopes were close to zero and changed direction
under some participant-deletion and leave-one-site-out checks, although no
deletion changed a slope by the material 0.25-hour threshold. Chest
longest-period/MEQ had a singular maximum-likelihood interaction fit; its
additive main-effect fit remained estimable, but the interaction diagnostic is
qualified. The registered longest-period outcome was unavailable in the
gap-timing-unaware prepared artifact at both placements and for both
instruments. Fixed-effect matrices were full rank and Hessians positive
definite for all stored fits. Continuous-time AR(1) corrections changed slopes
by at most 0.016 h.

```{r}
#| label: fig-h09-near-eye-diagnostics
#| fig-cap: "Near-eye residual-versus-fitted and normal-score Q-Q diagnostics for the six analysed timing outcomes and both chronotype instruments. Visual review supported acceptable response/residual distribution and heteroscedasticity assessments."
#| fig-alt: "Twenty-four near-eye panels show residuals versus fitted timing in the upper half and normal-score Q-Q plots in the lower half. Some first-above-250 and L10 panels depart more at the tails than other outcomes, while the overall patterns were judged acceptable."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H09",
  "H09_diagnostics_near_eye.png"
))
```

```{r}
#| label: fig-h09-chest-diagnostics
#| fig-cap: "Chest residual-versus-fitted and normal-score Q-Q diagnostics for the six analysed timing outcomes and both chronotype instruments. Visual review supported acceptable response/residual distribution and heteroscedasticity assessments."
#| fig-alt: "Twenty-four chest panels show residuals versus fitted timing in the upper half and normal-score Q-Q plots in the lower half. Some first-above-250 and L10 panels depart more at the tails than other outcomes, while the overall patterns were judged acceptable."
#| out-width: "100%"
#| fig-align: center

knitr::include_graphics(file.path(
  root,
  "artifacts",
  "10_figures",
  "H09",
  "H09_diagnostics_chest.png"
))
```

The [complete assessment registry](../../artifacts/08_diagnostics/H09/H09_diagnostic_assessment_registry.csv),
[visual-adjudication mapping](../../artifacts/08_diagnostics/H09/H09_diagnostic_author_adjudication.csv),
[participant-influence summary](../../artifacts/08_diagnostics/H09/H09_participant_influence_summary.csv),
and [leave-one-site-out summary](../../artifacts/08_diagnostics/H09/H09_leave_one_site_out_summary.csv)
retain the thresholds, evidence, and final assessment basis.

## Dataset sensitivity

For the four registered outcomes available in both datasets, primary and
gap-timing-unaware models were compared on identical participant-day keys.
Fifteen of 16 instrument-by-placement comparisons were stable within model
uncertainty. Chest MCTQ L10 was precision-sensitive, with overlapping 95%
intervals but a change in whether the interval excluded zero. The registered
longest-period midpoint could not be evaluated because the gap-timing-unaware
artifact contains neither that metric nor the endpoints needed to derive it.

```{r}
#| label: tbl-h09-gap-sensitivity
#| tbl-cap: "Registered dataset-sensitivity comparisons requiring qualification."

gap_sensitivity |>
  filter(
    .data$metric_id != "mean_timing_above_250",
    .data$common_sample_stability != "stable within model uncertainty"
  ) |>
  arrange(.data$metric_order, .data$instrument_id, .data$placement_label) |>
  transmute(
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    `Primary effect (95% CI), h` = format_effect(
      .data$primary_common_estimate,
      .data$primary_common_conf_low,
      .data$primary_common_conf_high
    ),
    `Gap-timing-unaware effect (95% CI), h` = format_effect(
      .data$gap_common_estimate,
      .data$gap_common_conf_low,
      .data$gap_common_conf_high
    ),
    `Common participants / days / observations / hours / sites` =
      format_sample(
        .data$primary_common_participants,
        .data$primary_common_participant_days,
        .data$primary_common_observations,
        .data$primary_common_derivation_hours,
        .data$primary_common_sites
      ),
    Assessment = str_to_sentence(.data$common_sample_stability)
  ) |>
  h09_gt(
    note = paste(
      "The other 15 estimable registered comparisons were stable within",
      "model uncertainty on exact common keys. Non-estimable rows retain",
      "the missing prepared-data limitation rather than an imputed result."
    ),
    size = 9
  )
```

The [complete dataset-sensitivity table](../../artifacts/09_tables/H09/H09_gap_timing_unaware_sensitivity.csv)
contains all exact samples, intervals, family-adjusted decisions, and row-key
identities.

## Other structured sensitivities

Within-site photoperiod adjustment was stable for 21 of 24 targets; its two
direction-sensitive results were the near-zero near-eye last-timing slopes,
and one chest mean-timing estimate was precision-sensitive. Equal-weight
participant summaries were stable for 20 of 24 targets; direction changes
were confined to near-zero last-timing or chest longest-period slopes. All 24
continuous-time AR(1) checks were stable. The stricter afternoon cut used to
linearize L10 changed slopes by at most 0.037 h, and all four L10 checks were
acceptable.

```{r}
#| label: tbl-h09-sensitivity-summary
#| tbl-cap: "Summary of structured sensitivity classifications."

tibble(
  Sensitivity = c(
    "Within-site photoperiod adjustment",
    "Equal-weight participant summary",
    "Continuous-time AR(1)",
    "L10 clock cut"
  ),
  `Targets assessed` = c(
    nrow(photoperiod),
    nrow(participant_summary),
    nrow(ar1),
    nrow(l10_cut)
  ),
  Result = c(
    sprintf(
      "%d stable; %d direction-sensitive; %d precision-sensitive",
      sum(photoperiod$stability_classification ==
        "stable within model uncertainty"),
      sum(photoperiod$stability_classification == "direction-sensitive"),
      sum(photoperiod$stability_classification == "precision-sensitive")
    ),
    sprintf(
      "%d stable; %d direction-sensitive",
      sum(participant_summary$stability_classification ==
        "stable within model uncertainty"),
      sum(participant_summary$stability_classification ==
        "direction-sensitive")
    ),
    sprintf(
      "%d stable; maximum slope change %.3f h",
      sum(ar1$stability_classification == "stable within model uncertainty"),
      max(abs(ar1$ar1_effect_difference))
    ),
    sprintf(
      "%d acceptable; maximum slope change %.3f h",
      sum(l10_cut$clock_cut_assessment == "acceptable"),
      max(abs(l10_cut$estimate_difference))
    )
  )
) |>
  h09_gt(size = 10)
```

The registered fifth outcome remained the midpoint of the longest continuous
period. Mean timing across all supported exposure above 250 lx melEDI is a
distinct sensitivity estimand and was not substituted for that outcome. Its
near-eye MEQ estimate was −0.169 h per 10 MEQ points (95% CI −0.334 to −0.003;
raw *p* = 0.041), but this raw result is outside the registered multiplicity
families and does not alter the confirmatory conclusion.

```{r}
#| label: tbl-h09-mean-timing-sensitivity
#| tbl-cap: "Distinct mean-timing sensitivity outside the registered five-metric families."

mean_timing |>
  arrange(.data$placement, .data$instrument_id) |>
  transmute(
    Placement = .data$placement_label,
    Instrument = .data$instrument_name,
    `Effect in local clock hours (95% CI)` = format_effect(
      .data$estimate,
      .data$conf_low,
      .data$conf_high
    ),
    `Raw p` = format_p_cell(
      .data$main_p_raw,
      .data$main_p_raw < 0.05
    ),
    `Participants / days / observations / hours / sites` = format_sample(
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$derivation_hours,
      .data$sites
    )
  ) |>
  h09_gt(
    note = paste(
      "Intervals are 95% Wald confidence intervals. Raw p-values are bold",
      "when raw p < 0.050. No BH-adjusted p-value is assigned because this",
      "distinct estimand is outside the registered five-metric families."
    ),
    size = 10
  )
```

[Photoperiod](../../artifacts/09_tables/H09/H09_photoperiod_sensitivity.csv),
[participant-summary](../../artifacts/09_tables/H09/H09_participant_summary_sensitivity.csv),
[continuous-time AR(1)](../../artifacts/09_tables/H09/H09_ar1_sensitivity.csv),
[L10 clock-cut](../../artifacts/09_tables/H09/H09_l10_cut_sensitivity.csv),
and [mean-timing](../../artifacts/09_tables/H09/H09_fifth_outcome_sensitivity.csv)
source tables retain exact samples and 95% confidence intervals.

## Interpretation

The primary near-eye evidence supports an association between chronotype and
the timing of several, but not all, features of personal light exposure.
Participants with later corrected midsleep tended to have later M10, L10, and
first-above-250 timing; participants with greater morning preference tended to
show the corresponding earlier timing. The agreement of these distinct
chronotype constructs strengthens the directional interpretation without
making them interchangeable.

The data did not support an association with last-above-250 timing or the
registered longest-period midpoint, nor did they support site-specific
heterogeneity after multiplicity adjustment. Chest evidence was complementary
and broadly consistent, but it neither replaces the primary near-eye results
nor establishes placement equivalence. The analysis is observational and
between participants; it does not establish that chronotype causes a change
in personal light exposure timing.

The conclusion should be read alongside the direction instability of the
near-zero near-eye last-timing estimates, the qualified chest
longest-period/MEQ interaction fit, and the unavailable longest-period
gap-timing-unaware comparison. These limitations do not change the supported
M10, L10, and first-timing pattern, but they constrain claims about the
unsupported outcomes and cross-dataset robustness.

## Reproducibility

All displayed scientific results are read from the accepted, hash-audited H09
outputs; this report does not refit an inferential model. The
[analysis-preparation and provenance companion](../../audit/hypotheses/H09/H09_analysis_preparation.html)
documents the path from verified inputs to these outputs. Consequential
calculations were produced in R 4.6.1. The [master result table](../../artifacts/09_tables/H09/H09_model_results_master.csv),
[complete diagnostic registry](../../artifacts/08_diagnostics/H09/H09_diagnostic_assessment_registry.csv),
[figure manifest and alt text](../../artifacts/12_manifests/H09/H09_figure_manifest.csv),
and reader-report manifest provide the numerical and file-level audit trail.

```{r}
#| label: h09-session-information
#| code-fold: true
#| code-summary: "R session information"

sessionInfo()
```
