## Supplementary materials

**Supplementary Methods: Preregistration deviations** is supplied as a separate Word document (`preregistration-deviations.docx`). It describes all departures from the registered plan and the associated qualifications.

::: {.content-visible when-format="html"}

[Download Supplementary Methods: Preregistration deviations](preregistration-deviations.docx), or [read the deviations online](preregistration-deviations.qmd).

:::

Near-eye and complementary chest measurements remain separate sensor-position estimands. Common-sample analyses use the same participants and participant-days at both positions and fit the positions separately. They do not pool positions, test equivalence, or establish a universal correction. Detailed model checks, sensitivities, exact formulas, and package versions are documented in the analysis pages. The [project repository](https://github.com/tscnlab/ZaunerEtAl_bioRxiv_2026) provides the data, code and reproduction instructions.

## Supplementary figures

### Supplementary Figure S1. Near-eye metric distributions

<figure id="fig-s1" class="display-figure">
<img src="`r publication_image("figure-s1")`" alt="Distributions of near-eye light-exposure metrics by country-coded study site.">
<figcaption><strong>Supplementary Figure S1.</strong> Detailed near-eye metric distributions by country-coded study site.</figcaption>
</figure>

### Supplementary Figure S2. Worked metric derivation

<figure id="fig-s2" class="display-figure">
<img src="`r publication_image("figure-s2")`" alt="Example time series linked to the daily time-above-250-lx metric.">
<figcaption><strong>Supplementary Figure S2.</strong> Examples linking 30-minute time series to the daily time-above-250-lx metric.</figcaption>
</figure>

### Supplementary Figure S3. Latitude and civil-photoperiod measurement range

<figure id="fig-s3" class="display-figure">
<img src="`r publication_image("figure-s3")`" alt="Observed civil photoperiod and theoretical bounds by absolute latitude.">
<figcaption><strong>Supplementary Figure S3.</strong> Observed civil photoperiod and theoretical bounds by absolute latitude.</figcaption>
</figure>

### Supplementary Figure S4. Site-average recommendation adherence

<figure id="fig-s4" class="display-figure">
<img src="`r publication_image("figure-s4")`" alt="Six modelled recommendation-adherence levels with 95% confidence intervals across the Daytime, Pre-sleep and Sleep windows. Colours and point shapes distinguish work and free days. The primary free-day estimates are lower for daytime and sleep and higher for pre-sleep. Sleep describes the device-recorded bedside environment.">
<figcaption><strong>Supplementary Figure S4.</strong> Site-average recommendation adherence on work and free days. Daytime, Pre-sleep and Sleep identify the Brown et al. recommendation windows. Each point gives equal weight to nine sites; bars are 95% confidence intervals. The wake-start date labels the preceding sleep, daytime and following pre-sleep windows. Sleep describes the device-recorded bedside environment, not ocular exposure or a whole-bedroom measurement. The pre-sleep contrast remains positive but its interval includes zero under the 80% coverage restriction; temporal dependence remains unresolved. The coverage sensitivity is reported in the Results and Methods.</figcaption>
</figure>

### Supplementary Figure S5. Site-specific adherence and day-type differences

<figure id="fig-s5" class="display-figure">
<img src="`r publication_image("figure-s5a")`" alt="Supplementary Figure S5A. Work-day adherence for all nine country-coded sites in three recommendation-window panels. Filled diamonds identify FDR-retained site-minus-average comparisons; open circles do not. Bars are 95% confidence intervals and the long-dashed blue line is the equal-site work-day mean, including the indexed site.">
<figcaption><strong>Supplementary Figure S5A.</strong> Site-specific work-day recommendation adherence. Filled diamonds identify site-minus-average comparisons retained after FDR adjustment; bars are 95% confidence intervals. The full explanation accompanies part B.</figcaption>
</figure>

::: {.content-visible when-format="docx"}

### Supplementary Figure S5. Site-specific adherence (continued)

:::

<figure id="fig-s5-part-b" class="display-figure">
<img src="`r publication_image("figure-s5b")`" alt="Supplementary Figure S5B. All 27 site-specific free-minus-work differences with 95% confidence intervals. Filled diamonds identify FDR-retained differences from zero. Asterisks identify departures from the window-specific equal-site difference in a separate FDR family. Dotted lines mark zero and long-dashed blue lines mark the equal-site reference.">
<figcaption><strong>Supplementary Figure S5.</strong> Site-specific recommendation adherence, shown as two separate images. <strong>A</strong>, Work-day adherence, with filled diamonds identifying site-minus-average differences that met their FDR criterion. <strong>B</strong>, Free-minus-work differences, with filled diamonds identifying differences from zero and asterisks identifying departures from the window-specific equal-site difference in a separate 27-comparison FDR family. Daytime, Pre-sleep and Sleep identify the Brown et al. recommendation windows. Points are estimates and bars are 95% confidence intervals. Each equal-site reference includes the indexed site. Sites are components of one pooled observational model, not independent replications or causal effects of location. Day type follows the wake-start date; sleep is the preceding bedside sleep-environment window. Temporal dependence remains unresolved, and the pre-sleep coverage qualification applies.</figcaption>
</figure>

### Supplementary Figure S6. Anonymous participant recommendation-adherence profiles

<figure id="fig-s6" class="display-figure">
<img src="`r publication_image("figure-s6")`" alt="Participant-average recommendation adherence across the Daytime, Pre-sleep and Sleep Brown et al. recommendation windows. Colored points show anonymous participants' equal-cycle means, faint lines connect the same participant across windows, half-violin shapes show the window distributions, and dark bars with white points show interquartile ranges and medians. The lines are not time trajectories or ranks.">
<figcaption><strong>Supplementary Figure S6.</strong> Participant-average recommendation adherence across the Brown et al. recommendation windows. Daytime, Pre-sleep, and Sleep are the window labels. Each colored point is one anonymous participant's equal-cycle mean within that window. Faint lines connect the same anonymous participant across the three window summaries; they are not time trajectories or ranks. Half-violin shapes show the window distributions, and dark bars with white points show interquartile ranges and medians. Because each window has its own Brown et al. threshold and direction, absolute percentages should be interpreted within window.</figcaption>
</figure>

### Supplementary Figure S7. Geographic and civil-photoperiod support

<figure id="fig-s7" class="display-figure">
<img src="`r publication_image("figure-s7")`" alt="FDR support matrix for site, civil photoperiod, latitude and site-versus-latitude adequacy across near-eye exposure metrics.">
<figcaption><strong>Supplementary Figure S7.</strong> Geographic and civil-photoperiod associations across near-eye personal light-exposure metrics. Evidence retained after FDR adjustment for overall site, civil photoperiod, latitude, and site-versus-linear-latitude adequacy. These descriptive associations do not identify a mechanism.</figcaption>
</figure>

### Supplementary Figure S8. Nonlinear civil-photoperiod associations and fitted slopes

<figure id="fig-s8" class="display-figure">
<img src="`r publication_image("figure-s8")`" alt="Near-eye metric response smooths and their fitted slopes across observed civil photoperiod, arranged as paired panels.">
<figcaption><strong>Supplementary Figure S8.</strong> Nonlinear civil-photoperiod associations and their fitted slopes. A qualifying transition denotes an increase followed by a sustained near-flat tail under the specified slope rule. These descriptive associations do not identify a mechanism, and no physiological or environmental ceiling was identified.</figcaption>
</figure>

### Supplementary Figure S9. Light-source category and local-clock pattern of near-eye melanopic EDI

<figure id="fig-s9" class="display-figure">
<img src="`r publication_image("figure-s9")`" alt="Four-panel display of fitted near-eye melanopic EDI, light-source ratios, participant-hour support across local clock time, and light-source-by-site estimates.">
<figcaption><strong>Supplementary Figure S9.</strong> Near-eye light exposure by reported light source across time of day and study site. <strong>A–C</strong>, Exploratory time-of-day analysis. Panel A shows expected one-hour melanopic EDI with clock-specific 95% confidence intervals; the dashed line is the site-average local-clock smooth. Panel B shows each light-source curve relative to that smooth, with 1 as the reference. Panel C shows available participant-hours; open circles indicate locally sparse support and grey gaps indicate no observations. <strong>D</strong>, Site-average category estimates and site-specific deviation ratios from the light-source-by-site interaction model, with participant-cluster-robust 95% confidence intervals, FDR-adjusted <em>p</em> values, and exact support. The analysis includes 17,935 participant-hours from 140 participants, covering 801 participant-days and nine sites. Panels A–C and panel D answer different questions and are not numerically interchangeable. Estimates are observational. During reported sleep, measurements describe the bedside sleep environment.</figcaption>
</figure>

### Supplementary Figure S10. Paired sensor-position hourly associations

<figure id="fig-s10" class="display-figure">
<img src="`r publication_image("figure-s10")`" alt="Paired near-eye and chest hourly associations estimated separately in the same participants and participant-days.">
<figcaption><strong>Supplementary Figure S10.</strong> Near-eye and complementary chest associations estimated separately for the same participants and participant-days. The sets of supported participant-hours differ between sensor positions, so the display is not an observation-level identity comparison and does not test equivalence.</figcaption>
</figure>

### Supplementary Figure S11. Free-versus-Work hourly timing

<figure id="fig-s11" class="display-figure">
<img src="`r publication_image("figure-s11")`" alt="Exploratory Free-versus-Work hourly fitted exposure and response-scale ratio curves with participant-hour support.">
<figcaption><strong>Supplementary Figure S11.</strong> Exploratory Free-versus-Work hourly fitted exposure and response-scale ratio curves with 95% confidence intervals and participant-hour support.</figcaption>
</figure>

### Supplementary Figure S12. Active-versus-Sedentary hourly timing

<figure id="fig-s12" class="display-figure">
<img src="`r publication_image("figure-s12")`" alt="Exploratory Active-versus-Sedentary hourly fitted exposure and response-scale ratio curves with participant-hour support.">
<figcaption><strong>Supplementary Figure S12.</strong> Exploratory Active-versus-Sedentary hourly fitted exposure and response-scale ratio curves with 95% confidence intervals and participant-hour support.</figcaption>
</figure>

### Supplementary Figure S13. Site-specific hourly routine associations

<figure id="fig-s13" class="display-figure">
<img src="`r publication_image("figure-s13")`" alt="Three aligned site-specific forest-plot panels show Free versus Work day, Active versus Sedentary daily activity status, and previous-sleep ratios with 95% confidence intervals across nine country-coded study sites. Filled points mark associations retained after a separate nine-site FDR adjustment within each predictor, and dashed lines mark site-average estimates.">
<figcaption><strong>Supplementary Figure S13.</strong> Site-specific estimates from the predictor-by-site interaction model. Filled and hollow points distinguish associations that did and did not meet the 0.050 criterion after a separate nine-site FDR adjustment within each predictor. The 95% confidence intervals are unadjusted; dashed lines show site-average estimates calculated with equal weight for each of the nine sites. Sites are components of one pooled model, not independent replications or causal effects of location.</figcaption>
</figure>

### Supplementary Figure S14. Light-exposure behaviour and awareness

<figure id="fig-s14" class="display-figure">
<img src="`r publication_image("figure-s14")`" alt="Near-eye associations across Light Exposure Behaviour Assessment factors and personal light-exposure metrics.">
<figcaption><strong>Supplementary Figure S14.</strong> Near-eye associations per one participant-level standard deviation of each Light Exposure Behaviour Assessment factor. None of the 68 tests retained FDR-adjusted support. Four sleep-environment cells were unfit for inference and are labelled Unfit with estimates suppressed. Lack of adjusted support is inconclusive rather than proof of no association.</figcaption>
</figure>

### Supplementary Figure S15. Visual light sensitivity

<figure id="fig-s15" class="display-figure">
<img src="`r publication_image("figure-s15")`" alt="Near-eye site-average associations between VLSQ-8 and nine light-exposure metrics.">
<figcaption><strong>Supplementary Figure S15.</strong> Near-eye site-average associations between the Visual Light Sensitivity Questionnaire-8 and nine personal light-exposure metrics. Associations compare scores separated by one participant-level standard deviation. Horizontal bars are 95% Wald confidence intervals. None of the nine associations retained FDR-adjusted support.</figcaption>
</figure>

### Supplementary Figure S16. Adjusted chronotype associations

<figure id="fig-s16" class="display-figure">
<img src="`r publication_image("figure-s16")`" alt="Study-site-adjusted chronotype associations with the analysed timing metrics, with estimates and confidence intervals.">
<figcaption><strong>Supplementary Figure S16.</strong> Chronotype and timing of personal light exposure. Study-site-adjusted associations of sleep-timing-based corrected midsleep on Free days and questionnaire-based morningness-eveningness with the analysed timing metrics. Points are estimates and bars are 95% confidence intervals; the two chronotype instruments remain separate. Near-eye estimates provide ocular-exposure evidence; chest estimates, where shown, are complementary non-ocular evidence.</figcaption>
</figure>

### Supplementary Figure S17. Observed timing across chronotype and study sites

<figure id="fig-s17" class="display-figure">
<img src="`r publication_image("figure-s17")`" alt="Observed timing across chronotype and study sites. Observed participant-day timing values across chronotype and country-coded study sites in the corresponding model samples, arranged by timing metric.">
<figcaption><strong>Supplementary Figure S17.</strong> Observed timing across chronotype and study sites. Observed participant-day timing values across chronotype and country-coded study sites for the samples used in the corresponding models. Each point represents one participant-day in a fitted sample. This figure is descriptive and does not replace the adjusted models or estimate independent site effects. Near-eye estimates provide ocular-exposure evidence; chest estimates, where shown, are complementary non-ocular evidence.</figcaption>
</figure>

### Supplementary Figure S18. Age and biological sex

<figure id="fig-s18" class="display-figure">
<img src="`r publication_image("figure-s18")`" alt="Three-panel figure with uppercase left-side panel tags showing age distributions, FDR-retained associations, and descriptive site-specific age-by-site estimates.">
<figcaption><strong>Supplementary Figure S18.</strong> Participant age distributions, the 11 main associations retained after FDR adjustment, and descriptive site-specific estimates for two retained age-by-site interactions in the complementary chest analysis. The site-specific interaction estimates are descriptive model contrasts, not independent site tests. Biological sex, coded Female or Male, was analysed; gender was recorded separately and was not analysed.</figcaption>
</figure>

### Supplementary Figure S19. Biological-sex-specific daily curves

<figure id="fig-s19" class="display-figure">
<img src="`r publication_image("figure-s19")`" alt="Biological-sex-specific near-eye fitted curves and Female-to-Male ratio with 95% confidence intervals.">
<figcaption><strong>Supplementary Figure S19.</strong> Near-eye biological-sex-specific fitted melanopic equivalent daylight illuminance curves and Female-to-Male shifted-value ratio. Ribbons and contrast intervals are participant-cluster-robust 95% confidence intervals at the displayed clock times. Grey shading gives site-average civil-night context with each site weighted equally and is not a model covariate. Biological sex, coded Female or Male, was analysed; gender was recorded separately and was not analysed.</figcaption>
</figure>

## Supplementary tables

### Supplementary Table S1. Descriptive sample flow

::: {#supp-table-s1 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S1, descriptive sample flow"}

```{r}
#| label: display-table-s1
#| echo: false

read_publication_table("table-s1")
```

**Supplementary Table S1.** Participant, participant-day, and one-minute observation counts across the roster, sensor-specific completeness and all-zero screens, final descriptive datasets, and paired main subset. Each row names its denominator explicitly. An excluded participant-day is not necessarily an excluded participant.
:::

### Supplementary Table S2. Full near-eye metric dictionary and distributions

::: {#supp-table-s2 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S2, near-eye metric dictionary and distributions"}

```{r}
#| label: display-table-s2
#| echo: false

read_publication_table("table-s2")
```

**Supplementary Table S2.** Definitions, exact support, median (interquartile range) summaries, and density displays for the 17 near-eye personal light-exposure metrics overall and by site. Values use three decimal places where applicable. Metrics are distinct summaries and should not be interpreted as interchangeable outcomes.
:::

### Supplementary Table S3. Descriptive recommendation-range fractions

::: {#supp-table-s3 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S3, descriptive recommendation-range fractions"}

```{r}
#| label: display-table-s3
#| echo: false

read_publication_table("table-s3")
```

**Supplementary Table S3.** Fractions of valid near-eye minutes meeting each Brown et al. recommendation range, with explicit numerators and denominators. Daytime, Pre-sleep, and Sleep identify the recommendation windows. The Sleep summaries use near-eye illuminance recorded during sleep and describe the bedside sleep environment rather than verified ocular exposure. These pooled-minute summaries are descriptive and differ from the modelled recommendation-window-period estimates in main Table 2.
:::

### Supplementary Table S4. Exploratory cross-window recommendation-adherence associations

::: {#supp-table-s4 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S4, cross-window recommendation-adherence associations"}

```{r}
#| label: display-table-s4
#| echo: false

read_publication_table("table-s4")
```

**Supplementary Table S4.** Exploratory within-participant and between-participant associations across Brown et al. recommendation windows. Results are percentage-point differences per 10 percentage points higher Daytime adherence. The four tests form one FDR family. The within-participant day-level claim is withheld because serial dependence remains unresolved. Between-participant results describe observed monitoring-period averages and do not establish rankings, stable traits, or causal effects.
:::

### Supplementary Table S5. Near-eye fitted-curve dispersion and full-model R² allocation

::: {#supp-table-s5 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S5, near-eye fitted-curve dispersion and full-model R squared allocation"}

```{r}
#| label: display-table-s5
#| echo: false

read_publication_table("table-s5")
```

**Supplementary Table S5.** Near-eye fitted-curve dispersion and Shapley allocation of full-model in-sample R². Fitted-curve variation is in squared log10(melanopic EDI + 0.1 lx) prediction units; dispersion and R²-share ratios are unitless. The allocation includes the shared local-clock curve in every component model as the baseline. Intervals are 95% hierarchical cluster-bootstrap percentile intervals from `r format(bootstrap_count(2000L), big.mark = ",")` replicates and are conditional on the fitted models. Dispersion and R² allocation describe different estimands and should not be added or interpreted causally.
:::

### Supplementary Table S6. Chest fitted-curve dispersion and full-model R² allocation

::: {#supp-table-s6 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S6, chest fitted-curve dispersion and full-model R squared allocation"}

```{r}
#| label: display-table-s6
#| echo: false

read_publication_table("table-s6")
```

**Supplementary Table S6.** Complementary chest fitted-curve dispersion and Shapley allocation of full-model in-sample R². The quantities, bootstrap intervals and interpretation follow Supplementary Table S5; chest and near-eye measurements remain distinct sensor-position estimands.
:::

### Supplementary Table S7. Geographic and photoperiod associations

::: {#supp-table-s7 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S7, near-eye geographic and photoperiod associations"}

```{r}
#| label: display-table-s7
#| echo: false

read_publication_table("table-s7")
```

**Supplementary Table S7.** Overall-site, civil-photoperiod and absolute-latitude associations for the 17 near-eye personal light-exposure metrics. Each displayed *p* value is FDR-adjusted within its stated 17-test family. Site and latitude were evaluated in separate models because absolute latitude is fixed within and confounded with study site.
:::

### Supplementary Table S8. Nonlinear civil-photoperiod classifications

::: {#supp-table-s8 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S8, near-eye nonlinear civil-photoperiod classifications"}

```{r}
#| label: display-table-s8
#| echo: false

read_publication_table("table-s8")
```

**Supplementary Table S8.** Derivative-based classifications for nine near-eye personal light-exposure metrics across the observed civil-photoperiod range. A qualifying transition required the preceding fitted-slope interval to be wholly above zero, the next interval to include zero and every later interval through the recorded maximum to remain zero-compatible. These classifications are descriptive and do not identify a physiological or environmental ceiling.
:::

### Supplementary Table S9. Participant-hour routine associations

::: {#supp-table-s9 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S9, participant-hour routine associations"}

```{r}
#| label: display-table-s9
#| echo: false

read_publication_table("table-s9")
```

**Supplementary Table S9.** Common-association estimates from the participant-hour analysis of Free versus Work day, Active versus Sedentary daily activity status and each additional hour of previous-night sleep duration. Ratios and participant-cluster-robust 95% confidence intervals come from the fixed-site model; the three *p* values form one FDR-adjusted family. Predictor-by-site differences are shown in [Supplementary Figure S13](#fig-s13).
:::

### Supplementary Table S10. Person-level evidence synthesis

::: {#supp-table-s10 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S10, person-level evidence synthesis"}

```{r}
#| label: display-table-s10
#| echo: false

read_publication_table("table-s10")
```

**Supplementary Table S10.** Summary of person-level results in the near-eye analyses. Each row identifies the analysis-specific FDR-correction set or decision structure applied to that result. Results that did not meet the applicable FDR criterion are inconclusive rather than evidence of no association. Four light-behaviour sleep-environment cells were unfit for inference. The two chronotype instruments were analysed separately. Biological sex and gender were recorded separately; gender was not analysed. The complete available-data curve difference was attenuated and not retained in the activity-complete near-eye analysis.
:::

### Supplementary Table S11. Light-exposure behaviour and awareness

:::: {#supp-table-s11 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S11, light-exposure behaviour and awareness"}

```{r}
#| label: display-table-s11a
#| echo: false

read_publication_table("table-s11a")
```

::: {.content-visible when-format="docx"}

{{< pagebreak >}}

:::

```{r}
#| label: display-table-s11b
#| echo: false

read_publication_table("table-s11b")
```
::::

### Supplementary Table S12. Visual light sensitivity

::: {#supp-table-s12 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S12, visual light sensitivity"}

```{r}
#| label: display-table-s12
#| echo: false

read_publication_table("table-s12")
```
:::

### Supplementary Table S13. Chronotype and timing

::: {#supp-table-s13 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S13, chronotype and timing"}

```{r}
#| label: display-table-s13
#| echo: false

read_publication_table("table-s13")
```
:::

### Supplementary Table S14. Age and biological sex

::: {#supp-table-s14 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S14, age and biological sex"}

```{r}
#| label: display-table-s14
#| echo: false

read_publication_table("table-s14")
```
:::

### Supplementary Table S15. Biological-sex-specific daily curves

::: {#supp-table-s15 .supplementary-table .column-page-left tabindex="0" aria-label="Supplementary Table S15, biological-sex-specific daily curves and global tests"}

```{r}
#| label: display-table-s15
#| echo: false

read_publication_table("table-s15")
```
:::
