---
title: "Personal light exposure: Data preparation"
subtitle: "Individual, behavioural, and environmental determinants of personal light exposure in daily life: a multi-country wearable and experience-sampling study"
format: 
  html:
    code-tools: true
    code-link: true
---

## Preface

This supplement documents the import and preprocessing of personal light exposure data for the MeLiDos field study analysis. The goal is to derive at three sets of data:

1. An unaggregated set of light exposure data, both for the eye-level and chest-level wearing position where
  - Values > 120 000 lx are removed (set to `NA`)
  - non-wear periods, that are not also sleep periods, are removed
2. A further processed set (from 1.) where
  - hours with less than 50% data availability are removed
  - days with less than 80% data availability (after the previous step) are removed
3. A further processed set (from 2.) where personal light exposure metrics are calculated
  a. per participant:
  - dynamics-based:
      - *IS*: interdaily stability
      - *IV*: intradaily variability
  b. per participant-day:
  - level-based:
      - `Mean`: geometric mean of melanopic EDI (lx)
      - `M10mean`: geometric mean of melanopic EDI during 10 brightest hours of the day (lx)
      - `L5mean`: geometric mean of melanopic EDI during 5 darkest hours of the day (lx)
  - duration-based:  
      - `TAT1000`: time above 1000 lx melanopic EDI (minutes)
      - `TAT250`: time above 250 lx melanopic EDI (during wake) (hours)
      - `TBT10`: time below 10 lx melanopic EDI (during evening, i.e. 3 hours before sleep) (hours)
      - `TBT1`: time below 1 lx melanopic EDI (during sleep) (hours)
      - `PAT250`: longest period above 250lx melanopic EDI (minutes)
  - timing-based:
      - `FLIT250`: first time above 250 lx melanopic EDI (HH:MM)
      - `LLIT250`: last time above 250 lx melanopic EDI (HH:MM)
      - `M10`: midpoint of brightest 10 hours (HH:MM)
      - `L5`: midpoint of darkest 5 hours (HH:MM)
  - exposure-history-based:
      - `Dose`: melanopic EDI dose (lx·h)
  - spectrum-based:  
      - `MDER`: melanopic daylight efficacy ratio
  c. per participant-hour:
  - level-based:
      - `Mean`:  geometric mean of melanopic EDI (lx)

## Setup

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(LightLogR)
library(melidosData)
library(cowplot)
```

## Importing data

### Light

In the first step, we import the unaggregated light eposure data for the glasses and the chest. These will be loaded from the [GitHub project page](https://github.com/MeLiDosProject). These data have been imported already, trimmed by trial dates, checked for irregular data and gaps. All of them run on a `10 second` interval.

```{r}
#| message: false
light_glasses <- load_data("light_glasses")
light_chest <- load_data("light_chest")
```

### Sleep

The sleep data comes from a morning sleep diary. They contain both sleep and wake times.

```{r}
#| message: false
sleepdiary <- load_data("sleepdiaries")
```

```{r}
# sleepdiary <- sleepdiary |> flatten_data()
# trial_times <- load_data("trial_times") |> flatten_data()
```

```{r}
sleep_tester <- function(sleepdiary, trial_times, filter) {
  trial_times <- trial_times |> filter(site == filter) |> drop_na(start, end)
  
  sleepdiary |> 
  group_by(Id) |> 
  filter(site == filter) |> 
  ggplot(aes(y = Id)) +
  geom_linerange(aes(xmin = sleepprep, xmax = wake),
                 col = "red", linewidth = 2) +
  geom_point(data = trial_times, aes(x = start), col = "green3", size = 4) +
  geom_point(data = trial_times, aes(x = end), col = "green3", size = 4) +
  facet_wrap(~Id, ncol = 1, scales = "free") +
  scale_x_datetime(date_breaks = "24 hours",
                   date_labels = "%H:%M %D",
                   limits = LightLogR::Datetime_limits,
                   )+
  labs(y = NULL) +
  # theme_cowplot() +
  theme_sub_strip(background = element_blank(),
                  text = element_blank()) +
  coord_cartesian(clip = "off")
}
```

```{r}
#| fig-height: 10
# sleep_tester(sleepdiary, trial_times, "KNUST")
```

### Wear log

The non-wear data comes from an app-based wear log that participants filled in whenever they removed or put on the device(s). 

```{r}
#| message: false
wearlog <- load_data("wearlog")
# wearlog <- wearlog |> flatten_data()
```

```{r}
wearlog_tester <- function(wearlog, trial_times, filter) {
  trial_times <- trial_times |> filter(site == filter) |> drop_na(start, end)
  wearlog |> 
  group_by(Id) |> 
  filter(site == filter) |> 
  ggplot(aes(y = Id)) +
  geom_linerange(aes(xmin = start, xmax = end, colour = state), linewidth = 2) +
  geom_point(data = trial_times, aes(x = start), col = "green3", size = 4) +
  geom_point(data = trial_times, aes(x = end), col = "green3", size = 4) +
  facet_wrap(~Id, ncol = 1, scales = "free") +
  scale_x_datetime(date_breaks = "24 hours",
                   date_labels = "%H:%M %D",
                   limits = LightLogR::Datetime_limits,
                   )+
  labs(y = NULL) +
  # theme_cowplot() +
  theme_sub_strip(background = element_blank(),
                  text = element_blank()) +
  coord_cartesian(clip = "off")
}

```

```{r}
#| fig-height: 10
# wearlog_tester(wearlog, trial_times, "KNUST")
```

### Combine light exposure data with log and diary data

For the sleepdiary, a selection of sleep and wake times will be added. For the wearlog, the start and end times of a removal, as well as the type (`state`) of removal will be used.

### Preparation

```{r}
#| message: false

sleepdiary_adj <-
  sleepdiary |>
  map(\(x) x |>
        select(Id, sleepprep, wake) |>
        group_by(Id) |>
        pivot_longer(-Id, names_to = "sleep", values_to = "Datetime") |>
        sc2interval(Statechange.colname = sleep, starting.state = "wake") |>
        sleep_int2Brown(sleep.state = "sleepprep", Brown.day = "wake", 
                        Brown.evening = "pre-sleep", Brown.night = "sleep") |>
        mutate(sleep = case_when(is.na(sleep) & State.Brown == "pre-sleep" ~ "wake",
                                 .default = sleep))
  )

wearlog_adj <-
  wearlog |>  
    map(\(x) x |> select(Id, start, end, wear = state))
```

### Combination

```{r}
light_chest_expanded <- 
light_chest |> 
  imap(\(x, idx) x |>
    select(Id, Datetime, MEDI, LIGHT, is.implicit) |>
    add_states(sleepdiary_adj[[idx]], start = Interval, end = Interval) |>
    add_states(wearlog_adj[[idx]]) |> 
    aggregate_Datetime("1 min", type = "floor")
  )

light_glasses_expanded <- 
light_glasses |> 
  imap(\(x, idx) x |> 
    select(Id, Datetime, MEDI, LIGHT, is.implicit) |> 
    add_states(sleepdiary_adj[[idx]], start = Interval, end = Interval) |> 
    add_states(wearlog_adj[[idx]]) |> 
    aggregate_Datetime("1 min", type = "floor")
  )
```

## Preprocessing step 1

In this step, we will remove non-wear instances that are neither marked as sleep, nor fall into a sleep window (sleepdiary). We will also add photoperiod, and remove instances ≥ 1.2*10^5 lx melanopic EDI.

```{r}
light_chest_expanded <- 
light_chest_expanded |> 
  imap(\(x, idx) x |> 
    mutate(MEDI = replace_when(MEDI, 
                               wear == "off" & (State.Brown != "sleep" | is.na(State.Brown)) ~ NA,
                               MEDI >= 120000 ~ NA
                               )) |> 
      add_photoperiod(melidos_coordinates[[idx]])
  )

light_glasses_expanded <- 
light_glasses_expanded |> 
  imap(\(x, idx) x |> 
    mutate(MEDI = replace_when(MEDI, 
                               wear == "off" & (State.Brown != "sleep" | is.na(State.Brown)) ~ NA,
                               MEDI >= 120000 ~ NA
                               )) |> 
      add_photoperiod(melidos_coordinates[[idx]])
  )

light_chest_processed1 <- 
  structure(light_chest_expanded, class = "melidos_data") |> flatten_data() |> group_by(site, Id)

light_glasses_processed1 <- 
  structure(light_glasses_expanded, class = "melidos_data") |> flatten_data() |> group_by(site, Id)
```

**This marks our first dataset**

```{r}
save(light_chest_processed1, file = "data/preprocessed_chest_1.RData")
save(light_glasses_processed1, file = "data/preprocessed_glasses_1.RData")
```

## Preprocessing step 2

Here we further process the data:
  - hours with less than 50% data availability are removed
  - days with less than 80% data availability (after the previous step) are removed

```{r}
light_chest_processed2 <- 
  light_chest_processed1 |> 
  cut_Datetime(unit = "1 hour", group_by = TRUE, type = "floor") |> 
  remove_partial_data(MEDI, threshold.missing = 0.5) |> 
  ungroup(Datetime.rounded) |> 
  select(-Datetime.rounded) |> 
  add_Date_col(group.by = TRUE) |> 
  gap_handler(full.days = TRUE) |> 
  remove_partial_data(MEDI, threshold.missing = 0.2) |> 
  ungroup(Date)

light_glasses_processed2 <- 
  light_glasses_processed1 |> 
  cut_Datetime(unit = "1 hour", group_by = TRUE, type = "floor") |> 
  remove_partial_data(MEDI, threshold.missing = 0.5) |> 
  ungroup(Datetime.rounded) |> 
  select(-Datetime.rounded) |> 
  add_Date_col(group.by = TRUE) |> 
  gap_handler(full.days = TRUE) |> 
  remove_partial_data(MEDI, threshold.missing = 0.2) |> 
  ungroup(Date)
```

**This marks our second dataset**

```{r}
save(light_chest_processed2, file = "data/preprocessed_chest_2.RData")
save(light_glasses_processed2, file = "data/preprocessed_glasses_2.RData")
```

## Preprocessing step 3

### Per participant metric

  - dynamics-based:
      - *IS*: interdaily stability
      - *IV*: intradaily variability
      
```{r}
#| warning: false
metric_chest_participant <- 
  light_chest_processed2 |> 
  summarize(
    interdaily_stability(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
    intradaily_variability(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
  )

metric_glasses_participant <- 
  light_glasses_processed2 |> 
  summarize(
    interdaily_stability(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
    intradaily_variability(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
  )
```

### Per participant-day metric

  - level-based:
      - `Mean`: geometric mean of melanopic EDI (lx)
      - `M10mean`: geometric mean of melanopic EDI during 10 brightest hours of the day (lx)
      - `L5mean`: geometric mean of melanopic EDI during 5 darkest hours of the day (lx)
  - duration-based:  
      - `TAT1000`: time above 1000 lx melanopic EDI (minutes)
      - `TAT250`: time above 250 lx melanopic EDI (during wake) (hours)
      - `TBT10`: time below 10 lx melanopic EDI (during evening, i.e. 3 hours before sleep) (hours)
      - `TBT1`: time below 1 lx melanopic EDI (during sleep) (hours)
      - `PAT250`: longest period above 250lx melanopic EDI (minutes)
  - timing-based:
      - `FLIT250`: first time above 250 lx melanopic EDI (HH:MM)
      - `LLIT250`: last time above 250 lx melanopic EDI (HH:MM)
      - `M10`: midpoint of brightest 10 hours (HH:MM)
      - `L10`: midpoint of darkest 10 hours (HH:MM)
  - exposure-history-based:
      - `Dose`: melanopic EDI dose (lx·h)
  - spectrum-based:  
      - `MDER`: melanopic daylight efficacy ratio

```{r}
metric_chest_participantday <-
  light_chest_processed2 |> 
  group_by(Date, .add = TRUE) |> 
  summarize(
    Mean = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated(),
    bright_dark_period(log_zero_inflated(MEDI), Datetime, "brightest", "10 hours", na.rm = TRUE, as.df = TRUE),
    bright_dark_period(log_zero_inflated(MEDI), Datetime, "darkest", "10 hours", loop = TRUE, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "above", 1000, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    period_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    timing_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    dose(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
    MDER = mean(MEDI / LIGHT, na.rm = TRUE),
    .groups = "drop_last"
  ) |> 
  mutate(MDER = ifelse(abs(MDER) == Inf, NA, MDER),
         across(c(brightest_10h_mean, darkest_10h_mean), exp_zero_inflated)
  )

metric_glasses_participantday <-
  light_glasses_processed2 |> 
  group_by(Date, .add = TRUE) |> 
  summarize(
    Mean = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated(),
    bright_dark_period(log_zero_inflated(MEDI), Datetime, "brightest", "10 hours", na.rm = TRUE, as.df = TRUE),
    bright_dark_period(log_zero_inflated(MEDI), Datetime, "darkest", "10 hours", loop = TRUE, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "above", 1000, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    period_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    timing_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    dose(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
    MDER = mean(MEDI / LIGHT, na.rm = TRUE),
    .groups = "drop_last"
  ) |> 
  mutate(MDER = ifelse(abs(MDER) == Inf, NA, MDER),
         across(c(brightest_10h_mean, darkest_10h_mean), exp_zero_inflated)
  )
```


```{r}
metric_chest_participantday2 <- 
  light_chest_processed2 |> 
  group_by(Date, State.Brown, .add = TRUE) |> 
  summarize(
    duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "below", 10, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "below", 1, na.rm = TRUE, as.df = TRUE),
    .groups = "drop_last"
  ) |> 
  pivot_longer(-c(site:State.Brown), names_to = "metric") |> 
  filter_out(
    !(State.Brown == "wake" & metric == "duration_above_250"),
    !(State.Brown == "pre-sleep" & metric == "duration_below_10"),
    !(State.Brown == "sleep" & metric == "duration_below_1"),
    ) |>
  drop_na(State.Brown) |> 
  unite(metric, metric, State.Brown) |> 
  pivot_wider(id_cols = c(site, Id, Date), values_from = value, names_from = metric)

metric_glasses_participantday2 <- 
  light_glasses_processed2 |> 
  group_by(Date, State.Brown, .add = TRUE) |> 
  summarize(
    duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "below", 10, na.rm = TRUE, as.df = TRUE),
    duration_above_threshold(MEDI, Datetime, "below", 1, na.rm = TRUE, as.df = TRUE),
    .groups = "drop_last"
  ) |> 
  pivot_longer(-c(site:State.Brown), names_to = "metric") |> 
  filter_out(
    !(State.Brown == "wake" & metric == "duration_above_250"),
    !(State.Brown == "pre-sleep" & metric == "duration_below_10"),
    !(State.Brown == "sleep" & metric == "duration_below_1"),
    ) |>
  drop_na(State.Brown) |> 
  unite(metric, metric, State.Brown) |> 
  pivot_wider(id_cols = c(site, Id, Date), values_from = value, names_from = metric)

```

```{r}
metric_chest_participantday <-
  metric_chest_participantday |> 
  left_join(metric_chest_participantday2, 
            by = c("site", "Id", "Date"))

metric_glasses_participantday <-
  metric_glasses_participantday |> 
  left_join(metric_glasses_participantday2, 
            by = c("site", "Id", "Date"))
```

### Per participant-hour metric

  - level-based:
      - `Mean`:  geometric mean of melanopic EDI (lx)

```{r}
metric_chest_participanthour <- 
  light_chest_processed2 |> 
  aggregate_Datetime(
    "30 mins",
    type = "floor",
    numeric.handler = \(x) x |> mean(na.rm = TRUE),
    geo.MEDI = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated()
  )|>
  add_Date_col(group.by = TRUE) |> 
  mutate(static = all(MEDI == MEDI[1])) |> 
  filter_out(static) |> 
  select(-static) |> 
  ungroup(Date)

metric_glasses_participanthour <- 
  light_glasses_processed2 |> 
  aggregate_Datetime(
    "30 mins",
    type = "floor",
    numeric.handler = \(x) x |> mean(na.rm = TRUE),
    geo.MEDI = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated()
  )|>
  add_Date_col(group.by = TRUE) |> 
  mutate(static = all(MEDI == MEDI[1])) |> 
  filter_out(static) |> 
  select(-static) |> 
  ungroup(Date)
```

**This marks our third dataset**

```{r}
save(metric_chest_participanthour, 
     metric_chest_participant, 
     metric_chest_participantday, 
     file = "data/metrics_separate_chest.RData")
save(metric_glasses_participanthour, 
     metric_glasses_participant, 
     metric_glasses_participantday, 
     file = "data/metrics_separate_glasses.RData")
```


## Session info

```{r}
sessionInfo()
```

