---
title: "Descriptives"
subtitle: "Individual, behavioural, and environmental determinants of personal light exposure in daily life: a multi-country wearable and experience-sampling study"
author: "Johannes Zauner"
format: 
  html:
    toc: true
    code-tools: true
    code-link: true
    page-layout: full
---

## Preface

This document contains descriptive plots and tables of the dataset.

## Setup

```{r}
#| message: false
#| warning: false
#| filename: Setup
library(tidyverse)
library(melidosData)
library(LightLogR)
library(magick)
library(cowplot)
library(ggridges)
library(gt)
library(rnaturalearth)
library(rnaturalearthdata)
library(legendry)
library(ggrepel)
library(sf)
library(patchwork)
library(glue)
library(rlang)
library(gtsummary)
source("scripts/helpers.R")
source("scripts/Fig2.R")
source("scripts/Fig3.R")
```

## Load data

```{r}
#| message: false
load("data/metrics_glasses.RData")
load("data/metrics_chest.RData")
demographics <- load_data("demographics") |> flatten_data()
chronotype <- load_data("chronotype") |> flatten_data()
sleep <- load_data("sleepdiaries") |> flatten_data()
wearlog <- load_data("wearlog") |> flatten_data()
load("data/preprocessed_glasses_1.RData")
load("data/preprocessed_chest_1.RData")
load("data/preprocessed_glasses_2.RData")
load("data/preprocessed_chest_2.RData")
```

## Table 1: Participant and site characteristics

### Prepare data

#### Site and city

```{r}
tbl1_country <-
  melidos_countries |> 
  enframe("site", "country")

tbl1_city <-
  melidos_cities |> 
  enframe("site", "city")
```

#### Participant N

```{r}
tbl1_partN <-
  demographics |> 
    count(site, name = "partN") |> 
    add_row(site = "Overall", demographics |> count(name = "partN"))

glasses_partN <- 
  light_glasses_processed1 |> summarize(Id = unique(Id), .groups = "drop")

glasses_partN <-
  glasses_partN |> 
    count(site, name = "partN_glasses") |> 
    add_row(site = "Overall", glasses_partN |> count(name = "partN_glasses"))

chest_partN <- 
  light_chest_processed1 |> summarize(Id = unique(Id), .groups = "drop")

chest_partN <-
  chest_partN |> 
    count(site, name = "partN_chest") |> 
    add_row(site = "Overall", chest_partN |> count(name = "partN_chest"))

tbl1_partN <- 
tbl1_partN |> 
  left_join(glasses_partN, by = "site") |> 
  left_join(chest_partN, by = "site") |> 
    replace_na(list(partN_chest = 0))
```

#### Participant-days N

```{r}
glasses_daysN <- 
  light_glasses_processed1 |> 
  add_Date_col(group.by = TRUE) |>
  summarize(Date = unique(Date), .groups = "drop")

tbl1_daysN <-
  glasses_daysN

glasses_daysN <-
  glasses_daysN |> 
    count(site, name = "daysN_glasses") |> 
    add_row(site = "Overall", glasses_daysN |> count(name = "daysN_glasses"))

chest_daysN <- 
  light_chest_processed1  |> 
  add_Date_col(group.by = TRUE) |> 
  summarize(Date = unique(Date), .groups = "drop")

tbl1_daysN <-
  tbl1_daysN |> full_join(chest_daysN, by = names(tbl1_daysN))

chest_daysN <-
  chest_daysN |> 
    count(site, name = "daysN_chest") |> 
    add_row(site = "Overall", chest_daysN |> count(name = "daysN_chest"))

tbl1_daysN <- 
tbl1_daysN |> 
  count(site, name = "daysN") |> 
  add_row(site = "Overall", tbl1_daysN |> count(name = "daysN")) |> 
  left_join(glasses_daysN, by = "site") |> 
  left_join(chest_daysN, by = "site") |> 
    replace_na(list(daysN_chest = 0))
```


#### Weekend/weekdays N

```{r}
glasses_weekdayN <-
  light_glasses_processed1 |> 
  add_Date_col(group.by = TRUE) |>
  mutate(Weekend = wday(Date, week_start = 1) >5) |> 
  distinct(site, Id, Date, Weekend) 

tbl1_weekdayN <-
  glasses_weekdayN

glasses_weekdayN <-
  glasses_weekdayN |> 
  ungroup(Date) |> 
  count(Weekend)|> 
    ungroup() |> 
    count(site, Weekend, name = "weekdayN_glasses", wt = n) |> 
    pivot_wider(id_cols = site, 
                names_from = Weekend, 
                values_from = weekdayN_glasses) |> 
    rename_with(\(x) x |> replace_values("FALSE" ~ "weekdayN_glasses", "TRUE" ~ "weekendN_glasses"))

glasses_weekdayN <-
  glasses_weekdayN |> 
    add_row(site = "Overall", glasses_weekdayN |> summarize(across(-site, sum)))

chest_weekdayN <-
  light_chest_processed1 |> 
  add_Date_col(group.by = TRUE) |>
  mutate(Weekend = wday(Date, week_start = 1) >5) |> 
  distinct(site, Id, Date, Weekend) 

tbl1_weekdayN <-
  tbl1_weekdayN |> full_join(chest_weekdayN, by = names(tbl1_weekdayN))

chest_weekdayN <-
  chest_weekdayN |> 
  ungroup(Date) |> 
  count(Weekend)|> 
    ungroup() |> 
    count(site, Weekend, name = "weekdayN_chest", wt = n) |> 
    pivot_wider(id_cols = site, 
                names_from = Weekend, 
                values_from = weekdayN_chest) |> 
    rename_with(\(x) x |> replace_values("FALSE" ~ "weekdayN_chest", "TRUE" ~ "weekendN_chest"))

chest_weekdayN <-
  chest_weekdayN |> 
    add_row(site = "Overall", chest_weekdayN |> summarize(across(-site, sum)))

tbl1_weekdayN <-
tbl1_weekdayN |> 
  ungroup(Date) |> 
  count(Weekend) |> 
  ungroup() |> 
  count(site, Weekend, name = "weekdayN", wt = n) |> 
  pivot_wider(id_cols = site, 
                names_from = Weekend, 
                values_from = weekdayN) |> 
    rename_with(\(x) x |> replace_values("FALSE" ~ "weekdayN", "TRUE" ~ "weekendN"))

tbl1_weekdayN <-
tbl1_weekdayN |> 
  add_row(site = "Overall", tbl1_weekdayN |> summarize(across(-site, sum))) |> 
  left_join(glasses_weekdayN, by = "site") |> 
  left_join(chest_weekdayN, by = "site") |> 
    replace_na(list(weekdayN_chest = 0))
```

#### Geolocation coordinates

```{r}
tbl1_location <-
  melidos_coordinates |> 
  enframe("site", "coordinates") |> 
  mutate(coordinates = map(coordinates, format_coordinates) |> unlist())
```

#### Photoperiod

```{r}
phot_glasses <- 
light_glasses_processed1 |> 
  add_Date_col(group.by = TRUE) |> 
  select(Id, Date, photoperiod) |> 
  distinct(site, Id, Date, .keep_all = TRUE)

phot_chest <- 
light_chest_processed1 |> 
  add_Date_col(group.by = TRUE) |> 
  select(Id, Date, photoperiod) |> 
  distinct(site, Id, Date, .keep_all = TRUE)

phot_dates <- 
  full_join(
    phot_glasses,
    phot_chest,
    by = names(phot_glasses)
  )

tbl1_photoperiod <-
  phot_dates |> 
  ungroup(Date, Id) |> 
  summarize(
    phot = median(photoperiod),
    phot_p05 = quantile(photoperiod, p=0.05),
    phot_p95 = quantile(photoperiod, p=0.95)
  ) |> 
  rbind(phot_dates |> 
  ungroup() |> 
  summarize(
    site = "Overall",
    phot = median(photoperiod),
    phot_p05 = quantile(photoperiod, p=0.05),
    phot_p95 = quantile(photoperiod, p=0.95)
  ))
```

#### Data collection dates

```{r}
tbl1_dates <- 
phot_dates |> 
  ungroup(Id, Date) |> 
  summarize(date_median = median(Date),
            date_min = min(Date),
            date_max = max(Date)
            ) |> 
  rbind(phot_dates |> 
  ungroup() |> 
  summarize(
    site = "Overall",
    date_median = median(Date),
            date_min = min(Date),
            date_max = max(Date)
  ))
```

#### Age, gender, employment status, chronotype

```{r}
partial_demographics <- 
demographics |> 
  left_join(chronotype |> select(Id, meq_type, meq, msf_sc, sjl), by = "Id") |> 
  mutate(across(c(sex, gender), forcats::fct_drop),
         meq_type = meq_type |> 
           fct_collapse(
             Morning = c("Definitely morning type", "Moderately morning type"),
             Evening = c("Definitely evening type", "Moderately evening type"),
                        ),
         employment_status = employment_status |> 
           fct_collapse(
             `Fully` = c("Full time employed", "Not employed but studying or in training", "Studying and employed"),
             `Partially` = c("Part time employed", "Marginally employed (Minijob)"),
             None = c("Not employed")
                        )
         ) |> 
  select(-c(Id, native_language, language_specification, comments))

factor_vars <- c("sex", "gender", "employment_status", "meq_type")

tbl1_demographics <-
  partial_demographics |> 
  reframe(
    .by = site,
    age_median = median(age) |> round(),
    age_min = min(age),
    age_max = max(age),
    across(c(meq),
           .fns = list(
             median = ~median(.x, na.rm =TRUE)|> round(),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE)|> round(),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)|> round()
           )), 
    across(c(sjl, msf_sc),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(all_of(factor_vars),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             })
  ) |> 
  rbind(
    partial_demographics |> 
    reframe(
    site = "Overall",
    age_median = median(age) |> round(),
    age_min = min(age),
    age_max = max(age),
    across(c(meq),
           .fns = list(
             median = ~median(.x, na.rm =TRUE)|> round(),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE)|> round(),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)|> round()
           )), 
    across(c(sjl, msf_sc),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(all_of(factor_vars),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             })
  )) |> 
  unnest(all_of(factor_vars))
```

#### Sleep duration

```{r}
tbl1_sleep <- 
sleep |> 
  mutate(daytype = daytype2 |> fct_relabel(\(x) x |> str_extract("work|free"))) |> 
  select(site, sleep_duration, daytype) |> 
  summarize(
    across(where(is.difftime),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(where(is.factor),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             }),
    .by = site
  ) |> 
  rbind(
    sleep |> 
  mutate(daytype = daytype2 |> fct_relabel(\(x) x |> str_extract("work|free"))) |> 
  select(site, sleep_duration, daytype) |> 
  summarize(
    site = "Overall",
    across(where(is.difftime),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(where(is.factor),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             })
  )
  ) |> 
  unnest(where(is_tibble))
```

####  Non-wear time

```{r}
#| fig-height: 30
#| fig-width: 10
tbl1_wear <- 
light_glasses_processed1 |> 
  rbind(light_glasses_processed1) |>
  filter(!is.implicit) |>
  distinct(site, Id, Datetime, .keep_all = TRUE) |>
  group_by(site, Id, nonwear = wear == "off" & sleep != "sleepprep") |>
  mutate(
    nonwear = replace_na(nonwear, FALSE)
  ) |>
  durations() |>  
  pivot_wider(id_cols = c(site, Id), names_from = "nonwear", values_from = duration) |> 
  rowwise() |> 
  mutate(total = sum(dplyr::c_across(c(`FALSE`,`TRUE`)), na.rm = TRUE) |> as.duration()) |> 
  group_by(site) |> 
  summarize(
    nonwear = sum(`TRUE`, na.rm = TRUE) |> as.duration(),
    total = sum(total, na.rm = TRUE) |> as.duration(),
    nonwear_pct = nonwear/total
  ) |> 
  rbind(
    light_glasses_processed1 |> 
  rbind(light_glasses_processed1) |>
  filter(!is.implicit) |>
  distinct(site, Id, Datetime, .keep_all = TRUE) |>
  group_by(site, Id, nonwear = wear == "off" & sleep != "sleepprep") |>
  mutate(
    nonwear = replace_na(nonwear, FALSE)
  ) |>
  durations() |>  
  pivot_wider(id_cols = c(site, Id), names_from = "nonwear", values_from = duration) |> 
  rowwise() |> 
  mutate(total = sum(dplyr::c_across(c(`FALSE`,`TRUE`)), na.rm = TRUE) |> as.duration()) |> 
  ungroup() |> 
  summarize(
    site = "Overall",
    nonwear = sum(`TRUE`, na.rm = TRUE) |> as.duration(),
    total = sum(total, na.rm = TRUE) |> as.duration(),
    nonwear_pct = nonwear/total
  )
  )
  
```

#### Days included in calculations

```{r}
glasses_validdaysN <- 
  light_glasses_processed2 |> 
  add_Date_col(group.by = TRUE) |>
  summarize(Date = unique(Date), .groups = "drop")

tbl1_validdaysN <-
  glasses_validdaysN

glasses_validdaysN <-
  glasses_validdaysN |> 
    count(site, name = "validdaysN_glasses") |> 
    add_row(site = "Overall", glasses_validdaysN |> count(name = "validdaysN_glasses"))

chest_validdaysN <- 
  light_chest_processed1  |> 
  add_Date_col(group.by = TRUE) |> 
  summarize(Date = unique(Date), .groups = "drop")

tbl1_validdaysN <-
  tbl1_validdaysN |> full_join(chest_validdaysN, by = names(tbl1_validdaysN))

chest_validdaysN <-
  chest_validdaysN |> 
    count(site, name = "validdaysN_chest") |> 
    add_row(site = "Overall", chest_validdaysN |> count(name = "validdaysN_chest"))

tbl1_validdaysN <- 
tbl1_validdaysN |> 
  count(site, name = "validdaysN") |> 
  add_row(site = "Overall", tbl1_validdaysN |> count(name = "validdaysN")) |> 
  left_join(glasses_validdaysN, by = "site") |> 
  left_join(chest_validdaysN, by = "site") |> 
    replace_na(list(validdaysN_chest = 0))
```

### Combine data

```{r}
tbl1_data <-
  tbl1_country |> 
  add_row(site = "Overall") |> 
    left_join(by = "site", tbl1_city) |> 
    left_join(by = "site", tbl1_location) |> 
    left_join(by = "site", tbl1_partN) |> 
    left_join(by = "site", tbl1_daysN) |> 
    left_join(by = "site", tbl1_weekdayN) |> 
    left_join(by = "site", tbl1_wear) |> 
    left_join(by = "site", tbl1_validdaysN) |> 
    left_join(by = "site", tbl1_dates) |> 
    left_join(by = "site", tbl1_photoperiod) |> 
    left_join(by = "site", tbl1_demographics) |> 
    left_join(by = "site", tbl1_sleep)
```

```{r}
tbl1_data_formatted <- 
tbl1_data |> 
  gt() |> 
  cols_merge(starts_with("partN"), pattern = "{1}<br>(**g:{2}**, c:{3})") |> 
  cols_merge(starts_with("daysN"), pattern = "{1}d<br>(**g:{2}d**, c:{3}d)") |> 
  cols_merge(starts_with("date_"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("validdaysN"), pattern = "{1}d<br>(**g:{2}d**, c:{3}d)") |> 
  cols_merge(starts_with("phot"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("age"), pattern = "{1}yr<br>({2}yr, {3}yr)") |> 
  cols_merge(starts_with("meq_type_"), pattern = "<<{1}m/>>{2}i<</{3}e>>") |> 
  cols_merge(c("meq_median","meq_p05", "meq_p95"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("sjl"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("sex_"), pattern = "{1}F/{2}M") |> 
  cols_merge(starts_with("gender_"), pattern = "{1}W/{2}M<</{3}NB>>") |> 
  cols_merge(starts_with("nonwear"), pattern = "{1} ({2})") |> 
  cols_merge(starts_with("msf_sc"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("sleep_duration"), pattern = "{1}\n({2}, {3})") |> 
  cols_merge(starts_with("employment_status"), pattern = "{1}F<</{2}P>><</{3}N>>") |> 
  cols_merge(c(weekdayN, weekendN), pattern = "{1}wd / {2}we") |> 
  cols_merge(c(daytype_free, daytype_work), pattern = "{2}work / {1}free") |> 
  cols_hide(c(starts_with("weekdayN_"), starts_with("weekendN_"), daytype_NA)) |> 
  fmt_percent(nonwear_pct, decimals = 1) |> 
  fmt_integer(c("meq_median","meq_p05", "meq_p95")) |> 
  fmt_date(starts_with("date_"), date_style = "y.mn.day") |> 
  fmt(starts_with("msf_sc"), fns = \(x) x |> strptime(format = "%H:%M:%S") |> format(format = "%H:%M")) |> 
  fmt_duration(all_of(c("total")),
               input_units = "secs", max_output_units = 2
               ) |> 
  fmt_duration(all_of(c("nonwear", "phot", "phot_p05", "phot_p95", 
                        "sjl_median", "sjl_p05", "sjl_p95", 
                        "sleep_duration_median", "sleep_duration_p05", "sleep_duration_p95")),
               input_units = "secs", max_output_units = 2
               ) |> 
  cols_move(nonwear, after = total) |> 
  cols_move(daytype_free, after = weekdayN) |> 
  cols_move(c(meq_median, sjl_median), after = meq_type_Morning) |> 
  cols_label_with(everything(), fn= str_to_sentence) |> 
  as.data.frame() |> 
  select(site:partN, daysN, weekdayN, daytype_free, total, nonwear, validdaysN, 
         date_median, phot, age_median, sex_Female, gender_Woman,
         employment_status_Fully, meq_type_Morning, msf_sc_median, meq_median, sjl_median, 
         sleep_duration_median)

tbl1_transposed <- 
tbl1_data_formatted |> 
  t()
  
colnames(tbl1_transposed) <- tbl1_transposed[1,]

tbl1_transposed <- 
tbl1_transposed |> 
  as_tibble(rownames = "description") |> 
  mutate(
    description = replace_values(
      description,
      "site" ~ "Institution",
      "country" ~ "Country",
      "city" ~ "City",
      "coordinates" ~ "Coordinates",
    "partN" ~ "Participants",
    "daysN" ~ "Participant-days",
    "weekdayN" ~ "Weekday / weekend",
    "daytype_free" ~ "Workday / free",
    "date_median" ~ "Collection dates",
    "total" ~ "Participant time",
    "nonwear" ~ "Nonwear time (%)",
    "validdaysN" ~ "Complete days (>80% data)",
    "phot" ~ "Photoperiod",
    "age_median" ~ "Age",
    "meq_median" ~ "Chronotype (MEQ-score)",
    "sex_Female" ~ "Sex",
    "gender_Woman" ~ "Gender",
    "employment_status_Fully" ~ "Employment status",
    "sleep_duration_median" ~ "Sleep duration",
    "meq_type_Morning" ~ "Chronotype (MEQ-based)",
    "sjl_median" ~ "Social jetlag",
    "msf_sc_median" ~ "Chronotype (MCTQ-score)"
    ),
    type = c(rep("Site specifics", 4),rep("Data collection", 9), rep("Participant information", 9))
  ) |> 
  relocate(Overall, all_of(names(melidos_cities) |> sort()))

```

### Finish table

```{r}
tbl1 <- 
tbl1_transposed |> 
  group_by(type) |> 
  gt(rowname_col = "description") |> 
  # cols_move(Overall, 1) |> 
  sub_missing() |> 
  gt_multiple(names(melidos_cities), style_tab) |> 
  tab_style(style = cell_text(align = "center"), 
            locations = list(cells_body(),
                             cells_column_labels())) |> 
  tab_style(style = cell_text(weight = "bold", ), 
            locations = list(cells_row_groups(),
                             cells_column_labels())) |> 
  fmt_markdown() |> 
  tab_footnote(
    "g: glasses-mounted measurement device; c: chest-level measurement device",
    locations = cells_stub(c("Participants", "Participant-days", "Complete days (>80% data)"))
  ) |> 
  tab_footnote(
    "Median (minimum, maximum)",
    locations = cells_stub(c("Collection dates", "Age"))
  ) |> 
  tab_footnote(
    "Median (5% percentile, 95% percentile)",
    locations = cells_stub(c("Photoperiod", "Chronotype (MCTQ-score)", 
                             "Chronotype (MEQ-score)", "Social jetlag",
                             "Sleep duration"))
  ) |> 
  tab_footnote(
    "w: weeks, d: days, h: hours, m: minutes, s: seconds",
    locations = cells_stub(c("Participant time", "Participant-days", "Nonwear time (%)", 
                             "Complete days (>80% data)", "Photoperiod", "Social jetlag",
                             "Sleep duration"))
  ) |> 
  tab_footnote(
    "F: female, M: male(sex)/man(gender), W: woman, NB: non-binary",
    locations = cells_stub(c("Sex", "Gender"))
  ) |> 
  tab_footnote(
    "F: Fully employed OR Studying and employed OR Not employed but studying or in training; P: Part time employed OR Marginally employed (Minijob); N: Unemployed",
    locations = cells_stub(c("Employment status"))
  ) |> 
  tab_footnote(
    "Following the MEQ score; m: morning type; i: intermediary type; e: evening type",
    locations = cells_stub(c("Chronotype (MEQ-based)"))
  ) |> 
  tab_footnote(
    "MEQ: Morningness-Eveningness-Questionnaire (circadian preference); MCTQ: Munich Chronotype Questionnaire (midsleep on free days, sleep corrected",
    locations = cells_stub(c("Chronotype (MEQ-score)", "Chronotype (MCTQ-score)"))
  ) |> 
  site_conv_gt(rev = FALSE)

gtsave(tbl1, "tables/tbl1.png", vwidth = 1200)
```


### Reduced table A

```{r}
tbl1_red <- 
tbl1_transposed |> 
  filter_out(
    description %in% 
      c("Chronotype (MEQ-score)", "Gender", "Collection dates", 
        "Weekday / weekend", "Workday / free", "Social jetlag", "Sleep duration")
    ) |> 
    group_by(type) |> 
  gt(rowname_col = "description") |> 
  # cols_move(Overall, 1) |> 
  sub_missing() |> 
  gt_multiple(names(melidos_cities), style_tab) |> 
  tab_style(style = cell_text(align = "center"), 
            locations = list(cells_body(),
                             cells_column_labels())) |> 
  tab_style(style = cell_text(weight = "bold", ), 
            locations = list(cells_row_groups(),
                             cells_column_labels())) |> 
  fmt_markdown() |> 
  tab_footnote(
    "g: glasses-mounted measurement device; c: chest-level measurement device",
    locations = cells_stub(c("Participants", "Participant-days", "Complete days (>80% data)"))
  ) |> 
  tab_footnote(
    "Median (minimum, maximum)",
    locations = cells_stub(any_of(c("Collection dates", "Age")))
  ) |> 
  tab_footnote(
    "Median (5% percentile, 95% percentile)",
    locations = cells_stub(any_of(c("Photoperiod", "Chronotype (MCTQ-score)", 
                             "Chronotype (MEQ-score)", "Social jetlag",
                             "Sleep duration")))
  ) |> 
  tab_footnote(
    "w: weeks, d: days, h: hours, m: minutes, s: seconds",
    locations = cells_stub(any_of(c("Participant time", "Participant-days", "Nonwear time (%)", 
                             "Complete days (>80% data)", "Photoperiod", "Social jetlag",
                             "Sleep duration")))
  ) |> 
  tab_footnote(
    "F: female, M: male(sex)",
    locations = cells_stub(any_of(c("Sex", "Gender")))
  ) |> 
  tab_footnote(
    "F: Fully employed OR Studying and employed OR Not employed but studying or in training; P: Part time employed OR Marginally employed (Minijob); N: Unemployed",
    locations = cells_stub(c("Employment status"))
  ) |> 
  tab_footnote(
    "Following the MEQ score; m: morning type; i: intermediary type; e: evening type",
    locations = cells_stub(c("Chronotype (MEQ-based)"))
  ) |> 
  tab_footnote(
    "MEQ: Morningness-Eveningness-Questionnaire (circadian preference); MCTQ: Munich Chronotype Questionnaire (midsleep on free days, sleep corrected)",
    locations = cells_stub(any_of(c("Chronotype (MEQ-score)", "Chronotype (MCTQ-score)")))
  ) |> 
  site_conv_gt(rev = FALSE)

tbl1_red

gtsave(tbl1_red, "tables/tbl1_red.png", vwidth = 1200)
```


### Reduced table B

```{r}
demographics |> 
  mutate(sex = sex |> fct_drop(),
         Female = sex == "Female",
         Participants = TRUE) |> 
  tbl_summary(by = site, include = c(Participants, Female, age),
              statistic = list(Participants ~ "{n}",
                               age ~ "{mean} ±{sd}"
                               )) |> 
  add_overall() |> 
  as_gt() |> 
  as.data.frame() |> 
  select(variable, stat_0:last_col()) |> 
  rename_with(\(x) c("Variable", "Overall", names(melidos_cities) |> sort())) |> 
  mutate(Variable = replace_values(Variable, "age" ~ "Age")) |> 
  gt() |> 
  gt_multiple(names(melidos_cities), style_tab) |> 
  site_conv_gt() |> 
  tab_header(md("**MeLiDos sites and demographics**")) |> 
  cols_width(
    Overall ~ px(90),
    Variable ~ px(100),
    everything() ~ px(85)) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    cells_column_labels()
  ) |> 
  gtsave("tables/tbl1_short.png")
  
```


## Table 2: Metric distribution

### Prepare data

```{r}
load("data/H1_results.RData")
```

```{r}
metrics <-
metrics |> 
  select(name, metric_type, data,  response, engine) |> 
  filter(!is.na(response)) |>
  mutate(scaling = recode_values(response,
                                 "log_zero_inflated(metric)" ~ "log_10_zero_inflated",
                                 default = "identity") |> 
           replace_when(
             engine == "glmmTMB" ~ "log_10_zero_inflated"
           )) |> 
  select(-c(response, engine))
```

### Calculate site metrics

```{r}
tbl2_data <- 
metrics |> 
  mutate(data = map2(data, name, 
                    \(data, name) if(name != "dose") return(data) else {
                      data |> mutate(metric = metric/1000)
                    }),
         summary = map(data, 
                      \(x) {
                        x |> 
                          summarize(
                            .by = site,
                            median = median(metric, na.rm = TRUE),
                            mean = mean(metric, na.rm = TRUE),
                            sd = sd(metric, na.rm = TRUE),
                            p25 = quantile(metric, na.rm = TRUE, p = 0.25),
                            p75 = quantile(metric, na.rm = TRUE, p = 0.75),
                            n = sum(!is.na(metric))
                          )
                          }
                      ),
         summary_overall = map(data, 
                      \(x) {
                        x |> 
                          summarize(
                            site = "Overall",
                            median = median(metric, na.rm = TRUE),
                            mean = mean(metric, na.rm = TRUE),
                            sd = sd(metric, na.rm = TRUE),
                            p25 = quantile(metric, na.rm = TRUE, p = 0.25),
                            p75 = quantile(metric, na.rm = TRUE, p = 0.75),
                            n = sum(!is.na(metric))
                          )
                          }
         ),
         summary = map2(summary, summary_overall, rbind)
  ) |> 
  select(-summary_overall)
```

### Add plots

```{r}
tbl2_data <- 
tbl2_data |>
  mutate(plot = map2(data, scaling, ridges_function2),
         type2 = metric_type
         ) |> 
  arrange(metric_type, name)
```

### Constructing table

```{r}
tbl2_data_formatted <- 
tbl2_data |>
  select(-data) |>
  unnest(summary) |> 
  pivot_wider(id_cols = c(name:scaling, type2), values_from = median:n, names_from = site) |> 
  rename_with(\(y) str_remove(y, "median_")) |> 
  group_by(metric_type) |> 
  gt(rowname_col = "name") |> 
  gt_multiple(c(names(melidos_cities), "Overall"), merge_desc_columns) |> 
  cols_hide(type2) |> 
  fmt_number(columns = !starts_with("n_"), rows = type2 %in% c("dynamics", "spectrum")) |> 
  fmt_number(columns = !starts_with("n_"), rows = type2 %in% c("exposure history"),
             decimals = 2) |> 
  fmt_number(columns = !starts_with("n_"), rows = type2 %in% c("level"),
             decimals = 1) |> 
  fmt(columns = !c(starts_with("n_"), name:scaling, type2),
      rows = type2 %in% c("timing", "duration"), 
      fns = \(x) {
        x |> hms::hms(hours = _) |> strptime(format = "%H:%M:%S") |> format(format = "%H:%M")
        }
      ) |> 
  as.data.frame() |> 
  select(1:Overall)
```

### Finish table

```{r}
tbl2 <- 
tbl2_data_formatted |> 
  mutate(metric_type = metric_type |> str_to_title(),
         across(c(name, scaling),
                \(x) x |> str_to_title() |>  str_replace_all("_", " ")),
         Unit = c(rep("Hrs:Mins (duration)", 5),
                  rep(NA, 2),
                  "klx·h",
                  rep("lx", 3),
                  NA,
                  rep("HH:MM (time-of-day)", 5)
                  ),
         .after = name) |> 
  group_by(metric_type) |> 
  gt(rowname_col = "name") |> 
  cols_hide(type2) |> 
  sub_values(values = "Mder", replacement = "MDER") |> 
  tab_header("Metric descriptive summary (glasses)") |> 
  gt_multiple(names(melidos_cities), style_tab) |> 
  cols_move(Overall, scaling) |> 
  cols_label(
    scaling = "Scaling"
  ) |> 
  tab_footnote(
    md("**Median** (25% percentile, 75% percentile), *Mean ± standard deviation* , <span style = 'color:grey'>n=observations</span>")
  )|> 
  tab_footnote(
    "Scaling refers to the distributional characteristics of the data and the scaling of the Plot column. Metric-values are on a linear scale. Log 10 zero inflated: logarithmic scaling (base 10), but also contains zero values (offset +0.1); Identity: linear scale",
    locations = cells_column_labels(scaling)
  ) |>
  cols_add(Plot = tbl2_data$plot |> 
             gt::ggplot_image(
                       height = gt::px(80), 
                       aspect_ratio = 2.4
                     )) |> 
  cols_label(
    Plot = "Distribution"
  ) |> 
  tab_footnote(
    md("Red horizontal lines indicate the median of a site. If the scaling column indicates *Log 10 zero inflated* then values were transformed accordingly before plotting"),
    locations = cells_column_labels(Plot)
  ) |>
  fmt_markdown() |> 
  cols_width(
             c(Overall, matches(names(melidos_cities))) ~ px(120),
             Plot ~ px(200),
             everything() ~ px(100)) |> 
  sub_missing() |> 
    tab_style(style = cell_text(weight = "bold"),
            locations = list(
              cells_column_labels(),
              cells_row_groups(),
              cells_title()
            )) |> 
  cols_move(scaling, UCR) |> 
  site_conv_gt(rev = FALSE)
```

```{r}
tbl2
gtsave(tbl2, "tables/tbl2.png", vwidth = 1750)
```

## Figure: Overview

### Panel A: Protocol

```{r}
path <- "assets/2026-03-30_MeLiDos_Protocol.png"
P_A <- 
ggdraw() +
  draw_image(path)
P_A
```

### Panel B: Overview


```{r}
P_B <-
  light_glasses_processed2 |> 
  rbind(light_chest_processed2) |> 
  distinct(site, Id, Datetime) |>
  mutate(site = fct(site, levels = names(melidos_cities))) |>
  site_conv_mutate() |> 
  gg_overview(Id.colname = site, gap.data = tibble(),
              color = site) +
  scale_color_manual(values = melidos_colors) +
  guides(color = "none") +
  labs(y = NULL, x = "Collection dates")
P_B
```

### Panel C: Location

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
    
    countries_colors <- tibble(
      country = melidos_countries,
      color   = melidos_colors,
      stringsAsFactors = FALSE
    )
    
    world$color <- ifelse(
      world$name %in% countries_colors$country,
      countries_colors$country[match(world$name, countries_colors$country)],
      NA
    )
    
    # 3) Locations in geographic CRS (EPSG:4326)
    location_info <- tibble(
      country  = melidos_countries,
      location = melidos_cities,
      lat      = melidos_coordinates |> map(1) |> unlist(),
      lon      = melidos_coordinates |> map(2) |> unlist(),
      color    = melidos_countries,
      stringsAsFactors = FALSE
    ) |> 
      rowwise() |> 
      mutate(
        coordinate_string = format_coordinates(c(lat,lon)),
        label = 
          paste0(location, ", ", country, " (", coordinate_string, ")")
        )
    
    locations <- st_as_sf(location_info, coords = c("lon", "lat"), crs = 4326)
    
    # 4) Project both layers to a planar CRS
    # robinson_crs <- paste0("+proj=", map_projection)
    robinson_crs <- "+proj=eqc"
    world_proj     <- st_transform(world,    crs = robinson_crs)
    locations_proj <- st_transform(locations, crs = robinson_crs)
    
    bb <- st_bbox(world_proj)
    y_offset <- 0.08 * (bb["ymax"] - bb["ymin"]) 
    
    # 5) Plot
P_C <-
      ggplot() +
      geom_sf(data = world_proj, aes(fill = color),
              color = NA, size = 0.25, alpha = 0.5, show.legend = FALSE) +
      geom_sf(data = locations_proj, aes(fill = color),
              shape = 21, color = "black", size = 3, stroke = 0.2) +
      geom_label_repel(
        data = locations_proj,
        stat = "sf_coordinates",
        label.size = 0,
        aes(label = label, fill = color, geometry = geometry),
        color = c(rep("black", 8), "white"),
        min.segment.length = 0.75,
        size = 3,
        alpha = 0.8,
        box.padding = 0.35,    # space between labels
        max.overlaps = 20,
        point.padding = 0.25,  # space from point to label
        force = 2.5,             # stronger repulsion
        force_pull = 0.25,      # weaker pull back to point
        seed = 123
      ) +
      scale_fill_manual(values = 
                          countries_colors |> 
                          select(country, color) |> 
                          deframe()) +
      theme_cowplot() +
      theme(legend.position = "none") +
      theme_sub_axis(line = element_blank()) +
      labs(x = NULL, y = NULL) +
      coord_sf(expand = FALSE)
P_C
```

### Panel D: Doubleplot

```{r}
source("scripts/Brown_bracket.R")
```

```{r}
P_D_data <-
      light_glasses_processed2 |>
      ungroup() |>
      mutate(across(where(is.POSIXct),
                    \(x) {
                      date(x) <- "2025-04-03"
                      x }),
             across(c(dawn, dusk),
                    LightLogR:::datetime_to_circular)
             ) |>
      aggregate_Datetime(
        unit = "15 mins",
        type = "floor",
        numeric.handler = \(x) median(x, na.rm = TRUE),
        upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
        upper75 = quantile(MEDI, 0.875, na.rm = TRUE),
        upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
        lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
        lower75 = quantile(MEDI, 0.125, na.rm = TRUE),
        lower95= quantile(MEDI, 0.025, na.rm = TRUE),
      ) |> 
        mutate(
          across(c(dawn, dusk), LightLogR:::circular_to_hms),
          across(c(dawn, dusk), 
                 \(x) {
                      x <- x |> strptime("%H:%M:%S", tz = tz(Datetime))
                      date(x) <- "2025-04-03"
                      x |> as.POSIXct() }),
               )

P_D_data_site <- 
    light_glasses_processed2 |>
    ungroup(Id) |> 
    select(site, Datetime, MEDI) |> 
    aggregate_Date(
        unit = "15 mins",
        type = "floor",
        numeric.handler = \(x) {
          median(x, na.rm = TRUE)
          },
        date.handler = \(x) median(x, na.rm = TRUE),
        upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
        upper75 = quantile(MEDI, 0.875, na.rm = TRUE),
        upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
        lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
        lower75 = quantile(MEDI, 0.125, na.rm = TRUE),
        lower95= quantile(MEDI, 0.025, na.rm = TRUE),
      ) |> 
    mutate(
      across(where(is.POSIXct),
                    \(x) {
                      date(x) <- "2025-04-03"
                      x })
    )

P_D_data_site <- 
  P_D_data_site |> 
  rbind(
    P_D_data_site |> 
    mutate(
      across(where(is.POSIXct),
                    \(x) {
                      date(x) <- "2025-04-04"
                      x })
    )  
    )

# site_color <- ggsci::pal_jco()(1)
site_color <- "black"

P_D <- 
P_D_data |> 
      mutate(sleep = replace_values(sleep, "wake" ~ NA),
             ) |> 
      gg_doubleplot(geom = "blank",
                    facetting = FALSE,
                    # jco_color = FALSE,
                    x.axis.label = "Local time (HH:MM)",
                    y.axis.label = "Melanopic EDI (lx)",
                    aes_fill = site
                    ) |> 
      gg_photoperiod(fill = "darkblue", alpha = 0.1) |> 
      gg_states(sleep, aes_fill = sleep, ymax = -0.1, fill = "red", alpha = 1,
                on.top = TRUE) +
      geom_ribbon(aes(ymin = lower95, ymax = upper95), alpha = 0.4) +
      geom_ribbon(aes(ymin = lower75, ymax = upper75), alpha = 0.4) +
      geom_ribbon(aes(ymin = lower50, ymax = upper50), alpha = 0.4) +
      geom_line(data = P_D_data_site |>       site_conv_mutate(),
                aes(y=MEDI, colour = site), linewidth = 1) +
      geom_line(aes(y = MEDI), linewidth = 2) +
      map(c(1,10,250), 
          \(x) geom_hline(aes(yintercept = x), col = "grey", linetype = "dashed")
      ) +
      scale_color_manual(values = melidos_colors) +
      coord_cartesian(ylim = c(0, 100000)) +
      guides(y = guide_axis_stack(Brown_bracket, "axis"), fill = "none",
             color = "none") +
      # labs(x = NULL)
      labs(
        caption = glue(
          "<i>daytime</i>, <i>evening</i>, and <i>sleep</i> indicate 
          recommendations for healthy light exposure (Brown et al., 2022). 
          <br><b>Median</b> with <b style = 'color:{alpha(site_color, 
          alpha = 0.75)}'>50%</b>, <b style = 'color:{alpha(site_color, 
          alpha = 0.5)}'>75%</b>, or <b style = 'color:{alpha(site_color, 
          alpha = 0.25)}'>95%</b> of data. The horizontal bars show
          <b style = 'color:red'>average sleep times</b>. The vertical bars show 
          <b style = 'color:{alpha('darkblue', alpha = 0.4)}'>average nighttimes</b>."
        )
      ) +
      theme(plot.caption = ggtext::element_markdown())

P_D
```

### Panel E: Photoperiods

```{r}
tz <- light_glasses_processed2 |> pull(Datetime) |> tz()

photoperiods <- metrics_chest$data[[3]] |> pull(photoperiod)

photoperiods_range <- photoperiods |> range() |> round(1)

photoperiods_seq <- 
      inject(seq(!!!photoperiods_range, by = 1) |> 
               round(1))

P_E_data <-
      metrics_glasses$data[[3]] |> 
      rbind(metrics_chest$data[[3]]) |> 
      distinct(site, Id, Date, .keep_all = TRUE) |> 
      mutate(site = site |> fct_rev()) |> 
      site_conv_mutate()

P_E <- 
P_E_data |>   
      ggplot(aes(x=photoperiod)) +
      geom_boxplot(aes(y= site, col = site), width = 0.25,
                   position = position_nudge(y = -0.25)) +
      geom_density_ridges(aes(y=site, fill = site),
                          bandwidth = 0.25,
                          alpha = 0.8) +
      scale_x_continuous(breaks = 10:21) +
      scale_color_manual(values = melidos_colors) +
      scale_fill_manual(values = melidos_colors) +
      labs( 
        y = NULL, 
        x = "Photoperiod (hours)"
        ) +
      theme_cowplot() +
      guides(
        fill = "none", color = "none") +
      coord_cartesian(xlim = photoperiods_range) +
      theme(axis.title.x = ggtext::element_markdown())
P_E
```


### Plot composition

```{r}
#| fig-height: 15
#| fig-width: 16
P_A / 
  ((P_C + P_B) + plot_layout(widths = c(3,2))) / 
  ((P_E + P_D) + plot_layout(widths = c(2,3))) + 
  plot_layout(heights = c(1.3,1,1)) +
  plot_annotation(tag_levels = "A")

ggsave("figures/Fig1.pdf", height = 10, width = 10.5, scale = 1.5)
ggsave("figures/Fig1.png", height = 10, width = 10.5, scale = 1.5)
```

## Figure: Site plots

```{r}
Fig2_data <-
      light_glasses_processed2 |>
      ungroup(Id) |>
      mutate(across(where(is.POSIXct),
                    \(x) {
                      date(x) <- "2025-04-03"
                      x }),
             across(c(dawn, dusk),
                    LightLogR:::datetime_to_circular)
             ) |>
      aggregate_Datetime(
        unit = "15 mins",
        type = "floor",
        numeric.handler = \(x) median(x, na.rm = TRUE),
        upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
        upper75 = quantile(MEDI, 0.875, na.rm = TRUE),
        upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
        lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
        lower75 = quantile(MEDI, 0.125, na.rm = TRUE),
        lower95= quantile(MEDI, 0.025, na.rm = TRUE),
      ) |> 
        mutate(
          across(c(dawn, dusk), LightLogR:::circular_to_hms),
          across(c(dawn, dusk), 
                 \(x) {
                      x <- x |> strptime("%H:%M:%S", tz = tz(Datetime))
                      date(x) <- "2025-04-03"
                      x |> as.POSIXct() }),
               )
```

```{r}
#| fig-width: 18
#| fig-height: 12
# Fig2 <- 
Fig2_data |> 
      mutate(sleep = replace_values(sleep, "wake" ~ NA),
             ) |> 
      site_conv_mutate(rev = FALSE) |> 
      gg_doubleplot(geom = "blank",
                    facetting = FALSE,
                    jco_color = FALSE,
                    x.axis.label = "Local time (HH:MM)",
                    y.axis.label = "Melanopic EDI (lx)",
                    aes_fill = site
                    ) |> 
      gg_photoperiod() |> 
      gg_states(sleep, aes_fill = sleep, ymax = -0.1, fill = "red", alpha = 1,
                on.top = TRUE) +
      facet_wrap(~site, nrow = 3, ncol = 3, dir = "lt") + 
      geom_ribbon(aes(ymin = lower95, ymax = upper95, fill = site), alpha = 0.4) +
      geom_ribbon(aes(ymin = lower75, ymax = upper75, fill = site), alpha = 0.4) +
      geom_ribbon(aes(ymin = lower50, ymax = upper50, fill = site), alpha = 0.4) +
      geom_line(aes(y = MEDI, col = site), linewidth = 0.5, layout = "fixed") +
      geom_line(aes(y = MEDI), linewidth = 1.5) +
      map(c(1,10,250), 
          \(x) geom_hline(aes(yintercept = x), col = "grey", linetype = "dashed")
      ) +
      scale_color_manual(values = melidos_colors) +
      scale_fill_manual(values = melidos_colors) +
      coord_cartesian(ylim = c(0, 100000)) +
      guides(y = guide_axis_stack(Brown_bracket, "axis"), fill = "none",
             color = "none") +
      # labs(x = NULL)
      labs(
        caption = glue(
          "<i>daytime</i>, <i>evening</i>, and <i>sleep</i> indicate 
          recommendations for healthy light exposure (Brown et al., 2022). 
          <br><b>Median</b> with <b style = 'color:{alpha(site_color, 
          alpha = 0.75)}'>50%</b>, <b style = 'color:{alpha(site_color, 
          alpha = 0.5)}'>75%</b>, or <b style = 'color:{alpha(site_color, 
          alpha = 0.25)}'>95%</b> of data. The horizontal bars show
          <b style = 'color:red'>average sleep times</b>. The vertical bars show 
          <b style = 'color:grey'>average nighttimes</b>."
        )
      ) +
      theme(plot.caption = ggtext::element_markdown(),
            panel.spacing.x = unit(30, "pt"))

ggsave("figures/Fig2.pdf", height = 8, width = 12, scale = 1.5)
ggsave("figures/Fig2.png", height = 8, width = 12, scale = 1.5)

```

```{r}
walk(names(melidos_cities), 
     \(x) Fig2_data |> 
            Fig2_plot_individual(x) |> 
            ggsave(plot = _, 
                   filename = str_c("figures/Fig2_individual/Fig2_",x ,".pdf"), 
                   width = 8, 
                   height = 5
                   ) 
     )

```

## Figure: Metric plots

```{r}
Fig3_data <- 
tbl2_data |> 
  mutate(name = 
           name |> 
           str_to_sentence() |> 
           str_replace_all("_", " ") |> 
           str_replace("(1000$|250 |250$|10 |1 )",
                       "\\1 lx mel EDI ") |> 
           str_replace("Mder", "MDER"),
         name = name |> 
           replace_when(
             metric_type == "duration" ~ str_c(name, " (HH:MM)"),
             metric_type == "level" ~ str_c(name, " (lx; from geometric mean)"),
             metric_type == "exposure history" ~ str_c(name, " (klx·h)"),
             metric_type == "timing" ~ str_c(name, " (HH:MM)")
           ),
         data = data |> map2(metric_type, \(x,y) {
           if(y %in% c("duration", "timing")) {
           x |> mutate(metric = metric*60*60)
         } else x
         })
         )

Fig3_data <- 
  Fig3_data |> 
  mutate(
    plot = pmap(list(data, scaling, metric_type, name), Fig3_plot)
  )
```

```{r}
#| fig-height: 20
#| fig-width: 20
#| message: false
#| warning: false
Fig3 <-
  Fig3_data |> 
  filter_out(name == "Darkest 10h mean (lx; from geometric mean)") |> 
  pull(plot) |> 
  reduce(\(plots, plot) `+`(plots, plot))
Fig3 + plot_annotation(tag_levels = "A") &
  theme(
    plot.tag = element_text(face = "bold")
  )

ggsave("figures/Fig3.pdf", width = 10, height = 10, scale = 2)
ggsave("figures/Fig3.png", width = 10, height = 10, scale = 2)
```

```{r}
walk2(Fig3_data$plot, 
      LETTERS[1:17], 
      \(x,y) ggsave(str_c("figures/Fig3_individual/Fig3_", 
                          y, 
                          ".pdf"), 
                    x+ theme_sub_axis_left(text = element_text()), 
                    width = 3.5,
                    height = 5)
      )
```


## Figure: latitude x photoperiod

```{r}
span_phot <- tibble(Datetime = as.POSIXct(ymd("2025-01-01"), tz = "UTC") + ddays(0:364))

phot_potential <-
0:60 |> 
  map(
   \(x) span_phot |> 
        extract_photoperiod(c(x,0)) |> 
        group_by(lat) |> 
        summarize(min = min(photoperiod), 
                  max = max(photoperiod))
) |> list_rbind()

phot_potential_range <- 
  phot_potential |> summarize(min = min(min), max = max(max)) |> unlist()

P_E_data |> 
  mutate(site = fct_rev(site)) |> 
  # site_conv_mutate(rev = TRUE) |>
  ggplot(aes(y = lat)) + 
  geom_jitter(aes(col = site, x = photoperiod), width = 0, height = 0.5, alpha = 0.4) +
  geom_density_ridges(aes(y=lat, fill = site, x = photoperiod),
                      scale = 2,
                          position = position_nudge(y = 1),
                          bandwidth = 0.15,
                          alpha = 0.8) +
  geom_ribbon(data = phot_potential, aes(xmin = 0, xmax = min), fill = "black") +
  geom_ribbon(data = phot_potential, aes(xmin = max, xmax = 24), fill = "black") +
  annotate("text", y = 3, x = 15.2, label = "possible photoperiods", color = "white",
           hjust = 0) +
  annotate("curve", y = 3, x = 15,
           xend = 13.3, yend = 5.1,
           curvature = -0.1,
           arrow = arrow(type = "closed",
                         length = unit(0.2, "cm")),
           color = "white") +
  theme_cowplot() + 
  guides(color = "none") +
  coord_cartesian(ylim = c(0, 63), expand = FALSE, xlim = phot_potential_range) +
  scale_color_manual(values = melidos_colors) +
  scale_fill_manual(values = melidos_colors) +
  labs(color = "Site", x = "Photoperiod (hr)", y = "absolute Latitude (°)",
       caption = "A jitter of height 0.5 was applied to points.\nPhotoperiod is defined as the timespan of daily sun elevation above -6°") +
  theme( 
    # panel.background = element_rect(fill = "black"),
        legend.position = "inside",
        legend.text = element_text(colour = "white"),
        legend.position.inside = c(0.65,0.4))

ggsave("figures/photoperiod_potential_ridges.pdf", width = 6, height = 6)
ggsave("figures/photoperiod_potential_ridges.png", width = 6, height = 6)
```

## Table: Personal light exposure vs. recommendations

```{r}
Brown.by.site <-
light_glasses_processed2 |> 
  Brown2reference(
    Brown.day = "wake",
    Brown.evening = "pre-sleep",
    Brown.night = "sleep"
  ) |> 
  drop_na(Reference.check) |> 
  group_by(site, State.Brown, Reference.check) |> 
  durations() |> 
  pivot_wider(names_from = Reference.check, names_prefix = "check_", values_from = duration) |> 
  mutate(adhered = check_TRUE,
         duration = check_FALSE + check_TRUE,
         adherence = check_TRUE / duration)

Brown.by.site.total <-
Brown.by.site |> 
  group_by(site) |> 
  summarize(adhered = sum(adhered) |> as.duration(),
            duration = sum(duration) |> as.duration(),
            adherence = adhered/duration,
            State.Brown = "total")

Brown.by.site <- 
Brown.by.site |> 
  select(site, State.Brown, adhered, duration, adherence) |> 
  bind_rows(Brown.by.site.total) |> 
  pivot_wider(names_from = State.Brown, values_from = c(adhered, duration, adherence))

Brown_missing <-
  light_glasses_processed2 |> 
  Brown2reference(
    Brown.day = "wake",
    Brown.evening = "pre-sleep",
    Brown.night = "sleep"
  ) |> 
  group_by(site) |> 
  durations(Reference.check, show.missing = TRUE) |> 
  mutate(missing_pct = missing/total) |> 
  select(site, missing_pct, missing, total)

Brown_overall <- 
Brown.by.site |> 
  left_join(Brown_missing, by = "site") |> 
  ungroup() |> 
  summarize(site = "Overall", 
            adherence_wake = sum(adhered_wake)/sum(duration_wake),
            `adherence_pre-sleep` = sum(`adhered_pre-sleep`)/sum(`duration_pre-sleep`),
            adherence_sleep = sum(adhered_sleep)/sum(duration_sleep),
            adherence_total = sum(adhered_total)/sum(duration_total),
            fraction_wake = sum(duration_wake)/sum(total),
            `fraction_pre-sleep` = sum(`duration_pre-sleep`)/sum(total),
            fraction_sleep = sum(duration_sleep)/sum(total),
            fraction_missing = sum(missing)/sum(total))

Brown_summary <-
Brown.by.site |> 
  left_join(Brown_missing, by = "site") |> 
  mutate(
    fraction_wake = (duration_wake)/(total),
    `fraction_pre-sleep` = (`duration_pre-sleep`)/(total),
    fraction_sleep = (duration_sleep)/(total),
    fraction_missing = (missing)/(total)
  ) |> 
  select(-starts_with("adhered_"), -starts_with("duration_"), -total, -missing_pct, -missing) |> 
  site_conv_mutate(rev = FALSE) |> 
  arrange(site)
```


```{r}
source("scripts/RQ2_specific.R")
Brown_table <- 
bind_rows(Brown_overall, Brown_summary) |> 
  gt() |> 
  gt::rows_add(site = NA, .after = 1)|> 
  tab_style(
    style = list(css(`padding-top` = "0px",
            `padding-bottom` = "0px"),
            cell_fill("lightgrey")
            ),
    locations = cells_body(rows = 2)) |>
  sub_missing(missing_text = "") |> 
  fmt_percent(decimals = 0) |> 
  tab_spanner("Adherence to recommendations during:", starts_with("adherence_")) |> 
  tab_spanner("Percentage of measured time:", starts_with("fraction_")) |> 
  cols_label_with(fn = \(x) str_remove(x, "adherence_|fraction_") |> str_to_sentence()) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_body(site),
      cells_column_labels(),
      cells_column_spanners(),
      cells_body(rows = 1)
    )
  ) |> 
  gt_multiple(
    melidos_colors |> names(),
    style_rows
  ) |>
  tab_style(
    style = cell_borders("right", "lightgrey", weight = "2px"),
    locations = cells_body(adherence_total)
  ) |> 
  tab_style(
    style = cell_text(align = "center"),
    list(cells_body(), cells_column_labels())
  ) |> 
  tab_footnote(
    "Missingness can result from either missing sleep-wake diary data for a given day, and/or missing or removed measurements of melanopic EDI",
    cells_column_labels(fraction_missing)
  ) |> 
  tab_footnote(
    "Recommendations for healthy lighting refer to Brown et al. (2022). Wake ≥ 250lx melanopic EDI, pre-sleep ≤ 10lx melanopic EDI, sleep ≤ 1lx melanopic EDI",
    cells_column_spanners(1)
  ) |> 
  tab_footnote(
    "Total is a time-weighted average",
    cells_column_labels(adherence_total)
  ) 
  
```


```{r}
Brown_table
gtsave(Brown_table, "tables/Brown_table.png")
```


And for chest-level data: 

```{r}
Brown.by.site <-
light_chest_processed2 |> 
  Brown2reference(
    Brown.day = "wake",
    Brown.evening = "pre-sleep",
    Brown.night = "sleep"
  ) |> 
  drop_na(Reference.check) |> 
  group_by(site, State.Brown, Reference.check) |> 
  durations() |> 
  pivot_wider(names_from = Reference.check, names_prefix = "check_", values_from = duration) |> 
  mutate(adhered = check_TRUE,
         duration = check_FALSE + check_TRUE,
         adherence = check_TRUE / duration)

Brown.by.site.total <-
Brown.by.site |> 
  group_by(site) |> 
  summarize(adhered = sum(adhered) |> as.duration(),
            duration = sum(duration) |> as.duration(),
            adherence = adhered/duration,
            State.Brown = "total")

Brown.by.site <- 
Brown.by.site |> 
  select(site, State.Brown, adhered, duration, adherence) |> 
  bind_rows(Brown.by.site.total) |> 
  pivot_wider(names_from = State.Brown, values_from = c(adhered, duration, adherence))

Brown_missing <-
  light_glasses_processed2 |> 
  Brown2reference(
    Brown.day = "wake",
    Brown.evening = "pre-sleep",
    Brown.night = "sleep"
  ) |> 
  group_by(site) |> 
  durations(Reference.check, show.missing = TRUE) |> 
  mutate(missing_pct = missing/total) |> 
  select(site, missing_pct, missing, total)

Brown_overall <- 
Brown.by.site |> 
  left_join(Brown_missing, by = "site") |> 
  ungroup() |> 
  summarize(site = "Overall", 
            adherence_wake = sum(adhered_wake)/sum(duration_wake),
            `adherence_pre-sleep` = sum(`adhered_pre-sleep`)/sum(`duration_pre-sleep`),
            adherence_sleep = sum(adhered_sleep)/sum(duration_sleep),
            adherence_total = sum(adhered_total)/sum(duration_total),
            fraction_wake = sum(duration_wake)/sum(total),
            `fraction_pre-sleep` = sum(`duration_pre-sleep`)/sum(total),
            fraction_sleep = sum(duration_sleep)/sum(total),
            fraction_missing = sum(missing)/sum(total))

Brown_summary <-
Brown.by.site |> 
  left_join(Brown_missing, by = "site") |> 
  mutate(
    fraction_wake = (duration_wake)/(total),
    `fraction_pre-sleep` = (`duration_pre-sleep`)/(total),
    fraction_sleep = (duration_sleep)/(total),
    fraction_missing = (missing)/(total)
  ) |> 
  select(-starts_with("adhered_"), -starts_with("duration_"), -total, -missing_pct, -missing) |> 
  site_conv_mutate(rev = FALSE) |> 
  arrange(site)
```


```{r}
source("scripts/RQ2_specific.R")
Brown_table <- 
bind_rows(Brown_overall, Brown_summary) |> 
  gt() |> 
  gt::rows_add(site = NA, .after = 1)|> 
  tab_style(
    style = list(css(`padding-top` = "0px",
            `padding-bottom` = "0px"),
            cell_fill("lightgrey")
            ),
    locations = cells_body(rows = 2)) |>
  sub_missing(missing_text = "") |> 
  fmt_percent(decimals = 0) |> 
  tab_spanner("Adherence to recommendations during:", starts_with("adherence_")) |> 
  tab_spanner("Percentage of measured time:", starts_with("fraction_")) |> 
  cols_label_with(fn = \(x) str_remove(x, "adherence_|fraction_") |> str_to_sentence()) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_body(site),
      cells_column_labels(),
      cells_column_spanners(),
      cells_body(rows = 1)
    )
  ) |> 
  gt_multiple(
    melidos_colors |> names(),
    style_rows
  ) |>
  tab_style(
    style = cell_borders("right", "lightgrey", weight = "2px"),
    locations = cells_body(adherence_total)
  ) |> 
  tab_style(
    style = cell_text(align = "center"),
    list(cells_body(), cells_column_labels())
  ) |> 
  tab_footnote(
    "Missingness can result from either missing sleep-wake diary data for a given day, and/or missing or removed measurements of melanopic EDI",
    cells_column_labels(fraction_missing)
  ) |> 
  tab_footnote(
    "Recommendations for healthy lighting refer to Brown et al. (2022). Wake ≥ 250lx melanopic EDI, pre-sleep ≤ 10lx melanopic EDI, sleep ≤ 1lx melanopic EDI",
    cells_column_spanners(1)
  ) |> 
  tab_footnote(
    "Total is a time-weighted average",
    cells_column_labels(adherence_total)
  ) 
  
```

```{r}
Brown_table
gtsave(Brown_table, "tables/chest/Brown_table.png")
```

## Figure: Timeseries to metrics

We start by downloading the german datasets from MeLiDos

```{r}
#load data
load("data/metrics_separate_glasses.RData")
#only german sites
data <- metric_glasses_participanthour |> filter(site %in% c("TUM", "BAUA", "MPI"))
#only participants with at least 6 days of data
data <- data |> sample_groups(condition = length(unique(Date)) >=6)
```

## Panel A

```{r}
symlog <- function(base = 10, thr = 1, scale = 1){
  trans <- function(x)
    ifelse(abs(x) < thr, x, sign(x) * 
             (thr + scale * (log(sign(x) * x / thr, base))))
}
```

```{r}
data_Fig1 <- 
data |> 
  add_Date_col(as.count = TRUE) |>
  ungroup() |>
  mutate(across(where(is.POSIXct),
                \(x)  as_datetime(paste0(min(as_date(Datetime)) + Date, " ", hms::as_hms(x)))
            )
         ) |>
  group_by(Id) |>
  # sample_groups(sample = 36:40) |> 
  filter(Id %in% c(
    "BAUA_S003", "BAUA_S009", "BAUA_S022", "MPI_S226", "MPI_S227",
    "TUM_S009", "MPI_S205"
  )) |> 
  filter_Date(start = "2023-08-16", length = "5 days") |> 
  group_by(Id, Date) |> 
  mutate(MEDI_sym = symlog()(MEDI),
         dur = duration_above_threshold(MEDI, Datetime, threshold = 250, na.rm = TRUE)) |> 
  group_by(Id)
```


```{r}
#| fig-width: 8
#| fig-height: 7.5
Fig1_PanelA <-
  data_Fig1 |> 
  ungroup() |> 
  mutate(Id = fct_reorder(Id, dur)) |> 
  group_by(Id) |> 
  drop_na(MEDI) |> 
  gg_days(geom = "ribbon", 
          # aes_col = MEDI >= 250, 
          aes_col = Id,
          aes_fill = Id, 
          x.axis.breaks = \(x) Datetime_breaks(x),
          x.axis.format = "%H:%M %a", 
          y.axis.breaks = c(0, 10, 250, 10000),
          # group = consecutive_id(MEDI >=250),
          facetting = FALSE,
          linewidth = 1, alpha = 0.25) |> 
  gg_photoperiod() + 
  facet_wrap(~Id, ncol = 1) +
  theme_sub_strip(text = element_blank(), placement = "inside",
                  background = element_blank(),
                  ) +
  guides(col = "none", fill = "none") +
  theme_sub_plot(margin = margin(10, 30, 10, 10)) +
  theme(
    panel.grid.major.x = 
        ggplot2::element_line(color = "grey80", linetype = 2, linewidth = 0.75),
      panel.grid.minor.x = 
        ggplot2::element_line(color = "grey80", linetype = 1, linewidth = 0.75)
  ) +
  geom_hline(yintercept = 250, linetype = 2) +
    annotate(geom = "text", y = 2000, x = as_datetime("2023-08-16 12:00:00"), 
             label = "recommended daytime minimum", hjust = 0, vjust = 0,
             fontface = "bold",
             layout = 1) + 
    annotate("curve", y = 4000, x = as_datetime("2023-08-16 11:30:00"),
           xend = as_datetime("2023-08-16 8:00:00"), yend = 250,
           curvature = 0.4, layout = 1,
           arrow = arrow(type = "closed",
                         length = unit(0.2, "cm")))

```

```{r}
data_Fig1_sum <- 
  data_Fig1 |> 
  filter(photoperiod.state == "day") |> 
  add_Date_col(as.wday = TRUE) |> 
  group_by(Id, Date) |> 
  summarize(
    duration_above_threshold(MEDI, Datetime, 
                             threshold = 250, na.rm = TRUE, as.df = TRUE),
    .groups = "drop"
  )
```

```{r}
Fig1_PanelC <-
data_Fig1_sum |> 
  ggplot(aes(x = Date, y = duration_above_250)) +
  geom_boxplot(width= 0.5) + 
  theme_cowplot() +
  scale_y_time(labels = scales::label_time("%H")) +
  labs(y = expression(TAT[250] ~ "(Hrs)"), x = NULL) +
  scale_x_discrete(expand = FALSE) +
  coord_cartesian(xlim = c(0.5,5.5))
```

```{r}
Fig1_PanelB <-
data_Fig1_sum |> 
  mutate(Id = fct_reorder(Id, duration_above_250, .desc = TRUE)) |> 
  ggplot(aes(x = duration_above_250, y = Id, fill = fct_rev(Id))) +
  geom_boxplot(alpha = 0.25) + 
  theme_cowplot() +
  ggsci::scale_fill_jco() +
  scale_x_time(labels = scales::label_time("%H")) +
  labs(x = expression(TAT[250] ~ "(Hrs)"), y = NULL) +
  guides(fill = "none") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank())
```

```{r}
Fig1_PanelD <-
data_Fig1_sum |> 
  mutate(Id = fct_reorder(Id, duration_above_250)) |> 
  ggplot(aes(x = duration_above_250)) +
  geom_histogram(
    # aes(fill = Id),
    binwidth = 3600, 
    fill = "black",
    color = "white",
    # alpha = 0.75
    ) +
  theme_cowplot() + 
  ggsci::scale_fill_jco() +
  labs(x = NULL, y = NULL) +
    theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank()) +
    scale_x_time(labels = scales::label_time("%H")) +
  guides(fill = "none")

```

```{r}
Fig1_PanelD2 <-
data_Fig1_sum |> 
  mutate(Id = fct_reorder(Id, duration_above_250)) |> 
  ggplot(aes(y=fct_rev(Id), x = Date)) +
  geom_point(aes(size = duration_above_250, col = Id), alpha = 0.5) +
  ggsci::scale_color_jco() +
  theme_cowplot() +
  labs(y = NULL, x = NULL) +
  theme(
    axis.text.y = element_blank(),
        axis.ticks = element_blank(),
    axis.line.y = element_blank()
  ) +
  scale_x_discrete(breaks = "Fri", labels = expression(Size: TAT[250])) +
  scale_size_continuous(range = c(1,6)) +
  guides(color = "none", size = "none")
```

```{r}
#| fig-width: 10
#| fig-height: 9

layout <- 
"AAAAB
AAAAB
AAAAB
AAAAB
CCCCD"

Fig1_PanelA + Fig1_PanelB + Fig1_PanelC + Fig1_PanelD2 +
  plot_layout(design = layout) +
  plot_annotation(
    tag_levels = "A",
    caption = paste(
      "TAT = Time above threshold, during daytime. Threshold is provided in the subscript and indicates melanopic EDI (lx).\nMelanopic EDI values are 30-minute averages at eye-level. ID's: ",
      paste(
    c("BAUA_S003", "BAUA_S009", "BAUA_S022", "MPI_S205", "MPI_S226", "MPI_S227",
    "TUM_S009"), collapse =  ", ")
    # "\nData source: https://github.com/MeLiDosProject"
  ))  &
  theme(
    plot.tag.position = c(-0.025, 1.025)
  )

ggsave("figures/Timeseries_to_metrics.png", height = 9, width = 10)
```

## Session Info

```{r}
sessionInfo()
```

