---
title: "Descriptives"
subtitle: "Individual, behavioural, and environmental determinants of personal light exposure in daily life: a multi-country wearable and experience-sampling study"
author: "Johannes Zauner"
format: 
  html:
    toc: true
    code-tools: true
    code-link: true
    page-layout: full
---

## Preface

This document contains descriptive plots and tables of the dataset.

## Setup

```{r}
#| message: false
#| warning: false
#| filename: Setup
library(tidyverse)
library(melidosData)
library(LightLogR)
library(cowplot)
library(ggridges)
library(gt)
library(gtsummary)
source("scripts/helpers.R")
```

## Load data

```{r}
#| message: false
load("data/metrics_glasses.RData")
load("data/metrics_chest.RData")
demographics <- load_data("demographics") |> flatten_data()
chronotype <- load_data("chronotype") |> flatten_data()
sleep <- load_data("sleepdiaries") |> flatten_data()
wearlog <- load_data("wearlog") |> flatten_data()
load("data/preprocessed_glasses_1.RData")
load("data/preprocessed_chest_1.RData")
load("data/preprocessed_glasses_2.RData")
load("data/preprocessed_chest_2.RData")
```

## Table1: Participant and site characteristics

### Prepare data

#### Site and city

```{r}
tbl1_country <-
  melidos_countries |> 
  enframe("site", "country")

tbl1_city <-
  melidos_cities |> 
  enframe("site", "city")
```

#### Participant N

```{r}
tbl1_partN <-
  demographics |> 
    count(site, name = "partN") |> 
    add_row(site = "Overall", demographics |> count(name = "partN"))

glasses_partN <- 
  light_glasses_processed1 |> summarize(Id = unique(Id), .groups = "drop")

glasses_partN <-
  glasses_partN |> 
    count(site, name = "partN_glasses") |> 
    add_row(site = "Overall", glasses_partN |> count(name = "partN_glasses"))

chest_partN <- 
  light_chest_processed1 |> summarize(Id = unique(Id), .groups = "drop")

chest_partN <-
  chest_partN |> 
    count(site, name = "partN_chest") |> 
    add_row(site = "Overall", chest_partN |> count(name = "partN_chest"))

tbl1_partN <- 
tbl1_partN |> 
  left_join(glasses_partN, by = "site") |> 
  left_join(chest_partN, by = "site") |> 
    replace_na(list(partN_chest = 0))
```

#### Participant-days N

```{r}
glasses_daysN <- 
  light_glasses_processed1 |> 
  add_Date_col(group.by = TRUE) |>
  summarize(Date = unique(Date), .groups = "drop")

tbl1_daysN <-
  glasses_daysN

glasses_daysN <-
  glasses_daysN |> 
    count(site, name = "daysN_glasses") |> 
    add_row(site = "Overall", glasses_daysN |> count(name = "daysN_glasses"))

chest_daysN <- 
  light_chest_processed1  |> 
  add_Date_col(group.by = TRUE) |> 
  summarize(Date = unique(Date), .groups = "drop")

tbl1_daysN <-
  tbl1_daysN |> full_join(chest_daysN, by = names(tbl1_daysN))

chest_daysN <-
  chest_daysN |> 
    count(site, name = "daysN_chest") |> 
    add_row(site = "Overall", chest_daysN |> count(name = "daysN_chest"))

tbl1_daysN <- 
tbl1_daysN |> 
  count(site, name = "daysN") |> 
  add_row(site = "Overall", tbl1_daysN |> count(name = "daysN")) |> 
  left_join(glasses_daysN, by = "site") |> 
  left_join(chest_daysN, by = "site") |> 
    replace_na(list(daysN_chest = 0))
```


#### Weekend/weekdays N

```{r}
glasses_weekdayN <-
  light_glasses_processed1 |> 
  add_Date_col(group.by = TRUE) |>
  mutate(Weekend = wday(Date, week_start = 1) >5) |> 
  distinct(site, Id, Date, Weekend) 

tbl1_weekdayN <-
  glasses_weekdayN

glasses_weekdayN <-
  glasses_weekdayN |> 
  ungroup(Date) |> 
  count(Weekend)|> 
    ungroup() |> 
    count(site, Weekend, name = "weekdayN_glasses", wt = n) |> 
    pivot_wider(id_cols = site, 
                names_from = Weekend, 
                values_from = weekdayN_glasses) |> 
    rename_with(\(x) x |> replace_values("FALSE" ~ "weekdayN_glasses", "TRUE" ~ "weekendN_glasses"))

glasses_weekdayN <-
  glasses_weekdayN |> 
    add_row(site = "Overall", glasses_weekdayN |> summarize(across(-site, sum)))

chest_weekdayN <-
  light_chest_processed1 |> 
  add_Date_col(group.by = TRUE) |>
  mutate(Weekend = wday(Date, week_start = 1) >5) |> 
  distinct(site, Id, Date, Weekend) 

tbl1_weekdayN <-
  tbl1_weekdayN |> full_join(chest_weekdayN, by = names(tbl1_weekdayN))

chest_weekdayN <-
  chest_weekdayN |> 
  ungroup(Date) |> 
  count(Weekend)|> 
    ungroup() |> 
    count(site, Weekend, name = "weekdayN_chest", wt = n) |> 
    pivot_wider(id_cols = site, 
                names_from = Weekend, 
                values_from = weekdayN_chest) |> 
    rename_with(\(x) x |> replace_values("FALSE" ~ "weekdayN_chest", "TRUE" ~ "weekendN_chest"))

chest_weekdayN <-
  chest_weekdayN |> 
    add_row(site = "Overall", chest_weekdayN |> summarize(across(-site, sum)))

tbl1_weekdayN <-
tbl1_weekdayN |> 
  ungroup(Date) |> 
  count(Weekend) |> 
  ungroup() |> 
  count(site, Weekend, name = "weekdayN", wt = n) |> 
  pivot_wider(id_cols = site, 
                names_from = Weekend, 
                values_from = weekdayN) |> 
    rename_with(\(x) x |> replace_values("FALSE" ~ "weekdayN", "TRUE" ~ "weekendN"))

tbl1_weekdayN <-
tbl1_weekdayN |> 
  add_row(site = "Overall", tbl1_weekdayN |> summarize(across(-site, sum))) |> 
  left_join(glasses_weekdayN, by = "site") |> 
  left_join(chest_weekdayN, by = "site") |> 
    replace_na(list(weekdayN_chest = 0))
```

#### Geolocation coordinates

```{r}
tbl1_location <-
  melidos_coordinates |> 
  enframe("site", "coordinates") |> 
  mutate(coordinates = map(coordinates, format_coordinates) |> unlist())
```

#### Photoperiod

```{r}
phot_glasses <- 
light_glasses_processed1 |> 
  add_Date_col(group.by = TRUE) |> 
  select(Id, Date, photoperiod) |> 
  distinct(site, Id, Date, .keep_all = TRUE)

phot_chest <- 
light_chest_processed1 |> 
  add_Date_col(group.by = TRUE) |> 
  select(Id, Date, photoperiod) |> 
  distinct(site, Id, Date, .keep_all = TRUE)

phot_dates <- 
  full_join(
    phot_glasses,
    phot_chest,
    by = names(phot_glasses)
  )

tbl1_photoperiod <-
  phot_dates |> 
  ungroup(Date, Id) |> 
  summarize(
    phot = median(photoperiod),
    phot_p05 = quantile(photoperiod, p=0.05),
    phot_p95 = quantile(photoperiod, p=0.95)
  ) |> 
  rbind(phot_dates |> 
  ungroup() |> 
  summarize(
    site = "Overall",
    phot = median(photoperiod),
    phot_p05 = quantile(photoperiod, p=0.05),
    phot_p95 = quantile(photoperiod, p=0.95)
  ))
```

#### Data collection dates

```{r}
tbl1_dates <- 
phot_dates |> 
  ungroup(Id, Date) |> 
  summarize(date_median = median(Date),
            date_min = min(Date),
            date_max = max(Date)
            ) |> 
  rbind(phot_dates |> 
  ungroup() |> 
  summarize(
    site = "Overall",
    date_median = median(Date),
            date_min = min(Date),
            date_max = max(Date)
  ))
```

#### Age, gender, employment status, chronotype

```{r}
partial_demographics <- 
demographics |> 
  left_join(chronotype |> select(Id, meq_type, meq, msf_sc, sjl), by = "Id") |> 
  mutate(across(c(sex, gender), forcats::fct_drop),
         meq_type = meq_type |> 
           fct_collapse(
             Morning = c("Definitely morning type", "Moderately morning type"),
             Evening = c("Definitely evening type", "Moderately evening type"),
                        ),
         employment_status = employment_status |> 
           fct_collapse(
             `Fully` = c("Full time employed", "Not employed but studying or in training", "Studying and employed"),
             `Partially` = c("Part time employed", "Marginally employed (Minijob)"),
             None = c("Not employed")
                        )
         ) |> 
  select(-c(Id, native_language, language_specification, comments))

factor_vars <- c("sex", "gender", "employment_status", "meq_type")

tbl1_demographics <-
  partial_demographics |> 
  reframe(
    .by = site,
    age_median = median(age) |> round(),
    age_min = min(age),
    age_max = max(age),
    across(c(meq),
           .fns = list(
             median = ~median(.x, na.rm =TRUE)|> round(),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE)|> round(),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)|> round()
           )), 
    across(c(sjl, msf_sc),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(all_of(factor_vars),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             })
  ) |> 
  rbind(
    partial_demographics |> 
    reframe(
    site = "Overall",
    age_median = median(age) |> round(),
    age_min = min(age),
    age_max = max(age),
    across(c(meq),
           .fns = list(
             median = ~median(.x, na.rm =TRUE)|> round(),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE)|> round(),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)|> round()
           )), 
    across(c(sjl, msf_sc),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(all_of(factor_vars),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             })
  )) |> 
  unnest(all_of(factor_vars))
```

#### Sleep duration

```{r}
tbl1_sleep <- 
sleep |> 
  mutate(daytype = daytype2 |> fct_relabel(\(x) x |> str_extract("work|free"))) |> 
  select(site, sleep_duration, daytype) |> 
  summarize(
    across(where(is.difftime),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(where(is.factor),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             }),
    .by = site
  ) |> 
  rbind(
    sleep |> 
  mutate(daytype = daytype2 |> fct_relabel(\(x) x |> str_extract("work|free"))) |> 
  select(site, sleep_duration, daytype) |> 
  summarize(
    site = "Overall",
    across(where(is.difftime),
           .fns = list(
             median = ~median(.x, na.rm =TRUE),
             p05 = ~quantile(.x, p=0.05, na.rm = TRUE),
             p95 = ~quantile(.x, p=0.95, na.rm = TRUE)
           )), 
      across(where(is.factor),
             \(x) {
               column <- cur_column()
               count(cur_data(), pick(column)) |> 
                 pivot_wider(names_from = cur_column(), 
                             values_from = "n", 
                             names_prefix = paste0(cur_column(), "_"))
             })
  )
  ) |> 
  unnest(where(is_tibble))
```

####  Non-wear time

```{r}
#| fig-height: 30
#| fig-width: 10
tbl1_wear <- 
light_glasses_processed1 |> 
  rbind(light_glasses_processed1) |>
  filter(!is.implicit) |>
  distinct(site, Id, Datetime, .keep_all = TRUE) |>
  group_by(site, Id, nonwear = wear == "off" & sleep != "sleepprep") |>
  mutate(
    nonwear = replace_na(nonwear, FALSE)
  ) |>
  durations() |>  
  pivot_wider(id_cols = c(site, Id), names_from = "nonwear", values_from = duration) |> 
  rowwise() |> 
  mutate(total = sum(dplyr::c_across(c(`FALSE`,`TRUE`)), na.rm = TRUE) |> as.duration()) |> 
  group_by(site) |> 
  summarize(
    nonwear = sum(`TRUE`, na.rm = TRUE) |> as.duration(),
    total = sum(total, na.rm = TRUE) |> as.duration(),
    nonwear_pct = nonwear/total
  ) |> 
  rbind(
    light_glasses_processed1 |> 
  rbind(light_glasses_processed1) |>
  filter(!is.implicit) |>
  distinct(site, Id, Datetime, .keep_all = TRUE) |>
  group_by(site, Id, nonwear = wear == "off" & sleep != "sleepprep") |>
  mutate(
    nonwear = replace_na(nonwear, FALSE)
  ) |>
  durations() |>  
  pivot_wider(id_cols = c(site, Id), names_from = "nonwear", values_from = duration) |> 
  rowwise() |> 
  mutate(total = sum(dplyr::c_across(c(`FALSE`,`TRUE`)), na.rm = TRUE) |> as.duration()) |> 
  ungroup() |> 
  summarize(
    site = "Overall",
    nonwear = sum(`TRUE`, na.rm = TRUE) |> as.duration(),
    total = sum(total, na.rm = TRUE) |> as.duration(),
    nonwear_pct = nonwear/total
  )
  )
  
```

#### Days included in calculations

```{r}
glasses_validdaysN <- 
  light_glasses_processed2 |> 
  add_Date_col(group.by = TRUE) |>
  summarize(Date = unique(Date), .groups = "drop")

tbl1_validdaysN <-
  glasses_validdaysN

glasses_validdaysN <-
  glasses_validdaysN |> 
    count(site, name = "validdaysN_glasses") |> 
    add_row(site = "Overall", glasses_validdaysN |> count(name = "validdaysN_glasses"))

chest_validdaysN <- 
  light_chest_processed1  |> 
  add_Date_col(group.by = TRUE) |> 
  summarize(Date = unique(Date), .groups = "drop")

tbl1_validdaysN <-
  tbl1_validdaysN |> full_join(chest_validdaysN, by = names(tbl1_validdaysN))

chest_validdaysN <-
  chest_validdaysN |> 
    count(site, name = "validdaysN_chest") |> 
    add_row(site = "Overall", chest_validdaysN |> count(name = "validdaysN_chest"))

tbl1_validdaysN <- 
tbl1_validdaysN |> 
  count(site, name = "validdaysN") |> 
  add_row(site = "Overall", tbl1_validdaysN |> count(name = "validdaysN")) |> 
  left_join(glasses_validdaysN, by = "site") |> 
  left_join(chest_validdaysN, by = "site") |> 
    replace_na(list(validdaysN_chest = 0))
```

### Combine data

```{r}
tbl1_data <-
  tbl1_country |> 
  add_row(site = "Overall") |> 
    left_join(by = "site", tbl1_city) |> 
    left_join(by = "site", tbl1_location) |> 
    left_join(by = "site", tbl1_partN) |> 
    left_join(by = "site", tbl1_daysN) |> 
    left_join(by = "site", tbl1_weekdayN) |> 
    left_join(by = "site", tbl1_wear) |> 
    left_join(by = "site", tbl1_validdaysN) |> 
    left_join(by = "site", tbl1_dates) |> 
    left_join(by = "site", tbl1_photoperiod) |> 
    left_join(by = "site", tbl1_demographics) |> 
    left_join(by = "site", tbl1_sleep)
```

```{r}
tbl1_data_formatted <- 
tbl1_data |> 
  gt() |> 
  cols_merge(starts_with("partN"), pattern = "{1}<br>(**g:{2}**, c:{3})") |> 
  cols_merge(starts_with("daysN"), pattern = "{1}d<br>(**g:{2}d**, c:{3}d)") |> 
  cols_merge(starts_with("date_"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("validdaysN"), pattern = "{1}d<br>(**g:{2}d**, c:{3}d)") |> 
  cols_merge(starts_with("phot"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("age"), pattern = "{1}yr<br>({2}yr, {3}yr)") |> 
  cols_merge(starts_with("meq_type_"), pattern = "<<{1}m/>>{2}i<</{3}e>>") |> 
  cols_merge(c("meq_median","meq_p05", "meq_p95"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("sjl"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("sex_"), pattern = "{1}F/{2}M") |> 
  cols_merge(starts_with("gender_"), pattern = "{1}W/{2}M<</{3}NB>>") |> 
  cols_merge(starts_with("nonwear"), pattern = "{1} ({2})") |> 
  cols_merge(starts_with("msf_sc"), pattern = "{1}<br>({2}, {3})") |> 
  cols_merge(starts_with("sleep_duration"), pattern = "{1}\n({2}, {3})") |> 
  cols_merge(starts_with("employment_status"), pattern = "{1}F<</{2}P>><</{3}N>>") |> 
  cols_merge(c(weekdayN, weekendN), pattern = "{1}wd / {2}we") |> 
  cols_merge(c(daytype_free, daytype_work), pattern = "{2}work / {1}free") |> 
  cols_hide(c(starts_with("weekdayN_"), starts_with("weekendN_"), daytype_NA)) |> 
  fmt_percent(nonwear_pct, decimals = 1) |> 
  fmt_integer(c("meq_median","meq_p05", "meq_p95")) |> 
  fmt_date(starts_with("date_"), date_style = "y.mn.day") |> 
  fmt(starts_with("msf_sc"), fns = \(x) x |> strptime(format = "%H:%M:%S") |> format(format = "%H:%M")) |> 
  fmt_duration(all_of(c("total")),
               input_units = "secs", max_output_units = 2
               ) |> 
  fmt_duration(all_of(c("nonwear", "phot", "phot_p05", "phot_p95", 
                        "sjl_median", "sjl_p05", "sjl_p95", 
                        "sleep_duration_median", "sleep_duration_p05", "sleep_duration_p95")),
               input_units = "secs", max_output_units = 2
               ) |> 
  cols_move(nonwear, after = total) |> 
  cols_move(daytype_free, after = weekdayN) |> 
  cols_move(c(meq_median, sjl_median), after = meq_type_Morning) |> 
  cols_label_with(everything(), fn= str_to_sentence) |> 
  as.data.frame() |> 
  select(site:partN, daysN, weekdayN, daytype_free, total, nonwear, validdaysN, 
         date_median, phot, age_median, sex_Female, gender_Woman,
         employment_status_Fully, meq_type_Morning, msf_sc_median, meq_median, sjl_median, 
         sleep_duration_median)

tbl1_transposed <- 
tbl1_data_formatted |> 
  t()
  
colnames(tbl1_transposed) <- tbl1_transposed[1,]

tbl1_transposed <- 
tbl1_transposed[-1,] |> 
  as_tibble(rownames = "description") |> 
  mutate(
    description = replace_values(
      description,
      "country" ~ "Country",
      "city" ~ "City",
      "coordinates" ~ "Coordinates",
    "partN" ~ "Participants",
    "daysN" ~ "Participant-days",
    "weekdayN" ~ "Weekday / weekend",
    "daytype_free" ~ "Workday / free",
    "date_median" ~ "Collection dates",
    "total" ~ "Participant time",
    "nonwear" ~ "Nonwear time (%)",
    "validdaysN" ~ "Complete days (>80% data)",
    "phot" ~ "Photoperiod",
    "age_median" ~ "Age",
    "meq_median" ~ "Chronotype (MEQ-score)",
    "sex_Female" ~ "Sex",
    "gender_Woman" ~ "Gender",
    "employment_status_Fully" ~ "Employment status",
    "sleep_duration_median" ~ "Sleep duration",
    "meq_type_Morning" ~ "Chronotype (MEQ-based)",
    "sjl_median" ~ "Social jetlag",
    "msf_sc_median" ~ "Chronotype (MCTQ-score)"
    ),
    type = c(rep("Site specifics", 3),rep("Data collection", 9), rep("Participant information", 9))
  ) |> 
  relocate(Overall, all_of(names(melidos_cities) |> sort()))

```

### Finish table

```{r}
tbl1 <- 
tbl1_transposed |> 
  group_by(type) |> 
  gt(rowname_col = "description") |> 
  # cols_move(Overall, 1) |> 
  sub_missing() |> 
  gt_multiple(names(melidos_cities), style_tab) |> 
  tab_style(style = cell_text(align = "center"), 
            locations = list(cells_body(),
                             cells_column_labels())) |> 
  tab_style(style = cell_text(weight = "bold", ), 
            locations = list(cells_row_groups(),
                             cells_column_labels())) |> 
  fmt_markdown() |> 
  tab_footnote(
    "g: glasses-mounted measurement device; c: chest-level measurement device",
    locations = cells_stub(c("Participants", "Participant-days", "Complete days (>80% data)"))
  ) |> 
  tab_footnote(
    "Median (minimum, maximum)",
    locations = cells_stub(c("Collection dates", "Age"))
  ) |> 
  tab_footnote(
    "Median (5% percentile, 95% percentile)",
    locations = cells_stub(c("Photoperiod", "Chronotype (MCTQ-score)", 
                             "Chronotype (MEQ-score)", "Social jetlag",
                             "Sleep duration"))
  ) |> 
  tab_footnote(
    "w: weeks, d: days, h: hours, m: minutes, s: seconds",
    locations = cells_stub(c("Participant time", "Participant-days", "Nonwear time (%)", 
                             "Complete days (>80% data)", "Photoperiod", "Social jetlag",
                             "Sleep duration"))
  ) |> 
  tab_footnote(
    "F: female, M: male(sex)/man(gender), W: woman, NB: non-binary",
    locations = cells_stub(c("Sex", "Gender"))
  ) |> 
  tab_footnote(
    "F: Fully employed OR Studying and employed OR Not employed but studying or in training; P: Part time employed OR Marginally employed (Minijob); N: Unemployed",
    locations = cells_stub(c("Employment status"))
  ) |> 
  tab_footnote(
    "Following the MEQ score; m: morning type; i: intermediary type; e: evening type",
    locations = cells_stub(c("Chronotype (MEQ-based)"))
  ) |> 
  tab_footnote(
    "MEQ: Morningness-Eveningness-Questionnaire (circadian preference); MCTQ: Munich Chronotype Questionnaire (midsleep on free days, sleep corrected",
    locations = cells_stub(c("Chronotype (MEQ-score)", "Chronotype (MCTQ-score)"))
  )

tbl1

gtsave(tbl1, "tables/tbl1.png", vwidth = 1200)
```

## Table2: Metric distribution

### Prepare data

```{r}
load("data/H1_results.RData")
```

```{r}
metrics <-
metrics |> 
  select(name, metric_type, data,  response, engine) |> 
  filter(!is.na(response)) |>
  mutate(scaling = recode_values(response,
                                 "log_zero_inflated(metric)" ~ "log_zero_inflated",
                                 default = "identity") |> 
           replace_when(
             engine == "glmmTMB" ~ "log_zero_inflated"
           )) |> 
  select(-c(response, engine))
```

### Calculate site metrics

```{r}
tbl2_data <- 
metrics |> 
  mutate(summary = map(data, 
                      \(x) {
                        x |> 
                          summarize(
                            .by = site,
                            median = median(metric, na.rm = TRUE),
                            sd = sd(metric, na.rm = TRUE),
                            p25 = quantile(metric, na.rm = TRUE, p = 0.25),
                            p75 = quantile(metric, na.rm = TRUE, p = 0.75),
                            n = sum(!is.na(metric))
                          )
                          }
                      ),
         summary_overall = map(data, 
                      \(x) {
                        x |> 
                          summarize(
                            site = "Overall",
                            median = median(metric, na.rm = TRUE),
                            sd = sd(metric, na.rm = TRUE),
                            p25 = quantile(metric, na.rm = TRUE, p = 0.25),
                            p75 = quantile(metric, na.rm = TRUE, p = 0.75),
                            n = sum(!is.na(metric))
                          )
                          }
         ),
         summary = map2(summary, summary_overall, rbind)
  ) |> 
  select(-summary_overall)
```

### Add plots

```{r}
tbl2_data <- 
tbl2_data |>
  mutate(plot = map2(data, scaling, ridges_function2),
         type2 = metric_type
         ) |> 
  arrange(metric_type, name)
```

### Constructing table

```{r}
tbl2_data_formatted <- 
tbl2_data |>
  select(-data) |>
  unnest(summary) |> 
  pivot_wider(id_cols = c(name:scaling, type2), values_from = median:n, names_from = site) |> 
  rename_with(\(y) str_remove(y, "median_")) |> 
  group_by(metric_type) |> 
  gt(rowname_col = "name") |> 
  gt_multiple(c(names(melidos_cities), "Overall"), merge_desc_columns) |> 
  cols_hide(type2) |> 
  fmt_number(columns = !starts_with("n_"), rows = type2 %in% c("dynamics", "spectrum")) |> 
  fmt_number(columns = !starts_with("n_"), rows = type2 %in% c("exposure history"),
             decimals = 0) |> 
  fmt_number(columns = !starts_with("n_"), rows = type2 %in% c("level"),
             decimals = 1) |> 
  fmt_duration(
      columns = !starts_with("n_"),
      rows = type2 == "duration",
      input_units = "hours",
      max_output_units = 2
    ) |> 
  fmt(columns = !c(starts_with("n_"), name:scaling, type2),
      rows = type2 == "timing", 
      fns = \(x) {
        x |> hms::hms(hours = _) |> strptime(format = "%H:%M:%S") |> format(format = "%H:%M")
        }
      ) |> 
  as.data.frame() |> 
  select(1:Overall)
```

### Finish table

```{r}
tbl2 <- 
tbl2_data_formatted |> 
  mutate(metric_type = metric_type |> str_to_title(),
         across(c(name, scaling),
                \(x) x |> str_to_title() |>  str_replace_all("_", " "))) |> 
  group_by(metric_type) |> 
  gt(rowname_col = "name") |> 
  cols_hide(type2) |> 
  sub_values(values = "Mder", replacement = "MDER") |> 
  tab_header("Metric descriptive summary (glasses)") |> 
  gt_multiple(names(melidos_cities), style_tab) |> 
  cols_move(Overall, scaling) |> 
  tab_style(style = cell_text(weight = "bold"),
            locations = list(
              cells_column_labels(),
              cells_row_groups(),
              cells_title()
            )) |> 
  cols_label(
    scaling = "Scaling"
  ) |> 
  tab_footnote(
    "Median ± standard deviation (25% percentile, 75% percentile), n=observations"
  )|> 
  tab_footnote(
    "Scaling refers to the distribution of the data and the scaling of the plot",
    locations = cells_column_labels(scaling)
  ) |> 
  cols_add(Plot = tbl2_data$plot |> 
             gt::ggplot_image(
                       height = gt::px(80), 
                       aspect_ratio = 2
                     )) |> 
  fmt_markdown()
```

```{r}
tbl2
gtsave(tbl2, "tables/tbl2.png", vwidth = 1700)
```

