---
title: "Research Question 1"
subtitle: "Do sites differ in objectively measured personal light exposure? If so, does an individual’s behaviour and micro-environment have a stronger effect than site?"
author: "Johannes Zauner"
format: 
  html:
    code-tools: true
    code-link: true
---

## Preface

This document contains the analysis for Research Question 1, as defined in the [preregistration](	https://aspredicted.org/te3zw2.pdf).

> **Research Question 1**: \
Do sites differ in objectively measured personal light exposure? If so, does an individual’s behaviour and micro-environment have a stronger effect than site?
> 
> *H1*: There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod.
> 
> *H2*: The variance of personal light exposure patterns (hourly geometric mean of melanopic EDI) by participants within sites is greater than the variance between sites.

## Setup

In this section, we load in the [preprocessed](data_preparation.qmd) and [prepared](metric_preparation.qmd) data, site data, and load required packages.

```{r}
#| message: false
#| warning: false
#| filename: Setup
library(tidyverse)
library(melidosData)
library(LightLogR)
library(mgcv)
library(lme4)
library(performance)
library(cowplot)
library(ggridges)
source("scripts/fitting.R")
```

```{r}
#| message: false
#| filename: Load data
load("data/metrics_glasses.RData")
```

## Hypothesis 1

### Details

Hypothesis 1 states:

> There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod.

*Type of model:* LMM

*Base formula:* $\text{Metric} \sim \text{Site} +  \text{Photoperiod} + \text{Latitude} + (1|\text{Site}:\text{Participant})$

*Additional notes:* 
- For dynamics-based metrics LM without Participant random effect.
- Photoperiod will only be applied for daytime and nighttime duration metrics.

*Specific formulas:* 
- $\text{Metric} \sim \text{Site} + \text{Photoperiod} + \text{Latitude}$ (for dynamics-based metrics)

*False-discovery-rate correction:*
`r nrow(metrics)` (number of metrics that are compared)

### Preparation

```{r}
metrics <-
  metrics |> 
  mutate(H1_formula = case_when(type == "participant" ~ 
                                  list(metric ~ site + lat + photoperiod),
                                type == "participant-day" ~ 
                                  list(metric ~ site + lat + photoperiod + (1| site:Id))),
         H1_formula0 = case_when(type == "participant" ~ 
                                   list(metric ~ lat + photoperiod),
                                 type == "participant-day" ~ 
                                  list(metric ~ lat + photoperiod + (1| site:Id)))
         )
```


### Analysis

```{r}
model_tbl <- tibble::tibble(
  name = c(
    "interdaily_stability",
    "intradaily_variability",
    "Mean",
    "dose",
    "MDER",
    "darkest_10h_mean",
    "brightest_10h_mean",
    "duration_above_1000",
    "duration_above_250",
    "duration_below_10",
    "duration_below_1",
    "duration_below_1_sleep",
    "period_above_250",
    "brightest_10h_midpoint",
    "brightest_10h_onset",
    "brightest_10h_offset",
    "darkest_10h_midpoint",
    "darkest_10h_onset",
    "darkest_10h_offset",
    "mean_timing_above_250",
    "first_timing_above_250",
    "last_timing_above_250"
  ),

  
  
  
  
  
    engine = c(
    "lm",
    "glm",
    "glmer",
    "glmer",
    "lmer",
    ""
  ),
  response = c(
    "qlogis(metric)",
    ""
  ),
  family = list(
    NULL,                        # lmer Gaussian
    Gamma(link = "log"),
    NULL,                        # e.g. log-transform beforehand
    Gamma(link = "log"),
    Gamma(link = "log")
  ),
  transform = list(
    qlogis,                      # or your own bounded-data transform
    log,
    log,
    identity,
    identity
  )
)
```


### Results


## Session info

```{r}
sessionInfo()
```

