---
title: "Research Question 1"
subtitle: "Do sites differ in objectively measured personal light exposure? If so, does an individual’s behaviour and micro-environment have a stronger effect than site?"
author: "Johannes Zauner"
format: 
  html:
    code-tools: true
    code-link: true
---

## Preface

This document contains the analysis for Research Question 1, as defined in the [preregistration](	https://aspredicted.org/te3zw2.pdf).

> **Research Question 1**: \
Do sites differ in objectively measured personal light exposure? If so, does an individual’s behaviour and micro-environment have a stronger effect than site?
> 
> *H1*: There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod.
> 
> *H2*: The variance of personal light exposure patterns (hourly geometric mean of melanopic EDI) by participants within sites is greater than the variance between sites.

## Setup

In this section, we load in the preprocessed data, site data, and load required packages.

```{r}
#| message: false
#| warning: false
#| filename: Setup
library(tidyverse)
library(melidosData)
library(LightLogR)
library(mgcv)
library(lme4)
library(performance)
library(cowplot)
library(ggridges)
```


```{r}
#| message: false
#| filename: Load data
load("data/metrics_glasses.RData")
load("data/preprocessed_glasses_2.RData")
demographics <- load_data("demographics") |> flatten_data()
```

## Preparation

### Context

```{r}
#| filename: Prepare photoperiods
photoperiod_daily <- 
light_glasses_processed2 |> 
  group_by(site, Id, Date) |> 
  summarize(photoperiod = mean(photoperiod, na.rm = TRUE),
            .groups = "drop_last")

photoperiod_participant <- 
light_glasses_processed2 |> 
  group_by(site, Id) |> 
  summarize(photoperiod = mean(photoperiod, na.rm = TRUE),
            .groups = "drop_last")
```

```{r}
#| filename: Prepare contextual info
demographics <- 
  demographics |> 
  select(site, Id, age, sex, gender, employment_status)

context <- 
  tibble(site = names(melidos_cities),
         country = unname(melidos_countries),
         tz = unname(melidos_tzs),
         city = unname(melidos_cities),
         ) |> 
  left_join(
    melidos_coordinates |> map(enframe) |> list_rbind(names_to = "site") |> pivot_wider() |> rename(lat = `1`, lon = `2`),
    by = "site"
  )

demographics <- 
  demographics |> 
  left_join(context, by = join_by("site"))

```

### Metrics

In this section, we will combine all the metrics into one data-frame

```{r}
#| filename: Prepare participant metrics
metrics <- 
metric_glasses_participant |> 
  left_join(demographics, by = c("site", "Id")) |> 
  left_join(photoperiod_participant, by = c("site", "Id")) |> 
  pivot_longer(c(interdaily_stability, intradaily_variability), names_to = "name", values_to = "metric") |> 
  ungroup() |> 
  nest(.by = name) |> 
  mutate(type = "participant")
```

```{r}
#| filename: Prepare participant-day metrics
metrics <-
  rbind(
    metrics,
    metric_glasses_participantday |>
      #converting datetimes to times
      Datetime2Time() |>
      mutate(
        #converting times and durations to decimal hours
        across(c(where(hms::is_hms), where(is.duration)), 
               \(x) as.numeric(x) / 60 / 60),
        #centering nighttime variables on midnight
        across(
          c(darkest_5h_offset, darkest_5h_onset, darkest_5h_midpoint),
          \(x) ifelse(x > 12, x - 24, x)
        )
      ) |>
      left_join(demographics, by = c("site", "Id")) |> 
      left_join(photoperiod_daily, by = c("site", "Id", "Date")) |> 
      pivot_longer(c(Mean:duration_above_250_wake), names_to = "name", values_to = "metric") |>
      ungroup() |>
      nest(.by = name) |> 
      mutate(type = "participant-day")
  )
```

### Distributional characteristics

```{r}
#| fig-width: 18
#| fig-height: 15
#| message: false
metrics |> 
  unnest(data) |> 
  ggplot(aes(x=metric)) +
  geom_density_ridges(aes(y=site, fill = site))+
  # geom_histogram(aes(y = after_stat(ncount))) +
  facet_wrap(~ name, scales = "free") +
  scale_fill_manual(values = melidos_colors) +
  theme_cowplot()

ggsave("figures/exploration/metric_distribution.pdf", width = 18, height = 15)
```


## Hypothesis 1

### Details

Hypothesis 1 states:

> There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod.

*Type of model:* LMM

*Base formula:* $\text{Metric} \sim \text{Site} +  \text{Photoperiod} + \text{Latitude} + (1|\text{Site}:\text{Participant})$

*Additional notes:* 
- For dynamics-based metrics LM without Participant random effect.
- Photoperiod will only be applied for daytime and nighttime duration metrics.

*Specific formulas:* 
- $\text{Metric} \sim \text{Site} + \text{Photoperiod} + \text{Latitude}$ (for dynamics-based metrics)

*False-discovery-rate correction:*
`r nrow(metrics)` (number of metrics that are compared)

### Preparation

```{r}


metrics <-
  metrics |> 
  mutate(H1_formula = case_when(type == "participant" ~ 
                                  list(metric ~ site + lat + photoperiod),
                                type == "participant-day" ~ 
                                  list(metric ~ site + lat + photoperiod + (1| site:Id))),
         H1_formula0 = case_when(type == "participant" ~ 
                                   list(metric ~ lat + photoperiod),
                                 type == "participant-day" ~ 
                                  list(metric ~ lat + photoperiod + (1| site:Id)))
         )
```


### Analysis

```{r}
model_tbl <- tibble::tibble(
  name = c(
    "interdaily_stability",
    "intradaily_variability",
    "Mean",
    "duration_above_250",
    "dose"
  ),
  engine = c(
    "lmer",
    "glmer",
    "lmer",
    "glmer",
    "glmer"
  ),
  family = list(
    NULL,                        # lmer Gaussian
    Gamma(link = "log"),
    NULL,                        # e.g. log-transform beforehand
    Gamma(link = "log"),
    Gamma(link = "log")
  ),
  transform = list(
    qlogis,                      # or your own bounded-data transform
    log,
    log,
    identity,
    identity
  )
)
```


### Results


## Session info

```{r}
sessionInfo()
```

