---
title: "Research Question 1"
subtitle: "Do sites differ in objectively measured personal light exposure? If so, does an individual’s behaviour and micro-environment have a stronger effect than site?"
author: "Johannes Zauner"
format: 
  html:
    toc: true
    code-tools: true
    code-link: true
    page-layout: full
---

## Preface

This document contains the analysis for Research Question 1, as defined in the [preregistration](	https://aspredicted.org/te3zw2.pdf).

> **Research Question 1**: \
Do sites differ in objectively measured personal light exposure? If so, does an individual’s behaviour and micro-environment have a stronger effect than site?
> 
> *H1*: There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod.
> 
> *H2*: The variance of personal light exposure patterns (hourly geometric mean of melanopic EDI) by participants within sites is greater than the variance between sites.

## Setup

In this section, we load in the [preprocessed](data_preparation.qmd) and [prepared](metric_preparation.qmd) data, site data, and load required packages.

```{r}
#| message: false
#| warning: false
#| filename: Setup
library(tidyverse)
library(melidosData)
library(LightLogR)
library(mgcv)
library(lme4)
library(performance)
library(cowplot)
library(ggridges)
library(broom)
library(broom.mixed)
library(gt)
library(performance)
library(glmmTMB)
library(gtsummary)
source("scripts/fitting.R")
source("scripts/summaries.R")
```

```{r}
#| message: false
#| filename: Load data
load("data/metrics_glasses.RData")
metrics <- metrics_glasses
H1 <- "Hypothesis: There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod."
sig.level <- 0.05

#for chest:
# load("data/metrics_chest.RData")
# metrics <- metrics_chest
# melidos_cities <- melidos_cities[names(melidos_cities)!="MPI"]
```

## Hypothesis 1

### Details

Hypothesis 1 states:

> There is a significant difference in personal light-exposure metrics across sites, even after accounting for latitude and photoperiod.

*Type of model:* LMM

*Base formula:* $\text{Metric} \sim \text{Site} +  \text{Photoperiod} + (1|\text{Site}:\text{Participant})$

*Additional notes:* 
- For dynamics-based metrics LM without Participant random effect.
- Because latitude and site are each individual values per site (i.e., each site has its unique latitude), they have to be modelled separately and will be compared via AIC.

*Specific formulas:* 
- $\text{Metric} \sim \text{Site} + \text{Photoperiod} $ (for dynamics-based metrics)
- $\text{Metric} \sim \text{Latitude} +  \text{Photoperiod} + (1|\text{Site}:\text{Participant})$ (for the latitude version of the base formula)
- $\text{Metric} \sim \text{Photoperiod} + (1|\text{Site}) + (1|\text{Site}:\text{Participant})$ (to check whether site is better represented as a random effect)

*False-discovery-rate correction:*
17 (number of metrics that are compared)

### Preparation

```{r}
#| filename: Model setup based on distribution
model_tbl <- tibble::tribble(
    ~name, ~engine, ~ response, ~family,
    "interdaily_stability", "lm", "qlogis(metric)", gaussian(),
    "intradaily_variability", "lm", "metric", gaussian(),
    "Mean", "lmer", "log_zero_inflated(metric)", gaussian(),
    "dose", "lmer", "log_zero_inflated(metric)", gaussian(),
    "MDER", "lmer", "metric", gaussian(),
    "darkest_10h_mean", "lmer", "log_zero_inflated(metric)", gaussian(),
    "brightest_10h_mean", "lmer", "log_zero_inflated(metric)", gaussian(),
    "duration_above_1000", "glmmTMB", "metric", tweedie(link = "log"),
    "duration_above_250", "lmer", NA, gaussian(),
    "duration_below_10", "lmer", NA, gaussian(),
    "duration_below_1", "lmer", NA, gaussian(),
    "duration_above_250_wake", "glmmTMB", "metric", tweedie(link = "log"),
    "duration_below_10_pre-sleep", "glmmTMB", "metric", tweedie(link = "log"),
    "duration_below_1_sleep", "glmmTMB", "metric", tweedie(link = "log"),
    "period_above_250", "lmer", "log_zero_inflated(metric)", gaussian(),
    "brightest_10h_midpoint", "lmer", "metric", gaussian(),
    "brightest_10h_onset", "lmer", NA, gaussian(),
    "brightest_10h_offset", "lmer", NA, gaussian(),
    "darkest_10h_midpoint", "lmer", "metric", gaussian(),
    "darkest_10h_onset", "lmer", NA, gaussian(),
    "darkest_10h_offset", "lmer", NA, gaussian(),
    "mean_timing_above_250", "lmer", "metric", gaussian(),
    "first_timing_above_250", "lmer", "metric", gaussian(),
    "last_timing_above_250", "lmer", "metric", gaussian()
  )
```

```{r}
#| filename: Formula setup for inference
metrics <-
  metrics |> 
  mutate(H1_1 = case_when(type == "participant" ~ 
                                  "~ site + photoperiod",
                                type == "participant-day" ~ 
                                  "~ site + photoperiod + (1| site:Id)"),
         H1_0 = case_when(type == "participant" ~ 
                                   "~ 1 + photoperiod",
                                 type == "participant-day" ~ 
                                  "~ 1 + photoperiod + (1| site:Id)"),
         H1_lat = case_when(type == "participant" ~ 
                                   "~ lat + photoperiod",
                                 type == "participant-day" ~ 
                                  "~ lat + photoperiod + (1|site:Id)"),
         # H1_phot = case_when(type == "participant" ~ 
         #                           "~ photoperiod",
         #                         type == "participant-day" ~ 
         #                          "~ photoperiod + (1|site:Id)"),
         H1_rand = case_when(type == "participant" ~ 
                                   NA,
                                 type == "participant-day" ~ 
                                  "~ 1 + photoperiod + (1|site/Id)")
         ) |> 
  left_join(model_tbl, by = "name")

H1_formulas <-  names(metrics) |> subset(grepl("H1_", names(metrics)))

metrics <- 
  metrics |> relocate(all_of(H1_formulas), .after = family)
H1_fdr_n <- metrics |> filter(!is.na(response)) |> nrow()
```

```{r}
#| echo: false
metrics |> 
  select(-data) |>
  filter(!is.na(response)) |> 
  arrange(metric_type, name) |> 
  mutate(family = map(family, 1),
         rowname = seq_along(1:length(name))) |> 
  group_by(metric_type) |> 
  gt() |> 
  cols_label_with(fn = \(x) x |> str_to_title() |> str_replace_all("_", " ")) |> 
  sub_missing(missing_text = "") |> 
  sub_values(values = "NULL", replacement = "") |> 
  fmt(name, fns = \(x) x |> str_to_title() |>  str_replace_all("_", " ")) |> 
  sub_values(values = "MDER", replacement = "MDER") |> 
  tab_header("Hypothesis 1: metric model specifications",
             subtitle = H1) |> 
  tab_footnote("Confirmatory analysis will use a p-value correction through the false-discovery rate for n=17 comparisons") |> 
  tab_footnote("These are the models used for the confirmatory analysis. Each H1 1 and H1 0 pair are compaired with likelihood-ratio tests to determine whether site is meaningful", locations = cells_column_labels(c(H1_1, H1_0))) |> 
  tab_footnote("These models will be checked via AIC against H1 1 to see whether there is indeed a site difference beyond the specific difference in the model. Also, this model is compared against the H1 0 model", locations = cells_column_labels(c("H1_lat",  "H1_rand")))
  # tab_footnote("These models will be checked via AIC against H1 1 to see whether there is indeed a site difference beyond the specific difference in the model. Also, this model is compared against the H1 0 model", locations = cells_column_labels(c("H1_lat", "H1_phot", "H1_rand")))
```


```{r}
#| filename: Formula creation
metrics <-
  metrics |> 
  mutate(across(all_of(H1_formulas),
                \(x) map2(x, response, \(y,z) make_formula(z, y))
                )
         )
```

```{r}
#| filename: Offset zero-values in mixed gamma models
 metrics <- 
  metrics |>
  mutate(data = map2(data, engine, \(data, engine) {
    if(engine == "glmer") {
      data |> mutate(metric = metric + 0.1)
    } else data
  }))
```


### Analysis

#### Fit models

```{r}
#| filename: Model fit
metrics <-
  metrics |> 
  mutate(
    across(all_of(H1_formulas),
           \(x) pmap(list(data, x, engine, family), fit_model),
           .names = "{.col}_model")
    )
```

#### Main model diagnostics

```{r}
#| filename: Create diagnostic plots for main model
#| message: false
metrics <-
  metrics |> 
  mutate(
    across(all_of(paste0(H1_formulas[1], "_model")),
           \(x) map(x, diagnostics),
           .names = "{str_remove(.col, '_model')}_diagnostics")
    )
```

Model diagnostics overall look good to OK. Anomalies are found but deemed non-critical. Alternative error families, like the `Tweedie` distribution have been used as well. For `Tweedie` models, the homogeneity of variance does not show in the plots (duration metrics), but was checked interactively. 

::: {.panel-tabset}

```{r, results='asis', fig.width = 8, fig.height = 12}
walk2(metrics$name, metrics$H1_1_diagnostics, function(name, plot) {
  
  # skip NULL plots
  if (is.null(plot)) return()
  
  # format tab title
  tab_title <- name |>
    str_to_title() |>
    str_replace_all("_", " ")
  
  # emit markdown for tab
  cat("\n\n##### ", tab_title, "\n\n", sep = "")
  
  # print the plot
  print(plot)
})
```

:::

### Model comparisons

```{r}
#| filename: Model comparisons
#| message: false
metrics <- 
  metrics |> 
  mutate(
    H1_1_comp = map2(H1_0_model, H1_1_model, model_anova),
    H1_1_p = map2(H1_1_comp, engine, model_comp_p, n = H1_fdr_n) |> list_c(),
    H1_1_sig = H1_1_p <= sig.level,
    
    H1_lat_comp = map2(H1_0_model, H1_lat_model, model_anova),
    H1_lat_p = map2(H1_lat_comp, engine, model_comp_p, n = H1_fdr_n) |> list_c(),
    H1_lat_sig = H1_lat_p <= sig.level,
    
    # H1_phot_comp = map2(H1_0_model, H1_phot_model, model_anova),
    # H1_phot_p = map2(H1_phot_comp, engine, model_comp_p, n = H1_fdr_n) |> list_c(),
    # H1_phot_sig = H1_phot_p <= sig.level,
    
    H1_rand_comp = map2(H1_0_model, H1_rand_model, model_anova),
    H1_rand_p = map2(H1_rand_comp, engine, model_comp_p, n = H1_fdr_n) |> list_c(),
    H1_rand_sig = H1_rand_p <= sig.level,
    
    AICs = pmap(list(H1_0_model, H1_1_model, list(NULL), H1_lat_model, H1_rand_model), model_AIC),
    
    summary = pmap(list(H1_1_sig, H1_0_model, H1_1_model), model_summary),
    summary_lat = pmap(list(H1_lat_sig, H1_0_model, H1_lat_model), model_summary),
    # summary_phot = pmap(list(H1_phot_sig, H1_0_model, H1_phot_model),
    #                     model_summary),
    summary_rand = pmap(list(H1_rand_sig, H1_0_model, H1_rand_model),
                        model_summary),
    
    table = pmap(list(name, H1_1_p, summary, metric_type, 
                           response, family), model_table),
    table_lat = pmap(list(name, H1_lat_p, summary_lat, metric_type, 
                           response, family), model_table),
    # table_phot = pmap(list(name, H1_phot_p, summary_phot, metric_type, 
    #                        response, family), model_table),
    table_rand = pmap(list(name, H1_rand_p, summary_rand, metric_type, 
                           response, family),model_table),
    table = pmap(list(table, table_lat, AICs, "lat", "H1_1", "H1_lat"),
                 H1_combining_tables),
    # table = pmap(list(table, table_phot, AICs, "photoperiod", "H1_1", "H1_phot"),
    #              H1_combining_tables),
    table = pmap(list(table, table_rand, AICs, "SD Site", "H1_1", "H1_rand"),
                 H1_combining_tables),
    table = map(table, \(x) if(is.null(x)) return() else x |>
                  mutate(across(any_of("lat"), \(x) x*10)))
    )
```

### Results

```{r}
#| message: false
#| warning: false
H1_tab <- H1_table(metrics, H1_fdr_n)
H1_tab
gtsave(H1_tab, "tables/H1_phot.png", vwidth = 1700)
```

### Interpretation

## Session info

```{r}
sessionInfo()
```

