---
title: "H06 sensitivity: employment eligibility in the near-eye sample"
subtitle: "Isolated participant-eligibility sensitivity for scientific and author review"
date: "2026-09-02"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: true
    code-summary: "Show audit code"
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r}
#| label: setup

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- locate_project_root()
source(file.path(project_root, "scripts/pipeline/p_value_display.R"))

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("This sensitivity report requires R 4.6.1", call. = FALSE)
}

artifact <- function(stage, filename) {
  file.path(
    project_root,
    "artifacts",
    stage,
    "H06",
    "employment_eligibility_sensitivity",
    filename
  )
}

read_h06_sensitivity <- function(stage, filename) {
  readr::read_csv(
    artifact(stage, filename),
    show_col_types = FALSE,
    na = ""
  )
}

scenario_contract <- read_h06_sensitivity(
  "06_model_data",
  "H06_employment_eligibility_scenario_contract.csv"
)
sample_flow <- read_h06_sensitivity(
  "06_model_data",
  "H06_employment_eligibility_sample_flow.csv"
)
formula_registry <- read_h06_sensitivity(
  "06_model_data",
  "H06_employment_eligibility_formula_registry.csv"
)
input_audit <- read_h06_sensitivity(
  "06_model_data",
  "H06_employment_eligibility_input_audit.csv"
)
package_versions <- read_h06_sensitivity(
  "06_model_data",
  "H06_employment_eligibility_package_versions.csv"
)
effects <- read_h06_sensitivity(
  "09_tables",
  "H06_employment_eligibility_effects.csv"
)
effect_comparison <- read_h06_sensitivity(
  "09_tables",
  "H06_employment_eligibility_effect_comparison.csv"
)
heterogeneity <- read_h06_sensitivity(
  "09_tables",
  "H06_employment_eligibility_heterogeneity_comparison.csv"
)
site_comparison <- read_h06_sensitivity(
  "09_tables",
  "H06_employment_eligibility_site_comparison.csv"
)
fit_diagnostics <- read_h06_sensitivity(
  "08_diagnostics",
  "H06_employment_eligibility_fit_diagnostics.csv"
)
calibration <- read_h06_sensitivity(
  "08_diagnostics",
  "H06_employment_eligibility_residual_calibration.csv"
)
residual_acf <- read_h06_sensitivity(
  "08_diagnostics",
  "H06_employment_eligibility_residual_acf.csv"
)

site_registry <- readr::read_csv(
  file.path(project_root, "config/site_display_registry.csv"),
  show_col_types = FALSE
) |>
  dplyr::arrange(.data$display_order)

predictor_labels <- c(
  work_free_day = "Free day versus work day",
  activity_status = "Active versus sedentary",
  previous_sleep_duration_centered_h =
    "Previous sleep duration, per additional hour"
)

display_predictor <- function(value) {
  output <- unname(predictor_labels[value])
  output[is.na(output)] <- value[is.na(output)]
  output
}

format_ratio <- function(estimate, low, high) {
  sprintf("%.2f × (95%% CI %.2f–%.2f)", estimate, low, high)
}

format_p <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  ifelse(display$p_bold, paste0("**", display$p_display, "**"), display$p_display)
}

compact_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(4),
      source_notes.font.size = gt::px(10)
    )
  if (!is.null(source_note)) {
    output <- gt::tab_source_note(output, source_note = source_note)
  }
  output
}

calibration_ranges <- calibration |>
  dplyr::summarise(
    calibration_low = min(.data$observed_to_fitted_ratio),
    calibration_high = max(.data$observed_to_fitted_ratio),
    .by = "model_role"
  )

acf_summary <- residual_acf |>
  dplyr::filter(.data$lag_hours %in% c(1, 6)) |>
  dplyr::select(
    .data$model_role,
    .data$lag_hours,
    .data$residual_correlation
  ) |>
  tidyr::pivot_wider(
    names_from = .data$lag_hours,
    values_from = .data$residual_correlation,
    names_prefix = "lag_"
  )

diagnostic_summary <- fit_diagnostics |>
  dplyr::left_join(calibration_ranges, by = "model_role") |>
  dplyr::left_join(acf_summary, by = "model_role")

stopifnot(
  nrow(scenario_contract) == 1L,
  scenario_contract$scenario_id == "H06-S-EMP-NE",
  identical(sample_flow$one_hour_observations, c(16596, 15871)),
  identical(sample_flow$participants, c(137, 131)),
  nrow(effect_comparison) == 3L,
  all(effect_comparison$stability_classification == "stable"),
  nrow(heterogeneity) == 3L,
  !any(heterogeneity$adjusted_conclusion_changed),
  nrow(site_comparison) == 27L,
  all(fit_diagnostics$numerical_gate_pass),
  all(input_audit$hash_verified)
)
```

# Question and fixed sensitivity contract

The accepted H06 analysis estimates average supported-hour near-eye melEDI
associations with day type, activity status, and previous sleep duration. This
sensitivity asks whether those conclusions change after restricting the fitted
sample to participants who met the registered employment requirement.

::: {.callout-note title="Answer in brief"}

Excluding six participants recorded as not employed or marginally employed
left 15,871 supported hours from 684 participant-days and 131 participants at
all nine sites. The free/work, active/sedentary, and previous-sleep results were
all **stable within model uncertainty**. Day-type heterogeneity across sites
also remained supported after false-discovery-rate adjustment. Both fitted
models passed the numerical checks. The sensitivity therefore does not change
the accepted near-eye conclusions.

:::

Employment status is the only changed analytical axis. Participants recorded
as `Not employed` or `Marginally employed (Minijob)` were removed. Students and
trainees were not excluded solely for that status, and age above 65 years was
not used as an exclusion rule. The outcome, supported-hour analysis unit,
predictors, reference categories, study sites, model formulas, equal-site
standardization, covariance estimator, confidence level, and multiplicity
families remain unchanged. Chest, paired-placement, gap-timing-unaware, and
exploratory clock-time models were not refitted.

```{r}
#| label: tbl-h06-emp-scenario-contract
#| tbl-cap: "The single analytical axis changed in the employment-eligibility sensitivity."

scenario_contract |>
  dplyr::transmute(
    Scenario = .data$scenario_label,
    `Changed axis` = .data$primary_axis_changed,
    `Exact change` = .data$change_from_primary,
    `Held fixed` = .data$invariants,
    `Comparison population` = .data$sample_strategy
  ) |>
  compact_gt()
```

# Exact sample change

The eligibility restriction removed 725 supported hours and 31
participant-days from six participants. This retained 95.6% of supported
hours, 95.7% of participant-days, and 95.6% of participants. All nine sites and
both levels of day type and activity status remained. The smallest retained
site-by-day-type and site-by-activity cells contained 187 and 180 supported
hours, respectively.

```{r}
#| label: tbl-h06-emp-sample-flow
#| tbl-cap: "Exact fitted samples before and after the employment-eligibility restriction."

sample_flow |>
  dplyr::transmute(
    Sample = .data$scenario_label,
    `Supported hours` = .data$one_hour_observations,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Hours retained` = .data$retained_hour_fraction
  ) |>
  compact_gt(
    source_note = paste(
      "The restriction removed four participants recorded as marginally",
      "employed and two recorded as not employed. Participant identifiers",
      "are not included in this report."
    )
  ) |>
  gt::fmt_integer(
    columns = c(
      `Supported hours`, `Participant-days`, Participants, Sites
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_percent(columns = `Hours retained`, decimals = 1)
```

# Models and inference

The same two accepted quasi-Poisson log-mean models were refitted. Study site
uses sum contrasts. Day type and activity use their first levels, work day and
sedentary, as references. The reported ratios give each study site equal
weight. Participant-cluster HC3 covariance estimates use 130 denominator
degrees of freedom in the restricted sample.

```{r}
#| label: tbl-h06-emp-formulas
#| tbl-cap: "Exact evaluated Wilkinson formulas for the employment-eligible near-eye sample."

formula_registry |>
  dplyr::transmute(
    Model = dplyr::recode(
      .data$model_role,
      additive = "Average association",
      full = "Predictor-by-site interaction"
    ),
    Formula = paste0("`", .data$wilkinson_formula, "`"),
    Family = paste(.data$family, "with", .data$link, "link"),
    Inference = .data$covariance
  ) |>
  compact_gt() |>
  gt::fmt_markdown(columns = "Formula")
```

False-discovery-rate adjustment was repeated within the same three-member
families used by the frozen near-eye analysis: the three average associations,
the three practical standardized contrasts, and the three predictor-by-site
omnibus tests. The site-specific summaries below are descriptive estimates
with pointwise 95% CIs. They do not create another testing family.

# Near-eye sensitivity results

The employment restriction produces little change in the free/work ratio. The
active/sedentary ratio is modestly smaller, but remains compatible with the
primary estimate and retains the same direction and adjusted conclusion.
Previous sleep remains unsupported. The project-defined comparison rule
classifies all three results as stable within model uncertainty.

```{r}
#| label: tbl-h06-emp-effect-comparison
#| tbl-cap: "Primary and employment-eligible near-eye associations."

effect_comparison |>
  dplyr::left_join(
    effects |>
      dplyr::select(
        .data$predictor_id,
        sensitivity_adjusted_significant =
          .data$adjusted_significant_0_05
      ),
    by = "predictor_id",
    relationship = "one-to-one"
  ) |>
  dplyr::transmute(
    Predictor = display_predictor(.data$predictor_id),
    Primary = format_ratio(
      .data$reference_ratio,
      .data$reference_conf_low,
      .data$reference_conf_high
    ),
    `Employment eligible` = format_ratio(
      .data$alternative_ratio,
      .data$alternative_conf_low,
      .data$alternative_conf_high
    ),
    `FDR-adjusted p` = format_p(
      .data$alternative_p_adjusted,
      .data$sensitivity_adjusted_significant
    ),
    Classification = tools::toTitleCase(
      .data$stability_classification
    )
  ) |>
  compact_gt(
    source_note = paste(
      "Ratios compare free with work days, active with sedentary days,",
      "or a one-hour increase in previous sleep. Adjusted p-values are",
      "from the employment-eligible near-eye sample."
    )
  ) |>
  gt::fmt_markdown(columns = `FDR-adjusted p`)
```

The day-type predictor-by-site test remained false-discovery-rate supported,
with adjusted *p* = 0.003 compared with 0.017 in the primary sample. The
activity and previous-sleep interaction tests remained unsupported. Thus, the
existing qualification that the day-type association differs among sites is
unchanged.

```{r}
#| label: tbl-h06-emp-heterogeneity-comparison
#| tbl-cap: "Predictor-by-site tests before and after the employment-eligibility restriction."

heterogeneity |>
  dplyr::transmute(
    Predictor = display_predictor(.data$predictor_id),
    `Primary test` = sprintf(
      "F(%d, %d) = %.2f",
      .data$primary_df1,
      .data$primary_df2,
      .data$primary_f_statistic
    ),
    `Primary FDR-adjusted p` = format_p(
      .data$primary_p_adjusted,
      .data$primary_adjusted_supported
    ),
    `Employment-eligible test` = sprintf(
      "F(%d, %d) = %.2f",
      .data$sensitivity_df1,
      .data$sensitivity_df2,
      .data$sensitivity_f_statistic
    ),
    `Employment-eligible FDR-adjusted p` = format_p(
      .data$sensitivity_p_adjusted,
      .data$sensitivity_adjusted_supported
    ),
    `Adjusted conclusion changed` = dplyr::if_else(
      .data$adjusted_conclusion_changed,
      "Yes",
      "No"
    )
  ) |>
  compact_gt(
    source_note = paste(
      "The three predictor-by-site tests form one FDR family within each",
      "sample. Threshold status is interpreted with the site-specific",
      "magnitudes, intervals, sample composition, and diagnostics."
    )
  ) |>
  gt::fmt_markdown(
    columns = c(
      `Primary FDR-adjusted p`,
      `Employment-eligible FDR-adjusted p`
    )
  )
```

<details>
<summary>Show the prespecified site-specific summaries</summary>

All nine day-type site summaries and all nine activity site summaries are
stable within model uncertainty. Eight of nine previous-sleep summaries are
also stable. Kumasi is precision-sensitive: its pointwise sleep interval
changes from 0.73–1.14 in the primary sample to 0.70–0.95 after the employment
restriction. This local interval is descriptive, unadjusted, and does not
change the overall previous-sleep conclusion.

```{r}
#| label: tbl-h06-emp-site-comparison
#| tbl-cap: "Descriptive site-specific ratios before and after the employment-eligibility restriction."

site_comparison |>
  dplyr::left_join(
    site_registry |>
      dplyr::select(.data$site, .data$display_order, .data$display_name),
    by = "site",
    relationship = "many-to-one"
  ) |>
  dplyr::arrange(
    match(.data$predictor_id, names(predictor_labels)),
    .data$display_order
  ) |>
  dplyr::transmute(
    Predictor = display_predictor(.data$predictor_id),
    Site = .data$display_name,
    Primary = format_ratio(
      .data$primary_ratio,
      .data$primary_conf_low,
      .data$primary_conf_high
    ),
    `Employment eligible` = format_ratio(
      .data$sensitivity_ratio,
      .data$sensitivity_conf_low,
      .data$sensitivity_conf_high
    ),
    Classification = dplyr::recode(
      .data$detailed_stability_classification,
      `stable within model uncertainty` = "Stable",
      `precision-sensitive` = "Quantitatively sensitive",
      `magnitude-sensitive` = "Quantitatively sensitive",
      `direction-sensitive` = "Qualitatively sensitive",
      `non-estimable` = "Non-estimable"
    )
  ) |>
  compact_gt(
    groupname_col = "Predictor",
    source_note = paste(
      "Pointwise 95% CIs are descriptive and unadjusted. No site-specific",
      "multiplicity family was added for this sensitivity."
    )
  )
```

</details>

# Model checks and limitations

Both models converged, retained full-rank designs, and produced finite,
positive-definite HC3 covariance matrices. The numerical gate passed for both.
The additive model retained residual mean-variance structure and within-person
temporal correlation similar to the primary model. These known limitations
remain qualifications rather than newly introduced failures.

```{r}
#| label: tbl-h06-emp-diagnostics
#| tbl-cap: "Numerical and residual checks for the employment-eligible models."

diagnostic_summary |>
  dplyr::transmute(
    Model = dplyr::recode(
      .data$model_role,
      additive = "Average association",
      full = "Predictor-by-site interaction"
    ),
    Converged = dplyr::if_else(.data$converged, "Yes", "No"),
    `Full rank` = dplyr::if_else(.data$full_rank, "Yes", "No"),
    `Positive-definite HC3` = dplyr::if_else(
      .data$hc3_covariance_positive_definite,
      "Yes",
      "No"
    ),
    `HC3 condition number` = .data$hc3_covariance_condition_number,
    Dispersion = .data$dispersion,
    `Exact-zero share` = .data$exact_zero_fraction,
    `Calibration range` = sprintf(
      "%.2f–%.2f",
      .data$calibration_low,
      .data$calibration_high
    ),
    `Residual correlation, lag 1` = .data$lag_1,
    `Residual correlation, lag 6` = .data$lag_6,
    `Largest cluster score share` = .data$maximum_cluster_score_share,
    Gate = dplyr::if_else(.data$numerical_gate_pass, "PASS", "FAIL")
  ) |>
  compact_gt(
    source_note = paste(
      "Calibration is descriptive and in-sample. Residual correlations use",
      "the retained true-time sequences."
    )
  ) |>
  gt::fmt_number(
    columns = c(
      `HC3 condition number`, Dispersion,
      `Residual correlation, lag 1`, `Residual correlation, lag 6`,
      `Largest cluster score share`
    ),
    decimals = 3
  ) |>
  gt::fmt_percent(columns = `Exact-zero share`, decimals = 1)
```

This analysis isolates employment eligibility, but it does not reconstruct a
hypothetical sample in which the excluded participants had instead met the
employment criterion. The comparison therefore measures the combined effect
of applying the registered rule and changing participant composition. Because
the model specification itself is identical, a separate common-sample model
would duplicate the restricted fit and would not separate those mechanisms.

# Reproducibility and review gate

The calculation used R 4.6.1. All frozen inputs, accepted comparator outputs,
modeling functions, and protected main H06 report files matched their recorded
SHA-256 identities before fitting. The exclusion audit stores only hashed
participant keys and is not displayed here. No accepted H06 model, result,
figure, report, or website page was changed.

```{r}
#| label: tbl-h06-emp-input-audit
#| tbl-cap: "Frozen inputs and protected H06 identities verified before fitting."

input_audit |>
  dplyr::transmute(
    Role = gsub("_", " ", .data$input_role),
    Path = paste0("`", .data$relative_path, "`"),
    `SHA-256` = paste0(substr(.data$observed_sha256, 1, 12), "…"),
    Verified = dplyr::if_else(.data$hash_verified, "Yes", "No")
  ) |>
  compact_gt() |>
  gt::fmt_markdown(columns = "Path")
```

```{r}
#| label: tbl-h06-emp-software
#| tbl-cap: "R and package versions used for the sensitivity."

package_versions |>
  dplyr::rename(Package = .data$package, Version = .data$version) |>
  compact_gt()
```

Source files for independent review:

- [Exact sample flow](../../../../artifacts/06_model_data/H06/employment_eligibility_sensitivity/H06_employment_eligibility_sample_flow.csv)
- [Common-effect comparison](../../../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_effect_comparison.csv)
- [Predictor-by-site comparison](../../../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_heterogeneity_comparison.csv)
- [Site-specific comparison](../../../../artifacts/09_tables/H06/employment_eligibility_sensitivity/H06_employment_eligibility_site_comparison.csv)
- [Model diagnostics](../../../../artifacts/08_diagnostics/H06/employment_eligibility_sensitivity/H06_employment_eligibility_fit_diagnostics.csv)
- [Non-circular analysis manifest](../../../../artifacts/12_manifests/H06/employment_eligibility_sensitivity/H06_employment_eligibility_analysis_manifest.csv)

This report stops at **H06-S-EMP-NE-REVIEW**. Its results require independent
scientific and author acceptance before any integration into the main H06
findings or preparation reports.
