---
title: "H07 Stage 1: audit and analysis plan"
subtitle: "Absolute latitude, photoperiod, collection period, and possible exposure plateaus"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-overflow: wrap
    df-print: paged
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

explicit_root <- Sys.getenv("NATHEALTH_PROJECT_ROOT", unset = "")
root_candidates <- unique(normalizePath(
  c(explicit_root[nzchar(explicit_root)], ".", "..", "../..", "../../.."),
  winslash = "/",
  mustWork = FALSE
))
root_matches <- root_candidates[file.exists(file.path(
  root_candidates,
  "preregistration",
  "AsPredicted #273407.pdf"
))]
if (length(root_matches) < 1L) {
  stop("Could not locate the project root", call. = FALSE)
}
root <- root_matches[[1L]]

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
if (dir.exists(project_library)) {
  .libPaths(c(project_library, .libPaths()))
}

suppressPackageStartupMessages({
  library(digest)
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(purrr)
  library(readr)
  library(rvest)
  library(scales)
  library(stringr)
  library(tibble)
  library(tidyr)
})

# Keep this bounded render from writing the user's global Sass cache.
options(sass.cache = FALSE)

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H07 Stage 1 must run under R 4.6.1", call. = FALSE)
}

h07_path <- function(...) file.path(root, ...)
h07_sha256 <- function(path) {
  digest::digest(file = path, algo = "sha256", serialize = FALSE)
}
h07_formula_text <- function(x) paste(deparse(x), collapse = " ")
h07_count_cell <- function(participants, participant_days, sites = NULL) {
  if (is.null(sites)) {
    paste0(participants, " / ", participant_days)
  } else {
    paste0(participants, " / ", participant_days, " / ", sites)
  }
}
h07_range_text <- function(x, digits = 2) {
  if (length(x) == 0L || !any(is.finite(x))) return("None")
  paste0(
    formatC(min(x, na.rm = TRUE), format = "f", digits = digits),
    "–",
    formatC(max(x, na.rm = TRUE), format = "f", digits = digits)
  )
}
h07_gt <- function(data, title = NULL, note = NULL, font_size = 11) {
  out <- data |>
    gt() |>
    opt_align_table_header(align = "left") |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(font_size),
      data_row.padding = px(3),
      column_labels.font.weight = "bold",
      source_notes.font.size = px(max(9, font_size - 2))
    )
  if (!is.null(title)) out <- out |> tab_header(title = md(title))
  if (!is.null(note)) out <- out |> tab_source_note(md(note))
  out
}

site_display <- read_csv(
  h07_path("config", "site_display_registry.csv"),
  show_col_types = FALSE
) |>
  arrange(display_order)
site_levels <- site_display$site

metric_ids <- c(
  "daily_geometric_mean_medi",
  "m10_mean_medi",
  "l10_mean_medi",
  "duration_above_1000",
  "duration_above_250_wake",
  "duration_below_10_pre_sleep",
  "duration_below_1_sleep_environment",
  "longest_bout_above_250",
  "dose_time_sensitive_corrected_medi"
)

metric_source_map <- tribble(
  ~metric_order, ~metric_id, ~source_column, ~v0_name,
  1L, "daily_geometric_mean_medi", "daily_geometric_mean_medi_lx", "Mean",
  2L, "m10_mean_medi", "m10_mean_medi_lx", "brightest_10h_mean",
  3L, "l10_mean_medi", "l10_mean_medi_lx", "darkest_10h_mean",
  4L, "duration_above_1000", "duration_above_1000_h", "duration_above_1000",
  5L, "duration_above_250_wake", "duration_above_250_wake_h", "duration_above_250_wake",
  6L, "duration_below_10_pre_sleep", "duration_below_10_pre_sleep_h", "duration_below_10_pre-sleep",
  7L, "duration_below_1_sleep_environment", "duration_below_1_sleep_environment_h", "duration_below_1_sleep",
  8L, "longest_bout_above_250", "longest_bout_above_250_h", "period_above_250",
  9L, "dose_time_sensitive_corrected_medi", "dose_corrected_medi_lx_h", "dose"
)

h05_metric_contract <- read_csv(
  h07_path("artifacts", "06_model_data", "H05", "H05_metric_registry.csv"),
  show_col_types = FALSE
) |>
  filter(metric_id %in% metric_ids)

metric_contract <- metric_source_map |>
  left_join(
    h05_metric_contract |>
      select(
        metric_id,
        response_family,
        response_transform,
        effect_scale,
        lower_bound,
        upper_bound,
        audit_upper_threshold,
        diagnostic_note,
        manuscript_name,
        abbreviation,
        manuscript_category,
        display_unit,
        variant_label,
        value_definition
      ),
    by = "metric_id"
  ) |>
  arrange(metric_order)

stopifnot(
  nrow(metric_contract) == 9L,
  identical(metric_contract$metric_id, metric_ids),
  !anyNA(metric_contract$manuscript_name),
  !anyNA(metric_contract$response_family)
)
```

::: {.callout-warning}
## Historical Stage 1 gate and post-Stage 2 amendment

This document originally reconstructed the signed H07 contract, the submitted
V0 model and GAM-derivative workflow, all stated deviations, candidate inputs,
and the pre-fit support and identifiability problem. Its evaluated cells create
formula objects and descriptive design summaries only; this Stage 1 document
does **not** fit an inferential model.

The owner approved the original gate as `H07-001` and later reopened its
ceiling and support components after reviewing Stage 2. The post-result author
revision is recorded as `H07-002`. Any computation-heavy bootstrap or
simulation still requires a 50- or 100-replicate pilot, displayed runtime
projection, and separate production approval.
:::

# How the revised ceiling question will be assessed

::: {.callout-important title="H07-002: revised derivative-defined plateau assessment"}
The owner has simplified the H07 question to whether a plateau-like
photoperiod association is visible in the pooled observed data. For each of
the nine metrics and each placement separately, the revised analysis:

1. uses the already fitted site- and participant-adjusted photoperiod GAM;
2. evaluates the first derivative of `s(photoperiod_hours)` at 100 equally
   spaced points from the exact metric-specific minimum to maximum recorded
   photoperiod;
3. identifies a pointwise detected increase when the derivative's 95%
   interval is wholly above zero;
4. locates the first subsequent point whose interval contains zero; and
5. classifies a **derivative-defined plateau pattern** only when every later
   point remains compatible with zero through the maximum recorded
   photoperiod.

The primary implementation uses central finite differences, forward/backward
differences at the observed boundaries, a 0.01 h step, and unconditional
covariance where available. The exact V0 forward, conditional, pointwise
settings are a method sensitivity. The transition is reported as the grid
bracket between the final detected-increase point and the first zero-compatible
point.

The pooled-support, one-hour, and leave-one-site-out rules no longer determine
eligibility. They remain diagnostics describing which sites and collection
periods contribute to different curve regions and how site omission changes
the fitted association.

This rule was chosen after inspection of the Stage 2 support results and is
therefore a post-result descriptive definition. Its pointwise intervals have
no curve-wide or cross-metric multiplicity control. An interval containing
zero does not demonstrate practical equivalence, so the classification is not
proof of an asymptote, a mechanistic ceiling, or a causal photoperiod effect.
Site, latitude, photoperiod, and collection period limitations remain explicit
but do not suppress the requested classification.

The revised implementation incorporates H11's computational lesson without
assuming that H11's method transfers to H07. H07 has 655--816 primary near-eye
participant-days and 743--902 complementary chest participant-days per metric,
not H11's large half-hourly Gaussian/AR(1) problem. The proposed primary engine
is therefore exact `mgcv::gam(method = "REML")`. No primary step requires
`bam(discrete = FALSE)`. A checkpointed `mgcv::bam(method = "fREML",
discrete = TRUE)` fit is only a computational fallback after a timed exact-fit
attempt and must be checked against the exact GAM for fitted values, effective
degrees of freedom, conditional AIC, supported predictions, and derivatives.

H11's participant-cluster sandwich was validated for a Gaussian identity-link
model after AR whitening. Its proposed family-aware extension to Tweedie is a
new method, not an inherited result. H07 will use maintained `mgcv` inference
as the primary conditional analysis and place any custom sandwich or
participant-cluster refit bootstrap behind the H07-G7 method-validation pilot
gate. Material residual dependence or failed identifiability prevents a
confirmatory support claim until that gate is resolved.
:::

# Scope, terminology, and analytical boundary

H07 asks whether nine personal light-exposure metrics are associated with
absolute latitude and civil photoperiod, whether the joint association is
identifiable given the sites and collection periods, and whether any
well-supported association can defensibly be called a plateau. The nine
outcomes enter independently of H01 fitted results, p-values, or support
classifications. H01 is used here only to reconstruct V0's historical
significance-dependent selection and to preserve the accepted metric and
reporting terminology.

Near-eye measurement is primary. Chest measurement is complementary. Direct
placement comparisons use the same participants and participant-days at both
placements and retain separate models; placements are never pooled and no
placement-equivalence claim is planned.

The predefined preparation sensitivity is the **gap-timing-unaware dataset**.
It still passed the general 50%-per-hour and 80%-per-day coverage rules. The
name means that the timing of the remaining missing observations was not used
for an additional metric-specific adjustment; it does not mean that gaps,
missingness, or coverage were ignored. For this one-time explanation, the
primary could be interpreted as a time-sensitive primary metric dataset
because its approved preparation uses the timing of remaining missing
observations where relevant. From here onward it is called simply the
**primary dataset**.

Reader-facing language uses **melEDI**, “period” rather than “bout,” and
“bedside sleep-environment exposure” for the sleep metric. No temperature,
weather, health outcome, clinical endpoint, unsupported predictor, causal
effect, or physiological mechanism is introduced.

# Registered H07 and decision provenance

## Exact registered hypothesis

The signed preregistration states:

> **H7: There is a ceiling (nonlinear) effect of absolute latitude and
> photoperiod on level-, duration-, and exposure-history-based light metrics.**

The signed generic model is
`Metric ~ te(Latitude, Photoperiod) + s(Site, bs="re") +
s(Participant, bs="re")`. The supporting developed analysis plan repeats the
same tensor-product GAMM and expressly notes that identifiability may be a
problem when latitude and photoperiod are both present. The registration also
places chest first, uses near-eye as a robustness repeat, requires FDR control
within hypothesis families, and uses a two-AIC-point rule for GAMM model
comparison. The approved placement decision now makes near eye primary and
chest complementary; that reversal is disclosed rather than hidden.

```{r}
#| label: registered-h07-formula
#| echo: true
#| code-fold: false

registered_h07_formula <- stats::as.formula(
  "Metric ~ te(Latitude, Photoperiod) + s(Site, bs = 're') + s(Participant, bs = 're')"
)
registered_h07_formula
```

## Evidence and accepted decisions

```{r}
evidence_inventory <- tribble(
  ~evidence, ~role,
  "preregistration/AsPredicted #273407.pdf", "Signed wording, tensor formula, placement, AIC, diagnostics, AR1, and FDR commitments",
  "preregistration/Analysis_plan.docx", "Developed plan and advance warning about latitude–photoperiod identifiability",
  "_deviations.qmd", "Registered removal of latitude, H01-dependent outcome selection, and revised photoperiod-only wording",
  "audit/hypotheses/H05-H08_migration_map.md", "Exact V0 reconstruction, defect map, proposed H07 repair, and major gates",
  "audit/hypotheses/H03-H11_gated_workflow.qmd", "Four-stage workflow and outcome-blinded Stage 1 boundary",
  "audit/decisions/metric_validity.md", "Admissible metric constructs and limitations",
  "audit/decisions/metric_implementation_parameters.md", "Support, gap, dose, window, and period implementation",
  "audit/decisions/placement_decision.md", "Near-eye primary, chest complementary, paired/common direct comparison",
  "audit/decisions/site_display_conventions.md", "Submitted site names, north-to-south order, and colours",
  "audit/decisions/model_reporting.md", "Exact samples, confidence intervals, diagnostics, and model provenance",
  "audit/decisions/p_value_display.md", "REPORT-008 three-decimal p-value presentation",
  "RQ2.qmd; RQ2_chest.qmd; scripts/RQ2_specific.R", "Executable V0 model selection and GAM-derivative implementation",
  "docs/RQ2.html; docs/RQ2_chest.html", "Rendered V0 retained-model tables and claims",
  "audit/hypotheses/H11/02_inferential_method_amendment.qmd; audit/hypotheses/H11/02_inferential_method_decision.md; scripts/hypotheses/H11/h11_stage2.R", "Accepted H11 replacement of unfinished ML/discrete-false fits by post-fit participant-cluster-robust whole-curve inference from one fREML/discrete model; methodological precedent only",
  "manuscript/reviewer_comments_NatMed.rtf; audit/ledgers/reviewer_feedback_disposition.csv", "Collection period, sparse site, placement, sample, multiplicity, and causal-language constraints",
  "audit/hypotheses/H01/; audit/handoffs/H01_worker_handoff.md", "Historical V0 selection dependency and accepted terminology only; no H01 fitted result selects H07 outcomes"
)

evidence_inventory |>
  h07_gt(
    title = "Evidence and decisions controlling H07 Stage 1",
    note = "Scientific design calculations below use R 4.6.1. No H01 estimate or p-value enters outcome selection.",
    font_size = 10
  ) |>
  cols_width(evidence ~ px(360), role ~ px(700))
```

## Registered and reconstructed deviations

```{r}
deviation_map <- tribble(
  ~component, ~registered_or_accepted, ~v0_or_legacy_change, ~Stage2_disposition,
  "Outcome set", "All nine level-, duration-, and exposure-history metrics", "H01 photoperiod support selected eight near-eye and six chest candidates; H07 AIC then selected six and four", "Fit all nine independently of H01; retain a reason-coded non-estimable row rather than shorten the family",
  "Predictor surface", "Tensor product of absolute latitude and photoperiod", "Latitude was removed because the observed two-dimensional space was sparse", "Attempt the registered tensor on exact metric rows; failure is non-identifiability, not permission for a ceiling claim",
  "Placement", "Chest main; near eye robustness", "Separate near-eye and chest code paths", "Approved near-eye primary, chest complementary, paired/common direct comparison",
  "Response model", "Metric-specific distribution not fully enumerated", "Every metric used log10(value + 0.1) Gaussian GAM", "Use the accepted metric-specific response family, transform, link, and common-family failure rule",
  "Repeated days", "Participant random effect", "Participant random smooth, with no interpreted date-lag diagnostic", "Explicit participant nested in site; interpret date-lag residual dependence",
  "Model comparison", "Two-AIC-point GAMM rule and FDR within H07", "Conditional delta-AIC was called significance; no nine-value p-vector", "Retain audited conditional AIC as secondary relative-support evidence; use maintained approximate GAM tests in complete nine-value BH families; never call an AIC threshold significance",
  "Plateau", "Ceiling/nonlinear wording without an operational margin", "Last pointwise-positive derivative boundary was treated as a ceiling", "H07-002 applies a post-result positive-to-zero-compatible-through-the-end derivative rule; report only a descriptive derivative-defined pattern",
  "Figures", "Not specified", "Daily mean duplicated and M10 omitted in both Fig14 assemblies", "Address metrics by ID; show every intended panel once and persist exact source rows"
)

deviation_map |>
  h07_gt(
    title = "H07 deviation and repair map",
    font_size = 9
  ) |>
  cols_width(
    component ~ px(135),
    registered_or_accepted ~ px(310),
    v0_or_legacy_change ~ px(330),
    Stage2_disposition ~ px(390)
  )
```

# Exact V0 reconstruction

## V0 formula objects and fitting rule

V0 filtered a global H01 result object by `H1_phot_sig` and metric category,
then supplied the following exact formulas to `mgcv::bam()`. Both fits used
`discrete = TRUE`, `method = "fREML"`, and ten threads. Every response was
`log10(metric + 0.1)`, including durations with a point mass at zero.

```{r}
#| label: v0-h07-formulas
#| echo: true
#| code-fold: false

v0_h07_formulas <- list(
  H7_full = stats::as.formula(
    "log_zero_inflated(metric) ~ s(photoperiod) + s(site, bs = 're') + s(Id, bs = 're')"
  ),
  H7_null = stats::as.formula(
    "log_zero_inflated(metric) ~ s(site, bs = 're') + s(Id, bs = 're')"
  )
)

enframe(v0_h07_formulas, name = "formula_id", value = "formula") |>
  mutate(formula = map_chr(formula, h07_formula_text))
```

V0 calculated `dAIC = AIC(full) - AIC(null)` and retained a model when
`dAIC < -2`. This is a relative-support convention, not a significance test.
For an `mgcv::gam` or `mgcv::bam` object, `AIC()` dispatches through
`logLik.gam()` and uses the model's conditional likelihood with effective
degrees of freedom; it does **not** compare the stored fREML smoothing
criterion. V0's values are therefore potentially interpretable as conditional
AIC evidence if the compared objects used identical response rows, family and
link, weights, offset, likelihood implementation, and fitting engine, and if
the smoothing-uncertainty-corrected `edf2` provenance can be verified. Those
historical object-level checks are not currently documented, so V0 conditional
AIC is classified as **provenance unverified**, not inherently invalid. Its
outcome-selection dependence and “significant” label remain invalid.

## Historical two-stage outcome selection and samples

```{r}
v0_candidate_registry <- tribble(
  ~placement, ~metric_id,
  "Near eye", "daily_geometric_mean_medi",
  "Near eye", "m10_mean_medi",
  "Near eye", "l10_mean_medi",
  "Near eye", "duration_above_1000",
  "Near eye", "longest_bout_above_250",
  "Near eye", "dose_time_sensitive_corrected_medi",
  "Near eye", "duration_below_10_pre_sleep",
  "Near eye", "duration_above_250_wake",
  "Chest", "daily_geometric_mean_medi",
  "Chest", "m10_mean_medi",
  "Chest", "duration_above_1000",
  "Chest", "longest_bout_above_250",
  "Chest", "dose_time_sensitive_corrected_medi",
  "Chest", "duration_above_250_wake"
)

read_v0_metric_list <- function(path, placement_label) {
  environment <- new.env(parent = emptyenv())
  object_names <- load(path, envir = environment)
  metric_list <- environment[[object_names[[1L]]]]

  metric_list |>
    inner_join(
      metric_source_map |> select(metric_id, metric_order, v0_name),
      by = c("name" = "v0_name")
    ) |>
    inner_join(
      v0_candidate_registry |> filter(placement == placement_label),
      by = "metric_id"
    ) |>
    transmute(
      placement = placement_label,
      metric_order,
      metric_id,
      data
    )
}

v0_frames <- bind_rows(
  read_v0_metric_list(h07_path("data", "metrics_glasses.RData"), "Near eye"),
  read_v0_metric_list(h07_path("data", "metrics_chest.RData"), "Chest")
) |>
  mutate(
    sample = map(data, function(frame) {
      frame |>
        filter(
          is.finite(metric),
          is.finite(photoperiod),
          !is.na(site),
          !is.na(Id)
        ) |>
        summarise(
          participants = n_distinct(site, Id),
          participant_days = n(),
          observations = n(),
          sites = n_distinct(site)
        )
    })
  ) |>
  select(-data) |>
  unnest(sample)

extract_v0_h07_aic <- function(path, placement_label) {
  document <- read_html(path)
  tables <- html_elements(document, "table")
  parsed <- map(tables, ~ html_table(.x, fill = TRUE))
  table_index <- which(map_lgl(parsed, ~ "dAIC" %in% names(.x)))
  if (length(table_index) != 1L) {
    stop("Could not uniquely identify the V0 H07 AIC table", call. = FALSE)
  }
  parsed[[table_index]] |>
    transmute(
      placement = placement_label,
      v0_name = name,
      dAIC = as.numeric(str_replace_all(dAIC, "−", "-"))
    ) |>
    left_join(
      metric_source_map |> select(metric_id, v0_name),
      by = "v0_name"
    )
}

v0_retained <- bind_rows(
  extract_v0_h07_aic(h07_path("docs", "RQ2.html"), "Near eye"),
  extract_v0_h07_aic(h07_path("docs", "RQ2_chest.html"), "Chest")
)

v0_selection <- v0_frames |>
  left_join(v0_retained, by = c("placement", "metric_id")) |>
  left_join(
    metric_contract |> select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  mutate(
    sample = h07_count_cell(participants, participant_days, sites),
    retained = if_else(is.finite(dAIC), "Yes", "No"),
    dAIC_display = if_else(is.finite(dAIC), sprintf("%.2f", dAIC), "—")
  ) |>
  select(metric_order, Metric = manuscript_name, placement, sample, retained, dAIC_display) |>
  pivot_wider(
    names_from = placement,
    values_from = c(sample, retained, dAIC_display),
    names_glue = "{placement} — {.value}"
  ) |>
  arrange(metric_order)

v0_selection |>
  select(-metric_order) |>
  h07_gt(
    title = "Historical V0 H01-selected candidates and second-stage AIC retention",
    note = "Sample cells are participants / participant-days / sites. A dash means that H01 excluded the metric before H07 or H07 did not retain it. These historical fitted results are reconstructed only for comparison; they do not select the repaired H07 outcomes.",
    font_size = 9
  ) |>
  cols_width(
    Metric ~ px(250),
    contains("sample") ~ px(145),
    contains("retained") ~ px(95),
    contains("dAIC") ~ px(95)
  )
```

The near-eye screen began with eight metrics and retained six. The chest
screen began with six and retained four. Time below 1 lx melEDI during sleep
never entered either V0 candidate set. These nested result-dependent screens
cannot define the repaired outcome family.

## Exact V0 GAM-derivative method

The user correctly identified the derivative workflow as the substantive V0
method. The helper did not merely draw a smooth: it converted the final grid
location with evidence of a positive derivative into the reported ceiling
hour.

```{r}
#| label: v0-h07-derivative-rule
#| echo: true
#| code-fold: false

v0_derivative_call <- quote(
  deriv <- gratia::derivatives(model, select = "s(photoperiod)")
)
v0_boundary_rule <- quote(
  deriv_signif <- deriv |>
    dplyr::filter(.lower_ci > 0) |>
    dplyr::slice_tail(n = 1)
)

v0_derivative_call
v0_boundary_rule
```

With the pinned V0 `gratia` 0.11.2 defaults, this means a 100-point grid,
first-order forward finite differences with `eps = 1e-7`, 95% **pointwise**
confidence intervals, conditional smoothing-parameter uncertainty, and no
simultaneous curve control. `slice_tail()` selects the final positive-
derivative point even if positive regions are separated rather than sustained.

```{r}
v0_derivative_defaults <- tribble(
  ~component, ~v0_setting, ~consequence,
  "Derivative target", "s(photoperiod) only", "No latitude derivative or joint-surface plateau was evaluated",
  "Grid", "n = 100", "The labelled hour depends on a software-generated grid",
  "Finite difference", "Forward; eps = 1e-7", "No stability check across difference rule or step size",
  "Uncertainty", "95% pointwise confidence interval", "Repeated grid-point inspection is uncontrolled",
  "Boundary", "Last row with lower CI > 0", "Marks the end of detected increase, not equivalence to zero",
  "Sustained interval", "None", "A later increase, decrease, reversal, or edge loss can be mislabelled as flattening",
  "Support guard", "None", "The boundary can occur where only sparse sites or collection periods remain",
  "Influence", "None", "No leave-one-site-out stability requirement"
)

v0_derivative_defaults |>
  h07_gt(
    title = "Interpretation of the V0 derivative algorithm",
    note = "A non-significant derivative, an AIC threshold, and a visually flat smooth are each insufficient evidence of a ceiling.",
    font_size = 10
  ) |>
  cols_width(component ~ px(170), v0_setting ~ px(245), consequence ~ px(560))
```

V0 then described the last positive-derivative locations as ceilings around
14.7–16.4 h for level metrics and 16.6 h for time above 1,000 lx melEDI. The
same near-eye and chest figure assembly duplicated daily mean and omitted M10.
Every V0 threshold-hour and cross-placement ceiling claim is reopened.

# Direct inputs, metric contract, and candidate samples

## Input identities

```{r}
base_manifest <- read_csv(
  h07_path("artifacts", "12_manifests", "base_model_data_artifacts.csv"),
  show_col_types = FALSE
)
gap_manifest <- read_csv(
  h07_path("artifacts", "12_manifests", "manuscript_prepared_data_artifacts.csv"),
  show_col_types = FALSE
)
solar_manifest <- read_csv(
  h07_path("artifacts", "12_manifests", "site_solar_context_artifacts.csv"),
  show_col_types = FALSE
)

direct_inputs <- bind_rows(
  base_manifest |>
    filter(artifact_id %in% c(
      "glasses_participant_day_enriched",
      "chest_participant_day_enriched"
    )) |>
    transmute(
      input = if_else(
        artifact_id == "glasses_participant_day_enriched",
        "Primary near-eye participant-day metrics",
        "Complementary chest participant-day metrics"
      ),
      path,
      expected_sha256 = sha256,
      producer,
      producer_r = r_version
    ),
  gap_manifest |>
    filter(artifact_id == "participant_day_metrics_rds") |>
    transmute(
      input = "Gap-timing-unaware participant-day metrics",
      path,
      expected_sha256 = sha256,
      producer,
      producer_r = "4.6.1"
    ),
  solar_manifest |>
    filter(artifact_type == "site_solar_context_rds") |>
    transmute(
      input = "Site/date civil-photoperiod context",
      path,
      expected_sha256 = sha256,
      producer,
      producer_r = r_version
    )
) |>
  mutate(
    absolute_path = file.path(root, path),
    exists = file.exists(absolute_path)
  )

stopifnot(all(direct_inputs$exists))

direct_inputs <- direct_inputs |>
  mutate(
    observed_sha256 = map_chr(absolute_path, h07_sha256),
    status = if_else(observed_sha256 == expected_sha256, "PASS", "FAIL")
  )

stopifnot(all(direct_inputs$status == "PASS"))

direct_inputs |>
  transmute(
    Input = input,
    Path = path,
    `SHA-256 prefix` = substr(observed_sha256, 1L, 16L),
    Producer = producer,
    R = producer_r,
    Status = status
  ) |>
  h07_gt(
    title = "Hash-verified direct H07 Stage 1 inputs",
    note = "The Stage 1 render reads these files but does not write a shared model-data artifact.",
    font_size = 9
  ) |>
  cols_width(
    Input ~ px(235),
    Path ~ px(390),
    `SHA-256 prefix` ~ px(135),
    Producer ~ px(300),
    R ~ px(65),
    Status ~ px(65)
  )
```

## Nine-metric response contract

All nine planned outcomes remain in scope even when a metric-specific model
is non-estimable. Whole-day metrics describe the accepted hybrid measurement
record, not uninterrupted 24-hour ocular exposure. The longest continuous
period is an observed lower bound unless its boundaries are exactly
identified. Corrected dose is available only when its relevance-support guard
passes.

```{r}
metric_contract_display <- metric_contract |>
  transmute(
    Order = metric_order,
    Metric = manuscript_name,
    Category = str_to_sentence(str_replace_all(manuscript_category, "-", " ")),
    Unit = display_unit,
    `Prepared construct` = variant_label,
    `Response model` = case_when(
      response_family == "gaussian" & response_transform == "log10_offset_0.1" ~
        "Gaussian; log10(value + 0.1)",
      response_family == "gaussian" & response_transform == "identity" ~
        "Gaussian; identity",
      response_family == "tweedie_log" ~
        "mgcv::tw; log link; power estimated in [1.01, 1.99]",
      TRUE ~ paste(response_family, response_transform)
    ),
    `Practical scale` = if_else(effect_scale == "ratio", "Ratio", "Difference"),
    `Required diagnostic/sensitivity` = diagnostic_note
  )

metric_contract_display |>
  h07_gt(
    title = "Proposed H07 outcome and response-family contract",
    note = "The accepted H01/H05 family package is inherited for cross-hypothesis continuity, not selected from H07 results. A family or link failure opens one common predeclared replacement gate; it never triggers a result-specific switch.",
    font_size = 9
  ) |>
  cols_width(
    Order ~ px(45),
    Metric ~ px(220),
    Category ~ px(120),
    Unit ~ px(60),
    `Prepared construct` ~ px(210),
    `Response model` ~ px(175),
    `Practical scale` ~ px(90),
    `Required diagnostic/sensitivity` ~ px(350)
  )
```

## Candidate-row construction

```{r}
main_near_eye <- readRDS(h07_path(
  "artifacts", "06_model_data", "base",
  "metrics_glasses_participant_day_enriched.rds"
))
main_chest <- readRDS(h07_path(
  "artifacts", "06_model_data", "base",
  "metrics_chest_participant_day_enriched.rds"
))
solar_context <- readRDS(h07_path(
  "artifacts", "06_model_data", "context", "site_solar_context.rds"
))

read_main_long <- function(data, placement_label) {
  data |>
    select(
      site,
      Id,
      local_date,
      latitude_deg,
      photoperiod_hours,
      all_of(metric_source_map$source_column)
    ) |>
    pivot_longer(
      cols = all_of(metric_source_map$source_column),
      names_to = "source_column",
      values_to = "value"
    ) |>
    left_join(metric_source_map, by = "source_column") |>
    transmute(
      data_scenario = "Primary dataset",
      placement = placement_label,
      site,
      Id,
      local_date = as.Date(local_date),
      abs_latitude_deg = abs(latitude_deg),
      photoperiod_hours,
      metric_order,
      metric_id,
      value
    )
}

main_long <- bind_rows(
  read_main_long(main_near_eye, "Near eye"),
  read_main_long(main_chest, "Chest")
)

gap_long <- readRDS(h07_path(
  "artifacts", "06_model_data", "scenarios", "manuscript_prepared_data",
  "participant_day_metrics.rds"
)) |>
  filter(metric_id %in% metric_ids) |>
  left_join(
    solar_context |>
      select(site, local_date, latitude_deg, photoperiod_hours),
    by = c("site", "local_date"),
    relationship = "many-to-one"
  ) |>
  left_join(
    metric_source_map |> select(metric_id, metric_order),
    by = "metric_id"
  ) |>
  transmute(
    data_scenario = "Gap-timing-unaware dataset",
    placement = if_else(position == "glasses", "Near eye", "Chest"),
    site,
    Id,
    local_date = as.Date(local_date),
    abs_latitude_deg = abs(latitude_deg),
    photoperiod_hours,
    metric_order,
    metric_id,
    value = manuscript_prepared_value
  )

candidate_long <- bind_rows(main_long, gap_long)

stopifnot(
  !anyDuplicated(candidate_long[c(
    "data_scenario", "placement", "site", "Id", "local_date", "metric_id"
  )]),
  sum(is.na(candidate_long$abs_latitude_deg)) == 0L,
  sum(is.na(candidate_long$photoperiod_hours)) == 0L
)

summarise_candidate <- function(data) {
  data |>
    filter(!is.na(value)) |>
    group_by(data_scenario, placement, metric_id, metric_order) |>
    summarise(
      participants = n_distinct(site, Id),
      participant_days = n(),
      observations = n(),
      sites = n_distinct(site),
      .groups = "drop"
    )
}

all_available_candidates <- summarise_candidate(candidate_long)

paired_wide <- candidate_long |>
  filter(!is.na(value)) |>
  select(
    data_scenario,
    site,
    Id,
    local_date,
    metric_id,
    metric_order,
    placement,
    value
  ) |>
  pivot_wider(names_from = placement, values_from = value) |>
  filter(!is.na(`Near eye`), !is.na(Chest))

paired_candidates <- paired_wide |>
  group_by(data_scenario, metric_id, metric_order) |>
  summarise(
    participants = n_distinct(site, Id),
    participant_days = n(),
    observations = n(),
    sites = n_distinct(site),
    .groups = "drop"
  )
```

These are candidate complete-case frames before response transformation,
design-rank checks, and model diagnostics. Stage 2 must persist the exact key
of every fitted full/reduced pair; prepared counts cannot substitute for exact
fitted samples.

## Placement and preparation samples

```{r}
candidate_sample_table <- function(scenario_label) {
  all_available_candidates |>
    filter(data_scenario == scenario_label) |>
    left_join(
      metric_contract |> select(metric_id, manuscript_name),
      by = "metric_id"
    ) |>
    mutate(cell = h07_count_cell(participants, participant_days, sites)) |>
    select(metric_order, Metric = manuscript_name, placement, cell) |>
    pivot_wider(names_from = placement, values_from = cell) |>
    arrange(metric_order) |>
    select(Metric, `Near eye`, Chest)
}

candidate_sample_table("Primary dataset") |>
  h07_gt(
    title = "Primary-dataset all-available candidate samples",
    note = "Cells are participants / participant-days / sites. Near eye is primary; chest is complementary and not an unmatched direct placement comparison.",
    font_size = 10
  ) |>
  cols_width(Metric ~ px(310), `Near eye` ~ px(175), Chest ~ px(175))
```

```{r}
candidate_sample_table("Gap-timing-unaware dataset") |>
  h07_gt(
    title = "Gap-timing-unaware all-available candidate samples",
    note = "Cells are participants / participant-days / sites. The same approved H07 implementation is applied; no result-dependent family or formula switch is allowed.",
    font_size = 10
  ) |>
  cols_width(Metric ~ px(310), `Near eye` ~ px(175), Chest ~ px(175))
```

```{r}
paired_candidates |>
  left_join(
    metric_contract |> select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  mutate(cell = h07_count_cell(participants, participant_days, sites)) |>
  select(metric_order, Metric = manuscript_name, data_scenario, cell) |>
  pivot_wider(names_from = data_scenario, values_from = cell) |>
  arrange(metric_order) |>
  select(Metric, `Primary dataset`, `Gap-timing-unaware dataset`) |>
  h07_gt(
    title = "Paired/common candidates for separate near-eye and chest fits",
    note = "Cells are paired participants / paired participant-days / sites. Both placement models use these exact keys for direct effect comparison; MPI has no chest data.",
    font_size = 10
  ) |>
  cols_width(
    Metric ~ px(310),
    `Primary dataset` ~ px(190),
    `Gap-timing-unaware dataset` ~ px(220)
  )
```

## Metric-specific site support

```{r}
candidate_site_matrix <- function(placement_label) {
  candidate_long |>
    filter(
      data_scenario == "Primary dataset",
      placement == placement_label,
      !is.na(value)
    ) |>
    group_by(metric_id, metric_order, site) |>
    summarise(
      participants = n_distinct(Id),
      participant_days = n(),
      .groups = "drop"
    ) |>
    left_join(
      metric_contract |> select(metric_id, manuscript_name),
      by = "metric_id"
    ) |>
    left_join(site_display, by = "site") |>
    mutate(cell = h07_count_cell(participants, participant_days)) |>
    select(
      metric_order,
      Metric = manuscript_name,
      display_order,
      display_name,
      cell
    ) |>
    arrange(metric_order, display_order) |>
    select(-display_order) |>
    pivot_wider(names_from = display_name, values_from = cell) |>
    arrange(metric_order) |>
    select(-metric_order)
}
```

::: {.panel-tabset}

### Near eye — primary

```{r}
candidate_site_matrix("Near eye") |>
  h07_gt(
    title = "Metric-specific primary near-eye candidate support by submitted site",
    note = "Cells are participants / participant-days. UCR has only six near-eye participants overall; sleep-, wake-, and dose-specific frames are smaller for several metrics.",
    font_size = 8
  ) |>
  cols_width(Metric ~ px(225), everything() ~ px(105))
```

### Chest — complementary

```{r}
candidate_site_matrix("Chest") |>
  h07_gt(
    title = "Metric-specific complementary chest candidate support by submitted site",
    note = "Cells are participants / participant-days. MPI is absent because no chest dataset exists for that site.",
    font_size = 8
  ) |>
  cols_width(Metric ~ px(225), everything() ~ px(112))
```

:::

# Predictor identifiability and collection-period support

## Site, latitude, dates, and photoperiod

```{r}
design_rows <- bind_rows(
  main_near_eye |>
    transmute(
      placement = "Near eye",
      site,
      Id,
      local_date = as.Date(local_date),
      abs_latitude_deg = abs(latitude_deg),
      photoperiod_hours
    ),
  main_chest |>
    transmute(
      placement = "Chest",
      site,
      Id,
      local_date = as.Date(local_date),
      abs_latitude_deg = abs(latitude_deg),
      photoperiod_hours
    )
)

collection_support <- design_rows |>
  group_by(placement, site) |>
  summarise(
    participants = n_distinct(Id),
    participant_days = n(),
    observed_dates = n_distinct(local_date),
    first_date = min(local_date),
    last_date = max(local_date),
    abs_latitude_deg = first(abs_latitude_deg),
    photoperiod_min = min(photoperiod_hours),
    photoperiod_max = max(photoperiod_hours),
    photoperiod_span = photoperiod_max - photoperiod_min,
    .groups = "drop"
  ) |>
  left_join(site_display, by = "site") |>
  arrange(
    factor(placement, levels = c("Near eye", "Chest")),
    display_order
  )

collection_support |>
  transmute(
    Placement = placement,
    Site = display_name,
    `Absolute latitude` = sprintf("%.2f°", abs_latitude_deg),
    Participants = participants,
    `Participant-days` = participant_days,
    `Observed dates` = observed_dates,
    `Collection range` = paste(first_date, last_date, sep = "–"),
    `Photoperiod range` = sprintf(
      "%.2f–%.2f h (span %.2f)",
      photoperiod_min,
      photoperiod_max,
      photoperiod_span
    )
  ) |>
  h07_gt(
    title = "Collection periods and predictor support by placement and submitted site",
    note = "The first-to-last range does not imply uninterrupted sampling. Observed-date counts make sparse or separated collection periods visible.",
    font_size = 9
  ) |>
  tab_row_group(label = "Near eye — primary", rows = Placement == "Near eye") |>
  tab_row_group(label = "Chest — complementary", rows = Placement == "Chest") |>
  cols_hide(Placement) |>
  cols_width(
    Site ~ px(185),
    `Absolute latitude` ~ px(105),
    Participants ~ px(80),
    `Participant-days` ~ px(100),
    `Observed dates` ~ px(95),
    `Collection range` ~ px(210),
    `Photoperiod range` ~ px(225)
  )
```

Absolute latitude is one constant per site. Photoperiod is deterministically
derived from site coordinates and local date. Thus site, latitude, season,
and collection period are not independent design factors. Photoperiod varies
within some sites, but UCR and KNUST cover almost no within-site range in the
near-eye data, whereas northern sites cover much wider but different calendar
periods.

```{r}
#| label: fig-h07-joint-support
#| fig-cap: "Observed civil-photoperiod support by submitted site and placement. Site labels include absolute latitude; point size represents participant-days on each observed local date."
#| fig-alt: "Two panels show photoperiod on the horizontal axis and the nine study sites in north-to-south order on the vertical axis. Northern European sites span broad but different photoperiod ranges, the German sites occupy overlapping latitude bands with distinct collection periods, and the Costa Rican and Ghanaian sites have very narrow ranges. The chest panel omits Tübingen because MPI has no chest data."
#| fig-width: 10.5
#| fig-height: 6.4

site_label_map <- collection_support |>
  distinct(site, display_name, abs_latitude_deg) |>
  mutate(label = paste0(display_name, " — ", sprintf("%.1f°", abs_latitude_deg))) |>
  select(site, label) |>
  deframe()

date_support <- design_rows |>
  group_by(placement, site, local_date, abs_latitude_deg, photoperiod_hours) |>
  summarise(
    participant_days = n(),
    participants = n_distinct(Id),
    .groups = "drop"
  ) |>
  mutate(
    site = factor(site, levels = rev(site_levels)),
    placement = factor(placement, levels = c("Near eye", "Chest"))
  )

ggplot(
  date_support,
  aes(
    x = photoperiod_hours,
    y = site,
    colour = site,
    size = participant_days
  )
) +
  geom_point(alpha = 0.58) +
  facet_wrap(~ placement, ncol = 2) +
  scale_colour_manual(
    values = setNames(site_display$color_hex, site_display$site),
    guide = "none",
    drop = FALSE
  ) +
  scale_y_discrete(labels = site_label_map, drop = FALSE) +
  scale_size_area(max_size = 5.2, breaks = c(1, 5, 10, 20)) +
  scale_x_continuous(
    breaks = seq(10, 20, by = 2),
    limits = c(9.8, 20.8)
  ) +
  labs(
    x = "Civil photoperiod (hours)",
    y = NULL,
    size = "Participant-days\non date"
  ) +
  theme_minimal(base_size = 11.5) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.spacing.x = grid::unit(1.4, "lines"),
    axis.text.y = element_text(size = 9),
    strip.text = element_text(face = "bold")
  )
```

## Structural rank and overlap audit

```{r}
design_rank_one <- function(data, placement_label) {
  frame <- data |>
    filter(placement == placement_label) |>
    mutate(site = factor(site, levels = site_levels[site_levels %in% site]))

  formulas <- list(
    `Site + photoperiod` = ~ site + photoperiod_hours,
    `Site + latitude + photoperiod` = ~ site + abs_latitude_deg + photoperiod_hours,
    `Latitude × photoperiod` = ~ abs_latitude_deg * photoperiod_hours
  )

  imap_dfr(formulas, function(formula, label) {
    matrix <- model.matrix(formula, data = frame)
    tibble(
      Placement = placement_label,
      Design = label,
      Columns = ncol(matrix),
      Rank = qr(matrix)$rank,
      Deficiency = ncol(matrix) - qr(matrix)$rank
    )
  })
}

design_rank <- bind_rows(
  design_rank_one(design_rows, "Near eye"),
  design_rank_one(design_rows, "Chest")
)

design_rank |>
  h07_gt(
    title = "Outcome-blinded linear design-rank audit",
    note = "The site-plus-latitude design is rank deficient because latitude is exactly determined by site. A penalized random-site tensor may return a fit, but software convergence alone cannot separate site from latitude.",
    font_size = 10
  ) |>
  cols_width(
    Placement ~ px(110),
    Design ~ px(290),
    Columns ~ px(90),
    Rank ~ px(90),
    Deficiency ~ px(100)
  )
```

```{r}
predictor_ranges <- design_rows |>
  group_by(placement, site) |>
  summarise(
    lower = min(photoperiod_hours),
    upper = max(photoperiod_hours),
    .groups = "drop"
  )

overlap_summary <- function(data, placement_label) {
  ranges <- data |>
    filter(placement == placement_label) |>
    arrange(match(site, site_levels))
  pairs <- combn(seq_len(nrow(ranges)), 2L)
  overlap <- apply(pairs, 2L, function(index) {
    min(ranges$upper[index]) - max(ranges$lower[index])
  })
  tibble(
    Placement = placement_label,
    Sites = nrow(ranges),
    `All-site common lower` = max(ranges$lower),
    `All-site common upper` = min(ranges$upper),
    `All-site common width` = min(ranges$upper) - max(ranges$lower),
    `Overlapping site pairs` = sum(overlap >= 0),
    `All site pairs` = length(overlap),
    `Median positive overlap` = median(overlap[overlap >= 0])
  )
}

bind_rows(
  overlap_summary(predictor_ranges, "Near eye"),
  overlap_summary(predictor_ranges, "Chest")
) |>
  mutate(
    across(
      c(
        `All-site common lower`,
        `All-site common upper`,
        `All-site common width`,
        `Median positive overlap`
      ),
      ~ sprintf("%.2f h", .x)
    ),
    `Pairwise overlap` = paste0(
      `Overlapping site pairs`, " / ", `All site pairs`
    )
  ) |>
  select(
    Placement,
    Sites,
    `All-site common lower`,
    `All-site common upper`,
    `All-site common width`,
    `Pairwise overlap`,
    `Median positive overlap`
  ) |>
  h07_gt(
    title = "Photoperiod overlap across sites",
    note = "A negative all-site width means that no photoperiod interval is observed at every site. Pairwise overlap does not create local replication of latitude because each latitude still belongs to one site.",
    font_size = 10
  ) |>
  cols_width(Placement ~ px(110), everything() ~ px(155))
```

The registered tensor therefore combines nine latitude stripes rather than a
dense two-dimensional design. A fitted tensor can borrow smoothness across
empty regions, but those interpolations are not observations. The high-
photoperiod edge is especially vulnerable to northern-site and collection-
period leverage.

## Metric-specific upper-tail support

```{r}
upper_tail_support <- candidate_long |>
  filter(
    data_scenario == "Primary dataset",
    placement == "Near eye",
    !is.na(value)
  ) |>
  group_by(metric_id, metric_order) |>
  group_modify(function(data, key) {
    threshold <- quantile(data$photoperiod_hours, probs = 0.90, names = FALSE)
    upper <- data |> filter(photoperiod_hours >= threshold)
    tibble(
      threshold = threshold,
      participants = n_distinct(upper$site, upper$Id),
      participant_days = nrow(upper),
      sites = n_distinct(upper$site),
      site_list = paste(
        site_display$display_name[
          site_display$site %in% unique(upper$site)
        ],
        collapse = "; "
      )
    )
  }) |>
  ungroup() |>
  left_join(
    metric_contract |> select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  arrange(metric_order)

upper_tail_support |>
  transmute(
    Metric = manuscript_name,
    `Upper-decile threshold` = sprintf("%.2f h", threshold),
    Participants = participants,
    `Participant-days` = participant_days,
    Sites = sites,
    `Sites represented` = site_list
  ) |>
  h07_gt(
    title = "Primary near-eye support at each metric's upper photoperiod decile",
    note = "The upper decile is a descriptive edge-support audit, not a plateau threshold. Participant-days repeat within participants and do not substitute for independent sites.",
    font_size = 9
  ) |>
  cols_width(
    Metric ~ px(250),
    `Upper-decile threshold` ~ px(140),
    Participants ~ px(85),
    `Participant-days` ~ px(105),
    Sites ~ px(65),
    `Sites represented` ~ px(390)
  )
```

# Revised derivative estimand and the H07-002 decision

## Association estimand

For metric (m), let μₘ(ℓ, p) be the population response-scale mean from
the registered tensor at absolute latitude ℓ and photoperiod (p), with site
and participant random effects set to zero. The primary registered association
estimand is the joint contribution of the latitude–photoperiod tensor over
the observed metric-specific support, conditional on the registered random
effects. It is evaluated by a complete nine-metric model-evidence package:

1. the maintained `summary.gam()` approximate whole-term test of the complete
   latitude–photoperiod tensor, assembled into the
   `H07-F1-registered-association` BH family;
2. a prespecified approximate nested `anova.gam()` comparison of the tensor
   against its bilinear latitude-by-photoperiod subspace, assembled into the
   `H07-F3-registered-nonlinear-shape` BH family;
3. conditional AIC differences for the full-versus-null and
   full-versus-bilinear candidate pairs as secondary relative-support
   evidence, never as p-values; and
4. response-scale predicted surfaces and directional partial derivatives with
   95% uncertainty only where the design-support rule permits them.

The tests are conditional, approximate GAM inference and not
participant-cluster-robust by construction. They are released only if the
registered tensor is identifiable and residual dependence is not materially
inconsistent with the fitted participant random effect. The nested shape test
also requires identical response rows, weights, offsets, family/link, and
fitting method; a genuinely nested coefficient space; and greater effective
degrees of freedom in the tensor model. Failure of any condition makes the
affected p-value non-estimable rather than prompting a favourable alternative.

The exact REML tensor, null, and bilinear models are fitted on one locked
metric-specific frame. `AIC()` is audited as `mgcv` conditional AIC, including
whether `edf2` is available, and the historical two-point convention is
reported only as a descriptive relative-support label. The association test
asks whether the tensor term is zero; the shape comparison asks whether the
supported linear-predictor surface requires more than a bilinear form. Because
every approved link is monotone, a constant linear predictor is also a
constant response mean, but a bilinear linear predictor need not be bilinear
after back-transformation. Practical estimates and derivatives are therefore
reported on the response scale. None of these procedures establishes a
plateau.

If the tensor is non-identifiable, a separately approved photoperiod-only
smooth can estimate a narrower, adapted association conditional on site. It
does not become the registered H07 result and cannot identify a latitude
effect.

## Derivative-defined plateau estimand

V0 evaluated the first derivative of the photoperiod smooth. H07-002 retains
that direction and defines the requested visible plateau pattern for metric
$m$ and placement $q$ on a 100-point metric-specific grid. Let
$d_{mq}(p)$ be the first derivative of the fitted
`s(photoperiod_hours)` component on its model scale and let
$[L_{mq}(p), U_{mq}(p)]$ be its pointwise 95% interval. A
derivative-defined plateau pattern is present when:

1. at one grid point, $L_{mq}(p) > 0$, indicating a pointwise detected
   increase;
2. the immediately following grid point has
   $L_{mq}(p) \leq 0 \leq U_{mq}(p)$; and
3. every later grid point through the exact maximum recorded photoperiod also
   has an interval containing zero.

The transition is interval-censored by the grid and is reported as the
bracket between the final detected-increase point and the first
zero-compatible point. A later pointwise detected increase or decrease makes
the rule fail. No minimum tail length, site-overlap threshold, or
leave-one-site-out stability requirement is imposed.

## Equivalence and terminology boundary

No accepted source supplies a minimum practically important remaining slope
for any of the nine heterogeneous metrics. H07-002 intentionally replaces the
former equivalence estimand with a pointwise non-detection rule. Instrument
resolution, model standard errors, a percentage chosen for convenience, and
the two-AIC-point rule still cannot supply an equivalence margin.

```{r}
equivalence_margin_audit <- metric_contract |>
  transmute(
    Metric = manuscript_name,
    `Practical derivative scale` = case_when(
      effect_scale == "ratio" ~ paste0("Response ratio per photoperiod hour (", display_unit, ")"),
      TRUE ~ paste0("Difference per photoperiod hour (", display_unit, ")")
    ),
    `External or scientific margin` = "Not available",
    `H07-002 disposition` = "Assess a descriptive derivative-defined pattern"
  )

equivalence_margin_audit |>
  h07_gt(
    title = "Metric-specific equivalence-margin audit",
    note = "The revised pointwise rule is not an equivalence test. An interval containing zero does not show that the remaining slope is scientifically negligible.",
    font_size = 9
  ) |>
  cols_width(
    Metric ~ px(260),
    `Practical derivative scale` ~ px(330),
    `External or scientific margin` ~ px(180),
    `H07-002 disposition` ~ px(320)
  )
```

::: {.callout-important}
## H07-002 interpretation

Stage 2 may report that a metric does or does not meet the requested
**derivative-defined plateau pattern**. Because the decision is based on
pointwise non-detection and was adopted after viewing Stage 2 output, it must
remain labelled descriptive and post-result. It must not be translated into
proof that the derivative is zero, a biological maximum has been reached, or
the association generalizes independently of site and collection period.
:::

## Former sustained-support rule: diagnostic only

The original H07-001 design-support rule was defined from predictor geometry
and outcome availability without examining metric values. It required:

1. evaluate a 0.1 h photoperiod grid;
2. a site contributes at a grid point only when the point lies within that
   metric/site's 5th–95th photoperiod percentiles and the ±0.5 h local window
   contains at least five participants and 15 participant-days;
3. a pooled point requires at least three sites, 20 participants, 60
   participant-days, and no site contributing more than 50% of local days;
4. the pooled rule must still pass after removal of every site; and
5. a sustained interval requires at least 1.0 continuous h of eligible grid
   points.

H07-002 removes these thresholds as eligibility conditions. Their calculations
are retained below to document cross-site overlap and upper-tail composition;
they no longer censor the derivative grid or determine the revised plateau
classification.

```{r}
aggregate_local_support <- function(data) {
  if (nrow(data) == 0L) {
    return(c(
      sites = 0,
      participants = 0,
      participant_days = 0,
      max_site_share = NA_real_,
      passes = 0
    ))
  }
  site_n <- nrow(data)
  participants_n <- sum(data$participants)
  days_n <- sum(data$participant_days)
  max_share <- if (days_n > 0) max(data$participant_days / days_n) else NA_real_
  passes <-
    site_n >= 3L &&
    participants_n >= 20L &&
    days_n >= 60L &&
    is.finite(max_share) &&
    max_share <= 0.50
  c(
    sites = site_n,
    participants = participants_n,
    participant_days = days_n,
    max_site_share = max_share,
    passes = as.numeric(passes)
  )
}

support_at_grid_point <- function(data, grid_value) {
  by_site <- data |>
    group_by(site) |>
    summarise(
      q05 = quantile(photoperiod_hours, 0.05, names = FALSE),
      q95 = quantile(photoperiod_hours, 0.95, names = FALSE),
      participants = n_distinct(
        Id[abs(photoperiod_hours - grid_value) <= 0.5]
      ),
      participant_days = sum(
        abs(photoperiod_hours - grid_value) <= 0.5
      ),
      .groups = "drop"
    ) |>
    mutate(
      site_supported =
        grid_value >= q05 &
        grid_value <= q95 &
        participants >= 5L &
        participant_days >= 15L
    ) |>
    filter(site_supported)

  pooled <- aggregate_local_support(by_site)
  omitted_sites <- unique(data$site)
  loso_passes <- all(vapply(
    omitted_sites,
    function(omitted_site) {
      aggregate_local_support(
        by_site |> filter(site != omitted_site)
      )[["passes"]] == 1
    },
    logical(1)
  ))

  tibble(
    photoperiod_hours = grid_value,
    sites = pooled[["sites"]],
    participants = pooled[["participants"]],
    participant_days = pooled[["participant_days"]],
    max_site_share = pooled[["max_site_share"]],
    pooled_eligible = pooled[["passes"]] == 1,
    loso_eligible = loso_passes
  )
}

primary_near_support_grid <- candidate_long |>
  filter(
    data_scenario == "Primary dataset",
    placement == "Near eye",
    !is.na(value)
  ) |>
  group_by(metric_id, metric_order) |>
  group_modify(function(data, key) {
    grid <- seq(
      floor(min(data$photoperiod_hours) * 10) / 10,
      ceiling(max(data$photoperiod_hours) * 10) / 10,
      by = 0.1
    )
    map_dfr(grid, ~ support_at_grid_point(data, .x))
  }) |>
  ungroup()

support_run_points <- primary_near_support_grid |>
  group_by(metric_id) |>
  arrange(photoperiod_hours, .by_group = TRUE) |>
  mutate(
    run_id = cumsum(
      loso_eligible != lag(loso_eligible, default = FALSE)
    )
  ) |>
  ungroup() |>
  filter(loso_eligible)

support_runs <- if (nrow(support_run_points) == 0L) {
  tibble(
    metric_id = character(),
    run_id = integer(),
    lower = double(),
    upper = double(),
    span = double()
  )
} else {
  support_run_points |>
    group_by(metric_id, run_id) |>
    summarise(
      lower = min(photoperiod_hours),
      upper = max(photoperiod_hours),
      span = upper - lower,
      .groups = "drop"
    ) |>
    filter(span >= 1.0 - 1e-8)
}

support_rule_audit <- primary_near_support_grid |>
  group_by(metric_id, metric_order) |>
  summarise(
    pooled_range = h07_range_text(
      photoperiod_hours[pooled_eligible],
      digits = 1
    ),
    loso_range = h07_range_text(
      photoperiod_hours[loso_eligible],
      digits = 1
    ),
    .groups = "drop"
  ) |>
  left_join(
    support_runs |>
      group_by(metric_id) |>
      summarise(
        sustained_intervals = paste0(
          sprintf("%.1f–%.1f h", lower, upper),
          collapse = "; "
        ),
        .groups = "drop"
      ),
    by = "metric_id"
  ) |>
  mutate(
    pooled_range = if_else(pooled_range == "None", "None", paste0(pooled_range, " h")),
    loso_range = if_else(loso_range == "None", "None", paste0(loso_range, " h")),
    sustained_intervals = replace_na(sustained_intervals, "None")
  ) |>
  left_join(
    metric_contract |> select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  arrange(metric_order)

support_rule_audit |>
  transmute(
    Metric = manuscript_name,
    `Pooled eligible grid range` = pooled_range,
    `LOSO-stable grid range` = loso_range,
    `LOSO-stable interval ≥1.0 h` = sustained_intervals
  ) |>
  h07_gt(
    title = "Outcome-blinded diagnostic application of the former support rule",
    note = "Only missing-versus-available outcome status and predictor geometry are used. H07-002 retains this result as a limitation display, not an eligibility gate for the revised derivative classification.",
    font_size = 10
  ) |>
  cols_width(
    Metric ~ px(300),
    `Pooled eligible grid range` ~ px(190),
    `LOSO-stable grid range` ~ px(190),
    `LOSO-stable interval ≥1.0 h` ~ px(220)
  )
```

The primary near-eye design does not pass the former cross-site LOSO rule.
Under H07-002 this shows that the upper curve cannot be interpreted as a
site-replicated photoperiod response; it does not prevent asking whether the
requested pattern is visible in the pooled fitted association.

# Exact proposed model objects — no fits

## Response construction and identifiers

Stage 2 will create `response_value` before fitting:

- `log10(value + 0.1)` for daily mean, M10, L10, longest period, and corrected
  dose, fitted with a Gaussian family;
- the natural non-negative value for the three Tweedie-log duration models,
  fitted with `mgcv::tw(link = "log", a = 1.01, b = 1.99)` so the
  compound-Poisson--gamma power is estimated inside fixed prespecified bounds;
  and
- the natural-hour value for pre-sleep time, fitted with identity Gaussian.

The Tweedie contract is not justified by zero values and right skew alone. It
assumes a compound occurrence--severity process for time above 1,000 lx melEDI,
time above 250 lx melEDI during wake, and time below 1 lx melEDI during sleep.
Stage 2 must report the fitted power, dispersion, distance from both power
bounds, observed-versus-simulated zero frequency, and positive-tail
calibration. A failure opens only the already accepted common-family decision
gate; it cannot trigger a result-specific switch. If a two-part occurrence and
positive-duration model is later judged scientifically necessary, its exact
form must be approved before outcome-specific H07 results are inspected.

In the installed `mgcv` 1.9-4 environment, Stage 2 will attach `mgcv` before
fitting `tw()` models and record that package state. The exact proposed family
calls are visible below; this cell constructs calls but fits no model.

```{r}
#| label: proposed-h07-family-calls
#| echo: true
#| code-fold: false

proposed_h07_family_calls <- list(
  gaussian_identity = quote(stats::gaussian(link = "identity")),
  tweedie_log_estimated_power = quote(
    mgcv::tw(link = "log", a = 1.01, b = 1.99)
  )
)

enframe(
  proposed_h07_family_calls,
  name = "family_id",
  value = "call"
) |>
  mutate(call = map_chr(call, h07_formula_text))
```

`site_participant` is `interaction(site, Id, drop = TRUE)` after the submitted
site levels are fixed. Every model frame retains the original value, practical
unit, transformation, local date, absolute latitude, photoperiod, site,
participant, and exact row key.

## Registered tensor and comparison formulas

The main tensor uses conservative fixed marginal basis dimensions
`k = c(4, 5)`. The expanded `k = c(5, 8)` formula is a mandatory basis-
stability diagnostic, not an alternative selected by its p-value. The main
tensor has a nominal marginal-basis product of 4 x 5 = 20 coefficients before
constraints; the expanded diagnostic has 5 x 8 = 40. Their information demand
is judged against only nine near-eye or eight chest latitude/site stripes and
the metric-specific joint support, never against the participant-day row count
alone. `te()` is retained because the complete joint surface is the estimand
and no separate marginal smooths are included. The explicitly displayed `tp`
margins are a pre-result basis decision; they will not be replaced after
seeing results.

Every proposed primary H07 model uses `mgcv::gam(method = "REML")`. The null
and bilinear formulas remain visible because they are the exact registered
comparison structures and are fitted on the identical metric-specific rows
for conditional AIC and the prespecified approximate nested comparison. The
fixed-site version tests sensitivity to treating only nine site levels as a
penalized random smooth; it cannot identify a latitude effect conditional on
a saturated site factor and is interpreted only as a stability diagnostic.
No numerical fit overrides the rank, concurvity, support-surface, or
leave-one-site-out identifiability gates.

```{r}
#| label: proposed-h07-registered-formulas
#| echo: true
#| code-fold: false

proposed_h07_registered_formulas <- list(
  registered_null = stats::as.formula(
    "response_value ~ s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  registered_linear_surface = stats::as.formula(
    "response_value ~ abs_latitude_deg * photoperiod_hours + s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  registered_tensor = stats::as.formula(
    "response_value ~ te(abs_latitude_deg, photoperiod_hours, k = c(4, 5), bs = c('tp', 'tp')) + s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  registered_tensor_expanded_basis = stats::as.formula(
    "response_value ~ te(abs_latitude_deg, photoperiod_hours, k = c(5, 8), bs = c('tp', 'tp')) + s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  registered_tensor_fixed_site = stats::as.formula(
    "response_value ~ te(abs_latitude_deg, photoperiod_hours, k = c(4, 5), bs = c('tp', 'tp')) + site + s(site_participant, bs = 're')"
  )
)

enframe(
  proposed_h07_registered_formulas,
  name = "formula_id",
  value = "formula"
) |>
  mutate(formula = map_chr(formula, h07_formula_text))
```

## Conditional photoperiod-only adaptation

These formulas are used only if the registered tensor fails the predeclared
identifiability criteria and the author approves H07-G2. The adapted question
is a site-adjusted photoperiod association over observed collection periods;
it is not the registered joint latitude–photoperiod hypothesis.

```{r}
#| label: proposed-h07-adapted-formulas
#| echo: true
#| code-fold: false

proposed_h07_adapted_formulas <- list(
  adapted_photoperiod_linear = stats::as.formula(
    "response_value ~ photoperiod_hours + s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  adapted_photoperiod_smooth = stats::as.formula(
    "response_value ~ s(photoperiod_hours, k = 6, bs = 'tp') + s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  adapted_photoperiod_expanded_basis = stats::as.formula(
    "response_value ~ s(photoperiod_hours, k = 10, bs = 'tp') + s(site, bs = 're') + s(site_participant, bs = 're')"
  ),
  adapted_photoperiod_fixed_site = stats::as.formula(
    "response_value ~ s(photoperiod_hours, k = 6, bs = 'tp') + site + s(site_participant, bs = 're')"
  )
)

enframe(
  proposed_h07_adapted_formulas,
  name = "formula_id",
  value = "formula"
) |>
  mutate(formula = map_chr(formula, h07_formula_text))
```

The registered null formula is reused as the adapted null. Primary, chest,
paired/common, gap-timing-unaware, exact-period, observed-dose, and leave-one-
site-out runs reuse these formula objects; only the predeclared rows or metric
variant change.

## Comparison and estimation contract

```{r}
all_proposed_formulas <- c(
  proposed_h07_registered_formulas,
  proposed_h07_adapted_formulas
)

comparison_registry <- tribble(
  ~comparison_id, ~full_formula, ~reduced_formula, ~question, ~implementation, ~role,
  "registered_joint_association", "registered_tensor", "registered_null", "Is the complete supported latitude–photoperiod tensor zero?", "Approximate summary.gam() whole-term test from the full REML fit; reduced REML fit used for conditional AIC only", "Primary registered association",
  "registered_flexible_shape", "registered_tensor", "registered_linear_surface", "Does the surface depart from its bilinear latitude-by-photoperiod subspace?", "Prespecified approximate anova.gam() nested comparison plus conditional AIC on identical rows", "Shape evidence; not a ceiling test",
  "registered_basis_stability", "registered_tensor_expanded_basis", "registered_tensor", "Are fitted surfaces and derivatives stable to the predeclared expanded basis?", "Separate exact REML diagnostic fit; no p-value search", "Basis diagnostic",
  "registered_site_representation", "registered_tensor_fixed_site", "registered_tensor", "Is the surface materially dependent on random- versus fixed-site representation?", "Separate exact REML diagnostic fit; no nested test", "Site-representation diagnostic",
  "adapted_photoperiod_association", "adapted_photoperiod_smooth", "registered_null", "If approved, is the complete supported photoperiod smooth zero?", "Approximate summary.gam() whole-term test from the full REML fit; reduced REML fit used for conditional AIC only", "Conditional adapted association",
  "adapted_photoperiod_shape", "adapted_photoperiod_smooth", "adapted_photoperiod_linear", "If approved, does the curve depart from a straight-line subspace?", "Prespecified approximate anova.gam() nested comparison plus conditional AIC on identical rows", "Adapted shape evidence; not a ceiling test",
  "adapted_basis_stability", "adapted_photoperiod_expanded_basis", "adapted_photoperiod_smooth", "Is the adapted curve stable to the predeclared expanded basis?", "Separate exact REML diagnostic fit; no p-value search", "Basis diagnostic",
  "adapted_site_representation", "adapted_photoperiod_fixed_site", "adapted_photoperiod_smooth", "Is the adapted curve materially dependent on site representation?", "Separate exact REML diagnostic fit; no nested test", "Site-representation diagnostic"
) |>
  mutate(
    full = map_chr(full_formula, ~ h07_formula_text(all_proposed_formulas[[.x]])),
    reduced = map_chr(reduced_formula, ~ h07_formula_text(all_proposed_formulas[[.x]]))
  )

comparison_registry |>
  transmute(
    Comparison = comparison_id,
    `Full formula` = full,
    `Reduced/comparison formula` = reduced,
    Question = question,
    Role = role
  ) |>
  h07_gt(
    title = "Exact H07 comparison registry",
    note = "Every primary model uses mgcv::gam(method = 'REML') on one locked metric-specific frame. Conditional AIC is relative-support evidence, not a test. Approximate p-values are withheld when identifiability, nesting, effective-df ordering, or residual-dependence conditions fail.",
    font_size = 8
  ) |>
  cols_width(
    Comparison ~ px(165),
    `Full formula` ~ px(315),
    `Reduced/comparison formula` ~ px(300),
    Question ~ px(300),
    Role ~ px(175)
  )
```

## H11 computational lesson and H07 inference-validation gate

H11 established a project-local workaround after its large planned ML,
`discrete = FALSE` fits ran for hours without a durable checkpoint. Its
accepted test used an existing Gaussian identity-link fREML/discrete fit,
AR-whitened raw-residual scores, participant CR1 covariance, `Vp - Ve`, and a
finite-cluster fractional-rank F reference. Those identities were verified for
that H11 model. They do not establish a general score, bread, covariance, or
reference distribution for H07's exact REML Gaussian-transformed and
Tweedie-log models.

H07 therefore adopts only the portable lessons: checkpoint long fits; avoid an
unnecessary `bam(discrete = FALSE)` branch; define the scientific contrast
before results; and do not confuse failed computation with evidence. The
primary H07 p-values use maintained `mgcv` interfaces under the explicit
conditional-inference limitations above. The ordinary term test is not called
cluster robust, and it is not interpreted when residual dependence,
concurvity, or support failure invalidates that approximation.

If Stage 2 residual diagnostics show material within-participant dependence,
the affected confirmatory support label is withheld. A participant-cluster
repair can proceed only through a displayed 50- or 100-replicate pilot and a
new author decision. The preferred pilot is a participant-cluster refit
bootstrap that repeats the complete prespecified fit and test path. A custom
sandwich pilot is permissible only if all of the following are demonstrated:

1. the Gaussian identity-link implementation reproduces the accepted H11
   score/bread result when AR whitening is absent or deliberately matched;
2. Gaussian transformed-response and Tweedie-log scores and bread are checked
   against numerical likelihood derivatives and the model's working weights;
3. smoothing-parameter conditioning and any `Vp - Ve` addition are reported
   separately rather than assumed to be generally conservative;
4. covariance eigenvalues, participant meat shares, contrast rank, finite-
   cluster denominator degrees of freedom, and representation invariance pass
   prespecified tolerances; and
5. no production p-value relies on the unexported `mgcv:::testStat()` interface
   without a version-frozen validation against a maintained calculation.

Neither repair solves shared-site dependence, the nine/eight-site
latitude--site confounding, or sparse joint support. Fixed-site and
leave-one-site-out analyses remain mandatory limitations and influence checks.

The preregistered two-AIC-point rule and V0's fREML `dAIC` values are retained
as conditional-AIC relative-support evidence. New H07 conditional AIC is
computed from the exact REML candidate fits only after all commensurability and
`edf2` checks pass. No ML likelihood-ratio branch or AIC-derived significance
label is proposed.

# Multiplicity, derivatives, and uncertainty

## Fixed families

```{r}
#| label: h07-bh-function
#| echo: true
#| code-fold: false

h07_adjust_complete_family <- function(p_raw) {
  stopifnot(length(p_raw) == 9L)
  stats::p.adjust(p_raw, method = "BH", n = 9L)
}
```

```{r}
multiplicity_registry <- tribble(
  ~family_id, ~placement_or_scenario, ~members, ~status, ~interpretation,
  "H07-F1-registered-association", "Primary near eye", 9L, "Active", "Approximate whole-tensor term tests; complete BH vector",
  "H07-F3-registered-nonlinear-shape", "Primary near eye", 9L, "Active secondary", "Approximate prespecified nested shape tests; not a plateau family",
  "H07-F1C-registered-association", "Complementary chest", 9L, "Complementary", "Same conditional whole-term tests; cannot override primary",
  "H07-F3C-registered-nonlinear-shape", "Complementary chest", 9L, "Complementary", "Same conditional nonlinear-shape tests; cannot override primary",
  "H07-F1P-registered-association", "Paired/common placement", 9L, "Sensitivity", "Separate placement-specific association families on identical keys",
  "H07-F3P-registered-nonlinear-shape", "Paired/common placement", 9L, "Sensitivity", "Separate placement-specific shape families on identical keys",
  "H07-F1G-registered-association", "Gap-timing-unaware dataset", 9L, "Sensitivity", "Scenario-specific association family",
  "H07-F3G-registered-nonlinear-shape", "Gap-timing-unaware dataset", 9L, "Sensitivity", "Scenario-specific nonlinear-shape family",
  "H07-F1A-adapted-association", "Conditional adapted runs", 9L, "Inactive pending H07-G2", "Separately labelled association family; never substituted silently for the registered family",
  "H07-F3A-adapted-nonlinear-shape", "Conditional adapted runs", 9L, "Inactive pending H07-G2", "Separately labelled shape family; not a plateau family",
  "H07-F2R-derivative-pattern", "Primary near eye", 9L, "Descriptive post-result", "Pointwise derivative transition; no curve-wide or cross-metric adjustment",
  "H07-F2RC-derivative-pattern", "Complementary chest", 9L, "Descriptive post-result", "Same rule; cannot override the primary placement"
)

multiplicity_registry |>
  h07_gt(
    title = "Proposed H07 multiplicity families",
    note = "Association and nonlinear shape retain their complete nine-member families. H07-002's derivative pattern is a post-result descriptive classification based on unadjusted pointwise intervals, not a multiplicity-controlled test family.",
    font_size = 9
  ) |>
  cols_width(
    family_id ~ px(235),
    placement_or_scenario ~ px(210),
    members ~ px(70),
    status ~ px(145),
    interpretation ~ px(430)
  )
```

All reader-facing p-values follow REPORT-008: leading zero, three decimal
places, `<0.001` below that threshold, and bold only when significant under
the explicitly labelled raw or adjusted rule. Full precision remains in
source CSVs.

## Derivative and prediction contract

Stage 2 will recreate V0's pointwise derivative rule exactly for historical
comparison, including its 100-point default grid. Under H07-002, the revised
derivative assessment will:

- address metrics by ID and show each intended metric once;
- evaluate `gratia::derivatives()` for `s(photoperiod_hours)` on the model/link
  scale at 100 equally spaced points from the exact metric-specific minimum to
  maximum recorded photoperiod;
- use central differences with forward/backward boundary handling, a 0.01 h
  step, pointwise 95% intervals, and `unconditional = TRUE` where available;
- classify the first positive-to-zero-compatible transition only when all
  later pointwise intervals contain zero through the recorded maximum;
- report the transition as a grid bracket and report the remaining tail span;
- retain the pooled-support and leave-one-site-out results as diagnostics, not
  masks or eligibility criteria;
- reproduce the V0 forward, `eps = 1e-7`, conditional pointwise settings as a
  method sensitivity; and
- call the result a derivative-defined plateau pattern, explicitly not an
  equivalence test, asymptote, mechanistic ceiling, or causal effect.

The requested rule uses analytic pointwise intervals and therefore requires no
production simulation. A simultaneous-band sensitivity would require a new
100-replicate full-range pilot and separate production approval; it would test
a stricter curve-wide question and would not turn non-detection into
equivalence.

# Diagnostics, influence, and sensitivity contract

## Mandatory diagnostics and stop rules

```{r}
diagnostic_registry <- tribble(
  ~domain, ~required_check, ~interpretation_or_action,
  "Input and exact samples", "Hash identity; key uniqueness; full/diagnostic-fit row identity; participants, participant-days, observations, sites overall/by site", "Any mismatch aborts the affected analysis; unavailable support is never coded as zero",
  "Joint support", "Metric-specific site/date/latitude/photoperiod map; local support; empty surface; interpolation distance; boundary composition", "Under H07-002 this limits transportability and independent photoperiod interpretation but does not mask the requested pooled derivative classification",
  "Identifiability", "Design rank; tensor coefficient rank; gratia::model_concurvity(); condition; unique latitude stripes; site-versus-latitude competition", "A numerical fit or small p-value is not enough; material confounding makes the tensor non-identifiable",
  "Basis and smoothing", "Effective degrees of freedom; mgcv::gam.check() k-index and residual panels; 20/40 nominal tensor products; main versus expanded basis; smoothing-parameter stability", "Material curve/derivative change or inadequate k makes shape non-estimable pending a new gate",
  "Convergence", "Optimizer return; gradient; Hessian; finite coefficients; gam.check(); reproducibility from a fresh R session", "Failure is reported as non-estimable; do not simplify after seeing the result",
  "Gaussian distribution", "Name the residual type; gratia::appraise(); residual-versus-fitted; Q-Q; tails; heteroscedasticity; transformed prediction range; influence", "A common response-family failure reopens the metric family contract",
  "Tweedie distribution", "mgcv::tw power and dispersion with distances to [1.01, 1.99] bounds; named deviance/randomized residuals; observed-versus-simulated zeros and positive-tail summaries; natural-scale prediction bounds", "Rootograms are unsupported for Tweedie in gratia 0.11.2; strong joint zero/tail failure reopens one common predeclared family alternative, not a result-specific switch",
  "Physical and construct bounds", "Non-negativity; 24 h full-day bound; state-specific denominators; pre-sleep >6 h audit; dose support; exact-period bounds", "Impossible prediction or construct violation triggers model review",
  "Random effects", "Site and participant variance/edf; boundary shrinkage; nine/eight-site instability", "Do not remove participant clustering to obtain a p-value; report weak site variance identification",
  "Temporal dependence", "Named residual ACF within participant and contiguous-date sequence, reset at participant and date gaps; site summaries; consecutive-day runs; gaps; short sequences", "Never pool unrelated series into one ACF; material residual dependence withholds confirmatory support and opens the pilot gate",
  "Influence", "Participant leverage; response-boundary leverage; leave-one-site-out models; high-photoperiod site leverage", "A site-sensitive pattern remains reportable under H07-002 only as a pooled descriptive association, not a general or mechanistic ceiling",
  "Conditional inference", "summary.gam whole-term extraction; anova.gam nesting and effective-df ordering; smoothing-parameter conditioning; residual-dependence gate; exact nine-value vectors", "Failed identifiability, nesting, edf ordering, or residual-dependence conditions make the affected p-value non-estimable",
  "Conditional AIC", "Identical rows, response, family/link, weights, offset, engine, likelihood; logLik.gam dispatch; edf2 availability; delta sign", "Report only relative support within the prespecified pair; never significance, a p-value, or ceiling evidence",
  "Optional cluster repair", "Pilot-only cluster refit/bootstrap or validated family-specific scores and bread; covariance eigenvalues; rank/df; participant meat shares; maintained-reference check", "No custom sandwich or production resampling result is released before the pilot and separate approval",
  "Fitting provenance", "Exact gam/REML primary; exact family/link/formula; durable checkpoint; timed attempt; if invoked, bam/fREML/discrete fallback with nthreads and BLAS record plus exact-fit comparison", "Any ML, bam/discrete-false, unvalidated fallback, or result-selected branch aborts release pending a new author gate",
  "Derivative inference", "First derivative of s(photoperiod_hours); 100-point exact-range grid; central plus boundary differences; pointwise interval; unconditional covariance; V0-setting sensitivity", "Report the requested positive-to-zero-compatible tail as a post-result derivative-defined pattern, never equivalence or an asymptote",
  "Placement", "Same estimand and exact paired keys; separate fits; component intervals", "No pooling, unmatched identity plot, placement equivalence, or chest override",
  "Data sensitivity", "Scenario-specific and exact common samples; association, curve, diagnostics, support, and influence stability", "Classify stable, magnitude-sensitive, precision-sensitive, direction-sensitive, support-sensitive, or non-estimable",
  "Multiplicity", "Complete nine-row raw vectors; independent R reproduction of BH; REPORT-008 display", "Scalar adjustment, shortened vectors, or recycled q-values aborts release"
)

diagnostic_registry |>
  h07_gt(
    title = "Mandatory Stage 2 diagnostic interpretation",
    note = "A diagnostic is incomplete until the report explains what it means for the scientific estimand. Automatic plots and software warnings are evidence inputs, not final classifications.",
    font_size = 8
  ) |>
  cols_width(
    domain ~ px(150),
    required_check ~ px(520),
    interpretation_or_action ~ px(540)
  )
```

## Fixed sensitivity matrix

```{r}
sensitivity_matrix <- tribble(
  ~order, ~sensitivity, ~change, ~held_fixed, ~role, ~reopen_trigger,
  1L, "Primary near eye", "None", "Approved metric, tensor, support, and family contracts", "Primary registered analysis", "Any mandatory fit or identifiability failure",
  2L, "Complementary chest", "Placement and available site set", "Metric, formula, family, and diagnostic rules", "Complementary", "Material disagreement with near eye",
  3L, "Paired/common placement", "Restrict both placements to identical eight-site participant-days", "Separate fits and common effect scale", "Direct placement concordance", "Direction/magnitude incompatibility or invalid direct scale",
  4L, "Gap-timing-unaware dataset", "Prepared metric values and available days", "H07 implementation, support rule, and formulas", "Preparation sensitivity", "Material estimate, interval, support, influence, or conclusion change",
  5L, "Exact common preparation sample", "Restrict primary and gap-timing-unaware runs to identical keys", "Formula, response family, and predictors", "Value-versus-availability separation", "Scenario difference persists or reverses on common rows",
  6L, "Longest period exact-only", "Use only days whose observed lower and possible upper bounds coincide", "Formula and family", "Period-identifiability sensitivity", "Material curve, interval, or support change",
  7L, "Observed dose", "Replace corrected dose with observed dose on exact common rows", "Formula and family", "Correction-dependence sensitivity", "Material curve, interval, or support change",
  8L, "Expanded basis", "Use the displayed larger k", "Rows, family, random effects, and predictors", "Basis diagnostic", "Material curve/derivative change",
  9L, "Fixed-site representation", "Replace the penalized site smooth by fixed site", "Rows, participant effect, and predictor smooth", "Site-stability diagnostic", "Material shape change or exposed confounding",
  10L, "Leave one site out", "Remove one whole site per refit", "Response and approved formula structure", "Influence diagnostic", "Association/shape driven by one site or support collapses",
  11L, "Conditional photoperiod-only", "Remove latitude tensor and fit displayed site-adjusted photoperiod formulas", "Nine outcomes, rows, family, and multiplicity", "Adapted analysis only after H07-G2", "Cannot be interpreted as registered H07 or as a ceiling",
  12L, "Temporal-precision repair", "Participant-cluster resampling or approved correlation sensitivity if required", "Estimand, formula, outcome, and support", "Conditional heavy-compute sensitivity", "Pilot indicates failure, instability, or prohibitive runtime"
)

sensitivity_matrix |>
  select(-order) |>
  h07_gt(
    title = "Proposed H07 sensitivity battery",
    note = "Sensitivities are interpreted for stability and are never searched for a preferred p-value. No season, temperature, weather, holiday, activity, or other unsupported predictor is added.",
    font_size = 8
  ) |>
  cols_width(
    sensitivity ~ px(210),
    change ~ px(330),
    held_fixed ~ px(300),
    role ~ px(190),
    reopen_trigger ~ px(300)
  )
```

Collection period is handled as a design limitation and support display, not
as an improvised categorical season covariate. Within a site, photoperiod is a
deterministic function of date; across sites, calendar periods, routines,
weather, holidays, and latitude differ. The fitted association is therefore
observational and collection-period-specific. It is not a seasonal, climatic,
behavioural, or causal effect.

# Stage 2 implementation and V0-comparison plan

After approval, Stage 2 will:

1. build an H07-owned model-row artifact from the four verified inputs, with
   exact keys, response transforms, site order, `site_participant`, metric
   variants, exclusions, and predictor support;
2. attempt the registered tensor for every one of the nine metrics without
   consulting H01 support or H07 results;
3. fit and checkpoint the exact full, null, and bilinear
   `mgcv::gam(method = "REML")` candidates on each locked metric frame; extract
   the maintained whole-term and prespecified nested approximate tests and
   audited conditional AIC; run no ML or `bam(discrete = FALSE)` branch;
4. if an exact fit is not computationally practicable, display its timed
   attempt before invoking the `bam(method = "fREML", discrete = TRUE)`
   fallback as a provisional pilot; no fallback result is reportable until the
   same candidate on identical rows is benchmarked against an exact fit for
   fitted values, effective degrees of freedom, conditional AIC, supported
   predictions, and derivatives, or a new author gate explicitly revises that
   requirement;
5. report exact fitted participants, participant-days, observations, and
   sites overall and by site, plus every 95% confidence interval and
   REPORT-008 p-value;
6. interpret convergence, distributional, residual, physical-bound,
   temporal, identifiability, basis, support-surface, random-effect, and
   influence diagnostics rather than merely displaying them;
7. run near-eye primary, chest complementary, paired/common placement,
   gap-timing-unaware, exact-common preparation, exact-period, observed-dose,
   basis, site-representation, and leave-one-site-out branches;
8. classify an unidentifiable tensor as such and run the photoperiod-only
   adaptation only if H07-G2 is approved;
9. reproduce V0's H01-dependent candidate set, fREML AIC screen, default
   pointwise derivative grid, and last-positive rule, then apply H07-002's
   positive-to-zero-compatible-through-the-end rule to the adapted models and
   separate changes caused by outcome selection, preparation, response family,
   tensor scope, comparison method, and derivative settings;
10. recreate recognizable V0 near-eye and chest tables/figures while correcting
   the duplicated mean, missing M10, positional indexing, terminology,
   readability, and absent source data; and
11. stop before any production participant-cluster bootstrap, custom sandwich
    release, or simultaneous-band simulation
    after displaying the required 50/100-replicate pilot and runtime estimate.

Under H07-002, Stage 2 reports a pointwise derivative-defined plateau pattern
when the exact operational rule is met. It does not use a non-significant
whole tensor, AIC threshold, or visual flattening for that classification and
does not interpret the pattern as equivalence, an asymptote, or a causal or
mechanistic ceiling.

# Original Stage 1 decisions and post-Stage 2 amendment

The owner approved H07-G1 through H07-G16 as decision `H07-001`. After viewing
the Stage 2 support output, the owner approved `H07-002`, which supersedes the
ceiling disposition in H07-G1, the eligibility role of H07-G9, and the
derivative prohibition in H07-G13. The remaining original decisions are
retained below as the historical implementation record.

The original pre-Stage 2 decision package was:

1. **H07-G1 — ceiling claim (superseded by H07-002):** accept the recommendation that no defensible
   metric-specific equivalence margin is available, disable
   `H07-F2-plateau-equivalence`, and prohibit ceiling/plateau claims for all
   nine outcomes. Any external margins must be supplied and approved before
   Stage 2 results are examined.
2. **H07-G2 — conditional adaptation:** if the registered tensor is classified
   non-identifiable by the fixed diagnostic rule, approve a separately
   labelled photoperiod-only adapted analysis. It cannot replace or be called
   the registered H07 test.
3. **H07-G3 — outcome set:** include all nine planned metrics independently of
   H01/H02 fitted results and retain reason-coded non-estimable family members.
4. **H07-G4 — estimands and maintained inference:** approve the conditional
   approximate `summary.gam()` whole-tensor test, the prespecified approximate
   nested `anova.gam()` departure from the bilinear subspace, audited
   conditional AIC as secondary relative-support evidence, and practical-scale
   directional derivatives restricted to observed support. None is a plateau
   test, and failed identifiability or residual-dependence gates withhold a
   support claim.
5. **H07-G5 — response package:** inherit the accepted H01/H05 metric-specific
   transforms, families, links, practical scales, and common-family failure
   rule; implement estimated-power `mgcv::tw(link = "log", a = 1.01,
   b = 1.99)` for the three Tweedie metrics with power/dispersion/boundary and
   joint zero/positive-tail diagnostics; reject V0's
   one-log-Gaussian-model-for-all implementation.
6. **H07-G6 — formulas and basis:** approve the exact displayed Wilkinson
   formulas, `te()` with `tp` margins, nominal 20/40 main/expanded tensor
   products, participant nested in site, fixed-site diagnostic, and exact
   `gam(method = "REML")` as the primary engine. Approve checkpointed
   `bam(method = "fREML", discrete = TRUE)` only as the documented fallback
   after a timed attempt; treat it as provisional until the same candidate and
   rows pass the exact-fit stability check.
7. **H07-G7 — H11-informed computation and repair gate:** remove new ML and
   `bam(discrete = FALSE)` branches; do not assume H11's Gaussian/AR-whitened
   CR1 construction transfers to transformed Gaussian or Tweedie H07 models;
   use maintained conditional GAM inference as primary and require a displayed
   50/100-replicate method-validation pilot plus separate approval before any
   participant-cluster refit bootstrap or custom sandwich becomes reportable.
   Retain both V0 and new audited conditional AIC as relative-support evidence,
   never significance.
8. **H07-G8 — multiplicity:** approve separate complete nine-member BH
   families for association and nonlinear shape in registered primary and
   separately named complementary/sensitivity runs; prohibit scalar
   adjustment and family shortening.
9. **H07-G9 — support rule (eligibility role superseded by H07-002):** approve or revise the 0.1 h grid, central 90%
   site range, ±0.5 h window, 5-participant/15-day site threshold,
   3-site/20-participant/60-day pooled threshold, 50% site-share cap, full
   leave-one-site-out requirement, and 1.0 h sustained interval before any
   outcome result is examined.
10. **H07-G10 — placement and samples:** near-eye all-available is primary;
    chest all-available is complementary; direct placement comparison uses
    separate fits on exact paired/common participant-days; no pooling or
    equivalence language.
11. **H07-G11 — preparation and metric sensitivities:** approve the
    gap-timing-unaware scenario-specific and exact-common comparisons,
    exact-period sensitivity, observed-dose sensitivity, and no result-based
    scenario selection.
12. **H07-G12 — collection period and predictors:** treat collection date,
    site, latitude, and photoperiod confounding as a limitation/support audit;
    add no season category, temperature, weather, holiday, behavioural, or
    other unsupported predictor.
13. **H07-G13 — V0 derivatives (revised by H07-002):** reproduce V0's exact default pointwise
    derivative and last-positive rule only as a historical comparator; use
    central `gratia::partial_derivatives()` and
    `gratia::response_derivatives()` or full-predictor covariance-draw finite
    differences, supported grids, recorded uncertainty settings, and
    simultaneous bands for audited shape interpretation; never relabel
    end-of-detected-increase as a ceiling.
14. **H07-G14 — diagnostics and stop rules:** approve the full diagnostic
    registry, non-identifiable/non-estimable dispositions, fixed basis/site
    sensitivities, participant influence, and leave-one-site-out checks.
15. **H07-G15 — temporal dependence and computation:** interpret date-lag
    residual dependence within participant-specific contiguous-date series;
    no primary step may require `bam(discrete = FALSE)`; material residual
    dependence withholds confirmatory support; if a resampling, custom
    covariance, or correlation sensitivity is required, run only a 50- or
    100-replicate pilot and obtain separate production approval.
16. **H07-G16 — reporting language:** report observational associations and
    collection-period limitations; make no causal, clinical, health-outcome,
    placement-equivalence, seasonal-mechanism, or unsupported ceiling claim.

::: {.callout-caution}
## Approval status

`H07-001` authorized Stage 2. `H07-002` subsequently reopened and revised the
plateau and support gates. The revised Stage 2 report must now be reviewed
again before Stage 3 begins. No production bootstrap or simultaneous-band
simulation is authorized by either decision.
:::

# Reproducibility record

```{r}
quarto_version <- tryCatch(
  system2("quarto", "--version", stdout = TRUE, stderr = FALSE)[1L],
  error = function(error) "Unavailable"
)

tibble(
  Item = c(
    "R",
    "Quarto",
    "mgcv",
    "gratia",
    "dplyr",
    "ggplot2",
    "gt",
    "digest",
    "rvest",
    "Stage 1 inferential fits",
    "Stage 1 production simulations"
  ),
  `Version or value` = c(
    as.character(getRversion()),
    quarto_version,
    as.character(packageVersion("mgcv")),
    as.character(packageVersion("gratia")),
    as.character(packageVersion("dplyr")),
    as.character(packageVersion("ggplot2")),
    as.character(packageVersion("gt")),
    as.character(packageVersion("digest")),
    as.character(packageVersion("rvest")),
    "0",
    "0"
  )
) |>
  h07_gt(
    title = "Stage 1 execution environment",
    note = "Rendered in a fresh R 4.6.1 session with the existing project library. No package was installed or updated.",
    font_size = 10
  ) |>
  cols_width(Item ~ px(300), `Version or value` ~ px(210))
```

The bounded render command is:

```bash
quarto render audit/hypotheses/H07/01_audit_and_plan.qmd --to html
```
