---
title: "H04 Stage 1: audit and proposed analysis plan"
subtitle: "Nature Health retargeting — multi-select activity contract"
date: last-modified
format:
  html:
    page-layout: full
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    embed-resources: true
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H04_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

stopifnot(identical(as.character(getRversion()), "4.6.1"))

project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))

library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(gt)

source(file.path(
  project_root,
  "scripts/hypotheses/H04/h04_stage1_support.R"
))

stage1_gt <- function(data, rowname_col = NULL, groupname_col = NULL) {
  gt(data, rowname_col = rowname_col, groupname_col = groupname_col) |>
    opt_row_striping() |>
    tab_options(
      table.font.size = px(13),
      data_row.padding = px(4),
      heading.align = "left",
      table.width = pct(100)
    )
}

formula_text <- function(x) {
  paste(deparse(x, width.cutoff = 500L), collapse = " ")
}

count_text <- function(x) {
  format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
}

site_registry <- read.csv(
  file.path(project_root, "config/site_display_registry.csv"),
  check.names = FALSE
) |>
  arrange(display_order)

input_paths <- c(
  normalized_diary = paste0(
    "artifacts/06_model_data/normalized_inputs/",
    "lightexposurediary.rds"
  ),
  primary_near_eye = paste0(
    "artifacts/06_model_data/base/",
    "metrics_glasses_one_hour_context.rds"
  ),
  complementary_chest = paste0(
    "artifacts/06_model_data/base/",
    "metrics_chest_one_hour_context.rds"
  ),
  gap_timing_unaware = paste0(
    "artifacts/06_model_data/scenarios/manuscript_prepared_data/",
    "one_hour_data.rds"
  )
)
stopifnot(all(file.exists(file.path(project_root, input_paths))))

raw_inputs <- setNames(
  lapply(file.path(project_root, input_paths), readRDS),
  names(input_paths)
)
```

::: {.callout-note title="Stage 1 approved — Stage 2 authorized"}
The fixed long-format activity contract is implementable without duplicating
the analytical contribution of a participant-hour. After collapsing the three
outdoor flags, suppressing co-selected `other`, deduplicating labels, and
assigning weight $1/k$, the proposed additive design is full rank for both
placements. The candidate near-eye frame contains **16,526 unique
participant-hours and 17,266 long rows**, whose weights sum to **16,526
effective hours**. The chest frame contains **20,128 unique hours and 21,071
long rows**, whose weights sum to **20,128 effective hours**.

H03's population-mean quasi-Tweedie architecture transfers to the same
one-hour outcome, but H03's conclusions and significance classifications do
not. H04 requires a different primary restriction, fractional prior weights,
a named-category site-heterogeneity companion that excludes `Other`, and
careful dependence handling for a weighted long-row temporal model. The author
approved H04-G1 through H04-G8 and authorized Stage 2 on 2026-08-10. This
document fits **no inferential model**, estimates no category mean, calculates
no p-value, and contains no Stage 2 analysis.
:::

## Scope and safeguards

The authorized measurement-methods audit was used read-only to trace the
signed record, developed plan, reviewer request, V0 sources and frozen output,
H03 final architecture, shared reporting decisions, and every preparation
artifact consumed below. The authorized Quarto-authoring workflow was used to
make the proposed contract executable and renderable. All scientific data
checks run in R 4.6.1 with the synchronized project library.

This is an outcome-blind support and estimability audit. It reads the outcome
only to determine whether the approved one-hour value is available, finite, or
exactly zero; it does not summarize outcome magnitude by activity, fit a model,
or inspect an activity association. Candidate counts are not substitutes for
the exact fitted samples that Stage 2 must report.

## Preregistered contract

The signed AsPredicted record states exactly:

> **H4: Hourly self-reported activity categories predict hourly geometric mean melanopic EDI.**

— `preregistration/AsPredicted #273407.pdf`, PDF p. 1.

The signed model on PDF p. 2 is:

```text
melEDI ~ Predictor + (Predictor | Site) + (1 | Site:Participant)
```

For H04, `Predictor` is the self-reported activity category. The developed
analysis plan restates a relationship between hourly activities and hourly
geometric melEDI and records the possibility that the site random-slope route
may be unstable. It does not define how multiple simultaneous activities
should contribute. The owner-fixed contract in this task resolves that
construct and weighting question before any new H04 inference.

### Evidence identities

```{r}
#| label: tbl-evidence-identities
#| tbl-cap: "Controlling H04 evidence and prepared inputs, hashed at render time."

evidence_paths <- c(
  signed_preregistration = "preregistration/AsPredicted #273407.pdf",
  developed_analysis_plan = "preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx",
  preregistration_contract = "audit/evidence/preregistration_contract.md",
  reviewer_feedback = "audit/ledgers/reviewer_feedback_disposition.csv",
  migration_map = "audit/hypotheses/H01-H04_migration_map.md",
  category_support_gate = "audit/hypotheses/H03_H04_category_support_gate.md",
  gated_workflow = "audit/hypotheses/H03-H11_gated_workflow.qmd",
  h03_reader_report = "notebooks/hypotheses/H03.qmd",
  h03_preparation_report = "audit/hypotheses/H03/H03_analysis_preparation.qmd",
  h03_worker_handoff = "audit/handoffs/H03_worker_handoff.md",
  v0_near_eye_source = "RQ2.qmd",
  v0_chest_source = "RQ2_chest.qmd",
  v0_near_eye_result = "docs/RQ2.html",
  v0_chest_result = "docs/RQ2_chest.html",
  model_reporting = "audit/decisions/model_reporting.md",
  site_display = "audit/decisions/site_display_conventions.md",
  symlog_display = "audit/decisions/reader_facing_symlog_scale.md",
  paired_display = "audit/decisions/paired_placement_comparison_display.md",
  gap_terminology = "audit/decisions/gap_timing_unaware_dataset_terminology.md",
  bootstrap_policy = "audit/decisions/bootstrap_execution_policy.md",
  metric_registry = "config/metric_display_registry.csv",
  site_registry = "config/site_display_registry.csv",
  normalized_diary = input_paths[["normalized_diary"]],
  primary_near_eye = input_paths[["primary_near_eye"]],
  complementary_chest = input_paths[["complementary_chest"]],
  gap_timing_unaware = input_paths[["gap_timing_unaware"]]
)

stopifnot(all(file.exists(file.path(project_root, evidence_paths))))

evidence_manifest <- tibble(
  evidence = names(evidence_paths),
  path = unname(evidence_paths),
  sha256 = vapply(
    file.path(project_root, unname(evidence_paths)),
    digest::digest,
    character(1),
    algo = "sha256",
    file = TRUE
  )
)

evidence_manifest |>
  mutate(
    path = gsub("/", paste0("/", "\u200b"), path, fixed = TRUE),
    sha256 = gsub("(.{16})", paste0("\\1", "\u200b"), sha256, perl = TRUE)
  ) |>
  stage1_gt() |>
  cols_width(
    evidence ~ pct(24),
    path ~ pct(42),
    sha256 ~ pct(34)
  )
```

The current Preparation 06 H04 support files were also reconciled with the
normalized diary: activity flag support, flag-count distribution, activity
patterns, participant-hour cardinality, local-hour cardinality, site
cardinality, and option accounting. They are upstream evidence, not a
replacement for applying the owner-fixed collapse and specificity rule to the
current analytical inputs.

## What transfers from H03

```{r}
#| label: tbl-h03-transfer
#| tbl-cap: "H03 components audited for H04 transfer."

h03_transfer <- tribble(
  ~component, ~h04_decision, ~reason,
  "One-hour zero-aware geometric outcome", "Transfer",
  "Same approved metric construct, 30-valid-minute support rule, and exact-zero support.",
  "Population-mean quasi-Tweedie GLM", "Transfer, with H04 weights",
  "Avoids returning to the rejected singular mixed-model route; H04 adds prior weight 1/k.",
  "Participant-clustered HC1 inference", "Transfer",
  "A site-prefixed participant cluster contains same-hour replicated memberships and longitudinal rows.",
  "Finite-cluster t/F reference", "Transfer",
  "Use denominator df equal to participant clusters minus one for intervals and robust tests.",
  "Equal-site link-scale standardization", "Transfer",
  "Average fitted log means equally over observed sites, then back-transform once.",
  "Primary reference", "Change",
  "At home is H04's reader-facing reference; H03's indoor-electric reference is irrelevant.",
  "Primary omnibus restriction", "Change",
  "Five named categories are tested for equality while Other remains unrestricted.",
  "Site heterogeneity", "Transfer as a separate companion, redesign categories",
  "Use named categories only and a support-gated fallback; never require an Other interaction cell.",
  "Temporal GAMM", "Transfer mean structure; adapt dependence handling",
  "Use the same weighted long representation as the primary model and H03's sum-to-zero activity smooth; prevent concurrent rows from becoming false AR neighbours.",
  "Near-eye primary; chest complementary", "Transfer",
  "Also fit the exact paired/common sample when estimable.",
  "Diagnostics, sensitivities, provenance, and claim limits", "Transfer and extend",
  "Add fractional-weight, concurrent-row, and unique-hour diagnostics; retain observational wording."
)

stage1_gt(h03_transfer) |>
  cols_width(
    component ~ pct(24),
    h04_decision ~ pct(22),
    reason ~ pct(54)
  )
```

No H03 estimate, interval, p-value, diagnostic assessment, significance label,
or substantive conclusion is transferred.

## V0 reconstruction

V0 is the submitted implementation in `RQ2.qmd` and `RQ2_chest.qmd`, with its
frozen HTML, table, and figure outputs. It is a comparison target, not an
approved analysis.

### Exact V0 formulas — reconstructed, not fitted

```{r}
#| label: tbl-v0-formulas
#| tbl-cap: "Evaluated V0 Wilkinson formulas. No model is fitted."
#| code-fold: false

v0_formulas <- list(
  registered_structural = melEDI ~ activity +
    (activity | Site) + (1 | Site:Participant),
  v0_full = geo.MEDI ~ site * activity + (1 | Id),
  v0_site_only = geo.MEDI ~ site + (1 | Id),
  v0_additive = geo.MEDI ~ site + activity + (1 | Id),
  v0_activity_only = geo.MEDI ~ activity + (1 | Id)
)

v0_formula_contract <- enframe(
  v0_formulas,
  name = "model_id",
  value = "formula"
) |>
  mutate(formula = vapply(formula, formula_text, character(1)))

stage1_gt(v0_formula_contract) |>
  cols_width(model_id ~ pct(23), formula ~ pct(77))
```

V0 pivoted all eight activity flags to long rows, retained every selected
flag, deduplicated only before the outdoor collapse, collapsed outdoor work,
outdoor free time, and active travel after pivoting, removed `other`, and used
every remaining row with unit weight. Consequently, a multi-select hour
contributed more than once, and two outdoor source flags from the same hour
could remain as duplicate collapsed outdoor rows. The full V0 mixed model used
`glmmTMB`, a Tweedie log link, maximum likelihood, a participant random
intercept, fixed site, activity, and their interaction. Site was sum-coded in
three of the four fits.

```{r}
#| label: tbl-v0-denominator-audit
#| tbl-cap: "Frozen V0 H04 model rows and the different denominator printed in its footnote."

v0_denominator_audit <- tribble(
  ~placement, ~frozen_h04_model_rows, ~printed_participant_hours, ~status,
  "Near-eye", 16801L, 16774L,
  "Model rows are selection-level rows; the footnote reused the H03 denominator.",
  "Chest", 20327L, 18391L,
  "Model rows are selection-level rows; the footnote reused the H03 denominator."
)

stage1_gt(v0_denominator_audit) |>
  fmt_integer(columns = c(frozen_h04_model_rows, printed_participant_hours))
```

The stated confirmatory V0 comparison was site-only versus the full
site-by-activity model. It jointly added the activity block and all
activity-by-site interactions, so it did not isolate the overall activity
association. The V0 activity-by-site screen also overwrote `activity` with the
H03 light-source factor. Its activity contrasts used Dunnett-X while the table
described them as false-discovery-rate adjusted. The temporal block used a
different arithmetic outcome and inherited objects from another analysis.

The recognizable V0 presentation is a site-by-activity display paired with
the H03 site-by-light-source display. Stage 2 should preserve that recognizably
faceted comparison, but it must add 95% confidence intervals, correct metric
and site terminology, correct denominators, readable mathematical text, and
source data. V0 conclusions and bolding are not evidence for the repaired
analysis.

## Fixed activity representation

### Category dictionary

```{r}
#| label: tbl-activity-dictionary
#| tbl-cap: "Owner-fixed H04 activity dictionary and collapse."

activity_dictionary <- tribble(
  ~display_order, ~reader_label, ~source_flags, ~rule,
  1L, "Sleeping", "act_sleep", "Retain as named category.",
  2L, "At home", "act_home", "Retain; model and reader-facing comparison reference.",
  3L, "On the road with public transport/car", "act_road_vehicle", "Retain as named category.",
  4L, "Working in the office/from home", "act_working_indoor", "Retain as named category.",
  5L, "Outdoors", "act_road_open; act_working_outdoor; act_free_outdoor",
  "Combine the three flags before pivoting, then deduplicate the collapsed label.",
  6L, "Other/unspecified activity", "act_other",
  "Retain only when it is the sole selection; show quietly and make no scientific claim."
)

stage1_gt(activity_dictionary) |>
  cols_width(
    display_order ~ pct(8),
    reader_label ~ pct(24),
    source_flags ~ pct(30),
    rule ~ pct(38)
  )
```

No free text is inspected or recoded. A free-text taxonomy would require a
separate privacy-aware, outcome-blind upstream approval.

### Exact hour-to-long implementation

```{r}
#| label: h04-primary-transformation
#| code-fold: false

diary_primary <- h04_prepare_diary(
  raw = raw_inputs[["normalized_diary"]],
  root = project_root,
  retain_coselected_other = FALSE
)

weight_check <- diary_primary$long |>
  group_by(site, Id, local_date, clock_minute) |>
  summarise(
    retained_distinct_labels = n_distinct(activity_code),
    long_rows = n(),
    k = first(k),
    weight_sum = sum(activity_weight),
    .groups = "drop"
  )

stopifnot(
  all(weight_check$retained_distinct_labels == weight_check$long_rows),
  all(weight_check$long_rows == weight_check$k),
  all(abs(weight_check$weight_sum - 1) < 1e-12),
  !anyDuplicated(
    diary_primary$long[c(
      "site", "Id", "local_date", "clock_minute", "activity_code"
    )]
  )
)

h04_activity_registry() |>
  arrange(display_order) |>
  stage1_gt()
```

The evaluated helper performs these operations in order:

1. require one eligible normalized diary row per site, participant, local
   date, and clock hour;
2. combine the three outdoor flags into one Boolean before pivoting;
3. deduplicate collapsed labels within the participant-hour;
4. suppress `other` when at least one named category is selected, but retain it
   when it is the only selection;
5. pivot one row per retained distinct label;
6. set $k$ to the number of retained distinct labels and assign every long
   row prior analytic weight $1/k$; and
7. assert that every retained participant-hour has distinct labels and weights
   summing exactly to one.

Hours with no retained category are not interpreted as “no activity.” They
remain visible in sample flow and do not enter the primary category model.

```{r}
#| label: tbl-diary-transformation-flow
#| tbl-cap: "Eligible normalized diary hours after the fixed collapse and Other specificity rule."

diary_transformation_flow <- diary_primary$eligible |>
  count(k, activity_observation_status, name = "participant_hours") |>
  arrange(k, activity_observation_status)

stage1_gt(diary_transformation_flow) |>
  fmt_integer(columns = participant_hours)
```

The eligible diary contains 2,226 hours with all eight flags missing and 431
hours with flags observed but no category selected. They are distinct forms of
missing activity information. The collapse leaves at most three primary
labels per hour under the main Other rule.

::: {.callout-important title="The weighting defines the estimand"}
The primary estimand is the average melEDI associated with each reported
collapsed activity category, with a participant-hour's contribution shared
across its concurrently reported categories. Fractional long-row weighting is
therefore not a variance-only adjustment: it changes the estimand from V0's
unweighted **selection-level** analysis. The estimates are associations with
reported categories, not causal effects and not mutually exclusive
behavioural states.
:::

## Candidate support before fitting

### Placement sample flow

```{r}
#| label: prepare-placement-candidates
#| code-fold: false

near_eye <- h04_prepare_primary_placement(
  raw = raw_inputs[["primary_near_eye"]],
  diary = diary_primary,
  placement = "Near-eye",
  root = project_root
)

chest <- h04_prepare_primary_placement(
  raw = raw_inputs[["complementary_chest"]],
  diary = diary_primary,
  placement = "Chest",
  root = project_root
)

placement_bundles <- list(near_eye = near_eye, chest = chest)

candidate_sample_flow <- bind_rows(lapply(
  placement_bundles,
  h04_sample_flow
))

stopifnot(
  identical(candidate_sample_flow$model_candidate_unique_hours, c(16526L, 20128L)),
  identical(candidate_sample_flow$generated_long_rows, c(17266L, 21071L)),
  all(abs(
    candidate_sample_flow$model_candidate_unique_hours -
      candidate_sample_flow$effective_weighted_hours
  ) < 1e-10)
)
```

```{r}
#| label: tbl-candidate-sample-flow
#| tbl-cap: "All-available H04 candidate flow. Counts precede model-specific complete-case checks."

candidate_sample_flow |>
  rename_with(
    ~ gsub("_", " ", .x, fixed = TRUE)
  ) |>
  stage1_gt() |>
  fmt_integer(columns = where(is.numeric))
```

```{r}
#| label: tbl-no-activity-flow
#| tbl-cap: "Outcome-available, diary-matched hours excluded because no activity category is retained."

no_activity_flow <- map_dfr(placement_bundles, function(bundle) {
  bundle$joined |>
    filter(
      outcome_available,
      diary_key_matched,
      !retained_activity_available
    ) |>
    count(placement, activity_observation_status, name = "unique_hours")
})

stage1_gt(no_activity_flow) |>
  fmt_integer(columns = unique_hours)
```

Near eye has 1,801 candidate-outcome hours with all activity flags missing and
145 with observed flags but no selected category. Chest has 146 and 145,
respectively. None is recoded as a substantive activity.

### Distribution of retained labels per hour

```{r}
#| label: tbl-k-distribution
#| tbl-cap: "Distribution of k in each placement-specific candidate frame."

k_distribution <- bind_rows(lapply(
  placement_bundles,
  h04_k_distribution
))

stage1_gt(k_distribution) |>
  fmt_integer(columns = c(unique_hours, participants, participant_days)) |>
  fmt_number(columns = percent_of_candidate_hours, decimals = 2) |>
  cols_label(percent_of_candidate_hours = "Percent of candidate hours")
```

Thus 716 near-eye hours and 918 chest hours contribute to two or three
categories. Participant clustering must encompass both those same-hour rows
and repeated longitudinal hours; long rows must never be treated as
independent observations.

### Weighted totals by placement, site, and category

```{r}
#| label: tbl-weighted-placement-support
#| tbl-cap: "Placement totals after fractional weighting."

weighted_support <- lapply(placement_bundles, h04_weighted_support)

bind_rows(lapply(weighted_support, `[[`, "placement")) |>
  select(-summary_level) |>
  stage1_gt() |>
  fmt_integer(columns = c(unique_hours, long_rows, participants, participant_days)) |>
  fmt_number(columns = effective_weighted_hours, decimals = 1)
```

```{r}
#| label: tbl-weighted-site-support
#| tbl-cap: "Candidate support and fractional weighted totals by submitted-manuscript site."

weighted_site_support <- bind_rows(lapply(weighted_support, `[[`, "site")) |>
  mutate(site = as.character(site)) |>
  left_join(
    site_registry |>
      select(site, display_order, display_name),
    by = "site"
  ) |>
  arrange(placement, display_order) |>
  select(
    placement,
    site = display_name,
    participants,
    participant_days,
    unique_hours,
    long_rows,
    effective_weighted_hours
  )

stage1_gt(weighted_site_support, groupname_col = "placement") |>
  fmt_integer(columns = c(participants, participant_days, unique_hours, long_rows)) |>
  fmt_number(columns = effective_weighted_hours, decimals = 1)
```

```{r}
#| label: tbl-weighted-category-support
#| tbl-cap: "Candidate support and fractional weighted totals by collapsed activity category."

weighted_category_support <- bind_rows(lapply(
  weighted_support,
  `[[`,
  "category"
)) |>
  arrange(placement, display_order) |>
  transmute(
    placement,
    category = as.character(activity),
    participants,
    participant_days,
    sites,
    unique_hours,
    long_rows,
    effective_weighted_hours
  )

stage1_gt(weighted_category_support, groupname_col = "placement") |>
  fmt_integer(columns = c(participants, participant_days, sites, unique_hours, long_rows)) |>
  fmt_number(columns = effective_weighted_hours, decimals = 1)
```

All six categories have pooled support in both placements. Other/unspecified
activity is shown for accounting, but its heterogeneity and lack of a coherent
substantive meaning preclude a scientific claim.

### Sparsity, overlap, rank, and estimability

The site-specific support screen transfers H03's results-blind thresholds:
at least 20 unique hours, five participants, and five participants shared with
the At-home reference. These are reporting and architecture gates, not causal
positivity assumptions.

```{r}
#| label: tbl-unsupported-cells
#| tbl-cap: "Observed cells that fail at least one transferred site-specific support threshold."

cell_support <- bind_rows(lapply(
  placement_bundles,
  h04_cell_support
))

unsupported_cells <- cell_support |>
  filter(!support_rule) |>
  mutate(
    site = as.character(site),
    activity = as.character(activity)
  ) |>
  left_join(
    site_registry |>
      transmute(site, site_order = display_order, display_name),
    by = "site"
  ) |>
  arrange(placement, site_order, activity) |>
  select(
    placement,
    site = display_name,
    activity,
    unique_hours,
    effective_weighted_hours,
    participants,
    participant_days,
    shared_with_home_participants
  )

stage1_gt(unsupported_cells, groupname_col = "placement") |>
  fmt_integer(columns = c(
    unique_hours,
    participants,
    participant_days,
    shared_with_home_participants
  )) |>
  fmt_number(columns = effective_weighted_hours, decimals = 1)
```

KNUST × Other is observed, not structurally empty, but has only three
near-eye and four chest participants and fails the overlap rule. Near-eye
UCR × Other and UCR × Outdoors also fail. All other current cells pass the
declared thresholds.

```{r}
#| label: tbl-rank-audit
#| tbl-cap: "Exact candidate design ranks under sum-coded site and At-home-referenced activity."

rank_audit <- bind_rows(lapply(
  placement_bundles,
  h04_design_rank
))

heterogeneity_rank <- bind_rows(lapply(
  placement_bundles,
  h04_heterogeneity_rank
))

empty_cells <- bind_rows(lapply(
  placement_bundles,
  h04_empty_cells
))

stopifnot(
  all(rank_audit$full_rank),
  all(heterogeneity_rank$full_rank),
  nrow(empty_cells) == 0L,
  identical(
    rank_audit$test_df[rank_audit$design_id == "primary_full"],
    c(4L, 4L)
  ),
  identical(
    rank_audit$test_df[
      rank_audit$design_id == "secondary_six_category_null"
    ],
    c(5L, 5L)
  )
)

rank_audit |>
  select(
    placement,
    design_id,
    formula,
    rows,
    columns,
    rank,
    rank_deficiency,
    test_df,
    full_rank
  ) |>
  stage1_gt(groupname_col = "placement") |>
  fmt_integer(columns = c(rows, columns, rank, rank_deficiency, test_df))
```

```{r}
#| label: tbl-heterogeneity-rank
#| tbl-cap: "Rank-complete named-category site-heterogeneity designs."

stage1_gt(heterogeneity_rank, groupname_col = "placement") |>
  fmt_integer(columns = c(
    unique_hours,
    long_rows,
    columns,
    rank,
    rank_deficiency
  ))
```

The additive full design has rank 14/14 near eye and 13/13 chest. The primary
five-named-category restriction has four degrees of freedom; the secondary
six-category restriction has five. The five-named-category interaction is
rank 45/45 near eye and 40/40 chest. A four-category core fallback containing
At home, Sleeping, vehicle/public-transport travel, and indoor/home working is
rank 36/36 near eye and 32/32 chest; every one of those current site cells
passes support. Core fallback hours are retained only when **all** their
selected categories are core, so their fractional weights still sum to one.

Rank completeness is necessary but not sufficient. In particular, it does
not rehabilitate sparse Other cells. Other never enters the heterogeneity
companion.

## Proposed primary analysis

### Exact Wilkinson formulas — evaluated, not fitted

```{r}
#| label: tbl-proposed-formulas
#| tbl-cap: "Exact Wilkinson formulas proposed for H04 and every comparison model."
#| code-fold: false

proposed_formulas <- list(
  primary_full = geo_medi_1h ~ site + activity,
  primary_five_named_null = geo_medi_1h ~ site + other_indicator,
  secondary_six_category_null = geo_medi_1h ~ site,
  heterogeneity_five_named_full = geo_medi_1h ~ site * activity_named,
  heterogeneity_five_named_additive = geo_medi_1h ~ site + activity_named,
  heterogeneity_core_full = geo_medi_1h ~ site * activity_core,
  heterogeneity_core_additive = geo_medi_1h ~ site + activity_core,
  heterogeneity_observed_cell_contingency =
    geo_medi_1h ~ 0 + site_activity_named_cell,
  temporal_no_activity = stats::as.formula(paste0(
    "geo_medi_1h ~ s(time_hour, bs = 'cc', k = 12) + ",
    "s(time_hour, site, bs = 'sz', k = 12) + ",
    "s(time_hour, participant, bs = 'fs', k = 10) + ",
    "s(participant_day, bs = 're')"
  )),
  temporal_activity_long = stats::as.formula(paste0(
    "geo_medi_1h ~ s(time_hour, bs = 'cc', k = 12) + ",
    "s(time_hour, activity, bs = 'sz', k = 12) + ",
    "s(time_hour, site, bs = 'sz', k = 12) + ",
    "s(time_hour, participant, bs = 'fs', k = 10) + ",
    "s(participant_day, bs = 're')"
  ))
)

proposed_formula_contract <- enframe(
  proposed_formulas,
  name = "model_id",
  value = "formula"
) |>
  mutate(formula = vapply(formula, formula_text, character(1)))

stage1_gt(proposed_formula_contract) |>
  cols_width(model_id ~ pct(28), formula ~ pct(72))
```

```{r}
#| label: tbl-temporal-basis-contract
#| tbl-cap: "Constructor-verified basis contract for the proposed temporal smooths."
#| code-fold: false

temporal_basis_probe <- tibble(
  time_hour = rep(seq(0, 23, length.out = 24), 3),
  activity = factor(rep(c("A", "B", "C"), each = 24)),
  site = factor(rep(c("A", "B", "C"), each = 24))
)
temporal_knots <- list(time_hour = c(0, 24))

activity_sz_probe <- mgcv::smoothCon(
  mgcv::s(time_hour, activity, bs = "sz", k = 12),
  data = temporal_basis_probe,
  knots = temporal_knots,
  absorb.cons = FALSE
)[[1]]
site_sz_probe <- mgcv::smoothCon(
  mgcv::s(time_hour, site, bs = "sz", k = 12),
  data = temporal_basis_probe,
  knots = temporal_knots,
  absorb.cons = FALSE
)[[1]]
global_cc_probe <- mgcv::smoothCon(
  mgcv::s(time_hour, bs = "cc", k = 12),
  data = temporal_basis_probe,
  knots = temporal_knots,
  absorb.cons = FALSE
)[[1]]

stopifnot(
  identical(activity_sz_probe$base.bs, "tp.smooth.spec"),
  identical(activity_sz_probe$base$bs, "tprs.smooth"),
  identical(site_sz_probe$base.bs, "tp.smooth.spec"),
  identical(site_sz_probe$base$bs, "tprs.smooth"),
  inherits(global_cc_probe, "cyclic.smooth")
)

temporal_basis_contract <- tribble(
  ~component, ~exact_smooth, ~constructed_class, ~marginal_basis, ~midnight_constraint,
  "Global time", "s(time_hour, bs = 'cc', k = 12)",
  class(global_cc_probe)[1], "cyclic cubic regression spline", "Endpoint equality imposed",
  "Activity deviation", "s(time_hour, activity, bs = 'sz', k = 12)",
  class(activity_sz_probe)[1],
  paste(activity_sz_probe$base.bs, activity_sz_probe$base$bs, sep = " / "),
  "No endpoint equality",
  "Site deviation", "s(time_hour, site, bs = 'sz', k = 12)",
  class(site_sz_probe)[1],
  paste(site_sz_probe$base.bs, site_sz_probe$base$bs, sep = " / "),
  "No endpoint equality"
)

stage1_gt(temporal_basis_contract) |>
  cols_width(
    component ~ pct(15),
    exact_smooth ~ pct(30),
    constructed_class ~ pct(15),
    marginal_basis ~ pct(22),
    midnight_constraint ~ pct(18)
  )
```

This constructor-only probe fits no model and uses no research outcome. Under
R 4.6.1 and synchronized `mgcv` 1.9-4, both `sz` specifications resolve to
`tp.smooth.spec`/`tprs.smooth`; only the global `cc` term resolves to a cyclic
smooth.

Stage 2 must pass these formula objects unchanged to the fitting functions and
assert that each fitted object's formula matches the registry. The same
primary formulas apply to both placements and all non-temporal sensitivities;
only the exact frame or prior weights change.

### Mean model, weights, and clustered inference

The proposed primary fit is:

```r
stats::glm(
  formula = proposed_formulas$primary_full,
  data = model_frame,
  family = statmod::tweedie(var.power = 1.539919, link.power = 0),
  weights = activity_weight,
  control = stats::glm.control(epsilon = 1e-10, maxit = 100L),
  model = TRUE,
  x = TRUE,
  y = TRUE
)
```

This is a population-mean quasi-Tweedie GLM with log link. Site is fixed and
sum-coded; activity is treatment-coded with At home as the first level. The
working variance power 1.539919 is proposed to transfer unchanged from H03
because the outcome and measurement support are identical; it is not
re-estimated from H04 category results. The variance-power values 1.30 and
1.80 remain prespecified diagnostic robustness checks rather than competing
result-selected models.

The prior analytic weight is exactly `activity_weight = 1/k` for every main
long row. Robust covariance is:

```r
sandwich::vcovCL(
  fit,
  cluster = model_frame$participant,
  type = "HC1",
  cadjust = TRUE,
  fix = FALSE
)
```

`participant` is the site-prefixed participant identifier. Its cluster score
contains all concurrent membership rows from an hour and every repeated hour
from that participant. No row-level independence claim is made. Confidence
intervals use a $t$ reference and robust joint tests use an $F$ reference,
both with denominator degrees of freedom $G-1$, where $G$ is the number of
participant clusters in the exact fitted frame.

### Primary five-named-category test

The full six-category fit remains unchanged. The null formula
`geo_medi_1h ~ site + other_indicator` constrains Sleeping, At home, vehicle or
public-transport travel, working, and Outdoors to the same fitted mean while
allowing Other/unspecified activity its own coefficient. The robust test is
therefore not a likelihood-ratio comparison and does not test Other.

```{r}
#| label: tbl-primary-contrast
#| tbl-cap: "Exact four-row joint restriction for equality of the five named categories."
#| code-fold: false

primary_contrasts <- h04_contrast_registry(near_eye)
primary_restriction <- primary_contrasts$primary_five_named_equality
secondary_restriction <- primary_contrasts$secondary_six_category_equality

stopifnot(
  qr(primary_restriction)$rank == 4L,
  qr(secondary_restriction)$rank == 5L,
  all(primary_restriction[, grepl("Other", colnames(primary_restriction))] == 0)
)

primary_restriction |>
  as.data.frame() |>
  rownames_to_column("restriction") |>
  select(restriction, where(~ any(.x != 0))) |>
  stage1_gt()
```

If $\widehat{\boldsymbol{\beta}}$ is the full coefficient vector, $R$ is
the displayed four-row matrix, and $V_{\mathrm{HC1}}$ is the
participant-cluster covariance, Stage 2 computes

$$
F =
\frac{
  (R\widehat{\boldsymbol{\beta}})^\mathsf{T}
  (R V_{\mathrm{HC1}} R^\mathsf{T})^{-1}
  (R\widehat{\boldsymbol{\beta}})
}{4},
\qquad
F \sim F_{4,G-1}.
$$

The single raw p-value is the primary H04 decision (`H04-F1`, alpha 0.05). The
secondary full six-category omnibus adds the Other-versus-home row, uses
$F_{5,G-1}$, and is labelled secondary (`H04-F1b`). It cannot replace or
reinterpret the five-named-category test.

### Equal-site standardized category estimands

For every category $c$ and each site observed in that placement, Stage 2
constructs the exact design row $x_{sc}$, calculates the fitted log mean, and
uses

$$
\bar{\eta}_c =
\frac{1}{S}\sum_{s=1}^{S}
x_{sc}^\mathsf{T}\widehat{\boldsymbol{\beta}},
\qquad
\bar{\mu}_c = \exp\!\left(\bar{\eta}_c\right).
$$

There is one back-transformation after equal-site averaging. Thus
$\bar{\mu}_c$ is the equal-site geometric mean of site-specific fitted means,
not their response-scale arithmetic mean. Standardization uses the sites
actually observed for that placement; it does not invent an absent chest
site.

Reader-facing output uses the owner-fixed display order and reports, with 95%
confidence intervals:

- all six standardized category means in melEDI lx;
- four named-category ratios versus At home, with one complete four-member BH
  family (`H04-F2`), alongside raw p-values;
- the corresponding absolute differences in melEDI lx, as magnitude summaries
  using the same robust coefficient covariance; and
- Other/unspecified activity separately and quietly, without a scientific
  claim or inferential contrast family membership.

The reference is written **At home**, not “control” or a mutually exclusive
awake state. Sleeping describes the reported bedside sleep environment. Raw
and adjusted p-values follow REPORT-008: three decimals, `<0.001` below that
threshold, explicit labels, and bold only when the stated raw or adjusted rule
is met.

## Separate site heterogeneity

The primary additive mean is not replaced by a full interaction merely to
resemble V0. Heterogeneity is a separate exploratory companion on named
categories:

1. remove Other-only hours from the companion, leaving each retained hour's
   named-category weights summing to one;
2. attempt the rank-complete five-named-category full versus additive
   comparison shown in the formula registry;
3. before reading any interaction estimate or p-value, require convergence,
   full rank, finite positive restriction covariance, condition number at most
   $10^{10}$, maximum participant cluster-score share at most 0.50, maximum
   cluster-leverage share at most 0.20, and finite intervals for supported
   site/category cells;
4. suppress every site-specific contrast that fails the 20-hour, five-person,
   five-shared-person support rule; and
5. if the full five-category gate fails, use the predeclared four-category
   core-only fallback. If that also fails, classify heterogeneity as
   non-estimable rather than forcing an interaction.

The five-category companion includes Outdoors but never Other, so KNUST ×
Other is not a parameter or claim target. The core fallback also removes hours
with any selected non-core label before fitting; it does not leave partial
hour weights behind. If a future sensitivity produces a structurally empty
named cell, the observed-cell formula in the registry replaces the literal
interaction and its robust additivity restriction is built from only observed
cells.

`H04-F3` is one raw robust heterogeneity omnibus. `H04-F4` contains all and
only support-eligible site deviations under the architecture selected by the
results-blind gate and applies one vector-wide BH adjustment. No unsupported
cell is silently imputed, pooled, or highlighted.

## Exploratory temporal context

Following author review, the temporal proposal now uses the **same long rows,
activity factor, and exact `activity_weight = 1/k` prior weights as the
primary model**. It also restores the accepted H03 mean structure: a global
cyclic time smooth plus the sum-to-zero factor smooths
`s(time_hour, activity, bs = "sz", k = 12)` and
`s(time_hour, site, bs = "sz", k = 12)`. With `xt` omitted, both `sz`
smooths use the H03-approved default thin-plate marginal basis in the
synchronized `mgcv` version. This gives activity- and site-associated
deviations from the global daily pattern while preserving the full six-category
fit. Other remains in the fitted activity smooth but is displayed quietly and
is not a scientific claim target.

Only the global time smooth is cyclic. The thin-plate activity and site
deviations may therefore differ between 0:00 and 24:00, so the resulting
category- and site-specific curves may separate across midnight. Stage 2 must
quantify and display that endpoint separation and must not draw a wrap-around
segment implying continuity. This is an accepted consequence of the approved
thin-plate `sz` basis, not a reason to substitute the previously unapproved
cyclic marginal basis.

Fractional weighting keeps the temporal mean estimator aligned with the
primary representation: a multi-select participant-hour appears once for each
retained distinct activity, but its long-row weights sum exactly to one. The
no-activity comparator uses the identical long frame, weights, and dependence
boundaries. No multi-membership covariates or widened one-row approximation are
used.

The long representation requires one H04-specific dependence adaptation. Rows
from concurrent activities have the same timestamp and outcome, so they must
not be placed next to one another in a single participant-day AR stream.
Instead, Stage 2 constructs **activity-specific temporal runs** within each
participant-day. Within a run, each clock hour is unique. A new run begins at
an activity boundary, participant/day boundary, daylight-saving repeated hour,
unsupported hour, non-consecutive clock hour, or any other recorded gap. Thus
AR(1) correlation is applied only between consecutive hours for the same
reported activity; concurrent activity rows are never treated as lag-one
neighbours.

Both temporal formulas use `mgcv::bam`, raw `geo_medi_1h`,
`mgcv::Tweedie(p = 1.539919, link = "log")`, the exact fractional prior
weights, fREML, `discrete = TRUE`, one thread, and knots
`time_hour = c(0, 24)`. The H03 participant factor smooth, participant-day
random effect, thin-plate activity and site sum-to-zero smooths, and cyclic
global smooth are retained. Population curves exclude participant and
participant-day smooths and use equal-site standardization.

The activity-specific AR runs solve the zero-lag adjacency problem but do not,
by themselves, claim that concurrent rows are independent. If the temporal
model passes Stage 2 diagnostics and is retained for durable reporting,
uncertainty bands must use a site-stratified participant-cluster bootstrap that
resamples complete participant histories, thereby preserving concurrent rows
and longitudinal dependence together. Stage 2 may inspect only a bounded
50- or 100-replicate pilot and must then report pilot output and estimated full
runtime at the separate production-computation gate. Model-based `bam`
intervals may be used for construction diagnostics but not for a reader-facing
temporal claim.

The temporal analysis remains exploratory context, not a causal activity
effect. It is retained only if it genuinely contextualizes time of day and
passes fractional-weight, run-boundary, global-smooth cyclicity, thin-plate
basis, residual autocorrelation, local-support, and participant-level stability
checks.
Approximate smooth p-values do not become the primary H04 test, and a
pointwise band does not justify a claimed significant period.

## Planned diagnostics and acceptability decisions

```{r}
#| label: tbl-diagnostic-plan
#| tbl-cap: "Mandatory Stage 2 diagnostics for the weighted duplicated-membership representation."

diagnostic_plan <- tribble(
  ~diagnostic, ~unit, ~acceptable_rule,
  "Fractional-weight integrity", "Unique participant-hour",
  "Distinct labels; row count equals k; finite positive 1/k weights; weight sum exactly one.",
  "Model frame and rank", "Long row and coefficient design",
  "No unintended complete-case loss; full design and restriction rank; exact participants, days, hours, rows, and weighted hours reconciled.",
  "IRLS and coefficient covariance", "Fit and participant cluster",
  "Converged finite coefficients; finite HC1 covariance; tested restriction covariance positive definite with condition number <= 1e10.",
  "Cluster influence", "Site-prefixed participant",
  "Report cluster score, leverage, and leave-one-participant refits; architecture gate requires score share <= 0.50 and leverage share <= 0.20.",
  "Mean-variance and residual structure", "Weighted long row and unique hour",
  "Inspect Pearson/deviance residuals versus fitted mean and support; aggregate weighted residual summaries by unique hour so duplicate memberships are not counted as independent.",
  "Category and site calibration", "Weighted category/site bins",
  "Observed versus fitted weighted means with 95% uncertainty; no material systematic pattern hidden by sparse cells.",
  "Exact zeros and positive tail", "Unique participant-hour",
  "Report zero prevalence and positive-tail behavior by placement, category, site, and fitted-mean bin; state that a quasi mean model does not model zero probability directly.",
  "Dependence", "Unique participant-hour sequence",
  "Report residual ACF within valid participant-day runs; clustered inference remains primary even if serial correlation is visible.",
  "Working variance power", "Identical fitted frame",
  "Compare p = 1.539919 with fixed 1.30 and 1.80 diagnostic fits; do not select power by H04 significance.",
  "Temporal adequacy", "Fractionally weighted activity-long frame",
  "Check activity and site `sz` k-index/EDF, global-smooth cyclic endpoint agreement, quantified 0:00-versus-24:00 separation for the thin-plate deviations without requiring equality, activity-specific run uniqueness and boundaries, concurrent-row preservation, residual ACF after AR(1), local clock/category support, and participant-bootstrap pilot stability if the model is retained."
)

stage1_gt(diagnostic_plan) |>
  cols_width(
    diagnostic ~ pct(22),
    unit ~ pct(22),
    acceptable_rule ~ pct(56)
  )
```

Every diagnostic gets a plain-language interpretation and an explicit
`acceptable`, `acceptable with limitation`, or `not acceptable` assessment.
An unacceptable primary model reopens the Stage 2 gate; it is not hidden by a
sensitivity or a visually familiar figure.

## Prespecified sensitivities and placement evidence

```{r}
#| label: sensitivity-support
#| include: false

scenario_support <- function(
  bundle,
  scenario,
  row_filter = identity,
  unweighted = FALSE
) {
  data <- row_filter(bundle$long)
  tibble(
    scenario = scenario,
    placement = unique(bundle$long$placement),
    participants = n_distinct(data$participant),
    participant_days = n_distinct(data$participant_day),
    sites = n_distinct(data$site),
    unique_hours = n_distinct(data$hour_id),
    long_rows = nrow(data),
    effective_weighted_hours = if (unweighted) {
      nrow(data)
    } else {
      sum(data$activity_weight)
    }
  )
}

diary_other_retained <- h04_prepare_diary(
  raw = raw_inputs[["normalized_diary"]],
  root = project_root,
  retain_coselected_other = TRUE
)

near_eye_other_retained <- h04_prepare_primary_placement(
  raw = raw_inputs[["primary_near_eye"]],
  diary = diary_other_retained,
  placement = "Near-eye",
  root = project_root
)

chest_other_retained <- h04_prepare_primary_placement(
  raw = raw_inputs[["complementary_chest"]],
  diary = diary_other_retained,
  placement = "Chest",
  root = project_root
)

sensitivity_candidate_support <- bind_rows(
  scenario_support(near_eye, "Primary fractional representation"),
  scenario_support(chest, "Primary fractional representation"),
  scenario_support(
    near_eye,
    "Exactly one retained category",
    row_filter = function(x) filter(x, k == 1L)
  ),
  scenario_support(
    chest,
    "Exactly one retained category",
    row_filter = function(x) filter(x, k == 1L)
  ),
  scenario_support(
    near_eye_other_retained,
    "Retain co-selected Other"
  ),
  scenario_support(
    chest_other_retained,
    "Retain co-selected Other"
  ),
  scenario_support(
    near_eye,
    "Exclude Other-only hours",
    row_filter = function(x) filter(x, activity_code != "other")
  ),
  scenario_support(
    chest,
    "Exclude Other-only hours",
    row_filter = function(x) filter(x, activity_code != "other")
  ),
  scenario_support(
    near_eye,
    "Unweighted long rows",
    unweighted = TRUE
  ),
  scenario_support(
    chest,
    "Unweighted long rows",
    unweighted = TRUE
  )
)

key <- h04_key_columns()
common_keys <- inner_join(
  near_eye$long |>
    distinct(across(all_of(key))),
  chest$long |>
    distinct(across(all_of(key))),
  by = key
)

paired_support <- map_dfr(placement_bundles, function(bundle) {
  data <- semi_join(bundle$long, common_keys, by = key)
  tibble(
    placement = unique(data$placement),
    participants = n_distinct(data$participant),
    participant_days = n_distinct(data$participant_day),
    sites = n_distinct(data$site),
    unique_hours = n_distinct(data$hour_id),
    long_rows = nrow(data),
    effective_weighted_hours = sum(data$activity_weight)
  )
})

gap_near_eye <- h04_prepare_gap_placement(
  raw = raw_inputs[["gap_timing_unaware"]],
  diary = diary_primary,
  position = "glasses",
  placement = "Near-eye",
  root = project_root
)

gap_chest <- h04_prepare_gap_placement(
  raw = raw_inputs[["gap_timing_unaware"]],
  diary = diary_primary,
  position = "chest",
  placement = "Chest",
  root = project_root
)

gap_support <- bind_rows(
  h04_gap_sample_flow(gap_near_eye),
  h04_gap_sample_flow(gap_chest)
)
```

```{r}
#| label: tbl-sensitivity-support
#| tbl-cap: "Candidate support for the fixed representation sensitivities. No model is fitted."

stage1_gt(sensitivity_candidate_support, groupname_col = "scenario") |>
  fmt_integer(columns = c(
    participants,
    participant_days,
    sites,
    unique_hours,
    long_rows
  )) |>
  fmt_number(columns = effective_weighted_hours, decimals = 1)
```

```{r}
#| label: tbl-paired-support
#| tbl-cap: "Exact near-eye/chest common candidate keys before model fitting."

stopifnot(
  identical(paired_support$unique_hours, c(14308L, 14308L)),
  identical(paired_support$long_rows, c(15001L, 15001L)),
  all(abs(paired_support$effective_weighted_hours - 14308) < 1e-10)
)

stage1_gt(paired_support) |>
  fmt_integer(columns = c(
    participants,
    participant_days,
    sites,
    unique_hours,
    long_rows
  )) |>
  fmt_number(columns = effective_weighted_hours, decimals = 1)
```

```{r}
#| label: tbl-gap-support
#| tbl-cap: "Candidate support in the gap-timing-unaware dataset."

stage1_gt(gap_support) |>
  fmt_integer(columns = where(is.numeric))
```

The required sensitivity interpretations are fixed before fitting:

- **Exactly one retained category:** restrict after the collapse and Other
  rule to `k == 1`; this includes Other-only hours but removes all concurrent
  named memberships.
- **Retain co-selected Other:** include Other alongside named categories and
  recompute $k$ and all `1/k` weights. This adds 67 near-eye and 119 chest
  long rows without changing unique hours.
- **Exclude Other-only:** remove the 391 near-eye and 631 chest Other-only
  candidate hours; the remaining long-row weights still sum to one per hour.
- **Unweighted long rows:** set every long-row weight to one. Its effective
  totals become 17,266 near eye and 21,071 chest; this deliberately recovers a
  selection-level estimand and is not a robustness-equivalent reweighting.
- **Gap-timing-unaware dataset:** use the same transformation and formulas on
  the manuscript-prepared hourly data. At first reader-facing use, explain
  that it passed the same coverage threshold but did not preserve the timing
  of unsupported within-hour gaps.
- **Placement-matched/common sample:** fit near-eye and chest models
  separately on the same 14,308 participant-hour keys, 110 participants, 625
  participant-days, eight sites, 15,001 long rows, and 14,308 weighted hours.
  This does not pool placements or treat them as independent additions.
- **Chest-level evidence:** fit the complementary chest analysis on all
  eligible chest data and on the exact common sample. It is complementary
  measurement-location evidence, not a second confirmatory test.

Stage 2 must compare direction, magnitude, 95% intervals, diagnostics, and the
predeclared decision across these scenarios. If a contrast is non-estimable,
the reason remains visible rather than being replaced by a nearby scenario.

## Required Stage 2 reporting and V0 comparison

Stage 2 must reproduce the recognizable H04 result structure with corrected
data and inference. It must produce:

- exact formula, family, link, working power, prior-weight, cluster, and
  finite-df registries;
- exact fitted model frames and sample flow with participants,
  participant-days, unique participant-hours, long rows, effective weighted
  hours, sites, categories, and category-by-site support;
- six equal-site standardized means, four named-versus-home ratios and
  differences, all 95% confidence intervals, and explicitly labelled raw and
  adjusted p-values;
- the primary five-named robust omnibus, secondary six-category omnibus, and
  separately gated heterogeneity test;
- primary near-eye, complementary chest, exact paired/common-sample, and
  gap-timing-unaware evidence;
- every mandatory diagnostic with plain-language interpretation and an
  acceptability decision;
- H04-specific source-data CSVs for every durable table and figure; and
- a result-difference table against V0 that classifies denominator, estimate,
  interval, p-value, adjusted decision, diagnostic, and claim changes.

The result-difference audit must explicitly compare:

```{r}
#| label: tbl-v0-comparison-requirements
#| tbl-cap: "Required H04 V0-to-repaired comparison fields."

v0_comparison_requirements <- tribble(
  ~comparison, ~v0_state, ~repaired_state,
  "Analytical unit", "Unweighted selected activity row", "Participant-hour contribution shared by 1/k",
  "Categories", "Five levels; Other removed", "Six-level fit; five interpretable categories in primary test; Other unrestricted",
  "Outdoor handling", "Collapsed after pivot; duplicate collapsed labels possible", "Collapsed and deduplicated before pivot",
  "Near-eye denominator", "16,801 model rows; 16,774 printed", "Exact participants, days, unique hours, long rows, and weighted hours",
  "Chest denominator", "20,327 model rows; 18,391 printed", "Exact participants, days, unique hours, long rows, and weighted hours",
  "Primary test", "Site-only versus activity-by-site mixed model", "Four-restriction participant-cluster robust Wald F",
  "Other", "Removed", "Quiet fitted category, unrestricted under primary null; no claim",
  "Heterogeneity", "Included in confirmatory comparison", "Separate named-category support-gated robust companion",
  "Uncertainty", "Asymptotic mixed-model output; CIs missing from reader table", "Participant-cluster HC1 with finite-cluster t/F and 95% CIs",
  "Multiplicity", "Dunnett-X labelled FDR; hidden by-groups", "Declared complete vectors with exact raw/BH labels",
  "Activity-by-site screen", "Actually used H03 light-source factor", "Correct H04 weighted descriptives and support only",
  "Temporal outcome/representation", "Arithmetic outcome and inherited single-category objects", "Fractionally weighted long-row geometric outcome with thin-plate activity and site `sz` smooths and activity-specific AR runs",
  "Claims", "Causal-sounding and significance-led", "Observational association, magnitude, uncertainty, and limitations"
)

stage1_gt(v0_comparison_requirements) |>
  cols_width(
    comparison ~ pct(20),
    v0_state ~ pct(38),
    repaired_state ~ pct(42)
  )
```

Reader-facing figures use submitted site names, order, and colours. Any
zero-containing log-normal-like melEDI display uses
`LightLogR::symlog_trans(base = 10, thr = 1, scale = 1)` while source data
retain untransformed melEDI lx. Figure QA checks clipping, wrapping, panel
balance, final physical size, and publication-scale legibility. The Stage 3
paired near-eye/chest display uses separately fitted estimates on identical
common keys only when the estimands are meaningfully aligned.

No temperature or unsupported predictor is added. A participant-cluster
bootstrap is proposed only if the revised temporal analysis passes diagnostics
and is retained for durable reporting. Stage 2 may run only a 50- or
100-replicate pilot, show its output and estimated full runtime, and stop at the
unnumbered production-computation gate before any full run. No other bootstrap
or simulation is proposed.

## Audit findings and author decisions

```{r}
#| label: tbl-stage1-findings
#| tbl-cap: "H04 Stage 1 findings and proposed disposition."

stage1_findings <- tribble(
  ~finding_id, ~severity, ~status, ~finding, ~proposed_disposition,
  "H04-001", "Major", "Confirmed V0 defect",
  "Multi-select hours received multiple unit-weight contributions.",
  "Use the fixed collapse, deduplication, Other rule, and 1/k prior weights.",
  "H04-002", "Major", "Confirmed V0 defect",
  "The V0 primary comparison combined activity association and site heterogeneity.",
  "Use a four-restriction primary robust F test and a separate heterogeneity companion.",
  "H04-003", "Major", "Confirmed V0 defect",
  "V0 removed Other and could duplicate collapsed outdoor labels.",
  "Retain Other only under the fixed specificity rule; collapse and deduplicate before pivot.",
  "H04-004", "Major", "Confirmed V0 defect",
  "V0 activity-by-site screening used the light-source factor.",
  "Replace with H04-specific weighted support/descriptives; do not use the faulty screen.",
  "H04-005", "Major", "Confirmed reporting defect",
  "V0 printed H03 denominators rather than H04 model rows and called selection rows participant-hours.",
  "Report exact participants, days, unique hours, long rows, and weighted hours for each fit.",
  "H04-006", "Major", "Confirmed inferential defect",
  "Dunnett-X contrasts were labelled as FDR-adjusted.",
  "Use complete declared BH vectors and retain V0 values only in the difference audit.",
  "H04-007", "Moderate", "Current support fact",
  "Other is pooled-estimable but support-inadequate at UCR and KNUST; near-eye UCR Outdoors is also sparse.",
  "Keep Other in the additive fit, exclude it from heterogeneity, and suppress unsupported site contrasts.",
  "H04-008", "Moderate", "Current design fact",
  "Primary and named heterogeneity designs are full rank in both placements.",
  "Proceed only after approval and retain the results-blind covariance/influence gate.",
  "H04-009", "Major", "Temporal design defect",
  "A naive long-row AR sequence would treat concurrent activity memberships with identical timestamps and outcomes as serial neighbours.",
  "Use exact 1/k weighted long rows, H03-style thin-plate activity and site `sz` smooths, activity-specific AR runs, and participant-cluster bootstrap uncertainty if the temporal model is retained.",
  "H04-010", "Moderate", "Open model decision",
  "The registered singular mixed-model route is not supported merely for symmetry with V0.",
  "Transfer H03's population-mean quasi-Tweedie architecture with H04-specific weights and restrictions."
)

stage1_gt(stage1_findings) |>
  cols_width(
    finding_id ~ pct(9),
    severity ~ pct(9),
    status ~ pct(16),
    finding ~ pct(31),
    proposed_disposition ~ pct(35)
  )
```

The owner-fixed category and weighting contract is not reopened. The owner
approved H04-G1 through H04-G8 and authorized Stage 2 on 2026-08-10:

```{r}
#| label: tbl-author-decisions
#| tbl-cap: "Decisions requested at the H04 Stage 1 gate."

author_decisions <- tribble(
  ~gate_id, ~decision_requested, ~status,
  "H04-G1", "Approve the additive population-mean quasi-Tweedie GLM with working power 1.539919, log link, exact 1/k prior weights, sum-coded fixed site, and participant-clustered HC1 covariance.", "APPROVED 2026-08-10",
  "H04-G2", "Approve the four-restriction robust F test for equality of the five named categories while Other remains unrestricted, plus the secondary five-restriction six-category omnibus.", "APPROVED 2026-08-10",
  "H04-G3", "Approve At home as the comparison reference and equal-site log-mean standardization with means, ratios, differences, and 95% confidence intervals.", "APPROVED 2026-08-10",
  "H04-G4", "Approve one four-member BH family for named-versus-home ratios, one raw heterogeneity omnibus, and one complete BH family of support-eligible site deviations.", "APPROVED 2026-08-10",
  "H04-G5", "Approve the separate five-named-category site-heterogeneity gate, the core-only four-category fallback, and non-estimable classification if both architectures fail.", "APPROVED 2026-08-10",
  "H04-G6", "Approve the fractionally weighted long-row H03-style temporal specification with thin-plate activity and site `sz` smooths, an exclusively cyclic global time smooth, accepted midnight separation, activity-specific AR runs, and participant-cluster bootstrap uncertainty if retained, as exploratory context only.", "APPROVED 2026-08-10",
  "H04-G7", "Approve the exactly-one, co-selected-Other, Other-only exclusion, unweighted-long, gap-timing-unaware, paired/common, and chest sensitivity battery exactly as defined above.", "APPROVED 2026-08-10",
  "H04-G8", "Approve the V0 comparison and reporting contract, including corrected denominators, source data, REPORT-008 p-values, symlog display, and final-size figure QA.", "APPROVED 2026-08-10"
)

stage1_gt(author_decisions) |>
  cols_width(
    gate_id ~ pct(9),
    decision_requested ~ pct(67),
    status ~ pct(24)
  )
```

## Reproducibility and Stage 1 stop

```{r}
#| label: tbl-session
#| tbl-cap: "R environment used for this Stage 1 audit."

package_names <- c(
  "dplyr", "tidyr", "purrr", "tibble", "gt", "digest",
  "statmod", "sandwich", "mgcv"
)

session_contract <- tibble(
  component = c("R", package_names),
  version = c(
    as.character(getRversion()),
    vapply(
      package_names,
      function(package) as.character(utils::packageVersion(package)),
      character(1)
    )
  )
)

stage1_gt(session_contract)
```

```{r}
#| label: stage1-assertions
#| include: false

stage1_objects <- mget(ls(envir = knitr::knit_global()), envir = knitr::knit_global())
inferential_classes <- c(
  "glm", "lm", "gam", "bam", "glmmTMB", "merMod", "lmerMod", "glmerMod"
)
is_inferential_fit <- function(object) {
  any(vapply(
    inferential_classes,
    function(class_name) inherits(object, class_name),
    logical(1)
  ))
}

stopifnot(
  !any(vapply(stage1_objects, is_inferential_fit, logical(1))),
  all(abs(weight_check$weight_sum - 1) < 1e-12),
  all(rank_audit$full_rank),
  all(heterogeneity_rank$full_rank),
  nrow(empty_cells) == 0L,
  all(candidate_sample_flow$exact_zero_unique_hours > 0L),
  all(abs(
    paired_support$effective_weighted_hours - paired_support$unique_hours
  ) < 1e-10),
  identical(as.character(getRversion()), "4.6.1")
)

stage1_assertion_record <- tribble(
  ~assertion, ~status,
  "No inferential fit object exists in the render", "PASS",
  "Every retained primary hour has distinct labels and weights summing to one", "PASS",
  "Primary and comparison designs are full rank", "PASS",
  "Five-named and core heterogeneity designs are full rank", "PASS",
  "No current site-by-category cell is structurally empty", "PASS",
  "Paired placement keys and weighted totals match exactly", "PASS",
  "Temporal activity/site `sz` marginals are thin plate and only global time is cyclic", "PASS",
  "Scientific audit ran in R 4.6.1", "PASS"
)
```

```{r}
#| label: tbl-stage1-assertions
#| tbl-cap: "Render-time H04 Stage 1 assertions."

stage1_gt(stage1_assertion_record)
```

::: {.callout-note title="Stage 1 approval recorded"}
The full H04 Stage 1 contract, including revised H04-G6, was approved on
2026-08-10. No new H04 inferential model was fit in Stage 1. Stage 2 was
explicitly authorized in this H04 task.
:::
