---
title: "H04 Stage 2: implementation, diagnostics, and V0 comparison"
subtitle: "Nature Health analysis audit — author result and Stage 2 gate"
date: "2026-08-11"
page-layout: full
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
editor: source
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 140,
  dplyr.summarise.inform = FALSE
)

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H04_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- locate_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

stopifnot(identical(as.character(getRversion()), "4.6.1"))
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
stopifnot(dir.exists(project_library))
.libPaths(c(project_library, .libPaths()))

library(dplyr)
library(tidyr)
library(readr)
library(tibble)
library(gt)

source(file.path(project_root, "scripts/pipeline/p_value_display.R"))

artifact <- function(...) file.path(project_root, "artifacts", ...)
read_h04 <- function(...) {
  readr::read_csv(artifact(...), show_col_types = FALSE, na = "")
}

approvals <- read_h04("06_model_data", "H04", "H04_author_approvals.csv")
input_audit <- read_h04("06_model_data", "H04", "H04_input_audit.csv")
formulas <- read_h04("06_model_data", "H04", "H04_formula_registry.csv")
samples <- read_h04("06_model_data", "H04", "H04_model_frame_index.csv")
sample_flow <- read_h04("06_model_data", "H04", "H04_sample_flow.csv")
k_distribution <- read_h04("06_model_data", "H04", "H04_k_distribution.csv")
weighted_support <- read_h04(
  "06_model_data", "H04", "H04_weighted_support_totals.csv"
)
category_support <- read_h04(
  "06_model_data", "H04", "H04_category_support.csv"
)
site_support <- read_h04(
  "06_model_data", "H04", "H04_site_category_support.csv"
)
primary <- read_h04("09_tables", "H04", "H04_primary_category_estimands.csv")
all_categories <- read_h04("09_tables", "H04", "H04_all_category_estimands.csv")
tests <- read_h04(
  "09_tables", "H04", "H04_primary_and_heterogeneity_tests.csv"
)
site_activity <- read_h04(
  "09_tables", "H04", "H04_site_activity_estimands.csv"
)
paired <- read_h04("09_tables", "H04", "H04_paired_placement_estimands.csv")
sensitivity <- read_h04(
  "09_tables", "H04", "H04_sensitivity_comparison.csv"
)
sensitivity_tests <- read_h04(
  "09_tables", "H04", "H04_sensitivity_omnibus_tests.csv"
)
diagnostics <- read_h04("08_diagnostics", "H04", "H04_model_diagnostics.csv")
assessments <- read_h04(
  "08_diagnostics", "H04", "H04_diagnostic_assessments.csv"
)
influence <- read_h04(
  "09_tables", "H04", "H04_influence_category_refits.csv"
)
influence_registry <- read_h04(
  "08_diagnostics", "H04", "H04_influence_refit_registry.csv"
)
heterogeneity_gate <- read_h04(
  "08_diagnostics", "H04", "H04_heterogeneity_architecture_gate.csv"
)
v0_models <- read_h04("09_tables", "H04", "H04_v0_bridge_models.csv")
v0_tests <- read_h04("09_tables", "H04", "H04_v0_bridge_tests.csv")
v0_ratios <- read_h04("09_tables", "H04", "H04_v0_bridge_ratios.csv")
v0_comparison <- read_h04(
  "09_tables", "H04", "H04_v0_denominator_and_test_comparison.csv"
)
temporal_summary <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_model_summary.csv"
)
temporal_retention <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_retention_decision.csv"
)
temporal_basis <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_basis_contract.csv"
)
temporal_run <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_run_diagnostics.csv"
)
temporal_k <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_k_check.csv"
)
temporal_acf <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_residual_acf.csv"
)
temporal_activity_midnight <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_activity_midnight_diagnostics.csv"
)
temporal_site_midnight <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_site_midnight_diagnostics.csv"
)
temporal_uncertainty <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_uncertainty_contract.csv"
)
bootstrap_supersession <- read_h04(
  "08_diagnostics", "H04", "H04_temporal_bootstrap_supersession.csv"
)
software <- read_h04("09_tables", "H04", "H04_software_environment.csv")

format_p_cell <- function(value, significant = FALSE) {
  ifelse(
    is.na(value),
    "—",
    {
      display <- nh_p_value_display(value, significant = significant)
      ifelse(
        display$p_bold,
        paste0("**", display$p_display, "**"),
        display$p_display
      )
    }
  )
}
format_mean_ci <- function(estimate, low, high) {
  sprintf("%.1f (%.1f to %.1f)", estimate, low, high)
}
format_ratio_ci <- function(estimate, low, high) {
  sprintf("%.3f (%.3f to %.3f)", estimate, low, high)
}
format_difference_ci <- function(estimate, low, high) {
  sprintf("%.1f (%.1f to %.1f)", estimate, low, high)
}
audit_gt <- function(data, title, subtitle = NULL, groupname_col = NULL) {
  gt::gt(data, groupname_col = groupname_col) |>
    gt::tab_header(title = title, subtitle = subtitle) |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(4),
      heading.align = "left",
      table.width = gt::pct(100)
    )
}

main_samples <- samples |>
  filter(.data$scenario_id == "primary_dataset") |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
main_flow <- sample_flow |>
  filter(!is.na(.data$outcome_grid_hours)) |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement)
named_sensitivity <- sensitivity |>
  filter(.data$activity_code %in% c(
    "sleeping", "road_vehicle", "working_indoor", "outdoors"
  ))
overall_assessment <- assessments |>
  filter(.data$diagnostic == "Overall primary mean-model assessment")
primary_tests <- tests |>
  filter(.data$test_id %in% c("H04-F1", "H04-F1b"))
heterogeneity_tests <- tests |>
  filter(.data$test_id == "H04-F3")

stopifnot(
  nrow(approvals) == 8L,
  all(approvals$status == "APPROVED"),
  all(input_audit$hash_verified),
  nrow(main_samples) == 2L,
  nrow(primary) == 12L,
  nrow(primary_tests) == 4L,
  nrow(heterogeneity_tests) == 2L,
  all(primary_tests$status == "ESTIMABLE"),
  all(heterogeneity_gate$gate_pass),
  all(overall_assessment$assessment == "ACCEPTABLE WITH LIMITATION"),
  all(named_sensitivity$stability == "stable"),
  all(sensitivity_tests$status == "ESTIMABLE"),
  all(temporal_retention$retained_for_context),
  all(temporal_run$concurrent_rows_never_lag_neighbours),
  all(temporal_run$hours_with_row_k_mismatch == 0L),
  all(temporal_run$maximum_hour_weight_error < 1e-10),
  all(temporal_basis$cyclic == c(TRUE, FALSE, FALSE)),
  all(temporal_uncertainty$interval_type ==
    "pointwise_95_percent_conditional"),
  all(temporal_uncertainty$resampling_replicates == 0L),
  all(!temporal_uncertainty$simultaneous_band),
  all(!temporal_uncertainty$curve_wide_inference),
  nrow(bootstrap_supersession) == 2L
)
```

## Scope and Stage 2 answer

The preregistered hypothesis is:

> **H4: Hourly self-reported activity categories predict hourly geometric mean melanopic EDI.**

This report implements the author-approved repair. It does not treat the
activity labels as mutually exclusive behavioural states and does not estimate
causal effects. Each estimate is the average melEDI associated with a reported
collapsed activity category after sharing a multi-select participant-hour's
contribution equally across its retained categories.

::: {.callout-important title="Stage 2 result gate"}
The primary near-eye robust equality test rejects equality of the five
interpretable activity categories, $F_{4,125} =$ `r sprintf("%.2f", primary_tests$f_statistic[primary_tests$placement == "Near-eye" & primary_tests$test_id == "H04-F1"])`,
with **raw p < 0.001** under the prespecified 0.05 omnibus decision rule. The estimates show large, precisely estimated
differences in magnitude, but the conclusion is an observational association:
reported activity category is associated with one-hour zero-aware geometric
mean melEDI. Chest evidence is complementary and points in the same broad
direction. Other/unspecified activity remains fitted but supports no scientific
claim.

Both exploratory temporal models pass their construction and added-context
gates. Following the author's 2026-08-11 amendment, their ribbons use the same
uncertainty implementation as H03: model-based pointwise 95% intervals from
the accepted GAMM's fitted-coefficient covariance. They are not simultaneous
bands and support no curve-wide significance claim. No bootstrap or simulation
enters the reported temporal results.
:::

No H03 result or significance classification was transferred. H03 supplied
the population-mean architecture; all H04 support, coefficients, restrictions,
intervals, diagnostics, and classifications were recomputed from H04 inputs.

## Fixed representation and exact fitted samples

The three outdoor flags—active travel, outdoor work, and outdoor free time—were
collapsed before pivoting. Duplicate collapsed labels were removed within an
hour. Co-selected Other was then suppressed in favour of named categories;
Other was retained only when it was the sole selection. A fitted hour with
$k$ retained distinct labels generated $k$ long rows, each with prior weight
$1/k$. Hours with no selected category are not “no activity”; they were kept
out of the category model and remain visible in sample flow. No diary free
text was inspected or recoded.

```{r}
#| label: tbl-stage2-sample-flow
#| tbl-cap: "Exact primary sample flow; hours lacking a retained category are accounted for but are not modelled as no activity."

main_flow |>
  transmute(
    Placement = as.character(.data$placement),
    `Outcome-grid hours` = .data$outcome_grid_hours,
    `Outcome-available hours` = .data$outcome_available_hours,
    `No eligible diary key` = .data$outcome_hours_without_eligible_diary_key,
    `No retained activity` = .data$outcome_hours_with_no_retained_activity,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Unique fitted hours` = .data$model_candidate_unique_hours,
    `Generated long rows` = .data$generated_long_rows,
    `Effective weighted hours` = .data$effective_weighted_hours
  ) |>
  audit_gt(
    "Primary sample flow",
    "Near-eye is primary; chest is complementary"
  ) |>
  fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  cols_width(Placement ~ pct(13), everything() ~ pct(9.7))
```

The near-eye model contains 126 participants, 724 participant-days, 16,526
unique participant-hours, 17,266 long rows, and 16,526 effective weighted
hours. The chest model contains 150 participants, 875 participant-days, 20,128
unique participant-hours, 21,071 long rows, and 20,128 effective weighted
hours.

```{r}
#| label: tbl-stage2-k-distribution
#| tbl-cap: "Number of retained distinct activity labels per fitted participant-hour."

k_distribution |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement, .data$k) |>
  transmute(
    Placement = as.character(.data$placement),
    k = .data$k,
    `Unique participant-hours` = .data$unique_participant_hours,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Percent of fitted hours` = .data$percent_of_fitted_hours
  ) |>
  audit_gt(
    "Distribution of k",
    "Each long row receives weight 1/k"
  ) |>
  fmt_integer(columns = c(k, `Unique participant-hours`, Participants, `Participant-days`), use_seps = TRUE) |>
  fmt_number(columns = `Percent of fitted hours`, decimals = 1) |>
  sub_missing()
```

```{r}
#| label: tbl-stage2-category-support
#| tbl-cap: "Observed category support and weighted totals in the exact fitted samples."

category_support |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement, .data$display_order) |>
  transmute(
    Placement = as.character(.data$placement),
    Category = .data$activity,
    `Unique hours` = .data$unique_participant_hours,
    `Long rows` = .data$long_rows,
    `Weighted hours` = .data$effective_weighted_hours,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites
  ) |>
  audit_gt(
    "Category support",
    "Weighted hours need not equal long rows when categories were co-selected"
  ) |>
  fmt_integer(columns = c(`Unique hours`, `Long rows`, Participants, `Participant-days`, Sites), use_seps = TRUE) |>
  fmt_number(columns = `Weighted hours`, decimals = 1, use_seps = TRUE) |>
  cols_width(Placement ~ pct(10), Category ~ pct(30), everything() ~ pct(10))
```

```{r}
#| label: tbl-stage2-site-support
#| tbl-cap: "Site-level support and weighted totals in the exact fitted samples."

weighted_support |>
  filter(.data$summary_level == "site") |>
  left_join(
    site_support |>
      distinct(
        .data$site,
        `Submitted-manuscript site` = .data$display_name,
        site_order = .data$display_order.y
      ),
    by = "site",
    relationship = "many-to-one"
  ) |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement, .data$site_order) |>
  transmute(
    Placement = as.character(.data$placement),
    Site = .data$`Submitted-manuscript site`,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Unique hours` = .data$unique_participant_hours,
    `Long rows` = .data$long_rows,
    `Weighted hours` = .data$effective_weighted_hours
  ) |>
  audit_gt(
    "Site support",
    "Sites follow submitted-manuscript names and order"
  ) |>
  fmt_integer(
    columns = c(Participants, `Participant-days`, `Unique hours`, `Long rows`),
    use_seps = TRUE
  ) |>
  fmt_number(columns = `Weighted hours`, decimals = 1, use_seps = TRUE)
```

Site-level weighted totals are retained in
`artifacts/06_model_data/H04/H04_weighted_support_totals.csv`, and complete
site-by-category support is retained in
`artifacts/06_model_data/H04/H04_site_category_support.csv`. KNUST ×
Other/unspecified activity remains observed in the additive primary frame but
is support-inadequate for a cell-specific heterogeneity estimate; no such term
was forced into the separate heterogeneity companion.

## Exact implemented models

### Wilkinson formulas supplied to the fitters

The following cell is evaluated during rendering but fits no model. It
reconstructs every fitted or prespecified fallback formula and verifies the
literal formula text against the durable Stage 2 registry.

```{r}
#| label: tbl-stage2-formulas
#| tbl-cap: "Exact Wilkinson formulas supplied to H04 fitting functions."
#| code-fold: false
#| echo: true

implemented_formulas <- list(
  primary_full = geo_medi_1h ~ site + activity,
  primary_five_named_null = geo_medi_1h ~ site + other_indicator,
  secondary_six_category_null = geo_medi_1h ~ site,
  heterogeneity_five_named_full = geo_medi_1h ~ site * activity_named,
  heterogeneity_five_named_additive = geo_medi_1h ~ site + activity_named,
  heterogeneity_core_full = geo_medi_1h ~ site * activity_core,
  heterogeneity_core_additive = geo_medi_1h ~ site + activity_core,
  v0_full = geo_medi_1h ~ site * activity_v0 + (1 | Id),
  v0_site_only = geo_medi_1h ~ site + (1 | Id),
  temporal_no_activity = stats::as.formula(paste0(
    "geo_medi_1h ~ s(time_hour, bs = 'cc', k = 12) + ",
    "s(time_hour, site, bs = 'sz', k = 12) + ",
    "s(time_hour, participant, bs = 'fs', k = 10) + ",
    "s(participant_day, bs = 're')"
  )),
  temporal_activity_long = stats::as.formula(paste0(
    "geo_medi_1h ~ s(time_hour, bs = 'cc', k = 12) + ",
    "s(time_hour, activity, bs = 'sz', k = 12) + ",
    "s(time_hour, site, bs = 'sz', k = 12) + ",
    "s(time_hour, participant, bs = 'fs', k = 10) + ",
    "s(participant_day, bs = 're')"
  ))
)

formula_text <- function(formula) {
  paste(deparse(formula, width.cutoff = 500L), collapse = " ")
}
implemented_formula_registry <- enframe(
  implemented_formulas,
  name = "formula_id",
  value = "formula"
) |>
  mutate(formula = vapply(.data$formula, formula_text, character(1)))

stopifnot(
  identical(implemented_formula_registry$formula_id, formulas$formula_id),
  identical(
    gsub("'", "\"", implemented_formula_registry$formula, fixed = TRUE),
    formulas$formula
  )
)

implemented_formula_registry |>
  audit_gt("Implemented formula registry") |>
  cols_width(formula_id ~ pct(25), formula ~ pct(75))
```

The following evaluated expressions record the exact calls without executing
another fit inside the report. The archived model objects assert the same
formula text.

```{r}
#| label: tbl-stage2-model-calls
#| tbl-cap: "Exact fitting calls used by the Stage 2 implementation; expressions are displayed, not re-fitted."
#| code-fold: false
#| echo: true

model_call_registry <- list(
  primary_quasi_tweedie = quote(stats::glm(
    formula = implemented_formulas$primary_full,
    data = model_frame,
    family = statmod::tweedie(var.power = 1.539919, link.power = 0),
    weights = analysis_weight,
    control = stats::glm.control(epsilon = 1e-10, maxit = 100L),
    model = TRUE,
    x = TRUE,
    y = TRUE
  )),
  participant_cluster_hc1 = quote(sandwich::vcovCL(
    fitted_primary_model,
    cluster = model_frame$participant,
    type = "HC1",
    cadjust = TRUE,
    fix = FALSE
  )),
  v0_bridge = quote(glmmTMB::glmmTMB(
    formula = implemented_formulas$v0_full,
    data = v0_frame,
    family = glmmTMB::tweedie(link = "log"),
    REML = FALSE,
    contrasts = list(
      site = stats::contr.sum,
      activity_v0 = stats::contr.treatment
    )
  )),
  temporal_preliminary = quote(mgcv::bam(
    formula = implemented_formulas$temporal_activity_long,
    data = temporal_frame,
    family = mgcv::Tweedie(p = 1.539919, link = "log"),
    weights = analysis_weight,
    method = "fREML",
    discrete = TRUE,
    rho = 0,
    AR.start = temporal_frame$AR_start,
    knots = list(time_hour = c(0, 24)),
    nthreads = 1L,
    gc.level = 1L,
    drop.unused.levels = TRUE
  )),
  temporal_final = quote(mgcv::bam(
    formula = implemented_formulas$temporal_activity_long,
    data = temporal_frame,
    family = mgcv::Tweedie(p = 1.539919, link = "log"),
    weights = analysis_weight,
    method = "fREML",
    discrete = TRUE,
    rho = estimated_rho,
    AR.start = temporal_frame$AR_start,
    knots = list(time_hour = c(0, 24)),
    nthreads = 1L,
    gc.level = 1L,
    drop.unused.levels = TRUE
  ))
)

enframe(model_call_registry, name = "model", value = "call") |>
  mutate(call = vapply(.data$call, function(x) {
    paste(deparse(x, width.cutoff = 500L), collapse = " ")
  }, character(1))) |>
  audit_gt("Implemented model calls") |>
  cols_width(model ~ pct(22), call ~ pct(78))
```

### Mean model, weights, and inference

The primary and complementary mean models use a quasi-Tweedie variance power
of 1.539919 and log link. The prior weight is the exact `analysis_weight = 1/k`.
The same site-prefixed participant cluster contains concurrent category rows
from an hour and all longitudinal rows from that participant. The covariance
is participant-clustered HC1 with cluster adjustment and no positive-definite
repair. Intervals use $t_{G-1}$; robust joint tests use $F_{q,G-1}$, where $G$
is the exact number of participant clusters.

For the primary five-category test, the unrestricted six-category model is
unchanged. If $R$ is the four-row restriction matrix that compares Sleeping,
vehicle/public-transport travel, indoor/home working, and Outdoors with At
home, then

$$
F = \frac{
  (R\widehat{\boldsymbol{\beta}})^\mathsf{T}
  (R V_{\mathrm{HC1}} R^\mathsf{T})^{-1}
  (R\widehat{\boldsymbol{\beta}})
}{4},
\qquad F \sim F_{4,G-1}.
$$

Every Other/unspecified coefficient in $R$ is zero. The secondary five-row
restriction adds Other versus At home and is reported only as `H04-F1b`.

Equal-site category means average site-specific fitted log means with weight
$1/S$ and back-transform once:

$$
\bar{\eta}_c = \frac{1}{S}\sum_{s=1}^{S}
x_{sc}^{\mathsf{T}}\widehat{\boldsymbol{\beta}},
\qquad
\bar{\mu}_c = \exp\!\left(\bar{\eta}_c\right).
$$

Thus $\bar{\mu}_c$ is the geometric mean of the site-specific fitted means,
not their response-scale arithmetic mean.

## Primary near-eye and complementary chest results

```{r}
#| label: tbl-stage2-primary-estimates
#| tbl-cap: "Equal-site standardized activity-category estimates with participant-cluster robust 95% confidence intervals."

primary_display <- primary |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    mean_ci = format_mean_ci(
      .data$standardized_mean_lx,
      .data$mean_conf_low_lx,
      .data$mean_conf_high_lx
    ),
    ratio_ci = ifelse(
      is.na(.data$ratio_to_home),
      "—",
      format_ratio_ci(
        .data$ratio_to_home,
        .data$ratio_conf_low,
        .data$ratio_conf_high
      )
    ),
    difference_ci = ifelse(
      is.na(.data$difference_from_home_lx),
      "—",
      format_difference_ci(
        .data$difference_from_home_lx,
        .data$difference_conf_low_lx,
        .data$difference_conf_high_lx
      )
    ),
    raw_p = ifelse(
      .data$inferential_role == "NAMED_VERSUS_HOME",
      format_p_cell(.data$p_raw, significant = FALSE),
      "—"
    ),
    adjusted_p = format_p_cell(
      .data$p_adjusted,
      significant = !is.na(.data$p_adjusted) & .data$p_adjusted < 0.05
    ),
    role = ifelse(
      .data$activity_code == "other",
      "Display only; no claim",
      ifelse(.data$activity_code == "home", "Reference", "H04-F2")
    )
  ) |>
  arrange(.data$placement, .data$display_order) |>
  transmute(
    Placement = as.character(.data$placement),
    Category = .data$activity,
    `Mean melEDI lx (95% CI)` = .data$mean_ci,
    `Ratio vs At home (95% CI)` = .data$ratio_ci,
    `Difference, lx (95% CI)` = .data$difference_ci,
    `Raw p` = .data$raw_p,
    `BH-adjusted p` = .data$adjusted_p,
    Role = .data$role
  )

primary_display |>
  audit_gt(
    "Activity-associated melEDI",
    "At home is the reference; H04-F2 decisions use BH-adjusted p; Other is display-only"
  ) |>
  fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`)) |>
  cols_width(
    Placement ~ pct(7),
    Category ~ pct(18),
    `Mean melEDI lx (95% CI)` ~ pct(16),
    `Ratio vs At home (95% CI)` ~ pct(16),
    `Difference, lx (95% CI)` ~ pct(16),
    `Raw p` ~ pct(7),
    `BH-adjusted p` ~ pct(8),
    Role ~ pct(12)
  )
```

Near eye, the equal-site standardized mean was 74.7 lx at home, 7.0 lx while
sleeping, 332.1 lx during vehicle/public-transport travel, 202.0 lx during
indoor/home working, and 864.8 lx outdoors. Relative to At home, the named
ratios were 0.093, 4.446, 2.705, and 11.578, respectively. Chest estimates were
7.7, 87.6, 575.9, 210.0, and 1,203.0 lx in the same display order. These are
average associations with reported categories, not transitions, individual
effects, or causal contrasts.

```{r}
#| label: tbl-stage2-primary-tests
#| tbl-cap: "Primary five-named-category and secondary six-category robust omnibus tests."

primary_tests |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    p = format_p_cell(.data$p_raw, significant = .data$p_raw < 0.05),
    Test = ifelse(
      .data$test_id == "H04-F1",
      "Five named categories; Other unrestricted",
      "All six categories (secondary)"
    )
  ) |>
  arrange(.data$placement, .data$test_id) |>
  transmute(
    Placement = as.character(.data$placement),
    Family = .data$test_id,
    Test,
    Restrictions = .data$restrictions,
    `F statistic` = .data$f_statistic,
    `Numerator df` = .data$restrictions,
    `Denominator df` = .data$denominator_df,
    `Raw p` = .data$p,
    Status = .data$status
  ) |>
  audit_gt(
    "Robust category tests",
    "Finite-cluster F inference; alpha = 0.05"
  ) |>
  fmt_number(columns = `F statistic`, decimals = 2) |>
  fmt_integer(columns = c(Restrictions, `Numerator df`, `Denominator df`)) |>
  fmt_markdown(columns = `Raw p`) |>
  cols_width(Test ~ pct(31), everything() ~ pct(9.8))
```

```{r}
#| label: fig-stage2-primary
#| fig-cap: "Equal-site standardized mean melEDI and named-category ratios versus At home."
#| fig-alt: "Two-panel figure. The left panel shows standardized mean melEDI with 95% confidence intervals for six activity categories at near-eye and chest placements. Sleeping is lowest and outdoors highest. The right panel shows four named-category ratios versus At home on a logarithmic scale; all confidence intervals exclude one. Other is shown only in the mean panel and is labelled display-only."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_primary_category_estimates.png"
))
```

## Separate site heterogeneity

The additive primary model was not replaced by an interaction. The separate
heterogeneity analysis excluded Other and tested only observed named-category
support. Both placements passed the results-blind covariance and influence
gate for the full five-named-category architecture; the four-category core
fallback was therefore not selected.

```{r}
#| label: tbl-stage2-heterogeneity
#| tbl-cap: "Support-gated site heterogeneity and robust omnibus tests."

heterogeneity_tests |>
  left_join(
    heterogeneity_gate |>
      select(
        "placement",
        "architecture",
        "full_rank",
        "restriction_covariance_condition_number",
        "maximum_cluster_score_share",
        "maximum_cluster_leverage_share",
        "gate_pass"
      ),
    by = c("placement", "selected_architecture" = "architecture")
  ) |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    p = format_p_cell(.data$p_raw, significant = .data$p_raw < 0.05)
  ) |>
  arrange(.data$placement) |>
  transmute(
    Placement = as.character(.data$placement),
    Architecture = .data$selected_architecture,
    Restrictions = .data$restrictions,
    `F statistic` = .data$f_statistic,
    `Denominator df` = .data$denominator_df,
    `Raw p` = .data$p,
    `Covariance condition` = .data$restriction_covariance_condition_number,
    `Max score share` = .data$maximum_cluster_score_share,
    `Max leverage share` = .data$maximum_cluster_leverage_share,
    `Gate pass` = .data$gate_pass
  ) |>
  audit_gt(
    "Five-named-category site heterogeneity",
    "Exploratory H04-F3; Other is excluded"
  ) |>
  fmt_number(
    columns = c(`F statistic`, `Covariance condition`),
    decimals = 2
  ) |>
  fmt_number(columns = c(`Max score share`, `Max leverage share`), decimals = 3) |>
  fmt_integer(columns = c(Restrictions, `Denominator df`)) |>
  fmt_markdown(columns = `Raw p`)
```

```{r}
#| label: tbl-stage2-site-deviation-family
#| tbl-cap: "Complete support-gated H04-F4 family accounting."

site_activity |>
  group_by(.data$placement) |>
  summarise(
    `H04-F4 members` = sum(!is.na(.data$site_deviation_p_adjusted)),
    `BH significant` = sum(
      .data$site_deviation_p_adjusted < 0.05,
      na.rm = TRUE
    ),
    `Other supported members` = sum(
      .data$site_deviation_p_adjusted >= 0.05,
      na.rm = TRUE
    ),
    `Support non-estimable` = sum(
      .data$reporting_status == "SUPPORT_NON_ESTIMABLE"
    ),
    `Other-category rows excluded` = sum(
      .data$reporting_status == "NOT_IN_SELECTED_ARCHITECTURE"
    ),
    .groups = "drop"
  ) |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement) |>
  rename(Placement = placement) |>
  audit_gt(
    "Site-deviation multiplicity accounting",
    "One complete BH family per placement; adjusted alpha = 0.050"
  ) |>
  fmt_integer(columns = c(
    `H04-F4 members`,
    `BH significant`,
    `Other supported members`,
    `Support non-estimable`,
    `Other-category rows excluded`
  ))
```

```{r}
#| label: fig-stage2-site-heterogeneity
#| fig-cap: "Observed site-specific activity estimates and the complete support-gated H04-F4 decisions from the accepted heterogeneity companion."
#| fig-alt: "Faceted point-and-interval plot of fitted melEDI by submitted-manuscript site and named activity category, with site colours fixed by the project registry. Filled circles mark supported site deviations with BH-adjusted p below 0.050 in the complete placement-specific family, open circles mark other supported estimates, and one cross marks the support-non-estimable near-eye San José Outdoors cell. Other is absent."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_site_activity_estimates.png"
))
```

This is evidence that activity-associated means vary among sites, not a basis
for selecting or ranking individual sites post hoc. The complete H04-F4 vector
contains 44 near-eye and 40 chest support-eligible deviations; 17 and 9,
respectively, pass the labelled BH-adjusted 0.050 rule. All raw and adjusted
p-values are retained in `H04_site_activity_estimands.csv` and the paired
figure-source CSV. Unsupported cells are labelled non-estimable rather than
filled or pooled.

## Diagnostics and acceptability

```{r}
#| label: tbl-stage2-diagnostic-assessment
#| tbl-cap: "Predeclared diagnostic assessments with plain-language interpretation."

assessments |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement, match(
    .data$diagnostic,
    unique(assessments$diagnostic)
  )) |>
  transmute(
    Placement = as.character(.data$placement),
    Diagnostic = .data$diagnostic,
    Evidence = .data$evidence,
    Assessment = .data$assessment,
    Interpretation = .data$interpretation
  ) |>
  audit_gt(
    "Primary mean-model diagnostics",
    "Acceptability applies to the stated population-average estimand"
  ) |>
  cols_width(
    Placement ~ pct(8),
    Diagnostic ~ pct(20),
    Evidence ~ pct(25),
    Assessment ~ pct(15),
    Interpretation ~ pct(32)
  )
```

Both models converged at full rank and had finite positive-definite registered
restriction covariance. Fractional weights summed to one within every hour.
No participant crossed the predeclared architecture-gate influence limits.
The primary mean models are nevertheless assessed **acceptable with
limitation**, not simply acceptable:

- exact zeros comprise 28.9% of near-eye and 29.4% of chest fitted hours; the
  quasi mean model retains them but does not separately model zero probability;
- absolute residual magnitude remains associated with fitted magnitude
  (Spearman 0.483 near eye and 0.524 chest); and
- unique-hour residual lag-one correlation remains 0.445 near eye and 0.498
  chest. Participant clustering encompasses the complete longitudinal record,
  but it does not make the serial structure disappear.

```{r}
#| label: fig-stage2-diagnostics
#| fig-cap: "Primary-model calibration and unique-hour residual autocorrelation."
#| fig-alt: "Diagnostic figure with calibration and residual autocorrelation panels for near-eye and chest models. Calibration varies across fitted-value bins, and residual correlation is positive at short lags for both placements, with the largest correlation at lag one."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_primary_diagnostics.png"
))
```

The bounded deletion audit refit each placement after deleting every site and
after deleting each of the five highest-score participants. No primary
decision changed. Across named contrasts, the maximum relative ratio change
in absolute magnitude was `r sprintf("%.1f", max(abs(influence$ratio_relative_change_percent[influence$inferential_role == "NAMED_VERSUS_HOME"]), na.rm = TRUE))`%.
This supports decision stability while flagging that some magnitude estimates
remain participant- or site-sensitive.

## Prespecified sensitivities and paired/common evidence

Long-row weighting is part of the estimand. Replacing $1/k$ by unit weight
changes the question from one contribution per participant-hour to an
unweighted selection-level analysis in which multi-select hours contribute
more. That run is a sensitivity, not an interchangeable implementation.

```{r}
#| label: tbl-stage2-sensitivity-samples
#| tbl-cap: "Exact fitted samples for every prespecified sensitivity."

samples |>
  filter(.data$scenario_id != "primary_dataset") |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    scenario_id = factor(
      .data$scenario_id,
      levels = c(
        "exactly_one_category",
        "retain_coselected_other",
        "exclude_other_only",
        "unweighted_long_rows",
        "gap_timing_unaware",
        "paired_common_sample",
        "working_power_1_30",
        "working_power_1_80"
      )
    )
  ) |>
  arrange(.data$scenario_id, .data$placement) |>
  transmute(
    Scenario = as.character(.data$scenario_id),
    Placement = as.character(.data$placement),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Unique hours` = .data$unique_participant_hours,
    `Long rows` = .data$long_rows,
    `Weighted hours` = .data$effective_weighted_hours
  ) |>
  audit_gt(
    "Sensitivity sample registry",
    "Unweighted-long weighted hours equal long rows by design"
  ) |>
  fmt_integer(
    columns = c(Participants, `Participant-days`, `Unique hours`, `Long rows`),
    use_seps = TRUE
  ) |>
  fmt_number(columns = `Weighted hours`, decimals = 0, use_seps = TRUE) |>
  cols_width(Scenario ~ pct(24), everything() ~ pct(12.7))
```

```{r}
#| label: tbl-stage2-sensitivity-stability
#| tbl-cap: "Stability of the four named-category ratios versus At home."

named_sensitivity |>
  group_by(.data$placement, .data$scenario_id) |>
  summarise(
    `Named contrasts` = dplyr::n(),
    Stable = sum(.data$stability == "stable"),
    `Maximum absolute relative change, %` = max(
      abs(.data$ratio_relative_change_percent),
      na.rm = TRUE
    ),
    `All directions concordant` = all(.data$direction_concordant),
    `All primary ratios inside sensitivity CI` = all(
      .data$primary_ratio_inside_sensitivity_interval
    ),
    .groups = "drop"
  ) |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$scenario_id, .data$placement) |>
  transmute(
    Scenario = .data$scenario_id,
    Placement = as.character(.data$placement),
    `Named contrasts`,
    Stable,
    `Maximum absolute relative change, %`,
    `All directions concordant`,
    `Primary ratio in sensitivity CI` =
      .data$`All primary ratios inside sensitivity CI`
  ) |>
  audit_gt(
    "Sensitivity stability",
    "Stability is descriptive and cannot upgrade the primary claim"
  ) |>
  fmt_integer(columns = c(`Named contrasts`, Stable)) |>
  fmt_number(columns = `Maximum absolute relative change, %`, decimals = 1)
```

```{r}
#| label: tbl-stage2-sensitivity-tests
#| tbl-cap: "Five-named-category robust omnibus results in each prespecified sensitivity."

sensitivity_tests |>
  filter(.data$test_id == "H04-F1") |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    p = format_p_cell(.data$p_raw, significant = FALSE)
  ) |>
  arrange(.data$scenario_id, .data$placement) |>
  transmute(
    Scenario = .data$scenario_id,
    Placement = as.character(.data$placement),
    Restrictions = .data$restrictions,
    `F statistic` = .data$f_statistic,
    `Denominator df` = .data$denominator_df,
    `Raw p` = .data$p,
    Status = .data$status
  ) |>
  audit_gt(
    "Sensitivity omnibus tests",
    "Unadjusted diagnostic results; no new confirmatory decision family"
  ) |>
  fmt_integer(columns = c(Restrictions, `Denominator df`)) |>
  fmt_number(columns = `F statistic`, decimals = 2) |>
  fmt_markdown(columns = `Raw p`) |>
  cols_width(Scenario ~ pct(28), everything() ~ pct(12))
```

All four named contrasts were classified stable in every required sensitivity
for both placements. The gap-timing-unaware runs are explicitly named as such;
they are not called “gap-aware” or “corrected”. Excluding Other-only hours
produces a five-category model, so its five-category omnibus is a sensitivity
test and is not relabelled as the secondary six-category omnibus.

The exact paired/common frame contains 110 participants, 625 participant-days,
14,308 unique participant-hours, 15,001 long rows, and 14,308 effective weighted
hours at each placement.

```{r}
#| label: fig-stage2-paired
#| fig-cap: "Near-eye and chest named-category ratios versus At home in the exact paired/common sample."
#| fig-alt: "Point-and-interval plot comparing four equal-site standardized activity-category ratios versus At home at near-eye and chest placements in the same 110 participants and 14,308 participant-hours. Sleeping ratios are below one, whereas vehicle or public-transport travel, indoor or home working, and Outdoors ratios are above one. Participant-cluster-robust 95% confidence intervals are displayed for each separately fitted placement."
#| out-width: "90%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_paired_placement_comparison.png"
))
```

The paired display compares separately fitted named-category ratios on
identical participant-hour keys. It is not a within-hour difference model and
supplies no unplanned placement-difference p-value.

## Exploratory temporal context

### Approved thin-plate `sz` construction

The activity-long temporal model uses the same rows and $1/k$ weights as the
mean model. Its global `s(time_hour, bs = "cc", k = 12)` is cyclic. The
activity and site `sz` terms omit `xt`, and the synchronized `mgcv` constructor
resolves both to the approved thin-plate marginal basis. They are not cyclic;
separation at midnight is therefore accepted, quantified, and displayed with
open endpoints and no wrap-around line.

```{r}
#| label: tbl-stage2-temporal-basis
#| tbl-cap: "Constructor-verified temporal basis contract."

temporal_basis |>
  transmute(
    Component = .data$component,
    `Constructed class` = .data$constructed_class,
    `Marginal basis` = .data$marginal_basis,
    Cyclic = .data$cyclic
  ) |>
  audit_gt(
    "Temporal basis contract",
    "Only the global time component is cyclic"
  ) |>
  cols_width(Component ~ pct(20), `Constructed class` ~ pct(24), `Marginal basis` ~ pct(42), Cyclic ~ pct(14))
```

Activity-specific runs ensure that concurrent rows with the same timestamp
never become lag-one neighbours. A run restarts at participant-day, activity,
DST repeated-hour, non-consecutive-hour, or gap boundaries. The no-activity
comparator uses the identical long frame, weights, and run construction.

```{r}
#| label: tbl-stage2-temporal-retention
#| tbl-cap: "Predeclared temporal retention decision."

temporal_retention |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement) |>
  transmute(
    Placement = as.character(.data$placement),
    `Minimum k-index` = .data$minimum_finite_k_index,
    `Activity-model AIC` = .data$activity_model_aic,
    `No-activity AIC` = .data$no_activity_model_aic,
    `Delta AIC` = .data$activity_model_delta_aic,
    `Activity deviance explained` = .data$activity_model_deviance_explained,
    `Comparator deviance explained` = .data$no_activity_model_deviance_explained,
    `Residual lag 1` = .data$residual_lag1,
    `Sparse clock/category cells` = .data$locally_sparse_clock_activity_cells,
    `Max activity midnight ratio` = .data$maximum_activity_endpoint_ratio,
    `Max site midnight ratio` = .data$maximum_site_endpoint_ratio,
    Retained = .data$retained_for_context,
    Assessment = .data$assessment
  ) |>
  audit_gt(
    "Temporal adequacy and added-context gate",
    "AIC is used only for this exploratory comparator"
  ) |>
  fmt_number(
    columns = c(
      `Minimum k-index`,
      `Activity deviance explained`,
      `Comparator deviance explained`,
      `Residual lag 1`
    ),
    decimals = 3
  ) |>
  fmt_number(
    columns = c(`Activity-model AIC`, `No-activity AIC`, `Delta AIC`),
    decimals = 1,
    use_seps = TRUE
  ) |>
  fmt_number(
    columns = c(`Max activity midnight ratio`, `Max site midnight ratio`),
    decimals = 2
  ) |>
  fmt_integer(columns = `Sparse clock/category cells`)
```

Near eye, the activity model improves AIC by 3,215 and explains 84.8% versus
81.8% deviance; chest improves AIC by 4,984 and explains 83.2% versus 79.5%.
The minimum finite k-indices are 0.872 and 0.863. Final residual lag-one
correlations are 0.176 and 0.167. Both models therefore add context and are
retained **acceptable with limitation**. The 41 near-eye and 40 chest locally
sparse clock/category cells, residual dependence, and potentially large
thin-plate midnight separation remain explicit limitations.
Both retained activity-model passes completed without warnings. The chest
no-activity comparator's preliminary zero-rho fit warned that its algorithm
did not converge, but its final fixed-rho fit converged without a warning; the
comparator is used only for this exploratory added-context check.

```{r}
#| label: fig-stage2-temporal-near
#| fig-cap: "Exploratory near-eye activity-associated time-of-day means with model-based pointwise 95% intervals and weighted support."
#| fig-alt: "Six activity-specific near-eye melEDI curves over local time with pale pointwise 95 percent ribbons, each above a matching weighted-support histogram. Curves have open points at 0 and 24 hours to expose non-cyclic activity deviations. Magenta bars mark locally sparse clock-by-activity cells."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_temporal_near_eye.png"
))
```

```{r}
#| label: fig-stage2-temporal-chest
#| fig-cap: "Exploratory chest activity-associated time-of-day means with model-based pointwise 95% intervals and weighted support."
#| fig-alt: "Six activity-specific chest melEDI curves over local time with pale pointwise 95 percent ribbons, each above a matching weighted-support histogram. Curves have open points at 0 and 24 hours to expose non-cyclic activity deviations. Magenta bars mark locally sparse clock-by-activity cells."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "10_figures", "H04", "H04_temporal_chest.png"
))
```

Approximate smooth p-values are not used as the H04 primary test. The temporal
curves are descriptive exploratory context and cannot support a causal or
“significant period” claim.

### H03-aligned pointwise uncertainty

For each time-by-activity curve, the fitted population design rows are averaged
equally across the observed sites on the log-mean scale after setting the
participant and participant-day smooth contributions to zero. If
$\bar{\boldsymbol{x}}_{tc}$ is that design row, $\widehat{\boldsymbol\beta}$
is the accepted final GAMM's coefficient vector, and
$\widehat{\boldsymbol V}_{\beta}$ is its fitted-coefficient covariance, the
reported interval is

$$
\exp\left\{
\bar{\boldsymbol{x}}_{tc}^{\mathsf T}\widehat{\boldsymbol\beta}
\mathbin{\pm}
1.96\sqrt{
\bar{\boldsymbol{x}}_{tc}^{\mathsf T}
\widehat{\boldsymbol V}_{\beta}
\bar{\boldsymbol{x}}_{tc}
}
\right\}.
$$

This is the H03 implementation: `stats::vcov(fit, unconditional = TRUE)` with
the fitted `Vp` matrix only as an error fallback, followed by one
back-transformation. It conditions on the fitted smoothing structure. The
ribbons are pointwise 95% intervals, not simultaneous bands, and no
time-specific hypothesis test is performed. This permits no curve-wide inference.

```{r}
#| label: tbl-stage2-temporal-uncertainty
#| tbl-cap: "H03-aligned temporal uncertainty contract."

temporal_uncertainty |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement) |>
  transmute(
    Placement = as.character(.data$placement),
    `Covariance source` = .data$covariance_source,
    Interval = paste0(
      "Pointwise 95% (critical value ",
      sprintf("%.2f", .data$interval_critical_value),
      ")"
    ),
    `Resampling replicates` = .data$resampling_replicates,
    `Inferential boundary` = paste(
      "No simultaneous band; no time-specific or curve-wide inference"
    )
  ) |>
  audit_gt(
    "Temporal uncertainty",
    "Model-based pointwise intervals; no bootstrap or simulation"
  ) |>
  fmt_integer(columns = `Resampling replicates`) |>
  cols_width(
    Placement ~ pct(11),
    `Covariance source` ~ pct(26),
    Interval ~ pct(23),
    `Resampling replicates` ~ pct(13),
    `Inferential boundary` ~ pct(27)
  )
```

::: {.callout-note title="Superseded bootstrap pilot preserved only for provenance"}
The bootstrap pilot was stopped before its target after the author selected
the H03 uncertainty implementation. Its `r sum(bootstrap_supersession$completed_checkpoint_files)`
completed checkpoint files are retained in the H04 model-artifact directory
so the attempted computation remains auditable. They are excluded from H04
estimates, intervals, figures, tests, and scientific claims. No production
bootstrap is required or authorized.
:::

## V0 reconstruction and comparison

V0 used a unit-weight long-row analysis, removed Other, could retain duplicate
collapsed outdoor rows, and fit a Tweedie mixed model with the joint
site-by-activity interaction. Its site-only versus full likelihood-ratio test
jointly added four activity coefficients and all supported site interactions;
it is not a test of the repaired H04 five-category equality hypothesis.

```{r}
#| label: tbl-stage2-v0-denominators
#| tbl-cap: "Frozen V0, current-input V0 bridge, and repaired H04 analytical denominators."

v0_comparison |>
  mutate(placement = factor(.data$placement, levels = c("Near-eye", "Chest"))) |>
  arrange(.data$placement, factor(
    .data$implementation,
    levels = c(
      "Frozen V0 model",
      "V0 bridge on current normalized inputs",
      "Approved repaired H04"
    )
  )) |>
  transmute(
    Placement = as.character(.data$placement),
    Implementation = .data$implementation,
    `Long rows` = .data$long_rows,
    `Unique participant-hours` = .data$unique_participant_hours,
    `Effective weighted hours` = .data$effective_weighted_hours,
    Categories = .data$categories,
    `Test definition` = .data$test_definition
  ) |>
  audit_gt(
    "V0 denominator and test comparison",
    "The current-input bridge is not the frozen submitted result"
  ) |>
  fmt_integer(
    columns = c(
      `Long rows`, `Unique participant-hours`,
      `Effective weighted hours`, Categories
    ),
    use_seps = TRUE
  ) |>
  sub_missing() |>
  cols_width(
    Placement ~ pct(9),
    Implementation ~ pct(23),
    `Test definition` ~ pct(38),
    everything() ~ pct(7.5)
  )
```

The frozen near-eye and chest V0 records contain 16,801 and 20,327 model rows.
The bridge rebuilt on current normalized inputs contains 16,925 rows/16,135
unique hours near eye and 20,489 rows/19,497 unique hours at chest. These are
not the frozen V0 denominators and are kept separate. The repaired H04 models
contain more long rows because Other-only hours remain, but their effective
weighted hours equal unique participant-hours by construction.

```{r}
#| label: tbl-stage2-v0-tests
#| tbl-cap: "Current-input reconstruction of V0's joint category-plus-interaction likelihood-ratio test."

v0_tests |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    p = format_p_cell(.data$p_raw, significant = FALSE)
  ) |>
  arrange(.data$placement) |>
  transmute(
    Placement = as.character(.data$placement),
    `Restrictions added` = .data$restrictions_added,
    `Likelihood-ratio chi-square` = .data$likelihood_ratio_chisq,
    `Raw p` = .data$p,
    Interpretation = .data$interpretation
  ) |>
  audit_gt(
    "V0 bridge joint tests",
    "Comparator only; no repaired-H04 decision or significance styling"
  ) |>
  fmt_integer(columns = `Restrictions added`) |>
  fmt_number(columns = `Likelihood-ratio chi-square`, decimals = 2) |>
  fmt_markdown(columns = `Raw p`) |>
  cols_width(Interpretation ~ pct(52), everything() ~ pct(12))
```

```{r}
#| label: tbl-stage2-v0-bridge
#| tbl-cap: "Recognizable V0 category-versus-home ratios rebuilt on current normalized inputs."

v0_ratios |>
  mutate(
    placement = factor(.data$placement, levels = c("Near-eye", "Chest")),
    ratio_ci = format_ratio_ci(
      .data$ratio_to_home,
      .data$conf_low,
      .data$conf_high
    )
  ) |>
  arrange(.data$placement, .data$contrast) |>
  transmute(
    Placement = as.character(.data$placement),
    `V0 contrast` = .data$contrast,
    `Ratio (95% model-based CI)` = .data$ratio_ci,
    Role = "V0 bridge only; no repaired-H04 decision"
  ) |>
  audit_gt(
    "V0 bridge ratios",
    "Current-input reconstruction; not the approved estimand or inference"
  ) |>
  cols_width(Placement ~ pct(12), `V0 contrast` ~ pct(26), `Ratio (95% model-based CI)` ~ pct(25), Role ~ pct(37))
```

The reconstructed V0 joint tests and familiar ratios are reported only to
explain why the repaired estimates differ. They do not inherit V0's p-values,
significance labels, or causal-sounding wording.

## Reproducibility, source data, and reporting limits

All scientific transformations, fits, diagnostics, estimates, and displays
were generated with R 4.6.1 and the synchronized project library. No package
was installed or updated. Input identities were hash-verified before fitting.
The model objects, exact frames, restrictions, diagnostics, source-data CSVs,
figures, checkpointed pilot replicates, and software versions are H04-scoped
artifacts.

```{r}
#| label: tbl-stage2-software
#| tbl-cap: "R and consequential package versions used for H04 Stage 2."

software |>
  transmute(Component = .data$component, Version = .data$version) |>
  audit_gt("Software environment") |>
  cols_width(Component ~ pct(60), Version ~ pct(40))
```

Durable figure/source pairs are:

- `H04_primary_category_estimates.*` with `H04_primary_category_figure.csv`;
- `H04_site_activity_estimates.*` with `H04_site_activity_figure.csv`;
- `H04_paired_placement_comparison.*` with
  `H04_paired_placement_figure.csv`;
- `H04_primary_diagnostics.*` with the calibration and ACF CSVs;
- `H04_temporal_near_eye.*` and `H04_temporal_chest.*` with exact
  placement-specific curve and weighted-support CSVs; and
- the temporal uncertainty contract and bootstrap-supersession registry, which
  make the pointwise-only reporting boundary and the unused pilot evidence
  explicit.

The primary inference is population-average and observational. Fractional
membership is a transparent allocation rule, not proof that the labels fully
capture behaviour. Self-report timing and category ambiguity remain. Other is
heterogeneous and non-substantive. Equal-site standardization describes the
observed site set and does not generalize automatically to other populations.
Robust covariance protects the registered mean-model tests against arbitrary
within-participant correlation under the working mean model; it does not cure
mean misspecification, sparse cells, or influential clusters.

## Stage 2 stop

The Stage 2 implementation is complete. No Stage 3 reader report has been
created. The following decisions require explicit author approval:

```{r}
#| label: tbl-stage2-decisions
#| tbl-cap: "Decisions requested at the H04 Stage 2 gate."

tribble(
  ~gate_id, ~decision_requested,
  "H04-S2-G1",
  paste(
    "Approve the Stage 2 primary result and claim scope: an observational",
    "association between reported activity category and one-hour melEDI,",
    "with Other excluded from scientific claims."
  ),
  "H04-S2-G2",
  paste(
    "Approve the separate five-named-category site-heterogeneity result and",
    "its support-limited exploratory presentation."
  ),
  "H04-S2-G3",
  paste(
    "Approve the primary diagnostics as acceptable with limitation and retain",
    "the full sensitivity and influence qualifications."
  ),
  "H04-S2-G4",
  paste(
    "Approve the exploratory temporal presentation with H03-aligned",
    "model-based pointwise 95% intervals, no simultaneous bands or",
    "curve-wide inference, and the stated residual, support, and midnight",
    "limitations."
  ),
  "H04-S2-G5",
  paste(
    "Authorize Stage 3 as a",
    "standalone reader-facing H04 report with no V0 or construction-history",
    "language."
  )
) |>
  audit_gt(
    "H04 Stage 2 gate",
    "Stage 3 is not authorized by this report"
  ) |>
  cols_width(gate_id ~ pct(13), decision_requested ~ pct(87))
```

**STOP.** Await explicit author approval. Do not create the Stage 3 reader
report or begin Stage 4 from this document.
