---
title: "H01 analysis preparation and provenance"
subtitle: "From verified personal light-exposure metrics to reported results"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(readr)
  library(scales)
  library(stringr)
  library(tibble)
  library(tidyr)
})

root_candidates <- unique(c(
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(file.path(getwd(), "../../.."), winslash = "/", mustWork = FALSE)
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(path, "scripts/hypotheses/H01/h01_contract.R"))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(root.dir = root)

source(file.path(root, "scripts/pipeline/paths_io.R"))
source(file.path(root, "scripts/pipeline/p_value_display.R"))
source(file.path(root, "scripts/hypotheses/H01/h01_contract.R"))

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H01 preparation requires R 4.6.1", call. = FALSE)
}

read_h01_csv <- function(path) {
  readr::read_csv(file.path(root, path), show_col_types = FALSE, progress = FALSE)
}

relative_path <- function(path) {
  sub(
    paste0("^", gsub("([][{}()+*^$|\\\\?.])", "\\\\\\1", root), "/"),
    "",
    normalizePath(path, winslash = "/", mustWork = TRUE)
  )
}

prep_gt <- function(table) {
  table |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(12),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "bold",
      source_notes.font.size = gt::px(10)
    )
}

format_count <- function(value) {
  ifelse(
    is.finite(value),
    format(value, big.mark = ",", scientific = FALSE, trim = TRUE),
    "Unavailable"
  )
}

format_range <- function(lower, upper, suffix = "") {
  ifelse(
    is.finite(lower) & is.finite(upper),
    paste0(format_count(lower), "–", format_count(upper), suffix),
    "Unavailable"
  )
}

verify_artifact_manifest <- function(path) {
  manifest <- read_h01_csv(path)
  stopifnot(all(c("path", "sha256") %in% names(manifest)))
  files <- file.path(root, manifest$path)
  stopifnot(all(file.exists(files)))
  observed <- unname(vapply(files, artifact_sha256, character(1)))
  stopifnot(identical(observed, unname(manifest$sha256)))
  manifest
}

input_contract <- h01_input_contract(root)
metric_registry <- h01_metric_registry()
family_registry <- h01_family_registry()
run_registry <- h01_run_registry()
invisible(h01_validate_registry(metric_registry))

main_manifest_path <- "artifacts/12_manifests/H01_model_data_artifacts.csv"
gap_manifest_path <-
  "artifacts/12_manifests/H01_manuscript_prepared_data_artifacts.csv"
bootstrap_manifest_path <- paste0(
  "artifacts/12_manifests/",
  "H01_response_family_bootstrap_production_artifacts.csv"
)
stage3_manifest_path <-
  "artifacts/12_manifests/H01_stage3_reporting_artifacts.csv"

stopifnot(
  identical(
    artifact_sha256(file.path(root, main_manifest_path)),
    input_contract$main$manifest_sha256
  ),
  identical(
    artifact_sha256(file.path(root, gap_manifest_path)),
    input_contract$manuscript_prepared_data$manifest_sha256
  )
)

main_manifest <- verify_artifact_manifest(main_manifest_path)
gap_manifest <- verify_artifact_manifest(gap_manifest_path)
bootstrap_manifest <- verify_artifact_manifest(bootstrap_manifest_path)
stage3_manifest <- verify_artifact_manifest(stage3_manifest_path)

main_rows <- read_h01_csv("artifacts/06_model_data/H01/model_rows.csv")
main_sample_flow <- read_h01_csv("artifacts/06_model_data/H01/sample_flow.csv")
main_exclusions <- read_h01_csv("artifacts/06_model_data/H01/exclusion_reasons.csv")
metric_contract <- read_h01_csv("artifacts/06_model_data/H01/metric_contract.csv")
exact_samples <- read_h01_csv(
  "artifacts/09_tables/H01/stage3/H01_stage3_exact_samples.csv"
)
exact_samples_by_site <- read_h01_csv(
  "artifacts/09_tables/H01/H01_exact_samples_by_site.csv"
)
formula_manifest <- read_h01_csv(
  "artifacts/09_tables/H01/stage3/H01_stage3_formula_manifest.csv"
)
diagnostics <- read_h01_csv(
  "artifacts/08_diagnostics/H01/H01_model_diagnostics.csv"
)
bootstrap_audit <- read_h01_csv(
  "artifacts/08_diagnostics/H01/H01_r2_bootstrap_audit.csv"
)
site_registry <- read_h01_csv("config/site_display_registry.csv") |>
  arrange(.data$display_order)

primary_near_id <- "main__glasses__all_available"
primary_chest_id <- "main__chest__all_available"
placement_labels <- c(glasses = "Near eye", chest = "Chest")
placement_colors <- c("Near eye" = "#0072B2", "Chest" = "#D55E00")
placement_shapes <- c("Near eye" = 16, "Chest" = 17)

source_dir <- file.path(root, "artifacts/11_source_data/H01/preparation")
dir.create(source_dir, recursive = TRUE, showWarnings = FALSE)
```

::: {.callout-note title="About this analysis record"}
This page explains how verified personal light-exposure metrics were turned
into the H01 results. It describes the analytical samples, transformations,
response families, model implementation, diagnostics, sensitivity branches,
and the connection between code and stored outputs. The companion
[H01 results report](../../../notebooks/hypotheses/H01.html) presents the
scientific findings.

Rendering verifies bounded file identities and calculates descriptive
summaries from the H01-specific model rows and sample-flow records. It does
**not** call `lm()`, `lmer()`, `glmmTMB()`, `emmeans()`, `predict()`, or
`simulate()`, refit any model, regenerate predictions, or rerun a bootstrap.
Model estimates, 95% confidence intervals, diagnostics, influence results,
and bootstrap summaries are read from the stored outputs named below.
:::

## Purpose

H01 asks whether personal light exposure differs among study sites. The
analysis treats near-eye measurements as primary and chest measurements as
complementary evidence. It evaluates 17 prespecified light-exposure metrics,
while distinguishing an overall site effect from photoperiod and absolute
latitude.

This record explains **which data and code produced those results, why each
operation was needed, what it contributed, and what evidence shows that it
behaved as intended**. The sections follow the data-to-result sequence.

## Analysis map

```{mermaid}
flowchart TD
  A["Verified near-eye and chest metric data"] --> B["Input identity, schema, key,<br/>unit, and construct checks"]
  B --> C["Eight H01 model-row scenarios<br/>preparation × placement × sample"]
  C --> D["Seventeen response-specific<br/>model frames"]
  D --> E["One shared model implementation<br/>for both preparation rules"]
  E --> F["Site, photoperiod, latitude,<br/>and adequacy tests"]
  E --> G["From-mean site contrasts<br/>and 95% intervals"]
  E --> H["R² summaries and<br/>joint bootstrap intervals"]
  E --> I["Diagnostics, influence,<br/>and sensitivity branches"]
  F --> J["Stored tables, figures,<br/>source data, and manifests"]
  G --> J
  H --> J
  I --> J
  J --> K["H01 results report"]
  L["Automated identity and structure checks"] -. verify .-> B
  L -. verify .-> E
  L -. verify .-> J
```

The arrows show dependencies. They do not imply that this page repeats the
fitting, prediction, or resampling operations.

## What is calculated when this page opens

```{r}
#| label: tbl-h01-prep-boundary
#| tbl-cap: "Calculations performed when this page is rendered."

tibble::tribble(
  ~Category, ~Calculated_on_render, ~Explanation,
  "File verification", "Yes", "Checks the current H01 input, production-bootstrap, and reporting files against their recorded SHA-256 identities.",
  "Descriptive sample summaries", "Yes", "Summarises already prepared H01 model rows, fitted samples, and sample retention without changing membership.",
  "Model fitting and comparison", "No", "Reads stored model records and inferential tables.",
  "Predictions and marginal means", "No", "Reads stored estimates and contrasts produced by the model workflow.",
  "Bootstrap uncertainty", "No", "Reads production draws, audit records, and 95% interval summaries.",
  "Diagnostics and influence", "No", "Reads stored residual, prediction-bound, participant-influence, and leave-one-site-out outputs."
) |>
  gt::gt() |>
  gt::cols_width(
    Category ~ gt::pct(22),
    Calculated_on_render ~ gt::pct(17),
    Explanation ~ gt::pct(61)
  ) |>
  prep_gt()
```

## 1. Data entering the analysis

### Verified inputs

Two prepared datasets enter the analysis. The primary dataset could be
interpreted as a **time-sensitive primary metric dataset** because the timing
of remaining missing observations is used where it matters for the metric.
The predefined sensitivity is the **gap-timing-unaware dataset**. It still
passed the general 50%-per-hour and 80%-per-day coverage rules, but the timing
of remaining missing observations was not used for an additional
metric-specific adjustment. This name does not mean that gaps, missingness,
or coverage were ignored.

```{r}
#| label: tbl-h01-prep-inputs
#| tbl-cap: "Verified model-data inputs entering the H01 implementation."

input_summary <- tibble::tribble(
  ~Input, ~Path, ~Manifest, ~Manifest_sha256, ~Status,
  "Primary dataset", relative_path(input_contract$main$path), main_manifest_path, input_contract$main$manifest_sha256, "PASS",
  "Gap-timing-unaware dataset", relative_path(input_contract$manuscript_prepared_data$path), gap_manifest_path, input_contract$manuscript_prepared_data$manifest_sha256, "PASS"
) |>
  mutate(`SHA-256 prefix` = substr(.data$Manifest_sha256, 1L, 16L)) |>
  select(.data$Input, .data$Path, .data$Manifest, .data$`SHA-256 prefix`, .data$Status)

input_summary |>
  gt::gt() |>
  gt::cols_width(
    Input ~ gt::pct(20),
    Path ~ gt::pct(34),
    Manifest ~ gt::pct(31),
    `SHA-256 prefix` ~ gt::pct(11),
    Status ~ gt::pct(4)
  ) |>
  gt::tab_source_note(
    "Both manifests were checked row by row against the current files during rendering."
  ) |>
  prep_gt()
```

<details>
<summary>Show the complete pinned manifest identities</summary>

- Primary model-data manifest:
  `r input_contract$main$manifest_sha256`
- Gap-timing-unaware model-data manifest:
  `r input_contract$manuscript_prepared_data$manifest_sha256`
- Shared H01 implementation contract:
  `r input_contract$implementation_contract_sha256`
- Shared implementation source:
  `r input_contract$shared_implementation_sha256`

</details>

### What these files contain

Each H01 data bundle contains the 17-metric contract, long model rows, sample
flow, exclusions, predictor centres, scenario availability, a variable
dictionary, and input provenance. The primary long table contains
`r format_count(nrow(main_rows))` rows and 50 named fields. Its composite key
is scenario, placement, sample definition, metric, site, participant, and—when
the outcome is daily—local date. The construction code verifies uniqueness
within the appropriate analytical unit before a model frame is written.

The model rows retain the original metric value and units, the response
construct, the metric-specific support denominator when available, site and
participant identifiers, photoperiod, latitude, and the centred predictors
supplied to the models. Missing support is never converted to zero exposure.

### Why prepared records did not enter every model frame

The same metric-availability rules are applied before the site and latitude
models are constructed. The table counts primary-dataset records excluded
from each placement's all-available frame; it lists each record once by using
the site-plus-photoperiod frame, because the corresponding latitude frame
applies the same outcome-availability screen.

```{r}
#| label: tbl-h01-prep-exclusions
#| tbl-cap: "Reasons prepared primary-dataset records did not enter an all-available fitted model frame."

exclusion_labels <- c(
  below_light_profile_support = "Insufficient illuminance-profile support",
  below_medi_profile_support = "Insufficient melEDI-profile support",
  incomplete_state_domain = "Incomplete required state domain",
  insufficient_days = "Too few contributing days",
  insufficient_relevance_support = "Insufficient metric-relevant support",
  insufficient_state_support = "Insufficient state-specific support",
  low_relevance_coverage = "Low metric-relevant coverage",
  no_state_window = "Required state window unavailable",
  no_threshold_event = "Required threshold event unavailable"
)

primary_exclusion_summary <- main_exclusions |>
  filter(
    .data$data_scenario_id == "main",
    .data$scenario == "all_available",
    .data$model_question == "site_and_photoperiod"
  ) |>
  mutate(
    reason_id = sub("^metric_not_estimable:", "", .data$exclusion_reason),
    Reason = unname(exclusion_labels[.data$reason_id]),
    Placement = unname(placement_labels[.data$placement])
  ) |>
  group_by(.data$Placement, .data$Reason) |>
  summarise(
    `Excluded records` = sum(.data$rows),
    `Affected metrics` = n_distinct(.data$metric_id),
    .groups = "drop"
  ) |>
  arrange(
    factor(.data$Placement, c("Near eye", "Chest")),
    desc(.data$`Excluded records`),
    .data$Reason
  )

stopifnot(!anyNA(primary_exclusion_summary$Reason))

primary_exclusion_summary |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_number(
    columns = c(`Excluded records`, `Affected metrics`),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Reason ~ gt::pct(58),
    `Excluded records` ~ gt::pct(23),
    `Affected metrics` ~ gt::pct(19)
  ) |>
  gt::tab_source_note(
    "Counts are outcome-specific and therefore are not unique participants or participant-days across metrics."
  ) |>
  prep_gt()
```

## 2. The 17-metric response package

```{r}
#| label: tbl-h01-prep-metric-contract
#| tbl-cap: "Outcome contract applied to the 17 H01 metrics."

metric_registry |>
  left_join(
    metric_contract |>
      select(
        .data$metric_id,
        .data$manuscript_name,
        .data$manuscript_category,
        .data$display_unit
      ),
    by = "metric_id"
  ) |>
  transmute(
    Order = .data$metric_order,
    Category = str_to_sentence(str_replace_all(.data$manuscript_category, "-", " ")),
    Metric = .data$manuscript_name,
    `Model unit` = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    Transformation = recode(
      .data$response_transform,
      logit = "Logit",
      identity = "Identity",
      log10_offset_0.1 = "log10(value + 0.1)",
      clock_hours = "Linear clock hours",
      clock_hours_midnight_after_16 = "Linear clock hours; subtract 24 strictly after 16:00"
    ),
    Family = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie with log link"
    ),
    `Reported effect` = recode(
      .data$effect_scale,
      difference = "Difference",
      ratio = "Ratio",
      odds_ratio = "Odds ratio"
    )
  ) |>
  gt::gt(groupname_col = "Category", rowname_col = "Metric") |>
  gt::cols_hide(columns = "Order") |>
  gt::cols_width(
    Metric ~ gt::pct(27),
    `Model unit` ~ gt::pct(13),
    Transformation ~ gt::pct(25),
    Family ~ gt::pct(20),
    `Reported effect` ~ gt::pct(15)
  ) |>
  prep_gt()
```

### Construct-specific rules

```{r}
#| label: tbl-h01-prep-construct-rules
#| tbl-cap: "Metric-specific rules needed to preserve the intended response constructs."

tibble::tribble(
  ~Metric, ~Rule, ~Why_it_is_needed, ~Audit_evidence,
  "Time below 10 lx melEDI before sleep", "Calendar-day cumulative duration; values strictly above six hours trigger an audit warning but are not truncated", "Two separate three-hour pre-sleep windows can contribute to one calendar day when sleep timing changes", "H01_gate_a_pre_sleep_calendar_day.csv",
  "Midpoint of the darkest 10 hours", "Subtract 24 hours only from values strictly after 16:00", "Places the observed nighttime midpoint distribution on a continuous linear scale without a noon discontinuity", "H01_gate_b_l10_conversion.csv",
  "Longest continuous period above 250 lx melEDI", "Retain the prepared period definition; repeat the analysis using exactly identified periods", "Distinguishes the primary duration construct from uncertainty at period boundaries", "Stored exactly-identified-period sensitivity outputs",
  "Timing metrics", "Model on a prespecified continuous linear clock scale", "The observed distributions are unimodal on the selected cut and do not require a circular likelihood", "Stored timing-range and residual diagnostics"
) |>
  gt::gt() |>
  gt::cols_width(
    Metric ~ gt::pct(23),
    Rule ~ gt::pct(30),
    Why_it_is_needed ~ gt::pct(34),
    Audit_evidence ~ gt::pct(13)
  ) |>
  prep_gt()
```

## 3. Model-frame and sample construction

### Eight analysis scenarios

The same implementation creates two placements × two prepared datasets × two
sample definitions. The all-available near-eye primary dataset is the primary
analysis. Chest is fitted separately. Placement-matched samples retain the
same participants and participant-days at both placements for direct
near-eye–chest comparison; they are not pooled models.

```{r}
#| label: tbl-h01-prep-scenarios
#| tbl-cap: "H01 model-row scenarios created before fitting."

scenario_summary <- exact_samples |>
  group_by(
    .data$run_id,
    .data$data_scenario_id,
    .data$placement,
    .data$sample_scenario
  ) |>
  summarise(
    Metrics = n_distinct(.data$metric_id),
    `Participants (range)` = format_range(
      min(.data$participants, na.rm = TRUE),
      max(.data$participants, na.rm = TRUE)
    ),
    `Participant-days (range)` = format_range(
      min(.data$participant_days, na.rm = TRUE),
      max(.data$participant_days, na.rm = TRUE)
    ),
    Sites = max(.data$sites, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    Dataset = recode(
      .data$data_scenario_id,
      main = "Primary dataset",
      manuscript_prepared_data = "Gap-timing-unaware dataset"
    ),
    Placement = unname(placement_labels[.data$placement]),
    Sample = recode(
      .data$sample_scenario,
      all_available = "All available",
      paired_common_sample = "Placement-matched"
    )
  ) |>
  arrange(
    factor(.data$Dataset, c("Primary dataset", "Gap-timing-unaware dataset")),
    factor(.data$Placement, c("Near eye", "Chest")),
    factor(.data$Sample, c("All available", "Placement-matched"))
  ) |>
  select(
    .data$Dataset,
    .data$Placement,
    .data$Sample,
    .data$Metrics,
    .data$`Participants (range)`,
    .data$`Participant-days (range)`,
    .data$Sites
  )

scenario_summary |>
  gt::gt(groupname_col = "Dataset") |>
  gt::cols_width(
    Placement ~ gt::pct(14),
    Sample ~ gt::pct(20),
    Metrics ~ gt::pct(9),
    `Participants (range)` ~ gt::pct(22),
    `Participant-days (range)` ~ gt::pct(25),
    Sites ~ gt::pct(10)
  ) |>
  prep_gt()
```

### Exact primary fitted samples

Participant-level dynamics models have one fitted outcome per participant;
their participant-day count records the days contributing to that summary.
Participant-day models have one fitted outcome per retained participant-day.
Support hours are reported only when the stored preparation carries a valid
metric-specific denominator.

```{r}
#| label: prepare-h01-sample-source
#| include: false

primary_sample_source <- exact_samples |>
  filter(.data$run_id %in% c(primary_near_id, primary_chest_id)) |>
  transmute(
    run_id = .data$run_id,
    placement = unname(placement_labels[.data$placement]),
    metric_order = .data$metric_order,
    metric_id = .data$metric_id,
    metric = .data$manuscript_name,
    analysis_unit = .data$analysis_unit_label,
    participants = .data$participants,
    participant_days = .data$participant_days,
    observations = .data$observations,
    support_hours = .data$derivation_support_hours,
    sites = .data$sites,
    status = .data$sample_status
  ) |>
  arrange(.data$metric_order, factor(.data$placement, c("Near eye", "Chest")))

readr::write_csv(
  primary_sample_source,
  file.path(source_dir, "H01_preparation_fitted_sample_support.csv")
)

primary_sample_table <- function(placement_name) {
  primary_sample_source |>
    filter(.data$placement == .env$placement_name) |>
    transmute(
      Metric = .data$metric,
      `Model unit` = .data$analysis_unit,
      Participants = .data$participants,
      `Participant-days` = .data$participant_days,
      Observations = .data$observations,
      `Support hours` = if_else(
        is.finite(.data$support_hours),
        formatC(.data$support_hours, format = "f", digits = 1, big.mark = ","),
        "Unavailable"
      ),
      Sites = .data$sites
    )
}
```

::: {.panel-tabset}

#### Near eye

```{r}
#| label: tbl-h01-prep-primary-near-samples
#| tbl-cap: "Exact fitted sample for each primary near-eye outcome."

primary_sample_table("Near eye") |>
  gt::gt(rowname_col = "Metric") |>
  gt::fmt_number(
    columns = c(Participants, `Participant-days`, Observations, Sites),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Metric ~ gt::pct(34),
    `Model unit` ~ gt::pct(14),
    Participants ~ gt::pct(11),
    `Participant-days` ~ gt::pct(13),
    Observations ~ gt::pct(11),
    `Support hours` ~ gt::pct(12),
    Sites ~ gt::pct(5)
  ) |>
  prep_gt()
```

#### Chest

```{r}
#| label: tbl-h01-prep-primary-chest-samples
#| tbl-cap: "Exact fitted sample for each complementary chest outcome."

primary_sample_table("Chest") |>
  gt::gt(rowname_col = "Metric") |>
  gt::fmt_number(
    columns = c(Participants, `Participant-days`, Observations, Sites),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Metric ~ gt::pct(34),
    `Model unit` ~ gt::pct(14),
    Participants ~ gt::pct(11),
    `Participant-days` ~ gt::pct(13),
    Observations ~ gt::pct(11),
    `Support hours` ~ gt::pct(12),
    Sites ~ gt::pct(5)
  ) |>
  prep_gt()
```

:::

```{r}
#| label: fig-h01-prep-sample-support
#| fig-cap: "Contributing participant-days for each primary near-eye and complementary chest model. Participant-level dynamics outcomes use all contributing days to form one fitted observation per participant."
#| fig-alt: "A horizontal dot plot shows contributing participant-days for all 17 metrics. Near-eye values range from 655 to 816 days and chest values from 743 to 902 days, with lower support for several sleep, wake, and timing metrics."
#| fig-width: 10.5
#| fig-height: 7.8

sample_plot <- primary_sample_source |>
  mutate(
    metric = factor(.data$metric, levels = rev(unique(.data$metric))),
    placement = factor(.data$placement, levels = c("Near eye", "Chest"))
  )

ggplot(
  sample_plot,
  aes(
    x = .data$participant_days,
    y = .data$metric,
    colour = .data$placement,
    shape = .data$placement
  )
) +
  geom_point(size = 2.7, position = position_dodge(width = 0.45)) +
  scale_colour_manual(values = placement_colors) +
  scale_shape_manual(values = placement_shapes) +
  scale_x_continuous(labels = label_number(big.mark = ",")) +
  labs(
    x = "Contributing participant-days",
    y = NULL,
    colour = "Placement",
    shape = "Placement"
  ) +
  theme_minimal(base_size = 11.5) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 9)
  )
```

[Source data for @fig-h01-prep-sample-support](H01_analysis_preparation_files/source-data/H01_preparation_fitted_sample_support.csv)

### Retention from prepared rows to fitted rows

```{r}
#| label: prepare-h01-retention-source
#| include: false

retention_source <- main_sample_flow |>
  filter(
    .data$data_scenario_id == "main",
    .data$scenario == "all_available",
    .data$scope == "overall",
    .data$stage %in% c("prepared_rows", "site_photoperiod_model"),
    .data$status == "AVAILABLE"
  ) |>
  select(
    .data$placement,
    .data$metric_order,
    .data$metric_id,
    .data$stage,
    .data$model_observations
  ) |>
  pivot_wider(names_from = .data$stage, values_from = .data$model_observations) |>
  left_join(
    metric_contract |>
      select(.data$metric_id, metric = .data$manuscript_name),
    by = "metric_id"
  ) |>
  mutate(
    placement = unname(placement_labels[.data$placement]),
    retention = .data$site_photoperiod_model / .data$prepared_rows
  ) |>
  arrange(.data$metric_order, factor(.data$placement, c("Near eye", "Chest")))

readr::write_csv(
  retention_source,
  file.path(source_dir, "H01_preparation_model_frame_retention.csv")
)
```

```{r}
#| label: fig-h01-prep-model-frame-retention
#| fig-cap: "Percentage of prepared metric rows retained in the complete site-plus-photoperiod model frame. Losses reflect outcome-specific availability and complete-case predictors; they are not post-fit exclusions."
#| fig-alt: "A horizontal dot plot shows model-frame retention for all 17 outcomes. Many outcomes retain all prepared rows, while sleep, wake, dose, and timing metrics retain smaller percentages. Near-eye and chest patterns are similar but not identical."
#| fig-width: 10.5
#| fig-height: 7.8

retention_plot <- retention_source |>
  mutate(
    metric = factor(.data$metric, levels = rev(unique(.data$metric))),
    placement = factor(.data$placement, levels = c("Near eye", "Chest"))
  )

ggplot(
  retention_plot,
  aes(
    x = .data$retention,
    y = .data$metric,
    colour = .data$placement,
    shape = .data$placement
  )
) +
  geom_vline(xintercept = 1, colour = "grey55", linetype = "dashed") +
  geom_point(size = 2.7, position = position_dodge(width = 0.45)) +
  scale_colour_manual(values = placement_colors) +
  scale_shape_manual(values = placement_shapes) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0.75, 1.005),
    breaks = seq(0.75, 1, by = 0.05)
  ) +
  labs(
    x = "Prepared rows retained in fitted model",
    y = NULL,
    colour = "Placement",
    shape = "Placement"
  ) +
  theme_minimal(base_size = 11.5) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(size = 9)
  )
```

[Source data for @fig-h01-prep-model-frame-retention](H01_analysis_preparation_files/source-data/H01_preparation_model_frame_retention.csv)

### Site-specific support

The complete per-metric, per-site sample record is stored in
[`H01_exact_samples_by_site.csv`](../../../artifacts/09_tables/H01/H01_exact_samples_by_site.csv).
The compact table below shows the range across the 17 outcomes without hiding
that some outcome-specific model frames are smaller.

```{r}
#| label: tbl-h01-prep-site-sample-ranges
#| tbl-cap: "Range of fitted sample support across the 17 primary-dataset outcomes at each site."

site_sample_ranges <- exact_samples_by_site |>
  filter(.data$run_id %in% c(primary_near_id, primary_chest_id)) |>
  left_join(
    site_registry |>
      select(.data$site, .data$display_name, .data$display_order),
    by = "site"
  ) |>
  group_by(.data$placement, .data$display_order, .data$display_name) |>
  summarise(
    Metrics = n_distinct(.data$metric_id),
    `Participants (range)` = format_range(
      min(.data$participants, na.rm = TRUE),
      max(.data$participants, na.rm = TRUE)
    ),
    `Participant-days (range)` = format_range(
      min(.data$participant_days, na.rm = TRUE),
      max(.data$participant_days, na.rm = TRUE)
    ),
    `Observations (range)` = format_range(
      min(.data$observations, na.rm = TRUE),
      max(.data$observations, na.rm = TRUE)
    ),
    .groups = "drop"
  ) |>
  mutate(Placement = unname(placement_labels[.data$placement])) |>
  arrange(factor(.data$Placement, c("Near eye", "Chest")), .data$display_order) |>
  select(
    .data$Placement,
    Site = .data$display_name,
    .data$Metrics,
    .data$`Participants (range)`,
    .data$`Participant-days (range)`,
    .data$`Observations (range)`
  )

site_sample_ranges |>
  gt::gt(groupname_col = "Placement", rowname_col = "Site") |>
  gt::cols_width(
    Site ~ gt::pct(22),
    Metrics ~ gt::pct(10),
    `Participants (range)` ~ gt::pct(22),
    `Participant-days (range)` ~ gt::pct(24),
    `Observations (range)` ~ gt::pct(22)
  ) |>
  prep_gt()
```

## 4. Model construction

### Exact formula set

The implementation uses ordinary Gaussian linear models for participant-level
outcomes and Gaussian linear mixed models or Tweedie generalized mixed models
for participant-day outcomes. Participant-day models include a participant
random intercept. Site and latitude models are fitted separately so that a
nine-level site factor is never interpreted as a single latitude coefficient.

```{r}
#| label: tbl-h01-prep-formulas
#| tbl-cap: "Exact Wilkinson formulas supplied to the H01 fitted and comparison models."

formula_manifest |>
  distinct(.data$analysis_unit, .data$model_name, .data$formula) |>
  mutate(
    `Model unit` = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    Model = str_to_sentence(str_replace_all(.data$model_name, "_", " ")),
    Formula = paste0("`", str_squish(.data$formula), "`")
  ) |>
  select(.data$`Model unit`, .data$Model, .data$Formula) |>
  gt::gt(groupname_col = "Model unit") |>
  gt::fmt_markdown(columns = "Formula") |>
  gt::cols_width(Model ~ gt::pct(22), Formula ~ gt::pct(78)) |>
  prep_gt()
```

The complete site model estimates site and photoperiod jointly. The reduced
site model removes site; the reduced photoperiod model removes photoperiod.
The latitude model substitutes centred absolute latitude, expressed per 10°,
for the site factor while retaining the same photoperiod term. All comparison
fits use maximum likelihood; final Gaussian mixed models are refitted with
restricted maximum likelihood for coefficient and variance reporting.

### Central settings

```{r}
#| label: tbl-h01-prep-model-settings
#| tbl-cap: "Model settings shared across the H01 analysis scenarios."

tibble::tribble(
  ~Setting, ~Value, ~Contribution,
  "Primary placement", "Near eye", "Keeps measurements nearer the eye primary; chest remains separate complementary evidence.",
  "Predictor centres", "Recorded separately within the permitted analysis frame", "Makes zero correspond to the recorded analysis-specific mean while preserving original photoperiod and latitude fields.",
  "Participant-level engine", "lm, maximum likelihood", "Fits one outcome per participant without a redundant participant random effect.",
  "Gaussian participant-day engine", "lmer; ML for comparisons, REML for final estimates", "Accounts for repeated participant-days with a participant random intercept.",
  "Tweedie participant-day engine", "glmmTMB, log link, maximum likelihood", "Handles non-negative continuous outcomes with a mass at zero without silently changing the response family.",
  "Site marginalisation", "Equal weight for every represented site", "Defines the overall site mean and prevents large sites from dominating the site contrasts.",
  "Confidence level", "95%", "Used for model effects, site contrasts, R² summaries, and sensitivity estimates.",
  "Bootstrap target", "At least 1,000 successful joint refits per estimable run and metric", "Propagates joint refit uncertainty into marginal R², conditional R², participant-associated share, and term part-R².",
  "Model implementation", input_contract$model_implementation_id, "The identical implementation is used for the primary and gap-timing-unaware datasets."
) |>
  gt::gt() |>
  gt::cols_width(
    Setting ~ gt::pct(22),
    Value ~ gt::pct(32),
    Contribution ~ gt::pct(46)
  ) |>
  prep_gt()
```

## 5. Estimands, uncertainty, and multiplicity

```{r}
#| label: tbl-h01-prep-estimands
#| tbl-cap: "H01 estimands and the operation that produces each reported quantity."

tibble::tribble(
  ~Question, ~Estimand, ~Comparison, ~Reported_scale,
  "Overall site", "Joint contribution of the represented site factor conditional on photoperiod", "Complete site model versus model without site", "One model-level p-value per metric",
  "Photoperiod", "Change per hour of photoperiod conditional on site", "Complete site model versus model without photoperiod", "Difference, ratio, or odds ratio with 95% CI",
  "Latitude", "Change per 10° absolute latitude conditional on photoperiod", "Latitude model versus model without latitude", "Difference, ratio, or odds ratio with 95% CI",
  "Site versus latitude adequacy", "Whether the site factor retains information beyond a linear absolute-latitude term in the same frame", "Complete site model versus latitude model", "One model-level p-value per metric",
  "Site-specific deviation", "Each site minus or divided by the equally weighted overall site mean", "Hierarchical emmeans follow-up only after a supported overall site test", "Difference or ratio with 95% CI and within-metric adjustment",
  "Variation represented", "Marginal R², conditional R², participant-associated share, and non-overlapping term part-R²", "Stored joint bootstrap refits", "Estimate with 95% bootstrap interval"
) |>
  gt::gt() |>
  gt::cols_width(
    Question ~ gt::pct(19),
    Estimand ~ gt::pct(34),
    Comparison ~ gt::pct(29),
    Reported_scale ~ gt::pct(18)
  ) |>
  prep_gt()
```

The four model-level questions form four separate complete 17-test
Benjamini–Hochberg families. Adjustment is never performed over only the
metrics that happened to have small raw p-values.

```{r}
#| label: tbl-h01-prep-multiplicity
#| tbl-cap: "Multiplicity families fixed for the H01 response package."

family_registry |>
  transmute(
    Order = .data$family_order,
    Family = .data$family_label,
    Comparison = str_to_sentence(str_replace_all(.data$comparison, "_", " ")),
    Method = .data$adjustment_method,
    Tests = .data$family_n
  ) |>
  gt::gt() |>
  gt::fmt_number(columns = c(Order, Tests), decimals = 0) |>
  gt::cols_width(
    Order ~ gt::pct(8),
    Family ~ gt::pct(28),
    Comparison ~ gt::pct(39),
    Method ~ gt::pct(12),
    Tests ~ gt::pct(13)
  ) |>
  prep_gt()
```

## 6. Bootstrap, diagnostics, and influence

### Production joint refits

```{r}
#| label: tbl-h01-prep-bootstrap-audit
#| tbl-cap: "Stored production-bootstrap audit across the eight H01 analysis scenarios."

bootstrap_audit |>
  group_by(.data$run_id) |>
  summarise(
    Metrics = n_distinct(.data$metric_id),
    `Minimum successful refits` = min(.data$successful_refits, na.rm = TRUE),
    `Maximum failed refits` = max(.data$failed_refits, na.rm = TRUE),
    `Maximum warning refits` = max(.data$warning_refits, na.rm = TRUE),
    Status = if_else(all(.data$status == "PASS"), "PASS", "CHECK"),
    .groups = "drop"
  ) |>
  left_join(
    run_registry |>
      select(.data$run_id, .data$data_scenario_id, .data$placement, .data$sample_scenario),
    by = "run_id"
  ) |>
  mutate(
    Dataset = recode(
      .data$data_scenario_id,
      main = "Primary dataset",
      manuscript_prepared_data = "Gap-timing-unaware dataset"
    ),
    Placement = unname(placement_labels[.data$placement]),
    Sample = recode(
      .data$sample_scenario,
      all_available = "All available",
      paired_common_sample = "Placement-matched"
    )
  ) |>
  arrange(
    factor(.data$Dataset, c("Primary dataset", "Gap-timing-unaware dataset")),
    factor(.data$Placement, c("Near eye", "Chest")),
    factor(.data$Sample, c("All available", "Placement-matched"))
  ) |>
  select(
    .data$Dataset,
    .data$Placement,
    .data$Sample,
    .data$Metrics,
    .data$`Minimum successful refits`,
    .data$`Maximum failed refits`,
    .data$`Maximum warning refits`,
    .data$Status
  ) |>
  gt::gt(groupname_col = "Dataset") |>
  gt::fmt_number(
    columns = c(
      Metrics,
      `Minimum successful refits`,
      `Maximum failed refits`,
      `Maximum warning refits`
    ),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Placement ~ gt::pct(13),
    Sample ~ gt::pct(18),
    Metrics ~ gt::pct(8),
    `Minimum successful refits` ~ gt::pct(19),
    `Maximum failed refits` ~ gt::pct(15),
    `Maximum warning refits` ~ gt::pct(16),
    Status ~ gt::pct(11)
  ) |>
  prep_gt()
```

The refit draws are stored separately for every run and metric. They are not
recreated for table or figure revisions. The bootstrap outputs and the
production execution record are covered by
`H01_response_family_bootstrap_production_artifacts.csv` and the larger model
result manifest.

### Diagnostic and influence chain

```{r}
#| label: tbl-h01-prep-diagnostic-map
#| tbl-cap: "Stored diagnostic outputs and the question each addresses."

tibble::tribble(
  ~Output, ~Question_answered, ~Interpretive_limit,
  "H01_model_diagnostics.csv", "Did each fit converge, pass its family-specific residual checks, and respect physical prediction bounds?", "Flags require response-specific interpretation; a warning is not automatically a failed model.",
  "Metric-specific diagnostic plot data and PNG files", "What residual shape, fitted-value pattern, and simulated-diagnostic behaviour produced a model-level assessment?", "Representative plots illustrate the stored fit and do not re-estimate it.",
  "H01_participant_influence.csv", "How much did fixed-effect estimates change when each participant was omitted and the model refitted?", "A large dfbeta indicates sensitivity to a participant, not necessarily data error.",
  "H01_latitude_leave_one_site_out.csv", "Did the latitude estimate depend strongly on one represented site?", "Only eight or nine sites are available, so leave-one-site-out changes both range and leverage.",
  "H01_r2_bootstrap_audit.csv", "Were the required successful joint refits obtained for every estimable target?", "Assesses execution and interval support, not whether the model is scientifically correct.",
  "Final response-family diagnostic evidence", "Did the retained final response families satisfy the common diagnostic screen in both prepared datasets and placements?", "Documents the final response-family decision; it is not a new fit during rendering."
) |>
  gt::gt() |>
  gt::cols_width(
    Output ~ gt::pct(28),
    Question_answered ~ gt::pct(40),
    Interpretive_limit ~ gt::pct(32)
  ) |>
  prep_gt()
```

The results report shows the corresponding diagnostic assessments and
representative plots. This page records where those assessments came from.
For the two primary all-available placements, the stored diagnostic table
contains `r nrow(diagnostics |> filter(.data$run_id %in% c(primary_near_id, primary_chest_id)))`
metric-level records.

## 7. Sensitivity branches

```{r}
#| label: tbl-h01-prep-sensitivity-map
#| tbl-cap: "H01 sensitivity analyses and what each branch changes."

tibble::tribble(
  ~Branch, ~Changed, ~Held_constant, ~Stored_evidence,
  "Complementary chest", "Measurement placement", "Metric definitions, response-family registry, formulas, multiplicity, and reporting scales", "Chest models, tables, diagnostics, and source data",
  "Placement-matched sample", "Restricts each metric to observations available for both near-eye and chest placement", "Separate placement-specific models and identical model structure", "Paired/common-sample models and paired placement table/figure",
  "Gap-timing-unaware dataset", "Metric preparation does not use the timing of remaining gaps for an additional metric-specific adjustment", "General 50%-per-hour and 80%-per-day coverage, implementation, formulas, and multiplicity", "Eight same-model sensitivity runs",
  "Noon cut for darkest-10-hour midpoint", "Subtracts 24 hours strictly after noon rather than strictly after 16:00", "Outcome records, model implementation, and response family", "Stored noon-cut model frames, estimates, and support decisions",
  "Exactly identified light periods", "Restricts the longest continuous period metric to exactly identified periods", "Response family, predictors, model implementation, and reporting scale", "Stored period sensitivity models and comparison table",
  "Preregistered predictor scope", "Omits the added photoperiod covariate where it was not in the preregistered formula", "Metric frames, site coding, latitude coding, response families", "Stored scope-sensitivity models and comparison table",
  "Equal-site marginalisation", "Uses equal rather than observed site weights", "Fitted models and contrasts", "Stored marginalisation comparison",
  "Participant and site influence", "Deletes one participant or one site at a time", "Outcome and model specification", "Stored dfbeta and leave-one-site-out results"
) |>
  gt::gt() |>
  gt::cols_width(
    Branch ~ gt::pct(21),
    Changed ~ gt::pct(30),
    Held_constant ~ gt::pct(31),
    Stored_evidence ~ gt::pct(18)
  ) |>
  prep_gt()
```

## 8. Code map and execution order

### Reusable modules

```{r}
#| label: tbl-h01-prep-module-map
#| tbl-cap: "H01 function modules used by the executable analysis scripts."

tibble::tribble(
  ~Module, ~Responsibility, ~Why_separate,
  "h01_contract.R", "Seventeen-metric registry, four multiplicity families, eight-run registry, pinned inputs, exact formulas, and reproducible seeds", "Provides one declared contract to every H01 operation.",
  "h01_modeling.R", "Response transformation, model-frame construction, lm/lmer/glmmTMB fitting, comparisons, emmeans contrasts, diagnostics, influence, and R² extraction", "Applies the same scientific implementation to every dataset, placement, and sample.",
  "paths_io.R", "Atomic artifact writing and SHA-256 calculation", "Keeps file identity and scientific modelling responsibilities distinct.",
  "p_value_display.R", "Three-decimal reader-facing p-value formatting", "Separates display formatting from full-precision inference."
) |>
  gt::gt() |>
  gt::cols_width(
    Module ~ gt::pct(21),
    Responsibility ~ gt::pct(47),
    Why_separate ~ gt::pct(32)
  ) |>
  prep_gt()
```

### Executable entry points

```{r}
#| label: tbl-h01-prep-script-map
#| tbl-cap: "Executable H01 scripts in analytical order."

tibble::tribble(
  ~Order, ~Entry_point, ~Contribution, ~Main_outputs, ~Run_when_page_renders,
  1, "run_h01_models.R", "Reads the verified H01 bundles, constructs response-specific frames, fits the declared model set, runs diagnostics and registered sensitivities, and writes model/result artifacts.", "Models, diagnostics, exact samples, effects, contrasts, R² draws, and model-results manifest", "No",
  2, "assess_h01_response_family_candidates.R", "Applies the common diagnostic screen to the response-family review and records the final response-family selection evidence.", "Response-family evidence and selection records", "No",
  3, "archive_h01_pre_sleep_tweedie_baseline.R", "Preserves the replaced pre-sleep response-family outputs before the selected identity-scale implementation is written.", "Checksummed superseded-output archive", "No",
  4, "run_h01_response_family_bootstrap_production.R", "Completes the production joint-refit bootstrap for the selected pre-sleep response family without changing unaffected draw files.", "Production draws, checkpoints, audit, provenance, and manifest", "No",
  5, "build_h01_reporting_inputs.R", "Assembles audited model results into stable reporting tables and source data.", "Reporting tables, figures, source data, and manifest", "No",
  6, "build_h01_stage3_reporting_inputs.R", "Builds the standalone result-report tables, diagnostic displays, paired-placement display, and accessibility metadata from stored outputs.", "Reader tables, figures, source data, and reporting manifest", "No",
  7, "notebooks/hypotheses/H01.qmd", "Assembles the standalone H01 scientific results report from verified reporting artifacts.", "H01.html", "No",
  8, "H01_analysis_preparation.qmd", "Explains and displays the data-to-result chain shown on this page.", "H01_analysis_preparation.html and two descriptive source-data CSVs", "Yes",
  9, "build_h01_preparation_report_manifest.R", "Records source, rendered output, displayed dependencies, and environment identities.", "H01_preparation_report_artifacts.csv", "No",
  10, "build_h01_worker_manifest.R", "Builds the final non-circular H01 file inventory.", "H01_worker_artifacts.csv", "No"
) |>
  gt::gt() |>
  gt::cols_width(
    Order ~ gt::pct(6),
    Entry_point ~ gt::pct(23),
    Contribution ~ gt::pct(36),
    Main_outputs ~ gt::pct(26),
    Run_when_page_renders ~ gt::pct(9)
  ) |>
  prep_gt()
```

### Commands

The principal commands were run from the repository root in fresh R 4.6.1
sessions using the activated project library. They are shown for
reproducibility; this page does not execute the computationally intensive
commands.

```bash
Rscript --vanilla scripts/hypotheses/H01/run_h01_models.R
Rscript --vanilla scripts/hypotheses/H01/assess_h01_response_family_candidates.R
Rscript --vanilla scripts/hypotheses/H01/archive_h01_pre_sleep_tweedie_baseline.R
Rscript --vanilla scripts/hypotheses/H01/run_h01_response_family_bootstrap_production.R
Rscript --vanilla scripts/hypotheses/H01/build_h01_reporting_inputs.R
Rscript --vanilla scripts/hypotheses/H01/build_h01_stage3_reporting_inputs.R
quarto --profile nathealth render notebooks/hypotheses/H01.qmd --to html
quarto --profile nathealth render \
  audit/hypotheses/H01/H01_analysis_preparation.qmd --to html
Rscript --vanilla scripts/hypotheses/H01/build_h01_preparation_report_manifest.R
Rscript --vanilla scripts/environment/run_renv_status_safe.R <project-root> 60 true
Rscript --vanilla scripts/hypotheses/H01/build_h01_worker_manifest.R
```

## 9. Output provenance and file identities

The manifest chain records output paths, SHA-256 identities, byte sizes,
producing code, and R version. The table verifies the manifests that lead from
the H01 data bundles to the standalone report. The final H01 inventory is
built after the reader-facing pages so that it can contain their final
identities without creating a circular checksum dependency.

```{r}
#| label: tbl-h01-prep-manifests
#| tbl-cap: "Manifests linking H01 code to stored scientific and reporting outputs."

manifest_registry <- tibble::tribble(
  ~Component, ~Path,
  "Primary model-data bundle", main_manifest_path,
  "Gap-timing-unaware model-data bundle", gap_manifest_path,
  "Complete model and diagnostic output inventory", "artifacts/12_manifests/H01_model_results_artifacts.csv",
  "Production response-family bootstrap", bootstrap_manifest_path,
  "Audited reporting inputs", "artifacts/12_manifests/H01_reporting_artifacts.csv",
  "Standalone results-report inputs", stage3_manifest_path
) |>
  rowwise() |>
  mutate(
    Rows = nrow(read_h01_csv(.data$Path)),
    `Manifest SHA-256 prefix` = substr(
      artifact_sha256(file.path(root, .data$Path)),
      1L,
      16L
    )
  ) |>
  ungroup()

manifest_registry |>
  gt::gt() |>
  gt::fmt_number(columns = "Rows", decimals = 0, use_seps = TRUE) |>
  gt::cols_width(
    Component ~ gt::pct(31),
    Path ~ gt::pct(48),
    Rows ~ gt::pct(8),
    `Manifest SHA-256 prefix` ~ gt::pct(13)
  ) |>
  prep_gt()
```

## 10. How to examine or reproduce H01

A reader can inspect the chain in four passes:

1. compare `h01_contract.R` with the two pinned model-data manifests to confirm
   the 17 responses, eight runs, transformations, and formulas;
2. compare the sample and retention displays above with `model_rows.csv`,
   `sample_flow.csv`, and `H01_exact_samples_by_site.csv`;
3. follow the producing script and SHA-256 identity in each manifest, using
   the stored model, bootstrap, diagnostic, table, and source-data files; and
4. read the [H01 results report](../../../notebooks/hypotheses/H01.html)
   alongside its full-precision source tables rather than inferring numeric
   values from figure pixels.

Changes to shared metric preparation must be made upstream and propagated to
every affected hypothesis. A display-only change can instead be made from the
stored reporting objects without refitting the models or repeating the
bootstrap.

::: {.callout-tip title="Reader takeaway"}
The H01 result comes from two verified model-data bundles, a single declared
17-response implementation, exact response-specific model frames, four
complete multiplicity families, stored joint-refit bootstrap draws, and
separate diagnostic and sensitivity records. This page connects those pieces
without changing the scientific result.
:::

## Environment

```{r}
#| label: tbl-h01-prep-environment
#| tbl-cap: "Environment used to render this page."

quarto_version <- tryCatch(
  system2("quarto", "--version", stdout = TRUE, stderr = FALSE)[1L],
  error = function(error) "Recorded by render command"
)

tibble::tibble(
  Component = c(
    "R", "dplyr", "ggplot2", "gt", "lme4", "glmmTMB", "emmeans",
    "Quarto", "renv.lock SHA-256 prefix"
  ),
  Version = c(
    as.character(getRversion()),
    as.character(packageVersion("dplyr")),
    as.character(packageVersion("ggplot2")),
    as.character(packageVersion("gt")),
    as.character(packageVersion("lme4")),
    as.character(packageVersion("glmmTMB")),
    as.character(packageVersion("emmeans")),
    quarto_version,
    substr(artifact_sha256(file.path(root, "renv.lock")), 1L, 16L)
  )
) |>
  gt::gt() |>
  gt::cols_width(Component ~ gt::pct(65), Version ~ gt::pct(35)) |>
  prep_gt()
```
