---
title: "Alternative preprocessing analysis"
---

The primary analysis derives its metrics from the downloaded recordings using
explicit rules for missing intervals, temporal support, and measurement states.
This sensitivity analysis asks how the results change when the metrics instead
start from a fixed prepared baseline that is less explicit about gap timing.
The model specifications are held fixed within each hypothesis.

## Fixed local inputs

The six files in `data/alternative-baseline/` contain near-eye and chest data:

- `metrics_glasses.RData` and `metrics_chest.RData` contain participant and daily metrics.
- `metrics_separate_glasses.RData` and `metrics_separate_chest.RData` contain the prepared 30-minute values.
- `preprocessed_glasses_2.RData` and `preprocessed_chest_2.RData` contain the minute series used to derive hourly means and MDER.

[Preparation of the analysis datasets](../preparation/06-analysis-datasets.qmd)
imports these local inputs on every render. It checks their object structure,
applies the metric crosswalk, derives hourly means, and recalculates MDER as the
mean of positive, finite simultaneous minute ratios. MDER requires at least
720 viable ratios on a complete 1,440-minute local-clock grid. Other baseline
metrics retain their input values.

The prepared baseline is the input to this alternative analysis. The primary
analysis is reproduced from the downloaded recordings. No network download
is needed when either analysis is rendered.

## Compare metric values and sample support

The following comparison uses the model-ready inputs from the same render.
It reports available rows in each scenario and differences on their common
sample. Participant-level metrics use one row per participant; daily metrics
use one row per participant and date. Missing alternatives remain explicit.

```{r}
#| label: alternative-setup
library(dplyr)
library(tibble)
library(gt)
source("scripts/project.R")
analysis_setup()
source("scripts/hypotheses/H01/h01_contract.R")
source("scripts/hypotheses/H01/h01_modeling.R")
root <- getOption("nh.root")
inputs <- h01_input_contract(root)
primary <- readRDS(inputs$main$path)
alternative <- readRDS(inputs$alternative_preprocessing$path)
registry <- h01_metric_registry()
```

```{r}
#| label: alternative-metric-comparison
comparisons <- list()
for (placement in c("glasses", "chest")) {
  for (i in seq_len(nrow(registry))) {
    spec <- registry[i, , drop = FALSE]
    main_frame <- h01_prepare_model_frame(primary, spec, placement, "all_available")
    alternative_frame <- h01_prepare_model_frame(alternative, spec, placement, "all_available")
    common <- if (nrow(main_frame) && nrow(alternative_frame)) {
      inner_join(
        select(main_frame, site, participant_key, local_date, primary_value = value),
        select(alternative_frame, site, participant_key, local_date, alternative_value = value),
        by = c("site", "participant_key", "local_date"),
        relationship = "one-to-one")
    } else {
      tibble(primary_value = numeric(), alternative_value = numeric())
    }
    difference <- common$alternative_value - common$primary_value
    comparisons[[paste(placement, spec$metric_id)]] <- tibble(
      placement = placement, metric_id = spec$metric_id,
      analysis_unit = spec$analysis_unit,
      primary_rows = nrow(main_frame), alternative_rows = nrow(alternative_frame),
      common_rows = nrow(common),
      mean_alternative_minus_primary = if (length(difference)) mean(difference) else NA_real_,
      maximum_absolute_difference = if (length(difference)) max(abs(difference)) else NA_real_)
  }
}
comparison <- bind_rows(comparisons)
metric_labels <- readr::read_csv("config/metric_display_registry.csv",
  show_col_types = FALSE) |>
  select(metric_id, manuscript_name, display_unit)
comparison |>
  left_join(metric_labels, by = "metric_id", relationship = "many-to-one") |>
  mutate(
    placement = recode(placement, glasses = "Near eye", chest = "Chest"),
    analysis_unit = recode(analysis_unit,
      participant = "Participant", participant_day = "Participant-day"),
    display_unit = if_else(display_unit == "clock time", "min", display_unit)
  ) |>
  select(placement, manuscript_name, display_unit, analysis_unit,
    primary_rows, alternative_rows, common_rows,
    mean_alternative_minus_primary, maximum_absolute_difference) |>
  gt(groupname_col = "placement", rowname_col = "manuscript_name") |>
  cols_label(display_unit = "Unit", analysis_unit = "Analysis unit",
    primary_rows = "Primary rows", alternative_rows = "Alternative rows",
    common_rows = "Common rows", mean_alternative_minus_primary = "Mean difference",
    maximum_absolute_difference = "Maximum absolute difference") |>
  fmt_number(columns = c(mean_alternative_minus_primary, maximum_absolute_difference), decimals = 3)
```

Differences in available rows describe changes in the analysis sample.
Differences on common rows describe changes in the untransformed metric values,
calculated as alternative minus primary. Timing values are stored as minutes
since midnight, so their differences use that linear scale, not the shortest
circular distance across midnight. These summaries alone do not assess the
stability of model estimates. The individual
hypothesis pages fit both scenarios where the alternative metric is available
and report their estimates and uncertainty.

```{r}
#| label: alternative-export
#| code-fold: true
readr::write_csv(comparison,
  "results/csv/source_data/alternative_preprocessing_metric_comparison.csv")
```

[Download the metric and sample comparison](../results/csv/source_data/alternative_preprocessing_metric_comparison.csv)
for the full-precision values and metric identifiers.
