---
title: "Descriptive tables and figures"
subtitle: "Study sample, light exposure distributions, and temporal context"
engine: knitr
format:
  html:
    toc: true
    page-layout: full
execute:
  echo: true
---

The near-eye measurements provide the primary description of personal light
exposure. Chest measurements provide a complementary description. They are
summarised separately throughout.

## Libraries and prepared inputs

The preparation pages supply screened minute records, metric grids, participant
information, diary windows, and solar context. Shared functions implement the
specified summaries and figure styling; the calculations are called below in
the order that their results are used.

```{r}
#| label: descriptive-setup
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(gt)
if (!requireNamespace("rnaturalearthdata", quietly = TRUE)) {
  stop("Restore the project library with renv::restore(); the local Natural Earth map data are required.")
}
source("scripts/project.R")
analysis_setup()
root <- getOption("nh.root")
source("scripts/pipeline/paths_io.R")
source("scripts/pipeline/assertions.R")
source("scripts/pipeline/site_solar_context.R")
source("scripts/descriptives/descriptive_contract.R")
configure_descriptive_site_display(root)
source("scripts/descriptives/build_descriptive_data.R")
source("scripts/descriptives/build_descriptive_displays.R")
source("scripts/descriptives/build_publication_tables.R")
source("scripts/descriptives/plot_descriptive_figures.R")
source("scripts/descriptives/plot_descriptive_displays.R")
source("scripts/descriptives/sample_counts.R")
paths <- descriptive_paths(root)
for (path in paths[-1]) dir.create(path, recursive = TRUE, showWarnings = FALSE)
```

## Astronomical context and input validation

Calculate civil-photoperiod bounds across the 2025 calendar year at absolute latitudes from 0 to 60 degrees. These bounds provide context for the observed collection dates; they do not depend on fitted models.

```{r}
#| label: descriptive-solar-context
span_photoperiod <- tibble::tibble(
  Datetime = as.POSIXct("2025-01-01", tz = "UTC") +
    as.difftime(0:364, units = "days")
)
photoperiod_bounds <- dplyr::bind_rows(lapply(0:60, function(latitude) {
  values <- LightLogR::extract_photoperiod(
    span_photoperiod,
    c(latitude, 0)
  )
  tibble::tibble(
    absolute_latitude_deg = latitude,
    minimum_possible_photoperiod_hours = min(
      as.numeric(values$photoperiod),
      na.rm = TRUE
    ),
    maximum_possible_photoperiod_hours = max(
      as.numeric(values$photoperiod),
      na.rm = TRUE
    ),
    calendar_year = 2025L,
    solar_depression_deg = 6
  )
}))


inputs <- load_descriptive_inputs(root, photoperiod_bounds)
validate_descriptive_inputs(inputs)
write_descriptive_csv(photoperiod_bounds, file.path(paths$source_dir, "photoperiod_latitude_bounds.csv"))
photoperiod_bounds |> filter(absolute_latitude_deg %% 10 == 0) |> gt()
```

## Participants, collection dates, and coverage

Summarise the complete roster and the screened samples separately. Paired counts are calculated from participant-days observed at both sensor positions.

```{r}
#| label: descriptive-sample-calculations
collection_days <- build_collection_days(inputs)
available_collection_days <- build_available_collection_days(inputs)
site_sample <- build_site_sample_characteristics(inputs, collection_days, available_collection_days)
participant_characteristics <- build_participant_characteristics(inputs)
participant_table_data <- build_participant_site_display(inputs, site_sample, available_collection_days)
participant_manuscript_table_data <- participant_site_manuscript_data(participant_table_data)
sample_contract <- build_sample_count_contract(inputs)
site_display <- descriptive_site_display_registry()
write_descriptive_csv(sample_contract, file.path(paths$table_dir, "sample_counts.csv"))
sample_contract |> gt()
```

## Light metrics and daily profiles

Derive metric summaries with their analysis-specific denominators and circular timing summaries. The MDER calculation uses the explained alternative baseline and viable minute-ratio support. Profile summaries use the screened minute measurements.

```{r}
#| label: descriptive-light-calculations
metric_values <- build_metric_values(inputs)
metric_summary <- build_metric_summary(metric_values)
metric_plot_values <- build_metric_plot_values(metric_values)
metric_table_data <- build_metric_display(metric_summary)
profiles <- build_profile_sources(inputs)
recommendation <- build_recommendation_context(inputs)
recommendation_table_data <- build_recommendation_display(recommendation)
metric_summary |> select(placement, metric_id, n_participants, n_participant_days, n_observations) |> head()
```

## Sources for the illustrative figures

The seven-participant illustration uses the local alternative-preprocessing baseline and recomputes duration above 250 lx from the displayed daytime samples. The world map comes from the installed Natural Earth data package.

```{r}
#| label: descriptive-figure-data
time_series <- build_time_series_display_sources(root)
latitude_source <- build_latitude_photoperiod_source(collection_days)
collection_counts <- build_collection_date_counts(available_collection_days)
collection_intervals <- build_collection_intervals(available_collection_days, pause_days = 6L)
protocol_flow <- build_protocol_flow(inputs, collection_days)
site_locations <- build_site_location_source(inputs)
world_map <- build_world_map_source()
figure_alt <- build_display_figure_alt_text(site_sample, metric_summary, time_series, latitude_source)
time_series$selected |> gt()
```

## Save numerical results and figure data

The following CSV files contain the numerical tables and exact inputs to each figure. The editable publication tables are saved automatically when displayed below.

```{r}
#| label: descriptive-exports
table_outputs <- list(
  site_sample_characteristics = site_sample,
  participant_characteristics_primary = participant_characteristics,
  metric_distribution_summary = metric_summary,
  recommendation_context_near_eye = recommendation,
  participant_site_characteristics_display = participant_table_data,
  participant_site_characteristics_manuscript_display = participant_manuscript_table_data,
  metric_descriptive_summary_display = metric_table_data,
  recommendation_context_display = recommendation_table_data)
source_outputs <- list(
  collection_days = collection_days, available_collection_days = available_collection_days,
  collection_date_counts = collection_counts, collection_intervals = collection_intervals,
  protocol_flow = protocol_flow, site_locations = site_locations, world_map_wkt = world_map,
  profile_summary = profiles$profile, profile_context_bands = profiles$state,
  profile_average_periods = profiles$period, metric_plot_values = metric_plot_values,
  time_series_display_selection = time_series$selected,
  time_series_display_30_minute = time_series$series,
  time_series_display_states = time_series$states, time_series_display_metrics = time_series$metrics,
  latitude_photoperiod = latitude_source, figure_alt_text = figure_alt)
for (name in names(table_outputs)) write_descriptive_csv(table_outputs[[name]], file.path(paths$table_dir, paste0(name, ".csv")))
for (name in names(source_outputs)) write_descriptive_csv(source_outputs[[name]], file.path(paths$source_dir, paste0(name, ".csv")))

```

```{r}
#| label: descriptive-alt-text
get_alt <- function(figure_id) figure_alt$short_alt_text[match(figure_id, figure_alt$figure_id)]
```
The study sites are displayed in the registered reader order, with country
codes and fixed colours: `r paste(site_display$display_name, collapse = ", ")`.
The same labels, order, and colours are used throughout the tables and figures.

The normalized participant roster contains **191 participants**. After the
pre-specified ≥80% full-day completeness rule and exact-all-zero screen, the
primary near eye dataset contains **141 participants, 816 participant-days, and
1,175,160 one-minute real observations**. The complementary chest dataset
contains **154 participants, 902 participant-days, and 1,298,880 one-minute
real observations**. The paired subset contains **112 participants and 643
participant-days**.

## Data coverage and measurement placements

Two of 818 otherwise eligible near eye participant-days and three of 905
otherwise eligible chest participant-days were exact-all-zero melEDI days and
were excluded. A participant was not removed merely because one of their days
was excluded.

```{r}
#| label: tbl-descriptive-sample-flow
#| tbl-cap: "Descriptive sample after each fixed coverage step."

sample_contract_before_gt <- sample_contract
sample_contract_key_before_gt <- sample_contract |>
  dplyr::select(step, placement)

sample_flow_display <- sample_contract |>
  dplyr::mutate(
    dplyr::across(
      c(participants, participant_days, one_minute_real_observations),
      ~ ifelse(is.na(.x), "Not applicable", format(.x, big.mark = ","))
    )
  ) |>
  dplyr::rename(
    Step = step,
    Placement = placement,
    Participants = participants,
    `Participant-days` = participant_days,
    `One-minute real observations` = one_minute_real_observations
  )

sample_flow_gt <- sample_flow_display |>
  gt::gt(rowname_col = "Step", groupname_col = "Placement") |>
  gt::cols_align(
    align = "right",
    columns = c(
      Participants,
      `Participant-days`,
      `One-minute real observations`
    )
  ) |>
  gt::tab_options(
    table.width = gt::pct(100),
    table.font.size = gt::px(12),
    container.overflow.x = "auto"
  )

stopifnot(
  inherits(sample_flow_gt, "gt_tbl"),
  identical(sample_contract, sample_contract_before_gt),
  identical(
    sample_contract |> dplyr::select(step, placement),
    sample_contract_key_before_gt
  ),
  identical(dim(sample_flow_display), c(8L, 5L))
)

sample_flow_gt
```

Collection dates and civil photoperiod use the union of every recorded
non-all-zero participant-day across both placements. Screened-day counts retain
the primary-analysis completeness rule. Declared non-wear is reported for near
eye only. Participant information covers the complete normalized roster.

## Main descriptive figure and table

### Figure 1: study overview

The five-panel overview combines the study protocol, site map, collection
timing, photoperiod coverage, and a repeated 48-hour primary near eye profile.
The second 24-hour cycle is a display duplicate, not additional data. Collection
spans are shown as rectangles and split only when a site has at least six
consecutive dates without any available data.

```{r}
#| label: fig-descriptive-overview
#| fig-cap: "Study protocol, country-coded sites, collection dates, observed civil photoperiod, and the pooled primary near eye 48-hour profile."
#| fig-alt: "Five-panel study overview showing the protocol, country-coded study sites and coordinates, roster-wide collection dates, site photoperiod distributions, and a repeated 15-minute 48-hour primary near eye melEDI profile with central 50% and 90% value bands."
#| out-width: "100%"

plot <- make_overview_display_figure(file.path(root, "assets/study-protocol.png"), site_locations, world_map, collection_intervals, available_collection_days, profiles$profile, profiles$state, profiles$period)
save_descriptive_figure(plot, file.path(paths$figure_dir, "descriptive_overview"), width = 10.5, height = 10, dpi = 300, scale = 1.5)
include_project_graphics(file.path(paths$figure_dir, "descriptive_overview.png"))
```

Panel B shows country and coordinate information while retaining the registered
site names and colours. The black and site-coloured profiles use
`LightLogR::aggregate_Datetime()` in 15-minute floor bins after removing
participant grouping; the numeric handler is the median with `na.rm = TRUE`.
Nested grey ribbons show the central 50% and 90% of eligible one-minute values
in each overall bin, with direct arrow labels. They are value intervals, not
confidence intervals. The red band from the lower plot edge to −0.1 shows
average diary sleep, and blue shading shows average civil night; declared
non-wear is not displayed. For the non-negative melEDI values, the symlog axis
is linear from 0 to 1 lx and logarithmic above 1 lx; tick labels remain in lx.
The axis is capped above 10,000 lx to reserve less space for unsupported
extremes.

### Table 1: participant and site characteristics

@tbl-participant-site uses site columns and identifies whether each quantity
belongs to near eye, chest, paired, roster, or diary data.

```{r}
#| label: tbl-participant-site
#| tbl-cap: "Participant, site, collection, sleep, chronotype, and wear characteristics."

participant_gt <- build_participant_site_publication_gt(participant_table_data)
participant_gt
```

#### Reduced participant-characteristics table

The reduced participant-characteristics table provides the concise row
selection used alongside the full table. It omits Morningness–Eveningness
Questionnaire score, gender, collection dates, weekday/weekend counts,
workday/free-day diaries, social jetlag, and sleep duration; all remaining rows
retain their registered order and styling.

```{r}
#| label: tbl-participant-site-manuscript
#| tbl-cap: "Reduced participant and site characteristics."

participant_manuscript_gt <-
  build_participant_site_manuscript_publication_gt(participant_table_data)
participant_manuscript_gt
```

## Light distributions and temporal context

### Table 2: metric distributions

@tbl-near-eye-metrics presents the registered metric categories and order, ten
site columns, scaling column, and miniature distribution column. Every
numerical cell states `N`, the number of participants, and `d`, the number of
participant-days with a finite value. The site columns have equal widths and
each median, interval, mean ± SD, and size line is kept intact. A compact line
under each metric name states what the metric captures and why that construct
is physiologically or health-relevant; these notes are context, not health-
effect estimates from this descriptive dataset.

Duration and clock values are displayed to the nearest minute, with exact half-minute ties rounded to the even minute. Numerical CSV exports retain the full precision.

MDER is the arithmetic mean of viable one-minute melEDI / photopic illuminance
ratios, with both channels finite and strictly positive. Each day
is evaluated on a complete 1,440-minute local wall-clock grid and retained
with at least 720 viable ratios (the inclusive 50% rule). Fall-back duplicate
local minutes are averaged within channel before the ratio is formed;
spring-forward absent minutes remain missing.

For the L10 mean, source windows verified as entirely zero are retained as
exact 0 lx. Machine-precision residuals from the averaging calculation are
not interpreted as positive exposure.

```{r}
#| label: tbl-near-eye-metrics
#| tbl-cap: "Primary near eye melEDI metric distributions overall and by site."

metric_gt <- build_metric_publication_gt(metric_table_data, metric_plot_values)
metric_gt
```

The table displays the metric summaries calculated above.
Clock-time means and
middle-50% intervals are circular, rather than treating 23:59 and 00:01 as far
apart. The miniature plots are redundant visual summaries: the adjacent cells
remain the authoritative values and provide the exact denominators.

### Figure 2: site-specific profiles

The primary near eye profiles use a 3 × 3 double-plot layout and the registered
site order. Each site uses the same 15-minute
`aggregate_Datetime()` median with nested central 50%, 75%, and 95% value
intervals. The interval legend shows the corresponding ribbon opacities. The
red mean-sleep strip extends from the lower plot edge to −0.1.

```{r}
#| label: fig-near-eye-site-profiles
#| fig-cap: "Pooled primary near eye melEDI profiles by site, with each 24-hour cycle repeated once."
#| fig-alt: "Nine country-coded site panels repeat pooled 15-minute primary near eye 24-hour melEDI profiles across 48 hours with nested central 50%, 75%, and 95% value bands and average civil-night and sleep periods."
#| out-width: "100%"

plot <- make_site_profile_display_figure(profiles$profile, profiles$state, profiles$period, "near_eye")
save_descriptive_figure(plot, file.path(paths$figure_dir, "near_eye_site_profiles"), width = 6.692913385826772, height = 6.889763779527559, dpi = 450, scale = 1)
include_project_graphics(file.path(paths$figure_dir, "near_eye_site_profiles.png"))
```

The complementary chest profile uses the same visual grammar in a complete
four-by-two layout. Tübingen (DE) is omitted because the complementary chest
dataset contains no eligible MPI site panel; the remaining sites retain their
registered order and colours.

```{r}
#| label: fig-chest-site-profiles
#| fig-cap: "Complementary chest melEDI profiles by available site, with each 24-hour cycle repeated once."
#| fig-alt: "Eight country-coded site panels repeat complementary 15-minute chest 24-hour melEDI profiles across 48 hours with nested central 50%, 75%, and 95% value bands and average civil-night and sleep periods. Tübingen (DE) is absent because there are no eligible chest days at MPI."
#| out-width: "100%"

plot <- make_site_profile_display_figure(profiles$profile, profiles$state, profiles$period, "chest")
save_descriptive_figure(plot, file.path(paths$figure_dir, "chest_site_profiles"), width = 6.692913385826772, height = 5.511811023622047, dpi = 450, scale = 1)
include_project_graphics(file.path(paths$figure_dir, "chest_site_profiles.png"))
```

Chest values describe the environment at the chest sensor. They are not
treated as equivalent to near eye or ocular exposure.

### Figure 3: near eye metric distributions

@fig-near-eye-metric-distributions uses a 4 × 4 ridge-plus-box layout. Site
names remain visible in the first column of panels, so colour is not the only
way to identify a distribution.

```{r}
#| label: fig-near-eye-metric-distributions
#| fig-cap: "Sixteen primary near eye melEDI metric distributions by site."
#| fig-alt: "A four-by-four grid of duration, dynamics, dose, level, spectrum, and circular timing metric distributions. Each panel contains labelled site ridges and white boxplots; red lines mark site medians."
#| out-width: "100%"

plot <- make_metric_distributions_display_figure(metric_plot_values)
save_descriptive_figure(plot, file.path(paths$figure_dir, "near_eye_metric_distributions"), width = 6.692913385826772, height = 6.692913385826772, dpi = 450, scale = 1)
include_project_graphics(file.path(paths$figure_dir, "near_eye_metric_distributions.png"))
```

### Figure 4: from time series to a daily metric

@fig-time-series-to-metrics uses seven fixed German-site examples over five
aligned study days (days 2–6) and orders participants from top to bottom by
their median TAT250 across those days. The source CSV preserves each
participant’s actual local dates, while the visible x-axis uses a Wednesday to
Sunday alignment.

```{r}
#| label: fig-time-series-to-metrics
#| fig-cap: "Seven fixed examples link alternative-preprocessing 30-minute near eye melEDI series to daily duration above 250 lx during daytime."
#| fig-alt: "Four panels connect five protocol weekdays of alternative-preprocessing 30-minute near eye light samples for seven fixed participants to TAT250 values calculated from those displayed daytime samples, using participant boxplots, weekday boxplots, and a participant-by-weekday bubble grid."
#| out-width: "100%"

plot <- make_time_series_display_figure(time_series$series, time_series$states, time_series$metrics)
save_descriptive_figure(plot, file.path(paths$figure_dir, "time_series_to_metrics"), width = 6.692913385826772, height = 6.0236220472440944, dpi = 300, scale = 1)
include_project_graphics(file.path(paths$figure_dir, "time_series_to_metrics.png"))
```

Panel A uses floor-aligned 30-minute near eye arithmetic means from the local alternative-preprocessing baseline.
Panels B–D recalculate daily TAT250 from
exactly those displayed daytime samples with
`LightLogR::duration_above_threshold()`. This figure illustrates the fixed
examples and does not estimate a population weekday effect.

### Figure 5: latitude and observed photoperiod

The latitude–photoperiod figure provides supplementary context for the 816
primary near eye participant-days. Black ribbons mark civil-photoperiod
combinations outside the site-specific theoretical bounds. The density curves
and observations are drawn first and the black feasibility curtains are drawn
over them. Points use deterministic two-dimensional jitter to avoid artificial
bands without changing any underlying photoperiod or latitude.

```{r}
#| label: fig-latitude-photoperiod
#| fig-cap: "Observed civil photoperiod and theoretical bounds by absolute latitude for primary near eye participant-days."
#| fig-alt: "Observed civil photoperiod distributions are plotted against absolute study-site latitude for 816 primary near eye participant-days, with black ribbons marking impossible photoperiods and a country-coded, registered-colour legend identifying the nine sites."
#| out-width: "100%"

plot <- make_latitude_photoperiod_display_figure(latitude_source, photoperiod_bounds)
save_descriptive_figure(plot, file.path(paths$figure_dir, "latitude_photoperiod_diagnostic"), width = 6.692913385826772, height = 6.692913385826772, dpi = 300, scale = 1)
include_project_graphics(file.path(paths$figure_dir, "latitude_photoperiod_diagnostic.png"))
```

### Light exposure in Brown et al. recommendation windows

@tbl-recommendation-context uses site rows and recommendation-window columns.
The Brown et al. recommendation windows are Daytime ≥250 lx melanopic EDI,
Pre-sleep ≤10 lx melanopic EDI, and Sleep ≤1 lx melanopic EDI
[@brown2022]. The results are deliberately described as *fractions of valid
measured minutes in a recommended range*, not as adherence, compliance, or
evidence of a biological response.

```{r}
#| label: tbl-recommendation-context
#| tbl-cap: "Fractions of primary near eye minutes in the Brown et al. recommended ranges, with explicit denominators."

recommendation_gt <- build_recommendation_publication_gt(
  recommendation_table_data
)
recommendation_gt
```

Each percentage is followed by one grey numerator/denominator count. Daytime,
Pre-sleep, and Sleep cells divide minutes within the applicable recommended
range by all valid one-minute observations in that recommendation window; the
combined column pools all classified valid minutes. The four state-share
columns divide minutes in the displayed state by all eligible real minutes.
Diary-defined sleep combines the near eye sensor while awake with a bedside
sensor during sleep; the Sleep window therefore describes the bedside sleep
environment rather than direct ocular exposure.

## Data-quality and denominator qualifications

- Participant and participant-day displays distinguish the complete roster,
  primary near eye, complementary chest, and paired subsets. Participant time
  uses the screened primary near eye basis and is limited to its two largest
  time units.
- Site labels follow the country-coded registry, including Tübingen (DE) for
  MPI and Munich (DE) for TUM.
- Screened days report the result after the completeness and all-zero screens.
  Civil photoperiod uses the available site-and-solar context across the whole
  roster.
- Declared non-wear is available for near eye only. Chest values are retained
  solely as complementary environmental measurements.
- Clock-time chronotype and timing summaries are circular.
  Recommendation-window percentages retain their displayed state-specific or
  eligible-minute denominators.


## Reproduction outputs

`results/tables/descriptives/` contains the numerical summaries;
`results/csv/source_data/descriptives/` contains figure data; and
`results/images/descriptives/` contains the rendered figures.
The [preparation pages](../preparation/01-import-and-alignment.qmd) explain
the downloaded inputs, alignment, coverage rules, and metric derivation.
