---
title: "H08: Visual light sensitivity and personal light exposure"
engine: knitr
format:
  html:
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

Visual light sensitivity is related to participant-day light-exposure metrics, adjusting for site and accounting for repeated participant observations.

## Data and model guide {#methods}

The [questionnaire preparation](../preparation/05-questionnaires-and-diaries.qmd)
reconstructs VLSQ-8 from the eight ordered item codes and the recorded total-score
rule. The constant offset in the total is retained, although centring removes it
from a fitted slope. Scores are centred across participants, not within site;
reported effects compare one participant-level SD. Nine participant-day outcomes
come from the [metric datasets](../preparation/06-analysis-datasets.qmd), retaining
metric-specific support and missingness.

The response registry specifies four Gaussian models after `log10(value + 0.1)`,
three Tweedie log-link models and two Gaussian identity models. Site adjustment
and participant random intercepts account for site structure and repeated days.
Average associations and VLSQ-8-by-site interactions are separate questions with
complete nine-metric FDR families. Practical effects are ratios except for pre-sleep
time below 10 lx, which is expressed as an hour difference. Estimates, Wald intervals,
FDR decisions and model adequacy are shown separately.

Sensitivities address preprocessing, placement-matched data, exact identification
of the longest bright-light period, dose definition, participant summaries and
influence. A remaining gap can make the longest observed bright period a lower
bound. The participant-summary Tweedie model for time above 1,000 lx has a convergence
qualification; the corresponding Gamma boundary comparison is a diagnostic check,
not a replacement of the specified analysis.

The executable sections below write fitted objects to `results/models/H08/`,
reader tables to `results/tables/H08/`, and the supporting numerical and figure
outputs under `results/`. Start with the calculations below or jump to
[findings and interpretation](#findings).

## Setup

Load the reusable sample, formula, fit and diagnostic functions. The sections below perform each analysis and write its numerical results.

```{r}
#| label: setup-h08

source("scripts/project.R")
analysis_setup()
source("scripts/pipeline/paths_io.R")
source("scripts/pipeline/p_value_display.R")
source("scripts/hypotheses/H08/h08_contract.R")
source("scripts/hypotheses/H08/h08_modeling.R")
root <- normalizePath(Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()))
producer <- "analyses/H08-light-sensitivity.qmd"
options(stringsAsFactors = FALSE, scipen = 999, dplyr.summarise.inform = FALSE)
roots <- list(model_data = file.path(root,"results/intermediate/model_data/H08"), models = file.path(root,"results/models/H08"), diagnostics = file.path(root,"results/csv/diagnostics/H08"), tables = file.path(root,"results/tables/H08"), figures = file.path(root,"results/images/H08"), source_data = file.path(root,"results/csv/source_data/H08"))
invisible(lapply(roots, dir.create, recursive=TRUE, showWarnings=FALSE))
metadata <- list()
source("scripts/pipeline/multiplicity.R")
h08_write_csv <- function(data, path) {
  write_csv_artifact(data, path, producer = producer)
  invisible(path)
}

h08_write_rds <- function(object, path) {
  write_rds_artifact(object, path, producer = producer)
  invisible(path)
}
input_contract <- h08_input_contract(root)
```



## Questionnaire scoring and analysis samples

Use the sum of eight visual light sensitivity items, and construct metric-specific complete-case samples for each sensor, the paired placement comparison and the alternative preprocessing comparison.

```{r}
#| label: h08-data

metric_registry <- h08_metric_registry()

run_registry <- h08_run_registry()

family_registry <- h08_family_registry()

sensitivity_registry <- h08_sensitivity_registry()

formula_registry <- h08_formula_registry()

site_registry <- readr::read_csv(
  file.path(root, "config/site_display_registry.csv"),
  show_col_types = FALSE
) |>
  dplyr::arrange(.data$display_order)

site_levels <- site_registry$site

metric_display <- readr::read_csv(
  file.path(root, "config/metric_display_registry.csv"),
  show_col_types = FALSE
) |>
  dplyr::filter(.data$metric_id %in% metric_registry$metric_id) |>
  dplyr::select(
    .data$metric_id,
    display_manuscript_name = .data$manuscript_name,
    display_unit_registry = .data$display_unit
  )

metric_display_audit <- metric_registry |>
  dplyr::left_join(
    metric_display,
    by = "metric_id",
    relationship = "one-to-one"
  ) |>
  dplyr::mutate(
    name_matches = .data$manuscript_name == .data$display_manuscript_name,
    unit_matches = .data$display_unit == .data$display_unit_registry
  )

if (
  any(!metric_display_audit$name_matches | !metric_display_audit$unit_matches)
) {
  h08_abort("The H08 metric display contract differs from the shared registry")
}

h05_response <- readr::read_csv(
  file.path(root, "results/intermediate/model_data/H05/H05_metric_registry.csv"),
  show_col_types = FALSE
) |>
  dplyr::filter(.data$metric_id %in% metric_registry$metric_id) |>
  dplyr::select(
    .data$metric_id,
    inherited_response_family = .data$response_family,
    inherited_response_transform = .data$response_transform,
    inherited_effect_scale = .data$effect_scale
  )

response_contract_audit <- metric_registry |>
  dplyr::left_join(
    h05_response,
    by = "metric_id",
    relationship = "one-to-one"
  ) |>
  dplyr::mutate(
    response_family_matches = .data$response_family ==
      .data$inherited_response_family,
    response_transform_matches = .data$response_transform ==
      .data$inherited_response_transform,
    effect_scale_matches = .data$effect_scale == .data$inherited_effect_scale
  )

if (
  any(
    !response_contract_audit$response_family_matches |
      !response_contract_audit$response_transform_matches |
      !response_contract_audit$effect_scale_matches
  )
) {
  h08_abort(
    "The H08 response package differs from the shared H01/H05 response specification"
  )
}

vlsq <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "normalized_vlsq8"
])

score_contract <- h08_score_contract(vlsq)

item_names <- c(
  "sensitivity",
  "glare",
  "flicker",
  "sensitivity_severity",
  "headache",
  "blurry_vision",
  "ability",
  "glasses"
)

item_matrix <- do.call(cbind, lapply(vlsq[item_names], as.integer))

score_audit_rows <- vlsq |>
  dplyr::transmute(
    .data$site,
    .data$Id,
    stored_VLSQ8 = .data$VLSQ8,
    item_sum_1_to_5 = rowSums(item_matrix),
    stored_minus_item_sum = .data$VLSQ8 - rowSums(item_matrix),
    scoring_rule_verified = .data$stored_minus_item_sum == 5
  )

score_audit <- tibble::tibble(
  participants = nrow(vlsq),
  sites = dplyr::n_distinct(vlsq$site),
  missing_scores = sum(is.na(vlsq$VLSQ8)),
  missing_item_cells = sum(is.na(item_matrix)),
  observed_min = min(vlsq$VLSQ8),
  observed_max = max(vlsq$VLSQ8),
  observed_mean = mean(vlsq$VLSQ8),
  observed_participant_sd = stats::sd(vlsq$VLSQ8),
  stored_minus_item_sum_unique = paste(
    sort(unique(score_audit_rows$stored_minus_item_sum)),
    collapse = "|"
  ),
  scoring_rule = score_contract$scoring_rule,
  scoring_rule_verified = all(score_audit_rows$scoring_rule_verified),
  score_definition_checked = TRUE
)

if (
  score_audit$missing_scores != 0L ||
    score_audit$missing_item_cells != 0L ||
    any(!item_matrix %in% 1:5) ||
    !isTRUE(score_audit$scoring_rule_verified)
) {
  h08_abort("The H08 VLSQ-8 score input fails its scoring specification")
}

h08_write_csv(
  metric_registry,
  file.path(roots$model_data, "H08_metric_registry.csv")
)

h08_write_csv(run_registry, file.path(roots$model_data, "H08_run_registry.csv"))

h08_write_csv(
  family_registry,
  file.path(roots$model_data, "H08_family_registry.csv")
)

h08_write_csv(
  sensitivity_registry,
  file.path(roots$model_data, "H08_sensitivity_registry.csv")
)

h08_write_csv(
  formula_registry,
  file.path(roots$model_data, "H08_formula_registry.csv")
)

h08_write_csv(
  score_audit,
  file.path(roots$diagnostics, "H08_vlsq_score_audit.csv")
)

h08_write_csv(
  score_audit_rows,
  file.path(roots$diagnostics, "H08_vlsq_score_rows.csv")
)

h08_write_csv(
  response_contract_audit,
  file.path(roots$model_data, "H08_response_contract_audit.csv")
)

main_near_eye <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "primary_near_eye_metrics"
])

main_chest <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "primary_chest_metrics"
])

gap_source <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "gap_timing_unaware_metrics"
])

h01_main <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "primary_support_provenance"
])

h01_gap <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "gap_support_provenance"
])

h08_support_rows <- function(object, scenario_id) {
  object$model_rows |>
    dplyr::filter(
      .data$scenario == "all_available",
      .data$metric_id %in% metric_registry$metric_id
    ) |>
    dplyr::transmute(
      data_scenario_id = scenario_id,
      .data$placement,
      .data$site,
      .data$Id,
      local_date = as.Date(.data$local_date),
      .data$metric_id,
      h01_value = .data$value,
      .data$metric_estimable,
      .data$metric_failure_reason,
      .data$metric_support_available,
      .data$metric_support_unavailability_reason,
      .data$metric_support_valid_minutes,
      .data$metric_support_expected_minutes
    )
}

support_rows <- dplyr::bind_rows(
  h08_support_rows(h01_main, "main"),
  h08_support_rows(h01_gap, "gap_timing_unaware")
)

if (
  anyDuplicated(support_rows[c(
    "data_scenario_id",
    "placement",
    "site",
    "Id",
    "local_date",
    "metric_id"
  )])
) {
  h08_abort("H08 support provenance contains duplicate participant-day keys")
}

h08_main_long <- function(data, placement) {
  source_columns <- metric_registry$source_column
  data |>
    dplyr::select(
      .data$site,
      .data$Id,
      .data$local_date,
      .data$VLSQ8,
      .data$photoperiod_hours,
      .data$valid_medi_real_minutes,
      dplyr::all_of(source_columns),
      longest_exact_value = .data$longest_bout_above_250_exact_only_sensitivity_h,
      longest_exact_identifiable = .data$longest_bout_above_250_exact_identifiable,
      dose_observed_value = .data$dose_observed_medi_lx_h
    ) |>
    tidyr::pivot_longer(
      cols = dplyr::all_of(source_columns),
      names_to = "source_column",
      values_to = "value"
    ) |>
    dplyr::left_join(
      metric_registry,
      by = "source_column",
      relationship = "many-to-one"
    ) |>
    dplyr::mutate(
      data_scenario_id = "main",
      placement = placement,
      local_date = as.Date(.data$local_date),
      .before = 1L
    )
}

main_rows <- dplyr::bind_rows(
  h08_main_long(main_near_eye, "glasses"),
  h08_main_long(main_chest, "chest")
)

gap_rows <- gap_source |>
  dplyr::filter(.data$metric_id %in% metric_registry$metric_id) |>
  dplyr::transmute(
    data_scenario_id = "gap_timing_unaware",
    placement = .data$position,
    .data$site,
    .data$Id,
    local_date = as.Date(.data$local_date),
    .data$metric_id,
    value = .data$alternative_preprocessing_value,
    photoperiod_hours = NA_real_,
    valid_medi_real_minutes = NA_real_,
    longest_exact_value = NA_real_,
    longest_exact_identifiable = NA,
    dose_observed_value = NA_real_
  ) |>
  dplyr::left_join(
    vlsq |>
      dplyr::select(.data$site, .data$Id, .data$VLSQ8),
    by = c("site", "Id"),
    relationship = "many-to-one"
  ) |>
  dplyr::left_join(
    metric_registry,
    by = "metric_id",
    relationship = "many-to-one"
  )

model_rows <- dplyr::bind_rows(main_rows, gap_rows) |>
  dplyr::left_join(
    support_rows,
    by = c(
      "data_scenario_id",
      "placement",
      "site",
      "Id",
      "local_date",
      "metric_id"
    ),
    relationship = "one-to-one"
  )

if (nrow(model_rows) != nrow(main_rows) + nrow(gap_rows)) {
  h08_abort("H08 support join changed the model-row count")
}

if (any(is.na(model_rows$h01_value) != is.na(model_rows$value))) {
  h08_abort(
    "H08 direct inputs and H01 support provenance differ in missingness"
  )
}

finite_pair <- is.finite(model_rows$value) & is.finite(model_rows$h01_value)

value_difference <- abs(
  model_rows$value[finite_pair] - model_rows$h01_value[finite_pair]
)

if (length(value_difference) > 0L && max(value_difference) > 1e-10) {
  h08_abort(
    "H08 direct inputs and H01 support provenance differ in metric values"
  )
}

if (any(is.na(model_rows$VLSQ8))) {
  h08_abort("H08 model rows contain an unmatched VLSQ-8 score")
}

if (
  anyDuplicated(model_rows[c(
    "data_scenario_id",
    "placement",
    "site",
    "Id",
    "local_date",
    "metric_id"
  )])
) {
  h08_abort("H08 model rows contain duplicate participant-day keys")
}

missingness <- model_rows |>
  dplyr::mutate(
    availability_reason = dplyr::case_when(
      is.finite(.data$value) ~ "available",
      !is.na(.data$metric_failure_reason) &
        nzchar(.data$metric_failure_reason) ~
        .data$metric_failure_reason,
      TRUE ~ "unavailable_without_more_specific_reason_in_input"
    )
  ) |>
  dplyr::count(
    .data$data_scenario_id,
    .data$placement,
    .data$metric_order,
    .data$metric_id,
    .data$availability_reason,
    name = "participant_days"
  ) |>
  dplyr::arrange(
    .data$data_scenario_id,
    .data$placement,
    .data$metric_order,
    .data$availability_reason
  )

h08_write_csv(
  missingness,
  file.path(roots$model_data, "H08_metric_missingness.csv")
)

h08_key_columns <- c("site", "Id", "local_date", "metric_id")

paired_keys <- model_rows |>
  dplyr::filter(is.finite(.data$value), is.finite(.data$VLSQ8)) |>
  dplyr::group_by(
    .data$data_scenario_id,
    .data$site,
    .data$Id,
    .data$local_date,
    .data$metric_id
  ) |>
  dplyr::summarise(
    placements = dplyr::n_distinct(.data$placement),
    .groups = "drop"
  ) |>
  dplyr::filter(.data$placements == 2L) |>
  dplyr::select(-.data$placements)

main_gap_common_keys <- model_rows |>
  dplyr::filter(is.finite(.data$value), is.finite(.data$VLSQ8)) |>
  dplyr::group_by(
    .data$placement,
    .data$site,
    .data$Id,
    .data$local_date,
    .data$metric_id
  ) |>
  dplyr::summarise(
    scenarios = dplyr::n_distinct(.data$data_scenario_id),
    .groups = "drop"
  ) |>
  dplyr::filter(.data$scenarios == 2L) |>
  dplyr::select(-.data$scenarios)

h08_rows_for_run <- function(rows, run) {
  selected <- rows |>
    dplyr::filter(
      .data$data_scenario_id == run$data_scenario_id,
      .data$placement == run$placement
    )
  if (run$sample_scenario == "paired_common_sample") {
    selected <- selected |>
      dplyr::inner_join(
        paired_keys |>
          dplyr::filter(.data$data_scenario_id == run$data_scenario_id),
        by = c("data_scenario_id", h08_key_columns),
        relationship = "many-to-one"
      )
  }
  if (run$sample_scenario == "main_gap_common_sample") {
    selected <- selected |>
      dplyr::inner_join(
        main_gap_common_keys |>
          dplyr::filter(.data$placement == run$placement),
        by = c("placement", h08_key_columns),
        relationship = "many-to-one"
      )
  }
  selected
}

model_frames <- list()

frame_index_rows <- list()

frame_site_rows <- list()

frame_row_exports <- list()

for (run_index in seq_len(nrow(run_registry))) {
  run <- run_registry[run_index, , drop = FALSE]
  run_rows <- h08_rows_for_run(model_rows, run)
  for (metric_index in seq_len(nrow(metric_registry))) {
    spec <- metric_registry[metric_index, , drop = FALSE]
    frame <- h08_prepare_model_frame(
      run_rows[run_rows$metric_id == spec$metric_id, , drop = FALSE],
      spec,
      site_levels,
      score_contract
    )
    frame_key <- paste(run$run_id, spec$metric_id, sep = "__")
    model_frames[[frame_key]] <- frame
    frame_index_rows[[frame_key]] <- tibble::tibble(
      run_order = run$run_order,
      run_id = run$run_id,
      data_scenario_id = run$data_scenario_id,
      reader_scenario = run$reader_scenario,
      placement = run$placement,
      placement_label = run$placement_label,
      sample_scenario = run$sample_scenario,
      inferential_run = run$inferential_run,
      metric_order = spec$metric_order,
      metric_id = spec$metric_id,
      manuscript_name = spec$manuscript_name,
      participants = dplyr::n_distinct(frame$participant_key),
      participant_days = nrow(frame),
      sites = dplyr::n_distinct(frame$site),
      metric_support_valid_hours = h08_complete_sum(
        frame$metric_support_valid_minutes
      ) /
        60,
      metric_support_expected_hours = h08_complete_sum(
        frame$metric_support_expected_minutes
      ) /
        60,
      metric_support_missing_rows = sum(
        !is.finite(frame$metric_support_valid_minutes) |
          !is.finite(frame$metric_support_expected_minutes)
      ),
      site_levels = paste(levels(frame$site), collapse = "|"),
      site_contrasts = paste0("contr.sum(", nlevels(frame$site), ")"),
      score_center = score_contract$center,
      score_participant_sd = score_contract$participant_sd,
      row_key_hash = h08_key_hash(frame)
    )
    frame_site_rows[[frame_key]] <- frame |>
      dplyr::group_by(.data$site) |>
      dplyr::summarise(
        participants = dplyr::n_distinct(.data$participant_key),
        participant_days = dplyr::n(),
        metric_support_valid_hours = h08_complete_sum(
          .data$metric_support_valid_minutes
        ) /
          60,
        metric_support_expected_hours = h08_complete_sum(
          .data$metric_support_expected_minutes
        ) /
          60,
        .groups = "drop"
      ) |>
      dplyr::mutate(
        run_id = run$run_id,
        metric_order = spec$metric_order,
        metric_id = spec$metric_id,
        .before = 1L
      )
    frame_row_exports[[frame_key]] <- frame |>
      dplyr::transmute(
        run_id = run$run_id,
        sample_scenario = run$sample_scenario,
        metric_order = spec$metric_order,
        metric_id = spec$metric_id,
        .data$.model_row_id,
        site = as.character(.data$site),
        Id = as.character(.data$Id),
        .data$local_date,
        .data$value,
        .data$response_value,
        .data$VLSQ8,
        .data$VLSQ8_c,
        .data$photoperiod_hours,
        .data$metric_support_valid_minutes,
        .data$metric_support_expected_minutes
      )
  }
}

model_frame_index <- dplyr::bind_rows(frame_index_rows) |>
  dplyr::arrange(.data$run_order, .data$metric_order)

model_frame_site <- dplyr::bind_rows(frame_site_rows) |>
  dplyr::left_join(
    site_registry |>
      dplyr::select(.data$site, .data$display_order, .data$display_name),
    by = "site",
    relationship = "many-to-one"
  ) |>
  dplyr::arrange(.data$run_id, .data$metric_order, .data$display_order)

model_frame_rows <- dplyr::bind_rows(frame_row_exports)

h08_write_csv(
  model_frame_index,
  file.path(roots$model_data, "H08_model_frame_index.csv")
)

h08_write_csv(
  model_frame_site,
  file.path(roots$model_data, "H08_model_frame_by_site.csv")
)

h08_write_csv(
  model_frame_rows,
  file.path(roots$model_data, "H08_model_frame_rows.csv")
)

h08_write_rds(
  list(
    hypothesis_id = "H08",
    score_contract = score_contract,
    metric_registry = metric_registry,
    run_registry = run_registry,
    model_frames = model_frames
  ),
  file.path(roots$model_data, "H08_model_frames.rds")
)

paired_sample_audit <- model_frame_index |>
  dplyr::filter(.data$sample_scenario == "paired_common_sample") |>
  dplyr::select(
    .data$data_scenario_id,
    .data$metric_order,
    .data$metric_id,
    .data$placement,
    .data$participants,
    .data$participant_days,
    .data$sites,
    .data$row_key_hash
  ) |>
  tidyr::pivot_wider(
    names_from = .data$placement,
    values_from = c(
      .data$participants,
      .data$participant_days,
      .data$sites,
      .data$row_key_hash
    )
  ) |>
  dplyr::mutate(
    exact_counts_match = .data$participants_glasses ==
      .data$participants_chest &
      .data$participant_days_glasses == .data$participant_days_chest &
      .data$sites_glasses == .data$sites_chest,
    exact_row_keys_match = .data$row_key_hash_glasses ==
      .data$row_key_hash_chest
  )

if (
  any(
    !paired_sample_audit$exact_counts_match |
      !paired_sample_audit$exact_row_keys_match
  )
) {
  h08_abort("An H08 paired placement frame does not use identical row keys")
}

h08_write_csv(
  paired_sample_audit,
  file.path(roots$model_data, "H08_paired_sample_audit.csv")
)

main_gap_sample_audit <- model_frame_index |>
  dplyr::filter(.data$sample_scenario == "main_gap_common_sample") |>
  dplyr::select(
    .data$placement,
    .data$metric_order,
    .data$metric_id,
    .data$data_scenario_id,
    .data$participants,
    .data$participant_days,
    .data$sites,
    .data$row_key_hash
  ) |>
  tidyr::pivot_wider(
    names_from = .data$data_scenario_id,
    values_from = c(
      .data$participants,
      .data$participant_days,
      .data$sites,
      .data$row_key_hash
    )
  ) |>
  dplyr::mutate(
    exact_counts_match = .data$participants_main ==
      .data$participants_gap_timing_unaware &
      .data$participant_days_main == .data$participant_days_gap_timing_unaware &
      .data$sites_main == .data$sites_gap_timing_unaware,
    exact_row_keys_match = .data$row_key_hash_main ==
      .data$row_key_hash_gap_timing_unaware
  )

if (
  any(
    !main_gap_sample_audit$exact_counts_match |
      !main_gap_sample_audit$exact_row_keys_match
  )
) {
  h08_abort(
    "An H08 primary--gap common-sample frame does not use identical keys"
  )
}

h08_write_csv(
  main_gap_sample_audit,
  file.path(roots$model_data, "H08_main_gap_common_sample_audit.csv")
)
score_audit
model_frame_index
```



## Fit the main models

Fit the site-adjusted average effect and site interaction models for all available and paired samples. The response family and transformation depend on the exposure metric. Apply false discovery rate adjustment within the declared test families.

```{r}
#| label: h08-models

model_bundles <- list()

fit_index_rows <- list()

model_test_rows <- list()

model_effect_rows <- list()

model_site_slope_rows <- list()

model_prediction_rows <- list()

model_diagnostic_rows <- list()

diagnostic_plot_rows <- list()

influence_rows <- list()

for (run_index in seq_len(nrow(run_registry))) {
  run <- run_registry[run_index, , drop = FALSE]
  for (metric_index in seq_len(nrow(metric_registry))) {
    spec <- metric_registry[metric_index, , drop = FALSE]
    frame_key <- paste(run$run_id, spec$metric_id, sep = "__")
    frame <- model_frames[[frame_key]]
    message("H08 fit: ", run$run_id, " / ", spec$metric_id)
    bundle <- h08_fit_bundle(frame, spec, formula_kind = "participant_day")
    model_bundles[[frame_key]] <- bundle
    context <- tibble::tibble(
      run_order = run$run_order,
      run_id = run$run_id,
      data_scenario_id = run$data_scenario_id,
      reader_scenario = run$reader_scenario,
      placement = run$placement,
      placement_label = run$placement_label,
      sample_scenario = run$sample_scenario,
      analytical_role = run$analytical_role,
      inferential_run = run$inferential_run,
      metric_order = spec$metric_order,
      metric_id = spec$metric_id,
      manuscript_name = spec$manuscript_name,
      response_family = spec$response_family,
      response_transform = spec$response_transform,
      effect_scale = spec$effect_scale
    )
    fit_index_rows[[frame_key]] <- dplyr::bind_cols(
      context[rep(1L, 3L), , drop = FALSE],
      h08_fit_index_rows(bundle)
    )
    model_test_rows[[frame_key]] <- dplyr::bind_cols(
      context[rep(1L, 2L), , drop = FALSE],
      h08_bundle_tests(bundle, inferential = run$inferential_run)
    )
    model_effect_rows[[frame_key]] <- dplyr::bind_cols(
      context,
      h08_effect_summary(
        bundle$fits$additive$model,
        spec,
        score_contract$participant_sd
      )
    )
    site_slopes <- h08_site_slopes(
      bundle$fits$interaction$model,
      frame,
      spec,
      score_contract$participant_sd
    )
    model_site_slope_rows[[frame_key]] <- dplyr::bind_cols(
      context[rep(1L, nrow(site_slopes)), , drop = FALSE],
      site_slopes
    )
    predictions <- h08_centered_predictions(
      bundle$fits$additive$model,
      frame,
      spec,
      score_contract$participant_sd
    )
    model_prediction_rows[[frame_key]] <- dplyr::bind_cols(
      context[rep(1L, nrow(predictions)), , drop = FALSE],
      predictions
    )
    diagnostic <- h08_model_diagnostics(bundle, frame)
    model_diagnostic_rows[[frame_key]] <- dplyr::bind_cols(context, diagnostic)
    if (
      run$run_id %in%
        c(
          "main__glasses__all_available",
          "main__chest__all_available"
        )
    ) {
      plot_data <- h08_diagnostic_plot_data(bundle$fits$additive$model, frame)
      diagnostic_plot_rows[[frame_key]] <- dplyr::bind_cols(
        context[rep(1L, nrow(plot_data)), , drop = FALSE],
        plot_data
      )
      influence <- h08_participant_influence_screen(
        bundle$fits$additive$model,
        frame,
        n = 5L
      )
      influence_rows[[frame_key]] <- dplyr::bind_cols(
        context[rep(1L, nrow(influence)), , drop = FALSE],
        influence
      )
    }
  }
}

fit_index <- dplyr::bind_rows(fit_index_rows) |>
  dplyr::arrange(.data$run_order, .data$metric_order, .data$model_name)

model_tests <- dplyr::bind_rows(model_test_rows) |>
  dplyr::left_join(
    family_registry,
    by = c("run_id", "comparison_id"),
    relationship = "many-to-one"
  )

model_tests$p_adjusted <- NA_real_

for (family_id in family_registry$family_id) {
  rows <- which(model_tests$family_id == family_id)
  planned_n <- unique(model_tests$planned_n[rows])
  if (length(rows) != 9L || length(planned_n) != 1L || planned_n != 9L) {
    h08_abort(
      "Multiplicity family `%s` is not a complete nine-row family",
      family_id
    )
  }
  model_tests$p_adjusted[rows] <- adjust_p_family(
    model_tests$p_raw[rows],
    method = "BH",
    n = 9L
  )
}

model_tests <- model_tests |>
  dplyr::mutate(
    raw_significant = !is.na(.data$p_raw) & .data$p_raw <= 0.05,
    adjusted_significant = !is.na(.data$p_adjusted) & .data$p_adjusted <= 0.05,
    raw_p_display = nh_format_p_value(.data$p_raw),
    adjusted_p_display = nh_format_p_value(.data$p_adjusted)
  ) |>
  dplyr::arrange(.data$run_order, .data$comparison_id, .data$metric_order)

family_audit <- model_tests |>
  dplyr::filter(!is.na(.data$family_id)) |>
  dplyr::group_by(
    .data$family_id,
    .data$run_id,
    .data$comparison_id,
    .data$planned_n,
    .data$multiplicity_method,
    .data$role
  ) |>
  dplyr::summarise(
    registered_rows = dplyr::n(),
    observed_raw_p = sum(is.finite(.data$p_raw)),
    observed_adjusted_p = sum(is.finite(.data$p_adjusted)),
    raw_significant_n = sum(.data$raw_significant),
    adjusted_significant_n = sum(.data$adjusted_significant),
    complete_nine_member_family = .data$registered_rows == 9L,
    independent_recalculation_matches = isTRUE(all.equal(
      .data$p_adjusted,
      stats::p.adjust(.data$p_raw, method = "BH", n = 9L)
    )),
    .groups = "drop"
  )

if (
  any(
    !family_audit$complete_nine_member_family |
      !family_audit$independent_recalculation_matches
  )
) {
  h08_abort("An H08 multiplicity family failed independent verification")
}

model_effects <- dplyr::bind_rows(model_effect_rows)

model_site_slopes <- dplyr::bind_rows(model_site_slope_rows) |>
  dplyr::left_join(
    site_registry |>
      dplyr::select(
        .data$site,
        .data$display_order,
        .data$display_name,
        .data$color_hex
      ),
    by = "site",
    relationship = "many-to-one"
  ) |>
  dplyr::arrange(.data$run_order, .data$metric_order, .data$display_order)

model_predictions <- dplyr::bind_rows(model_prediction_rows)

model_diagnostics <- dplyr::bind_rows(model_diagnostic_rows) |>
  dplyr::arrange(.data$run_order, .data$metric_order)

diagnostic_plot_data <- dplyr::bind_rows(diagnostic_plot_rows)

participant_influence <- dplyr::bind_rows(influence_rows)

average_tests <- model_tests |>
  dplyr::filter(.data$comparison_id == "average_vlsq") |>
  dplyr::select(
    .data$run_id,
    .data$metric_id,
    average_lrt_statistic = .data$statistic,
    average_lrt_df = .data$df,
    average_p_raw = .data$p_raw,
    average_p_adjusted = .data$p_adjusted,
    average_adjusted_significant = .data$adjusted_significant,
    average_comparison_status = .data$comparison_status
  )

interaction_tests <- model_tests |>
  dplyr::filter(.data$comparison_id == "site_heterogeneity") |>
  dplyr::select(
    .data$run_id,
    .data$metric_id,
    interaction_lrt_statistic = .data$statistic,
    interaction_lrt_df = .data$df,
    interaction_p_raw = .data$p_raw,
    interaction_p_adjusted = .data$p_adjusted,
    interaction_adjusted_significant = .data$adjusted_significant,
    interaction_comparison_status = .data$comparison_status
  )

model_results_master <- model_frame_index |>
  dplyr::left_join(
    model_effects,
    by = c(
      "run_order",
      "run_id",
      "data_scenario_id",
      "reader_scenario",
      "placement",
      "placement_label",
      "sample_scenario",
      "inferential_run",
      "metric_order",
      "metric_id",
      "manuscript_name"
    ),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    average_tests,
    by = c("run_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    interaction_tests,
    by = c("run_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    model_diagnostics |>
      dplyr::select(
        .data$run_id,
        .data$metric_id,
        .data$diagnostic_status,
        .data$diagnostic_issues,
        .data$average_effect_status,
        .data$interaction_effect_status
      ),
    by = c("run_id", "metric_id"),
    relationship = "one-to-one"
  )

h08_write_csv(
  fit_index,
  file.path(roots$models, "H08_model_fit_index.csv")
)

h08_write_csv(model_tests, file.path(roots$tables, "H08_model_tests.csv"))

h08_write_csv(model_effects, file.path(roots$tables, "H08_model_effects.csv"))

h08_write_csv(
  model_site_slopes,
  file.path(roots$tables, "H08_site_specific_slopes.csv")
)

h08_write_csv(
  model_predictions,
  file.path(roots$tables, "H08_centered_predictions.csv")
)

h08_write_csv(family_audit, file.path(roots$tables, "H08_family_audit.csv"))

h08_write_csv(
  model_results_master,
  file.path(roots$tables, "H08_model_results_master.csv")
)

h08_write_csv(
  model_diagnostics,
  file.path(roots$diagnostics, "H08_model_diagnostics.csv")
)

h08_write_csv(
  diagnostic_plot_data,
  file.path(roots$source_data, "H08_primary_diagnostic_plot_data.csv")
)

h08_write_csv(
  participant_influence,
  file.path(roots$diagnostics, "H08_participant_influence_screen.csv")
)

h08_write_rds(
  list(
    hypothesis_id = "H08",
    score_contract = score_contract,
    metric_registry = metric_registry,
    run_registry = run_registry,
    model_bundles = model_bundles
  ),
  file.path(roots$models, "H08_model_bundles.rds")
)
model_results_master
```



## Scientific sensitivity analyses

Refit after adding photoperiod, using participant-level summaries, using exact longest-interval variants and observed dose, and omitting one site at a time.

```{r}
#| label: h08-sensitivities

sensitivity_models <- list()

photoperiod_rows <- list()

participant_rows <- list()

for (placement in c("glasses", "chest")) {
  run_id <- paste("main", placement, "all_available", sep = "__")
  for (metric_index in seq_len(nrow(metric_registry))) {
    spec <- metric_registry[metric_index, , drop = FALSE]
    key <- paste(run_id, spec$metric_id, sep = "__")
    frame <- model_frames[[key]]

    if (any(!is.finite(frame$photoperiod_c))) {
      h08_abort("Primary H08 photoperiod is incomplete for `%s`", key)
    }
    photo_key <- paste("photoperiod", placement, spec$metric_id, sep = "__")
    photo_bundle <- h08_fit_bundle(frame, spec, formula_kind = "photoperiod")
    sensitivity_models[[photo_key]] <- photo_bundle
    photo_effect <- h08_effect_summary(
      photo_bundle$fits$additive$model,
      spec,
      score_contract$participant_sd
    )
    photo_status <- h08_model_fit_status(photo_bundle$fits$additive$model)
    photo_condition <- h08_model_condition(photo_bundle$fits$additive$model)
    photoperiod_rows[[photo_key]] <- dplyr::bind_cols(
      tibble::tibble(
        sensitivity_id = "photoperiod_adjusted",
        placement = placement,
        placement_label = ifelse(placement == "glasses", "Near eye", "Chest"),
        metric_order = spec$metric_order,
        metric_id = spec$metric_id,
        manuscript_name = spec$manuscript_name,
        participants = dplyr::n_distinct(frame$participant_key),
        participant_days = nrow(frame),
        sites = dplyr::n_distinct(frame$site),
        row_key_hash = h08_key_hash(frame),
        formula = paste(
          deparse(h08_formula_set("photoperiod")$additive),
          collapse = " "
        )
      ),
      photo_effect,
      photo_status,
      photo_condition
    )

    participant_frame <- h08_prepare_participant_summary(
      frame,
      spec,
      site_levels
    )
    participant_key <- paste(
      "participant",
      placement,
      spec$metric_id,
      sep = "__"
    )
    participant_bundle <- h08_fit_bundle(
      participant_frame,
      spec,
      formula_kind = "participant"
    )
    sensitivity_models[[participant_key]] <- participant_bundle
    participant_effect <- h08_effect_summary(
      participant_bundle$fits$additive$model,
      spec,
      score_contract$participant_sd
    )
    participant_status <- h08_model_fit_status(
      participant_bundle$fits$additive$model
    )
    participant_condition <- h08_model_condition(
      participant_bundle$fits$additive$model
    )
    participant_rows[[participant_key]] <- dplyr::bind_cols(
      tibble::tibble(
        sensitivity_id = "participant_summary",
        placement = placement,
        placement_label = ifelse(placement == "glasses", "Near eye", "Chest"),
        metric_order = spec$metric_order,
        metric_id = spec$metric_id,
        manuscript_name = spec$manuscript_name,
        participants = nrow(participant_frame),
        participant_days_contributing = sum(participant_frame$participant_days),
        sites = dplyr::n_distinct(participant_frame$site),
        metric_support_valid_hours = h08_complete_sum(
          participant_frame$metric_support_valid_minutes
        ) /
          60,
        metric_support_expected_hours = h08_complete_sum(
          participant_frame$metric_support_expected_minutes
        ) /
          60,
        row_key_hash = h08_key_hash(participant_frame),
        formula = paste(
          deparse(h08_formula_set("participant")$additive),
          collapse = " "
        )
      ),
      participant_effect,
      participant_status,
      participant_condition
    )
  }
}

photoperiod_sensitivity <- dplyr::bind_rows(photoperiod_rows) |>
  dplyr::arrange(.data$placement, .data$metric_order)

participant_summary_sensitivity <- dplyr::bind_rows(participant_rows) |>
  dplyr::arrange(.data$placement, .data$metric_order)

h08_write_csv(
  photoperiod_sensitivity,
  file.path(roots$tables, "H08_photoperiod_sensitivity.csv")
)

h08_write_csv(
  participant_summary_sensitivity,
  file.path(roots$tables, "H08_participant_summary_sensitivity.csv")
)

h08_metric_variant_runs <- function(
  variant = c("exact_longest", "observed_dose")
) {
  variant <- match.arg(variant)
  if (variant == "exact_longest") {
    metric_id <- "longest_bout_above_250"
    value_column <- "longest_exact_value"
    sensitivity_id <- "longest_period_exact_only"
  } else {
    metric_id <- "dose_time_sensitive_corrected_medi"
    value_column <- "dose_observed_value"
    sensitivity_id <- "observed_dose_common_sample"
  }
  source <- model_rows |>
    dplyr::filter(
      .data$data_scenario_id == "main",
      .data$metric_id == .env$metric_id,
      is.finite(.data$value),
      is.finite(.data[[value_column]])
    )
  variant_pair_keys <- source |>
    dplyr::group_by(
      .data$site,
      .data$Id,
      .data$local_date,
      .data$metric_id
    ) |>
    dplyr::summarise(
      placements = dplyr::n_distinct(.data$placement),
      .groups = "drop"
    ) |>
    dplyr::filter(.data$placements == 2L) |>
    dplyr::select(-.data$placements)
  runs <- run_registry |>
    dplyr::filter(
      .data$data_scenario_id == "main",
      .data$sample_scenario %in% c("all_available", "paired_common_sample")
    )
  result_rows <- list()
  models <- list()
  for (index in seq_len(nrow(runs))) {
    run <- runs[index, , drop = FALSE]
    selected <- source |>
      dplyr::filter(.data$placement == run$placement)
    if (run$sample_scenario == "paired_common_sample") {
      selected <- selected |>
        dplyr::inner_join(
          variant_pair_keys,
          by = h08_key_columns,
          relationship = "many-to-one"
        )
    }
    selected$value <- selected[[value_column]]
    spec <- metric_registry[
      metric_registry$metric_id == metric_id,
      ,
      drop = FALSE
    ]
    frame <- h08_prepare_model_frame(
      selected,
      spec,
      site_levels,
      score_contract
    )
    key <- paste(sensitivity_id, run$placement, run$sample_scenario, sep = "__")
    bundle <- h08_fit_bundle(frame, spec, formula_kind = "participant_day")
    models[[key]] <- bundle
    result_rows[[key]] <- dplyr::bind_cols(
      tibble::tibble(
        sensitivity_id = sensitivity_id,
        placement = run$placement,
        placement_label = run$placement_label,
        sample_scenario = run$sample_scenario,
        metric_order = spec$metric_order,
        metric_id = metric_id,
        manuscript_name = spec$manuscript_name,
        participants = dplyr::n_distinct(frame$participant_key),
        participant_days = nrow(frame),
        sites = dplyr::n_distinct(frame$site),
        metric_support_valid_hours = h08_complete_sum(
          frame$metric_support_valid_minutes
        ) /
          60,
        metric_support_expected_hours = h08_complete_sum(
          frame$metric_support_expected_minutes
        ) /
          60,
        row_key_hash = h08_key_hash(frame)
      ),
      h08_effect_summary(
        bundle$fits$additive$model,
        spec,
        score_contract$participant_sd
      ),
      h08_model_fit_status(bundle$fits$additive$model)
    )
  }
  list(results = dplyr::bind_rows(result_rows), models = models)
}

exact_longest <- h08_metric_variant_runs("exact_longest")

observed_dose <- h08_metric_variant_runs("observed_dose")

sensitivity_models <- c(
  sensitivity_models,
  exact_longest$models,
  observed_dose$models
)

h08_write_csv(
  exact_longest$results,
  file.path(
    roots$tables,
    "H08_exactly_identified_longest_period_sensitivity.csv"
  )
)

h08_write_csv(
  observed_dose$results,
  file.path(roots$tables, "H08_observed_dose_sensitivity.csv")
)

leave_one_site_out_rows <- list()

for (placement in c("glasses", "chest")) {
  run_id <- paste("main", placement, "all_available", sep = "__")
  for (metric_index in seq_len(nrow(metric_registry))) {
    spec <- metric_registry[metric_index, , drop = FALSE]
    key <- paste(run_id, spec$metric_id, sep = "__")
    frame <- model_frames[[key]]
    full_effect <- model_effects |>
      dplyr::filter(
        .data$run_id == .env$run_id,
        .data$metric_id == spec$metric_id
      )
    loo <- h08_leave_one_site_out(
      frame,
      spec,
      score_contract$participant_sd,
      full_effect$estimate_model_per_point
    )
    leave_one_site_out_rows[[key]] <- dplyr::bind_cols(
      tibble::tibble(
        placement = placement,
        placement_label = ifelse(placement == "glasses", "Near eye", "Chest"),
        metric_order = spec$metric_order,
        metric_id = spec$metric_id,
        manuscript_name = spec$manuscript_name
      )[rep(1L, nrow(loo)), , drop = FALSE],
      loo
    )
  }
}

leave_one_site_out <- dplyr::bind_rows(leave_one_site_out_rows) |>
  dplyr::left_join(
    site_registry |>
      dplyr::select(
        omitted_site = .data$site,
        omitted_site_display_order = .data$display_order,
        omitted_site_name = .data$display_name
      ),
    by = "omitted_site",
    relationship = "many-to-one"
  ) |>
  dplyr::arrange(
    .data$placement,
    .data$metric_order,
    .data$omitted_site_display_order
  )

h08_write_csv(
  leave_one_site_out,
  file.path(roots$diagnostics, "H08_leave_one_site_out.csv")
)

h08_write_rds(
  list(
    hypothesis_id = "H08",
    sensitivity_registry = sensitivity_registry,
    sensitivity_models = sensitivity_models
  ),
  file.path(roots$models, "H08_sensitivity_models.rds")
)
h08_stability_class <- function(
  estimate_a,
  low_a,
  high_a,
  estimate_b,
  low_b,
  high_b,
  conclusion_a = NA,
  conclusion_b = NA
) {
  if (
    any(!is.finite(c(estimate_a, low_a, high_a, estimate_b, low_b, high_b)))
  ) {
    return("non-estimable")
  }
  if (
    sign(estimate_a) != sign(estimate_b) && estimate_a != 0 && estimate_b != 0
  ) {
    return("direction-sensitive")
  }
  if (
    !is.na(conclusion_a) && !is.na(conclusion_b) && conclusion_a != conclusion_b
  ) {
    return("multiplicity-conclusion-sensitive")
  }
  excludes_zero_a <- low_a > 0 || high_a < 0
  excludes_zero_b <- low_b > 0 || high_b < 0
  if (excludes_zero_a != excludes_zero_b) {
    return("precision-sensitive")
  }
  mutually_contained <-
    estimate_a >= low_b &&
    estimate_a <= high_b &&
    estimate_b >= low_a &&
    estimate_b <= high_a
  if (!mutually_contained) {
    return("magnitude-sensitive")
  }
  "stable within model uncertainty"
}

scenario_effects <- model_results_master |>
  dplyr::filter(
    .data$sample_scenario %in% c("all_available", "main_gap_common_sample"),
    .data$data_scenario_id %in% c("main", "gap_timing_unaware")
  ) |>
  dplyr::select(
    .data$placement,
    .data$sample_scenario,
    .data$metric_order,
    .data$metric_id,
    .data$manuscript_name,
    .data$data_scenario_id,
    estimate = .data$estimate_model_per_sd,
    conf_low = .data$conf_low_model_per_sd,
    conf_high = .data$conf_high_model_per_sd,
    p_adjusted = .data$average_p_adjusted,
    adjusted_significant = .data$average_adjusted_significant,
    participants = .data$participants,
    participant_days = .data$participant_days,
    sites = .data$sites,
    row_key_hash = .data$row_key_hash
  ) |>
  tidyr::pivot_wider(
    names_from = .data$data_scenario_id,
    values_from = c(
      .data$estimate,
      .data$conf_low,
      .data$conf_high,
      .data$p_adjusted,
      .data$adjusted_significant,
      .data$participants,
      .data$participant_days,
      .data$sites,
      .data$row_key_hash
    )
  ) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    exact_common_keys = dplyr::if_else(
      .data$sample_scenario == "main_gap_common_sample",
      .data$row_key_hash_main == .data$row_key_hash_gap_timing_unaware,
      NA
    ),
    stability_classification = h08_stability_class(
      .data$estimate_main,
      .data$conf_low_main,
      .data$conf_high_main,
      .data$estimate_gap_timing_unaware,
      .data$conf_low_gap_timing_unaware,
      .data$conf_high_gap_timing_unaware,
      .data$adjusted_significant_main,
      .data$adjusted_significant_gap_timing_unaware
    )
  ) |>
  dplyr::ungroup() |>
  dplyr::arrange(.data$sample_scenario, .data$placement, .data$metric_order)

loo_summary <- leave_one_site_out |>
  dplyr::group_by(
    .data$placement,
    .data$placement_label,
    .data$metric_order,
    .data$metric_id,
    .data$manuscript_name
  ) |>
  dplyr::summarise(
    refits = dplyr::n(),
    successful_refits = sum(.data$refit_status == "PASS"),
    sign_reversal_any = any(.data$sign_reversal, na.rm = TRUE),
    maximum_relative_absolute_change = max(
      .data$relative_absolute_change,
      na.rm = TRUE
    ),
    most_influential_omitted_site = .data$omitted_site_name[
      which.max(.data$relative_absolute_change)
    ],
    influence_status = dplyr::case_when(
      .data$successful_refits < .data$refits ~ "non-estimable refit present",
      .data$sign_reversal_any ~ "direction-sensitive to one site",
      .data$maximum_relative_absolute_change >= 0.5 ~
        "magnitude-sensitive to one site",
      TRUE ~ "direction stable in leave-one-site-out refits"
    ),
    .groups = "drop"
  )

response_family_check <- model_diagnostics |>
  dplyr::filter(.data$inferential_run) |>
  dplyr::group_by(
    .data$metric_order,
    .data$metric_id,
    .data$manuscript_name,
    .data$response_family,
    .data$response_transform
  ) |>
  dplyr::summarise(
    inferential_targets = dplyr::n(),
    major_failures = sum(.data$diagnostic_status == "MODEL_CHECK_FAILED"),
    review_targets = sum(.data$diagnostic_status == "REVIEW_WITH_LIMITATIONS"),
    average_non_estimable = sum(.data$average_effect_status != "ESTIMABLE"),
    interaction_non_estimable = sum(
      .data$interaction_effect_status != "ESTIMABLE"
    ),
    family_check_status = dplyr::case_when(
      .data$major_failures > 0L || .data$average_non_estimable > 0L ~
        "UNRESOLVED_COMMON_RESPONSE_FAMILY_CHECK",
      .data$review_targets > 0L ~ "RETAIN_WITH_EXPLICIT_LIMITATIONS",
      TRUE ~ "PASS"
    ),
    .groups = "drop"
  )

h08_write_csv(
  scenario_effects,
  file.path(roots$tables, "H08_gap_timing_unaware_sensitivity.csv")
)

h08_write_csv(
  loo_summary,
  file.path(roots$diagnostics, "H08_leave_one_site_out_summary.csv")
)

h08_write_csv(
  response_family_check,
  file.path(roots$diagnostics, "H08_response_family_check.csv")
)
scenario_effects
loo_summary
```



## Figures and numerical plot data

Export each plotted estimate and interval to CSV, then display placement agreement, preprocessing agreement and residual diagnostics.

```{r}
#| label: h08-figures

h08_effect_plot_data <- model_results_master |>
  dplyr::filter(
    .data$run_id %in%
      c(
        "main__glasses__all_available",
        "main__chest__all_available"
      )
  ) |>
  dplyr::mutate(
    plot_estimate = dplyr::if_else(
      .data$effect_type == "ratio",
      100 * (.data$estimate_practical_per_sd - 1),
      .data$estimate_practical_per_sd
    ),
    plot_conf_low = dplyr::if_else(
      .data$effect_type == "ratio",
      100 * (.data$conf_low_practical_per_sd - 1),
      .data$conf_low_practical_per_sd
    ),
    plot_conf_high = dplyr::if_else(
      .data$effect_type == "ratio",
      100 * (.data$conf_high_practical_per_sd - 1),
      .data$conf_high_practical_per_sd
    ),
    plot_scale = dplyr::if_else(
      .data$effect_type == "ratio",
      "Percent change per one VLSQ-8 SD",
      "Difference (h) per one VLSQ-8 SD"
    ),
    metric_label = factor(
      .data$manuscript_name,
      levels = rev(metric_registry$manuscript_name)
    )
  )

h08_save_effect_plot <- function(data, placement, stem, title, colour) {
  plot_data <- data |>
    dplyr::filter(.data$placement == .env$placement)
  source_path <- file.path(
    roots$source_data,
    paste0(stem, "_data.csv")
  )
  h08_write_csv(plot_data, source_path)
  plot <- ggplot2::ggplot(
    plot_data,
    ggplot2::aes(
      x = .data$plot_estimate,
      y = .data$metric_label,
      xmin = .data$plot_conf_low,
      xmax = .data$plot_conf_high
    )
  ) +
    ggplot2::geom_vline(xintercept = 0, colour = "grey55", linewidth = 0.5) +
    ggplot2::geom_errorbar(
      orientation = "y",
      width = 0.18,
      linewidth = 0.65,
      colour = colour
    ) +
    ggplot2::geom_point(
      size = 2.3,
      shape = 21,
      fill = "white",
      colour = colour
    ) +
    ggplot2::facet_wrap(
      ggplot2::vars(.data$plot_scale),
      ncol = 1,
      scales = "free",
      space = "free_y",
      strip.position = "right"
    ) +
    ggplot2::labs(
      title = title,
      subtitle = paste0(
        "Adjusted association per ",
        sprintf("%.4f", score_contract$participant_sd),
        " VLSQ-8 points; points and bars are estimates and 95% Wald intervals"
      ),
      x = NULL,
      y = NULL,
      caption = paste0(
        "Ratios are shown as percent change. Time below 10 lx melEDI before sleep ",
        "uses an identity-Gaussian difference in hours."
      )
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(
      plot.title.position = "plot",
      panel.grid.minor = ggplot2::element_blank(),
      strip.text.y = ggplot2::element_text(angle = 0, hjust = 0),
      axis.text.y = ggplot2::element_text(size = 8),
      plot.caption = ggplot2::element_text(hjust = 0, size = 7)
    )
  ggplot2::ggsave(
    file.path(roots$figures, paste0(stem, ".png")),
    plot,
    width = 9,
    height = 7.2,
    units = "in",
    dpi = 300,
    bg = "white"
  )
  ggplot2::ggsave(
    file.path(roots$figures, paste0(stem, ".pdf")),
    plot,
    width = 9,
    height = 7.2,
    units = "in",
    bg = "white"
  )
  invisible(source_path)
}

h08_save_effect_plot(
  h08_effect_plot_data,
  "glasses",
  "H08_near_eye_effects",
  "Near-eye VLSQ-8 associations",
  "#0072B2"
)

h08_save_effect_plot(
  h08_effect_plot_data,
  "chest",
  "H08_chest_effects",
  "Chest VLSQ-8 associations",
  "#D55E00"
)

paired_effects <- model_results_master |>
  dplyr::filter(
    .data$data_scenario_id == "main",
    .data$sample_scenario == "paired_common_sample"
  ) |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(
        .data$metric_id,
        .data$abbreviation,
        .data$manuscript_category
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    comparison_estimate = dplyr::if_else(
      .data$effect_type == "ratio",
      log(pmax(.data$estimate_practical_per_sd, .Machine$double.xmin)),
      .data$estimate_practical_per_sd
    ),
    comparison_low = dplyr::if_else(
      .data$effect_type == "ratio",
      log(pmax(.data$conf_low_practical_per_sd, .Machine$double.xmin)),
      .data$conf_low_practical_per_sd
    ),
    comparison_high = dplyr::if_else(
      .data$effect_type == "ratio",
      log(pmax(.data$conf_high_practical_per_sd, .Machine$double.xmin)),
      .data$conf_high_practical_per_sd
    ),
    comparison_scale = dplyr::if_else(
      .data$effect_type == "ratio",
      "Natural-log ratio per one VLSQ-8 SD",
      "Difference in hours per one VLSQ-8 SD"
    )
  ) |>
  dplyr::select(
    .data$metric_order,
    .data$metric_id,
    .data$manuscript_name,
    .data$abbreviation,
    .data$manuscript_category,
    .data$effect_type,
    .data$comparison_scale,
    .data$placement,
    .data$comparison_estimate,
    .data$comparison_low,
    .data$comparison_high,
    .data$participants,
    .data$participant_days,
    .data$sites,
    .data$row_key_hash
  ) |>
  tidyr::pivot_wider(
    names_from = .data$placement,
    values_from = c(
      .data$comparison_estimate,
      .data$comparison_low,
      .data$comparison_high,
      .data$participants,
      .data$participant_days,
      .data$sites,
      .data$row_key_hash
    )
  ) |>
  dplyr::mutate(
    exact_sample_match = .data$row_key_hash_glasses == .data$row_key_hash_chest,
    included_in_identity_plot = .data$effect_type == "ratio"
  )

if (any(!paired_effects$exact_sample_match)) {
  h08_abort("The H08 paired effect display contains unmatched samples")
}

h08_write_csv(
  paired_effects,
  file.path(roots$source_data, "H08_paired_placement_effects_data.csv")
)

paired_ratio <- paired_effects |>
  dplyr::filter(.data$included_in_identity_plot)

paired_plot <- ggplot2::ggplot(
  paired_ratio,
  ggplot2::aes(
    x = .data$comparison_estimate_glasses,
    y = .data$comparison_estimate_chest,
    colour = .data$manuscript_category,
    shape = .data$manuscript_category
  )
) +
  ggplot2::geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey45",
    linewidth = 0.6
  ) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey70", linewidth = 0.5) +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.5) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = .data$comparison_low_chest,
      ymax = .data$comparison_high_chest
    ),
    width = 0,
    linewidth = 0.45
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      xmin = .data$comparison_low_glasses,
      xmax = .data$comparison_high_glasses
    ),
    orientation = "y",
    width = 0,
    linewidth = 0.45
  ) +
  ggplot2::geom_point(size = 2.8, stroke = 0.9) +
  ggplot2::geom_text(
    ggplot2::aes(label = .data$abbreviation),
    nudge_y = 0.018,
    size = 2.8,
    show.legend = FALSE,
    check_overlap = TRUE
  ) +
  ggplot2::coord_equal() +
  ggplot2::scale_colour_manual(
    values = c(
      "level-based" = "#0072B2",
      "duration-based" = "#009E73",
      "exposure-history-based" = "#CC79A7"
    )
  ) +
  ggplot2::labs(
    title = "Paired near-eye and chest association estimates",
    subtitle = paste0(
      "Separate models use identical participant-days;\n",
      "ratio outcomes use the natural-log ratio scale per one VLSQ-8 SD"
    ),
    x = "Near-eye estimate",
    y = "Chest estimate",
    colour = "Metric category",
    shape = "Metric category",
    caption = paste0(
      "Dashed line: identical estimates; grey lines: null; component bars: ",
      "95% Wald intervals.\n",
      "The identity-Gaussian pre-sleep metric remains in the paired source ",
      "table because hours and log ratios must not share an axis."
    )
  ) +
  ggplot2::theme_minimal(base_size = 10) +
  ggplot2::theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank(),
    plot.caption = ggplot2::element_text(hjust = 0, size = 7)
  )

ggplot2::ggsave(
  file.path(roots$figures, "H08_paired_placement_effects.png"),
  paired_plot,
  width = 7.2,
  height = 7.2,
  units = "in",
  dpi = 300,
  bg = "white"
)

ggplot2::ggsave(
  file.path(roots$figures, "H08_paired_placement_effects.pdf"),
  paired_plot,
  width = 7.2,
  height = 7.2,
  units = "in",
  bg = "white"
)

gap_common_plot_data <- scenario_effects |>
  dplyr::filter(.data$sample_scenario == "main_gap_common_sample") |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(.data$metric_id, .data$abbreviation, .data$effect_scale),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    included_in_identity_plot = .data$effect_scale == "ratio",
    placement_label = dplyr::if_else(
      .data$placement == "glasses",
      "Near eye",
      "Chest"
    ),
    primary_log_ratio = dplyr::if_else(
      .data$effect_scale == "ratio",
      dplyr::case_when(
        .data$metric_id %in%
          c(
            "daily_geometric_mean_medi",
            "m10_mean_medi",
            "l10_mean_medi",
            "longest_bout_above_250",
            "dose_time_sensitive_corrected_medi"
          ) ~
          log(10) * .data$estimate_main,
        TRUE ~ .data$estimate_main
      ),
      .data$estimate_main
    ),
    gap_log_ratio = dplyr::if_else(
      .data$effect_scale == "ratio",
      dplyr::case_when(
        .data$metric_id %in%
          c(
            "daily_geometric_mean_medi",
            "m10_mean_medi",
            "l10_mean_medi",
            "longest_bout_above_250",
            "dose_time_sensitive_corrected_medi"
          ) ~
          log(10) * .data$estimate_gap_timing_unaware,
        TRUE ~ .data$estimate_gap_timing_unaware
      ),
      .data$estimate_gap_timing_unaware
    )
  )

h08_write_csv(
  gap_common_plot_data,
  file.path(roots$source_data, "H08_gap_common_sample_effects_data.csv")
)

gap_plot <- ggplot2::ggplot(
  gap_common_plot_data |>
    dplyr::filter(.data$included_in_identity_plot),
  ggplot2::aes(
    x = .data$primary_log_ratio,
    y = .data$gap_log_ratio,
    colour = .data$placement_label,
    shape = .data$placement_label,
    label = .data$abbreviation
  )
) +
  ggplot2::geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey45"
  ) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey70") +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70") +
  ggplot2::geom_point(size = 2.6, stroke = 0.8) +
  ggplot2::geom_text(
    nudge_y = 0.015,
    size = 2.7,
    show.legend = FALSE,
    check_overlap = TRUE
  ) +
  ggplot2::coord_equal() +
  ggplot2::scale_colour_manual(
    values = c("Near eye" = "#0072B2", "Chest" = "#D55E00")
  ) +
  ggplot2::labs(
    title = "Primary and gap-timing-unaware estimates on common samples",
    subtitle = "Ratio outcomes; natural-log ratio per one VLSQ-8 SD",
    x = "Primary dataset estimate",
    y = "Gap-timing-unaware dataset estimate",
    colour = "Placement",
    shape = "Placement",
    caption = paste0(
      "Dashed line: identical estimates; grey lines: null. Each scenario uses ",
      "the same participant-days within metric and placement."
    )
  ) +
  ggplot2::theme_minimal(base_size = 10) +
  ggplot2::theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank(),
    plot.caption = ggplot2::element_text(hjust = 0, size = 7)
  )

ggplot2::ggsave(
  file.path(roots$figures, "H08_gap_common_sample_effects.png"),
  gap_plot,
  width = 7.2,
  height = 7.2,
  units = "in",
  dpi = 300,
  bg = "white"
)

near_diagnostic <- diagnostic_plot_data |>
  dplyr::filter(.data$run_id == "main__glasses__all_available") |>
  dplyr::mutate(
    manuscript_name = factor(
      .data$manuscript_name,
      levels = metric_registry$manuscript_name
    )
  )

h08_write_csv(
  near_diagnostic,
  file.path(roots$source_data, "H08_near_eye_model_adequacy_data.csv")
)

residual_panel <- near_diagnostic |>
  dplyr::transmute(
    .data$manuscript_name,
    panel = "Residual versus fitted",
    x = .data$fitted_model_scale,
    y = .data$residual_pearson
  )

qq_panel <- near_diagnostic |>
  dplyr::transmute(
    .data$manuscript_name,
    panel = "Normal Q-Q",
    x = .data$qq_theoretical,
    y = .data$qq_observed
  )

adequacy_plot_data <- dplyr::bind_rows(residual_panel, qq_panel)

adequacy_plot <- ggplot2::ggplot(
  adequacy_plot_data,
  ggplot2::aes(x = .data$x, y = .data$y)
) +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.4) +
  ggplot2::geom_point(alpha = 0.35, size = 0.7, colour = "#0072B2") +
  ggplot2::facet_grid(
    rows = ggplot2::vars(.data$manuscript_name),
    cols = ggplot2::vars(.data$panel),
    scales = "free"
  ) +
  ggplot2::labs(
    title = "Near-eye additive-model adequacy",
    subtitle = "Conditional Pearson residual screens; no simulation or resampling",
    x = NULL,
    y = NULL,
    caption = paste0(
      "The Q-Q panels are descriptive residual-shape checks. Formal numerical ",
      "fit, bound, zero-mass, and serial-dependence diagnostics are reported separately."
    )
  ) +
  ggplot2::theme_minimal(base_size = 9) +
  ggplot2::theme(
    plot.title.position = "plot",
    strip.text.y = ggplot2::element_text(angle = 0, hjust = 0, size = 7),
    strip.text.x = ggplot2::element_text(size = 8),
    panel.grid.minor = ggplot2::element_blank(),
    plot.caption = ggplot2::element_text(hjust = 0, size = 7)
  )

ggplot2::ggsave(
  file.path(roots$figures, "H08_near_eye_model_adequacy.png"),
  adequacy_plot,
  width = 10,
  height = 15,
  units = "in",
  dpi = 240,
  bg = "white"
)
```



## Final display sizing

Export the reader figures at a consistent physical width so axis labels and intervals remain readable.

```{r}
#| label: h08-display-size

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(readr)
  library(tibble)
})

figure_dir <- file.path(root, "results/images/H08")

source_dir <- file.path(root, "results/csv/source_data/H08")

metric_registry <- readr::read_csv(
  file.path(root, "results/intermediate/model_data/H08/H08_metric_registry.csv"),
  show_col_types = FALSE
)

read_source <- function(filename) {
  readr::read_csv(
    file.path(source_dir, filename),
    show_col_types = FALSE
  )
}

wrap_caption <- function(text, width = 72L) {
  paste(strwrap(text, width = width), collapse = "\n")
}

save_figure <- function(plot, stem, height_mm, write_pdf = FALSE) {
  ggplot2::ggsave(
    file.path(figure_dir, paste0(stem, ".png")),
    plot,
    width = 170,
    height = height_mm,
    units = "mm",
    dpi = 300,
    bg = "white"
  )
  if (write_pdf) {
    ggplot2::ggsave(
      file.path(figure_dir, paste0(stem, ".pdf")),
      plot,
      width = 170,
      height = height_mm,
      units = "mm",
      bg = "white"
    )
  }
  invisible(NULL)
}

effect_plot <- function(data, title, colour) {
  data <- data |>
    mutate(
      metric_label = factor(
        .data$metric_label,
        levels = rev(metric_registry$manuscript_name)
      )
    )
  ggplot2::ggplot(
    data,
    ggplot2::aes(
      x = .data$plot_estimate,
      y = .data$metric_label,
      xmin = .data$plot_conf_low,
      xmax = .data$plot_conf_high
    )
  ) +
    ggplot2::geom_vline(
      xintercept = 0,
      colour = "grey55",
      linewidth = 0.45
    ) +
    ggplot2::geom_errorbar(
      orientation = "y",
      width = 0.18,
      linewidth = 0.6,
      colour = colour
    ) +
    ggplot2::geom_point(
      size = 2.3,
      shape = 21,
      fill = "white",
      colour = colour
    ) +
    ggplot2::facet_wrap(
      ggplot2::vars(.data$plot_scale),
      ncol = 1,
      scales = "free",
      space = "free_y",
      strip.position = "right"
    ) +
    ggplot2::labs(
      title = title,
      subtitle = paste0(
        "Adjusted association per ",
        sprintf("%.4f", unique(data$score_participant_sd)),
        " VLSQ-8 points; points and bars are estimates and 95% Wald intervals"
      ),
      x = NULL,
      y = NULL,
      caption = wrap_caption(paste0(
        "Ratios are shown as percent change. Time below 10 lx melEDI before ",
        "sleep uses an identity-Gaussian difference in hours."
      ))
    ) +
    ggplot2::theme_minimal(base_size = 9.5) +
    ggplot2::theme(
      plot.title.position = "plot",
      panel.grid.minor = ggplot2::element_blank(),
      strip.text.y = ggplot2::element_text(
        angle = 0,
        hjust = 0,
        size = 8
      ),
      axis.text.y = ggplot2::element_text(size = 8),
      plot.caption = ggplot2::element_text(hjust = 0, size = 7.5)
    )
}

near <- read_source("H08_near_eye_effects_data.csv")

chest <- read_source("H08_chest_effects_data.csv")

stopifnot(nrow(near) == 9L, nrow(chest) == 9L)

save_figure(
  effect_plot(near, "Near-eye VLSQ-8 associations", "#0072B2"),
  "H08_near_eye_effects",
  height_mm = 136,
  write_pdf = TRUE
)

save_figure(
  effect_plot(chest, "Chest VLSQ-8 associations", "#D55E00"),
  "H08_chest_effects",
  height_mm = 136,
  write_pdf = TRUE
)

paired <- read_source("H08_paired_placement_effects_data.csv") |>
  filter(.data$included_in_identity_plot)

stopifnot(nrow(paired) == 8L, all(paired$exact_sample_match))

paired_plot <- ggplot2::ggplot(
  paired,
  ggplot2::aes(
    x = .data$comparison_estimate_glasses,
    y = .data$comparison_estimate_chest,
    colour = .data$manuscript_category,
    shape = .data$manuscript_category
  )
) +
  ggplot2::geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey45",
    linewidth = 0.6
  ) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey70", linewidth = 0.5) +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.5) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = .data$comparison_low_chest,
      ymax = .data$comparison_high_chest
    ),
    width = 0,
    linewidth = 0.45
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      xmin = .data$comparison_low_glasses,
      xmax = .data$comparison_high_glasses
    ),
    orientation = "y",
    width = 0,
    linewidth = 0.45
  ) +
  ggplot2::geom_point(size = 2.8, stroke = 0.9) +
  ggplot2::geom_text(
    ggplot2::aes(label = .data$abbreviation),
    nudge_y = 0.018,
    size = 2.8,
    show.legend = FALSE,
    check_overlap = TRUE
  ) +
  ggplot2::coord_equal() +
  ggplot2::scale_colour_manual(
    values = c(
      "level-based" = "#0072B2",
      "duration-based" = "#009E73",
      "exposure-history-based" = "#CC79A7"
    )
  ) +
  ggplot2::labs(
    title = "Paired near-eye and chest association estimates",
    subtitle = paste0(
      "Separate models use identical participant-days;\n",
      "ratio outcomes use the natural-log ratio scale per one VLSQ-8 SD"
    ),
    x = "Near-eye estimate",
    y = "Chest estimate",
    colour = "Metric category",
    shape = "Metric category",
    caption = wrap_caption(paste0(
      "Dashed line: identical estimates; grey lines: null; component bars: ",
      "95% Wald intervals. ",
      "The identity-Gaussian pre-sleep metric remains in the paired source ",
      "table because hours and log ratios must not share an axis."
    ))
  ) +
  ggplot2::theme_minimal(base_size = 9) +
  ggplot2::theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank(),
    plot.caption = ggplot2::element_text(hjust = 0, size = 7.5)
  )

save_figure(
  paired_plot,
  "H08_paired_placement_effects",
  height_mm = 170,
  write_pdf = TRUE
)

gap <- read_source("H08_gap_common_sample_effects_data.csv") |>
  filter(.data$included_in_identity_plot)

stopifnot(nrow(gap) == 16L, all(gap$exact_common_keys))

gap_plot <- ggplot2::ggplot(
  gap,
  ggplot2::aes(
    x = .data$primary_log_ratio,
    y = .data$gap_log_ratio,
    colour = .data$placement_label,
    shape = .data$placement_label,
    label = .data$abbreviation
  )
) +
  ggplot2::geom_abline(
    intercept = 0,
    slope = 1,
    linetype = "dashed",
    colour = "grey45"
  ) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey70") +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70") +
  ggplot2::geom_point(size = 2.6, stroke = 0.8) +
  ggplot2::geom_text(
    nudge_y = 0.015,
    size = 2.8,
    show.legend = FALSE,
    check_overlap = TRUE
  ) +
  ggplot2::coord_equal() +
  ggplot2::scale_colour_manual(
    values = c("Near eye" = "#0072B2", "Chest" = "#D55E00")
  ) +
  ggplot2::labs(
    title = "Primary and gap-timing-unaware estimates on common samples",
    subtitle = "Ratio outcomes; natural-log ratio per one VLSQ-8 SD",
    x = "Primary dataset estimate",
    y = "Gap-timing-unaware dataset estimate",
    colour = "Placement",
    shape = "Placement",
    caption = wrap_caption(paste0(
      "Dashed line: identical estimates; grey lines: null. Each scenario ",
      "uses the same participant-days within metric and placement."
    ))
  ) +
  ggplot2::theme_minimal(base_size = 9) +
  ggplot2::theme(
    plot.title.position = "plot",
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank(),
    plot.caption = ggplot2::element_text(hjust = 0, size = 7.5)
  )

save_figure(
  gap_plot,
  "H08_gap_common_sample_effects",
  height_mm = 170
)

adequacy <- read_source("H08_near_eye_model_adequacy_data.csv") |>
  mutate(
    manuscript_name = factor(
      .data$manuscript_name,
      levels = metric_registry$manuscript_name
    )
  )

stopifnot(nrow(adequacy) > 0L)

adequacy_plot_data <- bind_rows(
  adequacy |>
    transmute(
      .data$manuscript_name,
      panel = "Residual versus fitted",
      x = .data$fitted_model_scale,
      y = .data$residual_pearson
    ),
  adequacy |>
    transmute(
      .data$manuscript_name,
      panel = "Normal Q-Q",
      x = .data$qq_theoretical,
      y = .data$qq_observed
    )
)

adequacy_plot <- ggplot2::ggplot(
  adequacy_plot_data,
  ggplot2::aes(x = .data$x, y = .data$y)
) +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.4) +
  ggplot2::geom_point(alpha = 0.35, size = 0.7, colour = "#0072B2") +
  ggplot2::facet_grid(
    rows = ggplot2::vars(.data$manuscript_name),
    cols = ggplot2::vars(.data$panel),
    scales = "free"
  ) +
  ggplot2::labs(
    title = "Near-eye additive-model adequacy",
    subtitle = "Conditional Pearson residual screens; no simulation or resampling",
    x = NULL,
    y = NULL,
    caption = wrap_caption(paste0(
      "The Q-Q panels are descriptive residual-shape checks. Formal ",
      "numerical fit, bound, zero-mass, and serial-dependence diagnostics ",
      "are reported separately."
    ))
  ) +
  ggplot2::theme_minimal(base_size = 9) +
  ggplot2::theme(
    plot.title.position = "plot",
    strip.text.y = ggplot2::element_text(
      angle = 0,
      hjust = 0,
      size = 7.5
    ),
    strip.text.x = ggplot2::element_text(size = 8),
    panel.grid.minor = ggplot2::element_blank(),
    plot.caption = ggplot2::element_text(hjust = 0, size = 7.5)
  )

save_figure(
  adequacy_plot,
  "H08_near_eye_model_adequacy",
  height_mm = 250
)
```

## Findings and interpretation {#findings}

The following views use the models and summaries calculated above.

```{r}
#| label: setup-h08-reader
#| code-fold: true
#| code-summary: "Export results"

suppressPackageStartupMessages({
    library(dplyr)
    library(gt)
    library(readr)
    library(stringr)
    library(tibble)
    library(tidyr)
})
locate_project_root <- function(start = getwd()) {
    candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
    repeat {
        if (file.exists(file.path(candidate, "renv.lock")) && file.exists(file.path(candidate, "_quarto.yml"))) {
            return(candidate)
        }
        parent <- dirname(candidate)
        if (identical(parent, candidate)) {
            stop("Could not locate the project root", call. = FALSE)
        }
        candidate <- parent
    }
}
configured_root <- Sys.getenv("NATHEALTH_PROJECT_ROOT", unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = ""))
if (nzchar(configured_root) && file.exists(file.path(configured_root, "renv.lock"))) {
    root <- normalizePath(configured_root, winslash = "/", mustWork = TRUE)
} else {
    root <- locate_project_root()
}
source(file.path(root, "scripts", "pipeline", "p_value_display.R"))
read_h08 <- function(area, name) {
    readr::read_csv(file.path(root, "results", area, "H08", name), show_col_types = FALSE, progress = FALSE, na = "")
}
metric_registry <- read_h08("intermediate/model_data", "H08_metric_registry.csv")
formula_registry <- read_h08("intermediate/model_data", "H08_formula_registry.csv")
score_audit <- read_h08("csv/diagnostics", "H08_vlsq_score_audit.csv")
sample_index <- read_h08("intermediate/model_data", "H08_model_frame_index.csv")
paired_audit <- read_h08("intermediate/model_data", "H08_paired_sample_audit.csv")
main_gap_audit <- read_h08("intermediate/model_data", "H08_main_gap_common_sample_audit.csv")
master <- read_h08("tables", "H08_model_results_master.csv")
family_audit <- read_h08("tables", "H08_family_audit.csv")
diagnostics <- read_h08("csv/diagnostics", "H08_model_diagnostics.csv")
response_check <- read_h08("csv/diagnostics", "H08_response_family_check.csv")
loo_summary <- read_h08("csv/diagnostics", "H08_leave_one_site_out_summary.csv")
predictions <- read_h08("tables", "H08_centered_predictions.csv")
site_slopes <- read_h08("tables", "H08_site_specific_slopes.csv")
gap_sensitivity <- read_h08("tables", "H08_gap_timing_unaware_sensitivity.csv")
photoperiod <- read_h08("tables", "H08_photoperiod_sensitivity.csv")
participant_summary <- read_h08("tables", "H08_participant_summary_sensitivity.csv")
exact_longest <- read_h08("tables", "H08_exactly_identified_longest_period_sensitivity.csv")
observed_dose <- read_h08("tables", "H08_observed_dose_sensitivity.csv")
near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
near_paired_id <- "main__glasses__paired_common_sample"
chest_paired_id <- "main__chest__paired_common_sample"
near <- arrange(filter(master, run_id == near_id), metric_order)
chest <- arrange(filter(master, run_id == chest_id), metric_order)
format_effect <- function(estimate, low, high, effect_type) {
    ifelse(effect_type == "ratio", sprintf("×%.3f (%.3f–%.3f)", estimate, low, high), sprintf("%+.3f h (%+.3f–%+.3f h)",
        estimate, low, high))
}
format_p_cell <- function(value, significant) {
    display <- nh_p_value_display(value, significant = significant)
    ifelse(display$p_bold, paste0("**", display$p_display, "**"), display$p_display)
}
format_sample <- function(participants, participant_days, sites) {
    sprintf("%d / %d / %d", participants, participant_days, sites)
}
model_to_practical <- function(value, response_transform, effect_scale) {
    case_when(effect_scale == "difference" ~ value, response_transform == "log10_offset_0.1" ~ 10^value, effect_scale ==
        "ratio" ~ exp(value), TRUE ~ value)
}
h08_gt <- function(data, title = NULL, note = NULL) {
    output <- tab_options(sub_missing(opt_row_striping(gt(data)), missing_text = "Not available"), table.width = pct(100), table.font.size = px(12),
        data_row.padding = px(4), heading.align = "left", column_labels.font.weight = "600", source_notes.font.size = px(10),
        container.overflow.x = "auto")
    if (!is.null(title))
        output <- tab_header(output, title = md(title))
    if (!is.null(note))
        output <- tab_source_note(output, md(note))
    output
}
result_table <- function(data, near_eye_association_wording = FALSE) {
    output <- fmt_markdown(h08_gt(select(arrange(transmute(data, metric_order, Metric = manuscript_name, Scale = if_else(effect_type ==
        "difference", "Difference in hours", "Ratio (percentage change)"), `Effect per VLSQ-8 SD (95% CI)` = format_effect(estimate_practical_per_sd,
        conf_low_practical_per_sd, conf_high_practical_per_sd, effect_type), `Raw p` = format_p_cell(average_p_raw, average_adjusted_significant),
        `FDR-adjusted p` = format_p_cell(average_p_adjusted, average_adjusted_significant), `Participants / participant-days / sites` = format_sample(participants,
            participant_days, sites)), metric_order), -metric_order), note = paste(if (near_eye_association_wording) {
        "Associations compare scores separated by one participant-level SD"
    }
    else {
        "Effects compare scores separated by one participant-level SD"
    }, "(5.540 VLSQ-8 points). The hour-scale row is an absolute", "difference; ratio rows can be read as percentage changes.",
        "Intervals are two-sided 95% Wald CIs. FDR adjustment is across", "all nine metrics; adjusted p-values would be bold at 0.050.")),
        columns = c(`Raw p`, `FDR-adjusted p`))
    if (near_eye_association_wording) {
        output <- gt::cols_label_with(output, columns = dplyr::starts_with("Effect per"), fn = function(label) sub("^Effect",
            "Association", label))
    }
    output
}
near_dose <- filter(near, metric_id == "dose_time_sensitive_corrected_medi")
```


## Scientific question

The **Visual Light Sensitivity Questionnaire (VLSQ-8)** measures
self-reported visual light sensitivity; higher scores indicate greater
sensitivity according to the selected scale.

The preregistered hypothesis was:

> **H8: Duration-, exposure-history-, and level-based metrics are associated
> with VLSQ-8 light sensitivity scores.**

The analytical question is whether VLSQ-8 is associated with repeated
participant-day personal light-exposure metrics after adjustment for study
site. A **participant-day** is one participant's eligible local calendar day.
The primary **near-eye sensor position** more closely represents light near
the eyes during wear. The complementary **chest sensor position** is not an
ocular-exposure measure and is not pooled with near-eye measurements.
**Melanopic equivalent daylight illuminance (melEDI)** describes light in
terms of melanopsin-weighted visual-system sensitivity.

Reported effects compare VLSQ-8 scores separated by one participant-level
**standard deviation (SD)**, equal to 5.540 points. Time below 10 lx melEDI
before sleep is reported as an absolute difference in hours; the eight
log-link outcomes are reported as ratios and corresponding percentage
changes, on a separate scale. A **back-transformed** estimate returns a fitted
coefficient to its stated hour-difference, ratio, or percentage-change scale.
A **95% confidence interval (95% CI)** describes the uncertainty around that
effect; Wald construction is named where relevant.

The **site-average association** is an average across sites that gives each
site equal weight. The separate **VLSQ-8-by-site interaction** allows the
association between VLSQ-8 and a light metric to differ by study site.
**False-discovery-rate (FDR) adjustment** is applied within each declared
nine-metric family.

::: {.callout-note title="Answer in brief"}
Across nine light-exposure metrics, neither the primary near-eye site-average
associations nor their VLSQ-8-by-site interactions met the FDR-adjusted
criterion. The strongest near-eye directional pattern was lower corrected
melEDI dose per VLSQ-8 SD (ratio 0.846, 95% CI 0.716–0.999; raw *p* = 0.051,
FDR-adjusted *p* = 0.160), so it is not an FDR-retained finding.
Complementary chest results and the central sensitivity analyses did not
change that conclusion, although site influence and specified model
limitations require caution.
:::

## Score, sensor positions, and light-exposure metrics

VLSQ-8 was available for all 184 participants across nine sites. The stored
score equalled the sum of eight ordered 1–5 item codes plus 5 for every
participant. Scores ranged from 13 to 39, with a mean of 21.598 and a
participant SD of 5.540. The raw score was centred at the observed mean; no
within-site standardisation was used.

Nine participant-day outcomes covered light level, duration, continuous
periods, and exposure history. All use manuscript metric names and melEDI
units. Current definitions are documented in [Preparation 04](../preparation/04-light-metrics.qmd),
and the model-ready inputs in [Preparation 06](../preparation/06-analysis-datasets.qmd).
The longest continuous period is an observed lower bound unless its start and
end are both exactly identified; that stricter definition is examined as a
sensitivity analysis.

The predefined preparation sensitivity is called the
**gap-timing-unaware dataset**. It still passed the general 50%-per-hour and
80%-per-day coverage rules. The name means that the timing of the remaining
missing observations is not used for an additional metric-specific adjustment;
it does not mean that gaps, missingness, or coverage were ignored. For this
one-time first explanation, the primary could be interpreted as a
time-sensitive primary metric dataset because its preparation uses
the timing of remaining missing observations where that timing is relevant.
From here onward, it is called simply the **primary dataset**.

```{r}
#| label: tbl-h08-metrics
#| tbl-cap: "H08 light-exposure metrics, response specifications, and reported effect scales."

metric_registry |>
  arrange(metric_order) |>
  transmute(
    Metric = manuscript_name,
    Category = manuscript_category,
    Unit = display_unit,
    Model = recode(
      response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    Transform = recode(
      response_transform,
      log10_offset_0.1 = "log10(value + 0.1)",
      identity = "None"
    ),
    `Reported effect` = recode(
      effect_scale,
      ratio = "Ratio",
      difference = "Difference in hours"
    )
  ) |>
  h08_gt()
```

## Statistical models

Participant-day models include a participant random intercept nested within
site. This participant **random effect** represents remaining
between-participant variation after site and VLSQ-8 are considered. Fixed site
effects use the configured site display order and explicit sum contrasts, so the
additive VLSQ-8 coefficient is the site-average slope rather than the slope at
one reference site. The site-average likelihood-ratio test compares the
site-only and additive models. A separate test compares the additive and
interaction models to ask whether the VLSQ-8 association differs by site.

The following evaluated cell constructs every exact Wilkinson formula used
for the principal comparisons and the two structured sensitivities, then
presents their exact evaluated strings in one semantic table.

```{r}
#| label: tbl-h08-formulas
#| tbl-cap: "Exact evaluated Wilkinson formulas used in H08."
#| echo: true
#| code-fold: false

site_only <- stats::as.formula(
  "response_value ~ site + (1 | site:Id)"
)
additive <- stats::as.formula(
  "response_value ~ site + VLSQ8_c + (1 | site:Id)"
)
interaction <- stats::as.formula(
  "response_value ~ site * VLSQ8_c + (1 | site:Id)"
)
photoperiod_site_only <- stats::as.formula(
  "response_value ~ site + photoperiod_c + (1 | site:Id)"
)
photoperiod_additive <- stats::as.formula(
  "response_value ~ site + photoperiod_c + VLSQ8_c + (1 | site:Id)"
)
photoperiod_interaction <- stats::as.formula(
  "response_value ~ site * VLSQ8_c + photoperiod_c + (1 | site:Id)"
)
participant_site_only <- stats::as.formula(
  "participant_response ~ site"
)
participant_additive <- stats::as.formula(
  "participant_response ~ site + VLSQ8_c"
)
participant_interaction <- stats::as.formula(
  "participant_response ~ site * VLSQ8_c"
)

formula_names <- c(
  "site_only",
  "additive",
  "interaction",
  "photoperiod_site_only",
  "photoperiod_additive",
  "photoperiod_interaction",
  "participant_site_only",
  "participant_additive",
  "participant_interaction"
)
model_formula_strings <- c(
  "response_value ~ site + (1 | site:Id)",
  "response_value ~ site + VLSQ8_c + (1 | site:Id)",
  "response_value ~ site * VLSQ8_c + (1 | site:Id)",
  "response_value ~ site + photoperiod_c + (1 | site:Id)",
  "response_value ~ site + photoperiod_c + VLSQ8_c + (1 | site:Id)",
  "response_value ~ site * VLSQ8_c + photoperiod_c + (1 | site:Id)",
  "participant_response ~ site",
  "participant_response ~ site + VLSQ8_c",
  "participant_response ~ site * VLSQ8_c"
)
formula_objects_before_display <- mget(formula_names, inherits = FALSE)
formula_display <- tibble::tibble(
  Model = formula_names,
  `Evaluated Wilkinson formula` = unname(vapply(
    formula_objects_before_display,
    function(formula) {
      paste(deparse(formula, width.cutoff = 500L), collapse = " ")
    },
    character(1)
  ))
)
formula_table <- formula_display |>
  gt::gt(rowname_col = "Model") |>
  gt::opt_row_striping() |>
  gt::sub_missing(missing_text = "Not available") |>
  gt::cols_width(`Evaluated Wilkinson formula` ~ gt::pct(100)) |>
  gt::tab_options(
    table.width = gt::pct(100),
    table.font.size = gt::px(12),
    data_row.padding = gt::px(4),
    column_labels.font.weight = "600",
    container.overflow.x = "auto"
  )
formula_objects_after_display <- mget(formula_names, inherits = FALSE)
stopifnot(
  length(formula_objects_before_display) == 9L,
  identical(names(formula_objects_before_display), formula_names),
  identical(names(formula_objects_after_display), formula_names),
  identical(formula_objects_before_display, formula_objects_after_display),
  identical(formula_display$Model, formula_names),
  identical(
    formula_display$`Evaluated Wilkinson formula`,
    model_formula_strings
  ),
  inherits(formula_table, "gt_tbl")
)
formula_table
```

Site-average association and VLSQ-8-by-site interaction tests were adjusted
separately.
Near-eye and chest placements and the primary and gap-timing-unaware datasets
also had separate families, giving eight complete nine-metric FDR families.
Same-participant, same-day sensor and exact-sample sensitivities report
estimates and 95% CIs without creating additional *p*-value screens.

```{r}
#| label: tbl-h08-families
#| tbl-cap: "Complete H08 multiplicity families."

family_audit |>
  transmute(
    Family = family_id,
    Role = role,
    `Observed / planned tests` = sprintf(
      "%d / %d",
      observed_raw_p,
      planned_n
    ),
    Method = "FDR adjustment",
    `Raw p < 0.05` = raw_significant_n,
    `FDR-adjusted p < 0.05` = adjusted_significant_n
  ) |>
  h08_gt(
    note = paste(
      "All eight families were complete and independently verified.",
      "No FDR-adjusted p-value met the 0.050 criterion."
    )
  )
```

## Exact fitted samples

The primary near-eye models used 139–141 participants, 655–816
participant-days, and all nine sites, depending on the metric. Complementary
chest models used 153–154 participants, 743–902 participant-days, and eight
sites. Every model used its metric-specific complete frame.

```{r}
#| label: tbl-h08-primary-samples
#| tbl-cap: "Exact primary-dataset fitted samples by metric and placement."

sample_index |>
  filter(run_id %in% c(near_id, chest_id)) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Placement = placement_label,
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Placement) |>
  pivot_wider(names_from = Placement, values_from = Sample) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Cells are participants / participant-days / sites.",
      "Exact sample counts are retained in the source data."
    )
  )
```

The complete sample index is available as
[source data](../results/intermediate/model_data/H08/H08_model_frame_index.csv),
with [site-specific cells](../results/intermediate/model_data/H08/H08_model_frame_by_site.csv).

## Primary near-eye results

None of the nine average near-eye associations remained statistically
supported after the complete FDR adjustment. Several point estimates suggested
lower bright-light exposure at higher VLSQ-8 scores, but their uncertainty and
family-adjusted evidence do not support a confirmatory claim.

The corrected melEDI-dose ratio was 0.846 per VLSQ-8 SD (95% CI
0.716–0.999). Its likelihood-ratio raw *p* was 0.051 and FDR-adjusted *p* was
0.160. The Wald interval and likelihood-ratio test use different inferential
constructions; the complete-family conclusion is therefore inconclusive, not
statistically significant.

```{r}
#| label: fig-h08-near-eye-effects
#| fig-cap: "Primary near-eye site-average associations between VLSQ-8 and nine personal light-exposure metrics. Effects compare scores separated by one participant-level SD. The upper panel is an absolute difference in hours for time below 10 lx melEDI before sleep; the lower panel shows ratios and corresponding percentage changes for eight log-link outcomes. Horizontal bars are 95% Wald CIs; none of the nine associations retained FDR-adjusted support."
#| fig-alt: "Two separated forest-plot panels show nine primary near-eye site-average associations per VLSQ-8 standard deviation. The upper panel shows an absolute hour difference for time below 10 lx melEDI before sleep, with an interval crossing zero. The lower panel shows ratios and percentage changes for eight outcomes; most estimates indicate decreases, but none retained false-discovery-rate-adjusted support."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H08",
  "H08_near_eye_effects.png"
))
```

```{r}
#| label: tbl-h08-near-eye-results
#| tbl-cap: "Primary near-eye site-average VLSQ-8 associations, reported scales, 95% CIs, p-values, and exact fitted samples."

result_table(near, near_eye_association_wording = TRUE)
```

Figure and result sources are available as
[figure source data](../results/csv/source_data/H08/H08_near_eye_effects_data.csv)
and the [complete numerical result table](../results/tables/H08/H08_model_results_master.csv).

### Centred prediction contrast

Predictions at half an SD below and above the observed VLSQ-8 mean provide a
natural-scale description of the fitted contrast. They are not additional
hypothesis tests.

```{r}
#| label: tbl-h08-near-eye-predictions
#| tbl-cap: "Primary near-eye predictions at half an SD below and above the VLSQ-8 mean."

predictions |>
  filter(run_id == near_id) |>
  mutate(
    Point = recode(
      prediction_id,
      centre_minus_half_sd = "Lower score",
      centre_plus_half_sd = "Higher score"
    ),
    Prediction = sprintf(
      "%.3f (%.3f–%.3f)",
      estimate,
      conf_low,
      conf_high
    )
  ) |>
  select(metric_order, Metric = manuscript_name, Point, VLSQ8, Prediction) |>
  arrange(metric_order, Point) |>
  pivot_wider(
    names_from = Point,
    values_from = c(VLSQ8, Prediction),
    names_glue = "{Point}: {.value}"
  ) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Predictions are on each metric's natural scale with 95% delta-method",
      "confidence intervals and physical lower bounds applied."
    )
  )
```

## Complementary chest results

The complementary chest analysis repeated the same score, metric definitions,
response models, fixed-site adjustment, and nine-test family. No average chest
association remained after its separate FDR adjustment. The directional
pattern broadly resembled the near-eye result but does not create a second
primary finding.

```{r}
#| label: fig-h08-chest-effects
#| fig-cap: "Complementary chest site-average associations between VLSQ-8 and nine personal light-exposure metrics. Effects compare scores separated by one participant-level SD. The hour-difference and ratio outcomes remain on separate scales; horizontal bars are 95% Wald CIs. None retained FDR-adjusted support."
#| fig-alt: "Two separated forest-plot panels show nine complementary chest site-average associations per VLSQ-8 standard deviation. The upper hour-difference outcome has an interval crossing zero. Most lower-panel ratio estimates are below one, but none retained false-discovery-rate-adjusted support."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H08",
  "H08_chest_effects.png"
))
```

```{r}
#| label: tbl-h08-chest-results
#| tbl-cap: "Complementary chest VLSQ-8 associations, 95% confidence intervals, p-values, and exact fitted samples."

result_table(chest)
```

[Chest figure source data](../results/csv/source_data/H08/H08_chest_effects_data.csv)
contain the plotted values and intervals.

## VLSQ-8-by-site interactions

The separate interaction tests ask whether the association between VLSQ-8 and
each light metric differs across study sites. None of the nine near-eye or
nine chest VLSQ-8-by-site interaction tests remained after their
placement-specific FDR adjustments. The site-specific slopes are therefore
descriptive components of globally inconclusive interaction tests, not
independently supported site findings.

```{r}
#| label: tbl-h08-interactions
#| tbl-cap: "VLSQ-8-by-site interaction tests."

bind_rows(near, chest) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `LRT statistic (df)` = sprintf(
      "%.3f (%d)",
      interaction_lrt_statistic,
      interaction_lrt_df
    ),
    `Raw p` = format_p_cell(
      interaction_p_raw,
      interaction_adjusted_significant
    ),
    `FDR-adjusted p` = format_p_cell(
      interaction_p_adjusted,
      interaction_adjusted_significant
    )
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Near-eye and chest interaction tests form separate complete",
      "nine-metric families. No FDR-adjusted p-value met 0.050."
    )
  ) |>
  fmt_markdown(columns = c(`Raw p`, `FDR-adjusted p`))
```

All site-specific slopes and 95% intervals, using study site names and
order, are available in the
[site-slope source table](../results/tables/H08/H08_site_specific_slopes.csv).

## Same-participant, same-day sensor comparison

For the **paired-placement comparison**, the same participants and
participant-days are used at both sensor positions, and the near-eye and chest
models are fitted separately. It is not an equivalence test or a direct test
of a placement effect: it does not pool positions or estimate a within-day
difference. Ratio outcomes are shown on a common natural-log scale; the
pre-sleep hour difference remains in the adjacent table because the two units
must not share an axis.

```{r}
#| label: fig-h08-paired-placement
#| fig-cap: "Near-eye and chest VLSQ-8 associations on identical participant-day samples. The dashed diagonal marks equal point estimates and grey lines mark the null; it is not an equivalence analysis."
#| fig-alt: "Equal-axis scatter plot compares eight ratio-scale near-eye estimates on the horizontal axis with chest estimates on the vertical axis. Component 95% confidence intervals, horizontal and vertical null lines, and a dashed identity line are shown. Most estimates lie below zero on the log-ratio scale, but proximity to the diagonal does not establish equivalence."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H08",
  "H08_paired_placement_effects.png"
))
```

```{r}
#| label: tbl-h08-paired-placement
#| tbl-cap: "Exactly paired near-eye and chest component estimates."

master |>
  filter(run_id %in% c(near_paired_id, chest_paired_id)) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Placement = placement_label,
    Effect = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Placement) |>
  pivot_wider(
    names_from = Placement,
    values_from = c(Effect, Sample),
    names_glue = "{Placement}: {.value}"
  ) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Effects compare scores separated by one VLSQ-8 SD and include 95%",
      "Wald confidence intervals. Sample cells are participants /",
      "participant-days / sites."
    )
  )
```

[Paired-placement figure and table source data](../results/csv/source_data/H08/H08_paired_placement_effects_data.csv)
include the exact sample counts.

## Model checks and influence

All 36 primary-dataset additive and interaction fits converged, had
positive-definite Hessians, were nonsingular, and retained full-rank fixed
effects. No primary or complementary effect was non-estimable. The common
model checks retained all nine prespecified response models; review flags
require qualification, not a data-dependent change of response family.

```{r}
#| label: tbl-h08-response-check
#| tbl-cap: "Response-model dispositions across the four primary-dataset inferential targets."

response_check |>
  arrange(metric_order) |>
  transmute(
    Metric = manuscript_name,
    `Response / transform` = paste(
      recode(
        response_family,
        gaussian = "Gaussian",
        tweedie_log = "Tweedie, log link"
      ),
      recode(
        response_transform,
        log10_offset_0.1 = "log10(value + 0.1)",
        identity = "untransformed"
      ),
      sep = " / "
    ),
    `Targets reviewed` = inferential_targets,
    `Review flags` = review_targets,
    Decision = str_to_sentence(str_replace_all(
      family_check_status,
      "_",
      " "
    ))
  ) |>
  h08_gt(
    note = paste(
      "The targets are near-eye and chest site-average and interaction fits.",
      "No major failure or non-estimable target occurred."
    )
  )
```

Three limitations are material. Darkest 10 h mean showed increasing residual
spread with fitted values (absolute-residual versus fitted Spearman
correlations 0.341 near eye and 0.349 chest). Time below 1 lx melEDI during
sleep contained four exact zeros at each placement, whereas the fitted
Tweedie models assigned near-zero expected counts. Corrected melEDI dose
passed the numerical checks, but its value depends on the timing of remaining
missing observations and is interpreted with the dataset and observed-dose
sensitivities below.

```{r}
#| label: tbl-h08-flagged-diagnostics
#| tbl-cap: "Primary-dataset model checks requiring explicit qualification."

diagnostics |>
  filter(
    run_id %in% c(near_id, chest_id),
    diagnostic_status != "PASS"
  ) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `Residual |fit| Spearman` = sprintf(
      "%+.3f",
      residual_absolute_fitted_spearman
    ),
    `Observed / expected zeros` = if_else(
      response_family == "tweedie_log",
      sprintf("%d / %.3g", observed_zero_n, expected_zero_n),
      "Not applicable"
    ),
    Decision = str_to_sentence(str_replace_all(
      diagnostic_issues,
      "_",
      " "
    ))
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "All convergence, Hessian, singularity, fixed-rank, serial-correlation,",
      "and prediction-bound fields are retained in the complete source table."
    )
  )
```

```{r}
#| label: fig-h08-model-adequacy
#| fig-cap: "Selected primary near-eye model checks for all nine metrics. Panels retain their model-specific transformed or Pearson-residual scales and are diagnostic rather than effect displays."
#| fig-alt: "Multi-panel diagnostic figure for nine primary near-eye models shows residual-versus-fitted patterns and distribution checks. Darkest 10 hour mean has the clearest increase in residual spread, and the sleep-below-1-lux model has a small observed zero mass not reproduced by its Tweedie fit."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H08",
  "H08_near_eye_model_adequacy.png"
))
```

The [complete model-check table](../results/csv/diagnostics/H08/H08_model_diagnostics.csv)
and [model-check figure source data](../results/csv/source_data/H08/H08_near_eye_model_adequacy_data.csv)
contain the full checks.

All 153 planned leave-one-site-out refits succeeded. Some estimates changed
direction or magnitude when one site was omitted; this reinforces the
inconclusive interpretation but does not identify a site for exclusion. No
site or participant was removed after influence review.

```{r}
#| label: tbl-h08-site-influence
#| tbl-cap: "Leave-one-site-out stability of the primary-dataset average associations."

loo_summary |>
  arrange(metric_order, placement_label) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    Refits = sprintf("%d / %d", successful_refits, refits),
    `Sign reversal` = if_else(sign_reversal_any, "Yes", "No"),
    `Largest relative |change|` = sprintf(
      "%.2f",
      maximum_relative_absolute_change
    ),
    `Most influential omitted site` = most_influential_omitted_site,
    Interpretation = influence_status
  ) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Relative changes can be large when the full-sample estimate is close",
      "to zero; they do not by themselves identify an erroneous site."
    )
  )
```

The ranked
[participant-influence screen](../results/csv/diagnostics/H08/H08_participant_influence_screen.csv)
records the influence check and does not define an automatic deletion rule.

## Dataset sensitivity

The strongest comparison fits the primary and gap-timing-unaware values on
identical participant-day keys for every metric and placement. All 18 of
these exact-common-sample comparisons were stable within model uncertainty.
The scenario-specific all-available families also retained the same
FDR-adjusted conclusion. Near-eye wake time above 250 lx melEDI and corrected
melEDI dose were precision-sensitive in the all-available comparison because
zero-exclusion by a component 95% interval changed; all chest comparisons
were stable. Neither classification creates a new hypothesis test.

```{r}
#| label: tbl-h08-gap-common-sample
#| tbl-cap: "Primary and gap-timing-unaware associations on exact common participant-day samples."

gap_common_display <- gap_sensitivity |>
  filter(sample_scenario == "main_gap_common_sample") |>
  left_join(
    metric_registry |>
      select(metric_id, response_transform, effect_scale),
    by = "metric_id"
  ) |>
  mutate(
    estimate_main_display = model_to_practical(
      estimate_main,
      response_transform,
      effect_scale
    ),
    conf_low_main_display = model_to_practical(
      conf_low_main,
      response_transform,
      effect_scale
    ),
    conf_high_main_display = model_to_practical(
      conf_high_main,
      response_transform,
      effect_scale
    ),
    estimate_gap_display = model_to_practical(
      estimate_gap_timing_unaware,
      response_transform,
      effect_scale
    ),
    conf_low_gap_display = model_to_practical(
      conf_low_gap_timing_unaware,
      response_transform,
      effect_scale
    ),
    conf_high_gap_display = model_to_practical(
      conf_high_gap_timing_unaware,
      response_transform,
      effect_scale
    )
  )

stopifnot(
  nrow(gap_common_display) == 18L,
  all(gap_common_display$exact_common_keys),
  all(
    gap_common_display$stability_classification ==
      "stable within model uncertainty"
  )
)

gap_common_display |>
  transmute(
    metric_order,
    Placement = if_else(placement == "glasses", "Near eye", "Chest"),
    Metric = manuscript_name,
    `Primary effect (95% CI)` = format_effect(
      estimate_main_display,
      conf_low_main_display,
      conf_high_main_display,
      effect_scale
    ),
    `Gap-timing-unaware effect (95% CI)` = format_effect(
      estimate_gap_display,
      conf_low_gap_display,
      conf_high_gap_display,
      effect_scale
    ),
    Sample = format_sample(
      participants_main,
      participant_days_main,
      sites_main
    ),
    Stability = stability_classification
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Effects compare scores separated by one VLSQ-8 SD and include 95%",
      "Wald intervals. Samples are participants / participant-days / sites."
    )
  )
```

```{r}
#| label: fig-h08-gap-common-sample
#| fig-cap: "Primary and gap-timing-unaware VLSQ-8 associations on identical participant-day samples, shown separately for near-eye and chest placements."
#| fig-alt: "Paired comparison of primary and gap-timing-unaware estimates on identical samples for nine metrics at each placement. Estimates are generally close; all eighteen comparisons are classified as stable within model uncertainty."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H08",
  "H08_gap_common_sample_effects.png"
))
```

[Dataset-sensitivity figure source data](../results/csv/source_data/H08/H08_gap_common_sample_effects_data.csv)
and the [complete comparison table](../results/tables/H08/H08_gap_timing_unaware_sensitivity.csv)
retain the sample counts and all-available classifications.

## Other structured sensitivities

Linear centred photoperiod was added only as a sensitivity covariate, with
site retained and no photoperiod interaction. Participant-summary models give
each participant one outcome value and use the corresponding fixed-effects
model. All 36 additive sensitivity fits converged and retained full-rank
fixed effects. Their estimates were broadly compatible with the
participant-day results and do not form new multiplicity families.

```{r}
#| label: tbl-h08-photoperiod-participant-summary
#| tbl-cap: "Primary, photoperiod-adjusted, and participant-summary estimates."

primary_sensitivity_display <- bind_rows(near, chest) |>
  select(
    placement,
    placement_label,
    metric_order,
    metric_id,
    manuscript_name,
    effect_type,
    primary_estimate = estimate_practical_per_sd,
    primary_low = conf_low_practical_per_sd,
    primary_high = conf_high_practical_per_sd
  ) |>
  left_join(
    photoperiod |>
      select(
        placement,
        metric_id,
        photo_estimate = estimate_practical_per_sd,
        photo_low = conf_low_practical_per_sd,
        photo_high = conf_high_practical_per_sd,
        photo_converged = converged,
        photo_hessian = positive_definite_hessian,
        photo_full_rank = fixed_full_rank
      ),
    by = c("placement", "metric_id")
  ) |>
  left_join(
    participant_summary |>
      select(
        placement,
        metric_id,
        participant_estimate = estimate_practical_per_sd,
        participant_low = conf_low_practical_per_sd,
        participant_high = conf_high_practical_per_sd,
        participant_converged = converged,
        participant_hessian = positive_definite_hessian,
        participant_full_rank = fixed_full_rank
      ),
    by = c("placement", "metric_id")
  )

stopifnot(
  nrow(primary_sensitivity_display) == nrow(near) + nrow(chest),
  !anyDuplicated(primary_sensitivity_display[c("placement", "metric_id")])
)

sensitivity_fit_qualification <- function(converged, hessian, full_rank) {
  issues <- c(
    if (!isTRUE(converged)) "optimizer did not converge",
    if (!isTRUE(hessian)) "non-positive-definite Hessian",
    if (!isTRUE(full_rank)) "rank-deficient fixed effects"
  )
  paste(issues, collapse = "; ")
}
photo_qualification <- mapply(
  sensitivity_fit_qualification,
  primary_sensitivity_display$photo_converged,
  primary_sensitivity_display$photo_hessian,
  primary_sensitivity_display$photo_full_rank,
  USE.NAMES = FALSE
)
participant_qualification <- mapply(
  sensitivity_fit_qualification,
  primary_sensitivity_display$participant_converged,
  primary_sensitivity_display$participant_hessian,
  primary_sensitivity_display$participant_full_rank,
  USE.NAMES = FALSE
)
fit_qualification <- mapply(function(photo, participant) {
  issues <- c(
    if (nzchar(photo)) paste0("Photoperiod: ", photo),
    if (nzchar(participant)) paste0("Participant summary: ", participant)
  )
  if (length(issues)) paste(issues, collapse = ". ") else "None"
}, photo_qualification, participant_qualification, USE.NAMES = FALSE)

primary_sensitivity_display |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `Primary effect (95% CI)` = format_effect(
      primary_estimate,
      primary_low,
      primary_high,
      effect_type
    ),
    `Photoperiod-adjusted effect (95% CI)` = format_effect(
      photo_estimate,
      photo_low,
      photo_high,
      effect_type
    ),
    `Participant-summary effect (95% CI)` = format_effect(
      participant_estimate,
      participant_low,
      participant_high,
      effect_type
    ),
    `Fit qualifications` = fit_qualification
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "All entries are effects per one VLSQ-8 SD with 95% Wald intervals.",
      "Sensitivity-specific samples are retained in source data.",
      "Intervals from fits with convergence or Hessian qualifications are",
      "shown for transparency and should not be treated as reliable inference."
    )
  )
```

The
[photoperiod-adjusted](../results/tables/H08/H08_photoperiod_sensitivity.csv)
and
[participant-summary](../results/tables/H08/H08_participant_summary_sensitivity.csv)
tables contain exact sample counts.

### Metric-definition sensitivities

The primary longest-period metric is an observed lower bound. Requiring exact
identification reduced the sample and produced imprecise estimates: near-eye
ratio 0.958 (95% CI 0.856–1.072) and chest ratio 0.939 (95% CI
0.855–1.032). On the exactly paired placement sample, the ratios were 0.943
near eye and 0.952 chest, with both intervals crossing one.

The observed-dose comparison holds the corrected-dose fitted sample fixed and
replaces the value with uncorrected observed dose. It is descriptive and has
no additional *p*-value family. The all-available near-eye ratio was 0.842
(95% CI 0.713–0.995); its paired-sample interval crossed one, and the chest
all-available interval included one. These component intervals do not
override the primary FDR family.

```{r}
#| label: tbl-h08-metric-definition-sensitivities
#| tbl-cap: "Exactly identified longest-period and observed-dose sensitivities."

bind_rows(
  exact_longest |>
    mutate(Sensitivity = "Exactly identified longest period"),
  observed_dose |>
    mutate(Sensitivity = "Observed dose on corrected-dose sample")
) |>
  transmute(
    metric_order,
    Sensitivity,
    Placement = placement_label,
    `Sample basis` = recode(
      sample_scenario,
      all_available = "All available",
      paired_common_sample = "Paired placement"
    ),
    `Effect per VLSQ-8 SD (95% CI)` = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Sensitivity, `Sample basis`, Placement) |>
  select(-metric_order) |>
  h08_gt(
    note = paste(
      "Intervals are 95% Wald intervals. These descriptive sensitivities",
      "do not create additional p-value screens."
    )
  )
```

Exact sample counts are available for the
[longest-period](../results/tables/H08/H08_exactly_identified_longest_period_sensitivity.csv)
and [observed-dose](../results/tables/H08/H08_observed_dose_sensitivity.csv)
sensitivities.

## Interpretation

On the selected score, metric, sample, and model definitions, the observed
data provide no FDR-adjusted evidence that VLSQ-8 score is associated with any
of the nine primary near-eye personal light-exposure metrics. The
complementary chest analyses and the central sensitivity analyses do not
materially strengthen that evidence. Several estimates point toward lower
bright-light exposure at higher scores, but their uncertainty, complete-family
adjustment, and site influence make that directional pattern inconclusive.

This conclusion is not evidence that every true association is exactly zero,
and the paired-placement comparison is not an equivalence test. The analysis
is observational: it does not establish that visual light sensitivity changes
personal light exposure, nor does it support physiological or health-effect
claims.

## Preregistration deviations {#h08-preregistration-deviations}

- [H08 primary test estimand](../preregistration-deviations.qmd#h08-primary-test-estimand): each metric's
  study-site-average VLSQ-8 association is estimated separately from its
  VLSQ-8-by-site interaction;
- [H08 multiplicity](../preregistration-deviations.qmd#h08-multiplicity): average associations
  and site interactions use separate complete nine-metric FDR families, with
  no retained result in either family.
