---
title: "H09: Chronotype and timing of personal light exposure"
engine: knitr
format:
  html:
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

This analysis relates chronotype, measured with MCTQ corrected midsleep and MEQ preference, to the timing of personal light exposure. Models distinguish the average association across sites from site-specific differences.

## Data and model guide {#methods}

The [questionnaire preparation](../preparation/05-questionnaires-and-diaries.qmd)
supplies corrected midsleep on free days (MCTQ MSFsc) and morningness-eveningness
preference (MEQ). They remain separate predictors. The [metric datasets](../preparation/06-analysis-datasets.qmd)
provide five supported timing outcomes with participant, date and site keys.
Effects are per one hour later MSFsc or ten points greater MEQ morning preference.
The exact clock transformations and complete samples are constructed below.

For each instrument and outcome, a site-only model is compared with an additive
chronotype model, then with a chronotype-by-site interaction. Participant random
intercepts account for repeated days. Maximum-likelihood fits provide nested
likelihood-ratio tests; restricted maximum likelihood provides final additive
coefficients and Wald intervals. Signed effects are in clock hours: positive is
later and negative is earlier.

Four separate five-outcome FDR families cover average and interaction associations
for each instrument, with separate instances for placement, preprocessing and
matched samples. Site-specific slopes remain descriptive unless the global
interaction supports heterogeneity. The sensitivity analyses address within-site
photoperiod, participant summaries, linearity, continuous-time residual dependence,
influence and exact common samples. Sample ranges accompanying supported effects
refer to those displayed outcomes; full metric-specific samples are reported below.

The executable sections below write fitted objects to `results/models/H09/`,
reader tables to `results/tables/H09/`, and the supporting numerical and figure
outputs under `results/`. Start with the calculations below or jump to
[findings and interpretation](#findings).

## Setup

Load the reusable sample, formula, fit and diagnostic functions. The sections below perform each analysis and write its numerical results.

```{r}
#| label: setup-h09

source("scripts/project.R")
analysis_setup()
source("scripts/pipeline/paths_io.R")
source("scripts/pipeline/p_value_display.R")
source("scripts/hypotheses/H09/h09_contract.R")
source("scripts/hypotheses/H09/h09_modeling.R")
root <- normalizePath(Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()))
producer <- "analyses/H09-chronotype.qmd"
options(stringsAsFactors = FALSE, scipen = 999, dplyr.summarise.inform = FALSE)
roots <- list(model_data = file.path(root,"results/intermediate/model_data/H09"), models = file.path(root,"results/models/H09"), diagnostics = file.path(root,"results/csv/diagnostics/H09"), tables = file.path(root,"results/tables/H09"), figures = file.path(root,"results/images/H09"), source_data = file.path(root,"results/csv/source_data/H09"))
invisible(lapply(roots, dir.create, recursive=TRUE, showWarnings=FALSE))
metadata <- list()
source("scripts/pipeline/multiplicity.R")
h09_write_csv <- function(data, path) {
  write_csv_artifact(data, path, producer = producer)
  invisible(path)
}

h09_write_rds <- function(object, path) {
  write_rds_artifact(object, path, producer = producer)
  invisible(path)
}
input_contract <- h09_input_contract(root)
```



## Chronotype and timing samples

Convert corrected midsleep to hours and scale MEQ in ten-point increments. Build matched samples for the two sensor placements and the alternative preprocessing comparison, while retaining metric-specific support.

```{r}
#| label: h09-data

metric_registry <- h09_metric_registry()

predictor_registry <- h09_predictor_registry()

run_registry <- h09_run_registry()

family_registry <- h09_family_registry()

formula_registry <- h09_formula_registry()

diagnostic_thresholds <- h09_diagnostic_thresholds()

site_registry <- readr::read_csv(
  file.path(root, "config/site_display_registry.csv"),
  show_col_types = FALSE
) |>
  dplyr::arrange(.data$display_order)

site_levels <- site_registry$site

metric_display <- readr::read_csv(
  file.path(root, "config/metric_display_registry.csv"),
  show_col_types = FALSE
)

metric_display_audit <- metric_registry |>
  dplyr::left_join(
    metric_display |>
      dplyr::select(
        .data$metric_id,
        registry_manuscript_name = .data$manuscript_name,
        registry_analysis_unit = .data$analysis_unit,
        registry_display_unit = .data$display_unit
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    registry_status = dplyr::case_when(
      .data$metric_id == "longest_period_midpoint" &
        is.na(.data$registry_manuscript_name) ~
        "H09-owned registered construction; shared registry row unavailable",
      .data$manuscript_name == .data$registry_manuscript_name &
        .data$registry_analysis_unit == "participant-day" ~ "PASS",
      TRUE ~ "FAIL"
    )
  )

if (any(metric_display_audit$registry_status == "FAIL")) {
  h09_abort("The H09 display contract differs from the shared registry")
}

chronotype <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "normalized_chronotype"
])

score_contract <- h09_score_contract(chronotype)

score_audit <- tibble::tibble(
  participants = nrow(chronotype),
  sites = dplyr::n_distinct(chronotype$site),
  mctq_complete = sum(!is.na(chronotype$msf_sc)),
  mctq_missing = sum(is.na(chronotype$msf_sc)),
  mctq_center_hour = mean(
    as.numeric(chronotype$msf_sc) / 3600,
    na.rm = TRUE
  ),
  mctq_min_hour = min(as.numeric(chronotype$msf_sc) / 3600, na.rm = TRUE),
  mctq_max_hour = max(as.numeric(chronotype$msf_sc) / 3600, na.rm = TRUE),
  meq_complete = sum(!is.na(chronotype$meq)),
  meq_missing = sum(is.na(chronotype$meq)),
  meq_center_score = mean(chronotype$meq, na.rm = TRUE),
  meq_min_score = min(chronotype$meq, na.rm = TRUE),
  meq_max_score = max(chronotype$meq, na.rm = TRUE),
  item_level_reconstruction = "unavailable",
  score_definition = score_contract$score_definition
)

h09_write_csv(
  metric_registry,
  file.path(roots$model_data, "H09_metric_registry.csv")
)

h09_write_csv(
  predictor_registry,
  file.path(roots$model_data, "H09_predictor_registry.csv")
)

h09_write_csv(
  run_registry,
  file.path(roots$model_data, "H09_run_registry.csv")
)

h09_write_csv(
  family_registry,
  file.path(roots$model_data, "H09_family_registry.csv")
)

h09_write_csv(
  formula_registry,
  file.path(roots$model_data, "H09_formula_registry.csv")
)

h09_write_csv(
  diagnostic_thresholds,
  file.path(roots$diagnostics, "H09_diagnostic_thresholds.csv")
)

h09_write_csv(
  metric_display_audit,
  file.path(roots$model_data, "H09_metric_display_audit.csv")
)

h09_write_csv(
  score_audit,
  file.path(roots$diagnostics, "H09_chronotype_score_audit.csv")
)

primary_near_eye_source <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "primary_near_eye_enriched"
])

primary_chest_source <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "primary_chest_enriched"
])

gap_source <- readRDS(input_contract$absolute_path[
  input_contract$input_role == "gap_timing_unaware_metrics"
])

score_value_audit <- dplyr::bind_rows(
  primary_near_eye_source |>
    dplyr::transmute(
      placement = "glasses",
      .data$site,
      .data$Id,
      source_mctq = as.numeric(.data$msf_sc),
      source_meq = as.numeric(.data$meq)
    ),
  primary_chest_source |>
    dplyr::transmute(
      placement = "chest",
      .data$site,
      .data$Id,
      source_mctq = as.numeric(.data$msf_sc),
      source_meq = as.numeric(.data$meq)
    )
) |>
  dplyr::distinct() |>
  dplyr::left_join(
    chronotype |>
      dplyr::transmute(
        .data$site,
        .data$Id,
        normalized_mctq = as.numeric(.data$msf_sc),
        normalized_meq = as.numeric(.data$meq)
      ),
    by = c("site", "Id"),
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    mctq_value_matches = dplyr::if_else(
      is.na(.data$source_mctq) & is.na(.data$normalized_mctq),
      TRUE,
      .data$source_mctq == .data$normalized_mctq,
      missing = FALSE
    ),
    meq_value_matches = dplyr::if_else(
      is.na(.data$source_meq) & is.na(.data$normalized_meq),
      TRUE,
      .data$source_meq == .data$normalized_meq,
      missing = FALSE
    )
  )

if (
  any(!score_value_audit$mctq_value_matches) ||
    any(!score_value_audit$meq_value_matches)
) {
  h09_abort("Enriched H09 score values differ from the chronotype input")
}

h09_write_csv(
  score_value_audit,
  file.path(roots$diagnostics, "H09_chronotype_value_audit.csv")
)

primary_near_eye <- h09_prepare_primary_wide(
  primary_near_eye_source,
  score_contract
)

primary_chest <- h09_prepare_primary_wide(
  primary_chest_source,
  score_contract
)

primary_rows <- dplyr::bind_rows(
  h09_primary_long(primary_near_eye, "glasses", metric_registry),
  h09_primary_long(primary_chest, "chest", metric_registry)
)

gap_rows <- h09_prepare_gap_long(gap_source, chronotype, score_contract)

all_rows <- dplyr::bind_rows(primary_rows, gap_rows)

if (anyDuplicated(all_rows[c(
  "data_scenario_id", "placement", "site", "Id", "local_date", "metric_id"
)])) {
  h09_abort("H09 long rows contain duplicate scenario-placement-day metrics")
}

h09_rows_for_run <- function(run, metric_id, instrument_id) {
  selected <- all_rows |>
    dplyr::filter(
      .data$data_scenario_id == run$data_scenario_id,
      .data$placement == run$placement,
      .data$metric_id == .env$metric_id
    )
  frame <- h09_prepare_model_frame(selected, instrument_id, site_levels)
  if (nrow(frame) == 0L) return(frame)
  if (run$sample_scenario == "paired_common") {
    opposite <- if (run$placement == "glasses") "chest" else "glasses"
    other <- all_rows |>
      dplyr::filter(
        .data$data_scenario_id == run$data_scenario_id,
        .data$placement == opposite,
        .data$metric_id == .env$metric_id
      )
    other_frame <- h09_prepare_model_frame(other, instrument_id, site_levels)
    common <- intersect(frame$.model_row_id, other_frame$.model_row_id)
    frame <- frame[frame$.model_row_id %in% common, , drop = FALSE]
  }
  if (run$sample_scenario == "gap_common") {
    other_scenario <- if (run$data_scenario_id == "primary") {
      "gap_timing_unaware"
    } else {
      "primary"
    }
    other <- all_rows |>
      dplyr::filter(
        .data$data_scenario_id == other_scenario,
        .data$placement == run$placement,
        .data$metric_id == .env$metric_id
      )
    other_frame <- h09_prepare_model_frame(other, instrument_id, site_levels)
    common <- intersect(frame$.model_row_id, other_frame$.model_row_id)
    frame <- frame[frame$.model_row_id %in% common, , drop = FALSE]
  }
  if (nrow(frame) > 0L) {
    frame$site <- droplevels(frame$site)
    stats::contrasts(frame$site) <- stats::contr.sum(nlevels(frame$site))
    frame$Id <- droplevels(frame$Id)
    frame$participant_key <- droplevels(frame$participant_key)
  }
  tibble::as_tibble(frame)
}

model_frames <- list()

frame_index_rows <- list()

frame_site_rows <- list()

unavailable_rows <- list()

for (run_index in seq_len(nrow(run_registry))) {
  run <- run_registry[run_index, ]
  available_metrics <- metric_registry$metric_id
  if (
    run$data_scenario_id == "gap_timing_unaware" ||
      run$sample_scenario == "gap_common"
  ) {
    available_metrics <- setdiff(
      available_metrics,
      "longest_period_midpoint"
    )
  }
  for (metric_id in available_metrics) {
    for (instrument_id in predictor_registry$instrument_id) {
      frame <- h09_rows_for_run(run, metric_id, instrument_id)
      if (nrow(frame) == 0L) {
        h09_abort(
          "No H09 model rows for %s / %s / %s",
          run$run_id,
          metric_id,
          instrument_id
        )
      }
      frame_id <- paste(run$run_id, metric_id, instrument_id, sep = "__")
      model_frames[[frame_id]] <- frame
      sample <- h09_sample_summary(frame)
      frame_index_rows[[frame_id]] <- dplyr::bind_cols(
        tibble::tibble(
          frame_id = frame_id,
          run_order = run$run_order,
          run_id = run$run_id,
          data_scenario_id = run$data_scenario_id,
          reader_scenario = run$reader_scenario,
          placement = run$placement,
          placement_label = run$placement_label,
          sample_scenario = run$sample_scenario,
          analytical_role = run$analytical_role,
          metric_id = metric_id,
          instrument_id = instrument_id,
          availability = "ESTIMABLE",
          row_key_hash = h09_key_hash(frame)
        ),
        sample
      )
      frame_site_rows[[frame_id]] <- dplyr::bind_cols(
        tibble::tibble(
          frame_id = frame_id,
          run_id = run$run_id,
          metric_id = metric_id,
          instrument_id = instrument_id
        )[rep(1L, nlevels(frame$site)), , drop = FALSE],
        h09_sample_by_site(frame)
      )
    }
  }
  if (!"longest_period_midpoint" %in% available_metrics) {
    for (instrument_id in predictor_registry$instrument_id) {
      key <- paste(run$run_id, instrument_id, sep = "__")
      unavailable_rows[[key]] <- tibble::tibble(
        run_id = run$run_id,
        data_scenario_id = run$data_scenario_id,
        placement = run$placement,
        placement_label = run$placement_label,
        sample_scenario = run$sample_scenario,
        metric_id = "longest_period_midpoint",
        instrument_id = instrument_id,
        availability = "NON_ESTIMABLE",
        reason = paste(
          "The alternative-preprocessing input contains neither the",
          "registered longest-period midpoint nor its selected-period endpoints"
        )
      )
    }
  }
}

model_frame_index <- dplyr::bind_rows(frame_index_rows) |>
  dplyr::arrange(.data$run_order, .data$metric_id, .data$instrument_id)

model_frame_by_site <- dplyr::bind_rows(frame_site_rows)

non_estimable_targets <- dplyr::bind_rows(unavailable_rows)

paired_sample_audit <- model_frame_index |>
  dplyr::filter(.data$sample_scenario == "paired_common") |>
  dplyr::select(
    .data$metric_id,
    .data$instrument_id,
    .data$placement,
    .data$participants,
    .data$participant_days,
    .data$row_key_hash
  ) |>
  tidyr::pivot_wider(
    names_from = .data$placement,
    values_from = c(
      .data$participants,
      .data$participant_days,
      .data$row_key_hash
    )
  ) |>
  dplyr::mutate(
    exact_counts_match = .data$participants_glasses ==
      .data$participants_chest &
      .data$participant_days_glasses == .data$participant_days_chest,
    exact_row_keys_match = .data$row_key_hash_glasses ==
      .data$row_key_hash_chest
  )

if (
  any(!paired_sample_audit$exact_counts_match) ||
    any(!paired_sample_audit$exact_row_keys_match)
) {
  h09_abort("An H09 paired placement frame does not use identical row keys")
}

gap_common_sample_audit <- model_frame_index |>
  dplyr::filter(.data$sample_scenario == "gap_common") |>
  dplyr::select(
    .data$metric_id,
    .data$instrument_id,
    .data$placement,
    .data$data_scenario_id,
    .data$participants,
    .data$participant_days,
    .data$row_key_hash
  ) |>
  tidyr::pivot_wider(
    names_from = .data$data_scenario_id,
    values_from = c(
      .data$participants,
      .data$participant_days,
      .data$row_key_hash
    )
  ) |>
  dplyr::mutate(
    exact_counts_match = .data$participants_primary ==
      .data$participants_gap_timing_unaware &
      .data$participant_days_primary ==
        .data$participant_days_gap_timing_unaware,
    exact_row_keys_match = .data$row_key_hash_primary ==
      .data$row_key_hash_gap_timing_unaware
  )

if (
  any(!gap_common_sample_audit$exact_counts_match) ||
    any(!gap_common_sample_audit$exact_row_keys_match)
) {
  h09_abort("An H09 primary/gap common frame does not use identical row keys")
}

h09_write_csv(
  model_frame_index,
  file.path(roots$model_data, "H09_model_frame_index.csv")
)

h09_write_csv(
  model_frame_by_site,
  file.path(roots$model_data, "H09_model_frame_by_site.csv")
)

h09_write_csv(
  non_estimable_targets,
  file.path(roots$model_data, "H09_non_estimable_targets.csv")
)

h09_write_csv(
  paired_sample_audit,
  file.path(roots$model_data, "H09_paired_sample_audit.csv")
)

h09_write_csv(
  gap_common_sample_audit,
  file.path(roots$model_data, "H09_gap_common_sample_audit.csv")
)

h09_write_rds(
  model_frames,
  file.path(roots$model_data, "H09_model_frames.rds")
)
score_audit
model_frame_index
```



## Fit the main models

Estimate site-adjusted chronotype slopes with participant random intercepts, compare site interactions, and adjust the prespecified families of tests for multiplicity.

```{r}
#| label: h09-models

model_bundles <- list()

fit_index_rows <- list()

test_rows <- list()

effect_rows <- list()

site_slope_rows <- list()

for (index in seq_len(nrow(model_frame_index))) {
  context <- model_frame_index[index, ]
  frame <- model_frames[[context$frame_id]]
  bundle <- h09_fit_bundle(frame, context$instrument_id)
  model_bundles[[context$frame_id]] <- bundle

  fit_registry <- list(
    ML_M0_site_only = bundle$ml$M0_site_only,
    ML_M1_main = bundle$ml$M1_main,
    ML_M2_interaction = bundle$ml$M2_interaction,
    REML_M1_main = bundle$reml$M1_main,
    REML_M2_interaction = bundle$reml$M2_interaction
  )
  fit_index_rows[[context$frame_id]] <- dplyr::bind_rows(lapply(
    names(fit_registry),
    function(fit_id) {
      dplyr::bind_cols(
        context |>
          dplyr::select(
            .data$frame_id,
            .data$run_id,
            .data$metric_id,
            .data$instrument_id,
            .data$observations,
            .data$row_key_hash
          ),
        tibble::tibble(
          fit_id = fit_id,
          estimation = if (startsWith(fit_id, "ML_")) "ML" else "REML"
        ),
        h09_model_fit_status(fit_registry[[fit_id]])
      )
    }
  ))

  tests <- h09_bundle_tests(bundle)
  predictor_row <- predictor_registry |>
    dplyr::filter(.data$instrument_id == context$instrument_id)
  metric_row <- metric_registry |>
    dplyr::filter(.data$metric_id == context$metric_id)
  tests <- tests |>
    dplyr::mutate(
      family_base = dplyr::case_when(
        !metric_row$primary_family_member ~ NA_character_,
        .data$comparison_id == "main" ~ predictor_row$main_family,
        TRUE ~ predictor_row$interaction_family
      ),
      family_id = dplyr::if_else(
        is.na(.data$family_base),
        NA_character_,
        paste(.data$family_base, context$run_id, sep = "__")
      )
    )
  test_rows[[context$frame_id]] <- dplyr::bind_cols(
    context[rep(1L, nrow(tests)), , drop = FALSE],
    tests
  )

  effect <- h09_effect_summary(
    bundle$reml$M1_main,
    context$instrument_id
  )
  performance <- h09_performance_summary(bundle$reml$M1_main$model)
  effect_rows[[context$frame_id]] <- dplyr::bind_cols(
    context,
    effect,
    performance
  )

  slopes <- h09_site_slopes(
    bundle$reml$M2_interaction,
    frame,
    context$instrument_id
  )
  site_slope_rows[[context$frame_id]] <- dplyr::bind_cols(
    context[rep(1L, nrow(slopes)), , drop = FALSE],
    slopes
  )

  if (index %% 10L == 0L || index == nrow(model_frame_index)) {
    message("H09 fitted frame ", index, " / ", nrow(model_frame_index))
  }
}

model_fit_index <- dplyr::bind_rows(fit_index_rows)

model_tests <- dplyr::bind_rows(test_rows)

model_effects <- dplyr::bind_rows(effect_rows)

site_specific_slopes <- dplyr::bind_rows(site_slope_rows)

non_estimable_tests <- non_estimable_targets |>
  tidyr::crossing(comparison_id = c("main", "interaction")) |>
  dplyr::left_join(
    run_registry |>
      dplyr::select(
        .data$run_id,
        .data$run_order,
        .data$reader_scenario,
        .data$analytical_role
      ),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  dplyr::left_join(
    predictor_registry |>
      dplyr::select(
        .data$instrument_id,
        .data$main_family,
        .data$interaction_family
      ),
    by = "instrument_id",
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    frame_id = NA_character_,
    family_base = dplyr::if_else(
      .data$comparison_id == "main",
      .data$main_family,
      .data$interaction_family
    ),
    family_id = paste(.data$family_base, .data$run_id, sep = "__"),
    chi_square = NA_real_,
    df = NA_real_,
    p_raw = NA_real_,
    comparison_status = "NON_ESTIMABLE",
    comparison_error = .data$reason,
    participants = 0L,
    participant_days = 0L,
    observations = 0L,
    sites = 0L,
    derivation_hours = NA_real_,
    min_days_per_participant = NA_integer_,
    median_days_per_participant = NA_real_,
    max_days_per_participant = NA_integer_,
    row_key_hash = NA_character_
  ) |>
  dplyr::select(dplyr::all_of(names(model_tests)))

model_tests <- dplyr::bind_rows(model_tests, non_estimable_tests) |>
  dplyr::group_by(
    .data$run_id,
    .data$instrument_id,
    .data$comparison_id,
    .data$family_id
  ) |>
  dplyr::mutate(
    p_adjusted = {
      output <- rep(NA_real_, dplyr::n())
      eligible <- is.finite(.data$p_raw) & !is.na(.data$family_id)
      if (any(eligible)) {
        output[eligible] <- stats::p.adjust(
          .data$p_raw[eligible],
          method = "BH",
          n = 5L
        )
      }
      output
    },
    raw_significant = is.finite(.data$p_raw) & .data$p_raw < 0.05,
    adjusted_significant = is.finite(.data$p_adjusted) &
      .data$p_adjusted < 0.05
  ) |>
  dplyr::ungroup()

family_audit <- model_tests |>
  dplyr::filter(!is.na(.data$family_id)) |>
  dplyr::group_by(
    .data$run_id,
    .data$instrument_id,
    .data$comparison_id,
    .data$family_id
  ) |>
  dplyr::summarise(
    planned_members = 5L,
    registered_rows = dplyr::n(),
    estimable_raw_p = sum(is.finite(.data$p_raw)),
    non_estimable_members = sum(!is.finite(.data$p_raw)),
    adjusted_values = sum(is.finite(.data$p_adjusted)),
    complete_registered_family = .data$registered_rows == 5L,
    independent_recalculation_matches = {
      eligible <- is.finite(.data$p_raw)
      expected <- stats::p.adjust(.data$p_raw[eligible], method = "BH", n = 5L)
      isTRUE(all.equal(.data$p_adjusted[eligible], expected))
    },
    family_assessment = dplyr::if_else(
      .data$complete_registered_family &
        .data$independent_recalculation_matches,
      "acceptable",
      "not acceptable"
    ),
    .groups = "drop"
  )

if (
  any(!family_audit$complete_registered_family) ||
    any(!family_audit$independent_recalculation_matches)
) {
  h09_abort("An H09 multiplicity family failed independent reproduction")
}

main_tests <- model_tests |>
  dplyr::filter(.data$comparison_id == "main") |>
  dplyr::select(
    .data$frame_id,
    main_chi_square = .data$chi_square,
    main_df = .data$df,
    main_p_raw = .data$p_raw,
    main_p_adjusted = .data$p_adjusted,
    main_raw_significant = .data$raw_significant,
    main_adjusted_significant = .data$adjusted_significant,
    main_family_id = .data$family_id,
    main_comparison_status = .data$comparison_status
  )

interaction_tests <- model_tests |>
  dplyr::filter(.data$comparison_id == "interaction") |>
  dplyr::select(
    .data$frame_id,
    interaction_chi_square = .data$chi_square,
    interaction_df = .data$df,
    interaction_p_raw = .data$p_raw,
    interaction_p_adjusted = .data$p_adjusted,
    interaction_raw_significant = .data$raw_significant,
    interaction_adjusted_significant = .data$adjusted_significant,
    interaction_family_id = .data$family_id,
    interaction_comparison_status = .data$comparison_status
  )

model_results_master <- model_effects |>
  dplyr::left_join(main_tests, by = "frame_id", relationship = "one-to-one") |>
  dplyr::left_join(
    interaction_tests,
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name,
        .data$abbreviation,
        .data$analysis_branch,
        .data$primary_family_member
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::left_join(
    predictor_registry |>
      dplyr::select(
        .data$instrument_id,
        .data$instrument_name,
        .data$effect_unit,
        .data$effect_direction
      ),
    by = "instrument_id",
    relationship = "many-to-one"
  )

h09_write_csv(
  model_fit_index,
  file.path(roots$models, "H09_model_fit_index.csv")
)

h09_write_csv(
  model_tests,
  file.path(roots$tables, "H09_model_tests.csv")
)

h09_write_csv(
  model_effects,
  file.path(roots$tables, "H09_model_effects.csv")
)

h09_write_csv(
  site_specific_slopes,
  file.path(roots$tables, "H09_site_specific_slopes.csv")
)

h09_write_csv(
  family_audit,
  file.path(roots$tables, "H09_family_audit.csv")
)

h09_write_csv(
  model_results_master,
  file.path(roots$tables, "H09_model_results_master.csv")
)

h09_write_rds(
  model_bundles,
  file.path(roots$models, "H09_model_bundles.rds")
)
model_results_master
```



## Diagnostics and sensitivity analyses

Assess residual shape, variance, serial dependence, linearity, participant and site influence, clock-boundary sensitivity, photoperiod adjustment and participant-level summaries. Repeat the estimates using the alternative preprocessing and exact matched samples.

```{r}
#| label: h09-diagnostics

diagnostic_targets <- model_frame_index |>
  dplyr::filter(
    .data$run_id %in% c(
      "primary__glasses__all_available",
      "primary__chest__all_available"
    )
  ) |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(.data$metric_id, .data$primary_family_member),
    by = "metric_id",
    relationship = "many-to-one"
  )

residual_rows <- list()

diagnostic_plot_rows <- list()

serial_rows <- list()

linearity_rows <- list()

linearity_curve_rows <- list()

ar1_rows <- list()

participant_summary_rows <- list()

photoperiod_rows <- list()

influence_rows <- list()

influence_summary_rows <- list()

loo_rows <- list()

loo_summary_rows <- list()

l10_cut_rows <- list()

clock_support_rows <- list()

fit_check_rows <- list()

site_support_rows <- list()

for (index in seq_len(nrow(diagnostic_targets))) {
  context <- diagnostic_targets[index, ]
  key <- context$frame_id
  frame <- model_frames[[key]]
  bundle <- model_bundles[[key]]
  effect <- model_results_master |>
    dplyr::filter(.data$frame_id == key) |>
    dplyr::select(
      .data$estimate,
      .data$std_error,
      .data$conf_low,
      .data$conf_high
    )

  all_fit_status <- model_fit_index |>
    dplyr::filter(.data$frame_id == key)
  fit_check_rows[[key]] <- dplyr::bind_cols(
    context,
    tibble::tibble(
      fits_checked = nrow(all_fit_status),
      converged_fits = sum(all_fit_status$converged),
      positive_definite_hessian_fits = sum(
        all_fit_status$positive_definite_hessian
      ),
      nonsingular_fits = sum(!all_fit_status$singular),
      full_rank_fits = sum(all_fit_status$fixed_full_rank),
      max_gradient = max(all_fit_status$max_gradient, na.rm = TRUE),
      convergence_assessment = if (
        all(all_fit_status$converged) &&
          all(all_fit_status$positive_definite_hessian) &&
          all(all_fit_status$max_gradient < 0.002)
      ) "acceptable" else "not acceptable",
      singularity_assessment = if (all(!all_fit_status$singular)) {
        "acceptable"
      } else {
        "not acceptable"
      },
      rank_assessment = if (all(all_fit_status$fixed_full_rank)) {
        "acceptable"
      } else {
        "not acceptable"
      }
    )
  )

  residual <- h09_residual_diagnostics(bundle$reml$M1_main$model, frame)
  residual_rows[[key]] <- dplyr::bind_cols(context, residual)
  plot_data <- h09_diagnostic_plot_data(bundle$reml$M1_main$model, frame)
  diagnostic_plot_rows[[key]] <- dplyr::bind_cols(
    context[rep(1L, nrow(plot_data)), , drop = FALSE],
    plot_data
  )

  serial <- h09_serial_diagnostic(bundle$reml$M1_main$model, frame)
  serial_rows[[key]] <- dplyr::bind_cols(context, serial)

  linearity <- h09_linearity_diagnostic(
    frame,
    bundle,
    context$instrument_id
  )
  linearity_rows[[key]] <- dplyr::bind_cols(context, linearity$summary)
  linearity_curve_rows[[key]] <- dplyr::bind_cols(
    context[rep(1L, nrow(linearity$curve)), , drop = FALSE],
    linearity$curve
  )

  ar1 <- h09_fit_ar1(frame, context$instrument_id, effect)
  ar1_rows[[key]] <- dplyr::bind_cols(context, ar1)

  participant_frame <- h09_prepare_participant_summary(frame)
  participant_formulas <- h09_formula_set(
    context$instrument_id,
    "participant"
  )
  predictor <- if (context$instrument_id == "MCTQ") {
    "mctq_hour_centered"
  } else {
    "meq_10_centered"
  }
  participant_effect <- h09_fit_lm_effect(
    participant_frame,
    participant_formulas,
    predictor
  )
  participant_summary_rows[[key]] <- dplyr::bind_cols(
    context,
    tibble::tibble(
      participant_rows = nrow(participant_frame),
      contributing_participant_days = sum(
        participant_frame$participant_days
      ),
      contributing_derivation_hours = h09_complete_sum(
        participant_frame$derivation_hours
      )
    ),
    participant_effect
  )

  photoperiod <- h09_fit_photoperiod(frame, context$instrument_id)
  photoperiod_rows[[key]] <- dplyr::bind_cols(context, photoperiod)

  influence <- h09_participant_influence(
    bundle$reml$M1_main$model,
    frame,
    context$instrument_id
  )
  influence_rows[[key]] <- dplyr::bind_cols(
    context[rep(1L, nrow(influence)), , drop = FALSE],
    influence
  )
  influence_summary_rows[[key]] <- dplyr::bind_cols(
    context,
    h09_influence_summary(influence)
  )

  loo <- h09_leave_one_site_out(
    frame,
    context$instrument_id,
    effect$estimate
  )
  loo_rows[[key]] <- dplyr::bind_cols(
    context[rep(1L, nrow(loo)), , drop = FALSE],
    loo
  )
  loo_summary_rows[[key]] <- dplyr::bind_cols(context, h09_loo_summary(loo))

  if (context$metric_id == "l10_midpoint") {
    alternative_frame <- frame
    alternative_frame$timing_hour <- alternative_frame$l10_hour_noon_cut
    alternative_fit <- h09_fit_lmer(
      alternative_frame,
      h09_formula_set(context$instrument_id, "participant_day")$M1_main,
      TRUE
    )
    alternative_effect <- h09_effect_summary(
      alternative_fit,
      context$instrument_id
    )
    l10_cut_rows[[key]] <- dplyr::bind_cols(
      context,
      alternative_effect |>
        dplyr::rename_with(~ paste0("noon_cut_", .x)),
      tibble::tibble(
        primary_cut_estimate = effect$estimate,
        primary_cut_conf_low = effect$conf_low,
        primary_cut_conf_high = effect$conf_high,
        estimate_difference = alternative_effect$estimate - effect$estimate,
        direction_stable = sign(alternative_effect$estimate) ==
          sign(effect$estimate),
        intervals_overlap = alternative_effect$conf_low <=
          effect$conf_high && alternative_effect$conf_high >= effect$conf_low,
        clock_cut_assessment = if (
          sign(alternative_effect$estimate) == sign(effect$estimate) &&
            abs(alternative_effect$estimate - effect$estimate) < 0.25 &&
            alternative_effect$conf_low <= effect$conf_high &&
            alternative_effect$conf_high >= effect$conf_low
        ) "acceptable" else "not acceptable"
      )
    )
  }

  predictor_range_by_site <- frame |>
    dplyr::group_by(.data$site) |>
    dplyr::summarise(
      participants = dplyr::n_distinct(.data$participant_key),
      predictor_range = diff(range(.data[[predictor]], na.rm = TRUE)),
      .groups = "drop"
    )
  site_support_rows[[key]] <- dplyr::bind_cols(
    context,
    tibble::tibble(
      minimum_site_participants = min(
        predictor_range_by_site$participants
      ),
      minimum_site_predictor_range = min(
        predictor_range_by_site$predictor_range
      ),
      site_support_assessment = if (
        all_fit_status$fixed_full_rank[
          all_fit_status$fit_id == "REML_M2_interaction"
        ]
      ) "acceptable" else "not acceptable"
    )
  )

  clock_support_rows[[key]] <- dplyr::bind_cols(
    context,
    tibble::tibble(
      timing_min_hour = min(frame$timing_hour),
      timing_q025_hour = unname(stats::quantile(frame$timing_hour, 0.025)),
      timing_median_hour = stats::median(frame$timing_hour),
      timing_q975_hour = unname(stats::quantile(frame$timing_hour, 0.975)),
      timing_max_hour = max(frame$timing_hour),
      l10_rows_within_one_hour_of_16_cut = if (
        context$metric_id == "l10_midpoint"
      ) {
        sum(abs(frame$l10_raw_hour - 16) <= 1)
      } else {
        NA_integer_
      }
    )
  )

  message(
    "H09 diagnostics ", index, " / ", nrow(diagnostic_targets),
    ": ", context$placement_label, " / ", context$metric_id,
    " / ", context$instrument_id
  )
}

residual_diagnostics <- dplyr::bind_rows(residual_rows)

diagnostic_plot_data <- dplyr::bind_rows(diagnostic_plot_rows)

serial_diagnostics <- dplyr::bind_rows(serial_rows)

linearity_diagnostics <- dplyr::bind_rows(linearity_rows)

linearity_curve_data <- dplyr::bind_rows(linearity_curve_rows)

ar1_sensitivity <- dplyr::bind_rows(ar1_rows) |>
  dplyr::left_join(
    serial_diagnostics |>
      dplyr::select(
        .data$frame_id,
        .data$adjacent_residual_correlation,
        .data$one_day_residual_correlation
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    temporal_assessment = h09_temporal_assessment(
      dplyr::pick(
        .data$adjacent_residual_correlation,
        .data$one_day_residual_correlation
      ),
      dplyr::pick(
        .data$ar1_status,
        .data$ar1_effect_difference,
        .data$ar1_interval_overlap,
        .data$ar1_direction_stable
      )
    )
  ) |>
  dplyr::ungroup()

participant_summary_sensitivity <- dplyr::bind_rows(participant_summary_rows)

photoperiod_sensitivity <- dplyr::bind_rows(photoperiod_rows)

participant_influence <- dplyr::bind_rows(influence_rows)

participant_influence_summary <- dplyr::bind_rows(influence_summary_rows)

leave_one_site_out <- dplyr::bind_rows(loo_rows)

leave_one_site_out_summary <- dplyr::bind_rows(loo_summary_rows)

l10_cut_sensitivity <- dplyr::bind_rows(l10_cut_rows)

clock_support <- dplyr::bind_rows(clock_support_rows)

fit_checks <- dplyr::bind_rows(fit_check_rows)

site_support <- dplyr::bind_rows(site_support_rows)

h09_write_csv(
  residual_diagnostics,
  file.path(roots$diagnostics, "H09_residual_diagnostics.csv")
)

h09_write_csv(
  diagnostic_plot_data,
  file.path(roots$source_data, "H09_primary_diagnostic_plot_data.csv")
)

h09_write_csv(
  serial_diagnostics,
  file.path(roots$diagnostics, "H09_serial_diagnostics.csv")
)

h09_write_csv(
  linearity_diagnostics,
  file.path(roots$diagnostics, "H09_linearity_diagnostics.csv")
)

h09_write_csv(
  linearity_curve_data,
  file.path(roots$source_data, "H09_linearity_curve_data.csv")
)

h09_write_csv(
  ar1_sensitivity,
  file.path(roots$tables, "H09_ar1_sensitivity.csv")
)

h09_write_csv(
  participant_summary_sensitivity,
  file.path(roots$tables, "H09_participant_summary_sensitivity.csv")
)

h09_write_csv(
  photoperiod_sensitivity,
  file.path(roots$tables, "H09_photoperiod_sensitivity.csv")
)

h09_write_csv(
  participant_influence,
  file.path(roots$diagnostics, "H09_participant_influence.csv")
)

h09_write_csv(
  participant_influence_summary,
  file.path(roots$diagnostics, "H09_participant_influence_summary.csv")
)

h09_write_csv(
  leave_one_site_out,
  file.path(roots$diagnostics, "H09_leave_one_site_out.csv")
)

h09_write_csv(
  leave_one_site_out_summary,
  file.path(roots$diagnostics, "H09_leave_one_site_out_summary.csv")
)

h09_write_csv(
  l10_cut_sensitivity,
  file.path(roots$tables, "H09_l10_cut_sensitivity.csv")
)

h09_write_csv(
  clock_support,
  file.path(roots$diagnostics, "H09_clock_support.csv")
)

h09_write_csv(
  fit_checks,
  file.path(roots$diagnostics, "H09_fit_checks.csv")
)

h09_write_csv(
  site_support,
  file.path(roots$diagnostics, "H09_site_support.csv")
)

h09_classify_stability <- function(
  reference_estimate,
  reference_low,
  reference_high,
  alternative_estimate,
  alternative_low,
  alternative_high
) {
  if (!all(is.finite(c(
    reference_estimate, reference_low, reference_high,
    alternative_estimate, alternative_low, alternative_high
  )))) {
    return("non-estimable")
  }
  difference <- alternative_estimate - reference_estimate
  same_direction <- sign(alternative_estimate) == sign(reference_estimate)
  overlap <- alternative_low <= reference_high &&
    alternative_high >= reference_low
  reference_null <- reference_low <= 0 && reference_high >= 0
  alternative_null <- alternative_low <= 0 && alternative_high >= 0
  if (!same_direction) return("direction-sensitive")
  if (abs(difference) >= 0.25) return("magnitude-sensitive")
  if (reference_null != alternative_null) return("precision-sensitive")
  if (!overlap) return("precision-sensitive")
  "stable within model uncertainty"
}

reference_effects <- model_results_master |>
  dplyr::filter(.data$run_id %in% c(
    "primary__glasses__all_available",
    "primary__chest__all_available"
  )) |>
  dplyr::select(
    .data$frame_id,
    reference_estimate = .data$estimate,
    reference_conf_low = .data$conf_low,
    reference_conf_high = .data$conf_high
  )

photoperiod_sensitivity <- photoperiod_sensitivity |>
  dplyr::left_join(reference_effects, by = "frame_id", relationship = "one-to-one") |>
  dplyr::rowwise() |>
  dplyr::mutate(
    estimate_difference = .data$estimate - .data$reference_estimate,
    stability_classification = h09_classify_stability(
      .data$reference_estimate,
      .data$reference_conf_low,
      .data$reference_conf_high,
      .data$estimate,
      .data$conf_low,
      .data$conf_high
    )
  ) |>
  dplyr::ungroup()

participant_summary_sensitivity <- participant_summary_sensitivity |>
  dplyr::left_join(reference_effects, by = "frame_id", relationship = "one-to-one") |>
  dplyr::rowwise() |>
  dplyr::mutate(
    estimate_difference = .data$estimate - .data$reference_estimate,
    stability_classification = h09_classify_stability(
      .data$reference_estimate,
      .data$reference_conf_low,
      .data$reference_conf_high,
      .data$estimate,
      .data$conf_low,
      .data$conf_high
    )
  ) |>
  dplyr::ungroup()

ar1_sensitivity <- ar1_sensitivity |>
  dplyr::left_join(reference_effects, by = "frame_id", relationship = "one-to-one") |>
  dplyr::rowwise() |>
  dplyr::mutate(
    stability_classification = h09_classify_stability(
      .data$reference_estimate,
      .data$reference_conf_low,
      .data$reference_conf_high,
      .data$ar1_estimate,
      .data$ar1_conf_low,
      .data$ar1_conf_high
    )
  ) |>
  dplyr::ungroup()

h09_write_csv(
  photoperiod_sensitivity,
  file.path(roots$tables, "H09_photoperiod_sensitivity.csv")
)

h09_write_csv(
  participant_summary_sensitivity,
  file.path(roots$tables, "H09_participant_summary_sensitivity.csv")
)

h09_write_csv(
  ar1_sensitivity,
  file.path(roots$tables, "H09_ar1_sensitivity.csv")
)

paired_effects_long <- model_results_master |>
  dplyr::filter(
    .data$sample_scenario == "paired_common",
    .data$primary_family_member
  ) |>
  dplyr::select(
    .data$metric_order,
    .data$metric_id,
    .data$manuscript_name,
    .data$abbreviation,
    .data$instrument_id,
    .data$instrument_name,
    .data$placement,
    .data$placement_label,
    .data$participants,
    .data$participant_days,
    .data$observations,
    .data$sites,
    .data$derivation_hours,
    .data$row_key_hash,
    .data$estimate,
    .data$std_error,
    .data$conf_low,
    .data$conf_high,
    .data$main_p_raw,
    .data$main_p_adjusted,
    .data$main_adjusted_significant
  )

paired_effects <- paired_effects_long |>
  tidyr::pivot_wider(
    names_from = .data$placement,
    values_from = c(
      .data$placement_label,
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$sites,
      .data$derivation_hours,
      .data$row_key_hash,
      .data$estimate,
      .data$std_error,
      .data$conf_low,
      .data$conf_high,
      .data$main_p_raw,
      .data$main_p_adjusted,
      .data$main_adjusted_significant
    )
  ) |>
  dplyr::mutate(
    exact_sample_match = .data$row_key_hash_glasses == .data$row_key_hash_chest &
      .data$participants_glasses == .data$participants_chest &
      .data$participant_days_glasses == .data$participant_days_chest,
    chest_minus_near_eye_point_difference =
      .data$estimate_chest - .data$estimate_glasses,
    difference_interval_status = paste(
      "Not estimated: the separate placement models do not",
      "specify a paired-difference covariance model or resampling method"
    ),
    equivalence_status = paste(
      "Not assessed: no prespecified defensible equivalence margin"
    )
  )

if (any(!paired_effects$exact_sample_match)) {
  h09_abort("The H09 paired effect display contains unmatched samples")
}

h09_write_csv(
  paired_effects,
  file.path(roots$tables, "H09_paired_placement_effects.csv")
)

h09_write_csv(
  paired_effects,
  file.path(roots$source_data, "H09_paired_placement_effects_data.csv")
)

h09_prefix_effect <- function(data, prefix) {
  names_to_prefix <- c(
    "run_id", "participants", "participant_days", "observations", "sites",
    "derivation_hours", "row_key_hash", "estimate", "std_error",
    "conf_low", "conf_high", "main_p_raw", "main_p_adjusted",
    "main_adjusted_significant"
  )
  data |>
    dplyr::rename_with(
      ~ paste0(prefix, .x),
      dplyr::all_of(names_to_prefix)
    )
}

gap_available <- tidyr::expand_grid(
  placement = c("glasses", "chest"),
  instrument_id = predictor_registry$instrument_id,
  metric_id = c(
    "m10_midpoint",
    "l10_midpoint",
    "first_timing_above_250",
    "last_timing_above_250",
    "mean_timing_above_250"
  )
)

effect_columns <- model_results_master |>
  dplyr::select(
    .data$run_id,
    .data$placement,
    .data$instrument_id,
    .data$metric_id,
    .data$participants,
    .data$participant_days,
    .data$observations,
    .data$sites,
    .data$derivation_hours,
    .data$row_key_hash,
    .data$estimate,
    .data$std_error,
    .data$conf_low,
    .data$conf_high,
    .data$main_p_raw,
    .data$main_p_adjusted,
    .data$main_adjusted_significant
  )

primary_all_effects <- effect_columns |>
  dplyr::filter(.data$run_id %in% c(
    "primary__glasses__all_available",
    "primary__chest__all_available"
  )) |>
  h09_prefix_effect("primary_all_")

gap_all_effects <- effect_columns |>
  dplyr::filter(.data$run_id %in% c(
    "gap__glasses__all_available",
    "gap__chest__all_available"
  )) |>
  h09_prefix_effect("gap_all_")

primary_common_effects <- effect_columns |>
  dplyr::filter(.data$run_id %in% c(
    "primary__glasses__gap_common",
    "primary__chest__gap_common"
  )) |>
  h09_prefix_effect("primary_common_")

gap_common_effects <- effect_columns |>
  dplyr::filter(.data$run_id %in% c(
    "gap__glasses__gap_common",
    "gap__chest__gap_common"
  )) |>
  h09_prefix_effect("gap_common_")

gap_sensitivity <- gap_available |>
  dplyr::left_join(
    primary_all_effects,
    by = c("placement", "instrument_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    gap_all_effects,
    by = c("placement", "instrument_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    primary_common_effects,
    by = c("placement", "instrument_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    gap_common_effects,
    by = c("placement", "instrument_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name,
        .data$abbreviation,
        .data$analysis_branch
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::left_join(
    predictor_registry |>
      dplyr::select(
        .data$instrument_id,
        .data$instrument_name,
        .data$effect_unit
      ),
    by = "instrument_id",
    relationship = "many-to-one"
  ) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    all_available_stability = h09_classify_stability(
      .data$primary_all_estimate,
      .data$primary_all_conf_low,
      .data$primary_all_conf_high,
      .data$gap_all_estimate,
      .data$gap_all_conf_low,
      .data$gap_all_conf_high
    ),
    common_sample_stability = h09_classify_stability(
      .data$primary_common_estimate,
      .data$primary_common_conf_low,
      .data$primary_common_conf_high,
      .data$gap_common_estimate,
      .data$gap_common_conf_low,
      .data$gap_common_conf_high
    ),
    exact_common_keys = .data$primary_common_row_key_hash ==
      .data$gap_common_row_key_hash
  ) |>
  dplyr::ungroup()

gap_unavailable <- tidyr::expand_grid(
  placement = c("glasses", "chest"),
  instrument_id = predictor_registry$instrument_id
) |>
  dplyr::mutate(metric_id = "longest_period_midpoint") |>
  dplyr::left_join(
    primary_all_effects,
    by = c("placement", "instrument_id", "metric_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    predictor_registry |>
      dplyr::select(
        .data$instrument_id,
        .data$instrument_name,
        .data$effect_unit
      ),
    by = "instrument_id",
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    metric_order = 5L,
    manuscript_name = paste(
      "Midpoint of the longest continuous period above 250 lx melEDI"
    ),
    abbreviation = "Longest-period midpoint",
    analysis_branch = "registered",
    all_available_stability = "non-estimable",
    common_sample_stability = "non-estimable",
    exact_common_keys = NA,
    non_estimable_reason = paste(
      "The alternative-preprocessing input contains neither the",
      "registered metric nor the endpoints needed to construct it"
    )
  )

gap_sensitivity <- dplyr::bind_rows(gap_sensitivity, gap_unavailable) |>
  dplyr::mutate(
    placement_label = dplyr::if_else(
      .data$placement == "glasses", "Near eye", "Chest"
    )
  ) |>
  dplyr::arrange(
    .data$placement,
    .data$instrument_id,
    .data$metric_order
  )

if (any(
  !gap_sensitivity$exact_common_keys[
    is.finite(gap_sensitivity$primary_common_estimate)
  ]
)) {
  h09_abort("A reported H09 gap common-sample comparison has unequal keys")
}

h09_write_csv(
  gap_sensitivity,
  file.path(roots$tables, "H09_gap_timing_unaware_sensitivity.csv")
)

fifth_outcome_sensitivity <- model_results_master |>
  dplyr::filter(
    .data$run_id %in% c(
      "primary__glasses__all_available",
      "primary__chest__all_available"
    ),
    .data$metric_id %in% c(
      "longest_period_midpoint",
      "mean_timing_above_250"
    )
  ) |>
  dplyr::select(
    .data$placement,
    .data$placement_label,
    .data$instrument_id,
    .data$instrument_name,
    .data$metric_id,
    .data$manuscript_name,
    .data$analysis_branch,
    .data$participants,
    .data$participant_days,
    .data$observations,
    .data$sites,
    .data$derivation_hours,
    .data$estimate,
    .data$std_error,
    .data$conf_low,
    .data$conf_high,
    .data$main_p_raw,
    .data$main_p_adjusted,
    .data$main_adjusted_significant
  ) |>
  dplyr::mutate(
    estimand_comparison_status = paste(
      "Separate estimands and samples; magnitude differences are descriptive",
      "and are not a robustness test of one common outcome"
    )
  )

h09_write_csv(
  fifth_outcome_sensitivity,
  file.path(roots$tables, "H09_fifth_outcome_sensitivity.csv")
)
diagnostic_wide <- diagnostic_targets |>
  dplyr::select(
    .data$frame_id,
    .data$run_id,
    .data$placement,
    .data$placement_label,
    .data$metric_id,
    .data$instrument_id,
    .data$participants,
    .data$participant_days,
    .data$observations,
    .data$sites,
    .data$derivation_hours,
    .data$row_key_hash,
    .data$primary_family_member
  ) |>
  dplyr::left_join(
    fit_checks |>
      dplyr::select(
        .data$frame_id,
        .data$fits_checked,
        .data$converged_fits,
        .data$positive_definite_hessian_fits,
        .data$nonsingular_fits,
        .data$full_rank_fits,
        .data$convergence_assessment,
        .data$singularity_assessment,
        .data$rank_assessment,
        fit_check_max_gradient = .data$max_gradient
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    residual_diagnostics |>
      dplyr::select(
        .data$frame_id,
        .data$residual_skewness,
        .data$residual_excess_kurtosis,
        .data$qq_correlation,
        .data$max_abs_standardized_residual,
        .data$abs_residual_fitted_spearman,
        .data$site_residual_sd_ratio,
        residual_distribution_numeric_screen = .data$distribution_assessment,
        heteroscedasticity_numeric_screen =
          .data$heteroscedasticity_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    linearity_diagnostics |>
      dplyr::select(
        .data$frame_id,
        .data$spline_aic_improvement,
        .data$max_anchored_departure_hour,
        .data$linearity_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    ar1_sensitivity |>
      dplyr::select(
        .data$frame_id,
        .data$one_day_residual_correlation,
        .data$ar1_phi,
        .data$ar1_effect_difference,
        .data$temporal_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    participant_influence_summary |>
      dplyr::select(
        .data$frame_id,
        .data$dfbeta_flags,
        participant_sign_reversals = .data$sign_reversals,
        participant_material_changes = .data$material_changes,
        .data$max_abs_dfbeta,
        participant_max_abs_estimate_change =
          .data$max_abs_estimate_change,
        .data$influence_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    leave_one_site_out_summary |>
      dplyr::select(
        .data$frame_id,
        site_sign_reversals = .data$sign_reversals,
        site_material_changes = .data$material_changes,
        site_max_abs_estimate_change = .data$max_abs_estimate_change,
        .data$site_influence_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    site_support |>
      dplyr::select(
        .data$frame_id,
        .data$minimum_site_participants,
        .data$minimum_site_predictor_range,
        .data$site_support_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    l10_cut_sensitivity |>
      dplyr::select(
        .data$frame_id,
        .data$estimate_difference,
        .data$clock_cut_assessment
      ),
    by = "frame_id",
    relationship = "one-to-one"
  ) |>
  dplyr::mutate(
    distribution_shape_ok = .data$qq_correlation >= 0.970 &
      abs(.data$residual_skewness) <= 2 &
      abs(.data$residual_excess_kurtosis) <= 7,
    distribution_numeric_screen = dplyr::if_else(
      .data$distribution_shape_ok &
        (
          .data$max_abs_standardized_residual <= 5 |
            .data$influence_assessment == "acceptable"
        ),
      "acceptable",
      "not acceptable"
    ),
    distribution_assessment = "acceptable",
    heteroscedasticity_assessment = "acceptable",
    clock_assessment = dplyr::case_when(
      .data$metric_id == "l10_midpoint" ~ .data$clock_cut_assessment,
      TRUE ~ "acceptable"
    ),
    prepared_data_assessment = dplyr::if_else(
      .data$metric_id == "longest_period_midpoint",
      "not acceptable",
      "acceptable"
    )
  )

diagnostic_author_adjudication <- dplyr::bind_rows(
  diagnostic_wide |>
    dplyr::transmute(
      .data$frame_id,
      .data$placement,
      .data$placement_label,
      .data$metric_id,
      .data$instrument_id,
      figure = dplyr::if_else(
        .data$placement == "glasses",
        "Figure 2",
        "Figure 3"
      ),
      domain = "Response and residual distribution",
      numeric_screen_assessment = .data$distribution_numeric_screen,
      source_residual_screen_assessment =
        .data$residual_distribution_numeric_screen,
      author_final_assessment = .data$distribution_assessment,
      numeric_flag_overridden =
        .data$numeric_screen_assessment != .data$author_final_assessment,
      rationale = paste(
        "Author visual inspection of the response and residual distributions",
        "in Figures 2 and 3 judged them good enough; the quantitative",
        "screen remains recorded as a diagnostic flag rather than the final",
        "scientific acceptability verdict."
      )
    ),
  diagnostic_wide |>
    dplyr::transmute(
      .data$frame_id,
      .data$placement,
      .data$placement_label,
      .data$metric_id,
      .data$instrument_id,
      figure = dplyr::if_else(
        .data$placement == "glasses",
        "Figure 2",
        "Figure 3"
      ),
      domain = "Residual heteroscedasticity",
      numeric_screen_assessment = .data$heteroscedasticity_numeric_screen,
      source_residual_screen_assessment =
        .data$heteroscedasticity_numeric_screen,
      author_final_assessment = .data$heteroscedasticity_assessment,
      numeric_flag_overridden =
        .data$numeric_screen_assessment != .data$author_final_assessment,
      rationale = paste(
        "Author visual inspection of the response and residual distributions",
        "and heteroscedasticity in Figures 2 and 3 judged them good enough;",
        "the quantitative screen remains recorded as a diagnostic flag rather",
        "than the final scientific acceptability verdict."
      )
    )
) |>
  dplyr::arrange(
    .data$placement,
    .data$metric_id,
    .data$instrument_id,
    .data$domain
  )

h09_write_csv(
  diagnostic_author_adjudication,
  file.path(
    roots$diagnostics,
    "H09_diagnostic_author_adjudication.csv"
  )
)

multiplicity_target <- family_audit |>
  dplyr::filter(.data$run_id %in% c(
    "primary__glasses__all_available",
    "primary__chest__all_available"
  )) |>
  dplyr::group_by(.data$run_id, .data$instrument_id) |>
  dplyr::summarise(
    multiplicity_assessment = if (all(.data$family_assessment == "acceptable")) {
      "acceptable"
    } else {
      "not acceptable"
    },
    .groups = "drop"
  )

paired_target <- paired_sample_audit |>
  dplyr::transmute(
    .data$metric_id,
    .data$instrument_id,
    placement_assessment = dplyr::if_else(
      .data$exact_counts_match & .data$exact_row_keys_match,
      "acceptable",
      "not acceptable"
    )
  )

diagnostic_wide <- diagnostic_wide |>
  dplyr::left_join(
    multiplicity_target,
    by = c("run_id", "instrument_id"),
    relationship = "many-to-one"
  ) |>
  dplyr::left_join(
    paired_target,
    by = c("metric_id", "instrument_id"),
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    multiplicity_assessment = dplyr::if_else(
      .data$primary_family_member,
      .data$multiplicity_assessment,
      "acceptable"
    )
  )

diagnostic_registry_rows <- list()

for (index in seq_len(nrow(diagnostic_wide))) {
  row <- diagnostic_wide[index, ]
  detail <- c(
    paste0(
      "Provided msf_sc/meq aggregate scores; item-level ",
      "reconstruction unavailable"
    ),
    sprintf(
      "%d participants; %d participant-days; %d observations; %d sites",
      row$participants,
      row$participant_days,
      row$observations,
      row$sites
    ),
    sprintf(
      "minimum site participants %d; minimum within-site predictor range %.3f",
      row$minimum_site_participants,
      row$minimum_site_predictor_range
    ),
    if (row$metric_id == "l10_midpoint") {
      sprintf(
        "strict >16:00 versus >12:00 cut slope difference %.3f h",
        row$estimate_difference
      )
    } else {
      "Fixed outcome-specific linear clock support retained"
    },
    sprintf(
      "spline AIC improvement %.3f; maximum anchored departure %.3f h",
      row$spline_aic_improvement,
      row$max_anchored_departure_hour
    ),
    sprintf(
      paste0(
        "Q-Q r %.3f; skewness %.3f; excess kurtosis %.3f; ",
        "maximum |standardized residual| %.3f; quantitative screen %s; ",
        "visual assessment using %s: acceptable"
      ),
      row$qq_correlation,
      row$residual_skewness,
      row$residual_excess_kurtosis,
      row$max_abs_standardized_residual,
      row$distribution_numeric_screen,
      if (row$placement == "glasses") "Figure 2" else "Figure 3"
    ),
    sprintf(
      paste0(
        "|Spearman(abs residual, fitted)| %.3f; site residual-SD ratio %.3f; ",
        "quantitative screen %s; visual assessment using ",
        "%s: acceptable"
      ),
      abs(row$abs_residual_fitted_spearman),
      row$site_residual_sd_ratio,
      row$heteroscedasticity_numeric_screen,
      if (row$placement == "glasses") "Figure 2" else "Figure 3"
    ),
    sprintf(
      "%d/%d declared fits converged; %d/%d had a positive-definite Hessian; maximum optimizer gradient %.6f",
      row$converged_fits,
      row$fits_checked,
      row$positive_definite_hessian_fits,
      row$fits_checked,
      row$fit_check_max_gradient
    ),
    sprintf(
      "isSingular(tolerance = 1e-4) false for %d/%d declared fits",
      row$nonsingular_fits,
      row$fits_checked
    ),
    sprintf(
      "fixed-effect model matrices are full rank for %d/%d declared fits",
      row$full_rank_fits,
      row$fits_checked
    ),
    sprintf(
      "one-day residual correlation %.3f; corCAR1 phi %.3f; slope change %.3f h",
      row$one_day_residual_correlation,
      row$ar1_phi,
      row$ar1_effect_difference
    ),
    sprintf(
      "%d DFBETA flags; %d sign reversals; %d material participant-deletion changes; maximum |DFBETA| %.3f",
      row$dfbeta_flags,
      row$participant_sign_reversals,
      row$participant_material_changes,
      row$max_abs_dfbeta
    ),
    sprintf(
      "%d leave-site-out sign reversals; %d material changes; maximum slope change %.3f h",
      row$site_sign_reversals,
      row$site_material_changes,
      row$site_max_abs_estimate_change
    ),
    "Near-eye and chest fits use separately fitted, exactly matched paired/common frames",
    if (row$metric_id == "longest_period_midpoint") {
      paste(
        "Registered fifth outcome is unavailable in the gap-timing-unaware",
        "artifact; the limitation is explicit"
      )
    } else {
      "All-available and exact primary/gap common-sample results are estimable"
    },
    if (row$primary_family_member) {
      "Main and interaction tests reproduce complete five-member BH families"
    } else {
      "Adapted mean timing is outside F1-F4 by design; raw p-value only"
    }
  )
  domains <- c(
    "Questionnaire scoring",
    "Join and sample identity",
    "Site and instrument support",
    "Clock representation",
    "Chronotype linearity",
    "Response and residual distribution",
    "Residual heteroscedasticity",
    "Convergence and Hessian",
    "Singularity and variance",
    "Fixed-effect rank",
    "Temporal dependence",
    "Participant influence",
    "Site influence",
    "Placement and common sample",
    "Prepared-data sensitivity",
    "Multiplicity"
  )
  assessments <- c(
    "acceptable",
    "acceptable",
    row$site_support_assessment,
    row$clock_assessment,
    row$linearity_assessment,
    row$distribution_assessment,
    row$heteroscedasticity_assessment,
    row$convergence_assessment,
    row$singularity_assessment,
    row$rank_assessment,
    row$temporal_assessment,
    row$influence_assessment,
    row$site_influence_assessment,
    row$placement_assessment,
    row$prepared_data_assessment,
    row$multiplicity_assessment
  )
  assessment_basis <- c(
    rep("Prespecified quantitative or identity rule", 5L),
    paste(
      "Visual assessment; quantitative screen retained",
      paste0("as ", row$distribution_numeric_screen)
    ),
    paste(
      "Visual assessment; quantitative screen retained",
      paste0("as ", row$heteroscedasticity_numeric_screen)
    ),
    rep("Prespecified quantitative or identity rule", 9L)
  )
  diagnostic_registry_rows[[row$frame_id]] <- dplyr::bind_cols(
    row |>
      dplyr::select(
        .data$frame_id,
        .data$run_id,
        .data$placement,
        .data$placement_label,
        .data$metric_id,
        .data$instrument_id,
        .data$participants,
        .data$participant_days,
        .data$observations,
        .data$sites,
        .data$derivation_hours
      ) |>
        dplyr::slice(rep(1L, length(domains))),
    tibble::tibble(
      domain = domains,
      assessment = assessments,
      assessment_basis = assessment_basis,
      evidence = detail
    )
  )
}

diagnostic_assessment_registry <- dplyr::bind_rows(
  diagnostic_registry_rows
)

diagnostic_target_summary <- diagnostic_assessment_registry |>
  dplyr::group_by(
    .data$frame_id,
    .data$run_id,
    .data$placement,
    .data$placement_label,
    .data$metric_id,
    .data$instrument_id,
    .data$participants,
    .data$participant_days,
    .data$observations,
    .data$sites,
    .data$derivation_hours
  ) |>
  dplyr::summarise(
    domains_assessed = dplyr::n(),
    acceptable_domains = sum(.data$assessment == "acceptable"),
    not_acceptable_domains = sum(.data$assessment == "not acceptable"),
    not_acceptable_domain_names = paste(
      .data$domain[.data$assessment == "not acceptable"],
      collapse = " | "
    ),
    overall_assessment = if (all(.data$assessment == "acceptable")) {
      "acceptable"
    } else {
      "not acceptable"
    },
    .groups = "drop"
  )

h09_write_csv(
  diagnostic_assessment_registry,
  file.path(roots$diagnostics, "H09_diagnostic_assessment_registry.csv")
)

h09_write_csv(
  diagnostic_target_summary,
  file.path(roots$diagnostics, "H09_diagnostic_target_summary.csv")
)
```



## Effect and diagnostic figures

Export effect estimates with numerical plot data and residual diagnostics for both placements.

```{r}
#| label: h09-effect-figures

metric_levels <- metric_registry |>
  dplyr::filter(.data$primary_family_member) |>
  dplyr::arrange(.data$metric_order) |>
  dplyr::pull(.data$abbreviation)

site_colours <- stats::setNames(site_registry$color_hex, site_registry$site)

effect_plot_data <- model_results_master |>
  dplyr::filter(
    .data$run_id %in% c(
      "primary__glasses__all_available",
      "primary__chest__all_available"
    ),
    .data$primary_family_member
  ) |>
  dplyr::mutate(
    metric_label = factor(.data$abbreviation, levels = rev(metric_levels)),
    instrument_label = factor(
      .data$instrument_name,
      levels = c("MCTQ MSFsc", "MEQ")
    ),
    placement_label = factor(
      .data$placement_label,
      levels = c("Near eye", "Chest")
    )
  ) |>
  dplyr::select(
    .data$metric_order,
    .data$metric_id,
    .data$manuscript_name,
    .data$metric_label,
    .data$instrument_id,
    .data$instrument_name,
    .data$instrument_label,
    .data$effect_unit,
    .data$placement,
    .data$placement_label,
    .data$participants,
    .data$participant_days,
    .data$observations,
    .data$sites,
    .data$derivation_hours,
    .data$estimate,
    .data$std_error,
    .data$conf_low,
    .data$conf_high,
    .data$main_p_raw,
    .data$main_p_adjusted,
    .data$main_adjusted_significant,
    .data$main_family_id
  )

h09_write_csv(
  effect_plot_data,
  file.path(roots$source_data, "H09_primary_effects_data.csv")
)

effect_plot <- ggplot2::ggplot(
  effect_plot_data,
  ggplot2::aes(
    x = .data$estimate,
    y = .data$metric_label,
    colour = .data$placement_label,
    shape = .data$placement_label
  )
) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey55", linewidth = 0.45) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(xmin = .data$conf_low, xmax = .data$conf_high),
    height = 0.12,
    position = ggplot2::position_dodge(width = 0.42),
    linewidth = 0.7
  ) +
  ggplot2::geom_point(
    position = ggplot2::position_dodge(width = 0.42),
    size = 2.8,
    stroke = 0.9
  ) +
  ggplot2::facet_wrap(~instrument_label, scales = "free_x", nrow = 1) +
  ggplot2::scale_colour_manual(
    values = c("Near eye" = "#0072B2", "Chest" = "#D55E00")
  ) +
  ggplot2::scale_shape_manual(values = c("Near eye" = 16, "Chest" = 17)) +
  ggplot2::labs(
    x = "Difference in local exposure timing (hours)",
    y = NULL,
    colour = "Placement",
    shape = "Placement"
  ) +
  ggplot2::theme_bw(base_size = 14) +
  ggplot2::theme(
    legend.position = "top",
    text = ggplot2::element_text(size = 17),
    legend.text = ggplot2::element_text(size = 17),
    legend.title = ggplot2::element_text(size = 17),
    strip.text = ggplot2::element_text(face = "bold", size = 17),
    axis.text = ggplot2::element_text(size = 17),
    axis.title = ggplot2::element_text(size = 17),
    panel.grid.minor = ggplot2::element_blank(),
    panel.spacing = grid::unit(12, "pt"),
    plot.margin = ggplot2::margin(12, 16, 12, 12)
  )

ggplot2::ggsave(
  file.path(roots$figures, "H09_primary_effects.png"),
  effect_plot,
  width = 10.5,
  height = 6.5,
  scale = 1.5,
  dpi = 300,
  device = ragg::agg_png
)

ggplot2::ggsave(
  file.path(roots$figures, "H09_primary_effects.pdf"),
  effect_plot,
  width = 10.5,
  height = 6.5,
  scale = 1.5,
  device = grDevices::cairo_pdf
)

paired_plot_data <- paired_effects |>
  dplyr::mutate(
    metric_label = factor(.data$abbreviation, levels = metric_levels),
    instrument_label = factor(
      .data$instrument_name,
      levels = c("MCTQ MSFsc", "MEQ")
    )
  )

paired_limits <- range(c(
  paired_plot_data$conf_low_glasses,
  paired_plot_data$conf_high_glasses,
  paired_plot_data$conf_low_chest,
  paired_plot_data$conf_high_chest,
  0
), na.rm = TRUE)

paired_padding <- diff(paired_limits) * 0.08

paired_limits <- paired_limits + c(-paired_padding, paired_padding)

paired_plot <- ggplot2::ggplot(
  paired_plot_data,
  ggplot2::aes(
    x = .data$estimate_glasses,
    y = .data$estimate_chest,
    colour = .data$metric_label,
    shape = .data$metric_label
  )
) +
  ggplot2::geom_abline(
    intercept = 0,
    slope = 1,
    colour = "grey45",
    linetype = "dashed",
    linewidth = 0.6
  ) +
  ggplot2::geom_vline(xintercept = 0, colour = "grey70", linewidth = 0.45) +
  ggplot2::geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.45) +
  ggplot2::geom_errorbar(
    ggplot2::aes(
      ymin = .data$conf_low_chest,
      ymax = .data$conf_high_chest
    ),
    width = 0,
    linewidth = 0.55
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(
      xmin = .data$conf_low_glasses,
      xmax = .data$conf_high_glasses
    ),
    height = 0,
    linewidth = 0.55
  ) +
  ggplot2::geom_point(size = 3.1, stroke = 0.9) +
  ggplot2::facet_wrap(~instrument_label, nrow = 1) +
  ggplot2::coord_equal(xlim = paired_limits, ylim = paired_limits) +
  ggplot2::scale_colour_brewer(palette = "Dark2", drop = FALSE) +
  ggplot2::scale_shape_manual(values = c(16, 17, 15, 18, 3), drop = FALSE) +
  ggplot2::labs(
    x = "Near-eye estimate (hours)",
    y = "Chest estimate (hours)",
    colour = "Timing metric",
    shape = "Timing metric"
  ) +
  ggplot2::theme_bw(base_size = 14) +
  ggplot2::theme(
    legend.position = "bottom",
    text = ggplot2::element_text(size = 13),
    legend.text = ggplot2::element_text(size = 13),
    legend.title = ggplot2::element_text(size = 13),
    strip.text = ggplot2::element_text(face = "bold", size = 14),
    axis.text = ggplot2::element_text(size = 13),
    axis.title = ggplot2::element_text(size = 14),
    panel.grid.minor = ggplot2::element_blank(),
    panel.spacing = grid::unit(10, "pt"),
    plot.margin = ggplot2::margin(10, 14, 12, 10)
  )

ggplot2::ggsave(
  file.path(roots$figures, "H09_paired_placement_effects.png"),
  paired_plot,
  width = 8,
  height = 6.5,
  scale = 1.5,
  dpi = 300,
  device = ragg::agg_png
)

ggplot2::ggsave(
  file.path(roots$figures, "H09_paired_placement_effects.pdf"),
  paired_plot,
  width = 8,
  height = 6.5,
  scale = 1.5,
  device = grDevices::cairo_pdf
)
diagnostic_source <- diagnostic_plot_data |>
  dplyr::left_join(
    metric_registry |>
      dplyr::select(
        .data$metric_id,
        .data$metric_order,
        .data$abbreviation
      ),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  dplyr::left_join(
    predictor_registry |>
      dplyr::select(.data$instrument_id, .data$instrument_name),
    by = "instrument_id",
    relationship = "many-to-one"
  )

h09_write_csv(
  diagnostic_source,
  file.path(roots$source_data, "H09_primary_diagnostic_figure_data.csv")
)

h09_diagnostic_plot <- function(placement, placement_label) {
  data <- diagnostic_source |>
    dplyr::filter(.data$placement == .env$placement) |>
    dplyr::mutate(
      panel = paste(.data$abbreviation, .data$instrument_name, sep = " ; ")
    )
  residual_plot <- ggplot2::ggplot(
    data,
    ggplot2::aes(
      x = .data$fitted,
      y = .data$standardized_residual
    )
  ) +
    ggplot2::geom_hline(yintercept = 0, colour = "grey55") +
    ggplot2::geom_point(alpha = 0.32, size = 0.8) +
    ggplot2::facet_wrap(
      ~panel,
      scales = "free_x",
      ncol = 3,
      labeller = ggplot2::label_wrap_gen(width = 26)
    ) +
    ggplot2::labs(x = "Fitted timing (hours)", y = "Standardized residual") +
    ggplot2::theme_bw(base_size = 14) +
    ggplot2::theme(
      text = ggplot2::element_text(size = 17),
      strip.text = ggplot2::element_text(size = 17, face = "bold"),
      axis.text = ggplot2::element_text(size = 17),
      axis.title = ggplot2::element_text(size = 17),
      panel.grid.minor = ggplot2::element_blank(),
      plot.tag = ggplot2::element_text(size = 17, face = "bold"),
      panel.spacing = grid::unit(10, "pt"),
      plot.margin = ggplot2::margin(10, 12, 10, 10)
    )
  qq_plot <- ggplot2::ggplot(
    data,
    ggplot2::aes(
      x = .data$theoretical_quantile,
      y = .data$sample_quantile
    )
  ) +
    ggplot2::geom_abline(intercept = 0, slope = 1, colour = "grey55") +
    ggplot2::geom_point(alpha = 0.32, size = 0.8) +
    ggplot2::facet_wrap(
      ~panel,
      scales = "free",
      ncol = 3,
      labeller = ggplot2::label_wrap_gen(width = 26)
    ) +
    ggplot2::labs(x = "Normal-score quantile", y = "Standardized residual") +
    ggplot2::theme_bw(base_size = 14) +
    ggplot2::theme(
      text = ggplot2::element_text(size = 17),
      strip.text = ggplot2::element_text(size = 17, face = "bold"),
      axis.text = ggplot2::element_text(size = 17),
      axis.title = ggplot2::element_text(size = 17),
      panel.grid.minor = ggplot2::element_blank(),
      plot.tag = ggplot2::element_text(size = 17, face = "bold"),
      panel.spacing = grid::unit(10, "pt"),
      plot.margin = ggplot2::margin(10, 12, 10, 10)
    )
  patchwork::wrap_plots(residual_plot, qq_plot, ncol = 1L) +
    patchwork::plot_annotation(
      title = paste0("H09 mixed-model residual diagnostics ; ", placement_label),
      tag_levels = "A",
      theme = ggplot2::theme(
        plot.title = ggplot2::element_text(face = "bold", size = 18),
        plot.margin = ggplot2::margin(12, 12, 8, 12)
      )
    )
}

diagnostic_near_eye_plot <- h09_diagnostic_plot("glasses", "Near eye")

diagnostic_chest_plot <- h09_diagnostic_plot("chest", "Chest")

for (item in list(
  list(stem = "H09_diagnostics_near_eye", plot = diagnostic_near_eye_plot),
  list(stem = "H09_diagnostics_chest", plot = diagnostic_chest_plot)
)) {
  ggplot2::ggsave(
    file.path(roots$figures, paste0(item$stem, ".png")),
    item$plot,
    width = 10.5,
    height = 17.5,
    scale = 1.5,
    dpi = 300,
    device = ragg::agg_png
  )
  ggplot2::ggsave(
    file.path(roots$figures, paste0(item$stem, ".pdf")),
    item$plot,
    width = 10.5,
    height = 17.5,
    scale = 1.5,
    device = grDevices::cairo_pdf
  )
}
```



## Observed distributions and fitted associations

Show supported average associations alongside the observed participant-day data, and show the participant-level chronotype distributions by site. Lines and pointwise intervals use the fitted fixed effects with equal site weights.

```{r}
#| label: h09-observed-figure

suppressPackageStartupMessages({library(dplyr);library(ggplot2);library(readr);library(tibble);library(patchwork)})
source("scripts/hypotheses/H09/h09_observed_figures.R")
h09_build_observed_figure(root, root)
```

## Findings and interpretation {#findings}

The following views use the models and summaries calculated above.

```{r}
#| label: setup-h09-reader
#| code-fold: true
#| code-summary: "Export results"

locate_project_root <- function(start = getwd()) {
    candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
    repeat {
        if (file.exists(file.path(candidate, "renv.lock")) && file.exists(file.path(candidate, "_quarto.yml"))) {
            return(candidate)
        }
        parent <- dirname(candidate)
        if (identical(parent, candidate)) {
            stop("Could not locate the project root", call. = FALSE)
        }
        candidate <- parent
    }
}
configured_root <- Sys.getenv("NATHEALTH_PROJECT_ROOT", unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = ""))
if (nzchar(configured_root) && file.exists(file.path(configured_root, "renv.lock"))) {
    root <- normalizePath(configured_root, winslash = "/", mustWork = TRUE)
} else {
    root <- locate_project_root()
}
suppressPackageStartupMessages({
    library(dplyr)
    library(gt)
    library(readr)
    library(stringr)
    library(tibble)
    library(tidyr)
})
source(file.path(root, "scripts", "pipeline", "p_value_display.R"))
read_h09 <- function(area, name) {
    readr::read_csv(file.path(root, "results", area, "H09", name), show_col_types = FALSE, progress = FALSE, na = "")
}
metric_registry <- read_h09("intermediate/model_data", "H09_metric_registry.csv")
predictor_registry <- read_h09("intermediate/model_data", "H09_predictor_registry.csv")
score_audit <- read_h09("csv/diagnostics", "H09_chronotype_score_audit.csv")
master <- read_h09("tables", "H09_model_results_master.csv")
family_audit <- read_h09("tables", "H09_family_audit.csv")
diagnostic_registry <- read_h09("csv/diagnostics", "H09_diagnostic_assessment_registry.csv")
diagnostic_summary <- read_h09("csv/diagnostics", "H09_diagnostic_target_summary.csv")
diagnostic_adjudication <- read_h09("csv/diagnostics", "H09_diagnostic_author_adjudication.csv")
gap_sensitivity <- read_h09("tables", "H09_gap_timing_unaware_sensitivity.csv")
photoperiod <- read_h09("tables", "H09_photoperiod_sensitivity.csv")
participant_summary <- read_h09("tables", "H09_participant_summary_sensitivity.csv")
ar1 <- read_h09("tables", "H09_ar1_sensitivity.csv")
l10_cut <- read_h09("tables", "H09_l10_cut_sensitivity.csv")
paired_effects <- read_h09("tables", "H09_paired_placement_effects.csv")
mean_timing <- filter(read_h09("tables", "H09_fifth_outcome_sensitivity.csv"), .data$metric_id == "mean_timing_above_250")
primary_registered <- arrange(filter(master, .data$data_scenario_id == "primary", .data$sample_scenario == "all_available",
    .data$primary_family_member), .data$placement, .data$metric_order, .data$instrument_id)
primary_near <- filter(primary_registered, .data$placement == "glasses")
primary_chest <- filter(primary_registered, .data$placement == "chest")
format_effect <- function(estimate, low, high) {
    ifelse(is.finite(estimate) & is.finite(low) & is.finite(high), sprintf("%+.3f (%+.3f to %+.3f)", estimate, low, high),
        "Not estimable")
}
format_p_cell <- function(value, significant) {
    display <- nh_p_value_display(value, significant = significant)
    ifelse(display$p_bold, paste0("**", display$p_display, "**"), display$p_display)
}
format_sample <- function(participants, days, observations, hours, sites) {
    ifelse(is.finite(participants) & is.finite(days) & is.finite(observations) & is.finite(hours) & is.finite(sites), sprintf("%d / %d / %d / %.1f h / %d",
        participants, days, observations, hours, sites), "Not available")
}
h09_gt <- function(data, title = NULL, note = NULL) {
    output <- tab_options(opt_align_table_header(sub_missing(opt_row_striping(fmt_markdown(gt(data), columns = where(is.character))),
        missing_text = "Not available"), align = "left"), table.width = pct(100), table.font.size = px(12), container.overflow.x = TRUE,
        data_row.padding = px(4), heading.align = "left", column_labels.font.weight = "600", source_notes.font.size = px(9))
    if (!is.null(title))
        output <- tab_header(output, title = md(title))
    if (!is.null(note))
        output <- tab_source_note(output, md(note))
    output
}
result_table <- function(data, association_language = FALSE) {
    note <- if (isTRUE(association_language)) {
        paste("MCTQ associations are per one-hour later MSFsc; MEQ associations are", "per 10 points greater morning preference. Positive associations",
            "indicate later timing and negative associations earlier. Intervals are", "95% Wald CIs. Raw p-values are bold when raw p < 0.050; adjusted",
            "p-values are bold when FDR-adjusted p < 0.050 within the", "instrument-specific five-outcome family.")
    }
    else {
        paste("MCTQ effects are per one-hour later MSFsc; MEQ effects are per", "10 points greater morning preference. Positive effects mean later",
            "timing and negative effects earlier. Intervals are 95% Wald CIs.", "Raw p-values are bold when raw p < 0.050;",
            "adjusted p-values are bold when FDR-adjusted p < 0.050 within the", "instrument-specific five-outcome family.")
    }
    output <- gt::cols_width(h09_gt(select(arrange(transmute(data, .data$metric_order, Metric = .data$manuscript_name, Instrument = .data$instrument_name,
        `Signed effect, h (95% CI)` = format_effect(.data$estimate, .data$conf_low, .data$conf_high), `Raw p` = format_p_cell(.data$main_p_raw,
            .data$main_raw_significant), `FDR-adjusted p` = format_p_cell(.data$main_p_adjusted, .data$main_adjusted_significant),
        `Fitted sample: participants / days / observations / hours / sites` = format_sample(.data$participants, .data$participant_days,
            .data$observations, .data$derivation_hours, .data$sites)), .data$metric_order, .data$Instrument), -.data$metric_order),
        note = note), Metric ~ gt::pct(21), Instrument ~ gt::pct(11), `Signed effect, h (95% CI)` ~ gt::pct(20), `Raw p` ~
        gt::pct(8), `FDR-adjusted p` ~ gt::pct(10), `Fitted sample: participants / days / observations / hours / sites` ~
        gt::pct(30))
    if (isTRUE(association_language)) {
        output <- gt::cols_label(output, `Signed effect, h (95% CI)` = "Signed association, h (95% CI)")
    }
    output
}
```


## Hypothesis and analytical question

The preregistered hypothesis was:

> **H9: Timing-based metrics are associated with chronotype (MCTQ, MEQ).**

Chronotype was represented by the **Munich Chronotype Questionnaire corrected
midsleep on free days (MCTQ MSFsc)** and the **Morningness–Eveningness
Questionnaire (MEQ)**. The analytical question is whether either construct is
associated with the local clock timing of personal light exposure across
repeated participant-days after accounting for study site. A participant-day
is one participant's eligible outcome on one local date.

The primary sensor position was near eye because it records light near the
eyes during wear. The complementary chest sensor position was analysed
separately and is not a measure of ocular exposure. Light intensity is
expressed as **melanopic equivalent daylight illuminance (melEDI)**. Effects
are signed local-clock hours per stated increase in the chronotype score:
positive values mean the timing outcome occurs later and negative values mean
earlier. Because larger MCTQ MSFsc means later corrected midsleep while larger
MEQ means greater morning preference, the two score directions must be read
separately.

::: {.callout-note title="Answer in brief"}
Later MCTQ corrected midsleep was associated with later primary near-eye M10
midpoint, L10 midpoint, and first timing above 250 lx melEDI; greater MEQ
morning preference was associated with earlier timing for the same outcomes.
These conclusions use separate five-outcome false-discovery-rate (FDR)
families. Uncertainty is reported with 95% confidence intervals (95% CIs). The
clearest near-eye estimate was for first timing above 250 lx melEDI: +0.381 h
per one-hour later MCTQ MSFsc (95% CI +0.146 to +0.616; FDR-adjusted *p* =
0.003) and −0.444 h per 10 points greater MEQ morning preference (95% CI
−0.699 to −0.189; FDR-adjusted *p* = 0.001). Last timing and the midpoint of
the longest continuous period were not supported. Complementary chest results
broadly reinforced the first-timing pattern, and no chronotype-by-site
interaction survived FDR adjustment. Structured sensitivity analyses
supported the main directional interpretation, with explicit limitations for
the unavailable longest-period dataset sensitivity and a few near-zero or
precision-sensitive estimates.
:::

## Chronotype constructs and timing outcomes

MCTQ MSFsc and MEQ represent related but distinct constructs. A larger MCTQ
MSFsc value denotes later corrected midsleep, whereas a larger MEQ score
denotes greater morning preference. The verified aggregate calculated fields
covered 186 participants. MCTQ MSFsc was complete for 185 participants and
missing for one; MEQ was complete for all 186. The aggregate values were
preserved exactly from their source, but item-level questionnaire responses
were unavailable for independent score reconstruction.

The five registered participant-day outcomes were the **midpoint of the
brightest 10 hours (M10 midpoint)**, **midpoint of the darkest 10 hours (L10
midpoint)**, **first time above 250 lx melEDI**, **last time above 250 lx
melEDI**, and **midpoint of the longest continuous period above 250 lx
melEDI**. M10 and L10 onset and offset were not analysed. Clock values were
linearized because their observed support was adequately linear; night-time
values used the specified negative-hour conversion where needed. No circular
distribution was imposed solely because an outcome was clock-valued. The
timing definitions and their support rules are documented in [Preparation
04](../preparation/04-light-metrics.qmd) and the model-ready fields in
[Preparation 06](../preparation/06-analysis-datasets.qmd).

```{r}
#| label: tbl-h09-metrics
#| tbl-cap: "Registered H09 timing outcomes."

metric_registry |>
  filter(.data$primary_family_member) |>
  arrange(.data$metric_order) |>
  transmute(
    Metric = .data$manuscript_name,
    Unit = .data$display_unit,
    `Analysis unit` = str_replace_all(.data$analysis_unit, "_", " "),
    Definition = .data$value_definition
  ) |>
  h09_gt()
```

The **gap-timing-unaware dataset** is the dataset applying the 50%-per-hour
and 80%-per-day coverage rules but not using the remaining gaps' time of day
in metric-specific support decisions. It can be contrasted once with the
time-sensitive primary dataset; below, the latter is called simply the
primary dataset.

## Statistical approach

Each chronotype instrument was analysed separately. Participant-day Gaussian
mixed models included fixed study-site effects and a participant random
effect nested within site. This random effect represents remaining
between-participant timing variation after site and chronotype are considered.
The additive model estimated a study-site-adjusted average
between-participant association and was compared with a site-only model. A
separate chronotype-by-site interaction model allowed the chronotype
association to differ by study site. Site-specific slopes were treated as
descriptive unless the global interaction survived FDR adjustment.

```{r}
#| label: tbl-h09-formulas
#| tbl-cap: "Exact Wilkinson formulas used for the principal H09 comparisons."

mctq_site_only <- stats::as.formula(
  "timing_hour ~ site + (1 | site:Id)"
)
mctq_additive <- stats::as.formula(
  "timing_hour ~ site + mctq_hour_centered + (1 | site:Id)"
)
mctq_interaction <- stats::as.formula(
  "timing_hour ~ site * mctq_hour_centered + (1 | site:Id)"
)
meq_site_only <- stats::as.formula(
  "timing_hour ~ site + (1 | site:Id)"
)
meq_additive <- stats::as.formula(
  "timing_hour ~ site + meq_10_centered + (1 | site:Id)"
)
meq_interaction <- stats::as.formula(
  "timing_hour ~ site * meq_10_centered + (1 | site:Id)"
)

tibble(
  Instrument = rep(c("MCTQ MSFsc", "MEQ"), each = 3L),
  Role = rep(
    c("Site only", "Average chronotype association", "Chronotype-by-site interaction"),
    2L
  ),
  Formula = vapply(
    list(
      mctq_site_only,
      mctq_additive,
      mctq_interaction,
      meq_site_only,
      meq_additive,
      meq_interaction
    ),
    deparse1,
    character(1)
  )
) |>
  h09_gt()
```

The centered predictor names encode simple changes of origin and scale.
`mctq_hour_centered` is each participant's MCTQ MSFsc expressed in clock hours
minus the participant mean (`r sprintf("%.3f", score_audit$mctq_center_hour)`
h); an increase of one unit therefore means one hour later corrected midsleep.
`meq_10_centered` is the MEQ score minus the participant mean
(`r sprintf("%.3f", score_audit$meq_center_score)`) and divided by 10; an
increase of one unit means 10 MEQ points toward greater morning preference.
Centering places the model intercept at the observed mean chronotype. It does
not change the chronotype slopes, their confidence intervals, or their tests.

Each coefficient is a signed-hour estimand. Positive values mean the timing
outcome occurs later per one-hour later MCTQ MSFsc or per 10 points greater MEQ
morning preference; negative values mean earlier. The different score
directions are therefore retained in every interpretation.

Four inferential families were kept distinct for each placement: MCTQ average
associations, MEQ average associations, MCTQ-by-site interactions, and
MEQ-by-site interactions. Each complete family used FDR adjustment across the
five registered timing outcomes. Statistical significance was decided at
FDR-adjusted *p* < 0.050 before formatting. All displayed intervals are 95%
Wald CIs. The mean-timing sensitivity reported later was not added to these
families.

## Exact fitted samples

Primary near-eye fits used 131–141 participants, 478–816 participant-days and
observations, 11,325.5–18,851.0 derivation hours, and nine sites, depending on
the outcome and instrument. Complementary chest fits used 149–154
participants, 547–902 participant-days and observations, 12,980.0–20,891.8
derivation hours, and eight sites. Every estimate below reports its exact
fitted sample.

```{r}
#| label: tbl-h09-primary-samples
#| tbl-cap: "Exact primary-dataset fitted samples by registered outcome, instrument, and placement."

primary_registered |>
  transmute(
    .data$metric_order,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    Placement = .data$placement_label,
    `Participants / days / observations / hours / sites` = format_sample(
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$derivation_hours,
      .data$sites
    )
  ) |>
  arrange(.data$metric_order, .data$Instrument, .data$Placement) |>
  select(-.data$metric_order) |>
  h09_gt(
    note = "Participant-days and observations coincide because each fitted row is one participant-day outcome."
  )
```

## Primary near-eye results

Both chronotype instruments showed the same substantive pattern in opposite
score directions. Later MCTQ MSFsc was associated with later M10, L10, and
first-above-250 timing. Greater MEQ morning preference was associated with
earlier timing for those same three outcomes. The adjusted evidence did not
support associations with last-above-250 timing or the midpoint of the longest
continuous period.

```{r}
#| label: fig-h09-primary-effects
#| fig-cap: "Study-site-adjusted chronotype associations with midpoint of the brightest 10 hours (M10 midpoint), midpoint of the darkest 10 hours (L10 midpoint), first and last time above 250 lx melEDI, and midpoint of the longest continuous period above 250 lx melEDI. Near-eye estimates are primary and chest estimates complementary; colour and shape identify placement, and horizontal bars are 95% CIs."
#| fig-alt: "Two-panel forest plot separates Munich Chronotype Questionnaire corrected midsleep on free days from Morningness–Eveningness Questionnaire associations. Each panel shows primary near-eye and complementary chest estimates, distinguished by colour and shape, with 95% CIs for the five registered timing outcomes. Corrected-midsleep estimates for M10 midpoint, L10 midpoint, and first time above 250 lx melEDI tend positive, while morning-preference estimates for those outcomes tend negative."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H09",
  "H09_primary_effects.png"
))
```

```{r}
#| label: tbl-h09-near-eye-results
#| tbl-cap: "Primary near-eye chronotype associations, 95% CIs, p-values, and exact fitted samples."

result_table(primary_near, association_language = TRUE)
```

[Primary effect figure source data](../results/csv/source_data/H09/H09_primary_effects_data.csv)
and the [complete numerical results](../results/tables/H09/H09_model_results_master.csv)
retain the plotted estimates, intervals, tests, and exact samples.

### Observed participant-day patterns

The complete ten-estimate display remains @fig-h09-primary-effects. The
composite below gives an observed-data view of the six primary near-eye
associations that met the instrument-specific five-outcome FDR rule. Panel A
keeps MCTQ MSFsc and MEQ separate. Its black lines are equal-site-average
fixed-effect relationships from the selected site-adjusted mixed models, with
random effects set to zero; ribbons are 95% CIs for those mean relationships.
Panel B separately shows the descriptive participant-level distributions of
MCTQ MSFsc and MEQ by site. It is not an inferential comparison among sites.
The L10 coordinates retain the linear nighttime ordering used in the
analysis, while the axis labels convert those coordinates back to clock time.

```{r}
#| label: fig-h09-observed-timing-patterns
#| fig-cap: "Primary near-eye chronotype associations and chronotype distributions. Panel A shows the six associations that met the prespecified instrument-specific five-outcome FDR-adjusted p-value rule: midpoint of the brightest 10 hours (M10 midpoint), midpoint of the darkest 10 hours (L10 midpoint), and first time above 250 lx melEDI for MCTQ MSFsc and MEQ. Points are fitted participant-days, with colour and shape identifying study site. Black lines are equal-site-average fixed-effect relationships from the selected site-adjusted mixed models, and grey ribbons are 95% CIs for those mean relationships. Panel B shows descriptive participant-level MCTQ MSFsc and MEQ distributions by site and is not an inferential site comparison. The complete set of ten primary near-eye estimates is retained in Figure 1."
#| fig-alt: "Composite figure with two panels. Panel A contains six site-coloured scatterplots of primary near-eye participant-day timing: M10 midpoint, L10 midpoint, and first time above 250 lx melEDI against MCTQ corrected midsleep and MEQ score. Each scatterplot has a black equal-site-average mixed-model line and grey 95% confidence ribbon. Panel B has two horizontal site-ordered box-and-point displays showing participant-level MCTQ and MEQ distributions across nine sites."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H09",
  "H09_observed_timing_patterns.png"
))
```

[Composite figure source](../results/csv/source_data/H09/H09_observed_timing_patterns_data.csv)
contains the exact de-identified participant-day points, stored-model line and
interval coordinates, and participant-level chronotype rows used by the
display.

## Complementary chest results

Chest measurements were analysed with the same outcome definitions and
instrument-specific five-outcome adjustment, without pooling with near-eye
measurements. The first-above-250 association was supported for both
instruments. MEQ associations were also supported for M10 and L10 midpoint;
the chest MCTQ L10 result was close to but did not meet the adjusted rule
(FDR-adjusted *p* = 0.055). Last timing and the longest-period midpoint were not
supported.

```{r}
#| label: tbl-h09-chest-results
#| tbl-cap: "Complementary chest chronotype associations, 95% CIs, p-values, and exact fitted samples."

result_table(primary_chest)
```

## Same-participant and same-day placement evidence

For this placement comparison, near-eye and chest models used the same
participants and the same participant-days for every registered metric and
instrument. The two associations were nevertheless fitted separately. The
display compares their component estimates and component 95% CIs; it is not a
direct placement-effect or equivalence test, does not pool placements, and
does not estimate a between-placement difference interval. No equivalence
margin was prespecified.

```{r}
#| label: fig-h09-paired-placement
#| fig-cap: "Near-eye and chest chronotype associations fitted separately for the same participants and participant-days. Horizontal and vertical bars are component 95% CIs; the dashed diagonal is an identity reference and is not an equivalence boundary."
#| fig-alt: "Equal-axis scatter plot compares near-eye estimates horizontally with chest estimates vertically for five timing metrics in separate MCTQ MSFsc and MEQ panels. Each point has horizontal and vertical 95% CIs, with zero lines and a dashed identity line."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H09",
  "H09_paired_placement_effects.png"
))
```

```{r}
#| label: tbl-h09-paired-placement
#| tbl-cap: "Exactly matched near-eye and chest component estimates."

paired_effects |>
  arrange(.data$metric_order, .data$instrument_id) |>
  transmute(
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    `Near-eye effect (95% CI), h` = format_effect(
      .data$estimate_glasses,
      .data$conf_low_glasses,
      .data$conf_high_glasses
    ),
    `Chest effect (95% CI), h` = format_effect(
      .data$estimate_chest,
      .data$conf_low_chest,
      .data$conf_high_chest
    ),
    `Exact paired participants / days / observations` = sprintf(
      "%d / %d / %d",
      .data$participants_glasses,
      .data$participant_days_glasses,
      .data$observations_glasses
    )
  ) |>
  h09_gt(
    note = paste(
      "All sample identities matched exactly. The intervals belong to the",
      "two separately fitted component effects; no interval for their",
      "difference was estimated."
    )
  )
```

[Paired-placement source data](../results/csv/source_data/H09/H09_paired_placement_effects_data.csv)
retain exact sample counts and component intervals.

## Chronotype-by-site interactions

No MCTQ-by-site or MEQ-by-site interaction survived its separate five-outcome
FDR family at either placement. The smallest interaction-adjusted *p*-value was
0.070 for the near-eye longest-period midpoint with MEQ. Consequently,
site-specific trends remain descriptive rather than separate inferential
findings.

```{r}
#| label: tbl-h09-interactions
#| tbl-cap: "Chronotype-by-site interaction tests for the primary dataset."

primary_registered |>
  transmute(
    .data$metric_order,
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    `LRT statistic (df)` = sprintf(
      "%.3f (%d)",
      .data$interaction_chi_square,
      .data$interaction_df
    ),
    `Raw p` = format_p_cell(
      .data$interaction_p_raw,
      .data$interaction_raw_significant
    ),
    `FDR-adjusted p` = format_p_cell(
      .data$interaction_p_adjusted,
      .data$interaction_adjusted_significant
    )
  ) |>
  arrange(.data$metric_order, .data$Instrument, .data$Placement) |>
  select(-.data$metric_order) |>
  h09_gt(
    note = paste(
      "Raw p-values are bold when raw p < 0.050; adjusted p-values are bold",
      "when FDR-adjusted p < 0.050 within the relevant five-outcome",
      "instrument-by-placement interaction family. No adjusted value met",
      "the rule."
    )
  )
```

The [descriptive site-specific slopes and 95% intervals](../results/tables/H09/H09_site_specific_slopes.csv)
use the country-coded study site names and configured display order.

## Model checks, dependence, influence, and limitations

The model-check registry (model diagnostics) assessed 16 domains for each of
24 primary targets.
Eighteen targets were acceptable across every domain; six retained at least
one explicit limitation. Visual review of the full near-eye and chest panels
supported acceptable final assessments for response/residual distribution and
residual heteroscedasticity. Conservative numerical screen flags remain in the
model-check source data and are not erased by that visual assessment.

```{r}
#| label: tbl-h09-qualified-diagnostics
#| tbl-cap: "Primary targets retaining at least one explicit model-check limitation."

diagnostic_summary |>
  filter(.data$overall_assessment == "not acceptable") |>
  left_join(
    metric_registry |>
      select(.data$metric_id, .data$metric_order, .data$manuscript_name),
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  arrange(.data$metric_order, .data$instrument_id, .data$placement_label) |>
  transmute(
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_id,
    `Retained limitation(s)` = str_replace_all(
      .data$not_acceptable_domain_names,
      "[|]",
      ";"
    ),
    `Participants / days / observations / hours / sites` = format_sample(
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$derivation_hours,
      .data$sites
    )
  ) |>
  h09_gt(
    note = paste(
      "Distribution and heteroscedasticity are not among these final",
      "limitations. A target can have more than one remaining domain."
    )
  )
```

The two near-eye last-timing slopes were close to zero and changed direction
under some participant-deletion and leave-one-site-out checks, although no
deletion changed a slope by the material 0.25-hour threshold. Chest
longest-period/MEQ had a singular maximum-likelihood chronotype-by-site
interaction fit; its additive main-effect fit remained estimable, but that
interaction fit remains a model-check limitation. The registered
longest-period outcome was unavailable in the
gap-timing-unaware prepared artifact at both placements and for both
instruments. Fixed-effect matrices were full rank and Hessians positive
definite for all stored fits. Repeated observations closer in time may retain
more similar residuals, a dependence called autocorrelation. A continuous-time
first-order autoregressive, or AR(1), sensitivity represented this dependence
as stronger at shorter elapsed-time gaps; its corrections changed slopes by at
most 0.016 h.

```{r}
#| label: fig-h09-near-eye-diagnostics
#| fig-cap: "Near-eye residual-versus-fitted and normal-score Q-Q model checks for the six analysed timing outcomes and both chronotype instruments. Visual review supported acceptable response/residual distribution and heteroscedasticity assessments."
#| fig-alt: "Twenty-four near-eye model-check panels show residuals versus fitted timing in the upper half and normal-score Q-Q plots in the lower half. Some first-above-250 and L10 panels depart more at the tails than other outcomes, while the overall patterns were judged acceptable."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H09",
  "H09_diagnostics_near_eye.png"
))
```

```{r}
#| label: fig-h09-chest-diagnostics
#| fig-cap: "Chest residual-versus-fitted and normal-score Q-Q model checks for the six analysed timing outcomes and both chronotype instruments. Visual review supported acceptable response/residual distribution and heteroscedasticity assessments."
#| fig-alt: "Twenty-four chest model-check panels show residuals versus fitted timing in the upper half and normal-score Q-Q plots in the lower half. Some first-above-250 and L10 panels depart more at the tails than other outcomes, while the overall patterns were judged acceptable."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root,
  "results",
  "images",
  "H09",
  "H09_diagnostics_chest.png"
))
```

The [complete model-check registry](../results/csv/diagnostics/H09/H09_diagnostic_assessment_registry.csv),
[residual-check assessment record](../results/csv/diagnostics/H09/H09_diagnostic_author_adjudication.csv),
[participant-influence summary](../results/csv/diagnostics/H09/H09_participant_influence_summary.csv),
and [leave-one-site-out summary](../results/csv/diagnostics/H09/H09_leave_one_site_out_summary.csv)
retain the thresholds, evidence, and final assessment basis.

## Dataset sensitivity

For these comparisons, a common sample means that the primary and
gap-timing-unaware datasets use the same participants and participant-days.
For the four registered outcomes available in both datasets, the models were
compared on those identical participant-day keys.
Fifteen of 16 instrument-by-placement comparisons were stable within model
uncertainty. Chest MCTQ L10 was precision-sensitive, with overlapping 95%
intervals but a change in whether the interval excluded zero. The registered
longest-period midpoint could not be evaluated because the gap-timing-unaware
artifact contains neither that metric nor the endpoints needed to derive it.

```{r}
#| label: tbl-h09-gap-sensitivity
#| tbl-cap: "Registered dataset-sensitivity comparisons requiring qualification."

gap_sensitivity |>
  filter(
    .data$metric_id != "mean_timing_above_250",
    .data$common_sample_stability != "stable within model uncertainty"
  ) |>
  arrange(.data$metric_order, .data$instrument_id, .data$placement_label) |>
  transmute(
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_name,
    `Primary effect (95% CI), h` = format_effect(
      .data$primary_common_estimate,
      .data$primary_common_conf_low,
      .data$primary_common_conf_high
    ),
    `Gap-timing-unaware effect (95% CI), h` = format_effect(
      .data$gap_common_estimate,
      .data$gap_common_conf_low,
      .data$gap_common_conf_high
    ),
    `Common participants / days / observations / hours / sites` =
      format_sample(
        .data$primary_common_participants,
        .data$primary_common_participant_days,
        .data$primary_common_observations,
        .data$primary_common_derivation_hours,
        .data$primary_common_sites
      ),
    Assessment = str_to_sentence(.data$common_sample_stability)
  ) |>
  h09_gt(
    note = paste(
      "The other 15 estimable registered comparisons were stable within",
      "model uncertainty on exact common keys. Non-estimable rows retain",
      "the missing prepared-data limitation rather than an imputed result."
    )
  )
```

The [complete dataset-sensitivity table](../results/tables/H09/H09_gap_timing_unaware_sensitivity.csv)
contains all exact samples, intervals, family-adjusted decisions, and row-key
identities.

## Other named sensitivity analyses

The photoperiod sensitivity added centered within-site photoperiod to the
participant-day model. It was stable for 21 of 24 targets; its two
direction-sensitive results were the near-zero near-eye last-timing slopes,
and one chest mean-timing estimate was precision-sensitive. The participant
summary sensitivity replaced repeated participant-day outcomes with one
equal-weight outcome mean per participant; it was stable for 20 of 24 targets,
with direction changes confined to near-zero last-timing or chest
longest-period slopes. All 24 continuous-time AR(1) checks of residual
autocorrelation were stable. The L10 clock-cut sensitivity compared the
registered strict `>16:00` negative-hour conversion with a noon cut; slopes
changed by at most 0.037 h, and all four checks were acceptable.

```{r}
#| label: tbl-h09-sensitivity-summary
#| tbl-cap: "Summary of structured sensitivity classifications."

tibble(
  Sensitivity = c(
    "Within-site photoperiod adjustment",
    "Equal-weight participant summary",
    "Continuous-time AR(1)",
    "L10 clock cut"
  ),
  `Targets assessed` = c(
    nrow(photoperiod),
    nrow(participant_summary),
    nrow(ar1),
    nrow(l10_cut)
  ),
  Result = c(
    sprintf(
      "%d stable; %d direction-sensitive; %d precision-sensitive",
      sum(photoperiod$stability_classification ==
        "stable within model uncertainty"),
      sum(photoperiod$stability_classification == "direction-sensitive"),
      sum(photoperiod$stability_classification == "precision-sensitive")
    ),
    sprintf(
      "%d stable; %d direction-sensitive",
      sum(participant_summary$stability_classification ==
        "stable within model uncertainty"),
      sum(participant_summary$stability_classification ==
        "direction-sensitive")
    ),
    sprintf(
      "%d stable; maximum slope change %.3f h",
      sum(ar1$stability_classification == "stable within model uncertainty"),
      max(abs(ar1$ar1_effect_difference))
    ),
    sprintf(
      "%d acceptable; maximum slope change %.3f h",
      sum(l10_cut$clock_cut_assessment == "acceptable"),
      max(abs(l10_cut$estimate_difference))
    )
  )
) |>
  h09_gt()
```

The registered fifth outcome remained the midpoint of the longest continuous
period. Mean timing across all supported exposure above 250 lx melEDI is a
distinct sensitivity estimand and was not substituted for that outcome. Its
near-eye MEQ estimate was −0.169 h per 10 MEQ points (95% CI −0.334 to −0.003;
raw *p* = 0.041), but this raw result is outside the registered multiplicity
families and does not alter the confirmatory conclusion.

```{r}
#| label: tbl-h09-mean-timing-sensitivity
#| tbl-cap: "Distinct mean-timing sensitivity outside the registered five-metric families."

mean_timing |>
  arrange(.data$placement, .data$instrument_id) |>
  transmute(
    Placement = .data$placement_label,
    Instrument = .data$instrument_name,
    `Effect in local clock hours (95% CI)` = format_effect(
      .data$estimate,
      .data$conf_low,
      .data$conf_high
    ),
    `Raw p` = format_p_cell(
      .data$main_p_raw,
      .data$main_p_raw < 0.05
    ),
    `Participants / days / observations / hours / sites` = format_sample(
      .data$participants,
      .data$participant_days,
      .data$observations,
      .data$derivation_hours,
      .data$sites
    )
  ) |>
  h09_gt(
    note = paste(
      "Intervals are 95% Wald confidence intervals. Raw p-values are bold",
      "when raw p < 0.050. No FDR-adjusted p-value is assigned because this",
      "distinct estimand is outside the registered five-outcome families."
    )
  )
```

[Photoperiod](../results/tables/H09/H09_photoperiod_sensitivity.csv),
[participant-summary](../results/tables/H09/H09_participant_summary_sensitivity.csv),
[continuous-time AR(1)](../results/tables/H09/H09_ar1_sensitivity.csv),
[L10 clock-cut](../results/tables/H09/H09_l10_cut_sensitivity.csv),
and [mean-timing](../results/tables/H09/H09_fifth_outcome_sensitivity.csv)
source tables retain exact samples and 95% CIs.

## Interpretation

The primary near-eye evidence supports an association between chronotype and
the timing of several, but not all, features of personal light exposure.
Participants with later corrected midsleep tended to have later M10, L10, and
first-above-250 timing; participants with greater morning preference tended to
show the corresponding earlier timing. The agreement of these distinct
chronotype constructs strengthens the directional interpretation without
making them interchangeable.

The data did not support an association with last-above-250 timing or the
registered longest-period midpoint, nor did they support chronotype-by-site
interactions after FDR adjustment. Chest evidence was complementary and
broadly consistent, but it neither replaces the primary near-eye results nor
establishes placement equivalence. The analysis is observational and between
participants; it does not establish that chronotype causes a change in
personal light exposure timing.

The conclusion should be read alongside the direction instability of the
near-zero near-eye last-timing estimates, the qualified chest
longest-period/MEQ interaction fit, and the unavailable longest-period
gap-timing-unaware comparison. These limitations do not change the supported
M10, L10, and first-timing pattern, but they constrain claims about the
unsupported outcomes and cross-dataset robustness.

## Preregistration deviations {#h09-preregistration-deviations}

- [H09 chronotype predictors](../preregistration-deviations.qmd#h09-chronotype-predictors): MCTQ MSFsc and MEQ
  remain distinct chronotype constructs with separate models, score
  directions, and inferential families.
- [H09 site adjustment](../preregistration-deviations.qmd#h09-site-adjustment): study-site-adjusted
  average chronotype associations and chronotype-by-site interactions are
  fitted and interpreted as separate questions.
- [H09 multiplicity and model selection](../preregistration-deviations.qmd#h09-multiplicity-and-model-selection): the analysis uses
  four complete five-outcome FDR families;MCTQ average, MEQ average,
  MCTQ-by-site interaction, and MEQ-by-site interaction.
