---
title: "H11: Biological sex and daily personal light-exposure patterns"
engine: knitr
format:
  html:
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

Biological-sex-specific daily light patterns are estimated on the same temporal model structure used for the geographic daily-pattern analysis. Participant-cluster-robust tests assess the complete Female-minus-Male curve, and pointwise intervals describe its local uncertainty.

## Data and model guide {#methods}

This analysis inherits the [H02 half-hour model frames](H02-daily-patterns.qmd#methods)
and joins biological sex from the prepared participant information. The inherited
coverage rules, exact-zero handling, `log10(melEDI + 0.1)` response and elapsed-time
AR sequence boundaries are retained. Near-eye and chest measurements are fitted
separately. Complete available-data samples and activity-complete samples answer
different comparison questions.

The nonlinear model uses a cyclic Male reference curve and a cyclic Female-minus-Male
shape deviation, together with a constant sex effect, sum-to-zero site deviations,
participant curves and participant-day intercepts. AR(1) represents remaining
short-lag dependence. Exact formulas, model settings and sample counts are displayed
below. The activity-adjusted model uses Home as its reference and shares a smoothing
parameter across the ordered activity deviations.

The complete-curve test jointly evaluates constant and time-varying sex contributions.
It uses participant-cluster CR1 covariance from AR-whitened scores, the smoothing-bias
correction and a finite-cluster fractional-rank F reference. Pointwise intervals
apply to individual clock bins and are not simultaneous day-wide bands. Conditional
effect sizes and their bootstrap uncertainty describe the fitted response structure.
Alternative preprocessing and the sequence of sample restriction followed by activity
adjustment assess stability. Attenuation after restriction and adjustment cannot by
itself identify a behavioural mechanism.

The executable sections below write fitted objects to `results/models/H11/`,
reader tables to `results/tables/H11/`, and the supporting numerical and figure
outputs under `results/`. Start with the calculations below or jump to
[findings and interpretation](#findings).

## Model structure and inputs

Use the regenerated half-hour H02 frames, measured biological-sex data and the inherited temporal formula. Near-eye and chest curves are fitted independently.

```{r}
#| label: h11-setup

source("scripts/project.R")
analysis_setup()
source("scripts/pipeline/paths_io.R")
source("scripts/pipeline/p_value_display.R")
root <- normalizePath(Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()))
producer <- "analyses/H11-sex-daily-patterns.qmd"
options(stringsAsFactors = FALSE, scipen = 999, dplyr.summarise.inform = FALSE)
roots <- list(model_data = file.path(root,"results/intermediate/model_data/H11"), models = file.path(root,"results/models/H11"), diagnostics = file.path(root,"results/csv/diagnostics/H11"), tables = file.path(root,"results/tables/H11"), figures = file.path(root,"results/images/H11"), source_data = file.path(root,"results/csv/source_data/H11"))
invisible(lapply(roots, dir.create, recursive=TRUE, showWarnings=FALSE))
metadata <- list()
suppressPackageStartupMessages({library(dplyr);library(gratia);library(mgcv);library(readr);library(tibble);library(tidyr)})
source("scripts/hypotheses/H02/h02_contract.R")
source("scripts/hypotheses/H02/h02_modeling.R")
source("scripts/hypotheses/H02/h02_dominance.R")
source("scripts/hypotheses/H11/h11_data.R")
source("scripts/hypotheses/H11/h11_modeling.R")
source("scripts/hypotheses/H11/h11_activity_context.R")
paths <- h11_paths(root)
h11_create_directories(paths)
write_model_csv <- function(data, path) { readr::write_csv(data, path, na = ""); invisible(path) }
```


## Fit sex-specific temporal curves

Fit preliminary curves to estimate within-sequence autocorrelation, then fit the final curves using the same sequence boundaries and run-specific correlation. Save exact samples and fitted model objects.

```{r}
#| label: h11-fit

registry <- h11_registry(root)

active_registry <- registry

formulas <- h11_formula_set(root)

formula_plan <- tidyr::crossing(
  registry |>
    dplyr::select(
      "run_id", "data_scenario_id", "placement", "analytical_role"
    ),
  tibble::tribble(
    ~model_id, ~formula_id, ~method, ~discrete, ~rho_rule, ~fit_role,
    "mpattern_preliminary_rho0_fREML", "proposed_mpattern", "fREML", TRUE,
    "rho = 0", "preliminary boundary-aware rho estimation",
    "mpattern_final_fREML", "proposed_mpattern", "fREML", TRUE,
    "one run-specific rho from preliminary full model", "final estimation",
    "robust_complete_curve_test", "proposed_mpattern", "post-fit robust Wald", NA,
    "retained final-model rho and AR.start whitening", "global inference",
    "robust_level_shape_tests", "proposed_mpattern", "post-fit robust Wald", NA,
    "retained final-model rho and AR.start whitening", "BH-adjusted decomposition inference",
    "m0_effect_size_fREML", "proposed_m0", "fREML", TRUE,
    "one run-specific rho from preliminary full model",
    "conditional effect-size baseline; fitted only when placement criterion opens"
  )
) |>
  dplyr::mutate(
    formula = vapply(
      .data$formula_id,
      function(id) h11_formula_text(formulas[[id]]),
      character(1)
    ),
    family = "Gaussian",
    link = "identity",
    knots = "time_hour = c(0, 24)",
    biological_sex_encoding = paste(
      "sex treatment contrast with Male reference; sex_smooth ordered",
      "Male then Female for cyclic Female-minus-Male deviation"
    )
  )

write_model_csv(
  formula_plan,
  file.path(paths$model_data, "formula_and_fit_settings.csv")
)

demographics <- readRDS(file.path(
  root,
  "results/intermediate/model_data/normalized_inputs/demographics.rds"
))

message("Beginning H11 analysis fits")

for (index in seq_len(nrow(active_registry))) {
  run <- active_registry[index, , drop = FALSE]
  message("H11 run: ", run$run_id)
  frame <- readRDS(run$frame_path)
  h11_fit_run(frame, demographics, run, formulas, paths)
  rm(frame)
  invisible(gc())
}

bundle_paths <- file.path(
  paths$models,
  registry$run_id,
  "fit_bundle.rds"
)

bundles <- lapply(bundle_paths, readRDS)

names(bundles) <- registry$run_id

sample_counts <- dplyr::bind_rows(lapply(bundles, `[[`, "sample"))

model_fit_summary <- dplyr::bind_rows(lapply(bundles, `[[`, "model_rows"))

rho_summary <- dplyr::bind_rows(lapply(bundles, `[[`, "rho_row"))

write_model_csv(sample_counts, file.path(paths$tables, "sample_counts.csv"))

write_model_csv(
  model_fit_summary,
  file.path(paths$tables, "model_fit_summary.csv")
)

write_model_csv(rho_summary, file.path(paths$tables, "rho_summary.csv"))
```


## Sensor-position samples

Describe whether a position-matched visual comparison is supported by the underlying samples. The numerical models continue to have position-specific scopes.

```{r}
#| label: h11-samples

h02_sample_counts <- utils::read.csv(
  file.path(root, "results/intermediate/model_data/H02/sample_counts.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

paired_run_ids <- c(
  "main__glasses__paired_common_sample",
  "main__chest__paired_common_sample"
)

paired_counts <- h02_sample_counts |>
  dplyr::filter(
    .data$run_id %in% paired_run_ids,
    .data$site == "ALL_SITES"
  ) |>
  dplyr::arrange(match(.data$run_id, paired_run_ids))

paired_frame_paths <- file.path(
  root,
  "results/intermediate/model_data/H02",
  paste0(paired_run_ids, ".rds")
)

paired_display_assessment <- tibble::tibble(
  all_available_near_eye_participants = sample_counts$participants[
    sample_counts$run_id == "main__glasses__all_available"
  ],
  all_available_chest_participants = sample_counts$participants[
    sample_counts$run_id == "main__chest__all_available"
  ],
  all_available_near_eye_participant_days = sample_counts$participant_days[
    sample_counts$run_id == "main__glasses__all_available"
  ],
  all_available_chest_participant_days = sample_counts$participant_days[
    sample_counts$run_id == "main__chest__all_available"
  ],
  all_available_near_eye_observations = sample_counts$observations_30_minute[
    sample_counts$run_id == "main__glasses__all_available"
  ],
  all_available_chest_observations = sample_counts$observations_30_minute[
    sample_counts$run_id == "main__chest__all_available"
  ],
  all_available_near_eye_sites = sample_counts$sites[
    sample_counts$run_id == "main__glasses__all_available"
  ],
  all_available_chest_sites = sample_counts$sites[
    sample_counts$run_id == "main__chest__all_available"
  ],
  stored_paired_frame_participants = paired_counts$participants[[1L]],
  stored_paired_frame_participant_days = paired_counts$participant_days[[1L]],
  stored_paired_frame_observations = paired_counts$observations_30_minute[[1L]],
  stored_paired_frame_sites = paired_counts$sites[[1L]],
  stored_h02_paired_frames_available = all(file.exists(paired_frame_paths)),
  h11_paired_fitted_outputs_available = any(grepl(
    "paired_common_sample",
    model_fit_summary$run_id,
    fixed = TRUE
  )),
  predeclared_scalar_temporal_estimand_available = FALSE,
  scalar_identity_plot_applicable = FALSE,
  applicability_reason = paste(
    "The fitted H11 analysis outputs use unmatched all-available samples,",
    "and the registered paired/common H11 sensitivity has not been fitted.",
    "The H11 result is a temporal curve without a predeclared scalar",
    "reduction, so an H05-style scalar identity scatter would be misleading."
  ),
  closest_valid_display = paste(
    "Retain separate primary near-eye and complementary chest curves in",
    "analysis without direct paired interpretation. A paired/common",
    "sensitivity would require matched placement curves",
    "or Female-minus-Male contrast curves; reconsider a clock-bin identity",
    "display only after exact estimand comparability is verified."
  ),
  paired_source_data_csv = NA_character_,
  source_data_status = paste(
    "Not created because no valid H11 paired fitted estimands exist at",
    "the analysed samples; unmatched outputs are not paired."
  )
)

write_model_csv(
  paired_display_assessment,
  file.path(paths$model_data, "paired_placement_display_assessment.csv")
)
```


## Robust tests, intervals and model diagnostics

Compute participant-cluster CR1 covariance with the smooth-uncertainty correction, global and level/shape tests, pointwise curves, residual checks and site-constraint diagnostics.

```{r}
#| label: h11-inference

curve_tables <- list()

contrast_tables <- list()

contrast_contracts <- list()

closure_tables <- list()

parametric_tables <- list()

acf_tables <- list()

residual_tables <- list()

smooth_tables <- list()

concurvity_tables <- list()

k_check_tables <- list()

site_constraint_tables <- list()

assessment_tables <- list()

comparison_tables <- list()

robust_diagnostic_tables <- list()

message("Calculating final-model contrasts and diagnostics")

for (index in seq_len(nrow(registry))) {
  run <- registry[index, , drop = FALSE]
  bundle <- bundles[[run$run_id]]
  data <- readRDS(bundle$frame_path)
  preliminary <- readRDS(bundle$model_paths$preliminary)
  final <- readRDS(bundle$model_paths$final)

  message("  curves and diagnostics: ", run$run_id)
  robust_context <- h11_robust_context(
    final,
    data,
    run$run_id
  )
  robust_test <- h11_robust_tests(
    final,
    data,
    run,
    robust_context
  )
  comparison_tables[[run$run_id]] <- robust_test$comparisons
  robust_diagnostic_tables[[run$run_id]] <- robust_test$diagnostics
  curve_contract <- h11_pointwise_curves(
    final,
    data,
    run$run_id,
    robust_context
  )
  curve_tables[[run$run_id]] <- curve_contract$curves
  contrast_tables[[run$run_id]] <- curve_contract$contrast
  contrast_contracts[[run$run_id]] <- curve_contract
  closure_tables[[run$run_id]] <- h11_curve_closure(
    final, data, run$run_id
  )
  parametric_tables[[run$run_id]] <- h11_parametric_sex(
    final, run$run_id, robust_context
  )
  acf_tables[[run$run_id]] <- dplyr::bind_rows(
    h11_residual_acf(
      preliminary, data, run$run_id, "preliminary_rho0_response"
    ),
    h11_residual_acf(
      final, data, run$run_id, "final_AR1_standardized"
    )
  )
  residual_tables[[run$run_id]] <- h11_residual_summary(
    final, data, run$run_id
  )
  smooth_tables[[run$run_id]] <- h11_smooth_summary(
    final, run$run_id
  )
  concurvity_tables[[run$run_id]] <- h11_concurvity(
    final, run$run_id
  )
  k_check_tables[[run$run_id]] <- h11_k_check(
    final,
    run$run_id,
    seed = h02_seed(run$run_id, 90L)
  )
  site_constraint_tables[[run$run_id]] <- h11_site_constraint(
    final, data, run$run_id
  )
  assessment_tables[[run$run_id]] <- h11_diagnostic_assessment(
    bundle$model_rows,
    acf_tables[[run$run_id]],
    closure_tables[[run$run_id]],
    site_constraint_tables[[run$run_id]],
    k_check_tables[[run$run_id]],
    run$run_id
  )

  rm(data, preliminary, final, robust_context, robust_test)
  invisible(gc())
}

sex_curves <- dplyr::bind_rows(curve_tables)

sex_contrasts <- dplyr::bind_rows(contrast_tables)

pointwise_segments <- h11_pointwise_segments(sex_contrasts)

curve_closure <- dplyr::bind_rows(closure_tables)

parametric_sex <- dplyr::bind_rows(parametric_tables)

residual_acf <- dplyr::bind_rows(acf_tables)

residual_summary <- dplyr::bind_rows(residual_tables)

smooth_summary <- dplyr::bind_rows(smooth_tables)

formal_concurvity <- dplyr::bind_rows(concurvity_tables)

k_check <- dplyr::bind_rows(k_check_tables)

site_constraint <- dplyr::bind_rows(site_constraint_tables)

diagnostic_assessment <- dplyr::bind_rows(assessment_tables)

model_comparisons <- h11_adjust_multiplicity(
  dplyr::bind_rows(comparison_tables)
)

robust_inference_diagnostics <- dplyr::bind_rows(robust_diagnostic_tables)

write_model_csv(
  model_comparisons,
  file.path(paths$tables, "model_comparisons.csv")
)

write_model_csv(
  robust_inference_diagnostics,
  file.path(paths$diagnostics, "robust_inference_diagnostics.csv")
)

write_model_csv(
  sex_curves,
  file.path(paths$source_data, "sex_specific_curves_pointwise.csv")
)

write_model_csv(
  sex_contrasts,
  file.path(paths$source_data, "female_minus_male_pointwise_contrasts.csv")
)

write_model_csv(
  pointwise_segments,
  file.path(paths$tables, "pointwise_descriptive_segments.csv")
)

write_model_csv(
  curve_closure,
  file.path(paths$diagnostics, "cyclic_curve_closure.csv")
)

write_model_csv(
  parametric_sex,
  file.path(paths$tables, "parametric_sex_component.csv")
)

write_model_csv(
  residual_acf,
  file.path(paths$diagnostics, "boundary_aware_residual_acf.csv")
)

write_model_csv(
  residual_summary,
  file.path(paths$diagnostics, "residual_summary.csv")
)

write_model_csv(
  smooth_summary,
  file.path(paths$diagnostics, "smooth_summary.csv")
)

write_model_csv(
  formal_concurvity,
  file.path(paths$diagnostics, "formal_concurvity.csv")
)

write_model_csv(
  k_check,
  file.path(paths$diagnostics, "basis_dimension_check.csv")
)

write_model_csv(
  site_constraint,
  file.path(paths$diagnostics, "site_sum_to_zero_constraint.csv")
)

write_model_csv(
  diagnostic_assessment,
  file.path(paths$diagnostics, "diagnostic_assessment.csv")
)
```


## Conditional effect sizes

When the raw global test is below 0.05, describe additional in-sample variation beyond the temporal baseline. Hierarchical resampling uses 50 replicates by default and supports descriptive uncertainty rather than a separate confirmatory test.

```{r}
#| label: h11-effect-sizes

main_global <- model_comparisons |>
  dplyr::filter(
    .data$data_scenario_id == "main",
    .data$comparison_role == "global"
  )

effect_condition <- main_global |>
  dplyr::transmute(
    run_id = .data$run_id,
    placement = .data$placement,
    family_id = .data$family_id,
    p_raw = .data$p_raw,
    p_adjusted_BH = .data$p_adjusted_BH,
    support_status = .data$support_status,
    effect_size_condition_open = .data$p_raw < 0.05,
    criterion_rule = paste(
      "Open within this main placement only when the retained one-test",
      "global participant-cluster-robust raw p < 0.050; chest cannot alter",
      "the primary near-eye criterion."
    )
  )

write_model_csv(
  effect_condition,
  file.path(paths$tables, "effect_size_condition.csv")
)

effect_model_rows <- list()

effect_bootstrap_tables <- list()

effect_variation_tables <- list()

effect_prediction_metadata <- list()

for (index in which(effect_condition$effect_size_condition_open)) {
  criterion <- effect_condition[index, , drop = FALSE]
  run <- registry[match(criterion$run_id, registry$run_id), , drop = FALSE]
  bundle <- bundles[[run$run_id]]
  data <- readRDS(bundle$frame_path)
  final <- readRDS(bundle$model_paths$final)
  effect_robust_context <- h11_robust_context(
    final,
    data,
    run$run_id
  )
  baseline <- h11_fit_and_save(
    formulas$proposed_m0,
    data,
    "fREML",
    bundle$rho,
    TRUE,
    "m0_effect_size_fREML",
    run$run_id,
    paths$models
  )
  effect_model_rows[[run$run_id]] <- h11_model_row(
    baseline,
    run,
    "proposed_m0"
  ) |>
    dplyr::mutate(
      fit_role = "conditional H02-like effect-size baseline",
      inferential_role = "descriptive fitted-model relevance only"
    )
  baseline_prediction <- as.numeric(stats::fitted(baseline$fit))
  full_prediction <- as.numeric(stats::fitted(final))
  predictions <- tibble::tibble(
    run_id = run$run_id,
    row_index = seq_len(nrow(data)),
    response = data$response,
    baseline_prediction = baseline_prediction,
    full_prediction = full_prediction
  )
  prediction_path <- file.path(
    paths$source_data,
    paste0(run$run_id, "__effect_size_fixed_predictions.rds")
  )
  saveRDS(predictions, prediction_path, compress = "xz")
  resampling <- h11_bootstrap_r2(
    response = data$response,
    baseline_prediction = baseline_prediction,
    full_prediction = full_prediction,
    data = data,
    replicates = bootstrap_count(50L),
    seed = h02_seed(run$run_id, 110L)
  )
  bootstrap_path <- file.path(
    paths$diagnostics,
    paste0(run$run_id, "__effect_size_bootstrap.rds")
  )
  saveRDS(resampling, bootstrap_path, compress = "xz")
  effect_bootstrap_tables[[run$run_id]] <- h11_effect_bootstrap_summary(
    resampling,
    run$run_id
  ) |>
    dplyr::mutate(
      placement = run$placement,
      participants = dplyr::n_distinct(data$participant),
      participant_days = dplyr::n_distinct(data$participant_day),
      observations_30_minute = nrow(data),
      sites = dplyr::n_distinct(data$site),
      .after = "run_id"
    )
  effect_variation_tables[[run$run_id]] <- h11_curve_variation(
    final,
    contrast_contracts[[run$run_id]],
    run$run_id,
    effect_robust_context
  ) |>
    dplyr::mutate(
      placement = run$placement,
      participants = dplyr::n_distinct(data$participant),
      participant_days = dplyr::n_distinct(data$participant_day),
      observations_30_minute = nrow(data),
      sites = dplyr::n_distinct(data$site),
      .after = "run_id"
    )
  effect_prediction_metadata[[run$run_id]] <- tibble::tibble(
    run_id = run$run_id,
    placement = run$placement,
    prediction_path = sub(paste0("^", root, "/"), "", prediction_path),
    bootstrap_path = sub(paste0("^", root, "/"), "", bootstrap_path),
    baseline_formula = h11_formula_text(formulas$proposed_m0),
    full_formula = h11_formula_text(formulas$proposed_mpattern),
    rho = bundle$rho,
    effect_size_scope = paste(
      "Exact one-block conditional Shapley/general-dominance allocation",
      "beyond the inherited H02 temporal baseline; row-weighted in-sample",
      "R-squared on log10(melEDI + 0.1 lx); fixed predictions."
    )
  )
  rm(
    data,
    final,
    baseline,
    resampling,
    predictions,
    effect_robust_context
  )
  invisible(gc())
}

effect_model_summary <- dplyr::bind_rows(effect_model_rows)

effect_size_bootstrap <- dplyr::bind_rows(effect_bootstrap_tables)

effect_curve_variation <- dplyr::bind_rows(effect_variation_tables)

effect_prediction_export_index <- dplyr::bind_rows(effect_prediction_metadata)

write_model_csv(
  effect_model_summary,
  file.path(paths$tables, "effect_size_model_fit_summary.csv")
)

write_model_csv(
  effect_size_bootstrap,
  file.path(paths$tables, "effect_size_bootstrap.csv")
)

write_model_csv(
  effect_curve_variation,
  file.path(paths$tables, "effect_size_curve_variation.csv")
)
```


## Alternative preprocessing

Compare complete-curve tests between primary preprocessing and the gap-timing-unaware dataset. These are sensitivity results with their own samples.

```{r}
#| label: h11-preprocessing

new_global <- model_comparisons |>
  dplyr::filter(.data$comparison_role == "global") |>
  dplyr::select(
    "run_id", "data_scenario_id", "placement", "analytical_role",
    "test_statistic", "test_df", "denominator_df",
    "reduced_AIC", "full_AIC", "delta_AIC_reduced_minus_full",
    "p_raw", "p_adjusted_BH", "support_status"
  ) |>
  dplyr::left_join(
    sample_counts |>
      dplyr::select(
        "run_id", "participants", "participant_days",
        "observations_30_minute", "sites"
      ),
    by = "run_id",
    relationship = "one-to-one"
  )

data_preparation_comparison <- new_global |>
  dplyr::select(
    "placement", "data_scenario_id", "test_statistic", "test_df",
    "denominator_df",
    "p_raw", "p_adjusted_BH", "support_status",
    "participants", "participant_days", "observations_30_minute", "sites"
  ) |>
  tidyr::pivot_wider(
    names_from = "data_scenario_id",
    values_from = c(
      "test_statistic", "test_df", "denominator_df", "p_raw", "p_adjusted_BH",
      "support_status", "participants", "participant_days",
      "observations_30_minute", "sites"
    ),
    names_sep = "__"
  ) |>
  dplyr::mutate(
    added_participant_days_main_minus_alternative =
      .data$participant_days__main -
      .data$participant_days__alternative_preprocessing,
    added_observations_main_minus_alternative =
      .data$observations_30_minute__main -
      .data$observations_30_minute__alternative_preprocessing,
    robust_statistic_change_main_minus_gap_timing_unaware =
      .data$test_statistic__main -
      .data$test_statistic__alternative_preprocessing,
    raw_p_change_main_minus_gap_timing_unaware =
      .data$p_raw__main -
      .data$p_raw__alternative_preprocessing,
    interpretation_scope = paste(
      "This contrast holds the specified H11 implementation fixed and varies",
      "the shared preparation scenario; it does not isolate one individual",
      "upstream correction."
    )
  )

write_model_csv(
  data_preparation_comparison,
  file.path(paths$tables, "data_preparation_comparison.csv")
)
```


## Restrict and adjust for immediate setting

Construct the exact diary-complete sample, then fit unadjusted and activity-adjusted curves on that same sample. Each multiselect hour uses its defined diary representation; restriction and adjustment are examined separately.

```{r}
#| label: h11-activity-fit

paths <- h11_activity_paths(root)

h11_activity_create_directories(paths)

write_activity_csv <- function(data, path) {
  readr::write_csv(data, path, na = "")
  invisible(path)
}

relative_path <- function(path) {
  sub(paste0("^", root, "/"), "", normalizePath(
    path,
    winslash = "/",
    mustWork = TRUE
  ))
}

registry <- tibble::tribble(
  ~run_id, ~placement, ~analytical_role, ~primary_frame_relative_path,
  "activity_context__glasses", "glasses", "primary near-eye sensitivity",
  "results/intermediate/model_data/H11/fitted/main__glasses__all_available__frame.rds",
  "activity_context__chest", "chest", "complementary chest sensitivity",
  "results/intermediate/model_data/H11/fitted/main__chest__all_available__frame.rds"
) |>
  dplyr::mutate(
    primary_frame_path = file.path(root, .data$primary_frame_relative_path)
  )

diary_path <- file.path(
  root,
  "results/intermediate/model_data/normalized_inputs/lightexposurediary.rds"
)

diary <- h11_activity_prepare_diary(readRDS(diary_path))

write_activity_csv(
  diary$diagnostic,
  file.path(paths$model_data, "normalized_diary_activity_diagnostic.csv")
)

prepared <- list()

sample_rows <- list()

support_rows <- list()

join_rows <- list()

boundary_rows <- list()

for (index in seq_len(nrow(registry))) {
  run <- registry[index, , drop = FALSE]
  primary_frame <- readRDS(run$primary_frame_path)
  attached <- h11_activity_attach(
    primary_frame,
    diary$data,
    run$run_id
  )
  data <- attached$data
  frame_path <- file.path(paths$model_data, paste0(run$run_id, "__frame.rds"))
  saveRDS(data, frame_path, compress = "xz")
  prepared[[run$run_id]] <- list(data = data, frame_path = frame_path)
  sample_rows[[run$run_id]] <- h11_activity_sample_row(
    data,
    run$run_id,
    run$placement
  )
  support_rows[[run$run_id]] <- h11_activity_support(
    data,
    run$run_id,
    run$placement
  )
  join_rows[[run$run_id]] <- attached$join_diagnostic |>
    dplyr::mutate(placement = run$placement, .after = "run_id")
  boundary_rows[[run$run_id]] <- data |>
    dplyr::count(.data$activity_AR_start_reason, name = "observations_30_minute") |>
    dplyr::mutate(
      run_id = run$run_id,
      placement = run$placement,
      AR_start = .data$activity_AR_start_reason != "continuous",
      .before = 1L
    )
  rm(primary_frame, attached)
}

sample_counts <- dplyr::bind_rows(sample_rows)

activity_support <- dplyr::bind_rows(support_rows)

join_diagnostic <- dplyr::bind_rows(join_rows)

boundary_diagnostic <- dplyr::bind_rows(boundary_rows)

write_activity_csv(
  sample_counts,
  file.path(paths$tables, "activity_common_sample_counts.csv")
)

write_activity_csv(
  activity_support,
  file.path(paths$tables, "activity_category_support_by_sex.csv")
)

write_activity_csv(
  join_diagnostic,
  file.path(paths$model_data, "activity_join_attrition.csv")
)

write_activity_csv(
  boundary_diagnostic,
  file.path(paths$model_data, "activity_AR_boundary_diagnostic.csv")
)

formulas <- h11_activity_formulas(root)

formula_settings <- tibble::tribble(
  ~formula_id, ~fit_role, ~rho_rule, ~fit_criterion,
  "restricted_unadjusted", "same-sample unadjusted H11 sensitivity",
  "rho from preliminary activity-adjusted model", "always",
  "activity_adjusted", "same-sample activity-adjusted H11 sensitivity",
  "rho from preliminary activity-adjusted model", "always",
  "activity_adjusted_no_sex", "conditional effect-size baseline",
  "same final rho", "only if adjusted global raw p < 0.050"
) |>
  dplyr::mutate(
    formula = vapply(
      .data$formula_id,
      function(id) h11_formula_text(formulas[[id]]),
      character(1)
    ),
    family = "Gaussian",
    link = "identity",
    method = "fREML",
    discrete = TRUE,
    knots = "time_hour = c(0, 24)",
    activity_encoding = paste(
      "Treatment-coded activity level with home reference; ordered-factor",
      "cyclic deviations with common smoothing parameter id = 2"
    ),
    interval_method = "participant-cluster-robust pointwise 95% intervals"
  )

write_activity_csv(
  formula_settings,
  file.path(paths$model_data, "formula_and_fit_settings.csv")
)

active_registry <- registry

message("Beginning discrete=TRUE activity-context fits")

for (index in seq_len(nrow(active_registry))) {
  run <- active_registry[index, , drop = FALSE]
  data <- prepared[[run$run_id]]$data
  preliminary <- h11_activity_fit_and_save(
    formula = formulas$activity_adjusted,
    data = data,
    method = "fREML",
    rho = 0,
    model_id = "activity_adjusted_preliminary_rho0_fREML",
    run_id = run$run_id,
    model_directory = paths$models
  )
  rho <- h02_estimate_rho(preliminary$fit, data)
  if (!is.finite(rho) || abs(rho) > 0.95) {
    h11_activity_abort("Invalid boundary-aware rho for %s", run$run_id)
  }
  unadjusted <- h11_activity_fit_and_save(
    formula = formulas$restricted_unadjusted,
    data = data,
    method = "fREML",
    rho = rho,
    model_id = "restricted_unadjusted_final_fREML",
    run_id = run$run_id,
    model_directory = paths$models
  )
  adjusted <- h11_activity_fit_and_save(
    formula = formulas$activity_adjusted,
    data = data,
    method = "fREML",
    rho = rho,
    model_id = "activity_adjusted_final_fREML",
    run_id = run$run_id,
    model_directory = paths$models
  )
  bundle <- list(
    run_id = run$run_id,
    placement = run$placement,
    analytical_role = run$analytical_role,
    frame_path = prepared[[run$run_id]]$frame_path,
    rho = rho,
    preliminary_path = preliminary$model_path,
    unadjusted_path = unadjusted$model_path,
    adjusted_path = adjusted$model_path,
    preliminary_metadata_path = preliminary$metadata_path,
    unadjusted_metadata_path = unadjusted$metadata_path,
    adjusted_metadata_path = adjusted$metadata_path,
    completed_at = format(Sys.time(), tz = "Europe/Berlin", usetz = TRUE)
  )
  saveRDS(
    bundle,
    file.path(paths$models, run$run_id, "activity_context_run_bundle.rds"),
    compress = "xz"
  )
  rm(data, preliminary, unadjusted, adjusted, bundle)
  invisible(gc())
}

bundle_paths <- file.path(
  paths$models,
  registry$run_id,
  "activity_context_run_bundle.rds"
)
```


## Activity-adjusted comparisons

Compute robust tests and contrast attenuation on the activity-complete samples. These associational comparisons do not identify causal mediation.

```{r}
#| label: h11-activity-inference

bundles <- lapply(bundle_paths, readRDS)

names(bundles) <- registry$run_id

model_rows <- list()

rho_rows <- list()

comparison_rows <- list()

robust_rows <- list()

curve_rows <- list()

contrast_rows <- list()

acf_rows <- list()

residual_rows <- list()

smooth_rows <- list()

concurvity_rows <- list()

k_rows <- list()

closure_rows <- list()

site_rows <- list()

assessment_rows <- list()

effect_variation_rows <- list()

message("Computing robust tests, pointwise intervals, and diagnostics")

for (index in seq_len(nrow(registry))) {
  run <- registry[index, , drop = FALSE]
  bundle <- bundles[[run$run_id]]
  data <- readRDS(bundle$frame_path)
  fit_results <- list(
    restricted_unadjusted = list(
      fit = readRDS(bundle$unadjusted_path),
      metadata = readRDS(bundle$unadjusted_metadata_path)
    ),
    activity_adjusted = list(
      fit = readRDS(bundle$adjusted_path),
      metadata = readRDS(bundle$adjusted_metadata_path)
    )
  )
  preliminary <- readRDS(bundle$preliminary_path)
  rho_rows[[run$run_id]] <- tibble::tibble(
    run_id = run$run_id,
    placement = run$placement,
    rho = bundle$rho,
    rho_source_model = "activity_adjusted_preliminary_rho0_fREML",
    rho_estimator = paste(
      "Boundary-aware lag-1 correlation of preliminary response residuals;",
      "held fixed across both exact-common-sample final fits"
    ),
    preliminary_lag1_response_correlation = unname(
      h02_boundary_lag_correlation(
        stats::residuals(preliminary, type = "response"),
        data$AR_start,
        lag = 1L
      )["correlation"]
    ),
    AR_sequences = sum(data$AR_start)
  )

  for (model_variant in names(fit_results)) {
    result <- fit_results[[model_variant]]
    fit <- result$fit
    diagnostic_run_id <- paste(run$run_id, model_variant, sep = "__")
    row <- h11_activity_model_row(
      result,
      run$run_id,
      run$placement,
      model_variant
    )
    model_rows[[diagnostic_run_id]] <- row
    context <- h11_robust_context(fit, data, diagnostic_run_id)
    robust <- h11_activity_robust_test(
      fit,
      data,
      run$run_id,
      run$placement,
      model_variant,
      context
    )
    comparison_rows[[diagnostic_run_id]] <- robust$comparison
    robust_rows[[diagnostic_run_id]] <- robust$diagnostics
    pointwise <- h11_activity_pointwise_curves(
      fit,
      data,
      run$run_id,
      run$placement,
      model_variant,
      context
    )
    curve_rows[[diagnostic_run_id]] <- pointwise$curves
    contrast_rows[[diagnostic_run_id]] <- pointwise$contrast
    closure <- h11_activity_curve_closure(
      fit,
      data,
      run$run_id,
      run$placement,
      model_variant
    )
    closure_rows[[diagnostic_run_id]] <- closure
    acf <- h11_residual_acf(
      fit,
      data,
      diagnostic_run_id,
      "final_AR1_standardized"
    ) |>
      dplyr::mutate(
        base_run_id = run$run_id,
        placement = run$placement,
        model_variant = model_variant,
        .after = "run_id"
      )
    acf_rows[[diagnostic_run_id]] <- acf
    residual_rows[[diagnostic_run_id]] <- h11_residual_summary(
      fit,
      data,
      diagnostic_run_id
    ) |>
      dplyr::mutate(
        base_run_id = run$run_id,
        placement = run$placement,
        model_variant = model_variant,
        .after = "run_id"
      )
    smooth_rows[[diagnostic_run_id]] <- h11_smooth_summary(
      fit,
      diagnostic_run_id
    ) |>
      dplyr::mutate(
        base_run_id = run$run_id,
        placement = run$placement,
        model_variant = model_variant,
        .after = "run_id"
      )
    concurvity_rows[[diagnostic_run_id]] <- h11_concurvity(
      fit,
      diagnostic_run_id
    ) |>
      dplyr::mutate(
        base_run_id = run$run_id,
        placement = run$placement,
        model_variant = model_variant,
        .after = "run_id"
      )
    k_check <- h11_k_check(
      fit,
      diagnostic_run_id,
      seed = h02_seed(diagnostic_run_id, 130L)
    ) |>
      dplyr::mutate(
        base_run_id = run$run_id,
        placement = run$placement,
        model_variant = model_variant,
        .after = "run_id"
      )
    k_rows[[diagnostic_run_id]] <- k_check
    site_constraint <- h11_activity_site_constraint(
      fit,
      data,
      run$run_id,
      run$placement,
      model_variant
    )
    site_rows[[diagnostic_run_id]] <- site_constraint
    assessment_rows[[diagnostic_run_id]] <-
      h11_activity_diagnostic_assessment(
        row,
        acf,
        closure,
        site_constraint,
        k_check,
        run$run_id,
        run$placement,
        model_variant
      )
    if (
      model_variant == "activity_adjusted" &&
        robust$comparison$p_raw < 0.05
    ) {
      effect_variation_rows[[run$run_id]] <-
        h11_curve_variation(
          fit,
          pointwise,
          diagnostic_run_id,
          context
        ) |>
        dplyr::mutate(
          base_run_id = run$run_id,
          placement = run$placement,
          model_variant = model_variant,
          participants = dplyr::n_distinct(data$participant),
          participant_days = dplyr::n_distinct(data$participant_day),
          observations_30_minute = nrow(data),
          sites = dplyr::n_distinct(data$site),
          .after = "run_id"
        )
    }
    rm(
      fit,
      context,
      robust,
      pointwise,
      closure,
      acf,
      k_check,
      site_constraint
    )
    invisible(gc())
  }
  rm(data, fit_results, preliminary)
  invisible(gc())
}

model_fit_summary <- dplyr::bind_rows(model_rows)

rho_summary <- dplyr::bind_rows(rho_rows)

comparisons <- dplyr::bind_rows(comparison_rows)

robust_diagnostics <- dplyr::bind_rows(robust_rows)

curves <- dplyr::bind_rows(curve_rows)

contrasts <- dplyr::bind_rows(contrast_rows)

residual_acf <- dplyr::bind_rows(acf_rows)

residual_summary <- dplyr::bind_rows(residual_rows)

smooth_summary <- dplyr::bind_rows(smooth_rows)

formal_concurvity <- dplyr::bind_rows(concurvity_rows)

k_check <- dplyr::bind_rows(k_rows)

curve_closure <- dplyr::bind_rows(closure_rows)

site_constraint <- dplyr::bind_rows(site_rows)

diagnostic_assessment <- dplyr::bind_rows(assessment_rows)

effect_curve_variation <- dplyr::bind_rows(effect_variation_rows)

attenuation <- h11_activity_attenuation(comparisons, contrasts)

write_activity_csv(
  model_fit_summary,
  file.path(paths$tables, "model_fit_summary.csv")
)

write_activity_csv(rho_summary, file.path(paths$tables, "rho_summary.csv"))

write_activity_csv(
  comparisons,
  file.path(paths$tables, "global_sex_curve_tests.csv")
)

write_activity_csv(
  attenuation,
  file.path(paths$tables, "same_sample_activity_attenuation.csv")
)

write_activity_csv(
  curves,
  file.path(paths$source_data, "sex_specific_curves_pointwise.csv")
)

write_activity_csv(
  contrasts,
  file.path(paths$source_data, "female_minus_male_pointwise_contrasts.csv")
)

write_activity_csv(
  robust_diagnostics,
  file.path(paths$diagnostics, "robust_inference_diagnostics.csv")
)

write_activity_csv(
  residual_acf,
  file.path(paths$diagnostics, "boundary_aware_residual_acf.csv")
)

write_activity_csv(
  residual_summary,
  file.path(paths$diagnostics, "residual_summary.csv")
)

write_activity_csv(
  smooth_summary,
  file.path(paths$diagnostics, "smooth_summary.csv")
)

write_activity_csv(
  formal_concurvity,
  file.path(paths$diagnostics, "formal_concurvity.csv")
)

write_activity_csv(
  k_check,
  file.path(paths$diagnostics, "basis_dimension_check.csv")
)

write_activity_csv(
  curve_closure,
  file.path(paths$diagnostics, "cyclic_curve_closure.csv")
)

write_activity_csv(
  site_constraint,
  file.path(paths$diagnostics, "site_sum_to_zero_constraint.csv")
)

write_activity_csv(
  diagnostic_assessment,
  file.path(paths$diagnostics, "diagnostic_assessment.csv")
)

message("Applying the specified conditional sex effect-size criterion")

effect_condition <- comparisons |>
  dplyr::filter(.data$model_variant == "activity_adjusted") |>
  dplyr::transmute(
    run_id = .data$run_id,
    placement = .data$placement,
    p_raw = .data$p_raw,
    p_adjusted_BH = .data$p_adjusted_BH,
    effect_size_condition_open = .data$p_raw < 0.05,
    criterion_rule = paste(
      "Open only when the activity-adjusted exploratory global robust raw",
      "p-value is below 0.050 within that placement"
    )
  )

write_activity_csv(
  effect_condition,
  file.path(paths$tables, "activity_adjusted_effect_size_condition.csv")
)

effect_model_rows <- list()

effect_bootstrap_rows <- list()

effect_prediction_rows <- list()

for (index in which(effect_condition$effect_size_condition_open)) {
  criterion <- effect_condition[index, , drop = FALSE]
  run <- registry[match(criterion$run_id, registry$run_id), , drop = FALSE]
  bundle <- bundles[[run$run_id]]
  data <- readRDS(bundle$frame_path)
  adjusted <- readRDS(bundle$adjusted_path)
  baseline <- h11_activity_fit_and_save(
    formula = formulas$activity_adjusted_no_sex,
    data = data,
    method = "fREML",
    rho = bundle$rho,
    model_id = "activity_adjusted_no_sex_effect_size_fREML",
    run_id = run$run_id,
    model_directory = paths$models
  )
  effect_model_rows[[run$run_id]] <- h11_activity_model_row(
    baseline,
    run$run_id,
    run$placement,
    "activity_adjusted_no_sex_effect_size_baseline"
  ) |>
    dplyr::mutate(
      inferential_role = "conditional descriptive effect-size baseline"
    )
  predictions <- tibble::tibble(
    run_id = run$run_id,
    row_index = seq_len(nrow(data)),
    response = data$response,
    baseline_prediction = as.numeric(stats::fitted(baseline$fit)),
    full_prediction = as.numeric(stats::fitted(adjusted))
  )
  prediction_path <- file.path(
    paths$source_data,
    paste0(run$run_id, "__activity_adjusted_effect_size_predictions.rds")
  )
  saveRDS(predictions, prediction_path, compress = "xz")
  resampling <- h11_bootstrap_r2(
    response = predictions$response,
    baseline_prediction = predictions$baseline_prediction,
    full_prediction = predictions$full_prediction,
    data = data,
    replicates = bootstrap_count(50L),
    seed = h02_seed(paste0(run$run_id, "__activity_adjusted"), 140L)
  )
  bootstrap_path <- file.path(
    paths$diagnostics,
    paste0(run$run_id, "__activity_adjusted_effect_size_bootstrap.rds")
  )
  saveRDS(resampling, bootstrap_path, compress = "xz")
  effect_bootstrap_rows[[run$run_id]] <- h11_effect_bootstrap_summary(
    resampling,
    run$run_id
  ) |>
    dplyr::mutate(
      placement = run$placement,
      participants = dplyr::n_distinct(data$participant),
      participant_days = dplyr::n_distinct(data$participant_day),
      observations_30_minute = nrow(data),
      sites = dplyr::n_distinct(data$site),
      .after = "run_id"
    )
  effect_prediction_rows[[run$run_id]] <- tibble::tibble(
    run_id = run$run_id,
    placement = run$placement,
    prediction_relative_path = relative_path(prediction_path),
    bootstrap_relative_path = relative_path(bootstrap_path),
    baseline_formula = h11_formula_text(formulas$activity_adjusted_no_sex),
    full_formula = h11_formula_text(formulas$activity_adjusted),
    rho = bundle$rho,
    effect_size_scope = paste(
      "Sex-block increment beyond the activity-adjusted inherited temporal",
      "baseline on the exact common sample; fixed in-sample predictions"
    )
  )
  rm(data, adjusted, baseline, predictions, resampling)
  invisible(gc())
}

effect_model_summary <- dplyr::bind_rows(effect_model_rows)

effect_size_bootstrap <- dplyr::bind_rows(effect_bootstrap_rows)

effect_prediction_export_index <- dplyr::bind_rows(effect_prediction_rows)

write_activity_csv(
  effect_model_summary,
  file.path(paths$tables, "activity_adjusted_effect_size_model_fit_summary.csv")
)

write_activity_csv(
  effect_size_bootstrap,
  file.path(paths$tables, "activity_adjusted_effect_size_bootstrap.csv")
)

write_activity_csv(
  effect_curve_variation,
  file.path(paths$tables, "activity_adjusted_effect_size_curve_variation.csv")
)

write_activity_csv(
  effect_prediction_export_index,
  file.path(paths$model_data, "activity_adjusted_effect_size_prediction_export_index.csv")
)
```


## Prepare daylight context

Generate equally site-weighted daylight context for the fitted participant-days. Daylight shading is descriptive and is not a fitted covariate.

```{r}
#| label: h11-display-context

source("scripts/hypotheses/H11/h11_reporting.R")

paths <- h11_paths(root)

h11_create_directories(paths)

registry <- h11_registry(root)

solar_context_rows <- list()

for (index in seq_len(nrow(registry))) {
  run <- registry[index, , drop = FALSE]
  frame <- readRDS(file.path(
    paths$model_data,
    paste0(run$run_id, "__frame.rds")
  ))
  base_path <- file.path(
    root,
    "results/intermediate/model_data/base",
    paste0("metrics_", run$placement, "_30_minute_context.rds")
  )
  base <- readRDS(base_path)
  fitted_days <- frame |>
    dplyr::distinct(.data$site, .data$Id, .data$local_date)
  solar <- base |>
    dplyr::select(
      "site", "Id", "local_date",
      "civil_dawn_wall_minute", "civil_dusk_wall_minute"
    ) |>
    dplyr::distinct() |>
    dplyr::semi_join(
      fitted_days,
      by = c("site", "Id", "local_date")
    ) |>
    dplyr::group_by(.data$site) |>
    dplyr::summarise(
      participant_days = dplyr::n(),
      mean_civil_dawn_hour = mean(.data$civil_dawn_wall_minute) / 60,
      mean_civil_dusk_hour = mean(.data$civil_dusk_wall_minute) / 60,
      .groups = "drop"
    ) |>
    dplyr::summarise(
      sites = dplyr::n(),
      equal_site_mean_civil_dawn_hour = mean(.data$mean_civil_dawn_hour),
      equal_site_mean_civil_dusk_hour = mean(.data$mean_civil_dusk_hour)
    ) |>
    dplyr::mutate(
      run_id = run$run_id,
      placement = run$placement,
      data_scenario_id = run$data_scenario_id,
      .before = 1L
    )
  solar_context_rows[[run$run_id]] <- solar
}

solar_context <- dplyr::bind_rows(solar_context_rows)

readr::write_csv(
  solar_context,
  file.path(paths$source_data, "equal_site_solar_context.csv"),
  na = ""
)


```


## Prepare result tables

Summarise the exact samples, global tests, decomposition, uncertainty and model qualifications in reusable tables. Each output is calculated from the preceding analysis, including the activity sensitivity.

```{r}
#| label: h11-reader-tables

verified_read <- function(relative_path) readr::read_csv(file.path(root, relative_path), show_col_types = FALSE)
activity_verified_read <- verified_read

model_data_dir <- file.path(root, "results/intermediate/model_data/H11/reader")

diagnostic_dir <- file.path(root, "results/csv/diagnostics/H11/reader")

table_dir <- file.path(root, "results/tables/H11/reader")

figure_dir <- file.path(root, "results/images/H11/reader")

source_dir <- file.path(root, "results/csv/source_data/H11/reader")

invisible(vapply(
  c(model_data_dir, diagnostic_dir, table_dir, figure_dir, source_dir),
  dir.create,
  logical(1),
  recursive = TRUE,
  showWarnings = FALSE
))

samples <- verified_read(
  "results/tables/H11/fitted/sample_counts.csv"
)

comparisons <- verified_read(
  "results/tables/H11/fitted/model_comparisons.csv"
)

parametric <- verified_read(
  "results/tables/H11/fitted/parametric_sex_component.csv"
)

segments <- verified_read(
  "results/tables/H11/fitted/pointwise_descriptive_segments.csv"
)

curve_variation <- verified_read(
  "results/tables/H11/fitted/effect_size_curve_variation.csv"
)

diagnostic_assessment <- verified_read(
  "results/csv/diagnostics/H11/fitted/diagnostic_assessment.csv"
)

robust_diagnostics <- verified_read(
  "results/csv/diagnostics/H11/fitted/robust_inference_diagnostics.csv"
)

residual_acf <- verified_read(
  "results/csv/diagnostics/H11/fitted/boundary_aware_residual_acf.csv"
)

residual_summary <- verified_read(
  "results/csv/diagnostics/H11/fitted/residual_summary.csv"
)

curves <- verified_read(
  "results/csv/source_data/H11/fitted/sex_specific_curves_pointwise.csv"
)

contrasts <- verified_read(
  "results/csv/source_data/H11/fitted/female_minus_male_pointwise_contrasts.csv"
)

solar_context <- verified_read(
  "results/csv/source_data/H11/fitted/equal_site_solar_context.csv"
)

formula_settings <- verified_read(
  "results/intermediate/model_data/H11/fitted/formula_and_fit_settings.csv"
)

paired_assessment <- verified_read(
  "results/intermediate/model_data/H11/fitted/paired_placement_display_assessment.csv"
)

activity_samples <- activity_verified_read(
  "results/tables/H11/activity_context/activity_common_sample_counts.csv"
)

activity_tests <- activity_verified_read(
  "results/tables/H11/activity_context/global_sex_curve_tests.csv"
)

activity_attenuation <- activity_verified_read(
  "results/tables/H11/activity_context/same_sample_activity_attenuation.csv"
)

activity_support <- activity_verified_read(
  "results/tables/H11/activity_context/activity_category_support_by_sex.csv"
)

activity_diagnostics <- activity_verified_read(
  "results/csv/diagnostics/H11/activity_context/diagnostic_assessment.csv"
)

activity_robust_diagnostics <- activity_verified_read(
  "results/csv/diagnostics/H11/activity_context/robust_inference_diagnostics.csv"
)

activity_contrasts <- activity_verified_read(
  paste0(
    "results/csv/source_data/H11/activity_context/",
    "female_minus_male_pointwise_contrasts.csv"
  )
)

activity_formula_settings <- activity_verified_read(
  "results/intermediate/model_data/H11/activity_context/formula_and_fit_settings.csv"
)

activity_effect_condition <- activity_verified_read(
  paste0(
    "results/tables/H11/activity_context/",
    "activity_adjusted_effect_size_condition.csv"
  )
)

run_labels <- tibble::tribble(
  ~run_id, ~dataset_label, ~placement_label, ~reader_role,
  "main__glasses__all_available", "Primary dataset", "Near eye", "Primary",
  "main__chest__all_available", "Primary dataset", "Chest", "Complementary",
  "alternative_preprocessing__glasses__all_available",
  "Gap-timing-unaware dataset", "Near eye", "Sensitivity",
  "alternative_preprocessing__chest__all_available",
  "Gap-timing-unaware dataset", "Chest", "Sensitivity"
)

with_reader_labels <- function(data) {
  data |>
    dplyr::left_join(run_labels, by = "run_id", relationship = "many-to-one") |>
    dplyr::relocate(
      dataset_label,
      placement_label,
      reader_role,
      .after = run_id
    )
}

format_clock <- function(minutes) {
  ifelse(
    minutes == 1440,
    "24:00",
    sprintf("%02d:%02d", (minutes %/% 60) %% 24, minutes %% 60)
  )
}

reader_samples <- samples |>
  with_reader_labels() |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    participants,
    female_participants,
    male_participants,
    participant_days,
    female_participant_days,
    male_participant_days,
    observations_30_minute,
    nominal_observation_hours,
    female_observations_30_minute,
    male_observations_30_minute,
    sites,
    AR_sequences
  )

reader_global <- comparisons |>
  dplyr::filter(.data$comparison_id == "complete_sex_curve") |>
  with_reader_labels() |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    estimand,
    test_method,
    covariance_method,
    F_statistic = test_statistic,
    fractional_numerator_df = test_df,
    denominator_df,
    p_raw,
    p_adjusted = p_adjusted_BH,
    observed_family_n,
    support_status
  )

reader_decomposition <- comparisons |>
  dplyr::filter(
    .data$comparison_role == "decomposition"
  ) |>
  with_reader_labels() |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    comparison_id,
    estimand,
    F_statistic = test_statistic,
    fractional_numerator_df = test_df,
    denominator_df,
    p_raw,
    p_adjusted = p_adjusted_BH,
    observed_family_n,
    support_status
  )

reader_parametric <- parametric |>
  dplyr::filter(startsWith(.data$run_id, "main__")) |>
  with_reader_labels() |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    estimand,
    estimate_eta,
    standard_error_participant_cluster_robust,
    lower_eta_95,
    upper_eta_95,
    female_to_male_shifted_ratio = shifted_ratio,
    ratio_lower_95 = shifted_ratio_lower_95,
    ratio_upper_95 = shifted_ratio_upper_95,
    interval_method
  )

reader_curve_variation <- curve_variation |>
  with_reader_labels() |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    participants,
    participant_days,
    observations_30_minute,
    sites,
    effect_size_id,
    definition,
    estimate,
    lower_95,
    upper_95,
    standard_error_participant_cluster_robust,
    unit,
    confidence_interval_method,
    inferential_role
  )

reader_pointwise_context <- contrasts |>
  with_reader_labels() |>
  dplyr::group_by(
    .data$run_id,
    .data$dataset_label,
    .data$placement_label,
    .data$reader_role
  ) |>
  dplyr::summarise(
    displayed_bins = dplyr::n(),
    pointwise_bins_excluding_one = sum(
      .data$pointwise_direction != "not_distinguishable_pointwise"
    ),
    minimum_ratio = min(.data$female_to_male_shifted_ratio),
    minimum_clock = format_clock(
      round(
        60 * .data$time_hour[
          which.min(.data$female_to_male_shifted_ratio)
        ]
      )
    ),
    minimum_ratio_lower_95 = .data$ratio_lower_pointwise_95[
      which.min(.data$female_to_male_shifted_ratio)
    ],
    minimum_ratio_upper_95 = .data$ratio_upper_pointwise_95[
      which.min(.data$female_to_male_shifted_ratio)
    ],
    maximum_ratio = max(.data$female_to_male_shifted_ratio),
    maximum_clock = format_clock(
      round(
        60 * .data$time_hour[
          which.max(.data$female_to_male_shifted_ratio)
        ]
      )
    ),
    maximum_ratio_lower_95 = .data$ratio_lower_pointwise_95[
      which.max(.data$female_to_male_shifted_ratio)
    ],
    maximum_ratio_upper_95 = .data$ratio_upper_pointwise_95[
      which.max(.data$female_to_male_shifted_ratio)
    ],
    interval_scope = dplyr::first(.data$interval_scope),
    .groups = "drop"
  )

reader_segments <- segments |>
  with_reader_labels() |>
  dplyr::mutate(
    start_local_clock = format_clock(.data$start_clock_bin),
    end_local_clock = format_clock(.data$end_clock_bin_exclusive)
  ) |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    pointwise_direction,
    start_local_clock,
    end_local_clock,
    bins_30_minute,
    minimum_ratio,
    maximum_ratio,
    minimum_pointwise_lower,
    maximum_pointwise_upper,
    interval_scope
  )

reader_diagnostic_assessment <- diagnostic_assessment |>
  with_reader_labels()

reader_robust_diagnostics <- robust_diagnostics |>
  with_reader_labels()

reader_residual_acf <- residual_acf |>
  with_reader_labels()

reader_primary_residual_summary <- residual_summary |>
  with_reader_labels() |>
  dplyr::filter(.data$dataset_label == "Primary dataset")

reader_curves <- curves |>
  with_reader_labels()

reader_contrasts <- contrasts |>
  with_reader_labels()

reader_solar_context <- solar_context |>
  with_reader_labels()

reader_formula <- formula_settings |>
  dplyr::filter(
    .data$data_scenario_id == "main",
    .data$placement == "glasses",
    .data$model_id == "mpattern_final_fREML"
  ) |>
  dplyr::distinct(
    formula_id,
    formula,
    family,
    link,
    method,
    discrete,
    rho_rule,
    knots,
    biological_sex_encoding
  )

reader_paired_assessment <- paired_assessment |>
  dplyr::transmute(
    stored_common_sample_participants = .data$stored_paired_frame_participants,
    stored_common_sample_participant_days =
      .data$stored_paired_frame_participant_days,
    stored_common_sample_observations =
      .data$stored_paired_frame_observations,
    stored_common_sample_sites = .data$stored_paired_frame_sites,
    common_sample_H11_fitted_outputs_available =
      .data$h11_paired_fitted_outputs_available,
    predeclared_scalar_temporal_estimand_available =
      .data$predeclared_scalar_temporal_estimand_available,
    scalar_identity_plot_applicable = .data$scalar_identity_plot_applicable,
    applicability_reason = .data$applicability_reason,
    closest_valid_display = .data$closest_valid_display,
    paired_source_data_created = !is.na(.data$paired_source_data_csv),
    source_data_status = .data$source_data_status
  )

activity_placement_labels <- c(
  glasses = "Near eye",
  chest = "Chest"
)

activity_model_labels <- c(
  restricted_unadjusted = "Activity-complete sample, unadjusted",
  activity_adjusted = "Same sample, activity-adjusted"
)

reader_activity_samples <- activity_samples |>
  dplyr::mutate(
    placement_label = unname(activity_placement_labels[.data$placement]),
    reader_role = dplyr::if_else(
      .data$placement == "glasses",
      "Primary near-eye sensitivity",
      "Complementary chest sensitivity"
    ),
    dataset_label = "Activity-complete sample",
    .after = "run_id"
  ) |>
  dplyr::select(
    run_id,
    dataset_label,
    placement_label,
    reader_role,
    participants,
    female_participants,
    male_participants,
    participant_days,
    female_participant_days,
    male_participant_days,
    observations_30_minute,
    nominal_observation_hours,
    female_observations_30_minute,
    male_observations_30_minute,
    sites,
    AR_sequences
  )

reader_activity_tests <- activity_tests |>
  dplyr::left_join(
    reader_activity_samples |>
      dplyr::select(
        run_id,
        participants,
        participant_days,
        observations_30_minute,
        sites
      ),
    by = "run_id",
    relationship = "many-to-one"
  ) |>
  dplyr::mutate(
    placement_label = unname(activity_placement_labels[.data$placement]),
    analysis_step = unname(activity_model_labels[.data$model_variant]),
    step_order = dplyr::if_else(
      .data$model_variant == "restricted_unadjusted",
      2L,
      3L
    )
  ) |>
  dplyr::transmute(
    run_id,
    placement_label,
    analysis_step,
    step_order,
    participants,
    participant_days,
    observations_30_minute,
    sites,
    F_statistic = .data$test_statistic,
    fractional_numerator_df = .data$test_df,
    denominator_df,
    p_raw,
    p_adjusted = .data$p_adjusted_BH,
    support_status,
    inferential_role
  )

reader_activity_original <- reader_global |>
  dplyr::filter(
    .data$dataset_label == "Primary dataset",
    .data$run_id %in% c(
      "main__glasses__all_available",
      "main__chest__all_available"
    )
  ) |>
  dplyr::left_join(
    reader_samples |>
      dplyr::select(
        run_id,
        participants,
        participant_days,
        observations_30_minute,
        sites
      ),
    by = "run_id",
    relationship = "one-to-one"
  ) |>
  dplyr::transmute(
    run_id,
    placement_label,
    analysis_step = "Primary all-available model",
    step_order = 1L,
    participants,
    participant_days,
    observations_30_minute,
    sites,
    F_statistic,
    fractional_numerator_df,
    denominator_df,
    p_raw,
    p_adjusted,
    support_status,
    inferential_role = dplyr::if_else(
      .data$placement_label == "Near eye",
      "Primary result",
      "Complementary result"
    )
  )

reader_activity_comparison <- dplyr::bind_rows(
  reader_activity_original,
  reader_activity_tests
) |>
  dplyr::arrange(
    factor(.data$placement_label, levels = c("Near eye", "Chest")),
    .data$step_order
  )

reader_activity_attenuation <- activity_attenuation |>
  dplyr::mutate(
    placement_label = unname(activity_placement_labels[.data$placement]),
    .after = "run_id"
  )

reader_activity_support <- activity_support |>
  dplyr::mutate(
    placement_label = unname(activity_placement_labels[.data$placement]),
    .after = "run_id"
  )

reader_activity_diagnostics <- activity_diagnostics |>
  dplyr::left_join(
    activity_robust_diagnostics |>
      dplyr::select(
        run_id,
        model_variant,
        maximum_participant_unscaled_meat_share,
        effective_participants_unscaled_meat_trace,
        delete_one_participant_p_minimum,
        delete_one_participant_p_maximum
      ),
    by = c("run_id", "model_variant"),
    relationship = "one-to-one"
  ) |>
  dplyr::mutate(
    placement_label = unname(activity_placement_labels[.data$placement]),
    analysis_step = unname(activity_model_labels[.data$model_variant]),
    .after = "run_id"
  )

reader_activity_contrasts <- activity_contrasts |>
  dplyr::mutate(
    placement_label = unname(activity_placement_labels[.data$placement]),
    analysis_step = unname(activity_model_labels[.data$model_variant]),
    .after = "run_id"
  )

reader_activity_pointwise_context <- reader_activity_contrasts |>
  dplyr::group_by(
    .data$run_id,
    .data$placement_label,
    .data$model_variant,
    .data$analysis_step
  ) |>
  dplyr::summarise(
    displayed_bins = dplyr::n(),
    pointwise_bins_excluding_one = sum(
      .data$pointwise_direction != "not_distinguishable_pointwise"
    ),
    minimum_ratio = min(.data$female_to_male_shifted_ratio),
    minimum_clock = format_clock(round(
      60 * .data$time_hour[which.min(.data$female_to_male_shifted_ratio)]
    )),
    minimum_ratio_lower_95 = .data$ratio_lower_pointwise_95[
      which.min(.data$female_to_male_shifted_ratio)
    ],
    minimum_ratio_upper_95 = .data$ratio_upper_pointwise_95[
      which.min(.data$female_to_male_shifted_ratio)
    ],
    maximum_ratio = max(.data$female_to_male_shifted_ratio),
    maximum_clock = format_clock(round(
      60 * .data$time_hour[which.max(.data$female_to_male_shifted_ratio)]
    )),
    maximum_ratio_lower_95 = .data$ratio_lower_pointwise_95[
      which.max(.data$female_to_male_shifted_ratio)
    ],
    maximum_ratio_upper_95 = .data$ratio_upper_pointwise_95[
      which.max(.data$female_to_male_shifted_ratio)
    ],
    interval_scope = dplyr::first(.data$interval_scope),
    .groups = "drop"
  )

reader_activity_formula <- activity_formula_settings |>
  dplyr::filter(
    .data$formula_id %in% c("restricted_unadjusted", "activity_adjusted")
  ) |>
  dplyr::mutate(
    analysis_step = unname(activity_model_labels[.data$formula_id]),
    .before = 1L
  )

write_reader(reader_samples, table_dir, "H11_reader_samples.csv")

write_reader(reader_global, table_dir, "H11_reader_global_tests.csv")

write_reader(
  reader_decomposition,
  table_dir,
  "H11_reader_level_shape_decomposition.csv"
)

write_reader(
  reader_parametric,
  table_dir,
  "H11_reader_parametric_level_estimates.csv"
)

write_reader(
  reader_curve_variation,
  table_dir,
  "H11_reader_curve_variation_effect_size.csv"
)

write_reader(
  reader_pointwise_context,
  table_dir,
  "H11_reader_pointwise_context.csv"
)

write_reader(
  reader_segments,
  table_dir,
  "H11_reader_pointwise_segments.csv"
)

write_reader(
  reader_diagnostic_assessment,
  diagnostic_dir,
  "H11_reader_diagnostic_assessment.csv"
)

write_reader(
  reader_robust_diagnostics,
  diagnostic_dir,
  "H11_reader_robust_diagnostics.csv"
)

write_reader(
  reader_residual_acf,
  diagnostic_dir,
  "H11_reader_residual_acf.csv"
)

write_reader(
  reader_primary_residual_summary,
  diagnostic_dir,
  "H11_reader_primary_residual_summary.csv"
)

write_reader(
  reader_formula,
  model_data_dir,
  "H11_reader_formula_registry.csv"
)

write_reader(
  reader_paired_assessment,
  model_data_dir,
  "H11_reader_placement_comparison_assessment.csv"
)

write_reader(
  reader_activity_formula,
  model_data_dir,
  "H11_reader_activity_formula_registry.csv"
)

write_reader(
  reader_curves,
  source_dir,
  "H11_reader_sex_specific_curves.csv"
)

write_reader(
  reader_contrasts,
  source_dir,
  "H11_reader_female_minus_male_contrasts.csv"
)

write_reader(
  reader_solar_context,
  source_dir,
  "H11_reader_solar_context.csv"
)

write_reader(
  reader_activity_samples,
  table_dir,
  "H11_reader_activity_samples.csv"
)

write_reader(
  reader_activity_comparison,
  table_dir,
  "H11_reader_activity_global_comparison.csv"
)

write_reader(
  reader_activity_attenuation,
  table_dir,
  "H11_reader_activity_attenuation.csv"
)

write_reader(
  reader_activity_support,
  table_dir,
  "H11_reader_activity_support.csv"
)

write_reader(
  reader_activity_pointwise_context,
  table_dir,
  "H11_reader_activity_pointwise_context.csv"
)

write_reader(
  reader_activity_diagnostics,
  diagnostic_dir,
  "H11_reader_activity_diagnostic_assessment.csv"
)

write_reader(
  reader_activity_contrasts,
  source_dir,
  "H11_reader_activity_female_minus_male_contrasts.csv"
)
```


## Draw daily curves and diagnostics

Draw sex-specific curves, uncertainty and residual summaries from the regenerated source tables. SVG and PNG versions are available for the manuscript and website.

```{r}
#| label: h11-figures

figure_dir <- file.path(root, "results/images/H11/reader")

dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)

read_reader <- function(...) {
  path <- file.path(root, ...)
  if (!file.exists(path)) {
    stop(paste("Missing H11 reader-display input:", path), call. = FALSE)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

curves <- read_reader(
  "results", "csv/source_data", "H11", "reader",
  "H11_reader_sex_specific_curves.csv"
)

contrasts <- read_reader(
  "results", "csv/source_data", "H11", "reader",
  "H11_reader_female_minus_male_contrasts.csv"
)

solar_context <- read_reader(
  "results", "csv/source_data", "H11", "reader",
  "H11_reader_solar_context.csv"
)

samples <- read_reader(
  "results", "tables", "H11", "reader",
  "H11_reader_samples.csv"
)

residual_acf <- read_reader(
  "results", "csv/diagnostics", "H11", "reader",
  "H11_reader_residual_acf.csv"
)

residual_summary <- read_reader(
  "results", "csv/diagnostics", "H11", "reader",
  "H11_reader_primary_residual_summary.csv"
)

activity_contrasts <- read_reader(
  "results", "csv/source_data", "H11", "reader",
  "H11_reader_activity_female_minus_male_contrasts.csv"
)

activity_samples <- read_reader(
  "results", "tables", "H11", "reader",
  "H11_reader_activity_samples.csv"
)

figure_spec <- list(
  intended_width_mm = 170,
  base_width_in = 10.5,
  curve_base_height_in = 11,
  diagnostic_base_height_in = 6.5,
  diagnostic_summary_base_height_in = 9.0,
  activity_base_height_in = 7.2,
  export_scale_multiplier = 1,
  smallest_essential_nominal_text_pt = 12,
  smallest_minor_nominal_text_pt = 11,
  raster_dpi = 300,
  a4_width_mm = 210,
  a4_height_mm = 297,
  a4_side_margin_mm = 20
)

figure_spec$export_width_in <-
  figure_spec$base_width_in * figure_spec$export_scale_multiplier

figure_spec$export_width_mm <- figure_spec$export_width_in * 25.4

figure_spec$display_reduction_factor <-
  figure_spec$intended_width_mm / figure_spec$export_width_mm

figure_spec$effective_final_text_pt <-
  figure_spec$smallest_essential_nominal_text_pt *
  figure_spec$display_reduction_factor

figure_spec$effective_final_minor_text_pt <-
  figure_spec$smallest_minor_nominal_text_pt *
  figure_spec$display_reduction_factor

sex_palette <- c(Female = "#CC79A7", Male = "#0072B2")

figure_registry <- tibble::tribble(
  ~figure_id, ~run_id, ~figure_kind, ~placement_title, ~filename,
  "primary_near_eye_curves", "main__glasses__all_available", "curve",
  "near-eye curves", "H11_reader_primary_near_eye_curves",
  "complementary_chest_curves", "main__chest__all_available", "curve",
  "chest curves", "H11_reader_complementary_chest_curves",
  "sensitivity_near_eye_curves",
  "alternative_preprocessing__glasses__all_available", "curve",
  "near-eye curves in the gap-timing-unaware dataset",
  "H11_reader_gap_timing_unaware_near_eye_curves",
  "sensitivity_chest_curves",
  "alternative_preprocessing__chest__all_available", "curve",
  "chest curves in the gap-timing-unaware dataset",
  "H11_reader_gap_timing_unaware_chest_curves",
  "activity_context_attenuation", "activity_context_common_sample", "activity",
  "activity-context sensitivity",
  "H11_reader_activity_context_female_to_male_curves",
  "primary_near_eye_residual_dependence", "main__glasses__all_available",
  "diagnostic", "near eye", "H11_reader_near_eye_residual_dependence",
  "complementary_chest_residual_dependence", "main__chest__all_available",
  "diagnostic", "chest", "H11_reader_chest_residual_dependence",
  "primary_residual_summary", "primary_common", "diagnostic_summary",
  "near eye and chest", "H11_reader_primary_residual_summary"
) |>
  dplyr::mutate(
    base_height_in = dplyr::case_when(
      .data$figure_kind == "curve" ~ figure_spec$curve_base_height_in,
      .data$figure_kind == "activity" ~ figure_spec$activity_base_height_in,
      .data$figure_kind == "diagnostic_summary" ~
        figure_spec$diagnostic_summary_base_height_in,
      TRUE ~ figure_spec$diagnostic_base_height_in
    ),
    export_height_in = .data$base_height_in *
      figure_spec$export_scale_multiplier,
    export_height_mm = .data$export_height_in * 25.4,
    intended_display_height_mm = figure_spec$intended_width_mm *
      .data$base_height_in / figure_spec$base_width_in
  )

plots <- vector("list", nrow(figure_registry))

names(plots) <- figure_registry$figure_id
for (index in seq_len(nrow(figure_registry))) {
  row <- figure_registry[index, , drop = FALSE]
  plot <- if (row$figure_kind == "curve") {
    build_curve_figure(row$run_id, row$placement_title)
  } else if (row$figure_kind == "activity") {
    build_activity_figure()
  } else if (row$figure_kind == "diagnostic_summary") {
    build_residual_summary_figure()
  } else {
    build_diagnostic_figure(row$run_id, row$placement_title)
  }
  plots[[row$figure_id]] <- plot

  png_path <- file.path(figure_dir, paste0(row$filename, ".png"))
  pdf_path <- file.path(figure_dir, paste0(row$filename, ".pdf"))
  ggplot2::ggsave(
    png_path,
    plot = plot,
    width = figure_spec$base_width_in,
    height = row$base_height_in,
    units = "in",
    scale = figure_spec$export_scale_multiplier,
    dpi = figure_spec$raster_dpi,
    bg = "white"
  )
  ggplot2::ggsave(
    pdf_path,
    plot = plot,
    width = figure_spec$base_width_in,
    height = row$base_height_in,
    units = "in",
    scale = figure_spec$export_scale_multiplier,
    device = grDevices::cairo_pdf,
    bg = "white"
  )

  svg_path <- file.path(figure_dir, paste0(row$filename, ".svg"))
  ggplot2::ggsave(svg_path, plot = plot, width = figure_spec$base_width_in, height = row$base_height_in, units = "in", scale = figure_spec$export_scale_multiplier, device = svglite::svglite, bg = "white")
}
```

## Findings and interpretation {#findings}

The following views use the models and summaries calculated above.

```{r}
#| label: h11-result-view-setup
#| code-fold: true
#| code-summary: "Export results"

suppressPackageStartupMessages({
    library(dplyr)
    library(gt)
    library(readr)
    library(tibble)
})
root <- normalizePath(Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()), winslash = "/", mustWork = TRUE)
source(file.path(root, "scripts", "pipeline", "p_value_display.R"))
read_h11 <- function(...) {
    readr::read_csv(file.path(root, ...), show_col_types = FALSE)
}
samples <- read_h11("results", "tables", "H11", "reader", "H11_reader_samples.csv")
global_tests <- read_h11("results", "tables", "H11", "reader", "H11_reader_global_tests.csv")
decomposition <- read_h11("results", "tables", "H11", "reader", "H11_reader_level_shape_decomposition.csv")
parametric_levels <- read_h11("results", "tables", "H11", "reader", "H11_reader_parametric_level_estimates.csv")
curve_effects <- read_h11("results", "tables", "H11", "reader", "H11_reader_curve_variation_effect_size.csv")
pointwise_context <- read_h11("results", "tables", "H11", "reader", "H11_reader_pointwise_context.csv")
pointwise_segments <- read_h11("results", "tables", "H11", "reader", "H11_reader_pointwise_segments.csv")
diagnostic_assessment <- read_h11("results", "csv/diagnostics", "H11", "reader", "H11_reader_diagnostic_assessment.csv")
robust_diagnostics <- read_h11("results", "csv/diagnostics", "H11", "reader", "H11_reader_robust_diagnostics.csv")
formula_registry <- read_h11("results", "intermediate/model_data", "H11", "reader", "H11_reader_formula_registry.csv")
placement_assessment <- read_h11("results", "intermediate/model_data", "H11", "reader", "H11_reader_placement_comparison_assessment.csv")
activity_formula_registry <- read_h11("results", "intermediate/model_data", "H11", "reader", "H11_reader_activity_formula_registry.csv")
activity_samples <- read_h11("results", "tables", "H11", "reader", "H11_reader_activity_samples.csv")
activity_comparison <- read_h11("results", "tables", "H11", "reader", "H11_reader_activity_global_comparison.csv")
activity_attenuation <- read_h11("results", "tables", "H11", "reader", "H11_reader_activity_attenuation.csv")
activity_pointwise_context <- read_h11("results", "tables", "H11", "reader", "H11_reader_activity_pointwise_context.csv")
activity_diagnostics <- read_h11("results", "csv/diagnostics", "H11", "reader", "H11_reader_activity_diagnostic_assessment.csv")
near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
sensitivity_near_id <- "alternative_preprocessing__glasses__all_available"
sensitivity_chest_id <- "alternative_preprocessing__chest__all_available"
near_global <- filter(global_tests, .data$run_id == .env$near_id)
chest_global <- filter(global_tests, .data$run_id == .env$chest_id)
near_context <- filter(pointwise_context, .data$run_id == .env$near_id)
chest_context <- filter(pointwise_context, .data$run_id == .env$chest_id)
sensitivity_near_context <- filter(pointwise_context, .data$run_id == .env$sensitivity_near_id)
sensitivity_chest_context <- filter(pointwise_context, .data$run_id == .env$sensitivity_chest_id)
near_activity_unadjusted <- filter(activity_comparison, .data$placement_label == "Near eye", .data$analysis_step == "Activity-complete sample, unadjusted")
near_activity_adjusted <- filter(activity_comparison, .data$placement_label == "Near eye", .data$analysis_step == "Same sample, activity-adjusted")
chest_activity_unadjusted <- filter(activity_comparison, .data$placement_label == "Chest", .data$analysis_step == "Activity-complete sample, unadjusted")
chest_activity_adjusted <- filter(activity_comparison, .data$placement_label == "Chest", .data$analysis_step == "Same sample, activity-adjusted")
near_activity_pointwise <- filter(activity_pointwise_context, .data$placement_label == "Near eye", .data$model_variant ==
    "activity_adjusted")
chest_activity_pointwise <- filter(activity_pointwise_context, .data$placement_label == "Chest", .data$model_variant == "activity_adjusted")
format_number <- function(value, digits = 3L) {
    formatC(value, digits = digits, format = "f", big.mark = ",")
}
format_p_cell <- function(value, significant) {
    display <- nh_p_value_display(value, significant = significant)
    if (display$p_bold[[1L]]) {
        paste0("**", display$p_display[[1L]], "**")
    }
    else {
        display$p_display[[1L]]
    }
}
format_p_vector <- function(value, significant) {
    vapply(seq_along(value), function(index) format_p_cell(value[[index]], significant[[index]]), character(1))
}
format_p_inline <- function(value, significant) {
    format_p_cell(value[[1L]], significant[[1L]])
}
format_ratio_ci <- function(estimate, low, high) {
    paste0(format_number(estimate), " (", format_number(low), "–", format_number(high), ")")
}
format_clock <- function(value) {
    substr(as.character(value), 1L, 5L)
}
h11_gt <- function(table, font_size = 12) {
    gt::tab_options(gt::sub_missing(gt::opt_row_striping(table), missing_text = "Not available"), table.width = gt::pct(100),
        container.overflow.x = TRUE, table.font.size = gt::px(font_size), data_row.padding = gt::px(5), column_labels.font.weight = "600",
        source_notes.font.size = gt::px(11))
}
sample_table <- function(data) {
    h11_gt(gt::cols_width(gt::fmt_integer(gt::gt(transmute(data, Placement = .data$placement_label, Role = .data$reader_role,
        Participants = .data$participants, `Female / Male participants` = paste0(.data$female_participants, " / ", .data$male_participants),
        `Participant-days` = .data$participant_days, `30-minute observations` = .data$observations_30_minute, Sites = .data$sites),
        rowname_col = "Placement"), columns = c(Participants, `Participant-days`, `30-minute observations`, Sites), use_seps = TRUE),
        Role ~ gt::pct(15), Participants ~ gt::pct(12), `Female / Male participants` ~ gt::pct(19), `Participant-days` ~
            gt::pct(15), `30-minute observations` ~ gt::pct(19), Sites ~ gt::pct(8)), 12)
}
global_test_table <- function(data) {
    h11_gt(gt::tab_source_note(gt::fmt_markdown(gt::fmt_number(gt::fmt_number(gt::gt(transmute(data, Placement = .data$placement_label,
        `F statistic` = .data$F_statistic, `Numerator df` = .data$fractional_numerator_df, `Denominator df` = .data$denominator_df,
        `Raw p` = format_p_vector(.data$p_raw, .data$p_raw < 0.05), `FDR-adjusted p` = format_p_vector(.data$p_adjusted,
            .data$p_adjusted < 0.05), Support = tools::toTitleCase(.data$support_status)), rowname_col = "Placement"), columns = c(`F statistic`,
        `Numerator df`), decimals = 3), columns = `Denominator df`, decimals = 0), columns = c(`Raw p`, `FDR-adjusted p`)),
        source_note = paste("The complete 48-bin Female-minus-Male curve is tested jointly.", "Raw p is bold at raw p < 0.050; FDR-adjusted p is bold",
            "independently at adjusted p < 0.050. The global family contains one test per", "placement and dataset, so its adjusted and raw values coincide.")),
        12)
}
principal_global_test_table <- function(test_data, sample_data) {
    placement_order <- c("Near eye", "Chest")
    primary_tests <- arrange(filter(test_data, .data$dataset_label == "Primary dataset"), match(.data$placement_label, .env$placement_order))
    primary_samples <- arrange(filter(sample_data, .data$dataset_label == "Primary dataset"), match(.data$placement_label,
        .env$placement_order))
    stopifnot(nrow(primary_tests) == 2L, nrow(primary_samples) == 2L, identical(primary_tests$placement_label, placement_order),
        identical(primary_samples$placement_label, placement_order), identical(primary_tests$run_id, primary_samples$run_id),
        identical(primary_tests$reader_role, c("Primary", "Complementary")))
    h11_gt(gt::tab_source_note(gt::tab_source_note(gt::cols_width(gt::fmt_markdown(gt::fmt_number(gt::fmt_number(gt::gt(transmute(primary_tests,
        Placement = .data$placement_label, Role = .data$reader_role, Test = "Complete Female-minus-Male 24-hour curve", `F statistic` = .data$F_statistic,
        `Numerator df` = .data$fractional_numerator_df, `Denominator df` = .data$denominator_df, `Raw p` = format_p_vector(.data$p_raw,
            .data$p_raw < 0.05), `FDR-adjusted p` = format_p_vector(.data$p_adjusted, .data$p_adjusted < 0.05), Support = tools::toTitleCase(.data$support_status)),
        rowname_col = "Placement"), columns = c(`F statistic`, `Numerator df`), decimals = 3), columns = `Denominator df`,
        decimals = 0), columns = c(`Raw p`, `FDR-adjusted p`)), Role ~ gt::pct(11), Test ~ gt::pct(23), `F statistic` ~ gt::pct(9),
        `Numerator df` ~ gt::pct(10), `Denominator df` ~ gt::pct(11), `Raw p` ~ gt::pct(8), `FDR-adjusted p` ~ gt::pct(11),
        Support ~ gt::pct(9)), source_note = paste0("Primary near-eye fitted sample: ", primary_samples$participants[[1L]],
        " participants, ", format(primary_samples$participant_days[[1L]], big.mark = ","), " participant-days, ", format(primary_samples$observations_30_minute[[1L]],
            big.mark = ","), " 30-minute observations, and ", primary_samples$sites[[1L]], " sites. Complementary chest fitted sample: ",
        primary_samples$participants[[2L]], " participants, ", format(primary_samples$participant_days[[2L]], big.mark = ","),
        " participant-days, ", format(primary_samples$observations_30_minute[[2L]], big.mark = ","), " 30-minute observations, and ",
        primary_samples$sites[[2L]], " sites.")), source_note = paste("Each row is one joint complete-curve test. Raw p is bold at raw p <",
        "0.050; FDR-adjusted p is bold independently at adjusted p < 0.050.", "Each labelled global family contains one test, so the stored raw and",
        "adjusted values coincide.")), 12)
}
decomposition_table <- function(data) {
    h11_gt(gt::tab_source_note(gt::fmt_markdown(gt::fmt_number(gt::gt(arrange(transmute(mutate(data, Component = recode(.data$comparison_id,
        parametric_level_component = "Level", cyclic_shape_component = "Shape")), Placement = .data$placement_label, Component = .data$Component,
        `F statistic` = .data$F_statistic, `Numerator df` = .data$fractional_numerator_df, `Raw p` = format_p_vector(.data$p_raw,
            .data$p_raw < 0.05), `FDR-adjusted p` = format_p_vector(.data$p_adjusted, .data$p_adjusted < 0.05), Conclusion = if_else(.data$p_adjusted <
            0.05, "Supported", "Not supported")), .data$Placement, factor(.data$Component, levels = c("Level", "Shape"))),
        rowname_col = "Component", groupname_col = "Placement"), columns = c(`F statistic`, `Numerator df`), decimals = 3),
        columns = c(`Raw p`, `FDR-adjusted p`)), source_note = paste("Within each placement and dataset, level and shape form a two-test",
        "false-discovery-rate family. Raw and adjusted p-values are labelled", "and bolded independently at their respective p < 0.050 rules.",
        "These secondary tests ask whether the globally tested curve can be", "resolved into an independently supported level or shape component;",
        "they do not retest or invalidate the separate global curve result.")), 12)
}
effect_size_table <- function(placement) {
    level <- filter(parametric_levels, .data$placement_label == .env$placement)
    variation <- filter(curve_effects, .data$placement_label == .env$placement)
    stopifnot(nrow(level) == 1L, nrow(variation) == 1L)
    h11_gt(gt::tab_source_note(gt::cols_width(gt::gt(tibble::tibble(Estimator = c("Conditional shifted-value ratio: Female / Male",
        "Across-clock fitted-curve variation"), `Estimate (95% CI)` = c(format_ratio_ci(level$female_to_male_shifted_ratio,
        level$ratio_lower_95, level$ratio_upper_95), paste0(format_number(variation$estimate, 4), " (", format_number(variation$lower_95,
        4), "–", format_number(variation$upper_95, 4), ")")), Unit = c("Shifted-value melEDI ratio", "Squared log10(melEDI + 0.1) units"),
        Interpretation = c(paste("Time-constant sex component conditional on the cyclic", "sex-specific deviation"), paste("Mean squared Female-minus-Male fitted-curve contrast across",
            "48 equally weighted clock bins"))), rowname_col = "Estimator"), `Estimate (95% CI)` ~ gt::pct(23), Unit ~ gt::pct(24),
        Interpretation ~ gt::pct(38)), source_note = paste("Intervals are participant-cluster robust. Fitted-curve variation",
        "is an effect-size description, not a percentage of outcome", "variation explained.")), 12)
}
diagnostic_table <- function() {
    h11_gt(gt::tab_source_note(gt::cols_label(gt::fmt_number(gt::fmt_number(gt::fmt_number(gt::gt(transmute(left_join(diagnostic_assessment,
        select(robust_diagnostics, run_id, maximum_participant_unscaled_meat_share, effective_participants_unscaled_meat_trace),
        by = "run_id", relationship = "one-to-one"), Dataset = .data$dataset_label, Placement = .data$placement_label, Assessment = tools::toTitleCase(.data$classification),
        `Final lag-1 correlation` = .data$final_boundary_aware_lag1_correlation, `Largest participant share` = 100 * .data$maximum_participant_unscaled_meat_share,
        `Effective participants` = .data$effective_participants_unscaled_meat_trace), rowname_col = "Placement", groupname_col = "Dataset"),
        columns = `Final lag-1 correlation`, decimals = 3), columns = `Largest participant share`, decimals = 1, suffixing = FALSE),
        columns = `Effective participants`, decimals = 1), `Largest participant share` = "Largest participant share (%)"),
        source_note = paste("Participant share is the largest contribution to the unscaled", "participant-cluster covariance meat trace. Effective participants",
            "is the corresponding trace-based concentration summary.")), 12)
}
activity_comparison_table <- function() {
    h11_gt(gt::tab_source_note(gt::fmt_markdown(gt::fmt_number(gt::fmt_number(gt::fmt_integer(gt::gt(transmute(activity_comparison,
        Placement = .data$placement_label, Analysis = .data$analysis_step, Participants = .data$participants, `Participant-days` = .data$participant_days,
        `30-minute observations` = .data$observations_30_minute, `F statistic` = .data$F_statistic, `Numerator df` = .data$fractional_numerator_df,
        `Denominator df` = .data$denominator_df, `Raw p` = format_p_vector(.data$p_raw, .data$p_raw < 0.05), `FDR-adjusted p` = format_p_vector(.data$p_adjusted,
            .data$p_adjusted < 0.05), Conclusion = if_else(.data$p_adjusted < 0.05, "Supported in this model", "Not supported in this model")),
        rowname_col = "Analysis", groupname_col = "Placement"), columns = c(Participants, `Participant-days`, `30-minute observations`),
        use_seps = TRUE), columns = c(`F statistic`, `Numerator df`), decimals = 3), columns = `Denominator df`, decimals = 0),
        columns = c(`Raw p`, `FDR-adjusted p`)), source_note = paste("The two activity-complete models use the same observations within",
        "placement; the retained all-available model uses the larger retained", "sample. Each row is a one-test family, so raw and FDR-adjusted p-values",
        "coincide. Raw and adjusted values are bolded independently at their", "respective p < 0.050 rules. The activity comparison is exploratory.")),
        12)
}
activity_diagnostic_table <- function() {
    h11_gt(gt::tab_source_note(gt::fmt_number(gt::fmt_number(gt::fmt_number(gt::gt(transmute(activity_diagnostics, Placement = .data$placement_label,
        Analysis = .data$analysis_step, Assessment = tools::toTitleCase(.data$classification), `Final lag-1 correlation` = .data$final_boundary_aware_lag1_correlation,
        `Largest participant share (%)` = 100 * .data$maximum_participant_unscaled_meat_share, `Effective participants` = .data$effective_participants_unscaled_meat_trace),
        rowname_col = "Analysis", groupname_col = "Placement"), columns = `Final lag-1 correlation`, decimals = 3), columns = `Largest participant share (%)`,
        decimals = 1, suffixing = FALSE), columns = `Effective participants`, decimals = 1), source_note = paste("Participant share and effective-participant summaries describe",
        "concentration of the participant-cluster robust covariance.")), 12)
}
```


## Question

The preregistered hypothesis was:

> “H11: Diurnal exposure patterns differ by sex.”

The recorded predictor is **biological sex**, coded Female or Male. Gender
identity is a distinct construct and was not analysed. The analytical question
is whether the fitted 24-hour personal melanopic equivalent daylight
illuminance (melEDI) profile differs between participants recorded as Female
and Male after accounting for site-specific temporal patterns, repeated
observations within participants and days, and short-lag residual dependence.

::: {.callout-tip title="Answer in brief"}
After false-discovery-rate (FDR) adjustment, the primary near-eye
Female-minus-Male curve differed globally (adjusted *p* =
`r format_p_inline(near_global$p_adjusted, near_global$p_adjusted < 0.05)`), and
the complementary chest curve also differed globally (adjusted *p* =
`r format_p_inline(chest_global$p_adjusted,
chest_global$p_adjusted < 0.05)`). The secondary level/shape analysis could not
cleanly assign either globally supported curve to an independently resolved
level or shape component; this does not overturn the global results. In the
exploratory activity-complete sample, near eye was already unsupported before
activity adjustment (raw *p* = `r format_p_inline(near_activity_unadjusted$p_raw,
near_activity_unadjusted$p_raw < 0.05)`) and remained unsupported after it
(raw *p* = `r format_p_inline(near_activity_adjusted$p_raw,
near_activity_adjusted$p_raw < 0.05)`). At chest, the corresponding raw *p*-value
changed from `r format_p_inline(chest_activity_unadjusted$p_raw,
chest_activity_unadjusted$p_raw < 0.05)` to
`r format_p_inline(chest_activity_adjusted$p_raw,
chest_activity_adjusted$p_raw < 0.05)`. This sensitivity is compatible with
attenuation but cannot identify activity as a mechanism. Model checks classified
all retained fits as acceptable with specified limitations.
:::

## What was analysed

The near-eye sensor position is primary because it measures the light field
closer to the eyes. The chest sensor position is analysed separately as
complementary evidence and is not a measure of ocular exposure. The placements
are neither pooled nor treated as independent replicates. The response is the
30-minute arithmetic mean melEDI, modelled as `log10(melEDI + 0.1)`.

The estimand is the complete fitted Female-minus-Male 24-hour curve at equal
clock-time weighting. Its site-average fitted mean averages across included
sites and gives each site equal weight. The participant is the independent
unit for robust inference. A participant-day is one participant's included
observations on one local calendar day.

```{r}
#| label: tbl-h11-primary-samples
#| tbl-cap: "Exact fitted samples for the primary near-eye and complementary chest analyses."

samples |>
  filter(.data$dataset_label == "Primary dataset") |>
  sample_table()
```

The primary near-eye model contains 141 participants (79 Female, 62 Male), 816
participant-days, 37,756 30-minute observations, and all nine sites. The chest
model contains 154 participants (86 Female, 68 Male), 902 participant-days,
41,842 observations, and eight sites. These are the retained temporal-model
samples; no extra H11-specific row deletion was applied.

::: {.callout-note title="Key terms"}
- **Site-average fitted mean:** the fitted mean averaged across included sites,
  with each site given equal weight.
- **Complete-curve test:** one joint test of the time-constant and time-varying
  Female-minus-Male contributions across the whole 24-hour curve.
- **Pointwise 95% confidence interval (95% CI):** Confidence intervals apply at the displayed clock times.
  These are pointwise, not simultaneous bands across the full day, and do not
  support a significant-period claim.
- **Shifted-value ratio:** both fitted melEDI values are shifted by the model's
  0.1-lx offset before the Female-to-Male ratio is calculated.
- **AR(1):** a representation of autocorrelation in which residual similarity
  is stronger for measurements closer together in time.
- **Symlog:** a display scale that is linear from 0 to 1 lx and base-10
  logarithmic above 1 lx, with labels retained in original units.
:::

### Model and inference

The nonlinear generalized additive model (GAM) allows the daily melEDI pattern
to vary smoothly and nonlinearly over clock time. Its cyclic common daily curve
and cyclic Female-specific deviation join smoothly across midnight. The model
also includes site-specific temporal deviations, participant-specific time
curves, and participant-day random effects. The latter two represent remaining
repeated-measure variation after the shared time, biological-sex, and site
structure is considered.

Autocorrelation is residual similarity between measurements close together in
time. An AR(1) structure represents this dependence as stronger at shorter
lags. Sequence starts prevent that correlation from crossing participant-day
or gap boundaries. The estimation and robust-covariance calculations appear
in [Robust tests, intervals and model diagnostics](#robust-tests-intervals-and-model-diagnostics).

The following evaluated R cell reads and prints the exact Wilkinson formula
supplied to every retained H11 full model. Here, `response` is the transformed
melEDI outcome and Male is the reference level.

```{r}
#| label: h11-exact-model-formula
#| echo: true
#| code-fold: false

h11_formula <- stats::as.formula(formula_registry$formula[[1L]])
h11_formula
```

The complete-curve, or global, test jointly evaluates the time-constant and
time-varying contributions to the 48-bin Female-minus-Male curve. It produces
one decision per placement and dataset. FDR adjustment is applied within
separately labelled test families. Each global family has
one test. The planned level and shape attribution tests form a separate
two-test FDR family within each placement. Those secondary tests do not repeat
the global test, and failure to resolve either component does not invalidate a
supported global curve.

All curve and contrast intervals are participant-cluster-robust pointwise 95%
CIs. Confidence intervals apply at the displayed clock times. They are not
simultaneous and do not support a significant-period claim.

```{r}
#| label: tbl-h11-global-tests
#| tbl-cap: "Global complete-curve tests and separately prespecified sensitivity decisions: primary near-eye and complementary chest analyses."

principal_global_test_table(global_tests, samples)
```

Both complete curves were supported. For near eye, raw *p* and FDR-adjusted
*p* were `r format_p_inline(near_global$p_raw, near_global$p_raw < 0.05)`.
For chest, raw *p* and FDR-adjusted *p* were
`r format_p_inline(chest_global$p_raw, chest_global$p_raw < 0.05)`. Within each
one-test global family, the stored raw and adjusted values are identical.

## Primary near-eye result

```{r}
#| label: fig-h11-near-eye-curves
#| fig-alt: "A two-panel near-eye figure. The upper panel shows Female and Male fitted melEDI curves across 24 hours with pink and blue lines and pointwise confidence ribbons on a symlog axis that is linear from zero to one lux and logarithmic above one lux; both rise in the morning and decline at night, with the Female curve lower around morning and late afternoon. The lower panel shows the Female-to-Male shifted-value ratio with a horizontal null line at one; the deepest ratio is about 0.47 near 08:45. Intervals are pointwise, not simultaneous."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_primary_near_eye_curves.png"
))
```

The minimum displayed ratio was
`r format_ratio_ci(near_context$minimum_ratio,
near_context$minimum_ratio_lower_95,
near_context$minimum_ratio_upper_95)` at
`r format_clock(near_context$minimum_clock)`. The maximum was
`r format_ratio_ci(near_context$maximum_ratio,
near_context$maximum_ratio_lower_95,
near_context$maximum_ratio_upper_95)` at
`r format_clock(near_context$maximum_clock)`. Pointwise intervals excluded one
in the displayed 30-minute bins from 07:00–10:00 and 16:00–18:30, with Female
lower. These clock ranges are descriptive collections of pointwise intervals;
they are not a familywise significant period.

### Conditional effect-size context

Because the global near-eye test was supported, the conditional effect-size
estimators are reported.

```{r}
#| label: tbl-h11-near-effect-size
#| tbl-cap: "Primary near-eye biological-sex effect-size estimators."

effect_size_table("Near eye")
```

The conditional level ratio is compatible with both a moderate lower Female
level and a negligible difference. The fitted-curve variation is small on the
model scale despite a statistically supported structured curve difference.

## Complementary chest result

```{r}
#| label: fig-h11-chest-curves
#| fig-alt: "A two-panel chest figure. The upper panel shows Female and Male fitted melEDI curves across 24 hours with pointwise ribbons on a symlog axis that is linear from zero to one lux and logarithmic above one lux. The lower panel shows an almost horizontal Female-to-Male shifted-value ratio near 0.79, below the null line at one; every displayed pointwise interval excludes one, but the intervals are not simultaneous."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_complementary_chest_curves.png"
))
```

The chest ratio was effectively time-constant at
`r format_ratio_ci(chest_context$minimum_ratio,
chest_context$minimum_ratio_lower_95,
chest_context$minimum_ratio_upper_95)`. All 48 displayed pointwise intervals
excluded one. Because the intervals are pointwise, this observation does not
establish a simultaneous all-day difference.

### Conditional effect-size context

```{r}
#| label: tbl-h11-chest-effect-size
#| tbl-cap: "Complementary chest biological-sex effect-size estimators."

effect_size_table("Chest")
```

The chest shifted-value ratio is visually almost time-constant. Its conditional
interpretation remains subordinate to the complete-curve result because the
model also contains the cyclic Female-minus-Male deviation.

## Can the global results be separated into level and shape?

These secondary attribution tests are subordinate to the complete-curve tests
in @tbl-h11-global-tests. They ask whether each globally supported curve can be
assigned cleanly to an independently supported level or shape component.

### Primary near eye

```{r}
#| label: tbl-h11-near-decomposition
#| tbl-cap: "Primary near-eye level and shape attribution."

decomposition |>
  filter(
    .data$run_id == .env$near_id,
    .data$dataset_label == "Primary dataset"
  ) |>
  decomposition_table()
```

The secondary attribution did not cleanly resolve the globally supported curve
into an independently supported level or shape component. The shape component
had raw *p* = **0.043** but FDR-adjusted *p* = 0.075; the level component had
raw and adjusted *p* = 0.075. The global result is not formed by adding or
combining these two *p*-values. Evidence can be distributed across correlated
level and shape directions, so this result expresses uncertainty about
attribution and does not make the global result invalid.

### Complementary chest

```{r}
#| label: tbl-h11-chest-decomposition
#| tbl-cap: "Complementary chest level and shape attribution."

decomposition |>
  filter(
    .data$run_id == .env$chest_id,
    .data$dataset_label == "Primary dataset"
  ) |>
  decomposition_table()
```

The chest global result is visually level-like, but the secondary attribution
did not resolve that result independently to the level term: its raw *p* =
**0.029** and FDR-adjusted *p* = 0.057. The shape component had raw and adjusted
*p* = 0.986. This limits component attribution and does not overturn the
separately supported complete-curve result.

The all-available near-eye and chest models use different samples, so plotting
their estimates against one another would confound placement with sample
composition. A stored common frame contains 112 participants, 643 participant-
days, 29,786 matched observations, and eight sites, but no retained H11 model
was fitted to that frame and no scalar temporal estimand was predeclared.
Therefore a matched identity scatterplot is not applicable. The separate
near-eye and chest curves above are the closest valid comparison; they do not
test placement differences or establish equivalence.

## Exploratory activity-context sensitivity

The retained global test remains unadjusted for activity. An exploratory
sensitivity asked whether the estimated Female-minus-Male curve changed after
adding the recorded hourly activity context. To separate adjustment from sample
selection, the retained H11 model was first refitted without activity on the
exact activity-complete sample and then refitted with activity on those same
observations.

Diary hours were eligible only when exactly one activity was selected. Hours
with multiple selections, no selection, all missing responses, or only an
unspecified “other” activity were excluded rather than duplicated across
categories. The retained contexts were home (reference), sleep, road or
vehicle, indoor work, and outdoor activity. A 30-minute exposure observation
was attached only when it lay wholly within one unique diary hour. This
restriction produced the following exact samples.

```{r}
#| label: tbl-h11-activity-samples
#| tbl-cap: "Exact fitted samples for the exploratory activity-context sensitivity."

sample_table(activity_samples)
```

The near-eye activity-complete sample contains 126 participants (71 Female, 55
Male), 724 participant-days, 30,499 30-minute observations, and nine sites. The
chest sample contains 150 participants (83 Female, 67 Male), 875 participant-
days, 36,711 observations, and eight sites.

The following evaluated R cell prints the exact Wilkinson formulas used for
the same-sample comparison. The adjusted model adds both the activity level and
activity-specific cyclic temporal deviations. A single placement-specific
AR(1) correlation parameter estimated from the preliminary adjusted model was
held fixed in both final fits so that the comparison did not change its
autocorrelation assumption.

```{r}
#| label: h11-activity-exact-model-formulas
#| echo: true
#| code-fold: false

activity_h11_formulas <- stats::setNames(
  lapply(activity_formula_registry$formula, stats::as.formula),
  activity_formula_registry$analysis_step
)
activity_h11_formulas
```

```{r}
#| label: tbl-h11-activity-global-comparison
#| tbl-cap: "Primary and exploratory activity-context global complete-curve tests."

activity_comparison_table()
```

For near eye, restricting the data to observations with eligible activity
information was already sufficient to change the global result from the
retained all-available raw *p* =
`r format_p_inline(near_global$p_raw, near_global$p_raw < 0.05)` to raw *p* =
`r format_p_inline(near_activity_unadjusted$p_raw,
near_activity_unadjusted$p_raw < 0.05)` before activity was added. The
activity-adjusted result was raw *p* =
`r format_p_inline(near_activity_adjusted$p_raw,
near_activity_adjusted$p_raw < 0.05)`. Therefore the near-eye comparison cannot
attribute the loss of threshold support to activity adjustment; sample
restriction precedes it.

At chest, the exact same-sample raw *p*-value changed from
`r format_p_inline(chest_activity_unadjusted$p_raw,
chest_activity_unadjusted$p_raw < 0.05)` without activity to
`r format_p_inline(chest_activity_adjusted$p_raw,
chest_activity_adjusted$p_raw < 0.05)` with activity. The adjusted value is
displayed as FDR-adjusted *p* = 0.050. This small threshold crossing is
compatible with attenuation, but it is not evidence that activity mediates or
explains the observed sex association.

```{r}
#| label: fig-h11-activity-context
#| fig-cap: "Exploratory activity-context sensitivity of the Female-to-Male shifted-value melEDI ratio. Solid blue curves show the unadjusted H11 model on the activity-complete sample; dashed orange curves show the activity-adjusted model on the identical observations. Ribbons are participant-cluster-robust pointwise 95% confidence intervals, not simultaneous bands. The fitted samples contain 126 near-eye participants with 30,499 observations and 150 chest participants with 36,711 observations."
#| fig-alt: "Two stacked panels compare same-sample unadjusted and activity-adjusted Female-to-Male ratio curves. Near eye shows a morning trough that becomes shallower after adjustment, with both wide pointwise ribbons crossing one. Chest ratios are nearly horizontal near 0.81 before and 0.83 after adjustment; the adjusted pointwise interval reaches the null ratio of one."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_activity_context_female_to_male_curves.png"
))
```

The activity-adjusted near-eye minimum ratio was
`r format_ratio_ci(near_activity_pointwise$minimum_ratio,
near_activity_pointwise$minimum_ratio_lower_95,
near_activity_pointwise$minimum_ratio_upper_95)` at
`r format_clock(near_activity_pointwise$minimum_clock)`. The activity-adjusted
chest ratio was approximately
`r format_ratio_ci(chest_activity_pointwise$minimum_ratio,
chest_activity_pointwise$minimum_ratio_lower_95,
chest_activity_pointwise$minimum_ratio_upper_95)`; its exact upper confidence
limit was 1.000034. No displayed activity-adjusted pointwise interval excluded
one at either placement. These intervals remain pointwise and do not define a
simultaneously significant period.

An additional activity-adjusted sex effect-size estimator was required only if
the adjusted global sex curve had raw *p* < 0.050. That condition was not met at
either placement, so no activity-adjusted effect-size estimate is reported.

```{r}
#| label: tbl-h11-activity-diagnostics
#| tbl-cap: "Model checks for the exploratory same-sample activity-context models."

activity_diagnostic_table()
```

All four same-sample fits were **acceptable with specified limitations**. They
converged without serious warnings, retained cyclic closure and the inherited
site constraint, and had no severe model-flexibility flag. Final boundary-aware
lag-1 residual correlations were 0.144 and 0.084 near eye and 0.131 and 0.075
at chest for the unadjusted and adjusted fits, respectively. The sensitivity
is limited to uniquely mapped single-activity diary hours and remains
observational; neither adjustment nor a change around a significance threshold
identifies a behavioural mechanism.

## Data-preparation sensitivity

A predefined sensitivity uses the **gap-timing-unaware dataset**. It still
passed the general 50%-per-hour and 80%-per-day coverage rules. “Gap-timing-
unaware” means that the timing of the remaining missing observations is not
used for metric-specific adjustment; gaps, missingness, and coverage were not
ignored. For this contrast only, the primary dataset could be interpreted as a
**time-sensitive primary metric dataset**. Below, it is called simply the
primary dataset.

```{r}
#| label: tbl-h11-sensitivity-samples
#| tbl-cap: "Exact fitted samples for the gap-timing-unaware sensitivity."

samples |>
  filter(.data$reader_role == "Sensitivity") |>
  sample_table()
```

The sensitivity retains 141 near-eye participants and 154 chest participants,
but contains 809 and 894 participant-days and 37,603 and 41,664 observations,
respectively.

```{r}
#| label: tbl-h11-sensitivity-global
#| tbl-cap: "Global complete-curve tests and separately prespecified sensitivity decisions: gap-timing-unaware dataset."

global_tests |>
  filter(.data$reader_role == "Sensitivity") |>
  global_test_table()
```

Near eye remained supported (raw and FDR-adjusted *p* = **0.026**) and chest
remained supported (raw and FDR-adjusted *p* = **0.037**).

```{r}
#| label: tbl-h11-sensitivity-decomposition
#| tbl-cap: "Level and shape decompositions in the gap-timing-unaware dataset."

decomposition |>
  filter(.data$reader_role == "Sensitivity") |>
  decomposition_table()
```

The secondary sensitivity decompositions likewise did not resolve the globally
supported curves into independently supported level or shape components after
FDR adjustment. This does not alter either sensitivity global result.

::: {.panel-tabset}

### Near eye

```{r}
#| label: fig-h11-sensitivity-near-eye
#| fig-alt: "A two-panel sensitivity figure for near-eye melEDI. Female and Male daily curves resemble those in the primary dataset on a symlog axis that is linear from zero to one lux and logarithmic above one lux. The Female-to-Male ratio reaches about 0.48 near 08:45, and pointwise intervals exclude one in morning and late-afternoon bins; these are not simultaneous intervals."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_gap_timing_unaware_near_eye_curves.png"
))
```

The minimum ratio was
`r format_ratio_ci(sensitivity_near_context$minimum_ratio,
sensitivity_near_context$minimum_ratio_lower_95,
sensitivity_near_context$minimum_ratio_upper_95)` at
`r format_clock(sensitivity_near_context$minimum_clock)`. The same displayed
pointwise clock ranges excluded one as in the primary dataset.

### Chest

```{r}
#| label: fig-h11-sensitivity-chest
#| fig-alt: "A two-panel sensitivity figure for chest melEDI. Female and Male curves have similar shape on a symlog axis that is linear from zero to one lux and logarithmic above one lux, while the Female-to-Male ratio is nearly constant at about 0.80. All displayed pointwise intervals exclude one, but they are not simultaneous."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_gap_timing_unaware_chest_curves.png"
))
```

The chest ratio was approximately
`r format_ratio_ci(sensitivity_chest_context$minimum_ratio,
sensitivity_chest_context$minimum_ratio_lower_95,
sensitivity_chest_context$minimum_ratio_upper_95)`, with all 48 pointwise
intervals excluding one. Again, this is not a simultaneous all-day claim.

:::

The central conclusion is stable to this preparation change: the complete
near-eye and chest curves remain supported. The secondary level/shape analyses
do not cleanly attribute either global result to one independently resolved
component.

## Model checks

The following model checks (model diagnostics) assess fit quality and the
remaining limitations of the retained models.

```{r}
#| label: tbl-h11-diagnostic-assessment
#| tbl-cap: "Model-check assessment for every fitted H11 model."

diagnostic_table()
```

All four models are **acceptable with specified limitations**. They converged
without serious fitting warnings; the cyclic sex curves close at midnight;
site deviations satisfy their sum-to-zero constraint; and no severe basis-
dimension flag was found. The retained AR structure reduced boundary-aware
lag-1 residual correlation from approximately 0.60–0.62 before adjustment to
0.070–0.083 in the final residuals.

::: {.panel-tabset}

### Near eye

```{r}
#| label: fig-h11-near-eye-residual-dependence
#| fig-cap: "Boundary-aware near-eye residual autocorrelation before and after the retained AR(1) adjustment. Correlations use only pairs within valid participant-day sequences."
#| fig-alt: "A near-eye autocorrelation plot compares the initial residual series with the final AR-adjusted series over six hours. Initial lag-one correlation is about 0.62 and declines across lags; final correlations are close to zero after lag one, whose value is about 0.08."
#| out-width: "95%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_near_eye_residual_dependence.png"
))
```

### Chest

```{r}
#| label: fig-h11-chest-residual-dependence
#| fig-cap: "Boundary-aware chest residual autocorrelation before and after the retained AR(1) adjustment. Correlations use only pairs within valid participant-day sequences."
#| fig-alt: "A chest autocorrelation plot compares the initial residual series with the final AR-adjusted series over six hours. Initial lag-one correlation is about 0.61 and declines across lags; final correlations are near zero after lag one, whose value is about 0.07."
#| out-width: "95%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_chest_residual_dependence.png"
))
```

:::

```{r}
#| label: fig-h11-primary-residual-summary
#| fig-cap: "Residual-distribution and scale diagnostics for the primary near-eye and complementary chest models. Panel A shows stored final standardized-residual quantiles overall and by recorded biological sex: thin lines span the 1st–99th percentiles, medium lines the 5th–95th percentiles, thick lines the interquartile range, and points the median. Panel B shows the stored correlation between absolute residual magnitude and the fitted value. Medians are close to zero, while broad tails and positive correlations of 0.197–0.240 show remaining heavy tails and heteroscedasticity."
#| fig-alt: "A two-part diagnostic figure for primary near-eye and chest models. Residual quantile intervals are centered close to zero overall and for Female and Male observations, but extend to about minus two and plus two at the first and ninety-ninth percentiles. A second panel shows positive correlations, around 0.20 to 0.24, between absolute residuals and fitted values for each placement and sex group, indicating residual scale dependence."
#| out-width: "100%"
#| fig-align: center

include_project_graphics(file.path(
  root, "results", "images", "H11", "reader",
  "H11_reader_primary_residual_summary.png"
))
```

[Residual-distribution diagnostic source data](../results/csv/diagnostics/H11/reader/H11_reader_primary_residual_summary.csv)

Together, the residual figures show the main continuous diagnostic outcomes:
the retained AR(1) structure removes most short-lag dependence, while the
final residuals remain centered near zero but retain heavy tails and
fitted-value-dependent spread. Convergence, cyclic closure, the site
constraint, and basis-dimension flags remain in the diagnostic table because
they are pass/fail or count outcomes rather than quantities that benefit from
a separate plot.

No single participant dominates the robust covariance: the largest
participant contribution is 2.8–3.0% near eye and 7.5–7.6% at chest. The
coefficient-held-fixed delete-one-participant covariance check retains the
global decision for every participant. This check probes covariance influence;
it is not a leave-one-participant-out model-refit sensitivity.

The remaining limitations are material but do not invalidate the specified
test. Exact zeros, heavy tails, heteroscedasticity, and concurvity remain under
the Gaussian log-scale mean model. Results are conditional on the observed
sites; the site sample is too small for broad site-population generalisation,
and leave-one-site-out refits remain unresolved. The registered hourly
geometric-mean outcome remains the documented preregistered specification but
is outside the retained H11 analysis rather than an outstanding sensitivity. A
fitted common-placement sample, an all-zero-inclusive preparation, and
alternative temporal model forms remain unresolved. The activity-context
sensitivity is now reported, but it is limited to uniquely mapped single-
activity diary hours and cannot distinguish covariate adjustment from selection
into the activity-complete sample without the explicit same-sample comparison
shown above.

## Interpretation

These data support a biological-sex difference in the complete fitted daily
melEDI profile at near-eye level, with complementary support at chest level.
Near-eye pointwise estimates indicate the largest relative separation around
the morning, whereas the chest contrast is nearly time-constant. However, the
planned component analysis cannot attribute the supported joint result
specifically to a level or shape component after multiplicity adjustment. That
attribution limitation does not weaken the definition or validity of the
separate global curve test.

Activity adjustment did not yield a supported adjusted global curve on the
activity-complete sample at either placement. For near eye, however, the
same-sample unadjusted curve was already unsupported, so the change cannot be
assigned to activity. For chest, the curve magnitude changed modestly and the
raw *p*-value moved just above 0.050. Neither pattern establishes mediation or
a causal activity pathway.

The effect-size summaries are modest on the fitted model scale, and all clock-
specific confidence intervals are pointwise. The result is therefore evidence
for an overall structured association, not proof of a significant clock period,
not an equivalence or placement-difference analysis, and not evidence that
biological sex causally determines light exposure. Differences may reflect
behavioural, occupational, environmental, or other correlated factors that
this observational model does not identify.

## Preregistration deviations {#h11-preregistration-deviations}

The analysis differs from, or makes operationally explicit, the preregistered
plan in the following ways. The preregistered hourly geometric-mean outcome
is replaced by thirty-minute arithmetic-mean melEDI, transformed as
log10(melEDI + 0.1), as explained below.

```{r}
#| label: tbl-h11-data-deviations
#| tbl-cap: "H11 data, placement, and estimand deviations or clarifications."

tibble::tribble(
  ~Aspect, ~Preregistered, ~Analysed, ~Consequence,
  "Outcome and epoch",
  "Hourly geometric-mean melEDI",
  "Thirty-minute arithmetic-mean melEDI, transformed as log10(melEDI + 0.1)",
  paste(
    "The retained H11 estimand, row count, scale, and dependence structure",
    "follow the specified H02 adaptation and therefore differ from",
    "preregistration; the registered hourly outcome remains the documented",
    "preregistered specification, not an outstanding H11 sensitivity."
  ),
  "Placement role",
  "Chest primary; near eye as robustness placement",
  "Near eye primary; chest complementary",
  "The measurement closer to the eye carries the primary interpretation.",
  "Day inclusion",
  "Coverage-based exclusions; no deletion solely because a binned profile is constant",
  "Inherited coverage rules plus upstream exclusion of otherwise eligible verified exact-all-zero device days; supported constant profiles retained",
  "The fitted samples reconcile to the retained temporal-analysis samples.",
  "Analysis hierarchy",
  "Repeated hourly observations with participant and site effects",
  "Thirty-minute observations nested in participant-day and participant, conditional on observed sites",
  "Participant-specific curves and participant-day intercepts change weighting and uncertainty."
) |>
  gt::gt(rowname_col = "Aspect") |>
  gt::cols_width(
    Preregistered ~ gt::pct(24),
    Analysed ~ gt::pct(32),
    Consequence ~ gt::pct(31)
  ) |>
  h11_gt(12)
```

```{r}
#| label: tbl-h11-model-deviations
#| tbl-cap: "H11 model, inference, and reporting deviations or clarifications."

tibble::tribble(
  ~Aspect, ~Preregistered, ~Analysed, ~Consequence,
  "Sex and time terms",
  "Parametric sex effect plus cyclic sex-specific time smooth",
  "Parametric Female level plus cyclic Female-minus-Male deviation; common daily curve retained",
  "The joint null covers both a time-constant level and a time-varying shape difference.",
  "Site and participant structure",
  "Site factor smooth, site random effect, and participant random-effect smooth",
  "Inherited sum-to-zero site temporal deviations, participant-specific time curves, and participant-day random intercept; no site random intercept",
  "Inference is conditional on these observed sites and the richer repeated-measure structure.",
  "Additional context",
  "No photoperiod-state or activity-adjusted causal estimand",
  "No photoperiod-state term or activity adjustment in the retained test; an exploratory same-sample activity-context adjustment was added",
  "Sample restriction and observational adjustment are shown separately; no behavioural, causal, or mediational interpretation is made.",
  "Inference and multiplicity",
  "Delta AIC of at least 2; AR(1) if needed; FDR if multiple tests",
  "Participant-cluster robust global curve test; separate two-test BH level/shape decomposition",
  "The confirmatory claim is the complete curve, not a selected pointwise interval or raw component p-value.",
  "Effect magnitude",
  "Not operationally defined",
  "Female-to-Male shifted ratios and robust fitted-curve variation, reported only after a supported global result",
  "Fitted-curve variation is not described as variance explained."
) |>
  gt::gt(rowname_col = "Aspect") |>
  gt::cols_width(
    Preregistered ~ gt::pct(24),
    Analysed ~ gt::pct(32),
    Consequence ~ gt::pct(31)
  ) |>
  h11_gt(12)
```

H11 estimates describe the retained cohort, which includes documented participants
outside the preregistered age or employment criteria. The targeted eligibility
checks in H10 and H06 assess their respective hypotheses; they do not provide an
eligibility-restricted estimate of the H11 daily curves.

- Sample eligibility and the scope of the completed sensitivities: [Inclusion criteria](../preregistration-deviations.qmd#inclusion-criteria).
- Outcome and epoch: [H11 outcome and epoch](../preregistration-deviations.qmd#h11-outcome-and-epoch).
- Placement role: [All hypotheses; placement](../preregistration-deviations.qmd#all-hypotheses-placement).
- Day inclusion: [H02 and H11 participant-day inclusion](../preregistration-deviations.qmd#h02-and-h11-participant-day-inclusion) and [Participant-day light-signal plausibility](../preregistration-deviations.qmd#participant-day-light-signal-plausibility).
- Analysis hierarchy: [H11 temporal and site smooths](../preregistration-deviations.qmd#h11-temporal-and-site-smooths) and [H11 participant hierarchy](../preregistration-deviations.qmd#h11-participant-hierarchy).
- Biological-sex and time terms: [H11 sex estimand](../preregistration-deviations.qmd#h11-sex-estimand).
- Site and participant structure: [H11 temporal and site smooths](../preregistration-deviations.qmd#h11-temporal-and-site-smooths) and [H11 participant hierarchy](../preregistration-deviations.qmd#h11-participant-hierarchy).
- Exploratory activity context: [H11 contextual activity analysis](../preregistration-deviations.qmd#h11-contextual-activity-analysis).
- Complete-curve inference and multiplicity: [H11 curve inference and multiplicity](../preregistration-deviations.qmd#h11-curve-inference-and-multiplicity).
