---
title: "Publication figures and tables"
css: ../styles/manuscript.css
format:
  html:
    code-fold: false
execute:
  echo: true
  freeze: false
---

The preceding pages prepare the data, fit the models, and display their results.
This page assembles the figures and tables used in the manuscript and its
supplementary material. Every input is generated earlier in the same project
render. The table objects retain the data and formatting shown on the analysis
pages; the paired CSV files make the displayed values available independently.

```{r}
#| label: publication-setup

source("scripts/project.R")
analysis_setup()
source("scripts/reporting/publication.R")
```

## Descriptive and hypothesis tables

The main participant table and the detailed supplementary results use the same
styled table objects that are displayed beside their analyses. The table below
identifies each source by its executable Quarto cell label. These are regenerated
objects, so changing an analysis propagates to the manuscript in the same render.

```{r}
#| label: publication-table-sources

source_tables <- tibble::tribble(
  ~publication_table, ~analysis_cell,
  "table-1", "tbl-participant-site-manuscript",
  "table-s1", "tbl-descriptive-sample-flow",
  "table-s2", "tbl-near-eye-metrics",
  "table-s3", "tbl-recommendation-context",
  "table-s7", "tbl-h01-primary-publication-summary",
  "table-s8", "tbl-h07-near-results",
  "table-s9", "tbl-h06-primary-effects",
  "table-s11a", "tbl-h05-near-results-a",
  "table-s11b", "tbl-h05-near-results-b",
  "table-s12", "tbl-h08-near-eye-results",
  "table-s13", "tbl-h09-near-eye-results",
  "table-s14", "tbl-h10-main-results",
  "table-s15", "tbl-h11-global-tests"
)
source_tables
```

```{r}
#| label: publication-share-tables

for (i in seq_len(nrow(source_tables))) {
  table <- read_analysis_table(source_tables$analysis_cell[[i]])
  if (source_tables$publication_table[[i]] == "table-s8") {
    table <- gt::text_transform(table, function(x) ifelse(x == ";", "Not applicable", x),
      locations = gt::cells_body())
  }
  save_publication_table(table, source_tables$publication_table[[i]])
}
```

Table 3 combines the metric definitions, descriptive distributions, fitted
associations and represented variation. Its inferential quantities are those
shown in the geographic analysis, and the supplementary tables retain the
complete geographic and photoperiod results. The compact table uses the same
regenerated distributions as the descriptive metric dictionary, with site colours
and no site-name labels inside the thumbnails. Decimal hours are used for duration
summaries, while timing summaries retain clock notation.

```{r}
#| label: tbl-publication-metric-context
#| tbl-cap: "Near-eye light-exposure metrics, their distributions and geographic context."

metric_context_table <- build_metric_publication_table()
save_publication_table(metric_context_table, "table-3")
publication_html_metric_table(metric_context_table)
```

## Recommendation adherence

The main adherence table combines descriptive pooled-minute fractions with
modelled recommendation-window estimates. These have different denominators.
Only the primary any-valid sample contributes the model estimates in this table;
coverage sensitivities remain on the recommendation-adherence analysis page.

```{r}
#| label: tbl-publication-recommendation-adherence
#| tbl-cap: "Recommendation adherence by recommendation window and day type."

adherence_table <- build_adherence_publication_table()
save_publication_table(adherence_table, "table-2")
adherence_table
```

The exploratory cross-window table separates within-participant contrasts from
between-participant differences. Its displayed effects come from the four-effect
FDR family and preserve the temporal-dependence qualification.

```{r}
#| label: tbl-publication-cross-window-associations
#| tbl-cap: "Exploratory cross-window adherence associations."

cross_window_table <- build_cross_window_table()
save_publication_table(cross_window_table, "table-s4")
cross_window_table
```

## Daily-pattern variation

Fitted-curve dispersion and Shapley allocation summarise different aspects of the
same daily-pattern models. Each position is assembled separately from its
regenerated estimates and bootstrap intervals. The bootstrap count follows the
central full or quick-run setting.

```{r}
#| label: tbl-publication-near-eye-variation
#| tbl-cap: "Near-eye fitted-curve dispersion and allocation of in-sample model fit."

near_eye_variation <- build_temporal_summary_table("glasses")
save_publication_table(near_eye_variation, "table-s5")
near_eye_variation
```

```{r}
#| label: tbl-publication-chest-variation
#| tbl-cap: "Complementary chest fitted-curve dispersion and allocation of in-sample model fit."

chest_variation <- build_temporal_summary_table("chest")
save_publication_table(chest_variation, "table-s6")
chest_variation
```

## Person-level synthesis

The synthesis selects the prespecified near-eye analyses. Retained associations,
confidence intervals, exact samples and FDR decisions are read from the current
model outputs. This keeps an inconclusive association distinct from evidence of
absence and keeps the two chronotype instruments separate.

```{r}
#| label: tbl-publication-person-level-synthesis
#| tbl-cap: "Person-level evidence synthesis."

person_level_table <- build_person_level_summary()
save_publication_table(person_level_table, "table-s10")
person_level_table
```

## Light-source and activity figure assembly

The temporal curve panels are generated on the light-source and activity pages.
Here, the site-average estimates, site-specific ratios, confidence intervals,
FDR decisions, and exact category support are read from their current CSV
exports and laid out as the fourth panel. The activity display keeps Other as
an additive descriptive estimate separate from its five-category interaction.

```{r}
#| label: publication-context-figures

source("scripts/reporting/context-figures.R")
build_light_source_publication_figure()
build_activity_publication_figure()
```

## Figures

The figure files below are produced by their respective analyses. This page
assigns manuscript filenames to the regenerated images. Vector images are used
when available; PNG versions provide a portable alternative for Word. Each
analysis also exports the exact data used to draw its figures.

```{r}
#| label: publication-figure-sources

figure_sources <- tibble::tribble(
  ~figure, ~source_pattern,
  "figure-1", "^descriptive_overview[.](svg|png)$",
  "figure-2", "^daily_patterns[.](svg|png)$",
  "figure-3", "^H04.*manuscript_figure3[.](svg|png)$",
  "figure-s1", "^near_eye_metric_distributions[.](svg|png)$",
  "figure-s2", "^time_series_to_metrics[.](svg|png)$",
  "figure-s3", "^latitude_photoperiod_diagnostic[.](svg|png)$",
  "figure-s4", "^main_adherence_levels[.](svg|png)$",
  "figure-s5a", "^main_site_workday[.](svg|png)$",
  "figure-s5b", "^main_site_free_work[.](svg|png)$",
  "figure-s6", "^participant_state_raincloud[.](svg|png)$",
  "figure-s7", "^H01.*model_support[.](svg|png)$",
  "figure-s8", "^H07.*smooth_derivative_pairs_near_eye[.](svg|png)$",
  "figure-s9", "^H03_light_source_patterns[.](svg|png)$",
  "figure-s10", "^H06_paired_placement_effects[.](svg|png)$",
  "figure-s11", "^H06_reader_temporal_day_type[.](svg|png)$",
  "figure-s12", "^H06_reader_temporal_activity[.](svg|png)$",
  "figure-s13", "^H06.*site_specific_significance_screen[.](svg|png)$",
  "figure-s14", "^H05_reader_near_eye_effects[.](svg|png)$",
  "figure-s15", "^H08_near_eye_effects[.](svg|png)$",
  "figure-s16", "^H09_primary_effects[.](svg|png)$",
  "figure-s17", "^H09_observed_timing_patterns[.](svg|png)$",
  "figure-s18", "^H10_age_site_significant_associations[.](svg|png)$",
  "figure-s19", "^H11_reader_primary_near_eye_curves[.](svg|png)$"
)
figure_sources
```

```{r}
#| label: publication-save-figures

figure_paths <- purrr::map2_dfr(
  figure_sources$figure,
  figure_sources$source_pattern,
  publish_figure
)
readr::write_csv(
  figure_paths,
  "results/csv/source_data/publication/figure-paths.csv"
)
figure_paths
```

The [manuscript](../index.qmd) and [supplementary material](../supplementary.qmd)
now read these tables and images. Their figures and tables do not require
pre-existing rendered pages or manual copying.


## Word table layout

The manuscript uses native, editable Word tables. After Quarto writes `docs/index.docx`,
the project hook `scripts/reporting/word-layout.R` normalises paragraph alignment,
keeps table rows together, repeats column headings and gives wide tables landscape
pages. It reads the newly rendered Word file and changes layout properties only;
the text, numerical cells and embedded figures are preserved. This step runs
automatically as part of `quarto render`.
