---
title: "Recommendation adherence"
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

This analysis estimates the proportion of valid minutes meeting the Brown recommendations within each behavioural window. It regenerates the fitted models, uncertainty, model checks, sensitivity analyses and publication displays from the prepared measurement data. Reusable R functions implement the likelihood and numerical methods; the cells below make the inputs, analytical sequence and outputs explicit.

```{r}
#| label: adherence-setup
source("scripts/project.R")
analysis_setup()
library(data.table)
setDTthreads(1L)
library(gt)
`%chin%` <- data.table::`%chin%`
for (helper in c("io", "periods", "boundary_model", "fit_model", "estimands", "diagnostics", "sensitivity_frames")) {
  source(file.path("scripts/models/brown", paste0(helper, ".R")))
}
```

## Inputs and recommendation windows

The inputs are the regenerated minute-level coverage data and normalized sleep diaries. Daytime requires melanopic EDI of at least 250 lx; the three hours before sleep require at most 10 lx; the sleep environment requires at most 1 lx. Counts retain the actual valid-minute denominator. Sleep preceding a waking date, that date’s Daytime, and the following Pre-sleep window form one cycle, whose day type comes from the waking-date diary entry.

```{r}
#| label: adherence-inputs
site_registry <- fread("config/site_display_registry.csv")
sleep <- as.data.table(readRDS("results/intermediate/model_data/normalized_inputs/sleepdiaries.rds"))
coverage_near_eye <- as.data.table(readRDS("results/intermediate/coverage/light_glasses_coverage.rds"))
coverage_chest <- as.data.table(readRDS("results/intermediate/coverage/light_chest_coverage.rds"))
required_coverage <- c(
  "site",
  "Id",
  "position",
  "datetime_utc",
  "local_date",
  "State.Brown",
  "MEDI_eligible",
  "sleep_source_row_start",
  "sleep_source_row_end"
)
required_sleep <- c(
  "site",
  "Id",
  "source_row",
  "wake_wall",
  "wake_utc",
  "sleepprep_utc",
  "daytype2",
  "site_timezone"
)
if (
  !all(required_coverage %chin% names(coverage_near_eye)) ||
    !all(required_coverage %chin% names(coverage_chest)) ||
    !all(required_sleep %chin% names(sleep))
) {
  stop("Required coverage or sleep fields are missing.", call. = FALSE)
}

sleep_map <- sleep[, .(
  site = as.character(site),
  Id = as.character(Id),
  source_row = as.integer(source_row),
  behavior_date = as.Date(wake_wall),
  daytype_source_value = as.character(daytype2),
  day_type = data.table::fcase(
    as.character(daytype2) == "a work day",
    "Work day",
    as.character(daytype2) == "a free day",
    "Free day",
    default = NA_character_
  ),
  site_timezone = as.character(site_timezone),
  wake_utc,
  sleepprep_utc
)]
if (anyDuplicated(sleep_map[, .(site, Id, source_row)])) {
  stop("The sleep source-row key is duplicated.", call. = FALSE)
}

state_labels <- c(
  wake = "Wake outside the three hours before sleep",
  `pre-sleep` = "Pre-sleep",
  sleep = "Sleep environment"
)
site_levels <- site_registry[order(display_order), as.character(site)]
near_eye <- build_period_data(coverage_near_eye, "glasses", include_calendar = TRUE)
chest <- build_period_data(coverage_chest, "chest")
variant_b <- "B_previous_sleep_wake_following_presleep"
variant_c <- "C_previous_presleep_sleep_then_wake"
b_any <- factorize_frame(near_eye$period[
  linkage_variant == variant_b & !is.na(day_type) & valid_minutes > 0L
])
b_80 <- b_any[b_any$support_fraction >= 0.80, , drop = FALSE]
frames <- list(B_any = b_any, B_80 = b_80)
stopifnot(all(vapply(frames, function(x) all(
  x$brown_yes >= 0 & x$brown_no >= 0 & x$brown_yes + x$brown_no == x$valid_minutes
), logical(1))))
flow <- rbindlist(lapply(names(frames), function(id) {
  x <- frames[[id]]
  data.table(sample_id = id, rows = nrow(x), participants = uniqueN(x$participant_id),
             cycles = uniqueN(x$behavioral_day_id), valid_minutes = sum(x$valid_minutes))
}))
lb_save_rds(frames, "frames/model_frames.rds")
lb_save_rds(list(near_eye = near_eye$period, chest = chest$period), "frames/window_periods.rds")
lb_write_csv(flow, "frames/sample_flow.csv")
flow
```

## Endpoint-inflated beta-binomial model

The mean includes the full window × site × day-type interaction and a participant random intercept. Separate mixture components account for all-zero and all-one periods, with the all-one component restricted to Pre-sleep and Sleep. Window-specific beta-binomial dispersion accounts for remaining count heterogeneity. The C++ files implement this likelihood and its integrated predictions through TMB. Compilation, exact probability and moment checks, and automatic-derivative checks run before fitting.

```{r}
#| label: adherence-likelihood
brown_compile("endpoint_inflated_betabinomial")
brown_compile("endpoint_inflated_estimands")
brown_compile("endpoint_inflated_sensitivity_estimands")
synthetic_design <- function(
  y,
  n,
  one_active,
  use_zero_component,
  eta_mu,
  eta_zero,
  eta_one,
  eta_disp
) {
  observations <- length(y)
  list(
    design = list(
      data = list(
        y = as.integer(y),
        n = as.integer(n),
        X_mu = matrix(1, observations, 1),
        X_zero = matrix(1, observations, 1),
        X_one = matrix(1, observations, 1),
        X_disp = matrix(1, observations, 1),
        Z_mu_part = matrix(1, observations, 1),
        part_index = rep(0L, observations),
        day_index = rep(0L, observations),
        one_active = as.integer(one_active),
        use_zero_component = as.integer(use_zero_component),
        use_mu_part_re = 0L,
        use_day_re = 0L,
        use_zero_re = 0L,
        use_one_re = 0L
      )
    ),
    parameters = list(
      beta_mu = eta_mu,
      beta_zero = eta_zero,
      beta_one = eta_one,
      beta_disp = eta_disp,
      b_mu_part = matrix(0, 1, 1),
      b_day = 0,
      b_zero_part = 0,
      b_one_part = 0,
      log_sd_mu_part = 0,
      log_sd_day = 0,
      log_sd_zero = 0,
      log_sd_one = 0
    )
  )
}
grid <- expand.grid(
  n = c(1L, 2L, 12L),
  eta_mu = c(-2, 0.5),
  one_active = c(FALSE, TRUE),
  use_zero = c(FALSE, TRUE)
)
exact <- lapply(seq_len(nrow(grid)), function(i) {
  g <- grid[i, ]
  s <- synthetic_design(
    0:g$n,
    rep(g$n, g$n + 1L),
    rep(g$one_active, g$n + 1L),
    g$use_zero,
    g$eta_mu,
    -1.1,
    -0.8,
    log(7)
  )
  object <- ba_boundary_make_object(s$design, s$parameters)
  objective_value <- object$fn()
  report <- object$report()
  weights <- ba_boundary_component_weights(-1.1, -0.8, g$one_active, g$use_zero)
  pmf <- ba_boundary_mixture_support(
    g$n,
    plogis(g$eta_mu),
    7,
    weights$pi_zero,
    weights$pi_one,
    weights$pi_beta
  )
  fraction <- pmf$y / g$n
  exact_mean <- sum(fraction * pmf$probability)
  exact_variance <- sum((fraction - exact_mean)^2 * pmf$probability)
  data.frame(
    case = i,
    n = g$n,
    one_active = g$one_active,
    use_zero = g$use_zero,
    normalization_error = abs(sum(pmf$probability) - 1),
    objective_error = abs(objective_value + sum(log(pmf$probability))),
    pmf_error = max(abs(exp(report$log_likelihood) - pmf$probability)),
    mean_error = max(abs(report$conditional_mean - exact_mean)),
    variance_error = max(abs(report$conditional_variance - exact_variance)),
    cdf_minimum_increment = min(diff(c(0, pmf$cdf))),
    cdf_terminal_error = abs(tail(pmf$cdf, 1L) - 1),
    zero_restriction_exact = g$use_zero || all(report$pi_zero == 0),
    one_restriction_exact = g$one_active || all(report$pi_one == 0)
  )
})
exact <- do.call(rbind, exact)
derivative <- synthetic_design(
  c(0L, 2L, 5L, 8L, 10L),
  rep(10L, 5L),
  c(FALSE, TRUE, TRUE, TRUE, TRUE),
  TRUE,
  -0.2,
  -1.1,
  -0.8,
  log(7)
)
object <- ba_boundary_make_object(derivative$design, derivative$parameters)
theta <- object$par
ad <- object$gr(theta)
fd <- vapply(
  seq_along(theta),
  function(i) {
    delta <- 1e-6 * (1 + abs(theta[i]))
    upper <- lower <- theta
    upper[i] <- upper[i] + delta
    lower[i] <- lower[i] - delta
    (object$fn(upper) - object$fn(lower)) / (2 * delta)
  },
  numeric(1)
)
derivatives <- data.frame(
  parameter = names(theta),
  automatic = ad,
  finite_difference = fd,
  absolute_difference = abs(ad - fd)
)
stopifnot(max(exact$normalization_error) < 1e-12,
          max(exact$objective_error) < 1e-7, max(exact$pmf_error) < 1e-12,
          max(exact$mean_error) < 1e-12, max(exact$variance_error) < 1e-12,
          min(exact$cdf_minimum_increment) >= -1e-14,
          max(exact$cdf_terminal_error) < 1e-12,
          all(exact$zero_restriction_exact & exact$one_restriction_exact),
          max(derivatives$absolute_difference) < 1e-4)
lb_write_csv(exact, "likelihood/probability_moment_checks.csv")
lb_write_csv(derivatives, "likelihood/derivative_checks.csv")
```

## Primary models

Both the any-valid-period sample and the sample with at least 80% window coverage use the same model. Starting values for the any-valid sample are calculated deterministically from its design. The 80% fit starts from that newly fitted model, matching participant effects by participant identifier. Numerical convergence, gradient, Hessian and variance checks are exported with each fit.

```{r}
#| label: adherence-primary-models
primary_models <- list()
for (sample_id in c("B_any", "B_80")) {
  design <- ba_boundary_build_design(frames[[sample_id]], "F3", "R3", "Q2", "Q1", "D0")
  initial <- ba_boundary_initial_parameters(design)
  if (sample_id == "B_80") {
    parent <- primary_models$B_any
    for (component in c("beta_mu", "beta_zero", "beta_one", "beta_disp",
                        "log_sd_mu_part", "log_sd_day", "log_sd_zero", "log_sd_one")) {
      stopifnot(length(initial[[component]]) == length(parent$parameter_list[[component]]))
      initial[[component]] <- parent$parameter_list[[component]]
    }
    participant_match <- match(design$levels$participant_id, parent$design_object$levels$participant_id)
    stopifnot(!anyNA(participant_match))
    initial$b_mu_part <- parent$parameter_list$b_mu_part[participant_match, , drop = FALSE]
  }
  model_id <- paste0("BA-LB-PRIMARY-", if (sample_id == "B_any") "ANY" else "80")
  fit <- ba_lb_fit_candidate(design, initial, model_id, sample_id)
  brown_save_fit(fit)
  primary_models[[sample_id]] <- fit
  print(fit$fit_diagnostics[, .(model_id, rows, participants, objective,
    maximum_absolute_gradient, fit_status, failure_components)])
  stopifnot(!fit$fit_diagnostics$structural_failure)
}
```

## Marginal adherence and contrasts

Predictions integrate over the participant effect, average the observed valid-minute denominators within each window/site/day-type cell, and weight the nine sites equally. The 15-node and 30-node quadrature results must agree within 0.05 percentage points. The primary Free-minus-Work contrasts use one three-window false-discovery-rate family per sample. Additional site and interaction contrasts retain their specified multiplicity families.

```{r}
#| label: adherence-primary-estimands
model_paths <- c(primary_any_valid = brown_file("models/BA-LB-PRIMARY-ANY.rds"),
                 support_80 = brown_file("models/BA-LB-PRIMARY-80.rds"))
site_registry <- data.table::fread("config/site_display_registry.csv")
site_registry <- site_registry[order(display_order)]
site_label <- stats::setNames(site_registry$display_name, site_registry$site)
state_label <- c(
  "Wake outside the three hours before sleep" = "Daytime",
  "Pre-sleep" = "Pre-sleep",
  "Sleep environment" = "Sleep"
)
day_label <- c("Work day" = "Work day", "Free day" = "Free day")

derived <- lapply(names(model_paths), function(sample_id) {
  derive_sample(sample_id, model_paths[[sample_id]])
})
names(derived) <- names(model_paths)
collect <- function(name)
  data.table::rbindlist(lapply(derived, `[[`, name), fill = TRUE)
cell_predictions <- collect("cell_predictions")
equal_site_means <- collect("equal_site_means")
quadrature <- collect("quadrature")
families <- lapply(paste0("m", seq_len(5L)), collect)
names(families) <- paste0("BA_M", seq_len(5L))
family_checks <- data.table::rbindlist(lapply(seq_along(families), function(i) {
  family_table <- as.data.frame(families[[i]])
  expected <- c(3L, 3L, 1L, 27L, 54L)[[i]]
  data.table::rbindlist(lapply(names(model_paths), function(sample) {
    z <- family_table[family_table[["sample_id"]] == sample, , drop = FALSE]
    data.table::data.table(
      family = names(families)[[i]],
      sample_id = sample,
      expected = expected,
      observed = nrow(z),
      passed = nrow(z) == expected &&
        all(is.finite(z$p_value)) &&
        all(is.finite(z$p_adjusted))
    )
  }))
}))
primary_cells <- derived$primary_any_valid$cell_predictions
key <- function(x) paste(x$analysis_state, x$site, x$day_type, sep = "||")
m1 <- families$BA_M1
coverage <- merge(
  m1[sample_id == "primary_any_valid"],
  m1[sample_id == "support_80"],
  by = c("analysis_state", "state_display", "family", "contrast", "weighting"),
  suffixes = c("_any_valid", "_80")
)
coverage[, `:=`(
  direction_preserved = sign(estimate_any_valid) == sign(estimate_80),
  interval_exclusion_any_valid = conf_low_any_valid > 0 |
    conf_high_any_valid < 0,
  interval_exclusion_80 = conf_low_80 > 0 | conf_high_80 < 0,
  absolute_shift_percentage_points = 100 * abs(estimate_80 - estimate_any_valid)
)]
coverage[,
  interval_exclusion_preserved := interval_exclusion_any_valid ==
    interval_exclusion_80
]
coverage[,
  passed := direction_preserved &
    interval_exclusion_preserved &
    absolute_shift_percentage_points <= 2
]

grid <- primary_cells
members <- unique(grid[, .(analysis_state, state_display, site, site_display)])
members[,
  state_order := match(state_display, c("Daytime", "Pre-sleep", "Sleep"))
]
members[, site_order := match(site, site_registry$site)]
data.table::setorder(members, state_order, site_order)
members[, member_id := sprintf("BA-M6-%02d", seq_len(.N))]
stopifnot(
  nrow(members) == 27L,
  !anyNA(members),
  !anyDuplicated(members[, .(analysis_state, site)])
)
contrast_matrix <- matrix(
  0,
  nrow = 27L,
  ncol = 54L,
  dimnames = list(members$member_id, grid$cell_id)
)
for (i in seq_len(nrow(members))) {
  state <- members$analysis_state[[i]]
  site <- members$site[[i]]
  state_free <- which(
    grid$analysis_state == state & grid$day_type == "Free day"
  )
  state_work <- which(
    grid$analysis_state == state & grid$day_type == "Work day"
  )
  site_free <- which(
    grid$analysis_state == state &
      grid$site == site &
      grid$day_type == "Free day"
  )
  site_work <- which(
    grid$analysis_state == state &
      grid$site == site &
      grid$day_type == "Work day"
  )
  stopifnot(
    length(state_free) == 9L,
    length(state_work) == 9L,
    length(site_free) == 1L,
    length(site_work) == 1L
  )
  contrast_matrix[i, state_free] <- -1 / 9
  contrast_matrix[i, state_work] <- 1 / 9
  contrast_matrix[i, site_free] <- contrast_matrix[i, site_free] + 1
  contrast_matrix[i, site_work] <- contrast_matrix[i, site_work] - 1
}
stopifnot(
  all(rowSums(abs(contrast_matrix) > 0) == 18L),
  max(abs(rowSums(contrast_matrix))) < 1e-12
)
m6 <- lapply(names(derived), function(sample) {
  stored <- derived[[sample]]
  stopifnot(identical(key(stored$cell_predictions), key(grid)))
  value <- stored$sd_report$value[seq_len(54L)]
  covariance <- stored$sd_report$cov[seq_len(54L), seq_len(54L), drop = FALSE]
  estimate <- as.numeric(contrast_matrix %*% value)
  variance <- diag(contrast_matrix %*% covariance %*% t(contrast_matrix))
  stopifnot(
    all(is.finite(estimate)),
    all(is.finite(variance)),
    all(variance > 0)
  )
  result <- data.table::copy(members)
  result[, `:=`(
    sample_id = sample,
    family = "BA-M6",
    estimate = estimate,
    standard_error = sqrt(variance)
  )]
  result[, `:=`(
    conf_low = estimate - stats::qnorm(0.975) * standard_error,
    conf_high = estimate + stats::qnorm(0.975) * standard_error
  )]
  member_key <- paste(result$analysis_state, result$site)
  from_m4 <- stored$m4$estimate[match(
    member_key,
    paste(stored$m4$analysis_state, stored$m4$site)
  )]
  from_m1 <- stored$m1$estimate[match(
    result$analysis_state,
    stored$m1$analysis_state
  )]
  m5_free <- stored$m5$estimate[match(
    paste(member_key, "Free day"),
    paste(stored$m5$analysis_state, stored$m5$site, stored$m5$day_type)
  )]
  m5_work <- stored$m5$estimate[match(
    paste(member_key, "Work day"),
    paste(stored$m5$analysis_state, stored$m5$site, stored$m5$day_type)
  )]
  result[, `:=`(
    from_M4_minus_M1 = from_m4 - from_m1,
    from_free_M5_minus_work_M5 = m5_free - m5_work
  )]
  result[, state_sum := sum(estimate), by = analysis_state]
  result[,
    reconciliation_passed := abs(estimate - from_M4_minus_M1) < 1e-12 &
      abs(estimate - from_free_M5_minus_work_M5) < 1e-12 &
      abs(state_sum) < 1e-12
  ]
  stopifnot(all(result$reconciliation_passed))
  if (sample == "primary_any_valid") {
    result[, statistic := estimate / standard_error]
    result[, p_value := 2 * stats::pnorm(abs(statistic), lower.tail = FALSE)]
    stopifnot(nrow(result) == 27L, all(is.finite(result$p_value)))
    result[, p_adjusted := stats::p.adjust(p_value, method = "BH")]
    result[, fdr_significant := p_adjusted < 0.05]
  }
  result
})
names(m6) <- names(derived)
m6_coverage <- merge(
  m6$primary_any_valid,
  m6$support_80,
  by = c(
    "member_id",
    "analysis_state",
    "state_display",
    "site",
    "site_display",
    "state_order",
    "site_order",
    "family"
  ),
  suffixes = c("_primary", "_80")
)
m6_coverage[, `:=`(
  direction_retained = sign(estimate_primary) == sign(estimate_80),
  interval_exclusion_retained = (conf_low_primary > 0 |
    conf_high_primary < 0) ==
    (conf_low_80 > 0 | conf_high_80 < 0),
  fully_estimable = is.finite(standard_error_primary) &
    is.finite(standard_error_80)
)]
m6_coverage[,
  significant_direction_changed := fdr_significant & !direction_retained
]
m6_coverage[,
  sensitivity_qualification_required := direction_retained &
    !interval_exclusion_retained
]

probabilities <- c(
  "adherence",
  "all_zero_probability",
  "all_one_probability",
  "mixed_probability",
  "extra_all_zero_probability",
  "extra_all_one_probability",
  "beta_binomial_component_probability"
)
probability_values <- as.matrix(cell_predictions[, ..probabilities])
stopifnot(all(quadrature$passed), all(family_checks$passed),
          all(vapply(m6, function(x) all(x$reconciliation_passed), logical(1))))
lb_save_rds(list(derived = derived, m6 = m6, contrast_matrix_M6 = contrast_matrix,
                coverage_stability = coverage, m6_coverage = m6_coverage), "estimands/boundary_estimands.rds")
lb_write_csv(cell_predictions, "estimands/cell_predictions.csv")
lb_write_csv(equal_site_means, "estimands/equal_site_means.csv")
lb_write_csv(quadrature, "estimands/quadrature_check.csv")
lb_write_csv(coverage, "estimands/coverage_stability.csv")
lb_write_csv(collect("compact_source"), "estimands/compact_table_source.csv")
for (name in names(families))
  lb_write_csv(families[[name]], paste0("multiplicity/", name, ".csv"))
lb_write_csv(m6$primary_any_valid, "multiplicity/BA_M6_primary.csv")
lb_write_csv(m6$support_80, "multiplicity/BA_M6_support80.csv")
lb_write_csv(m6_coverage, "multiplicity/BA_M6_coverage_stability.csv")
lb_write_csv(
  data.frame(
    member_id = members$member_id,
    contrast_matrix,
    check.names = FALSE
  ),
  "multiplicity/BA_M6_contrast_matrix.csv"
)
lb_write_csv(family_checks, "multiplicity/family_checks.csv")
```

## Model diagnostics

Randomized quantile residuals, simulated count distributions, boundary probabilities, calibration, and temporal residual associations assess model adequacy. The simulation seeds are fixed. The central quick profile reduces the usual 250 predictive simulations to 50.

```{r}
#| label: adherence-primary-diagnostics
primary_diagnostics <- lapply(names(model_paths), function(id) {
  result <- derive_diagnostics(id, model_paths[[id]])
  for (name in names(result)) {
    if (is.data.frame(result[[name]])) lb_write_csv(result[[name]], paste0("diagnostics/", id, "/", name, ".csv"))
  }
  lb_save_rds(result, paste0("diagnostics/", id, "/diagnostics.rds"))
  result
})
names(primary_diagnostics) <- names(model_paths)
rbindlist(lapply(primary_diagnostics, `[[`, "assessment")) |> gt()
```

## Coverage, thresholds and period eligibility

Eight alternatives vary coverage, very short windows, complete three-window cycles, observation of both day types, inclusive versus strict thresholds, and exclusion of all-zero periods. A ninth holds dispersion constant across windows. Each uses a fresh fit of the same endpoint-inflated model. If excluding zero periods makes the zero-inflation component unidentifiable, a conditional model disables only that component; its interpretation remains conditional on some adherent minutes.

```{r}
#| label: adherence-simple-sensitivities
source("scripts/models/brown/comparisons.R")
sensitivity_env <- new.env(parent = environment())
sys.source("scripts/models/brown/sensitivity_estimands.R", sensitivity_env)
primary_m1 <- families$BA_M1[sample_id == "primary_any_valid"]
sensitivity_frames <- lb_simple_sensitivity_frames(frames$B_any)
scenario_ids <- c("SENS-SUPPORT70", "SENS-SUPPORT90", "SENS-TINYGT5", "SENS-TINYGT30",
                  "SENS-COMPLETE-TRIADS", "SENS-BOTH-DAYTYPES", "SENS-STRICT", "SENS-EXCLUDE-ZERO")
names(sensitivity_frames) <- scenario_ids
sensitivity_frames$`SENS-DISPERSION-D1` <- frames$B_any
simple_fits <- list()
simple_comparisons <- list()
for (scenario in names(sensitivity_frames)) {
  design <- ba_boundary_build_design(sensitivity_frames[[scenario]], "F3", "R3", "Q2", "Q1",
    if (scenario == "SENS-DISPERSION-D1") "D1" else "D0")
  initial <- ba_boundary_initial_parameters(design)
  bundle <- ba_lb_fit_candidate(design, initial, paste0("BA-LB-", scenario), scenario)
  brown_save_fit(bundle)
  simple_fits[[scenario]] <- bundle
  if (scenario == "SENS-EXCLUDE-ZERO" && bundle$fit_diagnostics$structural_failure) {
    zero <- bundle$fixed_summary[bundle$fixed_summary$parameter == "beta_zero", ]
    stopifnot(all(design$data$y > 0L),
              bundle$fit_diagnostics$failure_components == "endpoint_separation",
              any(abs(zero$estimate) > 15 | zero$standard_error > 10))
    design$data$use_zero_component <- 0L
    design$specification$use_zero_component <- FALSE
    scenario <- "SENS-EXCLUDE-ZERO-NOZERO"
    bundle <- ba_lb_fit_candidate(design, initial, paste0("BA-LB-", scenario), scenario)
    brown_save_fit(bundle)
    simple_fits[[scenario]] <- bundle
  }
  if (!bundle$fit_diagnostics$structural_failure) {
    result <- sensitivity_env$derive_one(scenario, brown_file(paste0("models/", bundle$model_id, ".rds")))
    stopifnot(all(result$quadrature$passed))
    simple_comparisons[[scenario]] <- brown_export_estimands(result,
      paste0("sensitivities/simple_estimands/", scenario), primary_m1, scenario)
  }
}
lb_write_csv(rbindlist(simple_comparisons, fill = TRUE), "sensitivities/simple_comparisons.csv")
```

## Site and participant influence

Leave-one-site-out fits assess dependence on each site. Five participant deletions are selected by the maximum percentile across minute share, residual magnitudes and fixed-predictor variance measures. Each deletion refits the model with starting mean predictors projected onto the remaining design. The exported estimates retain their interval and directional comparison with the full sample.

```{r}
#| label: adherence-deletions
source("scripts/models/brown/deletion_initializer.R")
deletion_env <- new.env(parent = sensitivity_env)
sys.source("scripts/models/brown/deletion_estimands.R", deletion_env)
primary_bundle <- primary_models$B_any
primary_frame <- primary_bundle$design_object$frame
primary_parameters <- primary_bundle$parameter_list
primary_levels <- lapply(primary_frame[c("analysis_state", "site", "day_type")], levels)
screen <- as.data.frame(primary_diagnostics$primary_any_valid$participant_influence)
metrics <- c("valid_minute_share", "mean_absolute_pearson_residual", "maximum_absolute_pearson_residual",
             "mean_absolute_quantile_residual", "maximum_absolute_quantile_residual",
             "maximum_fixed_linear_predictor_variance", "sum_fixed_linear_predictor_variance")
percentiles <- vapply(metrics, function(m) frank(screen[[m]], ties.method = "average") / nrow(screen), numeric(nrow(screen)))
ordered_screen <- data.table(score = apply(percentiles, 1L, max), participant_id = screen$participant_id)
setorder(ordered_screen, -score, participant_id)
selected <- ordered_screen$participant_id[seq_len(5L)]
sites <- c("RISE", "THUAS", "BAUA", "MPI", "TUM", "FUSPCEU", "IZTECH", "UCR", "KNUST")
targets <- data.frame(scenario = c(paste0("LOSO-", sites), paste0("INFLUENCE-", seq_len(5L))),
                      variable = c(rep("site", length(sites)), rep("participant_id", 5L)),
                      target = c(sites, as.character(selected)))
deletion_comparisons <- list()
for (i in seq_len(nrow(targets))) {
  target <- targets[i, ]
  keep <- as.character(primary_frame[[target$variable]]) != target$target
  raw_frame <- primary_frame[keep, , drop = FALSE]
  design <- ba_boundary_build_design(raw_frame, "F3", "R3", "Q2", "Q1", "D0")
  initial <- prepare_start(raw_frame, design)
  stopifnot(max(abs(as.numeric(primary_bundle$design_object$data$X_mu[keep, , drop = FALSE] %*% primary_parameters$beta_mu) -
                    as.numeric(design$data$X_mu %*% initial$beta_mu))) <= 1e-8)
  bundle <- ba_lb_fit_candidate(design, initial, paste0("BA-LB-", target$scenario), target$scenario)
  brown_save_fit(bundle)
  if (!bundle$fit_diagnostics$structural_failure) {
    result <- deletion_env$brown_deletion_estimands(bundle, target$scenario)
    deletion_comparisons[[target$scenario]] <- brown_export_estimands(result,
      paste0("sensitivities/deletion_estimands/", target$scenario), primary_m1, target$scenario)
  }
}
lb_write_csv(rbindlist(deletion_comparisons, fill = TRUE), "sensitivities/deletion_comparisons.csv")
rbindlist(deletion_comparisons, fill = TRUE)[, .(scenario, state_display, estimate,
  conf_low, conf_high, shift_percentage_points, direction_retained)] |> gt()
```

## Alternative response distributions and weighting

An ordinary-binomial model, a non-inflated beta-binomial model and fractional models with equal-period or valid-minute weights assess distributional assumptions. Fractional uncertainty uses participant-cluster HC3 covariance. An additional summary weights the primary model’s site means by observed support. These are sensitivity estimates with no additional significance families.

```{r}
#| label: adherence-family-comparisons
source("scripts/models/brown/comparison_models.R")
comparison_env <- new.env(parent = environment())
sys.source("scripts/models/brown/comparison_estimands.R", comparison_env)
comparison_env$mean_result <- mean_result
comparison_data <- ba_prepare_factor_frame(frames$B_any)
family_models <- list()
for (scenario in c("DIAG-BINOMIAL", "B-BETA-BINOMIAL", "FRACTIONAL-EQUAL", "FRACTIONAL-MINUTE")) {
  model_id <- paste0("BA-LB-", scenario)
  if (grepl("FRACTIONAL", scenario)) {
    bundle <- fit_fractional(comparison_data, weighted = scenario == "FRACTIONAL-MINUTE", model_id = model_id)
  } else {
    started <- proc.time()[["elapsed"]]
    fitted <- if (scenario == "DIAG-BINOMIAL") ba_fit_binomial(comparison_data, ba_formula_f3_r0) else
      ba_fit_beta_binomial(comparison_data, ba_formula_f3_r0, ba_dispersion_state)
    assessment <- ba_assess_fit(fitted, comparison_data, model_id, "B_any", "BA-F3-R0", "D0",
                                proc.time()[["elapsed"]] - started,
                                family_name = if (scenario == "DIAG-BINOMIAL") "binomial" else "betabinomial")
    bundle <- list(model = fitted, data = comparison_data, fit_diagnostics = assessment$fit_diagnostics, random_sd = assessment$random_sd)
  }
  family_models[[scenario]] <- bundle
  lb_save_rds(bundle, paste0("models/", model_id, ".rds"))
  lb_write_csv(bundle$fit_diagnostics, paste0("models/", model_id, "_fit_diagnostics.csv"))
}
family_comparisons <- list()
family_status <- rbindlist(lapply(names(family_models), function(id) {
  bundle <- family_models[[id]]
  data.table(scenario = id, fit_status = bundle$fit_diagnostics$fit_status,
             hard_failure = bundle$fit_diagnostics$hard_failure, estimates_eligible = !bundle$fit_diagnostics$hard_failure)
}))
for (scenario in names(family_models)) {
  bundle <- family_models[[scenario]]
  if (bundle$fit_diagnostics$hard_failure) next
  cells <- if (grepl("FRACTIONAL", scenario)) comparison_env$integrated_cells_fractional(bundle) else
    comparison_env$integrated_cells_glmm(bundle$model, bundle$data, 30L)
  covariance <- cells$gradient %*% cells$variance_covariance %*% t(cells$gradient)
  contrasts <- lapply(levels(bundle$data$analysis_state), function(state) {
    w <- as.numeric(cells$grid$analysis_state == state & cells$grid$day_type == "Free day") / 9 -
      as.numeric(cells$grid$analysis_state == state & cells$grid$day_type == "Work day") / 9
    cbind(data.frame(analysis_state = state), mean_result(w, cells$probability, covariance))
  })
  family_comparisons[[scenario]] <- brown_compare(do.call(rbind, contrasts), primary_m1, scenario)
}
lb_write_csv(family_status, "sensitivities/family_grouping_weighting/family_status.csv")
lb_write_csv(rbindlist(family_comparisons, fill = TRUE), "sensitivities/family_grouping_weighting/family_comparison.csv")
grid <- derived$primary_any_valid$cell_predictions
value <- derived$primary_any_valid$sd_report$value[seq_len(54L)]
covariance <- derived$primary_any_valid$sd_report$cov[seq_len(54L), seq_len(54L)]
site_weights <- comparison_env$site_weight_vector(comparison_data, "observed_support")
weighted_contrasts <- lapply(levels(comparison_data$analysis_state), function(state) {
  w <- comparison_env$weights_for_cells(list(grid = grid), state, "Free day", site_weights) -
    comparison_env$weights_for_cells(list(grid = grid), state, "Work day", site_weights)
  cbind(data.frame(analysis_state = state), mean_result(w, value, covariance))
})
lb_write_csv(brown_compare(do.call(rbind, weighted_contrasts), primary_m1, "Observed support weighting"),
             "sensitivities/family_grouping_weighting/observed_support_weighted_contrasts.csv")
```

## Alternative grouping of sleep and wake windows

The alternative grouping associates the preceding Pre-sleep period with the subsequent Sleep and Daytime periods. This sensitivity changes the ownership of the Pre-sleep window while retaining physical measurements and thresholds. Both coverage samples are rebuilt from the same minute data and fitted with the endpoint-inflated model.

```{r}
#| label: adherence-grouping-sensitivity
c_any <- factorize_frame(near_eye$period[
  linkage_variant == variant_c & !is.na(day_type) & valid_minutes > 0L
])
c_frames <- list(primary_any_valid = c_any, support_80 = c_any[c_any$support_fraction >= .80, , drop = FALSE])
c_comparisons <- list()
for (id in names(c_frames)) {
  design <- ba_boundary_build_design(c_frames[[id]], "F3", "R3", "Q2", "Q1", "D0")
  bundle <- ba_lb_fit_candidate(design, ba_boundary_initial_parameters(design),
                                paste0("BA-LB-GROUPING-C-", id), id)
  brown_save_fit(bundle)
  stopifnot(!bundle$fit_diagnostics$structural_failure)
  result <- sensitivity_env$derive_one(id, brown_file(paste0("models/", bundle$model_id, ".rds")))
  reference <- families$BA_M1[sample_id == id]
  comparison <- brown_compare(result$m1, reference, paste0("Grouping C, ", id))
  comparison$sample_id <- id
  c_comparisons[[id]] <- comparison
}
lb_write_csv(rbindlist(c_comparisons, fill = TRUE), "sensitivities/family_grouping_weighting/B_versus_C_grouping.csv")
rbindlist(c_comparisons, fill = TRUE)[, .(sample_id, state_display, estimate,
  conf_low, conf_high, shift_percentage_points, direction_retained)] |> gt()
```

## Response-scale variance decomposition

The response-scale decomposition separates fixed-cell variation, participant heterogeneity and expected count variability. Shapley allocation averages over all orders of the three predictors. These are point estimates, calculated at 15 and 30 quadrature nodes to check numerical stability. The random-effect export contains only effects estimated by the model; unused likelihood placeholders are omitted.

```{r}
#| label: adherence-variance
source("scripts/models/brown/variance_decomposition.R")
models <- primary_models[c("B_any", "B_80")]
names(models) <- names(fin_samples)
results <- list()
for (nodes in c(15L, 30L))
  for (sample_id in names(fin_samples)) {
    results[[paste(sample_id, nodes, sep = "_")]] <- fin_derive(
      models[[sample_id]],
      frames[[fin_samples[[sample_id]]]],
      sample_id,
      nodes
    )
  }
collect <- function(name) do.call(rbind, lapply(results, `[[`, name))
decomposition <- collect("decomposition")
allocation <- collect("allocation")
subsets <- collect("subsets")
cells <- collect("cells")
permutations <- collect("permutations")

quadrature <- list()
for (sample_id in names(fin_samples))
  for (scope in unique(decomposition$decomposition)) {
    a <- decomposition[
      decomposition$sample_id == sample_id &
        decomposition$decomposition == scope &
        decomposition$quadrature_nodes == 15L,
    ]
    b <- decomposition[
      decomposition$sample_id == sample_id &
        decomposition$decomposition == scope &
        decomposition$quadrature_nodes == 30L,
    ]
    for (metric in c(
      "marginal_r2",
      "conditional_r2",
      "random_effect_increment",
      "observation_distribution_share"
    )) {
      delta <- abs(b[[metric]] - a[[metric]]) * 100
      quadrature[[length(quadrature) + 1L]] <- data.frame(
        sample_id = sample_id,
        decomposition = scope,
        metric = metric,
        value_15 = a[[metric]],
        value_30 = b[[metric]],
        difference_percentage_points = delta,
        defined = TRUE,
        passed = is.finite(delta) && delta <= 0.05
      )
    }
    for (player in unique(allocation$player[
      allocation$decomposition == scope
    ])) {
      a <- allocation[
        allocation$sample_id == sample_id &
          allocation$decomposition == scope &
          allocation$player == player &
          allocation$quadrature_nodes == 15L,
      ]
      b <- allocation[
        allocation$sample_id == sample_id &
          allocation$decomposition == scope &
          allocation$player == player &
          allocation$quadrature_nodes == 30L,
      ]
      for (metric in c("absolute_r2_contribution", "relative_weight_percent")) {
        defined <- if (metric == "relative_weight_percent")
          a$relative_weight_defined && b$relative_weight_defined else TRUE
        delta <- if (defined)
          abs(b[[metric]] - a[[metric]]) *
            if (metric == "relative_weight_percent") 1 else 100 else NA_real_
        pass <- if (defined) is.finite(delta) && delta <= 0.05 else
          !a$relative_weight_defined &&
            !b$relative_weight_defined &&
            is.na(a[[metric]]) &&
            is.na(b[[metric]])
        quadrature[[length(quadrature) + 1L]] <- data.frame(
          sample_id = sample_id,
          decomposition = scope,
          metric = paste(player, metric, sep = "::"),
          value_15 = a[[metric]],
          value_30 = b[[metric]],
          difference_percentage_points = delta,
          defined = defined,
          passed = pass
        )
      }
    }
  }
quadrature <- do.call(rbind, quadrature)
final_d <- decomposition[decomposition$quadrature_nodes == 30L, ]
final_a <- allocation[allocation$quadrature_nodes == 30L, ]
final_s <- subsets[subsets$quadrature_nodes == 30L, ]
final_c <- cells[cells$quadrature_nodes == 30L, ]
validation <- data.frame(
  check = c(
    "quadrature_all_metrics",
    "all_permutations",
    "two_global_and_six_within_decompositions",
    "complete_global_and_within_subset_exports",
    "108_reference_cells",
    "variance_total_identity",
    "shares_identity",
    "conditional_minus_marginal_random_increment",
    "finite_point_components",
    "relative_weights_not_clamped"
  ),
  passed = c(
    all(quadrature$passed),
    all(permutations$passed),
    nrow(final_d) == 8L,
    nrow(final_s) == 40L,
    nrow(final_c) == 108L,
    all(
      abs(
        final_d$total_variance -
          final_d$fixed_variance -
          final_d$random_variance -
          final_d$observation_variance
      ) <=
        1e-8
    ),
    all(
      abs(
        final_d$marginal_r2 +
          final_d$random_effect_increment +
          final_d$observation_distribution_share -
          1
      ) <=
        1e-8
    ),
    all(
      abs(
        final_d$conditional_r2 -
          final_d$marginal_r2 -
          final_d$random_effect_increment
      ) <=
        1e-8
    ),
    all(is.finite(as.matrix(final_d[c(
      "fixed_variance",
      "random_variance",
      "observation_variance",
      "total_variance"
    )]))),
    all(
      is.na(final_a$relative_weight_percent) == !final_a$relative_weight_defined
    )
  )
)
stopifnot(all(validation$passed))
lb_write_csv(validation, "r2/numerical_checks.csv")
lb_write_csv(final_d, "r2/variance_decomposition.csv")
lb_write_csv(final_a, "r2/shapley_allocation.csv")
lb_write_csv(final_s, "r2/subset_variance.csv")
lb_write_csv(final_c, "r2/cell_moments.csv")
lb_write_csv(quadrature, "r2/quadrature.csv")
random_effects <- rbindlist(lapply(names(models), function(sample) {
  as.data.table(models[[sample]]$random_sd)[active == TRUE, .(
    sample_id = sample, component, term, logit_standard_deviation = standard_deviation
  )]
}))
lb_write_csv(random_effects, "r2/random_effects.csv")
```

## Complementary chest placement

The chest comparison uses Daytime and Pre-sleep windows with the same waking-date ownership. It requires consecutive diary dates and verifies the observed UTC minute bounds. Sleep is not included because the chest logger does not provide the same sleep-environment measurement. A fresh non-inflated chest model supplies deterministic starting values before the endpoint-inflated chest model is fitted.

```{r}
#| label: adherence-chest
chest_env <- new.env(parent = environment())
sys.source("scripts/models/brown/chest.R", chest_env)
stopifnot(
  !anyDuplicated(sleep_map[, .(site, Id, source_row)]),
  !anyDuplicated(sleep_map[, .(site, Id, wake_utc)])
)
data.table::setorder(sleep_map, site, Id, wake_utc, source_row)
chronology <- data.table::copy(sleep_map)
chronology[,
  `:=`(
    next_chronological_source = data.table::shift(source_row, type = "lead"),
    next_chronological_date = data.table::shift(behavior_date, type = "lead"),
    next_chronological_sleep_utc = data.table::shift(
      sleepprep_utc,
      type = "lead"
    )
  ),
  by = .(site, Id)
]
coverage <- coverage_chest
built <- build_period_data(coverage, "chest", include_calendar = FALSE)
cycle <- built$cycle
anchor_key <- function(site, id, source) paste(site, id, source, sep = "\r")
index <- match(
  anchor_key(cycle$site, cycle$Id, cycle$behavior_source_row),
  anchor_key(chronology$site, chronology$Id, chronology$source_row)
)
stopifnot(!anyNA(index))
anchor <- chronology[index]
cycle[, `:=`(
  chronological_next_source = anchor$next_chronological_source,
  chronological_next_date = anchor$next_chronological_date,
  diary_date_gap_days = as.numeric(
    anchor$next_chronological_date - anchor$behavior_date
  ),
  chronology_exact = next_sleep_source_row == anchor$next_chronological_source &
    abs(as.numeric(current_wake_utc) - as.numeric(anchor$wake_utc)) < 0.5 &
    abs(
      as.numeric(next_sleepprep_utc) -
        as.numeric(anchor$next_chronological_sleep_utc)
    ) <
      0.5 &
    as.Date(behavior_date) == anchor$behavior_date,
  known_day_type = !is.na(day_type)
)]
stopifnot(!anyNA(cycle$chronology_exact), all(cycle$chronology_exact))
cycle[,
  date_link_eligible := is.finite(diary_date_gap_days) &
    diary_date_gap_days == 1
]
candidate <- data.table::copy(built$period[
  linkage_variant == "B_previous_sleep_wake_following_presleep" &
    raw_state %chin% c("wake", "pre-sleep")
])
index <- match(
  anchor_key(candidate$site, candidate$Id, candidate$behavior_source_row),
  anchor_key(cycle$site, cycle$Id, cycle$behavior_source_row)
)
stopifnot(!anyNA(index))
candidate[, `:=`(
  chronology_exact = cycle$chronology_exact[index],
  diary_date_gap_days = cycle$diary_date_gap_days[index],
  date_link_eligible = cycle$date_link_eligible[index]
)]
candidate[,
  exclusion_reason := data.table::fcase(
    !date_link_eligible,
    "missing_or_nonconsecutive_diary_date",
    is.na(day_type),
    "unknown_wake_anchor_day_type",
    valid_minutes == 0L,
    "no_valid_measurement",
    default = "eligible"
  )
]

period_key <- c(
  "site",
  "Id",
  "raw_state",
  "period_source_start",
  "period_source_end"
)
observed <- data.table::copy(built$minute[
  raw_state %chin% c("wake", "pre-sleep")
])
stopifnot(!anyDuplicated(observed[, .(site, Id, datetime_utc)]))
observed[, tick_numeric := as.numeric(datetime_utc)]
observed_bounds <- observed[,
  .(
    earliest_tick = min(tick_numeric),
    latest_tick = max(tick_numeric),
    projected_recount = .N,
    valid_recount = sum(valid_minute),
    yes_recount = sum(brown_check, na.rm = TRUE),
    integer_grid = all(is.finite(tick_numeric) & tick_numeric %% 60 == 0)
  ),
  by = period_key
]
timestamp_check <- merge(
  candidate,
  observed_bounds,
  by = period_key,
  all.x = TRUE,
  sort = FALSE
)
timestamp_check[,
  timestamp_and_count_exact := data.table::fifelse(
    projected_minutes == 0L,
    is.na(projected_recount),
    !is.na(projected_recount) &
      integer_grid &
      earliest_tick >= as.numeric(period_tick_start_utc) &
      latest_tick < as.numeric(period_tick_end_exclusive_utc) &
      projected_recount == projected_minutes &
      valid_recount == valid_minutes &
      yes_recount == brown_yes
  )
]
stopifnot(!anyNA(timestamp_check$timestamp_and_count_exact),
          all(timestamp_check$timestamp_and_count_exact))
lb_write_csv(timestamp_check, "placement/timestamp_count_checks.csv")
eligible <- candidate[exclusion_reason == "eligible"]
frame <- factorize_frame(
  eligible,
  allowed_states = unname(state_labels[c("wake", "pre-sleep")])
)
chest_design <- chest_env$make_chest_design(frame)
chest_c <- factorize_frame(chest$period[
  linkage_variant == variant_c & raw_state %chin% c("wake", "pre-sleep") &
  !is.na(day_type) & valid_minutes > 0L], allowed_states = unname(state_labels[c("wake", "pre-sleep")]))
chest_start_fit <- ba_fit_beta_binomial(ba_prepare_factor_frame(chest_c), ba_formula_chest, ba_dispersion_state)
chest_initial <- ba_boundary_initial_parameters(chest_design, initial_model = chest_start_fit)
chest_fit <- ba_lb_fit_candidate(chest_design, chest_initial, "BA-LB-CHEST-ANY", "B_chest")
brown_save_fit(chest_fit)
lb_save_rds(frame, "frames/chest.rds")
stopifnot(!chest_fit$fit_diagnostics$structural_failure)
chest_env$chest_design <- chest_fit$design_object
chest_env$parameter_list <- chest_fit$parameter_list
chest_env$covariance_fixed <- chest_fit$covariance_fixed
bundle <- chest_fit
covariance_fixed <- bundle$covariance_fixed
e15 <- chest_env$make_estimand_object(15L)
e30 <- chest_env$make_estimand_object(30L)
stopifnot(
  identical(names(e30$objective$par), names(bundle$optimizer$par)),
  max(abs(e30$objective$par - bundle$optimizer$par)) < 1e-10,
  all(is.finite(covariance_fixed)),
  min(eigen(covariance_fixed, symmetric = TRUE, only.values = TRUE)$values) > 0
)
point15 <- e15$objective$report(e15$objective$par)
point30 <- e30$objective$report(e30$objective$par)
uncertain <- TMB::sdreport(
  e30$objective,
  par.fixed = e30$objective$par,
  hessian.fixed = solve(covariance_fixed),
  getReportCovariance = TRUE
)
grid <- e30$grid
count <- nrow(grid)
stopifnot(
  count == 32L,
  nlevels(grid$site) == 8L,
  identical(
    names(uncertain$value),
    rep(
      c("cell_mean", "cell_pi_zero", "cell_pi_one", "cell_pi_beta", "cell_phi"),
      each = count
    )
  )
)
mean_estimate <- uncertain$value[seq_len(count)]
mean_covariance <- uncertain$cov[seq_len(count), seq_len(count), drop = FALSE]
stopifnot(
  all(is.finite(mean_covariance)),
  all(diag(mean_covariance) >= 0),
  max(abs(mean_covariance - t(mean_covariance))) < 1e-8
)
grid$cell_id <- seq_len(count)
cells <- cbind(
  grid,
  adherence = as.numeric(mean_estimate),
  standard_error = as.numeric(uncertain$sd[seq_len(count)])
)
cells$conf_low <- cells$adherence - stats::qnorm(0.975) * cells$standard_error
cells$conf_high <- cells$adherence + stats::qnorm(0.975) * cells$standard_error
states <- levels(grid$analysis_state)
contrast_matrix <- do.call(
  rbind,
  lapply(states, function(state) {
    as.numeric(grid$analysis_state == state & grid$day_type == "Free day") /
      8 -
      as.numeric(grid$analysis_state == state & grid$day_type == "Work day") / 8
  })
)
rownames(contrast_matrix) <- states
colnames(contrast_matrix) <- paste0("cell_", seq_len(count))
contrasts <- do.call(
  rbind,
  lapply(seq_along(states), function(i) {
    contrast <- contrast_matrix[i, ]
    estimate <- sum(contrast * mean_estimate)
    variance <- as.numeric(contrast %*% mean_covariance %*% contrast)
    stopifnot(length(variance) == 1L, is.finite(variance), variance >= 0)
    se <- sqrt(variance)
    data.frame(
      analysis_state = states[i],
      estimate = estimate,
      variance = variance,
      standard_error = se,
      conf_low = estimate - stats::qnorm(0.975) * se,
      conf_high = estimate + stats::qnorm(0.975) * se,
      sites = 8L,
      weighting = "equal_observed_eight_sites",
      placement = "chest",
      inferential_role = "sensitivity"
    )
  })
)
chest_comparison <- brown_compare(contrasts, primary_m1, "Complementary chest")
lb_write_csv(chest_comparison, "placement/chest_estimands/primary_comparison.csv")
lb_write_csv(cells, "placement/chest_estimands/cell_predictions.csv")
lb_save_rds(list(cells = cells, contrasts = contrasts, covariance = mean_covariance), "placement/chest_estimands/estimands.rds")
chest_diagnostics <- derive_diagnostics("B_chest", brown_file("models/BA-LB-CHEST-ANY.rds"))
for (name in names(chest_diagnostics)) if (is.data.frame(chest_diagnostics[[name]])) {
  lb_write_csv(chest_diagnostics[[name]], paste0("diagnostics/chest/", name, ".csv"))
}
```

## Local-calendar sensitivity

Each behavioural window is intersected with its true local calendar dates. Counts are reconstructed from the minute data at these intersections, keeping the parent window’s day type and cycle. The model adds a participant-calendar-day intercept to the beta-binomial comparison; uncertainty integrates the retained participant, cycle and calendar-day random effects.

```{r}
#| label: adherence-calendar
source("scripts/models/brown/calendar_counts.R")
parent_frame <- frames$B_any
counted <- lb_calendar_counts(parent_frame, coverage_near_eye)
lb_write_csv(counted$reconciliation, "calendar/count_reconciliation.csv")
lb_write_csv(counted$minute_identity, "calendar/minute_identity_checks.csv")
parents <- counted$parents
chunks <- counted$candidate_chunks
metadata_columns <- c(
  "parent_index",
  "linkage_variant",
  "site",
  "Id",
  "behavior_source_row",
  "behavior_date",
  "daytype_source_value",
  "day_type",
  "site_timezone",
  "raw_state",
  "period_source_start",
  "period_source_end",
  "period_start_utc",
  "period_end_utc",
  "period_tick_start_utc",
  "period_tick_end_exclusive_utc",
  "participant_id",
  "behavioral_day_id",
  "participant_state_id",
  "placement"
)
stopifnot(all(metadata_columns %in% names(parents)))
chunks <- merge(
  chunks,
  parents[, ..metadata_columns],
  by = "parent_index",
  all.x = TRUE,
  sort = TRUE
)
chunks[, `:=`(
  brown_no = valid_minutes - brown_yes,
  brown_fraction = data.table::fifelse(
    valid_minutes > 0,
    brown_yes / valid_minutes,
    NA_real_
  ),
  support_fraction = valid_minutes / expected_minutes,
  exact_zero = valid_minutes > 0 & brown_yes == 0,
  exact_one = valid_minutes > 0 & brown_yes == valid_minutes,
  participant_calendar_day_id = paste(site, Id, local_date, sep = "::"),
  exclusion_reason = data.table::fifelse(
    valid_minutes > 0L,
    "eligible",
    "no_valid_measurement"
  )
)]
chunks[,
  support_band := data.table::fcase(
    valid_minutes == 0L,
    "0_no_valid",
    support_fraction < 0.1,
    "1_gt0_lt10pct",
    support_fraction < 0.25,
    "2_10_to_lt25pct",
    support_fraction < 0.5,
    "3_25_to_lt50pct",
    support_fraction < 0.8,
    "4_50_to_lt80pct",
    default = "5_ge80pct"
  )
]
eligible <- data.table::copy(chunks[valid_minutes > 0L])
frame <- as.data.frame(eligible)
state_labels <- c(
  wake = "Wake outside the three hours before sleep",
  `pre-sleep` = "Pre-sleep",
  sleep = "Sleep environment"
)
frame$analysis_state <- factor(
  state_labels[frame$raw_state],
  levels = levels(parent_frame$analysis_state)
)
frame$site <- factor(frame$site, levels = levels(parent_frame$site))
frame$day_type <- factor(
  as.character(frame$day_type),
  levels = levels(parent_frame$day_type)
)
frame <- ba_prepare_factor_frame(frame)
calendar_frame <- frame
calendar_fit <- ba_fit_beta_binomial(calendar_frame, ba_formula_calendar, ba_dispersion_state)
calendar_assessment <- ba_assess_fit(calendar_fit, calendar_frame, "BA-LB-CALENDAR-CHUNKS",
  "B_calendar", "BA-CALENDAR-F3", "D0", NA_real_)
calendar_bundle <- list(model = calendar_fit, data = calendar_frame,
                         fit_diagnostics = calendar_assessment$fit_diagnostics, random_sd = calendar_assessment$random_sd)
lb_save_rds(calendar_bundle, "models/BA-LB-CALENDAR-CHUNKS.rds")
lb_write_csv(calendar_assessment$fit_diagnostics, "calendar/fit_diagnostics.csv")
if (!calendar_assessment$fit_diagnostics$hard_failure) {
  cells15 <- comparison_env$integrated_cells_glmm(calendar_fit, calendar_frame, 15L)
  cells30 <- comparison_env$integrated_cells_glmm(calendar_fit, calendar_frame, 30L)
  stopifnot(100 * max(abs(cells15$probability - cells30$probability)) <= .05)
  covariance <- cells30$gradient %*% cells30$variance_covariance %*% t(cells30$gradient)
  contrasts <- lapply(levels(calendar_frame$analysis_state), function(state) {
    w <- as.numeric(cells30$grid$analysis_state == state & cells30$grid$day_type == "Free day") / 9 -
      as.numeric(cells30$grid$analysis_state == state & cells30$grid$day_type == "Work day") / 9
    cbind(data.frame(analysis_state = state), mean_result(w, cells30$probability, covariance))
  })
  calendar_comparison <- brown_compare(do.call(rbind, contrasts), primary_m1, "Local calendar")
  lb_write_csv(calendar_comparison, "calendar/estimands/primary_comparison.csv")
  lb_save_rds(cells30, "calendar/estimands/cells.rds")
}
```

## Residual temporal dependence

Temporal sensitivity models represent dependence between a participant’s repeated measurements within a window using an Ornstein–Uhlenbeck process with actual elapsed days. The endpoint-inflated model is tried first. Only when it fails numerical or identifiability checks are the beta-binomial random-effects alternatives attempted in their specified order. Selection uses these model checks, and the chosen model is then integrated and diagnosed with the same seeds as the primary model.

```{r}
#| label: adherence-temporal
source("scripts/models/brown/temporal_models.R")
source("scripts/models/brown/temporal_estimands.R")
brown_compile("endpoint_inflated_betabinomial_ou")
brown_compile("endpoint_inflated_ou_estimands")
likelihood_validation <- lb_endpoint_ou$validate_exact_objective()
stopifnot(all(likelihood_validation$passed))
lb_write_csv(likelihood_validation, "likelihood/temporal_probability_checks.csv")
temporal_reports <- list()
temporal_comparisons <- list()
for (sample in c("ANY", "80")) {
  sample_id <- if (sample == "ANY") "B_any" else "B_80"
  input <- lb_temporal_prepare(frames[[sample_id]])
  lb_save_rds(input, paste0("frames/temporal_", sample, ".rds"))
  candidate_checks <- list()
  selected <- NULL
  for (route in c("ENDPOINT-R3", "BB-R0", "BB-R3")) {
    job <- list(model_id = paste0("BA-LB-TEMPORAL-", sample, "-", route),
                sample_id = paste0("B_temporal_", tolower(sample)), route = route)
    fit <- if (route == "ENDPOINT-R3") lb_temporal_fit_endpoint(input, primary_models[[sample_id]], job) else
      lb_temporal_fit_bb(input, job)
    path <- paste0("temporal/", sample, "/", route, "/model.rds")
    lb_save_rds(fit, path)
    candidate_checks[[route]] <- fit$fit_diagnostics
    lb_write_csv(fit$fit_diagnostics, paste0("temporal/", sample, "/", route, "/fit_diagnostics.csv"))
    if (!fit$fit_diagnostics$structural_failure) { selected <- fit; selected_path <- path; break }
  }
  lb_write_csv(rbindlist(candidate_checks, fill = TRUE), paste0("temporal/", sample, "/reporting/candidate_diagnostics.csv"))
  stopifnot(!is.null(selected))
  report <- lb_temporal_report_bundle(selected)
  temporal_reference_id <- if (sample == "ANY") "primary_any_valid" else "support_80"
  temporal_reference <- families$BA_M1[sample_id == temporal_reference_id]
  comparison <- brown_compare(report$contrasts, temporal_reference, paste0("Temporal ", sample, ": ", selected$route))
  temporal_reports[[sample]] <- report
  temporal_comparisons[[sample]] <- comparison
  lb_write_csv(comparison, paste0("temporal/", sample, "/reporting/primary_comparison.csv"))
  lb_save_rds(report, paste0("temporal/", sample, "/reporting/estimands.rds"))
  diagnostic_env <- new.env(parent = environment())
  sys.source("scripts/models/brown/diagnostics.R", diagnostic_env)
  diagnostic_env$conditional_predictions <- lb_temporal_predictions
  diagnostic <- diagnostic_env$derive_diagnostics(paste0("temporal_", sample), brown_file(selected_path))
  for (name in names(diagnostic)) if (is.data.frame(diagnostic[[name]])) {
    lb_write_csv(diagnostic[[name]], paste0("diagnostics/temporal_", sample, "/", name, ".csv"))
  }
}
```

## Reader tables and figures

The displays below use the same unrounded estimates exported to `results/csv/source_data/brown`. Their CSV files provide the values needed to verify the manuscript, while the fitted models and full covariance-based results remain in `results/models/brown`. All confidence intervals refer to the stated model and weighting scheme.

```{r}
#| label: adherence-reporting
source("scripts/models/brown/reporting.R")
source("scripts/models/brown/figures.R")
br_load_tables()
for (id in names(br_tables)) lb_write_csv(br_tables[[id]], paste0("table_", id, "_source.csv"))
dir.create("results/images/brown", recursive = TRUE, showWarnings = FALSE)
brown_figures <- brown_main_figures()
```

```{r}
#| label: tbl-adherence-levels
#| tbl-cap: "Equal-site marginal recommendation adherence by behavioural window and day type."
br_show("levels")
```

```{r}
#| label: tbl-adherence-primary
#| tbl-cap: "Primary Free-minus-Work contrasts, with a three-window false-discovery-rate family within each sample."
br_show("primary")
```

```{r}
#| label: tbl-adherence-coverage
#| tbl-cap: "Sensitivity of the primary contrasts to requiring at least 80% window coverage."
br_show("coverage")
```

```{r}
#| label: fig-adherence-levels
#| fig-alt: "Estimated recommendation adherence for daytime, pre-sleep and sleep, comparing work days with free days and showing 95% confidence intervals."
#| fig-cap: "Equal-site marginal recommendation adherence. Bars show 95% confidence intervals."
brown_figures[[1L]]
```

```{r}
#| label: fig-adherence-workday
#| fig-alt: "Work-day adherence estimates and 95% confidence intervals by site and behavioural window, with an equal-site reference."
#| fig-cap: "Work-day recommendation adherence by site and window, relative to the equal-site mean."
brown_figures[[2L]]
```

```{r}
#| label: fig-adherence-daytype
#| fig-alt: "Site-specific differences in adherence between free and work days, with 95% confidence intervals and false-discovery-rate markers."
#| fig-cap: "Site-specific Free-minus-Work differences in recommendation adherence."
brown_figures[[3L]]
```

```{r}
#| label: fig-adherence-coverage
#| fig-cap: "Free-minus-Work adherence contrasts under the any-valid and at least 80% coverage samples."
#| fig-alt: "Free-minus-Work adherence contrasts and 95% confidence intervals for daytime, pre-sleep and sleep, comparing the two coverage requirements."
brown_figures[[4L]]
```

```{r}
#| label: tbl-adherence-sensitivities
#| tbl-cap: "Changes in Free-minus-Work contrasts under the eligibility and threshold sensitivity analyses."
rbindlist(simple_comparisons, fill = TRUE)[, .(scenario, state_display, estimate, conf_low, conf_high,
  shift_percentage_points, direction_retained, interval_exclusion_retained)] |> gt()
```

```{r}
#| label: tbl-adherence-variance
#| tbl-cap: "Response-scale variance decomposition. These are point estimates."
final_d |> gt()
```

```{r}
#| label: tbl-adherence-shapley
#| tbl-cap: "Shapley allocation of fixed-cell variation among window, site and day type."
final_a |> gt()
```
