---
title: "Descriptive tables and figures"
subtitle: "Exact-layout replication with updated verified data and documented corrections"
engine: knitr
bibliography: ../bibliography.bib
format:
  html:
    toc: true
    toc-depth: 3
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

<style>
.cell-output-display { overflow-x: auto; }
.gt_table { margin-bottom: 0.75rem; }
.descriptive-callout { max-width: 95rem; }
.quarto-float-caption,
figure > figcaption,
.table-caption,
caption {
  text-align: left !important;
  margin-left: 0 !important;
}
</style>

## Scope and result

This document reproduces the submitted table styles, figure compositions, and
manuscript-scale canvases from the earlier manuscript-generating descriptive
analysis, but fills them with the updated verified data. It also repairs errors that
could not responsibly be copied forward. The main analytical measurement is
near-eye melEDI; chest measurement is shown only as a complementary description
and the two placements are never pooled.

The replicated layouts include the wide participant-and-site table, the wide
site-by-metric table with miniature distributions, the five-part study
overview, the 3 × 3 site profiles, the 4 × 4 metric grid, the four-panel
time-series explanation, the latitude–photoperiod diagnostic, and the
site-by-state light-context table.

```{r}
#| label: setup-descriptives

root <- Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
if (!nzchar(root)) root <- getwd()
root <- normalizePath(root, winslash = "/", mustWork = TRUE)

source(file.path(root, "scripts", "descriptives", "build_descriptives.R"))
descriptive_build <- load_descriptive_render_bundle(root)

read_descriptive_csv <- function(...) {
  readr::read_csv(
    file.path(root, ...),
    show_col_types = FALSE,
    progress = FALSE
  )
}

sample_contract <- read_descriptive_csv(
  "audit", "descriptives", "sample_count_contract.csv"
)
participant_table_data <- descriptive_build$tables[[
  "participant_site_characteristics_replica.csv"
]]
participant_manuscript_table_data <- descriptive_build$tables[[
  "participant_site_characteristics_manuscript_replica.csv"
]]
metric_table_data <- descriptive_build$tables[[
  "metric_descriptive_summary_replica.csv"
]]
recommendation_table_data <- descriptive_build$tables[[
  "recommendation_context_replica.csv"
]]
metric_plot_values <- descriptive_build$sources[["metric_plot_values.csv"]]
site_display <- descriptive_site_display_registry()
figure_alt <- descriptive_build$sources[["figure_alt_text.csv"]]
artifact_manifest <- descriptive_build$manifest

get_alt <- function(figure_id) {
  figure_alt$short_alt_text[match(figure_id, figure_alt$figure_id)]
}

include_descriptive_figure <- function(filename) {
  source_path <- file.path(
    root, "artifacts", "10_figures", "descriptives", filename
  )
  input_file <- knitr::current_input(dir = TRUE)
  if (!length(input_file) || !nzchar(input_file)) {
    stop("Cannot resolve the source document path for a descriptive figure")
  }
  source_relative_path <- xfun::relative_path(
    source_path,
    dirname(input_file),
    error = FALSE
  )
  knitr::include_graphics(
    source_relative_path,
    rel_path = FALSE,
    error = FALSE
  )
}
```

All 12 required shared manifests matched their pinned SHA-256 fingerprints
before any analytical input was read. The current display registry is
`config/site_display_registry.csv`; it supplies the labels, visible order, and
colours in every table and figure. The visible site sequence is
`r paste(site_display$display_name, collapse = ", ")`.

## Updated sample

The normalized participant roster contains **191 participants**. After the
pre-specified ≥80% full-day completeness rule and exact-all-zero screen, the
main near-eye dataset contains **141 participants, 816 participant-days, and
1,175,160 one-minute real observations**. The complementary chest dataset
contains **154 participants, 902 participant-days, and 1,298,880 one-minute
real observations**. The paired subset contains **112 participants and 643
participant-days**.

Two of 818 otherwise eligible near-eye participant-days and three of 905
otherwise eligible chest participant-days were exact-all-zero melEDI days and
were excluded. A participant was not removed merely because one of their days
was excluded.

```{r}
#| label: tbl-descriptive-sample-flow
#| tbl-cap: "Updated descriptive sample after each fixed coverage step."

sample_contract |>
  dplyr::mutate(
    dplyr::across(
      c(participants, participant_days, one_minute_real_observations),
      ~ ifelse(is.na(.x), "Not applicable", format(.x, big.mark = ","))
    )
  ) |>
  dplyr::rename(
    Stage = stage,
    Placement = placement,
    Participants = participants,
    `Participant-days` = participant_days,
    `One-minute real observations` = one_minute_real_observations
  ) |>
  knitr::kable(align = c("l", "l", "r", "r", "r"))
```

## Table 1: participant and site characteristics

@tbl-participant-site keeps the earlier wide site-column organization. Cells
now identify whether a quantity belongs to near-eye, chest, paired, roster, or
diary data. Participant information covers the complete normalized roster.
Collection dates and civil photoperiod use the union of every recorded
non-all-zero participant-day across both placements; screened-day counts
retain the main-analysis completeness rule.

```{r}
#| label: tbl-participant-site
#| tbl-cap: "Participant, site, collection, sleep, chronotype, and wear characteristics using updated verified data."

participant_gt <- build_participant_site_publication_gt(participant_table_data)
participant_gt
```

### Manuscript Table 1 variant

The rendered submitted manuscript used the reduced `tbl1_red` variant rather
than the expanded table above. The following table reproduces that exact row
selection with the updated data and corrections. As in the submitted variant,
it omits Morningness–Eveningness Questionnaire score, gender, collection
dates, weekday/weekend counts, workday/free-day diaries, social jetlag, and
sleep duration; all remaining rows retain their original order and styling.

```{r}
#| label: tbl-participant-site-manuscript
#| tbl-cap: "Submitted-manuscript Table 1 row selection reproduced with updated verified data."

participant_manuscript_gt <-
  build_participant_site_manuscript_publication_gt(participant_table_data)
participant_manuscript_gt
```

Important corrections embodied in @tbl-participant-site are:

- MPI is displayed as Tübingen and TUM as Munich, repairing their interchanged
  institution labels in the earlier table.
- Participant and participant-day rows show the whole available roster plus
  near-eye, chest, and paired subsets. Participant time retains the screened
  near-eye basis and is limited to its two largest time units.
- Declared non-wear is reported for near-eye only; the earlier chest display
  reused the near-eye input and has therefore been removed.
- Screened days show only the final result after the completeness and all-zero
  screens. Civil photoperiod comes from verified site-and-solar context across
  the whole available roster.
- Clock-time chronotype summaries are circular. During rebuilding, a
  `quantile()` argument typo in the first implementation was detected,
  documented, and corrected before this final table was generated.

## Table 2: metric distributions

@tbl-near-eye-metrics reproduces the old metric categories, order, ten site
columns, scaling column, and enlarged miniature distribution column. Every
numerical cell states `N`, the number of participants, and `d`, the number of
participant-days with a finite value. The site columns have equal widths and
each median, interval, mean ± SD, and size line is kept intact.

```{r}
#| label: tbl-near-eye-metrics
#| tbl-cap: "Near-eye melEDI metric distributions overall and by site using updated verified metric artifacts."

metric_gt <- build_metric_publication_gt(metric_table_data, metric_plot_values)
metric_gt
```

The numerical values in @tbl-near-eye-metrics come from the approved metric
artifacts; the table does not recalculate any metric. Clock-time means and
middle-50% intervals are circular, rather than treating 23:59 and 00:01 as far
apart. The miniature plots are redundant visual summaries: the adjacent cells
remain the authoritative values and provide the exact denominators.

## Figure 1: study overview

The study overview retains the submitted five-part composition: protocol,
site map, collection timing, photoperiod coverage, and a repeated 48-hour
near-eye profile. The second 24-hour cycle is a display duplicate, not
additional data.
Collection spans are shown as rectangles and split only when a site has at
least six consecutive dates without any available data.

```{r}
#| label: fig-descriptive-overview
#| fig-cap: "Study protocol, locations, collection dates, observed civil photoperiod, and the pooled main near-eye 48-hour double profile."
#| fig-alt: "Five-panel study overview with the protocol, world map, roster-wide collection dates, site photoperiod distributions, and a repeated 15-minute 48-hour near-eye melEDI profile."
#| out-width: "100%"

include_descriptive_figure("descriptive_overview.png")
```

The black and site-coloured profiles use `LightLogR::aggregate_Datetime()` in
15-minute floor bins after removing participant grouping; the numeric handler
is the median with `na.rm = TRUE`. The grey ribbon is the central 67% of
eligible one-minute values in each overall bin. It is a value interval, not a
confidence interval. The red band shows average diary sleep and the blue
shading shows average civil night; declared non-wear is not displayed. The
melEDI axis uses `LightLogR::symlog_trans(base = 10, thr = 1, scale = 1)`.

## Figure 2: site-specific profiles

The main near-eye profiles retain the earlier 3 × 3 double-plot layout and
registered site order. Each site uses the same 15-minute
`aggregate_Datetime()` median with nested central 50%, 75%, and 95% value
intervals. The interval legend shows the corresponding ribbon opacities.

```{r}
#| label: fig-near-eye-site-profiles
#| fig-cap: "Pooled main near-eye melEDI profiles by site, with each 24-hour cycle repeated once."
#| fig-alt: "Nine labelled site panels repeat pooled 15-minute near-eye 24-hour melEDI profiles across 48 hours with nested central 50%, 75%, and 95% value bands and average civil-night and sleep periods."
#| out-width: "100%"

include_descriptive_figure("near_eye_site_profiles.png")
```

The complementary chest profile uses the same visual grammar in a complete
four-by-two layout. Tübingen is
omitted because the updated chest main dataset contains no eligible MPI site
panel; the remaining sites retain their registered order and colours.

```{r}
#| label: fig-chest-site-profiles
#| fig-cap: "Complementary chest-level melEDI profiles by available site, with each 24-hour cycle repeated once."
#| fig-alt: "Eight labelled site panels repeat complementary 15-minute chest-level 24-hour melEDI profiles across 48 hours with nested central 50%, 75%, and 95% value bands and average civil-night and sleep periods. Tübingen is absent because there are no eligible chest main days at MPI."
#| out-width: "100%"

include_descriptive_figure("chest_site_profiles.png")
```

Chest values describe the environment at the chest sensor. They are not
treated as equivalent to near-eye or ocular exposure.

## Figure 3: reader-facing metric distributions

@fig-near-eye-metric-distributions reproduces the old 4 × 4 ordering and
ridge-plus-box construction. Site names remain visible in the first column of
panels, so colour is not the only way to identify a distribution.

```{r}
#| label: fig-near-eye-metric-distributions
#| fig-cap: "Sixteen familiar near-eye melEDI metric distributions by site."
#| fig-alt: "A four-by-four grid of duration, dynamics, dose, level, spectrum, and circular timing metric distributions. Each panel contains labelled site ridges and white boxplots; red lines mark site medians."
#| out-width: "100%"

include_descriptive_figure("near_eye_metric_distributions.png")
```

## Figure 4: from time series to a daily metric

The submitted explanatory figure used seven fixed German-site examples over
five aligned study days. @fig-time-series-to-metrics retains the same IDs and
the same submitted study days 2–6, then orders participants from top to bottom by their
median TAT250 across those days. The source CSV preserves each participant’s
actual local dates, while the visible x-axis retains the submitted Wednesday
to Sunday alignment.

```{r}
#| label: fig-time-series-to-metrics
#| fig-cap: "The same seven manuscript examples link stored 30-minute near-eye melEDI series from the gap-timing-unaware dataset to daily duration above 250 lx during daytime."
#| fig-alt: "Four panels connect five protocol weekdays of stored gap-timing-unaware near-eye light samples for seven fixed participants to TAT250 values calculated from those same displayed daytime samples, using participant boxplots, weekday boxplots, and a participant-by-weekday bubble grid."
#| out-width: "100%"

include_descriptive_figure("time_series_to_metrics.png")
```

Panel A uses the stored floor-aligned 30-minute near-eye arithmetic means from
the gap-timing-unaware dataset. Panels B–D recalculate daily TAT250 from
exactly those displayed daytime samples with
`LightLogR::duration_above_threshold()`. This figure illustrates the fixed
examples and does not estimate a population weekday effect.

## Figure 5: latitude and observed photoperiod

The latitude diagnostic is retained as supplementary context based on the 816
main near-eye participant-days. Its black impossible-region ribbons consume
the verified theoretical bounds from the current H1 reporting artifact, so
the panel follows the submitted and current H1 visual grammar without a new
calculation in the descriptive pipeline. The density curves and observations
are drawn first and the black feasibility curtains are drawn over them.

```{r}
#| label: fig-latitude-photoperiod
#| fig-cap: "Observed verified civil photoperiod and verified theoretical bounds by absolute latitude for main near-eye participant-days."
#| fig-alt: "Observed civil photoperiod distributions are plotted against absolute study-site latitude for 816 main near-eye participant-days, with black ribbons marking impossible photoperiods and a registered-colour legend identifying the nine sites."
#| out-width: "100%"

include_descriptive_figure("latitude_photoperiod_diagnostic.png")
```

## Light exposure in recommendation contexts

@tbl-recommendation-context preserves the earlier site-row and state-column
organization. The thresholds are wake ≥250 lx melEDI, before sleep ≤10 lx
melEDI, and sleep ≤1 lx melEDI [@brown2022]. The results are deliberately
described as *fractions of valid measured minutes in a recommended range*, not
as adherence, compliance, or evidence of a biological response.

```{r}
#| label: tbl-recommendation-context
#| tbl-cap: "Fractions of updated main near-eye minutes in the Brown et al. light-context ranges, with explicit denominators."

recommendation_gt <- build_recommendation_publication_gt(
  recommendation_table_data
)
recommendation_gt
```

Each percentage is followed by one grey `relevant / all` count. Wake,
before-sleep, and sleep cells divide minutes within the
recommended context by all valid one-minute observations in that diary state;
the combined column pools all classified valid minutes. The four state-share
columns divide minutes in the displayed state by all eligible real minutes.
Diary-defined sleep combines the
near-eye sensor while awake with a bedside sensor during sleep; the sleep row
therefore describes the sleep environment rather than direct ocular exposure.

## Exact comparison with the earlier render

The earlier HTML was read only as a record of displayed cells; no old `.RData`
workspace was loaded. Machine-readable comparisons cover every displayed
publication-table cell:

```{r}
#| label: tbl-descriptive-comparison-summary
#| tbl-cap: "Complete cell-level comparison with the earlier manuscript-generating tables."

comparison_files <- c(
  participant = "previous_table1_comparison.csv",
  metric = "previous_table2_comparison.csv",
  recommendation = "previous_recommendation_comparison.csv"
)
comparison_summary <- dplyr::bind_rows(lapply(names(comparison_files), function(name) {
  data <- read_descriptive_csv(
    "audit", "descriptives", comparison_files[[name]]
  )
  data |>
    dplyr::count(.data$comparison_status, name = "cells") |>
    dplyr::mutate(table = name, .before = 1L)
}))
comparison_summary |>
  dplyr::rename(
    Table = table,
    `Comparison status` = comparison_status,
    Cells = cells
  ) |>
  knitr::kable(align = c("l", "l", "r"))
```

The comparison summary above reports all current cells and classifies each
against the submitted render. Metric timing rows also changed to circular
summaries, while the contextual table reflects the updated eligible-day and
state denominators.

Every numerical cell, plus the explanation assigned to its difference, is in:

- `audit/descriptives/previous_table1_comparison.csv`
- `audit/descriptives/previous_table2_comparison.csv`
- `audit/descriptives/previous_recommendation_comparison.csv`

Every old table, figure, panel, and batch export is covered by
`audit/descriptives/output_difference_explanations.csv`. Redundant reduced
tables, letter-only metric exports, and individual site PDFs were not copied
as separate deliverables because they add no distinct construct; their values
and panels remain in the full replicated tables, composite figures, and exact
source CSVs.

### Stand-alone visual exports

The four publication tables are exported as PNG files with their registered
`gtsave()` viewport widths. Each of the six durable figures uses its registered,
content-specific canvas and export scale to create a 300-dpi PNG, a
maximum-quality JPEG, a tightly bounded one-page PDF, and an SVG. The final
files contain the figure only, not an A4 paper canvas. Each figure is then
checked at its intended publication width; main text normally resolves to
5--7 pt, central text to 7--9 pt, and only repeated non-essential detail may
use 3.5--5 pt. The following record gives the
retained original and rebuilt PNG paths, dimensions, and hashes needed for
direct side-by-side review:

```{r}
#| label: tbl-descriptive-visual-export-comparison
#| tbl-cap: "Original and rebuilt visual exports available for side-by-side comparison."

visual_export_comparison <- read_descriptive_csv(
  "audit", "descriptives", "visual_export_comparison.csv"
)
visual_export_comparison |>
  dplyr::transmute(
    Output = output_id,
    Type = output_type,
    Original = original_path,
    Rebuilt = rebuilt_path,
    `Original pixels` = paste0(original_width_px, " × ", original_height_px),
    `Rebuilt pixels` = paste0(rebuilt_width_px, " × ", rebuilt_height_px)
  ) |>
  knitr::kable(align = c("l", "l", "l", "l", "r", "r"))
```

The completed visual assessment, including intended visible differences, is
recorded in `audit/descriptives/visual_export_review.md`.

## Reproducibility and output contract

Each durable figure has a fixed filename and dimensions, 300-dpi PNG and JPEG
exports, a deterministic one-page PDF, an SVG vector master, accurate alt
text, and an exact source-data mapping with source SHA-256 values. The PDF
places the verified 300-dpi PNG on a page of the same physical dimensions so
that Unicode labels and the inspected layout are preserved exactly while the
file remains byte-reproducible. The complete output manifest is
`artifacts/12_manifests/descriptives/descriptive_artifacts.csv`; figure-to-source
mapping is `artifacts/12_manifests/descriptives/figure_source_data_map.csv`.
The DISPLAY-001 registry provenance, including its hash, is recorded in
`audit/descriptives/site_display_registry_provenance.csv`.
Figure 1 is the REPORT-011-PILOT-001 typography reference. Its approved
source-level text sizes are the default 14-pt cowplot base, 12-pt tick labels,
14-pt axis titles and tags, 11-pt caption, and 3-unit map labels. Approval of
these values does not close the figure-specific detail QA. The 10.5-by-10-inch
base
canvas is exported with the literal `ggplot2::ggsave(scale = 1.5)` argument to
15.75 by 15 inches. This external asset is displayed at `out-width: 100%`. A
direct-knitr equivalent would use `fig-width: 15.75`, `fig-height: 15`, and the
same separate `out-width`, without changing fonts or marks. Figures 2–5 retain
their content-specific export geometry. The export/display calculation and
optional A4 portrait QA mock-up with 20-mm margins are recorded in
`audit/descriptives/figure_readability_qa.csv`. The mock-up is an inspection
artifact only; final figure files are tightly bounded to the figure and are not
saved on an A4 canvas.

The four table PNG export settings are recorded in
`artifacts/12_manifests/descriptives/table_export_specifications.csv`.

```{r}
#| label: tbl-descriptive-figure-contract
#| tbl-cap: "Deterministic dimensions and formats for the replicated descriptive figures."

read_descriptive_csv(
  "artifacts", "12_manifests", "descriptives", "figure_specifications.csv"
) |>
  dplyr::select(
    figure_id, base_width_in, base_height_in,
    export_scale_multiplier, export_width_in, export_height_in,
    html_out_width, print_display_width_mm, display_reduction_factor,
    nominal_min_essential_text_pt, effective_min_essential_text_pt,
    physical_size_qa
  ) |>
  dplyr::rename(
    Figure = figure_id,
    `Base width (in)` = base_width_in,
    `Base height (in)` = base_height_in,
    `Export scale` = export_scale_multiplier,
    `Export width (in)` = export_width_in,
    `Export height (in)` = export_height_in,
    `HTML out-width` = html_out_width,
    `Print width (mm)` = print_display_width_mm,
    `Display reduction` = display_reduction_factor,
    `Nominal minimum text (pt)` = nominal_min_essential_text_pt,
    `Effective minimum text (pt)` = effective_min_essential_text_pt,
    `Physical-size QA` = physical_size_qa
  ) |>
  knitr::kable(
    align = c("l", "r", "r", "r", "r", "r", "l", "r", "r", "r", "r", "l")
  )
```

The complete descriptive build is run with:

```text
RENV_CONFIG_SANDBOX_ENABLED=FALSE Rscript scripts/descriptives/run_descriptives.R
```

This document alone is rendered with:

```text
quarto render notebooks/descriptives.qmd --profile nathealth
```

No hypothesis model, inferential p-value, or exploratory H03/H04 analysis is
created or duplicated here.
