---
title: "H07 analysis preparation and provenance"
subtitle: "From verified participant-day metric frames to nonlinear photoperiod curves and qualifying transitions"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-h07-preparation
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(knitr)
  library(readr)
  library(stringr)
  library(tibble)
  library(tidyr)
})

root_candidates <- unique(c(
  normalizePath(
    Sys.getenv("QUARTO_PROJECT_DIR", unset = getwd()),
    winslash = "/",
    mustWork = FALSE
  ),
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(
    file.path(getwd(), "../../.."),
    winslash = "/",
    mustWork = FALSE
  )
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(
        path,
        "scripts/hypotheses/H07/build_h07_preparation_artifacts.R"
      ))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(
  root.dir = file.path(root, "audit", "hypotheses", "H07")
)

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H07 preparation requires R 4.6.1", call. = FALSE)
}

read_h07_csv <- function(relative_path) {
  readr::read_csv(file.path(root, relative_path), show_col_types = FALSE)
}

prep_gt <- function(table, font_size = 12) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

format_count <- function(value) {
  format(value, big.mark = ",", scientific = FALSE, trim = TRUE)
}

format_range <- function(lower, upper, digits = 2) {
  ifelse(
    lower == upper,
    format(round(lower, digits), nsmall = digits),
    paste0(
      format(round(lower, digits), nsmall = digits),
      "–",
      format(round(upper, digits), nsmall = digits)
    )
  )
}

source_dir <- "artifacts/11_source_data/H07/preparation"
table_dir <- "artifacts/09_tables/H07"

input_identities <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_input_identities.csv"
))
frame_integrity <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_frame_integrity.csv"
))
sample_support <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_sample_support.csv"
))
metric_contract <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_metric_contract.csv"
))
site_support <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_site_support.csv"
))
site_ranges <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_site_photoperiod_ranges.csv"
))
diagnostic_summary <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_diagnostic_summary.csv"
))
pattern_summary <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_pattern_summary.csv"
))
formula_registry <- read_h07_csv(file.path(
  source_dir,
  "H07_preparation_formula_registry.csv"
))

derivative_settings <- read_h07_csv(file.path(
  table_dir,
  "H07_revised_derivative_settings.csv"
))
sensitivity_results <- read_h07_csv(file.path(
  table_dir,
  "H07_revised_sensitivity_plateau_comparison.csv"
))
sensitivity_samples <- read_h07_csv(file.path(
  table_dir,
  "H07_sensitivity_samples.csv"
))
sensitivity_registry <- read_h07_csv(file.path(
  table_dir,
  "H07_sensitivity_run_registry.csv"
))
model_form_results <- read_h07_csv(file.path(
  table_dir,
  "H07_revised_model_form_plateau_comparison.csv"
))
loso_results <- read_h07_csv(file.path(
  table_dir,
  "H07_revised_loso_plateau_influence_summary.csv"
))
tweedie_pilot <- read_h07_csv(file.path(
  table_dir,
  "H07_main_tweedie_distribution_pilot.csv"
))
residual_summary <- read_h07_csv(file.path(
  table_dir,
  "H07_main_residual_distribution_summary.csv"
))
session_record <- read_h07_csv(file.path(
  table_dir,
  "H07_session.csv"
))
figure_settings <- read_h07_csv(file.path(
  table_dir,
  "H07_revised_paired_figure_settings.csv"
))
metric011_summary <- read_h07_csv(
  "artifacts/12_manifests/H07/H07_METRIC-011_reconciliation_summary.csv"
)
metric011_comparison <- read_h07_csv(
  "artifacts/12_manifests/H07/H07_METRIC-011_scientific_comparison.csv"
)
site_registry <- read_h07_csv("config/site_display_registry.csv")

placement_levels <- c("near_eye", "chest")
placement_labels <- c(
  near_eye = "Near eye — primary",
  chest = "Chest — complementary"
)

frame_summary <- frame_integrity |>
  group_by(.data$placement) |>
  summarise(
    `Metric frames` = n(),
    `Identity checks passed` = sum(.data$frame_identity == "PASS"),
    `Schema checks passed` = sum(.data$schema_complete),
    `Unique-key checks passed` = sum(.data$duplicate_row_keys == 0L),
    `Response reconstructions passed` = sum(
      .data$response_reconstruction_maximum_absolute_error < 1e-12
    ),
    `Latitude fixed within site` = all(.data$latitude_constant_within_site),
    .groups = "drop"
  ) |>
  mutate(
    Placement = unname(placement_labels[.data$placement]),
    .before = 1L
  ) |>
  select(-.data$placement)

sample_ranges <- sample_support |>
  group_by(.data$placement, .data$placement_label) |>
  summarise(
    Participants = format_range(
      min(.data$participants),
      max(.data$participants),
      digits = 0
    ),
    `Participant-days` = format_range(
      min(.data$participant_days),
      max(.data$participant_days),
      digits = 0
    ),
    Observations = format_range(
      min(.data$observations),
      max(.data$observations),
      digits = 0
    ),
    Sites = format_range(min(.data$sites), max(.data$sites), digits = 0),
    `Civil photoperiod (h)` = format_range(
      min(.data$photoperiod_min),
      max(.data$photoperiod_max),
      digits = 2
    ),
    `Collection dates` = paste0(
      min(.data$first_date),
      " to ",
      max(.data$last_date)
    ),
    .groups = "drop"
  ) |>
  transmute(
    Placement = .data$placement_label,
    .data$Participants,
    .data$`Participant-days`,
    .data$Observations,
    .data$Sites,
    .data$`Civil photoperiod (h)`,
    .data$`Collection dates`
  )

sensitivity_labels <- tibble::tribble(
  ~run_id, ~Scenario,
  "paired__near_eye", "Paired/common sample — near eye",
  "paired__chest", "Paired/common sample — chest",
  "gap_timing_unaware__near_eye", "Gap-timing-unaware dataset — near eye",
  "gap_timing_unaware__chest", "Gap-timing-unaware dataset — chest",
  "prep_common_primary__near_eye", "Cross-dataset common rows, primary dataset — near eye",
  "prep_common_primary__chest", "Cross-dataset common rows, primary dataset — chest",
  "prep_common_gap__near_eye", "Cross-dataset common rows, gap-timing-unaware — near eye",
  "prep_common_gap__chest", "Cross-dataset common rows, gap-timing-unaware — chest",
  "exact_period__near_eye", "Exactly identified longest-period rows — near eye",
  "exact_period__chest", "Exactly identified longest-period rows — chest",
  "dose_common_corrected__near_eye", "Corrected melEDI dose on common rows — near eye",
  "dose_common_corrected__chest", "Corrected melEDI dose on common rows — chest",
  "dose_common_observed__near_eye", "Observed melEDI dose on common rows — near eye",
  "dose_common_observed__chest", "Observed melEDI dose on common rows — chest"
)

sensitivity_sample_summary <- sensitivity_samples |>
  left_join(sensitivity_labels, by = "run_id", relationship = "many-to-one") |>
  group_by(.data$run_id, .data$Scenario) |>
  summarise(
    Metrics = n(),
    Participants = format_range(
      min(.data$participants),
      max(.data$participants),
      digits = 0
    ),
    `Participant-days/observations` = format_range(
      min(.data$participant_days),
      max(.data$participant_days),
      digits = 0
    ),
    Sites = format_range(min(.data$sites), max(.data$sites), digits = 0),
    .groups = "drop"
  ) |>
  select(-.data$run_id)

sensitivity_classification <- sensitivity_results |>
  group_by(
    .data$data_scenario,
    .data$placement,
    .data$sensitivity
  ) |>
  summarise(
    Evaluable = sum(!is.na(.data$classification_agrees)),
    Agreements = sum(.data$classification_agrees, na.rm = TRUE),
    changed = paste(
      .data$manuscript_name[
        !is.na(.data$classification_agrees) &
          !.data$classification_agrees
      ],
      collapse = "; "
    ),
    .groups = "drop"
  ) |>
  mutate(
    Analysis = case_when(
      .data$sensitivity == "gap_timing_unaware" ~
        "Gap-timing-unaware dataset",
      .data$sensitivity == "paired_common_placement" ~
        "Paired/common placement sample",
      .data$sensitivity == "preparation_exact_common" &
        .data$data_scenario == "primary" ~
        "Cross-dataset common rows, primary dataset",
      .data$sensitivity == "preparation_exact_common" ~
        "Cross-dataset common rows, gap-timing-unaware dataset",
      .data$sensitivity == "longest_period_exact_only" ~
        "Exactly identified longest-period rows",
      .data$sensitivity == "observed_dose_common_sample" ~
        "Observed/corrected melEDI dose on common rows",
      TRUE ~ .data$sensitivity
    ),
    Placement = unname(placement_labels[.data$placement]),
    Agreement = paste0(.data$Agreements, "/", .data$Evaluable),
    `Changed classifications` = if_else(
      .data$changed == "",
      "None",
      .data$changed
    )
  ) |>
  select(
    .data$Analysis,
    .data$Placement,
    .data$Agreement,
    .data$`Changed classifications`
  )

model_form_summary <- model_form_results |>
  group_by(.data$model_id, .data$placement) |>
  summarise(
    Evaluable = sum(!is.na(.data$classification_agrees)),
    Agreements = sum(.data$classification_agrees, na.rm = TRUE),
    `Failed fits` = sum(grepl("^FAIL", .data$fit_status)),
    changed = paste(
      .data$manuscript_name[
        !is.na(.data$classification_agrees) &
          !.data$classification_agrees
      ],
      collapse = "; "
    ),
    .groups = "drop"
  ) |>
  mutate(
    Model = recode(
      .data$model_id,
      adapted_photoperiod_expanded_basis =
        "Broader photoperiod smooth (k = 10)",
      adapted_photoperiod_fixed_site = "Fixed site effects"
    ),
    Placement = unname(placement_labels[.data$placement]),
    Agreement = paste0(.data$Agreements, "/", .data$Evaluable),
    `Changed classifications` = if_else(
      .data$changed == "",
      "None",
      .data$changed
    )
  ) |>
  select(
    .data$Model,
    .data$Placement,
    .data$Agreement,
    .data$`Failed fits`,
    .data$`Changed classifications`
  )

site_display_lookup <- stats::setNames(
  site_registry$display_name,
  site_registry$site
)

display_omissions <- function(value) {
  vapply(
    value,
    function(one) {
      if (is.na(one) || one == "") {
        return("None")
      }
      codes <- strsplit(one, "; ", fixed = TRUE)[[1L]]
      labels <- unname(site_display_lookup[codes])
      labels[is.na(labels)] <- codes[is.na(labels)]
      paste(labels, collapse = "; ")
    },
    character(1L)
  )
}

stopifnot(
  nrow(input_identities) == 4L,
  all(input_identities$identity_status == "PASS"),
  nrow(frame_integrity) == 18L,
  all(frame_integrity$overall_status == "PASS"),
  nrow(sample_support) == 18L,
  nrow(metric_contract) == 9L,
  nrow(site_support) == 153L,
  nrow(site_ranges) == 17L,
  nrow(pattern_summary) == 18L,
  sum(
    pattern_summary$placement == "near_eye" &
      pattern_summary$derivative_defined_pattern
  ) == 6L,
  sum(
    pattern_summary$placement == "chest" &
      pattern_summary$derivative_defined_pattern
  ) == 7L,
  nrow(formula_registry) == 9L,
  nrow(figure_settings) == 2L,
  all(figure_settings$width_in == 9),
  all(figure_settings$height_in == 18),
  all(figure_settings$dpi == 270L),
  all(metric011_summary$status == "PASS"),
  metric011_summary$observed[
    metric011_summary$check_id == "SCIENTIFIC_CONCLUSION_CHANGED"
  ] == "FALSE",
  !any(metric011_comparison$changed)
)
```

## Scientific target

This companion documents the analysis underlying the
[H07 results report](../../../notebooks/hypotheses/H07.qmd). The result report's
[preregistration-deviation section](../../../notebooks/hypotheses/H07.qmd#h07-preregistration-deviations)
provides reciprocal navigation to the current scientific deviations.

The preregistered hypothesis was:

> “H7: There is a ceiling (nonlinear) effect of absolute latitude and photoperiod on level-, duration-, and exposure-history-based light metrics.”

**Melanopic equivalent daylight illuminance (melEDI)** is an illuminance-like
measure weighted for melanopic photoreception and expressed in lux (lx). The
**primary near-eye sensor position** places the wearable sensor near the eyes;
the **complementary chest sensor position** is torso-worn and is not ocular
exposure. A **participant-day** is one participant's metric value on one local
calendar date at one sensor position.

A **nonlinear generalized additive model (GAM) analysis** allows the
photoperiod association to bend rather than follow a straight line. The
**derivative** is the slope of the fitted curve: the change in the fitted
outcome for each additional hour of photoperiod at each point. H07 evaluates
it on the model's linear-predictor scale, in model-scale units per hour. A
**pointwise 95% confidence interval (95% CI)** describes uncertainty at one
photoperiod value; it is not a simultaneous band for the whole curve. **Overlap
among nonlinear model terms** makes their separate contributions harder to
distinguish; the technical term for this is *concurvity*.

The design cannot separate absolute latitude from site, because each site has
one latitude. The estimable target is therefore narrower: for each of nine
planned metrics and each placement, describe the pooled association with civil
photoperiod and ask whether the fitted smooth changes from a pointwise detected
increase to a first derivative whose pointwise interval contains zero at every
remaining grid point through the longest recorded photoperiod. This is a
qualifying transition from increase to a sustained near-flat tail.[^h07-prep-formal]
It is not an equivalence test and does not establish a physiological or
environmental ceiling.

[^h07-prep-formal]: The stored formal classification is
    **derivative-defined plateau pattern**. It is the three-part pointwise rule
    documented below, not a biological-ceiling, permanent-plateau, or
    equivalence claim.

### Current preregistration deviations {#h07-preparation-preregistration-deviations}

- [`DEV-016`](../../../notebooks/preregistration_deviations.qmd#dev-016): the
  registered latitude-by-photoperiod surface is non-identifiable here, so the
  accepted report retains all nine outcomes and presents a separately labelled
  site-adjusted nonlinear photoperiod analysis as descriptive evidence.
- [`DEV-033`](../../../notebooks/preregistration_deviations.qmd#dev-033): all
  nine eligible outcomes remain visible rather than being selected using H01
  significance or H07 AIC.
- [`DEV-034`](../../../notebooks/preregistration_deviations.qmd#dev-034):
  response specifications and model checks are metric-specific rather than one
  universal zero-aware log-Gaussian model.

```{r}
#| label: fig-h07-preparation-map
#| echo: false
#| results: asis

cat("```{mermaid}\n")
cat("flowchart TD\n")
cat("  A[\"Four verified participant-day and solar inputs\"] --> B[\"Nine metric-specific frames per placement\"]\n")
cat("  B --> C[\"Separate near-eye and chest nonlinear GAMs\"]\n")
cat("  C --> D[\"Pooled civil-photoperiod smooth\"]\n")
cat("  D --> E[\"100-point first-derivative grid\"]\n")
cat("  E --> F[\"Qualifying transition to a sustained zero-compatible tail\"]\n")
cat("  B --> G[\"Identity, sample, site, date, and support checks\"]\n")
cat("  C --> H[\"Model checks, including overlap among nonlinear terms (concurvity)\"]\n")
cat("  F --> I[\"Common-row, gap, metric-definition, model-form, and site-omission sensitivities\"]\n")
cat("```\n")
```

```{r}
#| label: tbl-h07-preparation-contributions
#| tbl-cap: "What each analytical component contributes to H07"

tibble::tribble(
  ~Component, ~Question, ~`Evidence retained`,
  "Input identity", "Are the consequential inputs unchanged?", "Four recorded and current SHA-256 identities",
  "Frame construction", "What was fitted for each metric and placement?", "Eighteen exact participant-day frames with schema, key, value, and count checks",
  "Site and date support", "Where do latitude, photoperiod, and collection period enter?", "Site-specific latitude, date, and civil-photoperiod ranges",
  "Pooled association", "Is a civil-photoperiod association visible?", "A separate REML smooth for each metric and placement",
  "Derivative pattern", "Does a detected increase become zero-compatible through the observed end?", "One declared 100-point, pointwise classification rule",
  "Model checks and sensitivities", "How fragile is the classification?", "Overlap among nonlinear terms (concurvity), residual, distributional, model-form, sample, placement, and site-omission checks"
) |>
  gt::gt() |>
  gt::cols_width(
    Component ~ gt::pct(18),
    Question ~ gt::pct(34),
    `Evidence retained` ~ gt::pct(48)
  ) |>
  prep_gt()
```

## Verified inputs and exact frames

### Consequential inputs

The primary near-eye and chest inputs are
`artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds`
and
`artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds`.
Site-by-date civil photoperiod comes from
`artifacts/06_model_data/context/site_solar_context.rds`. The sensitivity input
is the recorded `participant_day_metrics.rds` object identified below. The
downloadable identity table retains every exact repository-relative path.

```{r}
#| label: tbl-h07-preparation-input-identities
#| tbl-cap: "Consequential input identities verified before preparing this companion"

input_identities |>
  transmute(
    Input = .data$input_label,
    `Recorded SHA-256` = .data$recorded_sha256,
    `Current SHA-256` = .data$current_sha256,
    Status = .data$identity_status
  ) |>
  gt::gt() |>
  gt::tab_style(
    style = gt::cell_text(font = "monospace", size = gt::px(10)),
    locations = gt::cells_body(
      columns = c(`Recorded SHA-256`, `Current SHA-256`)
    )
  ) |>
  gt::cols_width(
    Input ~ gt::pct(25),
    `Recorded SHA-256` ~ gt::pct(32),
    `Current SHA-256` ~ gt::pct(32),
    Status ~ gt::pct(11)
  ) |>
  prep_gt(font_size = 11)
```

[Input-identity source data](../../../artifacts/11_source_data/H07/preparation/H07_preparation_input_identities.csv)

### Identity, schema, keys, and values

Each primary frame has one row per site, participant, local date, metric, and
placement. The unique key is `site + Id + local_date` within a metric and
placement. Required fields are site, participant, date, absolute latitude,
civil photoperiod, original metric value, transformed response, nested
participant identifier, metric identity, and run identity. All 18 frozen
frames match their recorded serialized-object hashes; their keys are unique,
their response transformations reconstruct exactly, and their stored counts
match the frames.

```{r}
#| label: tbl-h07-preparation-frame-integrity
#| tbl-cap: "Integrity checks for the 18 primary metric-specific model frames"

frame_summary |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::fmt_tf(columns = `Latitude fixed within site`) |>
  prep_gt()
```

[Frame-integrity source data](../../../artifacts/11_source_data/H07/preparation/H07_preparation_frame_integrity.csv)

No frame is selected using a fitted result from another hypothesis. Missing
metric values are removed within the relevant metric and placement, so the
sample is metric-specific. Civil photoperiod and latitude must be finite.
Temperature is not part of the contract and is not introduced.

## Candidate samples, exclusions, and support

The **gap-timing-unaware dataset** is a predefined sensitivity in which the
timing of remaining missing observations is not used for an additional
metric-specific correction. The general 50%-per-hour and 80%-per-day coverage
rules still apply, so gaps and missingness are not ignored. For this contrast,
the primary preparation could be interpreted as a **time-sensitive primary
metric dataset** because it uses gap timing where that timing is relevant to a
metric. Below, the main input is called simply the primary dataset.

### Exact fitted samples

The primary near-eye analysis contains 139–141 participants, 655–816
participant-days and observations, and nine sites, depending on the metric.
The complementary chest analysis contains 153–154 participants, 743–902
participant-days and observations, and eight sites. One participant-day is one
observation; therefore the participant-day and observation counts are equal.

```{r}
#| label: tbl-h07-preparation-sample-ranges
#| tbl-cap: "Range of exact primary fitted samples across the nine metrics"

sample_ranges |>
  gt::gt(rowname_col = "Placement") |>
  prep_gt()
```

```{r}
#| label: tbl-h07-preparation-metric-samples
#| tbl-cap: "Exact fitted sample for every metric and placement"

sample_support |>
  transmute(
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Observations = .data$observations,
    Sites = .data$sites,
    `Civil photoperiod (h)` = format_range(
      .data$photoperiod_min,
      .data$photoperiod_max,
      digits = 2
    )
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Participants, `Participant-days`, Observations, Sites)
  ) |>
  gt::cols_width(
    Metric ~ gt::pct(34),
    `Civil photoperiod (h)` ~ gt::pct(19)
  ) |>
  prep_gt(font_size = 11)
```

```{r}
#| label: fig-h07-prep-metric-sample-support
#| fig-cap: "Metric-specific fitted participant-day samples. Near-eye and chest placements were fitted separately; horizontal distance reflects different available participant-days and does not represent a paired placement effect."
#| fig-alt: "Dot plot with nine personal light-exposure metrics on the vertical axis and participant-day counts on the horizontal axis. Near-eye counts range from 655 to 816 and chest counts from 743 to 902. The smallest samples are for time below 10 lux melEDI before sleep, followed by time below 1 lux melEDI during sleep and melEDI dose."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H07/preparation/H07_preparation_metric_sample_support.png"
))
```

[Metric-sample source data](../../../artifacts/11_source_data/H07/preparation/H07_preparation_sample_support.csv)

### Site, photoperiod, latitude, and collection period

Absolute latitude is constant within every site in every primary frame. It is
therefore structurally inseparable from site. Civil photoperiod varies within
site over collection dates, but the site ranges overlap only partially and
collection periods differ by site. The pooled smooth can describe a visible
association in the combined data; it cannot isolate photoperiod from every
site-specific, seasonal, or collection-period difference.

```{r}
#| label: fig-h07-prep-site-photoperiod-ranges
#| fig-cap: "Recorded civil-photoperiod ranges by submitted-manuscript site for the complete mean-melEDI model frame. Endpoint circles mark the shortest and longest photoperiod. Tübingen (DE) has near-eye but no chest data."
#| fig-alt: "Two-panel horizontal range plot for near-eye and chest data across submitted-manuscript sites. High-latitude Borås (SE) spans roughly 12 to 20.5 hours; Delft (NL) and Dortmund (DE) span about 11.5 to 18.2 hours; Madrid (ES) spans about 10.3 to 12.4 hours; and San José (CR) and Kumasi (GH) have much narrower ranges near 13 and 12.7 hours. The site ranges overlap only partly; Tübingen (DE) appears only near eye."
#| out-width: "100%"

knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H07/preparation/H07_preparation_site_photoperiod_ranges.png"
))
```

[Site-range source data](../../../artifacts/11_source_data/H07/preparation/H07_preparation_site_photoperiod_ranges.csv)

```{r}
#| label: tbl-h07-preparation-site-support
#| tbl-cap: "Site-level support in the complete mean-melEDI frame"

site_ranges |>
  transmute(
    Placement = .data$placement_label,
    Site = .data$site_name,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Absolute latitude (°)` = .data$abs_latitude_deg,
    `Civil photoperiod (h)` = format_range(
      .data$photoperiod_min,
      .data$photoperiod_max,
      digits = 2
    ),
    `Collection dates` = paste0(.data$first_date, " to ", .data$last_date)
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(columns = c(Participants, `Participant-days`)) |>
  gt::fmt_number(columns = `Absolute latitude (°)`, decimals = 2) |>
  gt::cols_width(
    Site ~ gt::pct(17),
    `Civil photoperiod (h)` ~ gt::pct(18),
    `Collection dates` ~ gt::pct(26)
  ) |>
  prep_gt(font_size = 11)
```

The downloadable all-metric table contains 153 metric-by-placement-by-site
rows, preserving metric-specific participants, participant-days, dates,
latitude, and photoperiod ranges.

[All metric-specific site-support source data](../../../artifacts/11_source_data/H07/preparation/H07_preparation_site_support.csv)

The earlier pooled joint-support grid is retained as an identifiability
check, not as an eligibility rule. The qualifying-transition classification
uses the complete recorded metric-specific photoperiod range. Partial site
overlap therefore does not erase the association question, but it remains a
central limitation on interpretation and motivates the site-omission checks.

## Metrics, units, and response construction

All nine planned metrics enter independently. Reader-facing names use melEDI
and “period.” Internal identifiers are retained only for reproducibility.
Definitions and derivations are documented in
[Preparation 04](../../../notebooks/preparation/04_metric_derivation.qmd).

```{r}
#| label: tbl-h07-preparation-metric-contract
#| tbl-cap: "Response contract for all nine planned H07 metrics"

metric_contract |>
  mutate(
    `Model distribution` = recode(
      .data$response_family,
      gaussian = "Gaussian with identity link",
      tweedie_log = "Tweedie with log link"
    ),
    `Response transformation` = recode(
      .data$response_transform,
      log10_offset_0.1 = "log10(metric + 0.1)",
      identity = "None"
    )
  ) |>
  transmute(
    Order = .data$metric_order,
    Metric = .data$manuscript_name,
    Unit = .data$display_unit,
    .data$`Model distribution`,
    .data$`Response transformation`
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = Order) |>
  gt::cols_width(
    Metric ~ gt::pct(36),
    Unit ~ gt::pct(9),
    `Model distribution` ~ gt::pct(27),
    `Response transformation` ~ gt::pct(23)
  ) |>
  prep_gt(font_size = 11)
```

The three illuminance-like metrics, the longest continuous period above 250
lx melEDI, and melEDI dose use `log10(metric + 0.1)` with a Gaussian identity
model. Time above 1,000 lx melEDI, time above 250 lx melEDI during wake, and
time below 1 lx melEDI during sleep use a Tweedie distribution with log link.
Time below 10 lx melEDI before sleep is untransformed Gaussian. The response
reconstruction check confirms these rules exactly in every frozen frame.

## Temporal order and repeated participant-days

Dates are local calendar dates. Civil photoperiod is joined by site and local
date. Frames are ordered by site, participant, and date before fitting.
Participants contribute repeated days, so each model contains a random-effect
smooth for participant nested within site. Site also has its own random-effect
smooth. One participant-day contributes only one row per metric and placement;
there is no within-day residual time series in H07.

No independent temperature, season, or collection-period predictor is added.
Season and collection period remain embedded in the observed date and
photoperiod pattern. Consecutive-day residual correlations are descriptive
checks of remaining temporal dependence rather than a fitted autocorrelation
structure.

## Nonlinear model, curves, and derivative estimand

### Exact evaluated formulas

Every formula below is the exact Wilkinson formula supplied to the recorded
model fit. The reported pooled association uses:

```r
response_value ~ s(photoperiod_hours, k = 6, bs = "tp") +
  s(site, bs = "re") +
  s(site_participant, bs = "re")
```

```{r}
#| label: tbl-h07-preparation-formulas
#| tbl-cap: "Exact evaluated H07 Wilkinson formulas and their analytical roles"

formula_registry |>
  transmute(
    Model = .data$model_id,
    Role = .data$analysis_role,
    Formula = .data$formula
  ) |>
  gt::gt() |>
  gt::tab_style(
    style = gt::cell_text(font = "monospace", size = gt::px(10)),
    locations = gt::cells_body(columns = Formula)
  ) |>
  gt::cols_width(
    Model ~ gt::pct(21),
    Role ~ gt::pct(24),
    Formula ~ gt::pct(55)
  ) |>
  prep_gt(font_size = 10)
```

[Formula-registry source data](../../../artifacts/11_source_data/H07/preparation/H07_preparation_formula_registry.csv)

The fitting call was `mgcv::gam()` with REML, `na.action = stats::na.fail`,
unused factor levels dropped, and the default non-tracing `gam.control`. It was
not a `bam()` fit. Consequently, `discrete` has no role in H07 and no
`discrete = FALSE` alternative is required. The latitude–photoperiod tensor
models are identifiability checks rather than the source of the reported
qualifying-transition classifications.

```{r}
#| label: tbl-h07-preparation-fit-settings
#| tbl-cap: "Recorded fitting settings common to the H07 model family"

tibble::tribble(
  ~Setting, ~Value,
  "Fitting function", "mgcv::gam",
  "Smoothing-parameter method", "REML",
  "Missing-data action", "stats::na.fail after metric-specific frame construction",
  "Unused factor levels", "Dropped",
  "Reported photoperiod basis", "Thin-plate regression spline, k = 6",
  "Site structure", "Random-effect smooth for site",
  "Repeated participants", "Random-effect smooth for participant nested within site",
  "Placements", "Near eye and chest fitted separately",
  "Large-data discrete approximation", "Not applicable; bam() was not used"
) |>
  gt::gt() |>
  gt::cols_width(Setting ~ gt::pct(32), Value ~ gt::pct(68)) |>
  prep_gt()
```

### Association and derivative estimands

The association estimand is the fitted pooled civil-photoperiod smooth for one
metric and placement, conditional on the model structure. Response-scale
curves are used for scientific display: when a response was transformed for
fitting, the fitted quantity was converted back to the displayed metric unit.
The derivative estimand underlying the qualifying transition is the first
derivative of that smooth on the model's linear-predictor scale per hour of
civil photoperiod.

For each metric and placement, derivatives are evaluated at 100 equally spaced
points from the exact minimum to maximum recorded photoperiod. Central finite
differences use an increment of 0.01 h; the lower boundary uses a forward
difference and the upper boundary a backward difference. The pointwise 95%
interval uses the unconditional covariance matrix.

A pattern is present only when:

1. the immediately preceding grid point has a derivative interval wholly
   above zero;
2. the next point has a derivative interval containing zero; and
3. every later point through the recorded maximum also has an interval
   containing zero.

The transition is bracketed by the preceding and current grid points. A
non-significant derivative at one point, a visual flattening, or an AIC
threshold is not sufficient. The rule has no scientifically defensible
equivalence margin, no simultaneous curve band, and no curve-wide or
cross-metric multiplicity adjustment. It therefore supports only a
pointwise, descriptive qualifying-transition classification, not a ceiling
claim.

```{r}
#| label: tbl-h07-preparation-derivative-settings
#| tbl-cap: "Recorded derivative and qualifying-transition classification settings"

derivative_settings |>
  filter(.data$setting != "revision_id") |>
  transmute(Setting = .data$setting, Value = .data$value) |>
  gt::gt() |>
  gt::cols_width(Setting ~ gt::pct(26), Value ~ gt::pct(74)) |>
  prep_gt(font_size = 11)
```

Smooth-term p-values and latitude–photoperiod tensor tests are not released as
evidence because the design shows structural latitude/site aliasing and very
high overlap among nonlinear model terms (concurvity). No p-value enters the
reported qualifying-transition classification.

```{r}
#| label: tbl-h07-preparation-pattern-counts
#| tbl-cap: "Recorded qualifying-transition classifications carried into the results report"

pattern_summary |>
  group_by(.data$placement_label) |>
  summarise(
    Metrics = n(),
    `Pattern present` = sum(.data$derivative_defined_pattern),
    `Pattern absent` = sum(!.data$derivative_defined_pattern),
    .groups = "drop"
  ) |>
  transmute(
    Placement = .data$placement_label,
    .data$Metrics,
    .data$`Pattern present`,
    .data$`Pattern absent`
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::tab_source_note(
    gt::md(
      "Counts describe the declared pointwise pattern. Scientific interpretation and metric-level intervals are in the linked results report."
    )
  ) |>
  prep_gt()
```

## Model checks and support limits

All 18 reported photoperiod-smooth fits converged. Four smooths—three near-eye
and one chest—met the recorded basis-dimension flag criterion, motivating the
broader-basis sensitivity. Estimated overlap among nonlinear model terms
(concurvity) supplied by the nested participant smooth to the photoperiod
smooth was approximately 0.9985–0.9988. This is the principal identifiability
limitation: the pooled photoperiod shape is difficult to separate from which
participants contribute at which photoperiods.

```{r}
#| label: tbl-h07-preparation-diagnostic-summary
#| tbl-cap: "Summary of convergence, basis, overlap among nonlinear terms (concurvity), and temporal model checks"

diagnostic_summary |>
  mutate(
    Placement = unname(placement_labels[.data$placement])
  ) |>
  transmute(
    .data$Placement,
    `Fitted models` = .data$fitted_models,
    Converged = .data$converged_models,
    `Basis flags` = .data$basis_dimension_flags,
    `Participant concurvity` = format_range(
      .data$minimum_participant_concurvity,
      .data$maximum_participant_concurvity,
      digits = 4
    ),
    `Maximum |consecutive-day lag 1|` =
      .data$maximum_absolute_consecutive_day_lag1
  ) |>
  gt::gt(rowname_col = "Placement") |>
  gt::fmt_integer(columns = c(`Fitted models`, Converged, `Basis flags`)) |>
  gt::fmt_number(columns = `Maximum |consecutive-day lag 1|`, decimals = 3) |>
  prep_gt()
```

Residual summaries retained fitted-value dependence and tail checks for every
metric. **Residual autocorrelation** means that observations close in sequence
retain more similar residuals than observations farther apart.
Consecutive-day lag-1 correlations did not cross the prespecified 0.30
descriptive warning threshold, but that does not prove temporal independence.

Three metrics use a Tweedie response. A frozen 100-draw pilot checked zero
mass and positive-tail behaviour. The two above-threshold duration metrics
were broadly compatible with the fitted distributions. For time below 1 lx
melEDI during sleep, each placement had four observed zeros, the analytic
expected count was below 0.0001, and none of the 100 pilot draws contained a
zero. This distributional failure weakens any pattern for that metric; more
simulation cannot repair the model's inability to reproduce the observed zero
mass.

```{r}
#| label: tbl-h07-preparation-tweedie-pilot
#| tbl-cap: "Frozen 100-draw Tweedie distributional pilot"

tweedie_pilot |>
  transmute(
    Placement = unname(placement_labels[.data$placement]),
    Metric = .data$manuscript_name,
    Draws = .data$draws,
    `Observed zeros` = .data$observed_zero_count,
    `Analytic expected zeros` = .data$analytic_expected_zero_count,
    `Simulated zero range` = paste0(
      .data$simulated_zero_lower,
      "–",
      .data$simulated_zero_upper
    ),
    Status = .data$pilot_status
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(columns = c(Draws, `Observed zeros`)) |>
  gt::fmt_scientific(columns = `Analytic expected zeros`, decimals = 2) |>
  gt::cols_width(Metric ~ gt::pct(33), Status ~ gt::pct(20)) |>
  prep_gt(font_size = 11)
```

The pilot is displayed rather than expanded to production computation. No
bootstrap, simulation, or derivative calculation runs when this page renders.

## Sensitivity analyses and site influence

### Candidate samples

Each sensitivity changes a named feature. The gap-timing-unaware dataset omits
the additional metric-specific adjustment based on when remaining gaps occur
while retaining the coverage rules. The paired/common analysis keeps only
participant-days available at both sensor positions. The cross-dataset
common-row analyses restrict the primary and gap-timing-unaware definitions to
identical rows. The longest-period analysis keeps only days with an exactly
identified period, and the dose analysis compares observed and corrected
melEDI dose on identical rows. Model-form sensitivities increase
photoperiod-smooth flexibility from *k* = 6 to *k* = 10 or represent site with
fixed rather than random effects. Leave-one-site-out analyses omit one named
site at a time.

```{r}
#| label: tbl-h07-preparation-sensitivity-samples
#| tbl-cap: "Exact fitted sensitivity samples; ranges are metric-specific"

sensitivity_sample_summary |>
  gt::gt() |>
  gt::fmt_integer(columns = Metrics) |>
  gt::cols_width(
    Scenario ~ gt::pct(52),
    `Participant-days/observations` ~ gt::pct(20)
  ) |>
  prep_gt(font_size = 11)
```

### Classification stability

The gap-timing-unaware dataset reproduced all nine near-eye classifications
and eight of nine chest classifications. The paired/common sample reproduced
four of nine near-eye and six of nine chest classifications. Cross-dataset
common rows preserved all near-eye classifications and eight of nine chest
classifications. The exactly identified longest-period sensitivity changed
the near-eye classification. These comparisons show that several patterns are
sample- or definition-sensitive. Agreement concerns only the yes/no
qualifying-transition classification, not an identical transition point.

```{r}
#| label: tbl-h07-preparation-sensitivity-classifications
#| tbl-cap: "Agreement of sensitivity qualifying-transition classifications with the reported pooled photoperiod analysis"

sensitivity_classification |>
  gt::gt(groupname_col = "Analysis") |>
  gt::cols_width(
    Placement ~ gt::pct(18),
    Agreement ~ gt::pct(12),
    `Changed classifications` ~ gt::pct(54)
  ) |>
  prep_gt(font_size = 11)
```

```{r}
#| label: tbl-h07-preparation-model-form
#| tbl-cap: "Qualifying-transition stability under broader photoperiod smooths and fixed site effects"

model_form_summary |>
  gt::gt(groupname_col = "Model") |>
  gt::fmt_integer(columns = `Failed fits`) |>
  gt::cols_width(
    Placement ~ gt::pct(18),
    Agreement ~ gt::pct(12),
    `Failed fits` ~ gt::pct(12),
    `Changed classifications` ~ gt::pct(48)
  ) |>
  prep_gt(font_size = 11)
```

The broader `k = 10` smooth had one near-eye Hessian failure and reproduced 16
of 17 evaluable classifications. Fixed site effects reproduced 16 of 18.
Neither sensitivity resolves the underlying latitude/site identity problem.

### Leave-one-site-out influence

Every site represented in a metric frame was omitted once and the same
photoperiod model and derivative rule were applied to the reduced frame. This
is an influence check, not an eligibility criterion. Only three of nine
metrics per placement retained the same classification after every site
omission.

```{r}
#| label: tbl-h07-preparation-loso
#| tbl-cap: "Leave-one-site-out stability of the qualifying-transition classifications"

loso_results |>
  arrange(
    factor(.data$placement, levels = placement_levels),
    .data$metric_order
  ) |>
  transmute(
    Placement = unname(placement_labels[.data$placement]),
    Metric = .data$manuscript_name,
    `Matching omissions` = paste0(
      .data$omissions_matching_main,
      "/",
      .data$omissions
    ),
    `Discordant site omissions` = display_omissions(
      .data$discordant_omissions
    ),
    `Transition range when present (h)` = if_else(
      is.na(.data$plateau_start_min),
      "—",
      format_range(
        .data$plateau_start_min,
        .data$plateau_start_max,
        digits = 2
      )
    )
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::cols_width(
    Metric ~ gt::pct(31),
    `Matching omissions` ~ gt::pct(14),
    `Discordant site omissions` ~ gt::pct(33),
    `Transition range when present (h)` ~ gt::pct(22)
  ) |>
  prep_gt(font_size = 10)
```

The combination of incomplete site overlap, collection-period differences,
high overlap among nonlinear model terms (concurvity), and site-omission
changes means that a visible pooled pattern must not be interpreted as an
isolated effect of photoperiod or latitude.

## Technical provenance and bounded reproduction

::: {.callout-note title="Execution boundary"}
This companion reads frozen H07 inputs, participant-day model frames, fitted
output summaries, model checks, figures, and manifests. Its executable cells
perform identity, schema, key, value, and lightweight descriptive checks and
format tables. They do not fit or refit a model, predict a curve, calculate a
derivative, rerun a diagnostic simulation, bootstrap, or otherwise alter a
scientific result.
:::

The exact stored derivative method ID is
`REVISED_CENTRAL_UNCONDITIONAL_POINTWISE`.

### Numerical-zero provenance

The sealed `METRIC-011` numerical-zero normalization changed three near-eye
and five chest darkest-10-h-mean cells from a residual
`4.163336342344337e-17` lx to exact zero. The bounded reconciliation reopened
only L10 mean model branches, their declared model checks and sensitivities,
and false-discovery-rate (FDR) fields in families containing the L10 slot. It
preserved every non-L10 fit and raw test and changed no H07 scientific
classification, transition grid point, `REPORT-008` display value, or fitted
sample.

```{r}
#| label: tbl-h07-preparation-metric011
#| tbl-cap: "Numerical-zero reconciliation retained in technical H07 provenance"

metric011_summary |>
  filter(.data$check_id %in% c(
    "PRIMARY_CHANGED_CELLS",
    "PRIMARY_NEAR_EYE_CHANGED_CELLS",
    "PRIMARY_CHEST_CHANGED_CELLS",
    "NON_L10_MODEL_FRAME_FILES",
    "NON_L10_RAW_TABLE_CHECKS",
    "CLASSIFICATION_FIELD_CHANGES",
    "ADJUSTED_THREE_DECIMAL_CHANGES",
    "SCIENTIFIC_CONCLUSION_CHANGED",
    "BOUNDED_ARTIFACT_APPLY"
  )) |>
  transmute(
    Check = str_replace_all(.data$check_id, "_", " "),
    Observed = .data$observed,
    Expected = .data$expected,
    Status = .data$status
  ) |>
  gt::gt() |>
  gt::cols_width(
    Check ~ gt::pct(45),
    Observed ~ gt::pct(23),
    Expected ~ gt::pct(22),
    Status ~ gt::pct(10)
  ) |>
  prep_gt(font_size = 10)
```

[METRIC-011 reconciliation summary](../../../artifacts/12_manifests/H07/H07_METRIC-011_reconciliation_summary.csv)

### Code and output map

The following scripts constitute the current result-producing chain. The
analysis core constructs frames and fits models; later scripts read stored
fits to prepare diagnostics, sensitivities, derivative classifications, and
figures. The final two rows build descriptive provenance artifacts and render
this companion without refitting.

```{r}
#| label: tbl-h07-preparation-code-order
#| tbl-cap: "H07 code in execution order"

tibble::tribble(
  ~Order, ~Purpose, ~Path,
  1L, "Shared H07 metric, frame, formula, family, fit, and diagnostic functions", "scripts/hypotheses/H07/h07_stage2_core.R",
  2L, "Primary near-eye and chest metric-specific models", "scripts/hypotheses/H07/run_h07_stage2_main.R",
  3L, "Residual, concurvity, support, and distributional diagnostics", "scripts/hypotheses/H07/build_h07_stage2_main_diagnostics.R",
  4L, "Gap, paired/common-row, exact-period, and dose sensitivities", "scripts/hypotheses/H07/run_h07_stage2_sensitivities.R",
  5L, "Leave-one-site-out model fits", "scripts/hypotheses/H07/run_h07_stage2_loso.R",
  6L, "Sensitivity and influence summaries", "scripts/hypotheses/H07/build_h07_stage2_sensitivity_summaries.R",
  7L, "Broader-basis and fixed-site model-form summaries", "scripts/hypotheses/H07/build_h07_stage2_model_form_summaries.R",
  8L, "Declared first-derivative assessment", "scripts/hypotheses/H07/build_h07_stage2_revised_derivative_assessment.R",
  9L, "Paired fitted-smooth and derivative figures", "scripts/hypotheses/H07/build_h07_stage2_paired_smooth_derivative_figures.R",
  10L, "Bounded METRIC-011 L10-only reconciliation and artifact update", "scripts/hypotheses/H07/reconcile_h07_metric011.R",
  11L, "Standalone scientific results report", "notebooks/hypotheses/H07.qmd",
  12L, "Bounded descriptive source data and preparation figures", "scripts/hypotheses/H07/build_h07_preparation_artifacts.R",
  13L, "Analysis-preparation and provenance companion", "audit/hypotheses/H07/H07_analysis_preparation.qmd"
) |>
  gt::gt() |>
  gt::fmt_integer(columns = Order) |>
  gt::tab_style(
    style = gt::cell_text(font = "monospace", size = gt::px(10)),
    locations = gt::cells_body(columns = Path)
  ) |>
  gt::cols_width(
    Order ~ gt::pct(7),
    Purpose ~ gt::pct(40),
    Path ~ gt::pct(53)
  ) |>
  prep_gt(font_size = 10)
```

### Durable outputs

```{r}
#| label: tbl-h07-preparation-output-map
#| tbl-cap: "Principal durable H07 output families"

tibble::tribble(
  ~Output, ~Location, ~`Scientific role`,
  "Metric-specific frames", "artifacts/07_models/H07/frames/", "Exact fitted rows, values, keys, dates, latitude, and photoperiod",
  "Fit checkpoints", "artifacts/07_models/H07/fits/", "Formula, family, versions, frame hash, warnings, and fitted object",
  "Result and diagnostic tables", "artifacts/09_tables/H07/", "Samples, formulas, curves, derivatives, diagnostics, sensitivities, and influence",
  "Reader figures", "artifacts/08_figures/H07/", "Paired response-scale smooths and model-scale derivatives",
  "Preparation figures", "artifacts/10_figures/H07/preparation/", "Metric-specific sample and site-photoperiod support displays",
  "Preparation source data", "artifacts/11_source_data/H07/preparation/", "Exact paired data for preparation tables and figures",
  "METRIC-011 reconciliation", "artifacts/12_manifests/H07/H07_METRIC-011_*.csv", "Seals, bounded update inventory, non-L10 preservation, and unchanged-conclusion checks",
  "Preparation manifest", "artifacts/12_manifests/H07/H07_preparation_report_manifest.csv", "SHA-256, bytes, producer, and R-version inventory"
) |>
  gt::gt() |>
  gt::tab_style(
    style = gt::cell_text(font = "monospace", size = gt::px(10)),
    locations = gt::cells_body(columns = Location)
  ) |>
  gt::cols_width(
    Output ~ gt::pct(18),
    Location ~ gt::pct(38),
    `Scientific role` ~ gt::pct(44)
  ) |>
  prep_gt(font_size = 10)
```

Every preparation figure has a paired source CSV. The two results figures use
`H07_main_curve_points.csv`, `H07_revised_derivative_points.csv`,
`H07_revised_plateau_summary.csv`, and
`H07_revised_derivative_photoperiod_rows.csv`; their 9 × 18 inch, 270-dpi
export geometry is recorded in `H07_revised_paired_figure_settings.csv`.
Physical-size readability evidence is retained separately with the final
artifact identities.

### Computational environment and bounded reproduction

```{r}
#| label: tbl-h07-preparation-environment
#| tbl-cap: "Recorded computational environment for the fitted H07 outputs"

session_record |>
  transmute(
    `Analysis version` = .data$stage2_version,
    R = .data$r_version,
    mgcv = .data$mgcv_version,
    gratia = .data$gratia_version,
    dplyr = .data$dplyr_version
  ) |>
  gt::gt() |>
  prep_gt()
```

The project library and R 4.6.1 are authoritative. A bounded reproduction of
the display/provenance layer runs the preparation-artifact builder, renders
the H07 results and this companion, rebuilds the H07 preparation manifest,
and runs the H07-focused tests. It does not rerun the model fits or the frozen
100-draw distributional pilot.

::: {.callout-tip title="Interpretation boundary"}
The complete provenance supports one defensible statement: several pooled
civil-photoperiod smooths show a qualifying transition from increase to a
sustained near-flat tail within the recorded data. It does not support a
distinct latitude effect, a causal effect of photoperiod, a physiological or
environmental ceiling, near-eye/chest equivalence, or a clinical or
health-outcome claim. See the
[H07 results report](../../../notebooks/hypotheses/H07.qmd) for metric-level
estimates, 95% intervals, figures, model checks, and sensitivity
interpretation.
:::

```{r}
#| label: session-h07-preparation
#| include: false

sessionInfo()
```
