---
title: "H09 Stage 1: audit and analysis plan"
subtitle: "Chronotype and timing of personal light exposure"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-overflow: wrap
    df-print: paged
bibliography: ../../../bibliography.bib
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

root_candidates <- unique(normalizePath(
  c(".", "..", "../..", "../../.."),
  winslash = "/",
  mustWork = FALSE
))
root_matches <- root_candidates[file.exists(file.path(
  root_candidates,
  "preregistration",
  "AsPredicted #273407.pdf"
))]
if (length(root_matches) < 1L) {
  stop("Could not locate the project root", call. = FALSE)
}
root <- root_matches[[1L]]

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
if (dir.exists(project_library)) {
  .libPaths(c(project_library, .libPaths()))
}

suppressPackageStartupMessages({
  library(digest)
  library(dplyr)
  library(gt)
  library(purrr)
  library(readr)
  library(stringr)
  library(tibble)
  library(tidyr)
})

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H09 Stage 1 must run under R 4.6.1", call. = FALSE)
}

h09_path <- function(...) file.path(root, ...)
h09_sha256 <- function(path) {
  digest::digest(file = path, algo = "sha256", serialize = FALSE)
}
h09_formula_text <- function(x) paste(deparse(x), collapse = " ")
h09_participant_key <- function(data) paste(data$site, data$Id, sep = ":")
h09_day_key <- function(data) {
  paste(data$site, data$Id, data$local_date, sep = ":")
}
h09_gt <- function(data, title = NULL, note = NULL, font_size = 12) {
  out <- data |>
    gt() |>
    opt_align_table_header(align = "left") |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(font_size),
      data_row.padding = px(3),
      column_labels.font.weight = "bold",
      source_notes.font.size = px(max(9, font_size - 2))
    )
  if (!is.null(title)) out <- out |> tab_header(title = title)
  if (!is.null(note)) out <- out |> tab_source_note(note)
  out
}
h09_count_cell <- function(participants, participant_days, sites = NULL) {
  if (is.null(sites)) {
    paste0(participants, " / ", participant_days)
  } else {
    paste0(participants, " / ", participant_days, " / ", sites)
  }
}

local_midpoint_clock_minute <- function(onset, offset, timezone) {
  mapply(
    FUN = function(on, off, tz) {
      if (is.na(on) || is.na(off) || is.na(tz)) return(NA_real_)
      midpoint <- on + as.numeric(difftime(off, on, units = "secs")) / 2
      local <- as.POSIXlt(midpoint, tz = tz)
      local$hour * 60 + local$min + local$sec / 60
    },
    onset,
    offset,
    timezone,
    USE.NAMES = FALSE
  )
}

site_display <- read_csv(
  h09_path("config", "site_display_registry.csv"),
  show_col_types = FALSE
) |>
  arrange(display_order)
site_levels <- site_display$site

metric_display <- read_csv(
  h09_path("config", "metric_display_registry.csv"),
  show_col_types = FALSE
)

chronotype <- readRDS(h09_path(
  "artifacts", "06_model_data", "normalized_inputs", "chronotype.rds"
))
near_eye <- readRDS(h09_path(
  "artifacts", "06_model_data", "base",
  "metrics_glasses_participant_day_enriched.rds"
))
chest <- readRDS(h09_path(
  "artifacts", "06_model_data", "base",
  "metrics_chest_participant_day_enriched.rds"
))
gap_timing_unaware <- readRDS(h09_path(
  "artifacts", "06_model_data", "scenarios", "manuscript_prepared_data",
  "participant_day_metrics.rds"
))

stopifnot(
  nrow(chronotype) == 186L,
  !anyDuplicated(chronotype[c("site", "Id")]),
  !anyDuplicated(near_eye[c("site", "Id", "local_date")]),
  !anyDuplicated(chest[c("site", "Id", "local_date")]),
  !anyDuplicated(gap_timing_unaware[
    c("scenario_id", "position", "site", "Id", "local_date", "metric_id")
  ])
)

prepare_h09_primary <- function(data) {
  raw_l10_hour <- data$l10_midpoint_clock_minute / 60
  midpoint_longest <- local_midpoint_clock_minute(
    data$longest_bout_above_250_onset_utc,
    data$longest_bout_above_250_offset_utc,
    data$timezone
  ) / 60

  data |>
    mutate(
      mctq_hour = as.numeric(msf_sc) / 3600,
      mctq_hour_centered = mctq_hour -
        mean(as.numeric(chronotype$msf_sc) / 3600, na.rm = TRUE),
      meq_score = as.numeric(meq),
      meq_10_centered = (meq_score - mean(chronotype$meq, na.rm = TRUE)) / 10,
      m10_hour = m10_midpoint_clock_minute / 60,
      l10_hour = if_else(
        raw_l10_hour > 16,
        raw_l10_hour - 24,
        raw_l10_hour
      ),
      l10_hour_noon_cut = if_else(
        raw_l10_hour > 12,
        raw_l10_hour - 24,
        raw_l10_hour
      ),
      first_hour = first_timing_above_250_clock_minute / 60,
      last_hour = last_timing_above_250_clock_minute / 60,
      mean_hour = mean_timing_above_250_clock_minute / 60,
      longest_midpoint_observed_hour = midpoint_longest,
      longest_midpoint_exact_hour = if_else(
        longest_bout_above_250_exact_identifiable %in% TRUE,
        midpoint_longest,
        NA_real_
      ),
      photoperiod_within_site = photoperiod_hours -
        ave(
          photoperiod_hours,
          site,
          FUN = function(x) mean(x, na.rm = TRUE)
        )
    )
}

near_eye <- prepare_h09_primary(near_eye)
chest <- prepare_h09_primary(chest)

metric_contract <- tribble(
  ~metric_order, ~metric_id, ~metric_name, ~short_name, ~value_column, ~branch,
  1L, "m10_midpoint", "Midpoint of the brightest 10 hours", "M10 midpoint", "m10_hour", "Both branches",
  2L, "l10_midpoint", "Midpoint of the darkest 10 hours", "L10 midpoint", "l10_hour", "Both branches",
  3L, "first_timing_above_250", "First light timing above 250 lx melEDI", "First >250", "first_hour", "Both branches",
  4L, "last_timing_above_250", "Last light timing above 250 lx melEDI", "Last >250", "last_hour", "Both branches",
  5L, "mean_timing_above_250", "Mean timing of exposure above 250 lx melEDI", "Mean >250", "mean_hour", "Adapted/V0",
  6L, "longest_period_midpoint", "Midpoint of the longest continuous period above 250 lx melEDI", "Longest-period midpoint", "longest_midpoint_exact_hour", "Registered"
)

predictor_contract <- tribble(
  ~instrument, ~predictor_column, ~centered_column, ~effect_unit, ~direction,
  "MCTQ MSFsc", "mctq_hour", "mctq_hour_centered", "1 hour", "Later corrected midsleep",
  "MEQ", "meq_score", "meq_10_centered", "10 score points", "Greater morning preference"
)

summarise_candidate <- function(
  data,
  value_column,
  predictor_column,
  dataset,
  placement,
  metric_id,
  metric_name,
  branch,
  instrument
) {
  candidate <- data |>
    filter(
      !is.na(site),
      !is.na(Id),
      !is.na(local_date),
      is.finite(.data[[value_column]]),
      is.finite(.data[[predictor_column]])
    )
  days_per_participant <- table(h09_participant_key(candidate))

  tibble(
    dataset = dataset,
    placement = placement,
    metric_id = metric_id,
    metric = metric_name,
    branch = branch,
    instrument = instrument,
    participants = n_distinct(h09_participant_key(candidate)),
    participant_days = n_distinct(h09_day_key(candidate)),
    observations = nrow(candidate),
    sites = n_distinct(candidate$site),
    min_days = if (length(days_per_participant)) min(days_per_participant) else NA_integer_,
    median_days = if (length(days_per_participant)) median(days_per_participant) else NA_real_,
    max_days = if (length(days_per_participant)) max(days_per_participant) else NA_integer_
  )
}

primary_candidates <- pmap_dfr(
  expand_grid(
    placement = c("Near eye", "Chest"),
    metric_row = seq_len(nrow(metric_contract)),
    predictor_row = seq_len(nrow(predictor_contract))
  ),
  function(placement, metric_row, predictor_row) {
    data <- if (placement == "Near eye") near_eye else chest
    summarise_candidate(
      data = data,
      value_column = metric_contract$value_column[[metric_row]],
      predictor_column = predictor_contract$predictor_column[[predictor_row]],
      dataset = "Primary",
      placement = placement,
      metric_id = metric_contract$metric_id[[metric_row]],
      metric_name = metric_contract$metric_name[[metric_row]],
      branch = metric_contract$branch[[metric_row]],
      instrument = predictor_contract$instrument[[predictor_row]]
    )
  }
)

primary_site_candidates <- pmap_dfr(
  expand_grid(
    placement = c("Near eye", "Chest"),
    metric_row = seq_len(nrow(metric_contract)),
    predictor_row = seq_len(nrow(predictor_contract)),
    site = site_levels
  ),
  function(placement, metric_row, predictor_row, site) {
    data <- if (placement == "Near eye") near_eye else chest
    candidate <- data |>
      filter(
        .data$site == site,
        is.finite(.data[[metric_contract$value_column[[metric_row]]]]),
        is.finite(.data[[predictor_contract$predictor_column[[predictor_row]]]])
      )
    tibble(
      placement = placement,
      metric_id = metric_contract$metric_id[[metric_row]],
      instrument = predictor_contract$instrument[[predictor_row]],
      site = site,
      participants = n_distinct(h09_participant_key(candidate)),
      participant_days = n_distinct(h09_day_key(candidate))
    )
  }
)

paired_data <- near_eye |>
  select(
    site, Id, local_date,
    mctq_hour, meq_score,
    all_of(metric_contract$value_column)
  ) |>
  inner_join(
    chest |>
      select(
        site, Id, local_date,
        all_of(metric_contract$value_column)
      ),
    by = c("site", "Id", "local_date"),
    suffix = c("_near_eye", "_chest")
  )

paired_candidates <- pmap_dfr(
  expand_grid(
    metric_row = seq_len(nrow(metric_contract)),
    predictor_row = seq_len(nrow(predictor_contract))
  ),
  function(metric_row, predictor_row) {
    near_column <- paste0(metric_contract$value_column[[metric_row]], "_near_eye")
    chest_column <- paste0(metric_contract$value_column[[metric_row]], "_chest")
    predictor_column <- predictor_contract$predictor_column[[predictor_row]]
    candidate <- paired_data |>
      filter(
        is.finite(.data[[near_column]]),
        is.finite(.data[[chest_column]]),
        is.finite(.data[[predictor_column]])
      )
    days_per_participant <- table(h09_participant_key(candidate))
    tibble(
      metric_id = metric_contract$metric_id[[metric_row]],
      metric = metric_contract$metric_name[[metric_row]],
      branch = metric_contract$branch[[metric_row]],
      instrument = predictor_contract$instrument[[predictor_row]],
      participants = n_distinct(h09_participant_key(candidate)),
      participant_days = n_distinct(h09_day_key(candidate)),
      observations_per_separate_fit = nrow(candidate),
      sites = n_distinct(candidate$site),
      min_days = if (length(days_per_participant)) min(days_per_participant) else NA_integer_,
      median_days = if (length(days_per_participant)) median(days_per_participant) else NA_real_,
      max_days = if (length(days_per_participant)) max(days_per_participant) else NA_integer_
    )
  }
)

chronotype_join <- chronotype |>
  transmute(
    site,
    Id,
    mctq_hour = as.numeric(msf_sc) / 3600,
    meq_score = as.numeric(meq)
  )

gap_metric_contract <- metric_contract |>
  filter(metric_id != "longest_period_midpoint")

gap_timing_unaware <- gap_timing_unaware |>
  inner_join(chronotype_join, by = c("site", "Id")) |>
  rename(outcome_value = manuscript_prepared_value)

gap_candidates <- pmap_dfr(
  expand_grid(
    position = c("glasses", "chest"),
    metric_row = seq_len(nrow(gap_metric_contract)),
    predictor_row = seq_len(nrow(predictor_contract))
  ),
  function(position, metric_row, predictor_row) {
    metric_row_data <- gap_metric_contract[metric_row, ]
    data <- gap_timing_unaware |>
      filter(
        .data$position == position,
        metric_id == metric_row_data$metric_id
      )
    summarise_candidate(
      data = data,
      value_column = "outcome_value",
      predictor_column = predictor_contract$predictor_column[[predictor_row]],
      dataset = "Gap-timing-unaware",
      placement = if_else(position == "glasses", "Near eye", "Chest"),
      metric_id = metric_row_data$metric_id,
      metric_name = metric_row_data$metric_name,
      branch = metric_row_data$branch,
      instrument = predictor_contract$instrument[[predictor_row]]
    )
  }
)

load_v0_metric_object <- function(file_name) {
  env <- new.env(parent = emptyenv())
  object_name <- load(h09_path("data", file_name), envir = env)
  env[[object_name[[1L]]]]
}

v0_metric_names <- c(
  "brightest_10h_midpoint",
  "darkest_10h_midpoint",
  "mean_timing_above_250",
  "first_timing_above_250",
  "last_timing_above_250"
)

summarise_v0 <- function(file_name, placement) {
  metrics <- load_v0_metric_object(file_name)
  map_dfr(v0_metric_names, function(metric_name) {
    data <- metrics$data[[match(metric_name, metrics$name)]] |>
      rename(local_date = Date) |>
      left_join(chronotype_join, by = c("site", "Id"))
    summarise_candidate(
      data = data,
      value_column = "metric",
      predictor_column = "mctq_hour",
      dataset = "V0",
      placement = placement,
      metric_id = metric_name,
      metric_name = metric_name,
      branch = "Adapted/V0",
      instrument = "MCTQ MSFsc only"
    )
  })
}

v0_candidates <- bind_rows(
  summarise_v0("metrics_glasses.RData", "Near eye"),
  summarise_v0("metrics_chest.RData", "Chest")
)

questionnaire_site <- chronotype |>
  transmute(
    site,
    Id,
    mctq_hour = as.numeric(msf_sc) / 3600,
    meq_score = as.numeric(meq)
  ) |>
  group_by(site) |>
  summarise(
    participants = n_distinct(Id),
    MCTQ_complete = sum(is.finite(mctq_hour)),
    MCTQ_missing = sum(!is.finite(mctq_hour)),
    MCTQ_min_h = min(mctq_hour, na.rm = TRUE),
    MCTQ_max_h = max(mctq_hour, na.rm = TRUE),
    MEQ_complete = sum(is.finite(meq_score)),
    MEQ_missing = sum(!is.finite(meq_score)),
    MEQ_min = min(meq_score, na.rm = TRUE),
    MEQ_max = max(meq_score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  right_join(site_display, by = "site") |>
  arrange(display_order)

meq_type_support <- chronotype |>
  group_by(meq_type) |>
  summarise(
    participants = sum(!is.na(meq)),
    minimum_score = min(meq, na.rm = TRUE),
    maximum_score = max(meq, na.rm = TRUE),
    .groups = "drop"
  )

clock_support <- pmap_dfr(
  expand_grid(
    placement = c("Near eye", "Chest"),
    metric_row = seq_len(nrow(metric_contract))
  ),
  function(placement, metric_row) {
    data <- if (placement == "Near eye") near_eye else chest
    value <- data[[metric_contract$value_column[[metric_row]]]]
    value <- value[is.finite(value)]
    tibble(
      placement = placement,
      metric = metric_contract$short_name[[metric_row]],
      branch = metric_contract$branch[[metric_row]],
      complete_days = length(value),
      minimum_h = min(value),
      q05_h = unname(quantile(value, 0.05)),
      median_h = median(value),
      q95_h = unname(quantile(value, 0.95)),
      maximum_h = max(value)
    )
  }
)

longest_support <- bind_rows(
  "Near eye" = near_eye,
  "Chest" = chest,
  .id = "placement"
) |>
  group_by(placement) |>
  summarise(
    prepared_participant_days = n(),
    endpoint_available_days = sum(is.finite(longest_midpoint_observed_hour)),
    exact_identifiable_days = sum(
      is.finite(longest_midpoint_observed_hour) &
        longest_bout_above_250_exact_identifiable %in% TRUE
    ),
    censored_days = sum(
      is.finite(longest_midpoint_observed_hour) &
        longest_bout_above_250_censored %in% TRUE
    ),
    endpoint_missing_days = sum(!is.finite(longest_midpoint_observed_hour)),
    .groups = "drop"
  )

photoperiod_support <- bind_rows(
  "Near eye" = near_eye,
  "Chest" = chest,
  .id = "placement"
) |>
  group_by(placement, site) |>
  summarise(
    participant_days = sum(is.finite(photoperiod_hours)),
    minimum_h = min(photoperiod_hours, na.rm = TRUE),
    maximum_h = max(photoperiod_hours, na.rm = TRUE),
    within_site_range_h = maximum_h - minimum_h,
    .groups = "drop"
  ) |>
  left_join(site_display, by = "site") |>
  arrange(placement, display_order)

preparation_paths <- unique(c(
  list.files(
    h09_path("artifacts", "01_imported", "model_inputs", "cache"),
    pattern = "_chronotype_.*[.]RData$",
    full.names = TRUE
  ),
  h09_path(
    "artifacts", "01_imported", "model_inputs",
    c("column_audit.csv", "object_audit.csv")
  ),
  list.files(
    h09_path("artifacts", "06_model_data", "normalized_inputs", "audits"),
    pattern = "[.]csv$",
    full.names = TRUE
  ),
  h09_path(
    "artifacts", "06_model_data", "normalized_inputs", "chronotype.rds"
  ),
  h09_path(
    "artifacts", "05_metrics",
    c(
      "metrics_glasses_participant_day.rds",
      "metrics_chest_participant_day.rds",
      "metrics_glasses_admissibility.csv",
      "metrics_chest_admissibility.csv",
      "metrics_glasses_censoring_diagnostics.csv",
      "metrics_chest_censoring_diagnostics.csv",
      "metrics_glasses_gap_diagnostics.csv",
      "metrics_chest_gap_diagnostics.csv",
      "metrics_glasses_support_diagnostics.csv",
      "metrics_chest_support_diagnostics.csv"
    )
  ),
  h09_path(
    "artifacts", "06_model_data", "base",
    c(
      "metrics_glasses_participant_day_enriched.rds",
      "metrics_chest_participant_day_enriched.rds",
      "metadata_field_audit.csv",
      "metric_key_join_audit.csv",
      "metric_metadata_availability_audit.csv",
      "modality_key_join_audit.csv",
      "participant_availability_audit.csv"
    )
  ),
  h09_path(
    "artifacts", "06_model_data", "context",
    "site_solar_context_join_audit.csv"
  ),
  h09_path(
    "artifacts", "06_model_data", "scenarios", "manuscript_prepared_data",
    c(
      "participant_day_metrics.rds",
      "metric_crosswalk.csv",
      "source_manifest.csv",
      "correction_manifest.csv"
    )
  ),
  h09_path(
    "artifacts", "12_manifests",
    c(
      "model_input_acquisition.csv",
      "model_input_normalization.csv",
      "metric_artifacts.csv",
      "base_model_data_artifacts.csv",
      "manuscript_prepared_data_artifacts.csv"
    )
  )
))

stopifnot(all(file.exists(preparation_paths)))

preparation_inventory <- tibble(path = preparation_paths) |>
  mutate(
    relative_path = str_remove(path, fixed(paste0(root, "/"))),
    artifact_group = case_when(
      str_detect(relative_path, "model_inputs/cache") ~ "Pinned chronotype source cache",
      str_detect(relative_path, "01_imported/model_inputs") ~ "Imported-input audit",
      str_detect(relative_path, "normalized_inputs/audits") ~ "Normalization audit",
      str_detect(relative_path, "normalized_inputs/chronotype") ~ "Normalized chronotype",
      str_detect(relative_path, "05_metrics") ~ "Metric preparation/support",
      str_detect(relative_path, "06_model_data/base") ~ "Enriched model-ready data/audit",
      str_detect(relative_path, "06_model_data/context") ~ "Site/solar context audit",
      str_detect(relative_path, "scenarios/manuscript_prepared_data") ~
        "Gap-timing-unaware sensitivity",
      str_detect(relative_path, "12_manifests") ~ "Shared artifact manifest",
      TRUE ~ "Other"
    ),
    bytes = file.size(path),
    sha256 = map_chr(path, h09_sha256)
  ) |>
  select(artifact_group, relative_path, bytes, sha256) |>
  arrange(artifact_group, relative_path)
```

::: {.callout-warning}
## Stage 1 gate — no new H09 inferential model has been fitted

This document audits the registered H09 contract, reconstructs the submitted
implementation (V0), checks the prepared inputs and candidate samples, and
freezes a proposed estimand, model, multiplicity, diagnostic, and sensitivity
contract. The evaluated R code creates formula objects and descriptive audit
summaries only. It does **not** call `lm()`, `lmer()`, `glmer()`, `lme()`, or
another inferential fitting function. Stage 2 must not begin until the author
explicitly approves the decisions at the end.
:::

# Scope and analytical question

H09 asks whether participant-level chronotype is associated with the local
clock timing of personal light exposure across participant-days. MCTQ and MEQ
are not interchangeable measures:

- MCTQ `MSFsc` represents corrected midsleep on free days; a positive
  coefficient means that people with later corrected midsleep also tend to
  have later exposure timing.
- MEQ represents morningness-eveningness preference; a positive coefficient
  per 10 score points means that greater morning preference tends to be
  associated with later exposure timing. A negative coefficient means the
  opposite.

The outcomes are participant-day quantities, while both predictors occur once
per participant. The primary estimand is therefore a **between-participant,
site-adjusted association in mean daily exposure timing**, not a within-person
day-to-day effect and not a causal effect of changing chronotype. Repeated days
increase precision about each participant's exposure timing but do not create
new independent chronotype observations.

Near-eye measurements are primary. Chest measurements are complementary.
Placements are never pooled, and no equivalence claim will be made without a
prespecified margin.

At first use, the **gap-timing-unaware dataset** is the dataset applying the
50%-per-hour and 80%-per-day coverage rules but not using the remaining gaps'
time of day in metric-specific support decisions. This contrasts with the
time-sensitive primary dataset. From here onward, the latter is called simply
the **primary dataset**.

# Registered H09 and controlling evidence

## Exact registered hypothesis and model

The signed preregistration states:

> **H9: Timing-based metrics are associated with chronotype (MCTQ, MEQ).**

It registers the following five outcomes:

1. first time above 250 lx melanopic EDI;
2. last time above 250 lx melanopic EDI;
3. midpoint of the brightest 10 hours;
4. midpoint of the darkest 10 hours; and
5. midpoint of the longest period above 250 lx melanopic EDI.

It assigns H8–H10 the generic mixed model below, uses participant nested in
site, applies FDR control within hypothesis families, and names chest as
primary with glasses as a robustness repeat. The settled placement decision
reverses the last two roles: near eye is primary and chest complementary.

```{r}
#| label: registered-h09-formula
#| echo: true
#| code-fold: false

registered_h09_formula <- stats::as.formula(
  "metric ~ predictor * site + (1 | site:participant)"
)
registered_h09_formula
```

The preregistration does not say that one chronotype instrument may silently
replace the other, does not define how main associations and site
heterogeneity divide into multiplicity families, and does not state how
clock-valued responses should be unwrapped. Those choices must be fixed before
fitting.

## Evidence and decisions read

```{r}
evidence_inventory <- tribble(
  ~evidence, ~role,
  "preregistration/AsPredicted #273407.pdf", "Signed H09 wording, five outcomes, generic H8-H10 formula, exclusions, placement, FDR, and diagnostics",
  "preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx", "Expanded H09 wording and intended MCTQ/MEQ construct roles",
  "audit/evidence/preregistration_contract.md", "Extracted preregistration contract",
  "_deviations.qmd; audit/ledgers/deviation_register.csv", "DEV-008, DEV-037 to DEV-039, IMP-001, and IMP-009",
  "audit/hypotheses/H03-H11_gated_workflow.qmd; audit/decisions/h03_h11_gated_workflow.md", "Fixed four-stage workflow and Stage 1 no-result-seeking rule",
  "audit/hypotheses/H09-H11_migration_map.md", "Developed H09 reconstruction, gates, proposed models, diagnostics, and outputs",
  "audit/decisions/model_reporting.md", "Plain language, exact fitted samples, 95% intervals, placement, and sensitivity reporting",
  "audit/decisions/p_value_display_conventions.md", "Three-decimal p-value display and significance-before-formatting rule",
  "audit/decisions/site_display_conventions.md", "Submitted site names, order, and colours",
  "audit/decisions/paired_placement_comparison_display.md", "Separate placement fits on exact paired/common samples",
  "audit/decisions/bootstrap_execution_policy.md", "Pilot-before-production resampling gate",
  "audit/decisions/figure_readability_and_layout.md", "Final-size figure inspection and export scaling",
  "audit/decisions/reader_facing_symlog_scale.md", "Symlog display rule; not applicable to clock-time outcomes",
  "audit/decisions/gap_timing_unaware_dataset_terminology.md", "Required sensitivity terminology",
  "audit/decisions/metric_validity.md; audit/decisions/metric_implementation_parameters.md", "Supported windows, strict >250 threshold, gap termination, and boundary censoring",
  "audit/decisions/placement_decision.md", "Near-eye primary and chest complementary",
  "config/metric_display_registry.csv; config/site_display_registry.csv", "Manuscript metric and site display contracts",
  "RQ3.qmd; RQ3_chest.qmd; scripts/fitting.R; scripts/helpers.R", "Executable V0 formulas, tests, multiplicity helper, estimates, and plots",
  "data/metrics_glasses.RData; data/metrics_chest.RData", "Exact V0 timing frames",
  "tables/H9.png; tables/chest/H9.png; figures/Fig7.png; figures/chest/Fig7.png", "Rendered V0 outputs",
  "index.qmd; supplementary_information.qmd", "Submitted claims and figure/table placement",
  "manuscript/reviewer_comments_NatMed.rtf; audit/ledgers/reviewer_feedback_disposition.csv", "Sample, placement, interaction-power, multiplicity, seasonality, and claim constraints"
)

evidence_inventory |>
  h09_gt(
    title = "Evidence and decisions controlling H09 Stage 1",
    note = "The PDF and DOCX were rendered and visually inspected as well as text-indexed. No source document was modified.",
    font_size = 11
  ) |>
  cols_width(evidence ~ px(350), role ~ px(720))
```

## Preparation lineage and integrity

The auditable H09 lineage is:

```{r}
lineage <- tribble(
  ~step, ~input, ~output, ~verified_property,
  "Pinned acquisition", "Nine site chronotype releases", "Nine cached chronotype RData files", "Repository revision, source object, and checksum retained",
  "Normalization", "Cached chronotype objects", "normalized_inputs/chronotype.rds", "One site-participant row; original aggregate values and source fields retained",
  "Metric preparation", "Eligible one-minute/30-minute light profiles", "Participant-day timing metrics", "Supported M10/L10 windows, strict >250 events, gaps, and censoring tracked",
  "Model-data enrichment", "Participant-day metrics + normalized inputs + solar context", "Near-eye and chest enriched participant-day RDS", "Many-to-one questionnaire join and one row per site-participant-day",
  "Prepared-data sensitivity", "Manuscript-prepared timing metrics + normalized chronotype", "Gap-timing-unaware long RDS", "Separate scenario, placement, metric, and participant-day keys",
  "H09 Stage 1", "Approved shared artifacts only", "In-memory candidate frames", "No fitted or predicted result and no shared artifact mutation"
)

lineage |>
  h09_gt(
    title = "H09 preparation lineage",
    note = "The registered longest-period midpoint is not yet a shared prepared metric; the audit-only construction is described below.",
    font_size = 11
  )
```

The normalized chronotype artifact contains 186 unique participants from nine
sites. The primary enriched files contain 816 unique near-eye and 902 unique
chest participant-days. Near-eye includes all nine sites; chest includes eight
because MPI has no chest placement in this dataset. Every key checked above is
unique. The candidate joins therefore preserve a many-to-one
participant-day-to-questionnaire relationship.

<details>
<summary>Every preparation artifact consumed or traced for H09</summary>

```{r}
preparation_inventory |>
  mutate(
    bytes = scales::label_bytes()(bytes),
    sha256 = str_trunc(sha256, 20, ellipsis = "…")
  ) |>
  h09_gt(
    title = "H09 preparation artifact inventory",
    note = "The full SHA-256 values are computed in the evaluated object before display truncation. Hashing and structural checks are non-analytical provenance operations.",
    font_size = 10
  ) |>
  cols_width(
    artifact_group ~ px(190),
    relative_path ~ px(610),
    bytes ~ px(90),
    sha256 ~ px(170)
  )
```

</details>

# Chronotype provenance, scoring, and support

## MCTQ and MEQ are separate constructs

The registered default is to evaluate both instruments separately:

```{r}
construct_roles <- tribble(
  ~instrument, ~stored_field, ~construct, ~planned_unit, ~coefficient_interpretation, ~joint_use,
  "MCTQ", "msf_sc", "Sleep-corrected midpoint of sleep on free days", "Per 1 hour later MSFsc", "Difference in local exposure-timing hours associated with 1 hour later corrected midsleep", "Separate five-outcome family",
  "MEQ", "meq", "Morningness-eveningness preference score", "Per 10 points greater morningness", "Difference in local exposure-timing hours associated with 10 points greater morning preference", "Separate five-outcome family"
)

construct_roles |>
  rename(
    Instrument = instrument,
    `Stored field` = stored_field,
    Construct = construct,
    `Planned unit` = planned_unit,
    `Coefficient interpretation` = coefficient_interpretation,
    `Joint use` = joint_use
  ) |>
  h09_gt(
    title = "Registered chronotype constructs and practical scales",
    note = "No composite chronotype score and no mutually adjusted MCTQ-plus-MEQ primary model is proposed. Their distinct direction and meaning must remain explicit.",
    font_size = 11
  )
```

The fixed centring references, calculated once from the normalized
participant-level artifact rather than separately within outcome frames, are
`r sprintf("%.3f", mean(as.numeric(chronotype$msf_sc) / 3600, na.rm = TRUE))`
hours for MCTQ `MSFsc` and
`r sprintf("%.2f", mean(chronotype$meq, na.rm = TRUE))` points for MEQ.
Centring changes only the intercept and interaction parameterization, not the
main-effect unit.

## Aggregate-score audit

```{r}
questionnaire_site |>
  transmute(
    Site = display_name,
    Participants = participants,
    `MCTQ complete` = MCTQ_complete,
    `MCTQ missing` = MCTQ_missing,
    `MCTQ range (h)` = sprintf("%.2f–%.2f", MCTQ_min_h, MCTQ_max_h),
    `MEQ complete` = MEQ_complete,
    `MEQ missing` = MEQ_missing,
    `MEQ range` = paste0(MEQ_min, "–", MEQ_max)
  ) |>
  h09_gt(
    title = "Chronotype support by submitted site",
    note = "MCTQ has one missing aggregate score, at Madrid (ES); MEQ is complete for all 186 normalized participant rows.",
    font_size = 11
  )
```

The MCTQ range is
`r sprintf("%.2f–%.2f", min(as.numeric(chronotype$msf_sc) / 3600, na.rm = TRUE), max(as.numeric(chronotype$msf_sc) / 3600, na.rm = TRUE))`
hours after midnight. It does not wrap across midnight, so it is represented
linearly in hours. The MEQ range is
`r paste(range(chronotype$meq, na.rm = TRUE), collapse = "–")`. The stored
MEQ category labels form non-overlapping, monotonically ordered score bands:

```{r}
meq_type_support |>
  rename(
    `MEQ category` = meq_type,
    Participants = participants,
    Minimum = minimum_score,
    Maximum = maximum_score
  ) |>
  h09_gt(
    title = "Internal MEQ score-to-category consistency",
    note = "This checks the aggregate score/category relationship only; it does not reconstruct item scoring.",
    font_size = 11
  )
```

The source objects identify both `meq` and `msf_sc` as calculated fields, but
the acquired artifacts contain no item-level questionnaire responses,
missing-item rule, questionnaire version, calculation function, or
participant-level scoring audit. Consequently:

- provenance from each pinned source release to the preserved aggregate score
  is strong;
- range, missingness, key uniqueness, source revision, and internal MEQ
  category ordering are auditable; but
- the MCTQ correction and MEQ item scoring cannot be independently
  reconstructed from the files available to H09.

This is an upstream limitation, not permission to omit MEQ. Stage 2 should
remain blocked until the author either supplies/approves the score-construction
provenance or explicitly accepts the pinned aggregate calculated fields as the
analysis inputs. The requested shared action is recorded in
`audit/handoffs/H09_shared_change_request.md`.

# Timing outcomes and the open fifth-outcome decision

## Registered versus V0 metric set

```{r}
outcome_comparison <- tribble(
  ~order, ~registered_outcome, ~v0_adapted_outcome, ~status,
  1L, "Midpoint of the brightest 10 hours", "Midpoint of the brightest 10 hours", "Same construct; repaired supported window",
  2L, "Midpoint of the darkest 10 hours", "Midpoint of the darkest 10 hours", "Same construct; repaired midnight-wrapping window",
  3L, "First light timing above 250 lx melEDI", "First timing above 250 lx", "Same construct; repaired boundary-censoring rule",
  4L, "Last light timing above 250 lx melEDI", "Last timing above 250 lx", "Same construct; repaired boundary-censoring rule",
  5L, "Midpoint of the longest continuous period above 250 lx melEDI", "Mean timing of exposure above 250 lx melEDI", "Different estimands; DEV-008 remains an author gate"
)

outcome_comparison |>
  rename(
    Order = order,
    `Registered outcome` = registered_outcome,
    `V0/adapted outcome` = v0_adapted_outcome,
    Status = status
  ) |>
  h09_gt(
    title = "The fifth registered timing metric was replaced in V0",
    note = "Both alternatives are preserved at Stage 1. They are not aliases and must not be combined into a five-member family without an author decision.",
    font_size = 11
  )
```

Mean timing is a duration-weighted circular centre across **all** supported
exposure above 250 lx. The registered metric is the local clock midpoint of
one selected longest continuous period. They can differ even when both are
calculated perfectly.

The shared primary data contain the selected longest-period onset and offset,
exact-identifiability and censoring flags, and winner-selection rule, but no
approved longest-period midpoint field or metric-registry row. For this audit
only, a provisional midpoint was obtained by halving the selected UTC
onset-to-offset interval and converting that instant to the row's local time
zone. Only exact-identifiable periods are treated as candidate registered
outcomes:

```{r}
longest_support |>
  rename(
    Placement = placement,
    `Prepared days` = prepared_participant_days,
    `Endpoint available` = endpoint_available_days,
    `Exact-identifiable` = exact_identifiable_days,
    Censored = censored_days,
    `Endpoint missing` = endpoint_missing_days
  ) |>
  h09_gt(
    title = "Prepared support for a provisional registered longest-period midpoint",
    note = "The provisional values exist only in memory in this Stage 1 document and are not approved prepared results. Censored selected periods are excluded from the recommended registered branch.",
    font_size = 11
  )
```

The gap-timing-unaware artifact contains M10, L10, first, last, and mean
timing, but not the registered longest-period midpoint or its endpoints. If
the author selects the registered branch, shared preparation must add that
metric to both primary and sensitivity contracts before Stage 2. H09 will not
edit the central registry or shared preparation code.

## Clock representation

The settled rules are applied before any model:

- M10, first, last, mean timing, and the provisional registered midpoint are
  expressed as decimal local clock hours on their observed linear support.
- L10 uses the established nighttime conversion: values strictly after 16:00
  subtract 24 hours. The same rows with a noon cut are a named sensitivity.
- M10 and L10 retain midpoint and mean only; onset and offset are excluded.
- A circular distribution is not imposed merely because the response is
  clock-valued.
- If Stage 2's fixed cut-density and partial-residual checks contradict a
  usable linear representation, the affected metric is **not acceptable** and
  fitting stops for a renewed author gate; the representation will not be
  selected by p-value.

```{r}
clock_support |>
  mutate(across(ends_with("_h"), ~ sprintf("%.2f", .x))) |>
  rename(
    Placement = placement,
    Metric = metric,
    Branch = branch,
    `Complete days` = complete_days,
    Minimum = minimum_h,
    `5th percentile` = q05_h,
    Median = median_h,
    `95th percentile` = q95_h,
    Maximum = maximum_h
  ) |>
  h09_gt(
    title = "Observed candidate clock-hour support before modelling",
    note = "L10 is shown after the fixed >16:00 negative-hour conversion. The registered midpoint uses exact-identifiable periods only. These are descriptive support checks, not estimates of H09.",
    font_size = 11
  )
```

First timing includes values at 00:00, and last timing approaches 24:00.
Those rows require explicit boundary-censor reason checks in Stage 2. Their
remaining support is one-sided rather than wrapped: first timing extends
forward from 00:00, while last timing approaches midnight from earlier hours.

# Exact V0 reconstruction

## V0 formula objects and model comparisons

V0 used MCTQ `MSFsc` only and supplied the following five Wilkinson formulas
to `lmer()` for every timing metric:

```{r}
#| label: v0-h09-formulas
#| echo: true
#| code-fold: false

v0_h09_formulas <- list(
  H9_1_full = stats::as.formula(
    "metric ~ site * msf_sc + (1 | Id)"
  ),
  H9_0_site_only = stats::as.formula(
    "metric ~ site + (1 | Id)"
  ),
  H9_00_intercept_only = stats::as.formula(
    "metric ~ 1 + (1 | Id)"
  ),
  H9_ni_additive = stats::as.formula(
    "metric ~ site + msf_sc + (1 | Id)"
  ),
  H9_ns_chronotype_only = stats::as.formula(
    "metric ~ msf_sc + (1 | Id)"
  )
)

tibble(
  formula_id = names(v0_h09_formulas),
  formula = map_chr(v0_h09_formulas, h09_formula_text)
)
```

```{r}
v0_comparisons <- tribble(
  ~comparison, ~reduced, ~full, ~v0_role, ~audit_finding,
  "H9_1", "H9_00_intercept_only", "H9_ns_chronotype_only", "Reported chronotype test", "Omitted site; not the registered site-adjusted association",
  "H9_i", "H9_ni_additive", "H9_1_full", "Site-by-MCTQ interaction", "Coherent interaction comparison but wrong scalar multiplicity and only MCTQ",
  "H9_s", "H9_ns_chronotype_only", "H9_ni_additive", "Site main effect", "Not the H09 association and not the claimed nine-member interaction family"
) |>
  mutate(
    reduced_formula = map_chr(reduced, ~ h09_formula_text(v0_h09_formulas[[.x]])),
    full_formula = map_chr(full, ~ h09_formula_text(v0_h09_formulas[[.x]]))
  ) |>
  select(comparison, v0_role, reduced_formula, full_formula, audit_finding) |>
  rename(
    Comparison = comparison,
    `V0 role` = v0_role,
    `Reduced formula` = reduced_formula,
    `Full formula` = full_formula,
    `Audit finding` = audit_finding
  )

v0_comparisons |>
  h09_gt(
    title = "Exact V0 comparisons",
    note = "V0 then created a significance-selected model object, but reported slopes, confidence intervals, and marginal R² from the chronotype-only model regardless of that selection.",
    font_size = 10
  )
```

V0 claimed sum coding for site, but `fit_model()` passed no contrast argument
and RQ3 set no global contrast option. The H09 mixed models therefore used
default treatment coding. The fact that `Id` embeds site labels makes its
random-intercept groups numerically unique, but it does not make the supplied
formula the registered `site:participant` expression.

## V0 outcomes and candidate frames

```{r}
v0_candidates |>
  transmute(
    Placement = placement,
    Metric = str_replace_all(metric, "_", " ") |> str_to_sentence(),
    Instrument = instrument,
    Participants = participants,
    `Participant-days` = participant_days,
    Observations = observations,
    Sites = sites,
    `Days/participant` = paste0(min_days, "–", max_days)
  ) |>
  h09_gt(
    title = "Exact V0 candidate frames after the MCTQ join",
    note = "These counts reproduce V0's legacy metric frames and complete-case MCTQ join; V0 did not disclose them in its result table.",
    font_size = 11
  )
```

V0 retained M10 midpoint, L10 midpoint, mean timing above 250 lx, first timing,
and last timing. It omitted MEQ. Its displayed near-eye table reported MCTQ
slopes of 0.31, 0.40, 0.32, and 0.44 hours per one-hour later `MSFsc` for M10,
L10, mean timing, and first timing, respectively; last timing displayed
`p = 0.12`. The chest table displayed slopes of 0.31, 0.32, 0.39, and 0.57
hours and a last-timing p-value above 0.9. These are V0 claims under audit, not
accepted results.

## V0 multiplicity, figures, and provenance

For each comparison and metric, V0 extracted one p-value and called
`p.adjust(p, method = "fdr", n = 5)`. A scalar call has no rank information
from the other four p-values and is Bonferroni-like; it is not a five-member
vector-wide Benjamini-Hochberg adjustment. MEQ supplied no tests, and the
claimed nine-member site family was never constructed. Significance from
these displayed values must therefore be reopened.

V0's figure:

- omitted last timing after observing its result;
- plotted pooled raw points with an ordinary `lm` smooth rather than
  mixed-model predictions;
- ignored site adjustment and participant dependence in the smooth;
- used internal underscore labels; and
- had no paired source-data CSV.

The copied `assets/H9.png` and `tables/H9.png` are not byte-identical, while
`assets/Fig7.png` and the near-eye `figures/Fig7.png` are identical. No durable
model frame, fitted object, diagnostic interpretation, sample flow, or figure
source data links a submitted number back to one declared model.

## V0-to-audit resolution map

```{r}
v0_resolution <- tribble(
  ~v0_issue, ~consequence, ~stage2_resolution,
  "MEQ absent", "Registered construct omitted", "Fit and report separate MCTQ and MEQ model families unless the author approves a deviation",
  "Mean timing substituted for registered longest-period midpoint", "Fifth outcome estimand changed", "Apply the approved H09-G2 branch and retain the other as a named alternative",
  "Chronotype-only primary comparison", "No site adjustment", "Compare site-only with site-plus-chronotype on an identical frame",
  "Reported estimate came from chronotype-only fit", "Estimate/model mismatch", "Extract every estimate and interval from the declared final model",
  "Default treatment coding despite sum-coding claim", "Reference-dependent interaction interpretation", "Set and record explicit site contrasts and marginalization",
  "Scalar p-value adjustment", "Invalid family-wide FDR claim", "Assemble complete raw vectors, decide significance, then adjust and format",
  "Significance-driven selection", "Data-dependent reporting model", "Always report the site-adjusted main model; treat heterogeneity as a separate question",
  "No exact samples or diagnostics", "Precision and validity cannot be judged", "Persist frames, counts, convergence, residual, temporal, and influence assessments",
  "Pooled ordinary-lm display", "Plot does not represent fitted model", "Recreate the visual layout with all outcomes and declared model-based estimates",
  "Chest labelled robustness", "Placement hierarchy reversed by settled decision", "Near eye primary, chest complementary, plus paired/common-sample display"
)

v0_resolution |>
  rename(
    `V0 issue` = v0_issue,
    Consequence = consequence,
    `Stage 2 resolution` = stage2_resolution
  ) |>
  h09_gt(
    title = "Required V0 resolutions",
    font_size = 11
  )
```

# Proposed estimands, models, and comparisons

## Estimand hierarchy

For each author-approved five-outcome branch and each chronotype instrument:

1. **Primary main association:** the site-adjusted difference in mean local
   exposure timing associated with one hour later MCTQ `MSFsc` or 10 points
   greater MEQ morningness, using all eligible near-eye participant-days.
2. **Registered heterogeneity question:** whether that association differs
   across sites, tested separately from the main association.
3. **Complementary placement:** the same two questions for chest data, with
   no pooling and no equivalence claim.
4. **Paired/common sample:** the same placement-specific models fitted
   separately on exactly matched site-participant-date rows, followed by a
   direct descriptive estimate-difference display.
5. **Prepared-data sensitivity:** the same model and family contract on the
   gap-timing-unaware dataset and, where possible, an exact common sample with
   the primary dataset.
6. **Weighting/dependence sensitivity:** one participant-mean row per
   site-participant to assess whether unequal numbers of eligible days
   materially drive the participant-day result.

There is no within-participant chronotype estimand because neither MCTQ nor
MEQ varies by participant-day. Site interactions are secondary heterogeneity
tests; they do not determine which main model is reported.

## Exact primary formula objects — no fits

The future H09 frames will use the exact column names below. `timing_hour` is
the approved linearized outcome; `mctq_hour_centered` is centred at the fixed
participant-level MCTQ mean and remains a one-hour unit;
`meq_10_centered` is centred and divided by 10.

```{r}
#| label: proposed-h09-primary-formulas
#| echo: true
#| code-fold: false

proposed_h09_formulas <- list(
  M0_site_only = stats::as.formula(
    "timing_hour ~ site + (1 | site:Id)"
  ),
  M1_MCTQ_main = stats::as.formula(
    "timing_hour ~ site + mctq_hour_centered + (1 | site:Id)"
  ),
  M2_MCTQ_interaction = stats::as.formula(
    "timing_hour ~ site * mctq_hour_centered + (1 | site:Id)"
  ),
  M1_MEQ_main = stats::as.formula(
    "timing_hour ~ site + meq_10_centered + (1 | site:Id)"
  ),
  M2_MEQ_interaction = stats::as.formula(
    "timing_hour ~ site * meq_10_centered + (1 | site:Id)"
  )
)

tibble(
  formula_id = names(proposed_h09_formulas),
  formula = map_chr(proposed_h09_formulas, h09_formula_text)
)
```

All fixed-effect comparisons use the **same complete-case rows** for reduced
and full models and maximum likelihood. The approved final Gaussian identity
model is refitted with REML for coefficients, 95% confidence intervals,
variance components, ICC, and marginal/conditional R². Site uses the submitted
factor order and explicit `contr.sum` coding; the model manifest records the
contrast matrix. Main-model slopes are always reported, irrespective of
interaction significance.

## Exact photoperiod sensitivity formulas — no fits

Site absorbs between-site photoperiod differences and other fixed site
context. Because photoperiod also varies across measurement days, a narrowly
defined sensitivity adds only **within-site-centred daily photoperiod**. It is
not a new primary predictor and is never interacted with chronotype.

```{r}
#| label: proposed-h09-photoperiod-formulas
#| echo: true
#| code-fold: false

photoperiod_h09_formulas <- list(
  P0_site_photoperiod = stats::as.formula(
    "timing_hour ~ site + photoperiod_within_site + (1 | site:Id)"
  ),
  P1_MCTQ_main = stats::as.formula(
    "timing_hour ~ site + photoperiod_within_site + mctq_hour_centered + (1 | site:Id)"
  ),
  P2_MCTQ_interaction = stats::as.formula(
    "timing_hour ~ site * mctq_hour_centered + photoperiod_within_site + (1 | site:Id)"
  ),
  P1_MEQ_main = stats::as.formula(
    "timing_hour ~ site + photoperiod_within_site + meq_10_centered + (1 | site:Id)"
  ),
  P2_MEQ_interaction = stats::as.formula(
    "timing_hour ~ site * meq_10_centered + photoperiod_within_site + (1 | site:Id)"
  )
)

tibble(
  formula_id = names(photoperiod_h09_formulas),
  formula = map_chr(photoperiod_h09_formulas, h09_formula_text)
)
```

Temperature, weather, latitude, and unsupported predictors are not added.
Photoperiod has almost no within-site range at San José and limited range at
Kumasi, so its coefficient is a nuisance adjustment rather than a cross-site
seasonal claim.

## Exact linearity, participant-summary, and temporal formulas — no fits

Chronotype linearity is checked with a fixed three-degree-of-freedom natural
spline on the same model frame. It is a diagnostic comparison, not a route to
select whichever curve has a smaller p-value. The participant-summary
sensitivity averages eligible `timing_hour` values within each
site-participant before fitting one-row-per-participant Gaussian models.

```{r}
#| label: proposed-h09-diagnostic-formulas
#| echo: true
#| code-fold: false

diagnostic_h09_formulas <- list(
  D_MCTQ_spline = stats::as.formula(
    "timing_hour ~ site + splines::ns(mctq_hour_centered, df = 3) + (1 | site:Id)"
  ),
  D_MEQ_spline = stats::as.formula(
    "timing_hour ~ site + splines::ns(meq_10_centered, df = 3) + (1 | site:Id)"
  ),
  S0_participant_site_only = stats::as.formula(
    "participant_mean_timing_hour ~ site"
  ),
  S1_participant_MCTQ_main = stats::as.formula(
    "participant_mean_timing_hour ~ site + mctq_hour_centered"
  ),
  S2_participant_MCTQ_interaction = stats::as.formula(
    "participant_mean_timing_hour ~ site * mctq_hour_centered"
  ),
  S1_participant_MEQ_main = stats::as.formula(
    "participant_mean_timing_hour ~ site + meq_10_centered"
  ),
  S2_participant_MEQ_interaction = stats::as.formula(
    "participant_mean_timing_hour ~ site * meq_10_centered"
  )
)

tibble(
  formula_id = names(diagnostic_h09_formulas),
  formula = map_chr(diagnostic_h09_formulas, h09_formula_text)
)
```

If participant-day residual autocorrelation is material, the prespecified
temporal sensitivity uses `nlme::lme` with the same fixed main-effect formula,
a participant-within-site random intercept, and continuous-time AR(1)
correlation indexed by elapsed local days:

```{r}
#| label: proposed-h09-temporal-structures
#| echo: true
#| code-fold: false

temporal_h09_structures <- list(
  MCTQ_fixed = stats::as.formula(
    "timing_hour ~ site + mctq_hour_centered"
  ),
  MEQ_fixed = stats::as.formula(
    "timing_hour ~ site + meq_10_centered"
  ),
  random = stats::as.formula(
    "~ 1 | site/Id"
  ),
  correlation = stats::as.formula(
    "~ elapsed_day | site/Id"
  )
)

tibble(
  structure_id = names(temporal_h09_structures),
  formula = map_chr(temporal_h09_structures, h09_formula_text)
)
```

The temporal model is judged by coefficient and interval stability plus
residual autocorrelation, not by treating days as an additional scientific
predictor. If it cannot be estimated reliably, that failure is reported
rather than silently dropping the check.

## Comparison registry

```{r}
comparison_registry <- tribble(
  ~comparison_id, ~instrument, ~reduced_id, ~full_id, ~question, ~family,
  "C1-MCTQ-main", "MCTQ", "M0_site_only", "M1_MCTQ_main", "Site-adjusted average association", "H09-F1-MCTQ-main",
  "C2-MCTQ-interaction", "MCTQ", "M1_MCTQ_main", "M2_MCTQ_interaction", "Omnibus site heterogeneity", "H09-F3-MCTQ-interaction",
  "C3-MEQ-main", "MEQ", "M0_site_only", "M1_MEQ_main", "Site-adjusted average association", "H09-F2-MEQ-main",
  "C4-MEQ-interaction", "MEQ", "M1_MEQ_main", "M2_MEQ_interaction", "Omnibus site heterogeneity", "H09-F4-MEQ-interaction"
) |>
  mutate(
    reduced_formula = map_chr(
      reduced_id,
      ~ h09_formula_text(proposed_h09_formulas[[.x]])
    ),
    full_formula = map_chr(
      full_id,
      ~ h09_formula_text(proposed_h09_formulas[[.x]])
    )
  )

photoperiod_comparison_registry <- tribble(
  ~comparison_id, ~instrument, ~reduced_id, ~full_id, ~question,
  "P-C1-MCTQ-main", "MCTQ", "P0_site_photoperiod", "P1_MCTQ_main", "Photoperiod-adjusted main association",
  "P-C2-MCTQ-interaction", "MCTQ", "P1_MCTQ_main", "P2_MCTQ_interaction", "Photoperiod-adjusted site heterogeneity",
  "P-C3-MEQ-main", "MEQ", "P0_site_photoperiod", "P1_MEQ_main", "Photoperiod-adjusted main association",
  "P-C4-MEQ-interaction", "MEQ", "P1_MEQ_main", "P2_MEQ_interaction", "Photoperiod-adjusted site heterogeneity"
) |>
  mutate(
    reduced_formula = map_chr(
      reduced_id,
      ~ h09_formula_text(photoperiod_h09_formulas[[.x]])
    ),
    full_formula = map_chr(
      full_id,
      ~ h09_formula_text(photoperiod_h09_formulas[[.x]])
    )
  )

bind_rows(
  comparison_registry |>
    mutate(analysis = "Primary/complementary/paired/gap"),
  photoperiod_comparison_registry |>
    mutate(family = "Sensitivity family matching the corresponding run",
           analysis = "Photoperiod sensitivity")
) |>
  select(
    analysis, comparison_id, instrument, question,
    reduced_formula, full_formula, family
  ) |>
  rename(
    Analysis = analysis,
    Comparison = comparison_id,
    Instrument = instrument,
    Question = question,
    `Reduced formula` = reduced_formula,
    `Full formula` = full_formula,
    Family = family
  ) |>
  h09_gt(
    title = "Every proposed likelihood-ratio comparison",
    note = "Reduced and full fits use identical rows and ML. The spline, AR(1), leave-one-site-out, L10-cut, participant-summary, placement, and dataset checks are stability sensitivities rather than additional primary likelihood-ratio families.",
    font_size = 9
  ) |>
  cols_width(
    Analysis ~ px(145),
    Comparison ~ px(145),
    Instrument ~ px(70),
    Question ~ px(180),
    `Reduced formula` ~ px(310),
    `Full formula` ~ px(350),
    Family ~ px(200)
  )
```

## Multiplicity families

The proposed primary near-eye family partition is:

```{r}
multiplicity_registry <- tribble(
  ~family_id, ~members, ~test, ~decision_rule, ~role,
  "H09-F1-MCTQ-main", 5L, "M0 versus M1 for MCTQ, one per approved timing metric", "BH q < 0.05", "Primary",
  "H09-F2-MEQ-main", 5L, "M0 versus M1 for MEQ, one per approved timing metric", "BH q < 0.05", "Primary",
  "H09-F3-MCTQ-interaction", 5L, "M1 versus M2 omnibus MCTQ-by-site, one per metric", "BH q < 0.05", "Secondary registered heterogeneity",
  "H09-F4-MEQ-interaction", 5L, "M1 versus M2 omnibus MEQ-by-site, one per metric", "BH q < 0.05", "Secondary registered heterogeneity",
  "H09-F5-site-trends", 90L, "Five metrics × nine sites × two instruments, only if site trends are declared inferential", "BH q < 0.05", "Optional; descriptive is recommended"
)

multiplicity_registry |>
  rename(
    Family = family_id,
    Members = members,
    Test = test,
    `Decision rule` = decision_rule,
    Role = role
  ) |>
  h09_gt(
    title = "Proposed H09 multiplicity families",
    note = "Each family is assembled as a complete raw p-value vector before BH adjustment. An omnibus interaction is one test per metric, not nine tests. The chosen fifth-outcome branch supplies exactly one member to F1–F4.",
    font_size = 11
  )
```

Raw p-values and BH-adjusted q-values will both be retained. Significance is
decided numerically before display. Values are formatted to three decimals;
`<0.001` is used only when strictly smaller than 0.001, and bolding is limited
to values significant under the explicitly labelled family rule.

The recommended decision is to keep site-specific trends descriptive unless
the author has a scientific need for inferential site claims. If inferential,
all 90 prespecified primary near-eye trends remain in F5 even when individual
site support is weak; non-estimable members are reported, not removed to
shrink the family.

Complementary chest, paired/common-sample, gap-timing-unaware, photoperiod,
clock-cut, and other sensitivity runs repeat the same clearly suffixed family
definitions within that run. They do not alter the primary near-eye family
after results are known and are not used to create new confirmatory claims.

# Exact candidate samples

These are complete-case **candidate** frames before model construction,
fixed-effect rank checks, and diagnostic exclusions. The exact Stage 2 fitted
sample must be counted from `model.frame()` and may be smaller. Because every
H09 outcome is participant-day level, observations equal participant-days;
no hourly outcome is fitted.

## Primary all-available placement samples

```{r}
primary_candidates |>
  left_join(
    metric_contract |>
      select(metric_id, short_name),
    by = "metric_id"
  ) |>
  transmute(
    Placement = placement,
    Metric = short_name,
    Branch = branch,
    Instrument = instrument,
    Participants = participants,
    `Participant-days` = participant_days,
    Observations = observations,
    Sites = sites,
    `Days/participant` = paste0(min_days, "–", max_days)
  ) |>
  h09_gt(
    title = "Primary-dataset all-available candidate samples",
    note = "Both fifth-outcome alternatives remain visible. Near-eye is primary; chest is complementary. MCTQ loses the six days of the one participant with missing MSFsc in M10/L10 frames.",
    font_size = 10
  )
```

The registered exact-identifiable branch is materially smaller than the
adapted mean-timing branch: near-eye MCTQ has
`r primary_candidates |> filter(placement == "Near eye", metric_id == "longest_period_midpoint", instrument == "MCTQ MSFsc") |> pull(participants)`
participants and
`r primary_candidates |> filter(placement == "Near eye", metric_id == "longest_period_midpoint", instrument == "MCTQ MSFsc") |> pull(participant_days)`
participant-days, compared with
`r primary_candidates |> filter(placement == "Near eye", metric_id == "mean_timing_above_250", instrument == "MCTQ MSFsc") |> pull(participants)`
and
`r primary_candidates |> filter(placement == "Near eye", metric_id == "mean_timing_above_250", instrument == "MCTQ MSFsc") |> pull(participant_days)`,
respectively, for mean timing. This sample difference is part of the estimand
decision, not merely a precision detail.

## Site-specific candidate support

```{r}
candidate_site_matrix <- function(placement_value) {
  primary_site_candidates |>
    filter(placement == placement_value) |>
    left_join(
      metric_contract |>
        select(metric_id, short_name),
      by = "metric_id"
    ) |>
    left_join(
      site_display |>
        select(site, display_name, display_order),
      by = "site"
    ) |>
    mutate(
      row_label = paste(short_name, instrument, sep = " — "),
      cell = h09_count_cell(participants, participant_days)
    ) |>
    select(row_label, display_name, display_order, cell) |>
    arrange(row_label, display_order) |>
    select(-display_order) |>
    pivot_wider(names_from = display_name, values_from = cell)
}

candidate_site_matrix("Near eye") |>
  rename(`Metric — instrument` = row_label) |>
  h09_gt(
    title = "Primary near-eye candidates by submitted site",
    note = "Cells are participants / participant-days. Zeroes remain explicit.",
    font_size = 9
  )
```

```{r}
candidate_site_matrix("Chest") |>
  rename(`Metric — instrument` = row_label) |>
  h09_gt(
    title = "Complementary chest candidates by submitted site",
    note = "Cells are participants / participant-days. MPI is explicitly zero because no chest data are available.",
    font_size = 9
  )
```

Sparse site-specific chronotype ranges and the absence of MPI chest data
limit interaction precision. Omnibus interaction tests may remain estimable
even when individual site trends are too imprecise for reader-facing claims;
Stage 2 must show those intervals rather than infer robustness from a
non-significant omnibus test.

## Exact paired/common placement samples

```{r}
paired_candidates |>
  left_join(
    metric_contract |>
      select(metric_id, short_name),
    by = "metric_id"
  ) |>
  transmute(
    Metric = short_name,
    Branch = branch,
    Instrument = instrument,
    Participants = participants,
    `Participant-days` = participant_days,
    `Observations per separate fit` = observations_per_separate_fit,
    Sites = sites,
    `Days/participant` = paste0(min_days, "–", max_days)
  ) |>
  h09_gt(
    title = "Exactly paired near-eye/chest candidates",
    note = "The near-eye and chest models will be fitted separately on these identical keys. Paired status does not authorize pooling or equivalence claims.",
    font_size = 10
  )
```

## Gap-timing-unaware candidates

```{r}
gap_candidates |>
  left_join(
    metric_contract |>
      select(metric_id, short_name),
    by = "metric_id"
  ) |>
  transmute(
    Placement = placement,
    Metric = short_name,
    Branch = branch,
    Instrument = instrument,
    Participants = participants,
    `Participant-days` = participant_days,
    Observations = observations,
    Sites = sites,
    `Days/participant` = paste0(min_days, "–", max_days)
  ) |>
  h09_gt(
    title = "Gap-timing-unaware all-available candidate samples",
    note = "The registered longest-period midpoint is absent from this shared artifact, so no candidate row is fabricated. If that branch is approved, the sensitivity remains blocked pending shared preparation.",
    font_size = 10
  )
```

The scenario-specific comparison will report both each scenario's exact fitted
sample and the intersection of primary and gap-timing-unaware
site-participant-date keys. Stability classifications are fixed in advance:
stable, magnitude-sensitive, precision-sensitive, direction-sensitive, or
non-estimable.

## Daily photoperiod support

```{r}
photoperiod_support |>
  transmute(
    Placement = placement,
    Site = display_name,
    `Participant-days` = participant_days,
    `Minimum (h)` = sprintf("%.2f", minimum_h),
    `Maximum (h)` = sprintf("%.2f", maximum_h),
    `Within-site range (h)` = sprintf("%.2f", within_site_range_h)
  ) |>
  h09_gt(
    title = "Prepared daily photoperiod support",
    note = "Photoperiod is only a within-site-centred sensitivity. The very small range at some sites precludes strong photoperiod interpretation.",
    font_size = 10
  )
```

# Sensitivity and diagnostic contract

## Fixed sensitivity matrix

```{r}
sensitivity_registry <- tribble(
  ~order, ~analysis, ~change, ~held_fixed, ~role, ~failure_or_interpretation,
  1L, "Primary near eye", "None", "Approved fifth outcome, instruments, formula, families, and clock rules", "Primary", "Any mandatory fit or diagnostic failure makes the metric-instrument target not acceptable",
  2L, "Complementary chest", "Placement only", "All analytical choices", "Complementary", "Report agreement/disagreement; never call equivalent",
  3L, "Paired/common placement", "Restrict both placements to identical keys and fit separately", "Formulas, scales, and contrasts", "Placement-selection sensitivity", "Compare estimate difference and 95% interval where estimand-compatible",
  4L, "Gap-timing-unaware", "Prepared dataset only", "Instrument, outcome, formula, family, and clock rules", "Preparation sensitivity", "Registered fifth outcome remains unavailable until shared preparation",
  5L, "Fifth-outcome alternative", "Registered exact longest-period midpoint versus adapted mean timing", "First four metrics and model contract", "Estimand sensitivity", "Do not present the two as the same construct",
  6L, "Photoperiod", "Add within-site-centred daily photoperiod", "No new interaction or unsupported predictor", "Seasonal-context sensitivity", "Sparse within-site range limits interpretation",
  7L, "L10 clock cut", "Strict >12:00 subtracts 24 instead of strict >16:00", "Exact rows and model formula", "Clock-representation sensitivity", "Direction or material magnitude change reopens L10 interpretation",
  8L, "Clock support", "Fixed cut-density, edge rows, and partial-residual checks", "No p-value-driven representation choice", "Admissibility diagnostic", "Material discontinuity at cut stops the affected fit",
  9L, "Chronotype functional form", "Three-df natural spline", "Same rows, site, and random intercept", "Linearity diagnostic", "Material curvature makes the linear slope not acceptable without a new gate",
  10L, "Continuous-time AR(1)", "Add corCAR1 by participant and elapsed day", "Fixed main-effect terms and rows", "Temporal-dependence sensitivity", "Report estimate/interval and residual-ACF stability",
  11L, "Participant summary", "One equal-weight row per participant", "Outcome definition, site, and predictor scale", "Day-weighting sensitivity", "Material change indicates sensitivity to unequal eligible days",
  12L, "Leave one site out", "Remove one whole site per run", "Main model and scale", "Site-influence diagnostic", "A sign reversal or loss of practical magnitude driven by one site is influential",
  13L, "Participant influence", "Refit after prespecified influential-participant flags", "Main model and rows otherwise", "Influence diagnostic", "Report, never silently delete, influential participants",
  14L, "Boundary/exact support", "First/last boundary reasons; exact-only registered midpoint", "Threshold and gap rules", "Metric-support sensitivity", "Censored or unsupported event times are not recoded as observed events"
)

sensitivity_registry |>
  rename(
    Order = order,
    Analysis = analysis,
    Change = change,
    `Held fixed` = held_fixed,
    Role = role,
    `Failure or interpretation rule` = failure_or_interpretation
  ) |>
  h09_gt(
    title = "Prespecified H09 sensitivity battery",
    note = "No temperature or unsupported predictor is introduced. Sensitivities do not become confirmatory because a primary result is inconvenient.",
    font_size = 9
  )
```

No heavy resampling is planned initially. If a bootstrap or other heavy
resampling method becomes necessary after ordinary diagnostics, H09 will first
run only 50 or 100 replicates, show the pilot output and estimated production
runtime, and stop for separate production approval.

## Mandatory diagnostics and acceptable/not-acceptable rules

```{r}
diagnostic_registry <- tribble(
  ~domain, ~required_evidence, ~acceptable, ~not_acceptable,
  "Questionnaire scoring", "Pinned releases, aggregate ranges/categories, missingness, scoring provenance decision", "One unique row per participant; fixed scale and approved provenance", "Unknown or inconsistent scoring without explicit author acceptance",
  "Join and sample identity", "Key uniqueness, many-to-one join, reduced/full model-frame hashes, exact participants/days/observations/sites", "Identical reduced/full rows and no unexplained loss", "Row mismatch, duplicate inflation, or unreported complete-case loss",
  "Site/instrument support", "Counts and predictor range by submitted site", "Full fixed-effect rank and interpretable support", "Rank deficiency or site trend based on inadequate variation",
  "Clock representation", "Range, density at fixed cut, boundary reasons, L10 alternative cut", "Continuous support compatible with declared linearization", "Material wrap discontinuity or unsupported boundary treated as observed",
  "Chronotype linearity", "Partial residuals and fixed spline comparison on same frame", "No material curvature over supported predictor range", "Linear coefficient masks material curvature",
  "Response/residual distribution", "Residual-versus-fitted, Q-Q, scale-location, site/predictor panels, outlying residuals", "No consequential non-normality or heteroscedasticity for estimates/intervals", "Systematic pattern or interval distortion not resolved by declared model",
  "Convergence/Hessian", "Optimizer messages, gradient, Hessian/eigenvalues", "Converged, optimizer code 0, finite estimates, positive-definite Hessian", "Warning, material gradient, non-finite estimate, or non-positive-definite Hessian",
  "Singularity/variance", "Random-effect variance and isSingular tolerance 1e-4", "Non-singular or scientifically justified boundary with stable fixed effect", "Unresolved singularity that invalidates dependence/uncertainty claims",
  "Fixed-effect rank", "Model matrix rank and estimable coefficients", "Full rank for declared target", "Aliased interaction/trend or hidden coefficient removal",
  "Temporal dependence", "Within-participant residual ACF by lag plus corCAR1 sensitivity", "No material residual autocorrelation or stable corrected estimate/interval", "Material autocorrelation with unstable or unestimable correction",
  "Participant influence", "Cook/DFBETAs with 2/sqrt(N participants) flag plus transparent refits", "No single participant materially changes sign/practical conclusion", "Claim depends on one participant and is not calibrated",
  "Site influence", "Leave-one-site-out estimates/intervals for every main target", "Direction and practical interpretation are not driven by one site", "Sign reversal or material conclusion change on removing one site",
  "Placement/common sample", "All-available and exact paired separate fits; estimate difference with 95% interval", "Placement role and selection sensitivity transparently described", "Pooling, unsupported equivalence, or sample-change explanation absent",
  "Prepared-data sensitivity", "Scenario-specific and exact common samples with stability class", "Stable or explicitly bounded sensitivity", "Direction-sensitive/non-estimable result presented as robust",
  "Multiplicity", "Complete family vectors, independent vector-wide BH reproduction, pre-format significance flags", "Declared family length and adjusted values reproduce exactly", "Scalar adjustment, shortened family, or display-driven significance"
)

diagnostic_registry |>
  rename(
    Domain = domain,
    `Required evidence` = required_evidence,
    Acceptable = acceptable,
    `Not acceptable` = not_acceptable
  ) |>
  h09_gt(
    title = "Stage 2 diagnostic interpretation contract",
    note = "Every fitted metric × instrument × placement target receives an explicit acceptable/not-acceptable assessment. Producing a plot or software flag without interpretation is insufficient.",
    font_size = 9
  )
```

If a mandatory diagnostic is not acceptable and the prespecified sensitivity
does not resolve it, that target is reported as non-estimable or interpretation
limited. Stage 2 will not hide the failure, change family/link, discard a site,
or choose a clock representation from significance.

# Stage 2 implementation and V0-comparison plan

After approval, Stage 2 will:

1. materialize immutable H09-specific model frames from the approved shared
   inputs, including the exact predictor centring constants, clock
   transformations, outcome branch, site contrasts, and frame hashes;
2. fit the declared primary near-eye MCTQ and MEQ models, then the registered
   interaction comparisons, with one unchanged row set per comparison;
3. assemble complete multiplicity vectors before adjustment;
4. report raw and adjusted p-values, estimates in practical units, 95%
   confidence intervals, exact participants, participant-days, observations,
   sites, and days per participant for every fitted model;
5. interpret every convergence, distributional, residual, temporal, rank,
   influence, placement, and prepared-data diagnostic and mark each target
   explicitly **acceptable** or **not acceptable**;
6. fit complementary chest and exact paired/common-sample near-eye/chest
   analyses without pooling placements;
7. apply the approved sensitivity matrix, including the gap-timing-unaware
   dataset where the chosen outcome exists;
8. recreate V0 results and displays, distinguish genuine data/model changes
   from formatting corrections, and trace every submitted claim to one stored
   result row; and
9. stop again for explicit approval before any reader-facing Stage 3 report.

## Durable result and sample outputs

The planned H09-owned Stage 2 artifacts include:

```{r}
stage2_artifacts <- tribble(
  ~artifact, ~content,
  "H09_input_audit.csv", "Source hashes, questionnaire provenance status, key and join assertions",
  "H09_metric_contract.csv", "Approved five outcomes, alternative outcome, units, clock cuts, support rules",
  "H09_formula_manifest.csv", "Exact formulas, contrasts, likelihood mode, package versions, frame hashes",
  "H09_model_frames.rds", "Immutable metric × instrument × placement × scenario fitted frames",
  "H09_sample_flow.csv", "Exact participants, participant-days, observations, sites, exclusions, and reason codes",
  "H09_model_results.csv", "Estimates, standard errors, 95% intervals, raw p-values, q-values, effect scales",
  "H09_diagnostic_assessment.csv", "Evidence plus acceptable/not-acceptable interpretation for every target",
  "H09_sensitivity_comparison.csv", "Scenario-specific and exact-common-sample stability classifications",
  "H09_placement_comparison.csv", "Separate all-available and paired/common placement estimates and differences",
  "H09_v0_comparison.csv", "V0 versus audited data, samples, formulas, estimates, p-values, and conclusions",
  "H09_reporting_manifest.csv", "Figure/table paths, source-data partners, dimensions, display widths, and checksums"
)

stage2_artifacts |>
  rename(Artifact = artifact, Content = content) |>
  h09_gt(
    title = "Planned durable H09 Stage 2 artifacts",
    note = "Exact filenames may receive directory prefixes under artifacts/06_model_data/H09, artifacts/07_models/H09, artifacts/08_results/H09, and artifacts/12_manifests, but remain H09-owned.",
    font_size = 11
  )
```

## V0 table and figure recreation

V0's result table will be recreated in a recognizable metric-by-row layout,
but corrected to:

- include all five approved outcomes regardless of significance;
- include both MCTQ and MEQ in separately labelled panels or compact tables;
- show the declared main-model slope, 95% confidence interval, raw p-value,
  family-adjusted q-value, and exact sample;
- label near-eye as primary and chest as complementary; and
- link each durable table or figure to a paired source-data CSV.

V0 Figure 7's raw point/site structure will be retained only where it remains
readable and scientifically useful. Its pooled ordinary-regression smooth
will be replaced by estimates or predictions from the declared mixed model,
with 95% uncertainty. Last timing will not be omitted because of its result.
Manuscript metric and site names, submitted order, and submitted colours will
be used. A paired/common-sample estimate-difference display will be added only
for estimand-compatible placement comparisons.

Clock-time axes remain linear clock axes; the melEDI symlog display rule does
not apply. No H09 figure will be exported on an A4 canvas. Unless a narrower
journal placement is fixed, visual QA will inspect the actual HTML and a
170-mm-wide print simulation. The manifest will record base width/height,
literal export scale, export width/height, raster DPI, intended display width,
display reduction factor, smallest essential nominal text, and effective
final text size. Final-size inspection must find no clipping, overlap,
distortion, awkward line breaks, unreadable text, or excess whitespace.

# Decisions requiring explicit author approval

```{r}
decision_registry <- tribble(
  ~gate, ~decision, ~recommendation, ~consequence_if_changed,
  "H09-G1", "Chronotype instruments", "Restore both registered instruments as separate MCTQ and MEQ analyses; do not collapse them or choose by results", "Omitting either is DEV-037 and changes construct scope and multiplicity",
  "H09-G2", "Fifth timing outcome", "Use the registered exact-identifiable longest-period midpoint as the fifth primary outcome and retain mean timing above 250 lx as a named adapted sensitivity", "Choosing mean timing as primary approves DEV-008; choosing the registered branch requires shared preparation before Stage 2",
  "H09-G3", "Longest-period censoring", "For the registered midpoint, include only selected periods marked exact-identifiable; do not midpoint censored observed lower bounds", "A broader frame changes the event-time estimand and can bias midpoint timing",
  "H09-G4", "MCTQ/MEQ score provenance", "Require upstream scoring documentation or explicitly accept the pinned aggregate calculated fields before fitting", "Without acceptance, scoring validity is undetermined and Stage 2 remains blocked",
  "H09-G5", "Primary association model", "Use site-only versus site-plus-chronotype for the primary test and always report the site-adjusted M1 estimate; test M1 versus M2 separately for heterogeneity", "Returning to V0 omits site and creates estimate/model mismatch",
  "H09-G6", "Predictor scaling and contrasts", "MCTQ per 1 hour later MSFsc; MEQ per 10 points greater morningness; fixed global centring; submitted site order with explicit sum contrasts", "Different scales or contrasts change practical interpretation and interaction marginalization",
  "H09-G7", "Multiplicity", "Adopt F1 MCTQ-main, F2 MEQ-main, F3 MCTQ-interaction, and F4 MEQ-interaction as separate five-member BH families", "A combined ten-member family is coherent but changes the correction; scalar adjustment is not acceptable",
  "H09-G8", "Site-specific trends", "Keep site-specific trends descriptive; if inferential, activate prespecified 90-member H09-F5-site-trends", "Result-selected trends or a shortened family are not acceptable",
  "H09-G9", "Clock representation", "Use fixed linearizations described above, including strict >16:00 negative-hour L10 primary and noon-cut same-row sensitivity; stop if fixed diagnostics reject linear support", "A circular or alternate unwrapping model requires a renewed prefit gate",
  "H09-G10", "Photoperiod", "Keep site in every model and add only within-site-centred daily photoperiod as a named sensitivity", "Making photoperiod primary or adding temperature/unsupported predictors changes the registered estimand",
  "H09-G11", "Placements and samples", "Near-eye all-available primary, chest complementary, exact paired/common separate fits, no pooling or equivalence", "A different hierarchy changes the evidence claim and reviewer response",
  "H09-G12", "Dependence and stability", "Approve spline-linearity, participant-summary, continuous-time AR(1), participant influence, and leave-one-site-out checks with the fixed stop rules", "Dropping these leaves repeated-day, weighting, and site influence unresolved",
  "H09-G13", "Diagnostics", "Require the full acceptable/not-acceptable registry for every metric × instrument × placement target", "A result without interpreted diagnostics is not releasable",
  "H09-G14", "Multiplicity/report display", "Retain raw p and q; decide significance before three-decimal formatting; bold only under the labelled rule", "Display formatting must never determine significance",
  "H09-G15", "V0 recreation and outputs", "Recreate all V0 metrics without outcome selection, use declared model estimates, and pair every durable figure with source CSV and final-size QA", "Pooled lm smooths or untraceable copied assets are not acceptable",
  "H09-G16", "Heavy resampling", "No production resampling without a 50- or 100-replicate pilot, runtime estimate, and separate approval", "Production computation remains gated"
)

decision_registry |>
  rename(
    Gate = gate,
    Decision = decision,
    Recommendation = recommendation,
    `Consequence if changed` = consequence_if_changed
  ) |>
  h09_gt(
    title = "H09 Stage 1 author decisions",
    note = "The recommendations are fixed before H09 inferential results are inspected. Any change must be stated explicitly so the model and multiplicity contracts can be updated before fitting.",
    font_size = 9
  )
```

::: {.callout-important}
## Explicit Stage 2 approval required

Stage 1 ends here. No H09 inferential model, p-value, confidence interval,
prediction, bootstrap, or result-facing figure has been produced.

To approve the plan unchanged, reply **“Approve H09 Stage 1 as written.”**
Otherwise, identify the gate numbers to change. In particular, please decide:

1. whether the registered exact-identifiable longest-period midpoint or V0
   mean timing above 250 lx is the fifth primary outcome;
2. whether the pinned aggregate MCTQ/MEQ fields may be used without an
   item-level scoring reconstruction; and
3. whether site-specific trends remain descriptive or activate the
   90-member inferential family.

Stage 2 will begin only after those decisions and any required
coordinator-owned preparation are approved.
:::

# Reproducibility record

## Runtime and consequential package versions

```{r}
runtime_record <- tibble(
  component = c(
    "R", "digest", "dplyr", "gt", "purrr", "readr",
    "stringr", "tibble", "tidyr"
  ),
  version = c(
    as.character(getRversion()),
    map_chr(
      c("digest", "dplyr", "gt", "purrr", "readr",
        "stringr", "tibble", "tidyr"),
      ~ as.character(packageVersion(.x))
    )
  )
)

runtime_record |>
  rename(Component = component, Version = version) |>
  h09_gt(
    title = "Stage 1 runtime",
    note = "R 4.6.1 and the existing project library were used. No package was installed or updated, and no unbounded renv::status() was run.",
    font_size = 11
  )
```

## Direct analytical input checksums

```{r}
direct_inputs <- tribble(
  ~relative_path, ~role,
  "artifacts/06_model_data/normalized_inputs/chronotype.rds", "Participant-level MCTQ and MEQ",
  "artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds", "Primary near-eye timing outcomes and context",
  "artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds", "Complementary chest timing outcomes and context",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds", "Gap-timing-unaware sensitivity outcomes",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/metric_crosswalk.csv", "Sensitivity metric mapping",
  "config/metric_display_registry.csv", "Manuscript metric names and units",
  "config/site_display_registry.csv", "Submitted site names, order, and colours",
  "data/metrics_glasses.RData", "V0 near-eye reconstruction",
  "data/metrics_chest.RData", "V0 chest reconstruction",
  "RQ3.qmd", "V0 near-eye formulas, comparisons, table, and figure",
  "RQ3_chest.qmd", "V0 chest formulas, comparisons, table, and figure",
  "scripts/fitting.R", "V0 fitting and scalar p-adjust helper"
) |>
  mutate(
    path = h09_path(relative_path),
    bytes = file.size(path),
    sha256 = map_chr(path, h09_sha256)
  )

stopifnot(all(file.exists(direct_inputs$path)))

direct_inputs |>
  transmute(
    Input = relative_path,
    Role = role,
    Bytes = bytes,
    `SHA-256` = sha256
  ) |>
  h09_gt(
    title = "Direct H09 Stage 1 input identity",
    note = "The full preparation lineage and its additional hashes are listed earlier.",
    font_size = 9
  ) |>
  cols_width(Input ~ px(570), Role ~ px(300), Bytes ~ px(90), `SHA-256` ~ px(520))
```

## Commands and computation boundary

The hypothesis-scoped render command is:

```{bash}
#| eval: false
R_PROFILE_USER=/dev/null quarto render \
  audit/hypotheses/H09/01_audit_and_plan.qmd --to html
```

`R_PROFILE_USER=/dev/null` prevents the local project profile from waiting on
an unrelated renv sandbox lock; the setup cell still prepends the existing
R-4.6 project library and enforces R 4.6.1. This document reads shared files
and writes only its H09-owned HTML render. It does not refit, predict,
simulate, bootstrap, or change a central decision, ledger, registry, shared
preparation file, manuscript, or Quarto profile.

```{r}
#| label: session-info
sessionInfo()
```
