---
title: "H09 analysis preparation and provenance"
subtitle: "From verified chronotype fields and personal light-exposure timing metrics to reported associations"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-h09-preparation
#| include: false

root_candidates <- unique(c(
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(
    file.path(getwd(), "../../.."),
    winslash = "/",
    mustWork = FALSE
  )
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(
        path,
        "scripts/hypotheses/H09/h09_contract.R"
      ))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(root.dir = root)

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
if (dir.exists(project_library)) {
  .libPaths(c(project_library, .libPaths()))
}

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(knitr)
  library(readr)
  library(scales)
  library(stringr)
  library(tibble)
  library(tidyr)
})

source(file.path(root, "scripts/pipeline/paths_io.R"))

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H09 preparation requires R 4.6.1", call. = FALSE)
}

read_h09_csv <- function(area, name) {
  readr::read_csv(
    file.path(root, "artifacts", area, "H09", name),
    show_col_types = FALSE,
    progress = FALSE,
    na = ""
  )
}

prep_gt <- function(table, font_size = 11) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(max(9, font_size - 2))
    )
}

format_count <- function(value) {
  format(value, big.mark = ",", scientific = FALSE, trim = TRUE)
}

input_audit <- read_h09_csv("06_model_data", "H09_input_audit.csv")
metric_registry <- read_h09_csv(
  "06_model_data",
  "H09_metric_registry.csv"
)
predictor_registry <- read_h09_csv(
  "06_model_data",
  "H09_predictor_registry.csv"
)
formula_registry <- read_h09_csv(
  "06_model_data",
  "H09_formula_registry.csv"
)
run_registry <- read_h09_csv("06_model_data", "H09_run_registry.csv")
family_registry <- read_h09_csv(
  "06_model_data",
  "H09_family_registry.csv"
)
frame_index <- read_h09_csv(
  "06_model_data",
  "H09_model_frame_index.csv"
)
paired_audit <- read_h09_csv(
  "06_model_data",
  "H09_paired_sample_audit.csv"
)
gap_common_audit <- read_h09_csv(
  "06_model_data",
  "H09_gap_common_sample_audit.csv"
)
non_estimable <- read_h09_csv(
  "06_model_data",
  "H09_non_estimable_targets.csv"
)
fit_index <- read_h09_csv("07_models", "H09_model_fit_index.csv")
score_audit <- read_h09_csv(
  "08_diagnostics",
  "H09_chronotype_score_audit.csv"
)
diagnostic_registry <- read_h09_csv(
  "08_diagnostics",
  "H09_diagnostic_assessment_registry.csv"
)
diagnostic_summary <- read_h09_csv(
  "08_diagnostics",
  "H09_diagnostic_target_summary.csv"
)
diagnostic_adjudication <- read_h09_csv(
  "08_diagnostics",
  "H09_diagnostic_author_adjudication.csv"
)
residual_diagnostics <- read_h09_csv(
  "08_diagnostics",
  "H09_residual_diagnostics.csv"
)
model_effects <- read_h09_csv("09_tables", "H09_model_effects.csv")
model_tests <- read_h09_csv("09_tables", "H09_model_tests.csv")
family_audit <- read_h09_csv("09_tables", "H09_family_audit.csv")
gap_sensitivity <- read_h09_csv(
  "09_tables",
  "H09_gap_timing_unaware_sensitivity.csv"
)
photoperiod <- read_h09_csv(
  "09_tables",
  "H09_photoperiod_sensitivity.csv"
)
participant_summary <- read_h09_csv(
  "09_tables",
  "H09_participant_summary_sensitivity.csv"
)
ar1 <- read_h09_csv("09_tables", "H09_ar1_sensitivity.csv")
l10_cut <- read_h09_csv("09_tables", "H09_l10_cut_sensitivity.csv")
paired_effects <- read_h09_csv(
  "09_tables",
  "H09_paired_placement_effects.csv"
)
fifth_outcome <- read_h09_csv(
  "09_tables",
  "H09_fifth_outcome_sensitivity.csv"
)
execution <- read_h09_csv("12_manifests", "H09_execution_record.csv")
packages <- read_h09_csv("12_manifests", "H09_package_versions.csv")
figure_manifest <- read_h09_csv(
  "12_manifests",
  "H09_figure_manifest.csv"
)
reader_manifest <- read_h09_csv(
  "12_manifests",
  "H09_stage3_artifacts.csv"
)
site_registry <- readr::read_csv(
  file.path(root, "config/site_display_registry.csv"),
  show_col_types = FALSE
) |>
  arrange(.data$display_order)

primary_frames <- frame_index |>
  filter(
    .data$data_scenario_id == "primary",
    .data$sample_scenario == "all_available"
  ) |>
  left_join(
    metric_registry |>
      select(
        .data$metric_id,
        .data$metric_order,
        .data$manuscript_name,
        .data$primary_family_member
      ),
    by = "metric_id",
    relationship = "many-to-one"
  )
primary_registered_frames <- primary_frames |>
  filter(.data$primary_family_member)
```

::: {.callout-note title="About this analysis record"}
This page explains how verified MCTQ and MEQ fields and personal
light-exposure timing metrics were turned into the H09 results. The companion
[H09 results report](../../../notebooks/hypotheses/H09.html) presents the
scientific findings.

Rendering this page recalculates only bounded file-identity, schema, count,
key, and checksum checks and lightweight descriptions from frozen H09
indexes. It reads stored estimates, 95% confidence intervals, diagnostics,
influence summaries, sensitivities, and figures from their recorded files. It
does not fit or refit a model, calculate predictions, re-estimate diagnostics,
repeat sensitivity models, bootstrap, resample, or simulate data.
:::

## Purpose

H09 asks whether participant-level chronotype is associated with five
participant-day timing metrics of personal light exposure. MCTQ corrected
midsleep and MEQ morningness-eveningness preference remain distinct
predictors. Near-eye measurement is primary; chest measurement is
complementary and is never pooled with near-eye measurement.

The **gap-timing-unaware dataset** is the dataset applying the 50%-per-hour
and 80%-per-day coverage rules but not using the remaining gaps' time of day
in metric-specific support decisions. At this first explanation it can be
contrasted with the time-sensitive primary dataset; hereafter the latter is
called simply the primary dataset.

This page identifies the exact inputs, derived fields, analytical frames,
formulae, execution order, diagnostics, sensitivities, outputs, and file
identities behind the results. It documents provenance and does not create an
additional analysis.

## Analysis map

```{mermaid}
flowchart TD
  A["Pinned aggregate chronotype fields<br/>MCTQ MSFsc and MEQ"] --> C["Identity, range, missingness<br/>and value-preservation checks"]
  B["Prepared personal light exposure<br/>near eye and chest"] --> C
  C --> D["Six timing definitions<br/>five registered + one sensitivity"]
  D --> E["Ten run definitions<br/>dataset × placement × sample"]
  E --> F["108 metric × instrument<br/>model-frame identities"]
  F --> G["Centered predictors<br/>MCTQ per 1 h; MEQ per 10 points"]
  G --> H["Site-only, additive and<br/>site-interaction models"]
  H --> I["Effects, 95% intervals<br/>and complete-family BH adjustment"]
  H --> J["Residual, temporal, support,<br/>linearity and influence checks"]
  H --> K["Placement, preparation,<br/>photoperiod, summary and AR(1) checks"]
  I --> L["Reader tables and figures<br/>with exact source CSV files"]
  J --> L
  K --> L
  M["Bounded identity and<br/>structural verification"] -. verifies .-> C
  M -. verifies .-> F
  M -. verifies .-> L
```

The arrows show dependencies. They do not mean that this page repeats model
estimation.

## What is calculated when this page opens

```{r}
#| label: tbl-h09-prep-boundary
#| tbl-cap: "Calculations performed when this page is rendered."

tibble::tribble(
  ~Category, ~Calculated_on_render, ~Explanation,
  "File identity and schema", "Yes", "Checks recorded SHA-256 identities, required columns, row counts, and manifest entries.",
  "Sample descriptions", "Yes", "Summarises counts already stored in the H09 model-frame index; it does not recreate a frame.",
  "Model construction or estimation", "No", "Reads formula, fit, effect, and test registries without fitting or comparing a model.",
  "Diagnostics and sensitivities", "No", "Summarises recorded assessments; no residual, influence, temporal, or sensitivity calculation is repeated.",
  "Descriptive support figure", "Display only", "Plots exact participant and participant-day counts from the linked model-frame index."
) |>
  gt::gt() |>
  gt::cols_width(
    Category ~ gt::pct(24),
    Calculated_on_render ~ gt::pct(18),
    Explanation ~ gt::pct(58)
  ) |>
  prep_gt()
```

## 1. Verified inputs

### File identity

Eleven pinned inputs define the current result-producing chain: preparation
manifests, near-eye and chest participant-day data, the normalized chronotype
table, the preparation-sensitivity data, and the site and metric display
registries. Their complete 64-character SHA-256 values were recorded before
analysis.

```{r}
#| label: tbl-h09-prep-input-identities
#| tbl-cap: "Verified inputs used by the current H09 result-producing analysis."

result_input_roles <- c(
  "metric_manifest",
  "base_manifest",
  "primary_near_eye_context",
  "primary_chest_context",
  "primary_near_eye_enriched",
  "primary_chest_enriched",
  "normalized_chronotype",
  "gap_timing_unaware_metrics",
  "gap_manifest",
  "site_display_registry",
  "metric_display_registry"
)
input_labels <- c(
  metric_manifest = "Metric-artifact manifest",
  base_manifest = "Model-ready data manifest",
  primary_near_eye_context = "Primary near-eye context",
  primary_chest_context = "Complementary chest context",
  primary_near_eye_enriched = "Primary near-eye analytical fields",
  primary_chest_enriched = "Complementary chest analytical fields",
  normalized_chronotype = "Normalized MCTQ and MEQ fields",
  gap_timing_unaware_metrics = "Preparation-sensitivity metrics",
  gap_manifest = "Preparation-sensitivity manifest",
  site_display_registry = "Submitted site display registry",
  metric_display_registry = "Manuscript metric display registry"
)

verified_inputs <- input_audit |>
  filter(.data$input_role %in% result_input_roles) |>
  mutate(
    current_sha256 = vapply(
      file.path(root, .data$path),
      artifact_sha256,
      character(1)
    ),
    Input = unname(input_labels[.data$input_role]),
    Path = .data$path,
    `SHA-256 prefix` = substr(.data$current_sha256, 1L, 16L),
    Verified = if_else(
      .data$current_sha256 == .data$observed_sha256 &
        .data$hash_verified & .data$rows_verified,
      "PASS",
      "FAIL"
    )
  )
stopifnot(nrow(verified_inputs) == 11L, all(verified_inputs$Verified == "PASS"))

verified_inputs |>
  select("Input", "Path", "SHA-256 prefix", "Verified") |>
  gt::gt() |>
  gt::cols_width(
    Input ~ gt::pct(24),
    Path ~ gt::pct(54),
    `SHA-256 prefix` ~ gt::pct(15),
    Verified ~ gt::pct(7)
  ) |>
  gt::tab_source_note(
    "Complete identities remain in artifacts/06_model_data/H09/H09_input_audit.csv."
  ) |>
  prep_gt(10)
```

### Bounded integrity checks

The stored indexes establish the joins and analytical denominators without
reopening model objects. Chronotype is keyed by site and participant; model
frames are keyed by dataset, placement, site, participant, local date, metric,
and instrument.

```{r}
#| label: tbl-h09-prep-integrity-checks
#| tbl-cap: "Bounded integrity checks from stored H09 registries and indexes."

required_frame_columns <- c(
  "frame_id", "run_id", "placement", "metric_id", "instrument_id",
  "row_key_hash", "model_frame_hash", "participants", "participant_days",
  "observations", "sites", "derivation_hours"
)
integrity_checks <- tibble::tribble(
  ~Check, ~Expected, ~Observed, ~Status,
  "Current result-input identities", "11 matches", paste0(sum(verified_inputs$Verified == "PASS"), " matches"), if_else(all(verified_inputs$Verified == "PASS"), "PASS", "FAIL"),
  "Chronotype participant key", "186 unique participants", paste0(score_audit$participants, " participants"), if_else(score_audit$participants == 186L, "PASS", "FAIL"),
  "Chronotype missingness", "MCTQ 1; MEQ 0", paste0("MCTQ ", score_audit$mctq_missing, "; MEQ ", score_audit$meq_missing), if_else(score_audit$mctq_missing == 1L && score_audit$meq_missing == 0L, "PASS", "FAIL"),
  "Metric registry", "6 unique metrics; 5 registered", paste0(n_distinct(metric_registry$metric_id), " unique; ", sum(metric_registry$primary_family_member), " registered"), if_else(n_distinct(metric_registry$metric_id) == 6L && sum(metric_registry$primary_family_member) == 5L, "PASS", "FAIL"),
  "Run registry", "10 unique runs", paste0(n_distinct(run_registry$run_id), " unique runs"), if_else(n_distinct(run_registry$run_id) == 10L, "PASS", "FAIL"),
  "Model-frame index", "108 unique frames", paste0(nrow(frame_index), " rows"), if_else(nrow(frame_index) == 108L && !anyDuplicated(frame_index$frame_id), "PASS", "FAIL"),
  "Frame-index schema", "12 required fields", paste0(sum(required_frame_columns %in% names(frame_index)), " present"), if_else(all(required_frame_columns %in% names(frame_index)), "PASS", "FAIL"),
  "Stored fit index", "540 model records", paste0(nrow(fit_index), " records"), if_else(nrow(fit_index) == 540L, "PASS", "FAIL"),
  "Paired placement identities", "12 exact pairs", paste0(sum(paired_audit$exact_row_keys_match), " exact pairs"), if_else(all(paired_audit$exact_row_keys_match), "PASS", "FAIL"),
  "Primary–sensitivity common keys", "20 exact key records", paste0(sum(gap_common_audit$exact_row_keys_match, na.rm = TRUE), " exact key records"), if_else(sum(gap_common_audit$exact_row_keys_match, na.rm = TRUE) == 20L, "PASS", "FAIL"),
  "Unavailable prepared outcomes", "12 explicit records", paste0(nrow(non_estimable), " records"), if_else(nrow(non_estimable) == 12L, "PASS", "FAIL")
)
stopifnot(all(integrity_checks$Status == "PASS"))

integrity_checks |>
  gt::gt() |>
  gt::cols_width(
    Check ~ gt::pct(28),
    Expected ~ gt::pct(27),
    Observed ~ gt::pct(36),
    Status ~ gt::pct(9)
  ) |>
  prep_gt(10)
```

## 2. Chronotype predictors

### Aggregate-field provenance and missingness

The analysis uses the pinned aggregate calculated fields `msf_sc` and `meq`.
The available source does not contain item-level questionnaire responses, so
item scoring cannot be independently reconstructed. Value-preservation,
ranges, missingness, and joins were verified for the exact aggregate fields.
MCTQ and MEQ were not combined: MCTQ describes corrected sleep timing, whereas
MEQ describes morningness-eveningness preference.

```{r}
#| label: tbl-h09-prep-score-audit
#| tbl-cap: "Observed support and centering constants for the two chronotype fields."

tibble::tribble(
  ~Instrument, ~Construct, ~Complete, ~Missing, ~Minimum, ~Mean, ~Maximum, ~Coefficient_unit,
  "MCTQ MSFsc", "Corrected midsleep timing", score_audit$mctq_complete, score_audit$mctq_missing, score_audit$mctq_min_hour, score_audit$mctq_center_hour, score_audit$mctq_max_hour, "1 hour later",
  "MEQ", "Morningness-eveningness preference", score_audit$meq_complete, score_audit$meq_missing, score_audit$meq_min_score, score_audit$meq_center_score, score_audit$meq_max_score, "10 points greater morning preference"
) |>
  gt::gt() |>
  gt::fmt_number(columns = c(Minimum, Mean, Maximum), decimals = 3) |>
  gt::fmt_integer(columns = c(Complete, Missing)) |>
  gt::cols_label(Coefficient_unit = "Coefficient unit") |>
  prep_gt(10)
```

### Centering and scaling

The two model-column names, `mctq_hour_centered` and `meq_10_centered`, encode
only a change of origin and, for MEQ, a reporting scale:

- **MCTQ:** `mctq_hour_centered = MCTQ MSFsc in hours - 4.1135843`
- **MEQ:** `meq_10_centered = (MEQ score - 52.8602151) / 10`

Thus one MCTQ unit means one hour later corrected midsleep, while one MEQ unit
means 10 score points toward greater morning preference. A centered value of
zero is the observed participant mean. Centering changes the intercept's
reference point; it does not change the chronotype slope, its 95% confidence
interval, or its test.

```{r}
#| label: tbl-h09-prep-predictor-contract
#| tbl-cap: "Predictor fields, transformations, and coefficient interpretations."

predictor_registry |>
  transmute(
    Instrument = .data$instrument_name,
    `Source field` = .data$predictor_column,
    `Model field` = .data$centered_column,
    Transformation = if_else(
      .data$instrument_id == "MCTQ",
      sprintf("MCTQ hours − %.7f", score_audit$mctq_center_hour),
      sprintf("(MEQ − %.7f) / 10", score_audit$meq_center_score)
    ),
    `Coefficient unit` = .data$effect_unit,
    Interpretation = .data$effect_direction
  ) |>
  gt::gt() |>
  prep_gt(10)
```

## 3. Timing metrics and analytical samples

### Metric definitions

Five registered participant-day outcomes enter each confirmatory family. The
midpoint of the longest continuous period is calculated only when the selected
period is exactly identifiable from its recorded endpoints. Mean timing above
250 lx melEDI is a distinct sensitivity estimand and is not substituted for
the registered midpoint. M10 and L10 onset and offset are excluded.

Clock values were linearized because their observed support was adequately
linear. Night-time values use the specified negative-hour conversion where
applicable; no circular response distribution was imposed merely because an
outcome is clock-valued.

```{r}
#| label: tbl-h09-prep-metric-contract
#| tbl-cap: "H09 metric definitions, analytical roles, and units."

metric_registry |>
  arrange(.data$metric_order) |>
  transmute(
    Order = .data$metric_order,
    Metric = .data$manuscript_name,
    Role = if_else(
      .data$primary_family_member,
      "Registered family member",
      "Distinct sensitivity estimand"
    ),
    `Analysis unit` = str_replace_all(.data$analysis_unit, "_", " "),
    Unit = .data$display_unit,
    Definition = .data$value_definition
  ) |>
  gt::gt(rowname_col = "Order") |>
  prep_gt(9)
```

### Exact primary and complementary samples

Each metric and instrument uses its exact complete participant-day frame.
Primary near-eye frames contain 131–141 participants and 478–816
participant-days. Complementary chest frames contain 149–154 participants and
547–902 participant-days.

```{r}
#| label: tbl-h09-prep-primary-samples
#| tbl-cap: "Exact primary-dataset samples for the five registered timing outcomes."

primary_registered_frames |>
  arrange(
    .data$placement_label,
    .data$metric_order,
    .data$instrument_id
  ) |>
  transmute(
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Instrument = .data$instrument_id,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Observations = .data$observations,
    `Derivation hours` = .data$derivation_hours,
    Sites = .data$sites
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(
    columns = c(Participants, `Participant-days`, Observations, Sites),
    use_seps = TRUE
  ) |>
  gt::fmt_number(columns = `Derivation hours`, decimals = 1, use_seps = TRUE) |>
  prep_gt(9)
```

```{r}
#| label: fig-h09-prep-sample-support
#| fig-cap: "Metric- and instrument-specific participant and participant-day support in the primary near-eye and complementary chest analyses. Counts come directly from the frozen H09 model-frame index."
#| fig-alt: "Two stacked line-and-point panels show participant and participant-day counts across five registered timing metrics for MCTQ and MEQ. Chest counts are generally above near-eye counts; participant-day counts vary most for the longest-period midpoint."
#| fig-width: 6.692913
#| fig-height: 5.0
#| out-width: "100%"

sample_plot_data <- primary_registered_frames |>
  select(
    .data$metric_order,
    .data$placement_label,
    .data$instrument_id,
    .data$participants,
    .data$participant_days
  ) |>
  pivot_longer(
    cols = c("participants", "participant_days"),
    names_to = "quantity",
    values_to = "count"
  ) |>
  mutate(
    quantity = recode(
      .data$quantity,
      participants = "Participants",
      participant_days = "Participant-days"
    ),
    series = paste(.data$placement_label, .data$instrument_id, sep = " — ")
  )

ggplot(
  sample_plot_data,
  aes(
    x = .data$metric_order,
    y = .data$count,
    colour = .data$placement_label,
    linetype = .data$instrument_id,
    group = .data$series
  )
) +
  geom_line(linewidth = 0.65) +
  geom_point(size = 2) +
  facet_wrap(vars(.data$quantity), scales = "free_y", ncol = 1) +
  scale_colour_manual(values = c("Near eye" = "#0072B2", "Chest" = "#D55E00")) +
  scale_x_continuous(breaks = seq_len(5L)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x = "Registered metric order",
    y = NULL,
    colour = "Placement",
    linetype = "Instrument"
  ) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text = element_text(size = 8),
    axis.title.x = element_text(size = 9, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8.5)
  )
```

The figure uses the exact
[model-frame index](../../../artifacts/06_model_data/H09/H09_model_frame_index.csv)
as its paired source data. Metric order follows the preceding table.

### Placement and dataset common samples

Near-eye and chest effects were fitted separately on exactly matched
participant-day keys for the placement comparison. Primary and
gap-timing-unaware effects likewise use exact common keys where the same
estimand exists in both datasets. The registered longest-period midpoint is
unavailable in the preparation-sensitivity artifact and is not imputed.

```{r}
#| label: tbl-h09-prep-common-samples
#| tbl-cap: "Stored sample-identity checks for placement and dataset comparisons."

tibble::tribble(
  ~Comparison, ~Stored_pairs, ~Exact_keys, ~Qualification,
  "Near eye versus chest", nrow(paired_audit), sum(paired_audit$exact_row_keys_match), "Six metrics × two instruments; placements fitted separately.",
  "Primary versus gap-timing-unaware", nrow(gap_common_audit), sum(gap_common_audit$exact_row_keys_match, na.rm = TRUE), "Sixteen estimable exact pairs; four registered longest-period pairs unavailable."
) |>
  gt::gt() |>
  gt::fmt_integer(columns = c(Stored_pairs, Exact_keys)) |>
  gt::cols_label(
    Stored_pairs = "Stored pairs",
    Exact_keys = "Exact-key pairs"
  ) |>
  prep_gt(10)
```

## 4. Model construction and inference

### Exact formulae supplied to the models

For each metric, placement, dataset, sample, and chronotype instrument, the
site-only model is compared with an additive chronotype model. A second
comparison adds the site-by-chronotype interaction. Restricted maximum
likelihood additive fits provide coefficients and 95% Wald confidence
intervals; maximum-likelihood fits provide the nested likelihood-ratio tests.
A natural-spline fit is used only to assess whether the observed predictor
support is adequately represented as linear.

```{r}
#| label: tbl-h09-prep-primary-formulas
#| tbl-cap: "Exact stored Wilkinson formulae for the principal models and linearity check."

formula_registry |>
  filter(.data$analysis_kind == "approved_participant_day") |>
  mutate(
    Role = recode(
      .data$formula_id,
      M0_site_only = "Site-only comparison model",
      M1_main = "Additive chronotype model",
      M2_interaction = "Site-interaction model",
      D_spline = "Linearity diagnostic"
    )
  ) |>
  transmute(
    Instrument = .data$instrument,
    Role,
    `Formula supplied` = .data$formula
  ) |>
  gt::gt(groupname_col = "Instrument") |>
  prep_gt(9)
```

The sensitivity formulae retain the same chronotype definitions. Photoperiod
is centered within site and added without an unsupported interaction;
participant-summary models use one outcome mean per participant; continuous
time AR(1) fits use elapsed day only for the within-participant correlation
structure.

```{r}
#| label: tbl-h09-prep-sensitivity-formulas
#| tbl-cap: "Exact stored formula components for structured model sensitivities."

formula_registry |>
  filter(.data$analysis_kind %in% c(
    "photoperiod_sensitivity",
    "participant_summary_sensitivity",
    "continuous_time_ar1_sensitivity"
  )) |>
  mutate(
    Analysis = recode(
      .data$analysis_kind,
      photoperiod_sensitivity = "Within-site photoperiod",
      participant_summary_sensitivity = "Participant summary",
      continuous_time_ar1_sensitivity = "Continuous-time AR(1)"
    )
  ) |>
  transmute(
    Analysis,
    Instrument = .data$instrument,
    Component = .data$formula_id,
    `Formula supplied` = .data$formula
  ) |>
  gt::gt(groupname_col = "Analysis") |>
  prep_gt(8)
```

### Model records and intermediate quantities

The analysis driver wrote one immutable frame identity per estimable target
and five stored model records per frame. Effects and tests are kept in
separate tables before being assembled into the master result table.

```{r}
#| label: tbl-h09-prep-intermediate-artifacts
#| tbl-cap: "Intermediate H09 quantities persisted before reader reporting."

tibble::tribble(
  ~Intermediate, ~Rows_or_objects, ~Purpose, ~Path,
  "Model-frame index", nrow(frame_index), "Exact samples, row hashes, frame hashes, and derivation hours", "artifacts/06_model_data/H09/H09_model_frame_index.csv",
  "Model-frame archive", nrow(frame_index), "Frozen participant-day frames keyed by frame_id", "artifacts/06_model_data/H09/H09_model_frames.rds",
  "Model-fit index", nrow(fit_index), "Fit status, convergence, Hessian, singularity, and rank fields", "artifacts/07_models/H09/H09_model_fit_index.csv",
  "Effect table", nrow(model_effects), "Additive coefficients and 95% confidence intervals", "artifacts/09_tables/H09/H09_model_effects.csv",
  "Test table", nrow(model_tests), "Main and interaction likelihood-ratio tests", "artifacts/09_tables/H09/H09_model_tests.csv",
  "Family audit", nrow(family_audit), "Complete-family membership and independent BH verification", "artifacts/09_tables/H09/H09_family_audit.csv",
  "Diagnostic registry", nrow(diagnostic_registry), "Sixteen explicit domains for each primary target", "artifacts/08_diagnostics/H09/H09_diagnostic_assessment_registry.csv"
) |>
  gt::gt() |>
  gt::fmt_integer(columns = Rows_or_objects, use_seps = TRUE) |>
  gt::cols_label(Rows_or_objects = "Rows or objects") |>
  prep_gt(9)
```

### Multiplicity families

MCTQ main effects, MEQ main effects, MCTQ-by-site interactions, and MEQ-by-site
interactions form distinct five-outcome Benjamini–Hochberg families. Near-eye,
chest, paired, and preparation-sensitivity runs retain separate family
instances. Significance is decided from unrounded adjusted values before
three-decimal display. Site-specific slopes are descriptive unless the global
interaction family supports heterogeneity.

```{r}
#| label: tbl-h09-prep-families
#| tbl-cap: "Registered H09 multiplicity families."

family_registry |>
  transmute(
    Family = .data$family_id,
    Instrument = .data$instrument_id,
    Test = str_to_sentence(.data$test_kind),
    Comparison = .data$comparison,
    `Planned outcomes` = .data$planned_n,
    Adjustment = .data$adjustment_method,
    Rule = .data$decision_rule
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = `Planned outcomes`) |>
  prep_gt(10)
```

## 5. Diagnostics and influence

Sixteen domains were interpreted for each of 24 primary metric-by-instrument-
by-placement targets. Stored checks cover questionnaire provenance, sample and
join identity, site support, clock representation, linearity, residual
distribution and spread, convergence, singularity, rank, temporal dependence,
participant and site influence, placement identity, preparation sensitivity,
and multiplicity.

Visual inspection of the complete residual panels supported acceptable final
response/residual distribution and heteroscedasticity assessments. The
quantitative screen values remain stored. Eighteen targets are acceptable
across all domains; six retain an influence, interaction-fit, or unavailable
preparation-sensitivity limitation.

```{r}
#| label: tbl-h09-prep-diagnostic-summary
#| tbl-cap: "Stored diagnostic assessment summary and retained quantitative screens."

tibble::tribble(
  ~Record, ~Targets_or_rows, ~Acceptable, ~Not_acceptable_or_flagged, ~Interpretation,
  "Overall target assessment", nrow(diagnostic_summary), sum(diagnostic_summary$overall_assessment == "acceptable"), sum(diagnostic_summary$overall_assessment == "not acceptable"), "Final interpreted assessment across 16 domains.",
  "Integrated distribution screen", nrow(diagnostic_summary), sum(diagnostic_adjudication$domain == "Response and residual distribution" & diagnostic_adjudication$numeric_screen_assessment == "acceptable"), sum(diagnostic_adjudication$domain == "Response and residual distribution" & diagnostic_adjudication$numeric_screen_assessment == "not acceptable"), "Eight conservative flags retained before visual adjudication.",
  "Lower-level distribution screen", nrow(residual_diagnostics), sum(residual_diagnostics$distribution_assessment == "acceptable"), sum(residual_diagnostics$distribution_assessment == "not acceptable"), "Fourteen all-threshold flags retained, including outlier-only flags.",
  "Heteroscedasticity screen", nrow(residual_diagnostics), sum(residual_diagnostics$heteroscedasticity_assessment == "acceptable"), sum(residual_diagnostics$heteroscedasticity_assessment == "not acceptable"), "Four conservative flags retained before visual adjudication."
) |>
  gt::gt() |>
  gt::fmt_integer(
    columns = c(Targets_or_rows, Acceptable, Not_acceptable_or_flagged)
  ) |>
  gt::cols_label(
    Targets_or_rows = "Targets or rows",
    Not_acceptable_or_flagged = "Not acceptable or flagged"
  ) |>
  prep_gt(9)
```

The complete [diagnostic registry](../../../artifacts/08_diagnostics/H09/H09_diagnostic_assessment_registry.csv),
[screen-to-final mapping](../../../artifacts/08_diagnostics/H09/H09_diagnostic_author_adjudication.csv),
[participant-influence summary](../../../artifacts/08_diagnostics/H09/H09_participant_influence_summary.csv),
and [leave-one-site-out summary](../../../artifacts/08_diagnostics/H09/H09_leave_one_site_out_summary.csv)
preserve the values and assessment basis. No diagnostic flag automatically
removes an observation.

## 6. Sensitivity analyses

The stored sensitivity battery separates dataset preparation, sample
composition, seasonal timing, repeated-measure dependence, clock
linearization, placement, and the distinct mean-timing estimand. These are
comparisons of prespecified or approved analysis components, not new outcome
searches.

```{r}
#| label: tbl-h09-prep-sensitivity-map
#| tbl-cap: "H09 sensitivity branches, stored scope, and interpretive role."

registered_gap <- gap_sensitivity |>
  filter(.data$metric_id != "mean_timing_above_250")
estimable_registered_gap <- registered_gap |>
  filter(.data$metric_id != "longest_period_midpoint")

tibble::tribble(
  ~Sensitivity, ~Stored_scope, ~Question, ~Reporting_rule,
  "Gap-timing-unaware exact common sample", paste0(nrow(estimable_registered_gap), " estimable comparisons; ", sum(estimable_registered_gap$exact_common_keys), " exact key matches"), "Does the registered estimand change when remaining-gap time of day is not used?", "Primary and sensitivity fits remain separate; the registered longest-period outcome is unavailable.",
  "Exactly paired placement", paste0(nrow(paired_effects), " metric-by-instrument comparisons; all exact samples"), "Are near-eye and chest component estimates similar on identical participant-days?", "No pooling, difference interval, or equivalence claim.",
  "Within-site photoperiod", paste0(nrow(photoperiod), " stored fits"), "Does linear within-site photoperiod adjustment materially change the chronotype slope?", "Photoperiod enters only as a centered within-site additive covariate.",
  "Equal-weight participant summary", paste0(nrow(participant_summary), " stored fits"), "Does giving each participant one outcome summary change the pattern?", "Fixed-site participant-level sensitivity, reported separately.",
  "Continuous-time AR(1)", paste0(nrow(ar1), " stored fits"), "Does within-participant temporal correlation materially change the slope?", "Elapsed day enters only the correlation structure.",
  "L10 clock cut", paste0(nrow(l10_cut), " stored comparisons"), "Does the approved negative-hour conversion depend on the afternoon cut?", "Direction, interval overlap, and slope change are retained.",
  "Mean timing above 250 lx melEDI", paste0(nrow(fifth_outcome |> filter(.data$metric_id == "mean_timing_above_250")), " stored estimates"), "What does the distinct duration-weighted mean-timing estimand show?", "Outside the registered five-outcome families; not substituted for longest-period midpoint."
) |>
  gt::gt() |>
  gt::cols_width(
    Sensitivity ~ gt::pct(22),
    Stored_scope ~ gt::pct(20),
    Question ~ gt::pct(30),
    Reporting_rule ~ gt::pct(28)
  ) |>
  prep_gt(9)
```

Complete sensitivity records retain exact fitted samples, 95% confidence
intervals, and stability classifications:

- [dataset preparation](../../../artifacts/09_tables/H09/H09_gap_timing_unaware_sensitivity.csv);
- [paired placement](../../../artifacts/09_tables/H09/H09_paired_placement_effects.csv);
- [within-site photoperiod](../../../artifacts/09_tables/H09/H09_photoperiod_sensitivity.csv);
- [participant summaries](../../../artifacts/09_tables/H09/H09_participant_summary_sensitivity.csv);
- [continuous-time AR(1)](../../../artifacts/09_tables/H09/H09_ar1_sensitivity.csv);
- [L10 clock cut](../../../artifacts/09_tables/H09/H09_l10_cut_sensitivity.csv); and
- [registered midpoint and mean-timing estimands](../../../artifacts/09_tables/H09/H09_fifth_outcome_sensitivity.csv).

## 7. Code map and execution order

### Function modules

```{r}
#| label: tbl-h09-prep-code-map
#| tbl-cap: "H09 and shared modules used by the analytical and reporting chain."

tibble::tribble(
  ~Path, ~Responsibility, ~Scientific_role, ~Boundary,
  "scripts/pipeline/paths_io.R", "Typed artifact writing and SHA-256 helpers", "Records file identity, bytes, producer, and R version.", "No scientific model logic.",
  "scripts/pipeline/p_value_display.R", "P-value display rules", "Formats p-values after unrounded decisions are made.", "No inferential decision is recalculated for display.",
  "scripts/hypotheses/H09/h09_contract.R", "Inputs, predictors, metrics, runs, formulae, families, and thresholds", "Defines the complete H09 specification before fitting.", "Contains no fitted result.",
  "scripts/hypotheses/H09/h09_modeling.R", "Frame construction, model fitting, effects, diagnostics, and sensitivity helpers", "Implements scientific operations against the contract.", "Called only by the analysis driver, never by this page."
) |>
  gt::gt() |>
  prep_gt(9)
```

### Executable and reporting entry points

```{r}
#| label: tbl-h09-prep-script-map
#| tbl-cap: "H09 scripts and documents in analytical and reporting order."

tibble::tribble(
  ~Order, ~Entry_point, ~Reads, ~Writes_or_displays, ~Runs_scientific_models,
  1L, "run_h09_stage2.R", "Pinned inputs plus H09 contract and modeling functions", "Frames, stored fits, effects, tests, diagnostics, sensitivities, figures, source CSVs, and execution records", "Yes; bounded analysis driver",
  2L, "finalize_h09_figure_qa.R", "Figure manifest and final PNG/PDF exports", "Final-size visual-QA assessments", "No",
  3L, "test_h09_stage2.R", "Scientific and reporting artifacts", "Hypothesis-scoped assertions", "No",
  4L, "H09.qmd", "Frozen effects, tests, diagnostics, sensitivities, and figures", "Standalone scientific results report", "No",
  5L, "build_h09_stage3_manifest.R", "Reader report plus accepted H09 chain", "Reader-report checksum inventory", "No",
  6L, "H09_analysis_preparation.qmd", "Frozen registries, indexes, diagnostics, sensitivities, and manifests", "This preparation and provenance companion", "No",
  7L, "build_h09_preparation_report_manifest.R", "Final source, render, scripts, outputs, source data, and environment", "Non-circular preparation-report manifest", "No",
  8L, "test_h09_preparation_report.R", "Preparation source, render, reciprocal links, and manifest", "Hypothesis-scoped structural and provenance assertions", "No"
) |>
  gt::gt() |>
  gt::fmt_integer(columns = Order) |>
  gt::cols_label(
    Entry_point = "Entry point",
    Writes_or_displays = "Writes or displays",
    Runs_scientific_models = "Runs scientific models"
  ) |>
  prep_gt(8)
```

The scientific build was recorded under R 4.6.1 with the existing project
library. No production bootstrap, simulation, or other resampling was run.

```text
Rscript --vanilla scripts/hypotheses/H09/run_h09_stage2.R
Rscript --vanilla scripts/hypotheses/H09/finalize_h09_figure_qa.R
Rscript --vanilla tests/hypotheses/H09/test_h09_stage2.R
quarto render notebooks/hypotheses/H09.qmd --profile nathealth
quarto render audit/hypotheses/H09/H09_analysis_preparation.qmd --profile nathealth
```

## 8. Output and checksum provenance

The accepted reader-report manifest inventories code, tests, contracts,
scientific outputs, figures, source data, and the rendered result. On render,
this page verifies every manifest identity except the two continuing handoff
documents, which are intentionally updated at later review points.

```{r}
#| label: tbl-h09-prep-reader-manifest-check
#| tbl-cap: "Bounded verification of the accepted H09 reader-report manifest."

mutable_records <- c(
  "audit/handoffs/H09_worker_handoff.md",
  "audit/handoffs/H09_shared_change_request.md"
)
immutable_reader_manifest <- reader_manifest |>
  filter(!.data$path %in% mutable_records)
manifest_paths <- file.path(root, immutable_reader_manifest$path)
manifest_exists <- file.exists(manifest_paths)
manifest_hash_match <- rep(FALSE, length(manifest_paths))
manifest_hash_match[manifest_exists] <- vapply(
  manifest_paths[manifest_exists],
  artifact_sha256,
  character(1)
) == immutable_reader_manifest$sha256[manifest_exists]

manifest_check <- tibble::tibble(
  Check = c(
    "Immutable manifest paths exist",
    "Immutable SHA-256 identities match",
    "Manifest rows are unique",
    "Recorded R version"
  ),
  Observed = c(
    paste0(sum(manifest_exists), " / ", length(manifest_exists)),
    paste0(sum(manifest_hash_match), " / ", length(manifest_hash_match)),
    paste0(nrow(reader_manifest) - anyDuplicated(reader_manifest$path), " unique-row result"),
    paste(sort(unique(reader_manifest$r_version)), collapse = ", ")
  ),
  Status = c(
    if_else(all(manifest_exists), "PASS", "FAIL"),
    if_else(all(manifest_hash_match), "PASS", "FAIL"),
    if_else(!anyDuplicated(reader_manifest$path), "PASS", "FAIL"),
    if_else(all(reader_manifest$r_version == "4.6.1"), "PASS", "FAIL")
  )
)
stopifnot(all(manifest_check$Status == "PASS"))

manifest_check |>
  gt::gt() |>
  prep_gt(10)
```

```{r}
#| label: tbl-h09-prep-key-output-identities
#| tbl-cap: "Selected reader-report, result, diagnostic, sensitivity, and figure identities."

key_output_paths <- c(
  "notebooks/hypotheses/H09.qmd",
  "_build/nathealth/notebooks/hypotheses/H09.html",
  "artifacts/09_tables/H09/H09_model_results_master.csv",
  "artifacts/08_diagnostics/H09/H09_diagnostic_assessment_registry.csv",
  "artifacts/08_diagnostics/H09/H09_diagnostic_author_adjudication.csv",
  "artifacts/09_tables/H09/H09_gap_timing_unaware_sensitivity.csv",
  "artifacts/10_figures/H09/H09_primary_effects.png",
  "artifacts/11_source_data/H09/H09_primary_effects_data.csv",
  "artifacts/12_manifests/H09/H09_figure_manifest.csv",
  "artifacts/12_manifests/H09/H09_execution_record.csv"
)

reader_manifest |>
  filter(.data$path %in% key_output_paths) |>
  mutate(
    `SHA-256 prefix` = substr(.data$sha256, 1L, 16L),
    `Bytes` = .data$bytes
  ) |>
  select(Path = .data$path, `SHA-256 prefix`, Bytes) |>
  gt::gt() |>
  gt::fmt_integer(columns = Bytes, use_seps = TRUE) |>
  gt::tab_source_note(
    "Complete identities remain in artifacts/12_manifests/H09/H09_stage3_artifacts.csv."
  ) |>
  prep_gt(9)
```

The final [figure manifest](../../../artifacts/12_manifests/H09/H09_figure_manifest.csv)
records PNG/PDF paths, intended display width, export scaling, final-size text,
visual inspection, alt text, and paired source CSVs. The
[reader-report manifest](../../../artifacts/12_manifests/H09/H09_stage3_artifacts.csv)
provides the complete accepted result-chain inventory. A separate preparation
manifest seals this page after rendering, so the page never attempts to hash
its own final HTML during its creation.

## 9. How to examine or reproduce H09

1. Verify the 11 result-producing inputs against
   `H09_input_audit.csv`.
2. Inspect the chronotype support and centering constants in
   `H09_chronotype_score_audit.csv` and the exact value-preservation audit.
3. Inspect metric, predictor, run, formula, family, and threshold registries.
4. Inspect the 108-row model-frame index and exact row/frame hashes before
   opening the stored frame archive.
5. Re-run the bounded analysis driver only under R 4.6.1 and the existing
   project library when scientific reproduction is intended.
6. Compare effects, tests, complete-family adjustment, diagnostics,
   sensitivities, figures, and source CSVs with the reader report.
7. Verify the final reader and preparation manifests against the current
   checkout.

No package installation, lockfile update, full-project render, or unrestricted
environment scan is required for these checks.

## Environment

```{r}
#| label: tbl-h09-prep-execution
#| tbl-cap: "Recorded scientific execution environment."

execution |>
  transmute(
    `R version` = .data$r_version,
    Platform = .data$platform,
    `Model frames` = .data$model_frames,
    `Stored model objects` = .data$fitted_model_objects,
    `Primary diagnostic targets` = .data$diagnostic_targets,
    `Production resampling replicates` =
      .data$production_resampling_replicates,
    `Run completed (UTC)` = .data$run_completed_utc
  ) |>
  gt::gt() |>
  gt::fmt_integer(
    columns = c(
      `Model frames`,
      `Stored model objects`,
      `Primary diagnostic targets`,
      `Production resampling replicates`
    ),
    use_seps = TRUE
  ) |>
  prep_gt(9)
```

The complete [package-version record](../../../artifacts/12_manifests/H09/H09_package_versions.csv)
and project `renv.lock` identify the software environment. The preparation
bundle retains the declared PREP-003/FIND-044 qualification: not every shared
metric and support value has an independent current-manifest reconstruction.
That qualification is carried forward rather than repaired or hidden here.

```{r}
#| label: h09-preparation-session-information
#| code-fold: true
#| code-summary: "R session information"

sessionInfo()
```
