---
title: "H02 analysis preparation and provenance"
subtitle: "From verified 30-minute data to reported results"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
    lightbox: true
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(knitr)
  library(readr)
  library(scales)
  library(tibble)
  library(tidyr)
})

root_candidates <- unique(c(
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(
    file.path(getwd(), "../../.."),
    winslash = "/",
    mustWork = FALSE
  )
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(
        path,
        "scripts/hypotheses/H02/h02_contract.R"
      ))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(root.dir = root)

source(file.path(root, "scripts/pipeline/paths_io.R"))
source(file.path(root, "scripts/pipeline/p_value_display.R"))
source(file.path(root, "scripts/hypotheses/H02/h02_contract.R"))

if (getRversion() != "4.6.1") {
  stop("H02 preparation requires R 4.6.1", call. = FALSE)
}

# This validates the source paths and SHA-256 identities. It does not
# transform data or fit a model.
input_audit <- h02_validate_inputs(root)
spec <- h02_model_specification()

read_h02_csv <- function(path) {
  readr::read_csv(file.path(root, path), show_col_types = FALSE)
}

relative_path <- function(path) {
  sub(
    paste0(
      "^",
      gsub("([][{}()+*^$|\\\\?.])", "\\\\\\1", root),
      "/"
    ),
    "",
    normalizePath(path, winslash = "/", mustWork = TRUE)
  )
}

prep_gt <- function(table) {
  table |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(13),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "bold",
      source_notes.font.size = gt::px(11)
    )
}

format_count <- function(x) {
  format(x, big.mark = ",", scientific = FALSE, trim = TRUE)
}

placement_labels <- c(
  glasses = "Near eye",
  chest = "Chest"
)
placement_colors <- c(
  "Near eye" = "#0072B2",
  "Chest" = "#D55E00"
)

site_registry <- read_h02_csv("config/site_display_registry.csv") |>
  arrange(.data$display_order)

near_frame_path <- file.path(
  root,
  "artifacts/06_model_data/H02/main__glasses__all_available.rds"
)
chest_frame_path <- file.path(
  root,
  "artifacts/06_model_data/H02/main__chest__all_available.rds"
)
near_frame <- readRDS(near_frame_path)
chest_frame <- readRDS(chest_frame_path)
primary_frames <- bind_rows(
  near_frame |> mutate(Placement = "Near eye"),
  chest_frame |> mutate(Placement = "Chest")
) |>
  mutate(
    Placement = factor(
      .data$Placement,
      levels = c("Near eye", "Chest")
    )
  )

sample_counts <- read_h02_csv(
  "artifacts/06_model_data/H02/sample_counts.csv"
)
support_audit <- read_h02_csv(
  "artifacts/06_model_data/H02/support_audit.csv"
)
residual_acf <- read_h02_csv(
  "artifacts/08_diagnostics/H02/residual_acf.csv"
)
ar_boundary_audit <- read_h02_csv(
  "artifacts/08_diagnostics/H02/ar_boundary_audit.csv"
)
model_fit <- read_h02_csv(
  "artifacts/09_tables/H02/model_fit_summary.csv"
)
model_comparisons <- read_h02_csv(
  "artifacts/09_tables/H02/model_structure_comparisons.csv"
)

near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
```

## About this analysis record

The scientific question was whether daily patterns of **melanopic equivalent
daylight illuminance (melEDI)** differed among sites and whether differences
among people within sites were larger than the differences among sites. The
near-eye sensor position was primary because it measures light closer to the
eye, but it is not a direct retinal measurement. The chest sensor position was
analysed separately as complementary evidence and is not a measure of ocular
exposure.

A **participant-day** is one participant's observations on one local calendar
date. Participant-day counts therefore describe repeated recorded days, not
additional independent participants.

This page explains **which data and code produced the reported H02 result, why
each analytical step was needed, and what evidence shows that the processing
and model behaved as intended**. The companion
[H02 results report](../../../notebooks/hypotheses/H02.qmd) presents the
scientific findings.

::: {.callout-note title="How to use this provenance page"}
This page explains the input and sample checks, model construction, model
checks, and links between analysis code and accepted stored outputs. Rendering
verifies file identities. A **model frame** is the exact set of rows and
variables used for one fitted model. Rendering also calculates descriptive
summaries from the H02 model frames. It does **not** fit a model, estimate an
autocorrelation parameter, calculate model predictions, rerun a bootstrap, or
recompute a Shapley allocation; those results are read from the listed output
files.
The result report lists the [H02 preregistration
deviations](../../../notebooks/hypotheses/H02.qmd#h02-preregistration-deviations).
:::

## Analysis map

```{mermaid}
flowchart TD
  A["Verified 30-minute near-eye and chest data<br/>plus true-time provenance"] --> B["Input identity and content checks"]
  B --> C["H02-only model frames<br/>placement × preparation × sample"]
  C --> D["Primary near-eye model comparison<br/>selected temporal structure"]
  D --> E["Same selected structure<br/>applied to every model frame"]
  E --> F["Fitted curves and<br/>variation summaries"]
  E --> G["Model checks<br/>and influence assessment"]
  E --> H["Conditional Shapley allocation<br/>near eye + chest"]
  E --> I["Sensitivity analyses"]
  F --> J["Pointwise curves, tables,<br/>and source data"]
  G --> J
  H --> J
  I --> J
  J --> K["H02 results report"]
  L["Recorded file and contract checks"] -. verify .-> B
  L -. verify .-> C
  L -. verify .-> E
  L -. verify .-> J
```

The arrows represent data or fitted-output dependencies. They do not imply
that this Quarto document reruns the downstream analyses.

## Render boundary

```{r}
#| label: tbl-h02-prep-boundary
#| tbl-cap: "Calculations performed when this page is rendered."

tibble::tribble(
  ~Category, ~Calculated_on_render, ~Explanation,
  "Source-file verification", "Yes", "Confirms that the files entering H02 retain their recorded SHA-256 identities.",
  "Descriptive input summaries", "Yes", "Shows fitted samples, exact-zero observations, temporal support, and participant-day coverage.",
  "Model fitting and selection", "No", "Reads stored fit summaries and model records rather than creating another fit.",
  "Uncertainty resampling", "No", "Reads stored interval estimates and resampling summaries.",
  "Fitted curves and model checks", "No", "Displays or references outputs generated from the stored fitted models."
) |>
  gt::gt() |>
  gt::cols_label(
    Calculated_on_render = "Calculated when rendered"
  ) |>
  gt::cols_width(
    Category ~ gt::pct(22),
    Calculated_on_render ~ gt::pct(18),
    Explanation ~ gt::pct(60)
  ) |>
  prep_gt()
```

## 1. Data entering the analysis

### Verifying the source files

The analysis starts from named, versioned files rather than an informal copy
of the data. The function
`h02_validate_inputs()` in `scripts/hypotheses/H02/h02_contract.R` compares
the current files with pinned SHA-256 identities and checks the relevant base
and temporal-provenance manifest rows. Every fitting and figure-building
script performs this check before proceeding.

Two prepared datasets enter the analysis. The primary dataset could be
interpreted as a time-sensitive primary metric dataset because its preparation
uses the timing of remaining missing observations where that timing is
relevant to the metric. The predefined sensitivity is the
**gap-timing-unaware dataset**. It still passed the general 50%-per-hour and
80%-per-day coverage rules, but the timing of the remaining missing
observations was not used for an additional metric-specific adjustment. This
name does not mean that gaps, missingness, or coverage were ignored.

```{r}
#| label: tbl-h02-prep-inputs
#| tbl-cap: "Verified inputs used by the H02 analysis."

current_input_ids <- c(
  "main_glasses",
  "main_chest",
  "base_model_data_manifest",
  "manuscript_prepared",
  "wall_outcome_links",
  "true_utc_source_bins",
  "temporal_provenance_manifest"
)
input_role_labels <- c(
  main_glasses = "Verified near-eye 30-minute base data",
  main_chest = "Verified chest 30-minute base data",
  base_model_data_manifest = "Base-data verification manifest",
  manuscript_prepared = "Gap-timing-unaware dataset",
  wall_outcome_links = "Wall-clock to true-time links",
  true_utc_source_bins = "True-time source-bin ordering",
  temporal_provenance_manifest = "Temporal-provenance verification manifest"
)

input_audit |>
  filter(.data$input_id %in% current_input_ids) |>
  mutate(
    Input = unname(input_role_labels[.data$input_id]),
    Path = relative_path(.data$path),
    `SHA-256 prefix` = substr(.data$sha256, 1L, 16L),
    Verified = if_else(
      .data$hash_verified,
      "Verified",
      "Review needed"
    )
  ) |>
  select(.data$Input, .data$Path, .data$`SHA-256 prefix`, .data$Verified) |>
  gt::gt() |>
  gt::cols_width(
    Input ~ gt::pct(25),
    Path ~ gt::pct(50),
    `SHA-256 prefix` ~ gt::pct(17),
    Verified ~ gt::pct(8)
  ) |>
  gt::tab_source_note(
    "The complete 64-character identities are recorded in artifacts/06_model_data/H02/input_hashes.csv."
  ) |>
  prep_gt()
```

<details>
<summary>Show the exact current SHA-256 identities</summary>

```{r}
#| echo: false
#| results: asis

exact_hashes <- input_audit |>
  filter(.data$input_id %in% current_input_ids) |>
  transmute(line = paste0("- `", .data$input_id, "`: `", .data$sha256, "`"))
cat(paste(exact_hashes$line, collapse = "\n"))
```

</details>

### What the source data already contain

The source files already contain 30-minute arithmetic-mean melEDI,
coverage and admissibility information, local dates and wall-clock bins, and
participant/site identifiers. H02 does not recalculate these measurements.
It joins them to verified true-time provenance so that local wall time can
describe the daily profile while true UTC controls observation ordering.

The primary dataset admitted a 30-minute bin when at least 15 one-minute
melEDI values were valid. Unsupported bins remained missing; they were not
converted to darkness. The support audit below is an output of
`build_h02_model_data.R`, not a calculation introduced for this document.

```{r}
#| label: tbl-h02-prep-support
#| tbl-cap: "Observation support carried from the source data into the H02 model frames."

support_audit |>
  mutate(
    Preparation = recode(
      .data$data_scenario_id,
      main = "Primary dataset",
      manuscript_prepared_data = "Gap-timing-unaware dataset"
    ),
    Placement = unname(placement_labels[.data$placement]),
    `Grid rows` = format_count(.data$grid_rows),
    `Finite admissible` = format_count(.data$finite_admissible),
    `Unsupported/non-finite` = format_count(
      .data$unsupported_or_nonfinite
    )
  ) |>
  select(
    .data$Preparation,
    .data$Placement,
    .data$participants,
    .data$participant_days,
    .data$`Grid rows`,
    .data$`Finite admissible`,
    .data$`Unsupported/non-finite`
  ) |>
  gt::gt() |>
  gt::cols_label(
    participants = "Participants",
    participant_days = "Participant-days"
  ) |>
  gt::cols_align(
    align = "right",
    columns = c(
      participants,
      participant_days,
      `Grid rows`,
      `Finite admissible`,
      `Unsupported/non-finite`
    )
  ) |>
  prep_gt()
```

## 2. Fitted samples and model frames

### Code and purpose

`scripts/hypotheses/H02/build_h02_model_data.R` sources two function modules:

- `h02_contract.R` fixes the input identities, model specification, scenario
  registry, transform, and reproducible seeds;
- `h02_data.R` validates keys, joins temporal provenance, creates the fitted
  support, defines residual-sequence starts, and constructs the all-available
  and placement-matched samples.

The script writes eight H02-only frames: two preparation rules × two
placements × two sample definitions. The all-available near-eye frame is the
primary sample. The all-available chest frame is complementary. The
placement-matched frames retain only participant-day clock bins available for
both placements and are sensitivity analyses, not a pooled model.

Within each frame, participant identifiers are prefixed by site, and
participant-day identifiers combine site, participant, and local date. Rows are
ordered using verified true elapsed time within those groups; local wall-clock
time remains the coordinate used to describe the daily profile.

```{r}
#| label: tbl-h02-prep-scenarios
#| tbl-cap: "H02 model frames created before any model was fitted."

scenario_labels <- c(
  main__all_available = "Main, all available",
  main__paired_common_sample = "Main, placement-matched bins",
  manuscript_prepared_data__all_available =
    "Gap-timing-unaware dataset, all available",
  manuscript_prepared_data__paired_common_sample =
    "Gap-timing-unaware dataset, placement-matched bins"
)

sample_counts |>
  filter(.data$site == "ALL_SITES") |>
  tidyr::separate(
    .data$run_id,
    into = c("Preparation_id", "Placement_id", "Sample_id"),
    sep = "__",
    remove = FALSE
  ) |>
  mutate(
    Scenario_key = paste(.data$Preparation_id, .data$Sample_id, sep = "__"),
    Scenario = unname(scenario_labels[.data$Scenario_key]),
    Placement = unname(placement_labels[.data$Placement_id]),
    `Zero observations` = percent(
      .data$exact_zero_observations / .data$observations_30_minute,
      accuracy = 0.1
    )
  ) |>
  arrange(.data$Scenario, factor(.data$Placement, c("Near eye", "Chest"))) |>
  select(
    .data$Scenario,
    .data$Placement,
    .data$participants,
    .data$participant_days,
    .data$observations_30_minute,
    .data$sites,
    .data$`Zero observations`
  ) |>
  gt::gt() |>
  gt::cols_label(
    participants = "Participants",
    participant_days = "Participant-days",
    observations_30_minute = "30-minute observations",
    sites = "Sites"
  ) |>
  gt::fmt_number(
    columns = c(participants, participant_days, observations_30_minute, sites),
    decimals = 0,
    use_seps = TRUE
  ) |>
  prep_gt()
```

### Primary fitted samples by site

The next two tables are calculated directly from the H02-specific primary
frames during rendering. They are descriptive checks of the inputs to the
models, not model results. Site names, order, and colours come from the shared
display registry.

```{r}
#| label: prep-site-summary
#| include: false

site_summary <- primary_frames |>
  group_by(.data$Placement, .data$site) |>
  summarise(
    Participants = n_distinct(.data$participant_key),
    `Participant-days` = n_distinct(.data$participant_day_key),
    Observations = dplyr::n(),
    `Exact zero` = mean(.data$metric_value_lx == 0),
    `Median melEDI (lx)` = median(.data$metric_value_lx),
    `90th percentile (lx)` = quantile(.data$metric_value_lx, 0.90),
    .groups = "drop"
  ) |>
  left_join(
    site_registry |>
      select(.data$site, .data$display_name, .data$display_order),
    by = "site"
  ) |>
  arrange(.data$Placement, .data$display_order)
```

```{r}
#| label: tbl-h02-prep-near-site-sample
#| tbl-cap: "Primary near-eye model-frame distribution by site."

site_summary |>
  filter(.data$Placement == "Near eye") |>
  select(
    Site = .data$display_name,
    .data$Participants,
    .data$`Participant-days`,
    .data$Observations,
    .data$`Exact zero`,
    .data$`Median melEDI (lx)`,
    .data$`90th percentile (lx)`
  ) |>
  gt::gt() |>
  gt::fmt_percent(columns = `Exact zero`, decimals = 1) |>
  gt::fmt_number(
    columns = c(`Median melEDI (lx)`, `90th percentile (lx)`),
    decimals = 1,
    use_seps = TRUE
  ) |>
  gt::fmt_number(
    columns = c(Participants, `Participant-days`, Observations),
    decimals = 0,
    use_seps = TRUE
  ) |>
  prep_gt()
```

```{r}
#| label: tbl-h02-prep-chest-site-sample
#| tbl-cap: "Complementary chest model-frame distribution by site."

site_summary |>
  filter(.data$Placement == "Chest") |>
  select(
    Site = .data$display_name,
    .data$Participants,
    .data$`Participant-days`,
    .data$Observations,
    .data$`Exact zero`,
    .data$`Median melEDI (lx)`,
    .data$`90th percentile (lx)`
  ) |>
  gt::gt() |>
  gt::fmt_percent(columns = `Exact zero`, decimals = 1) |>
  gt::fmt_number(
    columns = c(`Median melEDI (lx)`, `90th percentile (lx)`),
    decimals = 1,
    use_seps = TRUE
  ) |>
  gt::fmt_number(
    columns = c(Participants, `Participant-days`, Observations),
    decimals = 0,
    use_seps = TRUE
  ) |>
  prep_gt()
```

### Why the response was transformed

The fitted response is
$\log_{10}(\mathrm{melEDI}+0.1\ \mathrm{lx})$. The offset retains exact-zero
observations and maps them to −1. The transformation compresses a range that
extends from darkness to tens of thousands of lux, but it does not make the
zero mass disappear. Because this discrete spike is much taller than any bin
among the positive measurements, combining both in one histogram hides the
shape of the positive distribution. The figure therefore reports the number
and percentage of exact zeros separately and uses its vertical scale for
positive melEDI observations only.

```{r}
#| label: prepare-h02-response-distribution
#| include: false

response_distribution_positive <- primary_frames |>
  filter(.data$metric_value_lx > 0) |>
  select(.data$Placement, .data$response)

response_distribution_zeros <- primary_frames |>
  group_by(.data$Placement) |>
  summarise(
    exact_zero_observations = sum(.data$metric_value_lx == 0),
    observations = dplyr::n(),
    exact_zero_percentage =
      100 * .data$exact_zero_observations / .data$observations,
    annotation = paste0(
      "Exact zero (maps to −1): ",
      format_count(.data$exact_zero_observations),
      " (",
      formatC(.data$exact_zero_percentage, format = "f", digits = 1),
      "%)"
    ),
    .groups = "drop"
  )

```

The accepted descriptive values are retained in the frozen
[positive-observation source data](../../../artifacts/11_source_data/H02/preparation_response_distribution_positive_observations.csv)
and [exact-zero summary](../../../artifacts/11_source_data/H02/preparation_response_distribution_exact_zero_summary.csv).

```{r}
#| label: fig-h02-prep-response-distribution
#| fig-cap: "Distribution of positive transformed responses entering the two primary models. Exact-zero observations, which map to −1, are reported separately inside each panel."
#| fig-alt: "Two histograms show the transformed positive melEDI observations for the near-eye and chest model frames on separately scaled vertical axes. Both distributions rise broadly between transformed values of zero and about 2.5 before declining. Panel annotations report that approximately one third of observations were exact zero and therefore mapped to minus one."
#| fig-width: 10
#| fig-height: 6.5

ggplot(
  response_distribution_positive,
  aes(x = .data$response, fill = .data$Placement)
) +
  geom_histogram(
    binwidth = 0.10,
    boundary = -1,
    colour = "white",
    linewidth = 0.12,
    show.legend = FALSE
  ) +
  geom_vline(xintercept = -1, colour = "grey25", linetype = "dashed") +
  geom_label(
    data = response_distribution_zeros,
    aes(x = -0.88, y = Inf, label = .data$annotation),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1.25,
    size = 3.4,
    linewidth = 0.2,
    label.padding = grid::unit(0.18, "lines"),
    fill = "white"
  ) +
  facet_wrap(vars(.data$Placement), ncol = 1, scales = "free_y") +
  scale_fill_manual(values = placement_colors) +
  scale_x_continuous(breaks = -1:5) +
  scale_y_continuous(
    labels = label_number(big.mark = ","),
    expand = expansion(mult = c(0, 0.12))
  ) +
  coord_cartesian(xlim = c(-1.05, 5), clip = "off") +
  labs(
    x = expression(log[10](melEDI + 0.1~lx)),
    y = "Positive 30-minute observations"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

The near-eye model frame contained `r format_count(response_distribution_zeros$exact_zero_observations[response_distribution_zeros$Placement == "Near eye"])`
exact-zero observations (`r formatC(response_distribution_zeros$exact_zero_percentage[response_distribution_zeros$Placement == "Near eye"], format = "f", digits = 1)`%),
and the chest frame contained `r format_count(response_distribution_zeros$exact_zero_observations[response_distribution_zeros$Placement == "Chest"])`
(`r formatC(response_distribution_zeros$exact_zero_percentage[response_distribution_zeros$Placement == "Chest"], format = "f", digits = 1)`%).
These observations remain in both fitted models; they are separated here only
to make the continuous part of the input distribution legible.

### Clock support and participant-day completeness

The temporal model uses a 48-bin wall-clock grid. Not every participant-day
has all 48 bins, because unsupported intervals remain absent. The heatmap
shows, for each site and clock bin, the percentage of fitted participant-days
that supplied an observation. It is calculated here from the two primary
model frames.

```{r}
#| label: fig-h02-prep-clock-support
#| fig-cap: "Clock-bin support in the primary near-eye and complementary chest model frames."
#| fig-alt: "Two heatmaps show the percentage of fitted participant-days contributing an observation at each half-hour of local time for every site. Most cells have high coverage, with modest site- and time-specific gaps."
#| fig-width: 11
#| fig-height: 7.5

site_day_denominators <- primary_frames |>
  group_by(.data$Placement, .data$site) |>
  summarise(
    participant_days = n_distinct(.data$participant_day_key),
    .groups = "drop"
  )

clock_support <- primary_frames |>
  count(.data$Placement, .data$site, .data$time_hour, name = "observed_days") |>
  left_join(site_day_denominators, by = c("Placement", "site")) |>
  mutate(coverage = .data$observed_days / .data$participant_days) |>
  left_join(
    site_registry |>
      select(.data$site, .data$display_name, .data$display_order),
    by = "site"
  ) |>
  mutate(
    display_name = factor(
      .data$display_name,
      levels = rev(site_registry$display_name)
    )
  )

ggplot(
  clock_support,
  aes(
    x = .data$time_hour,
    y = .data$display_name,
    fill = .data$coverage
  )
) +
  geom_tile() +
  facet_grid(
    rows = vars(.data$Placement),
    scales = "free_y",
    space = "free_y"
  ) +
  scale_x_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24),
    expand = c(0, 0)
  ) +
  scale_fill_viridis_c(
    option = "C",
    limits = c(0.75, 1),
    oob = scales::squish,
    breaks = seq(0.75, 1, by = 0.05),
    labels = label_percent(accuracy = 1)
  ) +
  guides(
    fill = guide_colorbar(
      title.position = "top",
      title.hjust = 0.5,
      label.position = "bottom",
      barwidth = grid::unit(9, "cm"),
      barheight = grid::unit(0.4, "cm")
    )
  ) +
  labs(
    x = "Local wall-clock time (hours)",
    y = NULL,
    fill = "Fitted days"
  ) +
  theme_minimal(base_size = 11.5) +
  theme(
    panel.grid = element_blank(),
    strip.text.y = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box.spacing = grid::unit(0.15, "cm")
  )
```

```{r}
#| label: fig-h02-prep-day-support
#| fig-cap: "Number of supported 30-minute observations per fitted participant-day."
#| fig-alt: "Two bar charts show that most fitted participant-days contributed 45 to 48 half-hour observations, while a smaller number contributed between 38 and 44 observations."
#| fig-width: 10
#| fig-height: 5.5

day_support <- primary_frames |>
  count(
    .data$Placement,
    .data$participant_day_key,
    name = "bins_30_minute"
  ) |>
  count(.data$Placement, .data$bins_30_minute, name = "participant_days")

ggplot(
  day_support,
  aes(
    x = .data$bins_30_minute,
    y = .data$participant_days,
    fill = .data$Placement
  )
) +
  geom_col(show.legend = FALSE) +
  facet_wrap(vars(.data$Placement), ncol = 1, scales = "free_y") +
  scale_fill_manual(values = placement_colors) +
  scale_x_continuous(breaks = seq(38, 48, by = 2)) +
  scale_y_continuous(labels = label_number(big.mark = ",")) +
  labs(
    x = "Supported 30-minute observations in a participant-day",
    y = "Participant-days"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())
```

## 3. Response transformation and nonlinear model construction

### Exact selected formula

H02 used a **nonlinear GAM analysis**. A generalized additive model (GAM)
allows the association with local clock time to bend across the day rather
than forcing it to follow a straight line. The fitted response is
`log10(melEDI + 0.1 lx)`; a **back-transformed** curve applies the inverse
transformation and returns to melEDI in lux.

`scripts/hypotheses/H02/h02_modeling.R` contains the fitting functions;
`h02_contract.R` supplies their fixed specification. The exact Wilkinson
formula was:

```r
response ~
  s(time_hour, bs = "cc", k = 12) +
  s(time_hour, site, bs = "sz", k = 12) +
  s(time_hour, participant, bs = "fs", k = 10) +
  s(participant_day, bs = "re")
```

The first smooth is the **global time effect**, the site-average daily curve:
it gives every site equal weight and joins smoothly at midnight. The `sz` term
estimates site-specific departures that sum to zero, so no site is treated as
the reference. The participant factor smooth represents remaining differences
in daily shape among people after the time and site patterns are considered.
The participant-day random effect represents remaining day-to-day level shifts
within a participant after those patterns are considered; it moves an entire
fitted day up or down.

Only the global time effect is forced to join at midnight in the primary
model. A sensitivity analysis also makes the site and participant deviations
cyclic.

### Model settings

```{r}
#| label: tbl-h02-prep-parameters
#| tbl-cap: "Model settings fixed before fitting and reused across H02 scenarios."

tibble::tribble(
  ~Parameter, ~Setting, ~Contribution,
  "Primary placement", "Near eye", "Measures light closer to the eye but not at the retina; chest is complementary, not ocular exposure, and never pooled.",
  "Outcome", spec$response_name, "Uses the verified 30-minute arithmetic-mean melEDI metric.",
  "Response scale", spec$response_transform, "Keeps exact zeros finite and compresses the exposure range.",
  "Clock coordinate", spec$clock_coordinate, "Describes local daily patterns; true UTC remains the ordering coordinate.",
  "Global time effect", "Cyclic cubic, k = 12, knots 0 and 24 h", "Joins smoothly at midnight.",
  "Site departure", "Sum-to-zero sz smooth, k = 12", "Creates identifiable deviations around the global time effect.",
  "Participant pattern", "Factor smooth, k = 10", "Uses one shared smoothing parameter for person-specific daily shapes.",
  "Participant-day", "Random intercept", "Separates day-level shifts from participant curve shape.",
  "Estimation", "mgcv::bam(), fREML, discrete = TRUE, one R thread", "Fits the large penalized model reproducibly in the project environment.",
  "Residual dependence", spec$autocorrelation, "Adds an AR(1) working correlation after estimating rho from a preliminary fit.",
  "AR boundaries", spec$autocorrelation_boundaries, "Prevents correlation from crossing independent days or time discontinuities.",
  "Variation intervals", spec$variation_ci, "Describes cluster-sampling uncertainty conditional on the fitted smooths.",
  "Multiplicity", "FDR adjustment", "One confirmatory near-eye omnibus site-pattern test.",
  "Dominance", spec$dominance_scope, "Allocates descriptive in-sample fit only for the primary near-eye and complementary chest models."
) |>
  gt::gt() |>
  gt::cols_width(
    Parameter ~ gt::pct(19),
    Setting ~ gt::pct(35),
    Contribution ~ gt::pct(46)
  ) |>
  prep_gt()
```

### How the selected model was applied

The primary near-eye model is fitted first by `run_h02_analysis.R`. Reduced
versions without the site
pattern, participant pattern, or participant-day intercept are fitted with
the same estimated AR parameter to compare model structures. The full
structure is retained, and the identical selected structure is then applied
to the other seven H02 frames. Each scenario estimates its own boundary-aware
AR parameter from its preliminary residuals.

For every fitted scenario, the script writes the selected model, a checkpoint,
site-curve predictions, fitted term contributions, residual summaries, and
the stored 2,000-draw hierarchical variation bootstrap. The following fit
table is read from `model_fit_summary.csv`; no fit is performed here.

```{r}
#| label: tbl-h02-prep-primary-fits
#| tbl-cap: "Recorded primary near-eye and complementary chest fits."

model_fit |>
  filter(
    .data$run_id %in% c(near_id, chest_id),
    .data$model_id %in% c("site_pattern_final_fREML", "site_pattern")
  ) |>
  group_by(.data$run_id) |>
  slice_tail(n = 1L) |>
  ungroup() |>
  mutate(
    Placement = recode(
      .data$run_id,
      !!near_id := "Near eye",
      !!chest_id := "Chest"
    )
  ) |>
  select(
    .data$Placement,
    Observations = .data$n,
    .data$total_edf,
    .data$rank,
    .data$coefficients,
    .data$rho,
    .data$convergence
  ) |>
  gt::gt() |>
  gt::cols_label(
    total_edf = "Total edf",
    rank = "Rank",
    coefficients = "Coefficients",
    rho = "AR rho",
    convergence = "Converged"
  ) |>
  gt::fmt_number(columns = total_edf, decimals = 1) |>
  gt::fmt_number(
    columns = c(Observations, rank, coefficients),
    decimals = 0,
    use_seps = TRUE
  ) |>
  gt::fmt_number(columns = rho, decimals = 3) |>
  prep_gt()
```

```{r}
#| label: tbl-h02-prep-structure-checks
#| tbl-cap: "Recorded primary near-eye structure comparisons."

comparison_labels <- c(
  site_pattern_vs_no_site = "Add site-pattern deviations",
  participant_pattern = "Add participant-pattern smooths",
  participant_day = "Add participant-day intercepts"
)

model_comparisons |>
  mutate(
    Comparison = unname(comparison_labels[.data$comparison_id]),
    `AIC improvement` = .data$delta_AIC_reduced_minus_full,
    Role = if_else(
      .data$comparison_id == "site_pattern_vs_no_site",
      "Confirmatory omnibus site-pattern test",
      "Model-structure check"
    ),
    `FDR-adjusted p` = nh_format_p_value(
      .data$p_adjusted,
      na_label = "Not applicable"
    ),
    p_bold = !is.na(.data$p_adjusted) & .data$p_adjusted < 0.05
  ) |>
  select(
    .data$Comparison,
    .data$Role,
    .data$`AIC improvement`,
    Statistic = .data$test_statistic,
    `Degrees of freedom` = .data$test_df,
    .data$`FDR-adjusted p`,
    .data$p_bold
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(`AIC improvement`, Statistic, `Degrees of freedom`),
    decimals = 1
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(
      columns = `FDR-adjusted p`,
      rows = p_bold
    )
  ) |>
  gt::cols_hide(columns = p_bold) |>
  gt::tab_source_note(
    paste(
      "Only the site-pattern comparison belongs to the single confirmatory",
      "family. Bold denotes FDR-adjusted p < 0.05; the other rows are",
      "model-structure checks without an adjusted-p decision."
    )
  ) |>
  prep_gt()
```

## 4. Ordering and residual dependence

### Why local time and true time are both needed

**Autocorrelation** means that adjacent 30-minute residuals can remain more
alike than residuals farther apart. The AR(1) structure represents this
correlation as strongest for adjacent observations and weaker at longer lags.
It must restart wherever two observations are not part of the same continuous
recorded sequence.

Local wall-clock time answers the scientific question about time of day. It
cannot by itself establish elapsed ordering across daylight-saving folds,
gaps, or missing bins. `h02_data.R` therefore carries the verified true-UTC
source-bin links into each model frame, orders observations by true elapsed
time, and defines a new residual sequence:

- at the start of every participant-day;
- after a missing or otherwise discontinuous 30-minute interval; and
- on both sides of a wall-clock outcome that has no one-to-one elapsed-time
  coordinate.

The table is read from the recorded boundary check. “Continuous” rows are
ordinary observations inside an existing sequence and are included only to
make the denominator visible.

```{r}
#| label: tbl-h02-prep-ar-boundaries
#| tbl-cap: "Recorded residual-sequence boundary reasons in the two primary model frames."

ar_reason_labels <- c(
  participant_day_start = "Participant-day start",
  elapsed_discontinuity = "Elapsed-time discontinuity",
  non_one_to_one_wall_outcome = "Non-one-to-one wall outcome",
  after_non_one_to_one_wall_outcome = "After non-one-to-one wall outcome",
  continuous = "Continuous within sequence"
)

ar_boundary_audit |>
  filter(.data$run_id %in% c(near_id, chest_id)) |>
  mutate(
    Placement = recode(
      .data$run_id,
      !!near_id := "Near eye",
      !!chest_id := "Chest"
    ),
    Reason = unname(ar_reason_labels[.data$ar_start_reason])
  ) |>
  select(.data$Reason, .data$Placement, .data$observations) |>
  pivot_wider(names_from = .data$Placement, values_from = .data$observations) |>
  gt::gt(rowname_col = "Reason") |>
  gt::fmt_number(columns = everything(), decimals = 0, use_seps = TRUE) |>
  prep_gt()
```

### What the AR step contributed

The preliminary residual lag-1 correlation supplied $\rho$, the fitted
strength of adjacent-observation correlation. The final fit used that $\rho$
with the verified sequence boundaries. The plot below is drawn from
the stored `residual_acf.csv`; it does not calculate residuals from a model.

```{r}
#| label: fig-h02-prep-ar-change
#| fig-cap: "Recorded residual autocorrelation before and after the boundary-aware AR(1) correction."
#| fig-alt: "Line charts for near-eye and chest fits show residual correlation across lags of 30 to 180 minutes. Correlation is high before AR correction and close to zero after correction."
#| fig-width: 10
#| fig-height: 5.5

acf_display <- residual_acf |>
  filter(.data$run_id %in% c(near_id, chest_id)) |>
  mutate(
    Placement = recode(
      .data$run_id,
      !!near_id := "Near eye",
      !!chest_id := "Chest"
    ),
    Stage = recode(
      .data$stage,
      preliminary_no_AR1 = "Before AR correction",
      final_AR1_standardized = "After AR correction"
    ),
    `Lag (minutes)` = .data$lag_30_minute_bins * 30
  )

ggplot(
  acf_display,
  aes(
    x = .data$`Lag (minutes)`,
    y = .data$correlation,
    colour = .data$Stage,
    linetype = .data$Stage
  )
) +
  geom_hline(yintercept = 0, colour = "grey60") +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_wrap(vars(.data$Placement)) +
  scale_colour_manual(
    values = c(
      "Before AR correction" = "#CC6677",
      "After AR correction" = "#4477AA"
    )
  ) +
  labs(x = "Lag (minutes)", y = "Boundary-aware residual correlation") +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank()
  )
```

This step substantially reduced average temporal dependence. It did not
guarantee that every participant-day became independent; the later model
checks therefore also summarized residual autocorrelation separately within
participant-days.

## 5. From fitted models to reported quantities

### Site curves and pointwise intervals

`h02_modeling.R` and `h02_figure4_pointwise.R` evaluate the stored model on a
common 48-bin clock grid. The global time effect is the **site-average curve**,
an average across sites that gives every site equal weight. A site curve is
that global effect plus the site's sum-to-zero departure. Participant and
participant-day terms are set to their reference contributions when
constructing site curves. A **pointwise 95% confidence interval (95% CI)**
describes uncertainty at one displayed clock time. It comes from the fitted
linear-predictor matrix and coefficient covariance with smoothing parameters
held fixed; it is not a simultaneous band for the full day and does not
multiplicity-control every red segment. The plotted curves are back-transformed
to melEDI in lux for reader interpretation.

`build_h02_figure4_replication.R` applies the same builder to near-eye and
chest models. It reads stored fits and contributions; it does not refit the
model. Its principal outputs are:

- `artifacts/10_figures/H02/figure4_exact_layout_replication.png`;
- `artifacts/10_figures/H02/figure4_exact_layout_replication_chest.png`;
- matching PDF files; and
- `artifacts/11_source_data/H02/figure4_exact_layout_source*.rds`.

### Fitted-curve variation

For every scenario, `run_h02_analysis.R` extracts the fitted global, site,
participant, and participant-day contributions. It then calculates precisely
defined curve-dispersion summaries on the modelling scale. Within-site
participant and day summaries use site-average weighting, so every site
contributes equally.

The stored 95% intervals resample sites, participants within sites, and
participant-days within participants 2,000 times. The fitted contributions
remain fixed. Thus, the bootstrap describes cluster-sampling uncertainty
conditional on the fitted smooth structure; it is not a model-refitting
bootstrap. The stored resampling draws are under
`artifacts/08_diagnostics/H02/*__variation_bootstrap.rds`.

### Shapley allocation of fitted-model relevance

`run_h02_dominance_analysis.R` is intentionally narrower: it runs only for
the main all-available near-eye and complementary chest samples. The global
time effect is mandatory in all subset models. The script fits every one of
the eight possible combinations of site pattern, participant pattern, and
participant-day shift, then allocates their increments in row-weighted
in-sample $R^2$ across all orders of entry. This **Shapley allocation** is the
average contribution of a component across all orders in which the components
could be added, so shared fitted-model information is allocated rather than
counted repeatedly. The implementation is also recorded technically as exact
general dominance.

The 2,000-draw hierarchical bootstrap resamples clusters while holding the
eight sets of fitted predictions fixed. The calculation is descriptive
in-sample relevance, not causal importance and not cross-validated predictive
importance. Stored results and execution evidence are under:

- `artifacts/09_tables/H02/dominance_*.csv`;
- `artifacts/11_source_data/H02/*__dominance_predictions.rds`;
- `artifacts/08_diagnostics/H02/*__dominance_bootstrap.rds`; and
- `artifacts/12_manifests/H02/H02_dominance_manifest.csv`.

## 6. Model checks and influence assessment

`run_h02_temporal_diagnostics.R` reads the stored selected models and model
frames. It verifies coefficient rank and convergence, basis dimensions,
sum-to-zero constraints, midnight continuity, boundary-aware residual
behaviour, concurvity, and participant/site influence. **Concurvity** is
overlap among nonlinear model terms that makes their separate contributions
harder to distinguish. `gratia` is used for the formal concurvity wrapper;
`mgcv` supplies the fitted GAM machinery and core model checks (diagnostics).

`build_h02_reader_diagnostics.R` converts the diagnostic source objects into
the compact near-eye and chest model-check figures used in the H02 results
report. Neither script changes the selected formula. Key accepted stored
outputs are:

```{r}
#| label: tbl-h02-prep-diagnostic-map
#| tbl-cap: "Model-check outputs and the question each one addresses."

tibble::tribble(
  ~Output, ~Question_answered, ~Interpretive_limit,
  "residual_summary.csv", "How large and distributionally regular were AR-standardized residuals?", "In-sample and conditional on the fitted model.",
  "cluster_residual_acf_summary.csv", "Did dependence remain within participant-days?", "Summarizes observed fitted residuals, not future data.",
  "basis_dimension_checks.csv", "Was the chosen temporal basis visibly too small?", "A non-significant check is not proof of the true basis.",
  "formal_concurvity.csv", "How strongly could smooth terms reproduce one another?", "High overlap cautions against isolated coefficient interpretation.",
  "sz_constraint_identifiability.csv", "Did site departures sum numerically to zero?", "Checks implementation, not substantive effect size.",
  "conditional_deletion_influence.csv", "How much did fixed-fit summaries change after deleting a participant or site?", "Does not refit after deletion.",
  "model_diagnostics_*_source.rds", "What data generated the reader-facing model-check figures?", "Source object, not a new model."
) |>
  gt::gt() |>
  gt::cols_label(
    Question_answered = "Question answered",
    Interpretive_limit = "Interpretive limit"
  ) |>
  gt::cols_width(
    Output ~ gt::pct(27),
    Question_answered ~ gt::pct(40),
    Interpretive_limit ~ gt::pct(33)
  ) |>
  prep_gt()
```

The complete model-check assessment, including residual plots and exact
near-eye/chest values, is presented in the
[H02 results report](../../../notebooks/hypotheses/H02.qmd). The present page
focuses on how those outputs were generated rather than repeating
their full interpretation.

## 7. Sensitivity analyses and limitations

The following sensitivity analyses each change one interpretable part of the
analysis.
They use separate output identities and do not silently replace the primary
model.

```{r}
#| label: tbl-h02-prep-sensitivity-map
#| tbl-cap: "H02 sensitivity analyses and what was held constant."

tibble::tribble(
  ~Branch, ~Changed, ~Held_constant, ~Entry_point,
  "Gap-timing-unaware dataset", "Prepared 30-minute dataset", "Formula, transform, rho algorithm, AR boundaries, interval algorithm", "run_h02_analysis.R",
  "Placement-matched bins", "The same participants, participant-days, and 30-minute observations are required at both sensor positions", "Placement-specific models and selected formula", "build_h02_model_data.R → run_h02_analysis.R",
  "Fully cyclic deviations", "Site and participant time deviations also use cyclic marginals", "Near-eye data, response transform, AR algorithm, variation definitions", "run_h02_formula_sensitivity.R",
  "Non-cyclic global-time-basis check", "Only the global time basis changes from cyclic cubic to thin plate", "Site, participant, day terms; sample; transform; AR algorithm", "run_h02_global_time_diagnostic_sensitivity.R",
  "Complementary chest", "Sensor position changes from near eye to chest", "Selected formula and implementation", "run_h02_analysis.R"
) |>
  gt::gt() |>
  gt::cols_label(
    Held_constant = "Held constant",
    Entry_point = "Entry point"
  ) |>
  gt::cols_width(
    Branch ~ gt::pct(20),
    Changed ~ gt::pct(27),
    Held_constant ~ gt::pct(33),
    Entry_point ~ gt::pct(20)
  ) |>
  prep_gt()
```

The global-time-basis check was fit without a bootstrap and retained the
cyclic global time effect because residual model checks did not materially
improve and midnight continuity worsened. Its tables and manifest are under
`artifacts/08_diagnostics/H02/global_time_basis_*`,
`artifacts/09_tables/H02/global_time_basis_model_comparison.csv`, and
`artifacts/12_manifests/H02/H02_global_time_basis_diagnostic_manifest.csv`.

These analyses remain conditional descriptions of the fitted samples. The
near-eye position is not a retinal measurement, the chest position is not
ocular exposure, exact zeros produce a prominent lower-bound response mass,
some within-day residual correlation remains, and overlap between global and
site nonlinear terms limits interpretation of isolated coefficients. Shapley
allocations describe in-sample fitted-model relevance, not causal or
cross-validated importance.

## Technical provenance and reproduction

The remaining sections record exact code paths, commands, file identities, and
software versions. They are subordinate provenance for readers who want to
audit or reproduce the scientific objects described above. The confirmatory
multiplicity calculation used the Benjamini-Hochberg false-discovery-rate
procedure recorded in the H02 specification.

### Code modules

```{r}
#| label: tbl-h02-prep-module-map
#| tbl-cap: "H02 function modules sourced by the executable entry points."

tibble::tribble(
  ~Module, ~Responsibility, ~Why_separate,
  "h02_contract.R", "Pinned inputs, model/scenario registry, formulas, transform, seeds", "One reproducible source of settings used throughout the analysis.",
  "h02_data.R", "Temporal joins, support checks, scenario frames, AR starts", "Separates data admissibility and ordering from model fitting.",
  "h02_modeling.R", "BAM fitting, rho estimation, predictions, variation and core model checks", "Applies the same scientific implementation across placements and scenarios.",
  "h02_dominance.R", "Eight-model design, prediction assembly, exact Shapley allocation", "Keeps the descriptive relevance estimand explicit and reproducible.",
  "h02_figure4_pointwise.R", "Shared near-eye/chest Figure 1 contract", "Prevents placement-specific plotting implementations from diverging."
) |>
  gt::gt() |>
  gt::cols_label(
    Why_separate = "Why separate"
  ) |>
  gt::cols_width(
    Module ~ gt::pct(22),
    Responsibility ~ gt::pct(43),
    Why_separate ~ gt::pct(35)
  ) |>
  prep_gt()
```

### Executable entry points

```{r}
#| label: tbl-h02-prep-script-map
#| tbl-cap: "Executable H02 scripts in their analytical order."

tibble::tribble(
  ~Order, ~Entry_point, ~Contribution, ~Main_outputs, ~Run_when_page_renders,
  1, "audit_submitted_h02.R", "Verifies an archived source used only for historical comparison; it does not contribute to the current estimates.", "audit/hypotheses/H02/submitted_recovery.csv", "No",
  2, "build_h02_model_data.R", "Validates inputs and writes the eight H02-only model frames.", "artifacts/06_model_data/H02", "No",
  3, "run_h02_analysis.R", "Selects/fits the temporal structure, extracts contributions, and runs variation bootstraps.", "models, model-check outputs, tables, source data, analysis manifest", "No",
  4, "run_h02_formula_sensitivity.R", "Fits the fully cyclic model-form sensitivity.", "formula-sensitivity outputs and manifest", "No",
  5, "run_h02_temporal_diagnostics.R", "Adds temporal, identifiability, concurvity, and influence checks.", "model-check CSVs and temporal-diagnostics manifest", "No",
  6, "run_h02_dominance_analysis.R", "Fits eight subset structures and calculates conditional Shapley allocations.", "dominance tables, predictions, bootstraps, manifest", "No",
  7, "build_h02_figure4_replication.R", "Builds matched near-eye/chest result figures from stored fits.", "Figure 1 assets and source data", "No",
  8, "build_h02_reader_diagnostics.R", "Builds compact model-check figures from stored models.", "model-check figures and source data", "No",
  9, "build_h02_paired_placement_display.R", "Validates the placement-matched sample and prepares near-eye/chest temporal curves from stored predictions.", "paired placement source data and manifest", "No",
  10, "run_h02_global_time_diagnostic_sensitivity.R", "Compares cyclic and thin-plate global time bases without bootstrapping.", "basis-comparison model checks and manifest", "No",
  11, "notebooks/hypotheses/H02.qmd", "Assembles the H02 results report from recorded artifacts.", "rendered H02 results page", "No",
  12, "H02_analysis_preparation.qmd", "Explains and displays the analysis chain shown on this page.", "rendered preparation and provenance page", "Yes",
  13, "build_h02_figure_readability_qa.R", "Records the final-size visual review of every reader-facing H02 figure.", "H02_figure_readability_qa.csv", "No",
  14, "build_h02_preparation_report_manifest.R", "Records the preparation-page sources, rendered outputs, and dependencies.", "H02_preparation_report_manifest.csv", "No",
  15, "build_h02_worker_manifest.R", "Builds the final non-circular H02 file inventory.", "H02_worker_output_hashes.csv", "No"
) |>
  gt::gt() |>
  gt::cols_label(
    Entry_point = "Entry point",
    Main_outputs = "Main outputs",
    Run_when_page_renders = "Run when page renders"
  ) |>
  gt::cols_width(
    Order ~ gt::pct(6),
    Entry_point ~ gt::pct(21),
    Contribution ~ gt::pct(34),
    Main_outputs ~ gt::pct(29),
    Run_when_page_renders ~ gt::pct(10)
  ) |>
  prep_gt()
```

### Commands

The principal commands were run from the repository root in fresh R 4.6.1
sessions using the activated project library. Computationally intensive analyses
are shown for reproducibility; this page does not execute them.

```bash
Rscript --vanilla scripts/hypotheses/H02/audit_submitted_h02.R
Rscript --vanilla scripts/hypotheses/H02/build_h02_model_data.R
Rscript --vanilla scripts/hypotheses/H02/run_h02_analysis.R
Rscript --vanilla scripts/hypotheses/H02/run_h02_formula_sensitivity.R
Rscript --vanilla scripts/hypotheses/H02/run_h02_temporal_diagnostics.R
Rscript --vanilla scripts/hypotheses/H02/run_h02_dominance_analysis.R
Rscript --vanilla scripts/hypotheses/H02/build_h02_figure4_replication.R
Rscript --vanilla scripts/hypotheses/H02/build_h02_reader_diagnostics.R
Rscript --vanilla scripts/hypotheses/H02/build_h02_paired_placement_display.R
Rscript --vanilla scripts/hypotheses/H02/run_h02_global_time_diagnostic_sensitivity.R
quarto --profile nathealth render notebooks/hypotheses/H02.qmd --to html
quarto --profile nathealth render \
  audit/hypotheses/H02/H02_analysis_preparation.qmd --to html
Rscript --vanilla scripts/hypotheses/H02/build_h02_figure_readability_qa.R
Rscript --vanilla scripts/hypotheses/H02/build_h02_preparation_report_manifest.R
Rscript --vanilla scripts/environment/run_renv_status_safe.R <project-root> 60 true
Rscript --vanilla scripts/hypotheses/H02/build_h02_worker_manifest.R
```

### Output manifests and file identities

Each analysis script writes a small file inventory, or *manifest*, containing
output paths, SHA-256 identities, byte sizes, the producing script, R version,
and, for fitted models, the exact analytical sample. These records connect the
code to the files displayed in the reports and make later changes detectable.
The table below verifies the manifests used to navigate that chain.

```{r}
#| label: tbl-h02-prep-manifests
#| tbl-cap: "Output manifests linking H02 code to accepted stored outputs."

manifest_registry <- tibble::tribble(
  ~Stage, ~Path,
  "H02 model frames", "artifacts/12_manifests/H02/H02_model_data_manifest.csv",
  "Selected-model analysis", "artifacts/12_manifests/H02/H02_analysis_manifest.csv",
  "Temporal model checks", "artifacts/12_manifests/H02/H02_temporal_diagnostics_manifest.csv",
  "Model-form sensitivity", "artifacts/12_manifests/H02/H02_formula_sensitivity_manifest.csv",
  "Conditional dominance", "artifacts/12_manifests/H02/H02_dominance_manifest.csv",
  "Result-figure construction", "artifacts/12_manifests/H02/figure4_replication_manifest.csv",
  "Reader model checks", "artifacts/12_manifests/H02/H02_reader_diagnostics_manifest.csv",
  "Paired placement display", "artifacts/12_manifests/H02/H02_paired_placement_display_manifest.csv",
  "Final-size figure QA", "artifacts/12_manifests/H02/H02_figure_readability_qa.csv",
  "Global-time-basis check", "artifacts/12_manifests/H02/H02_global_time_basis_diagnostic_manifest.csv"
) |>
  rowwise() |>
  mutate(
    Rows = nrow(readr::read_csv(file.path(root, .data$Path), show_col_types = FALSE)),
    `Manifest SHA-256 prefix` = substr(
      artifact_sha256(file.path(root, .data$Path)),
      1L,
      16L
    )
  ) |>
  ungroup()

manifest_registry |>
  gt::gt() |>
  gt::cols_label(Stage = "Output group") |>
  gt::fmt_number(columns = Rows, decimals = 0, use_seps = TRUE) |>
  gt::cols_width(
    Stage ~ gt::pct(24),
    Path ~ gt::pct(55),
    Rows ~ gt::pct(7),
    `Manifest SHA-256 prefix` ~ gt::pct(14)
  ) |>
  gt::tab_source_note(
    "The final H02 file inventory is created after both reader-facing pages so that it can include their rendered outputs. It is kept outside these output manifests to avoid a circular checksum dependency."
  ) |>
  prep_gt()
```

### How to examine or reproduce H02

A reader can examine the analysis in four passes:

1. inspect `h02_contract.R` and `input_hashes.csv` to confirm the frozen input
   files and model specification;
2. compare the descriptive samples and support displays above with
   `sample_counts.csv` and the model-frame RDS files;
3. follow the producing script named in each output manifest, checking that its
   recorded inputs and outputs match the quantity being estimated; and
4. read the H02 results report alongside the result tables, model-check outputs, and
   source-data objects rather than attempting to infer results from a figure
   image alone.

Changes to source data or processing rules should be made before the
hypothesis-specific model frames are built and then propagated to every
affected analysis. Changing H02 alone would otherwise break comparability with
the other hypotheses. Alternative H02 model choices should remain clearly
labelled sensitivity analyses until their estimands and model checks are
established.

::: {.callout-tip title="Reader takeaway"}
The H02 result was not generated inside a single opaque notebook. It came from
a verified set of source files, H02-only model frames, a reusable temporal-model
implementation, separately recorded inference and model checks, stored
cluster-bootstrap evidence, and checksummed outputs. This page shows how those
pieces connect without changing the reported scientific result.
:::

### Environment

```{r}
#| label: tbl-h02-prep-environment
#| tbl-cap: "Environment used to render this page."

environment_table <- tibble::tibble(
  Component = c(
    "R",
    "mgcv",
    "gratia",
    "dplyr",
    "ggplot2",
    "gt",
    "Quarto",
    "renv.lock SHA-256 prefix"
  ),
  Version = c(
    as.character(getRversion()),
    as.character(packageVersion("mgcv")),
    as.character(packageVersion("gratia")),
    as.character(packageVersion("dplyr")),
    as.character(packageVersion("ggplot2")),
    as.character(packageVersion("gt")),
    Sys.getenv("QUARTO_VERSION", unset = "Recorded by render command"),
    substr(artifact_sha256(file.path(root, "renv.lock")), 1L, 16L)
  )
)

environment_table |>
  gt::gt() |>
  gt::cols_width(Component ~ gt::pct(65), Version ~ gt::pct(35)) |>
  prep_gt()
```
