---
title: "H10 Stage 1: audit and proposed analysis"
subtitle: "Nature Health analysis audit — author approval gate"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

library(dplyr)
library(tidyr)
library(readr)
library(tibble)
library(gt)
library(xml2)
library(rvest)
library(digest)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        dir.exists(file.path(candidate, "audit"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H10_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
root_path <- function(...) file.path(project_root, ...)

audit_gt <- function(data, title, subtitle = NULL) {
  data |>
    gt() |>
    tab_header(
      title = md(title),
      subtitle = if (is.null(subtitle)) NULL else md(subtitle)
    ) |>
    opt_row_striping() |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(12),
      data_row.padding = px(3),
      heading.align = "left"
    )
}

placement_label <- function(x) {
  ifelse(x == "glasses", "Near-eye", "Chest")
}

scenario_label <- function(x) {
  recode(
    x,
    primary = "Primary dataset",
    gap_timing_unaware = "Gap-timing-unaware dataset"
  )
}

relative_path <- function(path) {
  path <- normalizePath(path, winslash = "/", mustWork = FALSE)
  prefix <- paste0(project_root, "/")
  ifelse(startsWith(path, prefix), substring(path, nchar(prefix) + 1L), path)
}

sha256_file <- function(path) {
  if (!file.exists(path)) return(NA_character_)
  digest(path, algo = "sha256", file = TRUE)
}

metric_registry <- read_csv(
  root_path("config", "metric_display_registry.csv"),
  show_col_types = FALSE
) |>
  filter(
    analytical_role == "primary_outcome",
    analysis_unit %in% c("participant-day", "participant")
  ) |>
  mutate(
    analysis_unit = recode(
      analysis_unit,
      `participant-day` = "participant_day",
      participant = "participant"
    )
  )

site_registry <- read_csv(
  root_path("config", "site_display_registry.csv"),
  show_col_types = FALSE
)

source_crosswalk <- tribble(
  ~metric_id, ~source_field,
  "interdaily_stability", "interdaily_stability",
  "intradaily_variability", "intradaily_variability",
  "daily_geometric_mean_medi", "daily_geometric_mean_medi_lx",
  "m10_mean_medi", "m10_mean_medi_lx",
  "l10_mean_medi", "l10_mean_medi_lx",
  "duration_above_1000", "duration_above_1000_h",
  "duration_above_250_wake", "duration_above_250_wake_h",
  "duration_below_10_pre_sleep", "duration_below_10_pre_sleep_h",
  "duration_below_1_sleep_environment", "duration_below_1_sleep_environment_h",
  "longest_bout_above_250", "longest_bout_above_250_h",
  "m10_midpoint", "m10_midpoint_clock_minute",
  "l10_midpoint", "l10_midpoint_clock_minute",
  "mean_timing_above_250", "mean_timing_above_250_clock_minute",
  "first_timing_above_250", "first_timing_above_250_clock_minute",
  "last_timing_above_250", "last_timing_above_250_clock_minute",
  "dose_time_sensitive_corrected_medi", "dose_corrected_medi_lx_h",
  "mder_ratio_of_integrals", "mder"
)

response_contract <- tribble(
  ~metric_id, ~response_construction, ~engine, ~family, ~link, ~practical_effect,
  "interdaily_stability", "qlogis(value)", "lm", "Gaussian", "identity", "difference in IS; transformed-scale estimate also retained",
  "intradaily_variability", "value", "lm", "Gaussian", "identity", "difference in IV",
  "daily_geometric_mean_medi", "log10(value + 0.1)", "lmer", "Gaussian", "identity", "ratio of shifted melEDI values and percent change",
  "m10_mean_medi", "log10(value + 0.1)", "lmer", "Gaussian", "identity", "ratio of shifted melEDI values and percent change",
  "l10_mean_medi", "log10(value + 0.1)", "lmer", "Gaussian", "identity", "ratio of shifted melEDI values and percent change",
  "duration_above_1000", "value", "glmmTMB", "Tweedie", "log", "duration ratio, percent change, and modelled hours/minutes",
  "duration_above_250_wake", "value", "glmmTMB", "Tweedie", "log", "duration ratio, percent change, and modelled hours/minutes",
  "duration_below_10_pre_sleep", "value", "lmer", "Gaussian", "identity", "difference in hours/minutes",
  "duration_below_1_sleep_environment", "value", "glmmTMB", "Tweedie", "log", "duration ratio, percent change, and modelled hours/minutes",
  "longest_bout_above_250", "log10(value + 0.1)", "lmer", "Gaussian", "identity", "ratio of shifted period durations and percent change",
  "m10_midpoint", "value", "lmer", "Gaussian", "identity", "clock-minute difference",
  "l10_midpoint", "ifelse(value > 960, value - 1440, value)", "lmer", "Gaussian", "identity", "clock-minute difference on the approved continuous range",
  "mean_timing_above_250", "value", "lmer", "Gaussian", "identity", "clock-minute difference",
  "first_timing_above_250", "value", "lmer", "Gaussian", "identity", "clock-minute difference",
  "last_timing_above_250", "value", "lmer", "Gaussian", "identity", "clock-minute difference",
  "dose_time_sensitive_corrected_medi", "log10(value + 0.1)", "lmer", "Gaussian", "identity", "ratio of shifted melEDI dose values and percent change",
  "mder_ratio_of_integrals", "value", "lmer", "Gaussian", "identity", "difference in MDER"
)

metric_contract <- metric_registry |>
  left_join(source_crosswalk, by = "metric_id") |>
  left_join(response_contract, by = "metric_id") |>
  arrange(match(metric_id, metric_registry$metric_id))

stopifnot(
  nrow(metric_contract) == 17L,
  !anyNA(metric_contract$source_field),
  !anyNA(metric_contract$response_construction),
  identical(sort(metric_contract$metric_id), sort(source_crosswalk$metric_id))
)

demographics <- readRDS(
  root_path("artifacts", "06_model_data", "normalized_inputs", "demographics.rds")
) |>
  select(site, Id, age, sex, gender)

stopifnot(
  nrow(demographics) == nrow(distinct(demographics, site, Id)),
  all(!is.na(demographics$site)),
  all(!is.na(demographics$Id))
)

read_primary_long <- function(placement) {
  daily_path <- root_path(
    "artifacts", "06_model_data", "base",
    paste0("metrics_", placement, "_participant_day_enriched.rds")
  )
  participant_path <- root_path(
    "artifacts", "06_model_data", "base",
    paste0("metrics_", placement, "_participant_enriched.rds")
  )

  daily_fields <- metric_contract |>
    filter(analysis_unit == "participant_day") |>
    pull(source_field)
  participant_fields <- metric_contract |>
    filter(analysis_unit == "participant") |>
    pull(source_field)

  daily <- readRDS(daily_path) |>
    select(
      site, Id, local_date, age, sex, gender,
      valid_medi_real_minutes, all_of(daily_fields)
    ) |>
    pivot_longer(
      cols = all_of(daily_fields),
      names_to = "source_field",
      values_to = "metric_value"
    ) |>
    left_join(
      metric_contract |>
        select(metric_id, analysis_unit, source_field),
      by = "source_field"
    ) |>
    mutate(
      placement = placement,
      data_scenario = "primary",
      contributing_days = NA_real_,
      derivation_support_hours = valid_medi_real_minutes / 60
    )

  participant <- readRDS(participant_path) |>
    select(
      site, Id, age, sex, gender, days, valid_hours,
      all_of(participant_fields)
    ) |>
    pivot_longer(
      cols = all_of(participant_fields),
      names_to = "source_field",
      values_to = "metric_value"
    ) |>
    left_join(
      metric_contract |>
        select(metric_id, analysis_unit, source_field),
      by = "source_field"
    ) |>
    mutate(
      local_date = as.Date(NA),
      placement = placement,
      data_scenario = "primary",
      contributing_days = as.numeric(days),
      derivation_support_hours = as.numeric(valid_hours)
    ) |>
    select(-days, -valid_hours)

  bind_rows(daily, participant) |>
    select(
      data_scenario, placement, analysis_unit, metric_id,
      site, Id, local_date, metric_value, age, sex, gender,
      contributing_days, derivation_support_hours
    )
}

primary_long <- bind_rows(
  read_primary_long("glasses"),
  read_primary_long("chest")
)

gap_daily <- readRDS(
  root_path(
    "artifacts", "06_model_data", "scenarios", "manuscript_prepared_data",
    "participant_day_metrics.rds"
  )
) |>
  filter(metric_id %in% metric_contract$metric_id) |>
  inner_join(demographics, by = c("site", "Id")) |>
  transmute(
    data_scenario = "gap_timing_unaware",
    placement = position,
    analysis_unit = "participant_day",
    metric_id,
    site,
    Id,
    local_date = as.Date(local_date),
    metric_value = as.numeric(manuscript_prepared_value),
    age,
    sex,
    gender,
    contributing_days = NA_real_,
    derivation_support_hours = NA_real_
  )

gap_participant <- readRDS(
  root_path(
    "artifacts", "06_model_data", "scenarios", "manuscript_prepared_data",
    "participant_metrics.rds"
  )
) |>
  filter(metric_id %in% metric_contract$metric_id) |>
  inner_join(demographics, by = c("site", "Id")) |>
  transmute(
    data_scenario = "gap_timing_unaware",
    placement = position,
    analysis_unit = "participant",
    metric_id,
    site,
    Id,
    local_date = as.Date(NA),
    metric_value = as.numeric(manuscript_prepared_value),
    age,
    sex,
    gender,
    contributing_days = NA_real_,
    derivation_support_hours = NA_real_
  )

all_long <- bind_rows(primary_long, gap_daily, gap_participant) |>
  mutate(
    biological_sex = factor(as.character(sex), levels = c("Male", "Female")),
    age_decade = as.numeric(age) / 10,
    response = case_when(
      metric_id == "interdaily_stability" ~ qlogis(metric_value),
      metric_id %in% c(
        "daily_geometric_mean_medi", "m10_mean_medi", "l10_mean_medi",
        "longest_bout_above_250", "dose_time_sensitive_corrected_medi"
      ) ~ log10(metric_value + 0.1),
      metric_id == "l10_midpoint" ~ if_else(
        metric_value > 960,
        metric_value - 1440,
        metric_value
      ),
      TRUE ~ metric_value
    ),
    candidate = is.finite(response) &
      is.finite(age_decade) &
      !is.na(site) &
      !is.na(Id) &
      !is.na(biological_sex)
  )

candidate_long <- all_long |>
  filter(candidate)

candidate_samples_direct <- candidate_long |>
  group_by(data_scenario, placement, analysis_unit, metric_id) |>
  summarise(
    observations = n(),
    participants = n_distinct(site, Id),
    sites = n_distinct(site),
    missing_age_rows = sum(is.na(age_decade)),
    missing_biological_sex_rows = sum(is.na(biological_sex)),
    .groups = "drop"
  )

design_audit <- candidate_long |>
  group_by(data_scenario, placement, analysis_unit, metric_id) |>
  group_modify(function(.x, .y) {
    participant_cells <- .x |>
      distinct(site, Id, biological_sex, age_decade)
    sex_counts <- participant_cells |>
      count(site, biological_sex, name = "n")
    age_support <- participant_cells |>
      group_by(site) |>
      summarise(
        age_range = max(age_decade * 10) - min(age_decade * 10),
        unique_ages = n_distinct(age_decade),
        .groups = "drop"
      )
    mm_age <- model.matrix(~ site * age_decade, data = .x)
    mm_sex <- model.matrix(~ site * biological_sex, data = .x)
    tibble(
      age_full_rank = qr(mm_age)$rank == ncol(mm_age),
      sex_full_rank = qr(mm_sex)$rank == ncol(mm_sex),
      min_site_age_range_years = min(age_support$age_range),
      min_unique_ages_per_site = min(age_support$unique_ages),
      min_site_sex_cell_participants = min(sex_counts$n),
      max_site_participant_share = max(table(participant_cells$site)) /
        nrow(participant_cells)
    )
  }) |>
  ungroup()

demographic_membership <- candidate_long |>
  distinct(
    data_scenario, placement, site, Id, age,
    biological_sex = biological_sex,
    gender = gender
  )

site_demographics <- demographic_membership |>
  mutate(
    biological_sex = as.character(biological_sex),
    gender = as.character(gender)
  ) |>
  group_by(data_scenario, placement, site) |>
  summarise(
    participants = n(),
    age_min = min(age),
    age_median = median(age),
    age_max = max(age),
    Female = sum(biological_sex == "Female"),
    Male = sum(biological_sex == "Male"),
    Woman = sum(gender == "Woman"),
    Man = sum(gender == "Man"),
    missing_age = sum(is.na(age)),
    missing_sex = sum(is.na(biological_sex)),
    missing_gender = sum(is.na(gender)),
    .groups = "drop"
  ) |>
  left_join(site_registry, by = "site") |>
  arrange(data_scenario, placement, display_order)

sex_gender_crosswalk <- demographic_membership |>
  mutate(
    biological_sex = as.character(biological_sex),
    gender = as.character(gender)
  ) |>
  count(data_scenario, placement, biological_sex, gender, name = "participants") |>
  arrange(data_scenario, placement, biological_sex, gender)

flow_paths <- c(
  primary = root_path("artifacts", "06_model_data", "H01", "sample_flow.csv"),
  gap_timing_unaware = root_path(
    "artifacts", "06_model_data", "H01", "scenarios",
    "manuscript_prepared_data", "sample_flow.csv"
  )
)

flow <- bind_rows(lapply(names(flow_paths), function(scenario_name) {
  read_csv(flow_paths[[scenario_name]], show_col_types = FALSE) |>
    mutate(data_scenario = scenario_name)
}))

flow_candidate <- flow |>
  filter(
    stage == "metric_estimable",
    scope == "overall",
    site == "ALL",
    scenario == "all_available"
  ) |>
  transmute(
    data_scenario,
    placement,
    metric_id,
    analysis_unit,
    observations = model_observations,
    participants,
    participant_days,
    contributing_participant_days,
    derivation_support_hours = metric_support_valid_hours,
    sites,
    status
  )

candidate_reconciliation <- candidate_samples_direct |>
  left_join(
    flow_candidate |>
      select(
        data_scenario, placement, metric_id,
        flow_observations = observations,
        flow_participants = participants,
        flow_sites = sites
      ),
    by = c("data_scenario", "placement", "metric_id")
  ) |>
  mutate(
    observations_match = observations == flow_observations,
    participants_match = participants == flow_participants,
    sites_match = sites == flow_sites
  )

paired_candidate <- flow |>
  filter(
    stage == "metric_estimable",
    scope == "overall",
    site == "ALL",
    scenario == "paired_common_sample"
  ) |>
  transmute(
    data_scenario,
    placement,
    metric_id,
    analysis_unit,
    observations = model_observations,
    participants,
    participant_days,
    contributing_participant_days,
    derivation_support_hours = metric_support_valid_hours,
    sites,
    status
  )

stopifnot(
  all(candidate_reconciliation$observations_match),
  all(candidate_reconciliation$participants_match),
  all(candidate_reconciliation$sites_match),
  all(candidate_samples_direct$missing_age_rows == 0L),
  all(candidate_samples_direct$missing_biological_sex_rows == 0L)
)

sample_cell <- function(
  analysis_unit, observations, participants, participant_days,
  contributing_participant_days, derivation_support_hours, sites, status
) {
  if (!identical(status, "AVAILABLE")) return(status)
  day_text <- if (analysis_unit == "participant") {
    paste0(contributing_participant_days, " contributing days")
  } else {
    paste0(participant_days, " participant-days")
  }
  hour_text <- if (is.na(derivation_support_hours)) {
    "support hours unavailable"
  } else {
    paste0(formatC(derivation_support_hours, format = "f", digits = 2, big.mark = ","), " support h")
  }
  paste0(
    participants, " participants; ", day_text, "; ", observations,
    " model rows; ", hour_text, "; ", sites, " sites"
  )
}

sample_display <- function(data) {
  data |>
    left_join(
      metric_contract |>
        select(metric_id, manuscript_name, abbreviation, analysis_unit),
      by = c("metric_id", "analysis_unit")
    ) |>
    rowwise() |>
    mutate(
      sample = sample_cell(
        analysis_unit, observations, participants, participant_days,
        contributing_participant_days, derivation_support_hours, sites, status
      )
    ) |>
    ungroup() |>
    select(data_scenario, placement, metric_id, manuscript_name, abbreviation, sample) |>
    pivot_wider(names_from = placement, values_from = sample) |>
    arrange(match(metric_id, metric_contract$metric_id))
}

all_available_display <- sample_display(flow_candidate)
paired_display <- sample_display(paired_candidate)

read_v0_tables <- function(path, placement) {
  page <- read_html(path)
  section <- xml_find_first(page, '//*[@id="hypothesis-10"]')
  tables <- html_elements(section, "table")
  stopifnot(length(tables) == 3L)

  overall_raw <- html_table(tables[[2L]], fill = TRUE)
  overall_header <- as.character(overall_raw[3L, ])
  overall <- overall_raw[seq.int(4L, nrow(overall_raw) - 1L), , drop = FALSE]
  names(overall) <- make.unique(overall_header)
  tibble(
    placement = placement,
    metric_type = overall[[1L]],
    metric_v0 = overall[[2L]],
    sex_adjusted_p_display = overall[[3L]],
    age_adjusted_p_display = overall[[4L]],
    age_site_adjusted_p_display = overall[[5L]]
  ) |>
    filter(nzchar(metric_type), nzchar(metric_v0))
}

v0_overall_results <- bind_rows(
  read_v0_tables(root_path("docs", "RQ3.html"), "glasses"),
  read_v0_tables(root_path("docs", "RQ3_chest.html"), "chest")
)
stopifnot(nrow(v0_overall_results) == 34L)

input_inventory <- tribble(
  ~input_id, ~role, ~path,
  "metric_registry", "Metric names, units, analysis units, and outcome roles", "config/metric_display_registry.csv",
  "site_registry", "Submitted-manuscript site names, order, and colours", "config/site_display_registry.csv",
  "normalized_demographics", "Unique site-participant age, biological sex, and gender fields", "artifacts/06_model_data/normalized_inputs/demographics.rds",
  "base_contract", "Hypothesis-neutral base-data contract", "artifacts/06_model_data/base/base_model_contract.csv",
  "base_input_provenance", "Provenance for base model data", "artifacts/06_model_data/base/input_provenance.csv",
  "primary_near_eye_participant_day", "Primary near-eye participant-day metrics with demographics", "artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds",
  "primary_chest_participant_day", "Complementary chest participant-day metrics with demographics", "artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds",
  "primary_near_eye_participant", "Primary near-eye participant-level metrics with demographics", "artifacts/06_model_data/base/metrics_glasses_participant_enriched.rds",
  "primary_chest_participant", "Complementary chest participant-level metrics with demographics", "artifacts/06_model_data/base/metrics_chest_participant_enriched.rds",
  "primary_metric_contract", "Approved 17-metric prepared-data contract, not H01 fitted results", "artifacts/06_model_data/H01/metric_contract.csv",
  "primary_sample_flow", "Metric-specific all-available and paired/common sample counts", "artifacts/06_model_data/H01/sample_flow.csv",
  "gap_participant_day", "Gap-timing-unaware participant-day metric values", "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
  "gap_participant", "Gap-timing-unaware participant-level metric values", "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_metrics.rds",
  "gap_source_manifest", "Source identity for the gap-timing-unaware preparation", "artifacts/06_model_data/scenarios/manuscript_prepared_data/source_manifest.csv",
  "gap_correction_manifest", "Bounded corrections in the gap-timing-unaware preparation", "artifacts/06_model_data/scenarios/manuscript_prepared_data/correction_manifest.csv",
  "gap_metric_crosswalk", "Legacy-to-canonical metric mapping", "artifacts/06_model_data/scenarios/manuscript_prepared_data/metric_crosswalk.csv",
  "gap_normalized_input_references", "Normalized metadata references used by the gap preparation", "artifacts/06_model_data/scenarios/manuscript_prepared_data/normalized_input_references.csv",
  "gap_variable_dictionary", "Gap-preparation variable definitions", "artifacts/06_model_data/scenarios/manuscript_prepared_data/variable_dictionary.csv",
  "gap_metric_contract", "Approved 17-metric gap-scenario contract", "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/metric_contract.csv",
  "gap_contract_equivalence", "Primary-versus-gap contract reconciliation", "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/contract_equivalence.csv",
  "gap_input_provenance", "Gap-scenario H01 input provenance without fitted H01 results", "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/input_provenance.csv",
  "gap_sample_flow", "Gap-scenario all-available and paired/common sample counts", "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/sample_flow.csv",
  "gap_timing_conversion_audit", "Clock-variable conversion audit", "artifacts/06_model_data/H01/scenarios/manuscript_prepared_data/timing_conversion_audit.csv",
  "metric_artifact_manifest", "Current metric-artifact identity", "artifacts/12_manifests/metric_artifacts.csv",
  "base_model_manifest", "Current base-model-data identity", "artifacts/12_manifests/base_model_data_artifacts.csv"
) |>
  mutate(
    absolute_path = root_path(path),
    exists = file.exists(absolute_path),
    bytes = if_else(exists, as.numeric(file.info(absolute_path)$size), NA_real_),
    sha256 = vapply(absolute_path, sha256_file, character(1L))
  )

stopifnot(all(input_inventory$exists))

v0_output_paths <- c(
  shared_table_target = root_path("tables", "H10.png"),
  manuscript_table_asset = root_path("assets", "H10.png"),
  rendered_table_asset = root_path("docs", "assets", "H10.png"),
  near_eye_figure_pdf = root_path("figures", "Fig8.pdf"),
  near_eye_figure_png = root_path("figures", "Fig8.png"),
  chest_figure_pdf = root_path("figures", "chest", "Fig8.pdf"),
  chest_figure_png = root_path("figures", "chest", "Fig8.png")
)

v0_output_identity <- tibble(
  output_id = names(v0_output_paths),
  path = vapply(v0_output_paths, relative_path, character(1L)),
  exists = file.exists(v0_output_paths),
  sha256 = vapply(v0_output_paths, sha256_file, character(1L))
)

manifest_reconciliation <- tribble(
  ~check_id, ~observed, ~status, ~consequence,
  "H10-specific prepared artifact", as.character(file.exists(root_path("artifacts/06_model_data/H10.rds"))), "MISSING", "Expected by the placeholder H10 notebook; must be built only after Stage 1 approval.",
  "Current metric manifest SHA-256", sha256_file(root_path("artifacts/12_manifests/metric_artifacts.csv")), "IDENTIFIED", "Current scientific identity is pinned.",
  "Current base-model manifest SHA-256", sha256_file(root_path("artifacts/12_manifests/base_model_data_artifacts.csv")), "IDENTIFIED", "Current downstream identity is pinned.",
  "Preparation 04 independent reconstruction for current manifest", "Open under PREP-003 / FIND-044", "OPEN", "Identity checks do not substitute for current-manifest independent scientific reconstruction.",
  "Preparation 06 gate versus current base objects", "Gate states 811/897 participant-days; current objects contain 816/902", "STALE", "Coordinator must reconcile the shared gate before H10 fitting."
)

registered_formula <- as.formula(
  "Metric ~ Predictor * Site + (1 | Site:Participant)"
)

v0_rhs <- tribble(
  ~model_id, ~rhs,
  "H10_1a", "~ site + age + (1 | Id)",
  "H10_1s", "~ site + sex + (1 | Id)",
  "H10_0", "~ site + (1 | Id)",
  "H10_a", "~ age + (1 | Id)",
  "H10_00", "~ 1 + (1 | Id)",
  "H10_ia", "~ site * age + (1 | Id)",
  "H10_is", "~ site * sex + (1 | Id)"
)

v0_response_map <- tribble(
  ~metric_group, ~analysis_unit, ~response_expression, ~engine, ~family,
  "Interdaily stability", "participant", "qlogis(metric)", "lm", "Gaussian identity",
  "Intradaily variability", "participant", "metric", "lm", "Gaussian identity",
  "Mean melEDI, M10mean, L10mean, melEDI dose, longest period", "participant_day", "log_zero_inflated(metric)", "lmer", "Gaussian identity",
  "M10/L10 midpoints, mean/first/last timing, MDER", "participant_day", "metric", "lmer", "Gaussian identity",
  "TAT1000, waking TAT250, pre-sleep TBT10, sleep-environment TBT1", "participant_day", "metric", "glmmTMB", "Tweedie log"
)

v0_formula_manifest <- crossing(v0_response_map, v0_rhs) |>
  mutate(
    rhs_supplied = if_else(
      analysis_unit == "participant",
      sub(" \\+ \\(1 \\| Id\\)$", "", rhs),
      rhs
    ),
    formula = paste(response_expression, rhs_supplied),
    evaluated_formula = lapply(formula, as.formula)
  )

participant_day_formulas <- list(
  M0 = as.formula("response ~ site + (1 | site:Id)"),
  M_age = as.formula("response ~ site + age_decade + (1 | site:Id)"),
  M_age_site = as.formula("response ~ site * age_decade + (1 | site:Id)"),
  M_sex = as.formula("response ~ site + biological_sex + (1 | site:Id)"),
  M_sex_site = as.formula("response ~ site * biological_sex + (1 | site:Id)")
)

participant_formulas <- list(
  M0 = as.formula("response ~ site"),
  M_age = as.formula("response ~ site + age_decade"),
  M_age_site = as.formula("response ~ site * age_decade"),
  M_sex = as.formula("response ~ site + biological_sex"),
  M_sex_site = as.formula("response ~ site * biological_sex")
)

comparison_manifest <- tribble(
  ~comparison_id, ~reduced_model, ~full_model, ~estimand, ~family_id_near_eye, ~family_id_chest,
  "AGE-MAIN", "M0", "M_age", "Common site-adjusted age association per 10 years", "H10-F1-age-main", "H10-C1-age-main",
  "SEX-MAIN", "M0", "M_sex", "Common site-adjusted Female-minus-Male association", "H10-F2-sex-main", "H10-C2-sex-main",
  "AGE-SITE", "M_age", "M_age_site", "Omnibus age-by-site heterogeneity", "H10-F3-age-site-interaction", "H10-C3-age-site-interaction",
  "SEX-SITE", "M_sex", "M_sex_site", "Omnibus biological-sex-by-site heterogeneity", "H10-F4-sex-site-interaction", "H10-C4-sex-site-interaction"
)

findings <- tribble(
  ~finding_id, ~severity, ~evidence_status, ~domain, ~finding, ~consequence, ~required_disposition,
  "H10-AUD-001", "critical", "confirmed", "Multiplicity", "V0 calls p.adjust() on one p-value at a time with n = 34; this is not vector-wide BH FDR control.", "All V0 adjusted p-values and significance labels are reopened.", "Assemble and adjust complete declared 17-member vectors once.",
  "H10-AUD-002", "critical", "confirmed", "Registered inference", "V0 fits the sex-by-site model but never compares, reports, or classifies it.", "Part of the registered contract is missing.", "Test and report the complete sex-by-site family or reason-code non-estimability.",
  "H10-AUD-003", "major", "confirmed", "Contrasts", "V0 says site is sum-coded, but the H10 fitting call supplies no site contrast and default treatment coding applies.", "The claimed average-effect interpretation is unsupported by the implementation.", "Set contrasts explicitly and store them with every fit.",
  "H10-AUD-004", "major", "confirmed", "Effect provenance", "V0 reports the age slope from the site-adjusted model but marginal R² from an age-only model.", "The estimate and fit statistic do not describe the same model.", "Compute every fit summary from the explicitly named reported model.",
  "H10-AUD-005", "major", "confirmed", "Outputs", "Near-eye and chest renders write tables/H10.png; the manuscript asset has a different hash from the shared producer target.", "Placement identity and manuscript provenance are unstable.", "Use deterministic H10/placement/scenario paths and paired source-data CSVs.",
  "H10-AUD-006", "major", "confirmed", "Tables", "The V0 chest table merges engine-specific interval columns inconsistently and leaves some selected estimates blank.", "Blank cells are an output defect, not zero effects.", "Use one typed estimate schema across engines and verify no lost estimate/interval.",
  "H10-AUD-007", "critical", "confirmed", "Diagnostics", "V0 performs no H10 model-specific convergence, distributional, residual, temporal, or influence assessment.", "The inferential adequacy of V0 is unknown.", "Run, interpret, and classify every required diagnostic by placement and metric.",
  "H10-AUD-008", "major", "confirmed", "Reporting", "V0 reports estimates only for selected age rows and gives no biological-sex contrasts or 95% intervals.", "Magnitude, precision, and null-result compatibility cannot be evaluated.", "Report every planned estimate and 95% interval, including unsupported tests.",
  "H10-AUD-009", "major", "confirmed", "Figures", "The V0 age figure is an unadjusted pooled smoother, labels duration as local time, and repeats participant ages across metric frames in its site panel.", "The display does not represent the fitted estimand and contains factual layout errors.", "Recreate V0, then replace it with model-based, correctly labelled, source-backed displays.",
  "H10-AUD-010", "major", "confirmed", "Construct validity", "V0 does not explain that biological sex and gender are separate normalized fields; one participant per placement has Male sex and Woman gender.", "Renaming or substituting gender would change contrast membership.", "Use only biological sex for registered inference and describe gender separately.",
  "H10-AUD-011", "major", "confirmed", "Design support", "All nominal interaction matrices are full rank, but the smallest near-eye site-by-sex cell has two participants and the narrowest within-site age range is seven years.", "Interaction estimates can be imprecise or site-driven despite nominal estimability.", "Retain sparse-cell, leverage, participant, and leave-one-site-out diagnostics.",
  "H10-AUD-012", "critical", "open provenance gap", "Preparation", "PREP-003/FIND-044 records no complete independent reconstruction against the current metric manifest.", "Current identities are known, but current scientific values/support classifications are not independently reconstructed in the stored audit record.", "Coordinator must verify or explicitly accept the qualification before Stage 2 fitting.",
  "H10-AUD-013", "major", "confirmed", "Preparation", "The shared Preparation 06 gate reports 811/897 participant-days, while current base objects contain 816/902.", "The shared gate text is stale relative to H10 inputs.", "Coordinator must reconcile the gate and pin the intended current base identity.",
  "H10-AUD-014", "moderate", "confirmed", "Paired placement", "Participant-level IS and IV have no approved paired/common values because those metrics require recomputation on identical paired day sets.", "A matched placement display is unavailable for IS/IV without expanding preparation work.", "Report paired/common evidence only where estimands permit; do not approximate IS/IV."
)

diagnostic_plan <- tribble(
  ~domain, ~planned_evidence, ~acceptance_rule,
  "Frame and joins", "Key uniqueness; many-to-one demographic join; exact row hashes; missingness; transform-domain checks; participants/days/rows/hours/sites", "Any unexplained row change, key duplication, missing predictor, invalid transformed value, or comparison-frame mismatch is not acceptable.",
  "Design and hierarchy", "Factor levels/references; design rank; aliased coefficients; site-by-sex cells; within-site age range; participant-day repetition", "Rank deficiency or an unidentifiable registered contrast is not acceptable; it remains visible as non-estimable rather than being simplified silently.",
  "Convergence and numerical stability", "Optimizer status, warnings, gradients, Hessian/eigenvalues, finite coefficients, Tweedie dispersion and power, random-intercept variance, singularity", "Unresolved convergence, non-positive Hessian, non-finite coefficient, or singular fit that changes inference is not acceptable.",
  "Distribution and residuals", "Response/zero mass; residual-versus-fitted; Q–Q/tails; scale-location; residuals by site, sex, age, and metric-specific support", "A material family/link violation or unexplained structure that invalidates model-based intervals is not acceptable; lesser departures are reported as limitations.",
  "Temporal dependence", "Date-ordered within-participant residual ACF using consecutive days only; lag distribution; missing-date boundaries", "Material remaining day-to-day dependence must be quantified. If inference changes under an approved participant-cluster sensitivity, the result is unstable and cannot be called acceptable without qualification.",
  "Influence and heterogeneity", "Participant influence; DFBETAs or deletion summaries; leave-one-site-out estimates; equal-site versus observed-sample marginalization", "A conclusion created by one participant/site or reversed by plausible marginalization is not robust; classify as acceptable with limitation or not acceptable according to magnitude.",
  "Inference and reporting", "ML comparison integrity; REML final Gaussian estimates; raw/BH p-values; 95% intervals; practical-scale conversions; deterministic outputs", "A comparison with unequal frames/methods, incorrect family membership, or unreconciled table/figure source is not acceptable for inference or publication."
)

sensitivity_plan <- tribble(
  ~scenario_id, ~role, ~sample_or_data_change, ~model_change, ~interpretive_rule,
  "S0-primary-all-available", "Primary near-eye; complementary chest", "All metric-eligible rows in the primary dataset", "None", "Controls the H10 conclusion; placements are not pooled.",
  "S1-gap-timing-unaware", "Required data sensitivity", "Use the gap-timing-unaware dataset", "Identical approved H10 implementation", "May qualify stability but cannot replace the primary conclusion.",
  "S2-paired-common-placement", "Placement comparability", "Same participants and participant-days at both placements, metric by metric", "Identical approved H10 implementation", "Show side-by-side estimates only where the estimand is reproducible; no equivalence claim and no pooling.",
  "S3-common-primary-gap", "Data-versus-model decomposition", "Exact keys common to primary and gap-timing-unaware data", "Identical approved H10 implementation", "Separates changed preparation from changed sample composition.",
  "S4-leave-one-site-out", "Influence diagnostic", "Repeat after omitting each site", "Same response and fixed/random structure", "No p-value screening; evaluate estimate, interval, and family-decision stability.",
  "S5-preregistered-exclusion", "Deviation sensitivity", "Apply the registered age/employment exclusions documented in the deviation record", "Same approved H10 implementation", "Shows whether approved inclusion deviations materially affect age or sex associations.",
  "S6-longest-period-exact", "Metric-validity sensitivity", "Use only exactly identifiable longest periods", "Same approved H10 implementation", "Required for the observed-lower-bound longest-period metric.",
  "S7-L10-noon-conversion", "Clock-construction sensitivity", "Use the approved noon conversion on identical L10 rows", "Same approved H10 implementation", "Assesses dependence on the L10 continuous-range convention.",
  "S8-MDER-support", "Support sensitivity", "Use approved alternative MDER support thresholds and relevant waking-only check", "Same approved H10 implementation", "Assesses tail/support dependence without redefining the primary metric.",
  "S9-response-family-alternative", "Deferred diagnostic gate", "Same rows", "Common approved alternative family/link for both data scenarios", "Not run unless a primary family fails diagnostics and the author approves a replacement."
)

author_decisions <- tribble(
  ~decision_id, ~decision_required, ~recommended_disposition, ~why_it_matters,
  "H10-G1", "Multiplicity partition", "Approve four separate complete 17-member BH families for age main, biological-sex main, age-by-site, and biological-sex-by-site tests; use separate analogous chest families.", "The registration does not uniquely decide whether age and sex share one family; V0's scalar n = 34 operation is invalid either way.",
  "H10-G2", "Primary estimand", "Approve the common site-adjusted association from M_age or M_sex as primary, with interactions treated as registered heterogeneity analyses. Qualify the common summary if heterogeneity is supported or influence is material.", "A joint predictor-plus-interaction test answers a different question and makes a single practical slope/contrast ambiguous.",
  "H10-G3", "Response models", "Approve the H01 response-family/transform package shown here as H10's prespecified starting package, without inheriting any H01 fit or result.", "Likelihood, effect scale, intervals, and diagnostics depend on this choice.",
  "H10-G4", "Repeated observations and random structure", "Use (1 | site:Id) for participant-day metrics and ordinary site-adjusted lm models for participant-level IS/IV; do not add an unsupported predictor random slope. Treat temporal dependence as a diagnostic gate.", "The registered hierarchy must represent repeated days, while one row per participant cannot support a participant random intercept.",
  "H10-G5", "Construct and contrasts", "Use age_decade = age/10; use only biological sex with levels Male then Female; report Female-minus-Male; set site contrasts explicitly; do not infer gender.", "Biological sex and gender are distinct source constructs and one observed record maps differently.",
  "H10-G6", "Placement roles", "Near-eye is primary and chest complementary; never pool placements; use paired/common evidence only where the same estimand can be constructed.", "Uneven placement coverage and reviewer requests require placement-specific samples and matched evidence.",
  "H10-G7", "Gap sensitivity", "Run the identical approved implementation on the gap-timing-unaware dataset and on exact common primary/gap keys.", "This distinguishes preparation changes from model changes and sample-composition changes.",
  "H10-G8", "Metric-specific sensitivities", "Approve longest-period exact-only, L10 noon-conversion, MDER-support, and documented preregistered-exclusion sensitivities; keep response-family alternatives behind a new gate.", "These are the pre-existing admissibility/deviation questions most likely to affect H10 interpretation.",
  "H10-G9", "Diagnostic disposition", "Require an explicit acceptable / acceptable with specified limitations / not acceptable classification for every metric-placement analysis using the table above.", "A nominally converged model is not sufficient evidence of inferential adequacy.",
  "H10-G10", "Uncertainty and practical scale", "Use 95% model-based intervals: t intervals for lm, fixed-effect covariance intervals for mixed models, and emmeans-based trends/contrasts; back-transform log/log-link effects. Use no bootstrap unless diagnostics require it.", "Every estimate needs precision on an interpretable scale without triggering unapproved heavy resampling.",
  "H10-G11", "Follow-up interaction summaries", "After an omnibus interaction, show all site-specific slopes/contrasts with 95% intervals as heterogeneity descriptions; do not create site-specific significance claims without a separately approved family.", "Selective site-wise testing would add undeclared multiplicity and overstate sparse cells.",
  "H10-G12", "Paired IS/IV", "Mark paired/common IS and IV unavailable unless an approved upstream process recomputes them from identical paired days; do not substitute the all-available participant values.", "Participant-level dynamics metrics change when their contributing day set changes.",
  "H10-G13", "Preparation provenance prerequisite", "Do not fit H10 until the coordinator reconciles the stale Preparation 06 gate and resolves or explicitly accepts PREP-003/FIND-044 for the current manifest.", "The present identity/provenance mismatch is shared and outside H10 ownership."
)

claim_source_matrix <- tribble(
  ~claim_or_requirement, ~controlling_source, ~stage1_disposition,
  "H10 hypothesis and registered interaction model", "Signed preregistration as reconstructed in audit/evidence/preregistration_contract.md", "Retain as the starting contract.",
  "Age and biological sex require declared families", "Preregistration contract and H09-H11 migration map", "Resolve H10-G1 before fitting.",
  "Near-eye primary; chest complementary", "audit/decisions/placement_decision.md", "Apply without pooling or equivalence language.",
  "17 metric names, units, and roles", "config/metric_display_registry.csv and approved H01 metric contract", "Use all 17 independent of observed results.",
  "Response family and transform package", "scripts/hypotheses/H01/h01_contract.R and artifacts/09_tables/H01/response_family_candidates/H01_response_family_candidate_selection.csv", "Approved for H10 under amended H10-G3; no H01 fitted result is inherited.",
  "Gap-timing-unaware terminology and role", "audit/decisions/gap_timing_unaware_dataset_terminology.md and manuscript_prepared_data_sensitivity.md", "Use as required sensitivity, not as the primary dataset.",
  "Exact samples and paired/common placement", "H01 primary and gap sample_flow.csv files plus direct H10 reconstruction", "Counts reconcile; IS/IV paired values remain unavailable.",
  "Model-specific samples, uncertainty, multiplicity, influence", "NM-R1-05, NM-R2-01, and NM-R2-03 in reviewer_feedback_disposition.csv", "Adopt throughout Stage 2.",
  "Submitted H10 numerical claims", "RQ3.qmd, RQ3_chest.qmd, docs renders, index.qmd, and frozen assets", "All claims reopened pending repaired analysis.",
  "Current metric provenance qualification", "audit/decisions/preparation04_version_specific_verification.md", "Block fitting until coordinator response under H10-G13."
)
```

## Stage 1 verdict

**The H10 V0 audit and proposed specification are complete, but H10 is not
ready for Stage 2.** No new H10 inferential model, prediction, simulation, or
bootstrap has been run. The evaluated R code in this report reads prepared
artifacts, reconstructs candidate samples and design matrices, parses frozen
V0 output, and constructs formula objects only.

The proposed repair retains all 17 registered metric outcomes, separates age
from the measured biological-sex variable, treats near-eye measurements as
primary and chest measurements as complementary, and restores both registered
predictor-by-site heterogeneity tests. The recommended primary estimands are
the common site-adjusted age association per 10 years and the common
site-adjusted Female-minus-Male association. Site interactions answer a
separate heterogeneity question.

Two gates remain:

1. the author must approve every item in the Stage 1 decision checklist; and
2. the coordinator must reconcile the shared preparation provenance issue
documented here and in `audit/handoffs/H10_shared_change_request.md`.

## Controlling evidence and preparation provenance

The audit used R 4.6.1 and the project library. It did not use another
language to calculate or adjudicate a scientific value. The direct
reconstruction agrees with the two approved sample-flow artifacts for all 68
placement × scenario × metric candidate frames on model rows, participants,
and sites.

```{r}
#| label: environment
#| echo: false

tibble(
  Component = c(
    "R", "dplyr", "tidyr", "readr", "gt", "xml2", "rvest", "digest"
  ),
  Version = c(
    R.version.string,
    as.character(packageVersion("dplyr")),
    as.character(packageVersion("tidyr")),
    as.character(packageVersion("readr")),
    as.character(packageVersion("gt")),
    as.character(packageVersion("xml2")),
    as.character(packageVersion("rvest")),
    as.character(packageVersion("digest"))
  )
) |>
  audit_gt("Analytical environment")
```

### Every preparation artifact consumed

The table includes every shared preparation artifact read directly for this
Stage 1 reconstruction or required to interpret its provenance. H01 files are
used only as the approved shared metric/sample contract; no H01 model object,
fit, p-value, significance decision, or reported result is consumed.

```{r}
#| label: preparation-input-inventory
#| echo: false

input_inventory |>
  transmute(
    Input = input_id,
    Role = role,
    Path = path,
    Bytes = bytes,
    `SHA-256` = sha256
  ) |>
  audit_gt(
    "H10 preparation input inventory",
    "All 25 direct/provenance inputs existed and were hashed at render time."
  ) |>
  fmt_integer(columns = Bytes) |>
  cols_width(Path ~ px(340), `SHA-256` ~ px(430))
```

### Shared provenance gate

The current metric manifest is identified, but PREP-003/FIND-044 explicitly
states that the stored complete independent reconstruction belongs to an
earlier metric manifest and earlier 811/897-day sample. In addition, the
shared Preparation 06 gate still describes 811 near-eye and 897 chest
participant-days, whereas the current base objects contain 816 and 902. This
is not evidence that the current values are wrong; it is a version-specific
verification and documentation gap. H10 cannot repair coordinator-owned
records.

```{r}
#| label: shared-provenance-gate
#| echo: false

manifest_reconciliation |>
  rename(
    Check = check_id,
    Observed = observed,
    Status = status,
    Consequence = consequence
  ) |>
  audit_gt("Current preparation identities and unresolved gates")
```

## Preregistered contract

The preregistered hypothesis is:

> Personal light exposure metrics depend on age and sex.

The registered per-predictor model is printed below as an evaluated R formula
object.

```{r}
#| label: registered-formula
#| code-summary: "Exact registered H10 Wilkinson formula"

registered_formula
```

The registered participant structure is one random intercept for participant
within site. For participant-day outcomes, the proposed implementation writes
that structure explicitly as `site:Id`. IS and IV contain one row per
participant, so an ordinary site-adjusted model is the necessary transparent
adaptation; adding a participant random intercept to one observation per
participant would be fictitious.

The registration names biological sex, not gender. No temperature term or
other unsupported predictor is introduced.

## Construct and design audit

The normalized demographic artifact contains separate `sex` and `gender`
fields. All participants contributing at least one H10 outcome have complete
age and biological sex. Only Female and Male are observed for biological sex;
only Woman and Man are observed for gender in these placement-specific H10
frames. One participant at each placement is recorded as Male for biological
sex and Woman for gender. H10 therefore cannot relabel or substitute gender
without changing the registered construct and contrast membership.

```{r}
#| label: sex-gender-crosswalk
#| echo: false

sex_gender_crosswalk |>
  filter(data_scenario == "primary") |>
  transmute(
    Placement = placement_label(placement),
    `Biological sex` = biological_sex,
    Gender = gender,
    Participants = participants
  ) |>
  audit_gt(
    "Biological sex and gender are separate source variables",
    "Aggregated primary-dataset counts; only biological sex enters H10 inference."
  ) |>
  fmt_integer(columns = Participants)
```

For modelling, `age_decade = age / 10`. Biological sex is an unordered factor
with levels Male then Female, and the practical contrast is explicitly
Female-minus-Male. Gender remains descriptive only.

### Site support

Submitted-manuscript site names, order, and colours are used below. Both
biological-sex categories occur at every represented site. Nevertheless, San
José contributes only six near-eye participants (four Female, two Male), and
the smallest within-site age range is seven years. Full-rank design matrices
do not remove the resulting precision and influence risk.

```{r}
#| label: site-demographic-support
#| echo: false

site_demographics |>
  filter(data_scenario == "primary") |>
  transmute(
    Placement = placement_label(placement),
    Site = display_name,
    Colour = paste0("<span style=\"color:", color_hex, ";\">●</span> ", color_hex),
    Participants = participants,
    `Age, years` = paste0(age_min, "–", age_max, " (median ", age_median, ")"),
    Female,
    Male,
    Woman,
    Man
  ) |>
  audit_gt(
    "Site-specific demographic support",
    "Participant membership is the union of the 17 H10 candidate outcomes."
  ) |>
  fmt_markdown(columns = Colour) |>
  fmt_integer(columns = c(Participants, Female, Male, Woman, Man))
```

```{r}
#| label: design-rank-summary
#| echo: false

design_audit |>
  group_by(data_scenario, placement) |>
  summarise(
    metrics = n(),
    all_age_interactions_full_rank = all(age_full_rank),
    all_sex_interactions_full_rank = all(sex_full_rank),
    minimum_within_site_age_range_years = min(min_site_age_range_years),
    minimum_unique_ages_within_site = min(min_unique_ages_per_site),
    minimum_site_sex_cell_participants = min(min_site_sex_cell_participants),
    maximum_single_site_participant_share = max(max_site_participant_share),
    .groups = "drop"
  ) |>
  transmute(
    Scenario = scenario_label(data_scenario),
    Placement = placement_label(placement),
    Metrics = metrics,
    `Age interaction full rank` = all_age_interactions_full_rank,
    `Sex interaction full rank` = all_sex_interactions_full_rank,
    `Smallest site age range, years` = minimum_within_site_age_range_years,
    `Fewest unique ages in a site` = minimum_unique_ages_within_site,
    `Smallest site × sex cell` = minimum_site_sex_cell_participants,
    `Largest site share` = maximum_single_site_participant_share
  ) |>
  audit_gt(
    "Nominal interaction estimability and sparse-support warnings",
    "Rank is necessary but not sufficient; influence and interval width remain Stage 2 gates."
  ) |>
  fmt_percent(columns = `Largest site share`, decimals = 1) |>
  fmt_integer(columns = c(
    Metrics, `Smallest site age range, years`,
    `Fewest unique ages in a site`, `Smallest site × sex cell`
  ))
```

## Exact candidate samples

The tables give the exact pre-fit candidate frame for every metric, placement,
and required data scenario after applying metric availability, the declared
response construction, complete age and biological sex, and valid site and
participant keys. Model rows are participants for IS/IV and participant-days
for all other metrics. Support hours describe the exposure records used to
derive a metric; they are not model rows.

### Primary dataset

```{r}
#| label: primary-candidate-samples
#| echo: false

all_available_display |>
  filter(data_scenario == "primary") |>
  transmute(
    Metric = manuscript_name,
    Abbreviation = abbreviation,
    `Near-eye candidate sample` = glasses,
    `Chest candidate sample` = chest
  ) |>
  audit_gt(
    "Exact all-available H10 candidate samples: primary dataset",
    "Every candidate frame is shared by its five proposed formulas within metric and placement."
  ) |>
  cols_width(
    Metric ~ px(250),
    `Near-eye candidate sample` ~ px(390),
    `Chest candidate sample` ~ px(390)
  )
```

### Gap-timing-unaware dataset

The **gap-timing-unaware dataset** is the dataset applying the 50%-per-hour
and 80%-per-day coverage rules but not using the remaining gaps' time of day
in metric-specific support decisions. Thereafter, the current time-sensitive
data are called simply the primary dataset. Exact metric-specific derivation
hours are unavailable in this retained sensitivity artifact and are labelled
as unavailable, never as zero.

```{r}
#| label: gap-candidate-samples
#| echo: false

all_available_display |>
  filter(data_scenario == "gap_timing_unaware") |>
  transmute(
    Metric = manuscript_name,
    Abbreviation = abbreviation,
    `Near-eye candidate sample` = glasses,
    `Chest candidate sample` = chest
  ) |>
  audit_gt(
    "Exact all-available H10 candidate samples: gap-timing-unaware dataset",
    "The same approved H10 implementation will be used; only the prepared data change."
  ) |>
  cols_width(
    Metric ~ px(250),
    `Near-eye candidate sample` ~ px(390),
    `Chest candidate sample` ~ px(390)
  )
```

### Placement-matched/common samples

The paired/common sample retains identical site-participant-date keys at both
placements for participant-day outcomes. IS and IV are marked unavailable:
their participant-level values would have to be recomputed from the identical
paired day set, and all-available values are not a valid substitute.

```{r}
#| label: paired-candidate-samples
#| echo: false

paired_display |>
  transmute(
    Scenario = scenario_label(data_scenario),
    Metric = manuscript_name,
    Abbreviation = abbreviation,
    `Near-eye paired/common sample` = glasses,
    `Chest paired/common sample` = chest
  ) |>
  audit_gt(
    "Exact placement-matched/common H10 candidate samples",
    "No placement-equivalence test or pooled-placement model is proposed."
  ) |>
  cols_width(
    Metric ~ px(235),
    `Near-eye paired/common sample` ~ px(360),
    `Chest paired/common sample` ~ px(360)
  )
```

## What V0 implemented and reported

V0 used seven formula structures for each of 17 outcomes. Participant-day
models used a participant random intercept; V0 removed it from the two
participant-level models. It fitted all seven structures but compared only:

- `H10_0` with `H10_1a` for the age association;
- `H10_0` with `H10_1s` for the biological-sex association; and
- `H10_1a` with `H10_ia` for age-by-site heterogeneity.

`H10_is` was fitted but never compared. V0 also fitted age-only and
intercept-only structures; the age-only structure was later used for the
reported marginal R² even though the slope came from the site-adjusted model.

### Exact V0 response and formula reconstruction

The first visible evaluated cell prints the exact V0 response/family map. The
second constructs the exact Wilkinson formulas produced by V0's
`make_formula(response, rhs)` operation. The printed manifest is grouped where
multiple metrics genuinely shared an identical response and model structure.

```{r}
#| label: v0-response-map
#| code-summary: "Exact V0 response, engine, and family map"

v0_response_map
```

```{r}
#| label: v0-formulas
#| code-summary: "Evaluated exact V0 Wilkinson formula manifest"

for (i in seq_len(nrow(v0_formula_manifest))) {
  cat(
    "\n", v0_formula_manifest$metric_group[[i]],
    " | ", v0_formula_manifest$model_id[[i]], "\n",
    sep = ""
  )
  print(v0_formula_manifest$evaluated_formula[[i]])
}
```

V0 compared Gaussian fixed effects through the generic engine without an
explicit ML refit, used the default factor contrasts, and delegated Tweedie
comparisons to the same generic `anova()` wrapper. Its claimed site sum coding
was therefore not supplied to these models.

### Recovered V0 result record

The next table reproduces all 34 frozen V0 display rows from the rendered
near-eye and chest documents. These strings are preserved as historical
output; they are not valid BH-adjusted values and no Stage 1 significance
decision is made from them.

```{r}
#| label: v0-overall-results
#| echo: false

v0_overall_results |>
  transmute(
    Placement = placement_label(placement),
    Category = metric_type,
    Metric = metric_v0,
    `V0 sex adjusted-p display` = sex_adjusted_p_display,
    `V0 age adjusted-p display` = age_adjusted_p_display,
    `V0 age × site adjusted-p display` = age_site_adjusted_p_display
  ) |>
  audit_gt(
    "Frozen V0 H10 overall-result display",
    "Historical strings only; scalar p.adjust(..., n = 34) did not perform vector BH correction."
  )
```

The near-eye V0 selected one age row, time above 1,000 lx melEDI, and printed
an age slope of 0.024 h/year with a 0.011–0.037 interval and marginal R² of
0.074. The manuscript converted the slope to 1.44 min/year and approximately
3%–4% per year. All those numerical and interpretive claims are reopened. The
chest V0 selected five age rows, but several log-transformed Gaussian estimates
were blank because the table expected interval-column names from a different
engine. A blank cell is not a zero estimate.

### V0 output identity and collision

```{r}
#| label: v0-output-identity
#| echo: false

v0_output_identity |>
  rename(
    Output = output_id,
    Path = path,
    Exists = exists,
    `SHA-256` = sha256
  ) |>
  audit_gt(
    "V0 H10 durable-output identity",
    "Near-eye and chest source notebooks wrote the same table target; copied assets are not a deterministic provenance chain."
  ) |>
  cols_width(Path ~ px(330), `SHA-256` ~ px(430))
```

### Audit findings

Each finding states its evidence status. “Confirmed” means directly
reproduced from source, stored output, or current prepared inputs. The open
provenance gap is a documented absence of matching verification, not a claim
that current scientific values are wrong.

```{r}
#| label: audit-findings
#| echo: false

findings |>
  rename(
    ID = finding_id,
    Severity = severity,
    `Evidence status` = evidence_status,
    Domain = domain,
    Finding = finding,
    Consequence = consequence,
    `Required disposition` = required_disposition
  ) |>
  audit_gt("H10 Stage 1 audit findings")
```

## Claim-to-source traceability

```{r}
#| label: claim-source-traceability
#| echo: false

claim_source_matrix |>
  rename(
    `Claim or requirement` = claim_or_requirement,
    `Controlling source` = controlling_source,
    `Stage 1 disposition` = stage1_disposition
  ) |>
  audit_gt("H10 claim-to-source traceability matrix")
```

Reviewer feedback NM-R1-05, NM-R2-01, and NM-R2-03 is operationalized through
model-specific sample counts, explicit placement roles, complete uncertainty,
four vector-wide multiplicity families, sparse-cell reporting, diagnostics,
and participant/site influence checks.

## Proposed estimands

### Primary age estimand

For each metric, the primary near-eye age estimand is the common conditional
association per 10-year increase in age from `M_age`, adjusted for site. It is
not causal and does not assert that every site has the same age distribution.
On an identity scale it is a difference; on a base-10 log scale it is a ratio
of shifted values; on a log-link scale it is a response ratio. Per-year values
and, where meaningful, hours/minutes or percent change are derived from the
same estimate and 95% confidence interval.

### Primary biological-sex estimand

For each metric, the primary near-eye biological-sex estimand is the common
site-adjusted Female-minus-Male contrast from `M_sex`. It uses the measured
biological-sex variable only. The model is associational; it does not estimate
a gender effect or a causal effect of sex.

### Site heterogeneity

`M_age_site` and `M_sex_site` address whether the corresponding association
varies across sites. They are separately registered heterogeneity analyses,
not replacements for the two primary common associations. Equal-site and
observed-sample marginal summaries will both be shown. If they materially
differ, or if an interaction is supported, the common association is retained
only as an explicitly qualified summary and site-specific estimates are shown
without selective site-wise significance claims.

Chest repeats the same estimands as complementary evidence. It cannot rescue
or overturn a primary near-eye decision. Placements are never pooled and no
equivalence statement is permitted without a prespecified defensible margin.

## Exact proposed models

Each metric-specific Stage 2 frame will contain one numeric column named
`response`, constructed exactly as shown in the response contract below. The
formula objects printed in the next two visible evaluated cells are therefore
the exact Wilkinson formulas that will be supplied unchanged to `lm`, `lmer`,
or `glmmTMB`. Only the engine/family/link and participant structure differ by
metric analysis unit.

```{r}
#| label: proposed-response-contract
#| echo: false

metric_contract |>
  transmute(
    Metric = manuscript_name,
    Abbreviation = abbreviation,
    `Analysis unit` = recode(
      analysis_unit,
      participant_day = "participant-day",
      participant = "participant"
    ),
    `response construction` = response_construction,
    Engine = engine,
    Family = family,
    Link = link,
    `Practical effect` = practical_effect
  ) |>
  audit_gt(
    "Proposed 17-metric response, family, link, and effect-scale contract",
    "This inherits approved shared metric decisions, not a fitted H01 result."
  )
```

```{r}
#| label: proposed-participant-day-formulas
#| code-summary: "Exact proposed participant-day Wilkinson formulas"

for (model_id in names(participant_day_formulas)) {
  cat("\n", model_id, "\n", sep = "")
  print(participant_day_formulas[[model_id]])
}
```

```{r}
#| label: proposed-participant-formulas
#| code-summary: "Exact proposed participant-level Wilkinson formulas"

for (model_id in names(participant_formulas)) {
  cat("\n", model_id, "\n", sep = "")
  print(participant_formulas[[model_id]])
}
```

The formula hierarchy is fixed: every interaction model contains its main
effects. `site` is explicitly sum-coded using the registry order;
`biological_sex` has levels Male then Female with a Female-minus-Male reporting
contrast. Formula, contrasts, factor levels, frame hash, family, link, engine,
method, package versions, warnings, convergence state, and exact sample are
stored with every fit.

### Every proposed comparison

The code cell below prints the exact reduced and full formula objects used by
every comparison for both analysis units. No other Stage 2 inferential model
is proposed.

```{r}
#| label: proposed-comparison-formulas
#| code-summary: "Exact reduced/full Wilkinson formulas for every proposed comparison"

for (i in seq_len(nrow(comparison_manifest))) {
  row <- comparison_manifest[i, ]
  cat("\n", row$comparison_id, " — participant-day\n", sep = "")
  print(participant_day_formulas[[row$reduced_model]])
  print(participant_day_formulas[[row$full_model]])
  cat(row$comparison_id, " — participant\n", sep = "")
  print(participant_formulas[[row$reduced_model]])
  print(participant_formulas[[row$full_model]])
}
```

```{r}
#| label: comparison-manifest
#| echo: false

comparison_manifest |>
  rename(
    Comparison = comparison_id,
    Reduced = reduced_model,
    Full = full_model,
    Estimand = estimand,
    `Near-eye family` = family_id_near_eye,
    `Chest family` = family_id_chest
  ) |>
  audit_gt(
    "Proposed comparison and multiplicity manifest",
    "Each family contains all 17 planned metrics, including reason-coded unavailable/non-estimable rows."
  )
```

## Fitting, inference, and multiplicity plan

For each metric, placement, and scenario, Stage 2 will:

1. construct one complete comparison frame before fitting and save its exact
   keys, exclusion reasons, response, age, biological sex, site, and hash;
2. assert that `M0`, `M_age`, `M_age_site`, `M_sex`, and `M_sex_site` use the
   identical frame within that metric-placement-scenario;
3. compare Gaussian models with differing fixed effects under maximum
   likelihood, then obtain final estimates from the approved REML fit;
4. compare Tweedie-log `glmmTMB` models by their common-frame maximum
   likelihood;
5. retain raw p-values for all 17 planned rows and call `p.adjust()` once on
   each complete Benjamini–Hochberg vector;
6. store family ID, planned size 17, non-missing rank, raw p-value, adjusted
   p-value, and reason-coded unavailable/non-estimable status;
7. report every planned age slope and biological-sex contrast with a 95%
   confidence interval, irrespective of statistical significance; and
8. derive figures and durable tables only from paired machine-readable source
   CSVs.

The recommended family partition is four separate 17-member near-eye
families: `H10-F1-age-main`, `H10-F2-sex-main`,
`H10-F3-age-site-interaction`, and `H10-F4-sex-site-interaction`. Chest uses
the four analogous `H10-C*` complementary families. Keeping age and sex main
effects in one 34-member family is a coherent alternative interpretation of
the registration, but it must be explicitly chosen before repaired results
are examined. Interactions must not be mixed into a main-effect family.

Significance is decided from the unrounded adjusted value before display.
Stage 2 will format p-values to three decimal places, use “<0.001” only when
the value is strictly smaller than 0.001, and bold only values significant
under the explicitly labelled rule. Raw and adjusted values remain distinct.

### Confidence intervals and practical effects

The proposed primary intervals are model-based 95% intervals: t-based for
`lm`, fixed-effect covariance intervals for `lmer` and `glmmTMB`, and the
corresponding `emmeans` trend or contrast interval on the declared scale.
Log-scale and log-link estimates are back-transformed with both endpoints.
Timing results are expressed in clock minutes; duration effects include
hours/minutes where the model scale permits; age is reported per decade and
also per year.

No bootstrap is planned at this gate. If diagnostics make participant
resampling or another heavy resampling method necessary, Stage 2 may run only
a 50- or 100-replicate pilot, show its output and projected production
runtime, and return for approval before any production computation.

## Diagnostics and acceptance assessment

Stage 2 must present, interpret, and classify each metric-placement analysis
as **acceptable**, **acceptable with specified limitations**, or **not
acceptable for inference**. No diagnostic is satisfied merely by producing a
plot.

```{r}
#| label: diagnostic-plan
#| echo: false

diagnostic_plan |>
  rename(
    Domain = domain,
    `Planned evidence` = planned_evidence,
    `Acceptance rule` = acceptance_rule
  ) |>
  audit_gt("Predeclared H10 diagnostic and acceptance plan")
```

The registered participant random intercept is the starting structure. Age
and biological sex do not vary within participant, so a participant-specific
random slope for either predictor is unsupported. Repeated-day dependence is
instead checked directly from date-ordered residuals. A model change prompted
by those diagnostics is a new gate, not an automatic tuning step.

## Sensitivity plan and placement display

```{r}
#| label: sensitivity-plan
#| echo: false

sensitivity_plan |>
  rename(
    Scenario = scenario_id,
    Role = role,
    `Sample/data change` = sample_or_data_change,
    `Model change` = model_change,
    `Interpretive rule` = interpretive_rule
  ) |>
  audit_gt("Predeclared H10 sensitivity and influence scenarios")
```

For estimand-compatible participant-day outcomes, the Stage 2 placement
display will put near-eye and chest estimates with 95% intervals side by side
for both the all-available and paired/common samples. It will not connect
unmatched samples, pool placements, test equivalence, or imply that chest and
near-eye estimates are independent participant samples. IS and IV receive no
paired/common point until those metrics are validly recomputed on identical
paired days.

## V0 recreation and reporting plan

Stage 2 will first recreate the recognizable V0 table and age display from
the frozen specification, then show justified corrections:

- deterministic, placement-specific paths rather than `tables/H10.png`;
- all 17 age and biological-sex estimates and intervals rather than
  selection-only rows;
- the previously omitted sex-by-site family;
- a model-based site-adjusted age display rather than a pooled unadjusted
  smoother;
- correct duration, timing, melEDI, MDER, IS, and IV labels;
- one participant per demographic display row rather than repeated metric
  copies;
- primary near-eye and complementary chest evidence with paired/common
  comparison where estimands permit; and
- paired source-data CSVs for every durable figure and table.

Figures will be designed for their intended publication display width, use the
approved export-scaling rule, and be inspected at final size for clipping,
overlap, distortion, awkward line breaks, and excess whitespace. Any
nonnegative strongly right-skewed melEDI-like distribution that includes zero
will use `LightLogR::symlog_trans(base = 10, thr = 1, scale = 1)`. A4 is only
a physical-size QA scaffold, never the design canvas.

## Author decisions required before Stage 2

Every item requires explicit approval. Approval authorizes only Stage 2 and
the recommended dispositions; it does not authorize a deferred response-family
change, heavy production resampling, Stage 3, Stage 4, shared-ledger edits, or
shared preparation changes.

```{r}
#| label: author-decisions
#| echo: false

author_decisions |>
  rename(
    ID = decision_id,
    `Decision required` = decision_required,
    `Recommended disposition` = recommended_disposition,
    `Why it matters` = why_it_matters
  ) |>
  audit_gt(
    "H10 Stage 1 author-approval checklist",
    "Stage 2 remains blocked until all 13 items are resolved and the shared preparation prerequisite is answered."
  )
```

## Gate status

**Stop here.** Do not build `artifacts/06_model_data/H10.rds`; fit `M0`,
`M_age`, `M_age_site`, `M_sex`, or `M_sex_site`; run a diagnostic alternative,
bootstrap, or sensitivity model; edit `notebooks/hypotheses/H10.qmd`; or
change any shared preparation artifact, configuration file, ledger,
manuscript, or bibliography.

To open Stage 2, the author should explicitly approve the H10 Stage 1 proposal
and items `H10-G1` through `H10-G13`, or list the item IDs that need revision.
Stage 2 fitting also remains conditional on the coordinator-owned response to
`audit/handoffs/H10_shared_change_request.md`.
