---
title: "H06 Stage 1: audit and proposed analysis"
subtitle: "Nature Health analysis audit — revised robust-mean architecture gate"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    df-print: paged
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

library(digest)
library(dplyr)
library(gt)
library(purrr)
library(tibble)
library(tidyr)
library(xml2)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

read_project_csv <- function(path) {
  read.csv(
    file.path(project_root, path),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

read_project_rds <- function(path) {
  readRDS(file.path(project_root, path))
}

sha256_file <- function(path) {
  digest::digest(
    file.path(project_root, path),
    algo = "sha256",
    file = TRUE
  )
}

compact_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(12),
      table.width = gt::pct(100),
      data_row.padding = gt::px(3),
      heading.align = "left",
      source_notes.font.size = gt::px(10)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

site_registry <- read_project_csv("config/site_display_registry.csv") |>
  arrange(display_order)
site_levels <- site_registry$site

metric_display <- read_project_csv("config/metric_display_registry.csv")
h05_metric_registry <- read_project_csv(
  "artifacts/06_model_data/H05/H05_metric_registry.csv"
)
h03_route_registry <- read_project_csv(
  "artifacts/06_model_data/H06/H06_h03_robust_route_pilot_registry.csv"
)
h03_route_diagnostics <- read_project_csv(
  "artifacts/08_diagnostics/H06/H06_h03_robust_route_pilot_fit_diagnostics.csv"
)
h03_route_mean_gate <- read_project_csv(
  "artifacts/08_diagnostics/H06/H06_h03_robust_route_pilot_mean_structure_gate.csv"
)
h03_route_power_sensitivity <- read_project_csv(
  "artifacts/08_diagnostics/H06/H06_h03_robust_route_pilot_power_sensitivity.csv"
)
h03_route_summary <- read_project_csv(
  "artifacts/09_tables/H06/H06_h03_robust_route_pilot_summary.csv"
)
two_part_formula_registry <- read_project_csv(
  "artifacts/06_model_data/H06/H06_exploratory_two_part_formula_registry.csv"
)

stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0"),
  identical(as.character(utils::packageVersion("melidosData")), "1.0.6")
)
```

# Stage 1 verdict

::: {.callout-warning title="Revised H06 Stage 1 gate — author approval required"}
**Option B remains the proposed estimand, but its inferential architecture is
reopened.** The earlier Stage 2 work showed that the approved
participant/participant-day/AR Tweedie mixed architecture cannot support H06
inference. The owner therefore asked H06 to check the accepted H03 solution
and run a small transfer pilot.

H03 did not repair its mixed model. It changed to a population-mean log-link
GLM and treated the participant as the inferential replicate through
participant-cluster sandwich covariance. The H06 transfer pilot supports the
same route. The proposed H06 primary is now a zero-inclusive
**quasi-Poisson log-mean model with participant-cluster HC3 covariance**, fixed
site, the V0-recognizable three-predictor interaction scaffold, and finite
`G - 1` reference degrees of freedom. It contains no participant random
intercept, participant-day random intercept, latent AR term, working-AR term,
or confirmatory local-clock term.

The fixed-power quasi-Tweedie version with `p = 1.8` is a working-variance
sensitivity. The H03 value `p = 1.539919` is retained only as an architecture
comparator and is not imported as an H06 parameter. A value of `p = 1.9`,
motivated by the earlier H06 mixed fits, failed numerically for near eye and
is rejected. No predictor estimate, confidence interval, omnibus statistic,
p-value, multiplicity decision, or scientific claim was inspected to choose
this architecture.

The pilot passes the numerical construction gate at both placements, but it
does not make the hours independent: residual lag-1 correlation remains
material, and targeted participant deletion moves a core coefficient by up
to about one full-model HC3 standard error. These are mandatory limitations
and influence checks, not reasons to add a post hoc clock term or delete a
participant. This report fits no model; it reads the frozen bounded-pilot
artifacts and stops for approval of revised gate **H06-G1A**.
:::

The owner previously selected the repaired deliberate hourly adaptation
(Option B). The signed record specifies daily light-exposure metrics, whereas
the submitted implementation (hereafter **V0**) fitted repeated hourly
geometric mean melEDI. That change alters the outcome, unit, weighting of
participant-days, temporal dependence, and multiplicity family. The two
analyses remain different questions even though the adapted question remains
the proposed H06 primary estimand.

The owner has clarified that the hourly choice was deliberate. DEV-015 gives
the recorded rationale: the candidate metric-by-predictor set was judged
insufficiently parsimonious and interpretable. The outcome adaptation is
therefore **not an accidental substitution and not, by itself, an
implementation error**. The intent clarification does not make the submitted
dependence structure, hierarchy, multiplicity, or reporting valid, and it
did not by itself resolve the final estimand. The owner has now resolved
H06-G1 in favor of the repaired hourly question.

The approved primary predictor set is work/free day, binary activity status,
and previous-night sleep duration. Binary activity is defined transparently
as **Sedentary (no reported exercise)** versus **Active (light, moderate, or
vigorous reported exercise)**. Weekday/weekend is a named sensitivity. The
other registered day-type, exercise, and sleep variables may be explored but
are excluded from primary inference and its multiplicity families. The
primary response is one supported one-hour zero-aware geometric mean melEDI.

The confirmatory mean model estimates an average association over retained
supported hours and intentionally has no local-clock fixed effect. It is
therefore an observed-supported-hour-weighted average, not a
clock-standardized effect. Participant-cluster covariance allows arbitrary
dependence among a participant's retained hours and days for uncertainty; it
does not whiten residuals or repair a misspecified conditional mean.
Time-of-day patterning remains reserved for the separate exploratory
GAM/GAMM, whose entire workflow must remain compatible with
`bam(discrete = TRUE)`.

No proposed H06 inferential model, sensitivity model, temporal model,
bootstrap, or new effect estimate is fitted in this render. The bounded
architecture pilot was run once in R 4.6.1 by
`scripts/hypotheses/H06/pilot_h06_h03_robust_mean_route.R`; this report reads
its frozen numerical-adequacy artifacts alongside the original Stage 1 source
verification, V0 reconstruction, join/cardinality checks, and descriptive
candidate counts.

## Re-audit clarification

The owner supplied the controlling clarifications and decisions in this H06
task on 2026-08-01. The coordinating task confirmed the first two
project-level dispositions:

1. DEV-015 is the controlling outcome-deviation record. Its rationale is
   deliberate parsimony and interpretability. The owner now selects its
   repaired hourly estimand as primary.
2. DEV-056 is already approved. The analytical source remains the immutable
   TUM exercise-diary pin under `melidosData` 1.0.6. H06 must not fetch remote
   HEAD, update the package, or rebuild shared preparation merely to look for
   a newer release. A different release would require a separately pinned and
   coordinated comparison because it would reopen shared preparation and
   downstream hypotheses.
3. Work/free day is primary, weekday/weekend is a sensitivity, and the
   four-level exercise source is collapsed to the binary activity status
   defined above. Other named measures are exploratory only.
4. Fixed site with separate hierarchy-preserving heterogeneity tests,
   sum-to-zero contrasts, and equal-site reader-facing marginalization is
   primary. The registered random-site structure remains visible only as the
   unselected registered reference after the mixed architecture failed.
5. Participant and participant-day intercepts plus gap-bounded within-day
   AR(1) were approved at the original gate, then failed the Stage 2
   adequacy checks. The 2026-08-10 H03-informed amendment replaces them with
   participant-cluster robust covariance in a population-mean GLM. A
   confirmatory local-clock effect remains explicitly rejected so the
   reported association is averaged over retained supported hours.
6. The response, multiplicity, placement, sensitivity, interval, diagnostic,
   failure, resampling, and V0 claim gates are approved as revised below.
7. The `KNUST_S005` sitting/reclining entry recorded as 3,600 minutes is
   interpreted for H06 exploratory work as 3,600 seconds, or 1 hour. The
   shared source remains unchanged and the one-row reinterpretation is
   provenance-labelled.

This revision supersedes any earlier Stage 1 wording that characterized the
hourly choice itself as an error. It does not excuse the separable V0 defects
documented below.

## Meaning of V0

**V0** means the submitted H06 code and frozen outputs in `RQ2.qmd`,
`RQ2_chest.qmd`, `tables/H6.docx`, `tables/chest/H6.docx`, and the linked
figures and manuscript prose. It does not mean the proposed main analysis or
the **gap-timing-unaware dataset** sensitivity. At this first use, that
comparison dataset still passed the general 50%-per-hour and 80%-per-day
coverage rules, but the timing of the remaining missing observations was not
used for an additional metric-specific adjustment. For contrast only here,
the new preparation could be interpreted as a time-sensitive primary metric
dataset. This report calls it simply the **primary dataset** from this point
onward.

# Preregistered contract

The signed AsPredicted record states, exactly:

> **H6: Day type, daily exercise, and sleep variables predict daily light exposure metrics.**

— `preregistration/AsPredicted #273407.pdf`, p. 1.

The registered model on p. 2 is:

```text
Metric ~ Measure + (Measure | Site) + (1 | Site:Participant)
```

The registered unit is a participant-day nested in participant and site. Day
type includes weekday/weekend and free/work day. The supporting analysis plan
also names sleep onset, final wake time, and sleep duration, and records daily
exercise intensity/location, active travel, sedentary time, and whether the
logger was worn during exercise. It does not state a complete numeric family
or how these measures should be partitioned for false-discovery-rate control.

```{r}
#| label: contract-source-verification

contract_sources <- tibble::tribble(
  ~source, ~expected_sha256, ~role,
  "preregistration/AsPredicted #273407.pdf",
  "fce192c88676602b72013e4872a03b535a2f4ef59d34de3a9e1c47527742b7ad",
  "Controlling hypothesis wording and model",
  "preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx",
  "6e56766c18c476826c9c798799aa925752f11e5278f196797a583a13cc61ddf8",
  "Supporting predictor definitions"
) |>
  mutate(
    observed_sha256 = map_chr(source, sha256_file),
    verified = observed_sha256 == expected_sha256
  )

stopifnot(all(contract_sources$verified))

contract_sources |>
  compact_gt(
    source_note = "SHA-256 identities are checked during render; neither source is modified."
  ) |>
  gt::cols_label(
    source = "Source",
    expected_sha256 = "Expected SHA-256",
    observed_sha256 = "Observed SHA-256",
    role = "Stage 1 role",
    verified = "Check"
  ) |>
  gt::fmt(columns = verified, fns = function(x) ifelse(x, "PASS", "FAIL"))
```

```{r}
#| label: deliberate-outcome-deviation

dev015 <- read_project_csv("audit/ledgers/deviation_register.csv") |>
  filter(deviation_id == "DEV-015")

stopifnot(
  nrow(dev015) == 1L,
  dev015$preregistered_or_expected == "Daily light-exposure metrics",
  dev015$observed_or_approved == "Hourly geometric mean melanopic EDI",
  dev015$rationale_or_evidence == paste(
    "The candidate metric-by-predictor model set was judged insufficiently",
    "parsimonious and interpretable"
  ),
  dev015$status == "open_reaudit"
)

dev015 |>
  select(
    deviation_id, status, preregistered_or_expected,
    observed_or_approved, rationale_or_evidence, likely_consequence
  ) |>
  compact_gt(
    source_note = paste0(
      "The owner confirms that this adaptation was deliberate. The ledger's ",
      "open_reaudit value predates the author decision. This H06-owned report ",
      "records Option B as selected and requests a coordinator gate update; it ",
      "does not edit the shared ledger."
    )
  ) |>
  gt::cols_label(
    deviation_id = "Deviation",
    status = "Current gate status",
    preregistered_or_expected = "Registered outcome",
    observed_or_approved = "Deliberate V0 outcome",
    rationale_or_evidence = "Recorded rationale",
    likely_consequence = "Consequence requiring audit"
  )
```

# Authoritative current inputs

## File identities, runtime, and release

```{r}
#| label: current-input-contract

input_contract <- tibble::tribble(
  ~role, ~relative_path, ~expected_sha256,
  "Primary near-eye participant-day metrics",
  "artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds",
  "fc80f8a2a94a447c5470db36c1ad1915868463501f04e9db3218f8156efc59d2",
  "Complementary chest participant-day metrics",
  "artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds",
  "ce81c159fe018f35bec9bff44c9cd12f91fd4be1ea2696c56a8e91e628a9e2d3",
  "Primary near-eye one-hour metrics",
  "artifacts/06_model_data/base/metrics_glasses_one_hour_context.rds",
  "cee864bb0e329b98444088b777c7245250be68111ec9f24db7c0e3f395ed6445",
  "Complementary chest one-hour metrics",
  "artifacts/06_model_data/base/metrics_chest_one_hour_context.rds",
  "50f50ca5d39d621d06f2794037c4b769e01fc7329a70432e79a51203727ed209",
  "Normalized exercise diary",
  "artifacts/06_model_data/normalized_inputs/exercisediary.rds",
  "5bffe44cdd9c65f3d1dc20575d10b3f0a403577f3cf9109e83cfea95a2ae7107",
  "Normalized sleep diary",
  "artifacts/06_model_data/normalized_inputs/sleepdiaries.rds",
  "110819d74503c895170552c1f2214aca83de51cadd23bde7525aa821ee0b0e15",
  "Base-model manifest",
  "artifacts/12_manifests/base_model_data_artifacts.csv",
  "142d1044dd7e6a566e45e3a5c56033b1e8251e100c0d7c7523564c824950a0e4",
  "Gap-timing-unaware participant-day metrics",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
  "28266064c6213eae6046ec6469d0dae39529f56e60933dcd581cf96c8590e1a4",
  "Gap-timing-unaware one-hour metrics",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/one_hour_data.rds",
  "3c9a44d67d3267a3daa2a1392b096bdd89d80d62155edc44bdfe6c048ac4c105"
) |>
  mutate(
    observed_sha256 = map_chr(relative_path, sha256_file),
    verified = observed_sha256 == expected_sha256
  )

stopifnot(all(input_contract$verified))

input_contract |>
  select(role, relative_path, observed_sha256, verified) |>
  compact_gt(
    source_note = paste0(
      "The current base manifest is the bundle approved in H02-INPUT-001; ",
      "H06 independently verifies its direct inputs."
    )
  ) |>
  gt::cols_label(
    role = "Role",
    relative_path = "Exact path",
    observed_sha256 = "SHA-256",
    verified = "Check"
  ) |>
  gt::fmt(columns = verified, fns = function(x) ifelse(x, "PASS", "FAIL"))
```

The current and coordinator-confirmed project release is `melidosData` 1.0.6
under R 4.6.1. This is the analytical source contract, not a claim about a
remote repository's HEAD. The owner and coordinator explicitly require H06
to retain this pinned release: fetching or substituting a newer remote release
would reopen shared preparation and downstream hypotheses. No fetch is needed
because the verified normalized artifact already contains the approved
correction. The older narrative file
`audit/reconciliation/preparation06/base_model_gate.md` still reports 811/897
participant-days and older hashes. The later approved transition
`H02-INPUT-001` and the current verified manifest report 816/902; the
narrative drift is recorded in the H06 shared-change request and is not used
as an input authority here.

## Corrected TUM exercise participant

```{r}
#| label: corrected-tum-participant

exercise_raw <- read_project_rds(
  "artifacts/06_model_data/normalized_inputs/exercisediary.rds"
)
sleep_raw <- read_project_rds(
  "artifacts/06_model_data/normalized_inputs/sleepdiaries.rds"
)

source_pin <- read_project_csv("config/model_input_source_pins.csv") |>
  filter(site == "TUM", modality == "exercisediary")

correction_dates <- as.Date("2024-05-13") + 0:6
tum_correction <- exercise_raw |>
  filter(
    site == "TUM",
    Date %in% correction_dates,
    Id %in% c("TUM_S001", "TUM_S101")
  ) |>
  count(Id, Date, name = "Diary rows") |>
  arrange(Date)

stopifnot(
  nrow(source_pin) == 1L,
  source_pin$commit == "618fda8521f3cf581661ceb2c026e3de7cd9ba82",
  source_pin$expected_sha256 ==
    "0db56324e9826d427c42984fdfc43eb1e7487211365f222a1c0b2ecff618f1ef",
  source_pin$doi == "10.5281/zenodo.16893901",
  nrow(tum_correction) == 7L,
  all(tum_correction$Id == "TUM_S001"),
  sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S101") == 0L
)

tum_correction |>
  mutate(
    `Participant identifier` = sub("TUM_", "", Id),
    `Source commit` = source_pin$commit,
    `Source DOI` = source_pin$doi
  ) |>
  select(Date, `Participant identifier`, `Diary rows`, `Source commit`, `Source DOI`) |>
  compact_gt(
    source_note = paste0(
      "The normalized input is read from the immutable current TUM source; ",
      "no H06 or normalization-time identifier rewrite is applied."
    )
  ) |>
  gt::fmt_date(columns = Date, date_style = "iso") |>
  gt::fmt_integer(columns = `Diary rows`)
```

The approved DEV-056 transition is already present: the corrected participant
is **S001, not S101**. This is a passed provenance check, not a discrepancy or
an upstream change request. Six of these seven diary dates have current
near-eye and chest participant-day metric rows; the first date has no eligible
light-metric day. H06 will join the corrected diary as provided, will repeat
the seven-S001/zero-S101 assertion at Stage 2, and will neither fetch nor edit
shared data.

# Predictor definitions, timing, and keys

## One-row-per-participant-day construction

```{r}
#| label: predictor-table-construction

clock_hour <- function(x) {
  as.numeric(format(x, "%H")) + as.numeric(format(x, "%M")) / 60
}

exercise_intensity_levels <- c(
  "No exercise",
  "Light",
  "Moderate",
  "Vigorous"
)

activity_status_levels <- c(
  "Sedentary (no reported exercise)",
  "Active (light, moderate, or vigorous exercise)"
)

exercise_day <- exercise_raw |>
  transmute(
    site,
    Id,
    local_date = Date,
    exercise_intensity = recode(
      as.character(intensity),
      "None of the above, I did not perform any type of physical activity" =
        "No exercise",
      "Light (causing small to no increases in heart rate and breathing, e.g. taking a stroll in the park)" =
        "Light",
      "Moderate (causing moderate increases in heart rate and breathing, e.g. cycling in the city)" =
        "Moderate",
      "Vigorous (causing large increases in heart rate and breathing, e.g. running)" =
        "Vigorous"
    ) |>
      factor(levels = exercise_intensity_levels),
    activity_status = case_when(
      is.na(exercise_intensity) ~ NA_character_,
      exercise_intensity == "No exercise" ~ activity_status_levels[[1]],
      TRUE ~ activity_status_levels[[2]]
    ) |>
      factor(levels = activity_status_levels),
    exercise_location = recode(
      as.character(location),
      "Outdoors (e.g. running, cycling in the city)" = "Outdoors",
      "Indoors (e.g. gym or home workout)" = "Indoors",
      "Both indoors and outdoors" = "Indoors and outdoors"
    ) |>
      factor(levels = c("Outdoors", "Indoors", "Indoors and outdoors")),
    active_commute_h = as.numeric(commute, units = "hours"),
    sedentary_source_minutes = as.numeric(sedentary, units = "mins"),
    sedentary_h_as_recorded = as.numeric(sedentary, units = "hours"),
    sedentary_h = if_else(
      site == "KNUST" &
        Id == "KNUST_S005" &
        local_date == as.Date("2024-11-04") &
        sedentary_source_minutes == 3600,
      1,
      sedentary_h_as_recorded
    ),
    sedentary_reinterpretation = if_else(
      site == "KNUST" &
        Id == "KNUST_S005" &
        local_date == as.Date("2024-11-04") &
        sedentary_source_minutes == 3600,
      "Author-approved: source numeral interpreted as 3,600 seconds (1 h)",
      NA_character_
    ),
    wore_light_logger_during_exercise = factor(
      light_glasses,
      levels = c(0, 1),
      labels = c("No", "Yes")
    ),
    exercise_source_row = source_row
  )

sedentary_reinterpreted_row <- exercise_day |>
  filter(!is.na(sedentary_reinterpretation))

stopifnot(
  nrow(sedentary_reinterpreted_row) == 1L,
  sedentary_reinterpreted_row$sedentary_h_as_recorded == 60,
  sedentary_reinterpreted_row$sedentary_h == 1
)

sleep_day <- sleep_raw |>
  transmute(
    site,
    Id,
    local_date = as.Date(wake_wall),
    work_free_day = factor(
      as.character(daytype2),
      levels = c("a work day", "a free day"),
      labels = c("Work day", "Free day")
    ),
    previous_sleep_duration_h = as.numeric(sleep_duration, units = "hours"),
    previous_sleep_duration_centered_h = previous_sleep_duration_h - 8,
    previous_sleep_onset_clock_h = clock_hour(sleep_onset_wall),
    previous_sleep_onset_after_noon_h =
      (previous_sleep_onset_clock_h - 12) %% 24,
    previous_sleep_onset_centered_h =
      previous_sleep_onset_after_noon_h - 11,
    wake_clock_h = clock_hour(wake_wall),
    wake_centered_h = wake_clock_h - 7,
    sleep_onset_date = as.Date(sleep_onset_wall),
    sleep_interval_analysis_eligible,
    sleep_interval_quarantined,
    sleep_source_row = source_row
  )

exercise_duplicate_keys <- exercise_day |>
  count(site, Id, local_date, name = "rows") |>
  filter(rows != 1L)
sleep_duplicate_keys <- sleep_day |>
  filter(!is.na(local_date)) |>
  count(site, Id, local_date, name = "rows") |>
  filter(rows != 1L)

stopifnot(
  nrow(exercise_duplicate_keys) == 0L,
  nrow(sleep_duplicate_keys) == 0L,
  sum(is.na(sleep_day$local_date)) == 1L,
  sum(sleep_day$sleep_interval_quarantined %in% TRUE) == 1L
)

key_audit <- tibble::tribble(
  ~check, ~result, ~interpretation,
  "Exercise diary participant-day duplicates",
  nrow(exercise_duplicate_keys),
  "Must be zero before a many-to-one join",
  "Sleep diary wake-date participant-day duplicates",
  nrow(sleep_duplicate_keys),
  "Must be zero before a many-to-one join",
  "Sleep rows without a wake date",
  sum(is.na(sleep_day$local_date)),
  "One MPI row is retained but quarantined",
  "TUM S101 exercise rows",
  sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S101"),
  "Must be zero",
  "TUM S001 corrected dates",
  nrow(tum_correction),
  "Must equal seven",
  "Author-approved 3,600-second reinterpretations",
  nrow(sedentary_reinterpreted_row),
  "Exactly one H06-local exploratory record"
)

key_audit |>
  compact_gt(
    source_note = "All joins proposed below assert a many-to-one diary relationship."
  ) |>
  gt::fmt_integer(columns = result) |>
  gt::cols_label(
    check = "Audit check",
    result = "Observed",
    interpretation = "Contract"
  )
```

The exercise date is the recorded diary `Date`. The sleep interval is attached
to the local calendar date of final wake, exactly matching V0's intended
“previous night” linkage but using the normalized local-wall representation.
The source-calculated sleep duration is retained exactly and expressed in
hours; it is not recalculated from timestamps. Sleep onset and wake are local
clock measures. Sleep onset is unwrapped at noon and centered at 23:00 only if
that measure is approved; wake is centered at 07:00. These centers change the
intercept, not the slope.

```{r}
#| label: previous-night-linkage

previous_night_linkage <- sleep_day |>
  mutate(
    `Onset-to-wake date difference` = as.integer(local_date - sleep_onset_date)
  ) |>
  count(`Onset-to-wake date difference`, name = "Diary rows", .drop = FALSE) |>
  arrange(`Onset-to-wake date difference`)

stopifnot(
  sum(previous_night_linkage$`Diary rows`, na.rm = TRUE) == nrow(sleep_day),
  all(
    previous_night_linkage$`Onset-to-wake date difference`[
      !is.na(previous_night_linkage$`Onset-to-wake date difference`)
    ] %in% c(0L, 1L)
  )
)

previous_night_linkage |>
  mutate(
    `Link meaning` = case_when(
      `Onset-to-wake date difference` == 0L ~
        "Sleep began after midnight and ended on the same local date",
      `Onset-to-wake date difference` == 1L ~
        "Sleep began on the preceding local date",
      TRUE ~ "Missing wake; quarantined"
    )
  ) |>
  compact_gt(
    source_note = "No eligible interval starts after its assigned wake date."
  ) |>
  gt::fmt_integer(columns = `Diary rows`)
```

## Exact construct map and resolved exploratory range issue

```{r}
#| label: predictor-construct-map

predictor_contract <- tibble::tribble(
  ~measure_id, ~reader_name, ~source_field, ~unit_or_levels, ~date_rule,
  ~proposed_role,
  "work_free_day", "Work day versus free day", "sleepdiaries$daytype2",
  "Work day (reference), free day", "Date of final wake",
  "Approved primary measure",
  "weekday_weekend", "Weekday versus weekend", "derived from local_date",
  "Weekday (reference), weekend", "Metric participant-day date",
  "Approved named sensitivity",
  "activity_status", "Daily activity status", "exercisediary$intensity",
  paste(activity_status_levels, collapse = "; "), "Recorded exercise Date",
  "Approved primary binary measure",
  "exercise_intensity", "Daily exercise intensity", "exercisediary$intensity",
  "No exercise (reference), light, moderate, vigorous", "Recorded exercise Date",
  "Source audit and exploratory detail only",
  "exercise_location", "Exercise location", "exercisediary$location",
  "Outdoors, indoors, indoors and outdoors; only recorded on exercise days",
  "Recorded exercise Date", "Exploratory only; outside primary inference",
  "active_commute_h", "Walking/cycling travel time", "exercisediary$commute",
  "Hours; source question records minutes", "Recorded exercise Date",
  "Exploratory only; MPI unavailable",
  "sedentary_h", "Sitting/reclining time", "exercisediary$sedentary",
  "Hours; one identified numeral is interpreted as 3,600 seconds = 1 h",
  "Recorded exercise Date", "Exploratory only with row-level provenance",
  "wore_light_logger_during_exercise", "Logger worn during exercise",
  "exercisediary$light_glasses", "No/yes; not applicable or missing on many days",
  "Recorded exercise Date", "Measurement-support sensitivity, not an exposure predictor",
  "previous_sleep_onset_centered_h", "Previous sleep onset", "sleepdiaries$sleep",
  "Local clock hours after noon, centered at 23:00", "Date of final wake",
  "Exploratory only",
  "wake_centered_h", "Final wake time", "sleepdiaries$wake",
  "Local clock hours, centered at 07:00", "Date of final wake",
  "Exploratory only",
  "previous_sleep_duration_centered_h", "Previous sleep duration",
  "sleepdiaries$sleep_duration", "Elapsed hours, centered at 8 h",
  "Date of final wake", "Approved primary measure"
)

predictor_contract |>
  compact_gt(
    source_note = paste0(
      "Free-text exercise type and translated type are excluded from the ",
      "analytical RDS and are not candidates. The four source intensity levels ",
      "remain visible for provenance; primary inference uses binary activity. ",
      "Temperature and other unsupported predictors are not added."
    )
  ) |>
  gt::cols_label(
    measure_id = "Measure ID",
    reader_name = "Reader-facing construct",
    source_field = "Authoritative field",
    unit_or_levels = "Operational value",
    date_rule = "Participant-day linkage",
    proposed_role = "Stage 1 proposal"
  )
```

The normalized exercise diary contains one source numeral of 3,600 under the
minutes-labelled sitting/reclining field for `KNUST_S005` on 2024-11-04. The
owner has specified that this numeral represents **3,600 seconds**, or **1
hour**. H06 preserves the shared value unchanged and creates one explicitly
labelled exploratory reinterpretation. This resolves the H06 gate for
exploratory sedentary-time work; sedentary time remains outside the primary
analysis. The maximum active-travel value is 700 minutes and is retained only
for prespecified exploratory range and influence diagnostics rather than
silently winsorized.

```{r}
#| label: predictor-range-audit

range_audit <- tibble::tribble(
  ~measure, ~minimum, ~median, ~maximum, ~missing, ~unit,
  "Active walking/cycling travel",
  min(exercise_day$active_commute_h, na.rm = TRUE),
  median(exercise_day$active_commute_h, na.rm = TRUE),
  max(exercise_day$active_commute_h, na.rm = TRUE),
  sum(is.na(exercise_day$active_commute_h)), "h",
  "Sitting/reclining as source-labelled",
  min(exercise_day$sedentary_h_as_recorded, na.rm = TRUE),
  median(exercise_day$sedentary_h_as_recorded, na.rm = TRUE),
  max(exercise_day$sedentary_h_as_recorded, na.rm = TRUE),
  sum(is.na(exercise_day$sedentary_h_as_recorded)), "h",
  "Sitting/reclining for H06 exploration",
  min(exercise_day$sedentary_h, na.rm = TRUE),
  median(exercise_day$sedentary_h, na.rm = TRUE),
  max(exercise_day$sedentary_h, na.rm = TRUE),
  sum(is.na(exercise_day$sedentary_h)), "h",
  "Previous sleep duration",
  min(sleep_day$previous_sleep_duration_h, na.rm = TRUE),
  median(sleep_day$previous_sleep_duration_h, na.rm = TRUE),
  max(sleep_day$previous_sleep_duration_h, na.rm = TRUE),
  sum(is.na(sleep_day$previous_sleep_duration_h)), "h",
  "Previous sleep onset after noon",
  min(sleep_day$previous_sleep_onset_after_noon_h, na.rm = TRUE),
  median(sleep_day$previous_sleep_onset_after_noon_h, na.rm = TRUE),
  max(sleep_day$previous_sleep_onset_after_noon_h, na.rm = TRUE),
  sum(is.na(sleep_day$previous_sleep_onset_after_noon_h)), "h after 12:00",
  "Final wake time",
  min(sleep_day$wake_clock_h, na.rm = TRUE),
  median(sleep_day$wake_clock_h, na.rm = TRUE),
  max(sleep_day$wake_clock_h, na.rm = TRUE),
  sum(is.na(sleep_day$wake_clock_h)), "clock h"
)

range_audit |>
  compact_gt(
    source_note = paste0(
      "These are source/derived-value diagnostics, not model estimates. The ",
      "source-labelled 60 h and author-approved exploratory 1 h values are ",
      "shown separately. One daytime sleep onset (10:10) and one late wake ",
      "(15:24) require exploratory influence checks."
    )
  ) |>
  gt::fmt_number(columns = c(minimum, median, maximum), decimals = 2) |>
  gt::fmt_integer(columns = missing) |>
  gt::cols_label(
    measure = "Measure",
    minimum = "Minimum",
    median = "Median",
    maximum = "Maximum",
    missing = "Missing rows",
    unit = "Unit"
  )
```

# H06-G1 candidate outcomes and samples

## Registered daily option: metric-specific analysis units

```{r}
#| label: daily-metric-registry

metric_columns <- c(
  daily_geometric_mean_medi = "daily_geometric_mean_medi_lx",
  m10_mean_medi = "m10_mean_medi_lx",
  l10_mean_medi = "l10_mean_medi_lx",
  duration_above_1000 = "duration_above_1000_h",
  duration_above_250_wake = "duration_above_250_wake_h",
  duration_below_10_pre_sleep = "duration_below_10_pre_sleep_h",
  duration_below_1_sleep_environment = "duration_below_1_sleep_environment_h",
  longest_bout_above_250 = "longest_bout_above_250_h",
  m10_midpoint = "m10_midpoint_clock_minute",
  l10_midpoint = "l10_midpoint_clock_minute",
  mean_timing_above_250 = "mean_timing_above_250_clock_minute",
  first_timing_above_250 = "first_timing_above_250_clock_minute",
  last_timing_above_250 = "last_timing_above_250_clock_minute",
  dose_time_sensitive_corrected_medi = "dose_corrected_medi_lx_h",
  mder_ratio_of_integrals = "mder"
)

daily_metric_contract <- h05_metric_registry |>
  filter(analysis_unit == "participant_day") |>
  mutate(
    source_column = unname(metric_columns[metric_id]),
    manuscript_name = ifelse(
      metric_id == "longest_bout_above_250",
      "Longest continuous period above 250 lx melEDI",
      manuscript_name
    )
  ) |>
  select(
    metric_order, metric_id, manuscript_name, display_unit,
    response_family, response_transform, effect_scale,
    diagnostic_note, source_column
  ) |>
  arrange(metric_order)

stopifnot(
  nrow(daily_metric_contract) == 15L,
  !anyNA(daily_metric_contract$source_column),
  all(daily_metric_contract$metric_id %in% names(metric_columns))
)

daily_metric_contract |>
  compact_gt(
    source_note = paste0(
      "The accepted H05 registry is reused only as a 17-metric naming and ",
      "implementation convention. H05 estimates, conclusions, and diagnostic ",
      "dispositions are not inherited. IS and IV are excluded here because ",
      "their analysis unit is participant, not participant-day."
    )
  ) |>
  gt::cols_hide(columns = c(metric_order, metric_id, source_column)) |>
  gt::cols_label(
    manuscript_name = "Metric",
    display_unit = "Unit",
    response_family = "Continuity family",
    response_transform = "Continuity transform",
    effect_scale = "Effect scale",
    diagnostic_note = "Required metric-specific check"
  )
```

These family and transformation entries are continuity candidates only if
H06-G1 selects the registered daily option. They do not transfer H05 model
adequacy to H06. Stage 2 must evaluate each selected outcome's distribution
and physical bounds on the exact H06 frame before any result is accepted.

## Exact candidate samples after one-to-one diary joins

```{r}
#| label: candidate-sample-construction

near_day <- read_project_rds(
  "artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds"
)
chest_day <- read_project_rds(
  "artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds"
)

make_daily_candidate <- function(metric_data, placement) {
  metric_data |>
    select(site, Id, local_date, all_of(unname(metric_columns))) |>
    left_join(
      sleep_day |>
        select(-sleep_onset_date),
      by = c("site", "Id", "local_date"),
      relationship = "many-to-one"
    ) |>
    left_join(
      exercise_day,
      by = c("site", "Id", "local_date"),
      relationship = "many-to-one"
    ) |>
    mutate(
      placement = placement,
      participant_key = paste(site, Id, sep = "::"),
      participant_day_key = paste(site, Id, local_date, sep = "::"),
      weekday_weekend = factor(
        ifelse(
          as.POSIXlt(local_date)$wday %in% c(0L, 6L),
          "Weekend",
          "Weekday"
        ),
        levels = c("Weekday", "Weekend")
      )
    )
}

daily_candidates <- bind_rows(
  make_daily_candidate(near_day, "Near-eye"),
  make_daily_candidate(chest_day, "Chest")
) |>
  mutate(
    placement = factor(placement, levels = c("Near-eye", "Chest")),
    site = factor(site, levels = site_levels)
  )

base_candidate_sample <- daily_candidates |>
  summarise(
    `Participant-days` = n(),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = placement
  )

stopifnot(
  base_candidate_sample$`Participant-days`[
    base_candidate_sample$placement == "Near-eye"
  ] == 816L,
  base_candidate_sample$`Participant-days`[
    base_candidate_sample$placement == "Chest"
  ] == 902L
)

base_candidate_sample |>
  compact_gt(
    source_note = paste0(
      "These are available participant-day inputs before a metric or predictor ",
      "complete-case rule. They are not fitted-model samples."
    )
  ) |>
  gt::fmt_integer(columns = c(`Participant-days`, Participants, Sites))
```

```{r}
#| label: predictor-coverage

predictor_columns <- c(
  work_free_day = "work_free_day",
  weekday_weekend = "weekday_weekend",
  activity_status = "activity_status",
  exercise_intensity = "exercise_intensity",
  exercise_location = "exercise_location",
  active_commute_h = "active_commute_h",
  sedentary_h = "sedentary_h",
  wore_light_logger_during_exercise = "wore_light_logger_during_exercise",
  previous_sleep_onset_centered_h = "previous_sleep_onset_centered_h",
  wake_centered_h = "wake_centered_h",
  previous_sleep_duration_centered_h = "previous_sleep_duration_centered_h"
)

predictor_coverage <- imap_dfr(
  predictor_columns,
  function(column, measure_id) {
    daily_candidates |>
      filter(!is.na(.data[[column]])) |>
      summarise(
        measure_id = measure_id,
        `Participant-days` = n(),
        Participants = n_distinct(participant_key),
        Sites = n_distinct(site),
        .by = placement
      )
  }
) |>
  left_join(
    predictor_contract |>
      select(measure_id, reader_name, proposed_role),
    by = "measure_id",
    relationship = "many-to-one"
  ) |>
  arrange(match(measure_id, names(predictor_columns)), placement)

predictor_coverage |>
  select(proposed_role, reader_name, placement, `Participant-days`, Participants, Sites) |>
  compact_gt(
    groupname_col = "proposed_role",
    source_note = paste0(
      "Completeness is evaluated on the available placement-specific day ",
      "table before metric-specific missingness."
    )
  ) |>
  gt::fmt_integer(columns = c(`Participant-days`, Participants, Sites)) |>
  gt::cols_label(
    reader_name = "Predictor",
    placement = "Placement"
  )
```

## Category and category-by-site cells

```{r}
#| label: primary-category-cells

category_cell_table <- function(data, column, measure_label, level_labels) {
  data |>
    transmute(
      placement,
      site,
      category = as.character(.data[[column]])
    ) |>
    mutate(category = ifelse(is.na(category), "Missing", category)) |>
    count(placement, site, category, name = "Participant-days") |>
    complete(
      nesting(placement, site),
      category = c(level_labels, "Missing"),
      fill = list(`Participant-days` = 0L)
    ) |>
    mutate(measure = measure_label)
}

primary_category_cells <- bind_rows(
  category_cell_table(
    daily_candidates,
    "work_free_day",
    "Work/free day",
    c("Work day", "Free day")
  ),
  category_cell_table(
    daily_candidates,
    "weekday_weekend",
    "Weekday/weekend",
    c("Weekday", "Weekend")
  ),
  category_cell_table(
    daily_candidates,
    "activity_status",
    "Activity status",
    activity_status_levels
  )
) |>
  left_join(
    site_registry |>
      select(site, display_order, display_name),
    by = "site",
    relationship = "many-to-one"
  ) |>
  arrange(measure, placement, display_order, match(category, c(
    "Work day", "Free day", "Weekday", "Weekend",
    activity_status_levels, "Missing"
  ))) |>
  transmute(
    Measure = measure,
    Placement = placement,
    Site = display_name,
    Category = category,
    `Participant-days`
  )

primary_category_cells |>
  compact_gt(
    groupname_col = "Measure",
    source_note = paste0(
      "Zero and missing cells are shown. Counts precede metric-specific ",
      "outcome missingness; Stage 2 must repeat this table for every fitted frame."
    )
  ) |>
  gt::fmt_integer(columns = `Participant-days`)
```

The binary activity status is fixed before fitting: no reported exercise maps
to sedentary, and light/moderate/vigorous reported exercise map to active.
This is the only primary activity contrast and is not changed after seeing
results. The original four-level source cells remain visible below for
provenance and exploratory interpretation; their sparse moderate/vigorous
site cells do not define the primary model.

```{r}
#| label: additional-exercise-category-cells

additional_exercise_cells <- bind_rows(
  category_cell_table(
    daily_candidates,
    "exercise_intensity",
    "Original exercise intensity (exploratory)",
    exercise_intensity_levels
  ),
  category_cell_table(
    daily_candidates,
    "exercise_location",
    "Exercise location",
    c("Outdoors", "Indoors", "Indoors and outdoors")
  ),
  category_cell_table(
    daily_candidates,
    "wore_light_logger_during_exercise",
    "Logger worn during exercise",
    c("No", "Yes")
  )
) |>
  left_join(
    site_registry |>
      select(site, display_order, display_name),
    by = "site",
    relationship = "many-to-one"
  ) |>
  arrange(measure, placement, display_order, category) |>
  transmute(
    Measure = measure,
    Placement = placement,
    Site = display_name,
    Category = category,
    `Participant-days`
  )

additional_exercise_cells |>
  compact_gt(
    groupname_col = "Measure",
    source_note = paste0(
      "Original intensity is retained for provenance/exploration. Location ",
      "and logger-worn fields are conditionally applicable and do not define ",
      "the same population as the binary primary activity measure."
    )
  ) |>
  gt::fmt_integer(columns = `Participant-days`)
```

## Metric-by-predictor and paired/common candidate samples

The table below is the exact **candidate** sample for the proposed three-core-
measure set under the registered daily option. Each row uses a non-missing
daily metric and one non-missing measure. The proposed reduced, additive, and
interaction models for that row must all use this same ordered frame. Final
fitted counts can differ only for a declared transformation or model-engine
exclusion and must be reported as such under REPORT-005.

```{r}
#| label: metric-predictor-samples

core_predictors <- c(
  work_free_day = "work_free_day",
  activity_status = "activity_status",
  previous_sleep_duration_centered_h = "previous_sleep_duration_centered_h"
)

daily_metric_long <- daily_candidates |>
  pivot_longer(
    cols = all_of(unname(metric_columns)),
    names_to = "source_column",
    values_to = "response_value"
  ) |>
  mutate(
    metric_id = names(metric_columns)[match(source_column, metric_columns)]
  ) |>
  left_join(
    daily_metric_contract |>
      select(metric_id, metric_order, manuscript_name),
    by = "metric_id",
    relationship = "many-to-one"
  )

metric_predictor_frames <- imap_dfr(
  core_predictors,
  function(column, measure_id) {
    daily_metric_long |>
      filter(!is.na(response_value), !is.na(.data[[column]])) |>
      transmute(
        placement,
        site,
        Id,
        local_date,
        participant_key,
        participant_day_key,
        metric_id,
        metric_order,
        manuscript_name,
        measure_id,
        predictor_value = as.character(.data[[column]])
      )
  }
)

all_available_samples <- metric_predictor_frames |>
  summarise(
    `Participant-days` = n(),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = c(measure_id, metric_id, metric_order, manuscript_name, placement)
  )

paired_samples <- metric_predictor_frames |>
  select(
    placement, site, Id, local_date, participant_key,
    metric_id, metric_order, manuscript_name, measure_id
  ) |>
  distinct() |>
  pivot_wider(
    names_from = placement,
    values_from = placement,
    values_fn = length,
    values_fill = 0,
    names_prefix = "has_"
  ) |>
  filter(`has_Near-eye` > 0L, has_Chest > 0L) |>
  summarise(
    `Paired participant-days` = n(),
    `Paired participants` = n_distinct(participant_key),
    `Paired sites` = n_distinct(site),
    .by = c(measure_id, metric_id, metric_order, manuscript_name)
  )

sample_contract <- all_available_samples |>
  pivot_wider(
    names_from = placement,
    values_from = c(`Participant-days`, Participants, Sites),
    names_glue = "{.value} ({placement})"
  ) |>
  left_join(
    paired_samples,
    by = c("measure_id", "metric_id", "metric_order", "manuscript_name"),
    relationship = "one-to-one"
  ) |>
  mutate(
    Measure = recode(
      measure_id,
      work_free_day = "Work/free day",
      activity_status = "Activity status",
      previous_sleep_duration_centered_h = "Previous sleep duration"
    )
  ) |>
  arrange(match(measure_id, names(core_predictors)), metric_order) |>
  select(
    Measure, Metric = manuscript_name,
    `Participant-days (Near-eye)`, `Participants (Near-eye)`, `Sites (Near-eye)`,
    `Participant-days (Chest)`, `Participants (Chest)`, `Sites (Chest)`,
    `Paired participant-days`, `Paired participants`, `Paired sites`
  )

stopifnot(nrow(sample_contract) == 45L)

sample_contract |>
  compact_gt(
    groupname_col = "Measure",
    source_note = paste0(
      "Near-eye is the proposed primary all-available sample. Chest is shown ",
      "for composition; the complementary inferential comparison uses the ",
      "paired/common sample and a separately refitted near-eye model."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::tab_spanner(
    label = "All available near-eye",
    columns = contains("Near-eye")
  ) |>
  gt::tab_spanner(
    label = "All available chest (context)",
    columns = contains("Chest")
  ) |>
  gt::tab_spanner(
    label = "Exact paired/common sample",
    columns = starts_with("Paired")
  )
```

# V0 reconstruction

## Frozen code and output identities

```{r}
#| label: v0-input-output-contract

v0_contract <- tibble::tribble(
  ~role, ~relative_path, ~expected_sha256,
  "Near-eye submitted H06 source", "RQ2.qmd",
  "df80d5d8fba3688ddc5ae8db13ddf32449fc1c34ca587eaa2b22e2057bb4dfe7",
  "Chest submitted H06 source", "RQ2_chest.qmd",
  "572a8a065dab30d1e31c98ad89727dcfddf8c99cc06333aa479284afea015f3d",
  "Near-eye frozen H06 table", "tables/H6.docx",
  "20a259c13c5d0637c11f3522ed6fdca12d18d8f7df27b6ab8b1bf5787c330b80",
  "Chest frozen H06 table", "tables/chest/H6.docx",
  "4db844c889df3ee0d628f2d017a82dacd81c87c2e3e1ac02b6694bb1c8f40893",
  "Near-eye frozen preprocessed light data", "data/preprocessed_glasses_2.RData",
  "0d00bac25d955d45447f74cc9c3ad6a5f5dfa1de1282d8d619c9468f72e0c430",
  "Chest frozen preprocessed light data", "data/preprocessed_chest_2.RData",
  "b289c17c64d1bd24166fa3873da9aa8f22a5f57c793a7282ecb201dc77dcd274"
) |>
  mutate(
    observed_sha256 = map_chr(relative_path, sha256_file),
    verified = observed_sha256 == expected_sha256
  )

stopifnot(all(v0_contract$verified))

v0_contract |>
  select(role, relative_path, observed_sha256, verified) |>
  compact_gt(
    source_note = "These frozen identities define V0; current primary inputs are not substituted into V0."
  ) |>
  gt::cols_label(
    role = "V0 role",
    relative_path = "Path",
    observed_sha256 = "SHA-256",
    verified = "Check"
  ) |>
  gt::fmt(columns = verified, fns = function(x) ifelse(x, "PASS", "FAIL"))
```

The deliberate hourly V0 outcome was implemented by inheriting an H03 object
rather than by an H06-specific producer. The outcome choice is intentional;
the unnecessary cross-hypothesis dependency is a separate implementation
issue. One-minute data were aggregated to hours without the current
30-valid-minute rule; an all-constant hourly day was removed; light-source
states were joined even though H06 did not use them; and the daily diaries
were repeated across every retained hour. V0 then used joint complete cases
for one hourly outcome and three predictors. That gives participant-days with
more retained hours more weight.

## Exact V0 formulas — reconstructed, not refitted

The following evaluated cell prints every V0 comparison formula from
`RQ2.qmd:1866–1870` and its near-literal chest copy. The formulas are objects,
not prose, and no `glmmTMB()` call is made.

```{r}
#| label: v0-formula-contract
#| code-fold: false
#| code-summary: "V0 formula reconstruction (always shown)"

v0_formulas <- list(
  full = stats::as.formula(
    "geo.MEDI ~ site * daytype + site * exercise + site * sleep_duration + (1 | Id)"
  ),
  no_daytype = stats::as.formula(
    "geo.MEDI ~ site * exercise + site * sleep_duration + (1 | Id)"
  ),
  no_exercise = stats::as.formula(
    "geo.MEDI ~ site * daytype + site * sleep_duration + (1 | Id)"
  ),
  no_sleep = stats::as.formula(
    "geo.MEDI ~ site * daytype + site * exercise + (1 | Id)"
  ),
  no_site = stats::as.formula(
    "geo.MEDI ~ daytype + exercise + sleep_duration + (1 | Id)"
  ),
  weekend_full = stats::as.formula(
    "geo.MEDI ~ site * Weekend + site * exercise + site * sleep_duration + (1 | Id)"
  ),
  activity_adjusted_full = stats::as.formula(
    "geo.MEDI ~ site * activity + site * daytype + site * exercise + site * sleep_duration + (1 | Id)"
  ),
  activity_adjusted_no_daytype = stats::as.formula(
    "geo.MEDI ~ site * activity + site * exercise + site * sleep_duration + (1 | Id)"
  ),
  activity_adjusted_no_exercise = stats::as.formula(
    "geo.MEDI ~ site * activity + site * daytype + site * sleep_duration + (1 | Id)"
  ),
  activity_adjusted_no_sleep = stats::as.formula(
    "geo.MEDI ~ site * activity + site * daytype + site * exercise + (1 | Id)"
  )
)

v0_formula_table <- enframe(v0_formulas, name = "formula_id", value = "formula") |>
  mutate(
    formula = map_chr(formula, function(x) paste(deparse(x), collapse = " "))
  )

v0_formula_table |>
  compact_gt(
    source_note = paste0(
      "The first four likelihood-ratio comparisons removed each predictor ",
      "together with all its site interactions. They were not pure main-effect tests."
    )
  ) |>
  gt::cols_label(formula_id = "V0 model", formula = "Evaluated Wilkinson formula")
```

All V0 confirmatory fits used a Tweedie log-link GLMM, maximum likelihood,
sum coding for site in the full/reduced models, and only a participant random
intercept. The no-site model did not use the same site contrast declaration.
There was no participant-day intercept, sequence break, or residual
autocorrelation structure.

## Frozen V0 numerical output

The frozen DOCX files contain the historical numbers below. The render opens
their OOXML with R, checks the key strings, and reports them as submitted
artifacts. It does not claim to reproduce the fitted objects: V0 did not
persist its exact joined model frame or H06 model objects, and its diary loads
were moving package calls.

```{r}
#| label: v0-frozen-output

read_docx_text <- function(path) {
  extraction_dir <- tempfile("h06_docx_")
  dir.create(extraction_dir)
  on.exit(unlink(extraction_dir, recursive = TRUE), add = TRUE)
  utils::unzip(
    file.path(project_root, path),
    files = "word/document.xml",
    exdir = extraction_dir
  )
  document <- xml2::read_xml(
    file.path(extraction_dir, "word/document.xml")
  )
  paste(xml2::xml_text(xml2::xml_find_all(document, "//w:t")), collapse = " ")
}

v0_near_text <- read_docx_text("tables/H6.docx")
v0_chest_text <- read_docx_text("tables/chest/H6.docx")

stopifnot(
  grepl("16'406 participant-hours", v0_near_text, fixed = TRUE),
  grepl("18'124 participant-hours", v0_chest_text, fixed = TRUE),
  grepl("75.79 lx", v0_near_text, fixed = TRUE),
  grepl("96.18 lx", v0_chest_text, fixed = TRUE),
  grepl("0.92×", v0_near_text, fixed = TRUE),
  grepl("0.92×", v0_chest_text, fixed = TRUE)
)

v0_output_summary <- tibble::tribble(
  ~placement, ~participant_hours, ~reference_melEDI_lx,
  ~free_vs_work_ratio, ~light_vs_none_ratio, ~moderate_vs_none_ratio,
  ~vigorous_vs_none_ratio, ~sleep_ratio_per_hour,
  ~conditional_r2, ~marginal_r2, ~participant_random_multiplier,
  "Near-eye", 16406L, 75.79, 0.92, 1.85, 1.74, 1.50, 0.92, 0.42, 0.15, 2.36,
  "Chest", 18124L, 96.18, 0.92, 1.49, 1.61, 1.21, 0.92, 0.37, 0.13, 2.26
)

v0_output_summary |>
  compact_gt(
    source_note = paste0(
      "Historical values transcribed from, and string-checked against, the ",
      "two frozen DOCX outputs. Reference: work day, no exercise, 7.9 h sleep. ",
      "No 95% confidence intervals were retained in the final V0 table."
    )
  ) |>
  gt::fmt_integer(columns = participant_hours, use_seps = TRUE) |>
  gt::fmt_number(
    columns = c(
      reference_melEDI_lx, free_vs_work_ratio, light_vs_none_ratio,
      moderate_vs_none_ratio, vigorous_vs_none_ratio,
      sleep_ratio_per_hour, conditional_r2, marginal_r2,
      participant_random_multiplier
    ),
    decimals = 2
  ) |>
  gt::cols_label(
    placement = "Placement",
    participant_hours = "Participant-hours",
    reference_melEDI_lx = "Reference melEDI (lx)",
    free_vs_work_ratio = "Free/work ratio",
    light_vs_none_ratio = "Light/none",
    moderate_vs_none_ratio = "Moderate/none",
    vigorous_vs_none_ratio = "Vigorous/none",
    sleep_ratio_per_hour = "Per-hour sleep ratio",
    conditional_r2 = "Conditional R²",
    marginal_r2 = "Marginal R²",
    participant_random_multiplier = "Displayed participant multiplier"
  )
```

V0 displayed raw omnibus p-values as if a three-value false-discovery-rate
adjustment had been applied, used shared-control defaults for exercise
contrasts, split site contrasts across hidden groups, and discarded intervals.
The site p-value in each H06 table came from the global `comp_site` last made
in H04 (`IMP-007`), not an H06 comparison. The displayed marginal-R²
subtractions overlap and are not additive “unique contributions.” All V0
claims—including the 8% sleep association, site-specific effects, the
weekday/weekend AIC statement, and activity-adjusted persistence—remain
provisional audit targets.

## Why current rows do not reproduce V0 rows

```{r}
#| label: current-hourly-candidate-sample

near_hour <- read_project_rds(
  "artifacts/06_model_data/base/metrics_glasses_one_hour_context.rds"
)
chest_hour <- read_project_rds(
  "artifacts/06_model_data/base/metrics_chest_one_hour_context.rds"
)

make_hourly_candidate <- function(hour_data, placement) {
  hour_data |>
    filter(bin_admissible, !is.na(zero_aware_geometric_mean_medi_lx)) |>
    select(
      site, Id, local_date, clock_minute,
      response_value = zero_aware_geometric_mean_medi_lx
    ) |>
    left_join(
      sleep_day |>
        select(
          site, Id, local_date, work_free_day,
          previous_sleep_duration_centered_h
        ),
      by = c("site", "Id", "local_date"),
      relationship = "many-to-one"
    ) |>
    left_join(
      exercise_day |>
        select(site, Id, local_date, activity_status, exercise_intensity),
      by = c("site", "Id", "local_date"),
      relationship = "many-to-one"
    ) |>
    filter(
      !is.na(work_free_day),
      !is.na(activity_status),
      !is.na(previous_sleep_duration_centered_h)
    ) |>
    mutate(
      placement = placement,
      participant_key = paste(site, Id, sep = "::"),
      participant_day_key = paste(site, Id, local_date, sep = "::")
    )
}

current_hourly_candidates <- bind_rows(
  make_hourly_candidate(near_hour, "Near-eye"),
  make_hourly_candidate(chest_hour, "Chest")
)

current_hourly_sample <- current_hourly_candidates |>
  summarise(
    `One-hour observations` = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = placement
  ) |>
  left_join(
    v0_output_summary |>
      select(placement, `Frozen V0 participant-hours` = participant_hours),
    by = "placement",
    relationship = "one-to-one"
  ) |>
  mutate(
    `Current minus V0 rows` =
      `One-hour observations` - `Frozen V0 participant-hours`
  )

current_hourly_sample |>
  compact_gt(
    source_note = paste0(
      "The current frame enforces at least 30 valid minutes and uses corrected, ",
      "pinned diaries. Its row difference confirms that a current-data refit ",
      "would be a new analysis, not an exact V0 reproduction."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE)
```

Because the hourly adaptation is the selected H06-G1 estimand, its exact
design cells are not deferred to Stage 2. Counts below are candidate support
only; no model has been fitted.

```{r}
#| label: hourly-candidate-design-cells

hourly_hours_per_day <- current_hourly_candidates |>
  count(placement, participant_day_key, name = "Supported hours") |>
  summarise(
    `Participant-days` = n(),
    `Minimum hours` = min(`Supported hours`),
    `Median hours` = median(`Supported hours`),
    `Maximum hours` = max(`Supported hours`),
    .by = placement
  )

hourly_site_support <- current_hourly_candidates |>
  summarise(
    `One-hour observations` = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    .by = c(placement, site)
  ) |>
  mutate(
    placement = factor(placement, levels = c("Near-eye", "Chest")),
    site = factor(site, levels = site_levels)
  ) |>
  arrange(placement, site)

hourly_daytype_cells <- current_hourly_candidates |>
  summarise(
    `One-hour observations` = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    .by = c(placement, site, work_free_day)
  ) |>
  tidyr::complete(
    nesting(placement, site),
    work_free_day = factor(
      c("Work day", "Free day"),
      levels = c("Work day", "Free day")
    ),
    fill = list(
      `One-hour observations` = 0L,
      `Participant-days` = 0L,
      Participants = 0L
    )
  ) |>
  mutate(
    placement = factor(placement, levels = c("Near-eye", "Chest")),
    site = factor(site, levels = site_levels)
  ) |>
  arrange(placement, site, work_free_day)

hourly_activity_cells <- current_hourly_candidates |>
  summarise(
    `One-hour observations` = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    .by = c(placement, site, activity_status)
  ) |>
  tidyr::complete(
    nesting(placement, site),
    activity_status = factor(
      activity_status_levels,
      levels = activity_status_levels
    ),
    fill = list(
      `One-hour observations` = 0L,
      `Participant-days` = 0L,
      Participants = 0L
    )
  ) |>
  mutate(
    placement = factor(placement, levels = c("Near-eye", "Chest")),
    site = factor(site, levels = site_levels)
  ) |>
  arrange(placement, site, activity_status)

hourly_sleep_support <- current_hourly_candidates |>
  distinct(
    placement, site, participant_day_key,
    previous_sleep_duration_centered_h
  ) |>
  summarise(
    `Participant-days` = n(),
    `Minimum previous sleep (h)` =
      min(previous_sleep_duration_centered_h + 8),
    `Median previous sleep (h)` =
      median(previous_sleep_duration_centered_h + 8),
    `Maximum previous sleep (h)` =
      max(previous_sleep_duration_centered_h + 8),
    .by = c(placement, site)
  ) |>
  mutate(
    placement = factor(placement, levels = c("Near-eye", "Chest")),
    site = factor(site, levels = site_levels)
  ) |>
  arrange(placement, site)

hourly_hours_per_day |>
  compact_gt(
    source_note = paste0(
      "Unsupported hours remain missing and break residual sequences; days ",
      "are not required to contribute equally many hourly rows."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE)

hourly_site_support |>
  compact_gt(
    groupname_col = "placement",
    source_note = "Submitted site names and order follow the shared display registry."
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  gt::cols_label(site = "Site")

hourly_daytype_cells |>
  compact_gt(
    groupname_col = "placement",
    source_note = "Zero cells remain visible; no level is merged after inspection."
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  gt::cols_label(site = "Site", work_free_day = "Day type")

hourly_activity_cells |>
  compact_gt(
    groupname_col = "placement",
    source_note = paste0(
      "The author-approved binary mapping is fixed before fitting. Original ",
      "four-level intensity cells remain available for source audit only."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE) |>
  gt::cols_label(site = "Site", activity_status = "Activity status")

hourly_sleep_support |>
  compact_gt(
    groupname_col = "placement",
    source_note = "Sleep support is one value per participant-day, not per hour."
  ) |>
  gt::fmt_integer(columns = `Participant-days`, use_seps = TRUE) |>
  gt::fmt_number(
    columns = c(
      `Minimum previous sleep (h)`, `Median previous sleep (h)`,
      `Maximum previous sleep (h)`
    ),
    decimals = 2
  ) |>
  gt::cols_label(site = "Site")
```

# H06-G1 author decision

## Option A — registered daily light-exposure metrics (not selected)

For each approved daily metric and measure, estimate the within-study
association between the observed participant-day context and the expected
daily metric while adjusting for site and repeated days within participant.
The main association is tested in the additive model; predictor-by-site
heterogeneity is tested separately. Day type and exercise are observed
contexts, and preceding sleep is an observed diary interval. None is described
as causing light exposure or as a physiological or health outcome.

This option gives each participant-day one model row, matches the signed unit,
and avoids weighting a day by its count of supported hours. It expands the
outcome family to 15 daily metrics and therefore requires an exact
measure-by-metric registry before fitting. The author did not select this
question for Stage 2. Its formulas and candidate samples remain visible as the
registered reference, not as a second primary analysis.

## Option B — deliberate hourly adaptation (**estimand retained; architecture revised**)

Use the current one-hour zero-aware geometric mean melEDI with at least 30
valid minutes in a population-mean log-link GLM. Participant-cluster sandwich
covariance treats participants, rather than repeated hours, as inferential
replicates and allows arbitrary dependence across their retained hours and
days. The confirmatory model does not adjust for local clock because the
author wants the average association over retained supported hours; the
exploratory GAM/GAMM separately addresses time-of-day patterning. This carries
forward the deliberate DEV-015 choice and its parsimony and interpretability
rationale. It is a formally adapted hypothesis about repeated hourly
exposure, not an error correction back to the registered daily-metric set. It
does not test the registered daily-metric estimand and cannot be called
“daily exposure” in results.

The current three-predictor complete-case candidates contain 16,596 near-eye
and 18,352 chest one-hour observations, not V0's 16,406 and 18,124. The
already-audited H06 true-time sequence fields remain necessary for residual
diagnostics and the exploratory GAM/GAMM; the marginal primary fit itself
contains no AR term.

The alternatives remain scientifically distinct. Option A asks how
observed day-level contexts relate to a family of daily exposure summaries.
Option B asks how those contexts relate to the expected hourly exposure level
across retained supported hours. The approved Option B preserves the
deliberate V0 question but gives days with more supported hours more
model-row weight and does not standardize the predictor effects to a common
clock-hour distribution. Option A gives each metric-specific participant-day
one row but creates a larger, heterogeneous outcome family.

# Proposed revised formula and comparison contracts — not fitted in this render

`response_value` below means the supported one-hour zero-aware geometric mean
melEDI in the proposed Option B formulas, or a transformed daily metric only
in the visible unselected registered-reference formulas.
Formula construction is evaluated in visible cells so that each proposed or
comparison model has an exact Wilkinson formula. No fitting function is
called.

## Registered daily option: core measures

```{r}
#| label: daily-core-formula-contract
#| code-fold: false
#| code-summary: "Daily core formula contract (always shown)"

formula_text <- function(x) paste(deparse(x), collapse = " ")

make_daily_formula_set <- function(measure) {
  list(
    registered_exact = stats::as.formula(paste0(
      "response_value ~ ", measure,
      " + (", measure, " | site) + (1 | site:Id)"
    )),
    fixed_site_reduced = stats::as.formula(
      "response_value ~ site + (1 | participant_key)"
    ),
    fixed_site_additive = stats::as.formula(paste0(
      "response_value ~ site + ", measure,
      " + (1 | participant_key)"
    )),
    fixed_site_interaction = stats::as.formula(paste0(
      "response_value ~ site * ", measure,
      " + (1 | participant_key)"
    )),
    daily_ar1_reduced = stats::as.formula(
      paste0(
        "response_value ~ site + (1 | participant_key) + ",
        "ar1(day_index_factor + 0 | day_sequence_id)"
      )
    ),
    daily_ar1_additive = stats::as.formula(paste0(
      "response_value ~ site + ", measure,
      " + (1 | participant_key) + ",
      "ar1(day_index_factor + 0 | day_sequence_id)"
    )),
    daily_ar1_interaction = stats::as.formula(paste0(
      "response_value ~ site * ", measure,
      " + (1 | participant_key) + ",
      "ar1(day_index_factor + 0 | day_sequence_id)"
    ))
  )
}

daily_core_formulas <- imap_dfr(
  core_predictors,
  function(measure, measure_id) {
    enframe(
      make_daily_formula_set(measure),
      name = "model_id",
      value = "formula"
    ) |>
      mutate(
        measure_id = measure_id,
        formula = map_chr(formula, formula_text),
        .before = 1
      )
  }
)

daily_core_formulas |>
  compact_gt(
    groupname_col = "measure_id",
    source_note = paste0(
      "Main association: fixed_site_reduced versus fixed_site_additive. ",
      "Heterogeneity: fixed_site_additive versus fixed_site_interaction. ",
      "The AR(1) rows are a predeclared temporal-dependence counterpart using ",
      "actual consecutive-day sequences; gaps start a new sequence."
    )
  ) |>
  gt::cols_label(
    model_id = "Model",
    formula = "Evaluated Wilkinson formula"
  )
```

The signed `registered_exact` model is retained as a faithful comparison. The
author-approved fixed-site hierarchy applies to the selected hourly
implementation because it defines explicit adjusted and site-heterogeneity
estimands. The registered random-site form is a registered-reference model. These
daily formulas remain visible only to document the unselected registered
question; the random-site form is not a sensitivity of the revised hourly
marginal model and does not define a Stage 2 daily-metric family.

Daily residual dependence is checked against actual local-date lags within
participant. The participant random intercept is always retained. If the
predeclared autocorrelation check is material, the metric-compatible AR(1)
counterpart becomes the stability model; a missing day breaks a sequence and
is not compressed. A result is not accepted merely because the simpler model
converges.

## Registered daily option: nonprimary named measures

The following exact formulas document the unselected registered daily scope.
They are not an inferential Stage 2 family. Under the selected hourly option,
weekday/weekend is a named sensitivity; location, active travel, the
author-reinterpreted sedentary time, sleep onset, and final wake may be used
only in explicitly labelled exploratory work. Logger-worn status remains a
measurement-support sensitivity.

```{r}
#| label: daily-additional-formula-contract
#| code-fold: false
#| code-summary: "Additional measure formula contract (always shown)"

additional_measure_columns <- c(
  weekday_weekend = "weekday_weekend",
  exercise_location = "exercise_location",
  active_commute_h = "active_commute_h",
  sedentary_h = "sedentary_h",
  previous_sleep_onset_centered_h = "previous_sleep_onset_centered_h",
  wake_centered_h = "wake_centered_h"
)

make_fixed_daily_formula_set <- function(measure) {
  list(
    reduced = stats::as.formula(
      "response_value ~ site + (1 | participant_key)"
    ),
    additive = stats::as.formula(paste0(
      "response_value ~ site + ", measure,
      " + (1 | participant_key)"
    )),
    interaction = stats::as.formula(paste0(
      "response_value ~ site * ", measure,
      " + (1 | participant_key)"
    ))
  )
}

daily_additional_formulas <- imap_dfr(
  additional_measure_columns,
  function(measure, measure_id) {
    enframe(
      make_fixed_daily_formula_set(measure),
      name = "model_id",
      value = "formula"
    ) |>
      mutate(
        measure_id = measure_id,
        formula = map_chr(formula, formula_text),
        .before = 1
      )
  }
)

daily_additional_formulas |>
  compact_gt(
    groupname_col = "measure_id",
    source_note = paste0(
      "Each comparison uses its own exact metric × measure frame. ",
      "Location is conditional on recorded exercise; active travel lacks MPI."
    )
  ) |>
  gt::cols_label(model_id = "Model", formula = "Evaluated Wilkinson formula")
```

## Deliberate hourly option: H03-informed population-mean hierarchy

For Option B, every supported participant-hour remains a model row, but the
participant is the inferential replication unit. Participant-cluster sandwich
covariance encompasses all retained hours and days from that participant; a
participant-day random intercept and within-day AR term are therefore not
added to the marginal primary model. `hour_sequence_id` and `AR_start` remain
essential for honest residual-lag diagnostics and the exploratory temporal
model. The confirmatory formulas deliberately contain no local-clock term:
their coefficients are averages over the supported hours in the fitted frame,
so participant-days with more retained hours contribute more model rows.
Stage 2 must display that composition and cannot describe the effects as
clock-standardized.

```{r}
#| label: hourly-formula-contract
#| code-fold: false
#| code-summary: "Hourly formula contract (always shown)"

hourly_formulas <- list(
  primary_additive = stats::as.formula(paste(
    "response_value ~ site + work_free_day + activity_status +",
    "previous_sleep_duration_centered_h"
  )),
  heterogeneity_full = stats::as.formula(paste(
    "response_value ~ site * work_free_day + site * activity_status +",
    "site * previous_sleep_duration_centered_h"
  )),
  weekend_additive = stats::as.formula(paste(
    "response_value ~ site + weekday_weekend + activity_status +",
    "previous_sleep_duration_centered_h"
  )),
  weekend_heterogeneity = stats::as.formula(paste(
    "response_value ~ site * weekday_weekend + site * activity_status +",
    "site * previous_sleep_duration_centered_h"
  ))
)

hourly_formula_table <- enframe(
  hourly_formulas,
  name = "model_id",
  value = "formula"
) |>
  mutate(formula = map_chr(formula, formula_text))

hourly_formula_table |>
  compact_gt(
    source_note = paste0(
      "The three main tests are coefficient-block Wald-F tests in the additive ",
      "model. The three heterogeneity tests are separate site-interaction-block ",
      "Wald-F tests in the full model. Weekend rows are estimate/CI stability ",
      "comparisons without a new significance family. All primary rows use one ",
      "frozen joint complete-case frame."
    )
  ) |>
  gt::cols_label(model_id = "Model", formula = "Evaluated Wilkinson formula")
```

```{r}
#| label: robust-mean-execution-contract
#| code-fold: false
#| code-summary: "Robust mean-model execution contract (always shown)"

robust_mean_contract <- tibble::tribble(
  ~component, ~approved_proposal,
  "Primary family and link",
  "stats::quasipoisson(link = 'log'); zero-inclusive expected hourly melEDI",
  "Primary covariance",
  "sandwich::vcovCL clustered by participant_key; HC3; cadjust = TRUE; fix = FALSE",
  "Finite-cluster reference",
  "one coefficient: t with G - 1 df; q restrictions: Wald F with q and G - 1 df",
  "Site coding and overall estimand",
  "sum-coded fixed site; equal-site linear-predictor average, then one exponential back-transformation",
  "Working-variance sensitivity",
  "same formulas and rows with statmod::tweedie(var.power = 1.8, link.power = 0)",
  "Covariance sensitivities",
  "HC1 and HC2; no result is rescued by eigenvalue repair",
  "Unavailable summaries",
  "no likelihood-ratio test, AIC, ICC, random-effect SD, mixed-model R-squared, or likelihood-based variance decomposition",
  "Temporal limitation",
  "residual lag by true-time sequence is reported; no confirmatory clock or AR term is added"
)

robust_mean_contract |>
  compact_gt(
    source_note = paste0(
      "The quasi family is a working mean-variance model, not a calibrated ",
      "generative model for exact zeros or within-participant trajectories."
    )
  ) |>
  gt::cols_label(
    component = "Contract component",
    approved_proposal = "Revised H06 proposal"
  )
```

The primary additive model supplies H06-F1 and the reader-facing average
effects. The full interaction model supplies H06-F2 and site-specific
descriptive estimates. They answer separate questions and are never combined
in one test. Quasi-likelihood model comparisons are not used. Each omnibus is
an explicit robust linear-hypothesis test whose restriction matrix, rank, and
covariance are stored. A non-positive-definite consequential covariance is
reported as non-estimable rather than repaired with `fix = TRUE`.

### Bounded H03-route transfer pilot

The pilot used the exact frozen H06 candidate rows and the full
V0-recognizable fixed scaffold. It inspected construction diagnostics only.
The primary additive formula was also fitted as a separate numerical gate.

```{r}
#| label: h03-route-pilot-summary
#| tbl-cap: "Bounded H03-route transfer pilot; numerical diagnostics only."

h03_route_summary |>
  transmute(
    Placement = recode(placement, glasses = "Near-eye", chest = "Chest"),
    Hours = observations,
    `Participant-days` = participant_days,
    Participants = participant_clusters,
    `Additive rank` = paste0(
      additive_design_rank, "/", additive_design_columns
    ),
    `Full rank` = full_rank,
    `HC3 covariance condition number` =
      hc3_covariance_condition_number,
    `Maximum cluster-score share` = maximum_cluster_score_share,
    `Maximum cluster-leverage share` = maximum_cluster_leverage_share,
    `Pearson residual lag 1` = pearson_lag1_correlation,
    `Maximum targeted-deletion shift (HC3 SE)` =
      maximum_abs_core_shift_full_hc3_se,
    Gate = numerical_gate_pass & additive_numerical_gate_pass
  ) |>
  compact_gt(
    source_note = paste0(
      "No effect estimate, interval, omnibus statistic, p-value, or claim was ",
      "used to select the route. Deletion shifts use the five participants ",
      "with the largest cluster-score shares; nobody was removed."
    )
  ) |>
  gt::fmt_integer(columns = c(Hours, `Participant-days`, Participants)) |>
  gt::fmt_number(
    columns = c(
      `HC3 covariance condition number`, `Maximum cluster-score share`,
      `Maximum cluster-leverage share`, `Pearson residual lag 1`,
      `Maximum targeted-deletion shift (HC3 SE)`
    ),
    decimals = 3
  ) |>
  gt::fmt(columns = Gate, fns = function(x) ifelse(x, "PASS", "FAIL"))
```

Both primary additive designs are full rank (12/12 near eye and 11/11 chest),
as are both full interaction designs (36/36 and 32/32). The unmodified HC3
covariances are finite and positive definite. Maximum participant
cluster-score shares are
`r sprintf("%.3f", h03_route_summary$maximum_cluster_score_share[h03_route_summary$placement == "glasses"])`
near eye and
`r sprintf("%.3f", h03_route_summary$maximum_cluster_score_share[h03_route_summary$placement == "chest"])`
at chest, below the predeclared 0.50 construction threshold but material near
eye. Residual lag 1 remains
`r sprintf("%.3f", h03_route_summary$pearson_lag1_correlation[h03_route_summary$placement == "glasses"])`
and
`r sprintf("%.3f", h03_route_summary$pearson_lag1_correlation[h03_route_summary$placement == "chest"])`;
participant clustering permits this dependence for covariance estimation but
does not remove it.

```{r}
#| label: h03-route-working-power-screen
#| tbl-cap: "Effect-blind working-variance screen."

h03_route_diagnostics |>
  transmute(
    Placement = recode(placement, glasses = "Near-eye", chest = "Chest"),
    Route = recode(
      route_id,
      quasi_poisson_p1 = "Quasi-Poisson, p = 1 (proposed primary)",
      h03_literal_p1_539919 = "Quasi-Tweedie, p = 1.539919 (H03 comparator)",
      h06_stable_p1_8 = "Quasi-Tweedie, p = 1.8 (sensitivity)",
      h06_v0_boundary_p1_9 = "Quasi-Tweedie, p = 1.9 (boundary check)"
    ),
    Iterations = iterations,
    Converged = converged,
    `Full rank` = full_rank,
    `HC3 covariance positive definite` =
      hc3_covariance_positive_definite,
    Warning = fit_warnings,
    Error = fit_error
  ) |>
  compact_gt(
    source_note = paste0(
      "The H03 numerical power is not imported as an H06 parameter. The ",
      "p = 1.9 near-eye fit failed with step truncation and NA/NaN/Inf in the ",
      "working system; p = 1.8 is the predeclared fixed-power sensitivity."
    ),
    groupname_col = "Placement"
  ) |>
  gt::fmt_integer(columns = Iterations)
```

Across the successful working-variance comparators, the largest shift in a
core or site-interaction coefficient relative to the proposed primary is
`r sprintf("%.2f", max(h03_route_power_sensitivity$maximum_abs_core_and_site_interaction_shift_selected_hc3_se, na.rm = TRUE))`
primary-model HC3 standard errors. This is small enough to retain the route,
but large enough that the `p = 1.8` sensitivity must be reported rather than
assumed equivalent. The result remains subject to the complete Stage 2
calibration, residual, influence, site-deletion, dataset, placement, and
multiplicity contract.

### Additional hourly analyses approved as exploratory only

The author approved the other named day-type, exercise, and sleep variables
for exploration, not for the primary model. The formulas below are therefore
one-at-a-time descriptive association models on their own exact samples. They
receive estimates and 95% CIs but no confirmatory p-value family. Exercise
location is conditional on a reported active day, so the binary activity term
is omitted from that formula because it is constant in that frame. The
four-level exercise-intensity field is not fitted: it is collapsed to the
binary status in all planned models.

```{r}
#| label: hourly-exploratory-predictor-formula-contract
#| code-fold: false
#| code-summary: "Exploratory predictor formula contract (always shown)"

hourly_exploratory_formulas <- list(
  exercise_location_active_days = stats::as.formula(paste(
    "response_value ~ site + work_free_day + exercise_location +",
    "previous_sleep_duration_centered_h"
  )),
  active_travel = stats::as.formula(paste(
    "response_value ~ site + work_free_day + activity_status +",
    "previous_sleep_duration_centered_h + active_commute_h"
  )),
  sedentary_time = stats::as.formula(paste(
    "response_value ~ site + work_free_day + activity_status +",
    "previous_sleep_duration_centered_h + sedentary_h"
  )),
  previous_sleep_onset = stats::as.formula(paste(
    "response_value ~ site + work_free_day + activity_status +",
    "previous_sleep_duration_centered_h +",
    "previous_sleep_onset_centered_h"
  )),
  final_wake = stats::as.formula(paste(
    "response_value ~ site + work_free_day + activity_status +",
    "previous_sleep_duration_centered_h + wake_centered_h"
  ))
)

hourly_exploratory_formula_table <- enframe(
  hourly_exploratory_formulas,
  name = "analysis_id",
  value = "formula"
) |>
  mutate(formula = map_chr(formula, formula_text))

hourly_exploratory_formula_table |>
  compact_gt(
    source_note = paste0(
      "Exploratory only; each row uses its own declared complete-case sample. ",
      "These use the same quasi-Poisson log-mean and participant-cluster HC3 ",
      "architecture as the proposed primary, without confirmatory p-values. ",
      "The KNUST_S005 sedentary source numeral is provenance-labelled and ",
      "interpreted only here as 3,600 seconds = 1 h."
    )
  ) |>
  gt::cols_label(
    analysis_id = "Exploratory analysis",
    formula = "Evaluated Wilkinson formula"
  )
```

## Exploratory time-of-day GAM/GAMM

The submitted exploratory work is retained only because it can contextualize
when work/free and exercise differences occur. It is separate from
preregistered inference. V0 used hourly arithmetic melEDI, a noncyclic time
smooth, one AR start per participant, pointwise red significance segments,
and variances of correlated smooth terms called “contributions.” None of
those choices is carried forward as confirmatory evidence.

```{r}
#| label: exploratory-gamm-formula-contract
#| code-fold: false
#| code-summary: "Exploratory GAMM formula contract (always shown)"

exploratory_formulas <- list(
  v0_gam = stats::as.formula(paste(
    "lzMEDI ~ s(Time, k = 12) + s(Time, daytype, bs = 'sz', k = 12) +",
    "s(Time, exercise, bs = 'sz', k = 12) +",
    "s(Time, site, bs = 'sz', k = 12) +",
    "s(Time, photoperiod.state, bs = 'sz', k = 12) +",
    "s(Time, Id, bs = 'fs') + s(Id_date, bs = 're')"
  )),
  selected_two_part_occurrence = stats::as.formula(
    two_part_formula_registry$selected_formula[
      two_part_formula_registry$component == "occurrence"
    ]
  ),
  selected_two_part_positive_magnitude = stats::as.formula(
    two_part_formula_registry$selected_formula[
      two_part_formula_registry$component == "positive_magnitude"
    ]
  )
)

gamm_execution_contract <- tibble::tribble(
  ~component, ~family_link, ~engine, ~method, ~basis_k, ~discrete,
  ~sequence_rule, ~prohibition,
  "Positive occurrence", "Binomial, logit", "mgcv::bam", "fREML", 16L, TRUE,
  paste0(
    "All hours; AR.start at every participant-day and unsupported-hour gap; ",
    "rho uses true-time order"
  ),
  "No step may require discrete = FALSE; no confirmatory or whole-curve test",
  "Conditional positive magnitude", "Gamma, log", "mgcv::bam", "fREML", 16L, TRUE,
  paste0(
    "Positive hours only; rebuild contiguous sequences after removing zeros; ",
    "AR never bridges an excluded zero hour"
  ),
  "No step may require discrete = FALSE; no confirmatory or whole-curve test"
)

enframe(exploratory_formulas, name = "model_id", value = "formula") |>
  mutate(formula = map_chr(formula, formula_text)) |>
  compact_gt(
    source_note = paste0(
      "V0 is shown only as historical reconstruction. The selected exploratory ",
      "branch separates exact-zero occurrence from positive magnitude, uses ",
      "cyclic overall and factor-deviation smooths, and remains distinct from ",
      "the proposed confirmatory population-mean model."
    )
  ) |>
  gt::cols_label(model_id = "Exploratory model", formula = "Evaluated formula")

gamm_execution_contract |>
  compact_gt(
    source_note = paste0(
      "The H11 computation lesson is a hard H06 boundary: Stage 2 may run the ",
      "exploratory GAM/GAMM only with this discrete-compatible workflow. If a ",
      "desired calculation cannot honor it, that calculation is omitted or ",
      "returned to the author as a separate method decision."
    )
  ) |>
  gt::cols_label(
    engine = "Engine",
    component = "Component",
    family_link = "Family/link",
    method = "Method",
    basis_k = "Basis k",
    discrete = "Discrete approximation",
    sequence_rule = "Temporal rule",
    prohibition = "Prohibited computation"
  )
```

The earlier offset-log Gaussian exploratory attempt is retained only in the
Stage 2 audit history because its exact zeros created an artificial response
floor and its cyclic closure was inadequate. The two-part branch is the only
H06 analysis that addresses time of day. It cannot be used to choose the
primary day-type definition or predictor family, cannot establish a causal
mechanism or “factors beyond activity,” and supplies no confirmatory p-value
family.

# Multiplicity and reporting contract

## Proposed complete families

```{r}
#| label: multiplicity-registry

multiplicity_registry <- tibble::tribble(
  ~family_id, ~scenario, ~content, ~planned_size,
  ~adjustment, ~proposal,
  "H06-F1-main",
  "Primary dataset, near-eye",
  "One hourly outcome × 3 core predictor omnibus main-association tests",
  3L, "BH over one complete vector", "Primary confirmatory family",
  "H06-F2-heterogeneity",
  "Primary dataset, near-eye",
  "One hourly outcome × 3 predictor-by-site omnibus tests",
  3L, "BH over one complete vector", "Separate heterogeneity family",
  "H06-F3-practical-contrasts",
  "Primary dataset, near-eye",
  "1 work/free + 1 active/sedentary + 1 sleep-duration contrast/slope",
  3L, "BH over one complete vector", "Prespecified reader-facing contrasts",
  "H06-F1-gap",
  "Gap-timing-unaware dataset, near-eye",
  "Same one hourly outcome × 3 core main tests",
  3L, "BH within this scenario", "Required data sensitivity",
  "H06-F2-gap",
  "Gap-timing-unaware dataset, near-eye",
  "Same one hourly outcome × 3 heterogeneity tests",
  3L, "BH within this scenario", "Required data sensitivity"
)

multiplicity_registry |>
  compact_gt(
    source_note = paste0(
      "These are the complete active families for the author-selected hourly ",
      "estimand. A planned non-estimable member remains with NA and a reason; ",
      "families are not shortened after result inspection. Paired/common ",
      "placement and other sensitivity estimates have 95% CIs but no ",
      "additional significance screen."
    )
  ) |>
  gt::fmt_integer(columns = planned_size) |>
  gt::cols_label(
    family_id = "Family",
    scenario = "Dataset/placement",
    content = "Complete vector",
    planned_size = "Planned size",
    adjustment = "Adjustment",
    proposal = "Role"
  )
```

The deliberate hourly adaptation has three main tests because it has one
outcome and three core predictors. Calling `p.adjust()` on one value, or
labelling the three raw V0 p-values as adjusted, is prohibited. Binary
activity creates one reader-facing active-versus-sedentary contrast, so
H06-F3 also contains three values. Weekday/weekend, working-variance and
covariance alternatives, paired placement, and the five
additional named predictor explorations are stability or exploratory outputs
with estimates and 95% CIs only; they do not expand a confirmatory family.
The unselected daily-metric estimand has no active Stage 2 family.

Site-specific estimates and 95% CIs are scientifically useful even when
their cells are sparse. The approved display does not attach a second
site-specific significance screen, and H06-F4 is therefore not activated.
This avoids selecting and adjusting only visually interesting sites.

## P-values, intervals, scales, and samples

Stage 2 must retain raw p-values at full precision and display them under
REPORT-008: leading zero, exactly three decimals, `<0.001` only when the value
is strictly smaller than 0.001, and bold only under the explicitly labelled
raw or adjusted significance rule. Every table retains family ID, planned
size, available rank, adjustment method, raw p-value, adjusted p-value, and
non-estimability reason.

All consequential effects, marginal means, contrasts, ratios, slopes, model
performance summaries, and placement comparisons receive 95% confidence
intervals where estimable. The model-scale covariance interval and its
response-scale transformation are reported together. A failed interval is
shown as unavailable with its reason; a standard error or p-value is not a
replacement.

Every fitted result reports its exact participants, participant-days, sites,
and—only for an hourly option—one-hour observations. Categorical analyses
also report every category and category-by-site cell. Derivation-support
hours under the daily option are labelled as support for the daily metric,
not model observations. Under the hourly option they are repeated model rows,
never independent observations. Full and reduced models must use identical
rows in identical order, verified by a stored frame hash.

# Required data sensitivity and placement comparison

## Candidate data-sensitivity samples

```{r}
#| label: gap-dataset-candidate-samples

gap_day_raw <- read_project_rds(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds"
)

gap_candidates <- gap_day_raw |>
  filter(metric_id %in% daily_metric_contract$metric_id) |>
  left_join(
    sleep_day |>
      select(
        site, Id, local_date, work_free_day,
        previous_sleep_duration_centered_h
      ),
    by = c("site", "Id", "local_date"),
    relationship = "many-to-one"
  ) |>
  left_join(
    exercise_day |>
      select(site, Id, local_date, activity_status, exercise_intensity),
    by = c("site", "Id", "local_date"),
    relationship = "many-to-one"
  ) |>
  mutate(
    placement = recode(
      position_role,
      near_eye_primary = "Near-eye",
      complementary_chest = "Chest"
    ),
    participant_key = paste(site, Id, sep = "::")
  )

gap_base_sample <- gap_candidates |>
  distinct(placement, site, Id, local_date, participant_key) |>
  summarise(
    `Participant-days` = n(),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = placement
  )

gap_core_coverage <- imap_dfr(
  core_predictors,
  function(column, measure_id) {
    gap_candidates |>
      filter(
        !is.na(manuscript_prepared_value),
        !is.na(.data[[column]])
      ) |>
      summarise(
        measure_id = measure_id,
        `Metric × participant-day rows` = n(),
        `Distinct participant-days` = n_distinct(paste(site, Id, local_date)),
        Participants = n_distinct(participant_key),
        Sites = n_distinct(site),
        .by = placement
      )
  }
) |>
  mutate(
    Measure = recode(
      measure_id,
      work_free_day = "Work/free day",
      activity_status = "Activity status",
      previous_sleep_duration_centered_h = "Previous sleep duration"
    )
  )

gap_base_sample |>
  compact_gt(
    source_note = paste0(
      "The retained comparison contains 811 near-eye and 897 chest ",
      "participant-days before metric- and predictor-specific selection."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric))

gap_core_coverage |>
  select(Measure, placement, everything(), -measure_id) |>
  compact_gt(
    groupname_col = "Measure",
    source_note = paste0(
      "There are 15 metric rows per fully complete participant-day. ",
      "Stage 2 will export the exact metric-specific sample and common-sample ",
      "difference for every fitted comparison."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::cols_label(placement = "Placement")
```

The hourly option uses the one-hour comparison artifact rather than deriving
hours from the participant-day table:

```{r}
#| label: gap-hourly-candidate-samples

gap_hour_raw <- read_project_rds(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/one_hour_data.rds"
)

gap_hour_candidates <- gap_hour_raw |>
  transmute(
    placement = recode(
      position_role,
      near_eye_primary = "Near-eye",
      complementary_chest = "Chest"
    ),
    site,
    Id,
    local_date,
    local_clock_label,
    response_value = medi_geometric_mean_lx
  ) |>
  left_join(
    sleep_day |>
      select(
        site, Id, local_date, work_free_day,
        previous_sleep_duration_centered_h
      ),
    by = c("site", "Id", "local_date"),
    relationship = "many-to-one"
  ) |>
  left_join(
    exercise_day |>
      select(site, Id, local_date, activity_status),
    by = c("site", "Id", "local_date"),
    relationship = "many-to-one"
  ) |>
  mutate(
    participant_key = paste(site, Id, sep = "::"),
    participant_day_key = paste(site, Id, local_date, sep = "::")
  ) |>
  filter(
    !is.na(response_value),
    !is.na(work_free_day),
    !is.na(activity_status),
    !is.na(previous_sleep_duration_centered_h)
  )

gap_hour_sample <- gap_hour_candidates |>
  summarise(
    `One-hour observations` = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = placement
  )

gap_hour_hours_per_day <- gap_hour_candidates |>
  count(placement, participant_day_key, name = "Supported hours") |>
  summarise(
    `Participant-days` = n(),
    `Minimum hours` = min(`Supported hours`),
    `Median hours` = median(`Supported hours`),
    `Maximum hours` = max(`Supported hours`),
    .by = placement
  )

gap_hour_category_cells <- bind_rows(
  gap_hour_candidates |>
    mutate(level = as.character(work_free_day)) |>
    summarise(
      `One-hour observations` = n(),
      `Participant-days` = n_distinct(participant_day_key),
      Participants = n_distinct(participant_key),
      .by = c(placement, level)
    ) |>
    mutate(Measure = "Work/free day", .before = level),
  gap_hour_candidates |>
    mutate(level = as.character(activity_status)) |>
    summarise(
      `One-hour observations` = n(),
      `Participant-days` = n_distinct(participant_day_key),
      Participants = n_distinct(participant_key),
      .by = c(placement, level)
    ) |>
    mutate(Measure = "Activity status", .before = level)
)

gap_hour_pairable_days <- gap_hour_candidates |>
  distinct(placement, participant_day_key) |>
  count(participant_day_key, name = "placements") |>
  filter(placements == 2L) |>
  pull(participant_day_key)

gap_hour_paired_sample <- gap_hour_candidates |>
  filter(participant_day_key %in% gap_hour_pairable_days) |>
  summarise(
    `One-hour observations` = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = placement
  )

gap_hour_sample |>
  compact_gt(
    source_note = paste0(
      "Exact three-core-predictor complete-case candidates for the deliberate ",
      "hourly option under the gap-timing-unaware dataset."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE)

gap_hour_hours_per_day |>
  compact_gt(
    source_note = "Remaining supported hours differ by participant-day."
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE)

gap_hour_category_cells |>
  compact_gt(
    groupname_col = "Measure",
    source_note = paste0(
      "Category support is shown before any fit; Stage 2 also exports every ",
      "category-by-site cell."
    )
  ) |>
  gt::fmt_integer(
    columns = c(`One-hour observations`, `Participant-days`, Participants),
    use_seps = TRUE
  ) |>
  gt::cols_label(level = "Level", placement = "Placement")

gap_hour_paired_sample |>
  compact_gt(
    source_note = paste0(
      "Paired/common support requires the same site × participant × date at ",
      "both placements; placement models remain separate."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric), use_seps = TRUE)
```

Within the H06-G1-selected estimand, the same formulas, transformations,
contrasts, multiplicity definitions, confidence-interval methods, placement
roles, and failure rules apply to both datasets. Only the prepared
measurement data differ. If a comparison model cannot be fitted, it remains
non-estimable; it is not simplified only in that dataset. The verified TUM
S001 correction is already present in the shared normalized diary used by the
comparison adapter; no remote fetch or local rewrite is permitted.

## Near-eye and chest

Near-eye is primary. Chest is complementary and is never pooled with
near-eye as if it supplied independent observations. For each selected
response-and-measure combination—one hourly outcome under Option B or each
daily metric under Option A:

- fit the all-available near-eye model;
- identify the exact `site × participant × local date` intersection with an
  admissible value at both placements;
- refit near-eye and chest separately on that paired/common sample with the
  same response definition, formula, reference levels, transformation, and
  contrasts; and
- report the paired counts, component estimates, and 95% CIs.

Stage 3 will assess whether an H05-style near-eye-versus-chest estimate plot
is valid under REPORT-009. It will use near-eye on the horizontal axis, chest
on the vertical axis, an identity line, null lines, equal geometry, and source
CSV data only when estimands and scales match. Closeness is concordance, not
equivalence. Sleep-period light outcomes are explicitly bedside
sleep-environment measurements at both placements, not worn ocular exposure.

# Diagnostics and failure rules for Stage 2

## Data and design integrity

- Reverify all input hashes, R/package versions, the S001 correction, and zero
  S101 rows before constructing H06 artifacts.
- Persist one predictor row per `site × participant × local date`; fail on a
  duplicate, missing key, row multiplication, or unrecorded recode.
- Retain the wake-date/previous-night audit, local-wall clock fields, true
  elapsed sleep duration, and the excluded free-text record.
- Export exact missingness, participants, days, metric support, category cells,
  and category-by-site cells before fitting.
- Verify design-matrix rank, factor references, sum-to-zero site contrasts,
  hierarchy, and full/reduced frame identity for every comparison.
- Report collinearity among the approved predictors and site. Do not add
  temperature, season, or an unregistered behavioural predictor to repair a
  collinearity problem.

## Convergence and distribution

- Record the GLM family/link, iterations, convergence, warnings, design rank,
  aliased and finite coefficients, fitted-value range, dispersion, and every
  failed refit.
- For HC3, HC1, and HC2, record the unmodified participant-cluster covariance,
  eigenvalues, rank, condition number, finite-cluster degrees of freedom,
  cluster-score concentration, and cluster leverage. A covariance that is not
  positive definite with `fix = FALSE` is non-estimable.
- Check mean-dependent residual spread, tails, fitted-decile calibration,
  exact-zero support, physical predictions, and influential units. The quasi
  family does not imply a reportable zero probability or calibrated outcome
  distribution; likelihood diagnostics and simulation-based generative claims
  are unavailable.
- Fit the fixed-power `p = 1.8` quasi-Tweedie sensitivity on identical rows and
  report coefficient/interval stability without treating its working variance
  as a second outcome model. The failed `p = 1.9` route is not retried or
  silently truncated.
- For timing outcomes, verify that the approved unwrapped range is continuous.
  If support crosses its boundary incompatibly, stop that metric rather than
  forcing a linear result.
- Retain the mandatory exact-identifiable-only sensitivity for the longest
  continuous period above 250 lx melEDI and use **period**, never “bout,” in
  visible text.

## Residual, temporal, and influence diagnostics

- Plot and summarize participant-level residuals and actual-date lag residual
  correlations for daily models; missing dates are gaps, not adjacent lags.
- For an hourly option, inspect supported hours per day, true-time sequence
  starts, and residual ACF by sequence. Participant-cluster covariance permits
  arbitrary within-participant dependence for uncertainty but does not whiten
  these residuals; correlation never carries across a participant-day or
  missing-hour gap in the diagnostic calculation.
- Display the retained local-clock composition by predictor and site, and plot
  residuals against local clock as an estimand/adequacy diagnostic. Because
  the author approved an observed-supported-hour-weighted average, do not add
  a confirmatory clock term after inspecting these plots. Material imbalance
  or residual time patterning is reported as a limitation and contextualized
  only in the separate exploratory GAM/GAMM.
- Examine HC0–HC3 covariance sensitivity, cluster scores, cluster leverage,
  targeted participant deletion, and every leave-one-site refit, with special
  attention to sparse binary activity cells, the corrected TUM days, the
  daytime sleep interval, the late wake interval, and extreme active-travel or
  sedentary values. Do not remove a unit solely because it changes
  significance.
- Compare equal-site and observed-sample marginalization. Reader-facing
  “overall” effects use the author-approved scheme and state it explicitly.
- For any exploratory GAM/GAMM, check cyclic closure, basis dimension,
  concurvity, residual dependence, influential participants/sites, and a
  whole-curve interval method supported directly by the discrete fit. The fit
  and every downstream computation must remain compatible with
  `bam(discrete = TRUE)`; a calculation that requires another fitting mode is
  not run.

## Heavy computation boundary

No resampling is run in Stage 1. If Stage 2 requires a cluster bootstrap,
simultaneous-curve bootstrap, or other heavy resampling, first run only 50 or
100 replicates, show the pilot output and measured runtime, estimate the full
runtime, and stop for explicit production approval. Pilot replicates cannot
support reportable inferential conclusions.

# Predeclared sensitivity menu

The recommended Stage 2 menu is limited to interpretable, prespecified axes:

1. the required gap-timing-unaware dataset comparison under the identical
   H06 implementation;
2. quasi-Poisson primary working variance versus fixed-power quasi-Tweedie
   `p = 1.8`, with HC1 and HC2 covariance comparisons to primary HC3;
3. work/free primary versus the approved weekday/weekend day-type
   sensitivity, reported as an estimate/CI stability comparison without a new
   significance screen;
4. near-eye all available versus paired/common near-eye and chest;
5. equal-site versus observed-sample standardization on identical fits;
6. exact-zero support, coverage thresholds, and declared measurement-support
   comparisons already fixed upstream;
7. the mandatory participant and site influence/deletion battery, reported as
   stability evidence rather than an exclusion rule;
8. the rejected participant/day/AR mixed architecture only as a frozen
   construction comparison, never as a routine sensitivity or rescue model;
9. residual local-clock and true-time sequence diagnostics without a
   confirmatory clock or AR term; and
10. the non-selected H06-G1 estimand only if the author separately requests a
    cross-estimand comparison; it remains explicitly different and is not
    treated as a routine robustness check of the selected question.

Each sensitivity reports exact samples, estimates and 95% CIs, diagnostic
status, main-minus-alternative differences, and one of the approved stability
classifications. Sensitivities do not accumulate undeclared significance
screens. Exercise type remains excluded free text. The one author-approved
sedentary reinterpretation may appear only in the labelled exploratory model,
never in the primary analysis. Temperature is not introduced. Exploratory
time-of-day work must use `bam(discrete = TRUE)` throughout; no model or
downstream computation requiring `discrete = FALSE` is part of this menu.

# Stage 1 findings

```{r}
#| label: stage1-findings

h06_findings <- tibble::tribble(
  ~finding_id, ~severity, ~status, ~finding, ~evidence, ~required_disposition,
  "H06-S1-001", "major", "author_approved_deliberate_deviation",
  "V0 deliberately replaced the registered daily-metric set with one repeated hourly geometric-melEDI outcome for parsimony and interpretability.",
  "Owner clarification; DEV-015; signed PDF; RQ2.qmd H06",
  "Implement the repaired hourly estimand as approved Option B, retain the daily question as a visible unselected reference, and do not characterize intent as an error.",
  "H06-S1-002", "major", "author_approved_scope",
  "V0 narrowed the registered predictor scope to work/free, exercise intensity, and sleep duration.",
  "Supporting plan; DEV-030",
  "Use work/free, binary activity, and previous sleep duration as the primary set; weekday/weekend is a sensitivity and other named measures are exploratory only.",
  "H06-S1-003", "major", "superseded_after_adequacy_failure",
  "V0 repeated daily predictors across hours without participant-day or residual temporal dependence.",
  "RQ2.qmd:1866–1907; DEV-031",
  "Use participant-cluster robust covariance in the revised population-mean model; retain true-time sequences for residual diagnostics and the exploratory GAMM, without a confirmatory local-clock term.",
  "H06-S1-004", "major", "author_approved_repair",
  "V0 likelihood-ratio tests removed a predictor and all its site interactions but were labelled main effects.",
  "V0 full/reduced formulas",
  "Test additive main association and interaction heterogeneity separately with hierarchy.",
  "H06-S1-005", "major", "author_approved_repair",
  "V0 multiplicity labels do not match the performed adjustments.",
  "DEV-032; V0 table code/output",
  "Use the five approved complete vector-wide BH registries with fixed planned sizes and no site-specific significance screen.",
  "H06-S1-006", "critical", "confirmed",
  "The displayed H06 site p-value was inherited from H04's stale comp_site object.",
  "IMP-007; RQ2.qmd:2137–2140",
  "Withdraw it and generate every H06 statistic from an H06-specific producer.",
  "H06-S1-007", "major", "confirmed",
  "V0 dropped 95% intervals and described overlapping marginal-R² differences as unique contributions.",
  "Frozen H6 tables; RQ2.qmd:1937–1955, 2030–2032",
  "Report intervals and valid model-performance summaries without additive decomposition language.",
  "H06-S1-008", "major", "author_approved_exploratory_repair",
  "The V0 GAM used a mismatched arithmetic outcome, noncyclic time, incomplete AR starts, and pointwise-selected curve segments.",
  "RQ2.qmd:2274–2659",
  "Retain only a repaired, separately labelled exploratory GAM/GAMM whose fit and downstream computations are compatible with bam(discrete = TRUE).",
  "H06-S1-009", "informational", "approved_verified_input",
  "The current TUM exercise diary assigns seven dates to S001 and zero rows to S101.",
  "Pinned source commit 618fda85…; normalized RDS; DEV-056",
  "Join the authoritative input unchanged, repeat the Stage 2 assertion, and do not fetch or rewrite it.",
  "H06-S1-010", "major", "resolved_for_h06_exploration",
  "One sitting/reclining value is 3,600 minutes in a participant-day.",
  "KNUST_S005, 2024-11-04, normalized exercise diary",
  "Leave the shared source unchanged; provenance-label the H06 exploratory interpretation as 3,600 seconds = 1 h and keep sedentary time outside primary inference.",
  "H06-S1-011", "moderate", "shared_change_requested",
  "The Preparation 06 base-model narrative retains old 811/897 counts and hashes after the approved 816/902 transition.",
  "base_model_gate.md versus H02-INPUT-001 and manifest 142d1044…",
  "Coordinator should refresh or supersede the stale narrative gate.",
  "H06-S1-012", "major", "confirmed",
  "V0 model frames and H06 model objects were not durably persisted; moving diary loads prevent an exact refit from being called a reproduction.",
  "RQ2.qmd; frozen DOCX outputs",
  "Treat frozen values as historical artifact reconstruction and compare Stage 2 against them explicitly.",
  "H06-S1-013", "major", "confirmed",
  "Sleep-period light is a bedside-environment construct, not worn ocular exposure.",
  "STATE-005; metric implementation contract",
  "Use the hybrid/bedside wording in every metric, result, figure, and claim.",
  "H06-S1-014", "major", "revised_architecture_proposed",
  "H03 resolved the analogous mixed-model failure by using a population-mean log-link GLM with participant-cluster robust covariance rather than another random-effects model.",
  "Accepted H03 Stage 1–4 reports and H03 worker cross-check; H06 scaffold pilots",
  "Use the same marginal inferential route for H06, while retaining H06-specific predictors, site hierarchy, covariance choice, and working-variance checks.",
  "H06-S1-015", "major", "bounded_pilot_pass",
  "The H06 quasi-Poisson additive and full fixed designs converge at full rank with finite positive-definite unmodified HC3 covariance at both placements.",
  "H06_h03_robust_route_pilot_mean_structure_gate.csv; H06_h03_robust_route_pilot_summary.csv",
  "Propose quasi-Poisson log-mean with participant-cluster HC3 as primary; require full influence, residual, site-deletion, data, placement, and multiplicity checks before any claim.",
  "H06-S1-016", "moderate", "working_variance_boundary_resolved",
  "A fixed quasi-Tweedie power of 1.8 is numerically stable at both placements, whereas p = 1.9 fails near eye; H03's p = 1.539919 is not an H06 parameter.",
  "H06_h03_robust_route_pilot_fit_diagnostics.csv; effect-blind power screen",
  "Use p = 1.8 only as the fixed working-variance sensitivity, reject p = 1.9, and report sensitivity without likelihood language."
)

h06_findings |>
  compact_gt(
    source_note = "No shared ledger status is changed by this H06-owned Stage 1 report."
  ) |>
  gt::cols_label(
    finding_id = "Finding",
    severity = "Severity",
    status = "Status",
    finding = "Audit finding",
    evidence = "Evidence",
    required_disposition = "Required disposition"
  )
```

# V0 claim disposition

| V0 result or claim | Stage 1 disposition | Requirement before future use |
|---|---|---|
| Work/free day p-value, R² and named site ratios | Provisional; hourly intent is valid but V0 implementation is not accepted | H06-specific sample/statistics; corrected hierarchy and dependence; multiplicity; 95% CIs; diagnostics |
| Exercise p-value, R² and four displayed intensity ratios | Provisional V0 history and not one-to-one comparable with the approved binary activity contrast | Reconstruct the four-level V0 result faithfully, then report the new active-versus-sedentary estimand with its own exact sample, interval, diagnostics, and multiplicity |
| 0.92 sleep ratio interpreted as 8% lower “daily exposure” | Provisional and outcome-mislabelled | Exact previous-night frame, interval, temporal diagnostics, and wording that identifies supported hourly melEDI |
| Site p-value | Invalid H06 provenance | Replace; never reuse the H04 object |
| “Unique contributions” of 0.01–0.05 | Unsupported interpretation | Do not reproduce as an additive variance decomposition |
| Weekend model approximately 33 AIC points worse | Provisional V0 exploration | Recompute only under an approved same-frame day-type comparison |
| Time-of-day work/free patterns | Exploratory only | Repaired geometric outcome, cyclic clock, sequence-aware residuals, and a whole-curve interval procedure compatible with `bam(discrete = TRUE)` |
| Associations remain after activity-context adjustment | Withdrawn | Requires a durable approved activity producer and observational, non-mechanistic wording |
| “Small but robust” | Withdrawn | Requires the complete approved Stage 2 sensitivity and diagnostic record |

No manuscript, supplementary, shared preparation, central ledger, or V0 file
is edited at Stage 1.

# Earlier author decisions and revised H06-G1A

The owner resolved the original 13 Stage 1 decisions on 2026-08-01. Decisions
about the estimand, outcome, predictors, day type, site display, multiplicity,
placement, claims, source pin, and exploratory boundaries remain in force.
The mixed-model architecture clauses are superseded by the 2026-08-10 bounded
pilot and now require the revised H06-G1A approval below.

1. **H06-G1 estimand — Option B approved.** The repaired deliberate hourly
   adaptation is primary. The registered daily-metric analysis remains a
   scientifically distinct, unselected reference rather than a routine
   sensitivity.
2. **Primary day type — approved as recommended.** Work/free is primary;
   weekday/weekend is an estimate/CI sensitivity without a new significance
   screen.
3. **Predictor scope — approved with an exploratory boundary.** The primary
   set is work/free, binary activity, and previous sleep duration. Exercise
   location, active travel, sedentary time, previous sleep onset, and final
   wake are exploratory only and are excluded from primary inference.
   Logger-worn status is a measurement-support sensitivity; free-text exercise
   type is excluded.
4. **Outcome — Option B approved.** The only primary response is supported
   one-hour zero-aware geometric mean melEDI. It is not called daily exposure.
5. **Site structure — approved as recommended.** Fixed site, sum-to-zero site
   contrasts, separate hierarchy-preserving heterogeneity tests, and equal-site
   reader-facing marginalization are primary. Observed-sample marginalization
   remains a comparison; the registered random-site mixed form is retained as
   historical construction evidence rather than a routine sensitivity.
6. **Repeated-day/time structure — revised by H06-G1A.** The original
   participant, participant-day, and gap-bounded AR(1) proposal failed. The
   revised primary uses participant-cluster HC3 covariance and finite-cluster
   reference distributions. True-time sequences remain diagnostic, and the
   exploratory GAM/GAMM alone addresses time of day.
7. **Response and clock rules — revised by H06-G1A.** Quasi-Poisson log-mean
   is proposed as primary; fixed-power quasi-Tweedie `p = 1.8` is the working-
   variance sensitivity; `p = 1.9` is rejected. The no-confirmatory-clock,
   sleep-onset unwrap, and 23:00/07:00/8 h center rules remain unchanged.
8. **Multiplicity — approved as recommended.** The active H06-F1–F3 and gap
   families are fixed above. Site-specific effects and paired/common placement
   comparisons receive estimates and 95% CIs without an additional p-value
   screen.
9. **Placement/sample strategy — approved.** All-available near-eye is
   primary; separately refitted paired/common near-eye and chest are
   complementary; all-available chest is contextual.
10. **Sensitivity set — revised by H06-G1A.** The bounded menu above applies;
    HC1/HC2, `p = 1.8`, weekday/weekend, data, standardization, influence, and
    placement comparisons do not create new significance families. The daily
    reference, additional predictor explorations, and GAM/GAMM remain outside
    primary inference.
11. **Intervals, diagnostics, failure rules, and resampling — revised by
    H06-G1A.** All consequential estimates receive 95% participant-cluster
    robust CIs. Unmodified covariance adequacy, cluster-score/leverage,
    targeted participant deletion, leave-site refits, residual lag, and mean
    calibration are mandatory. Any heavy resampling first stops after 50 or
    100 replicates with output and a full-runtime estimate.
12. **Claim gate — approved.** Every V0 number and narrative claim remains
    provisional until the Stage 2 comparison, exact sample, interval,
    multiplicity, and diagnostics are accepted. Deliberate intent is not an
    implementation-validity finding.
13. **Source/upstream disposition — resolved.** The first clause means that
    H06 uses the coordinator-approved immutable `melidosData` 1.0.6 pin rather
    than checking or fetching a newer remote release. Stage 2 will make only a
    read-only assertion of seven TUM S001 rows and zero S101 rows. For
    `KNUST_S005`, the source numeral 3,600 is author-interpreted as seconds,
    yielding 1 h in the H06 exploratory derivation only; the shared source is
    not edited. The stale Preparation 06 narrative remains a coordinator-owned
    documentation request and does not reopen the H06 data gate.

## Author decision requested: H06-G1A

Approval of H06-G1A means approving this complete revised architecture:

1. quasi-Poisson log-link additive mean model for H06-F1 and average effects;
2. separate full site-interaction model for H06-F2 and site-specific
   descriptive estimates;
3. participant-cluster HC3 covariance with `cadjust = TRUE`, `fix = FALSE`,
   and finite `G - 1` *t*/*F* inference;
4. fixed-power quasi-Tweedie `p = 1.8` plus HC1/HC2 as working-variance and
   covariance sensitivities, with no likelihood claims;
5. no participant, participant-day, latent-AR, working-AR, random-site, or
   confirmatory local-clock term in primary inference; and
6. mandatory residual, mean-calibration, participant-influence, site-deletion,
   data, placement, multiplicity, and V0-comparison checks before any claim.

**Current gate:** H06-G1A is awaiting explicit author approval. The earlier
Stage 2 artifacts remain diagnostic construction history and contain no
accepted H06 result. No production refit under this revised architecture,
Stage 3 notebook, Stage 4 preparation companion, manuscript edit, or central
ledger change is authorized by this report. After author approval, the
coordinator must record the revised gate before the H06 worker implements the
production marginal analysis and returns to the Stage 2 author-review gate.

# Reproducibility assertions and environment

```{r}
#| label: audit-assertions

stopifnot(
  nrow(daily_metric_contract) == 15L,
  nrow(sample_contract) == 45L,
  nrow(daily_core_formulas) == length(core_predictors) * 7L,
  nrow(daily_additional_formulas) == length(additional_measure_columns) * 3L,
  nrow(hourly_formula_table) == 4L,
  nrow(hourly_exploratory_formula_table) == 5L,
  nrow(robust_mean_contract) == 8L,
  nrow(two_part_formula_registry) == 2L,
  all(two_part_formula_registry$discrete),
  length(exploratory_formulas) == 3L,
  nrow(gamm_execution_contract) == 2L,
  nrow(multiplicity_registry) == 5L,
  nrow(h06_findings) == 16L,
  nrow(h03_route_registry) == 4L,
  nrow(h03_route_diagnostics) == 8L,
  nrow(h03_route_mean_gate) == 2L,
  nrow(h03_route_summary) == 2L,
  all(h03_route_summary$numerical_gate_pass),
  all(h03_route_summary$additive_numerical_gate_pass),
  all(h03_route_mean_gate$converged),
  all(h03_route_mean_gate$full_rank),
  all(h03_route_mean_gate$hc3_covariance_positive_definite),
  !h03_route_diagnostics$converged[
    h03_route_diagnostics$placement == "glasses" &
      h03_route_diagnostics$route_id == "h06_v0_boundary_p1_9"
  ],
  nrow(dev015) == 1L,
  dev015$status == "open_reaudit",
  all(input_contract$verified),
  all(contract_sources$verified),
  all(v0_contract$verified),
  nrow(exercise_duplicate_keys) == 0L,
  nrow(sleep_duplicate_keys) == 0L,
  all(
    as.character(exercise_day$activity_status[
      !is.na(exercise_day$exercise_intensity) &
        exercise_day$exercise_intensity == "No exercise"
    ]) == activity_status_levels[[1]]
  ),
  all(
    as.character(exercise_day$activity_status[
      !is.na(exercise_day$exercise_intensity) &
        exercise_day$exercise_intensity != "No exercise"
    ]) == activity_status_levels[[2]]
  ),
  nrow(sedentary_reinterpreted_row) == 1L,
  sedentary_reinterpreted_row$sedentary_source_minutes == 3600,
  sedentary_reinterpreted_row$sedentary_h == 1,
  !any(grepl("clock_hour", hourly_formula_table$formula, fixed = TRUE)),
  !any(grepl("participant_key", hourly_formula_table$formula, fixed = TRUE)),
  !any(grepl("participant_day", hourly_formula_table$formula, fixed = TRUE)),
  !any(grepl("ar1(", hourly_formula_table$formula, fixed = TRUE)),
  !any(grepl("exercise_intensity", hourly_formula_table$formula, fixed = TRUE)),
  !any(grepl(
    "exercise_intensity",
    hourly_exploratory_formula_table$formula,
    fixed = TRUE
  )),
  all(gamm_execution_contract$discrete),
  nrow(tum_correction) == 7L,
  all(tum_correction$Id == "TUM_S001"),
  sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S101") == 0L,
  current_hourly_sample$`One-hour observations`[
    current_hourly_sample$placement == "Near-eye"
  ] == 16596L,
  current_hourly_sample$`One-hour observations`[
    current_hourly_sample$placement == "Chest"
  ] == 18352L,
  gap_hour_sample$`One-hour observations`[
    gap_hour_sample$placement == "Near-eye"
  ] == 16329L,
  gap_hour_sample$`One-hour observations`[
    gap_hour_sample$placement == "Chest"
  ] == 18112L
)

environment_summary <- tibble::tibble(
  Component = c(
    "R",
    "Quarto",
    "melidosData",
    "dplyr",
    "gt",
    "digest",
    "xml2",
    "sandwich",
    "statmod",
    "Scientific model fits in this render",
    "Bounded architecture fits read by this render",
    "Analytical files written by R cells"
  ),
  Value = c(
    R.version.string,
    "1.9.37",
    as.character(utils::packageVersion("melidosData")),
    as.character(utils::packageVersion("dplyr")),
    as.character(utils::packageVersion("gt")),
    as.character(utils::packageVersion("digest")),
    as.character(utils::packageVersion("xml2")),
    as.character(utils::packageVersion("sandwich")),
    as.character(utils::packageVersion("statmod")),
    "None",
    "Frozen H06 pilot CSV/RDS artifacts produced separately in R 4.6.1",
    "None"
  )
)

environment_summary |>
  compact_gt(
    source_note = paste0(
      "Rendered from the project root with R 4.6.1 and the project library. ",
      "Input identities, commands, formulas, and consequential package ",
      "versions are visible in this audit."
    )
  )
```

The first project-default render attempt stopped before knitting because the
project `renv/activate.R` startup hook did not return. No document code had
run. The successful standalone render therefore bypassed only that startup
hook while pointing `R_LIBS_USER` directly at the existing project library;
no package was installed, updated, or resolved. The exact successful command
was:

```text
env R_PROFILE_USER=/dev/null R_ENVIRON_USER=/dev/null R_LIBS=renv/library/macos/R-4.6/aarch64-apple-darwin23 quarto render audit/hypotheses/H06/01_audit_and_plan.qmd --to html
```

This is the end of H06 Stage 1.
