---
title: "H06 Stage 2 amendment pilot: revised V0 scaffold models"
subtitle: "Feasibility assessment only — no H06 result is accepted"
date: "2026-08-06"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: true
    code-summary: "Show audit code"
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r}
#| label: setup

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- locate_project_root()
source(file.path(project_root, "scripts", "pipeline", "p_value_display.R"))
artifact <- function(...) file.path(project_root, "artifacts", ...)
read_h06 <- function(...) {
  readr::read_csv(artifact(...), show_col_types = FALSE, na = "")
}
compact_gt <- function(data, title, subtitle = NULL) {
  gt::gt(data) |>
    gt::tab_header(title = title, subtitle = subtitle) |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.align = "left",
      table.width = gt::pct(100)
    )
}
format_ratio <- function(estimate, low, high, digits = 2L) {
  sprintf(
    paste0("%.", digits, "f × (%.", digits, "f to %.", digits, "f)"),
    estimate,
    low,
    high
  )
}

registry <- read_h06(
  "06_model_data", "H06", "H06_v0_scaffold_pilot_registry.csv"
)
samples <- read_h06(
  "06_model_data", "H06", "H06_v0_scaffold_pilot_samples.csv"
)
fit_status <- read_h06(
  "08_diagnostics", "H06", "H06_v0_scaffold_pilot_fit_status.csv"
)
temporal <- read_h06(
  "08_diagnostics", "H06", "H06_v0_scaffold_pilot_temporal.csv"
)
dharma <- read_h06(
  "08_diagnostics", "H06", "H06_v0_scaffold_pilot_DHARMa_100.csv"
)
runtime <- read_h06(
  "08_diagnostics", "H06", "H06_v0_scaffold_pilot_simulation_runtime.csv"
)
assessment <- read_h06(
  "09_tables", "H06", "H06_v0_scaffold_pilot_assessment.csv"
)
effects <- read_h06(
  "09_tables", "H06", "H06_v0_scaffold_pilot_effects.csv"
)
robust_assessment <- read_h06(
  "09_tables", "H06", "H06_v0_scaffold_pilot_robust_assessment.csv"
)
robust_covariance <- read_h06(
  "08_diagnostics", "H06",
  "H06_v0_scaffold_pilot_robust_covariance_sensitivity.csv"
)
robust_calibration <- read_h06(
  "08_diagnostics", "H06",
  "H06_v0_scaffold_pilot_robust_calibration.csv"
)
robust_deletions <- read_h06(
  "08_diagnostics", "H06",
  "H06_v0_scaffold_pilot_robust_targeted_deletions.csv"
)
robust_leave_site <- read_h06(
  "08_diagnostics", "H06",
  "H06_v0_scaffold_pilot_robust_leave_site.csv"
)
figure_audit <- read_h06(
  "11_source_data", "H06",
  "H06_v0_scaffold_pilot_figure_readability.csv"
)
bam_gamm_audit <- read_h06(
  "08_diagnostics", "H06",
  "H06_v0_scaffold_pilot_bam_gamm_audit.csv"
)
bam_sequence_lag <- read_h06(
  "08_diagnostics", "H06",
  "H06_v0_scaffold_pilot_bam_per_sequence_residual_lag.csv"
)

candidate_labels <- c(
  tmb_v0_participant = "V0 participant intercept",
  tmb_participant_day = "Participant + participant-day intercepts",
  tmb_participant_latent_ar = "Participant + latent AR(1)",
  tmb_participant_day_latent_ar = "Participant + day + latent AR(1)",
  bam_participant_day_working_ar = "Participant + day + working AR(1)",
  glm_participant_cluster_robust = "Participant-cluster robust marginal"
)
effect_labels <- c(
  free_day_vs_work_day = "Free day / work day",
  active_vs_sedentary = "Active / sedentary",
  per_hour_previous_sleep = "Per hour previous sleep"
)
placement_labels <- c(glasses = "Near-eye", chest = "Chest")

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  nrow(registry) == 6L,
  all(registry$collapsed_activity),
  !any(registry$confirmatory_clock_term),
  all(registry$inferential_status ==
    "feasibility_pilot_not_accepted_inference"),
  nrow(samples) == 2L,
  all(samples$fixed_design_full_rank),
  identical(samples$one_hour_observations, c(16596, 18352)),
  identical(samples$participant_days, c(715, 789)),
  identical(samples$participants, c(137, 149)),
  identical(samples$sites, c(9, 8)),
  nrow(fit_status) == 12L,
  nrow(assessment) == 12L,
  nrow(effects) == 36L,
  all(dharma$simulations == 100L),
  all(runtime$production_status == "not_run_not_authorized"),
  all(robust_deletions$refit_converged),
  all(robust_leave_site$refit_converged),
  nrow(robust_assessment) == 2L,
  !any(robust_assessment$confirmatory_clock_term),
  nrow(figure_audit) == 1L,
  figure_audit$report_011_status == "PASS",
  figure_audit$effective_final_text_pt >= 7,
  nrow(bam_gamm_audit) == 2L,
  all(bam_gamm_audit$random_effect_penalties_effectively_absent),
  all(bam_gamm_audit$simulated_zero_fraction_100 > 0.8),
  all(bam_gamm_audit$observed_zero_fraction < 0.3),
  all(bam_gamm_audit$positive_tail_calibration ==
    "NOT TESTED in the stored pilot"),
  all(c("pearson", "AR_standardized") %in%
    unique(bam_sequence_lag$residual_type))
)
```

# Assessment at a glance

::: {.callout-important title="Pilot conclusion"}
The **V0 fixed-effects scaffold can be retained**, with exercise collapsed to
sedentary or active. The **V0 participant-only random structure cannot be
retained as the inferential basis**: it is nonsingular, but leaves substantial
within-day dependence and fails the simulated distributional checks at both
placements.

Of the six piloted dependence/error strategies, only the
**participant-cluster robust marginal log-mean model** remains structurally
estimable without requiring a full Tweedie outcome distribution. It is the
recommended candidate for a production amendment, preferably with HC3 as the
leverage-sensitive covariance specification. This is not yet an accepted H06
model or result. The candidate remains sensitive to individual participants,
especially for the chest activity estimate, and must pass the complete
predeclared Stage 2 sensitivity and reporting contract before any H06 claim or
Stage 3 transition.
:::

The pilot preserves the deliberate hourly outcome and approved average-effect
question. It does not reopen the registered daily-metric alternative, add a
local-clock effect, add temperature, change the immutable source pin, or alter
shared preparation. The local-clock profile below is a residual calibration
diagnostic only; the separate exploratory GAM/GAMM remains the analysis of
time of day.

# Fixed-effects scaffold and exact samples

Every candidate uses this exact hierarchy-preserving fixed scaffold:

```{r}
#| label: fixed-scaffold-visible
#| echo: true
#| results: hide

response_value ~
  site * work_free_day +
  site * activity_status +
  site * previous_sleep_duration_centered_h
```

`activity_status` has exactly two levels: sedentary (no reported exercise) and
active (light, moderate, or vigorous exercise). Submitted site ordering is
retained, with sum contrasts so each displayed main coefficient is the
equal-site average association represented by that fitted model. The mixed
models estimate conditional associations, whereas the cluster-robust model
estimates a population-average mean association; their numerical estimates
are therefore a model-dependence diagnostic rather than interchangeable
versions of one inferential result.

```{r}
#| label: exact-pilot-samples

samples |>
  dplyr::transmute(
    Placement = placement_labels[placement],
    Hours = one_hour_observations,
    `Participant-days` = participant_days,
    Participants = participants,
    Sites = sites,
    `Fixed design` = paste0(fixed_design_rank, "/", fixed_design_columns),
    `Frame SHA-256` = frame_sha256
  ) |>
  compact_gt(
    "Exact common pilot frames",
    "The same repaired three-predictor frame is used by all six candidates within each placement."
  ) |>
  gt::cols_width(`Frame SHA-256` ~ gt::px(235))
```

Near-eye contains 16,596 supported one-hour observations from 715
participant-days, 137 participants, and nine sites. Chest contains 18,352
hours from 789 participant-days, 149 participants, and eight sites. Both fixed
designs are full rank. These are the approved primary-dataset frames; no new
filter was introduced for the pilot.

# Six piloted strategies

```{r}
#| label: pilot-registry

registry |>
  dplyr::transmute(
    Candidate = candidate_labels[candidate_id],
    Engine = engine,
    `Outcome/mean specification` = response_family,
    `Dependence strategy` = dependence_strategy,
    Formula = paste0("`", formula, "`")
  ) |>
  compact_gt(
    "V0-scaffold candidate registry",
    "All formulas were evaluated under R 4.6.1. Every fit is a feasibility pilot, not accepted inference."
  ) |>
  gt::fmt_markdown(columns = Formula) |>
  gt::cols_width(
    Candidate ~ gt::px(155),
    Engine ~ gt::px(105),
    `Outcome/mean specification` ~ gt::px(130),
    `Dependence strategy` ~ gt::px(210),
    Formula ~ gt::px(390)
  )
```

The working-AR candidate uses `mgcv::bam(method = "fREML", discrete = TRUE)`
for both the initial and AR-aware fits. No model, prediction, interval,
simulation, or downstream computation in this pilot uses or requires
`discrete = FALSE`.

# Structural, temporal, and distributional screens

```{r}
#| label: candidate-screen

screen_table <- assessment |>
  dplyr::left_join(
    fit_status |>
      dplyr::select(
        placement, candidate_id, converged, positive_definite_hessian,
        singular, random_component_boundary, participant_intercept_sd,
        participant_day_intercept_sd, latent_ar_rho, working_ar_rho
      ),
    by = c("placement", "candidate_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::mutate(
    Placement = placement_labels[placement],
    Candidate = candidate_labels[candidate_id],
    Structural = dplyr::case_when(
      automated_pilot_screen == "structurally_estimable_marginal_candidate" ~
        "Estimable marginal candidate",
      structural_screen ~ "Pass",
      TRUE ~ "Fail: boundary/singular"
    ),
    Temporal = dplyr::case_when(
      is.na(temporal_screen) ~ "Not a model-validity test",
      temporal_screen ~ "Pass",
      TRUE ~ "Fail"
    ),
    Distribution = dplyr::case_when(
      is.na(distribution_screen) ~ "Not specified by design",
      distribution_screen ~ "Pass",
      TRUE ~ "Fail"
    ),
    `Lag-1 used for screen` = dplyr::if_else(
      is.finite(residual_lag_for_screen),
      sprintf("%.3f", residual_lag_for_screen),
      "—"
    )
  ) |>
  dplyr::select(
    Placement, Candidate, Structural, `Lag-1 used for screen`, Temporal,
    Distribution
  )

screen_table |>
  compact_gt(
    "Automated feasibility screens",
    "For BAM, the temporal screen uses AR-standardized residuals. For the marginal robust model, residual correlation is allowed and is handled by participant-cluster covariance."
  )
```

The two nonsingular Tweedie models—the V0 participant intercept and the added
participant-day intercept—still have lag-1 residual correlations of
0.286–0.356 and fail every applicable 100-simulation distribution screen.
Adding latent AR(1) removes most residual lag but collapses participant and,
where present, participant-day variance to the boundary. The discrete BAM
working-AR fit likewise collapses both random-effect penalties and fails its
simulated distribution checks; near-eye standardized lag-1 is 0.105, just
outside the declared 0.10 screen.

The GAMM-specific re-audit strengthens that conclusion. The pooled lag used
by the original automated screen hides a broad distribution across short
within-day series. The table therefore reports per-sequence medians and
interquartile ranges, along with the fitted Tweedie and smooth-penalty
parameters.

```{r}
#| label: bam-gamm-specific-audit

bam_gamm_audit |>
  dplyr::mutate(
    Placement = placement_labels[placement],
    `Tweedie p` = tweedie_power,
    `Dispersion phi` = scale_parameter_phi,
    `Working rho` = working_ar_rho,
    `Participant EDF` = participant_random_effect_edf,
    `Participant-day EDF` = participant_day_random_effect_edf,
    `Standardized lag-1, median (IQR)` = sprintf(
      "%.3f (%.3f to %.3f)",
      standardized_sequence_lag1_median,
      standardized_sequence_lag1_first_quartile,
      standardized_sequence_lag1_third_quartile
    ),
    `Observed / simulated zeros` = sprintf(
      "%.1f%% / %.1f%%",
      100 * observed_zero_fraction,
      100 * simulated_zero_fraction_100
    ),
    `In-sample deviance explained` = deviance_explained
  ) |>
  dplyr::select(
    Placement, `Tweedie p`, `Dispersion phi`, `Working rho`,
    `Participant EDF`, `Participant-day EDF`,
    `Standardized lag-1, median (IQR)`,
    `Observed / simulated zeros`, `In-sample deviance explained`
  ) |>
  compact_gt(
    "Stored BAM re-audit",
    paste(
      "No model was refitted and no simulation was added; all diagnostics",
      "come from the stored discrete BAM and its existing 100 simulations."
    )
  ) |>
  gt::fmt_number(
    columns = c(
      `Tweedie p`, `Dispersion phi`, `Working rho`,
      `Participant EDF`, `Participant-day EDF`,
      `In-sample deviance explained`
    ),
    decimals = 3
  )
```

The estimated Tweedie powers are about 1.90 within the default 1.01–1.99
search interval, with dispersion about 84–97. The participant and
participant-day smooth EDFs are below 0.001 at both placements: their
penalized contributions are effectively absent even though a classical
mixed-model singularity flag is not available for BAM. Median
AR-standardized within-sequence lag remains 0.199 near-eye and 0.167 chest,
with wide interquartile ranges. Thus the working AR term reduces pooled
dependence but does not establish homogeneous residual independence.

For a non-Gaussian BAM, `rho` is a working-residual GEE approximation rather
than a full joint or generative AR likelihood. The existing 100 custom
simulations likewise test the marginal conditional Tweedie response
distribution; they do not simulate a full AR process. That marginal test is
already decisive: the model generates 84.7% zero near-eye hours and 82.1%
zero chest hours, compared with 28.3% and 29.1% observed. Uniformity and
zero-mass raw diagnostic *p* values are zero, and outlier diagnostics are also
incompatible. Positive-tail calibration was not separately tested in the
stored pilot, which is an additional missing adequacy check rather than a
reason to rehabilitate the model. A compound-process Tweedie distribution is
not made scientifically compelling merely by zeros and skew in an hourly
zero-aware geometric-mean light metric.

```{r}
#| label: simulation-diagnostics

dharma |>
  dplyr::transmute(
    Placement = placement_labels[placement],
    Candidate = candidate_labels[candidate_id],
    Simulations = simulations,
    `Uniformity diagnostic p` = nh_format_p_value(uniformity_p),
    `Dispersion diagnostic p` = nh_format_p_value(dispersion_p),
    `Zero-mass diagnostic p` = nh_format_p_value(zero_inflation_p),
    `Outlier diagnostic p` = nh_format_p_value(outlier_p)
  ) |>
  compact_gt(
    "Bounded simulation diagnostics for generative candidates",
    "Exactly 100 conditional simulations per placement and candidate; displayed values are raw diagnostic p-values with no significance styling or H06 decision rule."
  )
```

The ten generative candidate-placement checks took
`r sprintf("%.1f", sum(runtime$elapsed_seconds))` seconds at 100 simulations.
A linear projection to 1,000 simulations is
`r sprintf("%.1f", sum(runtime$projected_1000_simulation_seconds))` seconds in
total, with a maximum of
`r sprintf("%.1f", max(runtime$projected_1000_simulation_seconds))` seconds for
one fit. No production simulation or bootstrap was run because every
generative candidate already fails clearly and no larger run was authorized.

# How the candidate estimates move

```{r}
#| label: pilot-effect-figure
#| fig-cap: "Equal-site average effect ratios and pilot 95% intervals under six V0-scaffold dependence/error strategies. These intervals are feasibility output, not accepted H06 inference."
#| fig-alt: "Six candidate models are compared for near-eye and chest outcomes across free versus work day, active versus sedentary, and previous sleep duration. Estimates vary materially across dependence strategies, particularly for previous sleep. The participant-cluster robust marginal candidate shows activity ratios above one at both placements, while free-day and sleep intervals include one."
#| out-width: "170mm"

knitr::include_graphics(
  artifact("10_figures", "H06", "H06_v0_scaffold_pilot_effects.png")
)
```

Figure source:
`artifacts/11_source_data/H06/H06_v0_scaffold_pilot_effects_figure.csv`.

The spread is scientifically consequential. Near-eye previous-sleep ratios
range from 0.808 under the singular latent-AR fit to 0.973 under the marginal
robust fit; the analogous chest range is 0.772 to 0.943. Activity ranges from
1.18 to 1.93 near-eye and 1.22 to 1.54 chest. This is model dependence, not a
menu from which to select a preferred finding.

The robust candidate's covariance-only leverage sensitivity is:

```{r}
#| label: robust-covariance-sensitivity

robust_covariance |>
  dplyr::filter(covariance_type %in% c("HC1", "HC3")) |>
  dplyr::mutate(
    Placement = placement_labels[placement],
    Effect = effect_labels[effect_id],
    Covariance = covariance_type,
    `Ratio (pilot 95% CI)` = format_ratio(
      estimate_response_ratio,
      conf_low_response_ratio,
      conf_high_response_ratio
    )
  ) |>
  dplyr::select(Placement, Effect, Covariance, `Ratio (pilot 95% CI)`) |>
  compact_gt(
    "Participant-cluster robust covariance sensitivity",
    "HC3 increases protection against influential clusters. Intervals use t critical values with participant clusters minus one degree of freedom and are not accepted inference."
  )
```

HC3 increases the largest target standard error by a factor of
`r sprintf("%.2f", robust_assessment$maximum_hc3_to_hc1_se_ratio[robust_assessment$placement == "glasses"])`
near-eye and
`r sprintf("%.2f", robust_assessment$maximum_hc3_to_hc1_se_ratio[robust_assessment$placement == "chest"])`
at chest. The activity intervals remain above one in this pilot; the free-day
and sleep intervals include one. No p-values or multiplicity decisions are
made here.

# Stress test of the robust marginal candidate

The quasi-Poisson variance is deliberately only a working variance:
deviance/residual-df is
`r sprintf("%.1f", robust_assessment$deviance_per_residual_df[robust_assessment$placement == "glasses"])`
near-eye and
`r sprintf("%.1f", robust_assessment$deviance_per_residual_df[robust_assessment$placement == "chest"])`
chest. Model-based quasi-Poisson standard errors are therefore not used. The
candidate relies on participant-cluster sandwich covariance and a log mean
model, not a claim that melEDI is Poisson distributed.

Mean calibration by fitted-value decile is comparatively close but not exact:
observed/fitted ratios range from
`r sprintf("%.3f", robust_assessment$fitted_decile_min_observed_to_fitted[robust_assessment$placement == "glasses"])`
to
`r sprintf("%.3f", robust_assessment$fitted_decile_max_observed_to_fitted[robust_assessment$placement == "glasses"])`
near-eye and
`r sprintf("%.3f", robust_assessment$fitted_decile_min_observed_to_fitted[robust_assessment$placement == "chest"])`
to
`r sprintf("%.3f", robust_assessment$fitted_decile_max_observed_to_fitted[robust_assessment$placement == "chest"])`
chest. The diagnostic local-clock ratios span roughly 0.001–3.017 near-eye
and 0.002–2.970 chest, confirming strong time-of-day patterning intentionally
left to the exploratory GAM/GAMM. This does not silently add a clock term or
change the approved observed-supported-hour-weighted estimand.

```{r}
#| label: targeted-influence-table

targeted_max <- robust_deletions |>
  dplyr::slice_max(
    abs(shift_in_full_robust_se),
    n = 1L,
    with_ties = FALSE,
    by = c(placement, effect_id)
  ) |>
  dplyr::mutate(
    Placement = placement_labels[placement],
    Effect = effect_labels[effect_id],
    `Deleted participant` = excluded_participant,
    `Full ratio` = sprintf("%.2f", exp(full_estimate_link)),
    `Deleted-participant ratio` = sprintf("%.2f", exp(reduced_estimate_link)),
    `Shift / full HC1 SE` = sprintf("%.2f", shift_in_full_robust_se)
  ) |>
  dplyr::select(
    Placement, Effect, `Deleted participant`, `Full ratio`,
    `Deleted-participant ratio`, `Shift / full HC1 SE`
  )

targeted_max |>
  compact_gt(
    "Largest exact shift among five score-ranked participant deletions",
    "This is a targeted influence screen, not an exclusion rule. All refits converged at full fixed-effect rank."
  )
```

The largest targeted deletion is the chest activity coefficient after
removing `THUAS_S011`: its shift is 1.46 full-model HC1 standard errors.
Near-eye's maximum is 0.77 standard errors. No participant is excluded; this
finding is a reason to use an influence-sensitive covariance and to retain the
full deterministic influence audit in production.

All nine near-eye and eight chest leave-one-site refits also converge at full
reduced-frame rank. The maximum absolute log-coefficient changes are 0.097
near-eye and 0.145 chest. Those deletions are screening diagnostics only:
removing a site changes the equal-site estimand to the average over the
remaining sites, so they are not common-estimand sensitivity estimates.

# Reader-facing figure verification

```{r}
#| label: report-011-audit

figure_audit |>
  dplyr::transmute(
    Figure = figure_id,
    `Native size` = paste0(native_width_mm, " × ", native_height_mm, " mm"),
    `Intended width` = paste0(intended_width_mm, " mm"),
    `Scale factor` = sprintf("%.2f", scale_factor),
    `Smallest nominal text` = paste0(
      smallest_essential_nominal_text_pt, " pt"
    ),
    `Effective text` = paste0(effective_final_text_pt, " pt"),
    Status = report_011_status
  ) |>
  compact_gt(
    "REPORT-011 physical-size verification",
    figure_audit$clipping_overlap_wrapping_distortion_balance
  )
```

The figure was assessed at 170 mm on a 210 × 297 mm A4 portrait page with
20-mm side margins. The 170-mm native width gives a scale factor of 1.00 and
the smallest essential text remains 7.5 pt. The archived physical-size page is
`artifacts/12_manifests/H06/qa/H06_v0_scaffold_pilot_effects_A4_preview.pdf`.

# Recommendation and remaining gate

The pilot supports a narrow conclusion about **model development**, not H06:

1. **Do not use the V0 participant-only mixed model as the revised basis.** It
   is nonsingular but does not repair repeated participant-days, serial
   dependence, or the outcome distribution.
2. **Do not use either latent-AR Tweedie model or the BAM working-AR model.**
   Their dependence components are not jointly identifiable on these frames,
   and their generative diagnostics fail.
3. **Take the participant-cluster robust marginal log-mean model forward as
   the only viable candidate.** Preserve the exact V0 fixed-effects scaffold,
   binary activity, sum-coded site hierarchy, observed-supported-hour average,
   and no confirmatory clock effect. Use participant-cluster HC3 covariance as
   the leverage-sensitive primary candidate, with HC1 shown as a covariance
   sensitivity.
4. Before it can replace the failed Stage 2 model, run the already-required
   robust additive and site-heterogeneity tests with REPORT-008 p-values and
   the existing multiplicity families; exact near-eye/chest and paired/common
   samples; gap-timing-unaware sensitivity; full deterministic influence;
   calibration; and recreated reader outputs. Any heavy resampling remains
   separately gated at 50 or 100 pilot replicates.

The author decision is therefore whether to authorize that narrowly specified
robust-marginal production amendment. Until then, this pilot remains
feasibility evidence, the original Stage 2 adequacy failure remains in force,
no H06 result or claim is accepted, and Stage 3 remains blocked.

# Reproducibility record

- R version: 4.6.1.
- Scientific pilot:
  `scripts/hypotheses/H06/pilot_h06_v0_scaffold_models.R`.
- Robust calibration/influence checks:
  `scripts/hypotheses/H06/assess_h06_v0_scaffold_robust_candidate.R`.
- Stored BAM GAMM re-audit:
  `scripts/hypotheses/H06/audit_h06_stored_bam_models.R`.
- Reader artifact builder:
  `scripts/hypotheses/H06/build_h06_v0_scaffold_pilot_reader_artifacts.R`.
- Fitted objects:
  `artifacts/07_models/H06/H06_v0_scaffold_pilot_models.rds`.
- Strengthened BAM diagnostics:
  `artifacts/08_diagnostics/H06/H06_v0_scaffold_pilot_bam_gamm_audit.csv`
  and the paired per-sequence lag CSV.
- Every output is labelled `feasibility_pilot_not_accepted_inference` or the
  corresponding pilot-diagnostic status.
