---
title: "H08 Stage 1: audit and analysis plan"
subtitle: "Visual light sensitivity and personal light exposure"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-overflow: wrap
    df-print: paged
bibliography: ../../../bibliography.bib
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

root_candidates <- unique(normalizePath(
  c(".", "..", "../..", "../../.."),
  winslash = "/",
  mustWork = FALSE
))
root_matches <- root_candidates[file.exists(file.path(
  root_candidates,
  "preregistration",
  "AsPredicted #273407.pdf"
))]
if (length(root_matches) < 1L) {
  stop("Could not locate the project root", call. = FALSE)
}
root <- root_matches[[1L]]

project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
if (dir.exists(project_library)) {
  .libPaths(c(project_library, .libPaths()))
}

suppressPackageStartupMessages({
  library(digest)
  library(dplyr)
  library(gt)
  library(purrr)
  library(readr)
  library(rvest)
  library(stringr)
  library(tibble)
  library(tidyr)
})

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H08 Stage 1 must run under R 4.6.1", call. = FALSE)
}

h08_path <- function(...) file.path(root, ...)
h08_sha256 <- function(path) {
  digest::digest(file = path, algo = "sha256", serialize = FALSE)
}
h08_formula_text <- function(x) paste(deparse(x), collapse = " ")
h08_count_cell <- function(participants, participant_days, sites = NULL) {
  if (is.null(sites)) {
    paste0(participants, " / ", participant_days)
  } else {
    paste0(participants, " / ", participant_days, " / ", sites)
  }
}
h08_gt <- function(data, title = NULL, note = NULL, font_size = 12) {
  out <- data |>
    gt() |>
    opt_align_table_header(align = "left") |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(font_size),
      data_row.padding = px(3),
      column_labels.font.weight = "bold",
      source_notes.font.size = px(max(9, font_size - 2))
    )
  if (!is.null(title)) out <- out |> tab_header(title = md(title))
  if (!is.null(note)) out <- out |> tab_source_note(md(note))
  out
}

site_display <- read_csv(
  h08_path("config", "site_display_registry.csv"),
  show_col_types = FALSE
) |>
  arrange(display_order)
site_levels <- site_display$site

metric_ids <- c(
  "daily_geometric_mean_medi",
  "m10_mean_medi",
  "l10_mean_medi",
  "duration_above_1000",
  "duration_above_250_wake",
  "duration_below_10_pre_sleep",
  "duration_below_1_sleep_environment",
  "longest_bout_above_250",
  "dose_time_sensitive_corrected_medi"
)

metric_source_map <- tribble(
  ~metric_order, ~metric_id, ~source_column, ~v0_name,
  1L, "daily_geometric_mean_medi", "daily_geometric_mean_medi_lx", "Mean",
  2L, "m10_mean_medi", "m10_mean_medi_lx", "brightest_10h_mean",
  3L, "l10_mean_medi", "l10_mean_medi_lx", "darkest_10h_mean",
  4L, "duration_above_1000", "duration_above_1000_h", "duration_above_1000",
  5L, "duration_above_250_wake", "duration_above_250_wake_h", "duration_above_250_wake",
  6L, "duration_below_10_pre_sleep", "duration_below_10_pre_sleep_h", "duration_below_10_pre-sleep",
  7L, "duration_below_1_sleep_environment", "duration_below_1_sleep_environment_h", "duration_below_1_sleep",
  8L, "longest_bout_above_250", "longest_bout_above_250_h", "period_above_250",
  9L, "dose_time_sensitive_corrected_medi", "dose_corrected_medi_lx_h", "dose"
)

metric_display <- read_csv(
  h08_path("config", "metric_display_registry.csv"),
  show_col_types = FALSE
) |>
  filter(metric_id %in% metric_ids)

h05_metric_contract <- read_csv(
  h08_path("artifacts", "06_model_data", "H05", "H05_metric_registry.csv"),
  show_col_types = FALSE
) |>
  filter(metric_id %in% metric_ids)

metric_contract <- metric_source_map |>
  left_join(metric_display, by = "metric_id") |>
  left_join(
    h05_metric_contract |>
      select(
        metric_id,
        response_family,
        response_transform,
        effect_scale,
        diagnostic_note
      ),
    by = "metric_id"
  ) |>
  arrange(metric_order)

stopifnot(
  nrow(metric_contract) == 9L,
  !anyNA(metric_contract$manuscript_name),
  !anyNA(metric_contract$response_family)
)
```

::: {.callout-warning}
## Stage 1 gate — no new H08 inferential model has been fitted

This document audits the registered contract, reconstructs the submitted
implementation (V0), verifies inputs and candidate samples, and fixes a
proposed model and sensitivity contract. It creates formula objects and
descriptive audit summaries only. Stage 2 must not begin until the author
explicitly approves the decisions in the final section. Any later
computation-heavy resampling first requires a 50- or 100-replicate pilot and a
separate production approval.
:::

# Scope and analytical question

The scientific question is whether participant-level self-reported visual
light sensitivity is associated with participant-day personal light-exposure
metrics. The predictor is measured once per participant; the outcomes repeat
across days. Near-eye measurements are primary. Chest measurements are
complementary, and direct placement comparisons use separately fitted models
on exactly paired participants and participant-days. No placement is pooled
with another.

Temperature, latitude, age, sex, chronotype, and other unregistered predictors
are not added. Site remains a fixed adjustment term. A narrowly defined
photoperiod-adjusted sensitivity is proposed below because photoperiod varies
over measurement days and is already a prepared study variable; it is not a
new primary predictor and has no interaction with VLSQ-8.

The predefined preparation sensitivity is called the
**gap-timing-unaware dataset**. It still passed the general 50%-per-hour and
80%-per-day coverage rules. The name means that the timing of the remaining
missing observations is not used for an additional metric-specific adjustment;
it does not mean that gaps, missingness, or coverage were ignored. For this
one-time first explanation, the primary could be interpreted as a
time-sensitive primary metric dataset because its approved preparation uses
the timing of remaining missing observations where that timing is relevant.
From here onward, it is called simply the **primary dataset**.

# Registered H08 and decision provenance

## Exact registered hypothesis

The signed preregistration states:

> **H8: Duration-, exposure-history-, and level-based metrics are associated
> with VLSQ-8 light sensitivity scores.**

It registers the generic model
`Metric ~ Predictor * Site + (1 | Site:Participant)`, likelihood-ratio
testing at 0.05, false-discovery-rate control within H08, chest as the main
placement, and glasses as a robustness repeat. The approved placement decision
reverses those roles: near eye is primary and chest is complementary. This is
an explicit deviation, not a reinterpretation of the signed text.

```{r}
#| label: registered-h08-formula
#| echo: true
#| code-fold: false

registered_h08_formula <- stats::as.formula(
  "response_value ~ VLSQ8 * site + (1 | site:Id)"
)
registered_h08_formula
```

The analysis plan restates H08 as a significant association and writes
`Metric ~ VLSQ-8 * Site + (1 | Site:Participant)`. Neither document separates
the across-site average association from site heterogeneity, defines a total
VLSQ-8 score algorithm, or enumerates the two nine-member multiplicity
vectors. Those choices are therefore made explicit at this gate.

## Evidence and controlling decisions read for this audit

```{r}
evidence_inventory <- tribble(
  ~evidence, ~role,
  "preregistration/AsPredicted #273407.pdf", "Signed H08 wording, formula, placement, exclusions, and FDR statement",
  "preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx", "Expanded H08 wording and intended site interaction",
  "audit/evidence/preregistration_contract.md", "Extracted hypothesis contract",
  "_deviations.qmd; audit/ledgers/deviation_register.csv", "Registered-versus-implemented deviations",
  "audit/hypotheses/H03-H11_gated_workflow.qmd", "Four-stage gate and no-result-seeking Stage 1 rule",
  "audit/hypotheses/H05-H08_migration_map.md", "V0 H08 reconstruction and migration decisions",
  "audit/decisions/model_reporting.md", "Plain language, 95% intervals, exact samples, p-value and placement reporting",
  "audit/decisions/p_value_display_conventions.md", "REPORT-008 p-value display",
  "audit/decisions/paired_placement_comparison_display.md", "REPORT-009 paired-placement display",
  "audit/decisions/site_display_conventions.md", "Submitted site names, order, and colours",
  "audit/decisions/gap_timing_unaware_dataset_terminology.md", "REPORT-010 sensitivity terminology",
  "audit/decisions/bootstrap_execution_policy.md", "Pilot-first heavy resampling",
  "audit/decisions/metric_validity.md", "Metric admissibility and construct qualifications",
  "audit/decisions/metric_implementation_parameters.md", "Fixed grid, support, gap, dose, and period rules",
  "audit/decisions/daily_coverage_denominator.md", "50%-hour/80%-day Rule A primary and registered alternatives",
  "audit/decisions/all_zero_medi_day_exclusion.md", "Exact all-zero day screen",
  "audit/decisions/saturation_boundary.md", "Strict melEDI operating boundary",
  "audit/decisions/minute_aggregation.md", "Source-epoch-aware one-minute values",
  "config/metric_display_registry.csv", "Submitted metric names and units",
  "config/site_display_registry.csv", "Submitted site names, order, and colours",
  "RQ3.qmd; RQ3_chest.qmd; scripts/fitting.R", "Executable V0 implementation",
  "docs/RQ3.html; docs/RQ3_chest.html", "Rendered V0 H08 result strings",
  "manuscript/reviewer_comments_NatMed.rtf; audit/ledgers/reviewer_feedback_disposition.csv", "Hierarchy, interaction-power, multiplicity, placement, and claim constraints"
)

evidence_inventory |>
  h08_gt(
    title = "Evidence and decisions controlling H08 Stage 1",
    note = "Scientific computations below use R 4.6.1; document and code inspection does not alter author files.",
    font_size = 11
  ) |>
  cols_width(evidence ~ px(330), role ~ px(650))
```

## Direct analytical inputs and integrity checks

```{r}
normalization_manifest <- read_csv(
  h08_path("artifacts", "12_manifests", "model_input_normalization.csv"),
  show_col_types = FALSE
)
base_manifest <- read_csv(
  h08_path("artifacts", "12_manifests", "base_model_data_artifacts.csv"),
  show_col_types = FALSE
)
gap_manifest <- read_csv(
  h08_path("artifacts", "12_manifests", "manuscript_prepared_data_artifacts.csv"),
  show_col_types = FALSE
)

direct_inputs <- bind_rows(
  normalization_manifest |>
    filter(artifact_id == "normalized_vlsq8") |>
    transmute(
      input = "Normalized VLSQ-8",
      path,
      expected_sha256 = sha256,
      producer,
      producer_r = r_version
    ),
  base_manifest |>
    filter(artifact_id %in% c(
      "glasses_participant_day_enriched",
      "chest_participant_day_enriched"
    )) |>
    transmute(
      input = if_else(
        artifact_id == "glasses_participant_day_enriched",
        "Primary near-eye participant-day metrics",
        "Complementary chest participant-day metrics"
      ),
      path,
      expected_sha256 = sha256,
      producer,
      producer_r = r_version
    ),
  gap_manifest |>
    filter(artifact_id == "participant_day_metrics_rds") |>
    transmute(
      input = "Gap-timing-unaware participant-day metrics",
      path,
      expected_sha256 = sha256,
      producer,
      producer_r = "4.6.1"
    )
) |>
  mutate(
    absolute_path = file.path(root, path),
    exists = file.exists(absolute_path),
    observed_sha256 = map_chr(absolute_path, h08_sha256),
    status = if_else(
      exists & observed_sha256 == expected_sha256,
      "PASS",
      "FAIL"
    )
  )

stopifnot(all(direct_inputs$status == "PASS"))

direct_inputs |>
  transmute(
    Input = input,
    Path = path,
    `SHA-256` = observed_sha256,
    Producer = producer,
    `Producer R` = producer_r,
    Status = status
  ) |>
  h08_gt(
    title = "Direct H08 analytical inputs",
    note = "Observed file hashes exactly match their preparation manifests.",
    font_size = 10
  ) |>
  cols_width(
    Input ~ px(210),
    Path ~ px(360),
    `SHA-256` ~ px(390),
    Producer ~ px(300),
    `Producer R` ~ px(80),
    Status ~ px(65)
  )
```

The normalized VLSQ-8 participant key audit passes at all nine sites, and the
normalization value-preservation audit reports no value or missingness
differences for any VLSQ-8 field. Base-model joins are many-to-one on
`site + Id`, preserve all participant-day keys, and report no unmatched rows.
The metric registry firewall also passes without recalculating outcomes.

# VLSQ-8 scale and scoring audit

## Raw-source preservation

```{r}
acquisition_manifest <- read_csv(
  h08_path("artifacts", "12_manifests", "model_input_acquisition.csv"),
  show_col_types = FALSE
) |>
  filter(modality == "vlsq8") |>
  arrange(match(site, site_levels))

read_source_vlsq <- function(cache_path, site_label) {
  environment <- new.env(parent = emptyenv())
  object_names <- load(cache_path, envir = environment)
  if (!identical(object_names, "vlsq8")) {
    stop("Unexpected VLSQ-8 source object", call. = FALSE)
  }
  environment[["vlsq8"]] |>
    mutate(site = site_label, .before = 1L)
}

source_vlsq <- map2_dfr(
  file.path(root, acquisition_manifest$cache_path),
  acquisition_manifest$site,
  read_source_vlsq
)
vlsq <- readRDS(h08_path(
  "artifacts", "06_model_data", "normalized_inputs", "vlsq8.rds"
))

item_names <- c(
  "sensitivity",
  "glare",
  "flicker",
  "sensitivity_severity",
  "headache",
  "blurry_vision",
  "ability",
  "glasses"
)
comparison_columns <- c("VLSQ8", item_names)

source_normalized_comparison <- source_vlsq |>
  select(site, Id, all_of(comparison_columns)) |>
  mutate(across(all_of(item_names), as.character)) |>
  left_join(
    vlsq |>
      select(site, Id, all_of(comparison_columns)) |>
      mutate(across(all_of(item_names), as.character)),
    by = c("site", "Id"),
    suffix = c("_source", "_normalized")
  )

field_differences <- map_int(comparison_columns, function(field) {
  source_value <- source_normalized_comparison[[paste0(field, "_source")]]
  normalized_value <- source_normalized_comparison[[paste0(field, "_normalized")]]
  sum(!(source_value == normalized_value | (is.na(source_value) & is.na(normalized_value))))
})

source_hash_audit <- acquisition_manifest |>
  transmute(
    site,
    source_commit = commit,
    cache_path,
    declared_sha256 = sha256,
    observed_sha256 = map_chr(file.path(root, cache_path), h08_sha256),
    rows = map_int(file.path(root, cache_path), function(path) {
      environment <- new.env(parent = emptyenv())
      object_names <- load(path, envir = environment)
      nrow(environment[[object_names[[1L]]]])
    }),
    status = if_else(declared_sha256 == observed_sha256, "PASS", "FAIL")
  )

stopifnot(
  nrow(vlsq) == 184L,
  nrow(source_vlsq) == 184L,
  !anyDuplicated(vlsq[c("site", "Id")]),
  all(field_differences == 0L),
  all(source_hash_audit$status == "PASS")
)

source_hash_audit |>
  left_join(site_display, by = "site") |>
  arrange(display_order) |>
  transmute(
    Site = display_name,
    `Pinned commit` = source_commit,
    Rows = rows,
    `Source SHA-256` = observed_sha256,
    Status = status
  ) |>
  h08_gt(
    title = "Pinned VLSQ-8 source files",
    note = "All nine cached source objects match the acquisition manifest; normalized score and item values reproduce them exactly.",
    font_size = 10
  ) |>
  cols_width(
    Site ~ px(170),
    `Pinned commit` ~ px(300),
    Rows ~ px(60),
    `Source SHA-256` ~ px(420),
    Status ~ px(65)
  )
```

## Reconstructed score rule

The source article introduced the eight questions but explicitly reported no
standard total-score algorithm; it analysed items after coding their five
ordered categories from 1 to 5 [@verriotto2017]. The present data nevertheless
contain a numeric field labelled “VLSQ-8 sum score (light sensitivity).” The
following audit reconstructs its relationship to the eight preserved items.

```{r}
vlsq_scoring_audit <- vlsq |>
  mutate(
    item_sum_1_to_5 = rowSums(across(all_of(item_names), as.integer)),
    stored_minus_item_sum = VLSQ8 - item_sum_1_to_5
  )

item_audit <- tibble(item = item_names) |>
  mutate(
    item_number = row_number(),
    item_label = map_chr(item, function(name) attr(vlsq[[name]], "label")),
    ordered_categories = map_chr(
      item,
      function(name) paste(levels(vlsq[[name]]), collapse = " | ")
    ),
    missing = map_int(item, function(name) sum(is.na(vlsq[[name]])))
  )

score_summary <- tibble(
  check = c(
    "Participants",
    "Stored-score missing values",
    "Item missing values across 1,472 cells",
    "Observed stored-score range",
    "Observed eight-item 1–5 sum range",
    "Unique stored-minus-sum values",
    "Stored score mean",
    "Stored score participant SD"
  ),
  result = c(
    nrow(vlsq),
    sum(is.na(vlsq$VLSQ8)),
    sum(is.na(vlsq[item_names])),
    paste(range(vlsq$VLSQ8), collapse = "–"),
    paste(range(vlsq_scoring_audit$item_sum_1_to_5), collapse = "–"),
    paste(unique(vlsq_scoring_audit$stored_minus_item_sum), collapse = ", "),
    sprintf("%.4f", mean(vlsq$VLSQ8)),
    sprintf("%.4f", sd(vlsq$VLSQ8))
  )
)

stopifnot(
  all(vlsq_scoring_audit$stored_minus_item_sum == 5),
  sum(is.na(vlsq$VLSQ8)) == 0L,
  sum(is.na(vlsq[item_names])) == 0L
)

score_summary |>
  h08_gt(
    title = "VLSQ-8 score reconstruction",
    note = "Every source and normalized row satisfies stored VLSQ8 = sum of the eight current 1–5 item codes + 5.",
    font_size = 11
  ) |>
  cols_width(check ~ px(440), result ~ px(220))
```

```{r}
item_audit |>
  transmute(
    Item = item_number,
    Question = item_label,
    `Preserved category order` = ordered_categories,
    Missing = missing
  ) |>
  h08_gt(
    title = "Preserved VLSQ-8 item contract",
    note = "Higher category codes indicate more frequent or more severe symptoms. No item is reverse-coded in the current sources.",
    font_size = 10
  ) |>
  cols_width(
    Item ~ px(45),
    Question ~ px(510),
    `Preserved category order` ~ px(420),
    Missing ~ px(65)
  )
```

::: {.callout-important}
## Scoring decision required

The stored total is internally reproducible and preserved identically from all
nine pinned sources, but its constant `+5` is not justified by the cited
instrument article or by a scoring artifact in this project. A constant shift
does not change a centered per-point slope or its test, but it changes the
reported scale, possible range, and reference value. Stage 2 should use the
stored `VLSQ8` only if the author confirms its upstream scoring provenance and
intended interpretation. Otherwise the scoring producer must be corrected and
re-pinned before fitting H08.
:::

For a common, placement-independent reference, the proposed centering constant
is the mean of all 184 pinned questionnaire rows:
`r sprintf("%.4f", mean(vlsq$VLSQ8))` points. The proposed one-participant-SD
contrast is `r sprintf("%.4f", sd(vlsq$VLSQ8))` points. No within-site
standardization is used.

## Participant-level score support

```{r}
main_near_eye <- readRDS(h08_path(
  "artifacts", "06_model_data", "base",
  "metrics_glasses_participant_day_enriched.rds"
))
main_chest <- readRDS(h08_path(
  "artifacts", "06_model_data", "base",
  "metrics_chest_participant_day_enriched.rds"
))

verify_vlsq_join <- function(data) {
  comparison <- data |>
    distinct(site, Id, VLSQ8) |>
    left_join(
      vlsq |> select(site, Id, normalized_VLSQ8 = VLSQ8),
      by = c("site", "Id")
    )
  all(!is.na(comparison$normalized_VLSQ8)) &&
    all(comparison$VLSQ8 == comparison$normalized_VLSQ8)
}

stopifnot(
  !anyDuplicated(main_near_eye[c("site", "Id", "local_date")]),
  !anyDuplicated(main_chest[c("site", "Id", "local_date")]),
  verify_vlsq_join(main_near_eye),
  verify_vlsq_join(main_chest)
)

score_site_support <- bind_rows(
  main_near_eye |> mutate(placement = "Near eye"),
  main_chest |> mutate(placement = "Chest")
) |>
  group_by(placement, site) |>
  summarise(
    participants = n_distinct(Id),
    participant_days = n(),
    score_min = min(VLSQ8),
    score_max = max(VLSQ8),
    score_sd = sd(unique(VLSQ8)),
    photoperiod_min = min(photoperiod_hours),
    photoperiod_max = max(photoperiod_hours),
    .groups = "drop"
  ) |>
  left_join(site_display, by = "site") |>
  arrange(factor(placement, levels = c("Near eye", "Chest")), display_order)

score_site_support |>
  transmute(
    Placement = placement,
    Site = display_name,
    Participants = participants,
    `Participant-days` = participant_days,
    `VLSQ-8 range` = paste0(score_min, "–", score_max),
    `VLSQ-8 SD` = sprintf("%.2f", score_sd),
    `Photoperiod range (h)` = sprintf("%.2f–%.2f", photoperiod_min, photoperiod_max)
  ) |>
  h08_gt(
    title = "Participant-level predictor support in the prepared placement datasets",
    note = "Days repeat outcomes; they do not create new independent VLSQ-8 observations. The near-eye UCR interaction cell has only six participants.",
    font_size = 10
  ) |>
  tab_row_group(label = "Near eye — primary", rows = Placement == "Near eye") |>
  tab_row_group(label = "Chest — complementary", rows = Placement == "Chest") |>
  cols_hide(Placement) |>
  cols_width(
    Site ~ px(190),
    Participants ~ px(90),
    `Participant-days` ~ px(105),
    `VLSQ-8 range` ~ px(105),
    `VLSQ-8 SD` ~ px(85),
    `Photoperiod range (h)` ~ px(150)
  )
```

The site interaction is therefore a secondary heterogeneity question. It has
nine site-specific slopes in the primary near-eye analysis but only six
independent UCR participants; the chest and paired analyses omit MPI. A full
interaction that is rank deficient, fails to converge, has an unidentified
variance component, or yields effectively unbounded site-slope uncertainty is
reported as non-estimable rather than simplified after seeing its result.

# Admissible outcomes and response models

The nine registered metric concepts remain admissible after the approved
preparation repairs. Whole-day metrics describe a hybrid record: near-eye
measurement during wake and pre-sleep, and bedside sleep-environment
measurement during diary-defined sleep. They are not continuous 24-hour
ocular-exposure metrics.

```{r}
metric_sensitivity <- tribble(
  ~metric_id, ~primary_value, ~fixed_sensitivity,
  "daily_geometric_mean_medi", "Supported hybrid-day geometric mean; no coverage scaling", "Response-family alternative only after a new gate",
  "m10_mean_medi", "Brightest supported 10 h geometric mean", "Response-family alternative only after a new gate",
  "l10_mean_medi", "Darkest supported midnight-wrapping 10 h geometric mean", "Zero-mass and common-family diagnostic",
  "duration_above_1000", "Actual supported time above 1,000 lx melEDI", "Physical-bound and response diagnostics",
  "duration_above_250_wake", "Actual supported waking near-eye/chest time above 250 lx melEDI", "State-support diagnostic",
  "duration_below_10_pre_sleep", "Calendar-day cumulative supported pre-sleep time below 10 lx melEDI", "Six-hour audit warning; no truncation",
  "duration_below_1_sleep_environment", "Supported bedside sleep-environment time below 1 lx melEDI", "Sleep-support and response diagnostics",
  "longest_bout_above_250", "Observed lower bound for the longest continuous period above 250 lx melEDI", "Exactly identifiable days only",
  "dose_time_sensitive_corrected_medi", "Time-sensitive corrected melEDI dose", "Observed dose on a common sample"
)

metric_contract_display <- metric_contract |>
  left_join(metric_sensitivity, by = "metric_id") |>
  transmute(
    Metric = manuscript_name,
    Category = manuscript_category,
    Unit = display_unit,
    `Primary prepared value` = primary_value,
    `Response model` = case_when(
      response_family == "gaussian" & response_transform == "log10_offset_0.1" ~
        "Gaussian; log10(value + 0.1)",
      response_family == "gaussian" & response_transform == "identity" ~
        "Gaussian; identity",
      response_family == "tweedie_log" ~ "Tweedie; log link",
      TRUE ~ paste(response_family, response_transform)
    ),
    `Practical effect` = case_when(
      effect_scale == "ratio" ~ "Ratio",
      TRUE ~ "Difference"
    ),
    `Fixed diagnostic or sensitivity` = fixed_sensitivity
  )

metric_contract_display |>
  h08_gt(
    title = "Proposed H08 metric and response-family contract",
    note = "The response package is inherited from the current H01/H05 metric contract. A family/link switch is a major gate and must apply consistently across datasets, placements, and samples.",
    font_size = 9
  ) |>
  cols_width(
    Metric ~ px(220),
    Category ~ px(105),
    Unit ~ px(65),
    `Primary prepared value` ~ px(340),
    `Response model` ~ px(170),
    `Practical effect` ~ px(85),
    `Fixed diagnostic or sensitivity` ~ px(260)
  )
```

The V0 Tweedie model for time below 10 lx melEDI before sleep is not retained:
the current cross-hypothesis response contract uses an identity-scale Gaussian
model after the H01 candidate assessment. No family is selected from the H08
association result.

## Metric-specific missingness in the primary prepared data

```{r}
read_admissibility <- function(path, placement_label) {
  read_csv(path, show_col_types = FALSE) |>
    filter(
      analysis_unit == "participant_day",
      metric %in% metric_ids
    ) |>
    mutate(placement = placement_label)
}

admissibility <- bind_rows(
  read_admissibility(
    h08_path("artifacts", "05_metrics", "metrics_glasses_admissibility.csv"),
    "Near eye"
  ),
  read_admissibility(
    h08_path("artifacts", "05_metrics", "metrics_chest_admissibility.csv"),
    "Chest"
  )
)

admissibility_summary <- admissibility |>
  group_by(placement, metric) |>
  summarise(
    prepared_days = n(),
    estimable_days = sum(estimable),
    unavailable_days = sum(!estimable),
    failure_reasons = {
      reasons <- sort(table(failure_reason[!estimable]), decreasing = TRUE)
      if (length(reasons) == 0L) "None" else
        paste0(names(reasons), ": ", as.integer(reasons), collapse = "; ")
    },
    .groups = "drop"
  ) |>
  rename(metric_id = metric) |>
  left_join(metric_contract |> select(metric_id, metric_order, manuscript_name), by = "metric_id") |>
  arrange(metric_order, factor(placement, levels = c("Near eye", "Chest")))

admissibility_summary |>
  transmute(
    Metric = manuscript_name,
    Placement = placement,
    `Prepared days` = prepared_days,
    `Candidate days` = estimable_days,
    Unavailable = unavailable_days,
    `Reason-coded unavailable days` = failure_reasons
  ) |>
  h08_gt(
    title = "Metric-specific availability before VLSQ-8 model fitting",
    note = "A failed state, support, or dose rule makes only that metric unavailable; it does not remove the participant-day from other outcomes.",
    font_size = 10
  ) |>
  cols_width(
    Metric ~ px(245),
    Placement ~ px(85),
    `Prepared days` ~ px(85),
    `Candidate days` ~ px(90),
    Unavailable ~ px(75),
    `Reason-coded unavailable days` ~ px(390)
  )
```

For the longest continuous period, all 816 near-eye and 902 chest days have an
observed lower bound, but only 500 near-eye and 564 chest days are exactly
identifiable. For melEDI dose, observed dose exists on all prepared days;
time-sensitive corrected dose exists on 761 near-eye and 851 chest days after
the fixed relevance-coverage guard. These differences are built into the
named common-sample metric sensitivities.

# Exact V0 reconstruction

## V0 formula objects and tests

```{r}
#| label: v0-h08-formulas
#| echo: true
#| code-fold: false

v0_h08_formulas <- list(
  H8_1_full = stats::as.formula(
    "response_value ~ site * VLSQ8 + (1 | Id)"
  ),
  H8_0_site_only = stats::as.formula(
    "response_value ~ site + (1 | Id)"
  ),
  H8_ni_additive = stats::as.formula(
    "response_value ~ site + VLSQ8 + (1 | Id)"
  ),
  H8_ns_vlsq_only = stats::as.formula(
    "response_value ~ VLSQ8 + (1 | Id)"
  )
)

v0_h08_formula_registry <- enframe(
  v0_h08_formulas,
  name = "formula_id",
  value = "formula"
) |>
  mutate(formula = map_chr(formula, h08_formula_text))

v0_h08_formula_registry
```

V0 compared:

1. `H8_0_site_only` with `H8_1_full` and displayed that joint addition of the
   VLSQ-8 main effect plus all site interactions as “the association”;
2. `H8_ni_additive` with `H8_1_full` for interaction heterogeneity; and
3. `H8_ns_vlsq_only` with `H8_ni_additive` for a site block.

The first comparison is not an across-site average association test. V0 also
claimed sum coding but neither assigned contrasts nor stored a contrast
attribute; R 4.5.0 used the session's default treatment coding. The random
intercept omitted the explicit `site:` nesting even though the registration
included it.

## V0 response and sample reconstruction

```{r}
v0_log_metrics <- c(
  "daily_geometric_mean_medi",
  "m10_mean_medi",
  "l10_mean_medi",
  "longest_bout_above_250",
  "dose_time_sensitive_corrected_medi"
)

read_v0_metrics <- function(path, placement_label) {
  environment <- new.env(parent = emptyenv())
  object_names <- load(path, envir = environment)
  metric_list <- environment[[object_names[[1L]]]]

  metric_list |>
    inner_join(metric_source_map, by = c("name" = "v0_name")) |>
    transmute(
      placement = placement_label,
      metric_order,
      metric_id,
      data = map2(data, metric_id, function(frame, id) {
        frame <- frame |>
          left_join(vlsq |> select(site, Id, VLSQ8), by = c("site", "Id"))
        frame$response_value <- if (id %in% v0_log_metrics) {
          log10(frame$metric + 0.1)
        } else {
          frame$metric
        }
        frame
      })
    ) |>
    unnest(data)
}

v0_data <- bind_rows(
  read_v0_metrics(h08_path("data", "metrics_glasses.RData"), "Near eye"),
  read_v0_metrics(h08_path("data", "metrics_chest.RData"), "Chest")
)

v0_sample <- v0_data |>
  filter(!is.na(response_value), !is.na(VLSQ8)) |>
  group_by(placement, metric_id, metric_order) |>
  summarise(
    participants = n_distinct(site, Id),
    participant_days = n(),
    sites = n_distinct(site),
    .groups = "drop"
  ) |>
  left_join(metric_contract |> select(metric_id, manuscript_name), by = "metric_id") |>
  mutate(cell = h08_count_cell(participants, participant_days, sites)) |>
  select(metric_order, Metric = manuscript_name, placement, cell) |>
  pivot_wider(names_from = placement, values_from = cell) |>
  arrange(metric_order)

v0_sample |>
  select(-metric_order) |>
  h08_gt(
    title = "Exact V0 candidate model frames after the VLSQ-8 join",
    note = "Cells are participants / participant-days / sites. V0 did not report these exact frames in its result table.",
    font_size = 10
  ) |>
  cols_width(Metric ~ px(290), `Near eye` ~ px(150), Chest ~ px(150))
```

V0 used Gaussian models after `log10(metric + 0.1)` for Mean melEDI, the
brightest and darkest 10 h means, the longest period, and dose. It used
Tweedie log-link models for all four state/threshold durations. It relied on
default `lmer()` estimation and `glmmTMB()` likelihoods, did not store exact
model objects, did not provide 95% confidence intervals, and inherited
diagnostics from a different hypothesis rather than interpreting H08 fits.

## V0 multiplicity and rendered results

The V0 helper extracted one p-value at a time and then called
`p.adjust(p, method = "fdr", n = 9)`. Because each call received a scalar,
this is not a nine-value Benjamini--Hochberg adjustment; it is equivalent to
`min(9 * p, 1)` for each metric. The raw p-values were not persisted.

```{r}
extract_v0_h08_result <- function(path, placement_label) {
  document <- read_html(path)
  tables <- html_elements(document, "table")
  titles <- map_chr(tables, function(table) {
    paste(html_text2(html_elements(table, ".gt_title")), collapse = " ")
  })
  table_index <- which(str_detect(titles, fixed("H8: Model results")))
  if (length(table_index) != 1L) {
    stop("Could not uniquely identify the V0 H08 result table", call. = FALSE)
  }
  result <- html_table(tables[[table_index]], fill = TRUE)
  result <- result[result[[1L]] %in% c("level", "duration", "exposure history"), 1:4]
  names(result) <- c("metric_type", "v0_name", "displayed_adjusted_p", "supported")
  result |>
    mutate(placement = placement_label) |>
    left_join(metric_source_map, by = "v0_name")
}

v0_results <- bind_rows(
  extract_v0_h08_result(h08_path("docs", "RQ3.html"), "Near eye"),
  extract_v0_h08_result(h08_path("docs", "RQ3_chest.html"), "Chest")
) |>
  left_join(metric_contract |> select(metric_id, manuscript_name), by = "metric_id") |>
  arrange(metric_order, factor(placement, levels = c("Near eye", "Chest")))

v0_results |>
  transmute(
    Metric = manuscript_name,
    Placement = placement,
    `V0 displayed scalar-adjusted p` = displayed_adjusted_p,
    `V0 supported` = supported
  ) |>
  h08_gt(
    title = "Only recoverable V0 H08 result precision",
    note = "These are rendered strings, not full-precision p-values. They do not comply with REPORT-008 and cannot support a numerical before/after estimate comparison.",
    font_size = 10
  ) |>
  cols_width(
    Metric ~ px(300),
    Placement ~ px(90),
    `V0 displayed scalar-adjusted p` ~ px(185),
    `V0 supported` ~ px(110)
  )
```

The manuscript statement that all H08 p-values exceeded 0.9 is not supported
by the chest rendering, which shows 0.4, 0.7, and 0.3 for three outcomes. More
importantly, neither placement's V0 display tested the proposed across-site
average estimand with a proper nine-value BH family. The submitted absence-of-
association claim is therefore reopened pending Stage 2.

## V0 defects that Stage 2 must resolve

```{r}
v0_findings <- tribble(
  ~finding, ~consequence, ~Stage2_resolution,
  "Joint VLSQ-8 plus interaction block displayed as the association", "Different null from an across-site average slope", "Primary additive-versus-site-only test; interaction tested separately",
  "Scalar p.adjust calls with n = 9", "Not vector-wide BH", "Assemble all nine raw p-values, then adjust once",
  "Claimed sum coding, default treatment coding used", "Main slope did not have the claimed equal-site meaning", "Explicit submitted site levels and contr.sum",
  "Random effect written as (1 | Id)", "Registered nesting not explicit", "Use (1 | site:Id) and verify key uniqueness",
  "No scale-scoring audit", "VLSQ-8 total provenance unresolved", "Author-confirm score rule before fitting",
  "No exact fitted samples", "Prepared counts substituted for model frames", "Persist and report exact rows, participants, days, and sites",
  "No estimates or 95% confidence intervals", "Magnitude and uncertainty unavailable", "Report raw-point, one-SD, and practical-scale effects with 95% CIs",
  "No H08 diagnostics", "Family, interaction, influence, and serial dependence unverified", "Run and interpret the declared H08 diagnostic battery",
  "Chest was preregistered primary", "Current placement roles differ from registration", "Disclose near-eye primary/chest complementary deviation",
  "Metric implementation predates approved repairs", "Metric values and availability differ", "Recreate V0 and compare with the repaired primary and gap-timing-unaware datasets"
)

v0_findings |>
  h08_gt(
    title = "V0-to-audited H08 resolution map",
    font_size = 10
  ) |>
  cols_width(
    finding ~ px(300),
    consequence ~ px(330),
    Stage2_resolution ~ px(440)
  )
```

# Proposed estimands, models, and tests

## Estimand hierarchy

1. **Primary near-eye estimand:** the equal-site average association between
   VLSQ-8 and each daily metric, conditional on fixed site, expressed per raw
   score point and per one participant-level SD. This is a between-participant
   association; repeated days improve outcome precision but do not increase
   the number of independent VLSQ-8 observations.
2. **Registered secondary estimand:** whether the VLSQ-8 slope differs among
   sites. The global interaction test is reported before any site-specific
   slope. Site-specific slopes receive estimates and 95% confidence intervals
   but no separate significance screen by default.
3. **Complementary chest estimand:** the same two questions among the eight
   chest sites, in separately named BH families. It cannot override the
   primary near-eye conclusion.
4. **Paired placement estimand:** near-eye and chest effects fitted separately
   on identical eight-site participant-days, for a direct concordance display.
   This is not pooling and is not an equivalence test.

## Exact proposed formula objects — no fits

`VLSQ8_c` is the stored score minus
`r sprintf("%.4f", mean(vlsq$VLSQ8))`. `photoperiod_c` is daily photoperiod in
hours minus 12. `participant_response` is the same approved response transform
or link applied to a participant's arithmetic mean of admissible natural-scale
daily metric values; it is used only for the unequal-day-weighting sensitivity.

```{r}
#| label: proposed-h08-formulas
#| echo: true
#| code-fold: false

proposed_h08_formulas <- list(
  primary_site_only = stats::as.formula(
    "response_value ~ site + (1 | site:Id)"
  ),
  primary_additive_vlsq = stats::as.formula(
    "response_value ~ site + VLSQ8_c + (1 | site:Id)"
  ),
  secondary_site_interaction = stats::as.formula(
    "response_value ~ site * VLSQ8_c + (1 | site:Id)"
  ),
  photoperiod_site_only = stats::as.formula(
    "response_value ~ site + photoperiod_c + (1 | site:Id)"
  ),
  photoperiod_additive_vlsq = stats::as.formula(
    "response_value ~ site + photoperiod_c + VLSQ8_c + (1 | site:Id)"
  ),
  photoperiod_site_interaction = stats::as.formula(
    "response_value ~ site * VLSQ8_c + photoperiod_c + (1 | site:Id)"
  ),
  participant_site_only = stats::as.formula(
    "participant_response ~ site"
  ),
  participant_additive_vlsq = stats::as.formula(
    "participant_response ~ site + VLSQ8_c"
  ),
  participant_site_interaction = stats::as.formula(
    "participant_response ~ site * VLSQ8_c"
  )
)

proposed_h08_formula_registry <- enframe(
  proposed_h08_formulas,
  name = "formula_id",
  value = "formula"
) |>
  mutate(formula = map_chr(formula, h08_formula_text))

proposed_h08_formula_registry
```

The primary, chest, paired, gap-timing-unaware, observed-dose, and exact-period
analyses all reuse the applicable `primary_site_only`,
`primary_additive_vlsq`, and `secondary_site_interaction` objects. Only the
input rows or declared outcome variant changes. Leave-one-site-out analyses
reuse the additive formulas after removing one whole site at a time.

## Contrasts, likelihoods, and comparison registry

```{r}
contrast_contract <- tibble(
  analysis = c("Primary near eye", "Complementary chest", "Paired/common sample"),
  site_levels = c(
    paste(site_display$display_name, collapse = " | "),
    paste(site_display$display_name[site_display$site != "MPI"], collapse = " | "),
    paste(site_display$display_name[site_display$site != "MPI"], collapse = " | ")
  ),
  contrast = c("contr.sum(9)", "contr.sum(8)", "contr.sum(8)"),
  slope_interpretation = c(
    "Equal average over nine submitted sites",
    "Equal average over the eight chest sites",
    "Equal average over the same eight paired sites"
  )
)

contrast_contract |>
  h08_gt(
    title = "Explicit site contrast contract",
    note = "Stage 2 assigns factor levels and contrast matrices directly; it does not rely on the session default.",
    font_size = 10
  ) |>
  cols_width(
    analysis ~ px(165),
    site_levels ~ px(620),
    contrast ~ px(95),
    slope_interpretation ~ px(300)
  )
```

```{r}
comparison_registry <- tribble(
  ~comparison_id, ~full_formula, ~reduced_formula, ~question, ~inferential_role,
  "average_vlsq", "primary_additive_vlsq", "primary_site_only", "Does a site-adjusted average VLSQ-8 association improve fit?", "Primary",
  "site_heterogeneity", "secondary_site_interaction", "primary_additive_vlsq", "Do VLSQ-8 slopes differ among sites?", "Registered secondary",
  "photoperiod_average", "photoperiod_additive_vlsq", "photoperiod_site_only", "Is the average association stable after linear photoperiod adjustment?", "Descriptive sensitivity",
  "photoperiod_heterogeneity", "photoperiod_site_interaction", "photoperiod_additive_vlsq", "Is slope heterogeneity stable after photoperiod adjustment?", "Descriptive sensitivity",
  "participant_average", "participant_additive_vlsq", "participant_site_only", "Is the average association stable with one outcome summary per participant?", "Descriptive sensitivity",
  "participant_heterogeneity", "participant_site_interaction", "participant_additive_vlsq", "Is heterogeneity stable with one outcome summary per participant?", "Descriptive sensitivity"
) |>
  mutate(
    full = map_chr(full_formula, ~ h08_formula_text(proposed_h08_formulas[[.x]])),
    reduced = map_chr(reduced_formula, ~ h08_formula_text(proposed_h08_formulas[[.x]]))
  )

comparison_registry |>
  transmute(
    Comparison = comparison_id,
    `Full formula` = full,
    `Reduced formula` = reduced,
    Question = question,
    Role = inferential_role
  ) |>
  h08_gt(
    title = "Nested comparison contract",
    note = "Gaussian fixed-effect comparisons use maximum likelihood (REML = FALSE); final Gaussian estimates may use REML after the structure is fixed. Tweedie-log comparisons and estimates use maximum likelihood.",
    font_size = 9
  ) |>
  cols_width(
    Comparison ~ px(145),
    `Full formula` ~ px(315),
    `Reduced formula` ~ px(285),
    Question ~ px(350),
    Role ~ px(145)
  )
```

Every nested comparison uses exactly the same rows, response definition,
transformation, family, link, contrast matrix, random-effect structure, and
optimizer settings. A missing full or reduced fit leaves the planned family
member as `NA` with a reason; it does not shorten the family.

## Multiplicity families

```{r}
#| label: h08-multiplicity-function
#| echo: true
#| code-fold: false

h08_adjust_nine <- function(raw_p) {
  stopifnot(length(raw_p) == 9L)
  stats::p.adjust(raw_p, method = "BH", n = 9L)
}

multiplicity_registry <- tribble(
  ~family_id, ~placement_or_scenario, ~comparison, ~planned_n, ~role,
  "H08-F1-main-near-eye-average", "Primary near eye", "average_vlsq", 9L, "Primary decision family",
  "H08-F2-main-near-eye-heterogeneity", "Primary near eye", "site_heterogeneity", 9L, "Registered secondary family",
  "H08-C1-main-chest-average", "Complementary chest", "average_vlsq", 9L, "Complementary family",
  "H08-C2-main-chest-heterogeneity", "Complementary chest", "site_heterogeneity", 9L, "Complementary family",
  "H08-G1-gap-near-eye-average", "Gap-timing-unaware near eye", "average_vlsq", 9L, "Sensitivity family",
  "H08-G2-gap-near-eye-heterogeneity", "Gap-timing-unaware near eye", "site_heterogeneity", 9L, "Sensitivity family",
  "H08-GC1-gap-chest-average", "Gap-timing-unaware chest", "average_vlsq", 9L, "Complementary sensitivity family",
  "H08-GC2-gap-chest-heterogeneity", "Gap-timing-unaware chest", "site_heterogeneity", 9L, "Complementary sensitivity family"
)

multiplicity_registry
```

Each listed family is assembled as a complete nine-row registry and adjusted
once. Raw and BH-adjusted p-values remain at full precision in source
artifacts; displays use REPORT-008: a leading zero and three decimals, with
values strictly below 0.001 shown as `<0.001`. Significance is determined from
full precision and only the value governed by the declared rule is bold.

Paired/common-sample, photoperiod, participant-summary, observed-dose,
exact-period, leave-one-site-out, and other diagnostic sensitivities report
estimates and 95% confidence intervals without a second significance screen.
If inferential p-values are later requested for any of them, that requires a
new, separately named complete nine-member family approved before fitting.

## Practical estimands and intervals

```{r}
practical_estimands <- tribble(
  ~estimand_id, ~reader_facing_quantity, ~calculation_scale, ~interval,
  "raw_point", "Change per one VLSQ-8 point", "Difference for identity Gaussian; 10^beta ratio for log10(value + 0.1); exp(beta) ratio for Tweedie-log", "Two-sided 95% model interval",
  "participant_sd", paste0("Change per ", sprintf("%.4f", sd(vlsq$VLSQ8)), " VLSQ-8 points (one participant SD)"), "Same transform/link, with beta multiplied by the fixed SD", "Two-sided 95% model interval",
  "one_sd_contrast", "Adjusted outcome at centre - 0.5 SD versus centre + 0.5 SD", "Back-transformed marginal predictions, equally weighted over sites", "Two-sided 95% interval for each prediction and contrast",
  "site_slopes", "Site-specific VLSQ-8 slopes", "Same raw-point and one-SD scales; estimates follow the global interaction", "Two-sided 95% intervals; no default site-level p screen",
  "paired_placement", "Matched near-eye versus chest one-SD effects", "Separately fitted on identical participant-days and a common effect scale", "Component 95% intervals in the plot or adjacent table"
)

practical_estimands |>
  h08_gt(
    title = "Consequential H08 estimates required in Stage 2",
    note = "No effect estimate is calculated in Stage 1.",
    font_size = 10
  ) |>
  cols_width(
    estimand_id ~ px(135),
    reader_facing_quantity ~ px(310),
    calculation_scale ~ px(510),
    interval ~ px(310)
  )
```

Wald model intervals are proposed as the primary Stage 2 interval method,
with scale and transformation stated. If diagnostics require resampling, the
pilot-first policy applies before any production run. A p-value never replaces
an interval, and an inconclusive interval is not described as evidence of no
association.

# Exact candidate samples

## Candidate-frame construction

```{r}
read_main_long <- function(data, placement_label) {
  data |>
    select(
      site,
      Id,
      local_date,
      VLSQ8,
      photoperiod_hours,
      all_of(metric_source_map$source_column)
    ) |>
    pivot_longer(
      cols = all_of(metric_source_map$source_column),
      names_to = "source_column",
      values_to = "value"
    ) |>
    left_join(metric_source_map, by = "source_column") |>
    mutate(
      data_scenario = "Primary dataset",
      placement = placement_label
    )
}

main_long <- bind_rows(
  read_main_long(main_near_eye, "Near eye"),
  read_main_long(main_chest, "Chest")
)

gap_long <- readRDS(h08_path(
  "artifacts", "06_model_data", "scenarios", "manuscript_prepared_data",
  "participant_day_metrics.rds"
)) |>
  filter(metric_id %in% metric_ids) |>
  transmute(
    data_scenario = "Gap-timing-unaware dataset",
    placement = if_else(position == "glasses", "Near eye", "Chest"),
    site,
    Id,
    local_date = as.Date(local_date),
    metric_id,
    value = manuscript_prepared_value
  ) |>
  left_join(vlsq |> select(site, Id, VLSQ8), by = c("site", "Id")) |>
  left_join(metric_source_map |> select(metric_id, metric_order), by = "metric_id")

candidate_long <- bind_rows(
  main_long |>
    select(
      data_scenario,
      placement,
      site,
      Id,
      local_date,
      metric_id,
      metric_order,
      value,
      VLSQ8
    ),
  gap_long
)

stopifnot(
  !anyDuplicated(candidate_long[c(
    "data_scenario", "placement", "site", "Id", "local_date", "metric_id"
  )]),
  sum(is.na(main_long$VLSQ8)) == 0L,
  sum(is.na(gap_long$VLSQ8)) == 0L
)

summarise_candidate <- function(data) {
  data |>
    filter(!is.na(value), !is.na(VLSQ8)) |>
    group_by(data_scenario, placement, metric_id, metric_order) |>
    summarise(
      participants = n_distinct(site, Id),
      participant_days = n(),
      sites = n_distinct(site),
      score_min = min(VLSQ8),
      score_max = max(VLSQ8),
      .groups = "drop"
    )
}

all_available_candidates <- summarise_candidate(candidate_long)

paired_wide <- candidate_long |>
  filter(!is.na(value), !is.na(VLSQ8)) |>
  select(
    data_scenario,
    site,
    Id,
    local_date,
    metric_id,
    metric_order,
    placement,
    value,
    VLSQ8
  ) |>
  pivot_wider(names_from = placement, values_from = value) |>
  filter(!is.na(`Near eye`), !is.na(Chest))

paired_candidates <- paired_wide |>
  group_by(data_scenario, metric_id, metric_order) |>
  summarise(
    participants = n_distinct(site, Id),
    participant_days = n(),
    sites = n_distinct(site),
    score_min = min(VLSQ8),
    score_max = max(VLSQ8),
    .groups = "drop"
  )
```

These are candidate complete-case frames before response transformation and
model-specific rank/convergence checks. Stage 2 must persist the exact row key
of every fitted full/reduced pair; these prepared counts cannot substitute for
the fitted-model sample.

## Primary dataset: all-available placement samples

```{r}
all_available_candidates |>
  filter(data_scenario == "Primary dataset") |>
  left_join(metric_contract |> select(metric_id, manuscript_name), by = "metric_id") |>
  mutate(cell = h08_count_cell(participants, participant_days, sites)) |>
  select(metric_order, Metric = manuscript_name, placement, cell) |>
  pivot_wider(names_from = placement, values_from = cell) |>
  arrange(metric_order) |>
  select(Metric, `Near eye`, Chest) |>
  h08_gt(
    title = "Primary-dataset all-available candidate samples",
    note = "Cells are participants / participant-days / sites. Near eye is primary; chest is complementary and not a matched placement comparison.",
    font_size = 10
  ) |>
  cols_width(Metric ~ px(310), `Near eye` ~ px(165), Chest ~ px(165))
```

```{r}
candidate_site_matrix <- function(data_scenario_label, placement_label) {
  candidate_long |>
    filter(
      data_scenario == data_scenario_label,
      placement == placement_label,
      !is.na(value),
      !is.na(VLSQ8)
    ) |>
    group_by(metric_id, metric_order, site) |>
    summarise(
      participants = n_distinct(Id),
      participant_days = n(),
      .groups = "drop"
    ) |>
    left_join(metric_contract |> select(metric_id, manuscript_name), by = "metric_id") |>
    left_join(site_display, by = "site") |>
    mutate(cell = h08_count_cell(participants, participant_days)) |>
    select(metric_order, Metric = manuscript_name, display_order, display_name, cell) |>
    arrange(metric_order, display_order) |>
    select(-display_order) |>
    pivot_wider(names_from = display_name, values_from = cell) |>
    arrange(metric_order) |>
    select(-metric_order)
}

candidate_site_matrix("Primary dataset", "Near eye") |>
  h08_gt(
    title = "Primary near-eye candidate samples by submitted site",
    note = "Cells are participants / participant-days. These counts define the interaction support that must be re-reported for each exact fitted model.",
    font_size = 8
  ) |>
  cols_width(Metric ~ px(230), everything() ~ px(105))
```

```{r}
candidate_site_matrix("Primary dataset", "Chest") |>
  h08_gt(
    title = "Complementary chest candidate samples by submitted site",
    note = "Cells are participants / participant-days. MPI has no chest dataset and is absent by design.",
    font_size = 8
  ) |>
  cols_width(Metric ~ px(230), everything() ~ px(112))
```

## Exactly paired near-eye and chest samples

```{r}
paired_candidates |>
  left_join(metric_contract |> select(metric_id, manuscript_name), by = "metric_id") |>
  mutate(cell = h08_count_cell(participants, participant_days, sites)) |>
  select(metric_order, Metric = manuscript_name, data_scenario, cell) |>
  pivot_wider(names_from = data_scenario, values_from = cell) |>
  arrange(metric_order) |>
  select(
    Metric,
    `Primary dataset`,
    `Gap-timing-unaware dataset`
  ) |>
  h08_gt(
    title = "Paired/common-sample candidates for separate placement fits",
    note = "Cells are paired participants / paired participant-days / sites. Both placements use these exact same keys for REPORT-009 comparisons.",
    font_size = 10
  ) |>
  cols_width(
    Metric ~ px(300),
    `Primary dataset` ~ px(180),
    `Gap-timing-unaware dataset` ~ px(210)
  )
```

## Gap-timing-unaware dataset sensitivity

The sensitivity term and its 50%-per-hour/80%-per-day qualification were
defined at first use in the scope section. The same approved H08 implementation
is now applied to that dataset and compared with the primary dataset.

```{r}
all_available_candidates |>
  filter(data_scenario == "Gap-timing-unaware dataset") |>
  left_join(metric_contract |> select(metric_id, manuscript_name), by = "metric_id") |>
  mutate(cell = h08_count_cell(participants, participant_days, sites)) |>
  select(metric_order, Metric = manuscript_name, placement, cell) |>
  pivot_wider(names_from = placement, values_from = cell) |>
  arrange(metric_order) |>
  select(Metric, `Near eye`, Chest) |>
  h08_gt(
    title = "Gap-timing-unaware all-available candidate samples",
    note = "Cells are participants / participant-days / sites. The same H08 formulas, families, score reference, contrasts, and diagnostics are applied.",
    font_size = 10
  ) |>
  cols_width(Metric ~ px(310), `Near eye` ~ px(165), Chest ~ px(165))
```

The sensitivity comparison must report both scenario-specific samples and the
exact common sample shared with the primary dataset. A change in estimate,
uncertainty, multiplicity-adjusted conclusion, or site/placement pattern is
classified without selecting a dataset from its p-value.

# Site and photoperiod handling

Site is retained as a fixed categorical adjustment because H08 is a
multisite between-participant association and because the registered
interaction is defined by site. Explicit sum contrasts make the additive
VLSQ-8 coefficient an equal-site average rather than an observation-weighted
average. Near-eye and chest all-available models average over different site
sets; only the paired/common eight-site fits support a direct placement plot.

Photoperiod is not part of the primary H08 registration. It is proposed as a
single linear daily adjustment in a named sensitivity to check whether
seasonal measurement timing changes the VLSQ-8 slope. Site stays in that model;
there is no VLSQ-8-by-photoperiod interaction, no latitude substitution, and
no temperature or other added covariate. The sensitivity uses the same metric-
specific rows as its unadjusted counterpart because prepared photoperiod is
complete. A material change reopens interpretation; it does not make the
photoperiod model primary after results are seen.

# Diagnostics and sensitivity battery

## Fixed sensitivity matrix

```{r}
sensitivity_matrix <- tribble(
  ~order, ~sensitivity, ~change, ~held_fixed, ~inferential_role, ~reopen_trigger,
  1L, "Primary near eye", "None", "Approved metric and H08 contracts", "Primary", "Any mandatory fit or diagnostic failure",
  2L, "Complementary chest", "Placement only", "Formula, score, family, estimands", "Complementary", "Material disagreement with near eye",
  3L, "Paired/common placement", "Restrict both placements to identical eight-site days", "Separate fits and common effect scale", "Placement concordance", "Direction/magnitude incompatibility or invalid direct scale",
  4L, "Gap-timing-unaware dataset", "Prepared metric values and available days", "New H08 implementation in both datasets", "Data-preparation sensitivity", "Material estimate, interval, conclusion, site, or placement change",
  5L, "Photoperiod-adjusted", "Add centred linear photoperiod", "Site, score, metric, rows, family", "Descriptive model sensitivity", "Material VLSQ-8 slope change",
  6L, "Participant summary", "One natural-scale daily mean per participant, then approved response representation", "Site and VLSQ-8 estimand", "Unequal-day-weighting sensitivity", "Material slope or heterogeneity change",
  7L, "Longest period exact-only", "Use only lower bounds equal to upper bounds", "Formula and family", "Metric-identifiability sensitivity", "Material slope or interval change",
  8L, "Observed dose", "Replace corrected dose by observed dose on exact common rows", "Formula and family", "Correction-dependence sensitivity", "Material slope or interval change",
  9L, "Leave one site out", "Remove one whole site per run", "Additive model and response contract", "Influence diagnostic", "Conclusion driven by one site",
  10L, "All-zero-day inclusion / Rules B and C", "Approved day-eligibility alternatives when H08-ready metric artifacts exist", "Prespecified thresholds and model contract", "Deferred prepared-data sensitivity", "Material sample or result change"
)

sensitivity_matrix |>
  select(-order) |>
  h08_gt(
    title = "Proposed H08 sensitivity battery",
    note = "Sensitivities are interpreted for stability; they are not searched for a preferred result. Rules B/C and all-zero-inclusive H08 metrics are not current direct H08 inputs and cannot be fitted until their prepared metric artifacts exist.",
    font_size = 9
  ) |>
  cols_width(
    sensitivity ~ px(210),
    change ~ px(330),
    held_fixed ~ px(300),
    inferential_role ~ px(180),
    reopen_trigger ~ px(280)
  )
```

## Planned diagnostics and failure rules

```{r}
diagnostic_registry <- tribble(
  ~domain, ~required_check, ~interpretation_or_action,
  "Score and design", "Pinned score rule; missing items; range; centre/SD; within-site range and SD; design rank", "Do not fit until score provenance is approved; non-estimable interaction cells remain explicit",
  "Exact samples", "Full/reduced row identity; participants, participant-days, sites, and contributing metric hours overall/by site", "Mismatch aborts comparison; unavailable hours are labelled unavailable, never zero",
  "Missingness", "Metric reason codes; score missingness; days per participant by score and site", "Report differential availability; participant-summary sensitivity checks unequal day weighting",
  "Gaussian fits", "Convergence, singularity, residual-versus-fitted, Q-Q, spread, tails, influence, transformed prediction range", "A family failure opens a common response-family gate; no result-specific switch",
  "Tweedie-log fits", "Convergence, Hessian, power/dispersion, observed versus simulated zero mass, tails, response and state-duration prediction bounds", "Strong failure makes the result non-estimable pending one common approved replacement",
  "Random effects", "Participant variance, boundary/singularity, conditional residual pattern", "Do not delete the random intercept merely to obtain a p-value",
  "Interaction", "Rank, Hessian, condition, site-slope SEs/CIs, leverage, independent participants per site", "If unstable, retain the additive average question and report heterogeneity as non-estimable",
  "Serial dependence", "Residual correlation by true date lag within participant; long-gap and short-sequence support", "Interpret participant-day precision cautiously; predeclare any correlation/cluster-resampling sensitivity before use",
  "Influence", "Participant leverage and leave-one-site-out additive slopes", "A single-site-driven conclusion reopens the claim gate",
  "Metric bounds", "Non-negative outcomes; 24 h full-day bound; state-duration denominators; pre-sleep >6 h audit; dose correction guard", "Impossible predictions or construct violations trigger model review",
  "Longest period", "Observed lower/possible upper bounds and exact-identifiable-only sample", "Report the primary explicitly as an observed lower bound",
  "Placement", "Same effect scale and exact paired keys for near-eye/chest comparison", "No unmatched identity plot and no equivalence language",
  "Data sensitivity", "Scenario-specific and exact common samples; effect and interval stability", "Classify stable, magnitude-sensitive, precision-sensitive, direction-sensitive, or non-estimable",
  "Multiplicity", "Complete nine-row raw vectors and independent R reproduction of BH adjustment", "Scalar adjustment, family shortening, or reused q-values aborts release"
)

diagnostic_registry |>
  h08_gt(
    title = "Mandatory Stage 2 diagnostic interpretation",
    note = "A check is not complete merely because software produced a plot or warning flag; the H08 report must state what the diagnostic means for the result.",
    font_size = 9
  ) |>
  cols_width(
    domain ~ px(145),
    required_check ~ px(550),
    interpretation_or_action ~ px(540)
  )
```

If participant-level cluster resampling, simulated residual envelopes, or any
other heavy refitting is required, Stage 2 first runs 50 successful replicates
(100 only if practical) with production code. The pilot is labelled
**PILOT — NOT FOR INFERENCE OR MANUSCRIPT REPORTING** and reports failures,
warnings, elapsed time, full-runtime projection, memory/checkpoint behavior,
and preview outputs before requesting production approval.

# Stage 2 output and V0 comparison plan

After approval, Stage 2 will:

1. build a durable H08 model-frame artifact from the four hash-verified direct
   inputs, with score constants, factor levels, contrasts, exact row keys,
   metric variants, and exclusion reasons;
2. fit the approved near-eye, complementary chest, and paired/common models;
3. assemble complete nine-member raw and BH-adjusted p-value vectors for the
   approved inferential families;
4. report every consequential estimate with a 95% confidence interval and
   practical raw-point/one-SD interpretation;
5. interpret every required diagnostic and mark a failed target non-estimable
   rather than silently changing its family or rows;
6. reproduce the V0 model construction and rendered strings, then separate
   differences caused by data/metric preparation, score/model specification,
   placement role, and multiplicity;
7. recreate readable near-eye and chest H08 outputs using submitted metric and
   site conventions; and
8. run the same approved H08 implementation on the gap-timing-unaware dataset,
   including scenario-specific and exact common-sample comparisons.

All reader-facing figures will have informative alt text and an exact source
CSV. If fitted paired/common effects share the required scale and estimand,
Stage 3 will use the REPORT-009 near-eye-versus-chest display with identity and
null lines; otherwise it will give the closest valid paired table or effect
plot and explain why an identity plot is not applicable.

# Decisions requiring explicit author approval

Please approve, revise, or reject each item before Stage 2:

1. **H08-G1 — VLSQ-8 score:** confirm the pinned stored score as the intended
   predictor and provide/approve its upstream provenance. The reconstructed
   rule is `sum(eight 1–5 item codes) + 5`; the cited article supplies no
   standard total-score algorithm.
2. **H08-G2 — score scaling:** centre at 21.5978, the mean of all 184 pinned
   questionnaire rows; report effects per raw point and per 5.5398 points
   (one participant SD); do not standardize within site.
3. **H08-G3 — primary estimand:** use the site-adjusted additive VLSQ-8 slope
   as the primary across-site association, rather than V0's joint main-plus-
   interaction block.
4. **H08-G4 — heterogeneity:** test the registered site interaction separately
   and treat it as secondary; site slopes have estimates and 95% intervals but
   no default coefficient-wise p-value family.
5. **H08-G5 — metric set:** approve the exact nine repaired outcomes, including
   corrected dose and the longest continuous period as an observed lower bound;
   exclude timing, dynamics, MDER, and unsupported metrics from H08.
6. **H08-G6 — response package:** inherit the current H01/H05 response family,
   transform, link, and common-family failure rule shown above. In particular,
   pre-sleep time uses identity Gaussian rather than V0 Tweedie-log.
7. **H08-G7 — analysis unit:** retain participant-day outcomes with a
   participant nested-in-site random intercept; recognize the participant as
   the independent predictor unit; use the one-summary-per-participant model
   only as an unequal-day-weighting sensitivity.
8. **H08-G8 — site and photoperiod:** use explicit submitted site order and sum
   contrasts; keep photoperiod out of the primary model and include only the
   declared linear adjustment sensitivity. Add no temperature, latitude, or
   other covariate.
9. **H08-G9 — multiplicity:** approve separate complete nine-member BH families
   for the primary near-eye average and interaction tests, with separately
   named complementary/sensitivity families as listed. Prohibit scalar
   `p.adjust()` calls.
10. **H08-G10 — placement and samples:** near eye all-available is primary;
    chest all-available is complementary; direct placement comparison uses
    separate models on exact paired/common participant-days; no pooling.
11. **H08-G11 — practical reporting:** require raw-point, one-SD, and centred
    one-SD contrast estimates with 95% intervals; use ratios for linked or
    log-transformed outcomes and differences for identity outcomes.
12. **H08-G12 — sensitivities:** approve the fixed matrix, including the
    gap-timing-unaware dataset, photoperiod adjustment, participant summary,
    exact-period, observed-dose, leave-one-site-out, and deferred approved
    coverage/all-zero scenarios when their H08-ready metrics exist.
13. **H08-G13 — diagnostics and stop rules:** approve the full diagnostic
    registry, the common response-family gate, and non-estimable disposition
    for unstable interaction or model targets.
14. **H08-G14 — inference language:** describe supported results as
    associations, not physiological effects or health outcomes; do not turn an
    inconclusive result into evidence of absence.

::: {.callout-caution}
## Explicit approval required

Stage 1 ends here. No Stage 2 model, p-value, confidence interval, diagnostic
simulation, or result figure will be produced until the author explicitly
approves the H08-G1 through H08-G14 contract (or supplies revisions).
:::

# Reproducibility record

```{r}
reproducibility_record <- tibble(
  item = c(
    "R",
    "Quarto",
    "dplyr",
    "tidyr",
    "readr",
    "gt",
    "digest",
    "rvest",
    "Stage 1 inferential fits"
  ),
  version_or_value = c(
    as.character(getRversion()),
    "1.9.37",
    as.character(packageVersion("dplyr")),
    as.character(packageVersion("tidyr")),
    as.character(packageVersion("readr")),
    as.character(packageVersion("gt")),
    as.character(packageVersion("digest")),
    as.character(packageVersion("rvest")),
    "0"
  )
)

reproducibility_record |>
  h08_gt(
    title = "H08 Stage 1 execution environment",
    note = "Render command: quarto render audit/hypotheses/H08/01_audit_and_plan.qmd --to html. Direct input hashes are reported above.",
    font_size = 10
  ) |>
  cols_width(item ~ px(240), version_or_value ~ px(260))
```
