---
title: "H08 Stage 2: implementation, diagnostics, and V0 comparison"
subtitle: "Visual light sensitivity and personal light exposure — author result gate"
date: "2026-08-12"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    embed-resources: true
    code-fold: true
    code-summary: "Show audit code"
    code-overflow: wrap
    df-print: paged
bibliography: ../../../bibliography.bib
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r}
#| label: setup
#| include: false

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

configured_root <- Sys.getenv("NATHEALTH_PROJECT_ROOT", unset = "")
if (
  nzchar(configured_root) &&
    file.exists(file.path(configured_root, "renv.lock")) &&
    file.exists(file.path(configured_root, "_quarto.yml"))
) {
  root <- normalizePath(configured_root, winslash = "/", mustWork = TRUE)
} else {
  root <- locate_project_root()
}
project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
if (dir.exists(project_library)) {
  .libPaths(c(project_library, .libPaths()))
}

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(readr)
  library(stringr)
  library(tibble)
  library(tidyr)
})

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H08 Stage 2 report must run under R 4.6.1", call. = FALSE)
}

source(file.path(root, "scripts", "pipeline", "p_value_display.R"))

h08_artifact <- function(...) file.path(root, "artifacts", ...)
read_h08 <- function(...) {
  readr::read_csv(h08_artifact(...), show_col_types = FALSE, na = "")
}

approvals <- read_h08("06_model_data", "H08", "H08_author_approvals.csv")
input_audit <- read_h08("06_model_data", "H08", "H08_input_audit.csv")
metric_registry <- read_h08("06_model_data", "H08", "H08_metric_registry.csv")
response_contract <- read_h08(
  "06_model_data", "H08", "H08_response_contract_audit.csv"
)
sample_index <- read_h08("06_model_data", "H08", "H08_model_frame_index.csv")
sample_by_site <- read_h08(
  "06_model_data", "H08", "H08_model_frame_by_site.csv"
)
paired_audit <- read_h08("06_model_data", "H08", "H08_paired_sample_audit.csv")
main_gap_common_audit <- read_h08(
  "06_model_data", "H08", "H08_main_gap_common_sample_audit.csv"
)
score_audit <- read_h08("08_diagnostics", "H08", "H08_vlsq_score_audit.csv")
diagnostics <- read_h08("08_diagnostics", "H08", "H08_model_diagnostics.csv")
response_gate <- read_h08(
  "08_diagnostics", "H08", "H08_response_family_gate.csv"
)
loo_summary <- read_h08(
  "08_diagnostics", "H08", "H08_leave_one_site_out_summary.csv"
)
master <- read_h08("09_tables", "H08", "H08_model_results_master.csv")
family_audit <- read_h08("09_tables", "H08", "H08_family_audit.csv")
predictions <- read_h08("09_tables", "H08", "H08_centered_predictions.csv")
site_slopes <- read_h08("09_tables", "H08", "H08_site_specific_slopes.csv")
gap_sensitivity <- read_h08(
  "09_tables", "H08", "H08_gap_timing_unaware_sensitivity.csv"
)
photoperiod <- read_h08(
  "09_tables", "H08", "H08_photoperiod_sensitivity.csv"
)
participant_summary <- read_h08(
  "09_tables", "H08", "H08_participant_summary_sensitivity.csv"
)
exact_longest <- read_h08(
  "09_tables", "H08", "H08_exactly_identified_longest_period_sensitivity.csv"
)
observed_dose <- read_h08(
  "09_tables", "H08", "H08_observed_dose_sensitivity.csv"
)
v0_reproduction <- read_h08(
  "09_tables", "H08", "H08_v0_reproduction.csv"
)
v0_comparison <- read_h08(
  "09_tables", "H08", "H08_v0_to_new_comparison.csv"
)
metric011_result_comparison <- read_h08(
  "09_tables", "H08", "H08_metric011_result_comparison.csv"
)
metric011_display_invariance <- read_h08(
  "09_tables", "H08", "H08_metric011_display_invariance.csv"
)
metric011_bh_recalculation <- read_h08(
  "09_tables", "H08", "H08_metric011_bh_recalculation.csv"
)
metric011_reconciliation <- read_h08(
  "12_manifests", "H08", "H08_metric011_reconciliation.csv"
)
execution_environment <- read_h08(
  "12_manifests", "H08", "H08_execution_environment.csv"
)

stopifnot(
  nrow(input_audit) == 20L,
  all(input_audit$hash_verified),
  nrow(metric011_result_comparison) == 2L,
  !any(metric011_result_comparison$author_review_required),
  nrow(metric011_display_invariance) == 6L,
  all(metric011_display_invariance$display_identical),
  nrow(metric011_bh_recalculation) == 36L,
  all(metric011_bh_recalculation$adjusted_display_identical),
  all(metric011_bh_recalculation$adjusted_conclusion_identical),
  nrow(metric011_reconciliation) == 33L,
  all(metric011_reconciliation$invariant_verified)
)

main_near <- master |>
  filter(run_id == "main__glasses__all_available") |>
  arrange(metric_order)
main_chest <- master |>
  filter(run_id == "main__chest__all_available") |>
  arrange(metric_order)
main_primary <- bind_rows(main_near, main_chest)

near_average_significant <- sum(
  main_near$average_adjusted_significant,
  na.rm = TRUE
)
near_interaction_significant <- sum(
  main_near$interaction_adjusted_significant,
  na.rm = TRUE
)
chest_average_significant <- sum(
  main_chest$average_adjusted_significant,
  na.rm = TRUE
)
chest_interaction_significant <- sum(
  main_chest$interaction_adjusted_significant,
  na.rm = TRUE
)

format_p_cell <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  ifelse(
    display$p_bold,
    paste0("**", display$p_display, "**"),
    display$p_display
  )
}

format_effect <- function(estimate, low, high, effect_type) {
  ifelse(
    effect_type == "ratio",
    sprintf("%.3f× (%.3f to %.3f)", estimate, low, high),
    sprintf("%+.3f h (%+.3f to %+.3f)", estimate, low, high)
  )
}

model_to_practical <- function(value, response_transform, effect_scale) {
  case_when(
    effect_scale == "difference" ~ value,
    response_transform == "log10_offset_0.1" ~ 10^value,
    effect_scale == "ratio" ~ exp(value),
    TRUE ~ value
  )
}

format_sample <- function(participants, participant_days, sites) {
  sprintf("%d / %d / %d", participants, participant_days, sites)
}

h08_gt <- function(data, title, subtitle = NULL, note = NULL, size = 11) {
  out <- data |>
    gt() |>
    opt_align_table_header(align = "left") |>
    opt_row_striping() |>
    tab_options(
      table.width = pct(100),
      table.font.size = px(size),
      data_row.padding = px(3),
      heading.align = "left",
      source_notes.font.size = px(max(9, size - 2))
    )
  if (is.null(subtitle)) {
    out <- out |> tab_header(title = md(title))
  } else {
    out <- out |> tab_header(title = md(title), subtitle = md(subtitle))
  }
  if (!is.null(note)) out <- out |> tab_source_note(md(note))
  out
}

result_display <- function(data) {
  data |>
    transmute(
      Order = metric_order,
      Metric = manuscript_name,
      `Effect per 5.540 VLSQ-8 points (95% CI)` = format_effect(
        estimate_practical_per_sd,
        conf_low_practical_per_sd,
        conf_high_practical_per_sd,
        effect_type
      ),
      `Raw p` = format_p_cell(average_p_raw, average_adjusted_significant),
      `BH-adjusted p` = format_p_cell(
        average_p_adjusted,
        average_adjusted_significant
      ),
      `Participants / participant-days / sites` = format_sample(
        participants,
        participant_days,
        sites
      ),
      Diagnostics = if_else(
        diagnostic_status == "PASS",
        "Pass",
        str_replace_all(diagnostic_issues, "_", " ")
      )
    )
}
```

::: {.callout-note}
## Accepted Stage 2 package with bounded METRIC-011 maintenance

The author accepted this Stage 2 result package before the Stage 3 and Stage 4
reports were created. On 2026-08-12, shared decision METRIC-011 normalised eight
L10 participant-day round-off remnants to exact numerical zero. The bounded
H08 reseal refitted only the inherited primary-dataset L10 branches and their
specified diagnostics and sensitivities, then recalculated the four affected
nine-member BH families. It did not rerun the full analysis or any resampling.
The retained results, displayed values, diagnostic dispositions, sensitivity
classifications, and scientific claims remained unchanged.
:::

# Hypothesis and analytical question

The signed preregistration states:

> **H8: Duration-, exposure-history-, and level-based metrics are associated
> with VLSQ-8 light sensitivity scores.**

The audited question is whether participant-level VLSQ-8 scores are associated
with repeated participant-day personal light-exposure metrics after adjustment
for submitted site. Near-eye measurements are primary; chest measurements are
complementary. Temperature and other unsupported predictors are absent.

The predefined preparation sensitivity is called the
**gap-timing-unaware dataset**. It still passed the general 50%-per-hour and
80%-per-day coverage rules. The name means that the timing of the remaining
missing observations is not used for an additional metric-specific adjustment;
it does not mean that gaps, missingness, or coverage were ignored. For this
one-time first explanation, the primary could be interpreted as a
time-sensitive primary metric dataset because its approved preparation uses
the timing of remaining missing observations where that timing is relevant.
From here onward, it is called simply the **primary dataset**.

# Frozen implementation contract

All 14 Stage 1 decisions are recorded as approved, and all 20 sealed inputs
match their pinned SHA-256 values.

```{r}
#| label: approvals-and-inputs

bind_rows(
  tibble(
    Check = "Approved H08-G decisions",
    Result = sprintf("%d / %d", sum(approvals$approved), nrow(approvals))
  ),
  tibble(
    Check = "Pinned input hashes verified",
    Result = sprintf(
      "%d / %d",
      sum(input_audit$hash_verified),
      nrow(input_audit)
    )
  )
) |>
  h08_gt(
    title = "Implementation preconditions",
    note = "The accepted analysis and the bounded METRIC-011 reseal stop immediately if an approval, input pin, or invariance check fails."
  )
```

## METRIC-011 exact-zero reseal

The controlling [METRIC-011 decision](../../decisions/l10_numerical_zero_normalization.md)
classifies floating-point remnants that are numerically zero as exact zero.
Its independently verified evidence identified eight affected primary L10
means: three near-eye and five chest participant-days changed from
`4.163336342344337e-17` lx to `0` lx. All samples, non-L10 values,
hourly and 30-minute values, gap-timing-unaware values, V0 outputs, and the
accepted exactly identified longest-period and observed-dose sensitivities
were protected.

```{r}
#| label: tbl-h08-metric011-reseal

tibble::tribble(
  ~Check, ~Observed, ~Interpretation,
  "Primary L10 source cells", "8: 3 near eye; 5 chest", "Exact-zero normalisation only; no sample changed",
  "Primary-dataset L10 model bundles", "6", "All-available, paired-placement, and preparation-common samples for both placements",
  "Inherited L10 sensitivity bundles", "4", "Photoperiod and participant-summary fits for both placements",
  "L10 leave-one-site-out refits", "17", "All completed; influence interpretation unchanged",
  "Affected BH-family rows", sprintf("%d across %d families", nrow(metric011_bh_recalculation), n_distinct(metric011_bh_recalculation$family_id)), sprintf("%d adjusted values differed below display precision; no displayed p-value or decision changed", sum(!metric011_bh_recalculation$adjusted_p_identical)),
  "Reader-display checks", sprintf("%d / %d PASS", sum(metric011_display_invariance$display_identical), nrow(metric011_display_invariance)), "Effects, p-values, diagnostics, and sensitivity summaries retained their accepted display",
  "Protected-scope reconciliation", sprintf("%d / %d PASS", sum(metric011_reconciliation$invariant_verified), nrow(metric011_reconciliation)), "Every untargeted row, field, model object, and protected artifact matched its accepted identity"
) |>
  h08_gt(
    title = "Bounded METRIC-011 reseal",
    note = "The three sub-display adjusted-p differences comprise two L10 values and one M10 family derivative. None changed a three-decimal display or BH-adjusted conclusion."
  )
```

The [result comparison](../../../artifacts/09_tables/H08/H08_metric011_result_comparison.csv),
[complete-family BH record](../../../artifacts/09_tables/H08/H08_metric011_bh_recalculation.csv),
[display-invariance record](../../../artifacts/09_tables/H08/H08_metric011_display_invariance.csv),
and [33-scope reconciliation](../../../artifacts/12_manifests/H08/H08_metric011_reconciliation.csv)
provide the machine-readable audit trail. Because no retained result or claim
changed, the existing author approval remains applicable.

## VLSQ-8 scoring and scaling

The participant-level predictor is complete for all 184 participants across
nine sites. The stored score equals the sum of the eight ordered 1–5 item
codes plus 5 for every participant. Its observed range is 13–39, its mean is
21.597826, and its participant standard deviation is 5.539840. Models use the
centred raw score; practical effects compare scores separated by one such
standard deviation. No within-site standardisation was applied.

```{r}
#| label: score-audit-table

score_audit |>
  transmute(
    Participants = participants,
    Sites = sites,
    `Missing scores` = missing_scores,
    `Missing item cells` = missing_item_cells,
    Range = sprintf("%d–%d", observed_min, observed_max),
    Mean = sprintf("%.6f", observed_mean),
    SD = sprintf("%.6f", observed_participant_sd),
    `Stored − item sum` = stored_minus_item_sum_unique,
    Verified = scoring_rule_verified
  ) |>
  h08_gt(
    title = "VLSQ-8 score audit",
    note = "The +5 scoring constant is an author-approved project convention; it is not claimed as a published standard total-score algorithm."
  )
```

## Exact fitted formulas

The following evaluated cell is the complete formula contract. Site uses the
submitted order and explicit sum contrasts. `site:Id` is the participant
random-intercept grouping factor; participants remain nested within site.

```{r}
#| label: exact-fitted-formulas
#| echo: true
#| code-fold: false

site_only <- stats::as.formula(
  "response_value ~ site + (1 | site:Id)"
)
additive <- stats::as.formula(
  "response_value ~ site + VLSQ8_c + (1 | site:Id)"
)
interaction <- stats::as.formula(
  "response_value ~ site * VLSQ8_c + (1 | site:Id)"
)
photoperiod_site_only <- stats::as.formula(
  "response_value ~ site + photoperiod_c + (1 | site:Id)"
)
photoperiod_additive <- stats::as.formula(
  "response_value ~ site + photoperiod_c + VLSQ8_c + (1 | site:Id)"
)
photoperiod_interaction <- stats::as.formula(
  "response_value ~ site * VLSQ8_c + photoperiod_c + (1 | site:Id)"
)
participant_site_only <- stats::as.formula(
  "participant_response ~ site"
)
participant_additive <- stats::as.formula(
  "participant_response ~ site + VLSQ8_c"
)
participant_interaction <- stats::as.formula(
  "participant_response ~ site * VLSQ8_c"
)

site_only
additive
interaction
photoperiod_site_only
photoperiod_additive
photoperiod_interaction
participant_site_only
participant_additive
participant_interaction
```

The primary average association is the likelihood-ratio comparison of
`site_only` with `additive`. Site heterogeneity is the separate comparison of
`additive` with `interaction`. These tests answer different questions and are
adjusted in different complete nine-member Benjamini–Hochberg families.

## Metrics, response families, and practical estimands

```{r}
#| label: response-contract-table

response_contract |>
  arrange(metric_order) |>
  transmute(
    Order = metric_order,
    Metric = manuscript_name,
    Category = manuscript_category,
    Unit = display_unit,
    `Response / transform` = paste(response_family, response_transform, sep = " / "),
    `Effect shown` = if_else(
      effect_scale == "ratio",
      "Ratio per VLSQ-8 SD",
      "Hour difference per VLSQ-8 SD"
    ),
    `Participant-day value` = primary_value_definition
  ) |>
  h08_gt(
    title = "Nine-member H08 response contract",
    note = "melEDI and manuscript-facing period terminology are used throughout."
  )
```

# Exact fitted samples

Every row entering each model was fixed before fitting. The compact tables
below report participants / participant-days / sites. Row-key and full
model-frame SHA-256 values are preserved in the source artifact.

## All available samples by dataset and placement

```{r}
#| label: all-available-samples

sample_index |>
  filter(sample_scenario == "all_available") |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Dataset = reader_scenario,
    Placement = placement_label,
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Dataset, Placement) |>
  select(-metric_order) |>
  pivot_wider(names_from = c(Dataset, Placement), values_from = Sample) |>
  h08_gt(
    title = "Exact all-available fitted samples",
    note = "Cells are participants / participant-days / sites. Each metric has its own complete fitted frame."
  )
```

## Primary-dataset site cells

```{r}
#| label: primary-site-cells

site_levels <- sample_by_site |>
  distinct(display_order, display_name) |>
  arrange(display_order) |>
  pull(display_name)

sample_by_site |>
  filter(run_id %in% c(
    "main__glasses__all_available",
    "main__chest__all_available"
  )) |>
  mutate(
    Placement = if_else(
      run_id == "main__glasses__all_available",
      "Near eye",
      "Chest"
    ),
    Site = factor(display_name, levels = site_levels),
    Cell = sprintf("%d / %d", participants, participant_days)
  ) |>
  select(metric_order, metric_id, Placement, Site, Cell) |>
  left_join(
    metric_registry |>
      select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  select(metric_order, Metric = manuscript_name, Placement, Site, Cell) |>
  arrange(metric_order, Placement, Site) |>
  pivot_wider(names_from = Site, values_from = Cell, values_fill = "—") |>
  select(-metric_order) |>
  h08_gt(
    title = "Primary-dataset fitted site cells",
    note = "Cells are participants / participant-days. Borås has no chest observations; the chest model therefore contains eight sites. Submitted names and order are retained.",
    size = 10
  )
```

## Paired-placement and cross-dataset common samples

The paired-placement frames contain exactly the same participant-day keys for
near eye and chest in all 18 metric-by-dataset checks. The cross-dataset frames
likewise contain exactly the same keys for the primary and gap-timing-unaware
fits in all 18 placement-by-metric checks.

```{r}
#| label: common-sample-audits

paired_display <- paired_audit |>
  left_join(
    metric_registry |>
      select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Comparison = if_else(
      data_scenario_id == "main",
      "Primary paired placement",
      "Gap-timing-unaware paired placement"
    ),
    Sample = format_sample(
      participants_glasses,
      participant_days_glasses,
      sites_glasses
    ),
    `Exact keys` = exact_row_keys_match
  )

cross_dataset_display <- main_gap_common_audit |>
  left_join(
    metric_registry |>
      select(metric_id, manuscript_name),
    by = "metric_id"
  ) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Comparison = paste(
      if_else(placement == "glasses", "Near eye", "Chest"),
      "primary–gap common"
    ),
    Sample = format_sample(
      participants_main,
      participant_days_main,
      sites_main
    ),
    `Exact keys` = exact_row_keys_match
  )

bind_rows(paired_display, cross_dataset_display) |>
  arrange(metric_order, Comparison) |>
  select(-metric_order) |>
  pivot_wider(
    names_from = Comparison,
    values_from = c(Sample, `Exact keys`),
    names_glue = "{Comparison}: {.value}"
  ) |>
  h08_gt(
    title = "Exact common-sample audits",
    note = "Sample cells are participants / participant-days / sites. TRUE means both compared frames had identical sorted row keys and hashes.",
    size = 10
  )
```

[Download the complete model-frame index](../../../artifacts/06_model_data/H08/H08_model_frame_index.csv)
and [all fitted site cells](../../../artifacts/06_model_data/H08/H08_model_frame_by_site.csv).

# Primary near-eye results

No average near-eye association survived the complete nine-member BH family
(`r near_average_significant` adjusted-positive results), and no near-eye
site-by-VLSQ-8 interaction survived its separate family
(`r near_interaction_significant` adjusted-positive results). The practical
estimates below are still informative about direction and uncertainty; they
are not evidence that every true effect is exactly null.

```{r}
#| label: primary-near-eye-results

result_display(main_near) |>
  select(-Order) |>
  h08_gt(
    title = "Primary near-eye average VLSQ-8 associations",
    subtitle = "Practical estimate, 95% Wald CI, REPORT-008 p-values, and exact fitted sample",
    note = "Raw p-values compare site-only with additive models. BH adjustment is across all nine metrics. Significant adjusted p-values would be bold; none is bold."
  ) |>
  fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
```

The melEDI-dose estimate was a ratio of 0.846 per VLSQ-8 SD (95% CI
0.716–0.999), while its likelihood-ratio raw p-value was 0.051 and its
BH-adjusted p-value was 0.160. The Wald interval and likelihood-ratio test use
different inferential constructions; the complete family decision is
therefore **inconclusive**, not statistically significant. The same family
qualification applies to all unadjusted patterns.

```{r}
#| label: fig-near-eye-effects
#| fig-cap: "Primary near-eye average associations between VLSQ-8 and the nine personal light-exposure metrics. Points show effects per one participant-level VLSQ-8 SD; horizontal lines are 95% Wald confidence intervals."
#| fig-alt: "Forest plot with nine near-eye light-exposure metrics. Most ratio estimates are below one, but every interval reaches or crosses one except the melEDI-dose interval, which ends just below one. The hour-difference interval for time below 10 lx melEDI before sleep crosses zero. No BH-adjusted test is significant."
#| out-width: "100%"

knitr::include_graphics(h08_artifact(
  "10_figures", "H08", "H08_near_eye_effects.png"
))
```

[Source data for the near-eye figure](../../../artifacts/11_source_data/H08/H08_near_eye_effects_data.csv)

# Complementary chest results

Chest results are complementary rather than a second primary screen. No
average chest association survived its nine-member family
(`r chest_average_significant` adjusted-positive results), and no chest
interaction survived its separate family (`r chest_interaction_significant`
adjusted-positive results).

```{r}
#| label: complementary-chest-results

result_display(main_chest) |>
  select(-Order) |>
  h08_gt(
    title = "Complementary chest average VLSQ-8 associations",
    subtitle = "Practical estimate, 95% Wald CI, REPORT-008 p-values, and exact fitted sample",
    note = "The chest placement is complementary. BH adjustment is separate from the primary near-eye family. Significant adjusted p-values would be bold; none is bold."
  ) |>
  fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
```

```{r}
#| label: fig-chest-effects
#| fig-cap: "Complementary chest average associations between VLSQ-8 and the nine personal light-exposure metrics. Points show effects per one participant-level VLSQ-8 SD; horizontal lines are 95% Wald confidence intervals."
#| fig-alt: "Forest plot with nine chest light-exposure metrics. Most ratio estimates are below one. Several intervals approach one, but the BH-adjusted association tests are not significant. The hour-difference interval for time below 10 lx melEDI before sleep crosses zero."
#| out-width: "100%"

knitr::include_graphics(h08_artifact(
  "10_figures", "H08", "H08_chest_effects.png"
))
```

[Source data for the chest figure](../../../artifacts/11_source_data/H08/H08_chest_effects_data.csv)

# Interaction structure and estimands

## Global site-heterogeneity tests

The interaction comparison asks whether the VLSQ-8 slope varies by submitted
site after the average slope has been added. It is not the primary average
association test.

```{r}
#| label: interaction-tests

main_primary |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `LRT df` = interaction_lrt_df,
    `Raw p` = format_p_cell(
      interaction_p_raw,
      interaction_adjusted_significant
    ),
    `BH-adjusted p` = format_p_cell(
      interaction_p_adjusted,
      interaction_adjusted_significant
    ),
    Status = interaction_effect_status
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "Separate site-by-VLSQ-8 heterogeneity families",
    note = "Each placement has its own complete nine-member BH family. Significant adjusted p-values would be bold; none is bold."
  ) |>
  fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
```

## Site-specific slopes

Site-specific slopes are secondary estimates from the interaction models. They
are shown with 95% intervals for transparency, not as new site-wise hypothesis
families and not as evidence for site differences when the global
heterogeneity test is inconclusive.

```{r}
#| label: site-specific-slopes

site_slopes |>
  filter(run_id %in% c(
    "main__glasses__all_available",
    "main__chest__all_available"
  )) |>
  transmute(
    metric_order,
    site_order = display_order,
    Metric = manuscript_name,
    Site = display_name,
    Placement = placement_label,
    Effect = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    )
  ) |>
  arrange(metric_order, site_order, Placement) |>
  select(-metric_order, -site_order) |>
  pivot_wider(names_from = Placement, values_from = Effect, values_fill = "—") |>
  h08_gt(
    title = "Site-specific VLSQ-8 slopes",
    subtitle = "Effect per one VLSQ-8 SD (95% Wald CI)",
    note = "Ratios use ×; the pre-sleep metric is an hour difference. Borås has no chest estimate.",
    size = 10
  )
```

## Centred low-to-high predictions

The table compares predictions at half an SD below and half an SD above the
observed score mean, averaging the sum-contrast site effect. These are
predictions, not additional tests.

```{r}
#| label: centered-predictions

predictions |>
  filter(run_id %in% c(
    "main__glasses__all_available",
    "main__chest__all_available"
  )) |>
  mutate(
    Score = sprintf("%.3f", VLSQ8),
    Prediction = sprintf("%.3f (%.3f to %.3f)", estimate, conf_low, conf_high),
    Point = recode(
      prediction_id,
      centre_minus_half_sd = "Lower score",
      centre_plus_half_sd = "Higher score"
    )
  ) |>
  select(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    Point,
    Score,
    Prediction
  ) |>
  arrange(metric_order, Placement, Point) |>
  pivot_wider(
    names_from = Point,
    values_from = c(Score, Prediction),
    names_glue = "{Point}: {.value}"
  ) |>
  select(-metric_order) |>
  h08_gt(
    title = "Centred VLSQ-8 prediction contrast",
    note = "Predictions are on each metric's natural scale and include 95% delta-method intervals with physical lower bounds applied."
  )
```

# Paired-placement comparison

Near-eye and chest models were fitted separately on exactly matched
participant-day keys. The display compares estimates; it does not pool
placements, estimate within-day placement differences, or claim equivalence.

```{r}
#| label: paired-placement-table

master |>
  filter(run_id %in% c(
    "main__glasses__paired_common_sample",
    "main__chest__paired_common_sample"
  )) |>
  transmute(
    metric_order,
    Metric = manuscript_name,
    Placement = placement_label,
    Effect = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites)
  ) |>
  arrange(metric_order, Placement) |>
  pivot_wider(
    names_from = Placement,
    values_from = c(Effect, Sample),
    names_glue = "{Placement}: {.value}"
  ) |>
  select(-metric_order) |>
  h08_gt(
    title = "Exactly paired near-eye and chest estimates",
    note = "Effects are per one VLSQ-8 SD with 95% Wald CIs. Sample cells are participants / participant-days / sites. This comparison carries estimates and intervals only; it is not an extra p-value screen."
  )
```

The identity-line plot contains only ratio-scale metrics. The pre-sleep metric
has an additive hour-difference scale and is therefore kept in the adjacent
table rather than mixed into that geometry.

```{r}
#| label: fig-paired-placement
#| fig-cap: "Near-eye and chest VLSQ-8 associations estimated on identical participant-day samples. Ratio-scale metrics only; diagonal line marks equal point estimates, not an equivalence region."
#| fig-alt: "Scatter plot comparing paired-sample near-eye ratios on the horizontal axis with chest ratios on the vertical axis for eight metrics. Points cluster near the diagonal and below one, with no equivalence bounds. The additive pre-sleep hour-difference outcome is excluded from this plot and reported in the adjacent table."
#| out-width: "88%"

knitr::include_graphics(h08_artifact(
  "10_figures", "H08", "H08_paired_placement_effects.png"
))
```

[Source data for the paired-placement figure](../../../artifacts/11_source_data/H08/H08_paired_placement_effects_data.csv)

# Diagnostics and response-family gate

All 36 primary-dataset additive and interaction fits converged, had
positive-definite Hessians, were nonsingular, and retained full-rank fixed
effects. No primary or complementary effect was declared non-estimable. The
diagnostic gate retained all nine approved response specifications; three need
explicit limitations rather than a post-result family switch.

```{r}
#| label: response-family-gate

response_gate |>
  arrange(metric_order) |>
  transmute(
    Order = metric_order,
    Metric = manuscript_name,
    `Response / transform` = paste(response_family, response_transform, sep = " / "),
    `Inferential targets` = inferential_targets,
    `Major failures` = major_failures,
    `Review targets` = review_targets,
    Decision = str_replace_all(family_gate_status, "_", " ")
  ) |>
  select(-Order) |>
  h08_gt(
    title = "Common response-family gate",
    note = "A review flag triggers explicit qualification; it does not permit data-dependent response-family switching."
  )
```

```{r}
#| label: primary-diagnostics

diagnostics |>
  filter(run_id %in% c(
    "main__glasses__all_available",
    "main__chest__all_available"
  )) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    `Maximum gradient` = sprintf(
      "%.2g",
      pmax(additive_max_gradient, interaction_max_gradient, na.rm = TRUE)
    ),
    `Residual |fit| Spearman` = sprintf(
      "%+.3f",
      residual_absolute_fitted_spearman
    ),
    `Consecutive-day residual r` = sprintf(
      "%+.3f",
      pooled_consecutive_day_residual_correlation
    ),
    `Observed / expected zeros` = if_else(
      response_family == "tweedie_log",
      sprintf("%d / %.3g", observed_zero_n, expected_zero_n),
      "—"
    ),
    `Bound check` = prediction_bound_status,
    `Diagnostic decision` = if_else(
      diagnostic_status == "PASS",
      "PASS",
      str_replace_all(diagnostic_issues, "_", " ")
    )
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "Interpreted primary-dataset diagnostics",
    note = "Serial correlations are descriptive pooled checks. Zero expectations are analytical compound-Poisson probabilities from the fitted Tweedie model; no simulation was used.",
    size = 10
  )
```

The main limitations are specific:

- Darkest 10 h mean showed increasing residual spread with fitted values
  (absolute-residual versus fitted Spearman correlations 0.341 near eye and
  0.349 chest). Its direction and intervals should therefore be read with
  extra caution.
- Time below 1 lx melEDI during sleep contained four observed exact zeros at
  each placement, whereas the fitted compound-Poisson model assigned an
  extremely small expected zero count. This is a conspicuous relative
  discrepancy but only four rows in each large fitted frame. The model is
  retained under the approved common-family gate, with the zero-mass mismatch
  stated explicitly.
- The corrected melEDI-dose fits passed the main numerical checks, but the
  correction is timing-dependent and is therefore interpreted alongside the
  exact common-sample dataset and observed-dose sensitivities below.

```{r}
#| label: fig-model-adequacy
#| fig-cap: "Selected near-eye model-adequacy checks for all nine metrics. Panels retain the model-specific transformed or Pearson-residual scales and are diagnostic rather than effect displays."
#| fig-alt: "Multi-panel diagnostic figure for the nine primary near-eye models, showing residual-versus-fitted patterns and distribution checks. Darkest 10 h mean has the clearest increase in residual spread; the sleep-below-1-lux model has a small observed zero mass not reproduced by its Tweedie fit."
#| out-width: "100%"

knitr::include_graphics(h08_artifact(
  "10_figures", "H08", "H08_near_eye_model_adequacy.png"
))
```

[Source data for the adequacy figure](../../../artifacts/11_source_data/H08/H08_near_eye_model_adequacy_data.csv)

## Site and participant influence

All 153 planned leave-one-site-out refits succeeded. Some point estimates were
direction- or magnitude-sensitive to omitting one site, which reinforces the
inconclusive family-level interpretation. No site was removed. The participant
screen ranks the five largest residual profiles for each main
placement-by-metric fit; it is a transparent review list, not a deletion rule,
and no participant was excluded after seeing it.

```{r}
#| label: leave-one-site-out-summary

loo_summary |>
  arrange(metric_order, placement_label) |>
  transmute(
    metric_order,
    Placement = placement_label,
    Metric = manuscript_name,
    Refits = sprintf("%d / %d", successful_refits, refits),
    `Sign reversal` = sign_reversal_any,
    `Largest relative |change|` = sprintf(
      "%.2f",
      maximum_relative_absolute_change
    ),
    `Most influential omitted site` = most_influential_omitted_site,
    Interpretation = influence_status
  ) |>
  select(-metric_order) |>
  h08_gt(
    title = "Leave-one-site-out stability",
    note = "Relative changes can be large when the full-sample estimate is close to zero. They do not by themselves identify an erroneous site."
  )
```

[Download the participant influence screen](../../../artifacts/08_diagnostics/H08/H08_participant_influence_screen.csv).

# Gap-timing-unaware dataset sensitivity

The strongest dataset comparison uses identical row keys across datasets,
placement, and metric. All 18 exact-common-sample estimates were classified as
stable within model uncertainty. The scenario-specific all-available families
had no BH-adjusted conclusion change. Within those all-available fits,
near-eye wake time above 250 lx melEDI and corrected melEDI dose were
precision-sensitive because a 95% interval's zero-exclusion changed; all chest
comparisons were stable within model uncertainty.

```{r}
#| label: gap-common-sample-table

gap_common_display <- gap_sensitivity |>
  filter(sample_scenario == "main_gap_common_sample") |>
  left_join(
    metric_registry |>
      select(metric_id, response_transform, effect_scale),
    by = "metric_id"
  ) |>
  mutate(
    estimate_main_display = model_to_practical(
      estimate_main,
      response_transform,
      effect_scale
    ),
    conf_low_main_display = model_to_practical(
      conf_low_main,
      response_transform,
      effect_scale
    ),
    conf_high_main_display = model_to_practical(
      conf_high_main,
      response_transform,
      effect_scale
    ),
    estimate_gap_display = model_to_practical(
      estimate_gap_timing_unaware,
      response_transform,
      effect_scale
    ),
    conf_low_gap_display = model_to_practical(
      conf_low_gap_timing_unaware,
      response_transform,
      effect_scale
    ),
    conf_high_gap_display = model_to_practical(
      conf_high_gap_timing_unaware,
      response_transform,
      effect_scale
    )
  )

gap_common_display |>
  transmute(
    metric_order,
    Placement = if_else(placement == "glasses", "Near eye", "Chest"),
    Metric = manuscript_name,
    `Primary effect (95% CI)` = format_effect(
      estimate_main_display,
      conf_low_main_display,
      conf_high_main_display,
      effect_scale
    ),
    `Gap-timing-unaware effect (95% CI)` = format_effect(
      estimate_gap_display,
      conf_low_gap_display,
      conf_high_gap_display,
      effect_scale
    ),
    Sample = format_sample(
      participants_main,
      participant_days_main,
      sites_main
    ),
    `Exact row keys` = exact_common_keys,
    Stability = stability_classification
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "Exact-common-sample dataset sensitivity",
    note = "Effects are per one VLSQ-8 SD with 95% Wald CIs. These comparisons carry no additional p-value screen."
  )
```

```{r}
#| label: fig-gap-common-sample
#| fig-cap: "Primary and gap-timing-unaware VLSQ-8 associations on exactly matched participant-day samples, shown separately for near-eye and chest placements."
#| fig-alt: "Paired comparison of primary and gap-timing-unaware estimates on identical samples for nine metrics at each placement. Estimates are generally close; all eighteen comparisons are classified as stable within model uncertainty."
#| out-width: "100%"

knitr::include_graphics(h08_artifact(
  "10_figures", "H08", "H08_gap_common_sample_effects.png"
))
```

[Source data for the common-sample dataset figure](../../../artifacts/11_source_data/H08/H08_gap_common_sample_effects_data.csv)

```{r}
#| label: all-available-gap-stability

gap_sensitivity |>
  filter(sample_scenario == "all_available") |>
  count(
    Placement = if_else(placement == "glasses", "Near eye", "Chest"),
    Stability = stability_classification,
    name = "Metrics"
  ) |>
  h08_gt(
    title = "Scenario-specific all-available stability classifications",
    note = "These fits can differ in both metric values and available rows; the exact-common-sample table above isolates value-definition changes."
  )
```

# Other prespecified sensitivities

## Photoperiod adjustment and participant summaries

Linear centred photoperiod was added only as a sensitivity covariate, with site
retained and no photoperiod interaction. Participant-summary models give each
participant one outcome value and use the corresponding fixed-effects model.
All 36 additive sensitivity fits converged and retained full-rank fixed
effects. Their estimates were broadly compatible with the participant-day
results; they are not new multiplicity screens.

```{r}
#| label: photoperiod-participant-summary

bind_rows(
  photoperiod |>
    mutate(Sensitivity = "Photoperiod-adjusted"),
  participant_summary |>
    mutate(
      Sensitivity = "Participant summary",
      participant_days = participant_days_contributing
    )
) |>
  transmute(
    metric_order,
    Sensitivity,
    Placement = placement_label,
    Metric = manuscript_name,
    `Effect per VLSQ-8 SD (95% CI)` = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites),
    `Converged / full rank` = paste(converged, fixed_full_rank, sep = " / ")
  ) |>
  arrange(metric_order, Sensitivity, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "Photoperiod-adjusted and participant-summary estimates",
    note = "Effects have 95% Wald CIs. Sample cells are participants / contributing participant-days / sites. TRUE / TRUE means convergence and full fixed-effect rank."
  )
```

## Exactly identified longest period and observed dose

The primary longest-period metric is an observed lower bound. Requiring exact
identification reduced the available sample and did not yield a precise
association: near eye 0.958× (95% CI 0.856–1.072) and chest 0.939× (95% CI
0.855–1.032). On the exactly paired sample, estimates were 0.943× near eye and
0.952× chest, with both intervals crossing one.

The observed-dose comparison holds the corrected-dose fitted sample fixed and
replaces the value with uncorrected observed dose. It is descriptive. The
all-available near-eye estimate was 0.842× (95% CI 0.713–0.995), but the paired
near-eye interval crossed one; the chest all-available interval also reached
one. These sensitivities do not override the primary BH family.

```{r}
#| label: exact-longest-observed-dose

bind_rows(
  exact_longest |>
    mutate(Sensitivity = "Exactly identified longest period"),
  observed_dose |>
    mutate(Sensitivity = "Observed dose on corrected-dose sample")
) |>
  transmute(
    metric_order,
    Sensitivity,
    Placement = placement_label,
    `Sample basis` = recode(
      sample_scenario,
      all_available = "All available",
      paired_common_sample = "Paired placement"
    ),
    `Effect per VLSQ-8 SD (95% CI)` = format_effect(
      estimate_practical_per_sd,
      conf_low_practical_per_sd,
      conf_high_practical_per_sd,
      effect_type
    ),
    Sample = format_sample(participants, participant_days, sites),
    Status = effect_status
  ) |>
  arrange(metric_order, Sensitivity, `Sample basis`, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "Metric-definition sensitivities",
    note = "Estimates have 95% Wald CIs and do not carry an additional p-value family."
  )
```

## Deferred upstream-dependent scenarios

Approved Rule B/Rule C daily-denominator variants and the all-zero-inclusive
scenario are not H08-ready direct inputs. They remain deferred rather than
being approximated within H08. Stage 2 does not require an upstream change
request because the available, hash-verified sensitivity battery addresses the
current result gate. If the author makes those scenarios mandatory before
Stage 3, H08 will write `audit/handoffs/H08_shared_change_request.md` and stop
for the upstream producer.

# Multiplicity audit

All eight inferential families contain exactly nine observed raw and adjusted
p-values. Independent BH recalculation matched the stored results. No family
had an adjusted p-value below 0.05. Paired/common-sample and other sensitivity
estimates carry 95% intervals but no additional p-value screen.

```{r}
#| label: multiplicity-family-audit

family_audit |>
  transmute(
    Family = family_id,
    Role = role,
    `Observed / planned` = sprintf("%d / %d", observed_raw_p, planned_n),
    Method = multiplicity_method,
    `Raw p < 0.05` = raw_significant_n,
    `Adjusted p < 0.05` = adjusted_significant_n,
    `Complete family` = complete_nine_member_family,
    `Independent check` = independent_recalculation_matches
  ) |>
  h08_gt(
    title = "Eight complete nine-member H08 families",
    note = "REPORT-008 uses a leading zero and three decimals, with values below 0.001 shown as <0.001. Bold is reserved for the adjusted decision; none qualifies here."
  )
```

# V0 reconstruction and reconciliation

V0 is reproducible from the archived RData files and rendered outputs. It used
chest as the main placement, near eye as a robustness repeat, default treatment
coding, `(1 | Id)`, and a joint score-plus-site-interaction block as its
association test. It then called `p.adjust()` on one scalar at a time with
`n = 9`. That scalar operation reproduces the displayed strings but is not a
nine-member FDR adjustment. A consistent vector BH calculation is retained in
the audit artifact for comparison.

```{r}
#| label: v0-reproduction-table

v0_reproduction |>
  transmute(
    metric_order,
    Placement = placement,
    Metric = manuscript_name,
    `V0 sample` = format_sample(participants, participant_days, sites),
    `V0 joint raw p` = nh_format_p_value(association_p_raw),
    `V0 scalar display` = displayed_adjusted_p,
    `Scalar display supported` = displayed_supported,
    `Consistent vector BH p` = nh_format_p_value(consistent_vector_bh_p)
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "Exact V0 reconstruction",
    note = "Near-eye V0 displayed >0.9 for every metric. Chest displayed 0.4, 0.7, then >0.9 except 0.3 for pre-sleep; the archived render strings and scalar calculation agree."
  )
```

```{r}
#| label: v0-new-comparison

v0_comparison |>
  transmute(
    metric_order,
    Placement = placement,
    Metric = manuscript_name,
    `V0 joint raw p` = nh_format_p_value(association_p_raw),
    `Audited average raw p` = nh_format_p_value(new_average_p_raw),
    `Audited average BH p` = nh_format_p_value(new_average_p_adjusted),
    `Audited effect (95% CI)` = format_effect(
      new_estimate_practical_per_sd,
      new_conf_low_practical_per_sd,
      new_conf_high_practical_per_sd,
      new_effect_type
    ),
    `Participant-day change` = participant_day_change,
    `Adjusted conclusion changed` = new_conclusion_differs_from_v0_scalar
  ) |>
  arrange(metric_order, Placement) |>
  select(-metric_order) |>
  h08_gt(
    title = "V0-to-audited result comparison",
    note = "The tests are not numerically interchangeable: V0 tests the joint score-plus-interaction block, whereas the audited average test adds one centred score slope after site. No audited BH-adjusted conclusion is positive."
  )
```

The audited implementation therefore does not convert a V0 positive result
into a negative one. Rather, it replaces a nonstandard display adjustment and
a differently defined joint test with the approved average-slope and
heterogeneity families. The defensible Stage 2 conclusion is that the observed
data do not provide multiplicity-adjusted evidence of an association between
VLSQ-8 score and any of the nine near-eye metrics; chest and the sensitivity
battery do not materially strengthen that evidence. Several estimates lean
toward less bright-light exposure at higher scores, but their family-adjusted
evidence and site stability are inconclusive. These are associations, not
physiological responses or health effects.

# Reproducibility and artifact map

The Stage 2 runner fitted all approved models in R 4.6.1, wrote frozen model
frames and bundles, and exported tables, diagnostics, figures, source data,
and manifests. It used maintained `reformulas` formula helpers. The later
METRIC-011 script performed only the bounded L10 maintenance described above
and preserved all other model objects. Neither run used bootstrap, simulation,
or other resampling, so the pilot-runtime gate was not triggered.

```{r}
#| label: execution-environment

execution_environment |>
  h08_gt(
    title = "Recorded execution environment",
    note = "Consequential package versions, commands, and input files are recorded in the H08 manifest artifacts."
  )
```

Primary entry points and frozen outputs:

- [Stage 2 runner](../../../scripts/hypotheses/H08/run_h08_stage2.R)
- [Bounded METRIC-011 reseal](../../../scripts/hypotheses/H08/reseal_h08_l10_metric011.R)
- [H08 contract](../../../scripts/hypotheses/H08/h08_contract.R)
- [H08 modelling helpers](../../../scripts/hypotheses/H08/h08_modeling.R)
- [Focused Stage 2 test](../../../tests/hypotheses/H08/test_h08_stage2.R)
- [Focused METRIC-011 test](../../../tests/hypotheses/H08/test_h08_metric011_reseal.R)
- [Master result table](../../../artifacts/09_tables/H08/H08_model_results_master.csv)
- [Complete diagnostic table](../../../artifacts/08_diagnostics/H08/H08_model_diagnostics.csv)
- [Model-frame rows](../../../artifacts/06_model_data/H08/H08_model_frame_rows.csv)
- [Stage 2 artifact manifest](../../../artifacts/12_manifests/H08/H08_stage2_artifacts.csv)
- [Figure manifest and alt text](../../../artifacts/12_manifests/H08/H08_figure_manifest.csv)

# Stage 2 package status

::: {.callout-important}
## Accepted result package

The author accepted the following as one Stage 2 result package:

1. Accept the eight complete nine-member BH families and the conclusion that
   none has an adjusted p-value below 0.05.
2. Accept the primary near-eye and complementary chest practical estimates,
   95% Wald confidence intervals, exact fitted samples, and separate
   site-heterogeneity interpretation.
3. Accept the retained response families with explicit limitations for
   Darkest 10 h mean residual spread, sleep-below-1-lx zero mass, and the
   timing-dependent corrected dose.
4. Accept the leave-one-site-out interpretation: some estimates are
   site-sensitive, no site is removed, and the overall result remains
   inconclusive.
5. Accept the exact paired-placement display without pooling or equivalence
   claims and without an additional p-value screen.
6. Accept the gap-timing-unaware, photoperiod, participant-summary,
   exactly-identified-longest-period, and observed-dose sensitivity
   interpretations.
7. Accept the V0 reconstruction and the explanation that V0 and the audited
   model test different null hypotheses and use different multiplicity
   procedures.
8. Accept the plain-language claim boundary: no multiplicity-adjusted evidence
   for an H08 association, with uncertain directional patterns and no causal,
   physiological, or health-effect claim.

Stage 3 and Stage 4 were subsequently completed and accepted. The Stage 3
notebook includes the REPORT-012 **Results in brief** callout directly after
the hypothesis and analytical question. METRIC-011 did not alter that callout
or any of the multiplicity, uncertainty, chest, or sensitivity qualifications.
:::
