---
title: "H08 analysis preparation and provenance"
subtitle: "From verified VLSQ-8 scores and personal light-exposure metrics to reported associations"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-h08-preparation
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(knitr)
  library(readr)
  library(scales)
  library(tibble)
  library(tidyr)
})

root_candidates <- unique(c(
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(
    file.path(getwd(), "../../.."),
    winslash = "/",
    mustWork = FALSE
  )
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(
        path,
        "scripts/hypotheses/H08/h08_contract.R"
      ))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(root.dir = root)

source(file.path(root, "scripts/pipeline/paths_io.R"))

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H08 preparation requires R 4.6.1", call. = FALSE)
}

read_h08_csv <- function(relative_path) {
  readr::read_csv(
    file.path(root, relative_path),
    show_col_types = FALSE,
    progress = FALSE,
    na = ""
  )
}

prep_gt <- function(table, font_size = 12) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(max(10, font_size - 2))
    )
}

format_count <- function(value) {
  format(value, big.mark = ",", scientific = FALSE, trim = TRUE)
}

placement_colors <- c(
  "Near eye" = "#0072B2",
  "Chest" = "#D55E00"
)

input_audit <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_input_audit.csv"
)
metric_registry <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_metric_registry.csv"
)
formula_registry <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_formula_registry.csv"
)
run_registry <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_run_registry.csv"
)
frame_index <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_model_frame_index.csv"
)
response_contract <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_response_contract_audit.csv"
)
paired_audit <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_paired_sample_audit.csv"
)
main_gap_audit <- read_h08_csv(
  "artifacts/06_model_data/H08/H08_main_gap_common_sample_audit.csv"
)
score_audit <- read_h08_csv(
  "artifacts/08_diagnostics/H08/H08_vlsq_score_audit.csv"
)
diagnostics <- read_h08_csv(
  "artifacts/08_diagnostics/H08/H08_model_diagnostics.csv"
)
response_gate <- read_h08_csv(
  "artifacts/08_diagnostics/H08/H08_response_family_gate.csv"
)
loo_summary <- read_h08_csv(
  "artifacts/08_diagnostics/H08/H08_leave_one_site_out_summary.csv"
)
family_audit <- read_h08_csv(
  "artifacts/09_tables/H08/H08_family_audit.csv"
)
gap_sensitivity <- read_h08_csv(
  "artifacts/09_tables/H08/H08_gap_timing_unaware_sensitivity.csv"
)
photoperiod <- read_h08_csv(
  "artifacts/09_tables/H08/H08_photoperiod_sensitivity.csv"
)
participant_summary <- read_h08_csv(
  "artifacts/09_tables/H08/H08_participant_summary_sensitivity.csv"
)
exact_longest <- read_h08_csv(
  paste0(
    "artifacts/09_tables/H08/",
    "H08_exactly_identified_longest_period_sensitivity.csv"
  )
)
observed_dose <- read_h08_csv(
  "artifacts/09_tables/H08/H08_observed_dose_sensitivity.csv"
)
sample_support <- read_h08_csv(
  "artifacts/11_source_data/H08/H08_preparation_sample_support.csv"
)
score_distribution <- read_h08_csv(
  paste0(
    "artifacts/11_source_data/H08/",
    "H08_preparation_vlsq_score_distribution.csv"
  )
)
score_site_support <- read_h08_csv(
  paste0(
    "artifacts/11_source_data/H08/",
    "H08_preparation_vlsq_site_support.csv"
  )
)
metric_availability <- read_h08_csv(
  paste0(
    "artifacts/11_source_data/H08/",
    "H08_preparation_metric_availability.csv"
  )
)
site_support <- read_h08_csv(
  "artifacts/11_source_data/H08/H08_preparation_site_support.csv"
)
site_support_summary <- read_h08_csv(
  paste0(
    "artifacts/11_source_data/H08/",
    "H08_preparation_site_support_summary.csv"
  )
)
site_registry <- read_h08_csv("config/site_display_registry.csv") |>
  arrange(.data$display_order)

model_frame_archive <- readRDS(file.path(
  root,
  "artifacts/06_model_data/H08/H08_model_frames.rds"
))
model_frames <- model_frame_archive$model_frames

near_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
```

::: {.callout-note title="About this analysis record"}
This page explains how verified VLSQ-8 questionnaire scores and personal
light-exposure metrics were turned into the H08 results. The companion
[H08 results report](../../../notebooks/hypotheses/H08.html) presents the
scientific findings.

Rendering this page recalculates only file-identity checks, schema and key
checks, and lightweight descriptive summaries from frozen H08 model frames.
It reads stored fits, estimates, 95% confidence intervals, diagnostics,
influence screens, and sensitivity outputs from their recorded files. It does
not fit or refit a model, calculate predictions, re-estimate residual
diagnostics, rerun site-omission analyses, bootstrap, resample, or simulate
data.
:::

## Purpose

H08 asks whether nine duration-, exposure-history-, and level-based personal
light-exposure metrics are associated with VLSQ-8 visual-light-sensitivity
scores. Near-eye measurement is primary because it more closely represents
light near the eyes during wear. All-available chest measurement is
complementary and is not pooled with near-eye measurement. Every inferential
outcome has one row per participant-day.

The predefined preparation sensitivity uses the **gap-timing-unaware
dataset**. It still passed the general requirement of at least 50% valid
minutes per hour and at least 80% valid hours per day, but the timing of the
remaining missing observations was not used for an additional metric-specific
adjustment. For contrast at this first explanation, the primary could be
interpreted as a **time-sensitive primary metric dataset**. Hereafter it is
called simply the primary dataset.

This page identifies the exact data, score rule, samples, formulae, code, and
stored checks behind the accepted report. It documents provenance; it does
not create an alternate analysis.

## Analysis map

```{mermaid}
flowchart TD
  A["Verified VLSQ-8 scores<br/>184 participants"] --> C["Identity, score, schema,<br/>key and support checks"]
  B["Verified personal light exposure<br/>near eye and chest"] --> C
  C --> D["Twelve run definitions<br/>dataset × placement × sample"]
  D --> E["Nine metric-specific<br/>participant-day frames per run"]
  E --> F["Centred VLSQ-8 predictor<br/>one participant SD = 5.540 points"]
  F --> G["Site-only, additive and<br/>site-interaction models"]
  G --> H["Practical effects, 95% intervals<br/>and complete-family BH adjustment"]
  G --> I["Residual, support, serial,<br/>influence and response checks"]
  G --> J["Placement, preparation,<br/>photoperiod and summary sensitivities"]
  H --> K["Tables, accessible figures<br/>and exact source CSV files"]
  I --> K
  J --> K
  K --> L["H08 results report"]
  M["Automated identity and<br/>structural checks"] -. verify .-> C
  M -. verify .-> E
  M -. verify .-> K
```

The arrows show dependencies. They do not mean that this page repeats model
estimation.

## What is calculated when this page opens

```{r}
#| label: tbl-h08-prep-boundary
#| tbl-cap: "Calculations performed when this page is rendered."

tibble::tribble(
  ~Category, ~Calculated_on_render, ~Explanation,
  "File identity and schema", "Yes", "Checks recorded SHA-256 identities, required columns, unique keys, and archive dimensions.",
  "Descriptive support", "Yes", "Summarises stored score counts, metric-specific samples, missing-value reasons, and site contributions.",
  "Model construction and estimation", "No", "Reads the formula registry, model-frame archive, stored estimates, and model records.",
  "Diagnostics and sensitivities", "No", "Reads recorded diagnostic and sensitivity outputs; no check or sensitivity is repeated.",
  "Reader figures", "Display only", "Draws three descriptive figures from the linked, exact source-data CSV files."
) |>
  gt::gt() |>
  gt::cols_width(
    Category ~ gt::pct(24),
    Calculated_on_render ~ gt::pct(18),
    Explanation ~ gt::pct(58)
  ) |>
  prep_gt()
```

## 1. Verified inputs

### File identity

The result-producing analysis begins from nine current inputs whose SHA-256
identities were pinned before the model frames were constructed. Four
historical reconstruction inputs are retained in the complete audit record but
are not used to produce the current results and are therefore omitted below.

```{r}
#| label: tbl-h08-prep-input-identities
#| tbl-cap: "Verified inputs used by the current H08 result-producing analysis."

result_input_roles <- c(
  "normalized_vlsq8",
  "primary_near_eye_metrics",
  "primary_chest_metrics",
  "gap_timing_unaware_metrics",
  "primary_support_provenance",
  "gap_support_provenance",
  "site_display_registry",
  "metric_display_registry",
  "h05_response_contract"
)
input_labels <- c(
  normalized_vlsq8 = "Normalised VLSQ-8 scores",
  primary_near_eye_metrics = "Primary near-eye metrics",
  primary_chest_metrics = "Complementary chest metrics",
  gap_timing_unaware_metrics = "Preparation-sensitivity metrics",
  primary_support_provenance = "Primary metric-support provenance",
  gap_support_provenance = "Preparation-sensitivity support provenance",
  site_display_registry = "Submitted site display registry",
  metric_display_registry = "Manuscript metric display registry",
  h05_response_contract = "Inherited response-family contract"
)

verified_inputs <- input_audit |>
  filter(.data$input_role %in% result_input_roles) |>
  mutate(
    current_sha256 = vapply(
      file.path(root, .data$path),
      artifact_sha256,
      character(1)
    ),
    Input = unname(input_labels[.data$input_role]),
    Path = .data$path,
    `SHA-256 prefix` = substr(.data$current_sha256, 1L, 16L),
    Verified = if_else(
      .data$current_sha256 == .data$expected_sha256,
      "PASS",
      "FAIL"
    )
  )
stopifnot(nrow(verified_inputs) == 9L, all(verified_inputs$Verified == "PASS"))

verified_inputs |>
  select("Input", "Path", "SHA-256 prefix", "Verified") |>
  gt::gt() |>
  gt::cols_width(
    Input ~ gt::pct(24),
    Path ~ gt::pct(54),
    `SHA-256 prefix` ~ gt::pct(15),
    Verified ~ gt::pct(7)
  ) |>
  gt::tab_source_note(
    "Complete 64-character identities remain in artifacts/06_model_data/H08/H08_input_audit.csv."
  ) |>
  prep_gt(11)
```

### Schema, key, content, and value checks

VLSQ-8 is keyed by site and participant. Every model frame is keyed by site,
participant, local date, and metric. The 108-row model-frame index identifies
one exact frame for each of 12 runs and nine metrics.

```{r}
#| label: tbl-h08-prep-integrity-checks
#| tbl-cap: "Bounded integrity checks performed from stored H08 inputs and model frames."

required_frame_columns <- c(
  ".model_row_id", "site", "Id", "participant_key", "local_date",
  "VLSQ8", "VLSQ8_c", "value", "response_value", "metric_id"
)
frames_with_complete_schema <- sum(vapply(
  model_frames,
  function(frame) all(required_frame_columns %in% names(frame)),
  logical(1)
))
frame_duplicate_keys <- sum(vapply(
  model_frames,
  function(frame) anyDuplicated(frame$.model_row_id) > 0L,
  logical(1)
))
index_key_rows <- frame_index |>
  distinct(.data$run_id, .data$metric_id) |>
  nrow()
response_contract_verified <- all(
  response_contract$response_family_matches &
    response_contract$response_transform_matches &
    response_contract$effect_scale_matches
)

integrity_checks <- tibble::tribble(
  ~Check, ~Expected, ~Observed, ~Status,
  "Current input identities", "9 matches", paste0(sum(verified_inputs$Verified == "PASS"), " matches"), if_else(all(verified_inputs$Verified == "PASS"), "PASS", "FAIL"),
  "VLSQ-8 participant key", "184 participants; no missing score or item", paste0(score_audit$participants, " participants; ", score_audit$missing_scores, " missing scores; ", score_audit$missing_item_cells, " missing item cells"), if_else(score_audit$participants == 184L && score_audit$missing_scores == 0L && score_audit$missing_item_cells == 0L, "PASS", "FAIL"),
  "VLSQ-8 reconstruction", "Stored sum rule verified", if_else(score_audit$scoring_rule_verified, "Verified", "Not verified"), if_else(score_audit$scoring_rule_verified, "PASS", "FAIL"),
  "Metric registry", "9 unique participant-day metrics", paste0(n_distinct(metric_registry$metric_id), " unique metrics"), if_else(n_distinct(metric_registry$metric_id) == 9L && all(metric_registry$analysis_unit == "participant_day"), "PASS", "FAIL"),
  "Run registry", "12 unique runs", paste0(n_distinct(run_registry$run_id), " unique runs"), if_else(n_distinct(run_registry$run_id) == 12L, "PASS", "FAIL"),
  "Stored model frames", "108 frames", paste0(length(model_frames), " frames"), if_else(length(model_frames) == 108L, "PASS", "FAIL"),
  "Frame schema and row keys", "108 complete; 0 duplicate-key frames", paste0(frames_with_complete_schema, " complete; ", frame_duplicate_keys, " duplicate-key frames"), if_else(frames_with_complete_schema == 108L && frame_duplicate_keys == 0L, "PASS", "FAIL"),
  "Run–metric index", "108 unique cells", paste0(index_key_rows, " unique cells"), if_else(index_key_rows == 108L && nrow(frame_index) == 108L, "PASS", "FAIL"),
  "Inherited response contract", "9 exact family, transform, and scale matches", paste0(sum(response_contract$response_family_matches & response_contract$response_transform_matches & response_contract$effect_scale_matches), " exact matches"), if_else(response_contract_verified, "PASS", "FAIL"),
  "Paired placement keys", "18 exact pairs", paste0(sum(paired_audit$exact_row_keys_match), " exact pairs"), if_else(all(paired_audit$exact_row_keys_match), "PASS", "FAIL"),
  "Primary–sensitivity common keys", "18 exact pairs", paste0(sum(main_gap_audit$exact_row_keys_match), " exact pairs"), if_else(all(main_gap_audit$exact_row_keys_match), "PASS", "FAIL")
)
stopifnot(all(integrity_checks$Status == "PASS"))

integrity_checks |>
  gt::gt() |>
  gt::cols_width(
    Check ~ gt::pct(26),
    Expected ~ gt::pct(29),
    Observed ~ gt::pct(36),
    Status ~ gt::pct(9)
  ) |>
  prep_gt(11)
```

## 2. VLSQ-8 predictor

### Scale and score reconstruction

VLSQ-8 contains eight ordered item codes from 1 to 5. For every participant,
the stored score equalled the item sum plus 5. The constant offset does not
affect a centred slope, but retaining it is necessary to reproduce the
approved 13–39 score scale. All 184 participants had complete item and total
scores. The score was centred at its observed participant mean, 21.598, and
reported effects compare scores separated by the participant SD of 5.540
points. Scores were not standardised within site.

```{r}
#| label: tbl-h08-prep-score-audit
#| tbl-cap: "VLSQ-8 score reconstruction and observed support."

score_audit |>
  transmute(
    Participants = .data$participants,
    Sites = .data$sites,
    `Observed range` = paste0(.data$observed_min, "–", .data$observed_max),
    Mean = .data$observed_mean,
    `Participant SD` = .data$observed_participant_sd,
    `Missing scores` = .data$missing_scores,
    `Missing item cells` = .data$missing_item_cells,
    `Stored minus item sum` = .data$stored_minus_item_sum_unique,
    Verified = if_else(.data$scoring_rule_verified, "PASS", "FAIL")
  ) |>
  gt::gt() |>
  gt::fmt_number(columns = c(Mean, `Participant SD`), decimals = 3) |>
  gt::fmt_integer(
    columns = c(
      Participants, Sites, `Missing scores`, `Missing item cells`,
      `Stored minus item sum`
    )
  ) |>
  prep_gt(11)
```

```{r}
#| label: fig-h08-prep-vlsq-distribution
#| fig-cap: "Distribution of the verified VLSQ-8 scores. The bars sum to all 184 questionnaire participants."
#| fig-alt: "Bar chart of VLSQ-8 participant counts from score 13 to 39. Scores cluster around the low twenties, with fewer participants at the highest values; all 184 participants are represented."
#| fig-width: 6.692913
#| fig-height: 4.6
#| out-width: "100%"

ggplot(
  score_distribution,
  aes(x = .data$stored_VLSQ8, y = .data$participants)
) +
  geom_col(fill = "#0072B2", width = 0.85) +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.06))) +
  labs(x = "VLSQ-8 score", y = "Participants") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 9, face = "bold")
  )
```

The bars use the exact aggregated
[figure source data](../../../artifacts/11_source_data/H08/H08_preparation_vlsq_score_distribution.csv).
The corresponding
[site-specific score support](../../../artifacts/11_source_data/H08/H08_preparation_vlsq_site_support.csv)
retains the submitted site order, participant counts, ranges, means, and SDs.

## 3. Metrics and analytical samples

### Metric definitions and missing values

The nine outcomes retain the manuscript names and melEDI units. Unsupported
values remain missing; they are never converted to darkness or zero. Four
metrics use Gaussian models after `log10(value + 0.1)` transformation, three
non-negative duration metrics use Tweedie models with a log link, and two use
Gaussian identity models. Practical effects are ratios except for time below
10 lx melEDI before sleep, which is an hour difference.

The longest continuous period above 250 lx melEDI can be a lower bound when a
remaining gap interrupts the observed period. A separate exact-identification
sensitivity retains only participant-days for which both ends of the longest
period are observed. The corrected melEDI dose uses the registered
time-sensitive value; observed uncorrected dose is retained as a descriptive
sensitivity on the same fitted sample.

```{r}
#| label: tbl-h08-prep-metric-contract
#| tbl-cap: "Metric-specific analysis level, response model, transformation, and reported effect scale."

metric_registry |>
  arrange(.data$metric_order) |>
  transmute(
    Order = .data$metric_order,
    Metric = .data$manuscript_name,
    Category = .data$manuscript_category,
    `Analysis level` = "Participant-day",
    `Response model` = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    Transformation = recode(
      .data$response_transform,
      identity = "Identity",
      log10_offset_0.1 = "log10(value + 0.1)"
    ),
    `Reported effect` = recode(
      .data$effect_scale,
      difference = "Difference in hours",
      ratio = "Ratio"
    ),
    Unit = .data$display_unit
  ) |>
  gt::gt(rowname_col = "Order") |>
  gt::cols_width(
    Metric ~ gt::pct(27),
    Category ~ gt::pct(13),
    `Analysis level` ~ gt::pct(13),
    `Response model` ~ gt::pct(14),
    Transformation ~ gt::pct(15),
    `Reported effect` ~ gt::pct(12),
    Unit ~ gt::pct(6)
  ) |>
  prep_gt(10)
```

The preparation source records each available metric value and each explicit
reason for an unavailable participant-day. The compact table below reports
the stored counts without treating unavailable values as observations.

```{r}
#| label: tbl-h08-prep-availability
#| tbl-cap: "Stored metric availability by dataset and placement."

metric_availability |>
  mutate(
    Availability = if_else(
      .data$availability_reason == "available",
      "Available",
      "Unavailable with recorded reason"
    )
  ) |>
  group_by(
    .data$preparation,
    .data$placement_label,
    .data$Availability
  ) |>
  summarise(`Participant-days` = sum(.data$participant_days), .groups = "drop") |>
  gt::gt(groupname_col = "preparation") |>
  gt::cols_label(
    placement_label = "Placement",
    Availability = "Metric-value status"
  ) |>
  gt::fmt_integer(columns = `Participant-days`, use_seps = TRUE) |>
  gt::tab_source_note(
    "Counts sum metric-specific participant-days and therefore are not counts of unique days across metrics. Exact reason-level counts remain in the linked source."
  ) |>
  prep_gt(11)
```

The [availability source](../../../artifacts/11_source_data/H08/H08_preparation_metric_availability.csv)
retains each metric, dataset, placement, availability reason, and count.

### Exact primary and complementary samples

Each model used its metric-specific complete frame. Primary near-eye samples
contained 139–141 participants, 655–816 participant-days, and nine sites.
Complementary chest samples contained 153–154 participants, 743–902
participant-days, and eight sites.

```{r}
#| label: tbl-h08-prep-primary-samples
#| tbl-cap: "Exact all-available primary near-eye and complementary chest samples for each metric."

sample_support |>
  mutate(
    Placement = factor(
      .data$placement_label,
      levels = c("Near eye", "Chest")
    )
  ) |>
  transmute(
    Placement,
    Order = .data$metric_order,
    Metric = .data$manuscript_name,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    Sites = .data$sites,
    `Valid support hours` = .data$metric_support_valid_hours,
    `Expected support hours` = .data$metric_support_expected_hours
  ) |>
  gt::gt(groupname_col = "Placement", rowname_col = "Order") |>
  gt::fmt_integer(
    columns = c(
      Participants, `Participant-days`, Sites, `Valid support hours`,
      `Expected support hours`
    ),
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Metric ~ gt::pct(35),
    Participants ~ gt::pct(12),
    `Participant-days` ~ gt::pct(14),
    Sites ~ gt::pct(8),
    `Valid support hours` ~ gt::pct(15),
    `Expected support hours` ~ gt::pct(16)
  ) |>
  prep_gt(10)
```

The [exact table source](../../../artifacts/11_source_data/H08/H08_preparation_sample_support.csv)
also retains run, metric, placement, row-key, and frame identities.

```{r}
#| label: fig-h08-prep-sample-support
#| fig-cap: "Metric-specific participant and participant-day support in the primary near-eye and complementary chest analyses. Metric numbers follow the preceding table."
#| fig-alt: "Two line-and-point panels compare near-eye and chest support across nine ordered metrics. Participant counts are 139 to 141 near eye and 153 to 154 at the chest. Participant-day counts vary more, from 655 to 816 near eye and 743 to 902 at the chest."
#| fig-width: 6.692913
#| fig-height: 4.9
#| out-width: "100%"

sample_plot_data <- sample_support |>
  select(
    "metric_order", "placement_label", "participants", "participant_days"
  ) |>
  pivot_longer(
    cols = c("participants", "participant_days"),
    names_to = "quantity",
    values_to = "count"
  ) |>
  mutate(
    quantity = recode(
      .data$quantity,
      participants = "Participants",
      participant_days = "Participant-days"
    )
  )

ggplot(
  sample_plot_data,
  aes(
    x = .data$metric_order,
    y = .data$count,
    colour = .data$placement_label,
    group = .data$placement_label
  )
) +
  geom_line(linewidth = 0.65) +
  geom_point(size = 2) +
  facet_wrap(vars(.data$quantity), scales = "free_y", ncol = 1) +
  scale_colour_manual(values = placement_colors) +
  scale_x_continuous(breaks = seq_len(9L)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "Metric order", y = NULL, colour = "Placement") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text = element_text(size = 8),
    axis.title.x = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 8.5),
    legend.position = "top"
  )
```

The plot uses the same
[exact source data](../../../artifacts/11_source_data/H08/H08_preparation_sample_support.csv)
as the sample table.

### Site contributions

The primary analysis represents all nine submitted-manuscript sites. The
complementary chest analysis represents eight because chest data are absent
for Tübingen. Site-specific participant and day counts vary by metric
eligibility, so the ranges below come from the nine exact model frames and
not from a pooled approximation.

```{r}
#| label: tbl-h08-prep-site-support
#| tbl-cap: "Per-site sample ranges across the nine all-available metric frames."

site_support_summary |>
  transmute(
    Analysis = recode(
      .data$placement_label,
      `Near eye` = "Primary near eye",
      Chest = "Complementary chest"
    ),
    `Site order` = .data$display_order,
    Site = .data$site_name,
    `Metric frames` = .data$metric_frames,
    `Participants, min–max` = paste0(
      .data$participants_min,
      "–",
      .data$participants_max
    ),
    `Participant-days, min–max` = paste0(
      format_count(.data$participant_days_min),
      "–",
      format_count(.data$participant_days_max)
    )
  ) |>
  gt::gt(groupname_col = "Analysis", rowname_col = "Site order") |>
  gt::fmt_integer(columns = `Metric frames`) |>
  gt::cols_width(
    Site ~ gt::pct(30),
    `Metric frames` ~ gt::pct(17),
    `Participants, min–max` ~ gt::pct(23),
    `Participant-days, min–max` ~ gt::pct(30)
  ) |>
  gt::tab_source_note(
    "Sites follow the submitted-manuscript order. The complete metric-by-site table retains every exact count."
  ) |>
  prep_gt(11)
```

```{r}
#| label: fig-h08-prep-site-range
#| fig-cap: "Range of participant-day contributions by site across the nine primary near-eye and complementary chest metric frames. Site order and colours follow the submitted manuscript registry."
#| fig-alt: "Two panels show horizontal ranges of participant-day counts for each submitted site across nine metrics. The near-eye panel has nine sites and the chest panel eight. Tübingen contributes only near-eye data; San José has the largest chest contribution."
#| fig-width: 6.692913
#| fig-height: 5.9
#| out-width: "100%"

site_levels <- rev(site_registry$display_name)
site_colors <- site_registry$color_hex
names(site_colors) <- site_registry$display_name
site_plot_data <- site_support_summary |>
  mutate(
    site_name = factor(.data$site_name, levels = site_levels),
    Analysis = recode(
      .data$placement_label,
      `Near eye` = "Primary near eye",
      Chest = "Complementary chest"
    )
  )

ggplot(site_plot_data, aes(y = .data$site_name, colour = .data$site_name)) +
  geom_segment(
    aes(
      x = .data$participant_days_min,
      xend = .data$participant_days_max,
      yend = .data$site_name
    ),
    linewidth = 1.9,
    lineend = "round"
  ) +
  geom_point(aes(x = .data$participant_days_min), size = 2.2) +
  geom_point(aes(x = .data$participant_days_max), size = 2.2) +
  facet_wrap(vars(.data$Analysis), ncol = 2) +
  scale_colour_manual(values = site_colors, drop = FALSE) +
  scale_x_continuous(labels = scales::label_comma()) +
  labs(x = "Participant-days across metric-specific frames", y = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text = element_text(size = 8),
    axis.title.x = element_text(size = 9, face = "bold"),
    legend.position = "none"
  )
```

The plotted ranges have
[exact source data](../../../artifacts/11_source_data/H08/H08_preparation_site_support_summary.csv).
The complete
[metric-by-site source](../../../artifacts/11_source_data/H08/H08_preparation_site_support.csv)
contains all 153 cells across the 18 main all-available metric frames.

## 4. Model construction and inference

### Exact formulae supplied to the models

The primary specification uses fixed site with sum-to-zero contrasts and a
participant random intercept nested within site. The additive coefficient is
therefore a site-adjusted average VLSQ-8 slope rather than a slope at one
reference site. A likelihood-ratio comparison of the site-only and additive
models tests the average association. Comparing the additive and interaction
models tests site heterogeneity. Photoperiod adjustment and participant-level
summary models are structured sensitivities.

The table retains the exact evaluated Wilkinson formulae used for the current
analysis and its two structured sensitivity branches.

```{r}
#| label: tbl-h08-prep-exact-formulas
#| tbl-cap: "Exact evaluated Wilkinson formulae supplied to current H08 models and structured sensitivities."

analysis_labels <- c(
  approved_participant_day = "Participant-day analysis",
  photoperiod_sensitivity = "Photoperiod sensitivity",
  participant_summary_sensitivity = "Participant-summary sensitivity"
)
role_labels <- c(
  site_only = "Site-only comparison",
  additive = "Additive VLSQ-8 model",
  interaction = "Site-interaction model"
)

formula_registry |>
  filter(.data$analysis_kind %in% names(analysis_labels)) |>
  mutate(
    Analysis = unname(analysis_labels[.data$analysis_kind]),
    `Model role` = unname(role_labels[.data$formula_id]),
    Formula = paste0("`", .data$formula, "`")
  ) |>
  select("Analysis", "Model role", "Formula") |>
  gt::gt(groupname_col = "Analysis") |>
  gt::fmt_markdown(columns = Formula) |>
  gt::cols_width(
    `Model role` ~ gt::pct(25),
    Formula ~ gt::pct(75)
  ) |>
  prep_gt(11)
```

### Engines and estimation settings

```{r}
#| label: tbl-h08-prep-model-settings
#| tbl-cap: "Model engines, comparisons, and reporting settings used for H08."

tibble::tribble(
  ~Component, ~Setting, ~Reason,
  "Gaussian outcomes", "lme4::lmer, maximum likelihood for model comparisons", "Allows likelihood-ratio comparison of nested participant-day models on one likelihood basis.",
  "Non-negative duration outcomes", "glmmTMB::glmmTMB, Tweedie family with log link, maximum likelihood", "Accommodates right-skewed, non-negative durations while retaining a participant random intercept.",
  "Site adjustment", "Fixed submitted site order; sum-to-zero contrasts", "Estimates an across-site average VLSQ-8 slope and avoids choosing a reference-site slope.",
  "Repeated observations", "Random intercept for participant nested within site", "Accounts for multiple participant-days from the same person.",
  "Average-association test", "Likelihood-ratio test: site-only versus additive", "Tests whether VLSQ-8 adds information after site adjustment.",
  "Site-heterogeneity test", "Likelihood-ratio test: additive versus site interaction", "Tests whether the VLSQ-8 slope differs by site.",
  "Practical estimand", "Effect per one participant VLSQ-8 SD", "Reports ratios or an hour difference for a 5.540-point score contrast.",
  "Uncertainty", "Two-sided 95% Wald confidence interval", "Keeps uncertainty on the same practical scale as the effect.",
  "Multiplicity", "Benjamini–Hochberg within each complete nine-test family", "Separates placement, dataset, and average-versus-heterogeneity decisions."
) |>
  gt::gt() |>
  gt::cols_width(
    Component ~ gt::pct(22),
    Setting ~ gt::pct(35),
    Reason ~ gt::pct(43)
  ) |>
  prep_gt(10)
```

### Multiplicity families

Average-association and site-heterogeneity tests were adjusted separately.
Near-eye and chest placements and the primary and preparation-sensitivity
datasets were also separate. This gives eight complete nine-test
Benjamini–Hochberg families. The family adjustment was calculated from the
unrounded p-values; display rounding did not affect decisions.

```{r}
#| label: tbl-h08-prep-family-audit
#| tbl-cap: "Complete-family multiplicity audit for the eight declared H08 inferential families."

family_audit |>
  transmute(
    Family = .data$family_id,
    Role = .data$role,
    Comparison = recode(
      .data$comparison_id,
      average_vlsq = "Average association",
      site_heterogeneity = "Site heterogeneity"
    ),
    `Planned / observed` = paste0(
      .data$planned_n,
      " / ",
      .data$observed_raw_p
    ),
    Method = .data$multiplicity_method,
    `Raw p < 0.05` = .data$raw_significant_n,
    `Retained at BH 0.05` = .data$adjusted_significant_n,
    Verified = if_else(
      .data$complete_nine_member_family &
        .data$independent_recalculation_matches,
      "PASS",
      "FAIL"
    )
  ) |>
  gt::gt() |>
  gt::tab_source_note(
    "All eight families contain nine tests; no adjusted p-value met the 0.050 criterion."
  ) |>
  prep_gt(9)
```

## 5. Diagnostics and influence

Every current additive and interaction fit was checked for convergence,
Hessian status, singularity, fixed-effect rank, residual behaviour, response
support, prediction bounds, and serial dependence. Exact response zeros and
the expected zero mass were compared for Tweedie outcomes. The checks retain
three explicit limitations: increasing residual spread for darkest 10 h mean,
a small observed zero mass not reproduced for sleep time below 1 lx melEDI,
and corrected-dose sensitivity to remaining-gap timing. No limitation was used
as a significance screen.

```{r}
#| label: tbl-h08-prep-response-gate
#| tbl-cap: "Stored response-family diagnostic disposition for the nine H08 metrics."

response_gate |>
  transmute(
    Order = .data$metric_order,
    Metric = .data$manuscript_name,
    Model = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    `Inferential targets` = .data$inferential_targets,
    `Major failures` = .data$major_failures,
    `Review targets` = .data$review_targets,
    `Average non-estimable` = .data$average_non_estimable,
    `Interaction non-estimable` = .data$interaction_non_estimable,
    Disposition = recode(
      .data$family_gate_status,
      PASS = "Pass",
      RETAIN_WITH_EXPLICIT_LIMITATIONS = "Retain with explicit limitations"
    )
  ) |>
  gt::gt(rowname_col = "Order") |>
  gt::cols_width(
    Metric ~ gt::pct(30),
    Model ~ gt::pct(15),
    `Inferential targets` ~ gt::pct(10),
    `Major failures` ~ gt::pct(9),
    `Review targets` ~ gt::pct(9),
    `Average non-estimable` ~ gt::pct(9),
    `Interaction non-estimable` ~ gt::pct(9),
    Disposition ~ gt::pct(18)
  ) |>
  prep_gt(9)
```

All 153 planned leave-one-site-out refits succeeded. Some estimates changed
direction or magnitude after one site was omitted, which supports cautious,
inconclusive interpretation. It does not identify a site for removal. The
participant-influence screen likewise ranks observations for review and is
not a deletion rule; no site or participant was removed after inspection.

```{r}
#| label: tbl-h08-prep-site-influence
#| tbl-cap: "Leave-one-site-out stability of primary-dataset average associations."

loo_summary |>
  arrange(.data$metric_order, .data$placement_label) |>
  transmute(
    Order = .data$metric_order,
    Placement = .data$placement_label,
    Metric = .data$manuscript_name,
    Refits = paste0(.data$successful_refits, " / ", .data$refits),
    `Sign reversal` = if_else(.data$sign_reversal_any, "Yes", "No"),
    `Largest relative absolute change` = sprintf(
      "%.2f",
      .data$maximum_relative_absolute_change
    ),
    `Most influential omitted site` = .data$most_influential_omitted_site,
    Interpretation = .data$influence_status
  ) |>
  gt::gt(rowname_col = "Order") |>
  gt::tab_source_note(
    "Relative changes can be large when the full-sample estimate is close to zero; they do not by themselves identify an erroneous site."
  ) |>
  prep_gt(9)
```

The [complete diagnostic table](../../../artifacts/08_diagnostics/H08/H08_model_diagnostics.csv)
and [participant-influence screen](../../../artifacts/08_diagnostics/H08/H08_participant_influence_screen.csv)
retain the model-level evidence.

## 6. Sensitivity analyses

Sensitivity branches preserve the accepted estimand unless their purpose
explicitly changes sample construction, adjustment, or metric definition.
They describe stability and do not promote an association that was not
retained in the primary multiplicity family.

```{r}
#| label: tbl-h08-prep-sensitivity-map
#| tbl-cap: "H08 sensitivity branches, exact stored scope, and interpretive role."

gap_common <- gap_sensitivity |>
  filter(.data$sample_scenario == "main_gap_common_sample")
photo_summary <- photoperiod |>
  summarise(
    fits = n(),
    converged = sum(.data$converged),
    full_rank = sum(.data$fixed_full_rank)
  )
participant_summary_counts <- participant_summary |>
  summarise(
    fits = n(),
    converged = sum(.data$converged),
    full_rank = sum(.data$fixed_full_rank)
  )

tibble::tribble(
  ~Analysis, ~Exact_stored_scope, ~Purpose, ~Interpretive_rule,
  "Paired/common placement", paste0(nrow(paired_audit), " exact metric-by-dataset pairs; all row keys matched"), "Separates placement from unequal all-available samples.", "Compare separately fitted matched effects; do not infer equivalence.",
  "Preparation sensitivity on common samples", paste0(nrow(gap_common), " metric-by-placement comparisons; all exact common keys; all classified stable within model uncertainty"), "Checks sensitivity to metric-specific use of remaining-gap timing.", "Uses identical participant-day keys within metric and placement.",
  "Preparation sensitivity on all available samples", "18 metric-by-placement comparisons with separate complete nine-test families", "Checks the combined effect of preparation and resulting sample availability.", "Family-adjusted conclusions remain scenario-specific.",
  "Photoperiod adjustment", paste0(photo_summary$fits, " fits; ", photo_summary$converged, " converged and ", photo_summary$full_rank, " full rank"), "Adds centred linear photoperiod while retaining site.", "No photoperiod interaction and no new multiplicity family.",
  "Participant summaries", paste0(participant_summary_counts$fits, " fits; ", participant_summary_counts$converged, " converged and ", participant_summary_counts$full_rank, " full rank"), "Gives each participant one outcome value.", "Uses fixed site and reports descriptive compatibility with participant-day estimates.",
  "Exactly identified longest period", paste0(nrow(exact_longest), " placement-by-sample estimates"), "Removes lower-bound ambiguity when a gap can interrupt the longest continuous period.", "Retains only participant-days with an exactly observed period.",
  "Observed melEDI dose", paste0(nrow(observed_dose), " placement-by-sample estimates on corrected-dose fitted frames"), "Separates observed uncorrected dose from the corrected dose definition.", "Descriptive 95% intervals; no additional p-value screen.",
  "Leave one site out", paste0(sum(loo_summary$successful_refits), " successful refits"), "Assesses dependence on one submitted site.", "A stability description, not an exclusion or significance rule."
) |>
  gt::gt() |>
  gt::cols_label(
    Exact_stored_scope = "Exact stored scope",
    Interpretive_rule = "Interpretive rule"
  ) |>
  gt::cols_width(
    Analysis ~ gt::pct(20),
    Exact_stored_scope ~ gt::pct(28),
    Purpose ~ gt::pct(24),
    Interpretive_rule ~ gt::pct(28)
  ) |>
  prep_gt(9)
```

The paired-placement display in the results report puts near eye on the
x-axis and chest on the y-axis, includes only exact common participant-day
samples, and uses equal axis geometry, an identity line, and null lines. It
describes agreement and differences without claiming equivalence. The stored
component estimates and 95% confidence intervals were not recomputed for the
display.

Complete sensitivity records are available for the
[preparation comparison](../../../artifacts/09_tables/H08/H08_gap_timing_unaware_sensitivity.csv),
[photoperiod adjustment](../../../artifacts/09_tables/H08/H08_photoperiod_sensitivity.csv),
[participant summaries](../../../artifacts/09_tables/H08/H08_participant_summary_sensitivity.csv),
[exactly identified longest period](../../../artifacts/09_tables/H08/H08_exactly_identified_longest_period_sensitivity.csv),
and [observed melEDI dose](../../../artifacts/09_tables/H08/H08_observed_dose_sensitivity.csv).

## 7. Code map and execution order

### Function modules

```{r}
#| label: tbl-h08-prep-module-map
#| tbl-cap: "Function modules used by the H08 executable analysis."

tibble::tribble(
  ~Module, ~Reads_or_defines, ~Contribution, ~Why_separate,
  "scripts/pipeline/paths_io.R", "Artifact-writing and SHA-256 helpers", "Writes typed artifacts with hashes, byte sizes, producers, and R version.", "Keeps identity rules common across analyses.",
  "scripts/pipeline/multiplicity.R", "Complete-family adjustment helpers", "Applies complete nine-value Benjamini–Hochberg vectors and verifies family size.", "Separates multiplicity from individual model code.",
  "scripts/pipeline/p_value_display.R", "P-value display rules", "Formats p-values without changing unrounded inferential decisions.", "Keeps display policy separate from inference.",
  "scripts/hypotheses/H08/h08_contract.R", "Score, metrics, runs, formulae, families, and pinned inputs", "Defines exact H08 estimands and analysis registries.", "Provides one auditable specification before fitting.",
  "scripts/hypotheses/H08/h08_modeling.R", "Frame construction, engines, effects, diagnostics, and sensitivities", "Implements scientific operations against the contract.", "Keeps reusable operations outside the executable driver."
) |>
  gt::gt() |>
  gt::cols_width(
    Module ~ gt::pct(25),
    Reads_or_defines ~ gt::pct(23),
    Contribution ~ gt::pct(30),
    Why_separate ~ gt::pct(22)
  ) |>
  prep_gt(10)
```

### Executable entry points

```{r}
#| label: tbl-h08-prep-script-map
#| tbl-cap: "H08 scripts and documents in analytical and reporting order."

tibble::tribble(
  ~Order, ~Entry_point, ~Reads, ~Produces, ~Runs_when_page_renders,
  1L, "run_h08_stage2.R", "Verified score and metric inputs, H08 contract and functions", "108 frames, fitted models, effects, diagnostics, influence and sensitivity records", "No",
  2L, "rebuild_h08_reader_figures.R", "Stored H08 figure-source CSV files", "Five result figures exported natively at 170 mm and their display manifest", "No",
  3L, "notebooks/hypotheses/H08.qmd", "Stored scientific outputs and reader assets", "The standalone H08 results report", "No",
  4L, "build_h08_preparation_artifacts.R", "Frozen H08 frames, registries, samples and scores", "Six descriptive source-data CSV files for this page", "No; run before rendering",
  5L, "H08_analysis_preparation.qmd", "Stored H08 artifacts and descriptive source data", "This preparation and provenance companion", "Yes",
  6L, "build_h08_figure_physical_size_qa.R", "Eight final reader figures and their physical metadata", "A4 inspection sheets and a figure readability record", "No; run after rendering",
  7L, "build_h08_preparation_report_manifest.R", "Final source, website copy, HTML, scripts, outputs, figures, source data, and QA", "Non-circular preparation-report manifest", "No; run after rendering and QA"
) |>
  gt::gt(rowname_col = "Order") |>
  gt::cols_width(
    Entry_point ~ gt::pct(24),
    Reads ~ gt::pct(27),
    Produces ~ gt::pct(35),
    Runs_when_page_renders ~ gt::pct(14)
  ) |>
  prep_gt(9)
```

The scientific model run used R 4.6.1 and the project library. These commands
are shown for reproducibility; this page does not execute them.

```bash
Rscript --vanilla scripts/hypotheses/H08/run_h08_stage2.R
Rscript --vanilla scripts/hypotheses/H08/rebuild_h08_reader_figures.R
quarto render notebooks/hypotheses/H08.qmd --profile nathealth
Rscript --vanilla scripts/hypotheses/H08/build_h08_preparation_artifacts.R
quarto render audit/hypotheses/H08/H08_analysis_preparation.qmd --profile nathealth
Rscript --vanilla scripts/hypotheses/H08/build_h08_figure_physical_size_qa.R
```

## 8. Output and claim provenance

The frame index records exact participant, participant-day, site, row-key,
and frame identities for every run and metric. The model manifest connects
each frame to its site-only, additive, and interaction objects. Stored result,
diagnostic, and sensitivity tables then connect those fits to report tables
and figures. Every reader figure links to an exact source CSV so it need not be
digitised.

```{r}
#| label: tbl-h08-prep-output-identities
#| tbl-cap: "Principal stored H08 outputs and their current identities."

output_registry <- tibble::tribble(
  ~Output, ~Path, ~Producer,
  "Model-frame archive", "artifacts/06_model_data/H08/H08_model_frames.rds", "run_h08_stage2.R",
  "Model-frame index", "artifacts/06_model_data/H08/H08_model_frame_index.csv", "run_h08_stage2.R",
  "Inferential model archive", "artifacts/07_models/H08/H08_inferential_model_objects.rds", "run_h08_stage2.R",
  "Model manifest", "artifacts/07_models/H08/H08_model_manifest.csv", "run_h08_stage2.R",
  "Master results", "artifacts/09_tables/H08/H08_model_results_master.csv", "run_h08_stage2.R",
  "Model diagnostics", "artifacts/08_diagnostics/H08/H08_model_diagnostics.csv", "run_h08_stage2.R",
  "Site-omission summary", "artifacts/08_diagnostics/H08/H08_leave_one_site_out_summary.csv", "run_h08_stage2.R",
  "Paired placement source", "artifacts/11_source_data/H08/H08_paired_placement_effects_data.csv", "run_h08_stage2.R",
  "Figure readability record", "artifacts/12_manifests/H08/H08_figure_physical_size_qa.csv", "build_h08_figure_physical_size_qa.R",
  "H08 results report", "_build/nathealth/notebooks/hypotheses/H08.html", "notebooks/hypotheses/H08.qmd"
) |>
  filter(file.exists(file.path(root, .data$Path))) |>
  rowwise() |>
  mutate(
    `SHA-256 prefix` = substr(
      artifact_sha256(file.path(root, .data$Path)),
      1L,
      16L
    ),
    Bytes = file.info(file.path(root, .data$Path))$size
  ) |>
  ungroup()

output_registry |>
  gt::gt() |>
  gt::fmt_integer(columns = Bytes, use_seps = TRUE) |>
  gt::cols_width(
    Output ~ gt::pct(20),
    Path ~ gt::pct(45),
    Producer ~ gt::pct(19),
    `SHA-256 prefix` ~ gt::pct(12),
    Bytes ~ gt::pct(8)
  ) |>
  prep_gt(9)
```

The preparation-report manifest is created only after the final profile render
and physical-size figure inspection. It records complete identities for the
authoring source, byte-identical website source copy, rendered HTML, page
assets, H08 scripts and tests, scientific outputs, source data, shared
configuration snapshot, decision records, and R environment. It excludes
itself and the handoff that quotes its identity, avoiding a circular checksum
chain.

## 9. How to examine or reproduce H08

A reader can follow the analysis in five passes:

1. inspect `H08_input_audit.csv`, `h08_contract.R`, and the score, metric,
   formula, and run registries to verify frozen inputs and specifications;
2. compare the exact samples above with `H08_model_frame_index.csv` and the
   108 frames in `H08_model_frames.rds`;
3. follow `run_h08_stage2.R` through `h08_modeling.R` to connect each frame to
   its stored object, practical effect, 95% interval, likelihood-ratio test,
   diagnostic, and sensitivity record;
4. inspect all eight complete nine-value multiplicity families before
   interpreting an individual association; and
5. read the [H08 results report](../../../notebooks/hypotheses/H08.html)
   alongside its numerical tables, accessible figures, and exact source CSV
   files.

This preparation record documents a frozen analysis. If an input, score rule,
metric definition, sample rule, response model, or estimand changes, the
affected scientific pipeline must be rerun before this page is refreshed.

## Environment

```{r}
#| label: tbl-h08-prep-environment
#| tbl-cap: "Consequential R environment recorded for H08 analysis and reporting."

package_names <- c(
  "dplyr", "readr", "digest", "lme4", "glmmTMB", "ggplot2", "gt"
)
package_versions <- vapply(
  package_names,
  function(package) as.character(utils::packageVersion(package)),
  character(1)
)

bind_rows(
  tibble::tibble(
    Component = "R",
    Version = as.character(getRversion()),
    Role = "Authoritative scientific computation and Quarto rendering"
  ),
  tibble::tibble(
    Component = package_names,
    Version = unname(package_versions),
    Role = c(
      "Data manipulation",
      "Typed CSV input",
      "SHA-256 and model-frame identities",
      "Gaussian mixed models",
      "Tweedie mixed models",
      "Figures",
      "Reader-facing tables"
    )
  )
) |>
  gt::gt(rowname_col = "Component") |>
  gt::cols_width(
    Version ~ gt::pct(18),
    Role ~ gt::pct(62)
  ) |>
  gt::tab_source_note(
    "The project renv.lock records the full dependency set; artifact manifests record R 4.6.1 with each output identity."
  ) |>
  prep_gt()
```

```{r}
#| label: session-info-h08-preparation
#| include: false

sessionInfo()
```
