---
title: "H03 near-eye no-time GAMM test balloon"
subtitle: "Conditional mixed-model output and diagnostics"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: true
    code-summary: "Show R code"
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
input_path <- normalizePath(knitr::current_input(dir = TRUE), winslash = "/")
root <- normalizePath(
  file.path(dirname(input_path), "..", "..", "..", ".."),
  winslash = "/",
  mustWork = TRUE
)
project_library <- file.path(
  root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

read_h03_csv <- function(relative_path) {
  readr::read_csv(file.path(root, relative_path), show_col_types = FALSE)
}

model_summary <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_model_summary.csv"
)
model_runs <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_model_runs.csv"
)
category_means <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_category_means.csv"
)
category_ratios <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_category_ratios.csv"
)
variance_components <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_variance_components.csv"
)
smooth_table <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_smooth_table.csv"
)
r_squared <- read_h03_csv(
  "artifacts/09_tables/H03/H03_no_time_gamm_near_eye_pilot_r_squared.csv"
)
residual_acf <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_no_time_gamm_near_eye_pilot_residual_acf.csv"
)
zero_structure <- read_h03_csv(
  "artifacts/08_diagnostics/H03/H03_no_time_gamm_near_eye_pilot_zero_structure.csv"
)

fmt_int <- function(x) format(round(x), big.mark = ",", scientific = FALSE)
fmt_num <- function(x, digits = 2) {
  formatC(x, format = "f", digits = digits, big.mark = ",")
}
fmt_ci <- function(estimate, lower, upper, digits = 1) {
  paste0(
    fmt_num(estimate, digits),
    " (",
    fmt_num(lower, digits),
    " to ",
    fmt_num(upper, digits),
    ")"
  )
}
fmt_variance <- function(x) {
  ifelse(
    abs(x) < 0.001,
    formatC(x, format = "e", digits = 2),
    fmt_num(x, 3)
  )
}
fmt_flexible <- function(x) {
  ifelse(
    abs(x) < 0.001 | abs(x) >= 10000,
    formatC(x, format = "e", digits = 2),
    fmt_num(x, 3)
  )
}
fmt_flexible_ci <- function(estimate, lower, upper) {
  paste0(
    fmt_flexible(estimate),
    " (",
    fmt_flexible(lower),
    " to ",
    fmt_flexible(upper),
    ")"
  )
}

effective_thread_note <- if (grepl("openMP not available", model_summary$warnings)) {
  paste(
    "The call used `nthreads = 2`, but this local mgcv build reported that",
    "OpenMP was unavailable and therefore executed the computation on one thread."
  )
} else {
  "The call used `nthreads = 2` without a threading warning."
}
```

::: {.callout-important}
## Answer in brief

The two-step model is **computationally estimable**: both runs reached full
convergence and retained all `r model_summary$rank` of
`r model_summary$coefficients` coefficients. The AR1 correction reduced
boundary-aware lag-1 residual correlation from
`r fmt_num(model_runs$lag1_residual_correlation[model_runs$stage == "First run: no AR1"], 3)`
to `r fmt_num(model_summary$pearson_residual_lag1, 3)`. However, the corrected
model is **not non-singular in the practical mixed-model sense**: its
participant-day variance fell to
`r formatC(model_summary$participant_day_variance, format = "e", digits = 2)`
and its effective df to `r fmt_num(model_summary$participant_day_edf, 3)` of
792. The Gaussian residuals are nearly symmetric but retain heavy tails and
some short-lag dependence. The supplied construction therefore fits and
converges, but its participant-day term is unsupported after AR1 correction;
it should not replace the accepted H03 primary analysis unchanged.
:::

## Scope

This volatile test balloon fits **one near-eye mean structure in two runs** to
the accepted H03 primary sample: first without residual correlation, then once
with AR1 correction. It does not fit the chest placement, time-of-day terms,
site interactions, alternative distributions, or sensitivity models. It is a
conditional mixed-model construction exercise and does not replace the
accepted population-mean clustered-GLM analysis.

The sample contains `r fmt_int(model_summary$observations)` participant-hours
from `r model_summary$participants` participants and
`r model_summary$participant_days` participant-days across
`r model_summary$sites` sites. All seven prespecified reported primary-light-source
categories are retained.

## Exact model

```{r}
#| label: model-call
#| eval: false
no_time_gamm <- mgcv::bam(
  LightLogR::log_zero_inflated(geo_medi_1h) ~
    light_source +
    site +
    s(participant, bs = "re") +
    s(participant_day, bs = "re"),
  data = near_eye_primary,
  method = "fREML",
  discrete = TRUE,
  nthreads = 2L,
  gc.level = 1L,
  drop.unused.levels = TRUE
)

rho <- unname(h03_boundary_lag_correlation(
  residuals(no_time_gamm, type = "response"),
  near_eye_primary$AR_start,
  lag = 1L
)[["correlation"]])
rho <- pmin(0.95, pmax(-0.95, rho))

no_time_gamm_ar1 <- mgcv::bam(
  LightLogR::log_zero_inflated(geo_medi_1h) ~
    light_source +
    site +
    s(participant, bs = "re") +
    s(participant_day, bs = "re"),
  data = near_eye_primary,
  method = "fREML",
  discrete = TRUE,
  rho = rho,
  AR.start = near_eye_primary$AR_start,
  nthreads = 2L,
  gc.level = 1L,
  drop.unused.levels = TRUE
)
```

```{r}
#| label: model-specification-table
specification_table <- tibble::tribble(
  ~Component, ~Specification,
  "Outcome", "log10(exact zero-aware one-hour geometric melEDI + 0.1)",
  "Fixed effects", "Seven-level reported primary light source + nine-level site",
  "Random effects", "Participant intercept + participant-day intercept, represented as mgcv random-effect smooths",
  "Family and link", "Gaussian identity (bam() default) on the shifted-log outcome",
  "Estimation", "bam(), fREML, discrete fitting; first run rho = 0",
  "Second-run AR1", paste0("rho = ", fmt_num(model_summary$rho_estimate, 6), ", estimated from boundary-aware lag-1 first-run response residuals; AR starts reset at declared sequence boundaries"),
  "Category estimand", "Sites weighted equally; participant and participant-day effects set to zero; back-transformed only after averaging on the link scale",
  "Indoor contrasts", "Each non-reference category divided by indoor electric light; six P-values adjusted together using Benjamini-Hochberg",
  "Deliberately omitted", "Time of day, site interaction, chest placement, alternative families, and sensitivities"
)
knitr::kable(
  specification_table,
  align = c("l", "l"),
  caption = "Model specification used in the bounded test balloon."
)
```

`r effective_thread_note` The recorded elapsed time was
`r fmt_num(model_summary$elapsed_seconds, 1)` seconds.

```{r}
#| label: model-run-table
run_table <- model_runs |>
  dplyr::transmute(
    Run = .data$stage,
    `AR1 rho` = fmt_num(.data$rho, 3),
    Convergence = .data$convergence,
    Rank = paste0(.data$rank, "/", .data$coefficients),
    `Lag-1 residual correlation` = fmt_num(
      .data$lag1_residual_correlation,
      3
    ),
    `Participant-day variance` = fmt_variance(
      .data$participant_day_variance
    ),
    `Participant-day edf` = fmt_num(.data$participant_day_edf, 3),
    `Boundary signal` = ifelse(
      .data$participant_day_boundary,
      "Yes",
      "No"
    )
  )
knitr::kable(
  run_table,
  align = c("l", "r", "l", "r", "r", "r", "r", "l"),
  caption = "First-run estimation and the single requested AR1-corrected second run."
)
```

## Conditional category results

These second-run AR1-corrected estimates are fixed-effect predictions for a participant and
participant-day whose random effects are zero. Site coefficients are averaged
with the same weight for every site, and the resulting log-scale estimate is
back-transformed once at the end with `10^x - 0.1`. They should therefore be read as
**equal-weighted-across-sites conditional means**, not as population-marginal
means from the accepted primary clustered GLM.

```{r}
#| label: category-mean-table
means_table <- category_means |>
  dplyr::transmute(
    Category = .data$light_source,
    `Expected melEDI, lx (95% CI)` = fmt_ci(
      .data$expected_mel_edi_lx,
      .data$conf_low_lx,
      .data$conf_high_lx,
      1
    )
  )
knitr::kable(
  means_table,
  align = c("l", "r"),
  caption = paste(
    "Conditional one-hour melEDI means with pointwise 95% confidence",
    "intervals; sites receive equal weight."
  )
)
```

```{r}
#| label: indoor-ratio-table
ratio_table <- category_ratios |>
  dplyr::transmute(
    Category = .data$light_source,
    `Shifted-geometric factor to indoor electric (95% CI)` = fmt_ci(
      .data$ratio_to_indoor,
      .data$conf_low,
      .data$conf_high,
      2
    ),
    `BH-adjusted P` = ifelse(
      .data$significant_bh,
      paste0("**", .data$p_display_bh, "**"),
      .data$p_display_bh
    )
  )
knitr::kable(
  ratio_table,
  align = c("l", "r", "r"),
  escape = FALSE,
  caption = paste(
    "Conditional multiplicative factors for (melEDI + 0.1) relative to indoor electric light. Confidence intervals are",
    "unadjusted; P-values form one six-contrast Benjamini-Hochberg family."
  )
)
```

```{r}
#| label: result-figure
#| fig-cap: "Conditional near-eye category means and ratios from the no-time GAMM test balloon. The melEDI display uses the approved display-only symmetric-log transformation while retaining original-unit labels and untransformed source values."
#| out-width: "100%"
knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_no_time_gamm_near_eye_pilot_results.png"
))
```

## Random-effect and overall fit summaries

```{r}
#| label: variance-component-table
variance_table <- variance_components |>
  dplyr::filter(.data$.component != "scale") |>
  dplyr::left_join(
    smooth_table |>
      dplyr::select(.data$stage, .data$smooth, .data$edf, .data$Ref.df),
    by = c("stage", ".component" = "smooth"),
    relationship = "one-to-one"
  ) |>
  dplyr::mutate(
    Stage = dplyr::recode(
      .data$stage,
      `First run: no AR1` = "First: no AR1",
      `Second run: AR1 corrected` = "Second: AR1 corrected"
    ),
    Component = dplyr::recode(
      .data$.component,
      `s(participant)` = "Participant intercept",
      `s(participant_day)` = "Participant-day intercept"
    )
  ) |>
  dplyr::transmute(
    .data$Stage,
    .data$Component,
    `Variance` = fmt_variance(.data$.variance),
    `SD (95% CI)` = fmt_flexible_ci(
      .data$.std_dev,
      .data$.lower_ci,
      .data$.upper_ci
    ),
    `Effective df` = fmt_num(.data$edf, 3),
    `Reference df` = fmt_int(.data$Ref.df)
  )
knitr::kable(
  variance_table,
  align = c("l", "l", "r", "r", "r", "r"),
  caption = "Random-intercept variance components on the log10(melEDI + 0.1) scale before and after AR1 correction."
)
```

Before AR1 correction, the participant-day variance was
`r fmt_num(model_runs$participant_day_variance[model_runs$stage == "First run: no AR1"], 3)`
with effective df
`r fmt_num(model_runs$participant_day_edf[model_runs$stage == "First run: no AR1"], 1)`.
After correction it fell to
`r formatC(model_summary$participant_day_variance, format = "e", digits = 2)`
with effective df `r fmt_num(model_summary$participant_day_edf, 3)`. This is
effective boundary collapse: the AR1 process and participant-day intercept
compete for within-day dependence, and the corrected fit assigns essentially
none of the remaining variation to the day intercept. The optimizer can fully
converge while this mixed-model component is unsupported.

The model-level adjusted R² reported by `summary.gam()` was
`r fmt_num(model_summary$adjusted_r_squared, 3)`, and deviance explained was
`r scales::percent(model_summary$deviance_explained, accuracy = 0.1)`. These
summaries include the fitted random-effect smooths.

```{r}
#| label: squared-error-r2-table
r2_table <- r_squared |>
  dplyr::mutate(
    Prediction = dplyr::recode(
      .data$prediction,
      `fixed effects only` = "Fixed effects only (random effects = 0)",
      `full conditional fit` = "Full conditional fit"
    ),
    Weighting = dplyr::recode(
      .data$weighting,
      `participant-hour weighted` = "Every participant-hour weighted equally",
      `participant balanced` = "Every participant weighted equally"
    )
  ) |>
  dplyr::transmute(
    .data$Prediction,
    .data$Weighting,
    `Shifted-log squared-error R²` = fmt_num(.data$r_squared, 3)
  )
knitr::kable(
  r2_table,
  align = c("l", "l", "r"),
  caption = "Descriptive in-sample squared-error summaries; these are not partial-effect R² statistics."
)
```

The participant-hour version lets a participant with more retained hours
contribute more to squared error. The participant-balanced version first gives
each participant the same total weight, regardless of their number of retained
hours. Their near-identical values here indicate that unequal numbers of hours
per participant do not materially drive this particular descriptive summary.

## Diagnostics

```{r}
#| label: diagnostic-summary-table
diagnostic_table <- tibble::tribble(
  ~Domain, ~Observed, ~Interpretation,
  "Optimizer", "Full convergence in both runs", "Pass: both optimizers report full convergence",
  "Coefficient rank", paste0(model_summary$rank, "/", model_summary$coefficients), "Pass: full rank",
  "Smoothing Hessian", paste0("minimum eigenvalue ", formatC(model_summary$smoothing_hessian_minimum_eigenvalue, format = "e", digits = 2)), "Pass: positive definite",
  "Participant-day boundary", paste0("variance = ", formatC(model_summary$participant_day_variance, format = "e", digits = 2), "; edf = ", fmt_num(model_summary$participant_day_edf, 3), "/792"), "Fail: the second-run participant-day intercept is effectively collapsed",
  "Residual scale pattern", paste0("Spearman rho(|AR-standardized residual|, fitted) = ", fmt_num(model_summary$absolute_residual_fitted_spearman, 3)), "Mild concern: remaining fitted-value dependence",
  "Residual shape", paste0("skewness = ", fmt_num(model_summary$residual_skewness, 3), "; excess kurtosis = ", fmt_num(model_summary$residual_excess_kurtosis, 3)), "Center is nearly symmetric; tails remain heavier than Gaussian",
  "Observed zeros", paste0(scales::percent(model_summary$observed_zero_fraction, accuracy = 0.1), " map to -1 after transformation"), "The transform retains zeros, but the continuous Gaussian likelihood does not model their point probability",
  "Within-sequence dependence", paste0("lag 1: ", fmt_num(model_runs$lag1_residual_correlation[model_runs$stage == "First run: no AR1"], 3), " before to ", fmt_num(model_summary$pearson_residual_lag1, 3), " after AR1 (", fmt_int(model_summary$pearson_residual_lag1_pairs), " pairs)"), "Improved substantially, but non-trivial short-lag correlation remains"
)
knitr::kable(
  diagnostic_table,
  align = c("l", "l", "l"),
  caption = "Diagnostic interpretation for the single fitted model."
)
```

```{r}
#| label: diagnostic-figure
#| fig-cap: "Diagnostics for the two-run no-time near-eye GAMM. Panel A retains individual AR-standardized residual points and a binned center/interquartile overlay. Panel B is a visual Gaussian reference, not a formal normality test. Panel C compares boundary-aware correlations before and after AR1 correction. Panel D shows the shifted-log response and its observed zero spike at -1."
#| out-width: "100%"
knitr::include_graphics(file.path(
  root,
  "artifacts/10_figures/H03/H03_no_time_gamm_near_eye_pilot_diagnostics.png"
))
```

The AR1 correction reduces the one-hour residual correlation from
`r fmt_num(residual_acf$correlation[residual_acf$stage == "First run: no AR1" & residual_acf$lag == 1], 3)`
to
`r fmt_num(residual_acf$correlation[residual_acf$stage == "Second run: AR1 corrected" & residual_acf$lag == 1], 3)`.
It remains `r fmt_num(residual_acf$correlation[residual_acf$stage == "Second run: AR1 corrected" & residual_acf$lag == 3], 3)`
at three hours. The second run therefore improves but does not completely
whiten the residual sequence. At the same time, the participant-day random
intercept collapses, so the reduction cannot be described as a clean fit of
both dependence components.

## Decision from the test balloon

1. **Can it be fitted?** Yes, on all
   `r fmt_int(model_summary$observations)` near-eye participant-hours.
2. **Does the no-AR first run and AR1-corrected second run converge and retain
   full rank?** Yes.
3. **Is there a singularity/boundary signal?** Yes. After AR1 correction, the
   participant-day variance and effective df are effectively zero.
4. **Does AR1 resolve temporal dependence?** Partly. Lag-1 correlation falls
   from `r fmt_num(model_runs$lag1_residual_correlation[model_runs$stage == "First run: no AR1"], 3)`
   to `r fmt_num(model_summary$pearson_residual_lag1, 3)`, but is not eliminated.
5. **Are the diagnostics sufficiently acceptable?** Not for this exact full
   random-effects specification. The shifted-log Gaussian residuals are
   reasonably regular in their center, but the participant-day boundary
   collapse, remaining temporal dependence, and heavier-than-Gaussian tails
   prevent calling the model cleanly adequate.

The natural next model-construction question is whether the AR1-corrected
model should omit the now-unsupported participant-day intercept. That model was
**not** fitted here, because it would be a third model beyond the requested
two-run test balloon.

No additional distribution, time-of-day, placement, or sensitivity model was
fitted for this document.

## Provenance

The model was fitted with R `r as.character(getRversion())`, `mgcv`
`r as.character(utils::packageVersion("mgcv"))`, and `gratia`
`r as.character(utils::packageVersion("gratia"))`. The producing script is
`scripts/hypotheses/H03/run_h03_no_time_gamm_near_eye_pilot.R`; its artifact
manifest is
`artifacts/12_manifests/H03/H03_no_time_gamm_near_eye_pilot_manifest.csv`.
