---
title: "Import and align light recordings"
---

The analysis starts from the downloaded MeLiDos releases in `data/downloaded/recordings`. Each file is selected by site, measurement type, and the release commit in `config/site_sources.csv`. The computation below requires no download and does not use a stored analysis result.

```{r}
#| label: import-setup
#| echo: false
source("scripts/project.R")
analysis_setup()
for (helper in c("paths_io", "assertions", "time_axes", "time_support", "state_alignment", "aggregation_coverage", "import_sources", "build_import_alignment")) {
  source(file.path("scripts/pipeline", paste0(helper, ".R")))
}
root <- project_root()
paths <- pipeline_paths(root)
ensure_pipeline_directories(paths)
site_sources <- read_site_sources("config/site_sources.csv")
```

## Source recordings and diary intervals

The site time zone determines how recorded wall-clock labels are converted to actual UTC instants. Sleep-diary intervals run from attempted sleep preparation to waking; the three-hour pre-sleep interval is treated separately. Wear logs identify off-body periods. Keeping explicit, half-open intervals avoids filling gaps by carrying a state forward.

```{r}
#| label: import-state-intervals
state_pairs <- list()
source_inventory <- list()
for (site in site_sources$site) {
  site_row <- site_sources[site_sources$site == site, ]
  sleep_source <- recording_source(site_sources, site, "sleepdiaries")
  wear_source <- recording_source(site_sources, site, "wearlog")
  state_pairs[[site]] <- prepare_state_interval_pair(
    sleep_source$data, wear_source$data, site, site_row$timezone
  )
  for (kind in c("sleep", "wear")) {
    input <- if (kind == "sleep") sleep_source else wear_source
    intervals <- decorate_state_intervals(
      state_pairs[[site]][[kind]], site, site_row$timezone, kind,
      input$specification, input$downloaded_source
    )
    write_rds_artifact(intervals, file.path(paths$aligned, "state_intervals",
      paste0(site, "_", kind, "_intervals.rds")), "preparation/01-import-and-alignment.qmd")
  }
  source_inventory[[site]] <- tibble::tibble(
    site = site, repository = site_row$repository, release = site_row$commit,
    doi = site_row$doi, timezone = site_row$timezone,
    sleep_intervals = nrow(state_pairs[[site]]$sleep$intervals),
    wear_intervals = nrow(state_pairs[[site]]$wear$intervals)
  )
}
dplyr::bind_rows(source_inventory)
```

These interval counts describe the information available for aligning each recording. They do not select analysis days; coverage is evaluated in the next document.

## One-minute measurements and measurement context

Each native recording is aggregated to one minute only when every expected subepoch supplies a finite value. Both actual elapsed time and local wall-clock time are retained. The alignment joins are checked for duplicate keys and preserve the number of measurement rows. Values at or above 100,000 lx melEDI exceed the device operating boundary and are marked unusable.

```{r}
#| label: import-recordings
aligned_paths <- list(glasses = character(), chest = character())
recording_summary <- list()
for (site in site_sources$site) {
  site_row <- site_sources[site_sources$site == site, ]
  placements <- c("glasses", if (isTRUE(site_row$has_chest)) "chest")
  for (placement in placements) {
    input <- recording_source(site_sources, site, paste0("light_", placement))
    validate_light_stream(input$data, site, placement, site_row$timezone)
    annotated <- annotate_time_axes(input$data, timezone = site_row$timezone,
                                    datetime_col = "Datetime", site = site)
    annotated$position <- placement
    assert_unique_key(annotated, c("site", "Id", "position", "datetime_utc"))
    minute_data <- aggregate_native_epoch_to_minute(
      annotated, signal_cols = c("MEDI", "LIGHT"),
      id_cols = c("site", "Id", "position"),
      expected_subepochs = NULL, minimum_finite_subepochs = NULL,
      minimum_finite_fraction = 1, datetime_utc_col = "datetime_utc",
      datetime_wall_col = "datetime_wall", implicit_col = "is.implicit"
    )
    minute_data$source_repository <- input$specification$repository
    minute_data$source_commit <- input$specification$commit
    minute_data$source_doi <- input$specification$doi
    minute_data$source_sha256 <- input$downloaded_source$sha256
    write_rds_artifact(minute_data, file.path(paths$imported, "minute",
      paste0(site, "_", placement, "_minute.rds")), "preparation/01-import-and-alignment.qmd")
    minute_data$Datetime <- minute_data$datetime_utc
    sleep_joined <- attach_states_checked(minute_data, state_pairs[[site]]$sleep$intervals,
                                         stream_keys = "Id")
    wear_joined <- attach_states_checked(sleep_joined, state_pairs[[site]]$wear$intervals,
                                        stream_keys = "Id")
    aligned <- derive_measurement_context(dplyr::ungroup(wear_joined),
      placement = placement, brown_state_col = "State.Brown", wear_col = "wear",
      medi_col = "MEDI", light_col = "LIGHT", saturation_threshold = 100000)
    aligned <- dplyr::ungroup(aligned)
    validate_aligned_stream(aligned, 100000)
    stopifnot(nrow(aligned) == nrow(minute_data))
    path <- file.path(paths$aligned, "site_position", paste0(site, "_", placement, "_aligned.rds"))
    write_rds_artifact(aligned, path, "preparation/01-import-and-alignment.qmd")
    aligned_paths[[placement]] <- c(aligned_paths[[placement]], path)
    recording_summary[[paste(site, placement)]] <- tibble::tibble(
      site = site, placement = placement, participants = dplyr::n_distinct(aligned$Id),
      recorded_minutes = nrow(aligned), finite_melEDI = sum(is.finite(aligned$MEDI))
    )
    rm(input, annotated, minute_data, sleep_joined, wear_joined, aligned)
    invisible(gc())
  }
}
dplyr::bind_rows(recording_summary)
```

## Analysis inputs

The next document receives a chronologically ordered table for each sensor placement. Site and participant identifiers remain part of every measurement key.

```{r}
#| label: import-combined-outputs
#| code-fold: true
for (placement in names(aligned_paths)) {
  combined <- dplyr::bind_rows(lapply(aligned_paths[[placement]], readRDS)) |>
    dplyr::arrange(.data$site, .data$Id, .data$datetime_utc)
  assert_unique_key(combined, c("site", "Id", "position", "datetime_utc"))
  write_rds_artifact(combined, file.path(paths$aligned,
    paste0("light_", placement, "_aligned.rds")), "preparation/01-import-and-alignment.qmd")
  rm(combined)
  invisible(gc())
}
write_csv_artifact(dplyr::bind_rows(recording_summary),
  file.path(paths$diagnostics, "recording_summary.csv"), "preparation/01-import-and-alignment.qmd")
```
