---
title: "Reference profiles and temporal support"
---

Reference profiles describe when light normally occurs and support correction for missing periods. They are estimated before participant-day metrics. Profiles weight days equally within participants and then participants equally. Separate profiles describe full-day, waking, pre-sleep, and sleep-environment measurements for each sensor placement and light channel.

```{r}
#| label: setup
#| echo: false
source("scripts/project.R")
analysis_setup()
for (helper in c("paths_io", "assertions", "time_axes", "time_support", "state_alignment", "aggregation_coverage", "reference_profiles", "build_reference_profiles")) {
  source(file.path("scripts/pipeline", paste0(helper, ".R")))
}
root <- project_root()
paths <- pipeline_paths(root)
ensure_pipeline_directories(paths)
```
## Learn the profiles

Thirty-minute bins contain the profiles. The full-day pooled and leave-one-site-out profiles require at least 20 participants; state-specific and site-specific profiles require at least five. The same input grid also provides the probability of exceeding 250 lx melEDI in each time bin, used for timing-support calculations.

```{r}
#| label: learn-reference-profiles
profile_sets <- list()
distribution_sets <- list()
domains <- c("full_day", "wake", "pre-sleep", "sleep")
for (placement in c("glasses", "chest")) {
  coverage <- read_rds_artifact(file.path(paths$coverage,
    paste0("light_", placement, "_coverage.rds")), "data.frame")
  validate_reference_profile_input(coverage, placement)
  wall_planes <- stats::setNames(lapply(domains, function(state_domain) {
    collapse_reference_profile_wall_minutes(coverage, state_domain)
  }), domains)
  wall_minutes <- wall_planes$full_day
  placement_profiles <- list()
  for (signal in c("MEDI", "LIGHT")) {
    for (state_domain in domains) {
      training <- profile_training_rows(wall_planes[[state_domain]],
        signal = signal, state_domain = state_domain,
        skeleton_wall_minutes = wall_minutes)
      placement_profiles[[paste(signal, state_domain, sep = "/")]] <-
        learn_reference_profile_set(
          data = training, value_col = "profile_value", participant_col = "Id",
          participant_day_col = "local_date", site_col = "site",
          clock_bin_col = "clock_bin", strata = c("placement", "state_domain", "signal"),
          bin_minutes = 30L, variants = c("pooled", "site_specific", "leave_one_site_out"),
          pooled_full_day_min = 20L, leave_one_site_out_full_day_min = 20L,
          state_specific_min = 5L, site_specific_min = 5L)
    }
  }
  distribution_training <- timing_distribution_training_rows(wall_minutes)
  distribution_sets[[placement]] <- learn_exceedance_distribution_profile_set(
    data = distribution_training, value_col = "profile_value", participant_col = "Id",
    participant_day_col = "local_date", site_col = "site", clock_minute_col = "clock_minute",
    clock_bin_col = "clock_bin", strata = c("placement", "state_domain", "signal"),
    bin_minutes = 30L, threshold = 250,
    variants = c("pooled", "site_specific", "leave_one_site_out"),
    pooled_full_day_min = 20L, leave_one_site_out_full_day_min = 20L,
    state_specific_min = 5L, site_specific_min = 5L)
  profile_sets[[placement]] <- dplyr::bind_rows(placement_profiles)
  rm(coverage, wall_planes, wall_minutes, placement_profiles, distribution_training)
  invisible(gc())
}
profiles <- dplyr::bind_rows(profile_sets) |>
  dplyr::arrange(.data$profile_scope, .data$profile_variant, .data$placement,
    .data$state_domain, .data$signal, .data$clock_bin)
distribution_profiles <- dplyr::bind_rows(distribution_sets) |>
  dplyr::arrange(.data$profile_scope, .data$profile_variant, .data$placement,
    .data$state_domain, .data$signal, .data$clock_bin)
attr(profiles, "profile_learning") <- "fixed_once_before_participant_day_metrics"
attr(profiles, "profile_strata") <- c("placement", "state_domain", "signal")
attr(profiles, "bin_minutes") <- 30L
attr(distribution_profiles, "profile_learning") <- "fixed_once_before_participant_day_metrics"
attr(distribution_profiles, "profile_strata") <- c("placement", "state_domain", "signal")
attr(distribution_profiles, "bin_minutes") <- 30L
attr(distribution_profiles, "distribution_profile_type") <- "participant_balanced_valid_minute_exceedance_probability"
attr(distribution_profiles, "threshold_lx") <- 250
attr(distribution_profiles, "comparison") <- "strict_greater_than"
summarise_profile_support(profiles)
```
## Match temporal support to each metric

A relevance map identifies which observed portions of a day support a particular metric. Timing uses the exceedance probability; dose uses the light profile. Missing periods remain missing and are never filled with profile values.

```{r}
#| label: derive-relevance-maps
maps <- derive_metric_relevance_maps(profiles,
  distribution_profiles = distribution_profiles, signal_col = "signal",
  clock_bin_col = "clock_bin", strata = c("placement", "state_domain"), zero_offset = 0.1)
attr(maps, "profile_learning") <- "fixed_once_before_participant_day_metrics"
attr(maps, "bin_minutes") <- 30L
validate_reference_profile_outputs(profiles, distribution_profiles, maps,
  placements = c("glasses", "chest"), bin_minutes = 30L)
summarise_relevance_map_support(maps)
```

```{r}
#| label: save-reference-profiles
#| code-fold: true
write_result_pair(profiles, file.path(paths$profiles, "reference_profiles"))
write_result_pair(distribution_profiles, file.path(paths$profiles, "timing_exceedance_distributions"))
write_result_pair(maps, file.path(paths$profiles, "metric_relevance_maps"))
write_csv_artifact(summarise_profile_support(profiles), file.path(paths$profiles,
  "reference_profile_support.csv"), "preparation/03-reference-profiles.qmd")
```
