---
title: "Example participant days"
---

One reproducibly selected participant-day from each site illustrates the near-eye light recording, diary-defined sleep, and reported non-wear periods. Days must satisfy the coverage rules and contain one identifiable sleep-preparation transition and one waking transition. A fixed seed controls the selection. Site labels are joined from `config/site_metadata.csv`.

```{r}
#| label: setup
#| echo: false
source("scripts/project.R")
analysis_setup()
for (helper in c("paths_io", "assertions", "prepared_day_showcase")) {
  source(file.path("scripts/pipeline", paste0(helper, ".R")))
}
root <- project_root()
paths <- pipeline_paths(root)
ensure_pipeline_directories(paths)
```

```{r}
#| label: select-example-days
coverage <- readRDS(file.path(paths$coverage, "light_glasses_coverage.rds"))
context <- readRDS(file.path(paths$model_data, "context", "site_solar_context.rds"))
site_metadata <- utils::read.csv("config/site_metadata.csv", stringsAsFactors = FALSE,
  check.names = FALSE, fileEncoding = "UTF-8")
coverage <- dplyr::left_join(coverage,
  dplyr::select(site_metadata, site, location),
  by = "site", relationship = "many-to-one")
examples <- derive_prepared_day_showcase(coverage, context, site_metadata,
  seed = prepared_day_showcase_seed())
examples$selected_summary
```
The vertical extent of each light trace shows measured melEDI, while the overlaid intervals distinguish sleep and non-wear. Civil dawn and dusk provide site-specific astronomical context.

```{r}
#| label: fig-example-days
#| fig-cap: "Near-eye light exposure, diary-defined sleep and non-wear for one selected participant-day per site. Civil dawn and dusk mark the local solar context."
plot <- make_prepared_day_showcase_plot(examples$plot_data)
plot
```

```{r}
#| label: save-example-days
#| code-fold: true
write_csv_artifact(examples$selected_summary, file.path(paths$diagnostics,
  "prepared_day_showcase", "selected_days.csv"), "preparation/07-example-days.qmd")
write_csv_artifact(examples$eligible_counts, file.path(paths$diagnostics,
  "prepared_day_showcase", "eligible_day_counts.csv"), "preparation/07-example-days.qmd")
write_csv_artifact(examples$source_data, file.path(paths$source_data,
  "prepared_day_showcase.csv"), "preparation/07-example-days.qmd")
write_csv_artifact(examples$plot_data, file.path(paths$source_data,
  "prepared_day_showcase_plot.csv"), "preparation/07-example-days.qmd")
write_prepared_day_showcase_plot(plot, file.path(paths$figures, "prepared_day_showcase.png"),
  device = ragg::agg_png)
write_prepared_day_showcase_plot(plot, file.path(paths$figures, "prepared_day_showcase.svg"),
  device = svglite::svglite)
```
