---
title: "H06 daily metrics — Stage 1 audit and proposed analysis"
subtitle: "Preregistration-aligned participant-day estimand; no inferential fitting"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    df-print: paged
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

library(digest)
library(dplyr)
library(gt)
library(purrr)
library(tibble)
library(tidyr)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

read_project_csv <- function(path) {
  read.csv(
    file.path(project_root, path),
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
}

read_project_rds <- function(path) {
  readRDS(file.path(project_root, path))
}

sha256_file <- function(path) {
  digest::digest(
    file.path(project_root, path),
    algo = "sha256",
    file = TRUE
  )
}

compact_gt <- function(data, source_note = NULL, groupname_col = NULL) {
  output <- gt::gt(data, groupname_col = groupname_col) |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(12),
      table.width = gt::pct(100),
      data_row.padding = gt::px(3),
      heading.align = "left",
      source_notes.font.size = gt::px(10)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

finite_min <- function(x) {
  x <- x[is.finite(x)]
  if (length(x)) min(x) else NA_real_
}

finite_median <- function(x) {
  x <- x[is.finite(x)]
  if (length(x)) median(x) else NA_real_
}

finite_max <- function(x) {
  x <- x[is.finite(x)]
  if (length(x)) max(x) else NA_real_
}

site_registry <- read_project_csv("config/site_display_registry.csv") |>
  arrange(display_order)
site_levels <- site_registry$site

h05_metric_registry <- read_project_csv(
  "artifacts/06_model_data/H05/H05_metric_registry.csv"
)
metric_display <- read_project_csv("config/metric_display_registry.csv")

stopifnot(
  getRversion() >= numeric_version("4.6.1"),
  getRversion() < numeric_version("4.7.0"),
  identical(as.character(utils::packageVersion("melidosData")), "1.0.6")
)
```

# Stage 1 verdict

::: {.callout-warning title="H06-D-G1 — author approval is required before any fit"}
This document defines a scientifically separate **participant-day daily-metric
analysis**. Its proposed primary unit is one metric-specific participant-day,
not one supported hour. It gives each eligible day one row, does not weight a
day by its number of supported hours, and is not a sensitivity analysis of the
approved hourly H06 analysis.

The current data and contract checks pass for planning, subject to the
PREP-003/FIND-044 provenance qualification. The signed random-site/random-slope
formula is retained exactly, but with only nine near-eye sites (and eight chest
sites) it is not proposed as the primary inferential model. Fixed site with a
participant random intercept is the defensible primary hierarchy. No model,
bootstrap, omnibus test, contrast, confidence interval, p-value, or scientific
association estimate is fitted or calculated in this report.

Stage 2 is blocked until the author resolves the decisions listed at
**H06-D-G1** at the end of this report.
:::

# Registered scientific contract

The signed AsPredicted record states, exactly:

> **H6: Day type (weekday/weekend; free/work day), daily exercise, and sleep
> variables predict daily light exposure metrics.**

— `preregistration/AsPredicted #273407.pdf`, p. 1.

The registered model on p. 2 is:

```text
Metric ~ Measure + (Measure | Site) + (1 | Site:Participant)
```

The supporting analysis plan clarifies the recorded measures:

- day type: weekday/weekend and free/work day;
- sleep: sleep onset, final wake time, and duration;
- exercise: type/intensity, location, time in transit without cover, sitting
  time, and whether the device was worn while exercising.

The registration does not declare response families, links, transformations,
reference levels, a complete multiplicity partition, or a rule for choosing
between the two day-type variables. Those are genuine Stage 1 decisions; they
cannot be inferred from V0.

```{r}
#| label: contract-source-verification

contract_sources <- tibble::tribble(
  ~source, ~expected_sha256, ~role,
  "preregistration/AsPredicted #273407.pdf",
  "fce192c88676602b72013e4872a03b535a2f4ef59d34de3a9e1c47527742b7ad",
  "Controlling hypothesis wording and model",
  "preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx",
  "6e56766c18c476826c9c798799aa925752f11e5278f196797a583a13cc61ddf8",
  "Supporting outcome and predictor definitions"
) |>
  mutate(
    observed_sha256 = map_chr(source, sha256_file),
    verified = observed_sha256 == expected_sha256
  )

stopifnot(all(contract_sources$verified))

contract_sources |>
  select(source, role, observed_sha256, verified) |>
  compact_gt(
    source_note = "SHA-256 identities are evaluated at render; neither source is modified."
  ) |>
  gt::cols_label(
    source = "Signed source",
    role = "Contract role",
    observed_sha256 = "Observed SHA-256",
    verified = "Check"
  ) |>
  gt::fmt(columns = verified, fns = function(x) ifelse(x, "PASS", "FAIL"))
```

## Daily estimand versus V0 and the approved hourly H06 analysis

| Feature | Submitted V0 | Approved hourly H06 | This H06 daily variant |
|---|---|---|---|
| Response | Hourly geometric mean melEDI | Supported-hour zero-aware geometric mean melEDI | One of 15 supported participant-day metrics |
| Row | Participant-hour | Participant-hour | Metric-specific participant-day |
| Contribution of a day | More retained hours give more rows | Targets the supported-hour population | Exactly one row when that metric and predictor are admissible |
| Dependence target | Participant intercept only; no day/serial structure | Separate participant-cluster robust hourly analysis | Participant random intercept; actual-date daily dependence assessed |
| Primary scientific meaning | Association with repeated hourly melEDI | Average association over supported hours | Association with the distribution of a daily light-exposure metric |
| Relationship to this report | Historical implementation comparator | Separate approved estimand, untouched here | Standalone preregistration-aligned variant |

V0 repeated each day-level diary value over the available hourly outcome rows,
jointly modeled free/work day, four-level exercise intensity, and previous-night
sleep duration, and thereby weighted days by retained hours. The approved H06
analysis repaired that hourly question. Neither hourly estimate is numerically
interchangeable with a daily-metric estimate: the outcome, weighting,
dependence, multiplicity family, and claim all differ.

This daily variant therefore does not modify, replace, call, or depend on any
unfinished output in the main H06 analysis directories.

# Current input and provenance contract

```{r}
#| label: current-input-contract

input_contract <- tibble::tribble(
  ~role, ~relative_path, ~expected_sha256,
  "Near-eye participant-day metrics",
  "artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds",
  "fc80f8a2a94a447c5470db36c1ad1915868463501f04e9db3218f8156efc59d2",
  "Chest participant-day metrics",
  "artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds",
  "ce81c159fe018f35bec9bff44c9cd12f91fd4be1ea2696c56a8e91e628a9e2d3",
  "Normalized exercise diary",
  "artifacts/06_model_data/normalized_inputs/exercisediary.rds",
  "5bffe44cdd9c65f3d1dc20575d10b3f0a403577f3cf9109e83cfea95a2ae7107",
  "Normalized sleep diary",
  "artifacts/06_model_data/normalized_inputs/sleepdiaries.rds",
  "110819d74503c895170552c1f2214aca83de51cadd23bde7525aa821ee0b0e15",
  "Current metric-artifact manifest",
  "artifacts/12_manifests/metric_artifacts.csv",
  "6ec3185620d921e1f81b8464e850249613deb823988187d58794b3471303f4e8",
  "Current base-model manifest",
  "artifacts/12_manifests/base_model_data_artifacts.csv",
  "142d1044dd7e6a566e45e3a5c56033b1e8251e100c0d7c7523564c824950a0e4",
  "Gap-timing-unaware participant-day scenario",
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds",
  "28266064c6213eae6046ec6469d0dae39529f56e60933dcd581cf96c8590e1a4"
) |>
  mutate(
    observed_sha256 = map_chr(relative_path, sha256_file),
    verified = observed_sha256 == expected_sha256
  )

stopifnot(all(input_contract$verified))

input_contract |>
  select(role, relative_path, observed_sha256, verified) |>
  compact_gt(
    source_note = paste0(
      "Current preparation is used unchanged under PREP06-BASE-002. ",
      "The metric manifest retains the PREP-003/FIND-044 qualification."
    )
  ) |>
  gt::cols_label(
    role = "Role",
    relative_path = "Exact path",
    observed_sha256 = "SHA-256",
    verified = "Check"
  ) |>
  gt::fmt(columns = verified, fns = function(x) ifelse(x, "PASS", "FAIL"))
```

The approved PREP06-BASE-002 gate permits current downstream work. Its
qualification remains consequential: the current metric artifacts were not
all independently reconstructed under the exact current manifest/sample
identity, so PREP-003/FIND-044 remains open. This is a provenance limitation,
not evidence that the current values are wrong. Stage 2 must carry it into the
model and claim records.

The current construct is a supported **hybrid 24-hour measurement**: near-eye
measurements during wake and bedside sleep-environment measurements during
sleep. It must not be called continuous ocular exposure. The current primary
coverage rule intentionally uses the fixed 1,440-minute denominator under
DEV-055; the literal sleep-excluded preregistration denominator remains a
registered sensitivity elsewhere in the prepared metric system. Metrics are
read unchanged; preparation is not rebuilt here.

## Immutable diary checks

```{r}
#| label: diary-contract

exercise_raw <- read_project_rds(
  "artifacts/06_model_data/normalized_inputs/exercisediary.rds"
)
sleep_raw <- read_project_rds(
  "artifacts/06_model_data/normalized_inputs/sleepdiaries.rds"
)

clock_hour <- function(x) {
  as.numeric(format(x, "%H")) + as.numeric(format(x, "%M")) / 60
}

exercise_intensity_levels <- c(
  "No exercise", "Light", "Moderate", "Vigorous"
)
activity_status_levels <- c("Sedentary", "Active")

exercise_day <- exercise_raw |>
  transmute(
    site,
    Id,
    local_date = Date,
    exercise_intensity = recode(
      as.character(intensity),
      "None of the above, I did not perform any type of physical activity" =
        "No exercise",
      "Light (causing small to no increases in heart rate and breathing, e.g. taking a stroll in the park)" =
        "Light",
      "Moderate (causing moderate increases in heart rate and breathing, e.g. cycling in the city)" =
        "Moderate",
      "Vigorous (causing large increases in heart rate and breathing, e.g. running)" =
        "Vigorous"
    ) |>
      factor(levels = exercise_intensity_levels),
    activity_status = case_when(
      is.na(exercise_intensity) ~ NA_character_,
      exercise_intensity == "No exercise" ~ "Sedentary",
      TRUE ~ "Active"
    ) |>
      factor(levels = activity_status_levels),
    exercise_location = recode(
      as.character(location),
      "Outdoors (e.g. running, cycling in the city)" = "Outdoors",
      "Indoors (e.g. gym or home workout)" = "Indoors",
      "Both indoors and outdoors" = "Indoors and outdoors"
    ) |>
      factor(levels = c("Outdoors", "Indoors", "Indoors and outdoors")),
    active_commute_h = as.numeric(commute, units = "hours"),
    sedentary_source_minutes = as.numeric(sedentary, units = "mins"),
    sedentary_h_as_recorded = as.numeric(sedentary, units = "hours"),
    sedentary_h = if_else(
      site == "KNUST" &
        Id == "KNUST_S005" &
        local_date == as.Date("2024-11-04") &
        sedentary_source_minutes == 3600,
      1,
      sedentary_h_as_recorded
    ),
    sedentary_reinterpretation = if_else(
      site == "KNUST" &
        Id == "KNUST_S005" &
        local_date == as.Date("2024-11-04") &
        sedentary_source_minutes == 3600,
      "Author-approved: source numeral interpreted as 3,600 seconds (1 h)",
      NA_character_
    ),
    wore_light_logger_during_exercise = factor(
      light_glasses,
      levels = c(0, 1),
      labels = c("No", "Yes")
    ),
    exercise_source_row = source_row
  )

sleep_day <- sleep_raw |>
  transmute(
    site,
    Id,
    local_date = as.Date(wake_wall),
    work_free_day = factor(
      as.character(daytype2),
      levels = c("a work day", "a free day"),
      labels = c("Work day", "Free day")
    ),
    previous_sleep_duration_h = as.numeric(sleep_duration, units = "hours"),
    previous_sleep_duration_centered_h = previous_sleep_duration_h - 8,
    previous_sleep_onset_clock_h = clock_hour(sleep_onset_wall),
    previous_sleep_onset_after_noon_h =
      (previous_sleep_onset_clock_h - 12) %% 24,
    previous_sleep_onset_centered_h =
      previous_sleep_onset_after_noon_h - 11,
    wake_clock_h = clock_hour(wake_wall),
    wake_centered_h = wake_clock_h - 7,
    sleep_onset_date = as.Date(sleep_onset_wall),
    sleep_interval_analysis_eligible,
    sleep_interval_quarantined,
    sleep_source_row = source_row
  )

exercise_duplicate_keys <- exercise_day |>
  count(site, Id, local_date, name = "rows") |>
  filter(rows != 1L)
sleep_duplicate_keys <- sleep_day |>
  filter(!is.na(local_date)) |>
  count(site, Id, local_date, name = "rows") |>
  filter(rows != 1L)

stopifnot(
  nrow(exercise_duplicate_keys) == 0L,
  nrow(sleep_duplicate_keys) == 0L,
  sum(is.na(sleep_day$local_date)) == 1L,
  sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S001") == 7L,
  sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S101") == 0L,
  sum(!is.na(exercise_day$sedentary_reinterpretation)) == 1L
)

key_audit <- tibble::tribble(
  ~check, ~observed, ~required,
  "Exercise participant-day duplicate keys", nrow(exercise_duplicate_keys), "0",
  "Sleep wake-date participant-day duplicate keys", nrow(sleep_duplicate_keys), "0",
  "Quarantined sleep rows without a wake date", sum(is.na(sleep_day$local_date)), "1",
  "Current TUM S001 exercise rows", sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S001"), "7",
  "Current TUM S101 exercise rows", sum(exercise_raw$site == "TUM" & exercise_raw$Id == "TUM_S101"), "0",
  "H06-local 3,600-second interpretations", sum(!is.na(exercise_day$sedentary_reinterpretation)), "1"
)

key_audit |>
  compact_gt(
    source_note = paste0(
      "The KNUST_S005 2024-11-04 source numeral remains unchanged upstream. ",
      "Only the exploratory sedentary variable carries the row-level 3,600 s = 1 h interpretation."
    )
  ) |>
  gt::fmt_integer(columns = observed) |>
  gt::cols_label(check = "Audit check", observed = "Observed", required = "Contract")
```

Sleep is linked to the participant-day on which final wake occurred. The
source-calculated duration is retained in hours, not recomputed from timestamps.
Sleep onset is unwrapped at noon and centered at 23:00; wake is centered at
07:00. The exercise date is the recorded diary date. The immutable current
`melidosData` 1.0.6 source contains seven corrected TUM S001 exercise rows and
zero S101 rows, as required by DEV-056.

# Outcome registry and exact support

```{r}
#| label: metric-registry

metric_columns <- c(
  daily_geometric_mean_medi = "daily_geometric_mean_medi_lx",
  m10_mean_medi = "m10_mean_medi_lx",
  l10_mean_medi = "l10_mean_medi_lx",
  duration_above_1000 = "duration_above_1000_h",
  duration_above_250_wake = "duration_above_250_wake_h",
  duration_below_10_pre_sleep = "duration_below_10_pre_sleep_h",
  duration_below_1_sleep_environment = "duration_below_1_sleep_environment_h",
  longest_bout_above_250 = "longest_bout_above_250_h",
  m10_midpoint = "m10_midpoint_clock_minute",
  l10_midpoint = "l10_midpoint_clock_minute",
  mean_timing_above_250 = "mean_timing_above_250_clock_minute",
  first_timing_above_250 = "first_timing_above_250_clock_minute",
  last_timing_above_250 = "last_timing_above_250_clock_minute",
  dose_time_sensitive_corrected_medi = "dose_corrected_medi_lx_h",
  mder_ratio_of_integrals = "mder"
)

metric_contract <- h05_metric_registry |>
  filter(analysis_unit == "participant_day") |>
  mutate(
    source_column = unname(metric_columns[metric_id]),
    manuscript_name = if_else(
      metric_id == "longest_bout_above_250",
      "Longest continuous period above 250 lx melEDI",
      manuscript_name
    )
  ) |>
  arrange(metric_order)

stopifnot(
  nrow(metric_contract) == 15L,
  !anyNA(metric_contract$source_column),
  identical(metric_contract$metric_id, names(metric_columns)),
  all(metric_display$metric_id[
    metric_display$analytical_role == "primary_outcome" &
      metric_display$analysis_unit == "participant-day"
  ] %in% metric_contract$metric_id)
)

metric_contract |>
  select(
    manuscript_name, display_unit, response_family, response_transform,
    effect_scale, diagnostic_note
  ) |>
  compact_gt(
    source_note = paste0(
      "These are H05 continuity candidates only. H05 estimates, adequacy ",
      "decisions, and conclusions do not transfer to H06. IS and IV are ",
      "excluded because their analysis unit is participant, not participant-day."
    )
  ) |>
  gt::cols_label(
    manuscript_name = "Metric",
    display_unit = "Unit",
    response_family = "Candidate family/link",
    response_transform = "Candidate transformation",
    effect_scale = "Practical contrast scale",
    diagnostic_note = "Mandatory H06 check"
  )
```

The source column retains the historical internal identifier
`longest_bout_above_250_h`; all reader-facing text uses **period**, not “bout”.
melEDI is used throughout reader-facing material.

```{r}
#| label: metric-support-contract

metric_support_contract <- tibble::tribble(
  ~metric_id, ~support_and_admissibility,
  "daily_geometric_mean_medi",
  "Supported Rule-A hybrid day; current zero-aware daily geometric mean; non-negative.",
  "m10_mean_medi",
  "Brightest supported 600-minute window: 20 consecutive 30-minute bins; deterministic ties; non-negative.",
  "l10_mean_medi",
  "Darkest supported 600-minute window: midnight wrapping permitted; deterministic ties; non-negative.",
  "duration_above_1000",
  "Supported full-day hybrid intervals; true no-event duration is zero; physical range 0–24 h.",
  "duration_above_250_wake",
  "Supported waking-state domain; complete/adequate state support; duration cannot exceed supported waking domain.",
  "duration_below_10_pre_sleep",
  "Supported three-hour pre-sleep state window; reason-coded if absent/incomplete; six-hour audit flag retained.",
  "duration_below_1_sleep_environment",
  "Supported bedside sleep-environment domain; duration cannot exceed supported sleep domain.",
  "longest_bout_above_250",
  "Gaps and invalid observations terminate runs; primary is an observed lower bound; exact-identifiable-only sensitivity mandatory.",
  "m10_midpoint",
  "Midpoint of the supported M10 window; clock minute in [0, 1440).",
  "l10_midpoint",
  "Midpoint of the midnight-wrapping supported L10 window; clock minute in [0, 1440).",
  "mean_timing_above_250",
  "Circular, interval-duration-weighted mean above 250 lx melEDI; adequate relevance support; no event gives NA.",
  "first_timing_above_250",
  "First supported event above 250 lx melEDI; left boundary censoring is retained explicitly; no event gives NA.",
  "last_timing_above_250",
  "Last supported event above 250 lx melEDI; right boundary censoring is retained explicitly; no event gives NA.",
  "dose_time_sensitive_corrected_medi",
  "Time-sensitive fixed-profile melEDI dose; adequate relevance coverage; correction capped at 1.25; non-negative.",
  "mder_ratio_of_integrals",
  "Ratio of paired melEDI and illuminance integrals on the same intervals; ≥0.80 ordinary, melEDI-profile, and illuminance-profile support; positive illuminance integral; unscaled and unweighted."
) |>
  left_join(
    metric_contract |>
      select(metric_order, metric_id, manuscript_name, display_unit),
    by = "metric_id",
    relationship = "one-to-one"
  ) |>
  arrange(metric_order)

stopifnot(nrow(metric_support_contract) == 15L)

metric_support_contract |>
  select(manuscript_name, display_unit, support_and_admissibility) |>
  compact_gt(
    source_note = paste0(
      "Current prepared values are not recomputed. Exact-zero days already ",
      "follow the approved all-zero screen. Metric-specific failure reasons ",
      "remain missing, never recoded as zero."
    )
  ) |>
  gt::cols_label(
    manuscript_name = "Metric",
    display_unit = "Unit",
    support_and_admissibility = "Current support/admissibility contract"
  )
```

# Predictor map and proposed inferential roles

```{r}
#| label: predictor-contract

predictor_contract <- tibble::tribble(
  ~predictor_order, ~predictor_id, ~reader_name, ~source_field, ~type,
  ~registered_scope, ~proposed_role, ~reference_or_increment,
  1L, "work_free_day", "Work day versus free day", "sleepdiaries$daytype2",
  "Categorical", "Day type: free/work day", "Core primary candidate",
  "Work day reference; free day contrast",
  2L, "weekday_weekend", "Weekday versus weekend", "derived from metric local_date",
  "Categorical", "Day type: weekday/weekend", "Registered named sensitivity",
  "Weekday reference; weekend contrast",
  3L, "activity_status", "Daily activity status", "exercisediary$intensity",
  "Categorical", "Daily exercise", "Core primary candidate",
  "Sedentary reference; Active contrast",
  4L, "exercise_intensity", "Daily exercise intensity", "exercisediary$intensity",
  "Categorical", "Exercise type/intensity", "Registered exploratory detail",
  "No exercise reference; light/moderate/vigorous",
  5L, "exercise_location", "Exercise location", "exercisediary$location",
  "Categorical", "Exercise location", "Registered exploratory; exercise days",
  "Outdoors reference; conditional on recorded exercise/location",
  6L, "active_commute_h", "Walking/cycling travel time", "exercisediary$commute",
  "Continuous", "Time in transit without cover", "Registered exploratory; MPI unavailable",
  "Per 1 h; zero retained; no extrapolation beyond observed support",
  7L, "sedentary_h", "Sitting/reclining time", "exercisediary$sedentary",
  "Continuous", "Sitting time", "Registered exploratory with row provenance",
  "Per 1 h; one H06-local 3,600 s = 1 h interpretation",
  8L, "wore_light_logger_during_exercise", "Logger worn during exercise",
  "exercisediary$light_glasses", "Categorical", "Logger worn while exercising",
  "Measurement-support sensitivity", "No reference; Yes contrast; not an exposure determinant",
  9L, "previous_sleep_onset_centered_h", "Previous sleep onset", "sleepdiaries$sleep",
  "Continuous", "Sleep onset", "Registered exploratory",
  "Per 1 h on noon-unwrapped clock, centered at 23:00",
  10L, "wake_centered_h", "Final wake time", "sleepdiaries$wake",
  "Continuous", "Wake time", "Registered exploratory",
  "Per 1 h local clock, centered at 07:00",
  11L, "previous_sleep_duration_centered_h", "Previous-night sleep duration",
  "sleepdiaries$sleep_duration", "Continuous", "Sleep duration",
  "Core primary candidate", "Per 1 h, centered at 8 h"
)

predictor_contract |>
  select(
    reader_name, source_field, registered_scope, proposed_role,
    reference_or_increment
  ) |>
  compact_gt(
    source_note = paste0(
      "The current analytical RDS has no separate free-text exercise-type ",
      "field. The four structured intensity responses are the recoverable ",
      "type/intensity measure. Temperature and unsupported predictors are excluded."
    )
  ) |>
  gt::cols_label(
    reader_name = "Predictor",
    source_field = "Current field",
    registered_scope = "Registered construct",
    proposed_role = "H06-D-G1 proposal",
    reference_or_increment = "Coding / practical increment"
  )
```

The recommended primary set is the provisional three-predictor set:
work/free day, binary Sedentary/Active status, and previous-night sleep
duration. This recommendation is not treated as signed fact: weekday/weekend,
sleep onset, wake, and the additional recorded exercise fields are also in the
registered scope. H06-D-G1 must decide their roles before fitting. Each
predictor is proposed as a separate `Measure` analysis, matching the signed
formula; a jointly adjusted three-predictor model would be a distinct
exploratory estimand and is not the primary contract.

# Candidate participant-day frames — no fitting

## Exact joins and metric reconciliation

```{r}
#| label: candidate-frame-construction

near_day <- read_project_rds(
  "artifacts/06_model_data/base/metrics_glasses_participant_day_enriched.rds"
)
chest_day <- read_project_rds(
  "artifacts/06_model_data/base/metrics_chest_participant_day_enriched.rds"
)

make_daily_candidate <- function(metric_data, placement) {
  metric_data |>
    select(site, Id, local_date, all_of(unname(metric_columns))) |>
    left_join(
      sleep_day |>
        select(-sleep_onset_date),
      by = c("site", "Id", "local_date"),
      relationship = "many-to-one"
    ) |>
    left_join(
      exercise_day,
      by = c("site", "Id", "local_date"),
      relationship = "many-to-one"
    ) |>
    mutate(
      placement = placement,
      participant_key = paste(site, Id, sep = "::"),
      participant_day_key = paste(site, Id, local_date, sep = "::"),
      weekday_weekend = factor(
        if_else(
          as.POSIXlt(local_date)$wday %in% c(0L, 6L),
          "Weekend",
          "Weekday"
        ),
        levels = c("Weekday", "Weekend")
      )
    )
}

daily_candidates <- bind_rows(
  make_daily_candidate(near_day, "Near-eye"),
  make_daily_candidate(chest_day, "Chest")
) |>
  mutate(
    placement = factor(placement, levels = c("Near-eye", "Chest")),
    site = factor(site, levels = site_levels)
  )

base_candidate_sample <- daily_candidates |>
  summarise(
    `Participant-days` = n(),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    .by = placement
  )

stopifnot(
  base_candidate_sample$`Participant-days`[
    base_candidate_sample$placement == "Near-eye"
  ] == 816L,
  base_candidate_sample$`Participant-days`[
    base_candidate_sample$placement == "Chest"
  ] == 902L
)

base_candidate_sample |>
  compact_gt(
    source_note = paste0(
      "Available current participant-day inputs before metric- or predictor-specific ",
      "admissibility. These are not fitted samples."
    )
  ) |>
  gt::fmt_integer(columns = c(`Participant-days`, Participants, Sites))
```

The near-eye frame contains all nine submitted sites. The chest frame contains
eight because Izmir has no chest placement. Participant identifiers are used
only to construct nested keys and counts; they are not displayed.

```{r}
#| label: metric-artifact-reconciliation

metric_audit_long <- bind_rows(
  read_project_csv("artifacts/05_metrics/metrics_glasses_values_long.csv") |>
    mutate(placement = "Near-eye"),
  read_project_csv("artifacts/05_metrics/metrics_chest_values_long.csv") |>
    mutate(placement = "Chest")
) |>
  filter(
    analysis_unit == "participant_day",
    profile_variant == "pooled",
    metric %in% names(metric_columns)
  ) |>
  mutate(
    local_date = as.Date(local_date),
    placement = factor(placement, levels = c("Near-eye", "Chest"))
  ) |>
  rename(metric_id = metric, audited_value = value)

daily_metric_long <- daily_candidates |>
  pivot_longer(
    cols = all_of(unname(metric_columns)),
    names_to = "source_column",
    values_to = "response_value"
  ) |>
  mutate(
    metric_id = names(metric_columns)[match(source_column, metric_columns)]
  ) |>
  left_join(
    metric_contract,
    by = "metric_id",
    relationship = "many-to-one",
    suffix = c("", "_registry")
  ) |>
  left_join(
    metric_audit_long |>
      select(
        placement, site, Id, local_date, metric_id, audited_value, units,
        state_domain, estimable, failure_reason, ordinary_support,
        relevance_support, valid_minutes, expected_minutes,
        left_censored, right_censored, any_censored,
        state_domain_complete, medi_profile_support, light_profile_support,
        minimum_support, support_threshold_enforced,
        estimate_interpretation, missing_invalid_breaks_runs,
        exact_identifiable
      ),
    by = c("placement", "site", "Id", "local_date", "metric_id"),
    relationship = "one-to-one"
  )

stopifnot(
  nrow(metric_audit_long) == 15L * (816L + 902L),
  nrow(daily_metric_long) == nrow(metric_audit_long),
  all(is.na(daily_metric_long$audited_value) ==
    is.na(daily_metric_long$response_value)),
  all(
    is.na(daily_metric_long$audited_value) |
      abs(daily_metric_long$audited_value -
        daily_metric_long$response_value) < .Machine$double.eps^0.5
  ),
  all(is.na(daily_metric_long$response_value) == !daily_metric_long$estimable)
)

reconciliation_audit <- tibble::tribble(
  ~check, ~observed, ~required,
  "Selected participant-day metric rows", nrow(daily_metric_long),
  as.character(15L * (816L + 902L)),
  "Selected metrics", n_distinct(daily_metric_long$metric_id), "15",
  "Duplicate metric × placement × participant-day keys",
  daily_metric_long |>
    count(placement, site, Id, local_date, metric_id) |>
    filter(n != 1L) |>
    nrow(),
  "0",
  "Wide/long missingness mismatches",
  sum(is.na(daily_metric_long$audited_value) !=
    is.na(daily_metric_long$response_value)),
  "0",
  "Wide/long finite-value mismatches",
  sum(
    is.finite(daily_metric_long$audited_value) &
      abs(daily_metric_long$audited_value -
        daily_metric_long$response_value) >= .Machine$double.eps^0.5
  ),
  "0"
)

reconciliation_audit |>
  compact_gt(
    source_note = paste0(
      "All 25,770 selected values are reconciled to the current metric ",
      "long-form support/admissibility artifact; no metric is recalculated."
    )
  ) |>
  gt::fmt_integer(columns = observed) |>
  gt::cols_label(check = "Audit check", observed = "Observed", required = "Contract")
```

## Metric-specific admissibility and distribution screen

```{r}
#| label: metric-support-audit

metric_support_audit <- daily_metric_long |>
  summarise(
    input_days = n(),
    estimable_days = sum(estimable),
    nonestimable_days = sum(!estimable),
    zero_values = sum(response_value == 0, na.rm = TRUE),
    first_left_censored = sum(left_censored %in% TRUE, na.rm = TRUE),
    last_right_censored = sum(right_censored %in% TRUE, na.rm = TRUE),
    period_censored = sum(any_censored %in% TRUE, na.rm = TRUE),
    period_exact_identifiable = sum(exact_identifiable %in% TRUE, na.rm = TRUE),
    valid_minutes_min = finite_min(valid_minutes[estimable]),
    valid_minutes_median = finite_median(valid_minutes[estimable]),
    valid_minutes_max = finite_max(valid_minutes[estimable]),
    expected_minutes_min = finite_min(expected_minutes[estimable]),
    expected_minutes_max = finite_max(expected_minutes[estimable]),
    units_in_artifact = paste(sort(unique(na.omit(units))), collapse = "; "),
    .by = c(
      placement, metric_order, metric_id, manuscript_name,
      display_unit, response_family, response_transform
    )
  ) |>
  arrange(metric_order, placement)

metric_support_audit |>
  select(
    placement, manuscript_name, input_days, estimable_days,
    nonestimable_days, zero_values, first_left_censored,
    last_right_censored, period_censored, period_exact_identifiable,
    units_in_artifact
  ) |>
  compact_gt(
    groupname_col = "placement",
    source_note = paste0(
      "Directional timing censor flags are audited separately. The generic ",
      "any_censored field is intentionally used by current code for the ",
      "continuous-period exact-identifiable sensitivity, not as their union."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::cols_label(
    manuscript_name = "Metric",
    input_days = "Input days",
    estimable_days = "Admissible",
    nonestimable_days = "Unavailable",
    zero_values = "Exact zeros",
    first_left_censored = "Left-censored",
    last_right_censored = "Right-censored",
    period_censored = "Period censored",
    period_exact_identifiable = "Period exact",
    units_in_artifact = "Artifact unit"
  )
```

The exact zeros in L10 mean, threshold durations, and the longest continuous
period are valid observations when support is adequate; they are not missing
values. A timing value at clock minute zero is midnight, not “no event”.
No-event timing rows are reason-coded `NA`.

```{r}
#| label: nonestimability-reasons

metric_failure_reasons <- daily_metric_long |>
  filter(!estimable) |>
  mutate(failure_reason = replace_na(failure_reason, "[Missing reason]")) |>
  count(
    placement, metric_order, metric_id, manuscript_name, failure_reason,
    name = "participant_days"
  ) |>
  arrange(metric_order, placement, desc(participant_days))

metric_failure_reasons |>
  select(placement, manuscript_name, failure_reason, participant_days) |>
  compact_gt(
    groupname_col = "placement",
    source_note = paste0(
      "All metric unavailability is reason-coded. Stage 2 must preserve these ",
      "rows as unavailable and report each exact fitted sample."
    )
  ) |>
  gt::fmt_integer(columns = participant_days) |>
  gt::cols_label(
    manuscript_name = "Metric",
    failure_reason = "Current reason code",
    participant_days = "Participant-days"
  )
```

## Every metric × predictor × placement candidate sample

```{r}
#| label: candidate-sample-audit

predictor_columns <- setNames(
  predictor_contract$predictor_id,
  predictor_contract$predictor_id
)

candidate_missingness <- imap_dfr(
  predictor_columns,
  function(column, predictor_id) {
    daily_metric_long |>
      mutate(
        response_available = is.finite(response_value),
        predictor_available = !is.na(.data[[column]]) &
          if (is.numeric(.data[[column]])) {
            is.finite(.data[[column]])
          } else {
            TRUE
          },
        status = case_when(
          response_available & predictor_available ~ "Complete candidate",
          !response_available & predictor_available ~ "Metric unavailable only",
          response_available & !predictor_available ~ "Predictor unavailable only",
          TRUE ~ "Both unavailable"
        )
      ) |>
      summarise(
        predictor_id = predictor_id,
        base_days = n(),
        complete_days = sum(status == "Complete candidate"),
        metric_unavailable_only = sum(status == "Metric unavailable only"),
        predictor_unavailable_only = sum(status == "Predictor unavailable only"),
        both_unavailable = sum(status == "Both unavailable"),
        participants = n_distinct(
          participant_key[status == "Complete candidate"]
        ),
        sites = n_distinct(site[status == "Complete candidate"]),
        .by = c(placement, metric_order, metric_id, manuscript_name)
      )
  }
) |>
  left_join(
    predictor_contract,
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, metric_order, placement)

candidate_available_rows <- imap_dfr(
  predictor_columns,
  function(column, predictor_id) {
    daily_metric_long |>
      filter(is.finite(response_value), !is.na(.data[[column]])) |>
      transmute(
        placement, site, Id, local_date, participant_key,
        participant_day_key, metric_order, metric_id, manuscript_name,
        predictor_id, predictor_value = as.character(.data[[column]])
      )
  }
)

paired_samples <- candidate_available_rows |>
  distinct(
    placement, site, Id, local_date, participant_key,
    metric_order, metric_id, manuscript_name, predictor_id
  ) |>
  count(
    site, Id, local_date, participant_key, metric_order, metric_id,
    manuscript_name, predictor_id,
    name = "placements"
  ) |>
  filter(placements == 2L) |>
  summarise(
    paired_days = n(),
    paired_participants = n_distinct(participant_key),
    paired_sites = n_distinct(site),
    .by = c(predictor_id, metric_order, metric_id, manuscript_name)
  )

candidate_samples <- candidate_missingness |>
  left_join(
    paired_samples,
    by = c("predictor_id", "metric_order", "metric_id", "manuscript_name"),
    relationship = "many-to-one"
  ) |>
  mutate(
    across(
      c(paired_days, paired_participants, paired_sites),
      ~ replace_na(.x, 0L)
    )
  )

stopifnot(
  nrow(candidate_samples) == 11L * 15L * 2L,
  all(with(
    candidate_samples,
    complete_days + metric_unavailable_only +
      predictor_unavailable_only + both_unavailable == base_days
  ))
)

predictor_coverage_summary <- candidate_samples |>
  summarise(
    `Candidate days, minimum` = min(complete_days),
    `Candidate days, maximum` = max(complete_days),
    `Participants, minimum` = min(participants),
    `Participants, maximum` = max(participants),
    `Sites, minimum` = min(sites),
    `Sites, maximum` = max(sites),
    `Paired days, minimum` = min(paired_days),
    `Paired days, maximum` = max(paired_days),
    .by = c(predictor_order, reader_name, proposed_role, placement)
  ) |>
  arrange(predictor_order, placement)

predictor_coverage_summary |>
  select(-predictor_order) |>
  compact_gt(
    groupname_col = "proposed_role",
    source_note = paste0(
      "Ranges are across the 15 metric-specific frames. Paired means the same ",
      "participant-day has an admissible value at both placements and the predictor."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::cols_label(reader_name = "Predictor", placement = "Placement")
```

The complete exact audit table below has 330 rows: 15 metrics × 11 predictors
× 2 placements. Missingness categories are disjoint and sum to the placement
base on every row. The paired/common counts are computed from matching
site–participant–date keys; they are not the minimum of two marginal samples.

```{r}
#| label: exact-candidate-sample-table

candidate_samples |>
  select(
    `Predictor` = reader_name,
    `Role` = proposed_role,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Base days` = base_days,
    `Complete candidate days` = complete_days,
    `Metric unavailable only` = metric_unavailable_only,
    `Predictor unavailable only` = predictor_unavailable_only,
    `Both unavailable` = both_unavailable,
    `Participants` = participants,
    `Sites` = sites,
    `Paired/common days` = paired_days,
    `Paired/common participants` = paired_participants,
    `Paired/common sites` = paired_sites
  )
```

## Exact category cells and continuous-predictor support

```{r}
#| label: category-cell-audit

category_levels <- list(
  work_free_day = c("Work day", "Free day"),
  weekday_weekend = c("Weekday", "Weekend"),
  activity_status = c("Sedentary", "Active"),
  exercise_intensity = exercise_intensity_levels,
  exercise_location = c("Outdoors", "Indoors", "Indoors and outdoors"),
  wore_light_logger_during_exercise = c("No", "Yes")
)

make_category_cells <- function(column, predictor_id, defined_levels) {
  observed <- daily_metric_long |>
    filter(is.finite(response_value)) |>
    mutate(
      category = as.character(.data[[column]]),
      category = replace_na(category, "[Missing predictor]")
    ) |>
    summarise(
      participant_days = n(),
      participants = n_distinct(participant_key),
      sites = n_distinct(site),
      .by = c(
        placement, metric_order, metric_id, manuscript_name, category
      )
    )

  tidyr::expand_grid(
    placement = factor(
      c("Near-eye", "Chest"),
      levels = c("Near-eye", "Chest")
    ),
    metric_contract |>
      select(metric_order, metric_id, manuscript_name),
    category = c(defined_levels, "[Missing predictor]")
  ) |>
    left_join(
      observed,
      by = c(
        "placement", "metric_order", "metric_id", "manuscript_name",
        "category"
      ),
      relationship = "one-to-one"
    ) |>
    mutate(
      predictor_id = predictor_id,
      across(
        c(participant_days, participants, sites),
        ~ replace_na(.x, 0L)
      )
    )
}

category_cells <- imap_dfr(
  category_levels,
  function(defined_levels, predictor_id) {
    make_category_cells(predictor_id, predictor_id, defined_levels)
  }
) |>
  left_join(
    predictor_contract,
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, metric_order, placement, category)

stopifnot(nrow(category_cells) == 630L)

category_cells |>
  select(
    `Predictor` = reader_name,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Category` = category,
    `Participant-days` = participant_days,
    `Participants` = participants,
    `Sites` = sites
  )
```

The table includes explicit zero cells and missing-predictor cells for every
categorical metric × predictor × placement combination. Exercise location and
logger-worn status are conditionally applicable fields and therefore do not
target the same population as binary activity status.

```{r}
#| label: continuous-predictor-audit

continuous_predictors <- predictor_contract |>
  filter(type == "Continuous") |>
  pull(predictor_id)

continuous_distributions <- map_dfr(
  continuous_predictors,
  function(predictor_id) {
    daily_metric_long |>
      filter(
        is.finite(response_value),
        is.finite(.data[[predictor_id]])
      ) |>
      summarise(
        predictor_id = predictor_id,
        participant_days = n(),
        participants = n_distinct(participant_key),
        sites = n_distinct(site),
        minimum = min(.data[[predictor_id]]),
        q1 = quantile(.data[[predictor_id]], 0.25, names = FALSE),
        median = median(.data[[predictor_id]]),
        q3 = quantile(.data[[predictor_id]], 0.75, names = FALSE),
        maximum = max(.data[[predictor_id]]),
        .by = c(placement, metric_order, metric_id, manuscript_name)
      )
  }
) |>
  left_join(
    predictor_contract,
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, metric_order, placement)

continuous_distributions |>
  select(
    `Predictor` = reader_name,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Participant-days` = participant_days,
    `Participants` = participants,
    `Sites` = sites,
    `Minimum` = minimum,
    `Q1` = q1,
    `Median` = median,
    `Q3` = q3,
    `Maximum` = maximum
  )
```

No continuous variable will be winsorized or trimmed after seeing results.
The recorded 700-minute active-travel value and other tails require influence
diagnostics. The H06-local sedentary reinterpretation is used only if that
exploratory predictor is approved.

# Exact proposed model hierarchy

The response is transformed according to the metric registry before formula
evaluation. Each metric × predictor × placement comparison uses one frozen,
ordered complete-case frame for its reduced, additive, and heterogeneity
models. A comparison never gains rows merely because one term was removed.

```{r}
#| label: exact-daily-formula-contract
#| code-fold: false
#| code-summary: "Exact H06 daily formulas (always shown)"

formula_text <- function(x) paste(deparse(x), collapse = " ")

make_daily_formula_set <- function(measure) {
  list(
    registered_exact = stats::as.formula(paste0(
      "response_value ~ ", measure,
      " + (", measure, " | site) + (1 | site:Id)"
    )),
    fixed_site_reduced = stats::as.formula(
      "response_value ~ site + (1 | participant_key)"
    ),
    fixed_site_additive = stats::as.formula(paste0(
      "response_value ~ site + ", measure,
      " + (1 | participant_key)"
    )),
    fixed_site_heterogeneity = stats::as.formula(paste0(
      "response_value ~ site * ", measure,
      " + (1 | participant_key)"
    )),
    daily_ar1_reduced = stats::as.formula(paste0(
      "response_value ~ site + (1 | participant_key) + ",
      "ar1(day_index_factor + 0 | day_sequence_id)"
    )),
    daily_ar1_additive = stats::as.formula(paste0(
      "response_value ~ site + ", measure,
      " + (1 | participant_key) + ",
      "ar1(day_index_factor + 0 | day_sequence_id)"
    )),
    daily_ar1_heterogeneity = stats::as.formula(paste0(
      "response_value ~ site * ", measure,
      " + (1 | participant_key) + ",
      "ar1(day_index_factor + 0 | day_sequence_id)"
    ))
  )
}

daily_formula_registry <- imap_dfr(
  predictor_columns,
  function(measure, predictor_id) {
    enframe(
      make_daily_formula_set(measure),
      name = "model_id",
      value = "formula"
    ) |>
      mutate(
        predictor_id = predictor_id,
        formula = map_chr(formula, formula_text),
        .before = 1
      )
  }
) |>
  left_join(
    predictor_contract |>
      select(predictor_order, predictor_id, reader_name, proposed_role),
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, match(model_id, names(make_daily_formula_set("x"))))

daily_formula_registry |>
  select(
    `Predictor` = reader_name,
    `Proposed role` = proposed_role,
    `Model` = model_id,
    `Evaluated Wilkinson formula` = formula
  ) |>
  compact_gt(
    groupname_col = "Predictor",
    source_note = paste0(
      "registered_exact reproduces the signed random-site/random-slope form. ",
      "The primary association comparison is fixed_site_reduced versus ",
      "fixed_site_additive; site heterogeneity compares fixed_site_additive ",
      "with fixed_site_heterogeneity. AR(1) rows are conditional counterparts."
    )
  )
```

## Use of the hierarchy

1. **Exact registered formula.** Fit only as a preregistered benchmark when
   its design matrix is full rank and the engine returns a converged,
   nonsingular solution. It is never silently replaced or described as the
   primary model.
2. **Fixed-site reduced model.** Same exact frame, site fixed effects, and
   participant random intercept; no predictor.
3. **Fixed-site additive model.** Proposed primary association model. Its
   comparison with the reduced model supplies one metric-specific omnibus
   association test.
4. **Fixed-site predictor-by-site model.** Separately tests whether the
   association differs across submitted sites. It does not replace the
   equal-site average association.
5. **Actual-date, gap-aware daily AR(1) counterpart.** Uses `glmmTMB`-style
   `ar1()` terms only when consecutive-day support is adequate and residual
   temporal dependence triggers it. Missing calendar dates begin a new
   `day_sequence_id`; dates are never compressed into artificial adjacency.

For Gaussian comparisons, fixed-effect likelihood-ratio comparisons use ML;
the selected final Gaussian model is refitted with REML for estimation. For
Tweedie-log models, comparisons use the same likelihood family and exact
frame. Family changes are not chosen by significance.

## Random-site estimability before fitting

```{r}
#| label: random-site-design-audit

make_site_category_support <- function(column, predictor_id, defined_levels) {
  complete_rows <- daily_metric_long |>
    filter(is.finite(response_value), !is.na(.data[[column]])) |>
    transmute(
      placement, site, metric_order, metric_id, manuscript_name,
      category = as.character(.data[[column]])
    )

  site_grid <- complete_rows |>
    distinct(placement, site, metric_order, metric_id, manuscript_name) |>
    tidyr::crossing(category = defined_levels)

  site_cells <- complete_rows |>
    count(
      placement, site, metric_order, metric_id, manuscript_name, category,
      name = "days"
    )

  site_grid |>
    left_join(
      site_cells,
      by = c(
        "placement", "site", "metric_order", "metric_id",
        "manuscript_name", "category"
      ),
      relationship = "one-to-one"
    ) |>
    mutate(days = replace_na(days, 0L)) |>
    summarise(
      predictor_id = predictor_id,
      sites_in_frame = n_distinct(site),
      sites_with_required_variation = sum(tapply(days > 0L, site, all)),
      structural_zero_cells = sum(days == 0L),
      minimum_site_rows_or_cell = ifelse(
        any(days > 0L), min(days[days > 0L]), NA_integer_
      ),
      .by = c(placement, metric_order, metric_id, manuscript_name)
    )
}

categorical_site_support <- imap_dfr(
  category_levels,
  function(defined_levels, predictor_id) {
    make_site_category_support(
      predictor_id,
      predictor_id,
      defined_levels
    )
  }
)

continuous_site_support <- map_dfr(
  continuous_predictors,
  function(predictor_id) {
    daily_metric_long |>
      filter(
        is.finite(response_value),
        is.finite(.data[[predictor_id]])
      ) |>
      summarise(
        site_rows = n(),
        distinct_values = n_distinct(.data[[predictor_id]]),
        within_site_sd = ifelse(
          n() > 1L,
          stats::sd(.data[[predictor_id]]),
          NA_real_
        ),
        .by = c(
          placement, metric_order, metric_id, manuscript_name, site
        )
      ) |>
      summarise(
        predictor_id = predictor_id,
        sites_in_frame = n_distinct(site),
        sites_with_required_variation = sum(
          distinct_values >= 2L &
            is.finite(within_site_sd) & within_site_sd > 0
        ),
        minimum_site_rows_or_cell = min(site_rows),
        structural_zero_cells = NA_integer_,
        .by = c(placement, metric_order, metric_id, manuscript_name)
      )
  }
)

random_site_design <- bind_rows(
  categorical_site_support,
  continuous_site_support
) |>
  left_join(
    predictor_contract |>
      mutate(
        random_coefficient_dimension = if_else(
          type == "Categorical",
          map_int(predictor_id, ~ length(category_levels[[.x]])),
          2L
        ),
        random_covariance_parameters =
          random_coefficient_dimension *
          (random_coefficient_dimension + 1L) / 2L
      ) |>
      select(
        predictor_order, predictor_id, reader_name, type,
        random_coefficient_dimension, random_covariance_parameters
      ),
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, metric_order, placement)

random_site_summary <- random_site_design |>
  summarise(
    `Site counts across metrics` = paste(
      sort(unique(sites_in_frame)), collapse = ", "
    ),
    `Minimum sites with required variation` =
      min(sites_with_required_variation),
    `Random coefficients` = first(random_coefficient_dimension),
    `Random covariance parameters` = first(random_covariance_parameters),
    .by = c(predictor_order, reader_name, type, placement)
  ) |>
  arrange(predictor_order, placement)

random_site_summary |>
  select(-predictor_order) |>
  compact_gt(
    source_note = paste0(
      "A binary or continuous random slope requires three site-level ",
      "variance/covariance parameters; four-level intensity requires ten. ",
      "This count precedes convergence and singularity diagnostics."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::cols_label(reader_name = "Predictor", placement = "Placement")
```

The exact registered specification is **not reliably estimable as the primary
random-effects distribution with only nine near-eye site clusters** (and only
eight chest clusters). A numerical fit is not the same as defensible
estimation. Even binary/continuous measures ask nine clusters to identify
three site-level covariance parameters; four-level exercise intensity asks
for ten. Some exploratory fields also lose within-site variation or contain
structural site-by-category zeros on particular metric frames.

The exact model will therefore be a labelled registered benchmark only. A
converged, nonsingular benchmark may be reported descriptively beside the
fixed-site result. A singular, rank-deficient, non-positive-Hessian, or failed
fit is recorded as `NON_ESTIMABLE`; it is not simplified post hoc and is not
used for a claim. Fixed site is the proposed primary structure because the
nine submitted sites are the scientific target and equal-site contrasts can
be defined directly.

The complete metric-specific design audit follows.

```{r}
#| label: exact-random-site-design-table

random_site_design |>
  select(
    `Predictor` = reader_name,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Sites in frame` = sites_in_frame,
    `Sites with required variation` = sites_with_required_variation,
    `Structural zero site-level cells` = structural_zero_cells,
    `Minimum site rows / positive cell` = minimum_site_rows_or_cell,
    `Random coefficients` = random_coefficient_dimension,
    `Random covariance parameters` = random_covariance_parameters
  )
```

## Primary estimands, contrasts, and marginalization

For each metric and predictor, the proposed main estimand is the
**equal-site average adjusted association** in the fixed-site additive model,
conditional on a participant random intercept and averaged with equal weight
over the submitted site levels in that fitted frame. This gives each site
equal interpretive weight; it is not a participant-weighted or day-weighted
site average. Site fixed effects use sum-to-zero contrasts in the model frame;
reader-facing marginal means and contrasts use explicit equal site weights.

- Work/free: free day minus or divided by work day, depending on the response
  scale.
- Weekday/weekend: weekend versus weekday.
- Binary activity: Active versus Sedentary.
- Exercise intensity: one omnibus measure test; displayed level contrasts use
  no exercise as the reference but do not create extra confirmatory p-values.
- Exercise location: conditional exercise-day contrasts with outdoors as the
  reference.
- Device worn during exercise: Yes versus No as a measurement-support
  sensitivity, not a determinant claim.
- Continuous measures: slope for a one-hour increment. Sleep duration is
  centered at 8 h, onset at 23:00 on a noon-unwrapped axis, and wake at 07:00.

For a Tweedie log-link response, the practical effect is a ratio of expected
metric values. For identity Gaussian responses it is an absolute difference
in hours, clock hours, or MDER. For `log10(metric + 0.1)` responses, the exact
coefficient contrast is multiplicative on the **offset scale**; Stage 2 must
also report equal-site model-based marginal means in the original unit and
must not call that coefficient an exact ratio of unoffset metric values.
Clock-time predictions are returned modulo 24 h; L10 uses the registered
strict-after-16:00 unwrap before fitting.

The heterogeneity estimand is the omnibus change from the additive to the
predictor-by-site model on the same frame. Site-specific contrasts are shown
with 95% confidence intervals only after that hierarchy has been respected;
they are observational descriptions, not nine independent replications.

## Actual-date temporal support

```{r}
#| label: daily-sequence-audit

make_sequence_support <- function(column, predictor_id) {
  rows <- daily_metric_long |>
    filter(is.finite(response_value), !is.na(.data[[column]])) |>
    distinct(
      placement, metric_order, metric_id, manuscript_name,
      site, participant_key, local_date
    ) |>
    arrange(
      placement, metric_order, metric_id, participant_key, local_date
    ) |>
    group_by(
      placement, metric_order, metric_id, manuscript_name, participant_key
    ) |>
    mutate(
      day_gap = as.integer(local_date - lag(local_date)),
      sequence_number = cumsum(is.na(day_gap) | day_gap != 1L)
    ) |>
    ungroup()

  rows |>
    count(
      placement, metric_order, metric_id, manuscript_name,
      participant_key, sequence_number,
      name = "sequence_days"
    ) |>
    summarise(
      predictor_id = predictor_id,
      sequences = n(),
      adjacent_pairs = sum(pmax(sequence_days - 1L, 0L)),
      participants_with_adjacent_pair =
        n_distinct(participant_key[sequence_days >= 2L]),
      sequences_at_least_3_days = sum(sequence_days >= 3L),
      maximum_sequence_days = max(sequence_days),
      .by = c(placement, metric_order, metric_id, manuscript_name)
    )
}

temporal_support <- imap_dfr(
  predictor_columns,
  make_sequence_support
) |>
  left_join(
    predictor_contract,
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, metric_order, placement)

temporal_support_summary <- temporal_support |>
  summarise(
    `Adjacent pairs, minimum` = min(adjacent_pairs),
    `Adjacent pairs, maximum` = max(adjacent_pairs),
    `Participants with adjacency, minimum` =
      min(participants_with_adjacent_pair),
    `Participants with adjacency, maximum` =
      max(participants_with_adjacent_pair),
    `Sequences of ≥3 days, minimum` = min(sequences_at_least_3_days),
    `Longest sequence` = max(maximum_sequence_days),
    .by = c(predictor_order, reader_name, proposed_role, placement)
  ) |>
  arrange(predictor_order, placement)

temporal_support_summary |>
  select(-predictor_order) |>
  compact_gt(
    groupname_col = "proposed_role",
    source_note = paste0(
      "Adjacency means exactly one calendar day. Missing dates break ",
      "sequences. Counts are ranges across the 15 metric-specific frames."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::cols_label(reader_name = "Predictor", placement = "Placement")
```

All proposed core frames exceed the prespecified minimum support for an AR(1)
counterpart (at least 100 true adjacent pairs and 20 participants with an
adjacent pair). That establishes feasibility, not necessity. The AR(1) model
is triggered by the residual-dependence rule below; it is not fitted merely
because consecutive dates exist.

```{r}
#| label: exact-temporal-support-table

temporal_support |>
  select(
    `Predictor` = reader_name,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Sequences` = sequences,
    `True adjacent pairs` = adjacent_pairs,
    `Participants with adjacent pair` = participants_with_adjacent_pair,
    `Sequences of ≥3 days` = sequences_at_least_3_days,
    `Longest sequence, days` = maximum_sequence_days
  )
```

## Response-family screen on every exact candidate frame

```{r}
#| label: response-frame-screen

transform_response <- function(value, response_transform) {
  switch(
    response_transform,
    identity = value,
    log10_offset_0.1 = log10(value + 0.1),
    clock_hours = value / 60,
    clock_hours_midnight_after_16 = {
      hours <- value / 60
      ifelse(hours > 16, hours - 24, hours)
    },
    stop("Unknown response transformation: ", response_transform)
  )
}

make_response_frame_audit <- function(column, predictor_id) {
  daily_metric_long |>
    filter(is.finite(response_value), !is.na(.data[[column]])) |>
    group_by(
      placement, metric_order, metric_id, manuscript_name,
      response_family, response_transform, lower_bound, upper_bound,
      audit_upper_threshold
    ) |>
    summarise(
      predictor_id = predictor_id,
      participant_days = n(),
      zero_values = sum(response_value == 0),
      zero_fraction = mean(response_value == 0),
      minimum = min(response_value),
      median = median(response_value),
      maximum = max(response_value),
      lower_bound_violations = ifelse(
        all(is.na(lower_bound)),
        0L,
        sum(response_value < first(lower_bound) * ifelse(
          grepl("^clock_hours", first(response_transform)), 60, 1
        ))
      ),
      upper_bound_violations = ifelse(
        all(is.na(upper_bound)),
        0L,
        ifelse(
          grepl("^clock_hours", first(response_transform)),
          sum(response_value >= first(upper_bound) * 60),
          sum(response_value > first(upper_bound))
        )
      ),
      audit_upper_flags = ifelse(
        all(is.na(audit_upper_threshold)),
        0L,
        sum(response_value > first(audit_upper_threshold))
      ),
      support_bound_violations = ifelse(
        first(metric_id) %in% c(
          "duration_above_1000", "duration_above_250_wake",
          "duration_below_10_pre_sleep",
          "duration_below_1_sleep_environment"
        ),
        sum(
          is.finite(expected_minutes) &
            response_value > expected_minutes / 60 + 1e-10
        ),
        0L
      ),
      transformed_nonfinite = sum(!is.finite(
        transform_response(response_value, first(response_transform))
      )),
      .groups = "drop"
    )
}

response_frame_audit <- imap_dfr(
  predictor_columns,
  make_response_frame_audit
) |>
  left_join(
    predictor_contract,
    by = "predictor_id",
    relationship = "many-to-one"
  ) |>
  arrange(predictor_order, metric_order, placement)

stopifnot(
  nrow(response_frame_audit) == 11L * 15L * 2L,
  all(response_frame_audit$lower_bound_violations == 0L),
  all(response_frame_audit$upper_bound_violations == 0L),
  all(response_frame_audit$support_bound_violations == 0L),
  all(response_frame_audit$transformed_nonfinite == 0L)
)

response_screen_summary <- response_frame_audit |>
  summarise(
    `Candidate days, minimum` = min(participant_days),
    `Candidate days, maximum` = max(participant_days),
    `Zero fraction, minimum` = min(zero_fraction),
    `Zero fraction, maximum` = max(zero_fraction),
    `Bound violations` = sum(
      lower_bound_violations + upper_bound_violations +
        support_bound_violations
    ),
    `Nonfinite transformed values` = sum(transformed_nonfinite),
    .by = c(metric_order, manuscript_name, response_family, response_transform)
  ) |>
  arrange(metric_order)

response_screen_summary |>
  select(-metric_order) |>
  compact_gt(
    source_note = paste0(
      "Ranges cover all 11 predictor definitions and both placements. ",
      "Passing this source-scale screen does not establish residual or ",
      "distributional adequacy; those require Stage 2 fits."
    )
  ) |>
  gt::fmt_integer(
    columns = c(`Candidate days, minimum`, `Candidate days, maximum`,
      `Bound violations`, `Nonfinite transformed values`)
  ) |>
  gt::fmt_percent(
    columns = c(`Zero fraction, minimum`, `Zero fraction, maximum`),
    decimals = 1
  ) |>
  gt::cols_label(
    manuscript_name = "Metric",
    response_family = "Candidate family/link",
    response_transform = "Candidate transformation"
  )
```

The H05 registry provides defensible **candidates**, not transferred adequacy:

- Gaussian `log10(metric + 0.1)` for daily mean melEDI, M10/L10 means, the
  observed longest period, and corrected dose;
- Tweedie with a log link for the three zero-containing threshold durations
  selected for that family;
- identity Gaussian for pre-sleep time below 10 lx melEDI and MDER;
- Gaussian clock-hour models for timing metrics, with the strict-after-16:00
  L10 unwrap.

The 330 exact candidate frames contain no source-bound violation and no
nonfinite proposed transform. That is only the entry gate. Zero mass, tails,
heteroscedasticity, physical predictions, circular boundary behavior, and
residual dependence remain fit-level gates. A failed candidate is not retained
because H05 used it. A two-part, hurdle, alternative clock unwrap, or other
family change requires a bounded production-code pilot and an H06-D-G2
amendment before full fitting.

```{r}
#| label: exact-response-frame-table

response_frame_audit |>
  select(
    `Predictor` = reader_name,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Candidate family/link` = response_family,
    `Transformation` = response_transform,
    `Participant-days` = participant_days,
    `Exact zeros` = zero_values,
    `Zero fraction` = zero_fraction,
    `Minimum` = minimum,
    `Median` = median,
    `Maximum` = maximum,
    `Lower-bound violations` = lower_bound_violations,
    `Upper-bound violations` = upper_bound_violations,
    `Support-bound violations` = support_bound_violations,
    `Nonfinite transforms` = transformed_nonfinite
  )
```

# Placement and dataset scenarios

The author-approved placement hierarchy is:

1. **Near-eye all available — primary.** Each metric × predictor uses every
   admissible near-eye participant-day in that exact frame.
2. **Paired/common near-eye and chest — complementary.** Refit the near-eye
   and chest models on the same participant-day keys and compare estimates,
   uncertainty, direction, and practical magnitude.
3. **Chest all available — contextual complementary analysis.** Report the
   full chest sample and estimate/95% CI without treating it as a second
   confirmatory discovery screen.

This current placement hierarchy is a declared deviation from the signed
chest-primary wording; it follows the shared placement decision. Neither
placement is treated as a calibration gold standard. Comparisons are about
placement-specific measurement contexts, not measurement error in one device.

## Gap-timing-unaware dataset sensitivity

At first use, the **gap-timing-unaware dataset** means the prepared comparison
dataset that still applies the general hourly/day coverage rules but does not
use the timing of remaining missing observations for the additional
metric-specific adjustment. It is not called “unadjusted”, “raw”, or “original”.

```{r}
#| label: gap-dataset-candidate-audit

gap_day_raw <- read_project_rds(
  "artifacts/06_model_data/scenarios/manuscript_prepared_data/participant_day_metrics.rds"
)

gap_metric_long <- gap_day_raw |>
  filter(metric_id %in% metric_contract$metric_id) |>
  transmute(
    placement = factor(
      recode(
        position_role,
        near_eye_primary = "Near-eye",
        complementary_chest = "Chest"
      ),
      levels = c("Near-eye", "Chest")
    ),
    site,
    Id,
    local_date = as.Date(local_date),
    metric_id,
    response_value = manuscript_prepared_value
  ) |>
  left_join(
    metric_contract,
    by = "metric_id",
    relationship = "many-to-one"
  ) |>
  left_join(
    sleep_day |>
      select(-sleep_onset_date),
    by = c("site", "Id", "local_date"),
    relationship = "many-to-one"
  ) |>
  left_join(
    exercise_day,
    by = c("site", "Id", "local_date"),
    relationship = "many-to-one"
  ) |>
  mutate(
    site = factor(site, levels = site_levels),
    participant_key = paste(site, Id, sep = "::"),
    participant_day_key = paste(site, Id, local_date, sep = "::"),
    weekday_weekend = factor(
      if_else(
        as.POSIXlt(local_date)$wday %in% c(0L, 6L),
        "Weekend",
        "Weekday"
      ),
      levels = c("Weekday", "Weekend")
    )
  )

gap_scope <- gap_metric_long |>
  summarise(
    rows = n(),
    `Participant-days` = n_distinct(participant_day_key),
    Participants = n_distinct(participant_key),
    Sites = n_distinct(site),
    Metrics = n_distinct(metric_id),
    .by = placement
  )

stopifnot(
  gap_scope$`Participant-days`[gap_scope$placement == "Near-eye"] == 811L,
  gap_scope$`Participant-days`[gap_scope$placement == "Chest"] == 897L,
  all(gap_scope$Metrics == 15L),
  nrow(gap_metric_long) == 15L * (811L + 897L)
)

gap_scope |>
  compact_gt(
    source_note = paste0(
      "Counts precede metric-value and predictor completeness. The comparison ",
      "dataset is not substituted into the primary frame."
    )
  ) |>
  gt::fmt_integer(columns = where(is.numeric))
```

```{r}
#| label: gap-exact-sample-audit

gap_candidate_missingness <- imap_dfr(
  predictor_columns,
  function(column, predictor_id) {
    gap_metric_long |>
      mutate(
        response_available = is.finite(response_value),
        predictor_available = !is.na(.data[[column]]) &
          if (is.numeric(.data[[column]])) {
            is.finite(.data[[column]])
          } else {
            TRUE
          },
        status = case_when(
          response_available & predictor_available ~ "Complete candidate",
          !response_available & predictor_available ~ "Metric unavailable only",
          response_available & !predictor_available ~ "Predictor unavailable only",
          TRUE ~ "Both unavailable"
        )
      ) |>
      summarise(
        predictor_id = predictor_id,
        base_days = n(),
        complete_days = sum(status == "Complete candidate"),
        metric_unavailable_only = sum(status == "Metric unavailable only"),
        predictor_unavailable_only = sum(status == "Predictor unavailable only"),
        both_unavailable = sum(status == "Both unavailable"),
        participants = n_distinct(
          participant_key[status == "Complete candidate"]
        ),
        sites = n_distinct(site[status == "Complete candidate"]),
        .by = c(placement, metric_order, metric_id, manuscript_name)
      )
  }
) |>
  left_join(
    predictor_contract,
    by = "predictor_id",
    relationship = "many-to-one"
  )

gap_available_rows <- imap_dfr(
  predictor_columns,
  function(column, predictor_id) {
    gap_metric_long |>
      filter(is.finite(response_value), !is.na(.data[[column]])) |>
      distinct(
        placement, site, Id, local_date, participant_key,
        metric_order, metric_id, manuscript_name
      ) |>
      mutate(predictor_id = predictor_id)
  }
)

gap_paired_samples <- gap_available_rows |>
  count(
    site, Id, local_date, participant_key, metric_order, metric_id,
    manuscript_name, predictor_id,
    name = "placements"
  ) |>
  filter(placements == 2L) |>
  summarise(
    paired_days = n(),
    paired_participants = n_distinct(participant_key),
    paired_sites = n_distinct(site),
    .by = c(predictor_id, metric_order, metric_id, manuscript_name)
  )

gap_candidate_samples <- gap_candidate_missingness |>
  left_join(
    gap_paired_samples,
    by = c("predictor_id", "metric_order", "metric_id", "manuscript_name"),
    relationship = "many-to-one"
  ) |>
  mutate(
    across(
      c(paired_days, paired_participants, paired_sites),
      ~ replace_na(.x, 0L)
    )
  ) |>
  arrange(predictor_order, metric_order, placement)

stopifnot(
  nrow(gap_candidate_samples) == 11L * 15L * 2L,
  all(with(
    gap_candidate_samples,
    complete_days + metric_unavailable_only +
      predictor_unavailable_only + both_unavailable == base_days
  ))
)

gap_candidate_samples |>
  select(
    `Predictor` = reader_name,
    `Role` = proposed_role,
    `Metric` = manuscript_name,
    `Placement` = placement,
    `Base rows` = base_days,
    `Complete candidate days` = complete_days,
    `Metric unavailable only` = metric_unavailable_only,
    `Predictor unavailable only` = predictor_unavailable_only,
    `Both unavailable` = both_unavailable,
    `Participants` = participants,
    `Sites` = sites,
    `Paired/common days` = paired_days,
    `Paired/common participants` = paired_participants,
    `Paired/common sites` = paired_sites
  )
```

Stage 2 applies the same approved formula, coding, family, diagnostic, and
multiplicity contracts to this sensitivity, with scenario-specific exact
samples. Primary and gap-timing-unaware results are compared on both their
all-available samples and their common participant-day keys. The sensitivity
has its own BH vectors; its p-values are never pooled with the primary vector.

Other metric-implementation sensitivities already fixed upstream—especially
the exactly identified continuous-period subset and declared MDER/coverage
support variants—are reported as stability checks without expanding the
primary discovery family. No new metric is derived inside H06_daily.

# Multiplicity contract

The signed instruction is false-discovery-rate control within a declared
hypothesis family. The unit of multiplicity here is therefore the vector of
**15 metric-specific omnibus tests for one predictor**, not the coefficients
inside one model and not one isolated scalar p-value.

```{r}
#| label: multiplicity-registry

multiplicity_registry <- predictor_contract |>
  filter(predictor_id != "wore_light_logger_during_exercise") |>
  transmute(
    predictor_order,
    predictor_id,
    predictor = reader_name,
    proposed_role,
    association_family = sprintf("H06-D-A%02d", predictor_order),
    association_members = 15L,
    heterogeneity_family = sprintf("H06-D-H%02d", predictor_order),
    heterogeneity_members = 15L,
    activation = case_when(
      proposed_role == "Core primary candidate" ~
        "Proposed active confirmatory family after H06-D-G1",
      proposed_role == "Registered named sensitivity" ~
        "Separate registered sensitivity family if author approves p-values",
      TRUE ~
        "Dormant; activate prospectively only if author promotes this predictor"
    )
  ) |>
  arrange(predictor_order)

multiplicity_registry |>
  select(-predictor_order, -predictor_id) |>
  compact_gt(
    groupname_col = "proposed_role",
    source_note = paste0(
      "Each family contains one omnibus association or heterogeneity p-value ",
      "for each of the 15 metrics. Logger-worn status has no determinant-test family."
    )
  ) |>
  gt::fmt_integer(columns = c(association_members, heterogeneity_members)) |>
  gt::cols_label(
    predictor = "Predictor",
    association_family = "Association BH family",
    association_members = "Slots",
    heterogeneity_family = "Heterogeneity BH family",
    heterogeneity_members = "Slots",
    activation = "Gate status"
  )
```

For every activated family, Stage 2 constructs a named 15-slot vector in
metric-registry order and applies `stats::p.adjust(p, method = "BH", n = 15)`.
Non-estimable tests remain named `NA` slots and are reported; they are not
dropped to make a smaller-looking family. The association vector contains the
reduced-versus-additive omnibus test. The heterogeneity vector separately
contains the additive-versus-interaction omnibus test. Multi-level predictor
coefficients and pairwise practical contrasts do not enter merely because
they were estimated.

The proposed active confirmatory association families are work/free day,
Sedentary/Active status, and previous-night sleep duration, each across 15
metrics. Their three heterogeneity vectors are separate. Weekday/weekend is
registered but proposed as a named sensitivity with its own vector only if the
author wants inferential p-values. Other registered measures remain
exploratory unless prospectively promoted at H06-D-G1. Near-eye primary and
gap-timing-unaware sensitivity vectors are adjusted separately. Complementary
chest and paired-placement comparisons are proposed as estimates with 95%
confidence intervals, not additional significance screens.

All tables report raw and BH-adjusted p-values in separate labelled columns.
P-values are formatted to three decimals, with `<0.001` below that threshold.
Bold is applied only under the explicitly labelled raw or adjusted decision
rule; an omnibus decision is never inferred from a bold coefficient.

# Diagnostic and failure contract

Every diagnostic is evaluated on the exact metric × predictor × placement ×
dataset frame used for its estimate. “Acceptable” below means acceptable for
the specified claim; it does not mean the model is true. A failed diagnostic
cannot be repaired by silently deleting a participant, site, category, date,
or zero.

```{r}
#| label: diagnostic-decision-rules

diagnostic_rules <- tibble::tribble(
  ~domain, ~required_check, ~acceptable, ~not_acceptable_or_action,
  "Frame and joins",
  "One row per site–participant–date; finite response transform; frozen ordered frame and hash",
  "No duplicate key, no unlabelled exclusion, no bound/transform violation; reduced/additive/interaction use identical rows",
  "Any mismatch is a pipeline failure; do not fit or report",
  "Fixed-effect design",
  "Model-matrix rank, reference levels, site × category cells, within-site continuous variation",
  "Full column rank; every categorical interaction level represented within every site in the exact frame; positive within-site variation for continuous heterogeneity",
  "Aliased column or missing within-site contrast: heterogeneity model is NON_ESTIMABLE; additive model is separately assessed",
  "Convergence",
  "Optimizer status, finite coefficients/SEs, scaled gradient, Hessian",
  "Convergence code 0; finite estimates/SEs; positive-definite Hessian; for lme4 max |scaled gradient| ≤0.002",
  "Any optimizer error, non-positive Hessian, nonfinite estimate/SE, or gradient >0.002 is not acceptable",
  "Singularity",
  "Participant intercept and registered random-site benchmark",
  "lme4 isSingular(tol = 1e-4) is FALSE; variances finite; registered-site correlations not at boundary",
  "Singular fit, zero/near-zero required variance, or |random-effect correlation| ≥0.98: label benchmark NON_ESTIMABLE; never simplify it silently",
  "Gaussian residual shape",
  "Q–Q, residual-versus-fitted, scale-location, residual by site/category/date",
  "No structural pattern; <1% standardized residuals exceed |4|; |Spearman residual-spread trend| <0.20",
  "Trend 0.20–0.40 or tail excess is a limitation and triggers candidate comparison; ≥0.40 or persistent site/category structure rejects the candidate",
  "Tweedie distribution",
  "Power/dispersion, simulated residual uniformity/dispersion/outliers, zeros, upper tail",
  "Finite 1 < power < 2; observed zero fraction and tail summaries lie within the central 95% simulation envelope; no systematic residual pattern",
  "Boundary power, failed simulation, observed zero/tail outside envelope, or systematic residual structure rejects the candidate",
  "Exact zero mass",
  "Zero proportion by site/predictor and residual pile at the transformed zero",
  "<5% is routine; 5–10% is reported; ≥10% requires a prespecified two-part/discrete-mass pilot before accepting a one-part log-Gaussian model",
  "If the pilot changes direction by >1 primary-model SE, reverses direction, or the one-part model remains miscalibrated, the one-part family is not acceptable",
  "Physical bounds",
  "Original-scale point predictions over the observed design and support-specific duration limits",
  "No more than 1% of point predictions outside a physical bound by >0.05 display units; timing returned to [0, 24)",
  "More than 1% impossible predictions, or any severe violation, rejects the family/link for that metric",
  "Clock outcomes",
  "Transformed distribution, cut-point pile, circular residuals, alternative unwrap screen",
  "Smallest circular arc containing 95% of values and transformed central span are each ≤18 h, with no split cluster across the cut point",
  "Boundary split, >18 h central span, or circular residual pattern rejects linear Gaussian clock modeling and triggers a gated timing alternative",
  "Continuous-period identification",
  "Observed-lower-bound primary versus exactly identified subset",
  "Same direction and absolute effect shift <1 primary-model SE; censoring pattern reported by site/predictor",
  "Shift 1–2 SE is a major limitation; >2 SE or direction reversal prevents an unconditional period claim",
  "Participant influence",
  "Leave-one-participant refits; coefficient/contrast shift on primary scale",
  "Maximum absolute shift <1 full-model SE, no direction reversal, no adjusted-family decision changed",
  "Shift 1–2 SE is influential and reported; >2 SE, reversal, or repeated decision changes rejects a stable claim",
  "Site influence",
  "Leave-one-site refits for all submitted sites; equal-site estimand recalculated",
  "Direction retained and shift <1 full-model SE for every deletion",
  "Shift 1–2 SE limits generality; >2 SE or reversal means the cross-site claim is not acceptable",
  "Daily temporal dependence",
  "Residual correlation only for true one-day lags within participant; site-specific screen",
  "Pooled |lag-1 r| <0.20 and every site |r| <0.30",
  "Pooled 0.20–0.40 or any site ≥0.30 triggers AR(1); pooled >0.40 without an adequate AR counterpart rejects the independent-residual result",
  "Gap-aware AR(1)",
  "At least 100 adjacent pairs and 20 participants; convergence; |rho| away from boundary",
  "Support threshold met, convergence gates pass, |rho| <0.95, and association is stable within 1 primary-model SE",
  "Sparse support, failed convergence, |rho| ≥0.95, >2-SE change, or direction reversal means temporal stability is unresolved",
  "Placement",
  "All-available near-eye, paired/common near-eye/chest, and all-available chest",
  "Exact samples shown; direction and practical magnitude compared with 95% CIs; no gold-standard language",
  "Direction reversal or >2-SE paired-frame shift prevents a placement-general claim and is reported as measurement-context heterogeneity",
  "Multiplicity",
  "Named 15-slot raw and adjusted vectors; omnibus provenance",
  "Every slot present in registry order; BH uses n = 15; NA retained; raw/adjusted decisions labelled",
  "Scalar adjustment, coefficient stuffing, dropped NA slots, or family assembled after results invalidates inference",
  "Resampling",
  "Production-code 50/100-resample pilot, failure rate, runtime, Monte Carlo precision",
  "Pilot output shown and runtime approved before full run; checkpoint/restart deterministic; failed draws reason-coded",
  "No full heavy batch before approval; excessive failure or unstable Monte Carlo error reopens the resampling plan"
)

diagnostic_rules |>
  compact_gt(
    source_note = paste0(
      "Thresholds are prospective Stage 1 decision rules. Diagnostic p-values, ",
      "if produced, are screens and are interpreted with the graphical and ",
      "quantitative criteria rather than as isolated truth tests."
    )
  ) |>
  gt::cols_label(
    domain = "Domain",
    required_check = "Required check",
    acceptable = "Acceptable",
    not_acceptable_or_action = "Not acceptable / required action"
  )
```

The current source audit already passes uniqueness, wide/long value
reconciliation, bounds, and transform finiteness on every candidate frame.
It does **not** pre-pass convergence, residual, zero-mass, influence, temporal,
or prediction diagnostics, because those require fitted models. The ≥10%
zero-mass rule prospectively identifies L10 mean as requiring a bounded
one-part versus two-part/discrete-mass pilot before its final family can be
accepted; this conclusion follows source values, not H06 association results.

Stage 2 records every check as `ACCEPTABLE`, `ACCEPTABLE_WITH_LIMITATION`,
`NOT_ACCEPTABLE`, or `NOT_ESTIMABLE`, with a plain-language explanation. A
model batch is not complete until every metric has a disposition.

# Reporting and durable outputs

Stage 2 must report, for every model and scenario:

- the exact fitted participant-days, participants, sites, category cells,
  placement, dataset scenario, response family/link/transformation, formula,
  engine, optimizer, R/package versions, and model-frame hash;
- equal-site original-scale marginal means or slopes, practical contrasts,
  95% confidence intervals, raw omnibus p-values, and the explicitly named
  BH-adjusted family value where applicable;
- convergence, singularity, residual, zero-mass, bounds, influence,
  temporal-dependence, site-deletion, and metric-specific support assessments
  in plain language;
- near-eye all available, paired/common near-eye and chest, all-available chest
  context, and gap-timing-unaware results without mixing their samples;
- observational “associated with” language only—never causal effects, health
  outcomes, or unsupported generalization.

Reader-facing zero-containing strongly right-skewed quantities such as melEDI
may use `LightLogR::symlog_trans(base = 10, thr = 1, scale = 1)`. The source
and model-frame values remain untransformed except for the declared response
transformation. Figures use the submitted site names, order, and colours from
`config/site_display_registry.csv`, publication-scale typography, final-size
clipping/wrapping/panel-balance checks, informative alt text, and a paired
source-data CSV. Compact `gt` tables use manuscript metric names and units.

# Comparison with V0

## Implementation-level comparison

| Component | V0 implementation | Proposed H06 daily implementation | Consequence |
|---|---|---|---|
| Outcome | One hourly geometric mean melEDI outcome | Fifteen distinct participant-day metrics | Results answer different scientific questions |
| Formula | `geo.MEDI ~ site*daytype + site*exercise + site*sleep_duration + (1|Id)` | Separate metric × measure reduced/additive/heterogeneity hierarchy plus registered benchmark | No V0 coefficient maps one-to-one to the daily family |
| Predictor scope | Free/work, four-level exercise intensity, sleep duration jointly; weekday/weekend post hoc | Exact registered scope recovered; three-core proposal pending H06-D-G1 | V0 does not settle omitted registered variables |
| Day weighting | One row per retained hour | One row per admissible metric-specific day | Days with more supported hours no longer receive more rows |
| Dependence | Participant intercept only | Participant intercept plus actual-date residual screen and conditional gap-aware daily AR(1) | Standard errors and target differ |
| Site | Fixed predictor interactions in V0; stale displayed site p-value | Equal-site fixed-site association and separate heterogeneity; exact random-site benchmark visible | Site interpretation is rebuilt, not copied |
| Multiplicity | Raw block p-values labelled as FDR; site p-value inherited from H04 | Named 15-slot association and heterogeneity BH vectors | V0 significance labels cannot be retained |
| Uncertainty | Final table omitted 95% CIs | 95% CIs mandatory for every estimate | Daily claims require interval-based magnitude |
| Placement | Signed chest primary but near-eye-focused reporting | Current approved near-eye primary with paired/common and chest complementary | Placement roles are transparent |

## Claim-level comparison

V0 hourly ratios, site contrasts, and reported “unique” variance fractions are
historical descriptors only. They cannot be reworded as effects on daily mean,
M10/L10, threshold durations, timing, dose, MDER, or the longest period. V0's
unadjusted predictor-block p-values must not be carried forward as FDR results,
and its displayed site p-value is not an H06 model result. The daily analysis
may support only metric-specific observational statements on its own exact
frames, with 95% CIs and vector-wide multiplicity.

Stage 2 will compare implementations and claims side by side and recreate
recognizable daily-metric tables/figures where the new estimand permits. It
will explicitly mark V0 claims as retained, narrowed, replaced, or unsupported.
It will not imply that an hourly ratio and a daily-metric contrast should be
numerically similar.

# Measurement-methods audit findings

```{r}
#| label: audit-findings

audit_findings <- tibble::tribble(
  ~finding_id, ~severity, ~confidence, ~status, ~finding, ~required_disposition,
  "H06-D-FIND-001", "Major", "High", "Confirmed",
  "The signed H06 outcome/unit is participant-day daily metrics; V0 and the approved main H06 are hourly estimands.",
  "Keep H06_daily standalone; never describe it as an hourly sensitivity or replace main H06.",
  "H06-D-FIND-002", "Major", "High", "Author decision required",
  "The signed predictor scope is broader than the provisional work/free, binary activity, and sleep-duration set.",
  "Approve the primary/sensitivity/exploratory partition at H06-D-G1 before fitting.",
  "H06-D-FIND-003", "Major", "High", "Confirmed design limitation",
  "Nine near-eye and eight chest sites do not support a defensible primary random-site/random-slope distribution, especially for multi-level predictors.",
  "Use the exact registered formula only as a labelled estimability benchmark; use fixed site as proposed primary.",
  "H06-D-FIND-004", "Moderate", "High", "Open upstream qualification",
  "Current metric/base artifacts are approved under PREP06-BASE-002, but PREP-003/FIND-044 remains attached to the exact current metric manifest.",
  "Carry the provenance qualification through models, tables, figures, and claims; do not rebuild preparation locally.",
  "H06-D-FIND-005", "Assurance", "High", "Passed",
  "All 25,770 selected metric rows reconcile exactly between enriched and long-form artifacts; diary keys are unique; the current TUM source has seven S001 and zero S101 rows.",
  "Repeat these assertions in production code and freeze the exact frame hashes.",
  "H06-D-FIND-006", "Moderate", "High", "Prespecified model gate",
  "Several outcomes contain exact zeros and L10 mean crosses the ≥10% zero-mass pilot threshold on current candidate frames.",
  "Run the bounded production-code zero-mass family pilot before accepting its final one-part family.",
  "H06-D-FIND-007", "Major", "High", "Confirmed V0 reporting defect",
  "V0 labels raw predictor-block p-values as FDR and displays a site p-value produced by H04; final intervals are absent.",
  "Do not propagate V0 inferential labels; rebuild H06_daily multiplicity and uncertainty from owned outputs."
)

audit_findings |>
  compact_gt(
    source_note = paste0(
      "Severity concerns the validity of an H06 daily claim. Assurance is a ",
      "passed control, not evidence that later models will be adequate."
    )
  ) |>
  gt::cols_label(
    finding_id = "Finding",
    severity = "Severity",
    confidence = "Confidence",
    status = "Status",
    finding = "Evidence-backed finding",
    required_disposition = "Required disposition"
  )
```

# Four-stage workflow and ownership

## Stage 2 — only after H06-D-G1 approval

Stage 2 will use only `scripts/hypotheses/H06_daily/`,
`tests/hypotheses/H06_daily/`, and `artifacts/*/H06_daily/`. It will construct
immutable model-frame registries, run the approved hierarchy, compare with V0,
and report all diagnostics and scenarios above. It will not read unfinished
main-H06 model outputs or write to any main-H06 path.

A small production-code pilot precedes any heavy batch. It must include
representatives of the log-offset Gaussian, zero-heavy log-offset Gaussian,
Tweedie-log, identity Gaussian, and clock-time families. Any bootstrap or
resampling begins with 50 or 100 draws. Pilot outputs, empirical failure rate,
checkpoint behavior, and projected full runtime are shown to the author; the
full run waits for explicit approval and will not compete with an active heavy
main-H06 job.

## Stage 3 — only after Stage 2/production approval

Create `notebooks/hypotheses/H06_daily.qmd` as a standalone reader-facing
daily-metric report with an “Answer in brief” callout, no V0 or construction-
history language, and no modification of `notebooks/hypotheses/H06.qmd`.

## Stage 4 — only after Stage 3 approval

Create
`audit/hypotheses/H06_daily/H06_daily_analysis_preparation.qmd` as the bounded
scientific preparation/provenance companion. Website integration is not
requested unless the coordinator and author later decide that this variant
belongs in the submission-facing site.

# H06-D-G1 author decisions

::: {.callout-important title="Stop gate — approve or amend every item before Stage 2"}
1. **Primary predictor set.** Approve the recommendation of work/free day,
   binary Sedentary/Active status, and previous-night sleep duration as three
   separate primary measures.
2. **Registered remainder.** Approve weekday/weekend as a named registered
   sensitivity; decide whether exercise intensity, exercise location, active
   travel, sedentary time, sleep onset, and wake are exploratory or promoted
   prospectively. Confirm logger-worn status as measurement support only.
3. **Separate versus joint estimand.** Approve separate one-measure models as
   the signed primary contract; retain any mutually adjusted three-core model
   as explicitly exploratory.
4. **Site structure.** Approve fixed site with equal-site marginalization as
   primary, predictor-by-site heterogeneity as a separate hierarchy, and the
   signed random-site/random-slope formula as benchmark-only with
   `NON_ESTIMABLE` permitted.
5. **Response families.** Approve the 15 H05-derived continuity candidates
   subject to fresh H06 diagnostics, including the mandatory bounded L10
   zero-mass pilot and gated family amendment on failure.
6. **Timing outcomes.** Approve the clock transforms and the prospective
   18-hour/circular-boundary failure rule; a failed timing candidate pauses for
   an approved alternative rather than being forced through Gaussian fitting.
7. **Temporal rule.** Approve the true-date residual thresholds and the
   gap-aware daily AR(1) counterpart only when triggered and adequately
   supported.
8. **Multiplicity.** Approve one 15-slot BH association family per activated
   predictor, separate 15-slot heterogeneity families, `n = 15` with retained
   NA slots, and no inferential p-values for dormant exploratory predictors.
9. **Placement.** Approve near-eye all available as primary, paired/common
   near-eye/chest as complementary, and all-available chest as context without
   a second discovery screen.
10. **Dataset/metric sensitivities.** Approve the gap-timing-unaware dataset
    comparison, current/common samples, the exactly identified period subset,
    and other fixed prepared support variants as stability analyses.
11. **KNUST provenance.** Confirm the H06-local interpretation of the named
    2024-11-04 sedentary source numeral as 3,600 seconds = 1 h only if that
    exploratory variable is used.
12. **Diagnostic failure rules.** Approve the explicit acceptable/limitation/
    not-acceptable thresholds, including participant/site deletion and
    physical-prediction gates.
13. **V0 claims.** Approve the rule that no V0 hourly estimate or inferential
    label transfers numerically; Stage 2 will disposition each historical
    claim against daily evidence.
14. **Compute gate.** Approve a small production-code family pilot first and a
    separate approval after the projected runtime is known for any heavy full
    batch or resampling.

**No inferential fitting begins until H06-D-G1 is explicitly approved.**
:::

```{r}
#| label: runtime-record

runtime_record <- tibble::tibble(
  item = c(
    "R", "melidosData", "dplyr", "gt", "digest", "Quarto engine"
  ),
  version = c(
    as.character(getRversion()),
    as.character(utils::packageVersion("melidosData")),
    as.character(utils::packageVersion("dplyr")),
    as.character(utils::packageVersion("gt")),
    as.character(utils::packageVersion("digest")),
    "knitr"
  )
)

runtime_record |>
  compact_gt(
    source_note = paste0(
      "Render-time analytical environment. This Stage 1 document performs ",
      "data/contract audits only and contains no fitting call."
    )
  ) |>
  gt::cols_label(item = "Component", version = "Version")
```
