---
title: "H06 daily metrics — Stage 2 bounded pilot"
subtitle: "Daily-family checks and the exploratory 30-minute time-of-day GAMM amendment"
date: last-modified
engine: knitr
format:
  html:
    toc: true
    toc-depth: 4
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    df-print: paged
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 130,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("melidosData")), "1.0.6")
)

source(file.path(
  project_root,
  "scripts/hypotheses/H06_daily/h06_daily_contract.R"
))

read_h06d <- function(stage, file) {
  readr::read_csv(
    file.path(project_root, "artifacts", stage, "H06_daily", file),
    show_col_types = FALSE,
    na = ""
  )
}

compact_gt <- function(data, title = NULL, subtitle = NULL, source_note = NULL) {
  output <- data |>
    gt::gt() |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(12),
      table.width = gt::pct(100),
      data_row.padding = gt::px(3),
      heading.align = "left",
      source_notes.font.size = gt::px(10)
    )
  if (!is.null(title)) {
    output <- output |>
      gt::tab_header(
        title = gt::md(title),
        subtitle = if (is.null(subtitle)) NULL else gt::md(subtitle)
      )
  }
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

status_gt <- function(data, title, subtitle = NULL, source_note = NULL) {
  output <- compact_gt(data, title, subtitle, source_note)
  if ("Status" %in% names(data)) {
    output <- output |>
      gt::data_color(
        columns = "Status",
        fn = function(value) {
          dplyr::case_when(
            grepl("^PASS", value) ~ "#D9F2E6",
            grepl("WARN|TRIGGER|RETAIN", value) ~ "#FFF1CC",
            grepl("FAIL|NON_ESTIMABLE", value) ~ "#FADBD8",
            TRUE ~ "#EEEEEE"
          )
        }
      )
  }
  output
}

input_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_pilot_input_manifest.csv"
)
output_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_pilot_output_manifest.csv"
)
software <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_pilot_software_manifest.csv"
)
daily_registry <- read_h06d(
  "06_model_data",
  "H06_daily_stage2_pilot_registry.csv"
)
temporal_formula_registry <- read_h06d(
  "06_model_data",
  "H06_daily_30_minute_temporal_formula_registry.csv"
)
daily_diagnostics <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_pilot_daily_diagnostics.csv"
)
daily_verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_pilot_daily_verdict.csv"
)
daily_preview <- read_h06d(
  "09_tables",
  "H06_daily_stage2_pilot_effect_preview.csv"
)
l10_family <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_zero_mass_family_pilot.csv"
)
temporal_component <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_component_diagnostics.csv"
)
temporal_calibration <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_calibration.csv"
)
temporal_k <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_k_check.csv"
)
temporal_verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_diagnostic_verdict.csv"
)
temporal_closure <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_cyclic_closure.csv"
)
temporal_site_acf <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_site_residual_acf.csv"
)
temporal_concurvity <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_concurvity_fallback.csv"
)
temporal_boundaries <- read_h06d(
  "08_diagnostics",
  "H06_daily_30_minute_temporal_ar_boundaries.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_pilot_runtime_projection.csv"
)

input_hashes <- vapply(
  file.path(project_root, input_manifest$relative_path),
  digest::digest,
  character(1),
  algo = "sha256",
  file = TRUE
)
output_hashes <- vapply(
  file.path(project_root, output_manifest$relative_path),
  digest::digest,
  character(1),
  algo = "sha256",
  file = TRUE
)
stopifnot(
  identical(unname(input_hashes), input_manifest$sha256),
  identical(unname(output_hashes), output_manifest$sha256)
)
```

# Pilot verdict

::: {.callout-important title="H06-D-G2P — stop before any full production batch"}
The Stage 1 plan is author-approved, and the requested exploratory 30-minute
time-of-day GAMM has been added as a separate estimand. The bounded pilot does
**not** justify a full run yet.

- Representative fixed-site daily models converged, but the exact signed
  random-site/random-slope benchmark was singular and is therefore
  `NON_ESTIMABLE` in the pilot.
- The mandatory L10 check found 13.3% exact zeros. Its one-part and two-part
  point estimates had the same direction but materially different magnitude,
  so the one-part family is not accepted.
- The 30-minute one-part Tweedie GAMM is rejected because its zero probability
  is badly miscalibrated, its `k = 12` time functions are capacity-limited, and
  material residual autocorrelation remains after the working AR correction.
- A two-part binomial/Gamma architecture is more defensible, converges, and is
  exactly cyclic at midnight. Its current `k = 12` positive component and
  residual dependence still fail the production-effect gate. No day-type,
  activity, or sleep curve is interpreted from this pilot.
- Six temporal `bam()` fits on the primary near-eye frame took 15.0 minutes.
  A straight six-frame fit-only projection is about 79 minutes; measured
  overhead suggests allowing roughly 90 minutes before any uncertainty or
  deletion work.

The recommended next action is one further **near-eye-only repair pilot**, not
the full Stage 2 batch.
:::

# What was approved and what was added

The author approved every H06-D-G1 item on 2026-08-11. The approval is recorded
in `audit/hypotheses/H06_daily/H06_daily_stage1_gate_and_stage2_transition.md`.
The amendment asks how the same three core day contexts relate to the supported
30-minute geometric-mean melEDI metric over local clock time.

```{r}
#| label: tbl-estimand-separation
#| tbl-cap: "The three H06 estimands remain scientifically separate."

tibble::tribble(
  ~Analysis, ~Model_row, ~Question, ~Day_weighting, ~Role,
  "H06 daily metrics", "One metric-specific participant-day",
  "How each day-level context is associated with each of 15 daily metrics",
  "Each admissible metric-day contributes one row",
  "Preregistration-aligned primary variant",
  "Approved hourly H06", "One supported one-hour outcome",
  "How contexts are associated with expected hourly melEDI",
  "Days with more supported hours contribute more rows",
  "Separate approved analysis; no estimate transfers",
  "Exploratory 30-minute GAMM", "One supported 30-minute bin",
  "How context-associated profiles vary over local clock time",
  "Days with more supported 30-minute bins contribute more rows",
  "Exploratory amendment; outside daily BH families"
) |>
  compact_gt(
    source_note = paste0(
      "The 30-minute response is the zero-aware within-bin geometric mean. ",
      "A raw-scale model mean is therefore the conditional arithmetic mean ",
      "of that geometric-mean metric."
    )
  )
```

# Immutable inputs and exact frame

All pilot inputs match their stored SHA-256 identities. Preparation was not
rebuilt. The diary remains the corrected `melidosData` 1.0.6 source with seven
TUM S001 exercise rows and zero S101 rows. The H06-local KNUST sedentary-time
interpretation was not used because sedentary duration is not a pilot
predictor.

```{r}
#| label: tbl-pilot-inputs
#| tbl-cap: "Immutable pilot inputs."

input_manifest |>
  transmute(
    Input = .data$input_id,
    Path = .data$relative_path,
    `SHA-256` = substr(.data$sha256, 1, 12),
    `Size (MiB)` = .data$bytes / 1024^2
  ) |>
  compact_gt(
    source_note = "The shortened hashes are display-only; full hashes are stored and verified during render."
  ) |>
  gt::fmt_number(columns = `Size (MiB)`, decimals = 2)
```

The 30-minute complete-core near-eye frame contains **33,057** bins from 715
participant-days, 137 participants, and nine sites. It contains 10,225 exact
zeros (30.9%). The positive-magnitude component contains 22,832 bins. No fitted
row lacks true-time provenance.

```{r}
#| label: tbl-ar-boundaries
#| tbl-cap: "Why each 30-minute AR sequence starts."

temporal_boundaries |>
  transmute(
    Component = recode(
      .data$component,
      all_rows = "All supported bins",
      positive_rows = "Positive bins"
    ),
    Reason = recode(
      .data$ar_start_reason,
      participant_day_start = "Participant-day start",
      elapsed_discontinuity = "Unsupported/removed interval or elapsed gap",
      non_one_to_one_wall_outcome = "Non-one-to-one fall-back outcome",
      after_non_one_to_one_wall_outcome = "After fall-back outcome",
      continuous = "Continuous 30-minute sequence"
    ),
    Observations = .data$observations
  ) |>
  compact_gt(
    source_note = paste0(
      "Positive-magnitude sequences are reconstructed after zero removal. ",
      "AR(1) never bridges a zero, unsupported bin, day boundary, elapsed gap, ",
      "or non-one-to-one fall-back outcome."
    )
  ) |>
  gt::fmt_integer(columns = Observations)
```

# Exact model contracts

## Daily hierarchy exercised by the pilot

The daily pilot used the same generic formulas that production would use.
They are shown here with work/free day as the evaluated `Measure`; production
substitutes each metric-specific response and one approved predictor at a time.

```{r}
#| label: daily-formulas-visible
#| echo: true
#| code-fold: false

daily_formulas <- h06d_daily_formula_set("work_free_day")
daily_formula_table <- tibble::tibble(
  model_id = names(daily_formulas),
  formula = vapply(
    daily_formulas,
    function(formula) paste(deparse(formula), collapse = " "),
    character(1)
  )
)
```

```{r}
#| label: tbl-daily-formulas
#| tbl-cap: "Evaluated daily pilot formulas."

daily_formula_table |>
  transmute(
    Role = recode(
      .data$model_id,
      registered_random_site = "Exact signed benchmark",
      fixed_site_reduced = "Fixed-site reduced",
      fixed_site_additive = "Fixed-site additive",
      fixed_site_heterogeneity = "Fixed-site predictor-by-site"
    ),
    `Wilkinson formula` = .data$formula
  ) |>
  compact_gt(
    source_note = paste0(
      "The exact signed benchmark is never silently simplified. ",
      "The fixed-site additive model remains the proposed primary hierarchy."
    )
  )
```

## Exploratory 30-minute formulas

The H02 precedent supplies 30-minute midpoints, local-clock cyclicity, true UTC
ordering, participant-day/gap resets, fREML, and `bam(discrete = TRUE)`. The H03
precedent supplies a raw-outcome mean candidate and global-plus-context/site
time structure. H06-daily adds separate work/free and activity deviations plus
a cyclic varying-coefficient sleep term.

```{r}
#| label: temporal-formulas-visible
#| echo: true
#| code-fold: false

evaluated_temporal_formulas <- h06d_temporal_formula_registry()
```

```{r}
#| label: tbl-temporal-formulas
#| tbl-cap: "Evaluated exploratory temporal formulas supplied to bam()."

evaluated_temporal_formulas |>
  transmute(
    Candidate = recode(
      .data$candidate_id,
      one_part_tweedie = "One-part Tweedie",
      two_part_occurrence = "Two-part occurrence",
      two_part_positive = "Two-part positive magnitude"
    ),
    Component = .data$component,
    `Family / link` = .data$response_family,
    `Wilkinson formula` = .data$formula
  ) |>
  compact_gt(
    source_note = paste0(
      "Every clock smooth and factor/site deviation is explicitly cyclic. ",
      "The numeric-by sleep smooth is a time-varying linear association per ",
      "one-hour longer previous sleep, centered at eight hours."
    )
  )
```

All temporal fits use one R thread, `method = "fREML"`, `discrete = TRUE`, and
clock knots at 0 and 24 h. The global time smooth, work/free deviations,
activity deviations, and site deviations use `k = 12`; participant-specific
cyclic curves use `k = 8`; participant-day shifts use a random-effect smooth.
No observation weight is supplied.

For the one-part candidate, Tweedie power was estimated in the preliminary
no-AR fit and fixed at `r sprintf("%.4f", temporal_component$tweedie_power[temporal_component$candidate_id == "one_part_tweedie"])`
in the final fit. For all non-Gaussian candidates, `rho` acts on working
residuals as a GEE approximation. It is not an exact non-Gaussian likelihood
AR(1).

# Daily-family pilot

Four response-family problems were exercised without running the 15 × 3
production grid.

```{r}
#| label: tbl-daily-pilot-registry
#| tbl-cap: "Representative daily-family pilot registry."

daily_registry |>
  transmute(
    Pilot = .data$pilot_id,
    Metric = .data$metric_id,
    Predictor = .data$predictor_id,
    Family = .data$response_family,
    Transform = .data$response_transform,
    Purpose = .data$purpose
  ) |>
  compact_gt()
```

```{r}
#| label: tbl-daily-verdict
#| tbl-cap: "Daily-family pilot verdict."

daily_verdict |>
  transmute(
    Problem = .data$pilot_problem,
    Evidence = .data$evidence,
    Status = .data$status,
    `Required action` = .data$required_action
  ) |>
  status_gt(
    "Daily-family gate",
    source_note = paste0(
      "These are pilot acceptability decisions, not scientific association ",
      "decisions. The exact random-site/random-slope benchmark is retained ",
      "verbatim and labelled NON_ESTIMABLE."
    )
  )
```

The following values show that the production extraction path works. They are
**engineering pilot output**, not H06-daily results: no 15-slot BH vectors,
full diagnostic suite, placement comparison, deletion analysis, or sensitivity
set has been run.

```{r}
#| label: tbl-daily-effect-preview
#| tbl-cap: "Engineering-only effect extraction preview."

pilot_labels <- tibble::tribble(
  ~pilot_id, ~Metric, ~Contrast, ~Scale,
  "gaussian_offset", "Daily geometric mean melEDI",
  "Free day versus work day", "Ratio on melEDI + 0.1 lx scale",
  "tweedie_duration", "Time above 1,000 lx melEDI",
  "Active versus Sedentary", "Expected-duration ratio",
  "gaussian_identity", "Time below 10 lx melEDI before sleep",
  "Per one-hour longer previous sleep", "Difference (h)",
  "l10_one_part", "Darkest 10 h mean melEDI",
  "Free day versus work day", "Ratio on melEDI + 0.1 lx scale"
)

daily_preview |>
  left_join(pilot_labels, by = "pilot_id") |>
  transmute(
    Metric,
    Contrast,
    Scale,
    Estimate = .data$display_estimate,
    `Lower 95%` = .data$display_lower_95,
    `Upper 95%` = .data$display_upper_95,
    Role = .data$inferential_role
  ) |>
  compact_gt(
    source_note = "Intervals only verify extraction; they are not multiplicity-adjusted production intervals."
  ) |>
  gt::fmt_number(columns = c(Estimate, `Lower 95%`, `Upper 95%`), decimals = 3)
```

## Mandatory L10 zero-mass check

L10 contains 13.3% exact zeros in the representative frame. The one-part and
two-part free/work point contrasts agree in direction but not magnitude: the
equal-site mean ratio is `r sprintf("%.3f", l10_family$one_part_free_vs_work_ratio)`
under the one-part shifted-log fit and
`r sprintf("%.3f", l10_family$two_part_free_vs_work_ratio)` under the two-part
pilot. The positive-only model also retains true-date residual lag-1
correlation of -0.340. The current one-part L10 family is therefore not
accepted; the two-part positive component needs the prespecified daily AR
counterpart before production.

```{r}
#| label: tbl-l10-family
#| tbl-cap: "Bounded L10 one-part versus two-part point comparison."

l10_family |>
  transmute(
    `One-part work-day mean (lx)` =
      .data$`one_part_equal_site_mean_lx__Work day`,
    `One-part free-day mean (lx)` =
      .data$`one_part_equal_site_mean_lx__Free day`,
    `One-part ratio` = .data$one_part_free_vs_work_ratio,
    `Two-part work-day mean (lx)` =
      .data$`two_part_equal_site_mean_lx__Work day`,
    `Two-part free-day mean (lx)` =
      .data$`two_part_equal_site_mean_lx__Free day`,
    `Two-part ratio` = .data$two_part_free_vs_work_ratio,
    `Direction agrees` = .data$direction_agrees
  ) |>
  compact_gt(
    source_note = paste0(
      "Point comparison only. Random effects are excluded and submitted sites ",
      "receive equal weight. No uncertainty interval was simulated."
    )
  ) |>
  gt::fmt_number(
    columns = where(is.numeric),
    decimals = 3
  )
```

# Exploratory 30-minute GAMM family pilot

## Fit and numerical summary

```{r}
#| label: tbl-temporal-fit-summary
#| tbl-cap: "Near-eye 30-minute temporal pilot fit summary."

temporal_component |>
  transmute(
    Candidate = if_else(
      .data$candidate_id == "one_part_tweedie",
      "One-part Tweedie",
      "Two-part"
    ),
    Component = .data$component,
    Bins = .data$observations,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `rho` = .data$rho,
    `Fit time (s)` = .data$total_elapsed_seconds,
    Convergence = .data$convergence,
    `Hessian min` = .data$smoothing_hessian_minimum_eigenvalue,
    `Adjusted R²` = .data$adjusted_r_squared,
    `Deviance explained` = .data$deviance_explained,
    `Final residual lag-1` = .data$standardized_residual_lag1
  ) |>
  compact_gt(
    source_note = paste0(
      "Adjusted R² and deviance explained are in-sample fit summaries, not ",
      "effect tests or out-of-sample validation."
    )
  ) |>
  gt::fmt_integer(columns = c(Bins, `Participant-days`, Participants, Sites)) |>
  gt::fmt_number(
    columns = c(`rho`, `Adjusted R²`, `Deviance explained`,
      `Final residual lag-1`),
    decimals = 3
  ) |>
  gt::fmt_number(columns = `Fit time (s)`, decimals = 1) |>
  gt::fmt_scientific(columns = `Hessian min`, decimals = 2)
```

The one-part fit estimated `p = 1.7546`. Its fitted compound distribution
implies 41.1% zeros versus 30.9% observed. Across fitted-mean deciles, observed
to predicted zero-frequency ratios range from 0.048 to 1.181; that is a
structural calibration failure, not a cosmetic residual issue.

The two-part occurrence model reproduces the global zero fraction by design,
but its lowest predicted-positive-probability deciles remain miscalibrated.
The positive Gamma model underpredicts the highest fitted-mean decile
(observed/predicted ratio 1.287). These are explicit limitations, not evidence
about any predictor.

```{r}
#| label: tbl-calibration-summary
#| tbl-cap: "Temporal calibration ranges across ten fitted-value bins."

temporal_calibration |>
  summarise(
    `Minimum observed/predicted` = min(.data$observed_to_predicted),
    `Maximum observed/predicted` = max(.data$observed_to_predicted),
    `Minimum bin n` = min(.data$observations),
    `Maximum bin n` = max(.data$observations),
    .by = "component"
  ) |>
  transmute(
    Component = recode(
      .data$component,
      one_part_tweedie_zero_probability = "One-part zero probability",
      two_part_positive_occurrence_probability = "Two-part positive probability",
      two_part_positive_magnitude_mean = "Two-part positive magnitude"
    ),
    `Minimum observed/predicted`,
    `Maximum observed/predicted`,
    `Minimum bin n`,
    `Maximum bin n`
  ) |>
  compact_gt() |>
  gt::fmt_number(
    columns = c(`Minimum observed/predicted`, `Maximum observed/predicted`),
    decimals = 3
  ) |>
  gt::fmt_integer(columns = c(`Minimum bin n`, `Maximum bin n`))
```

## Basis, cyclicity, identifiability, and residual dependence

At `k = 12`, the one-part global clock smooth uses 9.94 of ten available
effective degrees of freedom and the sleep varying-coefficient term uses 9.97
of 11. The positive Gamma global smooth uses 9.75/10 and sleep uses 9.29/11.
Their deterministic `k` indices are below 0.90. A larger basis must therefore
be tested before any profile is interpreted.

```{r}
#| label: tbl-temporal-k
#| tbl-cap: "Deterministic basis-capacity screen; no permutation p-value was run."

temporal_k |>
  filter(
    .data$term %in% c(
      "s(time_hour)",
      "s(time_hour,work_free_day)",
      "s(time_hour,activity_status)",
      "s(time_hour):previous_sleep_duration_centered_h"
    )
  ) |>
  transmute(
    Candidate = if_else(
      .data$candidate_id == "one_part_tweedie",
      "One-part Tweedie",
      paste("Two-part", .data$component)
    ),
    Term = recode(
      .data$term,
      `s(time_hour)` = "Global time",
      `s(time_hour,work_free_day)` = "Work/free deviations",
      `s(time_hour,activity_status)` = "Activity deviations",
      `s(time_hour):previous_sleep_duration_centered_h` = "Sleep slope over time"
    ),
    `k-prime` = .data$k_prime,
    `Effective df` = .data$effective_df,
    `k-index` = .data$k_index
  ) |>
  compact_gt(
    source_note = paste0(
      "A fixed analytical seed selected the 5,000-row k-check subsample. ",
      "n.rep = 0, so no stochastic permutation p-value was computed."
    )
  ) |>
  gt::fmt_number(columns = c(`k-prime`, `Effective df`, `k-index`), decimals = 3)
```

All global, context, sleep, and site components close at midnight to numerical
precision. The maximum 24:00 versus 00:00 link difference is below
`r format(max(temporal_closure$maximum_absolute_endpoint_difference_link), scientific = TRUE, digits = 2)`.
The fitted-term multiple-R² fallback is at most
`r sprintf("%.3f", max(temporal_concurvity$fitted_term_multiple_r_squared, na.rm = TRUE))`.
This is reassuring but is not the full `mgcv::concurvity()` decomposition.

```{r}
#| label: tbl-temporal-closure
#| tbl-cap: "Cyclic closure and bounded fitted-term identifiability checks."

temporal_closure |>
  left_join(
    temporal_concurvity |>
      summarise(
        `Maximum fitted-term multiple R²` = max(
          .data$fitted_term_multiple_r_squared,
          na.rm = TRUE
        ),
        .by = c("candidate_id", "component")
      ),
    by = c("candidate_id", "component")
  ) |>
  transmute(
    Candidate = if_else(
      .data$candidate_id == "one_part_tweedie",
      "One-part Tweedie",
      "Two-part"
    ),
    Component = .data$component,
    `Endpoint pairs` = .data$endpoint_pairs,
    `Maximum endpoint difference` =
      .data$maximum_absolute_endpoint_difference_link,
    `Closure pass` = .data$cyclic_closure_pass,
    `Maximum fitted-term multiple R²`
  ) |>
  compact_gt(
    source_note = paste0(
      "Random participant and participant-day terms are excluded from the ",
      "population-component closure check. The multiple-R² diagnostic uses a ",
      "fixed 5,000-row sample and is explicitly a fallback, not full concurvity."
    )
  ) |>
  gt::fmt_integer(columns = `Endpoint pairs`) |>
  gt::fmt_scientific(columns = `Maximum endpoint difference`, decimals = 2) |>
  gt::fmt_number(columns = `Maximum fitted-term multiple R²`, decimals = 3)
```

The estimated working AR coefficients are 0.305 for one-part Tweedie, 0.261
for occurrence, and 0.308 for positive magnitude. Final standardized-residual
lag-1 correlations remain 0.315, 0.228, and 0.251. Site-specific maxima are
0.369, 0.307, and 0.304, respectively. The current correction is therefore
incomplete under the approved residual thresholds.

```{r}
#| label: tbl-site-acf
#| tbl-cap: "Final AR-standardized residual lag-1 correlation by submitted site."

temporal_site_acf |>
  mutate(
    Model = case_when(
      .data$candidate_id == "one_part_tweedie" ~ "One-part Tweedie",
      .data$component == "occurrence" ~ "Two-part occurrence",
      TRUE ~ "Two-part positive magnitude"
    )
  ) |>
  select(.data$site, .data$Model, .data$residual_lag1) |>
  pivot_wider(names_from = .data$Model, values_from = .data$residual_lag1) |>
  rename(Site = .data$site) |>
  compact_gt(
    source_note = paste0(
      "Each correlation uses adjacent residual pairs within verified AR ",
      "sequences only; no pair crosses a series boundary."
    )
  ) |>
  gt::fmt_number(columns = -Site, decimals = 3)
```

## Explicit diagnostic verdict

```{r}
#| label: tbl-temporal-verdict
#| tbl-cap: "Exploratory temporal diagnostic verdict."

temporal_verdict |>
  transmute(
    Candidate = .data$candidate,
    Domain = .data$domain,
    Evidence = .data$evidence,
    Status = .data$status,
    Consequence = .data$consequence,
    `Required action` = .data$required_action
  ) |>
  status_gt(
    "No temporal effect claim passes this pilot gate",
    source_note = paste0(
      "The one-part Tweedie family is rejected. The two-part architecture is ",
      "retained only for a repair pilot; current predictor smooths and ",
      "approximate p-values are not interpreted."
    )
  )
```

# Runtime and compute gate

```{r}
#| label: tbl-runtime
#| tbl-cap: "Measured and projected Stage 2 runtime."

runtime |>
  transmute(
    Component = .data$component,
    Scope = .data$pilot_scope,
    Units = .data$production_units,
    `Measured/projected time` = if_else(
      is.na(.data$observed_or_projected_seconds),
      "Not timed",
      if_else(
        .data$observed_or_projected_seconds < 60,
        sprintf("%.1f s", .data$observed_or_projected_seconds),
        sprintf("%.1f min", .data$observed_or_projected_seconds / 60)
      )
    ),
    Basis = .data$basis,
    Status = .data$approval_status
  ) |>
  compact_gt(
    source_note = paste0(
      "The six-frame temporal projection is fit-only and linear. It excludes ",
      "uncertainty simulation, deletion refits, rendering, and failed-fit ",
      "overhead. A 50/100-replicate pilot remains mandatory before any heavy ",
      "resampling."
    )
  )
```

The main H06 task was idle at its own approval gate during this run. The pilot
used one R thread and did not compete with an active heavy batch.

# Proposed bounded repair pilot

The diagnostics support the following next step, still near-eye only:

1. **Reject one-part Tweedie.** Do not carry its `p = 1.7546` fit into
   production.
2. **Retain the two-part architecture provisionally.** Keep separate binomial
   occurrence and Gamma positive-magnitude components.
3. **Increase temporal capacity to `k = 16`.** Apply it to the global,
   work/free, activity, sleep-slope, and site clock terms; retain participant
   cyclic curves at `k = 8` and participant-day random intercepts.
4. **Keep separate additive context deviations.** This preserves direct
   work/free and activity estimands. A four-level work/free × activity curve is
   a different interaction estimand and is not recommended unless the author
   explicitly prefers it.
5. **Add one H02-style diagnostic comparator.** Fit
   `log10(geometric-mean melEDI + 0.1)` with Gaussian identity and the same mean
   structure. This gives `bam()` its exact Gaussian AR(1) correction but changes
   the estimand to the conditional shifted-log mean; it cannot silently replace
   the raw-scale two-part target.
6. **Apply the existing residual rule.** If final pooled |lag-1| remains at or
   above 0.20 or any site remains at or above 0.30, retain the temporal analysis
   as descriptive context only and make no curve-wide inferential claim.
7. **Repair the daily family pilots in the same bounded run.** Fit the
   actual-date gap-aware daily AR counterpart for the identity-Gaussian
   pre-sleep duration and for the positive L10 component. Do not start the full
   15 × 3 grid yet.

At the observed pilot rate, the repair should be budgeted at approximately
20–30 minutes. It will use production code, no bootstrap/simulation, and stop
again with diagnostics and a revised runtime estimate.

# H06-D-G2P author decisions

::: {.callout-important title="Stop gate — explicit approval required"}
Please approve or amend these decisions before any further fit:

1. Reject the one-part Tweedie temporal family.
2. Authorize the near-eye-only two-part `k = 16` repair pilot.
3. Retain separate additive work/free and activity time deviations, with the
   cyclic one-hour sleep-duration slope, rather than a four-level
   work/free × activity interaction curve.
4. Authorize the H02-style shifted-log Gaussian comparator as a diagnostic
   estimand only.
5. Authorize the two bounded daily AR repairs for pre-sleep duration and L10
   positive magnitude.
6. Continue to withhold the full daily grid, chest/paired/gap temporal fits,
   deletion batches, and all resampling until the repair pilot passes and its
   runtime is approved.

No H06-daily or temporal association is accepted at this gate.
:::

# Reproducibility record

```{r}
#| label: tbl-software
#| tbl-cap: "Authoritative pilot software."

software |>
  transmute(Item = .data$item, Version = .data$version, Role = .data$role) |>
  compact_gt(
    source_note = paste0(
      "No package was installed or updated. Model objects, exact frames, ",
      "diagnostics, inputs, code hashes, and artifact hashes are stored only ",
      "under H06_daily-owned paths."
    )
  )
```

The output manifest verifies `r nrow(output_manifest)` frozen pilot artifacts.
No shared preparation, main-H06, manuscript, bibliography, Quarto
configuration, or `renv.lock` file was modified.
