---
title: "H06 daily metrics — corrected temporal pilot reconciliation"
subtitle: "H02-aligned arithmetic-mean melEDI GAMM with working AR(1) correction"
date: 2026-08-11
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("mgcv")), "1.9.4"),
  identical(as.character(packageVersion("melidosData")), "1.0.6")
)

read_h06d <- function(stage, file) {
  readr::read_csv(
    file.path(project_root, "artifacts", stage, "H06_daily", file),
    show_col_types = FALSE,
    na = ""
  )
}
sha256 <- function(path) {
  digest::digest(path, algo = "sha256", file = TRUE)
}
verify_manifest <- function(path, allowed_mismatch = character()) {
  manifest <- readr::read_csv(path, show_col_types = FALSE, na = "")
  observed <- vapply(
    file.path(project_root, manifest$relative_path),
    sha256,
    character(1)
  )
  checked <- manifest |>
    mutate(
      observed_sha256 = unname(observed),
      hash_match = .data$sha256 == .data$observed_sha256
    )
  mismatched_paths <- checked$relative_path[!checked$hash_match]
  stopifnot(setequal(mismatched_paths, allowed_mismatch))
  checked
}

manifest_names <- c(
  "H06_daily_temporal_h02_input_manifest.csv",
  "H06_daily_temporal_h02_code_manifest.csv",
  "H06_daily_temporal_h02_output_manifest.csv",
  "H06_daily_temporal_h02_basis_repair_input_manifest.csv",
  "H06_daily_temporal_h02_basis_repair_code_manifest.csv",
  "H06_daily_temporal_h02_basis_repair_output_manifest.csv",
  "H06_daily_temporal_h02_reconciliation_input_manifest.csv",
  "H06_daily_temporal_h02_reconciliation_code_manifest.csv",
  "H06_daily_temporal_h02_reconciliation_output_manifest.csv",
  "H06_daily_temporal_h02_production_gate_input_manifest.csv",
  "H06_daily_temporal_h02_production_gate_code_manifest.csv",
  "H06_daily_temporal_h02_production_gate_output_manifest.csv",
  "H06_daily_temporal_h02_upstream_repin_input_manifest.csv",
  "H06_daily_temporal_h02_upstream_repin_code_manifest.csv",
  "H06_daily_temporal_h02_upstream_repin_output_manifest.csv"
)
manifest_paths <- file.path(
  project_root,
  "artifacts/12_manifests/H06_daily",
  manifest_names
)
selected_input_manifest_path <- manifest_paths[
  manifest_names == "H06_daily_temporal_h02_input_manifest.csv"
]
selected_input_check <- verify_manifest(
  selected_input_manifest_path,
  allowed_mismatch = c(
    "artifacts/06_model_data/base/metrics_glasses_30_minute_context.rds",
    "artifacts/12_manifests/base_model_data_artifacts.csv",
    "artifacts/12_manifests/metric_artifacts.csv",
    "audit/decisions/preparation06_current_base_model_gate.md"
  )
)
qualified_input_mismatch <- selected_input_check |>
  filter(!.data$hash_match)
expected_qualified_mismatch <- tibble::tribble(
  ~relative_path, ~sha256, ~observed_sha256,
  "artifacts/06_model_data/base/metrics_glasses_30_minute_context.rds",
  "afa5a23308744ae495ef07a521c99e11bd7296aa855c5cb773f71f2b68eeb8e5",
  "0ad121f104af7ae885016e4dd82f62e94e94daabd01128ed8c76407bfbc54249",
  "artifacts/12_manifests/base_model_data_artifacts.csv",
  "142d1044dd7e6a566e45e3a5c56033b1e8251e100c0d7c7523564c824950a0e4",
  "6cfbfb18a2f6b1e31613e3fba803c3fbf64eb63e37155ae4fffd437a213b3ad0",
  "artifacts/12_manifests/metric_artifacts.csv",
  "6ec3185620d921e1f81b8464e850249613deb823988187d58794b3471303f4e8",
  "7bb2830c695e5600f0877acca500a187a040fe727a029bdb069f69099abe4f43",
  "audit/decisions/preparation06_current_base_model_gate.md",
  "4bc007db330b953a090985765b090d2a166515bbf056f4a56d3c44ea01426944",
  "789c1b2e0e52f1a22ed407096f0ebe1c239ee1c954197a880c19a9c32780626e"
)
stopifnot(
  nrow(qualified_input_mismatch) == 4L,
  nrow(dplyr::anti_join(
    qualified_input_mismatch |>
      select(.data$relative_path, .data$sha256, .data$observed_sha256),
    expected_qualified_mismatch,
    by = c("relative_path", "sha256", "observed_sha256")
  )) == 0L,
  nrow(dplyr::anti_join(
    expected_qualified_mismatch,
    qualified_input_mismatch |>
      select(.data$relative_path, .data$sha256, .data$observed_sha256),
    by = c("relative_path", "sha256", "observed_sha256")
  )) == 0L
)
verified_manifests <- lapply(
  setdiff(manifest_paths, selected_input_manifest_path),
  verify_manifest
)
stopifnot(all(vapply(
  verified_manifests,
  function(manifest) all(manifest$hash_match),
  logical(1)
)))

input_manifest <- selected_input_check |>
  select(-dplyr::all_of(c("observed_sha256", "hash_match")))
upstream_repin_comparison <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_upstream_repin_frame_comparison.csv"
)
upstream_repin_drift <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_upstream_repin_drift_registry.csv"
)
stopifnot(
  nrow(upstream_repin_comparison) == 6L,
  all(upstream_repin_comparison$exact_row_value_and_attribute_identity),
  all(!upstream_repin_comparison$mder_read_or_used),
  all(!upstream_repin_comparison$frame_written),
  all(!upstream_repin_comparison$model_refitted),
  all(upstream_repin_comparison$qualification == "PASS"),
  nrow(upstream_repin_drift) == 5L,
  all(upstream_repin_drift$current_hash_verified),
  all(upstream_repin_drift$qualification == "PASS_NO_REFIT")
)
software <- read_h06d(
  "12_manifests",
  "H06_daily_temporal_h02_software_manifest.csv"
)
formula_registry <- read_h06d(
  "06_model_data",
  "H06_daily_temporal_h02_formula_registry.csv"
)
specification <- read_h06d(
  "06_model_data",
  "H06_daily_temporal_h02_specification.csv"
)
support <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_support_overall.csv"
)
context <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_context_cells.csv"
)
sleep_support <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_sleep_support_by_site.csv"
)
boundaries <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_ar_boundaries.csv"
)
component <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_component_diagnostics.csv"
)
smooth_registry <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_smooth_registry.csv"
)
basis <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_basis_repair_comparison.csv"
)
basis_models <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_basis_repair_model_diagnostics.csv"
)
reconciled <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_reconciled_verdict.csv"
)
gate_contract <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_production_gate_contract.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_reconciled_runtime.csv"
)

stopifnot(
  nrow(formula_registry) == 1L,
  nrow(component) == 1L,
  nrow(basis) == 1L,
  all(smooth_registry$xt_is_null),
  identical(
    basis$basis_repair_status,
    "K12_RETAINED_AFTER_K16_SENSITIVITY"
  ),
  identical(
    reconciled$assessment[reconciled$domain == "Overall corrected pilot gate"],
    "ACCEPTABLE_FOR_EFFECT_EXTRACTION_AFTER_APPROVAL"
  )
)

compact_gt <- function(data, title = NULL, subtitle = NULL, source_note = NULL) {
  output <- data |>
    gt::gt() |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.font.size = gt::px(12),
      table.width = gt::pct(100),
      data_row.padding = gt::px(3),
      heading.align = "left",
      source_notes.font.size = gt::px(10),
      column_labels.font.weight = "bold"
    )
  if (!is.null(title)) {
    output <- output |>
      gt::tab_header(
        title = gt::md(title),
        subtitle = if (is.null(subtitle)) NULL else gt::md(subtitle)
      )
  }
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

status_gt <- function(data, title = NULL, subtitle = NULL, source_note = NULL) {
  output <- compact_gt(data, title, subtitle, source_note)
  status_column <- intersect(c("Assessment", "Status"), names(data))
  if (length(status_column) == 1L) {
    output <- output |>
      gt::data_color(
        columns = dplyr::all_of(status_column),
        fn = function(value) {
          dplyr::case_when(
            grepl(
              "acceptable|closed|completed|retained",
              value,
              ignore.case = TRUE
            ) ~ "#D9F2E6",
            grepl(
              "pending|withheld|hold|not authorized",
              value,
              ignore.case = TRUE
            ) ~ "#FFF1CC",
            TRUE ~ "#EEEEEE"
          )
        }
      )
  }
  output
}
```

# Reconciliation verdict

::: {.callout-important title="H06-D-G2P-H02 — production approved with pointwise uncertainty"}
The corrected exploratory temporal pilot is diagnostically acceptable after
one bounded basis-capacity sensitivity. The selected model uses the 30-minute
**arithmetic mean** melEDI response and the H02 shifted-log Gaussian
architecture. Only its global time smooth is cyclic.

The author approved the four context-association functions and six declared
placement/scenario frames. Uncertainty is restricted to pointwise 95%
confidence intervals. Heavy deletion or resampling remains separately
withheld, and no day-type, activity, or sleep association had been extracted
when this pilot verdict was made.
:::

The temporal sections of `02_stage2_pilot.qmd` and its HTML are superseded by
this reconciliation. Their zero-aware geometric-mean, Tweedie, two-part, and
cyclic-`sz`/`fs` proposals remain frozen engineering history only. The
participant-day pilot in that report is unchanged. MDER remains on its
separate upstream hold and is absent here.

::: {.callout-warning title="Qualified upstream manifest change; no local repinning"}
The 11 August shared Preparation 06 repin changed four historical near-eye
pilot pins: the direct 30-minute object, base-model manifest, metric manifest,
and Preparation 06 gate. Its production counterpart also changed the chest
30-minute object's file identity. The shared gate records zero changed cells
across the non-MDER metric fields.

A no-refit R 4.6.1 qualification reconstructed all six declared temporal
frames in memory from the current shared inputs. Every row, value, factor
attribute, and frame attribute exactly matched the frozen H06_daily production
frame. The qualification did not read or use MDER, write a model frame, or
refit a model. The existing temporal frames and fits therefore remain frozen;
this is not a local repin of MDER. MDER-specific work still waits for targeted
H06_daily instructions.
:::

```{r}
#| label: tbl-upstream-repin-qualification
#| tbl-cap: "No-refit qualification of the shared Preparation 06 repin against the six frozen temporal frames."

upstream_repin_comparison |>
  transmute(
    Frame = .data$run_id,
    Rows = .data$frozen_rows,
    `Exact current-to-frozen identity` = if_else(
      .data$exact_row_value_and_attribute_identity,
      "Pass",
      "Fail"
    ),
    `MDER used` = if_else(.data$mder_read_or_used, "Yes", "No"),
    `Model refitted` = if_else(.data$model_refitted, "Yes", "No")
  ) |>
  compact_gt(
    source_note = paste(
      "Exact identity includes rows, values, factor attributes, and frame",
      "attributes; the comparison used R 4.6.1."
    )
  )
```

```{r}
#| label: tbl-reconciled-verdict
#| tbl-cap: "Final diagnostic assessment of the corrected near-eye temporal pilot."

reconciled |>
  transmute(
    Domain = .data$domain,
    Evidence = .data$evidence,
    Assessment = recode(
      .data$assessment,
      ACCEPTABLE = "Acceptable",
      ACCEPTABLE_WITH_NUMERICAL_TOLERANCE =
        "Acceptable; numerical tolerance disclosed",
      ACCEPTABLE_AFTER_BOUNDED_SENSITIVITY =
        "Acceptable after bounded sensitivity",
      ACCEPTABLE_FOR_EXPLORATORY_MODEL =
        "Acceptable for exploratory model",
      ACCEPTABLE_FOR_EFFECT_EXTRACTION_AFTER_APPROVAL =
        "Acceptable for association extraction after approval"
    ),
    Implication = .data$implication
  ) |>
  status_gt(
    source_note = paste0(
      "‘Acceptable’ means adequate for the bounded exploratory claim under ",
      "the stated rules; it does not mean the model is true. No association ",
      "estimate or predictor p-value was inspected for this verdict."
    )
  )
```

# What is now being estimated

Let $Y$ be an admissible bin's 30-minute arithmetic mean melEDI in lx. The
model response is

$$
Z = \log_{10}(Y + 0.1\ \mathrm{lx}).
$$

The proposed exploratory estimand is the conditional expected transformed
profile

$$
\begin{aligned}
E[Z \mid\;& \text{local time, day type, daily activity,}\\
&\text{within- and between-person sleep duration,}\\
&\text{site, participant, participant-day}].
\end{aligned}
$$

For fitted linear predictor $\eta$, the reader-facing inverse transform is
$\max(0, 10^\eta - 0.1)$ lx. This is the back-transform of a mean log response:
it is geometric-mean-like for $Y + 0.1$ and, under the fitted Gaussian log
model, conditional-median-like. It is **not** the raw-scale arithmetic
expectation $E[Y]$.

```{r}
#| label: tbl-estimand-separation
#| tbl-cap: "The three H06 estimands are not numerically interchangeable."

tibble::tribble(
  ~Analysis, ~Row, ~Question, ~Weighting, ~Relationship,
  "Preregistration-aligned H06_daily",
  "One metric-specific participant-day",
  "How a day-level context is associated with a daily light-exposure metric",
  "Each admissible participant-day contributes one row per metric",
  "Primary scientific variant; not yet in production",
  "Approved hourly H06",
  "One supported hour",
  "How contexts are associated with repeated hourly melEDI",
  "Days contribute according to supported hourly rows",
  "Separate approved analysis; no estimate transfers",
  "Exploratory temporal GAMM",
  "One supported 30-minute bin",
  "How context-associated transformed profiles vary over local clock time",
  "Days contribute according to supported half-hour bins",
  "Separate exploratory amendment outside the daily BH families"
) |>
  compact_gt()
```

Day type is a Work day/Free day label, and activity is the day-level
Sedentary/Active label—not a contemporaneous half-hour state. Previous-night
sleep duration is separated into the participant's mean across equally
weighted fitted days, centred on an equally participant-weighted grand mean,
and the day's deviation from that participant mean. Thus the sleep functions
distinguish between-participant from within-participant associations.

All wording remains observational: these are associations with recorded
contexts, not causal effects or health outcomes.

# Exact H02 inheritance and H06 extension

The authoritative H02 specification and stored near-eye fit were pinned before
the pilot. The model did not copy H02's numerical $\rho$; it inherited the
algorithm and estimated $\rho$ anew on the complete H06_daily frame.

```{r}
#| label: tbl-h02-pins
#| tbl-cap: "Frozen H02 sources used for the structural cross-check."

input_manifest |>
  filter(.data$input_id %in% c(
    "h02_selected_specification",
    "h02_selected_near_eye_fit"
  )) |>
  transmute(
    Source = recode(
      .data$input_id,
      h02_selected_specification = "Selected temporal specification",
      h02_selected_near_eye_fit = "Stored selected near-eye fit"
    ),
    Path = .data$relative_path,
    `SHA-256` = .data$sha256
  ) |>
  compact_gt()
```

The exact evaluated corrected formula is:

```{r}
#| label: corrected-formula
#| echo: true

selected_formula <- stats::as.formula(formula_registry$formula[[1L]])
selected_formula
```

Only `s(time_hour, bs = "cc", k = 12)` is cyclic. The work/free, activity,
site, participant, and two sleep terms have no `xt` argument and use the
documented default thin-plate time marginals. This is the exact correction
requested after the H02 cross-check.

```{r}
#| label: tbl-smooth-registry
#| tbl-cap: "Fitted smooth bases and effective dimensions in the selected pilot."

smooth_registry |>
  transmute(
    Term = .data$term,
    `Smooth class` = .data$smooth_class,
    `Time marginal` = .data$declared_time_margin,
    `xt is NULL` = if_else(.data$xt_is_null, "Yes", "No"),
    `Coefficient columns` = .data$coefficient_columns,
    EDF = .data$effective_df
  ) |>
  compact_gt(
    source_note = paste0(
      "EDF is shown as a capacity diagnostic, not as evidence of a context ",
      "association. The participant factor smooth retains H02's k = 10."
    )
  ) |>
  gt::fmt_number(columns = EDF, decimals = 2)
```

The context smooths are additive and common across sites. The site smooth
allows each submitted site its own overall clock profile, but the selected
formula does **not** estimate context-by-site heterogeneity. Setting site,
participant, and participant-day terms to zero yields the sum-to-zero,
equal-site-centred display curve; it is not an observation-weighted site
average.

# Exact fitted frame and AR(1) construction

```{r}
#| label: tbl-support
#| tbl-cap: "Exact complete-core near-eye sample used by both corrected pilot fits."

support |>
  transmute(
    Placement = .data$placement,
    `30-minute bins` = .data$observations_30_minute,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Exact-zero bins` = .data$exact_zero_bins,
    `Exact-zero fraction` = .data$exact_zero_fraction,
    `AR sequences` = .data$ar_sequences,
    `Sleep grand mean (h)` = .data$sleep_grand_mean_h
  ) |>
  compact_gt() |>
  gt::fmt_percent(columns = `Exact-zero fraction`, decimals = 1) |>
  gt::fmt_number(columns = `Sleep grand mean (h)`, decimals = 2) |>
  gt::fmt_integer(columns = c(
    `30-minute bins`, `Participant-days`, Participants, Sites,
    `Exact-zero bins`, `AR sequences`
  ))
```

```{r}
#| label: tbl-context-cells
#| tbl-cap: "Joint day-type and daily-activity support in the fitted frame."

context |>
  transmute(
    `Day type` = .data$work_free_day,
    `Daily activity` = .data$activity_status,
    `30-minute bins` = .data$observations_30_minute,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites
  ) |>
  compact_gt() |>
  gt::fmt_integer(columns = where(is.numeric))
```

All four day-type-by-activity cells occur at all nine sites; the smallest has
91 participant-days. Of 137 participants, 135 have at least two fitted days
and within-participant sleep-duration variation. Every site has both within-
and between-participant sleep-duration variation in this frame.

Rows were ordered by site, participant, local date, true UTC start, and clock
bin. `AR.start` resets at the first row of each participant-day, around any
non-one-to-one wall-clock outcome, when elapsed-time provenance is missing, or
after any physical discontinuity. It is false only for exactly contiguous
true-time outcomes.

```{r}
#| label: tbl-ar-boundaries
#| tbl-cap: "Reasons for starting a new working-AR sequence."

boundaries |>
  transmute(
    Reason = recode(
      .data$ar_start_reason,
      participant_day_start = "Participant-day start",
      elapsed_discontinuity = "Elapsed-time gap or unsupported interval",
      non_one_to_one_wall_outcome = "Non-one-to-one wall-clock outcome",
      after_non_one_to_one_wall_outcome = "After non-one-to-one outcome",
      continuous = "Physically contiguous"
    ),
    Observations = .data$observations
  ) |>
  compact_gt() |>
  gt::fmt_integer(columns = Observations)
```

The exact full formula was first fitted with $\rho=0$. Response residuals from
adjacent rows within those sequences gave $\rho=0.579$, after which the
identical formula was refitted using that fixed value. The final standardized
residual lag-1 correlation was 0.090, and the largest absolute site-specific
value was 0.159.

# Why `k = 12` is retained

The initial automatic screen flagged work/free because its EDF used 94.6% of
the available basis dimension. `mgcv::k.check()` does not return a `k`-index
for this factor smooth, so EDF proximity alone could not establish inadequate
capacity. A single diagnostic model therefore increased only the work/free
smooth from `k = 12` to `k = 16`, retained the exact rows and fixed $\rho$, and
compared the Free-day-minus-Work-day curve on the 48 half-hour midpoints.

```{r}
#| label: tbl-basis-sensitivity
#| tbl-cap: "Bounded same-row work/free basis-capacity sensitivity."

basis |>
  transmute(
    `Same rows` = if_else(.data$exact_same_rows, "Yes", "No"),
    `Fixed rho` = .data$common_fixed_rho,
    `Selected k-prime / EDF` = sprintf(
      "%.0f / %.2f",
      .data$selected_work_k_prime,
      .data$selected_work_edf
    ),
    `Diagnostic k-prime / EDF` = sprintf(
      "%.0f / %.2f",
      .data$diagnostic_work_k_prime,
      .data$diagnostic_work_edf
    ),
    `Max |curve difference|` =
      .data$maximum_absolute_work_contrast_difference_link,
    `Curve-difference RMSE` = .data$work_contrast_difference_rmse_link,
    `Curve correlation` = .data$work_contrast_curve_correlation,
    `Global k-index` = .data$selected_global_k_index,
    `Global k-check p` = .data$selected_global_k_p_value,
    Status = recode(
      .data$basis_repair_status,
      K12_RETAINED_AFTER_K16_SENSITIVITY =
        "k = 12 retained after k = 16 sensitivity"
    )
  ) |>
  status_gt(
    source_note = paste0(
      "Prespecified stability limits were 0.05 maximum absolute link ",
      "difference and 0.02 curve RMSE. The p-value is a fixed-seed, ",
      "400-permutation basis diagnostic—not a context-association test."
    )
  ) |>
  gt::fmt_number(
    columns = c(
      `Fixed rho`, `Max |curve difference|`, `Curve-difference RMSE`,
      `Curve correlation`, `Global k-index`, `Global k-check p`
    ),
    decimals = 3
  )
```

The maximum link-scale difference was 0.035, the difference RMSE was 0.016,
and the curves correlated at 0.998. The diagnostic model's conditional AIC was
2.97 lower, but AIC was not a prespecified selection gate. Because the direct
curve-stability rules passed and the global `k` check was adequate, the
H02-comparable `k = 12` model is retained. The `k = 16` fit is not a production
model and generated no scientific result.

Both fits converged without model warnings. The selected model's smoothing
Hessian had a minimum relative eigenvalue of $-1.95 \times 10^{-15}$, a tiny
negative value at numerical tolerance; this is disclosed rather than silently
called positive definite. Residual diagnostics are acceptable for this
exploratory transformed-response architecture, with visible scale-pattern
limitations retained in any later report.

# Approved production estimands

The reader-facing production analysis extracts four distinct common
context-association functions:

1. Free day minus Work day over local clock time;
2. Active minus Sedentary daily status over local clock time;
3. the association per one-hour greater previous-night sleep duration within
   the same participant; and
4. the association per one-hour greater participant-mean sleep duration
   between participants.

The factor terms use sum-to-zero coding, but those directional contrasts fix
their reader-facing interpretation. Because the model is additive, the
work/free contrast does not depend on which activity level is used for its
construction, and vice versa. Population displays would marginalize sites and
the other binary context equally, set participant and participant-day terms to
zero, and evaluate sleep at its centred zero unless a sleep contrast is being
shown.

On the transformed scale, categorical differences and one-hour sleep slopes
can be exponentiated as ratios for $Y + 0.1$. Parallel inverse-transformed
profiles may be shown in lx, but neither representation is a raw-scale mean
association. The author approved pointwise 95% confidence intervals only.
Four whole-function tests form one explicitly exploratory BH family, separate
from every preregistration-aligned daily family. No coefficient enters that
family merely because it was estimated.

The declared production frames are near-eye and chest all available,
paired/common near-eye and chest, and gap-timing-unaware near-eye and chest.
Each frame must reconstruct its own sleep decomposition, re-estimate its own
$\rho$, retain exact samples, and repeat the full diagnostic rules. Chest is
complementary; it is not a second discovery screen.

# Runtime and compute boundary

```{r}
#| label: tbl-runtime
#| tbl-cap: "Measured pilot time and bounded production projection."

runtime |>
  transmute(
    Scope = .data$scope,
    Fits = .data$fits,
    Time = if_else(
      is.na(.data$minutes),
      "Not timed",
      if_else(
        .data$minutes < 1,
        sprintf("%.1f s", .data$seconds),
        sprintf("%.1f min", .data$minutes)
      )
    ),
    Basis = case_when(
      .data$scope == "Prediction, simultaneous-interval, or deletion work" ~
        "Pointwise intervals are model-based; heavy deletion or resampling remains separately gated",
      TRUE ~ .data$basis
    ),
    Status = case_when(
      .data$scope == "Corrected near-eye pilot plus basis sensitivity" ~
        "Completed",
      .data$scope == "Six-frame base-model batch" ~ "Approved",
      TRUE ~ "Heavy work withheld"
    )
  ) |>
  status_gt(
    source_note = paste0(
      "The six-frame estimate is linear and fit-only. It excludes prediction, ",
      "pointwise-interval extraction, deletion refits, rendering, failed-",
      "fit overhead, and any required repair."
    )
  )
```

The corrected near-eye pilot and basis sensitivity together took 3.4 minutes
on one R thread. The approved six-frame base-model projection is 10.1 minutes
for 12 fits. Pointwise intervals use model-based covariance rather than a
resampling batch. Before heavy resampling or a site-deletion batch, production
code must first run the required bounded representative or 50/100-resample
pilot and return for a separate compute decision.

# H06-D-G2P-H02 decision record

```{r}
#| label: tbl-gate-contract
#| tbl-cap: "Approved and withheld decisions after the corrected temporal pilot gate."

gate_contract |>
  transmute(
    ID = .data$decision_id,
    Decision = .data$decision,
    Contract = .data$approved_contract,
    Status = recode(
      .data$status,
      CLOSED_BY_AUTHOR_CONFIRMATION = "Closed by author confirmation",
      APPROVED = "Approved",
      APPROVED_POINTWISE_ONLY = "Approved: pointwise intervals only",
      WITHHELD_PENDING_PILOT = "Withheld pending pilot",
      UPSTREAM_HOLD = "Upstream hold"
    )
  ) |>
  status_gt()
```

::: {.callout-important title="Production transition recorded"}
H06-D-G2P-H02 items 02–05 are approved. Production uses pointwise 95%
confidence intervals only. Item 06 remains withheld until its own runtime
pilot, and MDER remains held under item 07.

The pilot gate itself inspected no context-association estimate or predictor
p-value; scientific interpretation begins only after the approved production
fits and diagnostics are complete.
:::

# Reproducibility and scope record

```{r}
#| label: tbl-software
#| tbl-cap: "Authoritative software for the selected corrected pilot."

software |>
  transmute(Item = .data$item, Version = .data$version, Role = .data$role) |>
  compact_gt(
    source_note = paste0(
      "All direct selected-pilot inputs and all code/output, basis-sensitivity, ",
      "reconciliation, and production-gate manifests were verified during ",
      "this render. The one ",
      "qualified shared-manifest change is disclosed above. No package was ",
      "installed or updated."
    )
  )
```

The corrected pilot used current prepared metrics and context artifacts
unchanged. It did not fetch or rebuild data. It did not fit or summarize MDER,
inspect predictor-association estimates, run chest/paired/gap models, perform
deletion or uncertainty resampling, edit shared preparation, touch main-H06
outputs, modify the manuscript, or request site integration.

No figure is produced at this gate, so no figure source-data CSV or alt text is
applicable. Any later durable figure must pair its source-data CSV with
publication-scale clipping, wrapping, panel-balance, final-size legibility,
and alt-text QA.
