---
title: "H06 daily metrics — exploratory temporal production"
subtitle: "H02-aligned 30-minute arithmetic-mean melEDI GAMM with working AR(1)"
date: 2026-08-11
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("mgcv")), "1.9.4"),
  identical(as.character(packageVersion("melidosData")), "1.0.6")
)

read_h06d <- function(stage, file) {
  readr::read_csv(
    file.path(project_root, "artifacts", stage, "H06_daily", file),
    show_col_types = FALSE,
    na = ""
  )
}
sha256 <- function(path) {
  digest::digest(path, algo = "sha256", file = TRUE)
}
verify_manifest <- function(path) {
  manifest <- readr::read_csv(path, show_col_types = FALSE, na = "")
  observed <- vapply(
    file.path(project_root, manifest$relative_path),
    sha256,
    character(1)
  )
  stopifnot(identical(unname(observed), manifest$sha256))
  manifest
}
p_display <- function(value) {
  ifelse(
    is.na(value),
    "—",
    ifelse(value < 0.001, "<0.001", sprintf("%.3f", value))
  )
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100)
    )
}

manifest_names <- c(
  "H06_daily_temporal_h02_production_gate_output_manifest.csv",
  "H06_daily_temporal_h02_production_frame_output_manifest.csv",
  "H06_daily_temporal_h02_production_model_output_manifest.csv",
  "H06_daily_temporal_h02_production_diagnostic_output_manifest.csv",
  "H06_daily_temporal_h02_production_inference_output_manifest.csv",
  "H06_daily_temporal_h02_figure_output_manifest.csv"
)
verified_manifests <- lapply(
  file.path(
    project_root,
    "artifacts/12_manifests/H06_daily",
    manifest_names
  ),
  verify_manifest
)

run_registry <- read_h06d(
  "06_model_data",
  "H06_daily_temporal_h02_production_run_registry.csv"
)
specification <- read_h06d(
  "06_model_data",
  "H06_daily_temporal_h02_specification.csv"
)
formula_registry <- read_h06d(
  "06_model_data",
  "H06_daily_temporal_h02_formula_registry.csv"
)
support <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_production_support_overall.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_production_fit_runtime.csv"
)
diagnostic_verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_production_diagnostic_verdict.csv"
)
tests <- read_h06d(
  "09_tables",
  "H06_daily_temporal_h02_production_whole_function_tests.csv"
)
function_summary <- read_h06d(
  "09_tables",
  "H06_daily_temporal_h02_production_context_function_summary.csv"
)
not_extracted <- read_h06d(
  "09_tables",
  "H06_daily_temporal_h02_production_not_extracted_frames.csv"
)
functions <- read_h06d(
  "11_source_data",
  "H06_daily_temporal_h02_production_context_functions.csv"
)
profiles <- read_h06d(
  "11_source_data",
  "H06_daily_temporal_h02_production_inverse_profiles.csv"
)
figure_qa <- read_h06d(
  "08_diagnostics",
  "H06_daily_temporal_h02_figure_readability_qa.csv"
)
figure_alt <- read_h06d(
  "11_source_data",
  "H06_daily_temporal_h02_figure_alt_text.csv"
)

overall <- diagnostic_verdict |>
  dplyr::filter(.data$domain == "Overall base-model diagnostic gate")
primary_tests <- tests |>
  dplyr::filter(.data$run_id == "primary__near_eye__all_available")
gap_tests <- tests |>
  dplyr::filter(
    .data$run_id == "gap_timing_unaware__near_eye__all_available"
  )

stopifnot(
  nrow(run_registry) == 6L,
  nrow(support) == 6L,
  nrow(formula_registry) == 1L,
  nrow(overall) == 6L,
  sum(overall$status == "ACCEPTABLE") == 5L,
  sum(overall$status == "NOT_ACCEPTABLE") == 1L,
  nrow(not_extracted) == 1L,
  not_extracted$run_id == "primary__chest__paired_common",
  nrow(functions) == 960L,
  nrow(profiles) == 1920L,
  identical(unique(functions$interval_type),
    "pointwise 95% confidence interval"
  ),
  identical(unique(profiles$interval_type),
    "pointwise 95% confidence interval"
  ),
  !any(grepl("simultaneous", functions$interval_type, ignore.case = TRUE)),
  !any(grepl("simultaneous", profiles$interval_type, ignore.case = TRUE)),
  nrow(primary_tests) == 4L,
  nrow(gap_tests) == 4L,
  all(primary_tests$bh_decision),
  all(gap_tests$bh_decision),
  sum(tests$raw_p_value_unclamped < 0, na.rm = TRUE) == 1L,
  all(figure_qa$verdict == "PASS"),
  nrow(figure_alt) == 3L
)

primary_summary <- function_summary |>
  dplyr::filter(.data$run_id == "primary__near_eye__all_available") |>
  dplyr::select(
    "estimand_order", "estimand_id", "estimand_label",
    "minimum_shifted_response_ratio", "maximum_shifted_response_ratio",
    "time_of_maximum_absolute_association_h",
    "ratio_at_maximum_absolute_association",
    "pointwise_bins_excluding_null"
  )
strongest_rows <- functions |>
  dplyr::filter(.data$run_id == "primary__near_eye__all_available") |>
  dplyr::inner_join(
    primary_summary |>
      dplyr::select(
        "estimand_id", "time_of_maximum_absolute_association_h"
      ),
    by = "estimand_id",
    relationship = "many-to-one"
  ) |>
  dplyr::filter(
    .data$time_hour == .data$time_of_maximum_absolute_association_h
  )
stopifnot(nrow(strongest_rows) == 4L)
```

::: {.callout-note appearance="simple" title="Answer in brief"}
The exploratory H02-aligned GAMM found time-varying associations between the
30-minute arithmetic-mean melEDI profile and each recorded day-level context.
All four near-eye whole-function tests remained below 0.001 after
Benjamini--Hochberg adjustment across the four prespecified tests. These are
associations on the fitted transformed scale, not daily-metric results, raw
arithmetic-mean effects, causal effects, or health outcomes.

Five of six prespecified frames passed the production diagnostics. The paired
chest frame failed the prespecified residual scale-pattern rule and was not
used for association extraction. Every displayed band is a **pointwise 95%
confidence interval**. No simultaneous interval was constructed or described.
:::

# Scope and estimand

This is a separate exploratory amendment to the preregistration-aligned
participant-day analysis. It uses one supported 30-minute bin per row and asks
how the transformed local-time profile varies with Work/Free day, the daily
Sedentary/Active label, and previous-night sleep duration. It is also separate
from the approved hourly H06 analysis.

Let $Y$ be a supported bin's 30-minute arithmetic mean melEDI in lx. The model
response is

$$
Z = \log_{10}\!\left(Y + 0.1\ \mathrm{lx}\right).
$$

The fitted target is $E[Z \mid \text{local time, context, site, participant,
participant-day}]$. The reader-facing inverse transform
$\max(0, 10^{\eta} - 0.1)$ lx is geometric-mean-like for $Y+0.1$ and
conditional-median-like under the Gaussian log model. It is **not** the
raw-scale arithmetic expectation $E[Y]$.

# Exact H02-aligned model

```{r}
#| label: exact-temporal-formula
#| echo: true

response ~
  s(time_hour, bs = "cc", k = 12) +
  s(time_hour, work_free_day, bs = "sz", k = 12) +
  s(time_hour, activity_status, bs = "sz", k = 12) +
  s(time_hour, by = sleep_between_h, k = 12) +
  s(time_hour, by = sleep_within_h, k = 12) +
  s(time_hour, site, bs = "sz", k = 12) +
  s(time_hour, participant, bs = "fs", k = 10) +
  s(participant_day, bs = "re")
```

Only the global clock smooth is cyclic. Every fitted `sz`, `fs`, and sleep
smooth has `xt = NULL`; their time marginals use mgcv's default thin-plate
basis. The Gaussian identity-link model used fREML, `discrete = TRUE`, one R
thread, 0/24 global knots, and H02's audited true-time sequence boundaries.

The working AR(1) coefficient was re-estimated for each H06_daily frame. The
identical full formula was first fitted with $\rho=0$; response residuals from
true-time-contiguous pairs gave the lag-1 estimate, clamped to
$[-0.95,0.95]$, and the formula was refitted with that fixed value. No H02
numerical $\rho$ was copied.

Previous-night sleep was decomposed within each exact fitted frame into the
participant mean across equally weighted days and the day's deviation from
that mean. The participant means were centred on an equally
participant-weighted grand mean. Thus the two sleep functions distinguish
between-participant from within-participant associations.

# Exact fitted samples

```{r}
#| label: tbl-temporal-samples
#| tbl-cap: "Exact samples and frame-specific working AR(1) coefficients."

sample_table <- support |>
  dplyr::left_join(
    runtime |>
      dplyr::select("run_id", "rho"),
    by = "run_id",
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    overall |>
      dplyr::select("run_id", "status"),
    by = "run_id",
    relationship = "one-to-one"
  ) |>
  dplyr::transmute(
    Dataset = .data$data_scenario_label,
    Placement = .data$placement_label,
    Sample = dplyr::recode(
      .data$sample_scenario,
      all_available = "All available",
      paired_common = "Paired/common"
    ),
    `30-min bins` = .data$observations_30_minute,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Exact zeros` = .data$exact_zero_bins,
    `Fixed rho` = .data$rho,
    Verdict = dplyr::recode(
      .data$status,
      ACCEPTABLE = "Acceptable",
      NOT_ACCEPTABLE = "Not acceptable"
    )
  )

sample_table |>
  gt::gt() |>
  gt::fmt_integer(
    columns = c(
      `30-min bins`, `Participant-days`, Participants, Sites, `Exact zeros`
    )
  ) |>
  gt::fmt_number(columns = `Fixed rho`, decimals = 3) |>
  compact_gt()
```

The six-frame batch required 12 BAM fits and `r sprintf("%.1f",
unique(runtime$full_batch_wall_seconds) / 60)` minutes on one R thread. The
primary near-eye frame reused its already verified selected pilot fit.

# Near-eye association functions

```{r}
#| label: fig-temporal-primary-profiles
#| fig-cap: "Equal-site-centred near-eye melEDI profiles by recorded context. Bands are pointwise 95% confidence intervals."
#| fig-alt: "Four-panel line plot of equal-site-centred near-eye melEDI over local clock time on a zero-preserving symlog display. Free-day exposure is higher around midnight but markedly lower than work-day exposure from morning into evening. Active days are above Sedentary days mainly during daylight hours. One hour greater previous-night sleep within or between participants corresponds mainly to lower fitted morning profiles. Shaded ribbons are pointwise 95% confidence intervals."
#| out-width: 100%

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H06_daily/",
  "H06_daily_temporal_h02_primary_profiles.png"
))
```

The melEDI display axis uses
`LightLogR::symlog_trans(base = 10, thr = 1, scale = 1)` only for
reader-facing visualization. Source values and model responses remain
untransformed beyond the explicitly fitted $\log_{10}(Y+0.1)$ response.

```{r}
#| label: fig-temporal-primary-functions
#| fig-cap: "Near-eye context-association functions as ratios for the shifted response Y + 0.1 lx. Bands are pointwise 95% confidence intervals."
#| fig-alt: "Four-panel ratio plot over local clock time. The fitted Free-day to Work-day shifted-response ratio falls from about 1.8 after midnight to 0.2 at 08:45. The Active-to-Sedentary ratio rises to about 1.6 in the afternoon. A one-hour within-participant sleep increase has ratios mostly between 0.76 and 1.05; the between-participant sleep ratio ranges from about 0.61 to 0.97. Ribbons are pointwise 95% confidence intervals and a dashed horizontal line marks one."
#| out-width: 100%

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H06_daily/",
  "H06_daily_temporal_h02_primary_context_functions.png"
))
```

```{r}
#| label: tbl-temporal-strongest-functions
#| tbl-cap: "Strongest fitted association on the shifted-response ratio scale."

strongest_rows |>
  dplyr::transmute(
    Estimand = .data$estimand_label,
    `Local time` = sprintf(
      "%02d:%02d",
      floor(.data$time_hour) %% 24,
      round((.data$time_hour %% 1) * 60)
    ),
    `Ratio for Y + 0.1` = .data$shifted_response_ratio,
    `Pointwise 95% CI` = sprintf(
      "%.2f to %.2f",
      .data$shifted_response_ratio_lower_95,
      .data$shifted_response_ratio_upper_95
    )
  ) |>
  gt::gt() |>
  gt::fmt_number(columns = `Ratio for Y + 0.1`, decimals = 2) |>
  compact_gt()
```

For the primary near-eye fit, the Free-day/Work-day shifted-response ratio was
lowest at 08:45 (`r sprintf("%.2f", primary_summary$minimum_shifted_response_ratio[
primary_summary$estimand_id == "free_minus_work"])`) and reached
`r sprintf("%.2f", primary_summary$maximum_shifted_response_ratio[
primary_summary$estimand_id == "free_minus_work"])` after midnight. The
Active/Sedentary ratio reached `r sprintf("%.2f",
primary_summary$maximum_shifted_response_ratio[
primary_summary$estimand_id == "active_minus_sedentary"])` at 15:15. One
hour greater sleep within participant was associated most strongly with a
ratio of `r sprintf("%.2f", primary_summary$ratio_at_maximum_absolute_association[
primary_summary$estimand_id == "sleep_within_plus_1h"])` at 00:15; the
between-participant sleep ratio was lowest at 08:45 (`r sprintf("%.2f",
primary_summary$minimum_shifted_response_ratio[
primary_summary$estimand_id == "sleep_between_plus_1h"])`). These summaries
describe fitted associations, not exposure effects caused by the contexts.

# Whole-function tests and multiplicity

```{r}
#| label: tbl-temporal-whole-function-tests
#| tbl-cap: "Exploratory whole-function tests and their prespecified four-test BH adjustment."

test_table <- dplyr::bind_rows(primary_tests, gap_tests) |>
  dplyr::mutate(
    Dataset = dplyr::recode(
      .data$data_scenario_id,
      primary = "Primary dataset",
      gap_timing_unaware = "Gap-timing-unaware dataset"
    ),
    Estimand = .data$estimand_label,
    `Raw p` = p_display(.data$raw_p_value),
    `BH-adjusted p` = p_display(.data$bh_adjusted_p_value),
    `BH decision` = ifelse(.data$bh_decision, "Retained", "Not retained")
  ) |>
  dplyr::select(
    Dataset, Estimand, `Raw p`, `BH-adjusted p`, `BH decision`
  )

test_table |>
  gt::gt() |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(
      columns = `BH-adjusted p`,
      rows = `BH decision` == "Retained"
    )
  ) |>
  gt::tab_footnote(
    footnote = paste(
      "Bold denotes retention under the explicitly labelled four-test BH",
      "rule. One mgcv approximate p-value was slightly negative from",
      "floating-point evaluation; it is preserved in source data and",
      "clamped to zero only for BH calculation and display."
    ),
    locations = gt::cells_column_labels(columns = `BH-adjusted p`)
  ) |>
  compact_gt()
```

The two BH families are separate: one for the four primary near-eye tests and
one for the four gap-timing-unaware near-eye tests. Chest and paired/common
frames are estimation-only complements. Neither family is combined with the
15 daily-metric families.

# Placement and dataset checks

The all-available chest fit was acceptable, and the paired/common near-eye fit
was acceptable. The paired/common chest fit was not acceptable because the
absolute-residual/fitted Spearman correlation was 0.313, above the
prespecified 0.300 limit. Its model object is retained for provenance, but no
context function, confidence interval, or p-value was inspected or extracted.

```{r}
#| label: fig-temporal-gap-sensitivity
#| fig-cap: "Primary and gap-timing-unaware near-eye association functions. Bands are pointwise 95% confidence intervals."
#| fig-alt: "Four-panel overlay of primary and gap-timing-unaware near-eye shifted-response ratio curves. The two datasets' curves and pointwise 95% confidence bands nearly overlap for day type, activity, and both sleep functions; the largest absolute ratio-curve difference is about 0.03."
#| out-width: 100%

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H06_daily/",
  "H06_daily_temporal_h02_near_eye_dataset_sensitivity.png"
))
```

The maximum absolute difference between the primary and
gap-timing-unaware near-eye ratio curves was below 0.031 for every function.
This is dataset stability for the exploratory fitted functions, not numerical
equivalence with the preregistration-aligned participant-day or approved
hourly H06 estimands.

# Diagnostic assessment

Five fitted frames met the prespecified frame, formula, convergence, basis,
residual, temporal-dependence, fixed-term identifiability, and cyclic-closure
rules. All six models converged without fitting warnings. The primary
near-eye residual Q--Q correlation was 0.965, its absolute-residual/fitted
Spearman correlation was 0.293, final pooled lag-1 correlation was 0.090, and
the largest absolute site lag-1 correlation was 0.159.

The smoothing Hessian's smallest relative eigenvalue in the primary near-eye
fit was $-1.95\times 10^{-15}$, within disclosed floating-point tolerance.
The retained global `k = 12` smooth had acceptable capacity after the bounded
`k = 16` work/free sensitivity already documented in the reconciliation
report.

Influence and site-deletion stability remain **withheld**, not passed. mgcv's
BAM object did not provide the required hat diagonal, and a deletion-refit
batch was outside the approved compute scope. Accordingly, the time-varying
associations may be reported as exploratory, but no site-deletion robustness
claim is made.

# Reproducibility and production closure

The six frame RDS files, six final BAM objects, diagnostics, pointwise source
data, tables, three figures, alt text, readability QA, and stage-specific
SHA-256 manifests are confined to H06_daily-owned paths. The figures were
exported at 170 mm with PNG, PDF, and SVG versions; all have paired source-data
CSVs and passed clipping, wrapping, panel-balance, and final-size legibility
checks.

This report implements the author's H06-D-G2P-H02 approval with the explicit
pointwise-only amendment. It does not request website integration, authorize a
simultaneous interval, authorize a deletion-refit batch, replace the standalone
daily-metric report, or transfer any estimate to the approved hourly H06
analysis.
