---
title: "H06 daily metrics — METRIC-010 MDER amendment"
subtitle: "Preregistration-aligned participant-day analysis under the mean-of-viable-ratios rule"
date: 2026-08-11
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("lme4")), "2.0.1"),
  identical(as.character(packageVersion("glmmTMB")), "1.1.14"),
  identical(as.character(packageVersion("melidosData")), "1.0.6")
)

source(file.path(
  project_root,
  "scripts/hypotheses/H06_daily/h06_daily_mder_metric010_contract.R"
))

read_h06d <- function(stage, file) {
  readr::read_csv(
    file.path(project_root, "artifacts", stage, "H06_daily", file),
    show_col_types = FALSE,
    na = ""
  )
}
sha256 <- function(path) {
  digest::digest(path, algo = "sha256", file = TRUE)
}
verify_manifest <- function(
  path,
  allowed_mismatch = character(),
  allowed_role_pattern = NULL
) {
  manifest <- readr::read_csv(path, show_col_types = FALSE, na = "")
  observed <- vapply(
    file.path(project_root, manifest$relative_path),
    sha256,
    character(1)
  )
  checked <- manifest |>
    dplyr::mutate(
      observed_sha256 = unname(observed),
      hash_match = .data$sha256 == .data$observed_sha256
    )
  if (!is.null(allowed_role_pattern)) {
    allowed_mismatch <- union(
      allowed_mismatch,
      manifest$relative_path[
        grepl(allowed_role_pattern, manifest$role)
      ]
    )
  }
  stopifnot(setequal(
    checked$relative_path[!checked$hash_match],
    allowed_mismatch
  ))
  checked
}
p_display <- function(value) {
  ifelse(
    is.na(value),
    "—",
    ifelse(value < 0.001, "<0.001", sprintf("%.3f", value))
  )
}
ci_display <- function(lower, upper, digits = 4L) {
  sprintf(
    paste0("%.", digits, "f to %.", digits, "f"),
    lower,
    upper
  )
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100)
    )
}

manifest_names <- c(
  "H06_daily_mder_metric010_static_input_manifest.csv",
  "H06_daily_mder_metric010_static_code_manifest.csv",
  "H06_daily_mder_metric010_static_output_manifest.csv",
  "H06_daily_mder_metric010_family_pilot_input_manifest.csv",
  "H06_daily_mder_metric010_family_pilot_code_manifest.csv",
  "H06_daily_mder_metric010_family_pilot_output_manifest.csv",
  "H06_daily_mder_metric010_influence_pilot_input_manifest.csv",
  "H06_daily_mder_metric010_influence_pilot_code_manifest.csv",
  "H06_daily_mder_metric010_influence_pilot_output_manifest.csv",
  "H06_daily_mder_metric010_production_input_manifest.csv",
  "H06_daily_mder_metric010_production_code_manifest.csv",
  "H06_daily_mder_metric010_production_output_manifest.csv",
  "H06_daily_mder_metric010_gap_refresh_input_manifest.csv",
  "H06_daily_mder_metric010_gap_refresh_code_manifest.csv",
  "H06_daily_mder_metric010_gap_refresh_output_manifest.csv",
  "H06_daily_mder_metric010_reader_input_manifest.csv",
  "H06_daily_mder_metric010_reader_code_manifest.csv",
  "H06_daily_mder_metric010_reader_output_manifest.csv"
)
pilot_historical_code_manifests <- c(
  "H06_daily_mder_metric010_family_pilot_code_manifest.csv",
  "H06_daily_mder_metric010_influence_pilot_code_manifest.csv"
)
frame_registry_path <- paste0(
  "artifacts/06_model_data/H06_daily/",
  "H06_daily_mder_metric010_frame_registry.csv"
)
verified_manifests <- Map(
  function(path, name) {
    verify_manifest(
      path,
      allowed_mismatch = if (name %in% pilot_historical_code_manifests) {
        paste0(
          "scripts/hypotheses/H06_daily/",
          c(
            "h06_daily_mder_metric010_contract.R",
            "h06_daily_mder_metric010_data.R",
            "h06_daily_mder_metric010_modeling.R"
          )
        )
      } else if (
        name == "H06_daily_mder_metric010_family_pilot_input_manifest.csv"
      ) {
        c(
          frame_registry_path,
          paste0(
            "artifacts/12_manifests/H06_daily/",
            "H06_daily_mder_metric010_static_output_manifest.csv"
          )
        )
      } else if (
        name == "H06_daily_mder_metric010_influence_pilot_input_manifest.csv"
      ) {
        frame_registry_path
      } else {
        character()
      },
      allowed_role_pattern = if (
        name == "H06_daily_mder_metric010_production_input_manifest.csv"
      ) {
        "^pre-refresh"
      } else {
        NULL
      }
    )
  },
  file.path(
    project_root,
    "artifacts/12_manifests/H06_daily",
    manifest_names
  ),
  manifest_names
)

input_contract <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_input_contract.csv"
)
metric_contract <- read_h06d(
  "06_model_data",
  "H06_daily_mder_metric010_metric_contract.csv"
)
support <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_support_summary.csv"
)
support_failures <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_support_failure_summary.csv"
)
static_verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_static_verdict.csv"
)
frames <- read_h06d(
  "06_model_data",
  "H06_daily_mder_metric010_frame_registry.csv"
)
effects <- read_h06d(
  "09_tables",
  "H06_daily_mder_metric010_effect_estimates.csv"
)
tests <- read_h06d(
  "09_tables",
  "H06_daily_mder_metric010_model_tests.csv"
)
bh_slots <- read_h06d(
  "09_tables",
  "H06_daily_mder_metric010_bh_slot_families.csv"
)
diagnostics <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_model_diagnostics.csv"
)
ar_counterparts <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_ar_counterparts.csv"
)
influence <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_influence_summary.csv"
)
production_verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_production_verdict.csv"
)
registered_benchmark <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_registered_benchmark.csv"
)
site_estimates <- read_h06d(
  "09_tables",
  "H06_daily_mder_metric010_site_heterogeneity_estimates.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_production_runtime.csv"
)
figure_qa <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_figure_readability_qa.csv"
)
frame_comparison <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_gap_refresh_frame_comparison.csv"
)
refresh_preservation <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_gap_refresh_preservation.csv"
)
refresh_provenance <- read_h06d(
  "08_diagnostics",
  "H06_daily_mder_metric010_gap_refresh_provenance.csv"
)

primary_effects <- effects |>
  dplyr::filter(
    .data$dataset_id == "primary",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available"
  )
primary_gaussian <- primary_effects |>
  dplyr::filter(.data$family == "Gaussian")
primary_t <- primary_effects |>
  dplyr::filter(.data$family == "Student-t")
primary_tests <- tests |>
  dplyr::filter(.data$dataset_id == "primary")
primary_diagnostics <- diagnostics |>
  dplyr::filter(
    .data$dataset_id == "primary",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available"
  )
primary_ar <- ar_counterparts |>
  dplyr::filter(
    .data$dataset_id == "primary",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available"
  )
gap_gaussian <- effects |>
  dplyr::filter(
    .data$dataset_id == "gap_timing_unaware",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available",
    .data$family == "Gaussian"
  )
gap_t <- effects |>
  dplyr::filter(
    .data$dataset_id == "gap_timing_unaware",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available",
    .data$family == "Student-t"
  )
gap_tests <- tests |>
  dplyr::filter(.data$dataset_id == "gap_timing_unaware")
gap_diagnostics <- diagnostics |>
  dplyr::filter(
    .data$dataset_id == "gap_timing_unaware",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available"
  )
gap_ar <- ar_counterparts |>
  dplyr::filter(
    .data$dataset_id == "gap_timing_unaware",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available"
  )

stopifnot(
  nrow(input_contract) == 16L,
  all(input_contract$verified),
  nrow(metric_contract) == 1L,
  metric_contract$metric_id == "mder_mean_of_viable_ratios",
  metric_contract$display_unit == "dimensionless",
  nrow(frames) == 36L,
  nrow(effects) == 72L,
  nrow(primary_gaussian) == 3L,
  nrow(primary_t) == 3L,
  all(primary_gaussian$lower_95 < 0),
  all(primary_gaussian$upper_95 > 0),
  all(primary_t$lower_95 < 0),
  all(primary_t$upper_95 > 0),
  all(primary_gaussian$student_t_shift_in_gaussian_se < 1),
  nrow(primary_tests) == 6L,
  nrow(bh_slots) == 180L,
  all(is.na(bh_slots$bh_adjusted_p_value)),
  all(bh_slots$family_slots_available == 1L),
  all(bh_slots$family_slots_required == 15L),
  all(bh_slots$family_status == "INCOMPLETE_1_OF_15_NO_BH_DECISION"),
  nrow(primary_diagnostics) == 3L,
  all(primary_diagnostics$gaussian_converged),
  all(primary_diagnostics$student_t_converged),
  all(primary_diagnostics$ar_acceptable),
  nrow(primary_ar) == 3L,
  all(primary_ar$ar_trigger),
  all(primary_ar$ar_acceptable),
  nrow(gap_gaussian) == 3L,
  nrow(gap_t) == 3L,
  isTRUE(all.equal(
    setNames(gap_gaussian$estimate, gap_gaussian$predictor_id)[c(
      "work_free_day",
      "activity_status",
      "previous_sleep_duration_centered_h"
    )],
    c(
      work_free_day = -0.001619123155536,
      activity_status = 0.009020499990256,
      previous_sleep_duration_centered_h = 0.001818193389273
    ),
    tolerance = 1e-12
  )),
  all(gap_gaussian$lower_95 < 0),
  all(gap_gaussian$upper_95 > 0),
  gap_t$lower_95[
    gap_t$predictor_id == "previous_sleep_duration_centered_h"
  ] > 0,
  nrow(gap_tests) == 6L,
  nrow(gap_diagnostics) == 3L,
  all(gap_diagnostics$diagnostic_disposition ==
    "ACCEPTABLE_WITH_HEAVY_TAIL_LIMITATION"),
  nrow(gap_ar) == 3L,
  all(gap_ar$ar_trigger),
  all(gap_ar$ar_acceptable),
  nrow(influence) == 3L,
  all(influence$estimate_magnitude_reportable),
  nrow(production_verdict) == 3L,
  all(production_verdict$claim_disposition != "NO_ASSOCIATION_CLAIM"),
  all(figure_qa$verdict == "PASS"),
  identical(support$available_days, c(702, 732, 687, 723)),
  identical(support$participants_with_value, c(137, 152, 137, 152)),
  support$exact_zeros[
    support$dataset_id == "gap_timing_unaware" &
      support$placement_id == "chest"
  ] == 0L,
  nrow(frame_comparison) == 36L,
  sum(!frame_comparison$object_identical) == 24L,
  all(refresh_preservation$preserved),
  all(refresh_provenance$verdict %in% c(
    "PASS",
    "INCOMPLETE_NO_BH_DECISION"
  ))
)

predictor_labels <- c(
  work_free_day = "Free day minus Work day",
  activity_status = "Active minus Sedentary",
  previous_sleep_duration_centered_h = "+1 h previous-night sleep"
)
```

::: {.callout-important appearance="simple" title="Answer in brief"}
Under the replacement METRIC-010 definition, the primary near-eye estimates
were -0.0013 MDER for Free day minus Work day (95% CI -0.0134 to
0.0108), 0.0083 for Active minus Sedentary (-0.0070 to 0.0235), and
0.0017 per one hour greater previous-night sleep (-0.0028 to 0.0063).
All three intervals include zero.

After the corrected gap-branch rebuild, the corresponding near-eye Gaussian
estimates are -0.0016, 0.0090, and 0.0018; their 95% intervals also
include zero. The mandatory Student-t sleep interval is only just above zero,
which is reported as tail-model sensitivity rather than a multiplicity-based
finding. The corrected gap samples contain 687 MDER days near eye and 723 at
the chest, with no finite nonpositive value.

The work/free and sleep estimates remain reportable as magnitudes, but their
directions reversed in at least one leave-one-site refit, so no stable
cross-site directional claim is defensible. The activity direction was stable,
but its generality is limited by site influence and all three Gaussian fits
retain a heavy-tail limitation. The mandatory Student-t and actual-date AR(1)
checks did not materially move the primary estimates.

These are **raw MDER-slot results only**. Each prespecified BH family currently
contains one of 15 required metric slots, so no adjusted p-value or
multiplicity-based decision exists. The associations are observational and do
not support causal or health claims.
:::

# Scope and supersession rule

This amendment analyzes MDER only. All MDER-independent H06_daily work is
frozen. No prior estimate, tail conclusion, device-day conclusion, or
time-profile threshold result from the superseded ratio-of-integrals
implementation is transferred.

The analysis unit is one admissible MDER-specific participant-day. Days are not
weighted by supported hours or by the number of viable minutes beyond the
admissibility threshold. This is the preregistration-aligned participant-day
route, not the approved hourly H06 model and not the exploratory 30-minute
GAMM.

# Exact replacement metric and provenance

Daily MDER is the arithmetic mean of viable one-minute melEDI/illuminance
ratios. A minute is viable only when both channels are finite and strictly
positive. The upstream preparation uses a complete 1,440-minute local
wall-clock grid, averages fall-back duplicate local minutes channel-wise
before forming a ratio, and leaves spring-forward absent minutes missing. MDER
is retained at **at least 720 viable ratios**, including exactly 720; otherwise
MDER alone is missing. There is no profile weighting, gating, scaling, or ratio
of daily integrals.

```{r}
#| label: tbl-mder-provenance
#| tbl-cap: "Pinned controlling METRIC-010 inputs."

input_contract |>
  dplyr::filter(.data$input_id %in% c(
    "metric010_decision", "metric_manifest", "mder_audit_manifest",
    "base_manifest", "gap_manifest", "h01_gap_manifest",
    "gap_repair_evidence", "gap_participant_day", "gap_mder_support"
  )) |>
  dplyr::transmute(
    Input = .data$input_id,
    Path = .data$relative_path,
    `SHA-256` = .data$observed_sha256,
    Verified = ifelse(.data$verified, "Yes", "No")
  ) |>
  gt::gt() |>
  gt::tab_style(
    style = gt::cell_text(font = "monospace", size = 10),
    locations = gt::cells_body(columns = `SHA-256`)
  ) |>
  compact_gt()
```

All 16 direct inputs matched their pinned SHA-256 identities. The controlling
decision is `audit/decisions/mder_mean_of_viable_ratios.md` with SHA-256
`1664347de976057807fcb5ac6e24bd4b66af1fe32378fabd1f2acafcfc9266de`.
The current metric, independent MDER-audit, base-model, gap-preparation, and
independent H01 gap manifests, together with the repair-evidence manifest and
both gap MDER objects, match the coordinator-supplied identities exactly. The
base input-bundle identity is
`168f25e18b6e494aa7a0272923041ad8249e25adb9ff3742d22e2f4cacf1bdf8`.

# Metric availability and exact fitted samples

```{r}
#| label: tbl-mder-availability
#| tbl-cap: "METRIC-010 availability before predictor-specific complete-case selection."

support |>
  dplyr::transmute(
    Dataset = .data$dataset_label,
    Placement = .data$placement,
    `Available days` = sprintf(
      "%s / %s",
      .data$available_days,
      .data$base_days
    ),
    `Participants with MDER` = .data$participants_with_value,
    `Sites with MDER` = .data$sites_with_value,
    `Exact zeros` = .data$exact_zeros
  ) |>
  gt::gt() |>
  gt::fmt_integer(
    columns = c(`Participants with MDER`, `Sites with MDER`, `Exact zeros`)
  ) |>
  compact_gt()
```

The primary values reproduce the independent verification: 702 of 816
near-eye participant-days from 137 participants and 732 of 902 chest days from
152 participants. Among retained primary days, viable-minute counts ranged
from 721 to 1,440 near eye and 721 to 1,334 at the chest.

The repaired gap branch contains 687 of 811 near-eye days from 137
participants and 723 of 897 chest days from 152 participants. Every retained
value is finite and strictly positive. The former THUAS chest zero is now
missing because it has no viable momentary ratios; no local repair was made in
H06_daily.

```{r}
#| label: tbl-mder-gap-support-flow
#| tbl-cap: "Exact repaired gap-branch MDER support flow."

support_failures |>
  dplyr::filter(.data$dataset_id == "gap_timing_unaware") |>
  dplyr::transmute(
    Placement = .data$placement,
    Status = dplyr::recode(
      .data$support_status,
      available = "Available",
      below_viable_ratio_fraction = "Below 720 viable ratios",
      no_viable_momentary_ratio = "No viable momentary ratios"
    ),
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites
  ) |>
  gt::gt(groupname_col = "Placement") |>
  gt::fmt_integer(columns = c(`Participant-days`, Participants, Sites)) |>
  compact_gt()
```

```{r}
#| label: tbl-mder-primary-samples
#| tbl-cap: "Exact all-available fitted samples for each predictor."

frames |>
  dplyr::filter(.data$sample_role == "all_available") |>
  dplyr::transmute(
    Dataset = dplyr::recode(
      .data$dataset_id,
      primary = "Primary",
      gap_timing_unaware = "Gap-timing-unaware"
    ),
    Placement = dplyr::recode(
      .data$placement_id,
      near_eye = "Near eye",
      chest = "Chest"
    ),
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites
  ) |>
  gt::gt(groupname_col = "Dataset") |>
  gt::fmt_integer(
    columns = c(`Participant-days`, Participants, Sites)
  ) |>
  compact_gt()
```

For primary near eye, the work/free and sleep models each used 679 days from
136 participants across nine sites; the activity model used 645 days from 133
participants across nine sites. These are predictor-specific complete-case
samples after the MDER support rule; no day was weighted by its viable-minute
count.

# Model hierarchy and estimands

For predictor `X`, the exact registered benchmark was retained unchanged:

```{r}
#| label: exact-mder-formulas
#| echo: true

registered_random_site <-
  response_value ~ X + (X | site) + (1 | site:participant_key)

fixed_site_additive <-
  response_value ~ site + X + (1 | participant_key)

fixed_site_heterogeneity <-
  response_value ~ site * X + (1 | participant_key)

gap_aware_ar1_counterpart <-
  response_value ~ site + X + (1 | participant_key) +
    ar1(day_index_factor + 0 | day_sequence_id)
```

The fixed-site additive Gaussian identity-link model is the selected primary
candidate. It estimates an absolute, dimensionless MDER difference common to
the nine submitted sites after fixed-site adjustment and accounts for repeated
days with a participant random intercept. Practical contrasts are Free day
minus Work day, Active minus Sedentary, and the slope per one hour greater
previous-night sleep duration. Work day, Sedentary, and 8 h sleep are the
registered reference codings.

The fixed-site interaction model separately describes predictor-by-site
heterogeneity. Its site-specific estimates do not replace the common primary
contrast. The exact random-site/random-slope formula is a labelled registered
benchmark only; it is never silently simplified or called the primary model.

# Primary near-eye results

```{r}
#| label: fig-mder-primary-effects
#| fig-cap: "Near-eye primary MDER estimates and mandatory heavy-tail sensitivities. Bars are 95% confidence intervals."
#| fig-alt: "Forest plot of three near-eye primary MDER associations with 95% confidence intervals and Student-t sensitivity estimates. The selected Gaussian estimates are minus 0.0013 for Free day minus Work day, 0.0083 for Active minus Sedentary, and 0.0017 per one hour greater previous-night sleep. Every selected and sensitivity interval crosses zero."
#| out-width: 100%

knitr::include_graphics(file.path(
  project_root,
  "artifacts/10_figures/H06_daily/",
  "H06_daily_mder_metric010_primary_effects.png"
))
```

```{r}
#| label: tbl-mder-primary-effects
#| tbl-cap: "Primary near-eye estimates with Gaussian selection and Student-t sensitivity."

primary_effects |>
  dplyr::transmute(
    Contrast = unname(predictor_labels[.data$predictor_id]),
    Model = dplyr::recode(
      .data$family,
      Gaussian = "Gaussian identity (selected)",
      `Student-t` = "Student-t identity (sensitivity)"
    ),
    Estimate = .data$estimate,
    `95% CI` = ci_display(.data$lower_95, .data$upper_95),
    `Days / participants / sites` = sprintf(
      "%d / %d / %d",
      .data$participant_days,
      .data$participants,
      .data$sites
    )
  ) |>
  gt::gt(groupname_col = "Contrast") |>
  gt::fmt_number(columns = Estimate, decimals = 4) |>
  compact_gt()
```

The Student-t estimates moved by 0.125, 0.628, and 0.645 primary Gaussian
standard errors for work/free, activity, and sleep, respectively. All shifts
were below the prespecified one-standard-error stability threshold, and all
Student-t intervals also included zero.

# Raw tests and incomplete multiplicity families

```{r}
#| label: tbl-mder-primary-tests
#| tbl-cap: "Raw MDER-slot tests; no BH-adjusted decision is available."

primary_tests |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Test = dplyr::recode(
      .data$test_type,
      association = "Common association",
      site_heterogeneity = "Predictor-by-site heterogeneity"
    ),
    `Raw p` = p_display(.data$raw_p_value),
    `BH-adjusted p` = "—",
    Status = "1 of 15 slots; no decision"
  ) |>
  gt::gt(groupname_col = "Predictor") |>
  compact_gt()
```

The raw common-association p-values were 0.830, 0.275, and 0.483. The raw
heterogeneity p-values were 0.005, 0.053, and 0.778. None is a multiplicity-
adjusted finding. For each activated predictor, association and heterogeneity
are separate named 15-metric BH vectors, and the gap-timing-unaware dataset has
its own separate vectors. Only the MDER slot is available in this amendment;
the other 14 slots are explicitly retained as unavailable, not silently
removed. Consequently, every adjusted value and every adjusted decision is
missing by design. No raw p-value is bolded or promoted to a primary decision.

# Family and residual diagnostics

```{r}
#| label: tbl-mder-family-diagnostics
#| tbl-cap: "Primary near-eye family diagnostics and sensitivity disposition."

primary_diagnostics |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Gaussian Q-Q r` = .data$gaussian_residual_qq_correlation,
    `|residual|-fitted rho` =
      .data$gaussian_absolute_residual_fitted_spearman,
    `|standardized residual| > 4` =
      .data$gaussian_standardized_residual_gt4_fraction,
    `Student-t df` = .data$student_t_t_degrees_of_freedom,
    `Student-t Q-Q r` =
      .data$student_t_quantile_residual_qq_correlation,
    Assessment = "Acceptable with heavy-tail limitation"
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(
      `Gaussian Q-Q r`, `|residual|-fitted rho`,
      `|standardized residual| > 4`, `Student-t df`, `Student-t Q-Q r`
    ),
    decimals = 3
  ) |>
  compact_gt()
```

All selected Gaussian and Student-t models converged, had positive-definite
Hessians, finite coefficients and standard errors, and nonsingular participant
random intercepts. Gaussian spread trends, extreme-residual fractions, and
physical-bound checks passed, but Gaussian Q--Q correlations of 0.870--0.879
failed the nominal 0.95 normal-tail rule. The Student-t quantile-residual Q--Q
correlations were 0.997. The Gaussian identity model is therefore reportable
only with the explicit heavy-tail limitation and mandatory Student-t
sensitivity; Gaussian adequacy is not claimed.

# Actual-date temporal check

```{r}
#| label: tbl-mder-ar-check
#| tbl-cap: "Primary near-eye actual-date residual screen and gap-aware AR(1) counterpart."

primary_ar |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `One-day pairs` = .data$adjacent_pairs,
    `Participants with pair` = .data$participants,
    `Pooled lag-1 r` = .data$independent_residual_lag1,
    `Maximum site |r|` =
      .data$independent_maximum_absolute_site_lag1,
    `AR(1) rho` = .data$ar_rho,
    `Effect shift (primary SE)` = .data$ar_effect_shift_in_primary_se,
    Assessment = "Stable sensitivity; structured-covariance singularity flagged"
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = c(`One-day pairs`, `Participants with pair`)) |>
  gt::fmt_number(
    columns = c(
      `Pooled lag-1 r`, `Maximum site |r|`, `AR(1) rho`,
      `Effect shift (primary SE)`
    ),
    decimals = 3
  ) |>
  compact_gt()
```

Each primary frame exceeded 100 true one-day pairs and 20 participants. The
pooled correlations were below 0.20 in absolute value, but site maxima near
0.39 crossed the 0.30 trigger, so actual-date, gap-aware AR(1) counterparts
were fitted. They converged with positive-definite Hessians, absolute rho below
0.95, and changed the associations by only 0.015--0.056 primary standard
errors.

The automated structured-covariance singularity diagnostic was `TRUE` for all
three AR(1) counterparts even though the participant and AR standard
deviations were finite. These models are therefore retained only as temporal-
stability sensitivities, not promoted as preferred primary models. The tiny
effect shifts support stability of the reported primary magnitudes, but no
claim of a fully adequate AR generative model is made.

# Influence and site heterogeneity

```{r}
#| label: tbl-mder-influence
#| tbl-cap: "Primary near-eye deletion diagnostics and claim disposition."

influence |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Maximum participant shift (SE)` =
      .data$maximum_participant_shift_in_full_se,
    `Participant reversals` = .data$participant_direction_reversal_count,
    `Maximum site shift (SE)` = .data$maximum_site_shift_in_full_se,
    `Most influential site` = .data$maximum_shift_site,
    `Site reversals` = .data$site_direction_reversal_count,
    Disposition = dplyr::recode(
      .data$influence_disposition,
      MAGNITUDE_REPORTABLE_CROSS_SITE_DIRECTION_UNSTABLE =
        "Magnitude reportable; cross-site direction unstable",
      MAGNITUDE_REPORTABLE_SITE_GENERALITY_LIMITED =
        "Magnitude reportable; site generality limited"
    )
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(
      `Maximum participant shift (SE)`, `Maximum site shift (SE)`
    ),
    decimals = 2
  ) |>
  gt::fmt_integer(
    columns = c(`Participant reversals`, `Site reversals`)
  ) |>
  compact_gt()
```

All 405 participant deletions and 27 site deletions fitted without failure.
Maximum participant shifts were below one primary standard error. Site shifts
were between 1.05 and 1.22 standard errors, which limits generality. The
work/free estimate reversed under deletion of KNUST, MPI, or RISE; the sleep
estimate reversed after deleting IZTECH. Their magnitudes and intervals are
reportable, but their cross-site directions are not stable. Activity had no
participant or site reversal, so its association estimate is reportable with
the stated site-influence and heavy-tail limitations.

The descriptive site-specific work/free estimates ranged from -0.0703 at
KNUST to 0.0460 at IZTECH; activity ranged from -0.0381 at BAUA to 0.0818
at KNUST; and sleep ranged from -0.0063 per hour at RISE to 0.0081 at
IZTECH. These are unadjusted heterogeneity descriptions, not separately tested
site claims.

# Registered benchmark

```{r}
#| label: tbl-mder-registered-benchmark
#| tbl-cap: "Exact registered random-site/random-slope benchmark disposition."

registered_benchmark |>
  dplyr::filter(.data$dataset_id == "primary") |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Formula = .data$formula,
    `Maximum |site correlation|` =
      .data$maximum_absolute_random_site_correlation,
    Converged = ifelse(.data$converged, "Yes", "No"),
    Singular = ifelse(.data$singular, "Yes", "No"),
    Disposition = dplyr::recode(
      .data$disposition,
      ESTIMABLE_BENCHMARK_ONLY = "Estimable benchmark only",
      NON_ESTIMABLE = "Non-estimable"
    )
  ) |>
  gt::gt() |>
  gt::fmt_number(columns = `Maximum |site correlation|`, decimals = 3) |>
  compact_gt()
```

The exact work/free and activity benchmarks were estimable but remain
benchmarks only. The sleep benchmark was singular, had a random-site
correlation at 1.000, and was classified non-estimable. It was not simplified
post hoc. The fixed-site route is the approved primary structure for the nine
submitted sites.

# Placement, paired/common, and dataset comparisons

```{r}
#| label: tbl-mder-placement-comparisons
#| tbl-cap: "Selected Gaussian placement estimates in the primary dataset."

effects |>
  dplyr::filter(
    .data$dataset_id == "primary",
    .data$family == "Gaussian",
    .data$sample_role %in% c("all_available", "paired_common")
  ) |>
  dplyr::transmute(
    Sample = dplyr::recode(
      .data$sample_role,
      all_available = "All available",
      paired_common = "Paired/common"
    ),
    Placement = dplyr::recode(
      .data$placement_id,
      near_eye = "Near eye",
      chest = "Chest"
    ),
    Contrast = unname(predictor_labels[.data$predictor_id]),
    Estimate = .data$estimate,
    `95% CI` = ci_display(.data$lower_95, .data$upper_95),
    `Days / participants / sites` = sprintf(
      "%d / %d / %d",
      .data$participant_days,
      .data$participants,
      .data$sites
    )
  ) |>
  gt::gt(groupname_col = "Sample") |>
  gt::fmt_number(columns = Estimate, decimals = 4) |>
  compact_gt()
```

All all-available and paired/common placement intervals included zero. The
paired/common work/free estimates were nearly identical near eye and at the
chest; activity and sleep estimates varied modestly by placement but remained
interval-compatible with zero. These are complementary placement analyses,
not replacements for near-eye primary.

```{r}
#| label: tbl-mder-dataset-comparisons
#| tbl-cap: "Near-eye primary versus gap-timing-unaware estimates on all-available and dataset-common samples."

effects |>
  dplyr::filter(
    .data$placement_id == "near_eye",
    .data$family == "Gaussian",
    .data$sample_role %in% c("all_available", "dataset_common")
  ) |>
  dplyr::transmute(
    Sample = dplyr::recode(
      .data$sample_role,
      all_available = "All available",
      dataset_common = "Primary/gap common"
    ),
    Dataset = dplyr::recode(
      .data$dataset_id,
      primary = "Primary",
      gap_timing_unaware = "Gap-timing-unaware"
    ),
    Contrast = unname(predictor_labels[.data$predictor_id]),
    Estimate = .data$estimate,
    `95% CI` = ci_display(.data$lower_95, .data$upper_95),
    `Days / participants / sites` = sprintf(
      "%d / %d / %d",
      .data$participant_days,
      .data$participants,
      .data$sites
    )
  ) |>
  gt::gt(groupname_col = "Sample") |>
  gt::fmt_number(columns = Estimate, decimals = 4) |>
  compact_gt()
```

On all available repaired near-eye days, the Gaussian estimates were
-0.0016 for work/free (95% CI -0.0140 to 0.0108), 0.0090 for activity
(-0.0066 to 0.0246), and 0.0018 per hour of previous-night sleep
(-0.0028 to 0.0065). All three intervals include zero. The exact
primary--gap common samples contain 664 work/free and sleep days and 630
activity days; on those rows the primary and gap estimates differ by at most
0.0004 MDER.

```{r}
#| label: tbl-mder-gap-family-results
#| tbl-cap: "Repaired gap near-eye estimates and mandatory heavy-tail sensitivities."

effects |>
  dplyr::filter(
    .data$dataset_id == "gap_timing_unaware",
    .data$placement_id == "near_eye",
    .data$sample_role == "all_available"
  ) |>
  dplyr::transmute(
    Contrast = unname(predictor_labels[.data$predictor_id]),
    Model = dplyr::recode(
      .data$family,
      Gaussian = "Gaussian identity (selected)",
      `Student-t` = "Student-t identity (sensitivity)"
    ),
    Estimate = .data$estimate,
    `95% CI` = ci_display(.data$lower_95, .data$upper_95, digits = 6L),
    `Days / participants / sites` = sprintf(
      "%d / %d / %d",
      .data$participant_days,
      .data$participants,
      .data$sites
    )
  ) |>
  gt::gt(groupname_col = "Contrast") |>
  gt::fmt_number(columns = Estimate, decimals = 6) |>
  compact_gt()
```

The Student-t shifts were 0.111, 0.543, and 0.651 Gaussian standard errors,
all below the prespecified one-standard-error threshold. The Student-t sleep
interval was 0.000003 to 0.006710, only just above zero; the corresponding
selected Gaussian interval includes zero. This tail-model sensitivity is
disclosed explicitly and does not create a BH-adjusted finding.

```{r}
#| label: tbl-mder-gap-tests
#| tbl-cap: "Repaired gap MDER-slot raw tests; no BH-adjusted decision is available."

gap_tests |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Test = dplyr::recode(
      .data$test_type,
      association = "Common association",
      site_heterogeneity = "Predictor-by-site heterogeneity"
    ),
    `Raw p` = p_display(.data$raw_p_value),
    `BH-adjusted p` = "—",
    Status = "1 of 15 slots; no decision"
  ) |>
  gt::gt(groupname_col = "Predictor") |>
  compact_gt()
```

The repaired gap raw common-association p-values are 0.804, 0.245, and 0.470.
The raw heterogeneity p-values are 0.005, 0.046, and 0.808. The work/free and
activity heterogeneity values are nominal raw signals only: their 15-slot
heterogeneity families remain incomplete, so neither has an adjusted p-value
or an inferential decision.

```{r}
#| label: tbl-mder-gap-diagnostics
#| tbl-cap: "Repaired gap near-eye residual, Student-t, and actual-date AR(1) checks."

gap_diagnostics |>
  dplyr::select(
    "predictor_id",
    "gaussian_residual_qq_correlation",
    "gaussian_absolute_residual_fitted_spearman",
    "student_t_quantile_residual_qq_correlation",
    "student_t_effect_shift_in_gaussian_se"
  ) |>
  dplyr::left_join(
    gap_ar |>
      dplyr::select(
        "predictor_id", "adjacent_pairs", "participants",
        "independent_residual_lag1",
        "independent_maximum_absolute_site_lag1",
        "ar_rho", "ar_effect_shift_in_primary_se", "ar_singular"
      ),
    by = "predictor_id",
    relationship = "one-to-one"
  ) |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Gaussian Q-Q r` = .data$gaussian_residual_qq_correlation,
    `|residual|-fitted rho` =
      .data$gaussian_absolute_residual_fitted_spearman,
    `Student-t Q-Q r` = .data$student_t_quantile_residual_qq_correlation,
    `One-day pairs / participants` = sprintf(
      "%d / %d",
      .data$adjacent_pairs,
      .data$participants
    ),
    `Pooled lag-1 r` = .data$independent_residual_lag1,
    `Maximum site |r|` =
      .data$independent_maximum_absolute_site_lag1,
    `AR(1) rho` = .data$ar_rho,
    `AR effect shift (SE)` = .data$ar_effect_shift_in_primary_se,
    Assessment = ifelse(
      .data$ar_singular,
      "Acceptable stability sensitivity; covariance singularity flagged",
      "Acceptable stability sensitivity"
    )
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(
      `Gaussian Q-Q r`, `|residual|-fitted rho`, `Student-t Q-Q r`,
      `Pooled lag-1 r`, `Maximum site |r|`, `AR(1) rho`,
      `AR effect shift (SE)`
    ),
    decimals = 3
  ) |>
  compact_gt()
```

All repaired gap Gaussian and Student-t fits passed convergence, spread,
extreme-residual, bounds, and sensitivity rules. Gaussian Q--Q correlations
of 0.870--0.880 retain the same heavy-tail limitation as the primary route;
Student-t Q--Q correlations were 0.996--0.997. Site lag maxima triggered all
three actual-date AR(1) counterparts. Their rho values ranged from -0.734
to -0.685, they shifted estimates by only 0.019--0.061 Gaussian standard
errors, and they carry the structured-covariance singularity flag. They remain
stability sensitivities only. The exact registered gap sleep benchmark is
non-estimable because its random-site correlation reached 1.000; it was not
simplified post hoc.

The repaired chest source has no finite nonpositive MDER value. The previous
THUAS zero and every other superseded gap result are excluded from this report;
the corrected THUAS day is missing under the no-viable-ratio rule.

# Bounded refresh and frozen-result integrity

```{r}
#| label: tbl-mder-gap-refresh-provenance
#| tbl-cap: "Bounded METRIC-010 gap-refresh execution and preservation record."

refresh_provenance |>
  dplyr::transmute(
    Domain = .data$domain,
    Observed = .data$observed,
    Verdict = .data$verdict
  ) |>
  gt::gt() |>
  gt::tab_style(
    style = gt::cell_fill(color = "#E8F3EC"),
    locations = gt::cells_body(
      columns = Verdict,
      rows = Verdict == "PASS"
    )
  ) |>
  compact_gt()
```

Exactly 24 MDER-dependent frames were refreshed: all 18 gap frames and the six
primary--gap common-sample frames whose membership necessarily changes with
gap MDER availability. The 12 primary all-available or placement-paired frames
are object-identical to their pre-refresh copies. Twelve additional production
preservation domains verify unchanged primary estimates, model tests,
heterogeneity rows, participant/site deletion results, claim verdicts, and
unaffected fitted-model serializations. No primary all-available,
placement-paired, or deletion model was refitted; the six primary--gap
common-sample fits were necessarily updated.

The prior production output manifest was reconstructed byte-for-byte from the
captured pre-refresh input rows and retained under a new immutable task-owned
path. The targeted refresh completed in 11.7 seconds and used no bootstrap,
simulation, or resampling.

# Comparison with V0 and the approved hourly analysis

V0 fitted one repeated hourly melEDI outcome and jointly entered context
predictors. This amendment fits one dimensionless daily MDER value per
admissible participant-day and one predictor-specific model at a time. V0 days
with more supported hours contributed more rows; here every admissible MDER day
contributes one row. V0 coefficients, p-values, and claims therefore do not
map numerically to these estimates.

The approved hourly H06 analysis also remains a separate estimand. This MDER
amendment is not its sensitivity analysis and does not replace it.

# Reproducibility and H06-D-G2-MDER-GAP decision record

The static audit, 36 exact frame RDS files, family pilot, 50-deletion runtime
pilot, original production batch, targeted 24-frame repair-dependent refresh,
conditional AR counterparts, registered benchmarks, frozen participant/site
deletion batch, tables, reader figure, paired source CSV, alt text, QA record,
and stage-specific SHA-256 manifests are confined to H06_daily-owned paths.
The original production script completed in 31.3 seconds; the targeted refresh
completed in 11.7 seconds. No bootstrap or heavy resampling was run.

The family and influence pilots retain their original output manifests. Their
historical code manifests pin the task-local contract, data adapter, and
modeling helper as they existed at pilot time. The family pilot also pins the
then-current frame registry and static output manifest; the influence pilot
pins the then-current frame registry. Those shared task-local files changed
only for the approved METRIC-010 source/support repin and repair-dependent
frames. The exact primary pilot frames and pilot outputs remain unchanged. The
report permits precisely these recorded historical mismatches. The current
production input manifest additionally retains the 14
pre-refresh output hashes and the prior output-manifest hash as historical
inputs. Those paths now correctly differ because the affected artifacts were
replaced; their prior manifest is reconstructed byte-for-byte and its identity
is verified by the dedicated gap-refresh provenance manifests. All current
static, production-output, refresh, and reader hashes match exactly.

The primary H06-D-G2-MDER decisions remain author-approved and unchanged. The
corrected gap branch stopped at **H06-D-G2-MDER-GAP**, and the author approved
the following items on 2026-08-11:

1. the repaired gap availability and exact predictor-specific samples;
2. the three selected Gaussian gap estimates and their explicit heavy-tail
   limitation;
3. disclosure that the Student-t sleep interval is only just above zero while
   the selected Gaussian interval includes zero;
4. treating the raw work/free and activity heterogeneity p-values as
   non-decisional because their 15-slot BH families remain incomplete;
5. retaining the singularly flagged gap AR(1) fits only as temporal-stability
   sensitivities and the singular gap sleep random-slope fit as a
   non-estimable registered benchmark; and
6. the verified preservation of every unaffected primary result, with no
   primary all-available, placement-paired, or deletion refit.

This approval closes and freezes the corrected MDER-specific branch. It does
not authorize the separate H06-D-G2P-AR repair decision, another model fit, the
remaining daily-metric production grid, a Stage 3 reader report, preparation
companion, site integration, shared preparation change, manuscript edit,
commit, or upload.
