---
title: "H06 daily metrics — bounded daily AR-repair pilot"
subtitle: "True-date temporal checks for two unresolved non-MDER response families"
date: 2026-08-11
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("lme4")), "2.0.1"),
  identical(as.character(packageVersion("glmmTMB")), "1.1.14"),
  identical(as.character(packageVersion("melidosData")), "1.0.6")
)

source(file.path(
  project_root,
  "scripts/hypotheses/H06_daily/h06_daily_daily_ar_repair.R"
))

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_h06d <- function(stage, file) {
  readr::read_csv(artifact(stage, file), show_col_types = FALSE, na = "")
}
sha256 <- function(path) {
  digest::digest(path, algo = "sha256", file = TRUE)
}
verify_relative_manifest <- function(manifest, hash_column = "sha256") {
  paths <- file.path(project_root, manifest$relative_path)
  stopifnot(all(file.exists(paths)))
  observed <- unname(vapply(paths, sha256, character(1)))
  stopifnot(identical(observed, manifest[[hash_column]]))
  invisible(TRUE)
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100)
    )
}
ci_text <- function(lower, upper, digits = 3L) {
  sprintf(
    paste0("%.", digits, "f to %.", digits, "f"),
    lower,
    upper
  )
}

input_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_daily_ar_repair_pilot_input_manifest.csv"
)
code_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_daily_ar_repair_pilot_code_manifest.csv"
)
output_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_daily_ar_repair_pilot_output_manifest.csv"
)
software_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_stage2_daily_ar_repair_pilot_software_manifest.csv"
)

stopifnot(
  all(input_manifest$verified),
  identical(input_manifest$expected_sha256, input_manifest$observed_sha256)
)
input_current_check <- input_manifest |>
  dplyr::mutate(
    current_observed_sha256 = unname(vapply(
      file.path(project_root, .data$relative_path),
      sha256,
      character(1)
    )),
    current_match = .data$expected_sha256 == .data$current_observed_sha256
  )
stopifnot(setequal(
  input_current_check$input_id[!input_current_check$current_match],
  "base_manifest"
))
verify_relative_manifest(code_manifest)
verify_relative_manifest(output_manifest)

installed_versions <- vapply(
  software_manifest$item[software_manifest$item != "R"],
  function(package) as.character(packageVersion(package)),
  character(1)
)
stopifnot(
  identical(
    unname(installed_versions),
    software_manifest$version[software_manifest$item != "R"]
  )
)

frame_registry <- read_h06d(
  "06_model_data",
  "H06_daily_stage2_daily_ar_repair_pilot_frame_registry.csv"
)
frame_equivalence <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_frame_equivalence.csv"
)
support <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_support.csv"
)
diagnostics <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_model_diagnostics.csv"
)
verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_verdict.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_runtime.csv"
)
l10_zero_audit <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_l10_zero_mass_audit.csv"
)
l10_near_zero <- read_h06d(
  "08_diagnostics",
  "H06_daily_stage2_daily_ar_repair_pilot_l10_near_zero_rows.csv"
)
effects <- read_h06d(
  "09_tables",
  "H06_daily_stage2_daily_ar_repair_pilot_effect_stability.csv"
)

ar_diagnostics <- diagnostics |>
  dplyr::filter(.data$model_id == "glmmTMB_gap_aware_ar1")
overall <- verdict |>
  dplyr::filter(.data$domain == "Overall repair gate")
failed_domains <- verdict |>
  dplyr::filter(!.data$passed, .data$domain != "Overall repair gate")

l10_frame <- readRDS(file.path(
  project_root,
  frame_registry$frame_relative_path[
    frame_registry$repair_id == "l10_positive_magnitude"
  ]
))
next_l10_positive_lx <- min(
  l10_frame$response_source[l10_frame$response_source > 1e-12]
)

stopifnot(
  nrow(frame_registry) == 2L,
  all(frame_equivalence$equivalence_pass),
  all(support$support_pass),
  nrow(overall) == 2L,
  all(overall$verdict == "NOT_ACCEPTABLE"),
  l10_zero_audit$positive_values_at_or_below_1e_12 == 3L
)

repair_labels <- c(
  pre_sleep_identity = "Pre-sleep duration × previous-night sleep",
  l10_positive_magnitude = "Positive L10 magnitude × work/free day"
)
model_labels <- c(
  frozen_lmer_independent = "Frozen independent model",
  glmmTMB_independent = "Independent glmmTMB calibration",
  glmmTMB_gap_aware_ar1 = "True-date, gap-aware AR(1)"
)
```

::: {.callout-important}
## Answer in brief

Neither repair is acceptable for production. The pre-sleep AR(1) model has a
non-positive-definite Hessian and false-convergence code even though its
effect direction, residual distribution, physical bounds, and post-AR
temporal screen pass. The positive-L10 model is singular, estimates its AR(1)
correlation at the boundary, leaves stronger residual temporal dependence,
and fails the Gaussian distributional rule.

The L10 audit also found three strictly positive values of
`4.163336e-17 lx`. The exact-zero split sends these values to the positive
component and maps them to `-16.381` under `log10()`. They were retained
unchanged, but their scientific classification must be decided by the shared
metric owner before L10 can proceed.

No association claim, predictor p-value, multiplicity decision, remaining
model-grid fit, deletion batch, bootstrap, or simulation was produced. This
document stops at **H06-D-G2P-AR**.
:::

# Scope and scientific boundary

This bounded pilot addresses two response-family problems identified in the
approved Stage 2 pilot:

1. whether an actual-date, gap-aware daily AR(1) counterpart is numerically
   defensible for duration below 10 lx melEDI before sleep; and
2. whether the same covariance construction repairs the positive-magnitude
   component of L10 mean melEDI.

Both analyses retain one metric-specific participant-day per row. A missing
calendar date starts a new sequence; observed days on either side of a gap are
never made adjacent by compressing time. This is not the repeated-hourly H06
analysis and not the separate exploratory 30-minute temporal GAMM.

The current post-MDER shared sources were compared analytically with the
frozen Stage 2 pilot frames before fitting. Participant-day keys, response and
predictor values, column order, and factor levels were identical for both
repairs. The MDER outcome was neither opened nor refitted.

::: {.callout-caution}
## Post-pilot shared-manifest qualification

All 12 inputs matched their controlling hashes when the four pilot fits were
run. At report render, 11 still matched. The broad shared
`base_model_data_artifacts.csv` path had changed from the controlling
`6cfbfb18...` identity to `b6fa2283...`; the directly used near-eye
participant-day source remains byte-identical at `fb04a84f...`, and both
task-owned fitted frames remain hash-identical to the verified pilot outputs.
The mismatch is recorded in the shared change request. This report can audit
the frozen fits, but no new fit or repinning is permitted until the preparation
coordinator reconciles the broad manifest identities.
:::

# Exact model and sequence contract

The independent calibration and the AR counterpart use Gaussian identity
models fitted by `glmmTMB` with REML. Site is fixed, participant has a random
intercept, and the AR term uses the within-sequence actual-day index. The
default Gaussian residual dispersion was retained because this was the exact
prespecified pilot contract.

```{r}
#| label: exact-repair-formulas
#| echo: true

pre_sleep_independent_formula <- h06d_ar_formula(
  "previous_sleep_duration_centered_h",
  ar = FALSE
)
pre_sleep_ar_formula <- h06d_ar_formula(
  "previous_sleep_duration_centered_h",
  ar = TRUE
)
l10_independent_formula <- h06d_ar_formula(
  "work_free_day",
  ar = FALSE
)
l10_ar_formula <- h06d_ar_formula(
  "work_free_day",
  ar = TRUE
)

pre_sleep_independent_formula
pre_sleep_ar_formula
l10_independent_formula
l10_ar_formula
```

The AR term is exactly
`ar1(day_index_factor + 0 | day_sequence_id)`. `day_index_factor` restarts at
one after every missing date, while `day_sequence_id` identifies that
contiguous participant sequence. The pilot did not estimate a pooled rho from
another metric or reuse any MDER correlation.

# Frozen/current equality and temporal support

```{r}
#| label: tbl-repair-support
#| tbl-cap: "Frozen/current frame equality and actual-date AR support."

support |>
  dplyr::left_join(
    frame_equivalence |>
      dplyr::select(
        "repair_id",
        "equivalence_pass",
        "old_normalized_sha256",
        "current_normalized_sha256"
      ),
    by = "repair_id"
  ) |>
  dplyr::transmute(
    Repair = unname(repair_labels[.data$repair_id]),
    `Frozen/current equal` = ifelse(.data$equivalence_pass, "Yes", "No"),
    Days = .data$observations,
    Participants = .data$participants,
    Sites = .data$sites,
    `Contiguous sequences` = .data$sequences,
    `True adjacent pairs` = .data$adjacent_pairs,
    `Participants with adjacency` =
      .data$participants_with_adjacent_pair,
    `Maximum sequence (days)` = .data$maximum_sequence_days
  ) |>
  gt::gt() |>
  compact_gt()
```

The pre-sleep frame contains 648 days from 139 participants at all nine sites,
with 450 true adjacent-day pairs. The positive-L10 frame contains 680 days
from 140 participants at all nine sites, with 484 pairs. Both exceed the
prespecified minimum of 100 pairs and 20 participants with adjacency.

# Engineering-only coefficient stability

```{r}
#| label: tbl-repair-effect-stability
#| tbl-cap: "Engineering comparison of frozen independent and AR-repair coefficients; these are not association results."

effects |>
  dplyr::filter(.data$model_id != "glmmTMB_independent") |>
  dplyr::mutate(
    result = dplyr::if_else(
      .data$repair_id == "pre_sleep_identity",
      sprintf(
        "%.3f h (95%% interval %s)",
        .data$display_estimate,
        ci_text(.data$display_lower_95, .data$display_upper_95)
      ),
      sprintf(
        "%.3f ratio (95%% interval %s)",
        .data$display_estimate,
        ci_text(.data$display_lower_95, .data$display_upper_95)
      )
    )
  ) |>
  dplyr::transmute(
    Repair = unname(repair_labels[.data$repair_id]),
    Model = unname(model_labels[.data$model_id]),
    `Engineering coefficient` = .data$result,
    `Shift from frozen model (SE)` =
      .data$effect_shift_in_frozen_lmer_se,
    `Direction retained` =
      ifelse(.data$direction_matches_frozen_lmer, "Yes", "No")
  ) |>
  gt::gt(groupname_col = "Repair") |>
  gt::fmt_number(columns = `Shift from frozen model (SE)`, decimals = 3) |>
  compact_gt()
```

The pre-sleep AR coefficient shifts by 0.130 frozen-model standard errors; the
positive-L10 coefficient shifts by 0.897. Both retain the frozen direction and
pass the prespecified less-than-one-standard-error engineering screen. That
does not rescue a model that fails numerical or residual diagnostics. The
intervals above are model-based pointwise engineering intervals only; no
predictor test or claim was extracted.

# Why neither AR repair passes

```{r}
#| label: tbl-ar-diagnostics
#| tbl-cap: "Prespecified diagnostics for the two actual-date, gap-aware AR(1) fits."

ar_diagnostics |>
  dplyr::transmute(
    Repair = unname(repair_labels[.data$repair_id]),
    Converged = ifelse(.data$converged, "Yes", "No"),
    `Positive-definite Hessian` =
      ifelse(.data$positive_definite_hessian, "Yes", "No"),
    Singular = ifelse(.data$singular, "Yes", "No"),
    rho = .data$ar_rho,
    `AR SD` = .data$ar_standard_deviation,
    `Participant SD` = .data$participant_standard_deviation,
    `Residual SD` = .data$residual_standard_deviation,
    `Post-AR pooled lag-1` = .data$true_date_residual_lag1,
    `Maximum |site lag-1|` =
      .data$maximum_absolute_site_residual_lag1,
    `Q-Q correlation` = .data$residual_qq_correlation,
    `Residual spread correlation` =
      .data$absolute_residual_fitted_spearman
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(
      rho,
      `AR SD`,
      `Participant SD`,
      `Residual SD`,
      `Post-AR pooled lag-1`,
      `Maximum |site lag-1|`,
      `Q-Q correlation`,
      `Residual spread correlation`
    ),
    decimals = 3
  ) |>
  compact_gt()
```

```{r}
#| label: tbl-failed-domains
#| tbl-cap: "Failed acceptance domains; every listed failure blocks production use."

failed_domains |>
  dplyr::transmute(
    Repair = unname(repair_labels[.data$repair_id]),
    `Failed domain` = .data$domain,
    Rule = .data$rule,
    Observed = .data$observed
  ) |>
  gt::gt(groupname_col = "Repair") |>
  compact_gt()
```

## Pre-sleep duration

The pre-sleep AR fit returned false-convergence code 8, two warnings, and a
non-positive-definite Hessian. Its estimated residual standard deviation
collapsed to `8.8e-05 h`, while separate participant and AR standard
deviations remained finite. The fitted rho was -0.085, and the post-AR pooled
and maximum site residual lag screens passed. Residual distribution and
physical-bound rules also passed. Nevertheless, the numerical failure is
decisive: this fit cannot be used for inference.

The collapse of the residual dispersion motivates, but does not itself
authorize, one narrowly specified no-nugget diagnostic using the same frame,
fixed effects, participant random intercept, and AR structure with
`dispformula = ~0`. That would test whether redundant Gaussian nugget and AR
variance components caused the failure. It would remain a diagnostic repair,
not an association result, until it passed the full gate.

## Positive L10 magnitude

The positive-L10 AR fit returned singular-convergence code 7. Rho was 1.000
to displayed precision, beyond the prespecified `|rho| < 0.95` rule. The
post-AR pooled residual lag was -0.491 and the maximum absolute site lag was
0.932, both worse than acceptable. Its Q-Q correlation was 0.808 and residual
spread correlation was -0.225, so the Gaussian positive-magnitude family also
failed independently of the AR boundary. Physical predictions remained
within the prespecified bounds, but that isolated pass cannot rescue the fit.

# L10 numerical zero-mass qualification

```{r}
#| label: tbl-l10-zero-audit
#| tbl-cap: "Exact-zero and numerical-near-zero audit for the complete representative L10 work/free frame."

l10_zero_audit |>
  dplyr::transmute(
    `Participant-days` = .data$participant_days,
    `Exact zeros` = .data$exact_zeros,
    `Exact-zero fraction` = .data$exact_zero_fraction,
    `Strictly positive values <= 1e-12 lx` =
      .data$positive_values_at_or_below_1e_12,
    `All values <= 1e-12 lx` = .data$values_at_or_below_1e_12,
    `Positive-component rows` = .data$positive_component_rows,
    `Minimum positive L10 (lx)` =
      .data$positive_component_minimum_lx,
    `Minimum log10 L10` = .data$positive_component_minimum_log10
  ) |>
  gt::gt() |>
  gt::fmt_percent(columns = `Exact-zero fraction`, decimals = 1) |>
  gt::fmt_scientific(columns = `Minimum positive L10 (lx)`, decimals = 3) |>
  gt::fmt_number(columns = `Minimum log10 L10`, decimals = 3) |>
  compact_gt()
```

```{r}
#| label: tbl-l10-near-zero-rows
#| tbl-cap: "The three strictly positive numerical-near-zero L10 rows retained by the exact-zero pilot contract."

l10_near_zero |>
  dplyr::transmute(
    Site = .data$site,
    Participant = .data$Id,
    Date = .data$local_date,
    `L10 mean melEDI (lx)` = .data$l10_mean_medi_lx,
    `log10 L10` = .data$log10_l10_mean_medi,
    `Day type` = .data$work_free_day
  ) |>
  gt::gt() |>
  gt::fmt_scientific(columns = `L10 mean melEDI (lx)`, decimals = 6) |>
  gt::fmt_number(columns = `log10 L10`, decimals = 3) |>
  compact_gt()
```

The complete work/free frame has 104 exact zeros among 784 participant-days.
The three rows above are strictly positive in the prepared metric and are
therefore retained by the exact `L10 > 0` component split. The next value above
the numerical-near-zero cluster is `r format(next_l10_positive_lx, scientific = TRUE, digits = 4)` lx. No
tolerance, rounding, deletion, replacement, or local reclassification was
applied.

This requires a shared metric decision. If the three values are floating-point
representations of zero, the rule must be implemented and independently
verified in shared preparation, followed by updated manifests and targeted
H06_daily repinning. If strict positivity is scientifically intended, the
values must remain and a separately approved robust positive-magnitude family
pilot must be specified. H06_daily must not choose a threshold locally.

# Runtime and withheld compute

```{r}
#| label: tbl-repair-runtime
#| tbl-cap: "Measured pilot runtime and bounded planning projections."

runtime |>
  dplyr::transmute(
    Component = .data$component,
    Scope = .data$scope,
    `Fit units` = .data$fit_units,
    `Seconds (measured or projected)` =
      .data$observed_or_projected_seconds,
    Basis = .data$basis,
    Disposition = .data$interpretation
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = `Seconds (measured or projected)`,
    decimals = 2
  ) |>
  gt::sub_missing(columns = everything(), missing_text = "Not run") |>
  compact_gt()
```

The four fitted models completed in less than one second of measured fitting
time. The remaining fit-only grid is therefore not blocked by raw runtime; it
is blocked by unresolved scientific and numerical contracts. Any deletion
diagnostic, bootstrap, or simulation batch remains outside this pilot and
would first require its own 50/100-run production-code runtime pilot and
approval.

# Reproducibility and H06-D-G2P-AR stop gate

All computations ran in R 4.6.1 using the synchronized project library. The
12 fit-time inputs, five executed code files, 16 scientific outputs, and
package versions are hash-pinned in H06_daily-owned manifests. Eleven of 12
inputs still match at report render; the one current broad base-manifest drift
is disclosed above and in the shared change request. The directly used data
source, both fitted frames, all executed code, and every scientific pilot
output still match exactly. The pilot fitted only one independent calibration
and one AR counterpart for each of the two prespecified frames. It did not fit
another response family, alter a shared metric, or update any multiplicity
slot.

This report stops at **H06-D-G2P-AR**. Before any further non-MDER daily model
is fitted, the author should explicitly decide:

1. whether to authorize one bounded no-nugget pre-sleep AR diagnostic with
   `dispformula = ~0`, followed by the unchanged full acceptance gate;
2. whether L10 should remain on hold pending the shared metric owner's
   classification of the three numerical-near-zero positive values;
3. if strict positivity is confirmed upstream, whether a new bounded robust
   positive-magnitude family pilot may be specified before any production use;
   and
4. whether the remaining daily-metric production grid should stay frozen
   until those two response-family issues are resolved.

The recommended decision is to authorize only the single pre-sleep diagnostic,
retain the L10 hold pending upstream classification, and keep the production
grid frozen. No Stage 3 reader report, Stage 4 preparation companion, website
integration, shared preparation edit, manuscript edit, commit, or upload is
authorized at this gate.
