---
title: "H06 daily metrics — pre-sleep no-nugget AR diagnostic"
subtitle: "Bounded H06-D-G2P-AR-NN temporal-sensitivity gate"
date: 2026-08-11
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

stopifnot(identical(as.character(getRversion()), "4.6.1"))
library(dplyr)
library(readr)

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_csv <- function(path) {
  readr::read_csv(path, show_col_types = FALSE, na = "")
}
render_table <- function(data, caption, digits = NULL) {
  arguments <- list(
    x = data,
    format = "html",
    caption = caption,
    escape = TRUE,
    align = "l"
  )
  if (!is.null(digits)) {
    arguments$digits <- digits
  }
  do.call(knitr::kable, arguments)
}

diagnostics <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_final_model_diagnostics.csv"
))
verdict <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_final_verdict.csv"
))
correction <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_audit_correction.csv"
))
support <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_support.csv"
))
equivalence <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_frame_equivalence.csv"
))
runtime <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_runtime.csv"
))
effects <- read_csv(artifact(
  "09_tables",
  "H06_daily_pre_sleep_no_nugget_effect_stability.csv"
))
repaired_inputs <- read_csv(artifact(
  "12_manifests",
  "H06_daily_pre_sleep_no_nugget_diagnostic_repaired_input_manifest.csv"
))
reference_provenance <- read_csv(artifact(
  "08_diagnostics",
  "H06_daily_pre_sleep_no_nugget_reference_provenance.csv"
))

stopifnot(
  nrow(diagnostics) == 1L,
  all(verdict$passed),
  identical(
    verdict$verdict[verdict$domain == "Overall no-nugget gate"],
    "ACCEPTABLE"
  ),
  runtime$fits_run[[1L]] == 1L,
  runtime$bootstrap_replicates[[1L]] == 0L,
  runtime$simulation_replicates[[1L]] == 0L,
  runtime$deletion_refits[[1L]] == 0L,
  runtime$remaining_grid_fits[[1L]] == 0L,
  all(repaired_inputs$verified),
  reference_provenance$serialization_byte_identical[[1L]],
  !reference_provenance$l10_field_in_executed_frame[[1L]],
  !reference_provenance$l10_text_in_executed_formula[[1L]]
)
```

::: {.callout-important}

## Answer in brief

The author approved **H06-D-G2P-AR-NN** on 2026-08-11. The single no-nugget
model is therefore frozen as **acceptable as a gap-aware temporal-stability
sensitivity**. Fixing rather than estimating the negligible Gaussian residual
dispersion resolved the earlier false convergence. Every prespecified
numerical, covariance, effect-stability, temporal, distributional, and
physical-bounds check passed.

The fitted association coefficient is shown only to verify stability. This
gate does not promote a new association finding or p-value. L10 and the
remaining daily-metric production grid remain on hold while the coordinator
implements the author-approved shared numerical-zero rule.

No association claim is made from this bounded diagnostic.

:::

# Scope and scientific boundary

The response is duration below 10 lx melEDI before sleep, in hours, with one
admissible participant-day per row. The predictor is previous-night sleep
duration, in hours. The model asks whether the earlier free-dispersion AR fit's
numerical failure disappears when the residual nugget is fixed, without
changing the response, sample, mean structure, participant clustering, or
calendar-gap definition.

The exact model was:

```r
response_value ~ site + previous_sleep_duration_centered_h +
  (1 | participant_key) +
  ar1(day_index_factor + 0 | day_sequence_id)
```

It used a Gaussian identity family, REML, and `dispformula = ~0`. In glmmTMB
1.1.14 this fixes the residual standard deviation at 0.000122070 h; it does not
estimate another dispersion coefficient.

Exactly one model was fitted. There was no independent-model refit, L10 or
MDER fit, remaining-grid fit, bootstrap, simulation, or deletion refit.

# Input identity and sample

The controlling repaired input manifest contains only current shared
provenance, the separately sealed pre-sleep frame, the author authorization,
and the standalone pre-sleep-only historical reference.

```{r}
#| label: tbl-no-nugget-input-summary
#| tbl-cap: "Controlling no-nugget input contract after the no-refit provenance-container repair."

input_summary <- repaired_inputs |>
  dplyr::transmute(
    Input = .data$input_id,
    Role = .data$role,
    Verified = ifelse(.data$verified, "Yes", "No")
  )
render_table(input_summary, caption = NULL)
```

The current reconstruction and authorized frozen frame remain exactly equal.

```{r}
#| label: tbl-no-nugget-sample
#| tbl-cap: "Exact model-frame and temporal-sequence support."

sample_table <- tibble::tibble(
  `Participant-days` = support$observations,
  Participants = support$participants,
  Sites = support$sites,
  `Gap-delimited sequences` = support$sequences,
  `True adjacent-day pairs` = support$adjacent_pairs,
  `Participants with adjacency` = support$participants_with_adjacent_pair,
  `Frame identity` = ifelse(
    equivalence$equivalence_pass,
    "Exact match",
    "Mismatch"
  )
)
render_table(sample_table, caption = NULL)
```

# Model and diagnostic result

The no-nugget model converged with a positive-definite Hessian, no warnings,
and no structured-covariance singularity. The estimated AR coefficient was
-0.085. Conditional residual dependence passed the prespecified rule.

```{r}
#| label: tbl-no-nugget-diagnostics
#| tbl-cap: "Selected model and residual diagnostics."

diagnostic_table <- diagnostics |>
  dplyr::transmute(
    Converged = ifelse(.data$converged, "Yes", "No"),
    `Positive-definite Hessian` = ifelse(
      .data$positive_definite_hessian,
      "Yes",
      "No"
    ),
    Singular = ifelse(.data$singular, "Yes", "No"),
    Warnings = .data$warning_count,
    `Participant SD (h)` = sprintf("%.3f", .data$participant_standard_deviation),
    `AR SD (h)` = sprintf("%.3f", .data$ar_standard_deviation),
    Rho = sprintf("%.3f", .data$ar_rho),
    `Pooled residual lag-1` = sprintf("%.3f", .data$true_date_residual_lag1),
    `Maximum absolute site lag-1` = sprintf(
      "%.3f",
      .data$maximum_absolute_site_residual_lag1
    ),
    `Q-Q correlation` = sprintf("%.3f", .data$residual_qq_correlation),
    `Spread Spearman` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `Outside-bound fraction` = sprintf(
      "%.3f",
      pmax(
        .data$conditional_bound_violation_fraction,
        .data$marginal_bound_violation_fraction
      )
    )
  )
render_table(diagnostic_table, caption = NULL)
```

All 12 substantive domains and the overall gate passed.

```{r}
#| label: tbl-no-nugget-verdict
#| tbl-cap: "Complete prespecified acceptance gate."

verdict_table <- verdict |>
  dplyr::transmute(
    Domain = .data$domain,
    Status = .data$verdict,
    Evidence = .data$observed
  )
render_table(verdict_table, caption = NULL)
```

# Effect-stability check

The diagnostic coefficient is virtually identical to the failed
free-dispersion AR estimate and remains close to the frozen independent-model
estimate. Its purpose here is engineering stability, not a separate
association conclusion.

```{r}
#| label: tbl-no-nugget-effect-stability
#| tbl-cap: "Engineering coefficient stability across the frozen and AR specifications; no p-value or multiplicity decision is produced."

model_labels <- c(
  frozen_lmer_independent = "Frozen independent model",
  failed_free_dispersion_ar1 = "Failed free-dispersion AR model",
  no_nugget_ar1 = "No-nugget AR diagnostic"
)
effect_table <- effects |>
  dplyr::mutate(Model = unname(model_labels[.data$model_id])) |>
  dplyr::transmute(
    Model,
    `Estimate (h per +1 h sleep)` = sprintf("%.3f", .data$estimate),
    `Model-based 95% interval (h)` = sprintf(
      "%.3f to %.3f",
      .data$lower_95,
      .data$upper_95
    ),
    `Shift from frozen model (SE)` = sprintf(
      "%.3f",
      .data$shift_from_frozen_lmer_se
    ),
    `Direction retained` = ifelse(
      .data$direction_matches_frozen_lmer,
      "Yes",
      "No"
    )
  )
render_table(effect_table, caption = NULL)
```

The no-nugget coefficient is -0.108 h per one-hour greater previous-night
sleep duration, with model-based 95% interval -0.157 to -0.058 h. It shifted
by 0.130 frozen-model standard errors and by only 0.000011 frozen-model
standard errors from the failed free-dispersion AR coefficient. These values
support temporal stability only; this bounded gate makes no inferential claim.

# Transparent parameter-map correction

The preliminary post-fit gate was falsely negative because it treated the
fixed dispersion value returned by `fixef()` as a freely estimated parameter.
The no-refit finalizer instead inspected glmmTMB's parameter map and optimized
vector.

```{r}
#| label: tbl-no-nugget-audit-correction
#| tbl-cap: "No-refit correction of the preliminary glmmTMB parameter-map audit."

correction_table <- correction |>
  dplyr::transmute(
    Issue = .data$issue,
    `Correct semantics` = .data$glmmTMB_semantics,
    Evidence = .data$evidence,
    Action = .data$action,
    Refit = ifelse(.data$model_refit, "Yes", "No"),
    `Scientific quantity changed` = ifelse(
      .data$scientific_quantity_changed,
      "Yes",
      "No"
    ),
    `Final verdict` = .data$final_verdict
  )
render_table(correction_table, caption = NULL)
```

The stored formula is `~0`; the dispersion map is entirely fixed; no
dispersion parameter occurs in the optimized vector; and the residual standard
deviation equals the package control value. Only the derived specification and
overall verdict rows changed. The model, coefficient, uncertainty, covariance
parameters, residuals, and other diagnostics did not.

# Standalone pre-sleep provenance repair

The initial process deserialized a sealed historical composite checkpoint and
accessed only `$fits$pre_sleep_identity`. The composite also contains a
historical L10 member. The coordinator confirmed that this did not create an
analytical L10 dependency because the L10 member was never accessed and the
executed frame, response, and formula contain no L10 field.

Before report finalization, the accessed pre-sleep member was sealed as a
standalone reference. Its version-3 serialized bytes are exactly identical to
the parent member. The controlling repaired input manifest now pins that
standalone object rather than treating the composite as a scientific input.

```{r}
#| label: tbl-no-nugget-reference-provenance
#| tbl-cap: "Standalone pre-sleep reference identity and composite-container disclosure."

reference_table <- reference_provenance |>
  dplyr::transmute(
    `Parent path` = .data$parent_relative_path,
    `Parent SHA-256` = .data$parent_sha256,
    `Exact subobject` = .data$exact_subobject_path,
    `Standalone path` = .data$standalone_relative_path,
    `Standalone SHA-256` = .data$standalone_file_sha256,
    `Serialized bytes identical` = ifelse(
      .data$serialization_byte_identical,
      "Yes",
      "No"
    ),
    `Frame identical to executed fit` = ifelse(
      .data$frame_identical_to_executed_model,
      "Yes",
      "No"
    ),
    `L10 field in executed frame` = ifelse(
      .data$l10_field_in_executed_frame,
      "Yes",
      "No"
    )
  )
render_table(reference_table, caption = NULL)
```

This cleanup did not rerun the model or alter any shared or L10 artifact.

# L10 and remaining-grid boundary

The author classified the three L10 values of
`4.163336342344337e-17 lx` as numerical zeros, plausibly introduced by the
10-hour averaging calculation. The coordinator's candidate shared rule resets
a tiny positive only when it is within the specified machine-precision
tolerance and every contributing source minute is exactly zero. Focused tests
pass for eight proven roundoff cells while genuinely low positive values are
preserved; the canonical rebuild and final manifest sealing remain in
progress. H06_daily has not changed, thresholded, discarded, modeled, or
repinned those values.

L10 response-family selection, L10 models, L10 multiplicity slots, L10 report
claims, and the remaining daily production grid remain on hold until the
coordinator supplies sealed shared identities and targeted repinning scope.

# Reproducibility and H06-D-G2P-AR-NN approval record

All computation used R 4.6.1 and the synchronized project library. The model
fit took 0.243 seconds. The original one-fit input/code/output/software
manifests, immutable preliminary audit, no-refit final audit, standalone-
reference repair, and controlling repaired input manifest preserve the full
execution chain.

The author approved **H06-D-G2P-AR-NN** as recommended on 2026-08-11. This
approval:

1. accept the no-nugget pre-sleep AR counterpart as an acceptable gap-aware
   temporal-stability sensitivity;
2. retain the independently specified fixed-site participant-intercept model
   as the primary route and not promote this diagnostic coefficient as a
   separate finding;
3. accept the transparent no-refit parameter-map and provenance-container
   corrections; and
4. retain the L10 and remaining-grid hold until the coordinator supplies the
   shared numerical-zero rebuild, downstream impact audit, and targeted
   repinning instructions.

This closes and freezes the bounded no-nugget diagnostic without authorizing
an additional fit, Stage 3/4 output, site integration, shared preparation
change, manuscript edit, commit, upload, or website integration. The L10 and
remaining-grid hold remains in force.
