---
title: "H06 daily metrics — METRIC-011 L10 amendment"
subtitle: "Two-part L10 analysis and estimability stop gate"
date: 2026-08-12
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
project_library <- file.path(
  project_root,
  "renv/library/macos/R-4.6/aarch64-apple-darwin23"
)
.libPaths(c(project_library, .libPaths()))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("lme4")), "2.0.1"),
  identical(as.character(packageVersion("glmmTMB")), "1.1.14")
)

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_h06d <- function(stage, file) {
  readr::read_csv(
    artifact(stage, file),
    show_col_types = FALSE,
    na = c("", "NA")
  )
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100),
      source_notes.font.size = gt::px(11)
    )
}
ratio_ci <- function(estimate, lower, upper) {
  ifelse(
    is.na(lower) | is.na(upper),
    sprintf("%.3f (interval withheld)", estimate),
    sprintf("%.3f (%.3f to %.3f)", estimate, lower, upper)
  )
}
yes_no <- function(value) ifelse(value, "Yes", "No")

input_contract <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_input_contract.csv"
)
static_verdict <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_static_verdict.csv"
)
exact_zeros <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_exact_zero_audit.csv"
)
change_inventory <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_metric011_scenario_component_change_inventory.csv"
)
samples <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_sample_summary.csv"
)
primary_verdict <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_primary_verdict.csv"
)
ar_counterparts <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_ar_counterparts.csv"
)
model_diagnostics <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_model_diagnostics.csv"
)
occurrence_cells <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_occurrence_site_cells.csv"
)
family_sensitivity <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_positive_family_sensitivity.csv"
)
influence <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_positive_influence_summary.csv"
)
map_coefficients <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_map_occurrence_coefficients.csv"
)
map_predictions <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_map_occurrence_predictions.csv"
)
registered_benchmark <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_registered_benchmark.csv"
)
runtime <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_production_runtime.csv"
)
preservation <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_metric011_protected_preservation_after_production.csv"
)
model_equivalence <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_no_refit_model_equivalence.csv"
)
figure_qa <- read_h06d(
  "08_diagnostics", "H06_daily_l10_metric011_figure_readability_qa.csv"
)
model_tests <- read_h06d(
  "09_tables", "H06_daily_l10_metric011_model_tests.csv"
)
effects <- read_h06d(
  "09_tables", "H06_daily_l10_metric011_effect_estimates.csv"
)
gap_comparison <- read_h06d(
  "09_tables", "H06_daily_l10_metric011_gap_comparison.csv"
)
bh_slots <- read_h06d(
  "09_tables", "H06_daily_l10_metric011_bh_slot_families.csv"
)
figure_source <- read_h06d(
  "11_source_data",
  "H06_daily_l10_metric011_positive_component_figure_source.csv"
)

primary_joint <- model_tests |>
  dplyr::filter(
    .data$run_id == "primary__near_eye__all_available",
    .data$component == "joint_two_part"
  )
gap_joint <- model_tests |>
  dplyr::filter(
    .data$run_id == "gap_timing_unaware__near_eye__all_available",
    .data$component == "joint_two_part"
  )

stopifnot(
  nrow(input_contract) == 13L,
  all(input_contract$identity_pass),
  static_verdict$component_frames == 36L,
  static_verdict$current_primary_changed_cells == 8L,
  static_verdict$gap_components_exactly_invariant,
  sum(exact_zeros$metric011_changed_source_cell) == 8L,
  nrow(change_inventory) == 36L,
  all(change_inventory$exact_frame_identity_to_pre_metric011[
    change_inventory$dataset_id == "gap_timing_unaware"
  ]),
  nrow(primary_joint) == 6L,
  nrow(gap_joint) == 6L,
  all(primary_joint$test_status == "NON_ESTIMABLE_COMPONENT_FAILURE"),
  all(gap_joint$test_status == "NON_ESTIMABLE_COMPONENT_FAILURE"),
  all(is.na(primary_joint$raw_p_value)),
  all(is.na(gap_joint$raw_p_value)),
  all(bh_slots$slot_status[bh_slots$metric_slot == 3L] ==
        "NON_ESTIMABLE_COMPONENT_FAILURE"),
  all(is.na(bh_slots$raw_p_value[bh_slots$metric_slot == 3L])),
  all(is.na(bh_slots$bh_adjusted_p_value)),
  all(bh_slots$family_slots_required == 15L),
  all(bh_slots$family_slots_available == 1L),
  all(preservation$preserved_byte_for_byte),
  nrow(model_equivalence) == 194L,
  all(model_equivalence$object_identical),
  all(figure_qa$disposition == "PASS"),
  identical(unique(figure_qa$interval_type), "pointwise, not simultaneous"),
  all(figure_source$interval_available[
    figure_source$family == "Gaussian primary component"
  ]),
  !any(figure_source$interval_available[
    figure_source$family == "Student-t sensitivity" &
      grepl("UNRESOLVED", figure_source$inferential_status)
  ])
)
```

::: {.callout-important}

## Answer in brief

The approved METRIC-011 normalization correctly reclassified eight proven
roundoff remnants as exact zero L10 mean melEDI values. Those participant-days
were retained: exact zeros entered the occurrence component and only the
strictly positive magnitude component excluded them.

The prespecified fixed-site occurrence model is nevertheless **not estimable
for any of the three primary near-eye context tests** because site-specific
zero patterns cause complete or quasi-complete separation. The work/free
actual-date AR(1) counterpart also failed at the correlation boundary
($\rho=-0.999964$). Consequently, the overall two-part L10 association and
site-heterogeneity tests are `NON_ESTIMABLE_COMPONENT_FAILURE`; their raw and
adjusted p-values are missing. The named L10 slot remains in each 15-metric
Benjamini--Hochberg family. No L10 joint association is claimed.

Among participant-days with L10 greater than zero, the Gaussian component can
be shown descriptively with pointwise 95% intervals, but it cannot replace the
joint estimand. Its work/free result reverses direction under the Student-t
sensitivity, the activity result has a 1.04-standard-error site-deletion shift,
and the sleep Student-t sensitivity fails its numerical gate. None supports a
standalone confirmatory or directional claim.

This report stops at **H06-D-G2P-L10-METRIC011** for an author decision. It
does not release the remaining daily-metric grid or alter the approved
pre-sleep no-nugget sensitivity.

:::

# Scope and controlling METRIC-011 rule

L10 mean melEDI is the arithmetic mean melEDI during the least-active 10-hour
period of an admissible participant-day, reported in lx. METRIC-011 changes a
derived L10 value to exact zero only when the value is within the shared
machine-precision rule **and every contributing source minute is exactly
zero**. It is not a fitted-model cutoff and does not round genuinely low
positive exposure.

The shared repair changed three near-eye and five chest participant-day cells.
All non-L10 scientific values, samples, hourly values, 30-minute values, and
gap-timing-unaware L10 values were unchanged. H06_daily consumed the sealed
artifacts; it did not rebuild preparation.

```{r}
#| label: tbl-l10-input-contract
#| tbl-cap: "Controlling METRIC-011 inputs. Every displayed identity was verified before fitting."

input_contract |>
  dplyr::transmute(
    Input = .data$input_id,
    Role = .data$role,
    `SHA-256 (first 12)` = substr(.data$expected_sha256, 1L, 12L),
    Verified = yes_no(.data$identity_pass)
  ) |>
  gt::gt() |>
  compact_gt()
```

# Estimand and model contract

Each model frame has one metric-specific participant-day per row. The
prespecified two-part analysis separates:

1. **Occurrence:** $P(\mathrm{L10}=0\mid X)$, using a binomial logit model in
   which an exact zero is the event; and
2. **Positive magnitude:**
   $E[\log_{10}(\mathrm{L10})\mid \mathrm{L10}>0,X]$, using a Gaussian
   identity mixed model, with a Student-t identity model as the mandatory
   response-family sensitivity.

The fixed-site additive formula used in both components was:

```r
response_value ~ site + predictor + (1 | participant_key)
```

The site-heterogeneity comparison used:

```r
response_value ~ site * predictor + (1 | participant_key)
```

An overall two-part likelihood-ratio test was to combine the component
likelihood-ratio statistics only if **both** components passed their
component-appropriate estimability and diagnostic gates. That condition did
not hold. The Gaussian positive-component ratio is
$10^{\widehat\beta}$: a conditional geometric-mean ratio among days with
L10 greater than zero, not a marginal mean ratio across all days.

The exact registered benchmark was retained without simplification:

```r
response_value ~ predictor + (predictor | site) +
  (1 | site:participant_key)
```

It remains a benchmark only. It was not promoted to replace the prespecified
fixed-site primary model.

# Exact samples and zero handling

The occurrence component keeps every model-eligible day. The positive
component excludes only exact-zero L10 days. The table gives exact fitted
support as days / participants / sites.

```{r}
#| label: tbl-l10-samples
#| tbl-cap: "Exact component samples for every authorized scenario and predictor."

scenario_labels <- c(
  primary__near_eye__all_available = "Primary near eye, all available",
  primary__chest__all_available = "Primary chest, all available",
  primary__near_eye__paired_common = "Primary near eye, paired/common",
  primary__chest__paired_common = "Primary chest, paired/common",
  gap_timing_unaware__near_eye__all_available =
    "Gap-timing-unaware near eye",
  gap_timing_unaware__chest__all_available =
    "Gap-timing-unaware chest"
)
predictor_labels <- c(
  work_free_day = "Work/free day",
  activity_status = "Activity status",
  previous_sleep_duration_centered_h = "Previous-night sleep"
)

sample_table <- samples |>
  dplyr::mutate(
    Scenario = unname(scenario_labels[.data$run_id]),
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Support = sprintf(
      "%d / %d / %d",
      .data$participant_days,
      .data$participants,
      .data$sites
    ),
    Component = dplyr::recode(
      .data$component,
      zero_occurrence = "Occurrence",
      positive_magnitude = "Positive magnitude"
    )
  ) |>
  dplyr::select(
    .data$run_order,
    .data$predictor_order,
    .data$Scenario,
    .data$Predictor,
    .data$Component,
    .data$Support
  ) |>
  tidyr::pivot_wider(names_from = .data$Component, values_from = .data$Support) |>
  dplyr::arrange(.data$run_order, .data$predictor_order) |>
  dplyr::select(-.data$run_order, -.data$predictor_order)

sample_table |>
  gt::gt(groupname_col = "Scenario") |>
  gt::tab_spanner(
    label = "Participant-days / participants / sites",
    columns = c("Occurrence", "Positive magnitude")
  ) |>
  compact_gt()
```

Across all source scenarios and placements, 493 exact-zero L10 rows were
audited. The eight METRIC-011 reclassifications are included among them. No
participant-day was discarded merely because L10 equals zero.

# Why the primary joint estimand is not available

The primary near-eye occurrence frames contain substantial overall zero mass,
but the event is absent in structurally important site cells. In particular,
UCR has no zero L10 day under either work/free category, either activity
category, or anywhere over the observed sleep-duration support. With site
fixed effects in a logistic mixed model, these empty event cells drive fixed
coefficients toward the boundary: maximum absolute coefficients were
16.41--16.93 and maximum standard errors were 5,003--6,833.

```{r}
#| label: tbl-l10-ucr-cells
#| tbl-cap: "Primary near-eye UCR occurrence support, showing the zero-free site cells central to separation."

occurrence_cells |>
  dplyr::filter(
    .data$run_id == "primary__near_eye__all_available",
    .data$site == "UCR"
  ) |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Level = .data$predictor_level,
    Days = .data$participant_days,
    `Exact-zero days` = .data$exact_zero_days,
    `Positive days` = .data$positive_days
  ) |>
  gt::gt() |>
  compact_gt()
```

```{r}
#| label: tbl-l10-primary-estimability
#| tbl-cap: "Primary near-eye component and joint-test dispositions. Missing p-values are intentional estimability decisions."

primary_test_table <- model_tests |>
  dplyr::filter(
    .data$run_id == "primary__near_eye__all_available",
    .data$test_type == "association"
  ) |>
  dplyr::mutate(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Component = dplyr::recode(
      .data$component,
      joint_two_part = "Overall two-part",
      zero_occurrence = "Occurrence",
      positive_magnitude = "Positive magnitude"
    ),
    Status = dplyr::recode(
      .data$test_status,
      NON_ESTIMABLE_COMPONENT_FAILURE = "Non-estimable: component failure",
      NON_ESTIMABLE_SEPARATION = "Non-estimable: separation",
      NON_ESTIMABLE_SEPARATION_AND_AR_FAILURE =
        "Non-estimable: separation and AR failure",
      NOT_ACCEPTABLE_FAMILY_SENSITIVITY =
        "Descriptive only: major family sensitivity",
      ACCEPTABLE_WITH_FAMILY_SENSITIVITY_LIMITATION =
        "Descriptive only: family sensitivity",
      NOT_ACCEPTABLE_T_SENSITIVITY_FAILED =
        "Descriptive only: Student-t fit failed"
    ),
    `Ordinary raw p-value` = ifelse(is.na(.data$raw_p_value), "—", "withheld")
  ) |>
  dplyr::select(
    .data$Predictor,
    .data$Component,
    .data$Status,
    .data$`Ordinary raw p-value`
  )

primary_test_table |>
  gt::gt(groupname_col = "Predictor") |>
  compact_gt()
```

The small ordinary likelihood-ratio p-values generated before the separation
gate were retained only in diagnostic provenance. They are **not accepted
results**, are not displayed as association evidence, and do not enter BH.

## Actual-date AR(1) checks

Actual-date AR(1) counterparts were fitted only when the prespecified residual
lag screen triggered and calendar adjacency was adequate. For work/free
occurrence, the fitted $\rho=-0.999964$ is a boundary failure; its likelihood
and coefficients are unusable for inference and do not repair separation. For
positive L10 with previous-night sleep, $\rho=0.836$ and the association
coefficient shifted by only 0.039 Gaussian standard errors; that temporal
sensitivity passed, but it cannot repair the missing occurrence component.

```{r}
#| label: tbl-l10-ar
#| tbl-cap: "Triggered primary near-eye actual-date AR(1) counterparts."

ar_counterparts |>
  dplyr::filter(
    .data$run_id == "primary__near_eye__all_available",
    .data$ar_trigger
  ) |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Component = dplyr::recode(
      .data$component,
      zero_occurrence = "Occurrence",
      positive_magnitude = "Positive magnitude"
    ),
    Rho = sprintf("%.3f", as.numeric(.data$ar_rho)),
    `Shift (primary SE)` = sprintf(
      "%.3f",
      as.numeric(.data$effect_shift_in_primary_se)
    ),
    Disposition = dplyr::recode(
      .data$disposition,
      NOT_ACCEPTABLE_RHO_BOUNDARY = "Not acceptable: rho boundary",
      ACCEPTABLE = "Acceptable temporal sensitivity"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

# Positive-magnitude component: descriptive only

The following ratios condition on L10 being strictly greater than zero. They
exclude zero days by definition and therefore do not answer the overall L10
question. All intervals are model-based **pointwise** 95% intervals; no
simultaneous interval was used.

```{r}
#| label: tbl-l10-primary-positive
#| tbl-cap: "Primary near-eye positive-component estimates. Ratios are conditional on L10 greater than zero."

primary_positive_table <- effects |>
  dplyr::filter(
    .data$run_id == "primary__near_eye__all_available",
    .data$component == "positive_magnitude",
    .data$family %in% c(
      "Gaussian identity on log10-positive L10",
      "Student-t identity on log10-positive L10 sensitivity"
    )
  ) |>
  dplyr::mutate(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Family = ifelse(
      grepl("Student-t", .data$family),
      "Student-t sensitivity",
      "Gaussian component"
    ),
    failed_t = grepl("Student-t", .data$family) &
      grepl("UNRESOLVED", .data$inferential_status),
    `Ratio (95% pointwise interval)` = ratio_ci(
      .data$estimate,
      ifelse(.data$failed_t, NA_real_, .data$lower_95),
      ifelse(.data$failed_t, NA_real_, .data$upper_95)
    ),
    Interpretation = dplyr::case_when(
      .data$predictor_id == "work_free_day" ~
        "Major family sensitivity; direction reverses",
      .data$predictor_id == "activity_status" ~
        "Family and site-influence limitations",
      TRUE ~ "Student-t numerical failure; sensitivity unresolved"
    )
  ) |>
  dplyr::select(
    .data$Predictor,
    .data$Family,
    .data$`Ratio (95% pointwise interval)`,
    .data$Interpretation
  )

primary_positive_table |>
  gt::gt(groupname_col = "Predictor") |>
  compact_gt()
```

For work/free day, the Gaussian ratio was 0.925 (0.759 to 1.128), whereas the
Student-t ratio was 1.061 (0.922 to 1.220): a 1.35-primary-standard-error shift
and a direction reversal. For activity, the response-family shift was 0.61
standard errors, but deleting MPI shifted the Gaussian estimate by 1.04
standard errors. For sleep, the Gaussian ratio per one hour longer previous
night was 0.775 (0.722 to 0.832); the Student-t fit reported false convergence
and a non-positive-definite Hessian, so its nominal interval is withheld and
the family sensitivity remains unresolved.

```{r}
#| label: fig-l10-positive-component-sensitivity
#| fig-cap: "Positive-component response-family and dataset sensitivity. Points are conditional geometric-mean ratios among participant-days with L10 greater than zero. Lines are model-based pointwise 95% intervals; invalid Student-t intervals are withheld. The vertical reference is one. These estimates do not represent the overall L10 association."
#| fig-alt: "Forest plot of Gaussian and Student-t estimates for the conditional geometric-mean L10 ratio among positive participant-days. Six scenarios are shown for each of three day-level contexts. Pointwise 95% intervals are drawn only for numerically acceptable fits; Student-t points without intervals identify failed sensitivity fits. The reference ratio is one. These positive-only estimates do not represent overall L10 associations."
#| out-width: 100%

knitr::include_graphics(artifact(
  "10_figures",
  "H06_daily_l10_metric011_positive_component_sensitivity.png"
))
```

## Influence diagnostics

The prespecified deletion audit fitted 440 positive-Gaussian deletion models
serially with no failures. It did **not** fit occurrence or joint deletion
models because those estimands were already non-estimable.

```{r}
#| label: tbl-l10-influence
#| tbl-cap: "Primary near-eye influence diagnostics for the positive Gaussian component only."

influence |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Participant deletions` = .data$participant_deletions,
    `Maximum participant shift (SE)` = sprintf(
      "%.2f",
      .data$maximum_participant_shift_in_full_se
    ),
    `Maximum site shift (SE)` = sprintf(
      "%.2f",
      .data$maximum_site_shift_in_full_se
    ),
    `Most influential site` = .data$maximum_shift_site,
    Disposition = dplyr::recode(
      .data$influence_disposition,
      ACCEPTABLE_FOR_DESCRIPTIVE_POSITIVE_COMPONENT =
        "Acceptable for descriptive positive component",
      MAJOR_INFLUENCE_LIMITATION_NO_DIRECTIONAL_CLAIM =
        "Major limitation; no directional claim"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

# Regularized occurrence sensitivity

The authorized diagnostic sensitivity used the same fixed-site occurrence
frame, formula, and contrasts with normal priors on the log-odds fixed effects.
The intercept was explicitly assigned a prior and the same prior was assigned
to all remaining conditional fixed effects. The prospective primary scale was
$N(0,3^2)$, with $N(0,1.5^2)$ and $N(0,6^2)$ checks.

All nine primary near-eye MAP fits converged. Predictor directions and
equal-site predicted zero probabilities changed little over the three scales,
but this does not turn a separated ordinary likelihood into an estimable
confirmatory test. The MAP coefficients, odds ratios, and predicted
probabilities are diagnostic only; there is no ordinary confidence interval,
p-value, or BH decision.

```{r}
#| label: tbl-l10-map
#| tbl-cap: "Regularized primary near-eye occurrence diagnostic across prior scales; no significance decision follows."

map_table <- map_coefficients |>
  dplyr::filter(.data$run_id == "primary__near_eye__all_available") |>
  dplyr::left_join(
    map_predictions |>
      dplyr::filter(.data$run_id == "primary__near_eye__all_available") |>
      dplyr::group_by(
        .data$predictor_id,
        .data$prior_standard_deviation
      ) |>
      dplyr::summarise(
        `Equal-site zero probability range` = sprintf(
          "%.3f to %.3f",
          min(.data$equal_site_probability_zero),
          max(.data$equal_site_probability_zero)
        ),
        .groups = "drop"
      ),
    by = c("predictor_id", "prior_standard_deviation")
  ) |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Prior SD` = .data$prior_standard_deviation,
    `MAP log-odds coefficient` = sprintf(
      "%.3f",
      .data$map_log_odds_coefficient
    ),
    `MAP odds ratio` = sprintf("%.3f", .data$map_odds_ratio),
    .data$`Equal-site zero probability range`
  )

map_table |>
  gt::gt(groupname_col = "Predictor") |>
  compact_gt()
```

# Complementary scenarios and gap-timing-unaware comparison

The same two-part contract was completed for chest all-available,
placement-paired/common near-eye and chest, and gap-timing-unaware near-eye and
chest frames. Any failed occurrence component forced the corresponding joint
test to missing. No complementary component estimate replaced a failed
primary slot.

All 12 prespecified primary and gap near-eye joint association/heterogeneity
tests are non-estimable. The gap L10 scientific values were exactly invariant
to METRIC-011; their fits were run only because no accepted production L10
component existed and the previously held branch had to be completed. They are
not described as a METRIC-011 scientific repair.

For the positive Gaussian component alone, the gap-minus-primary shifts were
0.45 standard errors for work/free day, 0.63 for activity status, and 0.68 for
previous-night sleep. These are descriptive conditional comparisons only.

```{r}
#| label: tbl-l10-gap
#| tbl-cap: "Positive-only primary versus gap-timing-unaware near-eye comparison. Joint two-part sensitivity remains non-estimable."

gap_comparison |>
  dplyr::filter(
    .data$family == "Gaussian identity on log10-positive L10"
  ) |>
  dplyr::transmute(
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Primary ratio (95% interval)` = ratio_ci(
      .data$estimate__primary,
      .data$lower_95__primary,
      .data$upper_95__primary
    ),
    `Gap-timing-unaware ratio (95% interval)` = ratio_ci(
      .data$estimate__gap_timing_unaware,
      .data$lower_95__gap_timing_unaware,
      .data$upper_95__gap_timing_unaware
    ),
    `Absolute shift (primary SE)` = sprintf(
      "%.2f",
      .data$absolute_gap_shift_in_primary_se
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

# Multiplicity disposition

L10 is metric slot 3 in each 15-slot primary association,
site-heterogeneity, and gap-timing-unaware family. Its joint raw p-value is
missing because a component failed; the slot is retained and the family size
remains 15. No isolated component p-value or regularized diagnostic result was
substituted.

Within this bounded amendment, only the already frozen MDER slot is otherwise
available. Thus every affected family is currently incomplete at 1 of 15, and
no BH-adjusted p-value or adjusted decision is produced.

```{r}
#| label: tbl-l10-bh
#| tbl-cap: "L10 multiplicity slots. The family denominator is retained despite non-estimability."

bh_slots |>
  dplyr::filter(.data$metric_slot == 3L) |>
  dplyr::transmute(
    Dataset = dplyr::recode(
      .data$dataset_id,
      primary = "Primary",
      gap_timing_unaware = "Gap-timing-unaware"
    ),
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Test = dplyr::recode(
      .data$test_type,
      association = "Association",
      site_heterogeneity = "Site heterogeneity"
    ),
    `L10 slot status` = "Non-estimable: component failure",
    `Raw p-value` = "—",
    `BH-adjusted p-value` = "—",
    `Available / required` = sprintf(
      "%d / %d",
      .data$family_slots_available,
      .data$family_slots_required
    )
  ) |>
  gt::gt(groupname_col = "Dataset") |>
  compact_gt()
```

# Diagnostics, preservation, and compute record

The work was bounded to the L10 branch. The production batch completed 18
scenario-by-predictor analyses in one serial R process. The production-code
pilot projected 1.59 minutes; the complete production process took 156.67
seconds, principally because sealing and the additional MAP diagnostics took
longer than the core fits. No parallel cluster was used. The 50-refit influence
pilot projected 0.40 minutes; the full 440-refit positive-only deletion batch
took 38.76 seconds.

The no-refit coordinator finalizer changed disposition strings only. All 194
serialized fitted-model subobjects were identical before and after that
change. All 294 pre-existing H06_daily scientific/report files in the
protected baseline—including the approved pre-sleep no-nugget frame,
reference, model, report, and manifests—remained byte-for-byte unchanged after
production. No non-L10 model was fitted or revised.

The figure is paired with PNG, PDF, and SVG exports, source-data CSV, alt text,
and a passing final-size readability record. Invalid Student-t intervals are
withheld rather than plotted. Every displayed interval is pointwise.

# Relationship to V0 and the other H06 estimands

V0 did not fit this two-part participant-day L10 model. The approved hourly H06
analysis estimates context associations in repeated hourly melEDI, and the
exploratory temporal GAMM estimates transformed 30-minute profiles over local
clock time. This amendment estimates a different daily-metric estimand with
one participant-day per row. Their coefficients, weighting, uncertainty, and
claims are not numerically interchangeable.

All language here is observational. Recorded work/free day, activity status,
and previous-night sleep duration are associated contexts, not randomized
exposures; no causal effect or health outcome is inferred.

# H06-D-G2P-L10-METRIC011 stop gate

The bounded L10 branch is complete and sealed. The recommended disposition is:

1. accept the primary occurrence and joint two-part L10 association and
   site-heterogeneity tests as **non-estimable**, with raw and adjusted
   p-values missing while retaining the named slot in each 15-metric family;
2. retain Gaussian and Student-t positive-component estimates only as
   explicitly conditional, descriptive diagnostics with the family,
   influence, and numerical limitations stated above;
3. retain the normal-prior fixed-site MAP occurrence models only as diagnostic
   sensitivities, with no ordinary interval, p-value, BH entry, or significance
   decision; and
4. either leave the L10 joint slot missing or explicitly authorize a new
   scientific amendment before changing the primary occurrence hierarchy.

Promoting the registered random-site benchmark, deleting UCR, pooling sites,
changing the fixed-site hierarchy, or adopting the regularized fit as primary
would change the estimand or inferential contract and is **not authorized** by
this gate.

No further L10 or remaining-grid work proceeds without an explicit author
decision at **H06-D-G2P-L10-METRIC011**.
