---
title: "H06 daily metrics — L10 shifted-log pilot"
subtitle: "H01-consistent one-part estimand and temporal stop gate"
date: 2026-08-12
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
.libPaths(c(
  file.path(project_root, "renv/library/macos/R-4.6/aarch64-apple-darwin23"),
  .libPaths()
))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("lme4")), "2.0.1"),
  identical(as.character(packageVersion("glmmTMB")), "1.1.14")
)

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_h06d <- function(stage, file) {
  readr::read_csv(
    artifact(stage, file),
    show_col_types = FALSE,
    na = c("", "NA")
  )
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100),
      source_notes.font.size = gt::px(11)
    )
}
format_ci <- function(estimate, lower, upper, digits = 3L) {
  sprintf(
    paste0("%.", digits, "f (%.", digits, "f to %.", digits, "f)"),
    estimate,
    lower,
    upper
  )
}
format_p <- function(value) {
  ifelse(
    is.na(value),
    "NA",
    ifelse(value < 0.001, "<0.001", sprintf("%.3f", value))
  )
}
yes_no <- function(value) ifelse(value, "Yes", "No")

predictor_labels <- tibble::tribble(
  ~predictor_id, ~predictor_label,
  "work_free_day", "Free day versus Work day",
  "activity_status", "Active versus Sedentary",
  "previous_sleep_duration_centered_h", "+1 h previous-night sleep"
)

samples <- read_h06d(
  "09_tables",
  "H06_daily_l10_shiftlog_pilot_samples.csv"
)
effects <- read_h06d(
  "09_tables",
  "H06_daily_l10_shiftlog_pilot_effects.csv"
)
model_tests <- read_h06d(
  "09_tables",
  "H06_daily_l10_shiftlog_pilot_model_tests.csv"
)
equal_site <- read_h06d(
  "09_tables",
  "H06_daily_l10_shiftlog_pilot_equal_site_estimates.csv"
)
diagnostics <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_diagnostics.csv"
)
zero_mass <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_zero_mass.csv"
)
family_sensitivity <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_family_sensitivity.csv"
)
lag <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_residual_lag.csv"
)
lag_by_site <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_site_residual_lag.csv"
)
post_ar_by_site <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_post_ar_site_residual_lag.csv"
)
ar <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_ar_counterparts.csv"
)
registered <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_registered_benchmark.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_runtime_projection.csv"
)
provenance <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_model_provenance.csv"
)
verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_pilot_verdict.csv"
)
formulas <- read_h06d(
  "09_tables",
  "H06_daily_l10_shiftlog_formula_registry.csv"
)
preservation <- read_h06d(
  "08_diagnostics",
  "H06_daily_l10_shiftlog_historical_preservation_after_pilot.csv"
)

stopifnot(
  nrow(samples) == 3L,
  nrow(effects) == 6L,
  nrow(model_tests) == 6L,
  all(model_tests$multiplicity_status == "PILOT_RAW_ONLY_NO_BH_UPDATE"),
  all(is.na(model_tests$bh_adjusted_p_value)),
  nrow(diagnostics) == 3L,
  nrow(ar) == 3L,
  all(ar$ar_disposition ==
    "NOT_ACCEPTABLE_POST_AR_RESIDUAL_DEPENDENCE"),
  all(verdict$overall_pilot_disposition == "NOT_ACCEPTABLE"),
  nrow(preservation) == 118L,
  all(preservation$preserved_after_pilot),
  nrow(provenance) == 3L,
  all(provenance$reuse_verified),
  !any(provenance$relabel_refit_performed)
)
```

::: {.callout-important}

## Answer in brief

The one-part shifted-log route **does solve the separation problem** that made
the accepted two-part L10 analysis non-estimable: all exact zeros remain in
one outcome, all three fixed-site model hierarchies are estimable, and the
three candidate additive models converge without singularity. It therefore
answers a coherent but different question about the fitted geometric mean of
`L10 mean melEDI + 0.1 lx`.

It does **not** pass the complete H06-daily pilot gate. All three residual-lag
screens triggered the prespecified actual-date AR(1) counterpart. Those AR
fits were numerically sound and had nonboundary correlations, but residual
dependence remained above the accepted limit after correction. Gaussian
residual shape was acceptable only with a limitation, exact zeros comprised
12.5%--13.6% of the fitted samples, activity and sleep had major Student-t
sensitivity, and the sleep model produced a small fraction of negative
individual back-transforms when the 0.1-lx offset was subtracted.

The recommendation is therefore: **DO NOT release the full six-scenario L10
shifted-log batch under the current temporal acceptance rule.** The raw pilot
tests below are diagnostic only (`PILOT_RAW_ONLY_NO_BH_UPDATE`); no
Benjamini--Hochberg field or historical L10 result was replaced. This report
stops at **H06-D-G2P-L10-SHIFTLOG** for an author decision.

:::

# Scope and amended estimand

The pilot implements H06-D-003 / CHG-111 for the primary near-eye,
all-available scenario and the three approved day-level predictors only. Let
$Y$ denote a participant-day's L10 mean melEDI in lx. The response is

$$
Z = \log_{10}\!\left(Y + 0.1\ \mathrm{lx}\right).
$$

An exact zero is retained and maps to $Z=-1$. For a fitted predictor
coefficient $\beta$, $10^\beta$ is a ratio of fitted geometric means of
$Y+0.1$ lx. It is **not** a ratio of unshifted raw L10, an arithmetic-mean
ratio, or the positive-only mean. A reader-scale fitted value is
$10^\eta-0.1$ lx; this report never clips a negative result to zero.

This response architecture is consistent with the H01 use of
`log10(melEDI + 0.1)`, but H01's model adequacy does not transfer. The current
H06-daily frames, predictors, repeated-day structure, zero mass, residuals,
and temporal dependence were audited independently.

The accepted H06-D-001/H06-D-002 two-part record remains correct for its own
occurrence-plus-positive-magnitude estimand and remains byte-for-byte frozen.
The present pilot does not reinterpret or overwrite its separation failure.

# Exact samples and provenance

```{r}
#| label: tbl-shiftlog-samples
#| tbl-cap: "Exact primary near-eye all-available pilot samples."

samples |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Exact zeros` = .data$exact_zeros,
    `Zero mass` = sprintf("%.1f%%", 100 * .data$exact_zero_fraction)
  ) |>
  gt::gt() |>
  compact_gt()
```

Each metric-specific participant-day contributes one row; days are not
weighted by the number of supported hours. The work/free and sleep frames each
contain 784 days from 141 participants at nine sites; the activity frame
contains 734 days from 137 participants at nine sites.

The three additive Gaussian REML objects were not refitted merely to relabel
them. They were reused only after exact verification of the current frame,
transformed response, formula, model matrix, contrasts, family, method,
optimizer, R version, package versions, warnings, convergence, and serialized
object identity. Their parent composite is pinned at SHA-256
`8dede8ff6e506b5dfdd9d7ff37dd675a7ee030c7361c39a16b97c69524ca9419`;
the exact subobject path and object hash for every predictor are recorded in
the model-provenance CSV.

For transparent development provenance, two preliminary compute-cleared
executions stopped before the output-writing stage because of implementation
errors in a row constructor and AR-sequence Boolean operation. A later
preliminary successful execution exposed an inappropriate generic determinant
singularity check for the structured AR covariance. The final run used the
same authorized data, formulas, and model scope with a structured-covariance-
aware diagnostic; only the final outputs are sealed here.

# Exact model hierarchy

The evaluated R objects below show the formulas used for every predictor by
substituting its registered column for `predictor`.

```{r}
#| label: visible-shiftlog-formulas
#| echo: true

fixed_site_reduced <- response_value ~ site + (1 | participant_key)
fixed_site_additive <- response_value ~ site + predictor +
  (1 | participant_key)
fixed_site_heterogeneity <- response_value ~ site * predictor +
  (1 | participant_key)
registered_benchmark <- response_value ~ predictor +
  (predictor | site) + (1 | site:participant_key)
triggered_actual_date_ar <- response_value ~ site + predictor +
  (1 | participant_key) +
  ar1(day_index_factor + 0 | day_sequence_id)

fixed_site_reduced
fixed_site_additive
fixed_site_heterogeneity
registered_benchmark
triggered_actual_date_ar
```

Site uses sum-to-zero contrasts. The reduced-versus-additive association and
additive-versus-heterogeneity comparisons use maximum likelihood on identical
rows; the selected additive Gaussian estimates use REML and equal-site
marginalization. The Student-t additive model is a distributional sensitivity.
The registered random-site/random-slope model remains a benchmark and is never
promoted when the fixed-site route fails. The actual-date AR counterpart is
fitted only after the existing lag trigger and starts a new sequence across
every missing calendar date.

# Pilot estimates and raw-only comparisons

```{r}
#| label: tbl-shiftlog-effects
#| tbl-cap: "Shifted-scale predictor contrasts with pointwise 95% confidence intervals."

effects |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::mutate(
    Model = ifelse(
      grepl("^Gaussian", .data$family),
      "Gaussian candidate",
      "Student-t sensitivity"
    ),
    `Shifted geometric-mean ratio (95% CI)` = format_ci(
      .data$shifted_geometric_mean_ratio,
      .data$ratio_lower_95,
      .data$ratio_upper_95
    )
  ) |>
  dplyr::select(
    Predictor = "predictor_label",
    "Model",
    "Shifted geometric-mean ratio (95% CI)"
  ) |>
  gt::gt(groupname_col = "Model") |>
  gt::tab_source_note(
    gt::md(
      "Ratios refer to fitted geometric means of L10 mean melEDI + 0.1 lx. Intervals are pointwise, not simultaneous."
    )
  ) |>
  compact_gt()
```

The Gaussian candidate ratios are 0.988 (0.918 to 1.065) for Free versus Work
day, 0.929 (0.845 to 1.021) for Active versus Sedentary, and 0.899 (0.876 to
0.923) per one-hour greater previous-night sleep duration. These are
observational associations on the shifted scale, not causal effects.

```{r}
#| label: tbl-equal-site-estimates
#| tbl-cap: "Equal-site typical fitted L10 values after the unclipped inverse transform."

equal_site |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    Target = .data$target,
    `Typical fitted value, lx (95% CI)` = format_ci(
      .data$equal_site_raw_back_transform_lx,
      .data$raw_back_transform_lower_95_lx,
      .data$raw_back_transform_upper_95_lx
    )
  ) |>
  gt::gt(groupname_col = "Predictor") |>
  gt::tab_source_note(
    gt::md(
      "Values use equal site weight and the unclipped transform 10^eta - 0.1 lx. They are geometric-mean-like typical values, not arithmetic means."
    )
  ) |>
  compact_gt()
```

```{r}
#| label: tbl-raw-pilot-tests
#| tbl-cap: "Raw pilot model-comparison p-values; no multiplicity update or decision."

model_tests |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    Test = dplyr::recode(
      .data$test_type,
      association = "Common association",
      site_heterogeneity = "Predictor-by-site heterogeneity"
    ),
    `Likelihood ratio` = sprintf("%.2f", .data$likelihood_ratio),
    df = .data$degrees_of_freedom,
    `Raw p` = format_p(.data$raw_p_value),
    Status = .data$multiplicity_status
  ) |>
  gt::gt(groupname_col = "Predictor") |>
  gt::tab_source_note(
    gt::md(
      "These values are diagnostic pilot outputs. No value is bolded, no Benjamini–Hochberg value was calculated, and no significance decision follows."
    )
  ) |>
  compact_gt()
```

The raw tests are shown because they confirm that the one-part hierarchy is
estimable. They must not be transferred into L10 slot 3 unless a later full
batch is explicitly authorized and passes its complete gates.

# Diagnostic assessment

## Gaussian residuals, zeros, and back-transform bounds

```{r}
#| label: tbl-gaussian-diagnostics
#| tbl-cap: "Per-predictor Gaussian diagnostics and explicit assessment."

diagnostics |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    `Q–Q correlation` = sprintf("%.3f", .data$residual_qq_correlation),
    `Variance ratio` = sprintf("%.2f", .data$residual_variance_ratio),
    `|residual|–fitted rho` = sprintf(
      "%.3f",
      .data$absolute_residual_fitted_spearman
    ),
    `|standardized residual| > 4` = sprintf(
      "%.2f%%",
      100 * .data$standardized_residual_gt4_fraction
    ),
    `Exact-zero mass` = sprintf("%.1f%%", 100 * .data$exact_zero_fraction),
    `Minimum inverse transform` = sprintf(
      "%.3f lx",
      .data$raw_back_transform_minimum_lx
    ),
    Assessment = ifelse(
      .data$bound_disposition == "ACCEPTABLE",
      "Acceptable only with Gaussian residual and zero-mass limitations",
      paste0(
        "Acceptable only with Gaussian residual, zero-mass, and negative ",
        "back-transform limitations"
      )
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

All three numerical hierarchies converge with positive-definite Hessians and
non-singular additive fits. Their Q--Q correlations are 0.976--0.979, but
heteroscedasticity diagnostics give variance ratios of 6.90--7.72 and
absolute-residual/fitted correlations of 0.337--0.344. Fewer than 0.6% of rows
have an absolute standardized residual above 4, so the predeclared severe-tail
rule is not crossed. The Gaussian working distribution is therefore
acceptable only with an explicit residual limitation.

Zeros are retained, as required, but form a discrete mass at $Z=-1$ in
12.5%--13.6% of rows. A shifted-log transformation can include those rows; it
does not make the mass Gaussian. This remains an open scientific limitation
and cannot be resolved by citing H01.

For the sleep frame, 2.30% of individual fitted points become slightly
negative after subtracting the 0.1-lx offset; the minimum is -0.018 lx and none
is below -0.05 lx. The equal-site target estimates shown above remain positive.
Negative fitted values are disclosed and never clipped.

## Student-t sensitivity

```{r}
#| label: tbl-family-sensitivity
#| tbl-cap: "Student-t effect shifts relative to the Gaussian candidate."

family_sensitivity |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    `Shift in Gaussian SE` = sprintf(
      "%.2f",
      .data$effect_shift_in_gaussian_se
    ),
    `Direction reversal` = yes_no(.data$direction_reversal),
    Assessment = dplyr::recode(
      .data$family_sensitivity_disposition,
      ACCEPTABLE_ONLY_WITH_FAMILY_SENSITIVITY_LIMITATION =
        "Acceptable with limitation",
      ACCEPTABLE_ONLY_WITH_MAJOR_FAMILY_SENSITIVITY_LIMITATION =
        "Acceptable only with major limitation"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

No direction reverses, but activity and sleep shift by 1.28 and 1.23 Gaussian
standard errors, respectively. Both therefore meet the existing major-
limitation rule. Work/free shifts by 0.58 standard errors and has an ordinary
family-sensitivity limitation. The Student-t fit is not a significance-selected
replacement.

## Actual-date temporal dependence

```{r}
#| label: tbl-ar-diagnostics
#| tbl-cap: "Triggered actual-date AR(1) counterparts and post-correction residual checks."

lag |>
  dplyr::select(
    "predictor_order",
    "predictor_id",
    pre_ar_lag = "residual_lag1",
    pre_ar_max_site = "maximum_absolute_site_lag1",
    "adjacent_pairs",
    "participants"
  ) |>
  dplyr::left_join(ar, by = c("predictor_order", "predictor_id")) |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    `True adjacent pairs` = .data$adjacent_pairs,
    `Pre-AR pooled lag` = sprintf("%.3f", .data$pre_ar_lag),
    `Pre-AR max site |lag|` = sprintf("%.3f", .data$pre_ar_max_site),
    rho = sprintf("%.3f", .data$ar_rho),
    `Post-AR pooled lag` = sprintf("%.3f", .data$post_ar_residual_lag1),
    `Post-AR max site |lag|` = sprintf(
      "%.3f",
      .data$post_ar_maximum_absolute_site_lag1
    ),
    `Effect shift, SE` = sprintf(
      "%.2f",
      .data$ar_effect_shift_in_gaussian_se
    ),
    Assessment = "Not acceptable: residual dependence persists"
  ) |>
  gt::gt() |>
  compact_gt()
```

The pooled pre-AR correlations are small (-0.095 to -0.097), but each screen
triggers because at least one site exceeds the prespecified absolute 0.30
threshold. MPI supplies the maximum in all three frames with 115--119 adjacent
pairs from 26 participants, so the trigger is not driven by a single pair or
participant.

All independent-engine calibrations and AR models converge with positive-
definite Hessians. The AR models have minimum random-effect SDs of
0.124--0.131, minimum scaled covariance eigenvalues of 0.083--0.097, and
$\rho=0.576$--0.609. The structured covariance is therefore not singular and
the correlation is not at the 0.95 boundary. The previous generic determinant
check was inappropriate because a covariance determinant shrinks with block
dimension even when the scaled eigenvalues are well separated from zero.

The failure is scientific rather than numerical: after AR correction, pooled
residual lag is -0.273 to -0.285 and maximum site absolute lag is 0.479--0.500.
All three exceed the accepted post-AR limits. Predictor estimates shift by at
most 0.31 Gaussian standard errors, but effect stability cannot override the
failed residual-dependence gate.

## Registered benchmark

```{r}
#| label: tbl-registered-benchmark
#| tbl-cap: "Exact registered random-site/random-slope benchmark status."

registered |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    Converged = yes_no(.data$converged),
    Singular = yes_no(.data$singular),
    `Maximum |random-site correlation|` = sprintf(
      "%.3f",
      .data$maximum_absolute_random_site_correlation
    ),
    Assessment = dplyr::recode(
      .data$disposition,
      ESTIMABLE_BENCHMARK_ONLY = "Estimable benchmark only",
      NON_ESTIMABLE_BENCHMARK = "Non-estimable benchmark"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

The work/free and activity benchmarks are estimable but remain benchmarks
only. The sleep benchmark is singular with a random-site correlation of 1.000
and is non-estimable. None is promoted into the fixed-site hierarchy.

# Per-predictor gate verdict

```{r}
#| label: tbl-pilot-verdict
#| tbl-cap: "Complete per-predictor pilot assessment."

verdict |>
  dplyr::left_join(predictor_labels, by = "predictor_id") |>
  dplyr::transmute(
    Predictor = .data$predictor_label,
    Numerical = "Acceptable",
    Residual = "Acceptable only with limitation",
    `Student-t` = ifelse(
      grepl("MAJOR", .data$family_sensitivity_disposition),
      "Major limitation",
      "Limitation"
    ),
    Temporal = "Not acceptable after triggered AR(1)",
    Overall = "NOT ACCEPTABLE",
    Recommendation = "Do not release full L10 shifted-log batch"
  ) |>
  gt::gt() |>
  compact_gt()
```

# Runtime, preservation, and bounded next step

The final successful pilot took `r sprintf("%.2f", runtime$pilot_wall_seconds)`
seconds of wall time. The three exact additive REML objects contributed zero
refit time; the remaining pilot models used
`r sprintf("%.2f", runtime$newly_fitted_model_seconds)` seconds in aggregate.
A deliberately conservative projection for all 18 predictor-by-scenario
analyses plus 440 later influence refits is
`r sprintf("%.1f", runtime$projected_full_seconds)` seconds
(`r sprintf("%.2f", runtime$projected_full_minutes)` minutes). This is
computationally bounded, but scientific authorization—not runtime—is the
controlling barrier.

All 118 H06-D-001/H06-D-002 historical report, model, diagnostic, manifest,
transition, and closure entries remain byte-for-byte identical. The pilot ran
under R 4.6.1 with the synchronized project library, serially in one process.
It ran no BH adjustment, deletion batch, bootstrap, simulation, remaining
daily-metric grid, main-H06 analysis, or shared-preparation operation.

## H06-D-G2P-L10-SHIFTLOG author gate

The recommended disposition is to accept the pilot verdict and stop this
shifted-log route under the current temporal rule. That preserves the accepted
two-part failure as the historical result and leaves L10 slot 3 missing.

If consistency with H01 is judged more important than the current H06-daily
post-AR residual threshold, moving forward would require a new explicit author
amendment. It would need to state the revised temporal estimand or acceptance
rule and how remaining residual dependence is handled; the present gate does
not authorize silently ignoring it. Only after that decision could the later
L10-only scope be considered: the same three predictors across primary
near-eye/chest all-available, paired/common near-eye/chest, and gap-timing-
unaware near-eye/chest scenarios, with Student-t, triggered AR, influence,
exact samples, and the dependent 15-slot BH fields.

Until such an amendment is explicit, **the full shifted-log batch, all BH
updates, the remaining daily production grid, Stage 3/4, main H06, and final
H06 version selection remain frozen.**
