---
title: "H06 daily metrics — remaining non-L10 bounded pilot"
subtitle: "Current-source reuse, clock-family diagnostics, and production runtime gate"
date: 2026-08-12
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
.libPaths(c(
  file.path(project_root, "renv/library/macos/R-4.6/aarch64-apple-darwin23"),
  .libPaths()
))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("lme4")), "2.0.1"),
  identical(as.character(packageVersion("glmmTMB")), "1.1.14")
)

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_h06d <- function(stage, file) {
  readr::read_csv(
    artifact(stage, file),
    show_col_types = FALSE,
    na = c("", "NA")
  )
}
sha256 <- function(path) digest::digest(path, algo = "sha256", file = TRUE)
verify_relative_manifest <- function(manifest, hash_column = "sha256") {
  paths <- file.path(project_root, manifest$relative_path)
  stopifnot(all(file.exists(paths)))
  actual <- unname(vapply(paths, sha256, character(1)))
  stopifnot(identical(actual, manifest[[hash_column]]))
  invisible(TRUE)
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100),
      source_notes.font.size = gt::px(11)
    )
}
format_seconds <- function(value) {
  dplyr::case_when(
    is.na(value) ~ "—",
    value < 60 ~ sprintf("%.1f s", value),
    TRUE ~ sprintf("%.1f min", value / 60)
  )
}
format_number <- function(value, digits = 3L) {
  ifelse(is.na(value), "—", sprintf(paste0("%.", digits, "f"), value))
}

input_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_non_l10_pilot_input_manifest.csv"
)
code_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_non_l10_pilot_code_manifest.csv"
)
output_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_non_l10_pilot_output_manifest.csv"
)
software_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_non_l10_pilot_software_manifest.csv"
)
stopifnot(
  all(input_manifest$verification_status == "PASS"),
  identical(input_manifest$expected_sha256, input_manifest$actual_sha256)
)
verify_relative_manifest(code_manifest)
verify_relative_manifest(output_manifest)

frame_inventory <- read_h06d(
  "06_model_data",
  "H06_daily_non_l10_pilot_frame_inventory.csv"
)
reconciliation <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_source_reconciliation.csv"
)
reuse <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_reuse_verification.csv"
)
diagnostics <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_timing_diagnostics.csv"
)
model_counts <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_timing_model_counts.csv"
)
registered <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_timing_registered_benchmarks.csv"
)
deletions <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_deletion_checkpoint.csv"
)
deletion_runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_deletion_runtime_by_class.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_runtime_and_projection.csv"
)
preservation <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_preservation_final.csv"
)
verdict <- read_h06d(
  "08_diagnostics",
  "H06_daily_non_l10_pilot_verdict.csv"
)

stopifnot(
  nrow(frame_inventory) == 468L,
  all(reconciliation$exact_reconciliation),
  nrow(reuse) == 3L,
  all(
    reuse$reuse_disposition ==
      "REUSE_EXACT_HISTORICAL_REML_ADDITIVE_OBJECT"
  ),
  nrow(diagnostics) == 15L,
  nrow(deletions) == 50L,
  all(preservation$identity_status == "BYTE_IDENTICAL"),
  verdict$bh_update[[1L]] == "NOT_RUN_NOT_AUTHORIZED",
  verdict$full_grid[[1L]] == "NOT_RUN_AWAITING_AUTHOR_APPROVAL"
)

predictor_labels <- c(
  work_free_day = "Work/free day",
  activity_status = "Daily activity status",
  previous_sleep_duration_centered_h = "Previous-night sleep duration"
)
class_labels <- c(
  gaussian_offset = "Shifted-log Gaussian",
  tweedie_log = "Tweedie log",
  gaussian_identity = "Identity Gaussian",
  ordinary_clock = "Ordinary clock-hour Gaussian",
  strict_clock = "Strict-after-16:00 clock-hour Gaussian"
)

metric_verdict <- diagnostics |>
  dplyr::summarise(
    predictor_cells = dplyr::n(),
    acceptable = sum(grepl("^ACCEPTABLE", .data$diagnostic_disposition)),
    gaussian_core_failures = sum(
      .data$diagnostic_disposition == "NOT_ACCEPTABLE_GAUSSIAN_CORE"
    ),
    temporal_failures = sum(
      .data$diagnostic_disposition == "NOT_ACCEPTABLE_TEMPORAL_STABILITY"
    ),
    ar_triggers = sum(.data$ar_trigger),
    .by = c("metric_slot", "manuscript_name")
  ) |>
  dplyr::arrange(.data$metric_slot)

deletion_summary <- deletions |>
  dplyr::summarise(
    refits = dplyr::n(),
    failures = sum(!.data$converged | !is.na(.data$fit_error)),
    warnings = sum(.data$warning_count),
    maximum_shift_se = max(.data$absolute_shift_in_full_se, na.rm = TRUE),
    direction_reversals = sum(.data$direction_reversal, na.rm = TRUE),
    .by = "class_id"
  )

all_deletion_projection <- runtime |>
  dplyr::filter(grepl("all_authorized_scenarios", .data$component))
base_projection_seconds <- runtime |>
  dplyr::filter(grepl("^full base-grid projection", .data$component)) |>
  dplyr::summarise(seconds = sum(.data$projected_seconds)) |>
  dplyr::pull("seconds")
```

::: {.callout-important title="H06-D-G2P-NONL10 — author stop gate"}
The bounded pilot is complete, but the remaining daily grid has **not** been
released for production. The source and preservation contracts pass, all 50
representative deletion refits complete, and all three predictor models for
mean timing above 250 lx melEDI pass the declared diagnostics. M10 midpoint,
L10 midpoint, first timing above 250 lx melEDI, and last timing above 250 lx
melEDI do not pass the current response-family or temporal-stability rules.
No BH family was assembled and no pilot association p-value is interpreted.
:::

# What this pilot did

This is the bounded pilot authorized by `H06-D-007`. It concerns the 13
unfinished **non-L10-mean, non-MDER** daily registry slots. It does not refit
L10 mean melEDI (slot 3), MDER (slot 15), the accepted 30-minute temporal
analysis, the pre-sleep no-nugget sensitivity, or the main hourly H06
analysis. The hourly H06 analysis remains primary; this branch remains
complementary, preregistration-oriented participant-day evidence.

The pilot performed four tasks:

1. verified the current shared inputs and the METRIC-010/METRIC-011 non-L10
   invariance evidence;
2. constructed the full candidate inventory without fitting the production
   grid;
3. fitted the five timing outcomes × three predictors on primary near-eye
   all-available frames; and
4. ran exactly 50 serial participant/site deletion refits across five
   consequential response-family classes.

There was no bootstrap, simulation, full production batch, BH update, or
reader-report work.

# Current-source and preservation contract

All `r nrow(input_manifest)` pinned inputs match their sealed SHA-256
identities. The inventory contains `r nrow(frame_inventory)` exact candidate
frames: 13 metrics × 3 predictors × 12 declared scenario/placement/sample
roles. Frame sizes range from `r min(frame_inventory$participant_days)` to
`r max(frame_inventory$participant_days)` participant-days.

The long CSV and wide RDS metric representations agree in missingness,
estimability, and values under the project numerical-representation contract.
They are not bit-identical after decimal parsing in
`r sum(reconciliation$representation_differences)` cells; the maximum scaled
difference is `r format(max(reconciliation$maximum_relative_difference), scientific = TRUE, digits = 3)`,
far below the prespecified $10^{-12}$ representation tolerance. This is a
serialization distinction, not a scientific-value difference.

```{r}
#| label: tbl-frame-inventory
#| tbl-cap: "Exact candidate-frame inventory by scenario role."

frame_inventory |>
  dplyr::summarise(
    `Metric × predictor frames` = dplyr::n(),
    `Participant-days, range` = sprintf(
      "%d–%d",
      min(.data$participant_days),
      max(.data$participant_days)
    ),
    `Participants, range` = sprintf(
      "%d–%d",
      min(.data$participants),
      max(.data$participants)
    ),
    Sites = paste(range(.data$sites), collapse = "–"),
    .by = c("run_order", "run_id", "analysis_role")
  ) |>
  dplyr::arrange(.data$run_order) |>
  dplyr::select(-"run_order") |>
  gt::gt() |>
  gt::cols_label(
    run_id = "Scenario",
    analysis_role = "Role"
  ) |>
  compact_gt()
```

The current state-support reconstruction qualification under
`PREP06-BASE-002` remains a provenance limitation. It is not treated as
evidence that these values are wrong.

All `r nrow(preservation)` pre-existing H06_daily files in the protected
scope remained byte-identical across the pilot. That includes the accepted
L10 and MDER histories, temporal outputs, and pre-sleep no-nugget object.

## Exact reuse of three earlier family pilots

The three representative additive objects were reused rather than refitted.
Their analytical frames, response/source vectors, fitted model frames,
formulas, design matrices, site contrasts, family/links, optimizers, warnings,
convergence states, and software identities are exact. The raw tibble object
is not byte-identical because the current wide source carries an updated
global `metric_settings` attribute after the shared MDER/L10 repins; removing
that unrelated global attribute leaves the complete analytical frame
identical. This exception is recorded rather than hidden.

```{r}
#| label: tbl-reuse
#| tbl-cap: "Exact analytical-object reuse verification."

reuse |>
  dplyr::transmute(
    Family = unname(class_labels[.data$reuse_id]),
    Metric = .data$metric_id,
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Days / participants / sites` = sprintf(
      "%d / %d / %d",
      .data$participant_days,
      .data$participants,
      .data$sites
    ),
    `Raw frame object` = ifelse(
      .data$raw_frame_object_identical,
      "Identical",
      "Global metric-settings attribute repinned"
    ),
    `Analytical model contract` = ifelse(
      .data$reuse_disposition ==
        "REUSE_EXACT_HISTORICAL_REML_ADDITIVE_OBJECT",
      "Exact — reused",
      "Failed"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

# Clock-family pilot

The clock outcomes use their already-approved linearized local-clock
estimands. M10 midpoint and the three above-250-lx timing metrics use ordinary
clock hours. L10 midpoint uses the strict after-16:00 unwrap. All 15 source
frames pass the source-clock span and cut-point rules; failures below arise
from response-distribution or residual temporal criteria, not an unnoticed
midnight split.

```{r}
#| label: tbl-clock-metric-verdict
#| tbl-cap: "Timing-family diagnostic verdict across the three predictors."

metric_verdict |>
  dplyr::transmute(
    Slot = .data$metric_slot,
    Outcome = .data$manuscript_name,
    `Predictor cells` = .data$predictor_cells,
    Acceptable = .data$acceptable,
    `Gaussian-core failure` = .data$gaussian_core_failures,
    `Temporal-stability failure` = .data$temporal_failures,
    `AR triggers` = .data$ar_triggers
  ) |>
  gt::gt() |>
  compact_gt()
```

Mean timing above 250 lx melEDI is acceptable for all three predictors. M10
midpoint and last timing above 250 lx melEDI pass the ordinary Gaussian checks
but their triggered AR counterparts leave excessive site-specific residual
dependence. L10 midpoint fails Gaussian residual-shape/tail rules in all three
cells; its activity cell also fails the AR criterion. First timing has two
Gaussian-core failures and one temporal-stability failure.

```{r}
#| label: tbl-clock-cell-diagnostics
#| tbl-cap: "Per-predictor timing diagnostics on primary near-eye frames."

diagnostics |>
  dplyr::transmute(
    Outcome = .data$manuscript_name,
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Days = .data$participant_days,
    `Q–Q r` = format_number(.data$residual_qq_correlation),
    `|spread trend|` = format_number(
      abs(.data$absolute_residual_fitted_spearman)
    ),
    `Tail > |4|` = sprintf(
      "%.1f%%",
      100 * .data$standardized_residual_gt4_fraction
    ),
    `Max site lag-1` = format_number(
      .data$independent_maximum_absolute_site_lag1
    ),
    `AR rho` = format_number(.data$ar_rho),
    `Post-AR max site lag-1` = format_number(
      .data$post_ar_maximum_absolute_site_lag1
    ),
    Verdict = dplyr::recode(
      .data$diagnostic_disposition,
      ACCEPTABLE = "Acceptable",
      NOT_ACCEPTABLE_GAUSSIAN_CORE = "Not acceptable — Gaussian core",
      NOT_ACCEPTABLE_TEMPORAL_STABILITY =
        "Not acceptable — temporal stability"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

The 15 cells attempted and fitted
`r sum(model_counts$model_components_fitted)` model components. Nine cells
triggered the actual-date, gap-aware AR counterpart; none of those nine met
the post-AR site-specific residual rule, and one was singular. The exact
registered random-site/random-slope benchmark was estimable in only
`r sum(registered$benchmark_disposition == "ESTIMABLE_BENCHMARK_ONLY")` of 15
cells. It therefore remains a benchmark rather than a substitute for the
approved fixed-site primary hierarchy.

The pipeline stored 30 reduced-versus-additive or additive-versus-site-
interaction pilot likelihood-ratio tests solely to verify production code.
They are raw pilot values: they are not displayed, interpreted, adjusted, or
used to select a model. No BH field exists in that table.

# Fifty-refit deletion pilot

All 50 scheduled refits completed serially: five participant and five site
deletions in each of five response-family classes. There were no fit failures,
optimizer warnings, or direction reversals. The strict-clock representative
is the already non-acceptable L10-midpoint work/free fit and is included only
to measure runtime; its deletion result cannot rescue its failed Gaussian
diagnostics.

```{r}
#| label: tbl-deletion-pilot
#| tbl-cap: "Representative participant/site deletion pilot."

deletion_summary |>
  dplyr::transmute(
    Class = unname(class_labels[.data$class_id]),
    Refits = .data$refits,
    Failures = .data$failures,
    Warnings = .data$warnings,
    `Maximum shift (full-model SE)` = sprintf(
      "%.2f",
      .data$maximum_shift_se
    ),
    `Direction reversals` = .data$direction_reversals,
    Assessment = dplyr::case_when(
      .data$class_id == "strict_clock" ~
        "1–2 SE major limitation; runtime-only failed family",
      .data$maximum_shift_se < 1 ~ "Acceptable in this representative pilot",
      TRUE ~ "Requires review"
    )
  ) |>
  gt::gt() |>
  compact_gt()
```

The largest shift among the four otherwise usable representative classes is
`r sprintf("%.2f", max(deletion_summary$maximum_shift_se[deletion_summary$class_id != "strict_clock"]))`
full-model SE. The strict-clock runtime representative shifts by
`r sprintf("%.2f", deletion_summary$maximum_shift_se[deletion_summary$class_id == "strict_clock"])`
SE under its largest tested deletion, reinforcing that it must not be promoted
from this pilot.

# Runtime and production boundary

The 15-cell timing pilot used `r format_seconds(runtime$measured_seconds[runtime$component == "15-cell primary near-eye clock-family pilot"])`.
The 50 fitted deletion models used a cumulative
`r format_seconds(runtime$measured_seconds[runtime$component == "five-class deletion pilot"])`.
Checkpoint output was written after every refit and round-tripped with all
task IDs 1–50.

Across all 468 candidate frames, a complete participant/site deletion battery
would require `r format(all_deletion_projection$projected_units, big.mark = ",")`
serial refits and is projected at
`r format_seconds(all_deletion_projection$projected_seconds)`. The base model
grid is projected at `r format_seconds(base_projection_seconds)`. Together
these imply a lower-bound fit time of
`r format_seconds(all_deletion_projection$projected_seconds + base_projection_seconds)`,
before triggered-AR overhead, diagnostics, source tables, and rendering.

```{r}
#| label: tbl-runtime
#| tbl-cap: "Measured pilot time and projected full-grid computation."

runtime |>
  dplyr::filter(
    .data$component %in% c(
      "15-cell primary near-eye clock-family pilot",
      "five-class deletion pilot"
    ) |
      grepl("all_authorized_scenarios", .data$component) |
      grepl("^full base-grid projection", .data$component)
  ) |>
  dplyr::transmute(
    Component = .data$component,
    `Completed units` = ifelse(
      .data$completed_units == 0,
      "—",
      format(.data$completed_units, big.mark = ",")
    ),
    `Measured time` = format_seconds(.data$measured_seconds),
    `Projected units` = ifelse(
      is.na(.data$projected_units),
      "—",
      format(.data$projected_units, big.mark = ",")
    ),
    `Projected time` = format_seconds(.data$projected_seconds),
    Scope = .data$projection_scope
  ) |>
  gt::gt() |>
  compact_gt()
```

# Recommendation and H06-D-G2P-NONL10 decisions

The recommended disposition is to **accept the bounded pilot but not release
the unchanged full 13-slot production batch**.

1. Accept the current-source contract, 468-frame inventory, exact reuse of the
   three historical analytical objects, and byte-for-byte preservation audit.
2. Accept mean timing above 250 lx melEDI as the only clock-family outcome
   currently cleared across all three predictor cells.
3. Keep M10 midpoint, L10 midpoint, first timing, and last timing on hold. A
   new bounded, prespecified repair pilot is needed before any of these can
   enter production; failed AR or Gaussian diagnostics must not be repaired by
   deletion, predictor-specific model selection, or an unregistered family.
4. Accept the 50-refit checkpoint/runtime evidence. The strict-clock 1.10-SE
   deletion shift is a major limitation attached to an already failed model,
   not an accepted influence result.
5. Keep BH unchanged. L10 mean remains the named missing slot 3, MDER remains
   frozen in slot 15, and all 12 completed families remain blocked until the
   production grid itself is authorized and complete.

Two scientifically coherent next actions would require explicit author
authorization: either (a) authorize production for the eight non-timing
outcomes plus mean timing while the four failed timing outcomes remain on a
separate repair hold, or (b) first authorize a bounded repair pilot for the
four failed timing outcomes and decide production only after that result. The
second route preserves a single production gate and is recommended.

::: {.callout-warning title="Stop"}
This report stops at `H06-D-G2P-NONL10`. It does not authorize the full
remaining grid, any BH update, Stage 3, Stage 4, changes to main H06, or final
H06 integration.
:::

# Reproducibility record

The authoritative command was:

```text
Rscript --vanilla scripts/hypotheses/H06_daily/run_h06_daily_non_l10_pilot.R
Rscript --vanilla scripts/hypotheses/H06_daily/postprocess_h06_daily_non_l10_pilot.R
```

R 4.6.1 and the synchronized project library were used. The input manifest
contains `r nrow(input_manifest)` identities, the code manifest contains
`r nrow(code_manifest)` identities, and the non-circular output manifest
contains `r nrow(output_manifest)` identities. The pilot report reads sealed
artifacts only and performs no model fit, diagnostic recomputation, deletion,
or multiplicity adjustment.

```{r}
#| label: tbl-software
#| tbl-cap: "Authoritative software identity."

software_manifest |>
  dplyr::select(Item = "item", Version = "version", Role = "role") |>
  gt::gt() |>
  compact_gt()
```
