---
title: "H06 daily metrics — bounded timing-repair pilot"
subtitle: "Participant-cluster HC3 candidate route and diagnostic sensitivities"
date: 2026-08-12
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H06_DAILY_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
.libPaths(c(
  file.path(project_root, "renv/library/macos/R-4.6/aarch64-apple-darwin23"),
  .libPaths()
))

library(digest)
library(dplyr)
library(gt)
library(knitr)
library(readr)
library(tibble)
library(tidyr)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  identical(as.character(packageVersion("sandwich")), "3.1.1"),
  identical(as.character(packageVersion("glmmTMB")), "1.1.14")
)

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_h06d <- function(stage, file) {
  readr::read_csv(
    artifact(stage, file),
    show_col_types = FALSE,
    na = c("", "NA")
  )
}
sha256 <- function(path) digest::digest(path, algo = "sha256", file = TRUE)
verify_relative_manifest <- function(manifest, hash_column = "sha256") {
  paths <- file.path(project_root, manifest$relative_path)
  stopifnot(all(file.exists(paths)))
  actual <- unname(vapply(paths, sha256, character(1)))
  stopifnot(identical(actual, manifest[[hash_column]]))
  invisible(TRUE)
}
compact_gt <- function(table) {
  table |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100),
      source_notes.font.size = gt::px(11)
    )
}
format_p <- function(value) {
  dplyr::case_when(
    is.na(value) ~ "—",
    value < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", value)
  )
}
format_class <- function(value) {
  dplyr::recode(
    value,
    STABLE = "Stable (<1 SE, no reversal)",
    SUBSTANTIAL_LIMITATION = "Major limitation (1–<2 SE)",
    UNSTABLE = "Unstable (reversal or ≥2 SE)",
    UNRESOLVED_NUMERICAL_FAILURE = "Unresolved — fit did not converge",
    UNRESOLVED_TERM_MISMATCH = "Unresolved — term mismatch"
  )
}

input_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_timing_repair_input_manifest.csv"
)
pipeline_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_timing_repair_pipeline_manifest.csv"
)
figure_manifest <- read_h06d(
  "12_manifests",
  "H06_daily_timing_repair_figure_manifest.csv"
)
stopifnot(
  all(input_manifest$verification_status == "PASS"),
  identical(input_manifest$expected_sha256, input_manifest$actual_sha256)
)
verify_relative_manifest(pipeline_manifest)
verify_relative_manifest(figure_manifest)

frame_contract <- read_h06d(
  "06_model_data",
  "H06_daily_timing_repair_frame_contract.csv"
)
candidate_models <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_candidate_models.csv"
)
candidate_cells <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_candidate_cells.csv"
)
student <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_student_t.csv"
)
ar <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_no_nugget_ar.csv"
)
sensitivity <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_sensitivity_stability.csv"
)
residual_summary <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_residual_plot_summary.csv"
)
preservation <- read_h06d(
  "08_diagnostics",
  "H06_daily_timing_repair_preservation_final.csv"
)
effects <- read_h06d(
  "09_tables",
  "H06_daily_timing_repair_effects.csv"
)
tests <- read_h06d(
  "09_tables",
  "H06_daily_timing_repair_raw_tests.csv"
)
verdict <- read_h06d(
  "09_tables",
  "H06_daily_timing_repair_outcome_verdict.csv"
)
runtime <- read_h06d(
  "09_tables",
  "H06_daily_timing_repair_runtime.csv"
)

stopifnot(
  nrow(frame_contract) == 12L,
  all(frame_contract$frame_verification_status == "PASS"),
  nrow(candidate_models) == 36L,
  all(candidate_models$candidate_model_gate_pass),
  all(candidate_models$positive_semidefinite),
  nrow(candidate_cells) == 12L,
  all(candidate_cells$candidate_gate_pass),
  nrow(student) == 36L,
  all(student$converged),
  nrow(ar) == 36L,
  nrow(sensitivity) == 48L,
  nrow(residual_summary) == 12L,
  nrow(effects) == 12L,
  nrow(tests) == 24L,
  all(tests$multiplicity_status == "PILOT_RAW_ONLY_NO_BH_UPDATE"),
  all(is.na(tests$adjusted_p_value)),
  nrow(verdict) == 4L,
  all(verdict$all_three_predictors_candidate_acceptable),
  nrow(preservation) == 571L,
  all(preservation$identity_status == "BYTE_IDENTICAL")
)

metric_labels <- c(
  m10_midpoint = "Midpoint of the brightest 10 hours",
  l10_midpoint = "Midpoint of the darkest 10 hours",
  first_timing_above_250 = "First timing above 250 lx melEDI",
  last_timing_above_250 = "Last timing above 250 lx melEDI"
)
predictor_labels <- c(
  work_free_day = "Work/free day",
  activity_status = "Daily activity status",
  previous_sleep_duration_centered_h = "Previous-night sleep duration"
)

association_tests <- tests |>
  dplyr::filter(.data$test_id == "association_predictor_block") |>
  dplyr::select("metric_id", "predictor_id", association_p = "raw_p_value")
heterogeneity_tests <- tests |>
  dplyr::filter(.data$test_id == "heterogeneity_predictor_by_site_block") |>
  dplyr::select("metric_id", "predictor_id", heterogeneity_p = "raw_p_value")

effect_table <- effects |>
  dplyr::left_join(
    association_tests,
    by = c("metric_id", "predictor_id"),
    relationship = "one-to-one"
  ) |>
  dplyr::left_join(
    heterogeneity_tests,
    by = c("metric_id", "predictor_id"),
    relationship = "one-to-one"
  )

sensitivity_table <- sensitivity |>
  dplyr::mutate(
    sensitivity_component = paste(.data$sensitivity_id, .data$component),
    display = format_class(.data$sensitivity_classification)
  ) |>
  dplyr::select("metric_id", "predictor_id", "sensitivity_component", "display") |>
  tidyr::pivot_wider(
    names_from = "sensitivity_component",
    values_from = "display"
  )

elapsed <- runtime$elapsed_seconds[
  runtime$component == "Complete bounded serial pilot"
]
projected <- runtime$projected_12_role_seconds[
  runtime$component == "Complete bounded serial pilot"
]
first_work_student_shift <- sensitivity$maximum_shift_in_hc3_se[
  sensitivity$metric_id == "first_timing_above_250" &
    sensitivity$predictor_id == "work_free_day" &
    sensitivity$sensitivity_id == "student_t_identity" &
    sensitivity$component == "predictor_additive"
]
l10_activity_ar_shift <- sensitivity$maximum_shift_in_hc3_se[
  sensitivity$metric_id == "l10_midpoint" &
    sensitivity$predictor_id == "activity_status" &
    sensitivity$sensitivity_id == "no_nugget_gap_aware_ar1" &
    sensitivity$component == "predictor_additive"
]
stopifnot(
  length(first_work_student_shift) == 1L,
  length(l10_activity_ar_shift) == 1L
)
```

::: {.callout-important title="H06-D-G2P-TIMING-REPAIR — author stop gate"}
The candidate participant-cluster HC3 route passes its source, design,
covariance, and numerical gates in all 12 cells. It does **not** make the
pilot results confirmatory: no BH family was updated and no pilot p-value is
accepted. The diagnostic evidence requires explicit author judgment. Overall
predictor associations are stable in 22 of 24 fitted Student-t/AR comparisons;
the other two shift by 1–<2 HC3 SE. Three additive AR fits do not converge.
Timing predictor-by-site heterogeneity is substantially less stable, with
coefficient reversals or nonconvergence across most cells.
:::

# What was repaired

The first non-L10 pilot used Gaussian mixed-model diagnostics that treated
remaining serial/site residual thresholds as acceptance requirements. This
bounded repair leaves every participant-day outcome and predictor unchanged.
It instead estimates the fixed mean structure with `stats::lm()` and uses
participant-cluster HC3 covariance:

```r
response_value ~ site
response_value ~ site + predictor
response_value ~ site * predictor

sandwich::vcovCL(
  model,
  cluster = ~ participant_key,
  type = "HC3",
  cadjust = TRUE,
  fix = FALSE
)
```

The covariance allows heteroscedasticity and arbitrary dependence among a
participant's retained days. Therefore, the residual-lag thresholds remain
useful descriptions but are not an independence requirement for this
candidate route. The Student-t and no-nugget AR models are diagnostic
sensitivities only; their p-values are never substituted.

# Exact frames and candidate-model gate

All `r nrow(input_manifest)` expanded input-contract rows verify. The 12
frame objects match their authorized hashes and retain nine sites, one row per
participant-day, sum-to-zero site contrasts, and the fixed clock encodings.
M10 midpoint, first timing, and last timing use the audited continuous clock
hours. L10 midpoint uses the fixed after-16:00 night-centred unwrap.

```{r}
#| label: tbl-frame-contract
#| tbl-cap: "Exact primary near-eye timing-repair frames."

frame_contract |>
  dplyr::transmute(
    Outcome = unname(metric_labels[.data$metric_id]),
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Clock support` = ifelse(
      .data$source_clock_acceptable,
      "Pass",
      "Fail"
    ),
    `Frame identity` = .data$frame_verification_status
  ) |>
  gt::gt(groupname_col = "Outcome") |>
  compact_gt()
```

All 36 reduced/additive/interaction candidate mean models are full rank. All
36 HC3 covariance matrices are finite, symmetric, positive semidefinite, and
usable without `fix = TRUE`; there are no covariance warnings. Maximum
observation leverage is `r sprintf("%.3f", max(candidate_models$maximum_observation_hat))`.
The largest participant cluster accounts for at most
`r sprintf("%.1f%%", 100 * max(candidate_models$maximum_cluster_hat_share))`
of total leverage in any fitted mean structure.

# Candidate estimates — pilot only

The following are pointwise 95% confidence intervals with the participant-
cluster-minus-one t reference. They are shown to verify the candidate route,
not to make scientific claims. All p-values are raw pilot values; none was
adjusted, bolded, or entered into a 15-slot family.

```{r}
#| label: tbl-candidate-effects
#| tbl-cap: "Candidate HC3 timing contrasts and raw pilot tests."

effect_table |>
  dplyr::transmute(
    Outcome = unname(metric_labels[.data$metric_id]),
    Predictor = unname(predictor_labels[.data$predictor_id]),
    Contrast = .data$contrast,
    `Estimate, h` = sprintf("%.2f", .data$estimate_hours),
    `Pointwise 95% CI, h` = sprintf(
      "%.2f to %.2f",
      .data$lower_95_hours,
      .data$upper_95_hours
    ),
    `Association p (raw pilot)` = format_p(.data$association_p),
    `Predictor × site p (raw pilot)` = format_p(.data$heterogeneity_p)
  ) |>
  gt::gt(groupname_col = "Outcome") |>
  gt::tab_source_note(
    "PILOT_RAW_ONLY_NO_BH_UPDATE; adjusted p-values are missing by design."
  ) |>
  compact_gt()
```

# Diagnostic sensitivities

All 36 Student-t fixed-mean fits converge with positive-definite Hessians. For
the overall predictor association, 11 of 12 shifts are below one HC3 SE.
First timing above 250 lx melEDI for work/free day shifts by
`r sprintf("%.2f", first_work_student_shift)` HC3 SE and is a major
limitation.

Nine of 12 additive no-nugget AR fits converge. Among them, eight overall
association shifts are stable. L10-midpoint activity shifts by
`r sprintf("%.2f", l10_activity_ar_shift)` HC3 SE and is a major limitation.
The M10-sleep, first-timing-work/free,
and last-timing-work/free additive AR checks did not converge; their shifts
are not used and their diagnostic status is unresolved.

```{r}
#| label: tbl-sensitivity
#| tbl-cap: "Effect-shift classifications for the two diagnostic sensitivities."

sensitivity_table |>
  dplyr::transmute(
    Outcome = unname(metric_labels[.data$metric_id]),
    Predictor = unname(predictor_labels[.data$predictor_id]),
    `Student-t association` =
      .data$`student_t_identity predictor_additive`,
    `AR association` =
      .data$`no_nugget_gap_aware_ar1 predictor_additive`,
    `Student-t site heterogeneity` =
      .data$`student_t_identity predictor_by_site_block`,
    `AR site heterogeneity` =
      .data$`no_nugget_gap_aware_ar1 predictor_by_site_block`
  ) |>
  gt::gt(groupname_col = "Outcome") |>
  compact_gt()
```

The site-heterogeneity sensitivity is weaker than the association sensitivity.
Ten of 12 Student-t interaction comparisons and five of seven converged AR
interaction comparisons contain at least one direction reversal under the
prespecified rule. Five AR interaction fits are unresolved because they do not
converge. This does not replace the robust candidate interaction test, but it
precludes an unqualified timing site-heterogeneity claim.

# Residual plots for author judgment

Each row is one predictor. The left panel is the candidate-model normal Q–Q
plot; the middle panel is candidate residual versus fitted clock hour; the
right panel shows residual pairs from the diagnostic no-nugget AR fit. Point
colours follow the submitted site palette. The image source data are paired
CSVs in `artifacts/10_source_data/H06_daily/`.

## Midpoint of the brightest 10 hours

```{r}
#| label: fig-residual-m10
#| fig-cap: "Residual diagnostics for the midpoint of the brightest 10 hours."
#| fig-alt: "Nine-panel diagnostic figure for M10 midpoint. Rows show work/free day, activity status, and previous-night sleep models. Q–Q plots show modest symmetric tail curvature; residual-versus-fitted plots show little systematic spread trend; diagnostic AR residual pairs have small pooled negative slopes but maximum site correlations between 0.368 and 0.542."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "08_diagnostics",
  "H06_daily_timing_repair_residuals_m10_midpoint.png"
))
```

## Midpoint of the darkest 10 hours

```{r}
#| label: fig-residual-l10
#| fig-cap: "Residual diagnostics for the midpoint of the darkest 10 hours."
#| fig-alt: "Nine-panel diagnostic figure for L10 midpoint. Rows show the three predictors. Q–Q plots have strong asymmetric extreme tails, with several standardized residuals around minus five or above six; residual-versus-fitted trends are modest; diagnostic AR residual pairs retain negative slopes and maximum site correlations from 0.404 to 0.729."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "08_diagnostics",
  "H06_daily_timing_repair_residuals_l10_midpoint.png"
))
```

## First timing above 250 lx melEDI

```{r}
#| label: fig-residual-first
#| fig-cap: "Residual diagnostics for first timing above 250 lx melEDI."
#| fig-alt: "Nine-panel diagnostic figure for first timing above 250 lx melEDI. Q–Q plots show heavy tails and separated extreme clusters, most visibly for work/free day and sleep; residual-versus-fitted plots show limited central trend but several large residuals; diagnostic AR residual pairs retain negative slopes and maximum site correlations from 0.375 to 0.443."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "08_diagnostics",
  "H06_daily_timing_repair_residuals_first_timing_above_250.png"
))
```

## Last timing above 250 lx melEDI

```{r}
#| label: fig-residual-last
#| fig-cap: "Residual diagnostics for last timing above 250 lx melEDI."
#| fig-alt: "Nine-panel diagnostic figure for last timing above 250 lx melEDI. Q–Q points lie comparatively close to the reference lines with mild tail departures; residual-versus-fitted plots show modest local structure; diagnostic AR residual pairs retain negative slopes and maximum site correlations from 0.605 to 0.677."
#| out-width: "100%"

knitr::include_graphics(artifact(
  "08_diagnostics",
  "H06_daily_timing_repair_residuals_last_timing_above_250.png"
))
```

The visual verdict is consistent with the numerical diagnostics: L10 and
first timing have the most consequential tail departures; M10 has milder
curvature; last timing is closest to linear in Q–Q space. All four retain
site-specific residual dependence after the diagnostic AR fit. For HC3 this
dependence is descriptive, while the interaction-coefficient instability is
a substantive limitation for site-heterogeneity interpretation.

# Runtime, preservation, and gate recommendation

The bounded serial pilot fitted 108 models in
`r sprintf("%.2f seconds", elapsed)`. A mechanical extension of this timing
route across the 12 already declared scenario roles is projected at
`r sprintf("%.1f seconds", projected)`. This projection does not authorize
production. The earlier 50-refit pilot remains the only deletion runtime
evidence; no deletion, bootstrap, or simulation was run here.

All `r nrow(preservation)` protected pre-existing H06_daily files remained
byte-identical. No L10-mean, MDER, temporal GAMM, pre-sleep, main-H06, shared,
Stage 3, or Stage 4 artifact changed. No BH family or adjusted p-value changed.

The recommended disposition is:

1. accept the participant-cluster HC3 route for later production of all four
   timing outcomes because all 12 candidate cells pass the uniform source,
   rank, covariance, and numerical contract;
2. carry the first-timing work/free Student-t shift and L10-activity AR shift
   as major limitations, and retain the three nonconverged additive AR checks
   as unresolved diagnostics;
3. retain the robust predictor-by-site tests and their named multiplicity
   slots if production is approved, but label timing site heterogeneity
   sensitivity-dependent and make no unqualified claim when diagnostic
   coefficients reverse or the AR fit does not converge; and
4. keep every pilot p-value nonconfirmatory until the complete production grid
   is explicitly authorized, fitted, diagnosed, and adjusted in its declared
   15-slot family.

::: {.callout-warning title="Stop — explicit author decision required"}
This report stops at `H06-D-G2P-TIMING-REPAIR`. Accepting the recommendation
would authorize neither computation nor reporting by itself. The next action
requires a separate author decision on the complete remaining non-L10
production grid, including its full diagnostic/deletion workload and the 12
declared BH families.
:::

# Reproducibility record

The authoritative scientific and display commands were:

```text
Rscript --vanilla scripts/hypotheses/H06_daily/run_h06_daily_timing_repair_pilot.R
Rscript --vanilla scripts/hypotheses/H06_daily/make_h06_daily_timing_repair_residual_plots.R
```

R 4.6.1 and the synchronized project library were used. The report reads only
sealed CSV/PNG artifacts and verifies their manifests. It performs no fit,
diagnostic recomputation, p-value adjustment, deletion, bootstrap, or
simulation.

```{r}
#| label: tbl-runtime
#| tbl-cap: "Measured bounded-pilot runtime."

runtime |>
  dplyr::transmute(
    Component = .data$component,
    `Models/calculations` = .data$model_fits,
    `Elapsed, s` = sprintf("%.3f", .data$elapsed_seconds),
    Role = .data$role,
    `12-role timing projection, s` = ifelse(
      is.na(.data$projected_12_role_seconds),
      "—",
      sprintf("%.1f", .data$projected_12_role_seconds)
    )
  ) |>
  gt::gt() |>
  compact_gt()
```
