---
title: "H06_daily: gap clock-hour repair"
subtitle: "Targeted task-owned sensitivity repair and H06-D-G2B author gate"
date: "2026-08-12"
engine: knitr
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit R code"
    code-tools: true
    embed-resources: true
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
  error: false
  freeze: false
tbl-cap-location: top
fig-cap-location: bottom
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "NATHEALTH_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) ||
    !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)
.libPaths(c(
  file.path(project_root, "renv/library/macos/R-4.6/aarch64-apple-darwin23"),
  .libPaths()
))

library(digest)
library(dplyr)
library(gt)
library(readr)
library(tibble)
library(tidyr)

artifact <- function(stage, file) {
  file.path(project_root, "artifacts", stage, "H06_daily", file)
}
read_h06d <- function(stage, file) {
  readr::read_csv(
    artifact(stage, file),
    show_col_types = FALSE,
    na = c("", "NA")
  )
}
sha256 <- function(path) {
  digest::digest(path, algo = "sha256", file = TRUE, serialize = FALSE)
}
compact_gt <- function(data, groupname_col = NULL, source_note = NULL) {
  output <- data |>
    gt::gt(groupname_col = groupname_col) |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(3),
      heading.padding = gt::px(4),
      column_labels.padding = gt::px(4),
      table.width = gt::pct(100),
      source_notes.font.size = gt::px(11),
      row_group.padding = gt::px(4)
    )
  if (!is.null(source_note)) {
    output <- output |> gt::tab_source_note(gt::md(source_note))
  }
  output
}
format_p <- function(value) {
  dplyr::case_when(
    is.na(value) ~ "—",
    value < 0.001 ~ "<0.001",
    TRUE ~ sprintf("%.3f", value)
  )
}
format_effect <- function(estimate, lower, upper) {
  sprintf("%.2f (%.2f to %.2f)", estimate, lower, upper)
}

classification <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_h01_classification.csv"
)
repaired_classification <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_90_cell_classification.csv"
)
class_summary <- read_h06d(
  "09_tables",
  "H06_daily_gap_clock_repair_classification_summary.csv"
)
bh <- read_h06d(
  "09_tables",
  "H06_daily_gap_clock_repair_bh_families.csv"
)
family_summary <- read_h06d(
  "09_tables",
  "H06_daily_gap_clock_repair_bh_family_summary.csv"
)
primary_results <- read_h06d(
  "09_tables",
  "H06_daily_gap_clock_repair_primary_gap_results.csv"
)
samples <- read_h06d(
  "09_tables",
  "H06_daily_gap_clock_repair_sample_summary.csv"
)
sidecars <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_sidecar_summary.csv"
)
scope_audit <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_scope_audit.csv"
)
visual_review <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_visual_review.csv"
)
plot_index <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_visual_plot_index.csv"
)
atlas_index <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_visual_atlas_index.csv"
)
scale_equivariance <- read_h06d(
  "09_tables",
  "H06_daily_gap_clock_repair_scale_equivariance.csv"
)
residual_scale <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_residual_scale_equivariance.csv"
)
protected <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_protected_final.csv"
)
g2a_outputs <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_g2a_outputs_final.csv"
)
runtime <- read_h06d(
  "08_diagnostics",
  "H06_daily_gap_clock_repair_runtime.csv"
)

stopifnot(
  identical(as.character(getRversion()), "4.6.1"),
  nrow(classification) == 468L,
  nrow(repaired_classification) == 90L,
  nrow(bh) == 180L,
  nrow(primary_results) == 15L,
  nrow(visual_review) == 90L,
  nrow(plot_index) == 90L,
  nrow(atlas_index) == 10L,
  all(scope_audit$verified),
  all(protected$identity_verified),
  all(g2a_outputs$identity_verified),
  !any(visual_review$numeric_threshold_set_verdict),
  !any(repaired_classification$hard_gate_failed),
  all(bh$multiplicity_family_construct_valid)
)

predictor_labels <- c(
  work_free_day = "Work/free day",
  activity_status = "Daily activity status",
  previous_sleep_duration_centered_h = "Previous-night sleep duration"
)
```

::: {.callout-important title="Answer in brief"}
The task-owned repair converted the five gap timing outcomes from their
audited clock-hour source representation to the minute-scale interface
expected by the frozen H06_daily adapter, after which the adapter converted
them exactly once back to modeled clock hours. The repair changed only the 90
authorized gap timing cells. All 90 now pass the hard construct, support, fit,
and estimability gates; visual review found reportable residual limitations
but no gross residual failure.

The six gap 15-slot BH families are interpretable again. In the repaired
near-eye all-available timing results, work/free day and previous-night sleep
duration were BH-supported for all five outcomes; daily activity status was
not BH-supported for any of the five. None of the five repaired
predictor-by-site timing tests was BH-supported. These are observational,
gap-timing-unaware sensitivity results and do not change the frozen primary
daily estimates or the selected hourly main H06 analysis.
:::

# Exact repair boundary

The owner authorized a targeted H06_daily-only repair after the H06-D-G2A
unit audit. No shared preparation file, other hypothesis, primary H06_daily
frame, L10 mean result, MDER model or raw result, temporal GAMM, pre-sleep
sensitivity, or hourly H06 artifact was changed.

The shared gap source stores the five outcomes in clock hours, including the
declared signed L10 representation. The frozen H06_daily adapter expects a
minute-scale input and divides by 60. The task-owned normalizer therefore
multiplies only these gap-source values by 60 immediately before that frozen
adapter. The modeled response is consequently the original clock-hour value,
not clock hours divided by 60. Outcomes, rows, dates, sites, predictors,
contrasts, sample roles, and response-family routes are unchanged.

```{r}
#| label: tbl-repair-scope
#| tbl-cap: "Targeted repair and preservation checks."

scope_audit |>
  transmute(
    Check = check,
    Expected = expected,
    Observed = observed,
    Verified = if_else(verified, "Yes", "No")
  ) |>
  compact_gt() |>
  cols_align("center", columns = c(Expected, Observed, Verified))
```

Exactly 378 non-L10 cells are frame-identical to the frozen grid. The repair
replaced 30 primary-gap raw tests—five metrics by three predictors by two test
types—and recomputed only the six gap families. The six primary families are
unchanged. L10 remains the named unavailable slot 3. MDER remains frozen in
slot 15; only its mathematically dependent gap-family BH fields could change.

# Repaired near-eye gap timing results

The near-eye all-available repaired frames contain 708–777 participant-days,
137–141 participants, and all nine sites, depending on outcome and predictor.
Effects are absolute clock-hour differences. For sleep duration, the effect is
per additional hour of the prespecified day-specific sleep-duration term.
Intervals are pointwise 95% confidence intervals; no simultaneous intervals
are used.

```{r}
#| label: tbl-gap-primary-results
#| tbl-cap: "Repaired gap-timing-unaware near-eye timing associations."

result_table <- primary_results |>
  mutate(
    Outcome = manuscript_name,
    Predictor = unname(predictor_labels[predictor_id]),
    Days = participant_days,
    Participants = participants,
    Sites = sites,
    `Difference in clock time, h (95% CI)` = format_effect(
      display_estimate,
      display_lower_95,
      display_upper_95
    ),
    `Raw p` = format_p(association_raw_p_value),
    `BH q` = format_p(association_bh_adjusted_p_value),
    bh_supported = association_bh_adjusted_p_value < 0.05
  )

result_table |>
  select(
    Outcome,
    Predictor,
    Days,
    Participants,
    Sites,
    `Difference in clock time, h (95% CI)`,
    `Raw p`,
    `BH q`,
    bh_supported
  ) |>
  compact_gt(
    groupname_col = "Outcome",
    source_note = paste(
      "BH q-values belong to separate 15-slot gap association families by",
      "predictor. Bold q-values meet the explicitly labelled BH q < 0.05",
      "rule. Every row remains acceptable with a visual residual limitation."
    )
  ) |>
  cols_hide(columns = bh_supported) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = `BH q`, rows = bh_supported)
  ) |>
  cols_align(
    "center",
    columns = c(Days, Participants, Sites, `Raw p`, `BH q`)
  )
```

Exact fitted samples for every complementary chest, paired/common, and
dataset-common cell are retained in
`artifacts/09_tables/H06_daily/H06_daily_gap_clock_repair_sample_summary.csv`;
the corresponding 90 point estimates and pointwise intervals are in
`artifacts/09_tables/H06_daily/H06_daily_gap_clock_repair_effects.csv`.

Across these five repaired timing outcomes:

- Free days were associated with later M10 midpoint, later L10 midpoint,
  later mean timing above 250 lx melEDI, later first timing above 250 lx
  melEDI, and earlier last timing above 250 lx melEDI. All five associations
  retained BH support.
- A longer previous-night sleep duration was associated with later M10 and
  L10 midpoints, later mean and first timing above 250 lx melEDI, and earlier
  last timing above 250 lx melEDI. All five associations retained BH support.
- Daily Active versus Sedentary status was not BH-supported for any of the
  five repaired outcomes. The last-timing contrast had raw p = 0.041 but
  BH q = 0.089, so it is not bolded and does not support a multiplicity-adjusted
  claim.

All wording describes associations with recorded day-level contexts. It does
not imply causal effects or health outcomes.

## Site heterogeneity

```{r}
#| label: tbl-gap-timing-heterogeneity
#| tbl-cap: "BH-adjusted site-heterogeneity tests for the five repaired timing outcomes."

primary_results |>
  transmute(
    Outcome = manuscript_name,
    Predictor = unname(predictor_labels[predictor_id]),
    `Heterogeneity BH q` = format_p(heterogeneity_bh_adjusted_p_value),
    supported = heterogeneity_bh_adjusted_p_value < 0.05
  ) |>
  compact_gt(groupname_col = "Outcome", source_note = paste(
    "No repaired timing heterogeneity q-value is below 0.05. The four HC3",
    "timing interaction routes remain sensitivity-dependent regardless of",
    "p-value; no unqualified site-interaction claim is permitted."
  )) |>
  cols_hide(columns = supported) |>
  cols_align("center", columns = `Heterogeneity BH q`)
```

No repaired timing predictor-by-site test was BH-supported. This does not
establish identical associations across sites; it means the fitted
heterogeneity block did not cross the prespecified multiplicity threshold.

# Complete 15-slot gap families

```{r}
#| label: tbl-gap-family-summary
#| tbl-cap: "Completed gap-timing-unaware BH families after the targeted repair."

family_summary |>
  filter(dataset_id == "gap_timing_unaware") |>
  transmute(
    Predictor = unname(predictor_labels[predictor_id]),
    Test = recode(
      test_type,
      association = "Association",
      site_heterogeneity = "Site heterogeneity"
    ),
    `Named slots` = named_slots,
    `Available raw tests` = available_raw_slots,
    `Named NA slots` = named_na_slots,
    `BH-supported slots` = fdr_supported_slots,
    `Claim-eligible slots` = claim_eligible_slots,
    `Family valid` = if_else(family_construct_valid, "Yes", "No")
  ) |>
  compact_gt(groupname_col = "Predictor", source_note = paste(
    "Each family retains all 15 named slots and uses stats::p.adjust with",
    "n = 15. Slot 3 (L10 mean) remains NA. Slot 15 (MDER) is frozen and is",
    "not counted as a new claim even if its derived q-value is below 0.05."
  )) |>
  cols_align(
    "center",
    columns = c(
      `Named slots`, `Available raw tests`, `Named NA slots`,
      `BH-supported slots`, `Claim-eligible slots`, `Family valid`
    )
  )
```

Across the complete gap families, 29 of the 39 estimable non-L10 association
slots and 11 of the 39 estimable non-L10 site-heterogeneity slots have BH
q-values below 0.05 and pass the H01-aligned hard gate. Site-interaction
claims from the HC3 timing route remain explicitly sensitivity-dependent.

# H01-aligned diagnostics

The residual rule was applied exactly as approved: centering,
homoscedasticity, and gross normality were judged visually from
residual-versus-fitted and normal Q-Q plots. Numeric summaries did not assign
any verdict. A gross visual problem could fail a cell; none was observed.

```{r}
#| label: tbl-final-classification
#| tbl-cap: "Final H01-aligned classification of the complete 468-cell non-L10 grid."

class_summary |>
  transmute(
    Class = h01_reader_label,
    Cells = cells,
    `Gross visual failures` = gross_visual_failures,
    `Hard-gate failures` = hard_gate_failures
  ) |>
  compact_gt() |>
  cols_align(
    "center",
    columns = c(Cells, `Gross visual failures`, `Hard-gate failures`)
  )
```

After repairing the true unit/construct defect, all 468 non-L10 production
cells are acceptable with limitations and none fails a hard gate. This does
not mean the diagnostics are clean: each cell has a recorded visual residual
limitation, and the sensitivity sidecars remain visible.

```{r}
#| label: tbl-repair-sidecars
#| tbl-cap: "Diagnostic and sensitivity sidecars for the 90 repaired cells."

sidecars |>
  transmute(
    Sidecar = sidecar,
    Classification = classification,
    Cells = cells,
    `Overall hard-gate input` = if_else(
      overall_hard_gate_input,
      "Only if FAIL_GROSS",
      "No"
    )
  ) |>
  compact_gt(groupname_col = "Sidecar") |>
  cols_align("center", columns = c(Cells, `Overall hard-gate input`))
```

For the 72 HC3 cells, the additive Student-t comparison was stable in 70 and
showed a substantial 1-to-less-than-2-SE shift in two. The gap-aware
no-nugget AR additive sidecar was stable in 42, substantially shifted in four,
unstable in five, and numerically unresolved in 21. For the 18 mixed-model
mean-timing cells, AR was not triggered in 11; three had unresolved AR
numerics, one was singular, and three retained post-AR residual dependence.
Deletion influence was stable in 67 cells, a substantial limitation in 19,
and unstable in four. Under the approved H01 architecture, these findings
must be reported but do not fail an otherwise estimable cell.

The repaired residuals and fitted values reproduce the historical
mis-scaled residual geometry after multiplication by 60 to within
`r sprintf("%.2g", max(residual_scale$fitted_scale_maximum_absolute_error))`
clock hours for fitted values and
`r sprintf("%.2g", max(residual_scale$standardized_residual_maximum_absolute_error))`
for standardized residuals. This scale-equivariance explains why the unit
repair restores the construct without erasing the previously observed
residual-shape limitations.

## Residual review atlas

The ten audit atlases below show all 90 repaired cells in indexed order, nine
cells per atlas. Within each atlas, panels run left-to-right and then
top-to-bottom. The canonical cell index records outcome, placement, sample,
predictor, plot SHA-256, paired source-data SHA-256, and cell-specific alt
text. The atlas is a navigation aid; the individual cell plots are the
full-resolution audit source.

```{r}
#| label: tbl-atlas-map
#| tbl-cap: "Residual atlas index."

atlas_map <- plot_index |>
  group_by(atlas_group_key) |>
  summarise(
    Cells = sprintf("%03d–%03d", min(repair_cell_index), max(repair_cell_index)),
    Outcomes = paste(unique(manuscript_name), collapse = "; "),
    `Atlas SHA-256` = first(atlas_sha256),
    .groups = "drop"
  ) |>
  mutate(Atlas = toupper(gsub("_", " ", atlas_group_key)), .before = 1L) |>
  select(-atlas_group_key)

atlas_map |>
  compact_gt() |>
  cols_width(
    Atlas ~ px(90),
    Cells ~ px(90),
    Outcomes ~ px(360),
    `Atlas SHA-256` ~ px(460)
  )
```

```{r}
#| label: fig-residual-atlases
#| results: asis

for (index in seq_len(nrow(atlas_index))) {
  row <- atlas_index[index, ]
  atlas_name <- toupper(gsub("_", " ", row$atlas_group_key))
  alt <- paste0(
    atlas_name,
    ", repaired residual diagnostics for cells ",
    sprintf("%03d", row$first_repair_cell),
    " through ",
    sprintf("%03d", row$last_repair_cell),
    ". Each cell shows standardized residuals versus fitted clock hours and ",
    "a normal Q-Q plot."
  )
  source_path <- file.path("..", "..", "..", row$atlas_relative_path)
  cat(sprintf("\n### %s: cells %03d–%03d\n\n", atlas_name,
              row$first_repair_cell, row$last_repair_cell))
  cat(sprintf(
    "![](%s){fig-alt=\"%s\" width=100%%}\n\n",
    source_path,
    alt
  ))
}
```

# Scientific interpretation and comparison boundary

This repair retains the intended gap-timing-unaware sensitivity: it asks
whether conclusions persist when the gap timing information is ignored while
keeping the participant-day outcome definitions and predictor hierarchy. It
does not rebuild shared preparation, alter the primary daily analysis, or
make the daily and hourly H06 estimands numerically interchangeable.

The corrected p-values are scale-equivariant to the historical values within
`r sprintf("%.2g", max(scale_equivariance$absolute_raw_p_difference))` for raw
p-values. That numerical similarity does not excuse the historical unit
error: clock hours divided by 60 were not a valid reader-facing construct.
The repaired frame and model objects are newly sealed, while the 378
unaffected cells remain frozen.

# Reproducibility and preservation

The repair ran serially in R `r as.character(getRversion())`. The 90 repaired
base cells and embedded sidecars took
`r sprintf("%.1f", runtime$wall_seconds[runtime$component == "90 repaired base cells and embedded sidecars"])`
seconds. The complete 12,837-refit participant/site deletion battery took
`r sprintf("%.1f", runtime$wall_seconds[runtime$component == "complete participant/site deletion battery"])`
seconds, with no failed refit or repeated checkpoint failure. No bootstrap or
simulation was run.

Final verification confirms:

- all 1,011 protected historical identities remain exact;
- all 997 frozen H06-D-G2A output identities remain exact;
- all 378 unaffected frame objects remain exact;
- the 30 repaired raw slots are the only replaced raw tests;
- the six gap families are the only recomputed BH families; and
- every output written by this amendment is under an H06_daily-owned path.

# H06-D-G2B-GAP-CLOCK-REPAIR author stop gate

The targeted repair is complete and stops here for explicit author review.
The recommended disposition is to accept the repaired gap clock-hour branch
as the intended gap-timing-unaware sensitivity, with all visual, AR,
response-family, influence, and timing-interaction limitations retained in
the result language.

Acceptance of this gate would permit the coordinator to close the Stage 2
repair record. It does not by itself authorize Stage 3 reader-report writing,
Stage 4 provenance writing, website integration, shared preparation edits,
main-H06 changes, a commit, or a push.
