---
title: "H06 daily-metric analysis preparation and provenance"
subtitle: "From participant-day light-exposure metrics and recorded contexts to complementary association estimates"
date: "2026-08-13"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: false
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
resources:
  - ../../../artifacts/10_figures/H06_daily/H06_daily_preparation_primary_sample_support.png
  - ../../../artifacts/10_figures/H06_daily/H06_daily_preparation_primary_sample_support.pdf
  - ../../../artifacts/10_figures/H06_daily/H06_daily_preparation_primary_sample_support.svg
  - ../../../artifacts/11_source_data/H06_daily/H06_daily_preparation_primary_sample_figure.csv
editor: source
---

```{r}
#| label: setup-h06-daily-preparation
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(gt)
  library(knitr)
  library(readr)
  library(tibble)
  library(tidyr)
})

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- locate_project_root()

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H06_daily preparation requires R 4.6.1", call. = FALSE)
}

read_h06d <- function(filename) {
  readr::read_csv(
    file.path(
      project_root,
      "artifacts/11_source_data/H06_daily",
      filename
    ),
    show_col_types = FALSE,
    na = c("", "NA")
  )
}

compact_gt <- function(
  data,
  groupname_col = NULL,
  source_note = NULL,
  font_size = 11.5
) {
  output <- if (inherits(data, "gt_tbl")) {
    data
  } else {
    gt::gt(data, groupname_col = groupname_col)
  }
  output <- output |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(10.5),
      row_group.padding = gt::px(4)
    )
  if (!is.null(source_note)) {
    output <- output |>
      gt::tab_source_note(gt::md(source_note))
  }
  output
}

format_range <- function(minimum, maximum) {
  ifelse(
    minimum == maximum,
    format(minimum, big.mark = ",", scientific = FALSE, trim = TRUE),
    paste0(
      format(minimum, big.mark = ",", scientific = FALSE, trim = TRUE),
      "–",
      format(maximum, big.mark = ",", scientific = FALSE, trim = TRUE)
    )
  )
}

input_provenance <- read_h06d(
  "H06_daily_preparation_input_provenance.csv"
)
metric_registry <- read_h06d(
  "H06_daily_preparation_metric_registry.csv"
)
predictor_registry <- read_h06d(
  "H06_daily_preparation_predictor_registry.csv"
)
model_routes <- read_h06d(
  "H06_daily_preparation_model_routes.csv"
)
exact_samples <- read_h06d(
  "H06_daily_preparation_exact_samples.csv"
)
sample_roles <- read_h06d(
  "H06_daily_preparation_sample_role_summary.csv"
)
primary_samples <- read_h06d(
  "H06_daily_preparation_primary_sample_by_metric.csv"
)
predictor_support <- read_h06d(
  "H06_daily_preparation_primary_predictor_support.csv"
)
site_support <- read_h06d(
  "H06_daily_preparation_primary_site_support.csv"
)
multiplicity <- read_h06d(
  "H06_daily_preparation_multiplicity_families.csv"
)
diagnostics <- read_h06d(
  "H06_daily_preparation_diagnostic_summary.csv"
)
sensitivity_registry <- read_h06d(
  "H06_daily_preparation_sensitivity_registry.csv"
)
temporal_support <- read_h06d(
  "H06_daily_preparation_temporal_support.csv"
)
code_map <- read_h06d("H06_daily_preparation_code_map.csv")
output_trace <- read_h06d("H06_daily_preparation_output_trace.csv")
figure_alt <- read_h06d("H06_daily_preparation_figure_alt_text.csv")

joint_registry <- readr::read_csv(
  file.path(
    project_root,
    paste0(
      "artifacts/06_model_data/H06_daily/",
      "H06_daily_joint_context_exploratory_frame_registry.csv"
    )
  ),
  show_col_types = FALSE
)
temporal_formula <- readr::read_csv(
  file.path(
    project_root,
    paste0(
      "artifacts/06_model_data/H06_daily/",
      "H06_daily_temporal_h02_formula_registry.csv"
    )
  ),
  show_col_types = FALSE
)

stopifnot(
  all(input_provenance$verification_status == "PASS"),
  all(input_provenance$expected_sha256 == input_provenance$observed_sha256),
  nrow(metric_registry) == 15L,
  identical(as.integer(metric_registry$metric_slot), 1:15),
  nrow(predictor_registry) == 3L,
  nrow(model_routes) == 15L,
  nrow(exact_samples) == 522L,
  nrow(sample_roles) == 12L,
  nrow(primary_samples) == 15L,
  nrow(site_support) == 405L,
  nrow(multiplicity) == 12L,
  nrow(diagnostics) == 25L,
  nrow(temporal_support) == 6L,
  nrow(joint_registry) == 14L,
  nrow(temporal_formula) == 1L,
  nrow(figure_alt) == 1L,
  all(metric_registry$analysis_unit == "participant-day"),
  all(multiplicity$named_slots == 15L),
  all(multiplicity$named_na_slots == 1L)
)
```

This page explains how the complementary H06 participant-day analysis was
prepared, fitted, checked, and translated into the reported tables and
figures. The selected [hourly H06 analysis](../../../notebooks/hypotheses/H06.qmd)
remains the main H06 result. A participant-day is one eligible local calendar
day from one participant. Each admissible light-exposure metric contributes at
most one row for that participant-day, so days are not weighted by their number
of supported hours.

The complementary [H06 daily-metric results report](../../../notebooks/hypotheses/H06_daily.qmd)
contains the association estimates, confidence intervals, multiplicity
decisions, site-specific patterns, complementary analyses, and the comparison
with the selected main hourly H06 analysis.

::: {.callout-note title="Execution boundary"}
This page verifies stored file identities and schemas and calculates only
bounded descriptive summaries of the already fitted samples. Model estimates,
predictions, residual correlation, p-values, false-discovery-rate (FDR)
adjustments, deletion analyses, and nonlinear clock-time functions are read
from stored outputs. FDR adjustment limits the expected proportion of false
discoveries within each defined family. Rendering this page does not fit or
refit a model, regenerate a
prediction, rerun a sensitivity analysis, resample, or simulate.
:::

## Analysis path

The primary path estimates each context separately for each daily metric. The
predictor-by-site interaction path allows that association to differ among
sites. Two exploratory paths then examine covariate stability and 30-minute
clock-time profiles without changing the primary participant-day estimand. The
nonlinear generalized additive mixed model (GAMM) analysis represents smooth
clock-time variation while accounting for repeated measurements. Its working
first-order autoregressive correlation, AR(1), allows adjacent residuals within
a participant-day sequence to be correlated. A pointwise 95% confidence
interval (95% CI) assesses one displayed estimate and is not a simultaneous
band. Model checks assess fit, observed support, residual behavior, and
sensitivity evidence.

```{mermaid}
flowchart TD
  A["Verified participant-day light-exposure metrics"] --> D["Metric-specific participant-day frames"]
  B["Work day or Free day and daily activity status"] --> D
  C["Previous-night sleep duration"] --> D
  D --> E["Fixed-site common-association model"]
  D --> F["Predictor-by-site interaction model"]
  D --> G["Exploratory jointly adjusted model"]
  H["Supported 30-minute arithmetic-mean melEDI"] --> I["Exploratory clock-time GAMM with working AR(1)"]
  E --> J["Contrasts, pointwise 95% CIs, FDR families, and model checks"]
  F --> J
  G --> K["Separate exploratory FDR families and stability checks"]
  I --> L["Pointwise clock-time functions and whole-function tests"]
  J --> M["Complementary H06 daily-metric results"]
  K --> M
  L --> M
```

```{r}
#| label: tbl-h06d-prep-execution-boundary
#| tbl-cap: "What this page reads and what it deliberately does not execute"

tibble::tribble(
  ~Activity, ~Status, ~Purpose,
  "Verify stored identities and schemas", "Executed", "Detects stale or mismatched inputs",
  "Summarize exact fitted rows", "Executed", "Makes metric-specific denominators visible",
  "Summarize category and site support", "Executed", "Shows the support behind fitted contrasts",
  "Fit or compare statistical models", "Not executed", "Stored fitted objects and results remain the source",
  "Generate predictions or contrasts", "Not executed", "Stored reader source data remain the source",
  "Calculate raw or adjusted p-values", "Not executed", "Stored complete-family results remain the source",
  "Rerun AR(1), deletion, resampling, or simulation", "Not executed", "Stored model-check sidecars remain the source"
) |>
  compact_gt()
```

## Exact inputs and analytical unit

### Verified provenance

The analysis consumed the current prepared metric and context artifacts. It did
not rebuild measurement preparation or alter a metric locally, except for the
documented task-local conversion of five gap-comparison clock outcomes from
minutes to hours. That conversion was limited to H06 daily-metric sensitivity
frames and left the primary, chest, paired, MDER, L10, temporal, and other
hypothesis artifacts unchanged.

```{r}
#| label: tbl-h06d-prep-input-provenance
#| tbl-cap: "Verified scientific and reporting input identities"

input_provenance |>
  transmute(
    Input = case_when(
      grepl("results source", .data$input_role, ignore.case = TRUE) ~ "Daily-metric results source",
      grepl("results HTML", .data$input_role, ignore.case = TRUE) ~ "Daily-metric results HTML",
      grepl("source-data manifest", .data$input_role, ignore.case = TRUE) ~ "Reader source-data manifest",
      grepl("output manifest", .data$input_role, ignore.case = TRUE) ~ gsub(" output manifest", " outputs", .data$input_role),
      grepl("render QA", .data$input_role, ignore.case = TRUE) ~ "Reader render verification",
      grepl("figure QA", .data$input_role, ignore.case = TRUE) ~ "Reader figure verification",
      grepl("focused test", .data$input_role, ignore.case = TRUE) ~ "Reader structural verification",
      grepl("transition", .data$input_role, ignore.case = TRUE) ~ "Scientific transition record",
      grepl("revision record", .data$input_role, ignore.case = TRUE) ~ "Reader revision record",
      TRUE ~ .data$input_role
    ),
    Path = paste0("`", .data$relative_path, "`"),
    `SHA-256 prefix` = substr(.data$observed_sha256, 1L, 12L),
    Status = .data$verification_status
  ) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "Path") |>
  compact_gt(font_size = 10.5)
```

The **gap-timing-unaware dataset** retains the general 50%-per-hour and
80%-per-day coverage rules. It differs because the timing of remaining gaps is
not used for the additional metric-specific adjustment applied in the primary
dataset. This comparison therefore tests the consequence of using remaining
gap timing, not the consequence of abandoning the shared coverage rules.

### Outcomes and participant-day rows

The 15 outcomes span light level, duration, timing, exposure history, and
spectral composition. Reader-facing units remain the metric's original units;
the stored source values are never overwritten by a display transformation.
“Period” denotes a continuous interval above a threshold.

```{r}
#| label: tbl-h06d-prep-metric-registry
#| tbl-cap: "Daily light-exposure registry, response treatment, and fitted route"

metric_registry |>
  transmute(
    Slot = .data$metric_slot,
    Metric = .data$manuscript_name,
    Abbreviation = .data$abbreviation,
    Unit = .data$display_unit,
    `Family and transformation` = paste(
      .data$response_family,
      .data$response_transform,
      sep = "; "
    ),
    `Fitted route` = .data$reader_route
  ) |>
  gt::gt() |>
  gt::tab_row_group(label = "Light level", rows = Slot %in% 1:3) |>
  gt::tab_row_group(label = "Duration and continuous period", rows = Slot %in% 4:8) |>
  gt::tab_row_group(label = "Clock timing", rows = Slot %in% 9:13) |>
  gt::tab_row_group(label = "Exposure history and spectrum", rows = Slot %in% 14:15) |>
  gt::row_group_order(groups = c(
    "Light level",
    "Duration and continuous period",
    "Clock timing",
    "Exposure history and spectrum"
  )) |>
  gt::cols_hide(columns = "Slot") |>
  compact_gt(font_size = 10.5)
```

For shifted-log outcomes, the fitted response is
$Z = \log_{10}(Y + 0.1)$, where $Y$ remains in the metric's stated unit, such
as lx for melEDI, h for continuous periods, or lx h for dose. The
back-transformed quantity is geometric-mean-like for $Y + 0.1$ in that unit;
it is not the raw arithmetic mean of $Y$. The main response treatments were:

- $\log_{10}(Y + 0.1)$ for zero-containing right-skewed level, continuous-
  period, and dose outcomes selected for the shifted-log Gaussian route;
- a Tweedie distribution with log link for non-negative duration outcomes with
  a point mass at zero and a continuous positive part;
- identity-scale Gaussian models for outcomes whose residual behavior was
  judged adequate on that scale; and
- clock hours for timing outcomes. L10 midpoint was unwrapped at 16:00 so that
  values after 16:00 and before the following morning occupy one continuous
  night-centred axis.

Exact zero handling was metric-specific. Exact zeros remained in every
one-part model whose response allowed zero. For the darkest 10 h mean, exact
zeros remained in the occurrence component and were excluded only from the
strictly positive magnitude component. Its joint two-part association was not
estimable because the occurrence component separated, so its named
multiplicity slot remains missing. Daily melanopic daylight efficacy ratio
(MDER) is the arithmetic mean of viable positive finite one-minute
melEDI/illuminance ratios. It requires at least 720 viable ratios on a complete
1,440-minute local wall-clock grid and is not a ratio of daily integrals.

### Recorded contexts

```{r}
#| label: tbl-h06d-prep-predictors
#| tbl-cap: "Recorded contexts and practical contrasts"

predictor_registry |>
  transmute(
    Context = .data$reader_name,
    Reference = .data$reference,
    `Reported contrast` = .data$interpretation
  ) |>
  compact_gt()
```

Work/Free and Sedentary/Active are participant-day labels, not
contemporaneous half-hour states. Previous-night sleep duration is linked from
the preceding eligible sleep period and centred at 8 h for fitting; a reported
coefficient represents one additional hour. The exploratory clock-time model
also separates each participant's mean previous sleep from the day's deviation
around that participant mean, distinguishing between-person from within-person
associations.

## Fitted samples and support

### Primary near-eye samples

The primary near-eye sample changes across metrics and predictors because a
row must have an admissible metric and the relevant recorded context. The exact
registry contains one row for every fitted metric, predictor, placement, sample
role, and dataset branch.

```{r}
#| label: fig-h06d-prep-primary-sample-support
#| fig-cap: "Exact participant-days in each primary near-eye metric-by-predictor fitted sample. Each point is the stored model denominator, not a newly selected sample."
#| fig-alt: !expr figure_alt$alt_text[[1L]]
#| out-width: "100%"

xfun::in_dir(
  file.path(project_root, "audit", "hypotheses", "H06_daily"),
  knitr::include_graphics(file.path(
    project_root,
    "artifacts/10_figures/H06_daily",
    "H06_daily_preparation_primary_sample_support.png"
  ))
)
```

```{r}
#| label: tbl-h06d-prep-primary-samples
#| tbl-cap: "Exact primary near-eye sample ranges across the three predictor-specific models"

primary_samples |>
  transmute(
    Slot = .data$metric_slot,
    Metric = .data$manuscript_name,
    `Participant-days` = format_range(
      .data$participant_days_min,
      .data$participant_days_max
    ),
    Participants = format_range(
      .data$participants_min,
      .data$participants_max
    ),
    Sites = format_range(.data$sites_min, .data$sites_max),
    `Exact source zeros` = format_range(
      .data$exact_source_zeros_min,
      .data$exact_source_zeros_max
    )
  ) |>
  gt::gt() |>
  gt::cols_hide(columns = "Slot") |>
  compact_gt(font_size = 10.5)
```

Every primary metric-by-predictor sample contains all nine sites. The smallest
samples occur where metric admissibility or context linkage is most demanding,
not because the fitting function silently removes rows. Complete frames were
constructed before fitting and their object identities were sealed.

### Placement and dataset roles

A common sample retains the same participant-days across the models,
placements, or dataset constructions being compared. The exploratory jointly
adjusted near-eye analysis uses 715 participant-days from 137 participants;
other common-sample roles use their exact metric-specific denominators.

```{r}
#| label: tbl-h06d-prep-sample-roles
#| tbl-cap: "Exact fitted-sample ranges by placement, dataset, and common-sample role"

sample_roles |>
  transmute(
    Scenario = .data$scenario_label,
    `Metric × predictor cells` = .data$analyzed_metric_predictor_cells,
    Metrics = .data$analyzed_metrics,
    `Participant-days` = format_range(
      .data$participant_days_min,
      .data$participant_days_max
    ),
    Participants = format_range(
      .data$participants_min,
      .data$participants_max
    ),
    Sites = format_range(.data$sites_min, .data$sites_max),
    Role = recode(
      .data$analysis_role,
      primary = "Primary",
      contextual_complement = "Complementary placement",
      paired_complement = "Paired/common placement",
      dataset_sensitivity = "Dataset sensitivity",
      dataset_sensitivity_context = "Complementary dataset sensitivity",
      dataset_paired_complement = "Paired/common dataset sensitivity",
      common_sample_primary = "Primary on common days",
      common_sample_gap = "Gap comparison on common days"
    )
  ) |>
  compact_gt(font_size = 10)
```

Near-eye all-available participant-days define the primary analysis. Chest
all-available estimates provide complementary non-ocular placement evidence.
Paired/common estimates compare placements on the same days. Dataset-common
estimates compare the primary and gap-timing-unaware constructions on identical
day keys. Only primary near-eye and gap-timing-unaware near-eye all-available
tests enter multiplicity families; the other roles are estimation-only.

### Predictor category and site support

```{r}
#| label: tbl-h06d-prep-category-support
#| tbl-cap: "Primary near-eye predictor support across the 15 metric-specific frames"

category_support <- predictor_support |>
  filter(.data$support_type == "category") |>
  summarise(
    metric_frames = n_distinct(.data$metric_slot),
    participant_days_min = min(.data$participant_days),
    participant_days_max = max(.data$participant_days),
    .by = c("predictor_order", "reader_name", "category")
  ) |>
  transmute(
    Context = .data$reader_name,
    Category = .data$category,
    `Metric frames` = .data$metric_frames,
    `Participant-days across metrics` = format_range(
      .data$participant_days_min,
      .data$participant_days_max
    )
  )

sleep_support <- predictor_support |>
  filter(.data$support_type == "continuous_range") |>
  summarise(
    metric_frames = n_distinct(.data$metric_slot),
    participant_days_min = min(.data$participant_days),
    participant_days_max = max(.data$participant_days),
    minimum_h = min(.data$previous_sleep_minimum_h),
    median_min_h = min(.data$previous_sleep_median_h),
    median_max_h = max(.data$previous_sleep_median_h),
    maximum_h = max(.data$previous_sleep_maximum_h)
  ) |>
  transmute(
    Context = "Previous-night sleep duration",
    Category = paste0(
      "Observed ", sprintf("%.1f", .data$minimum_h), "–",
      sprintf("%.1f", .data$maximum_h), " h; metric medians ",
      sprintf("%.1f", .data$median_min_h), "–",
      sprintf("%.1f", .data$median_max_h), " h"
    ),
    `Metric frames` = .data$metric_frames,
    `Participant-days across metrics` = format_range(
      .data$participant_days_min,
      .data$participant_days_max
    )
  )

bind_rows(category_support, sleep_support) |>
  compact_gt()
```

```{r}
#| label: tbl-h06d-prep-site-support
#| tbl-cap: "Primary near-eye participant-day support by site across all 45 metric-by-predictor fitted samples"

site_support |>
  summarise(
    fitted_cells = n(),
    participant_days_min = min(.data$participant_days),
    participant_days_max = max(.data$participant_days),
    .by = c("display_order", "display_name", "color_hex")
  ) |>
  arrange(.data$display_order) |>
  transmute(
    Site = paste0(
      "<span aria-hidden='true' style='color:",
      .data$color_hex,
      ";font-size:1.15em'>●</span> ",
      .data$display_name
    ),
    `Fitted cells` = .data$fitted_cells,
    `Participant-days across cells` = format_range(
      .data$participant_days_min,
      .data$participant_days_max
    )
  ) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "Site") |>
  compact_gt()
```

The site table reports observed support, not site-average analytical weighting.
Filled dots use consistent study-site colors to identify sites and do not
encode significance. Sum-to-zero site coding and site-average marginalization
give each included site equal weight in the reported common association even
though their observed sample sizes differ.

## Model construction and estimands

### Predictor-specific model hierarchy

For each metric and predictor, the reduced model contains site and the repeated-
participant structure. The additive model adds the predictor and estimates the
site-average common association, giving each site equal weight. The interaction
model adds the complete predictor-by-site block and tests whether the
association varies among sites. Work day and Sedentary are the categorical
references. All reported intervals are pointwise 95% CIs, not simultaneous
intervals.

```{r}
#| label: tbl-h06d-prep-model-hierarchy
#| tbl-cap: "Selected model routes and exact formula hierarchy"

model_routes |>
  distinct(
    .data$reader_route,
    .data$reduced_formula,
    .data$additive_formula,
    .data$heterogeneity_formula,
    .data$uncertainty
  ) |>
  transmute(
    Route = .data$reader_route,
    Reduced = paste0("`", .data$reduced_formula, "`"),
    Additive = paste0("`", .data$additive_formula, "`"),
    `Predictor-by-site interaction` = paste0("`", .data$heterogeneity_formula, "`"),
    Uncertainty = .data$uncertainty
  ) |>
  gt::gt() |>
  gt::fmt_markdown(columns = c("Reduced", "Additive", "Predictor-by-site interaction")) |>
  compact_gt(font_size = 10)
```

The four timing outcomes repaired to participant-cluster HC3 inference are M10
midpoint, L10 midpoint, first timing above 250 lx melEDI, and last timing above
250 lx melEDI. Their fixed-site linear models use participant-cluster HC3
covariance with finite-cluster adjustment and participant-clusters-minus-one
degrees of freedom. Other estimable daily outcomes use their selected Gaussian
or Tweedie mixed model with a participant random intercept, which allows
participants to have different baseline outcome levels.

The exact registered benchmark was retained without silent substitution:

```r
response_value ~ predictor + (predictor | site) +
  (1 | site:participant_key)
```

It is a benchmark because nine sites provide limited support for a random site
intercept, random predictor slope, and their correlation. The fixed-site route
is the reported estimand. For categorical predictors it is the Free/Work or
Active/Sedentary site-average contrast, with every site weighted equally. For
previous sleep it is the association per additional hour, also averaged equally
over sites. The
interaction route reports the complete site-specific association, not an
interaction adjustment that must be interpreted alone.

### Multiplicity

```{r}
#| label: tbl-h06d-prep-multiplicity
#| tbl-cap: "Complete primary and gap-timing-unaware false-discovery-rate families"

multiplicity |>
  mutate(
    Dataset = recode(
      .data$dataset_id,
      primary = "Primary near eye",
      gap_timing_unaware = "Gap-timing-unaware near eye"
    ),
    Test = recode(
      .data$test_type,
      association = "Common association",
      site_heterogeneity = "Predictor-by-site interaction"
    )
  ) |>
  transmute(
    Dataset,
    Predictor = .data$reader_name,
    Test,
    `Named slots` = .data$named_slots,
    Estimable = .data$estimable_slots,
    `Named missing` = .data$named_na_slots,
    `FDR-supported` = .data$fdr_supported_slots,
    `Claim-eligible` = .data$claim_eligible_slots
  ) |>
  gt::gt(groupname_col = "Dataset") |>
  compact_gt(font_size = 10.5)
```

There are 12 Benjamini–Hochberg FDR families: two datasets × three predictors
× two test types. Every family retains all 15 named metric slots. Darkest 10 h
mean melEDI remains the named missing slot because its joint test is
non-estimable; the family size is not reduced to 14. The frozen MDER raw test
occupies slot 15. Only adjusted fields that depend mathematically on the whole
family were updated when the remaining slots became available.

## Model checks, influence, and sensitivities

### Three-class model-check architecture

Hard failure was reserved for an unavailable or non-estimable base result,
convergence or Hessian failure, singularity, rank or covariance failure,
invalid timing support, observed-response support failure, or visually
confirmed gross residual behavior. Residual distribution was judged visually
from residual-versus-fitted and Q–Q displays for Gaussian or HC3 fits. Tweedie
fits used Pearson residual-versus-fitted displays and zero-mass and prediction-
support checks; normal Pearson residuals were not required.

All 468 non-L10 metric-by-predictor-by-scenario cells had complete visual
review and no gross visual or hard construct failure after the gap clock-unit
repair. Each was retained as acceptable with limitations. Autocorrelation,
participant or site deletion, exact-period, and response-family checks are
mandatory row-linked sidecars. They qualify interpretation but do not by
themselves fail an otherwise estimable result.

```{r}
#| label: tbl-h06d-prep-diagnostics
#| tbl-cap: "Stored model-check classifications and mandatory limitation sidecars"

diagnostics |>
  transmute(
    Domain = .data$domain,
    Classification = .data$reader_classification,
    Cells = .data$cells,
    Role = if_else(
      .data$hard_gate_input,
      "Overall classification input",
      "Nonblocking limitation sidecar"
    )
  ) |>
  gt::gt(groupname_col = "Domain") |>
  gt::fmt_integer(columns = "Cells") |>
  compact_gt(font_size = 10)
```

The sidecars make limitations visible rather than erasing an estimable
association. Examples include Student-t shifts, unresolved or singular AR
counterparts, residual dependence that remained after an AR fit, direction
changes under participant or site deletion, and exact-period sensitivity.
They are reported with the corresponding result and constrain generalization,
especially where site-interaction estimates are sensitivity-dependent.

### Prespecified and complementary branches

```{r}
#| label: tbl-h06d-prep-sensitivities
#| tbl-cap: "Stored sensitivity and complementary-analysis branches"

sensitivity_registry |>
  transmute(
    Branch = .data$branch,
    `Stored scope` = .data$stored_scope,
    Question = .data$scientific_question,
    Role = .data$inferential_role
  ) |>
  compact_gt(font_size = 10.5)
```

These branches answer distinct questions. The chest analysis is not an ocular
exposure estimate. Paired/common placement analyses are not direct sensor-
position effects. The gap-timing-unaware dataset is a preparation sensitivity,
and its own complete 15-slot near-eye families prevent a sensitivity result
from being folded into the primary family after inspection.

## Exploratory jointly adjusted daily models

The exploratory covariate-stability analysis places work/free day, activity,
and previous-night sleep in one model. Fourteen estimable metrics share the
same primary near-eye common frame of 715 participant-days from 137
participants and nine sites; the non-estimable darkest 10 h joint outcome is
retained as the named missing slot.

For mixed-model outcomes, the interaction specification is:

```r
response_value ~
  site * work_free_day +
  site * activity_status +
  site * previous_sleep_duration_centered_h +
  (1 | participant_key)
```

The repaired timing outcomes use the corresponding fixed-site linear model
without the participant random intercept and use participant-cluster HC3
covariance. Thus, each context is adjusted for the other two, while each
predictor may still vary by site. The six exploratory 15-slot families are
kept separate from the predictor-specific primary families.

```{r}
#| label: tbl-h06d-prep-joint-context
#| tbl-cap: "Common-frame support and route for the exploratory jointly adjusted daily analysis"

joint_registry |>
  transmute(
    Slot = .data$metric_slot,
    Metric = .data$manuscript_name,
    Family = paste(.data$response_family, .data$response_transform, sep = "; "),
    Route = recode(
      .data$route,
      mixed_model = "Mixed model",
      participant_cluster_HC3 = "Participant-cluster HC3"
    ),
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites
  ) |>
  gt::gt() |>
  gt::cols_hide(columns = "Slot") |>
  gt::fmt_integer(columns = c("Participant-days", "Participants", "Sites")) |>
  compact_gt(font_size = 10.5)
```

## Exploratory 30-minute clock-time GAMM

The clock-time analysis has a different row and estimand. Its outcome $Y$ is
the arithmetic mean melEDI in an admissible 30-minute bin. The model response
is $Z = \log_{10}(Y + 0.1\ \mathrm{lx})$. Exact zeros remain in this one-part
Gaussian model. The displayed inverse transform, $10^\eta - 0.1$, is a
back-transform of the conditional mean log response. It is geometric-mean-like
for $Y+0.1$ and is not raw-scale arithmetic $E[Y]$.

```{r}
#| label: tbl-h06d-prep-temporal-formula
#| tbl-cap: "Exact exploratory 30-minute response, model, and estimand"

temporal_formula |>
  transmute(
    Response = paste(.data$response_source, .data$response_transform, sep = "; "),
    Family = paste(.data$family, .data$link, sep = "; "),
    Formula = paste0("`", .data$formula, "`"),
    Estimand = .data$estimand,
    `Inverse-transform qualification` = .data$inverse_transform_qualification
  ) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "Formula") |>
  compact_gt(font_size = 10)
```

The model was fitted with `mgcv::bam()`, fREML, `discrete = TRUE`, one R
thread, and global 0/24 knots. The global clock smooth is cyclic cubic with
$k=12$. Work/free day, activity, and site use sum-to-zero factor smooths with
$k=12$ and the default thin-plate time marginal, not a cyclic marginal.
Participant uses a factor smooth with $k=10$, and participant-day uses a random
intercept smooth. Previous sleep is separated into between-participant and
within-participant smooth functions.

Rows are ordered by site, participant, local date, true source time, and clock
bin. A working AR(1) sequence restarts at each participant-day and at every
missing, non-one-to-one, or physically discontinuous true-time boundary. Each
frame's rho was estimated anew from its own preliminary full-model response
residuals and then fixed for the final fit.

```{r}
#| label: tbl-h06d-prep-temporal-support
#| tbl-cap: "Exact fitted support and working AR(1) parameter for the exploratory clock-time models"

temporal_support |>
  transmute(
    Dataset = .data$data_scenario_label,
    Placement = .data$placement_label,
    Sample = recode(
      .data$sample_scenario,
      all_available = "All available",
      paired_common = "Paired/common"
    ),
    `30-minute bins` = .data$observations_30_minute,
    `Participant-days` = .data$participant_days,
    Participants = .data$participants,
    Sites = .data$sites,
    `Exact zero bins` = .data$exact_zero_bins,
    `AR sequences` = .data$ar_sequences,
    Rho = .data$rho
  ) |>
  gt::gt(groupname_col = "Dataset") |>
  gt::fmt_integer(columns = c(
    "30-minute bins", "Participant-days", "Participants", "Sites",
    "Exact zero bins", "AR sequences"
  )) |>
  gt::fmt_number(columns = "Rho", decimals = 3) |>
  compact_gt(font_size = 10.5)
```

Whole-function tests and 95% intervals for clock-specific associations are
reported on the transformed scale. Every displayed interval is pointwise; no
simultaneous confidence band was constructed. The temporal analysis is
exploratory and does not replace a daily-metric slot or the selected hourly H06
analysis.

## Code, stored outputs, and reproducibility

### Execution order

```{r}
#| label: tbl-h06d-prep-code-map
#| tbl-cap: "H06 daily-metric modules in analytical order"

code_map |>
  transmute(
    Order = .data$sequence,
    Module = paste0("`", .data$module, "`"),
    Responsibility = .data$responsibility,
    `SHA-256 prefix` = substr(.data$sha256, 1L, 12L)
  ) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "Module") |>
  compact_gt(font_size = 10)
```

Contracts and frame construction are separated from fitting, model checking,
influence analysis, reader displays, and this provenance page. This separation
allows a stored result to be traced to the exact frame and script without
requiring the report renderer to execute scientific modeling code.

### Stored output trace

```{r}
#| label: tbl-h06d-prep-output-trace
#| tbl-cap: "Principal stored outputs used by the results report and this provenance companion"

output_trace |>
  transmute(
    Output = .data$role,
    Path = paste0("`", .data$relative_path, "`"),
    Type = .data$object_type,
    `SHA-256 prefix` = substr(.data$sha256, 1L, 12L),
    Bytes = .data$bytes
  ) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "Path") |>
  gt::fmt_integer(columns = "Bytes") |>
  compact_gt(font_size = 10)
```

Every durable reader figure has paired source data and alt text. The daily-grid
figures use original response units in labels. Reader-facing displays of
zero-containing, strongly right-skewed melEDI use the predefined symmetric-log
(symlog) display. It retains zero, is linear near zero, and uses base 10 with a
1 lx threshold, while the stored source values remain untransformed. P-values
are stored numerically, displayed to three
decimals with values below 0.001 shown as `<0.001`, and emphasized only under
their explicitly labelled raw or adjusted decision rule.

The synchronized environment uses R 4.6.1. Input and output manifests record
relative paths, SHA-256 identities, byte sizes, producing scripts, and
consequential package versions. The complete fitted-sample registry and the
figure source data are available as CSV files beside the model, model-check,
table, and figure artifacts.
