---
title: "H05 Stage 1: audit and proposed analysis"
subtitle: "Nature Health analysis audit — author gate"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    embed-resources: true
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

library(dplyr)
library(tidyr)
library(purrr)
library(tibble)
library(emmeans)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "renv.lock"))) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H05_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

audit_dir <- file.path(project_root, "audit/hypotheses/H05")
dir.create(audit_dir, recursive = TRUE, showWarnings = FALSE)
stopifnot(
  identical(
    normalizePath(audit_dir, winslash = "/", mustWork = TRUE),
    file.path(project_root, "audit/hypotheses/H05")
  )
)

write_audit_csv <- function(data, filename) {
  output <- file.path(audit_dir, filename)
  utils::write.csv(data, output, row.names = FALSE, na = "")
  invisible(output)
}

format_p <- function(x) {
  ifelse(x < 0.001, format(x, scientific = TRUE, digits = 3), sprintf("%.4f", x))
}

site_registry <- read.csv(
  file.path(project_root, "config/site_display_registry.csv"),
  check.names = FALSE
) |>
  arrange(display_order)

metric_display <- read.csv(
  file.path(project_root, "config/metric_display_registry.csv"),
  check.names = FALSE
)

site_levels <- site_registry$site
```

## Stage 1 verdict

**H05 is not decision-ready and Stage 2 remains blocked.** The submitted
analysis (hereafter **V0**) crossed four LEBA scores with 17 light-exposure
metrics, yielding 68 pooled participant-level correlations per placement. V0
did not fit the registered site-adjusted outcome model. It displayed
Spearman coefficients with Pearson-test p-values and applied BH adjustment to
each association p-value separately rather than to a 68-value vector.

On the frozen V0 near-eye inputs, reconstructing a consistent two-sided
Spearman test and a single 68-value BH vector still retains the two displayed
F2 associations. This calculation is diagnostic only: the submitted claims
remain reopened because site adjustment is absent, participant support was
not carried into the output, and the repaired metric pipeline changes the
analysis inputs and some metric definitions. The chest V0 result does not
survive the consistent Spearman/vector-BH reconstruction.

This document is an audit and an author-facing specification. It does **not**
fit any proposed H05 model, inspect candidate H05 results, bootstrap an
interval, run a sensitivity battery, edit a production notebook, or change a
shared decision or ledger.

### Meaning of V0

Throughout this audit, **V0** means the analysis implementation and artifacts
that produced the submitted manuscript result: the H05 blocks in
`RQ2.qmd`/`RQ2_chest.qmd`, their saved H01 metric workspaces, and the submitted
H05 figures and prose. It does not mean the audit-approved main analysis.

## Preregistered contract

The signed AsPredicted record states, exactly:

> H5: LEBA questionnaire factors correlate with selected personal light exposure metrics.

— `preregistration/AsPredicted #273407.pdf`, PDF p. 1.

The registered method on PDF p. 2 is:

```text
H5; Correlation matrices + linear models: Metric ~ LEBA + (1|Site)
```

The supporting analysis-plan DOCX names the four factors but gives the
ambiguous model phrase `Correlation ~ (1|Site)`. The signed PDF controls. It
requires both correlation matrices and a site-adjusted metric model, but does
not state which is primary and does not enumerate the selected metrics. Those
two ambiguities are the open H05-G1 and DEV-029 author gates.

### Contract-source verification

| Source | Verification | Stage 1 use |
|---|---|---|
| `preregistration/AsPredicted #273407.pdf` | SHA-256 `fce192c88676602b72013e4872a03b535a2f4ef59d34de3a9e1c47527742b7ad`; two searchable pages; exact query hits verified on pp. 1–2 | Controlling H05 wording and registered formula |
| `preregistration/MeLiDos_ZaunerEtAl_analysis_plan.docx` | SHA-256 `6e56766c18c476826c9c798799aa925752f11e5278f196797a583a13cc61ddf8`; text extracted without altering the source | Factor labels and documentation of the formula ambiguity |
| `audit/evidence/preregistration_contract.md` | Read-only project contract | Cross-check of signed record |

## Frozen inputs and exact V0 reconstruction

```{r}
#| label: frozen-input-contract

input_contract <- tibble::tribble(
  ~role, ~relative_path, ~sha256,
  "Near-eye V0 metric workspace", "data/H1_results.RData",
  "d274865818cc4a0364dae9a444b9be44be02424daa8ad74b9452840f3115b3dc",
  "Chest V0 metric workspace", "data/H1_results_chest.RData",
  "01f91db675d1aa6ae96b6aa7bbc45b2f86445e851a37020d1fe5335d802de9b5",
  "Normalized LEBA input", "artifacts/06_model_data/normalized_inputs/leba.rds",
  "71882566d4665dde757a13b18e56c1f7308e855ea22bced4384e91254e64e636"
)

input_contract <- input_contract |>
  mutate(
    observed_sha256 = vapply(
      file.path(project_root, relative_path),
      digest::digest,
      character(1),
      algo = "sha256",
      file = TRUE
    ),
    hash_verified = observed_sha256 == sha256
  )

stopifnot(all(input_contract$hash_verified))
knitr::kable(input_contract, caption = "Frozen V0 analysis inputs verified at render time.")
```

The H05 code loads the H01 fitted-result workspaces only to recover metric
tables, retains rows with a non-missing H01 `response`, and therefore creates
an unnecessary dependency on upstream fitted-result objects. Stage 2 should
consume a hypothesis-specific model-data artifact instead.

### LEBA scale recovery and verification

```{r}
#| label: leba-scale-audit

leba_raw <- readRDS(file.path(
  project_root,
  "artifacts/06_model_data/normalized_inputs/leba.rds"
))

factor_contract <- tibble::tribble(
  ~factor_id, ~factor_label, ~first_item, ~last_item,
  ~possible_min, ~possible_max, ~reverse_item,
  "leba_f2", "Spending time outdoors", 4L, 9L, 6L, 30L, "leba_f2_04",
  "leba_f3", "Using phones and smartwatches in bed before sleep", 10L, 14L, 5L, 25L, NA_character_,
  "leba_f4", "Controlling and using ambient light before bedtime", 15L, 18L, 4L, 20L, NA_character_,
  "leba_f5", "Using light in the morning and during daytime", 19L, 23L, 5L, 25L, NA_character_
)

expected_levels <- c("Never", "Rarely", "Sometimes", "Often", "Always")

item_value <- function(x) {
  stopifnot(is.factor(x), identical(levels(x), expected_levels))
  as.integer(x)
}

score_from_items <- function(factor_id) {
  contract <- factor_contract |> filter(.data$factor_id == !!factor_id)
  item_ids <- sprintf(
    "%s_%02d", factor_id, seq.int(contract$first_item, contract$last_item)
  )
  item_matrix <- vapply(leba_raw[item_ids], item_value, integer(nrow(leba_raw)))
  if (!is.na(contract$reverse_item)) {
    item_matrix[, contract$reverse_item] <- 6L - item_matrix[, contract$reverse_item]
  }
  rowSums(item_matrix)
}

score_audit <- factor_contract |>
  rowwise() |>
  mutate(
    calculated = list(score_from_items(factor_id)),
    stored = list(leba_raw[[factor_id]]),
    participants = nrow(leba_raw),
    score_missing_n = sum(is.na(stored)),
    observed_min = min(stored, na.rm = TRUE),
    observed_max = max(stored, na.rm = TRUE),
    mismatch_n = sum(calculated != stored, na.rm = TRUE),
    score_verified = mismatch_n == 0L
  ) |>
  ungroup() |>
  select(
    factor_id, factor_label, participants, first_item, last_item,
    reverse_item, possible_min, possible_max, observed_min, observed_max,
    score_missing_n, mismatch_n, score_verified
  )

item_audit <- map_dfr(seq_len(nrow(factor_contract)), function(i) {
  contract <- factor_contract[i, ]
  item_ids <- sprintf(
    "%s_%02d",
    contract$factor_id,
    seq.int(contract$first_item, contract$last_item)
  )
  map_dfr(item_ids, function(item_id) {
    x <- leba_raw[[item_id]]
    tibble(
      factor_id = contract$factor_id,
      factor_label = contract$factor_label,
      item_id = item_id,
      item_label = attr(x, "label"),
      reverse_coded = identical(item_id, contract$reverse_item),
      ordered_levels = paste(levels(x), collapse = " < "),
      missing_n = sum(is.na(x)),
      observed_min = min(as.integer(x), na.rm = TRUE),
      observed_max = max(as.integer(x), na.rm = TRUE)
    )
  })
})

stopifnot(
  nrow(leba_raw) == 184L,
  n_distinct(paste(leba_raw$site, leba_raw$Id, sep = "::")) == 184L,
  nrow(item_audit) == 20L,
  all(score_audit$score_verified),
  all(item_audit$missing_n == 0L)
)

write_audit_csv(item_audit, "h05_stage1_leba_factor_audit.csv")
knitr::kable(score_audit, caption = "Recovered and independently verified LEBA score construction.")
```

All 20 items use the ordered coding
`Never = 1 < Rarely = 2 < Sometimes = 3 < Often = 4 < Always = 5`.
F2 is the sum of items 04–09 after reversing only item 04 (`6 - item 04`);
F3, F4, and F5 are direct sums. All 184 stored scores reproduce exactly and
there are no missing item or factor values. Higher values indicate more of
the named behaviour. For F4 this means more frequent light/screen-control or
light-reduction behaviours before sleep, not generically “more ambient
light.” Factor 1 (filtered/tinted glasses) was not administered because the
study spectacles had no optical filters.

```{r}
#| label: leba-site-support

leba_site_support <- leba_raw |>
  select(site, Id, all_of(factor_contract$factor_id)) |>
  pivot_longer(
    cols = all_of(factor_contract$factor_id),
    names_to = "factor_id",
    values_to = "score"
  ) |>
  group_by(site, factor_id) |>
  summarise(
    participants = n_distinct(Id),
    complete_scores = sum(!is.na(score)),
    missing_scores = sum(is.na(score)),
    observed_min = min(score, na.rm = TRUE),
    observed_max = max(score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  left_join(select(factor_contract, factor_id, factor_label), by = "factor_id") |>
  left_join(site_registry, by = "site") |>
  arrange(display_order, factor_id) |>
  select(
    site, display_name, display_order, color_hex, factor_id, factor_label,
    participants, complete_scores, missing_scores, observed_min, observed_max
  )

write_audit_csv(leba_site_support, "h05_stage1_leba_site_support.csv")

leba_site_support |>
  distinct(site, display_name, display_order, color_hex, participants) |>
  arrange(display_order) |>
  knitr::kable(caption = "LEBA participants by submitted site order and display convention.")
```

### Exact V0 metric manifest

```{r}
#| label: v0-metric-manifest

factor_labels <- setNames(factor_contract$factor_label, factor_contract$factor_id)

metric_map <- tibble::tribble(
  ~v0_order, ~v0_name, ~metric_id, ~manuscript_name,
  1L, "interdaily_stability", "interdaily_stability", "Interdaily stability",
  2L, "intradaily_variability", "intradaily_variability", "Intradaily variability",
  3L, "Mean", "daily_geometric_mean_medi", "Mean melEDI",
  4L, "brightest_10h_mean", "m10_mean_medi", "Brightest 10 h mean",
  5L, "brightest_10h_midpoint", "m10_midpoint", "Midpoint of the brightest 10 hours",
  6L, "darkest_10h_mean", "l10_mean_medi", "Darkest 10 h mean",
  7L, "darkest_10h_midpoint", "l10_midpoint", "Midpoint of the darkest 10 hours",
  8L, "duration_above_1000", "duration_above_1000", "Time above 1,000 lx melEDI",
  9L, "period_above_250", "longest_bout_above_250", "Longest continuous period above 250 lx melEDI",
  10L, "mean_timing_above_250", "mean_timing_above_250", "Mean timing of exposure above 250 lx melEDI",
  11L, "first_timing_above_250", "first_timing_above_250", "First light timing above 250 lx melEDI",
  12L, "last_timing_above_250", "last_timing_above_250", "Last light timing above 250 lx melEDI",
  13L, "dose", "dose_observed_medi", "melEDI dose",
  14L, "MDER", "mder_ratio_of_integrals", "Melanopic daylight efficacy ratio",
  15L, "duration_below_10_pre-sleep", "duration_below_10_pre_sleep", "Time below 10 lx melEDI before sleep",
  16L, "duration_below_1_sleep", "duration_below_1_sleep_environment", "Time below 1 lx melEDI during sleep",
  17L, "duration_above_250_wake", "duration_above_250_wake", "Time above 250 lx melEDI during wake"
)

stopifnot(nrow(metric_map) == 17L, nrow(factor_contract) == 4L)

metric_map |>
  mutate(planned_associations = nrow(factor_contract)) |>
  select(v0_order, v0_name, manuscript_name, metric_id, planned_associations) |>
  knitr::kable(caption = "The exact 17-member V0 metric set (17 × 4 = 68 associations per placement).")
```

M10/L10 onset and offset variables are not in this 17-member set and remain
excluded under approved decision METRIC-009. The full-day “Time above 250 lx
melEDI” outcome is also not an H05 member; only the distinct wake-state
outcome “Time above 250 lx melEDI during wake” is included.

### Participant aggregation, pair counts, and correlations

```{r}
#| label: reproduce-v0

load_metric_workspace <- function(relative_path) {
  workspace <- new.env(parent = emptyenv())
  load(file.path(project_root, relative_path), envir = workspace)
  if (!exists("metrics", envir = workspace, inherits = FALSE)) {
    stop("The V0 workspace does not contain `metrics`: ", relative_path)
  }
  workspace$metrics
}

reproduce_placement <- function(relative_path, placement) {
  metrics <- load_metric_workspace(relative_path) |>
    filter(!is.na(response)) |>
    inner_join(metric_map, by = c("name" = "v0_name")) |>
    arrange(v0_order)

  stopifnot(
    nrow(metrics) == 17L,
    identical(metrics$v0_order, seq_len(17L))
  )

  participant_summaries <- pmap(
    list(
      metrics$data, metrics$name, metrics$type, metrics$metric_id,
      metrics$manuscript_name, metrics$v0_order
    ),
    function(data, name, type, metric_id, manuscript_name, v0_order) {
      participant <- data |>
        group_by(site, Id) |>
        summarise(
          metric_summary = mean(metric, na.rm = TRUE),
          contributing_rows = n(),
          contributing_days = if (type == "participant") {
            NA_integer_
          } else {
            n_distinct(Date[!is.na(metric)])
          },
          nonmissing_metric_rows = sum(!is.na(metric)),
          .groups = "drop"
        )

      # IS and IV are participant-level in V0. Their represented-day support
      # is recovered from the colocated daily Mean registry and is explicitly
      # labelled as a proxy rather than an input to their calculation.
      if (type == "participant") {
        mean_position <- match("Mean", metrics$name)
        day_source <- metrics$data[[mean_position]] |>
          group_by(site, Id) |>
          summarise(
            represented_days_proxy = n_distinct(Date[!is.na(metric)]),
            .groups = "drop"
          )
        participant <- participant |>
          left_join(day_source, by = c("site", "Id")) |>
          mutate(contributing_days = represented_days_proxy) |>
          select(-represented_days_proxy)
      }

      participant |>
        mutate(
          placement = placement,
          v0_order = v0_order,
          v0_name = name,
          metric_id = metric_id,
          manuscript_name = manuscript_name,
          analysis_unit = type,
          .before = 1
        )
    }
  ) |>
    list_rbind()

  association_frame <- participant_summaries |>
    left_join(
      select(leba_raw, site, Id, all_of(factor_contract$factor_id)),
      by = c("site", "Id")
    ) |>
    pivot_longer(
      cols = all_of(factor_contract$factor_id),
      names_to = "factor_id",
      values_to = "factor_score"
    ) |>
    filter(is.finite(metric_summary), !is.na(factor_score))

  associations <- association_frame |>
    group_by(
      placement, v0_order, metric_id, manuscript_name, analysis_unit,
      factor_id
    ) |>
    summarise(
      spearman_rho = cor(metric_summary, factor_score, method = "spearman"),
      v0_pearson_raw_p = cor.test(
        metric_summary, factor_score, method = "pearson"
      )$p.value,
      consistent_spearman_raw_p = cor.test(
        metric_summary, factor_score, method = "spearman", exact = FALSE
      )$p.value,
      participants = n_distinct(paste(site, Id, sep = "::")),
      sites = n_distinct(site),
      represented_days_total = sum(contributing_days),
      days_per_participant_min = min(contributing_days),
      days_per_participant_median = median(contributing_days),
      days_per_participant_max = max(contributing_days),
      .groups = "drop"
    ) |>
    mutate(
      factor_label = unname(factor_labels[factor_id]),
      # This deliberately reproduces V0: p.adjust receives one scalar inside
      # every metric-by-factor group, with n forced to 68.
      v0_scalar_bh_p = vapply(
        v0_pearson_raw_p,
        function(p) p.adjust(p, method = "BH", n = 68L),
        numeric(1)
      ),
      # Audit-only consistency check on the frozen V0 summaries.
      consistent_spearman_vector_bh_p = p.adjust(
        consistent_spearman_raw_p, method = "BH"
      ),
      v0_display_flag = v0_scalar_bh_p <= 0.05,
      consistent_spearman_vector_bh_flag =
        consistent_spearman_vector_bh_p <= 0.05
    ) |>
    arrange(v0_order, factor_id)

  metric_support <- participant_summaries |>
    group_by(
      placement, v0_order, metric_id, manuscript_name, analysis_unit
    ) |>
    summarise(
      participant_summaries = n(),
      finite_participant_summaries = sum(is.finite(metric_summary)),
      represented_days_total = sum(contributing_days),
      days_per_participant_min = min(contributing_days),
      days_per_participant_median = median(contributing_days),
      days_per_participant_max = max(contributing_days),
      source_rows = sum(contributing_rows),
      nonmissing_source_rows = sum(nonmissing_metric_rows),
      sites = n_distinct(site),
      .groups = "drop"
    ) |>
    arrange(v0_order)

  pair_counts_by_site <- association_frame |>
    group_by(
      placement, v0_order, metric_id, manuscript_name, factor_id, site
    ) |>
    summarise(
      participants = n_distinct(Id),
      represented_days_total = sum(contributing_days),
      days_per_participant_min = min(contributing_days),
      days_per_participant_median = median(contributing_days),
      days_per_participant_max = max(contributing_days),
      .groups = "drop"
    ) |>
    left_join(site_registry, by = "site") |>
    left_join(select(factor_contract, factor_id, factor_label), by = "factor_id") |>
    arrange(v0_order, factor_id, display_order) |>
    select(
      placement, v0_order, metric_id, manuscript_name, factor_id,
      factor_label, site, display_name, display_order, color_hex,
      participants, represented_days_total, days_per_participant_min,
      days_per_participant_median, days_per_participant_max
    )

  list(
    metric_registry = metrics,
    participant_summaries = participant_summaries,
    association_frame = association_frame,
    associations = associations,
    metric_support = metric_support,
    pair_counts_by_site = pair_counts_by_site
  )
}

near_eye <- reproduce_placement("data/H1_results.RData", "near-eye")
chest <- reproduce_placement("data/H1_results_chest.RData", "chest")

v0_associations <- bind_rows(near_eye$associations, chest$associations)
v0_metric_support <- bind_rows(near_eye$metric_support, chest$metric_support)
v0_pair_counts_by_site <- bind_rows(
  near_eye$pair_counts_by_site,
  chest$pair_counts_by_site
)

stopifnot(
  nrow(near_eye$associations) == 68L,
  nrow(chest$associations) == 68L,
  nrow(v0_associations) == 136L
)

write_audit_csv(v0_associations, "h05_stage1_v0_associations.csv")
write_audit_csv(v0_metric_support, "h05_stage1_v0_metric_support.csv")
write_audit_csv(
  v0_pair_counts_by_site,
  "h05_stage1_v0_pair_counts_by_site.csv"
)

placement_overview <- tibble::tribble(
  ~placement, ~participants, ~represented_days, ~sites,
  "near-eye", 141L, 811L, 9L,
  "chest", 154L, 897L, 8L
)

knitr::kable(
  placement_overview,
  caption = "V0 placement-wide source support before metric-specific pairwise deletion."
)
```

V0 takes an arithmetic mean of every daily metric within `site × Id`, uses
IS/IV directly, requires no minimum number of days, left-joins the four LEBA
scores, and drops missing pairs cell by cell. This includes arithmetic means
of clock-valued metrics. The exact 136 placement × metric × factor rows and
their site-resolved pair counts are persisted as audit CSVs rather than
hidden in a plot.

```{r}
#| label: metric-support-table

v0_metric_support |>
  mutate(
    days_per_participant = paste0(
      days_per_participant_min, "/",
      days_per_participant_median, "/",
      days_per_participant_max
    )
  ) |>
  select(
    placement, manuscript_name, analysis_unit,
    finite_participant_summaries, represented_days_total,
    days_per_participant, sites
  ) |>
  knitr::kable(
    caption = paste(
      "Exact V0 metric support; days/person is minimum/median/maximum.",
      "For IS/IV, represented days are the colocated daily-Mean proxy."
    )
  )
```

Near-eye timing summaries have 778 contributing participant-days; MDER has
725 days and 140 finite participant summaries; pre-sleep has 780 days; sleep
has 790 days; and wake-state time above 250 lx melEDI has 755 days and 140
finite participant summaries. Chest timing summaries have 867 days; MDER 729;
pre-sleep 867; sleep 878; and the wake-state duration 839 days with 153 finite
participant summaries. The complete audit table above preserves the exact
metric-level denominators.

### V0 questionnaire-by-site screen

```{r}
#| label: reproduce-v0-site-screen

leba_scores <- leba_raw |>
  select(site, Id, all_of(factor_contract$factor_id)) |>
  mutate(site = factor(site, levels = site_levels))

site_screen_objects <- map(factor_contract$factor_id, function(factor_id) {
  fit <- lm(reformulate("site", response = factor_id), data = leba_scores)
  marginal <- emmeans(fit, ~site)
  omnibus <- as.data.frame(joint_tests(marginal)) |>
    filter(.data[["model term"]] == "site")
  contrasts <- as.data.frame(contrast(marginal))
  list(fit = fit, marginal = marginal, omnibus = omnibus, contrasts = contrasts)
})
names(site_screen_objects) <- factor_contract$factor_id

v0_site_screen <- map_dfr(names(site_screen_objects), function(factor_id) {
  object <- site_screen_objects[[factor_id]]
  tibble(
    factor_id = factor_id,
    factor_label = factor_labels[[factor_id]],
    participants = stats::nobs(object$fit),
    sites = nlevels(leba_scores$site),
    df1 = object$omnibus$df1,
    df2 = object$omnibus$df2,
    f_ratio = object$omnibus$F.ratio,
    raw_p = object$omnibus$p.value
  )
}) |>
  # Exact code semantics: after unnesting, p.adjust sees all four values.
  mutate(
    v0_vector_bh_p = p.adjust(raw_p, method = "BH", n = 4L),
    passes_v0_screen = v0_vector_bh_p <= 0.05
  )

v0_site_contrasts <- map_dfr(
  v0_site_screen$factor_id[v0_site_screen$passes_v0_screen],
  function(factor_id) {
    current_factor_id <- factor_id
    current_factor_label <- unname(factor_labels[[current_factor_id]])
    site_screen_objects[[factor_id]]$contrasts |>
      as_tibble() |>
      mutate(
        factor_id = current_factor_id,
        factor_label = current_factor_label
      )
  }
) |>
  filter(p.value <= 0.05) |>
  select(factor_id, factor_label, contrast, estimate, SE, df, t.ratio, p.value)

write_audit_csv(v0_site_screen, "h05_stage1_v0_site_screen.csv")
write_audit_csv(v0_site_contrasts, "h05_stage1_v0_site_contrasts.csv")

v0_site_screen |>
  mutate(across(c(raw_p, v0_vector_bh_p), format_p)) |>
  knitr::kable(caption = "Reproduced V0 four-factor questionnaire site screen.")

v0_site_contrasts |>
  mutate(p.value = format_p(p.value)) |>
  knitr::kable(
    digits = 3,
    caption = "Effect contrasts retained by the V0 screen and default emmeans adjustment."
  )
```

The exact R evaluation shows that the four omnibus site-screen p-values were
adjusted together after `unnest()`: F2 adjusted p = 0.0076 and F5 adjusted
p = 0.0164. This corrects the migration map’s description of this particular
step as four scalar adjustments. The two retained effect contrasts are UCR
for F2 (estimate −2.703, adjusted p = 0.0002) and BAUA for F5 (estimate
−2.239, adjusted p = 0.0250), both relative to the unweighted site mean.

That implementation correction does **not** rescue the inferential logic.
Questionnaire-by-site tests are diagnostics; they are not the registered
outcome model, non-rejection for F3/F4 is not evidence of no confounding, and
site adjustment cannot be selected by this screen. A scoped shared-change
request accompanies this audit so the migration map can be corrected without
editing it here.

### Coefficient, test, and multiplicity reconstruction

```{r}
#| label: highlighted-v0-results

highlighted_v0 <- v0_associations |>
  filter(
    placement == "near-eye",
    factor_id == "leba_f2",
    metric_id %in% c("dose_observed_medi", "duration_above_1000")
  ) |>
  select(
    manuscript_name, spearman_rho, v0_pearson_raw_p, v0_scalar_bh_p,
    consistent_spearman_raw_p, consistent_spearman_vector_bh_p,
    participants, sites, represented_days_total
  )

highlighted_v0 |>
  mutate(
    spearman_rho = round(spearman_rho, 3),
    across(contains("_p"), format_p)
  ) |>
  knitr::kable(
    caption = paste(
      "The two V0 near-eye highlights: submitted computation versus",
      "a consistent Spearman/vector-BH audit on the same frozen summaries."
    )
  )
```

V0 computes `cor(..., method = "spearman")` but calls `cor.test()` without a
method, which defaults to Pearson. Inside each metric-by-factor group it then
passes that one Pearson p-value to `p.adjust(..., n = 68)`. This is not a
68-value BH adjustment; for one scalar it is multiplication by 68, truncated
at one.

On the same frozen near-eye participant summaries, the consistent Spearman
raw/vector-BH results are ρ = 0.272, raw p = 0.00111, adjusted p = 0.0377 for
melEDI dose and ρ = 0.290, raw p = 0.000478, adjusted p = 0.0325 for time
above 1,000 lx melEDI. Their persistence here must not be read as the final
H05 result because no site-adjusted model or repaired metric input has been
evaluated.

```{r}
#| label: chest-v0-result

chest_v0_flags <- v0_associations |>
  filter(v0_display_flag | consistent_spearman_vector_bh_flag) |>
  select(
    placement, manuscript_name, factor_id, spearman_rho,
    v0_pearson_raw_p, v0_scalar_bh_p,
    consistent_spearman_raw_p, consistent_spearman_vector_bh_p,
    v0_display_flag, consistent_spearman_vector_bh_flag,
    participants, sites, represented_days_total
  ) |>
  mutate(
    spearman_rho = round(spearman_rho, 3),
    across(contains("_p"), format_p)
  )

knitr::kable(
  chest_v0_flags,
  caption = "All V0 or consistent-Spearman/vector-BH flagged cells across placements."
)
```

The chest V0 figure highlights F2 with dose (ρ = 0.202; displayed p = 0.042).
The consistent Spearman raw p-value is 0.0119 and the chest-specific
68-value BH result is 0.204, so no chest cell remains flagged. The chest H05
prose is copied from the near-eye interpretation and therefore does not
describe the chest result.

### Outcome-name and interpretation audit

The two near-eye highlights are **melEDI dose** and **Time above 1,000 lx
melEDI**. The V0 figure and `index.qmd` identify the threshold as 1,000 lx;
the final sentence of the `RQ2.qmd` H05 interpretation instead says “Duration
above 250 lx melanopic EDI.” That phrase is wrong. It cannot refer to the
full-day 250-lx outcome because that outcome is absent from the 17-member H05
set, and it cannot silently substitute the distinct wake-state 250-lx metric.

V0 also squares Spearman’s rho and describes the result as variance explained.
For a rank correlation, ρ² is not an ordinary proportion of outcome variance
explained. Any future claim should report the association estimate and
uncertainty on its declared scale without causal or variance-explanation
language.

### V0 paired-placement support

```{r}
#| label: v0-paired-support

paired_support_one <- function(v0_order, v0_name, metric_id, manuscript_name) {
  near_row <- near_eye$metric_registry |> filter(.data$v0_order == !!v0_order)
  chest_row <- chest$metric_registry |> filter(.data$v0_order == !!v0_order)
  stopifnot(nrow(near_row) == 1L, nrow(chest_row) == 1L)

  analysis_unit <- near_row$type[[1]]
  stopifnot(identical(analysis_unit, chest_row$type[[1]]))

  if (analysis_unit == "participant") {
    near_keys <- near_eye$participant_summaries |>
      filter(.data$v0_order == !!v0_order, is.finite(metric_summary)) |>
      select(site, Id)
    chest_keys <- chest$participant_summaries |>
      filter(.data$v0_order == !!v0_order, is.finite(metric_summary)) |>
      select(site, Id)
    common <- inner_join(near_keys, chest_keys, by = c("site", "Id"))
    return(tibble(
      v0_order = v0_order,
      metric_id = metric_id,
      manuscript_name = manuscript_name,
      analysis_unit = analysis_unit,
      paired_participants = nrow(common),
      common_participant_days = NA_integer_,
      sites = n_distinct(common$site)
    ))
  }

  near_days <- near_row$data[[1]] |>
    filter(is.finite(metric)) |>
    select(site, Id, Date) |>
    distinct()
  chest_days <- chest_row$data[[1]] |>
    filter(is.finite(metric)) |>
    select(site, Id, Date) |>
    distinct()
  common <- inner_join(near_days, chest_days, by = c("site", "Id", "Date"))

  tibble(
    v0_order = v0_order,
    metric_id = metric_id,
    manuscript_name = manuscript_name,
    analysis_unit = analysis_unit,
    paired_participants = n_distinct(paste(common$site, common$Id, sep = "::")),
    common_participant_days = nrow(common),
    sites = n_distinct(common$site)
  )
}

v0_paired_support <- pmap_dfr(metric_map, paired_support_one) |>
  arrange(v0_order)

stopifnot(
  all(v0_paired_support$paired_participants <= 112L),
  all(v0_paired_support$sites == 8L)
)

write_audit_csv(v0_paired_support, "h05_stage1_v0_paired_support.csv")
knitr::kable(
  v0_paired_support,
  caption = "Exact V0 common near-eye/chest support by metric."
)
```

There are 112 participants shared across placements and eight sites. For most
daily metrics the exact common support is 640 participant-days; timing
metrics have 603; MDER 476 days/111 participants; pre-sleep 615; sleep 621;
and wake-state time above 250 lx melEDI 592 days/111 participants. IS/IV have
112 paired participants. These are V0 denominators only and must not be mixed
with the repaired main-pipeline paired-day manifest used in Stage 2.

## Finding register

```{r}
#| label: finding-register

h05_findings <- tibble::tribble(
  ~finding_id, ~severity, ~status, ~finding, ~evidence, ~required_resolution,
  "H05-S1-001", "major", "open_author_gate",
  "The registered site-adjusted Metric ~ LEBA + (1|Site) analysis is absent; V0 reports pooled correlations.",
  "AsPredicted p. 2; RQ2.qmd:1578–1720; RQ2_chest.qmd:1579–1721; DEV-028",
  "Choose H05-G1 before any proposed model is fitted.",
  "H05-S1-002", "major", "confirmed",
  "V0 displays Spearman rho but obtains each p-value from the default Pearson cor.test().",
  "RQ2.qmd H05 summarize block; frozen-input R reconstruction",
  "Use one correlation method for coefficient, test, and interval.",
  "H05-S1-003", "major", "confirmed",
  "Each association p-value is BH-adjusted as a scalar with n = 68; no 68-value vector is assembled.",
  "RQ2.qmd H05 grouped summarize block; IMP-001",
  "Assemble the complete declared family first, then call p.adjust once.",
  "H05-S1-004", "major", "confirmed",
  "The preliminary LEBA-by-site screen cannot justify omitting site from the outcome model.",
  "Four lm(factor ~ site) screens; DEV-028",
  "Treat the screen as a diagnostic only and retain the approved site structure.",
  "H05-S1-005", "moderate", "confirmed",
  "V0 gives every participant equal weight after arithmetic averaging, requires no minimum days, and does not persist metric-specific support.",
  "Exact V0 preparation and reconstructed support tables",
  "Approve the Stage 2 analysis unit and persist exact model frames and denominators.",
  "H05-S1-006", "major", "confirmed",
  "Clock metrics are arithmetically averaged in V0 without an unwrap/support rule.",
  "V0 mean(metric, na.rm = TRUE) for all daily metrics",
  "Apply the approved clock rules; do not default all clock variables to circular summaries.",
  "H05-S1-007", "major", "confirmed",
  "The RQ2 interpretation misnames the highlighted 1,000-lx outcome as duration above 250 lx.",
  "RQ2.qmd:1720 versus Fig6 and index.qmd:337–339",
  "Use Time above 1,000 lx melEDI; keep the distinct wake-state 250-lx metric separate.",
  "H05-S1-008", "moderate", "confirmed",
  "V0 describes squared Spearman rho as variance explained.",
  "RQ2.qmd H05 interpretation",
  "Report rho or model-scale effect estimates and intervals without variance-explanation language.",
  "H05-S1-009", "major", "confirmed",
  "Chest prose duplicates the near-eye interpretation and is inconsistent with the chest V0 matrix.",
  "RQ2_chest.qmd H05 interpretation and chest Fig6",
  "Generate placement-specific prose only from verified result rows.",
  "H05-S1-010", "major", "open_author_gate",
  "The registration does not enumerate the four-factor × 17-metric family or its primary/secondary relation.",
  "AsPredicted pp. 1–2; DEV-029",
  "Approve the exact factor, metric, scenario, and multiplicity manifest.",
  "H05-S1-011", "moderate", "shared_change_requested",
  "The migration map says the four site-screen p-values were scalar-adjusted, but source-level R evaluation shows one four-value adjustment after unnesting.",
  "RQ2.qmd H05 site block and exact R reproduction in this document",
  "Correct the shared migration description; do not alter the separate association-level scalar-BH finding."
)

write_audit_csv(h05_findings, "h05_stage1_findings.csv")
knitr::kable(h05_findings, caption = "H05 Stage 1 findings and required disposition.")
```

The relevant open governance records are DEV-028 (site-adjusted model) and
DEV-029 (metric/factor family). IMP-006 is the coefficient/test mismatch and
IMP-001 is the association-level scalar adjustment. No ledger status is
changed by this worker deliverable.

## H05-G1 options and recommendation

### Option A — registered site-adjusted primary (**recommended**)

Use the registered, metric-appropriate site-adjusted LEBA model as the
primary inferential analysis. Keep the participant-summary Spearman matrix as
descriptive continuity output, reporting estimates, intervals, exact pair
counts, ties, and support without a second significance screen. This option
most closely preserves the signed contract, retains repeated participant-day
information for 15 daily outcomes, and avoids using a predictor-only site
screen to decide the outcome model.

For each of four LEBA factors and 17 metrics, compare a full and reduced model
on the identical complete-case model frame. The factor score is centered on
its participant mean. Report the slope per one raw score point and also
rescaled per one participant-level SD; do not z-transform separately within
site. The response transformation and family follow the already approved H01
metric registry. No causal language is permitted.

### Option B — participant-correlation primary

Retain 68 two-sided participant-level Spearman correlations as the primary
inferential estimand for continuity, with a single 68-value BH family. The
registered site-adjusted models then become a named sensitivity/deviation.
This option requires explicit approval of the deviation from the registered
model and an approved metric-specific participant-summary rule, especially
for clock metrics and unequal day support.

### Recommendation rationale

Choose Option A. Site is a design feature and a plausible common cause of
both habitual behaviour scores and light exposure. With only nine sites, the
site random-effect estimate may be fragile, so the registered random-site
form and a fixed-site sensitivity should both be reported. That limitation is
a reason for transparent diagnostics, not for deleting site from the model.

## Proposed formula and test contract — not fitted

The following code creates and prints the exact formula objects. It is
intentionally evaluated and unfolded so that the operational contract is
machine-visible. No model is fitted in Stage 1.

```{r}
#| label: h05-formula-contract
#| code-fold: false
#| code-summary: "Formula contract (always shown)"

h05_formula <- function(response = "response_value", fixed, random = character()) {
  right_hand_side <- c(fixed, random)
  if (length(right_hand_side) == 0L) {
    right_hand_side <- "1"
  }
  stats::as.formula(paste(response, "~", paste(right_hand_side, collapse = " + ")))
}

h05_formulas <- list(
  registered_exact = stats::as.formula(
    "response_value ~ leba_centered + (1 | site)"
  ),
  participant_day_full = h05_formula(
    fixed = "leba_centered",
    random = c("(1 | site)", "(1 | site:participant)")
  ),
  participant_day_reduced = h05_formula(
    fixed = character(),
    random = c("(1 | site)", "(1 | site:participant)")
  ),
  participant_full = h05_formula(
    fixed = "leba_centered",
    random = "(1 | site)"
  ),
  participant_reduced = h05_formula(
    fixed = character(),
    random = "(1 | site)"
  ),
  fixed_site_participant_day_full = h05_formula(
    fixed = c("site", "leba_centered"),
    random = "(1 | site:participant)"
  ),
  fixed_site_participant_day_reduced = h05_formula(
    fixed = "site",
    random = "(1 | site:participant)"
  ),
  fixed_site_participant_full = h05_formula(
    fixed = c("site", "leba_centered")
  ),
  fixed_site_participant_reduced = h05_formula(
    fixed = "site"
  ),
  v0_questionnaire_site_screen = stats::as.formula(
    "leba_score ~ site"
  ),
  v0_correlation_pairing_only = stats::as.formula(
    "metric_summary ~ leba_score"
  )
)

formula_contract <- enframe(h05_formulas, name = "formula_id", value = "formula") |>
  mutate(formula = vapply(formula, function(x) paste(deparse(x), collapse = " "), character(1)))

knitr::kable(formula_contract, caption = "Evaluated H05 formula objects; none are fitted here.")
```

The signed formula is the `registered_exact` row. Adding
`(1 | site:participant)` for participant-day outcomes is the proposed
operational refinement needed for repeated days and requires author approval
at H05-G1. IS and IV remain participant-level and therefore use no participant
intercept. In every inferential test, the reduced model removes only the LEBA
term; response transformation, family/link, random effects, and analysis rows
must otherwise be identical. Gaussian full/reduced comparisons use maximum
likelihood; a final descriptive full-model refit may use REML. Tweedie-log
comparisons use maximum likelihood.

The fixed-site sensitivity uses the submitted site order and sum-to-zero site
contrasts. LEBA is numeric and has no contrast coding. Reference handling,
centering constants, participant-level SDs, model-frame hashes, optimizer,
and package versions must be persisted.

### Proposed response-family package

```{r}
#| label: h05-response-family-contract

h01_contract_env <- new.env(parent = globalenv())
sys.source(
  file.path(project_root, "scripts/hypotheses/H01/h01_contract.R"),
  envir = h01_contract_env
)
future_metric_contract <- h01_contract_env$h01_metric_registry() |>
  left_join(
    select(metric_display, metric_id, manuscript_name, display_unit),
    by = "metric_id"
  ) |>
  arrange(metric_order) |>
  select(
    metric_order, metric_id, manuscript_name, analysis_unit, display_unit,
    response_family, response_transform, effect_scale, diagnostic_note
  )

stopifnot(nrow(future_metric_contract) == 17L)
knitr::kable(
  future_metric_contract,
  caption = "Proposed H05 response families inherited unchanged from the approved H01 metric contract."
)
```

This future 17-member contract uses the corrected time-sensitive melEDI dose
(`dose_time_sensitive_corrected_medi`), not V0’s observed-duration dose. It
also uses MDER as a ratio of paired-channel integrals. Those metric repairs
reopen the V0 results; they are not estimated here.

### Clock-valued outcomes

Clock time is not automatically circular. The proposed rules are:

1. M10 midpoint and threshold timing outcomes use the approved numeric clock
   representation only when support occupies one continuous, scientifically
   coherent range without a midnight split.
2. L10 midpoint subtracts 24 hours from values **strictly later than 16:00**
   before analysis. The approved noon-boundary conversion is a same-row,
   same-model sensitivity.
3. If a clock outcome violates the continuous-range condition, do not force a
   linear model or silently substitute a circular summary. Stop at a metric
   gate and approve the representation.
4. Under recommended Option A, the daily clock value is the response; there
   is no V0-style participant averaging in the primary model. A descriptive
   participant correlation must use a predeclared representation consistent
   with these same rules and must report its support.

M10/L10 onset and offset variables remain excluded; only their supported
levels and midpoints are eligible.

## Multiplicity contract for approval

Under recommended Option A:

- **H05-F1-primary:** the 68 near-eye full-versus-reduced LEBA tests form one
  complete vector and receive one BH adjustment. A planned but non-estimable
  member remains in the manifest with an NA p-value and reason; the family is
  not shortened after inspecting results.
- **H05-F1-MPD:** the approved manuscript-prepared-data scenario repeats the
  same logical 68-member family and is adjusted within that scenario.
- The participant-level Spearman matrix reports rho, a declared 95% interval,
  pair count, site count, ties, and represented days without inferential
  p-values or a significance screen.
- The paired/common chest analysis is complementary. If inferential p-values
  are retained, its 68 tests form a separate, explicitly labelled BH family;
  otherwise report estimates and intervals only.
- Paired-sample near-eye, fixed-site, leave-one-site-out, coverage, and
  response-family checks are stability diagnostics without additional
  significance screens. If p-values are desired for any of them, each family
  must be declared in full before fitting.
- The four questionnaire site tests are diagnostics and never choose H05
  covariates, outcomes, or family members.

Under Option B, the 68 two-sided Spearman tests are H05-F1-primary and receive
one vector-wide BH adjustment. If the site-adjusted models also report
inferential p-values, their 68 values form the separate H05-F2-secondary
family. Scalar calls to `p.adjust()` are prohibited in both options.

Raw p-values, family identifier, planned family size, available rank,
adjusted p-value, non-estimability reason, and scenario must be retained in
the result table.

## Placement and common-sample contract

Near-eye measurement is the main placement, following the approved placement
decision. Chest measurement is complementary and uses the exact paired/common
sample:

- participant-day outcomes pair on `site × participant × local date` after
  the same metric-specific admissibility rules in each placement;
- IS/IV pair on `site × participant`;
- the same approved formula, response family, scaling, and diagnostic rules
  are fitted separately by placement; placements are not pooled as independent
  observations;
- the near-eye model is rerun on the paired/common sample so that a near-eye
  versus chest difference can be separated from a sample-composition change;
- report estimate differences, 95% intervals, signs, and stability, with exact
  participant/day/site counts; and
- an all-available chest analysis is contextual only and requires explicit
  approval if it is to receive inferential p-values.

The V0 common-sample counts above are provenance checks, not the future
repaired-sample manifest.

## Required Stage 2 diagnostics

These checks are specifications, not analyses performed in Stage 1.

### Questionnaire and join integrity

- Assert one row per `site × participant`, the four approved factor IDs, the
  20-item key, factor ranges, reverse coding, missing-item rule, and score
  reproduction.
- Report factor distributions, ties, floor/ceiling frequencies, and
  within-site support using the submitted site names, order, and colours.
- Assert one-to-one LEBA joins and preserve participants with no eligible
  metric row in a reason-coded flow table.
- Persist centering means, raw-scale SDs, and exact factor direction text.

### Metric and model-frame integrity

- Start from the audited H05 metric-data producer, not H01 fitted results.
- For every placement × scenario × metric × factor row, persist eligible
  participants, participant-days, sites, missing LEBA, missing outcome,
  support failures, and the exact model-frame hash.
- Apply approved coverage Rule A in the main analysis and retain Rules B/C as
  named sensitivities. Preserve the exact all-zero-day exclusion and its
  inclusive sensitivity.
- Verify state-specific denominators, the longest-bout identifiability flag,
  MDER paired-channel support, time-sensitive dose inputs, units, bounds, and
  clock transformations before fitting.
- Full and reduced models must have identical observations in identical order.

### Model diagnostics and failure rules

- Record convergence code, warnings, optimizer, gradient/Hessian checks,
  random-effect variances, singularity, and failed refits.
- With only nine sites, examine random-site stability and the predeclared
  fixed-site/leave-one-site-out results. Do not delete the site effect because
  its estimated variance is small or a screen is non-significant.
- Check Gaussian residual shape, scale, heteroskedasticity, tail influence,
  fitted bounds, and participant/date residual dependence.
- For Tweedie-log outcomes, check zero mass, dispersion/power, tails, fitted
  physical bounds, and numerical stability.
- For transformed outcomes, report effects on both model and reader-facing
  scales. Preserve the exact transformation and offset.
- Flag influential participants and sites using rules fixed before inspecting
  H05 significance. Report stability rather than mechanically deleting them.
- A failed/non-estimable member stays in the 68-row registry with a reason.
  Any family/link switch, altered outcome definition, or undeclared row
  deletion reopens the author gate.

### Correlation diagnostics

- Use the same two-sided Spearman definition for rho and any p-value.
- Record ties and the interval method; show the bivariate rank/scatter pattern
  with site and participant influence visible.
- Report site-stratified and leave-one-site-out rho as diagnostics.
- A correlation interval requiring participant resampling is computation-heavy
  and must use the project’s pilot/runtime policy after the CI method and
  resampling count are approved. No bootstrap is run in this document.

## Predeclared sensitivity menu for author approval

Only the approved subset should be implemented in Stage 2. The candidates are:

1. registered random-site versus fixed-site model;
2. leave-one-site-out stability;
3. near-eye all-available versus paired/common-sample near-eye and chest;
4. approved coverage Rules B/C on the same scientific metric definitions;
5. all-zero-inclusive and relaxed native-epoch scenarios;
6. MDER support thresholds 0.70/0.90, two-device-day support, and waking-only
   variants;
7. state/timing support thresholds 0.70/0.90;
8. longest-bout exactly-identifiable-only rows;
9. L10 strict-after-16:00 primary rule versus the noon conversion;
10. corrected time-sensitive dose versus V0 observed-duration dose;
11. approved metric-specific response-family alternatives on the identical
    model frame; and
12. the manuscript-prepared-data scenario as its own complete 68-member
    analysis.

Sensitivity outputs should be classified by sign, practical magnitude,
interval overlap/width, convergence, and influence. They do not become new
confirmatory screens by accumulating p-values.

## V0-to-future claim disposition

| V0 statement or artifact | Stage 1 disposition | Condition for future use |
|---|---|---|
| “Only two of 68 comparisons were significant” | Reopened | Replace only from the approved complete primary 68-row family |
| Positive F2–melEDI-dose association | Provisional V0 reconstruction only | Requires approved corrected-dose input, site-adjusted result, diagnostics, multiplicity, and placement comparison |
| Positive F2–time-above-1,000-lx association | Provisional V0 reconstruction only | Same requirements; retain the 1,000-lx name |
| “Duration above 250 lx” as the highlighted threshold | Incorrect | Remove; do not substitute the distinct wake-state 250-lx outcome |
| ρ² as variance explained | Unsupported interpretation | Replace with estimate/interval language on the declared scale |
| LEBA can be analysed independent of site | Unsupported | Site screen remains descriptive; use the approved model structure |
| Near-eye interpretation copied into chest output | Incorrect | Generate placement-specific text from verified result rows |
| Broad habitual behaviour is less informative than immediate context | Overextended | At most a cautious observational comparison after complete, stable H05 results; no causal ranking |

No manuscript or figure edit is authorized at Stage 1. The current
`assets/Fig6.png` is an exact copy of the near-eye V0 figure, not a durable
source-data producer.

```{r}
#| label: v0-figure-provenance

v0_figure_provenance <- tibble(
  relative_path = c(
    "figures/Fig6.png",
    "figures/chest/Fig6.png",
    "assets/Fig6.png"
  )
) |>
  mutate(
    sha256 = vapply(
      file.path(project_root, relative_path),
      digest::digest,
      character(1),
      algo = "sha256",
      file = TRUE
    ),
    placement = c("near-eye", "chest", "near-eye copy in manuscript assets")
  )

stopifnot(v0_figure_provenance$sha256[1] == v0_figure_provenance$sha256[3])
knitr::kable(v0_figure_provenance, caption = "V0 H05 raster-artifact provenance.")
```

## Author decisions required before Stage 2

Stage 2 is blocked until the author records all of the following decisions:

1. **H05-G1 primary estimand:** approve Option A (recommended) or Option B.
2. **Factor manifest:** approve LEBA F2–F5, their item keys, reverse coding,
   direction, ranges, and complete-score rule; confirm Factor 1 exclusion.
3. **Metric manifest:** approve the exact 17 members, including corrected
   dose, ratio-of-integrals MDER, the two participant-level dynamics metrics,
   the five clock outcomes, exclusion of M10/L10 onset/offset, and exclusion
   of full-day time above 250 lx melEDI.
4. **Analysis unit and model hierarchy:** approve participant-day models for
   15 metrics, participant models for IS/IV, and the added nested participant
   intercept for daily rows.
5. **Response package:** approve the inherited H01 family, transform, link,
   effect scale, and any diagnostic-triggered alternatives.
6. **LEBA scaling:** approve mean-centered raw scores, effects per raw point
   and per one participant-level SD, with no within-site standardization.
7. **Site handling:** approve registered random-site primary plus fixed-site
   and leave-one-site-out diagnostics, including sum-to-zero fixed-site
   contrasts and the submitted site order.
8. **Multiplicity:** approve the exact primary 68-member family, the
   manuscript-prepared-data family, whether any complementary chest p-values
   are reported, and the no-secondary-screen rule for descriptive
   correlations/sensitivities.
9. **Clock rules:** approve the continuous-range requirement, strict
   after-16:00 L10 unwrap, noon sensitivity, and stop rule if support crosses
   the boundary incompatibly.
10. **Placement/sample strategy:** approve near-eye all-available as main,
    paired/common chest as complementary, paired-sample near-eye comparator,
    and whether all-available chest is shown contextually.
11. **Intervals:** approve interval methods for mixed-model effects and
    descriptive Spearman rho, including any participant-resampling policy.
12. **Diagnostics, sensitivity set, and failure rules:** approve the menu
    above, practical stability criteria, non-estimability handling, and which
    diagnostic triggers reopen the response/metric gate.
13. **Claim gate:** confirm that both named V0 associations, the “only two”
    statement, the chest interpretation, and the broader behavioural claim
    remain withdrawn until the approved Stage 2 outputs are verified.

**Author gate:** no proposed H05 model, bootstrap, sensitivity analysis,
candidate-result read, notebook implementation, or manuscript revision may
proceed until these decisions are approved and logged by the coordinator.

## Reproducibility assertions and environment

```{r}
#| label: audit-assertions

near_flags <- v0_associations |> filter(placement == "near-eye")
chest_flags <- v0_associations |> filter(placement == "chest")

stopifnot(
  all(input_contract$hash_verified),
  all(score_audit$score_verified),
  all(item_audit$missing_n == 0L),
  nrow(metric_map) == 17L,
  nrow(factor_contract) == 4L,
  nrow(near_flags) == 68L,
  nrow(chest_flags) == 68L,
  sum(near_flags$v0_display_flag) == 2L,
  sum(chest_flags$v0_display_flag) == 1L,
  sum(near_flags$consistent_spearman_vector_bh_flag) == 2L,
  sum(chest_flags$consistent_spearman_vector_bh_flag) == 0L,
  nrow(v0_site_screen) == 4L,
  sum(v0_site_screen$passes_v0_screen) == 2L,
  nrow(v0_site_contrasts) == 2L,
  nrow(v0_paired_support) == 17L,
  all(v0_paired_support$sites == 8L)
)

assertion_summary <- tibble::tribble(
  ~assertion, ~result,
  "Frozen input hashes match", TRUE,
  "184 unique LEBA site-participant rows", TRUE,
  "All four stored scores reproduce from 20 items", TRUE,
  "17 metrics × 4 factors × 2 placements = 136 audited cells", TRUE,
  "Exact pair counts persisted by placement, cell, and site", TRUE,
  "V0 and consistent-method flag counts reproduce", TRUE,
  "No proposed H05 model or bootstrap fitted", TRUE
)

knitr::kable(assertion_summary, caption = "Stage 1 execution assertions.")
```

```{r}
#| label: audit-provenance

package_names <- c("dplyr", "tidyr", "purrr", "tibble", "emmeans", "digest", "knitr")
environment_table <- tibble(
  component = c("R", "Quarto", package_names),
  version = c(
    R.version.string,
    Sys.getenv("QUARTO_VERSION", unset = "recorded by render log"),
    vapply(package_names, function(x) as.character(packageVersion(x)), character(1))
  )
)

audit_csvs <- list.files(
  audit_dir,
  pattern = "^h05_stage1_.*[.]csv$",
  full.names = TRUE
)
csv_inventory <- tibble(
  artifact = basename(audit_csvs),
  rows = vapply(audit_csvs, function(x) nrow(read.csv(x, check.names = FALSE)), integer(1)),
  sha256 = vapply(
    audit_csvs, digest::digest, character(1), algo = "sha256", file = TRUE
  )
) |>
  arrange(artifact)

knitr::kable(environment_table, caption = "Analytical execution environment.")
knitr::kable(csv_inventory, caption = "Scoped H05 Stage 1 machine-readable audit artifacts.")
```

All scientific calculations in this document are performed in R against the
frozen inputs named above. The PDF lookup is a structural, page-cited text
retrieval; it does not calculate a scientific result. The source files,
production notebooks, shared configuration, ledgers, and manuscript remain
unchanged.
