---
title: "H05 Stage 2: implementation, diagnostics, and V0 comparison"
subtitle: "Nature Health analysis audit — author result gate"
date: "2026-08-01"
format:
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
    code-fold: true
    code-summary: "Show audit code"
    df-print: paged
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r}
#| label: setup

locate_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        file.exists(file.path(candidate, "_quarto.yml"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root", call. = FALSE)
    }
    candidate <- parent
  }
}

project_root <- locate_project_root()
source(file.path(project_root, "scripts", "pipeline", "p_value_display.R"))
artifact <- function(...) file.path(project_root, "artifacts", ...)
read_h05 <- function(...) {
  readr::read_csv(artifact(...), show_col_types = FALSE, na = "")
}

approvals <- read_h05(
  "06_model_data", "H05", "H05_author_approvals.csv"
)
input_audit <- read_h05(
  "06_model_data", "H05", "H05_input_audit.csv"
)
site_evidence <- read_h05(
  "06_model_data", "H05", "H05_site_structure_evidence.csv"
)
factor_registry <- read_h05(
  "06_model_data", "H05", "H05_factor_registry.csv"
)
metric_registry <- read_h05(
  "06_model_data", "H05", "H05_metric_registry.csv"
)
run_registry <- read_h05(
  "06_model_data", "H05", "H05_run_registry.csv"
)
samples <- read_h05(
  "06_model_data", "H05", "H05_model_frame_index.csv"
)
master <- read_h05(
  "09_tables", "H05", "H05_model_results_master.csv"
)
family_audit <- read_h05(
  "09_tables", "H05", "H05_family_audit.csv"
)
diagnostics <- read_h05(
  "08_diagnostics", "H05", "H05_model_diagnostics.csv"
)
random_site <- read_h05(
  "09_tables", "H05", "H05_random_site_sensitivity.csv"
)
loo_summary <- read_h05(
  "09_tables", "H05", "H05_leave_one_site_out_summary.csv"
)
paired <- read_h05(
  "09_tables", "H05", "H05_paired_placement_comparison.csv"
)
mpd <- read_h05(
  "09_tables", "H05", "H05_manuscript_prepared_comparison.csv"
)
exact_bout <- read_h05(
  "09_tables", "H05", "H05_exactly_identified_longest_bout_sensitivity.csv"
)
spearman <- read_h05(
  "09_tables", "H05", "H05_descriptive_spearman.csv"
)
v0 <- read_h05(
  "09_tables", "H05", "H05_v0_reproduction.csv"
)
v0_to_new <- read_h05(
  "09_tables", "H05", "H05_v0_to_new_comparison.csv"
)
diagnostic_plot_data <- read_h05(
  "11_source_data", "H05", "H05_primary_diagnostic_plot_data.csv"
)

primary <- master |>
  dplyr::filter(.data$run_id == "main__glasses__all_available")
chest <- master |>
  dplyr::filter(.data$run_id == "main__chest__all_available")
mpd_primary <- master |>
  dplyr::filter(
    .data$run_id ==
      "manuscript_prepared_data__glasses__all_available"
  )

duration_f2 <- primary |>
  dplyr::filter(
    .data$metric_id == "duration_above_1000",
    .data$factor_id == "leba_f2"
  )
dose_f2 <- primary |>
  dplyr::filter(
    .data$metric_id == "dose_time_sensitive_corrected_medi",
    .data$factor_id == "leba_f2"
  )

format_p <- function(value) {
  nh_format_p_value(value)
}

format_p_cell <- function(value, significant) {
  display <- nh_p_value_display(value, significant = significant)
  ifelse(
    display$p_bold,
    paste0("**", display$p_display, "**"),
    display$p_display
  )
}

format_effect <- function(estimate, low, high, type, digits = 3L) {
  ifelse(
    type %in% c("ratio", "odds_ratio"),
    sprintf(
      paste0("%.", digits, "f × (%.", digits, "f to %.", digits, "f)"),
      estimate,
      low,
      high
    ),
    sprintf(
      paste0("%+.", digits, "f (%+.", digits, "f to %+.", digits, "f)"),
      estimate,
      low,
      high
    )
  )
}

display_factor <- function(factor_id, factor_label) {
  paste0(toupper(sub("leba_", "", factor_id)), ": ", factor_label)
}

audit_gt <- function(data, title, subtitle = NULL) {
  gt::gt(data) |>
    gt::tab_header(title = title, subtitle = subtitle) |>
    gt::opt_row_striping() |>
    gt::tab_options(
      table.font.size = gt::px(12),
      data_row.padding = gt::px(4),
      heading.align = "left"
    )
}

primary_result_table <- function(factor_id) {
  primary |>
    dplyr::filter(.data$factor_id == factor_id) |>
    dplyr::mutate(
      `Effect per LEBA SD (95% CI)` = format_effect(
        .data$estimate_practical_per_sd,
        .data$conf_low_practical_per_sd,
        .data$conf_high_practical_per_sd,
        .data$effect_type
      ),
      `Raw p` = format_p_cell(.data$p_raw, FALSE),
      `BH-adjusted p` = format_p_cell(
        .data$p_adjusted,
        .data$p_adjusted <= 0.05
      ),
      Adequacy = dplyr::if_else(
        .data$metric_id == "duration_below_1_sleep_environment",
        "Unfit for inference (author disposition)",
        dplyr::recode(
          .data$model_adequacy,
          acceptable = "Acceptable",
          acceptable_with_specified_limitations = "Acceptable with limitations",
          not_acceptable = "Not acceptable"
        )
      )
    ) |>
    dplyr::select(
      Metric = .data$manuscript_name,
      `Effect per LEBA SD (95% CI)`,
      `Raw p`,
      `BH-adjusted p`,
      Adequacy
    ) |>
    audit_gt(
      title = display_factor(
        factor_id,
        primary$factor_label[match(factor_id, primary$factor_id)]
      ),
      subtitle = paste0(
        "Fixed-site primary model; one 68-test BH family across all four tabs. ",
        "BH-adjusted p is bold only at alpha = 0.050; Raw p has no bolding rule"
      )
    ) |>
    gt::fmt_markdown(columns = c(`Raw p`, `BH-adjusted p`))
}

stopifnot(
  nrow(approvals) == 13L,
  all(approvals$approved),
  nrow(primary) == 68L,
  nrow(chest) == 68L,
  nrow(mpd_primary) == 68L,
  all(family_audit$registry_rows == 68L),
  all(family_audit$vector_bh_verified),
  sum(primary$p_adjusted <= 0.05) == 0L,
  sum(primary$model_adequacy == "not_acceptable") == 0L
)
```

## Result at the Stage 2 gate

The signed hypothesis is:

> H5: LEBA questionnaire factors correlate with selected personal light exposure metrics.

The approved primary analysis provides **no association that survives the
complete 68-test Benjamini--Hochberg family**. All 68 fixed-site primary
models are estimable. The archived diagnostic classification contains 19
acceptable and 49 acceptable-with-limitations cells. Following author review,
the four sleep-environment cells are more conservatively designated **unfit
for inference**, leaving 45 retained cells with specified limitations.
The result does not support the submitted statement that exactly two H05
associations are significant.

The unfit classification is specific to the H05 sleep-environment response
and model structure. It does not determine whether the underlying metric is
suitable for another hypothesis that uses a different response variable or
model structure.

The two numerically strongest F2 results remain positive before multiplicity
adjustment:

- Time above 1,000 lx melEDI: ratio per participant SD of LEBA F2 =
  `r sprintf("%.3f", duration_f2$estimate_practical_per_sd)`
  (95% CI `r sprintf("%.3f", duration_f2$conf_low_practical_per_sd)` to
  `r sprintf("%.3f", duration_f2$conf_high_practical_per_sd)`), raw
  Raw p = `r format_p(duration_f2$p_raw)`, BH-adjusted
  *p* = `r format_p(duration_f2$p_adjusted)`.
- Corrected melEDI dose: ratio per participant SD of LEBA F2 =
  `r sprintf("%.3f", dose_f2$estimate_practical_per_sd)`
  (95% CI `r sprintf("%.3f", dose_f2$conf_low_practical_per_sd)` to
  `r sprintf("%.3f", dose_f2$conf_high_practical_per_sd)`), raw
  Raw p = `r format_p(dose_f2$p_raw)`, BH-adjusted
  *p* = `r format_p(dose_f2$p_adjusted)`.

These are observational estimates, not multiplicity-retained findings. The
duration model is acceptable. The dose model is acceptable with an ordinary
Gaussian residual warning. Both directions are stable in the registered
random-site, leave-one-site-out, paired-placement, and
  gap-timing-unaware dataset.

## Author decisions implemented

```{r}
#| label: approval-table

approvals |>
  dplyr::transmute(
    Decision = .data$decision_order,
    Gate = .data$decision_id,
    Disposition = .data$author_disposition,
    Approved = .data$approved
  ) |>
  audit_gt(
    "Stage 1 decisions implemented in Stage 2",
    "The author approved all 13 items on 2026-08-01"
  )
```

### Fixed site is primary

The author's conditional decision 7 resolves to fixed site as primary. In
the H01 main near-eye all-available evidence, every fixed-site model converged
with a positive-definite Hessian and none was singular. The corresponding
random-site descriptions included non-estimable and singular fits. H05
therefore uses submitted-order, sum-to-zero fixed-site contrasts for primary
inference and preserves the registered random-site formula as a named
sensitivity.

```{r}
#| label: site-evidence-table

site_evidence |>
  dplyr::transmute(
    `Fixed metrics` = .data$fixed_site_metrics,
    `Fixed converged` = .data$fixed_site_converged,
    `Fixed PD Hessian` = .data$fixed_site_positive_definite_hessian,
    `Fixed singular` = .data$fixed_site_singular,
    `Random pass` = .data$random_site_descriptive_pass,
    `Random unstable` = .data$random_site_descriptive_unstable,
    `Random non-estimable` = .data$random_site_non_estimable,
    `Random singular` = .data$random_site_singular
  ) |>
  audit_gt(
    "H01 evidence used to resolve H05 site handling",
    "Audit evidence only; H05 outcomes are fitted independently"
  )
```

### Clock and placement decisions

H05 uses the H01-approved continuous clock ranges and subtracts 24 hours from
darkest-10-hour midpoints strictly later than 16:00. The author explicitly
excluded a noon sensitivity because H01 had already resolved the issue. Main
near-eye all-available data are primary. All-available chest data form a
separate complementary 68-test family. Exact paired/common near-eye and chest
analyses are sensitivities. They are fitted separately; placements are never
treated as independent observations in one pooled model.

## Reproducible implementation

### Frozen inputs

```{r}
#| label: input-table

input_audit |>
  dplyr::transmute(
    Role = .data$input_role,
    Path = .data$path,
    `SHA-256` = .data$observed_sha256,
    Verified = .data$hash_verified
  ) |>
  audit_gt(
    "Frozen inputs and inherited analytical sources",
    "Every observed hash equals the approved hash"
  ) |>
  gt::cols_width(
    Path ~ gt::px(360),
    `SHA-256` ~ gt::px(470)
  )
```

All scientific calculations were run in R 4.6.1 from a clean
`Rscript --vanilla` session with the project R 4.6 library. The H05 driver
verified the two model-data artifacts, both model-data manifests, normalized
LEBA input, V0 workspaces, inherited H01 response/model functions, and the H01
fit result used only for the site-structure decision. The complete execution
command was:

```text
R_LIBS=renv/library/macos/R-4.6/aarch64-apple-darwin23 \
NATHEALTH_PROJECT_ROOT=<project-root> \
H05_STAGE=fit H05_RUN_LOO=true \
Rscript --vanilla scripts/hypotheses/H05/run_h05_stage2.R
```

### LEBA and model-frame contract

All 184 questionnaire records are unique by site and participant. All 20
ordered items and all four stored factor scores reproduce exactly; no item or
factor value is missing. F2 reverses only item 04. F3--F5 are direct sums.
Factor 1 remains excluded because it was not administered for the unfiltered
study spectacles.

For every exact run--metric--factor frame, LEBA is centered using each unique
participant once; repeated days do not weight the mean or SD. Effects are
reported per raw point and per that frame's participant SD. No within-site
standardization is used. Every frame's ordered row hash, centering mean, SD,
participants, participant-days, observations, and sites are persisted.

### Models, tests, and intervals

For 15 participant-day metrics, the primary full model is

$$
g(Y_{dij}) = \alpha + \mathrm{site}_j + \beta\,\mathrm{LEBA}_{ij,c}
  + b_{ij},
$$

with a participant random intercept. IS and IV are participant-level and use
the otherwise identical fixed-site linear model without a participant random
intercept. The response family, transformation, link, and reader-scale effect
follow the current approved H01 17-metric registry. Gaussian likelihood-ratio
comparisons use ML and final Gaussian mixed models use REML. Tweedie models use
a log link and ML.

Each inferential test compares the full model with a reduced model that
removes only LEBA on an identical ordered frame. Effect intervals are
two-sided 95% Wald intervals from the final model. The test (p)-value is the
one-degree-of-freedom full-versus-reduced likelihood-ratio test. Descriptive
Spearman coefficients use a model-compatible participant summary and an
explicit approximate Fisher-(z) 95% interval; they have no (p)-values or
significance screen.

```{r}
#| label: family-audit-table

family_audit |>
  dplyr::transmute(
    Family = .data$family_id,
    Planned = .data$planned_tests,
    Rows = .data$registry_rows,
    Observed = .data$observed_tests,
    `BH-adjusted p <= 0.050` = .data$passes_bh_0_05,
    `R vector check` = .data$vector_bh_verified
  ) |>
  audit_gt(
    "Complete multiplicity families",
    "No family was shortened after results were observed"
  )
```

## Exact primary samples

The four factors are complete, so within a run and metric all four models use
the same outcome rows. Participant-level IS/IV have 141 participants and nine
sites. Daily primary metrics range from 655 to 816 participant-days, with 139
to 141 participants and all nine sites.

```{r}
#| label: primary-sample-table-a

primary_samples <- samples |>
  dplyr::filter(.data$run_id == "main__glasses__all_available") |>
  dplyr::distinct(
    .data$metric_order,
    .data$manuscript_name,
    .data$analysis_unit,
    .data$observations,
    .data$participants,
    .data$participant_days,
    .data$represented_days,
    .data$sites
  ) |>
  dplyr::arrange(.data$metric_order)

primary_samples |>
  dplyr::filter(.data$metric_order <= 9L) |>
  dplyr::transmute(
    Metric = .data$manuscript_name,
    Unit = .data$analysis_unit,
    Observations = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Represented days` = .data$represented_days,
    Sites = .data$sites
  ) |>
  audit_gt("Primary near-eye model samples — metrics 1–9")
```

```{r}
#| label: primary-sample-table-b

primary_samples |>
  dplyr::filter(.data$metric_order >= 10L) |>
  dplyr::transmute(
    Metric = .data$manuscript_name,
    Unit = .data$analysis_unit,
    Observations = .data$observations,
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `Represented days` = .data$represented_days,
    Sites = .data$sites
  ) |>
  audit_gt("Primary near-eye model samples — metrics 10–17")
```

For paired IS/IV, H05 derives the exact common participant keys from the two
all-available placement frames because the upstream H01 paired artifact
conservatively omitted participant-level recomputation. Both placements use
the same 112 participants and eight sites. Daily paired models retain the
upstream exact common participant-date keys.

## Primary results

```{r}
#| label: primary-effect-figure
#| fig-cap: "Primary fixed-site effects per participant SD of each LEBA factor. Cell text is on the reader-facing response scale. Colours show model-scale direction and magnitude and are not directly comparable between outcomes. No cell has a black multiplicity-retained border."

knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_primary_effect_overview.png"
))
```

The smallest primary Raw p is
`r format_p(min(primary$p_raw))`; the smallest BH-adjusted p is
`r format_p(min(primary$p_adjusted))`. Thus, effect intervals that exclude the
null before multiplicity correction must not be described as confirmatory H05
findings. The full primary registry is shown below in four 17-row pieces.

::: {.panel-tabset}

### F2 — time outdoors

```{r}
#| label: primary-f2-table
primary_result_table("leba_f2")
```

### F3 — devices in bed

```{r}
#| label: primary-f3-table
primary_result_table("leba_f3")
```

### F4 — ambient light before bedtime

```{r}
#| label: primary-f4-table
primary_result_table("leba_f4")
```

### F5 — morning/daytime light use

```{r}
#| label: primary-f5-table
primary_result_table("leba_f5")
```

:::

## Diagnostics and model adequacy

```{r}
#| label: adequacy-figure
#| fig-cap: "Primary near-eye model assessment after author review. Green marks 19 acceptable cells, orange marks 45 retained cells with specified limitations, and red marks the four sleep-environment cells that are unfit for inference."

knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_primary_model_adequacy.png"
))
```

All fixed-site primary fits converge, have positive-definite Hessians where
applicable, and are nonsingular. No observed response violates its declared
physical bound, and no clock metric triggers the continuous-range stop rule.
The 49 limitation classifications comprise:

- 44 ordinary Gaussian residual warnings, generally driven by sensitive
  normality or moderate variance/tail diagnostics rather than fit failure;
- one ordinary Tweedie warning for F2 with time above 250 lx melEDI during
  wake (DHARMa uniformity Raw p = `r format_p(0.0064)`); and
- four material Tweedie warnings for time below 1 lx melEDI during sleep,
  one for each factor.

The sleep-environment models reproduce the inherited H01 problem: strong
DHARMa non-uniformity and zero-mass/outlier evidence, plus 183--187 fitted
values above participant-specific sleep support. H01's fixed candidate rule
found no adequate common replacement and retained Tweedie/log with an explicit
limitation. H05 therefore does not silently change the family. These four
models received the archived acceptable-with-limitations label. Following
author review, they are now designated **unfit for inference**: their
estimates and p-values remain archived for family transparency but are
suppressed from the reader-facing result matrices and support no claim.

```{r}
#| label: sleep-diagnostic-table

diagnostics |>
  dplyr::filter(
    .data$run_id == "main__glasses__all_available",
    .data$metric_id == "duration_below_1_sleep_environment"
  ) |>
  dplyr::transmute(
    Factor = display_factor(.data$factor_id, .data$factor_label),
    `DHARMa uniformity Raw p` = nh_format_p_value(
      .data$dharma_uniformity_p
    ),
    `DHARMa zero-mass Raw p` = nh_format_p_value(
      .data$dharma_zero_inflation_p
    ),
    `DHARMa outlier Raw p` = nh_format_p_value(.data$dharma_outlier_p),
    `Observed zero fraction` = .data$observed_zero_fraction,
    `Simulated zero fraction` = .data$simulated_zero_fraction,
    `Predicted above support` = .data$predicted_above_bound_n,
    `Reader disposition` = "Unfit for inference"
  ) |>
  audit_gt(
    "Unfit sleep-environment models",
    paste0(
      "Raw diagnostic p-values have no standalone alpha or bolding rule; ",
      "joint residual and support failures make these models unfit for inference"
    )
  ) |>
  gt::fmt_number(
    columns = c(`Observed zero fraction`, `Simulated zero fraction`),
    decimals = 4
  )
```

### Selected residual patterns

The following screens show the two strongest F2 models and the unfit
sleep-environment model. For the Tweedie model, the numerical DHARMa
tests above are the governing simulation-based diagnostics; the displayed
Pearson-residual plots are supporting pattern checks.

```{r}
#| label: selected-residual-fitted
#| fig-width: 10
#| fig-height: 7

selected_metrics <- c(
  "duration_above_1000",
  "dose_time_sensitive_corrected_medi",
  "duration_below_1_sleep_environment"
)
metric_labels <- primary |>
  dplyr::distinct(.data$metric_id, .data$manuscript_name)
selected_plot <- diagnostic_plot_data |>
  dplyr::filter(
    .data$factor_id == "leba_f2",
    .data$metric_id %in% selected_metrics
  )

selected_plot |>
  dplyr::filter(.data$panel == "residual_fitted") |>
  ggplot2::ggplot(ggplot2::aes(x = .data$x, y = .data$y)) +
  ggplot2::geom_hline(yintercept = 0, colour = "grey60") +
  ggplot2::geom_point(alpha = 0.35, size = 0.8) +
  ggplot2::geom_smooth(se = FALSE, method = "loess", colour = "#0072B2") +
  ggplot2::facet_wrap(~manuscript_name, scales = "free_x", ncol = 1) +
  ggplot2::labs(
    x = "Fitted value",
    y = "Standardized Pearson residual",
    title = "Selected F2 residual-versus-fitted screens"
  ) +
  ggplot2::theme_minimal(base_size = 10)
```

```{r}
#| label: selected-qq
#| fig-width: 10
#| fig-height: 7

selected_plot |>
  dplyr::filter(.data$panel == "normal_qq") |>
  ggplot2::ggplot(ggplot2::aes(x = .data$x, y = .data$y)) +
  ggplot2::geom_abline(slope = 1, intercept = 0, colour = "grey60") +
  ggplot2::geom_point(alpha = 0.35, size = 0.8) +
  ggplot2::facet_wrap(~manuscript_name, scales = "free", ncol = 1) +
  ggplot2::labs(
    x = "Theoretical normal quantile",
    y = "Observed standardized residual quantile",
    title = "Selected F2 residual quantile screens"
  ) +
  ggplot2::theme_minimal(base_size = 10)
```

## V0 reproduction and direct comparison

### Faithful and corrected V0 matrices

Stage 2 independently reconstructs all 136 V0 placement--metric--factor
cells from the two frozen workspaces. The faithful displays preserve V0's
Spearman coefficient paired with a default-Pearson test and cell-wise scalar
`p.adjust(..., n = 68)`. The corrected displays use a consistent two-sided
Spearman test and one vector-wide 68-cell BH adjustment. Labels and signed
colour improve readability; no scientific value is changed in the faithful
version.

::: {.panel-tabset}

### Near-eye faithful

```{r}
#| label: v0-near-faithful
#| fig-cap: "Faithful near-eye reconstruction. Cell annotations show Spearman rho and the Pearson-derived scalar-adjusted p; only values meeting its displayed alpha = 0.050 rule are bold."
#| fig-alt: "A four-by-seventeen near-eye matrix showing Spearman correlations and scalar-adjusted p-values formatted to three decimals. Two cells meet the displayed alpha rule and are bold."
knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_v0_near_eye_faithful.png"
))
```

### Near-eye corrected

```{r}
#| label: v0-near-corrected
#| fig-cap: "Corrected near-eye reconstruction. Cell annotations show Spearman rho and BH-adjusted p from one 68-test family; only BH-adjusted p at or below 0.050 is bold."
#| fig-alt: "A four-by-seventeen near-eye heatmap showing Spearman correlations and BH-adjusted p-values formatted to three decimals. Two cells meet the displayed alpha rule and are bold."
knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_v0_near_eye_corrected.png"
))
```

### Chest faithful

```{r}
#| label: v0-chest-faithful
#| fig-cap: "Faithful chest reconstruction. Cell annotations show Spearman rho and the Pearson-derived scalar-adjusted p; only values meeting its displayed alpha = 0.050 rule are bold."
#| fig-alt: "A four-by-seventeen chest matrix showing Spearman correlations and scalar-adjusted p-values formatted to three decimals. One cell meets the displayed alpha rule and is bold."
knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_v0_chest_faithful.png"
))
```

### Chest corrected

```{r}
#| label: v0-chest-corrected
#| fig-cap: "Corrected chest reconstruction. Cell annotations show Spearman rho and BH-adjusted p from one 68-test family; only BH-adjusted p at or below 0.050 is bold."
#| fig-alt: "A four-by-seventeen chest heatmap showing Spearman correlations and BH-adjusted p-values formatted to three decimals. No cell meets the displayed alpha rule."
knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_v0_chest_corrected.png"
))
```

:::

The faithful V0 near-eye matrix flags F2 with dose and time above 1,000 lx.
The consistent V0 calculation also retains those two, with BH-adjusted p =
`r format_p(0.03765590)` and `r format_p(0.03252746)`, respectively. The
faithful chest display flags F2 with dose; its consistent Spearman/vector-BH
value is BH-adjusted p = `r format_p(0.204)`, so no chest
cell remains. This reconciliation does not validate V0 because V0 omits the
registered site-adjusted outcome model and uses superseded metric inputs.

### The two V0 highlights under the approved analysis

```{r}
#| label: v0-highlight-comparison

v0_to_new |>
  dplyr::filter(
    .data$factor_id == "leba_f2",
    .data$v0_name %in% c("dose", "duration_above_1000")
  ) |>
  dplyr::mutate(
    corrected_bh_display = format_p_cell(
      .data$consistent_spearman_vector_bh,
      .data$consistent_spearman_vector_bh <= 0.05
    ),
    `V0 rho (BH-adjusted p)` = paste0(
      sprintf("%.3f", .data$spearman_rho),
      " (BH-adjusted p = ",
      .data$corrected_bh_display,
      ")"
    ),
    `New descriptive rho (95% CI)` = sprintf(
      "%.3f (%.3f to %.3f)",
      .data$new_descriptive_rho,
      .data$new_descriptive_rho_low,
      .data$new_descriptive_rho_high
    ),
    `Primary fixed-site effect per SD (95% CI)` = format_effect(
      .data$fixed_site_effect_per_sd,
      .data$fixed_site_effect_low,
      .data$fixed_site_effect_high,
      .data$effect_type
    ),
    `Primary Raw p` = format_p(.data$fixed_site_p_raw),
    `Primary BH-adjusted p` = format_p_cell(
      .data$fixed_site_p_adjusted,
      .data$fixed_site_p_adjusted <= 0.05
    )
  ) |>
  dplyr::select(
    Metric = .data$manuscript_name,
    `V0 rho (BH-adjusted p)`,
    `New descriptive rho (95% CI)`,
    `Primary fixed-site effect per SD (95% CI)`,
    `Primary Raw p`,
    `Primary BH-adjusted p`,
    Adequacy = .data$model_adequacy
  ) |>
  audit_gt(
    "V0 highlights versus the approved H05 implementation",
    paste0(
      "BH-adjusted p is bold only at alpha = 0.050; Primary Raw p and ",
      "descriptive rho have no separate significance styling"
    )
  ) |>
  gt::fmt_markdown(
    columns = c(`V0 rho (BH-adjusted p)`, `Primary BH-adjusted p`)
  )
```

The repaired participant-summary coefficients are close to V0 for these two
cells, but the approved estimand is the site-adjusted participant-day model.
Neither primary model survives the declared 68-test family. The submitted
phrase “duration above 250 lx” remains incorrect for the highlighted
1,000-lx outcome, and squared Spearman rho must not be described as variance
explained.

## Sensitivity analyses

### Registered random-site model

H05 fits the registered random-site form for all 136 main all-available cells
(68 per placement) without a second significance screen. Fifteen are
descriptively unstable: eight chest cells and seven near-eye cells, chiefly
zero or near-zero site variance. The fixed-site primary fits remain estimable.
The two leading F2 estimates are stable and slightly larger under random site.

```{r}
#| label: random-site-named-table

random_site |>
  dplyr::filter(
    .data$placement == "glasses",
    .data$factor_id == "leba_f2",
    .data$metric_id %in% c(
      "duration_above_1000",
      "dose_time_sensitive_corrected_medi"
    )
  ) |>
  dplyr::mutate(
    `Random-site effect per SD (95% CI)` = format_effect(
      .data$estimate_practical_per_sd,
      .data$conf_low_practical_per_sd,
      .data$conf_high_practical_per_sd,
      .data$effect_type
    )
  ) |>
  dplyr::select(
    Metric = .data$manuscript_name,
    `Site random-effect SD` = .data$site_random_effect_sd,
    `Random-site effect per SD (95% CI)`,
    Status = .data$random_site_status,
    Stability = .data$stability_class
  ) |>
  audit_gt(
    "Registered random-site sensitivity for the two leading F2 estimates",
    "No sensitivity p-values or significance screen"
  ) |>
  gt::fmt_number(columns = `Site random-effect SD`, decimals = 3)
```

### Leave one site out

All 612 primary leave-one-site-out refits pass fit checks. Across all 68
associations, 21 meet the conservative stable classification, 27 preserve
direction but change magnitude by more than 50%, and 20 change sign in at
least one omitted-site refit. Most sign changes concern weak, null-compatible
effects. Both leading F2 results preserve direction in all nine omissions;
their maximum relative changes are 22.8% and 22.9%.

```{r}
#| label: loo-named-table

loo_summary |>
  dplyr::filter(
    .data$factor_id == "leba_f2",
    .data$metric_id %in% c(
      "duration_above_1000",
      "dose_time_sensitive_corrected_medi"
    )
  ) |>
  dplyr::transmute(
    Metric = .data$manuscript_name,
    `Successful refits` = .data$successful_refits,
    `Sign-reversal sites` = .data$sign_reversal_sites,
    `Maximum relative change` = .data$maximum_relative_absolute_change,
    Stability = .data$stability_class
  ) |>
  audit_gt("Leave-one-site-out stability of the two leading F2 estimates") |>
  gt::fmt_percent(columns = `Maximum relative change`, decimals = 1)
```

### Paired/common placement analysis

Fifty-nine of 68 paired/common cells preserve direction and have overlapping
near-eye and chest component intervals; nine change direction. The exact
matched samples contain 110--112 participants and 505--643 participant-days
across eight sites for participant-day metrics; IS and IV use the same 112
participant rows at eight sites. Both
leading F2 estimates are concordant between placements. A valid interval for
the *difference* between correlated placement estimates requires a
participant-cluster resampling analysis; it is deliberately deferred rather
than approximated as independent.

```{r}
#| label: paired-figure
#| fig-cap: "Exact common-sample model-scale effects with near eye on x and chest on y. The dashed 45-degree identity line describes concordance, grey lines mark the null, and equal x/y geometry is used. Samples contain 110–112 participants, 505–643 matched participant-days, and eight sites; IS and IV use 112 participant rows. Closeness to identity is not evidence of equivalence."
#| fig-alt: "Four equal-axis scatterplot panels compare 17 matched near-eye and chest estimands for each LEBA factor. Near eye is horizontal, chest is vertical, a dashed identity line and horizontal and vertical null lines are visible, and nine points cross an axis. The exact matched samples contain 110 to 112 participants, 505 to 643 participant-days, and eight sites; IS and IV use 112 participant rows."

knitr::include_graphics(artifact(
  "10_figures", "H05", "H05_paired_placement_effects.png"
))
```

Exact plotted values, component confidence intervals, identifiers, and
metric-specific matched samples are retained in the
[paired figure source data](../../../artifacts/11_source_data/H05/H05_paired_effect_comparison_data.csv).

```{r}
#| label: paired-named-table

paired |>
  dplyr::filter(
    .data$factor_id == "leba_f2",
    .data$metric_id %in% c(
      "duration_above_1000",
      "dose_time_sensitive_corrected_medi"
    )
  ) |>
  dplyr::mutate(
    `Near-eye effect per SD (95% CI)` = format_effect(
      .data$estimate_practical_per_sd__glasses,
      .data$conf_low_practical_per_sd__glasses,
      .data$conf_high_practical_per_sd__glasses,
      .data$effect_type__glasses
    ),
    `Chest effect per SD (95% CI)` = format_effect(
      .data$estimate_practical_per_sd__chest,
      .data$conf_low_practical_per_sd__chest,
      .data$conf_high_practical_per_sd__chest,
      .data$effect_type__chest
    )
  ) |>
  dplyr::select(
    Metric = .data$manuscript_name,
    `Near-eye effect per SD (95% CI)`,
    `Chest effect per SD (95% CI)`,
    `Same direction` = .data$sign_concordant,
    Stability = .data$stability_class
  ) |>
  audit_gt(
    "Paired/common placement estimates for the leading F2 cells",
    "Component intervals are valid; the correlated difference interval is deferred"
  )
```

### All-available complementary chest family

The separate all-available chest family also has no BH-retained association.
Its smallest BH-adjusted p is `r format_p(min(chest$p_adjusted))`. F2 with
time above 1,000 lx and corrected dose yields reader-scale ratios per SD of
`r sprintf("%.3f", chest$estimate_practical_per_sd[chest$metric_id == "duration_above_1000" & chest$factor_id == "leba_f2"])`
and
`r sprintf("%.3f", chest$estimate_practical_per_sd[chest$metric_id == "dose_time_sensitive_corrected_medi" & chest$factor_id == "leba_f2"])`,
respectively, but neither is multiplicity-retained.

### Gap-timing-unaware dataset

The identical implementation applied to the gap-timing-unaware dataset has no
BH-retained result. This dataset still passed the general 50%-per-hour and
80%-per-day coverage rules, but the timing of remaining missing observations
was not used for additional metric-specific adjustment. Its smallest
BH-adjusted p is
`r format_p(min(mpd_primary$p_adjusted))`. Sixty-four of 68
primary-versus-alternative-preparation near-eye estimates preserve direction
and have overlapping component
intervals; four reverse direction, all among weak effects. The two leading F2
estimates are concordant and have overlapping intervals. This shows that their
positive directions are not created by the primary preparation rules, but the
complete multiplicity result remains null in both scenarios.

### Exactly identified longest-bout sensitivity

The mandatory exact-identifiability sensitivity retains 500 participant-days,
132 participants, and all nine sites. All four factor directions agree with
the all-available lower-bound analysis. For F2, the reader-scale ratio is
attenuated from
`r sprintf("%.3f", exact_bout$all_available_estimate_practical_per_sd[exact_bout$factor_id == "leba_f2"])`
to
`r sprintf("%.3f", exact_bout$estimate_practical_per_sd[exact_bout$factor_id == "leba_f2"])`
(95% CI
`r sprintf("%.3f", exact_bout$conf_low_practical_per_sd[exact_bout$factor_id == "leba_f2"])`
to
`r sprintf("%.3f", exact_bout$conf_high_practical_per_sd[exact_bout$factor_id == "leba_f2"])`).
This reinforces the decision not to promote the unadjusted all-available
longest-bout association.

## Claim disposition

| V0 claim or artifact | Stage 2 evidence | Proposed disposition |
|---|---|---|
| “Only two of 68 comparisons were significant” | Zero of 68 approved primary tests has BH-adjusted p at or below 0.050 | Withdraw and replace with a complete-family null result |
| Positive F2--dose association | Positive estimate and raw interval, but BH-adjusted p = 0.124; ordinary residual warning | Report only as a non-multiplicity-retained estimate if useful |
| Positive F2--time-above-1,000-lx association | Positive estimate and raw interval, but BH-adjusted p = 0.124; diagnostics and major stability checks pass | Report only as a non-multiplicity-retained estimate if useful |
| “Duration above 250 lx” for the highlighted threshold | The actual metric is time above 1,000 lx melEDI | Correct the metric name; do not substitute the distinct waking 250-lx metric |
| ρ² as explained variance | Rank correlation does not provide that interpretation | Remove |
| Site can be omitted | H01 and H05 random-site instability support fixed-site adjustment | Remove; retain site in the model |
| Chest duplicates the near-eye interpretation | Neither chest implementation has a retained result | Replace with placement-specific evidence |
| Habitual behaviour is broadly less informative than immediate context | H05 alone cannot rank causal or predictive importance | Do not make this claim from H05 |

The strongest defensible H05 conclusion is: **after fixed-site adjustment and
one complete 68-test BH correction, this dataset provides no
multiplicity-retained evidence that any of the four LEBA factors is associated
with the 17 selected personal light-exposure metrics.** Positive F2 estimates
for time above 1,000 lx and corrected melEDI dose are stable but exploratory
after multiplicity correction.

## Deferred analyses and production-computation decision

```{r}
#| label: deferred-analysis-table

tibble::tribble(
  ~Item, ~Status, ~Reason, ~Impact_on_current_gate,
  "Paired placement-difference cluster interval",
  "Deferred",
  paste0(
    "Requires participant-cluster bootstrap or an additional approved joint ",
    "model; independent-component approximation would be invalid"
  ),
  "None on the primary 68-test result",
  "Coverage Rules B/C and relaxed native-epoch scenarios",
  "Deferred",
  "Alternative prepared inputs are not present in the frozen H05 model-data package",
  "Could refine robustness, but cannot create an undeclared primary screen",
  "All-zero-inclusive and MDER/state-support threshold batteries",
  "Deferred",
  "Require upstream scenario production and are computationally broader",
  "No primary association is multiplicity-retained",
  "Alternative response-family fits",
  "Not repeated",
  paste0(
    "H01's fixed candidate assessment already retained the current sleep and ",
    "L10 families with limitations; a new family requires a major author gate"
  ),
  "Sleep results remain explicitly limited and support no claim",
  "L10 noon conversion",
  "Not run",
  "Explicitly excluded by the author after H01 resolution",
  "None",
  "Production bootstrap",
  "Not proposed",
  paste0(
    "Primary effects have declared model-based intervals; no computation-heavy ",
    "interval is required for the confirmatory result"
  ),
  "No production-computation gate is required; mandatory Stage 3 follows"
) |>
  audit_gt(
    "Deferred or unnecessary Stage 2 computations",
    "Production computation is a conditional gate between Stages 2 and 3, not Stage 3 itself"
  ) |>
  gt::cols_width(
    Item ~ gt::px(240),
    Status ~ gt::px(100),
    Reason ~ gt::px(430),
    Impact_on_current_gate ~ gt::px(320)
  )
```

No bootstrap pilot was run because the approved primary estimands have
model-based intervals and the confirmatory conclusion does not depend on a
paired placement-difference interval. If the author requires that difference
interval, a separate 50- or 100-successful-refit participant-cluster pilot,
runtime estimate, and approval gate are required before production.

## Stage 2 approval and transition to Stage 3

The author approved items 1--5 and 7 on 2026-08-01:

1. the zero-of-68 primary result and withdrawal of the two V0 significance
   claims;
2. the fixed-site effect and V0/corrected-matrix displays;
3. the original adequacy review, subsequently tightened at the Stage 3 author
   gate by designating the four sleep models unfit for inference;
4. the interpretation that the two leading F2 estimates are stable but not
   multiplicity-retained;
5. the complementary chest, paired/common, random-site, leave-one-site-out,
   gap-timing-unaware dataset, and exact-bout sensitivity
   interpretations;
7. the claim-disposition table and strongest defensible conclusion.

The former item 6 is superseded. The listed expensive analyses remain
deferred, and no production bootstrap or simulation is required for the
approved H05 inference. Production computation is an unnumbered conditional
gate between Stages 2 and 3; it is not Stage 3. H05 therefore proceeds to the
mandatory standalone reader-facing Stage 3 report and will stop again for
author approval before the scientific preparation/provenance description.

This transition is recorded centrally as `H05-001` in
`audit/decisions/h05_stage2_gate_and_stage3_transition.md`. The wording repair
does not change a model, estimate, interval, diagnostic, sensitivity result,
or claim.

## Reproducibility checks

The independent clean-session test verifies 544 fixed-site cells, three
complete 68-test families, identical full/reduced frames, 204 stored
inferential model bundles, 612 successful primary leave-one-site-out refits,
four exactly-identified longest-bout fits, exact paired IS/IV support, all V0
flag counts and coefficients, and all required figures.

```text
R_LIBS=renv/library/macos/R-4.6/aarch64-apple-darwin23 \
NATHEALTH_PROJECT_ROOT=<project-root> \
Rscript --vanilla tests/hypotheses/H05/test_h05_stage2.R
```

All H05 Stage 2 artifacts are inventoried with SHA-256 hashes under
`artifacts/12_manifests/H05/`. No model or sensitivity was rerun for the
workflow-numbering repair. Shared decisions and ledgers remain
coordinator-owned.
