---
title: "H05 analysis preparation and provenance"
subtitle: "From verified light-exposure metrics and LEBA scores to reported associations"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show the R code used for this display"
    code-tools: true
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  error: false
fig-cap-location: bottom
tbl-cap-location: top
---

```{r}
#| label: setup-h05-preparation
#| include: false

suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(gt)
  library(knitr)
  library(readr)
  library(scales)
  library(tibble)
  library(tidyr)
})

root_candidates <- unique(c(
  normalizePath(getwd(), winslash = "/", mustWork = TRUE),
  normalizePath(
    file.path(getwd(), "../../.."),
    winslash = "/",
    mustWork = FALSE
  )
))
is_project_root <- vapply(
  root_candidates,
  function(path) {
    file.exists(file.path(path, "renv.lock")) &&
      file.exists(file.path(
        path,
        "scripts/hypotheses/H05/h05_contract.R"
      ))
  },
  logical(1)
)
stopifnot(any(is_project_root))
root <- root_candidates[which(is_project_root)[1L]]
knitr::opts_knit$set(root.dir = root)

source(file.path(root, "scripts/pipeline/paths_io.R"))

if (!identical(as.character(getRversion()), "4.6.1")) {
  stop("H05 preparation requires R 4.6.1", call. = FALSE)
}

read_h05_csv <- function(relative_path) {
  readr::read_csv(file.path(root, relative_path), show_col_types = FALSE)
}

prep_gt <- function(table, font_size = 13) {
  table |>
    gt::opt_row_striping() |>
    gt::sub_missing(missing_text = "—") |>
    gt::tab_options(
      table.width = gt::pct(100),
      table.font.size = gt::px(font_size),
      data_row.padding = gt::px(4),
      column_labels.font.weight = "600",
      source_notes.font.size = gt::px(11)
    )
}

format_count <- function(value) {
  format(value, big.mark = ",", scientific = FALSE, trim = TRUE)
}

placement_labels <- c(
  glasses = "Near eye",
  chest = "Chest"
)
placement_colors <- c(
  "Near eye" = "#0072B2",
  "Chest" = "#D55E00"
)

input_audit <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_input_audit.csv"
)
factor_registry <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_factor_registry.csv"
)
leba_audit <- read_h05_csv(
  "artifacts/08_diagnostics/H05/H05_leba_score_audit.csv"
)
metric_registry <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_metric_registry.csv"
)
formula_registry <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_formula_registry.csv"
)
run_registry <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_run_registry.csv"
)
frame_index <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_model_frame_index.csv"
)
site_structure <- read_h05_csv(
  "artifacts/06_model_data/H05/H05_site_structure_evidence.csv"
)
diagnostics <- read_h05_csv(
  "artifacts/08_diagnostics/H05/H05_model_diagnostics.csv"
)
family_audit <- read_h05_csv(
  "artifacts/09_tables/H05/H05_family_audit.csv"
)
random_site <- read_h05_csv(
  "artifacts/09_tables/H05/H05_random_site_sensitivity.csv"
)
leave_one_site_out <- read_h05_csv(
  "artifacts/09_tables/H05/H05_leave_one_site_out_summary.csv"
)
paired_placement <- read_h05_csv(
  "artifacts/09_tables/H05/H05_paired_placement_comparison.csv"
)
gap_timing_comparison <- read_h05_csv(
  "artifacts/09_tables/H05/H05_manuscript_prepared_comparison.csv"
)
exact_period <- read_h05_csv(
  paste0(
    "artifacts/09_tables/H05/",
    "H05_exactly_identified_longest_bout_sensitivity.csv"
  )
)
sample_support <- read_h05_csv(
  "artifacts/11_source_data/H05/H05_preparation_sample_support.csv"
)
leba_distribution <- read_h05_csv(
  paste0(
    "artifacts/11_source_data/H05/",
    "H05_preparation_leba_score_distribution.csv"
  )
)
site_support <- read_h05_csv(
  "artifacts/11_source_data/H05/H05_preparation_site_support.csv"
)
site_support_summary <- read_h05_csv(
  paste0(
    "artifacts/11_source_data/H05/",
    "H05_preparation_site_support_summary.csv"
  )
)
site_registry <- read_h05_csv("config/site_display_registry.csv") |>
  arrange(.data$display_order)

model_frame_archive <- readRDS(file.path(
  root,
  "artifacts/06_model_data/H05/H05_model_frames.rds"
))
model_frames <- model_frame_archive$model_frames

primary_id <- "main__glasses__all_available"
chest_id <- "main__chest__all_available"
paired_near_id <- "main__glasses__paired_common_sample"
paired_chest_id <- "main__chest__paired_common_sample"
gap_near_id <- "manuscript_prepared_data__glasses__all_available"
sleep_metric_id <- "duration_below_1_sleep_environment"
```

::: {.callout-note title="About this analysis record"}
This page explains how verified light-exposure metrics and questionnaire scores
were turned into the H05 results. The companion
[H05 results report](../../../notebooks/hypotheses/H05.html) presents the
scientific findings.

Rendering this page recalculates only file-identity checks, schema and key
checks, and lightweight descriptive summaries from the frozen H05 model
frames. It reads the fitted-model archive, estimates, intervals, diagnostics,
influence screens, and sensitivity outputs from their recorded files. It does
not fit or refit a model, calculate predictions, re-estimate residual
diagnostics, rerun leave-one-site-out analyses, bootstrap, or simulate data.
:::

## Purpose

H05 asks whether four Light Exposure Behaviour Assessment (LEBA) questionnaire
factors are associated with 17 personal light-exposure metrics. The primary
analysis uses near-eye measurements; all-available chest measurements provide
complementary evidence. Participant-level metrics and participant-day metrics
use different model structures, but they share the same fixed-site adjustment,
questionnaire-score scaling, and multiplicity family.

This page identifies the data and code that produced those results, explains
why each analytical operation was necessary, and shows the stored evidence
used to check the analysis.

## Analysis map

```{mermaid}
flowchart TD
  A["Verified metric datasets<br/>near eye and chest"] --> C["Identity, schema, key,<br/>coverage and value checks"]
  B["Verified LEBA factor scores<br/>F2 to F5"] --> C
  C --> D["Eight run definitions<br/>placement × sample × preparation"]
  D --> E["Seventeen metric-specific frames<br/>participant or participant-day"]
  E --> F["Four centered LEBA predictors<br/>per exact model frame"]
  F --> G["Fixed-site full and reduced models<br/>68 associations per family"]
  G --> H["Effects, 95% intervals,<br/>and complete-family BH adjustment"]
  G --> I["Residual, response-support,<br/>temporal and influence checks"]
  G --> J["Placement, site-structure,<br/>site-omission and data sensitivities"]
  H --> K["Tables, figures and paired source data"]
  I --> K
  J --> K
  K --> L["H05 results report"]
  M["Automated structural and identity checks"] -. verify .-> C
  M -. verify .-> E
  M -. verify .-> K
```

The arrows describe dependencies. They do not mean that this page repeats the
model computation.

## What is calculated when this page opens

```{r}
#| label: tbl-h05-prep-boundary
#| tbl-cap: "Calculations performed when this page is rendered."

tibble::tribble(
  ~Category, ~Calculated_on_render, ~Explanation,
  "File identity and schema", "Yes", "Checks recorded SHA-256 identities, required columns, unique keys, and archive dimensions.",
  "Descriptive support", "Yes", "Summarises stored questionnaire-score counts, metric-specific samples, and site contributions.",
  "Model construction and estimation", "No", "Reads the formula registry, model manifest, fitted-model archive, and stored estimates.",
  "Diagnostics and sensitivities", "No", "Reads the recorded diagnostic and sensitivity outputs; no diagnostic or sensitivity is repeated.",
  "Reader figures", "Display only", "Draws three descriptive figures from the linked, exact source-data CSV files."
) |>
  gt::gt() |>
  gt::cols_width(
    Category ~ gt::pct(24),
    Calculated_on_render ~ gt::pct(18),
    Explanation ~ gt::pct(58)
  ) |>
  prep_gt()
```

## 1. Verified inputs

### File identity

The analysis begins from named files whose SHA-256 identities were pinned
before model construction. The table below is limited to inputs that determine
the current H05 results or the site-structure decision.

```{r}
#| label: tbl-h05-prep-input-identities
#| tbl-cap: "Verified inputs used by the H05 result-producing analysis."

result_input_roles <- c(
  "main",
  "main_manifest",
  "manuscript_prepared_data",
  "manuscript_prepared_data_manifest",
  "leba",
  "h01_contract_source",
  "h01_modeling_source",
  "h01_fit_results_site_evidence"
)
input_labels <- c(
  main = "Primary metric dataset",
  main_manifest = "Primary metric-data manifest",
  manuscript_prepared_data = "Gap-timing-unaware metric dataset",
  manuscript_prepared_data_manifest = "Gap-timing-unaware data manifest",
  leba = "Normalised LEBA questionnaire scores",
  h01_contract_source = "Shared metric contract",
  h01_modeling_source = "Shared metric and diagnostic functions",
  h01_fit_results_site_evidence = "Stored site-structure evidence"
)

verified_inputs <- input_audit |>
  filter(.data$input_role %in% result_input_roles) |>
  mutate(
    current_sha256 = vapply(
      file.path(root, .data$path),
      artifact_sha256,
      character(1)
    ),
    Input = unname(input_labels[.data$input_role]),
    Path = .data$path,
    `SHA-256 prefix` = substr(.data$current_sha256, 1L, 16L),
    Verified = if_else(
      .data$current_sha256 == .data$expected_sha256,
      "PASS",
      "FAIL"
    )
  )
stopifnot(nrow(verified_inputs) == 8L, all(verified_inputs$Verified == "PASS"))

verified_inputs |>
  select("Input", "Path", "SHA-256 prefix", "Verified") |>
  gt::gt() |>
  gt::cols_width(
    Input ~ gt::pct(25),
    Path ~ gt::pct(53),
    `SHA-256 prefix` ~ gt::pct(15),
    Verified ~ gt::pct(7)
  ) |>
  gt::tab_source_note(
    "Complete 64-character identities are retained in artifacts/06_model_data/H05/H05_input_audit.csv."
  ) |>
  prep_gt(12)
```

### Schema, key, content, and value checks

The LEBA table is keyed by site and participant identifier. Each metric frame
is keyed either by participant or by participant plus local date. The model
index adds run, metric, and factor identifiers, so every one of its 544 rows
refers to one unique analysis cell.

```{r}
#| label: tbl-h05-prep-integrity-checks
#| tbl-cap: "Bounded integrity checks performed from the stored H05 inputs and model frames."

frame_duplicate_keys <- sum(vapply(
  model_frames,
  function(frame) anyDuplicated(frame$.model_row_id) > 0L,
  logical(1)
))
required_frame_columns <- c(
  ".model_row_id", "site", "Id", "participant_key", "local_date",
  "participant_days_contributing", "value", "leba_f2", "leba_f3",
  "leba_f4", "leba_f5"
)
frames_with_complete_schema <- sum(vapply(
  model_frames,
  function(frame) all(required_frame_columns %in% names(frame)),
  logical(1)
))
frame_order_verified <- all(vapply(
  model_frames,
  function(frame) {
    expected_order <- order(
      frame$site,
      frame$participant_key,
      frame$local_date,
      na.last = TRUE
    )
    identical(expected_order, seq_len(nrow(frame)))
  },
  logical(1)
))
index_key_rows <- frame_index |>
  distinct(.data$run_id, .data$metric_id, .data$factor_id) |>
  nrow()

integrity_checks <- tibble::tribble(
  ~Check, ~Expected, ~Observed, ~Status,
  "Current result-input identities", "8 matches", paste0(sum(verified_inputs$Verified == "PASS"), " matches"), if_else(all(verified_inputs$Verified == "PASS"), "PASS", "FAIL"),
  "LEBA participant key", "184 unique site–participant rows", "184 unique rows; 0 duplicates", if_else(all(leba_audit$participants == 184L), "PASS", "FAIL"),
  "LEBA score reconstruction", "0 mismatches; 0 missing scores", paste0(sum(leba_audit$mismatch_n), " mismatches; ", sum(leba_audit$missing_scores), " missing"), if_else(sum(leba_audit$mismatch_n) == 0L && sum(leba_audit$missing_scores) == 0L, "PASS", "FAIL"),
  "Metric registry", "17 unique ordered metrics", paste0(n_distinct(metric_registry$metric_id), " unique metrics"), if_else(n_distinct(metric_registry$metric_id) == 17L, "PASS", "FAIL"),
  "Run registry", "8 unique runs", paste0(n_distinct(run_registry$run_id), " unique runs"), if_else(n_distinct(run_registry$run_id) == 8L, "PASS", "FAIL"),
  "Stored run–metric frames", "136 frames", paste0(length(model_frames), " frames"), if_else(length(model_frames) == 136L, "PASS", "FAIL"),
  "Frame schema and row keys", "136 complete; 0 duplicate keys", paste0(frames_with_complete_schema, " complete; ", frame_duplicate_keys, " duplicate-key frames"), if_else(frames_with_complete_schema == 136L && frame_duplicate_keys == 0L, "PASS", "FAIL"),
  "Frame ordering", "Site, participant, then local date", if_else(frame_order_verified, "Order retained", "Order differs"), if_else(frame_order_verified, "PASS", "FAIL"),
  "Run–metric–factor index", "544 unique cells", paste0(index_key_rows, " unique cells"), if_else(index_key_rows == 544L && nrow(frame_index) == 544L, "PASS", "FAIL")
)
stopifnot(all(integrity_checks$Status == "PASS"))

integrity_checks |>
  gt::gt() |>
  gt::cols_width(
    Check ~ gt::pct(27),
    Expected ~ gt::pct(29),
    Observed ~ gt::pct(35),
    Status ~ gt::pct(9)
  ) |>
  prep_gt(12)
```

## 2. Questionnaire predictors

### Factor construction

The four LEBA factors are complete-score sums. Ordered item responses run from
1 (“Never”) to 5 (“Always”). F2 contains one reverse-coded item; the other
factors use the recorded item direction. A factor score is non-estimable if
any contributing item is missing.

```{r}
#| label: tbl-h05-prep-factor-contract
#| tbl-cap: "LEBA factors and score-construction rules."

factor_registry |>
  transmute(
    Factor = toupper(sub("leba_", "", .data$factor_id)),
    Construct = .data$factor_label,
    Items = .data$last_item - .data$first_item + 1L,
    `Possible range` = paste0(.data$possible_min, "–", .data$possible_max),
    `Score rule` = .data$score_rule,
    `Missing-item rule` = .data$missing_item_rule
  ) |>
  gt::gt(rowname_col = "Factor") |>
  gt::cols_width(
    Construct ~ gt::pct(27),
    Items ~ gt::pct(7),
    `Possible range` ~ gt::pct(10),
    `Score rule` ~ gt::pct(31),
    `Missing-item rule` ~ gt::pct(25)
  ) |>
  prep_gt(12)
```

The recorded reconstruction found no missing score and no mismatch for any of
the 184 participants. Scores were not standardised within site. For each exact
model frame, the participant-level mean was subtracted from the raw factor
score, and the reported effect was expressed both per raw point and per one
participant-level standard deviation.

```{r}
#| label: tbl-h05-prep-leba-audit
#| tbl-cap: "Observed LEBA score support and reconstruction checks."

leba_audit |>
  transmute(
    Factor = toupper(sub("leba_", "", .data$factor_id)),
    Participants = .data$participants,
    `Observed range` = paste0(.data$observed_min, "–", .data$observed_max),
    `Unique scores` = .data$unique_scores,
    `Tied participants` = .data$tied_participants,
    `Missing scores` = .data$missing_scores,
    Mismatches = .data$mismatch_n,
    Verified = if_else(.data$score_verified, "PASS", "FAIL")
  ) |>
  gt::gt(rowname_col = "Factor") |>
  gt::fmt_integer(
    columns = c(
      Participants, `Unique scores`, `Tied participants`,
      `Missing scores`, Mismatches
    )
  ) |>
  prep_gt()
```

```{r}
#| label: fig-h05-prep-leba-distribution
#| fig-cap: "Distribution of the four verified LEBA factor scores. Each panel contains all 184 questionnaire participants."
#| fig-alt: "Four bar charts show participant counts at each observed LEBA score for F2 through F5. Each panel sums to 184 participants; the observed score ranges differ because the factors contain different numbers of items."
#| fig-width: 10
#| fig-height: 6.8

leba_plot_data <- leba_distribution |>
  mutate(
    factor_code = paste0("F", .data$factor_order + 1L),
    panel_label = paste0(.data$factor_code, ": ", .data$factor_label)
  )

ggplot(leba_plot_data, aes(x = .data$score, y = .data$participants)) +
  geom_col(fill = "#0072B2", width = 0.85) +
  facet_wrap(vars(.data$panel_label), scales = "free_x", ncol = 2) +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.06))) +
  labs(x = "LEBA factor score", y = "Participants") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    axis.title = element_text(face = "bold")
  )
```

The exact aggregated counts used by the bars are available as
[figure source data](../../../artifacts/11_source_data/H05/H05_preparation_leba_score_distribution.csv).

## 3. Metric construction and analytical samples

### Coverage, gaps, and metric support

The primary dataset applied the general requirement of at least 50% valid
minutes per hour and at least 80% valid hours per day, together with
metric-specific support rules where the timing of remaining gaps affects what
the metric can represent. Unsupported values remained missing rather than
being interpreted as darkness. Light-level metrics use melEDI in reader-facing
labels, with a 0.1 lx offset where a logarithmic transform requires it.

One predefined sensitivity uses the **gap-timing-unaware dataset**. It still
passed the general 50%-per-hour and 80%-per-day coverage rules, but the timing
of the remaining missing observations was not used for an additional
metric-specific adjustment. For contrast at this first explanation, the
primary dataset could be interpreted as a **time-sensitive primary metric
dataset**. Hereafter it is called simply the primary dataset.

The longest continuous period above 250 lx melEDI is an observed lower bound
when a gap can interrupt the period. Its exact-identifiability sensitivity
therefore retains only participant-days for which the longest **period** is
fully observed. The L10 midpoint retains its continuous range and its
strict-after-16:00 representation; no noon sensitivity was needed.

```{r}
#| label: tbl-h05-prep-metric-contract
#| tbl-cap: "Metric-specific analysis level, response model, transformation, and reported effect scale."

metric_registry |>
  transmute(
    Order = .data$metric_order,
    Metric = .data$manuscript_name,
    `Analysis level` = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    `Response model` = recode(
      .data$response_family,
      gaussian = "Gaussian",
      tweedie_log = "Tweedie, log link"
    ),
    Transformation = recode(
      .data$response_transform,
      identity = "Identity",
      logit = "Logit",
      log10_offset_0.1 = "log10(value + 0.1 lx)"
    ),
    `Reported effect` = recode(
      .data$effect_scale,
      difference = "Difference",
      ratio = "Ratio",
      odds_ratio = "Odds ratio"
    ),
    Unit = .data$display_unit
  ) |>
  gt::gt(rowname_col = "Order") |>
  gt::cols_width(
    Metric ~ gt::pct(31),
    `Analysis level` ~ gt::pct(13),
    `Response model` ~ gt::pct(16),
    Transformation ~ gt::pct(18),
    `Reported effect` ~ gt::pct(13),
    Unit ~ gt::pct(9)
  ) |>
  prep_gt(11)
```

### Model-frame construction

`h05_prepare_metric_rows()` selects one placement, preparation, sample rule,
and metric at a time. It applies metric eligibility, checks the participant or
participant-day key, joins LEBA by site and participant, excludes rows without
a complete factor score, fixes the registered site order, and sorts rows by
site, participant, and local date. Interdaily stability and intradaily
variability contribute one row per participant. The other 15 metrics
contribute one row per eligible participant-day.

The archive contains 136 run-by-metric frames (8 × 17). Joining the four LEBA
factors creates the 544 model-index rows without changing a metric frame's
participants or observations.

```{r}
#| label: tbl-h05-prep-primary-samples
#| tbl-cap: "Exact all-available primary near-eye and complementary chest samples for each metric."

sample_support |>
  mutate(
    Placement = factor(
      .data$placement_label,
      levels = c("Near eye", "Chest")
    ),
    Metric = .data$manuscript_name,
    Level = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    )
  ) |>
  select(
    "Placement", "metric_order", "Metric", "Level", "observations",
    "participants", "participant_days", "represented_days", "sites"
  ) |>
  gt::gt(groupname_col = "Placement", rowname_col = "metric_order") |>
  gt::cols_label(
    observations = "Observations",
    participants = "Participants",
    participant_days = "Participant-days",
    represented_days = "Days represented",
    sites = "Sites"
  ) |>
  gt::fmt_integer(
    columns = c(
      observations, participants, participant_days,
      represented_days, sites
    ),
    use_seps = TRUE
  ) |>
  gt::cols_width(
    Metric ~ gt::pct(33),
    Level ~ gt::pct(13),
    observations ~ gt::pct(11),
    participants ~ gt::pct(11),
    participant_days ~ gt::pct(12),
    represented_days ~ gt::pct(12),
    sites ~ gt::pct(8)
  ) |>
  prep_gt(11)
```

The corresponding [exact table source](../../../artifacts/11_source_data/H05/H05_preparation_sample_support.csv)
retains internal metric and placement identifiers as well as every displayed
count.

```{r}
#| label: fig-h05-prep-sample-support
#| fig-cap: "Metric-specific participant and represented-day support in the primary near-eye and complementary chest analyses. Metric numbers follow the preceding table."
#| fig-alt: "Two line-and-point panels compare near-eye and chest sample support across 17 ordered metrics. The participant panel ranges roughly from 139 to 154 people; represented-day support varies more strongly by metric, with near-eye generally contributing about 655 to 816 days and chest about 687 to 902 days."
#| fig-width: 11
#| fig-height: 6.5

sample_plot_data <- sample_support |>
  select(
    "metric_order", "manuscript_name", "placement_label",
    "participants", "represented_days"
  ) |>
  pivot_longer(
    cols = c("participants", "represented_days"),
    names_to = "quantity",
    values_to = "count"
  ) |>
  mutate(
    quantity = recode(
      .data$quantity,
      participants = "Participants",
      represented_days = "Days represented"
    )
  )

ggplot(
  sample_plot_data,
  aes(
    x = .data$metric_order,
    y = .data$count,
    colour = .data$placement_label,
    group = .data$placement_label
  )
) +
  geom_line(linewidth = 0.75) +
  geom_point(size = 2.3) +
  facet_wrap(vars(.data$quantity), scales = "free_y", ncol = 1) +
  scale_colour_manual(values = placement_colors) +
  scale_x_continuous(breaks = seq_len(17L)) +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(
    x = "Metric order",
    y = NULL,
    colour = "Placement"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    legend.position = "top"
  )
```

The figure uses the same
[paired source data](../../../artifacts/11_source_data/H05/H05_preparation_sample_support.csv)
as the exact sample table.

### Site contributions

The primary near-eye analysis represents all nine submitted-manuscript sites.
The complementary chest data represent eight. Site-specific counts vary with
metric eligibility; the displayed range is therefore taken across the 17
exact model frames, not from a pooled approximation.

```{r}
#| label: tbl-h05-prep-site-support
#| tbl-cap: "Per-site sample ranges across the 17 all-available metric frames."

site_support_summary |>
  transmute(
    Analysis = recode(
      .data$run_label,
      `Primary near-eye dataset` = "Primary near eye",
      `Complementary chest dataset` = "Complementary chest"
    ),
    `Site order` = .data$display_order,
    Site = .data$site_name,
    `Metric frames` = .data$metric_frames,
    `Participants, min–max` = paste0(
      .data$participants_min,
      "–",
      .data$participants_max
    ),
    `Days represented, min–max` = paste0(
      format_count(.data$represented_days_min),
      "–",
      format_count(.data$represented_days_max)
    )
  ) |>
  gt::gt(groupname_col = "Analysis", rowname_col = "Site order") |>
  gt::fmt_integer(columns = `Metric frames`) |>
  gt::cols_width(
    Site ~ gt::pct(30),
    `Metric frames` ~ gt::pct(18),
    `Participants, min–max` ~ gt::pct(23),
    `Days represented, min–max` ~ gt::pct(29)
  ) |>
  gt::tab_source_note(
    "Sites follow the submitted-manuscript order. The full metric-by-site table retains observations, participants, participant-days, and represented days."
  ) |>
  prep_gt(12)
```

```{r}
#| label: fig-h05-prep-site-range
#| fig-cap: "Range of participants contributed by each site across the 17 primary near-eye and complementary chest metric frames. Site order and colours follow the submitted manuscript registry."
#| fig-alt: "Two panels show horizontal participant-count ranges for each site across 17 metrics. The primary near-eye panel contains nine sites and the complementary chest panel eight. Tübingen and Madrid contribute the largest near-eye samples, while San José contributes the smallest."
#| fig-width: 11
#| fig-height: 7

site_levels <- rev(site_registry$display_name)
site_colors <- site_registry$color_hex
names(site_colors) <- site_registry$display_name
site_plot_data <- site_support_summary |>
  mutate(
    site_name = factor(.data$site_name, levels = site_levels),
    run_label = recode(
      .data$run_label,
      `Primary near-eye dataset` = "Primary near eye",
      `Complementary chest dataset` = "Complementary chest"
    )
  )

ggplot(site_plot_data, aes(y = .data$site_name, colour = .data$site_name)) +
  geom_segment(
    aes(
      x = .data$participants_min,
      xend = .data$participants_max,
      yend = .data$site_name
    ),
    linewidth = 2.2,
    lineend = "round"
  ) +
  geom_point(aes(x = .data$participants_min), size = 2.6) +
  geom_point(aes(x = .data$participants_max), size = 2.6) +
  facet_wrap(vars(.data$run_label), ncol = 2) +
  scale_colour_manual(values = site_colors, drop = FALSE) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  labs(x = "Participants across metric-specific frames", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    axis.title.x = element_text(face = "bold"),
    legend.position = "none"
  )
```

The plotted range has
[exact source data](../../../artifacts/11_source_data/H05/H05_preparation_site_support_summary.csv).
The complete
[metric-by-site source](../../../artifacts/11_source_data/H05/H05_preparation_site_support.csv)
contains 1,122 rows across every predefined run.

## 4. Model construction and inference

### Exact formulas supplied to the models

The analysis uses fixed site with sum-to-zero contrasts as the primary site
adjustment. Participant-day models add a participant random intercept because
several days can come from the same participant. Participant-level IS and IV
models contain one row per person and therefore do not add that intercept.

The formula registry below replaces the raw printed R list with a compact
reader-facing table while retaining the exact Wilkinson formulas evaluated by
the model functions.

```{r}
#| label: tbl-h05-prep-exact-formulas
#| tbl-cap: "Exact evaluated Wilkinson formulas supplied to the H05 full, reduced, and random-site sensitivity models."

formula_registry |>
  mutate(
    `Analysis level` = recode(
      .data$analysis_unit,
      participant = "Participant",
      participant_day = "Participant-day"
    ),
    `Model role` = recode(
      .data$formula_id,
      fixed_full = "Fixed-site full",
      fixed_reduced = "Fixed-site reduced",
      random_site = "Random-site sensitivity"
    ),
    Formula = paste0("`", .data$formula, "`")
  ) |>
  select("Analysis level", "Model role", "Formula") |>
  gt::gt(groupname_col = "Analysis level") |>
  gt::fmt_markdown(columns = Formula) |>
  gt::cols_width(
    `Model role` ~ gt::pct(24),
    Formula ~ gt::pct(76)
  ) |>
  prep_gt(12)
```

### Engines and estimation settings

```{r}
#| label: tbl-h05-prep-model-settings
#| tbl-cap: "Model engines and estimation settings used for H05."

tibble::tribble(
  ~Model_component, ~Engine, ~Estimation, ~Reason,
  "Participant-level Gaussian outcomes", "stats::lm", "Maximum likelihood-equivalent linear model", "One row per participant; site is a fixed effect.",
  "Participant-day Gaussian comparisons", "lme4::lmer", "Maximum likelihood", "Full-versus-reduced likelihood-ratio tests require a common likelihood basis.",
  "Participant-day Gaussian final effects", "lme4::lmer", "Restricted maximum likelihood", "Provides the stored final coefficient and 95% Wald interval after the comparison model is fixed.",
  "Participant-day positive outcomes", "glmmTMB::glmmTMB", "Tweedie family, log link, maximum likelihood", "Accommodates non-negative, right-skewed duration responses while retaining a participant random intercept.",
  "Random-site sensitivity", "lme4 or glmmTMB", "Final-model settings", "Replaces fixed site with a site random intercept without changing the response model.",
  "Optimisation", "nloptwrap or glmmTMB control", "Derivatives retained; up to 10,000 iterations/evaluations for Tweedie", "Makes convergence and Hessian evidence available for the stored diagnostic gate."
) |>
  gt::gt() |>
  gt::cols_width(
    Model_component ~ gt::pct(24),
    Engine ~ gt::pct(16),
    Estimation ~ gt::pct(22),
    Reason ~ gt::pct(38)
  ) |>
  prep_gt(11)
```

### Site specification

The fixed-site primary specification differs from the registered random-site
structure. The decision was based on the inherited 17-metric site-structure
evidence: every fixed-site model converged with a positive-definite Hessian and
none was singular, whereas the random-site descriptions contained unstable,
non-estimable, and singular fits. Keeping site fixed also estimates each
site's adjustment directly rather than relying on a variance component from
only nine sites. The random-site structure remains a sensitivity.

```{r}
#| label: tbl-h05-prep-site-structure
#| tbl-cap: "Evidence supporting fixed site as the primary specification."

site_structure |>
  transmute(
    Evidence = "Inherited 17-metric near-eye site-structure assessment",
    `Fixed-site models` = .data$fixed_site_metrics,
    `Fixed converged` = .data$fixed_site_converged,
    `Fixed positive Hessian` = .data$fixed_site_positive_definite_hessian,
    `Fixed singular` = .data$fixed_site_singular,
    `Random descriptive pass` = .data$random_site_descriptive_pass,
    `Random unstable` = .data$random_site_descriptive_unstable,
    `Random non-estimable` = .data$random_site_non_estimable,
    `Random singular` = .data$random_site_singular
  ) |>
  gt::gt() |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  prep_gt(11)
```

Across the H05 random-site sensitivity itself, 121 of 136 fits passed the
descriptive stability checks and 15 were unstable. This is consistent with
using fixed site for primary inference and random site only as a stability
assessment.

### Estimand, uncertainty, and multiplicity

For each metric and factor, the full model adds the centered LEBA score to the
reduced site-adjusted model. The raw inferential p-value comes from that
full-versus-reduced likelihood-ratio comparison. Each declared family contains
all 17 metrics × 4 factors = 68 tests, and Benjamini–Hochberg adjustment is
performed on the complete 68-value vector. Significance is decided at adjusted
*p* ≤ 0.05 before any display formatting.

Coefficients are converted to the metric's practical scale and reported per
one participant-level LEBA standard deviation. Gaussian identity outcomes are
differences, log-transformed light outcomes are ratios, and IS is an odds
ratio. Every reported interval is a two-sided 95% Wald interval from the final
model. No secondary significance screen is attached to descriptive
sensitivities.

```{r}
#| label: tbl-h05-prep-family-audit
#| tbl-cap: "Complete-family multiplicity audit for the three declared H05 inferential families."

family_labels <- c(
  `H05-F1-primary` = "Primary near eye",
  `H05-F2-complementary-chest` = "Complementary chest",
  `H05-F3-manuscript-prepared` = "Gap-timing-unaware near-eye sensitivity"
)

family_audit |>
  transmute(
    Family = unname(family_labels[.data$family_id]),
    `Planned tests` = .data$planned_tests,
    `Observed tests` = .data$observed_tests,
    `Adjusted tests` = .data$estimable_adjusted_tests,
    `Retained at BH 0.05` = .data$passes_bh_0_05,
    `Complete-vector check` = if_else(.data$vector_bh_verified, "PASS", "FAIL")
  ) |>
  gt::gt(rowname_col = "Family") |>
  gt::fmt_integer(columns = where(is.numeric)) |>
  gt::tab_source_note(
    "Zero of 68 primary associations survives Benjamini–Hochberg correction. The same zero-retained result occurs in the complementary and gap-timing-unaware families."
  ) |>
  prep_gt()
```

::: {.callout-note title="Why a less conservative adjustment can accompany fewer retained associations"}
For one unchanged vector of raw p-values, Benjamini–Hochberg cannot retain
fewer associations than Bonferroni at the same 0.05 threshold. The H05 family
above is based on likelihood-ratio p-values from the fixed-site models. It is
therefore not interchangeable with a family built from unadjusted rank
correlations or another raw test. Multiplicity adjustment and the construction
of the raw p-values are separate analytical decisions; the latter can change
which associations enter the smallest ranks before either correction is
applied.
:::

## 5. Diagnostics and influence

### Diagnostic sequence

Every stored final model was checked for convergence, Hessian status,
singularity, residual behaviour, response bounds, metric-specific timing
support, and within-participant serial dependence where applicable. Gaussian
models use transformed-residual and influence checks. Tweedie models add
simulation-based uniformity, dispersion, zero-support, and outlier checks.
Candidate influential participants were ranked by Cook's distance for ordinary
linear models or by maximum absolute Pearson residual for mixed models; this
screen did not silently remove observations.

The compact table below excludes the sleep-environment metric because that
outcome has a separate, stronger disposition immediately afterwards.

```{r}
#| label: tbl-h05-prep-diagnostic-overview
#| tbl-cap: "Stored diagnostic disposition for non-sleep-environment primary and complementary models."

diagnostic_overview <- diagnostics |>
  filter(
    .data$run_id %in% c(primary_id, chest_id),
    .data$metric_id != sleep_metric_id
  ) |>
  mutate(
    Placement = recode(
      .data$run_id,
      !!primary_id := "Primary near eye",
      !!chest_id := "Complementary chest"
    ),
    Disposition = recode(
      .data$model_adequacy,
      acceptable = "Acceptable",
      acceptable_with_specified_limitations = "Acceptable with stated limitations",
      not_acceptable = "Not acceptable"
    )
  ) |>
  count(.data$Placement, .data$Disposition, name = "Factor–metric models")

diagnostic_overview |>
  gt::gt(groupname_col = "Placement", rowname_col = "Disposition") |>
  gt::fmt_integer(columns = `Factor–metric models`) |>
  gt::tab_source_note(
    "Counts cover 16 metrics × 4 factors per placement. A limitation label is not a significance decision."
  ) |>
  prep_gt()
```

::: {.callout-note title="Sleep-environment outcome: unfit for H05 inference"}
The four primary near-eye models for **time below 1 lx melEDI during sleep**
are unfit for H05 inference. The observed response contains exact zeros, while
the fitted Tweedie simulations generated none; simulation-based uniformity and
outlier checks also showed material failure, and fitted values crossed the
admissible response support. Their coefficients, intervals, and p-values are
therefore suppressed from the H05 results. The complementary chest models show
the same material response-support and simulated-residual limitations and are
also not used as inferential evidence.

This disposition is specific to this outcome **as the H05 response under this
model structure**. It does not mean that the sleep-environment metric cannot be
analysed in another hypothesis that uses a different response variable,
estimand, or model structure.
:::

```{r}
#| label: tbl-h05-prep-sleep-disposition
#| tbl-cap: "Response-support evidence behind the unfit H05 sleep-environment disposition."

sleep_samples <- frame_index |>
  filter(
    .data$run_id %in% c(primary_id, chest_id),
    .data$metric_id == sleep_metric_id,
    .data$factor_id == "leba_f2"
  ) |>
  select("run_id", "observations", "participants", "participant_days", "sites")

sleep_disposition <- diagnostics |>
  filter(
    .data$run_id %in% c(primary_id, chest_id),
    .data$metric_id == sleep_metric_id
  ) |>
  group_by(.data$run_id) |>
  summarise(
    `Factor models` = n(),
    `Observed zero fraction` = first(.data$observed_zero_fraction),
    `Simulated zero fraction` = first(.data$simulated_zero_fraction),
    `Uniformity check` = "Material failure",
    `Prediction support` = "Material limitation",
    `H05 inference` = "Unfit",
    .groups = "drop"
  ) |>
  left_join(sleep_samples, by = "run_id", relationship = "one-to-one") |>
  mutate(
    Placement = recode(
      .data$run_id,
      !!primary_id := "Primary near eye",
      !!chest_id := "Complementary chest"
    )
  ) |>
  select(
    "Placement", "Factor models", "observations", "participants",
    "participant_days", "sites", "Observed zero fraction",
    "Simulated zero fraction", "Uniformity check", "Prediction support",
    "H05 inference"
  )

sleep_disposition |>
  gt::gt(rowname_col = "Placement") |>
  gt::cols_label(
    observations = "Observations",
    participants = "Participants",
    participant_days = "Participant-days",
    sites = "Sites"
  ) |>
  gt::fmt_integer(
    columns = c(
      `Factor models`, observations, participants, participant_days, sites
    ),
    use_seps = TRUE
  ) |>
  gt::fmt_percent(
    columns = c(`Observed zero fraction`, `Simulated zero fraction`),
    decimals = 2
  ) |>
  prep_gt(10)
```

## 6. Sensitivity analyses

Sensitivity branches preserve the accepted primary estimand unless their
purpose explicitly changes the sample or site representation. They are used to
describe stability, not to promote a non-retained primary association.

```{r}
#| label: tbl-h05-prep-sensitivity-map
#| tbl-cap: "H05 sensitivity branches, exact stored scope, and interpretive role."

random_counts <- random_site |>
  count(.data$random_site_status) |>
  tibble::deframe()
paired_sample_range <- frame_index |>
  filter(
    .data$run_id %in% c(paired_near_id, paired_chest_id),
    .data$factor_id == "leba_f2"
  ) |>
  summarise(
    participants_min = min(.data$participants),
    participants_max = max(.data$participants),
    days_min = min(.data$represented_days),
    days_max = max(.data$represented_days),
    sites_min = min(.data$sites),
    sites_max = max(.data$sites)
  )
gap_sample_range <- frame_index |>
  filter(
    .data$run_id == gap_near_id,
    .data$factor_id == "leba_f2"
  ) |>
  summarise(
    participants_min = min(.data$participants),
    participants_max = max(.data$participants),
    days_min = min(.data$represented_days),
    days_max = max(.data$represented_days),
    sites = max(.data$sites)
  )

sensitivity_map <- tibble::tribble(
  ~Analysis, ~Exact_stored_scope, ~Purpose, ~Interpretive_rule,
  "Paired/common near eye and chest",
  paste0(
    "68 matched factor–metric estimands; ",
    paired_sample_range$participants_min, "–",
    paired_sample_range$participants_max, " participants, ",
    paired_sample_range$days_min, "–", paired_sample_range$days_max,
    " represented days, ", paired_sample_range$sites_min, " sites"
  ),
  "Separates placement from unequal all-available samples.",
  "Compare only matched estimands; do not infer equivalence and do not fit a new model for the display.",
  "Random site",
  paste0(
    "136 main all-available fits: ",
    random_counts[["DESCRIPTIVE_PASS"]], " passed and ",
    random_counts[["DESCRIPTIVE_UNSTABLE"]], " were unstable"
  ),
  "Checks the registered site-random-intercept specification.",
  "Instability supports retaining fixed site as primary.",
  "Leave one site out",
  paste0(
    sum(leave_one_site_out$successful_refits),
    " primary fixed-site refits across 68 associations and nine omitted sites"
  ),
  "Identifies estimates whose direction or magnitude depends on one site.",
  "A stability description, not another significance family.",
  "Gap-timing-unaware near eye",
  paste0(
    "68 factor–metric cells; ", gap_sample_range$participants_min, "–",
    gap_sample_range$participants_max, " participants, ",
    gap_sample_range$days_min, "–", gap_sample_range$days_max,
    " represented days, ", gap_sample_range$sites, " sites"
  ),
  "Checks sensitivity to metric-specific handling of remaining gap timing.",
  "Uses the same model implementation and its own complete 68-test family.",
  "Exactly identified longest period",
  paste0(
    nrow(exact_period), " factor models; ",
    first(exact_period$participants), " participants, ",
    first(exact_period$participant_days), " participant-days, ",
    first(exact_period$sites), " sites"
  ),
  "Removes lower-bound ambiguity for the longest continuous period above 250 lx melEDI.",
  "Retain only participant-days whose longest period is exactly identified.",
  "Participant-level Spearman summaries",
  "Participant aggregates with 95% descriptive intervals",
  "Provides a direct rank-based description and site-stratified context.",
  "No p-values and no secondary significance screen."
)

sensitivity_map |>
  gt::gt() |>
  gt::cols_label(
    Exact_stored_scope = "Exact stored scope",
    Interpretive_rule = "Interpretive rule"
  ) |>
  gt::cols_width(
    Analysis ~ gt::pct(20),
    Exact_stored_scope ~ gt::pct(28),
    Purpose ~ gt::pct(24),
    Interpretive_rule ~ gt::pct(28)
  ) |>
  prep_gt(11)
```

The paired/common placement figure in the results report uses near eye on the
x-axis and chest on the y-axis, matched factor–metric estimands only, identity
and null lines, and equal axis geometry. Its exact source is
[H05_paired_effect_comparison_data.csv](../../../artifacts/11_source_data/H05/H05_paired_effect_comparison_data.csv).
It describes agreement and differences without claiming equivalence. The
stored component estimates come from the exact common sample; no model was
fitted solely for that plot.

Complete stored sensitivity records are available for the
[paired/common placement comparison](../../../artifacts/09_tables/H05/H05_paired_placement_comparison.csv),
[random-site analysis](../../../artifacts/09_tables/H05/H05_random_site_sensitivity.csv),
[leave-one-site-out analysis](../../../artifacts/09_tables/H05/H05_leave_one_site_out_summary.csv),
[gap-timing-unaware analysis](../../../artifacts/09_tables/H05/H05_manuscript_prepared_comparison.csv),
and [exactly identified longest-period analysis](../../../artifacts/09_tables/H05/H05_exactly_identified_longest_bout_sensitivity.csv).

## 7. Code map and execution order

### Function modules

```{r}
#| label: tbl-h05-prep-module-map
#| tbl-cap: "Function modules used by the H05 executable analysis."

tibble::tribble(
  ~Module, ~Reads_or_defines, ~Contribution, ~Why_separate,
  "scripts/pipeline/paths_io.R", "Artifact-writing and SHA-256 helpers", "Writes typed artifacts with hashes, byte sizes, producers, and R version.", "Keeps file identity rules common across analyses.",
  "scripts/pipeline/multiplicity.R", "Complete-family adjustment helpers", "Applies the 68-value Benjamini–Hochberg vectors and verifies family size.", "Separates multiplicity from individual model code.",
  "scripts/hypotheses/H01/h01_contract.R", "Shared 17-metric definitions", "Supplies metric transforms, admissible supports, display names, and units.", "Keeps the H05 responses identical to the verified H01 metric definitions.",
  "scripts/hypotheses/H01/h01_modeling.R", "Shared response and diagnostic functions", "Transforms responses, extracts practical-scale effects, and supplies common diagnostics.", "Prevents H05-specific implementations from drifting from H01.",
  "scripts/hypotheses/H05/h05_contract.R", "Factors, runs, formulas, families, pinned inputs", "Defines the exact H05 estimands and analysis registry.", "Provides one auditable specification before any model is run.",
  "scripts/hypotheses/H05/h05_modeling.R", "Frame construction, engines, summaries, diagnostics", "Implements the H05 models and sensitivity functions against the contract.", "Keeps reusable scientific operations outside the executable driver."
) |>
  gt::gt() |>
  gt::cols_width(
    Module ~ gt::pct(25),
    Reads_or_defines ~ gt::pct(24),
    Contribution ~ gt::pct(29),
    Why_separate ~ gt::pct(22)
  ) |>
  prep_gt(11)
```

### Executable entry points

```{r}
#| label: tbl-h05-prep-script-map
#| tbl-cap: "H05 scripts and documents in analytical and reporting order."

tibble::tribble(
  ~Order, ~Entry_point, ~Reads, ~Produces, ~Runs_when_page_renders,
  1L, "run_h05_stage2.R", "Verified metric datasets, LEBA scores, shared H01 functions, H05 contract", "136 frames, 204 inferential bundles, 544 core result cells, diagnostics, influence and sensitivity records", "No",
  2L, "build_h05_reader_artifacts.R", "Frozen H05 tables and diagnostic plot data", "Reader-facing result tables, seven figures, and exact paired figure source data", "No",
  3L, "notebooks/hypotheses/H05.qmd", "Stored scientific outputs and reader assets", "The standalone H05 results report", "No",
  4L, "build_h05_preparation_artifacts.R", "Frozen H05 frames, registries, samples, and LEBA scores", "Four descriptive source-data CSV files for this page", "No; run before rendering",
  5L, "H05_analysis_preparation.qmd", "Stored H05 artifacts and descriptive source data", "This scientific preparation and provenance companion", "Yes",
  6L, "build_h05_preparation_report_manifest.R", "Final source, website copy, HTML, scripts, scientific outputs, figures, and source data", "Non-circular preparation-report manifest", "No; run after rendering"
) |>
  gt::gt(rowname_col = "Order") |>
  gt::cols_width(
    Entry_point ~ gt::pct(23),
    Reads ~ gt::pct(27),
    Produces ~ gt::pct(36),
    Runs_when_page_renders ~ gt::pct(14)
  ) |>
  prep_gt(10)
```

The scientific model run was executed from the repository root in a clean R
4.6.1 session using the project library. The command is shown for
reproducibility; this page does not execute it.

```bash
Rscript --vanilla scripts/hypotheses/H05/run_h05_stage2.R
Rscript --vanilla scripts/hypotheses/H05/build_h05_reader_artifacts.R
quarto render notebooks/hypotheses/H05.qmd --profile nathealth
Rscript --vanilla scripts/hypotheses/H05/build_h05_preparation_artifacts.R
quarto render audit/hypotheses/H05/H05_analysis_preparation.qmd --profile nathealth
```

## 8. Output and claim provenance

The model-frame index records the exact sample and SHA-256 frame identity for
each run, metric, and factor. The model manifest connects each of those cells
to its full, reduced, and final model object. Result, diagnostic, and
sensitivity CSV files then connect the stored fits to the reader tables and
figures. Figure source-data files contain the exact values used by their plot
layers, so a reader does not need to digitise an image.

```{r}
#| label: tbl-h05-prep-output-identities
#| tbl-cap: "Principal stored H05 outputs and their current identities."

output_registry <- tibble::tribble(
  ~Output, ~Path, ~Producer,
  "Model-frame archive", "artifacts/06_model_data/H05/H05_model_frames.rds", "run_h05_stage2.R",
  "Model-frame index", "artifacts/06_model_data/H05/H05_model_frame_index.csv", "run_h05_stage2.R",
  "Inferential model archive", "artifacts/07_models/H05/H05_inferential_model_objects.rds", "run_h05_stage2.R",
  "Model manifest", "artifacts/07_models/H05/H05_model_manifest.csv", "run_h05_stage2.R",
  "Master results", "artifacts/09_tables/H05/H05_model_results_master.csv", "run_h05_stage2.R",
  "Model diagnostics", "artifacts/08_diagnostics/H05/H05_model_diagnostics.csv", "run_h05_stage2.R",
  "Random-site sensitivity", "artifacts/09_tables/H05/H05_random_site_sensitivity.csv", "run_h05_stage2.R",
  "Leave-one-site-out summary", "artifacts/09_tables/H05/H05_leave_one_site_out_summary.csv", "run_h05_stage2.R",
  "Paired placement source", "artifacts/11_source_data/H05/H05_paired_effect_comparison_data.csv", "build_h05_reader_artifacts.R",
  "Reader report", "_build/nathealth/notebooks/hypotheses/H05.html", "notebooks/hypotheses/H05.qmd"
) |>
  rowwise() |>
  mutate(
    `SHA-256 prefix` = substr(
      artifact_sha256(file.path(root, .data$Path)),
      1L,
      16L
    ),
    Bytes = file.info(file.path(root, .data$Path))$size
  ) |>
  ungroup()

output_registry |>
  gt::gt() |>
  gt::fmt_integer(columns = Bytes, use_seps = TRUE) |>
  gt::cols_width(
    Output ~ gt::pct(20),
    Path ~ gt::pct(46),
    Producer ~ gt::pct(18),
    `SHA-256 prefix` ~ gt::pct(12),
    Bytes ~ gt::pct(8)
  ) |>
  prep_gt(10)
```

The H05 preparation-report manifest is created after the final website render.
It records complete identities for the authoring source, the byte-identical
website QMD copy, rendered HTML, page assets, H05 scripts and tests, scientific
outputs, source data, shared configuration snapshot, and R environment. The
manifest excludes itself and the handoff that quotes its identity, preventing
a circular checksum chain.

## 9. How to examine or reproduce H05

A reader can follow the analysis in five passes:

1. inspect `H05_input_audit.csv`, `h05_contract.R`, and the factor, metric,
   formula, and run registries to verify the frozen inputs and specifications;
2. compare the exact samples above with `H05_model_frame_index.csv` and the
   136 frames in `H05_model_frames.rds`;
3. follow `run_h05_stage2.R` through `h05_modeling.R` to connect each frame to
   its model object, coefficient, interval, likelihood-ratio test, diagnostic,
   and sensitivity record;
4. inspect the complete 68-value family audit before interpreting any
   individual association; and
5. read the [H05 results report](../../../notebooks/hypotheses/H05.html)
   alongside its tables, figures, and paired source data.

The most important boundary is that the preparation record documents a frozen
analysis; it does not create an alternate result. If an input, metric rule,
response model, or estimand changes, the affected scientific pipeline must be
rerun before this page is refreshed.

## Environment

```{r}
#| label: tbl-h05-prep-environment
#| tbl-cap: "Consequential R environment recorded for H05 analysis and reporting."

package_names <- c(
  "dplyr", "readr", "digest", "lme4", "glmmTMB", "DHARMa", "ggplot2", "gt"
)
package_versions <- vapply(
  package_names,
  function(package) as.character(utils::packageVersion(package)),
  character(1)
)

bind_rows(
  tibble::tibble(
    Component = "R",
    Version = as.character(getRversion()),
    Role = "Authoritative scientific computation and Quarto rendering"
  ),
  tibble::tibble(
    Component = package_names,
    Version = unname(package_versions),
    Role = c(
      "Data manipulation",
      "Typed CSV input",
      "SHA-256 and model-frame identities",
      "Gaussian mixed models",
      "Tweedie mixed models",
      "Simulation-based residual diagnostics",
      "Figures",
      "Reader-facing tables"
    )
  )
) |>
  gt::gt(rowname_col = "Component") |>
  gt::cols_width(
    Version ~ gt::pct(18),
    Role ~ gt::pct(62)
  ) |>
  gt::tab_source_note(
    "The project renv.lock records the full dependency set; artifact manifests record R 4.6.1 with each output identity."
  ) |>
  prep_gt()
```

```{r}
#| label: session-info-h05-preparation
#| include: false

sessionInfo()
```
