---
title: "H11 Stage 1: audit and proposed analysis"
subtitle: "Nature Health analysis audit — author approval gate"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show audit code"
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

library(dplyr)
library(tibble)
library(gt)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        dir.exists(file.path(candidate, "audit"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H11_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

source(file.path(project_root, "scripts/hypotheses/H11/h11_stage1.R"))
stage1 <- h11_stage1_artifacts(project_root)
formulas <- h11_formula_set(project_root)

artifact_dir <- file.path(project_root, "artifacts/06_model_data/H11")
artifact_manifest <- read.csv(
  file.path(project_root, "artifacts/12_manifests/H11/H11_stage1_artifacts.csv"),
  stringsAsFactors = FALSE,
  check.names = FALSE
)
stopifnot(
  nrow(artifact_manifest) == length(stage1),
  all(stage1$input_provenance$hash_verified),
  all(stage1$candidate_sample_counts$missing_biological_sex_rows == 0L)
)

audit_gt <- function(data, title, subtitle = NULL) {
  data |>
    gt() |>
    tab_header(
      title = md(title),
      subtitle = if (is.null(subtitle)) NULL else md(subtitle)
    ) |>
    opt_row_striping() |>
    tab_options(
      table.width = pct(100),
      data_row.padding = px(4),
      heading.align = "left"
    )
}

placement_label <- function(x) {
  ifelse(x == "glasses", "Near-eye", "Chest")
}
```

## Stage 1 verdict

**The H11 audit and proposed specification are complete, but Stage 2 is
blocked pending explicit author approval.** No proposed H11 inferential model,
bootstrap, coefficient simulation, or sensitivity model has been fitted.

The recommended primary analysis asks whether the complete biological-sex
contribution---a constant level plus a cyclic time-varying deviation---improves
the finalized H02 temporal model. It uses near-eye measurements as primary and
recreates the same analysis for chest measurements as complementary evidence.
The proposed model sequence restores the registered parametric sex term,
inherits the accepted H02 clock-time, participant, participant-day, site, and
AR-boundary structure without reopening it, compares candidate mean structures
under one common correlation specification, and replaces pointwise period
claims with a global comparison and a simultaneous 95% curve band.

The current submitted-analysis claims are reopened. In particular, this audit
does not support the existing claims of a significant 08:00--10:00 interval,
a factor-two peak, less than 1% variance explained, a retained chest sex term,
or a behaviourally driven mechanism.

## Controlling evidence and provenance

The signed AsPredicted PDF was indexed in a temporary, non-repository
page-level index. Its SHA-256 is
`fce192c88676602b72013e4872a03b535a2f4ef59d34de3a9e1c47527742b7ad`;
both pages were text-searchable and the index verified against the source.
Exact searches located the hypothesis and outcome on PDF p. 1 and the model,
FDR, GAMM ΔAIC, diagnostics, and AR(1) rules on PDF p. 2.

The H02 handoff and 174-file manifest also match their authoritative hashes:

- handoff: `5f4bc01d7cb3d7c6e61fa1b70f2ed5fbdc4e6b76fed3542b315ba3f35b24157b`;
- manifest: `a978faedbbcbf337a2ed3bb624d5704c50c4929334f62a6c159f8417b65098c9`;
- selected temporal specification:
  `c3c95e97dbf7a260e4aa7513a15bb8c3c7e98bdb754d97c54700e4d82629310f`.

```{r}
#| label: key-input-provenance
#| echo: false

stage1$input_provenance |>
  filter(input_id %in% c(
    "signed_preregistration",
    "h02_worker_handoff",
    "h02_worker_manifest",
    "h02_selected_temporal_specification",
    "h11_developed_plan",
    "demographics",
    "v0_near_eye_render",
    "v0_chest_render"
  )) |>
  transmute(
    Source = input_id,
    Path = relative_path,
    `SHA-256` = observed_sha256,
    Verified = hash_verified
  ) |>
  audit_gt(
    "Key frozen inputs verified at render time",
    "The complete 44-row provenance table is exported with the Stage 1 artifacts."
  ) |>
  cols_width(Path ~ px(330), `SHA-256` ~ px(430))
```

The audit used R and did not use another language to calculate or adjudicate a
scientific result.

```{r}
#| label: environment
#| echo: false

stage1$environment |>
  setNames(c("Component", "Version")) |>
  audit_gt("Analytical environment")
```

## Preregistered contract

The exact signed H11 hypothesis is:

> H11: Diurnal exposure patterns differ by sex.

The registered H11 outcome is the hourly geometric mean of melanopic EDI. The
signed document places chest measurements first and repeats the analysis with
glasses-mounted measurements. The later project-wide approved placement rule
reverses those roles: near-eye is primary and chest is complementary.

The signed method is printed below as the actual R formula object recovered by
the Stage 1 formula builder.

```{r}
#| label: registered-formula
#| code-summary: "Registered H11 formula"

formulas$registered_h11
```

The signed general rules require FDR correction within hypothesis families,
GAMM model selection using a ΔAIC threshold of at least 2, residual,
homoscedasticity, and normality checks, and AR(1) errors when residual
autocorrelation is detected.

### Deviations relevant to H11

The shared deviation register already identifies `DEV-042` through `DEV-048`,
plus `DEV-019`, `IMP-005`, and `DOC-001`. The reader-facing shared deviation
narrative currently stops after H07; this H11 worker therefore proposes the
needed central entries in its handoff rather than editing a coordinator-owned
file.

The finalized H02 handoff now resolves parts of the earlier H11 migration
gate. H11 must use the supported 30-minute arithmetic-mean outcome as primary,
the H02 transform and basis dimensions, the H02 site/participant/day
hierarchy, and the H02 boundary-aware AR algorithm. The registered hourly
geometric-mean outcome remains a named sensitivity. Those inherited choices
are audited here but are not reselected using H11 results.

## The measured construct: biological sex, not gender

The normalized demographics artifact has separate `sex` and `gender` fields.
H11 uses only the measured biological-sex field. It is complete and contains
105 participants recorded as Female and 86 recorded as Male in the full
demographic source. Gender contains 106 Woman, 84 Man, and one Non-binary
record; two records do not follow a simple Female/Woman or Male/Man mapping.
No gender value is converted into sex, and gender does not enter an H11 model.

```{r}
#| label: sex-gender-crosswalk
#| echo: false

stage1$sex_gender_crosswalk |>
  rename(
    `Biological sex` = biological_sex,
    Gender = gender,
    Participants = participants
  ) |>
  audit_gt(
    "Separate source fields for biological sex and gender",
    "Aggregated counts only; the H11 predictor is biological sex."
  )
```

For modelling, `sex` is an unordered treatment-coded factor with Male as the
reference. A second variable, `sex_smooth`, is an ordered factor with levels
Male then Female. It is derived from the same biological-sex field solely to
request one Female-minus-Male cyclic deviation smooth from `mgcv`; it is not a
different construct.

## Exact candidate samples

### Reconciliation to H02

Joining biological sex by the unique `site × Id` key changes no row and
introduces no missing value. All three proposed comparison formulas therefore
use the same frame within a placement and scenario.

The primary model frames reconcile exactly to the accepted H02 counts:

- near-eye: **141 participants, 816 participant-days, 37,756 30-minute
  observations, and 9 sites**;
- chest: **154 participants, 902 participant-days, 41,842 30-minute
  observations, and 8 sites**.

```{r}
#| label: all-candidate-samples
#| echo: false

stage1$candidate_sample_counts |>
  mutate(
    Placement = placement_label(placement),
    Scenario = recode(
      data_scenario_id,
      main = "Main H02-prepared",
      manuscript_prepared_data = "Manuscript-prepared",
      registered_outcome_sensitivity = "Registered hourly sensitivity"
    ),
    Sample = recode(
      sample_scenario,
      all_available = "All available",
      paired_common_sample = "Paired/common",
      all_available_h02_day_domain = "All available H02 day domain"
    )
  ) |>
  select(
    Scenario, Placement, Sample, participants, participant_days,
    observations_or_hours, observation_unit, sites,
    female_participants, male_participants
  ) |>
  rename(
    Participants = participants,
    `Participant-days` = participant_days,
    `Rows entering each candidate model` = observations_or_hours,
    `Row unit` = observation_unit,
    Sites = sites,
    `Female participants` = female_participants,
    `Male participants` = male_participants
  ) |>
  audit_gt(
    "Exact candidate samples for M0, M_level, and M_pattern",
    "Within each row, all three proposed formulas share the exact frame."
  ) |>
  fmt_integer(columns = c(
    Participants, `Participant-days`, `Rows entering each candidate model`,
    Sites, `Female participants`, `Male participants`
  ))
```

The registered-outcome sensitivity is already countable without fitting a
model. On the accepted H02 participant-day domain it contains 18,953 supported
near-eye hours and 21,004 supported chest hours. Every accepted participant-day
has at least one supported hour. The hourly metric is used only when at least
30 of 60 one-minute intervals are valid; an under-supported raw geometric
summary remains masked from the candidate model value.

### Main-versus-manuscript common rows

The Stage 2 data-preparation comparison will use the same new implementation
on both inputs. Its exact common keys are fixed before modelling.

```{r}
#| label: main-manuscript-common
#| echo: false

stage1$main_manuscript_common_samples |>
  transmute(
    Placement = placement_label(placement),
    Participants = participants,
    `Participant-days` = participant_days,
    `Common 30-minute observations` = common_observations_30_minute,
    Sites = sites
  ) |>
  audit_gt(
    "Exact main/manuscript common sample",
    "Used to distinguish data-preparation changes from model changes in Stage 2."
  ) |>
  fmt_integer(columns = everything())
```

### Site-by-sex support

Both recorded biological-sex categories occur at every included site, but
some cells are small. In the primary near-eye frame, San José has four Female
and two Male participants; Munich has six Female and four Male participants.
This makes site influence and the common-across-sites sex assumption important
diagnostics. H11 does not introduce a sex-by-site smooth or interaction after
seeing results.

```{r}
#| label: site-sex-near-eye
#| echo: false

stage1$site_sex_support |>
  filter(placement == "glasses") |>
  transmute(
    Site = display_name,
    Colour = paste0(
      "<span style=\"color:", color_hex, ";\">●</span> ", color_hex
    ),
    `Biological sex` = biological_sex,
    Participants = participants,
    `Participant-days` = participant_days,
    `30-minute observations` = observations_30_minute
  ) |>
  audit_gt(
    "Primary near-eye site-by-sex support",
    "Submitted-manuscript site names and order; internal codes are not displayed."
  ) |>
  fmt_integer(columns = c(Participants, `Participant-days`, `30-minute observations`)) |>
  fmt_markdown(columns = Colour)
```

```{r}
#| label: site-sex-chest
#| echo: false

stage1$site_sex_support |>
  filter(placement == "chest") |>
  transmute(
    Site = display_name,
    Colour = paste0(
      "<span style=\"color:", color_hex, ";\">●</span> ", color_hex
    ),
    `Biological sex` = biological_sex,
    Participants = participants,
    `Participant-days` = participant_days,
    `30-minute observations` = observations_30_minute
  ) |>
  audit_gt(
    "Complementary chest site-by-sex support",
    "Tübingen is absent because the accepted chest frame contains eight sites."
  ) |>
  fmt_integer(columns = c(Participants, `Participant-days`, `30-minute observations`)) |>
  fmt_markdown(columns = Colour)
```

## What V0 estimated

V0 models the zero-aware base-10 log of the 30-minute arithmetic mean melEDI.
It includes common, photoperiod-state, site, participant, and participant-day
temporal terms. Its full model adds only a non-explicitly-cyclic
sum-to-zero sex smooth; it omits the registered parametric sex term. It fits
each formula first without AR(1), estimates a separate `rho` for each formula,
refits with participant-only `AR.start`, and compares the final fREML AICs.

The exact V0 formula objects recovered from `RQ3.qmd` and
`RQ3_chest.qmd` are printed below. The structures are identical between
placements, so one object per distinct structure is sufficient.

```{r}
#| label: v0-formulas
#| code-summary: "Exact V0 H11 formula objects"

for (formula_id in c(
  "v0_temporal_null",
  "v0_temporal_full",
  "v0_activity_only",
  "v0_activity_plus_sex"
)) {
  cat("\n", formula_id, "\n", sep = "")
  print(formulas[[formula_id]])
}
```

### Recovered V0 numerical facts

```{r}
#| label: v0-results
#| echo: false

stage1$v0_result_summary |>
  mutate(
    Placement = placement_label(placement),
    Comparison = comparison
  ) |>
  select(
    Placement, Comparison, participants, participant_days,
    observations_30_minute, reduced_AIC, full_AIC,
    delta_AIC_reduced_minus_full, smooth_sex_p, adjusted_R2
  ) |>
  rename(
    Participants = participants,
    `Participant-days` = participant_days,
    `30-minute observations` = observations_30_minute,
    `Reduced AIC` = reduced_AIC,
    `Full AIC` = full_AIC,
    `ΔAIC (reduced − full)` = delta_AIC_reduced_minus_full,
    `V0 sex-smooth p` = smooth_sex_p,
    `V0 adjusted R²` = adjusted_R2
  ) |>
  audit_gt(
    "Recovered V0 model summaries",
    "These values reproduce the rendered record; the AIC comparisons are not decision-valid."
  ) |>
  fmt_number(
    columns = c(`Reduced AIC`, `Full AIC`, `ΔAIC (reduced − full)`),
    decimals = 2
  ) |>
  fmt_number(columns = c(`V0 sex-smooth p`, `V0 adjusted R²`), decimals = 4)
```

V0 near-eye favours its full formula by about 7.82 AIC units, whereas chest
favours its null by about 3.13. The chest prose nevertheless says to retain
sex. The activity comparison similarly favours adding the sex smooth by only
about 1.74 AIC units near eye and favours omitting it by about 3.11 for chest.
These comparisons are descriptive recovery, not accepted inference, because
the compared fits use differing fREML mean structures and separately estimated
correlation parameters.

### Confirmed findings

```{r}
#| label: critical-findings
#| echo: false

stage1$findings |>
  filter(severity == "critical") |>
  select(finding_id, area, finding, stage1_disposition) |>
  rename(
    ID = finding_id,
    Area = area,
    Finding = finding,
    `Required disposition` = stage1_disposition
  ) |>
  audit_gt("Critical H11 findings")
```

```{r}
#| label: other-findings
#| echo: false

stage1$findings |>
  filter(severity != "critical") |>
  select(finding_id, severity, area, finding, stage1_disposition) |>
  rename(
    ID = finding_id,
    Severity = severity,
    Area = area,
    Finding = finding,
    `Required disposition` = stage1_disposition
  ) |>
  audit_gt("Major and minor H11 findings")
```

## Distinguishing data-preparation from model changes

Stage 2 will make three comparisons, in this order:

1. reproduce V0 from its frozen outputs and data description;
2. apply the new implementation to the manuscript-prepared frame, isolating
   model and inference changes as far as the frozen artifacts permit; and
3. compare that fit with the identical new implementation on the main H02
   frame, isolating data-preparation changes on both full and exact common
   samples.

| Element | V0 | Proposed Stage 2 | Change class |
|---|---|---|---|
| Near-eye frame | 141 participants / 809 days / 37,603 rows | 141 / 816 / 37,756 | Data preparation |
| Chest frame | 154 participants / 894 days / 41,664 rows | 154 / 902 / 41,842 | Data preparation |
| Sex construct | Source `sex`, but no parametric term | Biological `sex`; Male-reference level plus cyclic deviation | Model/estimand |
| H02 hierarchy | V0 working hierarchy and photoperiod-state term | Finalized H02 common/site/participant/day structure | Model/provenance |
| AR boundaries | Participant start only | Participant-day, omitted/discontinuous bin, and fall-back boundaries | Dependence implementation |
| Mean-structure comparison | Separate-rho fREML AIC | Common-rho ML likelihood/AIC comparison | Inference implementation |
| Time localization | Pointwise interval | Simultaneous 95% curve band | Multiplicity/inference |
| Activity claim | Sample-changing adjustment interpreted causally | Deferred exact-common-sample contextual sensitivity | Estimand/claim |

No new numerical result is available in Stage 1, so no direction, magnitude,
or significance comparison is made here.

## Proposed estimand and analysis unit

### Primary estimand

The primary estimand is the joint contribution of recorded biological sex to
the conditional 24-hour local-clock trajectory of
`log10(30-minute arithmetic-mean melEDI + 0.1 lx)` in the all-available
near-eye H02 frame. It combines:

- a constant Female-minus-Male level difference; and
- a cyclic Female-minus-Male deviation over clock time.

The global primary comparison is `M0` versus `M_pattern`. The model's
site-temporal term adjusts for the accepted site patterns; participant-specific
curves and participant-day intercepts account for repeated measurements. For
the displayed Female-minus-Male contrast, the same site, participant, and day
terms are used on both sides and therefore cancel. The sex effect is assumed
common across sites; site-by-sex interaction is not added post hoc.

The likelihood is row-based: the analysis unit is one supported 30-minute
observation. Repeated observations are nested in participant-days and
participants, and AR(1) dependence is allowed only within the uninterrupted
H02-defined sequences. The estimand is not a participant-balanced average and
is not causal.

### Practical scale

The fitted response is the conditional mean on the base-10 log scale. The
Female-minus-Male curve is reported both as a log-scale difference and as
`10^difference`, the ratio of the modelled shifted values
`(melEDI + 0.1 lx)`. Sex-specific curves are displayed after
`10^prediction - 0.1 lx`. These are back-transformed conditional log-scale
curves, not untransformed arithmetic means of raw one-minute measurements.

Every estimate will carry a 95% confidence interval. A claimed time interval
must be supported by the simultaneous band; pointwise bands may be displayed
only when explicitly labelled descriptive.

## Exact proposed models

The H02-selected formula is the exact source of `M0`. `M_level` adds the
parametric biological-sex term. `M_pattern` adds one explicitly cyclic
Female-minus-Male deviation by using the ordered `sex_smooth` factor. The
common curve remains the Male reference curve; this is the ordered-factor
reference/difference construction documented by the installed `mgcv` 1.9-4.

The same generic `response` formula is used for near-eye and chest, main and
manuscript-prepared data, and the registered hourly sensitivity because the
right-hand side is identical. Only the response construction, frame, and
placement-specific `rho` differ.

```{r}
#| label: proposed-formulas
#| code-summary: "Exact proposed H11 formula objects"

for (formula_id in c("proposed_m0", "proposed_mlevel", "proposed_mpattern")) {
  cat("\n", formula_id, "\n", sep = "")
  print(formulas[[formula_id]])
}
```

No proposed formula contains gender, photoperiod state, an activity term, or
a sex-by-site interaction.

### Formula-to-fit map

```{r}
#| label: formula-manifest
#| echo: false

stage1$formula_manifest |>
  filter(grepl("proposed_", formula_id, fixed = TRUE)) |>
  select(
    formula_id, analytical_role, family, link,
    estimation_and_correlation, contrasts, knots
  ) |>
  rename(
    `Formula ID` = formula_id,
    Role = analytical_role,
    Family = family,
    Link = link,
    `Estimation and correlation` = estimation_and_correlation,
    Contrasts = contrasts,
    Knots = knots
  ) |>
  audit_gt(
    "Proposed formula and fitting contract",
    "The full formula text is exported in the machine-readable manifest."
  )
```

## Proposed fitting and comparison sequence

For each placement and scenario, Stage 2 will:

1. join demographics many-to-one by `site × Id`, assert no row change or
   missing biological sex, and save the exact candidate frame;
2. preserve H02 row order, `time_hour`, `AR_start`, participant, and
   participant-day keys for the 30-minute analyses;
3. fit one `M_pattern` preliminary model with `method = "fREML"`, `rho = 0`,
   and the H02 knots to estimate a boundary-aware lag-1 `rho`;
4. hold that single `rho`, frame, row order, and `AR.start` vector fixed while
   fitting `M0`, `M_level`, and `M_pattern` with `method = "ML"` and
   `discrete = FALSE` for mean-structure comparison;
5. fail the comparison if model-frame hashes, row counts, sequence starts, or
   correlation settings differ;
6. report ML log likelihoods, effective degrees of freedom, likelihood-ratio
   statistics, raw and adjusted p-values, AIC, and
   `AIC(reduced) - AIC(full)`;
7. fit `M_pattern` with `method = "fREML"`, `discrete = TRUE`, and the same
   fixed `rho` for final estimation, regardless of whether the global result
   supports a sex contribution; and
8. persist every frame, fit, warning, smoothing parameter, `rho`, sequence
   count, diagnostic, contrast, and display-source row.

All fits use Gaussian errors with identity link,
`knots = list(time_hour = c(0, 24))`, one thread, and the installed
`mgcv` 1.9-4. H02's accepted `rho` values are not copied; `rho` is re-estimated
for each H11 placement/scenario from the H11 full preliminary residuals.

The registered GAMM rule is retained: the full model is AIC-supported only
when the reduced-minus-full ΔAIC is at least 2. The ML likelihood-ratio p-value
is also reported. If the AIC rule and global p-value disagree, the result is
labelled inconclusive rather than selecting the favourable criterion.

## Planned contrasts and multiplicity

| Output | Model comparison or contrast | Role | Multiplicity |
|---|---|---|---|
| Global sex contribution | `M0` versus `M_pattern` | Primary confirmatory near-eye test | `H11-F1-primary`, BH with planned n = 1 |
| Constant level | `M0` versus `M_level`; parametric Female-minus-Male estimate | Secondary decomposition | `H11-F2-level-shape`, BH across 2 tests |
| Time-varying shape | `M_level` versus `M_pattern` | Secondary decomposition | `H11-F2-level-shape`, BH across 2 tests |
| Female-minus-Male curve | Linear-predictor difference at identical site/participant/day settings | Practical localization | Simultaneous 95% band across the daily grid |
| Chest repeat | Same global and decomposition sequence | Complementary evidence, not a second primary rejection opportunity | `H11-C1-global`, BH n = 1; `H11-C2-level-shape`, BH n = 2 |
| Registered hourly and other sensitivities | Same predeclared outputs | Stability evidence only | Cannot replace the primary conclusion; p-values are descriptive, and any displayed decomposition pair is BH-adjusted with n = 2 within placement |

Near-eye and chest families are not pooled: chest cannot rescue or overturn
the primary rejection rule. Manuscript-prepared, paired/common, hourly,
leave-one-site-out, and activity outputs cannot create a new primary
rejection. No leave-one-site-out p-value screen is performed.

The simultaneous band will be calibrated from the full coefficient covariance,
including the parametric sex term, using a fixed-seed 10,000-draw coefficient
simulation on a 5-minute grid over `[0, 24)`. The critical value is the maximum
standardized absolute contrast across the grid. A 1-minute and 15-minute grid
check must not materially change a localized interval. This is a bounded,
non-bootstrap calculation and does not require a production-computation stage.
If implementation instead requires participant resampling or another heavy
bootstrap, only a 50- or 100-replicate pilot may run, and the task must return
to the author before a full run.

## Diagnostics and acceptance assessment

Stage 2 must show each diagnostic, interpret it in plain language, and classify
the near-eye and chest analyses separately as **acceptable**, **acceptable
with specified limitations**, or **not acceptable for inference**.

| Domain | Required check | Consequence |
|---|---|---|
| Data and design | Exact participants, participant-days, rows/hours, sites, site-by-sex cells, missingness before/after join, factor references, design rank | Any unexplained row change, missing sex, or rank deficiency stops inference |
| Temporal provenance | True-UTC sort, 30-minute or one-hour epoch, fold provenance, every H02 AR boundary, sequence counts, no crossing of a day/gap/non-one-to-one outcome | Any boundary violation is not acceptable |
| Cyclic topology | Predicted common and sex-difference curves agree at 0 and 24 h to numerical tolerance | Failure blocks time-of-day interpretation |
| Fit stability | Convergence state, warnings, finite coefficients, rank, smoothing parameters, effective degrees of freedom | Unresolved convergence/rank failure is not acceptable |
| Basis adequacy | `k` checks, EDF relative to available basis, residual pattern by clock time | Evidence of an undersized sex basis reopens the specification rather than triggering silent tuning |
| Residual distribution | Residual-versus-fitted, Q--Q/tails, scale-location, residuals by sex/site/time, exact-zero behaviour | Material heteroscedasticity or tails are described and can limit interval interpretation |
| Autocorrelation | Raw and standardized residual ACF within valid sequences, boundary-aware lag correlations, comparison with the H02 diagnostic profile | Material residual dependence is an explicit limitation or can make inference unacceptable |
| Concurvity | Full formal concurvity, with particular attention to common time, site time, and sex deviation | A poorly identified sex deviation prevents a sex-pattern claim |
| Influence | Participant/day summaries, site deletion, and leave-one-site-out sex curves without selecting a favourable site subset | A conclusion driven by one site is unstable and cannot be generalized across sites |
| Inference integrity | Common-frame/common-rho checks, AIC/LRT agreement status, raw/adjusted p-values, simultaneous-band coverage and grid check | Any provenance mismatch invalidates the comparison |
| Display integrity | Every plotted time maps one-to-one to exported source data; 95% intervals, placement, scale, and sample labels reconcile | Unreconciled figures or tables are not publishable |

The H02 diagnostic inheritance is explicit: its residual zero-bound,
heteroscedasticity/tail, residual-autocorrelation, and common/site-concurvity
limitations are not erased by reusing the model. H11 adds a sex-specific
identifiability question. A satisfactory H02 diagnostic profile does not by
itself establish that the H11 sex deviation is estimable.

## Sensitivity battery and V0 recreation plan

```{r}
#| label: sensitivity-plan
#| echo: false

stage1$sensitivity_plan |>
  rename(
    Scenario = scenario_id,
    Role = role,
    `Data change` = data_change,
    `Model change` = model_change,
    `Comparison strategy` = comparison_strategy,
    Stage = stage
  ) |>
  audit_gt(
    "Predeclared H11 sensitivity and diagnostic scenarios",
    "Expensive or underspecified side analyses remain behind their stated gate."
  )
```

The Stage 2 core includes the main near-eye and chest fits, the registered
hourly outcome, manuscript-prepared data, paired/common placement samples, and
the feasible leave-one-site-out diagnostic. The H02 model-form alternative and
activity analysis are deferred. The all-zero-inclusive scenario cannot be fit
until the coordinator supplies an approved H02-compatible sensitivity frame;
this worker will not reconstruct a shared input independently.

V0 figures will be recreated from their recovered specification before being
corrected. The near-eye and chest displays will retain the recognizable
sex-specific curve and contrast-panel layout, but the caption will refer to
biological-sex differences, not work/free days; pointwise red segments will
not be treated as inferential; and the new display will show 95% uncertainty,
exact samples, scale, outcome, and placement. V0 and new displays remain in the
Stage 2 audit only. The final reader-facing notebook will contain no V0 or
construction terminology.

The activity scenario, if later authorized, must use the audited H04 activity
dictionary and three fits on one exact activity-complete frame: unrestricted
main result restricted to the activity-complete sample, the same restricted
model without activity, and then the activity-adjusted model. It may support
only wording such as “attenuated after adjustment for recorded activity
context”; it cannot identify mediation or a behaviourally driven mechanism.

## Author decisions required before Stage 2

Every row below requires explicit approval. Approval of the table authorizes
implementation of the recommended dispositions, not any deferred analysis
whose row states that a later gate is required.

```{r}
#| label: author-decisions
#| echo: false

stage1$author_decisions |>
  rename(
    ID = decision_id,
    `Decision required` = decision_required,
    `Recommended disposition` = recommended_disposition,
    `Why it matters` = why_it_matters
  ) |>
  audit_gt(
    "H11 author-approval checklist",
    "Stage 2 remains blocked until all items are resolved."
  )
```

## Gate status

**Stop here.** Do not fit `M0`, `M_level`, or `M_pattern`; do not run the
coefficient simulation, activity model, model-form sensitivity, bootstrap, or
full sensitivity battery; do not edit `notebooks/hypotheses/H11.qmd`; and do
not change a shared ledger, configuration file, manuscript, or preparation
artifact until the author explicitly approves the Stage 1 specification in
this task.

To open Stage 2, the author should state that the H11 Stage 1 proposal and the
11-item approval checklist are approved, or list the item IDs that must be
changed.
