---
title: "H11 Stage 2: fast robust inferential-method amendment"
subtitle: "Nature Health analysis audit — author approval gate"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
    code-tools: true
    embed-resources: true
    html-math-method: katex
execute:
  echo: true
  warning: false
  message: false
  error: false
  freeze: false
---

```{r}
#| label: setup
#| include: false

options(
  stringsAsFactors = FALSE,
  scipen = 999,
  width = 120,
  dplyr.summarise.inform = FALSE
)

library(dplyr)
library(gt)
library(readr)
library(tibble)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (
      file.exists(file.path(candidate, "renv.lock")) &&
        dir.exists(file.path(candidate, "audit"))
    ) {
      return(candidate)
    }
    parent <- dirname(candidate)
    if (identical(parent, candidate)) {
      stop("Could not locate the project root from: ", start)
    }
    candidate <- parent
  }
}

project_root <- Sys.getenv(
  "H11_PROJECT_ROOT",
  unset = Sys.getenv("QUARTO_PROJECT_DIR", unset = "")
)
if (!nzchar(project_root) || !file.exists(file.path(project_root, "renv.lock"))) {
  project_root <- find_project_root()
}
project_root <- normalizePath(project_root, winslash = "/", mustWork = TRUE)

method_directory <- file.path(
  project_root,
  "artifacts/08_diagnostics/H11/stage2/method_amendment"
)
read_method_artifact <- function(filename) {
  path <- file.path(method_directory, filename)
  if (!file.exists(path) || file.info(path)$size == 0) {
    stop("Missing H11 method-amendment input: ", path)
  }
  readr::read_csv(path, show_col_types = FALSE)
}

results <- read_method_artifact("robust_test_results.csv")
diagnostics <- read_method_artifact("robust_diagnostics.csv")
runtime <- read_method_artifact("runtime_comparison.csv")
proposal <- read_method_artifact("method_proposal.csv")
manifest <- readr::read_csv(
  file.path(
    project_root,
    "artifacts/12_manifests/H11/H11_robust_method_feasibility_hashes.csv"
  ),
  show_col_types = FALSE
)

stopifnot(
  nrow(results) == 6L,
  nrow(diagnostics) == 2L,
  all(results$status == "method_feasibility_only_not_accepted_inference"),
  all(diagnostics$status == "method_feasibility_only_not_accepted_inference"),
  all(file.exists(file.path(project_root, manifest$path)))
)

amendment_gt <- function(data, title, subtitle = NULL) {
  data |>
    gt() |>
    tab_header(
      title = md(title),
      subtitle = if (is.null(subtitle)) NULL else md(subtitle)
    ) |>
    opt_row_striping() |>
    tab_options(
      table.width = pct(100),
      data_row.padding = px(4),
      heading.align = "left"
    )
}
```

::: {.callout-warning title="Decision required"}
The recommendation is ready, but it is **not yet the accepted H11 test**.
Approve the decision IDs at the end of this report before the stored
feasibility values are interpreted as results or used to open an effect-size
branch. No bootstrap or new model fit was run for this amendment.
:::

## Recommendation

Replace the unfinished `ML`, `discrete = FALSE` likelihood/AIC comparison
branch with **one joint participant-cluster-robust Wald test of the complete
Female-minus-Male temporal curve**, calculated from each already fitted final
`M_pattern` model.

This is the fastest scientifically defensible replacement found in the audit.
It tests the H11 biological-sex contribution directly, retains the finalized
H02 mean structure, clock-time basis, participant/day hierarchy, run-specific
AR(1) coefficient, and exact `AR.start` boundaries, and makes no invalid
restricted-likelihood comparison across different fixed-effect spaces.

The near-eye analysis remains primary. The identical procedure is used for
chest as complementary evidence. The measured construct remains biological
sex; gender is not substituted or analysed.

## Why the approved ML branch is being reopened

H02 fitted all temporal models with `method = "fREML"` and
`discrete = TRUE`; it did not run an equivalent `ML`, `discrete = FALSE`
comparison. The original H11 Stage 1 proposal added ML because `M0` and
`M_pattern` differ in their fixed mean space, which makes their fREML
likelihoods and AIC values unsuitable for formal comparison.

The H11 full-model fREML fits completed on the same order as H02. By contrast,
the exact ML comparison fits ran for hours without producing a durable model
checkpoint. Continuing that branch would require twelve such fits across the
four core placement/scenario runs. The computational failure does not make ML
scientifically invalid, but it makes the approved implementation operationally
unacceptable.

```{r}
#| label: runtime-comparison
#| echo: false

runtime |>
  mutate(
    Placement = recode(
      .data$placement,
      glasses = "Near eye",
      chest = "Chest"
    ),
    `Measured stage` = recode(
      .data$method_stage,
      `H11 final fREML/discrete=TRUE fit` =
        "H11 final fREML fit",
      `Proposed participant-cluster robust post-fit audit` =
        "Robust post-fit test plus feasibility diagnostics",
      `H02 comparable two-fit fREML/discrete=TRUE diagnostic` =
        "H02 comparable two-fit fREML diagnostic"
    ),
    `Elapsed time (seconds)` = round(.data$elapsed_seconds, 1)
  ) |>
  select(Placement, `Measured stage`, `Elapsed time (seconds)`) |>
  arrange(Placement, `Elapsed time (seconds)`) |>
  amendment_gt(
    "Measured durable runtimes",
    paste(
      "The proposed row includes the joint test and its feasibility checks;",
      "the production test alone is a subset of this time."
    )
  ) |>
  cols_width(`Measured stage` ~ px(430))
```

## Exact retained model

No inferential comparison model is added. The test uses the existing final
model fitted by `mgcv::bam()` with Gaussian identity link,
`method = "fREML"`, `discrete = TRUE`, the run-specific accepted `rho`, and
the exact boundary-aware `AR.start` vector.

```{r}
#| label: exact-retained-model-formula
#| eval: false

M_pattern <- response ~
  sex +
  s(time_hour, bs = "cc", k = 12) +
  s(time_hour, by = sex_smooth, bs = "cc", k = 12) +
  s(time_hour, site, bs = "sz", k = 12) +
  s(time_hour, participant, bs = "fs", k = 10) +
  s(participant_day, bs = "re")
```

`sex` is Male-reference treatment coding. `sex_smooth` is the same measured
biological-sex field represented as an ordered Male-then-Female factor, so the
model contains one cyclic Female-minus-Male time deviation in addition to the
parametric Female-minus-Male level.

## Estimand, null, and analysis unit

The estimand remains the complete conditional Female-minus-Male difference in
`log10(30-minute arithmetic-mean melEDI + 0.1 lx)` over a 24-hour clock day,
evaluated at the 48 observed half-hour bin midpoints and equally averaged over
the included fixed site support. Because the model has no sex-by-site term,
the contrast itself contains only the parametric sex coefficient and the
cyclic sex-deviation coefficients.

The global null is

$$
H_0: \Delta(t_j) = 0 \quad \text{for all } j=1,\ldots,48,
$$

where $\Delta(t_j)$ is Female minus Male at clock-bin midpoint $t_j$.

The observation is the analytical unit. The participant is the independent
inference cluster; participant-day random intercepts, participant factor
smooths, and the accepted boundary-aware AR(1) working correlation retain the
H02 repeated-measurement structure.

```{r}
#| label: exact-samples
#| echo: false

diagnostics |>
  transmute(
    Placement = recode(
      .data$placement,
      glasses = "Near eye (primary)",
      chest = "Chest (complementary)"
    ),
    Participants = .data$participants,
    `Participant-days` = .data$participant_days,
    `30-minute observations` = .data$observations_30_minute,
    Sites = .data$sites
  ) |>
  amendment_gt(
    "Exact main fitted-model samples",
    "These reproduce the accepted H02 samples after the biological-sex join."
  ) |>
  fmt_integer(columns = c(
    Participants,
    `Participant-days`,
    `30-minute observations`,
    Sites
  ))
```

## Robust covariance and finite-cluster reference

For each run, the proposed calculation is:

1. reconstruct the final-model design matrix in fitted row order;
2. apply the accepted run-specific AR(1) whitening transformation to both the
   design matrix and raw residuals, resetting at every accepted `AR.start`;
3. sum the resulting score contributions over all rows and days within each
   participant;
4. form a participant-cluster CR1 sandwich covariance using
   $G/(G-1)\,(N-1)/(N-\mathrm{edf})$;
5. add `Vp - Ve`, the same smoothing-bias covariance component used by
   `mgcv` for its Bayesian sandwich covariance; and
6. evaluate the joint curve contrast with `mgcv`'s fractional-rank Wald test
   and a finite-cluster F reference. The denominator degrees of freedom are
   $G-\lceil r\rceil$, where $G$ is the number of participants and $r$ is the
   numerator reference rank.

This covariance is robust to heteroskedasticity, non-Gaussian score variance,
and residual dependence remaining within participants after the working AR(1)
model. It does not require independent half-hour observations.

```{r}
#| label: feasibility-diagnostics
#| echo: false

diagnostics |>
  transmute(
    Placement = recode(
      .data$placement,
      glasses = "Near eye",
      chest = "Chest"
    ),
    `AR-whitened residual check` = format(
      .data$residual_contract_maximum_absolute_error,
      scientific = TRUE,
      digits = 2
    ),
    `Largest participant meat share` =
      .data$maximum_participant_unscaled_meat_share,
    `Effective participants (trace diagnostic)` =
      round(.data$effective_participants_unscaled_meat_trace, 1),
    `Curve/coefficient-block p difference` =
      .data$curve_vs_coefficient_block_p_absolute_difference
  ) |>
  amendment_gt(
    "Feasibility and influence diagnostics",
    paste(
      "P-values themselves remain sealed as feasibility-only outputs at this",
      "approval gate; the final column checks representation invariance."
    )
  ) |>
  fmt_percent(
    columns = `Largest participant meat share`,
    decimals = 1
  ) |>
  fmt_scientific(
    columns = `Curve/coefficient-block p difference`,
    decimals = 2
  )
```

The AR-whitened residual reconstruction matches the fitted standardized
residuals to machine precision. Every audited covariance is positive
semidefinite up to floating-point rounding. No participant contributes 10% or
more of the unscaled global-test meat. The 48-bin curve and the corresponding
11-coefficient sex block give practically identical feasibility values.

These checks support numerical stability; they do not replace the final-model
residual, basis-dimension, concurvity, cyclic-closure, and site-influence
diagnostics required in Stage 2.

## Alternatives rejected

| Alternative | Decision | Reason |
|---|---|---|
| Compare `M0` and `M_pattern` by fREML AIC or a restricted-likelihood ratio | Reject | Their fixed mean spaces differ, so this would exchange runtime for an invalid formal comparison. |
| Use only the ordinary `mgcv` parametric and smooth summary tests | Reject | They split the registered complete sex contribution and do not aggregate uncertainty at the participant level. |
| Treat half-hour observations as independent in a sandwich covariance | Reject | It discards the accepted participant/day hierarchy and is not robust to remaining within-participant dependence. |
| Participant-level permutation of biological sex | Reject | Biological sex was observed, not randomized; exchangeability under arbitrary relabelling is not justified. |
| Make a bootstrap the primary replacement | Defer | It is unnecessary for the global test and would introduce another compute gate. The analytic test has 141 or 154 participant clusters. |

## Multiplicity, intervals, and effect-size gate

The proposed multiplicity structure remains:

- one primary near-eye global test (`H11-F1-primary`, family size 1);
- a two-test near-eye level/shape decomposition with BH adjustment
  (`H11-F2-level-shape`);
- one complementary chest global test (`H11-C1-global`, family size 1); and
- a two-test chest level/shape decomposition with BH adjustment
  (`H11-C2-level-shape`).

Sensitivity p-values remain descriptive and cannot create or rescue a primary
near-eye result. Every displayed raw and BH-adjusted p-value will follow
REPORT-008 independently; full precision remains in scientific artifacts.

The author-approved interval amendment is retained: the curve display uses
pointwise, not simultaneous, 95% confidence intervals. For internal
coherence, this amendment proposes calculating those pointwise intervals from
the same participant-cluster-robust covariance and a participant-level
t-reference. They remain local descriptive intervals and will not define a
familywise-supported clock-time period.

The failed ML delta-AIC condition must be removed from the effect-size gate.
The H02-like descriptive effect-size branch opens for a placement only if its
global robust test has raw p < 0.050 in its one-test family. If it opens, one
fast `M0` fREML/discrete fit may be made solely to calculate the approved
in-sample fitted-prediction R-squared increment and curve-variation estimator.
That reduced fit is descriptive, not a likelihood comparison. Any hierarchical
effect-size bootstrap remains restricted to the already approved 50-replicate
pilot before a separate production-computation decision.

## Limits on the robustness claim

The test is asymptotic and linearized, conditional on the fitted cyclic basis,
smoothing-parameter selection, and accepted working AR structure. The
`Vp - Ve` addition is deliberately conservative about smooth estimation but
does not make basis selection random.

Participants are treated as independent clusters. Site remains fixed support;
with only nine near-eye and eight chest sites, this test cannot justify a
population-of-sites claim or absorb arbitrary dependence between participants
at one site. The planned leave-one-site-out analysis therefore remains a
required sensitivity before reader-facing interpretation.

The contrast is associational. It estimates temporal patterns by measured
biological sex and does not identify a causal effect of sex or a gender effect.

## Outcome-awareness disclosure

The exact feasibility calculation necessarily produced numerical test values.
Those values are stored at full precision in H11-scoped audit artifacts and
explicitly labelled `method_feasibility_only_not_accepted_inference`. They are
not displayed or interpreted in this proposal. The method was selected for
its estimand match, participant-level robustness, compatibility with the
accepted H02 structure, and measured runtime—not because of whether a
threshold was crossed.

The author should decide on the method before treating those stored values as
H11 results. This amendment will be recorded as an outcome-aware registered
deviation because the computation problem arose after final full-model fits
existed.

## Author approval checklist

Approve or revise each item:

1. **H11-METHOD-001:** remove the twelve planned `ML`, `discrete = FALSE`
   comparison fits and their delta-AIC support rule.
2. **H11-METHOD-002:** use the complete 48-bin Female-minus-Male curve as the
   global null, calculated from the existing final `M_pattern` fREML model.
3. **H11-METHOD-003:** use the AR-whitened participant-cluster CR1 sandwich,
   add `Vp - Ve`, and use the stated finite-cluster fractional-rank F reference.
4. **H11-METHOD-004:** retain the one-test global families and the two-test
   BH-adjusted level/shape decomposition families.
5. **H11-METHOD-005:** use participant-cluster-robust pointwise 95% curve
   intervals; do not infer a significant time period from them.
6. **H11-METHOD-006:** open the H02-like descriptive effect-size branch only
   when the placement-specific global robust p-value is below 0.050; fit the
   reduced fREML model only after that gate.
7. **H11-METHOD-007:** retain leave-one-site-out sensitivity and the explicit
   fixed-site, associational scope limitations.

**Stage 2 remains at the method-amendment author gate.** No feasibility value
will be promoted to an accepted H11 result until these decisions are explicitly
approved in this H11 task.
